### 2018-02-08 Bivariate_analysis/Bivariate_analysis.Rmd
Rmd file for BIVAN

---
title: "Bivariate_analysis on different datasets"
author: "Shyamolina Ghosh"
date: "February 8, 2018"
output: 
  html_document:
    fig_caption: yes
    fig_width: 5
    highlight: zenburn
    keep_tex: yes
    number_sections: yes
    df_print: kable # data frame printing option
fontsize: 11pt
---
updated on `r Sys.Date()`
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<mark>*Code preloaded*</mark>

```{r BIVAN, echo=F}
#----------------------
rm(list=ls())
graphics.off()
library(copula)
library(VineCopula)
#----------------------
source("OurBiCopSelect.R")
#----------------------
# This function takes a bivariable dataset (matrix or dataframe)
# Output : a plot(optional) and rankings as a dataframe

getcopula<-function(d,ploton){

  #get ranks
  v1<-d[,1]
  v2<-d[,2]

  #get ranks modified now
  if(max(rbind(v1,v2))>=1){
    v1<-pobs(v1)
    v2<-pobs(v2)
  }

  if(ploton==T){
    plot(v1,v2,type="p",col="red",xlab=expression(v[1]),ylab=expression(v[2]),cex.lab=1)
  }

  cop<-data.frame(v1=v1,v2=v2)
  return(cop)
}

#Test the function
#cop<-claytonCopula(8,2)
#d<-rCopula(1000,cop)
#cop_rank<-(getcopula(d=d,ploton=T))

#--------------------------------------------------------------------------------------------------
# This function fits the data within the given set of families
# Input : given data, a vector of possible copula families, indep_test_p threshold
# output : a list of two dataframe: 1. all possible info for bestfit cop, 
#                          2. info for all possible fit cop(copcode,copname,par,par1,logLik,AIC,BIC)

fit_BIV<-function(d,families,level=0.05){

  num_indep<-0

  m<-getcopula(d,ploton=F)
  u1<-m[,1]
  u2<-m[,2]

  ans<-OurBiCopSelect(u1,u2,families,AICBIC="AIC",level=0.05,numBS=4,gofnormal=T)
  if(ans$IndepTestRes<level){
    gfc_numBS<-100
    gfc_p_CvM<-ans$GofRes_CvM
    gfc_p_KS<-ans$GofRes_KS
    gfc_p_CvM_stat<-ans$GofRes_CvM_stat
    gfc_p_KS_stat<-ans$GofRes_KS_stat

    gfc_normal_p_CvM<-ans$GofRes_Normal_CvM
    gfc_normal_p_KS<-ans$GofRes_Normal_KS
    gfc_normal_p_CvM_stat<-ans$GofRes_Normal_CvM_stat
    gfc_normal_p_KS_stat<-ans$GofRes_Normal_KS_stat

    pmin<-min(ans$GofRes_CvM,ans$GofRes_KS,ans$GofRes_Normal_CvM,ans$GofRes_Normal_KS)
    if(pmin<0.2){
      ans2<-OurBiCopSelect(u1,u2,families,AICBIC="AIC",level=0.05,numBS=5,gofnormal=T)
      gfc_numBS<-1000
      gfc_p_CvM<-ans2$GofRes_CvM
      gfc_p_KS<-ans2$GofRes_KS
      gfc_p_CvM_stat<-ans$GofRes_CvM_stat
      gfc_p_KS_stat<-ans$GofRes_KS_stat

      gfc_normal_p_CvM<-ans2$GofRes_Normal_CvM
      gfc_normal_p_KS<-ans2$GofRes_Normal_KS
      gfc_normal_p_CvM_stat<-ans$GofRes_Normal_CvM_stat
      gfc_normal_p_KS_stat<-ans$GofRes_Normal_KS_stat
    }

    info<-ans$InfCritRes
    info_ord<-info[order(info[,6],decreasing = F),] # ordered according to min to max AIC

  }else{
    num_indep<-num_indep+1
    gfc_numBS<-Inf
    gfc_p_CvM<-Inf
    gfc_p_KS<-Inf
    gfc_p_CvM_stat<-Inf
    gfc_p_KS_stat<-Inf
    gfc_normal_p_CvM<-Inf 
    gfc_normal_p_KS<-Inf
    gfc_normal_p_CvM_stat<-Inf
    gfc_normal_p_KS_stat<-Inf
    info_ord<-data.frame(copname="Indep")
  }

#-----------------------------------------------------------------------------
# Save the results
  RES<-list(data.frame(num_indep_location=num_indep,
               gfc_numBS=gfc_numBS,
               gfc_p_CvM=gfc_p_CvM,
               gfc_p_KS=gfc_p_KS,
               gfc_p_CvM_stat=gfc_p_CvM_stat,
               gfc_p_KS_stat=gfc_p_KS_stat,
               gfc_normal_p_CvM=gfc_normal_p_CvM,
               gfc_normal_p_KS=gfc_normal_p_KS,
               gfc_normal_p_CvM_stat=gfc_normal_p_CvM_stat,
               gfc_normal_p_KS_stat=gfc_normal_p_KS_stat),
               info_ord=info_ord)

  return(RES)

}

#-------------------------------------------------------------------------------------
#Check the function
#families<-c(1,3,5)
#x<-fit_BIV(d,families)
#---------------------------------------------------------------------------------------------------------

```

# Testing code

* **About the dataset :** This data came from a clayton with param=8

```{r loaddatatest, echo=F}
# here you can directly load the bivariate data to the variable d
# or you can save the bivariate data to d after some data extraction
d<-readRDS("givendata.RDS")
```

* **Copula plot of the data :**
```{r callingcodetest, echo=F, cache=T, cache.extra=list(d)}
#Test the function
cop_rank<-(getcopula(d=d,ploton=T))

families<-c(1,3,5)
x<-fit_BIV(d=d,families)
```

* **Results :**
```{r restest, echo=F}
knitr::kable(x[[1]])
knitr::kable(x[[2]])
```

# Results from dataset1 : Prof. T. Loecke
<mark>*You have to edit this section*</mark>

* **About the dataset :** <mark>*Please provide a brief description*</mark>

```{r loaddata1, echo=F}
# here you can directly load the bivariate data to the variable d
# or you can save the bivariate data to d after some data extraction
d<-readRDS("givendata.RDS")
```

* **Copula plot of the data :**
```{r callingcode1, echo=F, cache=T, cache.extra=list(d)}
#Test the function
cop_rank<-(getcopula(d=d,ploton=T))

families<-c(1,3,5)
x<-fit_BIV(d=d,families)
```

* **Results :**
```{r res1, echo=F}
knitr::kable(x[[1]])
knitr::kable(x[[2]])
```

# Results from dataset2 : Prof. ____________

### 2018-02-08 All_analysis.Rmd
file management

--- 2018-02-08 Bivariate_analysis/Bivariate_analysis.Rmd
@@
 # or you can save the bivariate data to d after some data extraction
-d<-readRDS("givendata.RDS")
+d<-readRDS("../Data/givendata.RDS")
 ```

### 2018-02-08 All_analysis.Rmd
modified

--- 2018-02-08 Bivariate_analysis/Bivariate_analysis.Rmd
@@
 
-  ans<-OurBiCopSelect(u1,u2,families,AICBIC="AIC",level=0.05,numBS=4,gofnormal=T)
+  ans<-OurBiCopSelect(u1,u2,families,AICBIC="AIC",level=0.05,numBS=100,gofnormal=T)
   if(ans$IndepTestRes<level){
@@
     if(pmin<0.2){
-      ans2<-OurBiCopSelect(u1,u2,families,AICBIC="AIC",level=0.05,numBS=5,gofnormal=T)
+      ans2<-OurBiCopSelect(u1,u2,families,AICBIC="AIC",level=0.05,numBS=1000,gofnormal=T)
       gfc_numBS<-1000
@@
 # or you can save the bivariate data to d after some data extraction
-d<-readRDS("givendata.RDS")
+d<-readRDS("Data/givendata.RDS")
 ```

### 2018-02-12 All_analysis.Rmd
cleaning up and adding ancillary features, and a separate testing doc

@@
 ```{r BIVAN, echo=F}
-#----------------------
-rm(list=ls())
-graphics.off()
 library(copula)
-library(VineCopula)
-#----------------------
 source("OurBiCopSelect.R")
-#----------------------
+
 # This function takes a bivariable dataset (matrix or dataframe)

### 2018-02-12 Test_OurBiCopSelect.Rmd
cleaning up and adding ancillary features, and a separate testing doc

---
title: "For testing OurBiCopSelect.R"
author: "Daniel Reuman"
date: "February 12, 2018"
output: pdf_document
---

This is just for testing the `OurBiCopSelect.R` function. We test that
it runs and gives reasonable results by simulating data and seeing if 
you recover what kind of copula was used.

<!--Source in the function and a package, changeable constants-->
```{r}
source("OurBiCopSelect.R")
require(copula)
nbsmall<-8
pthresh<-0.2
nblarge<-15 #for bootstrapping, see below
numdat<-100 #amount of artificial data
```

<!--A function needed for caching-->
```{r echo=F}
mtime <- function(files)
{
  lapply(Sys.glob(files),function(x) X =   
          file.info(x)$mtime)
}
```

If real data come from a Clayton copula, independence should
be rejected, Clayton should be the AIC best, it should pass the
goodness of fit test, and the normal copula should fail the
goodness of fit test:
<!--Test for real data from a clayton-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,mtime("OutBiCopSelect.R"))}
families<-c(1,3:6,13,14,16)

#data are really from a clayton
cop<-claytonCopula(5,2)
dat<-rCopula(numdat,cop)
allresC<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresC
```

Same test, but data now come from a Gumbel copula:
<!--Test for real data from a gumbel-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,mtime("OutBiCopSelect.R"))}
#data are really from a gumbel
cop<-gumbelCopula(4,2)
dat<-rCopula(numdat,cop)
allresG<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresG
```

Same test, but data now come from a Frank copula:
<!--Test for real data from a frank-->
```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,mtime("OutBiCopSelect.R"))}
#data are really from a frank
cop<-frankCopula(5,2)
dat<-rCopula(numdat,cop)
allresF<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
                        numBSlarge=nblarge,pthresh=0.2)
allresF
```

### 2018-02-12 OurBiCopSelect.R

@@ -1,130 +1,172 @@
-#This function gives you all the info at a selective location pair
-OurBiCopSelect<-function(u1,u2,families,AICBIC="AIC",level=0.05,numBS,gofnormal){
-  #first test for independence
-  indeptest<-BiCopIndTest(u1,u2)$p.value
+library(VineCopula)
+
+#This function takes a sample from a bivariate distribution
+#consisting of values between 0 and 1 and:
+#a) tests for independence
+#b) if independence can be rejected, it fit a series of bivariate 
+#   copula models and provides fitted parameters, log likelihood, 
+#   AIC and BIC for each
+#c) for the best copula model (by AIC or BIC), the function tests
+#   for the goodness of fit in two ways
+#d) optionally tests for the goodness of fit of the normal copula
+#The function is a re-implementation of the BiCopSelect function
+#in the VineCopula package, re-done to ensure we could control
+#what we were getting.
+#
+#Args
+#u1, u2       The sample. These can be obtained, for instance, 
+#               by applying pobs in the VineCopula package to
+#               samples from any bivariate distribution or to 
+#               any bivariate dataset
+#level        Significance level to be used for the independence
+#               test
+#families     The copula families to use in b above. Uses codes
+#               as specified in the docs for BiCopEst in the 
+#               VineCopula package.
+#AICBIC       Which one to use for model selection.
+#numBSsmall
+#pthresh
+#numBSlarge   Number of bootstraps to use for goodness of fit.
+#               First, tests are run with numBSsmall, and if p-values
+#               are ever less than pthresh, tests are re-run with
+#               numBSlarge. Watch out, large values can mean long 
+#               run times.
+#gofnormal    T/F on whether a goodness of fit test for the normal
+#               copula should be run.
+#status       T/F - should status updates on the run be printed?
+#
[…]

### 2018-02-13 Test_OurBiCopSelect.Rmd
small changes

@@
 require(copula)
-nbsmall<-8
+nbsmall<-100
 pthresh<-0.2
-nblarge<-15 #for bootstrapping, see below
-numdat<-100 #amount of artificial data
+nblarge<-1000 #for bootstrapping, see below
+numdat<-500 #amount of artificial data
+seed<-101
 ```
@@
 <!--Test for real data from a clayton-->
-```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,mtime("OutBiCopSelect.R"))}
+```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
+set.seed(seed)
+
 families<-c(1,3:6,13,14,16)
@@
 <!--Test for real data from a gumbel-->
-```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,mtime("OutBiCopSelect.R"))}
+```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
+set.seed(seed)
+
 #data are really from a gumbel
-cop<-gumbelCopula(4,2)
+cop<-gumbelCopula(6,2)
 dat<-rCopula(numdat,cop)
@@
 <!--Test for real data from a frank-->
-```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,mtime("OutBiCopSelect.R"))}
+```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
+set.seed(seed)
+
 #data are really from a frank
@@
 ```
+
+Same test, but data now come from a Joe copula:
+<!--Test for real data from a joe-->
+```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
+set.seed(seed)
+
+#data are really from a joe
+cop<-joeCopula(6,2)
+dat<-rCopula(numdat,cop)
+allresJ<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
+                        numBSlarge=nblarge,pthresh=0.2)
+allresJ
+```

### 2018-02-13 All_analysis.Rmd
separate the plotter function and make the tester doc into the beginning of a sup mat

@@
 ---
-title: "Bivariate_analysis on different datasets"
-author: "Shyamolina Ghosh"
+title: "Exploring the importance of a complete statistical description of dependence in ecology and evolution"
+author: "Shyamolina Ghosh and Daniel Reuman"
 date: "February 8, 2018"
@@
 ---
-updated on `r Sys.Date()`
+
 ```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
+seed<-101
+families<-c(1,3:6,14)
+source("getcopula.R")
+source("OurBiCopSelect.R")
 ```
 
-<mark>*Code preloaded*</mark>
+Updated on `r Sys.Date()`
 
-```{r BIVAN, echo=F}
-library(copula)
-source("OurBiCopSelect.R")
+# Template analysis
 
-# This function takes a bivariable dataset (matrix or dataframe)
-# Output : a plot(optional) and rankings as a dataframe
+**About the dataset :** These data were simulated from a Clayton copula with parameter 8.
 
-getcopula<-function(d,ploton){
-
-  #get ranks
-  v1<-d[,1]
-  v2<-d[,2]
-
-  #get ranks modified now
-  if(max(rbind(v1,v2))>=1){
-    v1<-pobs(v1)
-    v2<-pobs(v2)
-  }
-
-  if(ploton==T){
-    plot(v1,v2,type="p",col="red",xlab=expression(v[1]),ylab=expression(v[2]),cex.lab=1)
-  }
-
-  cop<-data.frame(v1=v1,v2=v2)
-  return(cop)
-}
-
-#Test the function
-#cop<-claytonCopula(8,2)
-#d<-rCopula(1000,cop)
-#cop_rank<-(getcopula(d=d,ploton=T))
-
-#--------------------------------------------------------------------------------------------------
-# This function fits the data within the given set of families
-# Input : given data, a vector of possible copula families, indep_test_p threshold
-# output : a list of two dataframe: 1. all possible info for bestfit cop, 
-#                          2. info for all possible fit cop(copcode,copname,par,par1,logLik,AIC,BIC)
-
-fit_BIV<-function(d,families,level=0.05){
-
-  num_indep<-0
-
-  m<-getcopula(d,ploton=F)
-  u1<-m[,1]
-  u2<-m[,2]
-
-  ans<-OurBiCopSelect(u1,u2,families,AICBIC="AIC",level=0.05,numBS=100,gofnormal=T)
-  if(ans$IndepTestRes<level){
-    gfc_numBS<-100
-    gfc_p_CvM<-ans$GofRes_CvM
-    gfc_p_KS<-ans$GofRes_KS
-    gfc_p_CvM_stat<-ans$GofRes_CvM_stat
-    gfc_p_KS_stat<-ans$GofRes_KS_stat
-
-    gfc_normal_p_CvM<-ans$GofRes_Normal_CvM
-    gfc_normal_p_KS<-ans$GofRes_Normal_KS
-    gfc_normal_p_CvM_stat<-ans$GofRes_Normal_CvM_stat
-    gfc_normal_p_KS_stat<-ans$GofRes_Normal_KS_stat
-
-    pmin<-min(ans$GofRes_CvM,ans$GofRes_KS,ans$GofRes_Normal_CvM,ans$GofRes_Normal_KS)
-    if(pmin<0.2){
-      ans2<-OurBiCopSelect(u1,u2,families,AICBIC="AIC",level=0.05,numBS=1000,gofnormal=T)
-      gfc_numBS<-1000
-      gfc_p_CvM<-ans2$GofRes_CvM
-      gfc_p_KS<-ans2$GofRes_KS
-      gfc_p_CvM_stat<-ans$GofRes_CvM_stat
-      gfc_p_KS_stat<-ans$GofRes_KS_stat
-
-      gfc_normal_p_CvM<-ans2$GofRes_Normal_CvM
-      gfc_normal_p_KS<-ans2$GofRes_Normal_KS
-      gfc_normal_p_CvM_stat<-ans$GofRes_Normal_CvM_stat
-      gfc_normal_p_KS_stat<-ans$GofRes_Normal_KS_stat
-    }
-
-    info<-ans$InfCritRes
-    info_ord<-info[order(info[,6],decreasing = F),] # ordered according to min to max AIC
-
-  }else{
-    num_indep<-num_indep+1
-    gfc_numBS<-Inf
-    gfc_p_CvM<-Inf
-    gfc_p_KS<-Inf
-    gfc_p_CvM_stat<-Inf
-    gfc_p_KS_stat<-Inf
-    gfc_normal_p_CvM<-Inf 
-    gfc_normal_p_KS<-Inf
-    gfc_normal_p_CvM_stat<-Inf
-    gfc_normal_p_KS_stat<-Inf
-    info_ord<-data.frame(copname="Indep")
-  }
-
-#-----------------------------------------------------------------------------
-# Save the results
-  RES<-list(data.frame(num_indep_location=num_indep,
-               gfc_numBS=gfc_numBS,
-               gfc_p_CvM=gfc_p_CvM,
-               gfc_p_KS=gfc_p_KS,
-               gfc_p_CvM_stat=gfc_p_CvM_stat,
-               gfc_p_KS_stat=gfc_p_KS_stat,
-               gfc_normal_p_CvM=gfc_normal_p_CvM,
-               gfc_normal_p_KS=gfc_normal_p_KS,
-               gfc_normal_p_CvM_stat=gfc_normal_p_CvM_stat,
-               gfc_normal_p_KS_stat=gfc_normal_p_KS_stat),
-               info_ord=info_ord)
-
-  return(RES)
-
-}
-
-#-------------------------------------------------------------------------------------
-#Check the function
-#families<-c(1,3,5)
-#x<-fit_BIV(d,families)
-#---------------------------------------------------------------------------------------------------------
-
-```
-
-# Testing code
-
-* **About the dataset :** This data came from a clayton with param=8
-
-```{r loaddatatest, echo=F}
-# here you can directly load the bivariate data to the variable d
-# or you can save the bivariate data to d after some data extraction
+```{r load_data_test, echo=F}
+#Load your data - should be a data frame with 2 columns. Can contain NAs.
 d<-readRDS("Data/givendata.RDS")
@@
 
-* **Copula plot of the data :**
-```{r callingcodetest, echo=F, cache=T, cache.extra=list(d)}
-#Test the function
-cop_rank<-(getcopula(d=d,ploton=T))
+**Copula plot of the data :**
+```{r rankandplot_test, echo=F, cache=T, cache.extra=list(d)}
+#convert to a copula and plot
+v<-getcopula(d=d,rankon=F,ploton=T) #with real data, rankon should be T
+```
 
-families<-c(1,3,5)
+**Bivariate copula model selection and goodness of fit pipeline:**
+```{r bivmodselect_test, echo=F, cache=T, cache.extra=list(seed,d)}
 x<-fit_BIV(d=d,families)

### 2018-02-13 SupportingInformation.Rmd
separate the plotter function and make the tester doc into the beginning of a sup mat

--- 2018-02-13 Test_OurBiCopSelect.Rmd
@@
 ---
-title: "For testing OurBiCopSelect.R"
-author: "Daniel Reuman"
+title: "Supporting Information"
+author: "Shyamolina Ghosh and Daniel Reuman"
 date: "February 12, 2018"
@@
 
-This is just for testing the `OurBiCopSelect.R` function. We test that
+# Descriptions of data
+
+# Testing the ranking and bivariate copula display function
+
+This is just for testing the `getcopula` function.
+
+<!--Source in function and a package--->
+```{r}
+source("getcopula.R")
+require(copula)
+require(mvtnorm)
+```
+
+Plot data from a Clayton copula, which has lower tail dependence:
+```{r}
+copC<-claytonCopula(8,2)
+dC<-rCopula(1000,copC)
+vC<-getcopula(d=dC,rankon=F,ploton=T)
+```
+
+Plot data from a Joe copula, which has upper tail dependence:
+```{r}
+copJ<-joeCopula(8,2)
+dJ<-rCopula(1000,copJ)
+vJ<-getcopula(d=dJ,rankon=F,ploton=T)
+```
+
+Get data from a bivariate normal distribution and take ranks and
+then plot:
+```{r}
+sigma<-matrix(c(1,.7,.7,1),2,2)
+dN<-rmvnorm(1000,mean=c(0,0),sigma=sigma)
+vN<-getcopula(d=dN,rankon=T,ploton=T)
+```
+
+# Testing the bivariate model selection and goodness-of-fit pipeline
+
+This is just for testing the `OurBiCopSelect` function. We test that
 it runs and gives reasonable results by simulating data and seeing if 
@@
 ```
+
+# Testing the pipeline for examining tail dependence in bivariate data

### 2018-02-13 getcopula.R

@@ -1,0 +1,37 @@
+library(VineCopula)
+
+#This function takes a bivariable dataset (matrix or dataframe) and gets the
+#associated sample from the underlying copula by taking ranks.
+#
+#Args
+#d          An N by 2 matrix or data frame
+#rankon     Should ranks be taken? This could be F, for instance, if data 
+#             were simulated from a copula.
+#ploton     Should a plot of the copula be generated?
+#
+#Note: An NA means that whole row is automatically removed.
+#
+#Output : a plot(optional) and rankings as a matrix
+#
+getcopula<-function(d,rankon=T,ploton=F)
+{
+  #get rid of NAs
+  inds<-which(is.finite(d[,1]) && is.finite(d[,2]))
+  d<-d[inds,]
+  
+  #get ranks
+  if (rankon)
+  {
+    v<-VineCopula::pobs(d)
+  }else
+  {
+    v<-d
+  }
+  
+  #plot
+  if(ploton==T){
+    plot(v[,1],v[,2],type="p",col="red",xlab=expression(v[1]),ylab=expression(v[2]),cex.lab=1)
+  }
+  
+  return(v)
+}

### 2018-02-14 All_analysis.Rmd
added the analysis of Loecke data

@@
     fig_width: 5
-    highlight: zenburn
+    highlight: tango
     keep_tex: yes
@@
 
-```{r setup, include=FALSE}
+```{r setup, echo=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
@@
 
-**About the dataset :** These data were simulated from a Clayton copula with parameter 8.
+**About the dataset :**  
+These data were simulated from a Clayton copula with parameter 8.
 
+**Plot the raw data:**  
 ```{r load_data_test, echo=F}
@@
 d<-readRDS("Data/givendata.RDS")
+plot(d[,1],d[,2],type="p",col="red",xlab=expression(d[1]),ylab=expression(d[2]),cex.lab=1,
+     xlim=c(0,1),ylim=c(0,1),asp=1)
+rect(0,0,1,1)
 ```
 
-**Copula plot of the data :**
-```{r rankandplot_test, echo=F, cache=T, cache.extra=list(d)}
+**Copula plot of the data:**  
+```{r rankandplot_test, echo=F}
 #convert to a copula and plot
@@
 **Bivariate copula model selection and goodness of fit pipeline:**  
-```{r bivmodselect_test, echo=F, cache=T, cache.extra=list(seed,d)}
-x<-fit_BIV(d=d,families)
+```{r bivmodselect_test, echo=F, cache=T, cache.extra=list(seed,v,families)}
+set.seed(seed)
+BivMS_res_test<-OurBiCopSelect(v[,1],v[,2],families=families,gofnormal=FALSE,status=TRUE)
 ```
 
-* **Results :**
-```{r restest, echo=F}
-knitr::kable(x[[1]])
-knitr::kable(x[[2]])
+```{r}
+BivMS_res_test
 ```
 
-# Results from dataset1 : Prof. T. Loecke
-<mark>*You have to edit this section*</mark>
+# Results from dataset1 : T. Loecke
 
-* **About the dataset :** <mark>*Please provide a brief description*</mark>
+**About the dataset :**  
+Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description).
 
-```{r loaddata1, echo=F}
-# here you can directly load the bivariate data to the variable d
-# or you can save the bivariate data to d after some data extraction
-d<-readRDS("Data/givendata.RDS")
+**Plot the raw data:**  
+```{r load_data_Loecke, echo=F}
+#Load your data - should be a data frame with 2 columns. Can contain NAs.
+d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
+d<-d[,c("SOCstock100","TSNstock100")]
+plot(d[,1],d[,2],type="p",col="red",xlab=expression(d[1]),ylab=expression(d[2]),cex.lab=1)
+rect(0,0,1,1)
 ```
 
-* **Copula plot of the data :**
-```{r callingcode1, echo=F, cache=T, cache.extra=list(d)}
-#Test the function
-cop_rank<-(getcopula(d=d,ploton=T))
-
-families<-c(1,3,5)
-x<-fit_BIV(d=d,families)
+**Copula plot of the data:**  
+```{r rankandplot_Loecke, echo=F}
+#convert to a copula and plot
+v<-getcopula(d=d,rankon=T,ploton=T) #with real data, rankon should be T
 ```
 
-* **Results :**
-```{r res1, echo=F}
-knitr::kable(x[[1]])
-knitr::kable(x[[2]])
+**Bivariate copula model selection and goodness of fit pipeline:**  
+```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v)}
+set.seed(seed)
+families_Loecke<-c(1,3:10,13:14,16:20,104,114,204,214)
+BivMS_res_Loecke<-OurBiCopSelect(v[,1],v[,2],families=families_Loecke,
+                                 gofnormal=FALSE,status=TRUE)
 ```
 
-# Results from dataset2 : Prof. ____________
+```{r}
+BivMS_res_Loecke
+```
+
+None of the copulas we tried here was a good fit, but some are much 
+better than others, and the normal copula is not close to the best, 
+so that shows there is non-trivial (i.e., non-normal) copula structure.
+
+It looks like the data might be asymmetric about the y=x line, which
+might require a special copula to model. What is the significance,
+biologically?
+
+Strong dependence in the extreme right tails is evident. We could
+quantify this, and (probably) show it is significantly more than what one would expect from a null hypothesis of a normal copula with the same kendall as the data.
+What is the significance, biologically?

### 2018-02-14 SupportingInformation.Rmd
added the analysis of Loecke data

@@
 
+<!--A function needed for caching-->
+```{r echo=F}
+mtime <- function(files)
+{
+  lapply(Sys.glob(files),function(x) X =   
+          file.info(x)$mtime)
+}
+```
+
 # Descriptions of data
+Each data contributor writes one to a few paragraphs about their data.
 
@@
 seed<-101
-```
-
-<!--A function needed for caching-->
-```{r echo=F}
-mtime <- function(files)
-{
-  lapply(Sys.glob(files),function(x) X =   
-          file.info(x)$mtime)
-}
 ```
@@
 
+The goodness of fit test is quite slow when testing the fit of a
+normal copula, and the results are not great: often the normal copula
+cannot be rejected even when it is not the true copula, even though
+a comparison of A/BICs strongly supports the true copula over the normal.
+So I think the best way forward is to use `gofnormal=FALSE`.
+
 # Testing the pipeline for examining tail dependence in bivariate data

### 2018-02-14 getcopula.R

@@ -16,7 +16,7 @@ library(VineCopula)
 getcopula<-function(d,rankon=T,ploton=F)
 {
   #get rid of NAs
-  inds<-which(is.finite(d[,1]) && is.finite(d[,2]))
+  inds<-which(is.finite(d[,1]) & is.finite(d[,2]))
   d<-d[inds,]
   
   #get ranks
@@ -30,7 +30,10 @@ getcopula<-function(d,rankon=T,ploton=F)
   
   #plot
   if(ploton==T){
-    plot(v[,1],v[,2],type="p",col="red",xlab=expression(v[1]),ylab=expression(v[2]),cex.lab=1)
+    plot(v[,1],v[,2],type="p",col="red",xlab=expression(v[1]),ylab=expression(v[2]),cex.lab=1,
+         xlim=c(0,1),ylim=c(0,1),asp=1)
+    rect(0,0,1,1)
+    lines(c(0,1),c(0,1),type='l')
   }
   
   return(v)

### 2018-02-14 All_analysis.Rmd
mostly changes to implement the bivariate models selection on Terry's C/N data, and to randomize copula structure and see what it does to the distribution of some ecosystem function

@@
 source("OurBiCopSelect.R")
+```
+
+<!--A function needed for caching-->
+```{r echo=F}
+mtime <- function(files)
+{
+  lapply(Sys.glob(files),function(x) X =   
+          file.info(x)$mtime)
+}
 ```
@@
 None of the copulas we tried here was a good fit, but some are much 
-better than others, and the normal copula is not close to the best, 
-so that shows there is non-trivial (i.e., non-normal) copula structure.
+better than others, and the normal copula is not close to the best. 
+So that shows there is non-trivial (i.e., non-normal) copula structure, which is an interesting conclusion.
+It is more interesting than it would have been to specifically identify a copula that fits, so I am not too concered about our failure to do that.
 
 It looks like the data might be asymmetric about the y=x line, which
-might require a special copula to model. What is the significance,
+might require a special copula to model. What is the significance of this,
 biologically?
@@
 What is the significance, biologically?
+We could (probably) also show there is much more asymmetry of tail dependence (upper v.
+lower) than would be expected from a normal-copula null hypothesis.
+
+Probably the most promising next step is to get a function $f(N,C)$ from Amy or Terry (or someone else) which is the expected value of some ecosystem function (or other quantity of interest), supported by a publication.
+Then calculate its value on all the sites and look at the distribution and calculate, say, its skewness or some other property.
+Then get surrogate, randomized data with normal copula structure (or some other copula structure - try Clayton because it has the reverse tail dependence).
+Then calculate the distribution of $f(N,C)$ for each surrogate dataset and its skewness (or other statistic) and compare.
+I already have code for this (`copsurrog2d.R`).
+So let's try this.
+Here is a function:
+
+```{r ecofunc_Loecke, cache=T, cache.extra=list(d,mtime("copsurrog2d.R"))}
+#Expected value of some ecosystem function for a 
+#soil based on soil N and C
+#
+#Args
+#N    Vector of N values for a bunch of soils
+#C    Vector (same length as N) of C values for the same soils
+#
+#Output
+#Expected value of some ecosystem function for those soils (vector of the same length as N and C).
+#
+ecofunc<-function(N,C)
+{
+  #right now this is just a placeholder, a real formula and a specific ecosystem function of interest to be specified by Amy and Terry
+  return(N+C)
+}
+
+source("copsurrog2d.R")
+source("SkewnessAnd3CentMom.R")
+inds<-which(is.finite(d[,1]) & is.finite(d[,2]))
+d<-d[inds,]
+
+#compare to a normal-copula null
+cop<-normalCopula(.5,2)
+numresamp<-1000
+dsur<-copsurrog2d(m=d,targetcop=cop,numsurrog=numresamp)
+efval<-myskns(ecofunc(d[,2],d[,1]))
+
+efvalsurN<-NA*numeric(numresamp)
+for (counter in 1:numresamp)
+{
+  efvalsurN[counter]<-myskns(ecofunc(dsur[,2,counter],dsur[,1,counter]))
+}
+sum(efvalsurN<efval)/numresamp
+hist(efvalsurN,50)
+points(efval,0,col='red')
+
+#compare to a clayton-copula null
+cop<-claytonCopula(.5,2)
+numresamp<-1000
+dsur<-copsurrog2d(m=d,targetcop=cop,numsurrog=numresamp)
+efvalsurC<-NA*numeric(numresamp)
+for (counter in 1:numresamp)
+{
+  efvalsurC[counter]<-myskns(ecofunc(dsur[,2,counter],dsur[,1,counter]))
+}
+sum(efvalsurC<efval)/numresamp
+hist(efvalsurC,50)
+points(efval,0,col='red')
+
+#compare to a joe-copula null
+cop<-joeCopula(3,2)
+numresamp<-1000
+dsur<-copsurrog2d(m=d,targetcop=cop,numsurrog=numresamp)
+efvalsurJ<-NA*numeric(numresamp)
+for (counter in 1:numresamp)
+{
+  efvalsurJ[counter]<-myskns(ecofunc(dsur[,2,counter],dsur[,1,counter]))
+}
+sum(efvalsurJ<efval)/numresamp
+hist(efvalsurJ,50)
+points(efval,0,col='red')
+```

### 2018-02-14 copsurrog2d.R

@@ -52,47 +52,47 @@ copsurrog2d<-function(m,targetcop,numsurrog)
 }
 
 #***basic tests of the function
-datcop<-claytonCopula(5,2)
-numpts<-1000
-m<-rCopula(numpts,datcop)
-plot(m[,1],m[,2],type='p')
+#datcop<-claytonCopula(5,2)
+#numpts<-1000
+#m<-rCopula(numpts,datcop)
+#plot(m[,1],m[,2],type='p')
 
-tarcop<-gumbelCopula(3,2)
+#tarcop<-gumbelCopula(3,2)
 
-res<-copsurrog2d(m,tarcop,3)
-dim(res)
+#res<-copsurrog2d(m,tarcop,3)
+#dim(res)
 
-plot(res[,1,1],res[,2,1],type='p')
-BiCopGofTest(res[,1,1],res[,2,1],family=4)
-BiCopGofTest(res[,1,2],res[,2,2],family=4)
-BiCopGofTest(res[,1,3],res[,2,3],family=4)
+#plot(res[,1,1],res[,2,1],type='p')
+#BiCopGofTest(res[,1,1],res[,2,1],family=4)
+#BiCopGofTest(res[,1,2],res[,2,2],family=4)
+#BiCopGofTest(res[,1,3],res[,2,3],family=4)
 
-sum(sort(res[,1,1])==sort(m[,1]))
-sum(sort(res[,2,1])==sort(m[,2]))
-sum(sort(res[,1,2])==sort(m[,1]))
-sum(sort(res[,2,2])==sort(m[,2]))
-sum(sort(res[,1,3])==sort(m[,1]))
-sum(sort(res[,2,3])==sort(m[,2]))
+#sum(sort(res[,1,1])==sort(m[,1]))
+#sum(sort(res[,2,1])==sort(m[,2]))
+#sum(sort(res[,1,2])==sort(m[,1]))
+#sum(sort(res[,2,2])==sort(m[,2]))
+#sum(sort(res[,1,3])==sort(m[,1]))
+#sum(sort(res[,2,3])==sort(m[,2]))
 
 #***test to see whether changing the copula
 #markedly influences the distribution
 #of the mean
 
 #make the data
-datcop<-rotCopula(claytonCopula(5,2))
-d<-rCopula(10000,datcop)
-d<-qnorm(d,mean=0,sd=1)
+#datcop<-rotCopula(claytonCopula(5,2))
[…]

### 2018-02-15 All_analysis.Rmd
added a way of comparing the empirical cdf to that of a cloud of surrogates, and applied it

--- 2018-02-14 All_analysis.Rmd
@@
 source("OurBiCopSelect.R")
+source("copsurrog2d.R")
+source("SkewnessAnd3CentMom.R")
+source("empcdfcloud.R")
+```
+
+<!--A function needed for caching-->
+```{r echo=F}
+mtime <- function(files)
+{
+  lapply(Sys.glob(files),function(x) X =   
+          file.info(x)$mtime)
+}
 ```
@@
 
-**Plot the raw data:**  
+**Load and plot the raw data:**  
 ```{r load_data_test, echo=F}
@@
 **Bivariate copula model selection and goodness of fit pipeline:**  
-```{r bivmodselect_test, echo=F, cache=T, cache.extra=list(seed,v,families)}
+```{r bivmodselect_test, echo=F, cache=T, cache.extra=list(seed,v,families,mtime("OurBiCopSelect.R"))}
 set.seed(seed)
@@
 BivMS_res_test
+```
+
+**Surrogate datasets and tail dependence pipeline:**
+<!--Use normal-copula surrogates, calculate upper and lower tail dependence stats for
+real data and for the surrogates, and compare in the usual way. Also for the difference
+of the two stats (to measure asymmetry of tail dependence). Have to decide what tail 
+dependence stat to use.-->
+
+**Surrogate datasets and an "outcome" function:**
+You first choose an "outcome function" that takes as input the two variables of your 
+data and gives as output some third thing of interest. For instance, if the two columns
+are soil C and N, the outcome function might be expected soil respiration. Obviously
+you have to have a context and know some science about that context to write a function
+that makes sense. For testing the workflow, we just use the sum:
+
+```{r}
+ocf_test<-function(d)
+{
+  return(d[,1]+d[,2])
+}
+```
+
+Once you have such a function, you can compare the distribution of values of this
+function for your real data to what it would be if copula were the normal copula
+instead, but other statistical aspects are the same. We compare empirical cdfs in
+two ways:
+
+```{r outcomefunc_test, echo=F, cache=T, cache.extra=list(seed,d,mtime("copsurrog2d.R"),mtime("empcdfcloud.R"))}
+#get rid of NAs
+inds<-which(is.finite(d[,1]) & is.finite(d[,2]))
+d<-d[inds,]
+
+#normalize the data - won't do this with real data
+d<-qnorm(d)
+
+#get surrogate data
+cop<-normalCopula(.5,2)
+dsur<-copsurrog2d(m=d,targetcop=cop,numsurrog=1000)
+
+#compute the outcome function on d and dsur
+ocfd<-ocf_test(d)
+ocfsur<-apply(FUN=ocf_test,MARGIN=3,X=dsur)
+ecdfc_test<-empcdfcloud(ocfd,ocfsur,prob=c(0.005,.5,.995))
+```
+
+We can compare skewness of the outcome function to what you 
+would get if the copula were normal:
+
+```{r echo=F}
+realskew<-myskns(ocfd)
+surskew<-apply(FUN=myskns,MARGIN=2,X=ocfsur)
+sum(surskew>realskew)/length(surskew)
+hist(surskew,50)
+points(realskew,0,col='red')
+realskew
 ```
@@
 d<-d[,c("SOCstock100","TSNstock100")]
-plot(d[,1],d[,2],type="p",col="red",xlab=expression(d[1]),ylab=expression(d[2]),cex.lab=1)
+plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1)
 rect(0,0,1,1)
@@
 None of the copulas we tried here was a good fit, but some are much 
-better than others, and the normal copula is not close to the best, 
-so that shows there is non-trivial (i.e., non-normal) copula structure.
+better than others, and the normal copula is not close to the best. 
+So that shows there is non-trivial (i.e., non-normal) copula structure, which is an interesting conclusion.
+It is more interesting than it would have been to specifically identify a copula that fits, so I am not too concered about our failure to do that.
 
 It looks like the data might be asymmetric about the y=x line, which
-might require a special copula to model. What is the significance,
+might require a special copula to model. What is the significance of this,
 biologically?
 
-Strong dependence in the extreme right tails is evident. We could
+**Surrogate datasets and tail dependence pipeline:**
+<!--Use normal-copula surrogates, calculate upper and lower tail dependence stats for
+real data and for the surrogates, and compare in the usual way. Also for the difference
+of the two stats (to measure asymmetry of tail dependence). Have to decide what tail 
+dependence stat to use.-->
+
+Strong dependence in the extreme right tails is evident. We should
 quantify this, and (probably) show it is significantly more than what one would expect from a null hypothesis of a normal copula with the same kendall as the data.
 What is the significance, biologically?
+We could (probably) also show there is much more asymmetry of tail dependence (upper v.
+lower) than would be expected from a normal-copula null hypothesis.
+First do the parallel analysis for the test data, indicated but not completed above.
+
+**Surrogate datasets and an "outcome" function:**
+
+You first choose an "outcome function" that takes as input the two variables of your 
+data and gives as output some third thing of interest. This is a place holder now, 
+have asked for a real function from Terry and Amy. May be more than one for different
+ecosystem processes.
+```{r}
+ocf_Loecke<-function(d)
+{
+  C<-d[,1]
+  N<-d[,2]
+  return(C+N)
+}
+```
+
+Once you have such a function, you can compare the distribution of values of this
+function for your real data to what it would be if copula were the normal copula
+instead, but other statistical aspects are the same. We compare empirical cdfs in two ways:
+
+```{r outcomefunc_Loecke, echo=F, cache=T, cache.extra=list(seed,d,mtime("copsurrog2d.R"),mtime("empcdfcloud.R"))}
+#get rid of NAs
+inds<-which(is.finite(d[,1]) & is.finite(d[,2]))
+d<-d[inds,]
+
+#get surrogate data
+cop<-normalCopula(.5,2)
+dsur<-copsurrog2d(m=d,targetcop=cop,numsurrog=1000)
+
+#compute the outcome function on d and dsur
+ocfd<-ocf_Loecke(d)
+ocfsur<-apply(FUN=ocf_Loecke,MARGIN=3,X=dsur)
+ecdfc_Loecke<-empcdfcloud(ocfd,ocfsur,prob=c(.005,.5,.995))
+```
+
+We can compare skewness of the outcome function to what you 
+would get if the copula were normal:
+
+```{r echo=F}
+realskew<-myskns(ocfd)
+surskew<-apply(FUN=myskns,MARGIN=2,X=ocfsur)
+sum(surskew>realskew)/length(surskew)
+hist(surskew,50)
+points(realskew,0,col='red')
+realskew
+```
+
+We can compare skewness of the outcome function to what you 
+would get if the copula were Clayton:
+
+```{r echo=F}
+#get surrogate data
+cop<-claytonCopula(5,2)
+dsur<-copsurrog2d(m=d,targetcop=cop,numsurrog=1000)
+
+#compute the outcome function on d and dsur
+ocfd<-ocf_Loecke(d)
+ocfsur<-apply(FUN=ocf_Loecke,MARGIN=3,X=dsur)
+
+realskew<-myskns(ocfd)
+surskew<-apply(FUN=myskns,MARGIN=2,X=ocfsur)
+sum(surskew>realskew)/length(surskew)
+hist(surskew,50)
+points(realskew,0,col='red')
+realskew
+```
+
+We can compare skewness of the outcome function to what you 
+would get if the copula were Joe:
+
+```{r echo=F}
+#get surrogate data
+cop<-joeCopula(5,2)
+dsur<-copsurrog2d(m=d,targetcop=cop,numsurrog=1000)
+
+#compute the outcome function on d and dsur
+ocfd<-ocf_Loecke(d)
+ocfsur<-apply(FUN=ocf_Loecke,MARGIN=3,X=dsur)
+
+realskew<-myskns(ocfd)
+surskew<-apply(FUN=myskns,MARGIN=2,X=ocfsur)
+sum(surskew>realskew)/length(surskew)
+hist(surskew,50)
+points(realskew,0,col='red')
+realskew
+```

### 2018-02-15 empcdfcloud.R

@@ -1,0 +1,63 @@
+require(stats)
+
+#This function compares the empirical cdf of a univariate dataset to a cloud
+#of empirical cdfs of surrogate datasets of the same format, visually.
+#
+#Args
+#d          A numeric vector, assumed no NAs
+#m          A numeric matrix, number of rows equal to the length of d. Assumed
+#             no NAs.
+#
+#Output: a list with these named elements
+#dcdf       An ecdf object (see the ecdf function in the stats package)
+#x          A vector of x axis values spanning from min(d) to max(d)
+#mcdf       The cdfs of the columns of m evaluated at the values of x
+#prob       Quantiles to use in plotting
+#
+#A plot is generated showing the quantiles of the cloud of empirical cdfs of 
+#the columns of m, with the empirical cdf of d plotted on top in red.
+#
+#A second plot is generated showing the fraction of surrogate empirical cdfs
+#the real empirical cdf is greater than.
+#
+empcdfcloud<-function(d,m,probs=c(.005,.5,.995))
+{
+  #get the empirical cdf of d and plot
+  dcdf<-ecdf(d)
+  plot(dcdf,main='',col='red')
+  
+  #compute the surrogate empirical cdfs
+  x<-seq(from=min(d),to=max(d),length.out=250)
+  mcdf<-matrix(NA,length(x),dim(m)[2])
+  for (counter in 1:length(x))
+  {
+    mcdf[counter,]<-apply(FUN=sum,MARGIN=2,X=(m<=x[counter]))/length(d)
+  }
+  qcdf<-apply(FUN=quantile,X=mcdf,MARGIN=1,probs=probs)
+  
+  #add the quantiles to the plot
+  for (counter in 1:(dim(qcdf)[1]))
+  {
+    lines(x,qcdf[counter,],type='l',lty='solid',col='grey')
+  }
+
+  #make the second plot
+  dcdfy<-dcdf(x)
+  py<-apply(FUN=sum,X=(matrix(dcdfy,length(dcdfy),dim(mcdf)[2])>mcdf),MARGIN=1)/dim(mcdf)[2]
+  plot(x,py,type='l',xlab='x',ylab='Frac surrog cdfs less than data cdf',ylim=c(0,1))
+  lines(range(x),rep(.025,2),lty='dashed')
[…]

### 2018-02-16 All_analysis_theory.Rmd
Started a theory Rmd and also thought about the effects of tail dependence on extinction risk

---
title: "Exploring the importance of a complete statistical description of dependence in ecology and evolution: Theory"
author: "Shyamolina Ghosh and Daniel Reuman"
date: "February 8, 2018"
output: 
  pdf_document:
    fig_caption: yes
    fig_width: 5
    highlight: tango
    keep_tex: yes
    number_sections: yes
    df_print: kable # data frame printing option
fontsize: 11pt
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
families<-c(1,3:6,14)
source("getcopula.R")
source("OurBiCopSelect.R")
source("copsurrog2d.R")
source("SkewnessAnd3CentMom.R")
source("empcdfcloud.R")
```

# Simulations showing that copula structure matters for the distribution of the spatial mean

# Simulations showing that copula structure matters for Taylor's law

# Simulations exploring whether copula structure matters for extinction risk

We assume a 2-location Lewontin-Cohen model with dispersal,
$$\left[\begin{array}{cc}
p_{1,t+1} \\ 
p_{2,t+1}
\end{array}\right]=
\left[ \begin{array}{cc}
1-d & d \\
d & 1-d
\end{array} \right]
\left[\begin{array}{cc}
\lambda_{1,t}p_{1,t} \\
\lambda_{2,t}p_{2,t}
\end{array}\right],$$
where $\lambda_{i,t}=\exp(\epsilon_{i,t})$ and the 
$\epsilon_t=(\epsilon_{1,t},\epsilon_{2,t})$ are idd draws 
from a bivariate random variable representing the environment at 
the two locations. After each step, if $p_{i,t+1}<1$ it is set to $0$. We will experiment with keeping the Kendall 
correlation and the marginal distributions of $\epsilon$ constant 
but changing its copula structure to see how that affects the 
extinction risk.

```{r echo=F} 
#The simulator of the model described above.
#
#Args
#p0       A length 2 vector holding initial populations
#ns       An numsims by 2 by n array of the epsilons, where
#           n is the number of time steps you want
#d        The dispersal rate
#
#Output
#A numsims by 2 by n+1 array of populations
#
popsim<-function(p0,ns,d)
{
  numsims<-dim(ns)[1]
  n<-dim(ns)[3]

  #initialization
  res<-array(NA,c(numsims,2,n+1))
  res[,,1]<-rep(p0,each=numsims)

  #step the populations forward
  for (tct in 1:n)
  {
    #growth rates based on the noise
    lam1<-exp(ns[,1,tct])
    lam2<-exp(ns[,2,tct])

    #growth prior to dispersal
    res[,1,tct+1]<-lam1*res[,1,tct]
    res[,2,tct+1]<-lam2*res[,2,tct]

    #dispersal
    res[,1,tct+1]<-(1-d)*res[,1,tct+1]+d*res[,2,tct+1]
    res[,2,tct+1]<-(1-d)*res[,2,tct+1]+d*res[,1,tct+1]

    #see if there is extinction
    h<-res[,,tct+1]
    h[h<1]<-0
    res[,,tct+1]<-h
  }

  return(res)
}

extrisk<-function(sims)
{
  totpop<-apply(FUN=sum,X=sims,MARGIN=c(1,3))
  return(apply(FUN=sum,X=(totpop==0),MARGIN=2)/dim(sims)[1])
}
```

First use the extreme tail dependent random number generator in `ExtremeTailDep.R`, so there is extreme right tail dependence in the noise. This means that good years for population growth are prefectly correlated. Extinction risk through time is plotted in blue below. Then consider if the tail dependence is reverse (extreme left tail dependence), so bad years are perfectly correlated (plotted in red below).

```{r echo=F}
source("ExtremeTailDep.R")
numsims<-10000
numsteps<-100
d<-.1
ns1<-retd(n=numsteps*numsims,d=2,rl=1)
ns1<-array(ns1,c(numsteps,numsims,2))
ns1<-aperm(ns1,c(2,3,1))
pops<-popsim(p0=rep(25,2),ns=ns1,d=d)
risk1<-extrisk(pops)

ns2<-(-ns1)
pops<-popsim(p0=rep(25,2),ns=ns2,d=d)
risk2<-extrisk(pops)

plot(0:(length(risk1)-1),risk1,type='l',xlab='Time',ylab='Risk',ylim=c(0,max(risk1,risk2)),col='blue')
lines(0:(length(risk2)-1),risk2,type='l',col='red')
```

Risks are a bit higher for the left-tail dependence. This was for $d=$ `r d`. Try again with $d=0.25$.

```{r echo=F}
d<-.25
pops<-popsim(p0=rep(25,2),ns=ns1,d=d)
risk1<-extrisk(pops)

pops<-popsim(p0=rep(25,2),ns=ns2,d=d)
risk2<-extrisk(pops)

plot(0:(length(risk1)-1),risk1,type='l',xlab='Time',ylab='Risk',ylim=c(0,max(risk1,risk2)),col='blue')
lines(0:(length(risk2)-1),risk2,type='l',col='red')
```

This looks very similar. Try with $d=0$ to see what happens.

```{r echo=F}
d<-0
pops<-popsim(p0=rep(25,2),ns=ns1,d=d)
risk1<-extrisk(pops)

pops<-popsim(p0=rep(25,2),ns=ns2,d=d)
risk2<-extrisk(pops)

plot(0:(length(risk1)-1),risk1,type='l',xlab='Time',ylab='Risk',ylim=c(0,max(risk1,risk2)),col='blue')
lines(0:(length(risk2)-1),risk2,type='l',col='red')
```

Now let's try a model with 20 locations, back to $d=0.1$.

```{r echo=F} 
#The simulator of the multi-location model.
#
#Args
#p0       A length numlocs vector holding initial populations
#ns       An numsims by numlocs by numsteps array of the epsilons, 
#           where numsteps is the number of time steps you want
#d        The dispersal rate (sptially homogeneous)
#
#Output
#A numsims by numlocs by numsteps+1 array of populations
#
popsim_ml<-function(p0,ns,d)
{
  numsims<-dim(ns)[1]
  numlocs<-dim(ns)[2]
  numsteps<-dim(ns)[3]

  #initialization
  res<-array(NA,c(numsims,numlocs,numsteps+1))
  res[,,1]<-rep(p0,each=numsims)

  #step the populations forward
  for (tct in 1:numsteps)
  {
    #growth rates based on the noise
    lam<-exp(ns[,,tct])

    #growth prior to dispersal
    res[,,tct+1]<-lam*res[,,tct]
    res[,,tct+1]<-lam*res[,,tct]

    #dispersal - a fraction d of the total population 
    #disperses evenly everywhere
    totpop<-apply(FUN=sum,X=res[,,tct+1],MARGIN=1)
    totpop<-matrix(rep(totpop,times=numlocs),numsims,numlocs)
    res[,,tct+1]<-(d/numlocs)*totpop+(1-d)*res[,,tct+1]

    #see if there is extinction
    h<-res[,,tct+1]
    h[h<1]<-0
    res[,,tct+1]<-h
  }

  return(res)
}
```

```{r echo=F}
numsims<-10000
numsteps<-100
numlocs<-20
d<-.1
ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1)
ns1<-array(ns1,c(numsteps,numsims,numlocs))
ns1<-aperm(ns1,c(2,3,1))
pops<-popsim_ml(p0=rep(25,numlocs),ns=ns1,d=d)
risk1<-extrisk(pops)

ns2<-(-ns1)
pops<-popsim_ml(p0=rep(25,numlocs),ns=ns2,d=d)
risk2<-extrisk(pops)

plot(0:(length(risk1)-1),risk1,type='l',xlab='Time',ylab='Risk',ylim=c(0,max(risk1,risk2)),col='blue')
lines(0:(length(risk2)-1),risk2,type='l',col='red')
```

Now try with $d=0.25$:

```{r echo=F}
d<-.25
pops<-popsim_ml(p0=rep(25,numlocs),ns=ns1,d=d)
risk1<-extrisk(pops)

pops<-popsim_ml(p0=rep(25,numlocs),ns=ns2,d=d)
risk2<-extrisk(pops)

plot(0:(length(risk1)-1),risk1,type='l',xlab='Time',ylab='Risk',ylim=c(0,max(risk1,risk2)),col='blue')
lines(0:(length(risk2)-1),risk2,type='l',col='red')
```

And with $d=0.05$:

```{r echo=F}
d<-.05
pops<-popsim_ml(p0=rep(25,numlocs),ns=ns1,d=d)
risk1<-extrisk(pops)

pops<-popsim_ml(p0=rep(25,numlocs),ns=ns2,d=d)
risk2<-extrisk(pops)

plot(0:(length(risk1)-1),risk1,type='l',xlab='Time',ylab='Risk',ylim=c(0,max(risk1,risk2)),col='blue')
lines(0:(length(risk2)-1),risk2,type='l',col='red')
```

And with $d=0.01$:

```{r echo=F}
d<-.01
pops<-popsim_ml(p0=rep(25,numlocs),ns=ns1,d=d)
risk1<-extrisk(pops)

pops<-popsim_ml(p0=rep(25,numlocs),ns=ns2,d=d)
risk2<-extrisk(pops)

plot(0:(length(risk1)-1),risk1,type='l',xlab='Time',ylab='Risk',ylim=c(0,max(risk1,risk2)),col='blue')
lines(0:(length(risk2)-1),risk2,type='l',col='red')
```

And with $d=0$:

```{r echo=F}
d<-0
pops<-popsim_ml(p0=rep(25,numlocs),ns=ns1,d=d)
risk1<-extrisk(pops)

pops<-popsim_ml(p0=rep(25,numlocs),ns=ns2,d=d)
risk2<-extrisk(pops)

plot(0:(length(risk1)-1),risk1,type='l',xlab='Time',ylab='Risk',ylim=c(0,max(risk1,risk2)),col='blue')
lines(0:(length(risk2)-1),risk2,type='l',col='red')
```

This preliminary analysis seems to show copula structure can make a noticable difference on extinction risks, even controlling for marginals and correlations.

Recode carefully and systematically to explore dependence of the effect on parameters and model setup assumptions within a reasonable space of possibilities. E.g., consider a range of $d$, perhaps a range of numbers of locations. Can shorten runs to 25 years. Look at time series to make sure nothing crazy/unrealistic is happening. Think it through and figure out how to move this from a preliminary result to a real result. Maybe make dispersal different. Also, why is it dispersal is really making no difference, or not much, the way things are? Is this from a bug? If I decide to change how dispersal works in the multi-population model, one way to do it might be to assume the pops are located along a transect and dispersal only occurs to the one or two (depending on whether you are on the end or not) neighbors.

### 2018-02-16 ExtremeTailDep.R

@@ -1,0 +1,38 @@
+#Generates data with extreme tail dependence. Function name
+#stands for "extreme tail dependence".
+#
+#Args
+#n        Number of points to get
+#d        Dimension
+#rl       1 for right tail dependence, -1 for left tail dependence
+#
+#Output
+#A matrix of dimension n by d, marginals are standard normal.
+#
+retd<-function(n,d,rl)
+{
+  m<-matrix(rnorm(n*(d+1)),n,d+1)
+  flag<-(m[,1]>0) #a coin toss
+  m<-m[,2:(d+1)] #an n by d matrix of iid standard normals
+  
+  m[flag,]<-rep(abs(m[flag,1]),times=d)
+  m[!flag,]<-(-abs(m[!flag,]))
+  
+  if(rl==-1)
+  {
+    m<-(-m)  
+  }
+  
+  return(m)
+}
+
+##test
+#d<-retd(10000,2,1)
+#plot(d[,1],d[,2],type='p')
+#hist(d[,1],50)
+#hist(d[,2],50)
+#
+#d<-retd(10000,2,-1)
+#plot(d[,1],d[,2],type='p')
+#hist(d[,1],50)
+#hist(d[,2],50)

### 2018-02-19 Paper.Rmd
added paper outline in Rmd form

---
title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
author: "Shyamoline Ghosh, other authors, Daniel Reuman"
date: "February 18, 2018"
output: 
  pdf_document:
    fig_caption: yes
    fig_width: 5
    highlight: tango
    keep_tex: yes
    number_sections: yes
    df_print: kable # data frame printing option
fontsize: 11pt
---

# Front matter

*Affiliations:*

Reuman:

Ghosh:

*Possible alternative titles:*

- A complete statistical description of dependence between variables for ecology and related fields

*Keywords:*

# Abstract

# Introduction

*Goes in the last paragraph of the intro:*
We address the following questions. 1) Do datasets in ecology and related fields have non-Gausian copula srtructure? Do they show tail dependence distinct from that of a Gaussian copula, and in particular do they show asymmetric tail dependence? 2) If so, what are some possible causes of non-Gaussian copula structure? And 3) what are the consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?

# Methods

## Datasets

## Statistical methods

Shyamolina, please work on a description of the model-selection approach you have implemented and insert it here. You can just talk about how you go about doing model selection to decide on the best copula of several alternatives (and what those alternatives are) for a bivariate dataset, including the independence test step, the AIC step, and the goodness of fit test for the AIC-best copula. Omit the goodness of fit test for the normal copula, please, since I think we won't be using that.

Shyamolina and Dan to discuss which tail-dependence statistic(s) to use and then we can insert a description here.

Dan to insert a description of the surrogate approach here.

# Results

## Do datasets have non-Gaussian copula structure and tail dependence?

Model selection results:  
- Paragraph on model selection results for the soil C and N data from Terry
- Other stuff, including aphids, depending on what datasets we settle on

Tail dependence results
- again start with C/N data from Terry
- other stuff, inc. aphids

## What are some causes of non-Gaussian copula structure and tail dependence?

## What are the consequences of non-Gaussian copula structure and tail dependence?

# Discussion

### 2018-02-20 Paper.Rmd
filled in method sec : fitting copula

@@
 title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
-author: "Shyamoline Ghosh, other authors, Daniel Reuman"
+author: "Shyamolina Ghosh, other authors, Daniel Reuman"
 date: "February 18, 2018"
@@
 ---
+updated on `r Sys.Date()`
 
@@
 
-Shyamolina, please work on a description of the model-selection approach you have implemented and insert it here.
-You can just talk about how you go about doing model selection to decide on the best copula of several alternatives (and what those alternatives are) for a bivariate dataset, including the independence test step, the AIC step, and the goodness of fit test for the AIC-best copula.
-Omit the goodness of fit test for the normal copula, please, since I think we won't be using that.
+Shyamolina, please work on a description of the model-selection approach 
+you have implemented and insert it here. You can just talk about how you go
+about doing model selection to decide on the best copula of several alternatives
+(and what those alternatives are) for a bivariate dataset, including the 
+independence test step, the AIC step, and the goodness of fit 
+test for the AIC-best copula. Omit the goodness of fit test for 
+the normal copula, please, since I think we won't be using that.
 
+____________________________________________________
+
+Here we intend to study that if for a given bivariate dataset we are able to plot the data within a unit square box based on their rankings then does that data look like totally independent or do they assume a copula structure?
+If they do show some dependence then at best by which family of copula we can describe their dependence?
+To answer these questions, we first choose to pass our bivariate dataset through a p-value based independence-test using *BiCopIndTest* function from *VineCopula* package.
+For that we set the significance level equal to *0.05*.
+If independence is rejected then we perform a test based on maximum likelihood estimation which returns us estimated parameters and log-likelihood, AIC, BIC informations for a specific copula indicated by its familycode.
+We used *BiCopEst* function from *VineCopula* package for families like **N**ormal/Gaussian*{1}*, **C**layton*{3}*, **G**umbel*{4}*, **F**rank*{5}*, **J**oe*{6}*, **BB1***{7}*, **BB6***{8}*, **BB7***{9}*, **BB8***{10}*, **S**urvival **C**layton*{13}*, **S**urvival **G**umbel*{14}*, **S**urvival **J**oe*{16}*, **S**urvival **BB1***{17}*, **S**urvival **BB6***{18}*, **S**urvival **BB7***{19}* and **S**urvival **BB8***{20}*.
+The integer within the curly bracket *`{}`* indicates the familycode of a specific copula.
+We choose a best-fitted copula based on minimum AIC value and to verify that fitted copula-model we perform a goodness-of-fit test of our given data with that best-fitted copula-family by *BiCopGofTest* function from *VineCopula* package.
+For the goodness-of-fit test if any of these p-values (based on CvM or KS method) is less than a threshold *(=0.2)* we used *1000* number of bootstraps and *100* otherwise.
+We also keep the informations like p-values and associated statistics based on Cramer-von Mises test and Kolmogorov-Smirnov test for all next possible fitted copulas in addition to that best-fitted one.
+We choose a collection of possible good fitted copula by seeting a reasonable bar from the minimum AIC values obtained from bestfitted copula model.
+Say, if the AIC of any copula is atmost greater than the AIC value of the best-fitted one by a threshold value *(=2 unit)* then we consider that copula to be in the possible good-fit copula list.
+We can use that same approach for multivariable dataset too.
+A multivariate dataset when considered as pairwise behaves as a bivariate dataset and we can count how many times a specific copula family wins over others as a bestfit or possible good-fit copula model.
+If that data has a certain tail dependence then the winners mostly come from either {**C, SG, SJ, SBB8**} or {**G, SC, J, BB8**} category indicating lower and upper tail dependence respectively.
+
+----------------------------------------------------------------------
 Shyamolina and Dan to discuss which tail-dependence statistic(s) to use and then we can insert a description here.

### 2018-02-21 SupportingInformation.Rmd
mostly changes to implement the bivariate models selection on Terry's C/N data, and to randomize copula structure and see what it does to the distribution of some ecosystem function
a few changes to the structure of supporting information so it can show all the analyses we performed in the same structure as the paper outline

@@
 
-# Testing the ranking and bivariate copula display function
+# Question 1 analyses for each dataset
 
-This is just for testing the `getcopula` function.
+# Question 2 analyses for each dataset
+
+# Question 3 analyses for each dataset
+
+# Testing
+
+## Testing the ranking and bivariate copula display function
+
+This is for testing the `getcopula` function.
 
@@
 
-# Testing the bivariate model selection and goodness-of-fit pipeline
+## Testing the bivariate model selection and goodness-of-fit pipeline
 
-This is just for testing the `OurBiCopSelect` function. We test that
+This is for testing the `OurBiCopSelect` function. We test that
 it runs and gives reasonable results by simulating data and seeing if 
@@
 be rejected, Clayton should be the AIC best, it should pass the
-goodness of fit test, and the normal copula should fail the
-goodness of fit test:
+goodness of fit test, the normal copula should fail the
+goodness of fit test, and model-averaged tail dependence should
+indicate left-tail dependence:
 <!--Test for real data from a clayton-->
@@
 dat<-rCopula(numdat,cop)
+
+#do the analysis
 allresC<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
@@
 allresC
+
+#calculate model-averaged left- and right-tail dependence
+#DAN: ad this later
 ```
@@
 dat<-rCopula(numdat,cop)
+
+#do the analysis
 allresG<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
@@
 allresG
+
+#calculate model-averaged left- and right-tail dependence
+#DAN: ad this later
 ```
@@
 dat<-rCopula(numdat,cop)
+
+#do the analysis
 allresF<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
@@
 allresF
+
+#calculate model-averaged left- and right-tail dependence
+#DAN: ad this later
 ```
@@
 dat<-rCopula(numdat,cop)
+
+#do the analysis
 allresJ<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
@@
 allresJ
+
+#calculate model-averaged left- and right-tail dependence
+#DAN: ad this later
 ```
@@
 a comparison of A/BICs strongly supports the true copula over the normal.
-So I think the best way forward is to use `gofnormal=FALSE`.
+So proceed using `gofnormal=FALSE` when analyzing real data.
 
-# Testing the pipeline for examining tail dependence in bivariate data
+# Testing the pipeline for examining tail dependence in bivariate data using `copsurrog2d`

### 2018-02-22 SupportingInformation.Rmd
planning what to do in supp mat with aphid and plankton and C/N for model selection analysis

@@
 # Descriptions of data
-Each data contributor writes one to a few paragraphs about their data.
+<!--Each data contributor writes one to a few paragraphs about their data.-->
+
+## Soil C and N data
+
+## Aphid data
+<!--Shya to add some facts, make sure to name the species-->
+We threw out all locations that had fewer than 30 years data (Shya to make more precise, also add any other cleaning info you did, in summary/bullet form)
+
+## North Sea plankton data
+Make sure to name the species used
 
 # Question 1 analyses for each dataset
+
+## Soil C and N data
+<!--Shya to add model selection analysis of Terry data here. You just have to move it over from All_analysis.Rmd and edit the code. in a few ways:
+1) make the code get displayed at each step
+2) you have added some columns to the InfCritRes table so that will come out differently
+3) you can also report your result for tail dependence based on model averaging
+4) stop after that (e.g., don't put in the stuff about the outcome function, etc.-->
+
+## Aphid data, count
+<!--Shya to add model selection analysis here for aphid data, first the species for which you are using count data, then the one for ff data, just the model selection stuff, we'll do the other stuff with the tail dependence asymmetry statistics later-->
+
+## Aphid data, first flight
+<!--Shya also add as above-->
+
+## North sea plankton data
+<!--Shya also add as above-->
 

### 2018-02-23 SupportingInformation.Rmd
adding aphid and plankton north sea results for fitting

--- 2018-02-14 SupportingInformation.Rmd
@@
 
-# Descriptions of data
-Each data contributor writes one to a few paragraphs about their data.
-
-# Testing the ranking and bivariate copula display function
-
-This is just for testing the `getcopula` function.
-
-<!--Source in function and a package--->
-```{r}
+```{r setup, echo=F}
+knitr::opts_chunk$set(echo = TRUE)
+seed<-101
+families<-c(1,3:10,13,14,16:20)
 source("getcopula.R")
-require(copula)
-require(mvtnorm)
+source("OurBiCopSelect.R")
+source("FittingCopula_selective_loc.R")
 ```
 
-Plot data from a Clayton copula, which has lower tail dependence:
-```{r}
-copC<-claytonCopula(8,2)
-dC<-rCopula(1000,copC)
-vC<-getcopula(d=dC,rankon=F,ploton=T)
+# Descriptions of data
+<!--Each data contributor writes one to a few paragraphs about their data.-->
+
+## Soil C and N data
+
+## Aphid data
+<!--Shya to add some facts, make sure to name the species-->
+   * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
+   * We threw out all locations that had fewer than 30 years data 
+   * We use species=10 [Name : "Green spruce aphid"] for aphid-count data 
+   * We use species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data
+
+## North Sea plankton data
+
+   * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
+   * We threw out all locations that had fewer than 45 years data
+   * We use species=16 [Name : "Ceratium_furca"] for Plankton_North_Sea data
+
+# Question 1 analyses for each dataset
+
+## Soil C and N data
+<!--Shya to add model selection analysis of Terry data here. You just have to move it over from All_analysis.Rmd and edit the code. in a few ways:
+1) make the code get displayed at each step
+2) you have added some columns to the InfCritRes table so that will come out differently
+3) you can also report your result for tail dependence based on model averaging
+4) stop after that (e.g., don't put in the stuff about the outcome function, etc.-->
+**About the dataset :**  
+Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description).
+
+**Plot the raw data:**
+
+```{r load_data_Loecke, echo=F}
+#Load your data - should be a data frame with 2 columns. Can contain NAs.
+d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
+d<-d[,c("SOCstock100","TSNstock100")]
+plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1)
+rect(0,0,1,1)
 ```
 
-Plot data from a Joe copula, which has upper tail dependence:
-```{r}
-copJ<-joeCopula(8,2)
-dJ<-rCopula(1000,copJ)
-vJ<-getcopula(d=dJ,rankon=F,ploton=T)
+**Copula plot of the data:**  
+```{r rankandplot_Loecke, echo=F}
+#convert to a copula and plot
+v<-getcopula(d=d,rankon=T,ploton=T) #with real data, rankon should be T
 ```
 
-Get data from a bivariate normal distribution and take ranks and
-then plot:
-```{r}
-sigma<-matrix(c(1,.7,.7,1),2,2)
-dN<-rmvnorm(1000,mean=c(0,0),sigma=sigma)
-vN<-getcopula(d=dN,rankon=T,ploton=T)
+**Bivariate copula model selection and goodness of fit pipeline:**  
+```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v)}
+set.seed(seed)
+families_Loecke<-c(1,3:10,13:14,16:20)
+BivMS_res_Loecke<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=families_Loecke,level=0.05,AICBIC="AIC",
+                         numBSsmall=100,pthresh=0.2,numBSlarge=1000,
+                         gofnormal=F,status=T)
 ```
 
-# Testing the bivariate model selection and goodness-of-fit pipeline
-
-This is just for testing the `OurBiCopSelect` function. We test that
-it runs and gives reasonable results by simulating data and seeing if 
-you recover what kind of copula was used.
-
-<!--Source in the function and a package, changeable constants-->
 ```{r}
-source("OurBiCopSelect.R")
-require(copula)
-nbsmall<-100
-pthresh<-0.2
-nblarge<-1000 #for bootstrapping, see below
-numdat<-500 #amount of artificial data
-seed<-101
+BivMS_res_Loecke
+BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw # -ve value indicates UT dep?
+#or we should discard this result because none of the chosen model is a good fit for this dataset?
 ```
 
-If real data come from a Clayton copula, independence should
-be rejected, Clayton should be the AIC best, it should pass the
-goodness of fit test, and the normal copula should fail the
-goodness of fit test:
-<!--Test for real data from a clayton-->
-```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
-set.seed(seed)
+## Aphid data
+<!--Shya to add model selection analysis here for aphid data, first the species for which you are using count data, then the one for ff data, just the model selection stuff, we'll do the other stuff with the tail dependence asymmetry statistics later-->
+**Bivariate copula model selection and goodness of fit pipeline:** 
+```{r vivj_matrix, echo=F}
+# Select any two location pair [i,j] for a copula of species sp
+# This function gives you a matrix with vi and vj as two columns
+vivj_matrix<-function(d_allsp,sp,i,j){
 
-families<-c(1,3:6,13,14,16)
+  ds1<-d_allsp[[sp]][[i]]
+  ds2<-d_allsp[[sp]][[j]]
+  #----------------------------
+  a1<-ds1$Year[1]
+  a2<-ds2$Year[1]
+  a3<-ds1$Year[dim(ds1)[1]]
+  a4<-ds2$Year[dim(ds2)[1]]
+  year_s<-max(a1,a2)
+  year_e<-min(a3,a4)
+  ind_s1<-which(ds1$Year==year_s)
+  ind_s2<-which(ds2$Year==year_s)
+  ind_e1<-which(ds1$Year==year_e)
+  ind_e2<-which(ds2$Year==year_e)
+  ds1<-ds1[ind_s1:ind_e1,]
+  ds2<-ds2[ind_s2:ind_e2,]
+  # Omitting the years and data containing NA in either d1 or d2 
+  #from both d1 and d2
+  if(anyNA(ds1$Dat)==T | anyNA(ds2$Dat)==T){
+    ind_na1<-which(is.na(ds1$Dat))
+    ind_na2<-c(ind_na1,which(is.na(ds2$Dat)))
+    ind_na<-unique(ind_na2)
 
-#data are really from a clayton
-cop<-claytonCopula(5,2)
-dat<-rCopula(numdat,cop)
-allresC<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
-                        numBSlarge=nblarge,pthresh=0.2)
-allresC
+    d1Dat<-ds1$Dat[-ind_na]
+    d2Dat<-ds2$Dat[-ind_na]
+    Years<-ds1$Year[-ind_na]
+    d1<-data.frame(Year=Years,Dat=d1Dat)
+    d2<-data.frame(Year=Years,Dat=d2Dat)
+  } else {
+    d1<-ds1
+    d2<-ds2
+  }
+  #get ranks modified now
+  vi<-pobs(d1$Dat)
+  vj<-pobs(d2$Dat)
+  Years<-d1$Year
+  #-------------------------
+  #n_datapt<-length(vi)
+  #--------------------
+  #plot(vi,vj,type="p")
+  #-------------------------
+  mat<-as.matrix(cbind(vi,vj))
+  return(mat)  
+}
 ```
 
-Same test, but data now come from a Gumbel copula:
-<!--Test for real data from a gumbel-->
-```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
-set.seed(seed)
+```{r d_allsp_data_aphid, echo=F}
+d_allsp_data<-function(d0){
 
-#data are really from a gumbel
-cop<-gumbelCopula(6,2)
-dat<-rCopula(numdat,cop)
-allresG<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
-                        numBSlarge=nblarge,pthresh=0.2)
-allresG
+  name.aphid<-c("Apple-grass aphid","Bird cherry-oat aphid","Black bean aphid", "Blackberry-cereal aphid","Blackcurrant-sowthistle aphid",
+                "Corn leaf aphid","Currant-lettuce aphid","Damson-hop aphid","Grain aphid","Green spruce aphid",
+                "Leaf-curling plum aphid","Mealy cabbage aphid","Mealy plum aphid","Pea aphid","Peach-potato aphid",
+                "Potato aphid","Rose-grain aphid","Shallot aphid","Sycamore aphid","Willow-carrot aphid")
+  longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
+  lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)
+  loc_name<-c("Ayr","Broom's Barn","Dundee","Edinburgh","Hereford","Newcastle","Preston","Rothamsted",
+              "Starcross","Writtle","Wye")
+  d0[d0==-999]<-NA
+  Years<-1976:2010
+  d_allsp<-list() #this is a list with 20 elements, all the species
+  for (sp_c in 1:20)
+  {
+    d_thissp<-list() #a list with 11 locations for the current species
+    for (st_c in 1:11)
+    {
+      d_thissp[[st_c]]<-data.frame(Year=Years,Dat=d0[st_c,((sp_c-1)*35+1):(sp_c*35)])
+    }
+    d_allsp[[sp_c]]<-d_thissp
+  }
+  names(d_allsp)<-name.aphid
+
+  return(d_allsp=d_allsp)
+
+}
+
+# Aphid raw data for count
+d0_count<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
+
+# Aphid raw data for first flight
+d0_firstflight<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtsfirstflight141117.csv",header=F))
+
+# Aphid raw data for flight duration
+d0_flightduration<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))
+
 ```
 
-Same test, but data now come from a Frank copula:
-<!--Test for real data from a frank-->
-```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
+## Aphid data, count
+
+```{r aphid_count, echo=F, cache=T, cache.extra=list(seed,d0_count)}
 set.seed(seed)
-
-#data are really from a frank
-cop<-frankCopula(5,2)
-dat<-rCopula(numdat,cop)
-allresF<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
-                        numBSlarge=nblarge,pthresh=0.2)
-allresF
+cat(paste("start-time: ",Sys.time(),"\n"))
+RES_aphid_count<-RES_single_sp(d_allsp=d_allsp_data(d0_count),sp=10,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
+cat(paste("stop-time: ",Sys.time(),"\n"))
+resloc<-"Results/"
+saveRDS(RES_aphid_count,paste(resloc,file="AphidCopulaFit_selecloc_count_species_10.RDS",sep=""))
 ```
 
-Same test, but data now come from a Joe copula:
-<!--Test for real data from a joe-->
-```{r cache=T, cache.extra=list(nbsmall,nblarge,pthresh,numdat,seed,mtime("OurBiCopSelect.R"))}
-set.seed(seed)
+```{r}
+RES_aphid_count
+#no of locations showing greater lower tail dependence
+sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)
 
-#data are really from a joe
-cop<-joeCopula(6,2)
-dat<-rCopula(numdat,cop)
-allresJ<-OurBiCopSelect(dat[,1],dat[,2],families,numBSsmall=nbsmall,
-                        numBSlarge=nblarge,pthresh=0.2)
-allresJ
+#no of locations showing greater upper tail dependence
+sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)<0,na.rm=T)
+
 ```
 
-The goodness of fit test is quite slow when testing the fit of a
-normal copula, and the results are not great: often the normal copula
-cannot be rejected even when it is not the true copula, even though
-a comparison of A/BICs strongly supports the true copula over the normal.
-So I think the best way forward is to use `gofnormal=FALSE`.
+## Aphid data, first flight
 
-# Testing the pipeline for examining tail dependence in bivariate data
+```{r aphid_firstflight, echo=F, cache=T, cache.extra=list(seed,d0_firstflight)}
+set.seed(seed)
+cat(paste("start-time: ",Sys.time(),"\n"))
+RES_aphid_firstflight<-RES_single_sp(d_allsp=d_allsp_data(d0_firstflight),sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
+cat(paste("stop-time: ",Sys.time(),"\n"))
+resloc<-"Results/"
+saveRDS(RES_aphid_firstflight,paste(resloc,file="AphidCopulaFit_selecloc_firstflight_species_11.RDS",sep=""))
+```
+
+```{r}
+RES_aphid_firstflight
+#no of locations showing greater lower tail dependence
+sum((RES_aphid_firstflight$LTdep_AICw-RES_aphid_firstflight$UTdep_AICw)>0,na.rm=T)
+
+#no of locations showing greater upper tail dependence
+sum((RES_aphid_firstflight$LTdep_AICw-RES_aphid_firstflight$UTdep_AICw)<0,na.rm=T)
+
+```
+
+## North sea plankton data
+<!--Shya also add as above-->
+```{r d_allsp_data_plankton_north_sea, echo=F}
+d_allsp_data_plankton_north_sea<-function(d0){
+
+  name.plankton<-c('Calanus_I-IV','Para-Pseudocalanus_spp','Temora_longicornis','Acartia_spp_(unidentified)','Centropages_typicus',
+                   'Oithona_spp','Echinoderm_larvae','Calanus_finmarchicus','Calanus_helgolandicus','Metridia_lucens','Decapoda_larvae_(Total)',
+                   'Euphausiacea_Total','Thalassiosira_spp','Rhizosolenia_styliformis','Ceratium_fusus','Ceratium_furca','Ceratium_tripos',
+                   'Ceratium_macroceros','Rhizosolenia_alata_alata','Pseudocalanus_elongatus_Adult','Nitzschia_delicatissima','Nitzschia_seriata')
+
+  longs<-read.csv("Data/Plankton_North_Sea_data/boxcornerlongs201117.csv",header=F)
+  lats<-read.csv("Data/Plankton_North_Sea_data/boxcornerlats201117.csv",header=F)
+  longs<-longs[[1]]
+  lats<-lats[[1]]
+  d0[d0==-999]<-NA
+  Years<-1958:2013
+
+  d_allsp<-list() #this is a list with 20 elements, all the species
+  for (sp_c in 1:22)
+  {
+    d_thissp<-list() #a list with 11 locations for the current species
+    d1<-d0[((sp_c-1)*26+1):(sp_c*26),]
+    for (st_c in 1:26)
+    {
+      d_thissp[[st_c]]<-data.frame(Year=Years,Dat=d1[st_c,])
+    }
+    d_allsp[[sp_c]]<-d_thissp
+  }
+  names(d_allsp)<-name.plankton
+
+  return(d_allsp=d_allsp)
+
+}
+
+d0_plankton_north_sea<-as.matrix(read.csv("Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))
+
+```
+
+```{r plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,d0_plankton_north_sea)}
+set.seed(seed)
+cat(paste("start-time: ",Sys.time(),"\n"))
+RES_plankton_north_sea<-RES_single_sp(d_allsp=d_allsp_data_plankton_north_sea(d0_plankton_north_sea),sp=16,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=45)
+cat(paste("stop-time: ",Sys.time(),"\n"))
+resloc<-"Results/"
+saveRDS(RES_plankton_north_sea,paste(resloc,file="Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS",sep=""))
+```
+
+```{r}
+RES_plankton_north_sea
+#no of locations showing greater lower tail dependence
+sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)
+
+#no of locations showing greater upper tail dependence
+sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)<0,na.rm=T)
+
+```
+
+# Question 2 analyses for each dataset
+
+# Question 3 analyses for each dataset

### 2018-02-23 SupportingInformation.Rmd
organizational changes

@@
 date: "February 12, 2018"
-output: pdf_document
+output: 
+  pdf_document:
+    number_sections: yes
+    keep_tex: yes
 ---
@@
 
+<!--Setup-->
 ```{r setup, echo=F}
@@
 ## Soil C and N data
-
-## Aphid data
-<!--Shya to add some facts, make sure to name the species-->
-   * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
-   * We threw out all locations that had fewer than 30 years data 
-   * We use species=10 [Name : "Green spruce aphid"] for aphid-count data 
-   * We use species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data
-
-## North Sea plankton data
-
-   * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
-   * We threw out all locations that had fewer than 45 years data
-   * We use species=16 [Name : "Ceratium_furca"] for Plankton_North_Sea data
-
-# Question 1 analyses for each dataset
-
-## Soil C and N data
-<!--Shya to add model selection analysis of Terry data here. You just have to move it over from All_analysis.Rmd and edit the code. in a few ways:
-1) make the code get displayed at each step
-2) you have added some columns to the InfCritRes table so that will come out differently
-3) you can also report your result for tail dependence based on model averaging
-4) stop after that (e.g., don't put in the stuff about the outcome function, etc.-->
-**About the dataset :**  
 Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description).
 
-**Plot the raw data:**
-
+Plot the raw data:
 ```{r load_data_Loecke, echo=F}
@@
 
-**Copula plot of the data:**  
+Copula plot of the data:
 ```{r rankandplot_Loecke, echo=F}
@@
 
-**Bivariate copula model selection and goodness of fit pipeline:**  
+## Aphid data
+   * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
+   * We threw out all locations that had fewer than 30 years data 
+   * We use species=10 [Name : "Green spruce aphid"] for aphid-count data 
+   * We use species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data
+
+## North Sea plankton data
+   * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
+   * We threw out all locations that had fewer than 45 years data
+   * We use species=16 [Name : "Ceratium_furca"] for Plankton_North_Sea data
+
+# Question 1 analyses for each dataset
+
+Recall that question 1 is: Do datasets in ecology and related fields have non-Gausian copula srtructure?
+Do they show tail dependence distinct from that of a Gaussian copula?
+In particular do they show asymmetric tail dependence?
+
+## Details of methods
+
+### Model selection approach
+
+### Non-paramtric analysis of tail dependence
+
+## Soil C and N data
+
+### Model selection approach
+
 ```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v)}
@@
 families_Loecke<-c(1,3:10,13:14,16:20)
-BivMS_res_Loecke<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=families_Loecke,level=0.05,AICBIC="AIC",
-                         numBSsmall=100,pthresh=0.2,numBSlarge=1000,
-                         gofnormal=F,status=T)
+BivMS_res_Loecke<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
 ```
 
-```{r}
-BivMS_res_Loecke
-BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw # -ve value indicates UT dep?
-#or we should discard this result because none of the chosen model is a good fit for this dataset?
+The test of indepedence gave $p=$`r BivMS_res_Loecke$IndepTestRes`. We fitted several copulas
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas:
+
+```{r echo=F}
+h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+rownames(h)<-NULL
+h
 ```
 
+This shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+
+The model-averaged (using AIC) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged (again, AIC) upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.
+
+Although these results do convincingly show non-Gaussian copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+The next section looks at tail dependence in a non-parametric way.
+
+### Non-paramtric analysis of tail dependence
+
+<!--Shya could work on this. Select 1-2 statistics, maybe more, but try to cut some of them (looks like there will be 4). Apply to C/N data, get results, display in the same kind of way.-->
+
 ## Aphid data
-<!--Shya to add model selection analysis here for aphid data, first the species for which you are using count data, then the one for ff data, just the model selection stuff, we'll do the other stuff with the tail dependence asymmetry statistics later-->
-**Bivariate copula model selection and goodness of fit pipeline:** 
+<!--Shya to edit the prentation below to make it more parallel to the presentation of the model selection results for C/N data above-->
+
+### Model selection approach
+
 ```{r vivj_matrix, echo=F}
@@
 
-## Aphid data, count
+Count results
 
@@
 
-## Aphid data, first flight
+First flight results
 
@@
 
+### Non-paramtric analysis of tail dependence
+
+Cout results
+
+First flight results
+
 ## North sea plankton data
-<!--Shya also add as above-->
+
+<!--Shya to edit the prentation below to make it more parallel to the presentation of the model selection results for C/N data above-->
+
+### Model selection approach
+
 ```{r d_allsp_data_plankton_north_sea, echo=F}
@@
 
+### Non-paramtric analysis of tail dependence
+
 # Question 2 analyses for each dataset

### 2018-02-23 OurBiCopSelect.R

@@ -99,16 +99,16 @@ OurBiCopSelect<-function(u1,u2,families,level=0.05,AICBIC="AIC",
       InfCritRes$BICw[counter]<-exp(-0.5*(InfCritRes$BIC[counter]-min(InfCritRes$BIC)))
     }
 
-      InfCritRes$AICw<-InfCritRes$AICw/sum(InfCritRes$AICw)
-      InfCritRes$BICw<-InfCritRes$BICw/sum(InfCritRes$BICw)
+    InfCritRes$AICw<-InfCritRes$AICw/sum(InfCritRes$AICw)
+    InfCritRes$BICw<-InfCritRes$BICw/sum(InfCritRes$BICw)
 
-      # check : sum(InfCritRes$AICw)=1, sum(InfCritRes$BICw)=1
+    # check : sum(InfCritRes$AICw)=1, sum(InfCritRes$BICw)=1
       
-      relLTdep_AICw<-sum((InfCritRes$LTdep*InfCritRes$AICw)/sum(InfCritRes$AICw))
-      relUTdep_AICw<-sum((InfCritRes$UTdep*InfCritRes$AICw)/sum(InfCritRes$AICw))
-      
-      relLTdep_BICw<-sum((InfCritRes$LTdep*InfCritRes$BICw)/sum(InfCritRes$BICw))
-      relUTdep_BICw<-sum((InfCritRes$UTdep*InfCritRes$BICw)/sum(InfCritRes$BICw))
+    relLTdep_AICw<-sum((InfCritRes$LTdep*InfCritRes$AICw)/sum(InfCritRes$AICw))
+    relUTdep_AICw<-sum((InfCritRes$UTdep*InfCritRes$AICw)/sum(InfCritRes$AICw))
+    
+    relLTdep_BICw<-sum((InfCritRes$LTdep*InfCritRes$BICw)/sum(InfCritRes$BICw))
+    relUTdep_BICw<-sum((InfCritRes$UTdep*InfCritRes$BICw)/sum(InfCritRes$BICw))
     
     
     if (status) {cat(paste("Done: ",Sys.time(),"\n"))}

### 2018-02-24 SupportingInformation.Rmd
minor tweak

@@
 # Aphid raw data for count
-d0_count<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
+d0_count<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
 
 # Aphid raw data for first flight
-d0_firstflight<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtsfirstflight141117.csv",header=F))
+d0_firstflight<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsfirstflight141117.csv",header=F))
 
-# Aphid raw data for flight duration
-d0_flightduration<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))
+## Aphid raw data for flight duration
+#d0_flightduration<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtsflightdu#ration141117.csv",header=F))
 

### 2018-02-26 SupportingInformation.Rmd
trivial changes
fill in nonparametric stat section

--- 2018-02-23 SupportingInformation.Rmd
@@
+---
+output:
+  pdf_document: default
+---
 ---
@@
 date: "February 12, 2018"
+geometry: "left=1cm,right=1cm,top=2.5cm,bottom=2.8cm"
+
 output: 
@@
     keep_tex: yes
+    fig_caption: yes
 ---
 updated on `r Sys.Date()`
+
+```{r ,echo=FALSE}
+library(captioner)
+table_nums <- captioner(prefix = "Table")
+fig_nums <- captioner(prefix="Figure")
+```
 
@@
 ```{r setup, echo=F}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
 seed<-101
@@
 source("FittingCopula_selective_loc.R")
+source("aphid_data_calling.R")
+source("plankton_north_sea_data_calling.R")
+source("NonParamStat.R")
 ```
@@
 
-Plot the raw data:
-```{r load_data_Loecke, echo=F}
+```{r load_data_Loecke, echo=F, fig.cap="plot of the raw data", fig.height=5, fig.width=5}
 #Load your data - should be a data frame with 2 columns. Can contain NAs.
@@
 d<-d[,c("SOCstock100","TSNstock100")]
+fig_nums("soilCNraw",display=F)
 plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1)
@@
 
-Copula plot of the data:
-```{r rankandplot_Loecke, echo=F}
+```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C/N data",  fig.height=5, fig.width=5}
 #convert to a copula and plot
+fig_nums("soilCNcop",display=F)
 v<-getcopula(d=d,rankon=T,ploton=T) #with real data, rankon should be T
@@
 
-The test of indepedence gave $p=$`r BivMS_res_Loecke$IndepTestRes`. We fitted several copulas
+The test of indepedence gave *p=*`r BivMS_res_Loecke$IndepTestRes` . We fitted several copulas
 and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas:
 
-```{r echo=F}
+```{r soilCNfit,echo=F}
 h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-rownames(h)<-NULL
-h
+#rownames(h)<-NULL
+table_nums("soilCNfit",display=F)
+knitr::kable(h, caption = "Fitting info for soilC/N dataset")
 ```
 
-This shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+`r table_nums("soilCNfit",display="c")` shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
 
@@
 
+```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v)}
+set.seed(seed) 
+resloc<-"./Results/stat_results/stat_soilCN/"
+
+stat_soilCN<-copsync(v)
+saveRDS(stat_soilCN,paste(resloc,file="stat_soilCN.RDS",sep=''))
+sCN<-stat_soilCN[c(2:5,7,8,10:13)]
+stat<-unlist(sCN)
+```
+
+```{r table_stat_soilCN, echo=F, results="asis"}
+table_nums("table_stat_soilCN",display=F)
+
+knitr::kable(data.frame(stat=stat),caption = "Non-parametric stats on soil C/N dataset")
+```
+
+```{r ,echo=F}
+#sCN
+CorlmCoru<-stat_soilCN$Corl-stat_soilCN$Coru
+cat("Corl-Coru = ",CorlmCoru,"\n")
+PlmPu<-stat_soilCN$Pl-stat_soilCN$Pu
+cat("Pl-Pu = ",PlmPu,"\n")
+TlmTu<-stat_soilCN$Tl-stat_soilCN$Tu
+cat("Tl-Tu = ",TlmTu,"\n")
+D2umD2l<-stat_soilCN$D2u-stat_soilCN$D2l
+cat("D2u-D2l = ",D2umD2l,"\n")
+```
+
 <!--Shya could work on this. Select 1-2 statistics, maybe more, but try to cut some of them (looks like there will be 4). Apply to C/N data, get results, display in the same kind of way.-->
@@
 
-```{r vivj_matrix, echo=F}
-# Select any two location pair [i,j] for a copula of species sp
-# This function gives you a matrix with vi and vj as two columns
-vivj_matrix<-function(d_allsp,sp,i,j){
-
-  ds1<-d_allsp[[sp]][[i]]
-  ds2<-d_allsp[[sp]][[j]]
-  #----------------------------
-  a1<-ds1$Year[1]
-  a2<-ds2$Year[1]
-  a3<-ds1$Year[dim(ds1)[1]]
-  a4<-ds2$Year[dim(ds2)[1]]
-  year_s<-max(a1,a2)
-  year_e<-min(a3,a4)
-  ind_s1<-which(ds1$Year==year_s)
-  ind_s2<-which(ds2$Year==year_s)
-  ind_e1<-which(ds1$Year==year_e)
-  ind_e2<-which(ds2$Year==year_e)
-  ds1<-ds1[ind_s1:ind_e1,]
-  ds2<-ds2[ind_s2:ind_e2,]
-  # Omitting the years and data containing NA in either d1 or d2 
-  #from both d1 and d2
-  if(anyNA(ds1$Dat)==T | anyNA(ds2$Dat)==T){
-    ind_na1<-which(is.na(ds1$Dat))
-    ind_na2<-c(ind_na1,which(is.na(ds2$Dat)))
-    ind_na<-unique(ind_na2)
-
-    d1Dat<-ds1$Dat[-ind_na]
-    d2Dat<-ds2$Dat[-ind_na]
-    Years<-ds1$Year[-ind_na]
-    d1<-data.frame(Year=Years,Dat=d1Dat)
-    d2<-data.frame(Year=Years,Dat=d2Dat)
-  } else {
-    d1<-ds1
-    d2<-ds2
-  }
-  #get ranks modified now
-  vi<-pobs(d1$Dat)
-  vj<-pobs(d2$Dat)
-  Years<-d1$Year
-  #-------------------------
-  #n_datapt<-length(vi)
-  #--------------------
-  #plot(vi,vj,type="p")
-  #-------------------------
-  mat<-as.matrix(cbind(vi,vj))
-  return(mat)  
-}
-```
-
-```{r d_allsp_data_aphid, echo=F}
-d_allsp_data<-function(d0){
-
-  name.aphid<-c("Apple-grass aphid","Bird cherry-oat aphid","Black bean aphid", "Blackberry-cereal aphid","Blackcurrant-sowthistle aphid",
-                "Corn leaf aphid","Currant-lettuce aphid","Damson-hop aphid","Grain aphid","Green spruce aphid",
-                "Leaf-curling plum aphid","Mealy cabbage aphid","Mealy plum aphid","Pea aphid","Peach-potato aphid",
-                "Potato aphid","Rose-grain aphid","Shallot aphid","Sycamore aphid","Willow-carrot aphid")
-  longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
-  lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)
-  loc_name<-c("Ayr","Broom's Barn","Dundee","Edinburgh","Hereford","Newcastle","Preston","Rothamsted",
-              "Starcross","Writtle","Wye")
-  d0[d0==-999]<-NA
-  Years<-1976:2010
-  d_allsp<-list() #this is a list with 20 elements, all the species
-  for (sp_c in 1:20)
-  {
-    d_thissp<-list() #a list with 11 locations for the current species
-    for (st_c in 1:11)
-    {
-      d_thissp[[st_c]]<-data.frame(Year=Years,Dat=d0[st_c,((sp_c-1)*35+1):(sp_c*35)])
-    }
-    d_allsp[[sp_c]]<-d_thissp
-  }
-  names(d_allsp)<-name.aphid
-
-  return(d_allsp=d_allsp)
-
-}
-
+```{r read_aphid_data, echo=F}
 # Aphid raw data for count
-d0_count<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
+d0_count<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
+data_aphid_count<-d_allsp_data(d0_count)
 
 # Aphid raw data for first flight
-d0_firstflight<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtsfirstflight141117.csv",header=F))
+d0_firstflight<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsfirstflight141117.csv",header=F))
+data_aphid_ff<-d_allsp_data(d0_firstflight)
 
 # Aphid raw data for flight duration   
-d0_flightduration<-as.matrix(read.csv("Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))
+#d0_flightduration<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))
 
@@
 
-```{r aphid_count, echo=F, cache=T, cache.extra=list(seed,d0_count)}
+```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count)}
 set.seed(seed)
-cat(paste("start-time: ",Sys.time(),"\n"))
-RES_aphid_count<-RES_single_sp(d_allsp=d_allsp_data(d0_count),sp=10,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
-cat(paste("stop-time: ",Sys.time(),"\n"))
-resloc<-"Results/"
+#cat(paste("start-time: ",Sys.time(),"\n")) 
+RES_aphid_count<-RES_single_sp(d_allsp=data_aphid_count,sp=10,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
+#cat(paste("stop-time: ",Sys.time(),"\n"))
+resloc<-"./Results/fitting_results/"
 saveRDS(RES_aphid_count,paste(resloc,file="AphidCopulaFit_selecloc_count_species_10.RDS",sep=""))
@@
 
-```{r}
-RES_aphid_count
-#no of locations showing greater lower tail dependence
-sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)
+```{r read_RES_aphid_count, echo=F}
+RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
+```
 
-#no of locations showing greater upper tail dependence
-sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)<0,na.rm=T)
+```{r count_bestcop,echo=F,results="asis"}
+#library(pander)
+#pander(best_cop_LTdep, caption = "Hello caption!")
+#knitr::kable(RES_aphid_count$info_ord_copname[,,1],format="latex",longtable=F,caption = "Aphid_count : Best fit copula family for each selective location pair")
+
+table_nums("count_bestcop",display=F)
+knitr::kable(RES_aphid_count$info_ord_copname[,,1],caption = "Aphid_count : Best fit copula family for each selective location pair")
+```
+
+```{r count_AICw,echo=F,results="asis"}
+
+cat("\n\n\\pagebreak\n")
+
+table_nums("count_AICw",display=F)
+knitr::kable(RES_aphid_count$info_ord_AICw[,,1], caption = "Aphid_count : AICw for best fit copula family for each selective location pair")
+```
+
+```{r count_LTdep,echo=F,results="asis"}
+table_nums("count_LTdep",display=F)
+knitr::kable(RES_aphid_count$info_ord_LTdep[,,1], caption = "Aphid_count : Lower Tail dependence for best fit copula family for each selective location pair")
+```
+
+```{r count_UTdep,echo=F,results="asis"}
+table_nums("count_UTdep",display=F)
+knitr::kable(RES_aphid_count$info_ord_UTdep[,,1], caption = "Aphid_count : Upper Tail dependence for best fit copula family for each selective location pair")
+```
+
+```{r count_LTdep_AICw,echo=F,results="asis"}
+table_nums("count_LTdep_AICw",display=F)
+knitr::kable(RES_aphid_count$LTdep_AICw,caption = "Aphid_count : Lower Tail dependence statistic for best fit copula family for each selective location pair")
+```
+
+```{r count_UTdep_AICw,echo=F,results="asis"}
+table_nums("count_UTdep_AICw",display=F)
+knitr::kable(RES_aphid_count$UTdep_AICw,caption = "Aphid_count : Upper Tail dependence statistic for best fit copula family for each selective location pair")
 
@@
 
+```{r count_LTmUT,echo=F,results="asis"}
+table_nums("count_LTmUT",display=F)
+knitr::kable(round(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,5),caption = "Aphid_count : Lower Tail - Upper Tail dependence statistic for best fit copula family for each of the selective location pair")
+```
+
+```{r count_p_CvM,echo=F,results="asis"}
+table_nums("count_p_CvM",display=F)
+knitr::kable(RES_aphid_count$gfc_p_CvM, caption = "Aphid_count : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula")
+```
+
+```{r count_p_KS,echo=F,results="asis"}
+table_nums("count_p_KS",display=F)
+knitr::kable(RES_aphid_count$gfc_p_KS,caption = "Aphid_count : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula")
+```
+
+`r table_nums("count_bestcop",display="c")` shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in `r table_nums("count_LTdep_AICw",display="c")` and `r table_nums("count_UTdep_AICw",display="c")` respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper) in `r table_nums("count_LTmUT",display="c")` indicates it is different from 0 for most of the locations, answering the third part of question 1.
+
+These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)`** out of **`r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-RES_aphid_count$num_indep_loc_pair`** location pair with greater lower tail dependence.
+Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [`r table_nums("count_p_CvM",display="c")`] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [`r table_nums("count_p_KS",display="c")`] of the goodness of fit.
+
+The next section looks at tail dependence in a non-parametric way.
+
 First flight results
 
-```{r aphid_firstflight, echo=F, cache=T, cache.extra=list(seed,d0_firstflight)}
+```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff)} 
 set.seed(seed)
-cat(paste("start-time: ",Sys.time(),"\n"))
-RES_aphid_firstflight<-RES_single_sp(d_allsp=d_allsp_data(d0_firstflight),sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
-cat(paste("stop-time: ",Sys.time(),"\n"))
-resloc<-"Results/"
-saveRDS(RES_aphid_firstflight,paste(resloc,file="AphidCopulaFit_selecloc_firstflight_species_11.RDS",sep=""))
+#cat(paste("start-time: ",Sys.time(),"\n"))
+RES_aphid_ff<-RES_single_sp(d_allsp=data_aphid_ff,sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
+#cat(paste("stop-time: ",Sys.time(),"\n")) 
+resloc<-"./Results/fitting_results/"
+saveRDS(RES_aphid_ff,paste(resloc,file="AphidCopulaFit_selecloc_firstflight_species_11.RDS",sep=""))
 ```
 
-```{r}
-RES_aphid_firstflight
-#no of locations showing greater lower tail dependence
-sum((RES_aphid_firstflight$LTdep_AICw-RES_aphid_firstflight$UTdep_AICw)>0,na.rm=T)
+```{r read_RES_aphid_ff,echo=F}
+RES_aphid_ff<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_firstflight_species_11.RDS")
+```
 
-#no of locations showing greater upper tail dependence
-sum((RES_aphid_firstflight$LTdep_AICw-RES_aphid_firstflight$UTdep_AICw)<0,na.rm=T)
+```{r ff_bestcop,echo=F,results="asis"}
+table_nums("ff_bestcop",display=F)
+knitr::kable(RES_aphid_ff$info_ord_copname[,,1], caption = "Aphid_firstflight : Best fit copula family for each selective location pair")
+```
 
+```{r ff_AICw,echo=F,results="asis"}
+table_nums("ff_AICw",display=F)
+knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], caption = "Aphid_firstflight : AICw for best fit copula family for each selective location pair")
 ```
+
+```{r ff_LTdep,echo=F,results="asis"}
+cat("\n\n\\pagebreak\n")
+
+table_nums("ff_LTdep",display=F)
+knitr::kable(RES_aphid_ff$info_ord_LTdep[,,1], caption = "Aphid_firstflight : Lower Tail dependence for best fit copula family for each selective location pair")
+```
+
+```{r ff_UTdep,echo=F,results="asis"}
+table_nums("ff_UTdep",display=F)
+knitr::kable(RES_aphid_ff$info_ord_UTdep[,,1], caption = "Aphid_firstflight : Upper Tail dependence for best fit copula family for each selective location pair")
+```
+
+```{r ff_LTdep_AICw,echo=F,results="asis"}
+table_nums("ff_LTdep_AICw",display=F)
+knitr::kable(RES_aphid_ff$LTdep_AICw, caption = "Aphid_firstflight : Lower Tail dependence statistic for best fit copula family for each selective location pair")
+```
+
+```{r ff_UTdep_AICw,echo=F,results="asis"}
+table_nums("ff_UTdep_AICw",display=F)
+knitr::kable(RES_aphid_ff$UTdep_AICw, caption = "Aphid_firstflight : Upper Tail dependence statistic for best fit copula family for each selective location pair")
+```
+
+```{r ff_LTmUT,echo=F,results="asis"}
+table_nums("ff_LTmUT",display=F)
+knitr::kable(round(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,5), caption = "Aphid_firstflight : Lower Tail - Upper Tail dependence statistic for best fit copula family for each of the selective location pair")
+```
+
+```{r ff_p_CvM,echo=F,results="asis"}
+table_nums("ff_p_CvM",display=F)
+knitr::kable(RES_aphid_ff$gfc_p_CvM, caption = "Aphid_firstflight : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula")
+```
+
+```{r ff_p_KS,echo=F,results="asis"}
+table_nums("ff_p_KS",display=F)
+knitr::kable(RES_aphid_ff$gfc_p_KS, caption = "Aphid_firstflight : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula")
+```
+
+`r table_nums("ff_bestcop",display="c")`  shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in `r table_nums("ff_LTdep_AICw",display="c")` and `r table_nums("ff_UTdep_AICw",display="c")` respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper) in `r table_nums("ff_LTmUT",display="c")` indicates it is different from 0 for most of the locations, answering the third part of question 1.
+
+These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.
+Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [`r table_nums("ff_p_CvM",display="c")`] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [`r table_nums("ff_p_KS",display="c")`] of the goodness of fit.
 
@@
 
-Cout results
+# Count results
+```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count)}
+set.seed(seed) 
+resloc<-"./Results/stat_results/stat_aphid_count/"
 
-First flight results
+sp<-10
+good_loc <- good_loclist(d_allsp=data_aphid_count,sp=sp,data_pt_thrs=30)
+longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
+lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)
+
+stat_aphid_count<-multcall(d_allsp=data_aphid_count,sp=sp,lats=lats,longs=longs,pfname=paste(resloc,"Sp_",sp,sep=''),good_loc=good_loc)
+saveRDS(stat_aphid_count,paste(resloc,file="stat_aphid_count_sp_10.RDS",sep='')) 
+```
+
+```{r read_stat_aphid_count,echo=F}
+stat_aphid_count<-readRDS("./Results/stat_results/stat_aphid_count/stat_aphid_count_sp_10.RDS")
+```
+
+```{r count_resamp,echo=F,results="asis"}
+table_nums("count_resamp",display=F)
+knitr::kable(stat_aphid_count$numericdf, caption = "Aphid_count : resampling result for stat indicating lower tail dependence")
+```
+
+```{r count_allcop,echo=F}
+fig_nums("count_allcop",display=F)
+```
+![copula plot for aphid-count](./Results/stat_results/stat_aphid_count/Sp_10_AllCops.pdf)
+
+```{r count_spear,echo=F}
+fig_nums("count_spear",display=F)
+```
+![Plot for aphid-count : Spearman vs. Distance](./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf)
+
+```{r count_kend,echo=F}
+fig_nums("count_kend",display=F)
+```
+![Plot for aphid-count : Kendall vs. Distance](./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf)
+
+```{r count_Corl-Coru,echo=F}
+fig_nums("count_Corl-Coru",display=F)
+```
+![Corl-Coru vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf)
+
+```{r count_Pl-Pu,echo=F}
+fig_nums("count_Pl-Pu",display=F)
+```
+![Pl-Pu vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf)
+
+```{r count_Tl-Tu,echo=F}
+fig_nums("count_Tl-Tu",display=F)
+```
+![Tl-Tu vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Tl-Tu_vs_D.pdf)
+
+```{r count_D2u-D2l,echo=F}
+fig_nums("count_D2u-D2l",display=F)
+```
+![D2u-D2l vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf)
+
+```{r count_D2u/D2l,echo=F}
+fig_nums("count_D2u/D2l",display=F)
+```
+![D2u/D2l vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_D2u_by_D2l_vs_D.pdf)
+
+<!--
+```{r, echo=F,out.width = "8cm", out.height= "8cm", out.caption="adhjk"}
+knitr::include_graphics("./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf")
+knitr::include_graphics("./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf")
+```
+-->
+
+# First flight results
+
+```{r stat_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff)}
+set.seed(seed)
+resloc<-"./Results/stat_results/stat_aphid_ff/"
+
+sp<-11
+good_loc<-good_loclist(d_allsp=data_aphid_ff,sp=sp,data_pt_thrs=30)
+longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
+lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)
+
+stat_aphid_ff<-multcall(d_allsp=data_aphid_ff,sp=sp,lats=lats,longs=longs,pfname=paste(resloc,"Sp_",sp,sep=''),good_loc=good_loc) 
+saveRDS(stat_aphid_ff,paste(resloc,file="stat_aphid_ff_sp_11.RDS",sep='')) 
+```
+
+```{r read_stat_aphid_ff,echo=F}
+stat_aphid_ff<-readRDS("./Results/stat_results/stat_aphid_ff/stat_aphid_ff_sp_11.RDS")
+```
+
+```{r ff_resamp,echo=F,results="asis"}
+table_nums("ff_resamp",display=F)
+knitr::kable(stat_aphid_ff$numericdf, caption = "Aphid_firstflight : resampling result for stat indicating upper tail dependence")
+```
+
+```{r ff_allcop,echo=F}
+fig_nums("ff_allcop",display=F)
+```
+![copula plot for aphid-firstflight](./Results/stat_results/stat_aphid_ff/Sp_11_AllCops.pdf)
+
+```{r ff_spear,echo=F}
+fig_nums("ff_spear",display=F)
+```
+![Plot for aphid-firstflight : Spearman vs. Distance](./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf)
+
+```{r ff_kend,echo=F}
+fig_nums("ff_kend",display=F)
+```
+![Plot for aphid-firstflight : Kendall vs. Distance](./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf)
+
+```{r ff_Corl-Coru,echo=F}
+fig_nums("ff_Corl-Coru",display=F)
+```
+![Corl-Coru vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf)
+
+```{r ff_Pl-Pu,echo=F}
+fig_nums("ff_Pl-Pu",display=F)
+```
+![Pl-Pu vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf)
+
+```{r ff_Tl-Tu,echo=F}
+fig_nums("ff_Tl-Tu",display=F)
+```
+![Tl-Tu vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Tl-Tu_vs_D.pdf)
+
+```{r ff_D2u-D2l,echo=F}
+fig_nums("ff_D2u-D2l",display=F)
+```
+![D2u-D2l vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf)
+
+```{r ff_D2u/D2l,echo=F}
+fig_nums("ff_D2u/D2l",display=F)
+```
+![D2u/D2l vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_D2u_by_D2l_vs_D.pdf)
 
@@
 
-```{r d_allsp_data_plankton_north_sea, echo=F}
-d_allsp_data_plankton_north_sea<-function(d0){
-
-  name.plankton<-c('Calanus_I-IV','Para-Pseudocalanus_spp','Temora_longicornis','Acartia_spp_(unidentified)','Centropages_typicus',
-                   'Oithona_spp','Echinoderm_larvae','Calanus_finmarchicus','Calanus_helgolandicus','Metridia_lucens','Decapoda_larvae_(Total)',
-                   'Euphausiacea_Total','Thalassiosira_spp','Rhizosolenia_styliformis','Ceratium_fusus','Ceratium_furca','Ceratium_tripos',
-                   'Ceratium_macroceros','Rhizosolenia_alata_alata','Pseudocalanus_elongatus_Adult','Nitzschia_delicatissima','Nitzschia_seriata')
-
-  longs<-read.csv("Data/Plankton_North_Sea_data/boxcornerlongs201117.csv",header=F)
-  lats<-read.csv("Data/Plankton_North_Sea_data/boxcornerlats201117.csv",header=F)
-  longs<-longs[[1]]
-  lats<-lats[[1]]
-  d0[d0==-999]<-NA
-  Years<-1958:2013
-
-  d_allsp<-list() #this is a list with 20 elements, all the species
-  for (sp_c in 1:22)
-  {
-    d_thissp<-list() #a list with 11 locations for the current species
-    d1<-d0[((sp_c-1)*26+1):(sp_c*26),]
-    for (st_c in 1:26)
-    {
-      d_thissp[[st_c]]<-data.frame(Year=Years,Dat=d1[st_c,])
-    }
-    d_allsp[[sp_c]]<-d_thissp
-  }
-  names(d_allsp)<-name.plankton
-
-  return(d_allsp=d_allsp)
-
-}
-
-d0_plankton_north_sea<-as.matrix(read.csv("Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))
-
+```{r read_plankton_north_sea_data, echo=F}
+d0_plankton_north_sea<-as.matrix(read.csv("./Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))
+data_plankton_north_sea<-d_allsp_data_plankton_north_sea(d0_plankton_north_sea)
 ```
 
-```{r plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,d0_plankton_north_sea)}
+```{r RES_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea)}
 set.seed(seed)
-cat(paste("start-time: ",Sys.time(),"\n"))
-RES_plankton_north_sea<-RES_single_sp(d_allsp=d_allsp_data_plankton_north_sea(d0_plankton_north_sea),sp=16,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=45)
-cat(paste("stop-time: ",Sys.time(),"\n"))
-resloc<-"Results/"
+#cat(paste("start-time: ",Sys.time(),"\n")) 
+RES_plankton_north_sea<-RES_single_sp(d_allsp=data_plankton_north_sea,sp=16,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=45)
+#cat(paste("stop-time: ",Sys.time(),"\n"))    
+resloc<-"./Results/fitting_results/"
 saveRDS(RES_plankton_north_sea,paste(resloc,file="Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS",sep=""))
@@
 
-```{r}
-RES_plankton_north_sea
-#no of locations showing greater lower tail dependence
-sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)
-
-#no of locations showing greater upper tail dependence
-sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)<0,na.rm=T)
-
+```{r read_RES_plankton_north_sea,echo=F}
+RES_plankton_north_sea<-readRDS("./Results/fitting_results/Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS")
 ```
 
+```{r pns_bestcop,echo=F,results="asis"}
+#library(pander)
+#pander(best_cop_LTdep, caption = "Hello caption!")
+# alternatively,
+#pandoc.table(RES_plankton_north_sea$info_ord_AICw[,,1], caption="Plankton_north_sea : AICw for best fit copula family for each selective location pair")
+cat("\n\n\\pagebreak\n")
+
+table_nums("pns_bestcop",display=F)
+knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1], caption = "Plankton_north_sea : Best fit copula family for each selective location pair")
+```
+
+```{r pns_AICw,echo=F,results="asis"}
+table_nums("pns_AICw",display=F)
+knitr::kable(round(RES_plankton_north_sea$info_ord_AICw[,,1],3), caption = "Plankton_north_sea : AICw for best fit copula family for each selective location pair")
+```
+
+```{r pns_LTdep,echo=F,results="asis"}
+cat("\n\n\\pagebreak\n")
+
+table_nums("pns_LTdep",display=F)
+knitr::kable(round(RES_plankton_north_sea$info_ord_LTdep[,,1],3), caption = "Plankton_north_sea : Lower Tail dependence for best fit copula family for each selective location pair")
+```
+
+```{r pns_UTdep,echo=F,results="asis"}
+table_nums("pns_UTdep",display=F)
+knitr::kable(round(RES_plankton_north_sea$info_ord_UTdep[,,1],3), caption = "Plankton_north_sea : Upper Tail dependence for best fit copula family for each selective location pair")
+```
+
+```{r pns_LTdep_AICw,echo=F,results="asis"}
+cat("\n\n\\pagebreak\n")
+
+table_nums("pns_LTdep_AICw",display=F)
+knitr::kable(round(RES_plankton_north_sea$LTdep_AICw,3), caption = "Plankton_north_sea : Lower Tail dependence statistic for best fit copula family for each selective location pair")
+```
+
+```{r pns_UTdep_AICw,echo=F,results="asis"}
+table_nums("pns_UTdep_AICw",display=F)
+knitr::kable(round(RES_plankton_north_sea$UTdep_AICw,3), caption = "Plankton_north_sea : Upper Tail dependence statistic for best fit copula family for each selective location pair")
+```
+
+```{r pns_LTmUT,echo=F,results="asis"}
+cat("\n\n\\pagebreak\n")
+
+table_nums("pns_LTmUT",display=F)
+knitr::kable(round(RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw,3), caption = "Plankton_north_sea : Lower Tail - Upper Tail dependence statistic for best fit copula family for each of the selective location pair")
+```
+
+```{r pns_p_CvM,echo=F,results="asis"}
+table_nums("pns_p_CvM",display=F)
+knitr::kable(round(RES_plankton_north_sea$gfc_p_CvM,3), caption = "Plankton_north_sea : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula")
+```
+
+```{r pns_p_KS,echo=F,results="asis"}
+cat("\n\n\\pagebreak\n")
+
+table_nums("pns_p_KS",display=F)
+knitr::kable(round(RES_plankton_north_sea$gfc_p_KS,3), caption = "Plankton_north_sea : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula")
+```
+
+`r table_nums("pns_bestcop",display="c")` shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in   `r table_nums("pns_LTdep_AICw",display="c")` and `r table_nums("pns_UTdep_AICw",display="c")` respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper) in `r table_nums("pns_LTmUT",display="c")` indicates it is different from 0 for most of the locations, answering the third part of question 1.
+
+These results do convincingly show non-Gaussian copula structure, having `r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-RES_plankton_north_sea$num_indep_loc_pair` location pair more lower tail dependence.
+Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [`r table_nums("pns_p_CvM",display="c")`] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [`r table_nums("pns_p_KS",display="c")`] of the goodness of fit.
+
 ### Non-paramtric analysis of tail dependence
+```{r stat_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea)}
+set.seed(seed)
+resloc<-"./Results/stat_results/stat_plankton_north_sea/"
 
+sp<-16
+good_loc<-good_loclist(d_allsp=data_plankton_north_sea,sp=sp,data_pt_thrs=45)
+longs<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlongs201117.csv",header=F)
+lats<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlats201117.csv",header=F)
+longs<-longs[[1]] 
+lats<-lats[[1]]
+
+stat_plankton_north_sea<-multcall(d_allsp=data_plankton_north_sea,sp=sp,lats=lats,longs=longs,pfname=paste(resloc,"Sp_",sp,sep=''),good_loc=good_loc) 
+saveRDS(stat_plankton_north_sea,paste(resloc,file="stat_plankton_north_sea_sp_16.RDS",sep=''))
+```
+
+```{r read_stat_plankton_north_sea,echo=F}
+stat_plankton_north_sea<-readRDS("./Results/stat_results/stat_plankton_north_sea/stat_plankton_north_sea_sp_16.RDS")
+```
+
+```{r pns_resamp,echo=F,results="asis"}
+table_nums("count_resamp",display=F)
+knitr::kable(stat_plankton_north_sea$numericdf, caption = "Plankton_north_sea : resampling result for stat indicating lower tail dependence")
+```
+
+```{r pns_allcop,echo=F}
+fig_nums("pns_allcop",display=F)
+```
+![copula plot for Plankton_north_sea](./Results/stat_results/stat_plankton_north_sea/Sp_16_AllCops.pdf)
+
+```{r pns_spear,echo=F}
+fig_nums("pns_spear",display=F)
+```
+![Plot for Plankton_north_sea : Spearman vs. Distance](./Results/stat_results/stat_plankton_north_sea/Sp_16_Spearman_vs_D.pdf)
+
+```{r pns_kend,echo=F}
+fig_nums("pns_kend",display=F)
+```
+![Plot for Plankton_north_sea : Kendall vs. Distance](./Results/stat_results/stat_plankton_north_sea/Sp_16_Kendall_vs_D.pdf)
+
+```{r pns_Corl-Coru,echo=F}
+fig_nums("pns_Corl-Coru",display=F)
+```
+![Corl-Coru vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf)
+
+```{r pns_Pl-Pu,echo=F}
+fig_nums("pns_Pl-Pu",display=F)
+```
+![Pl-Pu vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf)
+
+```{r pns_Tl-Tu,echo=F}
+fig_nums("pns_Tl-Tu",display=F)
+```
+![Tl-Tu vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Tl-Tu_vs_D.pdf)
+
+```{r pns_D2u-D2l,echo=F}
+fig_nums("pns_D2u-D2l",display=F)
+```
+![D2u-D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf)
+
+```{r pns_D2u/D2l,echo=F}
+fig_nums("pns_D2u/D2l",display=F)
+```
+![D2u/D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u_by_D2l_vs_D.pdf)
+
+<!--
 # Question 2 analyses for each dataset
@@
 # Question 3 analyses for each dataset
+-->

### 2018-02-26 SupportingInformation.Rmd

@@
 ---
+
 updated on `r Sys.Date()`
 
+<!--Shya pls add a short description of what this chunk is for. Should it be in the setup chunk down below?-->
 ```{r ,echo=FALSE}
@@
 fig_nums <- captioner(prefix="Figure")
-```
-
-<!--A function needed for caching-->
-```{r echo=F}
-mtime <- function(files)
-{
-  lapply(Sys.glob(files),function(x) X =   
-          file.info(x)$mtime)
-}
 ```
@@
 families<-c(1,3:10,13,14,16:20)
+#A function needed for caching
+mtime <- function(files)
+{
+  lapply(Sys.glob(files),function(x) X =   
+          file.info(x)$mtime)
+}
 source("getcopula.R")
@@
 source("NonParamStat.R")
+source("copsurrog2d.R")
 ```
@@
 
-```{r load_data_Loecke, echo=F, fig.cap="plot of the raw data", fig.height=5, fig.width=5}
-#Load your data - should be a data frame with 2 columns. Can contain NAs.
+The raw data look like this:
+<!--Load the soil C and N data and make a plot-->
+```{r load_data_Loecke, echo=F, fig.cap="Plot of the raw soil C and N data.", fig.height=5, fig.width=5}
 d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
@@
 plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1)
-rect(0,0,1,1)
 ```
 
-```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C/N data",  fig.height=5, fig.width=5}
+The copula looks like this:
+<!--Plot the copula of the soil C and N data--> 
+```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of the soil C and N data",  fig.height=5, fig.width=5}
 #convert to a copula and plot
@@
 # Question 1 analyses for each dataset
-
 Recall that question 1 is: Do datasets in ecology and related fields have non-Gausian copula srtructure?
@@
 ### Model selection approach
-
-```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v)}
+```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v,mtime(list("OurBiCopSelect.R")))}
 set.seed(seed)
@@
 ```
+The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes`.
 
-The test of indepedence gave *p=*`r BivMS_res_Loecke$IndepTestRes` . We fitted several copulas
+We fitted several copulas
 and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas:
-
 ```{r soilCNfit,echo=F}
 h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-#rownames(h)<-NULL
 table_nums("soilCNfit",display=F)
-knitr::kable(h, caption = "Fitting info for soilC/N dataset")
+knitr::kable(h, caption = "Fitting info for soil C and N dataset.
+Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=Survival Clayton; SG=Survival Gumbel; SJ=Survival Joe.
+An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by 180 degrees).")
 ```
-
-`r table_nums("soilCNfit",display="c")` shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+`r table_nums("soilCNfit",display="c")` shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
 
@@
 
-```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v)}
+<!--DAN: wherever copsync is needs to be sourced and also added to the mlist below-->
+```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v,mlist(list("CopulaFunctions.R","copsurrog2d.R")))}
 set.seed(seed) 
@@
 
+#get the stats for the real data
 stat_soilCN<-copsync(v) 
@@
 stat<-unlist(sCN)
+
+#get surrogates and get stats for them
+cop<-normalCopula(.5,2)
+surv<-copsurrog2d(v,cop,1000)
+#DAN: once you get the code for copsync, apply it to each surrogate and keep the results
 ```
 
+<!--DAN: modify the table so it also shows tail asymmetry stats (differences) and also shows fractions of surrogates for both the tail dependence stats and the asymmetry stats that are greater than the real data stat-->
 ```{r table_stat_soilCN, echo=F, results="asis"}
@@
 
+<!--DAN: The below will no longer be necessary once you make the above changes-->
 ```{r ,echo=F}
@@
 ```
-
-<!--Shya could work on this. Select 1-2 statistics, maybe more, but try to cut some of them (looks like there will be 4). Apply to C/N data, get results, display in the same kind of way.-->
 

### 2018-02-27 SupportingInformation.Rmd
added functionality to the non-parametric test of tail dependence and asymmetry of tail dependence for soil C and N data

@@
 
+<!--Shya pls add a description of each of the statistics implemented in CopulaFunctions.R. As usual, please describe the stats with sufficient precision that they could be re-coded by the reader.-->
+
 ## Soil C and N data
@@
 ### Non-paramtric analysis of tail dependence
-
-<!--DAN: wherever copsync is needs to be sourced and also added to the mlist below-->
-```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v,mlist(list("CopulaFunctions.R","copsurrog2d.R")))}
+<!--Do the analysis-->
+```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v,mlist(list("CopulaFunctions.R","copsurrog2d.R","NonParamStat.R")))}
 set.seed(seed) 
@@
 stat<-unlist(sCN)
+stat<-c(stat[1:4],unname(stat[3]-stat[4]),stat[5:6],unname(stat[5]-stat[6]),stat[7:8],unname(stat[7]-stat[8]),stat[9:10],unname(stat[9]-stat[10]))
+names(stat)[c(5,8,11,14)]<-c("Corl_m_Coru","Pl_m_Pu","Tl_m_Tu","D2l_m_D2u")
 
@@
 cop<-normalCopula(.5,2)
-surv<-copsurrog2d(v,cop,1000)
-#DAN: once you get the code for copsync, apply it to each surrogate and keep the results
+numsurrog<-1000
+surv<-copsurrog2d(v,cop,numsurrog)
+statsurr<-matrix(NA,length(stat),numsurrog)
+rownames(statsurr)<-names(stat)
+for (counter in 1:dim(surv)[3])
+{
+  h<-copsync(surv[,,counter])
+  h<-unlist(h[c(2:5,7,8,10:13)])
+  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[9]-h[10])
+  statsurr[,counter]<-unname(h)
+}
+
+#now get fractions of surrogates lt the data and gt the data for each stat
+stattab<-data.frame(statname=names(stat),stat=unname(stat),frac_surr_gt=NA,frac_surr_lt=NA)
+for (counter in 1:length(stat))
+{
+  stattab[counter,3]<-sum(statsurr[counter,]>stat[counter])/numsurrog
+  stattab[counter,4]<-sum(statsurr[counter,]<stat[counter])/numsurrog
+}
 ```
 
-<!--DAN: modify the table so it also shows tail asymmetry stats (differences) and also shows fractions of surrogates for both the tail dependence stats and the asymmetry stats that are greater than the real data stat-->
+<!--Display the table generated above-->
 ```{r table_stat_soilCN, echo=F, results="asis"}
 table_nums("table_stat_soilCN",display=F)
-
-knitr::kable(data.frame(stat=stat),caption = "Non-parametric stats on soil C/N dataset")
-```
-
-<!--DAN: The below will no longer be necessary once you make the above changes-->
-```{r ,echo=F}
-#sCN
-CorlmCoru<-stat_soilCN$Corl-stat_soilCN$Coru
-cat("Corl-Coru = ",CorlmCoru,"\n")
-PlmPu<-stat_soilCN$Pl-stat_soilCN$Pu
-cat("Pl-Pu = ",PlmPu,"\n")
-TlmTu<-stat_soilCN$Tl-stat_soilCN$Tu
-cat("Tl-Tu = ",TlmTu,"\n")
-D2umD2l<-stat_soilCN$D2u-stat_soilCN$D2l
-cat("D2u-D2l = ",D2umD2l,"\n")
+knitr::kable(data.frame(stat=stattab),caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset.")
 ```

### 2018-02-27 SupportingInformation.Rmd
small further changes

@@
 ### Model selection approach
-```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v,mtime(list("OurBiCopSelect.R")))}
+```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v,mtime("OurBiCopSelect.R"))}
 set.seed(seed)
@@
 <!--Do the analysis-->
-```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v,mlist(list("CopulaFunctions.R","copsurrog2d.R","NonParamStat.R")))}
+```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v,mtime("CopulaFunctions.R"),mtime("copsurrog2d.R"),mtime("NonParamStat.R"))}
 set.seed(seed) 

### 2018-02-27 SupportingInformation.Rmd
it runs, also rearranged aphid count and ff results into separate sections

@@
----
-output:
-  pdf_document: default
----
 ---
@@
 ```
+
 `r table_nums("soilCNfit",display="c")` shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
@@
 
-## Aphid data
-<!--Shya to edit the prentation below to make it more parallel to the presentation of the model selection results for C/N data above-->
+## Aphid abundance data 
+<!--Add common and species name to header-->
 
@@
 ```
-
-Count results
 
@@
 ```{r count_AICw,echo=F,results="asis"}
-
-cat("\n\n\\pagebreak\n")
-
 table_nums("count_AICw",display=F)
@@
 
-First flight results
-
-```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff)} 
-set.seed(seed)
-#cat(paste("start-time: ",Sys.time(),"\n"))
-RES_aphid_ff<-RES_single_sp(d_allsp=data_aphid_ff,sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
-#cat(paste("stop-time: ",Sys.time(),"\n")) 
-resloc<-"./Results/fitting_results/"
-saveRDS(RES_aphid_ff,paste(resloc,file="AphidCopulaFit_selecloc_firstflight_species_11.RDS",sep=""))
-```
-
-```{r read_RES_aphid_ff,echo=F}
-RES_aphid_ff<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_firstflight_species_11.RDS")
-```
-
-```{r ff_bestcop,echo=F,results="asis"}
-table_nums("ff_bestcop",display=F)
-knitr::kable(RES_aphid_ff$info_ord_copname[,,1], caption = "Aphid_firstflight : Best fit copula family for each selective location pair")
-```
-
-```{r ff_AICw,echo=F,results="asis"}
-table_nums("ff_AICw",display=F)
-knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], caption = "Aphid_firstflight : AICw for best fit copula family for each selective location pair")
-```
-
-```{r ff_LTdep,echo=F,results="asis"}
-cat("\n\n\\pagebreak\n")
-
-table_nums("ff_LTdep",display=F)
-knitr::kable(RES_aphid_ff$info_ord_LTdep[,,1], caption = "Aphid_firstflight : Lower Tail dependence for best fit copula family for each selective location pair")
-```
-
-```{r ff_UTdep,echo=F,results="asis"}
-table_nums("ff_UTdep",display=F)
-knitr::kable(RES_aphid_ff$info_ord_UTdep[,,1], caption = "Aphid_firstflight : Upper Tail dependence for best fit copula family for each selective location pair")
-```
-
-```{r ff_LTdep_AICw,echo=F,results="asis"}
-table_nums("ff_LTdep_AICw",display=F)
-knitr::kable(RES_aphid_ff$LTdep_AICw, caption = "Aphid_firstflight : Lower Tail dependence statistic for best fit copula family for each selective location pair")
-```
-
-```{r ff_UTdep_AICw,echo=F,results="asis"}
-table_nums("ff_UTdep_AICw",display=F)
-knitr::kable(RES_aphid_ff$UTdep_AICw, caption = "Aphid_firstflight : Upper Tail dependence statistic for best fit copula family for each selective location pair")
-```
-
-```{r ff_LTmUT,echo=F,results="asis"}
-table_nums("ff_LTmUT",display=F)
-knitr::kable(round(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,5), caption = "Aphid_firstflight : Lower Tail - Upper Tail dependence statistic for best fit copula family for each of the selective location pair")
-```
-
-```{r ff_p_CvM,echo=F,results="asis"}
-table_nums("ff_p_CvM",display=F)
-knitr::kable(RES_aphid_ff$gfc_p_CvM, caption = "Aphid_firstflight : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula")
-```
-
-```{r ff_p_KS,echo=F,results="asis"}
-table_nums("ff_p_KS",display=F)
-knitr::kable(RES_aphid_ff$gfc_p_KS, caption = "Aphid_firstflight : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula")
-```
-
-`r table_nums("ff_bestcop",display="c")`  shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
-The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in `r table_nums("ff_LTdep_AICw",display="c")` and `r table_nums("ff_UTdep_AICw",display="c")` respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper) in `r table_nums("ff_LTmUT",display="c")` indicates it is different from 0 for most of the locations, answering the third part of question 1.
-
-These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.
-Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [`r table_nums("ff_p_CvM",display="c")`] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [`r table_nums("ff_p_KS",display="c")`] of the goodness of fit.
-
 ### Non-paramtric analysis of tail dependence
 
-# Count results
 ```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count)}
@@
 
-# First flight results
+## Aphid first flight data
+<!--Add species common name and scientific name to header-->
+
+### Model selection approach
+
+```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff)} 
+set.seed(seed)
+#cat(paste("start-time: ",Sys.time(),"\n"))
+RES_aphid_ff<-RES_single_sp(d_allsp=data_aphid_ff,sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
+#cat(paste("stop-time: ",Sys.time(),"\n")) 
+resloc<-"./Results/fitting_results/"
+saveRDS(RES_aphid_ff,paste(resloc,file="AphidCopulaFit_selecloc_firstflight_species_11.RDS",sep=""))
+```
+
+```{r read_RES_aphid_ff,echo=F}
+RES_aphid_ff<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_firstflight_species_11.RDS")
+```
+
+```{r ff_bestcop,echo=F,results="asis"}
+table_nums("ff_bestcop",display=F)
+knitr::kable(RES_aphid_ff$info_ord_copname[,,1], caption = "Aphid_firstflight : Best fit copula family for each selective location pair")
+```
+
+```{r ff_AICw,echo=F,results="asis"}
+table_nums("ff_AICw",display=F)
+knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], caption = "Aphid_firstflight : AICw for best fit copula family for each selective location pair")
+```
+
+```{r ff_LTdep,echo=F,results="asis"}
+table_nums("ff_LTdep",display=F)
+knitr::kable(RES_aphid_ff$info_ord_LTdep[,,1], caption = "Aphid_firstflight : Lower Tail dependence for best fit copula family for each selective location pair")
+```
+
+```{r ff_UTdep,echo=F,results="asis"}
+table_nums("ff_UTdep",display=F)
+knitr::kable(RES_aphid_ff$info_ord_UTdep[,,1], caption = "Aphid_firstflight : Upper Tail dependence for best fit copula family for each selective location pair")
+```
+
+```{r ff_LTdep_AICw,echo=F,results="asis"}
+table_nums("ff_LTdep_AICw",display=F)
+knitr::kable(RES_aphid_ff$LTdep_AICw, caption = "Aphid_firstflight : Lower Tail dependence statistic for best fit copula family for each selective location pair")
+```
+
+```{r ff_UTdep_AICw,echo=F,results="asis"}
+table_nums("ff_UTdep_AICw",display=F)
+knitr::kable(RES_aphid_ff$UTdep_AICw, caption = "Aphid_firstflight : Upper Tail dependence statistic for best fit copula family for each selective location pair")
+```
+
+```{r ff_LTmUT,echo=F,results="asis"}
+table_nums("ff_LTmUT",display=F)
+knitr::kable(round(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,5), caption = "Aphid_firstflight : Lower Tail - Upper Tail dependence statistic for best fit copula family for each of the selective location pair")
+```
+
+```{r ff_p_CvM,echo=F,results="asis"}
+table_nums("ff_p_CvM",display=F)
+knitr::kable(RES_aphid_ff$gfc_p_CvM, caption = "Aphid_firstflight : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula")
+```
+
+```{r ff_p_KS,echo=F,results="asis"}
+table_nums("ff_p_KS",display=F)
+knitr::kable(RES_aphid_ff$gfc_p_KS, caption = "Aphid_firstflight : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula")
+```
+
+`r table_nums("ff_bestcop",display="c")`  shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in `r table_nums("ff_LTdep_AICw",display="c")` and `r table_nums("ff_UTdep_AICw",display="c")` respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper) in `r table_nums("ff_LTmUT",display="c")` indicates it is different from 0 for most of the locations, answering the third part of question 1.
+
+These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.
+Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [`r table_nums("ff_p_CvM",display="c")`] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [`r table_nums("ff_p_KS",display="c")`] of the goodness of fit.
+
+### Non-parametric analysis of tail dependence
 
@@
 
-<!--
 # Question 2 analyses for each dataset
@@
 # Question 3 analyses for each dataset
--->

### 2018-02-27 SupportingInformation.Rmd
added both kendall and spearman preserving surrogte runs to the nonparametric tests of tail dependence for soil C and N data

@@
 
-#get surrogates and get stats for them
+#get surrogates and get stats for them, preserving kendall
 cop<-normalCopula(.5,2)
 numsurrog<-1000
-surv<-copsurrog2d(v,cop,numsurrog)
-statsurr<-matrix(NA,length(stat),numsurrog)
-rownames(statsurr)<-names(stat)
+surv<-copsurrog2d(v,cop,"kendall",numsurrog)
+statsurr_k<-matrix(NA,length(stat),numsurrog)
+rownames(statsurr_k)<-names(stat)
 for (counter in 1:dim(surv)[3])
@@
   h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[9]-h[10])
-  statsurr[,counter]<-unname(h)
+  statsurr_k[,counter]<-unname(h)
+}
+
+#same, preserving spearman
+surv<-copsurrog2d(v,cop,"spearman",numsurrog)
+statsurr_s<-matrix(NA,length(stat),numsurrog)
+rownames(statsurr_s)<-names(stat)
+for (counter in 1:dim(surv)[3])
+{
+  h<-copsync(surv[,,counter])
+  h<-unlist(h[c(2:5,7,8,10:13)])
+  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[9]-h[10])
+  statsurr_s[,counter]<-unname(h)
 }
@@
 #now get fractions of surrogates lt the data and gt the data for each stat
-stattab<-data.frame(statname=names(stat),stat=unname(stat),frac_surr_gt=NA,frac_surr_lt=NA)
+stattab<-data.frame(statname=names(stat),stat=unname(stat),frac_surr_gt_k=NA,frac_surr_lt_k=NA,frac_surr_gt_s=NA,frac_surr_lt_s=NA)
 for (counter in 1:length(stat))
 {
-  stattab[counter,3]<-sum(statsurr[counter,]>stat[counter])/numsurrog
-  stattab[counter,4]<-sum(statsurr[counter,]<stat[counter])/numsurrog
+  stattab[counter,3]<-sum(statsurr_k[counter,]>stat[counter])/numsurrog
+  stattab[counter,4]<-sum(statsurr_k[counter,]<stat[counter])/numsurrog
+  stattab[counter,5]<-sum(statsurr_s[counter,]>stat[counter])/numsurrog
+  stattab[counter,6]<-sum(statsurr_s[counter,]<stat[counter])/numsurrog
 }
@@
 ```{r table_stat_soilCN, echo=F, results="asis"}
+names(stattab)<-c("Statistic","Value on data","Frac. K surr >","Frac. K surr <","Frac. S surr >","Frac. S surr <")
 table_nums("table_stat_soilCN",display=F)
-knitr::kable(data.frame(stat=stattab),caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset.")
+knitr::kable(data.frame(stat=stattab),caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset.
+Names and values on the actual C and N data are listed in the first two columns.
+The third (respectively, fourth) column is the fraction of Kendall-preserving surrogate datasets for which the given statistic is greater than (resp., less than) its value on the real data.
+The last two columns are analogous but for Spearman-preserving surrogates.")
 ```

### 2018-02-28 SupportingInformation.Rmd
minor changes to chunk setup so reruns will be more efficient and minor formatting changes are easier

@@
 ### Non-paramtric analysis of tail dependence
-<!--Do the analysis-->
+<!--Do the intensive part of the analysis-->
 ```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v,mtime("CopulaFunctions.R"),mtime("copsurrog2d.R"),mtime("NonParamStat.R"))}
@@
 }
+```
 
+<!--Display the table generated above-->
+```{r table_stat_soilCN, echo=F, results="asis"}
 #now get fractions of surrogates lt the data and gt the data for each stat
@@
 }
-```
 
-<!--Display the table generated above-->
-```{r table_stat_soilCN, echo=F, results="asis"}
+#display
 names(stattab)<-c("Statistic","Value on data","Frac. K surr >","Frac. K surr <","Frac. S surr >","Frac. S surr <")

### 2018-02-28 SupportingInformation.Rmd
filled in nonparam stat portion

@@
 
-<!--Shya pls add a short description of what this chunk is for. Should it be in the setup chunk down below?-->
-```{r ,echo=FALSE}
+<!--Setup-->
+```{r setup, echo=F}
+knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
+
 library(captioner)
@@
 fig_nums <- captioner(prefix="Figure")
-```
 
-<!--Setup-->
-```{r setup, echo=F}
-knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
 seed<-101
@@
 
-The raw data look like this:
-<!--Load the soil C and N data and make a plot-->
-```{r load_data_Loecke, echo=F, fig.cap="Plot of the raw soil C and N data.", fig.height=5, fig.width=5}
+<!--The raw data look like this:-->
+```{r load_data_Loecke, echo=F, fig.cap="plot of the raw soil C and N data.", fig.height=5, fig.width=5}
 d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
@@
 
-The copula looks like this:
-<!--Plot the copula of the soil C and N data--> 
-```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of the soil C and N data",  fig.height=5, fig.width=5}
+<!--The copula looks like this:-->
+```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data",  fig.height=5, fig.width=5}
 #convert to a copula and plot
@@
 
+The raw data looks like : `r fig_nums("soilCNraw",display="c")`.
+
+The copula looks like : `r fig_nums("soilCNcop",display="c")`.
+
+\newpage
 ## Aphid data
@@
 <!--Shya pls add a description of each of the statistics implemented in CopulaFunctions.R. As usual, please describe the stats with sufficient precision that they could be re-coded by the reader.-->
+#### stat_CorlCoru
+```{r stat_CorlCoru_D2lD2u, echo=F}
+fig_nums("stat_CorlCoru_D2lD2u",display=F)
+```
+![schematic diagram for stat_CorlCoru and stat_D2lD2u](./CorlCoru_D2lD2u.jpg){ width=30%, height=30% }
+
+If we draw a ranking based copula plot of a dataset within a unit square box then the points are scattered within the box according to their joint-marginal distribution.
+Say, there are total $n$ number of datapoints and $vi$ and $vj$ are the ranks of the dataset taken at [i,j] location pair or for [i,j] observables.
+To measure their tail dependence, i.
+e., the points are concentrated in the left corner or right corner of the box, we will calculate our statistics on left and right side of $vi+vj=1$ line [see solid blue line in `r fig_nums("stat_CorlCoru_D2lD2u",display="c")`].
+Now, we define the total covariance of $(vi,vj)$ as;
+
+\begin{equation}\label{eq.CorlCoru}
+   C = \frac{1}{n-1}\sum_{k=1}^{n}\left [ vi(k)-\left<vi\right> \right ] \left[ vj(k)-\left<vi\right> \right ]
+\end{equation}
+
+where, $\left<vi\right>$ and $\left<vj\right>$ are the mean of $vi$ and $vj$ of the points within the box respectively.
+If we think the total covariance $C$ as sum of covariance of the points lie within $vi+vj<1$ region, namely, $Corl$ and for $vi+vj>1$ region, namely, $Coru$; then $C = Corl + Coru$.
+So, $Corl$ and $Coru$ can be defined by;
+
+\begin{equation}\label{eq.Corl}
+   Corl = \frac{1}{n-1}\sum_{k \epsilon (vi+vj<1)}\left [ vi(k)-\left<vi\right> \right ] \left[ vj(k)-\left<vj\right> \right ]
+\end{equation}
+
+and
+
+\begin{equation}\label{eq.Coru}
+   Coru = \frac{1}{n-1}\sum_{k \epsilon (vi+vj>1)}\left [ vi(k)-\left<vi\right> \right ] \left[ vj(k)-\left<vj\right> \right ]
+\end{equation}
+
+If we take their difference then that will tetll us which end of the copula is with stronger correlation.
+A copula with $(Corl-Coru)>0$ should have more lower tail dependence whereas $(Corl-Coru)<0$ indicates more upper tail dependence.
+Symmetric copula has $Corl = Coru$.
+
+#### stat_PlPu
+
+```{r stat_PlPu, echo=F}
+fig_nums("stat_PlPu",display=F)
+```
+![schematic diagram for stat_PlPu](./PlPu.jpg){ width=35%, height=35% }
+
+To start with, we consider the lower triangular region of  `r fig_nums("stat_PlPu",display="c")` for which $vi+vj<1$.
+First, we take $Sl_{p}$, the ratio of number of points within the shaded pentagon and number of points lie on the left side of $vi+vj=1$ line.
+This $Sl_{p}$ is a function of distance-parameter $h$ which has range from $0$ to $\sqrt2/2$ (= half of the length of the box-diagonal).
+Similarly, we calculate $Su_{p}(h)$ for region $vi+vj>1$.
+Then we considered the independent case as if the points within the box scattered uniformly.
+$Si_{p}$ represents the ratio of the pentagon and the area of right angled isosceles triangle of unit side.
+Now, we want to measure that how much $Sl_{p}$ and $Su_{p}$ deviate from the standard $Si_{p}$ line when they are plotted against $h$.
+Their deviations $Pl$ and $Pu$, representing our statistics to measure lower and upper tail dependence, are given by;
+
+\begin{equation}\label{eq.Pl}
+   Pl = \int_0^{\sqrt2/2} \left[ Sl_{p}(h) - Si_{p}(h) \right] \,dh
+\end{equation}
+
+\begin{equation}\label{eq.Pu}
+   Pu = \int_0^{\sqrt2/2} \left[ Su_{p}(h) - Si_{p}(h) \right] \,dh
+\end{equation}
+
+A copula having $(Pl-Pu)>0$ represents more lower tail dependence whereas a copula with negative $(Pl-Pu)$ represents more upper tail dependence.
+A copula with equal $Pl$ and $Pu$ value is considered as symmetric with no tail-dependence.
+
+#### stat_TlTu
+
+```{r stat_TlTu, echo=F}
+fig_nums("stat_TlTu",display=F)
+```
+![schematic diagram for stat_TlTu](./TlTu.jpg){ width=30%, height=30% }
+
+This statistic is an alternative kind of the previous statistics.
+Here, to measure tail dependence between two marginal distributions, we will calculate our statistics on left and right side of $vi+vj=1$ line [see `r fig_nums("stat_TlTu",display="c")`] of the unit square box.
+First, we take $Il$, the ratio of number of points within the shaded lower triangle and the number of points lie on the left side of $vi+vj=1$ line.
+This $Il$ is a function of
+box-corner to the box-diagonal distance $h$ which has range from $0$ to $\sqrt2/2$ (= half of the length of the box-diagonal).
+Thus $Il$ also lies in between the same range [$0,1$].
+Similarly, we calculate $Iu(h)$ for region $vi+vj>1$.
+Usual independent standard $Ii(h)$ represents the ratio of one shaded triange and the area of right angled isosceles triangle of unit side.
+Now, we want to measure that how much $Il(h)$ and $Iu(h)$ deviate from the standard $Ii(h)$ line when they are plotted against $h$.
+Their deviations $Tl$ and $Tu$, representing our statistics to measure lower and upper tail dependence, are given by;
+
+\begin{equation}\label{eq.Tl}
+   Tl = \int_0^{\sqrt2/2} \left[ Il(h) - Ii(h) \right] \,dh
+\end{equation}
+
+\begin{equation}\label{eq.Tu}
+   Tu = \int_0^{\sqrt2/2} \left[Iu(h) - Ii(h) \right] \,dh
+\end{equation}
+
+According to this statistic a copula with more lower tail dependence should have positive $(Tl-Tu)$ value whereas a copula with negative $(Tl-Tu)$ represents more upper tail dependence.
+A copula with equal $Tl$ and $Tu$ value is considered as a member of symmetric copula family with no tail-dependence.
+
+#### stat_D2lD2u
+
+In this statistic as usual we divide the unit square box into two region based on either side of $vi+vj=1$ line.
+For the lower triangular region, we calculate squared distance of each scattered point from $vi-vj=0$ [the dotted line in `r fig_nums("stat_CorlCoru_D2lD2u",display="c")`] and then take their average.
+This quantity (say, $D2l$) represents how much the points in the lower tail of a copula are condensed to the identity line.
+Similarly, we calculate $D2u$ for the upper triangular region.
+If for a given copula $(D2u-D2l)>0$, then it is with more lower tail dependence as the upper-end points are distantly spaced from the long axis of the copula.
+Conversely, $(D2u-D2l)<0$ means a copula with more upper tail dependence.
+If $D2l$ and $D2u$ are equal, then the data come from a symmetric copula.
 
@@
 stat<-unlist(sCN)
-stat<-c(stat[1:4],unname(stat[3]-stat[4]),stat[5:6],unname(stat[5]-stat[6]),stat[7:8],unname(stat[7]-stat[8]),stat[9:10],unname(stat[9]-stat[10]))
-names(stat)[c(5,8,11,14)]<-c("Corl_m_Coru","Pl_m_Pu","Tl_m_Tu","D2l_m_D2u")
+stat<-c(stat[1:4],unname(stat[3]-stat[4]),stat[5:6],unname(stat[5]-stat[6]),stat[7:8],unname(stat[7]-stat[8]),stat[9:10],unname(stat[10]-stat[9]))
+names(stat)[c(5,8,11,14)]<-c("Corl-Coru","Pl-Pu","Tl-Tu","D2u-D2l")
 
@@
   h<-unlist(h[c(2:5,7,8,10:13)])
-  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[9]-h[10])
+  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[10]-h[9])
   statsurr_k[,counter]<-unname(h)
@@
   h<-unlist(h[c(2:5,7,8,10:13)])
-  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[9]-h[10])
+  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[10]-h[9])
   statsurr_s[,counter]<-unname(h)
@@
 #display
-names(stattab)<-c("Statistic","Value on data","Frac. K surr >","Frac. K surr <","Frac. S surr >","Frac. S surr <")
+names(stattab)<-c("Statistic","Value on data","Frac_K_sur_gt","Frac_K_sur_lt","Frac_S_sur_gt","Frac_S_sur_lt")
 table_nums("table_stat_soilCN",display=F)
-knitr::kable(data.frame(stat=stattab),caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset.
+knitr::kable(data.frame(stattab),caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset.
 Names and values on the actual C and N data are listed in the first two columns.
@@
 ```
+
+\newpage
 
@@
 ```
-![Plot for aphid-count : Spearman vs. Distance](./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf)
+![Plot for aphid-count : Spearman vs. Distance](./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Plot for aphid-count : Kendall vs. Distance](./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf)
+![Plot for aphid-count : Kendall vs. Distance](./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Corl-Coru vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf)
+![Corl-Coru vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Pl-Pu vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf)
+![Pl-Pu vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Tl-Tu vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Tl-Tu_vs_D.pdf)
+![Tl-Tu vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Tl-Tu_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![D2u-D2l vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf)
+![D2u-D2l vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![D2u/D2l vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_D2u_by_D2l_vs_D.pdf)
+![D2u/D2l vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_D2u_by_D2l_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Plot for aphid-firstflight : Spearman vs. Distance](./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf)
+![Plot for aphid-firstflight : Spearman vs. Distance](./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Plot for aphid-firstflight : Kendall vs. Distance](./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf)
+![Plot for aphid-firstflight : Kendall vs. Distance](./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Corl-Coru vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf)
+![Corl-Coru vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Pl-Pu vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf)
+![Pl-Pu vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Tl-Tu vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Tl-Tu_vs_D.pdf)
+![Tl-Tu vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Tl-Tu_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![D2u-D2l vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf)
+![D2u-D2l vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![D2u/D2l vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_D2u_by_D2l_vs_D.pdf)
+![D2u/D2l vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_D2u_by_D2l_vs_D.pdf){ width=50%, height=50% }
 
@@
 
+\newpage
+
 ```{r pns_bestcop,echo=F,results="asis"}
@@
 #pandoc.table(RES_plankton_north_sea$info_ord_AICw[,,1], caption="Plankton_north_sea : AICw for best fit copula family for each selective location pair")
-cat("\n\n\\pagebreak\n")
+#cat("\n\n\\pagebreak\n")
 
@@
 
+\newpage
+
 ```{r pns_LTdep,echo=F,results="asis"}
-cat("\n\n\\pagebreak\n")
+#cat("\n\n\\pagebreak\n")
 
@@
 
+\newpage
+
 ```{r pns_LTdep_AICw,echo=F,results="asis"}
-cat("\n\n\\pagebreak\n")
+#cat("\n\n\\pagebreak\n")
 
@@
 
+\newpage
+
 ```{r pns_LTmUT,echo=F,results="asis"}
-cat("\n\n\\pagebreak\n")
+#cat("\n\n\\pagebreak\n")
 
@@
 
+\newpage
+
 ```{r pns_p_KS,echo=F,results="asis"}
-cat("\n\n\\pagebreak\n")
+#cat("\n\n\\pagebreak\n")
 
@@
 ```
-![Plot for Plankton_north_sea : Spearman vs. Distance](./Results/stat_results/stat_plankton_north_sea/Sp_16_Spearman_vs_D.pdf)
+![Plot for Plankton_north_sea : Spearman vs. Distance](./Results/stat_results/stat_plankton_north_sea/Sp_16_Spearman_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Plot for Plankton_north_sea : Kendall vs. Distance](./Results/stat_results/stat_plankton_north_sea/Sp_16_Kendall_vs_D.pdf)
+![Plot for Plankton_north_sea : Kendall vs. Distance](./Results/stat_results/stat_plankton_north_sea/Sp_16_Kendall_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Corl-Coru vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf)
+![Corl-Coru vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Pl-Pu vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf)
+![Pl-Pu vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![Tl-Tu vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Tl-Tu_vs_D.pdf)
+![Tl-Tu vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Tl-Tu_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![D2u-D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf)
+![D2u-D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
 
@@
 ```
-![D2u/D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u_by_D2l_vs_D.pdf)
+![D2u/D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u_by_D2l_vs_D.pdf){ width=50%, height=50% }
 

### 2018-03-01 SupportingInformation.Rmd
added outline of script for flexible nonparematric tail dependence stats

@@
 table_nums("table_stat_soilCN",display=F)
-knitr::kable(data.frame(stattab),caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset.
+knitr::kable(stattab,caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset.
 Names and values on the actual C and N data are listed in the first two columns.

### 2018-03-01 CopulaFunctions_flexible.R

@@ -1,0 +1,57 @@
+#Calculates the mean squared distance between copula
+#points between two bounds and the main diagonal of
+#the unit square.
+#
+#Args
+#vi, vj       Coordinates of points from a copula
+#lb, ub       Lower and upper bounds between 0 and 1
+#
+#Output
+#The mean squared distance.
+#
+#Examples
+#The earlier function D2lD2u should give the combined
+#results of D2(vi,vj,0,.5) and D2(vi,vj,0.5,1).
+#
+D2bds<-function(vi,vj,lb,ub)
+{
+  inds<-which(vi+vj>2*lb & vi+vj<2*ub)
+
+  if(length(inds)!=0)
+  { 
+    dsq<-0.5*(vi[inds]-vj[inds])^2
+    D2<-sum(dsq)/length(dsq)
+  }else
+  {
+    D2<-NA
+  }
+  
+  return(D2)
+}
+
+#Dan to write an analogous Cor function
+
+#Args
+#vi, vj       Coordinates of points from a copula
+#lb, ub       Lower and upper bounds between 0 and 1
+#
+#Output
+#
+
+Corbds<-function(vi,vj,lb,lu)
+{
+  
+}
+
+#Shya please write an analogous P function, fill in below.
+
+#Args
+#vi, vj       Coordinates of points from a copula
+#lb, ub       Lower and upper bounds between 0 and 1
+#
+#Output
+#
+Pbds<-function(vi,vj,lb,lu)
+{
+  #returns a single number, computes pretty fast
+}

### 2018-03-03 SupportingInformation.Rmd
changes in some chunks for caching

@@
 source("copsurrog2d.R")
+source("CopulaFunctions_flexible.R")
 ```
@@
 
-#### stat_TlTu
-
-```{r stat_TlTu, echo=F}
-fig_nums("stat_TlTu",display=F)
-```
-![schematic diagram for stat_TlTu](./TlTu.jpg){ width=30%, height=30% }
-
-This statistic is an alternative kind of the previous statistics.
-Here, to measure tail dependence between two marginal distributions, we will calculate our statistics on left and right side of $vi+vj=1$ line [see `r fig_nums("stat_TlTu",display="c")`] of the unit square box.
-First, we take $Il$, the ratio of number of points within the shaded lower triangle and the number of points lie on the left side of $vi+vj=1$ line.
-This $Il$ is a function of
-box-corner to the box-diagonal distance $h$ which has range from $0$ to $\sqrt2/2$ (= half of the length of the box-diagonal).
-Thus $Il$ also lies in between the same range [$0,1$].
-Similarly, we calculate $Iu(h)$ for region $vi+vj>1$.
-Usual independent standard $Ii(h)$ represents the ratio of one shaded triange and the area of right angled isosceles triangle of unit side.
-Now, we want to measure that how much $Il(h)$ and $Iu(h)$ deviate from the standard $Ii(h)$ line when they are plotted against $h$.
-Their deviations $Tl$ and $Tu$, representing our statistics to measure lower and upper tail dependence, are given by;
-
-\begin{equation}\label{eq.Tl}
-   Tl = \int_0^{\sqrt2/2} \left[ Il(h) - Ii(h) \right] \,dh
-\end{equation}
-
-\begin{equation}\label{eq.Tu}
-   Tu = \int_0^{\sqrt2/2} \left[Iu(h) - Ii(h) \right] \,dh
-\end{equation}
-
-According to this statistic a copula with more lower tail dependence should have positive $(Tl-Tu)$ value whereas a copula with negative $(Tl-Tu)$ represents more upper tail dependence.
-A copula with equal $Tl$ and $Tu$ value is considered as a member of symmetric copula family with no tail-dependence.
-
 #### stat_D2lD2u
@@
 <!--Do the intensive part of the analysis-->
-```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v,mtime("CopulaFunctions.R"),mtime("copsurrog2d.R"),mtime("NonParamStat.R"))}
+```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v,mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
 set.seed(seed) 
@@
 
-#get the stats for the real data
-stat_soilCN<-copsync(v) 
+stat_soilCN<-stat_flexible_fn(m=v,lb=0,ub=0.1)
 saveRDS(stat_soilCN,paste(resloc,file="stat_soilCN.RDS",sep=''))
-sCN<-stat_soilCN[c(2:5,7,8,10:13)]
-stat<-unlist(sCN)
-stat<-c(stat[1:4],unname(stat[3]-stat[4]),stat[5:6],unname(stat[5]-stat[6]),stat[7:8],unname(stat[7]-stat[8]),stat[9:10],unname(stat[10]-stat[9]))
-names(stat)[c(5,8,11,14)]<-c("Corl-Coru","Pl-Pu","Tl-Tu","D2u-D2l")
+
+stat<-unlist(stat_soilCN)
+stat<-c(stat[1:4],unname(stat[3]-stat[4]),stat[5:6],unname(stat[5]-stat[6]),stat[7:8],unname(stat[8]-stat[7]))
+names(stat)[c(5,8,11)]<-c("Corl-Coru","Pl-Pu","D2u-D2l")
 
@@
 {
-  h<-copsync(surv[,,counter])
-  h<-unlist(h[c(2:5,7,8,10:13)])
-  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[10]-h[9])
+  h<-stat_flexible_fn(m=surv[,,counter],lb=0,ub=0.1)
+  h<-unlist(h)
+  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[8]-h[7])
   statsurr_k[,counter]<-unname(h)
@@
 {
-  h<-copsync(surv[,,counter])
-  h<-unlist(h[c(2:5,7,8,10:13)])
-  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[10]-h[9])
+  h<-stat_flexible_fn(m=surv[,,counter],lb=0,ub=0.1)
+  h<-unlist(h)
+  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[8]-h[7])
   statsurr_s[,counter]<-unname(h)
@@
 
-```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count)}
+```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"),mtime("NonParamStat.R"))}
 set.seed(seed) 
@@
 
-```{r count_Tl-Tu,echo=F}
-fig_nums("count_Tl-Tu",display=F)
-```
-![Tl-Tu vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Tl-Tu_vs_D.pdf){ width=50%, height=50% }
-
 ```{r count_D2u-D2l,echo=F}
@@
 ![D2u-D2l vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
-
-```{r count_D2u/D2l,echo=F}
-fig_nums("count_D2u/D2l",display=F)
-```
-![D2u/D2l vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_D2u_by_D2l_vs_D.pdf){ width=50%, height=50% }
 
@@
 
-```{r stat_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff)}
+```{r stat_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"),mtime("NonParamStat.R"))}
+
 set.seed(seed)
@@
 
-```{r ff_Tl-Tu,echo=F}
-fig_nums("ff_Tl-Tu",display=F)
-```
-![Tl-Tu vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Tl-Tu_vs_D.pdf){ width=50%, height=50% }
-
 ```{r ff_D2u-D2l,echo=F}
@@
 ![D2u-D2l vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
-
-```{r ff_D2u/D2l,echo=F}
-fig_nums("ff_D2u/D2l",display=F)
-```
-![D2u/D2l vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_D2u_by_D2l_vs_D.pdf){ width=50%, height=50% }
 
@@
 ### Non-paramtric analysis of tail dependence
-```{r stat_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea)}
+```{r stat_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"),mtime("NonParamStat.R"))}
+
 set.seed(seed)
@@
 
-```{r pns_Tl-Tu,echo=F}
-fig_nums("pns_Tl-Tu",display=F)
-```
-![Tl-Tu vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Tl-Tu_vs_D.pdf){ width=50%, height=50% }
-
 ```{r pns_D2u-D2l,echo=F}
@@
 
-```{r pns_D2u/D2l,echo=F}
-fig_nums("pns_D2u/D2l",display=F)
-```
-![D2u/D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u_by_D2l_vs_D.pdf){ width=50%, height=50% }
-
 # Question 2 analyses for each dataset

### 2018-03-03 SupportingInformation.Rmd
additions to copula functions and application to C/N data

--- 2018-03-01 SupportingInformation.Rmd
@@
 
-<!--Setup-->
+<!--Basic setup-->
 ```{r setup, echo=F}
@@
 families<-c(1,3:10,13,14,16:20)
+
 #A function needed for caching
@@
 }
+
 source("getcopula.R")
@@
 ## Soil C and N data
-Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description).
 
@@
 fig_nums("soilCNcop",display=F)
-v<-getcopula(d=d,rankon=T,ploton=T) #with real data, rankon should be T
+v_CN<-getcopula(d=d,rankon=T,ploton=T) #with real data, rankon should be T
 ```
 
-The raw data looks like : `r fig_nums("soilCNraw",display="c")`.
+Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description).
+The raw data are shown in `r fig_nums("soilCNraw",display="c")`.
+The copula is shown in `r fig_nums("soilCNcop",display="c")`.
 
-The copula looks like : `r fig_nums("soilCNcop",display="c")`.
+<!--Shyamolina, let's not put in newpages and other spacing modifications until possibly at the end, once all the 
+content is essentially settled-->
 
-\newpage
 ## Aphid data
@@
 
-<!--Shya pls add a description of each of the statistics implemented in CopulaFunctions.R. As usual, please describe the stats with sufficient precision that they could be re-coded by the reader.-->
 #### stat_CorlCoru
@@
 ### Model selection approach
-```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v,mtime("OurBiCopSelect.R"))}
+```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"))}
 set.seed(seed)
 families_Loecke<-c(1,3:10,13:14,16:20) 
-BivMS_res_Loecke<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
+BivMS_res_Loecke<-OurBiCopSelect(u1=v_CN[,1],u2=v_CN[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
 ```
@@
 <!--Do the intensive part of the analysis-->
-```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v,mtime("CopulaFunctions.R"),mtime("copsurrog2d.R"),mtime("NonParamStat.R"))}
+```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
 set.seed(seed)
-resloc<-"./Results/stat_results/stat_soilCN/"
 
-#get the stats for the real data
-stat_soilCN<-copsync(v) 
-saveRDS(stat_soilCN,paste(resloc,file="stat_soilCN.RDS",sep=''))
-sCN<-stat_soilCN[c(2:5,7,8,10:13)]
-stat<-unlist(sCN)
-stat<-c(stat[1:4],unname(stat[3]-stat[4]),stat[5:6],unname(stat[5]-stat[6]),stat[7:8],unname(stat[7]-stat[8]),stat[9:10],unname(stat[10]-stat[9]))
-names(stat)[c(5,8,11,14)]<-c("Corl-Coru","Pl-Pu","Tl-Tu","D2u-D2l")
+#initialize
+npa_res_CN<-data.frame(Statistic=c("Kendall","Spearman","Corl","Coru","CorlmCoru","Corlt0p1","Corgt0p9","Corlt0p1mCorgt0p9","D2l","D2u","D2umD2l","D2lt0p1","D2gt0p9","D2lt0p9mD2gt0p1"),Value=NA,Frac_of_K_surr_gt_val=NA,Frac_of_S_surr_gt_val=NA)
 
-#get surrogates and get stats for them, preserving kendall
+#utility function
+calcstats<-function(v)
+{
+  h1<-Corbds(v[,1],v[,2],0,.5)
+  h2<-Corbds(v[,1],v[,2],0.5,1)
+  h3<-Corbds(v[,1],v[,2],0,.1)
+  h4<-Corbds(v[,1],v[,2],0.9,1)
+  h5<-D2bds(v[,1],v[,2],0,.5)
+  h6<-D2bds(v[,1],v[,2],0.5,1)
+  h7<-D2bds(v[,1],v[,2],0,.1)
+  h8<-D2bds(v[,1],v[,2],0.9,1)
+  return(c(cor(v[,1],v[,2],method="kendall"),
+           cor(v[,1],v[,2],method="spearman"),
+           h1,h2,h1-h2,
+           h3,h4,h3-h4,
+           h5,h6,h6-h5,
+           h7,h8,h8-h7
+  ))
+}
+
+#get value on the real data
+npa_res_CN[,"Value"]<-calcstats(v_CN)
+
+#Do the kendall surrogates
 cop<-normalCopula(.5,2)
 numsurrog<-1000
-surv<-copsurrog2d(v,cop,"kendall",numsurrog)
-statsurr_k<-matrix(NA,length(stat),numsurrog)
-rownames(statsurr_k)<-names(stat)
-for (counter in 1:dim(surv)[3])
+surv<-copsurrog2d(v_CN,cop,"kendall",numsurrog)
+res<-apply(FUN=calcstats,X=surv,MARGIN=3)
+for (counter in 1:dim(npa_res_CN)[1])
 {
-  h<-copsync(surv[,,counter])
-  h<-unlist(h[c(2:5,7,8,10:13)])
-  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[10]-h[9])
-  statsurr_k[,counter]<-unname(h)
+  npa_res_CN[counter,"Frac_of_K_surr_gt_val"]<-sum(res[counter,]>npa_res_CN[counter,"Value"])/numsurrog
 }
 
-#same, preserving spearman
-surv<-copsurrog2d(v,cop,"spearman",numsurrog)
-statsurr_s<-matrix(NA,length(stat),numsurrog)
-rownames(statsurr_s)<-names(stat)
-for (counter in 1:dim(surv)[3])
+#Do the spearman surrogates
+surv<-copsurrog2d(v_CN,cop,"spearman",numsurrog)
+res<-apply(FUN=calcstats,X=surv,MARGIN=3)
+for (counter in 1:dim(npa_res_CN)[1])
 {
-  h<-copsync(surv[,,counter])
-  h<-unlist(h[c(2:5,7,8,10:13)])
-  h<-c(h[1:4],h[3]-h[4],h[5:6],h[5]-h[6],h[7:8],h[7]-h[8],h[9:10],h[10]-h[9])
-  statsurr_s[,counter]<-unname(h)
+  npa_res_CN[counter,"Frac_of_S_surr_gt_val"]<-sum(res[counter,]>npa_res_CN[counter,"Value"])/numsurrog
 }
@@
 ```{r table_stat_soilCN, echo=F, results="asis"}
-#now get fractions of surrogates lt the data and gt the data for each stat
-stattab<-data.frame(statname=names(stat),stat=unname(stat),frac_surr_gt_k=NA,frac_surr_lt_k=NA,frac_surr_gt_s=NA,frac_surr_lt_s=NA)
-for (counter in 1:length(stat))
-{
-  stattab[counter,3]<-sum(statsurr_k[counter,]>stat[counter])/numsurrog
-  stattab[counter,4]<-sum(statsurr_k[counter,]<stat[counter])/numsurrog
-  stattab[counter,5]<-sum(statsurr_s[counter,]>stat[counter])/numsurrog
-  stattab[counter,6]<-sum(statsurr_s[counter,]<stat[counter])/numsurrog
-}
-
-#display
-names(stattab)<-c("Statistic","Value on data","Frac_K_sur_gt","Frac_K_sur_lt","Frac_S_sur_gt","Frac_S_sur_lt")
 table_nums("table_stat_soilCN",display=F)
-knitr::kable(stattab,caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset.
-Names and values on the actual C and N data are listed in the first two columns.
-The third (respectively, fourth) column is the fraction of Kendall-preserving surrogate datasets for which the given statistic is greater than (resp., less than) its value on the real data.
-The last two columns are analogous but for Spearman-preserving surrogates.")
+knitr::kable(npa_res_CN,caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset.
+Names of statistics and values on the actual C and N data are listed in the first two columns.
+The third (respectively, fourth) column is the fraction of Kendall-preserving (respectively, Spearman-preserving) surrogate datasets for which the given statistic is greater than its value on the real data.
+So if these quantities are smaller than 0.025 the statistic is significantly larger than expected based on null hypotheses of normal copulas; if larger than 0.975, significantly smaller.")
 ```

### 2018-03-03 CopulaFunctions_flexible.R

@@ -44,7 +44,7 @@ D2bds<-function(vi,vj,lb,ub)
 #The earlier function CorlCoru should give the combined
 #results of Corbds(vi,vj,0,.5) and Corbds(vi,vj,0.5,1).
 #
-Corbds<-function(vi,vj,lb,lu)
+Corbds<-function(vi,vj,lb,ub)
 {
   #get mean and variance
   vi_mean<-mean(vi)

### 2018-03-03 SupportingInformation.Rmd
added P functionality to C/N analysis

@@
 
-<!--Shyamolina, let's not put in newpages and other spacing modifications until possibly at the end, once all the 
-content is essentially settled-->
+<!--Shyamolina, let's not put in newpages and other spacing modifications until possibly at the end, once all the content is essentially settled-->
 
@@
 #initialize
-npa_res_CN<-data.frame(Statistic=c("Kendall","Spearman","Corl","Coru","CorlmCoru","Corlt0p1","Corgt0p9","Corlt0p1mCorgt0p9","D2l","D2u","D2umD2l","D2lt0p1","D2gt0p9","D2lt0p9mD2gt0p1"),Value=NA,Frac_of_K_surr_gt_val=NA,Frac_of_S_surr_gt_val=NA)
+npa_res_CN<-data.frame(Statistic=c("Kendall","Spearman",
+  "Corl","Coru","CorlmCoru","Corlt0p1","Corgt0p9","Corlt0p1mCorgt0p9",
+  "D2l","D2u","D2umD2l","D2lt0p1","D2gt0p9","D2lt0p9mD2gt0p1",
+  "Pl","Pu","PlmPu","Plt0p1","Pgt0p9","Plt0p1mPgt0p9"),
+  Value=NA,Frac_of_K_surr_gt_val=NA,Frac_of_S_surr_gt_val=NA)
 
@@
   h8<-D2bds(v[,1],v[,2],0.9,1)
+  h9<-Pbds(v[,1],v[,2],0,0.5)$abs_res
+  h10<-Pbds(v[,1],v[,2],0.5,1)$abs_res
+  h11<-Pbds(v[,1],v[,2],0,0.1)$abs_res
+  h12<-Pbds(v[,1],v[,2],0.9,1)$abs_res
   return(c(cor(v[,1],v[,2],method="kendall"),
@@
            h5,h6,h6-h5,
-           h7,h8,h8-h7
+           h7,h8,h8-h7,
+           h9,h10,h9-h10,
+           h11,h12,h11-h12
   ))
@@
 The third (respectively, fourth) column is the fraction of Kendall-preserving (respectively, Spearman-preserving) surrogate datasets for which the given statistic is greater than its value on the real data.
-So if these quantities are smaller than 0.025 the statistic is significantly larger than expected based on null hypotheses of normal copulas; if larger than 0.975, significantly smaller.")
+So if these quantities are smaller than 0.025 the statistic is significantly larger than expected based on a null hypotheses of normal copulas; if larger than 0.975, significantly smaller.")
 ```
+
+<!--Now need a chunk that calculated D2 for lb=0.1*n, ub=0.1*(n+1) for n=0,...,9, and get 
+a bunch of surrogates and calculates the same quantities and then makes a bunch of box-whisker
+plots and compares the real data to those for each of these lb/up limits. -->
 

### 2018-03-03 CopulaFunctions_flexible.R

@@ -287,9 +287,3 @@ stat_flexible_fn<-function(m,lb,ub){
 
 
 
-
-
-
-
-
->>>>>>> 0bb5df05a95de19346d23c80c1f76298ac06b307

### 2018-03-05 SupportingInformation.Rmd
added the main computational work for an analysis of C and N data that looks at deviations from a normal-copula null for multiple locations from the lower to upper tail

--- 2018-03-03 SupportingInformation.Rmd
@@
 <!--Do the intensive part of the analysis-->
-```{r stat_soilCN, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+```{r stat_soilCN_graphic_approach, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
 set.seed(seed)
 
-#initialize
-npa_res_CN<-data.frame(Statistic=c("Kendall","Spearman","Corl","Coru","CorlmCoru","Corlt0p1","Corgt0p9","Corlt0p1mCorgt0p9","D2l","D2u","D2umD2l","D2lt0p1","D2gt0p9","D2lt0p9mD2gt0p1"),Value=NA,Frac_of_K_surr_gt_val=NA,Frac_of_S_surr_gt_val=NA)
-
 #utility function
-calcstats<-function(v)
+calcstats<-function(v,f,nm)
 {
-  h1<-Corbds(v[,1],v[,2],0,.5)
-  h2<-Corbds(v[,1],v[,2],0.5,1)
-  h3<-Corbds(v[,1],v[,2],0,.1)
-  h4<-Corbds(v[,1],v[,2],0.9,1)
-  h5<-D2bds(v[,1],v[,2],0,.5)
-  h6<-D2bds(v[,1],v[,2],0.5,1)
-  h7<-D2bds(v[,1],v[,2],0,.1)
-  h8<-D2bds(v[,1],v[,2],0.9,1)
-  return(c(cor(v[,1],v[,2],method="kendall"),
-           cor(v[,1],v[,2],method="spearman"),
-           h1,h2,h1-h2,
-           h3,h4,h3-h4,
-           h5,h6,h6-h5,
-           h7,h8,h8-h7
-  ))
+  #calculate the stats
+  res<-c(f(v[,1],v[,2],0,.5),
+          f(v[,1],v[,2],0.5,1),
+          f(v[,1],v[,2],0,.25),
+          f(v[,1],v[,2],0.25,.5),
+          f(v[,1],v[,2],.5,.75),
+          f(v[,1],v[,2],.75,1))
+  for (lbd in seq(from=0,to=0.9,by=0.1))
+  {
+    res<-c(res,f(v[,1],v[,2],lbd,lbd+.1))
   }
 
-#get value on the real data
-npa_res_CN[,"Value"]<-calcstats(v_CN)
+  #name the result vector
+  h<-paste(nm,c(0,0.5,0,0.25,0.5,0.75,seq(from=0,to=0.9,by=0.1)),"to",
+                    c(0.5,1,0.25,0.5,0.75,1,seq(from=0,to=0.9,by=0.1)+.1),sep='')
+  h<-gsub(".","p",h,fixed=T)
+  names(res)<-h
+
+  return(res)
+}
 
@@
 numsurrog<-1000
-surv<-copsurrog2d(v_CN,cop,"kendall",numsurrog)
-res<-apply(FUN=calcstats,X=surv,MARGIN=3)
-for (counter in 1:dim(npa_res_CN)[1])
-{
-  npa_res_CN[counter,"Frac_of_K_surr_gt_val"]<-sum(res[counter,]>npa_res_CN[counter,"Value"])/numsurrog
-}
+surv_K<-copsurrog2d(v_CN,cop,"kendall",numsurrog)
 
 #Do the spearman surrogates
-surv<-copsurrog2d(v_CN,cop,"spearman",numsurrog)
-res<-apply(FUN=calcstats,X=surv,MARGIN=3)
-for (counter in 1:dim(npa_res_CN)[1])
+surv_S<-copsurrog2d(v_CN,cop,"spearman",numsurrog)
+
+#now work with the correlation stats
+corstats_d<-calcstats(v_CN,Corbds,"Cor")
+corstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Corbds,nm="Cor")
+corstats_K<-apply(FUN=quantile,X=corstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
+corstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Corbds,nm="Cor")
+corstats_S<-apply(FUN=quantile,X=corstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
+
+#now work with the D2 stats
+D2stats_d<-calcstats(v_CN,D2bds,"Dtwo")
+D2stats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=D2bds,nm="Dtwo")
+D2stats_K<-apply(FUN=quantile,X=D2stats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
+D2stats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=D2bds,nm="Dtwo")
+D2stats_S<-apply(FUN=quantile,X=D2stats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
+
+#now work with the P stats
+Pbds_wrap<-function(vi,vj,lb,ub)
 {
-  npa_res_CN[counter,"Frac_of_S_surr_gt_val"]<-sum(res[counter,]>npa_res_CN[counter,"Value"])/numsurrog
+  return(Pbds(vi,vj,lb,ub)$abs_res)
 }
-```
-
-<!--Display the table generated above-->
-```{r table_stat_soilCN, echo=F, results="asis"}
-table_nums("table_stat_soilCN",display=F)
-knitr::kable(npa_res_CN,caption = "Results of non-parametric tests of tail dependence and asymmetry of tail dependence for the soil C and N dataset.
-Names of statistics and values on the actual C and N data are listed in the first two columns.
-The third (respectively, fourth) column is the fraction of Kendall-preserving (respectively, Spearman-preserving) surrogate datasets for which the given statistic is greater than its value on the real data.
-So if these quantities are smaller than 0.025 the statistic is significantly larger than expected based on null hypotheses of normal copulas; if larger than 0.975, significantly smaller.")
+Pstats_d<-calcstats(v_CN,Pbds_wrap,"P")
+Pstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Pbds_wrap,nm="P")
+Pstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Pbds_wrap,nm="P")
 ```

### 2018-03-05 SupportingInformation.Rmd
just a comment on what to do next
completed but have not yet fully tested the changes that do a graphical version of the nonparametric tail dependence stuff for C and N data

@@
 
-#utility function
+#utility function for calculating all the stats we want
 calcstats<-function(v,f,nm)
@@
   #calculate the stats
-  res<-c(f(v[,1],v[,2],0,.5),
-          f(v[,1],v[,2],0.5,1),
-          f(v[,1],v[,2],0,.25),
-          f(v[,1],v[,2],0.25,.5),
-          f(v[,1],v[,2],.5,.75),
-          f(v[,1],v[,2],.75,1))
+  res<-c()
+  #f(v[,1],v[,2],0,.5),
+  #        f(v[,1],v[,2],0.5,1),
+  #        f(v[,1],v[,2],0,.25),
+  #        f(v[,1],v[,2],0.25,.5),
+  #        f(v[,1],v[,2],.5,.75),
+  #        f(v[,1],v[,2],.75,1))
   for (lbd in seq(from=0,to=0.9,by=0.1))
@@
   #name the result vector
-  h<-paste(nm,c(0,0.5,0,0.25,0.5,0.75,seq(from=0,to=0.9,by=0.1)),"to",
-                    c(0.5,1,0.25,0.5,0.75,1,seq(from=0,to=0.9,by=0.1)+.1),sep='')
+#  h<-paste(nm,c(0,0.5,0,0.25,0.5,0.75,seq(from=0,to=0.9,by=0.1)),"to",
+#                    c(0.5,1,0.25,0.5,0.75,1,seq(from=0,to=0.9,by=0.1)+.1),sep='')
+  h<-paste(nm,c(seq(from=0,to=0.9,by=0.1)),"to",
+                    c(seq(from=0,to=0.9,by=0.1)+.1),sep='')
   h<-gsub(".","p",h,fixed=T)
@@
 
+#utility function for calculating fractions of surrogates 
+#with a stat smaller than that of data, i.e., fraction of 
+#surrogates for which the stat value on the data is bigger
+fracwork<-function(dvals,surrvals)
+{
+  frac<-NA*numeric(length(corstats_d))
+  names(frac)<-names(dvals)
+  for (counter in 1:length(frac))
+  {
+    frac[counter]<-sum(surrvals[counter,]<dvals[counter])
+  }
+  return(frac)
+}
+
 #now work with the correlation stats
 corstats_d<-calcstats(v_CN,Corbds,"Cor")
+
 corstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Corbds,nm="Cor")
+corstats_frac_K<-fracwork(corstats_d,corstats_K)
 corstats_K<-apply(FUN=quantile,X=corstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
+
 corstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Corbds,nm="Cor")
+corstats_frac_S<-fracwork(corstats_d,corstats_S)
 corstats_S<-apply(FUN=quantile,X=corstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
@@
 D2stats_d<-calcstats(v_CN,D2bds,"Dtwo")
+
 D2stats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=D2bds,nm="Dtwo")
+D2stats_frac_K<-fracwork(D2stats_d,D2stats_K)
 D2stats_K<-apply(FUN=quantile,X=D2stats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
+
 D2stats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=D2bds,nm="Dtwo")
+D2stats_frac_S<-fracwork(D2stats_d,D2stats_S)
 D2stats_S<-apply(FUN=quantile,X=D2stats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
@@
 }
+
 Pstats_d<-calcstats(v_CN,Pbds_wrap,"P")
+
 Pstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Pbds_wrap,nm="P")
+Pstats_frac_K<-fracwork(Pstats_d,Pstats_K)
+Pstats_K<-apply(FUN=quantile,X=Pstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
+
 Pstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Pbds_wrap,nm="P")
+Pstats_frac_S<-fracwork(Pstats_d,Pstats_S)
+Pstats_S<-apply(FUN=quantile,X=Pstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
+```
+
+<!--Now do the plotting-->
+```{r plot_soilCN_graphic_approach, echo=F}
+x<-seq(from=0.05,to=0.95,by=0.1)
+xlimits<-c(0,1)
+
+#plotting layout, units inches
+xht<-0.5
+ywd<-0.5
+titleht<-.25
+panht<-1.5
+panwd<-3
+gap<-.05
+totwd<-ywd+2*panwd+2*gap
+totht<-xht+3*panht+3*gap+titleht
+pdf(file="CNdata_NonparamTailDep.pdf",width=totwd,height=totht)
+
+#plot kendall results on the left panels, spearman on right,
+#cor in top panels, then D2, then P on bottom panels
+
+#kendall, cor
+par(fig=c(ywd/totwd,
+          (ywd+panwd)/totwd,
+          (xht+2*panht+2*gap)/totht,
+          (xht+3*panht+2*gap)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
+ylimits_cor<-range(corstats_d,corstats_K,corstats_K)
+ylimits_cor[2]<-ylimits_cor[2]+.3*diff(ylimits_cor)
+plot(x,corstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_cor,
+     xaxt='n')
+mtext(side=3,line=0,text="Kendall-preserving surrogates")
+mtext(side=2,line=1,text="Partial correlation")
+axis(side=1,labels=F)
+lines(x,corstats_K[1,],type='l',lty='dotted')
+lines(x,corstats_K[2,],type='l',lty='dashed')
+lines(x,corstats_K[3,],type='l',lty='dashed')
+lines(x,corstats_K[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_cor[1],labels='A',cex=1.5,adj=c(.5,0))
+fracplot<-function(fracs,ylims)
+{
+  for (counter in 1:length(fracs))
+  {
+    ptxt<-''
+    if (fracs[counter]>.975*numsurrog)
+    {
+      ptxt<-paste(">",fracs[counter],sep='')
+    } 
+    if (fracs[counter]<.025*numsurrog)
+    {
+      ptxt<-paste("<",numsurrog-fracs[counter],sep='')
+    }
+    yht<-ylims[2]-.1*diff(ylims)
+    text(x[counter],yht,ptxt,adj=c(0.5,.5),cex=0.5,srt=90)
+  }
+}
+fracplot(corstats_frac_K,ylimits_cor)
+vlineplot<-function(x,ylims)
+{
+  for (counter in 1:length(x))
+  {
+    lines(rep(x[counter]-(x[2]-x[1])/2,2),ylims,type='l',lty='dotted')
+  }
+  lines(rep(x[length(x)]+(x[2]-x[1])/2,2),ylims,type='l',lty='dotted')
+}
+vlineplot(x,ylimits_cor)
+
+#spearman, cor
+par(fig=c((ywd+panwd+gap)/totwd,
+          (ywd+2*panwd+gap)/totwd,
+          (xht+2*panht+2*gap)/totht,
+          (xht+3*panht+2*gap)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
+plot(x,corstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_cor,
+     xaxt='n',yaxt='n')
+mtext(side=3,line=0,text="Spearman-preserving surrogates")
+axis(side=1,labels=F)
+axis(side=2,labels=F)
+lines(x,corstats_S[1,],type='l',lty='dotted')
+lines(x,corstats_S[2,],type='l',lty='dashed')
+lines(x,corstats_S[3,],type='l',lty='dashed')
+lines(x,corstats_S[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_cor[1],labels='B',cex=1.5,adj=c(.5,0))
+fracplot(corstats_frac_S,ylimits_cor)
+vlineplot(x,ylimits_cor)
+
+#kendall, D2
+par(fig=c(ywd/totwd,
+          (ywd+panwd)/totwd,
+          (xht+panht+gap)/totht,
+          (xht+2*panht+gap)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
+ylimits_D2<-range(D2stats_d,D2stats_K,D2stats_K)
+ylimits_D2[2]<-ylimits_D2[2]+.3*diff(ylimits_D2)
+plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2,
+     xaxt='n')
+mtext(side=2,line=1,text=expression(D[2]))
+axis(side=1,labels=F)
+lines(x,D2stats_K[1,],type='l',lty='dotted')
+lines(x,D2stats_K[2,],type='l',lty='dashed')
+lines(x,D2stats_K[3,],type='l',lty='dashed')
+lines(x,D2stats_K[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_D2[1],labels='C',cex=1.5,adj=c(.5,0))
+fracplot(D2stats_frac_K,ylimits_D2)
+vlineplot(x,ylimits_D2)
+
+#spearman, D2
+par(fig=c((ywd+panwd+gap)/totwd,
+          (ywd+2*panwd+gap)/totwd,
+          (xht+panht+gap)/totht,
+          (xht+2*panht+gap)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
+plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2,
+     xaxt='n',yaxt='n')
+axis(side=1,labels=F)
+axis(side=2,labels=F)
+lines(x,D2stats_S[1,],type='l',lty='dotted')
+lines(x,D2stats_S[2,],type='l',lty='dashed')
+lines(x,D2stats_S[3,],type='l',lty='dashed')
+lines(x,D2stats_S[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_D2[1],labels='D',cex=1.5,adj=c(.5,0))
+fracplot(D2stats_frac_S,ylimits_D2)
+vlineplot(x,ylimits_D2)
+
+#kendall, P
+par(fig=c(ywd/totwd,
+          (ywd+panwd)/totwd,
+          (xht)/totht,
+          (xht+panht)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
+ylimits_P<-range(Pstats_d,Pstats_K,Pstats_K)
+ylimits_P[2]<-ylimits_P[2]+.3*diff(ylimits_P)
+plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P)
+mtext(side=2,line=1,text="P")
+mtext(side=1,line=1,text="Diagonal slice")
+lines(x,Pstats_K[1,],type='l',lty='dotted')
+lines(x,Pstats_K[2,],type='l',lty='dashed')
+lines(x,Pstats_K[3,],type='l',lty='dashed')
+lines(x,Pstats_K[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_P[1],labels='E',cex=1.5,adj=c(.5,0))
+fracplot(Pstats_frac_K,ylimits_P)
+vlineplot(x,ylimits_P)
+
+#spearman, P
+par(fig=c((ywd+panwd+gap)/totwd,
+          (ywd+2*panwd+gap)/totwd,
+          (xht)/totht,
+          (xht+panht)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
+plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,
+     yaxt='n')
+mtext(side=1,line=1,text="Diagonal slice")
+axis(side=2,labels=F)
+lines(x,Pstats_S[1,],type='l',lty='dotted')
+lines(x,Pstats_S[2,],type='l',lty='dashed')
+lines(x,Pstats_S[3,],type='l',lty='dashed')
+lines(x,Pstats_S[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_P[1],labels='F',cex=1.5,adj=c(.5,0))
+fracplot(Pstats_frac_S,ylimits_P)
+vlineplot(x,ylimits_P)
+
+dev.off()
+
 ```

### 2018-03-05 SupportingInformation.Rmd
added some comments on things Shya can work on moving forward

@@
 # Question 1 analyses for each dataset
-Recall that question 1 is: Do datasets in ecology and related fields have non-Gausian copula srtructure?
+Recall that question 1 is: Do datasets in ecology and related fields have non-Gausian copula structure?
 Do they show tail dependence distinct from that of a Gaussian copula?
@@
 
-### Non-paramtric analysis of tail dependence
+### Non-parametric analysis of tail dependence
+
+<!--Shya to edit this description of the stats. Don't need subsections for each stat.-->
+<!--Eventually we will need to add methods on how we use these stats, which is different for multivariate and bivariate data-->
+<!--Eventually, for the multivariate data, we will need to demonstrate using a similar analysis to what is in mbsyn, that you can detect asymmetry of tail dependence using these stats - we have done this, but it needs to be adapted-->
 
@@
 dev.off()
-
+#DAN: have to make the plot actually appear in the Rmd output, with a caption and everything!
 ```
@@
 # Question 3 analyses for each dataset
+
+Recall question 3 is: What are the consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?
+
+## Aphid count
+
+<!--Shya to add results about skewness of spatial means-->
+
+## Aphid first flight
+
+<!--Shya to add results about skewness of spatial means-->
+
+## North Sea plankton
+
+<!--Shya to add results about skewness of spatial means-->

### 2018-03-06 SupportingInformation.Rmd
in the midst of adding asymmetry of tail dependence results to the C/N analysis

@@
 
-### Non-paramtric analysis of tail dependence
-<!--Do the intensive part of the analysis-->
+### Non-parametric analysis of tail dependence
+<!--Do the intensive part of the analysis, calculating stats on the data and surrogates-->
 ```{r stat_soilCN_graphic_approach, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
@@
 {
-  frac<-NA*numeric(length(corstats_d))
+  frac<-NA*numeric(length(dvals))
   names(frac)<-names(dvals)
@@
 corstats_d<-calcstats(v_CN,Corbds,"Cor")
-
 corstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Corbds,nm="Cor")
 corstats_frac_K<-fracwork(corstats_d,corstats_K)
+
+corlmcoru_d<-unname(corstats_d[1]-corstats_d[length(corstats_d)])
+corlmcoru_K<-corstats_K[,1]-corstats_K[,numsurrog]
+corlmcoru_frac_K<-sum(corlmcoru_K<corlmcoru_d)
 corstats_K<-apply(FUN=quantile,X=corstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
@@
 corstats_frac_S<-fracwork(corstats_d,corstats_S)
+
+corlmcoru_S<-corstats_S[,1]-corstats_S[,numsurrog]
+corlmcoru_frac_S<-sum(corlmcoru_S<corlmcoru_d)
 corstats_S<-apply(FUN=quantile,X=corstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
+
+#now work with the P stats
+Pbds_wrap<-function(vi,vj,lb,ub)
+{
+  return(Pbds(vi,vj,lb,ub)$abs_res)
+}
+
+Pstats_d<-calcstats(v_CN,Pbds_wrap,"P")
+Pstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Pbds_wrap,nm="P")
+Pstats_frac_K<-fracwork(Pstats_d,Pstats_K)
+
+PlmPu_d<-unname(Pstats_d[1]-Pstats_d[length(Pstats_d)])
+PlmPu_K<-Pstats_K[,1]-Pstats_K[,numsurrog]
+PlmPu_frac_K<-sum(PlmPu_K<PlmPu_d)
+Pstats_K<-apply(FUN=quantile,X=Pstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
+
+Pstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Pbds_wrap,nm="P")
+Pstats_frac_S<-fracwork(Pstats_d,Pstats_S)
+
+PlmPu_S<-Pstats_S[,1]-Pstats_S[,numsurrog]
+PlmPu_frac_S<-sum(PlmPu_S<PlmPu_d)
+Pstats_S<-apply(FUN=quantile,X=Pstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
 
@@
 D2stats_d<-calcstats(v_CN,D2bds,"Dtwo")
-
 D2stats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=D2bds,nm="Dtwo")
 D2stats_frac_K<-fracwork(D2stats_d,D2stats_K)
+
+D2lmD2u_d<-D2stats_d[1]-D2stats_d[length(D2stats_d)]
+D2lmD2u_K<-D2stats_K[,1]-D2stats_K[,numsurrog]
+D2lmD2u_frac_K<-sum(D2lmD2u_K<D2lmD2u_d)
 D2stats_K<-apply(FUN=quantile,X=D2stats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
@@
 D2stats_frac_S<-fracwork(D2stats_d,D2stats_S)
+
+D2lmD2u_S<-D2stats_S[,1]-D2stats_S[,numsurrog]
+D2lmD2u_frac_S<-sum(D2lmD2u_S<D2lmD2u_d)
 D2stats_S<-apply(FUN=quantile,X=D2stats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
-
-#now work with the P stats
-Pbds_wrap<-function(vi,vj,lb,ub)
-{
-  return(Pbds(vi,vj,lb,ub)$abs_res)
-}
-
-Pstats_d<-calcstats(v_CN,Pbds_wrap,"P")
-
-Pstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Pbds_wrap,nm="P")
-Pstats_frac_K<-fracwork(Pstats_d,Pstats_K)
-Pstats_K<-apply(FUN=quantile,X=Pstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
-
-Pstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Pbds_wrap,nm="P")
-Pstats_frac_S<-fracwork(Pstats_d,Pstats_S)
-Pstats_S<-apply(FUN=quantile,X=Pstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
-```
-
-<!--Now do the plotting-->
+```
+
+<!--Now save the plot to the results folder-->
 ```{r plot_soilCN_graphic_approach, echo=F}
@@
 totht<-xht+3*panht+3*gap+titleht
-pdf(file="CNdata_NonparamTailDep.pdf",width=totwd,height=totht)
+pdf(file="./Results/stat_results/stat_soilCN//CNdata_NonparamTailDep.pdf",width=totwd,height=totht)
 
 #plot kendall results on the left panels, spearman on right,
-#cor in top panels, then D2, then P on bottom panels
+#cor in top panels, then P, then D2 on bottom panels
 
@@
 
-#kendall, D2
+#kendall, P
 par(fig=c(ywd/totwd,
@@
 ylimits_D2[2]<-ylimits_D2[2]+.3*diff(ylimits_D2)
-plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2,
-     xaxt='n')
+ylimits_P<-range(Pstats_d,Pstats_K,Pstats_K)
+ylimits_P[2]<-ylimits_P[2]+.3*diff(ylimits_P)
+plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,xaxt='n')
+axis(side=1,labels=F)
+mtext(side=2,line=1,text="P")
+lines(x,Pstats_K[1,],type='l',lty='dotted')
+lines(x,Pstats_K[2,],type='l',lty='dashed')
+lines(x,Pstats_K[3,],type='l',lty='dashed')
+lines(x,Pstats_K[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_P[1],labels='C',cex=1.5,adj=c(.5,0))
+fracplot(Pstats_frac_K,ylimits_P)
+vlineplot(x,ylimits_P)
+
+#spearman, p
+par(fig=c((ywd+panwd+gap)/totwd,
+          (ywd+2*panwd+gap)/totwd,
+          (xht+panht+gap)/totht,
+          (xht+2*panht+gap)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
+plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,
+     xaxt='n',yaxt='n')
+axis(side=1,labels=F)
+axis(side=2,labels=F)
+lines(x,Pstats_S[1,],type='l',lty='dotted')
+lines(x,Pstats_S[2,],type='l',lty='dashed')
+lines(x,Pstats_S[3,],type='l',lty='dashed')
+lines(x,Pstats_S[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_P[1],labels='D',cex=1.5,adj=c(.5,0))
+fracplot(Pstats_frac_S,ylimits_P)
+vlineplot(x,ylimits_P)
+
+#kendall, D2
+par(fig=c(ywd/totwd,
+          (ywd+panwd)/totwd,
+          (xht)/totht,
+          (xht+panht)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
+plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2)
+mtext(side=1,line=1,text="Diagonal slice")
 mtext(side=2,line=1,text=expression(D[2]))
@@
 lines(x,D2stats_K[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_D2[1],labels='C',cex=1.5,adj=c(.5,0))
+text(xlimits[1],ylimits_D2[1],labels='E',cex=1.5,adj=c(.5,0))
 fracplot(D2stats_frac_K,ylimits_D2)
@@
 
-#spearman, D2
+#spearman, P
 par(fig=c((ywd+panwd+gap)/totwd,
           (ywd+2*panwd+gap)/totwd,
-          (xht+panht+gap)/totht,
-          (xht+2*panht+gap)/totht),
+          (xht)/totht,
+          (xht+panht)/totht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
 plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2,
-     xaxt='n',yaxt='n')
+     yaxt='n')
+mtext(side=1,line=1,text="Diagonal slice")
 axis(side=1,labels=F)
@@
 lines(x,D2stats_S[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_D2[1],labels='D',cex=1.5,adj=c(.5,0))
+text(xlimits[1],ylimits_D2[1],labels='F',cex=1.5,adj=c(.5,0))
 fracplot(D2stats_frac_S,ylimits_D2)
@@
 
+dev.off()
+```
+
+<!--Now include the plot in the Rmd output-->
+```{r show_plot_soildCN,echo=F}
+fig_nums("soilCN_nonparam",display=F)
+```
+![Nonparametric tests for tail dependence and other deviations from a normal copula for soil C and N data.
+As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
+When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than.](./Results/stat_results/stat_soilCN//CNdata_NonparamTailDep.pdf)
+
+<!--Now save a second plot to the results folder, this one about asymmetry of tail dependence-->
+```{r plot_soilCN_graphic_approach_2, echo=F}
+#plotting layout, units inches
+xht<-0.5
+ywd<-0.5
+titleht<-.25
+panht<-1.25
+panwd<-3
+gap<-.05
+totwd<-2*ywd+2*panwd+2*gap
+totht<-3*xht+3*panht+3*gap+titleht
+pdf(file="./Results/stat_results/stat_soilCN//CNdata_AsymmTailDep.pdf",width=totwd,height=totht)
+
+#plot kendall results on the left panels, spearman on right,
+#cor in top panels, then P, then D2 on bottom panels
+
+#kendall, cor
+par(fig=c(ywd/totwd,
+          (ywd+panwd)/totwd,
+          (3*xht+2*panht+2*gap)/totht,
+          (3*xht+3*panht+2*gap)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
+hist(corlmcoru_K,50)
+points(corlmcoru_d,0,col='red')
+mtext(side=3,line=0,text="Kendall-preserving surrogates")
+mtext(side=1,line=1,text="Partial correlation")
+#text(xlimits[1],ylimits_cor[1],labels='A',cex=1.5,adj=c(.5,0))
+fracplot<-function(fracs,ylims)
+{
+  for (counter in 1:length(fracs))
+  {
+    ptxt<-''
+    if (fracs[counter]>.975*numsurrog)
+    {
+      ptxt<-paste(">",fracs[counter],sep='')
+    } 
+    if (fracs[counter]<.025*numsurrog)
+    {
+      ptxt<-paste("<",numsurrog-fracs[counter],sep='')
+    }
+    yht<-ylims[2]-.1*diff(ylims)
+    text(x[counter],yht,ptxt,adj=c(0.5,.5),cex=0.5,srt=90)
+  }
+}
+fracplot(corstats_frac_K,ylimits_cor)
+vlineplot<-function(x,ylims)
+{
+  for (counter in 1:length(x))
+  {
+    lines(rep(x[counter]-(x[2]-x[1])/2,2),ylims,type='l',lty='dotted')
+  }
+  lines(rep(x[length(x)]+(x[2]-x[1])/2,2),ylims,type='l',lty='dotted')
+}
+vlineplot(x,ylimits_cor)
+
+#spearman, cor
+par(fig=c((ywd+panwd+gap)/totwd,
+          (ywd+2*panwd+gap)/totwd,
+          (xht+2*panht+2*gap)/totht,
+          (xht+3*panht+2*gap)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
+plot(x,corstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_cor,
+     xaxt='n',yaxt='n')
+mtext(side=3,line=0,text="Spearman-preserving surrogates")
+axis(side=1,labels=F)
+axis(side=2,labels=F)
+lines(x,corstats_S[1,],type='l',lty='dotted')
+lines(x,corstats_S[2,],type='l',lty='dashed')
+lines(x,corstats_S[3,],type='l',lty='dashed')
+lines(x,corstats_S[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_cor[1],labels='B',cex=1.5,adj=c(.5,0))
+fracplot(corstats_frac_S,ylimits_cor)
+vlineplot(x,ylimits_cor)
+
 #kendall, P
+par(fig=c(ywd/totwd,
+          (ywd+panwd)/totwd,
+          (xht+panht+gap)/totht,
+          (xht+2*panht+gap)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
+ylimits_D2<-range(D2stats_d,D2stats_K,D2stats_K)
+ylimits_D2[2]<-ylimits_D2[2]+.3*diff(ylimits_D2)
+ylimits_P<-range(Pstats_d,Pstats_K,Pstats_K)
+ylimits_P[2]<-ylimits_P[2]+.3*diff(ylimits_P)
+plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,xaxt='n')
+axis(side=1,labels=F)
+mtext(side=2,line=1,text="P")
+lines(x,Pstats_K[1,],type='l',lty='dotted')
+lines(x,Pstats_K[2,],type='l',lty='dashed')
+lines(x,Pstats_K[3,],type='l',lty='dashed')
+lines(x,Pstats_K[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_P[1],labels='C',cex=1.5,adj=c(.5,0))
+fracplot(Pstats_frac_K,ylimits_P)
+vlineplot(x,ylimits_P)
+
+#spearman, p
+par(fig=c((ywd+panwd+gap)/totwd,
+          (ywd+2*panwd+gap)/totwd,
+          (xht+panht+gap)/totht,
+          (xht+2*panht+gap)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
+plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,
+     xaxt='n',yaxt='n')
+axis(side=1,labels=F)
+axis(side=2,labels=F)
+lines(x,Pstats_S[1,],type='l',lty='dotted')
+lines(x,Pstats_S[2,],type='l',lty='dashed')
+lines(x,Pstats_S[3,],type='l',lty='dashed')
+lines(x,Pstats_S[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_P[1],labels='D',cex=1.5,adj=c(.5,0))
+fracplot(Pstats_frac_S,ylimits_P)
+vlineplot(x,ylimits_P)
+
+#kendall, D2
 par(fig=c(ywd/totwd,
@@
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-ylimits_P<-range(Pstats_d,Pstats_K,Pstats_K)
-ylimits_P[2]<-ylimits_P[2]+.3*diff(ylimits_P)
-plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P)
-mtext(side=2,line=1,text="P")
+plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2)
 mtext(side=1,line=1,text="Diagonal slice")
-lines(x,Pstats_K[1,],type='l',lty='dotted')
-lines(x,Pstats_K[2,],type='l',lty='dashed')
-lines(x,Pstats_K[3,],type='l',lty='dashed')
-lines(x,Pstats_K[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_P[1],labels='E',cex=1.5,adj=c(.5,0))
-fracplot(Pstats_frac_K,ylimits_P)
-vlineplot(x,ylimits_P)
+mtext(side=2,line=1,text=expression(D[2]))
+axis(side=1,labels=F)
+lines(x,D2stats_K[1,],type='l',lty='dotted')
+lines(x,D2stats_K[2,],type='l',lty='dashed')
+lines(x,D2stats_K[3,],type='l',lty='dashed')
+lines(x,D2stats_K[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_D2[1],labels='E',cex=1.5,adj=c(.5,0))
+fracplot(D2stats_frac_K,ylimits_D2)
+vlineplot(x,ylimits_D2)
 
@@
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,
+plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2,
      yaxt='n')
 mtext(side=1,line=1,text="Diagonal slice")
+axis(side=1,labels=F)
 axis(side=2,labels=F)
-lines(x,Pstats_S[1,],type='l',lty='dotted')
-lines(x,Pstats_S[2,],type='l',lty='dashed')
-lines(x,Pstats_S[3,],type='l',lty='dashed')
-lines(x,Pstats_S[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_P[1],labels='F',cex=1.5,adj=c(.5,0))
-fracplot(Pstats_frac_S,ylimits_P)
-vlineplot(x,ylimits_P)
+lines(x,D2stats_S[1,],type='l',lty='dotted')
+lines(x,D2stats_S[2,],type='l',lty='dashed')
+lines(x,D2stats_S[3,],type='l',lty='dashed')
+lines(x,D2stats_S[4,],type='l',lty='dotted')
+text(xlimits[1],ylimits_D2[1],labels='F',cex=1.5,adj=c(.5,0))
+fracplot(D2stats_frac_S,ylimits_D2)
+vlineplot(x,ylimits_D2)
 
 dev.off()
-#DAN: have to make the plot actually appear in the Rmd output, with a caption and everything!
 ```

### 2018-03-06 SupportingInformation.Rmd
done for now implementing asymetry of tail dependence stuff for C/N data, though there are a couple small things to work out still

@@
 corlmcoru_d<-unname(corstats_d[1]-corstats_d[length(corstats_d)])
-corlmcoru_K<-corstats_K[,1]-corstats_K[,numsurrog]
+corlmcoru_K<-corstats_K[1,]-corstats_K[length(corstats_d),]
 corlmcoru_frac_K<-sum(corlmcoru_K<corlmcoru_d)
@@
 
-corlmcoru_S<-corstats_S[,1]-corstats_S[,numsurrog]
+corlmcoru_S<-corstats_S[1,]-corstats_S[length(corlmcoru_d),]
 corlmcoru_frac_S<-sum(corlmcoru_S<corlmcoru_d)
@@
 PlmPu_d<-unname(Pstats_d[1]-Pstats_d[length(Pstats_d)])
-PlmPu_K<-Pstats_K[,1]-Pstats_K[,numsurrog]
+PlmPu_K<-Pstats_K[1,]-Pstats_K[length(Pstats_d),]
 PlmPu_frac_K<-sum(PlmPu_K<PlmPu_d)
@@
 
-PlmPu_S<-Pstats_S[,1]-Pstats_S[,numsurrog]
+PlmPu_S<-Pstats_S[1,]-Pstats_S[length(Pstats_d),]
 PlmPu_frac_S<-sum(PlmPu_S<PlmPu_d)
@@
 
-D2lmD2u_d<-D2stats_d[1]-D2stats_d[length(D2stats_d)]
-D2lmD2u_K<-D2stats_K[,1]-D2stats_K[,numsurrog]
-D2lmD2u_frac_K<-sum(D2lmD2u_K<D2lmD2u_d)
+D2umD2l_d<-D2stats_d[length(D2stats_d)]-D2stats_d[1]
+D2umD2l_K<-D2stats_K[length(D2stats_d),]-D2stats_K[1,]
+D2umD2l_frac_K<-sum(D2umD2l_K<D2umD2l_d)
 D2stats_K<-apply(FUN=quantile,X=D2stats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
@@
 
-D2lmD2u_S<-D2stats_S[,1]-D2stats_S[,numsurrog]
-D2lmD2u_frac_S<-sum(D2lmD2u_S<D2lmD2u_d)
+D2umD2l_S<-D2stats_S[length(D2stats_d),]-D2stats_S[1,]
+D2umD2l_frac_S<-sum(D2umD2l_S<D2umD2l_d)
 D2stats_S<-apply(FUN=quantile,X=D2stats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
@@
 
-<!--Now save a second plot to the results folder, this one about asymmetry of tail dependence-->
-```{r plot_soilCN_graphic_approach_2, echo=F}
-#plotting layout, units inches
-xht<-0.5
-ywd<-0.5
-titleht<-.25
-panht<-1.25
-panwd<-3
-gap<-.05
-totwd<-2*ywd+2*panwd+2*gap
-totht<-3*xht+3*panht+3*gap+titleht
-pdf(file="./Results/stat_results/stat_soilCN//CNdata_AsymmTailDep.pdf",width=totwd,height=totht)
-
-#plot kendall results on the left panels, spearman on right,
-#cor in top panels, then P, then D2 on bottom panels
-
-#kendall, cor
-par(fig=c(ywd/totwd,
-          (ywd+panwd)/totwd,
-          (3*xht+2*panht+2*gap)/totht,
-          (3*xht+3*panht+2*gap)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
-hist(corlmcoru_K,50)
-points(corlmcoru_d,0,col='red')
-mtext(side=3,line=0,text="Kendall-preserving surrogates")
-mtext(side=1,line=1,text="Partial correlation")
-#text(xlimits[1],ylimits_cor[1],labels='A',cex=1.5,adj=c(.5,0))
-fracplot<-function(fracs,ylims)
-{
-  for (counter in 1:length(fracs))
+<!--Now make a table, this one about asymmetry of tail dependence-->
+```{r plot_soilCN_asymmtry_taildep, echo=F}
+atdres<-data.frame(Statistic=c("Corl-Coru","Pl-Pu","D2u-D2l"),
+                   Kendall=rep("",3),
+                   Spearman=rep("",3),stringsAsFactors = F)
+fractxt<-function(fracval)
 {
   ptxt<-''
-    if (fracs[counter]>.975*numsurrog)
-    {
-      ptxt<-paste(">",fracs[counter],sep='')
-    } 
-    if (fracs[counter]<.025*numsurrog)
-    {
-      ptxt<-paste("<",numsurrog-fracs[counter],sep='')
-    }
-    yht<-ylims[2]-.1*diff(ylims)
-    text(x[counter],yht,ptxt,adj=c(0.5,.5),cex=0.5,srt=90)
-  }
-}
-fracplot(corstats_frac_K,ylimits_cor)
-vlineplot<-function(x,ylims)
-{
-  for (counter in 1:length(x))
-  {
-    lines(rep(x[counter]-(x[2]-x[1])/2,2),ylims,type='l',lty='dotted')
-  }
-  lines(rep(x[length(x)]+(x[2]-x[1])/2,2),ylims,type='l',lty='dotted')
-}
-vlineplot(x,ylimits_cor)
-
-#spearman, cor
-par(fig=c((ywd+panwd+gap)/totwd,
-          (ywd+2*panwd+gap)/totwd,
-          (xht+2*panht+2*gap)/totht,
-          (xht+3*panht+2*gap)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-plot(x,corstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_cor,
-     xaxt='n',yaxt='n')
-mtext(side=3,line=0,text="Spearman-preserving surrogates")
-axis(side=1,labels=F)
-axis(side=2,labels=F)
-lines(x,corstats_S[1,],type='l',lty='dotted')
-lines(x,corstats_S[2,],type='l',lty='dashed')
-lines(x,corstats_S[3,],type='l',lty='dashed')
-lines(x,corstats_S[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_cor[1],labels='B',cex=1.5,adj=c(.5,0))
-fracplot(corstats_frac_S,ylimits_cor)
-vlineplot(x,ylimits_cor)
-
-#kendall, P
-par(fig=c(ywd/totwd,
-          (ywd+panwd)/totwd,
-          (xht+panht+gap)/totht,
-          (xht+2*panht+gap)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-ylimits_D2<-range(D2stats_d,D2stats_K,D2stats_K)
-ylimits_D2[2]<-ylimits_D2[2]+.3*diff(ylimits_D2)
-ylimits_P<-range(Pstats_d,Pstats_K,Pstats_K)
-ylimits_P[2]<-ylimits_P[2]+.3*diff(ylimits_P)
-plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,xaxt='n')
-axis(side=1,labels=F)
-mtext(side=2,line=1,text="P")
-lines(x,Pstats_K[1,],type='l',lty='dotted')
-lines(x,Pstats_K[2,],type='l',lty='dashed')
-lines(x,Pstats_K[3,],type='l',lty='dashed')
-lines(x,Pstats_K[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_P[1],labels='C',cex=1.5,adj=c(.5,0))
-fracplot(Pstats_frac_K,ylimits_P)
-vlineplot(x,ylimits_P)
-
-#spearman, p
-par(fig=c((ywd+panwd+gap)/totwd,
-          (ywd+2*panwd+gap)/totwd,
-          (xht+panht+gap)/totht,
-          (xht+2*panht+gap)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,
-     xaxt='n',yaxt='n')
-axis(side=1,labels=F)
-axis(side=2,labels=F)
-lines(x,Pstats_S[1,],type='l',lty='dotted')
-lines(x,Pstats_S[2,],type='l',lty='dashed')
-lines(x,Pstats_S[3,],type='l',lty='dashed')
-lines(x,Pstats_S[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_P[1],labels='D',cex=1.5,adj=c(.5,0))
-fracplot(Pstats_frac_S,ylimits_P)
-vlineplot(x,ylimits_P)
-
-#kendall, D2
-par(fig=c(ywd/totwd,
-          (ywd+panwd)/totwd,
-          (xht)/totht,
-          (xht+panht)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2)
-mtext(side=1,line=1,text="Diagonal slice")
-mtext(side=2,line=1,text=expression(D[2]))
-axis(side=1,labels=F)
-lines(x,D2stats_K[1,],type='l',lty='dotted')
-lines(x,D2stats_K[2,],type='l',lty='dashed')
-lines(x,D2stats_K[3,],type='l',lty='dashed')
-lines(x,D2stats_K[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_D2[1],labels='E',cex=1.5,adj=c(.5,0))
-fracplot(D2stats_frac_K,ylimits_D2)
-vlineplot(x,ylimits_D2)
-
-#spearman, P
-par(fig=c((ywd+panwd+gap)/totwd,
-          (ywd+2*panwd+gap)/totwd,
-          (xht)/totht,
-          (xht+panht)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2,
-     yaxt='n')
-mtext(side=1,line=1,text="Diagonal slice")
-axis(side=1,labels=F)
-axis(side=2,labels=F)
-lines(x,D2stats_S[1,],type='l',lty='dotted')
-lines(x,D2stats_S[2,],type='l',lty='dashed')
-lines(x,D2stats_S[3,],type='l',lty='dashed')
-lines(x,D2stats_S[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_D2[1],labels='F',cex=1.5,adj=c(.5,0))
-fracplot(D2stats_frac_S,ylimits_D2)
-vlineplot(x,ylimits_D2)
-
-dev.off()
+  if (fracval>.975*numsurrog)
+  {
+    ptxt<-paste(">",fracval,sep='')
+  } 
+  if (fracval<.025*numsurrog)
+  {
+    ptxt<-paste("<",numsurrog-fracval,sep='')
+  }
+  return(ptxt)
+}
+atdres[1,2]<-fractxt(corlmcoru_frac_K)
+atdres[1,3]<-fractxt(corlmcoru_frac_S)
+atdres[2,2]<-fractxt(PlmPu_frac_K)
+atdres[2,3]<-fractxt(PlmPu_frac_S)
+atdres[3,2]<-fractxt(D2umD2l_frac_K)
+atdres[3,3]<-fractxt(D2umD2l_frac_S)
+atdres
 ```

### 2018-03-06 SupportingInformation.Rmd
modified nonparam stat description

@@
 
-#### stat_CorlCoru
-```{r stat_CorlCoru_D2lD2u, echo=F}
-fig_nums("stat_CorlCoru_D2lD2u",display=F)
+```{r stat_image, echo=F}
+fig_nums("stat_image",display=F)
 ```
-![schematic diagram for stat_CorlCoru and stat_D2lD2u](./CorlCoru_D2lD2u.jpg){ width=30%, height=30% }
+![schematic diagram for npa of tail dependence](./stat_image.jpg){ width=30%, height=30% }
 
-If we draw a ranking based copula plot of a dataset within a unit square box then the points are scattered within the box according to their joint-marginal distribution.
-Say, there are total $n$ number of datapoints and $vi$ and $vj$ are the ranks of the dataset taken at [i,j] location pair or for [i,j] observables.
-To measure their tail dependence, i.
-e., the points are concentrated in the left corner or right corner of the box, we will calculate our statistics on left and right side of $vi+vj=1$ line [see solid blue line in `r fig_nums("stat_CorlCoru_D2lD2u",display="c")`].
-Now, we define the total covariance of $(vi,vj)$ as;
+In a given ranking based copula plot of a dataset within a unit square box the data points are scattered according to their joint-marginal distribution.
+Say, there are total $n$ number of data points and $vi$ and $vj$ are the ranks of the dataset scaled down by factor $(n+1)$ taken at $(i,j)$ location pair or for $(i,j)$ observables.
+Now, to calculate our statistic we choose an arbitrary region of the box bounded by two parallel lines [blue line : $vi+vj=2l_{b}$, red line : $vi+vj=2u_{b}$ in `r fig_nums("stat_image",display="c")`] where $(l_{b},u_{b})$ has a range $[0,1]$ with the condition that $u_{b}$ must be greater than $l_{b}$.
+Now we will demonstrate 3 different type of stats :
 
-\begin{equation}\label{eq.CorlCoru}
-   C = \frac{1}{n-1}\sum_{k=1}^{n}\left [ vi(k)-\left<vi\right> \right ] \left[ vj(k)-\left<vi\right> \right ]
+$Cor$, the correlation between the ${n}'$ number of points lying in between those two parallel lines, is given by;
+
+\begin{equation}\label{eq.Cor}
+   Cor = \frac{1}{n-1}\sum_{k=1}^{{n}'} \left [ vi(k)-\left<vi\right> \right ] \left[ vj(k)-\left<vj\right> \right ]
 \end{equation}
 
-where, $\left<vi\right>$ and $\left<vj\right>$ are the mean of $vi$ and $vj$ of the points within the box respectively.
-If we think the total covariance $C$ as sum of covariance of the points lie within $vi+vj<1$ region, namely, $Corl$ and for $vi+vj>1$ region, namely, $Coru$; then $C = Corl + Coru$.
-So, $Corl$ and $Coru$ can be defined by;
+where, $\left<vi\right>$ and $\left<vj\right>$ are the arithmatic mean of $vi$ and $vj$ of $n$ data points within the box, respectively.
 
-\begin{equation}\label{eq.Corl}
-   Corl = \frac{1}{n-1}\sum_{k \epsilon (vi+vj<1)}\left [ vi(k)-\left<vi\right> \right ] \left[ vj(k)-\left<vj\right> \right ]
+We can shift that pair of lines along the diagonal line of the box and can get a sense how the points are spread along the diagonal and how the quantity $Cor$ changes.
+
+For multivariate datasets due to limited number of data points this chosen region is broad, i.
+e., we measured $Cor$ for $(l_{b}=0,u_{b}=0.5)$ and $(l_{b}=0.5,u_{b}=1)$ and stored as $Corl$ and $Coru$ respectively.
+If we take their difference then that will tell us which end of the copula is with stronger correlation.
+A copula with $(Corl-Coru)>0$ should have more lower tail dependence whereas $(Corl-Coru)<0$ indicates more upper tail dependence.
+Symmetric copula, e.
+g., (Normal or Frannk) has $Corl = Coru$.
+
+On a second thought, we imagine as if we progress from the diagonal towards the sides of the box with $h$ gradually increasing from $0$ to $h_{max}$, where $h_{max}$ is the half-length of the longer one between two parallel lines.
+Then we take $S$, the ratio of number of points which are within $h$ distance from the diagonal on both side and the total number of points lie within the shaded region.
+Now, we want to measure that how much $S$ deviates from the standard $Si$ line when they are plotted against $h$ where $Si$ is the ratio of the area covered by distance $h$ on both side and the area of shaded region.
+This deviation $P$, representing our statistics to measure tail dependence, is given by;
+
+\begin{equation}\label{eq.P}
+   P = \int_0^{h_{max}} \left[ S(h) - Si(h) \right] \,dh
 \end{equation}
 
-and
-
-\begin{equation}\label{eq.Coru}
-   Coru = \frac{1}{n-1}\sum_{k \epsilon (vi+vj>1)}\left [ vi(k)-\left<vi\right> \right ] \left[ vj(k)-\left<vj\right> \right ]
-\end{equation}
-
-If we take their difference then that will tetll us which end of the copula is with stronger correlation.
-A copula with $(Corl-Coru)>0$ should have more lower tail dependence whereas $(Corl-Coru)<0$ indicates more upper tail dependence.
-Symmetric copula has $Corl = Coru$.
-
-#### stat_PlPu
-
-```{r stat_PlPu, echo=F}
-fig_nums("stat_PlPu",display=F)
-```
-![schematic diagram for stat_PlPu](./PlPu.jpg){ width=35%, height=35% }
-
-To start with, we consider the lower triangular region of  `r fig_nums("stat_PlPu",display="c")` for which $vi+vj<1$.
-First, we take $Sl_{p}$, the ratio of number of points within the shaded pentagon and number of points lie on the left side of $vi+vj=1$ line.
-This $Sl_{p}$ is a function of distance-parameter $h$ which has range from $0$ to $\sqrt2/2$ (= half of the length of the box-diagonal).
-Similarly, we calculate $Su_{p}(h)$ for region $vi+vj>1$.
-Then we considered the independent case as if the points within the box scattered uniformly.
-$Si_{p}$ represents the ratio of the pentagon and the area of right angled isosceles triangle of unit side.
-Now, we want to measure that how much $Sl_{p}$ and $Su_{p}$ deviate from the standard $Si_{p}$ line when they are plotted against $h$.
-Their deviations $Pl$ and $Pu$, representing our statistics to measure lower and upper tail dependence, are given by;
-
-\begin{equation}\label{eq.Pl}
-   Pl = \int_0^{\sqrt2/2} \left[ Sl_{p}(h) - Si_{p}(h) \right] \,dh
-\end{equation}
-
-\begin{equation}\label{eq.Pu}
-   Pu = \int_0^{\sqrt2/2} \left[ Su_{p}(h) - Si_{p}(h) \right] \,dh
-\end{equation}
-
+Like we said before, for multivariate dataset we measure $Pl = P$ for $(l_{b}=0, u_{b}=0.5)$ and $Pu = P$ for $(l_{b}=0.5, u_{b}=1)$ and these $Pl$, $Pu$ represent the lower and upper tail statistic respectively.
 A copula having $(Pl-Pu)>0$ represents more lower tail dependence whereas a copula with negative $(Pl-Pu)$ represents more upper tail dependence.
@@
 
-#### stat_TlTu
-
-```{r stat_TlTu, echo=F}
-fig_nums("stat_TlTu",display=F)
-```
-![schematic diagram for stat_TlTu](./TlTu.jpg){ width=30%, height=30% }
-
-This statistic is an alternative kind of the previous statistics.
-Here, to measure tail dependence between two marginal distributions, we will calculate our statistics on left and right side of $vi+vj=1$ line [see `r fig_nums("stat_TlTu",display="c")`] of the unit square box.
-First, we take $Il$, the ratio of number of points within the shaded lower triangle and the number of points lie on the left side of $vi+vj=1$ line.
-This $Il$ is a function of
-box-corner to the box-diagonal distance $h$ which has range from $0$ to $\sqrt2/2$ (= half of the length of the box-diagonal).
-Thus $Il$ also lies in between the same range [$0,1$].
-Similarly, we calculate $Iu(h)$ for region $vi+vj>1$.
-Usual independent standard $Ii(h)$ represents the ratio of one shaded triange and the area of right angled isosceles triangle of unit side.
-Now, we want to measure that how much $Il(h)$ and $Iu(h)$ deviate from the standard $Ii(h)$ line when they are plotted against $h$.
-Their deviations $Tl$ and $Tu$, representing our statistics to measure lower and upper tail dependence, are given by;
-
-\begin{equation}\label{eq.Tl}
-   Tl = \int_0^{\sqrt2/2} \left[ Il(h) - Ii(h) \right] \,dh
-\end{equation}
-
-\begin{equation}\label{eq.Tu}
-   Tu = \int_0^{\sqrt2/2} \left[Iu(h) - Ii(h) \right] \,dh
-\end{equation}
-
-According to this statistic a copula with more lower tail dependence should have positive $(Tl-Tu)$ value whereas a copula with negative $(Tl-Tu)$ represents more upper tail dependence.
-A copula with equal $Tl$ and $Tu$ value is considered as a member of symmetric copula family with no tail-dependence.
-
-#### stat_D2lD2u
-
-In this statistic as usual we divide the unit square box into two region based on either side of $vi+vj=1$ line.
-For the lower triangular region, we calculate squared distance of each scattered point from $vi-vj=0$ [the dotted line in `r fig_nums("stat_CorlCoru_D2lD2u",display="c")`] and then take their average.
-This quantity (say, $D2l$) represents how much the points in the lower tail of a copula are condensed to the identity line.
-Similarly, we calculate $D2u$ for the upper triangular region.
+For our third statistic, we imagine the diagonal line of the box as the major axis of the copula.
+We calculate squared distance of each scattered point from $vi-vj=0$ [the dotted diagonal line in `r fig_nums("stat_image",display="c")`] and then take their average.
+This mean squared distance (say, $D2$) represents the average width of the copula for that shaded region.
+Similarly, we calculate $D2l$ and $D2u$ for the lower and upper tail of a copula for multivariate dataset with $(l_{b}=0, u_{b}=0.5)$ and $(l_{b}=0.5, u_{b}=1)$ respectively.
 If for a given copula $(D2u-D2l)>0$, then it is with more lower tail dependence as the upper-end points are distantly spaced from the long axis of the copula.
@@
 ![D2u-D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
-
-```{r pns_D2u/D2l,echo=F}
-fig_nums("pns_D2u/D2l",display=F)
-```
-![D2u/D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u_by_D2l_vs_D.pdf){ width=50%, height=50% }
 

### 2018-03-06 Paper.Rmd
editorial changes

@@
 If that data has a certain tail dependence then the winners mostly come from either {**C, SG, SJ, SBB8**} or {**G, SC, J, BB8**} category indicating lower and upper tail dependence respectively.
-
+<!--Shya please add some information on model-averaged tail dependence statistics-->
 ----------------------------------------------------------------------

### 2018-03-06 SupportingInformation.Rmd
editorial changes

@@
 <!--The raw data look like this:-->
-```{r load_data_Loecke, echo=F, fig.cap="plot of the raw soil C and N data.", fig.height=5, fig.width=5}
+```{r load_data_Loecke, echo=F, fig.cap="Plot of the raw soil C and N data.", fig.height=5, fig.width=5}
 d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
@@
 <!--The copula looks like this:-->
-```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data",  fig.height=5, fig.width=5}
+```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data.",  fig.height=5, fig.width=5}
 #convert to a copula and plot
 fig_nums("soilCNcop",display=F)
-v_CN<-getcopula(d=d,rankon=T,ploton=T) #with real data, rankon should be T
+v_CN<-getcopula(d=d,rankon=T,ploton=T) 
 ```
@@
 ### Model selection approach
+<!--This does the computations for this section-->
 ```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"))}
@@
 ```
-The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes`.
+The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes`. We fitted several copulas
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (`r table_nums("soilCNfit",display="c")`).
+The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
 
-We fitted several copulas
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas:
+<!--Display model selection results in a table-->
 ```{r soilCNfit,echo=F}
@@
 
-`r table_nums("soilCNfit",display="c")` shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
-
-The model-averaged (using AIC) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged (again, AIC) upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`.
+These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
 Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.

### 2018-03-06 SupportingInformation.Rmd
more editorial changes

@@
   res<-c()
-  #f(v[,1],v[,2],0,.5),
-  #        f(v[,1],v[,2],0.5,1),
-  #        f(v[,1],v[,2],0,.25),
-  #        f(v[,1],v[,2],0.25,.5),
-  #        f(v[,1],v[,2],.5,.75),
-  #        f(v[,1],v[,2],.75,1))
   for (lbd in seq(from=0,to=0.9,by=0.1))
@@
   #name the result vector
-#  h<-paste(nm,c(0,0.5,0,0.25,0.5,0.75,seq(from=0,to=0.9,by=0.1)),"to",
-#                    c(0.5,1,0.25,0.5,0.75,1,seq(from=0,to=0.9,by=0.1)+.1),sep='')
   h<-paste(nm,c(seq(from=0,to=0.9,by=0.1)),"to",
@@
 
-<!--Now save the plot to the results folder-->
+<!--Now create the plot and save it to the results folder-->
 ```{r plot_soilCN_graphic_approach, echo=F}
@@
 totht<-xht+3*panht+3*gap+titleht
-pdf(file="./Results/stat_results/stat_soilCN//CNdata_NonparamTailDep.pdf",width=totwd,height=totht)
+pdf(file="./Results/stat_results/stat_soilCN/CNdata_NonparamTailDep.pdf",width=totwd,height=totht)
 
@@
 
-<!--Now include the plot in the Rmd output-->
+As described in Methods, each of our statistics, as computed for the 
+ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
+of values of the same statistic (with the same ranges) on 1000 normal-copula 
+surrogate datasets with similar Kendall or Spearman correlations, in different 
+runs (`r figure_nums("soilCN_nonparam",display="c")`). Because the correlation 
+used in the Cor statistic is the Spearman correlation, Spearman-preserving surrogates
+should be used, so the top-left panel of `r figure_nums("soilCN_nonparam",display="c")`
+should be ignored (DAN: need to fix this). Results indicate
+DAN: continue here.
+
+<!--Now give the plot a number so it can be refered to-->
 ```{r show_plot_soildCN,echo=F}
@@
 As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
-When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than.](./Results/stat_results/stat_soilCN//CNdata_NonparamTailDep.pdf)
+When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than.](./Results/stat_results/stat_soilCN/CNdata_NonparamTailDep.pdf)
 

### 2018-03-06 SupportingInformation.Rmd
minor error in fig calling

@@
 surrogate datasets with similar Kendall or Spearman correlations, in different 
-runs (`r figure_nums("soilCN_nonparam",display="c")`). Because the correlation 
+runs (`r fig_nums("soilCN_nonparam",display="c")`). Because the correlation 
 used in the Cor statistic is the Spearman correlation, Spearman-preserving surrogates
-should be used, so the top-left panel of `r figure_nums("soilCN_nonparam",display="c")`
+should be used, so the top-left panel of `r fig_nums("soilCN_nonparam",display="c")`
 should be ignored (DAN: need to fix this). Results indicate

### 2018-03-06 SupportingInformation.Rmd
finished adding narative for soil CN nonparametric tail dependence stuff, running now

@@
 ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
-of values of the same statistic (with the same ranges) on 1000 normal-copula 
+of values of the same statistic with the same ranges computed on 1000 normal-copula 
 surrogate datasets with similar Kendall or Spearman correlations, in different 
@@
 should be used, so the top-left panel of `r fig_nums("soilCN_nonparam",display="c")`
-should be ignored (DAN: need to fix this). Results indicate
-DAN: continue here.
+should be ignored (DAN: need to fix this later). Results indicate that tail dependence of 
+data is stronger in both the lower and upper tails than would be expected under a 
+null hypothesis of a normal copula. The dependence between C and N is also 
+weaker in the middle of both marginal distributions
+than would be expected based on a normal-copula null model.
 
@@
 
-<!--Now make a table, this one about asymmetry of tail dependence-->
+We also tested statistically for asymmetry of tail dependence, which seems visible in `r table_nums("soilCNcop",display="c")`. 
+Results (`r table_nums("soilCN_asym",display="c")`) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data.
+
+<!--Now make and display a table, this one about asymmetry of tail dependence-->
 ```{r plot_soilCN_asymmtry_taildep, echo=F}
@@
   ptxt<-''
-  if (fracval>.975*numsurrog)
+  if (fracval>.5*numsurrog)
   {
@@
   } 
-  if (fracval<.025*numsurrog)
+  if (fracval<=.5*numsurrog)
   {
@@
 atdres[3,3]<-fractxt(D2umD2l_frac_S)
-atdres
+
+table_nums("soilCN_asym",display=F)
+knitr::kable(atdres, caption = "Results for tests of whether asymmetry of tail dependence is significant compared to a normal-copula null hypothesis.
+Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the ranges $0-0.1$ and $0.9-1$ (Methods).
+One thousand Kendall-preserving and 1000 Spearman-preserving surrogates were created, and the same statistics were computed for the surrogates.
+A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates.")
 ```

### 2018-03-06 SupportingInformation.Rmd
adding skewness plot

@@
 <!--Shya to add results about skewness of spatial means-->
+```{r skewness_count_plot, echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
+set.seed(seed)
+pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf")
+par(mfrow=c(2,1))
+numsurrog<-10000
+loclist<-good_loclist(d_allsp=data_aphid_count,sp=10,data_pt_thrs=30)
+answer1k<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
+                          corpres="kendall")
+
+p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
+
+mtext(paste0("(A) result based on kendall preserving normal surrogates, p = ",p_left),side=3,line=0.2,col="navyblue")
+
+answer1s<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
+                          corpres="spearman")
+
+p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog
+
+mtext(paste0("(B) result based on spearman preserving normal surrogates , p = ",p_left),side=3,line=0.2,col="navyblue")
+
+dev.off()
+```
+
+```{r skewness_count,echo=F}
+fig_nums("skewness_count",display=F)
+```
+![Skewness_count for species : Green spruce aphid](./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf){ width=50%, height=50% }
 
@@
 <!--Shya to add results about skewness of spatial means-->
+```{r skewness_ff_plot, echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
+set.seed(seed)
+pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf")
+par(mfrow=c(2,1))
+numsurrog<-10000
+loclist<-good_loclist(d_allsp=data_aphid_ff,sp=11,data_pt_thrs=30)
+
+answer2k<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
+                          corpres="kendall") 
+p_right<-sum(answer2k$surrogskw > answer2k$realskw)/numsurrog
+mtext(paste0("(A) result based on kendall preserving normal surrogates, p = ",p_right),side=3,line=0.2,col="navyblue")
+
+answer2s<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
+                          corpres="spearman") 
+p_right<-sum(answer2s$surrogskw > answer2s$realskw)/numsurrog
+mtext(paste0("(B) result based on spearman preserving normal surrogates , p = ",p_right),side=3,line=0.2,col="navyblue")
+
+dev.off()
+```
+
+```{r skewness_ff,echo=F}
+fig_nums("skewness_ff",display=F)
+```
+![Skewness_firstflight for species : Leaf-curling plum aphid](./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf){ width=50%, height=50% }
 
@@
 <!--Shya to add results about skewness of spatial means-->
+<!--Shya to add results about skewness of spatial means-->
+```{r skewness_pns_plot, echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
+set.seed(seed)
+pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf")
+par(mfrow=c(2,1))
+numsurrog<-10000
+loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=16,data_pt_thrs=45)
+answer3k<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
+                          corpres="kendall")
+
+p_left<-sum(answer3k$surrogskw < answer3k$realskw)/numsurrog
+
+mtext(paste0("(A) result based on kendall preserving normal surrogates, p = ",p_left),side=3,line=0.2,col="navyblue")
+
+answer3s<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
+                          corpres="spearman")
+
+p_left<-sum(answer3s$surrogskw < answer3s$realskw)/numsurrog
+
+mtext(paste0("(B) result based on spearman preserving normal surrogates , p = ",p_left),side=3,line=0.2,col="navyblue")
+
+dev.off()
+```
+
+```{r skewness_pns,echo=F}
+fig_nums("skewness_pns",display=F)
+```
+![Skewness_plankton_north_sea for species : Ceratium_furca](./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf){ width=50%, height=50% }

### 2018-03-07 SupportingInformation.Rmd
adding stat_testing plot

@@
 source("copsurrog2d.R")
+source("TestStats_multivar_dataset.R")
 ```
@@
 If $D2l$ and $D2u$ are equal, then the data come from a symmetric copula.
+
+### Stat-testing of Non-parametric analysis (for limited number of datapoints)
+```{r stat_testing_plot, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"))}
+set.seed(seed)
+
+resultsloc<-"./Results/stat_results/stat_testing/"
+callfn<-coplist_with_params(spearvals=seq(from=0,to=0.9,by=0.1))
+coplist<-callfn$coplist
+ncores<-3
+res_pt35_rankF<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=F,mc.cores=ncores)
+res_pt35_rankT<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=T,mc.cores=ncores)
+res_pt1000_rankF<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=F,mc.cores=ncores)
+res_pt1000_rankT<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=T,mc.cores=ncores)
+save(res_pt35_rankF,res_pt35_rankT,res_pt1000_rankF,res_pt1000_rankT,file=paste(resultsloc,"StatTestNumericResults.RData",sep=''))
+
+b<-callfn$paramlist
+plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b)
+plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b)
+plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b)
+plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b)
+
+```
+
+<!--Now give the plot a number so it can be refered to-->
+```{r stat_testing35,echo=F}
+fig_nums("stat_testing35",display=F)
+```
+![stat_testing_results_datapt_35](./Results/stat_results/stat_testing/res_pt35_rankT.pdf)
+
+```{r stat_testing1000,echo=F}
+fig_nums("stat_testing1000",display=F)
+```
+![stat_testing_results_datapt_1000](./Results/stat_results/stat_testing/res_pt1000_rankT.pdf)
 

### 2018-03-07 Tests.Rmd
The beginning of a unit test suite

---
title: "Unit tests"
author: "Shyamolina Ghosh and Daniel Reuman"
date: "March 7, 2018"
output: pdf_document
---

# Test flexible copula functions

<!--Set up some test data-->
```{r test_copula_function_flexible_data, echo=T}
#Set up some test data
dl<-seq(from=0.1,by=0.1,to=0.9)
ep<-c(.05,-.03,.02,-.1,-.09,.05,.1,-.08,.05)
vi<-dl+ep
vj<-dl-ep
```

<!--Test the cor stat-->
```{r test_copula_function_flexible_cor, echo=T}
#Test the cor stat
source("CopulaFunctions_flexible.R")

h<-Corbds(vi,vj,lb=.05,ub=.45)
inds<-1:4
ht<-sum((vi[inds]-mean(vi))*(vj[inds]-mean(vj)))/((length(vi)-1)*sqrt(var(vi)*var(vj)))

if (abs(h-ht)<1e-12)
{
  prin("passed")
}
```

<!--Test the D2 stat-->
```{r test_copula_function_flexible_D2, echo=T}
#Test the D2 stat
source("CopulaFunctions_flexible.R")

h<-D2bds(vi,vj,lb=.05,ub=.45)
if (abs(sum(ep[1:4]^2)/2-h)<1e-12)
{
  print("passed")
}
```

<!--Test the P stat-->
```{r test_copula_function_flexible_P, echo=T}
#Test the P stat
source("CopulaFunctions_flexible.R")

#Shya to write, see above to get an idea of what kind of test we need
```

# Test surrogate functions

<!--Shya to add headers for other things that need to be tested, and let's discuss before you add the tests themselves-->

### 2018-03-07 SupportingInformation.Rmd
don't know what changed but it is making me commit
sourcing skewness file

--- 2018-03-06 SupportingInformation.Rmd
@@
 ```{r setup, echo=F}
-knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
+knitr::opts_chunk$set(echo = TRUE, fig.pos = "H") #, warning=F)
 
@@
 source("copsurrog2d.R")
+source("TestStats_multivar_dataset.R")
+source("skewness_multivar_dataset.R")
 ```
@@
 
+### Stat-testing of Non-parametric analysis (for limited number of datapoints)
+```{r stat_testing_save_result, message=F, warning=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"))}
+set.seed(seed)
+
+resultsloc<-"./Results/stat_results/stat_testing/"
+callfn<-coplist_with_params(spearvals=seq(from=0,to=0.9,by=0.1))
+coplist<-callfn$coplist
+ncores<-3
+res_pt35_rankF<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=F,mc.cores=ncores)
+res_pt35_rankT<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=T,mc.cores=ncores)
+res_pt1000_rankF<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=F,mc.cores=ncores)
+res_pt1000_rankT<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=T,mc.cores=ncores)
+save(res_pt35_rankF,res_pt35_rankT,res_pt1000_rankF,res_pt1000_rankT,file=paste(resultsloc,"StatTestNumericResults.RData",sep=''))
+```
+
+```{r stat_testing_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,callfn,mtime("StatTestNumericResults.RData"))}
+set.seed(seed)
+resultsloc<-"./Results/stat_results/stat_testing/"
+b<-callfn$paramlist
+plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b)
+plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b)
+plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b)
+plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b)
+```
+
+<!--Now give the plot a number so it can be refered to-->
+```{r stat_testing35,echo=F}
+fig_nums("stat_testing35",display=F)
+```
+![stat_testing_results_datapt_35](./Results/stat_results/stat_testing/res_pt35_rankT.pdf)
+
+```{r stat_testing1000,echo=F}
+fig_nums("stat_testing1000",display=F)
+```
+![stat_testing_results_datapt_1000](./Results/stat_results/stat_testing/res_pt1000_rankT.pdf)
+
 ## Soil C and N data
@@
 <!--Now create the plot and save it to the results folder-->
-```{r plot_soilCN_graphic_approach, echo=F}
+```{r plot_soilCN_graphic_approach, results='hide', echo=F}
 x<-seq(from=0.05,to=0.95,by=0.1)
@@
 
-We also tested statistically for asymmetry of tail dependence, which seems visible in `r table_nums("soilCNcop",display="c")`. 
-Results (`r table_nums("soilCN_asym",display="c")`) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data.
+We also tested statistically for asymmetry of tail dependence, which seems visible in `r fig_nums("soilCNcop",display="c")`. 
+Results (`r table_nums("soilCNasym",display="c")`) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data.
 
@@
 <!--Shya to add results about skewness of spatial means-->
-```{r skewness_count_plot, echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
+```{r skewness_count_plot, results='hide',echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
 set.seed(seed)
@@
 <!--Shya to add results about skewness of spatial means-->
-```{r skewness_ff_plot, echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
+```{r skewness_ff_plot, results='hide',echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
 set.seed(seed)
@@
 <!--Shya to add results about skewness of spatial means-->
-```{r skewness_pns_plot, echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
+```{r skewness_pns_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
 set.seed(seed)

### 2018-03-08 Tests.Rmd
checking Pbds filled in

@@
 
+```{r plot_vivj,echo=T}
+plot_vivj<-function(vi,vj,lb,ub){
+ plot(vi,vj,col="red",xlim=c(0,1),ylim=c(0,1),asp=1,cex.axis=0.5)
+ axis(1, at = seq(0, 1, by = 0.1),cex.axis=0.5)
+ abline(a=2*lb,b=-1,col="blue")
+ abline(a=2*ub,b=-1,col="green4")
+ abline(a=0,b=1)
+ rect(0,0,1,1)
+}
+```
+
 <!--Test the cor stat-->
@@
 source("CopulaFunctions_flexible.R")
-
-h<-Corbds(vi,vj,lb=.05,ub=.45)
+lb=.05
+ub=.45
+h<-Corbds(vi,vj,lb=lb,ub=ub)
 inds<-1:4
@@
 {
-  prin("passed")
+  print("passed")
 }
@@
 source("CopulaFunctions_flexible.R")
-
-h<-D2bds(vi,vj,lb=.05,ub=.45)
-if (abs(sum(ep[1:4]^2)/2-h)<1e-12)
+lb=.05
+ub=.45
+h<-D2bds(vi,vj,lb=lb,ub=ub)
+if (abs((sum(ep[1:4]^2)/2)-h)<1e-12)
 {
@@
 source("CopulaFunctions_flexible.R")
+lb=0
+ub=.45
+vi<-c(vi[3],vi[7],vi[9])
+vj<-c(vj[3],vj[7],vj[9])
 
-#Shya to write, see above to get an idea of what kind of test we need
+# When two parallel boundary line are on the left side of vi+vj=1 line
+
+#(vi,vj) plot in unit box : lower tail
+plot_vivj(vi=vi,vj=vj,lb=lb,ub=ub)
+
+#calling P stat for lower tail
+hl<-Pbds(vi,vj,lb=lb,ub=ub)
+
+#plot the step fn
+plot(hl$dist_S,hl$S,type="b")
+
+#calculate the area under the curve 
+vertices<-unique(hl$dist_S[(hl$dist_S!=0)])
+area_S<-(vertices[2]-vertices[1])*1
+area_S
+
+area_S==hl$Au_S # it's a check
+
+hl$Au_Si
+
+#P stat value for lower tail
+(area_S-hl$Au_Si)==hl$abs_res
+
+# When two parallel boundary line are on the right side of vi+vj=1 line
+lb=0
+ub=.4 # if ub=0.45 then hu$Au_Si = hl$Au_Si : so, it's a check on Au_Si
+#(vi,vj) plot in unit box : upper tail
+plot_vivj(vi=vi,vj=vj,lb=(1-ub),ub=(1-lb))
+
+#calling P stat for upper tail
+hu<-Pbds(vi,vj,lb=(1-ub),ub=(1-lb))
+
+#plot the step fn
+plot(hu$dist_S,hu$S,type="b")
+
+#calculate the area under the curve 
+w1<-unique(hu$dist_S[(hu$dist_S!=0)])
+w1  #x-coord where the jump occurs
+h1<-unique(hu$S)
+h1  #y-coord where the jump occurs
+area_S<-((w1[2]-w1[1])*(h1[2]-h1[1]))+((w1[3]-w1[2])*(h1[3]-h1[1]))
+area_S
+
+area_S==hu$Au_S  # it's a check
+
+hu$Au_Si
+
+#P stat value for upper tail
+(area_S-hu$Au_Si)==hu$abs_res
+
+# When two parallel boundary lines are on either side of vi+vj=1 line
+lb=0.3
+ub=0.8
+
+#calling P stat for middle section
+hm<-Pbds(vi,vj,lb=lb,ub=ub)
+
+#(vi,vj) plot in unit box : middle section
+plot_vivj(vi=vi,vj=vj,lb=lb,ub=ub)
+
+#plot the step fn
+plot(hm$dist_S,hm$S,type="b")
+
+#calculate the area under the curve 
+w1<-unique(hm$dist_S[(hm$dist_S!=0)])
+w1  #x-coord where the jump occurs
+h1<-unique(hm$S)
+h1  #y-coord where the jump occurs
+area_S<-((w1[2]-w1[1])*(h1[2]-h1[1]))+((w1[3]-w1[2])*(h1[3]-h1[1]))
+area_S
+
+area_S==hm$Au_S # it's a check
+
+hm$Au_Si
+
+#P stat value for middle section
+(area_S-hm$Au_Si)==hm$abs_res
 ```

### 2018-03-08 Tests.Rmd
moved some testing from copsurrog2d.R to Tests.Rmd and made better comments for the new context. Have not run it yet.

@@
 
+# Test skewness and third central moment functions
+
+```{r}
+source("SkewnessAnd3CentMom.R")
+#Dan fill in - see if there is testing text in the file itself 
+```
+
 # Test surrogate functions
 
+<!--Test the copsurrog2d stat-->
+```{r test_copsurrog2d, echo=T}
+#Test copsurrog2d
+
+source("copsurrog2d.R")
+set.seed(101)
+
+#***basic tests of the function using kendall-preserving surrogates
+datcop<-claytonCopula(5,2)
+numpts<-1000
+m<-rCopula(numpts,datcop)
+plot(m[,1],m[,2],type='p')
+
+tarcop<-gumbelCopula(3,2)
+
+res<-copsurrog2d(m,tarcop,'kendall',3)
+dim(res) 
+if (sum(dim(res)==dim(m))==2)
+{
+  print("passed")
+} else
+{
+  print("dim of output is wrong")
+}
+
+cor(m[,1],m[,2],method='kendall') #this should be similar to the following 
+cor(res[,1,1],res[,2,1],method='kendall')
+cor(res[,1,2],res[,2,2],method='kendall')
+cor(res[,1,3],res[,2,3],method='kendall')
+
+plot(res[,1,1],res[,2,1],type='p') #should look like a Gumbel
+BiCopGofTest(res[,1,1],res[,2,1],family=4) #should indicate an acceptable fit
+BiCopGofTest(res[,1,2],res[,2,2],family=4) #should indicate an acceptable fit
+BiCopGofTest(res[,1,3],res[,2,3],family=4) #should indicate an acceptable fit
+
+if (sum(sort(res[,1,1])==sort(m[,1]))==numpts &&
+    sum(sort(res[,2,1])==sort(m[,2]))==numpts &&
+    sum(sort(res[,1,2])==sort(m[,1]))==numpts &&
+    sum(sort(res[,2,2])==sort(m[,2]))==numpts &&
+    sum(sort(res[,1,3])==sort(m[,1]))==numpts &&
+    sum(sort(res[,2,3])==sort(m[,2]))==numpts)
+{
+  print("passed")
+} else
+{
+  print("Columns of output of copsurrog should be permutations of the corresponding columns of the input")
+}
+
+#basic tests using spearman preserving surrogates
+res<-copsurrog2d(m,tarcop,'spearman',3)
+dim(res)
+if (sum(dim(res)==dim(m))==2)
+{
+  print("passed")
+} else
+{
+  print("dim of output is wrong")
+}
+
+cor(m[,1],m[,2],method='spearman') #this should be similar to the following 
+cor(res[,1,1],res[,2,1],method='spearman')
+cor(res[,1,2],res[,2,2],method='spearman')
+cor(res[,1,3],res[,2,3],method='spearman')
+
+plot(res[,1,1],res[,2,1],type='p') #should look like a Gumbel
+BiCopGofTest(res[,1,1],res[,2,1],family=4) #should indicate an acceptable fit
+BiCopGofTest(res[,1,2],res[,2,2],family=4) #should indicate an acceptable fit
+BiCopGofTest(res[,1,3],res[,2,3],family=4) #should indicate an acceptable fit
+
+if (sum(sort(res[,1,1])==sort(m[,1]))==numpts &&
+    sum(sort(res[,2,1])==sort(m[,2]))==numpts &&
+    sum(sort(res[,1,2])==sort(m[,1]))==numpts &&
+    sum(sort(res[,2,2])==sort(m[,2]))==numpts &&
+    sum(sort(res[,1,3])==sort(m[,1]))==numpts &&
+    sum(sort(res[,2,3])==sort(m[,2]))==numpts)
+{
+  print("passed")
+} else
+{
+  print("Columns of output of copsurrog should be permutations of the corresponding columns of the input")
+}
+```
+
+<!--test to see whether changing the copula
+markedly influences the distribution
+of the spatial mean-->
+```{r test_copsurrog2d_spatmean, echo=T, cache=T, cache.extra=list(mtime("copsurrog2d.R"),mtime("SkewnessAnd3CentMom.R"))}
+#***test to see whether changing the copula
+#markedly influences the distribution
+#of the spatial mean
+set.seed(101)
+
+#make the data
+datcop<-rotCopula(claytonCopula(5,2))
+d<-rCopula(10000,datcop)
+d<-qnorm(d,mean=0,sd=1)
+
+#change the copula
+tarcop<-claytonCopula(3,2)
+sur<-copsurrog2d(d,tarcop,"kendall",1000)
+
+#see how the skewness of the spatial average
+#is affected
+source("SkewnessAnd3CentMom.R")
+skdat<-myskns(apply(FUN=mean,X=d,MARGIN=1)) #skewness of spatial mean of data
+mnres<-apply(FUN=mean,X=sur,MARGIN=c(1,3)) 
+sksur<-apply(FUN=myskns,X=mnres,MARGIN=2) #skewnesses of spatial means of surrogate datasets
+sum(sksur<skdat)/length(sksur) #skewnesses of surrogs should be less than that of data
+range(sksur) #these should mostly be negative
+skdat #this should be positive
+
+#now do the same thing but preserve the spearman instead
+sur<-copsurrog2d(d,tarcop,"spearman",1000)
+skdat<-myskns(apply(FUN=mean,X=d,MARGIN=1))
+mnres<-apply(FUN=mean,X=sur,MARGIN=c(1,3))
+sksur<-apply(FUN=myskns,X=mnres,MARGIN=2)
+sum(sksur<skdat)/length(sksur) #skewnesses of surrogs should be less than that of data
+range(sksur) #these should mostly be negative
+```
+
 <!--Shya to add headers for other things that need to be tested, and let's discuss before you add the tests themselves-->

### 2018-03-08 copsurrog2d.R

@@ -71,80 +71,3 @@ copsurrog2d<-function(m,targetcop,corpres,numsurrog)
   return(surrogs)
 }
 
-# #***basic tests of the function
-# datcop<-claytonCopula(5,2)
-# numpts<-1000
-# m<-rCopula(numpts,datcop)
-# plot(m[,1],m[,2],type='p')
-# 
-# tarcop<-gumbelCopula(3,2)
-# 
-# res<-copsurrog2d(m,tarcop,'kendall',3)
-# dim(res)
-# cor(m[,1],m[,2],method='kendall')
-# cor(res[,1,1],res[,2,1],method='kendall')
-# cor(res[,1,2],res[,2,2],method='kendall')
-# cor(res[,1,3],res[,2,3],method='kendall')
-# 
-# plot(res[,1,1],res[,2,1],type='p')
-# BiCopGofTest(res[,1,1],res[,2,1],family=4)
-# BiCopGofTest(res[,1,2],res[,2,2],family=4)
-# BiCopGofTest(res[,1,3],res[,2,3],family=4)
-# 
-# sum(sort(res[,1,1])==sort(m[,1]))
-# sum(sort(res[,2,1])==sort(m[,2]))
-# sum(sort(res[,1,2])==sort(m[,1]))
-# sum(sort(res[,2,2])==sort(m[,2]))
-# sum(sort(res[,1,3])==sort(m[,1]))
-# sum(sort(res[,2,3])==sort(m[,2]))
-# 
-# res<-copsurrog2d(m,tarcop,'spearman',3)
-# dim(res)
-# cor(m[,1],m[,2],method='spearman')
-# cor(res[,1,1],res[,2,1],method='spearman')
-# cor(res[,1,2],res[,2,2],method='spearman')
-# cor(res[,1,3],res[,2,3],method='spearman')
-# 
-# plot(res[,1,1],res[,2,1],type='p')
-# BiCopGofTest(res[,1,1],res[,2,1],family=4)
-# BiCopGofTest(res[,1,2],res[,2,2],family=4)
-# BiCopGofTest(res[,1,3],res[,2,3],family=4)
-# 
-# sum(sort(res[,1,1])==sort(m[,1]))
-# sum(sort(res[,2,1])==sort(m[,2]))
-# sum(sort(res[,1,2])==sort(m[,1]))
-# sum(sort(res[,2,2])==sort(m[,2]))
[…]

### 2018-03-08 SupportingInformation.Rmd
put the mtime function in its own file

@@
 
-#A function needed for caching
-mtime <- function(files)
-{
-  lapply(Sys.glob(files),function(x) X =   
-          file.info(x)$mtime)
-}
-
+source("mtime.R") #A function needed for caching
 source("getcopula.R")

### 2018-03-08 Tests.Rmd
put the mtime function in its own file
minor tweaks to testing

@@
 
+<!--Basic setup-->
+```{r setup, echo=F}
+#Basic setup
+
+#A function needed for caching
+source("mtime.R")
+```
+
 # Test flexible copula functions
 
-<!--Set up some test data-->
+Set up some test data:
 ```{r test_copula_function_flexible_data, echo=T}
-#Set up some test data
 dl<-seq(from=0.1,by=0.1,to=0.9)
@@
 
+A plot function to use in testing:
 ```{r plot_vivj,echo=T}
@@
 
-<!--Test the cor stat-->
+Test the cor statistic:
 ```{r test_copula_function_flexible_cor, echo=T}
-#Test the cor stat
 source("CopulaFunctions_flexible.R")
@@
 ub=.45
+plot_vivj(vi,vj,lb,ub)
 h<-Corbds(vi,vj,lb=lb,ub=ub)
@@
 
-<!--Test the D2 stat-->
+Test the D2 statistic:
 ```{r test_copula_function_flexible_D2, echo=T}
-#Test the D2 stat
 source("CopulaFunctions_flexible.R")
@@
 
-<!--Test the P stat-->
+Test the P statistic:
 ```{r test_copula_function_flexible_P, echo=T}
-#Test the P stat
 source("CopulaFunctions_flexible.R")
-lb=0
-ub=.45
+
 vi<-c(vi[3],vi[7],vi[9])
@@
 
-# When two parallel boundary line are on the left side of vi+vj=1 line
-
-#(vi,vj) plot in unit box : lower tail
+#Test for when two parallel boundary lines are on the left side of vi+vj=1 line
+lb=0
+ub=.45
 plot_vivj(vi=vi,vj=vj,lb=lb,ub=ub)
-
-#calling P stat for lower tail
 hl<-Pbds(vi,vj,lb=lb,ub=ub)
@@
 
-# When two parallel boundary line are on the right side of vi+vj=1 line
+#Test for when two parallel boundary lines are on the right side of vi+vj=1 line
 lb=0
 ub=.4 # if ub=0.45 then hu$Au_Si = hl$Au_Si : so, it's a check on Au_Si
-#(vi,vj) plot in unit box : upper tail
 plot_vivj(vi=vi,vj=vj,lb=(1-ub),ub=(1-lb))
@@
 
-# When two parallel boundary lines are on either side of vi+vj=1 line
+#Test for when two parallel boundary lines are on opposite sides of vi+vj=1 line
 lb=0.3
@@
 
-<!--Test the copsurrog2d stat-->
+Test the copsurrog2d function:
 ```{r test_copsurrog2d, echo=T}
-#Test copsurrog2d
-
+library("VineCopula")
 source("copsurrog2d.R")
@@
 plot(m[,1],m[,2],type='p')
+rect(0,0,1,1)
 
@@
 dim(res) 
-if (sum(dim(res)==dim(m))==2)
+if (sum(dim(res)[1:2]==dim(m))==2)
 {
@@
 plot(res[,1,1],res[,2,1],type='p') #should look like a Gumbel
+rect(0,0,1,1)
 BiCopGofTest(res[,1,1],res[,2,1],family=4) #should indicate an acceptable fit
@@
 dim(res)
-if (sum(dim(res)==dim(m))==2)
+if (sum(dim(res)[1:2]==dim(m))==2)
 {
@@
 
-<!--test to see whether changing the copula
+Test to see whether changing the copula
 markedly influences the distribution
-of the spatial mean-->
+of the spatial mean:
 ```{r test_copsurrog2d_spatmean, echo=T, cache=T, cache.extra=list(mtime("copsurrog2d.R"),mtime("SkewnessAnd3CentMom.R"))}
-#***test to see whether changing the copula
-#markedly influences the distribution
-#of the spatial mean
 set.seed(101)

### 2018-03-08 Tests.Rmd
minor changes to Tests.Rmd

@@
 m<-rCopula(numpts,datcop)
-plot(m[,1],m[,2],type='p')
+plot(m[,1],m[,2],type='p',asp=1)
 rect(0,0,1,1)
@@
 
-plot(res[,1,1],res[,2,1],type='p') #should look like a Gumbel
+plot(res[,1,1],res[,2,1],type='p',asp=1) #should look like a Gumbel
 rect(0,0,1,1)
@@
 {
-  print("Columns of output of copsurrog should be permutations of the corresponding columns of the input")
+  print("Columns of output of copsurrog should be permutations of the corresponding 
+        columns of the input")
 }
 
-#basic tests using spearman preserving surrogates
+#***basic tests using spearman preserving surrogates
 res<-copsurrog2d(m,tarcop,'spearman',3)
@@
 mnres<-apply(FUN=mean,X=sur,MARGIN=c(1,3)) 
-sksur<-apply(FUN=myskns,X=mnres,MARGIN=2) #skewnesses of spatial means of surrogate datasets
+sksur<-apply(FUN=myskns,X=mnres,MARGIN=2) #skewnesses of spatial means of surrogate 
+                                          #datasets
 sum(sksur<skdat)/length(sksur) #skewnesses of surrogs should be less than that of data

### 2018-03-08 Tests.Rmd
tests for ncsurrog added

@@
 rect(0,0,1,1)
-BiCopGofTest(res[,1,1],res[,2,1],family=4) #should indicate an acceptable fit
+BiCopGofTest(res[,1,1],res[,2,1],family=4) #should indicate an acceptable fit; 
+                                            #family=4 is Gumbel
 BiCopGofTest(res[,1,2],res[,2,2],family=4) #should indicate an acceptable fit
@@
 
-Test to see whether changing the copula
+Test copsurrog2d to see whether changing the copula
 markedly influences the distribution
@@
 
+Test the ncsurrog function:
+```{r test_ncsurrog, echo=T, cache=T, cache.extra=list(mtime("ncsurrog.R"))}
+source("ncsurrog.R")
+set.seed(103)
+
+#test without NAs
+cop<-claytonCopula(5,3)
+numpts<-1000
+m<-rCopula(numpts,cop)
+plot(m[,1],m[,2],type='p',asp=1)
+rect(0,0,1,1)
+plot(m[,1],m[,3],type='p',asp=1)
+rect(0,0,1,1)
+res<-ncsurrog(m,"kendall",3)
+if (sum(dim(res)==c(1000,3,3))==3)
+{
+  print("passed")
+} else
+{
+  print("dimension wrong in ncsurrog output")
+}
+
+plot(res[,1,1],res[,2,1],type='p',asp=1)
+rect(0,0,1,1) #should look like a normal copula
+plot(res[,1,1],res[,3,1],type='p',asp=1)
+rect(0,0,1,1) #should look like a normal copula
+
+#these should show an adequate fit
+ncop<-normalCopula(.5,2)
+BiCopGofTest(res[,1,1],res[,2,1],family=1)
+BiCopGofTest(res[,1,2],res[,2,2],family=1)
+BiCopGofTest(res[,1,3],res[,2,3],family=1)
+
+#test with NAs
+m[sample(1:numpts,5),1]<-NA
+m[sample(1:numpts,5),2]<-NA
+m[sample(1:numpts,5),3]<-NA
+res<-ncsurrog(m,"kendall",3)
+if (sum(dim(res)==c(1000,3,3))==3)
+{
+  print("passed")
+} else
+{
+  print("dimension wrong in ncsurrog output")
+}
+
+sum(is.na(res[,1,1])) #these should all be 5
+sum(is.na(res[,2,1]))
+sum(is.na(res[,1,2]))
+sum(is.na(res[,2,2]))
+
+plot(res[,1,1],res[,2,1],type='p',asp=1)
+rect(0,0,1,1) #should look like a normal copula
+plot(res[,1,1],res[,3,1],type='p',asp=1)
+rect(0,0,1,1) #should look like a normal copula
+
+#these should show an adequate fit
+BiCopGofTest(res[,1,1],res[,2,1],family=1)
+BiCopGofTest(res[,1,2],res[,2,2],family=1)
+BiCopGofTest(res[,1,3],res[,2,3],family=1)
+```
+
 <!--Shya to add headers for other things that need to be tested, and let's discuss before you add the tests themselves-->

### 2018-03-08 ncsurrog.R

@@ -91,30 +91,3 @@ ncsurrog<-function(m,corpres,numsurrog)
   return(surrogs)
 }
 
-#test without NAs
-#cop<-claytonCopula(5,2)
-#numpts<-1000
-#m<-rCopula(numpts,cop)
-#plot(m[,1],m[,2],type='p')
-#res<-ncsurrog(m,3)
-#dim(res)
-#plot(res[,1,1],res[,2,1],type='p')
-#ncop<-normalCopula(.5,2)
-#BiCopGofTest(res[,1,1],res[,2,1],family=1)
-#BiCopGofTest(res[,1,2],res[,2,2],family=1)
-#BiCopGofTest(res[,1,3],res[,2,3],family=1)
-
-#test with NAs
-#m[sample(1:numpts,5),1]<-NA
-#m[sample(1:numpts,5),2]<-NA
-#res<-ncsurrog(m,3)
-#dim(res)
-#sum(is.na(res[,1,1]))
-#sum(is.na(res[,2,1]))
-#sum(is.na(res[,1,2]))
-#sum(is.na(res[,2,2]))
-#plot(res[,1,1],res[,2,1],type='p')
-#BiCopGofTest(res[,1,1],res[,2,1],family=1)
-#BiCopGofTest(res[,1,2],res[,2,2],family=1)
-#BiCopGofTest(res[,1,3],res[,2,3],family=1)
-

### 2018-03-08 Tests.Rmd
tests of the skewness and 3rd central moment function added

@@
 
-```{r}
+Test for unbiased estimator of third central moment:
+```{r test_myskewness, echo=F, cache=T, cache.extra=list(mtime("SkewnessAnd3CentMom.R"))}
 source("SkewnessAnd3CentMom.R")
-#Dan fill in - see if there is testing text in the file itself 
+set.seed(101)
+
+x<-rnorm(100000)
+my3cm(x) #should be close to 0
+
+x<-rpois(100000,5)
+my3cm(x) #should be close to 5, but not necessarily super close
+#since 3rd moments are hard to estimate accurately
+
+#see if it appears unbaised
+res<-NA*numeric(1000)
+for (counter in 1:1000)
+{
+  x<-rpois(100000,5)
+  res[counter]<-my3cm(x) 
+}
+hist(res)
+mean(res)
 ```
+
+No need to test the accompanying skewness function, since it is such a simple extension of the above.
 

### 2018-03-08 SkewnessAnd3CentMom.R

@@ -2,12 +2,15 @@
 #moment and skewness.
 
 #Computes 3rd central moment of the data using
-#the unbiased estimator.
-#
-#See
+#the unbiased estimator. See
 #http://mathworld.wolfram.com/SampleCentralMoment.html
 #http://mathworld.wolfram.com/h-Statistic.html
 #
+#Args
+#x      A numeric vector
+#
+#Output
+#The estimated third central moment
 my3cm<-function(x,na.rm=F)
 {
   if (na.rm==T)
@@ -22,10 +25,14 @@ my3cm<-function(x,na.rm=F)
 }
 
 #Computes skewness of the data, making use
-#of the function my3cm
-#
-#See
+#of the function my3cm. See
 #https://en.wikipedia.org/wiki/Skewness
+#
+#Args
+#x      A numeric vector
+#
+#Output
+#The estimated skewness
 #
 myskns<-function(x,na.rm=F)
 {

### 2018-03-08 Tests.Rmd
trivial

@@
 Test for unbiased estimator of third central moment:
-```{r test_myskewness, echo=F, cache=T, cache.extra=list(mtime("SkewnessAnd3CentMom.R"))}
+```{r test_myskewness, echo=T, cache=T, cache.extra=list(mtime("SkewnessAnd3CentMom.R"))}
 source("SkewnessAnd3CentMom.R")

### 2018-03-08 SupportingInformation.Rmd
filled in skewness and stat-testing description

@@
 ### Stat-testing of Non-parametric analysis (for limited number of datapoints)
+
+Here, we want to test our $3$ types of nonparametric tail dependent statistics on both small $(35)$ and large $(1000)$ number of datapoints drawn from $4$ types of bivariate copula, namely, *Clayton* , *Normal*, *Frank* and *Survival Clayton*.
+Whereas *Normal* and *Frank* have symmetric tail dependence, *Clayton* and *Survival Clayton* have lower and upper tail dependence, respectively.
+First we chose a set of parameters [using *iRho* from *Copula* package] for each type of copula for which spearman correlation will be $10$ equispaced number in between [$0,0.9$].
+Then for each parameter (or spearman value) of each type of copula we carried out $100$ simulation and calculated $(Corl , Coru, Pl, Pu, D2l, D2u)$ stats.
+We took mean of our statistic $Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ over $100$ simulations and In `r fig_nums("stat_testing35", display="c")`, we plotted those statistic against spearman correlation.
+We put the standard error bar around each point in the plot and these errors got minimized with increasing numbe of data points [`r fig_nums("stat_testing1000", display="c")`].
+Statistic $Corl - Coru$, $Pl-Pu$ and $D2u-D2l$ should be $0$ for spearman=$0$ for *Clayton* and *Survival Clayton* and should remain $0$ for all values of spearman for symmetric copula like *Normal* and *Frank*.
+To verify this, we executed t-test and report the $p$-values for each case and most of the case shows $p>0.05$.
+
 ```{r stat_testing_save_result, message=F, warning=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"))}
@@
 
+**Change of skewness : ** Here, we claim that because of significant tail dependence of the multivariate copula structure, the skewness of the time-series data averaged over all selective locations changed significantly from the individual skewness of time-series data taken at each pairwise location.
+Spatial mean of the time series data extracts that tail dependence of the multivariate data set, if any, and reflected in the averaged skewness.
+We tested this with null hypothesis $H_{0}$.
+If $H_{0}$ is true that means our multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
+So, first we generated $10000$ number of surrogate normal copula from our given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
+Then we plotted the distribution of spatial mean of each surrogate time series data set and the given multivariate data.
+If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence.
+
+To verify that we use aphid-count data, aphid-firstflight data and plankton-north-sea data.
+For aphid data we choose only those locations for which atleast $30$ number of data points present in each time series when taken as pairwise locations.
+For plankton-north-sea data that threshold on number of data points is $45$.
+
 ## Aphid count
 
+`r fig_nums("skewness_count",display="c")` shows that averaged skewness is much lower than expected if the aphid count data came from a normal copula.
+Here, $p$ value indicates the fraction of surrogate-skewness which are lower than the averaged skewness of the given data (red line in `r fig_nums("skewness_count",display="c")`).
+As $p<0.05$, we reject $H_{0}$.
+This result is consistent with the non-parametric analysis and model fitting results which indicated lower tail dependence.
+
 <!--Shya to add results about skewness of spatial means-->
-```{r skewness_count_plot, results='hide',echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
+```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
 set.seed(seed)
@@
 
+`r fig_nums("skewness_ff",display="c")` shows that averaged skewness is much higher than expected if the aphid first flight data came from a normal copula.
+Here, $p$ value indicates the fraction of surrogate-skewness which are greater than the averaged skewness of the given data (red line in `r fig_nums("skewness_ff",display="c")`).
+As $p<0.05$, we reject $H_{0}$ in support for upper tail dependence.
+
 <!--Shya to add results about skewness of spatial means-->
-```{r skewness_ff_plot, results='hide',echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
+```{r skewness_ff_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
 set.seed(seed)
@@
 
-<!--Shya to add results about skewness of spatial means-->
-<!--Shya to add results about skewness of spatial means-->
+`r fig_nums("skewness_pns_plot",display="c")` shows that averaged skewness is lower than expected if the plankton north sea data came from a normal copula.
+Our previous results on non-parametric analysis and model fitting showed lower tail dependence.
+Here, as $p>0.05$, we cannot reject $H_{0}$, .......???
+
 ```{r skewness_pns_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}

### 2018-03-09 Tests.Rmd
Editorial changes to Tests.Rmd

@@
 # Test flexible copula functions
+We here test the functions `Corbds`, `Pbds` and `D2bds`.
 
@@
 
-Test the cor statistic:
+Test the `Corbds` function:
 ```{r test_copula_function_flexible_cor, echo=T}
@@
 
-Test the D2 statistic:
+Test the `D2bds` function:
 ```{r test_copula_function_flexible_D2, echo=T}
@@
 
-Test the P statistic:
+Test the `Pbds` function:
 ```{r test_copula_function_flexible_P, echo=T}
@@
 
-#Test for when two parallel boundary lines are on the left side of vi+vj=1 line
+#Test the case in which the two parallel boundary lines are on the left side 
+#of the vi+vj=1 line
 lb=0
@@
 
-#plot the step fn
+#plot the step function output - should have one jump at a small distance
+#due to the one data point that is between the bound lines
 plot(hl$dist_S,hl$S,type="b")
 
-#calculate the area under the curve 
+#calculate the area under the curve, compare to what the function returns 
 vertices<-unique(hl$dist_S[(hl$dist_S!=0)])
 area_S<-(vertices[2]-vertices[1])*1
-area_S
+if (abs(area_S-hl$Au_S)<1e-12)
+{
+  print("passed")
+} else
+{
+  print("Au_S incorrectly computed")
+}
 
-area_S==hl$Au_S # it's a check
+#check P stat value for lower tail
+if (abs(area_S-hl$Au_Si-hl$abs_res)<1e-12)
+{
+  print("passed")
+} else
+{
+  print("Error in Pbds, the P stat")
+}
 
-hl$Au_Si
-
-#P stat value for lower tail
-(area_S-hl$Au_Si)==hl$abs_res
-
-#Test for when two parallel boundary lines are on the right side of vi+vj=1 line
+#Test the case in which the two parallel boundary lines are on the right side 
+#of the vi+vj=1 line
 lb=0
-ub=.4 # if ub=0.45 then hu$Au_Si = hl$Au_Si : so, it's a check on Au_Si
+ub=.45 
 plot_vivj(vi=vi,vj=vj,lb=(1-ub),ub=(1-lb))
@@
 
-#plot the step fn
+#hl$Au_Si should be the same as hu$Au_Si
+if (abs(hl$Au_Si-hu$Au_Si)<1e-12)
+{
+  print("passed")
+}else
+{
+  print("Error in Pbds, Au_Si")
+}
+
+#plot the step fn - should have two jumps
 plot(hu$dist_S,hu$S,type="b")
@@
 w1<-unique(hu$dist_S[(hu$dist_S!=0)])
-w1  #x-coord where the jump occurs
+w1  #x-coords where the jumps occur, and the max dist
 h1<-unique(hu$S)
-h1  #y-coord where the jump occurs
+h1  #y-coords of the above points
 area_S<-((w1[2]-w1[1])*(h1[2]-h1[1]))+((w1[3]-w1[2])*(h1[3]-h1[1]))
-area_S
-
-area_S==hu$Au_S  # it's a check
-
-hu$Au_Si
+if (abs(area_S-hu$Au_S)<1e-12)  
+{
+  print("passed")
+} else
+{
+  print("Au_S incorrectly computed")
+}
 
 #P stat value for upper tail
-(area_S-hu$Au_Si)==hu$abs_res
+if (abs(area_S-hu$Au_Si-hu$abs_res)<1e-12)
+{
+  print("passed")
+}else
+{
+  print("P stat incorrectly computed")
+}
 
-#Test for when two parallel boundary lines are on opposite sides of vi+vj=1 line
-lb=0.3
-ub=0.8
+#Test for the case in which the two parallel boundary lines are on opposite 
+#sides of vi+vj=1 line
+lb<-0.28
+ub<-0.8
 
@@
 area_S<-((w1[2]-w1[1])*(h1[2]-h1[1]))+((w1[3]-w1[2])*(h1[3]-h1[1]))
-area_S
-
-area_S==hm$Au_S # it's a check
-
-hm$Au_Si
+if (abs(area_S-hm$Au_S)<1e-12) 
+{
+  print("passed")
+}else
+{
+  print("Error")
+}
 
 #P stat value for middle section
-(area_S-hm$Au_Si)==hm$abs_res
+if (abs(area_S-hm$Au_Si-hm$abs_res)<1e-12)
+{
+  print("passed")
+}else
+{
+  print("Error")
+}
 ```
@@
 
-Test for unbiased estimator of third central moment:
+We here test the function `my3cm`, which is an unbiased estimator of the 3rd central moment:
 ```{r test_myskewness, echo=T, cache=T, cache.extra=list(mtime("SkewnessAnd3CentMom.R"))}
@@
 x<-rpois(100000,5)
-my3cm(x) #should be close to 5, but not necessarily super close
+#should be close to 5, but not necessarily super close
 #since 3rd moments are hard to estimate accurately
+my3cm(x)
 
@@
 
-Test the copsurrog2d function:
+Test the `copsurrog2d` function:
 ```{r test_copsurrog2d, echo=T}
@@
 
-Test copsurrog2d to see whether changing the copula
+Test `copsurrog2d` to see whether changing the copula
 markedly influences the distribution
@@
 
-Test the ncsurrog function:
+Test the `ncsurrog` function:
 ```{r test_ncsurrog, echo=T, cache=T, cache.extra=list(mtime("ncsurrog.R"))}

### 2018-03-09 Tests.Rmd
added tests for OurBiCopSelect

@@
 
+# Tests to model selection tool
+
+We here test the `OurBiCopSelect` function with simulated data.
+
+Data are simulated from a Clayton copula with parameter 8:
+```{r get_data_test, echo=T}
+source("getcopula.R")
+source("OurBiCopSelect.R")
+set.seed(101)
+cop<-claytonCopula(8,2)
+d<-rCopula(1000,cop)
+v<-getcopula(d=d,rankon=F,ploton=T) #with real data, rankon should be T
+```
+
+Now apply the function:
+```{r bivmodselect_test, echo=F, cache=T, cache.extra=list(d,mtime("OurBiCopSelect.R"))}
+families<-c(1,3,4,5,13)
+BivMS_res_test<-OurBiCopSelect(d[,1],d[,2],families=families,gofnormal=TRUE,status=TRUE)
+```
+
+Independence should be rejected (low $p$ value):
+```{r echo=T}
+BivMS_res_test$IndepTestRes
+```
+
+Model selection should support the actual model (Clayton):
+```{r echo=T}
+BivMS_res_test$InfCritRes
+```
+
+The actual model (Clayton) should not be rejected:
+```{r echo=T}
+BivMS_res_test$GofRes_CvM
+BivMS_res_test$GofRes_KS
+```
+
+Look at goodness-of-fit results for the normal copulas:
+```{r echo=T}
+BivMS_res_test$GofRes_Normal_CvM
+BivMS_res_test$GofRes_Normal_KS
+```
+
 <!--Shya to add headers for other things that need to be tested, and let's discuss before you add the tests themselves-->

### 2018-03-09 Tests.Rmd
editorial

@@
 
+We decided not to use the goodness-of-fit tests for normal copulas (`gofnormal=F` henceforth)
+because it takes so long, and the model-selection results give 
+sufficient indication that the normal copula is not well supported. 
+Finally, in other simulations we have discovered the goodness-of-fit
+tests for the normal copula have limited statistical power.
+
 <!--Shya to add headers for other things that need to be tested, and let's discuss before you add the tests themselves-->

### 2018-03-11 SupportingInformation.Rmd
filled in and cleaned

@@
     fig_caption: yes
+tables: True
 ---
@@
 ```{r setup, echo=F}
-knitr::opts_chunk$set(echo = TRUE, fig.pos = "H") #, warning=F)
-
-library(captioner)
-table_nums <- captioner(prefix = "Table")
-fig_nums <- captioner(prefix="Figure")
-
+knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
 seed<-101
@@
 source("FittingCopula_selective_loc.R")
+source("normal_cop_AICw.R")
 source("aphid_data_calling.R")
@@
 <!--The raw data look like this:-->
-```{r load_data_Loecke, echo=F, fig.cap="Plot of the raw soil C and N data.", fig.height=5, fig.width=5}
+```{r load_data_Loecke, echo=F, fig.cap="Plot of the raw soil C and N data.\\label{fig_soilCNraw}", fig.height=5, fig.width=5}
 d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
 d<-d[,c("SOCstock100","TSNstock100")]
-fig_nums("soilCNraw",display=F)
 plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1)
@@
 <!--The copula looks like this:-->
-```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data.",  fig.height=5, fig.width=5}
+```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data.\\label{fig_soilCNcop}",  fig.height=5, fig.width=5}
 #convert to a copula and plot
-fig_nums("soilCNcop",display=F)
 v_CN<-getcopula(d=d,rankon=T,ploton=T) 
@@
 
-Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description).
-The raw data are shown in `r fig_nums("soilCNraw",display="c")`.
-The copula is shown in `r fig_nums("soilCNcop",display="c")`.
+Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description). The raw data are shown in Figure \ref{fig_soilCNraw}. The copula is shown in \ref{fig_soilCNcop}.
 
@@
 
+Here we intend to study that if for a given bivariate dataset we are able to plot the data within a unit square box based on their rankings then does that data look like totally independent or do they assume a copula structure?
+If they do show some dependence then at best by which family of copula we can describe their dependence?
+To answer these questions, we first choose to pass our bivariate dataset through a p-value based independence-test using *BiCopIndTest* function from *VineCopula* package.
+For that we set the significance level equal to $0.05$.
+If independence is rejected then we perform a test based on maximum likelihood estimation which returns us estimated parameters and log-likelihood, AIC, BIC, lower and upper tail dependence informations for a specific copula indicated by its familycode.
+We used *BiCopEst* function from *VineCopula* package for families like **N**ormal/Gaussian*{1}*, **C**layton*{3}*, **G**umbel*{4}*, **F**rank*{5}*, **J**oe*{6}*, **BB1***{7}*, **BB6***{8}*, **BB7***{9}*, **BB8***{10}*, **S**urvival **C**layton*{13}*, **S**urvival **G**umbel*{14}*, **S**urvival **J**oe*{16}*, **S**urvival **BB1***{17}*, **S**urvival **BB6***{18}*, **S**urvival **BB7***{19}* and **S**urvival **BB8***{20}*.
+The integer within the curly bracket *`{}`* indicates the familycode of a specific copula.
+We chose a best-fitted copula based on minimum AIC value and to verify that fitted copula-model we performed a goodness-of-fit test of our given data with that best-fitted copula-family by [*MyBiCopGofTest*](https://github.com/sghosh89/BIVAN/blob/master/MyBiCopGofTest.R) function from [*BIVAN*](https://github.com/sghosh89/BIVAN) repository.
+For the goodness-of-fit test if any of these p-values (based on CvM or KS method) is less than a threshold ($=0.2$) we used $1000$ number of bootstraps and $100$ otherwise.
+We calculated the Akaike weights for each copula model we chose and used that to find the model averaging lower and upper tail dependence value.
+
 ### Non-parametric analysis of tail dependence
@@
 
-```{r stat_image, echo=F}
-fig_nums("stat_image",display=F)
-```
-![schematic diagram for npa of tail dependence](./stat_image.jpg){ width=30%, height=30% }
+![schematic diagram for npa of tail dependence\label{fig_stat_image}](./stat_image.jpg){ width=30%, height=30% }
 
@@
 Say, there are total $n$ number of data points and $vi$ and $vj$ are the ranks of the dataset scaled down by factor $(n+1)$ taken at $(i,j)$ location pair or for $(i,j)$ observables.
-Now, to calculate our statistic we choose an arbitrary region of the box bounded by two parallel lines [blue line : $vi+vj=2l_{b}$, red line : $vi+vj=2u_{b}$ in `r fig_nums("stat_image",display="c")`] where $(l_{b},u_{b})$ has a range $[0,1]$ with the condition that $u_{b}$ must be greater than $l_{b}$.
+Now, to calculate our statistic we choose an arbitrary region of the box bounded by two parallel lines [blue line : $vi+vj=2l_{b}$, red line : $vi+vj=2u_{b}$ in Figure \ref{fig_stat_image}] where $(l_{b},u_{b})$ has a range $[0,1]$ with the condition that $u_{b}$ must be greater than $l_{b}$.
 Now we will demonstrate 3 different type of stats :
@@
 For our third statistic, we imagine the diagonal line of the box as the major axis of the copula.
-We calculate squared distance of each scattered point from $vi-vj=0$ [the dotted diagonal line in `r fig_nums("stat_image",display="c")`] and then take their average.
+We calculate squared distance of each scattered point from $vi-vj=0$ [the dotted diagonal line in Figure \ref{fig_stat_image}] and then take their average.
 This mean squared distance (say, $D2$) represents the average width of the copula for that shaded region.
@@
 Then for each parameter (or spearman value) of each type of copula we carried out $100$ simulation and calculated $(Corl , Coru, Pl, Pu, D2l, D2u)$ stats.
-We took mean of our statistic $Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ over $100$ simulations and In `r fig_nums("stat_testing35", display="c")`, we plotted those statistic against spearman correlation.
-We put the standard error bar around each point in the plot and these errors got minimized with increasing numbe of data points [`r fig_nums("stat_testing1000", display="c")`].
+We took mean of our statistic $Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ over $100$ simulations and In Figure \ref{fig_stat_testing35}, we plotted those statistic against spearman correlation.
+We put the standard error bar around each point in the plot and these errors got minimized with increasing numbe of data points [Figure \ref{fig_stat_testing1000}].
 Statistic $Corl - Coru$, $Pl-Pu$ and $D2u-D2l$ should be $0$ for spearman=$0$ for *Clayton* and *Survival Clayton* and should remain $0$ for all values of spearman for symmetric copula like *Normal* and *Frank*.
@@
 
-<!--Now give the plot a number so it can be refered to-->
-```{r stat_testing35,echo=F}
-fig_nums("stat_testing35",display=F)
-```
-![stat_testing_results_datapt_35](./Results/stat_results/stat_testing/res_pt35_rankT.pdf)
-
-```{r stat_testing1000,echo=F}
-fig_nums("stat_testing1000",display=F)
-```
-![stat_testing_results_datapt_1000](./Results/stat_results/stat_testing/res_pt1000_rankT.pdf)
-
+![stat_testing_results_datapt_35\label{fig_stat_testing35}](./Results/stat_results/stat_testing/res_pt35_rankT.pdf)
+
+![stat_testing_results_datapt_1000\label{fig_stat_testing1000}](./Results/stat_results/stat_testing/res_pt1000_rankT.pdf)
+
+\newpage
 ## Soil C and N data
@@
 <!--This does the computations for this section-->
-```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"))}
+```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"))}
 set.seed(seed)
@@
 ```
+
+<!--Display model selection results in a table-->
+```{r tab_soilCNfit,echo=F,results='markup'}
+h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+knitr::kable(h,
+             format='pandoc',
+             caption = "Fitting info for soil C and N dataset.
+Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=Survival Clayton; SG=Survival Gumbel; SJ=Survival Joe.
+An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by 180 degrees).\\label{tab_soilCNfit}",
+             booktabs=T)
+```
+
 The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes`. We fitted several copulas
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (`r table_nums("soilCNfit",display="c")`).
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_soilCNfit}).
 The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
-
-<!--Display model selection results in a table-->
-```{r soilCNfit,echo=F}
-h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-table_nums("soilCNfit",display=F)
-knitr::kable(h, caption = "Fitting info for soil C and N dataset.
-Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=Survival Clayton; SG=Survival Gumbel; SJ=Survival Joe.
-An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by 180 degrees).")
-```
 
@@
 
-As described in Methods, each of our statistics, as computed for the 
-ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
-of values of the same statistic with the same ranges computed on 1000 normal-copula 
-surrogate datasets with similar Kendall or Spearman correlations, in different 
-runs (`r fig_nums("soilCN_nonparam",display="c")`). Because the correlation 
-used in the Cor statistic is the Spearman correlation, Spearman-preserving surrogates
-should be used, so the top-left panel of `r fig_nums("soilCN_nonparam",display="c")`
-should be ignored (DAN: need to fix this later). Results indicate that tail dependence of 
-data is stronger in both the lower and upper tails than would be expected under a 
-null hypothesis of a normal copula. The dependence between C and N is also 
-weaker in the middle of both marginal distributions
-than would be expected based on a normal-copula null model.
-
-<!--Now give the plot a number so it can be refered to-->
-```{r show_plot_soildCN,echo=F}
-fig_nums("soilCN_nonparam",display=F)
-```
 ![Nonparametric tests for tail dependence and other deviations from a normal copula for soil C and N data.
 As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
-When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than.](./Results/stat_results/stat_soilCN/CNdata_NonparamTailDep.pdf)
-
-We also tested statistically for asymmetry of tail dependence, which seems visible in `r fig_nums("soilCNcop",display="c")`. 
-Results (`r table_nums("soilCNasym",display="c")`) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data.
+When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than.\label{fig_soilCN_nonparam}](./Results/stat_results/stat_soilCN/CNdata_NonparamTailDep.pdf)
 
@@
 atdres[3,3]<-fractxt(D2umD2l_frac_S)
-
-table_nums("soilCN_asym",display=F)
-knitr::kable(atdres, caption = "Results for tests of whether asymmetry of tail dependence is significant compared to a normal-copula null hypothesis.
+```
+
+```{r tab_soilCN_asym, echo=F, results="markup"}
+knitr::kable(atdres,
+             format="pandoc",
+             caption = "Results for tests of whether asymmetry of tail dependence is significant compared to a normal-copula null hypothesis.
 Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the ranges $0-0.1$ and $0.9-1$ (Methods).
 One thousand Kendall-preserving and 1000 Spearman-preserving surrogates were created, and the same statistics were computed for the surrogates.
-A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates.")
-```
-
-## Aphid abundance data 
+A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates.\\label{tab_soilCN_asym}",
+             booktabs=T)
+```
+
+As described in Methods, each of our statistics, as computed for the 
+ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
+of values of the same statistic with the same ranges computed on 1000 normal-copula 
+surrogate datasets with similar Kendall or Spearman correlations, in different 
+runs (Figure \ref{fig_soilCN_nonparam}). Because the correlation 
+used in the Cor statistic is the Spearman correlation, Spearman-preserving surrogates
+should be used, so the top-left panel of Figure \ref{fig_soilCN_nonparam}
+should be ignored (DAN: need to fix this later). Results indicate that tail dependence of 
+data is stronger in both the lower and upper tails than would be expected under a 
+null hypothesis of a normal copula. The dependence between C and N is also 
+weaker in the middle of both marginal distributions
+than would be expected based on a normal-copula null model.
+
+We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_soilCNcop}. 
+Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data.
+
+## Aphid abundance data for 'Green spruce aphid'
 <!--Add common and species name to header-->
@@
 
-```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count)}
+```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("FittingCopula_selective_loc.R"))}
 set.seed(seed)
@@
 
-\newpage
-
-```{r count_bestcop,echo=F,results="asis"}
-#library(pander)
-#pander(best_cop_LTdep, caption = "Hello caption!")
-#knitr::kable(RES_aphid_count$info_ord_copname[,,1],format="latex",longtable=F,caption = "Aphid_count : Best fit copula family for each selective location pair")
-
-table_nums("count_bestcop",display=F)
-knitr::kable(RES_aphid_count$info_ord_copname[,,1],caption = "Aphid_count : Best fit copula family for each selective location pair")
-```
-
-```{r count_AICw,echo=F,results="asis"}
-table_nums("count_AICw",display=F)
-knitr::kable(RES_aphid_count$info_ord_AICw[,,1], caption = "Aphid_count : AICw for best fit copula family for each selective location pair")
-```
-
-```{r count_LTdep,echo=F,results="asis"}
-table_nums("count_LTdep",display=F)
-knitr::kable(RES_aphid_count$info_ord_LTdep[,,1], caption = "Aphid_count : Lower Tail dependence for best fit copula family for each selective location pair")
-```
-
-```{r count_UTdep,echo=F,results="asis"}
-table_nums("count_UTdep",display=F)
-knitr::kable(RES_aphid_count$info_ord_UTdep[,,1], caption = "Aphid_count : Upper Tail dependence for best fit copula family for each selective location pair")
-```
-
-```{r count_LTdep_AICw,echo=F,results="asis"}
-table_nums("count_LTdep_AICw",display=F)
-knitr::kable(RES_aphid_count$LTdep_AICw,caption = "Aphid_count : Lower Tail dependence statistic for best fit copula family for each selective location pair")
-```
-
-```{r count_UTdep_AICw,echo=F,results="asis"}
-table_nums("count_UTdep_AICw",display=F)
-knitr::kable(RES_aphid_count$UTdep_AICw,caption = "Aphid_count : Upper Tail dependence statistic for best fit copula family for each selective location pair")
-
-```
-
-```{r count_LTmUT,echo=F,results="asis"}
-table_nums("count_LTmUT",display=F)
-knitr::kable(round(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,5),caption = "Aphid_count : Lower Tail - Upper Tail dependence statistic for best fit copula family for each of the selective location pair")
-```
-
-```{r count_p_CvM,echo=F,results="asis"}
-table_nums("count_p_CvM",display=F)
-knitr::kable(RES_aphid_count$gfc_p_CvM, caption = "Aphid_count : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula")
-```
-
-```{r count_p_KS,echo=F,results="asis"}
-table_nums("count_p_KS",display=F)
-knitr::kable(RES_aphid_count$gfc_p_KS,caption = "Aphid_count : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula")
-```
-
-`r table_nums("count_bestcop",display="c")` shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
-The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in `r table_nums("count_LTdep_AICw",display="c")` and `r table_nums("count_UTdep_AICw",display="c")` respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper) in `r table_nums("count_LTmUT",display="c")` indicates it is different from 0 for most of the locations, answering the third part of question 1.
-
+<!--\newpage-->
+
+```{r tab_count_bestcop, echo=F, results='markup'}
+knitr::kable(RES_aphid_count$info_ord_copname[,,1], 
+             format='pandoc',
+             caption = "Aphid_count : best fit copula family for each selective location pair\\label{tab_count_bestcop}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_count_AICw, echo=F, results='markup'}
+knitr::kable(RES_aphid_count$info_ord_AICw[,,1], 
+             format='pandoc',
+             caption = "Aphid_count : AICw for best fit copula family for each selective location pair\\label{tab_count_AICw}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_count_p_CvM, echo=F, results='markup'}
+knitr::kable(RES_aphid_count$gfc_p_CvM,
+             format='pandoc',
+             caption = "Aphid_count : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_count_p_CvM}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_count_p_KS, echo=F, results='markup'}
+knitr::kable(RES_aphid_count$gfc_p_KS,
+             format='pandoc',
+             caption = "Aphid_count : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula\\label{tab_count_p_KS}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_count_AICw_normal, echo=F, results='markup'}
+knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
+             format='pandoc',
+             caption = "Aphid_count : AICw for 'Normal' copula family for each selective location pair\\label{tab_count_AICw_normal}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_count_LTdep_AICw, echo=F, results='markup'}
+knitr::kable(RES_aphid_count$LTdep_AICw,
+             format='pandoc',
+             caption = "Aphid_count : model averaged lower tail dependence statistic for best fit copula family for each selective location pair\\label{tab_count_LTdep_AICw}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_count_UTdep_AICw, echo=F, results='markup'}
+knitr::kable(RES_aphid_count$UTdep_AICw,
+             format='pandoc',
+             caption = "Aphid_count : model averaged upper tail dependence statistic for best fit copula family for each selective location pair\\label{tab_count_UTdep_AICw}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_count_LTmUT, echo=F, results='markup'}
+knitr::kable(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,
+             digits = 5,
+             format='pandoc',
+             caption = "Aphid_count : model averaged lower tail dependence statistic for best fit copula family for each selective location pair\\label{tab_count_LTmUT}",
+             booktabs=TRUE
+             )
+```
+
+Table \ref{tab_count_bestcop} shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+Table \ref{tab_count_AICw} represents the Akaike weights of the best-fit copula for each location pair.
+Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_count_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_count_p_KS}] of the goodness of fit with atleast
+`r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` % successful bootstrap.
+Table \ref{tab_count_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_count_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC.
+The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_count_LTdep_AICw} and Table \ref{tab_count_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_count_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1.
 These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)`** out of **`r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-RES_aphid_count$num_indep_loc_pair`** location pair with greater lower tail dependence.
-Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [`r table_nums("count_p_CvM",display="c")`] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [`r table_nums("count_p_KS",display="c")`] of the goodness of fit.
 
@@
 
-```{r count_resamp,echo=F,results="asis"}
-table_nums("count_resamp",display=F)
-knitr::kable(stat_aphid_count$numericdf, caption = "Aphid_count : resampling result for stat indicating lower tail dependence")
-```
-
-```{r count_allcop,echo=F}
-fig_nums("count_allcop",display=F)
-```
-![copula plot for aphid-count](./Results/stat_results/stat_aphid_count/Sp_10_AllCops.pdf)
-
-```{r count_spear,echo=F}
-fig_nums("count_spear",display=F)
-```
-![Plot for aphid-count : Spearman vs. Distance](./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf){ width=50%, height=50% }
-
-```{r count_kend,echo=F}
-fig_nums("count_kend",display=F)
-```
-![Plot for aphid-count : Kendall vs. Distance](./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf){ width=50%, height=50% }
-
-```{r count_Corl-Coru,echo=F}
-fig_nums("count_Corl-Coru",display=F)
-```
-![Corl-Coru vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
-
-```{r count_Pl-Pu,echo=F}
-fig_nums("count_Pl-Pu",display=F)
-```
-![Pl-Pu vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
-
-```{r count_D2u-D2l,echo=F}
-fig_nums("count_D2u-D2l",display=F)
-```
-![D2u-D2l vs. Distance plot for aphid-count showing lower tail dependence](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
-
-<!--
-```{r, echo=F,out.width = "8cm", out.height= "8cm", out.caption="adhjk"}
-knitr::include_graphics("./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf")
-knitr::include_graphics("./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf")
-```
--->
-
-## Aphid first flight data
+```{r tab_count_resamp,echo=F,results="markup"}
+knitr::kable(stat_aphid_count$numericdf,
+             format = "pandoc",
+             caption = "Aphid_count : resampling result for stat indicating lower tail dependence\\label{tab_count_resamp}",
+             booktabs=TRUE)
+```
+
+![copula plot for aphid-count\label{fig_count_allcop}](./Results/stat_results/stat_aphid_count/Sp_10_AllCops.pdf)
+
+![Plot for aphid-count : Spearman vs. Distance\label{fig_count_spear}](./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf){ width=50%, height=50% }
+
+![Plot for aphid-count : Kendall vs. Distance\label{fig_count_kend}](./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf){ width=50%, height=50% }
+
+![Corl-Coru vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Corl-Coru}](./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
+
+![Pl-Pu vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Pl-Pu}](./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
+
+![D2u-D2l vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_D2u-D2l}](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
+
+Table \ref{tab_count_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean.
+Fourth column named $fracgt0$ in Table \ref{tab_count_resamp} indicates the fraction of resampled mean greater then zero.
+$Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ statistic with positive mean and almost $100$% $fracgt0$ proves that *Green spruce aphid*-count data has significant lower tail dependence.
+
+## Aphid first flight data for 'Leaf-curling plum aphid' 
 <!--Add species common name and scientific name to header-->
@@
 
-```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff)} 
+```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("FittingCopula_selective_loc.R"))} 
 set.seed(seed)
@@
 
+<!--\newpage-->
+
+```{r tab_ff_bestcop, echo=F, results='markup'}
+knitr::kable(RES_aphid_ff$info_ord_copname[,,1], 
+             format='pandoc',
+             caption = "Aphid_firstflight : best fit copula family for each selective location pair\\label{tab_ff_bestcop}",
+             booktabs=TRUE
+             )
+```
+
 \newpage
-
-```{r ff_bestcop,echo=F,results="asis"}
-table_nums("ff_bestcop",display=F)
-knitr::kable(RES_aphid_ff$info_ord_copname[,,1], caption = "Aphid_firstflight : Best fit copula family for each selective location pair")
-```
-
-```{r ff_AICw,echo=F,results="asis"}
-table_nums("ff_AICw",display=F)
-knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], caption = "Aphid_firstflight : AICw for best fit copula family for each selective location pair")
-```
-
-```{r ff_LTdep,echo=F,results="asis"}
-table_nums("ff_LTdep",display=F)
-knitr::kable(RES_aphid_ff$info_ord_LTdep[,,1], caption = "Aphid_firstflight : Lower Tail dependence for best fit copula family for each selective location pair")
-```
-
-```{r ff_UTdep,echo=F,results="asis"}
-table_nums("ff_UTdep",display=F)
-knitr::kable(RES_aphid_ff$info_ord_UTdep[,,1], caption = "Aphid_firstflight : Upper Tail dependence for best fit copula family for each selective location pair")
-```
-
-```{r ff_LTdep_AICw,echo=F,results="asis"}
-table_nums("ff_LTdep_AICw",display=F)
-knitr::kable(RES_aphid_ff$LTdep_AICw, caption = "Aphid_firstflight : Lower Tail dependence statistic for best fit copula family for each selective location pair")
-```
-
-```{r ff_UTdep_AICw,echo=F,results="asis"}
-table_nums("ff_UTdep_AICw",display=F)
-knitr::kable(RES_aphid_ff$UTdep_AICw, caption = "Aphid_firstflight : Upper Tail dependence statistic for best fit copula family for each selective location pair")
-```
-
-```{r ff_LTmUT,echo=F,results="asis"}
-table_nums("ff_LTmUT",display=F)
-knitr::kable(round(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,5), caption = "Aphid_firstflight : Lower Tail - Upper Tail dependence statistic for best fit copula family for each of the selective location pair")
-```
-
-```{r ff_p_CvM,echo=F,results="asis"}
-table_nums("ff_p_CvM",display=F)
-knitr::kable(RES_aphid_ff$gfc_p_CvM, caption = "Aphid_firstflight : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula")
-```
-
-```{r ff_p_KS,echo=F,results="asis"}
-table_nums("ff_p_KS",display=F)
-knitr::kable(RES_aphid_ff$gfc_p_KS, caption = "Aphid_firstflight : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula")
-```
-
-`r table_nums("ff_bestcop",display="c")`  shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
-The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in `r table_nums("ff_LTdep_AICw",display="c")` and `r table_nums("ff_UTdep_AICw",display="c")` respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper) in `r table_nums("ff_LTmUT",display="c")` indicates it is different from 0 for most of the locations, answering the third part of question 1.
-
+```{r tab_ff_AICw, echo=F, results='markup'}
+knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], 
+             format='pandoc',
+             caption = "Aphid_firstflight : AICw for best fit copula family for each selective location pair\\label{tab_ff_AICw}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_ff_p_CvM, echo=F, results='markup'}
+knitr::kable(RES_aphid_ff$gfc_p_CvM,
+             format='pandoc',
+             caption = "Aphid_firstflight : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_ff_p_CvM}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_ff_p_KS, echo=F, results='markup'}
+knitr::kable(RES_aphid_ff$gfc_p_KS,
+             format='pandoc',
+             caption = "Aphid_firstflight : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula\\label{tab_ff_p_KS}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_ff_AICw_normal, echo=F, results='markup'}
+knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
+             format='pandoc',
+             caption = "Aphid_firstflight : AICw for 'Normal' copula family for each selective location pair\\label{tab_ff_AICw_normal}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_ff_LTdep_AICw, echo=F, results='markup'}
+knitr::kable(RES_aphid_ff$LTdep_AICw,
+             format='pandoc',
+             caption = "Aphid_firstflight : model averaged lower tail dependence statistic for best fit copula family for each selective location pair\\label{tab_ff_LTdep_AICw}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_ff_UTdep_AICw, echo=F, results='markup'}
+knitr::kable(RES_aphid_ff$UTdep_AICw,
+             format='pandoc',
+             caption = "Aphid_firstflight : model averaged upper tail dependence statistic for best fit copula family for each selective location pair\\label{tab_ff_UTdep_AICw}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_ff_LTmUT, echo=F, results='markup'}
+knitr::kable(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,
+             digits = 5,
+             format='pandoc',
+             caption = "Aphid_firstflight : model averaged lower tail dependence statistic for best fit copula family for each selective location pair\\label{tab_ff_LTmUT}",
+             booktabs=TRUE
+             )
+```
+
+Table \ref{tab_ff_bestcop} shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+Table \ref{tab_ff_AICw} represents the Akaike weights of the best-fit copula for each location pair.
+Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_ff_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_ff_p_KS}] of the goodness of fit with atleast
+`r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` % successful bootstrap.
+Table \ref{tab_ff_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_ff_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC.
+The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_ff_LTdep_AICw} and Table \ref{tab_ff_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_ff_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1.
 These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.
-Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [`r table_nums("ff_p_CvM",display="c")`] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [`r table_nums("ff_p_KS",display="c")`] of the goodness of fit.
 
@@
 
-```{r ff_resamp,echo=F,results="asis"}
-table_nums("ff_resamp",display=F)
-knitr::kable(stat_aphid_ff$numericdf, caption = "Aphid_firstflight : resampling result for stat indicating upper tail dependence")
-```
-
-```{r ff_allcop,echo=F}
-fig_nums("ff_allcop",display=F)
-```
-![copula plot for aphid-firstflight](./Results/stat_results/stat_aphid_ff/Sp_11_AllCops.pdf)
-
-```{r ff_spear,echo=F}
-fig_nums("ff_spear",display=F)
-```
-![Plot for aphid-firstflight : Spearman vs. Distance](./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf){ width=50%, height=50% }
-
-```{r ff_kend,echo=F}
-fig_nums("ff_kend",display=F)
-```
-![Plot for aphid-firstflight : Kendall vs. Distance](./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf){ width=50%, height=50% }
-
-```{r ff_Corl-Coru,echo=F}
-fig_nums("ff_Corl-Coru",display=F)
-```
-![Corl-Coru vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
-
-```{r ff_Pl-Pu,echo=F}
-fig_nums("ff_Pl-Pu",display=F)
-```
-![Pl-Pu vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
-
-```{r ff_D2u-D2l,echo=F}
-fig_nums("ff_D2u-D2l",display=F)
-```
-![D2u-D2l vs. Distance plot for aphid-firstflight showing upper tail dependence](./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
-
-## North sea plankton data
+```{r tab_ff_resamp,echo=F,results="markup"}
+knitr::kable(stat_aphid_ff$numericdf,
+             format = "pandoc",
+             caption = "Aphid_firstflight : resampling result for stat indicating upper tail dependence\\label{tab_ff_resamp}",
+             booktabs=TRUE)
+```
+
+![copula plot for aphid-firstflight\label{fig_ff_allcop}](./Results/stat_results/stat_aphid_ff/Sp_11_AllCops.pdf)
+
+![Plot for aphid-firstflight : Spearman vs. Distance\label{fig_ff_spear}](./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf){ width=50%, height=50% }
+
+![Plot for aphid-firstflight : Kendall vs. Distance\label{fig_ff_kend}](./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf){ width=50%, height=50% }
+
+![Corl-Coru vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Corl-Coru}](./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
+
+![Pl-Pu vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Pl-Pu}](./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
+
+![D2u-D2l vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_D2u-D2l}](./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
+
+Table \ref{tab_ff_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean.
+Fourth column named $fracgt0$ in Table \ref{tab_ff_resamp} indicates the fraction of resampled mean greater then zero.
+$Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ statistic with negative mean has all resampled mean less than $0$ for *Leaf-curling plum aphid*-firstflight data.
+This proves that there is a significant upper tail dependence in *Leaf-curling plum aphid*-firstflight data.
+
+## North sea plankton data for 'Ceratium furca'
 
@@
 
-```{r RES_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea)}
+```{r RES_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"), mtime("FittingCopula_selective_loc.R"))}
+
 set.seed(seed)
@@
 
+<!--\newpage-->
+```{r tab_pns_bestcop, echo=F, results='markup'}
+knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1],
+             format='pandoc',
+             caption = "Plankton_north_sea : Best fit copula family for each selective location pair\\label{tab_pns_bestcop}",
+             booktabs=TRUE
+             )
+```
+
 \newpage
-
-```{r pns_bestcop,echo=F,results="asis"}
-#library(pander)
-#pander(best_cop_LTdep, caption = "Hello caption!")
-# alternatively,
-#pandoc.table(RES_plankton_north_sea$info_ord_AICw[,,1], caption="Plankton_north_sea : AICw for best fit copula family for each selective location pair")
-#cat("\n\n\\pagebreak\n")
-
-table_nums("pns_bestcop",display=F)
-knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1], caption = "Plankton_north_sea : Best fit copula family for each selective location pair")
-```
-
-```{r pns_AICw,echo=F,results="asis"}
-table_nums("pns_AICw",display=F)
-knitr::kable(round(RES_plankton_north_sea$info_ord_AICw[,,1],3), caption = "Plankton_north_sea : AICw for best fit copula family for each selective location pair")
-```
-
-\newpage
-
-```{r pns_LTdep,echo=F,results="asis"}
-#cat("\n\n\\pagebreak\n")
-
-table_nums("pns_LTdep",display=F)
-knitr::kable(round(RES_plankton_north_sea$info_ord_LTdep[,,1],3), caption = "Plankton_north_sea : Lower Tail dependence for best fit copula family for each selective location pair")
-```
-
-```{r pns_UTdep,echo=F,results="asis"}
-table_nums("pns_UTdep",display=F)
-knitr::kable(round(RES_plankton_north_sea$info_ord_UTdep[,,1],3), caption = "Plankton_north_sea : Upper Tail dependence for best fit copula family for each selective location pair")
-```
-
-\newpage
-
-```{r pns_LTdep_AICw,echo=F,results="asis"}
-#cat("\n\n\\pagebreak\n")
-
-table_nums("pns_LTdep_AICw",display=F)
-knitr::kable(round(RES_plankton_north_sea$LTdep_AICw,3), caption = "Plankton_north_sea : Lower Tail dependence statistic for best fit copula family for each selective location pair")
-```
-
-```{r pns_UTdep_AICw,echo=F,results="asis"}
-table_nums("pns_UTdep_AICw",display=F)
-knitr::kable(round(RES_plankton_north_sea$UTdep_AICw,3), caption = "Plankton_north_sea : Upper Tail dependence statistic for best fit copula family for each selective location pair")
-```
-
-\newpage
-
-```{r pns_LTmUT,echo=F,results="asis"}
-#cat("\n\n\\pagebreak\n")
-
-table_nums("pns_LTmUT",display=F)
-knitr::kable(round(RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw,3), caption = "Plankton_north_sea : Lower Tail - Upper Tail dependence statistic for best fit copula family for each of the selective location pair")
-```
-
-```{r pns_p_CvM,echo=F,results="asis"}
-table_nums("pns_p_CvM",display=F)
-knitr::kable(round(RES_plankton_north_sea$gfc_p_CvM,3), caption = "Plankton_north_sea : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula")
-```
-
-\newpage
-
-```{r pns_p_KS,echo=F,results="asis"}
-#cat("\n\n\\pagebreak\n")
-
-table_nums("pns_p_KS",display=F)
-knitr::kable(round(RES_plankton_north_sea$gfc_p_KS,3), caption = "Plankton_north_sea : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula")
-```
-
-`r table_nums("pns_bestcop",display="c")` shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
-The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in   `r table_nums("pns_LTdep_AICw",display="c")` and `r table_nums("pns_UTdep_AICw",display="c")` respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper) in `r table_nums("pns_LTmUT",display="c")` indicates it is different from 0 for most of the locations, answering the third part of question 1.
-
+```{r tab_pns_AICw,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$info_ord_AICw[,,1],
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : AICw for best fit copula family for each selective location pair\\label{tab_pns_AICw}",
+             booktabs=TRUE)
+```
+
+```{r tab_pns_p_CvM,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$gfc_p_CvM,
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_pns_p_CvM}",
+             booktabs=TRUE)
+```
+
+<!--\newpage-->
+```{r tab_pns_p_KS,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$gfc_p_KS,
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula\\label{tab_pns_p_KS}",
+             booktabs=TRUE)
+```
+
+```{r tab_pns_AICw_normal,echo=F,results="markup"}
+knitr::kable(normal_cop_AICw_matrix(r=RES_plankton_north_sea),
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : AICw for 'Normal' copula family for each selective location pair\\label{tab_pns_AICw_normal}",
+             booktabs=TRUE)
+```
+
+```{r tab_pns_LTdep_AICw,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$LTdep_AICw,
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : model averaged lower tail dependence statistic for best fit copula family for each selective location pair\\label{tab_pns_LTdep_AICw}",
+             booktabs=TRUE)
+```
+
+```{r tab_pns_UTdep_AICw,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$UTdep_AICw,
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : model averaged upper tail dependence statistic for best fit copula family for each selective location pair\\label{tab_pns_UTdep_AICw}",
+             booktabs=TRUE)
+```
+
+<!--\newpage-->
+
+```{r tab_pns_LTmUT,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw,
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : model averaged lower tail - model averaged upper tail dependence statistic for best fit copula family for each of the selective location pair\\label{tab_pns_LTmUT}",
+             booktabs=TRUE)
+```
+
+Table \ref{tab_pns_bestcop} shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+Table \ref{tab_pns_AICw} represents the Akaike weights of the best-fit copula for each location pair.
+Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_pns_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_pns_p_KS}] of the goodness of fit with atleast
+`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` % successful bootstrap.
+Table \ref{tab_pns_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_pns_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC.
+The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_pns_LTdep_AICw} and Table \ref{tab_pns_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_pns_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1.
 These results do convincingly show non-Gaussian copula structure, having `r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-RES_plankton_north_sea$num_indep_loc_pair` location pair more lower tail dependence.
-Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [`r table_nums("pns_p_CvM",display="c")`] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [`r table_nums("pns_p_KS",display="c")`] of the goodness of fit.
 
@@
 
-```{r pns_resamp,echo=F,results="asis"}
-table_nums("count_resamp",display=F)
-knitr::kable(stat_plankton_north_sea$numericdf, caption = "Plankton_north_sea : resampling result for stat indicating lower tail dependence")
-```
-
-```{r pns_allcop,echo=F}
-fig_nums("pns_allcop",display=F)
-```
-![copula plot for Plankton_north_sea](./Results/stat_results/stat_plankton_north_sea/Sp_16_AllCops.pdf)
-
-```{r pns_spear,echo=F}
-fig_nums("pns_spear",display=F)
-```
-![Plot for Plankton_north_sea : Spearman vs. Distance](./Results/stat_results/stat_plankton_north_sea/Sp_16_Spearman_vs_D.pdf){ width=50%, height=50% }
-
-```{r pns_kend,echo=F}
-fig_nums("pns_kend",display=F)
-```
-![Plot for Plankton_north_sea : Kendall vs. Distance](./Results/stat_results/stat_plankton_north_sea/Sp_16_Kendall_vs_D.pdf){ width=50%, height=50% }
-
-```{r pns_Corl-Coru,echo=F}
-fig_nums("pns_Corl-Coru",display=F)
-```
-![Corl-Coru vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
-
-```{r pns_Pl-Pu,echo=F}
-fig_nums("pns_Pl-Pu",display=F)
-```
-![Pl-Pu vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
-
-```{r pns_D2u-D2l,echo=F}
-fig_nums("pns_D2u-D2l",display=F)
-```
-![D2u-D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
+```{r tab_pns_resamp,echo=F,results="markup"}
+knitr::kable(stat_plankton_north_sea$numericdf,
+             format = "pandoc",
+             caption = "Plankton_north_sea : resampling result for stat indicating lower tail dependence\\label{tab_pns_resamp}",
+             booktabs=TRUE)
+```
+
+![copula plot for Plankton_north_sea\label{fig_pns_allcop}](./Results/stat_results/stat_plankton_north_sea/Sp_16_AllCops.pdf)
+
+![Plot for Plankton_north_sea : Spearman vs. Distance\label{fig_pns_spear}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Spearman_vs_D.pdf){ width=50%, height=50% }
+
+![Plot for Plankton_north_sea : Kendall vs. Distance\label{fig_pns_kend}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Kendall_vs_D.pdf){ width=50%, height=50% }
+
+![Corl-Coru vs.
+Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_Corl-Coru}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
+
+![Pl-Pu vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_Pl-Pu}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
+
+![D2u-D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_D2u-D2l}](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
+
+Table \ref{tab_pns_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean.
+Fourth column named $fracgt0$ in Table \ref{tab_pns_resamp} indicates the fraction of resampled mean greater then zero.
+$Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ statistic with positive mean and almost $100$% $fracgt0$ proves that plankton species data for *Ceratium furca* has significant lower tail dependence.
 
@@
 
-`r fig_nums("skewness_count",display="c")` shows that averaged skewness is much lower than expected if the aphid count data came from a normal copula.
-Here, $p$ value indicates the fraction of surrogate-skewness which are lower than the averaged skewness of the given data (red line in `r fig_nums("skewness_count",display="c")`).
+Figure \ref{fig_skewness_count} shows that averaged skewness is much lower than expected if the aphid count data came from a normal copula.
+Here, $p$ value indicates the fraction of surrogate-skewness which are lower than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_count}).
 As $p<0.05$, we reject $H_{0}$.
@@
 
-```{r skewness_count,echo=F}
-fig_nums("skewness_count",display=F)
-```
-![Skewness_count for species : Green spruce aphid](./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf){ width=50%, height=50% }
+![Skewness_count for species : Green spruce aphid\label{fig_skewness_count}](./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf){ width=50%, height=50% }
 
@@
 
-`r fig_nums("skewness_ff",display="c")` shows that averaged skewness is much higher than expected if the aphid first flight data came from a normal copula.
-Here, $p$ value indicates the fraction of surrogate-skewness which are greater than the averaged skewness of the given data (red line in `r fig_nums("skewness_ff",display="c")`).
+Figure \ref{fig_skewness_ff} shows that averaged skewness is much higher than expected if the aphid first flight data came from a normal copula.
+Here, $p$ value indicates the fraction of surrogate-skewness which are greater than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_ff}).
 As $p<0.05$, we reject $H_{0}$ in support for upper tail dependence.
@@
 
-```{r skewness_ff,echo=F}
-fig_nums("skewness_ff",display=F)
-```
-![Skewness_firstflight for species : Leaf-curling plum aphid](./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf){ width=50%, height=50% }
+![Skewness_firstflight for species : Leaf-curling plum aphid\label{fig_skewness_ff}](./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf){ width=50%, height=50% }
 
@@
 
-`r fig_nums("skewness_pns_plot",display="c")` shows that averaged skewness is lower than expected if the plankton north sea data came from a normal copula.
+Figure \ref{fig_skewness_pns} shows that averaged skewness is lower than expected if the plankton north sea data came from a normal copula.
 Our previous results on non-parametric analysis and model fitting showed lower tail dependence.
@@
 
-```{r skewness_pns,echo=F}
-fig_nums("skewness_pns",display=F)
-```
-![Skewness_plankton_north_sea for species : Ceratium_furca](./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf){ width=50%, height=50% }
+![Skewness_plankton_north_sea for species : Ceratium_furca\label{fig_skewness_pns}](./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf){ width=50%, height=50% }

### 2018-03-12 SupportingInformation.Rmd
edited a methods section and associated comments in bits of code; notes to shya included on references and other bits to be added, please

@@
 <!--The copula looks like this:-->
-```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data.\\label{fig_soilCNcop}",  fig.height=5, fig.width=5}
+```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data.
+Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations (see YYY).\\label{fig_soilCNcop}",  fig.height=5, fig.width=5}
 #convert to a copula and plot
@@
 ```
+<!--Shya please insert an automatic reference to the methods section "Model selection approach" at the location YYY above.
+I could not easily find a way to do section references by number that was consistent with the figure and table reference system you already have, but perhaps I missed it.
+In the end I think we will need to be able to refer to sections as well as figures and tables by number.
+I can see there is a package called "bookdown" that may help but there may be something easier using similar tools to what you have already used for figures and tables.-->
 
@@
 
-Here we intend to study that if for a given bivariate dataset we are able to plot the data within a unit square box based on their rankings then does that data look like totally independent or do they assume a copula structure?
-If they do show some dependence then at best by which family of copula we can describe their dependence?
-To answer these questions, we first choose to pass our bivariate dataset through a p-value based independence-test using *BiCopIndTest* function from *VineCopula* package.
-For that we set the significance level equal to $0.05$.
-If independence is rejected then we perform a test based on maximum likelihood estimation which returns us estimated parameters and log-likelihood, AIC, BIC, lower and upper tail dependence informations for a specific copula indicated by its familycode.
-We used *BiCopEst* function from *VineCopula* package for families like **N**ormal/Gaussian*{1}*, **C**layton*{3}*, **G**umbel*{4}*, **F**rank*{5}*, **J**oe*{6}*, **BB1***{7}*, **BB6***{8}*, **BB7***{9}*, **BB8***{10}*, **S**urvival **C**layton*{13}*, **S**urvival **G**umbel*{14}*, **S**urvival **J**oe*{16}*, **S**urvival **BB1***{17}*, **S**urvival **BB6***{18}*, **S**urvival **BB7***{19}* and **S**urvival **BB8***{20}*.
-The integer within the curly bracket *`{}`* indicates the familycode of a specific copula.
-We chose a best-fitted copula based on minimum AIC value and to verify that fitted copula-model we performed a goodness-of-fit test of our given data with that best-fitted copula-family by [*MyBiCopGofTest*](https://github.com/sghosh89/BIVAN/blob/master/MyBiCopGofTest.R) function from [*BIVAN*](https://github.com/sghosh89/BIVAN) repository.
-For the goodness-of-fit test if any of these p-values (based on CvM or KS method) is less than a threshold ($=0.2$) we used $1000$ number of bootstraps and $100$ otherwise.
-We calculated the Akaike weights for each copula model we chose and used that to find the model averaging lower and upper tail dependence value.
+For a bivariate dataset $(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps, 
+most implemented using tools from the `VineCopula` package. Code combining these tools is in the
+[*OurBiCopSelect*](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [*BIVAN*](https://github.com/sghosh89/BIVAN) repository.
+
+First, we produced ranks 
+$(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
+$\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
+we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; further work was with the $u_i$ 
+and $v_i$. Ranks $\tilde{u}_i$ and $\tilde{v}_i$ were $1$, for the 
+smallest element, up to $n$, for the largest.
+
+Second, we did a test of the independence of the $u_i$ and $v_i$. Subsequent model 
+selection algorithms are ineffective if data could not be distinguished from independent data 
+since most or all bivariate copula families considered include the independent copula in the 
+family or as a boundary case. The independence test was as implemented in the `BiCopIndTest`
+function in `VineCopula`, the documentation of which reports the test uses the asymptotic normality of the statistic $|\tau|\sqrt{\frac{9n(n-1)}{2(2n+5)}}$, where $\tau$ is Kendall's tau statistic.
+If independence could be rejected ($0.05$ significance level), model selection proceeded.
+
+The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`.
+Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian*{1}*, **C**layton*{3}*, **G**umbel*{4}*, **F**rank*{5}*, **J**oe*{6}*, **BB1***{7}*, **BB6***{8}*, **BB7***{9}*, **BB8***{10}*, **S**urvival **C**layton*{13}*, **S**urvival **G**umbel*{14}*, **S**urvival **J**oe*{16}*, **S**urvival **BB1***{17}*, **S**urvival **BB6***{18}*, **S**urvival **BB7***{19}* and **S**urvival **BB8***{20}* copulas.
+Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC) and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas (REF).
+<!--Shya, we here need a reference to the Burnham and Anderson book-->
+Descriptions of these copula families can be found in introductory texts on copulas
+<!--Shya to please insert references for both the books on copulas we have-->
+and in references cited in `VineCopula`. The lower- and upper-tail dependence of the 
+best-fitting member of each family was also provided by `BiCopEst`.
+
+Model selection methods will give the *relative* support of several models, but will not,
+on their own, indicate whether any of the models selected is an objectively good fit (all 
+models considered might poorly describe the data). To test this, and as the fourth step of
+our procedure, we tested the goodness of fit of our best-AIC copula family using a 
+bootstrapping procedure implemented in `BiCopGofTest` of `VineCopula`. The method performed 
+two tests, one based on a Cramer-von Mises statistic, and one based on a 
+Kolmogorov-Smirnov statistic. For computational efficiency, 
+an initial run using $100$ bootstraps was performed, and if
+the $p$-value resulting from either test was less than $0.2$, tests were re-run with 
+$1000$ boostraps.
+
+Finally, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
+model-averaged lower- and upper-tail dependence values using standard model averaging 
+formulas
+<!--Shya to add reference to Burnham and Anderson, please-->
+; likewise for BIC.
 

### 2018-03-12 MyBiCopGofTest.R

@@ -1,6 +1,11 @@
 library(mvtnorm)
 library(VineCopula)
 source("preprocessors.R")
+
+#This is a slight modification of the BiCopGofTest of the VineCopula package,
+#changed to include additional output on numbers of successful bootstraps,
+#generated but not returned by the original function.
+
 #-------------------------------------
 MyBiCopGofTest<-function (u1, u2, family, par = 0, par2 = 0, method = "white", 
           max.df = 30, B = 100, obj = NULL) 

### 2018-03-12 OurBiCopSelect.R

@@ -1,5 +1,6 @@
 library(VineCopula)
 source("MyBiCopGofTest.R")
+
 #This function takes a sample from a bivariate distribution
 #consisting of values between 0 and 1 and:
 #a) tests for independence
@@ -10,8 +11,8 @@ source("MyBiCopGofTest.R")
 #   for the goodness of fit in two ways
 #d) optionally tests for the goodness of fit of the normal copula
 #The function is a re-implementation of the BiCopSelect function
-#in the VineCopula package, re-done to ensure we could control
-#what we were getting.
+#in the VineCopula package, re-done to ensure we could understand and
+#control what we were getting.
 #
 #Args
 #u1, u2       The sample. These can be obtained, for instance, 

### 2018-03-12 preprocessors.R

@@ -1,3 +1,5 @@
+#Tools copied from the VineCopula package for use by MyBiCopGofTest.
+
 preproc <- function(args, ..., na.txt = NULL) {
   # augment arguments
   args$na.txt <- na.txt

### 2018-03-12 SupportingInformation.Rmd
one more comment added
substantial editorial changes to the section describing the nonparametric stats, cor, P, D2

@@
 updated on `r Sys.Date()`
+
+<!--New latex commands-->
+\newcommand{\mean}{\operatorname{mean}}
+\newcommand{\var}{\operatorname{var}}
+\newcommand{\cor}{\operatorname{cor}}
+\newcommand{\P}{\operatorname{P}}
+\newcommand{\D2}{\operatorname{D}^2}
 
@@
 [*OurBiCopSelect*](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [*BIVAN*](https://github.com/sghosh89/BIVAN) repository. 
+<!--Shya, links to code are a good idea! Thanks for figuring this out and adding, we should do something like this in other methods sections.-->
 
@@
 
+<!--Shya please insert some additional material for multivariate datasets - perhaps
+you don't need much because basically you just ran the bivariate analysis on each pair
+of locations, but you need to say that and then maybe a few sentences about how 
+everything was tallied up after having done that-->
+
 ### Non-parametric analysis of tail dependence
 
-<!--Shya to edit this description of the stats. Don't need subsections for each stat.-->
-<!--Eventually we will need to add methods on how we use these stats, which is different for multivariate and bivariate data-->
-<!--Eventually, for the multivariate data, we will need to demonstrate using a similar analysis to what is in mbsyn, that you can detect asymmetry of tail dependence using these stats - we have done this, but it needs to be adapted-->
-
-![schematic diagram for npa of tail dependence\label{fig_stat_image}](./stat_image.jpg){ width=30%, height=30% }
-
-In a given ranking based copula plot of a dataset within a unit square box the data points are scattered according to their joint-marginal distribution.
-Say, there are total $n$ number of data points and $vi$ and $vj$ are the ranks of the dataset scaled down by factor $(n+1)$ taken at $(i,j)$ location pair or for $(i,j)$ observables.
-Now, to calculate our statistic we choose an arbitrary region of the box bounded by two parallel lines [blue line : $vi+vj=2l_{b}$, red line : $vi+vj=2u_{b}$ in Figure \ref{fig_stat_image}] where $(l_{b},u_{b})$ has a range $[0,1]$ with the condition that $u_{b}$ must be greater than $l_{b}$.
-Now we will demonstrate 3 different type of stats :
-
-$Cor$, the correlation between the ${n}'$ number of points lying in between those two parallel lines, is given by;
-
+![Schematic diagram supporting defintions of nonparametric tail dependence statistics.\label{fig_stat_image}](./stat_image.jpg){ width=30%, height=30% }
+<!--Shya, can you please change this figure so it shows u (horizontal) and v (vertical) on the axis labels instead of vi and vj?
+Often times computer variables do not make good mathematical notation, so I have changed this, I also made changes for similar reasons to other notation elsewhere.-->
+
+Left- and right-tail dependence quantify the extent to which two ranked
+variables $u_i$, $v_i$ for $i=1,...,n$ 
+(see section REF for defintions of $u_i$ and $v_i$)
+<!--Shya pls insert ref to the section immediately above this one-->
+are related in the left and right tails, respectively, of their 
+distributions (REF). 
+<!--Shya, pls cite the Nelson book-->
+Our nonparametric study of tail dependence uses three
+statistics which can quantify the extent to which $u_i$ and $v_i$ are 
+related in any part of their distributions. We here describe the 
+statistics. The statistics were formulated with positively associated
+variables in mind. All the variables we study are positively associated
+when they are significantly associated. So definitions and associated 
+diagrams will assume positive associations, though generalizations are 
+possible. Given two bounds 
+$0 leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
+$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
+fall (these are the blue and red lines, respectively, in 
+Figure \ref{fig_stat_image}). Our statistics quantify the dependence
+between $u_i$ and $v_i$ in the region bounded by these lines.
+Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
+dependence in the left parts of the distributions of $u$ and $v$, and 
+using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
+the right parts of the distributions.
+
+The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
+correlation of $u_i$ and $v_i$ that is attributable to the points 
+lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
 \begin{equation}\label{eq.Cor}
-   Cor = \frac{1}{n-1}\sum_{k=1}^{{n}'} \left [ vi(k)-\left<vi\right> \right ] \left[ vj(k)-\left<vj\right> \right ]
+   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
+   (u_i-\mean(u)) (v_i-\mean(v)),
 \end{equation}
-
-where, $\left<vi\right>$ and $\left<vj\right>$ are the arithmatic mean of $vi$ and $vj$ of $n$ data points within the box, respectively.
-
-We can shift that pair of lines along the diagonal line of the box and can get a sense how the points are spread along the diagonal and how the quantity $Cor$ changes.
-
-For multivariate datasets due to limited number of data points this chosen region is broad, i.
-e., we measured $Cor$ for $(l_{b}=0,u_{b}=0.5)$ and $(l_{b}=0.5,u_{b}=1)$ and stored as $Corl$ and $Coru$ respectively.
-If we take their difference then that will tell us which end of the copula is with stronger correlation.
-A copula with $(Corl-Coru)>0$ should have more lower tail dependence whereas $(Corl-Coru)<0$ indicates more upper tail dependence.
-Symmetric copula, e.
-g., (Normal or Frannk) has $Corl = Coru$.
-
-On a second thought, we imagine as if we progress from the diagonal towards the sides of the box with $h$ gradually increasing from $0$ to $h_{max}$, where $h_{max}$ is the half-length of the longer one between two parallel lines.
-Then we take $S$, the ratio of number of points which are within $h$ distance from the diagonal on both side and the total number of points lie within the shaded region.
-Now, we want to measure that how much $S$ deviates from the standard $Si$ line when they are plotted against $h$ where $Si$ is the ratio of the area covered by distance $h$ on both side and the area of shaded region.
-This deviation $P$, representing our statistics to measure tail dependence, is given by;
-
+where means and variances are computed using all $n$ data 
+points, but the
+sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
+$u_i+v_i < 2u_b$. 
+Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
+between $u$ and $v$ in the region given by the bounds.
+
+We also defined a statistic $\P_{l_b,u_b}$, using the same bounds
+$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
+$S(h)$ to be the  
+number of points $(u_i,v_i)$ within the bounds 
+and a distance less than $h$ from the line $v=u$, divided by the 
+total number of points within the bounds. This function is
+defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
+is half the longer of the two segments obtained by intersecting the
+bounds with the unit square. It is easy to see $S(0)=0$ and 
+$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
+of $h$, to be the area within the bounds and the unit square
+and within distance $h$ of the line $v=u$, divided by the area
+within the bounds and the unit square. This is the expected value of
+$S(h)$ for independent data. We then defined
 \begin{equation}\label{eq.P}
-   P = \int_0^{h_{max}} \left[ S(h) - Si(h) \right] \,dh
+   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
 \end{equation}
-
-Like we said before, for multivariate dataset we measure $Pl = P$ for $(l_{b}=0, u_{b}=0.5)$ and $Pu = P$ for $(l_{b}=0.5, u_{b}=1)$ and these $Pl$, $Pu$ represent the lower and upper tail statistic respectively.
-A copula having $(Pl-Pu)>0$ represents more lower tail dependence whereas a copula with negative $(Pl-Pu)$ represents more upper tail dependence.
-A copula with equal $Pl$ and $Pu$ value is considered as symmetric with no tail-dependence.
-
-For our third statistic, we imagine the diagonal line of the box as the major axis of the copula.
-We calculate squared distance of each scattered point from $vi-vj=0$ [the dotted diagonal line in Figure \ref{fig_stat_image}] and then take their average.
-This mean squared distance (say, $D2$) represents the average width of the copula for that shaded region.
-Similarly, we calculate $D2l$ and $D2u$ for the lower and upper tail of a copula for multivariate dataset with $(l_{b}=0, u_{b}=0.5)$ and $(l_{b}=0.5, u_{b}=1)$ respectively.
-If for a given copula $(D2u-D2l)>0$, then it is with more lower tail dependence as the upper-end points are distantly spaced from the long axis of the copula.
-Conversely, $(D2u-D2l)<0$ means a copula with more upper tail dependence.
-If $D2l$ and $D2u$ are equal, then the data come from a symmetric copula.
+As for $\cor_{l_b,u_b}$, larger values of $P_{l_b,u_b}$ 
+indicate stronger positive association
+between $u$ and $v$ in the region given by the bounds.
+
+The third statistic, $\D2_{l_b,u_b}(u,v)$, is the average squared 
+distance between points satisfying $u_i+v_i>2l_b$ and 
+$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
+$\P_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
+$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\D2_{l_b,u_b}$ is 
+then small. 
+Thus large values of $\cor_{l_b,u_b}$ and $\P_{l_b,u_b}$ indicate strong 
+dependence in the portions of the distributions of $u$ and $v$ given by 
+the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
+$\D2_{l_b,u_b}$ indicate strong dependence.
+
+For large datasets (i.e., datasets with large $n$),
+we can consider $l_b$ and $u_b$ close together without incurring undue
+sampling variation in our statistics, and we can consider 
+several different bands $(l_b,u_b)$ to understand how dependence
+varies in different parts of the distributions. But for datasets with 
+smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
+We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
+and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\P_l=\P_{0,.5}$,
+$\P_u=\P_{0.5,1}$, $\D2_l=\D2_{0,0.5}$, and $\D2_u=\D2_{0.5,1}$.
+
+The difference $\cor_l-\cor_u$ is expected to be positive if 
+dependence in the left halves of the distributions is stronger 
+than dependence in their right halves; likewise for $\P_l-\P_u$.
+This holds, for instance, for the Clayton copula (REF).
+<!--Shya to cite the Nelson book-->
+On the other hand, the difference $\D2_u-\D2_l$ is expected to be 
+positive under the same conditions, so we use these three differences
+to detect asymmetry of tail dependence.
+
+<!--Eventually we will need to add methods on how we use these stats, which is different for multivariate and bivariate data.
+For bivariate data, it involves surrogates, for multivariate data it involves resampling locations.
+So, in particular, Dan will have to insert descriptions of the resampling methods.-->
 
@@
 
-\newpage
 ## Soil C and N data

### 2018-03-12 SupportingInformation.Rmd
minor additional edits

@@
 <!--Shya to cite the Nelson book-->
-On the other hand, the difference $\D2_u-\D2_l$ is expected to be 
-positive under the same conditions, so we use these three differences
-to detect asymmetry of tail dependence.
+The difference $\D2_u-\D2_l$, in which the "upper" and
+"lower" statistics occur in opposite order, is also expected to be 
+positive under the same conditions. We use these three differences
+to detect asymmetry of tail dependence. Likewise, for large datasets,
+we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
+$\P_{l_b,u_b}-\P_{1-u_b,1-l_b}$ and
+$\D2_{1-u_b,1-l_b}-\D2_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
 

### 2018-03-12 SupportingInformation.Rmd
chnaged the name of \P because of latex error

@@
 \newcommand{\cor}{\operatorname{cor}}
-\newcommand{\P}{\operatorname{P}}
+\newcommand{\Ps}{\operatorname{P}}
 \newcommand{\D2}{\operatorname{D}^2}
@@
 
-We also defined a statistic $\P_{l_b,u_b}$, using the same bounds
+We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
 $u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
@@
 $u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
-$\P_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
+$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
 $u_i+v_i<2u_b$ cluster close to the line $v=u$, $\D2_{l_b,u_b}$ is 
 then small. 
-Thus large values of $\cor_{l_b,u_b}$ and $\P_{l_b,u_b}$ indicate strong 
+Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
 dependence in the portions of the distributions of $u$ and $v$ given by 
@@
 We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
-and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\P_l=\P_{0,.5}$,
-$\P_u=\P_{0.5,1}$, $\D2_l=\D2_{0,0.5}$, and $\D2_u=\D2_{0.5,1}$.
+and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
+$\Ps_u=\Ps_{0.5,1}$, $\D2_l=\D2_{0,0.5}$, and $\D2_u=\D2_{0.5,1}$.
 
@@
 dependence in the left halves of the distributions is stronger 
-than dependence in their right halves; likewise for $\P_l-\P_u$.
+than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
 This holds, for instance, for the Clayton copula (REF).
@@
 we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
-$\P_{l_b,u_b}-\P_{1-u_b,1-l_b}$ and
+$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
 $\D2_{1-u_b,1-l_b}-\D2_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.

### 2018-03-13 SupportingInformation.Rmd
annoying changes to get latex to behave

@@
 \newcommand{\Ps}{\operatorname{P}}
-\newcommand{\D2}{\operatorname{D}^2}
+\newcommand{\Dt}{\text{D}^2}
 
@@
 ```
+
 <!--Shya please insert an automatic reference to the methods section "Model selection approach" at the location YYY above.
@@
 $u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
-$S(h)$ to be the  
-number of points $(u_i,v_i)$ within the bounds 
+$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
 and a distance less than $h$ from the line $v=u$, divided by the 
@@
 
-The third statistic, $\D2_{l_b,u_b}(u,v)$, is the average squared 
+The third statistic, $\Dt_{l_b,u_b}(u,v)$, is the average squared 
 distance between points satisfying $u_i+v_i>2l_b$ and 
@@
 $\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
-$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\D2_{l_b,u_b}$ is 
+$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dt_{l_b,u_b}$ is 
 then small. 
@@
 the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
-$\D2_{l_b,u_b}$ indicate strong dependence.
+$\Dt_{l_b,u_b}$ indicate strong dependence.
 
@@
 and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
-$\Ps_u=\Ps_{0.5,1}$, $\D2_l=\D2_{0,0.5}$, and $\D2_u=\D2_{0.5,1}$.
+$\Ps_u=\Ps_{0.5,1}$, $\Dt_l=\Dt_{0,0.5}$, and $\Dt_u=\Dt_{0.5,1}$.
 
@@
 <!--Shya to cite the Nelson book-->
-The difference $\D2_u-\D2_l$, in which the "upper" and
+The difference $\Dt_u-\Dt_l$, in which the "upper" and
 "lower" statistics occur in opposite order, is also expected to be 
@@
 $\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
-$\D2_{1-u_b,1-l_b}-\D2_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
+$\Dt_{1-u_b,1-l_b}-\Dt_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
 

### 2018-03-13 SupportingInformation.Rmd
filled in cross ref, citations thing

@@
 tables: True
+
+csl: science.csl
+bibliography: REF_SI.bib
 ---
@@
 \newcommand{\Ps}{\operatorname{P}}
-\newcommand{\Dt}{\text{D}^2}
+\newcommand{\Dsq}{\operatorname{D}^2}
+<!--\newcommand{\D2}{\operatorname{D}^2} gives error msg : insert $ -->
 
@@
 
-# Descriptions of data
+# Descriptions of data \label{Description_data}
 <!--Each data contributor writes one to a few paragraphs about their data.-->
 
-## Soil C and N data
+## Soil C and N data  \label{SoilCN_data}
+
+Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description). The raw data are shown in Figure \ref{fig_soilCNraw}. The copula is shown in Figure \ref{fig_soilCNcop}.
 
 <!--The raw data look like this:-->
-```{r load_data_Loecke, echo=F, fig.cap="Plot of the raw soil C and N data.\\label{fig_soilCNraw}", fig.height=5, fig.width=5}
+```{r load_data_Loecke, echo=F, results='hide'}
+#fig.cap="Plot of the raw soil C and N data.\\label{fig_soilCNraw}", fig.height=3, fig.width=3}
 d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
 d<-d[,c("SOCstock100","TSNstock100")]
-plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1)
+pdf("./Results/SoilCNraw.pdf")
+plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1.5,cex.axis=1.5)
+dev.off()
 ```
@@
 <!--The copula looks like this:-->
-```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data.
-Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations (see YYY).\\label{fig_soilCNcop}",  fig.height=5, fig.width=5}
+```{r rankandplot_Loecke, echo=F, results='hide'}
+#fig.cap="Copula plot of soil C and N data.
+Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations (see Sec.
+\ref{Model_selection}).\\label{fig_soilCNcop}",  fig.height=5, fig.width=5}
+
 #convert to a copula and plot
+pdf("./Results/SoilCNcop.pdf")
 v_CN<-getcopula(d=d,rankon=T,ploton=T) 
-```
-
+dev.off()
+```
 <!--Shya please insert an automatic reference to the methods section "Model selection approach" at the location YYY above.
@@
 I can see there is a package called "bookdown" that may help but there may be something easier using similar tools to what you have already used for figures and tables.-->
-
-Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description). The raw data are shown in Figure \ref{fig_soilCNraw}. The copula is shown in \ref{fig_soilCNcop}.
-
-<!--Shyamolina, let's not put in newpages and other spacing modifications until possibly at the end, once all the content is essentially settled-->
-
-## Aphid data
+![Plot of the raw soil C and N data.\label{fig_soilCNraw}](./Results/SoilCNraw.pdf){ width=30%, height=30% }
+
+![Copula plot of soil C and N data.
+Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations (see Sec.
+\ref{Model_selection}).
+\label{fig_soilCNcop}](./Results/SoilCNcop.pdf){ width=30%, height=30% }
+
+## Aphid data \label{Aphid_data}
    * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
@@
 
-## North Sea plankton data
+## North Sea plankton data \label{Pns_data}
    * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
@@
 
-## Details of methods
-
-### Model selection approach
+## Details of methods \label{Method}
+
+### Model selection approach \label{Model_selection}
 
@@
 Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian*{1}*, **C**layton*{3}*, **G**umbel*{4}*, **F**rank*{5}*, **J**oe*{6}*, **BB1***{7}*, **BB6***{8}*, **BB7***{9}*, **BB8***{10}*, **S**urvival **C**layton*{13}*, **S**urvival **G**umbel*{14}*, **S**urvival **J**oe*{16}*, **S**urvival **BB1***{17}*, **S**urvival **BB6***{18}*, **S**urvival **BB7***{19}* and **S**urvival **BB8***{20}* copulas.
-Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC) and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas (REF).
+Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC) and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas @burnham2003_modelselection.
 <!--Shya, we here need a reference to the Burnham and Anderson book-->
-Descriptions of these copula families can be found in introductory texts on copulas
+Descriptions of these copula families can be found in introductory texts on copulas @burnham2003_modelselection, @nelsen2007_copula
 <!--Shya to please insert references for both the books on copulas we have-->
@@
 model-averaged lower- and upper-tail dependence values using standard model averaging 
-formulas
+formulas @burnham2003_modelselection
 <!--Shya to add reference to Burnham and Anderson, please-->
@@
 of locations, but you need to say that and then maybe a few sentences about how 
-everything was tallied up after having done that-->
-
-### Non-parametric analysis of tail dependence
-
-![Schematic diagram supporting defintions of nonparametric tail dependence statistics.\label{fig_stat_image}](./stat_image.jpg){ width=30%, height=30% }
+everything was tallied up after having done that ??? -->
+
+### Non-parametric analysis of tail dependence \label{NPA}
+
 <!--Shya, can you please change this figure so it shows u (horizontal) and v (vertical) on the axis labels instead of vi and vj?
@@
 variables $u_i$, $v_i$ for $i=1,...,n$ 
-(see section REF for defintions of $u_i$ and $v_i$)
+(see Sec. \ref{Model_selection} for defintions of $u_i$ and $v_i$)
 <!--Shya pls insert ref to the section immediately above this one-->
 are related in the left and right tails, respectively, of their 
-distributions (REF). 
+distributions @nelsen2007_copula. 
 <!--Shya, pls cite the Nelson book-->
@@
 possible. Given two bounds 
-$0 leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
+$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
 $u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
@@
 the right parts of the distributions.
+
+![Schematic diagram supporting defintions of nonparametric tail dependence statistics.\label{fig_stat_image}](./stat_image.jpg){ width=25%, height=25% }
 
@@
 \end{equation}
-As for $\cor_{l_b,u_b}$, larger values of $P_{l_b,u_b}$ 
+As for $\cor_{l_b,u_b}$, larger values of $Ps_{l_b,u_b}$ 
 indicate stronger positive association
@@
 
-The third statistic, $\Dt_{l_b,u_b}(u,v)$, is the average squared 
+The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
 distance between points satisfying $u_i+v_i>2l_b$ and 
@@
 $\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
-$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dt_{l_b,u_b}$ is 
+$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
 then small. 
@@
 the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
-$\Dt_{l_b,u_b}$ indicate strong dependence.
+$\Dsq_{l_b,u_b}$ indicate strong dependence.
 
@@
 and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
-$\Ps_u=\Ps_{0.5,1}$, $\Dt_l=\Dt_{0,0.5}$, and $\Dt_u=\Dt_{0.5,1}$.
+$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.
 
@@
 than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
-This holds, for instance, for the Clayton copula (REF).
+This holds, for instance, for the Clayton copula @nelsen2007_copula.
 <!--Shya to cite the Nelson book-->
-The difference $\Dt_u-\Dt_l$, in which the "upper" and
+The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
 "lower" statistics occur in opposite order, is also expected to be 
@@
 $\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
-$\Dt_{1-u_b,1-l_b}-\Dt_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
+$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
 
@@
 
-### Stat-testing of Non-parametric analysis (for limited number of datapoints)
+### Stat-testing of Non-parametric analysis (for limited number of datapoints) \label{Stat_testing}
 
@@
 Whereas *Normal* and *Frank* have symmetric tail dependence, *Clayton* and *Survival Clayton* have lower and upper tail dependence, respectively.
-First we chose a set of parameters [using *iRho* from *Copula* package] for each type of copula for which spearman correlation will be $10$ equispaced number in between [$0,0.9$].
+First we chose a set of parameters [using *iRho* from `Copula` package] for each type of copula for which spearman correlation will be $10$ equispaced number in between [$0,0.9$].
 Then for each parameter (or spearman value) of each type of copula we carried out $100$ simulation and calculated $(Corl , Coru, Pl, Pu, D2l, D2u)$ stats.
@@
 
-## Soil C and N data
-
-### Model selection approach
+## Soil C and N data \label{Ex_SoilCN}
+
+### Model selection approach \label{Ex_SoilCN_Model_selection}
 <!--This does the computations for this section-->
@@
 
-### Non-parametric analysis of tail dependence
+### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}
 <!--Do the intensive part of the analysis, calculating stats on the data and surrogates-->
@@
 
-## Aphid abundance data for 'Green spruce aphid'
+## Aphid abundance data for 'Green spruce aphid' \label{Ex_count}
 <!--Add common and species name to header-->
 
-### Model selection approach
+### Model selection approach \label{Ex_count_Model_selection}
 
@@
 ```
-
-<!--\newpage-->
 
@@
 
-### Non-paramtric analysis of tail dependence
+### Non-paramtric analysis of tail dependence \label{Ex_count_NPA}
 
@@
 
-## Aphid first flight data for 'Leaf-curling plum aphid' 
+## Aphid first flight data for 'Leaf-curling plum aphid' \label{Ex_ff}
 <!--Add species common name and scientific name to header-->
 
-### Model selection approach
+### Model selection approach \label{Ex_ff_Model_selection}
 
@@
 
-<!--\newpage-->
-
 ```{r tab_ff_bestcop, echo=F, results='markup'}
@@
 
-\newpage
 ```{r tab_ff_AICw, echo=F, results='markup'}
@@
 
-### Non-parametric analysis of tail dependence
+### Non-parametric analysis of tail dependence  \label{Ex_ff_NPA}
 
@@
 
-## North sea plankton data for 'Ceratium furca'
+## North sea plankton data for 'Ceratium furca' \label{Ex_pns}
 
@@
 
-### Model selection approach
+### Model selection approach \label{Ex_pns_Model_selection}
 
@@
 
-<!--\newpage-->
 ```{r tab_pns_bestcop, echo=F, results='markup'}
@@
 
-\newpage
 ```{r tab_pns_AICw,echo=F,results="markup"}
@@
 
-<!--\newpage-->
 ```{r tab_pns_p_KS,echo=F,results="markup"}
@@
 
-### Non-paramtric analysis of tail dependence
+### Non-paramtric analysis of tail dependence  \label{Ex_pns_NPA}
 ```{r stat_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea)}
@@
 
-## Aphid count
+## Aphid count \label{skewness_count}
 
@@
 
-<!--Shya to add results about skewness of spatial means-->
 ```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
@@
 
-## Aphid first flight
+## Aphid first flight \label{skewness_ff}
 
@@
 
-<!--Shya to add results about skewness of spatial means-->
 ```{r skewness_ff_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
@@
 
-## North Sea plankton
+## North Sea plankton \label{skewness_pns}
 
@@
 ![Skewness_plankton_north_sea for species : Ceratium_furca\label{fig_skewness_pns}](./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf){ width=50%, height=50% }
+
+## References:

### 2018-03-13 SupportingInformation.Rmd
inserting \newpage command to get clean table

@@
 <!--\newcommand{\D2}{\operatorname{D}^2} gives error msg : insert $ -->
+
+<!--
+```{r, echo=F}
+options(knitr.table.format = "html")
+```
+-->
 
@@
 ```{r tab_count_bestcop, echo=F, results='markup'}
+#library(dplyr)
 knitr::kable(RES_aphid_count$info_ord_copname[,,1], 
@@
              booktabs=TRUE
-             )
+             )#%>% # only with 'latex'/'html' format
+ # kableExtra::kable_styling(latex_options = c("striped","HOLD_position"))
 ```
@@
 
+\newpage
 ```{r tab_pns_AICw,echo=F,results="markup"}
@@
 
+\newpage
 ```{r tab_pns_p_CvM,echo=F,results="markup"}
@@
 
+\newpage
 ## References:

### 2018-03-13 SupportingInformation.Rmd
new file just begun for new surrogate functionality that will support a Taylor's law analyss. Also, minor changes to comments and section structure in supp mat to accommodate a TL analysis

--- 2018-03-13 SupportingInformation.Rmd
@@
 # Question 2 analyses for each dataset
+Recall that question 2 is: What are some possible causes of non-Gaussian copula structure?
 
 # Question 3 analyses for each dataset
-
-Recall question 3 is: What are the consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?
-
-**Change of skewness : ** Here, we claim that because of significant tail dependence of the multivariate copula structure, the skewness of the time-series data averaged over all selective locations changed significantly from the individual skewness of time-series data taken at each pairwise location.
+Recall that question 3 is: What are the consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?
+
+## Skewness of the spatial mean
+
+Here, we claim that because of significant tail dependence of the multivariate copula structure, the skewness of the time-series data averaged over all selective locations changed significantly from the individual skewness of time-series data taken at each pairwise location.
 Spatial mean of the time series data extracts that tail dependence of the multivariate data set, if any, and reflected in the averaged skewness.
@@
 
-## Aphid count \label{skewness_count}
+### Aphid count \label{skewness_count}
 
@@
 
-## Aphid first flight \label{skewness_ff}
+### Aphid first flight \label{skewness_ff}
 
@@
 
-## North Sea plankton \label{skewness_pns}
+### North Sea plankton \label{skewness_pns}
 
@@
 
-## References:
+## Extinction risk
+
+## Taylor's law
+
+# References:

### 2018-03-13 copsurrognd.R

@@ -1,0 +1,27 @@
+
+
+
+#This function takes a bunch of time series measured at the same
+#times in different locations and creates surrogate datasets ...
+#
+#This differs from copsurrog2d and ncsurrog because...
+#
+#Args
+#m          A N by n matrix, where N is the length of the time series 
+#           and n is the number of time series
+#targetcop  The target copula. This is a copula object from the 
+#             copula package. Must have dimension equal to dim(m)[2]
+
+#Output
+#
+
+copsurrognd<-function(m,tcop)
+{
+  N<-dim(m)[1]
+  n<-dim(m)[2]
+  
+  #error checking
+  #check the dimension of targetcop is right
+
+  
+}

### 2018-03-13 ncsurrog.R

@@ -2,9 +2,16 @@ library(copula)
 library(mvtnorm)
 
 #This function takes a bunch of time series measured at the same
-#time in different locations and creates surrogate datasets which
+#times in different locations and creates surrogate datasets which
 #are statistically similar except the copula has been randomized
-#to a normal copula.
+#to a copula in the normal family. "Statistcally similar" means the location
+#marginals are (exactly) the same and the pairwise kendall or 
+#spearman correlations of time series between locations are 
+#similar.
+#
+#This is better than copsurrod2d because it can do multivariate data
+#but it is worse that copsurrog2d because the target copula can only 
+#be a normal copula.
 #
 #Args
 #m          A N by n matrix, where N is the length of the time series 

### 2018-03-13 SupportingInformation.Rmd

--- 2018-03-13 SupportingInformation.Rmd
@@
 
+## Extinction risk
+
+## Taylor's law
+
 \newpage
-## References:
+# References:

### 2018-03-13 SupportingInformation.Rmd
adding line in Yaml chunk

@@
 tables: True
+link-citations: true
 
@@
 options(knitr.table.format = "html")
+
 ```

### 2018-03-13 SupportingInformation.Rmd
fixing typo

@@
 options(knitr.table.format = "html")
-
 ```
@@
 \end{equation}
-As for $\cor_{l_b,u_b}$, larger values of $Ps_{l_b,u_b}$ 
+As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
 indicate stronger positive association

### 2018-03-14 Tests.Rmd
added a new surrogate routine and tests for it

@@
 
-# Test flexible copula functions
+# Test flexible copula tail dependency statistics
 We here test the functions `Corbds`, `Pbds` and `D2bds`.
@@
 
-# Tests to model selection tool
+Test the `copsurrognd` function:
+```{r test_copsurrognd, echo=T, cache=T, cache.extra=list(mtime("copsurrognd.R"))}
+library("VineCopula")
+source("copsurrognd.R")
+set.seed(101)
+
+#basic tests
+datcop<-claytonCopula(3,dim=3)
+numpts<-1000
+m<-rCopula(numpts,datcop)
+plot(m[,1],m[,2],type='p',asp=1)
+rect(0,0,1,1)
+plot(m[,1],m[,3],type='p',asp=1)
+rect(0,0,1,1)
+plot(m[,2],m[,3],type='p',asp=1)
+rect(0,0,1,1)
+
+tarcop<-gumbelCopula(8,dim=3)
+
+res<-copsurrognd(m=m,targetcop=tarcop,numsurrog=5)
+if (sum(dim(res)[1:2]==dim(m))==2)
+{
+  print("passed")
+} else
+{
+  print("dim of output is wrong")
+}
+
+cor(res[,1,1],res[,2,1],method="kendall")
+cor(res[,1,1],res[,3,1],method="kendall")
+cor(res[,2,1],res[,3,1],method="kendall") #these should be similar to each other and to the following...
+
+gd<-rCopula(1000,tarcop)
+cor(gd[,1],gd[,2],method="kendall")
+cor(gd[,1],gd[,3],method="kendall")
+cor(gd[,2],gd[,3],method="kendall")
+
+#but not to these, because correlations match the target, not the original data
+cor(m[,1],m[,2],method='kendall')  
+cor(m[,1],m[,3],method='kendall')  
+cor(m[,2],m[,3],method='kendall')
+
+plot(res[,1,1],res[,2,1],type='p',asp=1) #should look like a Gumbel
+rect(0,0,1,1)
+plot(res[,1,1],res[,3,1],type='p',asp=1) #should look like a Gumbel
+rect(0,0,1,1)
+plot(res[,2,1],res[,3,1],type='p',asp=1) #should look like a Gumbel
+rect(0,0,1,1)
+BiCopGofTest(res[,1,1],res[,2,1],family=4) #should indicate an acceptable fit; 
+                                            #family=4 is Gumbel
+BiCopGofTest(res[,1,1],res[,3,1],family=4) #should indicate an acceptable fit
+BiCopGofTest(res[,2,1],res[,3,1],family=4) #should indicate an acceptable fit
+
+if (sum(sort(res[,1,1])==sort(m[,1]))==numpts &&
+    sum(sort(res[,2,1])==sort(m[,2]))==numpts &&
+    sum(sort(res[,3,1])==sort(m[,3]))==numpts &&
+    sum(sort(res[,1,2])==sort(m[,1]))==numpts &&
+    sum(sort(res[,2,2])==sort(m[,2]))==numpts &&
+    sum(sort(res[,3,2])==sort(m[,3]))==numpts)
+{
+  print("passed")
+} else
+{
+  print("Columns of output of copsurrog should be permutations of the corresponding 
+        columns of the input")
+}
+```
+
+# Tests of model selection tool
 

### 2018-03-14 SupportingInformation.Rmd
added a chunk in support of a Taylor's law analysis

@@
 
+<!--make surrogate datasets for TL analysis to follow-->
+```{r taylors_law_make_surrogs, echo=F, cache=T, cache.extra=list(seed,mtime("copsurrognd.R"))}
+source("copsurrognd.R")
+set.seed(seed)
+
+#***constants
+N<-100 #length of time series
+n<-25 #number of sampling locations
+numdat<-25 #the number of datasets to get from the normal copula
+rho<-0.7 #normal-copula parameter for dependence between locations
+          #this is a covariance, see below call to rmvnorm
+
+#***generate the starting data, which has normal copula structure
+#because it comes from a multivariate normal
+sig<-matrix(rho,n,n)
+diag(sig)<-1
+m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
+m<-array(m,c(N,numdat,n))
+m<-aperm(m,c(1,3,2))
+gshape<-7.5
+gscale<-1 #gamma parameters
+cop_to_marg_trans<-function(x)
+{
+  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
+                                  #should be something skewed
+}
+marg_to_cop_trans<-function(x) #transforms to copula space
+{
+  return(pgamma(x,shape=gshape,scale=gscale))
+}
+m<-cop_to_marg_trans(pnorm(m))
+
+#***surrogates using a Clayton copula, kendall preserving
+#get the Clayton parameter to use
+ncop<-normalCopula(rho,2)
+ccop<-claytonCopula(2,2)
+cparam<-iTau(ccop,tau(ncop))
+ccop<-claytonCopula(cparam,n)
+
+#generate the surrogates
+csurrog<-array(NA,dim(m))
+for (counter in 1:dim(m)[3])
+{
+  csurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=ccop,numsurrog=1)
+}
+
+# #test to see whether kendalls are similar between m and csurrog - 
+#only works at all as a test if N is large (I used 1000)
+# cor(m[,1,1],m[,2,1],method='kendall')
+# cor(csurrog[,1,1],csurrog[,2,1],method='kendall')
+# cor(m[,3,1],m[,4,1],method='kendall')
+# cor(csurrog[,3,1],csurrog[,4,1],method='kendall')
+# cor(m[,5,1],m[,6,1],method='kendall')
+# cor(csurrog[,5,1],csurrog[,6,1],method='kendall')
+# 
+# cor(m[,1,2],m[,2,2],method='kendall')
+# cor(csurrog[,1,2],csurrog[,2,2],method='kendall')
+# cor(m[,3,2],m[,4,2],method='kendall')
+# cor(csurrog[,3,2],csurrog[,4,2],method='kendall')
+# cor(m[,5,2],m[,6,2],method='kendall')
+# cor(csurrog[,5,2],csurrog[,6,2],method='kendall')
+
+#***surrogates using a Frank copula, kendall preserving
+#get the Frank parameter to use
+fcop<-frankCopula(4,2)
+fparam<-iTau(fcop,tau(ncop))
+fcop<-frankCopula(fparam,n)
+
+#generate the surrogates
+fsurrog<-array(NA,dim(m))
+for (counter in 1:dim(m)[3])
+{
+  fsurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=fcop,numsurrog=1)
+}
+
+#***surrogates using a survival Clayton, kendall preserving
+#take the clayton surrogates and transform them to the 
+#copula space, then flip, then transform back
+scsurrog<-cop_to_marg_trans((1-marg_to_cop_trans(csurrog)))
+
+# #test: should have identical taus as csurrog
+# cor(csurrog[,1,1],csurrog[,2,1],meth='kendall')
+# cor(scsurrog[,1,1],scsurrog[,2,1],meth='kendall')
+# cor(csurrog[,1,2],csurrog[,2,2],meth='kendall')
+# cor(scsurrog[,1,2],scsurrog[,2,2],meth='kendall')
+
+kend.surrog<-list(cla=csurrog,nor=m,fra=fsurrog,scl=scsurrog)
+
+#***surrogates using a Clayton copula, spearman preserving
+#get the Clayton parameter to use
+ccop<-claytonCopula(2,2)
+cparam<-iRho(ccop,rho(ncop))
+ccop<-claytonCopula(cparam,n)
+
+#generate the surrogates
+csurrog<-array(NA,dim(m))
+for (counter in 1:dim(m)[3])
+{
+  csurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=ccop,numsurrog=1)
+}
+
+#***surrogates using a Frank copula, spearman preserving
+#get the Frank parameter to use
+fcop<-frankCopula(4,2)
+fparam<-iRho(fcop,rho(ncop))
+fcop<-frankCopula(fparam,n)
+
+#generate the surrogates
+fsurrog<-array(NA,dim(m))
+for (counter in 1:dim(m)[3])
+{
+  fsurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=fcop,numsurrog=1)
+}
+
+#***surrogates using a survival Clayton, spearman preserving
+#take the clayton surrogates and transform them to the 
+#copula space, then flip, then transform back
+scsurrog<-cop_to_marg_trans((1-marg_to_cop_trans(csurrog)))
+
+spea.surrog<-list(cla=csurrog,nor=m,fra=fsurrog,scl=scsurrog)
+```
+
 \newpage

### 2018-03-14 All_analysis_theory.Rmd
building on taylor's law analysis

@@
     res[,,tct+1]<-lam*res[,,tct]
-    res[,,tct+1]<-lam*res[,,tct]
 
@@
 If I decide to change how dispersal works in the multi-population model, one way to do it might be to assume the pops are located along a transect and dispersal only occurs to the one or two (depending on whether you are on the end or not) neighbors.
+
+Shyamolina, can you please build on what I have started here in some of the ways listed?
+Use your judgement and explore whatever results seem promising and at least somewhat biologically relevant.
+Perhaps a good place to start would be generalizing popsim_ml to take a dispersal matrix which links each location to each other location in a general way.
+That will make it easier to then experiment with local dispersal, instead of the global dispersal I have implemented.

### 2018-03-14 SupportingInformation.Rmd
building on taylor's law analysis

@@
 
+```{r tl_function, echo=F}
+source("TaylorsLawTools.R")
+```
+
+```{r tl_analysis, echo=F, cache=T, cache.extra=list(mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
+#We are only doing spatial taylor's law because we know a priori temporal
+#won't give anything interesting for the surrogates we use
+kend.stats.res<-list()
+kend.points.res<-list()
+spea.stats.res<-list()
+spea.points.res<-list()
+for (counter in 1:length(kend.surrog))
+{
+  ks<-kend.surrog[[counter]]
+  ss<-spea.surrog[[counter]]
+  kend.stats.res[[counter]]<-matrix(NA,5,numdat)
+  spea.stats.res[[counter]]<-matrix(NA,5,numdat)
+  kend.points.res[[counter]]<-array(NA,c(N,2,numdat))
+  spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
+  for (sc in 1:numdat)
+  {
+    khold<-tl_stats(ks[,,sc])
+    kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
+    rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
+    kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)
+
+    shold<-tl_stats(ss[,,sc])
+    spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
+    rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
+    spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
+  }
+}
+names(kend.stats.res)<-names(kend.surrog)
+names(kend.points.res)<-names(kend.surrog)
+names(spea.stats.res)<-names(spea.surrog)
+names(spea.points.res)<-names(spea.surrog)
+
+#DAN: now you want to compare whether tl held, and the slope (statistically and visually)
+#across the different elements of kend.XX.res and spea.XX.res. See drawing of a figure layout on desk.
+
+```
+
 \newpage

### 2018-03-14 SupportingInformation.Rmd
minor changes

--- 2018-03-14 SupportingInformation.Rmd
@@
 
+<!--do a TL analysis on the data just created-->
+```{r tl_analysis, echo=F, cache=T, cache.extra=list(mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
+source("TaylorsLawTools.R")
+
+#We are only doing spatial taylor's law because we know a priori temporal
+#won't give anything interesting for the surrogates we use
+
+kend.stats.res<-list()
+kend.points.res<-list()
+spea.stats.res<-list()
+spea.points.res<-list()
+for (counter in 1:length(kend.surrog))
+{
+  ks<-kend.surrog[[counter]]
+  ss<-spea.surrog[[counter]]
+  kend.stats.res[[counter]]<-matrix(NA,5,numdat)
+  spea.stats.res[[counter]]<-matrix(NA,5,numdat)
+  kend.points.res[[counter]]<-array(NA,c(N,2,numdat))
+  spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
+  for (sc in 1:numdat)
+  {
+    khold<-tl_stats(ks[,,sc])
+    kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
+    rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
+    kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)
+
+    shold<-tl_stats(ss[,,sc])
+    spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
+    rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
+    spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
+  }
+}
+names(kend.stats.res)<-names(kend.surrog)
+names(kend.points.res)<-names(kend.surrog)
+names(spea.stats.res)<-names(spea.surrog)
+names(spea.points.res)<-names(spea.surrog)
+
+#DAN: now you want to compare whether tl held, and the slope (statistically and visually)
+#across the different elements of kend.XX.res and spea.XX.res. See drawing of a figure layout on desk.
+
+```
+
 \newpage

### 2018-03-16 SupportingInformation.Rmd
adding Ext_risk_function

--- 2018-03-14 SupportingInformation.Rmd
@@
 source("skewness_multivar_dataset.R")
+source("ExtremeTailDep.R")
 ```
@@
 First we chose a set of parameters [using *iRho* from `Copula` package] for each type of copula for which spearman correlation will be $10$ equispaced number in between [$0,0.9$].
-Then for each parameter (or spearman value) of each type of copula we carried out $100$ simulation and calculated $(Corl , Coru, Pl, Pu, D2l, D2u)$ stats.
-We took mean of our statistic $Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ over $100$ simulations and In Figure \ref{fig_stat_testing35}, we plotted those statistic against spearman correlation.
+Then for each parameter (or spearman value) of each type of copula we carried out $100$ simulation and calculated ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$) stats.
+We took mean of our statistic $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ over $100$ simulations and In Figure \ref{fig_stat_testing35}, we plotted those statistic against spearman correlation.
 We put the standard error bar around each point in the plot and these errors got minimized with increasing numbe of data points [Figure \ref{fig_stat_testing1000}].
-Statistic $Corl - Coru$, $Pl-Pu$ and $D2u-D2l$ should be $0$ for spearman=$0$ for *Clayton* and *Survival Clayton* and should remain $0$ for all values of spearman for symmetric copula like *Normal* and *Frank*.
+Statistic $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ should be $0$ for spearman=$0$ for *Clayton* and *Survival Clayton* and should remain $0$ for all values of spearman for symmetric copula like *Normal* and *Frank*.
 To verify this, we executed t-test and report the $p$-values for each case and most of the case shows $p>0.05$.
@@
 
-```{r stat_testing_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,callfn,mtime("StatTestNumericResults.RData"))}
+```{r stat_testing_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,callfn,mtime("StatTestNumericResults.RData"),mtime("TestStats_multivar_dataset.R"))}
 set.seed(seed)
@@
 mtext(side=1,line=1,text="Diagonal slice")
-mtext(side=2,line=1,text=expression(D[2]))
+mtext(side=2,line=1,text=expression(D^{2}))
 axis(side=1,labels=F)
@@
 ```{r plot_soilCN_asymmtry_taildep, echo=F}
-atdres<-data.frame(Statistic=c("Corl-Coru","Pl-Pu","D2u-D2l"),
+atdres<-data.frame(Statistic=c("$cor_l-cor_u$","$P_l-P_u$","$D^2_u-D^2_l$"),
                    Kendall=rep("",3),
@@
 
-```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count)}
+```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("NonParamStat.R"))}
 set.seed(seed) 
@@
 ```{r tab_count_resamp,echo=F,results="markup"}
-knitr::kable(stat_aphid_count$numericdf,
+s<-stat_aphid_count$numericdf[,2:5]
+rownames(s)<-c("spear","kend","cor~l~","cor~u~","P~l~","P~u~","$D^2$~l~","$D^2$~u~","cor~l~ - cor~u~","P~l~ - P~u~","$D^2$~u~ - $D^2$~l~")
+knitr::kable(s,
              format = "pandoc",
@@
 
-![Corl-Coru vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Corl-Coru}](./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
-
-![Pl-Pu vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Pl-Pu}](./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
-
-![D2u-D2l vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_D2u-D2l}](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
+![$cor_l$-$cor_u$ vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Corl-Coru}](./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
+
+![$P_l$-$P_u$ vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Pl-Pu}](./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
+
+![$D^2_u$-$D^2_l$ vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_D2u-D2l}](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
 
@@
 Fourth column named $fracgt0$ in Table \ref{tab_count_resamp} indicates the fraction of resampled mean greater then zero.
-$Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ statistic with positive mean and almost $100$% $fracgt0$ proves that *Green spruce aphid*-count data has significant lower tail dependence.
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ statistic with positive mean and almost $100$% $fracgt0$ proves that *Green spruce aphid*-count data has significant lower tail dependence.
 
@@
 
-```{r stat_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff)}
+```{r stat_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("NonParamStat.R"))}
 set.seed(seed)
@@
 ```{r tab_ff_resamp,echo=F,results="markup"}
-knitr::kable(stat_aphid_ff$numericdf,
+s<-stat_aphid_ff$numericdf[,2:5]
+rownames(s)<-c("spear","kend","cor~l~","cor~u~","P~l~","P~u~","$D^2$~l~","$D^2$~u~","cor~l~ - cor~u~","P~l~ - P~u~","$D^2$~u~ - $D^2$~l~")
+knitr::kable(s,
              format = "pandoc",
@@
 
-![Corl-Coru vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Corl-Coru}](./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
-
-![Pl-Pu vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Pl-Pu}](./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
-
-![D2u-D2l vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_D2u-D2l}](./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
+![$cor_l$-$cor_u$ vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Corl-Coru}](./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
+
+![$P_l$-$P_u$ vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Pl-Pu}](./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
+
+![$D^2_u$-$D^2_l$ vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_D2u-D2l}](./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
 
@@
 Fourth column named $fracgt0$ in Table \ref{tab_ff_resamp} indicates the fraction of resampled mean greater then zero.
-$Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ statistic with negative mean has all resampled mean less than $0$ for *Leaf-curling plum aphid*-firstflight data.
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ statistic with negative mean has all resampled mean less than $0$ for *Leaf-curling plum aphid*-firstflight data.
 This proves that there is a significant upper tail dependence in *Leaf-curling plum aphid*-firstflight data.
@@
 ### Non-paramtric analysis of tail dependence  \label{Ex_pns_NPA}
-```{r stat_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea)}
+```{r stat_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("NonParamStat.R"))}
 set.seed(seed)
@@
 ```{r tab_pns_resamp,echo=F,results="markup"}
-knitr::kable(stat_plankton_north_sea$numericdf,
+s<-stat_plankton_north_sea$numericdf[,2:5]
+rownames(s)<-c("spear","kend","cor~l~","cor~u~","P~l~","P~u~","$D^2$~l~","$D^2$~u~","cor~l~ - cor~u~","P~l~ - P~u~","$D^2$~u~ - $D^2$~l~")
+knitr::kable(s,
              format = "pandoc",
@@
 
-![Corl-Coru vs.
+![$cor_l$-$cor_u$ vs.
 Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_Corl-Coru}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
 
-![Pl-Pu vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_Pl-Pu}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
-
-![D2u-D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_D2u-D2l}](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
+![$P_l$-$P_u$ vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_Pl-Pu}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
+
+![$D^2_u$-$D^2_l$ vs.
+Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_D2u-D2l}](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
 
@@
 Fourth column named $fracgt0$ in Table \ref{tab_pns_resamp} indicates the fraction of resampled mean greater then zero.
-$Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ statistic with positive mean and almost $100$% $fracgt0$ proves that plankton species data for *Ceratium furca* has significant lower tail dependence.
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ statistic with positive mean and almost $100$% $fracgt0$ proves that plankton species data for *Ceratium furca* has significant lower tail dependence.
 
@@
 
-![Skewness_count for species : Green spruce aphid\label{fig_skewness_count}](./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf){ width=50%, height=50% }
+![Skewness_count for species : Green spruce aphid\label{fig_skewness_count}](./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf){ width=40%, height=40% }
 
@@
 
-![Skewness_firstflight for species : Leaf-curling plum aphid\label{fig_skewness_ff}](./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf){ width=50%, height=50% }
+![Skewness_firstflight for species : Leaf-curling plum aphid\label{fig_skewness_ff}](./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf){ width=40%, height=40% }
 
@@
 
-![Skewness_plankton_north_sea for species : Ceratium_furca\label{fig_skewness_pns}](./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf){ width=50%, height=50% }
+![Skewness_plankton_north_sea for species : Ceratium_furca\label{fig_skewness_pns}](./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf){ width=40%, height=40% }
 
 ## Extinction risk
+```{r fn_popsim_ml_D, echo=F}
+
+# A function to generate dispersal matrix D (a square matrix : numlocs by numlocs)
+Dispersal_mat<-function(numlocs,d){
+  D<-matrix(0,numlocs,numlocs)
+  # A companion matrix (delta) that indicates how "off" a diagonal is:
+  mat <- matrix(seq_len(numlocs ^ 2), ncol = numlocs)
+  delta <- row(mat) - col(mat)
+  #print(delta)
+
+  diag(D)<-(1-d)
+  D[delta ==1 | delta==-1] <- d/2
+
+  D[numlocs-1,numlocs]<-d
+  D[2,1]<-d
+
+  stopifnot(all(colSums(D)==1)) # this is a check
+  return(D)
+}
+#-----------------------------------------------------------------------------------
+#The simulator of the multi-location model with a dispersal matrix.
+#-------Args--------------------------------------------------
+#p0       A length numlocs vector holding initial populations
+#ns       An numsims by numlocs by numsteps array of the epsilons, 
+#           where numsteps is the number of time steps you want
+#
+#--------Output------------------------------------------------------
+#A numsims by numlocs by numsteps+1 array of populations
+#----------------------------------------------------------------------
+popsim_ml_D<-function(p0,ns,D,r,K){
+  numsims<-dim(ns)[1]
+  numlocs<-dim(ns)[2]
+  numsteps<-dim(ns)[3]
+
+  #initialization
+  res<-array(NA,c(numsims,numlocs,numsteps+1))
+  res[,,1]<-rep(p0,each=numsims)
+
+  #step the populations forward
+  for (tct in 1:numsteps){
+
+    #stochastic ricker model : if r=0 then back to LC model
+
+    #growth rates based on the noise
+    lam_sto<-exp(ns[,,tct])  #numsims by numlocs matrix
+    lam_ricker<-exp(r*(1-(res[,,tct]/K)))
+
+    #growth prior to dispersal
+    res[,,tct+1]<-lam_ricker*lam_sto*res[,,tct] #numsims by numlocs matrix
+
+    #after dispersal
+    tres<-t(res[,,tct+1]) # numlocs by numsims matrix after taking transpose
+    temp<-D%*%tres  # a matrix multiplication (numlocs by numlocs) * (numlocs by numsims)
+    res[,,tct+1]<-t(temp)
+
+    #see if there is extinction
+    h<-res[,,tct+1]
+    h[h<1]<-0
+    res[,,tct+1]<-h
+  }
+
+  return(res)
+}
+
+# function to calculate extinction risk
+extrisk<-function(sims){
+  totpop<-apply(FUN=sum,X=sims,MARGIN=c(1,3))
+  return(apply(FUN=sum,X=(totpop==0),MARGIN=2)/dim(sims)[1])
+}
+```
+
+```{r fn_plotter_ext_risk, echo=F}
+# function to optionally plot extinction risk agsinst time and give you back extinction risk after specified numsteps 
+plotter_ext_risk<-function(numsims,numsteps,numlocs,d,r,K,ploton){
+  D<-Dispersal_mat(numlocs = numlocs,d=d)
+  ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1)# a righttail dep matrix(numpoints by numlocs,     
+#                                                      numpoints=numsteps*numsims)
+
+  ns1<-array(ns1,c(numsteps,numsims,numlocs))# convert to an array (numsteps by numsims by numlocs)
+  ns1<-aperm(ns1,c(2,3,1)) # convert to an array (numsims by numlocs by numsteps)
+  pops1<-popsim_ml_D(p0=rep(25,numlocs),ns=ns1,D=D,r=r,K=K)
+  risk_right<-extrisk(pops1)
+
+  ns2<-(-ns1)
+  pops2<-popsim_ml_D(p0=rep(25,numlocs),ns=ns2,D=D,r=r,K=K)
+  risk_left<-extrisk(pops2)
+
+  if(ploton==T){
+   par(mfrow=c(2,1))
+    plot(0:(length(risk_right)-1),risk_right,type='l',
+         xlab='Time',ylab='Risk',col='blue',panel.first = grid())
+    mtext("blue : right tail-dep")
+    plot(0:(length(risk_left)-1),risk_left,type='l',xlab='Time',ylab='Risk',col='red',panel.first = grid())
+    mtext("red : left tail-dep")
+  }
+
+  return(list(risk_right_after_numsteps=risk_right[numsteps+1],
+              risk_left_after_numsteps=risk_left[numsteps+1]))
+
+}
+
+```
+
+```{r fn_varying_d, echo=F}
+varying_d<-function(numsims,numsteps,numlocs,r,K){
+  risk_right<-c()
+  risk_left<-c()
+  d_seq<-seq(from=0,to=1,by=0.01)
+  for(d in d_seq){
+    riskrl<- plotter_ext_risk(numsims=numsims,numsteps = numsteps,numlocs = numlocs,d=d,r=r,K=K,ploton=F)
+    risk_r<-riskrl$risk_right_after_numsteps
+    risk_l<-riskrl$risk_left_after_numsteps
+    risk_right<-c(risk_right,risk_r)
+    risk_left<-c(risk_left,risk_l)
+  }
+  par(mfrow=c(1,2))
+  plot(d_seq,risk_right,xlab='d',ylab='Risk_right',ylim=c(0,max(risk_right)),type="b",col="blue",panel.first = grid())
+  plot(d_seq,risk_left,xlab='d',ylab='Risk_left',ylim=c(0,max(risk_left)),type="b",col="red",panel.first = grid())
+  #mtext("red : left tail-dep, blue : right tail-dep")
+}
+```
+
+```{r fig_risk_K50, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=50, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_K50}", fig.height=4, fig.width=8}
+varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=50)
+```
+
+```{r fig_risk_K100, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_K100}", fig.height=4, fig.width=8}
+varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=100)
+```
+
+```{r fig_risk_K200, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=200, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_K200}", fig.height=4, fig.width=8}
+varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=200)
+```
+
+```{r fig_risk_numlocs5, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=5, numsteps=25, numsims=1000.\\label{fig_risk_numlocs5}", fig.height=4, fig.width=8}
+varying_d(numsims=1000,numsteps =25,numlocs=5,r=1,K=100)
+```
+
+```{r fig_risk_numlocs15, echo=F, fig.cap="Plot Extinction risk vs.
+d(degree of dispersal) for r=1, K=100, numlocs=15, numsteps=25, numsims=1000.\\label{fig_risk_numlocs15}", fig.height=4, fig.width=8}
+varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=100)
+```
+
+```{r fig_risk_numlocs25, echo=F, fig.cap="Plot Extinction risk vs.
+d(degree of dispersal) for r=1, K=100, numlocs=25, numsteps=25, numsims=1000.\\label{fig_risk_numlocs25}", fig.height=4, fig.width=8}
+varying_d(numsims=1000,numsteps =25,numlocs=15,r=1,K=100)
+```
+
+  * With increasing $K$, extinction risk decreases [for a fixed $d$ , $r$ and $numlocs$] but this effect is more prominent in
+case of left tail-dep data. [see Figure \ref{fig_risk_K50} - Figure \ref{fig_risk_K200}]
+
+  * With increasing $numlocs$, extinction risk decreases [for a fixed $d$ , $r$ and $K$] but this effect is more prominent in
+case of left tail-dep data. [see Figure \ref{fig_risk_numlocs5} - Figure \ref{fig_risk_numlocs25}]
+
+\newpage
 
@@
 
+<!--do a TL analysis on the data just created-->
+```{r tl_analysis, echo=F, cache=T, cache.extra=list(mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
+#source("TaylorsLawTools.R")
+
+#We are only doing spatial taylor's law because we know a priori temporal
+#won't give anything interesting for the surrogates we use
+
+#kend.stats.res<-list()
+#kend.points.res<-list()
+#spea.stats.res<-list()
+#spea.points.res<-list()
+#for (counter in 1:length(kend.surrog))
+#{
+ # ks<-kend.surrog[[counter]]
+#  ss<-spea.surrog[[counter]]
+ # kend.stats.res[[counter]]<-matrix(NA,5,numdat)
+#  spea.stats.res[[counter]]<-matrix(NA,5,numdat)
+ # kend.points.res[[counter]]<-array(NA,c(N,2,numdat))
+  #spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
+  #for (sc in 1:numdat)
+  #{
+   # khold<-tl_stats(ks[,,sc])
+  #  kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
+   # rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
+    #kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)
+
+    #shold<-tl_stats(ss[,,sc])
+    #spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
+    #rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
+    #spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
+  #}
+#}
+#names(kend.stats.res)<-names(kend.surrog)
+#names(kend.points.res)<-names(kend.surrog)
+#names(spea.stats.res)<-names(spea.surrog)
+#names(spea.points.res)<-names(spea.surrog)
+
+#DAN: now you want to compare whether tl held, and the slope (statistically and visually)
+#across the different elements of kend.XX.res and spea.XX.res. See drawing of a figure layout on desk.
+
+```
+
 \newpage

### 2018-03-17 SupportingInformation.Rmd
adding ext risk for LC model

@@
   par(mfrow=c(1,2))
+  plot(d_seq,risk_left,xlab='d',ylab='Risk_left',ylim=c(0,max(risk_left)),type="b",col="red",panel.first = grid())
   plot(d_seq,risk_right,xlab='d',ylab='Risk_right',ylim=c(0,max(risk_right)),type="b",col="blue",panel.first = grid())
-  plot(d_seq,risk_left,xlab='d',ylab='Risk_left',ylim=c(0,max(risk_left)),type="b",col="red",panel.first = grid())
   #mtext("red : left tail-dep, blue : right tail-dep")
@@
 
-```{r fig_risk_K50, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=50, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_K50}", fig.height=4, fig.width=8}
-varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=50)
-```
-
-```{r fig_risk_K100, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_K100}", fig.height=4, fig.width=8}
+### LC model :
+```{r fig_risk_r0_numlocs5, echo=F, fig.cap="LC model : Plot Extinction risk vs.
+d(degree of dispersal) for numlocs=5, numsteps=25, numsims=1000.\\label{fig_risk_r0_numlocs5}", fig.height=4, fig.width=8}
+varying_d(numsims=1000,numsteps =25,numlocs=5,r=0,K=50)
+```
+
+```{r fig_risk_r0_numlocs25, echo=F, fig.cap="LC model : Plot Extinction risk vs.
+d(degree of dispersal) for numlocs=25, numsteps=25, numsims=1000.\\label{fig_risk_r0_numlocs25}", fig.height=4, fig.width=8}
+varying_d(numsims=1000,numsteps =25,numlocs=25,r=0,K=50)
+```
+
+  *  As expected, when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
+With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] but this effect was more prominent in case of data came from right tail-dep environmental noise.
+[see Figure \ref{fig_risk_r0_numlocs5} - Figure \ref{fig_risk_r0_numlocs25}].
+
+### Ricker model :
+
+```{r fig_risk_r1_K100, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_r1_K100}", fig.height=4, fig.width=8}
 varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=100)
@@
 
-```{r fig_risk_K200, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=200, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_K200}", fig.height=4, fig.width=8}
+```{r fig_risk_r1_K200, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=200, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_r1_K200}", fig.height=4, fig.width=8}
 varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=200)
@@
 
-```{r fig_risk_numlocs5, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=5, numsteps=25, numsims=1000.\\label{fig_risk_numlocs5}", fig.height=4, fig.width=8}
+```{r fig_risk_r1_numlocs5, echo=F, fig.cap="Plot Extinction risk vs.
+d(degree of dispersal) for r=1, K=100, numlocs=5, numsteps=25, numsims=1000.\\label{fig_risk_r1_numlocs5}", fig.height=4, fig.width=8}
 varying_d(numsims=1000,numsteps =25,numlocs=5,r=1,K=100)
@@
 
-```{r fig_risk_numlocs15, echo=F, fig.cap="Plot Extinction risk vs.
-d(degree of dispersal) for r=1, K=100, numlocs=15, numsteps=25, numsims=1000.\\label{fig_risk_numlocs15}", fig.height=4, fig.width=8}
-varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=100)
-```
-
 ```{r fig_risk_numlocs25, echo=F, fig.cap="Plot Extinction risk vs.
-d(degree of dispersal) for r=1, K=100, numlocs=25, numsteps=25, numsims=1000.\\label{fig_risk_numlocs25}", fig.height=4, fig.width=8}
-varying_d(numsims=1000,numsteps =25,numlocs=15,r=1,K=100)
+d(degree of dispersal) for r=1, K=100, numlocs=25, numsteps=25, numsims=1000.\\label{fig_risk_r1_numlocs25}", fig.height=4, fig.width=8}
+varying_d(numsims=1000,numsteps =25,numlocs=25,r=1,K=100)
 ```
@@
   * With increasing $K$, extinction risk decreases [for a fixed $d$ , $r$ and $numlocs$] but this effect is more prominent in
-case of left tail-dep data. [see Figure \ref{fig_risk_K50} - Figure \ref{fig_risk_K200}]
+case of left tail-dep data. [see Figure \ref{fig_risk_r1_K100} - Figure \ref{fig_risk_r1_K200}]
 
   * With increasing $numlocs$, extinction risk decreases [for a fixed $d$ , $r$ and $K$] but this effect is more prominent in
-case of left tail-dep data. [see Figure \ref{fig_risk_numlocs5} - Figure \ref{fig_risk_numlocs25}]
+case of left tail-dep data. [see Figure \ref{fig_risk_r1_numlocs5} - Figure \ref{fig_risk_r1_numlocs25}]
 

### 2018-03-20 SupportingInformation.Rmd
cannot remember what I changed

--- 2018-03-14 SupportingInformation.Rmd
@@
 names(spea.points.res)<-names(spea.surrog)
-
-#DAN: now you want to compare whether tl held, and the slope (statistically and visually)
-#across the different elements of kend.XX.res and spea.XX.res. See drawing of a figure layout on desk.
-
+```
+
+<!--Now make some plots-->
+```{r taylor_plots, echo=T}
+#plot TL for clayton
+xlimits<-range(kend.points.res[[1]][,1,],kend.points.res[[2]][,1,],kend.points.res[[3]][,1,],kend.points.res[[4]][,1,])
+ylimits<-range(kend.points.res[[1]][,2,],kend.points.res[[2]][,2,],kend.points.res[[3]][,2,],kend.points.res[[4]][,2,])
+par(mfrow=c(1,1))
+plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
+```
+
+```{r, echo=T}
+#plot TL for normal copula
+plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
+```
+
+```{r, echo=T}
+#plot TL for frank
+plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
+```
+
+```{r, echo=T}
+#plot TL for survival clayton
+plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
 ```

### 2018-03-20 SupportingInformation.Rmd

--- 2018-03-17 SupportingInformation.Rmd
@@
 ```{r tl_analysis, echo=F, cache=T, cache.extra=list(mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
-#source("TaylorsLawTools.R")
-
-#We are only doing spatial taylor's law because we know a priori temporal
-#won't give anything interesting for the surrogates we use
-
-#kend.stats.res<-list()
-#kend.points.res<-list()
-#spea.stats.res<-list()
-#spea.points.res<-list()
-#for (counter in 1:length(kend.surrog))
-#{
+# #source("TaylorsLawTools.R")
+# 
+# #We are only doing spatial taylor's law because we know a priori temporal
+# #won't give anything interesting for the surrogates we use
+# 
+# kend.stats.res<-list()
+# kend.points.res<-list()
+# spea.stats.res<-list()
+# spea.points.res<-list()
+# for (counter in 1:length(kend.surrog))
+# {
 #   ks<-kend.surrog[[counter]]
@@
 #   kend.points.res[[counter]]<-array(NA,c(N,2,numdat))
-  #spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
-  #for (sc in 1:numdat)
-  #{
+#   spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
+#   for (sc in 1:numdat)
+#   {
 #     khold<-tl_stats(ks[,,sc])
@@
 #     rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
-    #kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)
-
-    #shold<-tl_stats(ss[,,sc])
-    #spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
-    #rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
-    #spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
-  #}
-#}
-#names(kend.stats.res)<-names(kend.surrog)
-#names(kend.points.res)<-names(kend.surrog)
-#names(spea.stats.res)<-names(spea.surrog)
-#names(spea.points.res)<-names(spea.surrog)
-
-#DAN: now you want to compare whether tl held, and the slope (statistically and visually)
-#across the different elements of kend.XX.res and spea.XX.res. See drawing of a figure layout on desk.
-
+#     kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)
+# 
+#     shold<-tl_stats(ss[,,sc])
+#     spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
+#     rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
+#     spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
+#   }
+# }
+# names(kend.stats.res)<-names(kend.surrog)
+# names(kend.points.res)<-names(kend.surrog)
+# names(spea.stats.res)<-names(spea.surrog)
+# names(spea.points.res)<-names(spea.surrog)
+```
+
+<!--Now make some plots-->
+```{r taylor_plots, echo=T}
+##plot TL for clayton
+#xlimits<-range(kend.points.res[[1]][,1,],kend.points.res[[2]][,1,],kend.points.res[[3]][,1,],kend.points.res[[4]][,1,])
+#ylimits<-range(kend.points.res[[1]][,2,],kend.points.res[[2]][,2,],kend.points.res[[3]][,2,],kend.points.res[[4]][,2,])
+#par(mfrow=c(1,1))
+#plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
+```
+
+```{r, echo=T}
+##plot TL for normal copula
+#plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
+```
+
+```{r, echo=T}
+#plot TL for frank
+#plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
+```
+
+```{r, echo=T}
+#plot TL for survival clayton
+#plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
 ```

### 2018-03-20 SupportingInformation.Rmd
activate Taylors law section

--- 2018-03-17 SupportingInformation.Rmd
@@
 tables: True
-link-citations: true
+link-citations: True
 
@@
 ```{r tl_analysis, echo=F, cache=T, cache.extra=list(mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
-#source("TaylorsLawTools.R")
-
-#We are only doing spatial taylor's law because we know a priori temporal
-#won't give anything interesting for the surrogates we use
-
-#kend.stats.res<-list()
-#kend.points.res<-list()
-#spea.stats.res<-list()
-#spea.points.res<-list()
-#for (counter in 1:length(kend.surrog))
-#{
- # ks<-kend.surrog[[counter]]
-#  ss<-spea.surrog[[counter]]
- # kend.stats.res[[counter]]<-matrix(NA,5,numdat)
-#  spea.stats.res[[counter]]<-matrix(NA,5,numdat)
- # kend.points.res[[counter]]<-array(NA,c(N,2,numdat))
-  #spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
-  #for (sc in 1:numdat)
-  #{
-   # khold<-tl_stats(ks[,,sc])
-  #  kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
-   # rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
-    #kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)
-
-    #shold<-tl_stats(ss[,,sc])
-    #spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
-    #rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
-    #spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
-  #}
-#}
-#names(kend.stats.res)<-names(kend.surrog)
-#names(kend.points.res)<-names(kend.surrog)
-#names(spea.stats.res)<-names(spea.surrog)
-#names(spea.points.res)<-names(spea.surrog)
-
-#DAN: now you want to compare whether tl held, and the slope (statistically and visually)
-#across the different elements of kend.XX.res and spea.XX.res. See drawing of a figure layout on desk.
-
+source("TaylorsLawTools.R")
+# 
+# #We are only doing spatial taylor's law because we know a priori temporal
+# #won't give anything interesting for the surrogates we use
+# 
+ kend.stats.res<-list()
+ kend.points.res<-list()
+ spea.stats.res<-list()
+ spea.points.res<-list()
+ for (counter in 1:length(kend.surrog))
+ {
+   ks<-kend.surrog[[counter]]
+   ss<-spea.surrog[[counter]]
+   kend.stats.res[[counter]]<-matrix(NA,5,numdat)
+   spea.stats.res[[counter]]<-matrix(NA,5,numdat)
+   kend.points.res[[counter]]<-array(NA,c(N,2,numdat))
+   spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
+   for (sc in 1:numdat)
+   {
+     khold<-tl_stats(ks[,,sc])
+     kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
+     rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
+     kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)
+# 
+     shold<-tl_stats(ss[,,sc])
+     spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
+     rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
+     spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
+   }
+ }
+ names(kend.stats.res)<-names(kend.surrog)
+ names(kend.points.res)<-names(kend.surrog)
+ names(spea.stats.res)<-names(spea.surrog)
+ names(spea.points.res)<-names(spea.surrog)
+```
+
+<!--Now make some plots-->
+```{r taylor_plots, echo=T}
+##plot TL for clayton
+xlimits<-range(kend.points.res[[1]][,1,],kend.points.res[[2]][,1,],kend.points.res[[3]][,1,],kend.points.res[[4]][,1,])
+ylimits<-range(kend.points.res[[1]][,2,],kend.points.res[[2]][,2,],kend.points.res[[3]][,2,],kend.points.res[[4]][,2,])
+par(mfrow=c(1,1))
+plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
+```
+
+```{r, echo=T}
+##plot TL for normal copula
+plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
+```
+
+```{r, echo=T}
+#plot TL for frank
+plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
+```
+
+```{r, echo=T}
+#plot TL for survival clayton
+plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
 ```

### 2018-03-20 SupportingInformation.Rmd
changing table captions

@@
              format='pandoc',
-             caption = "Aphid_count : model averaged lower tail dependence statistic for best fit copula family for each selective location pair\\label{tab_count_LTdep_AICw}",
+             caption = "Aphid_count : model averaged lower tail dependence statistic for each selective location pair\\label{tab_count_LTdep_AICw}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Aphid_count : model averaged upper tail dependence statistic for best fit copula family for each selective location pair\\label{tab_count_UTdep_AICw}",
+             caption = "Aphid_count : model averaged upper tail dependence statistic for each selective location pair\\label{tab_count_UTdep_AICw}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Aphid_count : model averaged lower tail dependence statistic for best fit copula family for each selective location pair\\label{tab_count_LTmUT}",
+             caption = "Aphid_count : model averaged lower tail dependence - model averaged upper tail dependence statistic for each selective location pair\\label{tab_count_LTmUT}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Aphid_firstflight : model averaged lower tail dependence statistic for best fit copula family for each selective location pair\\label{tab_ff_LTdep_AICw}",
+             caption = "Aphid_firstflight : model averaged lower tail dependence statistic for each selective location pair\\label{tab_ff_LTdep_AICw}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Aphid_firstflight : model averaged upper tail dependence statistic for best fit copula family for each selective location pair\\label{tab_ff_UTdep_AICw}",
+             caption = "Aphid_firstflight : model averaged upper tail dependence statistic for each selective location pair\\label{tab_ff_UTdep_AICw}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Aphid_firstflight : model averaged lower tail dependence statistic for best fit copula family for each selective location pair\\label{tab_ff_LTmUT}",
+             caption = "Aphid_firstflight : model averaged lower tail dependence - model averaged upper tail dependence statistic for each selective location pair\\label{tab_ff_LTmUT}",
              booktabs=TRUE
@@
              digits=3,
-             caption = "Plankton_north_sea : model averaged lower tail dependence statistic for best fit copula family for each selective location pair\\label{tab_pns_LTdep_AICw}",
+             caption = "Plankton_north_sea : model averaged lower tail dependence statistic for each selective location pair\\label{tab_pns_LTdep_AICw}",
              booktabs=TRUE)
@@
              digits=3,
-             caption = "Plankton_north_sea : model averaged upper tail dependence statistic for best fit copula family for each selective location pair\\label{tab_pns_UTdep_AICw}",
+             caption = "Plankton_north_sea : model averaged upper tail dependence statistic for each selective location pair\\label{tab_pns_UTdep_AICw}",
              booktabs=TRUE)
@@
              digits=3,
-             caption = "Plankton_north_sea : model averaged lower tail - model averaged upper tail dependence statistic for best fit copula family for each of the selective location pair\\label{tab_pns_LTmUT}",
+             caption = "Plankton_north_sea : model averaged lower tail - model averaged upper tail dependence statistic for each of the selective location pair\\label{tab_pns_LTmUT}",
              booktabs=TRUE)

### 2018-03-23 SupportingInformation.Rmd
minor correction

@@
 Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_count_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1.
-These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)`** out of **`r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-RES_aphid_count$num_indep_loc_pair`** location pair with greater lower tail dependence.
+These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)`** out of **`r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`** location pair with greater lower tail dependence.
 
@@
 Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_ff_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1.
-These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.
+These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.
 
@@
 Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_pns_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1.
-These results do convincingly show non-Gaussian copula structure, having `r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-RES_plankton_north_sea$num_indep_loc_pair` location pair more lower tail dependence.
+These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)`** out of **`r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`** location pair more lower tail dependence.
 

### 2018-03-27 SupportingInformation.Rmd
filled in ext risk section

@@
 source("ExtremeTailDep.R")
+source("mydispersal.R")
 ```
@@
 ## Extinction risk
-```{r fn_popsim_ml_D, echo=F}
-
-# A function to generate dispersal matrix D (a square matrix : numlocs by numlocs)
-Dispersal_mat<-function(numlocs,d){
-  D<-matrix(0,numlocs,numlocs)
-  # A companion matrix (delta) that indicates how "off" a diagonal is:
-  mat <- matrix(seq_len(numlocs ^ 2), ncol = numlocs)
-  delta <- row(mat) - col(mat)
-  #print(delta)
-
-  diag(D)<-(1-d)
-  D[delta ==1 | delta==-1] <- d/2
-
-  D[numlocs-1,numlocs]<-d
-  D[2,1]<-d
-
-  stopifnot(all(colSums(D)==1)) # this is a check
-  return(D)
-}
-#-----------------------------------------------------------------------------------
-#The simulator of the multi-location model with a dispersal matrix.
-#-------Args--------------------------------------------------
-#p0       A length numlocs vector holding initial populations
-#ns       An numsims by numlocs by numsteps array of the epsilons, 
-#           where numsteps is the number of time steps you want
-#
-#--------Output------------------------------------------------------
-#A numsims by numlocs by numsteps+1 array of populations
-#----------------------------------------------------------------------
-popsim_ml_D<-function(p0,ns,D,r,K){
-  numsims<-dim(ns)[1]
-  numlocs<-dim(ns)[2]
-  numsteps<-dim(ns)[3]
-
-  #initialization
-  res<-array(NA,c(numsims,numlocs,numsteps+1))
-  res[,,1]<-rep(p0,each=numsims)
-
-  #step the populations forward
-  for (tct in 1:numsteps){
-
-    #stochastic ricker model : if r=0 then back to LC model
-
-    #growth rates based on the noise
-    lam_sto<-exp(ns[,,tct])  #numsims by numlocs matrix
-    lam_ricker<-exp(r*(1-(res[,,tct]/K)))
-
-    #growth prior to dispersal
-    res[,,tct+1]<-lam_ricker*lam_sto*res[,,tct] #numsims by numlocs matrix
-
-    #after dispersal
-    tres<-t(res[,,tct+1]) # numlocs by numsims matrix after taking transpose
-    temp<-D%*%tres  # a matrix multiplication (numlocs by numlocs) * (numlocs by numsims)
-    res[,,tct+1]<-t(temp)
-
-    #see if there is extinction
-    h<-res[,,tct+1]
-    h[h<1]<-0
-    res[,,tct+1]<-h
-  }
-
-  return(res)
-}
-
-# function to calculate extinction risk
-extrisk<-function(sims){
-  totpop<-apply(FUN=sum,X=sims,MARGIN=c(1,3))
-  return(apply(FUN=sum,X=(totpop==0),MARGIN=2)/dim(sims)[1])
-}
-```
-
-```{r fn_plotter_ext_risk, echo=F}
-# function to optionally plot extinction risk agsinst time and give you back extinction risk after specified numsteps 
-plotter_ext_risk<-function(numsims,numsteps,numlocs,d,r,K,ploton){
-  D<-Dispersal_mat(numlocs = numlocs,d=d)
-  ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1)# a righttail dep matrix(numpoints by numlocs,     
-#                                                      numpoints=numsteps*numsims)
-
-  ns1<-array(ns1,c(numsteps,numsims,numlocs))# convert to an array (numsteps by numsims by numlocs)
-  ns1<-aperm(ns1,c(2,3,1)) # convert to an array (numsims by numlocs by numsteps)
-  pops1<-popsim_ml_D(p0=rep(25,numlocs),ns=ns1,D=D,r=r,K=K)
-  risk_right<-extrisk(pops1)
-
-  ns2<-(-ns1)
-  pops2<-popsim_ml_D(p0=rep(25,numlocs),ns=ns2,D=D,r=r,K=K)
-  risk_left<-extrisk(pops2)
-
-  if(ploton==T){
-   par(mfrow=c(2,1))
-    plot(0:(length(risk_right)-1),risk_right,type='l',
-         xlab='Time',ylab='Risk',col='blue',panel.first = grid())
-    mtext("blue : right tail-dep")
-    plot(0:(length(risk_left)-1),risk_left,type='l',xlab='Time',ylab='Risk',col='red',panel.first = grid())
-    mtext("red : left tail-dep")
-  }
-
-  return(list(risk_right_after_numsteps=risk_right[numsteps+1],
-              risk_left_after_numsteps=risk_left[numsteps+1]))
-
-}
-
-```
-
-```{r fn_varying_d, echo=F}
-varying_d<-function(numsims,numsteps,numlocs,r,K){
-  risk_right<-c()
-  risk_left<-c()
-  d_seq<-seq(from=0,to=1,by=0.01)
-  for(d in d_seq){
-    riskrl<- plotter_ext_risk(numsims=numsims,numsteps = numsteps,numlocs = numlocs,d=d,r=r,K=K,ploton=F)
-    risk_r<-riskrl$risk_right_after_numsteps
-    risk_l<-riskrl$risk_left_after_numsteps
-    risk_right<-c(risk_right,risk_r)
-    risk_left<-c(risk_left,risk_l)
-  }
-  par(mfrow=c(1,2))
-  plot(d_seq,risk_left,xlab='d',ylab='Risk_left',ylim=c(0,max(risk_left)),type="b",col="red",panel.first = grid())
-  plot(d_seq,risk_right,xlab='d',ylab='Risk_right',ylim=c(0,max(risk_right)),type="b",col="blue",panel.first = grid())
-  #mtext("red : left tail-dep, blue : right tail-dep")
-}
-```
 
 ### LC model :
-```{r fig_risk_r0_numlocs5, echo=F, fig.cap="LC model : Plot Extinction risk vs.
-d(degree of dispersal) for numlocs=5, numsteps=25, numsims=1000.\\label{fig_risk_r0_numlocs5}", fig.height=4, fig.width=8}
-varying_d(numsims=1000,numsteps =25,numlocs=5,r=0,K=50)
-```
-
-```{r fig_risk_r0_numlocs25, echo=F, fig.cap="LC model : Plot Extinction risk vs.
-d(degree of dispersal) for numlocs=25, numsteps=25, numsims=1000.\\label{fig_risk_r0_numlocs25}", fig.height=4, fig.width=8}
-varying_d(numsims=1000,numsteps =25,numlocs=25,r=0,K=50)
-```
 
@@
 
+```{r fig_risk_r0_numlocs5, echo=F, cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="LC model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0_numlocs5}", fig.height=4, fig.width=8}
+set.seed(seed)
+varying_d(numsims=10000,numsteps =25,numlocs=5,r=0,K=100)
+```
+
+```{r fig_risk_r0_numlocs25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="LC model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0_numlocs25}", fig.height=4, fig.width=8}
+set.seed(seed)
+varying_d(numsims=10000,numsteps =25,numlocs=25,r=0,K=100)
+```
+
 ### Ricker model :
 
-```{r fig_risk_r1_K100, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_r1_K100}", fig.height=4, fig.width=8}
-varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=100)
-```
-
-```{r fig_risk_r1_K200, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=200, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_r1_K200}", fig.height=4, fig.width=8}
-varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=200)
-```
-
-```{r fig_risk_r1_numlocs5, echo=F, fig.cap="Plot Extinction risk vs.
-d(degree of dispersal) for r=1, K=100, numlocs=5, numsteps=25, numsims=1000.\\label{fig_risk_r1_numlocs5}", fig.height=4, fig.width=8}
-varying_d(numsims=1000,numsteps =25,numlocs=5,r=1,K=100)
-```
-
-```{r fig_risk_numlocs25, echo=F, fig.cap="Plot Extinction risk vs.
-d(degree of dispersal) for r=1, K=100, numlocs=25, numsteps=25, numsims=1000.\\label{fig_risk_r1_numlocs25}", fig.height=4, fig.width=8}
-varying_d(numsims=1000,numsteps =25,numlocs=25,r=1,K=100)
-```
-
-  * With increasing $K$, extinction risk decreases [for a fixed $d$ , $r$ and $numlocs$] but this effect is more prominent in
-case of left tail-dep data. [see Figure \ref{fig_risk_r1_K100} - Figure \ref{fig_risk_r1_K200}]
-
-  * With increasing $numlocs$, extinction risk decreases [for a fixed $d$ , $r$ and $K$] but this effect is more prominent in
-case of left tail-dep data. [see Figure \ref{fig_risk_r1_numlocs5} - Figure \ref{fig_risk_r1_numlocs25}]
+ *  For ricker model when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
+For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
+
+  * With increasing $K$ from $100$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_r0.05} - Figure \ref{fig_risk_r0.05_K200}]
+
+ * With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] but this effect was more prominent in case of data came from left tail-dep environmental noise.
+[see Figure \ref{fig_risk_r1.25} - Figure \ref{fig_risk_r1.25_numlocs25}].
+
+```{r fig_risk_r0.05, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.05}", fig.height=4, fig.width=8}
+set.seed(seed)
+r<-0.05
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r0.05_K200, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.05_K200}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=0.05
+numlocs<-5
+K<-200
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r0.25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.25}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=0.25
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r0.5, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.5}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=0.5
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r0.6, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.6}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=0.6
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r0.7, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.7}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=0.7
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r0.75, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.75}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=0.75
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r0.9, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.9}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=0.9
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r1.0, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.0}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=1.0
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r1.25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.25}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=1.25
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r1.25_numlocs25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.25_numlocs25}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=1.25
+numlocs<-25
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r1.5, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.5}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=1.5
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r1.75, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.75}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=1.75
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
+
+```{r fig_risk_r1.95, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.95}", fig.height=4, fig.width=8}
+set.seed(seed)
+r=1.95
+numlocs<-5
+K<-100
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+```
 

### 2018-03-28 SupportingInformation.Rmd
editing, mostly the tests of the nonparametric stats

--- 2018-03-23 SupportingInformation.Rmd
@@
 most implemented using tools from the `VineCopula` package. Code combining these tools is in the
-[*OurBiCopSelect*](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [*BIVAN*](https://github.com/sghosh89/BIVAN) repository. 
+[`OurBiCopSelect`](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository. 
 <!--Shya, links to code are a good idea! Thanks for figuring this out and adding, we should do something like this in other methods sections.-->
@@
 The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`.
-Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian*{1}*, **C**layton*{3}*, **G**umbel*{4}*, **F**rank*{5}*, **J**oe*{6}*, **BB1***{7}*, **BB6***{8}*, **BB7***{9}*, **BB8***{10}*, **S**urvival **C**layton*{13}*, **S**urvival **G**umbel*{14}*, **S**urvival **J**oe*{16}*, **S**urvival **BB1***{17}*, **S**urvival **BB6***{18}*, **S**urvival **BB7***{19}* and **S**urvival **BB8***{20}* copulas.
-Akaike Information Criterion (AIC), Bayesian Information Criterion (BIC) and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas @burnham2003_modelselection.
-<!--Shya, we here need a reference to the Burnham and Anderson book-->
-Descriptions of these copula families can be found in introductory texts on copulas @burnham2003_modelselection, @nelsen2007_copula
-<!--Shya to please insert references for both the books on copulas we have-->
+Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulas.
+Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC) values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas @burnham2003_modelselection.
+Descriptions of these copula families can be found in introductory texts on copulas @nelsen2007_copula
+<!--Shya to please insert references for the book on copulas by Joe - it is on my desk, come get it please-->
 and in references cited in `VineCopula`. The lower- and upper-tail dependence of the 
@@
 model-averaged lower- and upper-tail dependence values using standard model averaging 
-formulas @burnham2003_modelselection
-<!--Shya to add reference to Burnham and Anderson, please-->
-; likewise for BIC.
+formulas @burnham2003_modelselection; likewise for BIC.
 
@@
 
-<!--Shya, can you please change this figure so it shows u (horizontal) and v (vertical) on the axis labels instead of vi and vj?
-Often times computer variables do not make good mathematical notation, so I have changed this, I also made changes for similar reasons to other notation elsewhere.-->
-
 Left- and right-tail dependence quantify the extent to which two ranked
@@
 (see Sec. \ref{Model_selection} for defintions of $u_i$ and $v_i$)
-<!--Shya pls insert ref to the section immediately above this one-->
 are related in the left and right tails, respectively, of their 
 distributions @nelsen2007_copula. 
-<!--Shya, pls cite the Nelson book-->
 Our nonparametric study of tail dependence uses three
@@
 
-<!--Eventually we will need to add methods on how we use these stats, which is different for multivariate and bivariate data.
-For bivariate data, it involves surrogates, for multivariate data it involves resampling locations.
-So, in particular, Dan will have to insert descriptions of the resampling methods.-->
-
-### Stat-testing of Non-parametric analysis (for limited number of datapoints) \label{Stat_testing}
-
 Here, we want to test our $3$ types of nonparametric tail dependent statistics on both small $(35)$ and large $(1000)$ number of datapoints drawn from $4$ types of bivariate copula, namely, *Clayton* , *Normal*, *Frank* and *Survival Clayton*.
@@
 
+<!--Shya, why is message=F and warning=F? This is risky unless you ran it already without these and 
+know what the warnings are and that we can ignore them. If this is the case, please add a comment here
+explaining why it is safe. If you do not have a complete understanding of the warnings and messages,
+please do not suppress them. Even if you do understand them, there are probably better ways to handle 
+it than supressing them.-->
 ```{r stat_testing_save_result, message=F, warning=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"))}
 set.seed(seed)
+
+#Shya, there are a lot of functions in TestStats_multivar_dataset, do we still need all of them? Please delete any that 
+#we no longer use! You need to keep code clean!
 
@@
 ```{r stat_testing_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,callfn,mtime("StatTestNumericResults.RData"),mtime("TestStats_multivar_dataset.R"))}
-set.seed(seed)
-resultsloc<-"./Results/stat_results/stat_testing/"
+
 b<-callfn$paramlist
-plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b)
-plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b)
-plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b)
-plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b)
+plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b,resultsloc)
+plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b,resultsloc)
+plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b,resultsloc)
+plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b,resultsloc)
 ```
@@
 ![stat_testing_results_datapt_1000\label{fig_stat_testing1000}](./Results/stat_results/stat_testing/res_pt1000_rankT.pdf)
+
+<!--Eventually we will need to add methods on how we use these stats, which is different for multivariate and bivariate data.
+For bivariate data, it involves surrogates, for multivariate data it involves resampling locations.
+So, in particular, Dan will have to insert descriptions of the resampling methods.-->
 

### 2018-03-28 CopulaFunctions_flexible.R

@@ -238,7 +238,6 @@ Pbds<-function(vi,vj,lb,ub){
  
 }
 
-#--------------------------------------------------
 
 
 

### 2018-03-28 TestStats_multivar_dataset.R

@@ -3,13 +3,8 @@
 library(copula)
 library(parallel)
 
-#resultsloc<-"./Results/stat_results/stat_testing/"
-
-#To do
-#1) Add a horizontal dashed line at p=0.05 for the Frank and Normal plots
-
 #***
-#First do the calculations to test the stats
+#calculations to test the stats
 #***
 
 #This function is the worker for testing the stats. It generates numpts
@@ -24,7 +19,7 @@ library(parallel)
 #rank     T or F, apply the ranking to sims or not?
 #
 #Output
-#A 12 by numsims matrix with all the values of all the stats on each sim
+#A 6 by numsims matrix with all the values of all the stats on each sim
 #
 worker<-function(cop,numpts,numsims,rank){
   
@@ -39,7 +34,6 @@ worker<-function(cop,numpts,numsims,rank){
     
     #if rank==T, rank the simulated points
     if (rank==T){
-      
       vi<-(rank(vi)/(length(vi)+1))   # gives the same value if you write as : vi<-pobs(vi)
       vj<-(rank(vj)/(length(vj)+1))   # gives the same value if you write as : vj<-pobs(vj)
     }
@@ -54,17 +48,18 @@ worker<-function(cop,numpts,numsims,rank){
   
   return(res)
 }
+
 #----------------------------------------------------------------------------------------------
 #set up list of copulas on which to call worker using mclapply, start by finding 
 #the copula parameters
-
+#
 # This function takes:
 # Input : spearvals : a sequence of values 0 to 1 which indicates spearman correlation
 # Output : a list of 2 things:
 #            1. coplist : a long list (length=4*length(spearvals)) of 4 types of copula : clayton, normal, frank, survival clayton
 #            2. paramlist : a long list (length=length(coplist)) of parameters of above 4 types of copula, this paramlist
 #                           indicates the parameters for which each copula should have that given specified spearman correlation.
-
+#
 coplist_with_params<-function(spearvals){
   
   taCcop<-claytonCopula(3,2)  # initialize by any old Clayton and Frank and normal copulas
@@ -82,13 +77,14 @@ coplist_with_params<-function(spearvals){
     Fparams[counter]<-iRho(taFcop,spearvals[counter])
   }
   
-  Cparams[1]<-0
-  Fparams[1]<-0
+  Cparams[spearvals==0]<-0
[…]

### 2018-03-28 SupportingInformation.Rmd
finished checking tests of nonparametric stats

@@
 This holds, for instance, for the Clayton copula @nelsen2007_copula.
-<!--Shya to cite the Nelson book-->
 The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
@@
 
-Here, we want to test our $3$ types of nonparametric tail dependent statistics on both small $(35)$ and large $(1000)$ number of datapoints drawn from $4$ types of bivariate copula, namely, *Clayton* , *Normal*, *Frank* and *Survival Clayton*.
-Whereas *Normal* and *Frank* have symmetric tail dependence, *Clayton* and *Survival Clayton* have lower and upper tail dependence, respectively.
-First we chose a set of parameters [using *iRho* from `Copula` package] for each type of copula for which spearman correlation will be $10$ equispaced number in between [$0,0.9$].
-Then for each parameter (or spearman value) of each type of copula we carried out $100$ simulation and calculated ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$) stats.
-We took mean of our statistic $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ over $100$ simulations and In Figure \ref{fig_stat_testing35}, we plotted those statistic against spearman correlation.
-We put the standard error bar around each point in the plot and these errors got minimized with increasing numbe of data points [Figure \ref{fig_stat_testing1000}].
-Statistic $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ should be $0$ for spearman=$0$ for *Clayton* and *Survival Clayton* and should remain $0$ for all values of spearman for symmetric copula like *Normal* and *Frank*.
-To verify this, we executed t-test and report the $p$-values for each case and most of the case shows $p>0.05$.
-
-<!--Shya, why is message=F and warning=F? This is risky unless you ran it already without these and 
-know what the warnings are and that we can ignore them. If this is the case, please add a comment here
-explaining why it is safe. If you do not have a complete understanding of the warnings and messages,
-please do not suppress them. Even if you do understand them, there are probably better ways to handle 
-it than supressing them.-->
-```{r stat_testing_save_result, message=F, warning=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"))}
-set.seed(seed)
-
-#Shya, there are a lot of functions in TestStats_multivar_dataset, do we still need all of them? Please delete any that 
-#we no longer use! You need to keep code clean!
+In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we 
+also tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-Ps_u$, and 
+$\Dsq_u-\Dsq_u$ revealed the asymmetry of tail dependence known to be present in data simulated
+from various copulas. Normal and Frank copula families 
+are known to have symmetric tail dependence,
+i.e., strengths of dependence in the lower and upper tails are the same, for all copulas
+in these families @nelsen2007_copula. Clayton copulas are known to have stronger lower- than 
+upper-tail dependence. And survival Clayton copulas, being rotations of Clayton copulas by 
+$180^\circ$, have stronger upper- than lower tail dependence @nelsen2007_copula.  
+Using the `iRho` function from the `copula` package, for each of the four families normal, 
+Frank, Clayton, and survival Clayton, we obtained parameters for which the expected value of 
+Spearman's $\rho$ was, separately, $0,0.1,0.2,...,0.9$. For each copula family and for each 
+of the selected parameter values we generated  $n$ points, 100 times, and thereby computed 100 values of the 
+statistics $\cor_l-\cor_u$, $\Ps_l-Ps_u$, and $\Dsq_u-\Dsq_u$. This exercise 
+was repeated, separately, 
+for $n=1000$ and for $n=35$, these cases being chosen, respectively, to reflect asymptotic behavior of the 
+statistics and their behavior on the small datasets which are common in some fields
+of ecology. The mean and standard error of the three difference statistics were computed for 
+each copula family, parameter value, and value of $n$, and results were plotted against 
+Spearman's rho (Figures \ref{fig_stat_testing35} and \ref{fig_stat_testing1000}), 
+the expectation being that the difference statistics would be positive for 
+Clayton copulas and negative for survival Clayton copulas (except for the case of $\rho=0$, a 
+boundary case of these families corresponding to independence), and not significantly different 
+from $0$ (we used a $t$-test with a significance threshold of $0.05$) for Frank and normal 
+copulas. These expectations were fullfilled (Figures \ref{fig_stat_testing35} and 
+\ref{fig_stat_testing1000}).
+
+<!--The message=F and warning=F flags are to suppress warnings that come when you pass a parameter
+of 0 to the Frank or Clayton copulas. These are technicallly boundary cases for these families.
+They correspond to the independent copula. The packages we use provide the indepdent copula, but
+issue a warning. Since anyway we want the independent copula, we can suppress the warning. We ran this
+without the warning and it is OK to supress it.-->
+```{r stat_testing_save_result, message=F, warning=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopupulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
+set.seed(seed)
 
 resultsloc<-"./Results/stat_results/stat_testing/"
-callfn<-coplist_with_params(spearvals=seq(from=0,to=0.9,by=0.1))
+spearvals<-seq(from=0,to=0.9,by=0.1)
+callfn<-coplist_with_params(spearvals=spearvals)
 coplist<-callfn$coplist
@@
 b<-callfn$paramlist
-plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b,resultsloc)
-plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b,resultsloc)
-plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b,resultsloc)
-plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b,resultsloc)
-```
-
-![stat_testing_results_datapt_35\label{fig_stat_testing35}](./Results/stat_results/stat_testing/res_pt35_rankT.pdf)
-
-![stat_testing_results_datapt_1000\label{fig_stat_testing1000}](./Results/stat_results/stat_testing/res_pt1000_rankT.pdf)
-
-<!--Eventually we will need to add methods on how we use these stats, which is different for multivariate and bivariate data.
+plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b,resultsloc,spearvals)
+plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b,resultsloc,spearvals)
+plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b,resultsloc,spearvals)
+plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b,resultsloc,spearvals)
+```
+
+<!--Shya, currently your plotter function displays a p-value for each point for the
+middle two columns of panels, but only a p-value for the left-most point for the left 
+column of panels and for the right-most column. Can you please make the left- and right-most
+columns follow the same format at the middle two columns, i.e., they need to have a right axis label from 0 to 1 showing p-value and then show all the p-value points for all values
+of Spearman's rho. -->
+
+![Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_l-\Dsq_u$ reveal the know asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
+See text for details.
+$n=35$.
+\label{fig_stat_testing35}](./Results/stat_results/stat_testing/res_pt35_rankT.pdf)
+
+![Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_l-\Dsq_u$ reveal the know asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
+See text for details.
+$n=1000$.\label{fig_stat_testing1000}](./Results/stat_results/stat_testing/res_pt1000_rankT.pdf)
+
+<!--Eventually we will need to add methods on how we actually use these stats, which is different for multivariate and bivariate data.
 For bivariate data, it involves surrogates, for multivariate data it involves resampling locations.
-So, in particular, Dan will have to insert descriptions of the resampling methods.-->
-
-## Soil C and N data \label{Ex_SoilCN}
+Dan will have to insert descriptions of the surrogate and resampling methods.-->
+
+## Results: soil C and N data \label{Ex_SoilCN}
 

### 2018-03-28 TestStats_multivar_dataset.R

@@ -79,7 +79,7 @@ coplist_with_params<-function(spearvals){
   
   Cparams[spearvals==0]<-0
   Fparams[spearvals==0]<-0
-  Cfparams<-Cparams #for the flipped Clayton, same as the Clayton
+  Cfparams<-Cparams #for the survival (flipped) Clayton, same as the Clayton
   
   #now the copulas themselves
   coplist<-list()
@@ -99,7 +99,7 @@ coplist_with_params<-function(spearvals){
     coplist<-c(coplist,frankCopula(Fparams[counter],2))
   }
   
-  #flipped Claytons
+  #flipped/survival Claytons
   for (counter in 1:length(Cfparams)){
     coplist<-c(coplist,rotCopula(claytonCopula(Cfparams[counter],2)))
   }
@@ -127,6 +127,7 @@ coplist_with_params<-function(spearvals){
 #      1. reslist : A list of 4*length(spearvals) matrices and each is a 6 by numsims matrix, 6 rows are : Corl, Coru, Pl, Pu, D2l, D2u
 #      2. numsims : The number of simulations to do and apply the statistics to
 #      3. j1, j2 :  1 to 6, index of those 6 stats
+#
 stat_list<-function(reslist,numsims,j1,j2){ 
   stat_sms<-matrix(NA,nrow=length(reslist),ncol=numsims)
   for (i in 1:length(reslist)){
@@ -135,8 +136,8 @@ stat_list<-function(reslist,numsims,j1,j2){
   return(stat_sms)
 }
 
-
 # This function gives mean,lowCI,upCI of a vector
+#
 mCI<-function(x){
   m<-mean(x)
   se<-sd(x)/sqrt(length(x))   # denominator should be sqrt(length(x)-1) ?????
@@ -149,6 +150,7 @@ mCI<-function(x){
 #                                                   4 : type of copula(C,N,F,SC)
 #                                                   length of spearvals =10 (here)
 #------------------------------------------------------------------------------------------------------------------------
+#
 M1mM2<-function(reslist,numsims){
   
   
@@ -185,12 +187,11 @@ M1mM2<-function(reslist,numsims){
   return(list(CorlmCoru=CorlmCoru,
               PlmPu=PlmPu,
               D2umD2l=D2umD2l))
-  
 }
 
 #-----------------------------------------------------------------------------------------
 # This is a t-test function which returns a p-value
-
+#
[…]

### 2018-03-28 SupportingInformation.Rmd
checked and edited up through but not including the aphid results

@@
 
+The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
+from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_soilCNfit}).
+The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+
 <!--Display model selection results in a table-->
@@
              caption = "Fitting info for soil C and N dataset.
-Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=Survival Clayton; SG=Survival Gumbel; SJ=Survival Joe.
-An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by 180 degrees).\\label{tab_soilCNfit}",
+Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=survival Clayton; SG=survival Gumbel; SJ=survival Joe.
+An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by $180^\\circ$).\\label{tab_soilCNfit}",
              booktabs=T)
 ```
-
-The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes`. We fitted several copulas
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_soilCNfit}).
-The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
 
@@
 
-Although these results do convincingly show non-Gaussian copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+Although these results do convincingly show non-Gaussian copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
 The next section looks at tail dependence in a non-parametric way.
@@
 
+As described in Methods, each of our statistics, as computed for the 
+ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
+of values of the same statistic with the same ranges computed on 1000 normal-copula 
+surrogate datasets with similar Kendall or Spearman correlations, in different 
+runs for Kendall and Spearman (Figure \ref{fig_soilCN_nonparam}). Because the correlation 
+used in the $\cor$ statistic is the Spearman correlation, Spearman-preserving surrogates
+should be used, so the top-right panel of Figure \ref{fig_soilCN_nonparam}
+is more reliable than the top left. Results indicate that tail dependence of 
+data is stronger in both the lower and upper tails than would be expected under a 
+null hypothesis of a normal copula. The dependence between C and N is also 
+weaker in the middle of the distributions
+than would be expected based on a normal-copula null model.
+
 ![Nonparametric tests for tail dependence and other deviations from a normal copula for soil C and N data.
 As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
-When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than.\label{fig_soilCN_nonparam}](./Results/stat_results/stat_soilCN/CNdata_NonparamTailDep.pdf)
+When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
+When $\cor$ values (respectively, $\Ps$ or $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, stronger than or weaker than for $\Ps$ or $\Dsq$) expected from a normal-copula null hypothesis.
+\label{fig_soilCN_nonparam}](./Results/stat_results/stat_soilCN/CNdata_NonparamTailDep.pdf)
+
+We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_soilCNcop}. 
+Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data.
 
@@
 ```{r plot_soilCN_asymmtry_taildep, echo=F}
-atdres<-data.frame(Statistic=c("$cor_l-cor_u$","$P_l-P_u$","$D^2_u-D^2_l$"),
+atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$","$\\Ps_{0,0.1}-\\Ps_{0.9,1}$","$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
                    Kendall=rep("",3),
@@
              format="pandoc",
-             caption = "Results for tests of whether asymmetry of tail dependence is significant compared to a normal-copula null hypothesis.
-Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the ranges $0-0.1$ and $0.9-1$ (Methods).
-One thousand Kendall-preserving and 1000 Spearman-preserving surrogates were created, and the same statistics were computed for the surrogates.
-A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates.\\label{tab_soilCN_asym}",
+             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.
+Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (Methods).
+One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates.
+A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.
+Recall that Spearman-preserving surrogates are more appropriate for the $\\cor$ statistic (Figure \\ref{fig_soilCN_nonparam} caption).\\label{tab_soilCN_asym}",
              booktabs=T)
@@
 
-As described in Methods, each of our statistics, as computed for the 
-ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
-of values of the same statistic with the same ranges computed on 1000 normal-copula 
-surrogate datasets with similar Kendall or Spearman correlations, in different 
-runs (Figure \ref{fig_soilCN_nonparam}). Because the correlation 
-used in the Cor statistic is the Spearman correlation, Spearman-preserving surrogates
-should be used, so the top-left panel of Figure \ref{fig_soilCN_nonparam}
-should be ignored (DAN: need to fix this later). Results indicate that tail dependence of 
-data is stronger in both the lower and upper tails than would be expected under a 
-null hypothesis of a normal copula. The dependence between C and N is also 
-weaker in the middle of both marginal distributions
-than would be expected based on a normal-copula null model.
-
-We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_soilCNcop}. 
-Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data.
-
-## Aphid abundance data for 'Green spruce aphid' \label{Ex_count}
-<!--Add common and species name to header-->
+## Results: aphid abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
 

### 2018-03-28 SupportingInformation.Rmd
teststats plot done with same format and etc.

@@
 ```
-<!--Shya please insert an automatic reference to the methods section "Model selection approach" at the location YYY above.
-I could not easily find a way to do section references by number that was consistent with the figure and table reference system you already have, but perhaps I missed it.
-In the end I think we will need to be able to refer to sections as well as figures and tables by number.
-I can see there is a package called "bookdown" that may help but there may be something easier using similar tools to what you have already used for figures and tables.-->
+
 ![Plot of the raw soil C and N data.\label{fig_soilCNraw}](./Results/SoilCNraw.pdf){ width=30%, height=30% }
@@
 [`OurBiCopSelect`](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
-<!--Shya, links to code are a good idea! Thanks for figuring this out and adding, we should do something like this in other methods sections.-->
 
@@
 Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC) values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas @burnham2003_modelselection.
-Descriptions of these copula families can be found in introductory texts on copulas @nelsen2007_copula
-<!--Shya to please insert references for the book on copulas by Joe - it is on my desk, come get it please-->
-and in references cited in `VineCopula`. The lower- and upper-tail dependence of the 
+Descriptions of these copula families can be found in introductory texts on copulas @nelsen2007_copula , @joe2014_dependence and in references cited in `VineCopula`.
+The lower- and upper-tail dependence of the
 best-fitting member of each family was also provided by `BiCopEst`.
@@
 without the warning and it is OK to supress it.-->
-```{r stat_testing_save_result, message=F, warning=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopupulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
+<!-- The lines you are talking about appeared as message not as warning(actualy no warning appered for this chunk), so we have to set message=F here-->
+```{r stat_testing_save_result, message=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopupulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
 set.seed(seed)
@@
 
-<!--Shya to edit the prentation below to make it more parallel to the presentation of the model selection results for C/N data above-->
-
 ### Model selection approach \label{Ex_pns_Model_selection}

### 2018-03-28 SupportingInformation.Rmd
gone through aphid count model selection section, editing

--- 2018-03-28 SupportingInformation.Rmd
@@
 
+<!--Load the raw data-->
 ```{r read_aphid_data, echo=F}
@@
 #d0_flightduration<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))
-
-```
-
+```
+
+<!--Do the calculations-->
 ```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("FittingCopula_selective_loc.R"))}
@@
 
+As described in Methods, model selection among copulas was done separately for each pair of the 10 
+sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns 
+corresponding to locations. Table \ref{tab_count_bestcop} shows that, for the large majority of location 
+pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of 
+question 1 for these data. Comparing the 
+AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{tab_count_AICw}) 
+to the AIC weight 
+of the normal copula family for the same pair (Table \ref{tab_count_AICw_normal}) confirms that the degree of support for normal copulas was usually not
+comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{tab_count_p_CvM}) in every case 
+failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
+the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{tab_count_p_KS}),
+demonstrating that the collection of copula families we used was sufficiently broad that the copulas of 
+all pairs of locations could be characterized by at least one of them.
+In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+
 ```{r read_RES_aphid_count, echo=F}
@@
              format='pandoc',
-             caption = "Aphid_count : best fit copula family for each selective location pair\\label{tab_count_bestcop}",
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid. Copulas 
+considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
+survival Joe, survival BB1, survival BB6, survival BB7, survival BB8 copulas.
+Abbreviations were given in Methods.
+NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Aphid_count : AICw for best fit copula family for each selective location pair\\label{tab_count_AICw}",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for abundances of the green spruce aphid.\\label{tab_count_AICw}",
              booktabs=TRUE
@@
 
+```{r tab_count_AICw_normal, echo=F, results='markup'}
+knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
+             format='pandoc',
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for abundances of the green spruce aphid, for comparison with Table \\ref{tab_count_AICw}.
+\\label{tab_count_AICw_normal}",
+             booktabs=TRUE
+             )
+```
+
 ```{r tab_count_p_CvM, echo=F, results='markup'}
@@
              format='pandoc',
-             caption = "Aphid_count : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_count_p_CvM}",
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for abundances of the green spruce aphid.
+Tests were based on bootstrapping as described in Methods.\\label{tab_count_p_CvM}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Aphid_count : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula\\label{tab_count_p_KS}",
+             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for abundances of the green spruce aphid.
+Tests were based on bootstrapping as described in Methods.\\label{tab_count_p_KS}",
              booktabs=TRUE
@@
 
-```{r tab_count_AICw_normal, echo=F, results='markup'}
-knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
-             format='pandoc',
-             caption = "Aphid_count : AICw for 'Normal' copula family for each selective location pair\\label{tab_count_AICw_normal}",
+The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{tab_count_LTdep_AICw} and \ref{tab_count_UTdep_AICw} respectively.
+These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data.
+Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.
+
+```{r tab_count_LTdep_AICw, echo=F, results='markup'}
+knitr::kable(RES_aphid_count$LTdep_AICw,
+             format='pandoc',
+             caption = "Model-averaged lower tail dependence statistic for each location pair for abundances of the green spruce aphid.\\label{tab_count_LTdep_AICw}",
              booktabs=TRUE
@@
 
-```{r tab_count_LTdep_AICw, echo=F, results='markup'}
-knitr::kable(RES_aphid_count$LTdep_AICw,
-             format='pandoc',
-             caption = "Aphid_count : model averaged lower tail dependence statistic for each selective location pair\\label{tab_count_LTdep_AICw}",
+```{r tab_count_UTdep_AICw, echo=F, results='markup'}
+knitr::kable(RES_aphid_count$UTdep_AICw,
+             format='pandoc',
+             caption = "Model averaged upper tail dependence statistic for each location pair for abundances of the green spruce aphid.\\label{tab_count_UTdep_AICw}",
              booktabs=TRUE
@@
 
-```{r tab_count_UTdep_AICw, echo=F, results='markup'}
-knitr::kable(RES_aphid_count$UTdep_AICw,
-             format='pandoc',
-             caption = "Aphid_count : model averaged upper tail dependence statistic for each selective location pair\\label{tab_count_UTdep_AICw}",
-             booktabs=TRUE
-             )
-```
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1.
 
@@
              format='pandoc',
-             caption = "Aphid_count : model averaged lower tail dependence - model averaged upper tail dependence statistic for each selective location pair\\label{tab_count_LTmUT}",
+             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for abundances of the green spruce aphid.\\label{tab_count_LTmUT}",
              booktabs=TRUE
@@
 ```
-
-Table \ref{tab_count_bestcop} shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
-Table \ref{tab_count_AICw} represents the Akaike weights of the best-fit copula for each location pair.
-Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_count_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_count_p_KS}] of the goodness of fit with atleast
-`r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` % successful bootstrap.
-Table \ref{tab_count_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_count_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC.
-The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_count_LTdep_AICw} and Table \ref{tab_count_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_count_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1.
-These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)`** out of **`r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`** location pair with greater lower tail dependence.
-
-The next section looks at tail dependence in a non-parametric way.
 

### 2018-03-29 SupportingInformation.Rmd
added descriptions of 2d surriogates and spatial resampling

@@
 \newcommand{\Dsq}{\operatorname{D}^2}
+\newcommand{\rank}{\operatorname{rank}}
 <!--\newcommand{\D2}{\operatorname{D}^2} gives error msg : insert $ -->
@@
 
+<!--Shya, you added a hyperlink that lead to some of our code earlier in this document,
+and I have added one here, and the knitted pdf has the hyperlink, but the text "basic unit tests" 
+is the same color as all the surrounding text. So the only way you can tell that there is a 
+hyperlink there is if you happen to mouse over the words. We need to have a color that all
+text with a hyperlink is rendered in, preferably automatically. Can you figure out how to 
+make that happen, both here and for the earlier hyperlinks?-->
 In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we 
@@
 
-<!--Shya, currently your plotter function displays a p-value for each point for the
-middle two columns of panels, but only a p-value for the left-most point for the left 
-column of panels and for the right-most column. Can you please make the left- and right-most
-columns follow the same format at the middle two columns, i.e., they need to have a right axis label from 0 to 1 showing p-value and then show all the p-value points for all values
-of Spearman's rho. -->
-
 ![Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_l-\Dsq_u$ reveal the know asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
@@
 
-<!--Eventually we will need to add methods on how we actually use these stats, which is different for multivariate and bivariate data.
-For bivariate data, it involves surrogates, for multivariate data it involves resampling locations.
-Dan will have to insert descriptions of the surrogate and resampling methods.-->
+For bivariate datasets such as the soil C and N data, we compared values of the statistics $\cor_{l_b,u_b}$,
+$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to distributions of values of the same 
+statistics computed on surrogate datasets that 
+were produced from the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
+the same marginal distributions as the empirical data and had very similar Kendall or Spearman correlation, but had
+normal copula structure, and thus our comparisons tested the null hypothesis that our statistics took values on the
+empirical data no different from what would have been expected if the copula structure of the data were normal, but the
+data were otherwise statistically unchanged.
+For bivariate datasets, surrogates were produced as follows.
+
+- Given data $(x_i,y_i)$ for $i=1,\ldots,n$ and a 
+one-parameter target copula family $\mathcal{B}(\theta)$ (here $\theta$ denotes the parameter), let $\tau$ be the
+Kendall correlation of the $x_i$ with the $y_i$ and find the parameter value $\theta_\tau$ for which 
+$\mathcal{B}(\theta_\tau)$ has Kendall correlation $\tau$ (this is possible for the one-parameter copula families of this 
+study using the `iTau` function of the `VineCopula` package). 
+- Generate data $(a_i,b_i)$, $i=1,\ldots,n$ from 
+$\mathcal{B}(\theta_\tau)$. 
+- Permute the ordered set $(x_1,\ldots,x_n)$ such that the permutation
+$(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ has $\rank(x_{\sigma_x(i)})$ equal to $\rank(a_i)$ for all $i$, where 
+$\rank(x_{\sigma_x(i)})$ is the rank of $x_{\sigma_x(i)}$ in the set $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$, and
+$\rank(a_i)$ is the rank of $a_i$ in the set $(a_1,\ldots,a_n)$ (here $\sigma_x$ is a permutation of the 
+indices $1,\ldots,n$). 
+- Likewise permute $(y_1,\ldots,y_n)$
+such that the permutation
+$(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ has $\rank(y_{\sigma_y(i)})$ equal to $\rank(b_i)$ for all $i$ ($\sigma_y$ is 
+another permutation of $1,\ldots,n$).
+- The surrogate dataset is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.
+
+Our code for this algorithm is [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R) in the 
+[`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
+
+Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
+is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a permutation 
+of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as those of the original data. 
+Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
+Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
+will be close to $\tau$ (differences arising only through sampling variation). If the instead of $\theta_\tau$
+a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
+as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
+sampling variation) the Spearman correlation of the empirical data.
+
+The comparison of one of the
+statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the empirical data, to the distribution
+of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
+provides a test of whether dependence in the part of the distribution specified by $l_b$ and $u_b$ is different from
+what would be expected from a null hypothesis of normal copula structure but unchanged marginals and unchanged 
+Kendall (or Spearman)
+correlation. Significant deviations correspond to deviations from normal copula structure. 
+The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, as computed on the 
+empirical data, to the distribution
+of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
+provides a test of asymmetry of tail dependence.
+
+For multivariate datasets, e.g., the aphid abundance or first flight data, each of which consists of time series 
+from 10 locations, a spatial resampling scheme was used instead of the surrogate approach described above.
+The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
+widely used spline correlogram method; the same scheme was also used by @walteretal2017 as part of a method for 
+performing model selection among matrix regression models. For each pair of locations $i \neq j$, the statistic
+of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
+$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The values $s_{i,j}$ were then arranged in
+a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
+$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the average strength of 
+dependence, across all pairs of distinct locations, for data in the portion of the distributions 
+given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
+asymmetry of tail dependence across all pairs of locations. 
+Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a distribution of bootstrapped values.
+Bootstrapping was performed by resampling the locations of measurement, with replacement; this corresponds to 
+resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
+@bjornstadfalck2001, @walteretal2017.
 

### 2018-03-29 SupportingInformation.Rmd
hyperlink color added

@@
 link-citations: True
+urlcolor : blue
 
@@
 
-<!--Shya, you added a hyperlink that lead to some of our code earlier in this document,
-and I have added one here, and the knitted pdf has the hyperlink, but the text "basic unit tests" 
-is the same color as all the surrounding text. So the only way you can tell that there is a 
-hyperlink there is if you happen to mouse over the words. We need to have a color that all
-text with a hyperlink is rendered in, preferably automatically. Can you figure out how to 
-make that happen, both here and for the earlier hyperlinks?-->
 In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we 

### 2018-03-30 SupportingInformation.Rmd
modify Ext risk section to include all plots

@@
 In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we 
-also tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-Ps_u$, and 
+also tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and 
 $\Dsq_u-\Dsq_u$ revealed the asymmetry of tail dependence known to be present in data simulated
@@
 of the selected parameter values we generated  $n$ points, 100 times, and thereby computed 100 values of the 
-statistics $\cor_l-\cor_u$, $\Ps_l-Ps_u$, and $\Dsq_u-\Dsq_u$. This exercise 
+statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_u$. This exercise 
 was repeated, separately, 
@@
 ### LC model :
+```{r fig_risk_r0_numlocs5, echo=F, cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="LC model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0_numlocs5}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+varying_d(numsims=10000,numsteps =25,numlocs=5,r=0,K=50,disp_everywhere=F)
+```
+
+```{r fig_risk_r0_numlocs25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="LC model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0_numlocs25}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+varying_d(numsims=10000,numsteps =25,numlocs=25,r=0,K=50,disp_everywhere=F)
+```
 
   *  As expected, when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
-With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] but this effect was more prominent in case of data came from right tail-dep environmental noise.
-[see Figure \ref{fig_risk_r0_numlocs5} - Figure \ref{fig_risk_r0_numlocs25}].
-
-```{r fig_risk_r0_numlocs5, echo=F, cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="LC model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0_numlocs5}", fig.height=4, fig.width=8}
-set.seed(seed)
-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0,K=100)
-```
-
-```{r fig_risk_r0_numlocs25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="LC model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0_numlocs25}", fig.height=4, fig.width=8}
-set.seed(seed)
-varying_d(numsims=10000,numsteps =25,numlocs=25,r=0,K=100)
-```
+With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_r0_numlocs5} - Figure \ref{fig_risk_r0_numlocs25}].
 
@@
 
-  * With increasing $K$ from $100$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_r0.05} - Figure \ref{fig_risk_r0.05_K200}]
-
- * With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] but this effect was more prominent in case of data came from left tail-dep environmental noise.
-[see Figure \ref{fig_risk_r1.25} - Figure \ref{fig_risk_r1.25_numlocs25}].
+  * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_r0.05} - Figure \ref{fig_risk_r0.05_K200}]
+
+ * With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_r0.05} and Figure \ref{fig_risk_r0.05_numlocs25}].
 
 ```{r fig_risk_r0.05, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.05}", fig.height=4, fig.width=8}
+d(degree of dispersal).\\label{fig_risk_r0.05}", fig.height=3.5, fig.width=8}
 set.seed(seed)
@@
 numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
 ```
@@
 ```{r fig_risk_r0.05_K200, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.05_K200}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=0.05
+d(degree of dispersal).\\label{fig_risk_r0.05_K200}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.05
 numlocs<-5
 K<-200
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
+```
+
+```{r fig_risk_r0.05_numlocs25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.05_numlocs25}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.05
+numlocs<-25
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
 ```
@@
 ```{r fig_risk_r0.25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.25}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=0.25
-numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+d(degree of dispersal).\\label{fig_risk_r0.25}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.25
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
 ```
@@
 ```{r fig_risk_r0.5, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.5}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=0.5
-numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+d(degree of dispersal).\\label{fig_risk_r0.5}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.5
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
 ```
@@
 ```{r fig_risk_r0.6, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.6}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=0.6
-numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+d(degree of dispersal).\\label{fig_risk_r0.6}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.6
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
 ```
@@
 ```{r fig_risk_r0.7, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.7}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=0.7
-numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+d(degree of dispersal).\\label{fig_risk_r0.7}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.7
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
 ```
@@
 ```{r fig_risk_r0.75, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.75}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=0.75
-numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+d(degree of dispersal).\\label{fig_risk_r0.75}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.75
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
 ```
@@
 ```{r fig_risk_r0.9, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.9}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=0.9
-numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+d(degree of dispersal).\\label{fig_risk_r0.9}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.9
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
 ```
@@
 ```{r fig_risk_r1.0, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.0}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=1.0
-numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+d(degree of dispersal).\\label{fig_risk_r1.0}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-1.0
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
 ```
@@
 ```{r fig_risk_r1.25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.25}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=1.25
-numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
-```
-
-```{r fig_risk_r1.25_numlocs25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.25_numlocs25}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=1.25
+d(degree of dispersal).\\label{fig_risk_r1.25}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-1.25
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
+```
+
+```{r fig_risk_r1.5, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.5}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-1.5
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
+```
+
+```{r fig_risk_r1.75, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.75}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-1.75
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
+```
+
+```{r fig_risk_r1.95, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.95}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-1.95
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
+```
+
+# Explore equal_dispersal everywhere
+
+### LC model :
+
+    * No effect here.
+
+### Ricker model :
+
+  * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected.
+[see Figure \ref{fig_risk_r0.05_disp_everywhere} - Figure \ref{fig_risk_r0.05_K200_disp_everywhere}]
+
+ * With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_r0.05_disp_everywhere} and Figure \ref{fig_risk_r0.05_numlocs25_disp_everywhere}].
+
+```{r fig_risk_r0.05_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.05_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.05
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r0.05_K200_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal); With increasing K from 50 to 200, extinction risk decreases.\\label{fig_risk_r0.05_K200_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.05
+numlocs<-5
+K<-200
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r0.05_numlocs25_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal); With increasing numlocs from 5 to 25, extinction risk decreases.\\label{fig_risk_r0.05_numlocs25_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.05
 numlocs<-25
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
-```
-
-```{r fig_risk_r1.5, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.5}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=1.5
-numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
-```
-
-```{r fig_risk_r1.75, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.75}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=1.75
-numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
-```
-
-```{r fig_risk_r1.95, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.95}", fig.height=4, fig.width=8}
-set.seed(seed)
-r=1.95
-numlocs<-5
-K<-100
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r0.25_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.25_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.25
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r0.5_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.5_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.5
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r0.6_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.6_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.6
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r0.7_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.7_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.7
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r0.75_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.75_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.75
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r0.9_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r0.9_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-0.9
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r1.0_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.0_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-1.0
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r1.25_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.25_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-1.25
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r1.5_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.5_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-1.5
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r1.75_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.75_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-1.75
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
+```
+
+```{r fig_risk_r1.95_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
+d(degree of dispersal).\\label{fig_risk_r1.95_disp_everywhere}", fig.height=3.5, fig.width=8}
+set.seed(seed)
+r<-1.95
+numlocs<-5
+K<-50
+varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
 ```

### 2018-03-30 SupportingInformation.Rmd
finished editing up through aphid count

--- 2018-03-29 SupportingInformation.Rmd
@@
 Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
-will be close to $\tau$ (differences arising only through sampling variation). If the instead of $\theta_\tau$
+will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
 a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
@@
 Kendall (or Spearman)
-correlation. Significant deviations correspond to deviations from normal copula structure. 
+correlation. Significant deviations correspond to deviations from normal copula structure. In particular, deviations 
+using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence different from that of a normal copula;
+likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different from that of a normal copula.  
 The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, as computed on the 
@@
 @bjornstadfalck2001, @walteretal2017. 
+<!--Shya, these citations render as (4), (5), whereas they should render as (4, 5), or better still, (Bjornstad & Falck, 2001; Walter et al, 2017) (the author-date callout style is better when space is not a limitation because it gives the reader more information without having to look in the bibliography).
+Can you fix this, please?-->
 
@@
 data is stronger in both the lower and upper tails than would be expected under a 
-null hypothesis of a normal copula. The dependence between C and N is also 
+null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
 weaker in the middle of the distributions
@@
 We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_soilCNcop}. 
-Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data.
+Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.
 
@@
 
+<!--Do the main computations for this section-->
 ```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("NonParamStat.R"))}
@@
 
+Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
+computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
+resampling scheme, as described in Methods (Table \ref{tab_count_resamp}). Confidence intervals of
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
+indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
+exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
+structure do not have, providing further evidence answering the third part of question 1.
+
+<!--Show the statistical results in a table-->
 ```{r tab_count_resamp,echo=F,results="markup"}
-s<-stat_aphid_count$numericdf[,2:5]
-rownames(s)<-c("spear","kend","cor~l~","cor~u~","P~l~","P~u~","$D^2$~l~","$D^2$~u~","cor~l~ - cor~u~","P~l~ - P~u~","$D^2$~u~ - $D^2$~l~")
+s<-stat_aphid_count$numericdf[,c(3,2,4)]
+colnames(s)<-c("Statistic","Lower 95\\% CI", "Mean", "Upper 95\\% CI")
+rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Psq_l - \\Psq_u$","$\\Dsq_u - \\Dsq_l$")
 knitr::kable(s,
              format = "pandoc",
-             caption = "Aphid_count : resampling result for stat indicating lower tail dependence\\label{tab_count_resamp}",
+             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for abundances of the green spruce aphid.\\label{tab_count_resamp}",
              booktabs=TRUE)
@@
 
+<!--Shya, I dont think we need this, so I have commented it out. If you disagree, let me know, otherwise
+please go ahead and delete it and this comment.
 ![copula plot for aphid-count\label{fig_count_allcop}](./Results/stat_results/stat_aphid_count/Sp_10_AllCops.pdf)
-
+-->
+
+<!--Shya, can you please organize these following plots into one plot with mutliple panels?-->
 ![Plot for aphid-count : Spearman vs. Distance\label{fig_count_spear}](./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf){ width=50%, height=50% }
@@
 ![$D^2_u$-$D^2_l$ vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_D2u-D2l}](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
-
-Table \ref{tab_count_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean.
-Fourth column named $fracgt0$ in Table \ref{tab_count_resamp} indicates the fraction of resampled mean greater then zero.
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ statistic with positive mean and almost $100$% $fracgt0$ proves that *Green spruce aphid*-count data has significant lower tail dependence.
 

### 2018-03-30 SupportingInformation.Rmd
fixed minor latex problems that prevented knitting

@@
 s<-stat_aphid_count$numericdf[,c(3,2,4)]
-colnames(s)<-c("Statistic","Lower 95\\% CI", "Mean", "Upper 95\\% CI")
-rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Psq_l - \\Psq_u$","$\\Dsq_u - \\Dsq_l$")
+colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
+rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
 knitr::kable(s,
@@
 
-## Aphid first flight data for 'Leaf-curling plum aphid' \label{Ex_ff}
-<!--Add species common name and scientific name to header-->
+## Results: aphid first flight data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}
 
@@
 
+<!--Do the analysis-->
 ```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("FittingCopula_selective_loc.R"))} 
@@
 
+As in section \ref{Ex_count}, results are presented in $10 \times 10$ matrix format with rows
+and columns corresponding to locations. Table \ref{tab_ff_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the
+
+several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+Table \ref{tab_ff_AICw} represents the Akaike weights of the best-fit copula for each location pair.
+Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_ff_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_ff_p_KS}] of the goodness of fit with atleast
+`r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` % successful bootstrap.
+Table \ref{tab_ff_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_ff_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC.
+The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_ff_LTdep_AICw} and Table \ref{tab_ff_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_ff_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1.
+These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.
+
 ```{r tab_ff_bestcop, echo=F, results='markup'}
@@
              format='pandoc',
-             caption = "Aphid_firstflight : best fit copula family for each selective location pair\\label{tab_ff_bestcop}",
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.
+Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations given in Methods.
+NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.
+\\label{tab_ff_bestcop}",
              booktabs=TRUE
@@
 ```
-
-Table \ref{tab_ff_bestcop} shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
-Table \ref{tab_ff_AICw} represents the Akaike weights of the best-fit copula for each location pair.
-Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_ff_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_ff_p_KS}] of the goodness of fit with atleast
-`r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` % successful bootstrap.
-Table \ref{tab_ff_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_ff_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC.
-The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_ff_LTdep_AICw} and Table \ref{tab_ff_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_ff_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1.
-These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.
 

### 2018-03-30 SupportingInformation.Rmd
finished editing through model selection results for aphid ff

@@
 
-As in section \ref{Ex_count}, results are presented in $10 \times 10$ matrix format with rows
-and columns corresponding to locations. Table \ref{tab_ff_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the
-
-several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
-Table \ref{tab_ff_AICw} represents the Akaike weights of the best-fit copula for each location pair.
-Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_ff_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_ff_p_KS}] of the goodness of fit with atleast
-`r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` % successful bootstrap.
-Table \ref{tab_ff_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_ff_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC.
-The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_ff_LTdep_AICw} and Table \ref{tab_ff_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_ff_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1.
-These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)`** out of **`r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair`** location pair more upper tail dependence.
+As in section \ref{Ex_count_Model_selection}, results are presented in $10 \times 10$ 
+matrix format with rows
+and columns corresponding to locations. Table \ref{tab_ff_bestcop} shows 
+that, for the large majority of location pairs, some non-normal copula 
+was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data. Comparing the AIC weight of the best-fitting 
+(lowest AIC) copula family for each pair
+of sampling locations (Table \ref{tab_ff_AICw}) to the AIC weight of the normal copula 
+family for the same pair (Table \ref{tab_ff_AICw_normal}) confirms that the
+degree of support for normal copulas was usually not comparable to the 
+degree of support for other families. Goodness of fit tests
+based on a Cramer-von Mises statistic (Table \ref{tab_ff_p_CvM}) usually failed to reject the 
+hypothesis that the AIC-best copula family
+was also an objectively adequate description of the data, and likewise for 
+goodness of fit tests based on a Kolmogorov-Smirnov
+statistic (Table \ref{tab_ff_p_KS}). In all cases at least `r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` percent 
+of bootstrap randomizations used for
+goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
 
@@
              format='pandoc',
-             caption = "Aphid_firstflight : AICw for best fit copula family for each selective location pair\\label{tab_ff_AICw}",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_AICw}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_ff_AICw_normal, echo=F, results='markup'}
+knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
+             format='pandoc',
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for first flight data for the leaf-curlung plum aphid, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Aphid_firstflight : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_ff_p_CvM}",
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for first flight data for the leaf-curling plum aphid.
+Tests were based on bootstrapping as described in Methods.\\label{tab_ff_p_CvM}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Aphid_firstflight : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula\\label{tab_ff_p_KS}",
+             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for first flight data for the leaf-curling plum aphid.
+Tests were based on bootstrapping as described in Methods.\\label{tab_ff_p_KS}",
              booktabs=TRUE
@@
 
-```{r tab_ff_AICw_normal, echo=F, results='markup'}
-knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
-             format='pandoc',
-             caption = "Aphid_firstflight : AICw for 'Normal' copula family for each selective location pair\\label{tab_ff_AICw_normal}",
-             booktabs=TRUE
-             )
-```
+The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and 
+the model-averaged (again, AIC based) upper-tail dependence statistic are in 
+Tables \ref{tab_ff_LTdep_AICw} and \ref{tab_ff_UTdep_AICw}, respectively.
+These values were exclusively positive, thereby differing from what a normal copula would give (i.e., $0$), and answering the second part of question 1 for these data - these data have greater tail dependence (lower and upper) than would be expected from a normal-copula nul hypothesis.
 
@@
              format='pandoc',
-             caption = "Aphid_firstflight : model averaged lower tail dependence statistic for each selective location pair\\label{tab_ff_LTdep_AICw}",
+             caption = "Model-averaged lower tail dependence statistic for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_LTdep_AICw}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Aphid_firstflight : model averaged upper tail dependence statistic for each selective location pair\\label{tab_ff_UTdep_AICw}",
+             caption = "Model-averaged upper tail dependence statistic for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_UTdep_AICw}",
              booktabs=TRUE
@@
 ```
+
+The difference between the model-averaged lower-tail dependence statistic and the 
+model-averaged upper-tail dependence statistic (lower minus upper), shown in 
+Table \ref{tab_ff_LTmUT}, was negative for `r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)` 
+out of `r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair` location pairs, answering the third 
+part of question 1. We note this result is opposite the analogous result for aphid abundance
+data for the green spruce aphid, i.e., whereas those abundance data showed stronger lower tail 
+dependendence than upper tail dependence, first flights for the leaf-curling plum aphid showed 
+stronger upper tail dependence than lower tail dependence.
 
@@
              format='pandoc',
-             caption = "Aphid_firstflight : model averaged lower tail dependence - model averaged upper tail dependence statistic for each selective location pair\\label{tab_ff_LTmUT}",
+             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_LTmUT}",
              booktabs=TRUE

### 2018-03-30 SupportingInformation.Rmd
finished editing through the aphid ff results

@@
 
-<!--Shya, can you please organize these following plots into one plot with mutliple panels?-->
+<!--Shya, can you please organize these following plots into one plot with mutliple panels? Once that is done, please add some text describing these results.-->
 ![Plot for aphid-count : Spearman vs. Distance\label{fig_count_spear}](./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf){ width=50%, height=50% }
@@
 
+<!--Do the analysis-->
 ```{r stat_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("NonParamStat.R"))}
@@
 
+Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman 
+correlations, were 
+computed across all pairs of sampling locations, and confidence intervals of these means were 
+computed via a spatial
+resampling scheme, as described in Methods (Table \ref{tab_count_resamp}).
+Confidence intervals of
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded 
+zero and mean values were positive,
+indicating tail dependence.
+Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
+exluded $0$ and were negative, indicating asymmetry of tail dependence, a feature 
+that data with normal copula 
+structure do not have, providing further evidence answering the third part of question 1.
+The contrast here with the analogous results for abundances of the green spruce aphid is 
+again illuminating.
+
+<!--display the main statistical results in a table-->
 ```{r tab_ff_resamp,echo=F,results="markup"}
-s<-stat_aphid_ff$numericdf[,2:5]
-rownames(s)<-c("spear","kend","cor~l~","cor~u~","P~l~","P~u~","$D^2$~l~","$D^2$~u~","cor~l~ - cor~u~","P~l~ - P~u~","$D^2$~u~ - $D^2$~l~")
+s<-stat_aphid_ff$numericdf[,c(3,2,4)]
+colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
+rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
 knitr::kable(s,
              format = "pandoc",
-             caption = "Aphid_firstflight : resampling result for stat indicating upper tail dependence\\label{tab_ff_resamp}",
+             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for first flight data for the leaf-curling plum aphid.\\label{tab_ff_resamp}",
              booktabs=TRUE)
@@
 
+<!--As for the count data, I don't think we need this. If you disagree, let me know, otherwise pls delete the commented text and this comment.
 ![copula plot for aphid-firstflight\label{fig_ff_allcop}](./Results/stat_results/stat_aphid_ff/Sp_11_AllCops.pdf)
-
+-->
+
+<!--Shya, can you please organize these following plots into one plot with mutliple panels? Once that is done, please add some text describing these results.-->
 ![Plot for aphid-firstflight : Spearman vs. Distance\label{fig_ff_spear}](./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf){ width=50%, height=50% }
@@
 
-Table \ref{tab_ff_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean.
-Fourth column named $fracgt0$ in Table \ref{tab_ff_resamp} indicates the fraction of resampled mean greater then zero.
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ statistic with negative mean has all resampled mean less than $0$ for *Leaf-curling plum aphid*-firstflight data.
-This proves that there is a significant upper tail dependence in *Leaf-curling plum aphid*-firstflight data.
+<!--Dan stopped editing here 2018 03 30-->
 

### 2018-03-30 SupportingInformation.Rmd
various changes

@@
 
-## Aphid data \label{Aphid_data}
+## Abundance data for the green spruce aphid and first flight date data for the leaf-curling plum aphid in the UK\label{Aphid_data}
    * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
@@
 
-## North Sea plankton data \label{Pns_data}
+## Abundance data for *Ceratium furca* in UK seas \label{Pns_data}
    * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
@@
    * We use species=16 [Name : "Ceratium_furca"] for Plankton_North_Sea data
+   * *Ceratium furca* is a dicoflagellate commonly found in UK seas
+   * note, when we edit this, we cannot call it the North Seas since data come from around the UK not only the North Sea, unless Shya threw out all grid cells that were outside the North Sea
 
@@
 
-## Results: aphid abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
+## Results: abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
 
@@
 
-## Results: aphid first flight data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}
+## Results: first flight date data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}
 
@@
 
-<!--Dan stopped editing here 2018 03 30-->
-
-## North sea plankton data for 'Ceratium furca' \label{Ex_pns}
+## Results: abundance data for *Ceratium furca* in UK seas \label{Ex_pns}
 
@@
 
+<!--Load the data-->
 ```{r read_plankton_north_sea_data, echo=F}
@@
 
+<!--Do the main analysis-->
 ```{r RES_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"), mtime("FittingCopula_selective_loc.R"))}
@@
 
-```{r tab_pns_bestcop, echo=F, results='markup'}
-knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1],
-             format='pandoc',
-             caption = "Plankton_north_sea : Best fit copula family for each selective location pair\\label{tab_pns_bestcop}",
-             booktabs=TRUE
-             )
-```
-
-\newpage
-```{r tab_pns_AICw,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$info_ord_AICw[,,1],
-             format="pandoc",
-             digits=3,
-             caption = "Plankton_north_sea : AICw for best fit copula family for each selective location pair\\label{tab_pns_AICw}",
-             booktabs=TRUE)
-```
-
-\newpage
-```{r tab_pns_p_CvM,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$gfc_p_CvM,
-             format="pandoc",
-             digits=3,
-             caption = "Plankton_north_sea : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_pns_p_CvM}",
-             booktabs=TRUE)
-```
-
-```{r tab_pns_p_KS,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$gfc_p_KS,
-             format="pandoc",
-             digits=3,
-             caption = "Plankton_north_sea : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula\\label{tab_pns_p_KS}",
-             booktabs=TRUE)
-```
-
-```{r tab_pns_AICw_normal,echo=F,results="markup"}
-knitr::kable(normal_cop_AICw_matrix(r=RES_plankton_north_sea),
-             format="pandoc",
-             digits=3,
-             caption = "Plankton_north_sea : AICw for 'Normal' copula family for each selective location pair\\label{tab_pns_AICw_normal}",
-             booktabs=TRUE)
-```
-
-```{r tab_pns_LTdep_AICw,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$LTdep_AICw,
-             format="pandoc",
-             digits=3,
-             caption = "Plankton_north_sea : model averaged lower tail dependence statistic for each selective location pair\\label{tab_pns_LTdep_AICw}",
-             booktabs=TRUE)
-```
-
-```{r tab_pns_UTdep_AICw,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$UTdep_AICw,
-             format="pandoc",
-             digits=3,
-             caption = "Plankton_north_sea : model averaged upper tail dependence statistic for each selective location pair\\label{tab_pns_UTdep_AICw}",
-             booktabs=TRUE)
-```
-
-```{r tab_pns_LTmUT,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw,
-             format="pandoc",
-             digits=3,
-             caption = "Plankton_north_sea : model averaged lower tail - model averaged upper tail dependence statistic for each of the selective location pair\\label{tab_pns_LTmUT}",
-             booktabs=TRUE)
-```
-
-Table \ref{tab_pns_bestcop} shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data.
+As in previous sections (see also Methods) model selection among copulas was done separately for 
+each pair of sampling locations. There were $14$ sampling locations, so results are presented in 
+$14 \times 14$ matrix format, with rows and columns corresponding to locations. 
+Table \ref{tab_pns_bestcop} shows that, for the large majority of location pairs, some 
+non-normal copula was a better fit, according to AIC, than was the normal copula, answering the 
+first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula 
+family for each pair of sampling locations (Table \ref{tab_pns_AICw}) to the AIC weight of
+the normal copula family for the same pair (Table \ref{tab_pns_AICw_normal})
+
  Table \ref{tab_pns_AICw} represents the Akaike weights of the best-fit copula for each location pair.
@@
 These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)`** out of **`r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`** location pair more lower tail dependence.
+
+```{r tab_pns_bestcop, echo=F, results='markup'}
+knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1],
+             format='pandoc',
+             caption = "Plankton_north_sea : Best fit copula family for each selective location pair\\label{tab_pns_bestcop}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_pns_AICw,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$info_ord_AICw[,,1],
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : AICw for best fit copula family for each selective location pair\\label{tab_pns_AICw}",
+             booktabs=TRUE)
+```
+
+```{r tab_pns_AICw_normal,echo=F,results="markup"}
+knitr::kable(normal_cop_AICw_matrix(r=RES_plankton_north_sea),
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : AICw for 'Normal' copula family for each selective location pair\\label{tab_pns_AICw_normal}",
+             booktabs=TRUE)
+```
+
+```{r tab_pns_p_CvM,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$gfc_p_CvM,
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_pns_p_CvM}",
+             booktabs=TRUE)
+```
+
+```{r tab_pns_p_KS,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$gfc_p_KS,
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula\\label{tab_pns_p_KS}",
+             booktabs=TRUE)
+```
+
+```{r tab_pns_LTdep_AICw,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$LTdep_AICw,
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : model averaged lower tail dependence statistic for each selective location pair\\label{tab_pns_LTdep_AICw}",
+             booktabs=TRUE)
+```
+
+```{r tab_pns_UTdep_AICw,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$UTdep_AICw,
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : model averaged upper tail dependence statistic for each selective location pair\\label{tab_pns_UTdep_AICw}",
+             booktabs=TRUE)
+```
+
+```{r tab_pns_LTmUT,echo=F,results="markup"}
+knitr::kable(RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw,
+             format="pandoc",
+             digits=3,
+             caption = "Plankton_north_sea : model averaged lower tail - model averaged upper tail dependence statistic for each of the selective location pair\\label{tab_pns_LTmUT}",
+             booktabs=TRUE)
+```
 

### 2018-03-30 SupportingInformation.Rmd
added back in a newpage command to get it to knit

@@
 title: "Supporting Information"
-author: "Shyamolina Ghosh and Daniel Reuman"
+author: "S. Ghosh, T. Loecke, C.P. Reid, Daniel Reuman"
 date: "February 12, 2018"
@@
 
+\newpage
 ## Results: abundance data for *Ceratium furca* in UK seas \label{Ex_pns}

### 2018-03-30 SupportingInformation.Rmd
figure presntn changed

--- 2018-03-30 SupportingInformation.Rmd
@@
 
-csl: science.csl
+#csl: science.csl
 bibliography: REF_SI.bib
@@
 Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulas.
-Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC) values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas @burnham2003_modelselection.
-Descriptions of these copula families can be found in introductory texts on copulas @nelsen2007_copula , @joe2014_dependence and in references cited in `VineCopula`.
+Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC) values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection].
+Descriptions of these copula families can be found in introductory texts on copulas [@nelsen2007_copula ; @joe2014_dependence]  and in references cited in `VineCopula`.
 The lower- and upper-tail dependence of the
@@
 model-averaged lower- and upper-tail dependence values using standard model averaging 
-formulas @burnham2003_modelselection; likewise for BIC.
+formulas [@burnham2003_modelselection]; likewise for BIC.
 
@@
 are related in the left and right tails, respectively, of their 
-distributions @nelsen2007_copula. 
+distributions [@nelsen2007_copula]. 
 Our nonparametric study of tail dependence uses three
@@
 than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
-This holds, for instance, for the Clayton copula @nelsen2007_copula.
+This holds, for instance, for the Clayton copula [@nelsen2007_copula].
 The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
@@
 i.e., strengths of dependence in the lower and upper tails are the same, for all copulas
-in these families @nelsen2007_copula. Clayton copulas are known to have stronger lower- than 
+in these families [@nelsen2007_copula]. Clayton copulas are known to have stronger lower- than 
 upper-tail dependence. And survival Clayton copulas, being rotations of Clayton copulas by 
-$180^\circ$, have stronger upper- than lower tail dependence @nelsen2007_copula.
+$180^\circ$, have stronger upper- than lower tail dependence [@nelsen2007_copula].
 Using the `iRho` function from the `copula` package, for each of the four families normal, 
@@
 from 10 locations, a spatial resampling scheme was used instead of the surrogate approach described above.
-The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
-widely used spline correlogram method; the same scheme was also used by @walteretal2017 as part of a method for 
+The scheme is identical to that proposed by [@bjornstadfalck2001] for application in their now 
+widely used spline correlogram method; the same scheme was also used by [@walteretal2017] as part of a method for 
 performing model selection among matrix regression models. For each pair of locations $i \neq j$, the statistic
@@
 resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
-@bjornstadfalck2001, @walteretal2017. 
-<!--Shya, these citations render as (4), (5), whereas they should render as (4, 5), or better still, (Bjornstad & Falck, 2001; Walter et al, 2017) (the author-date callout style is better when space is not a limitation because it gives the reader more information without having to look in the bibliography).
-Can you fix this, please?-->
+[@bjornstadfalck2001 ; @walteretal2017].
 
@@
 
-<!--Shya, I dont think we need this, so I have commented it out. If you disagree, let me know, otherwise
-please go ahead and delete it and this comment.
-![copula plot for aphid-count\label{fig_count_allcop}](./Results/stat_results/stat_aphid_count/Sp_10_AllCops.pdf)
--->
-
 <!--Shya, can you please organize these following plots into one plot with mutliple panels? Once that is done, please add some text describing these results.-->
-![Plot for aphid-count : Spearman vs. Distance\label{fig_count_spear}](./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf){ width=50%, height=50% }
-
-![Plot for aphid-count : Kendall vs. Distance\label{fig_count_kend}](./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf){ width=50%, height=50% }
-
-![$cor_l$-$cor_u$ vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Corl-Coru}](./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
-
-![$P_l$-$P_u$ vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Pl-Pu}](./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
-
-![$D^2_u$-$D^2_l$ vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_D2u-D2l}](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
+\begin{figure}[!h]
+\begin{center}
+\textbf{\;\;\;\;\;\;\;\ (A) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (B)} \\
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf}
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf}\\
+\textbf{\;\;\;\;\;\;\;\ (C) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (D)} \\
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf}
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf}\\
+\textbf{\;\;\;\;\;\;\;\ (E)} \\
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
+\caption{Results for Aphid-count}
+\end{center}\label{fig_aphid_count}
+\end{figure}
 
@@
 
-<!--As for the count data, I don't think we need this. If you disagree, let me know, otherwise pls delete the commented text and this comment.
-![copula plot for aphid-firstflight\label{fig_ff_allcop}](./Results/stat_results/stat_aphid_ff/Sp_11_AllCops.pdf)
--->
-
 <!--Shya, can you please organize these following plots into one plot with mutliple panels? Once that is done, please add some text describing these results.-->
-![Plot for aphid-firstflight : Spearman vs. Distance\label{fig_ff_spear}](./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf){ width=50%, height=50% }
-
-![Plot for aphid-firstflight : Kendall vs. Distance\label{fig_ff_kend}](./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf){ width=50%, height=50% }
-
-![$cor_l$-$cor_u$ vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Corl-Coru}](./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
-
-![$P_l$-$P_u$ vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Pl-Pu}](./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
-
-![$D^2_u$-$D^2_l$ vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_D2u-D2l}](./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
-
+\begin{figure}[!h]
+\begin{center}
+\textbf{\;\;\;\;\;\;\;\ (A) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (B)} \\
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf}
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf}\\
+\textbf{\;\;\;\;\;\;\;\ (C) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (D)} \\
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf}
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf}\\
+\textbf{\;\;\;\;\;\;\;\ (E)} \\
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
+\caption{Results for Aphid-ff}
+\end{center}\label{fig_aphid_ff}
+\end{figure}
+
+\newpage
 ## Results: abundance data for *Ceratium furca* in UK seas \label{Ex_pns}

### 2018-03-30 SupportingInformation.Rmd
I think I got checkpoint to work

--- 2018-03-30 SupportingInformation.Rmd
@@
 title: "Supporting Information"
-author: "S. Ghosh, T. Loecke, C.P. Reid, Daniel Reuman"
+author: "S. Ghosh, T. Loecke, C.P. Reid, D.C. Reuman"
 date: "February 12, 2018"
@@
 \newcommand{\rank}{\operatorname{rank}}
-<!--\newcommand{\D2}{\operatorname{D}^2} gives error msg : insert $ -->
-
-<!--
-```{r, echo=F}
-options(knitr.table.format = "html")
-```
--->
+
+<!--checkpoint package-->
+```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
+library(checkpoint)
+checkloc<-"./.checkpoint"
+if (!dir.exists(checkloc))
+{
+  dir.create(checkloc)
+  checkpoint("2018-03-29",checkpointLocation = checkloc)
+}
+```
 
@@
 knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
+
 seed<-101

### 2018-03-30 SupportingInformation.Rmd

--- 2018-03-30 SupportingInformation.Rmd
@@
 title: "Supporting Information"
-author: "Shyamolina Ghosh and Daniel Reuman"
+author: "S. Ghosh, T. Loecke, C.P. Reid, D.C. Reuman"
 date: "February 12, 2018"

### 2018-03-30 SupportingInformation.Rmd
there was a problem with checkpoint, it would reinstall with every knit. I think I have fixed it

@@
 library(checkpoint)
-checkloc<-"./.checkpoint"
-if (!dir.exists(checkloc))
-{
-  dir.create(checkloc)
-  checkpoint("2018-03-29",checkpointLocation = checkloc)
-}
+#checkloc<-tempdir()
+#if (!dir.exists(paste0(checkloc,"/.checkpoint/")))
+#{
+#  dir.create(paste0(checkloc,"/.checkpoint/"))
+#  checkpoint("2018-03-29",checkpointLocation = checkloc)
+#}
+if (!dir.exists("./.checkpoint/"))
+{
+  dir.create("./.checkpoint/")
+}
+checkpoint("2018-03-29",checkpointLocation = "./")
 ```

### 2018-04-01 SupportingInformation.Rmd
filled in fig description for count and ff results in .Rmd and minor plot label change in .R

@@
 
-### Non-paramtric analysis of tail dependence \label{Ex_count_NPA}
+### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
 
@@
 stat_aphid_count<-readRDS("./Results/stat_results/stat_aphid_count/stat_aphid_count_sp_10.RDS")
-```
-
-Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
-computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
-resampling scheme, as described in Methods (Table \ref{tab_count_resamp}). Confidence intervals of
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
-indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
-exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
-structure do not have, providing further evidence answering the third part of question 1.
-
-<!--Show the statistical results in a table-->
-```{r tab_count_resamp,echo=F,results="markup"}
-s<-stat_aphid_count$numericdf[,c(3,2,4)]
-colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
-rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
-knitr::kable(s,
-             format = "pandoc",
-             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for abundances of the green spruce aphid.\\label{tab_count_resamp}",
-             booktabs=TRUE)
 ```
@@
 \includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
-\caption{Results for Aphid-count}
-\end{center}\label{fig_aphid_count}
+\caption{Non-parametric tail statistics results for green spruce aphid abundance data plotted against geographical distance $(D)$ between two sampling locations; $\textbf{(A)}$ Spearman correlation vs.
+$D$, $\textbf{(B)}$ Kendall correlation vs.
+$D$, $\textbf{(C)}$ $\cor_l-\cor_u$ vs.
+$D$, $\textbf{(D)}$ $\Ps_l-\Ps_u$ vs.
+$D$, $\textbf{(E)}$ $\Dsq_u-\Dsq_l$ vs.
+$D$ plotted for all (10 $\times$ 10 locations) taken pairwise.
+\label{fig_aphid_count}}
+\end{center}
 \end{figure}
+
+We computed Spearman and Kendall correlation along with our three non-parametric statistics ($\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$) as described in Sec.
+\ref{NPA} for  green spruce aphid abundance data taken at all $10$ $\times$ $10$ location pairs (code given in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R)).
+We also constructed a $10$ $\times$ $10$ matrix using `gcdist` from `ncf` package where each element in that matrix represented the geographical distance between any two selected locations.
+Figure \ref{fig_aphid_count} presented plots for **(A)** Spearman, **(B)** Kendall, **(C)** $\cor_l-\cor_u$, **(D)** $\Ps_l-\Ps_u$ and **(E)** $\Dsq_u-\Dsq_l$ statistics against the geographical distance for all selected sampling location pairs.
+With increasing the distance *D*, the Spearman and Kendall correlation (hence the synchrony among time series data) between the location pairs decreased as expected.
+But one cannot conclude the tail dependence of that synchrony from Figure \ref{fig_aphid_count} **(A)** or **(B)**, whereas **(C - E)** clearly showed the lower tail dependence with most of the data points lied in the positive quadrant as described before in Sec.
+\ref{NPA}.
+
+Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
+computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
+resampling scheme, as described in Methods (Table \ref{tab_count_resamp}). Confidence intervals of
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
+indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
+exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
+structure do not have, providing further evidence answering the third part of question 1.
+
+<!--Show the statistical results in a table-->
+```{r tab_count_resamp,echo=F,results="markup"}
+s<-stat_aphid_count$numericdf[,c(3,2,4)]
+colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
+rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
+knitr::kable(s,
+             format = "pandoc",
+             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for abundances of the green spruce aphid.\\label{tab_count_resamp}",
+             booktabs=TRUE)
+```
 
@@
 Tables \ref{tab_ff_LTdep_AICw} and \ref{tab_ff_UTdep_AICw}, respectively.
-These values were exclusively positive, thereby differing from what a normal copula would give (i.e., $0$), and answering the second part of question 1 for these data - these data have greater tail dependence (lower and upper) than would be expected from a normal-copula nul hypothesis.
+These values were exclusively positive, thereby differing from what a normal copula would give (i.e., $0$), and answering the second part of question 1 for these data - these data have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
 
@@
 stat_aphid_ff<-readRDS("./Results/stat_results/stat_aphid_ff/stat_aphid_ff_sp_11.RDS")
-```
-
-Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman 
-correlations, were 
-computed across all pairs of sampling locations, and confidence intervals of these means were 
-computed via a spatial
-resampling scheme, as described in Methods (Table \ref{tab_count_resamp}).
-Confidence intervals of
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded 
-zero and mean values were positive,
-indicating tail dependence.
-Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
-exluded $0$ and were negative, indicating asymmetry of tail dependence, a feature 
-that data with normal copula 
-structure do not have, providing further evidence answering the third part of question 1.
-The contrast here with the analogous results for abundances of the green spruce aphid is 
-again illuminating.
-
-<!--display the main statistical results in a table-->
-```{r tab_ff_resamp,echo=F,results="markup"}
-s<-stat_aphid_ff$numericdf[,c(3,2,4)]
-colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
-rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
-knitr::kable(s,
-             format = "pandoc",
-             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for first flight data for the leaf-curling plum aphid.\\label{tab_ff_resamp}",
-             booktabs=TRUE)
 ```
@@
 \includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
-\caption{Results for Aphid-ff}
-\end{center}\label{fig_aphid_ff}
+\caption{Non-parametric tail statistics results for leaf-curling plum aphid's first flight date data plotted against geographical distance $(D)$ between two sampling locations; $\textbf{(A)}$ Spearman correlation vs.
+$D$, $\textbf{(B)}$ Kendall correlation vs.
+$D$, $\textbf{(C)}$ $\cor_l-\cor_u$ vs.
+$D$, $\textbf{(D)}$ $\Ps_l-\Ps_u$ vs.
+$D$, $\textbf{(E)}$ $\Dsq_u-\Dsq_l$ vs.
+$D$ plotted for all (10 $\times$ 10 locations) taken pairwise.
+\label{fig_aphid_ff}}
+\end{center}
 \end{figure}
+
+Figure \ref{fig_aphid_ff} represented plots for **(A)** Spearman correlation, **(B)** Kendall correlation, **(C)** $\cor_l-\cor_u$, **(D)** $\Ps_l-\Ps_u$ and **(E)** $\Dsq_u-\Dsq_l$ statistics, computed on leaf-curling plum aphid's first flight date data, against the geographical distance for all selected sampling location pairs.
+Likewise in Figure \ref{fig_aphid_count}, with increasing the distance *D*, the Spearman and Kendall correlation (hence the synchrony among time series data) between the location pairs decreased as expected.
+But here in contrast, Figure \ref{fig_aphid_ff}  **(C - E)** clearly showed the upper tail dependence with most of the data points lied in the negative quadrant as described before in Sec.
+\ref{NPA}.
+
+Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman 
+correlations, were 
+computed across all pairs of sampling locations, and confidence intervals of these means were 
+computed via a spatial
+resampling scheme, as described in Methods (Table \ref{tab_ff_resamp}).
+Confidence intervals of
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded 
+zero and mean values were positive,
+indicating tail dependence.
+Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
+exluded $0$ and were negative, indicating asymmetry of tail dependence, a feature 
+that data with normal copula 
+structure do not have, providing further evidence answering the third part of question 1.
+The contrast here with the analogous results for abundances of the green spruce aphid is 
+again illuminating.
+
+<!--display the main statistical results in a table-->
+```{r tab_ff_resamp,echo=F,results="markup"}
+s<-stat_aphid_ff$numericdf[,c(3,2,4)]
+colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
+rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
+knitr::kable(s,
+             format = "pandoc",
+             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for first flight data for the leaf-curling plum aphid.\\label{tab_ff_resamp}",
+             booktabs=TRUE)
+```
 

### 2018-04-01 NonParamStat.R

@@ -283,7 +283,7 @@ multcall<-function(d_allsp,sp,lats,longs,pfname,good_loc){
   #---------------------- PLOT :  spear_vs_D -------------------------------
   pdf(paste(pfname,"_Spearman_vs_D.pdf",sep=""),width=8, height=8)
   op<-par(mgp=c(3.5,1,0),mar=c(7,7,1,1))
-  plot(D,spear,xlab="D",ylab="spear",col=rgb(0.5,0,0,.2),pch=19,cex.lab=3,cex.axis=2)
+  plot(D,spear,xlab="D",ylab="Spearman",col=rgb(0.5,0,0,.2),pch=19,cex.lab=3,cex.axis=2)
   lines(range(D),c(0,0),type='l',lty='dashed')
   result<-resampmn(spear)
   #***DAN: fill in one line of numericdf here, the line for spear
@@ -300,7 +300,7 @@ multcall<-function(d_allsp,sp,lats,longs,pfname,good_loc){
   #--------------------- PLOT :  kend_vs_D ---------------------------------
   pdf(paste(pfname,"_Kendall_vs_D.pdf",sep=""),width=8, height=8)
   op<-par(mgp=c(3.5,1,0),mar=c(7,7,1,1))
-  plot(D,kend,xlab="D",ylab="kend",col=rgb(0,0,1,.2),pch=19,cex.lab=3,cex.axis=2)
+  plot(D,kend,xlab="D",ylab="Kendall",col=rgb(0,0,1,.2),pch=19,cex.lab=3,cex.axis=2)
   lines(range(D),c(0,0),type='l',lty='dashed')
   result<-resampmn(kend)
   i_name<-which(numericdf$Stat=="kend")

### 2018-04-01 SupportingInformation.Rmd
typo fixed and commented out checkpoint chunk

@@
 ```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
-library(checkpoint)
+#library(checkpoint)
+#---------------------------------------------------------------
 #checkloc<-tempdir()
@@
 #}
-if (!dir.exists("./.checkpoint/"))
-{
-  dir.create("./.checkpoint/")
-}
-checkpoint("2018-03-29",checkpointLocation = "./")
+#---------------------------------------------------------------
+#if (!dir.exists("./.checkpoint/")){
+#  dir.create("./.checkpoint/")
+#}
+#checkpoint("2018-03-29",checkpointLocation = "./")
 ```
@@
 
-### Non-paramtric analysis of tail dependence  \label{Ex_pns_NPA}
+### Non-parametric analysis of tail dependence  \label{Ex_pns_NPA}
 ```{r stat_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("NonParamStat.R"))}

### 2018-04-02 Tests.Rmd
adding tests for mydispersal.R

@@
 <!--Shya to add headers for other things that need to be tested, and let's discuss before you add the tests themselves-->
+
+# Tests for retd() function from ExtremeTailDep.R
+<!--DCR to add this section-->
+
+# Tests for mydispersal.R
+
+**Test for `Dispersal_mat` function**
+
+```{r test_for_Dispersal_mat}
+source("mydispersal.R")
+set.seed(101)
+D1<-Dispersal_mat(numlocs=3,d=0.4,disp_everywhere=F)
+D1  #matrix for dispersion between nearest neighbours in a linear chain model
+D2<-Dispersal_mat(numlocs=3,d=0.4,disp_everywhere=T) 
+D2  #matrix for dispersion between any nodes in a linear chain model
+if((isSymmetric.matrix(D1)==F) & (isSymmetric.matrix(D2)==T)){print("Dispersal_mat : passed")} 
+```
+
+**Test for `popsim_ml_D`  and `extrisk` function**
+<!-- I assume here that retd() function is already checked and passed the test
+#-------Args--------------------------------------------------
+#p0       A length numlocs vector holding initial populations
+#ns       An numsims by numlocs by numsteps array of the epsilons, 
+#           where numsteps is the number of time steps you want
+#
+#--------Output------------------------------------------------------
+#A numsims by numlocs by numsteps+1 array of populations
+#----------------------------------------------------------------------
+popsim_ml_D<-function(p0,ns,D,r,K)
+-->
+```{r test_for_popsim_ml_D, echo=T}
+source("mydispersal.R")
+set.seed(101)
+numsteps<-15
+numsims<-4
+numlocs<-3
+
+ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1) # output from ExtremeTailDep.R : 
+#                                                      a righttail dep matrix(numpoints by numlocs,     
+#                                                      numpoints=numsteps*numsims)
+dim(ns1)
+ns1<-array(ns1,c(numsteps,numsims,numlocs))# convert to an array (numsteps by numsims by numlocs)
+if(all(dim(ns1)==c(numsteps,numsims,numlocs))==T){print("passed")}
+ns1<-aperm(ns1,c(2,3,1)) # convert to an array (numsims by numlocs by numsteps)
+if(all(dim(ns1)==c(numsims,numlocs,numsteps))==T){print("passed")}
+
+D<-D1
+r<-0.5
+K<-10
+res<-popsim_ml_D(p0=rep(K,numlocs),ns=ns1,D=D1,r=r,K=K)
+res[,,1]
+res[,,2]
+res[,,numsteps+1]
+if(all(dim(res)==c(numsims,numlocs,numsteps+1))==T){print("popsim_ml_D : passed")}
+
+# test for extrisk function
+totpop<-apply(FUN=sum,X=res,MARGIN=c(1,3))
+totpop # numsims by numsteps+1 matrix
+if(all(totpop[,1]==(numlocs*K))==T){print("passed")}
+ext<-(apply(FUN=sum,X=(totpop==0),MARGIN=2)/dim(res)[1])
+ext
+if(length(ext)==numsteps+1){print("passed")}
+
+ans<-extrisk(sims=res)
+if(all(ext==ans)==T){print("extrisk : passed")}
+```

### 2018-04-02 SupportingInformation.Rmd
minor changes
edited some more in the aphid sections how that shya has fixed some plots and added accompanyng text. Can next start from plankton.

@@
    * *Ceratium furca* is a dicoflagellate commonly found in UK seas
-   * note, when we edit this, we cannot call it the North Seas since data come from around the UK not only the North Sea, unless Shya threw out all grid cells that were outside the North Sea
+   * note, when we edit this, we cannot call it the North Seas since data come also from other seas around the UK
+
+\newpage
 
@@
 
+\newpage
+
 ## Results: soil C and N data \label{Ex_SoilCN}
@@
 
+\newpage
+
 ## Results: abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
@@
 
-<!--Shya, can you please organize these following plots into one plot with mutliple panels? Once that is done, please add some text describing these results.-->
+Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
+computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
+resampling scheme, as described in Methods (Table \ref{tab_count_resamp}). Confidence intervals of
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
+indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
+exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
+structure do not have, providing further evidence answering the third part of question 1.
+
+<!--Show the statistical results in a table-->
+```{r tab_count_resamp,echo=F,results="markup"}
+s<-stat_aphid_count$numericdf[,c(3,2,4)]
+colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
+rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
+knitr::kable(s,
+             format = "pandoc",
+             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for abundances of the green spruce aphid.\\label{tab_count_resamp}",
+             booktabs=TRUE)
+```
+
+Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance
+between the locations (Figure \ref{fig_aphid_count} A,B) showed the typical decline of correlation with 
+distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting 
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of  
+locations shows these quantities also decline with distance (Figure \ref{fig_aphid_count} C-H).
+<!--DAN add reference and revisit the text here once Shya add these panels-->
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Figure \ref{fig_aphid_count} I-K)
+did not show a decline with distance, but did illustrate visually the statistical
+result of Table \ref{tab_count_resamp} that averages of these quantities across all pairs of sampling locations
+were significantly greater than zero, demonstrating asymmetry of tail dependence.
+Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in 
+the [`BIVAN`](https://github.com/sghosh89/BIVAN) package.
+
+<!--Shya, can you please add panels for corl, coru, Pl, Pu, D2l, D2u vs. distance, following the Spearman and Kendall
+panels? I think it will be the first time anyone has shown these, so we should show the decline with distance
+holds for these as well. That means you'll have 6 new panels, for a total of 11 panels. So it'll be a big figure, 
+possibly taking up a whole page, or most of one. Similar changes, please,
+to the analogous figs for the aphid ff and ceratium furca data.-->
 \begin{figure}[!h]
@@
 \includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
-\caption{Non-parametric tail statistics results for green spruce aphid abundance data plotted against geographical distance $(D)$ between two sampling locations; $\textbf{(A)}$ Spearman correlation vs.
-$D$, $\textbf{(B)}$ Kendall correlation vs.
-$D$, $\textbf{(C)}$ $\cor_l-\cor_u$ vs.
-$D$, $\textbf{(D)}$ $\Ps_l-\Ps_u$ vs.
-$D$, $\textbf{(E)}$ $\Dsq_u-\Dsq_l$ vs.
-$D$ plotted for all (10 $\times$ 10 locations) taken pairwise.
-\label{fig_aphid_count}}
+\caption{Non-parametric tail statistics results for green spruce aphid abundance data plotted against geographical distance ($D$) between two sampling locations.
+(A) Spearman correlation vs.
+$D$; (B) Kendall correlation vs.
+$D$; (C) $\cor_l$ vs.
+$D$; (D) $\cor_u$ vs.
+$D$; (E) $\Ps_l$ vs.
+$D$; (F) $\Ps_u$ vs.
+$D$; (G) $\Dsq_u$ vs.
+$D$;
+(H) $\Dsq_l$ vs. $D$; 
+(I) $\cor_l-\cor_u$ vs. $D$, (J) $\Ps_l-\Ps_u$ vs. $D$, (K) $\Dsq_u-\Dsq_l$ vs. $D$ plotted for all 
+pairs of locations. \label{fig_aphid_count}}
 \end{center}
@@
 
-We computed Spearman and Kendall correlation along with our three non-parametric statistics ($\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$) as described in Sec.
-\ref{NPA} for  green spruce aphid abundance data taken at all $10$ $\times$ $10$ location pairs (code given in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R)).
-We also constructed a $10$ $\times$ $10$ matrix using `gcdist` from `ncf` package where each element in that matrix represented the geographical distance between any two selected locations.
-Figure \ref{fig_aphid_count} presented plots for **(A)** Spearman, **(B)** Kendall, **(C)** $\cor_l-\cor_u$, **(D)** $\Ps_l-\Ps_u$ and **(E)** $\Dsq_u-\Dsq_l$ statistics against the geographical distance for all selected sampling location pairs.
-With increasing the distance *D*, the Spearman and Kendall correlation (hence the synchrony among time series data) between the location pairs decreased as expected.
-But one cannot conclude the tail dependence of that synchrony from Figure \ref{fig_aphid_count} **(A)** or **(B)**, whereas **(C - E)** clearly showed the lower tail dependence with most of the data points lied in the positive quadrant as described before in Sec.
-\ref{NPA}.
-
-Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
-computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
-resampling scheme, as described in Methods (Table \ref{tab_count_resamp}). Confidence intervals of
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
-indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
-exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
-structure do not have, providing further evidence answering the third part of question 1.
-
-<!--Show the statistical results in a table-->
-```{r tab_count_resamp,echo=F,results="markup"}
-s<-stat_aphid_count$numericdf[,c(3,2,4)]
-colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
-rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
-knitr::kable(s,
-             format = "pandoc",
-             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for abundances of the green spruce aphid.\\label{tab_count_resamp}",
-             booktabs=TRUE)
-```
+\newpage
 
@@
 
-<!--Shya, can you please organize these following plots into one plot with mutliple panels? Once that is done, please add some text describing these results.-->
-\begin{figure}[!h]
-\begin{center}
-\textbf{\;\;\;\;\;\;\;\ (A) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (B)} \\
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf}
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf}\\
-\textbf{\;\;\;\;\;\;\;\ (C) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (D)} \\
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf}
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf}\\
-\textbf{\;\;\;\;\;\;\;\ (E)} \\
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
-\caption{Non-parametric tail statistics results for leaf-curling plum aphid's first flight date data plotted against geographical distance $(D)$ between two sampling locations; $\textbf{(A)}$ Spearman correlation vs.
-$D$, $\textbf{(B)}$ Kendall correlation vs.
-$D$, $\textbf{(C)}$ $\cor_l-\cor_u$ vs.
-$D$, $\textbf{(D)}$ $\Ps_l-\Ps_u$ vs.
-$D$, $\textbf{(E)}$ $\Dsq_u-\Dsq_l$ vs.
-$D$ plotted for all (10 $\times$ 10 locations) taken pairwise.
-\label{fig_aphid_ff}}
-\end{center}
-\end{figure}
-
-Figure \ref{fig_aphid_ff} represented plots for **(A)** Spearman correlation, **(B)** Kendall correlation, **(C)** $\cor_l-\cor_u$, **(D)** $\Ps_l-\Ps_u$ and **(E)** $\Dsq_u-\Dsq_l$ statistics, computed on leaf-curling plum aphid's first flight date data, against the geographical distance for all selected sampling location pairs.
-Likewise in Figure \ref{fig_aphid_count}, with increasing the distance *D*, the Spearman and Kendall correlation (hence the synchrony among time series data) between the location pairs decreased as expected.
-But here in contrast, Figure \ref{fig_aphid_ff}  **(C - E)** clearly showed the upper tail dependence with most of the data points lied in the negative quadrant as described before in Sec.
-\ref{NPA}.
-
 Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
@@
 
+Plotting Spearman or Kendall correlations between first flight time series for pairs of locations as a function of great circle distance between the locations (Figure \ref{fig_aphid_ff} A, B) showed the typical (though slight, in this case) decline of correlation with increasing distance.
+Plotting $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$, and $\Dsq_l$ against distances between pairs of locations shows these quantities also decline with distance (Figure \ref{fig_aphid_ff} C-H).
+<!--DAN: make sure this is true when you see the plots again-->
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_aphid_ff} I-K) did not show a decline with distance, but did demonstrate visually the statistical result of Table \ref{tab_ff_resamp} that averages of these quantities across all pairs of sampling locations were significantly less than zero, showing asymmetry of tail dependence.
+
+<!--Shya, pls add 6 new panels as described in my comment near the analogous figure for aphid count-->
+\begin{figure}[!h]
+\begin{center}
+\textbf{\;\;\;\;\;\;\;\ (A) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (B)} \\
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf}
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf}\\
+\textbf{\;\;\;\;\;\;\;\ (C) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (D)} \\
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf}
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf}\\
+\textbf{\;\;\;\;\;\;\;\ (E)} \\
+\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
+\caption{Non-parametric tail statistics results for leaf-curling plum aphid's first flight date data plotted against geographical distance $(D)$ between two sampling locations.
+(A) Spearman correlation vs.
+$D$, (B) Kendall correlation vs.
+$D$, (C) $\cor_l$ vs.
+$D$, (D) $\cor_u$ vs.
+$D$; (E) $\Ps_l$ vs.
+$D$; (F) $\Ps_u$ vs.
+$D$; (G) $\Dsq_u$ vs.
+$D$; (H) $\Dsq_l$ vs.
+$D$; (I) $\cor_l-\cor_u$ vs.
+$D$; (J) $\Ps_l-\Ps_u$ vs.
+$D$; (K) $\Dsq_u-\Dsq_l$ vs.
+$D$ plotted for all pairs of locations.
+\label{fig_aphid_ff}}
+\end{center}
+\end{figure}
+
 \newpage
+
+<!--DAN: effectively stopped here 2018 04 01 10pm-->
+
 ## Results: abundance data for *Ceratium furca* in UK seas \label{Ex_pns}
@@
 
+\newpage
+
 # Question 2 analyses for each dataset
 Recall that question 2 is: What are some possible causes of non-Gaussian copula structure?
+
+\newpage
 
@@
 \newpage
+
 # References:

### 2018-04-02 SupportingInformation.Rmd
multipanel (11 fig) plot for aphid and plankton added

@@
 \begin{figure}[!h]
+\textbf{ \hspace{2.6 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.8 cm} (D)} \\
 \begin{center}
-\textbf{\;\;\;\;\;\;\;\ (A) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (B)} \\
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf}
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf}\\
-\textbf{\;\;\;\;\;\;\;\ (C) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (D)} \\
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf}
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf}\\
-\textbf{\;\;\;\;\;\;\;\ (E)} \\
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Corl_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Coru_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{2.6 cm} (E) \hspace{3.6 cm} (F) \hspace{3.8 cm} (G) \hspace{3.8 cm} (H)} \\
+\begin{center}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Pl_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Pu_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_D2u_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_D2l_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{4.8 cm} (I) \hspace{3.8 cm} (J) \hspace{4 cm} (K) } \\
+\begin{center}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
 \caption{Non-parametric tail statistics results for green spruce aphid abundance data plotted against geographical distance ($D$) between two sampling locations.
@@
 \begin{figure}[!h]
+\textbf{ \hspace{2.6 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.8 cm} (D)} \\
 \begin{center}
-\textbf{\;\;\;\;\;\;\;\ (A) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (B)} \\
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf}
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf}\\
-\textbf{\;\;\;\;\;\;\;\ (C) \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\ (D)} \\
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf}
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf}\\
-\textbf{\;\;\;\;\;\;\;\ (E)} \\
-\includegraphics[width=2in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Coru_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{2.6 cm} (E) \hspace{3.6 cm} (F) \hspace{3.8 cm} (G) \hspace{3.8 cm} (H)} \\
+\begin{center}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pu_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2l_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{4.8 cm} (I) \hspace{3.8 cm} (J) \hspace{4 cm} (K) } \\
+\begin{center}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
 \caption{Non-parametric tail statistics results for leaf-curling plum aphid's first flight date data plotted against geographical distance $(D)$ between two sampling locations.
@@
 
-![copula plot for Plankton_north_sea\label{fig_pns_allcop}](./Results/stat_results/stat_plankton_north_sea/Sp_16_AllCops.pdf)
-
-![Plot for Plankton_north_sea : Spearman vs. Distance\label{fig_pns_spear}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Spearman_vs_D.pdf){ width=50%, height=50% }
-
-![Plot for Plankton_north_sea : Kendall vs. Distance\label{fig_pns_kend}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Kendall_vs_D.pdf){ width=50%, height=50% }
-
-![$cor_l$-$cor_u$ vs.
-Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_Corl-Coru}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf){ width=50%, height=50% }
-
-![$P_l$-$P_u$ vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_Pl-Pu}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf){ width=50%, height=50% }
-
-![$D^2_u$-$D^2_l$ vs.
-Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_D2u-D2l}](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf){ width=50%, height=50% }
+\begin{figure}[!h]
+\textbf{ \hspace{2.6 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.8 cm} (D)} \\
+\begin{center}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Spearman_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Kendall_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Coru_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{2.6 cm} (E) \hspace{3.6 cm} (F) \hspace{3.8 cm} (G) \hspace{3.8 cm} (H)} \\
+\begin{center}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pu_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2l_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{4.8 cm} (I) \hspace{3.8 cm} (J) \hspace{4 cm} (K) } \\
+\begin{center}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf}
+\caption{Non-parametric tail statistics results for Ceratium furca abundance data plotted against geographical distance $(D)$ between two sampling locations.
+(A) Spearman correlation vs.
+$D$, (B) Kendall correlation vs.
+$D$, (C) $\cor_l$ vs.
+$D$, (D) $\cor_u$ vs.
+$D$; (E) $\Ps_l$ vs.
+$D$; (F) $\Ps_u$ vs.
+$D$; (G) $\Dsq_u$ vs.
+$D$; (H) $\Dsq_l$ vs.
+$D$; (I) $\cor_l-\cor_u$ vs.
+$D$; (J) $\Ps_l-\Ps_u$ vs.
+$D$; (K) $\Dsq_u-\Dsq_l$ vs.
+$D$ plotted for all pairs of locations.
+\label{fig_pns}}
+\end{center}
+\end{figure}
 

### 2018-04-02 SupportingInformation.Rmd
adding checkpoint chunk

@@
 ```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
-#library(checkpoint)
-#---------------------------------------------------------------
-#checkloc<-tempdir()
-#if (!dir.exists(paste0(checkloc,"/.checkpoint/")))
-#{
-#  dir.create(paste0(checkloc,"/.checkpoint/"))
-#  checkpoint("2018-03-29",checkpointLocation = checkloc)
-#}
-#---------------------------------------------------------------
-#if (!dir.exists("./.checkpoint/")){
-#  dir.create("./.checkpoint/")
-#}
-#checkpoint("2018-03-29",checkpointLocation = "./")
+library(checkpoint)
+if (!dir.exists("./.checkpoint/")){
+  dir.create("./.checkpoint/")
+}
+checkpoint("2018-03-29",checkpointLocation = "./")
 ```

### 2018-04-02 SupportingInformation.Rmd
minor editing changes after shya added the plots of the nonparam stats versus distance for the aphid datasets

@@
 library(checkpoint)
+library(captioner)
 if (!dir.exists("./.checkpoint/")){
@@
 distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting 
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of  
-locations shows these quantities also decline with distance (Figure \ref{fig_aphid_count} C-H).
-<!--DAN add reference and revisit the text here once Shya add these panels-->
+$\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of 
+locations shows these quantities also decline with distance (Figure \ref{fig_aphid_count} C-F). Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows
+these quantities increase with distance (as expected, given the opposite interpretation 
+these
+two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively).
 Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Figure \ref{fig_aphid_count} I-K)
@@
 
-<!--Shya, can you please add panels for corl, coru, Pl, Pu, D2l, D2u vs. distance, following the Spearman and Kendall
-panels? I think it will be the first time anyone has shown these, so we should show the decline with distance
-holds for these as well. That means you'll have 6 new panels, for a total of 11 panels. So it'll be a big figure, 
-possibly taking up a whole page, or most of one. Similar changes, please,
-to the analogous figs for the aphid ff and ceratium furca data.-->
 \begin{figure}[!h]
@@
 Plotting Spearman or Kendall correlations between first flight time series for pairs of locations as a function of great circle distance between the locations (Figure \ref{fig_aphid_ff} A, B) showed the typical (though slight, in this case) decline of correlation with increasing distance.
-Plotting $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$, and $\Dsq_l$ against distances between pairs of locations shows these quantities also decline with distance (Figure \ref{fig_aphid_ff} C-H).
-<!--DAN: make sure this is true when you see the plots again-->
+Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear in most cases to decline slightly with distance (Figure \ref{fig_aphid_ff} C-F).
+$\Dsq_u$ and $\Dsq_l$ appear to slightly increase with distance, as expected given their
+opposite meaning (Figure \ref{fig_aphid_ff} G-H). 
 Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_aphid_ff} I-K) did not show a decline with distance, but did demonstrate visually the statistical result of Table \ref{tab_ff_resamp} that averages of these quantities across all pairs of sampling locations were significantly less than zero, showing asymmetry of tail dependence.
 
-<!--Shya, pls add 6 new panels as described in my comment near the analogous figure for aphid count-->
 \begin{figure}[!h]
@@
 
-<!--DAN: effectively stopped here 2018 04 01 10pm-->
-
 ## Results: abundance data for *Ceratium furca* in UK seas \label{Ex_pns}

### 2018-04-02 SupportingInformation.Rmd
various changes made

--- 2018-04-02 SupportingInformation.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
-
 seed<-101
-families<-c(1,3:10,13,14,16:20)
-
+# families<-c(1,3:10,13,14,16:20)
 source("mtime.R") #A function needed for caching
-source("getcopula.R")
-source("OurBiCopSelect.R")
-source("FittingCopula_selective_loc.R")
-source("normal_cop_AICw.R")
-source("aphid_data_calling.R")
-source("plankton_north_sea_data_calling.R")
-source("NonParamStat.R")
-source("copsurrog2d.R")
-source("TestStats_multivar_dataset.R")
-source("skewness_multivar_dataset.R")
-source("ExtremeTailDep.R")
-source("mydispersal.R")
 ```
@@
 \ref{Model_selection}).\\label{fig_soilCNcop}",  fig.height=5, fig.width=5}
-
+source("getcopula.R")
 #convert to a copula and plot
@@
 <!-- The lines you are talking about appeared as message not as warning(actualy no warning appered for this chunk), so we have to set message=F here-->
-```{r stat_testing_save_result, message=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopupulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
+```{r stat_testing_save_and_plot_result, message=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
+source("TestStats_multivar_dataset.R")
 set.seed(seed)
@@
 save(res_pt35_rankF,res_pt35_rankT,res_pt1000_rankF,res_pt1000_rankT,file=paste(resultsloc,"StatTestNumericResults.RData",sep=''))
-```
-
-```{r stat_testing_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,callfn,mtime("StatTestNumericResults.RData"),mtime("TestStats_multivar_dataset.R"))}
 
@@
 <!--This does the computations for this section-->
-```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"))}
+```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("OurBiCopSelect.R")
 set.seed(seed)
@@
 ```{r stat_soilCN_graphic_approach, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+source("CopulaFunctions_flexible.R")
+source("copsurrog2d.R")
 set.seed(seed)
@@
 ```{r read_aphid_data, echo=F}
+source("aphid_data_calling.R")
+
 # Aphid raw data for count
@@
 <!--Do the calculations-->
-```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("FittingCopula_selective_loc.R"))}
+```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+
+source("FittingCopula_selective_loc.R")  # top most hierarchy source file
 set.seed(seed)
@@
 ```{r tab_count_AICw_normal, echo=F, results='markup'}
+source("normal_cop_AICw.R")
 knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
@@
 <!--Do the main computations for this section-->
-```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("NonParamStat.R"))}
+```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+
+source("good_loclist.R")
+source("NonParamStat.R")
 set.seed(seed) 
@@
 <!--Do the analysis-->
-```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("FittingCopula_selective_loc.R"))} 
+```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+
+source("FittingCopula_selective_loc.R")  # top most hierarchy source file
 set.seed(seed)
@@
 ```{r tab_ff_AICw_normal, echo=F, results='markup'}
+source("normal_cop_AICw.R")
 knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
@@
 <!--Do the analysis-->
-```{r stat_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("NonParamStat.R"))}
+```{r stat_aphid_ff, echo=F, cache=T,  cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+
+source("good_loclist.R")
+source("NonParamStat.R")
 set.seed(seed)
@@
 ```{r read_plankton_north_sea_data, echo=F}
+source("plankton_north_sea_data_calling.R")
 d0_plankton_north_sea<-as.matrix(read.csv("./Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))
@@
 <!--Do the main analysis-->
-```{r RES_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"), mtime("FittingCopula_selective_loc.R"))}
+```{r RES_plankton_north_sea, echo=F, cache=T, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+
+source("FittingCopula_selective_loc.R")  # top most hierarchy source file
 
@@
 ```{r tab_pns_AICw_normal,echo=F,results="markup"}
+source("normal_cop_AICw.R")
 knitr::kable(normal_cop_AICw_matrix(r=RES_plankton_north_sea),
@@
 ### Non-parametric analysis of tail dependence  \label{Ex_pns_NPA}
-```{r stat_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("NonParamStat.R"))}
+```{r stat_plankton_north_sea, echo=F, cache=T, cache=T,  cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+
+source("good_loclist.R")
+source("NonParamStat.R")
 set.seed(seed)
@@
 
-```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
-set.seed(seed)
+```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+
+source("good_loclist.R")
+source("skewness_multivar_dataset.R")
+set.seed(seed)
+
 pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf")
@@
 
-```{r skewness_ff_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
-set.seed(seed)
+```{r skewness_ff_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+
+source("good_loclist.R")
+source("skewness_multivar_dataset.R")
+set.seed(seed)
+
 pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf")
@@
 
-```{r skewness_pns_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,mtime("ncsurrog.R"),mtime("skewness_multivar_dataset.R"))}
-set.seed(seed)
+```{r skewness_pns_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+
+source("good_loclist.R")
+source("skewness_multivar_dataset.R")
+set.seed(seed)
+
 pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf")
@@
 d(degree of dispersal).\\label{fig_risk_r0_numlocs5}", fig.height=3.5, fig.width=8}
+source("mydispersal.R")
 set.seed(seed)
@@
 d(degree of dispersal).\\label{fig_risk_r0.05}", fig.height=3.5, fig.width=8}
+source("mydispersal.R")
 set.seed(seed)
@@
 d(degree of dispersal).\\label{fig_risk_r0.05_disp_everywhere}", fig.height=3.5, fig.width=8}
+source("mydispersal.R")
 set.seed(seed)
@@
 ```{r tl_analysis, echo=F, cache=T, cache.extra=list(mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
+source("copsurrognd.R")
 source("TaylorsLawTools.R")

### 2018-04-02 CopulaFunctions_flexible.R

@@ -82,12 +82,10 @@ Pbds<-function(vi,vj,lb,ub){
   
   # when two boundary lines are on the right side of vi+vj=1 line  
   if((2*lb>=1) & (2*ub >=1)){
-    #d_max<-lb*sqrt(2)
-    d_max=abs((lb-1)*sqrt(2))
+    d_max<-abs((lb-1)*sqrt(2))
     a<-abs(2*sqrt(2)*(ub-1))
     b<-2*d_max
     h<-abs(2*(ub-lb))/sqrt(2)
-    #deno<-2*(((lb-1)^2)-((ub-1)^2))
     deno<-0.5*(a+b)*h
     
     dist_Si1<-c()
@@ -120,7 +118,6 @@ Pbds<-function(vi,vj,lb,ub){
     a<-2*sqrt(2)*lb
     b<-2*d_max
     h<-abs(2*(ub-lb))/sqrt(2)
-    #deno<-2*((ub^2)-(lb^2))
     deno<-0.5*(a+b)*h
  
     dist_Si1<-c()

### 2018-04-03 SupportingInformation.Rmd
changes to deal with a warning when we try to fit a BB7 to negatively correlated data

@@
 <!-- The lines you are talking about appeared as message not as warning(actualy no warning appered for this chunk), so we have to set message=F here-->
-```{r stat_testing_save_and_plot_result, message=F, echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
+```{r stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
 source("TestStats_multivar_dataset.R")

### 2018-04-03 Tests.Rmd
changes to bivariate model selection to exclude negatively correlated pairs

@@
 families<-c(1,3,4,5,13)
-BivMS_res_test<-OurBiCopSelect(d[,1],d[,2],families=families,gofnormal=TRUE,status=TRUE)
+BivMS_res_test<-OurBiCopSelect(d[,1],d[,2],families=families,gofnormal=F,status=TRUE)
 ```
@@
 BivMS_res_test$IndepTestRes
+```
+
+Tau value should be positive:
+```{r echo=T}
+BivMS_res_test$TauVal
 ```
@@
 
-<!--Shya to add headers for other things that need to be tested, and let's discuss before you add the tests themselves-->
+Now test with negatively correlated data, which happens ocassionally:
+
+Data are simulated from a normal copula with parameter $-0.2$:
+```{r get_neg_dat, echo=T}
+set.seed(101)
+cop<-normalCopula(-0.2,2,dispstr="un")
+d<-rCopula(1000,cop)
+v<-getcopula(d=d,rankon=F,ploton=T) #with real data, rankon should be T
+```
+
+Make sure the correlatation is negative for these data:
+```{r test_neg_dat, echo=T}
+cor(v[,1],v[,2],method="kendall")
+```
+
+Call `OurBiCopSelect` - all is should do is test independence, reject independence,
+get tau and find it negative and then stop:
+```{r test_OurBiCopSelect_neg, echo=T, cache=T, cache.extra=list(v,mtime("OurBiCopSelect.R"))}
+res<-OurBiCopSelect(v[,1],v[,2],families=c(1,3,9),level=0.05,AICBIC="AIC",
+                         numBSsmall=100,pthresh=0.2,numBSlarge=1000,
+                         gofnormal=F,status=TRUE)
+res
+```
 

### 2018-04-03 SupportingInformation.Rmd
adding chunk for phylo_data

@@
 \label{fig_soilCNcop}](./Results/SoilCNcop.pdf){ width=30%, height=30% }
+
+## Phylogenetic data  \label{Phylo_data}
+```{r calculation_phylo_data, echo=F, results="hide"}
+source("getcopula.R")
+phylo_dat<-read.csv("./Data/Anolis/contrasts.csv",header = F)
+# Input :
+#        data : a dataframe : phylo_dat with dim.=[99 x 14]
+#        i, j : column indices of data frame to get bivariate data set, should be (i,j)<=13
+#        ploton : logical for option plots of raw data and it's copula representation
+# Output : 
+#        a cop rank matrix for the raw data
+phylo_dat_fn<-function(data,i,j,ploton=T){
+
+  data<-data[,2:dim(phylo_dat)[2]]
+  d<-data[,c(i,j)]
+
+  v_phylo<-getcopula(d=d,rankon = T,ploton = F)
+
+  if(ploton==T){
+    op<-par(mfrow=c(1,2))
+    plot(d[,1],d[,2],col="blue")
+    getcopula(d=d,rankon = T,ploton = T)
+    par(op)
+  }
+  return(v_phylo)
+}
+
+# This plot I create on temporary basis to see the copula structure between various column of the raw data
+# After I know which variables should I look into, I will modify this chunk and add model selection and NPA on that
+pdf("./Results/temp_phylo_plot.pdf",width=20, height=20)
+op<-par(mfrow=c(13,13))
+for(i in c(1:13)){
+  for(j in c(1:13)){
+    v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = F)
+    plot(v_phylo[,1],v_phylo[,2],col="red")
+    #mtext("(i,j)=(",i,", ",j,")")
+    mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 3,col="blue")
+  } 
+}
+par(op)
+dev.off()
+```
 

### 2018-04-04 SupportingInformation.Rmd
changing extinction risk figures

@@
 ### LC model :
-```{r fig_risk_r0_numlocs5, echo=F, cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="LC model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0_numlocs5}", fig.height=3.5, fig.width=8}
+
+  *  As expected, when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
+With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_r0_numlocs5} - Figure \ref{fig_risk_r0_numlocs25}].
+
+### Ricker model :
+
+ *  For ricker model when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
+For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
+
+  * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_r0.05} - Figure \ref{fig_risk_r0.05_K200}]
+
+ * With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_r0.05} and Figure \ref{fig_risk_r0.05_numlocs25}].
+
+```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+
 source("mydispersal.R")
 set.seed(seed)
-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0,K=50,disp_everywhere=F)
-```
-
-```{r fig_risk_r0_numlocs25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="LC model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0_numlocs25}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-varying_d(numsims=10000,numsteps =25,numlocs=25,r=0,K=50,disp_everywhere=F)
-```
-
-  *  As expected, when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
-With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_r0_numlocs5} - Figure \ref{fig_risk_r0_numlocs25}].
-
-### Ricker model :
-
- *  For ricker model when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
-For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
-
-  * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_r0.05} - Figure \ref{fig_risk_r0.05_K200}]
-
- * With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_r0.05} and Figure \ref{fig_risk_r0.05_numlocs25}].
-
-```{r fig_risk_r0.05, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.05}", fig.height=3.5, fig.width=8}
-source("mydispersal.R")
-set.seed(seed)
-r<-0.05
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r0.05_K200, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.05_K200}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.05
-numlocs<-5
-K<-200
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r0.05_numlocs25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.05_numlocs25}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.05
-numlocs<-25
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r0.25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.25}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.25
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r0.5, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.5}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.5
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r0.6, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.6}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.6
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r0.7, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.7}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.7
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r0.75, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.75}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.75
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r0.9, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.9}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.9
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r1.0, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.0}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-1.0
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r1.25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.25}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-1.25
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r1.5, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.5}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-1.5
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r1.75, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.75}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-1.75
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
-```
-
-```{r fig_risk_r1.95, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.95}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-1.95
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=F)
+
+rlist<-c(0,0.05,0.25,0.5,0.7,0.75,0.9,1.0,1.25,1.5,1.75,1.95)
+
+# plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
+pdf("./Results/ext_risk_local_disp.pdf",heigh=6,width=10)
+op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
+for(r in rlist){
+    cat("r=",r,"\n")
+    s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
+    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s$d_seq,s$risk_right,type="b",col="blue")
+    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+# plot for global dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
+pdf("./Results/ext_risk_global_disp.pdf",heigh=6,width=10)
+op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
+for(r in rlist){
+  cat("r=",r,"\n")
+  s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
+  plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d_seq,s$risk_right,type="b",col="blue")
+  legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+  mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+# plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
+pdf("./Results/ext_risk_numlocs_effect.pdf",heigh=6,width=6)
+op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
+for(r in c(0,0.25)){
+  cat("r=",r,"\n")
+  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
+  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=T,ploton=F)
+
+  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
+  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
+       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
+  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+  #mtext(paste0("B",1))
+}
+par(op)
+dev.off()
+
+# plot to see the effect of K for r=0.25 (Ricker model)
+pdf("./Results/ext_risk_K_effect.pdf",heigh=6,width=6)
+op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
+
+s1<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=F,ploton=F)
+s2<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=F,ploton=F)
+
+s3<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=T,ploton=F)
+s4<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=T,ploton=F)
+
+plot(s1$d_seq,s1$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+lines(s2$d_seq,s2$risk_left,type="b",col="orange")
+legend("topleft",title=("Noise : left tail dep., local disp."), c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+plot(s1$d_seq,s1$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+lines(s2$d_seq,s2$risk_right,type="b",col="skyblue")
+legend("topleft",title=("Noise : right tail dep., local disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+
+plot(s3$d_seq,s3$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+lines(s4$d_seq,s4$risk_left,type="b",col="orange")
+legend("topleft",title= ("Noise : left tail dep., global disp."),c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+plot(s3$d_seq,s3$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+lines(s4$d_seq,s4$risk_right,type="b",col="skyblue")
+legend("topleft",title=("Noise : right tail dep., global disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+
+par(op)
+dev.off()
 ```
@@
  * With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_r0.05_disp_everywhere} and Figure \ref{fig_risk_r0.05_numlocs25_disp_everywhere}].
-
-```{r fig_risk_r0.05_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.05_disp_everywhere}", fig.height=3.5, fig.width=8}
-source("mydispersal.R")
-set.seed(seed)
-r<-0.05
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r0.05_K200_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal); With increasing K from 50 to 200, extinction risk decreases.\\label{fig_risk_r0.05_K200_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.05
-numlocs<-5
-K<-200
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r0.05_numlocs25_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal); With increasing numlocs from 5 to 25, extinction risk decreases.\\label{fig_risk_r0.05_numlocs25_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.05
-numlocs<-25
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r0.25_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.25_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.25
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r0.5_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.5_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.5
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r0.6_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.6_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.6
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r0.7_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.7_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.7
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r0.75_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.75_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.75
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r0.9_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r0.9_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-0.9
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r1.0_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.0_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-1.0
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r1.25_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.25_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-1.25
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r1.5_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.5_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-1.5
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r1.75_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.75_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-1.75
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
-
-```{r fig_risk_r1.95_disp_everywhere, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model_disp_everywhere : Plot Extinction risk vs.
-d(degree of dispersal).\\label{fig_risk_r1.95_disp_everywhere}", fig.height=3.5, fig.width=8}
-set.seed(seed)
-r<-1.95
-numlocs<-5
-K<-50
-varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K,disp_everywhere=T)
-```
 

### 2018-04-05 SupportingInformation.Rmd
minor modification in ext risk plot

@@
 
-### LC model :
-
-  *  As expected, when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
-With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_r0_numlocs5} - Figure \ref{fig_risk_r0_numlocs25}].
-
-### Ricker model :
-
- *  For ricker model when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
-For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
-
-  * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_r0.05} - Figure \ref{fig_risk_r0.05_K200}]
-
- * With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_r0.05} and Figure \ref{fig_risk_r0.05_numlocs25}].
-
 ```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
@@
 for(r in rlist){
-    cat("r=",r,"\n")
+    #cat("r=",r,"\n")
     s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
@@
 for(r in rlist){
-  cat("r=",r,"\n")
+  #cat("r=",r,"\n")
   s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
@@
 # plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
-pdf("./Results/ext_risk_numlocs_effect.pdf",heigh=6,width=6)
+pdf("./Results/ext_risk_numlocs_effect_local_disp.pdf",heigh=6,width=6)
 op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
 for(r in c(0,0.25)){
-  cat("r=",r,"\n")
-  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
-  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=T,ploton=F)
+  #cat("r=",r,"\n")
+  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
+  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=F,ploton=F)
 
@@
 
+# plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
+pdf("./Results/ext_risk_numlocs_effect_global_disp.pdf",heigh=6,width=6)
+op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
+for(r in c(0,0.25)){
+  #cat("r=",r,"\n")
+  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
+  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=T,ploton=F)
+
+  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
+  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
+       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
+  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+  #mtext(paste0("B",1))
+}
+par(op)
+dev.off()
+
 # plot to see the effect of K for r=0.25 (Ricker model)
@@
 
-# Explore equal_dispersal everywhere
-
-### LC model :
-
-    * No effect here.
-
-### Ricker model :
-
-  * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected.
-[see Figure \ref{fig_risk_r0.05_disp_everywhere} - Figure \ref{fig_risk_r0.05_K200_disp_everywhere}]
-
- * With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_r0.05_disp_everywhere} and Figure \ref{fig_risk_r0.05_numlocs25_disp_everywhere}].
+### LC model (r=0) :
+
+  * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
+[see first plot of Figure \ref{fig_risk_local_disp} and Figure \ref{fig_risk_global_disp}]
+
+### Ricker model (r $\neq$ 0):
+
+ *  For ricker model when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
+For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
+[see Figure \ref{fig_risk_local_disp} - \ref{fig_risk_global_disp}]
+
+  * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]
+
+With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_numlocs_effect_local_disp} - Figure \ref{fig_risk_numlocs_effect_global_disp}] for both model.
+
+![Extinction_Risk_local_dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp}](./Results/ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Extinction_Risk_global_dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp}](./Results/ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Extinction_Risk_numlocs_effect_local_disp; numsteps=25, numsims=10000\label{fig_risk_numlocs_effect_local_disp}](./Results/ext_risk_numlocs_effect_local_disp.pdf){ width=40%, height=40% }
+
+![Extinction_Risk_numlocs_effect_global_disp; numsteps=25, numsims=10000\label{fig_risk_numlocs_effect_global_disp}](./Results/ext_risk_numlocs_effect_global_disp.pdf){ width=40%, height=40% }
+
+![Extinction_Risk_K_effect on Ricker model; r=0.25, numsteps=25, numsims=10000\label{fig_risk_K_effect}](./Results/ext_risk_K_effect.pdf){ width=40%, height=40% }
 

### 2018-04-05 SupportingInformation.Rmd
working on BB7 warning message, I think I fixed it

--- 2018-04-03 SupportingInformation.Rmd
@@
 <!--Do the main analysis-->
-```{r RES_plankton_north_sea, echo=F, cache=T, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+```{r RES_plankton_north_sea, echo=F, cache=T, results="hide", cache.extra=list(seed,data_plankton_north_sea,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
 

### 2018-04-05 Tests.Rmd
working on BB7 warning message, I think I fixed it

@@
 families<-c(1,3,4,5,13)
-BivMS_res_test<-OurBiCopSelect(d[,1],d[,2],families=families,gofnormal=F,status=TRUE)
+BivMS_res_test<-OurBiCopSelect(d[,1],d[,2],families=families,gofnormal=T,status=TRUE)
 ```

### 2018-04-05 FittingCopula_selective_loc.R

@@ -116,9 +116,12 @@ RES_single_sp<-function(sp,d_allsp,families,level,data_pt_thrs){
         #plot(u1,u2)
 #        cat("[i, j]=[",i," ,",j,"]","(good_loc[i],good_loc[j])=(",good_loc[i],",",good_loc[j],")","\n")
         
+#        print(paste("i=",i,"; j=",j,sep=""))
+#        print("About to call OurBiCopGofTest")
         ans<-OurBiCopSelect(u1,u2,families,level=0.05,AICBIC="AIC",
                             numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                             gofnormal=FALSE,status=FALSE)
+#        print("Finished OurBiCopGofTest")
         
         copdata_cor_Kend[i,j]<-ans$TauVal
         

### 2018-04-05 OurBiCopSelect.R

@@ -73,10 +73,14 @@ OurBiCopSelect<-function(u1,u2,families,level=0.05,AICBIC="AIC",
   IndepTestRes<-BiCopIndTest(u1,u2)$p.value
   if (status) {cat(paste("Done:",Sys.time(),"\n"))}
   tauval<-cor(u1,u2,method="kendall")
+
+#  print(paste0("IndepTestRes=",IndepTestRes,"; tauval=",tauval))
   
   #if independence rejected and tau>0, then get AICs for copulas and do 
   #goodness of fit stuff
   if (IndepTestRes<level && tauval>0){
+
+#    print("Entering main work if statement in BiCopGofTest")    
     
     #AIC/BIC stuff
     if (status) {cat(paste("Starting A/BIC model selection: ",Sys.time(),"\n"))}
@@ -92,6 +96,7 @@ OurBiCopSelect<-function(u1,u2,families,level=0.05,AICBIC="AIC",
                     AICw=NA,
                     BICw=NA)
     for (counter in 1:(dim(InfCritRes)[1])){
+#      print(paste0("About to call BiCopEst for counter=",counter))
       tres<-BiCopEst(u1,u2,family=InfCritRes[counter,1])
       InfCritRes$par1[counter]<-tres$par
       InfCritRes$par2[counter]<-tres$par2
@@ -133,16 +138,20 @@ OurBiCopSelect<-function(u1,u2,families,level=0.05,AICBIC="AIC",
       stop("Error in OurBiCopSelect: incorrect AICBIC")
     }
     Numboot<-numBSsmall
+#    print("About to call MyBiCopGofTest")
     gres<-MyBiCopGofTest(u1,u2,family=InfCritRes$copcode[ind],
                        method="kendall",B=numBSsmall)
+#    print("Finished calling MyBiCopGofTest")
     GofRes_CvM<-gres$p.value.CvM
     GofRes_KS<-gres$p.value.KS
     Numboot_success<-gres$B_success
     if (GofRes_CvM<pthresh || GofRes_KS<pthresh)
     {
       Numboot<-numBSlarge
+#      print("About to call MyBiCopGofTest second time")
       gres<-MyBiCopGofTest(u1,u2,family=InfCritRes$copcode[ind],
                          method="kendall",B=numBSlarge)
+#      print("Finished calling MyBiCopGofTest second time")
       GofRes_CvM<-gres$p.value.CvM
       GofRes_KS<-gres$p.value.KS
       Numboot_success<-gres$B_success

### 2018-04-08 SupportingInformation.Rmd
trivial
finshed editing though end of UK seas plankton sections

@@
              format='pandoc',
-             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_LTmUT}",
+             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_LTmUT}",
              booktabs=TRUE
@@
 
-As in previous sections (see also Methods) model selection among copulas was done separately for 
+As described in Methods and as in previous sections, model selection among copulas was done separately for 
 each pair of sampling locations. There were $14$ sampling locations, so results are presented in 
@@
 the normal copula family for the same pair (Table \ref{tab_pns_AICw_normal}) 
-
- Table \ref{tab_pns_AICw} represents the Akaike weights of the best-fit copula for each location pair.
-Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_pns_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_pns_p_KS}] of the goodness of fit with atleast
-`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` % successful bootstrap.
-Table \ref{tab_pns_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_pns_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC.
-The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_pns_LTdep_AICw} and Table \ref{tab_pns_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_pns_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1.
-These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)`** out of **`r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`** location pair more lower tail dependence.
+confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{tab_pns_p_CvM}) usually failed to reject the hypothesis that 
+the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a
+Kolmogorov-Smirnov statistic (Table \ref{tab_pns_p_KS}). In all cases at least
+`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit
+tests successfully lead to values of the statistic for comparison with its value on the real data.
 
@@
              format='pandoc',
-             caption = "Plankton_north_sea : Best fit copula family for each selective location pair\\label{tab_pns_bestcop}",
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundance data for *Ceratium furca* in UK seas.
+See previous sections for copula abbreviations.
+NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_pns_bestcop}",
              booktabs=TRUE
@@
 ```{r tab_pns_AICw,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$info_ord_AICw[,,1],
+h<-RES_plankton_north_sea$info_ord_AICw[,,1]
+h[is.infinite(h)]<-NA
+knitr::kable(h,
              format="pandoc",
              digits=3,
-             caption = "Plankton_north_sea : AICw for best fit copula family for each selective location pair\\label{tab_pns_AICw}",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for abundance data for *Ceratium furca* in UK seas.\\label{tab_pns_AICw}",
              booktabs=TRUE)
@@
 source("normal_cop_AICw.R")
-knitr::kable(normal_cop_AICw_matrix(r=RES_plankton_north_sea),
+h<-normal_cop_AICw_matrix(r=RES_plankton_north_sea)
+h[is.infinite(h)]<-NA
+knitr::kable(h,
              format="pandoc",
              digits=3,
-             caption = "Plankton_north_sea : AICw for 'Normal' copula family for each selective location pair\\label{tab_pns_AICw_normal}",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for abundance data for *Ceratium furca* in UK seas.\\label{tab_pns_AICw_normal}",
              booktabs=TRUE)
@@
 ```{r tab_pns_p_CvM,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$gfc_p_CvM,
+h<-RES_plankton_north_sea$gfc_p_CvM
+h[is.infinite(h)]<-NA
+knitr::kable(h,
              format="pandoc",
              digits=3,
-             caption = "Plankton_north_sea : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_pns_p_CvM}",
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for abundance data for *Ceratium furca* in UK seas.
+Tests were based on bootstrapping as described in Methods.\\label{tab_pns_p_CvM}",
              booktabs=TRUE)
@@
 ```{r tab_pns_p_KS,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$gfc_p_KS,
+h<-RES_plankton_north_sea$gfc_p_KS
+h[is.infinite(h)]<-NA
+knitr::kable(h,
              format="pandoc",
              digits=3,
-             caption = "Plankton_north_sea : p-values (Kolmogorov-Smirnov-based test) for lowest AIC based best fit copula\\label{tab_pns_p_KS}",
+             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of
+the goodness of fit of the lowest-AIC copula family for each pair of
+locations for abundance data for *Ceratium furca* in UK seas. Tests
+were based on bootstrapping as described in Methods.\\label{tab_pns_p_KS}",
              booktabs=TRUE)
@@
 
+The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{tab_pns_LTdep_AICw} and\ref{tab_pns_UTdep_AICw}, respectively.
+These values were exclusively positive, thereby
+differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
+have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
+
 ```{r tab_pns_LTdep_AICw,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$LTdep_AICw,
+h<-RES_plankton_north_sea$LTdep_AICw
+h[is.infinite(h)]<-NA
+knitr::kable(h,
              format="pandoc",
              digits=3,
-             caption = "Plankton_north_sea : model averaged lower tail dependence statistic for each selective location pair\\label{tab_pns_LTdep_AICw}",
+             caption = "Model-averaged lower tail dependence statistic for each location pair for abundance data for *Ceratium furca* in UK seas.\\label{tab_pns_LTdep_AICw}",
              booktabs=TRUE)
@@
 ```{r tab_pns_UTdep_AICw,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$UTdep_AICw,
+h<-RES_plankton_north_sea$UTdep_AICw
+h[is.infinite(h)]<-NA
+knitr::kable(h,
              format="pandoc",
              digits=3,
-             caption = "Plankton_north_sea : model averaged upper tail dependence statistic for each selective location pair\\label{tab_pns_UTdep_AICw}",
+             caption = "Model-averaged upper tail dependence statistic for each location pair for abundance data for *Ceratium furca* in UK seas.\\label{tab_pns_UTdep_AICw}",
              booktabs=TRUE)
@@
 
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
+(lower minus upper), shown in Table \ref{tab_pns_LTmUT}, was positive for some
+location pairs and negative for others, but was positive for 
+`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`
+locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
+
 ```{r tab_pns_LTmUT,echo=F,results="markup"}
-knitr::kable(RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw,
+h<-RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw
+h[is.nan(h)]<-NA
+knitr::kable(h,
              format="pandoc",
              digits=3,
-             caption = "Plankton_north_sea : model averaged lower tail - model averaged upper tail dependence statistic for each of the selective location pair\\label{tab_pns_LTmUT}",
+             caption = "Model averaged lower tail dependence minus model
+averaged upper tail dependence statistic for each location pair for *Ceratium furca* in UK seas.\\label{tab_pns_LTmUT}",
              booktabs=TRUE)
@@
 
+Mean values of our statistics, as well as mean Kendall and Spearman
+correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
+spatial resampling scheme, as described in Methods (Table \ref{tab_pns_resamp}).
+Confidence intervals of $\cor_l$","$\cor_u$","$\Ps_l$","$\Ps_u$","$\Dsq_l$","$\Dsq_u$"
+always excluded $0$ and mean values were positive, indicating tail dependence.
+Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
+also all were strictly positive, indicating asymmetry of tail dependence, a feature that data with normal copula
+structure do not have, providing further evidence answering the third part of question 1.
+
+<!--display the main statistical results in a table-->
 ```{r tab_pns_resamp,echo=F,results="markup"}
-s<-stat_plankton_north_sea$numericdf[,2:5]
-rownames(s)<-c("spear","kend","cor~l~","cor~u~","P~l~","P~u~","$D^2$~l~","$D^2$~u~","cor~l~ - cor~u~","P~l~ - P~u~","$D^2$~u~ - $D^2$~l~")
+s<-stat_plankton_north_sea$numericdf[,c(3,2,4)]
+colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
+rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
 knitr::kable(s,
              format = "pandoc",
-             caption = "Plankton_north_sea : resampling result for stat indicating lower tail dependence\\label{tab_pns_resamp}",
+             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for abundance data for *Ceratium furca* in UK seas.\\label{tab_pns_resamp}",
              booktabs=TRUE)
 ```
+
+Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
+between the locations (Figure \ref{fig_pns} A, B) showed the typical decline of correlation with increasing
+distance.
+Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Figure \ref{fig_pns} C-F).
+$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
+their opposite meaning (Figure \ref{fig_pns} G-H).
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_pns} I-K) did not show a decline with distance, but did demonstrate visuaally the
+statistical result of Table \ref{tab_pns_resamp} that averages of these 
+quantities across all pairs of sampling locations were significantly greater 
+than zero, showing asymmetry of tail dependence.
 
@@
 \includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf}
-\caption{Non-parametric tail statistics results for Ceratium furca abundance data plotted against geographical distance $(D)$ between two sampling locations.
+\caption{Non-parametric tail statistics results for *Ceratium furca* abundance data plotted against geographical distance $(D)$ between two sampling locations.
 (A) Spearman correlation vs.
@@
 
-Table \ref{tab_pns_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean.
-Fourth column named $fracgt0$ in Table \ref{tab_pns_resamp} indicates the fraction of resampled mean greater then zero.
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ statistic with positive mean and almost $100$% $fracgt0$ proves that plankton species data for *Ceratium furca* has significant lower tail dependence.
+<!--***DAN: stopped here 2018 04 07-->
 

### 2018-04-08 SupportingInformation.Rmd
minor typo fixed

@@
 
-The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{tab_pns_LTdep_AICw} and\ref{tab_pns_UTdep_AICw}, respectively.
+The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{tab_pns_LTdep_AICw} and \ref{tab_pns_UTdep_AICw}, respectively.
 These values were exclusively positive, thereby
@@
 spatial resampling scheme, as described in Methods (Table \ref{tab_pns_resamp}).
-Confidence intervals of $\cor_l$","$\cor_u$","$\Ps_l$","$\Ps_u$","$\Dsq_l$","$\Dsq_u$"
+Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
 always excluded $0$ and mean values were positive, indicating tail dependence.
@@
 their opposite meaning (Figure \ref{fig_pns} G-H).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_pns} I-K) did not show a decline with distance, but did demonstrate visuaally the
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_pns} I-K) did not show a decline with distance, but did demonstrate visually the
 statistical result of Table \ref{tab_pns_resamp} that averages of these 

### 2018-04-08 SupportingInformation.Rmd
minor edit

@@
    * We threw out all locations that had fewer than 45 years data
-   * We use species=16 [Name : "Ceratium_furca"] for Plankton_North_Sea data
+   * We use species=16 [Name : "Ceratium furca"] for Plankton_North_Sea data
    * *Ceratium furca* is a dicoflagellate commonly found in UK seas
@@
 \includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf}
-\caption{Non-parametric tail statistics results for *Ceratium furca* abundance data plotted against geographical distance $(D)$ between two sampling locations.
+\caption{Non-parametric tail statistics results for \textit{Ceratium furca} abundance data plotted against geographical distance $(D)$ between two sampling locations.
 (A) Spearman correlation vs.

### 2018-04-08 SupportingInformation.Rmd
typo

@@
 
-![Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_l-\Dsq_u$ reveal the know asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
+![Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the know asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
 See text for details.

### 2018-04-09 SupportingInformation.Rmd
a few developments in the Taylor's law stuff

--- 2018-04-08 SupportingInformation.Rmd
@@
 
-The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{tab_pns_LTdep_AICw} and \ref{tab_pns_UTdep_AICw}, respectively.
+The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{tab_pns_LTdep_AICw} and\ref{tab_pns_UTdep_AICw}, respectively.
 These values were exclusively positive, thereby
@@
 spatial resampling scheme, as described in Methods (Table \ref{tab_pns_resamp}).
-Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
+Confidence intervals of $\cor_l$","$\cor_u$","$\Ps_l$","$\Ps_u$","$\Dsq_l$","$\Dsq_u$"
 always excluded $0$ and mean values were positive, indicating tail dependence.
@@
 their opposite meaning (Figure \ref{fig_pns} G-H).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_pns} I-K) did not show a decline with distance, but did demonstrate visually the
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_pns} I-K) did not show a decline with distance, but did demonstrate visuaally the
 statistical result of Table \ref{tab_pns_resamp} that averages of these 
@@
 
-# Question 3 analyses for each dataset
+# Question 3 analyses for each dataset and theoretical analyses
 Recall that question 3 is: What are the consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?
@@
 #***generate the starting data, which has normal copula structure
-#because it comes from a multivariate normal
+#because it comes originally from a multivariate normal, before 
+#transformation
 sig<-matrix(rho,n,n)
@@
 <!--do a TL analysis on the data just created-->
-```{r tl_analysis, echo=F, cache=T, cache.extra=list(mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
+```{r tl_analysis, echo=F, cache=T, cache.extra=list(kend.surrog,spea.surrog,mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
 source("copsurrognd.R")
 source("TaylorsLawTools.R")
-# 
-# #We are only doing spatial taylor's law because we know a priori temporal
-# #won't give anything interesting for the surrogates we use
-# 
+
+#We are only doing spatial taylor's law because we know a priori temporal
+#won't give anything interesting for the surrogates we use
+
 kend.stats.res<-list()
@@
    kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)
-# 
+
    shold<-tl_stats(ss[,,sc])
@@
 
-<!--Now make some plots-->
-```{r taylor_plots, echo=T}
+<!--Now make some plots, these one for kendall surrogates-->
+```{r taylor_plots, echo=F}
 ##plot TL for clayton
@@
 ylimits<-range(kend.points.res[[1]][,2,],kend.points.res[[2]][,2,],kend.points.res[[3]][,2,],kend.points.res[[4]][,2,])
-par(mfrow=c(1,1))
-plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
-```
-
-```{r, echo=T}
+par(mfrow=c(1,1),mar=c(4,4,1,1) + 0.1)
+plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
+points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
+points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
+
 ##plot TL for normal copula
-plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
-```
-
-```{r, echo=T}
+plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
+points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
+points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
+
 #plot TL for frank
-plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
-```
-
-```{r, echo=T}
+plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
+points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
+points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
+
 #plot TL for survival clayton
-plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits)  
+plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
+points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
+points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
+
+#prepare for plots coming below
+s.lin<-cbind(kend.stats.res[[1]][1,],kend.stats.res[[2]][1,],kend.stats.res[[3]][1,],kend.stats.res[[4]][1,])
+s.hom<-cbind(kend.stats.res[[1]][2,],kend.stats.res[[2]][2,],kend.stats.res[[3]][2,],kend.stats.res[[4]][2,])
+s.rmse<-cbind(kend.stats.res[[1]][3,],kend.stats.res[[2]][3,],kend.stats.res[[3]][3,],kend.stats.res[[4]][3,])
+s.int<-cbind(kend.stats.res[[1]][4,],kend.stats.res[[2]][4,],kend.stats.res[[3]][4,],kend.stats.res[[4]][4,])
+s.slope<-cbind(kend.stats.res[[1]][5,],kend.stats.res[[2]][5,],kend.stats.res[[3]][5,],kend.stats.res[[4]][5,])
+
+#plot s.lin distr for each set of surrogates
+#***DAN: x axis needs to be redone to show the copula names used in the surrogates
+par(mfrow=c(1,1),mar=c(4,4,1,1) + 0.1)
+x<-1:4
+plot(x,s.lin[1,],type="b",ylim=range(s.lin),xlab="Set of surrogates",ylab="Linearity test p-vals")
+for (counter in 2:dim(s.lin)[1])
+{
+  lines(x,s.lin[counter,],type='b')
+}
+
+#plot s.hom distr for each set of surrogates
+#***DAN: x axis needs to be redone to show the copula names used in the surrogates
+x<-1:4
+plot(x,s.hom[1,],type="b",ylim=range(s.hom),xlab="Set of surrogates",ylab="Homoskedasticity test p-vals")
+for (counter in 2:dim(s.slope)[1])
+{
+  lines(x,s.hom[counter,],type='b')
+}
+
+#plot s.rmse distr for each set of surrogates
+#***DAN: x axis needs to be redone to show the copula names used in the surrogates
+x<-1:4
+plot(x,s.rmse[1,],type="b",ylim=range(s.rmse),xlab="Set of surrogates",ylab="Spatial TL rmse")
+for (counter in 2:dim(s.rmse)[1])
+{
+  lines(x,s.rmse[counter,],type='b')
+}
+
+#plot s.int distr for each set of surrogates
+#***DAN: x axis needs to be redone to show the copula names used in the surrogates
+x<-1:4
+plot(x,s.int[1,],type="b",ylim=range(s.int),xlab="Set of surrogates",ylab="Spatial TL intercept")
+for (counter in 2:dim(s.int)[1])
+{
+  lines(x,s.int[counter,],type='b')
+}
+
+#plot s.slope distr for each set of surrogates
+#***DAN: x axis needs to be redone to show the copula names used in the surrogates
+x<-1:4
+plot(x,s.slope[1,],type="b",ylim=range(s.slope),xlab="Set of surrogates",ylab="Spatial TL slope")
+for (counter in 2:dim(s.slope)[1])
+{
+  lines(x,s.slope[counter,],type='b')
+}
 ```

### 2018-04-09 SupportingInformation.Rmd
minor change in anolis data calculation

@@
 source("getcopula.R")
-phylo_dat<-read.csv("./Data/Anolis/contrasts.csv",header = F)
+phylo_dat<-read.csv("./Data/Anolis/contrasts.csv",header = T)
 # Input :

### 2018-04-11 SupportingInformation.Rmd
typo

@@
 also tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and 
-$\Dsq_u-\Dsq_u$ revealed the asymmetry of tail dependence known to be present in data simulated
+$\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated
 from various copulas. Normal and Frank copula families 
@@
 of the selected parameter values we generated  $n$ points, 100 times, and thereby computed 100 values of the 
-statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_u$. This exercise 
+statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$. This exercise 
 was repeated, separately, 

### 2018-04-12 SupportingInformation.Rmd
new sections added for causes after meeting between Dan and Shyamolina

@@
 
-# Question 2 analyses for each dataset
+# Question 2 analyses 
 Recall that question 2 is: What are some possible causes of non-Gaussian copula structure?
+
+## The Moran effect and copula structure in populations
+
+We show here that if a non-normal copula structure describes the dependence of an environmental
+variable between two locations, then there can be a non-normal copula structure for the
+dependence of populations in the locations that are influenced by the environmental variables.
+
+<!--Shya to implement the modelling we discussed on 2018 04 12 and make results appear here!-->
+
+## Asymmetric sensitivity and copula structure in populations
+
+<!--Another model, distinct from the above, Dan and Shya to discuss-->
+
+## Need to think about mechanisms for biogeochemical, evolutionary, and community data
 

### 2018-04-13 SupportingInformation.Rmd
adding chunk for moran effect

@@
 <!--Shya to implement the modelling we discussed on 2018 04 12 and make results appear here!-->
+```{r cause4copula}
+source("Cause4copula.R")
+set.seed(seed)
+#--------------------------------------------------------------------------------
+# when noise comes from a clayton cop with spearmancor=0.8
+s<-GetNoise(N=10000,fcode=3,corcoef=0.8,method="spearman",ploton=T)
+s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q,burnin=1000)
+#hist(s2[,1],breaks=1000) # it's normal
+s2_p<-pnorm(s2)
+#hist(s2_p[,1],breaks=1000) # it's uniform
+points(s2_p[,1],s2_p[,2],col="red")
+
+# Now compare between two correlations coming from noise_q and s2_p
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "spearman")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "spearman")$estimate
+
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "kendall")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "kendall")$estimate
+
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "pearson")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "pearson")$estimate
+#-----------------------------------------------------------------------------------------
+# when noise comes from a clayton cop with kendallcor=0.8
+s<-GetNoise(N=10000,fcode=3,corcoef=0.8,method="kendall",ploton=T)
+s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q,burnin=1000)
+#hist(s2[,1],breaks=1000) # it's normal
+s2_p<-pnorm(s2)
+#hist(s2_p[,1],breaks=1000) # it's uniform
+points(s2_p[,1],s2_p[,2],col="red")
+
+# Now compare between two correlations coming from noise_q and s2_p
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "spearman")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "spearman")$estimate
+
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "kendall")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "kendall")$estimate
+
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "pearson")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "pearson")$estimate
+#----------------------------------------------------------------------------------
+# when noise comes from a gumbel cop with spearmancor=0.8
+s<-GetNoise(N=10000,fcode=4,corcoef=0.8,method="spearman",ploton=T)
+s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q,burnin=1000)
+#hist(s2[,1],breaks=1000) # it's normal
+s2_p<-pnorm(s2)
+#hist(s2_p[,1],breaks=1000) # it's uniform
+points(s2_p[,1],s2_p[,2],col="red")
+
+# Now compare between two correlations coming from noise_q and s2_p
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "spearman")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "spearman")$estimate
+
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "kendall")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "kendall")$estimate
+
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "pearson")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "pearson")$estimate
+#------------------------------------------------------------------------------------
+# when noise comes from a gumbel cop with kendallcor=0.8
+s<-GetNoise(N=10000,fcode=4,corcoef=0.8,method="kendall",ploton=T)
+s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q,burnin=1000)
+#hist(s2[,1],breaks=1000) # it's normal
+s2_p<-pnorm(s2)
+#hist(s2_p[,1],breaks=1000) # it's uniform
+points(s2_p[,1],s2_p[,2],col="red")
+
+# Now compare between two correlations coming from noise_q and s2_p
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "spearman")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "spearman")$estimate
+
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "kendall")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "kendall")$estimate
+
+cor.test(s$noise_q[,1],s$noise_q[,2],method = "pearson")$estimate
+cor.test(s2_p[,1],s2_p[,2],method = "pearson")$estimate
+
+```
 

### 2018-04-13 SupportingInformation.Rmd

--- 2018-04-12 SupportingInformation.Rmd
@@
 <!--Shya to implement the modelling we discussed on 2018 04 12 and make results appear here!-->
+```{r cause4copula}
+source("Cause4copula.R")
+set.seed(seed)
+#--------------------------------------------------------------------------------
+#function to get a comparison table 
+comp<-function(s,s2_p){
+
+  comp<-matrix(NA,nrow=3,ncol=2)
+  rownames(comp)<-c("spearman","kendall","pearson")
+  colnames(comp)<-c("noise_q","pop_p")
+  # Now compare between two correlations coming from noise_q and s2_p
+  comp[1,1]<-unname(cor.test(s$noise_q[,1],s$noise_q[,2],method = "spearman")$estimate)
+  comp[1,2]<-unname(cor.test(s2_p[,1],s2_p[,2],method = "spearman")$estimate)
+
+  comp[2,1]<-unname(cor.test(s$noise_q[,1],s$noise_q[,2],method = "kendall")$estimate)
+  comp[2,2]<-unname(cor.test(s2_p[,1],s2_p[,2],method = "kendall")$estimate)
+
+  comp[3,1]<-unname(cor.test(s$noise_q[,1],s$noise_q[,2],method = "pearson")$estimate)
+  comp[3,2]<-unname(cor.test(s2_p[,1],s2_p[,2],method = "pearson")$estimate)
+
+  return(comp)
+}
+#--------------------------------------------------------------------------------
+# when noise comes from a clayton cop with spearmancor=0.8
+s<-GetNoise(N=10000,fcode=3,corcoef=0.8,method="spearman",ploton=T)
+s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q,burnin=1000)
+#hist(s2[,1],breaks=1000) # it's normal
+s2_p<-pnorm(s2)
+#hist(s2_p[,1],breaks=1000) # it's uniform
+points(s2_p[,1],s2_p[,2],col="red")
+comp(s=s,s2_p=s2_p)
+#-----------------------------------------------------------------------------------------
+# when noise comes from a clayton cop with kendallcor=0.8
+s<-GetNoise(N=10000,fcode=3,corcoef=0.8,method="kendall",ploton=T)
+s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q,burnin=1000)
+#hist(s2[,1],breaks=1000) # it's normal
+s2_p<-pnorm(s2)
+#hist(s2_p[,1],breaks=1000) # it's uniform
+points(s2_p[,1],s2_p[,2],col="red")
+comp(s=s,s2_p=s2_p)
+#----------------------------------------------------------------------------------
+# when noise comes from a gumbel cop with spearmancor=0.8
+s<-GetNoise(N=10000,fcode=4,corcoef=0.8,method="spearman",ploton=T)
+s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q,burnin=1000)
+#hist(s2[,1],breaks=1000) # it's normal
+s2_p<-pnorm(s2)
+#hist(s2_p[,1],breaks=1000) # it's uniform
+points(s2_p[,1],s2_p[,2],col="red")
+comp(s=s,s2_p=s2_p)
+#------------------------------------------------------------------------------------
+# when noise comes from a gumbel cop with kendallcor=0.8
+s<-GetNoise(N=10000,fcode=4,corcoef=0.8,method="kendall",ploton=T)
+s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q,burnin=1000)
+#hist(s2[,1],breaks=1000) # it's normal
+s2_p<-pnorm(s2)
+#hist(s2_p[,1],breaks=1000) # it's uniform
+points(s2_p[,1],s2_p[,2],col="red")
+comp(s=s,s2_p=s2_p)
+```
 

### 2018-04-14 SupportingInformation.Rmd
adding plot for different copula to find Moran causes

--- 2018-04-12 SupportingInformation.Rmd
@@
 <!--Shya to implement the modelling we discussed on 2018 04 12 and make results appear here!-->
+```{r cause4copula, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"))}
+source("Cause4copula.R")
+set.seed(seed)
+
+pdf("./Results/Cause4copula_results/Clayton_spearman.pdf")
+Plotter_Cause4copula_stat(numsim=500,fcode=3,method="spearman",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/Clayton_kendall.pdf")
+Plotter_Cause4copula_stat(numsim=500,fcode=3,method="spearman",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/Gumbel_spearman.pdf")
+Plotter_Cause4copula_stat(numsim=500,fcode=4,method="spearman",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/Gumbel_kendall.pdf")
+Plotter_Cause4copula_stat(numsim=500,fcode=4,method="kendall",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
+Plotter_Cause4copula_stat(numsim=500,fcode=13,method="spearman",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
+Plotter_Cause4copula_stat(numsim=500,fcode=13,method="kendall",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/SGumbel_spearman.pdf")
+Plotter_Cause4copula_stat(numsim=500,fcode=14,method="spearman",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/SGumbel_kendall.pdf")
+Plotter_Cause4copula_stat(numsim=500,fcode=14,method="kendall",lb=0,ub=0.1)
+dev.off()
+```
 

### 2018-04-14 Cause4copula.R

@@ -1,5 +1,6 @@
 library(copula)
 library(VineCopula)
+source("CopulaFunctions_flexible.R")
 #---------------------------------------------------------------------------------------
 # This function GetNoise generates noises with a copula structure as you specified
 # Input :
@@ -45,7 +46,7 @@ GetNoise<-function(N,fcode,corcoef,method,ploton){
 # apply qnorm on noisecop to get normal distribution of each marginal
   noise_q<-qnorm(noisecop)
   
-  return(list(noise=noisecop,noise_q=noise_q,param=param))
+  return(list(noise_c=noisecop,noise_q=noise_q,param=param))
 }
 
 # Check the function
@@ -68,12 +69,13 @@ GetNoise<-function(N,fcode,corcoef,method,ploton){
 #noise       An N by 2 matrix of environment variables in two habitat patches through time 
 #                             (this should be noise_q from output of GetNoise function)
 #burnin       The number of population times to throw away to get rid of transient behavior in the
-#               dynamics. Default 500.
+#               dynamics. Default 1000.
 
 #Output
-#An N+1 by 2 matrix of populations through time, p0 is the first row
+#An N+1 by 2 matrix of populations through time, p0 is the first row in copula space : pop_c
+# pop_q : this is similar but with normal distribution marginal
 
-Simulator_Cause4copula<-function(cons,p0=c(0,0),noise,burnin=500){
+Simulator_Cause4copula<-function(cons,p0=c(0,0),noise,burnin=1000){
   
   N<-dim(noise)[1]
   cons2<-sqrt(1-cons^2)
@@ -88,28 +90,287 @@ Simulator_Cause4copula<-function(cons,p0=c(0,0),noise,burnin=500){
     res<-res[-(1:burnin),]
   }
   
-  return(res)
+  res2<-pnorm(res) # convert into copula space
[…]

### 2018-04-15 SupportingInformation.Rmd
adding plots...still need to modify

@@
 set.seed(seed)
-
-pdf("./Results/Cause4copula_results/Clayton_spearman.pdf")
-Plotter_Cause4copula_stat(numsim=500,fcode=3,method="spearman",lb=0,ub=0.1)
+#---------------------------------------
+pdf("./Results/Cause4copula_results/compare_params_pGOF.pdf", width=12, height=6)
+op<-par(mfrow=c(2,4),mar=c(3,3.5,3,3.5), mgp=c(1.5,0.5,0))
+
+Plotter_Cause4copula_GOF(fcode=3,method = "spearman")
+Plotter_Cause4copula_GOF(fcode=13,method="spearman")
+Plotter_Cause4copula_GOF(fcode=4,method = "spearman")
+Plotter_Cause4copula_GOF(fcode=14,method="spearman")
+
+Plotter_Cause4copula_GOF(fcode=3,method = "kendall")
+Plotter_Cause4copula_GOF(fcode=13,method="kendall")
+Plotter_Cause4copula_GOF(fcode=4,method = "kendall")
+Plotter_Cause4copula_GOF(fcode=14,method="kendall")
+
+par(op)
+
+op2<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
+plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
+legend("top", c("noise","population"), col = c("red", "blue"),
+       cex = 0.8, pch = c(2, 6), lty = c(1, 1), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
+       bty = "n") 
+par(op2)
 dev.off()
-
-pdf("./Results/Cause4copula_results/Clayton_kendall.pdf")
-Plotter_Cause4copula_stat(numsim=500,fcode=3,method="spearman",lb=0,ub=0.1)
-dev.off()
-
-pdf("./Results/Cause4copula_results/Gumbel_spearman.pdf")
-Plotter_Cause4copula_stat(numsim=500,fcode=4,method="spearman",lb=0,ub=0.1)
-dev.off()
-
-pdf("./Results/Cause4copula_results/Gumbel_kendall.pdf")
-Plotter_Cause4copula_stat(numsim=500,fcode=4,method="kendall",lb=0,ub=0.1)
-dev.off()
-
-pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
-Plotter_Cause4copula_stat(numsim=500,fcode=13,method="spearman",lb=0,ub=0.1)
-dev.off()
-
-pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
-Plotter_Cause4copula_stat(numsim=500,fcode=13,method="kendall",lb=0,ub=0.1)
-dev.off()
-
-pdf("./Results/Cause4copula_results/SGumbel_spearman.pdf")
-Plotter_Cause4copula_stat(numsim=500,fcode=14,method="spearman",lb=0,ub=0.1)
-dev.off()
-
-pdf("./Results/Cause4copula_results/SGumbel_kendall.pdf")
-Plotter_Cause4copula_stat(numsim=500,fcode=14,method="kendall",lb=0,ub=0.1)
-dev.off()
+#-------------------------------------------------------------
+#pdf("./Results/Cause4copula_results/Clayton_spearman.pdf")
+#Plotter_Cause4copula_stat(numsim=500,fcode=3,method="spearman",lb=0,ub=0.1)
+#dev.off()
+
+#pdf("./Results/Cause4copula_results/Clayton_kendall.pdf")
+#Plotter_Cause4copula_stat(numsim=500,fcode=3,method="spearman",lb=0,ub=0.1)
+#dev.off()
+
+#pdf("./Results/Cause4copula_results/Gumbel_spearman.pdf")
+#Plotter_Cause4copula_stat(numsim=500,fcode=4,method="spearman",lb=0,ub=0.1)
+#dev.off()
+
+#pdf("./Results/Cause4copula_results/Gumbel_kendall.pdf")
+#Plotter_Cause4copula_stat(numsim=500,fcode=4,method="kendall",lb=0,ub=0.1)
+#dev.off()
+
+#pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
+#Plotter_Cause4copula_stat(numsim=500,fcode=13,method="spearman",lb=0,ub=0.1)
+#dev.off()
+
+#pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
+#Plotter_Cause4copula_stat(numsim=500,fcode=13,method="kendall",lb=0,ub=0.1)
+#dev.off()
+
+#pdf("./Results/Cause4copula_results/SGumbel_spearman.pdf")
+#Plotter_Cause4copula_stat(numsim=500,fcode=14,method="spearman",lb=0,ub=0.1)
+#dev.off()
+
+#pdf("./Results/Cause4copula_results/SGumbel_kendall.pdf")
+#Plotter_Cause4copula_stat(numsim=500,fcode=14,method="kendall",lb=0,ub=0.1)
+#dev.off()
 ```

### 2018-04-15 Cause4copula.R

@@ -69,13 +69,13 @@ GetNoise<-function(N,fcode,corcoef,method,ploton){
 #noise       An N by 2 matrix of environment variables in two habitat patches through time 
 #                             (this should be noise_q from output of GetNoise function)
 #burnin       The number of population times to throw away to get rid of transient behavior in the
-#               dynamics. Default 1000.
+#               dynamics. Default 500.
 
 #Output
 #An N+1 by 2 matrix of populations through time, p0 is the first row in copula space : pop_c
 # pop_q : this is similar but with normal distribution marginal
 
-Simulator_Cause4copula<-function(cons,p0=c(0,0),noise,burnin=1000){
+Simulator_Cause4copula<-function(cons,p0=c(0,0),noise,burnin=500){
   
   N<-dim(noise)[1]
   cons2<-sqrt(1-cons^2)
@@ -130,7 +130,81 @@ comp<-function(s,s2){
 #hist(s2$pop_c[,1],breaks=1000) # it's uniform
 #points(s2$pop_c[,1],s2$pop_c[,2],col="red")
 #comp(s=s,s2=s2)
+#---------------------------------------------------------
+#-----------------------------------------------------------------------------------------------
+#-----------------------------------------------------------------------------------------------
+Plotter_Cause4copula_GOF<-function(fcode,method){
+  
+  corcoef_list<-seq(from=0.1,to=0.9,by=0.1)
+  
+  # initialize
+  par_noise<-c()
+  par_pop<-c()
+  pval_min<-c()
+  
+  for(corcoef in corcoef_list){
+    
+    #cat("corcoef=",corcoef,"\n")
+    
+    # generate noise copula 
+    s<-GetNoise(N=5000,fcode=fcode,corcoef=corcoef,method=method,ploton=F)
+    # generate pop_copula
+    s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q,burnin=500)
+    
+    # compare btw s$noise_c and s2$pop_c
+    z<-BiCopEst(u1=s2$pop_c[,1],u2=s2$pop_c[,2],family=fcode)
+    
+    u1<-s2$pop_c[,1]
+    u2<-s2$pop_c[,2]
+    zf<-BiCopGofTest(u1,u2,family=fcode,method = "kendall",B=100)
[…]

### 2018-04-17 SupportingInformation.Rmd
plot fixing,adding chunk for BMR data

@@
 ```{r calculation_phylo_data, echo=F, results="hide"}
-source("getcopula.R")
+source("phylo_dat_fn.R")
 phylo_dat<-read.csv("./Data/Anolis/contrasts.csv",header = T)
-# Input :
-#        data : a dataframe : phylo_dat with dim.=[99 x 14]
-#        i, j : column indices of data frame to get bivariate data set, should be (i,j)<=13
-#        ploton : logical for option plots of raw data and it's copula representation
-# Output : 
-#        a cop rank matrix for the raw data
-phylo_dat_fn<-function(data,i,j,ploton=T){
-
-  data<-data[,2:dim(phylo_dat)[2]]
-  d<-data[,c(i,j)]
-
-  v_phylo<-getcopula(d=d,rankon = T,ploton = F)
-
-  if(ploton==T){
-    op<-par(mfrow=c(1,2))
-    plot(d[,1],d[,2],col="blue")
-    getcopula(d=d,rankon = T,ploton = T)
-    par(op)
-  }
-  return(v_phylo)
-}
 
@@
 ### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}
-<!--Do the intensive part of the analysis, calculating stats on the data and surrogates-->
-```{r stat_soilCN_graphic_approach, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
-source("CopulaFunctions_flexible.R")
-source("copsurrog2d.R")
+
+```{r stat_soilCN_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_CN,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+source("bivfunctionplot.R")
 set.seed(seed) 
-
-#utility function for calculating all the stats we want
-calcstats<-function(v,f,nm)
-{
-  #calculate the stats
-  res<-c()
-  for (lbd in seq(from=0,to=0.9,by=0.1))
-  {
-    res<-c(res,f(v[,1],v[,2],lbd,lbd+.1))
-  }
-
-  #name the result vector
-  h<-paste(nm,c(seq(from=0,to=0.9,by=0.1)),"to",
-                    c(seq(from=0,to=0.9,by=0.1)+.1),sep='')
-  h<-gsub(".","p",h,fixed=T)
-  names(res)<-h
-
-  return(res)
-}
-
-#Do the kendall surrogates
-cop<-normalCopula(.5,2)
-numsurrog<-1000
-surv_K<-copsurrog2d(v_CN,cop,"kendall",numsurrog)
-
-#Do the spearman surrogates
-surv_S<-copsurrog2d(v_CN,cop,"spearman",numsurrog)
-
-#utility function for calculating fractions of surrogates 
-#with a stat smaller than that of data, i.e., fraction of 
-#surrogates for which the stat value on the data is bigger
-fracwork<-function(dvals,surrvals)
-{
-  frac<-NA*numeric(length(dvals))
-  names(frac)<-names(dvals)
-  for (counter in 1:length(frac))
-  {
-    frac[counter]<-sum(surrvals[counter,]<dvals[counter])
-  }
-  return(frac)
-}
-
-#now work with the correlation stats
-corstats_d<-calcstats(v_CN,Corbds,"Cor")
-corstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Corbds,nm="Cor")
-corstats_frac_K<-fracwork(corstats_d,corstats_K)
-
-corlmcoru_d<-unname(corstats_d[1]-corstats_d[length(corstats_d)])
-corlmcoru_K<-corstats_K[1,]-corstats_K[length(corstats_d),]
-corlmcoru_frac_K<-sum(corlmcoru_K<corlmcoru_d)
-corstats_K<-apply(FUN=quantile,X=corstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
-
-corstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Corbds,nm="Cor")
-corstats_frac_S<-fracwork(corstats_d,corstats_S)
-
-corlmcoru_S<-corstats_S[1,]-corstats_S[length(corlmcoru_d),]
-corlmcoru_frac_S<-sum(corlmcoru_S<corlmcoru_d)
-corstats_S<-apply(FUN=quantile,X=corstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
-
-#now work with the P stats
-Pbds_wrap<-function(vi,vj,lb,ub)
-{
-  return(Pbds(vi,vj,lb,ub)$abs_res)
-}
-
-Pstats_d<-calcstats(v_CN,Pbds_wrap,"P")
-Pstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Pbds_wrap,nm="P")
-Pstats_frac_K<-fracwork(Pstats_d,Pstats_K)
-
-PlmPu_d<-unname(Pstats_d[1]-Pstats_d[length(Pstats_d)])
-PlmPu_K<-Pstats_K[1,]-Pstats_K[length(Pstats_d),]
-PlmPu_frac_K<-sum(PlmPu_K<PlmPu_d)
-Pstats_K<-apply(FUN=quantile,X=Pstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
-
-Pstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Pbds_wrap,nm="P")
-Pstats_frac_S<-fracwork(Pstats_d,Pstats_S)
-
-PlmPu_S<-Pstats_S[1,]-Pstats_S[length(Pstats_d),]
-PlmPu_frac_S<-sum(PlmPu_S<PlmPu_d)
-Pstats_S<-apply(FUN=quantile,X=Pstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
-
-#now work with the D2 stats
-D2stats_d<-calcstats(v_CN,D2bds,"Dtwo")
-D2stats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=D2bds,nm="Dtwo")
-D2stats_frac_K<-fracwork(D2stats_d,D2stats_K)
-
-D2umD2l_d<-D2stats_d[length(D2stats_d)]-D2stats_d[1]
-D2umD2l_K<-D2stats_K[length(D2stats_d),]-D2stats_K[1,]
-D2umD2l_frac_K<-sum(D2umD2l_K<D2umD2l_d)
-D2stats_K<-apply(FUN=quantile,X=D2stats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))
-
-D2stats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=D2bds,nm="Dtwo")
-D2stats_frac_S<-fracwork(D2stats_d,D2stats_S)
-
-D2umD2l_S<-D2stats_S[length(D2stats_d),]-D2stats_S[1,]
-D2umD2l_frac_S<-sum(D2umD2l_S<D2umD2l_d)
-D2stats_S<-apply(FUN=quantile,X=D2stats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
-```
-
-<!--Now create the plot and save it to the results folder-->
-```{r plot_soilCN_graphic_approach, results='hide', echo=F}
-x<-seq(from=0.05,to=0.95,by=0.1)
-xlimits<-c(0,1)
-
-#plotting layout, units inches
-xht<-0.5
-ywd<-0.5
-titleht<-.25
-panht<-1.5
-panwd<-3
-gap<-.05
-totwd<-ywd+2*panwd+2*gap
-totht<-xht+3*panht+3*gap+titleht
-pdf(file="./Results/stat_results/stat_soilCN/CNdata_NonparamTailDep.pdf",width=totwd,height=totht)
-
-#plot kendall results on the left panels, spearman on right,
-#cor in top panels, then P, then D2 on bottom panels
-
-#kendall, cor
-par(fig=c(ywd/totwd,
-          (ywd+panwd)/totwd,
-          (xht+2*panht+2*gap)/totht,
-          (xht+3*panht+2*gap)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
-ylimits_cor<-range(corstats_d,corstats_K,corstats_K)
-ylimits_cor[2]<-ylimits_cor[2]+.3*diff(ylimits_cor)
-plot(x,corstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_cor,
-     xaxt='n')
-mtext(side=3,line=0,text="Kendall-preserving surrogates")
-mtext(side=2,line=1,text="Partial correlation")
-axis(side=1,labels=F)
-lines(x,corstats_K[1,],type='l',lty='dotted')
-lines(x,corstats_K[2,],type='l',lty='dashed')
-lines(x,corstats_K[3,],type='l',lty='dashed')
-lines(x,corstats_K[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_cor[1],labels='A',cex=1.5,adj=c(.5,0))
-fracplot<-function(fracs,ylims)
-{
-  for (counter in 1:length(fracs))
-  {
-    ptxt<-''
-    if (fracs[counter]>.975*numsurrog)
-    {
-      ptxt<-paste(">",fracs[counter],sep='')
-    } 
-    if (fracs[counter]<.025*numsurrog)
-    {
-      ptxt<-paste("<",numsurrog-fracs[counter],sep='')
-    }
-    yht<-ylims[2]-.1*diff(ylims)
-    text(x[counter],yht,ptxt,adj=c(0.5,.5),cex=0.5,srt=90)
-  }
-}
-fracplot(corstats_frac_K,ylimits_cor)
-vlineplot<-function(x,ylims)
-{
-  for (counter in 1:length(x))
-  {
-    lines(rep(x[counter]-(x[2]-x[1])/2,2),ylims,type='l',lty='dotted')
-  }
-  lines(rep(x[length(x)]+(x[2]-x[1])/2,2),ylims,type='l',lty='dotted')
-}
-vlineplot(x,ylimits_cor)
-
-#spearman, cor
-par(fig=c((ywd+panwd+gap)/totwd,
-          (ywd+2*panwd+gap)/totwd,
-          (xht+2*panht+2*gap)/totht,
-          (xht+3*panht+2*gap)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-plot(x,corstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_cor,
-     xaxt='n',yaxt='n')
-mtext(side=3,line=0,text="Spearman-preserving surrogates")
-axis(side=1,labels=F)
-axis(side=2,labels=F)
-lines(x,corstats_S[1,],type='l',lty='dotted')
-lines(x,corstats_S[2,],type='l',lty='dashed')
-lines(x,corstats_S[3,],type='l',lty='dashed')
-lines(x,corstats_S[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_cor[1],labels='B',cex=1.5,adj=c(.5,0))
-fracplot(corstats_frac_S,ylimits_cor)
-vlineplot(x,ylimits_cor)
-
-#kendall, P
-par(fig=c(ywd/totwd,
-          (ywd+panwd)/totwd,
-          (xht+panht+gap)/totht,
-          (xht+2*panht+gap)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-ylimits_D2<-range(D2stats_d,D2stats_K,D2stats_K)
-ylimits_D2[2]<-ylimits_D2[2]+.3*diff(ylimits_D2)
-ylimits_P<-range(Pstats_d,Pstats_K,Pstats_K)
-ylimits_P[2]<-ylimits_P[2]+.3*diff(ylimits_P)
-plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,xaxt='n')
-axis(side=1,labels=F)
-mtext(side=2,line=1,text="P")
-lines(x,Pstats_K[1,],type='l',lty='dotted')
-lines(x,Pstats_K[2,],type='l',lty='dashed')
-lines(x,Pstats_K[3,],type='l',lty='dashed')
-lines(x,Pstats_K[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_P[1],labels='C',cex=1.5,adj=c(.5,0))
-fracplot(Pstats_frac_K,ylimits_P)
-vlineplot(x,ylimits_P)
-
-#spearman, p
-par(fig=c((ywd+panwd+gap)/totwd,
-          (ywd+2*panwd+gap)/totwd,
-          (xht+panht+gap)/totht,
-          (xht+2*panht+gap)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,
-     xaxt='n',yaxt='n')
-axis(side=1,labels=F)
-axis(side=2,labels=F)
-lines(x,Pstats_S[1,],type='l',lty='dotted')
-lines(x,Pstats_S[2,],type='l',lty='dashed')
-lines(x,Pstats_S[3,],type='l',lty='dashed')
-lines(x,Pstats_S[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_P[1],labels='D',cex=1.5,adj=c(.5,0))
-fracplot(Pstats_frac_S,ylimits_P)
-vlineplot(x,ylimits_P)
-
-#kendall, D2
-par(fig=c(ywd/totwd,
-          (ywd+panwd)/totwd,
-          (xht)/totht,
-          (xht+panht)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2)
-mtext(side=1,line=1,text="Diagonal slice")
-mtext(side=2,line=1,text=expression(D^{2}))
-axis(side=1,labels=F)
-lines(x,D2stats_K[1,],type='l',lty='dotted')
-lines(x,D2stats_K[2,],type='l',lty='dashed')
-lines(x,D2stats_K[3,],type='l',lty='dashed')
-lines(x,D2stats_K[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_D2[1],labels='E',cex=1.5,adj=c(.5,0))
-fracplot(D2stats_frac_K,ylimits_D2)
-vlineplot(x,ylimits_D2)
-
-#spearman, P
-par(fig=c((ywd+panwd+gap)/totwd,
-          (ywd+2*panwd+gap)/totwd,
-          (xht)/totht,
-          (xht+panht)/totht),
-    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
-plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2,
-     yaxt='n')
-mtext(side=1,line=1,text="Diagonal slice")
-axis(side=1,labels=F)
-axis(side=2,labels=F)
-lines(x,D2stats_S[1,],type='l',lty='dotted')
-lines(x,D2stats_S[2,],type='l',lty='dashed')
-lines(x,D2stats_S[3,],type='l',lty='dashed')
-lines(x,D2stats_S[4,],type='l',lty='dotted')
-text(xlimits[1],ylimits_D2[1],labels='F',cex=1.5,adj=c(.5,0))
-fracplot(D2stats_frac_S,ylimits_D2)
-vlineplot(x,ylimits_D2)
-
-dev.off()
+stat_soilCN_ga<-bivfunctionplot(v=v_CN,resloc="./Results/stat_results/stat_soilCN/",nametag="soilCN",numbin=10)
 ```
@@
 When $\cor$ values (respectively, $\Ps$ or $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, stronger than or weaker than for $\Ps$ or $\Dsq$) expected from a normal-copula null hypothesis.
-\label{fig_soilCN_nonparam}](./Results/stat_results/stat_soilCN/CNdata_NonparamTailDep.pdf)
+\label{fig_soilCN_nonparam}](./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf)
 
@@
 <!--Now make and display a table, this one about asymmetry of tail dependence-->
-```{r plot_soilCN_asymmtry_taildep, echo=F}
+```{r make_table_stat_fn, echo=F}
+
+fractxt<-function(fracval,numsurrog=1000){
+  ptxt<-''
+  if (fracval>.5*numsurrog){                        # why 0.5 factor?
+    ptxt<-paste(">",fracval,sep='')
+  } 
+  if (fracval<=.5*numsurrog){
+    ptxt<-paste("<",numsurrog-fracval,sep='')
+  }
+  return(ptxt)
+}
+
+make_table_stat_fn<-function(data){
+
  atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$","$\\Ps_{0,0.1}-\\Ps_{0.9,1}$","$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
@@
                    Spearman=rep("",3),stringsAsFactors = F)
-fractxt<-function(fracval)
-{
-  ptxt<-''
-  if (fracval>.5*numsurrog)
-  {
-    ptxt<-paste(">",fracval,sep='')
-  } 
-  if (fracval<=.5*numsurrog)
-  {
-    ptxt<-paste("<",numsurrog-fracval,sep='')
-  }
-  return(ptxt)
-}
-atdres[1,2]<-fractxt(corlmcoru_frac_K)
-atdres[1,3]<-fractxt(corlmcoru_frac_S)
-atdres[2,2]<-fractxt(PlmPu_frac_K)
-atdres[2,3]<-fractxt(PlmPu_frac_S)
-atdres[3,2]<-fractxt(D2umD2l_frac_K)
-atdres[3,3]<-fractxt(D2umD2l_frac_S)
+
+ atdres[1,2]<-fractxt(data$corlmcoru_frac_K)
+ atdres[1,3]<-fractxt(data$corlmcoru_frac_S)
+ atdres[2,2]<-fractxt(data$PlmPu_frac_K)
+ atdres[2,3]<-fractxt(data$PlmPu_frac_S)
+ atdres[3,2]<-fractxt(data$D2umD2l_frac_K)
+ atdres[3,3]<-fractxt(data$D2umD2l_frac_S)
+
+ return(atdres)
+
+}
 ```
@@
 ```{r tab_soilCN_asym, echo=F, results="markup"}
-knitr::kable(atdres,
+tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
+knitr::kable(tab_soilCN_asym,
              format="pandoc",
@@
              booktabs=T)
+```
+
+## Results: BMR data \label{Ex_BMR}
+
+```{r read_plot_BMR_data, echo=FALSE}
+set.seed(seed=101)
+source("phylo_dat_fn.R")
+#---------------------------------------------------------
+BMR_dat<-read.csv("./Data/BMR/bmr-contrasts.csv",header = T)
+#class(BMR_dat)
+v_BMR<-phylo_dat_fn(data=BMR_dat,i=4,j=3,ploton = F)
+
+pdf("./Results/BMR_results/BMR_vs_lnmass.pdf")
+plot(v_BMR[,1],v_BMR[,2],col="blue",xlab="ln(mass)",ylab="BMR") # copula plot btw ln(Mass) vs. BMR
+dev.off()
+```
+
+### Model selection approach \label{Ex_BMR_Model_selection}
+
+```{r bivMS_BMR, echo=F, cache=T, cache.extra=list(seed,v_BMR,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+set.seed(seed)
+source("OurBiCopSelect.R")
+#families<-c(1,3:10,13:14,16:20) 
+BivMS_BMR<-OurBiCopSelect(u1=v_BMR[,1],u2=v_BMR[,2],families=c(1,3:10,13:14,16:20),
+                          level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
+saveRDS(BivMS_BMR,file="./Results/BMR_results/BivMS_BMR_lnmass.RDS")
+```
+
+### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
+
+```{r stat_BMR_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_BMR,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+source("bivfunctionplot.R")
+set.seed(seed) 
+# Now do NPA for v_BMR
+# If "F"/"SBB8" is the best fit : then it should be symmetric tail dep (0) : Then still I need to do NPA ?
+# May be : because it will also tell me how they are different from "N" copula in different sections?
+source("bivfunctionplot.R")
+stat_BMR_ga<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR",numbin = 10)
+```
+
+```{r }
+tab_BMR_asym<-make_table_stat_fn(data=stat_BMR_ga)
+knitr::kable(tab_BMR_asym)
 ```
@@
 <!--Shya to implement the modelling we discussed on 2018 04 12 and make results appear here!-->
-```{r cause4copula, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"))}
+```{r cause4copula, echo=F, cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"))}
 source("Cause4copula.R")
-set.seed(seed)
+set.seed(seed=101)
 #---------------------------------------
@@
 plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-legend("top", c("noise","population"), col = c("red", "blue"),
-       cex = 0.8, pch = c(2, 6), lty = c(1, 1), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
+legend("top", c("noise","population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
+       cex = 0.8, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
        bty = "n") 
@@
 dev.off()
+
 #-------------------------------------------------------------

### 2018-04-17 bivfunctionplot.R

@@ -1,0 +1,297 @@
+source("CopulaFunctions_flexible.R")
+source("copsurrog2d.R")
+#-----------------------------------------------------------
+#utility function for calculating all the stats we want
+calcstats<-function(v,f,nm,numbin){
+  
+  #calculate the stats
+  res<-c()
+  bingap<-1/numbin
+  x<-seq(from=0,to=1,by=bingap)
+  for (i in 1:(length(x)-1)){
+    lbd<-x[i]
+    #print(lbd)
+    res<-c(res,f(v[,1],v[,2],lbd,lbd+bingap))
+  }
+  
+  x1<-head(x,n=-1)
+  x2<-tail(x,n=-1)
+  #name the result vector
+  h<-paste(nm,x1,"to",
+           x2,sep='')
+  h<-gsub(".","p",h,fixed=T)
+  names(res)<-h
+  
+  return(res)
+}
+#-----------------------------------------------------
+makeSurrog<-function(v,numsurrog=1000){
+  cop<-normalCopula(.5,2)
+  #numsurrog<-1000
+  surv_K<-copsurrog2d(v,cop,"kendall",numsurrog) #Do the kendall surrogates
+  surv_S<-copsurrog2d(v,cop,"spearman",numsurrog) #Do the spearman surrogates
+  return(list(surv_K=surv_K,
+              surv_S=surv_S))
+}
+#makeSurrog(v=v_CN)
+#------------------------------------------------------------------
+#utility function for calculating fractions of surrogates 
+#with a stat smaller than that of data, i.e., fraction of 
+#surrogates for which the stat value on the data is bigger
+fracwork<-function(dvals,surrvals){
+  frac<-NA*numeric(length(dvals))
+  names(frac)<-names(dvals)
+  for (counter in 1:length(frac)){
+    frac[counter]<-sum(surrvals[counter,]<dvals[counter])
+  }
+  return(frac)
+}
+#--------------------------------------------------------------------
+#now work with the P stats
[…]

### 2018-04-17 SupportingInformation.Rmd
minor change

@@
 
-![Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the know asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
+![Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
 See text for details.
@@
 
-```{r read_plot_BMR_data, echo=FALSE}
+```{r read_plot_BMR_data, echo=FALSE, results="hide"}
 set.seed(seed=101)

### 2018-04-17 SupportingInformation.Rmd
modify tables

@@
 
- atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$","$\\Ps_{0,0.1}-\\Ps_{0.9,1}$","$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
+ #atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$","$\\Ps_{0,0.1}-\\Ps_{0.9,1}$","$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
+ atdres<-data.frame(Statistic=c("$\\cor_{first.slice}-\\cor_{last.slice}$",
+                                "$\\Ps_{first.slice}-\\Ps_{last.slice}$",
+                                "$\\Dsq_{first.slice}-\\Dsq_{last.slice}$"),
                    Kendall=rep("",3),

### 2018-04-17 SupportingInformation.Rmd
Cause4copula plots are made

@@
 <!--Shya to implement the modelling we discussed on 2018 04 12 and make results appear here!-->
-```{r cause4copula, echo=F, cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"))}
+```{r cause4copula, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"))}
 source("Cause4copula.R")
@@
 
-Plotter_Cause4copula_GOF(fcode=3,method = "spearman")
-Plotter_Cause4copula_GOF(fcode=13,method="spearman")
-Plotter_Cause4copula_GOF(fcode=4,method = "spearman")
-Plotter_Cause4copula_GOF(fcode=14,method="spearman")
-
-Plotter_Cause4copula_GOF(fcode=3,method = "kendall")
-Plotter_Cause4copula_GOF(fcode=13,method="kendall")
-Plotter_Cause4copula_GOF(fcode=4,method = "kendall")
-Plotter_Cause4copula_GOF(fcode=14,method="kendall")
+Plotter_Cause4copula_GOF(fcode=3,method = "spearman",num_keep_last=500,BS=100)
+Plotter_Cause4copula_GOF(fcode=13,method="spearman",num_keep_last=500,BS=100)
+Plotter_Cause4copula_GOF(fcode=4,method = "spearman",num_keep_last=500,BS=100)
+Plotter_Cause4copula_GOF(fcode=14,method="spearman",num_keep_last=500,BS=100)
+
+Plotter_Cause4copula_GOF(fcode=3,method = "kendall",num_keep_last=500,BS=100)
+Plotter_Cause4copula_GOF(fcode=13,method="kendall",num_keep_last=500,BS=100)
+Plotter_Cause4copula_GOF(fcode=4,method = "kendall",num_keep_last=500,BS=100)
+Plotter_Cause4copula_GOF(fcode=14,method="kendall",num_keep_last=500,BS=100)
 
@@
 #-------------------------------------------------------------
-#pdf("./Results/Cause4copula_results/Clayton_spearman.pdf")
-#Plotter_Cause4copula_stat(numsim=500,fcode=3,method="spearman",lb=0,ub=0.1)
-#dev.off()
-
-#pdf("./Results/Cause4copula_results/Clayton_kendall.pdf")
-#Plotter_Cause4copula_stat(numsim=500,fcode=3,method="spearman",lb=0,ub=0.1)
-#dev.off()
-
-#pdf("./Results/Cause4copula_results/Gumbel_spearman.pdf")
-#Plotter_Cause4copula_stat(numsim=500,fcode=4,method="spearman",lb=0,ub=0.1)
-#dev.off()
-
-#pdf("./Results/Cause4copula_results/Gumbel_kendall.pdf")
-#Plotter_Cause4copula_stat(numsim=500,fcode=4,method="kendall",lb=0,ub=0.1)
-#dev.off()
-
-#pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
-#Plotter_Cause4copula_stat(numsim=500,fcode=13,method="spearman",lb=0,ub=0.1)
-#dev.off()
-
-#pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
-#Plotter_Cause4copula_stat(numsim=500,fcode=13,method="kendall",lb=0,ub=0.1)
-#dev.off()
-
-#pdf("./Results/Cause4copula_results/SGumbel_spearman.pdf")
-#Plotter_Cause4copula_stat(numsim=500,fcode=14,method="spearman",lb=0,ub=0.1)
-#dev.off()
-
-#pdf("./Results/Cause4copula_results/SGumbel_kendall.pdf")
-#Plotter_Cause4copula_stat(numsim=500,fcode=14,method="kendall",lb=0,ub=0.1)
-#dev.off()
+pdf("./Results/Cause4copula_results/Clayton_spearman.pdf")
+Plotter_Cause4copula_stat(numsim=50,fcode=3,method="spearman",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/Clayton_kendall.pdf")
+Plotter_Cause4copula_stat(numsim=50,fcode=3,method="spearman",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/Gumbel_spearman.pdf")
+Plotter_Cause4copula_stat(numsim=50,fcode=4,method="spearman",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/Gumbel_kendall.pdf")
+Plotter_Cause4copula_stat(numsim=50,fcode=4,method="kendall",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
+Plotter_Cause4copula_stat(numsim=50,fcode=13,method="spearman",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
+Plotter_Cause4copula_stat(numsim=50,fcode=13,method="kendall",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/SGumbel_spearman.pdf")
+Plotter_Cause4copula_stat(numsim=50,fcode=14,method="spearman",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/SGumbel_kendall.pdf")
+Plotter_Cause4copula_stat(numsim=50,fcode=14,method="kendall",lb=0,ub=0.1)
+dev.off()
 ```

### 2018-04-17 Cause4copula.R

@@ -110,43 +110,57 @@ MCI<-function(x){
 #                     for each of s$noise_c, s$noise_q and s2$pop_c, s2$pop_q matrix : (default 500)
 comp<-function(s,s2,num_keep_last=500){
   
-  s$noise_c<-tail(s$noise_c,num_keep_last)
-  s$noise_q<-tail(s$noise_q,num_keep_last)
-  s2$pop_c<-tail(s2$pop_c,num_keep_last)
-  s2$pop_q<-tail(s2$pop_q,num_keep_last)
+  mod_s_noise_c<-tail(s$noise_c,num_keep_last)
+  mod_s_noise_q<-tail(s$noise_q,num_keep_last)
+  mod_s2_pop_c<-tail(s2$pop_c,num_keep_last)
+  mod_s2_pop_q<-tail(s2$pop_q,num_keep_last)
   
   comp<-matrix(NA,nrow=3,ncol=2)
   rownames(comp)<-c("spearman","kendall","pearson")
   colnames(comp)<-c("cor_noise","cor_pop")
   
   # Now compare between spearman corln btw noise_c and pop_c
-  comp[1,1]<-cor(s$noise_c[,1],s$noise_c[,2],method = "spearman")
-  comp[1,2]<-cor(s2$pop_c[,1],s2$pop_c[,2],method = "spearman")
+  comp[1,1]<-cor(mod_s_noise_c[,1],mod_s_noise_c[,2],method = "spearman")
+  comp[1,2]<-cor(mod_s2_pop_c[,1],mod_s2_pop_c[,2],method = "spearman")
   
   # Now compare between kendall corln btw noise_c and pop_c
-  comp[2,1]<-cor(s$noise_c[,1],s$noise_c[,2],method = "kendall")
-  comp[2,2]<-cor(s2$pop_c[,1],s2$pop_c[,2],method = "kendall")
+  comp[2,1]<-cor(mod_s_noise_c[,1],mod_s_noise_c[,2],method = "kendall")
+  comp[2,2]<-cor(mod_s2_pop_c[,1],mod_s2_pop_c[,2],method = "kendall")
   
   # Now compare between pearson corln btw noise_q and pop_q
-  comp[3,1]<-cor(s$noise_q[,1],s$noise_q[,2],method = "pearson")
-  comp[3,2]<-cor(s2$pop_q[,1],s2$pop_q[,2],method = "pearson")
[…]

### 2018-04-17 SupportingInformation.Rmd
results for BMR data sets added

@@
   ptxt<-''
-  if (fracval>.5*numsurrog){                        # why 0.5 factor?
+  if (fracval>.5*numsurrog){                        # ?DCR why 0.5 factor?
     ptxt<-paste(">",fracval,sep='')
@@
 
- #atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$","$\\Ps_{0,0.1}-\\Ps_{0.9,1}$","$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
- atdres<-data.frame(Statistic=c("$\\cor_{first.slice}-\\cor_{last.slice}$",
-                                "$\\Ps_{first.slice}-\\Ps_{last.slice}$",
-                                "$\\Dsq_{first.slice}-\\Dsq_{last.slice}$"),
+ atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
+                                "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
+                                "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
+ #atdres<-data.frame(Statistic=c("$\\cor_{first.slice}-\\cor_{last.slice}$",
+ #                              "$\\Ps_{first.slice}-\\Ps_{last.slice}$",
+ #                               "$\\Dsq_{first.slice}-\\Dsq_{last.slice}$"),
                    Kendall=rep("",3),
@@
 ```{r read_plot_BMR_data, echo=FALSE, results="hide"}
-set.seed(seed=101)
+set.seed(seed)
 source("phylo_dat_fn.R")
@@
 pdf("./Results/BMR_results/BMR_vs_lnmass.pdf")
-plot(v_BMR[,1],v_BMR[,2],col="blue",xlab="ln(mass)",ylab="BMR") # copula plot btw ln(Mass) vs. BMR
-dev.off()
-```
+op<-par(mar=c(5,5,5,5))
+plot(v_BMR[,1],v_BMR[,2],col="red",xlab="ln(mass)",ylab="BMR",cex.lab=2,cex.axis=2) # copula plot btw ln(Mass) vs. BMR
+par(op)
+dev.off()
+```
+
+![Copula plot for BMR dataset \label{fig_BMR_cop}](./Results/BMR_results/BMR_vs_lnmass.pdf){ width=30%, height=30% }
 
@@
 saveRDS(BivMS_BMR,file="./Results/BMR_results/BivMS_BMR_lnmass.RDS")
+```
+
+<!--Display model selection results for BMR data in a table-->
+```{r tab_BMRfit, echo=F, results='markup'}
+h<-BivMS_BMR$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+knitr::kable(h,
+             format='pandoc',
+             caption = "Fitting info for BMR dataset. \\label{tab_BMRfit}",
+             booktabs=T)
 ```
@@
 
-```{r }
+![Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data.
+As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
+\label{fig_BMR_nonparam}](./Results/BMR_results/BMR_bivfunctionplot.pdf)
+
+```{r tab_BMR_asym, echo=F, results="markup"}
 tab_BMR_asym<-make_table_stat_fn(data=stat_BMR_ga)
-knitr::kable(tab_BMR_asym)
+knitr::kable(tab_BMR_asym,
+             format="pandoc",
+             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)?
+\\label{tab_BMR_asym}",
+             booktabs=T)
 ```
@@
 
+pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
+Plotter_Cause4copula_stat(numsim=50,fcode=13,method="spearman",lb=0,ub=0.1)
+dev.off()
+
+pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
+Plotter_Cause4copula_stat(numsim=50,fcode=13,method="kendall",lb=0,ub=0.1)
+dev.off()
+
 pdf("./Results/Cause4copula_results/Gumbel_spearman.pdf")
@@
 
-pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
-Plotter_Cause4copula_stat(numsim=50,fcode=13,method="spearman",lb=0,ub=0.1)
-dev.off()
-
-pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
-Plotter_Cause4copula_stat(numsim=50,fcode=13,method="kendall",lb=0,ub=0.1)
-dev.off()
-
 pdf("./Results/Cause4copula_results/SGumbel_spearman.pdf")
@@
 ```
+
+![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF.pdf)
+
+![Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_spear}](./Results/Cause4copula_results/Clayton_spearman.pdf)
+
+![Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_kend}](./Results/Cause4copula_results/Clayton_kendall.pdf)
+
+![Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_spear}](./Results/Cause4copula_results/SClayton_spearman.pdf)
+
+![Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}](./Results/Cause4copula_results/SClayton_kendall.pdf)
+
+![Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Gumbel** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_G_spear}](./Results/Cause4copula_results/Gumbel_spearman.pdf)
+
+![Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Gumbel** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_G_kend}](./Results/Cause4copula_results/Gumbel_kendall.pdf)
+
+![Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Gumbel** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SG_spear}](./Results/Cause4copula_results/SGumbel_spearman.pdf)
+
+![Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Gumbel** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SG_kend}](./Results/Cause4copula_results/SGumbel_kendall.pdf)
 

### 2018-04-19 SupportingInformation.Rmd
modified

@@
 
-Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description). The raw data are shown in Figure \ref{fig_soilCNraw}. The copula is shown in Figure \ref{fig_soilCNcop}.
+Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description).
+The raw data are shown in Figure \ref{fig_raw_cop_soilCN}(A).
+The copula is shown in Figure \ref{fig_raw_cop_soilCN}(B).
 
@@
 ```{r load_data_Loecke, echo=F, results='hide'}
-#fig.cap="Plot of the raw soil C and N data.\\label{fig_soilCNraw}", fig.height=3, fig.width=3}
+#fig.cap="Plot of the raw soil C and N data.\\label{fig_raw_cop_soilCN}(A)", fig.height=3, fig.width=3}
 d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
@@
 pdf("./Results/SoilCNraw.pdf")
+op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
 plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1.5,cex.axis=1.5)
+par(op)
 dev.off()
@@
 ```{r rankandplot_Loecke, echo=F, results='hide'}
-#fig.cap="Copula plot of soil C and N data.
-Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations (see Sec.
-\ref{Model_selection}).\\label{fig_soilCNcop}",  fig.height=5, fig.width=5}
 source("getcopula.R")
@@
 
-![Plot of the raw soil C and N data.\label{fig_soilCNraw}](./Results/SoilCNraw.pdf){ width=30%, height=30% }
-
-![Copula plot of soil C and N data.
-Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations (see Sec.
-\ref{Model_selection}).
-\label{fig_soilCNcop}](./Results/SoilCNcop.pdf){ width=30%, height=30% }
+\begin{figure}[!h]
+\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
+\begin{center}
+\includegraphics[width=3 in]{./Results/SoilCNraw.pdf}
+\includegraphics[width=3 in]{./Results/SoilCNcop.pdf}
+\caption{\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data.
+Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see Sec.
+\ref{Model_selection}).\label{fig_raw_cop_soilCN}}
+\end{center}
+\end{figure}
+
+## BMR data  \label{BMR_data}
+
+```{r read_plot_BMR_data, echo=FALSE, results="hide"}
+set.seed(seed)
+source("phylo_dat_fn.R")
+#---------------------------------------------------------
+BMR_dat<-read.csv("./Data/BMR/bmr-contrasts.csv",header = T)
+#class(BMR_dat)
+
+# Generating raw data plot btw ln(mass) and BMR from bmr-contrasts.csv
+pdf("./Results/BMR_results/rawdataplot_BMR_vs_lnmass.pdf")
+op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
+plot(BMR_dat$ln.mass,BMR_dat$BMR,type="p",col="red",xlab="ln(mass)",ylab="BMR",
+     cex.lab=1.5,cex.axis=1.5)
+par(op)
+dev.off()
+#--------------------------------------------------------------------------------------------
+# Generating raw data plot btw ln(mass) and BMR from bmr-contrasts.csv
+pdf("./Results/BMR_results/copula_BMR_vs_lnmass.pdf")
+v_BMR<-phylo_dat_fn(data=BMR_dat,i=4,j=3,ploton = T) 
+#plot(v_BMR[,1],v_BMR[,2],type="p",col="red",xlab="ln(mass)",ylab="BMR",
+#     cex.lab=1.5,cex.axis=1.5)
+dev.off()
+```
+
+\begin{figure}[!h]
+\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
+\begin{center}
+\includegraphics[width=3 in]{./Results/BMR_results/rawdataplot_BMR_vs_lnmass.pdf}
+\includegraphics[width=3 in]{./Results/BMR_results/copula_BMR_vs_lnmass.pdf}
+\caption{\textbf{(A)} Plot of the raw BMR vs.
+ln(mass) data \textbf{(B)} Copula plot for BMR dataset.
+Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see Sec.
+\ref{Model_selection}).\label{fig_BMR_raw_cop}}
+\end{center}
+\end{figure}
 
 ## Phylogenetic data  \label{Phylo_data}
+
 ```{r calculation_phylo_data, echo=F, results="hide"}
@@
 phylo_dat<-read.csv("./Data/Anolis/contrasts.csv",header = T)
-
-# This plot I create on temporary basis to see the copula structure between various column of the raw data
-# After I know which variables should I look into, I will modify this chunk and add model selection and NPA on that
-pdf("./Results/temp_phylo_plot.pdf",width=20, height=20)
-op<-par(mfrow=c(13,13))
-for(i in c(1:13)){
+#class(phylo_dat) #it's a data.frame
+#----------------------------------------------------------------------------------------------------
+# Generating all copula plots between different pathlength and traits 
+pdf("./Results/Anolis_results/contrasts_csv_results/contrasts_csv_pathlength_traits_cop_plot.pdf",width=20, height = 20)
+op<-par(mfrow=c(4,4))
+i<-1
   for(j in c(1:13)){
-    v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = F)
-    plot(v_phylo[,1],v_phylo[,2],col="red")
+    v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = T)
+    #plot(v_phylo[,1],v_phylo[,2],col="green4")
     #mtext("(i,j)=(",i,", ",j,")")
@@
   } 
-}
 par(op)
 dev.off()
-```
+#------------------------------------------------------------------------------------------------------------
+# Generating all 12 by 12 copula plots between different traits
+pdf("./Results/Anolis_results/contrasts_csv_results/contrasts_csv_btw_traits_cop_plot.pdf",width=20, height = 20)
+op<-par(mfrow=c(12,12))
+for(i in c(2:13)){
+  for(j in c(2:13)){
+    v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = T)
+    mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 3,col="blue")
+  } 
+}
+par(op)
+dev.off()
+#----------------------------------------------------------------------------------------
+# Generating raw data plot btw AVG.humerus and AVG.radius from contrasts.csv
+pdf("./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_humerus_radius.pdf")
+op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
+plot(phylo_dat$AVG.humerus,phylo_dat$AVG.radius,type="p",col="red",xlab="AVG.humerus",ylab="AVG.radius",
+     cex.lab=1.5,cex.axis=1.5)
+par(op)
+dev.off()
+#--------------------------------------------------------------------------------------------
+# Generating copula plot btw AVG.humerus and AVG.radius from contrasts.csv
+pdf("./Results/Anolis_results/contrasts_csv_results/copula_AVG_humerus_radius.pdf")
+v_contrasts_67<-phylo_dat_fn(data=phylo_dat,i=6,j=7,ploton = T)
+dev.off()
+```
+
+\begin{figure}[!h]
+\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
+\begin{center}
+\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_humerus_radius.pdf}
+\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_humerus_radius.pdf}
+\caption{\textbf{(A)} Plot of the raw AVG.radius vs.
+AVG.humerus phylogenetic traits.
+\textbf{(B)} Copula plot of AVG.radius vs.
+AVG.humerus phylogenetic traits.
+Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see Sec.
+\ref{Model_selection}).
+\label{fig_raw_cop_AVG_humerus_radius}}
+\end{center}
+\end{figure}
 
@@
 
-We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_soilCNcop}. 
+We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_raw_cop_soilCN}(B). 
 Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.
@@
 
- atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
-                                "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
-                                "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
- #atdres<-data.frame(Statistic=c("$\\cor_{first.slice}-\\cor_{last.slice}$",
- #                              "$\\Ps_{first.slice}-\\Ps_{last.slice}$",
- #                               "$\\Dsq_{first.slice}-\\Dsq_{last.slice}$"),
+ #atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
+ #                               "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
+ #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
+ atdres<-data.frame(Statistic=c("$\\cor_{first.slice}-\\cor_{last.slice}$",
+                               "$\\Ps_{first.slice}-\\Ps_{last.slice}$",
+                                "$\\Dsq_{first.slice}-\\Dsq_{last.slice}$"),
                    Kendall=rep("",3),
@@
 
+\newpage
+
 ## Results: BMR data \label{Ex_BMR}
-
-```{r read_plot_BMR_data, echo=FALSE, results="hide"}
-set.seed(seed)
-source("phylo_dat_fn.R")
-#---------------------------------------------------------
-BMR_dat<-read.csv("./Data/BMR/bmr-contrasts.csv",header = T)
-#class(BMR_dat)
-v_BMR<-phylo_dat_fn(data=BMR_dat,i=4,j=3,ploton = F)
-
-pdf("./Results/BMR_results/BMR_vs_lnmass.pdf")
-op<-par(mar=c(5,5,5,5))
-plot(v_BMR[,1],v_BMR[,2],col="red",xlab="ln(mass)",ylab="BMR",cex.lab=2,cex.axis=2) # copula plot btw ln(Mass) vs. BMR
-par(op)
-dev.off()
-```
-
-![Copula plot for BMR dataset \label{fig_BMR_cop}](./Results/BMR_results/BMR_vs_lnmass.pdf){ width=30%, height=30% }
 
@@
 
+The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_BMRfit}).
+The table shows that Frank is much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+
 <!--Display model selection results for BMR data in a table-->
@@
 
+The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was $0$ as expected, because the contributions came from Frank and SBB8 which are both symmetric copula.
+<!--$`r BivMS_BMR$relLTdep_AICw`$ and the model-averaged upper-tail dependence statistic was 
+$`r BivMS_BMR$relUTdep_AICw`$.
+These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper),
+$`r BivMS_BMR$relLTdep_AICw-BivMS_BMR$relUTdep_AICw`$, was different from 0, answering the third part of question 1. 
+-->
+We also tested that lowest-AIC copula Frank was a good-enough fit, giving $p$-value 
+$`r BivMS_BMR$GofRes_CvM`$ (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value $`r BivMS_BMR$GofRes_KS`$ in the Kolmogorov-Smirnov-based test of the goodness of fit.
+The next section looks at tail dependence in a non-parametric way.
+
 ### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
@@
 source("bivfunctionplot.R")
-stat_BMR_ga<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR",numbin = 10)
+stat_BMR_ga<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_10",numbin = 10)
+stat_BMR_ga_numbin5<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_5",numbin = 5)
 ```
@@
 As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
-\label{fig_BMR_nonparam}](./Results/BMR_results/BMR_bivfunctionplot.pdf)
+\label{fig_BMR_nonparam}](./Results/BMR_results/BMR_numbin_10_bivfunctionplot.pdf)
+
+![Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data.
+As described in Methods, each statistic was computed on the data using the ranges $0-0.2$, $0.2-0.4$, ..., $0.8-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
+\label{fig_BMR_nonparam5}](./Results/BMR_results/BMR_numbin_5_bivfunctionplot.pdf)
 
@@
 tab_BMR_asym<-make_table_stat_fn(data=stat_BMR_ga)
+tab_BMR_asym_numbin5<-make_table_stat_fn(data=stat_BMR_ga_numbin5)
 knitr::kable(tab_BMR_asym,
              format="pandoc",
-             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)?
-\\label{tab_BMR_asym}",
+             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)?, numbin=10 \\label{tab_BMR_asym}",
+             booktabs=T)
+
+knitr::kable(tab_BMR_asym_numbin5,
+             format="pandoc",
+             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)?, numbin=5 \\label{tab_BMR_asym_numbin5}",
+             booktabs=T)
+```
+
+\newpage
+
+## Results: contrasts.csv data \label{Ex_contrasts}
+
+### Model selection approach \label{Ex_contrasts_Model_selection}
+
+```{r bivMS_contrasts_67, echo=F, cache=T, cache.extra=list(seed,v_contrasts_67,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+set.seed(seed)
+# Now do Model Selection for [i,j]=[6,7]
+# 6 : AVG. humerus
+# 7 : AVG. radius
+source("OurBiCopSelect.R")
+families<-c(1,3:10,13:14,16:20)
+BivMS_contrasts_67<-OurBiCopSelect(u1=v_contrasts_67[,1],u2=v_contrasts_67[,2],families=families,
+                          level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
+                          gofnormal=F,status=F)
+saveRDS(BivMS_contrasts_67,file="./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
+```
+
+The test of indepedence gave $p=$ `r BivMS_contrasts_67$IndepTestRes` (to within the precision available 
+from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_contrasts_67fit}).
+The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+
+<!--Display model selection results for contrasts data in a table-->
+```{r tab_contrasts_67fit, echo=F, results='markup'}
+h<-BivMS_contrasts_67$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+knitr::kable(h,
+             format='pandoc',
+             caption = "Fitting info for contrasts between AVG.humerus and AVG.radius dataset. \\label{tab_contrasts_67fit}",
+             booktabs=T)
+```
+
+The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`.
+These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.
+
+These results do convincingly show non-Gaussian copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+The next section looks at tail dependence in a non-parametric way.
+
+### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}
+
+```{r stat_contrasts_67_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_contrasts_67,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+source("bivfunctionplot.R")
+set.seed(seed) 
+# Now do NPA for v_contrasts_67
+source("bivfunctionplot.R")
+stat_contrasts_67_ga<-bivfunctionplot(v=v_contrasts_67,resloc="./Results/Anolis_results/contrasts_csv_results/",
+                                      nametag="contrasts_67",numbin = 4)
+```
+
+![Nonparametric tests for tail dependence and other deviations from a normal copula for contrasts_67 data.
+As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
+\label{fig_BMR_nonparam}](./Results/Anolis_results/contrasts_csv_results/contrasts_67_bivfunctionplot.pdf)
+
+```{r tab_contrasts_67_asym, echo=F, results="markup"}
+tab_contrasts_67_asym<-make_table_stat_fn(data=stat_contrasts_67_ga)
+knitr::kable(tab_contrasts_67_asym,
+             format="pandoc",
+             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_contrasts_67_asym}",
              booktabs=T)

### 2018-04-19 getcopula.R

@@ -30,11 +30,13 @@ getcopula<-function(d,rankon=T,ploton=F)
   
   #plot
   if(ploton==T){
+    op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
     plot(v[,1],v[,2],type="p",col="red",xlab="u",ylab="v",
          cex.lab=2,cex.axis=1.5,
          xlim=c(0,1),ylim=c(0,1),asp=1)
     rect(0,0,1,1)
     lines(c(0,1),c(0,1),type='l')
+    par(op)
   }
   
   return(v)

### 2018-04-19 phylo_dat_fn.R

@@ -1,8 +1,12 @@
 source("getcopula.R")
 #--------------------------------------------------------------------------------------
-# This function (mainly written for phylodata with ontrasts)
+# This function (mainly written for phylodata for contrasts : containing path length and different traits)
 # takes a dataframe and remove the 1st column of that df and then calculates stuffs on that 
 # modified dataframe to return output
+
+# Note : to get copula between traits use usual data cleaning but to find copula between pathlength and trait 
+# take absolute value of trait only
+
 # Input :
 #        data : a dataframe 
 #        i, j : column indices of data frame to get bivariate data set
@@ -12,15 +16,27 @@ source("getcopula.R")
 phylo_dat_fn<-function(data,i,j,ploton=T){
   
   data<-data[,2:dim(data)[2]]
-  d<-data[,c(i,j)]
   
-  v_phylo<-getcopula(d=d,rankon = T,ploton = F)
-  
-  if(ploton==T){
-    op<-par(mfrow=c(1,2))
-    plot(d[,1],d[,2],col="blue")
-    getcopula(d=d,rankon = T,ploton = T)
-    par(op)
+  x<-c(i,j)
+  if(1 %in% x){
+    ind<-which(x!=1)
+    data[,x[ind]]<-abs(data[,x[ind]]) # replace by its absolute value if it's comparison btw pathlength & traits
+    d<-data[,c(i,j)] 
+  }else{
+    d<-data[,c(i,j)] 
   }
+
+  v_phylo<-getcopula(d=d,rankon = T,ploton = ploton)
+  
+  #if(ploton==T){
+    #op<-par(mfrow=c(1,2))
+    #plot(d[,1],d[,2],col="blue")
+    #getcopula(d=d,rankon = T,ploton = T)
+    #par(op)
+  #}
   return(v_phylo)
 }
+
+
+
+

### 2018-04-19 SupportingInformation.Rmd
cause4copula section modified

@@
 #class(phylo_dat) #it's a data.frame
+# no NA/missing value in this dataset
 #----------------------------------------------------------------------------------------------------
@@
     v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = T)
-    #plot(v_phylo[,1],v_phylo[,2],col="green4")
-    #mtext("(i,j)=(",i,", ",j,")")
     mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 3,col="blue")
@@
     v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = T)
-    mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 3,col="blue")
+    mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 4,col="blue",line=0.01)
   } 
@@
 #----------------------------------------------------------------------------------------
-# Generating raw data plot btw AVG.humerus and AVG.radius from contrasts.csv
-pdf("./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_humerus_radius.pdf")
+# Generating raw data plot btw AVG.ltoe.IV and AVG.humerus from contrasts.csv
+pdf("./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf")
 op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(phylo_dat$AVG.humerus,phylo_dat$AVG.radius,type="p",col="red",xlab="AVG.humerus",ylab="AVG.radius",
+plot(phylo_dat$AVG.ltoe.IV,phylo_dat$AVG.humerus,type="p",col="red",xlab="AVG.ltoe.IV",ylab="AVG.humerus",
      cex.lab=1.5,cex.axis=1.5)
@@
 #--------------------------------------------------------------------------------------------
-# Generating copula plot btw AVG.humerus and AVG.radius from contrasts.csv
-pdf("./Results/Anolis_results/contrasts_csv_results/copula_AVG_humerus_radius.pdf")
+# Generating copula plot btw AVG.ltoe.IV and AVG.humerus from contrasts.csv
+pdf("./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf")
 v_contrasts_67<-phylo_dat_fn(data=phylo_dat,i=6,j=7,ploton = T)
@@
 \begin{center}
-\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_humerus_radius.pdf}
-\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_humerus_radius.pdf}
-\caption{\textbf{(A)} Plot of the raw AVG.radius vs.
+\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
+\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
+\caption{\textbf{(A)} Plot of the raw AVG.ltoe.IV vs.
 AVG.humerus phylogenetic traits.
-\textbf{(B)} Copula plot of AVG.radius vs.
+\textbf{(B)} Copula plot of AVG.ltoe.IV vs.
 AVG.humerus phylogenetic traits.
@@
 \ref{Model_selection}).
-\label{fig_raw_cop_AVG_humerus_radius}}
+\label{fig_raw_cop_AVG_humerus_ltoe}}
 \end{center}
@@
 # Now do Model Selection for [i,j]=[6,7]
-# 6 : AVG. humerus
-# 7 : AVG. radius
+# 6 : AVG. ltoe.IV
+# 7 : AVG. humerus
 source("OurBiCopSelect.R")
@@
              format='pandoc',
-             caption = "Fitting info for contrasts between AVG.humerus and AVG.radius dataset. \\label{tab_contrasts_67fit}",
+             caption = "Fitting info for contrasts between AVG.ltoe.IV and AVG.humerus dataset. \\label{tab_contrasts_67fit}",
              booktabs=T)
@@
 # Now do NPA for v_contrasts_67
-source("bivfunctionplot.R")
 stat_contrasts_67_ga<-bivfunctionplot(v=v_contrasts_67,resloc="./Results/Anolis_results/contrasts_csv_results/",
@@
 
+Q1/2: How are copula structure of ecologically influential climate variables changing as part of climate change?
+We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the pops, but we don’t really know how a change in the environment cop structure would make what kind of change in the pop cop structure.
+How?
+
 ## The Moran effect and copula structure in populations
@@
 <!--Shya to implement the modelling we discussed on 2018 04 12 and make results appear here!-->
-```{r cause4copula, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"))}
+```{r cause4copula, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
 source("Cause4copula.R")
@@
 #---------------------------------------
-pdf("./Results/Cause4copula_results/compare_params_pGOF.pdf", width=12, height=6)
+pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf", width=12, height=6)
 op<-par(mfrow=c(2,4),mar=c(3,3.5,3,3.5), mgp=c(1.5,0.5,0))
 
-Plotter_Cause4copula_GOF(fcode=3,method = "spearman",num_keep_last=500,BS=100)
-Plotter_Cause4copula_GOF(fcode=13,method="spearman",num_keep_last=500,BS=100)
-Plotter_Cause4copula_GOF(fcode=4,method = "spearman",num_keep_last=500,BS=100)
-Plotter_Cause4copula_GOF(fcode=14,method="spearman",num_keep_last=500,BS=100)
-
-Plotter_Cause4copula_GOF(fcode=3,method = "kendall",num_keep_last=500,BS=100)
-Plotter_Cause4copula_GOF(fcode=13,method="kendall",num_keep_last=500,BS=100)
-Plotter_Cause4copula_GOF(fcode=4,method = "kendall",num_keep_last=500,BS=100)
-Plotter_Cause4copula_GOF(fcode=14,method="kendall",num_keep_last=500,BS=100)
+Plotter_Cause4copula_GOF(fcode=3,method = "spearman",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=13,method="spearman",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=4,method = "spearman",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=14,method="spearman",num_keep_last=500,BS=1000)
+
+Plotter_Cause4copula_GOF(fcode=3,method = "kendall",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=13,method="kendall",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=4,method = "kendall",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=14,method="kendall",num_keep_last=500,BS=1000)
 
@@
 dev.off()
+#----------------------------
+pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=12, height=6)
+op<-par(mfrow=c(2,4),mar=c(3,3.5,3,3.5), mgp=c(1.5,0.5,0))
+
+Plotter_Cause4copula_GOF(fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=13,method="spearman",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=14,method="spearman",num_keep_last=2500,BS=1000)
+
+Plotter_Cause4copula_GOF(fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=13,method="kendall",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(fcode=14,method="kendall",num_keep_last=2500,BS=1000)
+
+par(op)
+
+op2<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
+plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
+legend("top", c("noise","population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
+       cex = 0.8, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
+       bty = "n") 
+par(op2)
+dev.off()
 
@@
 
-![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF.pdf)
+![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf)
+
+![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **2500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf)
 

### 2018-04-19 Cause4copula.R

@@ -1,6 +1,7 @@
 library(copula)
 library(VineCopula)
 source("CopulaFunctions_flexible.R")
+source("MyBiCopGofTest.R")
 #---------------------------------------------------------------------------------------
 # This function GetNoise generates noises with a copula structure as you specified
 # Input :
@@ -107,7 +108,8 @@ MCI<-function(x){
 #      s : gives you [s$noise_c, s$noise_q : each is a N by 2 noise matrix] , param
 #      s2 : gives you [s2$pop_c, s2$pop_q : each is a N+1 by 2 noise matrix] 
 #     num_keep_last : number of rows you want to keep from bottom 
-#                     for each of s$noise_c, s$noise_q and s2$pop_c, s2$pop_q matrix : (default 500)
+#                     for each of s$noise_c, s$noise_q and s2$pop_c, s2$pop_q matrix : Default=500
+#
 comp<-function(s,s2,num_keep_last=500){
   
   mod_s_noise_c<-tail(s$noise_c,num_keep_last)
@@ -160,7 +162,7 @@ comp<-function(s,s2,num_keep_last=500){
 #  parameter of noise copula vs. correln coef (spearman/kendall)
 # BS : number of bootstrapps used for BiCopGOFTest
 
-Plotter_Cause4copula_GOF<-function(fcode,method,num_keep_last=500,BS){
+Plotter_Cause4copula_GOF<-function(fcode,method,num_keep_last,BS,ploton){
   
   corcoef_list<-seq(from=0.1,to=0.9,by=0.1)
   
@@ -170,6 +172,7 @@ Plotter_Cause4copula_GOF<-function(fcode,method,num_keep_last=500,BS){
   se_par_pop<-c()
   pval_CvM<-c()
   pval_KS<-c()
+  BS_success<-c()
   
   for(corcoef in corcoef_list){
     
@@ -186,15 +189,18 @@ Plotter_Cause4copula_GOF<-function(fcode,method,num_keep_last=500,BS){
     u2<-tail(s2$pop_c[,2],num_keep_last)
     
     z<-BiCopEst(u1,u2,family=fcode,se=T) # apply BiCopEst to get par_pop and se_par_pop for sample estimation
[…]

### 2018-04-19 SupportingInformation.Rmd
plot fixing

@@
 source("Cause4copula.R")
+#-----------------
 set.seed(seed=101)
@@
 
-par(op)
-
-op2<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
+#par(op)
+
+op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
 plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
 legend("top", c("noise","population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
-       cex = 0.8, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
+       cex = 1.5, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
        bty = "n") 
-par(op2)
-dev.off()
+par(op)
+dev.off()
+
 #----------------------------
@@
 
-par(op)
-
-op2<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
+#par(op)
+
+op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
 plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
 legend("top", c("noise","population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
-       cex = 0.8, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
+       cex = 1.5, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
        bty = "n") 
-par(op2)
+par(op)
 dev.off()

### 2018-04-19 Cause4copula.R

@@ -215,7 +215,9 @@ Plotter_Cause4copula_GOF<-function(fcode,method,num_keep_last,BS,ploton){
     op<-par(mar=c(5, 4, 4, 6) + 0.1)
     
     ## Plot first set of data and draw its axis
-    plot(corcoef_list, par_pop, pch=6, axes=FALSE, ylim=c(ceiling(min(0,par_pop-1.96*se_par_pop)),ceiling(max(par_noise,par_pop+1.96*se_par_pop))), xlab="", ylab="", 
+    plot(corcoef_list, par_pop, pch=6, axes=FALSE, 
+         ylim=c(ceiling(min(0,par_pop-1.96*se_par_pop)),ceiling(max(par_noise,par_pop+1.96*se_par_pop))),xlim=c(0,1),
+         xlab="", ylab="", 
          type="b",col="blue")#,main=BiCopName(family = fcode,short=F))
     arrows(corcoef_list,par_pop-1.96*se_par_pop,corcoef_list,par_pop+1.96*se_par_pop,length=0.03, angle=90, code=3, col='blue')
     points(corcoef_list,par_noise,pch=2,col="red",type="b")
@@ -228,7 +230,7 @@ Plotter_Cause4copula_GOF<-function(fcode,method,num_keep_last,BS,ploton){
     par(new=TRUE)
     
     ## Plot the second plot and put axis scale on right
-    plot(corcoef_list, pval_CvM, pch=16,  xlab="", ylab="", ylim=c(0,1), 
+    plot(corcoef_list, pval_CvM, pch=16,  xlab="", ylab="", ylim=c(0,1), xlim=c(0,1),
          axes=FALSE, type="p", col="magenta")
     points(corcoef_list, pval_KS, pch=16, col="green2")
     lines(range(0,1),c(0.05,0.05),type='l',lty='dashed',col='purple')
@@ -604,7 +606,6 @@ Plotter_Cause4copula_stat<-function(numsim=50,fcode,method,lb=0,ub=0.1){
   par(op2)
   
 }
-
 
 
 

### 2018-04-20 SupportingInformation.Rmd
modified

@@
 
-```{r read_plot_BMR_data, echo=FALSE, results="hide"}
+**bmr-contrasts.csv** :
+
+```{r read_BMR_data, echo=FALSE, results="markup"}
 set.seed(seed)
-source("phylo_dat_fn.R")
 #---------------------------------------------------------
@@
 #class(BMR_dat)
+knitr::kable(head(BMR_dat,3),format='pandoc',
+             caption = "First few rows of raw BMR data.\\label{tab_rawdata_BMR}",
+             booktabs=T) # display first few rows
+```
+
+This data contains BMR and ln(mass) for total `r nrow(BMR_dat)` contrasts.
+Table \ref{tab_rawdata_BMR} displays the raw data for first few contrasts.
+The raw data are shown in Figure \ref{fig_raw_cop_BMR}(A).
+The copula is shown in Figure \ref{fig_raw_cop_BMR}(B).
+
+```{r plot_BMR_data, echo=FALSE, results="hide"}
+set.seed(seed)
+source("phylo_dat_fn.R")
 
@@
 Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see Sec.
-\ref{Model_selection}).\label{fig_BMR_raw_cop}}
+\ref{Model_selection}).\label{fig_raw_cop_BMR}}
 \end{center}
@@
 
-## Phylogenetic data  \label{Phylo_data}
-
-```{r calculation_phylo_data, echo=F, results="hide"}
-source("phylo_dat_fn.R")
+## Phylogenetic data : Anolis  \label{Phylo_data}
+
+**contrasts.csv** :
+
+```{r read_phylo_data, echo=F, results="markup"}
+set.seed(seed)
 phylo_dat<-read.csv("./Data/Anolis/contrasts.csv",header = T)
@@
 # no NA/missing value in this dataset
+tab_phylo<-phylo_dat[,c(1,7,8)] 
+knitr::kable(head(tab_phylo,3),format='pandoc',
+             caption = "First few rows of raw phylo data.\\label{tab_rawdata_phylo}",
+             booktabs=T) # display first few rows
+```
+
+This data contains path length and different traits for total `r nrow(phylo_dat)` contrasts.
+Table \ref{tab_rawdata_phylo} displays the raw data for first few contrasts.
+The raw data are shown in Figure \ref{fig_raw_cop_AVG_humerus_ltoe}(A).
+The copula is shown in Figure \ref{fig_raw_cop_AVG_humerus_ltoe}(B).
+
+```{r plot_phylo_data, echo=F, results="hide"}
+set.seed(seed)
+source("phylo_dat_fn.R")
 #----------------------------------------------------------------------------------------------------
@@
 ```
+<!--****DAN: DAN suspects a bug in the code that generates this fig.
+because the red dot should appear outside the x's iff the text > 999 (or so,mething like that) appears in the bin, also check make_table_stat_fn r chunk-->
 
@@
                                "$\\Ps_{first.slice}-\\Ps_{last.slice}$",
-                                "$\\Dsq_{first.slice}-\\Dsq_{last.slice}$"),
+                                "$\\Dsq_{last.slice}-\\Dsq_{first.slice}$"),
                    Kendall=rep("",3),
@@
 <!--Shya to implement the modelling we discussed on 2018 04 12 and make results appear here!-->
-```{r cause4copula, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+```{r cause4copula_pGOF_500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
 source("Cause4copula.R")
@@
 dev.off()
-
+```
+
+```{r cause4copula_pGOF_2500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed=101)
+#---------------------------------------
+# I need to fix that SG_kendall ylimit error
 #----------------------------
-pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=12, height=6)
-op<-par(mfrow=c(2,4),mar=c(3,3.5,3,3.5), mgp=c(1.5,0.5,0))
-
-Plotter_Cause4copula_GOF(fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=13,method="spearman",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=14,method="spearman",num_keep_last=2500,BS=1000)
-
-Plotter_Cause4copula_GOF(fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=13,method="kendall",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=14,method="kendall",num_keep_last=2500,BS=1000)
-
+#pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=12, height=6)
+#op<-par(mfrow=c(2,4),mar=c(3,3.5,3,3.5), mgp=c(1.5,0.5,0))
+
+#Plotter_Cause4copula_GOF(fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(fcode=13,method="spearman",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(fcode=14,method="spearman",num_keep_last=2500,BS=1000)
+
+#Plotter_Cause4copula_GOF(fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(fcode=13,method="kendall",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(fcode=14,method="kendall",num_keep_last=2500,BS=1000)
+
+#op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
+#plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
+#legend("top", c("noise","population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
+#       cex = 1.5, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
+#       bty = "n") 
 #par(op)
-
-op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
-plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-legend("top", c("noise","population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
-       cex = 1.5, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
-       bty = "n") 
-par(op)
-dev.off()
-
-#-------------------------------------------------------------
+#dev.off()
+```
+
+cause4copula_stat_C_spear chunk: `r Sys.time()`
+
+```{r cause4copula_stat_C_spear, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed=101)
+#---------------------------------------
 pdf("./Results/Cause4copula_results/Clayton_spearman.pdf")
@@
 dev.off()
-
+```
+
+cause4copula_stat_C_kend chunk: `r Sys.time()`
+
+```{r cause4copula_stat_C_kend, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed=101)
+#---------------------------------------
 pdf("./Results/Cause4copula_results/Clayton_kendall.pdf")
@@
 dev.off()
-
+```
+
+cause4copula_stat_SC_spear chunk: `r Sys.time()`
+
+```{r cause4copula_stat_SC_spear, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed=101)
+#---------------------------------------
 pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
@@
 dev.off()
-
+```
+
+cause4copula_stat_SC_kend chunk: `r Sys.time()`
+
+```{r cause4copula_stat_SC_kend, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed=101)
+#----------------------
 pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
@@
 dev.off()
-
+```
+
+cause4copula_stat_G_spear chunk: `r Sys.time()`
+
+```{r cause4copula_stat_G_spear, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed=101)
+#----------------------
 pdf("./Results/Cause4copula_results/Gumbel_spearman.pdf")
@@
 dev.off()
-
+```
+
+cause4copula_stat_G_kend chunk: `r Sys.time()`
+
+```{r cause4copula_stat_G_kend, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed=101)
+#----------------------
 pdf("./Results/Cause4copula_results/Gumbel_kendall.pdf")
@@
 dev.off()
-
+```
+
+cause4copula_stat_SG_spear chunk: `r Sys.time()`
+
+```{r cause4copula_stat_SG_spear, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed=101)
+#----------------------
 pdf("./Results/Cause4copula_results/SGumbel_spearman.pdf")
@@
 dev.off()
-
+```
+
+cause4copula_stat_SG_kend chunk: `r Sys.time()`
+
+```{r cause4copula_stat_SG_kend, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed=101)
+#----------------------
 pdf("./Results/Cause4copula_results/SGumbel_kendall.pdf")

### 2018-04-20 SupportingInformation.Rmd
ylim NA plot fixed

@@
 
-Plotter_Cause4copula_GOF(fcode=3,method = "spearman",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=13,method="spearman",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=4,method = "spearman",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=14,method="spearman",num_keep_last=500,BS=1000)
-
-Plotter_Cause4copula_GOF(fcode=3,method = "kendall",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=13,method="kendall",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=4,method = "kendall",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(fcode=14,method="kendall",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=3,method = "spearman",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=13,method="spearman",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=4,method = "spearman",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=14,method="spearman",num_keep_last=500,BS=1000)
+
+Plotter_Cause4copula_GOF(N=10000,fcode=3,method = "kendall",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=13,method="kendall",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=4,method = "kendall",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=14,method="kendall",num_keep_last=500,BS=1000)
 
@@
 #---------------------------------------
-# I need to fix that SG_kendall ylimit error
+# I need to fix that SG_kendall ylimit error : I think I fixed it
 #----------------------------
-#pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=12, height=6)
-#op<-par(mfrow=c(2,4),mar=c(3,3.5,3,3.5), mgp=c(1.5,0.5,0))
-
-#Plotter_Cause4copula_GOF(fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(fcode=13,method="spearman",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(fcode=14,method="spearman",num_keep_last=2500,BS=1000)
-
-#Plotter_Cause4copula_GOF(fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(fcode=13,method="kendall",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(fcode=14,method="kendall",num_keep_last=2500,BS=1000)
-
-#op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
-#plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-#legend("top", c("noise","population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
-#       cex = 1.5, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
-#       bty = "n") 
-#par(op)
-#dev.off()
+pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=12, height=6)
+op<-par(mfrow=c(2,4),mar=c(3,3.5,3,3.5), mgp=c(1.5,0.5,0))
+
+Plotter_Cause4copula_GOF(N=10000,fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=13,method="spearman",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=14,method="spearman",num_keep_last=2500,BS=1000)
+
+Plotter_Cause4copula_GOF(N=10000,fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=13,method="kendall",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=10000,fcode=14,method="kendall",num_keep_last=2500,BS=1000)
+
+op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
+plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
+legend("top", c("noise","population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
+       cex = 1.5, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
+       bty = "n") 
+par(op)
+dev.off()
 ```
@@
 pdf("./Results/Cause4copula_results/Clayton_spearman.pdf")
-Plotter_Cause4copula_stat(numsim=50,fcode=3,method="spearman",lb=0,ub=0.1)
+Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/Clayton_kendall.pdf")
-Plotter_Cause4copula_stat(numsim=50,fcode=3,method="spearman",lb=0,ub=0.1)
+Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
-Plotter_Cause4copula_stat(numsim=50,fcode=13,method="spearman",lb=0,ub=0.1)
+Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=13,method="spearman",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
-Plotter_Cause4copula_stat(numsim=50,fcode=13,method="kendall",lb=0,ub=0.1)
+Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=13,method="kendall",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/Gumbel_spearman.pdf")
-Plotter_Cause4copula_stat(numsim=50,fcode=4,method="spearman",lb=0,ub=0.1)
+Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=4,method="spearman",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/Gumbel_kendall.pdf")
-Plotter_Cause4copula_stat(numsim=50,fcode=4,method="kendall",lb=0,ub=0.1)
+Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=4,method="kendall",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/SGumbel_spearman.pdf")
-Plotter_Cause4copula_stat(numsim=50,fcode=14,method="spearman",lb=0,ub=0.1)
+Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=14,method="spearman",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/SGumbel_kendall.pdf")
-Plotter_Cause4copula_stat(numsim=50,fcode=14,method="kendall",lb=0,ub=0.1)
-dev.off()
-```
-
-![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf)
-
-![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **2500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf)
+Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=14,method="kendall",lb=0,ub=0.1,num_keep_last=500)
+dev.off()
+```
+
+![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 10000 points are generated but we kept last **500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf)
+
+![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 10000 points are generated but we kept last **2500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_spear}](./Results/Cause4copula_results/Clayton_spearman.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_spear}](./Results/Cause4copula_results/Clayton_spearman.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_kend}](./Results/Cause4copula_results/Clayton_kendall.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_kend}](./Results/Cause4copula_results/Clayton_kendall.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_spear}](./Results/Cause4copula_results/SClayton_spearman.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_spear}](./Results/Cause4copula_results/SClayton_spearman.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}](./Results/Cause4copula_results/SClayton_kendall.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}](./Results/Cause4copula_results/SClayton_kendall.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Gumbel** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_G_spear}](./Results/Cause4copula_results/Gumbel_spearman.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Gumbel** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_G_spear}](./Results/Cause4copula_results/Gumbel_spearman.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Gumbel** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_G_kend}](./Results/Cause4copula_results/Gumbel_kendall.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Gumbel** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_G_kend}](./Results/Cause4copula_results/Gumbel_kendall.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Gumbel** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SG_spear}](./Results/Cause4copula_results/SGumbel_spearman.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Gumbel** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SG_spear}](./Results/Cause4copula_results/SGumbel_spearman.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Gumbel** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SG_kend}](./Results/Cause4copula_results/SGumbel_kendall.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Gumbel** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SG_kend}](./Results/Cause4copula_results/SGumbel_kendall.pdf)
 

### 2018-04-20 Cause4copula.R

@@ -107,10 +107,10 @@ MCI<-function(x){
 # Input:
 #      s : gives you [s$noise_c, s$noise_q : each is a N by 2 noise matrix] , param
 #      s2 : gives you [s2$pop_c, s2$pop_q : each is a N+1 by 2 noise matrix] 
-#     num_keep_last : number of rows you want to keep from bottom 
-#                     for each of s$noise_c, s$noise_q and s2$pop_c, s2$pop_q matrix : Default=500
+#     num_keep_last : an integer : number of rows you want to keep from bottom 
+#                     for each of s$noise_c, s$noise_q and s2$pop_c, s2$pop_q matrix 
 #
-comp<-function(s,s2,num_keep_last=500){
+comp<-function(s,s2,num_keep_last){
   
   mod_s_noise_c<-tail(s$noise_c,num_keep_last)
   mod_s_noise_q<-tail(s$noise_q,num_keep_last)
@@ -162,7 +162,7 @@ comp<-function(s,s2,num_keep_last=500){
 #  parameter of noise copula vs. correln coef (spearman/kendall)
 # BS : number of bootstrapps used for BiCopGOFTest
 
-Plotter_Cause4copula_GOF<-function(fcode,method,num_keep_last,BS,ploton){
+Plotter_Cause4copula_GOF<-function(N,fcode,method,num_keep_last,BS){
   
   corcoef_list<-seq(from=0.1,to=0.9,by=0.1)
   
@@ -174,12 +174,13 @@ Plotter_Cause4copula_GOF<-function(fcode,method,num_keep_last,BS,ploton){
   pval_KS<-c()
   BS_success<-c()
   
+ 
   for(corcoef in corcoef_list){
     
     #cat("corcoef=",corcoef,"\n")
     
     # generate noise copula 
-    s<-GetNoise(N=5000,fcode=fcode,corcoef=corcoef,method=method,ploton=F)
+    s<-GetNoise(N=N,fcode=fcode,corcoef=corcoef,method=method,ploton=F)
     # generate pop_copula
     s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q)
     
@@ -200,6 +201,7 @@ Plotter_Cause4copula_GOF<-function(fcode,method,num_keep_last,BS,ploton){
     
   }
   
+ 
   BS_success_percentage<-((min(BS_success))/BS)*100
   
[…]

### 2018-04-21 SupportingInformation.Rmd
temporary solution for ylim with se=NA

@@
 
-Plotter_Cause4copula_GOF(N=10000,fcode=3,method = "spearman",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=13,method="spearman",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=4,method = "spearman",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=14,method="spearman",num_keep_last=500,BS=1000)
-
-Plotter_Cause4copula_GOF(N=10000,fcode=3,method = "kendall",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=13,method="kendall",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=4,method = "kendall",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=14,method="kendall",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=500,BS=1000)
+
+Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=500,BS=1000)
 
@@
 
-Plotter_Cause4copula_GOF(N=10000,fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=13,method="spearman",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=14,method="spearman",num_keep_last=2500,BS=1000)
-
-Plotter_Cause4copula_GOF(N=10000,fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=13,method="kendall",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=10000,fcode=14,method="kendall",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=2500,BS=1000)
+
+Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
+Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=2500,BS=1000)
 
@@
 pdf("./Results/Cause4copula_results/Clayton_spearman.pdf")
-Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500)
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/Clayton_kendall.pdf")
-Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500)
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
-Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=13,method="spearman",lb=0,ub=0.1,num_keep_last=500)
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="spearman",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
-Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=13,method="kendall",lb=0,ub=0.1,num_keep_last=500)
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="kendall",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/Gumbel_spearman.pdf")
-Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=4,method="spearman",lb=0,ub=0.1,num_keep_last=500)
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=4,method="spearman",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/Gumbel_kendall.pdf")
-Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=4,method="kendall",lb=0,ub=0.1,num_keep_last=500)
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=4,method="kendall",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/SGumbel_spearman.pdf")
-Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=14,method="spearman",lb=0,ub=0.1,num_keep_last=500)
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=14,method="spearman",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 pdf("./Results/Cause4copula_results/SGumbel_kendall.pdf")
-Plotter_Cause4copula_stat(N=10000,numsim=50,fcode=14,method="kendall",lb=0,ub=0.1,num_keep_last=500)
-dev.off()
-```
-
-![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 10000 points are generated but we kept last **500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf)
-
-![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 10000 points are generated but we kept last **2500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf)
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=14,method="kendall",lb=0,ub=0.1,num_keep_last=500)
+dev.off()
+```
+
+![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf)
+
+![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **2500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf)
 

### 2018-04-21 Cause4copula.R

@@ -215,10 +215,10 @@ Plotter_Cause4copula_GOF<-function(N,fcode,method,num_keep_last,BS){
 #  if(ploton==T){
     ## add extra space to right margin of plot within frame
     op<-par(mar=c(5, 4, 4, 6) + 0.1)
-    
+    se_par_pop_lim<-max(se_par_pop,na.rm=T) # to remove NA from ylim
     ## Plot first set of data and draw its axis
     plot(corcoef_list, par_pop, pch=6, axes=FALSE, 
-         ylim=c(ceiling(min(0,par_pop-1.96*se_par_pop)),ceiling(max(par_noise,par_pop+1.96*se_par_pop))),xlim=c(0,1),
+         ylim=c(ceiling(min(0,par_pop-1.96*se_par_pop_lim)),ceiling(max(par_noise,par_pop+1.96*se_par_pop_lim))),xlim=c(0,1),
          xlab="", ylab="", 
          type="b",col="blue")#,main=BiCopName(family = fcode,short=F))
     arrows(corcoef_list,par_pop-1.96*se_par_pop,corcoef_list,par_pop+1.96*se_par_pop,length=0.03, angle=90, code=3, col='blue')

### 2018-04-23 SupportingInformation.Rmd
error bar plot fixing

@@
 #---------------------------------------
-pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf", width=12, height=6)
-op<-par(mfrow=c(2,4),mar=c(3,3.5,3,3.5), mgp=c(1.5,0.5,0))
+pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf", width=8, height=6)
+op<-par(mfrow=c(2,2))
 
@@
 Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=500,BS=1000)
 
@@
 Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=500,BS=1000)
 
@@
 
-op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
+op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
 plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-legend("top", c("noise","population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
-       cex = 1.5, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
+legend("top", c("noise", "population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
+       cex = 1.5, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
        bty = "n") 
@@
 #----------------------------
-pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=12, height=6)
-op<-par(mfrow=c(2,4),mar=c(3,3.5,3,3.5), mgp=c(1.5,0.5,0))
+pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=8, height=6)
+op<-par(mfrow=c(2,2))
 
@@
 Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=2500,BS=1000)
 
@@
 Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=2500,BS=1000)
 
@@
 plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-legend("top", c("noise","population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
-       cex = 1.5, pch = c(2, 6, 16, 16), lty = c(1, 1, NA, NA), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
+legend("top", c("noise", "population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
+       cex = 1.5, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
        bty = "n") 

### 2018-04-23 Cause4copula.R

@@ -94,6 +94,13 @@ Simulator_Cause4copula<-function(cons,p0=c(0,0),noise){
   
 } 
 #-----------------------------------------------------------------------------------------------
+# The following lines are examples to draw error bar without calling arrows
+#plot(x=c(1:5),y=c(1:5),ylim=c(0,6))
+#se<-c(1,0.5,NA,0,0.8)
+#segments(x,y-se,x,y+se)
+#epsilon <- 0.02
+#segments(x-epsilon,y-se,x+epsilon,y-se)
+#segments(x-epsilon,y+se,x+epsilon,y+se)
 #----------------------------------------------------------
 # This function gives mean,lowCI,upCI of a vector
 #
@@ -212,17 +219,21 @@ Plotter_Cause4copula_GOF<-function(N,fcode,method,num_keep_last,BS){
     warning("specify method",immediate.=T,call.=T)
   }
   
-#  if(ploton==T){
     ## add extra space to right margin of plot within frame
-    op<-par(mar=c(5, 4, 4, 6) + 0.1)
+    op<-par(mar=c(3.5, 4, 4, 6) + 0.1)
     se_par_pop_lim<-max(se_par_pop,na.rm=T) # to remove NA from ylim
     ## Plot first set of data and draw its axis
     plot(corcoef_list, par_pop, pch=6, axes=FALSE, 
-         ylim=c(ceiling(min(0,par_pop-1.96*se_par_pop_lim)),ceiling(max(par_noise,par_pop+1.96*se_par_pop_lim))),xlim=c(0,1),
+         #ylim=c(ceiling(min(0,par_pop-1.96*se_par_pop_lim)),
+         ylim=c(0,ceiling(max(par_noise,par_pop+1.96*se_par_pop_lim))),xlim=c(0,1),
          xlab="", ylab="", 
-         type="b",col="blue")#,main=BiCopName(family = fcode,short=F))
-    arrows(corcoef_list,par_pop-1.96*se_par_pop,corcoef_list,par_pop+1.96*se_par_pop,length=0.03, angle=90, code=3, col='blue')
-    points(corcoef_list,par_noise,pch=2,col="red",type="b")
[…]

### 2018-04-24 SupportingInformation.Rmd
minor change:typo fixed

@@
 
-cause4copula_stat_C_spear chunk: `r Sys.time()`
-
-```{r cause4copula_stat_C_spear, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+```{r cause4copula_stat, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
 source("Cause4copula.R")
@@
 dev.off()
-```
-
-cause4copula_stat_C_kend chunk: `r Sys.time()`
-
-```{r cause4copula_stat_C_kend, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed=101)
-#---------------------------------------
+
 pdf("./Results/Cause4copula_results/Clayton_kendall.pdf")
@@
 dev.off()
-```
-
-cause4copula_stat_SC_spear chunk: `r Sys.time()`
-
-```{r cause4copula_stat_SC_spear, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed=101)
-#---------------------------------------
+
 pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
@@
 dev.off()
-```
-
-cause4copula_stat_SC_kend chunk: `r Sys.time()`
-
-```{r cause4copula_stat_SC_kend, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed=101)
-#----------------------
+
 pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
@@
 dev.off()
-```
-
-cause4copula_stat_G_spear chunk: `r Sys.time()`
-
-```{r cause4copula_stat_G_spear, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed=101)
-#----------------------
+
 pdf("./Results/Cause4copula_results/Gumbel_spearman.pdf")
@@
 dev.off()
-```
-
-cause4copula_stat_G_kend chunk: `r Sys.time()`
-
-```{r cause4copula_stat_G_kend, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed=101)
-#----------------------
+
 pdf("./Results/Cause4copula_results/Gumbel_kendall.pdf")
@@
 dev.off()
-```
-
-cause4copula_stat_SG_spear chunk: `r Sys.time()`
-
-```{r cause4copula_stat_SG_spear, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed=101)
-#----------------------
+
 pdf("./Results/Cause4copula_results/SGumbel_spearman.pdf")
@@
 dev.off()
-```
-
-cause4copula_stat_SG_kend chunk: `r Sys.time()`
-
-```{r cause4copula_stat_SG_kend, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed=101)
-#----------------------
+
 pdf("./Results/Cause4copula_results/SGumbel_kendall.pdf")
@@
 Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}](./Results/Cause4copula_results/SClayton_kendall.pdf)
-
-![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Gumbel** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_G_spear}](./Results/Cause4copula_results/Gumbel_spearman.pdf)
-
-![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Gumbel** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_G_kend}](./Results/Cause4copula_results/Gumbel_kendall.pdf)
-
-![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Gumbel** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SG_spear}](./Results/Cause4copula_results/SGumbel_spearman.pdf)
-
-![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Gumbel** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SG_kend}](./Results/Cause4copula_results/SGumbel_kendall.pdf)
 

### 2018-04-24 Cause4copula.R

@@ -472,13 +472,13 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
   plot(corcoef_list,S_noise_mat[,2],cex=0.5,col="red",xlab=xlabel,ylab="Spearman",xlim=c(0,1),ylim=c(0,1))
   segments(corcoef_list,S_noise_mat[,1],corcoef_list,S_noise_mat[,3],col='red')
   bar_len<-0.02
-  segments(corcoef_list-bar_len,S_noise_mat[,1],corcoef_list+bar_len,S_noise_mat[,3],col='red')
-  segments(corcoef_list-bar_len,S_noise_mat[,1],corcoef_list+bar_len,S_noise_mat[,3],col='red')
+  segments(corcoef_list-bar_len,S_noise_mat[,1],corcoef_list+bar_len,S_noise_mat[,1],col='red')
+  segments(corcoef_list-bar_len,S_noise_mat[,3],corcoef_list+bar_len,S_noise_mat[,3],col='red')
   #arrows(corcoef_list,S_noise_mat[,1],corcoef_list,S_noise_mat[,3],length=0.03, angle=90, code=3, col='red')
   points(corcoef_list,S_pop_mat[,2],cex=0.5,col="blue")
   segments(corcoef_list,S_pop_mat[,1],corcoef_list,S_pop_mat[,3],col='blue')
-  segments(corcoef_list-bar_len,S_pop_mat[,1],corcoef_list+bar_len,S_pop_mat[,3],col='blue')
-  segments(corcoef_list-bar_len,S_pop_mat[,1],corcoef_list+bar_len,S_pop_mat[,3],col='blue')
+  segments(corcoef_list-bar_len,S_pop_mat[,1],corcoef_list+bar_len,S_pop_mat[,1],col='blue')
+  segments(corcoef_list-bar_len,S_pop_mat[,3],corcoef_list+bar_len,S_pop_mat[,3],col='blue')
   #arrows(corcoef_list,S_pop_mat[,1],corcoef_list,S_pop_mat[,3],length=0.03, angle=90, code=3, col='blue')
   points(corcoef_list,pval_S,col="purple")
   lines(range(0,1),c(0.05,0.05),type='l',lty='dashed',col='purple')
[…]

### 2018-04-24 Tests.Rmd
Tests for Cause4copula.R added

@@
 # Tests for retd() function from ExtremeTailDep.R
-<!--DCR to add this section-->
+
+DCR to add this section
 
@@
 ```
+
+# Tests for Cause4copula.R
+
+**Test for `GetNoise` function**
+
+```{r test_GetNoise, echo=T}
+set.seed(seed=101)
+source("Cause4copula.R")
+# Here we generate noise with 5000 points drawn from Clayton copula with Kendall's tau=0.5 
+s<-GetNoise(N=5000,fcode=3,corcoef=0.5,method="kendall",ploton=T) # plot in copula space
+s1<-s$noise_q
+hist(s1[,1],breaks=1000) # This should be normal distribution
+hist(s1[,2],breaks=1000) # This should be normal distribution
+plot(s1[,1],s1[,2],col="red") # plot after qnorm
+
+# Here we also checked C or SC copula 
+# with same spearman's rho/ kendall's tau have same parameter
+tgcop<-claytonCopula(3,2)
+tgcop2<-rotCopula(tgcop)
+iRho(tgcop,0.5)==iRho(tgcop2,0.5) # check if it is True?
+iTau(tgcop,0.5)==iTau(tgcop2,0.5) # check if it is True?
+```
+
+**Test for `Simulator_Cause4copula` and `comp` function**
+
+```{r test_Simulator_Cause4copula, echo=T}
+set.seed(seed=101)
+source("Cause4copula.R")
+# when noise comes from a clayton cop with spearmancor=0.8
+s<-GetNoise(N=5000,fcode=3,corcoef=0.8,method="spearman",ploton=T) # noise copula
+s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q)
+points(s2$pop_c[,1],s2$pop_c[,2],col="green") # population copula
+hist(s2$pop_q[,1],breaks=1000) # check if it's normal
+hist(s2$pop_c[,1],breaks=1000) # check if it's uniform
+num_keep_last<-500
+zz<-comp(s=s,s2=s2,num_keep_last=num_keep_last)
+zz$comp
+x<-c(dim(zz$last_num_keep_noise$last_num_keep_noise_c)[1],
+     dim(zz$last_num_keep_noise$last_num_keep_noise_q)[1],
+     dim(zz$last_num_keep_pop$last_num_keep_pop_c)[1],
+     dim(zz$last_num_keep_pop$last_num_keep_pop_c)[1])
+if(all(x==num_keep_last)){
+  print("passed")
+}else{
+  print("failed")
+}
+```

### 2018-04-25 SupportingInformation.Rmd
adding toc
adding chunk for Model selection and skewness chunk for methane flux data

@@
     fig_caption: yes
+
 tables: True
@@
 ---
+
+\tableofcontents
+\listoftables
+\listoffigures
 
@@
 
-**bmr-contrasts.csv** :
+**bmr-contrasts.csv** : Prof. M. T. Holder to add description here about this dataset.
 
@@
 
-**contrasts.csv** :
+**contrasts.csv** : Prof. M. T. Holder to add description here about this dataset.
 
@@
    * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
-   * We threw out all locations that had fewer than 30 years data 
-   * We use species=10 [Name : "Green spruce aphid"] for aphid-count data 
-   * We use species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data
+   * We threw out all locations that had fewer than 30 years data. 
+   * We use species=10 [Name : "Green spruce aphid"] for aphid-count data. 
+   * We use species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data.
 
@@
    * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
-   * We threw out all locations that had fewer than 45 years data
-   * We use species=16 [Name : "Ceratium furca"] for Plankton_North_Sea data
-   * *Ceratium furca* is a dicoflagellate commonly found in UK seas
-   * note, when we edit this, we cannot call it the North Seas since data come also from other seas around the UK
+   * We threw out all locations that had fewer than 45 years data.
+   * We use species=16 [Name : "Ceratium furca"] for Plankton_North_Sea data.
+   * *Ceratium furca* is a dicoflagellate commonly found in UK seas.
+   * note, when we edit this, we cannot call it the North Seas since data come also from other seas around the UK.
+
+## Methane-flux data  \label{Methane_data}
+   * Methane-flux at many locations (Prof. Terry to augment this description).
+   * We threw out all locations that had fewer than 50 dates data.
+
+```{r read_methane_data, echo=F, results="markup"}
+set.seed(seed)
+methane_raw<-readRDS("./Data/methane_data_for_Shyamolina.rds")
+#class(methane_raw) #it's a data.frame for multivariable data
+#knitr::kable(head(methane_raw,3),format='pandoc',
+#             caption = "First few rows of raw methane data.\\label{tab_rawdata_methane}",
+#             booktabs=T) # display first few rows
+```
 
@@
 <!--***DAN: stopped here 2018 04 07-->
+\newpage
+
+## Results: methane flux data \label{Ex_methane}
+
+### Model selection approach \label{Ex_methane_Model_selection}
+
+<!--Load and format the data-->
+```{r read_format_methane_data, echo=F}
+
+methane_raw<-readRDS("./Data/methane_data_for_Shyamolina.rds")
+
+myls <- vector("list", length = dim(methane_raw)[2]-1)
+for(i in 2:dim(methane_raw)[2]){
+  temp<-methane_raw[,c(1,i)]
+  colnames(temp)<-c("Year","Dat")
+  myls[[i-1]]<-temp
+}
+methane_data<-list(myls=myls)
+```
+
+<!--Do the main analysis-->
+```{r RES_methane, echo=F, cache=T, results="hide", cache.extra=list(seed,methane_data,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+
+source("FittingCopula_selective_loc.R")  # top most hierarchy source file
+
+set.seed(seed)
+#cat(paste("start-time: ",Sys.time(),"\n")) 
+RES_methane<-RES_single_sp(d_allsp=methane_data,sp=1,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=50)
+#cat(paste("stop-time: ",Sys.time(),"\n"))    
+resloc<-"./Results/fitting_results/"
+saveRDS(RES_methane,paste(resloc,file="MethaneFluxFit_selecloc_data_pt_thrs_50.RDS",sep=""))
+```
+
+```{r read_RES_methane,echo=F}
+RES_methane<-readRDS("./Results/fitting_results/MethaneFluxFit_selecloc_data_pt_thrs_50.RDS")
+```
+
+As described in Methods and as in previous sections, model selection among copulas was done separately for 
+each pair of sampling locations. There were $13$ sampling locations, so results are presented in 
+$13 \times 13$ matrix format, with rows and columns corresponding to locations. 
+Table \ref{tab_methane_bestcop} shows that, for the large majority of location pairs, some 
+non-normal copula was a better fit, according to AIC, than was the normal copula, answering the 
+first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula 
+family for each pair of sampling locations (Table \ref{tab_methane_AICw}) to the AIC weight of
+the normal copula family for the same pair (Table \ref{tab_methane_AICw_normal}) 
+confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{tab_methane_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{tab_methane_p_KS}).
+In all cases at least
+`r min((RES_methane$gfc_numBS_success/RES_methane$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+
+```{r tab_methane_bestcop, echo=F, results='markup'}
+knitr::kable(RES_methane$info_ord_copname[,,1],
+             format='pandoc',
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.
+See previous sections for copula abbreviations.
+NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_methane_bestcop}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_methane_AICw,echo=F,results="markup"}
+h<-RES_methane$info_ord_AICw[,,1]
+h[is.infinite(h)]<-NA
+knitr::kable(h,
+             format="pandoc",
+             digits=3,
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for methane-flux data.\\label{tab_methane_AICw}",
+             booktabs=TRUE)
+```
+
+```{r tab_methane_AICw_normal,echo=F,results="markup"}
+source("normal_cop_AICw.R")
+h<-normal_cop_AICw_matrix(r=RES_methane)
+h[is.infinite(h)]<-NA
+knitr::kable(h,
+             format="pandoc",
+             digits=3,
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for methane-flux data.\\label{tab_methane_AICw_normal}",
+             booktabs=TRUE)
+```
+
+```{r tab_methane_p_CvM,echo=F,results="markup"}
+h<-RES_methane$gfc_p_CvM
+h[is.infinite(h)]<-NA
+knitr::kable(h,
+             format="pandoc",
+             digits=3,
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for methane-flux data.
+Tests were based on bootstrapping as described in Methods.\\label{tab_methane_p_CvM}",
+             booktabs=TRUE)
+```
+
+```{r tab_methane_p_KS,echo=F,results="markup"}
+h<-RES_methane$gfc_p_KS
+h[is.infinite(h)]<-NA
+knitr::kable(h,
+             format="pandoc",
+             digits=3,
+             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of
+the goodness of fit of the lowest-AIC copula family for each pair of
+locations for methane-flux data. Tests were based on bootstrapping as described in Methods.\\label{tab_methane_p_KS}",
+             booktabs=TRUE)
+```
+
+The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{tab_methane_LTdep_AICw} and \ref{tab_methane_UTdep_AICw}, respectively.
+These values were exclusively positive, thereby
+differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
+have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
+
+```{r tab_methane_LTdep_AICw,echo=F,results="markup"}
+h<-RES_methane$LTdep_AICw
+h[is.infinite(h)]<-NA
+knitr::kable(h,
+             format="pandoc",
+             digits=3,
+             caption = "Model-averaged lower tail dependence statistic for each location pair for methane-flux data.\\label{tab_methane_LTdep_AICw}",
+             booktabs=TRUE)
+```
+
+```{r tab_methane_UTdep_AICw,echo=F,results="markup"}
+h<-RES_methane$UTdep_AICw
+h[is.infinite(h)]<-NA
+knitr::kable(h,
+             format="pandoc",
+             digits=3,
+             caption = "Model-averaged upper tail dependence statistic for each location pair for methane-flux data.\\label{tab_methane_UTdep_AICw}",
+             booktabs=TRUE)
+```
+
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
+(lower minus upper), shown in Table \ref{tab_methane_LTmUT}, was positive for some
+location pairs and negative for others, but was positive for 
+`r sum((RES_methane$LTdep_AICw-RES_methane$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_methane$gfc_p_CvM)[1]^2-dim(RES_methane$gfc_p_CvM)[1])-2*RES_methane$num_indep_loc_pair`
+locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
+
+```{r tab_methane_LTmUT,echo=F,results="markup"}
+h<-RES_methane$LTdep_AICw-RES_methane$UTdep_AICw
+h[is.nan(h)]<-NA
+knitr::kable(h,
+             format="pandoc",
+             digits=3,
+             caption = "Model averaged lower tail dependence minus model
+averaged upper tail dependence statistic for each location pair for methane-flux data.\\label{tab_methane_LTmUT}",
+             booktabs=TRUE)
+```
 
@@
 ![Skewness_plankton_north_sea for species : Ceratium_furca\label{fig_skewness_pns}](./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf){ width=40%, height=40% }
+
+### Methane-flux data \label{skewness_methane}
+
+Figure \ref{fig_skewness_methane} shows that averaged skewness is lower than expected if the plankton north sea data came from a normal copula.
+Our previous results on non-parametric analysis and model fitting showed lower tail dependence.
+As $p<0.05$, we reject $H_{0}$.
+This result is consistent with the non-parametric analysis and model fitting results which indicated lower tail dependence.
+
+```{r skewness_methane_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+
+source("good_loclist.R")
+source("skewness_multivar_dataset.R")
+set.seed(seed)
+
+pdf("./Results/skewness_results/skewness_methane/skewness_methane.pdf")
+par(mfrow=c(2,1))
+numsurrog<-10000
+loclist<-good_loclist(d_allsp=methane_data,sp=1,data_pt_thrs=50)
+
+answer1k<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
+                           corpres="kendall") 
+p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
+mtext(paste0("(A) result based on kendall preserving normal surrogates, p = ",p_left),side=3,line=0.2,col="navyblue")
+
+answer1s<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
+                           corpres="spearman") 
+p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog
+mtext(paste0("(B) result based on spearman preserving normal surrogates , p = ",p_left),side=3,line=0.2,col="navyblue")
+
+dev.off()
+```
+
+![Skewness_methane-flux data\label{fig_skewness_methane}](./Results/skewness_results/skewness_methane/skewness_methane.pdf){ width=40%, height=40% }
 

### 2018-04-26 SupportingInformation.Rmd
filled in description for Cause4copula part

@@
 ---
-title: "Supporting Information"
+title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
 author: "S. Ghosh, T. Loecke, C.P. Reid, D.C. Reuman"
@@
 set.seed(seed)
-methane_raw<-readRDS("./Data/methane_data_for_Shyamolina.rds")
+methane_raw<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")
 #class(methane_raw) #it's a data.frame for multivariable data
@@
 
-methane_raw<-readRDS("./Data/methane_data_for_Shyamolina.rds")
+methane_raw<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")
 
@@
 
+### Non-parametric analysis of tail dependence  \label{Ex_methane_NPA}
+```{r stat_methane, echo=F, cache=T, cache=T,  cache.extra=list(seed,methane_data,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+
+source("good_loclist.R")
+source("NonParamStat.R")
+set.seed(seed)
+resloc<-"./Results/stat_results/stat_methane/"
+
+data_pt_thrs<-50
+good_loc <- good_loclist(d_allsp=methane_data,sp=1,data_pt_thrs=data_pt_thrs)
+
+latlon<-readRDS("./Data/Methane_data/chamberLocations.rds")
+chamber_name<-colnames(methane_raw[,c(2:dim(methane_raw)[2])])
+ndim<-length(chamber_name)
+
+latlon_arranged<-data.frame(chamber_name=chamber_name,lat=NA*numeric(ndim),lon=NA*numeric(ndim))
+for(i in 1:ndim){
+  if (chamber_name[i] %in% latlon$chamber){
+    ind<-which(latlon$chamber==chamber_name[i])
+    latlon_arranged[i,2:3]<-unname(latlon[ind,2:3])
+  }
+}
+
+stat_methane<-multcall(d_allsp=methane_data,sp=1,lats=latlon_arranged$lat,longs=latlon_arranged$lon,
+                           pfname=paste(resloc,"data_pt_thrs",data_pt_thrs,sep=''),good_loc=good_loc)
+
+saveRDS(stat_methane,paste(resloc,file="stat_methane.RDS",sep='')) 
+```
+
+```{r read_stat_methane,echo=F}
+stat_methane<-readRDS("./Results/stat_results/stat_methane/stat_methane.RDS")
+```
+
+Mean values of our statistics, as well as mean Kendall and Spearman
+correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
+spatial resampling scheme, as described in Methods (Table \ref{tab_methane_resamp}).
+Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
+always excluded $0$ and mean values were positive, indicating tail dependence.
+But the confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ were marginal, indicating no significant asymmetry of tail dependence.
+
+<!--display the main statistical results in a table-->
+```{r tab_methane_resamp,echo=F,results="markup"}
+s<-stat_methane$numericdf[,c(3,2,4)]
+colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
+rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
+knitr::kable(s,
+             format = "pandoc",
+             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for methane-flux data.\\label{tab_methane_resamp}",
+             booktabs=TRUE)
+```
+
+Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
+between the locations (Figure \ref{fig_methane} A, B) showed the typical decline of correlation with increasing
+distance.
+Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Figure \ref{fig_methane} C-F).
+$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
+their opposite meaning (Figure \ref{fig_methane} G-H).
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_methane} I-K) did not show a decline with distance, but did demonstrate visually the
+statistical result of Table \ref{tab_methane_resamp} that averages of these 
+quantities across all pairs of sampling locations were more or less equally dispersed around the zero-line, showing no significant asymmetry of tail dependence.
+
+\begin{figure}[!h]
+\textbf{ \hspace{2.6 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.8 cm} (D)} \\
+\begin{center}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Spearman_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Kendall_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Corl_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Coru_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{2.6 cm} (E) \hspace{3.6 cm} (F) \hspace{3.8 cm} (G) \hspace{3.8 cm} (H)} \\
+\begin{center}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Pl_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Pu_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_D2l_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{4.8 cm} (I) \hspace{3.8 cm} (J) \hspace{4 cm} (K) } \\
+\begin{center}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Corl-Coru_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Pl-Pu_vs_D.pdf}
+\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u-D2l_vs_D.pdf}
+\caption{Non-parametric tail statistics results for \textit{Ceratium furca} abundance data plotted against geographical distance $(D)$ between two sampling locations.
+(A) Spearman correlation vs.
+$D$, (B) Kendall correlation vs.
+$D$, (C) $\cor_l$ vs.
+$D$, (D) $\cor_u$ vs.
+$D$; (E) $\Ps_l$ vs.
+$D$; (F) $\Ps_u$ vs.
+$D$; (G) $\Dsq_u$ vs.
+$D$; (H) $\Dsq_l$ vs.
+$D$; (I) $\cor_l-\cor_u$ vs.
+$D$; (J) $\Ps_l-\Ps_u$ vs.
+$D$; (K) $\Dsq_u-\Dsq_l$ vs.
+$D$ plotted for all pairs of locations.
+\label{fig_methane}}
+\end{center}
+\end{figure}
+
 \newpage
@@
 dependence of populations in the locations that are influenced by the environmental variables.
+
+To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
+\ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.
+
+\begin{equation}\label{eq.AR1}
+P_{i}(t+1)=\beta P_{i}(t)+\alpha \epsilon_{i}(t)
+\end{equation}
+
+To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$.
+We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches.
+Should their dependence be exactly the same as of the noises or not ?
+
+To test that our approach is of two way : first, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
+Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome which was a bivariate population-copula, with last $500$ simulated points for each habitat patch (see Figure \ref{fig_Cause4copula_params_GOF_500}).
+We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size and reported p-values for both statistics (Cramer-von Mises and Kolmogorov-Smirnov).
+p-values greater or equal to $0.05$ (above the dashed line in Figure \ref{fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters.
+Figure \ref{fig_Cause4copula_params_GOF_2500} represents similar figure as of Figure \ref{fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$.
+(For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
+
+Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see Sec.
+\ref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
+Then we plotted mean of each measured quantity with standard error against Spearman's Rho (or Kendall's Tau) of the noise copula family.
+Figure \ref{fig_Cause4copula_C_spear} and Figure \ref{fig_Cause4copula_C_kend} represent this picture for noise generating from a lower tail dependent Clayton copula whereas for Figure \ref{fig_Cause4copula_SC_spear} and Figure \ref{fig_Cause4copula_SC_kend} noise came from a upper tail dependent Survival Clayton family.
+We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach, here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise.
+(For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
 
@@
 pdf("./Results/Cause4copula_results/Clayton_kendall.pdf")
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500)
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="kendall",lb=0,ub=0.1,num_keep_last=500)
 dev.off()
@@
 
-![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf)
-
-![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **2500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf)
+![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF_500}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf)
+
+![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **2500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF_2500}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_spear}](./Results/Cause4copula_results/Clayton_spearman.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_spear}](./Results/Cause4copula_results/Clayton_spearman.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_kend}](./Results/Cause4copula_results/Clayton_kendall.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_kend}](./Results/Cause4copula_results/Clayton_kendall.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_spear}](./Results/Cause4copula_results/SClayton_spearman.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_spear}](./Results/Cause4copula_results/SClayton_spearman.pdf)
 
 ![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 10000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}](./Results/Cause4copula_results/SClayton_kendall.pdf)
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}](./Results/Cause4copula_results/SClayton_kendall.pdf)
 

### 2018-05-02 SupportingInformation.Rmd
this version contains the whole old description in a single file
savingmodel selection result for soilCN

@@
     fig_caption: yes
+    includes:
+      in_header: head_supp.sty
 
@@
 
-#csl: science.csl
-bibliography: REF_SI.bib
+csl: ecology-letters.csl
+bibliography: REF_ALL.bib
 ---
@@
 \listoffigures
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{section}{0}
+\renewcommand{\thesection}{S\arabic{section}}
 
@@
 
-# Question 1 analyses for each dataset
-Recall that question 1 is: Do datasets in ecology and related fields have non-Gausian copula structure?
-Do they show tail dependence distinct from that of a Gaussian copula?
-In particular do they show asymmetric tail dependence?
-
-## Details of methods \label{Method}
-
-### Model selection approach \label{Model_selection}
-
-For a bivariate dataset $(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps, 
-most implemented using tools from the `VineCopula` package. Code combining these tools is in the
-[`OurBiCopSelect`](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
-
-First, we produced ranks 
-$(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
-$\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
-we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; further work was with the $u_i$ 
-and $v_i$. Ranks $\tilde{u}_i$ and $\tilde{v}_i$ were $1$, for the 
-smallest element, up to $n$, for the largest.
-
-Second, we did a test of the independence of the $u_i$ and $v_i$. Subsequent model 
-selection algorithms are ineffective if data could not be distinguished from independent data 
-since most or all bivariate copula families considered include the independent copula in the 
-family or as a boundary case. The independence test was as implemented in the `BiCopIndTest`
-function in `VineCopula`, the documentation of which reports the test uses the asymptotic normality of the statistic $|\tau|\sqrt{\frac{9n(n-1)}{2(2n+5)}}$, where $\tau$ is Kendall's tau statistic.
-If independence could be rejected ($0.05$ significance level), model selection proceeded.
-
-The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`.
-Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulas.
-Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC) values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection].
-Descriptions of these copula families can be found in introductory texts on copulas [@nelsen2007_copula ; @joe2014_dependence]  and in references cited in `VineCopula`.
-The lower- and upper-tail dependence of the
-best-fitting member of each family was also provided by `BiCopEst`.
-
-Model selection methods will give the *relative* support of several models, but will not,
-on their own, indicate whether any of the models selected is an objectively good fit (all 
-models considered might poorly describe the data). To test this, and as the fourth step of
-our procedure, we tested the goodness of fit of our best-AIC copula family using a 
-bootstrapping procedure implemented in `BiCopGofTest` of `VineCopula`. The method performed 
-two tests, one based on a Cramer-von Mises statistic, and one based on a 
-Kolmogorov-Smirnov statistic. For computational efficiency, 
-an initial run using $100$ bootstraps was performed, and if
-the $p$-value resulting from either test was less than $0.2$, tests were re-run with 
-$1000$ boostraps.
-
-Finally, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
-model-averaged lower- and upper-tail dependence values using standard model averaging 
-formulas [@burnham2003_modelselection]; likewise for BIC.
-
-<!--Shya please insert some additional material for multivariate datasets - perhaps
-you don't need much because basically you just ran the bivariate analysis on each pair
-of locations, but you need to say that and then maybe a few sentences about how 
-everything was tallied up after having done that ??? -->
-
-### Non-parametric analysis of tail dependence \label{NPA}
-
-Left- and right-tail dependence quantify the extent to which two ranked
-variables $u_i$, $v_i$ for $i=1,...,n$ 
-(see Sec. \ref{Model_selection} for defintions of $u_i$ and $v_i$)
-are related in the left and right tails, respectively, of their 
-distributions [@nelsen2007_copula]. 
-Our nonparametric study of tail dependence uses three
-statistics which can quantify the extent to which $u_i$ and $v_i$ are 
-related in any part of their distributions. We here describe the 
-statistics. The statistics were formulated with positively associated
-variables in mind. All the variables we study are positively associated
-when they are significantly associated. So definitions and associated 
-diagrams will assume positive associations, though generalizations are 
-possible. Given two bounds 
-$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
-$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
-fall (these are the blue and red lines, respectively, in 
-Figure \ref{fig_stat_image}). Our statistics quantify the dependence
-between $u_i$ and $v_i$ in the region bounded by these lines.
-Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
-dependence in the left parts of the distributions of $u$ and $v$, and 
-using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
-the right parts of the distributions.
-
-![Schematic diagram supporting defintions of nonparametric tail dependence statistics.\label{fig_stat_image}](./stat_image.jpg){ width=25%, height=25% }
-
-The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
-correlation of $u_i$ and $v_i$ that is attributable to the points 
-lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
-\begin{equation}\label{eq.Cor}
-   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
-   (u_i-\mean(u)) (v_i-\mean(v)),
-\end{equation}
-where means and variances are computed using all $n$ data 
-points, but the
-sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
-$u_i+v_i < 2u_b$. 
-Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
-between $u$ and $v$ in the region given by the bounds.
-
-We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
-$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
-$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
-and a distance less than $h$ from the line $v=u$, divided by the 
-total number of points within the bounds. This function is
-defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
-is half the longer of the two segments obtained by intersecting the
-bounds with the unit square. It is easy to see $S(0)=0$ and 
-$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
-of $h$, to be the area within the bounds and the unit square
-and within distance $h$ of the line $v=u$, divided by the area
-within the bounds and the unit square. This is the expected value of
-$S(h)$ for independent data. We then defined
-\begin{equation}\label{eq.P}
-   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
-\end{equation}
-As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
-indicate stronger positive association
-between $u$ and $v$ in the region given by the bounds.
-
-The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
-distance between points satisfying $u_i+v_i>2l_b$ and 
-$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
-$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
-$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
-then small. 
-Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
-dependence in the portions of the distributions of $u$ and $v$ given by 
-the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
-$\Dsq_{l_b,u_b}$ indicate strong dependence.
-
-For large datasets (i.e., datasets with large $n$),
-we can consider $l_b$ and $u_b$ close together without incurring undue
-sampling variation in our statistics, and we can consider 
-several different bands $(l_b,u_b)$ to understand how dependence
-varies in different parts of the distributions. But for datasets with 
-smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
-We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
-and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
-$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.
-
-The difference $\cor_l-\cor_u$ is expected to be positive if 
-dependence in the left halves of the distributions is stronger 
-than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
-This holds, for instance, for the Clayton copula [@nelsen2007_copula].
-The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
-"lower" statistics occur in opposite order, is also expected to be 
-positive under the same conditions. We use these three differences
-to detect asymmetry of tail dependence. Likewise, for large datasets,
-we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
-$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
-$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
-
-In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we 
-also tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and 
-$\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated
-from various copulas. Normal and Frank copula families 
+# Testing non-parametric statistics for various symmetric and asymmetric copulas \label{Test_npa_stat}
+
+Normal and Frank copula families 
 are known to have symmetric tail dependence,
@@
 
-For bivariate datasets such as the soil C and N data, we compared values of the statistics $\cor_{l_b,u_b}$,
-$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to distributions of values of the same 
-statistics computed on surrogate datasets that 
-were produced from the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
-the same marginal distributions as the empirical data and had very similar Kendall or Spearman correlation, but had
-normal copula structure, and thus our comparisons tested the null hypothesis that our statistics took values on the
-empirical data no different from what would have been expected if the copula structure of the data were normal, but the
-data were otherwise statistically unchanged.
-For bivariate datasets, surrogates were produced as follows.
-
-- Given data $(x_i,y_i)$ for $i=1,\ldots,n$ and a 
-one-parameter target copula family $\mathcal{B}(\theta)$ (here $\theta$ denotes the parameter), let $\tau$ be the
-Kendall correlation of the $x_i$ with the $y_i$ and find the parameter value $\theta_\tau$ for which 
-$\mathcal{B}(\theta_\tau)$ has Kendall correlation $\tau$ (this is possible for the one-parameter copula families of this 
-study using the `iTau` function of the `VineCopula` package). 
-- Generate data $(a_i,b_i)$, $i=1,\ldots,n$ from 
-$\mathcal{B}(\theta_\tau)$. 
-- Permute the ordered set $(x_1,\ldots,x_n)$ such that the permutation
-$(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ has $\rank(x_{\sigma_x(i)})$ equal to $\rank(a_i)$ for all $i$, where 
-$\rank(x_{\sigma_x(i)})$ is the rank of $x_{\sigma_x(i)}$ in the set $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$, and
-$\rank(a_i)$ is the rank of $a_i$ in the set $(a_1,\ldots,a_n)$ (here $\sigma_x$ is a permutation of the 
-indices $1,\ldots,n$). 
-- Likewise permute $(y_1,\ldots,y_n)$
-such that the permutation
-$(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ has $\rank(y_{\sigma_y(i)})$ equal to $\rank(b_i)$ for all $i$ ($\sigma_y$ is 
-another permutation of $1,\ldots,n$).
-- The surrogate dataset is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.
-
-Our code for this algorithm is [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R) in the 
-[`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
-
-Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
-is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a permutation 
-of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as those of the original data. 
-Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
-Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
-will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
-a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
-as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
-sampling variation) the Spearman correlation of the empirical data.
-
-The comparison of one of the
-statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the empirical data, to the distribution
-of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
-provides a test of whether dependence in the part of the distribution specified by $l_b$ and $u_b$ is different from
-what would be expected from a null hypothesis of normal copula structure but unchanged marginals and unchanged 
-Kendall (or Spearman)
-correlation. Significant deviations correspond to deviations from normal copula structure. In particular, deviations 
-using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence different from that of a normal copula;
-likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different from that of a normal copula.  
-The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, as computed on the 
-empirical data, to the distribution
-of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
-provides a test of asymmetry of tail dependence.
-
-For multivariate datasets, e.g., the aphid abundance or first flight data, each of which consists of time series 
-from 10 locations, a spatial resampling scheme was used instead of the surrogate approach described above.
-The scheme is identical to that proposed by [@bjornstadfalck2001] for application in their now 
-widely used spline correlogram method; the same scheme was also used by [@walteretal2017] as part of a method for 
-performing model selection among matrix regression models. For each pair of locations $i \neq j$, the statistic
-of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
-$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The values $s_{i,j}$ were then arranged in
-a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
-$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the average strength of 
-dependence, across all pairs of distinct locations, for data in the portion of the distributions 
-given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
-asymmetry of tail dependence across all pairs of locations. 
-Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a distribution of bootstrapped values.
-Bootstrapping was performed by resampling the locations of measurement, with replacement; this corresponds to 
-resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
-[@bjornstadfalck2001 ; @walteretal2017].
-
 \newpage
 
+# Question 1 analyses for each dataset
+Recall that question 1 is: Do datasets in ecology and related fields have non-Gausian copula structure?
+Do they show tail dependence distinct from that of a Gaussian copula?
+In particular do they show asymmetric tail dependence?
+
+<!--
+## Details of methods \label{Method}
+
+### Model selection approach \label{Model_selection}
+
+### Non-parametric analysis of tail dependence \label{NPA}
+
+moved to Paper.Rmd
+-->
+
+<!--
 ## Results: soil C and N data \label{Ex_SoilCN}
@@
 ### Model selection approach \label{Ex_SoilCN_Model_selection}
+
+moved to Paper.Rmd
+-->
+
 <!--This does the computations for this section-->
@@
 BivMS_res_Loecke<-OurBiCopSelect(u1=v_CN[,1],u2=v_CN[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
-```
-
-The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
-from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_soilCNfit}).
-The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+saveRDS(BivMS_res_Loecke,file="./Results/fitting_results/BivMS_soilCN.RDS")
+```
 
@@
 ```
-
-The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`.
-These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.
-
-Although these results do convincingly show non-Gaussian copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
-The next section looks at tail dependence in a non-parametric way.
 
@@
 \newpage
-
-# References:
+# References

### 2018-05-02 Paper.Rmd
v1.description text moved from SupportingInformation.Rmd to Paper.Rmd

--- 2018-02-19 Paper.Rmd
@@
 title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
-author: "Shyamoline Ghosh, other authors, Daniel Reuman"
+author: "Shyamolina Ghosh, other authors, Daniel Reuman"
 date: "February 18, 2018"
+geometry: "left=1cm,right=1cm,top=2.5cm,bottom=2.8cm"
+
 output: 
   pdf_document:
+    number_sections: no
+    keep_tex: yes
     fig_caption: yes
-    fig_width: 5
-    highlight: tango
-    keep_tex: yes
-    number_sections: yes
-    df_print: kable # data frame printing option
-fontsize: 11pt
+    includes:
+      in_header: head_maintext.sty
+
+tables: True
+link-citations: True
+urlcolor : blue
+
+csl: ecology-letters.csl
+bibliography: REF_ALL.bib
 ---
+updated on `r Sys.Date()`
 
@@
 
+*Type of Article*
+
+- Ideas and Perspectives
+
+*The number of words in the abstract*
+
+*The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends)*
+
+*The number of words in each text box*
+
+*The number of references*
+
+*The number of figures, tables, and text boxes*
+
+*The name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent*
+
+\newpage
 # Abstract
 
+\newpage
 # Introduction
@@
 
-# Methods
+# Material and methods
 
@@
 
-## Statistical methods
+## Statistical methods  \label{Methods}
 
-Shyamolina, please work on a description of the model-selection approach you have implemented and insert it here.
-You can just talk about how you go about doing model selection to decide on the best copula of several alternatives (and what those alternatives are) for a bivariate dataset, including the independence test step, the AIC step, and the goodness of fit test for the AIC-best copula.
-Omit the goodness of fit test for the normal copula, please, since I think we won't be using that.
+### Model selection approach \label{Model_selection}
 
-Shyamolina and Dan to discuss which tail-dependence statistic(s) to use and then we can insert a description here.
+For a bivariate dataset $(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps, 
+most implemented using tools from the `VineCopula` package. Code combining these tools is in the
+[`OurBiCopSelect`](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
 
-Dan to insert a description of the surrogate approach here.
+First, we produced ranks 
+$(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
+$\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
+we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; further work was with the $u_i$ 
+and $v_i$. Ranks $\tilde{u}_i$ and $\tilde{v}_i$ were $1$, for the 
+smallest element, up to $n$, for the largest.
+
+Second, we did a test of the independence of the $u_i$ and $v_i$. Subsequent model 
+selection algorithms are ineffective if data could not be distinguished from independent data 
+since most or all bivariate copula families considered include the independent copula in the 
+family or as a boundary case. The independence test was as implemented in the `BiCopIndTest`
+function in `VineCopula`, the documentation of which reports the test uses the asymptotic normality of the statistic $|\tau|\sqrt{\frac{9n(n-1)}{2(2n+5)}}$, where $\tau$ is Kendall's tau statistic.
+If independence could be rejected ($0.05$ significance level), model selection proceeded.
+
+The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`.
+Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulas.
+Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC) values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection].
+Descriptions of these copula families can be found in introductory texts on copulas [@nelsen2007_copula ; @joe2014_dependence]  and in references cited in `VineCopula`.
+The lower- and upper-tail dependence of the
+best-fitting member of each family was also provided by `BiCopEst`.
+
+Model selection methods will give the *relative* support of several models, but will not,
+on their own, indicate whether any of the models selected is an objectively good fit (all 
+models considered might poorly describe the data). To test this, and as the fourth step of
+our procedure, we tested the goodness of fit of our best-AIC copula family using a 
+bootstrapping procedure implemented in `BiCopGofTest` of `VineCopula`. The method performed 
+two tests, one based on a Cramer-von Mises statistic, and one based on a 
+Kolmogorov-Smirnov statistic. For computational efficiency, 
+an initial run using $100$ bootstraps was performed, and if
+the $p$-value resulting from either test was less than $0.2$, tests were re-run with 
+$1000$ boostraps.
+
+Finally, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
+model-averaged lower- and upper-tail dependence values using standard model averaging 
+formulas [@burnham2003_modelselection]; likewise for BIC.
+
+<!--Shya please insert some additional material for multivariate datasets - perhaps
+you don't need much because basically you just ran the bivariate analysis on each pair
+of locations, but you need to say that and then maybe a few sentences about how 
+everything was tallied up after having done that ??? -->
+
+### Non-parametric analysis of tail dependence \label{NPA}
+
+Left- and right-tail dependence quantify the extent to which two ranked
+variables $u_i$, $v_i$ for $i=1,...,n$ 
+(see \nameref{Model_selection} for defintions of $u_i$ and $v_i$)
+are related in the left and right tails, respectively, of their 
+distributions [@nelsen2007_copula]. 
+Our nonparametric study of tail dependence uses three
+statistics which can quantify the extent to which $u_i$ and $v_i$ are 
+related in any part of their distributions. We here describe the 
+statistics. The statistics were formulated with positively associated
+variables in mind. All the variables we study are positively associated
+when they are significantly associated. So definitions and associated 
+diagrams will assume positive associations, though generalizations are 
+possible. Given two bounds 
+$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
+$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
+fall (these are the blue and red lines, respectively, in 
+Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
+between $u_i$ and $v_i$ in the region bounded by these lines.
+Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
+dependence in the left parts of the distributions of $u$ and $v$, and 
+using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
+the right parts of the distributions.
+
+![Schematic diagram supporting defintions of nonparametric tail dependence statistics.\label{fig_stat_image}](./stat_image.jpg){ width=25%, height=25% }
+
+The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
+correlation of $u_i$ and $v_i$ that is attributable to the points 
+lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
+\begin{equation}\label{eq.Cor}
+   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
+   (u_i-\mean(u)) (v_i-\mean(v)),
+\end{equation}
+where means and variances are computed using all $n$ data 
+points, but the
+sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
+$u_i+v_i < 2u_b$. 
+Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
+between $u$ and $v$ in the region given by the bounds.
+
+We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
+$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
+$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
+and a distance less than $h$ from the line $v=u$, divided by the 
+total number of points within the bounds. This function is
+defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
+is half the longer of the two segments obtained by intersecting the
+bounds with the unit square. It is easy to see $S(0)=0$ and 
+$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
+of $h$, to be the area within the bounds and the unit square
+and within distance $h$ of the line $v=u$, divided by the area
+within the bounds and the unit square. This is the expected value of
+$S(h)$ for independent data. We then defined
+\begin{equation}\label{eq.P}
+   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
+\end{equation}
+As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
+indicate stronger positive association
+between $u$ and $v$ in the region given by the bounds.
+
+The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
+distance between points satisfying $u_i+v_i>2l_b$ and 
+$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
+$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
+$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
+then small. 
+Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
+dependence in the portions of the distributions of $u$ and $v$ given by 
+the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
+$\Dsq_{l_b,u_b}$ indicate strong dependence.
+
+For large datasets (i.e., datasets with large $n$),
+we can consider $l_b$ and $u_b$ close together without incurring undue
+sampling variation in our statistics, and we can consider 
+several different bands $(l_b,u_b)$ to understand how dependence
+varies in different parts of the distributions. But for datasets with 
+smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
+We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
+and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
+$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.
+
+The difference $\cor_l-\cor_u$ is expected to be positive if 
+dependence in the left halves of the distributions is stronger 
+than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
+This holds, for instance, for the Clayton copula [@nelsen2007_copula].
+The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
+"lower" statistics occur in opposite order, is also expected to be 
+positive under the same conditions. We use these three differences
+to detect asymmetry of tail dependence. Likewise, for large datasets,
+we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
+$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
+$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
+
+In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we 
+also tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and 
+$\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated
+from various copulas (for details see Appendix \ref{SM-Test_npa_stat}).
+
+For bivariate datasets such as the soil C and N data, we compared values of the statistics $\cor_{l_b,u_b}$,
+$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to distributions of values of the same 
+statistics computed on surrogate datasets that 
+were produced from the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
+the same marginal distributions as the empirical data and had very similar Kendall or Spearman correlation, but had
+normal copula structure, and thus our comparisons tested the null hypothesis that our statistics took values on the
+empirical data no different from what would have been expected if the copula structure of the data were normal, but the
+data were otherwise statistically unchanged.
+For bivariate datasets, surrogates were produced as follows.
+
+- Given data $(x_i,y_i)$ for $i=1,\ldots,n$ and a 
+one-parameter target copula family $\mathcal{B}(\theta)$ (here $\theta$ denotes the parameter), let $\tau$ be the
+Kendall correlation of the $x_i$ with the $y_i$ and find the parameter value $\theta_\tau$ for which 
+$\mathcal{B}(\theta_\tau)$ has Kendall correlation $\tau$ (this is possible for the one-parameter copula families of this 
+study using the `iTau` function of the `VineCopula` package). 
+- Generate data $(a_i,b_i)$, $i=1,\ldots,n$ from 
+$\mathcal{B}(\theta_\tau)$. 
+- Permute the ordered set $(x_1,\ldots,x_n)$ such that the permutation
+$(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ has $\rank(x_{\sigma_x(i)})$ equal to $\rank(a_i)$ for all $i$, where 
+$\rank(x_{\sigma_x(i)})$ is the rank of $x_{\sigma_x(i)}$ in the set $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$, and
+$\rank(a_i)$ is the rank of $a_i$ in the set $(a_1,\ldots,a_n)$ (here $\sigma_x$ is a permutation of the 
+indices $1,\ldots,n$). 
+- Likewise permute $(y_1,\ldots,y_n)$
+such that the permutation
+$(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ has $\rank(y_{\sigma_y(i)})$ equal to $\rank(b_i)$ for all $i$ ($\sigma_y$ is 
+another permutation of $1,\ldots,n$).
+- The surrogate dataset is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.
+
+Our code for this algorithm is [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R) in the 
+[`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
+
+Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
+is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a permutation 
+of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as those of the original data. 
+Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
+Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
+will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
+a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
+as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
+sampling variation) the Spearman correlation of the empirical data.
+
+The comparison of one of the
+statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the empirical data, to the distribution
+of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
+provides a test of whether dependence in the part of the distribution specified by $l_b$ and $u_b$ is different from
+what would be expected from a null hypothesis of normal copula structure but unchanged marginals and unchanged 
+Kendall (or Spearman)
+correlation. Significant deviations correspond to deviations from normal copula structure. In particular, deviations 
+using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence different from that of a normal copula;
+likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different from that of a normal copula.  
+The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, as computed on the 
+empirical data, to the distribution
+of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
+provides a test of asymmetry of tail dependence.
+
+For multivariate datasets, e.g., the aphid abundance or first flight data, each of which consists of time series 
+from 10 locations, a spatial resampling scheme was used instead of the surrogate approach described above.
+The scheme is identical to that proposed by [@bjornstadfalck2001] for application in their now 
+widely used spline correlogram method; the same scheme was also used by [@walteretal2017] as part of a method for 
+performing model selection among matrix regression models. For each pair of locations $i \neq j$, the statistic
+of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
+$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The values $s_{i,j}$ were then arranged in
+a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
+$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the average strength of 
+dependence, across all pairs of distinct locations, for data in the portion of the distributions 
+given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
+asymmetry of tail dependence across all pairs of locations. 
+Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a distribution of bootstrapped values.
+Bootstrapping was performed by resampling the locations of measurement, with replacement; this corresponds to 
+resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
+[@bjornstadfalck2001 ; @walteretal2017].
+
+<!--Dan to insert a description of the surrogate approach here.-->
 
@@
 
-## Do datasets have non-Gaussian copula structure and tail dependence?
+## question 1
 
-Model selection results:  
-- Paragraph on model selection results for the soil C and N data from Terry
-- Other stuff, including aphids, depending on what datasets we settle on
+- Do datasets in ecology and related fields have non-Gausian copula structure?
+- Do they show tail dependence distinct from that of a Gaussian copula?
+- In particular do they show asymmetric tail dependence?
 
-Tail dependence results
-- again start with C/N data from Terry
-- other stuff, inc. aphids
+### soil C and N data \label{Ex_SoilCN}
 
-## What are some causes of non-Gaussian copula structure and tail dependence?
+```{r read_BivMS_soilCN_data,echo=F}
+BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
+```
 
-## What are the consequences of non-Gaussian copula structure and tail dependence?
+#### Model selection approach \label{Ex_SoilCN_Model_selection}
+
+The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
+from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}).
+The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+
+The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`.
+These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.
+
+Although these results do convincingly show non-Gaussian copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+The next section looks at tail dependence in a non-parametric way.
+
+#### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}
+
+As described in Methods, each of our statistics, as computed for the 
+ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
+of values of the same statistic with the same ranges computed on 1000 normal-copula 
+surrogate datasets with similar Kendall or Spearman correlations, in different 
+runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). Because the correlation 
+used in the $\cor$ statistic is the Spearman correlation, Spearman-preserving surrogates
+should be used, so the top-right panel of Fig. \ref{SM-fig_soilCN_nonparam}
+is more reliable than the top left. Results indicate that tail dependence of 
+data is stronger in both the lower and upper tails than would be expected under a 
+null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
+weaker in the middle of the distributions
+than would be expected based on a normal-copula null model.
+
+We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{SM-fig_raw_cop_soilCN}(B). 
+Results (Table \ref{SM-tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.
+
+### BMR data \label{Ex_BMR}
+
+```{r read_BivMS_BMR_data,echo=F}
+BivMS_BMR<-readRDS("./Results/BMR_results/BivMS_BMR_lnmass.RDS")
+```
+
+#### Model selection approach \label{Ex_BMR_Model_selection}
+
+The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_BMRfit}).
+The table shows that Frank is much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+
+The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was $0$ as expected, because the contributions came from Frank and SBB8 which are both symmetric copula.
+We also tested that lowest-AIC copula Frank was a good-enough fit, giving $p$-value $`r BivMS_BMR$GofRes_CvM`$ (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value $`r BivMS_BMR$GofRes_KS`$ in the Kolmogorov-Smirnov-based test of the goodness of fit.
+The next section looks at tail dependence in a non-parametric way.
+
+#### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
+<!--we have to add description here after DAN fixes bug(?) in bivfunctionplot.R-->
+
+### Anolis data \label{Ex_contrasts}
+
+```{r read_BivMS_contrasts_67_data,echo=F}
+BivMS_contrasts_67<-readRDS("./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
+```
+
+#### Model selection approach \label{Ex_contrasts_Model_selection}
+
+The test of indepedence gave $p=$ `r BivMS_contrasts_67$IndepTestRes` (to within the precision available 
+from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_contrasts_67fit}).
+The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+
+The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`.
+These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.
+
+These results do convincingly show non-Gaussian copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+The next section looks at tail dependence in a non-parametric way.
+
+#### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}
+<!--we have to add description here after DAN fixes bug(?) in bivfunctionplot.R-->
+
+### Abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
+
+```{r read_RES_aphid_count, echo=F}
+RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
+```
+
+#### Model selection approach \label{Ex_count_Model_selection}
+
+As described in Methods, model selection among copulas was done separately for each pair of the 10 
+sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns 
+corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location 
+pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of 
+question 1 for these data. Comparing the 
+AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) 
+to the AIC weight 
+of the normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for normal copulas was usually not
+comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
+failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
+the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}),
+demonstrating that the collection of copula families we used was sufficiently broad that the copulas of 
+all pairs of locations could be characterized by at least one of them.
+In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+
+The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively.
+These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data.
+Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.
+
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1.
+
+#### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
+
+Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
+computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
+resampling scheme, as described in Methods (Table \ref{SM-tab_count_resamp}). Confidence intervals of
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
+indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
+exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
+structure do not have, providing further evidence answering the third part of question 1.
+
+Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig.
+\ref{SM-fig_aphid_count} A,B) showed the typical decline of correlation with
+distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting 
+$\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig.
+\ref{SM-fig_aphid_count} C-F).
+Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows
+these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively).
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig.
+\ref{SM-fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical
+result of Table \ref{SM-tab_count_resamp} that averages of these quantities across all pairs of sampling locations
+were significantly greater than zero, demonstrating asymmetry of tail dependence.
+Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) package.
+
+### First flight date data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}
+
+```{r read_RES_aphid_ff,echo=F}
+RES_aphid_ff<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_firstflight_species_11.RDS")
+```
+
+#### Model selection approach \label{Ex_ff_Model_selection}
+
+As in section \nameref{Ex_count_Model_selection}, results are presented in $10 \times 10$ matrix format with rows
+and columns corresponding to locations.
+Table \ref{SM-tab_ff_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data.
+Comparing the AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_ff_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_ff_AICw_normal}) confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_ff_p_CvM}) usually failed to reject the
+hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for 
+goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_ff_p_KS}).
+In all cases at least `r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for
+goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+
+The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and 
+the model-averaged (again, AIC based) upper-tail dependence statistic are in 
+Tables \ref{SM-tab_ff_LTdep_AICw} and \ref{SM-tab_ff_UTdep_AICw}, respectively.
+These values were exclusively positive, thereby differing from what a normal copula would give (i.e., $0$), and answering the second part of question 1 for these data - these data have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
+
+The difference between the model-averaged lower-tail dependence statistic and the 
+model-averaged upper-tail dependence statistic (lower minus upper), shown in 
+Table \ref{SM-tab_ff_LTmUT}, was negative for `r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)` 
+out of `r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair` location pairs, answering the third part of question 1.
+We note this result is opposite the analogous result for aphid abundance
+data for the green spruce aphid, i.e., whereas those abundance data showed stronger lower tail dependendence than upper tail dependence, first flights for the leaf-curling plum aphid showed stronger upper tail dependence than lower tail dependence.
+
+#### Non-parametric analysis of tail dependence  \label{Ex_ff_NPA}
+
+Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman 
+correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were 
+computed via a spatial resampling scheme, as described in \nameref{Methods} (Table \ref{SM-tab_ff_resamp}).
+Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded 
+zero and mean values were positive, indicating tail dependence.
+Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always exluded $0$ and were negative, indicating asymmetry of tail dependence, a feature
+that data with normal copula structure do not have, providing further evidence answering the third part of question 1.
+The contrast here with the analogous results for abundances of the green spruce aphid is again illuminating.
+
+Plotting Spearman or Kendall correlations between first flight time series for pairs of locations as a function of great circle distance between the locations (Fig.
+\ref{SM-fig_aphid_ff} A, B) showed the typical (though slight, in this case) decline of correlation with increasing distance.
+Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear in most cases to decline slightly with distance (Fig.
+\ref{SM-fig_aphid_ff} C-F).
+$\Dsq_u$ and $\Dsq_l$ appear to slightly increase with distance, as expected given their
+opposite meaning (Fig. \ref{SM-fig_aphid_ff} G-H). 
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
+\ref{SM-fig_aphid_ff} I-K) did not show a decline with distance, but did demonstrate visually the statistical result of Table \ref{SM-tab_ff_resamp} that averages of these quantities across all pairs of sampling locations were significantly less than zero, showing asymmetry of tail dependence.
+
+### Abundance data for *Ceratium furca* in UK seas \label{Ex_pns}
+
+```{r read_RES_plankton_north_sea,echo=F}
+RES_plankton_north_sea<-readRDS("./Results/fitting_results/Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS")
+```
+
+#### Model selection approach \label{Ex_pns_Model_selection}
+
+As described in Methods and as in previous sections, model selection among copulas was done separately for 
+each pair of sampling locations.
+There were $14$ sampling locations, so results are presented in $14 \times 14$ matrix format, with rows and columns corresponding to locations.
+Table \ref{SM-tab_pns_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data.
+Comparing the AIC weight of the lowest-AIC copula family for each pair of sampling locations (Table \ref{SM-tab_pns_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_pns_AICw_normal})
+confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_pns_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_pns_p_KS}).
+In all cases at least
+`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+
+The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_pns_LTdep_AICw} and \ref{SM-tab_pns_UTdep_AICw}, respectively.
+These values were exclusively positive, thereby
+differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
+have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
+
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
+(lower minus upper), shown in Table \ref{SM-tab_pns_LTmUT}, was positive for some
+location pairs and negative for others, but was positive for 
+`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`
+locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
+
+#### Non-parametric analysis of tail dependence  \label{Ex_pns_NPA}
+
+Mean values of our statistics, as well as mean Kendall and Spearman
+correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
+spatial resampling scheme, as described in Methods (Table \ref{SM-tab_pns_resamp}).
+Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
+always excluded $0$ and mean values were positive, indicating tail dependence.
+Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
+also all were strictly positive, indicating asymmetry of tail dependence, a feature that data with normal copula
+structure do not have, providing further evidence answering the third part of question 1.
+
+Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
+between the locations (Fig. \ref{SM-fig_pns} A, B) showed the typical decline of correlation with increasing
+distance.
+Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig.
+\ref{SM-fig_pns} C-F).
+$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given their opposite meaning (Fig.
+\ref{SM-fig_pns} G-H).
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
+\ref{SM-fig_pns} I-K) did not show a decline with distance, but did demonstrate visually the
+statistical result of Table \ref{SM-tab_pns_resamp} that averages of these 
+quantities across all pairs of sampling locations were significantly greater 
+than zero, showing asymmetry of tail dependence.
+
+### Methane flux data \label{Ex_methane}
+
+```{r read_RES_methane,echo=F}
+RES_methane<-readRDS("./Results/fitting_results/MethaneFluxFit_selecloc_data_pt_thrs_50.RDS")
+```
+
+#### Model selection approach \label{Ex_methane_Model_selection}
+
+As described in Methods and as in previous sections, model selection among copulas was done separately for 
+each pair of sampling locations. There were $13$ sampling locations, so results are presented in 
+$13 \times 13$ matrix format, with rows and columns corresponding to locations. 
+Table \ref{SM-tab_methane_bestcop} shows that, for the large majority of location pairs, some 
+non-normal copula was a better fit, according to AIC, than was the normal copula, answering the 
+first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula 
+family for each pair of sampling locations (Table \ref{SM-tab_methane_AICw}) to the AIC weight of
+the normal copula family for the same pair (Table \ref{SM-tab_methane_AICw_normal}) 
+confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_methane_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_methane_p_KS}).
+In all cases at least
+`r min((RES_methane$gfc_numBS_success/RES_methane$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+
+The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_methane_LTdep_AICw} and \ref{SM-tab_methane_UTdep_AICw}, respectively.
+These values were exclusively positive, thereby
+differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
+have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
+
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
+(lower minus upper), shown in Table \ref{SM-tab_methane_LTmUT}, was positive for some
+location pairs and negative for others, but was positive for 
+`r sum((RES_methane$LTdep_AICw-RES_methane$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_methane$gfc_p_CvM)[1]^2-dim(RES_methane$gfc_p_CvM)[1])-2*RES_methane$num_indep_loc_pair`
+locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
+
+#### Non-parametric analysis of tail dependence  \label{Ex_methane_NPA}
+
+Mean values of our statistics, as well as mean Kendall and Spearman
+correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
+spatial resampling scheme, as described in Methods (Table \ref{SM-tab_methane_resamp}).
+Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
+always excluded $0$ and mean values were positive, indicating tail dependence.
+But the confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ were marginal, indicating no significant asymmetry of tail dependence.
+
+Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
+between the locations (Fig. \ref{SM-fig_methane} A, B) showed the typical decline of correlation with increasing
+distance.
+Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig.
+\ref{SM-fig_methane} C-F).
+$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
+their opposite meaning (Fig.
+\ref{SM-fig_methane} G-H).
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
+\ref{SM-fig_methane} I-K) did not show a decline with distance, but did demonstrate visually the
+statistical result of Table \ref{SM-tab_methane_resamp} that averages of these 
+quantities across all pairs of sampling locations were more or less equally dispersed around the zero-line, showing no significant asymmetry of tail dependence.
+
+<!--Question 2 analyses-->
+## Some possible causes of non-Gaussian copula structure
+
+Q1/2: How are copula structure of ecologically influential climate variables changing as part of climate change?
+We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the pops, but we don’t really know how a change in the environment cop structure would make what kind of change in the pop cop structure.
+How?
+
+### The Moran effect and copula structure in populations
+
+We show here that if a non-normal copula structure describes the dependence of an environmental
+variable between two locations, then there can be a non-normal copula structure for the
+dependence of populations in the locations that are influenced by the environmental variables.
+
+To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
+\ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.
+
+\begin{equation}\label{eq.AR1}
+P_{i}(t+1)=\beta P_{i}(t)+\alpha \epsilon_{i}(t)
+\end{equation}
+
+To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$.
+We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches.
+Should their dependence be exactly the same as of the noises or not ?
+
+To test that our approach is of two way : first, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
+Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome which was a bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
+\ref{SM-fig_Cause4copula_params_GOF_500}).
+We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size and reported p-values for both statistics (Cramer-von Mises and Kolmogorov-Smirnov).
+p-values greater or equal to $0.05$ (above the dashed line in Fig.
+\ref{SM-fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters.
+Fig.
+\ref{SM-fig_Cause4copula_params_GOF_2500} represents similar figure as of Fig.
+\ref{SM-fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$.
+(For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
+
+Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
+Then we plotted mean of each measured quantity with standard error against Spearman's Rho (or Kendall's Tau) of the noise copula family.
+Fig.
+\ref{SM-fig_Cause4copula_C_spear} and Fig.
+\ref{SM-fig_Cause4copula_C_kend} represent this picture for noise generating from a lower tail dependent Clayton copula whereas for Fig.
+\ref{SM-fig_Cause4copula_SC_spear} and Fig.
+\ref{SM-fig_Cause4copula_SC_kend} noise came from a upper tail dependent Survival Clayton family.
+We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach, here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise.
+(For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
+
+### Asymmetric sensitivity and copula structure in populations
+
+<!--Another model, distinct from the above, Dan and Shya to discuss-->
+
+### Need to think about mechanisms for biogeochemical, evolutionary, and community data
+
+<!--Question 3 analyses-->
+## Consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?
+
+### Effect on skewness of the spatial mean
+
+Here, we claim that because of significant tail dependence of the multivariate copula structure, the skewness of the time-series data averaged over all selective locations changed significantly from the individual skewness of time-series data taken at each pairwise location.
+Spatial mean of the time series data extracts that tail dependence of the multivariate data set, if any, and reflected in the averaged skewness.
+We tested this with null hypothesis $H_{0}$.
+If $H_{0}$ is true that means our multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
+So, first we generated $10000$ number of surrogate normal copula from our given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
+Then we plotted the distribution of spatial mean of each surrogate time series data set and the given multivariate data.
+If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence.
+
+To verify that we used aphid-count data, aphid-firstflight data, plankton data and methane-flux data.
+For aphid data we choose only those locations for which atleast $30$ number of data points present in each time series when taken as pairwise locations.
+For plankton-UK-sea data and methane-flux data that threshold on number of data points are $45$ and $50$, respectively.
+
+- Aphid count \label{skewness_count}
+
+Figure \ref{SM-fig_skewness_count} shows that averaged skewness is much lower than expected if the aphid count data came from a normal copula.
+Here, $p$ value indicates the fraction of surrogate-skewness which are lower than the averaged skewness of the given data (red line in Figure \ref{SM-fig_skewness_count}).
+As $p<0.05$, we reject $H_{0}$.
+This result is consistent with the non-parametric analysis and model fitting results which indicated lower tail dependence.
+
+- Aphid first flight \label{skewness_ff}
+
+Figure \ref{fig_skewness_ff} shows that averaged skewness is much higher than expected if the aphid first flight data came from a normal copula.
+Here, $p$ value indicates the fraction of surrogate-skewness which are greater than the averaged skewness of the given data (red line in Figure \ref{SM-fig_skewness_ff}).
+As $p<0.05$, we reject $H_{0}$ in support for upper tail dependence.
+
+- UK Sea plankton \label{skewness_pns}
+
+Figure \ref{SM-fig_skewness_pns} shows that averaged skewness is lower than expected if the plankton data came from a normal copula.
+Our previous results on non-parametric analysis and model fitting showed lower tail dependence.
+Here, as $p>0.05$, we cannot reject $H_{0}$, .......???
+
+- Methane-flux data \label{skewness_methane}
+
+Figure \ref{SM-fig_skewness_methane} shows that averaged skewness is lower than expected if the plankton north sea data came from a normal copula.
+Our previous results on non-parametric analysis and model fitting showed lower tail dependence.
+As $p<0.05$, we reject $H_{0}$.
+This result is consistent with the non-parametric analysis and model fitting results which indicated lower tail dependence.
+
+### Effect on extinction risk
+
+- LC model (r=0) :
+
+    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
+[see first plot of Fig.
+\ref{SM-fig_risk_local_disp} and Fig.
+\ref{SM-fig_risk_global_disp}]
+
+- Ricker model (r $\neq$ 0):
+
+    * For ricker model when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
+For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
+[see Fig.
+\ref{SM-fig_risk_local_disp} - \ref{SM-fig_risk_global_disp}]
+
+    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]
+
+With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig.
+\ref{SM-fig_risk_numlocs_effect_local_disp} - Fig.
+\ref{SM-fig_risk_numlocs_effect_global_disp}] for both model.
 
 # Discussion
+
+\newpage
+# References

### 2018-05-02 SupportingInformation.Rmd
v1.description text moved from SupportingInformation.Rmd to Paper.Rmd

@@
 
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{table}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\setcounter{section}{0}
-\renewcommand{\thesection}{S\arabic{section}}
-
 updated on `r Sys.Date()`
-
-<!--New latex commands-->
-\newcommand{\mean}{\operatorname{mean}}
-\newcommand{\var}{\operatorname{var}}
-\newcommand{\cor}{\operatorname{cor}}
-\newcommand{\Ps}{\operatorname{P}}  
-\newcommand{\Dsq}{\operatorname{D}^2}
-\newcommand{\rank}{\operatorname{rank}}
 
@@
 \includegraphics[width=3 in]{./Results/SoilCNcop.pdf}
-\caption{\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data.
+\caption[\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data]{\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data.
 Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see Sec.
@@
 \includegraphics[width=3 in]{./Results/BMR_results/copula_BMR_vs_lnmass.pdf}
-\caption{\textbf{(A)} Plot of the raw BMR vs.
+\caption[\textbf{(A)} Plot of the raw BMR vs.
+ln(mass) data \textbf{(B)} Copula plot for BMR dataset]{\textbf{(A)} Plot of the raw BMR vs.
 ln(mass) data \textbf{(B)} Copula plot for BMR dataset.
@@
 \includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
-\caption{\textbf{(A)} Plot of the raw AVG.ltoe.IV vs.
+\caption[\textbf{(A)} Plot of the raw AVG.ltoe.IV vs.
+AVG.humerus phylogenetic traits.
+\textbf{(B)} Copula plot of AVG.ltoe.IV vs.
+AVG.humerus phylogenetic traits]{\textbf{(A)} Plot of the raw AVG.ltoe.IV vs.
 AVG.humerus phylogenetic traits.
@@
 
-![Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
+\caption[Testing non-parametric statistics with datapoints $n=35$]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
 See text for details.
 $n=35$.
-\label{fig_stat_testing35}](./Results/stat_results/stat_testing/res_pt35_rankT.pdf)
-
-![Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
+\label{fig_stat_testing35}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
+\caption[Testing non-parametric statistics with datapoints $n=1000$]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
 See text for details.
-$n=1000$.\label{fig_stat_testing1000}](./Results/stat_results/stat_testing/res_pt1000_rankT.pdf)
-
-\newpage
-
+$n=1000$.\label{fig_stat_testing1000}}
+\end{center}
+\end{figure}
+
+<!--
 # Question 1 analyses for each dataset
@@
 
-<!--
 ## Details of methods \label{Method}
-
 ### Model selection approach \label{Model_selection}
-
 ### Non-parametric analysis of tail dependence \label{NPA}
-
 moved to Paper.Rmd
@@
 ## Results: soil C and N data \label{Ex_SoilCN}
-
 ### Model selection approach \label{Ex_SoilCN_Model_selection} 
-
-moved to Paper.Rmd
+### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}
+                    [[[[moved to Paper.Rmd]]]]
 -->
 
-<!--This does the computations for this section-->
+<!--This chunk is for model selection results of soilCN dataset-->
 ```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
-### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}
-
+<!--This chunk is for npa testing results of soilCN dataset-->
 ```{r stat_soilCN_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_CN,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
@@
 
-As described in Methods, each of our statistics, as computed for the 
-ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
-of values of the same statistic with the same ranges computed on 1000 normal-copula 
-surrogate datasets with similar Kendall or Spearman correlations, in different 
-runs for Kendall and Spearman (Figure \ref{fig_soilCN_nonparam}). Because the correlation 
-used in the $\cor$ statistic is the Spearman correlation, Spearman-preserving surrogates
-should be used, so the top-right panel of Figure \ref{fig_soilCN_nonparam}
-is more reliable than the top left. Results indicate that tail dependence of 
-data is stronger in both the lower and upper tails than would be expected under a 
-null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
-weaker in the middle of the distributions
-than would be expected based on a normal-copula null model.
-
-![Nonparametric tests for tail dependence and other deviations from a normal copula for soil C and N data.
-As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
+\caption[Testing non-parametric statistics for soilCN data]{Nonparametric tests for tail dependence and other deviations from a normal copula for soil C and N data.
+As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
 When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
 When $\cor$ values (respectively, $\Ps$ or $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, stronger than or weaker than for $\Ps$ or $\Dsq$) expected from a normal-copula null hypothesis.
-\label{fig_soilCN_nonparam}](./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf)
-
-We also tested statistically for asymmetry of tail dependence, which seems visible in Figure \ref{fig_raw_cop_soilCN}(B). 
-Results (Table \ref{tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.
+\label{fig_soilCN_nonparam}}
+\end{center}
+\end{figure}
 
@@
 
+<!--
 \newpage
-
 ## Results: BMR data \label{Ex_BMR}
-
 ### Model selection approach \label{Ex_BMR_Model_selection}
-
+### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
+                  ----moved to Paper.Rmd----
+-->
+
+<!--This chunk is for model selection results of BMR data-->
 ```{r bivMS_BMR, echo=F, cache=T, cache.extra=list(seed,v_BMR,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
-The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_BMRfit}).
-The table shows that Frank is much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
-
 <!--Display model selection results for BMR data in a table-->
@@
 
-The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was $0$ as expected, because the contributions came from Frank and SBB8 which are both symmetric copula.
-<!--$`r BivMS_BMR$relLTdep_AICw`$ and the model-averaged upper-tail dependence statistic was 
-$`r BivMS_BMR$relUTdep_AICw`$.
-These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper),
-$`r BivMS_BMR$relLTdep_AICw-BivMS_BMR$relUTdep_AICw`$, was different from 0, answering the third part of question 1. 
--->
-We also tested that lowest-AIC copula Frank was a good-enough fit, giving $p$-value 
-$`r BivMS_BMR$GofRes_CvM`$ (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value $`r BivMS_BMR$GofRes_KS`$ in the Kolmogorov-Smirnov-based test of the goodness of fit.
-The next section looks at tail dependence in a non-parametric way.
-
-### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
-
+<!--This chunk is for npa testing results of BMR dataset-->
 ```{r stat_BMR_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_BMR,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
@@
 
-![Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data.
-As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
-\label{fig_BMR_nonparam}](./Results/BMR_results/BMR_numbin_10_bivfunctionplot.pdf)
-
-![Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data.
-As described in Methods, each statistic was computed on the data using the ranges $0-0.2$, $0.2-0.4$, ..., $0.8-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
-\label{fig_BMR_nonparam5}](./Results/BMR_results/BMR_numbin_5_bivfunctionplot.pdf)
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/BMR_results/BMR_numbin_10_bivfunctionplot.pdf}
+\caption[Testing non-parametric statistics for BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data.
+As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
+\label{fig_BMR_nonparam}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/BMR_results/BMR_numbin_5_bivfunctionplot.pdf}
+\caption[Testing non-parametric statistics for BMR data with $numbin=5$]{Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data.
+As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.2$, $0.2-0.4$, ..., $0.8-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
+\label{fig_BMR_nonparam5}}
+\end{center}
+\end{figure}
 
@@
 
+<!--
 \newpage
-
 ## Results: contrasts.csv data \label{Ex_contrasts}
-
 ### Model selection approach \label{Ex_contrasts_Model_selection}
-
+### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}
+                 ----moved to Paper.Rmd----
+-->
+
+<!--This chunk is for model selection results of anolis data-->
 ```{r bivMS_contrasts_67, echo=F, cache=T, cache.extra=list(seed,v_contrasts_67,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
-The test of indepedence gave $p=$ `r BivMS_contrasts_67$IndepTestRes` (to within the precision available 
-from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{tab_contrasts_67fit}).
-The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
-
 <!--Display model selection results for contrasts data in a table-->
@@
 
-The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`.
-These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.
-
-These results do convincingly show non-Gaussian copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
-The next section looks at tail dependence in a non-parametric way.
-
-### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}
-
+<!--This chunk is for npa testing results of anolis dataset-->
 ```{r stat_contrasts_67_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_contrasts_67,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
@@
 
-![Nonparametric tests for tail dependence and other deviations from a normal copula for contrasts_67 data.
-As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
-\label{fig_BMR_nonparam}](./Results/Anolis_results/contrasts_csv_results/contrasts_67_bivfunctionplot.pdf)
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Anolis_results/contrasts_csv_results/contrasts_67_bivfunctionplot.pdf}
+\caption[Testing non-parametric statistics for anolis data]{Nonparametric tests for tail dependence and other deviations from a normal copula for anolis data.
+As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
+\label{fig_contrasts_67_nonparam}}
+\end{center}
+\end{figure}
+
+<!--
+{Nonparametric tests for tail dependence and other deviations from a normal copula for anolis data.
+As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for $1000$ normal-copula surrogate datasets (dashed lines show $0.025$ and $0.975$ quantiles of these results and dotted lines show $0.005$ and $0.995$ quantiles).
+\label{fig_contrasts_67_nonparam}}
+\end{center}
+\end{figure}
+-->
 
@@
 
+<!--
 \newpage
-
 ## Results: abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
-
 ### Model selection approach \label{Ex_count_Model_selection}
+### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
+                  ----moved to Paper.Rmd----
+-->
 
@@
 
-<!--Do the calculations-->
+<!--Do the model selection for aphid count data and then showing results in Tables-->
 ```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 saveRDS(RES_aphid_count,paste(resloc,file="AphidCopulaFit_selecloc_count_species_10.RDS",sep=""))
-```
-
-As described in Methods, model selection among copulas was done separately for each pair of the 10 
-sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns 
-corresponding to locations. Table \ref{tab_count_bestcop} shows that, for the large majority of location 
-pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of 
-question 1 for these data. Comparing the 
-AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{tab_count_AICw}) 
-to the AIC weight 
-of the normal copula family for the same pair (Table \ref{tab_count_AICw_normal}) confirms that the degree of support for normal copulas was usually not
-comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{tab_count_p_CvM}) in every case 
-failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
-the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{tab_count_p_KS}),
-demonstrating that the collection of copula families we used was sufficiently broad that the copulas of 
-all pairs of locations could be characterized by at least one of them.
-In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-
-```{r read_RES_aphid_count, echo=F}
-RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
 ```
@@
 
-The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{tab_count_LTdep_AICw} and \ref{tab_count_UTdep_AICw} respectively.
-These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data.
-Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.
-
 ```{r tab_count_LTdep_AICw, echo=F, results='markup'}
@@
 
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1.
-
 ```{r tab_count_LTmUT, echo=F, results='markup'}
@@
 
-### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
-
-<!--Do the main computations for this section-->
+<!--Do the main computations for non-parametric analysis of aphid count data-->
 ```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 ```
-
-```{r read_stat_aphid_count,echo=F}
-stat_aphid_count<-readRDS("./Results/stat_results/stat_aphid_count/stat_aphid_count_sp_10.RDS")
-```
-
-Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
-computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
-resampling scheme, as described in Methods (Table \ref{tab_count_resamp}). Confidence intervals of
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
-indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
-exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
-structure do not have, providing further evidence answering the third part of question 1.
 
@@
 
-Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance
-between the locations (Figure \ref{fig_aphid_count} A,B) showed the typical decline of correlation with 
-distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting 
-$\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of 
-locations shows these quantities also decline with distance (Figure \ref{fig_aphid_count} C-F). Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows
-these quantities increase with distance (as expected, given the opposite interpretation 
-these
-two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Figure \ref{fig_aphid_count} I-K)
-did not show a decline with distance, but did illustrate visually the statistical
-result of Table \ref{tab_count_resamp} that averages of these quantities across all pairs of sampling locations
-were significantly greater than zero, demonstrating asymmetry of tail dependence.
-Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in 
-the [`BIVAN`](https://github.com/sghosh89/BIVAN) package.
-
-\begin{figure}[!h]
-\textbf{ \hspace{2.6 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.8 cm} (D)} \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Corl_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Coru_vs_D.pdf}\\
-\end{center}
-\textbf{ \hspace{2.6 cm} (E) \hspace{3.6 cm} (F) \hspace{3.8 cm} (G) \hspace{3.8 cm} (H)} \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Pl_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Pu_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_D2u_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_D2l_vs_D.pdf}\\
-\end{center}
-\textbf{ \hspace{4.8 cm} (I) \hspace{3.8 cm} (J) \hspace{4 cm} (K) } \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
-\caption{Non-parametric tail statistics results for green spruce aphid abundance data plotted against geographical distance ($D$) between two sampling locations.
+\begin{figure}[!h]
+\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Spearman_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Kendall_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Corl_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Coru_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Pl_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Pu_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_D2u_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_D2l_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
+\caption[Non-parametric tail statistics results for green spruce aphid abundance data]{Non-parametric tail statistics results for green spruce aphid abundance data plotted against geographical distance ($D$) between two sampling locations.
 (A) Spearman correlation vs.
@@
 
+<!--
 \newpage
-
 ## Results: first flight date data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}
-
 ### Model selection approach \label{Ex_ff_Model_selection}
-
-<!--Do the analysis-->
+### Non-parametric analysis of tail dependence  \label{Ex_ff_NPA}
+                          ----moved to Paper.Rmd----
+-->
+
+<!--Do the model selection for aphid ff data and then showing results in Tables-->
 ```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 ```
-
-```{r read_RES_aphid_ff,echo=F}
-RES_aphid_ff<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_firstflight_species_11.RDS")
-```
-
-As in section \ref{Ex_count_Model_selection}, results are presented in $10 \times 10$ 
-matrix format with rows
-and columns corresponding to locations. Table \ref{tab_ff_bestcop} shows 
-that, for the large majority of location pairs, some non-normal copula 
-was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data. Comparing the AIC weight of the best-fitting 
-(lowest AIC) copula family for each pair
-of sampling locations (Table \ref{tab_ff_AICw}) to the AIC weight of the normal copula 
-family for the same pair (Table \ref{tab_ff_AICw_normal}) confirms that the
-degree of support for normal copulas was usually not comparable to the 
-degree of support for other families. Goodness of fit tests
-based on a Cramer-von Mises statistic (Table \ref{tab_ff_p_CvM}) usually failed to reject the 
-hypothesis that the AIC-best copula family
-was also an objectively adequate description of the data, and likewise for 
-goodness of fit tests based on a Kolmogorov-Smirnov
-statistic (Table \ref{tab_ff_p_KS}). In all cases at least `r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` percent 
-of bootstrap randomizations used for
-goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
 
@@
 
-The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and 
-the model-averaged (again, AIC based) upper-tail dependence statistic are in 
-Tables \ref{tab_ff_LTdep_AICw} and \ref{tab_ff_UTdep_AICw}, respectively.
-These values were exclusively positive, thereby differing from what a normal copula would give (i.e., $0$), and answering the second part of question 1 for these data - these data have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
-
 ```{r tab_ff_LTdep_AICw, echo=F, results='markup'}
@@
 
-The difference between the model-averaged lower-tail dependence statistic and the 
-model-averaged upper-tail dependence statistic (lower minus upper), shown in 
-Table \ref{tab_ff_LTmUT}, was negative for `r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)` 
-out of `r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair` location pairs, answering the third 
-part of question 1. We note this result is opposite the analogous result for aphid abundance
-data for the green spruce aphid, i.e., whereas those abundance data showed stronger lower tail 
-dependendence than upper tail dependence, first flights for the leaf-curling plum aphid showed 
-stronger upper tail dependence than lower tail dependence.
-
 ```{r tab_ff_LTmUT, echo=F, results='markup'}
@@
 
-### Non-parametric analysis of tail dependence  \label{Ex_ff_NPA}
-
-<!--Do the analysis-->
+<!--Do the main computations for non-parametric analysis of aphid ff data-->
 ```{r stat_aphid_ff, echo=F, cache=T,  cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 ```
-
-```{r read_stat_aphid_ff,echo=F}
-stat_aphid_ff<-readRDS("./Results/stat_results/stat_aphid_ff/stat_aphid_ff_sp_11.RDS")
-```
-
-Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman 
-correlations, were 
-computed across all pairs of sampling locations, and confidence intervals of these means were 
-computed via a spatial
-resampling scheme, as described in Methods (Table \ref{tab_ff_resamp}).
-Confidence intervals of
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded 
-zero and mean values were positive,
-indicating tail dependence.
-Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
-exluded $0$ and were negative, indicating asymmetry of tail dependence, a feature 
-that data with normal copula 
-structure do not have, providing further evidence answering the third part of question 1.
-The contrast here with the analogous results for abundances of the green spruce aphid is 
-again illuminating.
 
@@
 
-Plotting Spearman or Kendall correlations between first flight time series for pairs of locations as a function of great circle distance between the locations (Figure \ref{fig_aphid_ff} A, B) showed the typical (though slight, in this case) decline of correlation with increasing distance.
-Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear in most cases to decline slightly with distance (Figure \ref{fig_aphid_ff} C-F).
-$\Dsq_u$ and $\Dsq_l$ appear to slightly increase with distance, as expected given their
-opposite meaning (Figure \ref{fig_aphid_ff} G-H). 
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_aphid_ff} I-K) did not show a decline with distance, but did demonstrate visually the statistical result of Table \ref{tab_ff_resamp} that averages of these quantities across all pairs of sampling locations were significantly less than zero, showing asymmetry of tail dependence.
-
-\begin{figure}[!h]
-\textbf{ \hspace{2.6 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.8 cm} (D)} \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Coru_vs_D.pdf}\\
-\end{center}
-\textbf{ \hspace{2.6 cm} (E) \hspace{3.6 cm} (F) \hspace{3.8 cm} (G) \hspace{3.8 cm} (H)} \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pu_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2l_vs_D.pdf}\\
-\end{center}
-\textbf{ \hspace{4.8 cm} (I) \hspace{3.8 cm} (J) \hspace{4 cm} (K) } \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
-\caption{Non-parametric tail statistics results for leaf-curling plum aphid's first flight date data plotted against geographical distance $(D)$ between two sampling locations.
+\begin{figure}[!h]
+\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Spearman_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Kendall_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Coru_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Pu_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_D2l_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
+\caption[Non-parametric tail statistics results for leaf-curling plum aphid's first flight date data]{Non-parametric tail statistics results for leaf-curling plum aphid's first flight date data plotted against geographical distance $(D)$ between two sampling locations.
 (A) Spearman correlation vs.
@@
 
+<!--
 \newpage
-
 ## Results: abundance data for *Ceratium furca* in UK seas \label{Ex_pns}
-
 ### Model selection approach \label{Ex_pns_Model_selection}
+### Non-parametric analysis of tail dependence  \label{Ex_pns_NPA}
+                          ----moved to Paper.Rmd----
+-->
 
@@
 
-<!--Do the main analysis-->
+<!--Do the model selection for plankton data and then showing results in Tables-->
 ```{r RES_plankton_north_sea, echo=F, cache=T, results="hide", cache.extra=list(seed,data_plankton_north_sea,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 ```
-
-```{r read_RES_plankton_north_sea,echo=F}
-RES_plankton_north_sea<-readRDS("./Results/fitting_results/Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS")
-```
-
-As described in Methods and as in previous sections, model selection among copulas was done separately for 
-each pair of sampling locations. There were $14$ sampling locations, so results are presented in 
-$14 \times 14$ matrix format, with rows and columns corresponding to locations. 
-Table \ref{tab_pns_bestcop} shows that, for the large majority of location pairs, some 
-non-normal copula was a better fit, according to AIC, than was the normal copula, answering the 
-first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula 
-family for each pair of sampling locations (Table \ref{tab_pns_AICw}) to the AIC weight of
-the normal copula family for the same pair (Table \ref{tab_pns_AICw_normal}) 
-confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{tab_pns_p_CvM}) usually failed to reject the hypothesis that 
-the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a
-Kolmogorov-Smirnov statistic (Table \ref{tab_pns_p_KS}). In all cases at least
-`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit
-tests successfully lead to values of the statistic for comparison with its value on the real data.
 
@@
 
-The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{tab_pns_LTdep_AICw} and \ref{tab_pns_UTdep_AICw}, respectively.
-These values were exclusively positive, thereby
-differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
-have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
-
 ```{r tab_pns_LTdep_AICw,echo=F,results="markup"}
@@
 ```
-
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
-(lower minus upper), shown in Table \ref{tab_pns_LTmUT}, was positive for some
-location pairs and negative for others, but was positive for 
-`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`
-locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
 
@@
 
-### Non-parametric analysis of tail dependence  \label{Ex_pns_NPA}
+<!--Do the main computations for non-parametric analysis of plankton data-->
 ```{r stat_plankton_north_sea, echo=F, cache=T, cache=T,  cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 ```
-
-```{r read_stat_plankton_north_sea,echo=F}
-stat_plankton_north_sea<-readRDS("./Results/stat_results/stat_plankton_north_sea/stat_plankton_north_sea_sp_16.RDS")
-```
-
-Mean values of our statistics, as well as mean Kendall and Spearman
-correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
-spatial resampling scheme, as described in Methods (Table \ref{tab_pns_resamp}).
-Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
-always excluded $0$ and mean values were positive, indicating tail dependence.
-Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
-also all were strictly positive, indicating asymmetry of tail dependence, a feature that data with normal copula
-structure do not have, providing further evidence answering the third part of question 1.
 
@@
 
-Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
-between the locations (Figure \ref{fig_pns} A, B) showed the typical decline of correlation with increasing
-distance.
-Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Figure \ref{fig_pns} C-F).
-$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
-their opposite meaning (Figure \ref{fig_pns} G-H).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_pns} I-K) did not show a decline with distance, but did demonstrate visually the
-statistical result of Table \ref{tab_pns_resamp} that averages of these 
-quantities across all pairs of sampling locations were significantly greater 
-than zero, showing asymmetry of tail dependence.
-
-\begin{figure}[!h]
-\textbf{ \hspace{2.6 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.8 cm} (D)} \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Spearman_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Kendall_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Coru_vs_D.pdf}\\
-\end{center}
-\textbf{ \hspace{2.6 cm} (E) \hspace{3.6 cm} (F) \hspace{3.8 cm} (G) \hspace{3.8 cm} (H)} \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pu_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2l_vs_D.pdf}\\
-\end{center}
-\textbf{ \hspace{4.8 cm} (I) \hspace{3.8 cm} (J) \hspace{4 cm} (K) } \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf}
-\caption{Non-parametric tail statistics results for \textit{Ceratium furca} abundance data plotted against geographical distance $(D)$ between two sampling locations.
+\begin{figure}[!h]
+\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Spearman_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Kendall_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Coru_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pu_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2l_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf}
+\caption[Non-parametric tail statistics results for \textit{Ceratium furca} abundance data]{Non-parametric tail statistics results for \textit{Ceratium furca} abundance data plotted against geographical distance $(D)$ between two sampling locations.
 (A) Spearman correlation vs.
@@
 <!--***DAN: stopped here 2018 04 07-->
+
+<!--
 \newpage
-
 ## Results: methane flux data \label{Ex_methane}
-
 ### Model selection approach \label{Ex_methane_Model_selection}
+### Non-parametric analysis of tail dependence  \label{Ex_methane_NPA}
+                 ----moved to Paper.Rmd----
+-->
 
@@
 ```
-
-```{r read_RES_methane,echo=F}
-RES_methane<-readRDS("./Results/fitting_results/MethaneFluxFit_selecloc_data_pt_thrs_50.RDS")
-```
-
-As described in Methods and as in previous sections, model selection among copulas was done separately for 
-each pair of sampling locations. There were $13$ sampling locations, so results are presented in 
-$13 \times 13$ matrix format, with rows and columns corresponding to locations. 
-Table \ref{tab_methane_bestcop} shows that, for the large majority of location pairs, some 
-non-normal copula was a better fit, according to AIC, than was the normal copula, answering the 
-first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula 
-family for each pair of sampling locations (Table \ref{tab_methane_AICw}) to the AIC weight of
-the normal copula family for the same pair (Table \ref{tab_methane_AICw_normal}) 
-confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{tab_methane_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{tab_methane_p_KS}).
-In all cases at least
-`r min((RES_methane$gfc_numBS_success/RES_methane$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
 
@@
 
-The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{tab_methane_LTdep_AICw} and \ref{tab_methane_UTdep_AICw}, respectively.
-These values were exclusively positive, thereby
-differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
-have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
-
 ```{r tab_methane_LTdep_AICw,echo=F,results="markup"}
@@
 ```
-
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
-(lower minus upper), shown in Table \ref{tab_methane_LTmUT}, was positive for some
-location pairs and negative for others, but was positive for 
-`r sum((RES_methane$LTdep_AICw-RES_methane$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_methane$gfc_p_CvM)[1]^2-dim(RES_methane$gfc_p_CvM)[1])-2*RES_methane$num_indep_loc_pair`
-locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
 
@@
 
-### Non-parametric analysis of tail dependence  \label{Ex_methane_NPA}
+<!--Do the main computations for non-parametric analysis of methane-flux data-->
 ```{r stat_methane, echo=F, cache=T, cache=T,  cache.extra=list(seed,methane_data,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 ```
-
-Mean values of our statistics, as well as mean Kendall and Spearman
-correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
-spatial resampling scheme, as described in Methods (Table \ref{tab_methane_resamp}).
-Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
-always excluded $0$ and mean values were positive, indicating tail dependence.
-But the confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ were marginal, indicating no significant asymmetry of tail dependence.
 
@@
 
-Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
-between the locations (Figure \ref{fig_methane} A, B) showed the typical decline of correlation with increasing
-distance.
-Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Figure \ref{fig_methane} C-F).
-$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
-their opposite meaning (Figure \ref{fig_methane} G-H).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Figure \ref{fig_methane} I-K) did not show a decline with distance, but did demonstrate visually the
-statistical result of Table \ref{tab_methane_resamp} that averages of these 
-quantities across all pairs of sampling locations were more or less equally dispersed around the zero-line, showing no significant asymmetry of tail dependence.
-
-\begin{figure}[!h]
-\textbf{ \hspace{2.6 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.8 cm} (D)} \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Spearman_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Kendall_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Corl_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Coru_vs_D.pdf}\\
-\end{center}
-\textbf{ \hspace{2.6 cm} (E) \hspace{3.6 cm} (F) \hspace{3.8 cm} (G) \hspace{3.8 cm} (H)} \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Pl_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Pu_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_D2l_vs_D.pdf}\\
-\end{center}
-\textbf{ \hspace{4.8 cm} (I) \hspace{3.8 cm} (J) \hspace{4 cm} (K) } \\
-\begin{center}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Corl-Coru_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_Pl-Pu_vs_D.pdf}
-\includegraphics[width=1.8in]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u-D2l_vs_D.pdf}
-\caption{Non-parametric tail statistics results for \textit{Ceratium furca} abundance data plotted against geographical distance $(D)$ between two sampling locations.
+\begin{figure}[!h]
+\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.4 cm} (D)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Spearman_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Kendall_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Corl_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Coru_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{3.4 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.4 cm} (H)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Pl_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Pu_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_D2l_vs_D.pdf}\\
+\end{center}
+\textbf{ \hspace{5.4 cm} (I) \hspace{3.4 cm} (J) \hspace{3.2 cm} (K) } \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Corl-Coru_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_Pl-Pu_vs_D.pdf}
+\includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u-D2l_vs_D.pdf}
+\caption[Non-parametric tail statistics results for methane-flux data]{Non-parametric tail statistics results for methane-flux data plotted against geographical distance $(D)$ between two sampling locations.
 (A) Spearman correlation vs.
@@
 
-\newpage
-
+<!--
 # Question 2 analyses 
-Recall that question 2 is: What are some possible causes of non-Gaussian copula structure?
-
-Q1/2: How are copula structure of ecologically influential climate variables changing as part of climate change?
-We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the pops, but we don’t really know how a change in the environment cop structure would make what kind of change in the pop cop structure.
-How?
-
 ## The Moran effect and copula structure in populations 
-
-We show here that if a non-normal copula structure describes the dependence of an environmental
-variable between two locations, then there can be a non-normal copula structure for the
-dependence of populations in the locations that are influenced by the environmental variables.
-
-To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
-\ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.
-
-\begin{equation}\label{eq.AR1}
-P_{i}(t+1)=\beta P_{i}(t)+\alpha \epsilon_{i}(t)
-\end{equation}
-
-To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$.
-We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches.
-Should their dependence be exactly the same as of the noises or not ?
-
-To test that our approach is of two way : first, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
-Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome which was a bivariate population-copula, with last $500$ simulated points for each habitat patch (see Figure \ref{fig_Cause4copula_params_GOF_500}).
-We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size and reported p-values for both statistics (Cramer-von Mises and Kolmogorov-Smirnov).
-p-values greater or equal to $0.05$ (above the dashed line in Figure \ref{fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters.
-Figure \ref{fig_Cause4copula_params_GOF_2500} represents similar figure as of Figure \ref{fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$.
-(For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
-
-Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see Sec.
-\ref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
-Then we plotted mean of each measured quantity with standard error against Spearman's Rho (or Kendall's Tau) of the noise copula family.
-Figure \ref{fig_Cause4copula_C_spear} and Figure \ref{fig_Cause4copula_C_kend} represent this picture for noise generating from a lower tail dependent Clayton copula whereas for Figure \ref{fig_Cause4copula_SC_spear} and Figure \ref{fig_Cause4copula_SC_kend} noise came from a upper tail dependent Survival Clayton family.
-We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach, here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise.
-(For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
-
-<!--Shya to implement the modelling we discussed on 2018 04 12 and make results appear here!-->
+           ----moved to Paper.Rmd----
+-->
+
 ```{r cause4copula_pGOF_500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
-![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF_500}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf)
-
-![Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last **2500** points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF_2500}](./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf)
-
-![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_spear}](./Results/Cause4copula_results/Clayton_spearman.pdf)
-
-![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_C_kend}](./Results/Cause4copula_results/Clayton_kendall.pdf)
-
-![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Spearman's Rho** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_spear}](./Results/Cause4copula_results/SClayton_spearman.pdf)
-
-![Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from **Survival Clayton** copula and of the affected populations for 9 equidistant **Kendall's Tau** values in a range (0.1 to 0.9) (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}](./Results/Cause4copula_results/SClayton_kendall.pdf)
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf}
+\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{500} points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF_500}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf}
+\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF_2500}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Cause4copula_results/Clayton_spearman.pdf}
+\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_C_spear}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Cause4copula_results/Clayton_kendall.pdf}
+\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_C_kend}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_spearman.pdf}
+\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_SC_spear}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_kendall.pdf}
+\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}}
+\end{center}
+\end{figure}
+
+\newpage
 
@@
 
+<!--
 # Question 3 analyses for each dataset and theoretical analyses
 Recall that question 3 is: What are the consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?
-
 ## Skewness of the spatial mean
-
-Here, we claim that because of significant tail dependence of the multivariate copula structure, the skewness of the time-series data averaged over all selective locations changed significantly from the individual skewness of time-series data taken at each pairwise location.
-Spatial mean of the time series data extracts that tail dependence of the multivariate data set, if any, and reflected in the averaged skewness.
-We tested this with null hypothesis $H_{0}$.
-If $H_{0}$ is true that means our multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
-So, first we generated $10000$ number of surrogate normal copula from our given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
-Then we plotted the distribution of spatial mean of each surrogate time series data set and the given multivariate data.
-If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence.
-
-To verify that we use aphid-count data, aphid-firstflight data and plankton-north-sea data.
-For aphid data we choose only those locations for which atleast $30$ number of data points present in each time series when taken as pairwise locations.
-For plankton-north-sea data that threshold on number of data points is $45$.
-
 ### Aphid count \label{skewness_count}
-
-Figure \ref{fig_skewness_count} shows that averaged skewness is much lower than expected if the aphid count data came from a normal copula.
-Here, $p$ value indicates the fraction of surrogate-skewness which are lower than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_count}).
-As $p<0.05$, we reject $H_{0}$.
-This result is consistent with the non-parametric analysis and model fitting results which indicated lower tail dependence.
+### Aphid first flight \label{skewness_ff}
+### UK Sea plankton \label{skewness_pns}
+### Methane-flux data \label{skewness_methane}
+                   ----moved to Paper.Rmd----
+-->
 
@@
 
-![Skewness_count for species : Green spruce aphid\label{fig_skewness_count}](./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf){ width=40%, height=40% }
-
-### Aphid first flight \label{skewness_ff}
-
-Figure \ref{fig_skewness_ff} shows that averaged skewness is much higher than expected if the aphid first flight data came from a normal copula.
-Here, $p$ value indicates the fraction of surrogate-skewness which are greater than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_ff}).
-As $p<0.05$, we reject $H_{0}$ in support for upper tail dependence.
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf}
+\caption[Skewness count for species Green spruce aphid]{Skewness count for species Green spruce aphid\label{fig_skewness_count}}
+\end{center}
+\end{figure}
 
@@
 
-![Skewness_firstflight for species : Leaf-curling plum aphid\label{fig_skewness_ff}](./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf){ width=40%, height=40% }
-
-### North Sea plankton \label{skewness_pns}
-
-Figure \ref{fig_skewness_pns} shows that averaged skewness is lower than expected if the plankton north sea data came from a normal copula.
-Our previous results on non-parametric analysis and model fitting showed lower tail dependence.
-Here, as $p>0.05$, we cannot reject $H_{0}$, .......???
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf}
+\caption[Skewness firstflight for species Leaf-curling plum aphid]{Skewness firstflight for species Leaf-curling plum aphid\label{fig_skewness_ff}}
+\end{center}
+\end{figure}
 
@@
 
-![Skewness_plankton_north_sea for species : Ceratium_furca\label{fig_skewness_pns}](./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf){ width=40%, height=40% }
-
-### Methane-flux data \label{skewness_methane}
-
-Figure \ref{fig_skewness_methane} shows that averaged skewness is lower than expected if the plankton north sea data came from a normal copula.
-Our previous results on non-parametric analysis and model fitting showed lower tail dependence.
-As $p<0.05$, we reject $H_{0}$.
-This result is consistent with the non-parametric analysis and model fitting results which indicated lower tail dependence.
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf}
+\caption[Skewness-plankton for species Ceratium furca]{Skewness-plankton for species Ceratium furca\label{fig_skewness_pns}}
+\end{center}
+\end{figure}
 
@@
 
-![Skewness_methane-flux data\label{fig_skewness_methane}](./Results/skewness_results/skewness_methane/skewness_methane.pdf){ width=40%, height=40% }
-
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane.pdf}
+\caption[Skewness-methane-flux data]{Skewness-methane-flux data\label{fig_skewness_methane}}
+\end{center}
+\end{figure}
+
+<!--
 ## Extinction risk
+        ----moved to Paper.Rmd----
+-->
 
@@
 
-### LC model (r=0) :
-
-  * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
-[see first plot of Figure \ref{fig_risk_local_disp} and Figure \ref{fig_risk_global_disp}]
-
-### Ricker model (r $\neq$ 0):
-
- *  For ricker model when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
-For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
-[see Figure \ref{fig_risk_local_disp} - \ref{fig_risk_global_disp}]
-
-  * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]
-
-With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Figure \ref{fig_risk_numlocs_effect_local_disp} - Figure \ref{fig_risk_numlocs_effect_global_disp}] for both model.
-
 ![Extinction_Risk_local_dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp}](./Results/ext_risk_local_disp.pdf){ width=40%, height=40% }
@@
 
+<!-- Taylor's law section need to move to Paper.Rmd once it's completed-->
 ## Taylor's law
@@
 # References
+<div id="refs"></div>

### 2018-05-03 Paper.Rmd
making another plot for main text

@@
 updated on `r Sys.Date()`
-
+```{r setup_Paper, echo=F}
+knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
+seed<-101
+# families<-c(1,3:10,13,14,16:20)
+source("mtime.R") #A function needed for caching
+```
 # Front matter
@@
 fall (these are the blue and red lines, respectively, in 
-Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
+Fig. \ref{fig_stat_image}(A)). Our statistics quantify the dependence
 between $u_i$ and $v_i$ in the region bounded by these lines.
@@
 
-![Schematic diagram supporting defintions of nonparametric tail dependence statistics.\label{fig_stat_image}](./stat_image.jpg){ width=25%, height=25% }
+```{r fig_stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"),mtime("StatTestNumericResults.RData"))}
+set.seed(seed) 
+source("fig_Paper_stat_testing.R")
+load("./Results/stat_results/stat_testing/StatTestNumericResults.RData")
+resultsloc<-"./Results/stat_results/stat_testing/"
+
+copname<-"C"
+reslist<-res_pt35_rankT
+
+statname<-"CorlmCoru"
+call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
+statname<-"PlmPu"
+call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
+statname<-"D2umD2l"
+call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
+```
+
+\begin{figure}[!h]
+\textbf{ \hspace{6 cm} (A) \hspace{5.6 cm} (B)} \\
+\vspace{-0.6 cm}
+\begin{center}
+\hspace{0.5 cm}
+\includegraphics[width=5 cm,height=4.3 cm]{./stat_image.jpg} \hspace{0.6 cm}
+\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/CorlmCoru_C.pdf}
+\end{center}
+\textbf{ \hspace{6 cm} (C) \hspace{5.7 cm} (D)} \\
+\vspace{-0.6 cm}
+\begin{center}
+\hspace{0.2 cm}
+\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/PlmPu_C.pdf} \hspace{0.3 cm}
+\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/D2umD2l_C.pdf}
+\caption{(A) Schematic diagram supporting defintions of nonparametric tail dependence statistics; (B) Results of tests of $\cor_l-\cor_u$, (C) $\Ps_l-\Ps_u$, and (D) $\Dsq_u-\Dsq_l$ reveal the known asymmetry (more lower tail than upper) of tail dependence of the Clayton copula with $n=35$ data points.
+\label{fig_stat_image}}
+\end{center}
+\end{figure}
 
@@
 In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we 
-also tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and 
+also tested our statistics in Fig. \ref{fig_stat_image}(B - D) by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and 
 $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated
-from various copulas (for details see Appendix \ref{SM-Test_npa_stat}).
+from Clayton copulas (for details see Appendix \ref{SM-Test_npa_stat}).
 
 For bivariate datasets such as the soil C and N data, we compared values of the statistics $\cor_{l_b,u_b}$,
-$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to distributions of values of the same 
-statistics computed on surrogate datasets that 
-were produced from the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
-the same marginal distributions as the empirical data and had very similar Kendall or Spearman correlation, but had
-normal copula structure, and thus our comparisons tested the null hypothesis that our statistics took values on the
-empirical data no different from what would have been expected if the copula structure of the data were normal, but the
-data were otherwise statistically unchanged.
+$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to distributions of values of the same statistics computed on surrogate datasets that were produced from the empirical data by randomizing it in a special way.
+The surrogate/randomized datasets had exactly the same marginal distributions as the empirical data and had very similar Kendall or Spearman correlation, but had normal copula structure, and thus our comparisons tested the null hypothesis that our statistics took values on the empirical data no different from what would have been expected if the copula structure of the data were normal, but the data were otherwise statistically unchanged.
 For bivariate datasets, surrogates were produced as follows.

### 2018-05-03 fig_Paper_stat_testing.R

@@ -1,0 +1,98 @@
+#--------------------------------------------
+source("./TestStats_multivar_dataset.R")
+#--------------------------------------------
+
+# This function make stat_testing plot for a specified copula with a specific statistic at a time
+# It is actually the simpler version of "plotter_stat_testing" function from TestStats_multivar_dataset.R
+
+call_fig_Paper_stat_testing<-function(resultsloc,statname,copname,reslist,numsims=100){
+  
+  pdf(paste(resultsloc,statname,"_",copname,".pdf",sep=""),width=4, height=3)
+  a<-M1mM2(reslist=res_pt35_rankT,numsims)
+  spearvals<-seq(from=0,to=0.9,by=0.1)
+  callfn<-coplist_with_params(spearvals=spearvals)
+  b<-callfn$paramlist
+  
+  t1<-c(1,length(b[[1]])+1,length(b[[1]])+length(b[[2]])+1,length(b[[1]])+length(b[[2]])+length(b[[3]])+1)
+  t2<-c(length(b[[1]]),length(b[[1]])+length(b[[2]]),length(b[[1]])+length(b[[2]])+length(b[[3]]),length(b[[1]])+length(b[[2]])+length(b[[3]])+length(b[[4]]))
+  
+  
+  if(copname=="C"){
+    i<-1
+  }else if(copname=="N"){
+    i<-2
+  }else if(copname=="F"){
+    i<-3
+  }else if(copname=="SC"){
+    i<-4
+  }else{
+    i<-NA
+    warning("specify copname",immediate.=T,call.=T)
+  }
+  
+  is<-t1[i]
+  ie<-t2[i]
+  
+  t_CorlmCoru<-c()
+  t_PlmPu<-c()
+  t_D2umD2l<-c()
+  
+  for(ispearval in 1:length(spearvals)){
+    x1<-ttest(reslist,numsims,j1=1,j2=2,is,ie,ispearval)
+    x2<-ttest(reslist,numsims,j1=3,j2=4,is,ie,ispearval)
+    x3<-ttest(reslist,numsims,j1=6,j2=5,is,ie,ispearval)
+    
+    t_CorlmCoru<-c(t_CorlmCoru,x1)
+    t_PlmPu<-c(t_PlmPu,x2)
+    t_D2umD2l<-c(t_D2umD2l,x3)
+  }
+  
+  op<-par(mar=c(3.5,4,1,3.5),mgp=c(2.2,1,0))
+  
[…]

### 2018-05-03 Paper.Rmd
modified

@@
 
-In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we 
-also tested our statistics in Fig. \ref{fig_stat_image}(B - D) by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and 
-$\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated
-from Clayton copulas (for details see Appendix \ref{SM-Test_npa_stat}).
+In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we also tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas.
+Fig.
+\ref{fig_stat_image}(B - D) represent the results for Clayton copulas with $n=35$ data points and $100$ simulations (for details see Appendix \ref{SM-Test_npa_stat}).
 

### 2018-05-03 SupportingInformation.Rmd
modified

@@
 
+\newpage
 # Descriptions of data \label{Description_data}
@@
 
+\newpage
+
 <!--Display model selection results in a table-->
-```{r tab_soilCNfit,echo=F,results='markup'}
+```{r tab_soilCNfit,echo=F}
+knitr::opts_knit$set(kable.force.latex = TRUE)
 h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-knitr::kable(h,
-             format='pandoc',
+knitr::kable(h, longtable = T, booktabs = T, 
+             format = "latex",
+      caption.short = "Fitting info for soil C and N dataset",
       caption = "Fitting info for soil C and N dataset.
 Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=survival Clayton; SG=survival Gumbel; SJ=survival Joe.
-An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by $180^\\circ$).\\label{tab_soilCNfit}",
-             booktabs=T)
+An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by $180^\\circ$).\\label{tab_soilCNfit}")
 ```
@@
 
-```{r tab_soilCN_asym, echo=F, results="markup"}
+```{r tab_soilCN_asym,echo=F}
+knitr::opts_knit$set(kable.force.latex = TRUE)
 tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
-knitr::kable(tab_soilCN_asym,
-             format="pandoc",
+knitr::kable(tab_soilCN_asym, booktabs = T, 
+             format = "latex",
+      caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
       caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.
@@
 A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.
-Recall that Spearman-preserving surrogates are more appropriate for the $\\cor$ statistic (Figure \\ref{fig_soilCN_nonparam} caption).\\label{tab_soilCN_asym}",
-             booktabs=T)
+Recall that Spearman-preserving surrogates are more appropriate for the $\\cor$ statistic (Figure \\ref{fig_soilCN_nonparam} caption).\\label{tab_soilCN_asym}")
+#kable_styling(, position = "float_left")
 ```
@@
 
+\newpage
 <!--Display model selection results for BMR data in a table-->
@@
 
-```{r tab_count_bestcop, echo=F, results='markup'}
+```{r tab_count_bestcop, echo=F}
 #library(dplyr)
-knitr::kable(RES_aphid_count$info_ord_copname[,,1], 
-             format='pandoc',
+knitr::opts_knit$set(kable.force.latex = TRUE)
+knitr::kable(RES_aphid_count$info_ord_copname[,,1], longtable = T,
+             format='latex',
+             caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid.",
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid. Copulas 
@@
 survival Joe, survival BB1, survival BB6, survival BB7, survival BB8 copulas.
-Abbreviations were given in Methods.
+Abbreviations were given in \\nameref{MT-Methods}.
 NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
@@
 
-```{r tab_ff_bestcop, echo=F, results='markup'}
-knitr::kable(RES_aphid_ff$info_ord_copname[,,1], 
-             format='pandoc',
+```{r tab_ff_bestcop, echo=F}
+knitr::kable(RES_aphid_ff$info_ord_copname[,,1], longtable = T,
+             format='latex',
+             caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.
-Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations given in Methods.
+Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations given in \\nameref{MT-Methods}.
 NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.
@@
 
-```{r tab_pns_bestcop, echo=F, results='markup'}
-knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1],
-             format='pandoc',
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundance data for *Ceratium furca* in UK seas.
-See previous sections for copula abbreviations.
+```{r tab_pns_bestcop, echo=F}
+knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1], longtable=T,
+             format='latex', 
+             caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundance data for \\textit{Ceratium furca} in UK seas.",
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundance data for \\textit{Ceratium furca} in UK seas.
+See \\nameref{MT-Methods} for copula abbreviations.
 NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_pns_bestcop}",
@@
 
-```{r tab_methane_bestcop, echo=F, results='markup'}
-knitr::kable(RES_methane$info_ord_copname[,,1],
-             format='pandoc',
+```{r tab_methane_bestcop, echo=F}
+knitr::kable(RES_methane$info_ord_copname[,,1], longtable=T,
+             format='latex',
+             caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.",
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.
-See previous sections for copula abbreviations.
+See \\nameref{MT-Methods} for copula abbreviations.
 NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_methane_bestcop}",

### 2018-05-04 SupportingInformation.Rmd
small formatting change

@@
  #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
- atdres<-data.frame(Statistic=c("$\\cor_{first.slice}-\\cor_{last.slice}$",
+ atdres<-data.frame(Statistic=c("$\cor_{\text{first slice}}-\cor_{\text{last slice}}$",
                                "$\\Ps_{first.slice}-\\Ps_{last.slice}$",

### 2018-05-04 SupportingInformation.Rmd
fixing table as pandoc and \text{} format

--- 2018-05-02 SupportingInformation.Rmd
@@
 <!--Display model selection results in a table-->
-```{r tab_soilCNfit,echo=F,results='markup'}
+```{r tab_soilCNfit, echo=F, results="markup"}
+#knitr::opts_knit$set(kable.force.latex = TRUE)
 h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-knitr::kable(h,
-             format='pandoc',
+knitr::kable(h, #longtable = T, 
+             format = "pandoc",
+      #caption.short = "Fitting info for soil C and N dataset",
       caption = "Fitting info for soil C and N dataset.
@@
 An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by $180^\\circ$).\\label{tab_soilCNfit}",
-             booktabs=T)
+      booktabs = TRUE)
 ```
@@
  #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
- atdres<-data.frame(Statistic=c("$\\cor_{first.slice}-\\cor_{last.slice}$",
-                               "$\\Ps_{first.slice}-\\Ps_{last.slice}$",
-                                "$\\Dsq_{last.slice}-\\Dsq_{first.slice}$"),
+ atdres<-data.frame(Statistic=c("$\\cor_\\text{first slice}-\\cor_\\text{last slice}$",
+                               "$\\Ps_\\text{first slice}-\\Ps_\\text{last slice}$",
+                                "$\\Dsq_\\text{last slice}-\\Dsq_\\text{first slice}$"),
                    Kendall=rep("",3),
@@
 
-```{r tab_soilCN_asym, echo=F, results="markup"}
+```{r tab_soilCN_asym,echo=F,results="markup"}
+#knitr::opts_knit$set(kable.force.latex = TRUE)
 tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
-knitr::kable(tab_soilCN_asym,
-             format="pandoc",
+knitr::kable(tab_soilCN_asym, booktabs = T, 
+             format = "pandoc",
+      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
       caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.
@@
 
+\newpage
 <!--Display model selection results for contrasts data in a table-->
@@
 
-```{r tab_count_bestcop, echo=F, results='markup'}
+```{r tab_count_bestcop, echo=F, results="markup"}
 #library(dplyr)
-knitr::kable(RES_aphid_count$info_ord_copname[,,1], 
+#knitr::opts_knit$set(kable.force.latex = TRUE)
+knitr::kable(RES_aphid_count$info_ord_copname[,,1], #longtable = T,
              format='pandoc',
+             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid.",
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid. Copulas 
@@
 
-```{r tab_ff_bestcop, echo=F, results='markup'}
-knitr::kable(RES_aphid_ff$info_ord_copname[,,1], 
+```{r tab_ff_bestcop, echo=F, results="markup"}
+knitr::kable(RES_aphid_ff$info_ord_copname[,,1], #longtable = T,
              format='pandoc',
+             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.
-Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations given in Methods.
+Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations given in \\nameref{MT-Methods}.
 NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.
@@
 
-```{r tab_pns_bestcop, echo=F, results='markup'}
-knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1],
+```{r tab_pns_bestcop, echo=F, results="markup"}
+knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1], #longtable=T,
              format='pandoc', 
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundance data for *Ceratium furca* in UK seas.
-See previous sections for copula abbreviations.
+             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundance data for \\textit{Ceratium furca} in UK seas.",
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundance data for \\textit{Ceratium furca} in UK seas.
+See \\nameref{MT-Methods} for copula abbreviations.
 NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_pns_bestcop}",
@@
 
-```{r tab_methane_bestcop, echo=F, results='markup'}
-knitr::kable(RES_methane$info_ord_copname[,,1],
+```{r tab_methane_bestcop, echo=F, results="markup"}
+knitr::kable(RES_methane$info_ord_copname[,,1], #longtable=T,
              format='pandoc',
+             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.",
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.
-See previous sections for copula abbreviations.
+See \\nameref{MT-Methods} for copula abbreviations.
 NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_methane_bestcop}",

### 2018-05-06 SupportingInformation.Rmd
adding beta variation chunk for asym sensitivity analysis and some minor changes

@@
 #-----------------
-set.seed(seed=101)
+set.seed(seed)
 #---------------------------------------
@@
 <!--Another model, distinct from the above, Dan and Shya to discuss-->
+```{r asym_sens_beta_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+set.seed(seed)
+source("asym_sensitivity.R")
+
+#-----------------------------------------------------------------------------------
+# This function takes 
+# Input :
+#       N : number of global and local noise points you want
+#       rho : covariance of global noise matrix
+#       alpha : a number : local noise coeeficient
+#       beta : growth factor of pop
+#       decrease : logical (controls the nature of global noise)
+#       numkeep_last : number of points you want to keep for the pop matrix from the end
+#       ploton : logical for optional plot
+
+# Output :
+#     a list of 4 pop matrices 
+#                         1. (for beta=+ve, decrease=F)
+#                         2. (for beta=+ve, decrease=T)
+#                         3. (for beta=-ve, decrease=F)
+#                         4. (for beta=-ve, decrease=T)
+
+plotter_asym_sens_beta_varn<-function(N,rho,alpha,beta,numkeep_last,ploton){
+
+  pop<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
+  pop_d<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
+
+  neg_beta<- -beta
+  pop_neg_beta<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
+  pop_d_neg_beta<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
+
+  if(ploton==T){
+    op<-par(mfrow=c(2,2),mar=c(4,3,1.5,1),mgp=c(2,1,0))
+
+    plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
+    mtext(paste0("(A)"),side = 3, line=0.3, adj=0.3, col="black")
+    plot(pop_d[,1],pop_d[,2],col="blue",main = bquote(beta ==.(beta)))
+    mtext(paste0("(B)"),side = 3, line=0.3, adj=0.3, col="black")
+    plot(pop_neg_beta[,1],pop_neg_beta[,2],col="orange",main = bquote(beta ==.(neg_beta)))
+    mtext(paste0("(C)"),side = 3, line=0.3, adj=0.3, col="black")
+    plot(pop_d_neg_beta[,1],pop_d_neg_beta[,2],col="skyblue",main = bquote(beta ==.(neg_beta)))
+    mtext(paste0("(D)"),side = 3, line=0.3, adj=0.3, col="black")
+
+    par(op)
+  }
+
+  return(list(pop=pop,
+              pop_d=pop_d,
+              pop_neg_beta=pop_neg_beta,
+              pop_d_neg_beta=pop_d_neg_beta))
+
+}
+#--------------------------------------------------------------
+N<-10000
+rho<-0.8
+alpha<-0.2
+numkeep_last<-1000
+
+pdf("./Results/asym_sens_results/beta_0.1_asym_sen.pdf")
+ans_beta_0.1<-plotter_asym_sens_beta_varn(N=N,rho=rho,alpha=alpha,beta=0.1,numkeep_last=numkeep_last,ploton=T)
+dev.off()
+
+pdf("./Results/asym_sens_results/beta_0.9_asym_sens.pdf")
+ans_beta_0.9<-plotter_asym_sens_beta_varn(N=N,rho=rho,alpha=alpha,beta=0.9,numkeep_last=numkeep_last,ploton=T)
+dev.off()
+
+vlist_beta_0.1_beta_0.9<-c(ans_beta_0.1,ans_beta_0.9) # a list of 8 matrices
+
+summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
+rownames(summary_stat)<-c("bestcop_AICw","N_AIC_w","relLTdep_AICw","relUTdep_AIC_w","p_CvM","p_KS","Numboot_success")
+bg <- intToUtf8(0x03B2) 
+cname<-paste0(bg, c("_0.1","_0.1_decrease_T","_-0.1","_-0.1_decrease_T","_0.9","_0.9_decrease_T","_-0.9","_-0.9_decrease_T"))
+colnames(summary_stat)<-cname
+
+stat_npa_asym_beta_0.1_beta_0.9<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible beta variation
+
+for(ic in c(1:8)){
+  v<-vlist_beta_0.1_beta_0.9[[ic]]
+  res<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=c(1,3:10,13:14,16:20),level=0.05,
+                      AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
+                      gofnormal=F,status=F)
+
+  summary_stat[1,ic]<-max(res$InfCritRes$AICw)
+  summary_stat[2,ic]<-res$InfCritRes$AICw[1]
+  summary_stat[3,ic]<-res$relLTdep_AICw
+  summary_stat[4,ic]<-res$relUTdep_AICw
+  summary_stat[5,ic]<-res$GofRes_CvM
+  summary_stat[6,ic]<-res$GofRes_KS
+  summary_stat[7,ic]<-res$Numboot_success
+
+  stat_npa_asym_beta_0.1_beta_0.9[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cname[ic],numbin=10)
+
+}
+
+saveRDS(summary_stat,"./Results/asym_sens_results/beta_variation_summary_stat.RDS")
+saveRDS(stat_npa_asym_beta_0.1_beta_0.9,"./Results/asym_sens_results/stat_npa_asym_beta_0.1_beta_0.9.RDS")
+```
 

### 2018-05-07 SupportingInformation.Rmd
addingfig and tables from beta variation results : asym sensitivity

@@
 
-pdf("./Results/asym_sens_results/beta_0.1_asym_sen.pdf")
+pdf("./Results/asym_sens_results/beta_0.1_asym_sens.pdf")
 ans_beta_0.1<-plotter_asym_sens_beta_varn(N=N,rho=rho,alpha=alpha,beta=0.1,numkeep_last=numkeep_last,ploton=T)
@@
 bg <- intToUtf8(0x03B2) 
-cname<-paste0(bg, c("_0.1","_0.1_decrease_T","_-0.1","_-0.1_decrease_T","_0.9","_0.9_decrease_T","_-0.9","_-0.9_decrease_T"))
+cnm<-c("_0.1","_0.1_decrease_T","_-0.1","_-0.1_decrease_T","_0.9","_0.9_decrease_T","_-0.9","_-0.9_decrease_T")
+cname<-paste0(bg,cnm)
 colnames(summary_stat)<-cname
@@
 stat_npa_asym_beta_0.1_beta_0.9<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible beta variation
+names(stat_npa_asym_beta_0.1_beta_0.9)<-cname
 
@@
 saveRDS(stat_npa_asym_beta_0.1_beta_0.9,"./Results/asym_sens_results/stat_npa_asym_beta_0.1_beta_0.9.RDS")
+```
+
+![results for model selection \label{fig_ms_asym_beta_0.1}](./Results/asym_sens_results/beta_0.1_asym_sens.pdf){ width=40%, height=40% }
+
+![results for model selection \label{fig_ms_asym_beta_0.9}](./Results/asym_sens_results/beta_0.9_asym_sens.pdf){ width=40%, height=40% }
+
+```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
+
+knitr::kable(summary_stat, booktabs = T, 
+             format = "pandoc",
+      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
+      caption="summary of model selection results for asym. sensitivity \\label{tab_summary_stat_asym_sens}")
+```
+
+```{r tab_stat_npa_asym,echo=F,results="asis"}
+#knitr::opts_knit$set(kable.force.latex = TRUE)
+tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_beta_0.1_beta_0.9.RDS")
+
+for(i in c(1:8)){
+  temp<-make_table_stat_fn(data=tab_asym[[i]])
+  temp2<-knitr::kable(temp, booktabs = T, 
+             format = "pandoc",
+      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
+      caption =paste("beta",cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
+  print(temp2)
+  cat("\\newpage")
+}
 ```

### 2018-05-07 SupportingInformation.Rmd
temporary solution for unicode character in table_summary_stat_model selection results for asym sensitivity

@@
 ```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
-
+cnm2<-c("b_0.1","b_0.1d_T","b_-0.1","b_-0.1d_T","b_0.9","b_0.9d_T","b_-0.9","b_-0.9d_T")
+colnames(summary_stat)<-cnm2
 knitr::kable(summary_stat, booktabs = T, 
              format = "pandoc",
+             digits=3,
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
@@
 
+\newpage
 ```{r tab_stat_npa_asym,echo=F,results="asis"}

### 2018-05-08 SupportingInformation.Rmd
description of BMR data

@@
 
-**bmr-contrasts.csv** : Prof. M. T. Holder to add description here about this dataset.
+**bmr-contrasts.csv** : This contains the trait contrasts
+from phylogenetically independent contrasts [@SukumaranH2010] 
+of data reported by [@GenoudIM2017] on the basal metabolic rate (BMR)
+and mass of mammals species.
+Specifically, the "SELECT" BMR data in their database were used for
+ those species for which data were available.
+A phylogeny in NEXUS [@NEXUS] format was extracted from the 
+ supplementary information of [@GenoudIM2017].
+Version 4.3 of the DendroPy library [@SukumaranH2010] was used to calculate 
+  phylogenetic independent contrasts.
+For each contrast, the path length spanned by the contrast (sum of the branch
+  lengths) and the contrast for the mass, BMR and natural
+  logarithm of the mass were written as a row in the bmr-contrasts.csv file.
 

### 2018-05-08 SupportingInformation.Rmd
notes on Anolis and regenerated data from DendroPy 4.4.0

@@
  supplementary information of [@GenoudIM2017].
-Version 4.3 of the DendroPy library [@SukumaranH2010] was used to calculate 
-  phylogenetic independent contrasts.
+Version 4.4.0 of the DendroPy library [@SukumaranH2010] was used to calculate 
+  phylogenetic independent contrasts using Python 2.7.
 For each contrast, the path length spanned by the contrast (sum of the branch
@@
 
-**contrasts.csv** : Prof. M. T. Holder to add description here about this dataset.
+**contrasts.csv** : Raw data on measurements of anatomical
+features of several species of *Anolis* lizards was downloaded
+from the Dryad data package [@MahlerIRL2013Data] associated with
+[@MahlerIRL2013].
+The contrasts.csv file was produced by using version 4.4.0 of the DendroPy 
+  library [@SukumaranH2010]  to calculate 
+  phylogenetic independent contrasts on the the data in the GA_Anolis_traits.csv
+  file of Mahler et at. using their maximum clade credibility tree (GA_Anolis_MCC.nex)
+  as the phylogeny.
 

### 2018-05-09 SupportingInformation.Rmd
rearranging figure
calling noise mat outside the function and then passing as input

@@
 set.seed(seed)
+#-----------------------------------------------------------------------------------
+source("OurBiCopSelect.R")
+source("bivfunctionplot.R")
 source("asym_sensitivity.R")
-
 #-----------------------------------------------------------------------------------
@@
 # Input :
-#       N : number of global and local noise points you want
-#       rho : covariance of global noise matrix
+#       mg : global noise matrix (dim = N by 2)
+#       ml : local noise matrix (dim = N by 2)
 #       alpha : a number : local noise coeeficient
@@
 
-plotter_asym_sens_beta_varn<-function(N,rho,alpha,beta,numkeep_last,ploton){
-
-  pop<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
-  pop_d<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
+plotter_asym_sens_beta_varn<-function(mg,ml,alpha,beta,numkeep_last,ploton){
+
+  pop<-Simulator_asym_sens(mg=mg,ml=ml,alpha=alpha,beta=beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
+  pop_d<-Simulator_asym_sens(mg=mg,ml=ml,alpha=alpha,beta=beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
 
   neg_beta<- -beta
-  pop_neg_beta<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
-  pop_d_neg_beta<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
+  pop_neg_beta<-Simulator_asym_sens(mg=mg,ml=ml,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
+  pop_d_neg_beta<-Simulator_asym_sens(mg=mg,ml=ml,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
 
@@
 }
-#--------------------------------------------------------------
+#----------------------------------------------------------
 N<-10000
@@
 numkeep_last<-1000
+temp<-get_noise_glob_loc_mat(N=N,rho=rho)
+mg<-temp$mg
+ml<-temp$ml
 
 pdf("./Results/asym_sens_results/beta_0.1_asym_sens.pdf")
-ans_beta_0.1<-plotter_asym_sens_beta_varn(N=N,rho=rho,alpha=alpha,beta=0.1,numkeep_last=numkeep_last,ploton=T)
+ans_beta_0.1<-plotter_asym_sens_beta_varn(mg=mg,ml=ml,alpha=alpha,beta=0.1,numkeep_last=numkeep_last,ploton=T)
 dev.off()
@@
 pdf("./Results/asym_sens_results/beta_0.9_asym_sens.pdf")
-ans_beta_0.9<-plotter_asym_sens_beta_varn(N=N,rho=rho,alpha=alpha,beta=0.9,numkeep_last=numkeep_last,ploton=T)
+ans_beta_0.9<-plotter_asym_sens_beta_varn(mg=mg,ml=ml,alpha=alpha,beta=0.9,numkeep_last=numkeep_last,ploton=T)
 dev.off()
@@
 \newpage
+
 # References

### 2018-05-09 asym_sensitivity.R

@@ -2,9 +2,6 @@
 library(VineCopula)
 library(mvtnorm)
 #----------------------------
-source("OurBiCopSelect.R")
-source("bivfunctionplot.R")
-#----------------------------
 # This function takes a vector (x) and a logical tag and then transform them accordingly
 glob_noise_fn<-function(x,decrease){
   
@@ -45,10 +42,31 @@ glob_noise_fn<-function(x,decrease){
 #g_noise_dec<-apply(mg,MARGIN = 2,FUN = glob_noise_fn, decrease=T)
 #h_noise<-alpha*ml
 #-----------------------------------------------------------------------
+# This function generates global and local noise matrices 
+#            which will be called as input in next simulator function
+# Input:      
+#       N : number of global and local noise points you want
+#       rho : covariance of global noise matrix
+# Output:
+#       a list of two matrices (mg for global noise mat) and (ml for local noise matrix)
+get_noise_glob_loc_mat<-function(N,rho){
+  
+  sigma_glob<-matrix(c(1,rho,rho,1),2,2)
+  sigma_loc<-matrix(c(1,0,0,1),2,2)
+  
+  mg<-rmvnorm(N,mean=c(0,0),sigma=sigma_glob)
+  ml<-rmvnorm(N,mean=c(0,0),sigma=sigma_loc)
+  
+  return(list(mg=mg,
+              ml=ml))
+  
+}
+#----------------------------------------------------------------------------------
 # This function simulates the equation to test asym sensitivity :
 #      P(i,t+1)=beta*P(i,t)+globalnoise(i,t)+alpha*localnoise(i,t)
 # Input :
-#       N : number of global and local noise points you want
+#       mg : global noise matrix (dim = N by 2)
+#       ml : local noise matrix (dim = N by 2)
 #       rho : covariance of global noise matrix
 #       alpha : a number : local noise coeeficient
 #       beta : growth factor of pop
@@ -57,14 +75,9 @@ glob_noise_fn<-function(x,decrease){
 # Output :
 #     pop matrix with dim. = numkeep_last by 2
 
-Simulator_asym_sens<-function(N,rho,alpha,beta,p0=c(0,0),decrease,numkeep_last){
[…]

### 2018-05-12 SupportingInformation.Rmd
saving model selection results for asym sensitivity

@@
 
+ms_asym_beta_0.1_beta_0.9<-vector(mode = "list", length = 8) # initializing list to store model selection results for all possible beta variation
+names(ms_asym_beta_0.1_beta_0.9)<-cname
+
 stat_npa_asym_beta_0.1_beta_0.9<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible beta variation
@@
                       gofnormal=F,status=F)
+
+  ms_asym_beta_0.1_beta_0.9[[ic]]<-res
 
@@
 saveRDS(summary_stat,"./Results/asym_sens_results/beta_variation_summary_stat.RDS")
+saveRDS(ms_asym_beta_0.1_beta_0.9,"./Results/asym_sens_results/ms_asym_beta_0.1_beta_0.9.RDS")
 saveRDS(stat_npa_asym_beta_0.1_beta_0.9,"./Results/asym_sens_results/stat_npa_asym_beta_0.1_beta_0.9.RDS")

### 2018-05-14 SupportingInformation.Rmd
some changes

--- 2018-05-08 SupportingInformation.Rmd
@@
 <!--Another model, distinct from the above, Dan and Shya to discuss-->
-```{r asym_sens_beta_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
-set.seed(seed)
+```{r asym_sens_b_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+set.seed(seed)
+#-----------------------------------------------------------------------------------
+source("OurBiCopSelect.R")
+source("bivfunctionplot.R")
 source("asym_sensitivity.R")
-
 #-----------------------------------------------------------------------------------
@@
 # Input :
-#       N : number of global and local noise points you want
-#       rho : covariance of global noise matrix
-#       alpha : a number : local noise coeeficient
-#       beta : growth factor of pop
+#       mg : global noise matrix (dim = N by 2)
+#       ml : local noise matrix (dim = N by 2)
+#       a : a number : local noise coeeficient
+#       b : growth factor of pop
 #       decrease : logical (controls the nature of global noise)
@@
 #     a list of 4 pop matrices 
-#                         1. (for beta=+ve, decrease=F)
-#                         2. (for beta=+ve, decrease=T)
-#                         3. (for beta=-ve, decrease=F)
-#                         4. (for beta=-ve, decrease=T)
-
-plotter_asym_sens_beta_varn<-function(N,rho,alpha,beta,numkeep_last,ploton){
-
-  pop<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
-  pop_d<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
-
-  neg_beta<- -beta
-  pop_neg_beta<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
-  pop_d_neg_beta<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
+#                         1. (for b=+ve, decrease=F)
+#                         2. (for b=+ve, decrease=T)
+#                         3. (for b=-ve, decrease=F)
+#                         4. (for b=-ve, decrease=T)
+
+plotter_asym_sens_b_varn<-function(mg,ml,a,b,numkeep_last,ploton){
+
+  pop<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
+  pop_d<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
+
+  neg_b<- -b
+  pop_neg_b<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=neg_b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
+  pop_d_neg_b<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=neg_b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
 
@@
 
-    plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
-    mtext(paste0("(A)"),side = 3, line=0.3, adj=0.3, col="black")
-    plot(pop_d[,1],pop_d[,2],col="blue",main = bquote(beta ==.(beta)))
-    mtext(paste0("(B)"),side = 3, line=0.3, adj=0.3, col="black")
-    plot(pop_neg_beta[,1],pop_neg_beta[,2],col="orange",main = bquote(beta ==.(neg_beta)))
-    mtext(paste0("(C)"),side = 3, line=0.3, adj=0.3, col="black")
-    plot(pop_d_neg_beta[,1],pop_d_neg_beta[,2],col="skyblue",main = bquote(beta ==.(neg_beta)))
-    mtext(paste0("(D)"),side = 3, line=0.3, adj=0.3, col="black")
+    #plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
+    plot(pop[,1],pop[,2],col="red",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
+    mtext(expression(paste("(A) ","b = ",b,", g = ",g[1])),side = 3, line=0.3, adj=0.5, col="black")
+    plot(pop_d[,1],pop_d[,2],col="blue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
+    mtext(expression(paste("(B) ","b = ",b,", g = ",g[2])),side = 3, line=0.3, adj=0.5, col="black")
+    plot(pop_neg_b[,1],pop_neg_b[,2],col="orange",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
+    mtext(expression(paste("(C) ","b = ",b,", g = ",g[1])),side = 3, line=0.3, adj=0.5, col="black")
+    plot(pop_d_neg_b[,1],pop_d_neg_b[,2],col="skyblue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
+    mtext(expression(paste("(D) ","b = ",b,", g = ",g[2])),side = 3, line=0.3, adj=0.5, col="black")
 
@@
               pop_d=pop_d,
-              pop_neg_beta=pop_neg_beta,
-              pop_d_neg_beta=pop_d_neg_beta))
-
-}
-#--------------------------------------------------------------
+              pop_neg_b=pop_neg_b,
+              pop_d_neg_b=pop_d_neg_b))
+
+}
+#----------------------------------------------------------
 N<-10000
-rho<-0.8
-alpha<-0.2
+r<-0.8
+a<-0.2
 numkeep_last<-1000
-
-pdf("./Results/asym_sens_results/beta_0.1_asym_sens.pdf")
-ans_beta_0.1<-plotter_asym_sens_beta_varn(N=N,rho=rho,alpha=alpha,beta=0.1,numkeep_last=numkeep_last,ploton=T)
-dev.off()
-
-pdf("./Results/asym_sens_results/beta_0.9_asym_sens.pdf")
-ans_beta_0.9<-plotter_asym_sens_beta_varn(N=N,rho=rho,alpha=alpha,beta=0.9,numkeep_last=numkeep_last,ploton=T)
-dev.off()
-
-vlist_beta_0.1_beta_0.9<-c(ans_beta_0.1,ans_beta_0.9) # a list of 8 matrices
+temp<-get_noise_glob_loc_mat(N=N,r=r)
+mg<-temp$mg
+ml<-temp$ml
+
+pdf("./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf")
+ans_b_0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.1,numkeep_last=numkeep_last,ploton=T)
+dev.off()
+
+pdf("./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf")
+ans_b_0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.9,numkeep_last=numkeep_last,ploton=T)
+dev.off()
+
+vlist_b_0.1_b_0.9<-c(ans_b_0.1,ans_b_0.9) # a list of 8 matrices
 
 summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
-rownames(summary_stat)<-c("bestcop_AICw","N_AIC_w","relLTdep_AICw","relUTdep_AIC_w","p_CvM","p_KS","Numboot_success")
-bg <- intToUtf8(0x03B2) 
-cnm<-c("_0.1","_0.1_decrease_T","_-0.1","_-0.1_decrease_T","_0.9","_0.9_decrease_T","_-0.9","_-0.9_decrease_T")
-cname<-paste0(bg,cnm)
-colnames(summary_stat)<-cname
-
-stat_npa_asym_beta_0.1_beta_0.9<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible beta variation
-names(stat_npa_asym_beta_0.1_beta_0.9)<-cname
+rownames(summary_stat)<-c("bestcop_AICw","N_AICw","relLTdep_AICw","relUTdep_AICw","p_CvM","p_KS","Numboot_success")
+#bg <- intToUtf8(0x03B2) 
+cnm<-c("b_0.1_g~1~","b_0.1_g~2~","b_-0.1_g~1~","b_-0.1_g~2~","b_0.9_g~1~","b_0.9_g~2~","b_-0.9_g~1~","b_-0.9_g~2~")
+#cname<-paste0(bg,cnm)
+colnames(summary_stat)<-cnm
+
+ms_asym_b_0.1_b_0.9<-vector(mode = "list", length = 8) # initializing list to store model selection results for all possible b variation
+names(ms_asym_b_0.1_b_0.9)<-cnm
+
+stat_npa_asym_b_0.1_b_0.9<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible b variation
+names(stat_npa_asym_b_0.1_b_0.9)<-cnm
 
 for(ic in c(1:8)){
-  v<-vlist_beta_0.1_beta_0.9[[ic]]
+  v<-vlist_b_0.1_b_0.9[[ic]]
   res<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=c(1,3:10,13:14,16:20),level=0.05,
@@
                       gofnormal=F,status=F)
+
+  ms_asym_b_0.1_b_0.9[[ic]]<-res
 
@@
 
-  stat_npa_asym_beta_0.1_beta_0.9[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cname[ic],numbin=10)
-
-}
-
-saveRDS(summary_stat,"./Results/asym_sens_results/beta_variation_summary_stat.RDS")
-saveRDS(stat_npa_asym_beta_0.1_beta_0.9,"./Results/asym_sens_results/stat_npa_asym_beta_0.1_beta_0.9.RDS")
-```
-
-![results for model selection \label{fig_ms_asym_beta_0.1}](./Results/asym_sens_results/beta_0.1_asym_sens.pdf){ width=40%, height=40% }
-
-![results for model selection \label{fig_ms_asym_beta_0.9}](./Results/asym_sens_results/beta_0.9_asym_sens.pdf){ width=40%, height=40% }
+  stat_npa_asym_b_0.1_b_0.9[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cnm[ic],numbin=10)
+
+}
+
+saveRDS(summary_stat,"./Results/asym_sens_results/b_variation_summary_stat.RDS")
+saveRDS(ms_asym_b_0.1_b_0.9,"./Results/asym_sens_results/ms_asym_b_0.1_b_0.9.RDS")
+saveRDS(stat_npa_asym_b_0.1_b_0.9,"./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
+```
+
+![results for model selection \label{fig_ms_asym_b_0.1}](./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf){ width=40%, height=40% }
+
+![results for model selection \label{fig_ms_asym_b_0.9}](./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf){ width=40%, height=40% }
 
 ```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
-cnm2<-c("b_0.1","b_0.1d_T","b_-0.1","b_-0.1d_T","b_0.9","b_0.9d_T","b_-0.9","b_-0.9d_T")
-colnames(summary_stat)<-cnm2
 knitr::kable(summary_stat, booktabs = T, 
@@
 #knitr::opts_knit$set(kable.force.latex = TRUE)
-tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_beta_0.1_beta_0.9.RDS")
+tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
 
@@
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste("beta",cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
+      caption =paste(cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
   print(temp2)

### 2018-05-14 asym_sensitivity.R

@@ -26,32 +26,16 @@ glob_noise_fn<-function(x,decrease){
   }
 }
 #-----------------------------------------------------------------------
-# Check the function
-#N<-1000
-#rho<-0.8
-#del<-1 #???
-#alpha<-0.5
-  
-#sigma_glob<-matrix(c(1,rho,rho,1),2,2)
-#sigma_loc<-matrix(c(del,0,0,del),2,2)
-
-#mg<-rmvnorm(N,mean=c(0,0),sigma=sigma_glob)
-#ml<-rmvnorm(N,mean=c(0,0),sigma=sigma_loc)
-  
-#g_noise_inc<-apply(mg,MARGIN = 2,FUN = glob_noise_fn, decrease=F)
-#g_noise_dec<-apply(mg,MARGIN = 2,FUN = glob_noise_fn, decrease=T)
-#h_noise<-alpha*ml
-#-----------------------------------------------------------------------
 # This function generates global and local noise matrices 
 #            which will be called as input in next simulator function
 # Input:      
 #       N : number of global and local noise points you want
-#       rho : covariance of global noise matrix
+#       r : covariance of global noise matrix
 # Output:
 #       a list of two matrices (mg for global noise mat) and (ml for local noise matrix)
-get_noise_glob_loc_mat<-function(N,rho){
+get_noise_glob_loc_mat<-function(N,r){
   
-  sigma_glob<-matrix(c(1,rho,rho,1),2,2)
+  sigma_glob<-matrix(c(1,r,r,1),2,2)
   sigma_loc<-matrix(c(1,0,0,1),2,2)
   
   mg<-rmvnorm(N,mean=c(0,0),sigma=sigma_glob)
@@ -63,29 +47,29 @@ get_noise_glob_loc_mat<-function(N,rho){
 }
 #----------------------------------------------------------------------------------
 # This function simulates the equation to test asym sensitivity :
-#      P(i,t+1)=beta*P(i,t)+globalnoise(i,t)+alpha*localnoise(i,t)
+#      P(i,t+1)=b*P(i,t)+globalnoise(i,t)+a*localnoise(i,t)
 # Input :
 #       mg : global noise matrix (dim = N by 2)
 #       ml : local noise matrix (dim = N by 2)
-#       rho : covariance of global noise matrix
-#       alpha : a number : local noise coeeficient
-#       beta : growth factor of pop
+#       r : covariance of global noise matrix
[…]

### 2018-05-14 SupportingInformation.Rmd
minor change in plot

@@
     plot(pop[,1],pop[,2],col="red",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    mtext(expression(paste("(A) ","b = ",b,", g = ",g[1])),side = 3, line=0.3, adj=0.5, col="black")
+    title(expression(paste("(A) g = ",g[1]," , ")), adj=0.4)
+    mtext(paste("b = ",b),side = 3, line=0.2, adj=0.72, col="black")
     plot(pop_d[,1],pop_d[,2],col="blue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    mtext(expression(paste("(B) ","b = ",b,", g = ",g[2])),side = 3, line=0.3, adj=0.5, col="black")
+    title(expression(paste("(B) g = ",g[2]," , ")), adj=0.4)
+    mtext(paste("b = ",b),side = 3, line=0.2, adj=0.72, col="black")
     plot(pop_neg_b[,1],pop_neg_b[,2],col="orange",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    mtext(expression(paste("(C) ","b = ",b,", g = ",g[1])),side = 3, line=0.3, adj=0.5, col="black")
+    title(expression(paste("(C) g = ",g[1]," , ")), adj=0.4)
+    mtext(paste("b = ",neg_b),side = 3, line=0.2, adj=0.75, col="black")
     plot(pop_d_neg_b[,1],pop_d_neg_b[,2],col="skyblue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    mtext(expression(paste("(D) ","b = ",b,", g = ",g[2])),side = 3, line=0.3, adj=0.5, col="black")
+    title(expression(paste("(D) g = ",g[2]," , ")), adj=0.4)
+    mtext(paste("b = ",neg_b),side = 3, line=0.2, adj=0.75, col="black")
 

### 2018-05-15 Paper.Rmd
adding asym sens description : not completed yet

@@
 ### Asymmetric sensitivity and copula structure in populations
+<!--Another model, distinct from the above, Dan and Shya to discuss-->
+Here we intend to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) beyond a certain expectation threshold.
+To test this assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;
 
-<!--Another model, distinct from the above, Dan and Shya to discuss-->
+\begin{equation}\label{eq.AS}
+P_{i}(t+1) = b P_{i}(t) + g(\epsilon_{i}(t)) + h(\delta_{i}(t))
+\end{equation}
+
+$b$ was the linear growth rate.
+Global environmental fluctuation function $g(\epsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq.
+\ref{eq.AS_g}) where $\epsilon=(\epsilon_{1}, \epsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\epsilon}$.
+
+\begin{align}
+g(\epsilon) = \begin{cases} 
+g_{1} = \begin{cases}
+\epsilon & \text{if $\epsilon< 0$}; \\
+0 & \text{otherwise}.
+\end{cases}\\ \\
+g_{2} = \begin{cases}
+\epsilon & \text{if $\epsilon> 0$}; \\
+0 & \text{otherwise}\label{eq.AS_g}.
+\end{cases}
+\end{cases} \text{where } \epsilon = N(\textbf{0},\Sigma_{\epsilon})\;\ \text{with} \;\ \Sigma_{\epsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
+\end{align}
+
+Local environmental fluctuation function $h(\delta)$ varied according to Eq.
+\ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$.
+$\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$.
+
+\begin{equation}\label{eq.AS_h}
+h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
+\end{equation}
+
+We simulated our model (\ref{eq.AS}) for different values of $b$ with $a=0.2$, $r=0.8$.
+$N=10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ populations to plot the copula for a specific $b$ and a specific functional form of $g$.
+Fig.
+\ref{SM-fig_ms_asym_b_0.1} and Fig.
+\ref{SM-fig_ms_asym_b_0.9} show these results.
+We also carried out model selection and non-parametric statistical analysis with each copula of those figures and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_8}.
+Table \ref{SM-tab_summary_stat_asym_sens} shows that according to our model selection approach for $b=\pm0.1$, non-normal copula (with higher AIC weights) were always best-fit with some asymetric tail dependence.
 
@@
 
-Figure \ref{fig_skewness_ff} shows that averaged skewness is much higher than expected if the aphid first flight data came from a normal copula.
+Figure \ref{SM-fig_skewness_ff} shows that averaged skewness is much higher than expected if the aphid first flight data came from a normal copula.
 Here, $p$ value indicates the fraction of surrogate-skewness which are greater than the averaged skewness of the given data (red line in Figure \ref{SM-fig_skewness_ff}).
@@
 
-    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]
+    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{SM-fig_risk_K_effect}]
 

### 2018-05-15 SupportingInformation.Rmd
adding asym sens description : not completed yet

@@
 \caption[\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data]{\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data.
-Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see Sec.
-\ref{Model_selection}).\label{fig_raw_cop_soilCN}}
+Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_soilCN}}
 \end{center}
@@
 ln(mass) data \textbf{(B)} Copula plot for BMR dataset.
-Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see Sec.
-\ref{Model_selection}).\label{fig_raw_cop_BMR}}
+Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_BMR}}
 \end{center}
@@
 AVG.humerus phylogenetic traits.
-Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see Sec.
-\ref{Model_selection}).
+Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).
 \label{fig_raw_cop_AVG_humerus_ltoe}}
@@
 
-\newpage
-
+<!--
 ## Asymmetric sensitivity and copula structure in populations
-
-<!--Another model, distinct from the above, Dan and Shya to discuss-->
+           ----moved to Paper.Rmd----
+-->
+
 ```{r asym_sens_b_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
@@
 summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
-rownames(summary_stat)<-c("bestcop_AICw","N_AICw","relLTdep_AICw","relUTdep_AICw","p_CvM","p_KS","Numboot_success")
+rownames(summary_stat)<-c("bestcop_AICw","N_AICw","relLTdep_AICw","relUTdep_AICw","p_CvM","p_KS","BS_success%")
 #bg <- intToUtf8(0x03B2) 
@@
   summary_stat[6,ic]<-res$GofRes_KS
-  summary_stat[7,ic]<-res$Numboot_success
+  summary_stat[7,ic]<-(res$Numboot_success/res$Numboot)*100
 
@@
 
-![results for model selection \label{fig_ms_asym_b_0.1}](./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf){ width=40%, height=40% }
-
-![results for model selection \label{fig_ms_asym_b_0.9}](./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf){ width=40%, height=40% }
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
+\caption[asym sensitivity results : copula plot]{asym sensitivity results : copula plot \label{fig_ms_asym_b_0.1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf}
+\caption[asym sensitivity results : copula plot]{asym sensitivity results : copula plot \label{fig_ms_asym_b_0.9}}
+\end{center}
+\end{figure}
 
@@
 
-\newpage
-```{r tab_stat_npa_asym,echo=F,results="asis"}
+```{r read_tab_stat_npa_asym,echo=F,results="asis"}
 #knitr::opts_knit$set(kable.force.latex = TRUE)
@@
 
-for(i in c(1:8)){
-  temp<-make_table_stat_fn(data=tab_asym[[i]])
-  temp2<-knitr::kable(temp, booktabs = T, 
+#for(i in c(1:8)){
+#  temp<-make_table_stat_fn(data=tab_asym[[i]])
+#  temp2<-knitr::kable(temp, booktabs = T, 
+#             format = "pandoc",
+      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
+#      caption =paste(cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
+#  print(temp2)
+#  cat("\\newpage")
+#}
+```
+
+```{r tab_stat_npa_asym_1,echo=F,results="markup"}
+temp<-make_table_stat_fn(data=tab_asym[[1]])
+knitr::kable(temp, booktabs = T, 
              format = "pandoc",
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
-  print(temp2)
-  cat("\\newpage")
-}
+      caption =paste(cnm[1]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_1}"))
+```
+
+\newpage
+```{r tab_stat_npa_asym_2,echo=F,results="markup"}
+temp<-make_table_stat_fn(data=tab_asym[[2]])
+knitr::kable(temp, booktabs = T, 
+             format = "pandoc",
+      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
+      caption =paste(cnm[2]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_2}"))
+```
+
+\newpage
+```{r tab_stat_npa_asym_3,echo=F,results="markup"}
+temp<-make_table_stat_fn(data=tab_asym[[3]])
+knitr::kable(temp, booktabs = T, 
+             format = "pandoc",
+      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
+      caption =paste(cnm[3]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_3}"))
+```
+
+\newpage
+
+```{r tab_stat_npa_asym_4,echo=F,results="markup"}
+temp<-make_table_stat_fn(data=tab_asym[[4]])
+knitr::kable(temp, booktabs = T, 
+             format = "pandoc",
+      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
+      caption =paste(cnm[4]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_4}"))
+```
+
+```{r tab_stat_npa_asym_5,echo=F,results="markup"}
+temp<-make_table_stat_fn(data=tab_asym[[5]])
+knitr::kable(temp, booktabs = T, 
+             format = "pandoc",
+      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
+      caption =paste(cnm[5]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_5}"))
+```
+
+```{r tab_stat_npa_asym_6,echo=F,results="markup"}
+temp<-make_table_stat_fn(data=tab_asym[[6]])
+knitr::kable(temp, booktabs = T, 
+             format = "pandoc",
+      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
+      caption =paste(cnm[6]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_6}"))
+```
+
+```{r tab_stat_npa_asym_7,echo=F,results="markup"}
+temp<-make_table_stat_fn(data=tab_asym[[7]])
+knitr::kable(temp, booktabs = T, 
+             format = "pandoc",
+      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
+      caption =paste(cnm[7]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_7}"))
+```
+
+```{r tab_stat_npa_asym_8,echo=F,results="markup"}
+temp<-make_table_stat_fn(data=tab_asym[[8]])
+knitr::kable(temp, booktabs = T, 
+             format = "pandoc",
+      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
+      caption =paste(cnm[8]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_8}"))
 ```

### 2018-05-15 makefile.Rmd
a single rmd to compile both paper and supmat Rmds

---
title: "Makefile_linux, compiliation notes"
author: "Daniel Reuman"
date: ""
output: pdf_document
---

# Get tex files and pdf files for the supp mat, but the pdf files 
#are missing the external references

```{r tex_for_supp_mat}
rmarkdown::render(input="SupportingInformation.Rmd") 
```

# Get tex files and pdf files for the main text, but the pdf files 
#are missing the external references

```{r tex_for_main_text}
rmarkdown::render(input="Paper.Rmd") 
```

# Now recompile the latex

```{r recompile_latex}
#do it once
system2(command="pdflatex",args="SupportingInformation.tex",stdout=TRUE,stderr=TRUE)
system2(command="pdflatex",args="Paper.tex",stdout=TRUE,stderr=TRUE)

#do it again to make sure to get the external references right
system2(command="pdflatex",args="SupportingInformation.tex",stdout=TRUE,stderr=TRUE)
system2(command="pdflatex",args="Paper.tex",stdout=TRUE,stderr=TRUE)
```

### 2018-05-16 Paper.Rmd
some changes and asym sens description completed

@@
 knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
-seed<-101
-# families<-c(1,3:10,13,14,16:20)
-source("mtime.R") #A function needed for caching
+#seed<-101
+#source("mtime.R") #A function needed for caching
 ```
@@
 the right parts of the distributions.
-
-```{r fig_stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"),mtime("StatTestNumericResults.RData"))}
-set.seed(seed) 
-source("fig_Paper_stat_testing.R")
-load("./Results/stat_results/stat_testing/StatTestNumericResults.RData")
-resultsloc<-"./Results/stat_results/stat_testing/"
-
-copname<-"C"
-reslist<-res_pt35_rankT
-
-statname<-"CorlmCoru"
-call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
-statname<-"PlmPu"
-call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
-statname<-"D2umD2l"
-call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
-```
 
@@
 We simulated our model (\ref{eq.AS}) for different values of $b$ with $a=0.2$, $r=0.8$.
-$N=10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ populations to plot the copula for a specific $b$ and a specific functional form of $g$.
+$10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ time step of simulation to plot the population-copula for a specific $b$ and a specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
 Fig.
 \ref{SM-fig_ms_asym_b_0.1} and Fig.
-\ref{SM-fig_ms_asym_b_0.9} show these results.
+\ref{SM-fig_ms_asym_b_0.9} show these results for $b=\pm0.1$ and $b=\pm0.9$, respectively.
 We also carried out model selection and non-parametric statistical analysis with each copula of those figures and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_8}.
-Table \ref{SM-tab_summary_stat_asym_sens} shows that according to our model selection approach for $b=\pm0.1$, non-normal copula (with higher AIC weights) were always best-fit with some asymetric tail dependence.
+Table \ref{SM-tab_summary_stat_asym_sens} represent the model selection results with AIC weights for best fitted copula and normal copula, model averaged lower and upper tail dependence, p-values (both Cramer-von Mises statistic and Kolmogorov-Smirnov statistic) for goodness of fit test.
+In each case, successful bootstrap was $100\%$.
+According to our model selection approach, for $b=\pm0.1$, non-normal copula (with higher AIC weights) were always best-fit with some asymetric tail dependence whereas for $b\pm0.9$, most of the time normal was best.
+It is probably because the populations get long-time exposure to environmental changes for slower growth(or decay) rate and so global fluctuation can affect them more efficiently.
+Our non-parametric analysis results (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_8}) are consistent with this model selection approach.
 

### 2018-05-16 SupportingInformation.Rmd
some changes and asym sens description completed

@@
 
-<!--checkpoint package-->
-```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
-library(checkpoint)
-library(captioner)
-if (!dir.exists("./.checkpoint/")){
-  dir.create("./.checkpoint/")
-}
-checkpoint("2018-03-29",checkpointLocation = "./")
-```
+<!--checkpoint chunk is moved to makefile.Rmd-->
 
@@
 
+<!--This r-chunk generates figure necessary for maintext Paper-->
+```{r fig_stat_testing_for_maintext, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"),mtime("StatTestNumericResults.RData"))}
+set.seed(seed) 
+source("fig_Paper_stat_testing.R")
+#load("./Results/stat_results/stat_testing/StatTestNumericResults.RData")
+resultsloc<-"./Results/stat_results/stat_testing/"
+
+copname<-"C"
+reslist<-res_pt35_rankT
+
+statname<-"CorlmCoru"
+call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
+statname<-"PlmPu"
+call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
+statname<-"D2umD2l"
+call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
+```
+
 <!--

### 2018-05-16 makefile.Rmd
some changes and asym sens description completed

@@
 ---
-title: "Makefile_linux, compiliation notes"
+title: "Makefile, compiliation notes"
 author: "Daniel Reuman"
@@
 ---
+<!--checkpoint package-->
+```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
+library(checkpoint)
+library(captioner)
+library(rmarkdown)
+if (!dir.exists("./.checkpoint/")){
+  dir.create("./.checkpoint/")
+}
+checkpoint("2018-03-29",checkpointLocation = "./")
+```
 

### 2018-05-16 SupportingInformation.Rmd
bug fixed

@@
 tab_BMR_asym_numbin5<-make_table_stat_fn(data=stat_BMR_ga_numbin5)
+
 knitr::kable(tab_BMR_asym,

### 2018-05-16 bivfunctionplot.R

@@ -33,7 +33,7 @@ makeSurrog<-function(v,numsurrog=1000){
   return(list(surv_K=surv_K,
               surv_S=surv_S))
 }
-#makeSurrog(v=v_CN)
+#makeSurrog(v=v_BMR)
 #------------------------------------------------------------------
 #utility function for calculating fractions of surrogates 
 #with a stat smaller than that of data, i.e., fraction of 
@@ -46,6 +46,8 @@ fracwork<-function(dvals,surrvals){
   }
   return(frac)
 }
+#corstats_frac_K<-fracwork(corstats_d,corstats_K)
+
 #--------------------------------------------------------------------
 #now work with the P stats
 Pbds_wrap<-function(vi,vj,lb,ub){
@@ -63,6 +65,24 @@ fracplot<-function(x,fracs,ylims,numsurrog=1000){
     }
     yht<-ylims[2]-.1*diff(ylims)
     text(x[counter],yht,ptxt,adj=c(0.5,.5),cex=0.5,srt=90)
+  }
+}
+#-----------------------------
+#This function will do the same thing as of fracplot
+myfracplot<-function(x,stats_d,surrog_statsq,stats_frac,ylims,numsurrog=1000){
+  for (counter in 1:length(x)){
+    ptxt<-''
+    if(stats_d[counter]>surrog_statsq[3,][counter]){
+      ptxt<-paste(">",stats_frac[counter],sep='')
+    }
+    if(stats_d[counter]<surrog_statsq[2,][counter]){
+      ptxt<-paste("<",numsurrog-stats_frac[counter],sep='')
+    }
+    #ptxt2<-stats_frac[counter]
+    
+    yht<-ylims[2]-.1*diff(ylims)
+    text(x[counter],yht,ptxt,adj=c(0.5,.5),cex=0.5,srt=90)
+    #text(x[counter],yht,ptxt2,adj=c(0.5,.5),cex=0.5,srt=90)
   }
 }
 #----------------------------------------------------------------
@@ -87,16 +107,16 @@ bivfunctionplot<-function(v,resloc,nametag,numbin){
[…]

### 2018-05-17 SupportingInformation.Rmd
some minor changes made

@@
 op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(phylo_dat$AVG.ltoe.IV,phylo_dat$AVG.humerus,type="p",col="red",xlab="AVG.ltoe.IV",ylab="AVG.humerus",
+plot(phylo_dat$AVG.ltoe.IV,phylo_dat$AVG.humerus,type="p",col="red",xlab="contrasts for AVG.ltoe.IV",ylab="contrasts for AVG.humerus",
      cex.lab=1.5,cex.axis=1.5)
@@
 \includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
-\caption[\textbf{(A)} Plot of the raw AVG.ltoe.IV vs.
+\caption[\textbf{(A)} Plot of the contrasts for AVG.ltoe.IV vs.
 AVG.humerus phylogenetic traits.
-\textbf{(B)} Copula plot of AVG.ltoe.IV vs.
-AVG.humerus phylogenetic traits]{\textbf{(A)} Plot of the raw AVG.ltoe.IV vs.
-AVG.humerus phylogenetic traits.
-\textbf{(B)} Copula plot of AVG.ltoe.IV vs.
-AVG.humerus phylogenetic traits.
+\textbf{(B)} Copula plot of contrasts for AVG.ltoe.IV vs.
+contrasts for AVG.humerus phylogenetic traits]{\textbf{(A)} Plot of the contrasts for AVG.ltoe.IV vs.
+contrasts for AVG.humerus phylogenetic traits.
+\textbf{(B)} Copula plot of contrasts for AVG.ltoe.IV vs.
+contrasts for AVG.humerus phylogenetic traits.
 Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).
@@
       #caption.short = "Fitting info for soil C and N dataset",
-      caption = "Fitting info for soil C and N dataset.
+      caption = "Fitting info for **soil C and N dataset**.
 Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=survival Clayton; SG=survival Gumbel; SJ=survival Joe.
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
-\caption[Testing non-parametric statistics for soilCN data]{Nonparametric tests for tail dependence and other deviations from a normal copula for soil C and N data.
+\caption[Testing non-parametric statistics for soilCN data]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{soil C and N data}.
 As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
@@
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.
+      caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.
 Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (Methods).
 One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates.
-A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.
-Recall that Spearman-preserving surrogates are more appropriate for the $\\cor$ statistic (Figure \\ref{fig_soilCN_nonparam} caption).\\label{tab_soilCN_asym}")
+A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
 #kable_styling(, position = "float_left")
@@
 \includegraphics[width=16 cm]{./Results/BMR_results/BMR_numbin_10_bivfunctionplot.pdf}
-\caption[Testing non-parametric statistics for BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data.
+\caption[Testing non-parametric statistics for BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{BMR data}.
 As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
@@
 
+<!--
 \begin{figure}[!h]
@@
 \end{figure}
+-->
 
@@
              format="pandoc",
-             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)?, numbin=10 \\label{tab_BMR_asym}",
+             caption = "**BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis. \\label{tab_BMR_asym}",
              booktabs=T)
 
-knitr::kable(tab_BMR_asym_numbin5,
-             format="pandoc",
-             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)?, numbin=5 \\label{tab_BMR_asym_numbin5}",
-             booktabs=T)
+#knitr::kable(tab_BMR_asym_numbin5,
+#             format="pandoc",
+#             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)?, numbin=5 \\label{tab_BMR_asym_numbin5}",
+#             booktabs=T)
 ```
@@
 set.seed(seed) 
-# Now do NPA for v_contrasts_67
+# Now do NPA for v_contrasts_67 with numbin=4
 stat_contrasts_67_ga<-bivfunctionplot(v=v_contrasts_67,resloc="./Results/Anolis_results/contrasts_csv_results/",
                                       nametag="contrasts_67",numbin = 4) 
+```
+
+```{r stat_contrasts_67_graphic_approach2, echo=F, results="hide", cache=T, cache.extra=list(seed,v_contrasts_67,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+source("bivfunctionplot.R")
+set.seed(seed) 
+# Now do NPA for v_contrasts_67
+
+stat_contrasts_67_ga_numbin2<-bivfunctionplot(v=v_contrasts_67,
+                            resloc="./Results/Anolis_results/contrasts_csv_results/",
+                            nametag="contrasts_67_numbin2",numbin = 2)
+
+stat_contrasts_67_ga_numbin3<-bivfunctionplot(v=v_contrasts_67,
+                            resloc="./Results/Anolis_results/contrasts_csv_results/",
+                            nametag="contrasts_67_numbin2",numbin = 3)
+
+stat_contrasts_67_ga_numbin6<-bivfunctionplot(v=v_contrasts_67,
+                            resloc="./Results/Anolis_results/contrasts_csv_results/",
+                            nametag="contrasts_67_numbin6",numbin = 6)
+
+stat_contrasts_67_ga_numbin7<-bivfunctionplot(v=v_contrasts_67,
+                            resloc="./Results/Anolis_results/contrasts_csv_results/",
+                            nametag="contrasts_67_numbin7",numbin = 7)
+
+stat_contrasts_67_ga_numbin10<-bivfunctionplot(v=v_contrasts_67,
+                            resloc="./Results/Anolis_results/contrasts_csv_results/",
+                            nametag="contrasts_67_numbin10",numbin = 10)
 ```
@@
 ```
+
+<!--
+This chunks are just to see the tables for all numbin variation for anolis data set: It appears that numbin=4 is optimized, when numbin=2, >= 7, there is no significant asymetry in tail dep in all 3 types of stats.
+```{r tab_contrasts_67_asym67_numbin2, echo=F, results="markup"}
+#tab_contrasts_67_asym2<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin2)
+#knitr::kable(tab_contrasts_67_asym2,
+#            format="pandoc",
+#             caption = "numbin_2,anolis \\label{tab_contrasts_67_asym2}",
+#             booktabs=T)
+```
+
+```{r tab_contrasts_67_asym67_numbin3, echo=F, results="markup"}
+#tab_contrasts_67_asym3<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin3)
+#knitr::kable(tab_contrasts_67_asym3,
+#             format="pandoc",
+#             caption = "numbin_3,anolis \\label{tab_contrasts_67_asym3}",
+#             booktabs=T)
+```
+
+```{r tab_contrasts_67_asym67_numbin6, echo=F, results="markup"}
+#tab_contrasts_67_asym6<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin6)
+#knitr::kable(tab_contrasts_67_asym6,
+#             format="pandoc",
+#             caption = "numbin_6,anolis \\label{tab_contrasts_67_asym6}",
+#             booktabs=T)
+```
+
+```{r tab_contrasts_67_asym67_numbin7, echo=F, results="markup"}
+#tab_contrasts_67_asym7<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin7)
+#knitr::kable(tab_contrasts_67_asym7,
+#             format="pandoc",
+#             caption = "numbin_7,anolis \\label{tab_contrasts_67_asym7}",
+#             booktabs=T)
+```
+
+```{r tab_contrasts_67_asym67_numbin10, echo=F, results="markup"}
+#tab_contrasts_67_asym10<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin10)
+#knitr::kable(tab_contrasts_67_asym10,
+#             format="pandoc",
+#             caption = "numbin_10,anolis \\label{tab_contrasts_67_asym10}",
+#             booktabs=T)
+```
+
+-->
 

### 2018-05-18 SupportingInformation.Rmd
log transforming the BMR, not just the mass

@@
 For each contrast, the path length spanned by the contrast (sum of the branch
-  lengths) and the contrast for the mass, BMR and natural
+  lengths) and the contrast for the mass, natural logarithm of the BMR and natural
   logarithm of the mass were written as a row in the bmr-contrasts.csv file.

### 2018-05-18 SupportingInformation.Rmd
log transforming the BMR, not just the mass

--- 2018-05-16 SupportingInformation.Rmd
@@
 
-<!--checkpoint chunk is moved to makefile.Rmd-->
+<!--checkpoint package-->
+```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
+library(checkpoint)
+library(captioner)
+if (!dir.exists("./.checkpoint/")){
+  dir.create("./.checkpoint/")
+}
+checkpoint("2018-03-29",checkpointLocation = "./")
+```
 
@@
 \caption[\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data]{\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data.
-Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_soilCN}}
+Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see Sec.
+\ref{Model_selection}).\label{fig_raw_cop_soilCN}}
 \end{center}
@@
 ln(mass) data \textbf{(B)} Copula plot for BMR dataset.
-Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_BMR}}
+Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see Sec.
+\ref{Model_selection}).\label{fig_raw_cop_BMR}}
 \end{center}
@@
 
-\newpage
-
 \begin{figure}[!h]
@@
 AVG.humerus phylogenetic traits.
-Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).
+Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see Sec.
+\ref{Model_selection}).
 \label{fig_raw_cop_AVG_humerus_ltoe}}
@@
 
-<!--This r-chunk generates figure necessary for maintext Paper-->
-```{r fig_stat_testing_for_maintext, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"),mtime("StatTestNumericResults.RData"))}
-set.seed(seed) 
-source("fig_Paper_stat_testing.R")
-#load("./Results/stat_results/stat_testing/StatTestNumericResults.RData")
-resultsloc<-"./Results/stat_results/stat_testing/"
-
-copname<-"C"
-reslist<-res_pt35_rankT
-
-statname<-"CorlmCoru"
-call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
-statname<-"PlmPu"
-call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
-statname<-"D2umD2l"
-call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
-```
-
 <!--
@@
 tab_BMR_asym_numbin5<-make_table_stat_fn(data=stat_BMR_ga_numbin5)
-
 knitr::kable(tab_BMR_asym,
@@
 
-<!--
+\newpage
+
 ## Asymmetric sensitivity and copula structure in populations
-           ----moved to Paper.Rmd----
--->
-
-```{r asym_sens_b_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
-set.seed(seed)
-#-----------------------------------------------------------------------------------
-source("OurBiCopSelect.R")
-source("bivfunctionplot.R")
+
+<!--Another model, distinct from the above, Dan and Shya to discuss-->
+```{r asym_sens_beta_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+set.seed(seed)
 source("asym_sensitivity.R")
+
 #-----------------------------------------------------------------------------------
@@
 # Input :
-#       mg : global noise matrix (dim = N by 2)
-#       ml : local noise matrix (dim = N by 2)
-#       a : a number : local noise coeeficient
-#       b : growth factor of pop
+#       N : number of global and local noise points you want
+#       rho : covariance of global noise matrix
+#       alpha : a number : local noise coeeficient
+#       beta : growth factor of pop
 #       decrease : logical (controls the nature of global noise)
@@
 #     a list of 4 pop matrices 
-#                         1. (for b=+ve, decrease=F)
-#                         2. (for b=+ve, decrease=T)
-#                         3. (for b=-ve, decrease=F)
-#                         4. (for b=-ve, decrease=T)
-
-plotter_asym_sens_b_varn<-function(mg,ml,a,b,numkeep_last,ploton){
-
-  pop<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
-  pop_d<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
-
-  neg_b<- -b
-  pop_neg_b<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=neg_b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
-  pop_d_neg_b<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=neg_b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
+#                         1. (for beta=+ve, decrease=F)
+#                         2. (for beta=+ve, decrease=T)
+#                         3. (for beta=-ve, decrease=F)
+#                         4. (for beta=-ve, decrease=T)
+
+plotter_asym_sens_beta_varn<-function(N,rho,alpha,beta,numkeep_last,ploton){
+
+  pop<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
+  pop_d<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
+
+  neg_beta<- -beta
+  pop_neg_beta<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
+  pop_d_neg_beta<-Simulator_asym_sens(N=N,rho=rho,alpha=alpha,beta=neg_beta,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
 
@@
 
-    #plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
-    plot(pop[,1],pop[,2],col="red",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    title(expression(paste("(A) g = ",g[1]," , ")), adj=0.4)
-    mtext(paste("b = ",b),side = 3, line=0.2, adj=0.72, col="black")
-    plot(pop_d[,1],pop_d[,2],col="blue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    title(expression(paste("(B) g = ",g[2]," , ")), adj=0.4)
-    mtext(paste("b = ",b),side = 3, line=0.2, adj=0.72, col="black")
-    plot(pop_neg_b[,1],pop_neg_b[,2],col="orange",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    title(expression(paste("(C) g = ",g[1]," , ")), adj=0.4)
-    mtext(paste("b = ",neg_b),side = 3, line=0.2, adj=0.75, col="black")
-    plot(pop_d_neg_b[,1],pop_d_neg_b[,2],col="skyblue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    title(expression(paste("(D) g = ",g[2]," , ")), adj=0.4)
-    mtext(paste("b = ",neg_b),side = 3, line=0.2, adj=0.75, col="black")
+    plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
+    mtext(paste0("(A)"),side = 3, line=0.3, adj=0.3, col="black")
+    plot(pop_d[,1],pop_d[,2],col="blue",main = bquote(beta ==.(beta)))
+    mtext(paste0("(B)"),side = 3, line=0.3, adj=0.3, col="black")
+    plot(pop_neg_beta[,1],pop_neg_beta[,2],col="orange",main = bquote(beta ==.(neg_beta)))
+    mtext(paste0("(C)"),side = 3, line=0.3, adj=0.3, col="black")
+    plot(pop_d_neg_beta[,1],pop_d_neg_beta[,2],col="skyblue",main = bquote(beta ==.(neg_beta)))
+    mtext(paste0("(D)"),side = 3, line=0.3, adj=0.3, col="black")
 
@@
               pop_d=pop_d,
-              pop_neg_b=pop_neg_b,
-              pop_d_neg_b=pop_d_neg_b))
-
-}
-#----------------------------------------------------------
+              pop_neg_beta=pop_neg_beta,
+              pop_d_neg_beta=pop_d_neg_beta))
+
+}
+#--------------------------------------------------------------
 N<-10000
-r<-0.8
-a<-0.2
+rho<-0.8
+alpha<-0.2
 numkeep_last<-1000
-temp<-get_noise_glob_loc_mat(N=N,r=r)
-mg<-temp$mg
-ml<-temp$ml
-
-pdf("./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf")
-ans_b_0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.1,numkeep_last=numkeep_last,ploton=T)
-dev.off()
-
-pdf("./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf")
-ans_b_0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.9,numkeep_last=numkeep_last,ploton=T)
-dev.off()
-
-vlist_b_0.1_b_0.9<-c(ans_b_0.1,ans_b_0.9) # a list of 8 matrices
+
+pdf("./Results/asym_sens_results/beta_0.1_asym_sens.pdf")
+ans_beta_0.1<-plotter_asym_sens_beta_varn(N=N,rho=rho,alpha=alpha,beta=0.1,numkeep_last=numkeep_last,ploton=T)
+dev.off()
+
+pdf("./Results/asym_sens_results/beta_0.9_asym_sens.pdf")
+ans_beta_0.9<-plotter_asym_sens_beta_varn(N=N,rho=rho,alpha=alpha,beta=0.9,numkeep_last=numkeep_last,ploton=T)
+dev.off()
+
+vlist_beta_0.1_beta_0.9<-c(ans_beta_0.1,ans_beta_0.9) # a list of 8 matrices
 
 summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
-rownames(summary_stat)<-c("bestcop_AICw","N_AICw","relLTdep_AICw","relUTdep_AICw","p_CvM","p_KS","BS_success%")
-#bg <- intToUtf8(0x03B2) 
-cnm<-c("b_0.1_g~1~","b_0.1_g~2~","b_-0.1_g~1~","b_-0.1_g~2~","b_0.9_g~1~","b_0.9_g~2~","b_-0.9_g~1~","b_-0.9_g~2~")
-#cname<-paste0(bg,cnm)
-colnames(summary_stat)<-cnm
-
-ms_asym_b_0.1_b_0.9<-vector(mode = "list", length = 8) # initializing list to store model selection results for all possible b variation
-names(ms_asym_b_0.1_b_0.9)<-cnm
-
-stat_npa_asym_b_0.1_b_0.9<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible b variation
-names(stat_npa_asym_b_0.1_b_0.9)<-cnm
+rownames(summary_stat)<-c("bestcop_AICw","N_AIC_w","relLTdep_AICw","relUTdep_AIC_w","p_CvM","p_KS","Numboot_success")
+bg <- intToUtf8(0x03B2) 
+cnm<-c("_0.1","_0.1_decrease_T","_-0.1","_-0.1_decrease_T","_0.9","_0.9_decrease_T","_-0.9","_-0.9_decrease_T")
+cname<-paste0(bg,cnm)
+colnames(summary_stat)<-cname
+
+stat_npa_asym_beta_0.1_beta_0.9<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible beta variation
+names(stat_npa_asym_beta_0.1_beta_0.9)<-cname
 
 for(ic in c(1:8)){
-  v<-vlist_b_0.1_b_0.9[[ic]]
+  v<-vlist_beta_0.1_beta_0.9[[ic]]
   res<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=c(1,3:10,13:14,16:20),level=0.05,
@@
                       gofnormal=F,status=F)
-
-  ms_asym_b_0.1_b_0.9[[ic]]<-res
 
@@
   summary_stat[6,ic]<-res$GofRes_KS
-  summary_stat[7,ic]<-(res$Numboot_success/res$Numboot)*100
-
-  stat_npa_asym_b_0.1_b_0.9[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cnm[ic],numbin=10)
-
-}
-
-saveRDS(summary_stat,"./Results/asym_sens_results/b_variation_summary_stat.RDS")
-saveRDS(ms_asym_b_0.1_b_0.9,"./Results/asym_sens_results/ms_asym_b_0.1_b_0.9.RDS")
-saveRDS(stat_npa_asym_b_0.1_b_0.9,"./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
-```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
-\caption[asym sensitivity results : copula plot]{asym sensitivity results : copula plot \label{fig_ms_asym_b_0.1}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf}
-\caption[asym sensitivity results : copula plot]{asym sensitivity results : copula plot \label{fig_ms_asym_b_0.9}}
-\end{center}
-\end{figure}
+  summary_stat[7,ic]<-res$Numboot_success
+
+  stat_npa_asym_beta_0.1_beta_0.9[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cname[ic],numbin=10)
+
+}
+
+saveRDS(summary_stat,"./Results/asym_sens_results/beta_variation_summary_stat.RDS")
+saveRDS(stat_npa_asym_beta_0.1_beta_0.9,"./Results/asym_sens_results/stat_npa_asym_beta_0.1_beta_0.9.RDS")
+```
+
+![results for model selection \label{fig_ms_asym_beta_0.1}](./Results/asym_sens_results/beta_0.1_asym_sens.pdf){ width=40%, height=40% }
+
+![results for model selection \label{fig_ms_asym_beta_0.9}](./Results/asym_sens_results/beta_0.9_asym_sens.pdf){ width=40%, height=40% }
 
 ```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
+cnm2<-c("b_0.1","b_0.1d_T","b_-0.1","b_-0.1d_T","b_0.9","b_0.9d_T","b_-0.9","b_-0.9d_T")
+colnames(summary_stat)<-cnm2
 knitr::kable(summary_stat, booktabs = T, 
@@
 
-```{r read_tab_stat_npa_asym,echo=F,results="asis"}
+\newpage
+```{r tab_stat_npa_asym,echo=F,results="asis"}
 #knitr::opts_knit$set(kable.force.latex = TRUE)
-tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
-
-#for(i in c(1:8)){
-#  temp<-make_table_stat_fn(data=tab_asym[[i]])
-#  temp2<-knitr::kable(temp, booktabs = T, 
-#             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-#      caption =paste(cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
-#  print(temp2)
-#  cat("\\newpage")
-#}
-```
-
-```{r tab_stat_npa_asym_1,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[1]])
-knitr::kable(temp, booktabs = T, 
+tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_beta_0.1_beta_0.9.RDS")
+
+for(i in c(1:8)){
+  temp<-make_table_stat_fn(data=tab_asym[[i]])
+  temp2<-knitr::kable(temp, booktabs = T, 
              format = "pandoc",
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[1]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_1}"))
-```
-
-\newpage
-```{r tab_stat_npa_asym_2,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[2]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[2]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_2}"))
-```
-
-\newpage
-```{r tab_stat_npa_asym_3,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[3]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[3]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_3}"))
-```
-
-\newpage
-
-```{r tab_stat_npa_asym_4,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[4]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[4]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_4}"))
-```
-
-```{r tab_stat_npa_asym_5,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[5]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[5]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_5}"))
-```
-
-```{r tab_stat_npa_asym_6,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[6]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[6]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_6}"))
-```
-
-```{r tab_stat_npa_asym_7,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[7]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[7]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_7}"))
-```
-
-```{r tab_stat_npa_asym_8,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[8]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[8]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_8}"))
+      caption =paste("beta",cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
+  print(temp2)
+  cat("\\newpage")
+}
 ```
@@
 \newpage
-
 # References

### 2018-05-19 Paper.Rmd
BMR data calculation completed and description added

@@
 
-As described in Methods, each of our statistics, as computed for the 
-ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$, was compared to the distributions 
+As described in \nameref{Methods}, each of our statistics, as computed for the 
+ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
 of values of the same statistic with the same ranges computed on 1000 normal-copula 
 surrogate datasets with similar Kendall or Spearman correlations, in different 
-runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). Because the correlation 
+runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). 
+<!--
+Because the correlation 
 used in the $\cor$ statistic is the Spearman correlation, Spearman-preserving surrogates
 should be used, so the top-right panel of Fig. \ref{SM-fig_soilCN_nonparam}
-is more reliable than the top left. Results indicate that tail dependence of 
+is more reliable than the top left.
+--> 
+Results indicate that tail dependence of 
 data is stronger in both the lower and upper tails than would be expected under a 
@@
 ```{r read_BivMS_BMR_data,echo=F}
-BivMS_BMR<-readRDS("./Results/BMR_results/BivMS_BMR_lnmass.RDS")
+BivMS_BMR<-readRDS("./Results/BMR_results/BivMS_lnBMR_vs_lnmass.RDS")
 ```
@@
 and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_BMRfit}).
-The table shows that Frank is much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+The table shows that BB1 and SBB1 are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
 
-The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was $0$ as expected, because the contributions came from Frank and SBB8 which are both symmetric copula.
-We also tested that lowest-AIC copula Frank was a good-enough fit, giving $p$-value $`r BivMS_BMR$GofRes_CvM`$ (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value $`r BivMS_BMR$GofRes_KS`$ in the Kolmogorov-Smirnov-based test of the goodness of fit.
+The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was `r BivMS_BMR$relLTdep_AICw - BivMS_BMR$relUTdep_AICw`, indicating nearly symmetric tail dependence.
+BB1, being as a two parameter copula,  has both lower and upper tail dependence and depending upon which parameter is greater it shows comparatively greater lower or upper tail dependence.
+SBB1 is just opposite of BB1 in tail dependency.
+So, the result was expected, because the contributions came from more or less equally weighted BB1 and SBB1.
+We also tested that lowest-AIC copula BB1 was a good-enough fit, giving $p$-value `r BivMS_BMR$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_BMR$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
 The next section looks at tail dependence in a non-parametric way.
@@
 #### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
-<!--we have to add description here after DAN fixes bug(?) in bivfunctionplot.R-->
+
+As described in \nameref{Methods}, each of our statistics, as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig.
+\ref{SM-fig_BMR_nonparam}).
+
+Results indicate that dependence of data is different in some bins than would be expected under a 
+null hypothesis of a normal copula, answering the second part of question 1.
+But Table \ref{SM-tab_BMR_asym} indicates that tail dependence for the BMR data copula is significantly (95% CI) not different from symmetric normal copula and thus consistent with model selection results.
 

### 2018-05-19 SupportingInformation.Rmd
BMR data calculation completed and description added

--- 2018-05-18 SupportingInformation.Rmd
@@
 
-**bmr-contrasts.csv** : This contains the trait contrasts
+[**bmr-contrasts.csv**](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) : 
+This contains the trait contrasts
 from phylogenetically independent contrasts [@SukumaranH2010] 
@@
 # Generating raw data plot btw ln(mass) and BMR from bmr-contrasts.csv
-pdf("./Results/BMR_results/rawdataplot_BMR_vs_lnmass.pdf")
+pdf("./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf")
 op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(BMR_dat$ln.mass,BMR_dat$BMR,type="p",col="red",xlab="ln(mass)",ylab="BMR",
+plot(BMR_dat$ln.mass,BMR_dat$ln.BMR,type="p",col="red",xlab="ln(mass) of contrasts",ylab="ln(BMR) of contrasts",
      cex.lab=1.5,cex.axis=1.5)
@@
 #--------------------------------------------------------------------------------------------
-# Generating raw data plot btw ln(mass) and BMR from bmr-contrasts.csv
-pdf("./Results/BMR_results/copula_BMR_vs_lnmass.pdf")
+# Generating copula plot btw ln(mass) and ln(BMR) from bmr-contrasts.csv
+pdf("./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf")
 v_BMR<-phylo_dat_fn(data=BMR_dat,i=4,j=3,ploton = T) 
@@
 \begin{center}
-\includegraphics[width=3 in]{./Results/BMR_results/rawdataplot_BMR_vs_lnmass.pdf}
-\includegraphics[width=3 in]{./Results/BMR_results/copula_BMR_vs_lnmass.pdf}
-\caption[\textbf{(A)} Plot of the raw BMR vs.
-ln(mass) data \textbf{(B)} Copula plot for BMR dataset]{\textbf{(A)} Plot of the raw BMR vs.
-ln(mass) data \textbf{(B)} Copula plot for BMR dataset.
+\includegraphics[width=3 in]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
+\includegraphics[width=3 in]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
+\caption[\textbf{(A)} Plot of the raw ln(BMR) of contrasts vs.
+ln(mass) of contrasts \textbf{(B)} Copula plot for ln(BMR) vs.
+ln(mass) contrasts]{\textbf{(A)} Plot of the raw ln(BMR) of contrasts vs.
+ln(mass) of contrasts \textbf{(B)} Copula plot for ln(BMR) vs.
+ln(mass) contrasts.
 Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_BMR}}
@@
 \caption[Testing non-parametric statistics for soilCN data]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{soil C and N data}.
-As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
-When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
+As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
+When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
 When $\cor$ values (respectively, $\Ps$ or $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, stronger than or weaker than for $\Ps$ or $\Dsq$) expected from a normal-copula null hypothesis.
@@
       caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.
-Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (Methods).
+Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (\\nameref{MT-Methods}).
 One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates.
@@
                           level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
-saveRDS(BivMS_BMR,file="./Results/BMR_results/BivMS_BMR_lnmass.RDS")
+saveRDS(BivMS_BMR,file="./Results/BMR_results/BivMS_lnBMR_vs_lnmass.RDS")
 ```
@@
 # Now do NPA for v_BMR
-# If "F"/"SBB8" is the best fit : then it should be symmetric tail dep (0) : Then still I need to do NPA ?
-# May be : because it will also tell me how they are different from "N" copula in different sections?
 source("bivfunctionplot.R")
@@
 \caption[Testing non-parametric statistics for BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{BMR data}.
-As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
+As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
+When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
 \label{fig_BMR_nonparam}}

### 2018-05-19 Paper.Rmd
adding results and description for anolis data

@@
 #### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}
-<!--we have to add description here after DAN fixes bug(?) in bivfunctionplot.R-->
+
+As described in \nameref{Methods}, each of our statistics, are computed for $(l_b,u_b)$ in the 
+ranges $0-0.25$, $0.25-0.5$, $0.5-0.75$, $0.75-1$ (with $4$ number of bins).
+Anolis data set has lesser number of data compared to other two bivariate data set, so fewer number of bins were chosen to measure tail-dependence.
+Those statistics were compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different
+runs for Kendall and Spearman (Fig. \ref{SM-fig_contrasts_67_nonparam}).
+
+Results indicate that tail dependence of 
+data is stronger in both the lower and upper tails than would be expected under a 
+null hypothesis of a normal copula, answering the second part of question 1.
+
+We also tested statistically for asymmetry of tail dependence, which seems visible in Fig.
+\ref{SM-fig_raw_cop_AVG_humerus_ltoe}(B).
+Results (Table \ref{SM-tab_contrasts_67_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.
 

### 2018-05-19 SupportingInformation.Rmd
adding results and description for anolis data

@@
 [**bmr-contrasts.csv**](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) : 
-This contains the trait contrasts
-from phylogenetically independent contrasts [@SukumaranH2010] 
+This contains the trait contrasts from phylogenetically independent contrasts [@SukumaranH2010] 
 of data reported by [@GenoudIM2017] on the basal metabolic rate (BMR)
@@
 
-**contrasts.csv** : Raw data on measurements of anatomical
-features of several species of *Anolis* lizards was downloaded
+[**contrasts.csv**](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) : 
+Raw data on measurements of anatomical features of several species of *Anolis* lizards was downloaded
 from the Dryad data package [@MahlerIRL2013Data] associated with
@@
 This data contains path length and different traits for total `r nrow(phylo_dat)` contrasts.
-Table \ref{tab_rawdata_phylo} displays the raw data for first few contrasts.
+Table \ref{tab_rawdata_phylo} displays the raw data for first few contrasts for two selective phylogenetic traits, namely, AVG.ltoe.IV (average length of fourth hindtoe) and AVG.humerus (average length of humerus).
 The raw data are shown in Figure \ref{fig_raw_cop_AVG_humerus_ltoe}(A).
@@
              format='pandoc',
-             caption = "Fitting info for BMR dataset. \\label{tab_BMRfit}",
+             caption = "Fitting info for **BMR dataset**. \\label{tab_BMRfit}",
              booktabs=T)
@@
              format='pandoc',
-             caption = "Fitting info for contrasts between AVG.ltoe.IV and AVG.humerus dataset. \\label{tab_contrasts_67fit}",
+             caption = "Fitting info for **Anolis data set** with contrasts for AVG.ltoe.IV and AVG.humerus phylogenetic traits. \\label{tab_contrasts_67fit}",
              booktabs=T)
@@
 
+stat_contrasts_67_ga_numbin5<-bivfunctionplot(v=v_contrasts_67,
+                            resloc="./Results/Anolis_results/contrasts_csv_results/",
+                            nametag="contrasts_67_numbin5",numbin = 5)
+
 stat_contrasts_67_ga_numbin6<-bivfunctionplot(v=v_contrasts_67,
@@
 \caption[Testing non-parametric statistics for anolis data]{Nonparametric tests for tail dependence and other deviations from a normal copula for anolis data.
-As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
+As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.25$, $0.25-0.5$, $0.5-0.75$, $0.75-1$ for $4$ number of bins (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
 \label{fig_contrasts_67_nonparam}}
@@
 \end{figure}
-
-<!--
-{Nonparametric tests for tail dependence and other deviations from a normal copula for anolis data.
-As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for $1000$ normal-copula surrogate datasets (dashed lines show $0.025$ and $0.975$ quantiles of these results and dotted lines show $0.005$ and $0.995$ quantiles).
-\label{fig_contrasts_67_nonparam}}
-\end{center}
-\end{figure}
--->
 
@@
              format="pandoc",
-             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_contrasts_67_asym}",
+             caption = "**Anolis data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_contrasts_67_asym}",
              booktabs=T)
@@
 This chunks are just to see the tables for all numbin variation for anolis data set: It appears that numbin=4 is optimized, when numbin=2, >= 7, there is no significant asymetry in tail dep in all 3 types of stats.
+
 ```{r tab_contrasts_67_asym67_numbin2, echo=F, results="markup"}
@@
 #             caption = "numbin_3,anolis \\label{tab_contrasts_67_asym3}",
+#             booktabs=T)
+```
+
+```{r tab_contrasts_67_asym67_numbin5, echo=F, results="markup"}
+#tab_contrasts_67_asym5<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin5)
+#knitr::kable(tab_contrasts_67_asym5,
+#            format="pandoc",
+#             caption = "numbin_5,anolis \\label{tab_contrasts_67_asym5}",
 #             booktabs=T)
@@
              #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid.",
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid. Copulas 
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **abundances of the green spruce aphid**. Copulas 
 considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
@@
              format='pandoc',
-             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for abundances of the green spruce aphid.\\label{tab_count_AICw}",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_AICw}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for abundances of the green spruce aphid, for comparison with Table \\ref{tab_count_AICw}.
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **abundances of the green spruce aphid**, for comparison with Table \\ref{tab_count_AICw}.
 \\label{tab_count_AICw_normal}",
@@
              format='pandoc',
-             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for abundances of the green spruce aphid.
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **abundances of the green spruce aphid**.
 Tests were based on bootstrapping as described in Methods.\\label{tab_count_p_CvM}",
@@
              format='pandoc',
-             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for abundances of the green spruce aphid.
+             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **abundances of the green spruce aphid**.
 Tests were based on bootstrapping as described in Methods.\\label{tab_count_p_KS}",
@@
              format='pandoc',
-             caption = "Model-averaged lower tail dependence statistic for each location pair for abundances of the green spruce aphid.\\label{tab_count_LTdep_AICw}",
+             caption = "Model-averaged lower tail dependence statistic for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_LTdep_AICw}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Model averaged upper tail dependence statistic for each location pair for abundances of the green spruce aphid.\\label{tab_count_UTdep_AICw}",
+             caption = "Model averaged upper tail dependence statistic for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_UTdep_AICw}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for abundances of the green spruce aphid.\\label{tab_count_LTmUT}",
+             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_LTmUT}",
              booktabs=TRUE
@@
              format = "pandoc",
-             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for abundances of the green spruce aphid.\\label{tab_count_resamp}",
+             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **abundances of the green spruce aphid**.\\label{tab_count_resamp}",
              booktabs=TRUE)
@@
              #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **first flight data for the leaf-curling plum aphid**.
 Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations given in \\nameref{MT-Methods}.
@@
              format='pandoc',
-             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_AICw}",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_AICw}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for first flight data for the leaf-curlung plum aphid, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **first flight data for the leaf-curlung plum aphid**, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
              booktabs=TRUE
@@
 
+\newpage
 ```{r tab_ff_p_CvM, echo=F, results='markup'}
@@
              format='pandoc',
-             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for first flight data for the leaf-curling plum aphid.
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **first flight data for the leaf-curling plum aphid**.
 Tests were based on bootstrapping as described in Methods.\\label{tab_ff_p_CvM}",
@@
              format='pandoc',
-             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for first flight data for the leaf-curling plum aphid.
+             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **first flight data for the leaf-curling plum aphid**.
 Tests were based on bootstrapping as described in Methods.\\label{tab_ff_p_KS}",
@@
              format='pandoc',
-             caption = "Model-averaged lower tail dependence statistic for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_LTdep_AICw}",
+             caption = "Model-averaged lower tail dependence statistic for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_LTdep_AICw}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Model-averaged upper tail dependence statistic for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_UTdep_AICw}",
+             caption = "Model-averaged upper tail dependence statistic for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_UTdep_AICw}",
              booktabs=TRUE
@@
              format='pandoc',
-             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for first flight data for the leaf-curling plum aphid.\\label{tab_ff_LTmUT}",
+             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_LTmUT}",
              booktabs=TRUE
@@
              format = "pandoc",
-             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for first flight data for the leaf-curling plum aphid.\\label{tab_ff_resamp}",
+             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_resamp}",
              booktabs=TRUE)
@@
              #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundance data for \\textit{Ceratium furca} in UK seas.",
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundance data for \\textit{Ceratium furca} in UK seas.
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **abundance data for \\textit{Ceratium furca} in UK seas**.
 See \\nameref{MT-Methods} for copula abbreviations.
@@
              digits=3,
-             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for abundance data for *Ceratium furca* in UK seas.\\label{tab_pns_AICw}",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_AICw}",
              booktabs=TRUE)
@@
              digits=3,
-             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for abundance data for *Ceratium furca* in UK seas.\\label{tab_pns_AICw_normal}",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_AICw_normal}",
              booktabs=TRUE)
@@
              digits=3,
-             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for abundance data for *Ceratium furca* in UK seas.
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **abundance data for *Ceratium furca* in UK seas**.
 Tests were based on bootstrapping as described in Methods.\\label{tab_pns_p_CvM}",
@@
 the goodness of fit of the lowest-AIC copula family for each pair of
-locations for abundance data for *Ceratium furca* in UK seas. Tests
+locations for **abundance data for *Ceratium furca* in UK seas**. Tests
 were based on bootstrapping as described in Methods.\\label{tab_pns_p_KS}",
@@
              digits=3,
-             caption = "Model-averaged lower tail dependence statistic for each location pair for abundance data for *Ceratium furca* in UK seas.\\label{tab_pns_LTdep_AICw}",
+             caption = "Model-averaged lower tail dependence statistic for each location pair for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_LTdep_AICw}",
              booktabs=TRUE)
@@
              digits=3,
-             caption = "Model-averaged upper tail dependence statistic for each location pair for abundance data for *Ceratium furca* in UK seas.\\label{tab_pns_UTdep_AICw}",
+             caption = "Model-averaged upper tail dependence statistic for each location pair for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_UTdep_AICw}",
              booktabs=TRUE)
@@
              caption = "Model averaged lower tail dependence minus model
-averaged upper tail dependence statistic for each location pair for *Ceratium furca* in UK seas.\\label{tab_pns_LTmUT}",
+averaged upper tail dependence statistic for each location pair for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_LTmUT}",
              booktabs=TRUE)
@@
              format = "pandoc",
-             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for abundance data for *Ceratium furca* in UK seas.\\label{tab_pns_resamp}",
+             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_resamp}",
              booktabs=TRUE)
@@
              #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.",
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **methane-flux data**.
 See \\nameref{MT-Methods} for copula abbreviations.
@@
              digits=3,
-             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for methane-flux data.\\label{tab_methane_AICw}",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **methane-flux data**.\\label{tab_methane_AICw}",
              booktabs=TRUE)
@@
              digits=3,
-             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for methane-flux data.\\label{tab_methane_AICw_normal}",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **methane-flux data**.\\label{tab_methane_AICw_normal}",
              booktabs=TRUE)
@@
              digits=3,
-             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for methane-flux data.
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **methane-flux data**.
 Tests were based on bootstrapping as described in Methods.\\label{tab_methane_p_CvM}",
@@
 the goodness of fit of the lowest-AIC copula family for each pair of
-locations for methane-flux data. Tests were based on bootstrapping as described in Methods.\\label{tab_methane_p_KS}",
+locations for **methane-flux data**. Tests were based on bootstrapping as described in Methods.\\label{tab_methane_p_KS}",
              booktabs=TRUE)
@@
              digits=3,
-             caption = "Model-averaged lower tail dependence statistic for each location pair for methane-flux data.\\label{tab_methane_LTdep_AICw}",
+             caption = "Model-averaged lower tail dependence statistic for each location pair for **methane-flux data**.\\label{tab_methane_LTdep_AICw}",
              booktabs=TRUE)
@@
              digits=3,
-             caption = "Model-averaged upper tail dependence statistic for each location pair for methane-flux data.\\label{tab_methane_UTdep_AICw}",
+             caption = "Model-averaged upper tail dependence statistic for each location pair for **methane-flux data**.\\label{tab_methane_UTdep_AICw}",
              booktabs=TRUE)
@@
              caption = "Model averaged lower tail dependence minus model
-averaged upper tail dependence statistic for each location pair for methane-flux data.\\label{tab_methane_LTmUT}",
+averaged upper tail dependence statistic for each location pair for **methane-flux data**.\\label{tab_methane_LTmUT}",
              booktabs=TRUE)
@@
              format = "pandoc",
-             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for methane-flux data.\\label{tab_methane_resamp}",
+             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **methane-flux data**.\\label{tab_methane_resamp}",
              booktabs=TRUE)

### 2018-05-25 Paper.Rmd
adding description for effect on extrisk section

@@
 
+In order to study the effects of non-gaussian copula structure on extinction risk for population of a species at different locations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
+<!--did I write correct spelling for lewton model?-->
+
+\begin{equation}\label{eq.extrisk_model}
+P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
+\end{equation}
+
+$r$ is the liner growth rate and $K$ is the carrying capacity of the environment.
+$\epsilon$ is the environmental noise which may take strong right tail (or left tail) dependence.
+We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix.
+$D$ can take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch can interact with the rest of all (we called it as global dispersion).
+We simulated Eq.
+\ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $25$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;
+
 - LC model (r=0) :
 
-    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
+    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
 [see first plot of Fig.
@@
 
-    * For ricker model when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
+    * For ricker model when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
 For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
@@
 
-With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig.
+With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig.
 \ref{SM-fig_risk_numlocs_effect_local_disp} - Fig.

### 2018-06-08 SupportingInformation.Rmd
adding loecke_data description

@@
 
-Carbon and nitrogen in soil at many locations across CONUS (Terry to augment this description).
-The raw data are shown in Figure \ref{fig_raw_cop_soilCN}(A).
-The copula is shown in Figure \ref{fig_raw_cop_soilCN}(B).
+To explore the correlation of soil organic C and total soil N across a wide geographic range (conterminous
+US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
+database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
+stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the **6???** sites were included in this
+study.
+
+The raw data are shown in Figure \ref{fig_raw_cop_soilCN}(A). The copula is shown in Figure \ref{fig_raw_cop_soilCN}(B).
 
@@
 ## Methane-flux data  \label{Methane_data}
-   * Methane-flux at many locations (Prof. Terry to augment this description).
+
+Methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
+Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
+Additional information on the history of the site and soil types is detailed in @Jarecke2016.
+Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at **???(check after filtering to 50 fluxes)** locations across a hydrologic gradient (upland to seasonal wetland) at
+daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
+additional details???** on greenhouse gas flux methods for this site.
+
 * We threw out all locations that had fewer than 50 dates data.

### 2018-06-08 SupportingInformation.Rmd
minor change

@@
 database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
-stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the **6???** sites were included in this
+stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
 study.
@@
 Additional information on the history of the site and soil types is detailed in @Jarecke2016.
-Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at **???(check after filtering to 50 fluxes)** locations across a hydrologic gradient (upland to seasonal wetland) at
+Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at
 daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for

### 2018-06-08 Paper.Rmd
minor change

@@
 SBB1 is just opposite of BB1 in tail dependency.
-So, the result was expected, because the contributions came from more or less equally weighted BB1 and SBB1.
+So, the result was expected, because the contributions came from more or less equally weighted BB1 and SBB1 with nearly same tail dependencies.
 We also tested that lowest-AIC copula BB1 was a good-enough fit, giving $p$-value `r BivMS_BMR$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_BMR$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
@@
 null hypothesis of a normal copula, answering the second part of question 1.
-But Table \ref{SM-tab_BMR_asym} indicates that tail dependence for the BMR data copula is significantly (95% CI) not different from symmetric normal copula and thus consistent with model selection results.
+But Table \ref{SM-tab_BMR_asym} indicates that difference in tail dependence for the BMR data copula is significantly (95% CI) not different from that of the symmetric normal copula and thus consistent with model selection results.
 

### 2018-06-14 Paper.Rmd
changes in contrasts datasets handling

--- 2018-05-25 Paper.Rmd
@@
 and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_BMRfit}).
-The table shows that BB1 and SBB1 are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+The table shows that BB6 and BB8 are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
 
-The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was `r BivMS_BMR$relLTdep_AICw - BivMS_BMR$relUTdep_AICw`, indicating nearly symmetric tail dependence.
-BB1, being as a two parameter copula,  has both lower and upper tail dependence and depending upon which parameter is greater it shows comparatively greater lower or upper tail dependence.
-SBB1 is just opposite of BB1 in tail dependency.
-So, the result was expected, because the contributions came from more or less equally weighted BB1 and SBB1.
-We also tested that lowest-AIC copula BB1 was a good-enough fit, giving $p$-value `r BivMS_BMR$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_BMR$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was `r BivMS_BMR$relLTdep_AICw - BivMS_BMR$relUTdep_AICw`, indicating more upper tail dependence.
+We also tested that lowest-AIC copula BB8 was a good-enough fit, giving $p$-value `r BivMS_BMR$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_BMR$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
 The next section looks at tail dependence in a non-parametric way.
@@
 null hypothesis of a normal copula, answering the second part of question 1.
-But Table \ref{SM-tab_BMR_asym} indicates that tail dependence for the BMR data copula is significantly (95% CI) not different from symmetric normal copula and thus consistent with model selection results.
+Table \ref{SM-tab_BMR_asym} indicates stronger upper tail dependence for the BMR data copula which is significantly (95% CI) different from that of the symmetric normal copula and thus consistent with model selection results.
 
@@
 
-The test of indepedence gave $p=$ `r BivMS_contrasts_67$IndepTestRes` (to within the precision available 
+The test of indepedence gave $p=0$ <!--`r BivMS_contrasts_67$IndepTestRes` this is e-15 could not compile latex--> (to within the precision available 
 from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
 and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_contrasts_67fit}).
-The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+The table shows that several copulas (mostly J and SC) are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
 
-The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`.
+The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was <!--$\num{1e-10}$--> `r round(BivMS_contrasts_67$relLTdep_AICw,3)` <!--`r BivMS_contrasts_67$relLTdep_AICw` this is e-4 could not compile latex--> and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`.
 These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.

### 2018-06-14 SupportingInformation.Rmd
changes in contrasts datasets handling

@@
 ln(mass) of contrasts \textbf{(B)} Copula plot for ln(BMR) vs.
-ln(mass) contrasts]{\textbf{(A)} Plot of the raw ln(BMR) of contrasts vs.
-ln(mass) of contrasts \textbf{(B)} Copula plot for ln(BMR) vs.
-ln(mass) contrasts.
+ln(mass) contrasts taken as their absolute values]{\textbf{(A)} Plot of the raw ln(BMR) of contrasts vs.
+ln(mass) of contrasts \textbf{(B)} Copula plot for absolute values of ln(BMR) vs.
+ln(mass) for all $`r nrow(v_BMR)`$ contrasts.
 Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_BMR}}
@@
 \textbf{(B)} Copula plot of contrasts for AVG.ltoe.IV vs.
-contrasts for AVG.humerus phylogenetic traits]{\textbf{(A)} Plot of the contrasts for AVG.ltoe.IV vs.
+contrasts for AVG.humerus phylogenetic traits(taken as absolute values)]{\textbf{(A)} Plot of the contrasts for AVG.ltoe.IV vs.
 contrasts for AVG.humerus phylogenetic traits.
-\textbf{(B)} Copula plot of contrasts for AVG.ltoe.IV vs.
+\textbf{(B)} Copula plot of absolute values of contrasts for AVG.ltoe.IV vs.
 contrasts for AVG.humerus phylogenetic traits.
@@
                             resloc="./Results/Anolis_results/contrasts_csv_results/",
-                            nametag="contrasts_67_numbin2",numbin = 3)
+                            nametag="contrasts_67_numbin3",numbin = 3)
 

### 2018-06-14 phylo_dat_fn.R

@@ -6,34 +6,39 @@ source("getcopula.R")
 
 # Note : to get copula between traits use usual data cleaning but to find copula between pathlength and trait 
 # take absolute value of trait only
+# Also we keep the first trait [i. e., index=i of (i,j)] positive for all contrsts #asymmetric way
+
+#always take absolute values for all traits # symmetric way
 
 # Input :
 #        data : a dataframe 
 #        i, j : column indices of data frame to get bivariate data set
 #        ploton : logical for option plots of raw data and it's copula representation
+
 # Output : 
 #        a cop rank matrix for the raw data
+
 phylo_dat_fn<-function(data,i,j,ploton=T){
   
   data<-data[,2:dim(data)[2]]
   
   x<-c(i,j)
-  if(1 %in% x){
-    ind<-which(x!=1)
-    data[,x[ind]]<-abs(data[,x[ind]]) # replace by its absolute value if it's comparison btw pathlength & traits
-    d<-data[,c(i,j)] 
-  }else{
-    d<-data[,c(i,j)] 
-  }
+  
+  d<-data[,x]
+  d<-abs(d) #always take absolute values for comparison 
+  
+#  if(1 %in% x){
+#    ind<-which(x!=1)
+#    data[,x[ind]]<-abs(data[,x[ind]]) # replace by its absolute value if it's comparison btw pathlength & traits
+#    d<-data[,c(i,j)] 
+#  }else{
+#    d<-data[,c(i,j)]
+#    ni<-which(d[,1]<0) # negative row index of d matrix with respect to 1st column
+#    d[ni,]<- -d[ni,]   # making all trait positive always for the 1st column
+#  }
 
   v_phylo<-getcopula(d=d,rankon = T,ploton = ploton)
   
-  #if(ploton==T){
-    #op<-par(mfrow=c(1,2))
-    #plot(d[,1],d[,2],col="blue")
-    #getcopula(d=d,rankon = T,ploton = T)
-    #par(op)
-  #}
   return(v_phylo)
 }
 

### 2018-06-14 SupportingInformation.Rmd
minor edit in text

@@
 #class(BMR_dat)
-knitr::kable(head(BMR_dat,3),format='pandoc',
+knitr::kable(head(BMR_dat[,c(1,4,5)] ,3),format='pandoc',
              caption = "First few rows of raw BMR data.\\label{tab_rawdata_BMR}",
@@
 
-This data contains BMR and ln(mass) for total `r nrow(BMR_dat)` contrasts.
 Table \ref{tab_rawdata_BMR} displays the raw data for first few contrasts.
 The raw data are shown in Figure \ref{fig_raw_cop_BMR}(A).
-The copula is shown in Figure \ref{fig_raw_cop_BMR}(B).
+The copula [using absolute values of each column in Table \ref{tab_rawdata_BMR}] is shown in Figure \ref{fig_raw_cop_BMR}(B).
 
@@
 
-This data contains path length and different traits for total `r nrow(phylo_dat)` contrasts.
 Table \ref{tab_rawdata_phylo} displays the raw data for first few contrasts for two selective phylogenetic traits, namely, AVG.ltoe.IV (average length of fourth hindtoe) and AVG.humerus (average length of humerus).
 The raw data are shown in Figure \ref{fig_raw_cop_AVG_humerus_ltoe}(A).
-The copula is shown in Figure \ref{fig_raw_cop_AVG_humerus_ltoe}(B).
+The copula [taking absolute values of each column in Table \ref{tab_rawdata_phylo}] is shown in Figure \ref{fig_raw_cop_AVG_humerus_ltoe}(B).
 

### 2018-06-25 makefile.Rmd
trivial

--- 2018-05-15 makefile.Rmd
@@
 ---
+
+<!--checkpoint package-->
+```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
+library(checkpoint)
+library(captioner)
+library(rmarkdown)
+if (!dir.exists("./.checkpoint/")){
+  dir.create("./.checkpoint/")
+}
+checkpoint("2018-03-29",checkpointLocation = "./")
+```
 

### 2018-06-25 Paper.Rmd

@@
 knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
+options(scipen = 1, digits = 5) #This option round all numbers appeared in the inline r code upto 5th digit
 #seed<-101
@@
 
-The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was <!--$\num{1e-10}$--> `r round(BivMS_contrasts_67$relLTdep_AICw,3)` <!--`r BivMS_contrasts_67$relLTdep_AICw` this is e-4 could not compile latex--> and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`.
+The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`.
 These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.

### 2018-06-26 Paper.Rmd
edits

@@
 title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
-author: "Shyamolina Ghosh, other authors, Daniel Reuman"
+author: "Shyamolina Ghosh, Mark T. Holder, Terrance E. Loecke, Daniel C. Reuman"
 date: "February 18, 2018"
@@
 ---
+
+<!--**DAN: right now we sometimes use the terminology "normal copula" and sometimes 
+"Gaussin copula". Let's choose one and be consistent across the whole ms, including 
+main text and supp mat. I am leaning toward "Gaussin copula" but could be convinced 
+otherwise, it is just the consistency that I want to be sure we get to.-->
+
 updated on `r Sys.Date()`
+
 ```{r setup_Paper, echo=F}
@@
 ```
+
 # Front matter
@@
 
-Reuman:
+Ghosh: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
 
-Ghosh:
+Holder: Department of Ecology and Evolutionary Biology and Biodiversity Institute, University of Kansas, Lawrence, KS, 66045, USA
+
+Loecke: Environmental Studies Program and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA
+
+Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA.
 
@@
 
-## Statistical methods  \label{Methods}
+## Statistical methods for detecting copula structure \label{Methods}
+
+Question 1 from the Introduction was answered via statistical analysis of the
+data described above, using both model selection and nonparametric approaches.
 
@@
 
-For a bivariate dataset $(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps, 
+A model-selection approach based on likelihood and the Akaike Information Criterion (AIC) 
+and Bayesian Information Criterion (BIC) was used to
+help answer question 1 from the Introduction.
+
+For bivariate datasets 
+$(x_i,y_i)$ for $i=1,...,n$ such as the soil C and N data, model selection involved several steps, 
 most implemented using tools from the `VineCopula` package. Code combining these tools is in the
@@
 Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulas.
-Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC) values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection].
-Descriptions of these copula families can be found in introductory texts on copulas [@nelsen2007_copula ; @joe2014_dependence]  and in references cited in `VineCopula`.
+AIC and BIC values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection]. 
+Descriptions of these copula families can be found in introductory texts on copulas [@nelsen2007_copula ; @joe2014_dependence]  and in `VineCopula` documentation. 
 The lower- and upper-tail dependence of the 
@@
 
-<!--Shya please insert some additional material for multivariate datasets - perhaps
-you don't need much because basically you just ran the bivariate analysis on each pair
-of locations, but you need to say that and then maybe a few sentences about how 
-everything was tallied up after having done that ??? -->
+The first part or question 1 from the Introduction was answered for a bivariate dataset 
+$(x_i,y_i)$ for $i=1,...,n$ by comparing the AIC weights of the non-Gaussian copula
+families with that of the Gaussian family. The same was also done using BIC.
+The second part of question 1 was answered by comparing 
+the model-averaged lower- and upper-tail dependence
+values to the lower- and upper-tail dependence values of a Gaussian copula, both of which 
+are 0. The third part of question 1
+was answered by comparing the difference of the model-averaged lower- and upper-tail 
+dependence values to the analogous difference for a Gaussian copula, again 0.
 
-### Non-parametric analysis of tail dependence \label{NPA}
+For multivariate datasets, e.g., the aphid abundance or first flight data, 
+each of which consists of time series 
+from 10 locations...
+<!--***DAN: need to fill this in, prrobably enough to be brief, since these were just done 
+pairwise-->
+
+### Nonparametric analysis of tail dependence \label{NPA}
+
+In addition to the model selection approach described above, nonparametric 
+approaches were also used to answer question 1 from the introduction.
+Using two distinct methods, when results are consistent, provides greater confidence 
+that results are robust.
 
@@
 (see \nameref{Model_selection} for defintions of $u_i$ and $v_i$)
-are related in the left and right tails, respectively, of their 
-distributions [@nelsen2007_copula]. 
+are related to each other in the left and right tails, respectively, of their 
+distributions [@nelsen2007_copula, @joe2014_dependence]. 
 Our nonparametric study of tail dependence uses three
@@
 
-In addition to [basic unit tests](https://github.com/sghosh89/BIVAN/blob/master/Tests.Rmd), we also tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas.
+<!--I deleted the phrase about unit testing because unit testing should be mentioned in the supp mat-->
+We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas.
 Fig.
 \ref{fig_stat_image}(B - D) represent the results for Clayton copulas with $n=35$ data points and $100$ simulations (for details see Appendix \ref{SM-Test_npa_stat}).
+<!--***DAN: There is either too little or too much here in this 
+paragraph. The reader cannot get the main
+idea, really, of what was done for these tests, from this paragraph alone - they will have to
+look at the appendix. They need to be able to get the main idea. So you could either add detail,
+or else take away detail and just say something like "We tested our statistics by testing whether 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed asymmetry of tail dependence 
+known to be present in data simulated from various copulas (Appendix X)" and then stop there.
+Then you would remove panels B-D from Fig. 1. At the moment I prefer this second option, but 
+my preference is not strong, so if you prefer to keep the material and add detail that would be 
+fine.-->
 
-For bivariate datasets such as the soil C and N data, we compared values of the statistics $\cor_{l_b,u_b}$,
-$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to distributions of values of the same statistics computed on surrogate datasets that were produced from the empirical data by randomizing it in a special way.
-The surrogate/randomized datasets had exactly the same marginal distributions as the empirical data and had very similar Kendall or Spearman correlation, but had normal copula structure, and thus our comparisons tested the null hypothesis that our statistics took values on the empirical data no different from what would have been expected if the copula structure of the data were normal, but the data were otherwise statistically unchanged.
+For bivariate datasets, we compared values 
+of the statistics $\cor_{l_b,u_b}$,
+$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to 
+distributions of values of the same statistics computed on surrogate datasets that were produced from 
+the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
+the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
+correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
+but had Gaussian copula structure. Thus our comparisons tested the null hypothesis 
+that our statistics took values on the empirical data no different from what would have been expected 
+if the copula structure of the data were Gaussian, but the data were otherwise 
+statistically unchanged.
+Details of how surrogates were prroduced are in XYZ, and code is in ABC.
+<!--***DAN: I think the below material belongs in supp mat, because the reader can get the
+main idea via the above text, and only some readers will be interested in the detailed 
+algorithm. Please move it there, and then make the above links XYZ and ABC into live links
+to the appropriate locations in the sup mat.
+
 For bivariate datasets, surrogates were produced as follows.
@@
 Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
-is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a permutation 
-of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as those of the original data. 
+is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a 
+permutation 
+of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as 
+those of the original data. 
 Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
@@
 sampling variation) the Spearman correlation of the empirical data.  
+-->
 
 The comparison of one of the
-statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the empirical data, to the distribution
-of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
-provides a test of whether dependence in the part of the distribution specified by $l_b$ and $u_b$ is different from
-what would be expected from a null hypothesis of normal copula structure but unchanged marginals and unchanged 
+statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
+empirical data, to the distribution
+of its values computed on many surrogate datasets 
+provides a test of whether dependence in the part of the distribution specified 
+by $l_b$ and $u_b$ is different from
+what would be expected from a null hypothesis of normal copula structure but 
+unchanged marginals and unchanged 
 Kendall (or Spearman)
-correlation. Significant deviations correspond to deviations from normal copula structure. In particular, deviations 
-using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence different from that of a normal copula;
-likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different from that of a normal copula.  
-The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, as computed on the 
+correlation. Thus the comparison helps answer the first two parts of question 1 
+from the Introduction for bivariate datasets. 
+Significant deviations correspond to deviations from normal 
+copula structure. In particular, deviations 
+using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
+different from that of a normal copula;
+likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
+from that of a normal copula.
+
+The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
+as computed on the 
 empirical data, to the distribution
-of its values computed on many surrogate datasets produced using the normal copula family for $\mathcal{B}(\theta)$
-provides a test of asymmetry of tail dependence.
+of its values on many surrogate datasets 
+provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
+third part of question 1 from the Introduction for bivariate datasets.
 
-For multivariate datasets, e.g., the aphid abundance or first flight data, each of which consists of time series 
-from 10 locations, a spatial resampling scheme was used instead of the surrogate approach described above.
-The scheme is identical to that proposed by [@bjornstadfalck2001] for application in their now 
-widely used spline correlogram method; the same scheme was also used by [@walteretal2017] as part of a method for 
-performing model selection among matrix regression models. For each pair of locations $i \neq j$, the statistic
-of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
-$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The values $s_{i,j}$ were then arranged in
+For multivariate datasets, a spatial resampling scheme was used instead of the surrogate 
+approach described above.
+The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
+widely used spline correlogram method; the same scheme was also used by @walteretal2017 
+as part of a method for 
+performing model selection among matrix regression models. For each pair of 
+locations $i \neq j$, the statistic
+of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
+$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
+$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
+values $s_{ij}$ were then arranged in
 a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
-$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the average strength of 
+$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
+average strength of 
 dependence, across all pairs of distinct locations, for data in the portion of the distributions 
@@
 asymmetry of tail dependence across all pairs of locations. 
-Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a distribution of bootstrapped values.
-Bootstrapping was performed by resampling the locations of measurement, with replacement; this corresponds to 
+Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
+distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
+revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
+or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
+Gaussian null hypothesis, and answering question 1 from the Introduction.
+Bootstrapping was performed by resampling the locations of measurement, with replacement. 
+This corresponds to 
 resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
@@
 
-<!--Dan to insert a description of the surrogate approach here.-->
+## Theoretical models
+
+Question 2 from the Introduction was explored using simple population models.
 

### 2018-06-26 Paper.Rmd
adding response by **SG

@@
 otherwise, it is just the consistency that I want to be sure we get to.-->
+
+<!--**SG: I think normal copula would be better as "N" is the abbreviation for that but for "G" means gumbel not gaussian.
+If you agree with me, then I will made all "gaussian" as "normal" copula in both main and supmat.-->
 
@@
 fall (these are the blue and red lines, respectively, in 
-Fig. \ref{fig_stat_image}(A)). Our statistics quantify the dependence
+Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
 between $u_i$ and $v_i$ in the region bounded by these lines.
@@
 
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
+\caption{Schematic diagram supporting defintions of nonparametric tail dependence statistics \label{fig_stat_image}}
+\end{center}
+\end{figure}
+
+<!--
 \begin{figure}[!h]
@@
 <!--I deleted the phrase about unit testing because unit testing should be mentioned in the supp mat-->
-We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas.
-Fig.
-\ref{fig_stat_image}(B - D) represent the results for Clayton copulas with $n=35$ data points and $100$ simulations (for details see Appendix \ref{SM-Test_npa_stat}).
+We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas (for details see Appendix \ref{SM-Test_npa_stat}).
 <!--***DAN: There is either too little or too much here in this 
@@
 fine.-->
+<!--**SG: Agreed with 2nd option and did things as necessary-->
 
@@
 statistically unchanged.
-Details of how surrogates were prroduced are in XYZ, and code is in ABC.
+Details of how surrogates were prroduced are in Appendix \ref{SM-surrog_test}, and code is in [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R).
 <!--***DAN: I think the below material belongs in supp mat, because the reader can get the
@@
 -->
+<!--**SG: Done-->
 

### 2018-06-26 SupportingInformation.Rmd
trivial
adding response by **SG

@@
 updated on `r Sys.Date()`
-
-<!--checkpoint chunk is moved to makefile.Rmd-->
 
@@
 
+# Surrogate testing with bivariate datasets for non-parametric statistics \label{surrog_test}
+
+For bivariate datasets, surrogates were produced as follows.
+
+- Given data $(x_i,y_i)$ for $i=1,\ldots,n$ and a 
+one-parameter target copula family $\mathcal{B}(\theta)$ (here $\theta$ denotes the parameter), let $\tau$ be the
+Kendall correlation of the $x_i$ with the $y_i$ and find the parameter value $\theta_\tau$ for which 
+$\mathcal{B}(\theta_\tau)$ has Kendall correlation $\tau$ (this is possible for the one-parameter copula families of this 
+study using the `iTau` function of the `VineCopula` package). 
+- Generate data $(a_i,b_i)$, $i=1,\ldots,n$ from 
+$\mathcal{B}(\theta_\tau)$. 
+- Permute the ordered set $(x_1,\ldots,x_n)$ such that the permutation
+$(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ has $\rank(x_{\sigma_x(i)})$ equal to $\rank(a_i)$ for all $i$, where 
+$\rank(x_{\sigma_x(i)})$ is the rank of $x_{\sigma_x(i)}$ in the set $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$, and
+$\rank(a_i)$ is the rank of $a_i$ in the set $(a_1,\ldots,a_n)$ (here $\sigma_x$ is a permutation of the 
+indices $1,\ldots,n$). 
+- Likewise permute $(y_1,\ldots,y_n)$
+such that the permutation
+$(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ has $\rank(y_{\sigma_y(i)})$ equal to $\rank(b_i)$ for all $i$ ($\sigma_y$ is 
+another permutation of $1,\ldots,n$).
+- The surrogate dataset is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.
+
+Our code for this algorithm is [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R) in the 
+[`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
+
+Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
+is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a 
+permutation 
+of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as 
+those of the original data. 
+Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
+Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
+will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
+a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
+as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
+sampling variation) the Spearman correlation of the empirical data.
+
 <!--

### 2018-06-27 Paper.Rmd
adding summary tables for bi & multi-var datasets

@@
 
+```{r tab_bivar_summary, echo=F, results='markup'}
+
+tab_multivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
+knitr::kable(tab_bivar_summary, 
+             format='pandoc',
+             caption = "Summary of results for bivariate data-sets.\\label{tab_bivar_summary}",
+             booktabs=TRUE)
+```
+
+```{r tab_multivar_summary, echo=F, results='markup'}
+
+tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
+knitr::kable(tab_multivar_summary, 
+             format='pandoc',
+             caption = "Summary of results for multivariate data-sets.\\label{tab_multivar_summary}",
+             booktabs=TRUE)
+```
+
 <!--Question 2 analyses-->

### 2018-06-27 SupportingInformation.Rmd
adding summary tables for bi & multi-var datasets

@@
 
+```{r bivar_tab_summary_fn,echo=F}
+bivar_tab_summary_fn<-function(r,rs){
+
+  m<-matrix(NA,1,12)
+  colnames(m)<-c("Indep.p","AIC-best.copula","Highest.AICw",
+                 "N.AICw","p.CvM","p.KS",
+                 "LT_model.avg","UT_model.avg","(LT-UT)_model.avg",
+                 "Rank_Pl_surrog","Rank_Pu_surrog","Rank_Pl-Pu_surrog")
+  m<-as.data.frame(m)
+
+  m$Indep.p<-round(r$IndepTestRes,3)
+  m$`AIC-best.copula`<-r$InfCritRes$copname[which.min(r$InfCritRes$AIC)]
+  m$Highest.AICw<-round(max(r$InfCritRes$AICw),3)
+  m$N.AICw<-round(r$InfCritRes$AICw[1],3)
+  m$p.CvM<-r$GofRes_CvM
+  m$p.KS<-r$GofRes_KS
+  m$LT_model.avg<-round(r$relLTdep_AICw,3)
+  m$UT_model.avg<-round(r$relUTdep_AICw,3)
+  m$`(LT-UT)_model.avg`<-m$LT_model.avg-m$UT_model.avg
+  m$Rank_Pl_surrog<-NA
+  m$Rank_Pu_surrog<-NA
+  m$`Rank_Pl-Pu_surrog`<-NA
+  mt<-as.data.frame(t(m))
+  return(mt)
+
+}
+#----------------------------------
+#calling the above function on 3 multivariable datasets
+m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke)
+m2<-bivar_tab_summary_fn(r=BivMS_BMR)
+m3<-bivar_tab_summary_fn(r=BivMS_contrasts_67)
+tab_bivar_summary<-cbind(m1,m2,m3)
+colnames(tab_bivar_summary)<-c("SoilCN","BMR","Anolis")
+saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
+```
+```
+
+```{r multivar_tab_summary_fn,echo=F}
+source("normal_cop_AICw.R")
+
+multivar_tab_summary_fn<-function(r,rs){
+
+  m<-matrix(NA,1,12)
+  colnames(m)<-c("#pairs","#non-indep","#non-N",
+                 "Avg.best.AICw","Avg.N.AICw","#good.fit",
+                 "#LT>0","#UT>0","#LT-UT>0",
+                 "CI_Pl","CI_Pu","CI_Pl-Pu")
+  m<-as.data.frame(m)
+
+  normal_AICw<-normal_cop_AICw_matrix(r=r)
+
+  d<-dim(r$gfc_numBS)[1]
+  m$`#pairs`<-(d^2)-d
+  m$`#non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
+  m$`#non-N`<-sum(r$info_ord_copcode[,,1]!=1,na.rm = T)
+  r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN
+  m$Avg.best.AICw<-round(mean(r$info_ord_AICw[,,1],na.rm = T),3)
+  normal_AICw[normal_AICw==Inf]<-NaN
+  m$Avg.N.AICw<-round(mean(normal_AICw,na.rm=T),3)
+  temp<-which((r$gfc_p_CvM>0.01) & (r$gfc_p_KS>0.01),arr.ind=T)
+  m$`#good.fit`<-dim(temp)[1]
+  m$`#LT>0`<-sum(r$LTdep_AICw>0,na.rm=T)
+  m$`#UT>0`<-sum(r$UTdep_AICw>0,na.rm=T)
+  m$`#LT-UT>0`<-sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)
+  m$CI_Pl<-list(c(round(rs$numericdf[5,3],3),round(rs$numericdf[5,4],3)))
+  m$CI_Pu<-list(c(round(rs$numericdf[6,3],3),round(rs$numericdf[6,4],3)))
+  m$`CI_Pl-Pu`<-list(c(round(rs$numericdf[10,3],3),round(rs$numericdf[10,4],3)))
+
+  mt<-as.data.frame(t(m))
+  return(mt)
+
+}
+#----------------------------------
+#calling the above function on 4 multivariable datasets
+m1<-multivar_tab_summary_fn(r=RES_aphid_count,rs=stat_aphid_count)
+m2<-multivar_tab_summary_fn(r=RES_aphid_ff,rs=stat_aphid_ff)
+m3<-multivar_tab_summary_fn(r=RES_plankton_north_sea,rs=stat_plankton_north_sea)
+m4<-multivar_tab_summary_fn(r=RES_methane,rs=stat_methane)
+tab_multivar_summary<-cbind(m1,m2,m3,m4)
+colnames(tab_multivar_summary)<-c("Aphid_abundance","Aphid_firstflight","Plankton","Methane-flux")
+saveRDS(tab_multivar_summary,"./Results/tab_multivar_summary.RDS")
+```
+
 <!--

### 2018-06-27 Paper.Rmd
summary table and sample plots inclusion

@@
 
+\begin{figure}[!h]
+\textbf{ \hspace{4.3 cm} (A) \hspace{4.3 cm} (B) \hspace{4 cm} (C) } \\
+\vspace{-0.4 cm}
+\begin{center}
+\includegraphics[width=5 cm]{./Results/SoilCNraw.pdf}
+\includegraphics[width=5 cm]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
+\includegraphics[width=5 cm]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}\\
+\end{center}
+\textbf{ \hspace{4.3 cm} (D) \hspace{4.3 cm} (E) \hspace{4 cm} (F) } \\
+\vspace{-0.4 cm}
+\begin{center}
+\includegraphics[width=5 cm]{./Results/SoilCNcop.pdf}
+\includegraphics[width=5 cm]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
+\includegraphics[width=5 cm]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
+\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} ln(BMR) of contrasts vs.
+ln(mass) of contrasts, \textbf{(C)} contrasts for AVG.ltoe.IV vs.
+contrasts for AVG.humerus phylogenetic traits.
+Lower panel \textbf{(D - F)} shows copula plots for the corresponding datasets of the upper panel.
+These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of contrasts for last two datasets (see \nameref{Model_selection} of manuscript).
+\label{fig_biv_raw_cop}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
+\vspace{-0.4 cm}
+\begin{center}
+\includegraphics[width=7 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
+\includegraphics[width=7 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
+\end{center}
+\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
+\vspace{-0.4 cm}
+\begin{center}
+\includegraphics[width=7 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
+\includegraphics[width=7 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
+\caption{Upper panel shows raw data plots for \textbf{(A)} green spruce aphid-abundance data from location pair(4,11) and \textbf{(B)} leaf-curling plum aphid-firstflight data from location pair(2,5).
+Lower panel \textbf{(C - D)} shows copula plots for the corresponding datasets of the upper panel.
+These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of data-points available for both location at same year (see \nameref{Model_selection} of manuscript).
+\label{fig_aphid_raw_cop}}
+\end{center}
+\end{figure}
+
 # Results
@@
 
-tab_multivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
+tab_biivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
 knitr::kable(tab_bivar_summary, 

### 2018-06-27 SupportingInformation.Rmd
summary table and sample plots inclusion

@@
 ```
-```
 
@@
 
-  m<-matrix(NA,1,12)
+  m<-matrix(NA,1,17)
   colnames(m)<-c("#pairs","#non-indep","#non-N",
                  "Avg.best.AICw","Avg.N.AICw","#good.fit",
-                 "#LT>0","#UT>0","#LT-UT>0",
-                 "CI_Pl","CI_Pu","CI_Pl-Pu")
+                 "CI0.025_LT","CI0.975_LT","CI0.025_UT","CI0.975_UT",
+                 "#LT-UT>0",
+                 "CI0.025_Pl","CI0.975_Pl","CI0.025_Pu","CI0.975_Pu","CI0.025_Pl-Pu","CI0.975_Pl-Pu")
   m<-as.data.frame(m)
@@
   m$`#good.fit`<-dim(temp)[1]
-  m$`#LT>0`<-sum(r$LTdep_AICw>0,na.rm=T)
-  m$`#UT>0`<-sum(r$UTdep_AICw>0,na.rm=T)
+  t<-r$LTdep_AICw[is.finite(r$LTdep_AICw)]
+  CI_LT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
+  m$CI0.025_LT<-round(CI_LT[1],3)
+  m$CI0.975_LT<-round(CI_LT[2],3)
+  t<-r$UTdep_AICw[is.finite(r$UTdep_AICw)]
+  CI_UT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
+  m$CI0.025_UT<-round(CI_UT[1],3)
+  m$CI0.975_UT<-round(CI_UT[2],3)
   m$`#LT-UT>0`<-sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)
-  m$CI_Pl<-list(c(round(rs$numericdf[5,3],3),round(rs$numericdf[5,4],3)))
-  m$CI_Pu<-list(c(round(rs$numericdf[6,3],3),round(rs$numericdf[6,4],3)))
-  m$`CI_Pl-Pu`<-list(c(round(rs$numericdf[10,3],3),round(rs$numericdf[10,4],3)))
+  m$CI0.025_Pl<-round(rs$numericdf[5,3],3)
+  m$CI0.975_Pl<-round(rs$numericdf[5,4],3)
+  m$CI0.025_Pu<-round(rs$numericdf[6,3],3)
+  m$CI0.975_Pu<-round(rs$numericdf[6,4],3)
+  m$`CI0.025_Pl-Pu`<-round(rs$numericdf[10,3],3)
+  m$`CI0.975_Pl-Pu`<-list(round(rs$numericdf[10,4],3))
+  #m$`CI0.975_Pl-Pu`<-round(rs$numericdf[10,4],3)
+  #m$CI_Pl<-list(c(round(rs$numericdf[5,3],3),round(rs$numericdf[5,4],3)))
+  #m$CI_Pu<-list(c(round(rs$numericdf[6,3],3),round(rs$numericdf[6,4],3)))
+  #m$`CI_Pl-Pu`<-list(c(round(rs$numericdf[10,3],3),round(rs$numericdf[10,4],3)))
 
@@
 saveRDS(tab_multivar_summary,"./Results/tab_multivar_summary.RDS")
+```
+
+```{r sample_plot_aphid_count_ff,echo=F,results="hide"}
+#-------------------------------------------------------------------------------
+# code to plot raw data and copula plot for aphid count @location pair (4,11)
+# bestfit cop : SJ
+
+d1<-data_aphid_count$`Green spruce aphid`[[4]]
+d2<-data_aphid_count$`Green spruce aphid`[[11]]
+pdf("./Results/Aphid_count_raw_loc_4_11.pdf",height = 8,width=8)
+op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
+plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Aphid abundance from location 4",ylab="Aphid abundance from location 11",cex.lab=1.5,cex.axis=1.5)
+par(op)
+dev.off()
+
+source("./vivj_matrix.R")
+m<-vivj_matrix(d_allsp = data_aphid_count,sp=10,i=4,j=11)
+pdf("./Results/Aphid_count_cop_loc_4_11.pdf",height = 8,width=8)
+op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
+plot(m[,1],m[,2],type="p",col="red",xlab="u",ylab="v",
+     cex.lab=2,cex.axis=1.5,
+     xlim=c(0,1),ylim=c(0,1),asp=1)
+rect(0,0,1,1)
+lines(c(0,1),c(0,1),type='l')
+par(op)
+dev.off()
+
+# code to plot raw data and copula plot for aphid ff @location pair (2,5)
+# bestfit cop : J
+d1<-data_aphid_ff$`Leaf-curling plum aphid`[[2]]
+d2<-data_aphid_ff$`Leaf-curling plum aphid`[[5]]
+pdf("./Results/Aphid_ff_raw_loc_2_5.pdf",height = 8,width=8)
+op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
+plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Aphid first-flight from location 2",ylab="Aphid first-flight from location 5",cex.lab=1.5,cex.axis=1.5)
+par(op)
+dev.off()
+
+m<-vivj_matrix(d_allsp = data_aphid_ff,sp=11,i=2,j=5)
+pdf("./Results/Aphid_ff_cop_loc_2_5.pdf",height = 8,width=8)
+op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
+plot(m[,1],m[,2],type="p",col="red",xlab="u",ylab="v",
+     cex.lab=2,cex.axis=1.5,
+     xlim=c(0,1),ylim=c(0,1),asp=1)
+rect(0,0,1,1)
+lines(c(0,1),c(0,1),type='l')
+par(op)
+dev.off()
 ```

### 2018-06-27 Paper.Rmd
replaced Gaussian by Normal and some text addition

@@
 <!--**SG: I think normal copula would be better as "N" is the abbreviation for that but for "G" means gumbel not gaussian.
-If you agree with me, then I will made all "gaussian" as "normal" copula in both main and supmat.-->
+I made all "Gaussian" as "Normal" copula in both main and supmat.-->
 
@@
 1) Do datasets in ecology and related fields have non-Gausian copula srtructure?
-Do they show tail dependence distinct from that of a Gaussian copula, and in particular do they show asymmetric tail dependence?
-2) If so, what are some possible causes of non-Gaussian copula structure?
-And 3) what are the consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?
+Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence?
+2) If so, what are some possible causes of non-Normal copula structure?
+And 3) what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
 
@@
 The first part or question 1 from the Introduction was answered for a bivariate dataset 
-$(x_i,y_i)$ for $i=1,...,n$ by comparing the AIC weights of the non-Gaussian copula
-families with that of the Gaussian family. The same was also done using BIC.
+$(x_i,y_i)$ for $i=1,...,n$ by comparing the AIC weights of the non-Normal copula
+families with that of the Normal family. The same was also done using BIC.
 The second part of question 1 was answered by comparing 
 the model-averaged lower- and upper-tail dependence
-values to the lower- and upper-tail dependence values of a Gaussian copula, both of which 
+values to the lower- and upper-tail dependence values of a Normal copula, both of which 
 are 0. The third part of question 1
 was answered by comparing the difference of the model-averaged lower- and upper-tail 
-dependence values to the analogous difference for a Gaussian copula, again 0.
+dependence values to the analogous difference for a Normal copula, again 0.
 
 For multivariate datasets, e.g., the aphid abundance or first flight data, 
-each of which consists of time series 
-from 10 locations...
+each of which consists of time series from 10 locations, we carried out the same bivariate analysis
+for all possible pairwise combinations of those locations and compared AIC-weights and tail-dependence of best-fit(minimum AIC) copula with that of the Normal copula.
+We also counted for how many location pairs model -averaged lower tail dependence is greater than model-averaged upper-tail dependence or vice-versa.
+For example, for green-spruce aphid abundance data we found major lower tail dependence whereas leaf-curling plum aphid's first flight data had significant upper tail dependence at all location pairs (for details see Table \ref{tab_multivar_summary}).
 <!--***DAN: need to fill this in, prrobably enough to be brief, since these were just done 
 pairwise-->
+<!--**SG: done-->
 
@@
 correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
-but had Gaussian copula structure. Thus our comparisons tested the null hypothesis 
+but had Normal copula structure. Thus our comparisons tested the null hypothesis 
 that our statistics took values on the empirical data no different from what would have been expected 
-if the copula structure of the data were Gaussian, but the data were otherwise 
+if the copula structure of the data were Normal, but the data were otherwise 
 statistically unchanged.
@@
 or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
-Gaussian null hypothesis, and answering question 1 from the Introduction.
+Normal null hypothesis, and answering question 1 from the Introduction.
 Bootstrapping was performed by resampling the locations of measurement, with replacement. 
@@
 
-- Do datasets in ecology and related fields have non-Gausian copula structure?
-- Do they show tail dependence distinct from that of a Gaussian copula?
+- Do datasets in ecology and related fields have non-Normal copula structure?
+- Do they show tail dependence distinct from that of a Normal copula?
 - In particular do they show asymmetric tail dependence?
@@
 and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}).
-The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
 
 The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`.
-These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data.
 Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.
 
-Although these results do convincingly show non-Gaussian copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+Although these results do convincingly show non-Normal copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
 The next section looks at tail dependence in a non-parametric way.
@@
 and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_BMRfit}).
-The table shows that BB6 and BB8 are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+The table shows that BB6 and BB8 are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
 
@@
 and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_contrasts_67fit}).
-The table shows that several copulas (mostly J and SC) are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.
+The table shows that several copulas (mostly J and SC) are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
 
 The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`.
-These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data.
+These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data.
 Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.
 
-These results do convincingly show non-Gaussian copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+These results do convincingly show non-Normal copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
 The next section looks at tail dependence in a non-parametric way.
@@
 <!--Question 2 analyses-->
-## Some possible causes of non-Gaussian copula structure
+## Some possible causes of non-Normal copula structure
 
@@
 <!--Question 3 analyses-->
-## Consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?
+## Consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
 
@@
 
-In order to study the effects of non-gaussian copula structure on extinction risk for population of a species at different locations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
+In order to study the effects of non-Normal copula structure on extinction risk for population of a species at different locations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
 <!--did I write correct spelling for lewton model?-->

### 2018-06-27 SupportingInformation.Rmd
replaced Gaussian by Normal and some text addition

@@
 Recall that question 1 is: Do datasets in ecology and related fields have non-Gausian copula structure?
-Do they show tail dependence distinct from that of a Gaussian copula?
+Do they show tail dependence distinct from that of a Normal copula?
 In particular do they show asymmetric tail dependence?
@@
 # Question 3 analyses for each dataset and theoretical analyses
-Recall that question 3 is: What are the consequences of non-Gaussian copula structure and tail dependence for our understanding of biology and for applications?
+Recall that question 3 is: What are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
 ## Skewness of the spatial mean

### 2018-06-27 SupportingInformation.Rmd
changes in code

@@
 ```{r bivar_tab_summary_fn,echo=F}
-bivar_tab_summary_fn<-function(r,rs){
+bivar_tab_summary_fn<-function(r){
 
@@
                  "LT_model.avg","UT_model.avg","(LT-UT)_model.avg",
-                 "Rank_Pl_surrog","Rank_Pu_surrog","Rank_Pl-Pu_surrog")
+                 "Rank_Pl_K","Rank_Pu_K","Rank_Pl-Pu_K")
   m<-as.data.frame(m)
@@
   m$`(LT-UT)_model.avg`<-m$LT_model.avg-m$UT_model.avg
-  m$Rank_Pl_surrog<-NA
-  m$Rank_Pu_surrog<-NA
-  m$`Rank_Pl-Pu_surrog`<-NA
+  m$Rank_Pl_K<-r$Rank_Pl_K
+  m$Rank_Pu_K<-r$Rank_Pu_K
+  m$`Rank_Pl-Pu_K`<-r$Rank_PlmPu_K
   mt<-as.data.frame(t(m))

### 2018-06-28 SupportingInformation.Rmd
error fixed

@@
 ```{r bivar_tab_summary_fn,echo=F}
-bivar_tab_summary_fn<-function(r){
+bivar_tab_summary_fn<-function(r,rs){
 
@@
   m$`(LT-UT)_model.avg`<-m$LT_model.avg-m$UT_model.avg
-  m$Rank_Pl_K<-r$Rank_Pl_K
-  m$Rank_Pu_K<-r$Rank_Pu_K
-  m$`Rank_Pl-Pu_K`<-r$Rank_PlmPu_K
+  m$Rank_Pl_K<-rs$Rank_Pl_K
+  m$Rank_Pu_K<-rs$Rank_Pu_K
+  m$`Rank_Pl-Pu_K`<-rs$Rank_PlmPu_K
   mt<-as.data.frame(t(m))
@@
 #calling the above function on 3 multivariable datasets
-m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke)
-m2<-bivar_tab_summary_fn(r=BivMS_BMR)
-m3<-bivar_tab_summary_fn(r=BivMS_contrasts_67)
+m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke,rs=stat_soilCN_ga)
+m2<-bivar_tab_summary_fn(r=BivMS_BMR,rs=stat_BMR_ga)
+m3<-bivar_tab_summary_fn(r=BivMS_contrasts_67,rs=stat_contrasts_67_ga)
 tab_bivar_summary<-cbind(m1,m2,m3)

### 2018-06-29 SupportingInformation.Rmd
keeping backup for all dataset results from maintext to suppmat

@@
 <!--
+#Results section detailed description
+### soil C and N data \label{Ex_SoilCN}
+
+```{r read_BivMS_soilCN_data,echo=F}
+BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
+```
+
+#### Model selection approach \label{Ex_SoilCN_Model_selection}
+
+The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
+from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}).
+The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
+
+The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`.
+These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.
+
+Although these results do convincingly show non-Normal copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+The next section looks at tail dependence in a non-parametric way.
+
+#### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}
+
+As described in \nameref{Methods}, each of our statistics, as computed for the 
+ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
+of values of the same statistic with the same ranges computed on 1000 normal-copula 
+surrogate datasets with similar Kendall or Spearman correlations, in different 
+runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}).
+
+Results indicate that tail dependence of 
+data is stronger in both the lower and upper tails than would be expected under a 
+null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
+weaker in the middle of the distributions
+than would be expected based on a normal-copula null model.
+
+We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{SM-fig_raw_cop_soilCN}(B). 
+Results (Table \ref{SM-tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.
+
+### BMR data \label{Ex_BMR}
+
+```{r read_BivMS_BMR_data,echo=F}
+BivMS_BMR<-readRDS("./Results/BMR_results/BivMS_lnBMR_vs_lnmass.RDS")
+```
+
+#### Model selection approach \label{Ex_BMR_Model_selection}
+
+The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_BMRfit}).
+The table shows that BB6 and BB8 are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
+
+The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was `r BivMS_BMR$relLTdep_AICw - BivMS_BMR$relUTdep_AICw`, indicating more upper tail dependence.
+We also tested that lowest-AIC copula BB8 was a good-enough fit, giving $p$-value `r BivMS_BMR$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_BMR$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+The next section looks at tail dependence in a non-parametric way.
+
+#### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
+
+As described in \nameref{Methods}, each of our statistics, as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig.
+\ref{SM-fig_BMR_nonparam}).
+
+Results indicate that dependence of data is different in some bins than would be expected under a 
+null hypothesis of a normal copula, answering the second part of question 1.
+Table \ref{SM-tab_BMR_asym} indicates stronger upper tail dependence for the BMR data copula which is significantly (95% CI) different from that of the symmetric normal copula and thus consistent with model selection results.
+
+### Anolis data \label{Ex_contrasts}
+
+```{r read_BivMS_contrasts_67_data,echo=F}
+BivMS_contrasts_67<-readRDS("./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
+```
+
+#### Model selection approach \label{Ex_contrasts_Model_selection}
+
+The test of indepedence gave $p=0$  (to within the precision available 
+from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_contrasts_67fit}).
+The table shows that several copulas (mostly J and SC) are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
+
+The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`.
+These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data.
+Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.
+
+These results do convincingly show non-Normal copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+The next section looks at tail dependence in a non-parametric way.
+
+#### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}
+
+As described in \nameref{Methods}, each of our statistics, are computed for $(l_b,u_b)$ in the 
+ranges $0-0.25$, $0.25-0.5$, $0.5-0.75$, $0.75-1$ (with $4$ number of bins).
+Anolis data set has lesser number of data compared to other two bivariate data set, so fewer number of bins were chosen to measure tail-dependence.
+Those statistics were compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different
+runs for Kendall and Spearman (Fig. \ref{SM-fig_contrasts_67_nonparam}).
+
+Results indicate that tail dependence of 
+data is stronger in both the lower and upper tails than would be expected under a 
+null hypothesis of a normal copula, answering the second part of question 1.
+
+We also tested statistically for asymmetry of tail dependence, which seems visible in Fig.
+\ref{SM-fig_raw_cop_AVG_humerus_ltoe}(B).
+Results (Table \ref{SM-tab_contrasts_67_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.
+
+### Abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
+
+```{r read_RES_aphid_count, echo=F}
+RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
+```
+
+#### Model selection approach \label{Ex_count_Model_selection}
+
+As described in Methods, model selection among copulas was done separately for each pair of the 10 
+sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns 
+corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location 
+pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of 
+question 1 for these data. Comparing the 
+AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) 
+to the AIC weight 
+of the normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for normal copulas was usually not
+comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
+failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
+the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}),
+demonstrating that the collection of copula families we used was sufficiently broad that the copulas of 
+all pairs of locations could be characterized by at least one of them.
+In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+
+The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively.
+These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data.
+Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.
+
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1.
+
+#### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
+
+Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
+computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
+resampling scheme, as described in Methods (Table \ref{SM-tab_count_resamp}). Confidence intervals of
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
+indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
+exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
+structure do not have, providing further evidence answering the third part of question 1.
+
+Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig.
+\ref{SM-fig_aphid_count} A,B) showed the typical decline of correlation with
+distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting 
+$\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig.
+\ref{SM-fig_aphid_count} C-F).
+Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows
+these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively).
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig.
+\ref{SM-fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical
+result of Table \ref{SM-tab_count_resamp} that averages of these quantities across all pairs of sampling locations
+were significantly greater than zero, demonstrating asymmetry of tail dependence.
+Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) package.
+
+### First flight date data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}
+
+```{r read_RES_aphid_ff,echo=F}
+RES_aphid_ff<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_firstflight_species_11.RDS")
+```
+
+#### Model selection approach \label{Ex_ff_Model_selection}
+
+As in section \nameref{Ex_count_Model_selection}, results are presented in $10 \times 10$ matrix format with rows
+and columns corresponding to locations.
+Table \ref{SM-tab_ff_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data.
+Comparing the AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_ff_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_ff_AICw_normal}) confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_ff_p_CvM}) usually failed to reject the
+hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for 
+goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_ff_p_KS}).
+In all cases at least `r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for
+goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+
+The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and 
+the model-averaged (again, AIC based) upper-tail dependence statistic are in 
+Tables \ref{SM-tab_ff_LTdep_AICw} and \ref{SM-tab_ff_UTdep_AICw}, respectively.
+These values were exclusively positive, thereby differing from what a normal copula would give (i.e., $0$), and answering the second part of question 1 for these data - these data have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
+
+The difference between the model-averaged lower-tail dependence statistic and the 
+model-averaged upper-tail dependence statistic (lower minus upper), shown in 
+Table \ref{SM-tab_ff_LTmUT}, was negative for `r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)` 
+out of `r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair` location pairs, answering the third part of question 1.
+We note this result is opposite the analogous result for aphid abundance
+data for the green spruce aphid, i.e., whereas those abundance data showed stronger lower tail dependendence than upper tail dependence, first flights for the leaf-curling plum aphid showed stronger upper tail dependence than lower tail dependence.
+
+#### Non-parametric analysis of tail dependence  \label{Ex_ff_NPA}
+
+Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman 
+correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were 
+computed via a spatial resampling scheme, as described in \nameref{Methods} (Table \ref{SM-tab_ff_resamp}).
+Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded 
+zero and mean values were positive, indicating tail dependence.
+Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always exluded $0$ and were negative, indicating asymmetry of tail dependence, a feature
+that data with normal copula structure do not have, providing further evidence answering the third part of question 1.
+The contrast here with the analogous results for abundances of the green spruce aphid is again illuminating.
+
+Plotting Spearman or Kendall correlations between first flight time series for pairs of locations as a function of great circle distance between the locations (Fig.
+\ref{SM-fig_aphid_ff} A, B) showed the typical (though slight, in this case) decline of correlation with increasing distance.
+Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear in most cases to decline slightly with distance (Fig.
+\ref{SM-fig_aphid_ff} C-F).
+$\Dsq_u$ and $\Dsq_l$ appear to slightly increase with distance, as expected given their
+opposite meaning (Fig. \ref{SM-fig_aphid_ff} G-H). 
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
+\ref{SM-fig_aphid_ff} I-K) did not show a decline with distance, but did demonstrate visually the statistical result of Table \ref{SM-tab_ff_resamp} that averages of these quantities across all pairs of sampling locations were significantly less than zero, showing asymmetry of tail dependence.
+
+### Abundance data for *Ceratium furca* in UK seas \label{Ex_pns}
+
+```{r read_RES_plankton_north_sea,echo=F}
+RES_plankton_north_sea<-readRDS("./Results/fitting_results/Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS")
+```
+
+#### Model selection approach \label{Ex_pns_Model_selection}
+
+As described in Methods and as in previous sections, model selection among copulas was done separately for 
+each pair of sampling locations.
+There were $14$ sampling locations, so results are presented in $14 \times 14$ matrix format, with rows and columns corresponding to locations.
+Table \ref{SM-tab_pns_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data.
+Comparing the AIC weight of the lowest-AIC copula family for each pair of sampling locations (Table \ref{SM-tab_pns_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_pns_AICw_normal})
+confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_pns_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_pns_p_KS}).
+In all cases at least
+`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+
+The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_pns_LTdep_AICw} and \ref{SM-tab_pns_UTdep_AICw}, respectively.
+These values were exclusively positive, thereby
+differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
+have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
+
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
+(lower minus upper), shown in Table \ref{SM-tab_pns_LTmUT}, was positive for some
+location pairs and negative for others, but was positive for 
+`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`
+locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
+
+#### Non-parametric analysis of tail dependence  \label{Ex_pns_NPA}
+
+Mean values of our statistics, as well as mean Kendall and Spearman
+correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
+spatial resampling scheme, as described in Methods (Table \ref{SM-tab_pns_resamp}).
+Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
+always excluded $0$ and mean values were positive, indicating tail dependence.
+Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
+also all were strictly positive, indicating asymmetry of tail dependence, a feature that data with normal copula
+structure do not have, providing further evidence answering the third part of question 1.
+
+Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
+between the locations (Fig. \ref{SM-fig_pns} A, B) showed the typical decline of correlation with increasing
+distance.
+Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig.
+\ref{SM-fig_pns} C-F).
+$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given their opposite meaning (Fig.
+\ref{SM-fig_pns} G-H).
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
+\ref{SM-fig_pns} I-K) did not show a decline with distance, but did demonstrate visually the
+statistical result of Table \ref{SM-tab_pns_resamp} that averages of these 
+quantities across all pairs of sampling locations were significantly greater 
+than zero, showing asymmetry of tail dependence.
+
+### Methane flux data \label{Ex_methane}
+
+```{r read_RES_methane,echo=F}
+RES_methane<-readRDS("./Results/fitting_results/MethaneFluxFit_selecloc_data_pt_thrs_50.RDS")
+```
+
+#### Model selection approach \label{Ex_methane_Model_selection}
+
+As described in Methods and as in previous sections, model selection among copulas was done separately for 
+each pair of sampling locations. There were $13$ sampling locations, so results are presented in 
+$13 \times 13$ matrix format, with rows and columns corresponding to locations. 
+Table \ref{SM-tab_methane_bestcop} shows that, for the large majority of location pairs, some 
+non-normal copula was a better fit, according to AIC, than was the normal copula, answering the 
+first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula 
+family for each pair of sampling locations (Table \ref{SM-tab_methane_AICw}) to the AIC weight of
+the normal copula family for the same pair (Table \ref{SM-tab_methane_AICw_normal}) 
+confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_methane_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_methane_p_KS}).
+In all cases at least
+`r min((RES_methane$gfc_numBS_success/RES_methane$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+
+The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_methane_LTdep_AICw} and \ref{SM-tab_methane_UTdep_AICw}, respectively.
+These values were exclusively positive, thereby
+differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
+have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
+
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
+(lower minus upper), shown in Table \ref{SM-tab_methane_LTmUT}, was positive for some
+location pairs and negative for others, but was positive for 
+`r sum((RES_methane$LTdep_AICw-RES_methane$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_methane$gfc_p_CvM)[1]^2-dim(RES_methane$gfc_p_CvM)[1])-2*RES_methane$num_indep_loc_pair`
+locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
+
+#### Non-parametric analysis of tail dependence  \label{Ex_methane_NPA}
+
+Mean values of our statistics, as well as mean Kendall and Spearman
+correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
+spatial resampling scheme, as described in Methods (Table \ref{SM-tab_methane_resamp}).
+Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
+always excluded $0$ and mean values were positive, indicating tail dependence.
+But the confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ were marginal, indicating no significant asymmetry of tail dependence.
+
+Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
+between the locations (Fig. \ref{SM-fig_methane} A, B) showed the typical decline of correlation with increasing
+distance.
+Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig.
+\ref{SM-fig_methane} C-F).
+$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
+their opposite meaning (Fig.
+\ref{SM-fig_methane} G-H).
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
+\ref{SM-fig_methane} I-K) did not show a decline with distance, but did demonstrate visually the
+statistical result of Table \ref{SM-tab_methane_resamp} that averages of these 
+quantities across all pairs of sampling locations were more or less equally dispersed around the zero-line, showing no significant asymmetry of tail dependence.
+-->
+
+<!--
 # Question 2 analyses 

### 2018-06-30 Paper.Rmd
adding biv-soilCN results_Q1

@@
 
+To answer Q1, we choose one representative datasets for each category : with a bivariate (soil C and N data) and
+a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.
+
 ### soil C and N data \label{Ex_SoilCN}
@@
 ```{r read_BivMS_soilCN_data,echo=F}
+d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
+d<-d[,c("SOCstock100","TSNstock100")] #raw soilCN data
 BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
 ```
+
+As described in the \ref{SM-SoilCN_data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central
+pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig. \ref{fig_biv_raw_cop}(A) and (D),respectively.
+We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.
 
@@
 The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
-from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
+from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`).
+We then fitted several copulas (all $16$ families as mentioned in \nameref{Model_selection})
 and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}). 
+Based on AIC weights, `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the best fit.
 The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
@@
 Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.
+To be precise, this difference being negative, indicated stronger upper-tail dependence than lower.
 
@@
 
-As described in \nameref{Methods}, each of our statistics, as computed for the 
+As described in \nameref{NPA}, each of our statistics, as computed for the 
 ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
-of values of the same statistic with the same ranges computed on 1000 normal-copula 
+of values of the same statistic with the same ranges computed on 1000 Normal-copula 
 surrogate datasets with similar Kendall or Spearman correlations, in different 
-runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). 
-<!--
-Because the correlation 
-used in the $\cor$ statistic is the Spearman correlation, Spearman-preserving surrogates
-should be used, so the top-right panel of Fig. \ref{SM-fig_soilCN_nonparam}
-is more reliable than the top left.
---> 
-Results indicate that tail dependence of 
-data is stronger in both the lower and upper tails than would be expected under a 
+runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). Each bin in each subplot of Fig. \ref{SM-fig_soilCN_nonparam}
+contains a number which indicates how many surrogates out of 1000, are significantly different ($95\%$ confidence interval) from Normal-copula statistics.
+
+Results indicate that tail dependence of data is stronger in both the lower and upper tails than would be expected under a 
 null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
@@
 
-### BMR data \label{Ex_BMR}
+We summarized our results columnwise in Table \ref{tab_bivar_summary} for all bivariate datasets plotted in Fig. \ref{fig_biv_raw_cop}.
 
-```{r read_BivMS_BMR_data,echo=F}
-BivMS_BMR<-readRDS("./Results/BMR_results/BivMS_lnBMR_vs_lnmass.RDS")
-```
+The first $9$ rows of Table \ref{tab_bivar_summary} summarized model selection results [namely, (1) p-value from `BiCopIndTest`, (2) best copula based on minimum AIC, (3) AIC-weight of that best copula, (4) AIC-weight of Normal copula, (5) p-value from `BiCopGofTest` for Cramer-von Mises method, (6) p-value from `BiCopGofTest` for Kolmogorov-Smirnov method, (7) AIC-weights model averaged lower tail dependence, (8) AIC-weights model averaged upper tail dependence, (9)  AIC-weights model averaged (lower-upper) tail dependence].
+Last $3$ rows ($10^{th}$ to $12^{th}$) of Table \ref{tab_bivar_summary} summarized non-parametric statistics results for $P$ statistics calculated on two extreme bins ($\Ps_l$ for first bin, $\Ps_u$ for last bin) for each copula data.
+We computed $\Ps_l$, $\Ps_u$, $\Ps_l - \Ps_u$ for the actual copula data and for Kendall-correlation 
+preserving $1000$ Normal surrogates and calculated the rank for actual data as the position if all these $1001$ values were 
+arranged in ascending order. So, whenever those ranks are $>975$ or $<25$, that indicate a significantly ($95\%$ CI) different 
+tail-dependence measures based on Normal-null hypothesis.
 
-#### Model selection approach \label{Ex_BMR_Model_selection}
-
-The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_BMRfit}).
-The table shows that BB6 and BB8 are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
-
-The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was `r BivMS_BMR$relLTdep_AICw - BivMS_BMR$relUTdep_AICw`, indicating more upper tail dependence.
-We also tested that lowest-AIC copula BB8 was a good-enough fit, giving $p$-value `r BivMS_BMR$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_BMR$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
-The next section looks at tail dependence in a non-parametric way.
-
-#### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
-
-As described in \nameref{Methods}, each of our statistics, as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig.
-\ref{SM-fig_BMR_nonparam}).
-
-Results indicate that dependence of data is different in some bins than would be expected under a 
-null hypothesis of a normal copula, answering the second part of question 1.
-Table \ref{SM-tab_BMR_asym} indicates stronger upper tail dependence for the BMR data copula which is significantly (95% CI) different from that of the symmetric normal copula and thus consistent with model selection results.
-
-### Anolis data \label{Ex_contrasts}
-
-```{r read_BivMS_contrasts_67_data,echo=F}
-BivMS_contrasts_67<-readRDS("./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
-```
-
-#### Model selection approach \label{Ex_contrasts_Model_selection}
-
-The test of indepedence gave $p=0$ <!--`r BivMS_contrasts_67$IndepTestRes` this is e-15 could not compile latex--> (to within the precision available 
-from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_contrasts_67fit}).
-The table shows that several copulas (mostly J and SC) are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
-
-The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`.
-These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.
-
-These results do convincingly show non-Normal copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
-The next section looks at tail dependence in a non-parametric way.
-
-#### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}
-
-As described in \nameref{Methods}, each of our statistics, are computed for $(l_b,u_b)$ in the 
-ranges $0-0.25$, $0.25-0.5$, $0.5-0.75$, $0.75-1$ (with $4$ number of bins).
-Anolis data set has lesser number of data compared to other two bivariate data set, so fewer number of bins were chosen to measure tail-dependence.
-Those statistics were compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different
-runs for Kendall and Spearman (Fig. \ref{SM-fig_contrasts_67_nonparam}).
-
-Results indicate that tail dependence of 
-data is stronger in both the lower and upper tails than would be expected under a 
-null hypothesis of a normal copula, answering the second part of question 1.
-
-We also tested statistically for asymmetry of tail dependence, which seems visible in Fig.
-\ref{SM-fig_raw_cop_AVG_humerus_ltoe}(B).
-Results (Table \ref{SM-tab_contrasts_67_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.
-
-### Abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
-
-```{r read_RES_aphid_count, echo=F}
-RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
-```
-
-#### Model selection approach \label{Ex_count_Model_selection}
-
-As described in Methods, model selection among copulas was done separately for each pair of the 10 
-sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns 
-corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location 
-pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of 
-question 1 for these data. Comparing the 
-AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) 
-to the AIC weight 
-of the normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for normal copulas was usually not
-comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
-failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
-the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}),
-demonstrating that the collection of copula families we used was sufficiently broad that the copulas of 
-all pairs of locations could be characterized by at least one of them.
-In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-
-The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively.
-These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data.
-Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.
-
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1.
-
-#### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
-
-Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
-computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
-resampling scheme, as described in Methods (Table \ref{SM-tab_count_resamp}). Confidence intervals of
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
-indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
-exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
-structure do not have, providing further evidence answering the third part of question 1.
-
-Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig.
-\ref{SM-fig_aphid_count} A,B) showed the typical decline of correlation with
-distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting 
-$\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig.
-\ref{SM-fig_aphid_count} C-F).
-Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows
-these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig.
-\ref{SM-fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical
-result of Table \ref{SM-tab_count_resamp} that averages of these quantities across all pairs of sampling locations
-were significantly greater than zero, demonstrating asymmetry of tail dependence.
-Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) package.
-
-### First flight date data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}
-
-```{r read_RES_aphid_ff,echo=F}
-RES_aphid_ff<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_firstflight_species_11.RDS")
-```
-
-#### Model selection approach \label{Ex_ff_Model_selection}
-
-As in section \nameref{Ex_count_Model_selection}, results are presented in $10 \times 10$ matrix format with rows
-and columns corresponding to locations.
-Table \ref{SM-tab_ff_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data.
-Comparing the AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_ff_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_ff_AICw_normal}) confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_ff_p_CvM}) usually failed to reject the
-hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for 
-goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_ff_p_KS}).
-In all cases at least `r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for
-goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-
-The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and 
-the model-averaged (again, AIC based) upper-tail dependence statistic are in 
-Tables \ref{SM-tab_ff_LTdep_AICw} and \ref{SM-tab_ff_UTdep_AICw}, respectively.
-These values were exclusively positive, thereby differing from what a normal copula would give (i.e., $0$), and answering the second part of question 1 for these data - these data have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
-
-The difference between the model-averaged lower-tail dependence statistic and the 
-model-averaged upper-tail dependence statistic (lower minus upper), shown in 
-Table \ref{SM-tab_ff_LTmUT}, was negative for `r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)` 
-out of `r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair` location pairs, answering the third part of question 1.
-We note this result is opposite the analogous result for aphid abundance
-data for the green spruce aphid, i.e., whereas those abundance data showed stronger lower tail dependendence than upper tail dependence, first flights for the leaf-curling plum aphid showed stronger upper tail dependence than lower tail dependence.
-
-#### Non-parametric analysis of tail dependence  \label{Ex_ff_NPA}
-
-Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman 
-correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were 
-computed via a spatial resampling scheme, as described in \nameref{Methods} (Table \ref{SM-tab_ff_resamp}).
-Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded 
-zero and mean values were positive, indicating tail dependence.
-Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always exluded $0$ and were negative, indicating asymmetry of tail dependence, a feature
-that data with normal copula structure do not have, providing further evidence answering the third part of question 1.
-The contrast here with the analogous results for abundances of the green spruce aphid is again illuminating.
-
-Plotting Spearman or Kendall correlations between first flight time series for pairs of locations as a function of great circle distance between the locations (Fig.
-\ref{SM-fig_aphid_ff} A, B) showed the typical (though slight, in this case) decline of correlation with increasing distance.
-Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear in most cases to decline slightly with distance (Fig.
-\ref{SM-fig_aphid_ff} C-F).
-$\Dsq_u$ and $\Dsq_l$ appear to slightly increase with distance, as expected given their
-opposite meaning (Fig. \ref{SM-fig_aphid_ff} G-H). 
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
-\ref{SM-fig_aphid_ff} I-K) did not show a decline with distance, but did demonstrate visually the statistical result of Table \ref{SM-tab_ff_resamp} that averages of these quantities across all pairs of sampling locations were significantly less than zero, showing asymmetry of tail dependence.
-
-### Abundance data for *Ceratium furca* in UK seas \label{Ex_pns}
-
-```{r read_RES_plankton_north_sea,echo=F}
-RES_plankton_north_sea<-readRDS("./Results/fitting_results/Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS")
-```
-
-#### Model selection approach \label{Ex_pns_Model_selection}
-
-As described in Methods and as in previous sections, model selection among copulas was done separately for 
-each pair of sampling locations.
-There were $14$ sampling locations, so results are presented in $14 \times 14$ matrix format, with rows and columns corresponding to locations.
-Table \ref{SM-tab_pns_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data.
-Comparing the AIC weight of the lowest-AIC copula family for each pair of sampling locations (Table \ref{SM-tab_pns_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_pns_AICw_normal})
-confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_pns_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_pns_p_KS}).
-In all cases at least
-`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-
-The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_pns_LTdep_AICw} and \ref{SM-tab_pns_UTdep_AICw}, respectively.
-These values were exclusively positive, thereby
-differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
-have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
-
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
-(lower minus upper), shown in Table \ref{SM-tab_pns_LTmUT}, was positive for some
-location pairs and negative for others, but was positive for 
-`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`
-locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
-
-#### Non-parametric analysis of tail dependence  \label{Ex_pns_NPA}
-
-Mean values of our statistics, as well as mean Kendall and Spearman
-correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
-spatial resampling scheme, as described in Methods (Table \ref{SM-tab_pns_resamp}).
-Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
-always excluded $0$ and mean values were positive, indicating tail dependence.
-Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
-also all were strictly positive, indicating asymmetry of tail dependence, a feature that data with normal copula
-structure do not have, providing further evidence answering the third part of question 1.
-
-Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
-between the locations (Fig. \ref{SM-fig_pns} A, B) showed the typical decline of correlation with increasing
-distance.
-Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig.
-\ref{SM-fig_pns} C-F).
-$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given their opposite meaning (Fig.
-\ref{SM-fig_pns} G-H).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
-\ref{SM-fig_pns} I-K) did not show a decline with distance, but did demonstrate visually the
-statistical result of Table \ref{SM-tab_pns_resamp} that averages of these 
-quantities across all pairs of sampling locations were significantly greater 
-than zero, showing asymmetry of tail dependence.
-
-### Methane flux data \label{Ex_methane}
-
-```{r read_RES_methane,echo=F}
-RES_methane<-readRDS("./Results/fitting_results/MethaneFluxFit_selecloc_data_pt_thrs_50.RDS")
-```
-
-#### Model selection approach \label{Ex_methane_Model_selection}
-
-As described in Methods and as in previous sections, model selection among copulas was done separately for 
-each pair of sampling locations. There were $13$ sampling locations, so results are presented in 
-$13 \times 13$ matrix format, with rows and columns corresponding to locations. 
-Table \ref{SM-tab_methane_bestcop} shows that, for the large majority of location pairs, some 
-non-normal copula was a better fit, according to AIC, than was the normal copula, answering the 
-first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula 
-family for each pair of sampling locations (Table \ref{SM-tab_methane_AICw}) to the AIC weight of
-the normal copula family for the same pair (Table \ref{SM-tab_methane_AICw_normal}) 
-confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_methane_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_methane_p_KS}).
-In all cases at least
-`r min((RES_methane$gfc_numBS_success/RES_methane$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-
-The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_methane_LTdep_AICw} and \ref{SM-tab_methane_UTdep_AICw}, respectively.
-These values were exclusively positive, thereby
-differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
-have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
-
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
-(lower minus upper), shown in Table \ref{SM-tab_methane_LTmUT}, was positive for some
-location pairs and negative for others, but was positive for 
-`r sum((RES_methane$LTdep_AICw-RES_methane$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_methane$gfc_p_CvM)[1]^2-dim(RES_methane$gfc_p_CvM)[1])-2*RES_methane$num_indep_loc_pair`
-locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
-
-#### Non-parametric analysis of tail dependence  \label{Ex_methane_NPA}
-
-Mean values of our statistics, as well as mean Kendall and Spearman
-correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
-spatial resampling scheme, as described in Methods (Table \ref{SM-tab_methane_resamp}).
-Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
-always excluded $0$ and mean values were positive, indicating tail dependence.
-But the confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ were marginal, indicating no significant asymmetry of tail dependence.
-
-Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
-between the locations (Fig. \ref{SM-fig_methane} A, B) showed the typical decline of correlation with increasing
-distance.
-Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig.
-\ref{SM-fig_methane} C-F).
-$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
-their opposite meaning (Fig.
-\ref{SM-fig_methane} G-H).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
-\ref{SM-fig_methane} I-K) did not show a decline with distance, but did demonstrate visually the
-statistical result of Table \ref{SM-tab_methane_resamp} that averages of these 
-quantities across all pairs of sampling locations were more or less equally dispersed around the zero-line, showing no significant asymmetry of tail dependence.
+For all $3$ bivariate-datasets indicated by each column name (SoilCN, BMR, Anolis) 
+in Table \ref{tab_bivar_summary}, Normal copula were always a poor fit compared to 
+highest AIC-weight based asymmetric copulas - which answers the first part of Q1. 
+Each dataset presented there, has significant upper tail dependence 
+compared to Normal as indicated by negative AIC-weights model averaged 
+(lower-upper) tail dependence and very low ranks of actual copula 
+data for $\Ps_l - \Ps_u$ statistics, which answers the second and third part of Q1
+(For detailed results of each dataset see Tables \ref{SM-tab_soilCN_asym} - \ref{SM-tab_contrasts_67_asym} and 
+Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_contrasts_67_nonparam}).
 
@@
 
+### Abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
+
+```{r read_RES_aphid_count, echo=F}
+RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
+```
+
+#### Model selection approach \label{Ex_count_Model_selection}
+
+As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid abundance data, model selection among copulas was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations.
+Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question 1 for these data.
+Comparing the
+AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulas was usually not
+comparable to the degree of support for other families.
+Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
+failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
+the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulas of all pairs of locations could be characterized by at least one of them.
+In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+
+The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively.
+These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data.
+Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.
+
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1.
+Fig.
+\ref{fig_aphid_raw_cop}(A) and (C) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.
+
+#### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
+
+Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
+computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
+resampling scheme, as described in \nameref{NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
+indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
+exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question 1.
+
+Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig.
+\ref{SM-fig_aphid_count} A,B) showed the typical decline of correlation with distance between sampling locations that has been seen for numerous other datasets in earlier work.
+Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig.
+\ref{SM-fig_aphid_count} C-F).
+Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively).
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig.
+\ref{SM-fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical result of Table \ref{SM-tab_count_resamp} that averages of these quantities across all pairs of sampling locations were significantly greater than zero, demonstrating asymmetry of tail dependence, as also seen from model selection approach.
+Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
+
 ```{r tab_multivar_summary, echo=F, results='markup'}
@@
 ```
+
+We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets 
+(i.e., abundance data for green spruce aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
+firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
+abundance data for plankton from UK seas from $14$ sampling locations, 
+methane flux data from $13$ sampling locations). We counted for how many location pair (excluding self-comparison)
+copula are non-independent.
 

### 2018-06-30 SupportingInformation.Rmd
adding biv-soilCN results_Q1

@@
 As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
-When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
+When red cross was above (or below) the blue and green cross, text at the top of plots indicates the actual statistics was greater than (or less than) by those number of surrogate values.
 When $\cor$ values (respectively, $\Ps$ or $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, stronger than or weaker than for $\Ps$ or $\Dsq$) expected from a normal-copula null hypothesis.
@@
       caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.
-Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (\\nameref{MT-Methods}).
+Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\nameref{MT-NPA} in the manuscript).
 One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates.
@@
   m<-matrix(NA,1,17)
-  colnames(m)<-c("#pairs","#non-indep","#non-N",
-                 "Avg.best.AICw","Avg.N.AICw","#good.fit",
+  colnames(m)<-c("num_pairs","num_non-indep","num_non-N",
+                 "Avg.best.AICw","Avg.N.AICw","num_good.fit",
                  "CI0.025_LT","CI0.975_LT","CI0.025_UT","CI0.975_UT",
-                 "#LT-UT>0",
+                 "num_LT-UT>0",
                  "CI0.025_Pl","CI0.975_Pl","CI0.025_Pu","CI0.975_Pu","CI0.025_Pl-Pu","CI0.975_Pl-Pu")
@@
   d<-dim(r$gfc_numBS)[1]
-  m$`#pairs`<-(d^2)-d
-  m$`#non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
-  m$`#non-N`<-sum(r$info_ord_copcode[,,1]!=1,na.rm = T)
+  m$num_pairs<-(d^2)-d 
+  m$`num_non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
+  m$`num_non-N`<-sum(r$info_ord_copcode[,,1]!=1,na.rm = T)
   r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN
@@
   temp<-which((r$gfc_p_CvM>0.01) & (r$gfc_p_KS>0.01),arr.ind=T)
-  m$`#good.fit`<-dim(temp)[1]
+  m$num_good.fit<-dim(temp)[1]
   t<-r$LTdep_AICw[is.finite(r$LTdep_AICw)]
@@
   m$CI0.975_UT<-round(CI_UT[2],3)
-  m$`#LT-UT>0`<-sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)
+  m$`num_LT-UT>0`<-sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)
   m$CI0.025_Pl<-round(rs$numericdf[5,3],3)

### 2018-06-30 Paper.Rmd
aphid count and multivar table description added

@@
 Comparing the
-AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulas was usually not
+AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the Normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulas was usually not
 comparable to the degree of support for other families.
@@
 abundance data for plankton from UK seas from $14$ sampling locations, 
-methane flux data from $13$ sampling locations). We counted for how many location pair (excluding self-comparison)
-copula are non-independent.
+methane flux data from $13$ sampling locations). First $11$ rows in Table \ref{tab_multivar_summary} represent
+model selection results :
+(1) total number of location pairs (excluding self-comparison) considered, 
+(2) how many location pair copula are non-independent among them, 
+(3) number of non-Normal copula as best fit (based on highest AIC-weight),
+(4) average AIC-weights of best fit copula from all location pair,
+(5) average AIC-weights of Normal copula from all location pair,
+(6) number of location pair for which best fit copula have p-values > 0.01 
+(both Cramer-von Mises and Kolmogorov-Smirnov method)
+(7) lower and (8) upper quantile ($95\%$CI) of model averaged (AIC-weight based) lower tail dependence from all location pair,
+(9) lower and (10) upper quantile ($95\%$CI) of model averaged (AIC-weight based) upper tail dependence from all location pair,
+(11) number of location pair which showed stronger lower tail dependence than upper.
+
+Last $6$ rows of Table \ref{tab_multivar_summary} summarized the non-parametric analysis results 
+for $\Ps$ statistics : lower and upper quantile ($95\%$CI) of $\Ps_l$, $\Ps_u$, $\Ps_l-\Ps_u$ based on resampling
+location pair.
+
+Likewise, green-spruce aphid abundance data ($1^{st}$ column in Table \ref{tab_multivar_summary}), 
+plankton data from UK seas also showed major lower tail dependent spatial synchrony 
+($3^{rd}$ column in Table \ref{tab_multivar_summary}), whereas firstflight data for leaf-curling plum aphid 
+showed stronger upper tail dependence at every location pair ($2^{nd}$ column in Table \ref{tab_multivar_summary}).
+Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.
 

### 2018-07-11 SupportingInformation.Rmd
adding code for ceder creek data

@@
 \end{figure}
+
+## Ceder-creek data for e120 \label{ceder-creek_data}
+
+```{r read_ceder_creek_data, echo=F, results="hide"}
+set.seed(seed)
+#-------------------------------------------------
+# This function takes 
+# Input :
+#       db : a dataframe : smaller subset of ple_120 data to calculate avg. biomass for each plot
+#       dp : a dataframe : smaller subset of pce_120 data to calculate avg. shannon's H for those plots
+#       year : the year for which you want to calculate biomass and shannon's index data
+#       rawplot : logical : T for rawdata plot, F for copula plot
+#       ploton : logical for optional plotting
+
+# Output :
+#       a list of two data-frame:
+#                   tab_dbp : a data-frame with 3 column : (plot no , avg.H, avg.Biomass(g/m2))
+#                   cop_tab_dbp : a data-frame with 3 column : (plot no, pobs(avg.H), pobs(avg.Biomass(g/m2)))
+
+cop_HB<-function(db,dp,year,rawplot,ploton){
+  # db : biomass dataset for a specific year
+  # dp : percent cover dataset for a specific year
+
+  plots<-sort(intersect(db$Plot,dp$Plot)) # common plots for both datasets for a specific year
+  db<-subset(db,Plot%in%plots)
+  dp<-subset(dp,Plot%in%plots)
+
+  tab_dbp<-data.frame(Plot=plots,avg.H=NA,avg.Biomass=NA)
+  for(i in 1:length(plots)){
+    # calculate biomass for each plot (averaged over all strips for that plot)
+    x<-subset(db,Plot==plots[i])
+    tab_dbp$avg.Biomass[i]<-mean(x$Biomass)
+
+    # calculate shannon's index H for each plot (averaged over all Quadrates for that plot)
+    x<-subset(dp,Plot==plots[i])
+    uniq_sp<-sort(as.character(unique(x$Species)))  # These are the species in a particular plot
+    Q<-as.character(unique(x$Quadrat))
+
+    tab_each_sp<-data.frame(sp=uniq_sp,avg_cov=0,rel_avg_cov=0)
+
+    for(sp in 1:length(uniq_sp)){
+      my_sp<-uniq_sp[sp]
+      tab_temp<-data.frame(Q=Q,abs_cov=0)
+      ind<-which(x$Species==my_sp)
+      li<-length(ind)
+      tab_temp$abs_cov[1:li]<-x$Abscover[ind]
+      tab_each_sp$avg_cov[[sp]]<-mean(tab_temp$abs_cov) # avg. over 4Quadrates for a sp.
+    }
+
+    tab_each_sp$rel_avg_cov<-tab_each_sp$avg_cov/sum(tab_each_sp$avg_cov)
+    tab_dbp$avg.H[i]<- -sum((tab_each_sp$rel_avg_cov)*log(tab_each_sp$rel_avg_cov))
+
+  }
+
+  tab_dbp<-na.omit(tab_dbp)
+
+  cop_tab_dbp<-data.frame(Plot=tab_dbp$Plot,avg.H=NA,avg.Biomass=NA)
+  cop_tab_dbp$avg.H<-VineCopula::pobs(tab_dbp$avg.H)
+  cop_tab_dbp$avg.Biomass<-VineCopula::pobs(tab_dbp$avg.Biomass)
+
+  if(ploton==T){
+    if(rawplot==T){
+      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col=rgb(0,0,1,0.3),pch=19,cex.lab=1.5)
+      mtext(paste0("Year=",year))
+    }else{
+      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col=rgb(1,0,0,0.3),pch=19,cex.lab=1.5)
+      mtext(paste0("Year=",year))
+    }
+  }
+
+  return(list(tab_dbp=tab_dbp,
+         cop_tab_dbp=cop_tab_dbp))
+}
+#-----------------------------
+r_ple<-read.delim("./Data/Ceder_creek_e120/edited_ple120.txt")
+r_pce<-read.delim("./Data/Ceder_creek_e120/edited_pce120.txt")
+
+y1<-unique(r_ple$Year)
+y2<-unique(r_pce$Year)
+yr<-sort(intersect(y1,y2))
+
+a<-as.data.frame(table(sort(r_pce$Species)))
+b<-as.character(a$Var1)
+bad_sp<-c(b[51:53],b[156:158])
+
+ind<-which(r_pce$Species%in%bad_sp)
+r_pce_clean<-r_pce[-ind,]
+#--------------------------------------------------------------------------------------------
+pdf("./Results/Ceder_creek_results/raw_HB_all_yr.pdf",height = 6,width = 10)
+op<-par(mfrow=c(2,3),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
+for(i in 1:length(yr)){
+  db<-subset(r_ple,Year==yr[i])
+  dp<-subset(r_pce_clean,Year==yr[i])
+
+  sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
+}
+par(op)
+dev.off()
+#----------------------------------------------------------------------------------------
+cop_HB_all_yr<-vector("list",length(yr))
+names(cop_HB_all_yr)<-yr
+
+pdf("./Results/Ceder_creek_results/cop_HB_all_yr.pdf",height = 6,width = 10)
+op<-par(mfrow=c(2,3),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
+for(i in 1:length(yr)){
+  db<-subset(r_ple,Year==yr[i])
+  dp<-subset(r_pce_clean,Year==yr[i])
+
+  s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
+  cop_HB_all_yr[[i]]<-s$cop_tab_dbp[,2:3]
+}
+par(op)
+dev.off()
+
+saveRDS(cop_HB_all_yr,"./Results/Ceder_creek_results/cop_HB_all_yr.RDS")
+#---------------------------------------------------------------------------------------
+cop_HB_all_yr<-readRDS("./Results/Ceder_creek_results/cop_HB_all_yr.RDS")
+```
 
@@
 -->
+
+```{r bivmodselect_cc, echo=F, cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("OurBiCopSelect.R")
+set.seed(seed)
+
+BivMS_res_cc<-vector("list",length(cop_HB_all_yr))
+names(BivMS_res_cc)<-names(cop_HB_all_yr)
+
+for(i in 1:length(BivMS_res_cc)){
+  BivMS_res_cc[[i]]<-OurBiCopSelect(u1=cop_HB_all_yr[[i]]$avg.H,u2=cop_HB_all_yr[[i]]$avg.Biomass,
+                                    families=c(1,3:10,13:14,16:20),
+                                    level=0.05,AICBIC="AIC",
+                                    numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
+}
+saveRDS(BivMS_res_cc,file="./Results/Ceder_creek_results/BivMS_res_cc.RDS")
+```
+
+```{r stat_cc_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+source("bivfunctionplot.R")
+set.seed(seed)
+
+stat_cc_ga<-vector("list",length(cop_HB_all_yr))
+names(stat_cc_ga)<-names(cop_HB_all_yr)
+
+for(i in 1:length(stat_cc_ga)){
+  temp<-bivfunctionplot(v=cop_HB_all_yr[[i]],
+                             resloc="./Results/Ceder_creek_results/",
+                             nametag=paste0(names(cop_HB_all_yr)[i],"cc_numbin_10",sep=""),
+                             numbin = 10)
+  stat_cc_ga[[i]]<-temp
+}
+
+saveRDS(stat_cc_ga,"./Results/Ceder_creek_results/stat_cc_ga.RDS")
+```
 

### 2018-07-11 SupportingInformation.Rmd
NA error fixing

@@
 r_ple<-read.delim("./Data/Ceder_creek_e120/edited_ple120.txt")
+names(r_ple)[36]<-"Biomass"
+
 r_pce<-read.delim("./Data/Ceder_creek_e120/edited_pce120.txt")
@@
 # Now do NPA for v_BMR
-source("bivfunctionplot.R")
 stat_BMR_ga<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_10",numbin = 10)
@@
 saveRDS(BivMS_res_cc,file="./Results/Ceder_creek_results/BivMS_res_cc.RDS")
+BivMS_res_cc<-readRDS("./Results/Ceder_creek_results/BivMS_res_cc.RDS")
 ```

### 2018-07-11 CopulaFunctions_flexible.R

@@ -56,9 +56,12 @@ Corbds<-function(vi,vj,lb,ub)
   #compute the indices of the points between the bounds
   inds<-which(vi+vj>2*lb & vi+vj<2*ub)
   
+  if(length(inds)!=0){ 
   #get the portion of the Spearman
   res<-sum((vi[inds]-vi_mean)*(vj[inds]-vj_mean))/((length(vi)-1)*sqrt(var_vi*var_vj))
-
+  }else{
+   res<-NA
+  }
   return(res)  
 }
 
@@ -198,32 +201,45 @@ Pbds<-function(vi,vj,lb,ub){
   #---------------------------------------------------------
   
   inds<-which(vi+vj>2*lb & vi+vj<2*ub)
-  dist_sort<-sort(abs(vi[inds]-vj[inds])/sqrt(2))
   
-  dpt_c<-0
-  dpt_uniq<-c(0)
-  dist_sort_df<-as.data.frame(table(dist_sort))
-  for (i in 1:length(dist_sort_df$Freq)){
-    dpt_c<-dpt_c+(dist_sort_df$Freq[i]/length(dist_sort))
-    dpt_uniq<-c(dpt_uniq,dpt_c)
+  if(length(inds)!=0){
+    
+    dist_sort<-sort(abs(vi[inds]-vj[inds])/sqrt(2))
+    
+    dpt_c<-0
+    dpt_uniq<-c(0)
+    dist_sort_df<-as.data.frame(table(dist_sort))
+    for (i in 1:length(dist_sort_df$Freq)){
+      dpt_c<-dpt_c+(dist_sort_df$Freq[i]/length(dist_sort))
+      dpt_uniq<-c(dpt_uniq,dpt_c)
+    }
+    dpt_uniq<-c(dpt_uniq,1)
+    dist_dpt_uniq<-c(0,as.numeric(as.character(dist_sort_df$dist_sort)),d_max)
+    
+    nrep<-2
+    S<-rep(dpt_uniq,each=nrep)
+    dist_S<-rep(dist_dpt_uniq[2:length(dist_dpt_uniq)],each=nrep)
+    dist_S<-c(0,dist_S,d_max)
+    #plot(dist_S,S,type="l",col="red")
+    
+    #integrals under the functions S
+    Au_S<-0
+    
+    for(i in 1:(length(dpt_uniq)-1)){
[…]

### 2018-07-11 bivfunctionplot.R

@@ -42,7 +42,7 @@ fracwork<-function(dvals,surrvals){
   frac<-NA*numeric(length(dvals))
   names(frac)<-names(dvals)
   for (counter in 1:length(frac)){
-    frac[counter]<-sum(surrvals[counter,]<dvals[counter])
+    frac[counter]<-sum(surrvals[counter,]<dvals[counter],na.rm=T) # to handle NA's
   }
   return(frac)
 }
@@ -107,15 +107,15 @@ bivfunctionplot<-function(v,resloc,nametag,numbin){
   
   corlmcoru_d<-unname(corstats_d[1]-corstats_d[length(corstats_d)])
   corlmcoru_K<-corstats_K[1,]-corstats_K[length(corstats_d),]
-  corlmcoru_frac_K<-sum(corlmcoru_K<corlmcoru_d)
-  corstats_Kq<-apply(FUN=quantile,X=corstats_K,MARGIN=1,prob=c(.005,0.025,.975,.995))
+  corlmcoru_frac_K<-sum(corlmcoru_K<corlmcoru_d,na.rm=T)
+  corstats_Kq<-apply(FUN=quantile,X=corstats_K,MARGIN=1,prob=c(.005,0.025,.975,.995),na.rm=T)
   
   corstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Corbds,nm="Cor",numbin=numbin)
   corstats_frac_S<-fracwork(corstats_d,corstats_S)
   
   corlmcoru_S<-corstats_S[1,]-corstats_S[length(corstats_d),]
-  corlmcoru_frac_S<-sum(corlmcoru_S<corlmcoru_d)
-  corstats_Sq<-apply(FUN=quantile,X=corstats_S,MARGIN=1,prob=c(.005,0.025,.975,.995))
+  corlmcoru_frac_S<-sum(corlmcoru_S<corlmcoru_d,na.rm=T)
+  corstats_Sq<-apply(FUN=quantile,X=corstats_S,MARGIN=1,prob=c(.005,0.025,.975,.995),na.rm=T)
 
   #P stats
   Pstats_d<-calcstats(v=v,f=Pbds_wrap,nm="P",numbin=numbin)
[…]

### 2018-07-12 SupportingInformation.Rmd
edit make_table fn

@@
 
-fractxt<-function(fracval,numsurrog=1000){
+fractxt<-function(fracval,numsurrog){
   ptxt<-''
@@
 
- atdres[1,2]<-fractxt(data$corlmcoru_frac_K)
- atdres[1,3]<-fractxt(data$corlmcoru_frac_S)
- atdres[2,2]<-fractxt(data$PlmPu_frac_K)
- atdres[2,3]<-fractxt(data$PlmPu_frac_S)
- atdres[3,2]<-fractxt(data$D2umD2l_frac_K)
- atdres[3,3]<-fractxt(data$D2umD2l_frac_S)
+ atdres[1,2]<-fractxt(data$corlmcoru_frac_K,numsurrog = data$numsurrog_success_K)
+ atdres[1,3]<-fractxt(data$corlmcoru_frac_S,numsurrog = data$numsurrog_success_S)
+ atdres[2,2]<-fractxt(data$PlmPu_frac_K,numsurrog = data$numsurrog_success_K)
+ atdres[2,3]<-fractxt(data$PlmPu_frac_S,numsurrog = data$numsurrog_success_S)
+ atdres[3,2]<-fractxt(data$D2umD2l_frac_K,numsurrog = data$numsurrog_success_K)
+ atdres[3,3]<-fractxt(data$D2umD2l_frac_S,numsurrog = data$numsurrog_success_S)
 

### 2018-07-16 Paper.Rmd
ceder-creek description fig and description added

@@
 \begin{figure}[!h]
-\textbf{ \hspace{4.3 cm} (A) \hspace{4.3 cm} (B) \hspace{4 cm} (C) } \\
-\vspace{-0.4 cm}
+\textbf{ \hspace{3.2 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
+\vspace{-0.35 cm}
 \begin{center}
-\includegraphics[width=5 cm]{./Results/SoilCNraw.pdf}
-\includegraphics[width=5 cm]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
-\includegraphics[width=5 cm]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}\\
+\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
+\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
+\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
 \end{center}
-\textbf{ \hspace{4.3 cm} (D) \hspace{4.3 cm} (E) \hspace{4 cm} (F) } \\
-\vspace{-0.4 cm}
+\textbf{ \hspace{3.2 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
+\vspace{-0.35 cm}
 \begin{center}
-\includegraphics[width=5 cm]{./Results/SoilCNcop.pdf}
-\includegraphics[width=5 cm]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
-\includegraphics[width=5 cm]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
+\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
+\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
+\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
 \caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} ln(BMR) of contrasts vs.
@@
 contrasts for AVG.humerus phylogenetic traits.
-Lower panel \textbf{(D - F)} shows copula plots for the corresponding datasets of the upper panel.
-These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of contrasts for last two datasets (see \nameref{Model_selection} of manuscript).
+\textbf{(D)} above-ground plant biomass data vs.
+Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
+Lower panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel.
+These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of contrasts for next two datasets and number of plots for the last one (see \nameref{Model_selection} of manuscript).
 \label{fig_biv_raw_cop}}
@@
 
-For all $3$ bivariate-datasets indicated by each column name (SoilCN, BMR, Anolis) 
+For all $4$ bivariate-datasets indicated by each column name (SoilCN, BMR, Anolis, Ceder-creek) 
 in Table \ref{tab_bivar_summary}, Normal copula were always a poor fit compared to 
 highest AIC-weight based asymmetric copulas - which answers the first part of Q1. 
-Each dataset presented there, has significant upper tail dependence 
-compared to Normal as indicated by negative AIC-weights model averaged 
+Among four datasets presented there, first three have significant upper tail dependence whereas the last one has significant lower tail dependence compared to Normal.
+It was indicated by negative AIC-weights model averaged
 (lower-upper) tail dependence and very low ranks of actual copula 
-data for $\Ps_l - \Ps_u$ statistics, which answers the second and third part of Q1
-(For detailed results of each dataset see Tables \ref{SM-tab_soilCN_asym} - \ref{SM-tab_contrasts_67_asym} and 
-Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_contrasts_67_nonparam}).
+data for $\Ps_l - \Ps_u$ statistics for first three datasets and positive AIC-weights model averaged 
+(lower-upper) tail dependence and very high ranks of actual copula 
+data for $\Ps_l - \Ps_u$ statistics for the last one, which answers the second and third part of Q1
+(For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and 
+Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}).
 
@@
 
-tab_biivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
+tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
 knitr::kable(tab_bivar_summary, 

### 2018-07-16 SupportingInformation.Rmd
ceder-creek description fig and description added

@@
 
+These biodiversity experiment data sets, for [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and for [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120), are collected by Tilman et al.
+over many years.
+We used thsese datasets [@Tilman2001; @Tilman2006] to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m),
+for year 2000.
+
 ```{r read_ceder_creek_data, echo=F, results="hide"}
@@
     if(rawplot==T){
-      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col=rgb(0,0,1,0.3),pch=19,cex.lab=1.5)
-      mtext(paste0("Year=",year))
+      op<-par(mar=c(5.1, 4.1, 1.5, 2.1))
+      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col="red",pch=1,cex.lab=1.6,cex.axis=1.5)
+      mtext(paste0("Year=",year),cex=1.5)
+      par(op)
     }else{
-      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col=rgb(1,0,0,0.3),pch=19,cex.lab=1.5)
-      mtext(paste0("Year=",year))
+      op<-par(mar=c(5.1, 4.1, 1.5, 2.1))
+      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",col="red",pch=1,cex.lab=2,cex.axis=1.5,
+         xlim=c(0,1),ylim=c(0,1),asp=1)
+      rect(0,0,1,1)
+      lines(c(0,1),c(0,1),type='l')
+      mtext(paste0("Year=",year),cex=1.5)
+      par(op)
     }
@@
 dev.off()
+#-----------------------
+#raw data plot for year2000
+pdf("./Results/Ceder_creek_results/raw_HB_2000.pdf",height = 6,width = 6)
+i<-5
+db<-subset(r_ple,Year==yr[i])
+dp<-subset(r_pce_clean,Year==yr[i])
+sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
+dev.off()
 #----------------------------------------------------------------------------------------
@@
 dev.off()
+#-------------------------
+#cop data plot for year2000
+pdf("./Results/Ceder_creek_results/cop_HB_2000.pdf",height = 6,width = 6)
+i<-5
+db<-subset(r_ple,Year==yr[i])
+dp<-subset(r_pce_clean,Year==yr[i])
+s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
+cc_cop_2000<-s$cop_tab_dbp[,2:3]
+dev.off()
+#------------------
 
@@
 ```
+
+\begin{figure}[!h]
+\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
+\vspace{-0.2 cm}
+\begin{center}
+\includegraphics[width=3 in]{./Results/Ceder_creek_results/raw_HB_2000.pdf}
+\includegraphics[width=3 in]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
+\caption[\textbf{(A)} Plot of the plant-aboveground avg.
+Biomass vs.
+avg.
+Shannon's index (H).
+\textbf{(B)} Copula plot of plant-aboveground avg.
+Biomass vs.
+avg.
+Shannon's index.]{\textbf{(A)} Plot of the plant-aboveground avg.
+Biomass vs.
+avg.
+Shannon's index (H).
+\textbf{(B)} Copula plot of plant-aboveground avg.
+Biomass vs.
+avg.
+Shannon's index.
+Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(cc_cop_2000)`$ is the number of (9m by 9m)plots.
+(see \nameref{MT-Model_selection} of manuscript).
+\label{fig_raw_cop_cc2000}}
+\end{center}
+\end{figure}
 
@@
 
+<!--Display model selection results for Ceder-creek data for year 2000 in a table-->
+```{r tab_cc2000fit, echo=F, results='markup'}
+h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+knitr::kable(h,
+             format='pandoc',
+             caption = "Fitting info for **Ceder-creek dataset** for year 2000. \\label{tab_cc2000fit}",
+             booktabs=T)
+```
+
 ```{r stat_cc_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
@@
 saveRDS(stat_cc_ga,"./Results/Ceder_creek_results/stat_cc_ga.RDS")
+```
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_10_bivfunctionplot.pdf}
+\caption[Testing non-parametric statistics for Ceder-creek data with $numbin=10$ for year 2000]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{Ceder-creek data}.
+As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
+When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
+\label{fig_cc2000_nonparam}}
+\end{center}
+\end{figure}
+
+```{r tab_cc_asym_2000, echo=F, results="markup"}
+tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
+knitr::kable(tab_cc_asym_2000,
+             format="pandoc",
+             caption = "**Ceder-creek data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis on year 2000 \\label{tab_cc_asym_2000}",
+             booktabs=T)
 ```
@@
 m3<-bivar_tab_summary_fn(r=BivMS_contrasts_67,rs=stat_contrasts_67_ga)
-tab_bivar_summary<-cbind(m1,m2,m3)
-colnames(tab_bivar_summary)<-c("SoilCN","BMR","Anolis")
+m4<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
+tab_bivar_summary<-cbind(m1,m2,m3,m4)
+colnames(tab_bivar_summary)<-c("SoilCN","BMR","Anolis","Ceder-creek")
 saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")

### 2018-07-24 SupportingInformation.Rmd
Taylors law plot modified

@@
 #***constants
-N<-100 #length of time series
+N<-50 #length of time series
 n<-25 #number of sampling locations
-numdat<-25 #the number of datasets to get from the normal copula
+numdat<-10 #the number of datasets to get from the normal copula
 rho<-0.7 #normal-copula parameter for dependence between locations
@@
 diag(sig)<-1
-m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
+m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig) # each column of m has normal histogram
 m<-array(m,c(N,numdat,n))
-m<-aperm(m,c(1,3,2))
+m<-aperm(m,c(1,3,2)) # 2nd dim is normally distributed e.g. hist(m[,1,])
 gshape<-7.5
 gscale<-1 #gamma parameters
-cop_to_marg_trans<-function(x)
-{
+
+cop_to_marg_trans<-function(x){
   return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
@@
 }
-marg_to_cop_trans<-function(x) #transforms to copula space
-{
+
+marg_to_cop_trans<-function(x){ #transforms to copula space
   return(pgamma(x,shape=gshape,scale=gscale))
 }
+
+# pnorm(m) makes normal to uniform
+# then qgamma(...) makes uniform to gamma distributed
 m<-cop_to_marg_trans(pnorm(m))
@@
 #***surrogates using a Clayton copula, kendall preserving
+
 #get the Clayton parameter to use
@@
 csurrog<-array(NA,dim(m))
-for (counter in 1:dim(m)[3])
-{
+for (counter in 1:dim(m)[3]){
   csurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=ccop,numsurrog=1)
@@
 fsurrog<-array(NA,dim(m))
-for (counter in 1:dim(m)[3])
-{
+for (counter in 1:dim(m)[3]){
   fsurrog[,,counter]<-copsurrognd(m[,,counter],targetcop=fcop,numsurrog=1)
@@
 scsurrog<-cop_to_marg_trans((1-marg_to_cop_trans(csurrog)))
+
+# check : plot(csurrog[,1,1],csurrog[,2,1]) and plot(scsurrog[,1,1],scsurrog[,2,1])
 
@@
 spea.points.res<-list()
-for (counter in 1:length(kend.surrog))
-{
+for (counter in 1:length(kend.surrog)){
  ks<-kend.surrog[[counter]]
  ss<-spea.surrog[[counter]]
- kend.stats.res[[counter]]<-matrix(NA,5,numdat)
+ kend.stats.res[[counter]]<-matrix(NA,5,numdat) # to keep spatial TL stats
  spea.stats.res[[counter]]<-matrix(NA,5,numdat)
- kend.points.res[[counter]]<-array(NA,c(N,2,numdat))
+ kend.points.res[[counter]]<-array(NA,c(N,2,numdat)) # to keep log(mu), log(var) of spatial TL as a dataframe
  spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
- for (sc in 1:numdat)
- {
+ for (sc in 1:numdat){
    khold<-tl_stats(ks[,,sc])
@@
 <!--Now make some plots, these one for kendall surrogates-->
-```{r taylor_plots, echo=F}
-##plot TL for clayton
+```{r taylor_plots, echo=F, results="hide"}
 xlimits<-range(kend.points.res[[1]][,1,],kend.points.res[[2]][,1,],kend.points.res[[3]][,1,],kend.points.res[[4]][,1,])
 ylimits<-range(kend.points.res[[1]][,2,],kend.points.res[[2]][,2,],kend.points.res[[3]][,2,],kend.points.res[[4]][,2,])
-par(mfrow=c(1,1),mar=c(4,4,1,1) + 0.1)
+
+pdf("./Results/taylorslaw_results/SpatialTL_CNFSC.pdf",height=4,width=16)
+op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
+##plot TL for clayton
 plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
-     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
+     cex.axis=1.5,cex.lab=1.5)  
 points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
+mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black")
 
@@
 plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
-     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
+     cex.axis=1.5,cex.lab=1.5)  
 points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
+mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black")
 
@@
 plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
-     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
+     cex.axis=1.5,cex.lab=1.5)  
 points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
+mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black")
 
@@
 plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
-     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")))  
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
+     cex.axis=1.5,cex.lab=1.5)  
 points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
-
+mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black")
+
+par(op)
+dev.off()
+#---------------------------------------------------
 #prepare for plots coming below
@@
 
-#plot s.lin distr for each set of surrogates
+pdf("./Results/taylorslaw_results/SpatialTL_5stats.pdf",height=3,width=15)
+op<-par(mfrow=c(1,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
 #***DAN: x axis needs to be redone to show the copula names used in the surrogates
-par(mfrow=c(1,1),mar=c(4,4,1,1) + 0.1)
+
 x<-1:4
-plot(x,s.lin[1,],type="b",ylim=range(s.lin),xlab="Set of surrogates",ylab="Linearity test p-vals")
-for (counter in 2:dim(s.lin)[1])
-{
-  lines(x,s.lin[counter,],type='b')
+x2<-c("C", "N", "F", "SC")
+plot(x,s.lin[1,],type="l",col="grey",ylim=range(s.lin),xlab="Set of surrogates",ylab="Linearity test p-vals",
+     cex.axis=1.5,cex.lab=1.5,xaxt="n")
+points(x,s.lin[1,],col="red")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+for (counter in 2:dim(s.lin)[1]){
+  lines(x,s.lin[counter,],col="grey")
+  points(x,s.lin[counter,],col="red")
 }
@@
 x<-1:4
-plot(x,s.hom[1,],type="b",ylim=range(s.hom),xlab="Set of surrogates",ylab="Homoskedasticity test p-vals")
-for (counter in 2:dim(s.slope)[1])
-{
-  lines(x,s.hom[counter,],type='b')
+plot(x,s.hom[1,],type="l",col="grey",ylim=range(s.hom),xlab="Set of surrogates",
+     ylab="Homoskedasticity test p-vals",
+     cex.axis=1.5,cex.lab=1.5,xaxt="n")
+points(x,s.hom[1,],col="red")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+for (counter in 2:dim(s.slope)[1]){
+  lines(x,s.hom[counter,],col="grey")
+  points(x,s.hom[counter,],col="red")
 }
@@
 #plot s.rmse distr for each set of surrogates
-#***DAN: x axis needs to be redone to show the copula names used in the surrogates
 x<-1:4
-plot(x,s.rmse[1,],type="b",ylim=range(s.rmse),xlab="Set of surrogates",ylab="Spatial TL rmse")
-for (counter in 2:dim(s.rmse)[1])
-{
-  lines(x,s.rmse[counter,],type='b')
+plot(x,s.rmse[1,],type="l",col="grey",ylim=range(s.rmse),xlab="Set of surrogates",ylab="Spatial TL rmse",
+     cex.axis=1.5,cex.lab=1.5,xaxt="n")
+points(x,s.rmse[1,],col="red")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+for (counter in 2:dim(s.rmse)[1]){
+  lines(x,s.rmse[counter,],col="grey")
+  points(x,s.rmse[counter,],col="red")
 }
@@
 x<-1:4
-plot(x,s.int[1,],type="b",ylim=range(s.int),xlab="Set of surrogates",ylab="Spatial TL intercept")
-for (counter in 2:dim(s.int)[1])
-{
-  lines(x,s.int[counter,],type='b')
+plot(x,s.int[1,],type="l",col="grey",ylim=range(s.int),xlab="Set of surrogates",
+     ylab="Spatial TL intercept",
+     cex.axis=1.5,cex.lab=1.5,xaxt="n")
+points(x,s.int[counter,],col="red")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+for (counter in 2:dim(s.int)[1]){
+  lines(x,s.int[counter,],col="grey")
+  points(x,s.int[counter,],col="red")
 }
@@
 x<-1:4
-plot(x,s.slope[1,],type="b",ylim=range(s.slope),xlab="Set of surrogates",ylab="Spatial TL slope")
-for (counter in 2:dim(s.slope)[1])
-{
-  lines(x,s.slope[counter,],type='b')
-}
-```
+plot(x,s.slope[1,],type="l",col="grey",ylim=range(s.slope),xlab="Set of surrogates",
+     ylab="Spatial TL slope",
+     cex.axis=1.5,cex.lab=1.5,xaxt="n")
+points(x,s.slope[counter,],col="red")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+for (counter in 2:dim(s.slope)[1]){
+  lines(x,s.slope[counter,],col="grey")
+  points(x,s.slope[counter,],col="red")
+}
+par(op)
+dev.off()
+```
+
+\begin{figure}[!h]
+\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.1 cm} (D)} \\
+\vspace{-0.5 cm} 
+\begin{center}
+\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC.pdf}\\
+\end{center}
+\textbf{ \hspace{3.1 cm} (E) \hspace{2.4 cm} (F) \hspace{2.3 cm} (G) \hspace{2.3 cm} (H) \hspace{2.3 cm} (I)}\\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_5stats.pdf}
+\caption[Spatial Taylor's Law results]{Spatial Taylor's Law results (A) hjghj ; (B) fgjfgjfg. \label{fig_spTL}}
+\end{center}
+\end{figure}
 

### 2018-07-25 Paper.Rmd
adding summary for taylors law section

@@
 
+### Effects on Taylor's law
+
+Spatial Taylor's law, which shows linear relationship between `log`(variance) and `log`(mean) of population densities of a species taken from different locations, is a well-known pattern found in ecology and related fields.
+Here, we want to examine the nature of the slope and how it changes if population time series (taken as $50$ years) from different locations (say, $numloc=25$) have some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula.
+First, we simulated (N $\times$ n $\times$ ns) dimensional normal copula with covariance $\rho =0.7$, where $(N=50,n=25,ns=10)$ are the length of timeseries, number of sampling locations and number of replica for that copula, respectively.
+Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal.
+With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D).
+We also tested statistically that whether linearity, homoskedasticity is maintained in each copula dataset or not and plotted our results alongwith root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).
+
+\begin{figure}[!h]
+\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.1 cm} (D)} \\
+\vspace{-0.5 cm} 
+\begin{center}
+\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC.pdf}\\
+\end{center}
+\textbf{ \hspace{3.1 cm} (E) \hspace{2.4 cm} (F) \hspace{2.3 cm} (G) \hspace{2.3 cm} (H) \hspace{2.3 cm} (I)}\\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_5stats.pdf}
+\caption[Spatial Taylor's Law results]{Spatial Taylor's Law results: log(spatial variance) vs.
+log(spatial mean) for $ns=10$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of hhomoskedasticity, (G) root mean square error, (H) intercept, (I) slope of the first panel plots.
+\label{fig_spTL}}
+\end{center}
+\end{figure}
+
 # Discussion

### 2018-07-25 SupportingInformation.Rmd
adding summary for taylors law section

@@
 
-<!-- Taylor's law section need to move to Paper.Rmd once it's completed-->
-## Taylor's law
+<!-- Taylor's law section moved to Paper.Rmd -->
 
@@
 
-\begin{figure}[!h]
-\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.1 cm} (D)} \\
-\vspace{-0.5 cm} 
-\begin{center}
-\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC.pdf}\\
-\end{center}
-\textbf{ \hspace{3.1 cm} (E) \hspace{2.4 cm} (F) \hspace{2.3 cm} (G) \hspace{2.3 cm} (H) \hspace{2.3 cm} (I)}\\
-\vspace{-0.5 cm}
-\begin{center}
-\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_5stats.pdf}
-\caption[Spatial Taylor's Law results]{Spatial Taylor's Law results (A) hjghj ; (B) fgjfgjfg. \label{fig_spTL}}
-\end{center}
-\end{figure}
-
 \newpage

### 2018-07-26 SupportingInformation.Rmd
adding pedagog figure

@@
 source("mtime.R") #A function needed for caching
+```
+
+```{r call_cop_pedag_fig,echo=F,results="hide",cache=T, cache.extra=list(seed,mtime("copula_pedag_fig.R"))}
+set.seed(seed)
+source("./copula_pedag_fig.R")
+
+#cat(paste("start-time: ",Sys.time(),"\n"))
+df<-cop_est_2par(fam=7,p1range = c(0.01:7),p2range = c(1:7),incr = 0.005)
+saveRDS(df,"./Results/df_fam7_incr_0.005.RDS")
+#cat(paste("stop-time: ",Sys.time(),"\n"))
+
+df_fam7<-readRDS("./Results/df_fam7_incr_0.005.RDS")
+#range(df_fam7$LTmUT)
+
+pdf("copula_pedagog_fig_BB1_incr_0.005.pdf",height=8,width=20)
+op<-par(mfrow=c(2,5),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
+
+df_fam7_scr<-screening_df(df=df_fam7,ncut=5,Tau0=0.4,eps=1e-3,
+                               LTmUT_vector_given = F,my_LTmUT_vector = NA)
+
+df_fam7_screened<-df_fam7_scr$df_screened
+df_fam7_LTmUT_vector<-df_fam7_scr$LTmUT_vector
+
+for(i in 1:length(df_fam7_screened$LTmUT)){
+  c<-BiCopSim(500,family=7,par=df_fam7_screened$par[i],par2=df_fam7_screened$par2[i])
+  plot(c[,1],c[,2],col="red",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
+  mtext(paste0("tau = ",round(df_fam7_screened$Tau[i],4),
+               ", LTmUT = ",round(df_fam7_screened$LTmUT[i],4),sep=""),side=3)
+}
+
+df_fam7_scr<-screening_df(df=df_fam7,ncut=5,Tau0=0.8,eps=1e-3,
+                               LTmUT_vector_given = T,my_LTmUT_vector = df_fam7_LTmUT_vector)
+df_fam7_screened<-df_fam7_scr$df_screened
+
+for(i in 1:length(df_fam7_screened$LTmUT)){
+  c<-BiCopSim(500,family=7,par=df_fam7_screened$par[i],par2=df_fam7_screened$par2[i])
+  plot(c[,1],c[,2],col="red",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
+  mtext(paste0("tau = ",round(df_fam7_screened$Tau[i],4),
+               ", LTmUT = ",round(df_fam7_screened$LTmUT[i],4),sep=""),side=3)
+}
+
+par(op)
+dev.off()
 ```

### 2018-07-26 copula_pedag_fig.R

@@ -1,0 +1,73 @@
+library(VineCopula)
+# This function takes 
+# Input to create a seed copula:
+#        fam : family of considered bivariate copula with two parameters
+#        p1range : range of par
+#        p2range : range of par2
+#        incr : increment for parameter range
+# and gives 
+# Output after BiCopEst of that specified copula:
+#       a data.frame which gives info like estimated par,par2,kendall's tau,taildep
+
+cop_est_2par<-function(fam,p1range,p2range,incr){
+  par_range<-as.vector(seq(from=p1range[1],to=tail(p1range,1),by=incr))
+  par2_range<-as.vector(seq(from=p2range[1],to=tail(p2range,1),by=incr))
+  dm<-length(par_range)*length(par2_range)
+  df<-data.frame(par=NA*numeric(dm),par2=NA*numeric(dm),
+                 Tau=NA*numeric(dm),
+                 LT=NA*numeric(dm),
+                 UT=NA*numeric(dm),
+                 LTmUT=NA*numeric(dm))
+  count<-0
+  for(par in par_range){
+    for(par2 in par2_range){
+      count<-count+1
+      #cat("par=",par,"par2=",par2,"\n")
+      cop<-BiCop(family=fam,par=par,par2=par2)
+      df$par[count]<-cop$par
+      df$par2[count]<-cop$par2
+      df$Tau[count]<-cop$tau
+      df$LT[count]<-cop$taildep$lower 
+      df$UT[count]<-cop$taildep$upper 
+      df$LTmUT[count]<-(cop$taildep$lower - cop$taildep$upper)
+    }
+  }
+  
+  return(df)
+}
+#-----------------------------------------------
+# This function takes input :
+#            df : A dataframe which is the output of cop_est_2par function
+#            ncut : an integer which indicates the number of label in LTmUT range
+#            Tau0 : target Tau with which you want to plot your copula
+#            eps : error limit in Tau0
+#            LTmUT_vector_given : logical (if F then it will generate LTmUT_vector automatically from df)
+#            my_LTmUT_vector : a vector for LTmUT is given if the above argument is T, otherwise it's NA
+
+# gives output is a list of two:
+#            A screened version of input dataframe with specific LTmUT dep value
[…]

### 2018-07-26 SupportingInformation.Rmd
modified

@@
 
-```{r call_cop_pedag_fig,echo=F,results="hide",cache=T, cache.extra=list(seed,mtime("copula_pedag_fig.R"))}
+```{r call_cop_pedag_fig,echo=F,cache=T, cache.extra=list(seed,mtime("copula_pedag_fig.R"))}
 set.seed(seed)
@@
 #cat(paste("stop-time: ",Sys.time(),"\n"))
-
+```
+
+```{r cop_pedag_fig,echo=F,results="hide"}
 df_fam7<-readRDS("./Results/df_fam7_incr_0.005.RDS")

### 2018-07-28 SupportingInformation.Rmd
pedagog fig for incr=0.05

--- 2018-07-26 SupportingInformation.Rmd
@@
 
-```{r call_cop_pedag_fig,echo=F,results="hide",cache=T, cache.extra=list(seed,mtime("copula_pedag_fig.R"))}
+```{r call_cop_pedag_fig,echo=F,cache=T, cache.extra=list(seed,mtime("copula_pedag_fig.R")),results="hide"}
 set.seed(seed)
@@
 
-#cat(paste("start-time: ",Sys.time(),"\n"))
 df<-cop_est_2par(fam=7,p1range = c(0.01:7),p2range = c(1:7),incr = 0.005)
-saveRDS(df,"./Results/df_fam7_incr_0.005.RDS")
-#cat(paste("stop-time: ",Sys.time(),"\n"))
+saveRDS(df,"./Results/df_fam7_incr_0.01.RDS")
 

### 2018-07-31 SupportingInformation.Rmd
adding optimized copula_pedag_fig

@@
 
-```{r call_cop_pedag_fig,echo=F,cache=T, cache.extra=list(seed,mtime("copula_pedag_fig.R")),results="hide"}
-set.seed(seed)
-source("./copula_pedag_fig.R")
-
-df<-cop_est_2par(fam=7,p1range = c(0.01:7),p2range = c(1:7),incr = 0.005)
-saveRDS(df,"./Results/df_fam7_incr_0.01.RDS")
-
-df_fam7<-readRDS("./Results/df_fam7_incr_0.005.RDS")
-#range(df_fam7$LTmUT)
-
-pdf("copula_pedagog_fig_BB1_incr_0.005.pdf",height=8,width=20)
+```{r cop_pedag_fig,echo=F,results="hide"}
+set.seed(seed)
+library(VineCopula)
+f<-function(par,destau,desLTmUT,fam){
+  cop<-BiCop(family=fam,par=par[1],par2=par[2])
+  tau<-cop$tau
+  LT<-cop$taildep$lower
+  UT<-cop$taildep$upper
+  LTmUT<-LT-UT
+  z<-(tau-destau)^2
+  z<-z+(LTmUT-desLTmUT)^2
+  return(z)
+}
+#----------------------------------------------------------
+pdf("./Results/copula_pedagog_fig_BB1.pdf",height=8,width=20)
 op<-par(mfrow=c(2,5),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
 
-df_fam7_scr<-screening_df(df=df_fam7,ncut=5,Tau0=0.4,eps=1e-3,
-                               LTmUT_vector_given = F,my_LTmUT_vector = NA)
-
-df_fam7_screened<-df_fam7_scr$df_screened
-df_fam7_LTmUT_vector<-df_fam7_scr$LTmUT_vector
-
-for(i in 1:length(df_fam7_screened$LTmUT)){
-  c<-BiCopSim(500,family=7,par=df_fam7_screened$par[i],par2=df_fam7_screened$par2[i])
+desLTmUT_vector<-c(-0.4,-0.2,0,0.2,0.4)
+#parlist<-list(c(0.2,1.5),c(0.2,1.1),c(0.2,1.5),c(0.5,1.1),c(0.5,1.5))
+#for(i in 1:length(desLTmUT_vector)){
+#  print(i)
+#  iLmU<-desLTmUT_vector[i]
+#  call_optim<-optim(par=parlist[[i]],fn=f,destau=0.4,desLTmUT=iLmU,fam=7,method = "BFGS")
+#  call_optim$par
+#  c<-BiCopSim(N=500,family=7,par=call_optim$par[1],par2=call_optim$par[2])
+#  plot(c[,1],c[,2],col="blue",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
+#  mtext(paste0("tau = 0.4", ", LTmUT = ",iLmU,sep=""),side=3)
+#}
+
+for(i in desLTmUT_vector){
+  # print(i)
+  destau<-0.5
+  call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=i,fam=7,method = "BFGS")
+  obj<-BiCop(family=7,par=call_optim$par[1],par2=call_optim$par[2])
+  c<-BiCopSim(N=500,obj)
+  plot(c[,1],c[,2],col="blue",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
+  mtext(paste0("tau = ",round(obj$tau,5),", LTmUT = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
+}
+
+for(i in desLTmUT_vector){
+ # print(i)
+  destau<-0.8
+ call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=i,fam=7,method = "BFGS")
+ obj<-BiCop(family=7,par=call_optim$par[1],par2=call_optim$par[2])
+ c<-BiCopSim(N=500,obj)
  plot(c[,1],c[,2],col="red",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
-  mtext(paste0("tau = ",round(df_fam7_screened$Tau[i],4),
-               ", LTmUT = ",round(df_fam7_screened$LTmUT[i],4),sep=""),side=3)
-}
-
-df_fam7_scr<-screening_df(df=df_fam7,ncut=5,Tau0=0.8,eps=1e-3,
-                               LTmUT_vector_given = T,my_LTmUT_vector = df_fam7_LTmUT_vector)
-df_fam7_screened<-df_fam7_scr$df_screened
-
-for(i in 1:length(df_fam7_screened$LTmUT)){
-  c<-BiCopSim(500,family=7,par=df_fam7_screened$par[i],par2=df_fam7_screened$par2[i])
-  plot(c[,1],c[,2],col="red",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
-  mtext(paste0("tau = ",round(df_fam7_screened$Tau[i],4),
-               ", LTmUT = ",round(df_fam7_screened$LTmUT[i],4),sep=""),side=3)
+ mtext(paste0("tau = ",round(obj$tau,5),", LTmUT = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
 }

### 2018-08-13 Paper.Rmd
adding intro basic

@@
 
-*Goes in the last paragraph of the intro:*
-We address the following questions.
-1) Do datasets in ecology and related fields have non-Gausian copula srtructure?
+Identifying and quantifying dependencies is the core of modelling in many applications, frequently found in the
+finance and risk management. For many years, measuring and modeling dependencies has focused on the
+Pearson correlation because of its ease in use.
+But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence because it is a measure of linear association between two random variables.
+Therefore, it is better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation.
+However, using copula-based approach as a dependence measure reflects more knowledge about the dependence structure instead of summarizing in just a single number.
+The flexibility of copulas allows us to model more complex dependence structures by isolating its marginal from its joint distribution structures, with the capability in modeling a rich variety of tail behavior, such as tail dependence versus tail independence, and reflection symmetry versus reflection asymmetry.
+Tail dependence occurs when the correlation between two variables increases as we get "further" in the tail (either or both) of the distribution.
+For example, in the upper panel of Fig.
+\ref{fig_cop_pedag}, different dependence structures, with same Kendall correlation $\tau = 0.5$, were shown which became dominantly lower tail dependent (lower tail - upper tail dependence being positive) from dominantly upper tail dependent (lower tail - upper tail dependence being negative) as we moved along the row.
+Lower panel represents the similar scenario but with stronger correlation (Kendall's $\tau = 0.8$).
+To implement this, we need to select candidate bivariate copulas from many existing parametric copula families.
+A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulas to be selected from.
+Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), mentioned in the \nameref{Methods}.
+
+Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016] or a short period study of intersite dependence of rainfall data [@Serinaldi2008], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data.
+So, in ecological perspective, when we find a dominantly lower (upper) tail dependence structure, it means both variable are rare synchronously.
+If any one of them be environmental variable and the other is an observable (like abundance of a species) then it can also explain the causal relationship between them.
+Here, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets.
+In addition to a copula based \nameref{Model_selection}, we also carried out a \nameref{NPA} with three different statistics to verify our model selection based results.
+
+Basically, we want to address the following questions.
+1) Do datasets in ecology and related fields have non-Normal copula srtructure?
 Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence?
@@
 And 3) what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
+In \nameref{Material&Methods}, we described briefly about each datasets we used to answer these questions and provided a detailed descriptions of implemented methods.
+In the \nameref{Results} section, we summarized our finidings for each datasets.
+Next two sections are focused on possible causes and consequences of our analysis.
+Finally, we concluded in the \nameref{Discussion} section.
 
-# Material and methods
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
+\caption{Pedagogical figure showing BB1 copula with varying tail-dep difference but of same Kendall's tau along each row and with same tail-dep difference but of different Kendall's tau along each column.\label{fig_cop_pedag}}
+\end{center}
+\end{figure}
+
+# Material and methods \label{Material&Methods}
 
 ## Datasets
+<!-- each dataset's description is in suppmat file-->
 
@@
 
-## Theoretical models
-
-Question 2 from the Introduction was explored using simple population models.
+<!--Theoretical models
+Question 2 from the Introduction was explored using simple population models.-->
 
@@
 
-# Results
+# Results \label{Results}
 
@@
 
-# Discussion
+# Discussion \label{Discussion}
 

### 2018-08-13 SupportingInformation.Rmd
adding intro basic

@@
   plot(c[,1],c[,2],col="blue",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
-  mtext(paste0("tau = ",round(obj$tau,5),", LTmUT = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
+  mtext(paste0("tau = ",round(obj$tau,5),", ( LT - UT ) = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
+  legend("topleft",paste0("UT=",round(obj$taildep$upper,5)),bty="n",cex=1.5,text.col="magenta")
+  legend("bottomright",paste0("LT=",round(obj$taildep$lower,5)),bty="n",cex=1.5,text.col="magenta")
 }
@@
  plot(c[,1],c[,2],col="red",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
- mtext(paste0("tau = ",round(obj$tau,5),", LTmUT = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
+ mtext(paste0("tau = ",round(obj$tau,5),", ( LT - UT ) = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
+ legend("topleft",paste0("UT=",round(obj$taildep$upper,5)),bty="n",cex=1.5,text.col="magenta")
+ legend("bottomright",paste0("LT=",round(obj$taildep$lower,5)),bty="n",cex=1.5,text.col="magenta")
 }

### 2018-08-13 Paper.Rmd
modified

@@
 Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016] or a short period study of intersite dependence of rainfall data [@Serinaldi2008], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data.
-So, in ecological perspective, when we find a dominantly lower (upper) tail dependence structure, it means both variable are rare synchronously.
+So, in ecological perspective, when we find a dominantly lower (upper) tail dependence structure, it means both incidence are rare (abundant) synchronously.
 If any one of them be environmental variable and the other is an observable (like abundance of a species) then it can also explain the causal relationship between them.

### 2018-08-31 Paper.Rmd
modified data section

@@
 
-## Datasets
-<!-- each dataset's description is in suppmat file-->
+## Data \label{Data}
+
+For bivariate analysis we used four datasets which are briefly described as follows;
+to begin with **Soil C and N data**, to explore the correlation of soil organic C and total soil N across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
+database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
+stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
+study.
+
+Next, the [**BMR data**](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) contains the trait contrasts from phylogenetically independent contrasts [@SukumaranH2010] of data reported by [@GenoudIM2017] on the basal metabolic rate (BMR) and mass of mammals species.
+Specifically, the "SELECT" BMR data in their database were used for those species for which data were available.
+A phylogeny in NEXUS [@NEXUS] format was extracted from the supplementary information of [@GenoudIM2017].
+Version 4.4.0 of the DendroPy library [@SukumaranH2010] was used to calculate phylogenetic independent contrasts using Python 2.7.
+For each contrast, the path length spanned by the contrast (sum of the branch lengths) and the contrast for the mass, natural logarithm of the BMR and natural logarithm of the mass were written as a row in the bmr-contrasts.csv file.
+
+[**Third dataset**](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) was a raw data on measurements of anatomical features of several species of *Anolis* lizards, which was downloaded
+from the Dryad data package [@MahlerIRL2013Data] associated with [@MahlerIRL2013].
+The [contrasts.csv file](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) was produced by using version 4.4.0 of the DendroPy library [@SukumaranH2010]  to calculate phylogenetic independent contrasts on the the data in the GA_Anolis_traits.csv file of Mahler et at.
+using their maximum clade credibility tree (GA_Anolis_MCC.nex) as the phylogeny.
+
+Absolute values of traits for each contrast are considered to construct corresponding copula for second and third datasets.
+
+Fourth, we used **Ceder-creek data for e120** which resulted from a biodiversity experiment, carried out by Tilman and his co-warkers [@Tilman2001; @Tilman2006] over many years to collect [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120) data.
+These datasets were used to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m),
+for year 2000.
+
+The raw data and copula for each four datasets are shown along each column of Figure \ref{fig_biv_raw_cop}.
+
+For multivariate analysis, we applied four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014].
+These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species.
+We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony.
+In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}).
+Figure \ref{fig_aphid_raw_cop} represents raw and copula data plot for aphid count ("Green spruce aphid") and aphid-firstflight date ("Leaf-curling plum aphid") taken from two locations.
+
+Our third dataset for multivariate analysis is a plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013).
+These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
+The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
+unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
+location in the ocean.
+Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979].
+All the samples available for a given variable in a given area in a given month of a given year were
+averaged, producing a monthly time series for each year.
+Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).
+Monthly values were converted to yearly values by averaging over all months.
+This approach was used by [@Sheppard2017] earlier.
+But here we only allowed maximum 4 months in a year with missing data which were replaced by those medians.
+If a year got 5 or more missing monthly data value then we set NA value for that particular year.
+We used data for species=16 [Name : "Ceratium furca"] which is a dicoflagellate commonly found in UK seas.
+Among 26 locations, only those having atleast 45 years time series data, were considered (see Figure \ref{SM-fig_plankton_map}).
+
+Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
+Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
+Additional information on the history of the site and soil types is detailed in @Jarecke2016.
+Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at
+daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
+additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.
+
+\begin{figure}[!h]
+\textbf{ \hspace{3.2 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
+\vspace{-0.35 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
+\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
+\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
+\end{center}
+\textbf{ \hspace{3.2 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
+\vspace{-0.35 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
+\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
+\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
+\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} ln(BMR) of contrasts vs.
+ln(mass) of contrasts, \textbf{(C)} contrasts for AVG.ltoe.IV vs.
+contrasts for AVG.humerus phylogenetic traits.
+\textbf{(D)} above-ground plant biomass data vs.
+Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
+Lower panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel.
+These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of contrasts for next two datasets and number of plots for the last one (see \nameref{Model_selection} of manuscript).
+\label{fig_biv_raw_cop}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
+\vspace{-0.4 cm}
+\begin{center}
+\includegraphics[width=7 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
+\includegraphics[width=7 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
+\end{center}
+\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
+\vspace{-0.4 cm}
+\begin{center}
+\includegraphics[width=7 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
+\includegraphics[width=7 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
+\caption{Upper panel shows raw data plots for \textbf{(A)} green spruce aphid-abundance data from location pair(4,11) and \textbf{(B)} leaf-curling plum aphid-firstflight data from location pair(2,5).
+Lower panel \textbf{(C - D)} shows copula plots for the corresponding datasets of the upper panel.
+These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of data-points available for both location at same year (see \nameref{Model_selection} of manuscript).
+\label{fig_aphid_raw_cop}}
+\end{center}
+\end{figure}
 
@@
 Question 2 from the Introduction was explored using simple population models.-->
-
-\begin{figure}[!h]
-\textbf{ \hspace{3.2 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
-\vspace{-0.35 cm}
-\begin{center}
-\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
-\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
-\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
-\end{center}
-\textbf{ \hspace{3.2 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
-\vspace{-0.35 cm}
-\begin{center}
-\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
-\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
-\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
-\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} ln(BMR) of contrasts vs.
-ln(mass) of contrasts, \textbf{(C)} contrasts for AVG.ltoe.IV vs.
-contrasts for AVG.humerus phylogenetic traits.
-\textbf{(D)} above-ground plant biomass data vs.
-Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
-Lower panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel.
-These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of contrasts for next two datasets and number of plots for the last one (see \nameref{Model_selection} of manuscript).
-\label{fig_biv_raw_cop}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
-\vspace{-0.4 cm}
-\begin{center}
-\includegraphics[width=7 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
-\includegraphics[width=7 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
-\end{center}
-\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
-\vspace{-0.4 cm}
-\begin{center}
-\includegraphics[width=7 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
-\includegraphics[width=7 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
-\caption{Upper panel shows raw data plots for \textbf{(A)} green spruce aphid-abundance data from location pair(4,11) and \textbf{(B)} leaf-curling plum aphid-firstflight data from location pair(2,5).
-Lower panel \textbf{(C - D)} shows copula plots for the corresponding datasets of the upper panel.
-These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of data-points available for both location at same year (see \nameref{Model_selection} of manuscript).
-\label{fig_aphid_raw_cop}}
-\end{center}
-\end{figure}
 

### 2018-08-31 SupportingInformation.Rmd
modified data section

@@
 updated on `r Sys.Date()`
+\pagebreak
 
@@
 }
-#----------------------------------------------------------
+
 pdf("./Results/copula_pedagog_fig_BB1.pdf",height=8,width=20)
@@
 
-\newpage
-# Descriptions of data \label{Description_data}
-<!--Each data contributor writes one to a few paragraphs about their data.-->
-
-## Soil C and N data  \label{SoilCN_data}
-
-To explore the correlation of soil organic C and total soil N across a wide geographic range (conterminous
-US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
-database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
-stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
-study.
-
-The raw data are shown in Figure \ref{fig_raw_cop_soilCN}(A). The copula is shown in Figure \ref{fig_raw_cop_soilCN}(B).
-
 <!--The raw data look like this:-->
@@
 
-\begin{figure}[!h]
-\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
-\begin{center}
-\includegraphics[width=3 in]{./Results/SoilCNraw.pdf}
-\includegraphics[width=3 in]{./Results/SoilCNcop.pdf}
-\caption[\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data]{\textbf{(A)} Plot of the raw soil C and N data, \textbf{(B)} Copula plot of soil C and N data.
-Obtained by applying ranks separately to the C and N data, and then dividing the ranks by $n+1$ where $n = `r nrow(d)` $ is the number of sampling locations (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_soilCN}}
-\end{center}
-\end{figure}
-
-## BMR data  \label{BMR_data}
-
-[**bmr-contrasts.csv**](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) : 
-This contains the trait contrasts from phylogenetically independent contrasts [@SukumaranH2010] 
-of data reported by [@GenoudIM2017] on the basal metabolic rate (BMR)
-and mass of mammals species.
-Specifically, the "SELECT" BMR data in their database were used for
- those species for which data were available.
-A phylogeny in NEXUS [@NEXUS] format was extracted from the 
- supplementary information of [@GenoudIM2017].
-Version 4.4.0 of the DendroPy library [@SukumaranH2010] was used to calculate 
-  phylogenetic independent contrasts using Python 2.7.
-For each contrast, the path length spanned by the contrast (sum of the branch
-  lengths) and the contrast for the mass, natural logarithm of the BMR and natural
-  logarithm of the mass were written as a row in the bmr-contrasts.csv file.
-
 ```{r read_BMR_data, echo=FALSE, results="markup"}
@@
 #class(BMR_dat)
-knitr::kable(head(BMR_dat[,c(1,4,5)] ,3),format='pandoc',
-             caption = "First few rows of raw BMR data.\\label{tab_rawdata_BMR}",
-             booktabs=T) # display first few rows
-```
-
-Table \ref{tab_rawdata_BMR} displays the raw data for first few contrasts.
-The raw data are shown in Figure \ref{fig_raw_cop_BMR}(A).
-The copula [using absolute values of each column in Table \ref{tab_rawdata_BMR}] is shown in Figure \ref{fig_raw_cop_BMR}(B).
+#knitr::kable(head(BMR_dat[,c(1,4,5)] ,3),format='pandoc',
+#             caption = "First few rows of raw BMR data.\\label{tab_rawdata_BMR}",
+#             booktabs=T) # display first few rows
+```
 
@@
 
-\begin{figure}[!h]
-\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
-\begin{center}
-\includegraphics[width=3 in]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
-\includegraphics[width=3 in]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
-\caption[\textbf{(A)} Plot of the raw ln(BMR) of contrasts vs.
-ln(mass) of contrasts \textbf{(B)} Copula plot for ln(BMR) vs.
-ln(mass) contrasts taken as their absolute values]{\textbf{(A)} Plot of the raw ln(BMR) of contrasts vs.
-ln(mass) of contrasts \textbf{(B)} Copula plot for absolute values of ln(BMR) vs.
-ln(mass) for all $`r nrow(v_BMR)`$ contrasts.
-Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_BMR)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).\label{fig_raw_cop_BMR}}
-\end{center}
-\end{figure}
-
-## Phylogenetic data : Anolis  \label{Phylo_data}
-
-[**contrasts.csv**](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) : 
-Raw data on measurements of anatomical features of several species of *Anolis* lizards was downloaded
-from the Dryad data package [@MahlerIRL2013Data] associated with
-[@MahlerIRL2013].
-The contrasts.csv file was produced by using version 4.4.0 of the DendroPy 
-  library [@SukumaranH2010]  to calculate 
-  phylogenetic independent contrasts on the the data in the GA_Anolis_traits.csv
-  file of Mahler et at. using their maximum clade credibility tree (GA_Anolis_MCC.nex)
-  as the phylogeny.
-
 ```{r read_phylo_data, echo=F, results="markup"}
@@
 phylo_dat<-read.csv("./Data/Anolis/contrasts.csv",header = T)
-#class(phylo_dat) #it's a data.frame
-# no NA/missing value in this dataset
-tab_phylo<-phylo_dat[,c(1,7,8)] 
-knitr::kable(head(tab_phylo,3),format='pandoc',
-             caption = "First few rows of raw phylo data.\\label{tab_rawdata_phylo}",
-             booktabs=T) # display first few rows
-```
-
-Table \ref{tab_rawdata_phylo} displays the raw data for first few contrasts for two selective phylogenetic traits, namely, AVG.ltoe.IV (average length of fourth hindtoe) and AVG.humerus (average length of humerus).
-The raw data are shown in Figure \ref{fig_raw_cop_AVG_humerus_ltoe}(A).
-The copula [taking absolute values of each column in Table \ref{tab_rawdata_phylo}] is shown in Figure \ref{fig_raw_cop_AVG_humerus_ltoe}(B).
+#class(phylo_dat) #it's a data.frame : no NA/missing value in this dataset
+#tab_phylo<-phylo_dat[,c(1,7,8)] 
+#knitr::kable(head(tab_phylo,3),format='pandoc',
+#             caption = "First few rows of raw phylo data.\\label{tab_rawdata_phylo}",
+#             booktabs=T) # display first few rows
+```
 
@@
 ```
-
-\newpage
-
-\begin{figure}[!h]
-\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
-\begin{center}
-\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
-\includegraphics[width=3 in]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
-\caption[\textbf{(A)} Plot of the contrasts for AVG.ltoe.IV vs.
-AVG.humerus phylogenetic traits.
-\textbf{(B)} Copula plot of contrasts for AVG.ltoe.IV vs.
-contrasts for AVG.humerus phylogenetic traits(taken as absolute values)]{\textbf{(A)} Plot of the contrasts for AVG.ltoe.IV vs.
-contrasts for AVG.humerus phylogenetic traits.
-\textbf{(B)} Copula plot of absolute values of contrasts for AVG.ltoe.IV vs.
-contrasts for AVG.humerus phylogenetic traits.
-Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(v_contrasts_67)`$ is the number of contrasts (see \nameref{MT-Model_selection} of manuscript).
-\label{fig_raw_cop_AVG_humerus_ltoe}}
-\end{center}
-\end{figure}
-
-## Ceder-creek data for e120 \label{ceder-creek_data}
-
-These biodiversity experiment data sets, for [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and for [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120), are collected by Tilman et al.
-over many years.
-We used thsese datasets [@Tilman2001; @Tilman2006] to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m),
-for year 2000.
 
@@
 
+```{r map_plot_aphid_plankton,results="hide",echo=F}
+source("aphid_data_calling.R")
+source("good_loclist.R")
+library(maps)
+
+#----------------- Aphid map--------------------------------------------
+d0_count<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtscount141117.csv",header=F))
+data_aphid_count<-d_allsp_data(d0_count)
+
+d0_firstflight<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsfirstflight141117.csv",header=F))
+data_aphid_ff<-d_allsp_data(d0_firstflight)
+
+selective_loc1<-good_loclist(d_allsp=data_aphid_count,sp=10,data_pt_thrs=30)
+selective_loc2<-good_loclist(d_allsp=data_aphid_ff,sp=11,data_pt_thrs=30)
+all(selective_loc1==selective_loc2)==T
+
+longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
+lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)
+
+locno<-length(longs)
+df_loc_info_org<-data.frame(loc_no=c(1:length(lats)),longs=longs,lats=lats)
+df_loc_info<-df_loc_info_org[selective_loc1,]
+xlim=c(-10,2)
+ylim=c(50, 60)
+
+pdf("./Results/aphid_map_plot.pdf")
+map("world", fill=TRUE, col="grey84", bg="white",border=0, xlim=xlim, ylim=ylim)
+map.axes(cex.axis=2.5)
+points(x = df_loc_info_org$longs, y = df_loc_info_org$lats, col="red",pch=16)
+points(x = df_loc_info$longs, y = df_loc_info$lats, col="blue", pch=0, cex=1.5)
+text(df_loc_info_org$loc_no, x = df_loc_info_org$longs, y = df_loc_info_org$lats, col="red",  pos=1)
+dev.off()
+
+#--------------Plankton map----------------------------
+
+source("plankton_north_sea_data_calling.R")
+d0_plankton_north_sea<-as.matrix(read.csv("./Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))
+data_plankton_north_sea<-d_allsp_data_plankton_north_sea(d0_plankton_north_sea)
+
+longs<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlongs201117.csv",header=F)
+lats<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlats201117.csv",header=F)
+longs<-longs[[1]]
+lats<-lats[[1]]
+
+df_loc_info_org<-data.frame(loc_no=c(1:length(lats)),longs=longs,lats=lats)
+selective_loc<-good_loclist(d_allsp=data_plankton_north_sea,sp=16,data_pt_thrs=45)
+df_loc_info<-df_loc_info_org[selective_loc,]
+xlim=c(-12,10)
+ylim=c(48, 60)
+
+pdf("./Results/plankton_map_plot.pdf")
+map("world", fill=TRUE, col="grey84", bg="white",border=0, xlim=xlim, ylim=ylim)
+map.axes(cex.axis=2.5)
+abline(h=seq(from=ylim[1],to=ylim[2],by=2), v=seq(from=xlim[1],to=xlim[2],by=2), col="black", lty=2)
+points(x = df_loc_info_org$longs+1, y = df_loc_info_org$lats+1, col="red",pch=16)
+points(x = df_loc_info$longs+1, y = df_loc_info$lats+1, col="blue", pch=0, cex=1.5)
+text(df_loc_info_org$loc_no, x = df_loc_info_org$longs+1, y = df_loc_info_org$lats+1, col="red",  pos=1)
+dev.off()
+```
+
 \begin{figure}[!h]
-\textbf{ \hspace{5.5 cm} (A) \hspace{7 cm} (B)  } \\
-\vspace{-0.2 cm}
-\begin{center}
-\includegraphics[width=3 in]{./Results/Ceder_creek_results/raw_HB_2000.pdf}
-\includegraphics[width=3 in]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
-\caption[\textbf{(A)} Plot of the plant-aboveground avg.
-Biomass vs.
-avg.
-Shannon's index (H).
-\textbf{(B)} Copula plot of plant-aboveground avg.
-Biomass vs.
-avg.
-Shannon's index.]{\textbf{(A)} Plot of the plant-aboveground avg.
-Biomass vs.
-avg.
-Shannon's index (H).
-\textbf{(B)} Copula plot of plant-aboveground avg.
-Biomass vs.
-avg.
-Shannon's index.
-Obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n=`r nrow(cc_cop_2000)`$ is the number of (9m by 9m)plots.
-(see \nameref{MT-Model_selection} of manuscript).
-\label{fig_raw_cop_cc2000}}
+\begin{center}
+\includegraphics[width=8 cm,height=8 cm]{./Results/aphid_map_plot.pdf} 
+\caption{Locations of Rothamsted Insect Survey suction-trap across UK are indicated by the red circles 
+out of which only blue boxes are considered for our study based on having atleast 30 annual data points. \label{fig_aphid_map}}
 \end{center}
@@
 
-## Abundance data for the green spruce aphid and first flight date data for the leaf-curling plum aphid in the UK\label{Aphid_data}
-   * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
-   * We threw out all locations that had fewer than 30 years data. 
-   * We use species=10 [Name : "Green spruce aphid"] for aphid-count data. 
-   * We use species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data.
-
-## Abundance data for *Ceratium furca* in UK seas \label{Pns_data}
-   * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
-   * We threw out all locations that had fewer than 45 years data.
-   * We use species=16 [Name : "Ceratium furca"] for Plankton_North_Sea data.
-   * *Ceratium furca* is a dicoflagellate commonly found in UK seas.
-   * note, when we edit this, we cannot call it the North Seas since data come also from other seas around the UK.
-
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=8 cm,height=8 cm]{./Results/plankton_map_plot.pdf} 
+\caption{CPR data collected from 26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ on map) around UK sea (indicated by the red circles) but out of which only blue boxes are considered for our study based on having atleast 45 annual data points.
+\label{fig_plankton_map}}
+\end{center}
+\end{figure}
+
+<!--
 ## Methane-flux data  \label{Methane_data}
@@
 
-* We threw out all locations that had fewer than 50 dates data.
+* We threw out all locations that had fewer than 50 dates data.-->
 

### 2018-09-04 Paper.Rmd
modified

@@
 \newpage
-# Introduction
+# Introduction \label{Intro}
 
@@
 Basically, we want to address the following questions.
-1) Do datasets in ecology and related fields have non-Normal copula srtructure?
+\textbf{(Q1)} Do datasets in ecology and related fields have non-Normal copula srtructure?
 Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence?
-2) If so, what are some possible causes of non-Normal copula structure?
-And 3) what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
-In \nameref{Material&Methods}, we described briefly about each datasets we used to answer these questions and provided a detailed descriptions of implemented methods.
+\textbf{(Q2)} If so, what are some possible causes of non-Normal copula structure?
+And \textbf{(Q3)} what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
+In \nameref{Material&Methods}, we described briefly about each bivariate and multivariate datasets we used to answer these questions and provided a detailed descriptions of implemented methods.
 In the \nameref{Results} section, we summarized our finidings for each datasets.
-Next two sections are focused on possible causes and consequences of our analysis.
+Next two sections and are focused on possible causes and consequences of non-Normal copula structure in ecology and related field.
 Finally, we concluded in the \nameref{Discussion} section.
@@
 For bivariate analysis we used four datasets which are briefly described as follows;
-to begin with **Soil C and N data**, to explore the correlation of soil organic C and total soil N across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
+First, to explore the dependence of **soil organic C** and total **soil N** across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
 database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
@@
 Version 4.4.0 of the DendroPy library [@SukumaranH2010] was used to calculate phylogenetic independent contrasts using Python 2.7.
-For each contrast, the path length spanned by the contrast (sum of the branch lengths) and the contrast for the mass, natural logarithm of the BMR and natural logarithm of the mass were written as a row in the bmr-contrasts.csv file.
+For each contrast, the path length spanned by the contrast (sum of the branch lengths) and the contrast for the mass, natural logarithm of the BMR and natural logarithm of the mass were written as a row in the [bmr-contrasts.csv](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) file.
 
@@
 
-For multivariate analysis, we applied four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014].
+Similarly, we applied multivariate analysis on four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014].
 These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species.
@@
 
-Question 1 from the Introduction was answered via statistical analysis of the
+Question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} was answered via statistical analysis of the
 data described above, using both model selection and nonparametric approaches.
@@
 and Bayesian Information Criterion (BIC) was used to
-help answer question 1 from the Introduction.
+help answer \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
 
@@
 
-The first part or question 1 from the Introduction was answered for a bivariate dataset 
+The first part of \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} was answered for a bivariate dataset 
 $(x_i,y_i)$ for $i=1,...,n$ by comparing the AIC weights of the non-Normal copula
 families with that of the Normal family. The same was also done using BIC.
-The second part of question 1 was answered by comparing 
+The second part of \hyperref[Intro]{\textbf{(Q1)}} was answered by comparing 
 the model-averaged lower- and upper-tail dependence
 values to the lower- and upper-tail dependence values of a Normal copula, both of which 
-are 0. The third part of question 1
+are 0. The third part of \hyperref[Intro]{\textbf{(Q1)}}
 was answered by comparing the difference of the model-averaged lower- and upper-tail 
@@
 In addition to the model selection approach described above, nonparametric 
-approaches were also used to answer question 1 from the introduction.
+approaches were also used to answer \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
 Using two distinct methods, when results are consistent, provides greater confidence 
@@
 Kendall (or Spearman)
-correlation. Thus the comparison helps answer the first two parts of question 1 
-from the Introduction for bivariate datasets. 
+correlation. Thus the comparison helps answer the first two parts of question \hyperref[Intro]{\textbf{(Q1)}} 
+from the \nameref{Intro} for bivariate datasets. 
 Significant deviations correspond to deviations from normal 
@@
 provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
-third part of question 1 from the Introduction for bivariate datasets.
+third part of question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} for bivariate datasets.
 
@@
 or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
-Normal null hypothesis, and answering question 1 from the Introduction.
+Normal null hypothesis, and answering question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
 Bootstrapping was performed by resampling the locations of measurement, with replacement. 
@@
 
-## question 1
+<!--## question 1
 
@@
 - Do they show tail dependence distinct from that of a Normal copula?
-- In particular do they show asymmetric tail dependence?
+- In particular do they show asymmetric tail dependence?-->
 
-To answer Q1, we choose one representative datasets for each category : with a bivariate (soil C and N data) and
-a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.
+To answer \hyperref[Intro]{\textbf{(Q1)}}, we choose one representative datasets for each category : with a bivariate (soil C and N data) and a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.
 
-### soil C and N data \label{Ex_SoilCN}
+### Example for bivariate case : soil C and N data \label{Ex_SoilCN}
 
@@
 
-As described in the \ref{SM-SoilCN_data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central
-pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig. \ref{fig_biv_raw_cop}(A) and (D),respectively.
+As described in the \nameref{Data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
+\ref{fig_biv_raw_cop}(A) and (E),respectively.
 We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.
@@
 Based on AIC weights, `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the best fit.
-The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
+The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of \hyperref[Intro]{\textbf{(Q1)}} for these data.
 
 The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`.
-These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.
+These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of \hyperref[Intro]{\textbf{(Q1)}} for these data.
+Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}.
 To be precise, this difference being negative, indicated stronger upper-tail dependence than lower.
@@
 Results indicate that tail dependence of data is stronger in both the lower and upper tails than would be expected under a 
-null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
+null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}. The dependence between C and N is also 
 weaker in the middle of the distributions
@@
 
-We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{SM-fig_raw_cop_soilCN}(B). 
-Results (Table \ref{SM-tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.
+We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{fig_biv_raw_cop}(E). 
+Results (Table \ref{SM-tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}.
 
-We summarized our results columnwise in Table \ref{tab_bivar_summary} for all bivariate datasets plotted in Fig. \ref{fig_biv_raw_cop}.
+We summarized our results columnwise in Table \ref{tab_bivar_summary} for all bivariate datasets that were plotted in Fig. \ref{fig_biv_raw_cop}.
 
@@
 preserving $1000$ Normal surrogates and calculated the rank for actual data as the position if all these $1001$ values were 
-arranged in ascending order. So, whenever those ranks are $>975$ or $<25$, that indicate a significantly ($95\%$ CI) different 
+arranged in ascending order. So, whenever those ranks were $>975$ or $<25$, that indicated a significantly ($95\%$ CI) different 
 tail-dependence measures based on Normal-null hypothesis.
@@
 in Table \ref{tab_bivar_summary}, Normal copula were always a poor fit compared to 
-highest AIC-weight based asymmetric copulas - which answers the first part of Q1. 
+highest AIC-weight based asymmetric copulas - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}. 
 Among four datasets presented there, first three have significant upper tail dependence whereas the last one has significant lower tail dependence compared to Normal.
@@
 (lower-upper) tail dependence and very high ranks of actual copula 
-data for $\Ps_l - \Ps_u$ statistics for the last one, which answers the second and third part of Q1
+data for $\Ps_l - \Ps_u$ statistics for the last one, which answers the second and third part of \hyperref[Intro]{\textbf{(Q1)}}
 (For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and 
@@
 
-### Abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
+### Example for multivariate case : abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
 
@@
 As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid abundance data, model selection among copulas was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations.
-Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question 1 for these data.
+Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question \hyperref[Intro]{\textbf{(Q1)}} for these data.
 Comparing the
@@
 The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively.
-These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data.
+These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question \hyperref[Intro]{\textbf{(Q1)}} for these data.
 Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.
 
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1.
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}.
 Fig.
@@
 indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
-exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question 1.
+exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}.
 
@@
 <!--Question 2 analyses-->
-## Some possible causes of non-Normal copula structure
+## Some possible causes of non-Normal copula structure \label{Causes}
 
@@
 <!--Question 3 analyses-->
-## Consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
+## Consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? \label{Consequences}
 

### 2018-09-05 Paper.Rmd
modified

@@
 
-Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016] or a short period study of intersite dependence of rainfall data [@Serinaldi2008], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data.
+Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data.
 So, in ecological perspective, when we find a dominantly lower (upper) tail dependence structure, it means both incidence are rare (abundant) synchronously.
@@
 
-Q1/2: How are copula structure of ecologically influential climate variables changing as part of climate change?
-We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the pops, but we don’t really know how a change in the environment cop structure would make what kind of change in the pop cop structure.
-How?
+This section is dedicated to explore question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}.
+We want to find how are copula structure of ecologically influential climate variables changing as part of climate change?
+We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we don’t really know how a change in the environment copula structure would make what kind of change in the population copula structure.
 
@@
 
-We show here that if a non-normal copula structure describes the dependence of an environmental
+Our hypothesis is if a non-normal copula structure describes the dependence of an environmental
 variable between two locations, then there can be a non-normal copula structure for the
 dependence of populations in the locations that are influenced by the environmental variables.
+This assumption is inspired by the the idea of Moran-effect [@Moran1953] which was originally used to explain synchronization of widely dispersed populations.
 
@@
 Fig.
-\ref{SM-fig_Cause4copula_params_GOF_2500} represents similar figure as of Fig.
+\ref{SM-fig_Cause4copula_params_GOF_2500} represents similar feature as of Fig.
 \ref{SM-fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$.
@@
 We simulated our model (\ref{eq.AS}) for different values of $b$ with $a=0.2$, $r=0.8$.
-$10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ time step of simulation to plot the population-copula for a specific $b$ and a specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
+$10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ time step of simulated points to plot the population-copula for a specific $b$ and a specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
 Fig.
@@
 
+This section is attributed to the analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}. Here we summarized our findings as follows;
+
 ### Effect on skewness of the spatial mean
 
-Here, we claim that because of significant tail dependence of the multivariate copula structure, the skewness of the time-series data averaged over all selective locations changed significantly from the individual skewness of time-series data taken at each pairwise location.
-Spatial mean of the time series data extracts that tail dependence of the multivariate data set, if any, and reflected in the averaged skewness.
-We tested this with null hypothesis $H_{0}$.
-If $H_{0}$ is true that means our multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
-So, first we generated $10000$ number of surrogate normal copula from our given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
-Then we plotted the distribution of spatial mean of each surrogate time series data set and the given multivariate data.
-If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence.
+We hypothesized that because of distinct tail dependence of the multivariate copula structure, the skewness of the time-series data averaged over all selective locations changed significantly from the individual skewness of time-series data taken for each pairwise location.
+Spatial mean of the time series data extracted that tail dependence of the multivariate data set, if any, and reflected in the averaged skewness.
+We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
+So, first we generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
+Then we plotted the distribution of spatial mean of each surrogate time series data and of the given multivariate data.
+If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).
 
-To verify that we used aphid-count data, aphid-firstflight data, plankton data and methane-flux data.
-For aphid data we choose only those locations for which atleast $30$ number of data points present in each time series when taken as pairwise locations.
-For plankton-UK-sea data and methane-flux data that threshold on number of data points are $45$ and $50$, respectively.
-
-- Aphid count \label{skewness_count}
-
-Figure \ref{SM-fig_skewness_count} shows that averaged skewness is much lower than expected if the aphid count data came from a normal copula.
-Here, $p$ value indicates the fraction of surrogate-skewness which are lower than the averaged skewness of the given data (red line in Figure \ref{SM-fig_skewness_count}).
-As $p<0.05$, we reject $H_{0}$.
-This result is consistent with the non-parametric analysis and model fitting results which indicated lower tail dependence.
-
-- Aphid first flight \label{skewness_ff}
-
-Figure \ref{SM-fig_skewness_ff} shows that averaged skewness is much higher than expected if the aphid first flight data came from a normal copula.
-Here, $p$ value indicates the fraction of surrogate-skewness which are greater than the averaged skewness of the given data (red line in Figure \ref{SM-fig_skewness_ff}).
-As $p<0.05$, we reject $H_{0}$ in support for upper tail dependence.
-
-- UK Sea plankton \label{skewness_pns}
-
-Figure \ref{SM-fig_skewness_pns} shows that averaged skewness is lower than expected if the plankton data came from a normal copula.
-Our previous results on non-parametric analysis and model fitting showed lower tail dependence.
-Here, as $p>0.05$, we cannot reject $H_{0}$, .......???
-
-- Methane-flux data \label{skewness_methane}
-
-Figure \ref{SM-fig_skewness_methane} shows that averaged skewness is lower than expected if the plankton north sea data came from a normal copula.
-Our previous results on non-parametric analysis and model fitting showed lower tail dependence.
-As $p<0.05$, we reject $H_{0}$.
-This result is consistent with the non-parametric analysis and model fitting results which indicated lower tail dependence.
+To verify that we used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data.
+Figure \ref{SM-fig_skewness_count} showed that averaged skewness was much lower than expected if the aphid count data came from a normal copula.
+In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data (red line in Figure \ref{SM-fig_skewness_count}).
+As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results.
+On the other hand, Figure \ref{SM-fig_skewness_ff} demonstrated that averaged skewness was much higher than expected if the aphid first flight data came from a normal copula.
+As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red line in Figure \ref{SM-fig_skewness_ff}) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence.
+Figure \ref{SM-fig_skewness_pns} showed that averaged skewness was lower than expected if the plankton data came from a normal copula.
+Our previous results on non-parametric analysis and model fitting also supported lower tail dependence.
+But, as $p>0.05$, we could not reject $H_{0}$ in that case.
+Finally, for methane-flux data, the averaged skewness was lower than expected if the data came from a normal copula (Figure \ref{SM-fig_skewness_methane}).
+As $p<0.05$, we rejected $H_{0}$ and that result was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence.
 
@@
 
-In order to study the effects of non-Normal copula structure on extinction risk for population of a species at different locations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
+In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
 <!--did I write correct spelling for lewton model?-->
@@
 $r$ is the liner growth rate and $K$ is the carrying capacity of the environment.
-$\epsilon$ is the environmental noise which may take strong right tail (or left tail) dependence.
-We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix.
-$D$ can take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch can interact with the rest of all (we called it as global dispersion).
+$\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence.
+We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion).
 We simulated Eq.
@@
     * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
-[see first plot of Fig.
+[see first subplot of Fig.
 \ref{SM-fig_risk_local_disp} and Fig.
@@
 
-### Effects on Taylor's law
+### Effect on Taylor's law
 
-Spatial Taylor's law, which shows linear relationship between `log`(variance) and `log`(mean) of population densities of a species taken from different locations, is a well-known pattern found in ecology and related fields.
-Here, we want to examine the nature of the slope and how it changes if population time series (taken as $50$ years) from different locations (say, $numloc=25$) have some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula.
-First, we simulated (N $\times$ n $\times$ ns) dimensional normal copula with covariance $\rho =0.7$, where $(N=50,n=25,ns=10)$ are the length of timeseries, number of sampling locations and number of replica for that copula, respectively.
+Spatial Taylor's law, which shows linear relationship between `log`(variance) and `log`(mean) of population densities of a species from different locations, is a well-known pattern found in ecology and related fields [@Taylor1961, @Taylor1988, @Cohen2013, @Doring2015, @Kuo2016, @Xu2016, @Reuman2017].
+Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula.
+First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with covariance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of timeseries, number of sampling locations and number of replica for that copula, respectively.
 Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal.

### 2018-09-11 Paper.Rmd
modified@sept10

@@
 Here, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets.
-In addition to a copula based \nameref{Model_selection}, we also carried out a \nameref{NPA} with three different statistics to verify our model selection based results.
+In addition to a copula based \nameref{Model_selection}, we also carried out a \ref{SM-NPA} with three different statistics to verify our model selection based results.
 
@@
 
-For bivariate analysis we used four datasets which are briefly described as follows;
-First, to explore the dependence of **soil organic C** and total **soil N** across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
-database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
-stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
-study.
-
-Next, the [**BMR data**](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) contains the trait contrasts from phylogenetically independent contrasts [@SukumaranH2010] of data reported by [@GenoudIM2017] on the basal metabolic rate (BMR) and mass of mammals species.
-Specifically, the "SELECT" BMR data in their database were used for those species for which data were available.
-A phylogeny in NEXUS [@NEXUS] format was extracted from the supplementary information of [@GenoudIM2017].
-Version 4.4.0 of the DendroPy library [@SukumaranH2010] was used to calculate phylogenetic independent contrasts using Python 2.7.
-For each contrast, the path length spanned by the contrast (sum of the branch lengths) and the contrast for the mass, natural logarithm of the BMR and natural logarithm of the mass were written as a row in the [bmr-contrasts.csv](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) file.
-
-[**Third dataset**](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) was a raw data on measurements of anatomical features of several species of *Anolis* lizards, which was downloaded
-from the Dryad data package [@MahlerIRL2013Data] associated with [@MahlerIRL2013].
-The [contrasts.csv file](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) was produced by using version 4.4.0 of the DendroPy library [@SukumaranH2010]  to calculate phylogenetic independent contrasts on the the data in the GA_Anolis_traits.csv file of Mahler et at.
-using their maximum clade credibility tree (GA_Anolis_MCC.nex) as the phylogeny.
-
-Absolute values of traits for each contrast are considered to construct corresponding copula for second and third datasets.
-
-Fourth, we used **Ceder-creek data for e120** which resulted from a biodiversity experiment, carried out by Tilman and his co-warkers [@Tilman2001; @Tilman2006] over many years to collect [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120) data.
-These datasets were used to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m),
-for year 2000.
-
-The raw data and copula for each four datasets are shown along each column of Figure \ref{fig_biv_raw_cop}.
-
-Similarly, we applied multivariate analysis on four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014].
-These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species.
-We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony.
-In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}).
-Figure \ref{fig_aphid_raw_cop} represents raw and copula data plot for aphid count ("Green spruce aphid") and aphid-firstflight date ("Leaf-curling plum aphid") taken from two locations.
-
-Our third dataset for multivariate analysis is a plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013).
-These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
-The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
-unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
-location in the ocean.
-Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979].
-All the samples available for a given variable in a given area in a given month of a given year were
-averaged, producing a monthly time series for each year.
-Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).
-Monthly values were converted to yearly values by averaging over all months.
-This approach was used by [@Sheppard2017] earlier.
-But here we only allowed maximum 4 months in a year with missing data which were replaced by those medians.
-If a year got 5 or more missing monthly data value then we set NA value for that particular year.
-We used data for species=16 [Name : "Ceratium furca"] which is a dicoflagellate commonly found in UK seas.
-Among 26 locations, only those having atleast 45 years time series data, were considered (see Figure \ref{SM-fig_plankton_map}).
-
-Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
-Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
-Additional information on the history of the site and soil types is detailed in @Jarecke2016.
-Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at
-daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
-additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.
+We used four datasets in each category of bivariate (soil C and N data, BMR data, Anolis data, Ceder-creek experiment data) and multivariate (aphid-abundance data, aphid-firstflight data, plankton-abundance data and methane-flux data) examples.
+Details of each datasets used were given in \ref{SM-Data_detailed}.
 
@@
 
-### Nonparametric analysis of tail dependence \label{NPA}
+<!--### Nonparametric analysis of tail dependence \label{NPA}-->
 
-In addition to the model selection approach described above, nonparametric 
-approaches were also used to answer \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
+In addition to the model selection approach described above, **nonparametric approaches** (for details see Appendix \ref{SM-NPA}) to quantify lower ($\cor_l$, $\Ps_l$ and $\Dsq_l$ statistics) and upper tail dependence ($\cor_u$, $\Ps_u$ and $\Dsq_u$ statistics) were also used to answer \hyperref[Intro]{\textbf{(Q1)}}.
 Using two distinct methods, when results are consistent, provides greater confidence 
@@
 
-Left- and right-tail dependence quantify the extent to which two ranked
-variables $u_i$, $v_i$ for $i=1,...,n$ 
-(see \nameref{Model_selection} for defintions of $u_i$ and $v_i$)
-are related to each other in the left and right tails, respectively, of their 
-distributions [@nelsen2007_copula, @joe2014_dependence]. 
-Our nonparametric study of tail dependence uses three
-statistics which can quantify the extent to which $u_i$ and $v_i$ are 
-related in any part of their distributions. We here describe the 
-statistics. The statistics were formulated with positively associated
-variables in mind. All the variables we study are positively associated
-when they are significantly associated. So definitions and associated 
-diagrams will assume positive associations, though generalizations are 
-possible. Given two bounds 
-$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
-$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
-fall (these are the blue and red lines, respectively, in 
-Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
-between $u_i$ and $v_i$ in the region bounded by these lines.
-Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
-dependence in the left parts of the distributions of $u$ and $v$, and 
-using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
-the right parts of the distributions.
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
-\caption{Schematic diagram supporting defintions of nonparametric tail dependence statistics \label{fig_stat_image}}
-\end{center}
-\end{figure}
-
-<!--
-\begin{figure}[!h]
-\textbf{ \hspace{6 cm} (A) \hspace{5.6 cm} (B)} \\
-\vspace{-0.6 cm}
-\begin{center}
-\hspace{0.5 cm}
-\includegraphics[width=5 cm,height=4.3 cm]{./stat_image.jpg} \hspace{0.6 cm}
-\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/CorlmCoru_C.pdf}
-\end{center}
-\textbf{ \hspace{6 cm} (C) \hspace{5.7 cm} (D)} \\
-\vspace{-0.6 cm}
-\begin{center}
-\hspace{0.2 cm}
-\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/PlmPu_C.pdf} \hspace{0.3 cm}
-\includegraphics[width=6 cm]{./Results/stat_results/stat_testing/D2umD2l_C.pdf}
-\caption{(A) Schematic diagram supporting defintions of nonparametric tail dependence statistics; (B) Results of tests of $\cor_l-\cor_u$, (C) $\Ps_l-\Ps_u$, and (D) $\Dsq_u-\Dsq_l$ reveal the known asymmetry (more lower tail than upper) of tail dependence of the Clayton copula with $n=35$ data points.
-\label{fig_stat_image}}
-\end{center}
-\end{figure}
--->
-
-The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
-correlation of $u_i$ and $v_i$ that is attributable to the points 
-lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
-\begin{equation}\label{eq.Cor}
-   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
-   (u_i-\mean(u)) (v_i-\mean(v)),
-\end{equation}
-where means and variances are computed using all $n$ data 
-points, but the
-sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
-$u_i+v_i < 2u_b$. 
-Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
-between $u$ and $v$ in the region given by the bounds.
-
-We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
-$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
-$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
-and a distance less than $h$ from the line $v=u$, divided by the 
-total number of points within the bounds. This function is
-defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
-is half the longer of the two segments obtained by intersecting the
-bounds with the unit square. It is easy to see $S(0)=0$ and 
-$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
-of $h$, to be the area within the bounds and the unit square
-and within distance $h$ of the line $v=u$, divided by the area
-within the bounds and the unit square. This is the expected value of
-$S(h)$ for independent data. We then defined
-\begin{equation}\label{eq.P}
-   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
-\end{equation}
-As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
-indicate stronger positive association
-between $u$ and $v$ in the region given by the bounds.
-
-The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
-distance between points satisfying $u_i+v_i>2l_b$ and 
-$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
-$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
-$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
-then small. 
-Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
-dependence in the portions of the distributions of $u$ and $v$ given by 
-the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
-$\Dsq_{l_b,u_b}$ indicate strong dependence.
-
-For large datasets (i.e., datasets with large $n$),
-we can consider $l_b$ and $u_b$ close together without incurring undue
-sampling variation in our statistics, and we can consider 
-several different bands $(l_b,u_b)$ to understand how dependence
-varies in different parts of the distributions. But for datasets with 
-smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
-We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
-and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
-$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.
-
-The difference $\cor_l-\cor_u$ is expected to be positive if 
-dependence in the left halves of the distributions is stronger 
-than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
-This holds, for instance, for the Clayton copula [@nelsen2007_copula].
-The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
-"lower" statistics occur in opposite order, is also expected to be 
-positive under the same conditions. We use these three differences
-to detect asymmetry of tail dependence. Likewise, for large datasets,
-we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
-$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
-$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
-
-<!--I deleted the phrase about unit testing because unit testing should be mentioned in the supp mat-->
-We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas (for details see Appendix \ref{SM-Test_npa_stat}).
 <!--***DAN: There is either too little or too much here in this 
@@
 
-For bivariate datasets, we compared values 
-of the statistics $\cor_{l_b,u_b}$,
-$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to 
-distributions of values of the same statistics computed on surrogate datasets that were produced from 
-the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
-the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
-correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
-but had Normal copula structure. Thus our comparisons tested the null hypothesis 
-that our statistics took values on the empirical data no different from what would have been expected 
-if the copula structure of the data were Normal, but the data were otherwise 
-statistically unchanged.
-Details of how surrogates were prroduced are in Appendix \ref{SM-surrog_test}, and code is in [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R).
 <!--***DAN: I think the below material belongs in supp mat, because the reader can get the
@@
 
-The comparison of one of the
-statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
-empirical data, to the distribution
-of its values computed on many surrogate datasets 
-provides a test of whether dependence in the part of the distribution specified 
-by $l_b$ and $u_b$ is different from
-what would be expected from a null hypothesis of normal copula structure but 
-unchanged marginals and unchanged 
-Kendall (or Spearman)
-correlation. Thus the comparison helps answer the first two parts of question \hyperref[Intro]{\textbf{(Q1)}} 
-from the \nameref{Intro} for bivariate datasets. 
-Significant deviations correspond to deviations from normal 
-copula structure. In particular, deviations 
-using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
-different from that of a normal copula;
-likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
-from that of a normal copula.
-
-The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
-as computed on the 
-empirical data, to the distribution
-of its values on many surrogate datasets 
-provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
-third part of question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} for bivariate datasets.
-
-For multivariate datasets, a spatial resampling scheme was used instead of the surrogate 
-approach described above.
-The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
-widely used spline correlogram method; the same scheme was also used by @walteretal2017 
-as part of a method for 
-performing model selection among matrix regression models. For each pair of 
-locations $i \neq j$, the statistic
-of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
-$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
-$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
-values $s_{ij}$ were then arranged in
-a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
-$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
-average strength of 
-dependence, across all pairs of distinct locations, for data in the portion of the distributions 
-given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
-asymmetry of tail dependence across all pairs of locations. 
-Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
-distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
-revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
-or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
-Normal null hypothesis, and answering question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
-Bootstrapping was performed by resampling the locations of measurement, with replacement. 
-This corresponds to 
-resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
-[@bjornstadfalck2001 ; @walteretal2017].
-
 <!--Theoretical models
@@
 
-As described in the \nameref{Data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
-\ref{fig_biv_raw_cop}(A) and (E),respectively.
+As described in the Appendix \ref{SM-Data_detailed}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
+\ref{fig_biv_raw_cop}(A) and (E), respectively.
 We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.
@@
 
-As described in \nameref{NPA}, each of our statistics, as computed for the 
+As described in Appendix \ref{SM-NPA}, each of our statistics, as computed for the 
 ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
@@
 computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
-resampling scheme, as described in \nameref{NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
+resampling scheme, as described in \ref{SM-NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
 $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
@@
 This section is dedicated to explore question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}.
-We want to find how are copula structure of ecologically influential climate variables changing as part of climate change?
-We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we don’t really know how a change in the environment copula structure would make what kind of change in the population copula structure.
-
-### The Moran effect and copula structure in populations
-
-Our hypothesis is if a non-normal copula structure describes the dependence of an environmental
+We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change?
+<!--We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we don’t really know how a change in the environment copula structure would make what kind of change in the population copula structure.-->
+<!--### The Moran effect and copula structure in populations-->
+Firstly, Our hypothesis was if a non-normal copula structure describes the dependence of an environmental
 variable between two locations, then there can be a non-normal copula structure for the
 dependence of populations in the locations that are influenced by the environmental variables.
-This assumption is inspired by the the idea of Moran-effect [@Moran1953] which was originally used to explain synchronization of widely dispersed populations.
-
-To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
-\ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.
-
-\begin{equation}\label{eq.AR1}
-P_{i}(t+1)=\beta P_{i}(t)+\alpha \epsilon_{i}(t)
-\end{equation}
-
-To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$.
-We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches.
-Should their dependence be exactly the same as of the noises or not ?
-
-To test that our approach is of two way : first, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
-Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome which was a bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
-\ref{SM-fig_Cause4copula_params_GOF_500}).
-We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size and reported p-values for both statistics (Cramer-von Mises and Kolmogorov-Smirnov).
-p-values greater or equal to $0.05$ (above the dashed line in Fig.
-\ref{SM-fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters.
-Fig.
-\ref{SM-fig_Cause4copula_params_GOF_2500} represents similar feature as of Fig.
-\ref{SM-fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$.
-(For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
-
-Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
-Then we plotted mean of each measured quantity with standard error against Spearman's Rho (or Kendall's Tau) of the noise copula family.
-Fig.
-\ref{SM-fig_Cause4copula_C_spear} and Fig.
-\ref{SM-fig_Cause4copula_C_kend} represent this picture for noise generating from a lower tail dependent Clayton copula whereas for Fig.
-\ref{SM-fig_Cause4copula_SC_spear} and Fig.
-\ref{SM-fig_Cause4copula_SC_kend} noise came from a upper tail dependent Survival Clayton family.
-We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach, here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise.
-(For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
-
+This assumption was inspired by the the idea of Moran-effect [@Moran1953] which was originally used to explain synchronization of widely dispersed populations.
+We tested our hypothesis with a two habitat patch model (1st order AR method) in noisy environment and described our findings in Appendix \ref{SM-Cause_Moran}.
+<!--
 ### Asymmetric sensitivity and copula structure in populations
-<!--Another model, distinct from the above, Dan and Shya to discuss-->
-Here we intend to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) beyond a certain expectation threshold.
-To test this assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;
-
-\begin{equation}\label{eq.AS}
-P_{i}(t+1) = b P_{i}(t) + g(\epsilon_{i}(t)) + h(\delta_{i}(t))
-\end{equation}
-
-$b$ was the linear growth rate.
-Global environmental fluctuation function $g(\epsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq.
-\ref{eq.AS_g}) where $\epsilon=(\epsilon_{1}, \epsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\epsilon}$.
-
-\begin{align}
-g(\epsilon) = \begin{cases} 
-g_{1} = \begin{cases}
-\epsilon & \text{if $\epsilon< 0$}; \\
-0 & \text{otherwise}.
-\end{cases}\\ \\
-g_{2} = \begin{cases}
-\epsilon & \text{if $\epsilon> 0$}; \\
-0 & \text{otherwise}\label{eq.AS_g}.
-\end{cases}
-\end{cases} \text{where } \epsilon = N(\textbf{0},\Sigma_{\epsilon})\;\ \text{with} \;\ \Sigma_{\epsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
-\end{align}
-
-Local environmental fluctuation function $h(\delta)$ varied according to Eq.
-\ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$.
-$\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$.
-
-\begin{equation}\label{eq.AS_h}
-h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
-\end{equation}
-
-We simulated our model (\ref{eq.AS}) for different values of $b$ with $a=0.2$, $r=0.8$.
-$10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ time step of simulated points to plot the population-copula for a specific $b$ and a specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
-Fig.
-\ref{SM-fig_ms_asym_b_0.1} and Fig.
-\ref{SM-fig_ms_asym_b_0.9} show these results for $b=\pm0.1$ and $b=\pm0.9$, respectively.
-We also carried out model selection and non-parametric statistical analysis with each copula of those figures and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_8}.
-Table \ref{SM-tab_summary_stat_asym_sens} represent the model selection results with AIC weights for best fitted copula and normal copula, model averaged lower and upper tail dependence, p-values (both Cramer-von Mises statistic and Kolmogorov-Smirnov statistic) for goodness of fit test.
-In each case, successful bootstrap was $100\%$.
-According to our model selection approach, for $b=\pm0.1$, non-normal copula (with higher AIC weights) were always best-fit with some asymetric tail dependence whereas for $b\pm0.9$, most of the time normal was best.
-It is probably because the populations get long-time exposure to environmental changes for slower growth(or decay) rate and so global fluctuation can affect them more efficiently.
-Our non-parametric analysis results (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_8}) are consistent with this model selection approach.
+Another model, distinct from the above, Dan and Shya to discuss-->
+On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold.
+So, we considered another two-habitat patch model which was influenced by both local and global noises and found that a long time exposure to assymetric copula for global fluctuation shaped the population copula accordingly (for details, see Appendix \ref{SM-Cause_asym_sens}).
 
@@
 
-This section is attributed to the analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}. Here we summarized our findings as follows;
-
-### Effect on skewness of the spatial mean
-
-We hypothesized that because of distinct tail dependence of the multivariate copula structure, the skewness of the time-series data averaged over all selective locations changed significantly from the individual skewness of time-series data taken for each pairwise location.
-Spatial mean of the time series data extracted that tail dependence of the multivariate data set, if any, and reflected in the averaged skewness.
-We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
-So, first we generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
-Then we plotted the distribution of spatial mean of each surrogate time series data and of the given multivariate data.
-If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).
-
-To verify that we used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data.
-Figure \ref{SM-fig_skewness_count} showed that averaged skewness was much lower than expected if the aphid count data came from a normal copula.
-In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data (red line in Figure \ref{SM-fig_skewness_count}).
-As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results.
-On the other hand, Figure \ref{SM-fig_skewness_ff} demonstrated that averaged skewness was much higher than expected if the aphid first flight data came from a normal copula.
-As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red line in Figure \ref{SM-fig_skewness_ff}) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence.
-Figure \ref{SM-fig_skewness_pns} showed that averaged skewness was lower than expected if the plankton data came from a normal copula.
-Our previous results on non-parametric analysis and model fitting also supported lower tail dependence.
-But, as $p>0.05$, we could not reject $H_{0}$ in that case.
-Finally, for methane-flux data, the averaged skewness was lower than expected if the data came from a normal copula (Figure \ref{SM-fig_skewness_methane}).
-As $p<0.05$, we rejected $H_{0}$ and that result was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence.
-
-### Effect on extinction risk
-
-In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
-<!--did I write correct spelling for lewton model?-->
-
-\begin{equation}\label{eq.extrisk_model}
-P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
-\end{equation}
-
-$r$ is the liner growth rate and $K$ is the carrying capacity of the environment.
-$\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence.
-We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion).
-We simulated Eq.
-\ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $25$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;
-
-- LC model (r=0) :
-
-    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
-[see first subplot of Fig.
-\ref{SM-fig_risk_local_disp} and Fig.
-\ref{SM-fig_risk_global_disp}]
-
-- Ricker model (r $\neq$ 0):
-
-    * For ricker model when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
-For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
-[see Fig.
-\ref{SM-fig_risk_local_disp} - \ref{SM-fig_risk_global_disp}]
-
-    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{SM-fig_risk_K_effect}]
-
-With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig.
-\ref{SM-fig_risk_numlocs_effect_local_disp} - Fig.
-\ref{SM-fig_risk_numlocs_effect_global_disp}] for both model.
-
-### Effect on Taylor's law
-
-Spatial Taylor's law, which shows linear relationship between `log`(variance) and `log`(mean) of population densities of a species from different locations, is a well-known pattern found in ecology and related fields [@Taylor1961, @Taylor1988, @Cohen2013, @Doring2015, @Kuo2016, @Xu2016, @Reuman2017].
-Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula.
-First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with covariance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of timeseries, number of sampling locations and number of replica for that copula, respectively.
-Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal.
-With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D).
-We also tested statistically that whether linearity, homoskedasticity is maintained in each copula dataset or not and plotted our results alongwith root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).
-
-\begin{figure}[!h]
-\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.1 cm} (D)} \\
-\vspace{-0.5 cm} 
-\begin{center}
-\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC.pdf}\\
-\end{center}
-\textbf{ \hspace{3.1 cm} (E) \hspace{2.4 cm} (F) \hspace{2.3 cm} (G) \hspace{2.3 cm} (H) \hspace{2.3 cm} (I)}\\
-\vspace{-0.5 cm}
-\begin{center}
-\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_5stats.pdf}
-\caption[Spatial Taylor's Law results]{Spatial Taylor's Law results: log(spatial variance) vs.
-log(spatial mean) for $ns=10$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of hhomoskedasticity, (G) root mean square error, (H) intercept, (I) slope of the first panel plots.
-\label{fig_spTL}}
-\end{center}
-\end{figure}
+This section is attributed to analyze of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}.
+We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data : we tested with four multivariate datasets and found results consistent with model selection and nonparametric approach (For details see Appendix \ref{SM-Consequence_skewness}).
+We then showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}).
+Finally, we studied **(iii)** the effects of non-Normal copula on extinction risk with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}).
 

### 2018-09-11 SupportingInformation.Rmd
modified@sept10

@@
 
-\newpage
+# Description of Data \label{Data_detailed}
+
+For bivariate analysis we used four datasets which are briefly described as follows;
+First, to explore the dependence of **soil organic C** and total **soil N** across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
+database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
+stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
+study.
+
+Next, the [**BMR data**](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) contains the trait contrasts from phylogenetically independent contrasts [@SukumaranH2010] of data reported by [@GenoudIM2017] on the basal metabolic rate (BMR) and mass of mammals species.
+Specifically, the "SELECT" BMR data in their database were used for those species for which data were available.
+A phylogeny in NEXUS [@NEXUS] format was extracted from the supplementary information of [@GenoudIM2017].
+Version 4.4.0 of the DendroPy library [@SukumaranH2010] was used to calculate phylogenetic independent contrasts using Python 2.7.
+For each contrast, the path length spanned by the contrast (sum of the branch lengths) and the contrast for the mass, natural logarithm of the BMR and natural logarithm of the mass were written as a row in the [bmr-contrasts.csv](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) file.
+
+[**Third dataset**](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) was a raw data on measurements of anatomical features of several species of *Anolis* lizards, which was downloaded
+from the Dryad data package [@MahlerIRL2013Data] associated with [@MahlerIRL2013].
+The [contrasts.csv file](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) was produced by using version 4.4.0 of the DendroPy library [@SukumaranH2010]  to calculate phylogenetic independent contrasts on the the data in the GA_Anolis_traits.csv file of Mahler et at.
+using their maximum clade credibility tree (GA_Anolis_MCC.nex) as the phylogeny.
+
+Absolute values of traits for each contrast are considered to construct corresponding copula for second and third datasets.
+
+Fourth, we used **Ceder-creek data for e120** which resulted from a biodiversity experiment, carried out by Tilman and his co-warkers [@Tilman2001; @Tilman2006] over many years to collect [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120) data.
+These datasets were used to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m),
+for year 2000.
+
+<!--The raw data and copula for each four datasets are shown along each column of Figure \ref{MT-fig_biv_raw_cop}.-->
+
+Similarly, we applied multivariate analysis on four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014].
+These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species.
+We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony.
+In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{fig_aphid_map}).
+
+<!--Figure \ref{MT-fig_aphid_raw_cop} represents raw and copula data plot for aphid count ("Green spruce aphid") and aphid-firstflight date ("Leaf-curling plum aphid") taken from two locations.-->
+
+Our third dataset for multivariate analysis is a plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013).
+These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
+The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
+unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
+location in the ocean.
+Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979].
+All the samples available for a given variable in a given area in a given month of a given year were
+averaged, producing a monthly time series for each year.
+Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).
+Monthly values were converted to yearly values by averaging over all months.
+This approach was used by [@Sheppard2017] earlier.
+But here we only allowed maximum 4 months in a year with missing data which were replaced by those medians.
+If a year got 5 or more missing monthly data value then we set NA value for that particular year.
+We used data for species=16 [Name : "Ceratium furca"] which is a dicoflagellate commonly found in UK seas.
+Among 26 locations, only those having atleast 45 years time series data, were considered (see Figure \ref{fig_plankton_map}).
+
+Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
+Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
+Additional information on the history of the site and soil types is detailed in @Jarecke2016.
+Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at
+daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
+additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.
+
+# Nonparametric analysis of tail dependence \label{NPA}
+
+Left (or lower)- and right (or upper)-tail dependence quantify the extent to which two ranked
+variables $u_i$, $v_i$ for $i=1,...,n$ 
+(see \nameref{MT-Model_selection} for defintions of $u_i$ and $v_i$ in the manuscript)
+are related to each other in the left and right tails, respectively, of their 
+distributions [@nelsen2007_copula, @joe2014_dependence]. 
+Our nonparametric study of tail dependence uses three
+statistics which can quantify the extent to which $u_i$ and $v_i$ are 
+related in any part of their distributions. We here describe the 
+statistics. The statistics were formulated with positively associated
+variables in mind. All the variables we study are positively associated
+when they are significantly associated. So definitions and associated 
+diagrams will assume positive associations, though generalizations are 
+possible. Given two bounds 
+$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
+$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
+fall (these are the blue and red lines, respectively, in 
+Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
+between $u_i$ and $v_i$ in the region bounded by these lines.
+Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
+dependence in the left parts of the distributions of $u$ and $v$, and 
+using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
+the right parts of the distributions.
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
+\caption{Schematic diagram supporting defintions of nonparametric tail dependence statistics \label{fig_stat_image}}
+\end{center}
+\end{figure}
+
+The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
+correlation of $u_i$ and $v_i$ that is attributable to the points 
+lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
+\begin{equation}\label{eq.Cor}
+   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
+   (u_i-\mean(u)) (v_i-\mean(v)),
+\end{equation}
+where means and variances are computed using all $n$ data 
+points, but the
+sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
+$u_i+v_i < 2u_b$. 
+Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
+between $u$ and $v$ in the region given by the bounds.
+
+We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
+$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
+$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
+and a distance less than $h$ from the line $v=u$, divided by the 
+total number of points within the bounds. This function is
+defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
+is half the longer of the two segments obtained by intersecting the
+bounds with the unit square. It is easy to see $S(0)=0$ and 
+$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
+of $h$, to be the area within the bounds and the unit square
+and within distance $h$ of the line $v=u$, divided by the area
+within the bounds and the unit square. This is the expected value of
+$S(h)$ for independent data. We then defined
+\begin{equation}\label{eq.P}
+   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
+\end{equation}
+As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
+indicate stronger positive association
+between $u$ and $v$ in the region given by the bounds.
+
+The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
+distance between points satisfying $u_i+v_i>2l_b$ and 
+$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
+$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
+$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
+then small. 
+Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
+dependence in the portions of the distributions of $u$ and $v$ given by 
+the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
+$\Dsq_{l_b,u_b}$ indicate strong dependence.
+
+For large datasets (i.e., datasets with large $n$),
+we can consider $l_b$ and $u_b$ close together without incurring undue
+sampling variation in our statistics, and we can consider 
+several different bands $(l_b,u_b)$ to understand how dependence
+varies in different parts of the distributions. But for datasets with 
+smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
+We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
+and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
+$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.
+
+The difference $\cor_l-\cor_u$ is expected to be positive if 
+dependence in the left halves of the distributions is stronger 
+than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
+This holds, for instance, for the Clayton copula [@nelsen2007_copula].
+The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
+"lower" statistics occur in opposite order, is also expected to be 
+positive under the same conditions. We use these three differences
+to detect asymmetry of tail dependence. Likewise, for large datasets,
+we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
+$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
+$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
+
+We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas (for details see Appendix \ref{Test_npa_stat}).
+
+For bivariate datasets, we compared values 
+of the statistics $\cor_{l_b,u_b}$,
+$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to 
+distributions of values of the same statistics computed on surrogate datasets that were produced from 
+the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
+the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
+correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
+but had Normal copula structure. Thus our comparisons tested the null hypothesis 
+that our statistics took values on the empirical data no different from what would have been expected 
+if the copula structure of the data were Normal, but the data were otherwise 
+statistically unchanged.
+Details of how surrogates were prroduced are in Appendix \ref{surrog_test}, and code is in [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R).
+
+The comparison of one of the
+statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
+empirical data, to the distribution
+of its values computed on many surrogate datasets 
+provides a test of whether dependence in the part of the distribution specified 
+by $l_b$ and $u_b$ is different from
+what would be expected from a null hypothesis of normal copula structure but 
+unchanged marginals and unchanged 
+Kendall (or Spearman)
+correlation. Thus the comparison helps answer the first two parts of question \hyperref[MT-Intro]{\textbf{(Q1)}} 
+from the \nameref{MT-Intro} for bivariate datasets. 
+Significant deviations correspond to deviations from normal 
+copula structure. In particular, deviations 
+using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
+different from that of a normal copula;
+likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
+from that of a normal copula.
+
+The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
+as computed on the 
+empirical data, to the distribution
+of its values on many surrogate datasets 
+provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
+third part of question \hyperref[MT-Intro]{\textbf{(Q1)}} from the \nameref{MT-Intro} for bivariate datasets.
+
+For multivariate datasets, a spatial resampling scheme was used instead of the surrogate 
+approach described above.
+The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
+widely used spline correlogram method; the same scheme was also used by @walteretal2017 
+as part of a method for 
+performing model selection among matrix regression models. For each pair of 
+locations $i \neq j$, the statistic
+of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
+$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
+$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
+values $s_{ij}$ were then arranged in
+a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
+$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
+average strength of 
+dependence, across all pairs of distinct locations, for data in the portion of the distributions 
+given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
+asymmetry of tail dependence across all pairs of locations. 
+Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
+distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
+revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
+or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
+Normal null hypothesis, and answering question \hyperref[MT-Intro]{\textbf{(Q1)}} from the \nameref{MT-Intro}.
+Bootstrapping was performed by resampling the locations of measurement, with replacement. 
+This corresponds to 
+resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
+[@bjornstadfalck2001 ; @walteretal2017].
 
@@
       caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.
-Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\nameref{MT-NPA} in the manuscript).
+Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\ref{NPA}).
 One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates.
@@
 
+# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}
+
+To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
+\ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.
+
+\begin{equation}\label{eq.AR1}
+P_{i}(t+1)=\beta P_{i}(t)+\alpha \epsilon_{i}(t)
+\end{equation}
+
+To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$.
+We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches.
+Should their dependence be exactly the same as of the noises or not ?
+
+To test that our approach is of two way : first, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
+Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome which was a bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
+\ref{fig_Cause4copula_params_GOF_500}).
+We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size and reported p-values for both statistics (Cramer-von Mises and Kolmogorov-Smirnov).
+p-values greater or equal to $0.05$ (above the dashed line in Fig.
+\ref{fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters.
+Fig.
+\ref{fig_Cause4copula_params_GOF_2500} represents similar feature as of Fig.
+\ref{fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$.
+(For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
+
+Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \ref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
+Then we plotted mean of each measured quantity with standard error against Spearman's Rho (or Kendall's Tau) of the noise copula family.
+Fig.
+\ref{fig_Cause4copula_C_spear} and Fig.
+\ref{fig_Cause4copula_C_kend} represent this picture for noise generating from a lower tail dependent Clayton copula whereas for Fig.
+\ref{fig_Cause4copula_SC_spear} and Fig.
+\ref{fig_Cause4copula_SC_kend} noise came from a upper tail dependent Survival Clayton family.
+We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach, here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise.
+(For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
+
 <!--
@@
 
-\newpage
 ```{r tab_stat_npa_asym_2,echo=F,results="markup"}
@@
 
-\newpage
 ```{r tab_stat_npa_asym_3,echo=F,results="markup"}
@@
 
-\newpage
-
 ```{r tab_stat_npa_asym_4,echo=F,results="markup"}
@@
 
-## Need to think about mechanisms for biogeochemical, evolutionary, and community data
+# Cause of non-Normal copula structure : Asymmetric sensitivity and copula structure in populations \label{Cause_asym_sens}
+
+To test the assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;
+
+\begin{equation}\label{eq.AS}
+P_{i}(t+1) = b P_{i}(t) + g(\epsilon_{i}(t)) + h(\delta_{i}(t))
+\end{equation}
+
+$b$ was the linear growth rate.
+Global environmental fluctuation function $g(\epsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq.
+\ref{eq.AS_g}) where $\epsilon=(\epsilon_{1}, \epsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\epsilon}$.
+
+\begin{align}
+g(\epsilon) = \begin{cases} 
+g_{1} = \begin{cases}
+\epsilon & \text{if $\epsilon< 0$}; \\
+0 & \text{otherwise}.
+\end{cases}\\ \\
+g_{2} = \begin{cases}
+\epsilon & \text{if $\epsilon> 0$}; \\
+0 & \text{otherwise}\label{eq.AS_g}.
+\end{cases}
+\end{cases} \text{where } \epsilon = N(\textbf{0},\Sigma_{\epsilon})\;\ \text{with} \;\ \Sigma_{\epsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
+\end{align}
+
+Local environmental fluctuation function $h(\delta)$ varied according to Eq.
+\ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$.
+$\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$.
+
+\begin{equation}\label{eq.AS_h}
+h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
+\end{equation}
+
+We simulated our model (\ref{eq.AS}) for different values of $b$ with $a=0.2$, $r=0.8$.
+$10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ time step of simulated points to plot the population-copula for a specific $b$ and a specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
+Fig.
+\ref{fig_ms_asym_b_0.1} and Fig.
+\ref{fig_ms_asym_b_0.9} show these results for $b=\pm0.1$ and $b=\pm0.9$, respectively.
+We also carried out model selection and non-parametric statistical analysis with each copula of those figures and summarized our results in Tables \ref{tab_summary_stat_asym_sens} - \ref{tab_stat_npa_asym_8}.
+Table \ref{tab_summary_stat_asym_sens} represent the model selection results with AIC weights for best fitted copula and normal copula, model averaged lower and upper tail dependence, p-values (both Cramer-von Mises statistic and Kolmogorov-Smirnov statistic) for goodness of fit test.
+In each case, successful bootstrap was $100\%$.
+According to our model selection approach, for $b=\pm0.1$, non-normal copula (with higher AIC weights) were always best-fit with some asymetric tail dependence whereas for $b\pm0.9$, most of the time normal was best.
+It is probably because the populations get long-time exposure to environmental changes for slower growth(or decay) rate and so global fluctuation can affect them more efficiently.
+Our non-parametric analysis results (Tables \ref{tab_stat_npa_asym_1} -\ref{tab_stat_npa_asym_8}) are consistent with this model selection approach.
+
+#  Need to think about mechanisms for biogeochemical, evolutionary, and community data
 
@@
 \end{figure}
-
-<!--
-## Extinction risk
-        ----moved to Paper.Rmd----
--->
-
-```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
-
-source("mydispersal.R")
-set.seed(seed)
-
-rlist<-c(0,0.05,0.25,0.5,0.7,0.75,0.9,1.0,1.25,1.5,1.75,1.95)
-
-# plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-pdf("./Results/ext_risk_local_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
-for(r in rlist){
-    #cat("r=",r,"\n")
-    s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
-    mtext(paste0("r = ", r))
-}
-par(op)
-dev.off()
-
-# plot for global dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-pdf("./Results/ext_risk_global_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
-for(r in rlist){
-  #cat("r=",r,"\n")
-  s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
-  plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-  lines(s$d_seq,s$risk_right,type="b",col="blue")
-  legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
-  mtext(paste0("r = ", r))
-}
-par(op)
-dev.off()
-
-# plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
-pdf("./Results/ext_risk_numlocs_effect_local_disp.pdf",heigh=6,width=6)
-op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
-for(r in c(0,0.25)){
-  #cat("r=",r,"\n")
-  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
-  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=F,ploton=F)
-
-  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
-  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
-       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
-  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
-  #mtext(paste0("B",1))
-}
-par(op)
-dev.off()
-
-# plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
-pdf("./Results/ext_risk_numlocs_effect_global_disp.pdf",heigh=6,width=6)
-op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
-for(r in c(0,0.25)){
-  #cat("r=",r,"\n")
-  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
-  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=T,ploton=F)
-
-  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
-  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
-       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
-  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
-  #mtext(paste0("B",1))
-}
-par(op)
-dev.off()
-
-# plot to see the effect of K for r=0.25 (Ricker model)
-pdf("./Results/ext_risk_K_effect.pdf",heigh=6,width=6)
-op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
-
-s1<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=F,ploton=F)
-s2<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=F,ploton=F)
-
-s3<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=T,ploton=F)
-s4<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=T,ploton=F)
-
-plot(s1$d_seq,s1$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-lines(s2$d_seq,s2$risk_left,type="b",col="orange")
-legend("topleft",title=("Noise : left tail dep., local disp."), c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-plot(s1$d_seq,s1$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-lines(s2$d_seq,s2$risk_right,type="b",col="skyblue")
-legend("topleft",title=("Noise : right tail dep., local disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
-
-plot(s3$d_seq,s3$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-lines(s4$d_seq,s4$risk_left,type="b",col="orange")
-legend("topleft",title= ("Noise : left tail dep., global disp."),c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-plot(s3$d_seq,s3$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-lines(s4$d_seq,s4$risk_right,type="b",col="skyblue")
-legend("topleft",title=("Noise : right tail dep., global disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
-
-par(op)
-dev.off()
-```
-
-![Extinction_Risk_local_dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp}](./Results/ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Extinction_Risk_global_dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp}](./Results/ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Extinction_Risk_numlocs_effect_local_disp; numsteps=25, numsims=10000\label{fig_risk_numlocs_effect_local_disp}](./Results/ext_risk_numlocs_effect_local_disp.pdf){ width=40%, height=40% }
-
-![Extinction_Risk_numlocs_effect_global_disp; numsteps=25, numsims=10000\label{fig_risk_numlocs_effect_global_disp}](./Results/ext_risk_numlocs_effect_global_disp.pdf){ width=40%, height=40% }
-
-![Extinction_Risk_K_effect on Ricker model; r=0.25, numsteps=25, numsims=10000\label{fig_risk_K_effect}](./Results/ext_risk_K_effect.pdf){ width=40%, height=40% }
-
-\newpage
 
@@
 
+\begin{figure}[!h]
+\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.1 cm} (D)} \\
+\vspace{-0.5 cm} 
+\begin{center}
+\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC.pdf}\\
+\end{center}
+\textbf{ \hspace{3.1 cm} (E) \hspace{2.4 cm} (F) \hspace{2.3 cm} (G) \hspace{2.3 cm} (H) \hspace{2.3 cm} (I)}\\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_5stats.pdf}
+\caption[Spatial Taylor's Law results]{Spatial Taylor's Law results: log(spatial variance) vs.
+log(spatial mean) for $ns=10$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of hhomoskedasticity, (G) root mean square error, (H) intercept, (I) slope of the first panel plots.
+\label{fig_spTL}}
+\end{center}
+\end{figure}
+
+<!--
+## Extinction risk
+        ----moved to Paper.Rmd----
+-->
+
+```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+
+source("mydispersal.R")
+set.seed(seed)
+
+rlist<-c(0,0.05,0.25,0.5,0.7,0.75,0.9,1.0,1.25,1.5,1.75,1.95)
+
+# plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
+pdf("./Results/ext_risk_local_disp.pdf",heigh=6,width=10)
+op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
+for(r in rlist){
+    #cat("r=",r,"\n")
+    s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
+    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s$d_seq,s$risk_right,type="b",col="blue")
+    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+# plot for global dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
+pdf("./Results/ext_risk_global_disp.pdf",heigh=6,width=10)
+op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
+for(r in rlist){
+  #cat("r=",r,"\n")
+  s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
+  plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+  lines(s$d_seq,s$risk_right,type="b",col="blue")
+  legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+  mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+# plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
+pdf("./Results/ext_risk_numlocs_effect_local_disp.pdf",heigh=6,width=6)
+op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
+for(r in c(0,0.25)){
+  #cat("r=",r,"\n")
+  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
+  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=F,ploton=F)
+
+  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
+  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
+       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
+  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+  #mtext(paste0("B",1))
+}
+par(op)
+dev.off()
+
+# plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
+pdf("./Results/ext_risk_numlocs_effect_global_disp.pdf",heigh=6,width=6)
+op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
+for(r in c(0,0.25)){
+  #cat("r=",r,"\n")
+  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
+  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=T,ploton=F)
+
+  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
+  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
+       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
+  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+  #mtext(paste0("B",1))
+}
+par(op)
+dev.off()
+
+# plot to see the effect of K for r=0.25 (Ricker model)
+pdf("./Results/ext_risk_K_effect.pdf",heigh=6,width=6)
+op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
+
+s1<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=F,ploton=F)
+s2<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=F,ploton=F)
+
+s3<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=T,ploton=F)
+s4<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=T,ploton=F)
+
+plot(s1$d_seq,s1$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+lines(s2$d_seq,s2$risk_left,type="b",col="orange")
+legend("topleft",title=("Noise : left tail dep., local disp."), c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+plot(s1$d_seq,s1$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+lines(s2$d_seq,s2$risk_right,type="b",col="skyblue")
+legend("topleft",title=("Noise : right tail dep., local disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+
+plot(s3$d_seq,s3$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+lines(s4$d_seq,s4$risk_left,type="b",col="orange")
+legend("topleft",title= ("Noise : left tail dep., global disp."),c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+plot(s3$d_seq,s3$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+lines(s4$d_seq,s4$risk_right,type="b",col="skyblue")
+legend("topleft",title=("Noise : right tail dep., global disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+
+par(op)
+dev.off()
+```
+
+![Extinction_Risk_local_dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp}](./Results/ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Extinction_Risk_global_dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp}](./Results/ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Extinction_Risk_numlocs_effect_local_disp; numsteps=25, numsims=10000\label{fig_risk_numlocs_effect_local_disp}](./Results/ext_risk_numlocs_effect_local_disp.pdf){ width=40%, height=40% }
+
+![Extinction_Risk_numlocs_effect_global_disp; numsteps=25, numsims=10000\label{fig_risk_numlocs_effect_global_disp}](./Results/ext_risk_numlocs_effect_global_disp.pdf){ width=40%, height=40% }
+
+![Extinction_Risk_K_effect on Ricker model; r=0.25, numsteps=25, numsims=10000\label{fig_risk_K_effect}](./Results/ext_risk_K_effect.pdf){ width=40%, height=40% }
+
+# Consequence of non-Normal copula structure : effect on skewness of the spatial mean \label{Consequence_skewness}
+
+Because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, changed significantly from the individual skewness of time-series data.
+Spatial mean of the time series data extracted that tail dependence from the multivariate data set, if any, and reflected in the averaged skewness.
+We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
+So, first we generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
+Then we calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R).
+We made a histogram of surragate skewness and found if actual skewness were within $95\%$ quantile or not.
+If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).
+
+To verify that we used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data.
+Figure \ref{fig_skewness_count} showed that averaged skewness was much lower than expected if the aphid count data came from a normal copula.
+In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_count}).
+As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results.
+On the other hand, Figure \ref{fig_skewness_ff} demonstrated that averaged skewness was much higher than expected if the aphid first flight data came from a normal copula.
+As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red line in Figure \ref{fig_skewness_ff}) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence.
+Figure \ref{fig_skewness_pns} showed that averaged skewness was lower than expected if the plankton data came from a normal copula.
+Our previous results on non-parametric analysis and model fitting also supported lower tail dependence.
+But, as $p>0.05$, we could not reject $H_{0}$ in that case.
+Finally, for methane-flux data, the averaged skewness was lower than expected if the data came from a normal copula (Figure \ref{fig_skewness_methane}).
+As $p<0.05$, we rejected $H_{0}$ and that result was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence.
+
+# Consequence of non-Normal copula structure : Effect on Taylor's law \label{Consequence_TL}
+
+Spatial Taylor's law, which shows linear relationship between `log`(variance) and `log`(mean) of population densities of a species from different locations, is a well-known pattern found in ecology and related fields [@Taylor1961, @Taylor1988, @Cohen2013, @Doring2015, @Kuo2016, @Xu2016, @Reuman2017].
+Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula.
+First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with covariance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of timeseries, number of sampling locations and number of replica for that copula, respectively.
+Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal.
+With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D).
+We also tested statistically that whether linearity, homoskedasticity is maintained in each copula dataset or not and plotted our results alongwith root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).
+
+# Consequence of non-Normal copula structure : effect on extinction risk \label{Consequence_ext_risk}
+
+In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
+<!--did I write correct spelling for lewton model?-->
+
+\begin{equation}\label{eq.extrisk_model}
+P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
+\end{equation}
+
+$r$ is the liner growth rate and $K$ is the carrying capacity of the environment.
+$\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence.
+We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion).
+We simulated Eq.
+\ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $25$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;
+
+- LC model (r=0) :
+
+    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
+[see first subplot of Fig.
+\ref{fig_risk_local_disp} and Fig.
+\ref{fig_risk_global_disp}]
+
+- Ricker model (r $\neq$ 0):
+
+    * For ricker model when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
+For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
+[see Fig.
+\ref{fig_risk_local_disp} - \ref{fig_risk_global_disp}]
+
+    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]
+
+With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig.
+\ref{fig_risk_numlocs_effect_local_disp} - Fig.
+\ref{fig_risk_numlocs_effect_global_disp}] for both model.
+
 \newpage

### 2018-09-11 SupportingInformation.Rmd
modified

@@
 #cat(paste("start-time: ",Sys.time(),"\n")) 
-RES_aphid_count<-RES_single_sp(d_allsp=data_aphid_count,sp=10,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
+RES_aphid_count<-RES_single_sp(d_allsp=data_aphid_count,sp=10,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
+                               good_loc_given=F,good_loc=NA)
 #cat(paste("stop-time: ",Sys.time(),"\n"))
@@
 #cat(paste("start-time: ",Sys.time(),"\n"))
-RES_aphid_ff<-RES_single_sp(d_allsp=data_aphid_ff,sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30)
+RES_aphid_ff<-RES_single_sp(d_allsp=data_aphid_ff,sp=11,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
+                            good_loc_given=F,good_loc=NA)
 #cat(paste("stop-time: ",Sys.time(),"\n")) 
@@
 #cat(paste("start-time: ",Sys.time(),"\n")) 
-RES_plankton_north_sea<-RES_single_sp(d_allsp=data_plankton_north_sea,sp=16,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=45)
+RES_plankton_north_sea<-RES_single_sp(d_allsp=data_plankton_north_sea,sp=16,families=c(1,3:10,13,14,16:20),level=0.05,
+                                      data_pt_thrs=45,good_loc_given=F,good_loc=NA)
 #cat(paste("stop-time: ",Sys.time(),"\n"))    
@@
 #cat(paste("start-time: ",Sys.time(),"\n")) 
-RES_methane<-RES_single_sp(d_allsp=methane_data,sp=1,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=50)
+RES_methane<-RES_single_sp(d_allsp=methane_data,sp=1,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=50,
+                           good_loc_given=F,good_loc=NA)
 #cat(paste("stop-time: ",Sys.time(),"\n"))    

### 2018-09-14 Paper.Rmd
modified@Sept14

@@
 Here, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets.
-In addition to a copula based \nameref{Model_selection}, we also carried out a \ref{SM-NPA} with three different statistics to verify our model selection based results.
+In addition to a copula based \nameref{Model_selection}, we also carried out a \nameref{NPA} with three different statistics to verify our model selection based results.
 
@@
 
+Considered datasets were ranged widely from population ecology, community ecology , evolutionery and biogeochemical sphere. 
 We used four datasets in each category of bivariate (soil C and N data, BMR data, Anolis data, Ceder-creek experiment data) and multivariate (aphid-abundance data, aphid-firstflight data, plankton-abundance data and methane-flux data) examples.
-Details of each datasets used were given in \ref{SM-Data_detailed}.
+Details of each dataset used were given in \ref{SM-Data_detailed}.
 
@@
 
-<!--### Nonparametric analysis of tail dependence \label{NPA}-->
+### Nonparametric analysis of tail dependence \label{NPA}
 
-In addition to the model selection approach described above, **nonparametric approaches** (for details see Appendix \ref{SM-NPA}) to quantify lower ($\cor_l$, $\Ps_l$ and $\Dsq_l$ statistics) and upper tail dependence ($\cor_u$, $\Ps_u$ and $\Dsq_u$ statistics) were also used to answer \hyperref[Intro]{\textbf{(Q1)}}.
+In addition to the model selection approach described above, **nonparametric approaches** to quantify lower ($\cor_l$, $\Ps_l$ and $\Dsq_l$ statistics) and upper tail dependence ($\cor_u$, $\Ps_u$ and $\Dsq_u$ statistics) were also used to answer \hyperref[Intro]{\textbf{(Q1)}}.
 Using two distinct methods, when results are consistent, provides greater confidence
@@
 
-<!--***DAN: There is either too little or too much here in this 
-paragraph. The reader cannot get the main
-idea, really, of what was done for these tests, from this paragraph alone - they will have to
-look at the appendix. They need to be able to get the main idea. So you could either add detail,
-or else take away detail and just say something like "We tested our statistics by testing whether 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed asymmetry of tail dependence 
-known to be present in data simulated from various copulas (Appendix X)" and then stop there.
-Then you would remove panels B-D from Fig. 1. At the moment I prefer this second option, but 
-my preference is not strong, so if you prefer to keep the material and add detail that would be 
-fine.-->
-<!--**SG: Agreed with 2nd option and did things as necessary-->
+Left (or lower)- and right (or upper)-tail dependence quantify the extent to which two ranked
+variables $u_i$, $v_i$ for $i=1,...,n$ 
+(see \nameref{Model_selection} for defintions of $u_i$ and $v_i$)
+are related to each other in the left and right tails, respectively, of their 
+distributions [@nelsen2007_copula, @joe2014_dependence]. 
+Our nonparametric study of tail dependence uses three
+statistics which can quantify the extent to which $u_i$ and $v_i$ are 
+related in any part of their distributions. We here describe the 
+statistics. The statistics were formulated with positively associated
+variables in mind. All the variables we study are positively associated
+when they are significantly associated. So definitions and associated 
+diagrams will assume positive associations, though generalizations are 
+possible. Given two bounds 
+$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
+$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
+fall (these are the blue and red lines, respectively, in 
+Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
+between $u_i$ and $v_i$ in the region bounded by these lines.
+Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
+dependence in the left parts of the distributions of $u$ and $v$, and 
+using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
+the right parts of the distributions.
 
-<!--***DAN: I think the below material belongs in supp mat, because the reader can get the
-main idea via the above text, and only some readers will be interested in the detailed 
-algorithm. Please move it there, and then make the above links XYZ and ABC into live links
-to the appropriate locations in the sup mat.
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
+\caption{Schematic diagram supporting defintions of nonparametric tail dependence statistics \label{fig_stat_image}}
+\end{center}
+\end{figure}
 
-For bivariate datasets, surrogates were produced as follows.
+The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
+correlation of $u_i$ and $v_i$ that is attributable to the points 
+lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
+\begin{equation}\label{eq.Cor}
+   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
+   (u_i-\mean(u)) (v_i-\mean(v)),
+\end{equation}
+where means and variances are computed using all $n$ data 
+points, but the
+sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
+$u_i+v_i < 2u_b$. 
+Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
+between $u$ and $v$ in the region given by the bounds.
 
-- Given data $(x_i,y_i)$ for $i=1,\ldots,n$ and a 
-one-parameter target copula family $\mathcal{B}(\theta)$ (here $\theta$ denotes the parameter), let $\tau$ be the
-Kendall correlation of the $x_i$ with the $y_i$ and find the parameter value $\theta_\tau$ for which 
-$\mathcal{B}(\theta_\tau)$ has Kendall correlation $\tau$ (this is possible for the one-parameter copula families of this 
-study using the `iTau` function of the `VineCopula` package). 
-- Generate data $(a_i,b_i)$, $i=1,\ldots,n$ from 
-$\mathcal{B}(\theta_\tau)$. 
-- Permute the ordered set $(x_1,\ldots,x_n)$ such that the permutation
-$(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ has $\rank(x_{\sigma_x(i)})$ equal to $\rank(a_i)$ for all $i$, where 
-$\rank(x_{\sigma_x(i)})$ is the rank of $x_{\sigma_x(i)}$ in the set $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$, and
-$\rank(a_i)$ is the rank of $a_i$ in the set $(a_1,\ldots,a_n)$ (here $\sigma_x$ is a permutation of the 
-indices $1,\ldots,n$). 
-- Likewise permute $(y_1,\ldots,y_n)$
-such that the permutation
-$(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ has $\rank(y_{\sigma_y(i)})$ equal to $\rank(b_i)$ for all $i$ ($\sigma_y$ is 
-another permutation of $1,\ldots,n$).
-- The surrogate dataset is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.
+We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
+$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
+$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
+and a distance less than $h$ from the line $v=u$, divided by the 
+total number of points within the bounds. This function is
+defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
+is half the longer of the two segments obtained by intersecting the
+bounds with the unit square. It is easy to see $S(0)=0$ and 
+$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
+of $h$, to be the area within the bounds and the unit square
+and within distance $h$ of the line $v=u$, divided by the area
+within the bounds and the unit square. This is the expected value of
+$S(h)$ for independent data. We then defined
+\begin{equation}\label{eq.P}
+   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
+\end{equation}
+As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
+indicate stronger positive association
+between $u$ and $v$ in the region given by the bounds.
 
-Our code for this algorithm is [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R) in the 
-[`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
+The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
+distance between points satisfying $u_i+v_i>2l_b$ and 
+$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
+$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
+$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
+then small. 
+Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
+dependence in the portions of the distributions of $u$ and $v$ given by 
+the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
+$\Dsq_{l_b,u_b}$ indicate strong dependence.
 
-Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
-is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a 
-permutation 
-of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as 
-those of the original data. 
-Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
-Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
-will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
-a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
-as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
-sampling variation) the Spearman correlation of the empirical data.  
--->
-<!--**SG: Done-->
+For large datasets (i.e., datasets with large $n$),
+we can consider $l_b$ and $u_b$ close together without incurring undue
+sampling variation in our statistics, and we can consider 
+several different bands $(l_b,u_b)$ to understand how dependence
+varies in different parts of the distributions. But for datasets with 
+smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
+We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
+and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
+$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.
+
+The difference $\cor_l-\cor_u$ is expected to be positive if 
+dependence in the left halves of the distributions is stronger 
+than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
+This holds, for instance, for the Clayton copula [@nelsen2007_copula].
+The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
+"lower" statistics occur in opposite order, is also expected to be 
+positive under the same conditions. We use these three differences
+to detect asymmetry of tail dependence. Likewise, for large datasets,
+we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
+$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
+$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
+
+We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas (for details see Appendix \ref{SM-Test_npa_stat}).
+
+For bivariate datasets, we compared values 
+of the statistics $\cor_{l_b,u_b}$,
+$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to 
+distributions of values of the same statistics computed on surrogate datasets that were produced from 
+the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
+the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
+correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
+but had Normal copula structure. Thus our comparisons tested the null hypothesis 
+that our statistics took values on the empirical data no different from what would have been expected 
+if the copula structure of the data were Normal, but the data were otherwise 
+statistically unchanged.
+Details of how surrogates were produced are in Appendix \ref{SM-surrog_test}, and code is in [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R).
+
+The comparison of one of the
+statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
+empirical data, to the distribution
+of its values computed on many surrogate datasets 
+provides a test of whether dependence in the part of the distribution specified 
+by $l_b$ and $u_b$ is different from
+what would be expected from a null hypothesis of normal copula structure but 
+unchanged marginals and unchanged 
+Kendall (or Spearman)
+correlation. Thus the comparison helps answer the first two parts of question \hyperref[Intro]{\textbf{(Q1)}} 
+from the \nameref{Intro} for bivariate datasets. 
+Significant deviations correspond to deviations from normal 
+copula structure. In particular, deviations 
+using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
+different from that of a normal copula;
+likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
+from that of a normal copula.
+
+The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
+as computed on the 
+empirical data, to the distribution
+of its values on many surrogate datasets 
+provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
+third part of question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} for bivariate datasets.
+
+For multivariate datasets, a spatial resampling scheme was used instead of the surrogate 
+approach described above.
+The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
+widely used spline correlogram method; the same scheme was also used by @walteretal2017 
+as part of a method for 
+performing model selection among matrix regression models. For each pair of 
+locations $i \neq j$, the statistic
+of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
+$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
+$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
+values $s_{ij}$ were then arranged in
+a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
+$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
+average strength of 
+dependence, across all pairs of distinct locations, for data in the portion of the distributions 
+given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
+asymmetry of tail dependence across all pairs of locations. 
+Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
+distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
+revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
+or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
+Normal null hypothesis, and answering question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
+Bootstrapping was performed by resampling the locations of measurement, with replacement. 
+This corresponds to 
+resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
+[@bjornstadfalck2001 ; @walteretal2017].
 
@@
 
-As described in Appendix \ref{SM-NPA}, each of our statistics, as computed for the 
+As described in \nameref{NPA}, each of our statistics, as computed for the 
 ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
@@
 computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
-resampling scheme, as described in \ref{SM-NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
+resampling scheme, as described in \nameref{NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
 $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
@@
 
+```{r read_RES_temp_aphid, echo=F}
+RES_temp_aphid<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_winter_neg_temp.RDS")
+```
+
 This section is dedicated to explore question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}.
@@
 This assumption was inspired by the the idea of Moran-effect [@Moran1953] which was originally used to explain synchronization of widely dispersed populations.
-We tested our hypothesis with a two habitat patch model (1st order AR method) in noisy environment and described our findings in Appendix \ref{SM-Cause_Moran}.
-<!--
-### Asymmetric sensitivity and copula structure in populations
-Another model, distinct from the above, Dan and Shya to discuss-->
+We tested our hypothesis with a two habitat patch model (1st order AR method) in noisy environment (having non-Normal copula structure) and found that populations were shaped by similar copula structure as that of local noises ( for details see Appendix \ref{SM-Cause_Moran}).
+We also found the Moran effect on leaf-plum curling aphid's firstflight data induced by average winter temperature for the period of 56 years (1958-2013).
+We did pair wise model selection with negative average winter temperature (averaged over (n-1)$^{th}$ December to n$^{th}$ March temperature) from the selected 10 locations for which we did model selection with aphid-firstflight data.
+Negative winter temperature was considered for simulation to compare the aphid's firstflight results with temparature copula (aphid's firstflight date were negatively correlated with temperature or in other words positively correlated with negative temparature).
+Model selection results for negative average winter temperature in 10 $\times$ 10 matrices were represented in Table \ref{SM-tab_temp_aphid_bestcop} - Table \ref{SM-tab_temp_aphid_LTmUT}.
+In all cases `r min((RES_temp_aphid$gfc_numBS_success/RES_temp_aphid$gfc_numBS),na.rm=T)*100` $\%$ of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+Table \ref{SM-tab_temp_aphid_LTmUT} showed `r sum(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw <0, na.rm=T)` out of `r (dim(RES_temp_aphid$gfc_p_CvM)[1]^2-dim(RES_temp_aphid$gfc_p_CvM)[1])-2*RES_temp_aphid$num_indep_loc_pair` location pair showed upper tail dependence in temperature copula, which can be considered as the Moran effect results causing dominant upper tail dependence in the copulas with aphid's firstflight date (Table \ref{SM-tab_ff_LTmUT}).
+
 On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold.
@@
 This section is attributed to analyze of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}.
-We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data : we tested with four multivariate datasets and found results consistent with model selection and nonparametric approach (For details see Appendix \ref{SM-Consequence_skewness}).
+We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data.
+Spatial mean of the time series data extracted that tail dependence from the multivariate data set, if any, and reflected in the averaged skewness.
+We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula (for details see Appendix \ref{SM-Consequence_skewness}).
+If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).
+Figure \ref{fig_skewness_methane_maintext} showed asymmetric copula structure's effect on spatial skewness for biogeochemical methane-flux data.
+The result, showing significantly different skewness of spatial mean from that of Normal surrogates, was consistent with model selection and nonparametric approach which supported lower tail dependence copula for methane-flux data.
+
 We then showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}).
 Finally, we studied **(iii)** the effects of non-Normal copula on extinction risk with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}).
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/skewness_results/skewness_methane/skewness_methane.pdf}
+\caption[Skewness-methane-flux data]{Skewness of spatially averaged methane-flux data (red line) was compared with respect to that of 10000 Normal surrogates preserving $(A)$ Kendall correlation, and $(B)$ Spearman correlation.
+p-value < 0.05 in both case showed significant difference in skewness (as expected for lower tail dependent methane flux data from model selection results) from Normal surrogates.
+\label{fig_skewness_methane_maintext}}
+\end{center}
+\end{figure}
 

### 2018-09-14 SupportingInformation.Rmd
modified@Sept14

@@
 
+updated on `r Sys.Date()`
+
 \tableofcontents
@@
 
-updated on `r Sys.Date()`
 \pagebreak
@@
 
-# Nonparametric analysis of tail dependence \label{NPA}
+<!--# Nonparametric analysis of tail dependence \label{NPA}
 
@@
 resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
-[@bjornstadfalck2001 ; @walteretal2017].
+[@bjornstadfalck2001 ; @walteretal2017].-->
 
@@
       caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.
-Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\ref{NPA}).
+Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\nameref{MT-NPA}).
 One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates.
@@
 
-\newpage
 <!--Display model selection results for contrasts data in a table-->
@@
 
+\newpage
 <!--Display model selection results for Ceder-creek data for year 2000 in a table-->
@@
 #d0_flightduration<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))
+
+# Time series analysis aphids winter avg temp (here taken as -ve)
+temp_aphid_raw<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidmonthlytempdata.csv",header = F))
+
+temp_aphid<-vector("list",dim(temp_aphid_raw)[1])
+names(temp_aphid)<-paste("loc",c(1:dim(temp_aphid_raw)[1]),sep="")
+
+for(il in 1:length(temp_aphid)){
+
+  s<-matrix(temp_aphid_raw[il,],36,12,byrow = T)
+  rownames(s)<-c(1975:2010)
+  colnames(s)<-c("jan","feb","mar","apr","may","jun","jul","aug","sep","oct","nov","dec")
+
+  temp_aphid[[il]]<-s
+}
+
+winter_avg_temp<-matrix(NA,11,35)
+rownames(winter_avg_temp)<-paste("loc",c(1:dim(temp_aphid_raw)[1]),sep="")
+colnames(winter_avg_temp)<-c(1976:2010)
+
+for(il in 1:length(temp_aphid)){
+  temploc<-temp_aphid[[il]]
+  for(i in 2:dim(temploc)[1]){
+    t<-c(temploc[i-1,12],temploc[i,1:3]) # previous yr dec to present year march averaging temp.
+    winter_avg_temp[il,i-1]<-mean(t,na.rm = T)
+  }
+}
+
+aphid_winter_neg_temp<-vector("list",11) #initializing
+names(aphid_winter_neg_temp)<-rownames(winter_avg_temp)
+
+for(i in 1:11){
+  aphid_winter_neg_temp[[i]]<-data.frame(Year=c(1976:2010),Dat=-winter_avg_temp[i,])
+}
+
+temp_aphid<-list(aphid_winter_neg_temp)
+names(temp_aphid)<-"aphid_winter_neg_temp"
 ```
@@
 
+\newpage
+
 ```{r tab_ff_AICw_normal, echo=F, results='markup'}
@@
 
-\newpage
 ```{r tab_ff_p_CvM, echo=F, results='markup'}
@@
 
+```{r moran_effect_on_aphid_ff, echo=F, results="hide", cache=T, cache.extra=list(seed,temp_aphid,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+set.seed(seed)
+source("FittingCopula_selective_loc.R")  # top most hierarchy source file
+
+#cat(paste("start-time: ",Sys.time(),"\n"))
+RES_temp_aphid<-RES_single_sp(d_allsp=temp_aphid,sp=1,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
+                              good_loc_given=T,good_loc=c(2:11))
+#cat(paste("stop-time: ",Sys.time(),"\n")) 
+resloc<-"./Results/fitting_results/"
+saveRDS(RES_temp_aphid,paste(resloc,file="AphidCopulaFit_selecloc_winter_neg_temp.RDS",sep=""))
+
+LTmUT_temp<-(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw)
+
+sum(LTmUT_temp>0,na.rm = T)
+sum(LTmUT_temp<0,na.rm = T) #66 out of 90 : suggests UT in leaf-plum curling aphids firstflight data (90 out of 90 UT)
+```
+
+```{r tab_temp_aphid_bestcop, echo=F, results="markup"}
+knitr::kable(RES_temp_aphid$info_ord_copname[,,1], #longtable = T,
+             format='pandoc',
+             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **negative average winter temperature data (1958-2013)**.
+Copulas' abbreviations were given in \\nameref{MT-Methods}.
+NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.
+\\label{tab_temp_aphid_bestcop}",
+             booktabs=TRUE
+             )
+```
+
+\newpage
+
+```{r tab_temp_aphid_AICw, echo=F, results='markup'}
+knitr::kable(RES_temp_aphid$info_ord_AICw[,,1], 
+             format='pandoc',
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_AICw}",
+             booktabs=TRUE
+             )
+```
+
+\newpage
+
+```{r tab_temp_aphid_AICw_normal, echo=F, results='markup'}
+source("normal_cop_AICw.R")
+knitr::kable(normal_cop_AICw_matrix(r=RES_temp_aphid),
+             format='pandoc',
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **negative average winter temperature data**, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_temp_aphid_AICw_normal}",
+             booktabs=TRUE
+             )
+```
+
+\newpage
+```{r tab_temp_aphid_p_CvM, echo=F, results='markup'}
+knitr::kable(RES_temp_aphid$gfc_p_CvM,
+             format='pandoc',
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**.
+Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_CvM}",
+             booktabs=TRUE
+             )
+```
+
+\newpage
+```{r tab_temp_aphid_p_KS, echo=F, results='markup'}
+knitr::kable(RES_temp_aphid$gfc_p_KS,
+             format='pandoc',
+             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**.
+Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_KS}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_temp_aphid_LTdep_AICw, echo=F, results='markup'}
+knitr::kable(RES_temp_aphid$LTdep_AICw,
+             format='pandoc',
+             caption = "Model-averaged lower tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTdep_AICw}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_temp_aphid_UTdep_AICw, echo=F, results='markup'}
+knitr::kable(RES_temp_aphid$UTdep_AICw,
+             format='pandoc',
+             caption = "Model-averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_UTdep_AICw}",
+             booktabs=TRUE
+             )
+```
+
+```{r tab_temp_aphid_LTmUT, echo=F, results='markup'}
+knitr::kable(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw,
+             digits = 5,
+             format='pandoc',
+             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTmUT}",
+             booktabs=TRUE
+             )
+```
+
 # Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}
@@
 
-Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \ref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
+Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \nameref{MT-NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
 Then we plotted mean of each measured quantity with standard error against Spearman's Rho (or Kendall's Tau) of the noise copula family.
@@
 
-pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf")
-par(mfrow=c(2,1))
+pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf",height=6,width=16)
+par(mfrow=c(1,2))
 numsurrog<-10000
@@
 
-mtext(paste0("(A) result based on kendall preserving normal surrogates, p = ",p_left),side=3,line=0.2,col="navyblue")
+mtext(paste0("(A) p = ",p_left," (for Kendall preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
 
@@
 
-mtext(paste0("(B) result based on spearman preserving normal surrogates , p = ",p_left),side=3,line=0.2,col="navyblue")
+mtext(paste0("(B) p = ",p_left," (for Spearman preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
 
@@
 \begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf}
+\includegraphics[width=16 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf}
 \caption[Skewness count for species Green spruce aphid]{Skewness count for species Green spruce aphid\label{fig_skewness_count}}
@@
 
-pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf")
-par(mfrow=c(2,1))
+pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf",height=6,width=16)
+par(mfrow=c(1,2))
 numsurrog<-10000
@@
 p_right<-sum(answer2k$surrogskw > answer2k$realskw)/numsurrog
-mtext(paste0("(A) result based on kendall preserving normal surrogates, p = ",p_right),side=3,line=0.2,col="navyblue")
+mtext(paste0("(A) p = ",p_right," (for Kendall preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
 
@@
 p_right<-sum(answer2s$surrogskw > answer2s$realskw)/numsurrog
-mtext(paste0("(B) result based on spearman preserving normal surrogates , p = ",p_right),side=3,line=0.2,col="navyblue")
+mtext(paste0("(B) p = ",p_right," (for Spearman preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
 
@@
 \begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf}
+\includegraphics[width=16 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf}
 \caption[Skewness firstflight for species Leaf-curling plum aphid]{Skewness firstflight for species Leaf-curling plum aphid\label{fig_skewness_ff}}
@@
 
-pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf")
-par(mfrow=c(2,1))
+pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf",height=6,width=16)
+par(mfrow=c(1,2))
 numsurrog<-10000
@@
 \begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf}
+\includegraphics[width=16 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf}
 \caption[Skewness-plankton for species Ceratium furca]{Skewness-plankton for species Ceratium furca\label{fig_skewness_pns}}
@@
 
-pdf("./Results/skewness_results/skewness_methane/skewness_methane.pdf")
-par(mfrow=c(2,1))
+pdf("./Results/skewness_results/skewness_methane/skewness_methane.pdf",height=6,width=16)
+par(mfrow=c(1,2))
 numsurrog<-10000
@@
 p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
-mtext(paste0("(A) result based on kendall preserving normal surrogates, p = ",p_left),side=3,line=0.2,col="navyblue")
+mtext(paste0("(A) p = ",p_left," (for Kendall preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
 
@@
 p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog
-mtext(paste0("(B) result based on spearman preserving normal surrogates , p = ",p_left),side=3,line=0.2,col="navyblue")
+mtext(paste0("(B) p = ",p_left," (for Spearman preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
 
@@
 \begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane.pdf}
+\includegraphics[width=16 cm]{./Results/skewness_results/skewness_methane/skewness_methane.pdf}
 \caption[Skewness-methane-flux data]{Skewness-methane-flux data\label{fig_skewness_methane}}
@@
 \caption[Spatial Taylor's Law results]{Spatial Taylor's Law results: log(spatial variance) vs.
-log(spatial mean) for $ns=10$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of hhomoskedasticity, (G) root mean square error, (H) intercept, (I) slope of the first panel plots.
+log(spatial mean) for $ns=10$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of homoskedasticity, (G) root mean square error, (H) intercept, (I) slope of the first panel plots.
 \label{fig_spTL}}
@@
 Because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, changed significantly from the individual skewness of time-series data.
-Spatial mean of the time series data extracted that tail dependence from the multivariate data set, if any, and reflected in the averaged skewness.
-We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
-So, first we generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
+To test null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure, we generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
 Then we calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R).

### 2018-09-25 Paper.Rmd
modified

@@
 date: "February 18, 2018"
-geometry: "left=1cm,right=1cm,top=2.5cm,bottom=2.8cm"
+geometry: "left=1.5cm,right=1.5cm,top=2.5cm,bottom=2.8cm"
 
@@
 
-Considered datasets were ranged widely from population ecology, community ecology , evolutionery and biogeochemical sphere. 
+Considered datasets ranged widely from population ecology, community ecology , evolutionary and biogeochemical sphere. 
 We used four datasets in each category of bivariate (soil C and N data, BMR data, Anolis data, Ceder-creek experiment data) and multivariate (aphid-abundance data, aphid-firstflight data, plankton-abundance data and methane-flux data) examples.
@@
 \includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf}
-\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/birds/raw_birds_BMR.pdf}
 \includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
@@
 \includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf}
-\includegraphics[width=4 cm]{./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/birds/copula_birds_BMR.pdf}
 \includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
@@
 
-The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
+\textbf{(i)} The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
 correlation of $u_i$ and $v_i$ that is attributable to the points 
@@
 
-We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
+\textbf{(ii)} We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
 $u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
@@
 
-The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
+\textbf{(iii)} The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
 distance between points satisfying $u_i+v_i>2l_b$ and 
@@
 distributions of values of the same statistics computed on surrogate datasets that were produced from 
-the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
+the empirical data by randomizing it in a special way to have no tail dependence. The surrogate/randomized datasets had exactly 
 the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
@@
 
-For multivariate datasets, a spatial resampling scheme was used instead of the surrogate 
+For multivariate datasets, a spatial resampling scheme was used to calculate confidence intervals, instead of the surrogate 
 approach described above.
@@
 
-# Results \label{Results}
+# Application : copula structures in ecology, evolutionary and biogeochemical spheres \label{Results}
 
 <!--## question 1
-
 - Do datasets in ecology and related fields have non-Normal copula structure?
@@
 
-To answer \hyperref[Intro]{\textbf{(Q1)}}, we choose one representative datasets for each category : with a bivariate (soil C and N data) and a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.
+To answer \hyperref[Intro]{\textbf{(Q1)}}, we chose one representative datasets for each category : with a bivariate (soil C and N data) and a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.
 
-### Example for bivariate case : soil C and N data \label{Ex_SoilCN}
+## Example for bivariate case : soil C and N data \label{Ex_SoilCN}
 
@@
 
-#### Model selection approach \label{Ex_SoilCN_Model_selection}
+<!--#### Model selection approach \label{Ex_SoilCN_Model_selection}-->
 
-The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
-from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`).
+From the model selection approach, the test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`).
 We then fitted several copulas (all $16$ families as mentioned in \nameref{Model_selection})
@@
 These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of \hyperref[Intro]{\textbf{(Q1)}} for these data.
-Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}.
-To be precise, this difference being negative, indicated stronger upper-tail dependence than lower.
+Although these results do convincingly show non-Normal copula structure, we should caution that because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+Therefore, we also verified this finding with \nameref{NPA} approach (see Fig.
+\ref{SM-fig_soilCN_nonparam}), where each of our statistics ($\cor$, $\Ps$, $\Dsq$), as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 Normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig.
+\ref{SM-fig_soilCN_nonparam}).
+Each bin in each subplot of Fig.
+\ref{SM-fig_soilCN_nonparam} contains a number indicating how many surrogates out of 1000, are significantly different ($95\%$ confidence interval) from Normal-copula statistics.
+Those numbers clearly show that tail dependence of data is stronger in both the lower and upper tails than would be expected under a null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}.
 
-Although these results do convincingly show non-Normal copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
-The next section looks at tail dependence in a non-parametric way.
+Furthermore, the difference of model-averaged tail dependence statistics (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}.
+To be precise, this difference being negative, indicated stronger upper-tail dependence than lower, which was also supported by non-parametric results (see Table \ref{SM-tab_soilCN_asym}).
 
-#### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}
+<!--#### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}-->
 
-As described in \nameref{NPA}, each of our statistics, as computed for the 
-ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
-of values of the same statistic with the same ranges computed on 1000 Normal-copula 
-surrogate datasets with similar Kendall or Spearman correlations, in different 
-runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). Each bin in each subplot of Fig. \ref{SM-fig_soilCN_nonparam}
-contains a number which indicates how many surrogates out of 1000, are significantly different ($95\%$ confidence interval) from Normal-copula statistics.
+We summarized our results columnwise in Table \ref{tab_bivar_summary} for different kinds of bivariate datasets (biogeochemical, evolutionary and ecological data) that were plotted in Fig.
+\ref{fig_biv_raw_cop}.
 
-Results indicate that tail dependence of data is stronger in both the lower and upper tails than would be expected under a 
-null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}. The dependence between C and N is also 
-weaker in the middle of the distributions
-than would be expected based on a normal-copula null model.
-
-We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{fig_biv_raw_cop}(E). 
-Results (Table \ref{SM-tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}.
-
-We summarized our results columnwise in Table \ref{tab_bivar_summary} for all bivariate datasets that were plotted in Fig. \ref{fig_biv_raw_cop}.
-
-The first $9$ rows of Table \ref{tab_bivar_summary} summarized model selection results [namely, (1) p-value from `BiCopIndTest`, (2) best copula based on minimum AIC, (3) AIC-weight of that best copula, (4) AIC-weight of Normal copula, (5) p-value from `BiCopGofTest` for Cramer-von Mises method, (6) p-value from `BiCopGofTest` for Kolmogorov-Smirnov method, (7) AIC-weights model averaged lower tail dependence, (8) AIC-weights model averaged upper tail dependence, (9)  AIC-weights model averaged (lower-upper) tail dependence].
-Last $3$ rows ($10^{th}$ to $12^{th}$) of Table \ref{tab_bivar_summary} summarized non-parametric statistics results for $P$ statistics calculated on two extreme bins ($\Ps_l$ for first bin, $\Ps_u$ for last bin) for each copula data.
-We computed $\Ps_l$, $\Ps_u$, $\Ps_l - \Ps_u$ for the actual copula data and for Kendall-correlation 
-preserving $1000$ Normal surrogates and calculated the rank for actual data as the position if all these $1001$ values were 
-arranged in ascending order. So, whenever those ranks were $>975$ or $<25$, that indicated a significantly ($95\%$ CI) different 
-tail-dependence measures based on Normal-null hypothesis.
-
-For all $4$ bivariate-datasets indicated by each column name (SoilCN, BMR, Anolis, Ceder-creek) 
+For all $3$ bivariate-datasets indicated by each column name (SoilCN, Birds' BMR, Ceder-creek) 
 in Table \ref{tab_bivar_summary}, Normal copula were always a poor fit compared to 
 highest AIC-weight based asymmetric copulas - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}. 
-Among four datasets presented there, first three have significant upper tail dependence whereas the last one has significant lower tail dependence compared to Normal.
+Among three datasets presented there, first two have significant upper tail dependence whereas the last one has significant lower tail dependence compared to Normal.
 It was indicated by negative AIC-weights model averaged
 (lower-upper) tail dependence and very low ranks of actual copula 
-data for $\Ps_l - \Ps_u$ statistics for first three datasets and positive AIC-weights model averaged 
+data for $\Ps_l - \Ps_u$ statistics for first two datasets and positive AIC-weights model averaged 
 (lower-upper) tail dependence and very high ranks of actual copula 
@@
 
-```{r tab_bivar_summary, echo=F, results='markup'}
-
-tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
-knitr::kable(tab_bivar_summary, 
-             format='pandoc',
-             caption = "Summary of results for bivariate data-sets.\\label{tab_bivar_summary}",
-             booktabs=TRUE)
-```
-
-### Example for multivariate case : abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
+## Example for multivariate case : abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
 
@@
 
-#### Model selection approach \label{Ex_count_Model_selection}
+<!--#### Model selection approach \label{Ex_count_Model_selection}-->
 
-As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid abundance data, model selection among copulas was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations.
+As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid's abundance data, model selection among copulas was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations.
 Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question \hyperref[Intro]{\textbf{(Q1)}} for these data.
@@
 the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulas of all pairs of locations could be characterized by at least one of them.
-In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
+In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100`$\%$ of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
 
@@
 
-#### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
+<!--#### Non-parametric analysis of tail dependence \label{Ex_count_NPA}-->
 
-Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+As described in \nameref{NPA}, mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
 $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
 computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
-resampling scheme, as described in \nameref{NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
+resampling scheme (Table \ref{SM-tab_count_resamp}). Confidence intervals of
 $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
@@
 
-```{r tab_multivar_summary, echo=F, results='markup'}
-
-tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
-knitr::kable(tab_multivar_summary, 
-             format='pandoc',
-             caption = "Summary of results for multivariate data-sets.\\label{tab_multivar_summary}",
-             booktabs=TRUE)
-```
-
 We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets 
@@
 abundance data for plankton from UK seas from $14$ sampling locations, 
-methane flux data from $13$ sampling locations). First $11$ rows in Table \ref{tab_multivar_summary} represent
-model selection results :
-(1) total number of location pairs (excluding self-comparison) considered, 
-(2) how many location pair copula are non-independent among them, 
-(3) number of non-Normal copula as best fit (based on highest AIC-weight),
-(4) average AIC-weights of best fit copula from all location pair,
-(5) average AIC-weights of Normal copula from all location pair,
-(6) number of location pair for which best fit copula have p-values > 0.01 
-(both Cramer-von Mises and Kolmogorov-Smirnov method)
-(7) lower and (8) upper quantile ($95\%$CI) of model averaged (AIC-weight based) lower tail dependence from all location pair,
-(9) lower and (10) upper quantile ($95\%$CI) of model averaged (AIC-weight based) upper tail dependence from all location pair,
-(11) number of location pair which showed stronger lower tail dependence than upper.
-
-Last $6$ rows of Table \ref{tab_multivar_summary} summarized the non-parametric analysis results 
-for $\Ps$ statistics : lower and upper quantile ($95\%$CI) of $\Ps_l$, $\Ps_u$, $\Ps_l-\Ps_u$ based on resampling
-location pair.
+methane flux data from $13$ sampling locations).
 
@@
 
+```{r tab_bivar_summary, echo=F, results='markup'}
+
+tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
+knitr::kable(tab_bivar_summary, 
+             format='pandoc',
+             caption = "Summary of results for bivariate data-sets : \\textbf{Indep.p}: p-value from `BiCopIndTest`, \\textbf{AIC-best.copula}: best copula based on minimum AIC, \\textbf{Highest.AICw}: AIC-weight of that best copula, \\textbf{N.AICw}: AIC-weight of Normal copula, \\textbf{p.CvM}: p-value from `BiCopGofTest` for Cramer-von Mises method, \\textbf{p.KS}: p-value from `BiCopGofTest` for Kolmogorov-Smirnov method,\\textbf{LT.model.avg}: AIC-weights model averaged lower tail dependence, \\textbf{UT.model.avg}: AIC-weights model averaged upper tail dependence, \\textbf{(LT-UT).model.avg}:  AIC-weights model averaged (lower-upper) tail dependence.
+$\\Ps_l$, $\\Ps_u$, $\\Ps_l$ - $\\Ps_u$ for the actual copula data and for Kendall-correlation preserving $1000$ Normal surrogates were computed and the rank for actual data (\\textbf{Rank.Pl.K, Rank.Pu.K, Rank.Pl-Pu.K}) were calculated as the position if all these $1001$ values were arranged in ascending order.
+Ranks if were $>975$ or $<25$, then indicated a significantly ($95\\%$ CI) different tail-dependence measures based on Normal-null hypothesis.
+\\label{tab_bivar_summary}",
+             booktabs=TRUE)
+```
+
+```{r tab_multivar_summary, echo=F, results='markup'}
+
+tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
+knitr::kable(tab_multivar_summary, 
+             format='pandoc',
+             caption = "Summary of results for multivariate data-sets : \\textbf{num.pairs}: total number of location pairs (excluding self-comparison) considered, \\textbf{num.non-indep}: how many location pair copula are non-independent among them, \\textbf{num.non-N}: number of non-Normal copula as best fit (based on highest AIC-weight), \\textbf{Avg.best.AICw}: average AIC-weights of best fit copula from all location pair, \\textbf{Avg.N.AICw}: average AIC-weights of Normal copula from all location pair, \\textbf{num.good.fit}: number of location pair for which best fit copula have p-values $>$ 0.01
+(both Cramer-von Mises and Kolmogorov-Smirnov method), \\textbf{0.025-CI.LT}: lower and \\textbf{0.975-CI.LT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) lower tail dependence from all location pair, \\textbf{0.025-CI.UT}: lower and \\textbf{0.975-CI.UT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) upper tail dependence from all location pair, \\textbf{num.LT-UT>0}: number of location pair which showed stronger lower tail dependence than upper, Last $6$ rows summarized the non-parametric analysis results for $\\Ps$ statistics : lower and upper quantile ($95\\%$CI) of $\\Ps_l$, $\\Ps_u$, $\\Ps_l-\\Ps_u$ based on resampling location pair.\\label{tab_multivar_summary}",
+             booktabs=TRUE)
+```
+
 <!--Question 2 analyses-->
-## Some possible causes of non-Normal copula structure \label{Causes}
+# Some possible causes of non-Normal copula structure \label{Causes}
+
+This section is dedicated to explore question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}.
+We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change?
+
+\textbf{(i)} To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
+\ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.
+
+\begin{equation}\label{eq.AR1}
+P_{i}(t+1)=\beta P_{i}(t)+\alpha \epsilon_{i}(t)
+\end{equation}
+
+To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$.
+We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches.
+Should their dependence be exactly the same as of the noises or not ?
+
+To test that our approach is of two way : first, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
+Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome which was a bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
+\ref{fig_Cause4copula_params_GOF_500}).
+We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size and reported p-values for both statistics (Cramer-von Mises and Kolmogorov-Smirnov).
+p-values greater or equal to $0.05$ (above the dashed line in Fig.
+\ref{fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters.
+Fig.
+\ref{SM-fig_Cause4copula_params_GOF_2500} represents similar feature as of Fig.
+\ref{fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$.
+(For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf}
+\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{500} points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF_500}}
+\end{center}
+\end{figure}
+
+Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
+Then we plotted mean of each measured quantity with standard error against Kendall's Tau of the noise copula family.
+Fig.
+\ref{fig_Cause4copula_C_kend} and Fig.
+\ref{fig_Cause4copula_SC_kend} represent this picture for noise generating from a lower tail dependent Clayton and upper tail dependent Survival Clayton copula whereas for Fig.
+\ref{SM-fig_Cause4copula_C_spear} and Fig.
+\ref{SM-fig_Cause4copula_SC_spear} Spearman's Rho of noise copula were plotted along x-axis.
+We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach, here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise.
+(For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Cause4copula_results/Clayton_kendall.pdf}
+\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_C_kend}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_kendall.pdf}
+\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}}
+\end{center}
+\end{figure}
+
+\textbf{(ii)} On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold.
+So, we considered another two-habitat patch model which was influenced by both local and global noises and found that a long time exposure to assymetric copula for global fluctuation shaped the population copula accordingly (for details, see Appendix \ref{SM-Cause_asym_sens}).
+
+## Application : The Moran effect and copula structure in population
 
@@
 
-This section is dedicated to explore question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}.
-We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change?
 <!--We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we don’t really know how a change in the environment copula structure would make what kind of change in the population copula structure.-->
@@
 This assumption was inspired by the the idea of Moran-effect [@Moran1953] which was originally used to explain synchronization of widely dispersed populations.
-We tested our hypothesis with a two habitat patch model (1st order AR method) in noisy environment (having non-Normal copula structure) and found that populations were shaped by similar copula structure as that of local noises ( for details see Appendix \ref{SM-Cause_Moran}).
-We also found the Moran effect on leaf-plum curling aphid's firstflight data induced by average winter temperature for the period of 56 years (1958-2013).
+In the previous section, we already tested our hypothesis with a two habitat patch model (1st order AR method) in noisy environment (having non-Normal copula structure) and found that populations were shaped by similar copula structure as that of local noises.
+Here, we investigated the probable cause of upper tail dependence in the leaf-curling plum aphid's firstflight date throughout all the locations.
+We assumed that average winter temperature which might be spatially synchronized and had asymmetric tail dependence cause the synchronized delay in leaf-plum curling aphid's appearence.
+We tested our hypothesis and indeed found the Moran effect on leaf-plum curling aphid's firstflight data induced by average winter temperature for the period of 35 years (1976-2010).
 We did pair wise model selection with negative average winter temperature (averaged over (n-1)$^{th}$ December to n$^{th}$ March temperature) from the selected 10 locations for which we did model selection with aphid-firstflight data.
@@
 
-On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold.
-So, we considered another two-habitat patch model which was influenced by both local and global noises and found that a long time exposure to assymetric copula for global fluctuation shaped the population copula accordingly (for details, see Appendix \ref{SM-Cause_asym_sens}).
-
 ### Need to think about mechanisms for biogeochemical, evolutionary, and community data
@@
 <!--Question 3 analyses-->
-## Consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? \label{Consequences}
+# Consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? \label{Consequences}
 
-This section is attributed to analyze of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}.
+This section is attributed to analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}.
 We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data.

### 2018-09-25 SupportingInformation.Rmd
modified

@@
     fig_caption: yes
+    toc: True
     includes:
@@
 
-\tableofcontents
+<!--\tableofcontents-->
 \listoftables
@@
 
-```{r read_BMR_data, echo=FALSE, results="markup"}
-set.seed(seed)
-#---------------------------------------------------------
-BMR_dat<-read.csv("./Data/BMR/bmr-contrasts.csv",header = T)
-#class(BMR_dat)
-#knitr::kable(head(BMR_dat[,c(1,4,5)] ,3),format='pandoc',
-#             caption = "First few rows of raw BMR data.\\label{tab_rawdata_BMR}",
-#             booktabs=T) # display first few rows
-```
-
-```{r plot_BMR_data, echo=FALSE, results="hide"}
-set.seed(seed)
-source("phylo_dat_fn.R")
-
-# Generating raw data plot btw ln(mass) and BMR from bmr-contrasts.csv
-pdf("./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf")
+```{r read_BMR_birds_and_copulaplot, echo=F, results='hide'}
+set.seed(seed)
+raw_b<-readxl::read_xls("./Data/BMR/BirdBodyMassesMetabolicRates/BirdMassesAndMetabolicRatesMcNab2008.xls")
+raw_b<-raw_b[,1:3]
+raw_b<-na.omit(raw_b)
+raw_b<-raw_b[,2:3]
+raw_b<-as.data.frame(raw_b)
+raw_b[303,2]<-0.66  #to convert as numeric
+raw_b[,2]<-as.numeric(raw_b[,2])
+raw_b<-raw_b[-which(raw_b[,1]<0),]  # omit -ve mass?
+
+b<-log10(raw_b)
+colnames(b)<-c("log10M(g)","log10BMR(KJ/h)")
+pdf("./Results/BMR_results/birds/raw_birds_BMR.pdf")
 op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(BMR_dat$ln.mass,BMR_dat$ln.BMR,type="p",col="red",xlab="ln(mass) of contrasts",ylab="ln(BMR) of contrasts",
-     cex.lab=1.5,cex.axis=1.5)
+plot(b[,1],b[,2],type="p",col="red",xlab="log(Mass)",ylab="log(BMR)",cex.lab=1.5,cex.axis=1.5)
 par(op)
 dev.off()
-#--------------------------------------------------------------------------------------------
-# Generating copula plot btw ln(mass) and ln(BMR) from bmr-contrasts.csv
-pdf("./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf")
-v_BMR<-phylo_dat_fn(data=BMR_dat,i=4,j=3,ploton = T) 
-#plot(v_BMR[,1],v_BMR[,2],type="p",col="red",xlab="ln(mass)",ylab="BMR",
-#     cex.lab=1.5,cex.axis=1.5)
-dev.off()
-```
-
-```{r read_phylo_data, echo=F, results="markup"}
-set.seed(seed)
-phylo_dat<-read.csv("./Data/Anolis/contrasts.csv",header = T)
-#class(phylo_dat) #it's a data.frame : no NA/missing value in this dataset
-#tab_phylo<-phylo_dat[,c(1,7,8)] 
-#knitr::kable(head(tab_phylo,3),format='pandoc',
-#             caption = "First few rows of raw phylo data.\\label{tab_rawdata_phylo}",
-#             booktabs=T) # display first few rows
-```
-
-```{r plot_phylo_data, echo=F, results="hide"}
-set.seed(seed)
-source("phylo_dat_fn.R")
-#----------------------------------------------------------------------------------------------------
-# Generating all copula plots between different pathlength and traits 
-pdf("./Results/Anolis_results/contrasts_csv_results/contrasts_csv_pathlength_traits_cop_plot.pdf",width=20, height = 20)
-op<-par(mfrow=c(4,4))
-i<-1
-  for(j in c(1:13)){
-    v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = T)
-    mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 3,col="blue")
-  } 
-par(op)
-dev.off()
-#------------------------------------------------------------------------------------------------------------
-# Generating all 12 by 12 copula plots between different traits
-pdf("./Results/Anolis_results/contrasts_csv_results/contrasts_csv_btw_traits_cop_plot.pdf",width=20, height = 20)
-op<-par(mfrow=c(12,12))
-for(i in c(2:13)){
-  for(j in c(2:13)){
-    v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = T)
-    mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 4,col="blue",line=0.01)
-  } 
-}
-par(op)
-dev.off()
-#----------------------------------------------------------------------------------------
-# Generating raw data plot btw AVG.ltoe.IV and AVG.humerus from contrasts.csv
-pdf("./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf")
-op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(phylo_dat$AVG.ltoe.IV,phylo_dat$AVG.humerus,type="p",col="red",xlab="contrasts for AVG.ltoe.IV",ylab="contrasts for AVG.humerus",
-     cex.lab=1.5,cex.axis=1.5)
-par(op)
-dev.off()
-#--------------------------------------------------------------------------------------------
-# Generating copula plot btw AVG.ltoe.IV and AVG.humerus from contrasts.csv
-pdf("./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf")
-v_contrasts_67<-phylo_dat_fn(data=phylo_dat,i=6,j=7,ploton = T)
+
+source("getcopula.R")
+pdf("./Results/BMR_results/birds/copula_birds_BMR.pdf")
+cop_b<-getcopula(d=b,rankon=T,ploton=T) #convert to a copula and plot
 dev.off()
@@
 additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.
-
-<!--# Nonparametric analysis of tail dependence \label{NPA}
-
-Left (or lower)- and right (or upper)-tail dependence quantify the extent to which two ranked
-variables $u_i$, $v_i$ for $i=1,...,n$ 
-(see \nameref{MT-Model_selection} for defintions of $u_i$ and $v_i$ in the manuscript)
-are related to each other in the left and right tails, respectively, of their 
-distributions [@nelsen2007_copula, @joe2014_dependence]. 
-Our nonparametric study of tail dependence uses three
-statistics which can quantify the extent to which $u_i$ and $v_i$ are 
-related in any part of their distributions. We here describe the 
-statistics. The statistics were formulated with positively associated
-variables in mind. All the variables we study are positively associated
-when they are significantly associated. So definitions and associated 
-diagrams will assume positive associations, though generalizations are 
-possible. Given two bounds 
-$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
-$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
-fall (these are the blue and red lines, respectively, in 
-Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
-between $u_i$ and $v_i$ in the region bounded by these lines.
-Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
-dependence in the left parts of the distributions of $u$ and $v$, and 
-using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
-the right parts of the distributions.
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
-\caption{Schematic diagram supporting defintions of nonparametric tail dependence statistics \label{fig_stat_image}}
-\end{center}
-\end{figure}
-
-The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
-correlation of $u_i$ and $v_i$ that is attributable to the points 
-lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
-\begin{equation}\label{eq.Cor}
-   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
-   (u_i-\mean(u)) (v_i-\mean(v)),
-\end{equation}
-where means and variances are computed using all $n$ data 
-points, but the
-sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
-$u_i+v_i < 2u_b$. 
-Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
-between $u$ and $v$ in the region given by the bounds.
-
-We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
-$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
-$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
-and a distance less than $h$ from the line $v=u$, divided by the 
-total number of points within the bounds. This function is
-defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
-is half the longer of the two segments obtained by intersecting the
-bounds with the unit square. It is easy to see $S(0)=0$ and 
-$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
-of $h$, to be the area within the bounds and the unit square
-and within distance $h$ of the line $v=u$, divided by the area
-within the bounds and the unit square. This is the expected value of
-$S(h)$ for independent data. We then defined
-\begin{equation}\label{eq.P}
-   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
-\end{equation}
-As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
-indicate stronger positive association
-between $u$ and $v$ in the region given by the bounds.
-
-The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
-distance between points satisfying $u_i+v_i>2l_b$ and 
-$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
-$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
-$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
-then small. 
-Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
-dependence in the portions of the distributions of $u$ and $v$ given by 
-the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
-$\Dsq_{l_b,u_b}$ indicate strong dependence.
-
-For large datasets (i.e., datasets with large $n$),
-we can consider $l_b$ and $u_b$ close together without incurring undue
-sampling variation in our statistics, and we can consider 
-several different bands $(l_b,u_b)$ to understand how dependence
-varies in different parts of the distributions. But for datasets with 
-smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
-We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
-and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
-$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.
-
-The difference $\cor_l-\cor_u$ is expected to be positive if 
-dependence in the left halves of the distributions is stronger 
-than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
-This holds, for instance, for the Clayton copula [@nelsen2007_copula].
-The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
-"lower" statistics occur in opposite order, is also expected to be 
-positive under the same conditions. We use these three differences
-to detect asymmetry of tail dependence. Likewise, for large datasets,
-we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
-$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
-$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
-
-We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas (for details see Appendix \ref{Test_npa_stat}).
-
-For bivariate datasets, we compared values 
-of the statistics $\cor_{l_b,u_b}$,
-$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to 
-distributions of values of the same statistics computed on surrogate datasets that were produced from 
-the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
-the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
-correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
-but had Normal copula structure. Thus our comparisons tested the null hypothesis 
-that our statistics took values on the empirical data no different from what would have been expected 
-if the copula structure of the data were Normal, but the data were otherwise 
-statistically unchanged.
-Details of how surrogates were prroduced are in Appendix \ref{surrog_test}, and code is in [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R).
-
-The comparison of one of the
-statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
-empirical data, to the distribution
-of its values computed on many surrogate datasets 
-provides a test of whether dependence in the part of the distribution specified 
-by $l_b$ and $u_b$ is different from
-what would be expected from a null hypothesis of normal copula structure but 
-unchanged marginals and unchanged 
-Kendall (or Spearman)
-correlation. Thus the comparison helps answer the first two parts of question \hyperref[MT-Intro]{\textbf{(Q1)}} 
-from the \nameref{MT-Intro} for bivariate datasets. 
-Significant deviations correspond to deviations from normal 
-copula structure. In particular, deviations 
-using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
-different from that of a normal copula;
-likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
-from that of a normal copula.
-
-The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
-as computed on the 
-empirical data, to the distribution
-of its values on many surrogate datasets 
-provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
-third part of question \hyperref[MT-Intro]{\textbf{(Q1)}} from the \nameref{MT-Intro} for bivariate datasets.
-
-For multivariate datasets, a spatial resampling scheme was used instead of the surrogate 
-approach described above.
-The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
-widely used spline correlogram method; the same scheme was also used by @walteretal2017 
-as part of a method for 
-performing model selection among matrix regression models. For each pair of 
-locations $i \neq j$, the statistic
-of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
-$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
-$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
-values $s_{ij}$ were then arranged in
-a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
-$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
-average strength of 
-dependence, across all pairs of distinct locations, for data in the portion of the distributions 
-given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
-asymmetry of tail dependence across all pairs of locations. 
-Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
-distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
-revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
-or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
-Normal null hypothesis, and answering question \hyperref[MT-Intro]{\textbf{(Q1)}} from the \nameref{MT-Intro}.
-Bootstrapping was performed by resampling the locations of measurement, with replacement. 
-This corresponds to 
-resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
-[@bjornstadfalck2001 ; @walteretal2017].-->
 
@@
 
-<!--
-# Question 1 analyses for each dataset
-Recall that question 1 is: Do datasets in ecology and related fields have non-Gausian copula structure?
-Do they show tail dependence distinct from that of a Normal copula?
-In particular do they show asymmetric tail dependence?
-
-## Details of methods \label{Method}
-### Model selection approach \label{Model_selection}
-### Non-parametric analysis of tail dependence \label{NPA}
-moved to Paper.Rmd
--->
-
-<!--
-## Results: soil C and N data \label{Ex_SoilCN}
-### Model selection approach \label{Ex_SoilCN_Model_selection} 
-### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}
-                    [[[[moved to Paper.Rmd]]]]
--->
-
 <!--This chunk is for model selection results of soilCN dataset-->
@@
 
-<!--
-\newpage
-## Results: BMR data \label{Ex_BMR}
-### Model selection approach \label{Ex_BMR_Model_selection}
-### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
-                  ----moved to Paper.Rmd----
--->
-
-<!--This chunk is for model selection results of BMR data-->
-```{r bivMS_BMR, echo=F, cache=T, cache.extra=list(seed,v_BMR,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-set.seed(seed) 
+```{r bivmodselect_birds_BMR, echo=F, cache=T, cache.extra=list(seed,cop_b,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
 source("OurBiCopSelect.R")
-#families<-c(1,3:10,13:14,16:20) 
-BivMS_BMR<-OurBiCopSelect(u1=v_BMR[,1],u2=v_BMR[,2],families=c(1,3:10,13:14,16:20),
-                          level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
-saveRDS(BivMS_BMR,file="./Results/BMR_results/BivMS_lnBMR_vs_lnmass.RDS")
-```
-
-<!--Display model selection results for BMR data in a table-->
-```{r tab_BMRfit, echo=F, results='markup'}
-h<-BivMS_BMR$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+set.seed(seed)
+BivMS_birds<-OurBiCopSelect(u1=cop_b[,1],u2=cop_b[,2],families=c(1,3:10,13:14,16:20),
+                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
+                              gofnormal=F,status=F)
+saveRDS(BivMS_birds,file="./Results/BMR_results/birds/BivMS_birds.RDS")
+```
+
+<!--Display model selection results for BMR:birds data in a table-->
+```{r tab_bmr_birds_fit, echo=F, results='markup'}
+h<-BivMS_birds$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
 knitr::kable(h,
              format='pandoc',
-             caption = "Fitting info for **BMR dataset**. \\label{tab_BMRfit}",
+             caption = "Fitting info for **birds' BMR dataset** . \\label{tab_bmr_birds_fit}",
              booktabs=T)
@@
 
-<!--This chunk is for npa testing results of BMR dataset-->
-```{r stat_BMR_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_BMR,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+```{r stat_bmr_birds_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_b,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
 source("bivfunctionplot.R")
 set.seed(seed)
-# Now do NPA for v_BMR
-stat_BMR_ga<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_10",numbin = 10)
-stat_BMR_ga_numbin5<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_5",numbin = 5)
+stat_birds_ga<-bivfunctionplot(v=cop_b,resloc="./Results/BMR_results/birds/",nametag="birds_numbin_10",numbin = 10)
 ```
@@
 \begin{center}
-\includegraphics[width=16 cm]{./Results/BMR_results/BMR_numbin_10_bivfunctionplot.pdf}
-\caption[Testing non-parametric statistics for BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{BMR data}.
+\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
+\caption[Testing non-parametric statistics for Ceder-creek data with $numbin=10$ for year 2000]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{birds' BMR data}.
 As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
 When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
-\label{fig_BMR_nonparam}}
+\label{fig_bmr_birds_nonparam}}
 \end{center}
@@
 
-<!--
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/BMR_results/BMR_numbin_5_bivfunctionplot.pdf}
-\caption[Testing non-parametric statistics for BMR data with $numbin=5$]{Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data.
-As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.2$, $0.2-0.4$, ..., $0.8-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).
-\label{fig_BMR_nonparam5}}
-\end{center}
-\end{figure}
--->
-
-```{r tab_BMR_asym, echo=F, results="markup"}
-tab_BMR_asym<-make_table_stat_fn(data=stat_BMR_ga)
-tab_BMR_asym_numbin5<-make_table_stat_fn(data=stat_BMR_ga_numbin5)
-
-knitr::kable(tab_BMR_asym,
+```{r tab_bmr_birds_asym, echo=F, results="markup"}
+tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
+knitr::kable(tab_bmr_birds_asym,
              format="pandoc",
-             caption = "**BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis. \\label{tab_BMR_asym}",
+             caption = "**Birds' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis on year 2000 \\label{tab_bmr_birds_asym}",
              booktabs=T)
-
-#knitr::kable(tab_BMR_asym_numbin5,
-#             format="pandoc",
-#             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)?, numbin=5 \\label{tab_BMR_asym_numbin5}",
-#             booktabs=T)
-```
-
-<!--
-\newpage
-## Results: contrasts.csv data \label{Ex_contrasts}
-### Model selection approach \label{Ex_contrasts_Model_selection}
-### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}
-                 ----moved to Paper.Rmd----
--->
-
-<!--This chunk is for model selection results of anolis data-->
-```{r bivMS_contrasts_67, echo=F, cache=T, cache.extra=list(seed,v_contrasts_67,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-set.seed(seed)
-# Now do Model Selection for [i,j]=[6,7]
-# 6 : AVG. ltoe.IV
-# 7 : AVG. humerus
-source("OurBiCopSelect.R")
-families<-c(1,3:10,13:14,16:20)
-BivMS_contrasts_67<-OurBiCopSelect(u1=v_contrasts_67[,1],u2=v_contrasts_67[,2],families=families,
-                          level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
-                          gofnormal=F,status=F)
-saveRDS(BivMS_contrasts_67,file="./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
-```
-
-<!--Display model selection results for contrasts data in a table-->
-```{r tab_contrasts_67fit, echo=F, results='markup'}
-h<-BivMS_contrasts_67$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-knitr::kable(h,
-             format='pandoc',
-             caption = "Fitting info for **Anolis data set** with contrasts for AVG.ltoe.IV and AVG.humerus phylogenetic traits. \\label{tab_contrasts_67fit}",
-             booktabs=T)
-```
-
-<!--This chunk is for npa testing results of anolis dataset-->
-```{r stat_contrasts_67_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_contrasts_67,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
-source("bivfunctionplot.R")  
-set.seed(seed) 
-# Now do NPA for v_contrasts_67 with numbin=4
-stat_contrasts_67_ga<-bivfunctionplot(v=v_contrasts_67,resloc="./Results/Anolis_results/contrasts_csv_results/",
-                                      nametag="contrasts_67",numbin = 4) 
-```
-
-```{r stat_contrasts_67_graphic_approach2, echo=F, results="hide", cache=T, cache.extra=list(seed,v_contrasts_67,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
-source("bivfunctionplot.R")
-set.seed(seed) 
-# Now do NPA for v_contrasts_67
-
-stat_contrasts_67_ga_numbin2<-bivfunctionplot(v=v_contrasts_67,
-                            resloc="./Results/Anolis_results/contrasts_csv_results/",
-                            nametag="contrasts_67_numbin2",numbin = 2)
-
-stat_contrasts_67_ga_numbin3<-bivfunctionplot(v=v_contrasts_67,
-                            resloc="./Results/Anolis_results/contrasts_csv_results/",
-                            nametag="contrasts_67_numbin3",numbin = 3)
-
-stat_contrasts_67_ga_numbin5<-bivfunctionplot(v=v_contrasts_67,
-                            resloc="./Results/Anolis_results/contrasts_csv_results/",
-                            nametag="contrasts_67_numbin5",numbin = 5)
-
-stat_contrasts_67_ga_numbin6<-bivfunctionplot(v=v_contrasts_67,
-                            resloc="./Results/Anolis_results/contrasts_csv_results/",
-                            nametag="contrasts_67_numbin6",numbin = 6)
-
-stat_contrasts_67_ga_numbin7<-bivfunctionplot(v=v_contrasts_67,
-                            resloc="./Results/Anolis_results/contrasts_csv_results/",
-                            nametag="contrasts_67_numbin7",numbin = 7)
-
-stat_contrasts_67_ga_numbin10<-bivfunctionplot(v=v_contrasts_67,
-                            resloc="./Results/Anolis_results/contrasts_csv_results/",
-                            nametag="contrasts_67_numbin10",numbin = 10)
-```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/Anolis_results/contrasts_csv_results/contrasts_67_bivfunctionplot.pdf}
-\caption[Testing non-parametric statistics for anolis data]{Nonparametric tests for tail dependence and other deviations from a normal copula for anolis data.
-As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.25$, $0.25-0.5$, $0.5-0.75$, $0.75-1$ for $4$ number of bins (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
-\label{fig_contrasts_67_nonparam}}
-\end{center}
-\end{figure}
-
-```{r tab_contrasts_67_asym, echo=F, results="markup"}
-tab_contrasts_67_asym<-make_table_stat_fn(data=stat_contrasts_67_ga)
-knitr::kable(tab_contrasts_67_asym,
-             format="pandoc",
-             caption = "**Anolis data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_contrasts_67_asym}",
-             booktabs=T)
-```
-
-<!--
-This chunks are just to see the tables for all numbin variation for anolis data set: It appears that numbin=4 is optimized, when numbin=2, >= 7, there is no significant asymetry in tail dep in all 3 types of stats.
-
-```{r tab_contrasts_67_asym67_numbin2, echo=F, results="markup"}
-#tab_contrasts_67_asym2<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin2)
-#knitr::kable(tab_contrasts_67_asym2,
-#            format="pandoc",
-#             caption = "numbin_2,anolis \\label{tab_contrasts_67_asym2}",
-#             booktabs=T)
-```
-
-```{r tab_contrasts_67_asym67_numbin3, echo=F, results="markup"}
-#tab_contrasts_67_asym3<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin3)
-#knitr::kable(tab_contrasts_67_asym3,
-#             format="pandoc",
-#             caption = "numbin_3,anolis \\label{tab_contrasts_67_asym3}",
-#             booktabs=T)
-```
-
-```{r tab_contrasts_67_asym67_numbin5, echo=F, results="markup"}
-#tab_contrasts_67_asym5<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin5)
-#knitr::kable(tab_contrasts_67_asym5,
-#            format="pandoc",
-#             caption = "numbin_5,anolis \\label{tab_contrasts_67_asym5}",
-#             booktabs=T)
-```
-
-```{r tab_contrasts_67_asym67_numbin6, echo=F, results="markup"}
-#tab_contrasts_67_asym6<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin6)
-#knitr::kable(tab_contrasts_67_asym6,
-#             format="pandoc",
-#             caption = "numbin_6,anolis \\label{tab_contrasts_67_asym6}",
-#             booktabs=T)
-```
-
-```{r tab_contrasts_67_asym67_numbin7, echo=F, results="markup"}
-#tab_contrasts_67_asym7<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin7)
-#knitr::kable(tab_contrasts_67_asym7,
-#             format="pandoc",
-#             caption = "numbin_7,anolis \\label{tab_contrasts_67_asym7}",
-#             booktabs=T)
-```
-
-```{r tab_contrasts_67_asym67_numbin10, echo=F, results="markup"}
-#tab_contrasts_67_asym10<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin10)
-#knitr::kable(tab_contrasts_67_asym10,
-#             format="pandoc",
-#             caption = "numbin_10,anolis \\label{tab_contrasts_67_asym10}",
-#             booktabs=T)
-```
-
--->
+```
 
@@
 
-\newpage
 <!--Display model selection results for Ceder-creek data for year 2000 in a table-->
@@
 ```
-
-<!--
-\newpage
-## Results: abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
-### Model selection approach \label{Ex_count_Model_selection}
-### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
-                  ----moved to Paper.Rmd----
--->
 
@@
 
-<!--
-\newpage
-## Results: first flight date data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}
-### Model selection approach \label{Ex_ff_Model_selection}
-### Non-parametric analysis of tail dependence  \label{Ex_ff_NPA}
-                          ----moved to Paper.Rmd----
--->
-
 <!--Do the model selection for aphid ff data and then showing results in Tables-->
@@
 ```
-
-\newpage
 
@@
 
-<!--
-\newpage
-## Results: methane flux data \label{Ex_methane}
-### Model selection approach \label{Ex_methane_Model_selection}
-### Non-parametric analysis of tail dependence  \label{Ex_methane_NPA}
-                 ----moved to Paper.Rmd----
--->
-
 <!--Load and format the data-->
@@
                  "N.AICw","p.CvM","p.KS",
-                 "LT_model.avg","UT_model.avg","(LT-UT)_model.avg",
-                 "Rank_Pl_K","Rank_Pu_K","Rank_Pl-Pu_K")
+                 "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
+                 "Rank.Pl.K","Rank.Pu.K","Rank.Pl-Pu.K")
   m<-as.data.frame(m)
@@
   m$p.KS<-r$GofRes_KS
-  m$LT_model.avg<-round(r$relLTdep_AICw,3)
-  m$UT_model.avg<-round(r$relUTdep_AICw,3)
-  m$`(LT-UT)_model.avg`<-m$LT_model.avg-m$UT_model.avg
-  m$Rank_Pl_K<-rs$Rank_Pl_K
-  m$Rank_Pu_K<-rs$Rank_Pu_K
-  m$`Rank_Pl-Pu_K`<-rs$Rank_PlmPu_K
+  m$LT.model.avg<-round(r$relLTdep_AICw,3)
+  m$UT.model.avg<-round(r$relUTdep_AICw,3)
+  m$`(LT-UT).model.avg`<-m$LT.model.avg-m$UT.model.avg
+  m$Rank.Pl.K<-rs$Rank_Pl_K
+  m$Rank.Pu.K<-rs$Rank_Pu_K
+  m$`Rank.Pl-Pu.K`<-rs$Rank_PlmPu_K
   mt<-as.data.frame(t(m))
@@
 #----------------------------------
-#calling the above function on 3 multivariable datasets
+#calling the above function on bivariable datasets
 m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke,rs=stat_soilCN_ga)
-m2<-bivar_tab_summary_fn(r=BivMS_BMR,rs=stat_BMR_ga)
-m3<-bivar_tab_summary_fn(r=BivMS_contrasts_67,rs=stat_contrasts_67_ga)
-m4<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
-tab_bivar_summary<-cbind(m1,m2,m3,m4)
-colnames(tab_bivar_summary)<-c("SoilCN","BMR","Anolis","Ceder-creek")
+m2<-bivar_tab_summary_fn(r=BivMS_birds,rs=stat_birds_ga)
+m3<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
+tab_bivar_summary<-cbind(m1,m2,m3)
+colnames(tab_bivar_summary)<-c("SoilCN","Birds' BMR","Ceder-creek")
 saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
@@
   m<-matrix(NA,1,17)
-  colnames(m)<-c("num_pairs","num_non-indep","num_non-N",
-                 "Avg.best.AICw","Avg.N.AICw","num_good.fit",
-                 "CI0.025_LT","CI0.975_LT","CI0.025_UT","CI0.975_UT",
-                 "num_LT-UT>0",
-                 "CI0.025_Pl","CI0.975_Pl","CI0.025_Pu","CI0.975_Pu","CI0.025_Pl-Pu","CI0.975_Pl-Pu")
+  colnames(m)<-c("num.pairs","num.non-indep","num.non-N",
+                 "Avg.best.AICw","Avg.N.AICw","num.good.fit",
+                 "0.025-CI.LT","0.975-CI.LT","0.025-CI.UT","0.975-CI.UT",
+                 "num.LT-UT>0",
+                 "0.025-CI.Pl","0.975-CI-Pl","0.025-CI.Pu","0.975-CI.Pu","0.025-CI.Pl-Pu","0.975-CI.Pl-Pu")
   m<-as.data.frame(m)
@@
   d<-dim(r$gfc_numBS)[1]
-  m$num_pairs<-(d^2)-d 
-  m$`num_non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
-  m$`num_non-N`<-sum(r$info_ord_copcode[,,1]!=1,na.rm = T)
+  m$num.pairs<-(d^2)-d 
+  m$`num.non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
+  m$`num.non-N`<-sum(r$info_ord_copcode[,,1]!=1,na.rm = T)
   r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN
@@
   temp<-which((r$gfc_p_CvM>0.01) & (r$gfc_p_KS>0.01),arr.ind=T)
-  m$num_good.fit<-dim(temp)[1]
+  m$num.good.fit<-dim(temp)[1]
   t<-r$LTdep_AICw[is.finite(r$LTdep_AICw)]
   CI_LT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
-  m$CI0.025_LT<-round(CI_LT[1],3)
-  m$CI0.975_LT<-round(CI_LT[2],3)
+  m$`0.025-CI.LT`<-round(CI_LT[1],3)
+  m$`0.975-CI.LT`<-round(CI_LT[2],3)
   t<-r$UTdep_AICw[is.finite(r$UTdep_AICw)]
   CI_UT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
-  m$CI0.025_UT<-round(CI_UT[1],3)
-  m$CI0.975_UT<-round(CI_UT[2],3)
-  m$`num_LT-UT>0`<-sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)
-  m$CI0.025_Pl<-round(rs$numericdf[5,3],3)
-  m$CI0.975_Pl<-round(rs$numericdf[5,4],3)
-  m$CI0.025_Pu<-round(rs$numericdf[6,3],3)
-  m$CI0.975_Pu<-round(rs$numericdf[6,4],3)
-  m$`CI0.025_Pl-Pu`<-round(rs$numericdf[10,3],3)
-  m$`CI0.975_Pl-Pu`<-list(round(rs$numericdf[10,4],3))
+  m$`0.025-CI.UT`<-round(CI_UT[1],3)
+  m$`0.975-CI.UT`<-round(CI_UT[2],3)
+  m$`num.LT-UT>0`<-sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)
+  m$`0.025-CI.Pl`<-round(rs$numericdf[5,3],3)
+  m$`0.975-CI-Pl`<-round(rs$numericdf[5,4],3)
+  m$`0.025-CI.Pu`<-round(rs$numericdf[6,3],3)
+  m$`0.975-CI.Pu`<-round(rs$numericdf[6,4],3)
+  m$`0.025-CI.Pl-Pu`<-round(rs$numericdf[10,3],3)
+  m$`0.975-CI.Pl-Pu`<-list(round(rs$numericdf[10,4],3))
   #m$`CI0.975_Pl-Pu`<-round(rs$numericdf[10,4],3)
@@
 
-<!--
-#Results section detailed description
-### soil C and N data \label{Ex_SoilCN}
-
-```{r read_BivMS_soilCN_data,echo=F}
-BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
-```
-
-#### Model selection approach \label{Ex_SoilCN_Model_selection}
-
-The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available 
-from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}).
-The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
-
-The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`.
-These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.
-
-Although these results do convincingly show non-Normal copula structure, we should take the tail dependence results with a grain of salt because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
-The next section looks at tail dependence in a non-parametric way.
-
-#### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}
-
-As described in \nameref{Methods}, each of our statistics, as computed for the 
-ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
-of values of the same statistic with the same ranges computed on 1000 normal-copula 
-surrogate datasets with similar Kendall or Spearman correlations, in different 
-runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}).
-
-Results indicate that tail dependence of 
-data is stronger in both the lower and upper tails than would be expected under a 
-null hypothesis of a normal copula, answering the second part of question 1. The dependence between C and N is also 
-weaker in the middle of the distributions
-than would be expected based on a normal-copula null model.
-
-We also tested statistically for asymmetry of tail dependence, which seems visible in Fig. \ref{SM-fig_raw_cop_soilCN}(B). 
-Results (Table \ref{SM-tab_soilCN_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.
-
-### BMR data \label{Ex_BMR}
-
-```{r read_BivMS_BMR_data,echo=F}
-BivMS_BMR<-readRDS("./Results/BMR_results/BivMS_lnBMR_vs_lnmass.RDS")
-```
-
-#### Model selection approach \label{Ex_BMR_Model_selection}
-
-The test of indepedence gave $p=$ `r BivMS_BMR$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_BMRfit}).
-The table shows that BB6 and BB8 are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
-
-The model-averaged (AIC weights used for model averaging) lower-tail and upper-tail dependence statistic was `r BivMS_BMR$relLTdep_AICw - BivMS_BMR$relUTdep_AICw`, indicating more upper tail dependence.
-We also tested that lowest-AIC copula BB8 was a good-enough fit, giving $p$-value `r BivMS_BMR$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_BMR$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
-The next section looks at tail dependence in a non-parametric way.
-
-#### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
-
-As described in \nameref{Methods}, each of our statistics, as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig.
-\ref{SM-fig_BMR_nonparam}).
-
-Results indicate that dependence of data is different in some bins than would be expected under a 
-null hypothesis of a normal copula, answering the second part of question 1.
-Table \ref{SM-tab_BMR_asym} indicates stronger upper tail dependence for the BMR data copula which is significantly (95% CI) different from that of the symmetric normal copula and thus consistent with model selection results.
-
-### Anolis data \label{Ex_contrasts}
-
-```{r read_BivMS_contrasts_67_data,echo=F}
-BivMS_contrasts_67<-readRDS("./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
-```
-
-#### Model selection approach \label{Ex_contrasts_Model_selection}
-
-The test of indepedence gave $p=0$  (to within the precision available 
-from `BiCopIndTest` in the `VineCopula` package). We fitted several copulas
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_contrasts_67fit}).
-The table shows that several copulas (mostly J and SC) are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of question 1 for these data.
-
-The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_contrasts_67$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_contrasts_67$relUTdep_AICw`.
-These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of question 1 for these data.
-Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_contrasts_67$relLTdep_AICw-BivMS_contrasts_67$relUTdep_AICw`, was different from 0, answering the third part of question 1.
-
-These results do convincingly show non-Normal copula structure, giving $p$-value `r BivMS_contrasts_67$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_contrasts_67$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
-The next section looks at tail dependence in a non-parametric way.
-
-#### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}
-
-As described in \nameref{Methods}, each of our statistics, are computed for $(l_b,u_b)$ in the 
-ranges $0-0.25$, $0.25-0.5$, $0.5-0.75$, $0.75-1$ (with $4$ number of bins).
-Anolis data set has lesser number of data compared to other two bivariate data set, so fewer number of bins were chosen to measure tail-dependence.
-Those statistics were compared to the distributions of values of the same statistic with the same ranges computed on 1000 normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different
-runs for Kendall and Spearman (Fig. \ref{SM-fig_contrasts_67_nonparam}).
-
-Results indicate that tail dependence of 
-data is stronger in both the lower and upper tails than would be expected under a 
-null hypothesis of a normal copula, answering the second part of question 1.
-
-We also tested statistically for asymmetry of tail dependence, which seems visible in Fig.
-\ref{SM-fig_raw_cop_AVG_humerus_ltoe}(B).
-Results (Table \ref{SM-tab_contrasts_67_asym}) show that upper-tail dependence was, indeed, significantly stronger than lower-tail dependence in these data, answering the third part of question 1.
-
-### Abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
-
-```{r read_RES_aphid_count, echo=F}
-RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
-```
-
-#### Model selection approach \label{Ex_count_Model_selection}
-
-As described in Methods, model selection among copulas was done separately for each pair of the 10 
-sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns 
-corresponding to locations. Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location 
-pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of 
-question 1 for these data. Comparing the 
-AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) 
-to the AIC weight 
-of the normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for normal copulas was usually not
-comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
-failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
-the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}),
-demonstrating that the collection of copula families we used was sufficiently broad that the copulas of 
-all pairs of locations could be characterized by at least one of them.
-In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-
-The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively.
-These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data.
-Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.
-
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question 1.
-
-#### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
-
-Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
-computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
-resampling scheme, as described in Methods (Table \ref{SM-tab_count_resamp}). Confidence intervals of
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
-indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
-exluded $0$ and were positive, indicating asymmetry of tail dependence, a feature that data with normal copula 
-structure do not have, providing further evidence answering the third part of question 1.
-
-Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig.
-\ref{SM-fig_aphid_count} A,B) showed the typical decline of correlation with
-distance between sampling locations that has been seen for numerous other datasets in earlier work. Plotting 
-$\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig.
-\ref{SM-fig_aphid_count} C-F).
-Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows
-these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig.
-\ref{SM-fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical
-result of Table \ref{SM-tab_count_resamp} that averages of these quantities across all pairs of sampling locations
-were significantly greater than zero, demonstrating asymmetry of tail dependence.
-Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) package.
-
-### First flight date data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}
-
-```{r read_RES_aphid_ff,echo=F}
-RES_aphid_ff<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_firstflight_species_11.RDS")
-```
-
-#### Model selection approach \label{Ex_ff_Model_selection}
-
-As in section \nameref{Ex_count_Model_selection}, results are presented in $10 \times 10$ matrix format with rows
-and columns corresponding to locations.
-Table \ref{SM-tab_ff_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data.
-Comparing the AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_ff_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_ff_AICw_normal}) confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_ff_p_CvM}) usually failed to reject the
-hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for 
-goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_ff_p_KS}).
-In all cases at least `r min((RES_aphid_ff$gfc_numBS_success/RES_aphid_ff$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for
-goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-
-The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and 
-the model-averaged (again, AIC based) upper-tail dependence statistic are in 
-Tables \ref{SM-tab_ff_LTdep_AICw} and \ref{SM-tab_ff_UTdep_AICw}, respectively.
-These values were exclusively positive, thereby differing from what a normal copula would give (i.e., $0$), and answering the second part of question 1 for these data - these data have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
-
-The difference between the model-averaged lower-tail dependence statistic and the 
-model-averaged upper-tail dependence statistic (lower minus upper), shown in 
-Table \ref{SM-tab_ff_LTmUT}, was negative for `r sum((RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw)<0,na.rm=T)` 
-out of `r (dim(RES_aphid_ff$gfc_p_CvM)[1]^2-dim(RES_aphid_ff$gfc_p_CvM)[1])-2*RES_aphid_ff$num_indep_loc_pair` location pairs, answering the third part of question 1.
-We note this result is opposite the analogous result for aphid abundance
-data for the green spruce aphid, i.e., whereas those abundance data showed stronger lower tail dependendence than upper tail dependence, first flights for the leaf-curling plum aphid showed stronger upper tail dependence than lower tail dependence.
-
-#### Non-parametric analysis of tail dependence  \label{Ex_ff_NPA}
-
-Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman 
-correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were 
-computed via a spatial resampling scheme, as described in \nameref{Methods} (Table \ref{SM-tab_ff_resamp}).
-Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded 
-zero and mean values were positive, indicating tail dependence.
-Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always exluded $0$ and were negative, indicating asymmetry of tail dependence, a feature
-that data with normal copula structure do not have, providing further evidence answering the third part of question 1.
-The contrast here with the analogous results for abundances of the green spruce aphid is again illuminating.
-
-Plotting Spearman or Kendall correlations between first flight time series for pairs of locations as a function of great circle distance between the locations (Fig.
-\ref{SM-fig_aphid_ff} A, B) showed the typical (though slight, in this case) decline of correlation with increasing distance.
-Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear in most cases to decline slightly with distance (Fig.
-\ref{SM-fig_aphid_ff} C-F).
-$\Dsq_u$ and $\Dsq_l$ appear to slightly increase with distance, as expected given their
-opposite meaning (Fig. \ref{SM-fig_aphid_ff} G-H). 
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
-\ref{SM-fig_aphid_ff} I-K) did not show a decline with distance, but did demonstrate visually the statistical result of Table \ref{SM-tab_ff_resamp} that averages of these quantities across all pairs of sampling locations were significantly less than zero, showing asymmetry of tail dependence.
-
-### Abundance data for *Ceratium furca* in UK seas \label{Ex_pns}
-
-```{r read_RES_plankton_north_sea,echo=F}
-RES_plankton_north_sea<-readRDS("./Results/fitting_results/Plankton_North_Sea_CopulaFit_selecloc_species_16.RDS")
-```
-
-#### Model selection approach \label{Ex_pns_Model_selection}
-
-As described in Methods and as in previous sections, model selection among copulas was done separately for 
-each pair of sampling locations.
-There were $14$ sampling locations, so results are presented in $14 \times 14$ matrix format, with rows and columns corresponding to locations.
-Table \ref{SM-tab_pns_bestcop} shows that, for the large majority of location pairs, some non-normal copula was a better fit, according to AIC, than was the normal copula, answering the first part of question 1 for these data.
-Comparing the AIC weight of the lowest-AIC copula family for each pair of sampling locations (Table \ref{SM-tab_pns_AICw}) to the AIC weight of the normal copula family for the same pair (Table \ref{SM-tab_pns_AICw_normal})
-confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_pns_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_pns_p_KS}).
-In all cases at least
-`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-
-The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_pns_LTdep_AICw} and \ref{SM-tab_pns_UTdep_AICw}, respectively.
-These values were exclusively positive, thereby
-differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
-have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
-
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
-(lower minus upper), shown in Table \ref{SM-tab_pns_LTmUT}, was positive for some
-location pairs and negative for others, but was positive for 
-`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`
-locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
-
-#### Non-parametric analysis of tail dependence  \label{Ex_pns_NPA}
-
-Mean values of our statistics, as well as mean Kendall and Spearman
-correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
-spatial resampling scheme, as described in Methods (Table \ref{SM-tab_pns_resamp}).
-Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
-always excluded $0$ and mean values were positive, indicating tail dependence.
-Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
-also all were strictly positive, indicating asymmetry of tail dependence, a feature that data with normal copula
-structure do not have, providing further evidence answering the third part of question 1.
-
-Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
-between the locations (Fig. \ref{SM-fig_pns} A, B) showed the typical decline of correlation with increasing
-distance.
-Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig.
-\ref{SM-fig_pns} C-F).
-$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given their opposite meaning (Fig.
-\ref{SM-fig_pns} G-H).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
-\ref{SM-fig_pns} I-K) did not show a decline with distance, but did demonstrate visually the
-statistical result of Table \ref{SM-tab_pns_resamp} that averages of these 
-quantities across all pairs of sampling locations were significantly greater 
-than zero, showing asymmetry of tail dependence.
-
-### Methane flux data \label{Ex_methane}
-
-```{r read_RES_methane,echo=F}
-RES_methane<-readRDS("./Results/fitting_results/MethaneFluxFit_selecloc_data_pt_thrs_50.RDS")
-```
-
-#### Model selection approach \label{Ex_methane_Model_selection}
-
-As described in Methods and as in previous sections, model selection among copulas was done separately for 
-each pair of sampling locations. There were $13$ sampling locations, so results are presented in 
-$13 \times 13$ matrix format, with rows and columns corresponding to locations. 
-Table \ref{SM-tab_methane_bestcop} shows that, for the large majority of location pairs, some 
-non-normal copula was a better fit, according to AIC, than was the normal copula, answering the 
-first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula 
-family for each pair of sampling locations (Table \ref{SM-tab_methane_AICw}) to the AIC weight of
-the normal copula family for the same pair (Table \ref{SM-tab_methane_AICw_normal}) 
-confirms that the degree of support for normal copulas was usually not comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_methane_p_CvM}) usually failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_methane_p_KS}).
-In all cases at least
-`r min((RES_methane$gfc_numBS_success/RES_methane$gfc_numBS),na.rm=T)*100` percent of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-
-The model-averaged (averaging performed using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistics are in Tables \ref{SM-tab_methane_LTdep_AICw} and \ref{SM-tab_methane_UTdep_AICw}, respectively.
-These values were exclusively positive, thereby
-differing from what a normal copula would give (i.e., 0), and answering the second part of question 1 for these data - these data
-have greater tail dependence (lower and upper) than would be expected from a normal-copula null hypothesis.
-
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic
-(lower minus upper), shown in Table \ref{SM-tab_methane_LTmUT}, was positive for some
-location pairs and negative for others, but was positive for 
-`r sum((RES_methane$LTdep_AICw-RES_methane$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_methane$gfc_p_CvM)[1]^2-dim(RES_methane$gfc_p_CvM)[1])-2*RES_methane$num_indep_loc_pair`
-locations pairs. Thus lower tail dependence tended to be stronger than upper tail dependence, answering the third part of question 1.
-
-#### Non-parametric analysis of tail dependence  \label{Ex_methane_NPA}
-
-Mean values of our statistics, as well as mean Kendall and Spearman
-correlations, were computed across all pairs of sampling locations, and confidence intervals of these means were computed via a
-spatial resampling scheme, as described in Methods (Table \ref{SM-tab_methane_resamp}).
-Confidence intervals of $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$
-always excluded $0$ and mean values were positive, indicating tail dependence.
-But the confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ were marginal, indicating no significant asymmetry of tail dependence.
-
-Plotting Spearman or Kendall correlations between time series for pairs of locations as a function of great circle distance
-between the locations (Fig. \ref{SM-fig_methane} A, B) showed the typical decline of correlation with increasing
-distance.
-Plotting $\cor_l$ , $\cor_u$ , $\Ps_l$, and $\Ps_u$ against distances between pairs of locations shows these quantities also appear to decline with distance (Fig.
-\ref{SM-fig_methane} C-F).
-$\Dsq_u$ and $\Dsq_l$ appear to increase with distance, as expected given
-their opposite meaning (Fig.
-\ref{SM-fig_methane} G-H).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ against distance (Fig.
-\ref{SM-fig_methane} I-K) did not show a decline with distance, but did demonstrate visually the
-statistical result of Table \ref{SM-tab_methane_resamp} that averages of these 
-quantities across all pairs of sampling locations were more or less equally dispersed around the zero-line, showing no significant asymmetry of tail dependence.
--->
-
-<!--
-# Question 2 analyses 
-## The Moran effect and copula structure in populations 
-           ----moved to Paper.Rmd----
--->
-
 ```{r cause4copula_pGOF_500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 \begin{center}
-\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf}
-\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{500} points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF_500}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
 \includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf}
@@
 \begin{center}
-\includegraphics[width=16 cm]{./Results/Cause4copula_results/Clayton_kendall.pdf}
-\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_C_kend}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
 \includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_spearman.pdf}
@@
 Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_SC_spear}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_kendall.pdf}
-\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}}
 \end{center}
@@
 
-\newpage
-
 ```{r tab_temp_aphid_AICw, echo=F, results='markup'}
@@
 
+\newpage
 ```{r tab_temp_aphid_LTdep_AICw, echo=F, results='markup'}
@@
 
+\newpage
 ```{r tab_temp_aphid_UTdep_AICw, echo=F, results='markup'}
@@
 
+\newpage
 ```{r tab_temp_aphid_LTmUT, echo=F, results='markup'}
@@
 
-# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}
-
-To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
-\ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.
-
-\begin{equation}\label{eq.AR1}
-P_{i}(t+1)=\beta P_{i}(t)+\alpha \epsilon_{i}(t)
-\end{equation}
-
-To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$.
-We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches.
-Should their dependence be exactly the same as of the noises or not ?
-
-To test that our approach is of two way : first, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
-Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome which was a bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
-\ref{fig_Cause4copula_params_GOF_500}).
-We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size and reported p-values for both statistics (Cramer-von Mises and Kolmogorov-Smirnov).
-p-values greater or equal to $0.05$ (above the dashed line in Fig.
-\ref{fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters.
-Fig.
-\ref{fig_Cause4copula_params_GOF_2500} represents similar feature as of Fig.
-\ref{fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$.
-(For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
-
-Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \nameref{MT-NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
-Then we plotted mean of each measured quantity with standard error against Spearman's Rho (or Kendall's Tau) of the noise copula family.
-Fig.
-\ref{fig_Cause4copula_C_spear} and Fig.
-\ref{fig_Cause4copula_C_kend} represent this picture for noise generating from a lower tail dependent Clayton copula whereas for Fig.
-\ref{fig_Cause4copula_SC_spear} and Fig.
-\ref{fig_Cause4copula_SC_kend} noise came from a upper tail dependent Survival Clayton family.
-We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach, here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise.
-(For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
+<!--# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}-->
 
@@
 
+\newpage
+
 ```{r tab_stat_npa_asym_5,echo=F,results="markup"}
@@
 
+\newpage
 ```{r tab_stat_npa_asym_6,echo=F,results="markup"}
@@
 
+\newpage
 ```{r tab_stat_npa_asym_7,echo=F,results="markup"}
@@
 \newpage
-
-<!--
-# Question 3 analyses for each dataset and theoretical analyses
-Recall that question 3 is: What are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
-## Skewness of the spatial mean
-### Aphid count \label{skewness_count}
-### Aphid first flight \label{skewness_ff}
-### UK Sea plankton \label{skewness_pns}
-### Methane-flux data \label{skewness_methane}
-                   ----moved to Paper.Rmd----
--->
 

### 2018-10-01 Paper.Rmd
modified

@@
 Spatial mean of the time series data extracted that tail dependence from the multivariate data set, if any, and reflected in the averaged skewness.
-We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula (for details see Appendix \ref{SM-Consequence_skewness}).
+Because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, changed significantly from the individual skewness of time-series data.
+To test null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure, we generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
+Then we calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R).
+We made a histogram of surragate skewness and found if actual skewness were within $95\%$ quantile or not.
 If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).
-Figure \ref{fig_skewness_methane_maintext} showed asymmetric copula structure's effect on spatial skewness for biogeochemical methane-flux data.
-The result, showing significantly different skewness of spatial mean from that of Normal surrogates, was consistent with model selection and nonparametric approach which supported lower tail dependence copula for methane-flux data.
 
-We then showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}).
-Finally, we studied **(iii)** the effects of non-Normal copula on extinction risk with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}).
+We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}).
+For example, our model simulation showed lower tail dependent Clayton copula had much higher slope (spatial-Taylor's law) than symmetric Normal copula, which inturn had higher slope than upper tail dependent Survival Clayton family (Fig.
+\ref{SM-fig_spTL}).
+Finally, we studied **(iii)** the effects of non-Normal noise copula on extinction risk (calculated after 25 years) with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}).
+We considered patches were arranged in a linear chain fashion and the population may disperse to its nearerst two neighbors (local dispersion) or to any patches (global dispersion).
+For both model with both dispersion, increasing number of habitat patches and increasing carrying capacity, separately, had decreased extinction risk of the populations which was expected (Fig.
+\ref{SM-fig_risk_numlocs_effect_local_disp} - Fig.
+\ref{SM-fig_risk_K_effect}).
+For Lewton-Cohen model, we found greater extinction risk if populations were affected by left tail dependence environmental noise copula than right tail dependent one.
+For Ricker model, we got the similar observation for smaller values of $r$, the intrinsic growth rate, but after $r=0.7$, there was a turn-over in extinction risks, and right tail dependent noise copula caused population go extinct faster.
 
 \begin{figure}[!h]
+\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
+\vspace{-0.1 cm}
 \begin{center}
-\includegraphics[width=16 cm]{./Results/skewness_results/skewness_methane/skewness_methane.pdf}
-\caption[Skewness-methane-flux data]{Skewness of spatially averaged methane-flux data (red line) was compared with respect to that of 10000 Normal surrogates preserving $(A)$ Kendall correlation, and $(B)$ Spearman correlation.
-p-value < 0.05 in both case showed significant difference in skewness (as expected for lower tail dependent methane flux data from model selection results) from Normal surrogates.
-\label{fig_skewness_methane_maintext}}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_kendall.pdf}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_kendall.pdf}\\
+\end{center}
+\vspace{0.2 cm}
+\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
+\vspace{-0.1 cm}
+\begin{center}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_kendall.pdf}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_kendall.pdf}
+\caption{Skewness of spatially averaged $(A)$ Green-spruce aphid's abundance, $(B)$ Leaf-plum curling aphid's firstflight date, $(C)$ Ceratium furca's abundance and $(D)$ methane-flux data, were indicated by red line in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Kendall correlation).
+\label{fig_skewness_kendall}}
 \end{center}
 \end{figure}
+
+## Application : Change of skewness
+
+we used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data, to compute skewness of spatially averaged data.
+Figure \ref{fig_skewness_kendall} represents the results for each dataset with Kendall correlatiomn preserving surrogates.
+We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
+If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).
+
+Figure \ref{fig_skewness_kendall}(A) showed that averaged skewness was much lower than expected if the aphid count data came from a normal copula.
+In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_kendall} A).
+As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results.
+On the other hand, Figure \ref{fig_skewness_kendall}(B) demonstrated that averaged skewness was much higher than expected if the aphid first flight data came from a normal copula.
+As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red line in Figure \ref{fig_skewness_kendall} B) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence.
+Figure \ref{fig_skewness_kendall}(C) showed that averaged skewness was lower than expected if the plankton data came from a normal copula.
+Our previous results on non-parametric analysis and model fitting also supported lower tail dependence.
+But, as $p>0.05$, we could not reject $H_{0}$ in that case.
+Finally, for methane-flux data, the averaged skewness was lower than expected if the data came from a normal copula (Figure \ref{fig_skewness_kendall} D).
+As $p<0.05$, we rejected $H_{0}$ and that result was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence.
+
+We found the same result when we repeated our simulation with Spearman correlation preserving surrogates (see Fig. \ref{SM-fig_skewness_spearman}).
 

### 2018-10-01 SupportingInformation.Rmd
modified

@@
 names(temp_aphid)<-"aphid_winter_neg_temp"
+saveRDS(temp_aphid,"./Results/temp_aphid_winter_neg_temp_avg_data.RDS")
 ```
@@
 
-pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf",height=6,width=16)
-par(mfrow=c(1,2))
 numsurrog<-10000
 loclist<-good_loclist(d_allsp=data_aphid_count,sp=10,data_pt_thrs=30)
+
+pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_kendall.pdf",height=6,width=8)
+op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
 answer1k<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
                           corpres="kendall") 
-
 p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
-
-mtext(paste0("(A) p = ",p_left," (for Kendall preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
-
+mtext(paste0("p = ",p_left),side=3,line=-0.5,col="navyblue",cex = 2)
+par(op)
+dev.off()
+
+pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf",height=6,width=8)
+op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
 answer1s<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog
-
-mtext(paste0("(B) p = ",p_left," (for Spearman preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
-
-dev.off()
-```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf}
-\caption[Skewness count for species Green spruce aphid]{Skewness count for species Green spruce aphid\label{fig_skewness_count}}
-\end{center}
-\end{figure}
+mtext(paste0("p = ",p_left),side=3,line=-0.5,col="navyblue",cex = 2)
+par(op)
+dev.off()
+```
 
@@
 
-pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf",height=6,width=16)
-par(mfrow=c(1,2))
 numsurrog<-10000
@@
 
+pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_kendall.pdf",height=6,width=8)
+op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
 answer2k<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 p_right<-sum(answer2k$surrogskw > answer2k$realskw)/numsurrog
-mtext(paste0("(A) p = ",p_right," (for Kendall preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
-
+mtext(paste0("p = ",p_right),side=3,line=-0.5,col="navyblue",cex = 2)
+par(op)
+dev.off()
+
+pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf",height=6,width=8)
+op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
 answer2s<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 p_right<-sum(answer2s$surrogskw > answer2s$realskw)/numsurrog
-mtext(paste0("(B) p = ",p_right," (for Spearman preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
-
-dev.off()
-```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf}
-\caption[Skewness firstflight for species Leaf-curling plum aphid]{Skewness firstflight for species Leaf-curling plum aphid\label{fig_skewness_ff}}
-\end{center}
-\end{figure}
+mtext(paste0("p = ",p_right),side=3,line=-0.5,col="navyblue",cex = 2)
+par(op)
+dev.off()
+```
 
@@
 
-pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf",height=6,width=16)
-par(mfrow=c(1,2))
 numsurrog<-10000
 loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=16,data_pt_thrs=45)
+
+pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_kendall.pdf",height=6,width=8)
+op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
 answer3k<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
                           corpres="kendall") 
-
 p_left<-sum(answer3k$surrogskw < answer3k$realskw)/numsurrog
-
-mtext(paste0("(A) p = ",p_left," (for Kendall preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
-
+mtext(paste0("p = ",p_left),side=3,line=-0.5,col="navyblue",cex = 2)
+par(op)
+dev.off()
+
+pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf",height=6,width=8)
+op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
 answer3s<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
                           corpres="spearman") 
-
 p_left<-sum(answer3s$surrogskw < answer3s$realskw)/numsurrog
-
-mtext(paste0("(B) p = ",p_left," (for Spearman preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
-
-dev.off()
-```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf}
-\caption[Skewness-plankton for species Ceratium furca]{Skewness-plankton for species Ceratium furca\label{fig_skewness_pns}}
-\end{center}
-\end{figure}
+mtext(paste0("p = ",p_left),side=3,line=-0.5,col="navyblue",cex = 2)
+par(op)
+dev.off()
+```
 
@@
 
-pdf("./Results/skewness_results/skewness_methane/skewness_methane.pdf",height=6,width=16)
-par(mfrow=c(1,2))
 numsurrog<-10000
@@
 
+pdf("./Results/skewness_results/skewness_methane/skewness_methane_kendall.pdf",height=6,width=8)
+op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
 answer1k<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
-mtext(paste0("(A) p = ",p_left," (for Kendall preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
-
+mtext(paste0("p = ",p_left),side=3,line=-0.5,col="navyblue",cex = 2)
+par(op)
+dev.off()
+
+pdf("./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf",height=6,width=8)
+op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
 answer1s<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog
-mtext(paste0("(B) p = ",p_left," (for Spearman preserving Normal surrogates)"),side=3,line=0.2,col="navyblue")
-
+mtext(paste0("p = ",p_left),side=3,line=-0.5,col="navyblue",cex = 2)
+par(op)
 dev.off()
@@
 \begin{figure}[!h]
+\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
+\vspace{-0.1 cm}
 \begin{center}
-\includegraphics[width=16 cm]{./Results/skewness_results/skewness_methane/skewness_methane.pdf}
-\caption[Skewness-methane-flux data]{Skewness-methane-flux data\label{fig_skewness_methane}}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
+\end{center}
+\vspace{0.2 cm}
+\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
+\vspace{-0.1 cm}
+\begin{center}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
+\caption{Skewness of spatially averaged $(A)$ green-spruce aphid's abundance, $(B)$ leaf-curling plum aphid's firstflight date, $(C)$ ceratium furca's abundance and $(D)$ methane-flux data, were indicated by red line in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation).
+\label{fig_skewness_spearman}}
 \end{center}
@@
 n<-25 #number of sampling locations
-numdat<-10 #the number of datasets to get from the normal copula
+#numdat<-10 #the number of datasets to get from the normal copula
+numdat<-1000 #the number of datasets to get from the normal copula
 rho<-0.7 #normal-copula parameter for dependence between locations
@@
  ss<-spea.surrog[[counter]]
- kend.stats.res[[counter]]<-matrix(NA,5,numdat) # to keep spatial TL stats
- spea.stats.res[[counter]]<-matrix(NA,5,numdat)
+ kend.stats.res[[counter]]<-matrix(NA,6,numdat) # to keep spatial TL stats
+ spea.stats.res[[counter]]<-matrix(NA,6,numdat)
  kend.points.res[[counter]]<-array(NA,c(N,2,numdat)) # to keep log(mu), log(var) of spatial TL as a dataframe
@@
 <!--Now make some plots, these one for kendall surrogates-->
-```{r taylor_plots, echo=F, results="hide"}
+```{r taylor_plots_kendall, echo=F, results="hide"}
 xlimits<-range(kend.points.res[[1]][,1,],kend.points.res[[2]][,1,],kend.points.res[[3]][,1,],kend.points.res[[4]][,1,])
@@
 
-pdf("./Results/taylorslaw_results/SpatialTL_CNFSC.pdf",height=4,width=16)
+pdf("./Results/taylorslaw_results/SpatialTL_CNFSC_kendall.pdf",height=4,width=16)
 op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
@@
 s.slope<-cbind(kend.stats.res[[1]][5,],kend.stats.res[[2]][5,],kend.stats.res[[3]][5,],kend.stats.res[[4]][5,])
-
-pdf("./Results/taylorslaw_results/SpatialTL_5stats.pdf",height=3,width=15)
-op<-par(mfrow=c(1,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+s.quadcoeff<-cbind(kend.stats.res[[1]][6,],kend.stats.res[[2]][6,],kend.stats.res[[3]][6,],kend.stats.res[[4]][6,])
+
+pdf("./Results/taylorslaw_results/SpatialTL_6stats_kendall.pdf",height=3,width=18)
+op<-par(mfrow=c(1,6),mar=c(4.2,5,1,1),mgp=c(3,1,0))
 #***DAN: x axis needs to be redone to show the copula names used in the surrogates
 
-x<-1:4
+#x<-1:4
 x2<-c("C", "N", "F", "SC")
-plot(x,s.lin[1,],type="l",col="grey",ylim=range(s.lin),xlab="Set of surrogates",ylab="Linearity test p-vals",
-     cex.axis=1.5,cex.lab=1.5,xaxt="n")
-points(x,s.lin[1,],col="red")
+#plot(x,s.lin[1,],type="l",col="grey",ylim=range(s.lin),xlab="Set of surrogates",ylab="Linearity test p-vals",
+#     cex.axis=1.5,cex.lab=1.5,xaxt="n")
+#points(x,s.lin[1,],col="red")
+#axis(1, at = c(1:4), x2, cex.axis=1.5)
+#for (counter in 2:dim(s.lin)[1]){
+#  lines(x,s.lin[counter,],col="grey")
+#  points(x,s.lin[counter,],col="red")
+#}
+boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="Linearity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
 axis(1, at = c(1:4), x2, cex.axis=1.5)
-for (counter in 2:dim(s.lin)[1]){
-  lines(x,s.lin[counter,],col="grey")
-  points(x,s.lin[counter,],col="red")
-}
 
 #plot s.hom distr for each set of surrogates
-#***DAN: x axis needs to be redone to show the copula names used in the surrogates
-x<-1:4
-plot(x,s.hom[1,],type="l",col="grey",ylim=range(s.hom),xlab="Set of surrogates",
-     ylab="Homoskedasticity test p-vals",
-     cex.axis=1.5,cex.lab=1.5,xaxt="n")
-points(x,s.hom[1,],col="red")
+boxplot(s.hom , col=rainbow(4), xlab="Copula family", ylab="Homoskedasticity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
 axis(1, at = c(1:4), x2, cex.axis=1.5)
-for (counter in 2:dim(s.slope)[1]){
-  lines(x,s.hom[counter,],col="grey")
-  points(x,s.hom[counter,],col="red")
-}
 
 #plot s.rmse distr for each set of surrogates
-x<-1:4
-plot(x,s.rmse[1,],type="l",col="grey",ylim=range(s.rmse),xlab="Set of surrogates",ylab="Spatial TL rmse",
-     cex.axis=1.5,cex.lab=1.5,xaxt="n")
-points(x,s.rmse[1,],col="red")
+boxplot(s.rmse , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL rmse", cex.axis=1.5,cex.lab=1.5,xaxt="n")
 axis(1, at = c(1:4), x2, cex.axis=1.5)
-for (counter in 2:dim(s.rmse)[1]){
-  lines(x,s.rmse[counter,],col="grey")
-  points(x,s.rmse[counter,],col="red")
-}
 
 #plot s.int distr for each set of surrogates
-#***DAN: x axis needs to be redone to show the copula names used in the surrogates
-x<-1:4
-plot(x,s.int[1,],type="l",col="grey",ylim=range(s.int),xlab="Set of surrogates",
-     ylab="Spatial TL intercept",
-     cex.axis=1.5,cex.lab=1.5,xaxt="n")
-points(x,s.int[counter,],col="red")
+boxplot(s.int , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL intercept", cex.axis=1.5,cex.lab=1.5,xaxt="n")
 axis(1, at = c(1:4), x2, cex.axis=1.5)
-for (counter in 2:dim(s.int)[1]){
-  lines(x,s.int[counter,],col="grey")
-  points(x,s.int[counter,],col="red")
-}
 
 #plot s.slope distr for each set of surrogates
-#***DAN: x axis needs to be redone to show the copula names used in the surrogates
-x<-1:4
-plot(x,s.slope[1,],type="l",col="grey",ylim=range(s.slope),xlab="Set of surrogates",
-     ylab="Spatial TL slope",
-     cex.axis=1.5,cex.lab=1.5,xaxt="n")
-points(x,s.slope[counter,],col="red")
+boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL slope", cex.axis=1.5,cex.lab=1.5,xaxt="n")
 axis(1, at = c(1:4), x2, cex.axis=1.5)
-for (counter in 2:dim(s.slope)[1]){
-  lines(x,s.slope[counter,],col="grey")
-  points(x,s.slope[counter,],col="red")
-}
+
+#plot s.quadcoeff distr for each set of surrogates
+boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL quadratic coefficient", cex.axis=1.5,cex.lab=1.5,xaxt="n")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+
+par(op)
+dev.off()
+```
+
+<!--Now make some plots, these one for spearman surrogates-->
+```{r taylor_plots_spearman, echo=F, results="hide"}
+xlimits<-range(spea.points.res[[1]][,1,],spea.points.res[[2]][,1,],spea.points.res[[3]][,1,],spea.points.res[[4]][,1,])
+ylimits<-range(spea.points.res[[1]][,2,],spea.points.res[[2]][,2,],spea.points.res[[3]][,2,],spea.points.res[[4]][,2,])
+
+pdf("./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.pdf",height=4,width=16)
+op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
+##plot TL for clayton
+plot(spea.points.res[[1]][,1,],spea.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
+     cex.axis=1.5,cex.lab=1.5)  
+points(spea.points.res[[1]][,1,1],spea.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
+points(spea.points.res[[1]][,1,2],spea.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
+mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black")
+
+##plot TL for normal copula
+plot(spea.points.res[[2]][,1,],spea.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
+     cex.axis=1.5,cex.lab=1.5)  
+points(spea.points.res[[2]][,1,1],spea.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
+points(spea.points.res[[2]][,1,2],spea.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
+mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black")
+
+#plot TL for frank
+plot(spea.points.res[[3]][,1,],spea.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
+     cex.axis=1.5,cex.lab=1.5)  
+points(spea.points.res[[3]][,1,1],spea.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
+points(spea.points.res[[3]][,1,2],spea.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
+mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black")
+
+#plot TL for survival clayton
+plot(spea.points.res[[4]][,1,],spea.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+     xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
+     cex.axis=1.5,cex.lab=1.5)  
+points(spea.points.res[[4]][,1,1],spea.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
+points(spea.points.res[[4]][,1,2],spea.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
+mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black")
+
+par(op)
+dev.off()
+#---------------------------------------------------
+#prepare for plots coming below
+s.lin<-cbind(spea.stats.res[[1]][1,],spea.stats.res[[2]][1,],spea.stats.res[[3]][1,],spea.stats.res[[4]][1,])
+s.hom<-cbind(spea.stats.res[[1]][2,],spea.stats.res[[2]][2,],spea.stats.res[[3]][2,],spea.stats.res[[4]][2,])
+s.rmse<-cbind(spea.stats.res[[1]][3,],spea.stats.res[[2]][3,],spea.stats.res[[3]][3,],spea.stats.res[[4]][3,])
+s.int<-cbind(spea.stats.res[[1]][4,],spea.stats.res[[2]][4,],spea.stats.res[[3]][4,],spea.stats.res[[4]][4,])
+s.slope<-cbind(spea.stats.res[[1]][5,],spea.stats.res[[2]][5,],spea.stats.res[[3]][5,],spea.stats.res[[4]][5,])
+s.quadcoeff<-cbind(spea.stats.res[[1]][6,],spea.stats.res[[2]][6,],spea.stats.res[[3]][6,],spea.stats.res[[4]][6,])
+
+pdf("./Results/taylorslaw_results/SpatialTL_6stats_spearman.pdf",height=3,width=18)
+op<-par(mfrow=c(1,6),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+x2<-c("C", "N", "F", "SC")
+
+#plot s.lin distr for each set of surrogates
+boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="Linearity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+
+#plot s.hom distr for each set of surrogates
+boxplot(s.hom , col=rainbow(4), xlab="Copula family", ylab="Homoskedasticity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+
+#plot s.rmse distr for each set of surrogates
+boxplot(s.rmse , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL rmse", cex.axis=1.5,cex.lab=1.5,xaxt="n")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+
+#plot s.int distr for each set of surrogates
+boxplot(s.int , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL intercept", cex.axis=1.5,cex.lab=1.5,xaxt="n")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+
+#plot s.slope distr for each set of surrogates
+boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL slope", cex.axis=1.5,cex.lab=1.5,xaxt="n")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+
+#plot s.quadcoeff distr for each set of surrogates
+boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL quadratic coefficient", cex.axis=1.5,cex.lab=1.5,xaxt="n")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+
 par(op)
@@
 \begin{center}
-\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC.pdf}\\
+\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_kendall.pdf}\\
 \end{center}
-\textbf{ \hspace{3.1 cm} (E) \hspace{2.4 cm} (F) \hspace{2.3 cm} (G) \hspace{2.3 cm} (H) \hspace{2.3 cm} (I)}\\
+\textbf{ \hspace{2.8 cm} (E) \hspace{1.8 cm} (F) \hspace{1.8 cm} (G) \hspace{1.8 cm} (H) \hspace{1.6 cm} (I) \hspace{2 cm} (J)}\\
 \vspace{-0.5 cm}
 \begin{center}
-\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_5stats.pdf}
+\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_6stats_kendall.pdf}
 \caption[Spatial Taylor's Law results]{Spatial Taylor's Law results: log(spatial variance) vs.
-log(spatial mean) for $ns=10$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of homoskedasticity, (G) root mean square error, (H) intercept, (I) slope of the first panel plots.
+log(spatial mean) for $ns=10$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of homoskedasticity, (G) root mean square error, (H) intercept, (I) slope, (J) quadratic coefficient of regression line for first panel plot.
 \label{fig_spTL}}
@@
 
-# Consequence of non-Normal copula structure : effect on skewness of the spatial mean \label{Consequence_skewness}
-
-Because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, changed significantly from the individual skewness of time-series data.
-To test null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure, we generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
-Then we calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R).
-We made a histogram of surragate skewness and found if actual skewness were within $95\%$ quantile or not.
-If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).
-
-To verify that we used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data.
-Figure \ref{fig_skewness_count} showed that averaged skewness was much lower than expected if the aphid count data came from a normal copula.
-In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_count}).
-As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results.
-On the other hand, Figure \ref{fig_skewness_ff} demonstrated that averaged skewness was much higher than expected if the aphid first flight data came from a normal copula.
-As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red line in Figure \ref{fig_skewness_ff}) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence.
-Figure \ref{fig_skewness_pns} showed that averaged skewness was lower than expected if the plankton data came from a normal copula.
-Our previous results on non-parametric analysis and model fitting also supported lower tail dependence.
-But, as $p>0.05$, we could not reject $H_{0}$ in that case.
-Finally, for methane-flux data, the averaged skewness was lower than expected if the data came from a normal copula (Figure \ref{fig_skewness_methane}).
-As $p<0.05$, we rejected $H_{0}$ and that result was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence.
-
 # Consequence of non-Normal copula structure : Effect on Taylor's law \label{Consequence_TL}

### 2018-10-01 Cause4copula.R

@@ -468,7 +468,7 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
     warning("specify method",immediate.=T,call.=T)
   }
   
-  op<-par(mfrow=c(4,3),mar=c(3,3.5,3,3.5), mgp=c(1.5,0.5,0))
+  op<-par(mfrow=c(4,3),mar=c(2.6,3.5,2,3.5), mgp=c(1.5,0.5,0))
   plot(corcoef_list,S_noise_mat[,2],cex=0.5,col="red",xlab=xlabel,ylab="Spearman",xlim=c(0,1),ylim=c(0,1))
   segments(corcoef_list,S_noise_mat[,1],corcoef_list,S_noise_mat[,3],col='red')
   bar_len<-0.02

### 2018-10-01 TaylorsLawTools.R

@@ -25,12 +25,12 @@ tl_stats<-function(m)
 {
   sres<-worker_tl(m)
   s.stats<-sres$stats
-  names(s.stats)<-c("s.lin","s.hom","s.rmse","s.int","s.slope")
+  names(s.stats)<-c("s.lin","s.hom","s.rmse","s.int","s.slope","s.quad.coeff")
   s.points<-sres$points
   
   tres<-worker_tl(t(m))
   t.stats<-tres$stats
-  names(t.stats)<-c("t.lin","t.hom","t.rmse","t.int","t.slope")
+  names(t.stats)<-c("t.lin","t.hom","t.rmse","t.int","t.slope","t.quad.coeff")
   t.points<-tres$points
   
   return(list(s.stats=s.stats,t.stats=t.stats,
@@ -58,7 +58,8 @@ worker_tl<-function(m)
          unname(hhet[5][1,1]), #the p-value for a test of homoskedasticity
          sqrt(mean(resid(mod1)^2)), #the root mean squared error for the TL model
          unname(mod1$coefficients[1]), #the intercept
-         unname(mod1$coefficients[2])) #the slope
+         unname(mod1$coefficients[2]), #the slope
+         unname(mod2$coefficients[2]))  # the coefficient of quadratic term
   
   return(list(stats=res,points=data.frame(log10m=lms,log10v=lvs)))
 }

### 2018-10-01 skewness_multivar_dataset.R

@@ -56,7 +56,7 @@ skewness_testing<-function(ts_matrix,loclist,numsurrog,ploton,corpres){
   surrogskw<-apply(FUN=myskns,X=ts_surrog_avg,MARGIN=2)
   
   if(ploton==T){
-    hist(surrogskw,breaks=100,main="",xlab="surrogate_skewness")
+    hist(surrogskw,breaks=100,main="",xlab="surrogate_skewness",ylab="frequency",cex.lab=2)
     abline(v=realskw,col="red")
   }
   

### 2018-10-02 SupportingInformation.Rmd
adding chunk for skewness on real data calculation : aphids & plankton

@@
 
+```{r skewness_aphidcount_allsp_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+
+source("good_loclist.R")
+source("skewness_multivar_dataset.R")
+set.seed(seed)
+
+numsurrog<-10000
+
+splist<-c(1:20)
+skw_aphid_count_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+colnames(skw_aphid_count_spearman)<-c("realskw","p_left","p_right")
+rownames(skw_aphid_count_spearman)<-paste("sp",splist,sep="")
+pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_spearman.pdf",height=20,width=20)
+op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+for(i in splist){
+  cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="spearman") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_aphid_count_spearman$realskw[i]<-ans$realskw
+  skw_aphid_count_spearman$p_left[i]<-p_left
+  skw_aphid_count_spearman$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_aphid_count_spearman,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_spearman.RDS")
+
+skw_aphid_count_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+colnames(skw_aphid_count_kendall)<-c("realskw","p_left","p_right")
+rownames(skw_aphid_count_kendall)<-paste("sp",splist,sep="")
+pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_kendall.pdf",height=20,width=20)
+op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+for(i in splist){
+  #cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="kendall") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_aphid_count_kendall$realskw[i]<-ans$realskw
+  skw_aphid_count_kendall$p_left[i]<-p_left
+  skw_aphid_count_kendall$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_aphid_count_kendall,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_kendall.RDS")
+```
+
+```{r skewness_aphidff_allsp_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+
+source("good_loclist.R")
+source("skewness_multivar_dataset.R")
+set.seed(seed)
+
+numsurrog<-10000
+
+splist<-c(1:20)
+skw_aphid_ff_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+colnames(skw_aphid_ff_spearman)<-c("realskw","p_left","p_right")
+rownames(skw_aphid_ff_spearman)<-paste("sp",splist,sep="")
+pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_spearman.pdf",height=20,width=20)
+op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+for(i in splist){
+  cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="spearman") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_aphid_ff_spearman$realskw[i]<-ans$realskw
+  skw_aphid_ff_spearman$p_left[i]<-p_left
+  skw_aphid_ff_spearman$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_aphid_ff_spearman,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_spearman.RDS")
+
+skw_aphid_ff_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+colnames(skw_aphid_ff_kendall)<-c("realskw","p_left","p_right")
+rownames(skw_aphid_ff_kendall)<-paste("sp",splist,sep="")
+pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_kendall.pdf",height=20,width=20)
+op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+for(i in splist){
+  cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="kendall") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_aphid_ff_kendall$realskw[i]<-ans$realskw
+  skw_aphid_ff_kendall$p_left[i]<-p_left
+  skw_aphid_ff_kendall$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_aphid_ff_kendall,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_kendall.RDS")
+```
+
+```{r skewness_plankton_allsp_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+
+source("good_loclist.R")
+source("skewness_multivar_dataset.R")
+set.seed(seed)
+
+splist<-c(1:22)
+skw_plankton_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+colnames(skw_plankton_spearman)<-c("realskw","p_left","p_right")
+rownames(skw_plankton_spearman)<-paste("sp",splist,sep="")
+skw_plankton_kendall<-skw_plankton_spearman
+pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_spearman.pdf",height=24,width=24)
+op<-par(mfrow=c(4,6),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+for(i in splist){
+  #cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="spearman") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_plankton_spearman$realskw[i]<-ans$realskw
+  skw_plankton_spearman$p_left[i]<-p_left
+  skw_plankton_spearman$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_plankton_spearman,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_spearman.RDS")
+
+pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=20,width=20)
+op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+for(i in splist){
+  #cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="kendall") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_plankton_kendall$realskw[i]<-ans$realskw
+  skw_plankton_kendall$p_left[i]<-p_left
+  skw_plankton_kendall$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_plankton_kendall,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_kendall.RDS")
+```
+
 ```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}

### 2018-10-03 SupportingInformation.Rmd
modified

--- 2018-10-01 SupportingInformation.Rmd
@@
 saveRDS(temp_aphid,"./Results/temp_aphid_winter_neg_temp_avg_data.RDS")
+```
+
+```{r modelsel_and_npa_aphid_count_allsp, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+
+set.seed(seed)
+library(parallel)
+source("FittingCopula_selective_loc.R")
+RES_all_sp_aphid_count<-mclapply(X=c(1:20),FUN=RES_single_sp,d_allsp=data_aphid_count,
+                                 families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=35,
+                                 good_loc_given=F,good_loc=NA,mc.cores=3)
+saveRDS(RES_all_sp_aphid_count,"./Results/fitting_results/RES_all_sp_aphid_count.RDS")
+
+source("good_loclist.R")
+source("NonParamStat.R")
+resloc<-"./Results/stat_results/stat_aphid_count/"
+longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
+lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)
+
+stat_aphid_count_allsp<-vector("list",20)
+
+for(sp in 1:20){
+  nm<-paste("sp_",sp,sep="")
+  resloc2<-paste(resloc,nm,sep="")
+   if (!dir.exists(resloc2)){
+    dir.create(resloc2)
+  }
+  good_loc<-good_loclist(d_allsp=data_aphid_count,sp=sp,data_pt_thrs=30)
+
+  temp<-multcall(d_allsp=data_aphid_count,sp=sp,lats=lats,longs=longs,
+                 pfname=paste(resloc2,"/","Sp_",sp,sep=''),good_loc=good_loc)
+  stat_aphid_count_allsp[[sp]]<-temp
+}
+saveRDS(stat_aphid_count_allsp,paste(resloc,file="stat_aphid_count_all_sp.RDS",sep=''))
+
+summary_aphid_count_allsp<-as.data.frame(matrix(NA,nrow=20,ncol=8))
+colnames(summary_aphid_count_allsp)<-c("species","#LTmUT>0","#LTmUT<0","val_LTmUT>0","val_LTmUT<0",
+                                     "CorlmCoru_mean_CI","PlmPu_mean_CI","D2umD2l_mean_CI")
+
+for(sp in c(1:20)){
+  #cat("sp=",sp,"\n")
+  summary_aphid_count_allsp$species[sp]<-sp
+  LTmUT<-(RES_all_sp_aphid_count[[sp]]$LTdep_AICw - RES_all_sp_aphid_count[[sp]]$UTdep_AICw) # total loc pair 182
+  summary_aphid_count_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
+  summary_aphid_count_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
+  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
+  summary_aphid_count_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
+  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
+  summary_aphid_count_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])
+
+  summary_aphid_count_allsp$CorlmCoru_mean_CI[sp]<-list(
+    c(unname(round(stat_aphid_count_allsp[[sp]]$numericdf[9,c(2:4)],3))))
+  summary_aphid_count_allsp$PlmPu_mean_CI[sp]<-list(
+    c(unname(round(stat_aphid_count_allsp[[sp]]$numericdf[10,c(2:4)],3))))
+  summary_aphid_count_allsp$D2umD2l_mean_CI[sp]<-list(
+    c(unname(round(stat_aphid_count_allsp[[sp]]$numericdf[11,c(2:4)],3))))
+}
+saveRDS(summary_aphid_count_allsp,"./Results/summary_aphid_count_allsp.RDS")
 ```
@@
 \newpage
+<!--results='hide'-->
+```{r skewness_aphidcount_allsp_plot, echo=F, results="hide", warning=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+
+# For sp =1, sp=4 and sp=6 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite : you can see those msg by deleting results="hide" and warning=F option in r chunk
+
+source("good_loclist.R")
+source("skewness_multivar_dataset.R")
+set.seed(seed)
+
+numsurrog<-10000
+
+splist<-c(1:20)
+skw_aphid_count_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+colnames(skw_aphid_count_spearman)<-c("realskw","p_left","p_right")
+rownames(skw_aphid_count_spearman)<-paste("sp",splist,sep="")
+pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_spearman.pdf",height=20,width=20)
+op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+cat("Now for spearman preserving surrogs : aphid count","\n")
+for(i in splist){
+  cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="spearman") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_aphid_count_spearman$realskw[i]<-ans$realskw
+  skw_aphid_count_spearman$p_left[i]<-p_left
+  skw_aphid_count_spearman$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_aphid_count_spearman,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_spearman.RDS")
+
+skw_aphid_count_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+colnames(skw_aphid_count_kendall)<-c("realskw","p_left","p_right")
+rownames(skw_aphid_count_kendall)<-paste("sp",splist,sep="")
+pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_kendall.pdf",height=20,width=20)
+op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+cat("Now for kendall preserving surrogs : aphid count","\n")
+for(i in splist){
+  cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="kendall") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_aphid_count_kendall$realskw[i]<-ans$realskw
+  skw_aphid_count_kendall$p_left[i]<-p_left
+  skw_aphid_count_kendall$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_aphid_count_kendall,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_kendall.RDS")
+```
+
+```{r skewness_aphidff_allsp_plot, echo=F, results="hide", warning=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+
+# For sp =12 and sp=16 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite : you can see those msg by deleting results="hide" and warning=F option in r chunk
+
+source("good_loclist.R")
+source("skewness_multivar_dataset.R")
+set.seed(seed)
+
+numsurrog<-10000
+
+splist<-c(1:20)
+skw_aphid_ff_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+colnames(skw_aphid_ff_spearman)<-c("realskw","p_left","p_right")
+rownames(skw_aphid_ff_spearman)<-paste("sp",splist,sep="")
+pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_spearman.pdf",height=20,width=20)
+op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+cat("Now for spearman preserving surrogs : aphid ff","\n")
+for(i in splist){
+  cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="spearman") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_aphid_ff_spearman$realskw[i]<-ans$realskw
+  skw_aphid_ff_spearman$p_left[i]<-p_left
+  skw_aphid_ff_spearman$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_aphid_ff_spearman,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_spearman.RDS")
+
+skw_aphid_ff_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+colnames(skw_aphid_ff_kendall)<-c("realskw","p_left","p_right")
+rownames(skw_aphid_ff_kendall)<-paste("sp",splist,sep="")
+pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_kendall.pdf",height=20,width=20)
+op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+cat("Now for kendall preserving surrogs : aphid ff","\n")
+for(i in splist){
+  cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="kendall") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_aphid_ff_kendall$realskw[i]<-ans$realskw
+  skw_aphid_ff_kendall$p_left[i]<-p_left
+  skw_aphid_ff_kendall$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_aphid_ff_kendall,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_kendall.RDS")
+```
+
+```{r skewness_plankton_allsp_plot, echo=F, results="hide", cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+
+# There is no error msg or warning coming from ncsurrog : ncov semidefinite matrix for any species with spearman
+# or kendall preserving surrogates.
+
+source("good_loclist.R")
+source("skewness_multivar_dataset.R")
+set.seed(seed)
+
+splist<-c(1:22)
+skw_plankton_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+colnames(skw_plankton_spearman)<-c("realskw","p_left","p_right")
+rownames(skw_plankton_spearman)<-paste("sp",splist,sep="")
+skw_plankton_kendall<-skw_plankton_spearman
+pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_spearman.pdf",height=24,width=24)
+op<-par(mfrow=c(4,6),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+#cat("Now for spearman preserving surrogs : pns","\n")
+for(i in splist){
+  #cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="spearman") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_plankton_spearman$realskw[i]<-ans$realskw
+  skw_plankton_spearman$p_left[i]<-p_left
+  skw_plankton_spearman$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_plankton_spearman,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_spearman.RDS")
+
+pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=20,width=20)
+op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+#cat("Now for kendall preserving surrogs : pns","\n")
+for(i in splist){
+  #cat("sp=",i,"\n")
+  loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
+  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
+                        numsurrog=numsurrog,
+                        ploton=T,
+                        corpres="kendall") 
+  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+  skw_plankton_kendall$realskw[i]<-ans$realskw
+  skw_plankton_kendall$p_left[i]<-p_left
+  skw_plankton_kendall$p_right[i]<-p_right
+  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+}
+par(op)
+dev.off()
+saveRDS(skw_plankton_kendall,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_kendall.RDS")
+```
 
@@
 <!--do a TL analysis on the data just created-->
-```{r tl_analysis, echo=F, cache=T, cache.extra=list(kend.surrog,spea.surrog,mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
+```{r tl_analysis, echo=F, cache=T, cache.extra=list(kend.surrog,spea.surrog,mtime("mncurvat.R"),mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
 source("copsurrognd.R")
@@
  ss<-spea.surrog[[counter]]
- kend.stats.res[[counter]]<-matrix(NA,6,numdat) # to keep spatial TL stats
- spea.stats.res[[counter]]<-matrix(NA,6,numdat)
+ kend.stats.res[[counter]]<-matrix(NA,7,numdat) # to keep spatial TL stats
+ spea.stats.res[[counter]]<-matrix(NA,7,numdat)
  kend.points.res[[counter]]<-array(NA,c(N,2,numdat)) # to keep log(mu), log(var) of spatial TL as a dataframe
@@
 s.quadcoeff<-cbind(kend.stats.res[[1]][6,],kend.stats.res[[2]][6,],kend.stats.res[[3]][6,],kend.stats.res[[4]][6,])
-
-pdf("./Results/taylorslaw_results/SpatialTL_6stats_kendall.pdf",height=3,width=18)
-op<-par(mfrow=c(1,6),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+s.mncurv<-cbind(kend.stats.res[[1]][7,],kend.stats.res[[2]][7,],kend.stats.res[[3]][7,],kend.stats.res[[4]][7,])
+
+pdf("./Results/taylorslaw_results/SpatialTL_7stats_kendall.pdf",height=3,width=21)
+op<-par(mfrow=c(1,7),mar=c(4.2,5,1,1),mgp=c(3,1,0))
 #***DAN: x axis needs to be redone to show the copula names used in the surrogates
@@
 #plot s.quadcoeff distr for each set of surrogates
-boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL quadratic coefficient", cex.axis=1.5,cex.lab=1.5,xaxt="n")
+boxplot(s.quadcoeff , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL quadratic coefficient", cex.axis=1.5,cex.lab=1.5,xaxt="n")
+axis(1, at = c(1:4), x2, cex.axis=1.5)
+
+#plot s.mncurv distr for each set of surrogates
+boxplot(s.mncurv , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL mean curvature", cex.axis=1.5,cex.lab=1.5,xaxt="n")
 axis(1, at = c(1:4), x2, cex.axis=1.5)
@@
 s.quadcoeff<-cbind(spea.stats.res[[1]][6,],spea.stats.res[[2]][6,],spea.stats.res[[3]][6,],spea.stats.res[[4]][6,])
-
-pdf("./Results/taylorslaw_results/SpatialTL_6stats_spearman.pdf",height=3,width=18)
-op<-par(mfrow=c(1,6),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+s.mncurv<-cbind(spea.stats.res[[1]][7,],spea.stats.res[[2]][7,],spea.stats.res[[3]][7,],spea.stats.res[[4]][7,])
+
+pdf("./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf",height=3,width=21)
+op<-par(mfrow=c(1,7),mar=c(4.2,5,1,1),mgp=c(3,1,0))
 x2<-c("C", "N", "F", "SC")
@@
 \begin{center}
-\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_6stats_kendall.pdf}
+\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
 \caption[Spatial Taylor's Law results]{Spatial Taylor's Law results: log(spatial variance) vs.

### 2018-10-03 TaylorsLawTools.R

@@ -21,6 +21,7 @@
 #for spatial TL
 #4) similar for temporal
 #
+source("mncurvat.R")
 tl_stats<-function(m)
 {
   sres<-worker_tl(m)
@@ -54,12 +55,17 @@ worker_tl<-function(m)
   mod3<-lm(absresids~predicts)
   hhet<-anova(mod3)
   
+  cs<-unname(mod2$coefficients) #intercept,quadcoeff,lineartermcoef
+  coefs<-cs[c(2,3,1)]
+  mn_curv<-mncurvat(coefs = coefs,xvals=lms)
+  
   res<-c(unname(hlin[6][2,1]), #the p-value for the quadratic comparison
          unname(hhet[5][1,1]), #the p-value for a test of homoskedasticity
          sqrt(mean(resid(mod1)^2)), #the root mean squared error for the TL model
          unname(mod1$coefficients[1]), #the intercept
          unname(mod1$coefficients[2]), #the slope
-         unname(mod2$coefficients[2]))  # the coefficient of quadratic term
+         unname(mod2$coefficients[2]), #the coefficient of quadratic term
+         mn_curv)  #mean curvature of fitted quadratic polynomial
   
   return(list(stats=res,points=data.frame(log10m=lms,log10v=lvs)))
 }

### 2018-10-03 mncurvat.R

@@ -1,0 +1,16 @@
+#Gives the average curvature of a quadratic function f(x)=ax^2+bx+c
+#at a bunch of values of x
+#
+#coefs      c(a,b,c)
+#xvals      The values of x
+#
+#Returns one single number which is the average curvature
+
+mncurvat<-function(coefs,xvals){
+  a<-coefs[1]
+  b<-coefs[2]
+  c<-coefs[3]
+  kappa<-2*a/((1+(2*a*xvals+b)^2)^(3/2))
+  return(mean(kappa))
+}
+

### 2018-10-05 SupportingInformation.Rmd
adding modelsel and npa rtesults for all aphid sp : firstflight

@@
 cop_b<-getcopula(d=b,rankon=T,ploton=T) #convert to a copula and plot
+dev.off()
+```
+
+```{r read_BMR_mammals_and_copulaplot, echo=F, results='hide'}
+set.seed(seed)
+raw_m<-readxl::read_xls("./Data/BMR/MammalianBodyMassesMetabolicRates/MammalianMassesAndMetabolicRatesMcNab2008.xls")
+raw_m<-raw_m[,1:3]
+raw_m<-na.omit(raw_m)
+raw_m<-raw_m[,2:3]
+
+m<-log10(raw_m)
+colnames(m)<-c("log10M(g)","log10BMR(KJ/h)")
+m<-as.data.frame(m)
+
+pdf("./Results/BMR_results/mammals/raw_mammals_BMR.pdf")
+op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
+plot(m[,1],m[,2],type="p",col="red",xlab="log(Mass)",ylab="log(BMR)",cex.lab=1.5,cex.axis=1.5)
+par(op)
+dev.off()
+
+source("getcopula.R")
+pdf("./Results/BMR_results/mammals/copula_mammals_BMR.pdf")
+cop_m<-getcopula(d=m,rankon=T,ploton=T) #convert to a copula and plot
 dev.off()
@@
 \includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
-\caption[Testing non-parametric statistics for Ceder-creek data with $numbin=10$ for year 2000]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{birds' BMR data}.
+\caption[Testing non-parametric statistics for birds' BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{birds' BMR data}.
 As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
@@
              format="pandoc",
-             caption = "**Birds' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis on year 2000 \\label{tab_bmr_birds_asym}",
+             caption = "**Birds' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_birds_asym}",
+             booktabs=T)
+```
+
+```{r bivmodselect_mammals_BMR, echo=F, cache=T, cache.extra=list(seed,cop_m,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("OurBiCopSelect.R")
+set.seed(seed)
+BivMS_mammals<-OurBiCopSelect(u1=cop_m[,1],u2=cop_m[,2],families=c(1,3:10,13:14,16:20),
+                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
+                              gofnormal=F,status=F)
+saveRDS(BivMS_mammals,file="./Results/BMR_results/mammals/BivMS_mammals.RDS")
+```
+
+<!--Display model selection results for BMR:birds data in a table-->
+```{r tab_bmr_mammals_fit, echo=F, results='markup'}
+h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+knitr::kable(h,
+             format='pandoc',
+             caption = "Fitting info for **mammals' BMR dataset** . \\label{tab_bmr_birds_fit}",
+             booktabs=T)
+```
+
+```{r stat_bmr_mammals_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_m,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+source("bivfunctionplot.R")
+set.seed(seed)
+stat_mammals_ga<-bivfunctionplot(v=cop_m,resloc="./Results/BMR_results/mammals/",nametag="mammals_numbin_10",numbin = 10)
+```
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
+\caption[Testing non-parametric statistics for mammals' BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{mammals' BMR data}.
+As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
+When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
+\label{fig_bmr_mammals_nonparam}}
+\end{center}
+\end{figure}
+
+```{r tab_bmr_mammals_asym, echo=F, results="markup"}
+tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
+knitr::kable(tab_bmr_mammals_asym,
+             format="pandoc",
+             caption = "**Mammals' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_mammals_asym}",
              booktabs=T)
@@
 RES_all_sp_aphid_count<-mclapply(X=c(1:20),FUN=RES_single_sp,d_allsp=data_aphid_count,
-                                 families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=35,
+                                 families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
                                  good_loc_given=F,good_loc=NA,mc.cores=3)
@@
   summary_aphid_count_allsp$species[sp]<-sp
-  LTmUT<-(RES_all_sp_aphid_count[[sp]]$LTdep_AICw - RES_all_sp_aphid_count[[sp]]$UTdep_AICw) # total loc pair 182
+  LTmUT<-(RES_all_sp_aphid_count[[sp]]$LTdep_AICw - RES_all_sp_aphid_count[[sp]]$UTdep_AICw) 
   summary_aphid_count_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
@@
 \end{figure}
+
+```{r modelsel_and_npa_aphid_ff_allsp, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+
+set.seed(seed)
+library(parallel)
+source("FittingCopula_selective_loc.R")
+RES_all_sp_aphid_ff<-mclapply(X=c(1:20),FUN=RES_single_sp,d_allsp=data_aphid_ff,
+                                 families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
+                                 good_loc_given=F,good_loc=NA,mc.cores=3)
+saveRDS(RES_all_sp_aphid_ff,"./Results/fitting_results/RES_all_sp_aphid_ff.RDS")
+
+source("good_loclist.R")
+source("NonParamStat.R")
+resloc<-"./Results/stat_results/stat_aphid_ff/"
+longs<-c(-4.567,0.57,-3.069,-3.312,-2.637,-1.682,-2.763,-0.356,-3.454,0.427,0.939)
+lats<-c(55.477,52.26,56.457,55.949,52.125,55.213,53.854,51.807,50.628,51.733,51.185)
+
+stat_aphid_ff_allsp<-vector("list",20)
+
+for(sp in 1:20){
+  nm<-paste("sp_",sp,sep="")
+  resloc2<-paste(resloc,nm,sep="")
+   if (!dir.exists(resloc2)){
+    dir.create(resloc2)
+  }
+  good_loc<-good_loclist(d_allsp=data_aphid_ff,sp=sp,data_pt_thrs=30)
+
+  temp<-multcall(d_allsp=data_aphid_ff,sp=sp,lats=lats,longs=longs,
+                 pfname=paste(resloc2,"/","Sp_",sp,sep=''),good_loc=good_loc)
+  stat_aphid_ff_allsp[[sp]]<-temp
+}
+saveRDS(stat_aphid_ff_allsp,paste(resloc,file="stat_aphid_ff_all_sp.RDS",sep=''))
+
+summary_aphid_ff_allsp<-as.data.frame(matrix(NA,nrow=20,ncol=8))
+colnames(summary_aphid_ff_allsp)<-c("species","#LTmUT>0","#LTmUT<0","val_LTmUT>0","val_LTmUT<0",
+                                     "CorlmCoru_mean_CI","PlmPu_mean_CI","D2umD2l_mean_CI")
+
+for(sp in c(1:20)){
+  #cat("sp=",sp,"\n")
+  summary_aphid_ff_allsp$species[sp]<-sp
+  LTmUT<-(RES_all_sp_aphid_ff[[sp]]$LTdep_AICw - RES_all_sp_aphid_ff[[sp]]$UTdep_AICw) 
+  summary_aphid_ff_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
+  summary_aphid_ff_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
+  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
+  summary_aphid_ff_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
+  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
+  summary_aphid_ff_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])
+
+  summary_aphid_ff_allsp$CorlmCoru_mean_CI[sp]<-list(
+    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[9,c(2:4)],3))))
+  summary_aphid_ff_allsp$PlmPu_mean_CI[sp]<-list(
+    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[10,c(2:4)],3))))
+  summary_aphid_ff_allsp$D2umD2l_mean_CI[sp]<-list(
+    c(unname(round(stat_aphid_ff_allsp[[sp]]$numericdf[11,c(2:4)],3))))
+}
+saveRDS(summary_aphid_ff_allsp,"./Results/summary_aphid_ff_allsp.RDS")
+```
 
@@
 m2<-bivar_tab_summary_fn(r=BivMS_birds,rs=stat_birds_ga)
-m3<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
-tab_bivar_summary<-cbind(m1,m2,m3)
-colnames(tab_bivar_summary)<-c("SoilCN","Birds' BMR","Ceder-creek")
+m3<-bivar_tab_summary_fn(r=BivMS_mammals,rs=stat_mammals_ga)
+m4<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
+tab_bivar_summary<-cbind(m1,m2,m3,m4)
+colnames(tab_bivar_summary)<-c("SoilCN","Birds' BMR","Mammals' BMR","Ceder-creek")
 saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")

### 2018-10-05 Paper.Rmd
adding allsp_pns_rchunk and other modifications

@@
 
+<!--Moran effects-->
 \textbf{(i)} To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
@@
 Fig.
-\ref{fig_Cause4copula_C_kend} and Fig.
-\ref{fig_Cause4copula_SC_kend} represent this picture for noise generating from a lower tail dependent Clayton and upper tail dependent Survival Clayton copula whereas for Fig.
+\ref{fig_Cause4copula_C_kend} represents this picture for noise generating from a lower tail dependent Clayton copula (for upper tail dependent Survival Clayton copula see \ref{fig_Cause4copula_SC_kend}) whereas for Fig.
 \ref{SM-fig_Cause4copula_C_spear} and Fig.
@@
 
+<!--Asymmetric sensitivity and copula structure in populations-->
+\textbf{(ii)} On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold.
+
+To test the assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;
+
+\begin{equation}\label{eq.AS}
+P_{i}(t+1) = b P_{i}(t) + g(\epsilon_{i}(t)) + h(\delta_{i}(t))
+\end{equation}
+
+$b$ was the linear growth rate.
+Global environmental fluctuation function $g(\epsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq.
+\ref{eq.AS_g}) where $\epsilon=(\epsilon_{1}, \epsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\epsilon}$.
+
+\begin{align}
+g(\epsilon) = \begin{cases} 
+g_{1} = \begin{cases}
+\epsilon & \text{if $\epsilon< 0$}; \\
+0 & \text{otherwise}.
+\end{cases}\\ \\
+g_{2} = \begin{cases}
+\epsilon & \text{if $\epsilon> 0$}; \\
+0 & \text{otherwise}\label{eq.AS_g}.
+\end{cases}
+\end{cases} \text{where } \epsilon = N(\textbf{0},\Sigma_{\epsilon})\;\ \text{with} \;\ \Sigma_{\epsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
+\end{align}
+
+Local environmental fluctuation function $h(\delta)$ varied according to Eq.
+\ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$.
+$\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$.
+
+\begin{equation}\label{eq.AS_h}
+h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
+\end{equation}
+
+We simulated our model (\ref{eq.AS}) for $b=\pm0.1$ , $a=0.2$, $r=0.8$.
+$10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ time step of simulated points to plot the population-copula (Fig.
+\ref{fig_ms_asym_b_0.1}) for a specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
+We also carried out model selection and non-parametric statistical analysis with each copula of Fig.
+\ref{fig_ms_asym_b_0.1} and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_4}.
+Table \ref{SM-tab_summary_stat_asym_sens} represent the model selection results with AIC weights for best fitted copula and normal copula, model averaged lower and upper tail dependence, p-values (both Cramer-von Mises statistic and Kolmogorov-Smirnov statistic) for goodness of fit test.
+In each case, successful bootstrap was $100\%$.
+According to our model selection approach, non-normal copulae (with higher AIC weights) were always best-fit with some asymetric tail dependence.
+Our non-parametric analysis results (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_4}) are consistent with this model selection approach.
+
 \begin{figure}[!h]
 \begin{center}
-\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_kendall.pdf}
-\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}}
+\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
+\caption[asym sensitivity results : copula plot]{asym sensitivity results : copula plot \label{fig_ms_asym_b_0.1}}
 \end{center}
@@
 
-\textbf{(ii)} On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold.
-So, we considered another two-habitat patch model which was influenced by both local and global noises and found that a long time exposure to assymetric copula for global fluctuation shaped the population copula accordingly (for details, see Appendix \ref{SM-Cause_asym_sens}).
-
+<!--We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we don’t really know how a change in the environment copula structure would make what kind of change in the population copula structure.-->
+<!--
 ## Application : The Moran effect and copula structure in population
@@
 
-<!--We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we don’t really know how a change in the environment copula structure would make what kind of change in the population copula structure.-->
-<!--### The Moran effect and copula structure in populations-->
 Firstly, Our hypothesis was if a non-normal copula structure describes the dependence of an environmental
@@
 Table \ref{SM-tab_temp_aphid_LTmUT} showed `r sum(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw <0, na.rm=T)` out of `r (dim(RES_temp_aphid$gfc_p_CvM)[1]^2-dim(RES_temp_aphid$gfc_p_CvM)[1])-2*RES_temp_aphid$num_indep_loc_pair` location pair showed upper tail dependence in temperature copula, which can be considered as the Moran effect results causing dominant upper tail dependence in the copulas with aphid's firstflight date (Table \ref{SM-tab_ff_LTmUT}).
+-->
 
@@
 \begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_kendall.pdf}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_kendall.pdf}\\
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
 \end{center}
@@
 \begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_kendall.pdf}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_kendall.pdf}
-\caption{Skewness of spatially averaged $(A)$ Green-spruce aphid's abundance, $(B)$ Leaf-plum curling aphid's firstflight date, $(C)$ Ceratium furca's abundance and $(D)$ methane-flux data, were indicated by red line in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Kendall correlation).
-\label{fig_skewness_kendall}}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
+\caption{Skewness of spatially averaged $(A)$ green-spruce aphid's abundance, $(B)$ leaf-curling plum aphid's firstflight date, $(C)$ ceratium furca's abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation).
+\label{fig_skewness_spearman}}
 \end{center}
@@
 we used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data, to compute skewness of spatially averaged data.
-Figure \ref{fig_skewness_kendall} represents the results for each dataset with Kendall correlatiomn preserving surrogates.
+Figure \ref{fig_skewness_spearman} represents the results for each dataset with Kendall correlatiomn preserving surrogates.
 We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
@@
 
-Figure \ref{fig_skewness_kendall}(A) showed that averaged skewness was much lower than expected if the aphid count data came from a normal copula.
-In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_kendall} A).
+Figure \ref{fig_skewness_spearman}(A) showed that averaged skewness was much lower than expected if the aphid count data came from a normal copula.
+In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_spearman} A).
 As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results.
-On the other hand, Figure \ref{fig_skewness_kendall}(B) demonstrated that averaged skewness was much higher than expected if the aphid first flight data came from a normal copula.
-As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red line in Figure \ref{fig_skewness_kendall} B) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence.
-Figure \ref{fig_skewness_kendall}(C) showed that averaged skewness was lower than expected if the plankton data came from a normal copula.
+On the other hand, Figure \ref{fig_skewness_spearman}(B) demonstrated that averaged skewness was much higher than expected if the aphid first flight data came from a normal copula.
+As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red line in Figure \ref{fig_skewness_spearman} B) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence.
+Figure \ref{fig_skewness_spearman}(C) showed that averaged skewness was lower than expected if the plankton data came from a normal copula.
 Our previous results on non-parametric analysis and model fitting also supported lower tail dependence.
 But, as $p>0.05$, we could not reject $H_{0}$ in that case.
-Finally, for methane-flux data, the averaged skewness was lower than expected if the data came from a normal copula (Figure \ref{fig_skewness_kendall} D).
+Finally, for methane-flux data, the averaged skewness was lower than expected if the data came from a normal copula (Figure \ref{fig_skewness_spearman} D).
 As $p<0.05$, we rejected $H_{0}$ and that result was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence.
 
-We found the same result when we repeated our simulation with Spearman correlation preserving surrogates (see Fig. \ref{SM-fig_skewness_spearman}).
+We found the same result when we repeated our simulation with Spearman correlation preserving surrogates (see Fig. \ref{SM-fig_skewness_kendall}).
 

### 2018-10-05 SupportingInformation.Rmd
adding allsp_pns_rchunk and other modifications

@@
     fig_caption: yes
-    toc: True
     includes:
@@
 
-<!--\tableofcontents-->
+\pagenumbering{roman}
+\tableofcontents
 \listoftables
@@
 \pagebreak
+\pagenumbering{arabic}
 
@@
 
+```{r modelsel_and_npa_pns_allsp, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+
+set.seed(seed)
+library(parallel)
+source("FittingCopula_selective_loc.R")
+RES_all_sp_pns<-mclapply(X=c(1:22),FUN=RES_single_sp,d_allsp=data_plankton_north_sea,
+                              families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=45,
+                              good_loc_given=F,good_loc=NA,mc.cores=3)
+saveRDS(RES_all_sp_pns,"./Results/fitting_results/RES_all_sp_pns.RDS")
+
+source("good_loclist.R")
+source("NonParamStat.R")
+resloc<-"./Results/stat_results/stat_plankton_north_sea/"
+
+longs<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlongs201117.csv",header=F)
+lats<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlats201117.csv",header=F)
+longs<-longs[[1]] 
+lats<-lats[[1]]
+
+stat_plankton_north_sea_allsp<-vector("list",22)
+
+for(sp in 1:22){
+  nm<-paste("sp_",sp,sep="")
+  resloc2<-paste(resloc,nm,sep="")
+   if (!dir.exists(resloc2)){
+    dir.create(resloc2)
+  }
+  good_loc<-good_loclist(d_allsp=data_plankton_north_sea,sp=sp,data_pt_thrs=45)
+
+  temp<-multcall(d_allsp=data_plankton_north_sea,sp=sp,lats=lats,longs=longs,
+                                      pfname=paste(resloc2,"/","Sp_",sp,sep=''),good_loc=good_loc)
+  stat_plankton_north_sea_allsp[[sp]]<-temp
+}
+saveRDS(stat_plankton_north_sea_allsp,paste(resloc,file="stat_plankton_north_sea_allsp.RDS",sep=''))
+
+summary_pns_allsp<-as.data.frame(matrix(NA,nrow=22,ncol=8))
+colnames(summary_pns_allsp)<-c("species","#LTmUT>0","#LTmUT<0","val_LTmUT>0","val_LTmUT<0",
+                                     "CorlmCoru_mean_CI","PlmPu_mean_CI","D2umD2l_mean_CI")
+
+for(sp in c(1:22)){
+  #cat("sp=",sp,"\n")
+  summary_pns_allsp$species[sp]<-sp
+  LTmUT<-(RES_all_sp_pns[[sp]]$LTdep_AICw - RES_all_sp_pns[[sp]]$UTdep_AICw) 
+  summary_pns_allsp$`#LTmUT>0`[sp]<-sum(LTmUT>0,na.rm=T)
+  summary_pns_allsp$`#LTmUT<0`[sp]<-sum(LTmUT<0,na.rm=T)
+  ind_LTmUT_pos<-which(LTmUT>0,arr.ind = T)
+  summary_pns_allsp$`val_LTmUT>0`[sp]<-sum(LTmUT[ind_LTmUT_pos])
+  ind_LTmUT_neg<-which(LTmUT<0,arr.ind = T)
+  summary_pns_allsp$`val_LTmUT<0`[sp]<-sum(LTmUT[ind_LTmUT_neg])
+
+  summary_pns_allsp$CorlmCoru_mean_CI[sp]<-list(
+    c(unname(round(stat_plankton_north_sea_allsp[[sp]]$numericdf[9,c(2:4)],3))))
+  summary_pns_allsp$PlmPu_mean_CI[sp]<-list(
+    c(unname(round(stat_plankton_north_sea_allsp[[sp]]$numericdf[10,c(2:4)],3))))
+  summary_pns_allsp$D2umD2l_mean_CI[sp]<-list(
+    c(unname(round(stat_plankton_north_sea_allsp[[sp]]$numericdf[11,c(2:4)],3))))
+}
+saveRDS(summary_pns_allsp,"./Results/summary_pns_allsp.RDS")
+```
+
 <!--Do the model selection for plankton data and then showing results in Tables-->
@@
 \begin{center}
+\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_kendall.pdf}
+\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
 \includegraphics[width=16 cm]{./Results/Cause4copula_results/Clayton_spearman.pdf}
@@
 
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
-\caption[asym sensitivity results : copula plot]{asym sensitivity results : copula plot \label{fig_ms_asym_b_0.1}}
-\end{center}
-\end{figure}
-
+<!--
 \begin{figure}[!h]
@@
 \end{center}
-\end{figure}
+\end{figure}-->
 
 ```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
-knitr::kable(summary_stat, booktabs = T, 
+knitr::kable(summary_stat[,c(1:4)], booktabs = T, 
              format = "pandoc",
@@
 
-\newpage
-
+<!--
 ```{r tab_stat_npa_asym_5,echo=F,results="markup"}
@@
 
-\newpage
 ```{r tab_stat_npa_asym_6,echo=F,results="markup"}
@@
 
-\newpage
 ```{r tab_stat_npa_asym_7,echo=F,results="markup"}
@@
 ```
-
+-->
+<!--
 # Cause of non-Normal copula structure : Asymmetric sensitivity and copula structure in populations \label{Cause_asym_sens}
-
-To test the assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;
-
-\begin{equation}\label{eq.AS}
-P_{i}(t+1) = b P_{i}(t) + g(\epsilon_{i}(t)) + h(\delta_{i}(t))
-\end{equation}
-
-$b$ was the linear growth rate.
-Global environmental fluctuation function $g(\epsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq.
-\ref{eq.AS_g}) where $\epsilon=(\epsilon_{1}, \epsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\epsilon}$.
-
-\begin{align}
-g(\epsilon) = \begin{cases} 
-g_{1} = \begin{cases}
-\epsilon & \text{if $\epsilon< 0$}; \\
-0 & \text{otherwise}.
-\end{cases}\\ \\
-g_{2} = \begin{cases}
-\epsilon & \text{if $\epsilon> 0$}; \\
-0 & \text{otherwise}\label{eq.AS_g}.
-\end{cases}
-\end{cases} \text{where } \epsilon = N(\textbf{0},\Sigma_{\epsilon})\;\ \text{with} \;\ \Sigma_{\epsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
-\end{align}
-
-Local environmental fluctuation function $h(\delta)$ varied according to Eq.
-\ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$.
-$\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$.
-
-\begin{equation}\label{eq.AS_h}
-h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
-\end{equation}
-
-We simulated our model (\ref{eq.AS}) for different values of $b$ with $a=0.2$, $r=0.8$.
-$10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ time step of simulated points to plot the population-copula for a specific $b$ and a specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
-Fig.
-\ref{fig_ms_asym_b_0.1} and Fig.
-\ref{fig_ms_asym_b_0.9} show these results for $b=\pm0.1$ and $b=\pm0.9$, respectively.
-We also carried out model selection and non-parametric statistical analysis with each copula of those figures and summarized our results in Tables \ref{tab_summary_stat_asym_sens} - \ref{tab_stat_npa_asym_8}.
-Table \ref{tab_summary_stat_asym_sens} represent the model selection results with AIC weights for best fitted copula and normal copula, model averaged lower and upper tail dependence, p-values (both Cramer-von Mises statistic and Kolmogorov-Smirnov statistic) for goodness of fit test.
-In each case, successful bootstrap was $100\%$.
-According to our model selection approach, for $b=\pm0.1$, non-normal copula (with higher AIC weights) were always best-fit with some asymetric tail dependence whereas for $b\pm0.9$, most of the time normal was best.
-It is probably because the populations get long-time exposure to environmental changes for slower growth(or decay) rate and so global fluctuation can affect them more efficiently.
-Our non-parametric analysis results (Tables \ref{tab_stat_npa_asym_1} -\ref{tab_stat_npa_asym_8}) are consistent with this model selection approach.
+-->
 
@@
 \begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_kendall.pdf}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_kendall.pdf}\\
 \end{center}
@@
 \begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
-\caption{Skewness of spatially averaged $(A)$ green-spruce aphid's abundance, $(B)$ leaf-curling plum aphid's firstflight date, $(C)$ ceratium furca's abundance and $(D)$ methane-flux data, were indicated by red line in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation).
-\label{fig_skewness_spearman}}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_kendall.pdf}
+\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_kendall.pdf}
+\caption{Skewness of spatially averaged $(A)$ Green-spruce aphid's abundance, $(B)$ Leaf-plum curling aphid's firstflight date, $(C)$ Ceratium furca's abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Kendall correlation).
+\label{fig_skewness_kendall}}
 \end{center}
@@
 \begin{figure}[!h]
-\textbf{ \hspace{3.4 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.1 cm} (D)} \\
+\textbf{ \hspace{2.8 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.6 cm} (D)} \\
 \vspace{-0.5 cm} 
 \begin{center}
-\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_kendall.pdf}\\
+\includegraphics[width=18 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.pdf}\\
 \end{center}
-\textbf{ \hspace{2.8 cm} (E) \hspace{1.8 cm} (F) \hspace{1.8 cm} (G) \hspace{1.8 cm} (H) \hspace{1.6 cm} (I) \hspace{2 cm} (J)}\\
+\textbf{ \hspace{1.8 cm} (E) \hspace{1.6 cm} (F) \hspace{1.8 cm} (G) \hspace{1.8 cm} (H) \hspace{1.8 cm} (I) \hspace{1.8 cm} (J) \hspace{1.8 cm} (K)}\\
 \vspace{-0.5 cm}
 \begin{center}
-\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
+\includegraphics[width=18 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
 \caption[Spatial Taylor's Law results]{Spatial Taylor's Law results: log(spatial variance) vs.
-log(spatial mean) for $ns=10$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of homoskedasticity, (G) root mean square error, (H) intercept, (I) slope, (J) quadratic coefficient of regression line for first panel plot.
+log(spatial mean) for $ns=10$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of homoskedasticity, (G) root mean square error, (H) intercept, (I) slope, (J) quadratic coefficient, (K) mean curvature of regression line from first panel-plots.
 \label{fig_spTL}}
@@
 \end{figure}
+
+```{r spatialTL_on_realdata_aphid_count, results="hide", echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("TaylorsLawTools.R"),mtime("mncurvat.R"),mtime("ncsurrog.R"))}
+
+set.seed(seed)
+source("ncsurrog.R")
+source("TaylorsLawTools.R")
+source("good_loclist.R")
+
+TL_on_realdata<-function(data,numsurrog=1000,sp,data_pt_thrs,ploton,tag,labelon=T){
+
+  loclist<-good_loclist(d_allsp=data,sp=sp,data_pt_thrs=data_pt_thrs)
+  data<-data[[sp]]
+  df<-c()
+  for(i in loclist){
+    tem<-data[[i]]$Dat
+    df<-cbind(df,tem)
+  }
+  colnames(df)<-paste("loc",loclist,sep="")
+  df<-na.omit(df)
+
+  mat_real<-as.matrix(df)
+  check_rep<-function(y){all( abs(y - mean(y)) < 1e-8 )} #This fn checks for repeatation of same values in a vector
+  call_check_rep<-apply(mat_real,1,check_rep) # checks the same for mat_real matrix along all row
+
+  if(all(call_check_rep==F)==T){
+
+    realTL_stats<-tl_stats(m = mat_real)
+    real_spatialTLslope<-unname(realTL_stats$s.stats[5])
+    real_sTL_quadcoeff<-unname(realTL_stats$s.stats[6])
+    real_sTL_mncurv<-unname(realTL_stats$s.stats[7])
+
+    #generate the normal surrogates with spearman correlation preserving
+    array_surrog<-ncsurrog(m=mat_real,corpres="spearman",numsurrog=numsurrog)
+
+    bad_id<-c()
+    for(i in 1:numsurrog){
+      m<-array_surrog[,,i]
+      mean_m<-apply(X=m, MARGIN = 1, FUN = mean)
+      var_m<-apply(X=m, MARGIN = 1, FUN = var)
+      if(any(mean_m==0) | any(var_m==0)){
+        bad_id<-c(bad_id,i)
+      }
+    }
+
+    m<-array_surrog
+    if(length(bad_id)!=0){
+      m<-array_surrog[,,-bad_id]
+    }
+    ts<-apply(m,3,tl_stats)
+
+    good_surrog_slope<-c()
+    good_surrog_quadcoeff<-c()
+    good_surrog_mncurv<-c()
+    for(i in c(1:length(ts))){
+      good_surrog_slope<-c(good_surrog_slope,unname(ts[[i]]$s.stats[5]))
+      good_surrog_quadcoeff<-c(good_surrog_quadcoeff,unname(ts[[i]]$s.stats[6]))
+      good_surrog_mncurv<-c(good_surrog_mncurv,unname(ts[[i]]$s.stats[7]))
+    }
+
+    numsurrog_good<- numsurrog - length(bad_id)
+
+    p_left<-sum(good_surrog_slope < real_spatialTLslope)/numsurrog_good
+    p_right<-sum(good_surrog_slope > real_spatialTLslope)/numsurrog_good
+
+    p_left_quadcoeff<-sum(good_surrog_quadcoeff < real_sTL_quadcoeff)/numsurrog_good
+    p_right_quadcoeff<-sum(good_surrog_quadcoeff > real_sTL_quadcoeff)/numsurrog_good
+
+    p_left_mncurv<-sum(good_surrog_mncurv < real_sTL_mncurv)/numsurrog_good
+    p_right_mncurv<-sum(good_surrog_mncurv > real_sTL_mncurv)/numsurrog_good
+
+    if(ploton==T & tag=="slope"){
+      hist(good_surrog_slope,breaks=100,main="",xlab="surrogate_TL_slope",cex.axis=1.5,cex.lab=1.5)
+      points(x=real_spatialTLslope,y=0,col="red",pch=15)
+      if(labelon==T){
+        mtext(paste0("sp= ",sp,", p_left= ",round(p_left,3),", p_right=",round(p_right,3),sep=""),
+              side = 3, line=-2, adj=0, col="blue")
+      }
+    }
+
+    if(ploton==T & tag=="quadcoeff"){
+      hist(good_surrog_quadcoeff,breaks=100,main="",xlab="surrogate_TL_quadcoeff",cex.axis=1.5,cex.lab=1.5)
+      points(x=real_sTL_quadcoeff,y=0,col="red",pch=15)
+      if(labelon==T){
+        mtext(paste0("sp= ",sp,", p_left= ",round(p_left_quadcoeff,3),", p_right=",round(p_right_quadcoeff,3),sep=""),
+              side = 3, line=-2, adj=0, col="blue")
+      }
+    }
+
+    if(ploton==T & tag=="mncurv"){
+      hist(good_surrog_mncurv,breaks=100,main="",xlab="surrogate_TL_meanCurvature",cex.axis=1.5,cex.lab=1.5)
+      points(x=real_sTL_mncurv,y=0,col="red",pch=15)
+      if(labelon==T){
+        mtext(paste0("sp= ",sp,", p_left= ",round(p_left_mncurv,3),", p_right=",round(p_right_mncurv,3),sep=""),
+              side = 3, line=-2, adj=0, col="blue")
+      }
+    }
+
+  }else{
+    real_spatialTLslope<-NA
+    p_left<-NA
+    p_right<-NA
+    numsurrog_good<-NA
+    real_sTL_quadcoeff<-NA
+    p_left_quadcoeff<-NA
+    p_right_quadcoeff<-NA
+    real_sTL_mncurv<-NA
+    p_left_mncurv<-NA
+    p_right_mncurv<-NA
+  }
+  return(list(realTL_stats=realTL_stats,
+              real_spatialTLslope=real_spatialTLslope,
+              p_left=p_left,
+              p_right=p_right,
+              numsurrog_good=numsurrog_good,
+              real_sTL_quadcoeff=real_sTL_quadcoeff,
+              p_left_quadcoeff=p_left_quadcoeff,
+              p_right_quadcoeff=p_right_quadcoeff,
+              real_sTL_mncurv=real_sTL_mncurv,
+              p_left_mncurv=p_left_mncurv,
+              p_right_mncurv=p_right_mncurv,
+              surrogTL_stats=ts,
+              good_surrog_mncurv=good_surrog_mncurv))
+}
+
+#call on real data
+# For all sp
+splist<-c(1:20)
+tl_aphid_count<-as.data.frame(matrix(NA,nrow=length(splist),ncol=10))
+colnames(tl_aphid_count)<-c("real_sTL_slope","p_left","p_right","numsurrog_good",
+                            "real_sTL_quadcoeff","p_left_quadcoeff","p_right_quadcoeff",
+                            "real_sTL_mncurv","p_left_mncurv","p_right_mncurv")
+rownames(tl_aphid_count)<-paste("sp",splist,sep="")
+#pdf("aphid_count_allsp_spTLslope.pdf",height=20,width=20)
+#pdf("./Results/taylorslaw_results/aphid_count_allsp_sTL_quadcoeff.pdf",height=20,width=20)
+pdf("./Results/taylorslaw_results/aphid_count_allsp_sTL_mncurv.pdf",height=20,width=20)
+op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+for(i in splist){
+  #cat("sp=",i,"\n")
+  #ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="slope",labelon=T)
+  #ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="quadcoeff",labelon=T)
+  ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon=T)
+  tl_aphid_count$real_sTL_slope[i]<-ans$real_spatialTLslope
+  tl_aphid_count$p_left[i]<-ans$p_left
+  tl_aphid_count$p_right[i]<-ans$p_right
+  tl_aphid_count$numsurrog_good[i]<-ans$numsurrog_good
+  tl_aphid_count$real_sTL_quadcoeff[i]<-ans$real_sTL_quadcoeff
+  tl_aphid_count$p_left_quadcoeff[i]<-ans$p_left_quadcoeff
+  tl_aphid_count$p_right_quadcoeff[i]<-ans$p_right_quadcoeff
+  tl_aphid_count$real_sTL_mncurv[i]<-ans$real_sTL_mncurv
+  tl_aphid_count$p_left_mncurv[i]<-ans$p_left_mncurv
+  tl_aphid_count$p_right_mncurv[i]<-ans$p_right_mncurv
+}
+par(op)
+dev.off()
+saveRDS(tl_aphid_count,"./Results/taylorslaw_results/tl_aphid_count.RDS")
+```
+
+```{r spatialTL_on_realdata_aphid_count_sp6, results="hide", echo=F, cache=T, cache.extra=list(seed,data_aphid_count,TL_on_realdata,mtime("good_loclist.R"),mtime("TaylorsLawTools.R"),mtime("mncurvat.R"),mtime("ncsurrog.R"))}
+
+set.seed(seed)
+
+#Now for sp=6
+sp<-6
+pdf("./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf",height=5,width=5)
+ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=sp,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon = F)
+mtext(paste0("p= ",round(ans$p_left_mncurv,3)),side = 3, line=-2, adj=0.5, col="blue")
+dev.off()
+
+pdf("./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf",height=5,width=10)
+op<-op<-par(mfrow=c(1,2),mar=c(4,5,1,1),mgp=c(3,1,0))
+plot(ans$realTL_stats$s.points,xlab=expression(log[10]("spatial mean")),
+     ylab=expression(log[10]("spatial variance")),col="red",
+     cex.axis=1.5,cex.lab=1.5)
+mtext(paste0("mean curvature= ",round(unname(ans$realTL_stats$s.stats[7]),3)),side = 3, line=-2, adj=0.5, col="blue")
+
+target_mncurv<-0.1
+id<-which(abs(ans$good_surrog_mncurv-target_mncurv)<1e-4)[1]
+ans_s<-ans$surrogTL_stats
+plot(ans_s[[id]]$s.points,xlab=expression(log[10]("spatial mean")),
+     ylab=expression(log[10]("spatial variance")),
+     cex.axis=1.5,cex.lab=1.5)
+mtext(paste0("mean curvature= ",round(ans$good_surrog_mncurv[id],3)),side = 3, line=-2, adj=0.5, col="blue")
+
+par(op)
+dev.off()
+```
 

### 2018-10-05 TaylorsLawTools.R

@@ -26,12 +26,12 @@ tl_stats<-function(m)
 {
   sres<-worker_tl(m)
   s.stats<-sres$stats
-  names(s.stats)<-c("s.lin","s.hom","s.rmse","s.int","s.slope","s.quad.coeff")
+  names(s.stats)<-c("s.lin","s.hom","s.rmse","s.int","s.slope","s.quad.coeff","s.mncurv")
   s.points<-sres$points
   
   tres<-worker_tl(t(m))
   t.stats<-tres$stats
-  names(t.stats)<-c("t.lin","t.hom","t.rmse","t.int","t.slope","t.quad.coeff")
+  names(t.stats)<-c("t.lin","t.hom","t.rmse","t.int","t.slope","t.quad.coeff","t.mncurv")
   t.points<-tres$points
   
   return(list(s.stats=s.stats,t.stats=t.stats,

### 2018-10-05 skewness_multivar_dataset.R

@@ -57,7 +57,8 @@ skewness_testing<-function(ts_matrix,loclist,numsurrog,ploton,corpres){
   
   if(ploton==T){
     hist(surrogskw,breaks=100,main="",xlab="surrogate_skewness",ylab="frequency",cex.lab=2)
-    abline(v=realskw,col="red")
+    #abline(v=realskw,col="red")
+    points(x=realskw,y=0,col="red",pch=15)
   }
   
   return(list(surrogskw=surrogskw,realskw=unname(realskw)))

### 2018-10-05 SupportingInformation.Rmd
small change

@@
 lats<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlats201117.csv",header=F)
-longs<-longs[[1]]
-lats<-lats[[1]]
-
-df_loc_info_org<-data.frame(loc_no=c(1:length(lats)),longs=longs,lats=lats)
+longs_pns<-longs[[1]]
+lats_pns<-lats[[1]]
+
+df_loc_info_org<-data.frame(loc_no=c(1:length(lats_pns)),longs=longs_pns,lats=lats_pns)
 selective_loc<-good_loclist(d_allsp=data_plankton_north_sea,sp=16,data_pt_thrs=45)
@@
 
-```{r modelsel_and_npa_pns_allsp, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+```{r modelsel_and_npa_pns_allsp, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,longs_pns,lats_pns,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
 
@@
 
-longs<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlongs201117.csv",header=F)
-lats<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlats201117.csv",header=F)
-longs<-longs[[1]] 
-lats<-lats[[1]]
-
 stat_plankton_north_sea_allsp<-vector("list",22)
@@
 
-  temp<-multcall(d_allsp=data_plankton_north_sea,sp=sp,lats=lats,longs=longs,
+  temp<-multcall(d_allsp=data_plankton_north_sea,sp=sp,lats=lats_pns,longs=longs_pns,
                                       pfname=paste(resloc2,"/","Sp_",sp,sep=''),good_loc=good_loc)
@@
 <!--Do the main computations for non-parametric analysis of plankton data-->
-```{r stat_plankton_north_sea, echo=F, cache=T, cache=T,  cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+```{r stat_plankton_north_sea, echo=F, cache=T, cache=T,  cache.extra=list(seed,data_plankton_north_sea,longs_pns,lats_pns,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
 
@@
 good_loc<-good_loclist(d_allsp=data_plankton_north_sea,sp=sp,data_pt_thrs=45)
-longs<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlongs201117.csv",header=F)
-lats<-read.csv("./Data/Plankton_North_Sea_data/boxcornerlats201117.csv",header=F)
-longs<-longs[[1]] 
-lats<-lats[[1]]
-
-stat_plankton_north_sea<-multcall(d_allsp=data_plankton_north_sea,sp=sp,lats=lats,longs=longs,pfname=paste(resloc,"Sp_",sp,sep=''),good_loc=good_loc) 
+
+stat_plankton_north_sea<-multcall(d_allsp=data_plankton_north_sea,sp=sp,lats=lats_pns,longs=longs_pns,
+                                  pfname=paste(resloc,"Sp_",sp,sep=''),good_loc=good_loc) 
 saveRDS(stat_plankton_north_sea,paste(resloc,file="stat_plankton_north_sea_sp_16.RDS",sep=''))

### 2018-10-08 Paper.Rmd
modified

@@
 \begin{figure}[!h]
-\textbf{ \hspace{3.2 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
+\textbf{ \hspace{2.8 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
 \vspace{-0.35 cm}
@@
 \includegraphics[width=4 cm]{./Results/BMR_results/birds/raw_birds_BMR.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/mammals/raw_mammals_BMR.pdf}
 \includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
 \end{center}
-\textbf{ \hspace{3.2 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
+\textbf{ \hspace{2.8 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
 \vspace{-0.35 cm}
@@
 \includegraphics[width=4 cm]{./Results/BMR_results/birds/copula_birds_BMR.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/mammals/copula_mammals_BMR.pdf}
 \includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
-\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} ln(BMR) of contrasts vs.
-ln(mass) of contrasts, \textbf{(C)} contrasts for AVG.ltoe.IV vs.
-contrasts for AVG.humerus phylogenetic traits.
-\textbf{(D)} above-ground plant biomass data vs.
+\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs.
+log(mass) for birds, \textbf{(C)} log(BMR) vs.
+log(mass) for mammals, \textbf{(D)} above-ground plant biomass data vs.
 Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
 Lower panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel.
-These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of contrasts for next two datasets and number of plots for the last one (see \nameref{Model_selection} of manuscript).
+These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of species for next two datasets and number of plots for the last one (see \nameref{Model_selection} of manuscript).
 \label{fig_biv_raw_cop}}
@@
 \includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf}
-\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{500} points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF_500}}
+\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{500} points for both noise and population data for comparison).
+p-values $> 0.05$ (above horizontal dashed line) indicates good fit between noise and population copulae.
+\label{fig_Cause4copula_params_GOF_500}}
 \end{center}
@@
 
-Secondly, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
+Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
 Then we plotted mean of each measured quantity with standard error against Kendall's Tau of the noise copula family.
 Fig.
-\ref{fig_Cause4copula_C_kend} represents this picture for noise generating from a lower tail dependent Clayton copula (for upper tail dependent Survival Clayton copula see \ref{fig_Cause4copula_SC_kend}) whereas for Fig.
+\ref{fig_Cause4copula_C_kend} represents this picture for noise generating from a lower tail dependent Clayton copula (for upper tail dependent Survival Clayton copula see Fig.
+\ref{SM-fig_Cause4copula_SC_kend}) whereas for Fig.
 \ref{SM-fig_Cause4copula_C_spear} and Fig.
@@
 \caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_C_kend}}
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae.
+\label{fig_Cause4copula_C_kend}}
 \end{center}
@@
 \includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
-\caption[asym sensitivity results : copula plot]{asym sensitivity results : copula plot \label{fig_ms_asym_b_0.1}}
+\caption{Effects of asymmetric global noise copula on two habitat patch model (Eq.
+\ref{eq.AS}) with growth rate $b=\pm0.1$: copula of two populations ($P_{1}, P_{2}$) shaped according to the asymmetry of noise copula\label{fig_ms_asym_b_0.1}}
 \end{center}
@@
 To test null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure, we generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
-Then we calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R).
-We made a histogram of surragate skewness and found if actual skewness were within $95\%$ quantile or not.
-If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).
+Then we calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R)).
+We made a histogram of surragate skewness and found if actual skewness were within $95\%$ confidence interval or not.
+If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness than Normal copula).
 
 We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}).
-For example, our model simulation showed lower tail dependent Clayton copula had much higher slope (spatial-Taylor's law) than symmetric Normal copula, which inturn had higher slope than upper tail dependent Survival Clayton family (Fig.
-\ref{SM-fig_spTL}).
-Finally, we studied **(iii)** the effects of non-Normal noise copula on extinction risk (calculated after 25 years) with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}).
+For example, our model simulation showed lower tail dependent Clayton copula had much higher slope (spatial-Taylor's law) than symmetric Normal copula, which in turn had higher slope than upper tail dependent Survival Clayton family (Fig.
+\ref{SM-fig_spTL}(I)).
+Finally, we studied **(iii)** the effects of environmental noise, having a non-Normal copula, on extinction risk (calculated after 25 years) with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}).
 We considered patches were arranged in a linear chain fashion and the population may disperse to its nearerst two neighbors (local dispersion) or to any patches (global dispersion).
@@
 For Lewton-Cohen model, we found greater extinction risk if populations were affected by left tail dependence environmental noise copula than right tail dependent one.
-For Ricker model, we got the similar observation for smaller values of $r$, the intrinsic growth rate, but after $r=0.7$, there was a turn-over in extinction risks, and right tail dependent noise copula caused population go extinct faster.
+For Ricker model, we got the similar observation for smaller values of $r$, the intrinsic growth rate, but for higher values of $r$, there was a turn-over in extinction risks, and right tail dependent noise copula caused population go extinct faster.
 
@@
 we used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data, to compute skewness of spatially averaged data.
-Figure \ref{fig_skewness_spearman} represents the results for each dataset with Kendall correlatiomn preserving surrogates.
-We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
-If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness).
+Figure \ref{fig_skewness_spearman} represents the results for each dataset with Spearman correlation preserving surrogates.
+We tested null hypothesis $(H_{0})$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
+If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness than Normal copula).
 
-Figure \ref{fig_skewness_spearman}(A) showed that averaged skewness was much lower than expected if the aphid count data came from a normal copula.
-In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data (red line in Figure \ref{fig_skewness_spearman} A).
-As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results.
+Figure \ref{fig_skewness_spearman}(A) showed that averaged skewness (indicated by red dot) was much lower than expected if the aphid count data came from a normal copula.
+In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data.
+As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results (see first column of Table \ref{tab_multivar_summary}).
 On the other hand, Figure \ref{fig_skewness_spearman}(B) demonstrated that averaged skewness was much higher than expected if the aphid first flight data came from a normal copula.
-As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red line in Figure \ref{fig_skewness_spearman} B) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence.
+As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red dot in Figure \ref{fig_skewness_spearman} B) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence, which was also confirmed by model selection and non-parametric results(see second column of Table \ref{tab_multivar_summary}).
 Figure \ref{fig_skewness_spearman}(C) showed that averaged skewness was lower than expected if the plankton data came from a normal copula.
-Our previous results on non-parametric analysis and model fitting also supported lower tail dependence.
+Our previous results on non-parametric analysis and model fitting also supported lower tail dependence (see third column of Table \ref{tab_multivar_summary}).
 But, as $p>0.05$, we could not reject $H_{0}$ in that case.
 Finally, for methane-flux data, the averaged skewness was lower than expected if the data came from a normal copula (Figure \ref{fig_skewness_spearman} D).
-As $p<0.05$, we rejected $H_{0}$ and that result was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence.
+As $p<0.05$, we rejected $H_{0}$ and that result too was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence (see last column of Table \ref{tab_multivar_summary}).
 
-We found the same result when we repeated our simulation with Spearman correlation preserving surrogates (see Fig. \ref{SM-fig_skewness_kendall}).
+We carried out the same analysis with aphid abundance and aphid firstflight date data for all 20 aphid species and plankton abundance data for all 22 plankton species.
+In each study, we got $>5\%$ positive results which gave us confidence in validating significantly different skewness for asymmetric copula than Normal.
+
+We found the same result when we repeated our simulation with Kendall correlation preserving surrogates (see Fig. \ref{SM-fig_skewness_kendall}).
+
+## Application : Change of curvature of spatial Taylor's law regression line
+
+Spatial Taylor's law, which shows linear relationship between `log`(variance) and `log`(mean) of population densities of a species from different locations, is a well-known pattern found in ecology and related fields [@Taylor1961, @Taylor1988, @Cohen2013, @Doring2015, @Kuo2016, @Xu2016, @Reuman2017].
+We already showed through simulation (Appendix \ref{SM-Consequence_TL_simulation}) that for asymmetric copula this linear relationship somehow deviates compared to Normal copula which was presumed to have gamma marginals.
+We tested spatial Taylor's law with aphid's count data (for all 20 species) and found that the mean curvature of the regression line from `log`(variance) vs.
+`log`(mean) of population abundance data was significantly different from $10000$ Normal surrogates (Spearman-correlation and marginal distribution preserving) for $25\%$ of total species considered.
+For example, species "Corn leaf aphid" (having significantly upper tail dependence from model selection and non-parametric analysis) showed mean curvature of regression line close to zero, whereas it's one of the $10000$ Normal surrogate showed a higher curvature, indicating here Normal copula didn't hold linear relationship between `log`(spatial variance) and `log`(spatial mean) (see Fig \ref{fig_sTL_mncurv_sp6}).
+
+\begin{figure}[!h]
+\textbf{ \hspace{4 cm} (A) \hspace{4 cm} (B) \hspace{4 cm} (C)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=10 cm]{./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf}
+\includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
+\caption{log(spatial variance) and log(spatial mean) for $(A)$ corn-leaf aphid's abundance and for $(B)$ one of the $10000$ Normal surrogates (having same marginal distribution and Spearman correlation of actual data).
+$(C)$ Histogram of mean curvature of $10000$ normal surrogates showed that for leaf plum corn aphid's abundance data (with asymmetric tail dependence), Normal surrogates didn't follow linear relationship between log(spatial variance) and log(spatial mean).
+\label{fig_sTL_mncurv_sp6}}
+\end{center}
+\end{figure}
 

### 2018-10-08 SupportingInformation.Rmd
modified

@@
              format='pandoc',
-             caption = "Fitting info for **mammals' BMR dataset** . \\label{tab_bmr_birds_fit}",
+             caption = "Fitting info for **mammals' BMR dataset** . \\label{tab_bmr_mammals_fit}",
              booktabs=T)
@@
 \includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf}
-\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison) \label{fig_Cause4copula_params_GOF_2500}}
+\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison).
+p-values $> 0.05$ (above horizontal dashed line) indicates good fit between noise and population copulae.
+\label{fig_Cause4copula_params_GOF_2500}}
 \end{center}
@@
 \caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_SC_kend}}
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae.
+\label{fig_Cause4copula_SC_kend}}
 \end{center}
@@
 \caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_C_spear}}
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae.
+\label{fig_Cause4copula_C_spear}}
 \end{center}
@@
 \caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison) \label{fig_Cause4copula_SC_spear}}
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae.
+\label{fig_Cause4copula_SC_spear}}
 \end{center}
@@
 
+<!--
 ```{r tab_temp_aphid_bestcop, echo=F, results="markup"}
@@
 ```
-
+-->
 <!--# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}-->
@@
 
+\newpage
 ```{r tab_stat_npa_asym_4,echo=F,results="markup"}
@@
 -->
-
-#  Need to think about mechanisms for biogeochemical, evolutionary, and community data
 
@@
 
-pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=20,width=20)
-op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=24,width=24)
+op<-par(mfrow=c(4,6),mar=c(4.2,5,1,1),mgp=c(3,1,0))
 #cat("Now for kendall preserving surrogs : pns","\n")
@@
 pdf("./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf",height=5,width=5)
+op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
 ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=sp,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon = F)
 mtext(paste0("p= ",round(ans$p_left_mncurv,3)),side = 3, line=-2, adj=0.5, col="blue")
+par(op)
 dev.off()
@@
 
-# Consequence of non-Normal copula structure : Effect on Taylor's law \label{Consequence_TL}
-
-Spatial Taylor's law, which shows linear relationship between `log`(variance) and `log`(mean) of population densities of a species from different locations, is a well-known pattern found in ecology and related fields [@Taylor1961, @Taylor1988, @Cohen2013, @Doring2015, @Kuo2016, @Xu2016, @Reuman2017].
+# Consequence of non-Normal copula structure : Effect on Taylor's law \label{Consequence_TL_simulation}
+
 Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula.
@@
 With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D).
-We also tested statistically that whether linearity, homoskedasticity is maintained in each copula dataset or not and plotted our results alongwith root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).
+We also tested statistically that whether linearity, homoskedasticity is maintained in each copula dataset or not and plotted our results alongwith root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I).
+We did a quadratic regression with each simulated copula for all types and measured quadratic coefficient and mean curvature for the regression line which is plotted in Fig.\ref{fig_spTL}(J-K) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).
 

### 2018-10-08 Paper.Rmd
modify intro : still continue

@@
 
-Identifying and quantifying dependencies is the core of modelling in many applications, frequently found in the
-finance and risk management. For many years, measuring and modeling dependencies has focused on the
+Identifying and quantifying dependencies is the core of modelling in many applications, frequently found in finance and risk management.
+For many years, measuring and modeling dependencies has focused on the
 Pearson correlation because of its ease in use.
-But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence because it is a measure of linear association between two random variables.
-Therefore, it is better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation.
-However, using copula-based approach as a dependence measure reflects more knowledge about the dependence structure instead of summarizing in just a single number.
-The flexibility of copulas allows us to model more complex dependence structures by isolating its marginal from its joint distribution structures, with the capability in modeling a rich variety of tail behavior, such as tail dependence versus tail independence, and reflection symmetry versus reflection asymmetry.
+But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence because it is a measure of linear association between two variables.
+It may be better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation, when the marginal distribution of the variables are highly non-Gaussian.
+Copula, a ranked based dependence measurement between two variables with initially some non-uniformly distributed marginals, transforms the marginals into uniform distribution when joint distribution of ranked data is plotted.
+For example, in the first panel of Fig.
+\ref{fig_cop_pedag}, we showed the joint distribution of raw data (left) and copula data (right), where the marginals of raw data variables (x and y) were gamma distributed but in copula plot those marginals had uniform distribution.
+Copula plot was obtained by scaling individual ranked variables by a factor $(n+1)$ where $n$ was the number of sampling points, so that the copula was confined within a unit box.
+However, using a "copula-based approach" as a dependence measure reflects more knowledge about the dependence structure instead of summarizing in just a single number, the linear correlation coefficient.
+It tells us about the tail-dependence of the variables, i.e., the association between extreme values of two variables.
+If smaller values of each variable are more correlated to each other, then the corresponding copula structure is said to have "lower" tail-dependence as shown in the first panel of Fig.
+\ref{fig_cop_pedag} copula plot (right).
+Similarly, an "upper" tail-dependence copula indicates stronger correlation between higher extreme variables in the raw data.
+The flexibility of copulas allows us to model more complex dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence, and reflection symmetry versus reflection asymmetry.
 Tail dependence occurs when the correlation between two variables increases as we get "further" in the tail (either or both) of the distribution.
 For example, in the upper panel of Fig.
-\ref{fig_cop_pedag}, different dependence structures, with same Kendall correlation $\tau = 0.5$, were shown which became dominantly lower tail dependent (lower tail - upper tail dependence being positive) from dominantly upper tail dependent (lower tail - upper tail dependence being negative) as we moved along the row.
+\ref{fig_cop_pedag}, different dependence structures, with same Kendall correlation $\tau = 0.5$, are shown which vary between dominantly lower tail dependent (lower tail - upper tail dependence being positive) from dominantly upper tail dependent (lower tail - upper tail dependence being negative) as we move along the row.
 Lower panel represents the similar scenario but with stronger correlation (Kendall's $\tau = 0.8$).
 To implement this, we need to select candidate bivariate copulas from many existing parametric copula families.
-A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulas to be selected from.
-Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), mentioned in the \nameref{Methods}.
+<!--A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulas to be selected from.--> Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), mentioned in the \nameref{Methods}.
 
 Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data.
-So, in ecological perspective, when we find a dominantly lower (upper) tail dependence structure, it means both incidence are rare (abundant) synchronously.
-If any one of them be environmental variable and the other is an observable (like abundance of a species) then it can also explain the causal relationship between them.
+In ecological perspective, when we find a dominantly lower (upper) tail dependence structure between two species abundance timeseries, it means both incidence are rare (abundant) synchronously.
+If one variable is an environmental variable and another is an observable (like abundance of a species) then it can also explain the causal relationship between them.
 Here, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets.
@@
 \begin{center}
+\includegraphics[width=4 cm,height=4 cm]{./Results/rawdata.pdf}
+\includegraphics[width=4 cm,height=4 cm]{./Results/copdata.pdf}\\
 \includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}

### 2018-10-08 SupportingInformation.Rmd
modify intro : still continue

@@
 source("mtime.R") #A function needed for caching
+```
+
+```{r cop_fig_intro,echo=F,results="hide",cache=T, cache.extra=list(seed,mtime("copsurrognd.R"))}
+library(copula)
+library(VineCopula)
+library(mvtnorm)
+set.seed(seed)
+
+source("copsurrognd.R")
+
+#x <- read.table("./Data/dataset/x.txt")
+#y <- read.table("./Data/dataset/y.txt")
+#mat<-cbind(x,y)
+#colnames(mat)<-c("x","y")
+
+#***constants
+N<-200 #length of time series
+n<-2 #number of sampling locations
+numdat<-1 #the number of datasets to get from the normal copula
+rho<-0.8 #normal-copula parameter for dependence between locations
+#this is a covariance, see below call to rmvnorm
+
+#***generate the starting data, which has normal copula structure
+#because it comes originally from a multivariate normal, before 
+#transformation
+sig<-matrix(rho,n,n)
+diag(sig)<-1
+m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
+m<-array(m,c(N,numdat,n))
+m<-aperm(m,c(1,3,2)) # 2nd dim is normally distributed e.g. hist(m[,1,])
+gshape<-7.5
+gscale<-1 #gamma parameters
+
+cop_to_marg_trans<-function(x){
+  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
+  #should be something skewed
+}
+# pnorm(m) makes normal to uniform
+# then qgamma(...) makes uniform to gamma distributed
+m<-cop_to_marg_trans(pnorm(m))
+
+ccop<-claytonCopula(2,2)
+csurrog<-array(NA,dim(m))
+csurrog[,,1]<-copsurrognd(m[,,1],targetcop=ccop,numsurrog=1)
+mat<-csurrog[,,1]
+
+pdf("./Results/rawdata.pdf",height = 4, width = 4)
+scale<-2
+xhist <- hist(mat[,1], breaks=30, plot=FALSE)
+yhist <- hist(mat[,2], breaks=30, plot=FALSE) 
+top <- max(c(xhist$counts, yhist$counts)) 
+xrange <- range(mat[,1])
+yrange <- range(mat[,2])
+nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
+op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0))  
+plot(mat[,1], mat[,2], xlim=xrange, ylim=yrange, xlab="x", ylab="y", cex.lab=1.5) 
+op<-par(mar=c(0,3,1,1)) 
+barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
+op<-par(mar=c(3,0,1,1)) 
+barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
+par(op)
+dev.off()
+
+pdf("./Results/copdata.pdf",height = 4, width = 4)
+copmat<-pobs(mat)
+xhist <- hist(copmat[,1], breaks=30, plot=FALSE)
+yhist <- hist(copmat[,2], breaks=30, plot=FALSE) 
+top <- max(c(xhist$counts, yhist$counts)) 
+xrange <- c(0,1)
+yrange <- c(0,1)
+nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
+op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0)) 
+plot(copmat[,1], copmat[,2], xlim=xrange, ylim=yrange, xlab="u=rank(x)/(n+1)", ylab="v=rank(y)/(n+1)",cex.lab=1.5) 
+op<-par(mar=c(0,3,1,1)) 
+barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
+op<-par(mar=c(3,0,1,1)) 
+barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
+par(op)
+dev.off()
 ```
@@
 
+<!--
 Next, the [**BMR data**](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) contains the trait contrasts from phylogenetically independent contrasts [@SukumaranH2010] of data reported by [@GenoudIM2017] on the basal metabolic rate (BMR) and mass of mammals species.
@@
 Absolute values of traits for each contrast are considered to construct corresponding copula for second and third datasets.
-
+--> 
 Fourth, we used **Ceder-creek data for e120** which resulted from a biodiversity experiment, carried out by Tilman and his co-warkers [@Tilman2001; @Tilman2006] over many years to collect [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120) data.

### 2018-10-10 Paper.Rmd
intro completed

@@
 It may be better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation, when the marginal distribution of the variables are highly non-Gaussian.
-Copula, a ranked based dependence measurement between two variables with initially some non-uniformly distributed marginals, transforms the marginals into uniform distribution when joint distribution of ranked data is plotted.
+Copula [@joe2014_dependence; @nelsen2007_copula], a ranked based dependence measurement between two variables with initially some non-uniformly distributed marginals, transforms the marginals into uniform distribution when joint distribution of ranked data is plotted.
 For example, in the first panel of Fig.
@@
 Similarly, an "upper" tail-dependence copula indicates stronger correlation between higher extreme variables in the raw data.
-The flexibility of copulas allows us to model more complex dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence, and reflection symmetry versus reflection asymmetry.
 Tail dependence occurs when the correlation between two variables increases as we get "further" in the tail (either or both) of the distribution.
-For example, in the upper panel of Fig.
-\ref{fig_cop_pedag}, different dependence structures, with same Kendall correlation $\tau = 0.5$, are shown which vary between dominantly lower tail dependent (lower tail - upper tail dependence being positive) from dominantly upper tail dependent (lower tail - upper tail dependence being negative) as we move along the row.
+For example, in the middle panel of Fig.
+\ref{fig_cop_pedag}, different dependence structures, with same Kendall correlation $\tau = 0.5$, are shown which vary between dominantly upper tail dependent (lower tail - upper tail dependence being negative) to dominantly lower tail dependent (lower tail - upper tail dependence being positive) as we move along the row.
 Lower panel represents the similar scenario but with stronger correlation (Kendall's $\tau = 0.8$).
+So, the flexibility of copulas allows us to model more complex multivariate dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence, and reflection symmetry versus reflection asymmetry.
 To implement this, we need to select candidate bivariate copulas from many existing parametric copula families.
@@
 If one variable is an environmental variable and another is an observable (like abundance of a species) then it can also explain the causal relationship between them.
-Here, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets.
+On a similar note, if a species' abundance time series from different sampling locations shows multivariate dependence structure, then we can analyze them as pairwise bivariate copula and find if there is any tail dependence or not.
+If those copula came up with asymmetric tail dependence, then it can be viewed as a measure of "spatial synchrony" [@Liebhold2004; @walteretal2017], which is a celebrated phenomenon in population ecology.
+In this article, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets.
 In addition to a copula based \nameref{Model_selection}, we also carried out a \nameref{NPA} with three different statistics to verify our model selection based results.
 
-Basically, we want to address the following questions.
+We want to address the following questions.
 \textbf{(Q1)} Do datasets in ecology and related fields have non-Normal copula srtructure?
@@
 And \textbf{(Q3)} what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
-In \nameref{Material&Methods}, we described briefly about each bivariate and multivariate datasets we used to answer these questions and provided a detailed descriptions of implemented methods.
+In \nameref{Data}, we described briefly about each bivariate and multivariate datasets we used to answer these questions and provided a detailed descriptions of implemented methods in \nameref{Methods}.
 In the \nameref{Results} section, we summarized our finidings for each datasets.
-Next two sections and are focused on possible causes and consequences of non-Normal copula structure in ecology and related field.
+Next two sections are focused on possible causes and consequences of non-Normal copula structure in ecology and related field with applications shown in real data as well as some simulations.
 Finally, we concluded in the \nameref{Discussion} section.
@@
 \includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
-\caption{Pedagogical figure showing BB1 copula with varying tail-dep difference but of same Kendall's tau along each row and with same tail-dep difference but of different Kendall's tau along each column.\label{fig_cop_pedag}}
+\caption{First panel : raw data plot has non-uniform (gamma distributed) marginal whereas copula makes its marginal uniform and indicating lower tail dependence in joint distribution, Second and third panel : pedagogical figure showing BB1 copula with varying tail-dependence difference but of same Kendall's tau along each row and with same tail-dependence difference but of different Kendall's tau along each column.\label{fig_cop_pedag}}
 \end{center}
@@
 
-# Material and methods \label{Material&Methods}
+<!--# Material and methods \label{Material&Methods}-->
 
@@
 
+<!--
 Considered datasets ranged widely from population ecology, community ecology , evolutionary and biogeochemical sphere. 
 We used four datasets in each category of bivariate (soil C and N data, BMR data, Anolis data, Ceder-creek experiment data) and multivariate (aphid-abundance data, aphid-firstflight data, plankton-abundance data and methane-flux data) examples.
-Details of each dataset used were given in \ref{SM-Data_detailed}.
+Details of each dataset used were given in \ref{SM-Data_detailed}.-->
+
+For bivariate analysis we used four datasets which are briefly described as follows;
+\textbf{(i)} to explore the dependence of **soil organic C** and total **soil N** across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
+database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
+stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
+study.
+Then, we used extensive set of basal metabolic rate data for \textbf{(ii)} birds [@Mcnab_2009] and \textbf{(iii)} mammals [@Mcnab_2008] as our second and third bivariate datasets to explore the dependence between log(BMR) and log(mass) of respective species.
+\textbf{(iv)} We used [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120) data from a biodiversity experiment [@Tilman2001; @Tilman2006] to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m), for year 2000.
+
+Similarly, we applied multivariate analysis on four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014].
+These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species.
+We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony.
+In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}).
+
+Our third multivariate dataset is a plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013).
+These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
+<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
+unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
+location in the ocean.
+Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979].
+All the samples available for a given variable in a given area in a given month of a given year were
+averaged, producing a monthly time series for each year.
+Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).--> [@Sheppard2017] converted monthly values to yearly values by averaging over all months.
+But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available).
+We used data for species=16 [Name : "Ceratium furca"] from those locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}).
+
+Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
+Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
+Additional information on the history of the site and soil types is detailed in @Jarecke2016.
+Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016.
+**See and Smyth et al, pending for
+additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.
 
@@
 
+\newpage
 ```{r tab_bivar_summary, echo=F, results='markup'}
@@
 
-We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}).
+We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL_simulation}).
 For example, our model simulation showed lower tail dependent Clayton copula had much higher slope (spatial-Taylor's law) than symmetric Normal copula, which in turn had higher slope than upper tail dependent Survival Clayton family (Fig.

### 2018-10-10 SupportingInformation.Rmd
intro completed

--- 2018-10-08 SupportingInformation.Rmd
@@
 source("mtime.R") #A function needed for caching
+```
+
+```{r cop_fig_intro,echo=F,results="hide",cache=T, cache.extra=list(seed,mtime("copsurrognd.R"))}
+library(copula)
+library(VineCopula)
+library(mvtnorm)
+set.seed(seed)
+
+source("copsurrognd.R")
+
+#x <- read.table("./Data/dataset/x.txt")
+#y <- read.table("./Data/dataset/y.txt")
+#mat<-cbind(x,y)
+#colnames(mat)<-c("x","y")
+
+#***constants
+N<-200 #length of time series
+n<-2 #number of sampling locations
+numdat<-1 #the number of datasets to get from the normal copula
+rho<-0.8 #normal-copula parameter for dependence between locations
+#this is a covariance, see below call to rmvnorm
+
+#***generate the starting data, which has normal copula structure
+#because it comes originally from a multivariate normal, before 
+#transformation
+sig<-matrix(rho,n,n)
+diag(sig)<-1
+m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
+m<-array(m,c(N,numdat,n))
+m<-aperm(m,c(1,3,2)) # 2nd dim is normally distributed e.g. hist(m[,1,])
+gshape<-7.5
+gscale<-1 #gamma parameters
+
+cop_to_marg_trans<-function(x){
+  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
+  #should be something skewed
+}
+# pnorm(m) makes normal to uniform
+# then qgamma(...) makes uniform to gamma distributed
+m<-cop_to_marg_trans(pnorm(m))
+
+ccop<-claytonCopula(2,2)
+csurrog<-array(NA,dim(m))
+csurrog[,,1]<-copsurrognd(m[,,1],targetcop=ccop,numsurrog=1)
+mat<-csurrog[,,1]
+
+pdf("./Results/rawdata.pdf",height = 4, width = 4)
+scale<-2
+xhist <- hist(mat[,1], breaks=30, plot=FALSE)
+yhist <- hist(mat[,2], breaks=30, plot=FALSE) 
+top <- max(c(xhist$counts, yhist$counts)) 
+xrange <- range(mat[,1])
+yrange <- range(mat[,2])
+nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
+op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0))  
+plot(mat[,1], mat[,2], xlim=xrange, ylim=yrange, xlab="x", ylab="y", cex.lab=1.5)
+legend("bottomright","raw data plot",bty="n",cex=1.5,text.col="magenta")
+op<-par(mar=c(0,3,1,1)) 
+barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
+op<-par(mar=c(3,0,1,1)) 
+barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
+par(op)
+dev.off()
+
+pdf("./Results/copdata.pdf",height = 4, width = 4)
+copmat<-pobs(mat)
+xhist <- hist(copmat[,1], breaks=30, plot=FALSE)
+yhist <- hist(copmat[,2], breaks=30, plot=FALSE) 
+top <- max(c(xhist$counts, yhist$counts)) 
+xrange <- c(0,1)
+yrange <- c(0,1)
+nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
+op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0)) 
+plot(copmat[,1], copmat[,2], xlim=xrange, ylim=yrange, xlab="u=rank(x)/(n+1)", ylab="v=rank(y)/(n+1)",cex.lab=1.5)
+legend("bottomright","copula plot",bty="n",cex=1.5,text.col="magenta")
+op<-par(mar=c(0,3,1,1)) 
+barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
+op<-par(mar=c(3,0,1,1)) 
+barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
+par(op)
+dev.off()
 ```
@@
 
-<!--
-## Methane-flux data  \label{Methane_data}
-
-Methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
-Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
-Additional information on the history of the site and soil types is detailed in @Jarecke2016.
-Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at
-daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
-additional details???** on greenhouse gas flux methods for this site.
-
-* We threw out all locations that had fewer than 50 dates data.-->
-
 ```{r read_methane_data, echo=F, results="markup"}
@@
 
-# Description of Data \label{Data_detailed}
+<!--# Description of Data \label{Data_detailed}
 
@@
 
-<!--The raw data and copula for each four datasets are shown along each column of Figure \ref{MT-fig_biv_raw_cop}.-->
+The raw data and copula for each four datasets are shown along each column of Figure \ref{MT-fig_biv_raw_cop}.
 
@@
 
-<!--Figure \ref{MT-fig_aphid_raw_cop} represents raw and copula data plot for aphid count ("Green spruce aphid") and aphid-firstflight date ("Leaf-curling plum aphid") taken from two locations.-->
+Figure \ref{MT-fig_aphid_raw_cop} represents raw and copula data plot for aphid count ("Green spruce aphid") and aphid-firstflight date ("Leaf-curling plum aphid") taken from two locations.
 
@@
 daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
-additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.
+additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.-->
 
@@
 
+\newpage
 ```{r tab_soilCN_asym,echo=F,results="markup"}
@@
 tab_multivar_summary<-cbind(m1,m2,m3,m4)
-colnames(tab_multivar_summary)<-c("Aphid_abundance","Aphid_firstflight","Plankton","Methane-flux")
+colnames(tab_multivar_summary)<-c("Green-spruce aphid_count","Leaf-plum curling aphid_firstflight","C. furca_count","Methane-flux")
 saveRDS(tab_multivar_summary,"./Results/tab_multivar_summary.RDS")
@@
 
-\newpage
 ```{r tab_stat_npa_asym_4,echo=F,results="markup"}
@@
 
-\newpage
+\clearpage
 
 # References
+
+<!--
 <div id="refs"></div>
+-->

### 2018-10-10 Paper.Rmd
modified

@@
 \includegraphics[width=4 cm]{./Results/BMR_results/mammals/copula_mammals_BMR.pdf}
-\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}
+\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}\\
+\end{center}
+\textbf{ \hspace{2.8 cm} (I) \hspace{3.2 cm} (J) \hspace{3.2 cm} (K) \hspace{3.2 cm} (L) } \\
+\vspace{-0.35 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
+\includegraphics[width=4 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
+\includegraphics[width=4 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}
+\includegraphics[width=4 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
 \caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs.
@@
 Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
-Lower panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel.
+Middle panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel.
 These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of species for next two datasets and number of plots for the last one (see \nameref{Model_selection} of manuscript).
-\label{fig_biv_raw_cop}}
+Lower panel shows \textbf{(I)} raw data , \textbf{(J)} copula plot for  green spruce aphid's count data from location pair(4,11) and \textbf{(K)} raw data , \textbf{(L)} copula plot for leaf-curling plum aphid's firstflight data from location pair(2,5).
+These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of data-points available for both location at same year (see \nameref{Model_selection} of manuscript)\label{fig_biv_multi_raw_cop}}
 \end{center}
@@
 \end{center}
-\end{figure}
+\end{figure}-->
 
@@
 
-As described in the Appendix \ref{SM-Data_detailed}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
-\ref{fig_biv_raw_cop}(A) and (E), respectively.
+As described in the Appendix \nameref{Data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
+\ref{fig_biv_multi_raw_cop}(A) and (E), respectively.
 We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.
@@
 We summarized our results columnwise in Table \ref{tab_bivar_summary} for different kinds of bivariate datasets (biogeochemical, evolutionary and ecological data) that were plotted in Fig.
-\ref{fig_biv_raw_cop}.
+\ref{fig_biv_multi_raw_cop}(A-H).
 
@@
 Fig.
-\ref{fig_aphid_raw_cop}(A) and (C) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.
+\ref{fig_biv_multi_raw_cop}(A) and (B) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.
 
@@
 
-\newpage
 ```{r tab_bivar_summary, echo=F, results='markup'}

### 2018-10-10 SupportingInformation.Rmd
modified

@@
 op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Aphid abundance from location 4",ylab="Aphid abundance from location 11",cex.lab=1.5,cex.axis=1.5)
+plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Green-spruce aphid count from location 4",ylab="Green-spruce aphid count from location 11",cex.lab=1.5,cex.axis=1.5)
 par(op)
@@
 op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Aphid first-flight from location 2",ylab="Aphid first-flight from location 5",cex.lab=1.5,cex.axis=1.5)
+plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Leaf-plum curling aphid's first-flight from location 2",ylab="Leaf-plum curling aphid's first-flight from location 5",cex.lab=1.5,cex.axis=1.5)
 par(op)

### 2018-10-10 SupportingInformation.Rmd
modified

@@
 
-\newpage
 ```{r tab_soilCN_asym,echo=F,results="markup"}
@@
 
+\newpage
 <!--Display model selection results for BMR:birds data in a table-->
@@
 
+\newpage
 ```{r tab_count_AICw_normal, echo=F, results='markup'}
@@
 ```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
-knitr::kable(summary_stat[,c(1:4)], booktabs = T, 
+summary_stat<-summary_stat[,c(1:4)]
+rownames(summary_stat)<-c("bestcop.AICw","N.AICw","relLTdep.AICw","relUTdep.AICw","p.CvM","p.KS","BS.success.percent")
+knitr::kable(summary_stat, booktabs = T, 
              format = "pandoc",
@@
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption="summary of model selection results for asym. sensitivity \\label{tab_summary_stat_asym_sens}")
+      caption="summary of model selection results for asymmetric sensitivity : \\textbf{bestcop.AICw} and \\textbf{N.AICw} represent the AIC-weightage for best fit copula and Normal copula, \\textbf{relLTdep.AICw} and \\textbf{relLTdep.AICw} represent relative lower tail and upper tail dependence based on AIC-weightage respectively, \\textbf{p.CvM} and \\textbf{p.KS} are the p-values coming from goodness of fit test (Cramer-von Mises and Kolmogorov-Smirnov statistics based), \\textbf{BS.success.percent} percentage of successful bootstraps in goodness of fit test.
+Each column summarizes those statistics for each copula plotted in Fig.
+\\ref{MT-fig_ms_asym_b_0.1}.\\label{tab_summary_stat_asym_sens}")
 ```
@@
 
+\newpage
 ```{r tab_stat_npa_asym_1,echo=F,results="markup"}
@@
 
-\newpage
 <!--results='hide'-->
@@
 
-![Extinction_Risk_local_dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp}](./Results/ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Extinction_Risk_global_dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp}](./Results/ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Extinction_Risk_numlocs_effect_local_disp; numsteps=25, numsims=10000\label{fig_risk_numlocs_effect_local_disp}](./Results/ext_risk_numlocs_effect_local_disp.pdf){ width=40%, height=40% }
-
-![Extinction_Risk_numlocs_effect_global_disp; numsteps=25, numsims=10000\label{fig_risk_numlocs_effect_global_disp}](./Results/ext_risk_numlocs_effect_global_disp.pdf){ width=40%, height=40% }
-
-![Extinction_Risk_K_effect on Ricker model; r=0.25, numsteps=25, numsims=10000\label{fig_risk_K_effect}](./Results/ext_risk_K_effect.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion (see Eq.
+\ref{eq.extrisk_model}); numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp}](./Results/ext_risk_local_disp.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with global dispersion (see Eq.
+\ref{eq.extrisk_model}); numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp}](./Results/ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+![Increasing patch number allows metapopulation to disperse more efficiently and thereby decresing extinction risk calculated after 25 years; models with local dispersion with numsims=10000 (see Eq.
+\ref{eq.extrisk_model}).\label{fig_risk_numlocs_effect_local_disp}](./Results/ext_risk_numlocs_effect_local_disp.pdf){ width=40%, height=40% }
+
+![Increasing patch number allows metapopulation to disperse more efficiently and thereby decresing extinction risk calculated after 25 years; models with global dispersion with numsims=10000 (see Eq.
+\ref{eq.extrisk_model}).\label{fig_risk_numlocs_effect_global_disp}](./Results/ext_risk_numlocs_effect_global_disp.pdf){ width=40%, height=40% }
+
+![Extinction risk as calculated after 25 years on Ricker model (r=0.25, numsims=10000) decreases as carrying capacity $K$ increases for both local and global dispersion (see Eq.
+\ref{eq.extrisk_model}).\label{fig_risk_K_effect}](./Results/ext_risk_K_effect.pdf){ width=40%, height=40% }
 

### 2018-10-16 Paper.Rmd
fig arranged properly
reply : Lawrence comment...continued

@@
 Pearson correlation because of its ease in use.
-But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence because it is a measure of linear association between two variables.
+But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence (i.e., varying association between two variables) because it is a measure of linear association between two variables.
 It may be better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation, when the marginal distribution of the variables are highly non-Gaussian.
-Copula [@joe2014_dependence; @nelsen2007_copula], a ranked based dependence measurement between two variables with initially some non-uniformly distributed marginals, transforms the marginals into uniform distribution when joint distribution of ranked data is plotted.
+The copula [@joe2014_dependence; @nelsen2007_copula], a ranked based dependence measurement between two variables, transforms non-uniformly distributed marginals in original variables to uniformly distributed marginals when joint distribution of ranked data is plotted.
 For example, in the first panel of Fig.
@@
 Copula plot was obtained by scaling individual ranked variables by a factor $(n+1)$ where $n$ was the number of sampling points, so that the copula was confined within a unit box.
-However, using a "copula-based approach" as a dependence measure reflects more knowledge about the dependence structure instead of summarizing in just a single number, the linear correlation coefficient.
+From now, we will call it as "normalized ranking".
+Using a "copula-based approach" as a dependence measure reflects more knowledge about the dependence structure instead of summarizing in just a single number, the linear correlation coefficient.
 It tells us about the tail-dependence of the variables, i.e., the association between extreme values of two variables.
-If smaller values of each variable are more correlated to each other, then the corresponding copula structure is said to have "lower" tail-dependence as shown in the first panel of Fig.
+If smaller values of each variable are more correlated to each other, then the corresponding copula structure is said to have "lower tail-dependence" as shown in the first panel of Fig.
 \ref{fig_cop_pedag} copula plot (right).
-Similarly, an "upper" tail-dependence copula indicates stronger correlation between higher extreme variables in the raw data.
+Similarly, an "upper tail-dependence" copula indicates stronger correlation between higher extreme variables in the raw data.
 Tail dependence occurs when the correlation between two variables increases as we get "further" in the tail (either or both) of the distribution.
@@
 Lower panel represents the similar scenario but with stronger correlation (Kendall's $\tau = 0.8$).
-So, the flexibility of copulas allows us to model more complex multivariate dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence, and reflection symmetry versus reflection asymmetry.
-To implement this, we need to select candidate bivariate copulas from many existing parametric copula families.
+So, the flexibility of copulas allows us to model more complex multivariate dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence.<!--and reflection symmetry versus reflection asymmetry.--> To identify the kinds of copula structure found in real ecological data, we need to select candidate bivariate copulas from many existing parametric copula families.
 <!--A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulas to be selected from.--> Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), mentioned in the \nameref{Methods}.
@@
 Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data.
+
 In ecological perspective, when we find a dominantly lower (upper) tail dependence structure between two species abundance timeseries, it means both incidence are rare (abundant) synchronously.
-If one variable is an environmental variable and another is an observable (like abundance of a species) then it can also explain the causal relationship between them.
-On a similar note, if a species' abundance time series from different sampling locations shows multivariate dependence structure, then we can analyze them as pairwise bivariate copula and find if there is any tail dependence or not.
-If those copula came up with asymmetric tail dependence, then it can be viewed as a measure of "spatial synchrony" [@Liebhold2004; @walteretal2017], which is a celebrated phenomenon in population ecology.
+Again, if abundance of a species (say, species A) is associated with scarcity of another species (say, species B) then it can explain the causality or negative feedback of A on species B.
+If one variable is an environmental factor (like temperatute or rainfall) and another is an observable (like abundance of a species) then it can also explain the causal relationship between them.
+On a similar note, if a species' abundance time series from different sampling locations shows multivariate dependence structure, then we can analyze the "spatial data" as pairwise bivariate copula and find if there is any tail dependence or not.
+If those copula came up with asymmetric tail dependence, then it can be viewed as a new type to quantify "spatial synchrony" [@Liebhold2004; @walteretal2017], which is a celebrated phenomenon in population ecology.
 In this article, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets.
@@
 
-Similarly, we applied multivariate analysis on four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014].
+Similarly, we applied multivariate analysis on four spatial datasets; \textbf{(i)} aphid-count and \textbf{(ii)} aphid-firstflight date were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014].
 These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species.
@@
 In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}).
-
-Our third multivariate dataset is a plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013).
+\textbf{(iii)} Plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013).
 These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
@@
 We used data for species=16 [Name : "Ceratium furca"] from those locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}).
-
-Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
-Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
+\textbf{(iv)} Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
 Additional information on the history of the site and soil types is detailed in @Jarecke2016.
@@
 additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.
-
-\begin{figure}[!h]
-\textbf{ \hspace{2.8 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
-\vspace{-0.35 cm}
-\begin{center}
-\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/birds/raw_birds_BMR.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/mammals/raw_mammals_BMR.pdf}
-\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
-\end{center}
-\textbf{ \hspace{2.8 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
-\vspace{-0.35 cm}
-\begin{center}
-\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/birds/copula_birds_BMR.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/mammals/copula_mammals_BMR.pdf}
-\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}\\
-\end{center}
-\textbf{ \hspace{2.8 cm} (I) \hspace{3.2 cm} (J) \hspace{3.2 cm} (K) \hspace{3.2 cm} (L) } \\
-\vspace{-0.35 cm}
-\begin{center}
-\includegraphics[width=4 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
-\includegraphics[width=4 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
-\includegraphics[width=4 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}
-\includegraphics[width=4 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
-\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs.
-log(mass) for birds, \textbf{(C)} log(BMR) vs.
-log(mass) for mammals, \textbf{(D)} above-ground plant biomass data vs.
-Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
-Middle panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel.
-These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of sampling locations for soil C and N data and number of species for next two datasets and number of plots for the last one (see \nameref{Model_selection} of manuscript).
-Lower panel shows \textbf{(I)} raw data , \textbf{(J)} copula plot for  green spruce aphid's count data from location pair(4,11) and \textbf{(K)} raw data , \textbf{(L)} copula plot for leaf-curling plum aphid's firstflight data from location pair(2,5).
-These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of data-points available for both location at same year (see \nameref{Model_selection} of manuscript)\label{fig_biv_multi_raw_cop}}
-\end{center}
-\end{figure}
-
-<!--
-\begin{figure}[!h]
-\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
-\vspace{-0.4 cm}
-\begin{center}
-\includegraphics[width=7 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
-\includegraphics[width=7 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
-\end{center}
-\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
-\vspace{-0.4 cm}
-\begin{center}
-\includegraphics[width=7 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
-\includegraphics[width=7 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
-\caption{Upper panel shows raw data plots for \textbf{(A)} green spruce aphid-abundance data from location pair(4,11) and \textbf{(B)} leaf-curling plum aphid-firstflight data from location pair(2,5).
-Lower panel \textbf{(C - D)} shows copula plots for the corresponding datasets of the upper panel.
-These copula plots are obtained by applying ranks separately to each variable, and then dividing the ranks by $n+1$ where $n$ is the number of data-points available for both location at same year (see \nameref{Model_selection} of manuscript).
-\label{fig_aphid_raw_cop}}
-\end{center}
-\end{figure}-->
 
@@
 $\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
-we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; further work was with the $u_i$ 
+we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; further work was with the "normalized ranked variables" $u_i$ 
 and $v_i$. Ranks $\tilde{u}_i$ and $\tilde{v}_i$ were $1$, for the 
@@
 For multivariate datasets, e.g., the aphid abundance or first flight data, 
-each of which consists of time series from 10 locations, we carried out the same bivariate analysis
+each of which consists of time series from 10 selected locations, we carried out the same bivariate analysis
 for all possible pairwise combinations of those locations and compared AIC-weights and tail-dependence of best-fit(minimum AIC) copula with that of the Normal copula.
 We also counted for how many location pairs model -averaged lower tail dependence is greater than model-averaged upper-tail dependence or vice-versa.
-For example, for green-spruce aphid abundance data we found major lower tail dependence whereas leaf-curling plum aphid's first flight data had significant upper tail dependence at all location pairs (for details see Table \ref{tab_multivar_summary}).
-<!--***DAN: need to fill this in, prrobably enough to be brief, since these were just done 
-pairwise-->
-<!--**SG: done-->
+
+<!--For example, for green-spruce aphid abundance data we found major lower tail dependence whereas leaf-curling plum aphid's first flight data had significant upper tail dependence at all location pairs (for details see Table \ref{tab_multivar_summary}).
+***DAN: need to fill this in, prrobably enough to be brief, since these were just done 
+pairwise
+**SG: done-->
 
@@
 
-# Application : copula structures in ecology, evolutionary and biogeochemical spheres \label{Results}
-
+<!--# Application : copula structures in ecology, evolutionary and biogeochemical spheres \label{Results}-->
+# Application : biological dependencies \label{Results}
 <!--## question 1
@@
 
-## Example for bivariate case : soil C and N data \label{Ex_SoilCN}
+## Results for bivariate data \label{Result_biv}
+
+We summarized our results columnwise in Table \ref{tab_bivar_summary} for different kinds of bivariate datasets (biogeochemical, evolutionary and ecological data) that were plotted in Fig.
+\ref{fig_biv_multi_raw_cop}(A-H).
+For all $4$ bivariate-datasets indicated by each column name (SoilCN, Birds' BMR, Mammals' BMR, Ceder-creek) in Table \ref{tab_bivar_summary}, data were always significantly correlated as indicated by p-value (<0.05) from `BiCopIndTest`{`VineCopula`} (1$^{st}$ row of Table \ref{tab_bivar_summary}).
+Non-Normal copula were always emarged as best fit copula based on minimum AIC value (2$^{nd}$ row of Table \ref{tab_bivar_summary}).
+Normal copula were found as poor fit compared to highest AIC-weight based asymmetric copulas (3$^{rd}$-4$^{th}$ rows of Table \ref{tab_bivar_summary}) - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}.
+Goodness of fit of the best-fit asymmetric copulas were tested based on p-values from two statistics (Cramer-von Mises and Kolmogorov-Smirnov : 5$^{th}$-6$^{th}$ rows of Table \ref{tab_bivar_summary}).
+Among 4 datasets presented there, first three have significant upper tail dependence whereas the last one has significant lower tail dependence compared to Normal (7$^{th}$-9$^{th}$ rows of Table \ref{tab_bivar_summary}).
+It was indicated by negative AIC-weights model averaged (lower-upper) tail dependence and very low ranks of actual copula 
+data for $\Ps_l - \Ps_u$ statistics for first 3 datasets and positive AIC-weights model averaged 
+(lower-upper) tail dependence and very high ranks of actual copula data for $\Ps_l - \Ps_u$ statistics for the last one, which answers the second and third part of \hyperref[Intro]{\textbf{(Q1)}} (For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and Figs.
+\ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}).
+Our model-selection results were consistent with non-parametric analysis for all the datasets (compare last two rows of Table \ref{tab_bivar_summary}).
+We will now discuss briefly findings from one of our bivariate datasets as follows.
+
+### Example for bivariate case : soil C and N data \label{Ex_SoilCN}
 
@@
 
-As described in the Appendix \nameref{Data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
+As described in the \nameref{Data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
 \ref{fig_biv_multi_raw_cop}(A) and (E), respectively.
@@
 
-We summarized our results columnwise in Table \ref{tab_bivar_summary} for different kinds of bivariate datasets (biogeochemical, evolutionary and ecological data) that were plotted in Fig.
-\ref{fig_biv_multi_raw_cop}(A-H).
-
-For all $3$ bivariate-datasets indicated by each column name (SoilCN, Birds' BMR, Ceder-creek) 
-in Table \ref{tab_bivar_summary}, Normal copula were always a poor fit compared to 
-highest AIC-weight based asymmetric copulas - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}. 
-Among three datasets presented there, first two have significant upper tail dependence whereas the last one has significant lower tail dependence compared to Normal.
-It was indicated by negative AIC-weights model averaged
-(lower-upper) tail dependence and very low ranks of actual copula 
-data for $\Ps_l - \Ps_u$ statistics for first two datasets and positive AIC-weights model averaged 
-(lower-upper) tail dependence and very high ranks of actual copula 
-data for $\Ps_l - \Ps_u$ statistics for the last one, which answers the second and third part of \hyperref[Intro]{\textbf{(Q1)}}
-(For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and 
-Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}).
-
-## Example for multivariate case : abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
-
+## Results for multivariate spatial data \label{Result_multi}
 ```{r read_RES_aphid_count, echo=F}
@@
 ```
+
+We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets 
+(i.e., abundance data for green spruce aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
+firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
+abundance data for plankton from UK seas from $14$ sampling locations, 
+methane flux data from $13$ sampling locations).
+
+Likewise, green-spruce aphid abundance data ($1^{st}$ column in Table \ref{tab_multivar_summary}), 
+plankton data from UK seas also showed major lower tail dependent spatial synchrony 
+($3^{rd}$ column in Table \ref{tab_multivar_summary}), whereas firstflight data for leaf-curling plum aphid 
+showed stronger upper tail dependence at every location pair ($2^{nd}$ column in Table \ref{tab_multivar_summary}).
+Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.
+
+### Example for multivariate case : abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
 
@@
 
-We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets 
-(i.e., abundance data for green spruce aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
-firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
-abundance data for plankton from UK seas from $14$ sampling locations, 
-methane flux data from $13$ sampling locations).
-
-Likewise, green-spruce aphid abundance data ($1^{st}$ column in Table \ref{tab_multivar_summary}), 
-plankton data from UK seas also showed major lower tail dependent spatial synchrony 
-($3^{rd}$ column in Table \ref{tab_multivar_summary}), whereas firstflight data for leaf-curling plum aphid 
-showed stronger upper tail dependence at every location pair ($2^{nd}$ column in Table \ref{tab_multivar_summary}).
-Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.
-
 ```{r tab_bivar_summary, echo=F, results='markup'}
@@
              format='pandoc',
-             caption = "Summary of results for bivariate data-sets : \\textbf{Indep.p}: p-value from `BiCopIndTest`, \\textbf{AIC-best.copula}: best copula based on minimum AIC, \\textbf{Highest.AICw}: AIC-weight of that best copula, \\textbf{N.AICw}: AIC-weight of Normal copula, \\textbf{p.CvM}: p-value from `BiCopGofTest` for Cramer-von Mises method, \\textbf{p.KS}: p-value from `BiCopGofTest` for Kolmogorov-Smirnov method,\\textbf{LT.model.avg}: AIC-weights model averaged lower tail dependence, \\textbf{UT.model.avg}: AIC-weights model averaged upper tail dependence, \\textbf{(LT-UT).model.avg}:  AIC-weights model averaged (lower-upper) tail dependence.
-$\\Ps_l$, $\\Ps_u$, $\\Ps_l$ - $\\Ps_u$ for the actual copula data and for Kendall-correlation preserving $1000$ Normal surrogates were computed and the rank for actual data (\\textbf{Rank.Pl.K, Rank.Pu.K, Rank.Pl-Pu.K}) were calculated as the position if all these $1001$ values were arranged in ascending order.
-Ranks if were $>975$ or $<25$, then indicated a significantly ($95\\%$ CI) different tail-dependence measures based on Normal-null hypothesis.
+             caption = "Summary of results for bivariate data-sets : \\textbf{p.Independence.test}: p-value from `BiCopIndTest`, \\textbf{minimumAIC.bestfit.copula}: best fit copula based on minimum AIC, \\textbf{Bestfit.copula.AICw}: AIC-weight of that best fit copula, \\textbf{Normal.copula.AICw}: AIC-weight of Normal copula, \\textbf{p.CvM}: p-value from `BiCopGofTest` for Cramer-von Mises method, \\textbf{p.KS}: p-value from `BiCopGofTest` for Kolmogorov-Smirnov method,\\textbf{LT.model.avg}: AIC-weights model averaged lower tail dependence, \\textbf{UT.model.avg}: AIC-weights model averaged upper tail dependence, \\textbf{(LT-UT).model.avg}:  AIC-weights model averaged (lower-upper) tail dependence.
+Statistic $\\Ps_l$ - $\\Ps_u$ for the actual copula data and for Kendall-correlation preserving $1000$ Normal surrogates were computed and ranked in ascending order.
+Rank for real data \\textbf{Rank.(Pl-Pu)}, if were $>975$ or $<25$, then indicated a significantly ($95\\%$ CI) different lower or upper tail-dependent measures based on Normal-null hypothesis.
 \\label{tab_bivar_summary}",
@@
 
+\begin{figure}[!h]
+\textbf{ \hspace{2.8 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
+\vspace{-0.35 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/birds/raw_birds_BMR.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/mammals/raw_mammals_BMR.pdf}
+\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
+\end{center}
+\textbf{ \hspace{2.8 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
+\vspace{-0.35 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/birds/copula_birds_BMR.pdf}
+\includegraphics[width=4 cm]{./Results/BMR_results/mammals/copula_mammals_BMR.pdf}
+\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}\\
+\end{center}
+\textbf{ \hspace{2.8 cm} (I) \hspace{3.2 cm} (J) \hspace{3.2 cm} (K) \hspace{3.2 cm} (L) } \\
+\vspace{-0.35 cm}
+\begin{center}
+\includegraphics[width=4 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
+\includegraphics[width=4 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
+\includegraphics[width=4 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}
+\includegraphics[width=4 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
+\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs.
+log(mass) for birds, \textbf{(C)} log(BMR) vs.
+log(mass) for mammals, \textbf{(D)} above-ground plant biomass data vs.
+Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
+Middle panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel.
+Lower panel shows \textbf{(I)} raw data , \textbf{(J)} copula plot for  green spruce aphid's count data from location pair(4,11) and \textbf{(K)} raw data , \textbf{(L)} copula plot for leaf-curling plum aphid's firstflight data from location pair(2,5).\label{fig_biv_multi_raw_cop}}
+\end{center}
+\end{figure}
+
 <!--Question 2 analyses-->
@@
 This section is dedicated to explore question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}. 
-We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change?
+<!--We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change?-->
 
 <!--Moran effects-->
-\textbf{(i)} To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
+\textbf{(i)} We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches.
+Should their dependence be exactly the same as of the noises or not ?
+
+To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
 \ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.
@@
 To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$.
-We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches.
-Should their dependence be exactly the same as of the noises or not ?
 
-To test that our approach is of two way : first, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
+To test environmental effect on population copula, our approach is of two way :
+
+* First, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
 Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome which was a bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
@@
 
-Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
+* Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
 Then we plotted mean of each measured quantity with standard error against Kendall's Tau of the noise copula family.

### 2018-10-16 SupportingInformation.Rmd
reply : Lawrence comment...continued

@@
 
-  m<-matrix(NA,1,12)
-  colnames(m)<-c("Indep.p","AIC-best.copula","Highest.AICw",
-                 "N.AICw","p.CvM","p.KS",
+  m<-matrix(NA,1,10)
+  colnames(m)<-c("p.Independence.test","minimumAIC.bestfit.copula","Bestfit.copula.AICw",
+                 "Normal.copula.AICw","p.CvM","p.KS",
                  "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
-                 "Rank.Pl.K","Rank.Pu.K","Rank.Pl-Pu.K")
+                 "Rank.(Pl-Pu)")
   m<-as.data.frame(m)
 
-  m$Indep.p<-round(r$IndepTestRes,3)
-  m$`AIC-best.copula`<-r$InfCritRes$copname[which.min(r$InfCritRes$AIC)]
-  m$Highest.AICw<-round(max(r$InfCritRes$AICw),3)
-  m$N.AICw<-round(r$InfCritRes$AICw[1],3)
+  m$p.Independence.test<-round(r$IndepTestRes,3)
+  m$minimumAIC.bestfit.copula<-r$InfCritRes$copname[which.min(r$InfCritRes$AIC)]
+  m$Bestfit.copula.AICw<-round(max(r$InfCritRes$AICw),3)
+  m$Normal.copula.AICw<-round(r$InfCritRes$AICw[1],3)
   m$p.CvM<-r$GofRes_CvM
@@
   m$`(LT-UT).model.avg`<-m$LT.model.avg-m$UT.model.avg
-  m$Rank.Pl.K<-rs$Rank_Pl_K
-  m$Rank.Pu.K<-rs$Rank_Pu_K
-  m$`Rank.Pl-Pu.K`<-rs$Rank_PlmPu_K
+  m$`Rank.(Pl-Pu)`<-rs$Rank_PlmPu_K
   mt<-as.data.frame(t(m))
@@
 #---------------------------------------
-pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf", width=8, height=6)
-op<-par(mfrow=c(2,2))
+pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf", width=8, height=3)
+op<-par(mfrow=c(1,2))
 
 Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=500,BS=1000)
 #Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=500,BS=1000)
@@
 Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=500,BS=1000)
 #Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=500,BS=1000)
@@
 plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-legend("top", c("noise", "population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
-       cex = 1.5, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
+legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", "magenta", "green2"),
+       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
        bty = "n") 
@@
 #----------------------------
-pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=8, height=6)
+pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=8, height=3)
 op<-par(mfrow=c(2,2))
@@
 Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=2500,BS=1000)
 #Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
@@
 Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
-Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=2500,BS=1000)
 #Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
@@
 
-op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0), new = TRUE)
+op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
 plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-legend("top", c("noise", "population", "p_CvM", "p_KS"), col = c("red", "blue", "magenta", "green2"),
-       cex = 1.5, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0), 
+legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", "magenta", "green2"),
+       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
        bty = "n") 
@@
 #---------------------------------------
-pdf("./Results/Cause4copula_results/Clayton_spearman.pdf")
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500)
-dev.off()
-
-pdf("./Results/Cause4copula_results/Clayton_kendall.pdf")
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="kendall",lb=0,ub=0.1,num_keep_last=500)
-dev.off()
-
-pdf("./Results/Cause4copula_results/SClayton_spearman.pdf")
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="spearman",lb=0,ub=0.1,num_keep_last=500)
-dev.off()
-
-pdf("./Results/Cause4copula_results/SClayton_kendall.pdf")
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="kendall",lb=0,ub=0.1,num_keep_last=500)
-dev.off()
-
-pdf("./Results/Cause4copula_results/Gumbel_spearman.pdf")
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=4,method="spearman",lb=0,ub=0.1,num_keep_last=500)
-dev.off()
-
-pdf("./Results/Cause4copula_results/Gumbel_kendall.pdf")
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=4,method="kendall",lb=0,ub=0.1,num_keep_last=500)
-dev.off()
-
-pdf("./Results/Cause4copula_results/SGumbel_spearman.pdf")
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=14,method="spearman",lb=0,ub=0.1,num_keep_last=500)
-dev.off()
-
-pdf("./Results/Cause4copula_results/SGumbel_kendall.pdf")
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=14,method="kendall",lb=0,ub=0.1,num_keep_last=500)
-dev.off()
+resloc<-"./Results/Cause4copula_results/"
+if (!dir.exists(paste(resloc,"Cause4copula_stat_results",sep=""))){
+  dir.create(paste(resloc,"Cause4copula_stat_results",sep=""))
+  resloc<-"./Results/Cause4copula_results/Cause4copula_stat_results/"
+}
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)                    
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=4,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=4,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=14,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=14,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
 ```

### 2018-10-16 Cause4copula.R

@@ -276,8 +276,9 @@ Plotter_Cause4copula_GOF<-function(N,fcode,method,num_keep_last,BS){
 #       ub : upper bound for Non-parametric stat function (Default =0.1)
 #       num_keep_last : number of rows you want to keep from bottom 
 #                     for each of s$noise_c, s$noise_q and s2$pop_c, s2$pop_q matrix
+#       resloc : folder location where the plots should be saved
 #-----------------------------------------------------------------------------------------------
-Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_keep_last){
+Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_keep_last,resloc){
   
   corcoef_list<-seq(from=0.1,to=0.9,by=0.1)
   
@@ -467,8 +468,9 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
   }else{
     warning("specify method",immediate.=T,call.=T)
   }
-  
-  op<-par(mfrow=c(4,3),mar=c(2.6,3.5,2,3.5), mgp=c(1.5,0.5,0))
+ 
+  pdf(paste0(resloc,BiCopName(fcode,short=F),"_Spearman_vs_",xlabel,".pdf",sep=""),height=4,width=5)
+  op<-par(mar=c(2.6,3.5,2,3.5), mgp=c(1.5,0.5,0))
   plot(corcoef_list,S_noise_mat[,2],cex=0.5,col="red",xlab=xlabel,ylab="Spearman",xlim=c(0,1),ylim=c(0,1))
   segments(corcoef_list,S_noise_mat[,1],corcoef_list,S_noise_mat[,3],col='red')
   bar_len<-0.02
@@ -484,7 +486,20 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
   lines(range(0,1),c(0.05,0.05),type='l',lty='dashed',col='purple')
   axis(side=4,col='purple',col.axis="purple")
   mtext(side = 4, line = 1.5, 'p values', col='purple')
[…]

### 2018-10-17 Paper.Rmd
rearranging figures

@@
 
-* First, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence and Survival Clayton with upper tail dependence here) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
-Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome which was a bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
+* First, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
+Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
 \ref{fig_Cause4copula_params_GOF_500}).
-We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size and reported p-values for both statistics (Cramer-von Mises and Kolmogorov-Smirnov).
+We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size.
+We reported p-values from goodness of fit tests for both statistics (Cramer-von Mises and Kolmogorov-Smirnov).
 p-values greater or equal to $0.05$ (above the dashed line in Fig.
@@
 
-* Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
+* Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence difference (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
 Then we plotted mean of each measured quantity with standard error against Kendall's Tau of the noise copula family.
 Fig.
-\ref{fig_Cause4copula_C_kend} represents this picture for noise generating from a lower tail dependent Clayton copula (for upper tail dependent Survival Clayton copula see Fig.
-\ref{SM-fig_Cause4copula_SC_kend}) whereas for Fig.
-\ref{SM-fig_Cause4copula_C_spear} and Fig.
-\ref{SM-fig_Cause4copula_SC_spear} Spearman's Rho of noise copula were plotted along x-axis.
-We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach, here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise.
+\ref{fig_Cause4copula_C_kend} represents this picture for noise generating from a lower tail dependent Clayton copula.
+We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach (model selection method), here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise.
 (For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
+
+<!--
+whereas for Fig. \ref{SM-fig_Cause4copula_C_spear} and Fig. \ref{SM-fig_Cause4copula_SC_spear} Spearman's Rho of noise copula were plotted along x-axis. -->
 
@@
 \begin{center}
-\includegraphics[width=16 cm]{./Results/Cause4copula_results/Clayton_kendall.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=15 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/common_legend_cause4copula_stat.pdf}
 \caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.

### 2018-10-17 SupportingInformation.Rmd
rearranging figures

@@
 #---------------------------------------
-# I need to fix that SG_kendall ylimit error : I think I fixed it
-#----------------------------
+
 pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=8, height=3)
-op<-par(mfrow=c(2,2))
+op<-par(mfrow=c(1,2))
 
@@
   dir.create(paste(resloc,"Cause4copula_stat_results",sep=""))
-  resloc<-"./Results/Cause4copula_results/Cause4copula_stat_results/"
-}
+}
+resloc<-paste(resloc,"Cause4copula_stat_results/",sep="")
+
 Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
@@
 Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)                
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=4,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=4,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=14,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=14,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
 ```
@@
 \label{fig_Cause4copula_params_GOF_2500}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_kendall.pdf}
-\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
-p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae.
-\label{fig_Cause4copula_SC_kend}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/Cause4copula_results/Clayton_spearman.pdf}
-\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
-p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae.
-\label{fig_Cause4copula_C_spear}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/Cause4copula_results/SClayton_spearman.pdf}
-\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
-p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae.
-\label{fig_Cause4copula_SC_spear}}
 \end{center}

### 2018-10-19 Paper.Rmd
rearranging plots and some stuffs..continued

@@
 Fig.
-\ref{fig_Cause4copula_C_kend} represents this picture for noise generating from a lower tail dependent Clayton copula.
+\ref{fig_Cause4copula_C_kend} compares the statistics from population copula which was generated by noises drawn from lower tail dependent Clayton copula.
+It is possible to measure the statistics with variation of Spearman's Rho for any other noise copula family (e.g.
+see Fig.
+\ref{SM-fig_Cause4copula_SC_spear}).
 We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach (model selection method), here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise.
@@
 
-<!--
-whereas for Fig. \ref{SM-fig_Cause4copula_C_spear} and Fig. \ref{SM-fig_Cause4copula_SC_spear} Spearman's Rho of noise copula were plotted along x-axis. -->
-
 \begin{figure}[!h]
 \begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Spearman_vs_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=15 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/common_legend_cause4copula_stat.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
 \caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
-p-values $< 0.05$ (below horizontal dashed line) indicates significantly similar estimate between noise and population copulae.
+Spearman, Kendall, Pearson) and 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+p-values $< 0.05$ (below horizontal purple line) indicates significantly similar estimate between noise and population copulae.
 \label{fig_Cause4copula_C_kend}}
@@
 \begin{equation}\label{eq.AS}
-P_{i}(t+1) = b P_{i}(t) + g(\epsilon_{i}(t)) + h(\delta_{i}(t))
+P_{i}(t+1) = b P_{i}(t) + g(\varepsilon_{i}(t)) + h(\delta_{i}(t))
 \end{equation}
@@
 $b$ was the linear growth rate.
-Global environmental fluctuation function $g(\epsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq.
-\ref{eq.AS_g}) where $\epsilon=(\epsilon_{1}, \epsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\epsilon}$.
+Global environmental fluctuation function $g(\varepsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq.
+\ref{eq.AS_g}) where $\varepsilon=(\varepsilon_{1}, \varepsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\varepsilon}$.
 
 \begin{align}
-g(\epsilon) = \begin{cases} 
+g(\varepsilon) = \begin{cases} 
 g_{1} = \begin{cases}
-\epsilon & \text{if $\epsilon< 0$}; \\
+\varepsilon & \text{if $\varepsilon< 0$}; \\
 0 & \text{otherwise}.
@@
 g_{2} = \begin{cases}
-\epsilon & \text{if $\epsilon> 0$}; \\
+\varepsilon & \text{if $\varepsilon> 0$}; \\
 0 & \text{otherwise}\label{eq.AS_g}.
 \end{cases}
-\end{cases} \text{where } \epsilon = N(\textbf{0},\Sigma_{\epsilon})\;\ \text{with} \;\ \Sigma_{\epsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
+\end{cases} \text{where } \varepsilon = N(\textbf{0},\Sigma_{\varepsilon})\;\ \text{with} \;\ \Sigma_{\varepsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
 \end{align}

### 2018-10-19 SupportingInformation.Rmd
rearranging plots and some stuffs..continued

@@
 
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Spearman_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Kendall_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pearson_vs_Spearman's Rho.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Corl-Coru_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pl-Pu_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_D2u-D2l_vs_Spearman's Rho.pdf}
+\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from upper-tail dependent \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+p-values $< 0.05$ (below horizontal purple line) indicates significantly similar estimate between noise and population copulae.
+\label{fig_Cause4copula_SC_spear}}
+\end{center}
+\end{figure}
+
 ```{r moran_effect_on_aphid_ff, echo=F, results="hide", cache=T, cache.extra=list(seed,temp_aphid,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
-plotter_asym_sens_b_varn<-function(mg,ml,a,b,numkeep_last,ploton){
-
-  pop<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
-  pop_d<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
-
-  neg_b<- -b
-  pop_neg_b<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=neg_b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
-  pop_d_neg_b<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=neg_b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
+plotter_asym_sens_b_varn<-function(mg,ml,a,b,numkeep_last,ploton,resloc){
+
+  pop_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
+  pop_d_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
 
   if(ploton==T){
-    op<-par(mfrow=c(2,2),mar=c(4,3,1.5,1),mgp=c(2,1,0))
-
+    pdf(paste0(resloc,"a_",a,"_b_",b,"_r_",r,"_asym_sens.pdf",sep=""),height=5,width=8)
+    op<-par(mfrow=c(2,3),mar=c(4,4.2,1.5,1),mgp=c(2.5,0.6,0))
     #plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
-    plot(pop[,1],pop[,2],col="red",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    title(expression(paste("(A) g = ",g[1]," , ")), adj=0.4)
-    mtext(paste("b = ",b),side = 3, line=0.2, adj=0.72, col="black")
-    plot(pop_d[,1],pop_d[,2],col="blue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    title(expression(paste("(B) g = ",g[2]," , ")), adj=0.4)
-    mtext(paste("b = ",b),side = 3, line=0.2, adj=0.72, col="black")
-    plot(pop_neg_b[,1],pop_neg_b[,2],col="orange",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    title(expression(paste("(C) g = ",g[1]," , ")), adj=0.4)
-    mtext(paste("b = ",neg_b),side = 3, line=0.2, adj=0.75, col="black")
-    plot(pop_d_neg_b[,1],pop_d_neg_b[,2],col="skyblue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])))
-    title(expression(paste("(D) g = ",g[2]," , ")), adj=0.4)
-    mtext(paste("b = ",neg_b),side = 3, line=0.2, adj=0.75, col="black")
-
-    par(op)
+    pop<-pop_g$pop
+    geps<-pop_g$g_noise
+    plot(mg[,1],geps[,1],col="red",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=1")
+    plot(mg[,2],geps[,2],col="red",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=2")
+    plot(pop[,1],pop[,2],col="red",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2)
+    title(expression(paste("g = ",g[1]," , ")), adj=0.06,cex.main=1.5)
+    mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9, col="black")
+
+    pop_d<-pop_d_g$pop
+    geps<-pop_d_g$g_noise
+    plot(mg[,1],geps[,1],col="blue",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=1")
+    plot(mg[,2],geps[,2],col="blue",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=2")
+    plot(pop_d[,1],pop_d[,2],col="blue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2)
+    title(expression(paste("g = ",g[2]," , ")), adj=0.06,cex.main=1.5)
+    mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9, col="black")
+
+    par(op)
+    dev.off()
   }
@@
   return(list(pop=pop,
-              pop_d=pop_d,
-              pop_neg_b=pop_neg_b,
-              pop_d_neg_b=pop_d_neg_b))
+              pop_d=pop_d))
 
@@
 
-pdf("./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf")
-ans_b_0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.1,numkeep_last=numkeep_last,ploton=T)
-dev.off()
-
-pdf("./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf")
-ans_b_0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.9,numkeep_last=numkeep_last,ploton=T)
-dev.off()
-
-vlist_b_0.1_b_0.9<-c(ans_b_0.1,ans_b_0.9) # a list of 8 matrices
+resloc<-"./Results/asym_sens_results/"
+ans_b_0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
+ans_b_neg0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
+ans_b_0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.9,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
+ans_b_neg0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.9,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
+
+vlist_b_0.1_b_0.9<-c(ans_b_0.1,ans_b_neg0.1,ans_b_0.9,ans_b_neg0.9) # a list of 8 matrices
 

### 2018-10-19 asym_sensitivity.R

@@ -57,7 +57,9 @@ get_noise_glob_loc_mat<-function(N,r){
 #       decrease : logical (controls the nature of global noise)
 #       numkeep_last : number of points you want to keep for the pop matrix from the end
 # Output :
+#     A list of two :
 #     pop matrix with dim. = numkeep_last by 2
+#     g_noise : global noise
 
 Simulator_asym_sens<-function(mg,ml,a,b,p0=c(0,0),decrease,numkeep_last){
   
@@ -77,7 +79,8 @@ Simulator_asym_sens<-function(mg,ml,a,b,p0=c(0,0),decrease,numkeep_last){
   
   pop<-pobs(pop)
   
-  return(pop)
+  return(list(pop=pop,
+              g_noise=g_noise))
   
 } 
 #--------------------------------------------------------------------------------------

### 2018-10-19 Paper.Rmd
include scatter plot in cause4copula function

@@
 \includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf}
-\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{500} points for both noise and population data for comparison).
-p-values $> 0.05$ (above horizontal dashed line) indicates good fit between noise and population copulae.
+\caption{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{500} points for both noise and population data for comparison).
+p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise cpoula family.
 \label{fig_Cause4copula_params_GOF_500}}
@@
 
-* Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence difference (see \nameref{NPA}) for both bivariate noise copula and resultant population copula for $50$ simulations.
+* Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence difference (see \nameref{NPA}) for both bivariate noise copula and resultant population copula (following model \ref{eq.AR1}) for $50$ simulations.
 Then we plotted mean of each measured quantity with standard error against Kendall's Tau of the noise copula family.
 Fig.
-\ref{fig_Cause4copula_C_kend} compares the statistics from population copula which was generated by noises drawn from lower tail dependent Clayton copula.
+\ref{fig_Cause4copula_C_kend} compares the statistics from population copula and the copula made of noises drawn from lower tail dependent Clayton family.
 It is possible to measure the statistics with variation of Spearman's Rho for any other noise copula family (e.g.
@@
 <!--Asymmetric sensitivity and copula structure in populations-->
-\textbf{(ii)} On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold.
+\textbf{(ii)} On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold behaves differently.
 
@@
 
-We simulated our model (\ref{eq.AS}) for $b=\pm0.1$ , $a=0.2$, $r=0.8$.
+We simulated our model (\ref{eq.AS}) for $b=0.1$ , $a=0.2$, $r=0.8$.
 $10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ time step of simulated points to plot the population-copula (Fig.
 \ref{fig_ms_asym_b_0.1}) for a specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
+When asymmetric global noise
+had only non-positive contributions at two habitat patches ($g_{1}$ in the first panel of Fig.
+\ref{fig_ms_asym_b_0.1}), lower tail dependence appeared in the global noise copula and so thus in affected population copula (last subplot in the first panel of Fig.
+\ref{fig_ms_asym_b_0.1}).
+Similarly, when global noises having only non-negative contributions at two habitat patches ($g_{2}$ in the second panel of Fig.
+\ref{fig_ms_asym_b_0.1}), upper tail dependence appeared in the global noise copula and so thus in the outcome population copula (last subplot in the second panel of Fig.
+\ref{fig_ms_asym_b_0.1}).
 We also carried out model selection and non-parametric statistical analysis with each copula of Fig.
-\ref{fig_ms_asym_b_0.1} and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_4}.
-Table \ref{SM-tab_summary_stat_asym_sens} represent the model selection results with AIC weights for best fitted copula and normal copula, model averaged lower and upper tail dependence, p-values (both Cramer-von Mises statistic and Kolmogorov-Smirnov statistic) for goodness of fit test.
-In each case, successful bootstrap was $100\%$.
-According to our model selection approach, non-normal copulae (with higher AIC weights) were always best-fit with some asymetric tail dependence.
-Our non-parametric analysis results (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_4}) are consistent with this model selection approach.
+\ref{fig_ms_asym_b_0.1} and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_2}.
+According to our model selection approach (Table \ref{SM-tab_summary_stat_asym_sens}), non-normal copulae (with higher AIC weights) were always best-fit with similar asymetric tail dependence which gave the support of our hypothesis of the effect of asymmetric global noise copula on populations.
+Our non-parametric analysis results (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_2}) were consistent with the model selection approach.
 
@@
 \caption{Effects of asymmetric global noise copula on two habitat patch model (Eq.
-\ref{eq.AS}) with growth rate $b=\pm0.1$: copula of two populations ($P_{1}, P_{2}$) shaped according to the asymmetry of noise copula\label{fig_ms_asym_b_0.1}}
+\ref{eq.AS}) with growth rate $b=0.1$: copula of two populations ($P_{1}, P_{2}$) shaped according to the asymmetry of noise copula\label{fig_ms_asym_b_0.1}}
 \end{center}
@@
 We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data.
-Spatial mean of the time series data extracted that tail dependence from the multivariate data set, if any, and reflected in the averaged skewness.
 Because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, changed significantly from the individual skewness of time-series data.
@@
 \ref{SM-fig_spTL}(I)).
+
+In the following section we will show application of these ideas with real datasets.
+<!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion-----------
 Finally, we studied **(iii)** the effects of environmental noise, having a non-Normal copula, on extinction risk (calculated after 25 years) with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}).
@@
 For Ricker model, we got the similar observation for smaller values of $r$, the intrinsic growth rate, but for higher values of $r$, there was a turn-over in extinction risks, and right tail dependent noise copula caused population go extinct faster.
+-->
 
@@
 
-we used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data, to compute skewness of spatially averaged data.
-Figure \ref{fig_skewness_spearman} represents the results for each dataset with Spearman correlation preserving surrogates.
+We used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data, to compute skewness of spatially averaged data.
 We tested null hypothesis $(H_{0})$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
 If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness than Normal copula).
-
+We carried out the same analysis with aphid abundance and aphid firstflight date data for all 20 aphid species and plankton abundance data for all 22 plankton species.
+In each study, we got $>5\%$ positive results which gave us confidence in validating significantly different skewness for asymmetric copula than Normal.
+We found the same result when we repeated our simulation with Kendall correlation preserving surrogates (see Fig.
+\ref{SM-fig_skewness_kendall}).
+For example, Figure \ref{fig_skewness_spearman} represents the results for each dataset with Spearman correlation preserving surrogates.
 Figure \ref{fig_skewness_spearman}(A) showed that averaged skewness (indicated by red dot) was much lower than expected if the aphid count data came from a normal copula.
@@
 As $p<0.05$, we rejected $H_{0}$ and that result too was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence (see last column of Table \ref{tab_multivar_summary}).
-
-We carried out the same analysis with aphid abundance and aphid firstflight date data for all 20 aphid species and plankton abundance data for all 22 plankton species.
-In each study, we got $>5\%$ positive results which gave us confidence in validating significantly different skewness for asymmetric copula than Normal.
-
-We found the same result when we repeated our simulation with Kendall correlation preserving surrogates (see Fig. \ref{SM-fig_skewness_kendall}).
 

### 2018-10-19 SupportingInformation.Rmd
include scatter plot in cause4copula function

@@
 \includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf}
-\caption[Testing GOF between noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison).
-p-values $> 0.05$ (above horizontal dashed line) indicates good fit between noise and population copulae.
+\caption[Comparing noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison).
+p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise cpoula family.
 \label{fig_Cause4copula_params_GOF_2500}}
@@
 
-<!--
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf}
-\caption[asym sensitivity results : copula plot]{asym sensitivity results : copula plot \label{fig_ms_asym_b_0.9}}
-\end{center}
-\end{figure}-->
-
 ```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
-summary_stat<-summary_stat[,c(1:4)]
-rownames(summary_stat)<-c("bestcop.AICw","N.AICw","relLTdep.AICw","relUTdep.AICw","p.CvM","p.KS","BS.success.percent")
+summary_stat<-summary_stat[,c(1:2)]
+rownames(summary_stat)<-c("bestfit.copula.AICw","Normal.copula.AICw","LT.model.avg","UT.model.avg","p.CvM","p.KS","BS.success.percent")
 knitr::kable(summary_stat, booktabs = T, 
@@
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption="summary of model selection results for asymmetric sensitivity : \\textbf{bestcop.AICw} and \\textbf{N.AICw} represent the AIC-weightage for best fit copula and Normal copula, \\textbf{relLTdep.AICw} and \\textbf{relLTdep.AICw} represent relative lower tail and upper tail dependence based on AIC-weightage respectively, \\textbf{p.CvM} and \\textbf{p.KS} are the p-values coming from goodness of fit test (Cramer-von Mises and Kolmogorov-Smirnov statistics based), \\textbf{BS.success.percent} percentage of successful bootstraps in goodness of fit test.
+      caption="summary of model selection results for asymmetric sensitivity : \\textbf{bestfit.copula.AICw} and \\textbf{Normal.copula.AICw} represent the AIC-weightage for best fit copula and Normal copula, \\textbf{LT.model.avg} and \\textbf{UT.model.avg} represent AIC-weight based model averaged lower and upper tail dependence respectively, \\textbf{p.CvM} and \\textbf{p.KS} are the p-values coming from goodness of fit test (Cramer-von Mises and Kolmogorov-Smirnov statistics based), \\textbf{BS.success.percent} percentage of successful bootstraps in goodness of fit test.
 Each column summarizes those statistics for each copula plotted in Fig.
@@
 
-\newpage
 ```{r tab_stat_npa_asym_1,echo=F,results="markup"}
@@
 
-```{r tab_stat_npa_asym_3,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[3]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[3]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_3}"))
-```
-
-```{r tab_stat_npa_asym_4,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[4]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[4]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_4}"))
-```
-
-<!--
-```{r tab_stat_npa_asym_5,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[5]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[5]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_5}"))
-```
-
-```{r tab_stat_npa_asym_6,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[6]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[6]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_6}"))
-```
-
-```{r tab_stat_npa_asym_7,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[7]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[7]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_7}"))
-```
-
-```{r tab_stat_npa_asym_8,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[8]])
-knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[8]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_8}"))
-```
--->
 <!--
@@
 
-<!--results='hide'-->
 ```{r skewness_aphidcount_allsp_plot, echo=F, results="hide", warning=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
@@
 
-<!--
-## Extinction risk
-        ----moved to Paper.Rmd----
--->
-
 ```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
@@
 
+<!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion-->
+<!--
 ![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion (see Eq.
@@
 \ref{eq.extrisk_model}).\label{fig_risk_K_effect}](./Results/ext_risk_K_effect.pdf){ width=40%, height=40% }
+-->
 
@@
 
+<!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion------
+
 # Consequence of non-Normal copula structure : effect on extinction risk \label{Consequence_ext_risk}
@@
 In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
-<!--did I write correct spelling for lewton model?-->
 
@@
 \ref{fig_risk_numlocs_effect_global_disp}] for both model.
+-->
 

### 2018-10-19 Cause4copula.R

@@ -595,6 +595,21 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
   par(op)
   dev.off()
   
+  pdf(paste0(resloc,BiCopName(fcode,short=F),"_scatter_CorlmCoru_",xlabel,".pdf",sep=""),height=4,width=5)
+  op<-par(mar=c(4.5,6,2,1), mgp=c(3,0.5,0))
+  plot(CorlmCoru_noise_mat[,2],CorlmCoru_pop_mat[,2],cex=2,col="black",
+       xlab=expression("Cor"["l,noise"]-"Cor"["u,noise"]),
+       ylab=expression("Cor"["l,pop"]-"Cor"["u,pop"]),
+       cex.lab=2,cex.axis=1.5,
+       xlim=range(CorlmCoru_noise_mat[,2]),ylim=range(CorlmCoru_pop_mat[,2]))
+  dat<-data.frame(x1=CorlmCoru_noise_mat[,2],y1=CorlmCoru_pop_mat[,2])
+  mylm<-lm(y1~x1,data=dat)
+  abline(mylm,col="black")
+  c<-cor.test(dat$x1,dat$y1,method = "pearson",alternative = "t")
+  mtext(paste0("Pearson correlation = ",round(unname(c$estimate),3),", p = ",round(c$p.value,4),sep=""),cex=1.5,line=0.1)
+  par(op)
+  dev.off()
+  
   pdf(paste0(resloc,BiCopName(fcode,short=F),"_Pl_vs_",xlabel,".pdf",sep=""),height=4,width=5)
   op<-par(mar=c(3.5,4.5,2,3.5), mgp=c(1.9,0.5,0))
   plot(corcoef_list,Pl_noise_mat[,2],cex=0.5,col="red",xlab=xlabel,ylab=expression("P"["l"]),
@@ -670,6 +685,21 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
   par(op)
   dev.off()
   
+  pdf(paste0(resloc,BiCopName(fcode,short=F),"_scatter_PlmPu_",xlabel,".pdf",sep=""),height=4,width=5)
+  op<-par(mar=c(4.5,6,2,1), mgp=c(3,0.5,0))
+  plot(PlmPu_noise_mat[,2],PlmPu_pop_mat[,2],cex=2,col="black",
[…]

### 2018-10-20 Paper.Rmd
editting...continued

@@
 
-* Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence difference (see \nameref{NPA}) for both bivariate noise copula and resultant population copula (following model \ref{eq.AR1}) for $50$ simulations.
+* Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula (following model \ref{eq.AR1}) for $50$ simulations.
 Then we plotted mean of each measured quantity with standard error against Kendall's Tau of the noise copula family.
 Fig.
-\ref{fig_Cause4copula_C_kend} compares the statistics from population copula and the copula made of noises drawn from lower tail dependent Clayton family.
+\ref{SM-fig_Cause4copula_C_kend} compares the statistics from population copula and the copula made of noises drawn from lower tail dependent Clayton family.
 It is possible to measure the statistics with variation of Spearman's Rho for any other noise copula family (e.g.
@@
 (For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
+Linear regression between tail statistics from noise and population copula (Fig.
+\ref{fig_Cause4copula_scatter_C_kend}) confirmed that the asymmetric tail dependence in noise copula caused similar tail dependence in population copula.
 
@@
 \begin{center}
-\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/common_legend_cause4copula_stat.pdf}\\
-\vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Spearman_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Kendall_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
-\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
-p-values $< 0.05$ (below horizontal purple line) indicates significantly similar estimate between noise and population copulae.
-\label{fig_Cause4copula_C_kend}}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption{Linear regression between 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9 as shown in Fig.
+\ref{SM-fig_Cause4copula_C_kend}) showed significant pearson correlation, and henceforth confirmed asymmetric tail dependence in noise copula caused similar tail dependence in population copula.\label{fig_Cause4copula_scatter_C_kend}}
 \end{center}

### 2018-10-20 SupportingInformation.Rmd
editting...continued

@@
 \vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Spearman_vs_Spearman's Rho.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Kendall_vs_Spearman's Rho.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pearson_vs_Spearman's Rho.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Corl-Coru_vs_Spearman's Rho.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pl-Pu_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
+\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
+\label{fig_Cause4copula_C_kend}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Spearman_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Kendall_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pearson_vs_Spearman's Rho.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Corl_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Coru_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Corl-Coru_vs_Spearman's Rho.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pl_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pu_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pl-Pu_vs_Spearman's Rho.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_D2u_vs_Spearman's Rho.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_D2l_vs_Spearman's Rho.pdf}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_D2u-D2l_vs_Spearman's Rho.pdf}
-\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from upper-tail dependent \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
-p-values $< 0.05$ (below horizontal purple line) indicates significantly similar estimate between noise and population copulae.
+\caption[Comparing correlations and non-parametric statistic between upper tail dependent noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from upper tail dependent \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
 \label{fig_Cause4copula_SC_spear}}

### 2018-10-20 Cause4copula.R

@@ -597,11 +597,22 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
   
   pdf(paste0(resloc,BiCopName(fcode,short=F),"_scatter_CorlmCoru_",xlabel,".pdf",sep=""),height=4,width=5)
   op<-par(mar=c(4.5,6,2,1), mgp=c(3,0.5,0))
+  ylim1<-range(CorlmCoru_pop_mat[,1],CorlmCoru_pop_mat[,3])
+  xlim1<-range(CorlmCoru_noise_mat[,1],CorlmCoru_noise_mat[,3])
+  b_len<-diff(xlim1)/50
   plot(CorlmCoru_noise_mat[,2],CorlmCoru_pop_mat[,2],cex=2,col="black",
        xlab=expression("Cor"["l,noise"]-"Cor"["u,noise"]),
        ylab=expression("Cor"["l,pop"]-"Cor"["u,pop"]),
        cex.lab=2,cex.axis=1.5,
-       xlim=range(CorlmCoru_noise_mat[,2]),ylim=range(CorlmCoru_pop_mat[,2]))
+       xlim=xlim1,ylim=ylim1)
+  segments(CorlmCoru_noise_mat[,2],CorlmCoru_pop_mat[,1],CorlmCoru_noise_mat[,2],CorlmCoru_pop_mat[,3],col='blue')
+  segments(CorlmCoru_noise_mat[,2]-b_len,CorlmCoru_pop_mat[,1],CorlmCoru_noise_mat[,2]+b_len,CorlmCoru_pop_mat[,1],col='blue')
+  segments(CorlmCoru_noise_mat[,2]-b_len,CorlmCoru_pop_mat[,3],CorlmCoru_noise_mat[,2]+b_len,CorlmCoru_pop_mat[,3],col='blue')
+  
+  segments(CorlmCoru_noise_mat[,1],CorlmCoru_pop_mat[,2],CorlmCoru_noise_mat[,3],CorlmCoru_pop_mat[,2],col='red')
+  segments(CorlmCoru_noise_mat[,1],CorlmCoru_pop_mat[,2]-b_len,CorlmCoru_noise_mat[,1],CorlmCoru_pop_mat[,2]+b_len,col='red')
[…]

### 2018-10-21 Paper.Rmd
adding abstract

@@
 # Abstract
+
+We introduce a new rank based "copula" approach to model dependencies between ecological variables.
+Our method could overcome the shortcomings of linear correlation approach and give us new insight about "asymmetric tail dependence" found in ecology, biogeochemistry and related field.
+We focus on three objectives (1) Do datasets in ecology and related fields have non-Normal copula srtructure?
+Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric
+tail dependence? (2) If so, what are some possible causes of non-Normal copula structure? And (3) what are the
+consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
+We present several examples with real data which show asymmetric tail dependence in bivariate as well as multivariate non-Normal copulae.
+Asymmetric tail dependence can explain some important phenomena like "spatial synchrony", causal relationship between ecological variables (like species abundance and rainfall).
+We show "Moran-effect" and asymmetric sensitivity towards global environmental fluctuations could be the possible causes for non-Normal copula structures.
+We demonstrate with simulations and real data that non-Normal copulae have profound effects on spatial skewness of species abundances and slope from spatial Taylor's law.
+Our method is new and could have potential in exploring some other related phenomena like effect of climate change on extinction risks of metapopulations.
 
@@
 \includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf}
-\caption{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{500} points for both noise and population data for comparison).
+\caption{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison).
 p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise cpoula family.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption{Linear regression between 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9 as shown in Fig.
+\caption{Linear regression between 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from lower tail dependent Clayton copula and of the affected populations for 9 equidistant Kendall's Tau values in a range (0.1 to 0.9 as shown in Fig.
 \ref{SM-fig_Cause4copula_C_kend}) showed significant pearson correlation, and henceforth confirmed asymmetric tail dependence in noise copula caused similar tail dependence in population copula.\label{fig_Cause4copula_scatter_C_kend}}
@@
 <!--Asymmetric sensitivity and copula structure in populations-->
-\textbf{(ii)} On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold behaves differently.
+\textbf{(ii)} On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when beyond a certain expectation threshold it behaved differently.
 
@@
 In each study, we got $>5\%$ positive results which gave us confidence in validating significantly different skewness for asymmetric copula than Normal.
-We found the same result when we repeated our simulation with Kendall correlation preserving surrogates (see Fig.
-\ref{SM-fig_skewness_kendall}).
 For example, Figure \ref{fig_skewness_spearman} represents the results for each dataset with Spearman correlation preserving surrogates.
@@
 As $p<0.05$, we rejected $H_{0}$ and that result too was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence (see last column of Table \ref{tab_multivar_summary}).
+We found the same result when we repeated our simulation with Kendall correlation preserving surrogates (see Fig.
+\ref{SM-fig_skewness_kendall}).
 

### 2018-10-21 SupportingInformation.Rmd
adding abstract

@@
 \caption[Spatial Taylor's Law results]{Spatial Taylor's Law results: log(spatial variance) vs.
-log(spatial mean) for $ns=10$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of homoskedasticity, (G) root mean square error, (H) intercept, (I) slope, (J) quadratic coefficient, (K) mean curvature of regression line from first panel-plots.
+log(spatial mean) for $ns=1000$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of homoskedasticity, (G) root mean square error, (H) intercept, (I) slope, (J) quadratic coefficient, (K) mean curvature of regression line from first panel-plots.
 \label{fig_spTL}}

### 2018-10-21 Paper.Rmd
modify table...continued

@@
 
+<!--
 ```{r tab_bivar_summary, echo=F, results='markup'}
-
 tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
@@
              booktabs=TRUE)
+```
+-->
+
+<!--bivariate table summary in latex format-->
+```{r tab_bivar_summaryl, echo=F, results='asis',message=F}
+tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
+row.names(tab_bivar_summary)<-NULL
+library(kableExtra)
+library(dplyr)
+library(tibble)
+dt <- tibble(Description = c("1. p-value from independence test",
+           "2. best fit copula based on minimum AIC",
+           "3. AIC-weight of that best fit copula",
+           "4. AIC-weight of Normal copula",
+           "5. p-value from goodness of fit test for Cramer-von Mises method",
+           "6. p-value from goodness of fit test for Kolmogorov-Smirnov method",
+           "7. model averaged (AIC-weights based) lower tail dependence",
+           "8. model averaged (AIC-weights based) upper tail dependence",
+           "9. model averaged (AIC-weights based) (lower - upper) tail dependence",
+           "10. Rank of statistics (Pl-Pu) for real data in comparison to 1000 Normal surrogates")
+)
+tab_bivar_summary<-cbind(dt,tab_bivar_summary)
+knitr::kable(tab_bivar_summary, 
+             format="latex", 
+             caption = "Summary of results for bivariate data-sets\\label{tab_bivar_summary}",
+             booktabs=T, linesep = "\\addlinespace",
+             col.names = c(" ", "SoilCN", "Birds’ BMR", "Mammals’ BMR", "Ceder-creek")
+             )%>%column_spec(2:5,width="1.4cm")%>%
+             column_spec(1, width = "11cm")%>%kable_styling(latex_options = "basic",position = "center")
+```
+
+<!--multivariate table summary in latex format-->
+
+```{r tab_multivar_summaryl, echo=F, results='asis',message=F}
+tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
+tab_multivar_summary<-tab_multivar_summary[c(1:11,16,17),]
+row.names(tab_multivar_summary)<-NULL
+library(kableExtra)
+library(dplyr)
+library(tibble)
+dt <- tibble(
+#Items = c("num.pairs","num.non-indep", "num.non-N", "Avg.best.AICw", "Avg.N.AICw", "num.good.fit", "0.025-CI.LT",
+#          "0.975-CI.LT","0.025-CI.UT","0.975-CI.UT","num.LT-UT>0","0.025-CI.Pl-Pu","0.975-CI.Pl-Pu"),
+Description = c("1. total number of location pairs (excluding self-comparison)",
+           "2. how many location pair have non-independent copula structures",
+           "3. number of non-Normal copula as best fit",
+           "4. average AIC-weights of best fit copula from all location pair",
+           "5. average AIC-weights of Normal copula from all location pair",
+           "6. number of location pair having best fit copula with p-values > 0.01
+            (both Cramer-von Mises and Kolmogorov-Smirnov method)",
+           "7. 2.5 percentile of model-avg. lower tail-dep. from all location pair",
+           "8. 97.5 percentile of model-avg. lower tail-dep. from all location pair",
+           "9. 2.5 percentile of model-avg. upper tail-dep. from all location pair",
+           "10. 97.5 percentile of model-avg. upper tail-dep. from all location pair",
+           "11. number of location pair showing stronger lower tail dependence",
+           "12. 2.5 percentile of Pl-Pu statistics from all location pair",
+           "13. 97.5 percentile of Pl-Pu statistics from all location pair")
+)
+
+tab_multivar_summary<-cbind(dt,tab_multivar_summary)
+knitr::kable(tab_multivar_summary, 
+             format="latex", 
+             caption = "Summary of results for multivariate data-sets\\label{tab_multivar_summary}",
+             booktabs=T, linesep = "\\addlinespace",
+             col.names = c(" ","Green-spruce aphid count",
+                        "Leaf-plum curling aphid firstflight",
+                        "Ceratium furca count","Methane-flux")
+             )%>%column_spec(2:5,width="1.4cm")%>%
+             column_spec(1, width = "11cm")%>%kable_styling(latex_options = "basic",position = "center")
 ```
@@
 
+<!--what we did?-->
+In the earlier work, most of the studies in ecology focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of asymmetric tail-dependence and non-Normal copulae in this field.
+We
+
+<!--why is it important to study?-->
+
+<!--future work-->
+
 \newpage

### 2018-10-22 Paper.Rmd
fixing tables

@@
 Lower panel represents the similar scenario but with stronger correlation (Kendall's $\tau = 0.8$).
-So, the flexibility of copulas allows us to model more complex multivariate dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence.<!--and reflection symmetry versus reflection asymmetry.--> To identify the kinds of copula structure found in real ecological data, we need to select candidate bivariate copulas from many existing parametric copula families.
-<!--A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulas to be selected from.--> Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), mentioned in the \nameref{Methods}.
+So, the flexibility of copulae allows us to model more complex multivariate dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence.<!--and reflection symmetry versus reflection asymmetry.--> To identify the kinds of copula structure found in real ecological data, we need to select candidate bivariate copulae from many existing parametric copula families.
+<!--A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulae to be selected from.--> Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), mentioned in the \nameref{Methods}.
 
@@
 The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`.
-Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulas.
+Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulae.
 AIC and BIC values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection]. 
-Descriptions of these copula families can be found in introductory texts on copulas [@nelsen2007_copula ; @joe2014_dependence]  and in `VineCopula` documentation. 
+Descriptions of these copula families can be found in introductory texts on copulae [@nelsen2007_copula ; @joe2014_dependence]  and in `VineCopula` documentation. 
 The lower- and upper-tail dependence of the 
@@
 
-We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas (for details see Appendix \ref{SM-Test_npa_stat}).
+We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulae (for details see Appendix \ref{SM-Test_npa_stat}).
 
@@
 
-For multivariate datasets, a spatial resampling scheme was used to calculate confidence intervals, instead of the surrogate 
+For multivariate datasets, we used a spatial resampling scheme (for details see Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals, instead of the surrogate 
 approach described above.
-The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
-widely used spline correlogram method; the same scheme was also used by @walteretal2017 
-as part of a method for 
-performing model selection among matrix regression models. For each pair of 
-locations $i \neq j$, the statistic
-of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
-$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
-$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
-values $s_{ij}$ were then arranged in
-a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
-$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
-average strength of 
-dependence, across all pairs of distinct locations, for data in the portion of the distributions 
-given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
-asymmetry of tail dependence across all pairs of locations. 
-Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
-distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
-revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
-or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
-Normal null hypothesis, and answering question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
-Bootstrapping was performed by resampling the locations of measurement, with replacement. 
-This corresponds to 
-resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
-[@bjornstadfalck2001 ; @walteretal2017].
 
@@
 Non-Normal copula were always emarged as best fit copula based on minimum AIC value (2$^{nd}$ row of Table \ref{tab_bivar_summary}).
-Normal copula were found as poor fit compared to highest AIC-weight based asymmetric copulas (3$^{rd}$-4$^{th}$ rows of Table \ref{tab_bivar_summary}) - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}.
-Goodness of fit of the best-fit asymmetric copulas were tested based on p-values from two statistics (Cramer-von Mises and Kolmogorov-Smirnov : 5$^{th}$-6$^{th}$ rows of Table \ref{tab_bivar_summary}).
-Among 4 datasets presented there, first three have significant upper tail dependence whereas the last one has significant lower tail dependence compared to Normal (7$^{th}$-9$^{th}$ rows of Table \ref{tab_bivar_summary}).
-It was indicated by negative AIC-weights model averaged (lower-upper) tail dependence and very low ranks of actual copula 
-data for $\Ps_l - \Ps_u$ statistics for first 3 datasets and positive AIC-weights model averaged 
-(lower-upper) tail dependence and very high ranks of actual copula data for $\Ps_l - \Ps_u$ statistics for the last one, which answers the second and third part of \hyperref[Intro]{\textbf{(Q1)}} (For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and Figs.
+Normal copula were found as poor fit compared to highest AIC-weight based asymmetric copulae (3$^{rd}$ and 4$^{th}$ rows of Table \ref{tab_bivar_summary}) - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}.
+Goodness of fit of the best-fit asymmetric copulae were tested based on p-values from two statistics (Cramer-von Mises and Kolmogorov-Smirnov : 5$^{th}$ and 6$^{th}$ rows of Table \ref{tab_bivar_summary}, respectively).
+Among 4 datasets presented there, first three have significantly stronger upper tail dependence whereas the last one has significantly stronger lower tail dependence compared to Normal (compare 7$^{th}$ and 8$^{th}$ rows of Table \ref{tab_bivar_summary}).
+It was indicated by negative AIC-weights model averaged (lower-upper) tail dependence (9$^{th}$ row of Table \ref{tab_bivar_summary}) for first 3 datasets and positive AIC-weights model averaged
+(lower-upper) tail dependence for the last one, which answers the second and third part of \hyperref[Intro]{\textbf{(Q1)}} (For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and Figs.
 \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}).
-Our model-selection results were consistent with non-parametric analysis for all the datasets (compare last two rows of Table \ref{tab_bivar_summary}).
+We also reported the rank of ($\Ps_l - \Ps_u$) for the actual copula data in camparison to the same non-parametric statistics of Kendall-correlation preserving 1000 Normal surrogates.
+Ranks if were $>$ 975 or $<$ 25, then indicated a significantly (95$\%$ CI) different lower tail-dependence or upper tail-dependence measures, respectively, based on Normal-null hypothesis.
+We found our model-selection results were consistent with non-parametric analysis for all the datasets (compare last two rows of Table \ref{tab_bivar_summary}).
 We will now discuss briefly findings from one of our bivariate datasets as follows.
@@
 From the model selection approach, the test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`).
-We then fitted several copulas (all $16$ families as mentioned in \nameref{Model_selection})
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (Table \ref{SM-tab_soilCNfit}). 
+We then fitted several copulae (all $16$ families as mentioned in \nameref{Model_selection})
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulae (Table \ref{SM-tab_soilCNfit}). 
 Based on AIC weights, `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the best fit.
-The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of \hyperref[Intro]{\textbf{(Q1)}} for these data.
+The table shows that several copulae are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of \hyperref[Intro]{\textbf{(Q1)}} for these data.
 
@@
 We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets 
-(i.e., abundance data for green spruce aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
-firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, 
-abundance data for plankton from UK seas from $14$ sampling locations, 
+(i.e., abundance data for green spruce aphid and firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, abundance data for plankton from UK seas from $14$ sampling locations,
 methane flux data from $13$ sampling locations).
 
-Likewise, green-spruce aphid abundance data ($1^{st}$ column in Table \ref{tab_multivar_summary}), 
-plankton data from UK seas also showed major lower tail dependent spatial synchrony 
-($3^{rd}$ column in Table \ref{tab_multivar_summary}), whereas firstflight data for leaf-curling plum aphid 
-showed stronger upper tail dependence at every location pair ($2^{nd}$ column in Table \ref{tab_multivar_summary}).
+We found major occurence of non-independent copulae (row 2, Table \ref{tab_multivar_summary}) out of all pairwise locations (row 1, Table \ref{tab_multivar_summary}) considered.
+Non-Normal copulae, most of the time, emerged as best-fit based on minimum AIC out of 16 copula family considered (row 3, Table \ref{tab_multivar_summary}) answering the first part of question \hyperref[Intro]{\textbf{(Q1)}}.
+All four data sets showed Normal copulae were always poor fit in comparison to best-fit non-Normal copulae (compare their average AIC-weights from row 4 and row 5 of Table \ref{tab_multivar_summary}).
+We considered the "good-fit" location pair based on the p-values (>0.01) obtained from both Cramer-von Mises and Kolmogorov-Smirnov statistics (row 6, Table \ref{tab_multivar_summary}).
+Some spatial data (green spruce aphid's count, Ceratium furca's count and methane flux) showed stronger lower tail dependence (compare row 7 and row 8 from Table \ref{tab_multivar_summary}, also see row 9) whereas Leaf-plum curling's firstflight data showed stronger upper tail dependence which answered the second and third part of question \hyperref[Intro]{\textbf{(Q1)}}.
+Those spatial synchrony from model selection results were also verified by the non-parametric statistic which showed exclusively positive (or negative) confidence interval of the distribution $(\Ps_l - \Ps_u)$ estimated from all location pair for lower (or upper)- tail dependent spatial data (except methane-flux data which had marginal lower quantile).
 Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.
@@
 
-As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid's abundance data, model selection among copulas was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations.
+As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid's abundance data, model selection among copulae was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations.
 Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question \hyperref[Intro]{\textbf{(Q1)}} for these data.
 Comparing the
-AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the Normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulas was usually not
+AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the Normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulae was usually not
 comparable to the degree of support for other families.
@@
 failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
-the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulas of all pairs of locations could be characterized by at least one of them.
+the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulae of all pairs of locations could be characterized by at least one of them.
 In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100`$\%$ of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
@@
 Fig.
-\ref{fig_biv_multi_raw_cop}(A) and (B) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.
+\ref{fig_biv_multi_raw_cop}(I) and (J) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.
 
@@
 
-As described in \nameref{NPA}, mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
+As mentioned in \nameref{NPA} and Appendix \ref{SM-spatial_resamp}, mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
 $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
@@
 exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}.
-
-Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig.
-\ref{SM-fig_aphid_count} A,B) showed the typical decline of correlation with distance between sampling locations that has been seen for numerous other datasets in earlier work.
-Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig.
-\ref{SM-fig_aphid_count} C-F).
-Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig.
-\ref{SM-fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical result of Table \ref{SM-tab_count_resamp} that averages of these quantities across all pairs of sampling locations were significantly greater than zero, demonstrating asymmetry of tail dependence, as also seen from model selection approach.
-Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
+Visual illustration of this results were discussed in Appendix \ref{SM-visual_npa}.
 
@@
 dt <- tibble(Description = c("1. p-value from independence test",
-           "2. best fit copula based on minimum AIC",
+           "2. Best fit copula based on minimum AIC",
            "3. AIC-weight of that best fit copula",
@@
            "6. p-value from goodness of fit test for Kolmogorov-Smirnov method",
-           "7. model averaged (AIC-weights based) lower tail dependence",
-           "8. model averaged (AIC-weights based) upper tail dependence",
-           "9. model averaged (AIC-weights based) (lower - upper) tail dependence",
+           "7. Model averaged (AIC-weights based) lower tail dependence",
+           "8. Model averaged (AIC-weights based) upper tail dependence",
+           "9. Model averaged (AIC-weights based) (lower - upper) tail dependence",
            "10. Rank of statistics (Pl-Pu) for real data in comparison to 1000 Normal surrogates")
@@
 tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
-tab_multivar_summary<-tab_multivar_summary[c(1:11,16,17),]
+tab1<-tab_multivar_summary[c(1:7,9,11,12,14,16),] # see carefully if you rearrange tab_multivar_summary's row
+                                                   #in suppmat this will change
+tab2<-tab_multivar_summary[c(8,10,13,15,17),]  # see carefully if you rearrange tab_multivar_summary's row
+                                                   #in suppmat this will change
+
+two_tab_in1 <- as.data.frame(do.call(cbind, lapply(1:ncol(tab1), function(i) paste0(" (",tab1[c(7:8,10:12),i], " , ",tab2[ , i], ")") )))   # see carefully if you rearrange tab_multivar_summary's row in suppmat this will change
+
+colnames(two_tab_in1)<-colnames(tab_multivar_summary)
+rownames(two_tab_in1)<-c("CI_LT","CI_UT","CI_Pl","CI_Pu","CI_Pl-Pu")
+
+tab_multivar_summary<-rbind(tab_multivar_summary[c(1:6),],
+                            two_tab_in1[1:2,], #CI of LT, CI of UT
+                            tab_multivar_summary[11,],
+                            two_tab_in1[5,] #CI of Pl-Pu
+                           )
+
 row.names(tab_multivar_summary)<-NULL
@@
 dt <- tibble(
-#Items = c("num.pairs","num.non-indep", "num.non-N", "Avg.best.AICw", "Avg.N.AICw", "num.good.fit", "0.025-CI.LT",
-#          "0.975-CI.LT","0.025-CI.UT","0.975-CI.UT","num.LT-UT>0","0.025-CI.Pl-Pu","0.975-CI.Pl-Pu"),
-Description = c("1. total number of location pairs (excluding self-comparison)",
-           "2. how many location pair have non-independent copula structures",
-           "3. number of non-Normal copula as best fit",
-           "4. average AIC-weights of best fit copula from all location pair",
-           "5. average AIC-weights of Normal copula from all location pair",
-           "6. number of location pair having best fit copula with p-values > 0.01
-            (both Cramer-von Mises and Kolmogorov-Smirnov method)",
-           "7. 2.5 percentile of model-avg. lower tail-dep. from all location pair",
-           "8. 97.5 percentile of model-avg. lower tail-dep. from all location pair",
-           "9. 2.5 percentile of model-avg. upper tail-dep. from all location pair",
-           "10. 97.5 percentile of model-avg. upper tail-dep. from all location pair",
-           "11. number of location pair showing stronger lower tail dependence",
-           "12. 2.5 percentile of Pl-Pu statistics from all location pair",
-           "13. 97.5 percentile of Pl-Pu statistics from all location pair")
+            Description = c("1. Total number of location pairs (excluding self-comparison)",
+           "2. Number of location pair having non-independent copula structures",
+           "3. Number of non-Normal copula as best fit",
+           "4. Average AIC-weights of best fit copula from all location pair",
+           "5. Average AIC-weights of Normal copula from all location pair",
+           "6. Number of good-fit location pair*",
+           "7. 95% CI of model-avg. lower tail-dep. from all location pair",
+           "8. 95% CI of model-avg. upper tail-dep. from all location pair",
+           "9. Number of location pair showing stronger lower tail dependence",
+           "10. 95% CI of (Pl - Pu) from all location pair")
            )
@@
 tab_multivar_summary<-cbind(dt,tab_multivar_summary)
+
 knitr::kable(tab_multivar_summary, 
@@
              booktabs=T, linesep = "\\addlinespace",
-             col.names = c(" ","Green-spruce aphid count",
-                        "Leaf-plum curling aphid firstflight",
-                        "Ceratium furca count","Methane-flux")
-             )%>%column_spec(2:5,width="1.4cm")%>%
-             column_spec(1, width = "11cm")%>%kable_styling(latex_options = "basic",position = "center")
+             col.names = c(" ","Green-spruce aphid's count",
+                        "Leaf-plum curling aphid's firstflight",
+                        "Ceratium furca's count","Methane-flux")
+             )%>%column_spec(2:5,width="2.5cm")%>%
+             column_spec(1, width = "7cm")%>%kable_styling(latex_options = "basic",position = "center")%>%
+   add_footnote("Good fit was considered for best fit copula with p-values > 0.01 (both Cramer-von Mises and Kolmogorov-Smirnov method)", notation="symbol", threeparttable = T, escape = F)
 ```
@@
 In all cases `r min((RES_temp_aphid$gfc_numBS_success/RES_temp_aphid$gfc_numBS),na.rm=T)*100` $\%$ of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-Table \ref{SM-tab_temp_aphid_LTmUT} showed `r sum(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw <0, na.rm=T)` out of `r (dim(RES_temp_aphid$gfc_p_CvM)[1]^2-dim(RES_temp_aphid$gfc_p_CvM)[1])-2*RES_temp_aphid$num_indep_loc_pair` location pair showed upper tail dependence in temperature copula, which can be considered as the Moran effect results causing dominant upper tail dependence in the copulas with aphid's firstflight date (Table \ref{SM-tab_ff_LTmUT}).
+Table \ref{SM-tab_temp_aphid_LTmUT} showed `r sum(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw <0, na.rm=T)` out of `r (dim(RES_temp_aphid$gfc_p_CvM)[1]^2-dim(RES_temp_aphid$gfc_p_CvM)[1])-2*RES_temp_aphid$num_indep_loc_pair` location pair showed upper tail dependence in temperature copula, which can be considered as the Moran effect results causing dominant upper tail dependence in the copulae with aphid's firstflight date (Table \ref{SM-tab_ff_LTmUT}).
 -->

### 2018-10-22 SupportingInformation.Rmd
fixing tables

@@
 date: "February 12, 2018"
-geometry: "left=1cm,right=1cm,top=2.5cm,bottom=2.8cm"
+geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"
 
@@
 pdf("./Results/SoilCNraw.pdf")
-op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1.5,cex.axis=1.5)
+op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
+plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=2.5,cex.axis=1.5)
 par(op)
@@
 pdf("./Results/BMR_results/birds/raw_birds_BMR.pdf")
-op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(b[,1],b[,2],type="p",col="red",xlab="log(Mass)",ylab="log(BMR)",cex.lab=1.5,cex.axis=1.5)
+op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
+plot(b[,1],b[,2],type="p",col="red",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=1.5)
 par(op)
@@
 pdf("./Results/BMR_results/mammals/raw_mammals_BMR.pdf")
-op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(m[,1],m[,2],type="p",col="red",xlab="log(Mass)",ylab="log(BMR)",cex.lab=1.5,cex.axis=1.5)
+op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
+plot(m[,1],m[,2],type="p",col="red",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=1.5)
 par(op)
@@
     if(rawplot==T){
-      op<-par(mar=c(5.1, 4.1, 1.5, 2.1))
-      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col="red",pch=1,cex.lab=1.6,cex.axis=1.5)
-      mtext(paste0("Year=",year),cex=1.5)
+      op<-par(mar=c(5.1, 5.1, 1.5, 2.1))
+      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col="red",pch=1,cex.lab=2.5,cex.axis=1.5)
+      mtext(paste0("Year=",year),cex=2)
       par(op)
     }else{
-      op<-par(mar=c(5.1, 4.1, 1.5, 2.1))
-      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",col="red",pch=1,cex.lab=2,cex.axis=1.5,
+      op<-par(mar=c(5.1, 5.1, 1.5, 2.1))
+      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",col="red",pch=1,cex.lab=2.5,cex.axis=1.5,
          xlim=c(0,1),ylim=c(0,1),asp=1)
@@
       lines(c(0,1),c(0,1),type='l')
-      mtext(paste0("Year=",year),cex=1.5)
+      mtext(paste0("Year=",year),cex=2)
       par(op)
@@
 
+# Spatial resampling with multivariate datasets for non-parametric statistics \label{spatial_resamp}
+
+The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
+widely used spline correlogram method; the same scheme was also used by @walteretal2017 
+as part of a method for 
+performing model selection among matrix regression models. For each pair of 
+locations $i \neq j$, the statistic
+of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
+$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
+$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
+values $s_{ij}$ were then arranged in
+a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
+$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
+average strength of 
+dependence, across all pairs of distinct locations, for data in the portion of the distributions 
+given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
+asymmetry of tail dependence across all pairs of locations. 
+Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
+distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
+revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
+or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
+Normal null hypothesis, and answering question \hyperref[MT-Intro]{\textbf{(Q1)}} from the \nameref{MT-Intro}.
+Bootstrapping was performed by resampling the locations of measurement, with replacement. 
+This corresponds to 
+resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
+[@bjornstadfalck2001 ; @walteretal2017].
+
+# How to visualize non-parametric analysis results for multivariate spatial data? \label{visual_npa}
+
+The principle remains same but for example we chose one of spatial data set (green spruce aphid's count) to illustrate results from Table \ref{tab_count_resamp}.
+Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig.
+\ref{fig_aphid_count} A,B) showed the typical decline of correlation with distance between sampling locations that has been seen for numerous other datasets in earlier work.
+Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig.
+\ref{fig_aphid_count} C-F).
+Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively).
+Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig.
+\ref{fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical result of Table \ref{tab_count_resamp} that averages of these quantities across all pairs of sampling locations were significantly greater than zero, demonstrating asymmetry of tail dependence, as also seen from model selection approach.
+Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
+
 <!--This chunk is for model selection results of soilCN dataset-->
@@
 
-\newpage
 <!--Display model selection results for BMR:birds data in a table-->
@@
 
+\newpage
 ```{r tab_count_bestcop, echo=F, results="markup"}
@@
 
-\newpage
 ```{r tab_count_AICw_normal, echo=F, results='markup'}
@@
 
+\newpage
 ```{r tab_count_LTdep_AICw, echo=F, results='markup'}
@@
 
+\newpage
 ```{r tab_count_UTdep_AICw, echo=F, results='markup'}
@@
 
+\newpage
 ```{r tab_count_LTmUT, echo=F, results='markup'}
@@
 
+\newpage
 ```{r tab_ff_AICw_normal, echo=F, results='markup'}
@@
 
+\newpage
 <!--display the main statistical results in a table-->
@@
   m$`0.975-CI.Pl-Pu`<-list(round(rs$numericdf[10,4],3))
-  #m$`CI0.975_Pl-Pu`<-round(rs$numericdf[10,4],3)
   #m$CI_Pl<-list(c(round(rs$numericdf[5,3],3),round(rs$numericdf[5,4],3)))
@@
 pdf("./Results/Aphid_count_raw_loc_4_11.pdf",height = 8,width=8)
-op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Green-spruce aphid count from location 4",ylab="Green-spruce aphid count from location 11",cex.lab=1.5,cex.axis=1.5)
+op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
+plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Green-spruce aphid count (loc. 4)",ylab="Green-spruce aphid count (loc. 11)",cex.lab=2.5,cex.axis=1.5)
 par(op)
@@
 pdf("./Results/Aphid_count_cop_loc_4_11.pdf",height = 8,width=8)
-op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
+op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
 plot(m[,1],m[,2],type="p",col="red",xlab="u",ylab="v",
-     cex.lab=2,cex.axis=1.5,
+     cex.lab=2.5,cex.axis=1.5,
      xlim=c(0,1),ylim=c(0,1),asp=1)
@@
 pdf("./Results/Aphid_ff_raw_loc_2_5.pdf",height = 8,width=8)
-op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
-plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Leaf-plum curling aphid's first-flight from location 2",ylab="Leaf-plum curling aphid's first-flight from location 5",cex.lab=1.5,cex.axis=1.5)
+op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
+plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Leaf-plum curling's first-flight (loc. 2)",ylab="Leaf-plum curling's first-flight (loc. 5)",cex.lab=2.5,cex.axis=1.5)
 par(op)
@@
 pdf("./Results/Aphid_ff_cop_loc_2_5.pdf",height = 8,width=8)
-op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
+op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
 plot(m[,1],m[,2],type="p",col="red",xlab="u",ylab="v",
-     cex.lab=2,cex.axis=1.5,
+     cex.lab=2.5,cex.axis=1.5,
      xlim=c(0,1),ylim=c(0,1),asp=1)
@@
 
+\newpage
 # Consequence of non-Normal copula structure : Effect on Taylor's law \label{Consequence_TL_simulation}

### 2018-10-22 Paper.Rmd
adding discussion(primarily) and fixing figures

@@
 
-<!--what we did?-->
-In the earlier work, most of the studies in ecology focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of asymmetric tail-dependence and non-Normal copulae in this field.
-We
+<!--what we did and why is it important to study?-->
 
-<!--why is it important to study?-->
+In the earlier work, most of the studies in ecology focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values of them.
+In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field.
+We showed by examples of several real data set (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence is common in nature and could tell us about the nature of interaction between different ecological variables.
+Our method in this regard of two ways.
+First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family.
+These copulae were chosen carefully so that they covered the whole range from exclusively lower (Clayton, Survival Gumbel, Survival Joe, Survival BB6, Survival BB8) or upper tail dependence (Survival Clayton, Gumbel, Joe, BB6, BB8) to both tail dependence (BB1, Survival BB1, BB7, Survival BB7) in addition to two symmetric copulae (Normal and Frank).
+Copula having highest AIC-weight came as best-fit from this approach.
+Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), still we wanted to make sure the reliability of our model selection results.
+So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different non-parametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we innovated and we found they were consistent with the inferences we drew from model selction results.
+Our approach of "copula" could help to interprete the synchronous behavior of a species from several metapopulations.
+To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g.
+abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that metapopulations are simultaneously rare at those locations.
+So, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of planktons could lead to algal-bloom in the ocean.
+Copula could also gave us information about the causal relationship between ecological variables.
+For example, if maximum railfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be strongerly correlated then lower values.
+
+We explained the possible causes which induced asymmetric tail-dependence in copula.
+Borrowing the idea from "Moran's theorem", we demonstrated with simulations that two habitat patches when affected by asymmetric bivariate-noise copula showed similar tail-dependence asymmetry in the population copula from those patches (see Fig.
+\ref{fig_Cause4copula_params_GOF_500} and Fig.
+\ref{fig_Cause4copula_scatter_C_kend}).
+We also showed the effect of global environmental fluctuation (behaving asymmetrically beyond a certain threshold) on the meta-population copula which was shaped with similar tail dependence as of the initial global noise copula (Fig.
+\ref{fig_ms_asym_b_0.1}).
+We observed significant changes in skewness of spatially averaged timeseries and also in the curvature of spatial Taylor's law for copula having asymmetric tail-dependence than it would have been expected for Normal copula.
+For example, Fig.
+\ref{fig_skewness_spearman}(A) showed green spruce aphid's abundance timeseries had right skewed distribution but its skewness was signifiantly lower than it was expected if it would came from its Spearman correlation preserving Normal surrogates.
+This result indicates less extreme outbreaks of green spruce aphid around UK than it would have been with Normal copula structure.
+Similarly, methane fluxes showing lower skewness than Normal surrogates mean that the level of green house gas is under-control situation and monitoring this skewness over several decades may be helpful to understand the "green-house effect" on environment.
 
 <!--future work-->
+We think this copula approach have scope to investigate into the extinction risks of metapopulation due to climate change.
+To illustrate, lower-tail dependent environmental fluctuations may suppress the metapopulations synchronously by making some of the habitats go extinct.
+This could help in managing policy for conservation of endangered populations.
 

### 2018-10-22 SupportingInformation.Rmd
adding discussion(primarily) and fixing figures

@@
       op<-par(mar=c(5.1, 5.1, 1.5, 2.1))
-      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col="red",pch=1,cex.lab=2.5,cex.axis=1.5)
+      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col="red",pch=1,cex.lab=2,cex.axis=1.5)
       mtext(paste0("Year=",year),cex=2)
@@
       op<-par(mar=c(5.1, 5.1, 1.5, 2.1))
-      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",col="red",pch=1,cex.lab=2.5,cex.axis=1.5,
+      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",col="red",pch=1,cex.lab=2,cex.axis=1.5,
          xlim=c(0,1),ylim=c(0,1),asp=1)
@@
 op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Green-spruce aphid count (loc. 4)",ylab="Green-spruce aphid count (loc. 11)",cex.lab=2.5,cex.axis=1.5)
+plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Green-spruce aphid count (loc. 4)",ylab="Green-spruce aphid count (loc. 11)",cex.lab=2.5,cex.axis=2)
 par(op)
@@
 plot(m[,1],m[,2],type="p",col="red",xlab="u",ylab="v",
-     cex.lab=2.5,cex.axis=1.5,
+     cex.lab=3,cex.axis=2,
      xlim=c(0,1),ylim=c(0,1),asp=1)
@@
 op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Leaf-plum curling's first-flight (loc. 2)",ylab="Leaf-plum curling's first-flight (loc. 5)",cex.lab=2.5,cex.axis=1.5)
+plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Leaf-plum curling's first-flight (loc. 2)",ylab="Leaf-plum curling's first-flight (loc. 5)",cex.lab=2.5,cex.axis=2)
 par(op)
@@
 rownames(skw_aphid_count_spearman)<-paste("sp",splist,sep="")
-pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_spearman.pdf",height=20,width=20)
-op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_spearman.pdf",height=25,width=25)
+op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
 cat("Now for spearman preserving surrogs : aphid count","\n")
@@
 rownames(skw_aphid_count_kendall)<-paste("sp",splist,sep="")
-pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_kendall.pdf",height=20,width=20)
-op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_kendall.pdf",height=25,width=25)
+op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
 cat("Now for kendall preserving surrogs : aphid count","\n")
@@
 rownames(skw_aphid_ff_spearman)<-paste("sp",splist,sep="")
-pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_spearman.pdf",height=20,width=20)
-op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_spearman.pdf",height=25,width=25)
+op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
 cat("Now for spearman preserving surrogs : aphid ff","\n")
@@
 rownames(skw_aphid_ff_kendall)<-paste("sp",splist,sep="")
-pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_kendall.pdf",height=20,width=20)
-op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_kendall.pdf",height=25,width=25)
+op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
 cat("Now for kendall preserving surrogs : aphid ff","\n")
@@
 skw_plankton_kendall<-skw_plankton_spearman
-pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_spearman.pdf",height=24,width=24)
-op<-par(mfrow=c(4,6),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_spearman.pdf",height=30,width=30)
+op<-par(mfrow=c(4,6),mar=c(5,5,1,1),mgp=c(3,1,0))
 #cat("Now for spearman preserving surrogs : pns","\n")
@@
 
-pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=24,width=24)
-op<-par(mfrow=c(4,6),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=30,width=30)
+op<-par(mfrow=c(4,6),mar=c(5,5,1,1),mgp=c(3,1,0))
 #cat("Now for kendall preserving surrogs : pns","\n")
@@
 pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_kendall.pdf",height=6,width=8)
-op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
+op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
 answer1k<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf",height=6,width=8)
-op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
+op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
 answer1s<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_kendall.pdf",height=6,width=8)
-op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
+op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
 answer2k<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf",height=6,width=8)
-op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
+op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
 answer2s<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_kendall.pdf",height=6,width=8)
-op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
+op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
 answer3k<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf",height=6,width=8)
-op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
+op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
 answer3s<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 pdf("./Results/skewness_results/skewness_methane/skewness_methane_kendall.pdf",height=6,width=8)
-op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
+op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
 answer1k<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 pdf("./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf",height=6,width=8)
-op<-par(mar=c(3.5,3.5,1.8,0.1), mgp=c(2, 1, 0))
+op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
 answer1s<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
@@
 ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=sp,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon = F)
-mtext(paste0("p= ",round(ans$p_left_mncurv,3)),side = 3, line=-2, adj=0.5, col="blue")
+mtext(paste0("p= ",round(ans$p_left_mncurv,3)),side = 3, line=-2, adj=0.5, col="navyblue",cex=2)
 par(op)
@@
      cex.axis=1.5,cex.lab=1.5)
-mtext(paste0("mean curvature= ",round(unname(ans$realTL_stats$s.stats[7]),3)),side = 3, line=-2, adj=0.5, col="blue")
+mtext(paste0("mean curvature= ",round(unname(ans$realTL_stats$s.stats[7]),3)),side = 3, line=-2, adj=0, col="navyblue",cex=2)
 
@@
      cex.axis=1.5,cex.lab=1.5)
-mtext(paste0("mean curvature= ",round(ans$good_surrog_mncurv[id],3)),side = 3, line=-2, adj=0.5, col="blue")
+mtext(paste0("mean curvature= ",round(ans$good_surrog_mncurv[id],3)),side = 3, line=-2, adj=0, col="navyblue",cex=2)
 

### 2018-10-22 Cause4copula.R

@@ -616,6 +616,7 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
   dat<-data.frame(x1=CorlmCoru_noise_mat[,2],y1=CorlmCoru_pop_mat[,2])
   mylm<-lm(y1~x1,data=dat)
   abline(mylm,col="black")
+  lines(x=xlim1,y=ylim1,col="green")
   c<-cor.test(dat$x1,dat$y1,method = "pearson",alternative = "t")
   mtext(paste0("Pearson correlation = ",round(unname(c$estimate),3),", p = ",round(c$p.value,4),sep=""),cex=1.5,line=0.1)
   par(op)
@@ -720,6 +721,7 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
   dat<-data.frame(x1=PlmPu_noise_mat[,2],y1=PlmPu_pop_mat[,2])
   mylm<-lm(y1~x1,data=dat)
   abline(mylm,col="black")
+  lines(x=xlim1,y=ylim1,col="green")
   c<-cor.test(dat$x1,dat$y1,method = "pearson",alternative = "t")
   mtext(paste0("Pearson correlation = ",round(unname(c$estimate),3),", p = ",round(c$p.value,4),sep=""),cex=1.5,line=0.1)
   par(op)
@@ -819,6 +821,7 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
   dat<-data.frame(x1=D2umD2l_noise_mat[,2],y1=D2umD2l_pop_mat[,2])
   mylm<-lm(y1~x1,data=dat)
   abline(mylm,col="black")
+  lines(x=xlim1,y=ylim1,col="green")
   c<-cor.test(dat$x1,dat$y1,method = "pearson",alternative = "t")
   mtext(paste0("Pearson correlation = ",round(unname(c$estimate),3),", p = ",round(c$p.value,4),sep=""),cex=1.5,line=0.1)
   par(op)

### 2018-10-22 skewness_multivar_dataset.R

@@ -56,9 +56,9 @@ skewness_testing<-function(ts_matrix,loclist,numsurrog,ploton,corpres){
   surrogskw<-apply(FUN=myskns,X=ts_surrog_avg,MARGIN=2)
   
   if(ploton==T){
-    hist(surrogskw,breaks=100,main="",xlab="surrogate_skewness",ylab="frequency",cex.lab=2)
+    hist(surrogskw,breaks=100,main="",xlab="surrogate_skewness",ylab="frequency",cex.lab=2,cex.axis=2)
     #abline(v=realskw,col="red")
-    points(x=realskw,y=0,col="red",pch=15)
+    points(x=realskw,y=0,col="red",pch=15,cex=2)
   }
   
   return(list(surrogskw=surrogskw,realskw=unname(realskw)))

### 2018-10-22 Paper.Rmd
editing

@@
 \caption{Linear regression between 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from lower tail dependent Clayton copula and of the affected populations for 9 equidistant Kendall's Tau values in a range (0.1 to 0.9 as shown in Fig.
-\ref{SM-fig_Cause4copula_C_kend}) showed significant pearson correlation, and henceforth confirmed asymmetric tail dependence in noise copula caused similar tail dependence in population copula.\label{fig_Cause4copula_scatter_C_kend}}
+\ref{SM-fig_Cause4copula_C_kend}) showed significant pearson correlation, and henceforth confirmed asymmetric tail dependence in noise copula caused similar tail dependence in population copula.
+Regression line (black) differing from 1:1 line (green) indicates that asymmetric tail dependence in noise and population copulae are similar but not exactly same.\label{fig_Cause4copula_scatter_C_kend}}
 \end{center}

### 2018-10-29 Paper.Rmd
editting...

@@
 title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
-author: "Shyamolina Ghosh, Mark T. Holder, Terrance E. Loecke, Daniel C. Reuman"
+author: "Shyamolina Ghosh, Lawrence W Sheppard, Mark T. Holder, Terrance E. Loecke, Daniel C. Reuman"
 date: "February 18, 2018"
@@
 
+Sheppard: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
+
 Holder: Department of Ecology and Evolutionary Biology and Biodiversity Institute, University of Kansas, Lawrence, KS, 66045, USA
@@
 We introduce a new rank based "copula" approach to model dependencies between ecological variables.
-Our method could overcome the shortcomings of linear correlation approach and give us new insight about "asymmetric tail dependence" found in ecology, biogeochemistry and related field.
+Our method could overcome the shortcomings of the linear correlation approach and give us new insight about "asymmetric tail dependence" found in ecology, biogeochemistry and related field.
 We focus on three objectives (1) Do datasets in ecology and related fields have non-Normal copula srtructure?
@@
 tail dependence? (2) If so, what are some possible causes of non-Normal copula structure? And (3) what are the
-consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
+consequences of non-Normal copula structure and tail dependence for our understanding of biology and what are the applications?
 We present several examples with real data which show asymmetric tail dependence in bivariate as well as multivariate non-Normal copulae.
-Asymmetric tail dependence can explain some important phenomena like "spatial synchrony", causal relationship between ecological variables (like species abundance and rainfall).
-We show "Moran-effect" and asymmetric sensitivity towards global environmental fluctuations could be the possible causes for non-Normal copula structures.
-We demonstrate with simulations and real data that non-Normal copulae have profound effects on spatial skewness of species abundances and slope from spatial Taylor's law.
-Our method is new and could have potential in exploring some other related phenomena like effect of climate change on extinction risks of metapopulations.
+Asymmetric tail dependence can illuminate some important phenomena like "spatial synchrony" and causal relationships between ecological variables (like species abundance and rainfall).
+We show that the "Moran-effect" and asymmetric sensitivity towards global environmental fluctuations could be the possible causes for non-Normal copula structures.
+We demonstrate with simulations and real data that non-Normal copulae have profound effects on spatial skewness of species abundances and spatial Taylor's law slope values.
+Our method is new and could have potential when exploring some other related phenomena like effects of climate change on extinction risks of metapopulations.
 
@@
 
-Identifying and quantifying dependencies is the core of modelling in many applications, frequently found in finance and risk management.
+Identifying and quantifying dependencies is the core of modelling in many applications, frequently found in biology and elsewhere.
 For many years, measuring and modeling dependencies has focused on the
 Pearson correlation because of its ease in use.
-But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence (i.e., varying association between two variables) because it is a measure of linear association between two variables.
-It may be better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation, when the marginal distribution of the variables are highly non-Gaussian.
-The copula [@joe2014_dependence; @nelsen2007_copula], a ranked based dependence measurement between two variables, transforms non-uniformly distributed marginals in original variables to uniformly distributed marginals when joint distribution of ranked data is plotted.
+But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence (i.e., varying strength of association between two variables depending on their values) because it is a measure of linear association between two variables.
+It may be better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation when the marginal distribution of the variables are highly non-Gaussian.
+The copula [@joe2014_dependence; @nelsen2007_copula], a rank-based dependence measurement between two variables, transforms non-uniformly distributed marginals in original variables to uniformly distributed marginals when the joint distribution of ranked data is plotted.
 For example, in the first panel of Fig.
-\ref{fig_cop_pedag}, we showed the joint distribution of raw data (left) and copula data (right), where the marginals of raw data variables (x and y) were gamma distributed but in copula plot those marginals had uniform distribution.
-Copula plot was obtained by scaling individual ranked variables by a factor $(n+1)$ where $n$ was the number of sampling points, so that the copula was confined within a unit box.
-From now, we will call it as "normalized ranking".
-Using a "copula-based approach" as a dependence measure reflects more knowledge about the dependence structure instead of summarizing in just a single number, the linear correlation coefficient.
+\ref{fig_cop_pedag}, we showed the joint distribution of raw data (left) and copula data (right), where the marginals of raw data variables (x and y) were gamma distributed but in the copula plot those marginals had uniform distribution.
+The copula plot was obtained by scaling individual ranked variables by a factor $(n+1)$ where $n$ was the number of sampling points, so that the copula was confined within a unit box.
+From now, we will call this "normalized ranking".
+Using a "copula-based approach" as a dependence measure reflects more knowledge about the dependence structure than summarizing in just a single number, the linear correlation coefficient.
 It tells us about the tail-dependence of the variables, i.e., the association between extreme values of two variables.
@@
 So, the flexibility of copulae allows us to model more complex multivariate dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence.<!--and reflection symmetry versus reflection asymmetry.--> To identify the kinds of copula structure found in real ecological data, we need to select candidate bivariate copulae from many existing parametric copula families.
-<!--A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulae to be selected from.--> Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), mentioned in the \nameref{Methods}.
+<!--A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulae to be selected from.--> Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), described in the \nameref{Methods}.
 
-Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data.
+Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related fields and thus has potential scope to better capture the evolution of ecological time series data.
 
-In ecological perspective, when we find a dominantly lower (upper) tail dependence structure between two species abundance timeseries, it means both incidence are rare (abundant) synchronously.
-Again, if abundance of a species (say, species A) is associated with scarcity of another species (say, species B) then it can explain the causality or negative feedback of A on species B.
-If one variable is an environmental factor (like temperatute or rainfall) and another is an observable (like abundance of a species) then it can also explain the causal relationship between them.
-On a similar note, if a species' abundance time series from different sampling locations shows multivariate dependence structure, then we can analyze the "spatial data" as pairwise bivariate copula and find if there is any tail dependence or not.
-If those copula came up with asymmetric tail dependence, then it can be viewed as a new type to quantify "spatial synchrony" [@Liebhold2004; @walteretal2017], which is a celebrated phenomenon in population ecology.
+From the ecological perspective, when we find a dominantly lower (upper) tail dependence structure between two species abundance timeseries, it means both incidence are rare (abundant) synchronously.
+Again, if abundance of a species (say, species A) is associated with scarcity of another species (say, species B) then this may indicate causality, i.e.
+negative feedback of A on species B.
+If one variable is an environmental factor (like temperatute or rainfall) and another is an observable (like abundance of a species) then copula structure can also help explain the causal relationship between them.
+On a similar note, if a species' abundance time series from different sampling locations shows multivariate dependence structure, then we can analyze the "spatial data" using pairwise bivariate copulae and find if there is any tail dependence or not.
+If the copulae have asymmetric tail dependence, then it can be viewed as a new way to quantify "spatial synchrony" [@Liebhold2004; @walteretal2017], which is a celebrated phenomenon in population ecology.
 In this article, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets.
@@
 And \textbf{(Q3)} what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
-In \nameref{Data}, we described briefly about each bivariate and multivariate datasets we used to answer these questions and provided a detailed descriptions of implemented methods in \nameref{Methods}.
-In the \nameref{Results} section, we summarized our finidings for each datasets.
-Next two sections are focused on possible causes and consequences of non-Normal copula structure in ecology and related field with applications shown in real data as well as some simulations.
-Finally, we concluded in the \nameref{Discussion} section.
+In \nameref{Data}, we described briefly each of bivariate and multivariate datasets we used to answer these questions and we provide a detailed descriptions of the implemented methods in \nameref{Methods}.
+In the \nameref{Results} section, we summarized our findings for each dataset.
+The next two sections are focused on possible causes and consequences of non-Normal copula structure in ecology and related fields with applications shown in real data as well as some simulations.
+Finally, we conclude in the \nameref{Discussion} section.
 
@@
 - Do they show tail dependence distinct from that of a Normal copula?
-- In particular do they show asymmetric tail dependence?-->
+- In particular do they show asymmetric tail dependence?
 
 To answer \hyperref[Intro]{\textbf{(Q1)}}, we chose one representative datasets for each category : with a bivariate (soil C and N data) and a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.
+-->
 
@@
 For all $4$ bivariate-datasets indicated by each column name (SoilCN, Birds' BMR, Mammals' BMR, Ceder-creek) in Table \ref{tab_bivar_summary}, data were always significantly correlated as indicated by p-value (<0.05) from `BiCopIndTest`{`VineCopula`} (1$^{st}$ row of Table \ref{tab_bivar_summary}).
-Non-Normal copula were always emarged as best fit copula based on minimum AIC value (2$^{nd}$ row of Table \ref{tab_bivar_summary}).
-Normal copula were found as poor fit compared to highest AIC-weight based asymmetric copulae (3$^{rd}$ and 4$^{th}$ rows of Table \ref{tab_bivar_summary}) - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}.
+Non-Normal copulae always emerged as best fit copula based on minimum AIC value (2$^{nd}$ row of Table \ref{tab_bivar_summary}).
+The Normal copula was found as poor fit compared to highest AIC-weight based asymmetric copulae (3$^{rd}$ and 4$^{th}$ rows of Table \ref{tab_bivar_summary}) - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}.
 Goodness of fit of the best-fit asymmetric copulae were tested based on p-values from two statistics (Cramer-von Mises and Kolmogorov-Smirnov : 5$^{th}$ and 6$^{th}$ rows of Table \ref{tab_bivar_summary}, respectively).
@@
 
-### Example for bivariate case : soil C and N data \label{Ex_SoilCN}
+### Example for bivariate case : soil C and N data shows upper tail dependence\label{Ex_SoilCN}
 
@@
 
-As described in the \nameref{Data}, raw data plot for soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
+As described in the \nameref{Data}, raw data plot for RaCA database soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
 \ref{fig_biv_multi_raw_cop}(A) and (E), respectively.
@@
 Some spatial data (green spruce aphid's count, Ceratium furca's count and methane flux) showed stronger lower tail dependence (compare row 7 and row 8 from Table \ref{tab_multivar_summary}, also see row 9) whereas Leaf-plum curling's firstflight data showed stronger upper tail dependence which answered the second and third part of question \hyperref[Intro]{\textbf{(Q1)}}.
-Those spatial synchrony from model selection results were also verified by the non-parametric statistic which showed exclusively positive (or negative) confidence interval of the distribution $(\Ps_l - \Ps_u)$ estimated from all location pair for lower (or upper)- tail dependent spatial data (except methane-flux data which had marginal lower quantile).
+The spatial synchrony (from model selection) results were also verified by the non-parametric statistic which showed exclusively positive (or negative) confidence interval of the distribution $(\Ps_l - \Ps_u)$ estimated from all location pair for lower (or upper)- tail dependent spatial data (except methane-flux data which had marginal lower quantile).
 Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.
 
-### Example for multivariate case : abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
+### Example for multivariate case : abundance data for the green spruce aphid (*Elatobium abietinum*) shows spatial synchrony with lower tail dependence\label{Ex_count}
 
@@
 
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}.
+The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`) which indicates aphids are synchronously rare at most of the location pairs.
+The dominance of lower tail dependence in green spruce aphid's abundance data with asymmetric copula structures answers the third part of question \hyperref[Intro]{\textbf{(Q1)}}.
 Fig.
@@
 exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}.
-Visual illustration of this results were discussed in Appendix \ref{SM-visual_npa}.
+Visual illustration of those results is discussed in Appendix \ref{SM-visual_npa}.
 
@@
             Description = c("1. Total number of location pairs (excluding self-comparison)",
-           "2. Number of location pair having non-independent copula structures",
+           "2. Number of location pairs having non-independent copula structures",
            "3. Number of non-Normal copula as best fit",
@@
            "8. 95% CI of model-avg. upper tail-dep. from all location pair",
-           "9. Number of location pair showing stronger lower tail dependence",
+           "9. Number of location pairs showing stronger lower tail dependence",
            "10. 95% CI of (Pl - Pu) from all location pair")
@@
 
-This section is dedicated to explore question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}. 
+This section is dedicated to exploring question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}. 
 <!--We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change?-->
@@
 <!--Moran effects-->
-\textbf{(i)} We were interested to explore the basic idea of Moran's theorem that if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence and then what will be the outcome dependence in the populations at those habitat patches.
+\textbf{(i)} We were interested to extend the basic idea of Moran's theorem : if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence then what will be the outcome dependence in the populations at those habitat patches.
 Should their dependence be exactly the same as of the noises or not ?
@@
 
-To test environmental effect on population copula, our approach is of two way :
+To test environmental effect on population copula, our approach is of two-fold :
 
@@
 \ref{fig_Cause4copula_params_GOF_500}).
-We also tested the goodness of fit of the estimated parameters for population-copula and its standard error due to sample size.
+We also tested the goodness of fit of the estimated parameters for the population-copula and its standard error due to sample size.
 We reported p-values from goodness of fit tests for both statistics (Cramer-von Mises and Kolmogorov-Smirnov).
@@
 \caption{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison).
-p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise cpoula family.
+p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise copula family.
 \label{fig_Cause4copula_params_GOF_500}}
@@
 \ref{SM-fig_Cause4copula_SC_spear}).
-We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach (model selection method), here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which ensured the Moran effect on the population data caused by environmental noise.
+We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach (model selection method), here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which verified the Moran effect on the population data caused by environmental noise.
 (For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
@@
 \ref{SM-fig_Cause4copula_C_kend}) showed significant pearson correlation, and henceforth confirmed asymmetric tail dependence in noise copula caused similar tail dependence in population copula.
-Regression line (black) differing from 1:1 line (green) indicates that asymmetric tail dependence in noise and population copulae are similar but not exactly same.\label{fig_Cause4copula_scatter_C_kend}}
+Regression line (black) differing from 1:1 line (green) indicates that asymmetric tail dependence in noise and population copulae are similar but not exactly the same.\label{fig_Cause4copula_scatter_C_kend}}
 \end{center}
@@
 <!--Asymmetric sensitivity and copula structure in populations-->
-\textbf{(ii)} On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when beyond a certain expectation threshold it behaved differently.
+\textbf{(ii)} Secondly, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) which behaved differently beyond a certain expectation threshold (as shown in Eq.
+\ref{eq.AS_g} and Fig.
+\ref{fig_ms_asym_b_0.1}).
 
@@
 Global environmental fluctuation function $g(\varepsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq.
-\ref{eq.AS_g}) where $\varepsilon=(\varepsilon_{1}, \varepsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\varepsilon}$.
+\ref{eq.AS_g}, Fig.
+\ref{fig_ms_asym_b_0.1}) where $\varepsilon=(\varepsilon_{1}, \varepsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\varepsilon}$.
 
@@
 We simulated our model (\ref{eq.AS}) for $b=0.1$ , $a=0.2$, $r=0.8$.
-$10000$ points were generated from each local and global fluctuations to apply on the initial populations and then after simulation we kept only last $1000$ time step of simulated points to plot the population-copula (Fig.
-\ref{fig_ms_asym_b_0.1}) for a specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
+$10000$ points were generated and we kept only the last $1000$ points to plot the population-copula (Fig.
+\ref{fig_ms_asym_b_0.1}) for each specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
 When asymmetric global noise
@@
 \ref{fig_ms_asym_b_0.1}).
-Similarly, when global noises having only non-negative contributions at two habitat patches ($g_{2}$ in the second panel of Fig.
+Similarly, when global noises had only non-negative contributions at two habitat patches ($g_{2}$ in the second panel of Fig.
 \ref{fig_ms_asym_b_0.1}), upper tail dependence appeared in the global noise copula and so thus in the outcome population copula (last subplot in the second panel of Fig.
@@
 \includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
-\caption{Effects of asymmetric global noise copula on two habitat patch model (Eq.
+\caption{Effects of asymmetric global noise copula (the dependence of $g$ on $\varepsilon$) on two habitat patch model (Eq.
 \ref{eq.AS}) with growth rate $b=0.1$: copula of two populations ($P_{1}, P_{2}$) shaped according to the asymmetry of noise copula\label{fig_ms_asym_b_0.1}}
@@
 <!--Question 3 analyses-->
-# Consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications? \label{Consequences}
+# Consequences of non-Normal copula structure and tail dependence for our understanding of biology \label{Consequences}
 
-This section is attributed to analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}.
+This section is dedicated to analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}.
 We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data.
 Because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, changed significantly from the individual skewness of time-series data.
-To test null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure, we generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same.
-Then we calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R)).
-We made a histogram of surragate skewness and found if actual skewness were within $95\%$ confidence interval or not.
-If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness than Normal copula).
-
+We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
 We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL_simulation}).
@@
 We tested null hypothesis $(H_{0})$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
-If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness than Normal copula).
+We generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same and calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R)).
+We made a histogram of surrogate skewness and found if actual skewness was within the $95\%$ confidence interval or not.
+If the $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness than Normal copula).
 We carried out the same analysis with aphid abundance and aphid firstflight date data for all 20 aphid species and plankton abundance data for all 22 plankton species.
@@
 For example, Figure \ref{fig_skewness_spearman} represents the results for each dataset with Spearman correlation preserving surrogates.
-Figure \ref{fig_skewness_spearman}(A) showed that averaged skewness (indicated by red dot) was much lower than expected if the aphid count data came from a normal copula.
+Figure \ref{fig_skewness_spearman}(A) showed that averaged skewness (indicated by red dot) was much lower than expected if the green-spruce aphid count data came from a normal copula.
 In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data.
 As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results (see first column of Table \ref{tab_multivar_summary}).
-On the other hand, Figure \ref{fig_skewness_spearman}(B) demonstrated that averaged skewness was much higher than expected if the aphid first flight data came from a normal copula.
-As $p$ value (the fraction of surrogate-skewness which were greater than the averaged skewness of the given data - red dot in Figure \ref{fig_skewness_spearman} B) was less than $0.05$, we reject $H_{0}$ in support of upper tail dependence, which was also confirmed by model selection and non-parametric results(see second column of Table \ref{tab_multivar_summary}).
+On the other hand, Figure \ref{fig_skewness_spearman}(B) demonstrated that averaged skewness was significantly higher than expected if the leaf-plum curling aphid first flight data came from a normal copula, which was also confirmed by upper-tail dependence results from model selection and non-parametric results (see second column of Table \ref{tab_multivar_summary}).
 Figure \ref{fig_skewness_spearman}(C) showed that averaged skewness was lower than expected if the plankton data came from a normal copula.
@@
 But, as $p>0.05$, we could not reject $H_{0}$ in that case.
-Finally, for methane-flux data, the averaged skewness was lower than expected if the data came from a normal copula (Figure \ref{fig_skewness_spearman} D).
-As $p<0.05$, we rejected $H_{0}$ and that result too was consistent with the non-parametric analysis and model selection results which also favored lower tail dependence (see last column of Table \ref{tab_multivar_summary}).
-We found the same result when we repeated our simulation with Kendall correlation preserving surrogates (see Fig.
+Finally, for methane-flux data, the averaged skewness was found significantly lower in comparison to its Normal surrogates (Figure \ref{fig_skewness_spearman} D).
+We found the same result with all datasets when we repeated our simulation with Kendall correlation preserving surrogates (see Fig.
 \ref{SM-fig_skewness_kendall}).
@@
 
-Spatial Taylor's law, which shows linear relationship between `log`(variance) and `log`(mean) of population densities of a species from different locations, is a well-known pattern found in ecology and related fields [@Taylor1961, @Taylor1988, @Cohen2013, @Doring2015, @Kuo2016, @Xu2016, @Reuman2017].
-We already showed through simulation (Appendix \ref{SM-Consequence_TL_simulation}) that for asymmetric copula this linear relationship somehow deviates compared to Normal copula which was presumed to have gamma marginals.
-We tested spatial Taylor's law with aphid's count data (for all 20 species) and found that the mean curvature of the regression line from `log`(variance) vs.
+```{r read_tl_aphid_count_sp6, echo=F}
+tlsp6<-readRDS("./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
+```
+
+Spatial Taylor's law, the observed linear relationship between `log`(variance) and `log`(mean) of population densities of a species from different locations, is a well-known pattern found in ecology and related fields [@Taylor1961, @Taylor1988, @Cohen2013, @Doring2015, @Kuo2016, @Xu2016, @Reuman2017].
+We already showed through simulation (Appendix \ref{SM-Consequence_TL_simulation}) that for asymmetric copula this linear relationship somehow deviates compared to Normal copula for a distribution with gamma marginals.
+We tested spatial Taylor's law with aphid count data (for all 20 species) and found that the mean curvature of the regression line from `log`(variance) vs.
 `log`(mean) of population abundance data was significantly different from $10000$ Normal surrogates (Spearman-correlation and marginal distribution preserving) for $25\%$ of total species considered.
-For example, species "Corn leaf aphid" (having significantly upper tail dependence from model selection and non-parametric analysis) showed mean curvature of regression line close to zero, whereas it's one of the $10000$ Normal surrogate showed a higher curvature, indicating here Normal copula didn't hold linear relationship between `log`(spatial variance) and `log`(spatial mean) (see Fig \ref{fig_sTL_mncurv_sp6}).
+For example, the "Corn leaf aphid" (having significantly upper tail dependence from model selection and non-parametric analysis) showed mean curvature of regression line close to zero, whereas `r 10000 -(tlsp6$p_left_mncurv*10000)` of the $10000$ Normal surrogates showed higher curvature, indicating here Normal copula didn't hold linear relationship between `log`(spatial variance) and `log`(spatial mean) (see Fig \ref{fig_sTL_mncurv_sp6}).
 
@@
 \includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
-\caption{log(spatial variance) and log(spatial mean) for $(A)$ corn-leaf aphid's abundance and for $(B)$ one of the $10000$ Normal surrogates (having same marginal distribution and Spearman correlation of actual data).
+\caption{log(spatial variance) and log(spatial mean) for $(A)$ corn-leaf aphid's (\textit{Rhopalosiphum maidis}) abundance and for $(B)$ one of the $10000$ Normal surrogates (having same marginal distribution and Spearman correlation of actual data).
 $(C)$ Histogram of mean curvature of $10000$ normal surrogates showed that for leaf plum corn aphid's abundance data (with asymmetric tail dependence), Normal surrogates didn't follow linear relationship between log(spatial variance) and log(spatial mean).
@@
 
-In the earlier work, most of the studies in ecology focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values of them.
+Most studies in ecology have focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values.
 In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field.
-We showed by examples of several real data set (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence is common in nature and could tell us about the nature of interaction between different ecological variables.
-Our method in this regard of two ways.
+We showed by examples of several real data sets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables.
+We applied two methods.
 First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family.
@@
 Copula having highest AIC-weight came as best-fit from this approach.
-Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), still we wanted to make sure the reliability of our model selection results.
+Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results.
 So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different non-parametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we innovated and we found they were consistent with the inferences we drew from model selction results.
-Our approach of "copula" could help to interprete the synchronous behavior of a species from several metapopulations.
+Our approach could help to interpret the synchronous behavior of a species from several metapopulations.
 To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g.
 abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that metapopulations are simultaneously rare at those locations.
-So, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of planktons could lead to algal-bloom in the ocean.
-Copula could also gave us information about the causal relationship between ecological variables.
-For example, if maximum railfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be strongerly correlated then lower values.
+Conversely, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of plankton could lead to algal-bloom in the ocean.
+Copulae could also give us information about the causal relationship between ecological variables.
+For example, if maximum railfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongerly correlated than lower values.
 
-We explained the possible causes which induced asymmetric tail-dependence in copula.
-Borrowing the idea from "Moran's theorem", we demonstrated with simulations that two habitat patches when affected by asymmetric bivariate-noise copula showed similar tail-dependence asymmetry in the population copula from those patches (see Fig.
+We explained the possible causes which induced asymmetric tail-dependence in the copula.
+Borrowing the idea from "Moran's theorem", we demonstrated with simulations that two habitat patches when affected by bivariate noise with an asymmetric copula structure showed similar tail-dependence asymmetry in the population copula from those patches (see Fig.
 \ref{fig_Cause4copula_params_GOF_500} and Fig.
 \ref{fig_Cause4copula_scatter_C_kend}).
-We also showed the effect of global environmental fluctuation (behaving asymmetrically beyond a certain threshold) on the meta-population copula which was shaped with similar tail dependence as of the initial global noise copula (Fig.
+We also showed the effect of global environmental fluctuation (behaving asymmetrically beyond a certain threshold) on the meta-population copula which was shaped with similar tail dependence as the initial global noise copula (Fig.
 \ref{fig_ms_asym_b_0.1}).
-We observed significant changes in skewness of spatially averaged timeseries and also in the curvature of spatial Taylor's law for copula having asymmetric tail-dependence than it would have been expected for Normal copula.
+We observed significant changes in skewness of spatially averaged timeseries and also in the curvature of spatial Taylor's law for copula having asymmetric tail-dependence than would have been expected for Normal copula.
 For example, Fig.
-\ref{fig_skewness_spearman}(A) showed green spruce aphid's abundance timeseries had right skewed distribution but its skewness was signifiantly lower than it was expected if it would came from its Spearman correlation preserving Normal surrogates.
+\ref{fig_skewness_spearman}(A) showed green spruce aphid abundance timeseries had right skewed distribution but its skewness was signifiantly lower than was expected if it came from Spearman correlation preserving Normal surrogates.
 This result indicates less extreme outbreaks of green spruce aphid around UK than it would have been with Normal copula structure.

### 2018-10-29 SupportingInformation.Rmd
editting...

@@
 ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=sp,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon = F)
-mtext(paste0("p= ",round(ans$p_left_mncurv,3)),side = 3, line=-2, adj=0.5, col="navyblue",cex=2)
+saveRDS(ans,"./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
+mtext(paste0("p= ",ans$p_left_mncurv),side = 3, line=-2, adj=0.5, col="navyblue",cex=2)
 par(op)

### 2018-10-31 Paper.Rmd
editted all

@@
 We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony.
-In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}).
+In our case, for aphid-abundance data and for aphid-firstflight data, only 10 locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}) were considered.
 \textbf{(iii)} Plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013).
@@
 But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available).
-We used data for species=16 [Name : "Ceratium furca"] from those locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}).
+We used data from 14 locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}).
 \textbf{(iv)} Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
@@
 **See and Smyth et al, pending for
-additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.
+additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.
 
@@
 The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`.
-Families considered (with abbreviations and numeric codes here indicated) included the **N**ormal/Gaussian {1}, **C**layton {3}, **G**umbel {4}, **F**rank {5}, **J**oe {6}, **BB1** {7}, **BB6** {8}, **BB7** {9}, **BB8** {10}, **S**urvival **C**layton {13}, **S**urvival **G**umbel {14}, **S**urvival **J**oe {16}, **S**urvival **BB1** {17}, **S**urvival **BB6** {18}, **S**urvival **BB7** {19} and **S**urvival **BB8** {20} copulae.
+Families considered (with abbreviations here indicated) included the **N**ormal/Gaussian, **C**layton, **G**umbel, **F**rank, **J**oe, **BB1**, **BB6**, **BB7**, **BB8**, **S**urvival **C**layton , **S**urvival **G**umbel, **S**urvival **J**oe, **S**urvival **BB1**, **S**urvival **BB6**, **S**urvival **BB7** and **S**urvival **BB8** copulae.
+These copulae were chosen carefully so that they covered the whole range from exclusively lower (C, SG, SJ, SBB6, SBB8) or upper tail dependence (SC, G, J, BB6, BB8) to both tail dependence (BB1, SBB1, BB7, SBB7) in addition to two symmetric copulae (N, F).
 AIC and BIC values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection].
@@
 \end{equation}
-where means and variances are computed using all $n$ data 
+where means and sample variances are computed using all $n$ data 
 points, but the
@@
 Each bin in each subplot of Fig.
-\ref{SM-fig_soilCN_nonparam} contains a number indicating how many surrogates out of 1000, are significantly different ($95\%$ confidence interval) from Normal-copula statistics.
+\ref{SM-fig_soilCN_nonparam} contains a number indicating how many Normal surrogates out of 1000, are different ($95\%$ confidence interval) from Normal-copula statistics.
 Those numbers clearly show that tail dependence of data is stronger in both the lower and upper tails than would be expected under a null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}.
@@
 
-We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets 
-(i.e., abundance data for green spruce aphid and firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, abundance data for plankton from UK seas from $14$ sampling locations,
+We did model selection and non-parametric analysis with all 20 aphid and 22 plankton species and found our two approaches were consistent to each other.
+Here, to present our results we chose one species from each category.
+We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets (i.e., abundance data for green spruce aphid and firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, abundance data for plankton from UK seas from $14$ sampling locations,
 methane flux data from $13$ sampling locations).
@@
            "8. 95% CI of model-avg. upper tail-dep. from all location pair",
-           "9. Number of location pairs showing stronger lower tail dependence",
+           "9. Percentage of location pairs showing stronger lower tail dependence",
            "10. 95% CI of (Pl - Pu) from all location pair")
@@
 \end{center}
-\textbf{ \hspace{2.8 cm} (I) \hspace{3.2 cm} (J) \hspace{3.2 cm} (K) \hspace{3.2 cm} (L) } \\
+\textbf{ \hspace{7 cm} (I) \hspace{3.2 cm} (J)} \\
 \vspace{-0.35 cm}
 \begin{center}
-\includegraphics[width=4 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
-\includegraphics[width=4 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
-\includegraphics[width=4 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}
-\includegraphics[width=4 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
+\includegraphics[width=3.8 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
+\includegraphics[width=3.8 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
+\end{center}
+\textbf{ \hspace{7 cm} (K) \hspace{3.2 cm} (L)} \\
+\vspace{-0.35 cm}
+\begin{center}
+\includegraphics[width=3.8 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
+\includegraphics[width=3.8 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
 \caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs.
@@
 Middle panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel.
-Lower panel shows \textbf{(I)} raw data , \textbf{(J)} copula plot for  green spruce aphid's count data from location pair(4,11) and \textbf{(K)} raw data , \textbf{(L)} copula plot for leaf-curling plum aphid's firstflight data from location pair(2,5).\label{fig_biv_multi_raw_cop}}
+Lower panel shows \textbf{(I - J)} raw data plots and \textbf{(K - L)} copula plots for  green spruce aphid's count data from location pair(4,11) and for leaf-curling plum aphid's firstflight data from location pair(2,5).\label{fig_biv_multi_raw_cop}}
 \end{center}
@@
 
-* First, we calculated the parameters of the noise-copula (for example, we chose Clayton with lower tail dependence) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
+* First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2007_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
 Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
@@
 \caption{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison).
-p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise copula family.
+p-values $> 0.05$ (above horizontal dashed line) indicate good fit (not significantly different) of population copula with same noise copula family.
 \label{fig_Cause4copula_params_GOF_500}}
@@
 \includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
-\caption{Effects of asymmetric global noise copula (the dependence of $g$ on $\varepsilon$) on two habitat patch model (Eq.
+\caption{Effects of asymmetric global noise copula (resulting from the asymmetric dependence of $g$ on $\varepsilon$) on two habitat patch model (Eq.
 \ref{eq.AS}) with growth rate $b=0.1$: copula of two populations ($P_{1}, P_{2}$) shaped according to the asymmetry of noise copula\label{fig_ms_asym_b_0.1}}
@@
 This section is dedicated to analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}.
-We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data.
-Because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, changed significantly from the individual skewness of time-series data.
-We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
+We wanted to examine whether the asymmetric tail dependence had any **(i)** effects on skewness of spatially averaged time series data.
+We predicted that because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, should change significantly from the individual skewness of time-series data.
 We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL_simulation}).
@@
 \includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
-\caption{Skewness of spatially averaged $(A)$ green-spruce aphid's abundance, $(B)$ leaf-curling plum aphid's firstflight date, $(C)$ ceratium furca's abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation).
+\caption{Skewness of spatially averaged $(A)$ green-spruce aphid abundance, $(B)$ leaf-curling plum aphid firstflight date, $(C)$ Ceratium furca abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation).
 \label{fig_skewness_spearman}}
@@
 \includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
-\caption{log(spatial variance) and log(spatial mean) for $(A)$ corn-leaf aphid's (\textit{Rhopalosiphum maidis}) abundance and for $(B)$ one of the $10000$ Normal surrogates (having same marginal distribution and Spearman correlation of actual data).
-$(C)$ Histogram of mean curvature of $10000$ normal surrogates showed that for leaf plum corn aphid's abundance data (with asymmetric tail dependence), Normal surrogates didn't follow linear relationship between log(spatial variance) and log(spatial mean).
+\caption{log(spatial variance) and log(spatial mean) for $(A)$ corn-leaf aphid (\textit{Rhopalosiphum maidis}) abundance and for $(B)$ one of the $9961$ Normal surrogates (having same marginal distribution and Spearman correlation of actual data).
+$(C)$ Histogram of mean curvature of $10000$ normal surrogates showed that for corn-leaf aphid abundance data (with asymmetric tail dependence), Normal surrogates didn't follow linear relationship between log(spatial variance) and log(spatial mean).
 \label{fig_sTL_mncurv_sp6}}
@@
 First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family.
-These copulae were chosen carefully so that they covered the whole range from exclusively lower (Clayton, Survival Gumbel, Survival Joe, Survival BB6, Survival BB8) or upper tail dependence (Survival Clayton, Gumbel, Joe, BB6, BB8) to both tail dependence (BB1, Survival BB1, BB7, Survival BB7) in addition to two symmetric copulae (Normal and Frank).
 Copula having highest AIC-weight came as best-fit from this approach.
 Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results.
-So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different non-parametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we innovated and we found they were consistent with the inferences we drew from model selction results.
+So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different non-parametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selction results.
 Our approach could help to interpret the synchronous behavior of a species from several metapopulations.
@@
 This result indicates less extreme outbreaks of green spruce aphid around UK than it would have been with Normal copula structure.
-Similarly, methane fluxes showing lower skewness than Normal surrogates mean that the level of green house gas is under-control situation and monitoring this skewness over several decades may be helpful to understand the "green-house effect" on environment.
+Similarly, methane fluxes showing lower skewness than Normal surrogates mean that the high emmisions are un-correlated between locations and monitoring this skewness over several decades may be helpful to understand the "green-house effect" on environment.
 

### 2018-10-31 SupportingInformation.Rmd
editted all

@@
                  "0.025-CI.LT","0.975-CI.LT","0.025-CI.UT","0.975-CI.UT",
-                 "num.LT-UT>0",
+                 "percent.LT-UT>0",
                  "0.025-CI.Pl","0.975-CI-Pl","0.025-CI.Pu","0.975-CI.Pu","0.025-CI.Pl-Pu","0.975-CI.Pl-Pu")
@@
   m$`0.975-CI.UT`<-round(CI_UT[2],3)
-  m$`num.LT-UT>0`<-sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)
+  m$`percent.LT-UT>0`<-round((sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)/((d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)))*100,2)
   m$`0.025-CI.Pl`<-round(rs$numericdf[5,3],3)

### 2018-10-31 Cause4copula.R

@@ -235,7 +235,7 @@ Plotter_Cause4copula_GOF<-function(N,fcode,method,num_keep_last,BS){
     #arrows(corcoef_list,par_pop-1.96*se_par_pop,corcoef_list,par_pop+1.96*se_par_pop,length=0.03, angle=90, code=3, col='blue')
     points(corcoef_list,par_noise,pch=2,col="red")
     axis(2, col="black",las=1)  ## las=1 makes horizontal labels
-    mtext("Parameters",side=2,line=2.5)
+    mtext(expression(paste("Parameter (",theta,")")), side = 2, line = 2.5)
     mtext(paste0(BiCopName(family = fcode,short=T)," , min_BS_success = ",BS_success_percentage,"%"),side=3,line=0.3,cex=0.8)
     box()
     

### 2018-10-31 Paper.Rmd
modified

@@
 Similarly, we applied multivariate analysis on four spatial datasets; \textbf{(i)} aphid-count and \textbf{(ii)} aphid-firstflight date were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014].
-These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species.
+These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species (Table \ref{SM-tab_aphid_info}).
 We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony.
 In our case, for aphid-abundance data and for aphid-firstflight data, only 10 locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}) were considered.
-\textbf{(iii)} Plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013).
+\textbf{(iii)} Plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species (Table \ref{SM-tab_plankton_info}) and for 56 years (1958-2013).
 These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).

### 2018-10-31 SupportingInformation.Rmd
modified

@@
 
+```{r tab_aphid_info, echo=F, results='asis',message=F}
+library(tibble)
+aphid_info <- tibble(c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
+           "Corn leaf aphid","Currant lettuce aphid","Damson hop aphid","Grain aphid","Green spruce aphid",
+           "Leaf-curling plum aphid","Mealy cabbage aphid","Mealy plum aphid","Pea aphid","Peach potato aphid",
+           "Potato aphid","Rose grain aphid","Shallot aphid","Sycamore aphid","Willow carrot aphid"),
+           c("Rhopalosiphum insertum", "Rhopalosiphum padi", "Aphis fabae", "Sitobion fragariae", "Hyperomyzus lactucae",
+             "Rhopalosiphum maidis", "Nasonovia ribisnigri","Phorodon humuli","Sitobion avenae","Elatobium abietinum",
+             "Brachycaudus helichrysi","Brevicoryne brassicae","Hyalopterus pruni","Acyrthosiphon pisum","Myzus persicae",
+             "Macrosiphum euphorbiae","Metopolophium dirhodum","Myzus ascalonicus","Drepanosiphum platanoidis","Cavariella aegopodii")
+)
+knitr::kable(aphid_info, "latex", booktabs = T, linesep = "",
+      caption = "Species names of 20 aphids commonly found across UK \\label{tab_aphid_info}",
+      col.names = c("Common name", "Latin binomial"))
+```
+
 \begin{figure}[!h]
@@
 \end{figure}
+
+```{r tab_plankton_info, echo=F, results='asis',message=F}
+library(tibble)
+plankton_info <- tibble(c('Calanus I-IV','Para-Pseudocalanus spp','Temora longicornis','Acartia spp',
+                          'Centropages typicus',
+                          'Oithona spp','Echinoderm larvae','Calanus finmarchicus',
+                          'Calanus helgolandicus','Metridia lucens','Decapoda larvae',
+                   'Euphausiacea','Thalassiosira spp','Rhizosolenia styliformis',
+                   'Ceratium fusus','Ceratium furca','Ceratium tripos','Ceratium macroceros',
+                   'Rhizosolenia alata alata','Pseudocalanus elongatus Adult',
+                   'Nitzschia delicatissima','Nitzschia seriata')
+)
+knitr::kable(plankton_info, "latex", booktabs = T, linesep = "",
+      caption = "Species names of 22 planktons found across UK seas\\label{tab_plankton_info}",
+      col.names = c("Plankton sp. name"))
+```
 

### 2018-11-05 Paper.Rmd
changed...to be continued

@@
 urlcolor : blue
+indent : True
 
@@
 
-## Statistical methods for detecting copula structure \label{Methods}
+## Concepts for \hyperref[Intro]{\textbf{(Q1)}}: Statistical tools for detecting copula structure \label{Methods}
 
@@
 
-A model-selection approach based on likelihood and the Akaike Information Criterion (AIC) 
-and Bayesian Information Criterion (BIC) was used to
-help answer \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
-
+A model-selection approach based on likelihood and the Akaike Information Criterion (AIC) <!--and Bayesian Information Criterion (BIC)--> was used to help answer \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
 For bivariate datasets
 $(x_i,y_i)$ for $i=1,...,n$ such as the soil C and N data, model selection involved several steps, 
-most implemented using tools from the `VineCopula` package. Code combining these tools is in the
+most implemented using tools from the `VineCopula` package. <!--Code combining these tools is in the
 [`OurBiCopSelect`](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
-
-First, we produced ranks 
+to be added in suppmat-->First, we produced ranks
 $(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
 $\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
-we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; further work was with the "normalized ranked variables" $u_i$ 
-and $v_i$. Ranks $\tilde{u}_i$ and $\tilde{v}_i$ were $1$, for the 
-smallest element, up to $n$, for the largest.
-
-Second, we did a test of the independence of the $u_i$ and $v_i$. Subsequent model 
+we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; further work was with these "normalized ranked variables" $u_i$ 
+and $v_i$. Ranks $\tilde{u}_i$ and $\tilde{v}_i$ were $1$, for the smallest element, up to $n$, for the largest. Second, we did a test of the independence of the $u_i$ and $v_i$. Subsequent model 
 selection algorithms are ineffective if data could not be distinguished from independent data 
 since most or all bivariate copula families considered include the independent copula in the 
-family or as a boundary case. The independence test was as implemented in the `BiCopIndTest`
+family or as a boundary case. <!--The independence test was as implemented in the `BiCopIndTest`
 function in `VineCopula`, the documentation of which reports the test uses the asymptotic normality of the statistic $|\tau|\sqrt{\frac{9n(n-1)}{2(2n+5)}}$, where $\tau$ is Kendall's tau statistic.
-If independence could be rejected ($0.05$ significance level), model selection proceeded.
-
+to be added in suppmat-->If independence could be rejected ($0.05$ significance level), model selection proceeded.
 The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`.
@@
 Descriptions of these copula families can be found in introductory texts on copulae [@nelsen2007_copula ; @joe2014_dependence]  and in `VineCopula` documentation.
-The lower- and upper-tail dependence of the 
-best-fitting member of each family was also provided by `BiCopEst`.
+The lower- and upper-tail dependence of the best-fitting member of each family was also provided by `BiCopEst`.
+Finally, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
+model-averaged lower- and upper-tail dependence values using standard model averaging 
+formulas [@burnham2003_modelselection]; likewise for BIC.
 
@@
 
-Finally, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
-model-averaged lower- and upper-tail dependence values using standard model averaging 
-formulas [@burnham2003_modelselection]; likewise for BIC.
-
 The first part of \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} was answered for a bivariate dataset 
 $(x_i,y_i)$ for $i=1,...,n$ by comparing the AIC weights of the non-Normal copula
-families with that of the Normal family. The same was also done using BIC.
+families with that of the Normal family. <!--The same was also done using BIC.....no, we didn't report it-->
 The second part of \hyperref[Intro]{\textbf{(Q1)}} was answered by comparing 
@@
 We also counted for how many location pairs model -averaged lower tail dependence is greater than model-averaged upper-tail dependence or vice-versa.
-
-<!--For example, for green-spruce aphid abundance data we found major lower tail dependence whereas leaf-curling plum aphid's first flight data had significant upper tail dependence at all location pairs (for details see Table \ref{tab_multivar_summary}).
-***DAN: need to fill this in, prrobably enough to be brief, since these were just done 
-pairwise
-**SG: done-->
+Then, we reported as the percentage of location pairs showing stronger lower tail dependence out of total non-independent location pairs considered.
 
@@
 
-In addition to the model selection approach described above, **nonparametric approaches** to quantify lower ($\cor_l$, $\Ps_l$ and $\Dsq_l$ statistics) and upper tail dependence ($\cor_u$, $\Ps_u$ and $\Dsq_u$ statistics) were also used to answer \hyperref[Intro]{\textbf{(Q1)}}.
+In addition to the model selection approach described above, **nonparametric approaches** to quantify lower <!--($\cor_l$, $\Ps_l$ and $\Dsq_l$ statistics)--> and upper tail dependence <!--($\cor_u$, $\Ps_u$ and $\Dsq_u$ statistics)--> were also used to answer \hyperref[Intro]{\textbf{(Q1)}}.
 Using two distinct methods, when results are consistent, provides greater confidence
@@
 Our nonparametric study of tail dependence uses three
-statistics which can quantify the extent to which $u_i$ and $v_i$ are 
+statistics (i.e., $\cor$, $\Ps$ and $\Dsq$) which can quantify the extent to which $u_i$ and $v_i$ are 
 related in any part of their distributions. We here describe the 
@@
 
-\textbf{(i)} The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
+The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
 correlation of $u_i$ and $v_i$ that is attributable to the points 
@@
 
-\textbf{(ii)} We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
+We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
 $u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
@@
 
-\textbf{(iii)} The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
+The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
 distance between points satisfying $u_i+v_i>2l_b$ and 
@@
 $\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.
-
 We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulae (for details see Appendix \ref{SM-Test_npa_stat}).
@@
 distributions of values of the same statistics computed on surrogate datasets that were produced from 
-the empirical data by randomizing it in a special way to have no tail dependence. The surrogate/randomized datasets had exactly 
+the empirical data by randomizing it in a special way to have no tail dependence (for details see Appendix \ref{SM-surrog_test}). The surrogate/randomized datasets had exactly 
 the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
@@
 statistically unchanged.
-Details of how surrogates were produced are in Appendix \ref{SM-surrog_test}, and code is in [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R).
 
@@
 The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
-as computed on the 
-empirical data, to the distribution
-of its values on many surrogate datasets 
+as computed on the empirical data, to the distribution of its values on many surrogate datasets 
 provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
 third part of question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} for bivariate datasets.
-
 For multivariate datasets, we used a spatial resampling scheme (for details see Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals, instead of the surrogate
@@
 
-<!--Theoretical models
-Question 2 from the Introduction was explored using simple population models.-->
-
-<!--# Application : copula structures in ecology, evolutionary and biogeochemical spheres \label{Results}-->
-# Application : biological dependencies \label{Results}
+# Results for \hyperref[Intro]{\textbf{(Q1)}}: Ecological datasets often have non-Normal copula structures, including asymmetric tail dependence \label{Results}
 <!--## question 1
@@
 We found our model-selection results were consistent with non-parametric analysis for all the datasets (compare last two rows of Table \ref{tab_bivar_summary}).
-We will now discuss briefly findings from one of our bivariate datasets as follows.
 
@@
 
+We will now discuss briefly findings from one of our bivariate datasets as follows.
 As described in the \nameref{Data}, raw data plot for RaCA database soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
@@
 To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$.
-
 To test environmental effect on population copula, our approach is of two-fold :
 
-* First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2007_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
+First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2007_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
 Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
@@
 
-* Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula (following model \ref{eq.AR1}) for $50$ simulations.
+Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula (following model \ref{eq.AR1}) for $50$ simulations.
 Then we plotted mean of each measured quantity with standard error against Kendall's Tau of the noise copula family.

### 2018-11-05 SupportingInformation.Rmd
changed...to be continued

@@
 op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=2.5,cex.axis=1.5)
+plot(d[,1],d[,2],type="p",col="black",xlab="Soil C",ylab="Soil N",cex.lab=2.5,cex.axis=1.5)
 par(op)
@@
 op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(b[,1],b[,2],type="p",col="red",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=1.5)
+plot(b[,1],b[,2],type="p",col="black",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=1.5)
 par(op)
@@
 op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(m[,1],m[,2],type="p",col="red",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=1.5)
+plot(m[,1],m[,2],type="p",col="black",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=1.5)
 par(op)
@@
       op<-par(mar=c(5.1, 5.1, 1.5, 2.1))
-      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",col="red",pch=1,cex.lab=2,cex.axis=1.5)
+      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",pch=1,cex.lab=2,cex.axis=1.5)
       mtext(paste0("Year=",year),cex=2)
@@
       op<-par(mar=c(5.1, 5.1, 1.5, 2.1))
-      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",col="red",pch=1,cex.lab=2,cex.axis=1.5,
+      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",pch=1,cex.lab=2,cex.axis=1.5,
          xlim=c(0,1),ylim=c(0,1),asp=1)
@@
 <!--# Description of Data \label{Data_detailed}
-
-For bivariate analysis we used four datasets which are briefly described as follows;
-First, to explore the dependence of **soil organic C** and total **soil N** across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
-database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
-stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
-study.
-
-<!--
-Next, the [**BMR data**](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) contains the trait contrasts from phylogenetically independent contrasts [@SukumaranH2010] of data reported by [@GenoudIM2017] on the basal metabolic rate (BMR) and mass of mammals species.
-Specifically, the "SELECT" BMR data in their database were used for those species for which data were available.
-A phylogeny in NEXUS [@NEXUS] format was extracted from the supplementary information of [@GenoudIM2017].
-Version 4.4.0 of the DendroPy library [@SukumaranH2010] was used to calculate phylogenetic independent contrasts using Python 2.7.
-For each contrast, the path length spanned by the contrast (sum of the branch lengths) and the contrast for the mass, natural logarithm of the BMR and natural logarithm of the mass were written as a row in the [bmr-contrasts.csv](https://github.com/sghosh89/BIVAN/blob/master/Data/BMR/bmr-contrasts.csv) file.
-
-[**Third dataset**](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) was a raw data on measurements of anatomical features of several species of *Anolis* lizards, which was downloaded
-from the Dryad data package [@MahlerIRL2013Data] associated with [@MahlerIRL2013].
-The [contrasts.csv file](https://github.com/sghosh89/BIVAN/blob/master/Data/Anolis/contrasts.csv) was produced by using version 4.4.0 of the DendroPy library [@SukumaranH2010]  to calculate phylogenetic independent contrasts on the the data in the GA_Anolis_traits.csv file of Mahler et at.
-using their maximum clade credibility tree (GA_Anolis_MCC.nex) as the phylogeny.
-
-Absolute values of traits for each contrast are considered to construct corresponding copula for second and third datasets.
-
-Fourth, we used **Ceder-creek data for e120** which resulted from a biodiversity experiment, carried out by Tilman and his co-warkers [@Tilman2001; @Tilman2006] over many years to collect [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120) data.
-These datasets were used to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m),
-for year 2000.
-
-The raw data and copula for each four datasets are shown along each column of Figure \ref{MT-fig_biv_raw_cop}.
-
-Similarly, we applied multivariate analysis on four datasets; first two data sets [for aphid-count and aphid-firstflight date] were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014].
-These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species.
-We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony.
-In our case, we chose species=10 [Name : "Green spruce aphid"] for aphid-abundance data and species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data and considered only those locations which had atleast 30 years of data (see Figure \ref{fig_aphid_map}).
-
-Figure \ref{MT-fig_aphid_raw_cop} represents raw and copula data plot for aphid count ("Green spruce aphid") and aphid-firstflight date ("Leaf-curling plum aphid") taken from two locations.
-
-Our third dataset for multivariate analysis is a plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species and for 56 years (1958-2013).
-These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
-The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
-unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
-location in the ocean.
-Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979].
-All the samples available for a given variable in a given area in a given month of a given year were
-averaged, producing a monthly time series for each year.
-Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).
-Monthly values were converted to yearly values by averaging over all months.
-This approach was used by [@Sheppard2017] earlier.
-But here we only allowed maximum 4 months in a year with missing data which were replaced by those medians.
-If a year got 5 or more missing monthly data value then we set NA value for that particular year.
-We used data for species=16 [Name : "Ceratium furca"] which is a dicoflagellate commonly found in UK seas.
-Among 26 locations, only those having atleast 45 years time series data, were considered (see Figure \ref{fig_plankton_map}).
-
-Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the
-Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
-Additional information on the history of the site and soil types is detailed in @Jarecke2016.
-Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at
-daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
-additional details???** on greenhouse gas flux methods for this site. We threw out all locations that had fewer than 50 dates data.-->
+to be filled later
+--->
+
+<!--# Suppmat for model selection approach \label{MS_detailed}
+to be filled later
+--->
 
@@
 op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Green-spruce aphid count (loc. 4)",ylab="Green-spruce aphid count (loc. 11)",cex.lab=2.5,cex.axis=2)
+plot(d1$Dat,d2$Dat,type="p",xlab="Green-spruce aphid count (loc. 4)",ylab="Green-spruce aphid count (loc. 11)",cex.lab=2.5,cex.axis=2)
 par(op)
@@
 op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(m[,1],m[,2],type="p",col="red",xlab="u",ylab="v",
+plot(m[,1],m[,2],type="p",xlab="u",ylab="v",
      cex.lab=3,cex.axis=2,
@@
 op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(d1$Dat,d2$Dat,type="p",col="red",xlab="Leaf-plum curling's first-flight (loc. 2)",ylab="Leaf-plum curling's first-flight (loc. 5)",cex.lab=2.5,cex.axis=2)
+plot(d1$Dat,d2$Dat,type="p",xlab="Leaf-plum curling's first-flight (loc. 2)",ylab="Leaf-plum curling's first-flight (loc. 5)",cex.lab=2.5,cex.axis=2)
 par(op)
@@
     geps<-pop_g$g_noise
-    plot(mg[,1],geps[,1],col="red",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=1")
-    plot(mg[,2],geps[,2],col="red",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=2")
-    plot(pop[,1],pop[,2],col="red",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2)
+    plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=1")
+    plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=2")
+    plot(pop[,1],pop[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2)
     title(expression(paste("g = ",g[1]," , ")), adj=0.06,cex.main=1.5)
-    mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9, col="black")
+    mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)
 
@@
     geps<-pop_d_g$g_noise
-    plot(mg[,1],geps[,1],col="blue",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=1")
-    plot(mg[,2],geps[,2],col="blue",cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=2")
-    plot(pop_d[,1],pop_d[,2],col="blue",xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2)
+    plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=1")
+    plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=2")
+    plot(pop_d[,1],pop_d[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2)
     title(expression(paste("g = ",g[2]," , ")), adj=0.06,cex.main=1.5)
-    mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9, col="black")
+    mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)
 
@@
 p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
-mtext(paste0("p = ",p_left),side=3,line=-0.5,col="navyblue",cex = 2)
+mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
 par(op)
@@
 p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog
-mtext(paste0("p = ",p_left),side=3,line=-0.5,col="navyblue",cex = 2)
+mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
 par(op)
@@
 p_right<-sum(answer2k$surrogskw > answer2k$realskw)/numsurrog
-mtext(paste0("p = ",p_right),side=3,line=-0.5,col="navyblue",cex = 2)
+mtext(paste0("p = ",p_right),side=3,line=-0.5,cex = 2)
 par(op)
@@
 p_right<-sum(answer2s$surrogskw > answer2s$realskw)/numsurrog
-mtext(paste0("p = ",p_right),side=3,line=-0.5,col="navyblue",cex = 2)
+mtext(paste0("p = ",p_right),side=3,line=-0.5,cex = 2)
 par(op)
@@
 p_left<-sum(answer3k$surrogskw < answer3k$realskw)/numsurrog
-mtext(paste0("p = ",p_left),side=3,line=-0.5,col="navyblue",cex = 2)
+mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
 par(op)
@@
 p_left<-sum(answer3s$surrogskw < answer3s$realskw)/numsurrog
-mtext(paste0("p = ",p_left),side=3,line=-0.5,col="navyblue",cex = 2)
+mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
 par(op)
@@
 saveRDS(ans,"./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
-mtext(paste0("p= ",ans$p_left_mncurv),side = 3, line=-2, adj=0.5, col="navyblue",cex=2)
+mtext(paste0("p= ",ans$p_left_mncurv),side = 3, line=-2, adj=0.5, col="black",cex=2)
 par(op)
@@
 plot(ans$realTL_stats$s.points,xlab=expression(log[10]("spatial mean")),
-     ylab=expression(log[10]("spatial variance")),col="red",
+     ylab=expression(log[10]("spatial variance")),col="black",
      cex.axis=1.5,cex.lab=1.5)
-mtext(paste0("mean curvature= ",round(unname(ans$realTL_stats$s.stats[7]),3)),side = 3, line=-2, adj=0, col="navyblue",cex=2)
+mtext(paste0("mean curvature= ",round(unname(ans$realTL_stats$s.stats[7]),3)),side = 3, line=-2, adj=0, col="black",cex=2)
 
@@
      cex.axis=1.5,cex.lab=1.5)
-mtext(paste0("mean curvature= ",round(ans$good_surrog_mncurv[id],3)),side = 3, line=-2, adj=0, col="navyblue",cex=2)
+mtext(paste0("mean curvature= ",round(ans$good_surrog_mncurv[id],3)),side = 3, line=-2, adj=0, col="black",cex=2)
 

### 2018-11-05 getcopula.R

@@ -31,7 +31,7 @@ getcopula<-function(d,rankon=T,ploton=F)
   #plot
   if(ploton==T){
     op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-    plot(v[,1],v[,2],type="p",col="red",xlab="u",ylab="v",
+    plot(v[,1],v[,2],type="p",xlab="u",ylab="v",
          cex.lab=2.5,cex.axis=1.5,
          xlim=c(0,1),ylim=c(0,1),asp=1)
     rect(0,0,1,1)

### 2018-11-06 Paper.Rmd
changed...to be continued

@@
 \textbf{ \hspace{2.8 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
-\vspace{-0.35 cm}
+\vspace{-0.6 cm}
 \begin{center}
@@
 \textbf{ \hspace{2.8 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
-\vspace{-0.35 cm}
+\vspace{-0.6 cm}
 \begin{center}
@@
 \textbf{ \hspace{7 cm} (I) \hspace{3.2 cm} (J)} \\
-\vspace{-0.35 cm}
+\vspace{-0.6 cm}
 \begin{center}
-\includegraphics[width=3.8 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
-\includegraphics[width=3.8 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
+\includegraphics[width=4 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
+\includegraphics[width=4 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
 \end{center}
 \textbf{ \hspace{7 cm} (K) \hspace{3.2 cm} (L)} \\
-\vspace{-0.35 cm}
+\vspace{-0.6 cm}
 \begin{center}
-\includegraphics[width=3.8 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
-\includegraphics[width=3.8 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
+\includegraphics[width=4 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
+\includegraphics[width=4 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
 \caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs.
@@
 <!--Question 2 analyses-->
-# Some possible causes of non-Normal copula structure \label{Causes}
+# Concepts for \hyperref[Intro]{\textbf{(Q2)}}: possible causes of non-Normal copula structure \label{Causes}
 
 This section is dedicated to exploring question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}.
+We showed environmental copula structure and asymmetric sensitivities to the environment could produce population copula structures.
 <!--We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change?-->
@@
 \begin{center}
-\includegraphics[width=16 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
+\includegraphics[width=15 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
 \caption{Effects of asymmetric global noise copula (resulting from the asymmetric dependence of $g$ on $\varepsilon$) on two habitat patch model (Eq.
@@
 <!--Question 3 analyses-->
-# Consequences of non-Normal copula structure and tail dependence for our understanding of biology \label{Consequences}
+# Concepts for \hyperref[Intro]{\textbf{(Q3)}}: Consequences of non-Normal copula structure and tail dependence for our understanding of biology \label{Consequences}
 
@@
 We predicted that because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, should change significantly from the individual skewness of time-series data.
+We tested null hypothesis $(H_{0})$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
+We generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same and calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set in the next section (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R)).
+We made a histogram of surrogate skewness and found if actual skewness was within the $95\%$ confidence interval or not.
+If the $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness than Normal copula).
+
 We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL_simulation}).
@@
 \ref{SM-fig_spTL}(I)).
-
 In the following section we will show application of these ideas with real datasets.
@@
 \begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
+\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
+\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
 \end{center}
@@
 \begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
+\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
+\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
 \caption{Skewness of spatially averaged $(A)$ green-spruce aphid abundance, $(B)$ leaf-curling plum aphid firstflight date, $(C)$ Ceratium furca abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation).
@@
 
-## Application : Change of skewness
+## Results for \hyperref[Intro]{\textbf{(Q3)}}: Change of skewness
 
 We used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data, to compute skewness of spatially averaged data.
-We tested null hypothesis $(H_{0})$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
-We generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same and calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R)).
-We made a histogram of surrogate skewness and found if actual skewness was within the $95\%$ confidence interval or not.
-If the $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness than Normal copula).
-We carried out the same analysis with aphid abundance and aphid firstflight date data for all 20 aphid species and plankton abundance data for all 22 plankton species.
+We carried out the analysis mentioned in the concepts above, with aphid abundance and aphid firstflight date data for all 20 aphid species and plankton abundance data for all 22 plankton species.
 In each study, we got $>5\%$ positive results which gave us confidence in validating significantly different skewness for asymmetric copula than Normal.
@@
 
-## Application : Change of curvature of spatial Taylor's law regression line
+## Results for \hyperref[Intro]{\textbf{(Q3)}}: Change of curvature of spatial Taylor's law regression line
 
@@
 
-\newpage
+\clearpage
+
 # References

### 2018-11-06 SupportingInformation.Rmd
changed...to be continued

@@
 pdf("./Results/SoilCNraw.pdf")
-op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(d[,1],d[,2],type="p",col="black",xlab="Soil C",ylab="Soil N",cex.lab=2.5,cex.axis=1.5)
+op<-par(mar=c(6, 6, 0.2, 2.1))
+plot(d[,1],d[,2],type="p",col="black",xlab="Soil C",ylab="Soil N",cex.lab=2.5,cex.axis=2.5)
 par(op)
@@
 pdf("./Results/BMR_results/birds/raw_birds_BMR.pdf")
-op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(b[,1],b[,2],type="p",col="black",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=1.5)
+op<-par(mar=c(6, 6, 0.2, 2.1))
+plot(b[,1],b[,2],type="p",col="black",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=2.5)
 par(op)
@@
 pdf("./Results/BMR_results/mammals/raw_mammals_BMR.pdf")
-op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(m[,1],m[,2],type="p",col="black",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=1.5)
+op<-par(mar=c(6, 6, 0.2, 2.1))
+plot(m[,1],m[,2],type="p",col="black",xlab="log(Mass)",ylab="log(BMR)",cex.lab=2.5,cex.axis=2.5)
 par(op)
@@
     if(rawplot==T){
-      op<-par(mar=c(5.1, 5.1, 1.5, 2.1))
-      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg.H",ylab="avg.Biomass",pch=1,cex.lab=2,cex.axis=1.5)
-      mtext(paste0("Year=",year),cex=2)
+      op<-par(mar=c(6, 6, 0.2, 2.1))
+      plot(tab_dbp$avg.H,tab_dbp$avg.Biomass,xlab="avg. H",ylab="avg. Biomass",pch=1,cex.lab=2.5,cex.axis=2)
+      mtext(paste0("Year=",year),cex=1)
       par(op)
     }else{
-      op<-par(mar=c(5.1, 5.1, 1.5, 2.1))
-      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",pch=1,cex.lab=2,cex.axis=1.5,
+      op<-par(mar=c(6, 6, 0.2, 2.1))
+      plot(cop_tab_dbp$avg.H,cop_tab_dbp$avg.Biomass,xlab="u",ylab="v",pch=1,cex.lab=3,cex.axis=2,
          xlim=c(0,1),ylim=c(0,1),asp=1)
       rect(0,0,1,1)
-      lines(c(0,1),c(0,1),type='l')
-      mtext(paste0("Year=",year),cex=2)
+      lines(c(0,1),c(0,1),type='l',col="grey")
+      #mtext(paste0("Year=",year),cex=1)
       par(op)
@@
 pdf("./Results/Aphid_count_raw_loc_4_11.pdf",height = 8,width=8)
-op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(d1$Dat,d2$Dat,type="p",xlab="Green-spruce aphid count (loc. 4)",ylab="Green-spruce aphid count (loc. 11)",cex.lab=2.5,cex.axis=2)
+op<-par(mar=c(10,10, 0.2, 2.1),mgp=c(3,1,0))
+plot(d1$Dat,d2$Dat,type="p",xlab="",
+     ylab="Green-spruce aphid \ncount (loc. 11)",cex.lab=3.5,cex.axis=2.5)
+mtext(paste("Green-spruce aphid \ncount (loc. 4)"),line=-38,cex=3.5)
 par(op)
@@
 pdf("./Results/Aphid_count_cop_loc_4_11.pdf",height = 8,width=8)
-op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
+op<-par(mar=c(6,6, 0.2, 2.1),mgp=c(3,1,0))
 plot(m[,1],m[,2],type="p",xlab="u",ylab="v",
-     cex.lab=3,cex.axis=2,
+     cex.lab=4,cex.axis=3,
      xlim=c(0,1),ylim=c(0,1),asp=1)
 rect(0,0,1,1)
-lines(c(0,1),c(0,1),type='l')
+lines(c(0,1),c(0,1),type='l',col="grey")
 par(op)
@@
 pdf("./Results/Aphid_ff_raw_loc_2_5.pdf",height = 8,width=8)
-op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
-plot(d1$Dat,d2$Dat,type="p",xlab="Leaf-plum curling's first-flight (loc. 2)",ylab="Leaf-plum curling's first-flight (loc. 5)",cex.lab=2.5,cex.axis=2)
+op<-par(mar=c(10,10, 0.2, 2.1),mgp=c(3,1,0))
+plot(d1$Dat,d2$Dat,type="p",xlab="",ylab="Leaf-plum curling's \nfirst-flight (loc. 5)",cex.lab=3.5,cex.axis=2.5)
+mtext(paste("Leaf-plum curling's \nfirst-flight (loc. 2)"),line=-38,cex=3.5)
 par(op)
@@
 pdf("./Results/Aphid_ff_cop_loc_2_5.pdf",height = 8,width=8)
-op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
+op<-par(mar=c(6,6, 0.2, 2.1),mgp=c(3,1,0))
 plot(m[,1],m[,2],type="p",xlab="u",ylab="v",
-     cex.lab=3,cex.axis=2,
+     cex.lab=4,cex.axis=3,
      xlim=c(0,1),ylim=c(0,1),asp=1)
 rect(0,0,1,1)
-lines(c(0,1),c(0,1),type='l')
+lines(c(0,1),c(0,1),type='l',col="grey")
 par(op)

### 2018-11-06 getcopula.R

@@ -30,12 +30,12 @@ getcopula<-function(d,rankon=T,ploton=F)
   
   #plot
   if(ploton==T){
-    op<-par(mar=c(5.1, 5.1, 0.2, 2.1))
+    op<-par(mar=c(6,6, 0.2, 2.1))
     plot(v[,1],v[,2],type="p",xlab="u",ylab="v",
-         cex.lab=2.5,cex.axis=1.5,
+         cex.lab=3.5,cex.axis=2.5,
          xlim=c(0,1),ylim=c(0,1),asp=1)
     rect(0,0,1,1)
-    lines(c(0,1),c(0,1),type='l')
+    lines(c(0,1),c(0,1),type='l',col="grey")
     par(op)
   }
   

### 2018-11-06 Paper.Rmd
primarily completed editting...now it has the same structure as we discussed@Nov2

@@
 
-## Data \label{Data}
+# Data \label{Data}
 
@@
 
-## Concepts for \hyperref[Intro]{\textbf{(Q1)}}: Statistical tools for detecting copula structure \label{Methods}
+# Concepts for \hyperref[Intro]{\textbf{(Q1)}}: Statistical tools for detecting copula structure \label{Methods}
 
@@
 
-### Model selection approach \label{Model_selection}
+## Model selection approach \label{Model_selection}
 
@@
 
-### Nonparametric analysis of tail dependence \label{NPA}
+## Nonparametric analysis of tail dependence \label{NPA}
 
@@
 <!--Question 2 analyses-->
-# Concepts for \hyperref[Intro]{\textbf{(Q2)}}: possible causes of non-Normal copula structure \label{Causes}
+# Concepts and models for \hyperref[Intro]{\textbf{(Q2)}}: possible causes of non-Normal copula structure \label{Causes}
 
@@
 <!--We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change?--> 
-
 <!--Moran effects-->
-\textbf{(i)} We were interested to extend the basic idea of Moran's theorem : if noise $\epsilon_{1}(t)$ and $\epsilon_{2}(t)$ at two different habitat patches had some specific assymetric tail dependence then what will be the outcome dependence in the populations at those habitat patches.
+\textbf{(i)} We were interested to extend the basic idea of Moran's theorem : if environmental noises at two different habitat patches had some specific assymetric tail dependence then what will be the outcome dependence in the populations at those habitat patches.
 Should their dependence be exactly the same as of the noises or not ?
-
 To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
@@
 To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$.
-To test environmental effect on population copula, our approach is of two-fold :
+We showed with Eq.
+\ref{eq.AR1} in the next section that both model selection method and non-parametric analysis indicated similar tail dependence in population copula as that of noise copula.
+<!--Asymmetric sensitivity and copula structure in populations-->
+\textbf{(ii)} Secondly, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) which behaved differently beyond a certain expectation threshold (as shown in Eq.
+\ref{eq.AS_g}).
+To test the assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;
 
-First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2007_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
+\begin{equation}\label{eq.AS}
+P_{i}(t+1) = b P_{i}(t) + g(\varepsilon_{i}(t)) + h(\delta_{i}(t))
+\end{equation}
+
+$b$ was the linear growth rate.
+Global environmental fluctuation function $g(\varepsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq.
+\ref{eq.AS_g}, Fig.
+\ref{fig_ms_asym_b_0.1}) where $\varepsilon=(\varepsilon_{1}, \varepsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\varepsilon}$.
+Local environmental fluctuation function $h(\delta)$ varied according to Eq.
+\ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$.
+$\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$.
+With this model we showed tail dependence in population copula changed from lower to upper tail dependence as global noise copula $g(\varepsilon)$ changed its form from $g_{1}$ to $g_{2}$ (Eq.
+\ref{eq.AS_g}).
+
+\begin{align}
+g(\varepsilon) = \begin{cases} 
+g_{1} = \begin{cases}
+\varepsilon & \text{if $\varepsilon< 0$}; \\
+0 & \text{otherwise}.
+\end{cases}\\ \\
+g_{2} = \begin{cases}
+\varepsilon & \text{if $\varepsilon> 0$}; \\
+0 & \text{otherwise}\label{eq.AS_g}.
+\end{cases}
+\end{cases} \text{where } \varepsilon = N(\textbf{0},\Sigma_{\varepsilon})\;\ \text{with} \;\ \Sigma_{\varepsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
+\end{align}
+
+\begin{equation}\label{eq.AS_h}
+h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
+\end{equation}
+
+# Results for \hyperref[Intro]{\textbf{(Q2)}}: similar tail dependence in population copula as that of noise copula \label{Results_Q2}
+
+To test environmental effect on population copula from the model \ref{eq.AR1}, our approach is of two-fold : First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2007_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
 Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
@@
 
-<!--Asymmetric sensitivity and copula structure in populations-->
-\textbf{(ii)} Secondly, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) which behaved differently beyond a certain expectation threshold (as shown in Eq.
-\ref{eq.AS_g} and Fig.
-\ref{fig_ms_asym_b_0.1}).
-
-To test the assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;
-
-\begin{equation}\label{eq.AS}
-P_{i}(t+1) = b P_{i}(t) + g(\varepsilon_{i}(t)) + h(\delta_{i}(t))
-\end{equation}
-
-$b$ was the linear growth rate.
-Global environmental fluctuation function $g(\varepsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq.
-\ref{eq.AS_g}, Fig.
-\ref{fig_ms_asym_b_0.1}) where $\varepsilon=(\varepsilon_{1}, \varepsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\varepsilon}$.
-
-\begin{align}
-g(\varepsilon) = \begin{cases} 
-g_{1} = \begin{cases}
-\varepsilon & \text{if $\varepsilon< 0$}; \\
-0 & \text{otherwise}.
-\end{cases}\\ \\
-g_{2} = \begin{cases}
-\varepsilon & \text{if $\varepsilon> 0$}; \\
-0 & \text{otherwise}\label{eq.AS_g}.
-\end{cases}
-\end{cases} \text{where } \varepsilon = N(\textbf{0},\Sigma_{\varepsilon})\;\ \text{with} \;\ \Sigma_{\varepsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
-\end{align}
-
-Local environmental fluctuation function $h(\delta)$ varied according to Eq.
-\ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$.
-$\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$.
-
-\begin{equation}\label{eq.AS_h}
-h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
-\end{equation}
-
-We simulated our model (\ref{eq.AS}) for $b=0.1$ , $a=0.2$, $r=0.8$.
+To explore the effect of asymmetric sensitivity, we simulated our model (\ref{eq.AS}) for $b=0.1$ , $a=0.2$, $r=0.8$.
 $10000$ points were generated and we kept only the last $1000$ points to plot the population-copula (Fig.
 \ref{fig_ms_asym_b_0.1}) for each specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
-When asymmetric global noise
-had only non-positive contributions at two habitat patches ($g_{1}$ in the first panel of Fig.
+When asymmetric global noise had only non-positive contributions at two habitat patches ($g_{1}$ in the first panel of Fig.
 \ref{fig_ms_asym_b_0.1}), lower tail dependence appeared in the global noise copula and so thus in affected population copula (last subplot in the first panel of Fig.

### 2018-11-12 SupportingInformation.Rmd
adding tinytex library command

@@
 library(tibble)
-aphid_info <- tibble(c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
+aphid_info <- tibble(c1=c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
                        "Corn leaf aphid","Currant lettuce aphid","Damson hop aphid","Grain aphid","Green spruce aphid",
@@
                        "Potato aphid","Rose grain aphid","Shallot aphid","Sycamore aphid","Willow carrot aphid"),
-           c("Rhopalosiphum insertum", "Rhopalosiphum padi", "Aphis fabae", "Sitobion fragariae", "Hyperomyzus lactucae",
+                     c2=c("Rhopalosiphum insertum", "Rhopalosiphum padi", "Aphis fabae", "Sitobion fragariae", "Hyperomyzus lactucae",
                        "Rhopalosiphum maidis", "Nasonovia ribisnigri","Phorodon humuli","Sitobion avenae","Elatobium abietinum",

### 2018-11-12 SupportingInformation.Rmd
modified

--- 2018-11-06 SupportingInformation.Rmd
@@
 ```{r tab_aphid_info, echo=F, results='asis',message=F}
+library(tinytex)
 library(tibble)
-aphid_info <- tibble(c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
+aphid_info <- tibble(c1=c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
                        "Corn leaf aphid","Currant lettuce aphid","Damson hop aphid","Grain aphid","Green spruce aphid",

### 2018-11-26 Paper.Rmd
edited Data section

@@
 
-<!--
-Considered datasets ranged widely from population ecology, community ecology , evolutionary and biogeochemical sphere. 
-We used four datasets in each category of bivariate (soil C and N data, BMR data, Anolis data, Ceder-creek experiment data) and multivariate (aphid-abundance data, aphid-firstflight data, plankton-abundance data and methane-flux data) examples.
-Details of each dataset used were given in \ref{SM-Data_detailed}.-->
+\noindent We used datasets that pertain to multiple fields and levels of organization within or related to ecology, including environmental, species-trait, phenological, population, community, and ecosystem functioning data.
+Copula structure describing the relationship between atmospheric weather variables such as X and Y has been examined previously in the meteorological literature (e.g., REF).
+We therefore examined environmental variables from the soil instead, using the Rapid Carbon Assessment database (RaCA; @Wills2014).
+<!--***DAN: Add Wills and Loecke pending reference if published in time--> 
+The database comprises measurements of soil organic carbon and total soil nitrogen (milligrams C or N per m$^{2}$ of soil surface) 
+<!--***DAN: Shya, this used to say Mg C, but what is Mg?
+Do you mean mg for miligram?
+I have written it out, please check and correct if necessary.
+Also, I added per m^2 of soil surface, please check.-->
+from X 
+<!--***DAN: Shya, make this an autolink to the number of data points - 5930, but should be an autolink, so if the number changes it updates-->
+locations across the coterminous United States (Fig. X for sampling locations).
+<!--***DAN: Shya, please add a sup mat figure showing the sampling locations and a link to it here in place of the X-->
+Measurments were from the central pedon of a soil core, to 1 m depth.
+<!--***DAN: Terry to check the wording and accuracy here-->
+Species-trait data were average species basal metabolic rate (units) and body mass (units) data for X species of birds [@Mcnab_2009] and Y species of mammals [@Mcnab_2008].
+<!--***DAN: Shya to add units, and replace X and Y with autolinks that give the numbers-->
+These data have been much studied, but to our knowledge the copula structure of the dependence between metabolic rate and body mass has not been examined. 
+Species-trait data such as these reflect the coevolution of the two traits.
+Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
+<!--***DAN: Shya to add a figure shwoing the locations on a map of the UK and cite it here with an automatic citation-->
+across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 35 years from 1976 to 2010.
+These time series were computed from the the Rothamsted Insect Survey suction-trap dataset [@Macaulay1988, @Harrington2014].
+The first of our two population-level datasets was also derived from the RIS suction-trap data, and comprised total counts of aphids trapped for the same locations, species, and years; these are indicators of flying aphid abundance.
+Data pre-processing for the aphid data was the same as that of @Sheppard2016,
+<!--***DAN: Shya, let's discuss please - if we say this we have to be sure we did *exactly* the same pre-processing as Lawrence, whereas I seem to remember we did something slightly different-->
+but we exlcuded one of the 11 sites they used so as to include only sites with at least 30 years data.
+Our second population dataset comprised average annual abundance estimates for 14 locations (Fig.
+\ref{SM-fig_plankton_map})
+in the North Sea and British seas for 22 plankton taxa (Table \ref{SM-tab_plankton_info}) for the 56 years from 1958 to 2013.
+The locations are $2^\circ$ by $2^\circ$ grid cells.
+These data were obtained from the Continuous Plankton Recorder (CPR) dataset maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS), now part of the Marine Biological Association of the UK.
+Data pre-processing was similar but not identical to that of @Sheppard2017; see Appendix \ref{SM-Data_details} for details.
+Community-level data were plant aboveground biomass and Shannon's diversity index for X plots, each 9m by 9m, maintained at Cedar Creek Ecosystem Science Reserve and sampled in the year 2000 [@Tilman2001; @Tilman2006].
+Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or water surface and the troposphere, measured at 13 locations at daily to weekly intervals from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, Ohio [@Holland1999; @Jarecke2016].
+Each included location was measured on at least 50 dates.
+Sampling locations spanned a hydrologic gradient from upland to seasonal wetland.
+See Appendix \ref{SM-Data_details} for additional data details.
 
-For bivariate analysis we used four datasets which are briefly described as follows;
-\textbf{(i)} to explore the dependence of **soil organic C** and total **soil N** across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
-database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
-stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
-study.
-Then, we used extensive set of basal metabolic rate data for \textbf{(ii)} birds [@Mcnab_2009] and \textbf{(iii)} mammals [@Mcnab_2008] as our second and third bivariate datasets to explore the dependence between log(BMR) and log(mass) of respective species.
-\textbf{(iv)} We used [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120) data from a biodiversity experiment [@Tilman2001; @Tilman2006] to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m), for year 2000.
-
-Similarly, we applied multivariate analysis on four spatial datasets; \textbf{(i)} aphid-count and \textbf{(ii)} aphid-firstflight date were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014].
-These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species (Table \ref{SM-tab_aphid_info}).
-We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony.
-In our case, for aphid-abundance data and for aphid-firstflight data, only 10 locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}) were considered.
-\textbf{(iii)} Plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species (Table \ref{SM-tab_plankton_info}) and for 56 years (1958-2013).
-These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
-<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
-unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
-location in the ocean.
-Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979].
-All the samples available for a given variable in a given area in a given month of a given year were
-averaged, producing a monthly time series for each year.
-Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).--> [@Sheppard2017] converted monthly values to yearly values by averaging over all months.
-But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available).
-We used data from 14 locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}).
-\textbf{(iv)} Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
-Additional information on the history of the site and soil types is detailed in @Jarecke2016.
-Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016.
-**See and Smyth et al, pending for
-additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.
+The environmental, species-trait, and community datasets, as we studied them, are henceforth called the "bivariate" datasets in the sense that they comprise two quantities measured at different locations or for different species.
+The phenology and population datasets are "multivariate" in that they comprise, for each taxon, measurements at multiple locations; copula structure can be studied for each pair of locations.
+Likewise the the ecosystem functioning (methane flux) dataset was  multivariate in this sense.
 

### 2018-11-26 SupportingInformation.Rmd
edited Data section

@@
 --->
+
+# Details of the data \label{Data_details}
+
+<!--***DAN: This is a dumping group at the moment, but needs to be organized into a section that gives all the necessary details about the data-->
+
+<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
+unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
+location in the ocean.
+Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979].
+All the samples available for a given variable in a given area in a given month of a given year were
+averaged, producing a monthly time series for each year.
+Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).-->
+
+ [@Sheppard2017] converted monthly values to yearly values by averaging over all months.
+But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available).
+We used data from 14 locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}).
+
+ The Great Miami Wetland Mitigation Bank  was at $36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
+Additional information on the history of the site and soil types is detailed in @Jarecke2016.
+Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016.
+**See and Smyth et al, pending for
+additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.
 

### 2018-11-27 Paper.Rmd
reply to editted Data section : still need to discus

@@
 
-\noindent We used datasets that pertain to multiple fields and levels of organization within or related to ecology, including environmental, species-trait, phenological, population, community, and ecosystem functioning data.
+```{r read_somedata, echo=F, results='hide'}
+d_soilCN<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
+d_birdsBMR<-readRDS("./Data/BMR/BirdBodyMassesMetabolicRates/my_birdsBMR_data.RDS")
+d_mammalsBMR<-readRDS("./Data/BMR/MammalianBodyMassesMetabolicRates/my_mammalsBMR_data.RDS")
+```
+
+We used datasets that pertain to multiple fields and levels of organization within or related to ecology, including environmental, species-trait, phenological, population, community, and ecosystem functioning data.
 Copula structure describing the relationship between atmospheric weather variables such as X and Y has been examined previously in the meteorological literature (e.g., REF).
@@
 <!--***DAN: Add Wills and Loecke pending reference if published in time--> 
-The database comprises measurements of soil organic carbon and total soil nitrogen (milligrams C or N per m$^{2}$ of soil surface) 
-<!--***DAN: Shya, this used to say Mg C, but what is Mg?
-Do you mean mg for miligram?
-I have written it out, please check and correct if necessary.
-Also, I added per m^2 of soil surface, please check.-->
-from X 
-<!--***DAN: Shya, make this an autolink to the number of data points - 5930, but should be an autolink, so if the number changes it updates-->
+The database comprises measurements of soil organic carbon and total soil nitrogen (Mg C or N per m$^{2}$ of soil surface) 
+<!--***DAN: Shya, this used to say Mg C, but what is Mg? Do you mean mg for miligram? I have written it out, please check and correct if necessary. Also, I added per m^2 of soil surface, please check.
+***Shya: Mg means mega gram, So an "Mg C" is a ton of Carbon - the standard SI unit when it comes to talking about soil Carbon measurement.-->
+from `r nrow(d_soilCN)` 
+<!--***DAN: Shya, make this an autolink to the number of data points - 5930, but should be an autolink, so if the number changes it updates
+***Shya: Done.-->
 locations across the coterminous United States (Fig. X for sampling locations).
@@
 <!--***DAN: Terry to check the wording and accuracy here-->
-Species-trait data were average species basal metabolic rate (units) and body mass (units) data for X species of birds [@Mcnab_2009] and Y species of mammals [@Mcnab_2008].
+Species-trait data were average species basal metabolic rate (KJ per hour) and body mass (gram) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
 <!--***DAN: Shya to add units, and replace X and Y with autolinks that give the numbers-->
+<!--***Shya: These species number for birds and mammals came from the processed data without any NA and -ve(unrealistic) mass, is that okay? 
+for example, species of birds [@Mcnab_2009] raw data (the xls file you gave me) shows 712 species out of which some have NA in both mass and BMR column and one entry has -ve mass!!!-->
 These data have been much studied, but to our knowledge the copula structure of the dependence between metabolic rate and body mass has not been examined. 
@@
 Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
-<!--***DAN: Shya to add a figure shwoing the locations on a map of the UK and cite it here with an automatic citation-->
+<!--***DAN: Shya to add a figure shwoing the locations on a map of the UK and cite it here with an automatic citation
+***Shya: It is already in suppmat figure S1.-->
 across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 35 years from 1976 to 2010.

### 2018-11-28 Paper.Rmd
responses on the Data section, Shya please look at one comment in particular

@@
 ***Shya: Mg means mega gram, So an "Mg C" is a ton of Carbon - the standard SI unit when it comes to talking about soil Carbon measurement.-->
+<!--***DAN: Can you check this please?
+I do not think it can be right.
+The values on Fig.
+3a range from close to 0 to over 3000.
+3000 would be 3000 Mg C.
+I do not think a cubic meter of soil even weighs that much, including all the carbon and non-carbon components.-->
 from `r nrow(d_soilCN)` 
-<!--***DAN: Shya, make this an autolink to the number of data points - 5930, but should be an autolink, so if the number changes it updates
-***Shya: Done.-->
 locations across the coterminous United States (Fig. X for sampling locations).
@@
 for example, species of birds [@Mcnab_2009] raw data (the xls file you gave me) shows 712 species out of which some have NA in both mass and BMR column and one entry has -ve mass!!!-->
+<!--***DAN: Yes, that makes sense, but please go back to the data as published, which should be a supporting appendix to the original publications, and make sure I have not given you erroneous data-->
 These data have been much studied, but to our knowledge the copula structure of the dependence between metabolic rate and body mass has not been examined. 
@@
 Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
-<!--***DAN: Shya to add a figure shwoing the locations on a map of the UK and cite it here with an automatic citation
-***Shya: It is already in suppmat figure S1.-->
 across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 35 years from 1976 to 2010.
@@
 Data pre-processing for the aphid data was the same as that of @Sheppard2016,
-<!--***DAN: Shya, let's discuss please - if we say this we have to be sure we did *exactly* the same pre-processing as Lawrence, whereas I seem to remember we did something slightly different-->
 but we exlcuded one of the 11 sites they used so as to include only sites with at least 30 years data.

### 2018-11-28 Paper.Rmd
replied to edited Data section

@@
 <!--***DAN: Add Wills and Loecke pending reference if published in time--> 
-The database comprises measurements of soil organic carbon and total soil nitrogen (Mg C or N per m$^{2}$ of soil surface) 
+The database comprises measurements of soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil surface) 
 <!--***DAN: Shya, this used to say Mg C, but what is Mg? Do you mean mg for miligram? I have written it out, please check and correct if necessary. Also, I added per m^2 of soil surface, please check.
@@
 3000 would be 3000 Mg C.
-I do not think a cubic meter of soil even weighs that much, including all the carbon and non-carbon components.-->
+I do not think a cubic meter of soil even weighs that much, including all the carbon and non-carbon components.
+***Shya: I checked : it should be megagram per hectare, Wills2014 reference is for stock of Carbon, not for nitrogen, Prof.
+loecke told me nitrogen data is not published yet.
+He screened locations which had both soil C and N stocks after omiting NAs.-->
 from `r nrow(d_soilCN)` 
 locations across the coterminous United States (Fig. X for sampling locations).
-<!--***DAN: Shya, please add a sup mat figure showing the sampling locations and a link to it here in place of the X-->
+<!--***DAN: Shya, please add a sup mat figure showing the sampling locations and a link to it here in place of the X
+***Shya: once Prof. Loecke give me the lattitude and longitudes, I will do that.-->
 Measurments were from the central pedon of a soil core, to 1 m depth.
@@
 Species-trait data were average species basal metabolic rate (KJ per hour) and body mass (gram) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
-<!--***DAN: Shya to add units, and replace X and Y with autolinks that give the numbers-->
-<!--***Shya: These species number for birds and mammals came from the processed data without any NA and -ve(unrealistic) mass, is that okay? 
-for example, species of birds [@Mcnab_2009] raw data (the xls file you gave me) shows 712 species out of which some have NA in both mass and BMR column and one entry has -ve mass!!!-->
-<!--***DAN: Yes, that makes sense, but please go back to the data as published, which should be a supporting appendix to the original publications, and make sure I have not given you erroneous data-->
 These data have been much studied, but to our knowledge the copula structure of the dependence between metabolic rate and body mass has not been examined. 

### 2018-11-28 SupportingInformation.Rmd
reply to editted Data section : still need to discus
replied to edited Data section

@@
 raw_b<-as.data.frame(raw_b)
-raw_b[303,2]<-0.66  #to convert as numeric
-raw_b[,2]<-as.numeric(raw_b[,2])
-raw_b<-raw_b[-which(raw_b[,1]<0),]  # omit -ve mass?
+
+saveRDS(raw_b,"./Data/BMR/BirdBodyMassesMetabolicRates/my_birdsBMR_data.RDS")
 
@@
 raw_m<-raw_m[,2:3]
+
+saveRDS(raw_m,"./Data/BMR/MammalianBodyMassesMetabolicRates/my_mammalsBMR_data.RDS")
 

### 2018-12-01 Paper.Rmd
made at start at revising the introduction

@@
 ---
-title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
 author: "Shyamolina Ghosh, Lawrence W Sheppard, Mark T. Holder, Terrance E. Loecke, Daniel C. Reuman"
@@
 
+All branches of ecology and related areas study relationships among environmental and biological 
+variables. However, commonly used correlation- and regression-based approaches to studying 
+such relationships are limited - it is well understood that common statistical
+approaches may provide only a very partial description of a relationship. 
+For instance, datasets showing markedly different 
+relationships may nevertheless have the same correlation coefficient value:  
+Fig. \ref{fig_cop_pedag1}A-B show two such datasets. The panel A 
+(respectively, B) variables of Fig. \ref{fig_cop_pedag1} are principally 
+related in the left (respectively, right) portions of their distributions, a difference that may
+have ecological significance, as discussed below, but is not captured by standard
+correlation. Panels C and D 
+were obtained from A and B through monotonic transformation (Appendix X for details),
+<!--***DAN: Dan to go back and write the appendix-->
+which alters the Pearson, but not the Spearman correlation. Although correlation
+is only one of the numerous metrics for 
+studying relationships between variables, other commonly used metrics also 
+provide partial descriptions only.
+
+Well-developed statistical tools do exist (REFS), and are applied widely in other fields 
+(REFS), that provide a complete description of the dependence between two or more variables. 
+The goal of this paper is to illustrate the potential for broader application of these
+tools in ecology and related fields. The tools are based on the statistical concept of the *copula*, which we here introduce conceptually.
+
+  a) then do it. some things to do include...
+  b) the rank-based approach for data (use bivariate data)
+  c) the inverse cdf approach for the copula of a bivariate distribution
+  d) statement that copulas are also multivariate, though we won't explore that here (or maybe leave that for Discussion?)
+  e) a one- or two-sentence description of sklars theorem
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
+\caption{(A-F, I, J) Diverse bivariate datasets, many of which have the same Pearson (P)
+or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
+obtained from A (respectively, B) via monotonic functions that transformed the standard-
+normal marginal distributions of A and B to gamma (C, D) or uniform (E, F) marginals. 
+Details in Appendix X.\label{fig_cop_pedag1}}
+\end{center}
+\end{figure}
+<!--***DAN: Dan to go back and add an appendix with details as it says. Make sure
+to mention the shape and scale params of the gamma marginals in C, D were 2-->
+
 Identifying and quantifying dependencies is the core of modelling in many applications, frequently found in biology and elsewhere.
@@
 <!--***DAN: Add Wills and Loecke pending reference if published in time--> 
-The database comprises measurements of soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil surface) 
-<!--***DAN: Shya, this used to say Mg C, but what is Mg? Do you mean mg for miligram? I have written it out, please check and correct if necessary. Also, I added per m^2 of soil surface, please check.
-***Shya: Mg means mega gram, So an "Mg C" is a ton of Carbon - the standard SI unit when it comes to talking about soil Carbon measurement.-->
-<!--***DAN: Can you check this please?
-I do not think it can be right.
-The values on Fig.
-3a range from close to 0 to over 3000.
-3000 would be 3000 Mg C.
-I do not think a cubic meter of soil even weighs that much, including all the carbon and non-carbon components.
-***Shya: I checked : it should be megagram per hectare, Wills2014 reference is for stock of Carbon, not for nitrogen, Prof.
-loecke told me nitrogen data is not published yet.
-He screened locations which had both soil C and N stocks after omiting NAs.-->
-from `r nrow(d_soilCN)` 
+The database comprises measurements of soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
 locations across the coterminous United States (Fig. X for sampling locations).
@@
 Species-trait data were average species basal metabolic rate (KJ per hour) and body mass (gram) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
-These data have been much studied, but to our knowledge the copula structure of the dependence between metabolic rate and body mass has not been examined. 
+These data have been much studied, but to our knowledge the copula structure 
+of the dependence between metabolic rate and body mass has not been examined. 
 Species-trait data such as these reflect the coevolution of the two traits.
 Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
-across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 35 years from 1976 to 2010.
-These time series were computed from the the Rothamsted Insect Survey suction-trap dataset [@Macaulay1988, @Harrington2014].
-The first of our two population-level datasets was also derived from the RIS suction-trap data, and comprised total counts of aphids trapped for the same locations, species, and years; these are indicators of flying aphid abundance.
-Data pre-processing for the aphid data was the same as that of @Sheppard2016,
-but we exlcuded one of the 11 sites they used so as to include only sites with at least 30 years data.
-Our second population dataset comprised average annual abundance estimates for 14 locations (Fig.
-\ref{SM-fig_plankton_map})
-in the North Sea and British seas for 22 plankton taxa (Table \ref{SM-tab_plankton_info}) for the 56 years from 1958 to 2013.
-The locations are $2^\circ$ by $2^\circ$ grid cells.
-These data were obtained from the Continuous Plankton Recorder (CPR) dataset maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS), now part of the Marine Biological Association of the UK.
-Data pre-processing was similar but not identical to that of @Sheppard2017; see Appendix \ref{SM-Data_details} for details.
-Community-level data were plant aboveground biomass and Shannon's diversity index for X plots, each 9m by 9m, maintained at Cedar Creek Ecosystem Science Reserve and sampled in the year 2000 [@Tilman2001; @Tilman2006].
-Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or water surface and the troposphere, measured at 13 locations at daily to weekly intervals from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, Ohio [@Holland1999; @Jarecke2016].
-Each included location was measured on at least 50 dates.
+across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 
+35 years from 1976 to 2010. These time series were computed from the the Rothamsted Insect 
+Survey suction-trap dataset [@Macaulay1988, @Harrington2014]. The first of our two 
+population-level datasets was also derived from the RIS suction-trap data, and comprised 
+total counts of aphids trapped for the same locations, species, and years; these are 
+indicators of flying aphid abundance. Data pre-processing for the aphid data was the same as 
+that of @Sheppard2016,
+but we exlcuded one of the 11 sites they used so as to include only sites with at least 30 
+years data.  Our second population dataset comprised average annual abundance estimates 
+for 14 locations (Fig. \ref{SM-fig_plankton_map})
+in the North Sea and British seas for 22 plankton taxa (Table \ref{SM-tab_plankton_info}) for the 
+56 years from 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. These data 
+were obtained from the Continuous Plankton Recorder (CPR) dataset maintained by the Sir 
+Alistair Hardy Foundation for Ocean Science (SAHFOS), now part of the Marine Biological 
+Association of the UK. Data pre-processing was similar but not identical to that 
+of @Sheppard2017; see Appendix \ref{SM-Data_details} for details.
+Community-level data were plant aboveground biomass and Shannon's diversity index for X 
+<!--***DAN: Shya, please fill in X with an autolink-->
+plots, each 9m by 9m, maintained at Cedar Creek Ecosystem Science Reserve and sampled in 
+the year 2000 [@Tilman2001; @Tilman2006]. 
+Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
+water surface and the troposphere, measured at 13 locations at daily to weekly intervals 
+from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, 
+Ohio [@Holland1999; @Jarecke2016]. Each included location was measured on at least 50 dates. 
 Sampling locations spanned a hydrologic gradient from upland to seasonal wetland.
@@
 
-The environmental, species-trait, and community datasets, as we studied them, are henceforth called the "bivariate" datasets in the sense that they comprise two quantities measured at different locations or for different species.
-The phenology and population datasets are "multivariate" in that they comprise, for each taxon, measurements at multiple locations; copula structure can be studied for each pair of locations.
-Likewise the the ecosystem functioning (methane flux) dataset was  multivariate in this sense.
+The environmental, species-trait, and community datasets, as we studied them, are henceforth 
+called the "bivariate" datasets in the sense that they comprise two quantities measured at 
+different locations or for different species. The phenology and population datasets are 
+"multivariate" in that they comprise, for each taxon, measurements at multiple locations; 
+copula structure can be studied for each pair of locations. Likewise the the ecosystem 
+functioning (methane flux) dataset was  multivariate in this sense.
 
-# Concepts for \hyperref[Intro]{\textbf{(Q1)}}: Statistical tools for detecting copula structure \label{Methods}
+# Concepts for Q1: Statistical tools for detecting copula structure \label{Methods}
 
-Question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} was answered via statistical analysis of the
+Question Q1 from the Introduction was answered via statistical analysis of the
 data described above, using both model selection and nonparametric approaches.

### 2018-12-06 Paper.Rmd
include soilCN map locations

@@
 ```{r read_somedata, echo=F, results='hide'}
-d_soilCN<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
+d_soilCN<-readRDS("Data/dataForShymolinaCoord.rds")
+d_soilCN<-na.omit(d_soilCN)
 d_birdsBMR<-readRDS("./Data/BMR/BirdBodyMassesMetabolicRates/my_birdsBMR_data.RDS")
@@
 The database comprises measurements of soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
-locations across the coterminous United States (Fig. X for sampling locations).
+locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
 <!--***DAN: Shya, please add a sup mat figure showing the sampling locations and a link to it here in place of the X

### 2018-12-06 SupportingInformation.Rmd
made at start at revising the introduction
fixed minor error
include soilCN map locations

@@
 
+```{r cop_pedagog_fig1,echo=F,results="hide"}
+set.seed(seed)
+source("./FirstPedagogFig.R")
+```
+
 ```{r cop_pedag_fig,echo=F,results="hide"}
@@
 #fig.cap="Plot of the raw soil C and N data.\\label{fig_raw_cop_soilCN}(A)", fig.height=3, fig.width=3}
-d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
-d<-d[,c("SOCstock100","TSNstock100")]
+d<-readRDS("Data/dataForShymolinaCoord.rds")
+d<-d[,c("SOCstock100","TSNstock100","lat","long")]
+d<-na.omit(d)
 pdf("./Results/SoilCNraw.pdf")
@@
 pdf("./Results/SoilCNcop.pdf")
-v_CN<-getcopula(d=d,rankon=T,ploton=T) 
-dev.off()
-```
+v_CN<-getcopula(d=d[,c(1:2)],rankon=T,ploton=T) 
+dev.off()
+```
+
+```{r plot_SoilCNmapUSA, echo=F, results='hide'}
+#---------map of soilCN location on USA------------
+library(maps)
+pdf("./Results/SoilCNmapUSA.pdf",height=4,width=8)
+map("usa")
+points(x = d$long, y = d$lat, col="red",pch=16,cex=0.5)
+dev.off()
+```
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=8 cm,height=8 cm]{./Results/SoilCNmapUSA.pdf} 
+\caption{Locations of central pedons with soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil surface) from 5907 sampling sites across the coterminous United States.
+\label{fig_SoilCNmapUSA}}
+\end{center}
+\end{figure}
 

### 2018-12-06 Paper.Rmd
mostly edits to Intro

@@
 
-All branches of ecology and related areas study relationships among environmental and biological 
-variables. However, commonly used correlation- and regression-based approaches to studying 
+Very generally, all branches of ecology and related areas study relationships among environmental and biological 
+variables. However, commonly used correlation- and regression-based 
+approaches to studying 
 such relationships are limited - it is well understood that common statistical
-approaches may provide only a very partial description of a relationship. 
+approaches provide only a very partial description of a relationship between variables. 
 For instance, datasets showing markedly different 
 relationships may nevertheless have the same correlation coefficient value:  
-Fig. \ref{fig_cop_pedag1}A-B show two such datasets. The panel A 
-(respectively, B) variables of Fig. \ref{fig_cop_pedag1} are principally 
-related in the left (respectively, right) portions of their distributions, a difference that may
-have ecological significance, as discussed below, but is not captured by standard
+Fig. \ref{fig_cop_pedag1}A-B shows two such datasets. The variables of panel A 
+(respectively, B) are principally 
+related in the left (respectively, right) portions of their distributions, an asymmetric pattern of dependence in the distribution tails that may
+have ecological significance, as discussed below, but is not captured by 
 correlation. Panels C and D 
@@
 
-Well-developed statistical tools do exist (REFS), and are applied widely in other fields 
-(REFS), that provide a complete description of the dependence between two or more variables. 
-The goal of this paper is to illustrate the potential for broader application of these
-tools in ecology and related fields. The tools are based on the statistical concept of the *copula*, which we here introduce conceptually.
-
-  a) then do it. some things to do include...
-  b) the rank-based approach for data (use bivariate data)
-  c) the inverse cdf approach for the copula of a bivariate distribution
-  d) statement that copulas are also multivariate, though we won't explore that here (or maybe leave that for Discussion?)
-  e) a one- or two-sentence description of sklars theorem
+Well-developed statistical tools do exist, and are 
+applied widely in other fields, that provide a demonstrably complete 
+description of the dependence 
+between two or more variables [@joe2014_dependence; @nelsen2006_copula]. 
+The goal of this paper is to illustrate the potential for 
+broader application of these
+tools in ecology and related fields. The tools are based on 
+the statistical concept of the *copula*, which we here 
+introduce. From a data perspective, given 
+a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
+Fig. \ref{fig_cop_pedag1}, panel A or B), the associated copula
+structure is revealed by the bivariate plot of $u_i$ versus $v_i$, where
+$u_i$ is the rank of $x_i$ in the set of $x_j$ ($j=1,\ldots,n$), divided by
+$n+1$; and $v_i$ is defined in the same way but using the $y_i$. See, 
+for instance,
+Fig. \ref{fig_cop_pedag1}E-F, which show the copula structure of 
+A-B, respectively. Mathematically, if $(X,Y)$ is a continuous bivariate random
+variable and the cummulative distribution functions (cdfs) of $X$ and $Y$ are 
+$F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the copula structure
+of $(X,Y)$ is revealed by the random variable $(F_X(X),F_Y(Y))$.
+This random variable takes values in the unit square, 
+$[0,1] \times [0,1]$, and has uniform marginal distributions. Densities
+corresponding to the copulas associated 
+with Fig. \ref{fig_cop_pedag1}E-F are in Fig.
+\ref{fig_cop_pedag1}G-H, respectively. Strictly speaking, 
+the copula is actually defined as the bivariate 
+cdf of $(F_X(X),F_Y(Y))$ rather than as this random variable itself, 
+but this makes little conceptual difference. See @nelsen2006_copula and @joe2014_dependence
+for mathematical details. An important theorem of Abe Sklar [@sklar1959_theorem] 
+provides the result that copulas give a unique and complete 
+representation of dependence
+between variable: the original joint distribution of $(X,Y)$ 
+can be reconstructed from the 
+copula and the marginal distributions. Copula theory is also well developed for
+multivariate distributions.
+
+One way we may expect a copula approach  
+to better reflect ecological relationships has to do with Liebig's law of the
+minimum. If, for instance, the growth of a plant depends on soil 
+nitrogen, N, as well
+as on other factors, data on soil nitrogen and resulting growth rates may be 
+expected to look more like Fig. \ref{fig_cop_pedag1}A or E than like 
+Fig. \ref{fig_cop_pedag1}B or F. Nitrogen 
+will control plant growth, producing a strong
+relationship, only when it is limiting, in the left portion of the N and 
+plant-growth-rate distributions. 
+Nitrogen will be less related or unrelated to growth rates in the right tails of 
+the distributions, as pictured in Fig. \ref{fig_cop_pedag1}A and E.
+
+A second reason we may expect a copula-based approach to benefit 
+ecological research is that such an approach may help illuminate 
+causal mechanisms by which variables are related. For instance, if two species
+X and Y are strong competitors, we may expect abundances of the two species
+across quadrats or through time to be negatively related, as in 
+Fig. \ref{fig_cop_pedag1}I-J. If X is plotted on the horizontal axis
+and Y is plotted on the vertical, then the plot in Fig. \ref{fig_cop_pedag1}I 
+suggests X is the dominant 
+competitor: when X is abundant, Y is necessarily rare because it is suppressed 
+by X; whereas when X is rare, Y
+may be abundant or may also be relatively rare due to 
+factors other than X. Alternatively, Fig. \ref{fig_cop_pedag1}J suggests that Y
+is the dominant competitor.
+
+A third reason we may expect a copula approach to benefit ecological
+research is that prior work has demonstrated important copula 
+stucture in the spatial dependence of environental 
+variables [@Serinaldi2008; @She2018; @Goswami2018;
+@li2013]. Such dependence tends to beget dependence across space
+in fluctuations of populations or other ecosystem variables 
+(this is called \emph{synchrony}) which are influenced by the environmental 
+variables. This is called the \emph{Moran effect}. 
+If copula structure is transmitted, 
+possibly in modified form, 
+from environmental to ecological variables, then we may expect important 
+copula structure to be a common feature of dependencies through time 
+between ecological variables measured in different locations. 
+One may expect copula structure in environmental variables, in the form of 
+accentuated tail-dependence, if
+intense meteorological events are also spatially widespread.
+We will explore this as well as other potential mechanisms for
+copula structure in ecological variables.
+
+A final \emph{a priori} reason we may expect copula structure to be important
+in ecology has to do with spatially aggregated or averaged quantities. 
+Many ecological variables of applied importance 
+depend on the spatial average or sum of local 
+quantities. For instance, regional methane of CO$_2$ 
+fluxes are the sum of local fluxes; 
+and the total economic value of a fishery is the sum of catches from 
+local areas. We will explore how details, captured by the copula, of the 
+nature of the dependence between local quantities can influence the spatial 
+mean or sum and how this may influence human concerns and 
+higher organizational levels in 
+ecology. Illustrating the idea here, prior work demonstrated that the overall 
+reliability of wind-generated electricity depends sensitively on details
+of the dependence between wind speeds in multiple wind sites [@li2013]. 
+Spatially aggregated ecological variables may show a similar dependence, with
+potentially substantial consequences. For instance, if populations of a 
+pest species in different locations are all mutually strongly
+related to each other in their right tails, then local outbreaks tend strongly
+to occur as part of regional epidemics. Similarly, if populations
+of a crop resource species are strongly related in their left tails, then 
+episodes of local scarcity tend to happen simultaneously, creating regional 
+shortages.
+
+<!--DAN: edit and add references to the following-->
+Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related fields and thus has potential scope to better capture the evolution of ecological time series data.
 
@@
 \includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
-\caption{(A-F, I, J) Diverse bivariate datasets, many of which have the same Pearson (P)
+\caption{(A-F, I, J) Bivariate datasets showing diverse relationships between variables; many of these datasets nevertheless have the same Pearson (P)
 or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
 obtained from A (respectively, B) via monotonic functions that transformed the standard-
-normal marginal distributions of A and B to gamma (C, D) or uniform (E, F) marginals. 
-Details in Appendix X.\label{fig_cop_pedag1}}
+normal marginal distributions of A and B to gamma (C, D) or 
+uniform (E, F) marginals (details in Appendix X). (G, H) Contours 
+of log-scale density functions of Clayton (G) and survival-Clayton 
+(H) copulas with parameter 5 (see Methods). (I,J) Reflections
+of E, F, respectively, about a vertical axis. Subscripts h and v 
+stand for "horizontal" and "vertical". \label{fig_cop_pedag1}}
 \end{center}
@@
 
-Identifying and quantifying dependencies is the core of modelling in many applications, frequently found in biology and elsewhere.
-For many years, measuring and modeling dependencies has focused on the
-Pearson correlation because of its ease in use.
-But correlation can provide little information about the underlying dependence structure in the case of asymmetric dependence (i.e., varying strength of association between two variables depending on their values) because it is a measure of linear association between two variables.
-It may be better to use rank-based correlation measures such as Kendall’s tau and Spearman’s rho instead of Pearson correlation when the marginal distribution of the variables are highly non-Gaussian.
-The copula [@joe2014_dependence; @nelsen2007_copula], a rank-based dependence measurement between two variables, transforms non-uniformly distributed marginals in original variables to uniformly distributed marginals when the joint distribution of ranked data is plotted.
-For example, in the first panel of Fig.
-\ref{fig_cop_pedag}, we showed the joint distribution of raw data (left) and copula data (right), where the marginals of raw data variables (x and y) were gamma distributed but in the copula plot those marginals had uniform distribution.
-The copula plot was obtained by scaling individual ranked variables by a factor $(n+1)$ where $n$ was the number of sampling points, so that the copula was confined within a unit box.
-From now, we will call this "normalized ranking".
 Using a "copula-based approach" as a dependence measure reflects more knowledge about the dependence structure than summarizing in just a single number, the linear correlation coefficient.
@@
 <!--A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulae to be selected from.--> Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), described in the \nameref{Methods}.
-
-Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related fields and thus has potential scope to better capture the evolution of ecological time series data.
 
@@
 AIC and BIC values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection].
-Descriptions of these copula families can be found in introductory texts on copulae [@nelsen2007_copula ; @joe2014_dependence]  and in `VineCopula` documentation.
+Descriptions of these copula families can be found in introductory texts on copulae [@nelsen2006_copula ; @joe2014_dependence]  and in `VineCopula` documentation.
 The lower- and upper-tail dependence of the best-fitting member of each family was also provided by `BiCopEst`.
@@
 are related to each other in the left and right tails, respectively, of their 
-distributions [@nelsen2007_copula, @joe2014_dependence]. 
+distributions [@nelsen2006_copula, @joe2014_dependence]. 
 Our nonparametric study of tail dependence uses three
@@
 than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
-This holds, for instance, for the Clayton copula [@nelsen2007_copula].
+This holds, for instance, for the Clayton copula [@nelsen2006_copula].
 The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
@@
 
-To test environmental effect on population copula from the model \ref{eq.AR1}, our approach is of two-fold : First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2007_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
+To test environmental effect on population copula from the model \ref{eq.AR1}, our approach is of two-fold : First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2006_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
 Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.

### 2018-12-06 SupportingInformation.Rmd
mostly edits to Intro

@@
 i.e., strengths of dependence in the lower and upper tails are the same, for all copulas
-in these families [@nelsen2007_copula]. Clayton copulas are known to have stronger lower- than 
+in these families [@nelsen2006_copula]. Clayton copulas are known to have stronger lower- than 
 upper-tail dependence. And survival Clayton copulas, being rotations of Clayton copulas by 
-$180^\circ$, have stronger upper- than lower tail dependence [@nelsen2007_copula].
+$180^\circ$, have stronger upper- than lower tail dependence [@nelsen2006_copula].
 Using the `iRho` function from the `copula` package, for each of the four families normal, 

### 2018-12-07 Paper.Rmd
minor edit

@@
 d_mammalsBMR<-readRDS("./Data/BMR/MammalianBodyMassesMetabolicRates/my_mammalsBMR_data.RDS")
+d_cc<-readRDS("./Results/Ceder_creek_results/cop_HB_all_yr.RDS")
 ```
@@
 locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
-<!--***DAN: Shya, please add a sup mat figure showing the sampling locations and a link to it here in place of the X
-***Shya: once Prof. Loecke give me the lattitude and longitudes, I will do that.-->
 Measurments were from the central pedon of a soil core, to 1 m depth.
@@
 of @Sheppard2017; see Appendix \ref{SM-Data_details} for details.
-Community-level data were plant aboveground biomass and Shannon's diversity index for X 
-<!--***DAN: Shya, please fill in X with an autolink-->
+Community-level data were plant aboveground biomass and Shannon's diversity index for `r nrow(d_cc[[5]])` 
+<!--***DAN: Shya, please fill in X with an autolink
+***Shya: I mention only number of common plots taken from both biomass and percent cover datasets.-->
 plots, each 9m by 9m, maintained at Cedar Creek Ecosystem Science Reserve and sampled in 

### 2018-12-07 Paper.Rmd
most of the way done with the Intro for this round of edits at least

--- 2018-12-06 Paper.Rmd
@@
 ---
-author: "Shyamolina Ghosh, Lawrence W Sheppard, Mark T. Holder, Terrance E. Loecke, Daniel C. Reuman"
+title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
+author: "Shyamolina Ghosh, Lawrence W Sheppard, Mark T. Holder, Terrance E. Loecke, P. Chris Reid, Daniel C. Reuman"
 date: "February 18, 2018"
@@
 Loecke: Environmental Studies Program and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA
+
+Reid:
 
@@
 For instance, datasets showing markedly different 
-relationships may nevertheless have the same correlation coefficient value:  
+relationships may nevertheless have the same correlation coefficient value;  
 Fig. \ref{fig_cop_pedag1}A-B shows two such datasets. The variables of panel A 
@@
 studying relationships between variables, other commonly used metrics also 
-provide partial descriptions only.
+provide partial information only.
 
@@
 the statistical concept of the *copula*, which we here 
-introduce. From a data perspective, given 
+introduce. Copula approaches separate the information contained in a bivariate 
+random variable into two distinct parts: the information contained in the 
+marginal distributions (which says nothing about the dependence between the 
+variables), and the remaining information (which is solely about
+the dependence between the variables). From a data perspective, given 
 a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
-Fig. \ref{fig_cop_pedag1}, panel A or B), the associated copula
-structure is revealed by the bivariate plot of $u_i$ versus $v_i$, where
+Fig. \ref{fig_cop_pedag1}, panel A or B), the associated dependence
+or copula structure is revealed by the bivariate plot of $u_i$ versus $v_i$, where
 $u_i$ is the rank of $x_i$ in the set of $x_j$ ($j=1,\ldots,n$), divided by
@@
 Fig. \ref{fig_cop_pedag1}E-F, which show the copula structure of 
-A-B, respectively. Mathematically, if $(X,Y)$ is a continuous bivariate random
+A-B, respectively. Ranking serves to uniformize the 
+marginal distributions of the underlying random variables, isolating the 
+depedence structure only.
+Mathematically, if $(X,Y)$ is a continuous bivariate random
 variable and the cummulative distribution functions (cdfs) of $X$ and $Y$ are 
-$F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the copula structure
+$F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the dependence or copula structure
 of $(X,Y)$ is revealed by the random variable $(F_X(X),F_Y(Y))$.
@@
 but this makes little conceptual difference. See @nelsen2006_copula and @joe2014_dependence
-for mathematical details. An important theorem of Abe Sklar [@sklar1959_theorem] 
-provides the result that copulas give a unique and complete 
+for mathematical details. An important theorem of A. Sklar [@sklar1959_theorem] 
+states that copulas give a unique and complete 
 representation of dependence
-between variable: the original joint distribution of $(X,Y)$ 
-can be reconstructed from the 
-copula and the marginal distributions. Copula theory is also well developed for
+between variables, in the sense that the joint distribution of $(X,Y)$ 
+can be reconstructed completely from the 
+copula and the marginal distributions. 
+Copula theory is also well developed for
 multivariate distributions.
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
+\caption{(A-F, I, J) Bivariate datasets showing diverse relationships between variables; many of these datasets nevertheless have the same Pearson (P)
+or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
+obtained from A (respectively, B) via monotonic functions that transformed the standard-
+normal marginal distributions of A and B to gamma (C, D) or 
+uniform (E, F) marginals (details in Appendix X). (G, H) Contours 
+of log-scale density functions of Clayton (G) and survival-Clayton 
+(H) copulas with parameter 5 (see Methods). (I,J) Reflections
+of E, F, respectively, about a vertical axis. Subscripts h and v 
+stand for "horizontal" and "vertical". \label{fig_cop_pedag1}}
+\end{center}
+\end{figure}
+<!--***DAN: Dan to go back and add an appendix with details as it says. Make sure
+to mention the shape and scale params of the gamma marginals in C, D were 2-->
+
+A principal benefit of a copula approach to ecological data is 
+that such an approach can sensitively model and detect
+tail dependence and asymmetries of tail dependence. 
+Tail dependence is association between extreme values of two variables. 
+If smaller values of two positively associated variables are more 
+strongly associated than are larger values, the variables are said to
+have stronger lower- or left-tail dependence than right- or upper-tail dependence;
+and \emph{vice versa} if larger values are more strongly associated than smaller
+ones. Datasets of the same corelation can have a range of tail dependence 
+characteristics (Fig. \ref{fig_cop_pedag}), which can be quantified (Methods).
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
+\caption{(A-C) Data generated using the BB1 copula (Methods) with three different 
+sets of parameters chosen to produce a variety of 
+lower-tail (LT) and upper-tail (UT) dependence strengths (Methods), but the same 
+Kendall correlation (to two digits) in all cases. (D-F) Similar to A-C, but
+with stronger correlation. Asymmetry of tail dependence (LT-UT) is the same 
+within each column of panels.\label{fig_cop_pedag}}
+\end{center}
+\end{figure}
+<!--***DAN: Shya, please reduce it to three panels in each row (the middle one and the most extreme ones), while keeping the panels the same size, and label the panels A-F.
+Please also get rid of the title over each panel, and express the UT and LT values to two digits, and make the font slightly bigger for those as well as for the axis labels and numbers.
+Please also remove the
+color - we don't need it and it would cost a lot of money.-->
 
@@
 Nitrogen will be less related or unrelated to growth rates in the right tails of 
-the distributions, as pictured in Fig. \ref{fig_cop_pedag1}A and E.
+the distributions, as pictured in Fig. \ref{fig_cop_pedag1}A and E. The portion
+of the distributions that dominante the association is visible in the copula
+approach, but may not be reflected by correlation and other standard techniques.
 
@@
 may be abundant or may also be relatively rare due to 
-factors other than X. Alternatively, Fig. \ref{fig_cop_pedag1}J suggests that Y
-is the dominant competitor.
+limiting factors other than X. 
+Alternatively, Fig. \ref{fig_cop_pedag1}J suggests that Y
+is the dominant competitor. The copula approach again reveals which portions of the
+distributions are principally responsible for the association between variables.
 
@@
 variables [@Serinaldi2008; @She2018; @Goswami2018;
-@li2013]. Such dependence tends to beget dependence across space
+@li2013]. One may expect copula structure in environmental variables 
+to take the form of strong tail dependence if
+intense meteorological events are also widespread in their effects (as seems 
+frequently to be the case).
+Spatial dependence in an environmental variable 
+tends to beget dependence across space
 in fluctuations of populations or other ecosystem variables 
-(this is called \emph{synchrony}) which are influenced by the environmental 
-variables. This is called the \emph{Moran effect}. 
+(this is called \emph{spatial synchrony}) which are influenced by the environmental 
+variable. This is called the \emph{Moran effect}. 
 If copula structure is transmitted, 
@@
 from environmental to ecological variables, then we may expect important 
-copula structure to be a common feature of dependencies through time 
-between ecological variables measured in different locations. 
-One may expect copula structure in environmental variables, in the form of 
-accentuated tail-dependence, if
-intense meteorological events are also spatially widespread.
-We will explore this as well as other potential mechanisms for
-copula structure in ecological variables.
-
-A final \emph{a priori} reason we may expect copula structure to be important
+copula structure to be a common feature of spatial synchrony.
+Synchrony is a phenomenon of great interest in 
+ecology [@Liebhold2004; @walteretal2017].
+
+A fourth reason we may expect copula structure to be important
 in ecology has to do with spatially aggregated or averaged quantities. 
@@
 depend on the spatial average or sum of local 
-quantities. For instance, regional methane of CO$_2$ 
+quantities. For instance, regional methane and CO$_2$ 
 fluxes are the sum of local fluxes; 
 and the total economic value of a fishery is the sum of catches from 
-local areas. We will explore how details, captured by the copula, of the 
+local areas. We will explore how details, captured by copula methods, of the 
 nature of the dependence between local quantities can influence the spatial 
-mean or sum and how this may influence human concerns and 
+mean or sum, and how this may influence human concerns and 
 higher organizational levels in 
@@
 potentially substantial consequences. For instance, if populations of a 
-pest species in different locations are all mutually strongly
-related to each other in their right tails, then local outbreaks tend strongly
-to occur as part of regional epidemics. Similarly, if populations
-of a crop resource species are strongly related in their left tails, then 
-episodes of local scarcity tend to happen simultaneously, creating regional 
-shortages.
-
-<!--DAN: edit and add references to the following-->
+pest species in different locations are all positively asscoiated and also strongly
+related to each other in their right tails, then local outbreaks tend 
+to occur together, creating regional epidemics. If populations
+of a crop or other resource species are strongly 
+related in their left tails, then 
+episodes of local scarcity occur together, creating regional 
+shortages. On the other hand left-tail dependence in the pest species 
+and right-tail dependence in the crop, even if overall correlation were 
+unchanged, would have much more benign effects.
+
+<!--DAN: made it to here 2018 12 06
+edit and add references to the following-->
 Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related fields and thus has potential scope to better capture the evolution of ecological time series data.
 
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
-\caption{(A-F, I, J) Bivariate datasets showing diverse relationships between variables; many of these datasets nevertheless have the same Pearson (P)
-or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
-obtained from A (respectively, B) via monotonic functions that transformed the standard-
-normal marginal distributions of A and B to gamma (C, D) or 
-uniform (E, F) marginals (details in Appendix X). (G, H) Contours 
-of log-scale density functions of Clayton (G) and survival-Clayton 
-(H) copulas with parameter 5 (see Methods). (I,J) Reflections
-of E, F, respectively, about a vertical axis. Subscripts h and v 
-stand for "horizontal" and "vertical". \label{fig_cop_pedag1}}
-\end{center}
-\end{figure}
-<!--***DAN: Dan to go back and add an appendix with details as it says. Make sure
-to mention the shape and scale params of the gamma marginals in C, D were 2-->
-
-Using a "copula-based approach" as a dependence measure reflects more knowledge about the dependence structure than summarizing in just a single number, the linear correlation coefficient.
-It tells us about the tail-dependence of the variables, i.e., the association between extreme values of two variables.
-If smaller values of each variable are more correlated to each other, then the corresponding copula structure is said to have "lower tail-dependence" as shown in the first panel of Fig.
-\ref{fig_cop_pedag} copula plot (right).
-Similarly, an "upper tail-dependence" copula indicates stronger correlation between higher extreme variables in the raw data.
-Tail dependence occurs when the correlation between two variables increases as we get "further" in the tail (either or both) of the distribution.
-For example, in the middle panel of Fig.
-\ref{fig_cop_pedag}, different dependence structures, with same Kendall correlation $\tau = 0.5$, are shown which vary between dominantly upper tail dependent (lower tail - upper tail dependence being negative) to dominantly lower tail dependent (lower tail - upper tail dependence being positive) as we move along the row.
-Lower panel represents the similar scenario but with stronger correlation (Kendall's $\tau = 0.8$).
-So, the flexibility of copulae allows us to model more complex multivariate dependence structures by isolating marginal from joint distribution structures, with the capability of modeling a rich variety of tail behavior, such as tail dependence versus tail independence.<!--and reflection symmetry versus reflection asymmetry.--> To identify the kinds of copula structure found in real ecological data, we need to select candidate bivariate copulae from many existing parametric copula families.
-<!--A very flexible bivariate copula can itself become a welcome candidate in the pool of bivariate copulae to be selected from.--> Here, throughout our studies for different datasets we used a universal pool of 16 bivariate copula families (having exclusively lower, exclusively upper, both or neither tail dependence), described in the \nameref{Methods}.
-
-From the ecological perspective, when we find a dominantly lower (upper) tail dependence structure between two species abundance timeseries, it means both incidence are rare (abundant) synchronously.
-Again, if abundance of a species (say, species A) is associated with scarcity of another species (say, species B) then this may indicate causality, i.e.
-negative feedback of A on species B.
-If one variable is an environmental factor (like temperatute or rainfall) and another is an observable (like abundance of a species) then copula structure can also help explain the causal relationship between them.
-On a similar note, if a species' abundance time series from different sampling locations shows multivariate dependence structure, then we can analyze the "spatial data" using pairwise bivariate copulae and find if there is any tail dependence or not.
-If the copulae have asymmetric tail dependence, then it can be viewed as a new way to quantify "spatial synchrony" [@Liebhold2004; @walteretal2017], which is a celebrated phenomenon in population ecology.
  In this article, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets.
@@
 Finally, we conclude in the \nameref{Discussion} section.
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=4 cm,height=4 cm]{./Results/rawdata.pdf}
-\includegraphics[width=4 cm,height=4 cm]{./Results/copdata.pdf}\\
-\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
-\caption{First panel : raw data plot has non-uniform (gamma distributed) marginal whereas copula makes its marginal uniform and indicating lower tail dependence in joint distribution, Second and third panel : pedagogical figure showing BB1 copula with varying tail-dependence difference but of same Kendall's tau along each row and with same tail-dependence difference but of different Kendall's tau along each column.\label{fig_cop_pedag}}
-\end{center}
-\end{figure}
 

### 2018-12-07 Paper.Rmd
finished editing Intro for this round, some items for Shya to take a look at, please

@@
 a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
-Fig. \ref{fig_cop_pedag1}, panel A or B), the associated dependence
+Fig. \ref{fig_cop_pedag1}A-B), the associated dependence
 or copula structure is revealed by the bivariate plot of $u_i$ versus $v_i$, where
@@
 states that copulas give a unique and complete 
-representation of dependence
+representation of the dependence
 between variables, in the sense that the joint distribution of $(X,Y)$ 
@@
 strongly associated than are larger values, the variables are said to
-have stronger lower- or left-tail dependence than right- or upper-tail dependence;
+have stronger lower- or left-tail dependence than right- or upper-tail dependence
+(Fig. \ref{fig_cop_pedag1}A, E);
 and \emph{vice versa} if larger values are more strongly associated than smaller
-ones. Datasets of the same corelation can have a range of tail dependence 
+ones (Fig. \ref{fig_cop_pedag1}B, F). Datasets of the same 
+corelation can have a range of tail dependence 
 characteristics (Fig. \ref{fig_cop_pedag}), which can be quantified (Methods).
@@
 Please also remove the
-color - we don't need it and it would cost a lot of money.-->
+color - we don't need it and it would cost us a lot of money.-->
 
@@
 the distributions, as pictured in Fig. \ref{fig_cop_pedag1}A and E. The portion
-of the distributions that dominante the association is visible in the copula
+of the distributions that dominate the overall association is visible in the copula
 approach, but may not be reflected by correlation and other standard techniques.
@@
 X and Y are strong competitors, we may expect abundances of the two species
-across quadrats or through time to be negatively related, as in 
+across quadrats or through time within a single quadrat to be negatively related, as in 
 Fig. \ref{fig_cop_pedag1}I-J. If X is plotted on the horizontal axis
@@
 is the dominant competitor. The copula approach again reveals which portions of the
-distributions are principally responsible for the association between variables.
+distributions are principally responsible for the association between variables, and this 
+detail may help support or refute alternative mechanistic and causal hypotheses about 
+the relationship between the variables.
 
@@
 to take the form of strong tail dependence if
-intense meteorological events are also widespread in their effects (as seems 
-frequently to be the case).
+intense meteorological events are also widespread in their effects, as seems 
+frequently to be the case.
 Spatial dependence in an environmental variable 
@@
 Synchrony is a phenomenon of great interest in 
-ecology [@Liebhold2004; @walteretal2017].
+ecology [@Liebhold2004; @Sheppard2016; @walteretal2017].
 
@@
 reliability of wind-generated electricity depends sensitively on details
-of the dependence between wind speeds in multiple wind sites [@li2013]. 
+of the dependence between wind speeds at multiple generator sites [@li2013]. 
 Spatially aggregated ecological variables may show a similar dependence, with
 potentially substantial consequences. For instance, if populations of a 
-pest species in different locations are all positively asscoiated and also strongly
+pest species in different locations are all positively associated and also strongly
 related to each other in their right tails, then local outbreaks tend 
@@
 
-<!--DAN: made it to here 2018 12 06
-edit and add references to the following-->
-Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related fields and thus has potential scope to better capture the evolution of ecological time series data.
-
- In this article, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets.
-In addition to a copula based \nameref{Model_selection}, we also carried out a \nameref{NPA} with three different statistics to verify our model selection based results.
-
-We want to address the following questions.
-\textbf{(Q1)} Do datasets in ecology and related fields have non-Normal copula srtructure?
-Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric tail dependence?
-\textbf{(Q2)} If so, what are some possible causes of non-Normal copula structure?
-And \textbf{(Q3)} what are the consequences of non-Normal copula structure and tail dependence for our understanding of biology and for applications?
-In \nameref{Data}, we described briefly each of bivariate and multivariate datasets we used to answer these questions and we provide a detailed descriptions of the implemented methods in \nameref{Methods}.
-In the \nameref{Results} section, we summarized our findings for each dataset.
-The next two sections are focused on possible causes and consequences of non-Normal copula structure in ecology and related fields with applications shown in real data as well as some simulations.
-Finally, we conclude in the \nameref{Discussion} section.
-
-<!--# Material and methods \label{Material&Methods}-->
+We approached our overall goal of illustrating the potential of 
+copula approaches for understanding dependences in ecology by addressing the 
+folling specific questions. (Q1) Do datasets in ecology and related fields 
+have copula structure distinct from a Gaussian/normal distribution (here called 
+non-normal copula structure)? Do they show tail dependence distinct from that 
+of a normal distribution, and in particular do they show asymmetric tail 
+dependence? (Q2) If so, what are some possible causes/mechanisms of this 
+non-normal copula structure? And (Q3) what are the consequences of non-normal 
+copula structure and tail dependence for our understanding of biology and for 
+applications? To address Q1, we analyze several datasets, including 
+environmental, species-trait, phenological, population, community, and 
+ecosystem functioning data. We selected these datasets to span multiple 
+sub-fields and organizational levels of ecology, in an attempt to be 
+reasonably representative of the field as a whole. The second section of the 
+paper introduces the data, and the third and fourth sections use the data to 
+address Q1. The fifth and sixth sections, which address Q2, use simple models to 
+explore mechanisms. The seventh and eigth sections, which address Q3, use both
+the data section two as well as models to explore a variety of consequences
+of copula structure and tail dependence for applications and for our understanding
+of ecology. The ninth section is the Discussion. 
+Copula approaches have been used to great benefit in neuroscience [@onken2009], 
+bioinformatics 
+[@Kim2008], medical research [@Emura2016], direct study of environmental 
+variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], finance, and other fields. 
+To our knowledge, these approaches have been used seldom or never for ecological 
+research. We argue that the benefits to ecological research of a complete statistical 
+desription of dependence will be substantial, and we hope this paper stimulates the
+use of copula approaches across the discipline.
 

### 2018-12-08 Paper.Rmd
fixed pedagog fig2 as instructed

@@
 lower-tail (LT) and upper-tail (UT) dependence strengths (Methods), but the same 
-Kendall correlation (to two digits) in all cases. (D-F) Similar to A-C, but
+Kendall correlation (KC) (to two digits) in all cases. (D-F) Similar to A-C, but
 with stronger correlation. Asymmetry of tail dependence (LT-UT) is the same 
@@
 Please also remove the
-color - we don't need it and it would cost us a lot of money.-->
+color - we don't need it and it would cost us a lot of money.
+***Shya: Done.-->
 

### 2018-12-08 SupportingInformation.Rmd
fixed pedagog fig2 as instructed

@@
 
-pdf("./Results/copula_pedagog_fig_BB1.pdf",height=8,width=20)
-op<-par(mfrow=c(2,5),mar=c(3,3,3,3), mgp=c(1.5,0.5,0))
-
-desLTmUT_vector<-c(-0.4,-0.2,0,0.2,0.4)
-#parlist<-list(c(0.2,1.5),c(0.2,1.1),c(0.2,1.5),c(0.5,1.1),c(0.5,1.5))
-#for(i in 1:length(desLTmUT_vector)){
-#  print(i)
-#  iLmU<-desLTmUT_vector[i]
-#  call_optim<-optim(par=parlist[[i]],fn=f,destau=0.4,desLTmUT=iLmU,fam=7,method = "BFGS")
-#  call_optim$par
-#  c<-BiCopSim(N=500,family=7,par=call_optim$par[1],par2=call_optim$par[2])
-#  plot(c[,1],c[,2],col="blue",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
-#  mtext(paste0("tau = 0.4", ", LTmUT = ",iLmU,sep=""),side=3)
-#}
-
-for(i in desLTmUT_vector){
+pdf("./Results/copula_pedagog_fig_BB1.pdf",height=8,width=12)
+op<-par(mfrow=c(2,3),mar=c(8,8,1.5,3), mgp=c(5,2,0))
+
+desLTmUT_vector<-c(-0.4,0,0.4)
+
+for(i in c(1:length(desLTmUT_vector))){
   # print(i)
   destau<-0.5
-  call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=i,fam=7,method = "BFGS")
+  call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=desLTmUT_vector[i],fam=7,method = "BFGS")
   obj<-BiCop(family=7,par=call_optim$par[1],par2=call_optim$par[2])
   c<-BiCopSim(N=500,obj)
-  plot(c[,1],c[,2],col="blue",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
-  mtext(paste0("tau = ",round(obj$tau,5),", ( LT - UT ) = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
-  legend("topleft",paste0("UT=",round(obj$taildep$upper,5)),bty="n",cex=1.5,text.col="magenta")
-  legend("bottomright",paste0("LT=",round(obj$taildep$lower,5)),bty="n",cex=1.5,text.col="magenta")
-}
-
-for(i in desLTmUT_vector){
+  plot(c[,1],c[,2],col="grey",xlab="u",ylab="v",cex.lab=3.5,cex.axis=3.5)
+  legend("topleft",LETTERS[i],bty="n",cex=4,x.intersp=-0.5)
+  #mtext(paste0("( LT - UT ) = ",round((obj$taildep$lower - obj$taildep$upper),2),sep=""),side=3,cex=2)
+  legend("bottomright",c(paste0("KC= ", round(obj$tau,2)),
+                                    paste0("LT=",round(obj$taildep$lower,2)),
+                                    paste0("UT=",round(obj$taildep$upper,2))),bty="n",horiz = F,cex=2.5)
+  #legend("bottomright",paste0("LT=",round(obj$taildep$lower,2)),bty="n",cex=2)
+}
+
+for(i in c(1:length(desLTmUT_vector))){
  # print(i)
   destau<-0.8
- call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=i,fam=7,method = "BFGS")
+ call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=desLTmUT_vector[i],fam=7,method = "BFGS")
  obj<-BiCop(family=7,par=call_optim$par[1],par2=call_optim$par[2])
  c<-BiCopSim(N=500,obj)
- plot(c[,1],c[,2],col="red",xlab="u",ylab="v",cex.lab=2,cex.axis=1.5)
- mtext(paste0("tau = ",round(obj$tau,5),", ( LT - UT ) = ",round((obj$taildep$lower - obj$taildep$upper),5),sep=""),side=3)
- legend("topleft",paste0("UT=",round(obj$taildep$upper,5)),bty="n",cex=1.5,text.col="magenta")
- legend("bottomright",paste0("LT=",round(obj$taildep$lower,5)),bty="n",cex=1.5,text.col="magenta")
+ plot(c[,1],c[,2],col="grey",xlab="u",ylab="v",cex.lab=3.5,cex.axis=3.5)
+ legend("topleft",LETTERS[3+i],bty="n",cex=4,x.intersp=-0.5)
+ #mtext(paste0("( LT - UT ) = ",round((obj$taildep$lower - obj$taildep$upper),2),sep=""),side=3,cex=2)
+  legend("bottomright",c(paste0("KC= ", round(obj$tau,2)),
+                                    paste0("LT=",round(obj$taildep$lower,2)),
+                                    paste0("UT=",round(obj$taildep$upper,2))),bty="n",horiz = F,cex=2.5)
+ #legend("bottomright",paste0("LT=",round(obj$taildep$lower,2)),bty="n",cex=2)
 }

### 2018-12-13 Paper.Rmd
Third pedagogigical figure and working on Q1 sections

@@
 correlation. Panels C and D 
-were obtained from A and B through monotonic transformation (Appendix X for details),
+were obtained from A and B through monotonic transformation 
+(Appendix \ref{SM-Data_details} for details),
 <!--***DAN: Dan to go back and write the appendix-->
@@
 This random variable takes values in the unit square, 
-$[0,1] \times [0,1]$, and has uniform marginal distributions. Densities
-corresponding to the copulas associated 
+$[0,1] \times [0,1]$, and has uniform marginal distributions. Probability density 
+functions (pdfs) corresponding to the copulas associated 
 with Fig. \ref{fig_cop_pedag1}E-F are in Fig.
@@
 \includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
-\caption{(A-F, I, J) Bivariate datasets showing diverse relationships between variables; many of these datasets nevertheless have the same Pearson (P)
+\caption{(A-F, I, J) Bivariate datasets showing diverse relationships between 
+variables; many of these datasets nevertheless have the same Pearson (P)
 or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
@@
 normal marginal distributions of A and B to gamma (C, D) or 
-uniform (E, F) marginals (details in Appendix X). (G, H) Contours 
+uniform (E, F) marginals (details in Appendix \ref{SM-fig_cop_pedag1}). (G, H) Contours 
 of log-scale density functions of Clayton (G) and survival-Clayton 
-(H) copulas with parameter 5 (see Methods). (I,J) Reflections
+(H) copulas with parameter 5 (Appendix \ref{SM-copula_details}). (I,J) Reflections
 of E, F, respectively, about a vertical axis. Subscripts h and v 
@@
 \end{figure}
-<!--***DAN: Dan to go back and add an appendix with details as it says. Make sure
-to mention the shape and scale params of the gamma marginals in C, D were 2-->
+<!--***DAN: Dan to go back and add the appendices with details as it says. Make sure
+to mention in the first one 
+the shape and scale params of the gamma marginals in C, D were 2-->
 
@@
 \end{figure}
-<!--***DAN: Shya, please reduce it to three panels in each row (the middle one and the most extreme ones), while keeping the panels the same size, and label the panels A-F.
-Please also get rid of the title over each panel, and express the UT and LT values to two digits, and make the font slightly bigger for those as well as for the axis labels and numbers.
-Please also remove the
-color - we don't need it and it would cost us a lot of money.
-***Shya: Done.-->
+<!--***DAN: Shya, I like the grey color of your points on the above figure, 
+but I prefer the 
+solid plotting symbols and the size of plotting symbols used on the new 
+pedagogical figure that I made. Can you please change the code of both figures
+so we get grey, with the plotting symbols I used, on both plots? Likeiwse for the 
+new figure produced in ./ThirdPedagogFig.R, please. It will be
+good to use the same points, for consistency. Don't worry at this stage about
+making other figures, other than these three, consistent.-->
 
@@
 Community-level data were plant aboveground biomass and Shannon's diversity index for `r nrow(d_cc[[5]])` 
-<!--***DAN: Shya, please fill in X with an autolink
-***Shya: I mention only number of common plots taken from both biomass and percent cover datasets.-->
 plots, each 9m by 9m, maintained at Cedar Creek Ecosystem Science Reserve and sampled in 
@@
 
-# Concepts for Q1: Statistical tools for detecting copula structure \label{Methods}
+# Background concepts and methods for Q1: detecting copula structure \label{Methods}
+
+We addressed Q1 in part via a model selection procedure in which the normal copula
+and a variety of alternative copulas were fit to 
+our ecological datasets, and fits of the various copula models 
+were compared via the Akaike and Bayesian 
+Information Criteria (AIC and BIC). To make the results more easily interpretable
+to readers unfamiliar with copulas, we first introduce the normal copula and 
+some alternative copulas we used; and we operationally introduce measures of tail dependence
+and asymmetry of tail dependence. 
+Additional details are in Appendix \ref{SM-copula_details}. 
+Conceptually, a bivariate copula is the same as a probability density function 
+on the unit square with uniform maginals. The normal copula family is a one-parameter
+family of copulas, with parameter corresponding to the covariance of the related 
+bivariate normal distribtion and controlling the degree of dependence 
+(Fig. \ref{fig_cop_pedag3}A-B for an example).
+Left- and right-tail dependence are $0$ for a the normal copula, and this 
+serves as a reference for tail dependence in other copulas. The Clayton and 
+survival-Clayton families (Fig. \ref{fig_cop_pedag1}G-H) are two more one-parameter 
+families. The Clayton has positive lower-tail dependence ($0.87$ for
+the example of Fig. \ref{fig_cop_pedag1}G) and $0$ upper-tail dependence; the 
+survival-Clayton, which is a rotation of the Clayton by $180^\circ$, has 
+positive upper-tail dependence ($0.87$ in Fig. \ref{fig_cop_pedag1}H)
+and $0$ lower-tail dependence. For our model selection procedure (see below), we considered
+
+Families considered (with abbreviations here indicated) included the **N**ormal/Gaussian, **C**layton, **G**umbel, **F**rank, **J**oe, **BB1**, **BB6**, **BB7**, **BB8**, **S**urvival **C**layton , **S**urvival **G**umbel, **S**urvival **J**oe, **S**urvival **BB1**, **S**urvival **BB6**, **S**urvival **BB7** and **S**urvival **BB8** copulae.
+These copulae were chosen carefully so that they covered the whole range from exclusively lower (C, SG, SJ, SBB6, SBB8) or upper tail dependence (SC, G, J, BB6, BB8) to both tail dependence (BB1, SBB1, BB7, SBB7) in addition to two symmetric copulae (N, F).
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=3.5in]{./Results/PedagogFig3.pdf}
+\caption{Log-transformed pdfs (A, C) and samples from (B, D) example copulas.
+Shown are the normal copula with parameter $0.8$ and the Frank copula with 
+parameter $10$. See also Fig. \ref{fig_cop_pedag1}E-H for the Clayton and  
+survival-Clayton copulas with parameter 5, and Appendix. \label{fig_cop_pedag3}}
+\end{center}
+\end{figure}
 
@@
 
-# Results for \hyperref[Intro]{\textbf{(Q1)}}: Ecological datasets often have non-Normal copula structures, including asymmetric tail dependence \label{Results}
+# Results for Q1: Ecological datasets often have non-normal copula structure and asymmetric tail dependence \label{Results}
 <!--## question 1
@@
 <!--Question 2 analyses-->
-# Concepts and models for \hyperref[Intro]{\textbf{(Q2)}}: possible causes of non-Normal copula structure \label{Causes}
+# Concepts and models for Q2: possible causes of non-normal copula structure \label{Causes}
 
@@
 <!--Question 3 analyses-->
-# Concepts for \hyperref[Intro]{\textbf{(Q3)}}: Consequences of non-Normal copula structure and tail dependence for our understanding of biology \label{Consequences}
+# Concepts, methods and models for Q3: Consequences of non-normal copula structure and tail dependence for our understanding of biology \label{Consequences}
 

### 2018-12-14 SupportingInformation.Rmd
Third pedagogigical figure and working on Q1 sections
made pedagog figs consistent

@@
 
-```{r cop_pedagog_fig1,echo=F,results="hide"}
+```{r cop_pedagog_fig,echo=F,results="hide"}
 set.seed(seed)
 source("./FirstPedagogFig.R")
-```
-
-```{r cop_pedag_fig,echo=F,results="hide"}
-set.seed(seed)
-library(VineCopula)
-f<-function(par,destau,desLTmUT,fam){
-  cop<-BiCop(family=fam,par=par[1],par2=par[2])
-  tau<-cop$tau 
-  LT<-cop$taildep$lower
-  UT<-cop$taildep$upper
-  LTmUT<-LT-UT
-  z<-(tau-destau)^2
-  z<-z+(LTmUT-desLTmUT)^2
-  return(z)
-}
-
-pdf("./Results/copula_pedagog_fig_BB1.pdf",height=8,width=12)
-op<-par(mfrow=c(2,3),mar=c(8,8,1.5,3), mgp=c(5,2,0))
-
-desLTmUT_vector<-c(-0.4,0,0.4)
-
-for(i in c(1:length(desLTmUT_vector))){
-  # print(i)
-  destau<-0.5
-  call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=desLTmUT_vector[i],fam=7,method = "BFGS")
-  obj<-BiCop(family=7,par=call_optim$par[1],par2=call_optim$par[2])
-  c<-BiCopSim(N=500,obj)
-  plot(c[,1],c[,2],col="grey",xlab="u",ylab="v",cex.lab=3.5,cex.axis=3.5)
-  legend("topleft",LETTERS[i],bty="n",cex=4,x.intersp=-0.5)
-  #mtext(paste0("( LT - UT ) = ",round((obj$taildep$lower - obj$taildep$upper),2),sep=""),side=3,cex=2)
-  legend("bottomright",c(paste0("KC= ", round(obj$tau,2)),
-                                    paste0("LT=",round(obj$taildep$lower,2)),
-                                    paste0("UT=",round(obj$taildep$upper,2))),bty="n",horiz = F,cex=2.5)
-  #legend("bottomright",paste0("LT=",round(obj$taildep$lower,2)),bty="n",cex=2)
-}
-
-for(i in c(1:length(desLTmUT_vector))){
- # print(i)
-  destau<-0.8
- call_optim<-optim(par=c(0.2,1.5),fn=f,destau=destau,desLTmUT=desLTmUT_vector[i],fam=7,method = "BFGS")
- obj<-BiCop(family=7,par=call_optim$par[1],par2=call_optim$par[2])
- c<-BiCopSim(N=500,obj)
- plot(c[,1],c[,2],col="grey",xlab="u",ylab="v",cex.lab=3.5,cex.axis=3.5)
- legend("topleft",LETTERS[3+i],bty="n",cex=4,x.intersp=-0.5)
- #mtext(paste0("( LT - UT ) = ",round((obj$taildep$lower - obj$taildep$upper),2),sep=""),side=3,cex=2)
-  legend("bottomright",c(paste0("KC= ", round(obj$tau,2)),
-                                    paste0("LT=",round(obj$taildep$lower,2)),
-                                    paste0("UT=",round(obj$taildep$upper,2))),bty="n",horiz = F,cex=2.5)
- #legend("bottomright",paste0("LT=",round(obj$taildep$lower,2)),bty="n",cex=2)
-}
-
-par(op)
-dev.off()
+source("./SecondPedagogFig.R")
+source("./ThirdPedagogFig.R")
 ```
@@
 additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.
+
+# Transformations used in Fig. \ref{MT-fig_cop_pedag1}
+
+# Background on the copulas we fitted with data \label{copula_details}
 

### 2018-12-14 FirstPedagogFig.R

@@ -51,8 +51,7 @@ par(fig=c((ymarg_wd+numsp)/tot_wd,
           (xmarg_ht+5*pan_ht+4*gap+5*numsp)/tot_ht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
 rg<-max(abs(dcc5_nm))
-plot(dcc5_nm[,1],dcc5_nm[,2],type='p',pch=pchval,cex=cexvalpts,
-     xlim=c(-rg,rg),ylim=c(-rg,rg))
+plot(dcc5_nm[,1],dcc5_nm[,2],type='p',pch=pchval,cex=cexvalpts,xlim=c(-rg,rg),ylim=c(-rg,rg),col="grey")
 mtext(expression(z[v]),2,line=1.15)
 text(-rg,rg,"A",cex=cexvalpl,adj=c(0,1))
 text(rg,-rg+.25*rg,paste0("P=",round(cor(dcc5_nm[,1],dcc5_nm[,2]),2)),
@@ -67,8 +66,7 @@ par(fig=c((ymarg_wd+2*numsp+pan_wd+gap)/tot_wd,
           (xmarg_ht+5*pan_ht+4*gap+5*numsp)/tot_ht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
 rg<-max(abs(dsc5_nm))
-plot(dsc5_nm[,1],dsc5_nm[,2],type='p',pch=pchval,cex=cexvalpts,
-     xlim=c(-rg,rg),ylim=c(-rg,rg))
+plot(dsc5_nm[,1],dsc5_nm[,2],type='p',pch=pchval,cex=cexvalpts,xlim=c(-rg,rg),ylim=c(-rg,rg),col="grey")
 text(-rg,rg,"B",cex=cexvalpl,adj=c(0,1))
 text(rg,-rg+.25*rg,paste0("P=",round(cor(dsc5_nm[,1],dsc5_nm[,2]),2)),
      adj=c(1,0),cex=cexvaltxt)
@@ -83,8 +81,7 @@ par(fig=c((ymarg_wd+numsp)/tot_wd,
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
 rg1<-min(dcc5_gm,dsc5_gm)
 rg2<-max(dcc5_gm,dsc5_gm)
[…]

### 2018-12-14 Paper.Rmd
mostly working on code for a series of pedagog figs for the sup mat showing the different copulas we use

@@
 
-# Background concepts and methods for Q1: detecting copula structure \label{Methods}
-
-We addressed Q1 in part via a model selection procedure in which the normal copula
+# Background, concepts and methods for Q1: detecting copula structure in data \label{Methods}
+
+We addressed Q1 partly via a model selection procedure in which the normal copula
 and a variety of alternative copulas were fit to 
@@
 were compared via the Akaike and Bayesian 
-Information Criteria (AIC and BIC). To make the results more easily interpretable
+Information Criteria (AIC and BIC). To make the results interpretable
 to readers unfamiliar with copulas, we first introduce the normal copula and 
-some alternative copulas we used; and we operationally introduce measures of tail dependence
+some other copulas we used; and we operationally introduce measures of tail dependence
 and asymmetry of tail dependence. 
@@
 family of copulas, with parameter corresponding to the covariance of the related 
-bivariate normal distribtion and controlling the degree of dependence 
+bivariate normal distribtion, and controlling the degree of dependence 
 (Fig. \ref{fig_cop_pedag3}A-B for an example).
@@
 parameter $10$. See also Fig. \ref{fig_cop_pedag1}E-H for the Clayton and  
-survival-Clayton copulas with parameter 5, and Appendix. \label{fig_cop_pedag3}}
+survival-Clayton copulas with parameter 5, and Appendix \ref{SM-copula_details}
+for additional details. \label{fig_cop_pedag3}}
 \end{center}

### 2018-12-14 SupportingInformation.Rmd
mostly working on code for a series of pedagog figs for the sup mat showing the different copulas we use

--- 2018-12-08 SupportingInformation.Rmd
@@
 source("./FirstPedagogFig.R")
+source("./ThirdPedagogFig.R")
 ```
@@
 additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.
+
+# Transformations used in Fig. \ref{MT-fig_cop_pedag1}
+
+# Background on the copulas we fitted with data \label{copula_details}
+
+<!--****DAN: Need a 10-panel fig for each one-parameter family showing samples and 
+pdfs for 5 values of the parameter. May also need something for the two-parameter
+families, but not sure yet how to handle it. After doing the univariate families, then
+maybe it is enough to show just the pdf, in which case it can be done in a 
+5-by-5 array of panels for the 2-parameter families. That seems like it would work.-->
 

### 2018-12-14 MakeCopulaPedagogFigs.R

@@ -1,0 +1,280 @@
+#' Make a pedagogical plot for a univariate copula
+#'
+#' @param coplist A list of 5 copula objects from the copula package
+#' @param numpts Some panels show samples of this many points from the copulas
+#' @param corvals A vector of five correlations for the five panels, which are
+#' displayed on the plot
+#' @param cortype "S", "P" or "K" for Spearman, Pearson, Kendall
+#' @param thetavals A vector of five values of the copula parameter, again for display
+#' @param tot_wd Total width of the figure generated, in inches
+#' @param filename A filename, without extension, for saving in the working directory. A pdf 
+#' is generated and saved.
+#' 
+#' @note For my own use at present, does little to no error catching
+#' 
+#' @author Daniel Reuman \email{reuman@@ku.edu}
+
+muvcpp<-function(coplist,numpts,corvals,cortype,thetavals,tot_wd,filename)
+{
+  if (length(coplist)!=5)
+  {
+    stop("Error in muvcpp: only works with 5 copulas")
+  }
+
+  #***plot dimensions
+  xmarg_ht<-.25
+  ymarg_wd<-.25
+  numsp<-.2
+  gap<-0.1
+  labgap<-.225
+  titgap<-.15
+  pan_wd<-(tot_wd-ymarg_wd-numsp-5*gap)/5
+  if (pan_wd<=0){stop("Error in muvcpp: need a larger tot_wd")}
+  pan_ht<-pan_wd
+  tot_ht<-xmarg_ht+numsp+2*pan_ht+2*labgap+2*gap+titgap
+  pchval<-20
+  cexvalpts<-0.25
+  cexvaltxt<-0.75
+  cexvalpl<-1.5
+  pdf(file=paste0(filename,".pdf"),width=tot_wd,height=tot_ht)
+  
+  x<-seq(from=0,to=1,length.out=101)
+  y<-x
+  dcarg<-matrix(c(rep(x,times=length(y)),rep(y,each=length(x))),length(x)*length(y),2)
+  
[…]

### 2018-12-14 SupportingInformation.Rmd

--- 2018-12-14 SupportingInformation.Rmd
@@
 # Background on the copulas we fitted with data \label{copula_details}
+
+<!--****DAN: Need a 10-panel fig for each one-parameter family showing samples and 
+pdfs for 5 values of the parameter. May also need something for the two-parameter
+families, but not sure yet how to handle it. After doing the univariate families, then
+maybe it is enough to show just the pdf, in which case it can be done in a 
+5-by-5 array of panels for the 2-parameter families. That seems like it would work.-->
 

### 2018-12-16 Paper.Rmd
furter editing

@@
 author: "Shyamolina Ghosh, Lawrence W Sheppard, Mark T. Holder, Terrance E. Loecke, P. Chris Reid, Daniel C. Reuman"
-date: "February 18, 2018"
+date: "December 16, 2018"
 geometry: "left=1.5cm,right=1.5cm,top=2.5cm,bottom=2.8cm"
@@
 ---
-
-<!--**DAN: right now we sometimes use the terminology "normal copula" and sometimes 
-"Gaussin copula". Let's choose one and be consistent across the whole ms, including 
-main text and supp mat. I am leaning toward "Gaussin copula" but could be convinced 
-otherwise, it is just the consistency that I want to be sure we get to.-->
-
-<!--**SG: I think normal copula would be better as "N" is the abbreviation for that but for "G" means gumbel not gaussian.
-I made all "Gaussian" as "Normal" copula in both main and supmat.-->
 
@@
 and asymmetry of tail dependence. 
-Additional details are in Appendix \ref{SM-copula_details}. 
 Conceptually, a bivariate copula is the same as a probability density function 
@@
 (Fig. \ref{fig_cop_pedag3}A-B for an example).
-Left- and right-tail dependence are $0$ for a the normal copula, and this 
-serves as a reference for tail dependence in other copulas. The Clayton and 
+Left- and right-tail dependence are $0$ for the normal copula, and this 
+serves as a benchmark for tail dependence in other copulas - positive values of our
+tail dependence measure correspond to dependence stronger than that of a normal
+copula. The Clayton and 
 survival-Clayton families (Fig. \ref{fig_cop_pedag1}G-H) are two more one-parameter 
@@
 positive upper-tail dependence ($0.87$ in Fig. \ref{fig_cop_pedag1}H)
-and $0$ lower-tail dependence. For our model selection procedure (see below), we considered
-
-Families considered (with abbreviations here indicated) included the **N**ormal/Gaussian, **C**layton, **G**umbel, **F**rank, **J**oe, **BB1**, **BB6**, **BB7**, **BB8**, **S**urvival **C**layton , **S**urvival **G**umbel, **S**urvival **J**oe, **S**urvival **BB1**, **S**urvival **BB6**, **S**urvival **BB7** and **S**urvival **BB8** copulae.
-These copulae were chosen carefully so that they covered the whole range from exclusively lower (C, SG, SJ, SBB6, SBB8) or upper tail dependence (SC, G, J, BB6, BB8) to both tail dependence (BB1, SBB1, BB7, SBB7) in addition to two symmetric copulae (N, F).
+and $0$ lower-tail dependence. For our model selection procedure (see below), we used 
+several one- and two-parameter families of copulas exhibiting a wide variety of 
+positive lower-tail dependence values 
+(Clayton, survival-Gumbel, survival-Joe, survival-BB6, henceforth
+abbreviated, respectively, C, SG, SJ, SBB6) and upper-tail dependence values
+(survival-Clayton, Gumbel, Joe, BB6, henceforth abbreviated, respectively, 
+SC, G, J, BB6). We also considered families that could exhibit positive upper- or lower-tail
+dependence, or both, depending on parameters (BB1, survival-BB1, BB7, survival-BB7, henceforth 
+abbreviated, respectively, BB1, SBB1, BB7, SBB7). 
+If LT and UT are used to denote 
+our two measures of tail dependence, their difference is a measure of asymmetry of 
+tail dependence. We also used two families that were asymmetric in their non-tail parts, 
+but that had zero tail dependence (BB8, survival-BB8, abbreviated BB8 and SBB8).
+In addition to the normal copula, we used another symmetric copula,
+the Frank copula (abbreviated F; Fig. \ref{fig_cop_pedag3}C-D). 
+Copulas are implemented in the `copula` and `VineCopula` packages for the 
+R statistical computing language.
+Additional details are in Appendix \ref{SM-copula_details} and in introductory texts 
+on copulas [@nelsen2006_copula ; @joe2014_dependence] and in `VineCopula` documentation.
+Visualizations of all the copulas we used are in Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}.
 
@@
 
-Question Q1 from the Introduction was answered via statistical analysis of the
-data described above, using both model selection and nonparametric approaches.
-
-## Model selection approach \label{Model_selection}
-
-A model-selection approach based on likelihood and the Akaike Information Criterion (AIC) <!--and Bayesian Information Criterion (BIC)--> was used to help answer \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
-For bivariate datasets
-$(x_i,y_i)$ for $i=1,...,n$ such as the soil C and N data, model selection involved several steps, 
-most implemented using tools from the `VineCopula` package. <!--Code combining these tools is in the
-[`OurBiCopSelect`](https://github.com/sghosh89/BIVAN/blob/master/OurBiCopSelect.R) function and its dependencies in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
-to be added in suppmat-->First, we produced ranks
+Question Q1 was answered principally via analysis of the
+data using a model selection approach. For bivariate datasets 
+$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
+First, as in the Introduction, we produced ranks 
 $(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
 $\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
-we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; further work was with these "normalized ranked variables" $u_i$ 
-and $v_i$. Ranks $\tilde{u}_i$ and $\tilde{v}_i$ were $1$, for the smallest element, up to $n$, for the largest. Second, we did a test of the independence of the $u_i$ and $v_i$. Subsequent model 
+we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; subsequent analysis used these 
+*normalized ranks* $u_i$ and $v_i$. Ranks were $1$ for the smallest element in a set. 
+Second, we did a test of the independence of the $u_i$ and $v_i$ (implemented in X in the Y package). 
+<!--***DAN: X and Y to be flled in, can't remember right now what they are and don't want to stop
+to look it up-->
+Subsequent model 
 selection algorithms are ineffective if data could not be distinguished from independent data 
 since most or all bivariate copula families considered include the independent copula in the 
-family or as a boundary case. <!--The independence test was as implemented in the `BiCopIndTest`
-function in `VineCopula`, the documentation of which reports the test uses the asymptotic normality of the statistic $|\tau|\sqrt{\frac{9n(n-1)}{2(2n+5)}}$, where $\tau$ is Kendall's tau statistic.
-to be added in suppmat-->If independence could be rejected ($0.05$ significance level), model selection proceeded.
-The third step of model selection was to fit multiple bivariate copula families to the ranked data $(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`.
-Families considered (with abbreviations here indicated) included the **N**ormal/Gaussian, **C**layton, **G**umbel, **F**rank, **J**oe, **BB1**, **BB6**, **BB7**, **BB8**, **S**urvival **C**layton , **S**urvival **G**umbel, **S**urvival **J**oe, **S**urvival **BB1**, **S**urvival **BB6**, **S**urvival **BB7** and **S**urvival **BB8** copulae.
-These copulae were chosen carefully so that they covered the whole range from exclusively lower (C, SG, SJ, SBB6, SBB8) or upper tail dependence (SC, G, J, BB6, BB8) to both tail dependence (BB1, SBB1, BB7, SBB7) in addition to two symmetric copulae (N, F).
-AIC and BIC values and accompanying model weights ($\text{AIC}_{\text{w}}$, $\text{BIC}_{\text{w}}$) were generated using standard formulas [@burnham2003_modelselection].
-Descriptions of these copula families can be found in introductory texts on copulae [@nelsen2006_copula ; @joe2014_dependence]  and in `VineCopula` documentation.
-The lower- and upper-tail dependence of the best-fitting member of each family was also provided by `BiCopEst`.
-Finally, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
+family or as a boundary case. If independence could be rejected ($0.05$ significance level), model 
+selection proceeded. Third, we fit the bivariate copula families listed previously to the ranked data 
+$(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`, and 
+obtained AIC and BIC values and accompanying model weights, $\text{AIC}_{\text{w}}$ and 
+$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection]. The weight of a model represents the 
+relative support of data for the model; weights sum to 1 across models considered. The lower- and 
+upper-tail dependence of the best-fitting member of each family was also provided by `BiCopEst`. 
+Fourth, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
 model-averaged lower- and upper-tail dependence values using standard model averaging 
 formulas [@burnham2003_modelselection]; likewise for BIC.
-
 Model selection methods will give the *relative* support of several models, but will not,
-on their own, indicate whether any of the models selected is an objectively good fit (all 
-models considered might poorly describe the data). To test this, and as the fourth step of
-our procedure, we tested the goodness of fit of our best-AIC copula family using a 
+on their own, indicate whether any of the models selected is an objectively good fit: all 
+models considered might poorly describe the data. To test this, and as the final step of
+our procedure, we tested the goodness of fit of our AIC-best copula family using a 
 bootstrapping procedure implemented in `BiCopGofTest` of `VineCopula`. The method performed 
 two tests, one based on a Cramer-von Mises statistic, and one based on a 
-Kolmogorov-Smirnov statistic. For computational efficiency, 
+Kolmogorov-Smirnov statistic. For keep computation times reasonable, 
 an initial run using $100$ bootstraps was performed, and if
@@
 
-The first part of \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} was answered for a bivariate dataset 
-$(x_i,y_i)$ for $i=1,...,n$ by comparing the AIC weights of the non-Normal copula
-families with that of the Normal family. <!--The same was also done using BIC.....no, we didn't report it-->
-The second part of \hyperref[Intro]{\textbf{(Q1)}} was answered by comparing 
+The first part Q1 was answered for a bivariate dataset 
+by comparing the AIC weights of the non-normal copula
+families with that of the normal family. 
+The second part of Q1 was answered by comparing 
 the model-averaged lower- and upper-tail dependence
-values to the lower- and upper-tail dependence values of a Normal copula, both of which 
-are 0. The third part of \hyperref[Intro]{\textbf{(Q1)}}
+values to the lower- and upper-tail dependence values of a normal copula (which 
+are 0). The third part of Q1
 was answered by comparing the difference of the model-averaged lower- and upper-tail 
-dependence values to the analogous difference for a Normal copula, again 0.
-
-For multivariate datasets, e.g., the aphid abundance or first flight data, 
-each of which consists of time series from 10 selected locations, we carried out the same bivariate analysis
-for all possible pairwise combinations of those locations and compared AIC-weights and tail-dependence of best-fit(minimum AIC) copula with that of the Normal copula.
-We also counted for how many location pairs model -averaged lower tail dependence is greater than model-averaged upper-tail dependence or vice-versa.
-Then, we reported as the percentage of location pairs showing stronger lower tail dependence out of total non-independent location pairs considered.
+dependence values to the analogous difference for a normal copula, again 0.
+For multivariate datasets,  
+each of which consists of time series from multiple locations, we carried out the bivariate analysis
+for all possible pairwise combinations of locations and compared AIC weights and tail-dependence of 
+minimum AIC copulas with that of the normal copula. We also counted the number of location pairs 
+for which model-averaged lower-tail dependence was greater than model-averaged upper-tail dependence 
+or vice-versa, and reported as the percentage of pairs showing stronger lower-tail dependence out of 
+the total number non-independent location pairs considered.
+
+<!--***DAN: made it to here 2018 12 16-->
 

### 2018-12-16 SupportingInformation.Rmd
furter editing

--- 2018-12-14 SupportingInformation.Rmd
@@
 author: "S. Ghosh, T. Loecke, C.P. Reid, D.C. Reuman"
-date: "February 12, 2018"
+date: "December 16, 2018"
 geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"
@@
 source("./ThirdPedagogFig.R")
+source("./MakeCopulaPedagogFigs.R")
 ```
@@
 ```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=8 cm,height=8 cm]{./Results/SoilCNmapUSA.pdf} 
-\caption{Locations of central pedons with soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil surface) from 5907 sampling sites across the coterminous United States.
-\label{fig_SoilCNmapUSA}}
-\end{center}
-\end{figure}
 
@@
 
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=8 cm,height=8 cm]{./Results/aphid_map_plot.pdf} 
-\caption{Locations of Rothamsted Insect Survey suction-trap across UK are indicated by the red circles 
-out of which only blue boxes are considered for our study based on having atleast 30 annual data points. \label{fig_aphid_map}}
-\end{center}
-\end{figure}
-
 ```{r tab_plankton_info, echo=F, results='asis',message=F}
@@
 
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=8 cm,height=8 cm]{./Results/plankton_map_plot.pdf} 
-\caption{CPR data collected from 26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ on map) around UK sea (indicated by the red circles) but out of which only blue boxes are considered for our study based on having atleast 45 annual data points.
-\label{fig_plankton_map}}
-\end{center}
-\end{figure}
-
 ```{r read_methane_data, echo=F, results="markup"}
@@
 ```
-
-<!--# Description of Data \label{Data_detailed}
-to be filled later
---->
-
-<!--# Suppmat for model selection approach \label{MS_detailed}
-to be filled later
---->
 
@@
 
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=8 cm,height=8 cm]{./Results/SoilCNmapUSA.pdf} 
+\caption{Locations of central pedons with soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil surface) from 5907 sampling sites across the coterminous United States.
+\label{fig_SoilCNmapUSA}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=8 cm,height=8 cm]{./Results/aphid_map_plot.pdf} 
+\caption{Locations of Rothamsted Insect Survey suction-trap across UK are indicated by the red circles 
+out of which only blue boxes are considered for our study based on having atleast 30 annual data points. \label{fig_aphid_map}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=8 cm,height=8 cm]{./Results/plankton_map_plot.pdf} 
+\caption{CPR data collected from 26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ on map) around UK sea (indicated by the red circles) but out of which only blue boxes are considered for our study based on having atleast 45 annual data points.
+\label{fig_plankton_map}}
+\end{center}
+\end{figure}
+
 # Transformations used in Fig. \ref{MT-fig_cop_pedag1}
@@
 # Background on the copulas we fitted with data \label{copula_details}
+
+<!--Notes about individual copula parameter ranges:
+
+-->
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Normal.pdf}
+\caption{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
+S is Spearman correlation; p is the value of the parameter for the N family (it is a 
+one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively.\label{figped_N}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Clayton.pdf}
+\caption{Log-transformed pdfs (A-E) and samples (F-J) from example C copulas. S is Spearman correlation; p is the value of the parameter for the C family (it is a 
+one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively.\label{figped_C}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvClayton.pdf}
+\caption{Log-transformed pdfs (A-E) and samples (F-J) from example SC 
+copulas. S is Spearman correlation; p is the value of the parameter for the SC 
+family (it is a one-parameter family); and LT and UT are the measures of lower- and 
+upper-tail dependence, respectively.\label{figped_SC}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
+\caption{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
+Spearman correlation; p is the value of the parameter for the G family (it is a 
+one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively.\label{figped_G}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvGumbel.pdf}
+\caption{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
+copulas. S is Spearman correlation; p is the value of the parameter for the SG 
+family (it is a one-parameter family); and LT and UT are the measures of lower- and 
+upper-tail dependence, respectively.\label{figped_SG}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Joe.pdf}
+\caption{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; p is the value of the parameter for the J family (it is a 
+one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively.\label{figped_J}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvJoe.pdf}
+\caption{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
+copulas. K is Spearman correlation; p is the value of the parameter for the SJ 
+family (it is a one-parameter family); and LT and UT are the measures of lower- and 
+upper-tail dependence, respectively.\label{figped_SJ}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
+\caption{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; p is the value of the parameter for the F family (it is a 
+one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively.\label{figped_F}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB1.pdf}
+\caption{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
+denote the two parameters of the family (it is a 
+two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
+respectively.\label{figped_BB1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
+\caption{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; p1 ad p2 
+denote the two parameters of the family (it is a 
+two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
+respectively.\label{figped_SBB1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB6.pdf}
+\caption{Log-transformed pdfs for example BB6 copulas. K is Kendall correlation; p1 and p2
+denote the two parameters of the family (it is a 
+two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
+respectively.\label{figped_BB6}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
+\caption{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; p1 ad p2 
+denote the two parameters of the family (it is a 
+two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
+respectively.\label{figped_SBB6}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB7.pdf}
+\caption{Log-transformed pdfs for example BB7 copulas. K is Kendall correlation; p1 and p2
+denote the two parameters of the family (it is a 
+two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
+respectively.\label{figped_BB7}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
+\caption{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; p1 ad p2 
+denote the two parameters of the family (it is a 
+two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
+respectively.\label{figped_SBB7}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB8.pdf}
+\caption{Log-transformed pdfs for example BB8 copulas. K is Kendall correlation; p1 and p2
+denote the two parameters of the family (it is a 
+two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
+respectively.\label{figped_BB8}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
+\caption{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; p1 ad p2 
+denote the two parameters of the family (it is a 
+two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
+respectively.\label{figped_SBB8}}
+\end{center}
+\end{figure}
+
+<!--***DAN: made it to here 2018 12 16-->
 

### 2018-12-16 MakeCopulaPedagogFigs.R

@@ -56,7 +56,9 @@ muvcpp<-function(coplist,numpts,corvals,cortype,thetavals,tot_wd,filename)
   axis(1,labels=FALSE)
   mtext("v",2,1.15)
   mtext("A",3,.05,cex=cexvalpl,adj=0)
-  mtext(paste0(cortype,"=",round(spearvals[px+1],2),"; p=",round(thetavals[px+1],2)),3,1.25,cex=cexvaltxt)
+  mtext(paste0(cortype,"=",signif(spearvals[px+1],2),"; p=",signif(thetavals[px+1],2)),3,1.55,cex=cexvaltxt)
+  lbds<-unname(lambda(coplist[[px+1]]))
+  mtext(paste0("LT=",signif(lbds[1],2),"; UT=",signif(lbds[2],2)),3,1,cex=cexvaltxt)
   
   #top panel 2
   px<-1
@@ -72,7 +74,9 @@ muvcpp<-function(coplist,numpts,corvals,cortype,thetavals,tot_wd,filename)
   axis(1,labels=FALSE)
   axis(2,labels=FALSE)
   mtext("B",3,.05,cex=cexvalpl,adj=0)
-  mtext(paste0(cortype,"=",round(spearvals[px+1],2),"; p=",round(thetavals[px+1],2)),3,1.25,cex=cexvaltxt)
+  mtext(paste0(cortype,"=",signif(spearvals[px+1],2),"; p=",signif(thetavals[px+1],2)),3,1.55,cex=cexvaltxt)
+  lbds<-unname(lambda(coplist[[px+1]]))
+  mtext(paste0("LT=",signif(lbds[1],2),"; UT=",signif(lbds[2],2)),3,1,cex=cexvaltxt)
   
   #top panel 3
   px<-2
@@ -88,7 +92,9 @@ muvcpp<-function(coplist,numpts,corvals,cortype,thetavals,tot_wd,filename)
   axis(1,labels=FALSE)
   axis(2,labels=FALSE)
   mtext("C",3,.05,cex=cexvalpl,adj=0)
-  mtext(paste0(cortype,"=",round(spearvals[px+1],2),"; p=",round(thetavals[px+1],2)),3,1.25,cex=cexvaltxt)
+  mtext(paste0(cortype,"=",signif(spearvals[px+1],2),"; p=",signif(thetavals[px+1],2)),3,1.55,cex=cexvaltxt)
+  lbds<-unname(lambda(coplist[[px+1]]))
[…]

### 2018-12-17 Paper.Rmd
further edits

@@
 for all possible pairwise combinations of locations and compared AIC weights and tail-dependence of 
-minimum AIC copulas with that of the normal copula. We also counted the number of location pairs 
+minimum-AIC copulas with that of the normal copula. We also counted the number of location pairs 
 for which model-averaged lower-tail dependence was greater than model-averaged upper-tail dependence 
-or vice-versa, and reported as the percentage of pairs showing stronger lower-tail dependence out of 
+or vice-versa, and reported the percentage of pairs showing stronger lower-tail dependence out of 
 the total number non-independent location pairs considered.
 
-<!--***DAN: made it to here 2018 12 16-->
-
-## Nonparametric analysis of tail dependence \label{NPA}
-
-In addition to the model selection approach described above, **nonparametric approaches** to quantify lower <!--($\cor_l$, $\Ps_l$ and $\Dsq_l$ statistics)--> and upper tail dependence <!--($\cor_u$, $\Ps_u$ and $\Dsq_u$ statistics)--> were also used to answer \hyperref[Intro]{\textbf{(Q1)}}.
-Using two distinct methods, when results are consistent, provides greater confidence
-that results are robust.
-
-Left (or lower)- and right (or upper)-tail dependence quantify the extent to which two ranked
-variables $u_i$, $v_i$ for $i=1,...,n$ 
-(see \nameref{Model_selection} for defintions of $u_i$ and $v_i$)
-are related to each other in the left and right tails, respectively, of their 
-distributions [@nelsen2006_copula, @joe2014_dependence]. 
-Our nonparametric study of tail dependence uses three
-statistics (i.e., $\cor$, $\Ps$ and $\Dsq$) which can quantify the extent to which $u_i$ and $v_i$ are 
-related in any part of their distributions. We here describe the 
+In addition to our model selection approach, we also used nonparametric approaches to 
+quantify lower- and upper-tail dependence to answer Q1. Using two methods can provide greater confidence 
+that results are robust. We used three
+statistics <!--( $\cor$, $\Ps$ and $\Dsq$)--> which can quantify the extent to which 
+two normalized ranks $u_i$ and $v_i$ are 
+related in any part of their distributions; we here describe the 
 statistics. The statistics were formulated with positively associated
@@
 $0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
-$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
-fall (these are the blue and red lines, respectively, in 
-Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
+$u+v=2u_b$, which intersect the unit square (Fig. \ref{SM-fig_stat_image}). 
+Our statistics quantify the dependence
 between $u_i$ and $v_i$ in the region bounded by these lines.
@@
 the right parts of the distributions. 
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
-\caption{Schematic diagram supporting defintions of nonparametric tail dependence statistics \label{fig_stat_image}}
-\end{center}
-\end{figure}
-
-The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
+The first statistic, denoted $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
 correlation of $u_i$ and $v_i$ that is attributable to the points 
-lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
+lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$:
 \begin{equation}\label{eq.Cor}
-   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
-   (u_i-\mean(u)) (v_i-\mean(v)),
+   \cor_{l_b,u_b}(u,v) = \frac{\sum 
+   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}}.
 \end{equation}
-where means and sample variances are computed using all $n$ data 
+Here means and sample variances are computed using all $n$ data 
 points, but the
@@
 Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
-between $u$ and $v$ in the region given by the bounds.
-
-We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
-$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
-$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
-and a distance less than $h$ from the line $v=u$, divided by the 
-total number of points within the bounds. This function is
-defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
-is half the longer of the two segments obtained by intersecting the
-bounds with the unit square. It is easy to see $S(0)=0$ and 
-$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
-of $h$, to be the area within the bounds and the unit square
-and within distance $h$ of the line $v=u$, divided by the area
-within the bounds and the unit square. This is the expected value of
-$S(h)$ for independent data. We then defined
-\begin{equation}\label{eq.P}
-   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
-\end{equation}
-As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
-indicate stronger positive association
-between $u$ and $v$ in the region given by the bounds.
-
-The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
+between $u$ and $v$ in the region given by the bounds. The sum of 
+$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
+$\cor_{l_b_k,u_b_k}(u,v)$ for bounds $l_b_k,u_b_k$ that partition the
+interval $(0,1)$) equals the standard Spearman correlation.
+
+We also defined the statistic $\Ps_{l_b,u_b}$, which has a similar 
+interpretation to $\cor_{l_b_k,u_b_k}(u,v)$; we use both statistics 
+because in some instances $\Ps_{l_b,u_b}$ appears to have more power to
+reveal significant tail dependence and asymmetry of tail dependence. 
+<!--***DAN: need to make this a bit more precise, perhaps just enough to cite one of the supp mat sections where tests are done of these stats, but you need to review the tests first and what they show-->
+Our third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
 distance between points satisfying $u_i+v_i>2l_b$ and 
-$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
+$u_i+v_i<2u_b$ and the line $v=u$. 
+Unlike $\cor_{l_b,u_b}$ and 
 $\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
 $u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
-then small. 
+small in that case. 
 Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
@@
 $\Dsq_{l_b,u_b}$ indicate strong dependence.
+
+<!--***DAN: made it to here 2018 12 17 5pm-->
 

### 2018-12-17 SupportingInformation.Rmd
further edits

@@
 
-<!--***DAN: made it to here 2018 12 16-->
+# Nonparametric statistics
+
+<!--***DAN: Right now this is a dumping grounds for text taken from the main text.
+The text is valuable, but needs to be organized to fit together in this supp mat
+section-->
+
+We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
+$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
+$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
+and a distance less than $h$ from the line $v=u$, divided by the 
+total number of points within the bounds. This function is
+defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
+is half the longer of the two segments obtained by intersecting the
+bounds with the unit square. It is easy to see $S(0)=0$ and 
+$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
+of $h$, to be the area within the bounds and the unit square
+and within distance $h$ of the line $v=u$, divided by the area
+within the bounds and the unit square. This is the expected value of
+$S(h)$ for independent data. We then defined
+\begin{equation}\label{eq.P}
+   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
+\end{equation}
+As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
+indicate stronger positive association
+between $u$ and $v$ in the region given by the bounds.
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
+\caption{Schematic diagram supporting definitions of nonparametric tail dependence statistics. The red and blue lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
+\end{center}
+\end{figure}
 

### 2018-12-19 Paper.Rmd
further edits

@@
 were obtained from A and B through monotonic transformation 
-(Appendix \ref{SM-Data_details} for details),
+(Appendix \ref{SM-pedagog1_details} for details),
 <!--***DAN: Dan to go back and write the appendix-->
@@
 normal marginal distributions of A and B to gamma (C, D) or 
-uniform (E, F) marginals (details in Appendix \ref{SM-fig_cop_pedag1}). (G, H) Contours 
+uniform (E, F) marginals (details in Appendix \ref{SM-pedagog1_details}). (G, H) Contours 
 of log-scale density functions of Clayton (G) and survival-Clayton 
-(H) copulas with parameter 5 (Appendix \ref{SM-copula_details}). (I,J) Reflections
+(H) copulas with parameter 5 (Appendix \ref{SM-pedagog1_details}). (I,J) Reflections
 of E, F, respectively, about a vertical axis. Subscripts h and v 
@@
 <!--***DAN: Dan to go back and add the appendices with details as it says. Make sure
-to mention in the first one 
-the shape and scale params of the gamma marginals in C, D were 2-->
+to mention the shape and scale params of the gamma marginals in C, D were 2-->
 
@@
 In addition to our model selection approach, we also used nonparametric approaches to 
-quantify lower- and upper-tail dependence to answer Q1. Using two methods can provide greater confidence 
-that results are robust. We used three
+quantify lower- and upper-tail dependence and to answer Q1. Using two approaches can provide greater confidence 
+that results are robust. We used three nonparametric
 statistics <!--( $\cor$, $\Ps$ and $\Dsq$)--> which can quantify the extent to which 
@@
 diagrams will assume positive associations, though generalizations are 
-possible. Given two bounds 
+straightforward. Given two bounds 
 $0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
@@
 \end{equation}
-Here means and sample variances are computed using all $n$ data 
+\noindent Here means and sample variances are computed using all $n$ data 
 points, but the
@@
 $\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
-$\cor_{l_b_k,u_b_k}(u,v)$ for bounds $l_b_k,u_b_k$ that partition the
+$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
 interval $(0,1)$) equals the standard Spearman correlation.
 
-We also defined the statistic $\Ps_{l_b,u_b}$, which has a similar 
-interpretation to $\cor_{l_b_k,u_b_k}(u,v)$; we use both statistics 
+We also defined the statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
+which has a similar 
+interpretation to $\cor_{l_{b_k},u_{b_k}}(u,v)$; we use both statistics 
 because in some instances $\Ps_{l_b,u_b}$ appears to have more power to
-reveal significant tail dependence and asymmetry of tail dependence. 
+reveal tail dependence than does $\cor_{l_{b_k},u_{b_k}}(u,v)$. 
 <!--***DAN: need to make this a bit more precise, perhaps just enough to cite one of the supp mat sections where tests are done of these stats, but you need to review the tests first and what they show-->
@@
 $u_i+v_i<2u_b$ and the line $v=u$.
+<!--***DAN: Shyamolina, is this actually right, i.e., did we use the average, or did we 
+use the sum divided by n? Whichever way we did it is OK, but please check the code and revise the above 
+text if necessary and insert a comment answering this comment. I tried to check this, myself, 
+but I was having a hard time finding where the statistics are implemented. Please also
+indicate in your response comment what line of what file has the implementation of these
+statistics.-->
 Unlike $\cor_{l_b,u_b}$ and 
@@
 dependence in the portions of the distributions of $u$ and $v$ given by 
-the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
+the bounds, whereas small values of 
 $\Dsq_{l_b,u_b}$ indicate strong dependence.
 
-<!--***DAN: made it to here 2018 12 17 5pm-->
-
-For large datasets (i.e., datasets with large $n$),
-we can consider $l_b$ and $u_b$ close together without incurring undue
-sampling variation in our statistics, and we can consider 
+For large datasets (large $n$),
+we could use $l_b$ and $u_b$ close together without incurring undue
+sampling variation in our statistics, and we could consider 
 several different bands $(l_b,u_b)$ to understand how dependence
@@
 
+To test for asymmetry of dependence in upper and lower portions of distributions,
+we used differences of upper and lower statistics.
 The difference $\cor_l-\cor_u$ is expected to be positive if 
@@
 than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
-This holds, for instance, for the Clayton copula [@nelsen2006_copula].
-The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
-"lower" statistics occur in opposite order, is also expected to be 
-positive under the same conditions. We use these three differences
-to detect asymmetry of tail dependence. Likewise, for large datasets,
-we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
+The difference $\Dsq_u-\Dsq_l$ (note the opposite order), is also expected to be 
+positive under the same conditions. These statistics were used for smaller datasets.
+For large datasets, we used $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
 $\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
 $\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$. 
-We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulae (for details see Appendix \ref{SM-Test_npa_stat}).
+How close depended on $n$, and was chosen to ensure that sample variation 
+was not unduly large. We tested our statistics by testing whether 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of 
+tail dependence known to be present in data simulated from various copulas 
+(Appendix \ref{SM-Test_npa_stat}).
 
@@
 of the statistics $\cor_{l_b,u_b}$,
-$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to 
+$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
+$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
+$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ to 
 distributions of values of the same statistics computed on surrogate datasets that were produced from 
-the empirical data by randomizing it in a special way to have no tail dependence (for details see Appendix \ref{SM-surrog_test}). The surrogate/randomized datasets had exactly 
+the empirical data by randomizing it in a special way to have no 
+tail dependence (Appendix \ref{SM-surrog_test}). 
+The surrogate/randomized datasets had exactly 
 the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
 correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
-but had Normal copula structure. Thus our comparisons tested the null hypothesis 
+but had normal copula structure. Thus our comparisons tested the null hypothesis 
 that our statistics took values on the empirical data no different from what would have been expected 
-if the copula structure of the data were Normal, but the data were otherwise 
+if the copula structure of the data were normal, but the data were otherwise 
 statistically unchanged.
@@
 Kendall (or Spearman)
-correlation. Thus the comparison helps answer the first two parts of question \hyperref[Intro]{\textbf{(Q1)}} 
-from the \nameref{Intro} for bivariate datasets. 
+correlation. Thus the comparison helps answer the first two parts of Q1
+for bivariate datasets. 
 Significant deviations correspond to deviations from normal 
@@
 from that of a normal copula.
-
-The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
+The comparison of one of the statistics 
+$\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
+$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
+$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$,
 as computed on the empirical data, to the distribution of its values on many surrogate datasets 
-provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
-third part of question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} for bivariate datasets.
-For multivariate datasets, we used a spatial resampling scheme (for details see Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals, instead of the surrogate
+provides a test of asymmetry of dependence. Thus this comparison helps answer the
+third part of Q1 for bivariate 
+datasets. For multivariate datasets, we used a spatial resampling scheme 
+(Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals, instead of the surrogate 
 approach described above.
+
+<!--***DAN: If we end up being desperate to cut length, and assuming the nonparmetric results
+are sufficiently similar to model selection results, we may be able to put everything about 
+the nonparametric results in the supp mat and just say in the main text that we also used a 
+nonparametric approach and it gave the same general conclusions. Though it may take some work
+to split out the two approaches and all their results from each other.-->
+
+<!--***DAN: made it to here 2018 12 18 9pm-->
 

### 2018-12-19 SupportingInformation.Rmd
further edits

@@
 
+# Transformations and copulas used in Fig. \ref{MT-fig_cop_pedag1} \label{pedagog1_details}
+
+<!--***DAN: fill in later-->
+
+# Details of the data \label{Data_details}
+
+<!--***DAN: This is a dumping ground at the moment, but needs to be organized into a section that gives all the necessary details about the data-->
+
+<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
+unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
+location in the ocean.
+Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979].
+All the samples available for a given variable in a given area in a given month of a given year were
+averaged, producing a monthly time series for each year.
+Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).-->
+
+ [@Sheppard2017] converted monthly values to yearly values by averaging over all months.
+But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available).
+We used data from 14 locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}).
+
+ The Great Miami Wetland Mitigation Bank  was at $36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
+Additional information on the history of the site and soil types is detailed in @Jarecke2016.
+Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016.
+**See and Smyth et al, pending for
+additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.
+
 ```{r tab_aphid_info, echo=F, results='asis',message=F}
@@
 
-# Details of the data \label{Data_details}
-
-<!--***DAN: This is a dumping group at the moment, but needs to be organized into a section that gives all the necessary details about the data-->
-
-<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
-unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
-location in the ocean.
-Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979].
-All the samples available for a given variable in a given area in a given month of a given year were
-averaged, producing a monthly time series for each year.
-Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).-->
-
- [@Sheppard2017] converted monthly values to yearly values by averaging over all months.
-But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available).
-We used data from 14 locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}).
-
- The Great Miami Wetland Mitigation Bank  was at $36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
-Additional information on the history of the site and soil types is detailed in @Jarecke2016.
-Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016.
-**See and Smyth et al, pending for
-additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.
-
 \begin{figure}[!h]
@@
 
-# Transformations used in Fig. \ref{MT-fig_cop_pedag1}
-
 # Background on the copulas we fitted with data \label{copula_details}
 
-<!--Notes about individual copula parameter ranges:
-
+<!--Dumping ground at present, need to organize and fill in.
+
+Notes about individual copula parameter ranges:
 -->
@@
 
-# Nonparametric statistics
+# Nonparametric statistics \label{nonparam_stats}
 
@@
 
-# Testing non-parametric statistics for various symmetric and asymmetric copulas \label{Test_npa_stat}
+# Testing nonparametric statistics for various symmetric and asymmetric copulas \label{Test_npa_stat}
 

### 2018-12-19 Paper.Rmd
adding comment about 3rd npa stat

@@
 good to use the same points, for consistency. Don't worry at this stage about
-making other figures, other than these three, consistent.-->
+making other figures, other than these three, consistent.
+***Shya: Done.-->
 
@@
 indicate in your response comment what line of what file has the implementation of these
-statistics.-->
+statistics.
+***Shya: In code we used the sum of squared distances divided by n(=number of points lying within two boundaries) which I think okay to be called as "average-squared distance" or average of squared distances between points within the two boundaries.
+Reference : CopulaFunctions_flexible.R file, line 24.-->
 Unlike $\cor_{l_b,u_b}$ and 

### 2018-12-20 Paper.Rmd
minor edit

@@
 statistics.
-***Shya: In code we used the sum of squared distances divided by n(=number of points lying within two boundaries) which I think okay to be called as "average-squared distance" or average of squared distances between points within the two boundaries.
+***Shya: In code we used the sum of squared distances divided by n(=number of points lying within two boundaries) which I think okay to be called as "average-squared distance" or (to be precised) average of squared distances of points from 1:1 line.
 Reference : CopulaFunctions_flexible.R file, line 24.-->

### 2018-12-20 Paper.Rmd
further edits, some tasks for Shyamolina

--- 2018-12-19 Paper.Rmd
@@
 
-\begin{figure}[!h]
+\begin{figure}
 \begin{center}
@@
 \end{figure}
-<!--***DAN: Shya, I like the grey color of your points on the above figure, 
-but I prefer the 
-solid plotting symbols and the size of plotting symbols used on the new 
-pedagogical figure that I made. Can you please change the code of both figures
-so we get grey, with the plotting symbols I used, on both plots? Likeiwse for the 
-new figure produced in ./ThirdPedagogFig.R, please. It will be
-good to use the same points, for consistency. Don't worry at this stage about
-making other figures, other than these three, consistent.-->
 
@@
 
-<!--***DAN: made it to here 2018 12 18 9pm-->
-
 # Results for Q1: Ecological datasets often have non-normal copula structure and asymmetric tail dependence \label{Results}
-<!--## question 1
-- Do datasets in ecology and related fields have non-Normal copula structure?
-- Do they show tail dependence distinct from that of a Normal copula?
-- In particular do they show asymmetric tail dependence?
-
-To answer \hyperref[Intro]{\textbf{(Q1)}}, we chose one representative datasets for each category : with a bivariate (soil C and N data) and a multivariate (green spruce aphid-abundance data) dataset, we summarized our findings as follows.
--->
-
-## Results for bivariate data \label{Result_biv}
-
+
+```{r read_BivMS_soilCN_data,echo=F}
+d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
+d<-d[,c("SOCstock100","TSNstock100")] #raw soilCN data
+BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
+```
+
+```{r read_RES_aphid_count, echo=F}
+RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
+```
+
+Across essentially all the sub-fields and levels of organization within ecology that we 
+examined, copula structures were typically non-normal and variables exhibited asymmetric tail 
+dependence. To make results easier for readers who are new to copulas to absorb: we first 
+present results in detail for an example bivariate dataset; followed by a summary of results
+for all bivariate datasets; then we present results in 
+detail for an example multivariate dataset; followed by a summary of results 
+for all multivariate datasets.
+
+<!--bivariate example-->
+We will now discuss briefly findings from one of our bivariate datasets as follows.
+As described in the \nameref{Data}, raw data plot for RaCA database soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
+\ref{fig_biv_multi_raw_cop}(A) and (E), respectively.
+We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.
+
+<!--#### Model selection approach \label{Ex_SoilCN_Model_selection}-->
+
+From the model selection approach, the test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`).
+We then fitted several copulae (all $16$ families as mentioned in \nameref{Model_selection})
+and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulae (Table \ref{SM-tab_soilCNfit}). 
+Based on AIC weights, `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the best fit.
+The table shows that several copulae are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of \hyperref[Intro]{\textbf{(Q1)}} for these data.
+
+The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`.
+These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of \hyperref[Intro]{\textbf{(Q1)}} for these data.
+Although these results do convincingly show non-Normal copula structure, we should caution that because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
+Therefore, we also verified this finding with \nameref{NPA} approach (see Fig.
+\ref{SM-fig_soilCN_nonparam}), where each of our statistics ($\cor$, $\Ps$, $\Dsq$), as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 Normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig.
+\ref{SM-fig_soilCN_nonparam}).
+Each bin in each subplot of Fig.
+\ref{SM-fig_soilCN_nonparam} contains a number indicating how many Normal surrogates out of 1000, are different ($95\%$ confidence interval) from Normal-copula statistics.
+Those numbers clearly show that tail dependence of data is stronger in both the lower and upper tails than would be expected under a null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}.
+
+Furthermore, the difference of model-averaged tail dependence statistics (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}.
+To be precise, this difference being negative, indicated stronger upper-tail dependence than lower, which was also supported by non-parametric results (see Table \ref{SM-tab_soilCN_asym}).
+
+<!--bivariate summary-->
 We summarized our results columnwise in Table \ref{tab_bivar_summary} for different kinds of bivariate datasets (biogeochemical, evolutionary and ecological data) that were plotted in Fig.
@@
 
-### Example for bivariate case : soil C and N data shows upper tail dependence\label{Ex_SoilCN}
-
-```{r read_BivMS_soilCN_data,echo=F}
-d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
-d<-d[,c("SOCstock100","TSNstock100")] #raw soilCN data
-BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
-```
-
-We will now discuss briefly findings from one of our bivariate datasets as follows.
-As described in the \nameref{Data}, raw data plot for RaCA database soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
-\ref{fig_biv_multi_raw_cop}(A) and (E), respectively.
-We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.
-
-<!--#### Model selection approach \label{Ex_SoilCN_Model_selection}-->
-
-From the model selection approach, the test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`).
-We then fitted several copulae (all $16$ families as mentioned in \nameref{Model_selection})
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulae (Table \ref{SM-tab_soilCNfit}). 
-Based on AIC weights, `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the best fit.
-The table shows that several copulae are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of \hyperref[Intro]{\textbf{(Q1)}} for these data.
-
-The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`.
-These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of \hyperref[Intro]{\textbf{(Q1)}} for these data.
-Although these results do convincingly show non-Normal copula structure, we should caution that because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
-Therefore, we also verified this finding with \nameref{NPA} approach (see Fig.
-\ref{SM-fig_soilCN_nonparam}), where each of our statistics ($\cor$, $\Ps$, $\Dsq$), as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 Normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig.
-\ref{SM-fig_soilCN_nonparam}).
-Each bin in each subplot of Fig.
-\ref{SM-fig_soilCN_nonparam} contains a number indicating how many Normal surrogates out of 1000, are different ($95\%$ confidence interval) from Normal-copula statistics.
-Those numbers clearly show that tail dependence of data is stronger in both the lower and upper tails than would be expected under a null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}.
-
-Furthermore, the difference of model-averaged tail dependence statistics (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}.
-To be precise, this difference being negative, indicated stronger upper-tail dependence than lower, which was also supported by non-parametric results (see Table \ref{SM-tab_soilCN_asym}).
-
-<!--#### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}-->
-
-## Results for multivariate spatial data \label{Result_multi}
-```{r read_RES_aphid_count, echo=F}
-RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
-```
-
-We did model selection and non-parametric analysis with all 20 aphid and 22 plankton species and found our two approaches were consistent to each other.
-Here, to present our results we chose one species from each category.
-We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets (i.e., abundance data for green spruce aphid and firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, abundance data for plankton from UK seas from $14$ sampling locations,
-methane flux data from $13$ sampling locations).
-
-We found major occurence of non-independent copulae (row 2, Table \ref{tab_multivar_summary}) out of all pairwise locations (row 1, Table \ref{tab_multivar_summary}) considered.
-Non-Normal copulae, most of the time, emerged as best-fit based on minimum AIC out of 16 copula family considered (row 3, Table \ref{tab_multivar_summary}) answering the first part of question \hyperref[Intro]{\textbf{(Q1)}}.
-All four data sets showed Normal copulae were always poor fit in comparison to best-fit non-Normal copulae (compare their average AIC-weights from row 4 and row 5 of Table \ref{tab_multivar_summary}).
-We considered the "good-fit" location pair based on the p-values (>0.01) obtained from both Cramer-von Mises and Kolmogorov-Smirnov statistics (row 6, Table \ref{tab_multivar_summary}).
-Some spatial data (green spruce aphid's count, Ceratium furca's count and methane flux) showed stronger lower tail dependence (compare row 7 and row 8 from Table \ref{tab_multivar_summary}, also see row 9) whereas Leaf-plum curling's firstflight data showed stronger upper tail dependence which answered the second and third part of question \hyperref[Intro]{\textbf{(Q1)}}.
-The spatial synchrony (from model selection) results were also verified by the non-parametric statistic which showed exclusively positive (or negative) confidence interval of the distribution $(\Ps_l - \Ps_u)$ estimated from all location pair for lower (or upper)- tail dependent spatial data (except methane-flux data which had marginal lower quantile).
-Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.
-
-### Example for multivariate case : abundance data for the green spruce aphid (*Elatobium abietinum*) shows spatial synchrony with lower tail dependence\label{Ex_count}
-
-<!--#### Model selection approach \label{Ex_count_Model_selection}-->
-
+<!--multivariate example-->
 As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid's abundance data, model selection among copulae was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations.
@@
 Visual illustration of those results is discussed in Appendix \ref{SM-visual_npa}.
+
+<!--multivariate summary-->
+We did model selection and non-parametric analysis with all 20 aphid and 22 plankton species and found our two approaches were consistent to each other.
+Here, to present our results we chose one species from each category.
+We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets (i.e., abundance data for green spruce aphid and firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, abundance data for plankton from UK seas from $14$ sampling locations,
+methane flux data from $13$ sampling locations).
+
+We found major occurence of non-independent copulae (row 2, Table \ref{tab_multivar_summary}) out of all pairwise locations (row 1, Table \ref{tab_multivar_summary}) considered.
+Non-Normal copulae, most of the time, emerged as best-fit based on minimum AIC out of 16 copula family considered (row 3, Table \ref{tab_multivar_summary}) answering the first part of question \hyperref[Intro]{\textbf{(Q1)}}.
+All four data sets showed Normal copulae were always poor fit in comparison to best-fit non-Normal copulae (compare their average AIC-weights from row 4 and row 5 of Table \ref{tab_multivar_summary}).
+We considered the "good-fit" location pair based on the p-values (>0.01) obtained from both Cramer-von Mises and Kolmogorov-Smirnov statistics (row 6, Table \ref{tab_multivar_summary}).
+Some spatial data (green spruce aphid's count, Ceratium furca's count and methane flux) showed stronger lower tail dependence (compare row 7 and row 8 from Table \ref{tab_multivar_summary}, also see row 9) whereas Leaf-plum curling's firstflight data showed stronger upper tail dependence which answered the second and third part of question \hyperref[Intro]{\textbf{(Q1)}}.
+The spatial synchrony (from model selection) results were also verified by the non-parametric statistic which showed exclusively positive (or negative) confidence interval of the distribution $(\Ps_l - \Ps_u)$ estimated from all location pair for lower (or upper)- tail dependent spatial data (except methane-flux data which had marginal lower quantile).
+Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.
 
@@
 library(tibble)
-dt <- tibble(Description = c("1. p-value from independence test",
-           "2. Best fit copula based on minimum AIC",
-           "3. AIC-weight of that best fit copula",
-           "4. AIC-weight of Normal copula",
-           "5. p-value from goodness of fit test for Cramer-von Mises method",
-           "6. p-value from goodness of fit test for Kolmogorov-Smirnov method",
-           "7. Model averaged (AIC-weights based) lower tail dependence",
-           "8. Model averaged (AIC-weights based) upper tail dependence",
-           "9. Model averaged (AIC-weights based) (lower - upper) tail dependence",
-           "10. Rank of statistics (Pl-Pu) for real data in comparison to 1000 Normal surrogates")
+
+dt <- tibble(Description = c("1. P-value, independence test",
+           "2. Minimum-AIC copula",
+           "3. AIC weight of minimum-AIC copula",
+           "4. AIC weight of N copula",
+           "5. P-value, Cramer-von Mises goodness of fit test",
+           "6. P-value, Kolmogorov-Smirnov goodness of fit test",
+           "7. Model averaged lower-tail dependence",
+           "8. Model averaged upper-tail dependence",
+           "9. Model averaged lower- minus upper-tail dependence",
+           "10. Asymmetry rank in surrogates")
 )
@@
              format="latex",
-             caption = "Summary of results for bivariate data-sets\\label{tab_bivar_summary}",
+             caption = "Summary of Q1 results for bivariate datasets.
+The $p$-values in rows 5-6 and are for the minimum-AIC copula.
+Model averaging used for rows 7-9 was based on AIC weights.
+For row 10, the statistic $P_l-P_u$ was computed for real data and for 1000 surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented.
+\\label{tab_bivar_summary}",
              booktabs=T, linesep = "\\addlinespace",
-             col.names = c(" ", "SoilCN", "Birds’ BMR", "Mammals’ BMR", "Ceder-creek")
+             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar creek data")
              )%>%column_spec(2:5,width="1.4cm")%>%
-             column_spec(1, width = "11cm")%>%kable_styling(latex_options = "basic",position = "center")
+             column_spec(1, width = "8.5cm")%>%kable_styling(latex_options = "basic",position = "center")
 ```
+<!--***DAN: Shya, it says Pl-Pu above, but I suspect it was P_first slice - P_last slice, right?
+Can you please look back at the code, if necessary, and write a comment below this one 
+indicating what results are actually presented in the table?-->
 
 <!--multivariate table summary in latex format-->
-
 ```{r tab_multivar_summaryl, echo=F, results='asis',message=F}
@@
 
-\begin{figure}[!h]
-\textbf{ \hspace{2.8 cm} (A) \hspace{3.2 cm} (B) \hspace{3.2 cm} (C) \hspace{3.2 cm} (D) } \\
-\vspace{-0.6 cm}
+\begin{figure}
 \begin{center}
-\includegraphics[width=4 cm]{./Results/SoilCNraw.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/birds/raw_birds_BMR.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/mammals/raw_mammals_BMR.pdf}
-\includegraphics[width=4 cm]{./Results/Ceder_creek_results/raw_HB_2000.pdf}\\
-\end{center}
-\textbf{ \hspace{2.8 cm} (E) \hspace{3.2 cm} (F) \hspace{3.2 cm} (G) \hspace{3.2 cm} (H) } \\
-\vspace{-0.6 cm}
-\begin{center}
-\includegraphics[width=4 cm]{./Results/SoilCNcop.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/birds/copula_birds_BMR.pdf}
-\includegraphics[width=4 cm]{./Results/BMR_results/mammals/copula_mammals_BMR.pdf}
-\includegraphics[width=4 cm]{./Results/Ceder_creek_results/cop_HB_2000.pdf}\\
-\end{center}
-\textbf{ \hspace{7 cm} (I) \hspace{3.2 cm} (J)} \\
-\vspace{-0.6 cm}
-\begin{center}
-\includegraphics[width=4 cm]{./Results/Aphid_count_raw_loc_4_11.pdf}
-\includegraphics[width=4 cm]{./Results/Aphid_ff_raw_loc_2_5.pdf}\\
-\end{center}
-\textbf{ \hspace{7 cm} (K) \hspace{3.2 cm} (L)} \\
-\vspace{-0.6 cm}
-\begin{center}
-\includegraphics[width=4 cm]{./Results/Aphid_count_cop_loc_4_11.pdf}
-\includegraphics[width=4 cm]{./Results/Aphid_ff_cop_loc_2_5.pdf}
-\caption{Upper panel shows raw data plots for \textbf{(A)} soil C and N data, \textbf{(B)} log(BMR) vs.
-log(mass) for birds, \textbf{(C)} log(BMR) vs.
-log(mass) for mammals, \textbf{(D)} above-ground plant biomass data vs.
-Shannon's index (H) averaged over 9m by 9m plots for year 2000 of Ceder-creek experiment.
-Middle panel \textbf{(E - H)} shows copula plots for the corresponding datasets of the upper panel.
-Lower panel shows \textbf{(I - J)} raw data plots and \textbf{(K - L)} copula plots for  green spruce aphid's count data from location pair(4,11) and for leaf-curling plum aphid's firstflight data from location pair(2,5).\label{fig_biv_multi_raw_cop}}
+\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
+\caption{Upper panels shows raw data plots for (A) soil C and N data, (B) $log_{10}$(basal metabolic rate (BMR)) vs. log(body mass) 
+for birds, (C) log(BMR) vs.
+log(body mass) for mammals, and (D) above-ground plant biomass vs.
+Shannon's index (H) from Cedar Creek.
+Bottom panels show the corresponding copula plots.
+\label{fig_biv_multi_raw_cop}}
 \end{center}
 \end{figure}
+<!--***DAN: Shya, we need to produce a single-pdf figure using R, not use latex to
+cobble together separate pdfs for each panel. Also, I think we can dispense with the 
+plots of the multivariate data. I have created a script that will do both of these things, 
+and inserted a call to it in the sup mat Rmd. Right now it only plots dummy, filler data. 
+Please look at MakeBivarDataPlot.R (that's the script) and search for
+"Shya", and do as it says in those comments, so the actual data are imported into that
+script and plotted. I have changed the above so this new figure is plotted instead of the
+cobbled together old figures.-->
 

### 2018-12-20 SupportingInformation.Rmd
further edits, some tasks for Shyamolina

@@
 author: "S. Ghosh, T. Loecke, C.P. Reid, D.C. Reuman"
-date: "December 16, 2018"
 geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"
@@
 
-updated on `r Sys.Date()`
-
 \pagenumbering{roman}
 \tableofcontents
+\listoffigures
 \listoftables
-\listoffigures
 
@@
 source("./MakeCopulaPedagogFigs.R")
+source("./MakeBivarDataPlot.R")
 ```
@@
 knitr::kable(aphid_info, "latex", booktabs = T, linesep = "",
-      caption = "Species names of 20 aphids commonly found across UK \\label{tab_aphid_info}",
+      caption = "Species names of 20 aphids for which data were available \\label{tab_aphid_info}",
       col.names = c("Common name", "Latin binomial"))
@@
 knitr::kable(plankton_info, "latex", booktabs = T, linesep = "",
-      caption = "Species names of 22 planktons found across UK seas\\label{tab_plankton_info}",
-      col.names = c("Plankton sp. name"))
+      caption = "Species names of 22 plankton taxa for which data were available \\label{tab_plankton_info}",
+      col.names = c("Plankton taxon name"))
 ```
@@
 \begin{center}
-\includegraphics[width=8 cm,height=8 cm]{./Results/SoilCNmapUSA.pdf} 
-\caption{Locations of central pedons with soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil surface) from 5907 sampling sites across the coterminous United States.
-\label{fig_SoilCNmapUSA}}
+\includegraphics[width=16 cm,height=16 cm]{./Results/SoilCNmapUSA.pdf} 
+\caption[Sampling locations for C and N data]{Locations of central pedons sampled 
+for soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil 
+surface) at 5907 sites across the coterminous United States. \label{fig_SoilCNmapUSA}}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=8 cm,height=8 cm]{./Results/aphid_map_plot.pdf} 
-\caption{Locations of Rothamsted Insect Survey suction-trap across UK are indicated by the red circles 
-out of which only blue boxes are considered for our study based on having atleast 30 annual data points. \label{fig_aphid_map}}
+\includegraphics[width=16 cm,height=16 cm]{./Results/aphid_map_plot.pdf} 
+\caption[Sampling locations for aphid data]{Locations of the 
+Rothamsted Insect Survey 
+suction traps from which data were available are indicated by the red circles, 
+out of which only blue boxes had at least 30 years of data and were 
+used for our study. \label{fig_aphid_map}}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=8 cm,height=8 cm]{./Results/plankton_map_plot.pdf} 
-\caption{CPR data collected from 26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ on map) around UK sea (indicated by the red circles) but out of which only blue boxes are considered for our study based on having atleast 45 annual data points.
-\label{fig_plankton_map}}
+\includegraphics[width=16 cm,height=16 cm]{./Results/plankton_map_plot.pdf} 
+\caption[Sampling regions for CPR plankton data]{CPR data were available from 
+26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ regions on the map) around UK 
+seas (red circles), out of which only blue boxes had at least 45 years of data and
+were used for our study. \label{fig_plankton_map}}
 \end{center}
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Normal.pdf}
-\caption{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
-S is Spearman correlation; p is the value of the parameter for the N family (it is a 
-one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
-respectively.\label{figped_N}}
+\caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
+S is Spearman correlation; p is the value of the parameter for the N 
+family (it is a 
+one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively.\label{figped_N}}
 \end{center}
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Clayton.pdf}
-\caption{Log-transformed pdfs (A-E) and samples (F-J) from example C copulas. S is Spearman correlation; p is the value of the parameter for the C family (it is a 
+\caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C copulas. S is Spearman correlation; p is the value of the parameter for the C family (it is a 
 one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvClayton.pdf}
-\caption{Log-transformed pdfs (A-E) and samples (F-J) from example SC 
+\caption[Example SC copulas]{Log-transformed pdfs (A-E) and samples (F-J) 
+from example SC 
 copulas. S is Spearman correlation; p is the value of the parameter for the SC 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
-\caption{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
+\caption[Example G copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
 Spearman correlation; p is the value of the parameter for the G family (it is a 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvGumbel.pdf}
-\caption{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
+\caption[Example SG copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
 copulas. S is Spearman correlation; p is the value of the parameter for the SG 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Joe.pdf}
-\caption{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; p is the value of the parameter for the J family (it is a 
+\caption[Example J copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; p is the value of the parameter for the J family (it is a 
 one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvJoe.pdf}
-\caption{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
+\caption[Example SJ copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
 copulas. K is Spearman correlation; p is the value of the parameter for the SJ 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
-\caption{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; p is the value of the parameter for the F family (it is a 
+\caption[Example F copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; p is the value of the parameter for the F family (it is a 
 one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB1.pdf}
-\caption{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
+\caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
 denote the two parameters of the family (it is a 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
-\caption{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; p1 ad p2 
+\caption[Example SBB1 copulas]{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; p1 ad p2 
 denote the two parameters of the family (it is a 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB6.pdf}
-\caption{Log-transformed pdfs for example BB6 copulas. K is Kendall correlation; p1 and p2
+\caption[Example BB6 copulas]{Log-transformed pdfs for example BB6 copulas. K is Kendall correlation; p1 and p2
 denote the two parameters of the family (it is a 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
-\caption{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; p1 ad p2 
+\caption[Example SBB6 copulas]{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; p1 ad p2 
 denote the two parameters of the family (it is a 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB7.pdf}
-\caption{Log-transformed pdfs for example BB7 copulas. K is Kendall correlation; p1 and p2
+\caption[Example BB7 copulas]{Log-transformed pdfs for example BB7 copulas. K is Kendall correlation; p1 and p2
 denote the two parameters of the family (it is a 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
-\caption{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; p1 ad p2 
+\caption[Example SBB7 copulas]{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; p1 ad p2 
 denote the two parameters of the family (it is a 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB8.pdf}
-\caption{Log-transformed pdfs for example BB8 copulas. K is Kendall correlation; p1 and p2
+\caption[Example BB8 copulas]{Log-transformed pdfs for example BB8 copulas. K is Kendall correlation; p1 and p2
 denote the two parameters of the family (it is a 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
-\caption{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; p1 ad p2 
+\caption[Example SBB8 copulas]{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; p1 ad p2 
 denote the two parameters of the family (it is a 
@@
 \includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
-\caption{Schematic diagram supporting definitions of nonparametric tail dependence statistics. The red and blue lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
+\caption[For definitions of nonparametric statistics]{Schematic diagram supporting definitions of nonparametric tail dependence statistics.
+The red and blue lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
 \end{center}
@@
 knitr::kable(h, #longtable = T, 
-             format = "pandoc",
-      #caption.short = "Fitting info for soil C and N dataset",
-      caption = "Fitting info for **soil C and N dataset**.
+             format = "latex",
+      caption.short = "Fitting info for soil C and N dataset",
+      caption = "Fitting information for **soil C and N dataset**.
 Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=survival Clayton; SG=survival Gumbel; SJ=survival Joe.
@@
 knitr::kable(tab_soilCN_asym, booktabs = T, 
-             format = "pandoc",
+             format = "latex",
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
@@
 knitr::kable(h,
-             format='pandoc',
+             format='latex',
              caption = "Fitting info for **birds' BMR dataset** . \\label{tab_bmr_birds_fit}",
@@
 knitr::kable(tab_bmr_birds_asym,
-             format="pandoc",
+             format="latex",
              caption = "**Birds' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_birds_asym}",
@@
 knitr::kable(h,
-             format='pandoc',
+             format='latex',
              caption = "Fitting info for **mammals' BMR dataset** . \\label{tab_bmr_mammals_fit}",
@@
 knitr::kable(tab_bmr_mammals_asym,
-             format="pandoc",
+             format="latex",
              caption = "**Mammals' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_mammals_asym}",
@@
 knitr::kable(h,
-             format='pandoc',
+             format='latex',
              caption = "Fitting info for **Ceder-creek dataset** for year 2000. \\label{tab_cc2000fit}",
@@
 knitr::kable(tab_cc_asym_2000,
-             format="pandoc",
+             format="latex",
              caption = "**Ceder-creek data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis on year 2000 \\label{tab_cc_asym_2000}",
@@
 knitr::kable(RES_aphid_count$info_ord_copname[,,1], #longtable = T,
-             format='pandoc',
+             format='latex',
              #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid.",
@@
 knitr::kable(RES_aphid_count$info_ord_AICw[,,1], 
-             format='pandoc',
+             format='latex',
              caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_AICw}",
@@
 knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
-             format='pandoc',
+             format='latex',
              caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **abundances of the green spruce aphid**, for comparison with Table \\ref{tab_count_AICw}.
@@
 knitr::kable(RES_aphid_count$gfc_p_CvM,
-             format='pandoc',
+             format='latex',
              caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **abundances of the green spruce aphid**.
@@
 knitr::kable(RES_aphid_count$gfc_p_KS,
-             format='pandoc',
+             format='latex',
              caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **abundances of the green spruce aphid**.
@@
 knitr::kable(RES_aphid_count$LTdep_AICw,
-             format='pandoc',
+             format='latex',
              caption = "Model-averaged lower tail dependence statistic for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_LTdep_AICw}",
@@
 knitr::kable(RES_aphid_count$UTdep_AICw,
-             format='pandoc',
+             format='latex',
              caption = "Model averaged upper tail dependence statistic for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_UTdep_AICw}",
@@
              digits = 5,
-             format='pandoc',
+             format='latex',
              caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_LTmUT}",
@@
 knitr::kable(s,
-             format = "pandoc",
+             format = "latex",
              caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **abundances of the green spruce aphid**.\\label{tab_count_resamp}",
@@
 knitr::kable(RES_aphid_ff$info_ord_copname[,,1], #longtable = T,
-             format='pandoc',
+             format='latex',
              #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
@@
 knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], 
-             format='pandoc',
+             format='latex',
              caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_AICw}",
@@
 knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
-             format='pandoc',
+             format='latex',
              caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **first flight data for the leaf-curlung plum aphid**, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
@@
 knitr::kable(RES_aphid_ff$gfc_p_CvM,
-             format='pandoc',
+             format='latex',
              caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **first flight data for the leaf-curling plum aphid**.
@@
 knitr::kable(RES_aphid_ff$gfc_p_KS,
-             format='pandoc',
+             format='latex',
              caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **first flight data for the leaf-curling plum aphid**.
@@
 knitr::kable(RES_aphid_ff$LTdep_AICw,
-             format='pandoc',
+             format='latex',
              caption = "Model-averaged lower tail dependence statistic for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_LTdep_AICw}",
@@
 knitr::kable(RES_aphid_ff$UTdep_AICw,
-             format='pandoc',
+             format='latex',
              caption = "Model-averaged upper tail dependence statistic for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_UTdep_AICw}",
@@
              digits = 5,
-             format='pandoc',
+             format='latex',
              caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_LTmUT}",
@@
 knitr::kable(s,
-             format = "pandoc",
+             format = "latex",
              caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_resamp}",
@@
 knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1], #longtable=T,
-             format='pandoc', 
+             format='latex', 
              #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundance data for \\textit{Ceratium furca} in UK seas.",
@@
 knitr::kable(h,
-             format="pandoc",
+             format="latex",
              digits=3,
@@
 knitr::kable(s,
-             format = "pandoc",
+             format = "latex",
              caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_resamp}",
@@
 knitr::kable(RES_methane$info_ord_copname[,,1], #longtable=T,
-             format='pandoc',
+             format='latex',
              #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.",
@@
 knitr::kable(s,
-             format = "pandoc",
+             format = "latex",
              caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **methane-flux data**.\\label{tab_methane_resamp}",
@@
 knitr::kable(RES_temp_aphid$info_ord_copname[,,1], #longtable = T,
-             format='pandoc',
+             format='latex',
              #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
@@
 knitr::kable(RES_temp_aphid$info_ord_AICw[,,1], 
-             format='pandoc',
+             format='latex',
              caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_AICw}",
@@
 knitr::kable(normal_cop_AICw_matrix(r=RES_temp_aphid),
-             format='pandoc',
+             format='latex',
              caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **negative average winter temperature data**, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_temp_aphid_AICw_normal}",
@@
 knitr::kable(RES_temp_aphid$gfc_p_CvM,
-             format='pandoc',
+             format='latex',
              caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**.
@@
 knitr::kable(RES_temp_aphid$gfc_p_KS,
-             format='pandoc',
+             format='latex',
              caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**.
@@
 knitr::kable(RES_temp_aphid$LTdep_AICw,
-             format='pandoc',
+             format='latex',
              caption = "Model-averaged lower tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTdep_AICw}",
@@
 knitr::kable(RES_temp_aphid$UTdep_AICw,
-             format='pandoc',
+             format='latex',
              caption = "Model-averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_UTdep_AICw}",
@@
              digits = 5,
-             format='pandoc',
+             format='latex',
              caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTmUT}",
@@
 knitr::kable(summary_stat, booktabs = T, 
-             format = "pandoc",
+             format = "latex",
              digits=3,
@@
 #  temp2<-knitr::kable(temp, booktabs = T, 
-#             format = "pandoc",
+#             format = "latex",
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
@@
 knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
+             format = "latex",
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[1]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_1}"))
+      caption ="A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_1}")
 ```
@@
 knitr::kable(temp, booktabs = T, 
-             format = "pandoc",
+             format = "latex",
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption =paste(cnm[2]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_2}"))
+      caption ="A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_2}")
 ```

### 2018-12-20 MakeBivarDataPlot.R

@@ -1,0 +1,147 @@
+#***plotting dimensions, units inches
+xmarg_ht<-.45
+ymarg_wd<-.45
+gap<-0.1
+tot_wd<-6.5
+pan_wd<-(tot_wd-4*ymarg_wd-4*gap)/4
+pan_ht<-pan_wd
+tot_ht<-2*xmarg_ht+2*pan_ht+2*gap
+pchval<-20
+cexvalpts<-0.25
+cexvaltxt<-0.75
+cexvalpl<-1.5
+pdf(file="./Results/BivarDataPlot.pdf",width=tot_wd,height=tot_ht)
+
+#***top row of panels
+
+#panel 1
+px<-0
+py<-1
+par(fig=c(((px+1)*ymarg_wd+px*pan_wd+px*gap)/tot_wd,
+          ((px+1)*ymarg_wd+(px+1)*pan_wd+px*gap)/tot_wd,
+          ((py+1)*xmarg_ht+py*pan_ht+py*gap)/tot_ht,
+          ((py+1)*xmarg_ht+(py+1)*pan_ht+py*gap)/tot_ht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
+#Shya, please load the appropriate data here as d1, replacing the below 
+#nonsense filler row
+d1<-data.frame(x=1:3,y=1:3)
+d1c<-copula::pobs(d1)
+plot(d1$x,d1$y,type='p',pch=pchval,cex=cexvalpts,
+     col="grey")
+mtext("Soil C",1,1.15)
+mtext("Soil N",2,1.15)
+mtext("A",3,-1.4,cex=cexvalpl,adj=.05)
+
+#panel 2
+px<-1
+py<-1
+par(fig=c(((px+1)*ymarg_wd+px*pan_wd+px*gap)/tot_wd,
+          ((px+1)*ymarg_wd+(px+1)*pan_wd+px*gap)/tot_wd,
+          ((py+1)*xmarg_ht+py*pan_ht+py*gap)/tot_ht,
+          ((py+1)*xmarg_ht+(py+1)*pan_ht+py*gap)/tot_ht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=TRUE)
+#Shya, please load the appropriate data here as d2, replacing the below 
[…]

### 2018-12-20 MakeCopulaPedagogFigs.R

@@ -52,7 +52,8 @@ muvcpp<-function(coplist,numpts,corvals,cortype,thetavals,tot_wd,filename)
       mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
   z<-dCopula(dcarg,coplist[[px+1]],log=FALSE)
   z<-matrix(z,101,101)
-  contour(x,y,log10(z),nlevels=5,xlim=c(0,1),ylim=c(0,1),xaxt="n")
+  contour(x,y,log10(z),nlevels=5,xlim=c(0,1),ylim=c(0,1),xaxt="n",col="gray",labels="")
+  contour(x,y,log10(z),nlevels=5,lty=0,add=TRUE)
   axis(1,labels=FALSE)
   mtext("v",2,1.15)
   mtext("A",3,.05,cex=cexvalpl,adj=0)
@@ -70,7 +71,8 @@ muvcpp<-function(coplist,numpts,corvals,cortype,thetavals,tot_wd,filename)
       mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=TRUE)
   z<-dCopula(dcarg,coplist[[px+1]],log=FALSE)
   z<-matrix(z,101,101)
-  contour(x,y,log10(z),nlevels=5,xlim=c(0,1),ylim=c(0,1),xaxt="n",yaxt="n")
+  contour(x,y,log10(z),nlevels=5,xlim=c(0,1),ylim=c(0,1),xaxt="n",yaxt="n",col="gray",labels="")
+  contour(x,y,log10(z),nlevels=5,lty=0,add=TRUE)
   axis(1,labels=FALSE)
   axis(2,labels=FALSE)
   mtext("B",3,.05,cex=cexvalpl,adj=0)
@@ -88,7 +90,8 @@ muvcpp<-function(coplist,numpts,corvals,cortype,thetavals,tot_wd,filename)
       mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=TRUE)
   z<-dCopula(dcarg,coplist[[px+1]],log=FALSE)
   z<-matrix(z,101,101)
-  contour(x,y,log10(z),nlevels=5,xlim=c(0,1),ylim=c(0,1),xaxt="n",yaxt="n")
+  contour(x,y,log10(z),nlevels=5,xlim=c(0,1),ylim=c(0,1),xaxt="n",yaxt="n",col="gray",labels="")
[…]

### 2018-12-20 Paper.Rmd

@@
 indicate in your response comment what line of what file has the implementation of these
-statistics.-->
+statistics.
+***Shya: In code we used the sum of squared distances divided by n(=number of points lying within two boundaries) which I think okay to be called as "average-squared distance" or (to be precised) average of squared distances of points from 1:1 line.
+Reference : CopulaFunctions_flexible.R file, line 24.-->
 Unlike $\cor_{l_b,u_b}$ and 

### 2018-12-21 Paper.Rmd
further edits

@@
 provide partial information only.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
+\caption{(A-F, I, J) Bivariate datasets showing diverse relationships between 
+variables; many of these datasets nevertheless have the same Pearson (P)
+or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
+obtained from A (respectively, B) via monotonic functions that transformed the standard-
+normal marginal distributions of A and B to gamma (C, D) or 
+uniform (E, F) marginals (details in Appendix \ref{SM-pedagog1_details}). (G, H) Contours 
+of log-scale density functions of Clayton (G) and survival-Clayton 
+(H) copulas with parameter 5 (Appendix \ref{SM-pedagog1_details}). (I,J) Reflections
+of E, F, respectively, about a vertical axis. Subscripts h and v 
+stand for "horizontal" and "vertical". \label{fig_cop_pedag1}}
+\end{center}
+\end{figure}
+<!--***DAN: Dan to go back and add the appendices with details as it says. Make sure
+to mention the shape and scale params of the gamma marginals in C, D were 2-->
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
-\caption{(A-F, I, J) Bivariate datasets showing diverse relationships between 
-variables; many of these datasets nevertheless have the same Pearson (P)
-or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
-obtained from A (respectively, B) via monotonic functions that transformed the standard-
-normal marginal distributions of A and B to gamma (C, D) or 
-uniform (E, F) marginals (details in Appendix \ref{SM-pedagog1_details}). (G, H) Contours 
-of log-scale density functions of Clayton (G) and survival-Clayton 
-(H) copulas with parameter 5 (Appendix \ref{SM-pedagog1_details}). (I,J) Reflections
-of E, F, respectively, about a vertical axis. Subscripts h and v 
-stand for "horizontal" and "vertical". \label{fig_cop_pedag1}}
-\end{center}
-\end{figure}
-<!--***DAN: Dan to go back and add the appendices with details as it says. Make sure
-to mention the shape and scale params of the gamma marginals in C, D were 2-->
-
 A principal benefit of a copula approach to ecological data is 
@@
 <!--***DAN: X and Y to be flled in, can't remember right now what they are and don't want to stop
-to look it up-->
+to look it up, Shya, you could also fill these in if you get a chance-->
 Subsequent model 
@@
 ***Shya: In code we used the sum of squared distances divided by n(=number of points lying within two boundaries) which I think okay to be called as "average-squared distance" or (to be precised) average of squared distances of points from 1:1 line.
-Reference : CopulaFunctions_flexible.R file, line 24.-->
+Reference : CopulaFunctions_flexible.R file, line 24.
+***DAN: OK, good, I am leaving these comments here as a reminder-->
 Unlike $\cor_{l_b,u_b}$ and 
@@
 
-Across essentially all the sub-fields and levels of organization within ecology that we 
-examined, copula structures were typically non-normal and variables exhibited asymmetric tail 
-dependence. To make results easier for readers who are new to copulas to absorb: we first 
+Across essentially all the sub-fields and levels of organization within ecology for which we 
+examined a dataset, copula structures were typically non-normal and variables 
+commonly exhibited asymmetric tail 
+dependence. To make results easier to absorb for readers who are new to copulas: we first 
 present results in detail for an example bivariate dataset; followed by a summary of results
@@
 <!--bivariate example-->
-We will now discuss briefly findings from one of our bivariate datasets as follows.
-As described in the \nameref{Data}, raw data plot for RaCA database soil organic C and soil N stocks to 1m of soil depth from the central pedon of the `r nrow(d)` sites and their corresponding copula plot were presented in Fig.
-\ref{fig_biv_multi_raw_cop}(A) and (E), respectively.
-We carried out model selection analysis and non-parametric statistics to measure tail-dependence (if any) on this copula plot.
-
-<!--#### Model selection approach \label{Ex_SoilCN_Model_selection}-->
-
-From the model selection approach, the test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes` (to within the precision available from `BiCopIndTest` in the `VineCopula` package) that showed a significant correlation (Kendall's coefficient = `r round(BivMS_res_Loecke$TauVal,2)`).
-We then fitted several copulae (all $16$ families as mentioned in \nameref{Model_selection})
-and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulae (Table \ref{SM-tab_soilCNfit}). 
-Based on AIC weights, `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the best fit.
-The table shows that several copulae are much better supported (higher AIC and BIC weights) than the Normal copula, answering the first part of \hyperref[Intro]{\textbf{(Q1)}} for these data.
-
-The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`.
-These values are distinct from what a Normal copula would give (i.e., 0), answering the second part of \hyperref[Intro]{\textbf{(Q1)}} for these data.
-Although these results do convincingly show non-Normal copula structure, we should caution that because even the lowest-AIC copula was a poor fit, giving $p$-value `r BivMS_res_Loecke$GofRes_CvM` (to within the precision available from `BiCopGofTest` in the `VineCopula` package) in the Cramer-von Mises-based test of the goodness of its fit, and $p$-value `r BivMS_res_Loecke$GofRes_KS` in the Kolmogorov-Smirnov-based test of the goodness of fit.
-Therefore, we also verified this finding with \nameref{NPA} approach (see Fig.
-\ref{SM-fig_soilCN_nonparam}), where each of our statistics ($\cor$, $\Ps$, $\Dsq$), as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 Normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig.
+We discuss the example of the soil C and N data (see Data), pictured in 
+Fig. \ref{fig_biv_multi_raw_cop}A (raw data) and E (copula plot). The variables were 
+non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
+from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent
+(Fig. \ref{fig_biv_multi_raw_cop}E).
+The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted the 16 copulas
+mentioned in the previous section to the 
+data pictured in Fig. \ref{fig_biv_multi_raw_cop}E; AIC and BIC weights and corresponding
+lower- and upper-tail statistics for the fitted distributions were computed (Table \ref{SM-tab_soilCNfit}). 
+`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the 
+minimum-AIC copula, with AIC weight `r round(max(BivMS_res_Loecke$InfCritRes$AICw),2)`, whereas the 
+normal copula had AIC weight essentially 0, answering the 
+first part of Q1 for these data. 
+The model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
+and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for model averaging). 
+These values are distinct from what a normal copula would give, namely, 0, answering the second part 
+of Q1 for these data. The values are also substantially 
+distinct from each other, answering the third part of Q1.
+
+<!--***DAN: made it to here 2018 12 20 920pm-->
+
+Although these results do convincingly show non-normal copula structure, 
+we caution that the model-averaged tail-dependence statistics may be biased 
+because even the lowest-AIC copula 
+(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
+was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
+Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
+available from `BiCopGofTest` in the `VineCopula` package). However, nonparametric statistics verified
+both the asymmetry of tail dependence for the soil data and the conclusion that tail dependence 
+in the data was distinct from a normal copula (Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}).
+
+<!--***DAN: the above block of text is the first paart of a para, and i have edited it. The below
+block is the rest of the same para, and I have not-->
+
+where each of our statistics ($\cor$, $\Ps$, $\Dsq$), as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 Normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig.
 \ref{SM-fig_soilCN_nonparam}).
@@
 To be precise, this difference being negative, indicated stronger upper-tail dependence than lower, which was also supported by non-parametric results (see Table \ref{SM-tab_soilCN_asym}).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
+\caption{Upper panels shows raw data plots for (A) soil C and N data, (B) $log_{10}$(basal metabolic rate (BMR)) vs. log(body mass) 
+for birds, (C) log(BMR) vs.
+log(body mass) for mammals, and (D) above-ground plant biomass vs.
+Shannon's index (H) from Cedar Creek.
+Bottom panels show the corresponding copula plots.
+\label{fig_biv_multi_raw_cop}}
+\end{center}
+\end{figure}
+<!--***DAN: Shya, we need to produce a single-pdf figure using R, not use latex to
+cobble together separate pdfs for each panel. Also, I think we can dispense with the 
+plots of the multivariate data. I have created a script that will do both of these things, 
+and inserted a call to it in the sup mat Rmd. Right now it only plots dummy, filler data. 
+Please look at MakeBivarDataPlot.R (that's the script) and search for
+"Shya", and do as it says in those comments, so the actual data are imported into that
+script and plotted. I have changed the above so this new figure is plotted instead of the
+cobbled together old figures.-->
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
-\caption{Upper panels shows raw data plots for (A) soil C and N data, (B) $log_{10}$(basal metabolic rate (BMR)) vs. log(body mass) 
-for birds, (C) log(BMR) vs.
-log(body mass) for mammals, and (D) above-ground plant biomass vs.
-Shannon's index (H) from Cedar Creek.
-Bottom panels show the corresponding copula plots.
-\label{fig_biv_multi_raw_cop}}
-\end{center}
-\end{figure}
-<!--***DAN: Shya, we need to produce a single-pdf figure using R, not use latex to
-cobble together separate pdfs for each panel. Also, I think we can dispense with the 
-plots of the multivariate data. I have created a script that will do both of these things, 
-and inserted a call to it in the sup mat Rmd. Right now it only plots dummy, filler data. 
-Please look at MakeBivarDataPlot.R (that's the script) and search for
-"Shya", and do as it says in those comments, so the actual data are imported into that
-script and plotted. I have changed the above so this new figure is plotted instead of the
-cobbled together old figures.-->
-
 <!--Question 2 analyses-->

### 2018-12-21 SupportingInformation.Rmd
further edits

@@
 #knitr::opts_knit$set(kable.force.latex = TRUE)
-h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LT","UT")]
 knitr::kable(h, #longtable = T, 
              format = "latex",
-      caption.short = "Fitting info for soil C and N dataset",
-      caption = "Fitting information for **soil C and N dataset**.
-Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=survival Clayton; SG=survival Gumbel; SJ=survival Joe.
-An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by $180^\\circ$).\\label{tab_soilCNfit}",
+      caption.short = "Model fitting results for soil C and N dataset",
+      caption = "Results of fitting our 16 copulas to the soil C and N dataset.
+Copula name abbreviations are given in the main text.
+AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the copula family with the fitted parameters; UT = the same for upper-tail dependence.\\label{tab_soilCNfit}",
       booktabs = TRUE)

### 2018-12-21 SupportingInformation.Rmd
fixed minor error

@@
 #knitr::opts_knit$set(kable.force.latex = TRUE)
-h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LT","UT")]
+h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+colnames(h)<-c("copname","AICw","BICw","LT","UT")
 knitr::kable(h, #longtable = T, 

### 2018-12-21 Paper.Rmd
tasks done : plotting in single pdf for Bivariate data

@@
 *normalized ranks* $u_i$ and $v_i$. Ranks were $1$ for the smallest element in a set. 
-Second, we did a test of the independence of the $u_i$ and $v_i$ (implemented in X in the Y package). 
+Second, we did a test of the independence of the $u_i$ and $v_i$ (implemented in `BiCopIndTest` in the `VineCopula` package). 
 <!--***DAN: X and Y to be flled in, can't remember right now what they are and don't want to stop
-to look it up, Shya, you could also fill these in if you get a chance-->
+to look it up, Shya, you could also fill these in if you get a chance
+***Shya: Done.-->
 Subsequent model 
@@
 script and plotted. I have changed the above so this new figure is plotted instead of the
-cobbled together old figures.-->
+cobbled together old figures.
+***Shya: filled in MakeBivDataPlot.R : Done.-->
 
@@
 Can you please look back at the code, if necessary, and write a comment below this one 
-indicating what results are actually presented in the table?-->
+indicating what results are actually presented in the table?
+***Shya: Yes, you are right, here Pl-Pu means P_first slice - P_last slice. In table, item 10 : Assymetry rank is shown for Kendall preserving surroagates.
+Reference : bivfunctionplot.R file, lines 134,135,161-->
 

### 2018-12-21 SupportingInformation.Rmd
tasks done : plotting in single pdf for Bivariate data

@@
 sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
+saveRDS(sr,"./Results/Ceder_creek_results/raw_cop_cc_2000.RDS")
 dev.off()

### 2018-12-21 MakeBivarDataPlot.R

@@ -14,7 +14,7 @@ pdf(file="./Results/BivarDataPlot.pdf",width=tot_wd,height=tot_ht)
 
 #***top row of panels
 
-#panel 1
+#panel 1 : raw data plot for SoilCN
 px<-0
 py<-1
 par(fig=c(((px+1)*ymarg_wd+px*pan_wd+px*gap)/tot_wd,
@@ -22,17 +22,17 @@ par(fig=c(((px+1)*ymarg_wd+px*pan_wd+px*gap)/tot_wd,
           ((py+1)*xmarg_ht+py*pan_ht+py*gap)/tot_ht,
           ((py+1)*xmarg_ht+(py+1)*pan_ht+py*gap)/tot_ht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
-#Shya, please load the appropriate data here as d1, replacing the below 
-#nonsense filler row
-d1<-data.frame(x=1:3,y=1:3)
+d1<-readRDS("./Data/dataForShymolinaCoord.rds")
+d1<-d1[,c("SOCstock100","TSNstock100")]
+d1<-na.omit(d1)
 d1c<-copula::pobs(d1)
-plot(d1$x,d1$y,type='p',pch=pchval,cex=cexvalpts,
+plot(d1$SOCstock100,d1$TSNstock100,type='p',pch=pchval,cex=cexvalpts,
      col="grey")
 mtext("Soil C",1,1.15)
 mtext("Soil N",2,1.15)
 mtext("A",3,-1.4,cex=cexvalpl,adj=.05)
 
-#panel 2
+#panel 2 : raw data plot for Birds' BMR
 px<-1
 py<-1
 par(fig=c(((px+1)*ymarg_wd+px*pan_wd+px*gap)/tot_wd,
@@ -40,17 +40,17 @@ par(fig=c(((px+1)*ymarg_wd+px*pan_wd+px*gap)/tot_wd,
           ((py+1)*xmarg_ht+py*pan_ht+py*gap)/tot_ht,
           ((py+1)*xmarg_ht+(py+1)*pan_ht+py*gap)/tot_ht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=TRUE)
-#Shya, please load the appropriate data here as d2, replacing the below 
-#nonsense filler row. Please make sure you use log10 here.
-d2<-data.frame(x=1:3,y=1:3)
[…]

### 2018-12-22 Paper.Rmd
further edits

@@
 *normalized ranks* $u_i$ and $v_i$. Ranks were $1$ for the smallest element in a set. 
-Second, we did a test of the independence of the $u_i$ and $v_i$ (implemented in `BiCopIndTest` in the `VineCopula` package). 
-<!--***DAN: X and Y to be flled in, can't remember right now what they are and don't want to stop
-to look it up, Shya, you could also fill these in if you get a chance
-***Shya: Done.-->
+Second, we did a test of the independence of the $u_i$ and $v_i$ 
+(implemented in `BiCopIndTest` in the `VineCopula` package). 
 Subsequent model 
@@
 interpretation to $\cor_{l_{b_k},u_{b_k}}(u,v)$; we use both statistics 
-because in some instances $\Ps_{l_b,u_b}$ appears to have more power to
+because $\cor_{l_{b_k},u_{b_k}}(u,v)$ is more familiar, but 
+in some instances $\Ps_{l_b,u_b}$ appears to have more power to
 reveal tail dependence than does $\cor_{l_{b_k},u_{b_k}}(u,v)$. 
@@
 from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent
-(Fig. \ref{fig_biv_multi_raw_cop}E).
+(Fig. \ref{fig_biv_multi_raw_cop}A,E).
 The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted the 16 copulas
@@
 These values are distinct from what a normal copula would give, namely, 0, answering the second part 
-of Q1 for these data. The values are also substantially 
-distinct from each other, answering the third part of Q1.
-
-<!--***DAN: made it to here 2018 12 20 920pm-->
+of Q1 for these data. The values also differed substantially 
+from each other, answering the third part of Q1. The numbers reflect stronger upper- than lower-tail
+dependence for the soil C and N data, and this can be seen visually in the extreme tails 
+of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
+\caption{Upper panels show raw data plots for (A) soil C and N data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
+for birds, (C) $\log_{10}$(BMR) vs.
+$\log_{10}$(body mass) for mammals, and (D) above-ground plant biomass vs.
+Shannon's index (H) from Cedar Creek.
+Bottom panels show the corresponding copula plots.
+\label{fig_biv_multi_raw_cop}}
+\end{center}
+\end{figure}
 
@@
 available from `BiCopGofTest` in the `VineCopula` package). However, nonparametric statistics verified
-both the asymmetry of tail dependence for the soil data and the conclusion that tail dependence 
-in the data was distinct from a normal copula (Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}).
-
-<!--***DAN: the above block of text is the first paart of a para, and i have edited it. The below
-block is the rest of the same para, and I have not-->
-
-where each of our statistics ($\cor$, $\Ps$, $\Dsq$), as computed for the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions of values of the same statistic with the same ranges computed on 1000 Normal-copula surrogate datasets with similar Kendall or Spearman correlations, in different runs for Kendall and Spearman (Fig.
-\ref{SM-fig_soilCN_nonparam}).
-Each bin in each subplot of Fig.
-\ref{SM-fig_soilCN_nonparam} contains a number indicating how many Normal surrogates out of 1000, are different ($95\%$ confidence interval) from Normal-copula statistics.
-Those numbers clearly show that tail dependence of data is stronger in both the lower and upper tails than would be expected under a null hypothesis of a normal copula, answering the second part of \hyperref[Intro]{\textbf{(Q1)}}.
-
-Furthermore, the difference of model-averaged tail dependence statistics (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of \hyperref[Intro]{\textbf{(Q1)}}.
-To be precise, this difference being negative, indicated stronger upper-tail dependence than lower, which was also supported by non-parametric results (see Table \ref{SM-tab_soilCN_asym}).
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
-\caption{Upper panels shows raw data plots for (A) soil C and N data, (B) $log_{10}$(basal metabolic rate (BMR)) vs. log(body mass) 
-for birds, (C) log(BMR) vs.
-log(body mass) for mammals, and (D) above-ground plant biomass vs.
-Shannon's index (H) from Cedar Creek.
-Bottom panels show the corresponding copula plots.
-\label{fig_biv_multi_raw_cop}}
-\end{center}
-\end{figure}
-<!--***DAN: Shya, we need to produce a single-pdf figure using R, not use latex to
-cobble together separate pdfs for each panel. Also, I think we can dispense with the 
-plots of the multivariate data. I have created a script that will do both of these things, 
-and inserted a call to it in the sup mat Rmd. Right now it only plots dummy, filler data. 
-Please look at MakeBivarDataPlot.R (that's the script) and search for
-"Shya", and do as it says in those comments, so the actual data are imported into that
-script and plotted. I have changed the above so this new figure is plotted instead of the
-cobbled together old figures.
-***Shya: filled in MakeBivDataPlot.R : Done.-->
+both that tail dependence was distinct from that of a normal copula for the soil data, and that it 
+was asymmetric (Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}). 
+The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data
+were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates
+of the data,
+separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
+(Fig. \ref{SM-fig_soilCN_nonparam}). Results show that tail dependence of data is stronger in 
+both the lower and upper tails than would be expected under a null hypothesis of a normal copula.
+The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
+$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for real data were also compared to distributions of their values
+on 1000 Kendall- or Spearman-preserving N surrogates, separately in two comparisons for each
+of the $(l_b,u_b)$ ranges listed above (Table \ref{SM-tab_soilCN_asym}). Results showed that
+upper-tail dependence was significantly stronger than lower-tail dependence.
+
+<!--***DAN: made it to here 2018 12 21 at 10pm-->
 
@@
              caption = "Summary of Q1 results for bivariate datasets.
-The $p$-values in rows 5-6 and are for the minimum-AIC copula.
+The $p$-values in rows 5-6 are for the minimum-AIC copula.
 Model averaging used for rows 7-9 was based on AIC weights.
-For row 10, the statistic $P_l-P_u$ was computed for real data and for 1000 surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented.
+For row 10, the statistic $P_{0,u_b}-P_{1-u_b}$ was computed for real data and for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented.
+The value of $u_b$ depended on the size of the dataset, and can be found in Figs.
+\\ref{SM-fig_soilCN_nonparam} to \\ref{SM-fig_cc2000_nonparam} and Tables \\ref{SM-tab_soilCN_asym}, \\ref{SM-tab_bmr_birds_asym}, \\ref{SM-tab_bmr_mammals_asym}, and \\ref{SM-tab_cc_asym_2000}.
 \\label{tab_bivar_summary}",
              booktabs=T, linesep = "\\addlinespace",
-             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar creek data")
+             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar Creek data")
              )%>%column_spec(2:5,width="1.4cm")%>%
@@
 ***Shya: Yes, you are right, here Pl-Pu means P_first slice - P_last slice. In table, item 10 : Assymetry rank is shown for Kendall preserving surroagates.
-Reference : bivfunctionplot.R file, lines 134,135,161-->
+Reference : bivfunctionplot.R file, lines 134,135,161
+***DAN: OK, great. I have made changes to deal with this, but am leaving the comment for now
+as a reminder that we looked at this-->
 

### 2018-12-22 SupportingInformation.Rmd
further edits

@@
 source("./MakeCopulaPedagogFigs.R")
-source("./MakeBivarDataPlot.R")
 ```
@@
 dev.off()
+```
+
+```{r make_data_plot_fig,echo=FALSE,results="hide"}
+source("./MakeBivarDataPlot.R")
 ```
@@
       caption.short = "Model fitting results for soil C and N dataset",
-      caption = "Results of fitting our 16 copulas to the soil C and N dataset.
+      caption = "Results of fitting our 16 copula families to the soil C and N dataset.
 Copula name abbreviations are given in the main text.
-AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the copula family with the fitted parameters; UT = the same for upper-tail dependence.\\label{tab_soilCNfit}",
+AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.\\label{tab_soilCNfit}",
       booktabs = TRUE)
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
-\caption[Testing non-parametric statistics for soilCN data]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{soil C and N data}.
-As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
-When red cross was above (or below) the blue and green cross, text at the top of plots indicates the actual statistics was greater than (or less than) by those number of surrogate values.
-When $\cor$ values (respectively, $\Ps$ or $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, stronger than or weaker than for $\Ps$ or $\Dsq$) expected from a normal-copula null hypothesis.
+\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N copula structure, for soil C and N data.
+As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
+Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
+For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates.
+When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level).
+When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.
 \label{fig_soilCN_nonparam}}
@@
 \end{figure}
+<!--***DAN: Shya, can you please adjust this figure and the analogous 3 other figures for the other three bivariate datasets as follows: replace the y-axis label "Partial correlation" with "cor", for consistency with the other y-axis labels-->
 
@@
 tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
+tab_soilCN_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
 knitr::kable(tab_soilCN_asym, booktabs = T, 
              format = "latex",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption = "**Soil CN data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.
-Three statistics are listed in the left column that measured asymmetry of tail dependence by comparing the distribution ranges $0-0.1$ and $0.9-1$ (see \\nameref{MT-NPA}).
-One thousand Kendall-preserving and 1000 Spearman-preserving surrogates with normal copula structure were created, and the same statistics were computed for the surrogates.
-A table entry $<N$ indicates the value of the given statistic on the data was less than its value on $N$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
+      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
+      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
+The values of the listed statistics (corFS-corLS $=\\cor_{0,0.1}-\\cor_{0.9,1}$, PFS-PLS $=\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D2LS-D2FS $=\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons.
+A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
 #kable_styling(, position = "float_left")
@@
 h<-BivMS_birds$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+colnames(h)<-c("copname","AICw","BICw","LT","UT")
 knitr::kable(h,
              format='latex',
-             caption = "Fitting info for **birds' BMR dataset** . \\label{tab_bmr_birds_fit}",
+             caption.short = "Model fitting results for the bird body mass and BMR dataset",
+             caption = "Results of fitting our 16 copula families to the bird body mass and BMR dataset.
+Analogous to Fig.
+\\ref{tab_soilCNfit}, see caption of that figure for details.
+\\label{tab_bmr_birds_fit}",
              booktabs=T)
@@
 \includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
-\caption[Testing non-parametric statistics for birds' BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{birds' BMR data}.
-As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
-When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
+\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N copula structure, for bird body mass and BMR data.
+Same format as Fig.
+\ref{fig_soilCN_nonparam}, see caption of that figure.
 \label{fig_bmr_birds_nonparam}}
@@
 tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
+tab_bmr_birds_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
 knitr::kable(tab_bmr_birds_asym,
              format="latex",
-             caption = "**Birds' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_birds_asym}",
+             caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
+             caption = "Bird body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
+Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_bmr_birds_asym}",
              booktabs=T)
@@
 
-<!--Display model selection results for BMR:birds data in a table-->
+<!--Display model selection results for BMR:mammals data in a table-->
 ```{r tab_bmr_mammals_fit, echo=F, results='markup'}
 h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+colnames(h)<-c("copname","AICw","BICw","LT","UT")
 knitr::kable(h,
              format='latex',
-             caption = "Fitting info for **mammals' BMR dataset** . \\label{tab_bmr_mammals_fit}",
+             caption.short = "Model fitting results for the mammal body mass and BMR dataset",
+             caption = "Results of fitting our 16 copula families to the mammal body mass and BMR dataset.
+Analogous to Fig.
+\\ref{tab_soilCNfit}, see caption of that figure for details.
+\\label{tab_bmr_mammals_fit}",
              booktabs=T)
@@
 \includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
-\caption[Testing non-parametric statistics for mammals' BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{mammals' BMR data}.
-As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
-When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
+\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N copula structure, for mammal body mass and BMR data.
+Same format as Fig.
+\ref{fig_soilCN_nonparam}, see caption of that figure.
 \label{fig_bmr_mammals_nonparam}}
@@
 tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
+tab_bmr_mammals_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
 knitr::kable(tab_bmr_mammals_asym,
              format="latex",
-             caption = "**Mammals' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_bmr_mammals_asym}",
+             caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
+             caption = "Mammal body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
+Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.
+\\label{tab_bmr_mammals_asym}",
              booktabs=T)
@@
 h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+colnames(h)<-c("copname","AICw","BICw","LT","UT")
 knitr::kable(h,
              format='latex',
-             caption = "Fitting info for **Ceder-creek dataset** for year 2000. \\label{tab_cc2000fit}",
+             caption.short = "Model fitting results for the Cedar Creek dataset",
+             caption = "Results of fitting our 16 copula families to the Cedar Creek dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_cc2000fit}",
              booktabs=T)
@@
 \includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_10_bivfunctionplot.pdf}
-\caption[Testing non-parametric statistics for Ceder-creek data with $numbin=10$ for year 2000]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{Ceder-creek data}.
-As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).
-When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than.
-\label{fig_cc2000_nonparam}}
+\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N copula structure, for Cedar Creek data.
+Same format as Fig.
+\ref{fig_soilCN_nonparam}, see caption of that figure.\label{fig_cc2000_nonparam}}
 \end{center}
@@
 tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
+tab_cc_asym_2000[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
 knitr::kable(tab_cc_asym_2000,
              format="latex",
-             caption = "**Ceder-creek data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis on year 2000 \\label{tab_cc_asym_2000}",
+             caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
+             caption = "Cedar Creek dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
+Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_cc_asym_2000}",
              booktabs=T)

### 2018-12-22 SupportingInformation.Rmd
modified

@@
 \end{figure}
-<!--***DAN: Shya, can you please adjust this figure and the analogous 3 other figures for the other three bivariate datasets as follows: replace the y-axis label "Partial correlation" with "cor", for consistency with the other y-axis labels-->
+<!--***DAN: Shya, can you please adjust this figure and the analogous 3 other figures for the other three bivariate datasets as follows: replace the y-axis label "Partial correlation" with "cor", for consistency with the other y-axis labels.
+***Shya: Done.-->
 

### 2019-01-03 SupportingInformation.Rmd
comment out extinction risk section

@@
 
-source("mydispersal.R")
-set.seed(seed)
-
-rlist<-c(0,0.05,0.25,0.5,0.7,0.75,0.9,1.0,1.25,1.5,1.75,1.95)
+#source("mydispersal.R")
+#set.seed(seed)
+
+#rlist<-c(0,0.05,0.25,0.5,0.7,0.75,0.9,1.0,1.25,1.5,1.75,1.95)
 
 # plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-pdf("./Results/ext_risk_local_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
-for(r in rlist){
+#pdf("./Results/ext_risk_local_disp.pdf",heigh=6,width=10)
+#op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
+#for(r in rlist){
     #cat("r=",r,"\n")
-    s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
-    mtext(paste0("r = ", r))
-}
-par(op)
-dev.off()
+#    s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
+#    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+#    lines(s$d_seq,s$risk_right,type="b",col="blue")
+#    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+#    mtext(paste0("r = ", r))
+#}
+#par(op)
+#dev.off()
 
 # plot for global dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-pdf("./Results/ext_risk_global_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
-for(r in rlist){
+#pdf("./Results/ext_risk_global_disp.pdf",heigh=6,width=10)
+#op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
+#for(r in rlist){
   #cat("r=",r,"\n")
-  s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
-  plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-  lines(s$d_seq,s$risk_right,type="b",col="blue")
-  legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
-  mtext(paste0("r = ", r))
-}
-par(op)
-dev.off()
+#  s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
+#  plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+#  lines(s$d_seq,s$risk_right,type="b",col="blue")
+#  legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+#  mtext(paste0("r = ", r))
+#}
+#par(op)
+#dev.off()
 
 # plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
-pdf("./Results/ext_risk_numlocs_effect_local_disp.pdf",heigh=6,width=6)
-op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
-for(r in c(0,0.25)){
+#pdf("./Results/ext_risk_numlocs_effect_local_disp.pdf",heigh=6,width=6)
+#op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
+#for(r in c(0,0.25)){
   #cat("r=",r,"\n")
-  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
-  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=F,ploton=F)
-
-  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
-  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
-       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
-  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+#  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
+#  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=F,ploton=F)
+
+#  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+#  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
+#  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+#  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
+#       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+#  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
+#  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
   #mtext(paste0("B",1))
-}
-par(op)
-dev.off()
+#}
+#par(op)
+#dev.off()
 
 # plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
-pdf("./Results/ext_risk_numlocs_effect_global_disp.pdf",heigh=6,width=6)
-op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
-for(r in c(0,0.25)){
+#pdf("./Results/ext_risk_numlocs_effect_global_disp.pdf",heigh=6,width=6)
+#op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
+#for(r in c(0,0.25)){
   #cat("r=",r,"\n")
-  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
-  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=T,ploton=F)
-
-  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
-  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
-       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
-  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+#  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
+#  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=T,ploton=F)
+
+#  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+#  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
+#  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+#  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
+#       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+#  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
+#  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
   #mtext(paste0("B",1))
-}
-par(op)
-dev.off()
+#}
+#par(op)
+#dev.off()
 
 # plot to see the effect of K for r=0.25 (Ricker model)
-pdf("./Results/ext_risk_K_effect.pdf",heigh=6,width=6)
-op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
-
-s1<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=F,ploton=F)
-s2<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=F,ploton=F)
-
-s3<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=T,ploton=F)
-s4<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=T,ploton=F)
-
-plot(s1$d_seq,s1$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-lines(s2$d_seq,s2$risk_left,type="b",col="orange")
-legend("topleft",title=("Noise : left tail dep., local disp."), c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-plot(s1$d_seq,s1$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-lines(s2$d_seq,s2$risk_right,type="b",col="skyblue")
-legend("topleft",title=("Noise : right tail dep., local disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
-
-plot(s3$d_seq,s3$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-lines(s4$d_seq,s4$risk_left,type="b",col="orange")
-legend("topleft",title= ("Noise : left tail dep., global disp."),c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-plot(s3$d_seq,s3$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-lines(s4$d_seq,s4$risk_right,type="b",col="skyblue")
-legend("topleft",title=("Noise : right tail dep., global disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
-
-par(op)
-dev.off()
+#pdf("./Results/ext_risk_K_effect.pdf",heigh=6,width=6)
+#op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
+
+#s1<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=F,ploton=F)
+#s2<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=F,ploton=F)
+
+#s3<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=T,ploton=F)
+#s4<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=T,ploton=F)
+
+#plot(s1$d_seq,s1$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+#lines(s2$d_seq,s2$risk_left,type="b",col="orange")
+#legend("topleft",title=("Noise : left tail dep., local disp."), c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+#plot(s1$d_seq,s1$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+#lines(s2$d_seq,s2$risk_right,type="b",col="skyblue")
+#legend("topleft",title=("Noise : right tail dep., local disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+
+#plot(s3$d_seq,s3$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+#lines(s4$d_seq,s4$risk_left,type="b",col="orange")
+#legend("topleft",title= ("Noise : left tail dep., global disp."),c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
+
+#plot(s3$d_seq,s3$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
+#lines(s4$d_seq,s4$risk_right,type="b",col="skyblue")
+#legend("topleft",title=("Noise : right tail dep., global disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
+
+#par(op)
+#dev.off()
 ```

### 2019-01-06 ext_risk_copula.Rmd
variation of b instead of r for Msmith model

---
title: "ext_risk_copula"
author: "Shyamolina Ghosh"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
seed<-101
source("mtime.R")
source("mydispersal.R")
```

#Ricker model :
```{r fig_ricker, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)

rlist<-c(0,0.05,0.25,0.5,0.7,0.75,0.9,1.0,1.25,1.5,1.75,1.95)

# plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
pdf("./Results/ext_risk_copula/ricker_ext_risk_local_disp.pdf",heigh=6,width=10)
op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
for(r in rlist){
    #cat("r=",r,"\n")
    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=50,a=NA,b=NA,model="ricker",disp_everywhere=F,ploton = F)
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
    mtext(paste0("r = ", r))
}
par(op)
dev.off()

# plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
pdf("./Results/ext_risk_copula/ricker_ext_risk_global_disp.pdf",heigh=6,width=10)
op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
for(r in rlist){
    #cat("r=",r,"\n")
    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=50,a=NA,b=NA,model="ricker",disp_everywhere=T,ploton = F)
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
    mtext(paste0("r = ", r))
}
par(op)
dev.off()
```

![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp}](./Results/ext_risk_copula/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }

![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp}](./Results/ext_risk_copula/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }

#Hassell Model :

```{r fig_hassell, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)

rlist<-seq(from=6,to=11,by=0.5)
a<-0.5
b<-1.2
K_e<-((r^(1/b))-1)/a
rc<-((b/(b-1)))^b # critical value of r from monotonic to osc steady state approach transition

# plot for local dispersal in D matrix 
pdf("./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf",heigh=6,width=10)
op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
for(r in rlist){
    #cat("r=",r,"\n")
    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=K_e,a=a,b=b,model="hassell",disp_everywhere=F,ploton = F)
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
    mtext(paste0("r = ", r))
}
par(op)
dev.off()

# plot for local dispersal in D matrix 
pdf("./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf",heigh=6,width=10)
op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
for(r in rlist){
    #cat("r=",r,"\n")
    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=K_e,a=a,b=b,model="hassell",disp_everywhere=T,ploton = F)
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
    mtext(paste0("r = ", r))
}
par(op)
dev.off()

# plot for local dispersal in D matrix 
pdf("./Results/ext_risk_copula/hassell_ext_risk_global_disp2.pdf",heigh=6,width=10)
op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
for(r in rlist){
    #cat("r=",r,"\n")
    s<-varying_d(numsims=10000,numsteps =10,numlocs=15,r=r,K=K_e,a=a,b=b,model="hassell",disp_everywhere=T,ploton = F)
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
    mtext(paste0("r = ", r))
}
par(op)
dev.off()
```

![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Hassell model : a = 0.5, b = 1.2) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }

![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Hassell model : a = 0.5, b = 1.2) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }

#Maynard Smith Model :

```{r fig_msmith, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
set.seed(seed)

blist<-seq(from=1,to=2,by=0.2)
r<-5
b<-1.2
K_e<-((r-1)^(1/b))/a

# plot for local dispersal in D matrix 
pdf("./Results/ext_risk_copula/msmith_ext_risk_local_disp.pdf",heigh=6,width=10)
op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
for(b in blist){
    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=K_e,a=a,b=b,model="msmith",disp_everywhere=F,ploton = F)
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend(x=-0.06,y=0.4, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
    mtext(paste0("b = ", b))
}
par(op)
dev.off()

# plot for global dispersal in D matrix 
pdf("./Results/ext_risk_copula/msmith_ext_risk_global_disp.pdf",heigh=6,width=10)
op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
for(b in blist){
    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=K_e,a=a,b=b,model="msmith",disp_everywhere=T,ploton = F)
    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
    lines(s$d_seq,s$risk_right,type="b",col="blue")
    legend(x=-0.06,y=0.4, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
    mtext(paste0("b = ", b))
}
par(op)
dev.off()
```

![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r = 5) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp_msmith}](./Results/ext_risk_copula/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }

![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r=5) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp_msmith}](./Results/ext_risk_copula/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }

### 2019-01-06 1var_pop_models.R

@@ -49,12 +49,29 @@ Hassell<-function(r,a,b,p0,lensim){
 }
 
 #call the function
-Hassell(r=5, a=0.5, b=5, p0=0.1, lensim=200) #b~1.5 : contest, b~100 : scramble
+Hassell(r=6, a=0.5, b=2, p0=0.1, lensim=200) #b~1.5 : contest, b~100 : scramble
 
+Hassell(r=5, a=0.5, b=2.5, p0=0.1, lensim=200) # see change : r=5,a=0.5 and vary b 1.5 to 5.5
 #----------------------------------------------------------------------------------------------------
 # Exploring Maynard Smith model
+Msmith<-function(r,a,b,p0,lensim){
+  pop<-c(p0)
+  time<-c(0)
+  
+  for(it in c(1:lensim)){
+    pt<-(r*p0)/(1+((a*p0)^b))
+    time<-c(time,it)
+    pop<-c(pop,pt)
+    p0<-pt
+  }
+  
+  plot(time,pop,type="b")
+  
+}
 
-
+#call the function
+Msmith(r=2,a=0.5,b=2,p0=0.1,lensim=200) #see changes as r=5,a=0.5 and vary b : 1,1.5,2,4
+                                          #see changes as a=0.5,b=4 and vary r : 1.2, 1.5, 2
 
 
 

### 2019-01-06 mydispersal.R

@@ -46,7 +46,9 @@ Dispersal_mat<-function(numlocs,d,disp_everywhere){
 #p0       A length numlocs vector holding initial populations
 #ns       An numsims by numlocs by numsteps array of the epsilons, 
 #           where numsteps is the number of time steps you want
-#
+#D        Dispersal matrix
+#r,K,a,b  model parameters
+#model    a character specifying model name
 #--------Output------------------------------------------------------
 #A numsims by numlocs by numsteps+1 array of populations
 #----------------------------------------------------------------------
@@ -69,12 +71,15 @@ popsim_ml_D<-function(p0,ns,D,r,K,a,b,model){
     
     lam_ricker<-exp(r*(1-(res[,,tct]/K)))
     lam_hassell<-r/((1+res[,,tct])^b)
+    lam_msmith<-r/(1+((res[,,tct])^b))
     
     #growth prior to dispersal
     if(model=="ricker"){
       given_model<-lam_ricker
     }else if(model=="hassell"){
       given_model<-lam_hassell
+    }else if(model=="msmith"){
+      given_model<-lam_msmith
     }else{
       warning("model not specified",immediate.=T,call.=T)
     }
@@ -188,32 +193,3 @@ varying_d<-function(numsims,numsteps,numlocs,r,K,a,b,model,disp_everywhere,ploto
 
 }
 #------------------------------------------------------------------------------------------------------------
-# For LC/Ricker
-#set.seed(seed=101)
-#r<-1
-#a<-NA
-#b<-NA
-#varying_d(numsims=10000,numsteps=10,numlocs=5,r=r,K=50,a=a,b=b,model="ricker",disp_everywhere=T,ploton = T)
-
-
-# For Hassell
-#set.seed(seed=101)
-#r<-6
-#a<-0.5
-#b<-1.2
-#K_e<-((r^(1/b))-1)/a
-#K_e
-#rc<-((b/(b-1)))^b
-#rc
-#varying_d(numsims=10000,numsteps=10,numlocs=5,r=r,K=K_e,a=a,b=b,model="hassell",disp_everywhere=T,ploton = T)
-
-
-
-
-
-
-
-
-
-
-

### 2019-01-06 ext_risk_copula.Rmd
modified

@@
 
-rlist<-seq(from=6,to=11,by=0.5)
+rlist<-seq(from=5,to=10,by=1)
 a<-0.5
-b<-1.2
+b<-1.5
 K_e<-((r^(1/b))-1)/a
@@
 pdf("./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
+op<-par(mfrow=c(2,4),mar=c(5,4,1,1) + 0.2)
 for(r in rlist){
@@
 pdf("./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
+op<-par(mfrow=c(2,4),mar=c(5,4,1,1) + 0.2)
 for(r in rlist){
@@
 dev.off()
-
-# plot for local dispersal in D matrix 
-pdf("./Results/ext_risk_copula/hassell_ext_risk_global_disp2.pdf",heigh=6,width=10)
-op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
-for(r in rlist){
-    #cat("r=",r,"\n")
-    s<-varying_d(numsims=10000,numsteps =10,numlocs=15,r=r,K=K_e,a=a,b=b,model="hassell",disp_everywhere=T,ploton = F)
-    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-    lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
-    mtext(paste0("r = ", r))
-}
-par(op)
-dev.off()
 ```
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Hassell model : a = 0.5, b = 1.2) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Hassell model : a = 0.5, b = 1.5, rc=5.19) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Hassell model : a = 0.5, b = 1.2) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Hassell model : a = 0.5, b = 1.5, rc=5.19) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
 
@@
 r<-5
-b<-1.2
+a<-0.5
 K_e<-((r-1)^(1/b))/a
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r = 5) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp_msmith}](./Results/ext_risk_copula/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r = 5, bc=1.4) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp_msmith}](./Results/ext_risk_copula/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r=5) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp_msmith}](./Results/ext_risk_copula/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r=5, bc=1.4) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp_msmith}](./Results/ext_risk_copula/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }

### 2019-01-06 1var_pop_models.R

@@ -49,9 +49,9 @@ Hassell<-function(r,a,b,p0,lensim){
 }
 
 #call the function
-Hassell(r=6, a=0.5, b=2, p0=0.1, lensim=200) #b~1.5 : contest, b~100 : scramble
+Hassell(r=5, a=0.5, b=2.5, p0=0.1, lensim=200) #b~1.5 : contest, b~100 : scramble
 
-Hassell(r=5, a=0.5, b=2.5, p0=0.1, lensim=200) # see change : r=5,a=0.5 and vary b 1.5 to 5.5
+Hassell(r=4, a=0.5, b=5.1, p0=0.1, lensim=200) # see change : r=5,a=0.5 and vary b 1.5 to 5.5
 #----------------------------------------------------------------------------------------------------
 # Exploring Maynard Smith model
 Msmith<-function(r,a,b,p0,lensim){
@@ -70,7 +70,7 @@ Msmith<-function(r,a,b,p0,lensim){
 }
 
 #call the function
-Msmith(r=2,a=0.5,b=2,p0=0.1,lensim=200) #see changes as r=5,a=0.5 and vary b : 1,1.5,2,4
+Msmith(r=5,a=0.5,b=1.4,p0=0.1,lensim=200) #see changes as r=5,a=0.5 and vary b : 1,1.5,2,4
                                           #see changes as a=0.5,b=4 and vary r : 1.2, 1.5, 2
 
 

### 2019-01-07 ext_risk_copula.Rmd
added description for ricker model

@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp}](./Results/ext_risk_copula/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_ricker_risk_local_disp}](./Results/ext_risk_copula/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp}](./Results/ext_risk_copula/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/ext_risk_copula/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
+
+<!--
+In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
+
+\begin{equation}\label{eq.extrisk_model}
+P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
+\end{equation}
+
+$r$ is the liner growth rate and $K$ is the carrying capacity of the environment.
+$\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence.
+We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion).
+We simulated Eq.
+\ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $10$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;
+
+- LC model (r=0) :
+
+    * For LC model, as expected, when extinction risk calculated after $10$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
+[see first subplot of Fig.
+\ref{fig_ricker_risk_local_disp} and Fig.
+\ref{fig_ricker_risk_global_disp}]
+
+- Ricker model (r $\neq$ 0):
+
+    * For ricker model when extinction risk calculated after $10$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
+For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations (rc =1, under- to over-compensatory dynamics).
+[see Fig.
+\ref{fig_rikcer_risk_local_disp} - \ref{fig_ricker_risk_global_disp}]
+
+    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. (produce the figs if not done yet)
+
+With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig.
+\ref{fig_risk_numlocs_effect_local_disp} - Fig.
+\ref{fig_risk_numlocs_effect_global_disp}] for both model.
+(produce the figs if not done yet)
+-->
 
@@
     lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+    legend(x=-0.06,y=0.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
     mtext(paste0("r = ", r))

### 2019-01-09 Paper.Rmd
further edits to Q1 materials

@@
 For instance, datasets showing markedly different 
-relationships may nevertheless have the same correlation coefficient value;  
-Fig. \ref{fig_cop_pedag1}A-B shows two such datasets. The variables of panel A 
-(respectively, B) are principally 
-related in the left (respectively, right) portions of their distributions, an asymmetric pattern of dependence in the distribution tails that may
+relationships may nevertheless have the same correlation coefficient value 
+(Fig. \ref{fig_cop_pedag1}A-B). The variables of Fig. \ref{fig_cop_pedag1}A 
+(respectively, Fig. \ref{fig_cop_pedag1}B) are principally 
+related in the left (respectively, right) portions of their distributions, an asymmetric 
+pattern of dependence in the distribution tails that may
 have ecological significance, as discussed below, but is not captured by 
@@
 (Appendix \ref{SM-pedagog1_details} for details),
-<!--***DAN: Dan to go back and write the appendix-->
 which alters the Pearson, but not the Spearman correlation. Although correlation
@@
 or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
-obtained from A (respectively, B) via monotonic functions that transformed the standard-
-normal marginal distributions of A and B to gamma (C, D) or 
+obtained from A (respectively, B) via monotonic functions that transformed the 
+standard-normal marginal distributions of A and B to gamma (C, D) or 
 uniform (E, F) marginals (details in Appendix \ref{SM-pedagog1_details}). (G, H) Contours 
 of log-scale density functions of Clayton (G) and survival-Clayton 
-(H) copulas with parameter 5 (Appendix \ref{SM-pedagog1_details}). (I,J) Reflections
+(H) copulas with parameter 5 (see section on methods for Q1). (I,J) Reflections
 of E, F, respectively, about a vertical axis. Subscripts h and v 
@@
 \end{figure}
-<!--***DAN: Dan to go back and add the appendices with details as it says. Make sure
-to mention the shape and scale params of the gamma marginals in C, D were 2-->
 
@@
 or copula structure is revealed by the bivariate plot of $u_i$ versus $v_i$, where
-$u_i$ is the rank of $x_i$ in the set of $x_j$ ($j=1,\ldots,n$), divided by
+$u_i$ is the rank of $x_i$ in the set of the $x_j$ ($j=1,\ldots,n$), divided by
 $n+1$; and $v_i$ is defined in the same way but using the $y_i$. See, 
@@
 corelation can have a range of tail dependence 
-characteristics (Fig. \ref{fig_cop_pedag}), which can be quantified (Methods).
+characteristics (Fig. \ref{fig_cop_pedag}), which can be quantified (see section on
+methods for Q1).
 
@@
 \includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
-\caption{(A-C) Data generated using the BB1 copula (Methods) with three different 
+\caption{(A-C) Data generated using the BB1 copula (see section on methods for Q1) 
+with three different 
 sets of parameters chosen to produce a variety of 
-lower-tail (LT) and upper-tail (UT) dependence strengths (Methods), but the same 
-Kendall correlation (KC) (to two digits) in all cases. (D-F) Similar to A-C, but
+lower-tail (LT) and upper-tail (UT) dependence strengths (methods for Q1), but the same 
+Kendall correlation (K) (to two digits) in all cases. (D-F) Similar to A-C, but
 with stronger correlation. Asymmetry of tail dependence (LT-UT) is the same 
@@
 \end{figure}
+<!--***DAN: Shya, I changed the abbreviation for Kendall Correlation from KC to K in the 
+caption, please change in the code that generates the figures, as well-->
 
@@
 of the distributions that dominate the overall association is visible in the copula
-approach, but may not be reflected by correlation and other standard techniques.
+approach, but may not be reflected by correlation coefficients and other standard techniques.
 
@@
 ecological research is that such an approach may help illuminate 
-causal mechanisms by which variables are related. For instance, if two species
+the causal mechanisms through which variables are related. For instance, if two species
 X and Y are strong competitors, we may expect abundances of the two species
@@
 (this is called \emph{spatial synchrony}) which are influenced by the environmental 
-variable. This is called the \emph{Moran effect}. 
+variable. This process is called the \emph{Moran effect}. 
 If copula structure is transmitted, 
@@
 shortages. On the other hand left-tail dependence in the pest species 
-and right-tail dependence in the crop, even if overall correlation were 
+or right-tail dependence in the crop, even if overall correlation were 
 unchanged, would have much more benign effects.
 
-We approached our overall goal of illustrating the potential of 
+We approached our overall goal of exploring the potential of 
 copula approaches for understanding dependences in ecology by addressing the 
 folling specific questions. (Q1) Do datasets in ecology and related fields 
-have copula structure distinct from a Gaussian/normal distribution (here called 
+have copula structure distinct from that of a Gaussian/normal distribution (here called 
 non-normal copula structure)? Do they show tail dependence distinct from that 
@@
 explore mechanisms. The seventh and eigth sections, which address Q3, use both
-the data section two as well as models to explore a variety of consequences
+the data of section two as well as models to explore a variety of consequences
 of copula structure and tail dependence for applications and for our understanding
@@
 research. We argue that the benefits to ecological research of a complete statistical 
-desription of dependence will be substantial, and we hope this paper stimulates the
-use of copula approaches across the discipline.
+desription of dependence will be substantial, and we hope this paper helps stimulate the
+fruitful use of copula approaches across the discipline.
 
@@
 and asymmetry of tail dependence. 
-Conceptually, a bivariate copula is the same as a probability density function 
+Conceptually, a bivariate copula is essentially a probability density function 
 on the unit square with uniform maginals. The normal copula family is a one-parameter
-family of copulas, with parameter corresponding to the covariance of the related 
+family of copulas, with parameter corresponding to the correlation of the related 
 bivariate normal distribtion, and controlling the degree of dependence 
@@
 R statistical computing language.
-Additional details are in Appendix \ref{SM-copula_details} and in introductory texts 
+Additional details are in introductory texts 
 on copulas [@nelsen2006_copula ; @joe2014_dependence] and in `VineCopula` documentation.
@@
 parameter $10$. See also Fig. \ref{fig_cop_pedag1}E-H for the Clayton and  
-survival-Clayton copulas with parameter 5, and Appendix \ref{SM-copula_details}
-for additional details. \label{fig_cop_pedag3}}
+survival-Clayton copulas with parameter 5. Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}
+have visualizations of all the copulas we used, including more visualizations
+of the Clayton and survival-Clayton copulas.\label{fig_cop_pedag3}}
 \end{center}
@@
 selection algorithms are ineffective if data could not be distinguished from independent data 
-since most or all bivariate copula families considered include the independent copula in the 
+since most or all bivariate copula families we considered include the independent copula in the 
 family or as a boundary case. If independence could be rejected ($0.05$ significance level), model 
@@
 two tests, one based on a Cramer-von Mises statistic, and one based on a 
-Kolmogorov-Smirnov statistic. For keep computation times reasonable, 
+Kolmogorov-Smirnov statistic. To keep computation times reasonable, 
 an initial run using $100$ bootstraps was performed, and if
@@
 statistics. The statistics were formulated with positively associated
-variables in mind. All the variables we study are positively associated
-when they are significantly associated. So definitions and associated 
+variables in mind. All the variables we study were positively associated
+when they were significantly associated. So definitions and associated 
 diagrams will assume positive associations, though generalizations are 
@@
 
-We also defined the statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
+We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
 which has a similar 
@@
 reveal tail dependence than does $\cor_{l_{b_k},u_{b_k}}(u,v)$. 
-<!--***DAN: need to make this a bit more precise, perhaps just enough to cite one of the supp mat sections where tests are done of these stats, but you need to review the tests first and what they show-->
 Our third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
@@
 tail dependence known to be present in data simulated from various copulas 
-(Appendix \ref{SM-Test_npa_stat}).
+(Appendix \ref{SM-Test_npa_stat}, Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}).
 
@@
 what would be expected from a null hypothesis of normal copula structure but 
-unchanged marginals and unchanged 
+unchanged marginals and similar 
 Kendall (or Spearman)
@@
 dependence. To make results easier to absorb for readers who are new to copulas: we first 
-present results in detail for an example bivariate dataset; followed by a summary of results
+present results in detail for an example bivariate dataset; then we summarize results
 for all bivariate datasets; then we present results in 
-detail for an example multivariate dataset; followed by a summary of results 
+detail for an example multivariate dataset; then we summarize results 
 for all multivariate datasets.
@@
 <!--bivariate example-->
-We discuss the example of the soil C and N data (see Data), pictured in 
+We present the example of the soil C and N data (see Data), pictured in 
 Fig. \ref{fig_biv_multi_raw_cop}A (raw data) and E (copula plot). The variables were 
@@
 from each other, answering the third part of Q1. The numbers reflect stronger upper- than lower-tail
-dependence for the soil C and N data, and this can be seen visually in the extreme tails 
+dependence for the soil C and N data, and this can also be seen visually in the extreme tails 
 of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).
@@
 Bottom panels show the corresponding copula plots.
+See text for units used in upper panels.
 \label{fig_biv_multi_raw_cop}}
@@
 \end{figure}
+<!--***DAN: We need to change the point plotting on these panels and other figures to use
+translucent points. Right now you cannot tell how many points are piled up in some parts of
+these plots, and we can easily fix that-->
 
 Although these results do convincingly show non-normal copula structure, 
-we caution that the model-averaged tail-dependence statistics may be biased 
+we caution that the model-averaged tail-dependence statistics may have been biased 
 because even the lowest-AIC copula 
@@
 
-<!--***DAN: made it to here 2018 12 21 at 10pm-->
-
 <!--bivariate summary-->
-We summarized our results columnwise in Table \ref{tab_bivar_summary} for different kinds of bivariate datasets (biogeochemical, evolutionary and ecological data) that were plotted in Fig.
-\ref{fig_biv_multi_raw_cop}(A-H).
-For all $4$ bivariate-datasets indicated by each column name (SoilCN, Birds' BMR, Mammals' BMR, Ceder-creek) in Table \ref{tab_bivar_summary}, data were always significantly correlated as indicated by p-value (<0.05) from `BiCopIndTest`{`VineCopula`} (1$^{st}$ row of Table \ref{tab_bivar_summary}).
-Non-Normal copulae always emerged as best fit copula based on minimum AIC value (2$^{nd}$ row of Table \ref{tab_bivar_summary}).
-The Normal copula was found as poor fit compared to highest AIC-weight based asymmetric copulae (3$^{rd}$ and 4$^{th}$ rows of Table \ref{tab_bivar_summary}) - which answers the first part of \hyperref[Intro]{\textbf{(Q1)}}.
-Goodness of fit of the best-fit asymmetric copulae were tested based on p-values from two statistics (Cramer-von Mises and Kolmogorov-Smirnov : 5$^{th}$ and 6$^{th}$ rows of Table \ref{tab_bivar_summary}, respectively).
-Among 4 datasets presented there, first three have significantly stronger upper tail dependence whereas the last one has significantly stronger lower tail dependence compared to Normal (compare 7$^{th}$ and 8$^{th}$ rows of Table \ref{tab_bivar_summary}).
-It was indicated by negative AIC-weights model averaged (lower-upper) tail dependence (9$^{th}$ row of Table \ref{tab_bivar_summary}) for first 3 datasets and positive AIC-weights model averaged
-(lower-upper) tail dependence for the last one, which answers the second and third part of \hyperref[Intro]{\textbf{(Q1)}} (For detailed results of each dataset see Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and Figs.
-\ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}).
-We also reported the rank of ($\Ps_l - \Ps_u$) for the actual copula data in camparison to the same non-parametric statistics of Kendall-correlation preserving 1000 Normal surrogates.
-Ranks if were $>$ 975 or $<$ 25, then indicated a significantly (95$\%$ CI) different lower tail-dependence or upper tail-dependence measures, respectively, based on Normal-null hypothesis.
-We found our model-selection results were consistent with non-parametric analysis for all the datasets (compare last two rows of Table \ref{tab_bivar_summary}).
-
-<!--multivariate example-->
-As described in \nameref{Model_selection}, for multivariable dataset like green-spruce aphid's abundance data, model selection among copulae was done separately for each pair of the 10 sampling locations, thus results are presented in $10 \times 10$ matrix format with rows and columns corresponding to locations.
-Table \ref{SM-tab_count_bestcop} shows that, for the large majority of location pairs, some non-Normal copula was a better fit, according to maximum AIC-weights, than was the Normal copula, answering the first part of question \hyperref[Intro]{\textbf{(Q1)}} for these data.
-Comparing the
-AIC weight of the best-fitting (lowest AIC) copula family for each pair of sampling locations (Table \ref{SM-tab_count_AICw}) to the AIC weight of the Normal copula family for the same pair (Table \ref{SM-tab_count_AICw_normal}) confirms that the degree of support for Normal copulae was usually not
-comparable to the degree of support for other families.
-Goodness of fit tests based on a Cramer-von Mises statistic (Table \ref{SM-tab_count_p_CvM}) in every case 
-failed to reject the hypothesis that the AIC-best copula family was also an objectively adequate description of 
-the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of copula families we used was sufficiently broad that the copulae of all pairs of locations could be characterized by at least one of them.
-In all cases at least `r min((RES_aphid_count$gfc_numBS_success/RES_aphid_count$gfc_numBS),na.rm=T)*100`$\%$ of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-
-The model-averaged (averaging performed using AIC weights) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are in Tables \ref{SM-tab_count_LTdep_AICw} and \ref{SM-tab_count_UTdep_AICw} respectively.
-These values were exclusively positive, thereby differing from what a normal copula would give (i.e., 0), and answering the second part of question \hyperref[Intro]{\textbf{(Q1)}} for these data.
-Apparently these aphid data have greater tail dependence (both lower and upper) than would be expected from a normal-copula null hypothesis.
-
-The difference between the model-averaged lower-tail dependence statistic and the model-averaged upper-tail dependence statistic (lower minus upper), shown in Table \ref{SM-tab_count_LTmUT}, is positive for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`) which indicates aphids are synchronously rare at most of the location pairs.
-The dominance of lower tail dependence in green spruce aphid's abundance data with asymmetric copula structures answers the third part of question \hyperref[Intro]{\textbf{(Q1)}}.
-Fig.
-\ref{fig_biv_multi_raw_cop}(I) and (J) represnt exemplary plots for aphid abundance (green spruce aphid) raw data and its copula form at $(4,11)$ location pair which shows stronger lower tail dependence.
-
-<!--#### Non-parametric analysis of tail dependence \label{Ex_count_NPA}-->
+Table \ref{tab_bivar_summary} summarizes bivariate results for Q1. For all $4$ 
+bivariate datasets, the two variables were always significantly related (1$^{st}$ row of 
+Table \ref{tab_bivar_summary}). A non-normal copula 
+always emerged as the lowest-AIC (highest-AIC-weight) copula 
+(2$^{nd}$ row of Table \ref{tab_bivar_summary}). 
+The N copula was always a poor fit compared to the highest-AIC-weight copula 
+(3$^{rd}$ and 4$^{th}$ rows of Table \ref{tab_bivar_summary}), answering the first part of Q1. 
+Upper- and lower-tail statistics (7$^{th}$ and 8$^{th}$ rows of 
+Table \ref{tab_bivar_summary}) differed substantially from 0, the tail dependence statistics of 
+the N copula, answering the second part of Q1; and also differed substantially from each other
+(9$^{th}$ row of Table \ref{tab_bivar_summary}),
+answering the third part of Q1. The first three datasets had stronger upper- than lower-tail dependence, and the last dataset had the reverse. 
+Although these results convincingly show non-normal copula structure, the 
+model-averaged tail-dependence statistics may have been biased 
+because even the lowest-AIC copula was a poor fit in some cases 
+(5$^{th}$ and 6$^{th}$ rows of Table \ref{tab_bivar_summary}).
+But our nonparametric statistics confirmed the same conclusions (10$^{th}$ row of 
+Table \ref{tab_bivar_summary}).
+Detailed results for each dataset are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} 
+and Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.
+
+We present our analysis of the green spruce aphid abundance data (see Data).
+Independence was tested for each pair of the 10 sampling locations, and was 
+rejected in every case.
+<!--***DAN: Right, Shyamolina? I am assuming that because "Indep" does not 
+appear anywhere in Table S11, this is correct, but please confirm or correct-->
+Model selection among copulas was carried out separately for each pair of locations.
+Best-fitting (lowest-AIC) copulas were non-normal
+for the large majority of location pairs (Table \ref{SM-tab_count_bestcop}), and 
+AIC weights for the best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
+were typically substiantially 
+larger than those for the N copula (Table \ref{SM-tab_count_AICw_normal}), answering 
+the first part of Q1 from the Introduction for these data.
+Goodness of fit tests based on a Cramer-von Mises statistic 
+(Table \ref{SM-tab_count_p_CvM}) in every case 
+failed to reject the hypothesis that the AIC-best copula family was also 
+an objectively adequate description of 
+the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov 
+statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of 
+copula families we used was sufficiently broad that the copulas of all pairs of 
+locations could be characterized by at least one of them, for these data.
+Model-averaged lower- (Table \ref{SM-tab_count_LTdep_AICw}) and upper-tail
+(Table \ref{SM-tab_count_UTdep_AICw}) dependence statistics were exclusively 
+positive and were frequently substantially large, thereby differing 
+from what an N copula would give (i.e., 0), and answering the second part of 
+Q1 for these data. Apparently these aphid data have greater tail dependence 
+(both lower and upper) than would be expected from an N-copula null hypothesis. 
+The model-averaged lower-tail dependence statistic minus the model-averaged 
+upper-tail dependence statistic (Table \ref{SM-tab_count_LTmUT}) was positive 
+for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
+out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), 
+indicating that lower-tail dependence across space was much stronger than upper-tail 
+dependence. Thus the spatial synchrony of rarity in green spruce aphid is stronger 
+than the spatial synchrony of outbreak events. This also answers the third part of Q1.
+
+<!--Non-parametric analysis of tail dependence, green spruce abundance data-->
 
@@
 For row 10, the statistic $P_{0,u_b}-P_{1-u_b}$ was computed for real data and for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented.
-The value of $u_b$ depended on the size of the dataset, and can be found in Figs.
+If ranks were $>975$ or $<25$, asymmetry of tail dependence was significant compared to a N-copula null hypothesis.
+The value of $u_b$ used depended on the size of the dataset, and can be found in Figs.
 \\ref{SM-fig_soilCN_nonparam} to \\ref{SM-fig_cc2000_nonparam} and Tables \\ref{SM-tab_soilCN_asym}, \\ref{SM-tab_bmr_birds_asym}, \\ref{SM-tab_bmr_mammals_asym}, and \\ref{SM-tab_cc_asym_2000}.

### 2019-01-09 SupportingInformation.Rmd
further edits to Q1 materials

--- 2018-12-22 SupportingInformation.Rmd
@@
 
-# Transformations and copulas used in Fig. \ref{MT-fig_cop_pedag1} \label{pedagog1_details}
-
-<!--***DAN: fill in later-->
+# Transformations used in Fig. \ref{MT-fig_cop_pedag1} \label{pedagog1_details}
+
+The data of Fig. \ref{MT-fig_cop_pedag1}A-B were generated to have standard normal marginals.
+Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
+of both variables by the cdf of the standard normal distribution, and then by the inverse of 
+the cdf of the gamma distribution with shape and scale parameters 2. 
+The function `qgamma` in R was used.
+Fig. \ref{MT-fig_cop_pedag1}D can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
+Fig. \ref{MT-fig_cop_pedag1}E can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
+of both variables by the cdf of the standard normal distribution.
+Fig. \ref{MT-fig_cop_pedag1}F can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
 
@@
 
-# Background on the copulas we fitted with data \label{copula_details}
-
-<!--Dumping ground at present, need to organize and fill in.
-
-Notes about individual copula parameter ranges:
--->
+<!--Now the copula figures-->
 
@@
 
-<!--***DAN: Right now this is a dumping grounds for text taken from the main text.
-The text is valuable, but needs to be organized to fit together in this supp mat
-section-->
-
-We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
-$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
+We here define $\Ps_{l_b,u_b}$, using the bounds
+$u+v=2l_b$ and $u+v=2u_b$ defined in the main text, as follows. For a distance $h$, we define
 $S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
@@
 bounds with the unit square. It is easy to see $S(0)=0$ and 
-$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
-of $h$, to be the area within the bounds and the unit square
+$S(h_{\text{max}})=1$. We define $S_i(h)$, also for $h$ going from $0$
+to $h_{\text{max}}$, to be the area within the bounds and the unit square
 and within distance $h$ of the line $v=u$, divided by the area
 within the bounds and the unit square. This is the expected value of
-$S(h)$ for independent data. We then defined
+$S(h)$ for independent data. We then define
 \begin{equation}\label{eq.P}
-   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
+   P_{l_b,u_b} = \int_0^{h_{\text{max}}} ( S(h) - S_i(h) ) dh.
 \end{equation}
-As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
+\noindent As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
 indicate stronger positive association
@@
 
-# Testing nonparametric statistics for various symmetric and asymmetric copulas \label{Test_npa_stat}
-
-Normal and Frank copula families 
+# Testing the nonparametric statistics \label{Test_npa_stat}
+
+N and F copula families 
 are known to have symmetric tail dependence,
 i.e., strengths of dependence in the lower and upper tails are the same, for all copulas
-in these families [@nelsen2006_copula]. Clayton copulas are known to have stronger lower- than 
-upper-tail dependence. And survival Clayton copulas, being rotations of Clayton copulas by 
+in these families [@nelsen2006_copula]. C copulas are known to have stronger lower- than 
+upper-tail dependence, and SC copulas, being rotations of C copulas by 
 $180^\circ$, have stronger upper- than lower tail dependence [@nelsen2006_copula].
-Using the `iRho` function from the `copula` package, for each of the four families normal, 
-Frank, Clayton, and survival Clayton, we obtained parameters for which the expected value of 
+Using the `iRho` function from the `copula` package, for each of the four families N, 
+F, C, and SC, we obtained parameters for which the expected value of 
 Spearman's $\rho$ was, separately, $0,0.1,0.2,...,0.9$. For each copula family and for each 
-of the selected parameter values we generated  $n$ points, 100 times, and thereby computed 100 values of the 
+of the selected parameter values we generated  $n$ points, 100 times, and thereby 
+computed 100 values of the 
 statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$. This exercise 
 was repeated, separately, 
-for $n=1000$ and for $n=35$, these cases being chosen, respectively, to reflect asymptotic behavior of the 
+for $n=1000$ and for $n=35$, these cases being chosen, respectively, to reflect 
+asymptotic behavior of the 
 statistics and their behavior on the small datasets which are common in some fields
@@
 each copula family, parameter value, and value of $n$, and results were plotted against 
-Spearman's rho (Figures \ref{fig_stat_testing35} and \ref{fig_stat_testing1000}), 
+Spearman's rho (Figs \ref{fig_stat_testing35} and \ref{fig_stat_testing1000}), 
 the expectation being that the difference statistics would be positive for 
-Clayton copulas and negative for survival Clayton copulas (except for the case of $\rho=0$, a 
+C copulas, negative for SC copulas (except for the case of $\rho=0$, a 
 boundary case of these families corresponding to independence), and not significantly different 
-from $0$ (we used a $t$-test with a significance threshold of $0.05$) for Frank and normal 
-copulas. These expectations were fullfilled (Figures \ref{fig_stat_testing35} and 
+from $0$ for F and N 
+copulas.
+We used a $t$-test with a significance threshold of $0.05$ to judge signficance here.
+These expectations were fullfilled (Figs \ref{fig_stat_testing35} and 
 \ref{fig_stat_testing1000}).
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
-\caption[Testing non-parametric statistics with datapoints $n=35$]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
-See text for details.
+\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas.
+See Appendix  \ref{Test_npa_stat} for details.
 $n=35$.
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
-\caption[Testing non-parametric statistics with datapoints $n=1000$]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence for the normal and Frank copulas.
-See text for details.
+\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas.
+See Appendix  \ref{Test_npa_stat} for details.
 $n=1000$.\label{fig_stat_testing1000}}
@@
 
-Our code for this algorithm is [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R) in the 
-[`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
-
+\noindent Our code for this algorithm is `copsurrog2d` in the 
+`BIVAN` repository (https://github.com/sghosh89/BIVAN). 
 Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
@@
 
-# Spatial resampling with multivariate datasets for non-parametric statistics \label{spatial_resamp}
+# Spatial resampling with multivariate datasets for nonparametric statistics \label{spatial_resamp}
 
@@
 a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
-$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
+$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
 average strength of 
 dependence, across all pairs of distinct locations, for data in the portion of the distributions 
-given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
+given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
+the mean characterizes average
 asymmetry of tail dependence across all pairs of locations. 
-Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
+Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
 distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
-revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
+revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
 or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
-Normal null hypothesis, and answering question \hyperref[MT-Intro]{\textbf{(Q1)}} from the \nameref{MT-Intro}.
+N-copula null hypothesis, and answering question Q1 from the Introduction.
 Bootstrapping was performed by resampling the locations of measurement, with replacement. 
@@
 [@bjornstadfalck2001 ; @walteretal2017].
+
+<!--***DAN: all good up to here, except as marked (the data section)-->
 
@@
 ```
-<!--****DAN: DAN suspects a bug in the code that generates this fig.
-because the red dot should appear outside the x's iff the text > 999 (or so,mething like that) appears in the bin, also check make_table_stat_fn r chunk-->
 
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N copula structure, for soil C and N data.
+\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for soil C and N data.
 As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
@@
 \includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N copula structure, for bird body mass and BMR data.
+\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for bird body mass and BMR data.
 Same format as Fig.
@@
 \includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N copula structure, for mammal body mass and BMR data.
+\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for mammal body mass and BMR data.
 Same format as Fig.
@@
 \includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N copula structure, for Cedar Creek data.
+\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data.
 Same format as Fig.
@@
              format='latex',
-             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundances of the green spruce aphid.",
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **abundances of the green spruce aphid**. Copulas 
+             caption.short = "Best-fitting copulas, green spruce aphid abundance data",
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for green spruce aphid abundance data. Copulas 
 considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
-survival Joe, survival BB1, survival BB6, survival BB7, survival BB8 copulas.
-Abbreviations were given in \\nameref{MT-Methods}.
+survival Joe, survival BB1, survival BB6, survival BB7, and survival BB8 copulas, with abbreviations in the methods for Q1.
 NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
@@
 ```{r tab_count_AICw, echo=F, results='markup'}
-knitr::kable(RES_aphid_count$info_ord_AICw[,,1], 
+knitr::kable(round(RES_aphid_count$info_ord_AICw[,,1],3), 
              format='latex',
-             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_AICw}",
+             caption.short = "AIC weights, best-fitting copulas, green spruce aphid abundance data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for green spruce aphid abundance data.\\label{tab_count_AICw}",
              booktabs=TRUE
@@
 source("normal_cop_AICw.R")
-knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
+knitr::kable(round(normal_cop_AICw_matrix(r=RES_aphid_count),3),
              format='latex',
-             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **abundances of the green spruce aphid**, for comparison with Table \\ref{tab_count_AICw}.
+             caption.short = "AIC weights, N copulas, green spruce aphid abundance data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for green spruce aphid abundance data, for comparison with Table \\ref{tab_count_AICw}.
 \\label{tab_count_AICw_normal}",
@@
 ```{r tab_count_p_CvM, echo=F, results='markup'}
-knitr::kable(RES_aphid_count$gfc_p_CvM,
+knitr::kable(round(RES_aphid_count$gfc_p_CvM,3),
              format='latex',
-             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **abundances of the green spruce aphid**.
-Tests were based on bootstrapping as described in Methods.\\label{tab_count_p_CvM}",
+             caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for green spruce abundance data.\\label{tab_count_p_CvM}",
              booktabs=TRUE
@@
 ```{r tab_count_p_KS, echo=F, results='markup'}
-knitr::kable(RES_aphid_count$gfc_p_KS,
+knitr::kable(round(RES_aphid_count$gfc_p_KS,3),
              format='latex',
-             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **abundances of the green spruce aphid**.
-Tests were based on bootstrapping as described in Methods.\\label{tab_count_p_KS}",
+             caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
+             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for green spruce abundance data.\\label{tab_count_p_KS}",
              booktabs=TRUE
@@
 ```{r tab_count_LTdep_AICw, echo=F, results='markup'}
-knitr::kable(RES_aphid_count$LTdep_AICw,
+knitr::kable(round(RES_aphid_count$LTdep_AICw,3),
              format='latex',
-             caption = "Model-averaged lower tail dependence statistic for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_LTdep_AICw}",
+             caption.short="Lower-tail dependence, green spruce aphid abundance data",
+             caption = "Model-averaged lower-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_LTdep_AICw}",
              booktabs=TRUE
@@
 ```{r tab_count_UTdep_AICw, echo=F, results='markup'}
-knitr::kable(RES_aphid_count$UTdep_AICw,
+knitr::kable(round(RES_aphid_count$UTdep_AICw,3),
              format='latex',
-             caption = "Model averaged upper tail dependence statistic for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_UTdep_AICw}",
+             caption.short="Upper-tail dependence, green spruce aphid abundance data",
+             caption = "Model-averaged upper-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_UTdep_AICw}",
              booktabs=TRUE
@@
 ```{r tab_count_LTmUT, echo=F, results='markup'}
-knitr::kable(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,
+knitr::kable(round(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,3),
              digits = 5,
              format='latex',
-             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **abundances of the green spruce aphid**.\\label{tab_count_LTmUT}",
+             caption.short="Lower- minus upper-tail dependence, green spruce aphid abundance data",
+             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for green spruce aphid abundance data.\\label{tab_count_LTmUT}",
              booktabs=TRUE
@@
              format='latex',
-             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **first flight data for the leaf-curling plum aphid**.
-Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations given in \\nameref{MT-Methods}.
+             caption.short = "Best-fitting copulas, leaf-curling plum aphid first flight data",
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for leaf-curling plum aphid first flight data.
+Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations in the methods for Q1.
 NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.
@@
 ```{r tab_ff_AICw, echo=F, results='markup'}
-knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], 
+knitr::kable(round(RES_aphid_ff$info_ord_AICw[,,1],3), 
              format='latex',
-             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_AICw}",
+             caption.short="AIC weights, best-fitting copulas, leaf-curling plum aphid first flight data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_AICw}",
              booktabs=TRUE
@@
 source("normal_cop_AICw.R")
-knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
+knitr::kable(round(normal_cop_AICw_matrix(r=RES_aphid_ff),3),
              format='latex',
-             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **first flight data for the leaf-curlung plum aphid**, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
+             caption.short="AIC weights, N copulas, leaf-curling plum aphid first flight data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for leaf-curlung plum aphid first flight data, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
              booktabs=TRUE
@@
 ```{r tab_ff_p_CvM, echo=F, results='markup'}
-knitr::kable(RES_aphid_ff$gfc_p_CvM,
+knitr::kable(round(RES_aphid_ff$gfc_p_CvM,3),
              format='latex',
-             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **first flight data for the leaf-curling plum aphid**.
-Tests were based on bootstrapping as described in Methods.\\label{tab_ff_p_CvM}",
+             caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for leaf-curling plum aphid first flight data.\\label{tab_ff_p_CvM}",
              booktabs=TRUE
@@
 ```{r tab_ff_p_KS, echo=F, results='markup'}
-knitr::kable(RES_aphid_ff$gfc_p_KS,
+knitr::kable(round(RES_aphid_ff$gfc_p_KS,3),
              format='latex',
-             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **first flight data for the leaf-curling plum aphid**.
-Tests were based on bootstrapping as described in Methods.\\label{tab_ff_p_KS}",
+             caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
+             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for leaf-curling plum aphid first flight data.\\label{tab_ff_p_KS}",
              booktabs=TRUE
@@
 ```{r tab_ff_LTdep_AICw, echo=F, results='markup'}
-knitr::kable(RES_aphid_ff$LTdep_AICw,
+knitr::kable(round(RES_aphid_ff$LTdep_AICw,3),
              format='latex',
-             caption = "Model-averaged lower tail dependence statistic for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_LTdep_AICw}",
+             caption.short="Lower-tail dependence, leaf-curling plum aphid first flight data",
+             caption = "Model-averaged lower-tail dependence statistics for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_LTdep_AICw}",
              booktabs=TRUE
@@
 ```{r tab_ff_UTdep_AICw, echo=F, results='markup'}
-knitr::kable(RES_aphid_ff$UTdep_AICw,
+knitr::kable(round(RES_aphid_ff$UTdep_AICw,3),
              format='latex',
-             caption = "Model-averaged upper tail dependence statistic for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_UTdep_AICw}",
+             caption.short="Upper-tail dependence, leaf-curling plum aphid first flight data",
+             caption = "Model-averaged upper-tail dependence statistics for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_UTdep_AICw}",
              booktabs=TRUE
@@
 ```{r tab_ff_LTmUT, echo=F, results='markup'}
-knitr::kable(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,
+knitr::kable(round(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,3),
              digits = 5,
              format='latex',
-             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_LTmUT}",
+             caption.short="Lower- minus upper-tail dependence, leaf-curling plum aphid first flight data",
+             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for leaf-curling plum aphid first flight data.\\label{tab_ff_LTmUT}",
              booktabs=TRUE
@@
              format='latex', 
-             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for abundance data for \\textit{Ceratium furca} in UK seas.",
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **abundance data for \\textit{Ceratium furca} in UK seas**.
-See \\nameref{MT-Methods} for copula abbreviations.
+             caption.short = "Best-fitting copulas, \\textit{Ceratium furca} abundance data",
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for \\textit{Ceratium furca} abundance data.
+See methods for Q1 for copula abbreviations.
+Indep.=Independent.
 NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_pns_bestcop}",
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_AICw}",
+             caption.short="AIC weights, best-fitting copulas, \\textit{Ceratium furca} abundance data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_AICw}",
              booktabs=TRUE)
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_AICw_normal}",
+             caption.short="AIC weights, N copulas, \\textit{Ceratium furca} abundance data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for \\textit{Ceratium furca} abundance data, for comparison with Table \\ref{tab_pns_AICw}.\\label{tab_pns_AICw_normal}",
              booktabs=TRUE)
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **abundance data for *Ceratium furca* in UK seas**.
-Tests were based on bootstrapping as described in Methods.\\label{tab_pns_p_CvM}",
+             caption.short="Goodness of fit of best-fitting copulas, \\textit{Ceratium furca} abundance data",
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for \\textit{Ceratium furca} abundance data.\\label{tab_pns_p_CvM}",
              booktabs=TRUE)
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
+             caption.short="Goodness of fit of best-fitting copulas, \\textit{Ceratium furca} abundance data",
              caption = "$P$-value results of a Kolmogorov-Smirnov-based test of
 the goodness of fit of the lowest-AIC copula family for each pair of
-locations for **abundance data for *Ceratium furca* in UK seas**. Tests
-were based on bootstrapping as described in Methods.\\label{tab_pns_p_KS}",
+locations for \\textit{Ceratium furca} abundance data.\\label{tab_pns_p_KS}",
              booktabs=TRUE)
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "Model-averaged lower tail dependence statistic for each location pair for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_LTdep_AICw}",
+             caption.short="Lower-tail dependence, \\textit{Ceratium furca} abundance data",
+             caption = "Model-averaged lower-tail dependence statistics for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_LTdep_AICw}",
              booktabs=TRUE)
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "Model-averaged upper tail dependence statistic for each location pair for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_UTdep_AICw}",
+             caption.short="Upper-tail dependence, \\textit{Ceratium furca} abundance data",
+             caption = "Model-averaged upper-tail dependence statistics for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_UTdep_AICw}",
              booktabs=TRUE)
@@
 h[is.nan(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "Model averaged lower tail dependence minus model
-averaged upper tail dependence statistic for each location pair for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_LTmUT}",
+             caption.short="Lower- minus upper-tail dependence, \\textit{Ceratium furca} abundance data",
+             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for \\textit{Ceratium furca} abundance data.\\label{tab_pns_LTmUT}",
              booktabs=TRUE)
@@
 
-<!--***DAN: stopped here 2018 04 07-->
-
 <!--Load and format the data-->
@@
              format='latex',
-             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.",
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **methane-flux data**.
-See \\nameref{MT-Methods} for copula abbreviations.
+             caption.short = "Best-fitting copulas, methane flux data",
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane flux data.
+See methods for Q1 for copula abbreviations.
+Indep.=Independent.
 NA occurs along the diagonal because diagonal entries represent a sampling location compared to itself.\\label{tab_methane_bestcop}",
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **methane-flux data**.\\label{tab_methane_AICw}",
+             caption.short="AIC weights, best-fitting copulas, methane flux data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for methane flux data.\\label{tab_methane_AICw}",
              booktabs=TRUE)
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **methane-flux data**.\\label{tab_methane_AICw_normal}",
+             caption.short="AIC weights, N copulas, methane flux data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for methane flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
              booktabs=TRUE)
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **methane-flux data**.
-Tests were based on bootstrapping as described in Methods.\\label{tab_methane_p_CvM}",
+             caption.short="Goodness of fit of best-fitting copulas, methane flux data",
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for methane flux data.\\label{tab_methane_p_CvM}",
              booktabs=TRUE)
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
+             caption.short="Goodness of fit of best-fitting copulas, methane flux data",
              caption = "$P$-value results of a Kolmogorov-Smirnov-based test of
 the goodness of fit of the lowest-AIC copula family for each pair of
-locations for **methane-flux data**. Tests were based on bootstrapping as described in Methods.\\label{tab_methane_p_KS}",
+locations for methane flux data.\\label{tab_methane_p_KS}",
              booktabs=TRUE)
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "Model-averaged lower tail dependence statistic for each location pair for **methane-flux data**.\\label{tab_methane_LTdep_AICw}",
+             caption.short="Lower-tail dependence, methane flux data",
+             caption = "Model-averaged lower-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_LTdep_AICw}",
              booktabs=TRUE)
@@
 h[is.infinite(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "Model-averaged upper tail dependence statistic for each location pair for **methane-flux data**.\\label{tab_methane_UTdep_AICw}",
+             caption.short="Upper-tail dependence, methane flux data",
+             caption = "Model-averaged upper-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_UTdep_AICw}",
              booktabs=TRUE)
@@
 h[is.nan(h)]<-NA
-knitr::kable(h,
+knitr::kable(round(h,3),
              format="latex",
              digits=3,
-             caption = "Model averaged lower tail dependence minus model
-averaged upper tail dependence statistic for each location pair for **methane-flux data**.\\label{tab_methane_LTmUT}",
+             caption.short="Lower- minus upper-tail dependence, methane flux data",
+             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_LTmUT}",
              booktabs=TRUE)

### 2019-01-09 Paper.Rmd
replied to Q1 corrections

@@
 <!--***DAN: Shya, I changed the abbreviation for Kendall Correlation from KC to K in the 
-caption, please change in the code that generates the figures, as well-->
+caption, please change in the code that generates the figures, as well
+***Shya: Done.-->
 
@@
 <!--***DAN: Right, Shyamolina? I am assuming that because "Indep" does not 
-appear anywhere in Table S11, this is correct, but please confirm or correct-->
+appear anywhere in Table S11, this is correct, but please confirm or correct.
+***Shya: Yes, you are right, all the pairwise sampling locations for aphid data rejected independence test and show copula structures.-->
 Model selection among copulas was carried out separately for each pair of locations.

### 2019-01-09 SupportingInformation.Rmd
fixing multivar summary table bug

@@
   m$Avg.N.AICw<-round(mean(normal_AICw,na.rm=T),3)
-  temp<-which((r$gfc_p_CvM>0.01) & (r$gfc_p_KS>0.01),arr.ind=T)
+  pcvm<-r$gfc_p_CvM
+  pcvm[pcvm==Inf]<-NaN
+  pks<-r$gfc_p_KS
+  pks[pks==Inf]<-NaN
+  temp<-which((pcvm>0.01) & (pks>0.01),arr.ind=T)
   m$num.good.fit<-dim(temp)[1]

### 2019-01-10 Paper.Rmd
firther edits

--- 2019-01-09 Paper.Rmd
@@
 available from `BiCopGofTest` in the `VineCopula` package). However, nonparametric statistics verified
-both that tail dependence was distinct from that of a normal copula for the soil data, and that it 
+both that tail dependence was distinct from that of an N copula for the soil data, and that it 
 was asymmetric (Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}). 
@@
 (both lower and upper) than would be expected from an N-copula null hypothesis. 
-The model-averaged lower-tail dependence statistic minus the model-averaged 
-upper-tail dependence statistic (Table \ref{SM-tab_count_LTmUT}) was positive 
+Model-averaged lower-tail dependence statistics minus model-averaged 
+upper-tail dependence statistics (Table \ref{SM-tab_count_LTmUT}) were positive 
 for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
@@
 
-<!--Non-parametric analysis of tail dependence, green spruce abundance data-->
-
-As mentioned in \nameref{NPA} and Appendix \ref{SM-spatial_resamp}, mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
-computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
-resampling scheme (Table \ref{SM-tab_count_resamp}). Confidence intervals of
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
-indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
-exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}.
-Visual illustration of those results is discussed in Appendix \ref{SM-visual_npa}.
+<!--Non-parametric analysis of tail dependence, green spruce aphid abundance data-->
+Nonparametric statistics verified both that tail dependence was distinct from that 
+of an N copula for the green spruce aphid abundance data, and that it was asymmetric.
+Mean values across all pairs of locations of the statistics 
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ were positive and
+confidence intervals excluded $0$ (the value for an N copula), confirming tail 
+dependence distinct from that of an N copula (Table \ref{SM-tab_count_resamp}). 
+Mean values of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
+also were always positive and had confidence intervals that excluded $0$ 
+(Table \ref{SM-tab_count_resamp}), confirming asymmetry of tail dependence and 
+stronger lower- than upper-tail dependence. Each of the statistics 
+was plotted against geographic (great-circle) distances between sampling locations, 
+across all pairs of sampling locations (Fig. \ref{SM-fig_aphid_count}).
 
 <!--multivariate summary-->
+Table \ref{tab_multivar_summary} summarizes some of our results using our
+multivariate data. In addition to summaries of the green spruce aphid abundance data results 
+elaborated in detail above, the table also summarizes results of analyzing 
+leaf-curling plum aphid first flight data, abundance data for the dinoflagellate
+\textit{Ceratium furca}, and methan flux data (see Data).
+
 We did model selection and non-parametric analysis with all 20 aphid and 22 plankton species and found our two approaches were consistent to each other.
 Here, to present our results we chose one species from each category.
-We summarized our results columnwise in Table \ref{tab_multivar_summary} for all multivariable datasets (i.e., abundance data for green spruce aphid and firstflight date data for leaf-curling plum aphid from `r nrow(RES_aphid_count$gfc_numBS)` sampling locations, abundance data for plankton from UK seas from $14$ sampling locations,
-methane flux data from $13$ sampling locations).
 
@@
 Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.
+
+We also carried out the same analyses for abundance and first-flight data for the 18 
+aphids species for which we had data other than the green spruce and leaf-curling plum aphids,
+as well as for the 21 plankton taxa for which we had abundance data other 
+\textit{Ceratium furca} (results not shown). The conclusions were generally supported that 
+non-N copula structure and tail dependence, as well as asymmetry of tail dependence, are 
+very common.
 
@@
 dt <- tibble(
-            Description = c("1. Total number of location pairs (excluding self-comparison)",
-           "2. Number of location pairs having non-independent copula structures",
-           "3. Number of non-Normal copula as best fit",
-           "4. Average AIC-weights of best fit copula from all location pair",
-           "5. Average AIC-weights of Normal copula from all location pair",
-           "6. Number of good-fit location pair*",
-           "7. 95% CI of model-avg. lower tail-dep. from all location pair",
-           "8. 95% CI of model-avg. upper tail-dep. from all location pair",
-           "9. Percentage of location pairs showing stronger lower tail dependence",
-           "10. 95% CI of (Pl - Pu) from all location pair")
+            Description = c("1. Location pairs (excluding self comparisons)",
+           "2. Number of non-independent pairs",
+           "3. Number of pairs with non-N copula as best fit",
+           "4. Average AIC-weights of best-fitting copulas across location pairs",
+           "5. Average AIC-weights of N copula across pairs",
+           "6. Number of well-fitting location pairs",
+           "7. 2.5th and 97.5 quantiles of model-avg. lower-tail dependence",
+           "8. 2.5th and 97.5th quantiles of model-avg. upper-tail dependence",
+           "9. Percent pairs showing stronger lower tail dependence",
+           "10. 95% CI of average P_l - P_u across all location pair")
            )
@@
              format="latex", 
-             caption = "Summary of results for multivariate data-sets\\label{tab_multivar_summary}",
+             short.caption = "Summary of results for multivariate data sets"
+             caption = "Summary of results for multivariate data-sets.
+Rows 3-9 of the table were computed for the non-independent pairs (row 2) only; row 10 was for all location pairs.
+A well-fitting location pair (row 6) was one for which the best-fitting copula had $p$-values $>0.01$ for both the Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests.
+\\label{tab_multivar_summary}",
              booktabs=T, linesep = "\\addlinespace",
@@
              )%>%column_spec(2:5,width="2.5cm")%>%
-             column_spec(1, width = "7cm")%>%kable_styling(latex_options = "basic",position = "center")%>%
-   add_footnote("Good fit was considered for best fit copula with p-values > 0.01 (both Cramer-von Mises and Kolmogorov-Smirnov method)", notation="symbol", threeparttable = T, escape = F)
+             column_spec(1, width = "7cm")%>%kable_styling(latex_options = "basic",position = "center")
 ```

### 2019-01-10 SupportingInformation.Rmd
firther edits

--- 2019-01-09 SupportingInformation.Rmd
@@
 
-<!--***DAN: all good up to here, except as marked (the data section)-->
-
-# How to visualize non-parametric analysis results for multivariate spatial data? \label{visual_npa}
-
-The principle remains same but for example we chose one of spatial data set (green spruce aphid's count) to illustrate results from Table \ref{tab_count_resamp}.
-Plotting Spearman or Kendall correlation between aphid abundance time series for pairs of locations as a function of great circle distance between the locations (Fig.
-\ref{fig_aphid_count} A,B) showed the typical decline of correlation with distance between sampling locations that has been seen for numerous other datasets in earlier work.
-Plotting $\cor_l$, $\cor_u$, $\Ps_l$, and $\Ps_u$  against distance between pairs of locations shows these quantities also decline with distance (Fig.
-\ref{fig_aphid_count} C-F).
-Plotting $\Dsq_u$ and  $\Dsq_l$ against distance between pairs of locations shows these quantities increase with distance (as expected, given the opposite interpretation these two statistics have compared to $\cor_u$, $\Ps_u$ and $\cor_l$, $\Ps_l$, respectively).
-Plotting $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$ versus distance (Fig.
-\ref{fig_aphid_count} I-K) did not show a decline with distance, but did illustrate visually the statistical result of Table \ref{tab_count_resamp} that averages of these quantities across all pairs of sampling locations were significantly greater than zero, demonstrating asymmetry of tail dependence, as also seen from model selection approach.
-Code for these analyses is in [`NonParamStat.R`](https://github.com/sghosh89/BIVAN/blob/master/NonParamStat.R) in the [`BIVAN`](https://github.com/sghosh89/BIVAN) repository.
+<!--***DAN: text all good up to here, except as marked (the data section)-->
 
@@
 s<-stat_aphid_count$numericdf[,c(3,2,4)]
-colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
-rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
-knitr::kable(s,
+colnames(s)<-c("Lower 95% CI", "Mean", "Upper 95% CI")
+rownames(s)<-c("Spearman","Kendall","cor_l","cor_u","P_l","P_u","D^2_l","D^2_u","cor_l - cor_u","P_l - P_u","D^2_u - D^2_l")
+knitr::kable(round(s,3),
              format = "latex",
-             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **abundances of the green spruce aphid**.\\label{tab_count_resamp}",
+             caption.short="Results for nonparametric statistics, green spruce aphid abundance data",
+             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for green spruce aphid abundance data.\\label{tab_count_resamp}",
              booktabs=TRUE)
@@
 \includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
-\caption[Non-parametric tail statistics results for green spruce aphid abundance data]{Non-parametric tail statistics results for green spruce aphid abundance data plotted against geographical distance ($D$) between two sampling locations.
-(A) Spearman correlation vs.
-$D$; (B) Kendall correlation vs.
-$D$; (C) $\cor_l$ vs.
-$D$; (D) $\cor_u$ vs.
-$D$; (E) $\Ps_l$ vs.
-$D$; (F) $\Ps_u$ vs.
-$D$; (G) $\Dsq_u$ vs.
-$D$;
-(H) $\Dsq_l$ vs. $D$; 
-(I) $\cor_l-\cor_u$ vs. $D$, (J) $\Ps_l-\Ps_u$ vs. $D$, (K) $\Dsq_u-\Dsq_l$ vs. $D$ plotted for all 
-pairs of locations. \label{fig_aphid_count}}
+\caption[Nonparametric tail statistics versus distance between locations for green spruce aphid abundance data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for green spruce aphid abundance data.
+\label{fig_aphid_count}}
 \end{center}
@@
 s<-stat_aphid_ff$numericdf[,c(3,2,4)]
-colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
-rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
-knitr::kable(s,
+colnames(s)<-c("Lower 95% CI", "Mean", "Upper 95% CI")
+rownames(s)<-c("Spearman","Kendall","cor_l","cor_u","P_l","P_u","D^2_l","D^2_u","cor_l - cor_u","P_l - P_u","D^2_u - D^2_l")
+knitr::kable(round(s,3),
              format = "latex",
-             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **first flight data for the leaf-curling plum aphid**.\\label{tab_ff_resamp}",
+             caption.short="Results for nonparametric statistics, leaf-curling plum aphid first flight data",
+             caption = "Average values of statistics across all location pairs and confidence intervals based on spatial resampling for leaf-curling plum aphid first flight data.\\label{tab_ff_resamp}",
              booktabs=TRUE)
@@
 \includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
-\caption[Non-parametric tail statistics results for leaf-curling plum aphid's first flight date data]{Non-parametric tail statistics results for leaf-curling plum aphid's first flight date data plotted against geographical distance $(D)$ between two sampling locations.
-(A) Spearman correlation vs.
-$D$, (B) Kendall correlation vs.
-$D$, (C) $\cor_l$ vs.
-$D$, (D) $\cor_u$ vs.
-$D$; (E) $\Ps_l$ vs.
-$D$; (F) $\Ps_u$ vs.
-$D$; (G) $\Dsq_u$ vs.
-$D$; (H) $\Dsq_l$ vs.
-$D$; (I) $\cor_l-\cor_u$ vs.
-$D$; (J) $\Ps_l-\Ps_u$ vs.
-$D$; (K) $\Dsq_u-\Dsq_l$ vs.
-$D$ plotted for all pairs of locations.
+\caption[Nonparametric tail statistics versus distance between locations for leaf-curling plum aphid first flight data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for leaf-curling plum aphid first flight data.
 \label{fig_aphid_ff}}
@@
 s<-stat_plankton_north_sea$numericdf[,c(3,2,4)]
-colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
-rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
-knitr::kable(s,
+colnames(s)<-c("Lower 95% CI", "Mean", "Upper 95% CI")
+rownames(s)<-c("Spearman","Kendall","cor_l","cor_u","P_l","P_u","D^2_l","D^2_u","cor_l - cor_u","P_l - P_u","D^2_u - D^2_l")
+knitr::kable(round(s,3),
              format = "latex",
-             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **abundance data for *Ceratium furca* in UK seas**.\\label{tab_pns_resamp}",
+             caption.short="Results for nonparametric statistics, \\textit{Ceratium furca} abundance data",
+             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for \\textit{Ceratium furca} abundance data.\\label{tab_pns_resamp}",
              booktabs=TRUE)
@@
 \includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf}
-\caption[Non-parametric tail statistics results for \textit{Ceratium furca} abundance data]{Non-parametric tail statistics results for \textit{Ceratium furca} abundance data plotted against geographical distance $(D)$ between two sampling locations.
-(A) Spearman correlation vs.
-$D$, (B) Kendall correlation vs.
-$D$, (C) $\cor_l$ vs.
-$D$, (D) $\cor_u$ vs.
-$D$; (E) $\Ps_l$ vs.
-$D$; (F) $\Ps_u$ vs.
-$D$; (G) $\Dsq_u$ vs.
-$D$; (H) $\Dsq_l$ vs.
-$D$; (I) $\cor_l-\cor_u$ vs.
-$D$; (J) $\Ps_l-\Ps_u$ vs.
-$D$; (K) $\Dsq_u-\Dsq_l$ vs.
-$D$ plotted for all pairs of locations.
+\caption[Nonparametric tail statistics versus distance between locations for \textit{Ceratium furca} abundance data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for \textit{Ceratium furca} abundance data.
 \label{fig_pns}}
@@
 
-\newpage
 <!--display the main statistical results in a table-->
@@
 s<-stat_methane$numericdf[,c(3,2,4)]
-colnames(s)<-c("Lower 95\\% CI", "Mean", "Upper 95\\% CI")
-rownames(s)<-c("Spearman","Kendall","$\\cor_l$","$\\cor_u$","$\\Ps_l$","$\\Ps_u$","$\\Dsq_l$","$\\Dsq_u$","$\\cor_l - \\cor_u$","$\\Ps_l - \\Ps_u$","$\\Dsq_u - \\Dsq_l$")
-knitr::kable(s,
+colnames(s)<-c("Lower 95% CI", "Mean", "Upper 95% CI")
+rownames(s)<-c("Spearman","Kendall","cor_l","cor_u","P_l","P_u","D^2_l","D^2_u","cor_l - cor_u","P_l - P_u","D^2_u - D^2_l")
+knitr::kable(round(s,3),
              format = "latex",
-             caption = "Spatially averaged values of statistics and confidence intervals based on resampling for **methane-flux data**.\\label{tab_methane_resamp}",
+             caption.short="Results for nonparametric statistics, methane flux data",
+             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for methane flux data.\\label{tab_methane_resamp}",
              booktabs=TRUE)
@@
 \includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u-D2l_vs_D.pdf}
-\caption[Non-parametric tail statistics results for methane-flux data]{Non-parametric tail statistics results for methane-flux data plotted against geographical distance $(D)$ between two sampling locations.
-(A) Spearman correlation vs.
-$D$, (B) Kendall correlation vs.
-$D$, (C) $\cor_l$ vs.
-$D$, (D) $\cor_u$ vs.
-$D$; (E) $\Ps_l$ vs.
-$D$; (F) $\Ps_u$ vs.
-$D$; (G) $\Dsq_u$ vs.
-$D$; (H) $\Dsq_l$ vs.
-$D$; (I) $\cor_l-\cor_u$ vs.
-$D$; (J) $\Ps_l-\Ps_u$ vs.
-$D$; (K) $\Dsq_u-\Dsq_l$ vs.
-$D$ plotted for all pairs of locations.
+\caption[Nonparametric tail statistics versus distance between locations for methane-flux data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for methane-flux data.
 \label{fig_methane}}

### 2019-01-10 Paper.Rmd
surpressing caption.short for table2 as of now to knit

@@
              format="latex", 
-             short.caption = "Summary of results for multivariate data sets"
+             #short.caption = "Summary of results for multivariate data sets",
              caption = "Summary of results for multivariate data-sets.

### 2019-01-10 Paper.Rmd
table caption short fixed

--- 2019-01-10 Paper.Rmd
@@
              format="latex", 
-             short.caption = "Summary of results for multivariate data sets"
+             caption.short = "Summary of results for multivariate data sets",
              caption = "Summary of results for multivariate data-sets.

### 2019-01-10 SupportingInformation.Rmd
table caption short fixed

@@
 knitr::kable(h, #longtable = T, 
-             format = "latex",
+             format = "latex",linesep = "",
       caption.short = "Model fitting results for soil C and N dataset",
@@
 knitr::kable(tab_soilCN_asym, booktabs = T, 
-             format = "latex",
+             format = "latex",linesep = "",
       caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
@@
 knitr::kable(h,
-             format='latex',
+             format='latex',linesep = "",
              caption.short = "Model fitting results for the bird body mass and BMR dataset",
@@
 knitr::kable(tab_bmr_birds_asym,
-             format="latex",
+             format="latex",linesep = "",
              caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
@@
 knitr::kable(h,
-             format='latex',
+             format='latex',linesep = "",
              caption.short = "Model fitting results for the mammal body mass and BMR dataset",
@@
 knitr::kable(tab_bmr_mammals_asym,
-             format="latex",
+             format="latex",linesep = "",
              caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
@@
 knitr::kable(h,
-             format='latex',
+             format='latex',linesep = "",
              caption.short = "Model fitting results for the Cedar Creek dataset",
@@
 knitr::kable(tab_cc_asym_2000,
-             format="latex",
+             format="latex",linesep = "",
              caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
@@
 knitr::kable(RES_aphid_count$info_ord_copname[,,1], #longtable = T,
-             format='latex',
+             format='latex',linesep = "",
              caption.short = "Best-fitting copulas, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$info_ord_AICw[,,1],3), 
-             format='latex',
+             format='latex',linesep = "",
              caption.short = "AIC weights, best-fitting copulas, green spruce aphid abundance data",
@@
 knitr::kable(round(normal_cop_AICw_matrix(r=RES_aphid_count),3),
-             format='latex',
+             format='latex',linesep = "",
              caption.short = "AIC weights, N copulas, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$gfc_p_CvM,3),
-             format='latex',
+             format='latex',linesep = "",
              caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$gfc_p_KS,3),
-             format='latex',
+             format='latex',linesep = "",
              caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$LTdep_AICw,3),
-             format='latex',
+             format='latex',linesep = "",
              caption.short="Lower-tail dependence, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$UTdep_AICw,3),
-             format='latex',
+             format='latex',linesep = "",
              caption.short="Upper-tail dependence, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,3),
-             digits = 5,
+             digits = 5,linesep = "",
              format='latex',
@@
 knitr::kable(round(s,3),
-             format = "latex",
+             format = "latex",linesep = "",
              caption.short="Results for nonparametric statistics, green spruce aphid abundance data",
@@
 knitr::kable(RES_aphid_ff$info_ord_copname[,,1], #longtable = T,
-             format='latex',
+             format='latex',linesep = "",
              caption.short = "Best-fitting copulas, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(RES_aphid_ff$info_ord_AICw[,,1],3), 
-             format='latex',
+             format='latex',linesep = "",
              caption.short="AIC weights, best-fitting copulas, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(normal_cop_AICw_matrix(r=RES_aphid_ff),3),
-             format='latex',
+             format='latex',linesep = "",
              caption.short="AIC weights, N copulas, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(RES_aphid_ff$gfc_p_CvM,3),
-             format='latex',
+             format='latex',linesep = "",
              caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(RES_aphid_ff$gfc_p_KS,3),
-             format='latex',
+             format='latex',linesep = "",
              caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(RES_aphid_ff$LTdep_AICw,3),
-             format='latex',
+             format='latex',linesep = "",
              caption.short="Lower-tail dependence, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(RES_aphid_ff$UTdep_AICw,3),
-             format='latex',
+             format='latex',linesep = "",
              caption.short="Upper-tail dependence, leaf-curling plum aphid first flight data",
@@
              digits = 5,
-             format='latex',
+             format='latex',linesep = "",
              caption.short="Lower- minus upper-tail dependence, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(s,3),
-             format = "latex",
+             format = "latex",linesep = "",
              caption.short="Results for nonparametric statistics, leaf-curling plum aphid first flight data",
@@
 knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1], #longtable=T,
-             format='latex', 
+             format='latex', linesep = "",
              caption.short = "Best-fitting copulas, \\textit{Ceratium furca} abundance data",
@@
              format="latex",
-             digits=3,
+             digits=3,linesep = "",
              caption.short="AIC weights, best-fitting copulas, \\textit{Ceratium furca} abundance data",
@@
 knitr::kable(round(h,3),
-             format="latex",
+             format="latex",linesep = "",
              digits=3,
@@
 knitr::kable(round(s,3),
-             format = "latex",
+             format = "latex",linesep = "",
              caption.short="Results for nonparametric statistics, \\textit{Ceratium furca} abundance data",
@@
 knitr::kable(RES_methane$info_ord_copname[,,1], #longtable=T,
-             format='latex',
+             format='latex',linesep = "",
              caption.short = "Best-fitting copulas, methane flux data",
@@
 knitr::kable(round(s,3),
-             format = "latex",
+             format = "latex",linesep = "",
              caption.short="Results for nonparametric statistics, methane flux data",
@@
 knitr::kable(summary_stat, booktabs = T, 
-             format = "latex",
+             format = "latex",linesep = "",
              digits=3,
@@
 knitr::kable(temp, booktabs = T, 
-             format = "latex",
+             format = "latex",linesep = "",
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",

### 2019-01-10 Paper.Rmd
fixing all table appearence

--- 2019-01-10 Paper.Rmd
@@
            "9. Percent pairs showing stronger lower tail dependence",
-           "10. 95% CI of average P_l - P_u across all location pair")
+           "10. 2.5th and 97.5th quantiles of $<$P$_{l}$ - P${_u}$$>$ across all location pair")
            )
@@
 knitr::kable(tab_multivar_summary, 
-             format="latex", 
-             short.caption = "Summary of results for multivariate data sets"
+             format="latex", escape = F,
+             caption.short = "Summary of results for multivariate data sets",
              caption = "Summary of results for multivariate data-sets.

### 2019-01-10 SupportingInformation.Rmd
fixing all table appearence

@@
 s<-stat_aphid_count$numericdf[,c(3,2,4)]
-colnames(s)<-c("Lower 95% CI", "Mean", "Upper 95% CI")
-rownames(s)<-c("Spearman","Kendall","cor_l","cor_u","P_l","P_u","D^2_l","D^2_u","cor_l - cor_u","P_l - P_u","D^2_u - D^2_l")
-knitr::kable(round(s,3),
-             format = "latex",linesep = "",
+sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
+                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
+s<-cbind(sc,s)
+colnames(s)<-c("","2.5th quantile", "Mean", "97.5th quantile")
+knitr::kable(s,
+             format = "latex",linesep = "", escape = F,align="c",
              caption.short="Results for nonparametric statistics, green spruce aphid abundance data",
@@
 s<-stat_aphid_ff$numericdf[,c(3,2,4)]
-colnames(s)<-c("Lower 95% CI", "Mean", "Upper 95% CI")
-rownames(s)<-c("Spearman","Kendall","cor_l","cor_u","P_l","P_u","D^2_l","D^2_u","cor_l - cor_u","P_l - P_u","D^2_u - D^2_l")
-knitr::kable(round(s,3),
-             format = "latex",linesep = "",
+sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
+                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
+s<-cbind(sc,s)
+colnames(s)<-c("","2.5th quantile", "Mean", "97.5th quantile")
+knitr::kable(s,
+             format = "latex",linesep = "", escape = F,align="c",
              caption.short="Results for nonparametric statistics, leaf-curling plum aphid first flight data",
@@
 s<-stat_plankton_north_sea$numericdf[,c(3,2,4)]
-colnames(s)<-c("Lower 95% CI", "Mean", "Upper 95% CI")
-rownames(s)<-c("Spearman","Kendall","cor_l","cor_u","P_l","P_u","D^2_l","D^2_u","cor_l - cor_u","P_l - P_u","D^2_u - D^2_l")
-knitr::kable(round(s,3),
-             format = "latex",linesep = "",
+sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
+                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
+s<-cbind(sc,s)
+colnames(s)<-c("","2.5th quantile", "Mean", "97.5th quantile")
+knitr::kable(s,
+             format = "latex",linesep = "", escape = F,align="c",
              caption.short="Results for nonparametric statistics, \\textit{Ceratium furca} abundance data",
@@
 s<-stat_methane$numericdf[,c(3,2,4)]
-colnames(s)<-c("Lower 95% CI", "Mean", "Upper 95% CI")
-rownames(s)<-c("Spearman","Kendall","cor_l","cor_u","P_l","P_u","D^2_l","D^2_u","cor_l - cor_u","P_l - P_u","D^2_u - D^2_l")
-knitr::kable(round(s,3),
-             format = "latex",linesep = "",
+
+#rownames(s)<-c("Spearman","Kendall","cor_l","cor_u","P_l","P_u","D^2_l","D^2_u","cor_l - cor_u","P_l - P_u","D^2_u - D^2_l")
+
+sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
+                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
+s<-cbind(sc,s)
+colnames(s)<-c("","2.5th quantile", "Mean", "97.5th quantile")
+knitr::kable(s,
+             format = "latex",linesep = "", escape = F,align="c",
              caption.short="Results for nonparametric statistics, methane flux data",

### 2019-01-10 SupportingInformation.Rmd
fixing tables

@@
  #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
- atdres<-data.frame(Statistic=c("$\\cor_\\text{first slice}-\\cor_\\text{last slice}$",
-                               "$\\Ps_\\text{first slice}-\\Ps_\\text{last slice}$",
-                                "$\\Dsq_\\text{last slice}-\\Dsq_\\text{first slice}$"),
+ atdres<-data.frame(Statistic=c("cor$_{FS}$-cor$_{LS}$",
+                               "P$_{FS}$-P$_{LS}$",
+                                "D$_{LS}^2$-D$_{FS}^2$"),
                    Kendall=rep("",3),
@@
 tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
-tab_soilCN_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
-knitr::kable(tab_soilCN_asym, booktabs = T, 
+#tab_soilCN_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
+knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
              format = "latex",linesep = "",
@@
       caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
-The values of the listed statistics (corFS-corLS $=\\cor_{0,0.1}-\\cor_{0.9,1}$, PFS-PLS $=\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D2LS-D2FS $=\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons.
+The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons.
 A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
-#kable_styling(, position = "float_left")
 ```
@@
 tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
-tab_bmr_birds_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
+#tab_bmr_birds_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
 knitr::kable(tab_bmr_birds_asym,
-             format="latex",linesep = "",
+             format="latex",linesep = "",escape=F,
              caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
@@
 tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
-tab_bmr_mammals_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
+#tab_bmr_mammals_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
 knitr::kable(tab_bmr_mammals_asym,
-             format="latex",linesep = "",
+             format="latex",linesep = "",escape=F,
              caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
@@
 tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
-tab_cc_asym_2000[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
+#tab_cc_asym_2000[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
 knitr::kable(tab_cc_asym_2000,
-             format="latex",linesep = "",
+             format="latex",linesep = "",escape=F,
              caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
@@
 ```{r tab_count_AICw, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_count$info_ord_AICw[,,1],3), 
-             format='latex',linesep = "",
+knitr::kable(RES_aphid_count$info_ord_AICw[,,1], 
+             format='latex',linesep = "",digits=3,
              caption.short = "AIC weights, best-fitting copulas, green spruce aphid abundance data",
@@
 source("normal_cop_AICw.R")
-knitr::kable(round(normal_cop_AICw_matrix(r=RES_aphid_count),3),
-             format='latex',linesep = "",
+knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
+             format='latex',linesep = "",digits=3,
              caption.short = "AIC weights, N copulas, green spruce aphid abundance data",
@@
 ```{r tab_count_p_CvM, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_count$gfc_p_CvM,3),
-             format='latex',linesep = "",
+knitr::kable(RES_aphid_count$gfc_p_CvM,
+             format='latex',linesep = "",digits=3,
              caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
@@
 ```{r tab_count_p_KS, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_count$gfc_p_KS,3),
-             format='latex',linesep = "",
+knitr::kable(RES_aphid_count$gfc_p_KS,
+             format='latex',linesep = "",digits=3,
              caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
@@
 ```{r tab_count_LTdep_AICw, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_count$LTdep_AICw,3),
-             format='latex',linesep = "",
+knitr::kable(RES_aphid_count$LTdep_AICw,
+             format='latex',linesep = "",digits=3,
              caption.short="Lower-tail dependence, green spruce aphid abundance data",
@@
 ```{r tab_count_UTdep_AICw, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_count$UTdep_AICw,3),
-             format='latex',linesep = "",
+knitr::kable(RES_aphid_count$UTdep_AICw,
+             format='latex',linesep = "",digits=3,
              caption.short="Upper-tail dependence, green spruce aphid abundance data",
@@
 ```{r tab_count_LTmUT, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,3),
-             digits = 5,linesep = "",
+knitr::kable(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,
+             digits = 3,linesep = "",
              format='latex',
@@
 s<-cbind(sc,s)
-colnames(s)<-c("","2.5th quantile", "Mean", "97.5th quantile")
-knitr::kable(s,
+knitr::kable(s, digits = 3,
+             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
              format = "latex",linesep = "", escape = F,align="c",
@@
 ```{r tab_ff_AICw, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_ff$info_ord_AICw[,,1],3), 
-             format='latex',linesep = "",
+knitr::kable(RES_aphid_ff$info_ord_AICw[,,1], 
+             format='latex',linesep = "",digits=3,
              caption.short="AIC weights, best-fitting copulas, leaf-curling plum aphid first flight data",
@@
 source("normal_cop_AICw.R")
-knitr::kable(round(normal_cop_AICw_matrix(r=RES_aphid_ff),3),
-             format='latex',linesep = "",
+knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
+             format='latex',linesep = "",digits=3,
              caption.short="AIC weights, N copulas, leaf-curling plum aphid first flight data",
@@
 ```{r tab_ff_p_CvM, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_ff$gfc_p_CvM,3),
-             format='latex',linesep = "",
+knitr::kable(RES_aphid_ff$gfc_p_CvM,
+             format='latex',linesep = "",digits=3,
              caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
@@
 ```{r tab_ff_p_KS, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_ff$gfc_p_KS,3),
-             format='latex',linesep = "",
+knitr::kable(RES_aphid_ff$gfc_p_KS,
+             format='latex',linesep = "",digits=3,
              caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
@@
 ```{r tab_ff_LTdep_AICw, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_ff$LTdep_AICw,3),
-             format='latex',linesep = "",
+knitr::kable(RES_aphid_ff$LTdep_AICw,
+             format='latex',linesep = "",digits=3,
              caption.short="Lower-tail dependence, leaf-curling plum aphid first flight data",
@@
 ```{r tab_ff_UTdep_AICw, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_ff$UTdep_AICw,3),
-             format='latex',linesep = "",
+knitr::kable(RES_aphid_ff$UTdep_AICw,
+             format='latex',linesep = "",digits=3,
              caption.short="Upper-tail dependence, leaf-curling plum aphid first flight data",
@@
 ```{r tab_ff_LTmUT, echo=F, results='markup'}
-knitr::kable(round(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,3),
-             digits = 5,
+knitr::kable(RES_aphid_ff$LTdep_AICw-RES_aphid_ff$UTdep_AICw,
+             digits = 3,
              format='latex',linesep = "",
@@
 h[is.infinite(h)]<-NA
-knitr::kable(round(h,3),
+knitr::kable(h,
              format="latex",
@@
 h[is.infinite(h)]<-NA
-knitr::kable(round(h,3),
+knitr::kable(h,
              format="latex",linesep = "",
@@
 h[is.nan(h)]<-NA
-knitr::kable(round(h,3),
+knitr::kable(h,
              format="latex",linesep = "",
@@
 s<-cbind(sc,s)
-colnames(s)<-c("","2.5th quantile", "Mean", "97.5th quantile")
-knitr::kable(s,
+knitr::kable(s,digits = 3,
              format = "latex",linesep = "", escape = F,align="c",
+             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
              caption.short="Results for nonparametric statistics, methane flux data",
@@
 rownames(summary_stat)<-c("bestfit.copula.AICw","Normal.copula.AICw","LT.model.avg","UT.model.avg","p.CvM","p.KS","BS.success.percent")
-knitr::kable(summary_stat, booktabs = T, 
-             format = "latex",linesep = "",
-             digits=3,
+knitr::kable(summary_stat, booktabs = T, escape=F,
+             format = "latex", linesep = "",
+             digits=3, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$"),
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
@@
 knitr::kable(temp, booktabs = T, 
-             format = "latex",linesep = "",
+             format = "latex",linesep = "",escape=F,
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",

### 2019-01-10 Paper.Rmd
finished (for now) editing the Q1 sections and related materials in supp mat

--- 2019-01-10 Paper.Rmd
@@
 title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
-author: "Shyamolina Ghosh, Lawrence W Sheppard, Mark T. Holder, Terrance E. Loecke, P. Chris Reid, Daniel C. Reuman"
+author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, P. Chris Reid, Daniel C. Reuman"
 date: "December 16, 2018"
@@
 \end{figure}
-<!--***DAN: Shya, I changed the abbreviation for Kendall Correlation from KC to K in the 
-caption, please change in the code that generates the figures, as well
-***Shya: Done.-->
 
@@
 from each other, answering the third part of Q1. The numbers reflect stronger upper- than lower-tail
-dependence for the soil C and N data, and this can also be seen visually in the extreme tails 
-of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).
+dependence for the soil C and N data, and this can also be seen visually in extreme upper-right
+and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).
 
@@
 Table \ref{tab_bivar_summary} summarizes bivariate results for Q1. For all $4$ 
-bivariate datasets, the two variables were always significantly related (1$^{st}$ row of 
-Table \ref{tab_bivar_summary}). A non-normal copula 
+bivariate datasets, the two variables were always significantly related 
+(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
 always emerged as the lowest-AIC (highest-AIC-weight) copula 
-(2$^{nd}$ row of Table \ref{tab_bivar_summary}). 
+(Table \ref{tab_bivar_summary}, row 2). 
 The N copula was always a poor fit compared to the highest-AIC-weight copula 
-(3$^{rd}$ and 4$^{th}$ rows of Table \ref{tab_bivar_summary}), answering the first part of Q1. 
-Upper- and lower-tail statistics (7$^{th}$ and 8$^{th}$ rows of 
-Table \ref{tab_bivar_summary}) differed substantially from 0, the tail dependence statistics of 
+(Table \ref{tab_bivar_summary}, rows 3-4), answering the first part of Q1. 
+Upper- and lower-tail statistics (Table \ref{tab_bivar_summary}, rows 7-8) 
+differed substantially from 0, the tail dependence statistics of 
 the N copula, answering the second part of Q1; and also differed substantially from each other
-(9$^{th}$ row of Table \ref{tab_bivar_summary}),
-answering the third part of Q1. The first three datasets had stronger upper- than lower-tail dependence, and the last dataset had the reverse. 
+(Table \ref{tab_bivar_summary}, row 9),
+answering the third part of Q1. The first three datasets had stronger upper- 
+than lower-tail dependence, and the last dataset had the reverse. 
 Although these results convincingly show non-normal copula structure, the 
@@
 because even the lowest-AIC copula was a poor fit in some cases 
-(5$^{th}$ and 6$^{th}$ rows of Table \ref{tab_bivar_summary}).
-But our nonparametric statistics confirmed the same conclusions (10$^{th}$ row of 
-Table \ref{tab_bivar_summary}).
+(Table \ref{tab_bivar_summary}, rows 5-6).
+But our nonparametric statistics confirmed the same conclusions 
+(Table \ref{tab_bivar_summary}, row 10).
 Detailed results for each dataset are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} 
@@
 
-We present our analysis of the green spruce aphid abundance data (see Data).
+We present our analysis of the green spruce aphid abundance data (see Data)
+as an example multivariate analysis.
 Independence was tested for each pair of the 10 sampling locations, and was 
@@
 appear anywhere in Table S11, this is correct, but please confirm or correct.
-***Shya: Yes, you are right, all the pairwise sampling locations for aphid data rejected independence test and show copula structures.-->
+***Shya: Yes, you are right, all the pairwise sampling locations for aphid data rejected independence test and show copula structures.
+***DAN: Good, thanks, leaving these comments here as a reminder-->
 Model selection among copulas was carried out separately for each pair of locations.
@@
 <!--Non-parametric analysis of tail dependence, green spruce aphid abundance data-->
-Nonparametric statistics verified both that tail dependence was distinct from that 
-of an N copula for the green spruce aphid abundance data, and that it was asymmetric.
+Nonparametric statistics verified that tail dependence was asymmetric for the green 
+spruce aphid abundance data.
 Mean values across all pairs of locations of the statistics 
 $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ were positive and
-confidence intervals excluded $0$ (the value for an N copula), confirming tail 
-dependence distinct from that of an N copula (Table \ref{SM-tab_count_resamp}). 
+confidence intervals excluded $0$ (Table \ref{SM-tab_count_resamp}). 
 Mean values of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
@@
 multivariate data. In addition to summaries of the green spruce aphid abundance data results 
-elaborated in detail above, the table also summarizes results of analyzing 
+elaborated in detail above, the table also summarizes results for 
 leaf-curling plum aphid first flight data, abundance data for the dinoflagellate
-\textit{Ceratium furca}, and methan flux data (see Data).
-
-We did model selection and non-parametric analysis with all 20 aphid and 22 plankton species and found our two approaches were consistent to each other.
-Here, to present our results we chose one species from each category.
-
-We found major occurence of non-independent copulae (row 2, Table \ref{tab_multivar_summary}) out of all pairwise locations (row 1, Table \ref{tab_multivar_summary}) considered.
-Non-Normal copulae, most of the time, emerged as best-fit based on minimum AIC out of 16 copula family considered (row 3, Table \ref{tab_multivar_summary}) answering the first part of question \hyperref[Intro]{\textbf{(Q1)}}.
-All four data sets showed Normal copulae were always poor fit in comparison to best-fit non-Normal copulae (compare their average AIC-weights from row 4 and row 5 of Table \ref{tab_multivar_summary}).
-We considered the "good-fit" location pair based on the p-values (>0.01) obtained from both Cramer-von Mises and Kolmogorov-Smirnov statistics (row 6, Table \ref{tab_multivar_summary}).
-Some spatial data (green spruce aphid's count, Ceratium furca's count and methane flux) showed stronger lower tail dependence (compare row 7 and row 8 from Table \ref{tab_multivar_summary}, also see row 9) whereas Leaf-plum curling's firstflight data showed stronger upper tail dependence which answered the second and third part of question \hyperref[Intro]{\textbf{(Q1)}}.
-The spatial synchrony (from model selection) results were also verified by the non-parametric statistic which showed exclusively positive (or negative) confidence interval of the distribution $(\Ps_l - \Ps_u)$ estimated from all location pair for lower (or upper)- tail dependent spatial data (except methane-flux data which had marginal lower quantile).
-Methane flux data (last column of Table \ref{tab_multivar_summary}) did show that there were non-Normal copula as best fit for most of the location pair but its model-averaged lower tail dependence was not significantly stronger in non-parametric analysis results.
-
+\textit{Ceratium furca}, and methane-flux data (see Data). Results supported the 
+conclusions that non-N copula structure, non-N tail dependence, and asymmetric tail 
+dependence were common. Most locations pairs were non-independent 
+(Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
+pairs had best-fitting copula that was not the N copula (Table \ref{tab_multivar_summary}, 
+row 3), and AIC weights of best-fitting copulas were 
+substantially better than AIC weights of N copulas (Table \ref{tab_multivar_summary}, rows 4-5).
+For these data, best-fitting copulas were nearly always considered an adequate fit, 
+according to Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests 
+(Table \ref{tab_multivar_summary}, row 6). Some of the datasets (green spruce aphid
+abundance, \textit{Ceratium furca} abundance, methane flux) showed stronger 
+or marginally stronger lower- than upper-tail dependence 
+(Table \ref{tab_multivar_summary}, rows 7-9), whereas leaf-curling plum aphid first flight
+data showed the reverse. Asymmetry results were verified by nonparametric approaches 
+(e.g., Table \ref{tab_multivar_summary}, row 10). Detailed results are in 
+Tables \ref{SM-tab_count_bestcop}-\ref{SM-tab_methane_resamp} and 
+Figs \ref{SM-fig_aphid_count}-\ref{SM-fig_methane}.
 We also carried out the same analyses for abundance and first-flight data for the 18 
-aphids species for which we had data other than the green spruce and leaf-curling plum aphids,
+aphid species for which we had data other than the green spruce and leaf-curling plum aphids,
 as well as for the 21 plankton taxa for which we had abundance data other 
@@
              booktabs=T, linesep = "\\addlinespace",
-             col.names = c(" ","Green-spruce aphid's count",
-                        "Leaf-plum curling aphid's firstflight",
-                        "Ceratium furca's count","Methane-flux")
+             col.names = c(" ","Green spruce aphid abundance",
+                        "Leaf-curling plum aphid first flight",
+                        "Ceratium furca abundance","Methane flux")
              )%>%column_spec(2:5,width="2.5cm")%>%
@@
 ```
+
+<!--***DAN: made it to here, 2018 1 10, 1252pm-->
 

### 2019-01-10 SupportingInformation.Rmd
finished (for now) editing the Q1 sections and related materials in supp mat

--- 2019-01-10 SupportingInformation.Rmd
@@
 title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
-author: "S. Ghosh, T. Loecke, C.P. Reid, D.C. Reuman"
+author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, P. Chris Reid, Daniel C. Reuman"
 geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"
@@
 
-<!--***DAN: text all good up to here, except as marked (the data section)-->
+<!--***DAN: text all good up to here, except as marked (the data section). Also
+all figs and tables relevant to the stuff this far have been edited and look pretty
+good. This is as of 2018 01 10, 1255pm. So start editing from this point.-->
 
@@
 knitr::kable(h, #longtable = T, 
-             format = "latex",linesep = "",
+             format = "latex",
       caption.short = "Model fitting results for soil C and N dataset",
@@
 \end{figure}
-<!--***DAN: Shya, can you please adjust this figure and the analogous 3 other figures for the other three bivariate datasets as follows: replace the y-axis label "Partial correlation" with "cor", for consistency with the other y-axis labels.
-***Shya: Done.-->
 
@@
 knitr::kable(tab_soilCN_asym, booktabs = T, 
-             format = "latex",linesep = "",
+             format = "latex",
       caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
@@
 knitr::kable(h,
-             format='latex',linesep = "",
+             format='latex',
              caption.short = "Model fitting results for the bird body mass and BMR dataset",
@@
 knitr::kable(tab_bmr_birds_asym,
-             format="latex",linesep = "",
+             format="latex",
              caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
@@
 knitr::kable(h,
-             format='latex',linesep = "",
+             format='latex',
              caption.short = "Model fitting results for the mammal body mass and BMR dataset",
@@
 knitr::kable(tab_bmr_mammals_asym,
-             format="latex",linesep = "",
+             format="latex",
              caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
@@
 knitr::kable(h,
-             format='latex',linesep = "",
+             format='latex',
              caption.short = "Model fitting results for the Cedar Creek dataset",
@@
 knitr::kable(tab_cc_asym_2000,
-             format="latex",linesep = "",
+             format="latex",
              caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
@@
 
-\newpage
 ```{r tab_count_bestcop, echo=F, results="markup"}
@@
 knitr::kable(RES_aphid_count$info_ord_copname[,,1], #longtable = T,
-             format='latex',linesep = "",
+             format='latex',
              caption.short = "Best-fitting copulas, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$info_ord_AICw[,,1],3), 
-             format='latex',linesep = "",
+             format='latex',
              caption.short = "AIC weights, best-fitting copulas, green spruce aphid abundance data",
@@
 knitr::kable(round(normal_cop_AICw_matrix(r=RES_aphid_count),3),
-             format='latex',linesep = "",
+             format='latex',
              caption.short = "AIC weights, N copulas, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$gfc_p_CvM,3),
-             format='latex',linesep = "",
+             format='latex',
              caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$gfc_p_KS,3),
-             format='latex',linesep = "",
+             format='latex',
              caption.short="Goodness of fit of best-fitting copulas, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$LTdep_AICw,3),
-             format='latex',linesep = "",
+             format='latex',
              caption.short="Lower-tail dependence, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$UTdep_AICw,3),
-             format='latex',linesep = "",
+             format='latex',
              caption.short="Upper-tail dependence, green spruce aphid abundance data",
@@
 knitr::kable(round(RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw,3),
-             digits = 5,linesep = "",
+             digits = 5,
              format='latex',
@@
 knitr::kable(round(s,3),
-             format = "latex",linesep = "",
+             format = "latex",
              caption.short="Results for nonparametric statistics, green spruce aphid abundance data",
@@
 knitr::kable(RES_aphid_ff$info_ord_copname[,,1], #longtable = T,
-             format='latex',linesep = "",
+             format='latex',
              caption.short = "Best-fitting copulas, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(RES_aphid_ff$info_ord_AICw[,,1],3), 
-             format='latex',linesep = "",
+             format='latex',
              caption.short="AIC weights, best-fitting copulas, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(normal_cop_AICw_matrix(r=RES_aphid_ff),3),
-             format='latex',linesep = "",
+             format='latex',
              caption.short="AIC weights, N copulas, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(RES_aphid_ff$gfc_p_CvM,3),
-             format='latex',linesep = "",
+             format='latex',
              caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(RES_aphid_ff$gfc_p_KS,3),
-             format='latex',linesep = "",
+             format='latex',
              caption.short="Goodness of fit of best-fitting copulas, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(RES_aphid_ff$LTdep_AICw,3),
-             format='latex',linesep = "",
+             format='latex',
              caption.short="Lower-tail dependence, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(RES_aphid_ff$UTdep_AICw,3),
-             format='latex',linesep = "",
+             format='latex',
              caption.short="Upper-tail dependence, leaf-curling plum aphid first flight data",
@@
              digits = 5,
-             format='latex',linesep = "",
+             format='latex',
              caption.short="Lower- minus upper-tail dependence, leaf-curling plum aphid first flight data",
@@
 knitr::kable(round(s,3),
-             format = "latex",linesep = "",
+             format = "latex",
              caption.short="Results for nonparametric statistics, leaf-curling plum aphid first flight data",
@@
 knitr::kable(RES_plankton_north_sea$info_ord_copname[,,1], #longtable=T,
-             format='latex', linesep = "",
+             format='latex', 
              caption.short = "Best-fitting copulas, \\textit{Ceratium furca} abundance data",
@@
              format="latex",
-             digits=3,linesep = "",
+             digits=3,
              caption.short="AIC weights, best-fitting copulas, \\textit{Ceratium furca} abundance data",
@@
 knitr::kable(round(h,3),
-             format="latex",linesep = "",
+             format="latex",
              digits=3,
@@
 knitr::kable(round(s,3),
-             format = "latex",linesep = "",
+             format = "latex",
              caption.short="Results for nonparametric statistics, \\textit{Ceratium furca} abundance data",
@@
 knitr::kable(RES_methane$info_ord_copname[,,1], #longtable=T,
-             format='latex',linesep = "",
+             format='latex',
              caption.short = "Best-fitting copulas, methane flux data",
@@
 knitr::kable(round(s,3),
-             format = "latex",linesep = "",
+             format = "latex",
              caption.short="Results for nonparametric statistics, methane flux data",
@@
 knitr::kable(summary_stat, booktabs = T, 
-             format = "latex",linesep = "",
+             format = "latex",
              digits=3,
@@
 knitr::kable(temp, booktabs = T, 
-             format = "latex",linesep = "",
+             format = "latex",
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",

### 2019-01-10 Paper.Rmd

@@
 knitr::kable(tab_multivar_summary, 
-             format="latex", 
+             format="latex", escape = F,
              caption.short = "Summary of results for multivariate data sets",

### 2019-01-11 ext_risk_copula.Rmd
modified

@@
 
-rlist<-seq(from=5,to=10,by=1)
+rlist<-seq(from=1,to=10,by=1)
 a<-0.5
@@
 pdf("./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(2,4),mar=c(5,4,1,1) + 0.2)
+op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
 for(r in rlist){
@@
 pdf("./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(2,4),mar=c(5,4,1,1) + 0.2)
+op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
 for(r in rlist){
@@
 
-```{r fig_msmith, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r fig_msmith_r_vary, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+set.seed(seed)
+
+rlist<-seq(from=3,to=8,by=1)
+b<-1.4
+a<-0.5
+K_e<-((r-1)^(1/b))/a
+
+# plot for local dispersal in D matrix 
+pdf("./Results/ext_risk_copula/msmith_ext_risk_local_disp_r_variation.pdf",heigh=6,width=10)
+op<-par(mfrow=c(2,3),mar=c(5,4,1,1) + 0.2)
+for(r in rlist){
+    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=K_e,a=a,b=b,model="msmith",disp_everywhere=F,ploton = F)
+    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s$d_seq,s$risk_right,type="b",col="blue")
+    legend(x=-0.06,y=0.4, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+
+# plot for global dispersal in D matrix 
+pdf("./Results/ext_risk_copula/msmith_ext_risk_global_disp_r_variation.pdf",heigh=6,width=10)
+op<-par(mfrow=c(2,3),mar=c(5,4,1,1) + 0.2)
+for(r in rlist){
+    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=K_e,a=a,b=b,model="msmith",disp_everywhere=T,ploton = F)
+    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
+    lines(s$d_seq,s$risk_right,type="b",col="blue")
+    legend(x=-0.06,y=0.4, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+    mtext(paste0("r = ", r))
+}
+par(op)
+dev.off()
+```
+
+![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, b = 1.4, rc=4.5) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/ext_risk_copula/msmith_ext_risk_local_disp_r_variation.pdf){ width=40%, height=40% }
+
+![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, b=1.4, rc=4.5) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/ext_risk_copula/msmith_ext_risk_global_disp_r_variation.pdf){ width=40%, height=40% }
+
+```{r fig_msmith_b_vary, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
@@
 # plot for local dispersal in D matrix 
-pdf("./Results/ext_risk_copula/msmith_ext_risk_local_disp.pdf",heigh=6,width=10)
+pdf("./Results/ext_risk_copula/msmith_ext_risk_local_disp_b_variation.pdf",heigh=6,width=10)
 op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
@@
 # plot for global dispersal in D matrix 
-pdf("./Results/ext_risk_copula/msmith_ext_risk_global_disp.pdf",heigh=6,width=10)
+pdf("./Results/ext_risk_copula/msmith_ext_risk_global_disp_b_variation.pdf",heigh=6,width=10)
 op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r = 5, bc=1.4) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp_msmith}](./Results/ext_risk_copula/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r = 5, bc=1.4) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp_msmith_b_vary}](./Results/ext_risk_copula/msmith_ext_risk_local_disp_b_variation.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r=5, bc=1.4) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp_msmith}](./Results/ext_risk_copula/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r=5, bc=1.4) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp_msmith_b_vary}](./Results/ext_risk_copula/msmith_ext_risk_global_disp_b_variation.pdf){ width=40%, height=40% }

### 2019-01-11 1var_pop_models.R

@@ -49,9 +49,14 @@ Hassell<-function(r,a,b,p0,lensim){
 }
 
 #call the function
-Hassell(r=5, a=0.5, b=2.5, p0=0.1, lensim=200) #b~1.5 : contest, b~100 : scramble
+r=5
+a=0.5
+b=1.5
+K_e<-((r^(1/b))-1)/a
+K_e
+Hassell(r=r, a=a, b=b, p0=K_e, lensim=200) #b~1.5 : contest, b~100 : scramble
 
-Hassell(r=4, a=0.5, b=5.1, p0=0.1, lensim=200) # see change : r=5,a=0.5 and vary b 1.5 to 5.5
+Hassell(r=5, a=0.01, b=2.5, p0=0.1, lensim=200) # see change : r=5,a=0.5 and vary b 1.5 to 5.5
 #----------------------------------------------------------------------------------------------------
 # Exploring Maynard Smith model
 Msmith<-function(r,a,b,p0,lensim){
@@ -70,7 +75,7 @@ Msmith<-function(r,a,b,p0,lensim){
 }
 
 #call the function
-Msmith(r=5,a=0.5,b=1.4,p0=0.1,lensim=200) #see changes as r=5,a=0.5 and vary b : 1,1.5,2,4
+Msmith(r=4.5,a=0.5,b=1.4,p0=0.1,lensim=200) #see changes as r=5,a=0.5 and vary b : 1,1.5,2,4
                                           #see changes as a=0.5,b=4 and vary r : 1.2, 1.5, 2
 
 

### 2019-01-14 ext_risk_copula.Rmd
added results for ricker and hassell, need to fix msmith

--- 2019-01-07 ext_risk_copula.Rmd
@@
 
-rlist<-c(0,0.05,0.25,0.5,0.7,0.75,0.9,1.0,1.25,1.5,1.75,1.95)
+rlist<-c(0,0.25,0.5,0.75,1.0,1.25,1.5,1.75,1.95)
+numsims<-10000
+numsteps<-25
+numlocs<-5
+a<-NA
+b<-NA
+K<-50
+scl<-1
+ext_thrs<-K/10
 
 # plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-pdf("./Results/ext_risk_copula/ricker_ext_risk_local_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
+pdf("./Results/ext_risk_copula/ricker_ext_risk_local_disp.pdf",height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1,1) + 0.2)
 for(r in rlist){
     #cat("r=",r,"\n")
-    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=50,a=NA,b=NA,model="ricker",disp_everywhere=F,ploton = F)
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+                 scl=scl,model="ricker",disp_everywhere=F,ploton = F)
     plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
     lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n',cex=1.2)
     mtext(paste0("r = ", r))
@@
 # plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-pdf("./Results/ext_risk_copula/ricker_ext_risk_global_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
+pdf("./Results/ext_risk_copula/ricker_ext_risk_global_disp.pdf",height=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1,1) + 0.2)
 for(r in rlist){
     #cat("r=",r,"\n")
-    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=50,a=NA,b=NA,model="ricker",disp_everywhere=T,ploton = F)
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+                 scl=scl,model="ricker",disp_everywhere=T,ploton = F)
     plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
     lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
     mtext(paste0("r = ", r))
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_ricker_risk_local_disp}](./Results/ext_risk_copula/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker_risk_local_disp}](./Results/ext_risk_copula/ricker_ext_risk_local_disp.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/ext_risk_copula/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_ricker__risk_global_disp}](./Results/ext_risk_copula/ricker_ext_risk_global_disp.pdf){ width=40%, height=40% }
 
@@
 
-rlist<-seq(from=5,to=10,by=1)
+rlist<-seq(from=1.5,to=7.5,by=0.5)
+numsims<-10000
+numsteps<-25
+numlocs<-5
 a<-0.5
-b<-1.5
+b<-100
 K_e<-((r^(1/b))-1)/a
+scl<-1
+K<-K_e
+ext_thrs<-K/10
 rc<-((b/(b-1)))^b # critical value of r from monotonic to osc steady state approach transition
@@
 # plot for local dispersal in D matrix 
-pdf("./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(2,4),mar=c(5,4,1,1) + 0.2)
+pdf("./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf",heigh=9,width=15)
+op<-par(mfrow=c(3,5),mar=c(5,4,1,1) + 0.2)
 for(r in rlist){
     #cat("r=",r,"\n")
-    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=K_e,a=a,b=b,model="hassell",disp_everywhere=F,ploton = F)
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+                 scl=scl,model="hassell",disp_everywhere=F,ploton = F)
     plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
     lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend(x=-0.06,y=0.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+    legend("topleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
     mtext(paste0("r = ", r))
@@
 # plot for local dispersal in D matrix 
-pdf("./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(2,4),mar=c(5,4,1,1) + 0.2)
+pdf("./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf",heigh=9,width=15)
+op<-par(mfrow=c(3,5),mar=c(5,4,1,1) + 0.2)
 for(r in rlist){
     #cat("r=",r,"\n")
-    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=K_e,a=a,b=b,model="hassell",disp_everywhere=T,ploton = F)
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+                 scl=scl,model="hassell",disp_everywhere=T,ploton = F)
     plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
     lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend(x=-0.06,y=0.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
+    legend("topleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
     mtext(paste0("r = ", r))
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Hassell model : a = 0.5, b = 1.5, rc=5.19) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc=2.73) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Hassell model : a = 0.5, b = 1.5, rc=5.19) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc=2.73) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
 
@@
 
-```{r fig_msmith, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r fig_msmith_r_vary, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
 
-blist<-seq(from=1,to=2,by=0.2)
-r<-5
+rlist<-seq(from=1.1,to=1.9,by=0.2)
+numsims<-10000
+numsteps<-25
+numlocs<-5
 a<-0.5
+b<-4
 K_e<-((r-1)^(1/b))/a
+K<-K_e
+scl<-1
+ext_thrs<-K/10
 
 # plot for local dispersal in D matrix 
-pdf("./Results/ext_risk_copula/msmith_ext_risk_local_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
-for(b in blist){
-    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=K_e,a=a,b=b,model="msmith",disp_everywhere=F,ploton = F)
+pdf("./Results/ext_risk_copula/msmith_ext_risk_local_disp_r_variation.pdf",heigh=3,width=15)
+op<-par(mfrow=c(1,5),mar=c(5,4,1,1) + 0.2)
+for(r in rlist){
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+                 scl=scl,model="msmith",disp_everywhere=F,ploton = F)
     plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
     lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend(x=-0.06,y=0.4, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
-    mtext(paste0("b = ", b))
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
 }
@@
 # plot for global dispersal in D matrix 
-pdf("./Results/ext_risk_copula/msmith_ext_risk_global_disp.pdf",heigh=6,width=10)
-op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
-for(b in blist){
-    s<-varying_d(numsims=10000,numsteps =10,numlocs=5,r=r,K=K_e,a=a,b=b,model="msmith",disp_everywhere=T,ploton = F)
+pdf("./Results/ext_risk_copula/msmith_ext_risk_global_disp_r_variation.pdf",heigh=3,width=15)
+op<-par(mfrow=c(1,5),mar=c(5,4,1,1) + 0.2)
+for(r in rlist){
+    s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
+                 scl=scl,model="msmith",disp_everywhere=T,ploton = F)
     plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
     lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend(x=-0.06,y=0.4, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
-    mtext(paste0("b = ", b))
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    mtext(paste0("r = ", r))
 }
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r = 5, bc=1.4) with local dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_local_disp_msmith}](./Results/ext_risk_copula/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 0.5, b = 4, rc=1.5) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/ext_risk_copula/msmith_ext_risk_local_disp_r_variation.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 10 years (Maynard Smith model : a = 0.5, r=5, bc=1.4) with global dispersion; numlocs=5, numsteps=10, numsims=10000\label{fig_risk_global_disp_msmith}](./Results/ext_risk_copula/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 0.5, b=4, rc=1.5) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/ext_risk_copula/msmith_ext_risk_global_disp_r_variation.pdf){ width=40%, height=40% }

### 2019-01-16 ext_risk_copula.Rmd
fixed bug

@@
 b<-100
-K_e<-((r^(1/b))-1)/a
 scl<-1
-K<-K_e
-ext_thrs<-K/10
+
 rc<-((b/(b-1)))^b # critical value of r from monotonic to osc steady state approach transition
@@
     #cat("r=",r,"\n")
+    K_e<-((r^(1/b))-1)/a
+    K<-K_e
+    ext_thrs<-K/10
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc=2.73) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_local_disp.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc=2.73) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Hassell model : a = 0.5, b = 100, rc = 2.73) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_hassell}](./Results/ext_risk_copula/hassell_ext_risk_global_disp.pdf){ width=40%, height=40% }
 
@@
 
-```{r fig_msmith_r_vary, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r fig_msmith_r_vary_localdisp, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 set.seed(seed)
-
-rlist<-seq(from=1.1,to=1.9,by=0.2)
+rlist<-seq(from=1.2,to=2.8,by=0.2)
 numsims<-10000
@@
 numlocs<-5
-a<-0.5
-b<-4
+a<-1
+b<-3
+scl<-1
+
+# plot for local dispersal in D matrix 
+pdf("./Results/ext_risk_copula/msmith_ext_risk_local_disp.pdf",heigh=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1,1) + 0.2)
+for(r in rlist){
     K_e<-((r-1)^(1/b))/a
     K<-K_e
-scl<-1
     ext_thrs<-K/10
-
-# plot for local dispersal in D matrix 
-pdf("./Results/ext_risk_copula/msmith_ext_risk_local_disp_r_variation.pdf",heigh=3,width=15)
-op<-par(mfrow=c(1,5),mar=c(5,4,1,1) + 0.2)
-for(r in rlist){
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
@@
 # plot for global dispersal in D matrix 
-pdf("./Results/ext_risk_copula/msmith_ext_risk_global_disp_r_variation.pdf",heigh=3,width=15)
-op<-par(mfrow=c(1,5),mar=c(5,4,1,1) + 0.2)
+pdf("./Results/ext_risk_copula/msmith_ext_risk_global_disp.pdf",heigh=9,width=9)
+op<-par(mfrow=c(3,3),mar=c(5,4,1,1) + 0.2)
 for(r in rlist){
+    K_e<-((r-1)^(1/b))/a
+    K<-K_e
+    ext_thrs<-K/10
     s<-varying_d(numsims=numsims,numsteps=numsteps,numlocs=numlocs,r=r,K=K,a=a,b=b,ext_thrs=ext_thrs,
@@
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 0.5, b = 4, rc=1.5) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/ext_risk_copula/msmith_ext_risk_local_disp_r_variation.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 1.7) with local dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp_msmith_r_vary}](./Results/ext_risk_copula/msmith_ext_risk_local_disp.pdf){ width=40%, height=40% }
 
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 0.5, b=4, rc=1.5) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/ext_risk_copula/msmith_ext_risk_global_disp_r_variation.pdf){ width=40%, height=40% }
+![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Maynard Smith model : a = 1, b = 3, rc = 1.7) with global dispersion; numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp_msmith_r_vary}](./Results/ext_risk_copula/msmith_ext_risk_global_disp.pdf){ width=40%, height=40% }

### 2019-01-16 1var_pop_models.R

@@ -49,9 +49,9 @@ Hassell<-function(r,a,b,p0,lensim){
 }
 
 #call the function
-r=1
-a=0.5
-b=100
+r<-1.1
+a<-0.5
+b<-1
 K_e<-((r^(1/b))-1)/a
 K_e
 Hassell(r=r, a=a, b=b, p0=0.1, lensim=50) #b~1.5 : contest, b~100 : scramble
@@ -75,13 +75,48 @@ Msmith<-function(r,a,b,p0,lensim){
 }
 
 #call the function
-r<-2
-a<-0.5
+r<-2.8
+a<-1
 b<-3
 K_e<-((r-1)^(1/b))/a
 K_e
 Msmith(r=r,a=a,b=b,p0=K_e-0.1,lensim=200) #see changes as r=5,a=0.5 and vary b : 1,1.5,2,4
                                           #see changes as a=0.5,b=4 and vary r : 1.2, 1.5, 2
+#--------------------------------------------------------------------------------------------------
+
+# Exploring Verhulst model
+Verhulst<-function(r,p0,lensim){
+  pop<-c(p0)
+  time<-c(0)
+  # po<-rep(K,r)
+  # print(po)
+  
+  for(it in c(1:lensim)){
+    pt<-p0*(1+r*(1-p0))
+    time<-c(time,it)
+    pop<-c(pop,pt)
+    p0<-pt
+  }
+  
+  plot(time,pop,type="b")
+}
+
+#call the function
+Verhulst(r=1.2,p0=0.4,lensim=400) # eqm. point =1
+
+
+
+
+
+
+
+
+
+
+
+
+
+
 
 
 

### 2019-01-16 mydispersal.R

@@ -71,12 +71,15 @@ popsim_ml_D<-function(p0,ns,D,r,K,a,b,ext_thrs,model){
     lam_sto<-exp(ns[,,tct])  #numsims by numlocs matrix
     
     lam_ricker<-exp(r*(1-(res[,,tct]/K)))
+    lam_verhulst<-1+(r*(1-res[,,tct]))
     lam_hassell<-r/((1+(a*res[,,tct]))^b)
     lam_msmith<-r/(1+((a*res[,,tct])^b))
     
     #growth prior to dispersal
     if(model=="ricker"){
       given_model<-lam_ricker
+    }else if(model=="verhulst"){
+      given_model<-lam_verhulst
     }else if(model=="hassell"){
       given_model<-lam_hassell
     }else if(model=="msmith"){

### 2019-01-18 SupportingInformation.Rmd
test run with neg beta ar1 model

@@
 ```{r cause4copula_pGOF_500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
+#source("Cause4copula.R")
 #-----------------
@@
 #---------------------------------------
-pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf", width=8, height=3)
-op<-par(mfrow=c(1,2))
-
-Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=500,BS=1000)
+#pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf", width=8, height=3)
+#op<-par(mfrow=c(1,2))
+
+#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=500,BS=1000)
 #Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=500,BS=1000)
@@
 
-Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=500,BS=1000)
 #Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=500,BS=1000)
@@
 
-op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
-plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", "magenta", "green2"),
-       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
-       bty = "n") 
-par(op)
-dev.off()
+#op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
+#plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
+#legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", #"magenta", "green2"),
+#       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
+#       bty = "n") 
+#par(op)
+#dev.off()
 ```
@@
 ```{r cause4copula_pGOF_2500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
+#source("Cause4copula.R")
 #-----------------
-set.seed(seed)
+#set.seed(seed)
 #---------------------------------------
 
-pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=8, height=3)
-op<-par(mfrow=c(1,2))
-
-Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
+#pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=8, height=3)
+#op<-par(mfrow=c(1,2))
+
+#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
 #Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=2500,BS=1000)
@@
 
-Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
+#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
 #Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=2500,BS=1000)
@@
 
-op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
-plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", "magenta", "green2"),
-       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
-       bty = "n") 
-par(op)
-dev.off()
-```
-
-```{r cause4copula_stat, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+#op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
+#plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
+#legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", #"magenta", "green2"),
+#       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
+#       bty = "n") 
+#par(op)
+#dev.off()
+```
+
+```{r cause4copula_stat_clayton_ar1_posbeta, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+#source("Cause4copula.R")
+#-----------------
+#set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model 
+#resloc<-"./Results/Cause4copula_results/"
+#model<-"ar1"
+#params<-c(0.5) # this parameter is beta in ar1 model
+#tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,sep="")
+
+#if (!dir.exists(tempo)){
+#  dir.create(tempo)
+#}
+
+#resloc<-paste(tempo,"/",sep="")
+
+# for clayton with spearman and kendall method
+#Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+#                          params=params,p0=c(0,0),model="ar1")
+
+#Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+#                          params=params,p0=c(0,0),model="ar1")
+
+# for survival clayton with spearman and kendall method
+#Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+#                          params=params,p0=c(0,0),model="ar1")
+
+#Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+#                          params=params,p0=c(0,0),model="ar1")
+```
+
+```{r cause4copula_stat_clayton_ar1_negbeta, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
 source("Cause4copula.R")
@@
 #---------------------------------------
+# This chunk stores results for ar1 model 
 resloc<-"./Results/Cause4copula_results/"
-if (!dir.exists(paste(resloc,"Cause4copula_stat_results",sep=""))){
-  dir.create(paste(resloc,"Cause4copula_stat_results",sep=""))
-}
-resloc<-paste(resloc,"Cause4copula_stat_results/",sep="")
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc)                
-```
-
+model<-"ar1"
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+
+# for clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model="ar1")
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model="ar1")
+
+# for survival clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model="ar1")
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model="ar1")               
+```
+
+<!--
 \begin{figure}[!h]

### 2019-01-18 Cause4copula.R

@@ -68,7 +68,10 @@ GetNoise<-function(N,fcode,corcoef,method,ploton){
 #Simulates a model for use in understanding Moran effects on population structure
 #
 #Args
-#cons         An autocorrelation coefficient for the model (one number, |cons|<1)
+#params       a vector  with 
+#               if model == "ar1" : an autocorrelation coefficient for the model (one number, |cons|<1)
+#               if model == "ricker" : with two number, 1. r, 2. K
+                            
 #p0           Initial conditions - length 2 vector, default c(0,0)
 #noise       An N by 2 matrix of environment variables in two habitat patches through time 
 #                             (this should be noise_q from output of GetNoise function)
@@ -78,15 +81,25 @@ GetNoise<-function(N,fcode,corcoef,method,ploton){
 #           pop_c : An N+1 by 2 matrix of populations through time, p0 is the first row in copula space
 #           pop_q : this is similar but with normal distribution marginal
 
-Simulator_Cause4copula<-function(cons,p0=c(0,0),noise){
+Simulator_Cause4copula<-function(params,p0=c(0,0),noise,model){
   
   N<-dim(noise)[1]
-  cons2<-sqrt(1-cons^2)
   res<-matrix(NA,N+1,2)
   res[1,]<-p0
   
   for (counter in 2:(N+1)){
-    res[counter,]<-(res[counter-1,]*cons)+(noise[counter-1,]*cons2)
+    if(model=="ar1"){
+      beta<-params
+      alpha<-sqrt(1-beta^2)
+      res[counter,]<-(beta*res[counter-1,])+(alpha*noise[counter-1,])
+    }else if(model=="ricker"){
+      r<-params[1]
+      K<-params[2]
+      res[counter,]<-res[counter-1,]*exp((r*(1-(res[counter-1,]/K)))+noise[counter-1,])
+    }else{
+      warning("model not specified",immediate.=T,call.=T)
+    }
+    
   }
   
   res2<-pnorm(res) # convert into copula space
@@ -169,7 +182,7 @@ comp<-function(s,s2,num_keep_last){
 #  parameter of noise copula vs. correln coef (spearman/kendall)
 # BS : number of bootstrapps used for BiCopGOFTest
 
[…]

### 2019-01-18 SupportingInformation.Rmd
adding chunk for C, SC : ar1 results with beta = 0.5, -0.5

@@
 
-```{r cause4copula_stat_clayton_ar1_posbeta, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-#source("Cause4copula.R")
-#-----------------
-#set.seed(seed)
-#---------------------------------------
-# This chunk stores results for ar1 model 
-#resloc<-"./Results/Cause4copula_results/"
-#model<-"ar1"
-#params<-c(0.5) # this parameter is beta in ar1 model
-#tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,sep="")
-
-#if (!dir.exists(tempo)){
-#  dir.create(tempo)
-#}
-
-#resloc<-paste(tempo,"/",sep="")
-
-# for clayton with spearman and kendall method
-#Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-#                          params=params,p0=c(0,0),model="ar1")
-
-#Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-#                          params=params,p0=c(0,0),model="ar1")
-
-# for survival clayton with spearman and kendall method
-#Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-#                          params=params,p0=c(0,0),model="ar1")
-
-#Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-#                          params=params,p0=c(0,0),model="ar1")
-```
-
-```{r cause4copula_stat_clayton_ar1_negbeta, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+```{r cause4copula_stat_ar1_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
 source("Cause4copula.R")
@@
 model<-"ar1"
-params<-c(-0.5) # this parameter is beta in ar1 model
-tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,sep="")
+params<-c(0.5) # this parameter is beta in ar1 model
+fcode<-3
+
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
 
@@
 }
-
 resloc<-paste(tempo,"/",sep="")
-
 # for clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model="ar1")
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=3,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model="ar1")
 
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+# for clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model="ar1")
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model="ar1")
+```
+
+```{r cause4copula_stat_ar1_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model 
+resloc<-"./Results/Cause4copula_results/"
+model<-"ar1"
+params<-c(0.5) # this parameter is beta in ar1 model
+fcode<-13
+
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
 # for survival clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model="ar1")
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=13,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model="ar1")
+
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+# for survival clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model="ar1")
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model="ar1")

### 2019-01-18 SupportingInformation.Rmd
minor edit

@@
 #---------------------------------------
-# This chunk stores results for ar1 model 
+# This chunk stores results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-3
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
 resloc<-"./Results/Cause4copula_results/"
-model<-"ar1"
-params<-c(0.5) # this parameter is beta in ar1 model
-fcode<-3
 
@@
 
+# for negative beta
+resloc<-"./Results/Cause4copula_results/"
 params<-c(-0.5) # this parameter is beta in ar1 model
@@
 #---------------------------------------
-# This chunk stores results for ar1 model 
+# This chunk stores results for ar1 model : survival clayton
+model<-"ar1"
+fcode<-13
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
 resloc<-"./Results/Cause4copula_results/"
-model<-"ar1"
-params<-c(0.5) # this parameter is beta in ar1 model
-fcode<-13
-
 tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
@@
 
+# for negative beta
 params<-c(-0.5) # this parameter is beta in ar1 model
+
+resloc<-"./Results/Cause4copula_results/"
 tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")

### 2019-01-19 SupportingInformation.Rmd
adding cause4copula chunk for ar1_beta_(0.5 and -0.5) and ricker(varying r, K=100) for 6 uniparameter copula: C,SC,G,SG,J,SJ

@@
 Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=c(0,0),model="ar1")
+                          params=params,p0=c(0,0),model=model)
 
 Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=c(0,0),model="ar1")
+                          params=params,p0=c(0,0),model=model)
 
@@
 Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=c(0,0),model="ar1")
+                          params=params,p0=c(0,0),model=model)
 
 Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=c(0,0),model="ar1")
+                          params=params,p0=c(0,0),model=model)
 ```
@@
 Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=c(0,0),model="ar1")
+                          params=params,p0=c(0,0),model=model)
 
 Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=c(0,0),model="ar1")
-```
-
+                          params=params,p0=c(0,0),model=model)
+```
+
+```{r cause4copula_stat_ar1_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-4
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+resloc<-"./Results/Cause4copula_results/"
+
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+# for clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+# for negative beta
+resloc<-"./Results/Cause4copula_results/"
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+# for clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+```
+
+```{r cause4copula_stat_ar1_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : survival clayton
+model<-"ar1"
+fcode<-14
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+resloc<-"./Results/Cause4copula_results/"
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+# for survival clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+
+resloc<-"./Results/Cause4copula_results/"
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+# for survival clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+```
+
+```{r cause4copula_stat_ar1_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-6
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+resloc<-"./Results/Cause4copula_results/"
+
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+# for clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+# for negative beta
+resloc<-"./Results/Cause4copula_results/"
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+# for clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+```
+
+```{r cause4copula_stat_ar1_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : survival clayton
+model<-"ar1"
+fcode<-16
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+resloc<-"./Results/Cause4copula_results/"
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+# for survival clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+
+resloc<-"./Results/Cause4copula_results/"
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+# for survival clayton with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=c(0,0),model=model)
+```
+
+```{r cause4copula_stat_ricker_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-3
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(i in rlist){
+  params<-c(i,100) # this parameter is a vector with r and K
+  p0<-c(params[2],params[2])
+
+  resloc<-"./Results/Cause4copula_results/"
+
+  tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+  if (!dir.exists(tempo)){
+   dir.create(tempo)
+  }
+  resloc<-paste(tempo,"/",sep="")
+
+# for clayton with spearman and kendall method
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+}
+```
+
+```{r cause4copula_stat_ricker_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-13
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(i in rlist){
+  params<-c(i,100) # this parameter is a vector with r and K
+  p0<-c(params[2],params[2])
+
+  resloc<-"./Results/Cause4copula_results/"
+
+  tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+  if (!dir.exists(tempo)){
+   dir.create(tempo)
+  }
+  resloc<-paste(tempo,"/",sep="")
+
+# for clayton with spearman and kendall method
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+}
+```
+
+```{r cause4copula_stat_ricker_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-4
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(i in rlist){
+  params<-c(i,100) # this parameter is a vector with r and K
+  p0<-c(params[2],params[2])
+
+  resloc<-"./Results/Cause4copula_results/"
+
+  tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+  if (!dir.exists(tempo)){
+   dir.create(tempo)
+  }
+  resloc<-paste(tempo,"/",sep="")
+
+# for clayton with spearman and kendall method
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+}
+```
+
+```{r cause4copula_stat_ricker_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-14
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(i in rlist){
+  params<-c(i,100) # this parameter is a vector with r and K
+  p0<-c(params[2],params[2])
+
+  resloc<-"./Results/Cause4copula_results/"
+
+  tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+  if (!dir.exists(tempo)){
+   dir.create(tempo)
+  }
+  resloc<-paste(tempo,"/",sep="")
+
+# for clayton with spearman and kendall method
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+}
+```
+
+```{r cause4copula_stat_ricker_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-6
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(i in rlist){
+  params<-c(i,100) # this parameter is a vector with r and K
+  p0<-c(params[2],params[2])
+
+  resloc<-"./Results/Cause4copula_results/"
+
+  tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+  if (!dir.exists(tempo)){
+   dir.create(tempo)
+  }
+  resloc<-paste(tempo,"/",sep="")
+
+# for clayton with spearman and kendall method
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+}
+```
+
+```{r cause4copula_stat_ricker_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-16
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(i in rlist){
+  params<-c(i,100) # this parameter is a vector with r and K
+  p0<-c(params[2],params[2])
+
+  resloc<-"./Results/Cause4copula_results/"
+
+  tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+  if (!dir.exists(tempo)){
+   dir.create(tempo)
+  }
+  resloc<-paste(tempo,"/",sep="")
+
+# for clayton with spearman and kendall method
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+
+  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+                          params=params,p0=p0,model=model)
+}
+```
 <!--

### 2019-01-19 Cause4copula.R

@@ -87,22 +87,26 @@ Simulator_Cause4copula<-function(params,p0=c(0,0),noise,model){
   res<-matrix(NA,N+1,2)
   res[1,]<-p0
   
-  for (counter in 2:(N+1)){
+  
     if(model=="ar1"){
       beta<-params
       alpha<-sqrt(1-beta^2)
-      res[counter,]<-(beta*res[counter-1,])+(alpha*noise[counter-1,])
+      for (counter in 2:(N+1)){
+        res[counter,]<-(beta*res[counter-1,])+(alpha*noise[counter-1,])
+      }
     }else if(model=="ricker"){
       r<-params[1]
       K<-params[2]
-      res[counter,]<-res[counter-1,]*exp((r*(1-(res[counter-1,]/K)))+noise[counter-1,])
+      for (counter in 2:(N+1)){
+        res[counter,]<-res[counter-1,]*exp((r*(1-(res[counter-1,]/K)))+noise[counter-1,])
+      }
     }else{
       warning("model not specified",immediate.=T,call.=T)
     }
-    
-  }
   
-  res2<-pnorm(res) # convert into copula space
+  # convert into copula space
+  #res2<-pnorm(res) 
+  res2<-VineCopula::pobs(res)
   return(list(pop_c=res2,pop_q=res))
   
 } 
@@ -118,6 +122,7 @@ Simulator_Cause4copula<-function(params,p0=c(0,0),noise,model){
 # This function gives mean,lowCI,upCI of a vector
 #
 MCI<-function(x){
+  x<-x[is.finite(x)] # only considering finite values
   m<-mean(x)
   se<-sd(x)/sqrt(length(x))   
   return(c(m-1.96*se,m,m+1.96*se))
@@ -170,11 +175,26 @@ comp<-function(s,s2,num_keep_last){
 }
 #-------------------------------------------------------------------
 # when noise comes from a clayton cop with spearmancor=0.8
-#s<-GetNoise(N=2000,fcode=3,corcoef=0.8,method="spearman",ploton=T)
-#s2<-Simulator_Cause4copula(cons=0.5,p0=c(0,0),noise=s$noise_q)
[…]

### 2019-01-19 SupportingInformation.Rmd
fixed error

@@
 resloc<-paste(tempo,"/",sep="")
-# for clayton with spearman and kendall method
+# for gumbel with spearman and kendall method
 Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
@@
 resloc<-paste(tempo,"/",sep="")
-# for survival clayton with spearman and kendall method
+# for survival gumbel with spearman and kendall method
 Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
@@
 resloc<-paste(tempo,"/",sep="")
-# for clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=c(0,0),model=model)
+# for joe with kendall method only
 
@@
 resloc<-paste(tempo,"/",sep="")
-# for survival clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=c(0,0),model=model)
+# for survival joe with kendall method only
 
@@
 
-# for clayton with spearman and kendall method
+# for survival clayton with spearman and kendall method
   Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
@@
 
-# for clayton with spearman and kendall method
+# for gumbel with spearman and kendall method
   Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
@@
 
-# for clayton with spearman and kendall method
+# for survival gumbel with spearman and kendall method
   Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
@@
 
-# for clayton with spearman and kendall method
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
+# for joe with kendall method only
 
@@
 
-# for clayton with spearman and kendall method
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
+# for survival joe with kendall method only
 

### 2019-01-19 Cause4copula.R

@@ -38,7 +38,7 @@ GetNoise<-function(N,fcode,corcoef,method,ploton){
     param<-iTau(copula = tgcop, tau = corcoef)
   }else{
     param<-NA
-    warning("specify method",immediate.=T,call.=T)
+    warning("specify method compatible with copula family",immediate.=T,call.=T)
   }
   
   noisecop<-BiCopSim(N=N, family=fcode, par=param)
@@ -222,7 +222,7 @@ Plotter_Cause4copula_GOF<-function(N,fcode,method,num_keep_last,BS,params,p0,mod
     # generate noise copula 
     s<-GetNoise(N=N,fcode=fcode,corcoef=corcoef,method=method,ploton=F)
     # generate pop_copula
-    s2<-Simulator_Cause4copula(params=params,p0=cp0,noise=s$noise_q,model=model)
+    s2<-Simulator_Cause4copula(params=params,p0=p0,noise=s$noise_q,model=model)
     
     # compare btw parameters of s$noise_c and s2$pop_c
     

### 2019-01-22 SupportingInformation.Rmd
for rlist and Klist100,1000 for all uniparameter family

@@
 
+Klist<-c(100,1000)
+
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
 rlist<-seq(from=0.1,to=1.3,by=0.2)
-
+#rlist<-c(0.5,1.1)
+
+for(iK in Klist){
   for(i in rlist){
-  params<-c(i,100) # this parameter is a vector with r and K
+    params<-c(i,iK) # this parameter is a vector with r and K
     p0<-c(params[2],params[2])
@@
   }
+}
 ```
@@
 fcode<-13
-
+Klist<-c(100,1000)
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
@@
 
+for(iK in Klist){
   for(i in rlist){
-  params<-c(i,100) # this parameter is a vector with r and K
+    params<-c(i,iK) # this parameter is a vector with r and K
     p0<-c(params[2],params[2])
@@
 fcode<-4
-
+Klist<-c(100,1000)
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
@@
 fcode<-14
-
+Klist<-c(100,1000)
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
@@
 fcode<-6
-
+Klist<-c(100,1000)
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
@@
                           params=params,p0=p0,model=model)
+  }
 }
@@
 fcode<-16
-
+Klist<-c(100,1000)
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
@@
 
-# for survival joe with kendall method only
+  # for joe with kendall method only
 

### 2019-01-22 Cause4copula.R

@@ -1,5 +1,6 @@
 library(copula)
 library(VineCopula)
+library(e1071)
 source("CopulaFunctions_flexible.R")
 source("MyBiCopGofTest.R")
 #---------------------------------------------------------------------------------------
@@ -384,12 +385,18 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
     D2umD2l_noise<-c()
     D2umD2l_pop<-c()
     
+    all_numsim_pop_patch1<-c()# to store all pop_q from each numsim for 1st patch : this should be num_keep_last times numsim points
+    all_numsim_pop_patch2<-c()# to store all pop_q from each numsim for 2nd patch : this should be num_keep_last times numsim points
     
     for(i in 1:numsim){
       #cat("i=",i,"\n")
       s<-GetNoise(N=N,fcode=fcode,corcoef=corcoef,method=method,ploton=F)
       s2<-Simulator_Cause4copula(params=params,p0=p0,noise=s$noise_q,model=model)
       z<-comp(s=s,s2=s2,num_keep_last = num_keep_last)
+      
+      all_numsim_pop_patch1<-c(all_numsim_pop_patch1,z$last_num_keep_pop$last_num_keep_pop_q[,1])
+      all_numsim_pop_patch2<-c(all_numsim_pop_patch2,z$last_num_keep_pop$last_num_keep_pop_q[,2])
+      
       S_noise<-c(S_noise,z$comp$cor_noise[1])
       K_noise<-c(K_noise,z$comp$cor_noise[2])
       P_noise<-c(P_noise,z$comp$cor_noise[3])
@@ -465,6 +472,61 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,method,lb=0,ub=0.1,num_kee
       
       D2umD2l_noise<-c(D2umD2l_noise,temp_D2u_noise-temp_D2l_noise)
       D2umD2l_pop<-c(D2umD2l_pop,temp_D2u_pop-temp_D2l_pop)
+      
+    }
+    
+    
+    if(model=="ricker"){
+      
+      tempo<-paste(resloc,"hist_pop",sep="")
+      if (!dir.exists(tempo)){
+        dir.create(tempo)
+      }
+      
+      K<-params[2]
+      
+      # Plot of histogram of pop_q (last num_keep_last number of points) from all numsim simulations
[…]

### 2019-01-24 SupportingInformation.Rmd
modified with nsd=0.1,1 for standard deviation of noise

--- 2019-01-19 SupportingInformation.Rmd
@@
 fcode<-3
+nsd<-1
 
@@
 # for clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model=model)
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model=model)
@@
 fcode<-13
+nsd<-1
 
@@
 # for survival clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model=model)
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model=model)
@@
 fcode<-4
+nsd<-1
 
@@
 resloc<-paste(tempo,"/",sep="")
-# for clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+# for gumbel with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model=model)
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model=model)
@@
 fcode<-14
+nsd<-1
 
@@
 resloc<-paste(tempo,"/",sep="")
-# for survival clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+# for survival gumbel with spearman and kendall method
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model=model)
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model=model)
@@
 fcode<-6
+nsd<-1
 
@@
 resloc<-paste(tempo,"/",sep="")
-# for clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=c(0,0),model=model)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+# for joe with kendall method only
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model=model)
@@
 fcode<-16
+nsd<-1
 
@@
 resloc<-paste(tempo,"/",sep="")
-# for survival clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=c(0,0),model=model)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+# for survival joe with kendall method only
+
+Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=c(0,0),model=model)
@@
 
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
@@
 
+for(insd in nsdlist){
+  for(iK in Klist){
    for(i in rlist){
-  params<-c(i,100) # this parameter is a vector with r and K
+    params<-c(i,iK) # this parameter is a vector with r and K
     p0<-c(params[2],params[2])
@@
 
-  tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
 
@@
   # for clayton with spearman and kendall method
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=p0,model=model)
 
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=p0,model=model)
+  }
+ }
 }
@@
 #---------------------------------------
-# This chunk stores results for ar1 model : clayton
+# This chunk stores results for ar1 model : survival clayton
 
@@
 fcode<-13
-
+nsdlist<-c(0.1,1)
+Klist<-c(100)
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
@@
 
-# for clayton with spearman and kendall method
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+  # for survival clayton with spearman and kendall method
+    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=p0,model=model)
 
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=p0,model=model)
+  }
+ }
 }
@@
 #---------------------------------------
-# This chunk stores results for ar1 model : clayton
+# This chunk stores results for ar1 model : gumbel
 
@@
 fcode<-4
-
+nsdlist<-c(0.1,1)
+Klist<-c(100)
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
@@
 
-# for clayton with spearman and kendall method
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+  # for gumbel with spearman and kendall method
+    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=p0,model=model)
 
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=p0,model=model)
+  }
+ }
 }
@@
 #---------------------------------------
-# This chunk stores results for ar1 model : clayton
+# This chunk stores results for ar1 model : survival gumbel
 
@@
 fcode<-14
-
+nsdlist<-c(0.1,1)
+Klist<-c(100)
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
@@
 
-# for clayton with spearman and kendall method
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+  # for survival gumbel with spearman and kendall method
+    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=p0,model=model)
 
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=p0,model=model)
+  }
+ }
 }
@@
 #---------------------------------------
-# This chunk stores results for ar1 model : clayton
+# This chunk stores results for ar1 model : joe
 
@@
 fcode<-6
-
+nsdlist<-c(0.1,1)
+Klist<-c(100)
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
@@
 
-# for clayton with spearman and kendall method
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+  # for joe with spearman and kendall method
+
+    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=p0,model=model)
-
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
+  }
+ }
 }
@@
 #---------------------------------------
-# This chunk stores results for ar1 model : clayton
+# This chunk stores results for ar1 model : survival joe
 
@@
 fcode<-16
-
+nsdlist<-c(0.1,1)
+Klist<-c(100)
 # for varying r (@r=1 transition from under- to over-compensatory dynamics )
@@
 
-# for clayton with spearman and kendall method
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+  # for survival joe with spearman and kendall method
+
+    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
                           params=params,p0=p0,model=model)
-
-  Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
+  }
+ }
 }

### 2019-01-25 Paper.Rmd
updated Cause4copula section write up as instructed

@@
            "6. Number of well-fitting location pairs",
-           "7. 2.5th and 97.5 quantiles of model-avg. lower-tail dependence",
-           "8. 2.5th and 97.5th quantiles of model-avg. upper-tail dependence",
+           "7. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. lower-tail dependence",
+           "8. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. upper-tail dependence",
            "9. Percent pairs showing stronger lower tail dependence",
-           "10. 2.5th and 97.5th quantiles of $<$P$_{l}$ - P${_u}$$>$ across all location pair")
+           "10. 2.5$^{th}$ and 97.5$^{th}$ quantiles of $<$P$_{l}$ - P${_u}$$>$ across all location pair")
            )
@@
 We showed with Eq.
-\ref{eq.AR1} in the next section that both model selection method and non-parametric analysis indicated similar tail dependence in population copula as that of noise copula.
+\ref{eq.AR1} in the next section that non-parametric analysis indicated similar tail dependence in population copula as that of noise copula.
 <!--Asymmetric sensitivity and copula structure in populations-->
@@
 
-To test environmental effect on population copula from the model \ref{eq.AR1}, our approach is of two-fold : First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2006_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
+To test environmental effect on population copula from the model \ref{eq.AR1} with $\beta = 0.5$, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence statistics for both bivariate noise copula and resultant population copula (following model \ref{eq.AR1}) for $50$ simulations.
+Then we plotted mean of each measured quantity with standard error against Kendall's correlation of the noise copula family.
+Fig.
+\ref{SM-fig_Cause4copula_C_kend_ar1} compares the statistics from population copula and the copula made of noises drawn from Clayton (lower tail dependent copula family).
+We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which verified the Moran effect on the population data caused by environmental noise.
+(For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
+Linear regression between tail statistics from noise and population copula (Fig.
+\ref{fig_Cause4copula_scatter_C_kend_ar1}) confirmed that the asymmetric tail dependence in noise copula caused similar tail dependence in population copula.
+We did the same analysis with other asymmetric uniparameter copula family and got similar tail dependence in resultant population copula (see Figs.
+\ref{SM-fig_Cause4copula_SC_kend_ar1} - \ref{SM-fig_Cause4copula_SJ_kend_ar1}).
+We also measured tail-dependence between noise and population copula for non-linear stochastic Ricker model and found that our hypothesis of similar tail dependence held true for weak non-linearity regime (see Appendix \ref{SM-cause4copula_ricker}).
+
+<!--
+***SG : commented out this section as it gives the similar interpretation but in a more complex way.
+our approach is of two-fold : First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2006_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
 Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
@@
 \end{figure}
-
-Next, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence statistics (see \nameref{NPA}) for both bivariate noise copula and resultant population copula (following model \ref{eq.AR1}) for $50$ simulations.
-Then we plotted mean of each measured quantity with standard error against Kendall's Tau of the noise copula family.
-Fig.
-\ref{SM-fig_Cause4copula_C_kend} compares the statistics from population copula and the copula made of noises drawn from lower tail dependent Clayton family.
-It is possible to measure the statistics with variation of Spearman's Rho for any other noise copula family (e.g.
-see Fig.
-\ref{SM-fig_Cause4copula_SC_spear}).
-We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that like our first approach (model selection method), here also the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which verified the Moran effect on the population data caused by environmental noise.
-(For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
-Linear regression between tail statistics from noise and population copula (Fig.
-\ref{fig_Cause4copula_scatter_C_kend}) confirmed that the asymmetric tail dependence in noise copula caused similar tail dependence in population copula.
+-->
 
@@
 \begin{center}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
 \caption{Linear regression between 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from lower tail dependent Clayton copula and of the affected populations for 9 equidistant Kendall's Tau values in a range (0.1 to 0.9 as shown in Fig.
-\ref{SM-fig_Cause4copula_C_kend}) showed significant pearson correlation, and henceforth confirmed asymmetric tail dependence in noise copula caused similar tail dependence in population copula.
-Regression line (black) differing from 1:1 line (green) indicates that asymmetric tail dependence in noise and population copulae are similar but not exactly the same.\label{fig_Cause4copula_scatter_C_kend}}
+\ref{SM-fig_Cause4copula_C_kend_ar1}) showed significant pearson correlation, and henceforth confirmed asymmetric tail dependence in noise copula caused similar tail dependence in population copula for AR(1) model.
+Regression line (black) differing from 1:1 line (green) indicates that asymmetric tail dependence in noise and population copulae are similar but not exactly the same.\label{fig_Cause4copula_scatter_C_kend_ar1}}
 \end{center}

### 2019-01-25 SupportingInformation.Rmd
updated Cause4copula section write up as instructed

@@
 \end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/common_legend_cause4copula_stat.pdf}\\
+--->
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/common_legend_cause4copula_stat.pdf}\\
 \vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Spearman_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Kendall_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Corl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Coru_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pu_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2u_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2l_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
-\caption[Comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different \textbf{Kendall's Tau}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant \textbf{Kendall's Tau} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
+\caption[AR(1) model : comparing correlations and non-parametric statistic for between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{AR(1) model : results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
 p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
-\label{fig_Cause4copula_C_kend}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/common_legend_cause4copula_stat.pdf}\\
+\label{fig_Cause4copula_C_kend_ar1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
 \vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Spearman_vs_Spearman's Rho.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Kendall_vs_Spearman's Rho.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Pearson_vs_Spearman's Rho.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Corl_vs_Spearman's Rho.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Clayton_Coru_vs_Spearman's Rho.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Corl-Coru_vs_Spearman's Rho.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pl_vs_Spearman's Rho.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pu_vs_Spearman's Rho.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_Pl-Pu_vs_Spearman's Rho.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_D2u_vs_Spearman's Rho.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_D2l_vs_Spearman's Rho.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results/Survival Clayton_D2u-D2l_vs_Spearman's Rho.pdf}
-\caption[Comparing correlations and non-parametric statistic between upper tail dependent noise-copula (\textbf{Survival Clayton}) and population copula for different \textbf{Spearman's Rho}]{Results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from upper tail dependent \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant \textbf{Spearman's Rho} values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[AR(1) model : comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{AR(1) model : results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from upper tail dependent \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
 p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
-\label{fig_Cause4copula_SC_spear}}
-\end{center}
-\end{figure}
--->
+Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with pearson correlation p-value $< 0.05$, then they are significantly similar.\label{fig_Cause4copula_SC_kend_ar1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[AR(1) model : comparing correlations and non-parametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from upper tail dependent \textbf{Gumbel} copula.
+\label{fig_Cause4copula_G_kend_ar1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[AR(1) model : comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from lower tail dependent \textbf{Survival Gumbel} copula.
+\label{fig_Cause4copula_SG_kend_ar1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[AR(1) model : comparing correlations and non-parametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from upper tail dependent \textbf{Joe} copula \label{fig_Cause4copula_J_kend_ar1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[AR(1) model: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from lower tail dependent \textbf{Survival Joe} copula.
+\label{fig_Cause4copula_SJ_kend_ar1}}
+\end{center}
+\end{figure}
+
+# Similar tail dependence in population copula as that of noise copula for stochastic Ricker model \label{cause4copula_ricker}
+
+To find whether nature of tail dependence in noise copula preserved for affected population copula or it would be destroyed because of underlying non-linearity in the model, we simulated the following stochastic Ricker model with intrinsic growth rate $r = 0.5$ and carrying capacity $K = 100$ for two population patches ($i = 1, 2$);
+
+\begin{equation}\label{eq.sRicker}
+P_{i}(t+1)=P_{i}(t) exp \left[ r \left\{ 1-\frac{P_{i}(t)}{K} \right\} +\epsilon_{i}(t) \right]
+\end{equation}
+
+The noise ($\epsilon_{i}$), initially drawn from a bivariate asymmetric copula family, was transformed into a normally distribution with mean zero and standard deviation $0.1$ (considered as weak non-linear regime) or $1$ (considered as strongly non-linear), for code see [`GetNoise`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function.
+For weak non-linear regime, we found similar tail dependence in noise copula and the resultant population copula from the stochastic Ricker model (Figs.
+\ref{fig_Cause4copula_C_kend_sRicker} - \ref{fig_Cause4copula_SJ_kend_sRicker}) as we obtained in linear AR(1) model (see \nameref{MT-Results_Q2} in the maintext).
+But for strongly non-linear regime the tail-dependence between noise and population copula was not significantly similar (see the summary in Fig.
+\ref{fig_Cause4copula_kend_sRicker_nsd_1}).
+We repeated the whole analysis with $r > 1$ (i.e.
+Ricker model's over-compensatory dynamical regime) and came up with same conclusion that strong non-linearility could disturb the tail-dependence information translated from noise to population copula.
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
+Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with pearson correlation p-value $< 0.05$, then they are significantly similar.
+\label{fig_Cause4copula_C_kend_sRicker}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_C_kend_sRicker} but for noise generated from upper tail dependent \textbf{Survival Clayton} copula.
+\label{fig_Cause4copula_SC_kend_sRicker}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_C_kend_sRicker} but for noise generated from upper tail dependent \textbf{Gumbel} copula.
+\label{fig_Cause4copula_G_kend_sRicker}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_C_kend_sRicker} but for noise generated from lower tail dependent \textbf{Survival Gumbel} copula.
+\label{fig_Cause4copula_SG_kend_sRicker}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_C_kend_sRicker} but for noise generated from upper tail dependent \textbf{Joe} copula.
+\label{fig_Cause4copula_J_kend_sRicker}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_C_kend_sRicker} but for noise generated from lower tail dependent \textbf{Survival Joe} copula.
+\label{fig_Cause4copula_SJ_kend_sRicker}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}\\
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}\\
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}\\
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}\\
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}\\
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with strong non-linearity: comparing tail-dependence statistic between noise-copula and population copula for different Kendall's correlation]{Stochastic Ricker model with strong non-linearity: comparing tail-dependence statistic between noise-copula and population copula for different Kendall's correlation with noise generated from \textbf{Clayton} (row 1), \textbf{Survival Clayton} (row 2), \textbf{Gumbel} (row 3),\textbf{Survival Gumbel} (row 4),\textbf{Joe} (row 5),\textbf{Survival Joe} (row 6).
+\label{fig_Cause4copula_kend_sRicker_nsd_1}}
+\end{center}
+\end{figure}
 

### 2019-01-26 Paper.Rmd
edits I forgot to commit before

--- 2019-01-10 Paper.Rmd
@@
 Model averaging used for rows 7-9 was based on AIC weights.
-For row 10, the statistic $P_{0,u_b}-P_{1-u_b}$ was computed for real data and for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented.
+For row 10, the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented.
 If ranks were $>975$ or $<25$, asymmetry of tail dependence was significant compared to a N-copula null hypothesis.
@@
 <!--Question 2 analyses-->
-# Concepts and models for Q2: possible causes of non-normal copula structure \label{Causes}
-
-This section is dedicated to exploring question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}.
-We showed environmental copula structure and asymmetric sensitivities to the environment could produce population copula structures.
-<!--We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change?--> 
-<!--Moran effects-->
-\textbf{(i)} We were interested to extend the basic idea of Moran's theorem : if environmental noises at two different habitat patches had some specific assymetric tail dependence then what will be the outcome dependence in the populations at those habitat patches.
-Should their dependence be exactly the same as of the noises or not ?
+# Background, concepts and methods for Q2: causes/mechanisms of non-normal copula structure \label{Causes}
+
+Having demonstrated that non-N copula structure and asymmetric tail dependence are common
+features of ecological data, we explore via mathematical 
+models two possible mechanisms of this structure which 
+seem likely to operate in population systems. The first mechanism 
+is an extension of the well known Moran
+effect, and was summarized in the Introduction. 
+It explores the idea that non-N copula structure and asymmetric tail dependence 
+between environmental variables measured in two places may bring about similarly 
+structured dependence between ecological variables that are influenced by the 
+environmental variables. Let $P_i(t)$ be a population index for location $i$ ($i=1,2$),
+and assume the index follows AR(1) dynamics within a patch, 
+$P_i(t+1)=\beta P_{i}(t-1)+ \sqrt{1-\beta^2} \epsilon_i(t)$ where the $\epsilon_i(t)$ are 
+standard-normal random variables that are independent for distinct times, $t$, but correlated
+
 To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
@@
 <!--Asymmetric sensitivity and copula structure in populations-->
-\textbf{(ii)} Secondly, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) which behaved differently beyond a certain expectation threshold (as shown in Eq.
+\textbf{(ii)}
+
+The second mechanism, which was also summarized in the introduction,
+relates to Liebig's law of the minimum, or, more generally, to the potential for 
+environmental variables to influence ecological variables through nonlinear relationships.
+
+Secondly, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) which behaved differently beyond a certain expectation threshold (as shown in Eq.
 \ref{eq.AS_g}).

### 2019-01-26 Paper.Rmd

--- 2019-01-25 Paper.Rmd
@@
 We showed with Eq.
-\ref{eq.AR1} in the next section that non-parametric analysis indicated similar tail dependence in population copula as that of noise copula.
+\ref{eq.AR1} in the next section that both model selection method and non-parametric analysis indicated similar tail dependence in population copula as that of noise copula.
 <!--Asymmetric sensitivity and copula structure in populations-->
-\textbf{(ii)} Secondly, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) which behaved differently beyond a certain expectation threshold (as shown in Eq.
+\textbf{(ii)}
+
+The second mechanism, which was also summarized in the introduction,
+relates to Liebig's law of the minimum, or, more generally, to the potential for 
+environmental variables to influence ecological variables through nonlinear relationships.
+
+Secondly, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) which behaved differently beyond a certain expectation threshold (as shown in Eq.
 \ref{eq.AS_g}).

### 2019-01-26 ext_risk_copula.Rmd

@@
     lines(s$d_seq,s$risk_right,type="b",col="blue")
-    legend("topleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
+    legend("bottomleft", c("noise : left","noise : right"), lty=c(1,1), pch=c(1,1), col=c('red', 'blue'), horiz=T, bty='n', cex=1.2)
     mtext(paste0("r = ", r))

### 2019-01-28 SupportingInformation.Rmd
figure fixed

@@
 
-The noise ($\epsilon_{i}$), initially drawn from a bivariate asymmetric copula family, was transformed into a normally distribution with mean zero and standard deviation $0.1$ (considered as weak non-linear regime) or $1$ (considered as strongly non-linear), for code see [`GetNoise`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function.
-For weak non-linear regime, we found similar tail dependence in noise copula and the resultant population copula from the stochastic Ricker model (Figs.
-\ref{fig_Cause4copula_C_kend_sRicker} - \ref{fig_Cause4copula_SJ_kend_sRicker}) as we obtained in linear AR(1) model (see \nameref{MT-Results_Q2} in the maintext).
-But for strongly non-linear regime the tail-dependence between noise and population copula was not significantly similar (see the summary in Fig.
-\ref{fig_Cause4copula_kend_sRicker_nsd_1}).
+The noise ($\epsilon_{i}$), initially drawn from a bivariate asymmetric copula family, was transformed into a normally distribution with mean zero and standard deviation $0.1$ (considered as weak noise) or $1$ (considered as strong noise), for code see [`GetNoise`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function.
+For weak noise, we found similar tail dependence in noise copula and the resultant population copula from the stochastic Ricker model (Figs.
+\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} - \ref{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}) as we obtained in linear AR(1) model (see \nameref{MT-Results_Q2} in the maintext).
+But for strong noise the tail-dependence between noise and population copula was not significantly similar (Figs.
+\ref{fig_Cause4copula_C_kend_sRicker_nsd_1} - \ref{fig_Cause4copula_SJ_kend_sRicker_nsd_1}).
 We repeated the whole analysis with $r > 1$ (i.e.
-Ricker model's over-compensatory dynamical regime) and came up with same conclusion that strong non-linearility could disturb the tail-dependence information translated from noise to population copula.
-
+Ricker model's over-compensatory dynamical regime) and came up with same conclusion that strong noise with non-linearility could disturb the tail-dependence information translated from noise to population copula.
+
+<!--adding results for stochastic ricker with weak noise : nsd=0.1-->
 \begin{figure}[!h]
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak nonise: results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula (having normal marginal with standard deviation 0.1) and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
 p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
 Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with pearson correlation p-value $< 0.05$, then they are significantly similar.
-\label{fig_Cause4copula_C_kend_sRicker}}
+\label{fig_Cause4copula_C_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_C_kend_sRicker} but for noise generated from upper tail dependent \textbf{Survival Clayton} copula.
-\label{fig_Cause4copula_SC_kend_sRicker}}
+\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Survival Clayton} copula having normal marginal with standard deviation 0.1.
+\label{fig_Cause4copula_SC_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_C_kend_sRicker} but for noise generated from upper tail dependent \textbf{Gumbel} copula.
-\label{fig_Cause4copula_G_kend_sRicker}}
+\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Gumbel} copula having normal marginal with standard deviation 0.1.
+\label{fig_Cause4copula_G_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_C_kend_sRicker} but for noise generated from lower tail dependent \textbf{Survival Gumbel} copula.
-\label{fig_Cause4copula_SG_kend_sRicker}}
+\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Gumbel} copula having normal marginal with standard deviation 0.1.
+\label{fig_Cause4copula_SG_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_C_kend_sRicker} but for noise generated from upper tail dependent \textbf{Joe} copula.
-\label{fig_Cause4copula_J_kend_sRicker}}
+\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Joe} copula having normal marginal with standard deviation 0.1.
+\label{fig_Cause4copula_J_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak non-linearity: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak non-linearity: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_C_kend_sRicker} but for noise generated from lower tail dependent \textbf{Survival Joe} copula.
-\label{fig_Cause4copula_SJ_kend_sRicker}}
+\caption[Stochastic Ricker model with weak nonise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Joe} copula having normal marginal with standard deviation 0.1.
+\label{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}}
+\end{center}
+\end{figure}
+
+<!--adding results for stochastic ricker with strong noise : nsd=1-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with strong noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Clayton} copula  having normal marginal with standard deviation 1.
+\label{fig_Cause4copula_C_kend_sRicker_nsd_1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with strong noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_SC_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Survival Clayton} copula  having normal marginal with standard deviation 1.
+\label{fig_Cause4copula_SC_kend_sRicker_nsd_1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with strong noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_G_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Gumbel} copula  having normal marginal with standard deviation 1.
+\label{fig_Cause4copula_G_kend_sRicker_nsd_1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with strong noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_SG_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Gumbel} copula having normal marginal with standard deviation 1.
+\label{fig_Cause4copula_SG_kend_sRicker_nsd_1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with strong noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_J_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Joe} copula having normal marginal with standard deviation 1.
+\label{fig_Cause4copula_J_kend_sRicker_nsd_1}}
 \end{center}
@@
 \vspace{-0.4 cm}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}\\
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}\\
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}\\
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}\\
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}\\
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=4.5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with strong non-linearity: comparing tail-dependence statistic between noise-copula and population copula for different Kendall's correlation]{Stochastic Ricker model with strong non-linearity: comparing tail-dependence statistic between noise-copula and population copula for different Kendall's correlation with noise generated from \textbf{Clayton} (row 1), \textbf{Survival Clayton} (row 2), \textbf{Gumbel} (row 3),\textbf{Survival Gumbel} (row 4),\textbf{Joe} (row 5),\textbf{Survival Joe} (row 6).
-\label{fig_Cause4copula_kend_sRicker_nsd_1}}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Stochastic Ricker model with weak nonise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\ref{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Joe} copula having normal marginal with standard deviation 1.
+\label{fig_Cause4copula_SJ_kend_sRicker_nsd_1}}
 \end{center}

### 2019-01-31 Paper.Rmd
minor edit
minor edits

@@
 We introduce a new rank based "copula" approach to model dependencies between ecological variables.
-Our method could overcome the shortcomings of the linear correlation approach and give us new insight about "asymmetric tail dependence" found in ecology, biogeochemistry and related field.
-We focus on three objectives (1) Do datasets in ecology and related fields have non-Normal copula srtructure?
+Our method could overcome the shortcomings of the linear correlation approach and give us new insight about "asymmetric tail dependence" found in ecology, bio-geochemistry and related field.
+We focus on three objectives (1) Do data sets in ecology and related fields have non-Normal copula structure?
 Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric
@@
 We demonstrate with simulations and real data that non-Normal copulae have profound effects on spatial skewness of species abundances and spatial Taylor's law slope values.
-Our method is new and could have potential when exploring some other related phenomena like effects of climate change on extinction risks of metapopulations.
+Our method is new and could have potential when exploring some other related phenomena like effects of climate change on extinction risks of meta-populations.
 
@@
 approaches provide only a very partial description of a relationship between variables. 
-For instance, datasets showing markedly different 
+For instance, data sets showing markedly different 
 relationships may nevertheless have the same correlation coefficient value 
@@
 \includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
-\caption{(A-F, I, J) Bivariate datasets showing diverse relationships between 
-variables; many of these datasets nevertheless have the same Pearson (P)
+\caption{(A-F, I, J) Bivariate data sets showing diverse relationships between 
+variables; many of these data sets nevertheless have the same Pearson (P)
 or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
@@
 marginal distributions of the underlying random variables, isolating the 
-depedence structure only.
+dependence structure only.
 Mathematically, if $(X,Y)$ is a continuous bivariate random
-variable and the cummulative distribution functions (cdfs) of $X$ and $Y$ are 
+variable and the cumulative distribution functions (cdfs) of $X$ and $Y$ are 
 $F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the dependence or copula structure
@@
 and \emph{vice versa} if larger values are more strongly associated than smaller
-ones (Fig. \ref{fig_cop_pedag1}B, F). Datasets of the same 
-corelation can have a range of tail dependence 
+ones (Fig. \ref{fig_cop_pedag1}B, F). Data sets of the same 
+correlation can have a range of tail dependence 
 characteristics (Fig. \ref{fig_cop_pedag}), which can be quantified (see section on
@@
 
-For large datasets (large $n$),
+For large data sets (large $n$),
 we could use $l_b$ and $u_b$ close together without incurring undue
@@
 several different bands $(l_b,u_b)$ to understand how dependence
-varies in different parts of the distributions. But for datasets with 
+varies in different parts of the distributions. But for data sets with 
 smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
@@
 The difference $\Dsq_u-\Dsq_l$ (note the opposite order), is also expected to be 
-positive under the same conditions. These statistics were used for smaller datasets.
-For large datasets, we used $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
+positive under the same conditions. These statistics were used for smaller data sets.
+For large data sets, we used $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
 $\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
@@
 
-For bivariate datasets, we compared values 
+For bivariate data sets, we compared values 
 of the statistics $\cor_{l_b,u_b}$,
@@
 $\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ to 
-distributions of values of the same statistics computed on surrogate datasets that were produced from 
+distributions of values of the same statistics computed on surrogate data sets that were produced from 
 the empirical data by randomizing it in a special way to have no 
 tail dependence (Appendix \ref{SM-surrog_test}). 
-The surrogate/randomized datasets had exactly 
+The surrogate/randomized data sets had exactly 
 the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
@@
 empirical data, to the distribution
-of its values computed on many surrogate datasets 
+of its values computed on many surrogate data sets 
 provides a test of whether dependence in the part of the distribution specified 
@@
 correlation. Thus the comparison helps answer the first two parts of Q1
-for bivariate datasets. 
+for bivariate data sets. 
 Significant deviations correspond to deviations from normal 
@@
 $\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$,
-as computed on the empirical data, to the distribution of its values on many surrogate datasets 
+as computed on the empirical data, to the distribution of its values on many surrogate data sets 
 provides a test of asymmetry of dependence. Thus this comparison helps answer the
 third part of Q1 for bivariate 
-datasets. For multivariate datasets, we used a spatial resampling scheme 
+data sets. For multivariate data sets, we used a spatial resampling scheme 
 (Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals, instead of the surrogate 
@@
 Across essentially all the sub-fields and levels of organization within ecology for which we 
-examined a dataset, copula structures were typically non-normal and variables 
+examined a data set, copula structures were typically non-normal and variables 
 commonly exhibited asymmetric tail 
 dependence. To make results easier to absorb for readers who are new to copulas: we first 
-present results in detail for an example bivariate dataset; then we summarize results
-for all bivariate datasets; then we present results in 
-detail for an example multivariate dataset; then we summarize results 
-for all multivariate datasets.
+present results in detail for an example bivariate data set; then we summarize results
+for all bivariate data sets; then we present results in 
+detail for an example multivariate data set; then we summarize results 
+for all multivariate data sets.
 
@@
 Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
-available from `BiCopGofTest` in the `VineCopula` package). However, nonparametric statistics verified
+available from `BiCopGofTest` in the `VineCopula` package). However, non-parametric statistics verified
 both that tail dependence was distinct from that of an N copula for the soil data, and that it 
@@
 Table \ref{tab_bivar_summary} summarizes bivariate results for Q1. For all $4$ 
-bivariate datasets, the two variables were always significantly related 
+bivariate data sets, the two variables were always significantly related 
 (Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
@@
 (Table \ref{tab_bivar_summary}, row 9),
-answering the third part of Q1. The first three datasets had stronger upper- 
-than lower-tail dependence, and the last dataset had the reverse. 
+answering the third part of Q1. The first three data sets had stronger upper- 
+than lower-tail dependence, and the last data set had the reverse. 
 Although these results convincingly show non-normal copula structure, the 
@@
 (Table \ref{tab_bivar_summary}, rows 5-6).
-But our nonparametric statistics confirmed the same conclusions 
+But our non-parametric statistics confirmed the same conclusions 
 (Table \ref{tab_bivar_summary}, row 10).
-Detailed results for each dataset are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} 
+Detailed results for each data set are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} 
 and Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.
@@
 AIC weights for the best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
-were typically substiantially 
+were typically substantially 
 larger than those for the N copula (Table \ref{SM-tab_count_AICw_normal}), answering 
@@
 <!--Non-parametric analysis of tail dependence, green spruce aphid abundance data-->
-Nonparametric statistics verified that tail dependence was asymmetric for the green 
+Non-parametric statistics verified that tail dependence was asymmetric for the green 
 spruce aphid abundance data.
@@
 according to Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests 
-(Table \ref{tab_multivar_summary}, row 6). Some of the datasets (green spruce aphid
+(Table \ref{tab_multivar_summary}, row 6). Some of the data sets (green spruce aphid
 abundance, \textit{Ceratium furca} abundance, methane flux) showed stronger 
@@
 (Table \ref{tab_multivar_summary}, rows 7-9), whereas leaf-curling plum aphid first flight
-data showed the reverse. Asymmetry results were verified by nonparametric approaches 
+data showed the reverse. Asymmetry results were verified by non-parametric approaches 
 (e.g., Table \ref{tab_multivar_summary}, row 10). Detailed results are in 
@@
 For row 10, the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented.
-If ranks were $>975$ or $<25$, asymmetry of tail dependence was significant compared to a N-copula null hypothesis.
-The value of $u_b$ used depended on the size of the dataset, and can be found in Figs.
-\\ref{SM-fig_soilCN_nonparam} to \\ref{SM-fig_cc2000_nonparam} and Tables \\ref{SM-tab_soilCN_asym}, \\ref{SM-tab_bmr_birds_asym}, \\ref{SM-tab_bmr_mammals_asym}, and \\ref{SM-tab_cc_asym_2000}.
-\\label{tab_bivar_summary}",
+If ranks were $>975$ or $<25$, asymmetry of tail dependence was significant compared to a N-copula null hypothesis.",
              booktabs=T, linesep = "\\addlinespace",
@@
 \ref{eq.AS_g}).
-To test the assymetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;
+To test the asymmetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;
 
@@
 \ref{eq.AS_g}, Fig.
-\ref{fig_ms_asym_b_0.1}) where $\varepsilon=(\varepsilon_{1}, \varepsilon_{2})$ was an iid through time with mean zero and covariance matrix $\Sigma_{\varepsilon}$.
+\ref{fig_ms_asym_b_0.1}) where $\varepsilon=(\varepsilon_{1}, \varepsilon_{2})$ was an iid through time with mean zero and co-variance matrix $\Sigma_{\varepsilon}$.
 Local environmental fluctuation function $h(\delta)$ varied according to Eq.
 \ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$.
-$\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and covariance matrix $\Sigma_{\delta}$.
+$\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and co-variance matrix $\Sigma_{\delta}$.
 With this model we showed tail dependence in population copula changed from lower to upper tail dependence as global noise copula $g(\varepsilon)$ changed its form from $g_{1}$ to $g_{2}$ (Eq.
@@
 \ref{fig_Cause4copula_scatter_C_kend_ar1}) confirmed that the asymmetric tail dependence in noise copula caused similar tail dependence in population copula.
-We did the same analysis with other asymmetric uniparameter copula family and got similar tail dependence in resultant population copula (see Figs.
+We did the same analysis with other asymmetric uni-parameter copula family and got similar tail dependence in resultant population copula (see Figs.
 \ref{SM-fig_Cause4copula_SC_kend_ar1} - \ref{SM-fig_Cause4copula_SJ_kend_ar1}).
@@
 \caption{Linear regression between 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from lower tail dependent Clayton copula and of the affected populations for 9 equidistant Kendall's Tau values in a range (0.1 to 0.9 as shown in Fig.
-\ref{SM-fig_Cause4copula_C_kend_ar1}) showed significant pearson correlation, and henceforth confirmed asymmetric tail dependence in noise copula caused similar tail dependence in population copula for AR(1) model.
+\ref{SM-fig_Cause4copula_C_kend_ar1}) showed significant Pearson correlation, and henceforth confirmed asymmetric tail dependence in noise copula caused similar tail dependence in population copula for AR(1) model.
 Regression line (black) differing from 1:1 line (green) indicates that asymmetric tail dependence in noise and population copulae are similar but not exactly the same.\label{fig_Cause4copula_scatter_C_kend_ar1}}
@@
 \ref{fig_ms_asym_b_0.1} and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_2}.
-According to our model selection approach (Table \ref{SM-tab_summary_stat_asym_sens}), non-normal copulae (with higher AIC weights) were always best-fit with similar asymetric tail dependence which gave the support of our hypothesis of the effect of asymmetric global noise copula on populations.
+According to our model selection approach (Table \ref{SM-tab_summary_stat_asym_sens}), non-normal copulae (with higher AIC weights) were always best-fit with similar asymmetric tail dependence which gave the support of our hypothesis of the effect of asymmetric global noise copula on populations.
 Our non-parametric analysis results (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_2}) were consistent with the model selection approach.
@@
 We tested null hypothesis $(H_{0})$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
-We generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (spearman or kendall) same and calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set in the next section (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R)).
+We generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (Spearman or Kendall) same and calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set in the next section (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R)).
 We made a histogram of surrogate skewness and found if actual skewness was within the $95\%$ confidence interval or not.
@@
 \ref{SM-fig_spTL}(I)).
-In the following section we will show application of these ideas with real datasets.
+In the following section we will show application of these ideas with real data sets.
 <!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion-----------
@@
 \includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
-\caption{Skewness of spatially averaged $(A)$ green-spruce aphid abundance, $(B)$ leaf-curling plum aphid firstflight date, $(C)$ Ceratium furca abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation).
+\caption{Skewness of spatially averaged $(A)$ green-spruce aphid abundance, $(B)$ leaf-curling plum aphid first flight date, $(C)$ Ceratium furca abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation).
 \label{fig_skewness_spearman}}
@@
 
-We used our four multivariate datasets : aphid-count data, aphid-firstflight data, plankton data and methane-flux data, to compute skewness of spatially averaged data.
-We carried out the analysis mentioned in the concepts above, with aphid abundance and aphid firstflight date data for all 20 aphid species and plankton abundance data for all 22 plankton species.
+We used our four multivariate data sets : aphid-count data, aphid-first flight data, plankton data and methane-flux data, to compute skewness of spatially averaged data.
+We carried out the analysis mentioned in the concepts above, with aphid abundance and aphid first flight date data for all 20 aphid species and plankton abundance data for all 22 plankton species.
 In each study, we got $>5\%$ positive results which gave us confidence in validating significantly different skewness for asymmetric copula than Normal.
-For example, Figure \ref{fig_skewness_spearman} represents the results for each dataset with Spearman correlation preserving surrogates.
+For example, Figure \ref{fig_skewness_spearman} represents the results for each data set with Spearman correlation preserving surrogates.
 Figure \ref{fig_skewness_spearman}(A) showed that averaged skewness (indicated by red dot) was much lower than expected if the green-spruce aphid count data came from a normal copula.
@@
 Finally, for methane-flux data, the averaged skewness was found significantly lower in comparison to its Normal surrogates (Figure \ref{fig_skewness_spearman} D).
-We found the same result with all datasets when we repeated our simulation with Kendall correlation preserving surrogates (see Fig.
+We found the same result with all data sets when we repeated our simulation with Kendall correlation preserving surrogates (see Fig.
 \ref{SM-fig_skewness_kendall}).
@@
 Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results.
-So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different non-parametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selction results.
-Our approach could help to interpret the synchronous behavior of a species from several metapopulations.
+So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different non-parametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selection results.
+Our approach could help to interpret the synchronous behavior of a species from several meta-populations.
 To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g.
-abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that metapopulations are simultaneously rare at those locations.
+abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that meta-populations are simultaneously rare at those locations.
 Conversely, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of plankton could lead to algal-bloom in the ocean.
 Copulae could also give us information about the causal relationship between ecological variables.
-For example, if maximum railfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongerly correlated than lower values.
+For example, if maximum rainfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongly correlated than lower values.
 
@@
 \ref{fig_ms_asym_b_0.1}).
-We observed significant changes in skewness of spatially averaged timeseries and also in the curvature of spatial Taylor's law for copula having asymmetric tail-dependence than would have been expected for Normal copula.
+We observed significant changes in skewness of spatially averaged time series and also in the curvature of spatial Taylor's law for copula having asymmetric tail-dependence than would have been expected for Normal copula.
 For example, Fig.
-\ref{fig_skewness_spearman}(A) showed green spruce aphid abundance timeseries had right skewed distribution but its skewness was signifiantly lower than was expected if it came from Spearman correlation preserving Normal surrogates.
+\ref{fig_skewness_spearman}(A) showed green spruce aphid abundance time series had right skewed distribution but its skewness was significantly lower than was expected if it came from Spearman correlation preserving Normal surrogates.
 This result indicates less extreme outbreaks of green spruce aphid around UK than it would have been with Normal copula structure.
-Similarly, methane fluxes showing lower skewness than Normal surrogates mean that the high emmisions are un-correlated between locations and monitoring this skewness over several decades may be helpful to understand the "green-house effect" on environment.
+Similarly, methane fluxes showing lower skewness than Normal surrogates mean that the high emissions are un-correlated between locations and monitoring this skewness over several decades may be helpful to understand the "green-house effect" on environment.
 
 <!--future work-->
-We think this copula approach have scope to investigate into the extinction risks of metapopulation due to climate change.
-To illustrate, lower-tail dependent environmental fluctuations may suppress the metapopulations synchronously by making some of the habitats go extinct.
+We think this copula approach have scope to investigate into the extinction risks of meta-population due to climate change.
+To illustrate, lower-tail dependent environmental fluctuations may suppress the meta-populations synchronously by making some of the habitats go extinct.
 This could help in managing policy for conservation of endangered populations.
@@
 # References
+\setlength{\parindent}{-0.2in}
+\setlength{\leftskip}{0.2in}
+\setlength{\parskip}{1pt}
+\noindent

### 2019-01-31 SupportingInformation.Rmd
minor edits

@@
 But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available).
-We used data from 14 locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}).
+We used data from 14 locations which had at least 45 years time series data (see Figure \ref{SM-fig_plankton_map}).
 
@@
 **See and Smyth et al, pending for
-additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.
+additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had at least 50 dates data.
 
@@
 \includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
-\caption[For definitions of nonparametric statistics]{Schematic diagram supporting definitions of nonparametric tail dependence statistics.
+\caption[For definitions of non-parametric statistics]{Schematic diagram supporting definitions of non-parametric tail dependence statistics.
 The red and blue lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
@@
 asymptotic behavior of the 
-statistics and their behavior on the small datasets which are common in some fields
+statistics and their behavior on the small data sets which are common in some fields
 of ecology. The mean and standard error of the three difference statistics were computed for 
@@
 copulas.
-We used a $t$-test with a significance threshold of $0.05$ to judge signficance here.
-These expectations were fullfilled (Figs \ref{fig_stat_testing35} and 
+We used a $t$-test with a significance threshold of $0.05$ to judge significance here.
+These expectations were fulfilled (Figs \ref{fig_stat_testing35} and 
 \ref{fig_stat_testing1000}).
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
-\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas.
+\caption[Testing non-parametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas.
 See Appendix  \ref{Test_npa_stat} for details.
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
-\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas.
+\caption[Testing non-parametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas.
 See Appendix  \ref{Test_npa_stat} for details.
@@
 
-For bivariate datasets, surrogates were produced as follows.
+For bivariate data sets, surrogates were produced as follows.
 
@@
 another permutation of $1,\ldots,n$).
-- The surrogate dataset is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.
+- The surrogate data set is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.
 
@@
 Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
-Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
+Kendall correlation of the surrogate data set will be the same as that of the $a_i$ and $b_i$, and therefore 
 will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for soil C and N data.
+\caption[Non-parametric tests against an N-copula null, soil C and N data]{Non parametric tests for tail dependence and other deviations from N-copula structure, for soil C and N data.
 As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
@@
 \includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for bird body mass and BMR data.
+\caption[Non-parametric tests against an N-copula null, bird body mass and BMR data]{Non parametric tests for tail dependence and other deviations from N-copula structure, for bird body mass and BMR data.
 Same format as Fig.
@@
 \includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for mammal body mass and BMR data.
+\caption[Non-parametric tests against an N-copula null, mammal body mass and BMR data]{Non parametric tests for tail dependence and other deviations from N-copula structure, for mammal body mass and BMR data.
 Same format as Fig.
@@
 \includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data.
+\caption[Non-parametric tests against an N-copula null, Cedar Creek data]{Non parametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data.
 Same format as Fig.
@@
 \includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf}
-\caption[Nonparametric tail statistics versus distance between locations for green spruce aphid abundance data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for green spruce aphid abundance data.
+\caption[Non-parametric tail statistics versus distance between locations for green spruce aphid abundance data]{Non parametric tail statistics versus great-circle distance, $D$, between sampling locations for green spruce aphid abundance data.
 \label{fig_aphid_count}}
@@
 \includegraphics[width=4 cm]{./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf}
-\caption[Nonparametric tail statistics versus distance between locations for leaf-curling plum aphid first flight data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for leaf-curling plum aphid first flight data.
+\caption[Non-parametric tail statistics versus distance between locations for leaf-curling plum aphid first flight data]{Non parametric tail statistics versus great-circle distance, $D$, between sampling locations for leaf-curling plum aphid first flight data.
 \label{fig_aphid_ff}}
@@
 \includegraphics[width=4 cm]{./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf}
-\caption[Nonparametric tail statistics versus distance between locations for \textit{Ceratium furca} abundance data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for \textit{Ceratium furca} abundance data.
+\caption[Non-parametric tail statistics versus distance between locations for \textit{Ceratium furca} abundance data]{Non-parametric tail statistics versus great-circle distance, $D$, between sampling locations for \textit{Ceratium furca} abundance data.
 \label{fig_pns}}
@@
 \includegraphics[width=4 cm]{./Results/stat_results/stat_methane/data_pt_thrs50_D2u-D2l_vs_D.pdf}
-\caption[Nonparametric tail statistics versus distance between locations for methane-flux data]{Nonparametric tail statistics versus great-circle distance, $D$, between sampling locations for methane-flux data.
+\caption[Non-parametric tail statistics versus distance between locations for methane-flux data]{Non-parametric tail statistics versus great-circle distance, $D$, between sampling locations for methane-flux data.
 \label{fig_methane}}
@@
 p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
-Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with pearson correlation p-value $< 0.05$, then they are significantly similar.\label{fig_Cause4copula_SC_kend_ar1}}
+Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with Pearson correlation p-value $< 0.05$, then they are significantly similar.\label{fig_Cause4copula_SC_kend_ar1}}
 \end{center}
@@
 For weak noise, we found similar tail dependence in noise copula and the resultant population copula from the stochastic Ricker model (Figs.
-\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} - \ref{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}) as we obtained in linear AR(1) model (see \nameref{MT-Results_Q2} in the maintext).
+\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} - \ref{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}) as we obtained in linear AR(1) model (see \nameref{MT-Results_Q2} in the main text).
 But for strong noise the tail-dependence between noise and population copula was not significantly similar (Figs.
@@
 We repeated the whole analysis with $r > 1$ (i.e.
-Ricker model's over-compensatory dynamical regime) and came up with same conclusion that strong noise with non-linearility could disturb the tail-dependence information translated from noise to population copula.
+Ricker model's over-compensatory dynamical regime) and came up with same conclusion that strong noise with non-linearity could disturb the tail-dependence information translated from noise to population copula.
 
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak nonise: results of comparison between 3 types of correlation (i.e.
+\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: results of comparison between 3 types of correlation (i.e.
 Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula (having normal marginal with standard deviation 0.1) and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
 p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
-Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with pearson correlation p-value $< 0.05$, then they are significantly similar.
+Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with Pearson correlation p-value $< 0.05$, then they are significantly similar.
 \label{fig_Cause4copula_C_kend_sRicker_nsd_0.1}}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak nonise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Joe} copula having normal marginal with standard deviation 0.1.
@@
 Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula.
-First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with covariance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of timeseries, number of sampling locations and number of replica for that copula, respectively.
+First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with co-variance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of time series, number of sampling locations and number of replica for that copula, respectively.
 Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal.
 With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D).
-We also tested statistically that whether linearity, homoskedasticity is maintained in each copula dataset or not and plotted our results alongwith root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I).
+We also tested statistically that whether linearity, homoskedasticity is maintained in each copula data set or not and plotted our results along with root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I).
 We did a quadratic regression with each simulated copula for all types and measured quadratic coefficient and mean curvature for the regression line which is plotted in Fig.\ref{fig_spTL}(J-K) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).

### 2019-02-04 Paper.Rmd
various changes

--- 2019-01-26 Paper.Rmd
@@
 from environmental to ecological variables, then we may expect important 
-copula structure to be a common feature of spatial synchrony.
+copula structure to be a common feature of spatial synchrony of population, community, 
+biogeochemical and other ecological variables.
 Synchrony is a phenomenon of great interest in 
@@
 
-We used datasets that pertain to multiple fields and levels of organization within or related to ecology, including environmental, species-trait, phenological, population, community, and ecosystem functioning data.
-Copula structure describing the relationship between atmospheric weather variables such as X and Y has been examined previously in the meteorological literature (e.g., REF).
-We therefore examined environmental variables from the soil instead, using the Rapid Carbon Assessment database (RaCA; @Wills2014).
+We used datasets that pertain to multiple fields and levels of organization within 
+or related to ecology, including environmental, species-trait, phenological, 
+population, community, and ecosystem functioning data. Copula structure describing 
+the relationship between atmospheric weather variables such as X and Y has been 
+examined previously in the meteorological literature (e.g., REF). We therefore 
+examined environmental variables from the soil instead, using the Rapid Carbon 
+Assessment database (RaCA; @Wills2014).
 <!--***DAN: Add Wills and Loecke pending reference if published in time--> 
-The database comprises measurements of soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
+The database comprises measurements of soil organic carbon and total soil 
+nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
 locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
@@
 <!--***DAN: Terry to check the wording and accuracy here-->
-Species-trait data were average species basal metabolic rate (KJ per hour) and body mass (gram) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
+Species-trait data were average species basal metabolic rate (KJ per hour) and 
+body mass (gram) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
+and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
 These data have been much studied, but to our knowledge the copula structure 
@@
 models two possible mechanisms of this structure which 
-seem likely to operate in population systems. The first mechanism 
+seem likely to operate in diverse ecological scenarios. The first mechanism 
 is an extension of the well known Moran
-effect, and was summarized in the Introduction. 
+effect, and was summarized conceptually in the Introduction. 
 It explores the idea that non-N copula structure and asymmetric tail dependence 
@@
 and assume the index follows AR(1) dynamics within a patch, 
-$P_i(t+1)=\beta P_{i}(t-1)+ \sqrt{1-\beta^2} \epsilon_i(t)$ where the $\epsilon_i(t)$ are 
-standard-normal random variables that are independent for distinct times, $t$, but correlated
+$P_i(t+1)=\beta P_{i}(t-1)+ \sqrt{1-\beta^2} \epsilon_i(t)$. Here the $\epsilon_i(t)$ are 
+standard-normal random variables that are independent for distinct times, $t$, but 
+that are non-independent across patches.
+
+<!--***DAN: Note to self, this model, and also the other one we study in this section,
+are actually much more general than populatin. Should replace P_i with E_i where E_i is
+just some ecological variable. Could be deviaitons from a pop from carrying capacity, could be a 
+biogeochemical variable, could maybe even be an evolutinary variable. Anything that is
+affected by an environmental variable and could reasonably be modelled by an AR process. 
+So the generalization to biogeochemical and community
+data seems pretty straightforward. Need to think this through and reword the corresponding
+part of the intro to encompass this greater generality. How would the evolutionary case work?
+Well, suppose the two "patches" are actually traits, and the perturbations epsilon are somehow
+the "drift events" or something like that? This does not really make sense, strictly speaking,
+now that I am thinking about it, but maybe the model could be modified to make it make sense?
+At least I have straightforward generalization to biogeochem and community cases. 
+More on evolutionary: suppose we model traits as a bivariate random walk where the perturbations
+in the traits are related with a copula with some tail dependence. Suppose you have a tree and 
+you model evolution down the tree in this way and then look at the bivariate distribution of
+trait on the tips of the tree. What do you get? I guess you get some copula structure?-->
 

### 2019-02-04 Paper.Rmd
changed the description of what was done with the first set of models in the Q2 section

@@
 
-<!--***DAN: made it to here, 2018 1 10, 1252pm-->
-
 <!--Question 2 analyses-->
@@
 Having demonstrated that non-N copula structure and asymmetric tail dependence are common
-features of ecological data, we explore via mathematical 
+features of ecological data, we explored via mathematical 
 models two possible mechanisms of this structure which 
@@
 effect, and was summarized conceptually in the Introduction. 
-It explores the idea that non-N copula structure and asymmetric tail dependence 
-between environmental variables measured in two places may bring about similarly 
+It explores the hypothesis that non-N copula structure and asymmetric tail dependence 
+between environmental variables measured through time 
+in two places may bring about similarly 
 structured dependence between ecological variables that are influenced by the 
-environmental variables. Let $P_i(t)$ be a population index for location $i$ ($i=1,2$),
-and assume the index follows AR(1) dynamics within a patch, 
-$P_i(t+1)=\beta P_{i}(t-1)+ \sqrt{1-\beta^2} \epsilon_i(t)$. Here the $\epsilon_i(t)$ are 
+environmental variables.
+
+Let $E_i(t)$ be an ecological variable 
+measured in location $i$ ($i=1,2$).
+We initially assumed the variable follows AR(1) dynamics within a location, 
+$E_i(t+1)=\beta E_{i}(t-1)+ \sqrt{1-\beta^2} \epsilon_i(t)$. 
+We used $\beta=0.5$. Here the $\epsilon_i(t)$ represent environmental noise; they are 
 standard-normal random variables that are independent for distinct times, $t$, but 
-that are non-independent across patches.
+that were made to exhibit different kinds of 
+dependence across locations in different simulations (see below).
+The ecological variable, $E$, is very general, and could represent, for instance, deviations
+of a population density from a carrying capacity or average density, deviations of
+total plant community biomass or other community variable from an average value,
+or flux of a biogeochemical variable.
+
+To explore the importance of nonlinear dynamics for our hypothersis, we also considered 
+nonlinear population models based on $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right) \right)+\sigma\epsilon_i(t))$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
+$\sigma=1$. When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
+$K$, and the nonlinearities of the model have limited influence on dynamics. When
+$\sigma=1$, model dynamics are strongly nonlinear. We refer to these as
+the weak-noise and strong-noise cases. The $epsilon_i(t)$ were again standard
+normal and independent through time, but were related (see below) across locations.
+
+For each of the three model setups, for each of the copulas C, SC, G, SG, J, SJ,
+and for each Kendall correlation value $\tau=0.1,0.2,\ldots,0.9$, 
+we generated 5000 noise pairs $(\epsilon_1(t),\epsilon_2(t))$ from the
+bivariate random variable with the given copula family and the given $\tau$.
+We then used the generated noise to drive the model, and retained both the
+noise and population values for the final 500 
+time steps of the simulation. For each model, copula family, and $\tau$, the following
+statistics were computed for both the noises and populations: 
+Pearson, Spearman and Kendall correlations, $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, 
+$\Dsq_l$, $\Dsq_u$, $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$. 
+Values were plotted for noises and populations against $\tau$. If the hypothesis is
+correct that copula structure between locations of environmental noise can be 
+transmitted to produce similar dependence structure for ecological variables, 
+plots should be similar for populations and noises.
 
@@
 you model evolution down the tree in this way and then look at the bivariate distribution of
-trait on the tips of the tree. What do you get? I guess you get some copula structure?-->
-
-To begin with, we chose a two habitat patch ($i = 1, 2$) with AR(1) model [Eq.
-\ref{eq.AR1}] where population $P_{i}$ at time $(t+1)$ depends upon its past value at $(t)$ and the iid noise (assumed to be caused by environmental fluctuations) $\epsilon_{i} (t)$.
-
-\begin{equation}\label{eq.AR1}
-P_{i}(t+1)=\beta P_{i}(t)+\alpha \epsilon_{i}(t)
-\end{equation}
-
-To keep $Var(P_{i}(t+1))$ equal to unity, we set noise coefficient $\alpha$ in terms of AR coefficient $\beta$ as $\alpha = \sqrt(1-\beta^2)$.
-We showed with Eq.
-\ref{eq.AR1} in the next section that both model selection method and non-parametric analysis indicated similar tail dependence in population copula as that of noise copula.
-<!--Asymmetric sensitivity and copula structure in populations-->
-\textbf{(ii)}
+trait on the tips of the tree. What do you get? I guess you get some copula structure? maybe
+this thinking can lead to some isnights into how copula structure we observed in the species
+trait data might have come about?-->
 

### 2019-02-05 Paper.Rmd
edits of Q2 methods and results

@@
 Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
-available from `BiCopGofTest` in the `VineCopula` package). However, non-parametric statistics verified
+available from `BiCopGofTest` in the `VineCopula` package). However, nonparametric statistics verified
 both that tail dependence was distinct from that of an N copula for the soil data, and that it 
@@
 (Table \ref{tab_bivar_summary}, rows 5-6).
-But our non-parametric statistics confirmed the same conclusions 
+But our nonparametric statistics confirmed the same conclusions 
 (Table \ref{tab_bivar_summary}, row 10).
@@
 
-<!--Non-parametric analysis of tail dependence, green spruce aphid abundance data-->
-Non-parametric statistics verified that tail dependence was asymmetric for the green 
+<!--Nonparametric analysis of tail dependence, green spruce aphid abundance data-->
+Nonparametric statistics verified that tail dependence was asymmetric for the green 
 spruce aphid abundance data.
@@
 (Table \ref{tab_multivar_summary}, rows 7-9), whereas leaf-curling plum aphid first flight
-data showed the reverse. Asymmetry results were verified by non-parametric approaches 
+data showed the reverse. Asymmetry results were verified by nonparametric approaches 
 (e.g., Table \ref{tab_multivar_summary}, row 10). Detailed results are in 
@@
              caption = "Summary of results for multivariate data-sets : \\textbf{num.pairs}: total number of location pairs (excluding self-comparison) considered, \\textbf{num.non-indep}: how many location pair copula are non-independent among them, \\textbf{num.non-N}: number of non-Normal copula as best fit (based on highest AIC-weight), \\textbf{Avg.best.AICw}: average AIC-weights of best fit copula from all location pair, \\textbf{Avg.N.AICw}: average AIC-weights of Normal copula from all location pair, \\textbf{num.good.fit}: number of location pair for which best fit copula have p-values $>$ 0.01
-(both Cramer-von Mises and Kolmogorov-Smirnov method), \\textbf{0.025-CI.LT}: lower and \\textbf{0.975-CI.LT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) lower tail dependence from all location pair, \\textbf{0.025-CI.UT}: lower and \\textbf{0.975-CI.UT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) upper tail dependence from all location pair, \\textbf{num.LT-UT>0}: number of location pair which showed stronger lower tail dependence than upper, Last $6$ rows summarized the non-parametric analysis results for $\\Ps$ statistics : lower and upper quantile ($95\\%$CI) of $\\Ps_l$, $\\Ps_u$, $\\Ps_l-\\Ps_u$ based on resampling location pair.\\label{tab_multivar_summary}",
+(both Cramer-von Mises and Kolmogorov-Smirnov method), \\textbf{0.025-CI.LT}: lower and \\textbf{0.975-CI.LT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) lower tail dependence from all location pair, \\textbf{0.025-CI.UT}: lower and \\textbf{0.975-CI.UT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) upper tail dependence from all location pair, \\textbf{num.LT-UT>0}: number of location pair which showed stronger lower tail dependence than upper, Last $6$ rows summarized the nonparametric analysis results for $\\Ps$ statistics : lower and upper quantile ($95\\%$CI) of $\\Ps_l$, $\\Ps_u$, $\\Ps_l-\\Ps_u$ based on resampling location pair.\\label{tab_multivar_summary}",
              booktabs=TRUE)
@@
 For each of the three model setups, for each of the copulas C, SC, G, SG, J, SJ,
-and for each Kendall correlation value $\tau=0.1,0.2,\ldots,0.9$, 
-we generated 5000 noise pairs $(\epsilon_1(t),\epsilon_2(t))$ from the
+for each Kendall correlation value $\tau=0.1,0.2,\ldots,0.9$, and for each of
+$50$ replicate simulations, we generated 5000 noise pairs 
+$(\epsilon_1(t),\epsilon_2(t))$ from the
 bivariate random variable with the given copula family and the given $\tau$.
@@
 noise and population values for the final 500 
-time steps of the simulation. For each model, copula family, and $\tau$, the following
+time steps of the simulation. For each simulation, the following
 statistics were computed for both the noises and populations: 
@@
 
-To test environmental effect on population copula from the model \ref{eq.AR1} with $\beta = 0.5$, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of non-parametric tail-dependence statistics for both bivariate noise copula and resultant population copula (following model \ref{eq.AR1}) for $50$ simulations.
+The hypothesis that copula structure and tail dependence between environmental 
+variables measured through time in different locations can produce similar 
+structure in ecological variables influenced by the environmental variables was
+largely confirmed, though similarities between environmental- and 
+ecological-variable copula structure and tail dependencies can be reduced when the
+intrinsic dynamics of the ecological variable is strongly nonlinear.
+For simulations using AR(1) models (see section on methods for Q2), copula 
+statistics were always very similar for noise and populations 
+(Fig \ref{fig_Cause4copula_scatter_C_kend_ar1}, 
+\ref{SM-fig_Cause4copula_C_kend_ar1}-\ref{fig_Cause4copula_SJ_kend_ar1}).
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption{Example results showing similarity of copula statistics for 
+environmental-noise inputs and ecological-variable outputs of a dynamical model.  
+Asymmetric tail dependence was similar in both noise inputs and model outputs
+for this model. The 
+AR(1) model and a Clayton copula were used (see methods section for Q2).
+Each point is the mean across 50 replicate simulations for which the same 
+Kendall $\tau$ value was used. Error bars are standard errors 
+and panel headers give Pearson correlation results for
+the points. The regression line through the points (black line) was similar
+to the 1-1 line (green line). 
+See Fig. \ref{SM-fig_Cause4copula_C_kend_ar1} for additional results for the AR(1) model and Clayton copula, 
+Figs \ref{SM-fig_Cause4copula_SC_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1} for AR(1) results for other copula families, and 
+Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1} for
+the other models described in the methods for Q2.\label{fig_Cause4copula_scatter_C_kend_ar1}}
+\end{center}
+\end{figure}
+
+<!--***DAN: got to here-->
+
+To test environmental effect on population copula from the model \ref{eq.AR1} with $\beta = 0.5$, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of nonparametric tail-dependence statistics for both bivariate noise copula and resultant population copula (following model \ref{eq.AR1}) for $50$ simulations.
 Then we plotted mean of each measured quantity with standard error against Kendall's correlation of the noise copula family.
@@
 
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption{Linear regression between 3 non-parametric stats ($\cor_l - \cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) of environmental noise generated from lower tail dependent Clayton copula and of the affected populations for 9 equidistant Kendall's Tau values in a range (0.1 to 0.9 as shown in Fig.
-\ref{SM-fig_Cause4copula_C_kend_ar1}) showed significant Pearson correlation, and henceforth confirmed asymmetric tail dependence in noise copula caused similar tail dependence in population copula for AR(1) model.
-Regression line (black) differing from 1:1 line (green) indicates that asymmetric tail dependence in noise and population copulae are similar but not exactly the same.\label{fig_Cause4copula_scatter_C_kend_ar1}}
-\end{center}
-\end{figure}
-
 To explore the effect of asymmetric sensitivity, we simulated our model (\ref{eq.AS}) for $b=0.1$ , $a=0.2$, $r=0.8$.
@@
 \ref{fig_ms_asym_b_0.1}).
-We also carried out model selection and non-parametric statistical analysis with each copula of Fig.
+We also carried out model selection and nonparametric statistical analysis with each copula of Fig.
 \ref{fig_ms_asym_b_0.1} and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_2}.
 According to our model selection approach (Table \ref{SM-tab_summary_stat_asym_sens}), non-normal copulae (with higher AIC weights) were always best-fit with similar asymmetric tail dependence which gave the support of our hypothesis of the effect of asymmetric global noise copula on populations.
-Our non-parametric analysis results (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_2}) were consistent with the model selection approach.
+Our nonparametric analysis results (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_2}) were consistent with the model selection approach.
 
@@
 In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data.
-As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the non-parametric analysis and model fitting results (see first column of Table \ref{tab_multivar_summary}).
-On the other hand, Figure \ref{fig_skewness_spearman}(B) demonstrated that averaged skewness was significantly higher than expected if the leaf-plum curling aphid first flight data came from a normal copula, which was also confirmed by upper-tail dependence results from model selection and non-parametric results (see second column of Table \ref{tab_multivar_summary}).
+As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the nonparametric analysis and model fitting results (see first column of Table \ref{tab_multivar_summary}).
+On the other hand, Figure \ref{fig_skewness_spearman}(B) demonstrated that averaged skewness was significantly higher than expected if the leaf-plum curling aphid first flight data came from a normal copula, which was also confirmed by upper-tail dependence results from model selection and nonparametric results (see second column of Table \ref{tab_multivar_summary}).
 Figure \ref{fig_skewness_spearman}(C) showed that averaged skewness was lower than expected if the plankton data came from a normal copula.
-Our previous results on non-parametric analysis and model fitting also supported lower tail dependence (see third column of Table \ref{tab_multivar_summary}).
+Our previous results on nonparametric analysis and model fitting also supported lower tail dependence (see third column of Table \ref{tab_multivar_summary}).
 But, as $p>0.05$, we could not reject $H_{0}$ in that case.
@@
 `log`(mean) of population abundance data was significantly different from $10000$ Normal surrogates (Spearman-correlation and marginal distribution preserving) for $25\%$ of total species considered.
-For example, the "Corn leaf aphid" (having significantly upper tail dependence from model selection and non-parametric analysis) showed mean curvature of regression line close to zero, whereas `r 10000 -(tlsp6$p_left_mncurv*10000)` of the $10000$ Normal surrogates showed higher curvature, indicating here Normal copula didn't hold linear relationship between `log`(spatial variance) and `log`(spatial mean) (see Fig \ref{fig_sTL_mncurv_sp6}).
+For example, the "Corn leaf aphid" (having significantly upper tail dependence from model selection and nonparametric analysis) showed mean curvature of regression line close to zero, whereas `r 10000 -(tlsp6$p_left_mncurv*10000)` of the $10000$ Normal surrogates showed higher curvature, indicating here Normal copula didn't hold linear relationship between `log`(spatial variance) and `log`(spatial mean) (see Fig \ref{fig_sTL_mncurv_sp6}).
 
@@
 Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results.
-So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different non-parametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selection results.
+So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different nonparametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selection results.
 Our approach could help to interpret the synchronous behavior of a species from several meta-populations.

### 2019-02-05 SupportingInformation.Rmd
fixed typo
edits of Q2 methods and results

--- 2019-01-28 SupportingInformation.Rmd
@@
 \caption[For definitions of nonparametric statistics]{Schematic diagram supporting definitions of nonparametric tail dependence statistics.
-The red and blue lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
+The blue and red lines are, respectively, $u+v=2l_b$ and $u+v=2u_b$.\label{fig_stat_image}}
 \end{center}
@@
 
-# Surrogate testing with bivariate datasets for non-parametric statistics \label{surrog_test}
-
-For bivariate datasets, surrogates were produced as follows.
+# Surrogate testing with bivariate datasets for nonparametric statistics \label{surrog_test}
+
+For bivariate data sets, surrogates were produced as follows.
 
@@
 
-<!--Do the main computations for non-parametric analysis of aphid count data-->
+<!--Do the main computations for nonparametric analysis of aphid count data-->
 ```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 
-<!--Do the main computations for non-parametric analysis of aphid ff data-->
+<!--Do the main computations for nonparametric analysis of aphid ff data-->
 ```{r stat_aphid_ff, echo=F, cache=T,  cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 ### Model selection approach \label{Ex_pns_Model_selection}
-### Non-parametric analysis of tail dependence  \label{Ex_pns_NPA}
+### nonparametric analysis of tail dependence  \label{Ex_pns_NPA}
                           ----moved to Paper.Rmd----
@@
 
-<!--Do the main computations for non-parametric analysis of plankton data-->
+<!--Do the main computations for nonparametric analysis of plankton data-->
 ```{r stat_plankton_north_sea, echo=F, cache=T, cache=T,  cache.extra=list(seed,data_plankton_north_sea,longs_pns,lats_pns,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 
-<!--Do the main computations for non-parametric analysis of methane-flux data-->
+<!--Do the main computations for nonparametric analysis of methane-flux data-->
 ```{r stat_methane, echo=F, cache=T, cache=T,  cache.extra=list(seed,methane_data,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
-\caption[AR(1) model : comparing correlations and non-parametric statistic for between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{AR(1) model : results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+\caption[AR(1) model : comparing correlations and nonparametric statistic for between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{AR(1) model : results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 nonparametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
 p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[AR(1) model : comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{AR(1) model : results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from upper tail dependent \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+\caption[AR(1) model : comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{AR(1) model : results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 nonparametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from upper tail dependent \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
 p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
-Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with pearson correlation p-value $< 0.05$, then they are significantly similar.\label{fig_Cause4copula_SC_kend_ar1}}
+Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with Pearson correlation p-value $< 0.05$, then they are significantly similar.\label{fig_Cause4copula_SC_kend_ar1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[AR(1) model : comparing correlations and non-parametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
+\caption[AR(1) model : comparing correlations and nonparametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from upper tail dependent \textbf{Gumbel} copula.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[AR(1) model : comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
+\caption[AR(1) model : comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from lower tail dependent \textbf{Survival Gumbel} copula.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[AR(1) model : comparing correlations and non-parametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
+\caption[AR(1) model : comparing correlations and nonparametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from upper tail dependent \textbf{Joe} copula \label{fig_Cause4copula_J_kend_ar1}}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[AR(1) model: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
+\caption[AR(1) model: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from lower tail dependent \textbf{Survival Joe} copula.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak nonise: results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 non-parametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula (having normal marginal with standard deviation 0.1) and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
+\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: results of comparison between 3 types of correlation (i.e.
+Spearman, Kendall, Pearson) and 6 nonparametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula (having normal marginal with standard deviation 0.1) and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
 p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
-Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with pearson correlation p-value $< 0.05$, then they are significantly similar.
+Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with Pearson correlation p-value $< 0.05$, then they are significantly similar.
 \label{fig_Cause4copula_C_kend_sRicker_nsd_0.1}}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Survival Clayton} copula having normal marginal with standard deviation 0.1.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Gumbel} copula having normal marginal with standard deviation 0.1.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Gumbel} copula having normal marginal with standard deviation 0.1.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Joe} copula having normal marginal with standard deviation 0.1.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak nonise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Joe} copula having normal marginal with standard deviation 0.1.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with strong noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with strong noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Clayton} copula  having normal marginal with standard deviation 1.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with strong noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with strong noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_SC_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Survival Clayton} copula  having normal marginal with standard deviation 1.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with strong noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with strong noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_G_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Gumbel} copula  having normal marginal with standard deviation 1.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with strong noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with strong noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_SG_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Gumbel} copula having normal marginal with standard deviation 1.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with strong noise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with strong noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_J_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Joe} copula having normal marginal with standard deviation 1.
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak nonise: comparing correlations and non-parametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
+\caption[Stochastic Ricker model with weak nonise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
 \ref{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Joe} copula having normal marginal with standard deviation 1.

### 2019-02-05 Paper.Rmd
I have been editing Q2, the second class of models

@@
 
-# Background, concepts and methods for Q1: detecting copula structure in data \label{Methods}
+# Background, concepts and methods for Q1: Detecting copula structure in data \label{Methods}
 
@@
 <!--Question 2 analyses-->
-# Background, concepts and methods for Q2: causes/mechanisms of non-normal copula structure \label{Causes}
+# Background, concepts and methods for Q2: Causes and mechanisms of non-normal copula structure \label{Causes}
 
@@
 $\sigma=1$, model dynamics are strongly nonlinear. We refer to these as
-the weak-noise and strong-noise cases. The $epsilon_i(t)$ were again standard
+the weak-noise and strong-noise cases. The $\epsilon_i(t)$ were again standard
 normal and independent through time, but were related (see below) across locations.
@@
 <!--***DAN: Note to self, this model, and also the other one we study in this section,
-are actually much more general than populatin. Should replace P_i with E_i where E_i is
+are actually much more general than population. Should replace P_i with E_i where E_i is
 just some ecological variable. Could be deviaitons from a pop from carrying capacity, could be a 
@@
 
-The second mechanism, which was also summarized in the introduction,
-relates to Liebig's law of the minimum, or, more generally, to the potential for 
-environmental variables to influence ecological variables through nonlinear relationships.
-
-Secondly, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) which behaved differently beyond a certain expectation threshold (as shown in Eq.
-\ref{eq.AS_g}).
-To test the asymmetric sensitivity, we constructed a two habitat patch model for population $P_{i} (i=1,2)$ as follows;
-
-\begin{equation}\label{eq.AS}
-P_{i}(t+1) = b P_{i}(t) + g(\varepsilon_{i}(t)) + h(\delta_{i}(t))
-\end{equation}
-
-$b$ was the linear growth rate.
-Global environmental fluctuation function $g(\varepsilon)$ can take up any one form between $g_{1}$ and $g_{2}$ (Eq.
-\ref{eq.AS_g}, Fig.
-\ref{fig_ms_asym_b_0.1}) where $\varepsilon=(\varepsilon_{1}, \varepsilon_{2})$ was an iid through time with mean zero and co-variance matrix $\Sigma_{\varepsilon}$.
-Local environmental fluctuation function $h(\delta)$ varied according to Eq.
-\ref{eq.AS_h} where $a$ was the scaling parameter between $0$ to $1$.
-$\delta=(\delta_{1}, \delta_{2})$ was an iid across $i=(1,2)$ and time with mean zero and co-variance matrix $\Sigma_{\delta}$.
-With this model we showed tail dependence in population copula changed from lower to upper tail dependence as global noise copula $g(\varepsilon)$ changed its form from $g_{1}$ to $g_{2}$ (Eq.
-\ref{eq.AS_g}).
-
-\begin{align}
-g(\varepsilon) = \begin{cases} 
-g_{1} = \begin{cases}
-\varepsilon & \text{if $\varepsilon< 0$}; \\
-0 & \text{otherwise}.
-\end{cases}\\ \\
-g_{2} = \begin{cases}
-\varepsilon & \text{if $\varepsilon> 0$}; \\
-0 & \text{otherwise}\label{eq.AS_g}.
-\end{cases}
-\end{cases} \text{where } \varepsilon = N(\textbf{0},\Sigma_{\varepsilon})\;\ \text{with} \;\ \Sigma_{\varepsilon} = \begin{bmatrix} 1 & r \\ r & 1 \end{bmatrix}
-\end{align}
-
-\begin{equation}\label{eq.AS_h}
-h(\delta) = a \delta = aN(\textbf{0},\Sigma_{\delta})\;\ \text{with} \;\ \Sigma_{\delta} = \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}
-\end{equation}
-
-# Results for \hyperref[Intro]{\textbf{(Q2)}}: similar tail dependence in population copula as that of noise copula \label{Results_Q2}
+<!--***DAN: note to self. Suppose deviations from the 
+regression line for log met rate v log M are 
+more costly at one end than at another, for whatever reason. That is a proximate
+mechanism for asymmetric tail dependence-->
+
+The second mechanism we will explore, which was also summarized in the 
+Introduction,
+relates to Liebig's law of the minimum, or, more generally, to the fact that 
+environmental variables can often influence ecological variables through nonlinear 
+relationships. If an environmental variable influences an ecological variable
+more strongly in one of its tails, we explored using a model 
+whether the ecological variable
+could then exhibit asymmetric tail dependence between measurements made in two
+locations. Let $E_i(t)$ again be an ecological variable measured in locations
+$i=1,2$. Assume $E_i(t)$ has dynamics that follow 
+$E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where $g$ is a function
+describing how the environmental variable $\epsilon_i(t)$ influences
+$E_i(t)$, and the $\delta_i(t)$ are standard-normally distributed 
+local noise, independent across time and sampling locations. We used $b=0.1$ and
+$a=0.2$, and $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
+time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
+$\cov(\epsilon_1,\epsilon_2)=0.8$. We used $g$ equal to $g_1$ or $g_2$ in two different simulations:
+$g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise; and 
+$g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
+$g_1$ represent environmental effects that occur only for negative 
+values of the environmental variables, and $g_2$ represent environmental effects 
+that occur only for positive values. Note that both $(\epsilon_1,\epsilon_2)$
+and $(\delta_1,\delta_2)$ have N copula structure, being drawn from 
+bivariate normal distributions. Hence the Moran-like mechanism discussed
+previously cannot be operating in this model.
+
+We simulated the model once using $g_1$ and once using $g_2$, for 5000 time
+steps in both cases, retaining $E_1(t)$ and $E_2(t)$ for the final 500 time 
+steps of each simulation. We applied our model selection algorithms and 
+non-arametric statistics, which were previously applied to 
+data to answer Q1, to the 
+simulation outputs to discover if the mechanism built into the model could
+give rise to non-N copula structure and asymmetric tail dependence in the 
+ecological variable.
+
+# Results for Q2: Moran-like effects and asymmetric dependencies \label{Results_Q2}
 
@@
 (Fig \ref{fig_Cause4copula_scatter_C_kend_ar1}, 
-\ref{SM-fig_Cause4copula_C_kend_ar1}-\ref{fig_Cause4copula_SJ_kend_ar1}).
+\ref{SM-fig_Cause4copula_C_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1}),
+though there were significant but minor differences for some statistics
+and simulations.
 
@@
 
-<!--***DAN: got to here-->
-
-To test environmental effect on population copula from the model \ref{eq.AR1} with $\beta = 0.5$, we measured 3 types of correlations (Spearman, Kendall, Pearson) and different types of nonparametric tail-dependence statistics for both bivariate noise copula and resultant population copula (following model \ref{eq.AR1}) for $50$ simulations.
-Then we plotted mean of each measured quantity with standard error against Kendall's correlation of the noise copula family.
-Fig.
-\ref{SM-fig_Cause4copula_C_kend_ar1} compares the statistics from population copula and the copula made of noises drawn from Clayton (lower tail dependent copula family).
-We also carried out a paired t-test to compute the p-values of each statistic for noise and population data and found that the nature of tail dependence in noise and populations remained similar but they were not exactly the same except in Pearson correlation coefficient which verified the Moran effect on the population data caused by environmental noise.
-(For code see [`Plotter_Cause4copula_stat`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
-Linear regression between tail statistics from noise and population copula (Fig.
-\ref{fig_Cause4copula_scatter_C_kend_ar1}) confirmed that the asymmetric tail dependence in noise copula caused similar tail dependence in population copula.
-We did the same analysis with other asymmetric uni-parameter copula family and got similar tail dependence in resultant population copula (see Figs.
-\ref{SM-fig_Cause4copula_SC_kend_ar1} - \ref{SM-fig_Cause4copula_SJ_kend_ar1}).
-We also measured tail-dependence between noise and population copula for non-linear stochastic Ricker model and found that our hypothesis of similar tail dependence held true for weak non-linearity regime (see Appendix \ref{SM-cause4copula_ricker}).
-
-<!--
-***SG : commented out this section as it gives the similar interpretation but in a more complex way.
-our approach is of two-fold : First, we calculated the parameter values of the noise-copula (for example, we chose Clayton with lower tail dependence, in this case the parameter of interest is $\theta$, which determines the strength of association of the variables [@nelsen2006_copula]) using `iRho` or `iTau` function from `copula` package for Spearman's Rho or Kendall's Tau, respectively.
-Then we simulated our model (\ref{eq.AR1}) and estimated parameters of the outcome bivariate population-copula, with last $500$ simulated points for each habitat patch (see Fig.
-\ref{fig_Cause4copula_params_GOF_500}).
-We also tested the goodness of fit of the estimated parameters for the population-copula and its standard error due to sample size.
-We reported p-values from goodness of fit tests for both statistics (Cramer-von Mises and Kolmogorov-Smirnov).
-p-values greater or equal to $0.05$ (above the dashed line in Fig.
-\ref{fig_Cause4copula_params_GOF_500}) were considered as a good fit of estimated parameters.
-Fig.
-\ref{SM-fig_Cause4copula_params_GOF_2500} represents similar feature as of Fig.
-\ref{fig_Cause4copula_params_GOF_500} but here we considered last $2500$ simulated population data points for our parameters-estimation, thus making our comparison more sensitive on basis of p-value threshold at $0.05$.
-(For code see [`Plotter_Cause4copula_GOF`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function).
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf}
-\caption{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last 500 points for both noise and population data for comparison).
-p-values $> 0.05$ (above horizontal dashed line) indicate good fit (not significantly different) of population copula with same noise copula family.
-\label{fig_Cause4copula_params_GOF_500}}
-\end{center}
-\end{figure}
--->
-
-To explore the effect of asymmetric sensitivity, we simulated our model (\ref{eq.AS}) for $b=0.1$ , $a=0.2$, $r=0.8$.
-$10000$ points were generated and we kept only the last $1000$ points to plot the population-copula (Fig.
-\ref{fig_ms_asym_b_0.1}) for each specific functional form of $g$  (For code see [`Simulator_asym_sens`](https://github.com/sghosh89/BIVAN/blob/master/asym_sensitivity.R) function).
-When asymmetric global noise had only non-positive contributions at two habitat patches ($g_{1}$ in the first panel of Fig.
-\ref{fig_ms_asym_b_0.1}), lower tail dependence appeared in the global noise copula and so thus in affected population copula (last subplot in the first panel of Fig.
-\ref{fig_ms_asym_b_0.1}).
-Similarly, when global noises had only non-negative contributions at two habitat patches ($g_{2}$ in the second panel of Fig.
-\ref{fig_ms_asym_b_0.1}), upper tail dependence appeared in the global noise copula and so thus in the outcome population copula (last subplot in the second panel of Fig.
-\ref{fig_ms_asym_b_0.1}).
-We also carried out model selection and nonparametric statistical analysis with each copula of Fig.
-\ref{fig_ms_asym_b_0.1} and summarized our results in Tables \ref{SM-tab_summary_stat_asym_sens} - \ref{SM-tab_stat_npa_asym_2}.
-According to our model selection approach (Table \ref{SM-tab_summary_stat_asym_sens}), non-normal copulae (with higher AIC weights) were always best-fit with similar asymmetric tail dependence which gave the support of our hypothesis of the effect of asymmetric global noise copula on populations.
-Our nonparametric analysis results (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_2}) were consistent with the model selection approach.
+For weak noise and using our nonlinear model (methods for Q2), copula statistics
+were again very similar for noise and model outputs 
+(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}).
+Though there were again significant but small differences for some statistics
+and simulations, the overall hypothesis that copula structure in the noise translates
+to similar but not identical copula structure in output populations, at least
+as measured with the tail dependence and correlation statistics we used, was
+supported. Since many ecological models are nonlinear, this result provides the 
+important expectation that there will be a Moran-effect-like correspondence 
+between noise and model-output copula structure for typical ecological dynamics,
+as long as the noise is small enough that dynamics stay relatively close
+to model equilibrium. Results that hold for "weak noise" in this sense
+are common in ecology. 
+We repeated the analysis with $r=X$, with similar results. The deterministic 
+<!--***DAN: Shya, please replace X in the above line with the actual value(s)
+of r that you used.-->
+one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
+at $K$ when $r<1$ (so-called under-compensating dynamics), but exhibits 
+an oscillatory approach to a stable equilibrium at $K$ when $r>1$ (over-compensating
+dynamics).
+
+For strong noise and using our nonlinear model (methods for Q2), 
+copula statistics, generally speaking, were approximately similar 
+between noise and model
+outputs; however, similarity was reduced relative to the AR(1) 
+and weak-noise results, and, for a few simulations, tail dependence 
+asymmetry statistics
+even had opposite signs for noise and model outputs 
+(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). For instance, using a C copula with a large Kendall $\tau$,
+$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
+population model outputs; and likewise for SG and SJ copulas. Dynamics were
+strongly nonlinear for the model parameters we used, and nevertheless similarities
+between noise and model-output copula structure, according to the statistics 
+we used, were more noticable than discrepancies.
+We again repeated the analysis with $r=X$, with similar results.
+<!--***DAN: Right, Shya? Please replace X with the actual value(s)-->
+
+Our model with environmental effects which differed in the
+two tails produced simulated ecological variables with visually apparent asymmetry 
+of tail dependence and non-N copula structure (Fig. \ref{fig_ms_asym_b_0.1}).
+Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) and
+our nonparametric statistics (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_2}) statistically confirmed the visual result.
+We repeated our analyses with $b=-0.1, 0.9, -0.9$, with similar conclusions
+in all cases.
+<!--***DAN: Right Shya? I have written this, but I want you to show me these results
+(without putting them in the paper) to make sure the statement is accurate-->
 
@@
 \includegraphics[width=15 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
-\caption{Effects of asymmetric global noise copula (resulting from the asymmetric dependence of $g$ on $\varepsilon$) on two habitat patch model (Eq.
-\ref{eq.AS}) with growth rate $b=0.1$: copula of two populations ($P_{1}, P_{2}$) shaped according to the asymmetry of noise copula\label{fig_ms_asym_b_0.1}}
+\caption{If environmental effects operate asymmetrically in their tails on
+ecological variables, it can result in non-N copula structure and asymmetric 
+tail dependence across space in the ecological variables. Shown are the last
+$500$ points from simulations described in the text.\label{fig_ms_asym_b_0.1}}
 \end{center}
 \end{figure}
+<!--***DAN: Shya, pls remove the left 4 panels, and also the headers of the right
+two panels - we can put that information in the caption, as I have done-->
 

### 2019-02-05 SupportingInformation.Rmd
I have been editing Q2, the second class of models

@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
-\caption[AR(1) model : comparing correlations and nonparametric statistic for between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{AR(1) model : results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 nonparametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
-p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
-\label{fig_Cause4copula_C_kend_ar1}}
-\end{center}
-\end{figure}
+\caption[Comparison of copula structure of noise and model output, AR(1) model,
+C copula.]{Comparison of copula structure of noise inputs and model 
+outputs, AR(1) model, C copula. Red and blue points give 
+means over 50 replicate simulations of the statistics listed on the left axes, 
+red points are values for noise inputs and blue points are for
+model outputs. Error bars are standard errors. 
+$P$-values (purple points, right axis) are for a 
+paired $t$-test of the null hypothesis that the distributions of 50 noise and 
+output values have the same mean; $p$-values less than $0.05$ indicate that 
+copula structure as measured by the listed statistic differs significant
+between noise inputs and model outputs. See methods section for Q2 
+for additional details.\label{fig_Cause4copula_C_kend_ar1}}
+\end{center}
+\end{figure}
+<!--***DAN: Shya, can you please change the top legend of this and the similar-format
+figures so instead of saying "noise copula" for the red points and "population copula" 
+for the blue points it says "input noise copula" for the red points and "model output 
+copula" for the blue ones? This is because one of our models is much more general than
+just a population model, so I have changed the langauge in the text to reflect that.
+Please be aware that when you make this change it will mean the next knit will take 
+several hours, unless you go to the effort of reorganizing your code to put the sims
+and the plots in separate chunks. So maybe make this change right before you go home,
+so you can knit over night. Also please let me know, so I can knit overnight, too.
+
+At the same time that you do that above, please see if you can fix the alignment
+between the bottom three panels of Fig S32 (and subsequent analogous figures) so they
+line up with the panels above that.-->
 
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[AR(1) model : comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{AR(1) model : results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 nonparametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from upper tail dependent \textbf{Survival Clayton} copula and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
-p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
-Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with Pearson correlation p-value $< 0.05$, then they are significantly similar.\label{fig_Cause4copula_SC_kend_ar1}}
+\caption[Comparison of copula structure of noise and model output, AR(1) model, 
+SC copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+model, SC copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
+\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
+details.\label{fig_Cause4copula_SC_kend_ar1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[AR(1) model : comparing correlations and nonparametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from upper tail dependent \textbf{Gumbel} copula.
-\label{fig_Cause4copula_G_kend_ar1}}
+\caption[Comparison of copula structure of noise and model output, AR(1) model, 
+G copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+model, G copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
+\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
+details.\label{fig_Cause4copula_G_kend_ar1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[AR(1) model : comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from lower tail dependent \textbf{Survival Gumbel} copula.
-\label{fig_Cause4copula_SG_kend_ar1}}
+\caption[Comparison of copula structure of noise and model output, AR(1) model, 
+SG copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+model, SG copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
+\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
+details.\label{fig_Cause4copula_SG_kend_ar1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[AR(1) model : comparing correlations and nonparametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from upper tail dependent \textbf{Joe} copula \label{fig_Cause4copula_J_kend_ar1}}
+\caption[Comparison of copula structure of noise and model output, AR(1) model, 
+J copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+model, J copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
+\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
+details.\label{fig_Cause4copula_J_kend_ar1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[AR(1) model: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{AR(1) model: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_SC_kend_ar1} but for noise generated from lower tail dependent \textbf{Survival Joe} copula.
-\label{fig_Cause4copula_SJ_kend_ar1}}
-\end{center}
-\end{figure}
-
-# Similar tail dependence in population copula as that of noise copula for stochastic Ricker model \label{cause4copula_ricker}
-
-To find whether nature of tail dependence in noise copula preserved for affected population copula or it would be destroyed because of underlying non-linearity in the model, we simulated the following stochastic Ricker model with intrinsic growth rate $r = 0.5$ and carrying capacity $K = 100$ for two population patches ($i = 1, 2$);
-
-\begin{equation}\label{eq.sRicker}
-P_{i}(t+1)=P_{i}(t) exp \left[ r \left\{ 1-\frac{P_{i}(t)}{K} \right\} +\epsilon_{i}(t) \right]
-\end{equation}
-
-The noise ($\epsilon_{i}$), initially drawn from a bivariate asymmetric copula family, was transformed into a normally distribution with mean zero and standard deviation $0.1$ (considered as weak noise) or $1$ (considered as strong noise), for code see [`GetNoise`](https://github.com/sghosh89/BIVAN/blob/master/Cause4copula.R) function.
-For weak noise, we found similar tail dependence in noise copula and the resultant population copula from the stochastic Ricker model (Figs.
-\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} - \ref{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}) as we obtained in linear AR(1) model (see \nameref{MT-Results_Q2} in the main text).
-But for strong noise the tail-dependence between noise and population copula was not significantly similar (Figs.
-\ref{fig_Cause4copula_C_kend_sRicker_nsd_1} - \ref{fig_Cause4copula_SJ_kend_sRicker_nsd_1}).
-We repeated the whole analysis with $r > 1$ (i.e.
-Ricker model's over-compensatory dynamical regime) and came up with same conclusion that strong noise with non-linearity could disturb the tail-dependence information translated from noise to population copula.
+\caption[Comparison of copula structure of noise and model output, AR(1) model, 
+SJ copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+model, SJ copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
+\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
+details.\label{fig_Cause4copula_SJ_kend_ar1}}
+\end{center}
+\end{figure}
 
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: results of comparison between 3 types of correlation (i.e.
-Spearman, Kendall, Pearson) and 6 nonparametric stats ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_u$ and $\Dsq_l$) of environmental noise generated from lower tail dependent \textbf{Clayton} copula (having normal marginal with standard deviation 0.1) and of the affected populations for 9 equidistant Kendall's correlation ($\tau$) values in a range (0.1 to 0.9) (Initially $5000$ points are generated but we kept last $500$ points for both noise and population data for comparison).
-p-values from paired t-test if $> 0.05$ (above horizontal purple line) indicates exactly same estimate between noise and population copulae.
-Last row in the figure compared the linear regression line with the 1:1 green line and if noise and population tail-dependence stats came up with Pearson correlation p-value $< 0.05$, then they are significantly similar.
-\label{fig_Cause4copula_C_kend_sRicker_nsd_0.1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, weak noise, C copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, C copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Survival Clayton} copula having normal marginal with standard deviation 0.1.
-\label{fig_Cause4copula_SC_kend_sRicker_nsd_0.1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, weak noise, SC copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, SC copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Gumbel} copula having normal marginal with standard deviation 0.1.
-\label{fig_Cause4copula_G_kend_sRicker_nsd_0.1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, weak noise, G copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, G copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details.\label{fig_Cause4copula_G_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Gumbel} copula having normal marginal with standard deviation 0.1.
-\label{fig_Cause4copula_SG_kend_sRicker_nsd_0.1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, weak noise, SG copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, SG copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details. \label{fig_Cause4copula_SG_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Joe} copula having normal marginal with standard deviation 0.1.
-\label{fig_Cause4copula_J_kend_sRicker_nsd_0.1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, weak noise, J copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, J copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details.\label{fig_Cause4copula_J_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with weak noise: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Joe} copula having normal marginal with standard deviation 0.1.
-\label{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, weak noise, SJ copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, SJ copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details.\label{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with strong noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_C_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Clayton} copula  having normal marginal with standard deviation 1.
-\label{fig_Cause4copula_C_kend_sRicker_nsd_1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, strong noise, C copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, C copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with strong noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Clayton}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_SC_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Survival Clayton} copula  having normal marginal with standard deviation 1.
-\label{fig_Cause4copula_SC_kend_sRicker_nsd_1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, strong noise, SC copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, SC copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with strong noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_G_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Gumbel} copula  having normal marginal with standard deviation 1.
-\label{fig_Cause4copula_G_kend_sRicker_nsd_1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, strong noise, G copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, G copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details.\label{fig_Cause4copula_G_kend_sRicker_nsd_1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with strong noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Gumbel}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_SG_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Gumbel} copula having normal marginal with standard deviation 1.
-\label{fig_Cause4copula_SG_kend_sRicker_nsd_1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, strong noise, SG copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, SG copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details. \label{fig_Cause4copula_SG_kend_sRicker_nsd_1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with strong noise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_J_kend_sRicker_nsd_0.1} but for noise generated from upper tail dependent \textbf{Joe} copula having normal marginal with standard deviation 1.
-\label{fig_Cause4copula_J_kend_sRicker_nsd_1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, strong noise, J copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, J copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details.\label{fig_Cause4copula_J_kend_sRicker_nsd_1}}
 \end{center}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Stochastic Ricker model with weak nonise: comparing correlations and nonparametric statistic between noise-copula (\textbf{Survival Joe}) and population copula for different Kendall's correlation]{Stochastic Ricker model with strong noise: similar analysis as depicted in Fig.
-\ref{fig_Cause4copula_SJ_kend_sRicker_nsd_0.1} but for noise generated from lower tail dependent \textbf{Survival Joe} copula having normal marginal with standard deviation 1.
-\label{fig_Cause4copula_SJ_kend_sRicker_nsd_1}}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, strong noise, SJ copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, SJ copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details.\label{fig_Cause4copula_SJ_kend_sRicker_nsd_1}}
 \end{center}
@@
 #----------------------------------------------------------
-N<-10000
+N<-5000
 r<-0.8
 a<-0.2
-numkeep_last<-1000
+numkeep_last<-500
 temp<-get_noise_glob_loc_mat(N=N,r=r)

### 2019-02-05 Paper.Rmd
a few more edits on same section

@@
 <!--***DAN: Right Shya? I have written this, but I want you to show me these results
-(without putting them in the paper) to make sure the statement is accurate-->
+(without putting them in the paper) to make sure the statement is accurate.
+But we should put the figs in the sup mat, not the stats results, I think, but we could put the figs-->
 

### 2019-02-06 Paper.Rmd
code changed

@@
 time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
-$\cov(\epsilon_1,\epsilon_2)=0.8$. We used $g$ equal to $g_1$ or $g_2$ in two different simulations:
+$\\cov(\epsilon_1,\epsilon_2)=0.8$. We used $g$ equal to $g_1$ or $g_2$ in two different simulations:
 $g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise; and 

### 2019-02-06 SupportingInformation.Rmd
code changed

@@
 # for clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_C_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
                                            params=params,p0=c(0,0),model=model)
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+saveRDS(ar1_C_spear_beta0.5,paste(resloc,"ar1_C_spear_beta0.5.RDS",sep=""))
+
+ar1_C_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
                                           params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_C_kend_beta0.5,paste(resloc,"ar1_C_kend_beta0.5.RDS",sep=""))
 
@@
 # for clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_C_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                              num_keep_last=500,resloc=resloc,
                                               params=params,p0=c(0,0),model=model)
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+saveRDS(ar1_C_spear_negbeta0.5,paste(resloc,"ar1_C_spear_negbeta0.5.RDS",sep=""))
+
+ar1_C_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                               num_keep_last=500,resloc=resloc,
                                                params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_C_kend_negbeta0.5,paste(resloc,"ar1_C_kend_negbeta0.5.RDS",sep=""))
+```
+
+```{r plot_cause4copula_stat_ar1_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_C_spear_beta0.5,ar1_C_spear_negbeta0.5,ar1_C_kend_beta0.5,ar1_C_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-3
+resloc0<-"./Results/Cause4copula_results/"
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_C_spear_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_C_kend_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_C_spear_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_C_kend_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
 ```
@@
 # for survival clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_SC_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                            num_keep_last=500,resloc=resloc,
                                             params=params,p0=c(0,0),model=model)
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+saveRDS(ar1_SC_spear_beta0.5,paste(resloc,"ar1_SC_spear_beta0.5.RDS",sep=""))
+
+ar1_SC_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
                                            params=params,p0=c(0,0),model=model)
 
+saveRDS(ar1_SC_kend_beta0.5,paste(resloc,"ar1_SC_kend_beta0.5.RDS",sep=""))
 # for negative beta
@@
 # for survival clayton with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_SC_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                               num_keep_last=500,resloc=resloc,
                                                params=params,p0=c(0,0),model=model)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+saveRDS(ar1_SC_spear_negbeta0.5,paste(resloc,"ar1_SC_spear_negbeta0.5.RDS",sep=""))
+
+ar1_SC_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                              num_keep_last=500,resloc=resloc,
                                               params=params,p0=c(0,0),model=model)
+saveRDS(ar1_SC_kend_negbeta0.5,paste(resloc,"ar1_SC_kend_negbeta0.5.RDS",sep=""))
+```
+
+```{r plot_cause4copula_stat_ar1_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SC_spear_beta0.5,ar1_SC_spear_negbeta0.5,ar1_SC_kend_beta0.5,ar1_SC_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-13
+resloc0<-"./Results/Cause4copula_results/"
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SC_spear_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_SC_kend_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SC_spear_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_SC_kend_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
 ```
@@
 # for gumbel with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_G_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                              num_keep_last=500,resloc=resloc,
                                               params=params,p0=c(0,0),model=model)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+saveRDS(ar1_G_spear_beta0.5,paste(resloc,"ar1_G_spear_beta0.5.RDS",sep=""))
+
+ar1_G_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_G_kend_beta0.5,paste(resloc,"ar1_G_kend_beta0.5.RDS",sep=""))
 
@@
 # for gumbel with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_G_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                              num_keep_last=500,resloc=resloc,
                                               params=params,p0=c(0,0),model=model)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+saveRDS(ar1_G_spear_negbeta0.5,paste(resloc,"ar1_G_spear_negbeta0.5.RDS",sep=""))
+
+ar1_G_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_G_kend_negbeta0.5,paste(resloc,"ar1_G_kend_negbeta0.5.RDS",sep=""))
+```
+
+```{r plot_cause4copula_stat_ar1_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_G_spear_beta0.5,ar1_G_spear_negbeta0.5,ar1_G_kend_beta0.5,ar1_G_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-4
+resloc0<-"./Results/Cause4copula_results/"
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_G_spear_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_G_kend_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_G_spear_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_G_kend_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
 ```
@@
 # for survival gumbel with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_SG_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                              num_keep_last=500,resloc=resloc,
                                               params=params,p0=c(0,0),model=model)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+saveRDS(ar1_SG_spear_beta0.5,paste(resloc,"ar1_SG_spear_beta0.5.RDS",sep=""))
+
+ar1_SG_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_SG_kend_beta0.5,paste(resloc,"ar1_SG_kend_beta0.5.RDS",sep=""))
 
@@
 # for survival gumbel with spearman and kendall method
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_SG_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                              num_keep_last=500,resloc=resloc,
                                               params=params,p0=c(0,0),model=model)
-
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+saveRDS(ar1_SG_spear_negbeta0.5,paste(resloc,"ar1_SG_spear_negbeta0.5.RDS",sep=""))
+
+ar1_SG_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_SG_kend_negbeta0.5,paste(resloc,"ar1_SG_kend_negbeta0.5.RDS",sep=""))
+```
+
+```{r plot_cause4copula_stat_ar1_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SG_spear_beta0.5,ar1_SG_spear_negbeta0.5,ar1_SG_kend_beta0.5,ar1_SG_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-14
+resloc0<-"./Results/Cause4copula_results/"
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SG_spear_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_SG_kend_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SG_spear_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_SG_kend_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
 ```
@@
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_J_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_J_kend_beta0.5,paste(resloc,"ar1_J_kend_beta0.5.RDS",sep=""))
 
@@
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_J_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_J_kend_negbeta0.5,paste(resloc,"ar1_J_kend_negbeta0.5.RDS",sep=""))
+```
+
+```{r plot_cause4copula_stat_ar1_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_J_kend_beta0.5,ar1_J_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-6
+resloc0<-"./Results/Cause4copula_results/"
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_J_kend_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_J_kend_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
 ```
@@
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_SJ_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_SJ_kend_beta0.5,paste(resloc,"ar1_SJ_kend_beta0.5.RDS",sep=""))
 
@@
 
-Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
+ar1_SJ_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
                                              params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_SJ_kend_negbeta0.5,paste(resloc,"ar1_SJ_kend_negbeta0.5.RDS",sep=""))
+```
+
+```{r plot_cause4copula_stat_ar1_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SJ_kend_beta0.5,ar1_SJ_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-16
+resloc0<-"./Results/Cause4copula_results/"
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SJ_kend_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SJ_kend_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
 ```
@@
   # for clayton with spearman and kendall method
-    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
-
-    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
+                                           params=params,p0=c(0,0),model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_C_spear.RDS",sep=""))
+
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=c(0,0),model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_C_kend.RDS",sep=""))
+
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_C, echo=F, results="hide"}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-3
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_C_spear.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+    readres<-readRDS(paste(resloc,"ricker_C_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
   }
@@
   # for survival clayton with spearman and kendall method
-    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
-
-    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
+                                           params=params,p0=c(0,0),model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_SC_spear.RDS",sep=""))
+
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=c(0,0),model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_SC_kend.RDS",sep=""))
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_SC, echo=F, results="hide"}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-13
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_SC_spear.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+    readres<-readRDS(paste(resloc,"ricker_SC_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
   }
@@
   # for gumbel with spearman and kendall method
-    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
-
-    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
+                                           params=params,p0=c(0,0),model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_G_spear.RDS",sep=""))
+
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=c(0,0),model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_G_kend.RDS",sep=""))
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_G, echo=F, results="hide"}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-4
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_G_spear.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+    readres<-readRDS(paste(resloc,"ricker_G_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
   }
@@
   # for survival gumbel with spearman and kendall method
-    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
-
-    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
+                                           params=params,p0=c(0,0),model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_SG_spear.RDS",sep=""))
+
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=c(0,0),model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_SG_kend.RDS",sep=""))
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_SG, echo=F, results="hide"}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-14
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_SG_spear.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+    readres<-readRDS(paste(resloc,"ricker_SG_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
   }
@@
 
-    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=c(0,0),model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_J_kend.RDS",sep=""))
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_J, echo=F, results="hide"}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-6
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_J_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
   }
@@
 
-    Plotter_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,num_keep_last=500,resloc=resloc,
-                          params=params,p0=p0,model=model)
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=c(0,0),model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_J_kend.RDS",sep=""))
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_SJ, echo=F, results="hide"}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-16
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_SJ_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
   }
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
 \includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Joe/Survival Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
@@
   if(ploton==T){
-    pdf(paste0(resloc,"a_",a,"_b_",b,"_r_",r,"_asym_sens.pdf",sep=""),height=5,width=8)
-    op<-par(mfrow=c(2,3),mar=c(4,4.2,1.5,1),mgp=c(2.5,0.6,0))
+    pdf(paste0(resloc,"a_",a,"_b_",b,"_r_",r,"_asym_sens.pdf",sep=""),height=3,width=6)
+    op<-par(mfrow=c(1,2),mar=c(4,4.2,1.5,1),mgp=c(2.5,0.6,0))
     #plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
@@
     geps<-pop_g$g_noise
-    plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=1")
-    plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=2")
-    plot(pop[,1],pop[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2)
-    title(expression(paste("g = ",g[1]," , ")), adj=0.06,cex.main=1.5)
-    mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)
+    #plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=1")
+    #plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=2")
+    plot(pop[,1],pop[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2,col="grey",pch=20,cex=1.4)
+    #title(expression(paste("g = ",g[1]," , ")), adj=0.06,cex.main=1.5)
+    #mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)
 
@@
     geps<-pop_d_g$g_noise
-    plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=1")
-    plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=2")
-    plot(pop_d[,1],pop_d[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2)
-    title(expression(paste("g = ",g[2]," , ")), adj=0.06,cex.main=1.5)
-    mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)
+    #plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=1")
+    #plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=2")
+    plot(pop_d[,1],pop_d[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2,col="grey",pch=20,cex=1.4)
+    #title(expression(paste("g = ",g[2]," , ")), adj=0.06,cex.main=1.5)
+    #mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)
 
@@
 ```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
-summary_stat<-summary_stat[,c(1:2)]
+summary_stat<-summary_stat[c(1:6),]
 rownames(summary_stat)<-c("bestfit.copula.AICw","Normal.copula.AICw","LT.model.avg","UT.model.avg","p.CvM","p.KS","BS.success.percent")

### 2019-02-06 Cause4copula.R

@@ -315,7 +315,7 @@ Plotter_Cause4copula_GOF<-function(N,fcode,method,nsd,num_keep_last,BS,params,p0
 #       resloc : folder location where the plots should be saved
 #       params,p0,model : these are inputs for Simulator_Cause4copula function
 #-----------------------------------------------------------------------------------------------
-Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,nsd,method,lb=0,ub=0.1,num_keep_last,resloc,params,p0,model){
+Sim_Cause4copula_stat<-function(N,numsim=50,fcode,nsd,method,lb=0,ub=0.1,num_keep_last,resloc,params,p0,model){
   
   corcoef_list<-seq(from=0.1,to=0.9,by=0.1)
   
@@ -572,402 +572,46 @@ Plotter_Cause4copula_stat<-function(N,numsim=50,fcode,nsd,method,lb=0,ub=0.1,num
     
   }
   
-  if(method=="spearman"){
-    xlabel<-"Spearman's Rho"
-  }else if(method=="kendall"){
-    xlabel<-"Kendall's Tau"
-  }else{
-    warning("specify method",immediate.=T,call.=T)
-  }
-  
-  # Plotting Spearman correlation btw noise and pop copula
-  pdf(paste0(resloc,BiCopName(fcode,short=F),"_Spearman_vs_",xlabel,".pdf",sep=""),height=4,width=5)
-  op<-par(mar=c(3.5,4.5,2,3.5), mgp=c(1.9,0.5,0))
-  plot(corcoef_list,S_noise_mat[,2],cex=0.5,col="red",xlab=xlabel,ylab="Spearman",xlim=c(0,1),ylim=c(0,1),cex.lab=2,cex.axis=1.5)
-  segments(corcoef_list,S_noise_mat[,1],corcoef_list,S_noise_mat[,3],col='red')
-  bar_len<-0.02
-  segments(corcoef_list-bar_len,S_noise_mat[,1],corcoef_list+bar_len,S_noise_mat[,1],col='red')
-  segments(corcoef_list-bar_len,S_noise_mat[,3],corcoef_list+bar_len,S_noise_mat[,3],col='red')
[…]

### 2019-02-06 Plotter_Cause4copula_stat.R

@@ -1,0 +1,447 @@
+#This is a plotter function to plot the outputs from Sim_Cause4copula_stat function written in Cause4copula.R
+# Input :
+#       res : a big list which is the output from Sim_Cause4copula_stat function
+
+Plotter_Cause4copula_stat<-function(res){
+  
+  resloc<-res$resloc
+  method<-res$method
+  fcode<-res$fcode
+  corcoef_list<-res$corcoef_list
+  S_noise_mat<-res$S_noise_mat
+  K_noise_mat<-res$K_noise_mat
+  P_noise_mat<-res$P_noise_mat
+  S_pop_mat<-res$S_pop_mat
+  K_pop_mat<-res$K_pop_mat
+  P_pop_mat<-res$P_pop_mat
+  Corl_noise_mat<-res$Corl_noise_mat
+  Corl_pop_mat<-res$Corl_pop_mat
+  Coru_noise_mat<-res$Coru_noise_mat
+  Coru_pop_mat<-res$Coru_pop_mat
+  CorlmCoru_noise_mat<-res$CorlmCoru_noise_mat
+  CorlmCoru_pop_mat<-res$CorlmCoru_pop_mat
+  Pl_noise_mat<-res$Pl_noise_mat
+  Pl_pop_mat<-res$Pl_pop_mat
+  Pu_noise_mat<-res$Pu_noise_mat
+  Pu_pop_mat<-res$Pu_pop_mat
+  PlmPu_noise_mat<-res$PlmPu_noise_mat
+  PlmPu_pop_mat<-res$PlmPu_pop_mat
+  D2u_noise_mat<-res$D2u_noise_mat
+  D2u_pop_mat<-res$D2u_pop_mat
+  D2l_noise_mat<-res$D2l_noise_mat
+  D2l_pop_mat<-res$D2l_pop_mat
+  D2umD2l_noise_mat<-res$D2umD2l_noise_mat
+  D2umD2l_pop_mat<-res$D2umD2l_pop_mat
+  pval_S<-res$pval_S
+  pval_K<-res$pval_K
+  pval_P<-res$pval_P
+  pval_Corl<-res$pval_Corl
+  pval_Coru<-res$pval_Coru
+  pval_CorlmCoru<-res$pval_CorlmCoru
+  pval_Pl<-res$pval_Pl
+  pval_Pu<-res$pval_Pu
+  pval_PlmPu<-res$pval_PlmPu
[…]

### 2019-02-06 SupportingInformation.Rmd
adding cache tag

@@
 # This chunk stores results for ar1 model : clayton
+tag_ricker_C<-"999C" #this is a caching tag used later to cache output from present chunk
 
@@
 
-```{r plot_cause4copula_stat_ricker_C, echo=F, results="hide"}
+```{r plot_cause4copula_stat_ricker_C, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_C,mtime("Plotter_Cause4copula.R"))}
 
@@
 #---------------------------------------
+tag_ricker_SC<-"999SC" #this is a caching tag used later to cache output from present chunk
+
 # This chunk stores results for ar1 model : survival clayton
@@
 
-```{r plot_cause4copula_stat_ricker_SC, echo=F, results="hide"}
+```{r plot_cause4copula_stat_ricker_SC, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SC,mtime("Plotter_Cause4copula.R"))}
 
@@
 #---------------------------------------
-# This chunk plots results for ar1 model : clayton
 
@@
 #---------------------------------------
+tag_ricker_G<-"999G" #this is a caching tag used later to cache output from present chunk
+
 # This chunk stores results for ar1 model : gumbel
@@
 
-```{r plot_cause4copula_stat_ricker_G, echo=F, results="hide"}
+```{r plot_cause4copula_stat_ricker_G, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_G,mtime("Plotter_Cause4copula.R"))}
 
@@
 #---------------------------------------
+tag_ricker_SG<-"999SG" #this is a caching tag used later to cache output from present chunk
+
 # This chunk stores results for ar1 model : survival gumbel
@@
 
-```{r plot_cause4copula_stat_ricker_SG, echo=F, results="hide"}
+```{r plot_cause4copula_stat_ricker_SG, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SG,mtime("Plotter_Cause4copula.R"))}
 
@@
 #---------------------------------------
+tag_ricker_J<-"999J" #this is a caching tag used later to cache output from present chunk
+
 # This chunk stores results for ar1 model : joe
@@
 
-```{r plot_cause4copula_stat_ricker_J, echo=F, results="hide"}
+```{r plot_cause4copula_stat_ricker_J, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_J,mtime("Plotter_Cause4copula.R"))}
 
@@
 #---------------------------------------
+tag_ricker_SJ<-"999SJ" #this is a caching tag used later to cache output from present chunk
+
 # This chunk stores results for ar1 model : survival joe
@@
 
-```{r plot_cause4copula_stat_ricker_SJ, echo=F, results="hide"}
+```{r plot_cause4copula_stat_ricker_SJ, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SJ,mtime("Plotter_Cause4copula.R"))}
 

### 2019-02-06 SupportingInformation.Rmd
error fixed

@@
   # for clayton with spearman and kendall method
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                            num_keep_last=500,resloc=resloc,
-                                           params=params,p0=c(0,0),model=model)
+                                           params=params,p0=p0,model=model)
 
@@
 
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
                                           num_keep_last=500,resloc=resloc,
-                                          params=params,p0=c(0,0),model=model)
+                                          params=params,p0=p0,model=model)
 
@@
   # for survival clayton with spearman and kendall method
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                            num_keep_last=500,resloc=resloc,
-                                           params=params,p0=c(0,0),model=model)
+                                           params=params,p0=p0,model=model)
 
@@
   # for gumbel with spearman and kendall method
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                            num_keep_last=500,resloc=resloc,
-                                           params=params,p0=c(0,0),model=model)
+                                           params=params,p0=p0,model=model)
 
@@
   # for survival gumbel with spearman and kendall method
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
                                            num_keep_last=500,resloc=resloc,
-                                           params=params,p0=c(0,0),model=model)
+                                           params=params,p0=p0,model=model)
 

### 2019-02-06 SupportingInformation.Rmd
I fixed a bug (I think) that was preventing knitting

@@
 ```{r read_BMR_birds_and_copulaplot, echo=F, results='hide'}
+library(readxl)
 set.seed(seed)
@@
 
-    saveRDS(myres,paste(resloc,"ricker_J_kend.RDS",sep=""))
+    saveRDS(myres,paste(resloc,"ricker_SJ_kend.RDS",sep=""))
   }

### 2019-02-06 Paper.Rmd
minor changes to comments

@@
 steps of each simulation. We applied our model selection algorithms and 
-non-arametric statistics, which were previously applied to 
+non-parametric statistics, which were previously applied to 
 data to answer Q1, to the 
-simulation outputs to discover if the mechanism built into the model could
+simulation outputs to discover if the model could
 give rise to non-N copula structure and asymmetric tail dependence in the 
@@
 Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) and
-our nonparametric statistics (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_2}) statistically confirmed the visual result.
+our nonparametric statistics (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_2}) 
+statistically confirmed the visual result.
 We repeated our analyses with $b=-0.1, 0.9, -0.9$, with similar conclusions
@@
 (without putting them in the paper) to make sure the statement is accurate.
-But we should put the figs in the sup mat, not the stats results, I think, but we could put the figs-->
+But we should put the figs in the sup mat, not the stats results, I think, but we could put the figs. Or maybe we do want to put the stats. Let's look at them anyway. Maybe the sentence needs to
+be rewritten if the results for other b are actually different.-->
 

### 2019-02-06 SupportingInformation.Rmd
fixing error

@@
 summary_stat<-summary_stat[c(1:6),]
-rownames(summary_stat)<-c("bestfit.copula.AICw","Normal.copula.AICw","LT.model.avg","UT.model.avg","p.CvM","p.KS","BS.success.percent")
+rownames(summary_stat)<-c("bestfit.copula.AICw","Normal.copula.AICw","LT.model.avg","UT.model.avg","p.CvM","p.KS")
 knitr::kable(summary_stat, booktabs = T, escape=F,
              format = "latex", linesep = "",
-             digits=3, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$"),
+             digits=3, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$",
+                                   "b=-0.1, g$_{1}$","b=-0.1, g$_{2}$",
+                                   "b=0.9, g$_{1}$","b=0.9, g$_{2}$",
+                                   "b=-0.9, g$_{1}$","b=-0.9, g$_{2}$"),
       #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption="summary of model selection results for asymmetric sensitivity : \\textbf{bestfit.copula.AICw} and \\textbf{Normal.copula.AICw} represent the AIC-weightage for best fit copula and Normal copula, \\textbf{LT.model.avg} and \\textbf{UT.model.avg} represent AIC-weight based model averaged lower and upper tail dependence respectively, \\textbf{p.CvM} and \\textbf{p.KS} are the p-values coming from goodness of fit test (Cramer-von Mises and Kolmogorov-Smirnov statistics based), \\textbf{BS.success.percent} percentage of successful bootstraps in goodness of fit test.
+      caption="summary of model selection results for asymmetric sensitivity : \\textbf{bestfit.copula.AICw} and \\textbf{Normal.copula.AICw} represent the AIC-weightage for best fit copula and Normal copula, \\textbf{LT.model.avg} and \\textbf{UT.model.avg} represent AIC-weight based model averaged lower and upper tail dependence respectively, \\textbf{p.CvM} and \\textbf{p.KS} are the p-values coming from goodness of fit test (Cramer-von Mises and Kolmogorov-Smirnov statistics based).
 Each column summarizes those statistics for each copula plotted in Fig.

### 2019-02-07 SupportingInformation.Rmd
asym_sens : table with model sel and npa stat for all b and g1,g2 added in suppmat

@@
 #knitr::opts_knit$set(kable.force.latex = TRUE)
-tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
+#tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
 
@@
 
-```{r tab_stat_npa_asym_1,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[1]])
-knitr::kable(temp, booktabs = T, 
+```{r tab_stat_npa_asym,echo=F,results="markup"}
+library(kableExtra)
+
+tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
+tab_stat_npa_asym<-c()
+for(i in c(1:8)){
+  temp<-make_table_stat_fn(data=tab_asym[[i]])
+  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
+}
+
+knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
              format = "latex",linesep = "",escape=F,
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption ="A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_1}")
-```
-
-```{r tab_stat_npa_asym_2,echo=F,results="markup"}
-temp<-make_table_stat_fn(data=tab_asym[[2]])
-knitr::kable(temp, booktabs = T, 
-             format = "latex",linesep = "",escape=F,
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption ="A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence.\\label{tab_stat_npa_asym_2}")
+             caption.short = "Results for tests of whether asymmetry of tail dependence was significant
+                              compared to a normal-copula null hypothesis.",
+             caption ="A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more)
+                      than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ 
+                      or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail         
+                      dependence.\\label{tab_stat_npa_asym}")%>%
+             kable_styling()%>%
+             group_rows("b=0.1, g$_{1}$", 2,4, escape=F) %>%
+             group_rows("b=0.1, g$_{2}$", 5,7, escape=F) %>%
+             group_rows("b=-0.1, g$_{1}$", 8,10, escape=F) %>%
+             group_rows("b=-0.1, g$_{2}$", 11,13, escape=F) %>%
+             group_rows("b=0.9, g$_{1}$", 14,16, escape=F) %>%
+             group_rows("b=0.9, g$_{2}$", 17,19, escape=F) %>%
+             group_rows("b=-0.9, g$_{1}$", 20,22, escape=F) %>%
+             group_rows("b=-0.9, g$_{2}$", 23,25, escape=F)
 ```

### 2019-02-07 SupportingInformation.Rmd
commented the table:asym_npa results which could produce error: no sure

@@
 
-knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
-             format = "latex",linesep = "",escape=F,
-             caption.short = "Results for tests of whether asymmetry of tail dependence was significant
-                              compared to a normal-copula null hypothesis.",
-             caption ="A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more)
-                      than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ 
-                      or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail         
-                      dependence.\\label{tab_stat_npa_asym}")%>%
-             kable_styling()%>%
-             group_rows("b=0.1, g$_{1}$", 2,4, escape=F) %>%
-             group_rows("b=0.1, g$_{2}$", 5,7, escape=F) %>%
-             group_rows("b=-0.1, g$_{1}$", 8,10, escape=F) %>%
-             group_rows("b=-0.1, g$_{2}$", 11,13, escape=F) %>%
-             group_rows("b=0.9, g$_{1}$", 14,16, escape=F) %>%
-             group_rows("b=0.9, g$_{2}$", 17,19, escape=F) %>%
-             group_rows("b=-0.9, g$_{1}$", 20,22, escape=F) %>%
-             group_rows("b=-0.9, g$_{2}$", 23,25, escape=F)
+#knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
+#             format = "latex",linesep = "",escape=F,
+#             caption.short = "Results for tests of whether asymmetry of tail dependence was significant
+#                              compared to a normal-copula null hypothesis.",
+#             caption ="A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more)
+#                      than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ 
+#                      or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail         
+#                      dependence.\\label{tab_stat_npa_asym}")#%>%
+ #            kable_styling()%>%
+ #            group_rows("b=0.1, g$_{1}$", 2,4, escape=F) %>%
+ #            group_rows("b=0.1, g$_{2}$", 5,7, escape=F) %>%
+ #            group_rows("b=-0.1, g$_{1}$", 8,10, escape=F) %>%
+ #            group_rows("b=-0.1, g$_{2}$", 11,13, escape=F) %>%
+ #            group_rows("b=0.9, g$_{1}$", 14,16, escape=F) %>%
+ #            group_rows("b=0.9, g$_{2}$", 17,19, escape=F) %>%
+ #            group_rows("b=-0.9, g$_{1}$", 20,22, escape=F) %>%
+ #            group_rows("b=-0.9, g$_{2}$", 23,25, escape=F)
 ```

### 2019-02-07 Paper.Rmd
adding operator name in style file for \cov

@@
 time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
-$\\cov(\epsilon_1,\epsilon_2)=0.8$. We used $g$ equal to $g_1$ or $g_2$ in two different simulations:
+$\cov(\epsilon_1,\epsilon_2)=0.8$. We used $g$ equal to $g_1$ or $g_2$ in two different simulations:
 $g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise; and 

### 2019-02-07 Paper.Rmd
Now, I could knit makefile upto this : But still don't know how to deal with duplicate row indexing...so commented out

@@
 Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) and
-our nonparametric statistics (Tables \ref{SM-tab_stat_npa_asym_1} -\ref{SM-tab_stat_npa_asym_2}) 
+our nonparametric statistics (Tables \ref{SM-tab_stat_npa_asym}) 
 statistically confirmed the visual result.

### 2019-02-07 SupportingInformation.Rmd
Now, I could knit makefile upto this : But still don't know how to deal with duplicate row indexing...so commented out

--- 2019-02-07 SupportingInformation.Rmd
@@
 }
-
+gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.9, g$_{1}$","b=0.9, g$_{2}$","b=-0.9, g$_{1}$","b=-0.9, g$_{2}$")
 knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
@@
                       or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail         
-                      dependence.\\label{tab_stat_npa_asym}")%>%
-             kable_styling()%>%
-             group_rows("b=0.1, g$_{1}$", 2,4, escape=F) %>%
-             group_rows("b=0.1, g$_{2}$", 5,7, escape=F) %>%
-             group_rows("b=-0.1, g$_{1}$", 8,10, escape=F) %>%
-             group_rows("b=-0.1, g$_{2}$", 11,13, escape=F) %>%
-             group_rows("b=0.9, g$_{1}$", 14,16, escape=F) %>%
-             group_rows("b=0.9, g$_{2}$", 17,19, escape=F) %>%
-             group_rows("b=-0.9, g$_{1}$", 20,22, escape=F) %>%
-             group_rows("b=-0.9, g$_{2}$", 23,25, escape=F)
+                      dependence.\\label{tab_stat_npa_asym}")#%>%
+ #            kable_styling()%>%
+ #            group_rows(index=setNames(c(3,3,3,3,3,3,3,3), gtext))
+ #            group_rows("b=0.1, g$_{1}$", 2,4, escape=F) %>%
+ #            group_rows("b=0.1, g$_{2}$", 5,7, escape=F) %>%
+ #            group_rows("b=-0.1, g$_{1}$", 8,10, escape=F) %>%
+ #            group_rows("b=-0.1, g$_{2}$", 11,13, escape=F) %>%
+ #            group_rows("b=0.9, g$_{1}$", 14,16, escape=F) %>%
+ #            group_rows("b=0.9, g$_{2}$", 17,19, escape=F) %>%
+ #            group_rows("b=-0.9, g$_{1}$", 20,22, escape=F) %>%
+ #            group_rows("b=-0.9, g$_{2}$", 23,25, escape=F)
 ```

### 2019-02-07 SupportingInformation.Rmd
workaround to recent knitting and table labelling problems

--- 2019-02-07 SupportingInformation.Rmd
@@
 }
-
+gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.9, g$_{1}$","b=0.9, g$_{2}$","b=-0.9, g$_{1}$","b=-0.9, g$_{2}$")
+tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.9,0.9,-0.9,-0.9),each=3),
+  g=rep(c(1,2,1,2,1,2,1,2),each=3),tab_stat_npa_asym)
 knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,

### 2019-02-07 Paper.Rmd
trivial
minor edits to the Q2 sectios

@@
 total plant community biomass or other community variable from an average value,
-or flux of a biogeochemical variable.
-
-To explore the importance of nonlinear dynamics for our hypothersis, we also considered 
+or flux of a biogeochemical variable such as methane or CO$_2$..
+
+To explore the importance of nonlinear dynamics for our hypothesis, we also considered 
 nonlinear population models based on $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right) \right)+\sigma\epsilon_i(t))$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
@@
 $\sigma=1$, model dynamics are strongly nonlinear. We refer to these as
-the weak-noise and strong-noise cases. The $\epsilon_i(t)$ were again standard
+the weak-noise and strong-noise cases, though the importance of the noise here
+is that, when it is strong, it brings the nonlinearities of the model strongly into play. 
+The $\epsilon_i(t)$ were again standard
 normal and independent through time, but were related (see below) across locations.
@@
 
-The second mechanism we will explore, which was also summarized in the 
+The second mechanism we explored, which was also summarized in the 
 Introduction,
@@
 $E_i(t)$, and the $\delta_i(t)$ are standard-normally distributed 
-local noise, independent across time and sampling locations. We used $b=0.1$ and
+local noise, independent across time and sampling locations. We used $b=0.1, -0.1, 0.9, -0.9$ and
 $a=0.2$, and $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
 time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
-$\cov(\epsilon_1,\epsilon_2)=0.8$. We used $g$ equal to $g_1$ or $g_2$ in two different simulations:
+$\cov(\epsilon_1,\epsilon_2)=0.8$. Negative values of $b$ correspond to overcompensating dynamics 
+and positive values correspond to undercompensating dynamics; larger-magnitude $b$
+mean slower return time of the model to equilibrium after a perturbation. 
+We used $g$ equal to $g_1$ or $g_2$ in two different simulations:
 $g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise; and 
@@
 
-We simulated the model once using $g_1$ and once using $g_2$, for 5000 time
+For each value of $b$, we simulated the model once using $g_1$ and once using $g_2$, for 5000 time
 steps in both cases, retaining $E_1(t)$ and $E_2(t)$ for the final 500 time 
@@
 as long as the noise is small enough that dynamics stay relatively close
-to model equilibrium. Results that hold for "weak noise" in this sense
+to the model equilibrium. Results that hold for "weak noise" in this sense
 are common in ecology. 
@@
 <!--***DAN: Shya, please replace X in the above line with the actual value(s)
-of r that you used.-->
+of r that you used. But please be sure
+the results were "similar". I encourage you to show me the actual results if there is 
+any doubt in your mind.-->
 one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
-at $K$ when $r<1$ (so-called under-compensating dynamics), but exhibits 
-an oscillatory approach to a stable equilibrium at $K$ when $r>1$ (over-compensating
+at $K$ when $r<1$ (undercompensating dynamics), but exhibits 
+an oscillatory approach to $K$ when $r>1$ (overcompensating
 dynamics).
@@
 outputs; however, similarity was reduced relative to the AR(1) 
-and weak-noise results, and, for a few simulations, tail dependence 
+and weak-noise cases, and, for a few simulations, tail dependence 
 asymmetry statistics
@@
 We again repeated the analysis with $r=X$, with similar results.
-<!--***DAN: Right, Shya? Please replace X with the actual value(s)-->
+<!--***DAN: Right, Shya? Please replace X with the actual value(s), but please be sure
+the results were "similar". I encourage you to show me the actual results if there is 
+any doubt in your mind.-->
 
@@
 in all cases.
-<!--***DAN: Right Shya? I have written this, but I want you to show me these results
-(without putting them in the paper) to make sure the statement is accurate.
-But we should put the figs in the sup mat, not the stats results, I think, but we could put the figs. Or maybe we do want to put the stats. Let's look at them anyway. Maybe the sentence needs to
-be rewritten if the results for other b are actually different.-->
+<!--***DAN: Dan to rewrite this para after seeing the complete collection of results-->
 
@@
 \end{figure}
-<!--***DAN: Shya, pls remove the left 4 panels, and also the headers of the right
-two panels - we can put that information in the caption, as I have done-->
 

### 2019-02-07 Paper.Rmd
getting close to done (for now) with Q2, just need a couple small things from Shya

@@
 $E_i(t)$, and the $\delta_i(t)$ are standard-normally distributed 
-local noise, independent across time and sampling locations. We used $b=0.1, -0.1, 0.9, -0.9$ and
-$a=0.2$, and $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
-time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
-$\cov(\epsilon_1,\epsilon_2)=0.8$. Negative values of $b$ correspond to overcompensating dynamics 
+local noise, independent across time and sampling locations. We used $a=0.2$, and 
+$b=0.1, -0.1, 0.9, -0.9$ in different simulations. 
+Negative values of $b$ correspond to overcompensating dynamics 
 and positive values correspond to undercompensating dynamics; larger-magnitude $b$
 mean slower return time of the model to equilibrium after a perturbation. 
+The $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
+time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
+$\cov(\epsilon_1,\epsilon_2)=0.8$. 
 We used $g$ equal to $g_1$ or $g_2$ in two different simulations:
@@
 Our model with environmental effects which differed in the
-two tails produced simulated ecological variables with visually apparent asymmetry 
-of tail dependence and non-N copula structure (Fig. \ref{fig_ms_asym_b_0.1}).
+two tails could produce simulated ecological variables with visually apparent asymmetry 
+of tail dependence and non-N copula structure, depending on the value of $b$ used. 
+For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ (Fig. \ref{SM-fig_ms_asym_b_-0.1}), 
+for both $g_1$ and $g_2$,
+asymmetric tail dependence and non-N copula structure were visually apparent.
+Lower-tail (respectively, upper-tail) effects of the noise ($g_1$, resp., $g_2) 
+produced lower-tail (resp., upper-tail) dependence in the ecological variable.
 Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) and
 our nonparametric statistics (Tables \ref{SM-tab_stat_npa_asym}) 
-statistically confirmed the visual result.
-We repeated our analyses with $b=-0.1, 0.9, -0.9$, with similar conclusions
-in all cases.
-<!--***DAN: Dan to rewrite this para after seeing the complete collection of results-->
+statistically confirmed the visual result for $b=0.1, -0.1$. Results were more
+complex for $b=0.9$ or $-0.9$, indicating that strongly autocorrelated (or strongly 
+negatively autocorrelated dynamics) interact with the effect we have decribed.
+Ecological-variable copulas were strongly non-N for $g_1$, but not for $g_2$ 
+and asymmetries of tail dependence were not particularly noticable 
+(Figs \ref{SM-fig_ms_asym_b_0.9}-\ref{SM-fig_ms_asym_b_-0.9} and Tables 
+\ref{SM-tab_summary_stat_asym_sens}-\ref{SM-tab_stat_npa_asym}).
 

### 2019-02-07 SupportingInformation.Rmd
getting close to done (for now) with Q2, just need a couple small things from Shya

--- 2019-02-07 SupportingInformation.Rmd
@@
 summary_stat<-summary_stat[c(1:6),]
-rownames(summary_stat)<-c("bestfit.copula.AICw","Normal.copula.AICw","LT.model.avg","UT.model.avg","p.CvM","p.KS")
+rownames(summary_stat)<-c("Best-fit copula AICw","N-copula AICw","Model-avg. LT","Model-avg. UT","p, CvM","p, KS")
 knitr::kable(summary_stat, booktabs = T, escape=F,
@@
                                    "b=-0.9, g$_{1}$","b=-0.9, g$_{2}$"),
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-      caption="summary of model selection results for asymmetric sensitivity : \\textbf{bestfit.copula.AICw} and \\textbf{Normal.copula.AICw} represent the AIC-weightage for best fit copula and Normal copula, \\textbf{LT.model.avg} and \\textbf{UT.model.avg} represent AIC-weight based model averaged lower and upper tail dependence respectively, \\textbf{p.CvM} and \\textbf{p.KS} are the p-values coming from goodness of fit test (Cramer-von Mises and Kolmogorov-Smirnov statistics based).
-Each column summarizes those statistics for each copula plotted in Fig.
-\\ref{MT-fig_ms_asym_b_0.1}.\\label{tab_summary_stat_asym_sens}")
+      caption.short="Asymmetric sensitivity model, model selection results",
+      caption="Asymmetric sensitivity model, model selection results. 
+The output of simulations from the second model detailed in 
+the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
+\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.9}) were subjected to the model 
+selection procedures detailed in the 
+methods section for Q1. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
+N-copula AICw: the AIC weight of the N copula family; Model-avg. LT: the lower-tail 
+statistic, weighted-averaged across models using AICw; Model-avg. UT: same, but for the 
+upper-tail statistic; p, CvM: p-value for goodness of fit test of the best-fitting copula
+family using the test based on the Cramer-von Mises statistic; p, KS: p-value for goodness of fit 
+test of the best-fitting copula family using the test based on the Kolmogorov-Smirnov 
+statistic.\\label{tab_summary_stat_asym_sens}")
 ```
@@
 }
-
-#knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
-#             format = "latex",linesep = "",escape=F,
-#             caption.short = "Results for tests of whether asymmetry of tail dependence was significant
-#                              compared to a normal-copula null hypothesis.",
-#             caption ="A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more)
-#                      than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ 
-#                      or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail         
-#                      dependence.\\label{tab_stat_npa_asym}")#%>%
+gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.9, g$_{1}$","b=0.9, g$_{2}$","b=-0.9, g$_{1}$","b=-0.9, g$_{2}$")
+tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.9,0.9,-0.9,-0.9),each=3),
+  g=rep(c(1,2,1,2,1,2,1,2),each=3),tab_stat_npa_asym)
+knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
+             format = "latex",linesep = "",escape=F,
+             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
+             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
+The output of simulations from the second model detailed in 
+the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
+\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.9}) were subjected to nonparamtric statistical 
+analyses decribed in the methods section for Q1. The listed statistic was computed for the
+output of the model with the indicated b and g, and the value was compared to 1000 values
+of the same statistic computed on surrogate time series randomized to have N-copula structure
+(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
+given statistic on the data was less (respectively, more) than its value on $X$ of the 
+surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
+indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
+lower-tail dependence compared to an N-copula null hypothesis. `FS' stands for `first slice
+and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
+0.9 to 1.\\label{tab_stat_npa_asym}")#%>%
  #            kable_styling()%>%
+ #            group_rows(index=setNames(c(3,3,3,3,3,3,3,3), gtext))
  #            group_rows("b=0.1, g$_{1}$", 2,4, escape=F) %>%
@@
 
-<!--
-# Cause of non-Normal copula structure : Asymmetric sensitivity and copula structure in populations \label{Cause_asym_sens}
--->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=15 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$.\label{fig_ms_asym_b_-0.1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=15 cm]{./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.9$.\label{fig_ms_asym_b_0.9}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=15 cm]{./Results/asym_sens_results/a_0.2_b_-0.9_r_0.8_asym_sens.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.9$.\label{fig_ms_asym_b_-0.9}}
+\end{center}
+\end{figure}
 

### 2019-02-07 Paper.Rmd
fixed a table reference link that was not working

@@
 For row 10, the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented.
-If ranks were $>975$ or $<25$, asymmetry of tail dependence was significant compared to a N-copula null hypothesis.",
+If ranks were $>975$ or $<25$, asymmetry of tail dependence was significant compared to a N-copula null hypothesis.\\label{tab_bivar_summary}",
              booktabs=T, linesep = "\\addlinespace",

### 2019-02-07 Paper.Rmd
editing completed

--- 2019-02-07 Paper.Rmd
@@
 \begin{center}
-\includegraphics[width=15 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
 \caption{If environmental effects operate asymmetrically in their tails on
@@
 tail dependence across space in the ecological variables. Shown are the last
-$500$ points from simulations described in the text.\label{fig_ms_asym_b_0.1}}
+$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text.\label{fig_ms_asym_b_0.1}}
 \end{center}

### 2019-02-07 SupportingInformation.Rmd
editing completed

@@
 between the bottom three panels of Fig S32 (and subsequent analogous figures) so they
-line up with the panels above that.-->
+line up with the panels above that.
+***Shya: Done.-->
 
@@
 
+```{r overwritting_plots_asym_cop,echo=F,results="hide"}
+pdf("./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf",height=3,width=6)
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.3,0))
+pop<-ans_b_0.1$pop
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+text(0.1,0.95,"A",cex=2)
+pop<-ans_b_0.1$pop_d
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+text(0.1,0.95,"B",cex=2)
+par(op)
+dev.off()
+
+pdf("./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf",height=3,width=6)
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.3,0))
+pop<-ans_b_neg0.1$pop
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+text(0.1,0.95,"A",cex=2)
+pop<-ans_b_neg0.1$pop_d
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+text(0.1,0.95,"B",cex=2)
+par(op)
+dev.off()
+
+pdf("./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf",height=3,width=6)
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.3,0))
+pop<-ans_b_0.9$pop
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+text(0.1,0.95,"A",cex=2)
+pop<-ans_b_0.9$pop_d
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+text(0.1,0.95,"B",cex=2)
+par(op)
+dev.off()
+
+pdf("./Results/asym_sens_results/a_0.2_b_-0.9_r_0.8_asym_sens.pdf",height=3,width=6)
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.3,0))
+pop<-ans_b_neg0.9$pop
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+text(0.1,0.95,"A",cex=2)
+pop<-ans_b_neg0.9$pop_d
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+text(0.1,0.95,"B",cex=2)
+par(op)
+dev.off()
+```
+
 ```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
@@
 tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.9,0.9,-0.9,-0.9),each=3),
-  g=rep(c(1,2,1,2,1,2,1,2),each=3),tab_stat_npa_asym)
+  g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
 knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
@@
 indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
-lower-tail dependence compared to an N-copula null hypothesis. `FS' stands for `first slice
+lower-tail dependence compared to an N-copula null hypothesis. `FS' stands for `first slice'
 and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
@@
 \begin{center}
-\includegraphics[width=15 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
 \caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$.\label{fig_ms_asym_b_-0.1}}
@@
 \begin{center}
-\includegraphics[width=15 cm]{./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf}
 \caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.9$.\label{fig_ms_asym_b_0.9}}
@@
 \begin{center}
-\includegraphics[width=15 cm]{./Results/asym_sens_results/a_0.2_b_-0.9_r_0.8_asym_sens.pdf}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.9_r_0.8_asym_sens.pdf}
 \caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.9$.\label{fig_ms_asym_b_-0.9}}

### 2019-02-07 SupportingInformation.Rmd
minor plotting change

@@
 pdf("./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf",height=3,width=6)
-op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.3,0))
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
 pop<-ans_b_0.1$pop
@@
 pdf("./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf",height=3,width=6)
-op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.3,0))
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
 pop<-ans_b_neg0.1$pop
@@
 pdf("./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf",height=3,width=6)
-op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.3,0))
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
 pop<-ans_b_0.9$pop
@@
 pdf("./Results/asym_sens_results/a_0.2_b_-0.9_r_0.8_asym_sens.pdf",height=3,width=6)
-op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.3,0))
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
 pop<-ans_b_neg0.9$pop

### 2019-02-08 Paper.Rmd
minor typo

@@
 asymmetric tail dependence and non-N copula structure were visually apparent.
-Lower-tail (respectively, upper-tail) effects of the noise ($g_1$, resp., $g_2) 
+Lower-tail (respectively, upper-tail) effects of the noise ($g_1$, resp., $g_2$) 
 produced lower-tail (resp., upper-tail) dependence in the ecological variable.
@@
 complex for $b=0.9$ or $-0.9$, indicating that strongly autocorrelated (or strongly 
-negatively autocorrelated dynamics) interact with the effect we have decribed.
+negatively autocorrelated dynamics) interact with the effect we have described.
 Ecological-variable copulas were strongly non-N for $g_1$, but not for $g_2$ 

### 2019-02-08 Paper.Rmd
working on editing Q3 sections and accompanying sup mat material

--- 2019-02-07 Paper.Rmd
@@
 Having demonstrated that non-N copula structure and asymmetric tail dependence are common
-features of ecological data, we explored via mathematical 
+features of ecological data, we addressed Q2 by exploring via mathematical 
 models two possible mechanisms of this structure which 
@@
 more costly at one end than at another, for whatever reason. That is a proximate
-mechanism for asymmetric tail dependence-->
+mechanism for asymmetric tail dependence?-->
 
@@
 
-# Results for Q2: Moran-like effects and asymmetric dependencies \label{Results_Q2}
+# Results for Q2: Moran-like effects and asymmetric dependencies produce non-normal copula structure\label{Results_Q2}
 
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
+\includegraphics[width=15 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
 \caption{If environmental effects operate asymmetrically in their tails on
@@
 tail dependence across space in the ecological variables. Shown are the last
-$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text.\label{fig_ms_asym_b_0.1}}
+$500$ points from simulations described in the text.\label{fig_ms_asym_b_0.1}}
 \end{center}
@@
 <!--Question 3 analyses-->
-# Concepts, methods and models for Q3: Consequences of non-normal copula structure and tail dependence for our understanding of biology \label{Consequences}
+# Background, concepts and methods for Q3: Consequences of non-normal copula structure for our understanding of ecology \label{Consequences}
+
+We addressed Q3 using both models and data. The initial 
+idea was presented in the Introduction that
+the distribution (through time) of a spatially averaged quantity should depend on the 
+nature of the dependence between the local quantities being averaged, including their
+copula structure and tail dependence. For instance, suppose $E_i(t)$ are ecological
+variables measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and suppose 
+the distribution and temporal dynamics of the spatial average $\sum_i E_i(t)/N$ is of
+interest. The $E_i(t)$ could be, for instance, local abundances of a pest or 
+exploited species, or local fluxes of a greenhouse gas. If  $E_i$ and $E_j$ are 
+strongly right-tail dependent for all or most location pairs $i$ and $j$, then
+exceptionally large values tend to occur at the same time in all or most locations.
+This should tend to increase the skewness of the distribution through time of the
+spatial average. Using similar reasoning, left-tail dependence between local variables
+$E_i$ and $E_j$ will tend to decreases skewness of the spatial mean. Skewness of the 
+spatial mean corresponds to the "spikiness" of the spatial-mean time series. 
+The spatial-mean time series may be the quantity of principal importance for 
+cases such as pest or resource abundance or CO$_2$ or methane flux.
+
+We tested these ideas using the focal multivariate datasets of our earlier analyses:
+green spruce aphid abundance data, leaf-curling plum aphid first flight data, \emph{Ceratium
+furca} abundance data, and methane flux data (see Data and Table \ref{tab_multivar_summary}).
+For each dataset we calculated the spatial mean, and then the skewness (through time) of that
+mean. Then, for each dataset separately, we compared the value obtained 
+to a distribution of values 
+of the same quantity calculated for each of 10000 surrogate datasets that were 
+produced from the empirical data by randomizing
+it in a special way to have the copula structure of a multivariate-normal distribution
+(Appendix \ref{SM-Surrog_multi}). The surrogate/randomized datasets had exactly the same
+marginal distributions (i.e., distributions of values for each sampling location)
+as the original data. 
+They also had very similar Kendall or Spearman correlations (the algorithm had two versions,
+one for preserving each correlation coefficient) between pairs of sampling locations
+as the original data. But the surrogates had multivariate normal copula structure. 
+Our comparisons therefore tested the null hypothesis that the skewness of the spatial 
+mean took values on the empirical no different from what one would expect if the copula
+structure of the data were multivariate normal, but the data were otherwise statistically
+unchanged. Significant differences indicate that copula structure contributes to the skewness
+of the spatial mean, i.e., to its "spikiness."
+
+<!--***DAN: stopped here 2019 02 08 1106am central-->
 

### 2019-02-08 SupportingInformation.Rmd
working on editing Q3 sections and accompanying sup mat material

--- 2019-02-07 SupportingInformation.Rmd
@@
 
-\newpage
+# Surrogate testing with multivariate datasets\label{Surrog_multi}
+
+For multivariate datasets, surrogates were produced as follows:
+
+- Given data $x_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$, let $\tau_{ij}$ be the Kendall
+correlation of $x_i(t)$ and $x_j(t)$.  
+- Calculate the covariance matrix of a multivariate normal distribution with standard-normal
+marginals and with pairwise Kendall correlations equal to the $\tau_{ij}$. This is possible
+using the `iTau` function of the `VineCopula` package.  
+- Generate data $a_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$ by taking $T$ independent draws
+from the the multivariate normal distribution of the previous step.  
+- For any $i,t$ for which the datum $x_i(t)$ was missing, replace $a_i(t)$ by `NA`.
+- For each $k$ and $i$ for which $a_i(t)$ is not `NA`, replace the 
+$k$th-smallest (non-`NA`) element of the time series 
+$a_i(t)$, $t=1,\ldots,T$ by the $k$th smallest (non-`NA`) 
+element of $x_i(t)$, $t=1,\ldots,T$.
+Call the result $b_i(t)$. This is the surrogate dataset.
+
+Our code for this algorithm is `ncsurrog` in the `BIVAN` repository 
+(\url{https://github.com/sghosh89/BIVAN}). Because, for each $i$, the final time series 
+$b_i(t)$ is a permuted version of $x_i(t)$ (and with `NA`s for the same $t$), 
+the marginal distributions of the surrogate dataset (i.e., the distributions of 
+values for each sampling location, $i$) are exactly the same as for the original data.
+Because ranks are preserved in passing from $a_i(t)$ to $b_i(t)$, pairwise Kendall 
+correlations for the $b_i(t)$ will be the same as those of the $a_i(t)$. 
+Discrepancies will come only from the presence of NAs and from sampling variation, 
+and should be minor if the 
+numbers of NAs are not too large and $T$ is not too small.
+
 # Consequence of non-Normal copula structure : Effect on Taylor's law \label{Consequence_TL_simulation}

### 2019-02-11 Paper.Rmd
changes to make ncsurrog throw an error when it should

@@
 are common in ecology. 
-We repeated the analysis with $r=X$, with similar results. The deterministic 
-<!--***DAN: Shya, please replace X in the above line with the actual value(s)
-of r that you used. But please be sure
-the results were "similar". I encourage you to show me the actual results if there is 
-any doubt in your mind.-->
+We repeated the analysis with $r=1.3$, with similar results. The deterministic 
 one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
@@
 we used, were more noticable than discrepancies.
-We again repeated the analysis with $r=X$, with similar results.
-<!--***DAN: Right, Shya? Please replace X with the actual value(s), but please be sure
-the results were "similar". I encourage you to show me the actual results if there is 
-any doubt in your mind.-->
+We again repeated the analysis with $r=1.3$. Discrepancies between noise and population
+copula structure in this case were often glaring - apparently noise of standard deviation 1
+interacted very strongly with model nonlinearities when the model was in its overcompensatory 
+regime.
 
@@
 \end{figure}
-
-<!--We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we don’t really know how a change in the environment copula structure would make what kind of change in the population copula structure.-->
-<!--
-## Application : The Moran effect and copula structure in population
-
-```{r read_RES_temp_aphid, echo=F}
-RES_temp_aphid<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_winter_neg_temp.RDS")
-```
-
-Firstly, Our hypothesis was if a non-normal copula structure describes the dependence of an environmental
-variable between two locations, then there can be a non-normal copula structure for the
-dependence of populations in the locations that are influenced by the environmental variables.
-This assumption was inspired by the the idea of Moran-effect [@Moran1953] which was originally used to explain synchronization of widely dispersed populations.
-In the previous section, we already tested our hypothesis with a two habitat patch model (1st order AR method) in noisy environment (having non-Normal copula structure) and found that populations were shaped by similar copula structure as that of local noises.
-Here, we investigated the probable cause of upper tail dependence in the leaf-curling plum aphid's firstflight date throughout all the locations.
-We assumed that average winter temperature which might be spatially synchronized and had asymmetric tail dependence cause the synchronized delay in leaf-plum curling aphid's appearence.
-We tested our hypothesis and indeed found the Moran effect on leaf-plum curling aphid's firstflight data induced by average winter temperature for the period of 35 years (1976-2010).
-We did pair wise model selection with negative average winter temperature (averaged over (n-1)$^{th}$ December to n$^{th}$ March temperature) from the selected 10 locations for which we did model selection with aphid-firstflight data.
-Negative winter temperature was considered for simulation to compare the aphid's firstflight results with temparature copula (aphid's firstflight date were negatively correlated with temperature or in other words positively correlated with negative temparature).
-Model selection results for negative average winter temperature in 10 $\times$ 10 matrices were represented in Table \ref{SM-tab_temp_aphid_bestcop} - Table \ref{SM-tab_temp_aphid_LTmUT}.
-In all cases `r min((RES_temp_aphid$gfc_numBS_success/RES_temp_aphid$gfc_numBS),na.rm=T)*100` $\%$ of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data.
-Table \ref{SM-tab_temp_aphid_LTmUT} showed `r sum(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw <0, na.rm=T)` out of `r (dim(RES_temp_aphid$gfc_p_CvM)[1]^2-dim(RES_temp_aphid$gfc_p_CvM)[1])-2*RES_temp_aphid$num_indep_loc_pair` location pair showed upper tail dependence in temperature copula, which can be considered as the Moran effect results causing dominant upper tail dependence in the copulae with aphid's firstflight date (Table \ref{SM-tab_ff_LTmUT}).
--->
 

### 2019-02-11 SupportingInformation.Rmd
changes to make ncsurrog throw an error when it should

@@
 source("mtime.R") #A function needed for caching
+library(captioner)
 ```
@@
 
-\newpage
+<!--\newpage-->
 ```{r tab_count_LTdep_AICw, echo=F, results='markup'}
@@
 
-\newpage
+<!--\newpage-->
 ```{r tab_count_UTdep_AICw, echo=F, results='markup'}
@@
 
-\newpage
+<!--\newpage-->
 ```{r tab_count_LTmUT, echo=F, results='markup'}
@@
 
-\newpage
+<!--\newpage-->
 ```{r tab_ff_AICw_normal, echo=F, results='markup'}
@@
 \end{figure}
-
-<!--
-\newpage
-## Results: abundance data for *Ceratium furca* in UK seas \label{Ex_pns}
-### Model selection approach \label{Ex_pns_Model_selection}
-### nonparametric analysis of tail dependence  \label{Ex_pns_NPA}
-                          ----moved to Paper.Rmd----
--->
 
@@
 \end{figure}
---->
+-->
 
@@
 
-\newpage
+<!--\newpage-->
 ```{r tab_temp_aphid_p_CvM, echo=F, results='markup'}
@@
 
-\newpage
+<!--\newpage-->
 ```{r tab_temp_aphid_p_KS, echo=F, results='markup'}
@@
 
-\newpage
+<!--\newpage-->
 ```{r tab_temp_aphid_LTdep_AICw, echo=F, results='markup'}
@@
 
-\newpage
+<!--\newpage-->
 ```{r tab_temp_aphid_UTdep_AICw, echo=F, results='markup'}
@@
 
-\newpage
+<!--\newpage-->
 ```{r tab_temp_aphid_LTmUT, echo=F, results='markup'}
@@
 ```
--->
+
 <!--# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}-->
@@
 
-- Given data $x_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$, let $\tau_{ij}$ be the Kendall
+- Given data $x_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$, let $\tau_{ij}$ be the Spearman (or Kendall) 
 correlation of $x_i(t)$ and $x_j(t)$.
 - Calculate the covariance matrix of a multivariate normal distribution with standard-normal
-marginals and with pairwise Kendall correlations equal to the $\tau_{ij}$. This is possible
-using the `iTau` function of the `VineCopula` package.  
+marginals and with pairwise Spearman (or Kendall) correlations equal to the $\tau_{ij}$. This is possible
+using the `iRho` function (or the `iTau function) of the `VineCopula` package
 - Generate data $a_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$ by taking $T$ independent draws
@@
 
-Our code for this algorithm is `ncsurrog` in the `BIVAN` repository 
+\noindent Our code for this algorithm is `ncsurrog` in the `BIVAN` repository 
 (\url{https://github.com/sghosh89/BIVAN}). Because, for each $i$, the final time series 
@@
 values for each sampling location, $i$) are exactly the same as for the original data.
-Because ranks are preserved in passing from $a_i(t)$ to $b_i(t)$, pairwise Kendall 
+Because ranks are preserved in passing from $a_i(t)$ to $b_i(t)$, pairwise Spearman 
+(or Kendall) 
 correlations for the $b_i(t)$ will be the same as those of the $a_i(t)$. 
-Discrepancies will come only from the presence of NAs and from sampling variation, 
-and should be minor if the 
-numbers of NAs are not too large and $T$ is not too small.
+Discrepancies between the $\tau_{ij}$ and the pairwise Spearman (or Kendall) 
+correlations of the
+$b_i(t)$ will therefore come about only from the presence of NAs and from sampling variation. 
+These discrepancies should be minor if the 
+numbers of NAs are not too large and $T$ is not too small. The Kendall version of the
+algorithm throws an error if there are ties in any of the time series. The algorithm
+also throws an error in the event that
+the matrix generated in the second step is not positive semidefinite.
 

### 2019-02-11 makefile.Rmd
various changes
changes to make ncsurrog throw an error when it should

--- 2018-05-15 makefile.Rmd
@@
 ---
-title: "Makefile_linux, compiliation notes"
+title: "BIVAN makefile, compiliation notes"
 author: "Daniel Reuman"
@@
 
+<!--checkpoint package-->
+```{r checkpoint_chunk, echo=F, warning=F, message=F, results="hide"}
+library(checkpoint)
+if (!dir.exists("./.checkpoint/")){
+  dir.create("./.checkpoint/")
+}
+checkpoint("2018-03-29",checkpointLocation = "./")
+```
+
 # Get tex files and pdf files for the supp mat, but the pdf files 
@@
 ```{r tex_for_supp_mat}
+library(rmarkdown)
 rmarkdown::render(input="SupportingInformation.Rmd") 

### 2019-02-11 ncsurrog.R

@@ -1,5 +1,6 @@
 library(copula)
 library(mvtnorm)
+library(matrixcalc)
 
 #This function takes a bunch of time series measured at the same
 #times in different locations and creates surrogate datasets which
@@ -12,6 +13,10 @@ library(mvtnorm)
 #This is better than copsurrod2d because it can do multivariate data
 #but it is worse that copsurrog2d because the target copula can only 
 #be a normal copula.
+#
+#Notes: If the data have ties, Kendall cannot be used. In some cases
+#the algorithm does not work because an intermediate matrix which needs
+#to be positive semi-definite is not. In that case it throws an error.
 #
 #Args
 #m          A N by n matrix, where N is the length of the time series 
@@ -27,6 +32,16 @@ library(mvtnorm)
 ncsurrog<-function(m,corpres,numsurrog){
   N<-dim(m)[1]
   n<-dim(m)[2]
+  
+  #Some error checking
+  if (corpres!="kendall" && corpres!="spearman"){
+    stop("Error in ncsurrog: corpres must be 'kendall' or 'spearman'")
+  }
+  if (corpres=="kendall" && 
+      any(apply(X=m,MARGIN=2,FUN=function(x){length(x)-length(unique(x))})>0))
+  {
+    stop("Error in ncsurrog: ties in time series are not allowed if corpres is kendall")
+  }
   
   if (corpres=="kendall"){
     #get the kendall correlation matrix of the data
@@ -46,12 +61,11 @@ ncsurrog<-function(m,corpres,numsurrog){
     #scor
     ncov<-iRho(normalCopula(0,2),scor)
   }
-  if (corpres!="kendall" && corpres!="spearman"){
-    stop("Error in ncsurrog: corpres must be 'kendall' or 'spearman'")
-  }
   
-  if((any(eigen(ncov)$value<0))==T){
-    warning("Error in ncsurrog: ncov is not positive semidefinite",call.=T,immediate.=T)
+  #more error checking
+  if(!is.positive.semi.definite(ncov))
+  {
+    stop("Error in ncsurrog: ncov is not positive semidefinite")
   }
   
   #generate a bunch of mv normals in the shape of the final

### 2019-02-11 Paper.Rmd
after ncsurrog modification

@@
 We used our four multivariate data sets : aphid-count data, aphid-first flight data, plankton data and methane-flux data, to compute skewness of spatially averaged data.
-We carried out the analysis mentioned in the concepts above, with aphid abundance and aphid first flight date data for all 20 aphid species and plankton abundance data for all 22 plankton species.
+We carried out the analysis mentioned in the concepts above for spearman preserving surrogates, with aphid abundance and aphid first flight date data for all 20 aphid species and plankton abundance data for all 22 plankton species.
 In each study, we got $>5\%$ positive results which gave us confidence in validating significantly different skewness for asymmetric copula than Normal.
@@
 Finally, for methane-flux data, the averaged skewness was found significantly lower in comparison to its Normal surrogates (Figure \ref{fig_skewness_spearman} D).
-We found the same result with all data sets when we repeated our simulation with Kendall correlation preserving surrogates (see Fig.
+<!--We found the same result with all data sets when we repeated our simulation with Kendall correlation preserving surrogates (see Fig.
 \ref{SM-fig_skewness_kendall}).
+Now we did only with spearman preserving to handle ties and +ve semidefinite case-->
 
@@
 We tested spatial Taylor's law with aphid count data (for all 20 species) and found that the mean curvature of the regression line from `log`(variance) vs.
-`log`(mean) of population abundance data was significantly different from $10000$ Normal surrogates (Spearman-correlation and marginal distribution preserving) for $25\%$ of total species considered.
+`log`(mean) of population abundance data was significantly different from $10000$ Normal surrogates (Spearman-correlation and marginal distribution preserving) for $20\%$ of total species considered.
 For example, the "Corn leaf aphid" (having significantly upper tail dependence from model selection and nonparametric analysis) showed mean curvature of regression line close to zero, whereas `r 10000 -(tlsp6$p_left_mncurv*10000)` of the $10000$ Normal surrogates showed higher curvature, indicating here Normal copula didn't hold linear relationship between `log`(spatial variance) and `log`(spatial mean) (see Fig \ref{fig_sTL_mncurv_sp6}).

### 2019-02-11 SupportingInformation.Rmd
after ncsurrog modification

@@
 
-# For sp =1, sp=4 and sp=6 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite : you can see those msg by deleting results="hide" and warning=F option in r chunk
+# For sp =1, sp=4 and sp=6 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite
+# also it can't handle ties for "kendall" and code crashes
+# you can see those msg by deleting results="hide" and warning=F option in r chunk
 
@@
 
-skw_aphid_count_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
-colnames(skw_aphid_count_kendall)<-c("realskw","p_left","p_right")
-rownames(skw_aphid_count_kendall)<-paste("sp",splist,sep="")
-pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_kendall.pdf",height=25,width=25)
-op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
-cat("Now for kendall preserving surrogs : aphid count","\n")
-for(i in splist){
-  cat("sp=",i,"\n")
-  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
-  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
-                        numsurrog=numsurrog,
-                        ploton=T,
-                        corpres="kendall") 
-  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
-  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
-  skw_aphid_count_kendall$realskw[i]<-ans$realskw
-  skw_aphid_count_kendall$p_left[i]<-p_left
-  skw_aphid_count_kendall$p_right[i]<-p_right
-  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
-}
-par(op)
-dev.off()
-saveRDS(skw_aphid_count_kendall,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_kendall.RDS")
+#skw_aphid_count_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+#colnames(skw_aphid_count_kendall)<-c("realskw","p_left","p_right")
+#rownames(skw_aphid_count_kendall)<-paste("sp",splist,sep="")
+#pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_kendall.pdf",height=25,width=25)
+#op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
+#cat("Now for kendall preserving surrogs : aphid count","\n")
+#for(i in splist){
+#  cat("sp=",i,"\n")
+#  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
+#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
+#                        numsurrog=numsurrog,
+#                        ploton=T,
+#                        corpres="kendall") 
+#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+# p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+#  skw_aphid_count_kendall$realskw[i]<-ans$realskw
+#  skw_aphid_count_kendall$p_left[i]<-p_left
+#  skw_aphid_count_kendall$p_right[i]<-p_right
+#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+#}
+#par(op)
+#dev.off()
+#saveRDS(skw_aphid_count_kendall,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_kendall.RDS")
 ```
@@
 
-# For sp =12 and sp=16 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite : you can see those msg by deleting results="hide" and warning=F option in r chunk
+# For sp =12 and sp=16 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite,
+# also it can't handle ties for "kendall" and code crashes
+# you can see those msg by deleting results="hide" and warning=F option in r chunk
 
@@
 
-skw_aphid_ff_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
-colnames(skw_aphid_ff_kendall)<-c("realskw","p_left","p_right")
-rownames(skw_aphid_ff_kendall)<-paste("sp",splist,sep="")
-pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_kendall.pdf",height=25,width=25)
-op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
-cat("Now for kendall preserving surrogs : aphid ff","\n")
-for(i in splist){
-  cat("sp=",i,"\n")
-  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
-  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
-                        numsurrog=numsurrog,
-                        ploton=T,
-                        corpres="kendall") 
-  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
-  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
-  skw_aphid_ff_kendall$realskw[i]<-ans$realskw
-  skw_aphid_ff_kendall$p_left[i]<-p_left
-  skw_aphid_ff_kendall$p_right[i]<-p_right
-  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
-}
-par(op)
-dev.off()
-saveRDS(skw_aphid_ff_kendall,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_kendall.RDS")
+#skw_aphid_ff_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
+#colnames(skw_aphid_ff_kendall)<-c("realskw","p_left","p_right")
+#rownames(skw_aphid_ff_kendall)<-paste("sp",splist,sep="")
+#pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_kendall.pdf",height=25,width=25)
+#op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
+#cat("Now for kendall preserving surrogs : aphid ff","\n")
+#for(i in splist){
+#  cat("sp=",i,"\n")
+#  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
+#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
+#                        numsurrog=numsurrog,
+#                        ploton=T,
+#                        corpres="kendall") 
+#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+#  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+#  skw_aphid_ff_kendall$realskw[i]<-ans$realskw
+#  skw_aphid_ff_kendall$p_left[i]<-p_left
+#  skw_aphid_ff_kendall$p_right[i]<-p_right
+#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+#}
+#par(op)
+#dev.off()
+#saveRDS(skw_aphid_ff_kendall,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_kendall.RDS")
 ```
@@
 
-# There is no error msg or warning coming from ncsurrog : ncov semidefinite matrix for any species with spearman
-# or kendall preserving surrogates.
+# There is no error msg or warning coming from ncsurrog : ncov semidefinite matrix for any species with spearman or kendall preserving surrogates.
+# But there may be error handling ties with "kendall"
+# So,for safety, now we will only analyze with "spearman"
 
@@
 rownames(skw_plankton_spearman)<-paste("sp",splist,sep="")
-skw_plankton_kendall<-skw_plankton_spearman
+#skw_plankton_kendall<-skw_plankton_spearman
 pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_spearman.pdf",height=30,width=30)
@@
 
-pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=30,width=30)
-op<-par(mfrow=c(4,6),mar=c(5,5,1,1),mgp=c(3,1,0))
+#pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=30,width=30)
+#op<-par(mfrow=c(4,6),mar=c(5,5,1,1),mgp=c(3,1,0))
 #cat("Now for kendall preserving surrogs : pns","\n")
-for(i in splist){
+#for(i in splist){
   #cat("sp=",i,"\n")
-  loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
-  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
-                        numsurrog=numsurrog,
-                        ploton=T,
-                        corpres="kendall") 
-  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
-  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
-  skw_plankton_kendall$realskw[i]<-ans$realskw
-  skw_plankton_kendall$p_left[i]<-p_left
-  skw_plankton_kendall$p_right[i]<-p_right
-  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
-}
-par(op)
-dev.off()
-saveRDS(skw_plankton_kendall,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_kendall.RDS")
+ # loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
+#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
+#                        numsurrog=numsurrog,
+#                        ploton=T,
+#                        corpres="kendall") 
+#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
+#  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
+#  skw_plankton_kendall$realskw[i]<-ans$realskw
+#  skw_plankton_kendall$p_left[i]<-p_left
+#  skw_plankton_kendall$p_right[i]<-p_right
+#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
+#}
+#par(op)
+#dev.off()
+#saveRDS(skw_plankton_kendall,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_kendall.RDS")
 ```
@@
 ```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+
+# Only done for "Spearman" method
 
@@
 
-pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_kendall.pdf",height=6,width=8)
-op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer1k<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
-                          corpres="kendall") 
-p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
-mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
-par(op)
-dev.off()
+#pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_kendall.pdf",height=6,width=8)
+#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
+#answer1k<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
+#                          corpres="kendall") 
+#p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
+#mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
+#par(op)
+#dev.off()
 
@@
 
+# Only done for "Spearman" method
+
 source("good_loclist.R")
@@
 
-pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_kendall.pdf",height=6,width=8)
-op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer2k<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
-                          corpres="kendall") 
-p_right<-sum(answer2k$surrogskw > answer2k$realskw)/numsurrog
-mtext(paste0("p = ",p_right),side=3,line=-0.5,cex = 2)
-par(op)
-dev.off()
+#pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_kendall.pdf",height=6,width=8)
+#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
+#answer2k<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
+#                          corpres="kendall") 
+#p_right<-sum(answer2k$surrogskw > answer2k$realskw)/numsurrog
+#mtext(paste0("p = ",p_right),side=3,line=-0.5,cex = 2)
+#par(op)
+#dev.off()
 
@@
 
-pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_kendall.pdf",height=6,width=8)
-op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer3k<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
-                          corpres="kendall") 
-p_left<-sum(answer3k$surrogskw < answer3k$realskw)/numsurrog
-mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
-par(op)
-dev.off()
+#pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_kendall.pdf",height=6,width=8)
+#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
+#answer3k<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
+#                          corpres="kendall") 
+#p_left<-sum(answer3k$surrogskw < answer3k$realskw)/numsurrog
+#mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
+#par(op)
+#dev.off()
 
@@
 
-pdf("./Results/skewness_results/skewness_methane/skewness_methane_kendall.pdf",height=6,width=8)
-op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer1k<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
-                           corpres="kendall") 
-p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
-mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
-par(op)
-dev.off()
+#pdf("./Results/skewness_results/skewness_methane/skewness_methane_kendall.pdf",height=6,width=8)
+#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
+#answer1k<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
+#                           corpres="kendall") 
+#p_left<-sum(answer1k$surrogskw < answer1k$realskw)/numsurrog
+#mtext(paste0("p = ",p_left),side=3,line=-0.5,cex = 2)
+#par(op)
+#dev.off()
 
@@
 ```
-
-\begin{figure}[!h]
-\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
-\vspace{-0.1 cm}
-\begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_kendall.pdf}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_kendall.pdf}\\
-\end{center}
-\vspace{0.2 cm}
-\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
-\vspace{-0.1 cm}
-\begin{center}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_kendall.pdf}
-\includegraphics[width=8 cm]{./Results/skewness_results/skewness_methane/skewness_methane_kendall.pdf}
-\caption{Skewness of spatially averaged $(A)$ Green-spruce aphid's abundance, $(B)$ Leaf-plum curling aphid's firstflight date, $(C)$ Ceratium furca's abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Kendall correlation).
-\label{fig_skewness_kendall}}
-\end{center}
-\end{figure}
 

### 2019-02-12 Paper.Rmd
edits to Q3 sections

@@
 We again repeated the analysis with $r=1.3$. Discrepancies between noise and population
-copula structure in this case were often glaring - apparently noise of standard deviation 1
+copula structure in this case were often glaring. Apparently noise of standard deviation 1
 interacted very strongly with model nonlinearities when the model was in its overcompensatory 
@@
 
-We addressed Q3 using both models and data. The initial 
+We addressed Q3 using both models and data. An initial 
 idea was presented in the Introduction that
-the distribution (through time) of a spatially averaged quantity should depend on the 
+the distribution (through time) of a spatially averaged quantity should be influenced by the 
 nature of the dependence between the local quantities being averaged, including their
@@
 This should tend to increase the skewness of the distribution through time of the
-spatial average. Using similar reasoning, left-tail dependence between local variables
-$E_i$ and $E_j$ will tend to decreases skewness of the spatial mean. Skewness of the 
+spatial average. Similarly, left-tail dependence between local variables
+will tend to decreases skewness of the spatial mean. Skewness of the 
 spatial mean corresponds to the "spikiness" of the spatial-mean time series. 
@@
 as the original data. 
-They also had very similar Kendall or Spearman correlations (the algorithm had two versions,
-one for preserving each correlation coefficient) between pairs of sampling locations
+They also had very similar Spearman correlations between pairs of sampling locations
 as the original data. But the surrogates had multivariate normal copula structure. 
 Our comparisons therefore tested the null hypothesis that the skewness of the spatial 
-mean took values on the empirical no different from what one would expect if the copula
+mean took values on the empirical data no different from what one would expect if the copula
 structure of the data were multivariate normal, but the data were otherwise statistically
-unchanged. Significant differences indicate that copula structure contributes to the skewness
-of the spatial mean, i.e., to its "spikiness."
-
-<!--***DAN: stopped here 2019 02 08 1106am central-->
-
-This section is dedicated to analysis of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}.
-We wanted to examine whether the asymmetric tail dependence had any **(i)** effects on skewness of spatially averaged time series data.
-We predicted that because of overall distinct tail dependence of the multivariate copula structure, the skewness of the spatially-averaged time-series data, should change significantly from the individual skewness of time-series data.
-We tested null hypothesis $(H_{0})$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula.
-We generated $10000$ number of surrogate normal copula from a given multivariate copula keeping the marginal distribution and the correlation (Spearman or Kendall) same and calculated the skewness of spatially averaged surrogate time series data, as we did with the actual multivariate data set in the next section (For code see [skewness_multivar_dataset.R](https://github.com/sghosh89/BIVAN/blob/master/skewness_multivar_dataset.R)).
-We made a histogram of surrogate skewness and found if actual skewness was within the $95\%$ confidence interval or not.
-If the $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness than Normal copula).
+unchanged. Significant differences indicate that copula structure contributed 
+to the skewness of the spatial mean, i.e., to its "spikiness."
+
+For green spruce aphid abundance data, *Ceratium furca* abundance data, and methane-flux
+data, because these datasets exhibited lower-tail dependence 
+(Table \ref{tab_multivar_summary}), we compared empirical and surrogate skewness values
+via a one-tailed test in the left tail: the $p$-value result was simply the fraction
+of surrogate skewnesses less than the skewness computed from the empirical data. For
+leaf-curling plum aphid first flight data, because this dataset exhibited upper-tail 
+dependence (Table \ref{tab_multivar_summary}), we did the analogous one-tail test
+in the right tail.
+
+<!--Reminder that Table 2 says:
+Green spruce - lower tail
+Leaf-curling - upper tail
+Cerratium - lower tail
+methane - marginal lower tail-->
+
+<!--***DAN: stopped here in this section, 2019 02 12, 633am-->
 
@@
 In the following section we will show application of these ideas with real data sets.
+
 <!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion-----------
@@
 -->
+
+## Results for Q3: Tail dependence influences skewness of spatial average, Taylor's law, and extinction risk
 
@@
 \includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
-\caption{Skewness of spatially averaged $(A)$ green-spruce aphid abundance, $(B)$ leaf-curling plum aphid first flight date, $(C)$ Ceratium furca abundance and $(D)$ methane-flux data, were indicated by red dot in each plot and were compared with respect to skewness of spatially averaged 10000 Normal surrogates preserving Spearman correlation).
-\label{fig_skewness_spearman}}
+\caption{Skewness of spatially averaged green spruce aphid abundance $(A)$, leaf-curling plum 
+aphid first flight dates $(B)$, \emph{Ceratium furca} abundance $(C)$, and methane-flux
+$(D)$ time series compared to an multivariate 
+N-copula null hypothesis. Red dots are empirical 
+skewnesses; see text 
+and Appendix \ref{SM-Surrog_multi} for details of the null hypothesis. Results show a 
+tendency for skewness of the spatial average to be affected as hypothesized 
+by tail dependence. \label{fig_skewness_spearman}}
 \end{center}
 \end{figure}
-
-## Results for \hyperref[Intro]{\textbf{(Q3)}}: Change of skewness
+<!--***DAN: Shya, please replace the x axis label by "Skewness" and the y-axis label
+by "Frequency"-->
+
+<!--***DAN: stopped here in this section, 2019 02 12, 633am-->
 

### 2019-02-12 SupportingInformation.Rmd
edits to Q3 sections

@@
 
-```{r skewness_aphidcount_allsp_plot, echo=F, results="hide", warning=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+```{r skewness_aphidcount_allsp_plot, echo=F, results="hide", cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
 
@@
 
-```{r skewness_aphidff_allsp_plot, echo=F, results="hide", warning=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+```{r skewness_aphidff_allsp_plot, echo=F, results="hide", cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
 
@@
 element of $x_i(t)$, $t=1,\ldots,T$.
-Call the result $b_i(t)$. This is the surrogate dataset.
+Call the result $b_i(t)$. This is a surrogate dataset.
 
 \noindent Our code for this algorithm is `ncsurrog` in the `BIVAN` repository 
-(\url{https://github.com/sghosh89/BIVAN}). Because, for each $i$, the final time series 
+(\url{https://github.com/sghosh89/BIVAN}).
+The code for the Kendall version of the
+algorithm throws an error if there are ties in any of the time series. This is
+because the Kendall correlation in the very commonly used `cor` function in R 
+is the Kendall tau-b coefficient in the event of ties, and index which 
+differs in its definition from the commonly used tau-a that works when there are
+no ties; and `tau` and `iTau` are intended for continuous copulas, which cannot 
+produce data with ties. Our code also throws an error in the event that
+the matrix generated in the second step is not positive semidefinite. Because
+ties are common in our data, we only used the Spearman case in analyses using `ncsurrog`,
+though, for expansibility, the `ncsurrog` code itself is written to handle 
+the Kendall case as well when there are no ties.
+
+Because, for each $i$, the final time series 
 $b_i(t)$ is a permuted version of $x_i(t)$ (and with `NA`s for the same $t$), 
@@
 These discrepancies should be minor if the 
-numbers of NAs are not too large and $T$ is not too small. The Kendall version of the
-algorithm throws an error if there are ties in any of the time series. The algorithm
-also throws an error in the event that
-the matrix generated in the second step is not positive semidefinite.
+numbers of NAs are not too large and $T$ is not too small.
 

### 2019-02-12 SupportingInformation.Rmd
skewness plot axis labels fixed

@@
 op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer1s<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
+answer1s<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=F,
                           corpres="spearman") 
+hist(answer1s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
+#abline(v=realskw,col="red")
+points(x=answer1s$realskw,y=0,col="black",pch=15,cex=2)
 
@@
 op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer2s<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
+answer2s<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=F,
                           corpres="spearman")
+hist(answer2s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
+#abline(v=realskw,col="red")
+points(x=answer2s$realskw,y=0,col="black",pch=15,cex=2)
+
 p_right<-sum(answer2s$surrogskw > answer2s$realskw)/numsurrog
@@
 op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer3s<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
+answer3s<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=F,
                           corpres="spearman") 
+hist(answer3s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
+#abline(v=realskw,col="red")
+points(x=answer3s$realskw,y=0,col="black",pch=15,cex=2)
+
 p_left<-sum(answer3s$surrogskw < answer3s$realskw)/numsurrog
@@
 op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer1s<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
+answer1s<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=F,
                            corpres="spearman") 
+hist(answer1s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
+#abline(v=realskw,col="red")
+points(x=answer1s$realskw,y=0,col="black",pch=15,cex=2)
+
 p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog

### 2019-02-12 Paper.Rmd
skewness plot axis labels fixed
minor change

@@
 $(D)$ time series compared to an multivariate 
-N-copula null hypothesis. Red dots are empirical 
+N-copula null hypothesis. Black dots are empirical 
 skewnesses; see text 
@@
 <!--***DAN: Shya, please replace the x axis label by "Skewness" and the y-axis label
-by "Frequency"-->
+by "Frequency"
+***Shya: Done.-->
 
@@
 For example, Figure \ref{fig_skewness_spearman} represents the results for each data set with Spearman correlation preserving surrogates.
-Figure \ref{fig_skewness_spearman}(A) showed that averaged skewness (indicated by red dot) was much lower than expected if the green-spruce aphid count data came from a normal copula.
+Figure \ref{fig_skewness_spearman}(A) showed that averaged skewness (indicated by black dot) was much lower than expected if the green-spruce aphid count data came from a normal copula.
 In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data.

### 2019-02-12 SupportingInformation.Rmd
Table S1 made with italic font

@@
 library(tibble)
+library(kableExtra)
+library(dplyr)
 aphid_info <- tibble(c1=c("Apple grass aphid","Bird cherry oat aphid","Black bean aphid","Blackberry cereal aphid","Blackcurrant sowthistle aphid",
@@
 )
+
 knitr::kable(aphid_info, "latex", booktabs = T, linesep = "",
       caption = "Species names of 20 aphids for which data were available \\label{tab_aphid_info}",
-      col.names = c("Common name", "Latin binomial"))
+      col.names = c("", ""))%>%
+     column_spec(column = 2,italic=T)%>%add_header_above(header=c("Common name" = 1, "Latin binomial" = 1),align = "l",line=F)
 ```

### 2019-02-12 SupportingInformation.Rmd
Table S1 style modified

@@
       caption = "Species names of 20 aphids for which data were available \\label{tab_aphid_info}",
-      col.names = c("", ""))%>%
-     column_spec(column = 2,italic=T)%>%add_header_above(header=c("Common name" = 1, "Latin binomial" = 1),align = "l",line=F)
+      col.names = NULL)%>%
+     column_spec(column = 2,italic=T)%>%add_header_above(header=c("Common name" = 1, "Latin binomial" = 1))
 ```

### 2019-02-12 Paper.Rmd
finished editing Q3 for now (expect more changes later) several things for Shya to do, see comments in all docs

--- 2019-02-12 Paper.Rmd
@@
 
-### Need to think about mechanisms for biogeochemical, evolutionary, and community data
+<!--***DAN: Maybe need to think about mechanisms for biogeochemical, evolutionary, and community data-->
 
@@
 leaf-curling plum aphid first flight data, because this dataset exhibited upper-tail 
-dependence (Table \ref{tab_multivar_summary}), we did the analogous one-tail test
+dependence (Table \ref{tab_multivar_summary}), we did the analogous one-tailed test
 in the right tail.
 
-<!--Reminder that Table 2 says:
-Green spruce - lower tail
-Leaf-curling - upper tail
-Cerratium - lower tail
-methane - marginal lower tail-->
-
-<!--***DAN: stopped here in this section, 2019 02 12, 633am-->
-
-We also showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL_simulation}).
-For example, our model simulation showed lower tail dependent Clayton copula had much higher slope (spatial-Taylor's law) than symmetric Normal copula, which in turn had higher slope than upper tail dependent Survival Clayton family (Fig.
-\ref{SM-fig_spTL}(I)).
-In the following section we will show application of these ideas with real data sets.
+Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we tested whether
+the copula structure of dependence between population variables influences the
+spatial version of Taylor's
+law. Taylor's law is an extremely commonly observed and widely studied 
+[@Taylor1961, @Taylor1988, @Cohen2013, @Xu2016] empirical pattern that relates the 
+variances of groups, $g$, of population measurements to the means of these 
+groups via a power law, $v_g = a \times m_g^b$, or equivalently via a log-log linear 
+relationship, $\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
+the Taylor's law *exponent* or 
+*slope*. There are several versions of Taylor's law. For spatial Taylor's law,
+given population density or abundance data $x_i(t)$ measured in locations
+$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
+all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
+So the means and variances in the Tayor's law formula are computed across 
+space. Taylor's law has been 
+verified in data from hundreds of taxa and has been applied in numerous fields
+including fisheries management, estimation of species persistence times, and
+agriculture [@Cohen2015,@Reuman2017].
+
+We tested whether the linearity of the spatial Taylor's law relationship was partly
+due to the copula structure of the dependences between the populations sampled. 
+Using corn leaf aphid abundance data as an example, but also
+applying the same analysis to the abundance data we had for 19 other aphid species
+(see *Data*), we calculated the curvature (details in Appendix \ref{SM-TL_curvature}) 
+of the empirical Taylor's law relationship; Taylor's law hypothesizes that this 
+curvature should be close to 0. We then calculated the same curvature quantity for
+10000 surrogate datasets, using the same surrogate procedure described above in 
+this section, and compared the empirical Taylor's law curvature to the 
+distribution of surrogate curvatures using a two-tailed test. If the empirical
+curvature value was below the $5^{\text{th}}$ quantile of the surrogate distribution, 
+then copula structure contributes to the linearity of spatial Taylor's law 
+because randomizing copula structure while retaining other statistical properties
+of the data cause the Taylor's law relationship to become significantly more 
+curved.
+
+<!--***DAN: Shya,  want to add the extinction risk stuff back in, let's discuss-->
 
@@
 <!--***DAN: Shya, please replace the x axis label by "Skewness" and the y-axis label
-by "Frequency"-->
-
-<!--***DAN: stopped here in this section, 2019 02 12, 633am-->
-
-We used our four multivariate data sets : aphid-count data, aphid-first flight data, plankton data and methane-flux data, to compute skewness of spatially averaged data.
-We carried out the analysis mentioned in the concepts above for spearman preserving surrogates, with aphid abundance and aphid first flight date data for all 20 aphid species and plankton abundance data for all 22 plankton species.
-In each study, we got $>5\%$ positive results which gave us confidence in validating significantly different skewness for asymmetric copula than Normal.
-For example, Figure \ref{fig_skewness_spearman} represents the results for each data set with Spearman correlation preserving surrogates.
-Figure \ref{fig_skewness_spearman}(A) showed that averaged skewness (indicated by red dot) was much lower than expected if the green-spruce aphid count data came from a normal copula.
-In that example, $p$ value was denoted by the fraction of surrogate-skewness which were lower than the averaged skewness of the given data.
-As $p<0.05$, we rejected $H_{0}$ in support of significantly stronger lower tail dependence which was also consistent with the nonparametric analysis and model fitting results (see first column of Table \ref{tab_multivar_summary}).
-On the other hand, Figure \ref{fig_skewness_spearman}(B) demonstrated that averaged skewness was significantly higher than expected if the leaf-plum curling aphid first flight data came from a normal copula, which was also confirmed by upper-tail dependence results from model selection and nonparametric results (see second column of Table \ref{tab_multivar_summary}).
-Figure \ref{fig_skewness_spearman}(C) showed that averaged skewness was lower than expected if the plankton data came from a normal copula.
-Our previous results on nonparametric analysis and model fitting also supported lower tail dependence (see third column of Table \ref{tab_multivar_summary}).
-But, as $p>0.05$, we could not reject $H_{0}$ in that case.
-Finally, for methane-flux data, the averaged skewness was found significantly lower in comparison to its Normal surrogates (Figure \ref{fig_skewness_spearman} D).
-<!--We found the same result with all data sets when we repeated our simulation with Kendall correlation preserving surrogates (see Fig.
-\ref{SM-fig_skewness_kendall}).
-Now we did only with spearman preserving to handle ties and +ve semidefinite case-->
-
-## Results for \hyperref[Intro]{\textbf{(Q3)}}: Change of curvature of spatial Taylor's law regression line
+by "Frequency" in the above plot-->
+
+Skewness results were generally consistent with hypotheses. For the datasets that
+exhibited strong lower-tail (respectively, upper-tail) 
+dependence in earlier analyses (namely, the green spruce 
+aphid abundance data for lower-tail, and the leaf-curling plum aphid first flight data
+for upper-tail, Table 
+\ref{tab_multivar_summary}), the skewness, through time, of the spatial average of the 
+empirical data was less than (respectively, greater than) a significant fraction (i.e.,
+more than $95\%$) of the
+surrogate skewness values 
+(Fig. \ref{fig_skewness_spearman}A,B). For the datasets that demonstrated a 
+more moderate lower-tail dependence in earlier analyses (the 
+\emph{Ceratium furca} abundance and the methane data), skewness of the spatial
+average showed a non-significant tendency toward being less than surrogate skewnesses
+(Fig. \ref{fig_skewness_spearman}C), or was less than a marginally significant 
+fraction of surrogate skewnesses (Fig. \ref{fig_skewness_spearman}D).
 
@@
 
-Spatial Taylor's law, the observed linear relationship between `log`(variance) and `log`(mean) of population densities of a species from different locations, is a well-known pattern found in ecology and related fields [@Taylor1961, @Taylor1988, @Cohen2013, @Doring2015, @Kuo2016, @Xu2016, @Reuman2017].
-We already showed through simulation (Appendix \ref{SM-Consequence_TL_simulation}) that for asymmetric copula this linear relationship somehow deviates compared to Normal copula for a distribution with gamma marginals.
-We tested spatial Taylor's law with aphid count data (for all 20 species) and found that the mean curvature of the regression line from `log`(variance) vs.
-`log`(mean) of population abundance data was significantly different from $10000$ Normal surrogates (Spearman-correlation and marginal distribution preserving) for $20\%$ of total species considered.
-For example, the "Corn leaf aphid" (having significantly upper tail dependence from model selection and nonparametric analysis) showed mean curvature of regression line close to zero, whereas `r 10000 -(tlsp6$p_left_mncurv*10000)` of the $10000$ Normal surrogates showed higher curvature, indicating here Normal copula didn't hold linear relationship between `log`(spatial variance) and `log`(spatial mean) (see Fig \ref{fig_sTL_mncurv_sp6}).
+Results also suggested that copula structure contributes to the linearity of 
+spatial Taylor's law. For instance, the spatial Taylor's law relationship 
+for corn leaf aphid abundance data was significantly less curved than for 
+surrogate datasets (Fig. \ref{fig_sTL_mncurv_sp6}). Three of $20$ aphid species
+had cuvatures less than at least $95\%$ of surrogate datasets ($p=X$, $Y$, $Z$ for 
+the $A$, $B$ and $C$ species, respectively). 
+<!--***DAN: Shya, please fill in X, Y, Z, A, B, C with autolinks)-->
+Three such significant
+results of 20 species could have occurred by chance due to 
+multiple-testing error. However, for each of the 20 species, we computed the fraction
+of surrogate curvatures less than the curvature that came from real data, and we
+combined these $p$-values using Fisher's combined probability test to get an
+aggregate $p$-value of $0.019$ 
+<!--***DAN: Shya, see the Wikipedia article on Fisher's method to learn about it. Then 
+you need to put this analysis into the code suite and have the resulting p value
+typed in above be an autolink instead. Here is what I did: 
+1) I copied by hand all the pleft values on the pdf you sent me into a vector pleft. Instead you should create such a vector in your code from your test output, not typing in the results.
+2) I used the formula stat<--2*sum(log(pleft)) to get the Fisher method statistic
+3) I compared to a chi-squared distribution with 40 degress of freedom using 1-pchisq(stat,40) to get my p-value
+Please let me know if any of this is unclear.-->
+for the test of the null hypothesis that 
+copula structure did not contribute to linearity of Taylor's law for any of the 
+species. This result is usefully suggestive and can guide future work, but still does not
+provide certainty, because the Fisher method provides $p$-values that are too 
+small if its assumption of independent tests is violated. Though our 20 species
+have different host plants so may be independent for the purposes of our analyses, 
+it would be difficult to fully assess independence.
+<!--***DAN: Shya, can you remind me why the same analyses were not done for plankton,
+if there is a good reason? Or if you did it, what were the results?--> 
+We provide simulation results in Appendix \ref{SM-Consequence_TL_simulation}
+that robustly illustrate, at least in the model setup we used, 
+that copula structure influences spatial Taylor's law.
 
@@
 \includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
-\caption{log(spatial variance) and log(spatial mean) for $(A)$ corn-leaf aphid (\textit{Rhopalosiphum maidis}) abundance and for $(B)$ one of the $9961$ Normal surrogates (having same marginal distribution and Spearman correlation of actual data).
-$(C)$ Histogram of mean curvature of $10000$ normal surrogates showed that for corn-leaf aphid abundance data (with asymmetric tail dependence), Normal surrogates didn't follow linear relationship between log(spatial variance) and log(spatial mean).
-\label{fig_sTL_mncurv_sp6}}
+\caption{Mean curvature of the spatial Taylor's law relationship for 
+corn leaf aphid abundance
+data (A) was significantly less than for surrogates (B, C). The Taylor's law 
+relationship for one example surrogate (B) shows shows greater curvature than
+for data (A), and the significance of the difference was demonstrated
+by the fact that only $39$ of $10000$ surrogates had curvature less than that
+of the data (C). See Appendix \ref{SM-TL_curvature} for details on 
+how curvature was calculated.\label{fig_sTL_mncurv_sp6}}
 \end{center}
 \end{figure}
+<!--***DAN: Shya, pls make the x axis label on C read simply "Mean curvature"-->
+
+<!--***DAN: I want to put back in some extinction risk results. Let's talk about this, please, Shyamolina-->
 

### 2019-02-12 SupportingInformation.Rmd
finished editing Q3 for now (expect more changes later) several things for Shya to do, see comments in all docs

--- 2019-02-12 SupportingInformation.Rmd
@@
 ```
+<!--***DAN: These names differ from those used in Emma's 2016 Glob Change Biol paper.
+Need to check with Chris Reid. Have to get this right.-->
 
@@
 
-<!--\newpage-->
+<!--
 ```{r tab_temp_aphid_p_CvM, echo=F, results='markup'}
@@
 ```
-
-<!--\newpage-->
+-->
+
+<!--
 ```{r tab_temp_aphid_p_KS, echo=F, results='markup'}
@@
 ```
-
-<!--\newpage-->
+-->
+
+<!--
 ```{r tab_temp_aphid_LTdep_AICw, echo=F, results='markup'}
@@
 ```
-
-<!--\newpage-->
+-->
+
+<!--
 ```{r tab_temp_aphid_UTdep_AICw, echo=F, results='markup'}
@@
 ```
-
-<!--\newpage-->
+-->
+
+<!--
 ```{r tab_temp_aphid_LTmUT, echo=F, results='markup'}
@@
 ```
-
-<!--# Cause of non-Normal copula structure : the Moran effect and copula structure in populations \label{Cause_Moran}-->
-
-<!--
-## Asymmetric sensitivity and copula structure in populations
-           ----moved to Paper.Rmd----
 -->
@@
 op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer1s<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=F,
+answer1s<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=T,
                           corpres="spearman")
-hist(answer1s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
-#abline(v=realskw,col="red")
-points(x=answer1s$realskw,y=0,col="black",pch=15,cex=2)
 
@@
 op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer2s<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=F,
+answer2s<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=T,
                           corpres="spearman") 
-hist(answer2s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
-#abline(v=realskw,col="red")
-points(x=answer2s$realskw,y=0,col="black",pch=15,cex=2)
-
 p_right<-sum(answer2s$surrogskw > answer2s$realskw)/numsurrog
@@
 op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer3s<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=F,
+answer3s<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=T,
                           corpres="spearman") 
-hist(answer3s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
-#abline(v=realskw,col="red")
-points(x=answer3s$realskw,y=0,col="black",pch=15,cex=2)
-
 p_left<-sum(answer3s$surrogskw < answer3s$realskw)/numsurrog
@@
 op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
-answer1s<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=F,
+answer1s<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=T,
                            corpres="spearman") 
-hist(answer1s$surrogskw,breaks=100,main="",xlab="Skewness",ylab="Frequency",cex.lab=2,cex.axis=2,col="grey",border=F)
-#abline(v=realskw,col="red")
-points(x=answer1s$realskw,y=0,col="black",pch=15,cex=2)
-
 p_left<-sum(answer1s$surrogskw < answer1s$realskw)/numsurrog
@@
 
-# Consequence of non-Normal copula structure : Effect on Taylor's law \label{Consequence_TL_simulation}
-
+# Calculating curvature for Taylor's law \label{TL_curvature}
+
+<!--***DAN: Shya, I need a reminder how we did this, can you please tell me where the
+code is, then I can write this section. Feel free to take a stab at it yourself, too,
+though.-->
+
+# Effects of non-N copula structure on Taylor's law \label{Consequence_TL_simulation}
+
+<!--***DAN: Dan has not edited this section yet at all-->
 Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula.

### 2019-02-12 Paper.Rmd

@@
 \end{figure}
-<!--***DAN: Shya, please replace the x axis label by "Skewness" and the y-axis label
-by "Frequency" in the above plot-->
 

### 2019-02-12 SupportingInformation.Rmd

@@
 )
+
 knitr::kable(aphid_info, "latex", booktabs = T, linesep = "",
       caption = "Species names of 20 aphids for which data were available \\label{tab_aphid_info}",
-      col.names = c("Common name", "Latin binomial"))
+      col.names = NULL)%>%
+     column_spec(column = 2,italic=T)%>%add_header_above(header=c("Common name" = 1, "Latin binomial" = 1))
 ```

### 2019-02-13 Paper.Rmd
adding response

@@
 \begin{figure}[!h]
-\textbf{ \hspace{6 cm} (A) \hspace{6 cm} (B)} \\
-\vspace{-0.1 cm}
+\textbf{ \hspace{6.6 cm} (A) \hspace{4.6 cm} (B)} \\
+\vspace{-0.4 cm}
 \begin{center}
-\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
-\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
+\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf}
+\includegraphics[width=5 cm]{./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf}\\
 \end{center}
 \vspace{0.2 cm}
-\textbf{ \hspace{6 cm} (C) \hspace{6 cm} (D)} \\
-\vspace{-0.1 cm}
+\textbf{ \hspace{6.6 cm} (C) \hspace{4.6 cm} (D)} \\
+\vspace{-0.4 cm}
 \begin{center}
-\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
-\includegraphics[width=5.5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
+\includegraphics[width=5 cm]{./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf}
+\includegraphics[width=5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
 \caption{Skewness of spatially averaged green spruce aphid abundance $(A)$, leaf-curling plum 
@@
 tlsp6<-readRDS("./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
+tl_aphid_count<-readRDS("./Results/taylorslaw_results/tl_aphid_count.RDS")
+aphid_info<-readRDS("./Results/aphid_info.RDS")
+pleft<-tl_aphid_count$p_left_mncurv
+stat<--2*sum(log(pleft))
+pchi2<-1-pchisq(stat,40)
+idsp<-which(pleft<0.05)
+pleftgood<-pleft[idsp]
 ```
@@
 surrogate datasets (Fig. \ref{fig_sTL_mncurv_sp6}). Three of $20$ aphid species
-had cuvatures less than at least $95\%$ of surrogate datasets ($p=X$, $Y$, $Z$ for 
-the $A$, $B$ and $C$ species, respectively). 
-<!--***DAN: Shya, please fill in X, Y, Z, A, B, C with autolinks)-->
+had cuvatures less than at least $95\%$ of surrogate datasets ($p=$ `r pleftgood` for 
+the `r aphid_info$c1[idsp]` species, respectively). 
+<!--***DAN: Shya, please fill in X, Y, Z, A, B, C with autolinks)
+***Shya: Done.-->
 Three such significant
@@
 combined these $p$-values using Fisher's combined probability test to get an
-aggregate $p$-value of $0.019$ 
+aggregate $p$-value of `r pchi2`
 <!--***DAN: Shya, see the Wikipedia article on Fisher's method to learn about it. Then 
@@
 3) I compared to a chi-squared distribution with 40 degress of freedom using 1-pchisq(stat,40) to get my p-value
-Please let me know if any of this is unclear.-->
+Please let me know if any of this is unclear.
+***Shya: Done.-->
 for the test of the null hypothesis that 
@@
 <!--***DAN: Shya, can you remind me why the same analyses were not done for plankton,
-if there is a good reason? Or if you did it, what were the results?--> 
+if there is a good reason? Or if you did it, what were the results?
+***Shya : we did it, I have the pdf but I don't have the code where I called to generate the pdf,
+As we could not get any significant results for a single species (see attachment in the mail I sent you)
+we did not show the results.--> 
 We provide simulation results in Appendix \ref{SM-Consequence_TL_simulation}
@@
 for data (A), and the significance of the difference was demonstrated
-by the fact that only $39$ of $10000$ surrogates had curvature less than that
+by the fact that only `r (tlsp6$p_left_mncurv)*10000` of $10000$ surrogates had curvature less than that
 of the data (C). See Appendix \ref{SM-TL_curvature} for details on 
@@
 \end{figure}
-<!--***DAN: Shya, pls make the x axis label on C read simply "Mean curvature"-->
+<!--***DAN: Shya, pls make the x axis label on C read simply "Mean curvature"
+***sHYA: dONE.-->
 

### 2019-02-13 SupportingInformation.Rmd
adding response

@@
 
+saveRDS(aphid_info,"./Results/aphid_info.RDS")
 knitr::kable(aphid_info, "latex", booktabs = T, linesep = "",
@@
 
-pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf",height=6,width=8)
-op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
+pdf("./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10_spearman.pdf",height=5,width=5)
+op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
+#op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
 answer1s<-skewness_testing(ts_matrix=sp_data(sp=10,d_allsp=data_aphid_count),loclist=loclist,numsurrog=numsurrog,ploton=F,
@@
 
-pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf",height=6,width=8)
-op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
+pdf("./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11_spearman.pdf",height=5,width=5)
+op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
 answer2s<-skewness_testing(ts_matrix=sp_data(sp=11,d_allsp=data_aphid_ff),loclist=loclist,numsurrog=numsurrog,ploton=F,
@@
 
-pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf",height=6,width=8)
-op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
+pdf("./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16_spearman.pdf",height=5,width=5)
+op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
 answer3s<-skewness_testing(ts_matrix=sp_data(sp=16,d_allsp=data_plankton_north_sea),loclist=loclist,numsurrog=numsurrog,ploton=F,
@@
 
-pdf("./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf",height=6,width=8)
-op<-par(mar=c(5,5,1.8,0.1), mgp=c(3, 1, 0))
+pdf("./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf",height=5,width=5)
+op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
 answer1s<-skewness_testing(ts_matrix=sp_data(sp=1,d_allsp=methane_data),loclist=loclist,numsurrog=numsurrog,ploton=F,
@@
   ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon=T)
+  if(i==6){
+    saveRDS(ans,"./Results/taylorslaw_results/tl_aphid_count_sp6.RDS") #saving sp=6 results
+  }
   tl_aphid_count$real_sTL_slope[i]<-ans$real_spatialTLslope
@@
 
-```{r spatialTL_on_realdata_aphid_count_sp6, results="hide", echo=F, cache=T, cache.extra=list(seed,data_aphid_count,TL_on_realdata,mtime("good_loclist.R"),mtime("TaylorsLawTools.R"),mtime("mncurvat.R"),mtime("ncsurrog.R"))}
-
-set.seed(seed)
-
+```{r spatialTL_on_realdata_aphid_count_sp6, results="hide", echo=F}
 #Now for sp=6
-sp<-6
+
+ans<-readRDS("./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
+
 pdf("./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf",height=5,width=5)
 op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
-ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=sp,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon = F)
-saveRDS(ans,"./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
+hist(ans$good_surrog_mncurv,breaks=100,main="",xlab="Mean curvature",cex.axis=2,cex.lab=2,cex=2,col="grey",border=F)
+points(x=ans$real_sTL_mncurv,y=0,col="black",pch=15,cex=2)
 mtext(paste0("p= ",ans$p_left_mncurv),side = 3, line=-2, adj=0.5, col="black",cex=2)
@@
 pdf("./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf",height=5,width=10)
-op<-op<-par(mfrow=c(1,2),mar=c(4,5,1,1),mgp=c(3,1,0))
+op<-par(mfrow=c(1,2),mar=c(4,5,1,1),mgp=c(3,1,0))
 plot(ans$realTL_stats$s.points,xlab=expression(log[10]("spatial mean")),
-     ylab=expression(log[10]("spatial variance")),col="black",
-     cex.axis=1.5,cex.lab=1.5)
+     ylab=expression(log[10]("spatial variance")),col="grey",pch=20,cex=1.5,
+     cex.axis=2,cex.lab=2)
 mtext(paste0("mean curvature= ",round(unname(ans$realTL_stats$s.stats[7]),3)),side = 3, line=-2, adj=0, col="black",cex=2)
@@
 plot(ans_s[[id]]$s.points,xlab=expression(log[10]("spatial mean")),
-     ylab=expression(log[10]("spatial variance")),
-     cex.axis=1.5,cex.lab=1.5)
+     ylab=expression(log[10]("spatial variance")),col="grey",pch=20,cex=1.5,
+     cex.axis=2,cex.lab=2)
 mtext(paste0("mean curvature= ",round(ans$good_surrog_mncurv[id],3)),side = 3, line=-2, adj=0, col="black",cex=2)
@@
 code is, then I can write this section. Feel free to take a stab at it yourself, too,
-though.-->
+though.
+***Shya: What do you want to know? how did you calculate mean curvature and 
+slope for made up data or you want to see code with real data?
+
+In case of made up data :
+For each non-N copula family (C,F,SC) you generated surrogates
+having similar correlation (spearman/ kendall, though for 
+consistency we added plots in suppmat with spearman) that of N copula
+[using copsurrognd.R] : this thing is done in r chunk named "taylors_law_make_surrogs".
+
+Then you did spatial-TL calculation for each set of copula family in r chunk "tl_analysis"
+[using TaylorsLawTools.R which calls mncurvat.R code to calculate mean curvature]
+
+Next two r chunks : taylor_plots_kendall and taylor_plots_spearman generates plot with those data.
+
+For spatial TL calculation on real data :
+
+We used Aphid count all sp with spearman only, but to show a plot for example chose sp=6 (corn leaf aphid)
+as it had significant p_left value for mean curvature.
+code is in r chunk named "spatialTL_on_realdata_aphid_count".
+-->
 

### 2019-02-13 Paper.Rmd
part done moving TL material to supp mat, in midst of editing the text describing the sims done for TL

@@
 will tend to decreases skewness of the spatial mean. Skewness of the 
-spatial mean corresponds to the "spikiness" of the spatial-mean time series. 
-The spatial-mean time series may be the quantity of principal importance for 
+spatial mean corresponds to the "spikiness" of the spatial-mean time series or
+its tendency to crash to low levels. 
+The spatial-mean time series and its skewness may be quantities of 
+principal importance for 
 cases such as pest or resource abundance or CO$_2$ or methane flux.
@@
 
-Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we tested whether
-the copula structure of dependence between population variables influences the
-spatial version of Taylor's
-law. Taylor's law is an extremely commonly observed and widely studied 
-[@Taylor1961, @Taylor1988, @Cohen2013, @Xu2016] empirical pattern that relates the 
-variances of groups, $g$, of population measurements to the means of these 
-groups via a power law, $v_g = a \times m_g^b$, or equivalently via a log-log linear 
-relationship, $\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
-the Taylor's law *exponent* or 
-*slope*. There are several versions of Taylor's law. For spatial Taylor's law,
-given population density or abundance data $x_i(t)$ measured in locations
-$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
-all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
-So the means and variances in the Tayor's law formula are computed across 
-space. Taylor's law has been 
-verified in data from hundreds of taxa and has been applied in numerous fields
-including fisheries management, estimation of species persistence times, and
-agriculture [@Cohen2015,@Reuman2017].
-
 We tested whether the linearity of the spatial Taylor's law relationship was partly
@@
 For Ricker model, we got the similar observation for smaller values of $r$, the intrinsic growth rate, but for higher values of $r$, there was a turn-over in extinction risks, and right tail dependent noise copula caused population go extinct faster.
+-->
+
+<!-- 
+This extincton risk material had been commented out in the sup mat:
+
+Extinction risk results are not shown currently, it is introduced as a future idea in Discussion------
+
+# Consequence of non-Normal copula structure : effect on extinction risk \label{Consequence_ext_risk}
+
+In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
+
+\begin{equation}\label{eq.extrisk_model}
+P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
+\end{equation}
+
+$r$ is the liner growth rate and $K$ is the carrying capacity of the environment.
+$\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence.
+We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion).
+We simulated Eq.
+\ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $25$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;
+
+- LC model (r=0) :
+
+    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
+[see first subplot of Fig.
+\ref{fig_risk_local_disp} and Fig.
+\ref{fig_risk_global_disp}]
+
+- Ricker model (r $\neq$ 0):
+
+    * For ricker model when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
+For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
+[see Fig.
+\ref{fig_risk_local_disp} - \ref{fig_risk_global_disp}]
+
+    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]
+
+With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig.
+\ref{fig_risk_numlocs_effect_local_disp} - Fig.
+\ref{fig_risk_numlocs_effect_global_disp}] for both model.
 -->

### 2019-02-13 SupportingInformation.Rmd
part done moving TL material to supp mat, in midst of editing the text describing the sims done for TL

@@
 
-# Calculating curvature for Taylor's law \label{TL_curvature}
-
-<!--***DAN: Shya, I need a reminder how we did this, can you please tell me where the
+# Effects of non-N copula structure on Taylor's law \label{Consequence_TL_simulation}
+
+\noindent Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
+tested whether the copula structure of dependence between population variables 
+measured in different locations influences the spatial version of Taylor's
+law. Taylor's law is an extremely commonly observed and widely studied 
+[@Taylor1961, @Taylor1988, @Cohen2013, @Xu2016] empirical pattern that relates the 
+variances of groups, $g$, of population measurements to the means of the 
+groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
+$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
+the Taylor's law *exponent* or 
+*slope*. There are several versions of Taylor's law. For spatial Taylor's law,
+given population density or abundance data $x_i(t)$ measured in locations
+$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
+all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
+So means and variances are computed across space. Taylor's law has been 
+verified empirically for hundreds of taxa and has been applied in numerous fields
+including fisheries management, estimation of species persistence times, and
+agriculture [@Cohen2015,@Reuman2017].
+
+We examined using simulations whether copula structure influences spatial Taylor's law as 
+follows. First, an $n \times n$ covariance matrix $\Sigma$ was created
+with unit diagonal entries 
+and off-diagonal entries $\rho=0.7$; $n=25$ was the number of locations at which populations
+were sampled. Then $N \times 1000$ random vectors were generated from a multivariate normal 
+distribution with mean vector $(0,\ldots,0)$ and covariance matrix $\Sigma$. These 
+were organized into an $N \times n \times 1000$ array representing $1000$ simulations of 
+populations measured in $n$ locations over $N$ time steps. 
+The cumulative distribution function
+(cdf) of the standard normal distribution was applied to these numbers, followed by the 
+inverse of the cdf of a gamma distribution with shape and scale parameters $7.5$ and $1$. 
+These transformations made gamma distributed the marginal distributions of 
+each population sampling location in each simulation. 
+A skewed population marginal distribution is an 
+important factor in Taylor's law [@Cohen2015,@Reuman2017]. We refer henceforth to the 
+resulting $N \times n \times 1000$ array as our *N array* for this simulation exercise;
+copula structure of dependences between populations in different sampling locations
+was normal (N) for the simulations in this array.
+
+We then considered an $n$-dimensional C copula with parameter chosen so that the 
+Kendall correlations between each pair of the $n$ components were the same as those
+of the N distribution constructed above.
+
+the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula.
+First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with co-variance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of time series, number of sampling locations and number of replica for that copula, respectively.
+Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal.
+With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D).
+We also tested statistically that whether linearity, homoskedasticity is maintained in each copula data set or not and plotted our results along with root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I).
+We did a quadratic regression with each simulated copula for all types and measured quadratic coefficient and mean curvature for the regression line which is plotted in Fig.\ref{fig_spTL}(J-K) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).
+
+<!--
+Curvature calculations:
+***DAN: Shya, I need a reminder how we did this, can you please tell me where the
 code is, then I can write this section. Feel free to take a stab at it yourself, too,
@@
 
-# Effects of non-N copula structure on Taylor's law \label{Consequence_TL_simulation}
-
-<!--***DAN: Dan has not edited this section yet at all-->
-Here, we examined the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula.
-First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with co-variance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of time series, number of sampling locations and number of replica for that copula, respectively.
-Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal.
-With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D).
-We also tested statistically that whether linearity, homoskedasticity is maintained in each copula data set or not and plotted our results along with root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I).
-We did a quadratic regression with each simulated copula for all types and measured quadratic coefficient and mean curvature for the regression line which is plotted in Fig.\ref{fig_spTL}(J-K) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).
-
-<!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion------
-
-# Consequence of non-Normal copula structure : effect on extinction risk \label{Consequence_ext_risk}
-
-In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
-
-\begin{equation}\label{eq.extrisk_model}
-P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
-\end{equation}
-
-$r$ is the liner growth rate and $K$ is the carrying capacity of the environment.
-$\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence.
-We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion).
-We simulated Eq.
-\ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $25$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;
-
-- LC model (r=0) :
-
-    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
-[see first subplot of Fig.
-\ref{fig_risk_local_disp} and Fig.
-\ref{fig_risk_global_disp}]
-
-- Ricker model (r $\neq$ 0):
-
-    * For ricker model when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
-For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
-[see Fig.
-\ref{fig_risk_local_disp} - \ref{fig_risk_global_disp}]
-
-    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]
-
-With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig.
-\ref{fig_risk_numlocs_effect_local_disp} - Fig.
-\ref{fig_risk_numlocs_effect_global_disp}] for both model.
--->
-
 \clearpage

### 2019-02-13 Paper.Rmd
excised TL from main text

@@
 
-We tested whether the linearity of the spatial Taylor's law relationship was partly
-due to the copula structure of the dependences between the populations sampled. 
-Using corn leaf aphid abundance data as an example, but also
-applying the same analysis to the abundance data we had for 19 other aphid species
-(see *Data*), we calculated the curvature (details in Appendix \ref{SM-TL_curvature}) 
-of the empirical Taylor's law relationship; Taylor's law hypothesizes that this 
-curvature should be close to 0. We then calculated the same curvature quantity for
-10000 surrogate datasets, using the same surrogate procedure described above in 
-this section, and compared the empirical Taylor's law curvature to the 
-distribution of surrogate curvatures using a two-tailed test. If the empirical
-curvature value was below the $5^{\text{th}}$ quantile of the surrogate distribution, 
-then copula structure contributes to the linearity of spatial Taylor's law 
-because randomizing copula structure while retaining other statistical properties
-of the data cause the Taylor's law relationship to become significantly more 
-curved.
-
-<!--***DAN: Shya,  want to add the extinction risk stuff back in, let's discuss-->
+<!--***DAN: 
+1) Shya put concepts on why we would expect lt dependence of noise to mean
+more ext risk in a metapop context
+2) then methods for how we test that (present the model, say what sims you did-->
 
@@
 
-```{r read_tl_aphid_count_sp6, echo=F}
-tlsp6<-readRDS("./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
-tl_aphid_count<-readRDS("./Results/taylorslaw_results/tl_aphid_count.RDS")
-aphid_info<-readRDS("./Results/aphid_info.RDS")
-pleft<-tl_aphid_count$p_left_mncurv
-stat<--2*sum(log(pleft))
-pchi2<-1-pchisq(stat,40)
-idsp<-which(pleft<0.05)
-pleftgood<-pleft[idsp]
-```
-
-Results also suggested that copula structure contributes to the linearity of 
-spatial Taylor's law. For instance, the spatial Taylor's law relationship 
-for corn leaf aphid abundance data was significantly less curved than for 
-surrogate datasets (Fig. \ref{fig_sTL_mncurv_sp6}). Three of $20$ aphid species
-had cuvatures less than at least $95\%$ of surrogate datasets ($p=$ `r pleftgood` for 
-the `r aphid_info$c1[idsp]` species, respectively). 
-<!--***DAN: Shya, please fill in X, Y, Z, A, B, C with autolinks)
-***Shya: Done.-->
-Three such significant
-results of 20 species could have occurred by chance due to 
-multiple-testing error. However, for each of the 20 species, we computed the fraction
-of surrogate curvatures less than the curvature that came from real data, and we
-combined these $p$-values using Fisher's combined probability test to get an
-aggregate $p$-value of `r pchi2`
-<!--***DAN: Shya, see the Wikipedia article on Fisher's method to learn about it. Then 
-you need to put this analysis into the code suite and have the resulting p value
-typed in above be an autolink instead. Here is what I did: 
-1) I copied by hand all the pleft values on the pdf you sent me into a vector pleft. Instead you should create such a vector in your code from your test output, not typing in the results.
-2) I used the formula stat<--2*sum(log(pleft)) to get the Fisher method statistic
-3) I compared to a chi-squared distribution with 40 degress of freedom using 1-pchisq(stat,40) to get my p-value
-Please let me know if any of this is unclear.
-***Shya: Done.-->
-for the test of the null hypothesis that 
-copula structure did not contribute to linearity of Taylor's law for any of the 
-species. This result is usefully suggestive and can guide future work, but still does not
-provide certainty, because the Fisher method provides $p$-values that are too 
-small if its assumption of independent tests is violated. Though our 20 species
-have different host plants so may be independent for the purposes of our analyses, 
-it would be difficult to fully assess independence.
-<!--***DAN: Shya, can you remind me why the same analyses were not done for plankton,
-if there is a good reason? Or if you did it, what were the results?
-***Shya : we did it, I have the pdf but I don't have the code where I called to generate the pdf,
-As we could not get any significant results for a single species (see attachment in the mail I sent you)
-we did not show the results.--> 
-We provide simulation results in Appendix \ref{SM-Consequence_TL_simulation}
-that robustly illustrate, at least in the model setup we used, 
-that copula structure influences spatial Taylor's law.
-
-\begin{figure}[!h]
-\textbf{ \hspace{4 cm} (A) \hspace{4 cm} (B) \hspace{4 cm} (C)} \\
-\vspace{-0.5 cm}
-\begin{center}
-\includegraphics[width=10 cm]{./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf}
-\includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
-\caption{Mean curvature of the spatial Taylor's law relationship for 
-corn leaf aphid abundance
-data (A) was significantly less than for surrogates (B, C). The Taylor's law 
-relationship for one example surrogate (B) shows shows greater curvature than
-for data (A), and the significance of the difference was demonstrated
-by the fact that only `r (tlsp6$p_left_mncurv)*10000` of $10000$ surrogates had curvature less than that
-of the data (C). See Appendix \ref{SM-TL_curvature} for details on 
-how curvature was calculated.\label{fig_sTL_mncurv_sp6}}
-\end{center}
-\end{figure}
-<!--***DAN: Shya, pls make the x axis label on C read simply "Mean curvature"
-***sHYA: dONE.-->
-
-<!--***DAN: I want to put back in some extinction risk results. Let's talk about this, please, Shyamolina-->
+<!--***DAN: Shya to insert extinction risk results with one fig-->
 

### 2019-02-13 SupportingInformation.Rmd
excised TL from main text

@@
 
+<!--
+Stuff on TL cut from the main text
+From Q3 methods:
+We tested whether the linearity of the spatial Taylor's law relationship was partly
+due to the copula structure of the dependences between the populations sampled. 
+Using corn leaf aphid abundance data as an example, but also
+applying the same analysis to the abundance data we had for 19 other aphid species
+(see *Data*), we calculated the curvature (details in Appendix \ref{SM-TL_curvature}) 
+of the empirical Taylor's law relationship; Taylor's law hypothesizes that this 
+curvature should be close to 0. We then calculated the same curvature quantity for
+10000 surrogate datasets, using the same surrogate procedure described above in 
+this section, and compared the empirical Taylor's law curvature to the 
+distribution of surrogate curvatures using a two-tailed test. If the empirical
+curvature value was below the $5^{\text{th}}$ quantile of the surrogate distribution, 
+then copula structure contributes to the linearity of spatial Taylor's law 
+because randomizing copula structure while retaining other statistical properties
+of the data cause the Taylor's law relationship to become significantly more 
+curved.
+
+From Q3 results:
+```{r read_tl_aphid_count_sp6, echo=F}
+tlsp6<-readRDS("./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
+tl_aphid_count<-readRDS("./Results/taylorslaw_results/tl_aphid_count.RDS")
+aphid_info<-readRDS("./Results/aphid_info.RDS")
+pleft<-tl_aphid_count$p_left_mncurv
+stat<--2*sum(log(pleft))
+pchi2<-1-pchisq(stat,40)
+idsp<-which(pleft<0.05)
+pleftgood<-pleft[idsp]
+```
+
+Results also suggested that copula structure contributes to the linearity of 
+spatial Taylor's law. For instance, the spatial Taylor's law relationship 
+for corn leaf aphid abundance data was significantly less curved than for 
+surrogate datasets (Fig. \ref{fig_sTL_mncurv_sp6}). Three of $20$ aphid species
+had cuvatures less than at least $95\%$ of surrogate datasets ($p=$ `r pleftgood` for 
+the `r aphid_info$c1[idsp]` species, respectively). 
+***DAN: Shya, please fill in X, Y, Z, A, B, C with autolinks)
+***Shya: Done.
+Three such significant
+results of 20 species could have occurred by chance due to 
+multiple-testing error. However, for each of the 20 species, we computed the fraction
+of surrogate curvatures less than the curvature that came from real data, and we
+combined these $p$-values using Fisher's combined probability test to get an
+aggregate $p$-value of `r pchi2`
+***DAN: Shya, see the Wikipedia article on Fisher's method to learn about it. Then 
+you need to put this analysis into the code suite and have the resulting p value
+typed in above be an autolink instead. Here is what I did: 
+1) I copied by hand all the pleft values on the pdf you sent me into a vector pleft. Instead you should create such a vector in your code from your test output, not typing in the results.
+2) I used the formula stat<--2*sum(log(pleft)) to get the Fisher method statistic
+3) I compared to a chi-squared distribution with 40 degress of freedom using 1-pchisq(stat,40) to get my p-value
+Please let me know if any of this is unclear.
+***Shya: Done.
+for the test of the null hypothesis that 
+copula structure did not contribute to linearity of Taylor's law for any of the 
+species. This result is usefully suggestive and can guide future work, but still does not
+provide certainty, because the Fisher method provides $p$-values that are too 
+small if its assumption of independent tests is violated. Though our 20 species
+have different host plants so may be independent for the purposes of our analyses, 
+it would be difficult to fully assess independence.
+***DAN: Shya, can you remind me why the same analyses were not done for plankton,
+if there is a good reason? Or if you did it, what were the results?
+***Shya : we did it, I have the pdf but I don't have the code where I called to generate the pdf,
+As we could not get any significant results for a single species (see attachment in the mail I sent you)
+we did not show the results. 
+We provide simulation results in Appendix \ref{SM-Consequence_TL_simulation}
+that robustly illustrate, at least in the model setup we used, 
+that copula structure influences spatial Taylor's law.
+
+\begin{figure}[!h]
+\textbf{ \hspace{4 cm} (A) \hspace{4 cm} (B) \hspace{4 cm} (C)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=10 cm]{./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf}
+\includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
+\caption{Mean curvature of the spatial Taylor's law relationship for 
+corn leaf aphid abundance
+data (A) was significantly less than for surrogates (B, C). The Taylor's law 
+relationship for one example surrogate (B) shows shows greater curvature than
+for data (A), and the significance of the difference was demonstrated
+by the fact that only `r (tlsp6$p_left_mncurv)*10000` of $10000$ surrogates had curvature less than that
+of the data (C). See Appendix \ref{SM-TL_curvature} for details on 
+how curvature was calculated.\label{fig_sTL_mncurv_sp6}}
+\end{center}
+\end{figure}
+***DAN: Shya, pls make the x axis label on C read simply "Mean curvature"
+***sHYA: dONE.
+-->
+
 \clearpage

### 2019-02-13 Paper.Rmd
cleaned code for extinction risk on LCohen model

@@
 
-- Ricker model (r $\neq$ 0):
-
-    * For ricker model when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$.
-For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.
-[see Fig.
-\ref{fig_risk_local_disp} - \ref{fig_risk_global_disp}]
-
-    * With increasing $K$ from $50$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_K_effect}]
-
-With increasing number of patches from $5$ to $25$, extinction risk decreased [for a fixed $d$] ; [see Fig.
-\ref{fig_risk_numlocs_effect_local_disp} - Fig.
-\ref{fig_risk_numlocs_effect_global_disp}] for both model.
 -->

### 2019-02-13 SupportingInformation.Rmd
cleaned code for extinction risk on LCohen model

@@
 library(tibble)
-plankton_info <- tibble(c('Calanus I-IV','Para-Pseudocalanus spp','Temora longicornis','Acartia spp',
+plankton_info <- tibble(c1=c('Calanus stages I-IV','Para-pseudocalanus spp.','Temora longicornis','Acartia sp.',
                           'Centropages typicus',
-                          'Oithona spp','Echinoderm larvae','Calanus finmarchicus',
+                          'Oithona sp.','Echinoderm larvae','Calanus finmarchicus',
                           'Calanus helgolandicus','Metridia lucens','Decapoda larvae',
-                   'Euphausiacea','Thalassiosira spp','Rhizosolenia styliformis',
+                   'Euphausiids','Thalassiosira sp.','Rhizosolenia styliformis',
                    'Ceratium fusus','Ceratium furca','Ceratium tripos','Ceratium macroceros',
-                   'Rhizosolenia alata alata','Pseudocalanus elongatus Adult',
-                   'Nitzschia delicatissima','Nitzschia seriata')
+                   'Proboscia alata','Pseudocalanus sp.',
+                   'Pseudo-nitzschia delicatissima','Pseudo-nitzschia seriata')
 )
+cns<-c('Echinoderm larvae','Euphausiids','Decapoda larvae')
+'%ni%' <- Negate('%in%')
+idn<-which(plankton_info$c1%ni%cns)
 knitr::kable(plankton_info, "latex", booktabs = T, linesep = "",
       caption = "Species names of 22 plankton taxa for which data were available \\label{tab_plankton_info}",
-      col.names = c("Plankton taxon name"))
+      col.names = c("Plankton taxon name"))%>%row_spec(row=idn,italic=T)
 ```
@@
 
-#source("mydispersal.R")
-#set.seed(seed)
-
-#rlist<-c(0,0.05,0.25,0.5,0.7,0.75,0.9,1.0,1.25,1.5,1.75,1.95)
-
-# plot for local dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-#pdf("./Results/ext_risk_local_disp.pdf",heigh=6,width=10)
-#op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
-#for(r in rlist){
-    #cat("r=",r,"\n")
-#    s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
-#    plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-#    lines(s$d_seq,s$risk_right,type="b",col="blue")
-#    legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
-#    mtext(paste0("r = ", r))
-#}
-#par(op)
-#dev.off()
-
-# plot for global dispersal in D matrix ; r=0 for LC model, r!=0 for Ricker model
-#pdf("./Results/ext_risk_global_disp.pdf",heigh=6,width=10)
-#op<-par(mfrow=c(3,4),mar=c(5,4,1,1) + 0.2)
-#for(r in rlist){
-  #cat("r=",r,"\n")
-#  s<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
-#  plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5,cex.axis=1.5)
-#  lines(s$d_seq,s$risk_right,type="b",col="blue")
-#  legend(x=-0.06,y=1.1, c("noise : left","noise : right"), fill=c('red', 'blue'), horiz=T, bty='n')
-#  mtext(paste0("r = ", r))
-#}
-#par(op)
-#dev.off()
-
-# plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
-#pdf("./Results/ext_risk_numlocs_effect_local_disp.pdf",heigh=6,width=6)
-#op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
-#for(r in c(0,0.25)){
-  #cat("r=",r,"\n")
-#  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=F,ploton=F)
-#  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=F,ploton=F)
-
-#  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-#  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
-#  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-#  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
-#       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-#  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
-#  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
-  #mtext(paste0("B",1))
-#}
-#par(op)
-#dev.off()
-
-# plot to show the effect of increasing numlocs for r=0 (LC model), r=0.25 (Ricker model)
-#pdf("./Results/ext_risk_numlocs_effect_global_disp.pdf",heigh=6,width=6)
-#op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
-#for(r in c(0,0.25)){
-  #cat("r=",r,"\n")
-#  s5<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=r,K=50,disp_everywhere=T,ploton=F)
-#  s25<-varying_d(numsims=10000,numsteps =25,numlocs=25,r=r,K=50,disp_everywhere=T,ploton=F)
-
-#  plot(s5$d_seq,s5$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-#  lines(s25$d_seq,s25$risk_left,type="b",col="orange")
-#  legend("topleft", title=paste0("r=",r," , left tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-#  plot(s5$d_seq,s5$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",
-#       panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-#  lines(s25$d_seq,s25$risk_right,type="b",col="skyblue")
-#  legend("topleft",title=paste0("r=",r," , right tail dep. in noise"), c("numlocs=5","numlocs=25"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
-  #mtext(paste0("B",1))
-#}
-#par(op)
-#dev.off()
-
-# plot to see the effect of K for r=0.25 (Ricker model)
-#pdf("./Results/ext_risk_K_effect.pdf",heigh=6,width=6)
-#op<-par(mfrow=c(2,2),mar=c(4,4,1,1) + 0.1)
-
-#s1<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=F,ploton=F)
-#s2<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=F,ploton=F)
-
-#s3<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=50,disp_everywhere=T,ploton=F)
-#s4<-varying_d(numsims=10000,numsteps =25,numlocs=5,r=0.25,K=200,disp_everywhere=T,ploton=F)
-
-#plot(s1$d_seq,s1$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-#lines(s2$d_seq,s2$risk_left,type="b",col="orange")
-#legend("topleft",title=("Noise : left tail dep., local disp."), c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-#plot(s1$d_seq,s1$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-#lines(s2$d_seq,s2$risk_right,type="b",col="skyblue")
-#legend("topleft",title=("Noise : right tail dep., local disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
-
-#plot(s3$d_seq,s3$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="red",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-#lines(s4$d_seq,s4$risk_left,type="b",col="orange")
-#legend("topleft",title= ("Noise : left tail dep., global disp."),c("K=50","K=200"), fill=c('red', 'orange'),  horiz=F, bty='n')
-
-#plot(s3$d_seq,s3$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="blue",panel.first = grid(),xlab="d",ylab="ext_risk",cex.lab=1.5)
-#lines(s4$d_seq,s4$risk_right,type="b",col="skyblue")
-#legend("topleft",title=("Noise : right tail dep., global disp."),c("K=50","K=200"), fill=c('blue', 'skyblue'),  horiz=F, bty='n')
-
-#par(op)
-#dev.off()
+source("mydispersal.R")
+set.seed(seed)
+
+model<-"lcohen"
+numsims<-10000
+numsteps<-25
+params<-0
+ext_thrs<-1
+scl<-1
+p0<-50
+
+# plot for local dispersal in D matrix 
+pdf("./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf",heigh=5,width=10)
+op<-par(mfrow=c(1,2),mar=c(5,5,1.8,0.2),mgp=c(3,1,0))
+s<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
+s1<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
+plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="black",
+     xlab="d",ylab="Extinction risk",
+     cex.lab=1.5,cex.axis=1.5,cex=1,pch=16,lwd=1.5)
+lines(s1$d_seq,s1$risk_left,type="b",lwd=1.5,cex=1,pch=1)
+legend(x=0.3,y=0.9, title=paste0("r = ",params," , left tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
+       pch=c(16,1), lty=c(1,1),
+       horiz=F, bty='n')
+
+plot(s$d_seq,s$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="grey",
+     xlab="d",ylab="Extinction risk",
+     cex.lab=1.5,cex.axis=1.5,cex=1,pch=16,lwd=1.5)
+lines(s1$d_seq,s1$risk_right,type="b",lwd=1.5,cex=1,pch=1,col="grey")
+legend(x=0.3,y=0.9, title=paste0("r = ",params," , right tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
+       pch=c(16,1), lty=c(1,1), col=c("grey","grey"),
+       horiz=F, bty='n')
+
+par(op)
+dev.off()
+
+# plot for global dispersal in D matrix 
+# plot for local dispersal in D matrix 
+pdf("./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf",heigh=5,width=10)
+op<-par(mfrow=c(1,2),mar=c(5,5,1.8,0.2),mgp=c(3,1,0))
+s<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
+s1<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
+plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="black",xlab="d",ylab="Extinction risk",
+     cex.lab=1.5,cex.axis=1.5,cex=1,pch=16,lwd=1.5)
+lines(s1$d_seq,s1$risk_left,type="b",lwd=1.5,cex=1,pch=1)
+legend(x=0.3,y=0.9, title=paste0("r = ",params," , left tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
+       pch=c(16,1), lty=c(1,1),
+       horiz=F, bty='n')
+
+plot(s$d_seq,s$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="grey",xlab="d",ylab="Extinction risk",
+     cex.lab=1.5,cex.axis=1.5,cex=1,pch=16,lwd=1.5)
+lines(s1$d_seq,s1$risk_right,type="b",lwd=1.5,cex=1,pch=1,col="grey")
+legend(x=0.3,y=0.9, title=paste0("r = ",params," , right tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
+       pch=c(16,1), lty=c(1,1), col=c("grey","grey"),
+       horiz=F, bty='n')
+
+par(op)
+dev.off()
 ```

### 2019-02-13 mydispersal.R

@@ -53,7 +53,7 @@ Dispersal_mat<-function(numlocs,d,disp_everywhere){
 #--------Output------------------------------------------------------
 #A numsims by numlocs by numsteps+1 array of populations
 #----------------------------------------------------------------------
-popsim_ml_D<-function(p0,ns,D,r,K,a,b,ext_thrs,model){
+popsim_ml_D<-function(p0,ns,D,params,ext_thrs,model){
   numsims<-dim(ns)[1]
   numlocs<-dim(ns)[2]
   numsteps<-dim(ns)[3]
@@ -70,25 +70,12 @@ popsim_ml_D<-function(p0,ns,D,r,K,a,b,ext_thrs,model){
     #growth rates based on the noise
     lam_sto<-exp(ns[,,tct])  #numsims by numlocs matrix
     
-    lam_ricker<-exp(r*(1-(res[,,tct]/K)))
-    lam_verhulst<-1+(r*(1-res[,,tct]))
-    lam_hassell<-r/((1+(a*res[,,tct]))^b)
-    lam_msmith<-r/(1+((a*res[,,tct])^b))
-    
     #growth prior to dispersal
-    if(model=="ricker"){
-      given_model<-lam_ricker
-    }else if(model=="verhulst"){
-      given_model<-lam_verhulst
-    }else if(model=="hassell"){
-      given_model<-lam_hassell
-    }else if(model=="msmith"){
-      given_model<-lam_msmith
+    if(model=="lcohen"){
+      res[,,tct+1]<-lam_sto*exp(params)*res[,,tct] #numsims by numlocs matrix
     }else{
       warning("model not specified",immediate.=T,call.=T)
     }
-    
-    res[,,tct+1]<-given_model*lam_sto*res[,,tct] #numsims by numlocs matrix
     
     #after dispersal
     tres<-t(res[,,tct+1]) # numlocs by numsims matrix after taking transpose
@@ -117,15 +104,14 @@ extrisk<-function(sims){
 #       2) numsteps : an integer : number of time steps
 #       3) numlocs : an integer : number of locations
 #       4) D : dispersal matrix which is the output of Dispersal_mat function
-#       5) r : model parameter : growth rate
-#       6) K : Ricker model parameter : carrying capacity
-#       7), 8) a, b : model parameters for Hassell model 
-#       9) ext_thrs : extinction threshold below which populations go extinct
[…]

### 2019-02-13 Paper.Rmd
extinction risk results added

@@
 
-<!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion-----------
-Finally, we studied **(iii)** the effects of environmental noise, having a non-Normal copula, on extinction risk (calculated after 25 years) with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}).
-We considered patches were arranged in a linear chain fashion and the population may disperse to its nearerst two neighbors (local dispersion) or to any patches (global dispersion).
-For both model with both dispersion, increasing number of habitat patches and increasing carrying capacity, separately, had decreased extinction risk of the populations which was expected (Fig.
-\ref{SM-fig_risk_numlocs_effect_local_disp} - Fig.
-\ref{SM-fig_risk_K_effect}).
-For Lewton-Cohen model, we found greater extinction risk if populations were affected by left tail dependence environmental noise copula than right tail dependent one.
-For Ricker model, we got the similar observation for smaller values of $r$, the intrinsic growth rate, but for higher values of $r$, there was a turn-over in extinction risks, and right tail dependent noise copula caused population go extinct faster.
--->
-
-<!-- 
-This extincton risk material had been commented out in the sup mat:
-
-Extinction risk results are not shown currently, it is introduced as a future idea in Discussion------
-
-# Consequence of non-Normal copula structure : effect on extinction risk \label{Consequence_ext_risk}
-
-In order to study the effects of non-Normal copula structure on extinction risk for metapopulations with different dispersal rules, we examined the following generalized multiple patch model where local population dynamics at each patch were governed by the Lewton-Cohen model ($r=0$) or Ricker model ($r \neq 0$) as the case may be;
-
-\begin{equation}\label{eq.extrisk_model}
-P(t+1) = \exp\left [  r\left ( 1-\frac{P(t)}{K} \right ) + \epsilon(t) \right ] P(t)
-\end{equation}
-
-$r$ is the liner growth rate and $K$ is the carrying capacity of the environment.
-$\epsilon$ is the environmental noise which might take stronger right tail (means stronger upper tail)  or stronger left tail (means stronger lower tail) dependence.
-We then considered each population $P(t+1)$ was updated depending upon $D$, the dispersal matrix, which could take either forms; one case when each population patch on a linear chain interacted with its two nearest neighbors (we called it local dispersion) or when any patch was allowed to interact with the rest of all (we called it as global dispersion).
-We simulated Eq.
-\ref{eq.extrisk_model} for $10000$ simulations and calculated extinction risk after $25$ years with different possibilities (for code see [mydispersal.R](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R)) and summarized our findings as follows;
-
-- LC model (r=0) :
-
-    * For LC model, as expected, when extinction risk calculated after $25$ years, was plotted against each varying $d$ (=degree of dispersion), we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one.
-[see first subplot of Fig.
-\ref{fig_risk_local_disp} and Fig.
-\ref{fig_risk_global_disp}]
-
--->
+On a second thought, we addressed the question how asymmetry in 
+tail-dependence of environmental
+variable could influence the extinction risk from a meta-population
+context. We hypothesized to get higher 
+extinction risk if the environmental noises from multiple patches
+were more correlated to their smaller values 
+(resulting left-tail dependence
+in copula structure) than larger. This idea came from the context that
+"bad" years (e.g. extreme cold) would have adverse affect 
+on all metapopulation patches simultaneously, so there would 
+not be enough neighboring-populations to rescue the "extinct"
+patches. But in "good" years as the environmental variables 
+from multiple locations were not correlated strongly, some of
+extinct metapopulation-patches could be rescued by dispersal from
+other patches. We tested our hypothesis in the following section.
 
@@
 <!--***DAN: Shya to insert extinction risk results with one fig-->
+We tested our hypothesis of asymmetric copula-effect on extinction-risk
+as discussed in
+earlier section, with Lewinton-Cohen metapopulation
+model (Eq. \ref{eq.extrisk_LC}), where population density 
+$P$ could update exponentially depending on its intrinsic rate $r$.
+$\epsilon(t)$ represent environmental noise; they are standard-normal 
+random variables that are independent for distinct times, t, but 
+that were made to exhibit either extreme left tail or right tail dependence
+across patches.
+We considered metapopulation patches were arranged in
+a linear chain fashion
+and the population may disperse to its nearerst
+two neighbors (local dispersion) or to any patches (global dispersion); for code see [`mydispersal.R`](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R).
+
+\begin{equation}\label{eq.extrisk_LC}
+P(t+1) = \exp\left [  r + \epsilon(t) \right ] P(t)
+\end{equation}
+
+We simulated Eq. \ref{eq.extrisk_LC} with local dispersal for $r=0$ and calculated extinction risk after 25 years
+from $10000$ simulations for both left-tail and right-tail dependent environmental noise across $5$ patches. 
+Whenever population, at any patch after each year, went below the extinction threshold ($=1$) we considered 
+that patch for that time went extinct. We found, as we
+hypothesized, left tail dependent environmental fluctuations increased extinction risk of metapopulations than the right-tail
+dependent ones (Fig. \ref{fig_extrisk_lcohen_local}). When number of patches increased from $5$ to $25$, extinction risk 
+decreased in both cases. We repeated our simulations using same parameters with global dispersal among patches 
+and got similar results (see Fig. \ref{SM-fig_extrisk_lcohen_global}).
+
+\begin{figure}[!h]
+\textbf{ \hspace{6.6 cm} (A) \hspace{4.2 cm} (B)} \\
+\vspace{-0.8 cm}
+\begin{center}
+\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
+\caption{Extinction risk for Lewinton-Cohen metapopulation model was calculated after $25$ years with local dispersal
+among two nearest neighbors (Eq. \ref{eq.extrisk_LC}, $r=0$). $(A)$ Left-tail dependent environmental noise showed higher 
+extinction risk than $(B)$ right-tail dependent one, where both case showed extinction risk decreased with 
+increasing number of patches from $numlocs =5$ to $25$.\label{fig_extrisk_lcohen_local}}
+\end{center}
+\end{figure}
 

### 2019-02-13 SupportingInformation.Rmd
extinction risk results added

@@
 pdf("./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf",heigh=5,width=10)
-op<-par(mfrow=c(1,2),mar=c(5,5,1.8,0.2),mgp=c(3,1,0))
+op<-par(mfrow=c(1,2),mar=c(5,5,1.8,0.5),mgp=c(3,1,0))
 s<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
@@
      xlab="d",ylab="Extinction risk",
-     cex.lab=1.5,cex.axis=1.5,cex=1,pch=16,lwd=1.5)
-lines(s1$d_seq,s1$risk_left,type="b",lwd=1.5,cex=1,pch=1)
-legend(x=0.3,y=0.9, title=paste0("r = ",params," , left tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
-       pch=c(16,1), lty=c(1,1),
+     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
+lines(s1$d_seq,s1$risk_left,type="b",lwd=1.5,cex=1.2,pch=1)
+legend(x=0.2,y=0.9, title=paste0("r = ",params," , left tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
+       pch=c(16,1), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.2),
        horiz=F, bty='n')
@@
      xlab="d",ylab="Extinction risk",
-     cex.lab=1.5,cex.axis=1.5,cex=1,pch=16,lwd=1.5)
-lines(s1$d_seq,s1$risk_right,type="b",lwd=1.5,cex=1,pch=1,col="grey")
-legend(x=0.3,y=0.9, title=paste0("r = ",params," , right tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
-       pch=c(16,1), lty=c(1,1), col=c("grey","grey"),
+     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
+lines(s1$d_seq,s1$risk_right,type="b",lwd=1.5,cex=1.2,pch=1,col="grey")
+legend(x=0.2,y=0.9, title=paste0("r = ",params," , right tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
+       pch=c(16,1), lty=c(1,1), col=c("grey","grey"), cex=1.4,pt.cex=c(1.5,1.2),
        horiz=F, bty='n')
@@
 # plot for global dispersal in D matrix 
-# plot for local dispersal in D matrix 
 pdf("./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf",heigh=5,width=10)
@@
 plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="black",xlab="d",ylab="Extinction risk",
-     cex.lab=1.5,cex.axis=1.5,cex=1,pch=16,lwd=1.5)
-lines(s1$d_seq,s1$risk_left,type="b",lwd=1.5,cex=1,pch=1)
-legend(x=0.3,y=0.9, title=paste0("r = ",params," , left tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
-       pch=c(16,1), lty=c(1,1),
+     cex.lab=1.5,cex.axis=1.5,cex=1.5,pch=16,lwd=1.5)
+lines(s1$d_seq,s1$risk_left,type="b",lwd=1.5,cex=1.2,pch=1)
+legend(x=0.2,y=0.9, title=paste0("r = ",params," , left tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
+       pch=c(16,1), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.2),
        horiz=F, bty='n')
@@
 plot(s$d_seq,s$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="grey",xlab="d",ylab="Extinction risk",
-     cex.lab=1.5,cex.axis=1.5,cex=1,pch=16,lwd=1.5)
-lines(s1$d_seq,s1$risk_right,type="b",lwd=1.5,cex=1,pch=1,col="grey")
-legend(x=0.3,y=0.9, title=paste0("r = ",params," , right tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
-       pch=c(16,1), lty=c(1,1), col=c("grey","grey"),
+     cex.lab=1.5,cex.axis=1.5,cex=1.5,pch=16,lwd=1.5)
+lines(s1$d_seq,s1$risk_right,type="b",lwd=1.5,cex=1.2,pch=1,col="grey")
+legend(x=0.2,y=0.9, title=paste0("r = ",params," , right tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
+       pch=c(16,1), lty=c(1,1), col=c("grey","grey"), cex=1.4,pt.cex=c(1.5,1.2),
        horiz=F, bty='n')
@@
 
-<!-- Extinction risk results are not shown currently, it is introduced as a future idea in Discussion-->
-<!--
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with local dispersion (see Eq.
-\ref{eq.extrisk_model}); numlocs=5, numsteps=25, numsims=10000\label{fig_risk_local_disp}](./Results/ext_risk_local_disp.pdf){ width=40%, height=40% }
-
-![Effect of non-Normal environmental noise copula on extinction risk after 25 years (Lewton-Cohen model : $r=0$, Ricker model : $r\neq0$) with global dispersion (see Eq.
-\ref{eq.extrisk_model}); numlocs=5, numsteps=25, numsims=10000\label{fig_risk_global_disp}](./Results/ext_risk_global_disp.pdf){ width=40%, height=40% }
-
-![Increasing patch number allows metapopulation to disperse more efficiently and thereby decresing extinction risk calculated after 25 years; models with local dispersion with numsims=10000 (see Eq.
-\ref{eq.extrisk_model}).\label{fig_risk_numlocs_effect_local_disp}](./Results/ext_risk_numlocs_effect_local_disp.pdf){ width=40%, height=40% }
-
-![Increasing patch number allows metapopulation to disperse more efficiently and thereby decresing extinction risk calculated after 25 years; models with global dispersion with numsims=10000 (see Eq.
-\ref{eq.extrisk_model}).\label{fig_risk_numlocs_effect_global_disp}](./Results/ext_risk_numlocs_effect_global_disp.pdf){ width=40%, height=40% }
-
-![Extinction risk as calculated after 25 years on Ricker model (r=0.25, numsims=10000) decreases as carrying capacity $K$ increases for both local and global dispersion (see Eq.
-\ref{eq.extrisk_model}).\label{fig_risk_K_effect}](./Results/ext_risk_K_effect.pdf){ width=40%, height=40% }
--->
+\begin{figure}[!h]
+\textbf{ \hspace{7.3 cm} (A) \hspace{4 cm} (B)} \\
+\vspace{-0.8 cm}
+\begin{center}
+\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
+\caption{Analogous to figure \ref{MT-fig_extrisk_lcohen_local} in maintext but with global dispersion.\label{fig_extrisk_lcohen_global}}
+\end{center}
+\end{figure}
 

### 2019-02-18 Paper.Rmd
minor change
edited methods and background and some of the fig for extinction risk stuff

@@
 spatial average. Similarly, left-tail dependence between local variables
-will tend to decreases skewness of the spatial mean. Skewness of the 
-spatial mean corresponds to the "spikiness" of the spatial-mean time series or
-its tendency to crash to low levels. 
+will tend to decreases skewness of the spatial mean. Positive skewness of the 
+spatial mean corresponds to the "spikiness" of the spatial-mean time series; negative 
+skewness, if it occurs,  corresponds to a 
+tendency to crash to low levels but quickly recover. 
 The spatial-mean time series and its skewness may be quantities of 
@@
 to a distribution of values 
-of the same quantity calculated for each of 10000 surrogate datasets that were 
-produced from the empirical data by randomizing
-it in a special way to have the copula structure of a multivariate-normal distribution
-(Appendix \ref{SM-Surrog_multi}). The surrogate/randomized datasets had exactly the same
-marginal distributions (i.e., distributions of values for each sampling location)
+of the same quantity calculated for each of 10000 surrogate datasets. Surrogate 
+datasets were produced from the empirical data by randomizing
+it in a special way so that surrogate/randomized dataset then had the 
+copula structure of a multivariate-normal distribution
+(Appendix \ref{SM-Surrog_multi}). The surrogate datasets had exactly the same
+distributions of values for each sampling location as the original data. 
+They also had very similar Spearman correlations between pairs of sampling locations
 as the original data.  
-They also had very similar Spearman correlations between pairs of sampling locations
-as the original data. But the surrogates had multivariate normal copula structure. 
 Our comparisons therefore tested the null hypothesis that the skewness of the spatial 
@@
 structure of the data were multivariate normal, but the data were otherwise statistically
-unchanged. Significant differences indicate that copula structure contributed 
-to the skewness of the spatial mean, i.e., to its "spikiness."
+unchanged. Significant differences indicate that non-N copula structure in the data 
+contributed to the skewness of the spatial mean, i.e., to the "spikiness" of that
+time series.
 
@@
 
-<!--***DAN: 
-1) Shya put concepts on why we would expect lt dependence of noise to mean
-more ext risk in a metapop context
-2) then methods for how we test that (present the model, say what sims you did-->
-
-On a second thought, we addressed the question how asymmetry in 
-tail-dependence of environmental
-variable could influence the extinction risk from a meta-population
-context. We hypothesized to get higher 
-extinction risk if the environmental noises from multiple patches
-were more correlated to their smaller values 
-(resulting left-tail dependence
-in copula structure) than larger. This idea came from the context that
-"bad" years (e.g. extreme cold) would have adverse affect 
-on all metapopulation patches simultaneously, so there would 
-not be enough neighboring-populations to rescue the "extinct"
-patches. But in "good" years as the environmental variables 
-from multiple locations were not correlated strongly, some of
-extinct metapopulation-patches could be rescued by dispersal from
-other patches. We tested our hypothesis in the following section.
+We also examined whether the nature of the
+tail dependence of an environmental variable across space could influence the 
+extinction risk of a metapopulation. We hypothesized that environmental noises
+exhibiting greater left-tail dependence across different habitat patches would 
+cause higher metapopulation extinction risks because then bad years for the 
+component populations would occur simultaneously in many patches,
+reducing the potential for rescue effects. 
+We provided a test of the conceptual validity of this hypothesis using a metapopulation
+extension of the Lewinton-Cohen model,
+$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
+where $\vec{P}(t)$ is a vector of populations, the $i$th component $P_i(t)$ representing
+the population in the $i$th habitat patch ($i=1,\ldots,N$) at time $t$. 
+The $N \times N$ matrix $\lambda(t)$ was diagonal with $i$th diagonal entry 
+$\exp(r+\epsilon_i(t))$. Here $r$ is an intrinsic growth rate; we used $r=0$. The
+$\epsilon_i(t)$ represent environmental noises in each patch. They were standard-normally 
+distributed in each patch, were independent through time, and showed between-patch
+Spearman correlations X and Pearson correlations Y in all simulations, 
+<!--***Shya: Please fill in X and Y. If it was the Kendall that was used, please 
+change Spearman to Kendall. If anything is unclear here in what you should do, please ask. 
+If you don't know how to answer these questions, let me know and remind me where the 
+code is that you used to generate the noise. Actually, even if you can do this yourself,
+please insert a comment below this one indcating where the code is for model and noise-->
+but were made to exhibit right- or left-tail dependence between patches in
+different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
+matrix $D$ was a dispersal matrix. For simplicity, we considered habitat patches 
+to be arranged in a line, with populations dispersing either to nearest neighbors only,
+or globally, in different simulations (Appendix \ref{SM-ext_risk_model}). The dispersal 
+rate was $d$. After each time step, if the population in a patch was below $1$, it was
+set to $0$. We simulated the model 10000 times for each combination of parameters,
+starting from initial populations in each patch of X, 
+<!--***DAN: Shya, please fill in X-->
+and calculated extinction risk after $25$ time steps.
 
@@
 
-<!--***DAN: Shya to insert extinction risk results with one fig-->
-We tested our hypothesis of asymmetric copula-effect on extinction-risk
-as discussed in
-earlier section, with Lewinton-Cohen metapopulation
-model (Eq. \ref{eq.extrisk_LC}), where population density 
-$P$ could update exponentially depending on its intrinsic rate $r$.
-$\epsilon(t)$ represent environmental noise; they are standard-normal 
-random variables that are independent for distinct times, t, but 
-that were made to exhibit either extreme left tail or right tail dependence
-across patches.
-We considered metapopulation patches were arranged in
-a linear chain fashion
-and the population may disperse to its nearerst
-two neighbors (local dispersion) or to any patches (global dispersion); for code see [`mydispersal.R`](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R).
-
-\begin{equation}\label{eq.extrisk_LC}
-P(t+1) = \exp\left [  r + \epsilon(t) \right ] P(t)
-\end{equation}
-
-We simulated Eq. \ref{eq.extrisk_LC} with local dispersal for $r=0$ and calculated extinction risk after 25 years
-from $10000$ simulations for both left-tail and right-tail dependent environmental noise across $5$ patches. 
-Whenever population, at any patch after each year, went below the extinction threshold ($=1$) we considered 
-that patch for that time went extinct. We found, as we
+<!--***DAN: Dan has not edited the results for the extinction risk stuff yet-->
+ We found, as we
 hypothesized, left tail dependent environmental fluctuations increased extinction risk of metapopulations than the right-tail
@@
 \caption{Extinction risk for Lewinton-Cohen metapopulation model was calculated after $25$ years with local dispersal
-among two nearest neighbors (Eq. \ref{eq.extrisk_LC}, $r=0$). $(A)$ Left-tail dependent environmental noise showed higher 
+among two nearest neighbors (Eq. \ref{eq.extrisk_LC}, $r=0$, $d=$ degree of dispersal). 
+$(A)$ Left-tail dependent environmental noise showed higher 
 extinction risk than $(B)$ right-tail dependent one, where both case showed extinction risk decreased with 
@@
 \end{figure}
+<!--***DAN: Shya, can you please change this figure as follows:
+1) Make it only go to d=0.5. I think reviewers will complain that d close to 1 is a weird and 
+unrealistic. Anyway, results are the same for larger d, so we are not hiding anything
+2) the main comparison is between right and left tail dependence, so make the left panel show 
+the numlocs=5 case, with results for both right- and left-tail dependence of the noise shown. 
+And make the right panel show the numlocs=25 case with results for both right- and left-tail
+dependence of the noise shown.
+3) Right now you use grey dots on one panel and black on the other, but you can just use black 
+on both.
+4)  Use N instead of numlocs - in papers we use variable names like in math, not like in code.
+-->
 

### 2019-02-18 Paper.Rmd
response added

@@
 examined environmental variables from the soil instead, using the Rapid Carbon 
-Assessment database (RaCA; @Wills2014).
+Assessment database (RaCA; [@Wills2014]).
 <!--***DAN: Add Wills and Loecke pending reference if published in time--> 
@@
 35 years from 1976 to 2010. These time series were computed from the the Rothamsted Insect 
-Survey suction-trap dataset [@Macaulay1988, @Harrington2014]. The first of our two 
+Survey suction-trap dataset [@Macaulay1988; @Harrington2014]. The first of our two 
 population-level datasets was also derived from the RIS suction-trap data, and comprised 
@@
 $\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
-where $\vec{P}(t)$ is a vector of populations, the $i$th component $P_i(t)$ representing
-the population in the $i$th habitat patch ($i=1,\ldots,N$) at time $t$. 
-The $N \times N$ matrix $\lambda(t)$ was diagonal with $i$th diagonal entry 
+where $\vec{P}(t)$ is a vector of populations, the $i^{th}$ component $P_i(t)$ representing
+the population in the $i^{th}$ habitat patch ($i=1,\ldots,N$) at time $t$. 
+The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
 $\exp(r+\epsilon_i(t))$. Here $r$ is an intrinsic growth rate; we used $r=0$. The
@@
 Spearman correlations X and Pearson correlations Y in all simulations, 
-<!--***Shya: Please fill in X and Y. If it was the Kendall that was used, please 
+<!--***DAN: Please fill in X and Y. If it was the Kendall that was used, please 
 change Spearman to Kendall. If anything is unclear here in what you should do, please ask. 
@@
 code is that you used to generate the noise. Actually, even if you can do this yourself,
-please insert a comment below this one indcating where the code is for model and noise-->
+please insert a comment below this one indcating where the code is for model and noise
+***Shya: I don't know the answer, the code which generates extreme right or left tail dep noise 
+is written by you : ExtremeTailDep.R; there is no mention to preserve spearman or kendall or pearson.
+I called retd function from ExtremeTaildep.R in another script mydispersal.R line 116, 128 
+to generate right and left tail dep. noise.-->
 but were made to exhibit right- or left-tail dependence between patches in
@@
 \caption{Extinction risk for Lewinton-Cohen metapopulation model was calculated after $25$ years with local dispersal
-among two nearest neighbors (Eq. \ref{eq.extrisk_LC}, $r=0$, $d=$ degree of dispersal). 
-$(A)$ Left-tail dependent environmental noise showed higher 
-extinction risk than $(B)$ right-tail dependent one, where both case showed extinction risk decreased with 
-increasing number of patches from $numlocs =5$ to $25$.\label{fig_extrisk_lcohen_local}}
+among two nearest neighbors ($r=0$, $d=$ rate of dispersal). 
+Left-tail dependent environmental noise caused higher 
+extinction risk than right-tail dependent one for number of patches $(A)$ $N = 5$ and $(B)$ $N = 25$, respectively,
+where both case showed lower extinction risk with increasing number of patches.\label{fig_extrisk_lcohen_local}}
 \end{center}
@@
 4)  Use N instead of numlocs - in papers we use variable names like in math, not like in code.
--->
+***Shya: Done.-->
 

### 2019-02-18 SupportingInformation.Rmd
edited methods and background and some of the fig for extinction risk stuff
response added

@@
 
-```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+```{r sim_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
 
@@
 
+sl5<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
+saveRDS(sl5,"./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs5.RDS")
+
+sl25<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
+saveRDS(sl25,"./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs25.RDS")
+
+sg5<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
+saveRDS(sg5,"./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs5.RDS")
+
+sg25<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
+saveRDS(sg25,"./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs25.RDS")
+```
+
+```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,sl5,sl25,sg5,sg25)}
+sl5<-readRDS("./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs5.RDS")
+sl25<-readRDS("./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs25.RDS")
+sg5<-readRDS("./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs5.RDS")
+sg25<-readRDS("./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs25.RDS")
+
 # plot for local dispersal in D matrix 
 pdf("./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf",heigh=5,width=10)
-op<-par(mfrow=c(1,2),mar=c(5,5,1.8,0.5),mgp=c(3,1,0))
-s<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
-                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
-s1<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
-                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
-plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="black",
+op<-par(mfrow=c(1,2),mar=c(5,5,1.8,1),mgp=c(3,1,0))
+plot(sl5$d_seq[1:6],sl5$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
      xlab="d",ylab="Extinction risk",
      cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
-lines(s1$d_seq,s1$risk_left,type="b",lwd=1.5,cex=1.2,pch=1)
-legend(x=0.2,y=0.9, title=paste0("r = ",params," , left tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
-       pch=c(16,1), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.2),
+lines(sl5$d_seq[1:6],sl5$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
+legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 5"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
+       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
        horiz=F, bty='n')
 
-plot(s$d_seq,s$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="grey",
+plot(sl25$d_seq[1:6],sl25$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
      xlab="d",ylab="Extinction risk",
      cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
-lines(s1$d_seq,s1$risk_right,type="b",lwd=1.5,cex=1.2,pch=1,col="grey")
-legend(x=0.2,y=0.9, title=paste0("r = ",params," , right tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
-       pch=c(16,1), lty=c(1,1), col=c("grey","grey"), cex=1.4,pt.cex=c(1.5,1.2),
+lines(sl25$d_seq[1:6],sl25$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
+legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 25"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
+       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
        horiz=F, bty='n')
@@
 pdf("./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf",heigh=5,width=10)
-op<-par(mfrow=c(1,2),mar=c(5,5,1.8,0.2),mgp=c(3,1,0))
-s<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
-                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
-s1<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
-                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
-plot(s$d_seq,s$risk_left,xlim=c(0,1),ylim=c(0,1),type="b",col="black",xlab="d",ylab="Extinction risk",
-     cex.lab=1.5,cex.axis=1.5,cex=1.5,pch=16,lwd=1.5)
-lines(s1$d_seq,s1$risk_left,type="b",lwd=1.5,cex=1.2,pch=1)
-legend(x=0.2,y=0.9, title=paste0("r = ",params," , left tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
-       pch=c(16,1), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.2),
+op<-par(mfrow=c(1,2),mar=c(5,5,1.8,1),mgp=c(3,1,0))
+plot(sg5$d_seq[1:6],sg5$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
+     xlab="d",ylab="Extinction risk",
+     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
+lines(sg5$d_seq[1:6],sg5$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
+legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 5"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
+       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
        horiz=F, bty='n')
 
-plot(s$d_seq,s$risk_right,xlim=c(0,1),ylim=c(0,1),type="b",col="grey",xlab="d",ylab="Extinction risk",
-     cex.lab=1.5,cex.axis=1.5,cex=1.5,pch=16,lwd=1.5)
-lines(s1$d_seq,s1$risk_right,type="b",lwd=1.5,cex=1.2,pch=1,col="grey")
-legend(x=0.2,y=0.9, title=paste0("r = ",params," , right tail dep. in noise"), c("numlocs = 5","numlocs = 25"), 
-       pch=c(16,1), lty=c(1,1), col=c("grey","grey"), cex=1.4,pt.cex=c(1.5,1.2),
+plot(sg25$d_seq[1:6],sg25$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
+     xlab="d",ylab="Extinction risk",
+     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
+lines(sg25$d_seq[1:6],sg25$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
+legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 25"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
+       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
        horiz=F, bty='n')
-
 par(op)
@@
 \includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
-\caption{Analogous to figure \ref{MT-fig_extrisk_lcohen_local} in maintext but with global dispersion.\label{fig_extrisk_lcohen_global}}
+\caption{Analogous to figure \ref{MT-fig_extrisk_lcohen_local} in maintext but with global dispersal.\label{fig_extrisk_lcohen_global}}
 \end{center}
@@
 
+# Extinction risk noise and dispersal matrix \label{ext_risk_model}
+
+<!--How noise was generated ? code : ExtremeTailDep.R-->
+
+<!--How dispersal matrix was generated ? code : mydispersal.R-->
+A $N \times N$ dispersal matrix ($D$) was generated assuming that the $N$ number of metapopulation patches
+were arranged in a linear chain fashion. Say, population from $i^{th}$ patch could either disperse to its nearest
+connecting patches only (we called it as local dispersal) or could disperse to all other patches 
+equally (we called it as global dispersal) at an dispersal rate, $d$. In case of 
+local dispersal matrix the diagonal entries $D_{ii}$ would be equal to $d$ and its immediate neighbors
+$D_{ii\pm j ,\;\ j=1}$ would get an equal proportion from the rest of $(1-d)$. The other distant neighbors 
+$D_{ii\pm j ,\; j>1}$ were set to zero to execute local dispersal. For global dispersal matrix, $D_{ii}$
+would be $d$ and all the other entries were equal to $(1-d)/N$. Code to generate these dispersal matrices 
+was implemented in [Dispersal_mat](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R) function.
+
 # Effects of non-N copula structure on Taylor's law \label{Consequence_TL_simulation}
@@
 law. Taylor's law is an extremely commonly observed and widely studied 
-[@Taylor1961, @Taylor1988, @Cohen2013, @Xu2016] empirical pattern that relates the 
+[@Taylor1961; @Taylor1988; @Cohen2013; @Xu2016] empirical pattern that relates the 
 variances of groups, $g$, of population measurements to the means of the 
@@
 including fisheries management, estimation of species persistence times, and
-agriculture [@Cohen2015,@Reuman2017].
+agriculture [@Cohen2015; @Reuman2017].
 
@@
 A skewed population marginal distribution is an 
-important factor in Taylor's law [@Cohen2015,@Reuman2017]. We refer henceforth to the 
+important factor in Taylor's law [@Cohen2015; @Reuman2017]. We refer henceforth to the 
 resulting $N \times n \times 1000$ array as our *N array* for this simulation exercise;
@@
 ***DAN: Shya, pls make the x axis label on C read simply "Mean curvature"
-***sHYA: dONE.
+***Shya: dONE.
 -->

### 2019-02-19 Paper.Rmd
finished with a rount of editing the Q3 sections, except still need to do the TL sup mat section

@@
 component populations would occur simultaneously in many patches,
-reducing the potential for rescue effects. 
+reducing rescue effects. 
 We provided a test of the conceptual validity of this hypothesis using a metapopulation
@@
 $\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
-where $\vec{P}(t)$ is a vector of populations, the $i^{th}$ component $P_i(t)$ representing
-the population in the $i^{th}$ habitat patch ($i=1,\ldots,N$) at time $t$. 
+where $\vec{P}(t)$ is a vector of population densities, 
+the $i^{th}$ component $P_i(t)$ representing
+the density in the $i^{th}$ habitat patch ($i=1,\ldots,N$) at time $t$. 
 The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
-$\exp(r+\epsilon_i(t))$. Here $r$ is an intrinsic growth rate; we used $r=0$. The
+$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
 $\epsilon_i(t)$ represent environmental noises in each patch. They were standard-normally 
-distributed in each patch, were independent through time, and showed between-patch
-Spearman correlations X and Pearson correlations Y in all simulations, 
-<!--***DAN: Please fill in X and Y. If it was the Kendall that was used, please 
-change Spearman to Kendall. If anything is unclear here in what you should do, please ask. 
-If you don't know how to answer these questions, let me know and remind me where the 
-code is that you used to generate the noise. Actually, even if you can do this yourself,
-please insert a comment below this one indcating where the code is for model and noise
-***Shya: I don't know the answer, the code which generates extreme right or left tail dep noise 
-is written by you : ExtremeTailDep.R; there is no mention to preserve spearman or kendall or pearson.
-I called retd function from ExtremeTaildep.R in another script mydispersal.R line 116, 128 
-to generate right and left tail dep. noise.-->
+distributed in each patch, were independent through time, and showed the same 
+spatial correlations (to within sampling error) for every simulation, 
 but were made to exhibit right- or left-tail dependence between patches in
@@
 matrix $D$ was a dispersal matrix. For simplicity, we considered habitat patches 
-to be arranged in a line, with populations dispersing either to nearest neighbors only,
-or globally, in different simulations (Appendix \ref{SM-ext_risk_model}). The dispersal 
-rate was $d$. After each time step, if the population in a patch was below $1$, it was
+to be arranged in a line, with dispersal occurring either to nearest neighbors only,
+or globally, in different simulations. The dispersal 
+rate was $d$ (Appendix \ref{SM-ext_risk_model}). After each time step, 
+if the population density in a patch was below $1$ (an extinction threshold), it was
 set to $0$. We simulated the model 10000 times for each combination of parameters,
@@
 
-Skewness results were generally consistent with hypotheses. For the datasets that
+\noindent Skewness results were generally consistent with hypotheses. For the datasets that
 exhibited strong lower-tail (respectively, upper-tail) 
@@
 
-<!--***DAN: Dan has not edited the results for the extinction risk stuff yet-->
- We found, as we
-hypothesized, left tail dependent environmental fluctuations increased extinction risk of metapopulations than the right-tail
-dependent ones (Fig. \ref{fig_extrisk_lcohen_local}). When number of patches increased from $5$ to $25$, extinction risk 
-decreased in both cases. We repeated our simulations using same parameters with global dispersal among patches 
-and got similar results (see Fig. \ref{SM-fig_extrisk_lcohen_global}).
+As hypothesized, left-tail dependent environmental fluctuations increased metapopulation
+extinction risk for the spatially 
+extended Lewinton-Cohen model (Fig. \ref{fig_extrisk_lcohen_local}). 
+This was true for both $N=5$ and $N=25$ habitat patches, and for both local 
+(Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
+dispersal. We also examined the influence of non-N copula structure and tail dependence
+on Taylor's law in Appendix \ref{SM-Consequence_TL_simulation}.
 
@@
 \includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
-\caption{Extinction risk for Lewinton-Cohen metapopulation model was calculated after $25$ years with local dispersal
-among two nearest neighbors ($r=0$, $d=$ rate of dispersal). 
-Left-tail dependent environmental noise caused higher 
-extinction risk than right-tail dependent one for number of patches $(A)$ $N = 5$ and $(B)$ $N = 25$, respectively,
-where both case showed lower extinction risk with increasing number of patches.\label{fig_extrisk_lcohen_local}}
+\caption{Extinction risk for the metapopulation extension of the Lewinton-Cohen model, after 
+$25$ time steps, was higher for left-tail-dependent environmental noise. Dispersal was local,
+with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
+for $N=5$ (A) and $N=25$ (B).\label{fig_extrisk_lcohen_local}}
 \end{center}
 \end{figure}
-<!--***DAN: Shya, can you please change this figure as follows:
-1) Make it only go to d=0.5. I think reviewers will complain that d close to 1 is a weird and 
-unrealistic. Anyway, results are the same for larger d, so we are not hiding anything
-2) the main comparison is between right and left tail dependence, so make the left panel show 
-the numlocs=5 case, with results for both right- and left-tail dependence of the noise shown. 
-And make the right panel show the numlocs=25 case with results for both right- and left-tail
-dependence of the noise shown.
-3) Right now you use grey dots on one panel and black on the other, but you can just use black 
-on both.
-4)  Use N instead of numlocs - in papers we use variable names like in math, not like in code.
-***Shya: Done.-->
 

### 2019-02-19 SupportingInformation.Rmd
finished with a rount of editing the Q3 sections, except still need to do the TL sup mat section

@@
 
-<!-- Taylor's law section moved to Paper.Rmd -->
-
 <!--make surrogate datasets for TL analysis to follow-->
@@
 \includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
-\caption{Analogous to figure \ref{MT-fig_extrisk_lcohen_local} in maintext but with global dispersal.\label{fig_extrisk_lcohen_global}}
+\caption{Analogous to Fig. \ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal.\label{fig_extrisk_lcohen_global}}
 \end{center}
@@
 <!--How noise was generated ? code : ExtremeTailDep.R-->
+\noindent Noise was generated by first producing a 25 by $N$ matrix, $m$, of 
+independent standard-normal random draws. Then, each row of $m$ 
+was identified for a first set of modifications randomly 
+and independently with a $50\%$ chance; 
+for those rows identified, every element of the row was replaced with the 
+absolute value of the first element of the row (including the first element itself).
+For other rows, every element in the row was replaced by the negative of its
+own absolute value. This produces right-tail dependent noise, normally distributed
+for each location. If left-tail dependent noise is desired, take the negative of the
+entire matrix.
 
 <!--How dispersal matrix was generated ? code : mydispersal.R-->
-A $N \times N$ dispersal matrix ($D$) was generated assuming that the $N$ number of metapopulation patches
-were arranged in a linear chain fashion. Say, population from $i^{th}$ patch could either disperse to its nearest
-connecting patches only (we called it as local dispersal) or could disperse to all other patches 
-equally (we called it as global dispersal) at an dispersal rate, $d$. In case of 
-local dispersal matrix the diagonal entries $D_{ii}$ would be equal to $d$ and its immediate neighbors
-$D_{ii\pm j ,\;\ j=1}$ would get an equal proportion from the rest of $(1-d)$. The other distant neighbors 
-$D_{ii\pm j ,\; j>1}$ were set to zero to execute local dispersal. For global dispersal matrix, $D_{ii}$
-would be $d$ and all the other entries were equal to $(1-d)/N$. Code to generate these dispersal matrices 
-was implemented in [Dispersal_mat](https://github.com/sghosh89/BIVAN/blob/master/mydispersal.R) function.
+For local dispersal, the $N \times N$ dispersal matrix $D$ was generated by assuming, 
+for simplicity, that the $N$ habitat patches
+were arranged, evenly spaced, in a line. A fraction $d$ of each population dispersed during
+any given time step, equally distributed to the two or one nearest neighbors. 
+For global dispersal, the fraction $d$ that dispersed from each patch was divided up equally
+among all other patches.
 

### 2019-02-25 Paper.Rmd
typo fixed

@@
 We provided a test of the conceptual validity of this hypothesis using a metapopulation
-extension of the Lewinton-Cohen model,
+extension of the Lewontin-Cohen model,
 $\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
@@
 extinction risk for the spatially 
-extended Lewinton-Cohen model (Fig. \ref{fig_extrisk_lcohen_local}). 
+extended Lewontin-Cohen model (Fig. \ref{fig_extrisk_lcohen_local}). 
 This was true for both $N=5$ and $N=25$ habitat patches, and for both local 
@@
 \includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
-\caption{Extinction risk for the metapopulation extension of the Lewinton-Cohen model, after 
+\caption{Extinction risk for the metapopulation extension of the Lewontin-Cohen model, after 
 $25$ time steps, was higher for left-tail-dependent environmental noise. Dispersal was local,

### 2019-03-06 Paper.Rmd
Finished this round of edits except for the Discussion, which we are still working on. Some notes for Shya.

@@
 will tend to decreases skewness of the spatial mean. Positive skewness of the 
-spatial mean corresponds to the "spikiness" of the spatial-mean time series; negative 
-skewness, if it occurs,  corresponds to a 
-tendency to crash to low levels but quickly recover. 
+spatial mean corresponds to the "spikiness" of the spatial-mean time series. 
 The spatial-mean time series and its skewness may be quantities of 
@@
 datasets were produced from the empirical data by randomizing
-it in a special way so that surrogate/randomized dataset then had the 
+it in a special way so that surrogate/randomized dataset had the 
 copula structure of a multivariate-normal distribution
@@
 via a one-tailed test in the left tail: the $p$-value result was simply the fraction
-of surrogate skewnesses less than the skewness computed from the empirical data. For
+of surrogate skewnesses less than the skewness computed from the empirical data; the test
+examines whether lower-tail dependence made the spatial average have a significantly lower
+skewness than would have been expected without tail dependence. For
 leaf-curling plum aphid first flight data, because this dataset exhibited upper-tail 
 dependence (Table \ref{tab_multivar_summary}), we did the analogous one-tailed test
-in the right tail.
+in the right tail. The test examines whether upper-tail dependence made the spatial
+average of the first-flight data have significantly higher skewness than would have been
+expected without tail dependence.
 
@@
 
-## Results for Q3: Tail dependence influences skewness of spatial average, Taylor's law, and extinction risk
+## Results for Q3: Tail dependence influences skewness of spatial average, extinction risk, and Taylor's law
 
@@
 dispersal. We also examined the influence of non-N copula structure and tail dependence
-on Taylor's law in Appendix \ref{SM-Consequence_TL_simulation}.
+on Taylor's law in Appendices \ref{SM-Consequence_TL_simulation} and 
+\ref{SM-sect:copsurrognd}.
 
@@
 <!--what we did and why is it important to study?-->
-
+<!--***DAN: some notes by which to revise this para, Shya:
+1) no need to restate methods at all, the methods are not among our main results in this paper
+2) No real need to say much about why it is important to study what we stdied, that was done in the intro.
+Except insofar as saying what our main results are should make it clear it is an important thng to study, and see also further notes below.
+3) Some additional points below
+-->
 Most studies in ecology have focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values.
 In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field.
+<!--***DAN: Too discursive.
+We do not need to repeat that we introduced this.
+Just summarize our main results in pothy language!
+And then maybe emphasize the importance, or make any additional points about importance that could not be made in the Intro.
+This para needs to be hard-hitting and brief, saying all the important thigs we accomplished in short form, all in one place.
+See below for some examples.-->
 We showed by examples of several real data sets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables.
+<!--***DAN: How about simply: "We showed that non-N copula structure and asymmetric tail dependence are exceedingly common in datasets that range across multiple sub-disciplines of ecology." See if you can write 1-3 pithy, impactful sentences like this one for Q2, and 1-3 for Q3 as well.
+Look at the first para of Discussion for some of my recent papers to get examples of the writing techniques for how we have done this in the past.-->
 We applied two methods.
@@
 So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different nonparametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selection results.
+<!--***DAN: No need at all to summarize methods again, except in cases where the methods are 
+among the main products of the paper - not our case.
+So delete the above text.
+If you want to discuss our methods, do in later in the Discussion, not in this first paragraph, unless you may want to have a sentence saying something like "We synthesized and expanded upon methods for statistically assessing copula structure and tail dependence, and models for exploring the causes and consequences of copula strucuture in ecology and related fields."-->
 Our approach could help to interpret the synchronous behavior of a species from several meta-populations.
@@
 For example, if maximum rainfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongly correlated than lower values.
-
+<!--***DAN: I do not see how these examples go beyond the examples of the introduction. If they do, please clarify. If they don't, please remove.-->
+
+<!--***DAN: Shya, the below is more summary of results.
+Pls summarize very briefly in the first paragraph of the Discusion ONLY.
+So most of the below won't be needed, but some could be incorporated into the first Discussion para above.-->
 We explained the possible causes which induced asymmetric tail-dependence in the copula.
@@
 
-<!--future work-->
-We think this copula approach have scope to investigate into the extinction risks of meta-population due to climate change.
-To illustrate, lower-tail dependent environmental fluctuations may suppress the meta-populations synchronously by making some of the habitats go extinct.
-This could help in managing policy for conservation of endangered populations.
-
 \clearpage

### 2019-03-06 SupportingInformation.Rmd
Finished this round of edits except for the Discussion, which we are still working on. Some notes for Shya.

@@
 
-pdf("./Results/taylorslaw_results/SpatialTL_CNFSC_kendall.pdf",height=4,width=16)
+#pdf("./Results/taylorslaw_results/SpatialTL_CNFSC_kendall.pdf",height=4,width=16)
+png("./Results/taylorslaw_results/SpatialTL_CNFSC_kendall.png",width=16,height=4,units="in",res=400)
 op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
@@
 
-pdf("./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.pdf",height=4,width=16)
+#pdf("./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.pdf",height=4,width=16)
+png("./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png",height=4,width=16,units="in",res=400)
 op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
@@
 \begin{center}
-\includegraphics[width=18 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.pdf}\\
+\includegraphics[width=18 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png}\\
 \end{center}
@@
 \includegraphics[width=18 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
-\caption[Spatial Taylor's Law results]{Spatial Taylor's Law results: log(spatial variance) vs.
-log(spatial mean) for $ns=1000$ number of (A) Clayton, (B) Normal, (C) Frank, (D) Survival Clayton surrogates; (E) p-value for test of linearity, (F) p-value for test of homoskedasticity, (G) root mean square error, (H) intercept, (I) slope, (J) quadratic coefficient, (K) mean curvature of regression line from first panel-plots.
-\label{fig_spTL}}
+\caption[Spatial Taylor's law results]{Spatial Taylor's law results. Log(spatial variance) vs. 
+log(spatial mean) relationships for $1000$ simulations over $50$ time steps for $25$ locations 
+using Clayton (A), normal (B), Frank (C), and survival Clayton (D) copula structures between
+locations. Results from all $1000$ 
+simulations were plotted on the same axes for each panel, but two example simulations are 
+shown in red and green to help assess to what extent variation on the plots was between 
+simulations or within simulations. For each simulation, we tested the linearity 
+(E) and homoskedasticity (F) of the log(variance) vs. log(mean) plot for that simulation, 
+and quantified the root mean squared error of points from the linear regression line (G).
+We also calculated the linear regression intercept (H) and slope (I), the quadratic
+term of a quadratic regression through the points on the log(variance) vs. log(mean) plot
+(J), and the curvature of that quadratic regression (K). Distributions of values 
+across all 1000 simulations are displayed. Spearman-preserving surrogates were used, 
+though results using Kendall-preserving surrogates were very similar. See 
+Appendix \ref{Consequence_TL_simulation} for 
+simulation and plotting details.\label{fig_spTL}}
 \end{center}
@@
 - Calculate the covariance matrix of a multivariate normal distribution with standard-normal
-marginals and with pairwise Spearman (or Kendall) correlations equal to the $\tau_{ij}$. This is possible
-using the `iRho` function (or the `iTau function) of the `VineCopula` package
+marginals and with pairwise Spearman (or Kendall) correlations equal to the $\tau_{ij}$. This is possible.
+using the `iRho` function (or the `iTau` function) of the `VineCopula` package
 - Generate data $a_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$ by taking $T$ independent draws
-from the the multivariate normal distribution of the previous step.
+from the multivariate normal distribution of the previous step.
 - For any $i,t$ for which the datum $x_i(t)$ was missing, replace $a_i(t)$ by `NA`.
@@
 \noindent Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
-tested whether the copula structure of dependence between population variables 
+tested whether the copula structure of the dependences between population variables 
 measured in different locations influences the spatial version of Taylor's
@@
 the Taylor's law *exponent* or 
-*slope*. There are several versions of Taylor's law. For spatial Taylor's law,
+*slope*, and we refer to $\log(a)$ as the Taylor's law intercept. 
+There are several versions of Taylor's law. For spatial Taylor's law,
 given population density or abundance data $x_i(t)$ measured in locations
@@
 all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
-So means and variances are computed across space. Taylor's law has been 
+So means $m_g$ and variances $v_g$ are computed across space. We view the assumptions
+of Taylor's law to be validated for a dataset if the 
+$\log(v_g)$ versus $\log(m_g)$ scatter plot 
+for the dataset shows a linear, homoskedastic pattern.
+Taylor's law has been 
 verified empirically for hundreds of taxa and has been applied in numerous fields
@@
 
-We examined using simulations whether copula structure influences spatial Taylor's law as 
+We examined using simulations whether copula structure influences spatial Taylor's law, as 
 follows. First, an $n \times n$ covariance matrix $\Sigma$ was created
@@
 distribution with mean vector $(0,\ldots,0)$ and covariance matrix $\Sigma$. These 
-were organized into an $N \times n \times 1000$ array representing $1000$ simulations of 
-populations measured in $n$ locations over $N$ time steps. 
+were organized into an $N \times n \times 1000$ array representing $1000$ replicate 
+simulations of (temporally independent, in this example) 
+spatio-temporal population dynamics measured in 
+$n$ locations over $N$ time steps. We used $N=50$.
 The cumulative distribution function
@@
 inverse of the cdf of a gamma distribution with shape and scale parameters $7.5$ and $1$. 
-These transformations made gamma distributed the marginal distributions of 
-each population sampling location in each simulation. 
+These transformations made gamma distributed the marginal distributions of populations in 
+each sampling location for each simulation. 
 A skewed population marginal distribution is an 
 important factor in Taylor's law [@Cohen2015; @Reuman2017]. We refer henceforth to the 
-resulting $N \times n \times 1000$ array as our *N array* for this simulation exercise;
-copula structure of dependences between populations in different sampling locations
-was normal (N) for the simulations in this array.
-
-We then considered an $n$-dimensional C copula with parameter chosen so that the 
-Kendall correlations between each pair of the $n$ components were the same as those
-of the N distribution constructed above.
-
-the nature of the slope and how it changed if population time series (taken as $50$ years) from different locations (say, $numloc=25$) had some certain type of copula structure (namely, \textbf{C}, \textbf{F} and \textbf{SC}) other than the $\textbf{N}$ormal copula.
-First, we simulated (N $\times$ n $\times$ nr) dimensional normal copula with co-variance $\rho =0.7$, where $(N=50,n=25,nr=10)$ are the length of time series, number of sampling locations and number of replica for that copula, respectively.
-Then we generated Kendall-correlation preserving \textbf{C}layton, \textbf{F}rank and $\textbf{S}$urvival \textbf{C}layton surrogates with same dimension of \textbf{N}ormal.
-With these four different types of copula we computed `log`(spatial variance) and `log`(mean) and plotted in Fig.\ref{fig_spTL}(A-D).
-We also tested statistically that whether linearity, homoskedasticity is maintained in each copula data set or not and plotted our results along with root mean square error, intercept and slope in Fig.\ref{fig_spTL}(E-I).
-We did a quadratic regression with each simulated copula for all types and measured quadratic coefficient and mean curvature for the regression line which is plotted in Fig.\ref{fig_spTL}(J-K) (For code see [`copsurrognd.R`](https://github.com/sghosh89/BIVAN/blob/master/copsurrognd.R) and [`TaylorsLawTools.R`](https://github.com/sghosh89/BIVAN/blob/master/TaylorsLawTools.R)).
-
-<!--
-Curvature calculations:
-***DAN: Shya, I need a reminder how we did this, can you please tell me where the
-code is, then I can write this section. Feel free to take a stab at it yourself, too,
-though.
-***Shya: What do you want to know? how did you calculate mean curvature and 
-slope for made up data or you want to see code with real data?
-
-In case of made up data :
-For each non-N copula family (C,F,SC) you generated surrogates
-having similar correlation (spearman/ kendall, though for 
-consistency we added plots in suppmat with spearman) that of N copula
-[using copsurrognd.R] : this thing is done in r chunk named "taylors_law_make_surrogs".
-
-Then you did spatial-TL calculation for each set of copula family in r chunk "tl_analysis"
-[using TaylorsLawTools.R which calls mncurvat.R code to calculate mean curvature]
-
-Next two r chunks : taylor_plots_kendall and taylor_plots_spearman generates plot with those data.
-
-For spatial TL calculation on real data :
-
-We used Aphid count all sp with spearman only, but to show a plot for example chose sp=6 (corn leaf aphid)
-as it had significant p_left value for mean curvature.
-code is in r chunk named "spatialTL_on_realdata_aphid_count".
--->
-
-<!--
-Stuff on TL cut from the main text
-From Q3 methods:
-We tested whether the linearity of the spatial Taylor's law relationship was partly
-due to the copula structure of the dependences between the populations sampled. 
-Using corn leaf aphid abundance data as an example, but also
-applying the same analysis to the abundance data we had for 19 other aphid species
-(see *Data*), we calculated the curvature (details in Appendix \ref{SM-TL_curvature}) 
-of the empirical Taylor's law relationship; Taylor's law hypothesizes that this 
-curvature should be close to 0. We then calculated the same curvature quantity for
-10000 surrogate datasets, using the same surrogate procedure described above in 
-this section, and compared the empirical Taylor's law curvature to the 
-distribution of surrogate curvatures using a two-tailed test. If the empirical
-curvature value was below the $5^{\text{th}}$ quantile of the surrogate distribution, 
-then copula structure contributes to the linearity of spatial Taylor's law 
-because randomizing copula structure while retaining other statistical properties
-of the data cause the Taylor's law relationship to become significantly more 
-curved.
-
-From Q3 results:
-```{r read_tl_aphid_count_sp6, echo=F}
-tlsp6<-readRDS("./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
-tl_aphid_count<-readRDS("./Results/taylorslaw_results/tl_aphid_count.RDS")
-aphid_info<-readRDS("./Results/aphid_info.RDS")
-pleft<-tl_aphid_count$p_left_mncurv
-stat<--2*sum(log(pleft))
-pchi2<-1-pchisq(stat,40)
-idsp<-which(pleft<0.05)
-pleftgood<-pleft[idsp]
-```
-
-Results also suggested that copula structure contributes to the linearity of 
-spatial Taylor's law. For instance, the spatial Taylor's law relationship 
-for corn leaf aphid abundance data was significantly less curved than for 
-surrogate datasets (Fig. \ref{fig_sTL_mncurv_sp6}). Three of $20$ aphid species
-had cuvatures less than at least $95\%$ of surrogate datasets ($p=$ `r pleftgood` for 
-the `r aphid_info$c1[idsp]` species, respectively). 
-***DAN: Shya, please fill in X, Y, Z, A, B, C with autolinks)
-***Shya: Done.
-Three such significant
-results of 20 species could have occurred by chance due to 
-multiple-testing error. However, for each of the 20 species, we computed the fraction
-of surrogate curvatures less than the curvature that came from real data, and we
-combined these $p$-values using Fisher's combined probability test to get an
-aggregate $p$-value of `r pchi2`
-***DAN: Shya, see the Wikipedia article on Fisher's method to learn about it. Then 
-you need to put this analysis into the code suite and have the resulting p value
-typed in above be an autolink instead. Here is what I did: 
-1) I copied by hand all the pleft values on the pdf you sent me into a vector pleft. Instead you should create such a vector in your code from your test output, not typing in the results.
-2) I used the formula stat<--2*sum(log(pleft)) to get the Fisher method statistic
-3) I compared to a chi-squared distribution with 40 degress of freedom using 1-pchisq(stat,40) to get my p-value
-Please let me know if any of this is unclear.
-***Shya: Done.
-for the test of the null hypothesis that 
-copula structure did not contribute to linearity of Taylor's law for any of the 
-species. This result is usefully suggestive and can guide future work, but still does not
-provide certainty, because the Fisher method provides $p$-values that are too 
-small if its assumption of independent tests is violated. Though our 20 species
-have different host plants so may be independent for the purposes of our analyses, 
-it would be difficult to fully assess independence.
-***DAN: Shya, can you remind me why the same analyses were not done for plankton,
-if there is a good reason? Or if you did it, what were the results?
-***Shya : we did it, I have the pdf but I don't have the code where I called to generate the pdf,
-As we could not get any significant results for a single species (see attachment in the mail I sent you)
-we did not show the results. 
-We provide simulation results in Appendix \ref{SM-Consequence_TL_simulation}
-that robustly illustrate, at least in the model setup we used, 
-that copula structure influences spatial Taylor's law.
-
-\begin{figure}[!h]
-\textbf{ \hspace{4 cm} (A) \hspace{4 cm} (B) \hspace{4 cm} (C)} \\
-\vspace{-0.5 cm}
-\begin{center}
-\includegraphics[width=10 cm]{./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf}
-\includegraphics[width=5 cm]{./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf}\\
-\caption{Mean curvature of the spatial Taylor's law relationship for 
-corn leaf aphid abundance
-data (A) was significantly less than for surrogates (B, C). The Taylor's law 
-relationship for one example surrogate (B) shows shows greater curvature than
-for data (A), and the significance of the difference was demonstrated
-by the fact that only `r (tlsp6$p_left_mncurv)*10000` of $10000$ surrogates had curvature less than that
-of the data (C). See Appendix \ref{SM-TL_curvature} for details on 
-how curvature was calculated.\label{fig_sTL_mncurv_sp6}}
-\end{center}
-\end{figure}
-***DAN: Shya, pls make the x axis label on C read simply "Mean curvature"
-***Shya: dONE.
--->
+resulting $N \times n \times 1000$ array as our *N-copula population array* 
+for this simulation exercise;
+copula structure of dependences between populations in different locations
+was normal (N) for this array.
+
+We then generated an series of arrays that were the same dimensions ($N \times n \times 1000$)
+as the N-copula population array, and that were statistically similar to 
+the N-copula population array in all ways except the new
+arrays were constructed to have C (respectively, F, SC) copula structure for 
+the dependences between 
+populations in different sampling locations within a simulation. We 
+refer to the new arrays as our
+*C-copula*, *F-copula*, and *SC-copula population arrays*, respectively. The C-copula array was
+constructed as follows; the others were constructed analogously. First, we constructed an
+$n$-dimensional C copula with parameter chosen so that the 
+Spearman correlations between each pair of the $n$ components were the same as those
+of the N distribution constructed above. This was done using the `rho` and `iRho` functions 
+in the `copula` package in R. Because the C copula is a one-paremeter family, this 
+step relies on the fact that all pairs of locations were 
+equally correlated with each other in the original multivariate normal distribution
+constructed above. We then randomized each simulation in 
+the N-copula population array in a special way 
+(see Appendix \ref{sect:copsurrognd}) so that the resulting 
+surrogate/randomized dataset had the copula structure
+of the C copula just constructed. The randomization procedure worked 
+by permuting time series of values in the N-copula array, separately for each 
+sampling location and simulation. So population marginal distrbutions for each location
+and simulation were *exactly* the same as those in the N-copula population array. 
+Permutations were done in such a way
+that resulting pairwise Spearman correlations were the same as the C-copula 
+constructed above, and hence were the same as the N-copula array, 
+except for sampling variation. Thus, to summarize using mathematical notation,
+if $m^{(\text{N})}$ is our N-copula population array and $m^{(\text{C})}$ is our 
+C-copula population array, then,
+for any given $i$ and $j$, the sets $\{m^{(\text{N})}_{tij}: t=1,\ldots,N\}$ and 
+$\{m^{(\text{C})}_{tij}: t=1,\ldots,N\}$ are exactly the same (when considered 
+as unordered sets); 
+and for any given $i_1$, $i_2$ and $j$, the Spearman correlations $\cor_t(m^{(\text{C})}_{ti_1j},m^{(\text{C})}_{ti_2j})$
+and $\cor_t(m^{(\text{N})}_{ti_1j},m^{(\text{N})}_{ti_2j})$ are the same, to within 
+sampling variation. The whole exercise was repeated preserving Kendall instead of 
+Spearman correlations.
+
+A variety of Taylor's law statistics were computed for each simulation $m^{(X)}_{tij}$ 
+($i=1,\ldots,n$, $t=1,\ldots,N$) for each copula structure $X$ used. First, the $N=50$ 
+spatial means,
+$m_t$, and variances, $v_t$, were computed, and $\log(v_t)$ vs. $\log(m_t)$ plots 
+were considered. As a shorthand, denote $y=\log(v_t)$ and $x=\log(m_t)$ for a 
+single simulation. Linearity of the $y$ vs. $x$ relationship was 
+tested for each simulation by comparing the linear regression
+through these $50$ points to a quadratic alternative via an $F$-test, 
+producing a $p$-value result. If 
+these $p$-values were uniformly distributed across the unit interval for the 1000 replicate 
+simulations, it supported the linearity assumption of Taylor's law; whereas if 
+they were clustered toward smaller values
+the test tended to reject that assumption. We also tested, for each simulation, 
+the assumption of Taylor's 
+law that the $y$ vs. $x$ plot was homoskedastic: we regressed the absolute residuals of the 
+linear regression of $y$ versus $x$ against the predictions of that regression. 
+A significant $p$-value
+result of this test indicates heteroskedasticity. For each simulation we also computed the
+root mean squared error of $y$ vs. $x$ data from the linear regression
+of $y$ versus $x$, as well as the intercept and slope of the linear regression.
+Finally we recorded the quadratic coefficient of the regression of $y$ against
+$x$ and $x^2$, and the mean curvature of the quadratic regression equation itself
+across the values $x$. The distrubutions of all these statistics across 1000 replicate 
+simulations were compared for the C, N, F, and SC simulations to determine if copula structure 
+influences Taylor's law.
+
+Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
+linearity and homoskedasticity, by non-N copula structure. For N copula structure 
+(i.e., for simulations in our N-copula population array), $p$-values for the linearty
+and homoskedasticity tests were roughly uniformly distributed across replicate 
+simulations and Taylor's law appeared visually to be a reasonable approximation
+of the $\log(v_t)$ versus $\log(m_t)$ relationship (Fig. \ref{fig_spTL}B,E,F). 
+Furthermore, quadratic coefficients
+and curvature values were close to $0$ (Fig. \ref{fig_spTL}J,K), and root mean squared
+errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
+But linearity or homoskedasticity were violated more frequently for non-N copula structure
+(Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
+non-zero (Fig. \ref{fig_spTL}J,K); and root mean squared
+errors from the linear regression were much higher (Fig. \ref{fig_spTL}G). Slopes 
+and intercepts of the linear regression were also strongly affected by copula
+structure (Fig. \ref{fig_spTL}H,I), though some of the effect here was because
+linear regressions do not always adequately represent the $\log(v_t)$ versus $\log(m_t)$
+relationship when copula structure was not N.
+
+# Another tool for producing surrogates for multivariate datasets\label{sect:copsurrognd}
+
+\noindent Suppose one has data $x_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$, with
+no missing values and no ties, and an
+$N$-dimensional copula $C$. The following algorithm produces a surrogate dataset
+$y_i(t)$ ($i=1,\ldots,N$, $t=1,\ldots,T$) such that: 1) the Kendall and Spearman 
+correlations $\cor_t(y_i(t),y_j(t))$ are the same,
+up to sampling variation, as the Kendall and Spearman correlations, respectively, of the corresponding
+components of $C$, for all $i\neq j$; and 2) for any $i$, the values of the unordered set 
+$\{y_i(t):t=1,\ldots,T\}$ are the same as the values of the unordered set
+$\{x_i(t):t=1,\ldots,T\}$.
+
+- Generate data $a_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$ by taking $T$ independent 
+draws from $C$.
+- For each $k$ and $i$, replace the $k$th-smallest element of the time series $a_i(t)$, 
+$t=1,\ldots,T$ by the $k$th smallest element of $x_i(t)$, $t=1,\ldots,T$. Call the 
+result $y_i(t)$. This is the surrogate dataset.
 

### 2019-03-07 Paper.Rmd
revised the abstract

@@
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, P. Chris Reid, Daniel C. Reuman"
-date: "December 16, 2018"
+date: ""
 geometry: "left=1.5cm,right=1.5cm,top=2.5cm,bottom=2.8cm"
@@
 
-Reid:
-
-Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA.
+Reid: Marine Institute, Plymouth University, Drake Circus, Plymouth, United Kingdom; Mrine Biological Association of the UK, The Laboratory, Citadel Hill, Plymouth, United Kingdom
+
+Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA
 
@@
 
-We introduce a new rank based "copula" approach to model dependencies between ecological variables.
-Our method could overcome the shortcomings of the linear correlation approach and give us new insight about "asymmetric tail dependence" found in ecology, bio-geochemistry and related field.
-We focus on three objectives (1) Do data sets in ecology and related fields have non-Normal copula structure?
-Do they show tail dependence distinct from that of a Normal copula, and in particular do they show asymmetric
-tail dependence? (2) If so, what are some possible causes of non-Normal copula structure? And (3) what are the
-consequences of non-Normal copula structure and tail dependence for our understanding of biology and what are the applications?
-We present several examples with real data which show asymmetric tail dependence in bivariate as well as multivariate non-Normal copulae.
-Asymmetric tail dependence can illuminate some important phenomena like "spatial synchrony" and causal relationships between ecological variables (like species abundance and rainfall).
-We show that the "Moran-effect" and asymmetric sensitivity towards global environmental fluctuations could be the possible causes for non-Normal copula structures.
-We demonstrate with simulations and real data that non-Normal copulae have profound effects on spatial skewness of species abundances and spatial Taylor's law slope values.
-Our method is new and could have potential when exploring some other related phenomena like effects of climate change on extinction risks of meta-populations.
+All branches of ecology and related areas study relationships among environmental and biological variables.
+However, ubiquitously used approaches to studying such relationships, which are typically based on correlation and regression, are limited, providing only a small slice of the complex information content of the dependence.
+Well-developed statistical tools do exist that provide a demonstrably complete description of the dependence between two or more variables.
+The tools are based on the concept of the *copula*, and have been applied to great benefit in finance, neuroscience, and other fields, but not in ecology.
+We here examine the potential for copula approaches to improve our understanding across multiple disciplines of ecology, using environmental, species-trait, phenological, population, community, and ecosystem functioning data, as well as models that can be widely applied across ecology.
+We find that informative copula structure of dependence between variables is exceedingly common across all the disciplines we considered.
+An important feature in numerous datasets is asymmetric tail dependence, whereby two variables are more strongly related in their left compared to right tails, or vice versa.
+Copula structure and tail dependence can help reveal causal relationships between variables.
+Via models, we describe two mechanisms by which observed copula structures and tail dependence can arise in ecological data: a Moran-like effect, whereby dependence structures between environmental variables are inherited by ecological variables influenced by the environmental variables; and via asymmetric or nonlinear influences of environmental variables on ecological variables, such as under Liebig's law of the minimum.
+Via models and data, we describe important consequences of copula structure for applications and our understanding of ecology, including: impacts of copula structure on the stability through time of ecosystem service provisioning and spatially averaged population levels; and impacts on extinction risk.
+By documenting the importance of a complete statistical description of dependence, advancing conceptual frameworks, and demonstrating powerful methods, we aim to help elevate copula approaches to widespread use in studying ecological phenomena.
 

### 2019-03-07 Paper.Rmd
commenting out the paragraph for discussion as of now

@@
 copula approaches for understanding dependences in ecology by addressing the 
-folling specific questions. (Q1) Do datasets in ecology and related fields 
+following specific questions. (Q1) Do datasets in ecology and related fields 
 have copula structure distinct from that of a Gaussian/normal distribution (here called 
@@
 "multivariate" in that they comprise, for each taxon, measurements at multiple locations; 
-copula structure can be studied for each pair of locations. Likewise the the ecosystem 
+copula structure can be studied for each pair of locations. Likewise the ecosystem 
 functioning (methane flux) dataset was  multivariate in this sense.
@@
 
-The first part Q1 was answered for a bivariate dataset 
+The first part of Q1 was answered for a bivariate dataset 
 by comparing the AIC weights of the non-normal copula
@@
 or vice-versa, and reported the percentage of pairs showing stronger lower-tail dependence out of 
-the total number non-independent location pairs considered.
+the total number of non-independent location pairs considered.
 
@@
 (Table \ref{tab_bivar_summary}, rows 3-4), answering the first part of Q1. 
-Upper- and lower-tail statistics (Table \ref{tab_bivar_summary}, rows 7-8) 
+Lower- and upper-tail statistics (Table \ref{tab_bivar_summary}, rows 7-8) 
 differed substantially from 0, the tail dependence statistics of 
@@
 
-dt <- tibble(Description = c("1. P-value, independence test",
+dt <- tibble(Description = c("1. p-value, independence test",
            "2. Minimum-AIC copula",
@@
            "4. AIC weight of N copula",
-           "5. P-value, Cramer-von Mises goodness of fit test",
-           "6. P-value, Kolmogorov-Smirnov goodness of fit test",
+           "5. p-value, Cramer-von Mises goodness of fit test",
+           "6. p-value, Kolmogorov-Smirnov goodness of fit test",
            "7. Model averaged lower-tail dependence",
@@
 We initially assumed the variable follows AR(1) dynamics within a location, 
-$E_i(t+1)=\beta E_{i}(t-1)+ \sqrt{1-\beta^2} \epsilon_i(t)$. 
+$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$. 
 We used $\beta=0.5$. Here the $\epsilon_i(t)$ represent environmental noise; they are 
@@
 set to $0$. We simulated the model 10000 times for each combination of parameters,
-starting from initial populations in each patch of X, 
+starting from initial populations $p_0 = 50$ in each patch of $N$, 
 <!--***DAN: Shya, please fill in X-->
@@
 -->
+<!--
 Most studies in ecology have focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values.
 In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field.
+-->
 <!--***DAN: Too discursive.
@@
 See below for some examples.-->
-We showed by examples of several real data sets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables.
+
+<!--We showed by examples of several real data sets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables.-->
+
 <!--***DAN: How about simply: "We showed that non-N copula structure and asymmetric tail dependence are exceedingly common in datasets that range across multiple sub-disciplines of ecology." See if you can write 1-3 pithy, impactful sentences like this one for Q2, and 1-3 for Q3 as well.
 Look at the first para of Discussion for some of my recent papers to get examples of the writing techniques for how we have done this in the past.-->
-We applied two methods.
+<!--We applied two methods.
 First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family.
@@
 So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different nonparametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selection results.
+-->
 <!--***DAN: No need at all to summarize methods again, except in cases where the methods are 
@@
 If you want to discuss our methods, do in later in the Discussion, not in this first paragraph, unless you may want to have a sentence saying something like "We synthesized and expanded upon methods for statistically assessing copula structure and tail dependence, and models for exploring the causes and consequences of copula strucuture in ecology and related fields."-->
-Our approach could help to interpret the synchronous behavior of a species from several meta-populations.
+<!--Our approach could help to interpret the synchronous behavior of a species from several meta-populations.
 To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g.
@@
 For example, if maximum rainfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongly correlated than lower values.
+-->
 <!--***DAN: I do not see how these examples go beyond the examples of the introduction. If they do, please clarify. If they don't, please remove.-->
@@
 So most of the below won't be needed, but some could be incorporated into the first Discussion para above.-->
-We explained the possible causes which induced asymmetric tail-dependence in the copula.
+<!--We explained the possible causes which induced asymmetric tail-dependence in the copula.
 Borrowing the idea from "Moran's theorem", we demonstrated with simulations that two habitat patches when affected by bivariate noise with an asymmetric copula structure showed similar tail-dependence asymmetry in the population copula from those patches (see Fig.
@@
 Similarly, methane fluxes showing lower skewness than Normal surrogates mean that the high emissions are un-correlated between locations and monitoring this skewness over several decades may be helpful to understand the "green-house effect" on environment.
+-->
 

### 2019-03-07 Paper.Rmd
added acknowledgements

@@
 set to $0$. We simulated the model 10000 times for each combination of parameters,
-starting from initial populations $p_0 = 50$ in each patch of $N$, 
-<!--***DAN: Shya, please fill in X-->
+starting from initial populations $p_0 = 50$ in each patch, 
 and calculated extinction risk after $25$ time steps.
@@
 
+# Acknowledgments
+
+We thank the many contributors to the large datasets we have used; D.
+Stevens and P.
+Verrier for data extraction; and Jonathan Walter, Thomas Anderson, and Lei Zhao for helpful suggestions.
+We especially thank James Bell of the Rothamsted Insect Survey (RIS).
+The RIS is a National Capability of the United Kingdom, funded by the Biotechnology and Biological Sciences Research Council.
+SG, LWS and DCR were partly funded by US National Science Foundation grants 0075516 and 1442595, and the James S McDonnell Foundation.
+
 \clearpage

### 2019-03-08 Paper.Rmd
included changes @Bell's recommendation

@@
 35 years from 1976 to 2010. These time series were computed from the the Rothamsted Insect 
-Survey suction-trap dataset [@Macaulay1988; @Harrington2014]. The first of our two 
+Survey suction-trap dataset [@Harrington2014; @Bell2015]. The first of our two 
 population-level datasets was also derived from the RIS suction-trap data, and comprised 
 total counts of aphids trapped for the same locations, species, and years; these are 
-indicators of flying aphid abundance. Data pre-processing for the aphid data was the same as 
+indicators of aphid migration counts. Data pre-processing for the aphid data was the same as 
 that of @Sheppard2016,
@@
 We especially thank James Bell of the Rothamsted Insect Survey (RIS).
-The RIS is a National Capability of the United Kingdom, funded by the Biotechnology and Biological Sciences Research Council.
+The Rothamsted Insect Survey, a National Capability, is funded by the Biotechnology and Biological Sciences Research Council under the Core Capability Grant BBS/E/C/000J0200.
 SG, LWS and DCR were partly funded by US National Science Foundation grants 0075516 and 1442595, and the James S McDonnell Foundation.

### 2019-03-08 Paper.Rmd
Discussion : 1st para : summary added

@@
 
-<!--what we did and why is it important to study?-->
+We showed that non-N copula structure and asymmetric tail dependence are 
+exceedingly common in datasets that range across multiple sub-disciplines
+of ecology. Via basic mathematical models, we hypothesized two probable
+mechanisms causing those non-N copula structures with asymmetric tail-dependence. One
+mechanism was "Moran"-like effect where dependence structure between ecological
+variables from two locations was shaped by the dependence structure of 
+the environmental variables at those patches. Second mechanism, a mimic of 
+Liebig's law of the minimum, explored
+the asymmetric influence of environmental variables which 
+could limit the growth of populations when below (or beyond)
+a certain threshold. Using model simulations and real ecological data,
+we concluded asymmetric tail dependence feature of 
+non-N copula could affect (1) the skewness (henceforth, the stability)
+of spatially averaged quantities over years, (2) the extinction risk 
+and (3) the spatial Taylor's law slope of metapopulations.
+
 <!--***DAN: some notes by which to revise this para, Shya:
@@
 Look at the first para of Discussion for some of my recent papers to get examples of the writing techniques for how we have done this in the past.-->
+
 <!--We applied two methods.

### 2019-03-12 SupportingInformation.Rmd
typo fixed

@@
 $\mathcal{B}(\theta_\tau)$ has Kendall correlation $\tau$ (this is possible for the one-parameter copula families of this 
-study using the `iTau` function of the `VineCopula` package). 
+study using the `iTau` function of the `copula` package). 
 - Generate data $(a_i,b_i)$, $i=1,\ldots,n$ from 
@@
 marginals and with pairwise Spearman (or Kendall) correlations equal to the $\tau_{ij}$. This is possible.
-using the `iRho` function (or the `iTau` function) of the `VineCopula` package
+using the `iRho` function (or the `iTau` function) of the `copula` package
 - Generate data $a_i(t)$ for $i=1,\ldots,N$ and $t=1,\ldots,T$ by taking $T$ independent draws

### 2019-03-18 Paper.Rmd
thinkig about Discussion

@@
 
+<!--what we did and why is it important to study?-->
 We showed that non-N copula structure and asymmetric tail dependence are 

### 2019-03-21 Paper.Rmd
added place to brainstorm discussion para topics
general edits to paper

@@
 
-All branches of ecology and related areas study relationships among environmental and biological variables.
-However, ubiquitously used approaches to studying such relationships, which are typically based on correlation and regression, are limited, providing only a small slice of the complex information content of the dependence.
-Well-developed statistical tools do exist that provide a demonstrably complete description of the dependence between two or more variables.
-The tools are based on the concept of the *copula*, and have been applied to great benefit in finance, neuroscience, and other fields, but not in ecology.
-We here examine the potential for copula approaches to improve our understanding across multiple disciplines of ecology, using environmental, species-trait, phenological, population, community, and ecosystem functioning data, as well as models that can be widely applied across ecology.
-We find that informative copula structure of dependence between variables is exceedingly common across all the disciplines we considered.
-An important feature in numerous datasets is asymmetric tail dependence, whereby two variables are more strongly related in their left compared to right tails, or vice versa.
+All branches of ecology study relationships among environmental and biological variables.
+However, ubiquitously used approaches to studying such relationships, based on correlation and regression, provide only a small slice of the complex information content of the relationships.
+Statistical tools do exist that provide a complete description of relationships, based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, but not in ecology.
+We here examine the potential for copula approaches to improve our understanding of ecology.
+We find that informative copula structure in dependences between variables is exceedingly common across all the environmental, species-trait, phenological, population, community, and ecosystem functioning dataset we considered.
+Many datasets exhibited asymmetric tail dependence, whereby two variables are more strongly related in their left compared to right tails, or vice versa.
 Copula structure and tail dependence can help reveal causal relationships between variables.
-Via models, we describe two mechanisms by which observed copula structures and tail dependence can arise in ecological data: a Moran-like effect, whereby dependence structures between environmental variables are inherited by ecological variables influenced by the environmental variables; and via asymmetric or nonlinear influences of environmental variables on ecological variables, such as under Liebig's law of the minimum.
-Via models and data, we describe important consequences of copula structure for applications and our understanding of ecology, including: impacts of copula structure on the stability through time of ecosystem service provisioning and spatially averaged population levels; and impacts on extinction risk.
-By documenting the importance of a complete statistical description of dependence, advancing conceptual frameworks, and demonstrating powerful methods, we aim to help elevate copula approaches to widespread use in studying ecological phenomena.
+We describe two mechanisms by which observed copula structure and tail dependence can arise in ecological data: a Moran-like effect whereby dependence structures between environmental variables are inherited by ecological variables; and asymmetric or nonlinear influences of environments on ecological variables, such as under Liebig's law of the minimum.
+We describe consequences of copula structure, including impacts on extinction risk and the stability through time of ecosystem services.
+By documenting the importance of a complete statistical description of dependence, advancing conceptual frameworks, and demonstrating powerful methods, we aim to elevate copula approaches to widespread use in ecology.
 
@@
 Very generally, all branches of ecology and related areas study relationships among environmental and biological 
-variables. However, commonly used correlation- and regression-based 
+variables. However, commonly used correlation and regression 
 approaches to studying 
-such relationships are limited - it is well understood that common statistical
+such relationships are limited - common statistical
 approaches provide only a very partial description of a relationship between variables. 
-For instance, data sets showing markedly different 
+For instance, datasets showing markedly different 
 relationships may nevertheless have the same correlation coefficient value 
@@
 related in the left (respectively, right) portions of their distributions, an asymmetric 
-pattern of dependence in the distribution tails that may
+pattern of dependence that may
 have ecological significance, as discussed below, but is not captured by 
@@
 which alters the Pearson, but not the Spearman correlation. Although correlation
-is only one of the numerous metrics for 
-studying relationships between variables, other commonly used metrics also 
-provide partial information only.
+is only one of numerous metrics for 
+studying relationships between variables, other common metrics also 
+provide only partial information.
 
@@
 \includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
-\caption{(A-F, I, J) Bivariate data sets showing diverse relationships between 
-variables; many of these data sets nevertheless have the same Pearson (P)
+\caption{(A-F, I, J) Bivariate datasets showing diverse relationships between 
+variables; many of these datasets nevertheless have the same Pearson (P)
 or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
@@
 with Fig. \ref{fig_cop_pedag1}E-F are in Fig.
-\ref{fig_cop_pedag1}G-H, respectively. Strictly speaking, 
-the copula is actually defined as the bivariate 
-cdf of $(F_X(X),F_Y(Y))$ rather than as this random variable itself, 
-but this makes little conceptual difference. See @nelsen2006_copula and @joe2014_dependence
-for mathematical details. An important theorem of A. Sklar [@sklar1959_theorem] 
+\ref{fig_cop_pedag1}G-H, respectively. 
+See @nelsen2006_copula and @joe2014_dependence
+for general background. An theorem of A. Sklar 
 states that copulas give a unique and complete 
@@
 between variables, in the sense that the joint distribution of $(X,Y)$ 
-can be reconstructed completely from the 
-copula and the marginal distributions. 
+can be reconstructed from the 
+copula and the marginal distributions [@sklar1959_theorem]. 
 Copula theory is also well developed for
@@
 
-A principal benefit of a copula approach to ecological data is 
-that such an approach can sensitively model and detect
+A principal benefit of a copula approach is 
+that it can sensitively detect and model
 tail dependence and asymmetries of tail dependence. 
-Tail dependence is association between extreme values of two variables. 
+Tail dependence is association between extreme values of variables. 
 If smaller values of two positively associated variables are more 
@@
 and \emph{vice versa} if larger values are more strongly associated than smaller
-ones (Fig. \ref{fig_cop_pedag1}B, F). Data sets of the same 
+ones (Fig. \ref{fig_cop_pedag1}B, F). Datasets of the same 
 correlation can have a range of tail dependence 
-characteristics (Fig. \ref{fig_cop_pedag}), which can be quantified (see section on
+characteristics (Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see section on
 methods for Q1).
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
-\caption{(A-C) Data generated using the BB1 copula (see section on methods for Q1) 
-with three different 
-sets of parameters chosen to produce a variety of 
-lower-tail (LT) and upper-tail (UT) dependence strengths (methods for Q1), but the same 
-Kendall correlation (K) (to two digits) in all cases. (D-F) Similar to A-C, but
-with stronger correlation. Asymmetry of tail dependence (LT-UT) is the same 
-within each column of panels.\label{fig_cop_pedag}}
-\end{center}
-\end{figure}
-
 One way we may expect a copula approach 
-to better reflect ecological relationships has to do with Liebig's law of the
+to better reflect ecological relationships relates to Liebig's law of the
 minimum. If, for instance, the growth of a plant depends on soil 
 nitrogen, N, as well
-as on other factors, data on soil nitrogen and resulting growth rates may be 
+as on other factors, a plot of plant growth rates versus soil nitrogen may be 
 expected to look more like Fig. \ref{fig_cop_pedag1}A or E than like 
 Fig. \ref{fig_cop_pedag1}B or F. Nitrogen 
-will control plant growth, producing a strong
+will control plant growth, producing a clear
 relationship, only when it is limiting, in the left portion of the N and 
@@
 Nitrogen will be less related or unrelated to growth rates in the right tails of 
-the distributions, as pictured in Fig. \ref{fig_cop_pedag1}A and E. The portion
-of the distributions that dominate the overall association is visible in the copula
-approach, but may not be reflected by correlation coefficients and other standard techniques.
+the distributions, as in Fig. \ref{fig_cop_pedag1}A and E. The portion
+of the distributions that dominate the association is visible in the copula
+approach, but not via standard techniques.
 
@@
 ecological research is that such an approach may help illuminate 
-the causal mechanisms through which variables are related. For instance, if two species
+the causal mechanisms by which variables are related. For instance, if two species
 X and Y are strong competitors, we may expect abundances of the two species
-across quadrats or through time within a single quadrat to be negatively related, as in 
+across quadrats to be negatively related, as in 
 Fig. \ref{fig_cop_pedag1}I-J. If X is plotted on the horizontal axis
@@
 by X; whereas when X is rare, Y
-may be abundant or may also be relatively rare due to 
+may be abundant, or may also be relatively rare due to 
 limiting factors other than X. 
@@
 distributions are principally responsible for the association between variables, and this 
-detail may help support or refute alternative mechanistic and causal hypotheses about 
+detail may help support or refute alternative causal hypotheses about 
 the relationship between the variables.
@@
 We approached our overall goal of exploring the potential of 
-copula approaches for understanding dependences in ecology by addressing the 
+copula approaches in ecology by addressing the 
 following specific questions. (Q1) Do datasets in ecology and related fields 
@@
 ecosystem functioning data. We selected these datasets to span multiple 
-sub-fields and organizational levels of ecology, in an attempt to be 
-reasonably representative of the field as a whole. The second section of the 
+sub-fields and organizational levels of ecology. The second section of the 
 paper introduces the data, and the third and fourth sections use the data to 
@@
 the data of section two as well as models to explore a variety of consequences
-of copula structure and tail dependence for applications and for our understanding
-of ecology. The ninth section is the Discussion. 
+of copula structure and tail dependence for applications and for understanding
+ecology. The ninth section is the Discussion. 
 Copula approaches have been used to great benefit in neuroscience [@onken2009], 
@@
 variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], finance, and other fields. 
-To our knowledge, these approaches have been used seldom or never for ecological 
-research. We argue that the benefits to ecological research of a complete statistical 
-desription of dependence will be substantial, and we hope this paper helps stimulate the
-fruitful use of copula approaches across the discipline.
+To our knowledge, these approaches have been used seldom or never in ecology. 
+We argue that benefits to ecology of using a complete statistical 
+desription of dependence will be substantial.
 
@@
 
-We used datasets that pertain to multiple fields and levels of organization within 
-or related to ecology, including environmental, species-trait, phenological, 
+Datasets included environmental, species-trait, phenological, 
 population, community, and ecosystem functioning data. Copula structure describing 
-the relationship between atmospheric weather variables such as X and Y has been 
-examined previously in the meteorological literature (e.g., REF). We therefore 
+the relationship between atmospheric weather variables such as rainfall or wind speed 
+as measured in multiple locations has been 
+examined previously in the meteorological literature (e.g., @Serinaldi2008, @li2013). We therefore 
 examined environmental variables from the soil instead, using the Rapid Carbon 
-Assessment database (RaCA; [@Wills2014]).
+Assessment database (RaCA; @Wills2014).
 <!--***DAN: Add Wills and Loecke pending reference if published in time--> 
@@
 Species-trait data were average species basal metabolic rate (KJ per hour) and 
-body mass (gram) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
+body mass (grams) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
 and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
@@
 across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 
-35 years from 1976 to 2010. These time series were computed from the the Rothamsted Insect 
+35 years 1976 to 2010. These time series were computed from the Rothamsted Insect 
 Survey suction-trap dataset [@Harrington2014; @Bell2015]. The first of our two 
@@
 in the North Sea and British seas for 22 plankton taxa (Table \ref{SM-tab_plankton_info}) for the 
-56 years from 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. These data 
+56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. These data 
 were obtained from the Continuous Plankton Recorder (CPR) dataset maintained by the Sir 
@@
 
-The environmental, species-trait, and community datasets, as we studied them, are henceforth 
+The environmental, species-trait, and community datasets are henceforth 
 called the "bivariate" datasets in the sense that they comprise two quantities measured at 
 different locations or for different species. The phenology and population datasets are 
-"multivariate" in that they comprise, for each taxon, measurements at multiple locations; 
+"multivariate" in that they comprise, for each taxon, measurements through time at multiple locations; 
 copula structure can be studied for each pair of locations. Likewise the ecosystem 
@@
 and asymmetry of tail dependence. 
-Conceptually, a bivariate copula is essentially a probability density function 
-on the unit square with uniform maginals. The normal copula family is a one-parameter
+Conceptually, a bivariate copula is essentially a probability density function (pdf)
+on the unit square with uniform maginals. 
+Strictly speaking, 
+the copula is actually defined as the bivariate 
+cdf of the distribution rather than as the pdf or the random variable itself,
+but this makes little conceptual difference. 
+The normal copula family is a one-parameter
 family of copulas, with parameter corresponding to the correlation of the related 
@@
 serves as a benchmark for tail dependence in other copulas - positive values of our
-tail dependence measure correspond to dependence stronger than that of a normal
+tail dependence measures correspond to dependence stronger than that of a normal
 copula. The Clayton and 
@@
 $(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
-$\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$ and 
-we set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; subsequent analysis used these 
-*normalized ranks* $u_i$ and $v_i$. Ranks were $1$ for the smallest element in a set. 
+$\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$.
+Ranks were $1$ for the smallest element in a set. We then 
+set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; subsequent analysis used these 
+*normalized ranks* $u_i$ and $v_i$. 
 Second, we did a test of the independence of the $u_i$ and $v_i$ 
@@
 
-For large data sets (large $n$),
+For large datasets (large $n$),
 we could use $l_b$ and $u_b$ close together without incurring undue
@@
 several different bands $(l_b,u_b)$ to understand how dependence
-varies in different parts of the distributions. But for data sets with 
+varies in different parts of the distributions. But for datasets with 
 smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
@@
 The difference $\Dsq_u-\Dsq_l$ (note the opposite order), is also expected to be 
-positive under the same conditions. These statistics were used for smaller data sets.
-For large data sets, we used $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
+positive under the same conditions. These statistics were used for smaller datasets.
+For large datasets, we used $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
 $\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
@@
 
-For bivariate data sets, we compared values 
+For bivariate datasets, we compared values 
 of the statistics $\cor_{l_b,u_b}$,
@@
 $\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ to 
-distributions of values of the same statistics computed on surrogate data sets that were produced from 
+distributions of values of the same statistics computed on surrogate datasets that were produced from 
 the empirical data by randomizing it in a special way to have no 
 tail dependence (Appendix \ref{SM-surrog_test}). 
-The surrogate/randomized data sets had exactly 
+The surrogate/randomized datasets had exactly 
 the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
@@
 empirical data, to the distribution
-of its values computed on many surrogate data sets 
+of its values computed on many surrogate datasets 
 provides a test of whether dependence in the part of the distribution specified 
@@
 correlation. Thus the comparison helps answer the first two parts of Q1
-for bivariate data sets. 
+for bivariate datasets. 
 Significant deviations correspond to deviations from normal 
@@
 $\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$,
-as computed on the empirical data, to the distribution of its values on many surrogate data sets 
+as computed on the empirical data, to the distribution of its values on many surrogate datasets 
 provides a test of asymmetry of dependence. Thus this comparison helps answer the
 third part of Q1 for bivariate 
-data sets. For multivariate data sets, we used a spatial resampling scheme 
+datasets. For multivariate datasets, we used a spatial resampling scheme 
 (Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals, instead of the surrogate 
@@
 Across essentially all the sub-fields and levels of organization within ecology for which we 
-examined a data set, copula structures were typically non-normal and variables 
+examined a dataset, copula structures were typically non-normal and variables 
 commonly exhibited asymmetric tail 
 dependence. To make results easier to absorb for readers who are new to copulas: we first 
-present results in detail for an example bivariate data set; then we summarize results
-for all bivariate data sets; then we present results in 
-detail for an example multivariate data set; then we summarize results 
-for all multivariate data sets.
+present results in detail for an example bivariate dataset; then we summarize results
+for all bivariate datasets; then we present results in 
+detail for an example multivariate dataset; then we summarize results 
+for all multivariate datasets.
 
@@
 Table \ref{tab_bivar_summary} summarizes bivariate results for Q1. For all $4$ 
-bivariate data sets, the two variables were always significantly related 
+bivariate datasets, the two variables were always significantly related 
 (Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
@@
 (Table \ref{tab_bivar_summary}, row 9),
-answering the third part of Q1. The first three data sets had stronger upper- 
-than lower-tail dependence, and the last data set had the reverse. 
+answering the third part of Q1. The first three datasets had stronger upper- 
+than lower-tail dependence, and the last dataset had the reverse. 
 Although these results convincingly show non-normal copula structure, the 
@@
 (Table \ref{tab_bivar_summary}, row 10).
-Detailed results for each data set are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} 
+Detailed results for each dataset are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} 
 and Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.
@@
 according to Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests 
-(Table \ref{tab_multivar_summary}, row 6). Some of the data sets (green spruce aphid
+(Table \ref{tab_multivar_summary}, row 6). Some of the datasets (green spruce aphid
 abundance, \textit{Ceratium furca} abundance, methane flux) showed stronger 
@@
              format="latex", escape = F,
-             caption.short = "Summary of results for multivariate data sets",
+             caption.short = "Summary of results for multivariate datasets",
              caption = "Summary of results for multivariate data-sets.
@@
 
+Put ideas here for discussion paragraphs:
+1)
+
 <!--***DAN: some notes by which to revise this para, Shya:
@@
 
-<!--We showed by examples of several real data sets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables.-->
+<!--We showed by examples of several real datasets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables.-->
 
@@
 Stevens and P.
-Verrier for data extraction; and Jonathan Walter, Thomas Anderson, and Lei Zhao for helpful suggestions.
-We especially thank James Bell of the Rothamsted Insect Survey (RIS).
+Verrier for data extraction; and Joel E.
+Cohen, Jonathan Walter, Thomas Anderson, and Lei Zhao for helpful suggestions.
+We thank James Bell of the Rothamsted Insect Survey (RIS).
 The Rothamsted Insect Survey, a National Capability, is funded by the Biotechnology and Biological Sciences Research Council under the Core Capability Grant BBS/E/C/000J0200.
-SG, LWS and DCR were partly funded by US National Science Foundation grants 0075516 and 1442595, and the James S McDonnell Foundation.
+SG, LWS and DCR were partly funded by US National Science Foundation grants 0075516 and 1442595 and the James S McDonnell Foundation.
 

### 2019-03-21 SupportingInformation.Rmd
general edits to paper

@@
 Fig. \ref{MT-fig_cop_pedag1}F can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
+\caption[Lower- and upper-tail dependence can vary independently of correlation]{Lower- and upper-tail dependence can vary independently of correlation.
+(A-C) Data generated using the BB1 copula (see section on methods for Q1)
+with three different 
+sets of parameters chosen to produce a variety of 
+lower-tail (LT) and upper-tail (UT) dependence strengths (methods for Q1), but the same 
+Kendall correlation (K) (to two digits) in all cases. (D-F) Similar to A-C, but
+with stronger correlation. Asymmetry of tail dependence (LT-UT) is the same 
+within each column.\label{fig_cop_pedag}}
+\end{center}
+\end{figure}
 
@@
 asymptotic behavior of the 
-statistics and their behavior on the small data sets which are common in some fields
+statistics and their behavior on the small datasets which are common in some fields
 of ecology. The mean and standard error of the three difference statistics were computed for 
@@
 
-For bivariate data sets, surrogates were produced as follows.
+For bivariate datasets, surrogates were produced as follows.
 
@@
 another permutation of $1,\ldots,n$).
-- The surrogate data set is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.
+- The surrogate dataset is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.
 
@@
 Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
-Kendall correlation of the surrogate data set will be the same as that of the $a_i$ and $b_i$, and therefore 
+Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
 will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$

### 2019-03-22 Paper.Rmd
moving most of the nonparam stats results to sup mat

@@
 
-We addressed Q1 partly via a model selection procedure in which the normal copula
-and a variety of alternative copulas were fit to 
-our ecological datasets, and fits of the various copula models 
-were compared via the Akaike and Bayesian 
-Information Criteria (AIC and BIC). To make the results interpretable
+To make results interpretable
 to readers unfamiliar with copulas, we first introduce the normal copula and 
@@
 the copula is actually defined as the bivariate 
-cdf of the distribution rather than as the pdf or the random variable itself,
+cdf of the distribution rather than as the pdf,
 but this makes little conceptual difference. 
@@
 If LT and UT are used to denote 
-our two measures of tail dependence, their difference is a measure of asymmetry of 
+lower- and upper-tail dependence, their difference is a measure of asymmetry of 
 tail dependence. We also used two families that were asymmetric in their non-tail parts, 
@@
 on copulas [@nelsen2006_copula ; @joe2014_dependence] and in `VineCopula` documentation.
-Visualizations of all the copulas we used are in Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}.
+Visualizations of the copulas we used are in Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}.
 
@@
 survival-Clayton copulas with parameter 5. Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}
-have visualizations of all the copulas we used, including more visualizations
-of the Clayton and survival-Clayton copulas.\label{fig_cop_pedag3}}
+have visualizations of all the copulas we used.\label{fig_cop_pedag3}}
 \end{center}
@@
 
-Question Q1 was answered principally via analysis of the
-data using a model selection approach. For bivariate datasets 
+We addressed Q1 principally via a model selection procedure in which the normal copula
+the alternative copulas listed above were fit to 
+our ecological datasets, and fits of the various copula models 
+were compared via the Akaike and Bayesian 
+Information Criteria (AIC and BIC). For bivariate datasets 
 $(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
@@
 In addition to our model selection approach, we also used nonparametric approaches to 
-quantify lower- and upper-tail dependence and to answer Q1. Using two approaches can provide greater confidence 
-that results are robust. We used three nonparametric
-statistics <!--( $\cor$, $\Ps$ and $\Dsq$)--> which can quantify the extent to which 
+quantify lower- and upper-tail dependence and to answer Q1. Using two approaches can provide greater 
+confidence in results. We used three nonparametric
+statistics which can quantify the extent to which 
 two normalized ranks $u_i$ and $v_i$ are 
 related in any part of their distributions; we here describe the 
-statistics. The statistics were formulated with positively associated
+statistics operationally, with details in Appendix \ref{SM-nonparam_stats}.
+The statistics were formulated with positively associated
 variables in mind. All the variables we study were positively associated
 when they were significantly associated. So definitions and associated 
-diagrams will assume positive associations, though generalizations are 
-straightforward. Given two bounds 
+diagrams will assume positive associations. Given two bounds 
 $0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
@@
 using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
-the right parts of the distributions. 
-The first statistic, denoted $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
+the right parts of the distributions. The first statistic, denoted 
+$\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
 correlation of $u_i$ and $v_i$ that is attributable to the points 
-lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$:
-\begin{equation}\label{eq.Cor}
-   \cor_{l_b,u_b}(u,v) = \frac{\sum 
-   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}}.
-\end{equation}
-\noindent Here means and sample variances are computed using all $n$ data 
-points, but the
-sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
-$u_i+v_i < 2u_b$. 
+lying in the region given by the bounds.
 Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
-between $u$ and $v$ in the region given by the bounds. The sum of 
-$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
-$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
-interval $(0,1)$) equals the standard Spearman correlation.
-
+between $u$ and $v$ in the region.
 We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
 which has a similar 
-interpretation to $\cor_{l_{b_k},u_{b_k}}(u,v)$; we use both statistics 
-because $\cor_{l_{b_k},u_{b_k}}(u,v)$ is more familiar, but 
-in some instances $\Ps_{l_b,u_b}$ appears to have more power to
-reveal tail dependence than does $\cor_{l_{b_k},u_{b_k}}(u,v)$. 
+interpretation to $\cor_{l_{b},u_{b}}(u,v)$.
 Our third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
@@
 Unlike $\cor_{l_b,u_b}$ and 
-$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
-$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
-small in that case. 
-Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
-dependence in the portions of the distributions of $u$ and $v$ given by 
-the bounds, whereas small values of 
+$\Ps_{l_b,u_b}$, for which large values inidicate strong dependence
+in the portions of the the distributions of $u$ and $v$ given by 
+the bounds, small values of 
 $\Dsq_{l_b,u_b}$ indicate strong dependence. 
-
 For large datasets (large $n$),
@@
 sampling variation in our statistics, and we could consider 
-several different bands $(l_b,u_b)$ to understand how dependence
+multiple bands $(l_b,u_b)$ to understand how dependence
 varies in different parts of the distributions. But for datasets with 
@@
 $\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.
-
 To test for asymmetry of dependence in upper and lower portions of distributions,
-we used differences of upper and lower statistics.
-The difference $\cor_l-\cor_u$ is expected to be positive if 
-dependence in the left halves of the distributions is stronger 
-than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
-The difference $\Dsq_u-\Dsq_l$ (note the opposite order), is also expected to be 
-positive under the same conditions. These statistics were used for smaller datasets.
-For large datasets, we used $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
-$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
-$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$. 
-How close depended on $n$, and was chosen to ensure that sample variation 
-was not unduly large. We tested our statistics by testing whether 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of 
-tail dependence known to be present in data simulated from various copulas 
-(Appendix \ref{SM-Test_npa_stat}, Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}).
-
-For bivariate datasets, we compared values 
-of the statistics $\cor_{l_b,u_b}$,
-$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
-$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
-$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ to 
-distributions of values of the same statistics computed on surrogate datasets that were produced from 
-the empirical data by randomizing it in a special way to have no 
-tail dependence (Appendix \ref{SM-surrog_test}). 
-The surrogate/randomized datasets had exactly 
-the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
-correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
-but had normal copula structure. Thus our comparisons tested the null hypothesis 
-that our statistics took values on the empirical data no different from what would have been expected 
-if the copula structure of the data were normal, but the data were otherwise 
-statistically unchanged.
-
-The comparison of one of the
-statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
-empirical data, to the distribution
-of its values computed on many surrogate datasets 
-provides a test of whether dependence in the part of the distribution specified 
-by $l_b$ and $u_b$ is different from
-what would be expected from a null hypothesis of normal copula structure but 
-unchanged marginals and similar 
-Kendall (or Spearman)
-correlation. Thus the comparison helps answer the first two parts of Q1
-for bivariate datasets. 
-Significant deviations correspond to deviations from normal 
-copula structure. In particular, deviations 
-using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
-different from that of a normal copula;
-likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
-from that of a normal copula.
-The comparison of one of the statistics 
-$\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
-$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
-$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$,
-as computed on the empirical data, to the distribution of its values on many surrogate datasets 
-provides a test of asymmetry of dependence. Thus this comparison helps answer the
-third part of Q1 for bivariate 
-datasets. For multivariate datasets, we used a spatial resampling scheme 
-(Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals, instead of the surrogate 
-approach described above.
-
-<!--***DAN: If we end up being desperate to cut length, and assuming the nonparmetric results
-are sufficiently similar to model selection results, we may be able to put everything about 
-the nonparametric results in the supp mat and just say in the main text that we also used a 
-nonparametric approach and it gave the same general conclusions. Though it may take some work
-to split out the two approaches and all their results from each other.-->
+we used differences of upper and lower statistics: for smaller datasets, 
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ (note the opposite order
+in the last of these); for large datasets, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
+$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
+$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ (again note the opposite order) with $u_b$ close 
+to $0$.
+We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}).
 
@@
 
+Although these results do convincingly show non-normal copula structure, 
+we caution that the model-averaged tail-dependence statistics may have been biased 
+because even the lowest-AIC copula 
+(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
+was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
+Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
+available from `BiCopGofTest` in the `VineCopula` package). However, nonparametric statistics verified
+both that tail dependence was distinct from that of an N copula for the soil data, and that it 
+was asymmetric (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}, 
+Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}).
+
 \begin{figure}
@@
 \end{figure}
-<!--***DAN: We need to change the point plotting on these panels and other figures to use
-translucent points. Right now you cannot tell how many points are piled up in some parts of
-these plots, and we can easily fix that-->
-
-Although these results do convincingly show non-normal copula structure, 
-we caution that the model-averaged tail-dependence statistics may have been biased 
-because even the lowest-AIC copula 
-(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
-was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
-Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
-available from `BiCopGofTest` in the `VineCopula` package). However, nonparametric statistics verified
-both that tail dependence was distinct from that of an N copula for the soil data, and that it 
-was asymmetric (Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}). 
-The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data
-were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates
-of the data,
-separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
-(Fig. \ref{SM-fig_soilCN_nonparam}). Results show that tail dependence of data is stronger in 
-both the lower and upper tails than would be expected under a null hypothesis of a normal copula.
-The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
-$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for real data were also compared to distributions of their values
-on 1000 Kendall- or Spearman-preserving N surrogates, separately in two comparisons for each
-of the $(l_b,u_b)$ ranges listed above (Table \ref{SM-tab_soilCN_asym}). Results showed that
-upper-tail dependence was significantly stronger than lower-tail dependence.
 
 <!--bivariate summary-->
-Table \ref{tab_bivar_summary} summarizes bivariate results for Q1. For all $4$ 
+Table \ref{tab_bivar_summary} summarizes results for Q1 for the bivariate datasets. For all $4$ 
 bivariate datasets, the two variables were always significantly related 
@@
 But our nonparametric statistics confirmed the same conclusions 
-(Table \ref{tab_bivar_summary}, row 10).
+(Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}).
 Detailed results for each dataset are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} 
@@
 the first part of Q1 from the Introduction for these data.
-Goodness of fit tests based on a Cramer-von Mises statistic 
-(Table \ref{SM-tab_count_p_CvM}) in every case 
+Goodness of fit tests based on Cramer-von Mises or Kolmogorov-Smirnov statistics 
+(Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
 failed to reject the hypothesis that the AIC-best copula family was also 
 an objectively adequate description of 
-the data, and likewise for goodness of fit tests based on a Kolmogorov-Smirnov 
-statistic (Table \ref{SM-tab_count_p_KS}), demonstrating that the collection of 
+the data, demonstrating that the collection of 
 copula families we used was sufficiently broad that the copulas of all pairs of 
-locations could be characterized by at least one of them, for these data.
+locations could be characterized by at least one of them.
 Model-averaged lower- (Table \ref{SM-tab_count_LTdep_AICw}) and upper-tail
@@
 than the spatial synchrony of outbreak events. This also answers the third part of Q1.
-
-<!--Nonparametric analysis of tail dependence, green spruce aphid abundance data-->
 Nonparametric statistics verified that tail dependence was asymmetric for the green 
-spruce aphid abundance data.
-Mean values across all pairs of locations of the statistics 
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ were positive and
-confidence intervals excluded $0$ (Table \ref{SM-tab_count_resamp}). 
-Mean values of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
-also were always positive and had confidence intervals that excluded $0$ 
-(Table \ref{SM-tab_count_resamp}), confirming asymmetry of tail dependence and 
-stronger lower- than upper-tail dependence. Each of the statistics 
-was plotted against geographic (great-circle) distances between sampling locations, 
-across all pairs of sampling locations (Fig. \ref{SM-fig_aphid_count}).
+spruce aphid abundance data (Appendix \ref{SM-Q1multvarnonparam}).
 
@@
 data showed the reverse. Asymmetry results were verified by nonparametric approaches 
-(e.g., Table \ref{tab_multivar_summary}, row 10). Detailed results are in 
+<!--(e.g., Table \ref{tab_multivar_summary}, row 10)-->
+(Appendix \ref{SM-Q1multvarnonparam}). Detailed results are in 
 Tables \ref{SM-tab_count_bestcop}-\ref{SM-tab_methane_resamp} and 
@@
              format='pandoc',
-             caption = "Summary of results for bivariate data-sets : \\textbf{p.Independence.test}: p-value from `BiCopIndTest`, \\textbf{minimumAIC.bestfit.copula}: best fit copula based on minimum AIC, \\textbf{Bestfit.copula.AICw}: AIC-weight of that best fit copula, \\textbf{Normal.copula.AICw}: AIC-weight of Normal copula, \\textbf{p.CvM}: p-value from `BiCopGofTest` for Cramer-von Mises method, \\textbf{p.KS}: p-value from `BiCopGofTest` for Kolmogorov-Smirnov method,\\textbf{LT.model.avg}: AIC-weights model averaged lower tail dependence, \\textbf{UT.model.avg}: AIC-weights model averaged upper tail dependence, \\textbf{(LT-UT).model.avg}:  AIC-weights model averaged (lower-upper) tail dependence.
+             caption = "Summary of results for bivariate datasets : \\textbf{p.Independence.test}: p-value from `BiCopIndTest`, \\textbf{minimumAIC.bestfit.copula}: best fit copula based on minimum AIC, \\textbf{Bestfit.copula.AICw}: AIC-weight of that best fit copula, \\textbf{Normal.copula.AICw}: AIC-weight of Normal copula, \\textbf{p.CvM}: p-value from `BiCopGofTest` for Cramer-von Mises method, \\textbf{p.KS}: p-value from `BiCopGofTest` for Kolmogorov-Smirnov method,\\textbf{LT.model.avg}: AIC-weights model averaged lower tail dependence, \\textbf{UT.model.avg}: AIC-weights model averaged upper tail dependence, \\textbf{(LT-UT).model.avg}:  AIC-weights model averaged (lower-upper) tail dependence.
 Statistic $\\Ps_l$ - $\\Ps_u$ for the actual copula data and for Kendall-correlation preserving $1000$ Normal surrogates were computed and ranked in ascending order.
@@
              format='pandoc',
-             caption = "Summary of results for multivariate data-sets : \\textbf{num.pairs}: total number of location pairs (excluding self-comparison) considered, \\textbf{num.non-indep}: how many location pair copula are non-independent among them, \\textbf{num.non-N}: number of non-Normal copula as best fit (based on highest AIC-weight), \\textbf{Avg.best.AICw}: average AIC-weights of best fit copula from all location pair, \\textbf{Avg.N.AICw}: average AIC-weights of Normal copula from all location pair, \\textbf{num.good.fit}: number of location pair for which best fit copula have p-values $>$ 0.01
+             caption = "Summary of results for multivariate datasets : \\textbf{num.pairs}: total number of location pairs (excluding self-comparison) considered, \\textbf{num.non-indep}: how many location pair copula are non-independent among them, \\textbf{num.non-N}: number of non-Normal copula as best fit (based on highest AIC-weight), \\textbf{Avg.best.AICw}: average AIC-weights of best fit copula from all location pair, \\textbf{Avg.N.AICw}: average AIC-weights of Normal copula from all location pair, \\textbf{num.good.fit}: number of location pair for which best fit copula have p-values $>$ 0.01
 (both Cramer-von Mises and Kolmogorov-Smirnov method), \\textbf{0.025-CI.LT}: lower and \\textbf{0.975-CI.LT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) lower tail dependence from all location pair, \\textbf{0.025-CI.UT}: lower and \\textbf{0.975-CI.UT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) upper tail dependence from all location pair, \\textbf{num.LT-UT>0}: number of location pair which showed stronger lower tail dependence than upper, Last $6$ rows summarized the nonparametric analysis results for $\\Ps$ statistics : lower and upper quantile ($95\\%$CI) of $\\Ps_l$, $\\Ps_u$, $\\Ps_l-\\Ps_u$ based on resampling location pair.\\label{tab_multivar_summary}",
@@
 tab_bivar_summary<-cbind(dt,tab_bivar_summary)
-knitr::kable(tab_bivar_summary, 
+knitr::kable(tab_bivar_summary[1:9,], 
              format="latex",
@@
 The $p$-values in rows 5-6 are for the minimum-AIC copula.
-Model averaging used for rows 7-9 was based on AIC weights.
-For row 10, the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of real-data value in the distribution of surrogate values is presented.
-If ranks were $>975$ or $<25$, asymmetry of tail dependence was significant compared to a N-copula null hypothesis.\\label{tab_bivar_summary}",
+Model averaging used for rows 7-9 was based on AIC weights.\\label{tab_bivar_summary}",
              booktabs=T, linesep = "\\addlinespace",
@@
 ```
-<!--***DAN: Shya, it says Pl-Pu above, but I suspect it was P_first slice - P_last slice, right?
-Can you please look back at the code, if necessary, and write a comment below this one 
-indicating what results are actually presented in the table?
-***Shya: Yes, you are right, here Pl-Pu means P_first slice - P_last slice. In table, item 10 : Assymetry rank is shown for Kendall preserving surroagates.
-Reference : bivfunctionplot.R file, lines 134,135,161
-***DAN: OK, great. I have made changes to deal with this, but am leaving the comment for now
-as a reminder that we looked at this-->
 
@@
 
-knitr::kable(tab_multivar_summary, 
+knitr::kable(tab_multivar_summary[1:9,], 
              format="latex", escape = F,
              caption.short = "Summary of results for multivariate datasets",
-             caption = "Summary of results for multivariate data-sets.
+             caption = "Summary of results for multivariate datasets.
 Rows 3-9 of the table were computed for the non-independent pairs (row 2) only; row 10 was for all location pairs.

### 2019-03-22 SupportingInformation.Rmd
moving most of the nonparam stats results to sup mat

@@
 
-We here define $\Ps_{l_b,u_b}$, using the bounds
+\noindent We here define
+\begin{equation}\label{eq.Cor}
+   \cor_{l_b,u_b}(u,v) = \frac{\sum 
+   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}}.
+\end{equation}
+\noindent Here means and sample variances are computed using all $n$ data 
+points, but the
+sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
+$u_i+v_i < 2u_b$. 
+The sum of 
+$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
+$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
+interval $(0,1)$) equals the standard Spearman correlation.
+
+We also define $\Ps_{l_b,u_b}$, using the bounds
 $u+v=2l_b$ and $u+v=2u_b$ defined in the main text, as follows. For a distance $h$, we define
@@
 between $u$ and $v$ in the region given by the bounds.
+We use both statistics 
+because $\cor_{l_{b_k},u_{b_k}}(u,v)$ is more familiar, but 
+in some instances $\Ps_{l_b,u_b}$ appears to have more power to
+reveal tail dependence.
 
@@
 
+# Verifying Q1 results for bivariate datasets using nonparametric statistics \label{Q1bivarnonparam}
+
+For bivariate datasets, we compared values 
+of the statistics $\cor_{l_b,u_b}$,
+$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
+$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
+$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ to 
+distributions of values of the same statistics computed on surrogate datasets that were produced from 
+the empirical data by randomizing it in a special way to have no 
+tail dependence (Appendix \ref{surrog_test}). 
+The surrogate/randomized datasets had exactly 
+the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
+correlation (the sorrogate algorithm had two versions, one for preserving each correlation coefficient), 
+but had normal copula structure. Thus our comparisons tested the null hypothesis 
+that our statistics took values on the empirical data no different from what would have been expected 
+if the copula structure of the data were normal, but the data were otherwise 
+statistically unchanged. The comparison of one of the
+statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
+empirical data, to the distribution
+of its values computed on many surrogate datasets 
+provides a test of whether dependence in the part of the distribution specified 
+by $l_b$ and $u_b$ is different from
+what would be expected from a null hypothesis of normal copula structure but 
+unchanged marginals and similar 
+Kendall (or Spearman)
+correlation. Thus the comparison helps answer the first two parts of Q1
+for bivariate datasets. 
+Significant deviations correspond to deviations from normal 
+copula structure. In particular, deviations 
+using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
+different from that of a normal copula;
+likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
+from that of a normal copula.
+The comparison of one of the statistics 
+$\cor_{0,u_b}-\cor_{1-u_b,1}$, 
+$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
+$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$,
+as computed on the empirical data, to the distribution of its values on many surrogate datasets 
+provides a test of asymmetry of dependence. Thus this comparison helps answer the
+third part of Q1 for bivariate 
+datasets.
+
+The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
+C and N data
+were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates
+of the data,
+separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
+(Fig. \ref{fig_soilCN_nonparam}). Results show that tail dependence of data is stronger in 
+both the lower and upper tails than would be expected under a null hypothesis of a normal copula.
+The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
+$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
+on 1000 Kendall- or Spearman-preserving N surrogates, separately in two 
+comparisons (Table \ref{tab_soilCN_asym}). Results showed that
+upper-tail dependence was significantly stronger than lower-tail dependence.
+
+For all four bivariate datasets the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
+for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of the real-data value in 
+the distribution of surrogate values was computed. If ranks were $>975$ or $<25$, asymmetry of tail 
+dependence was significant compared to a N-copula null hypothesis. Ranks were 1, 1, 1, 998
+<!--***DAN: Shya, please make the quantities in the previous line automatic links-->
+for, respectively, the soil C and N data, bird masses and metabolic rates, mammal masses
+and metabolic rates, and the Cedar Creek data. Thus the first three datasets showed stronger 
+upper- than lower-tail dependence, and the Cedar Creek data showed the reverse, confirming the 
+results based on model selection presented in the main text.
+
 # Surrogate testing with bivariate datasets for nonparametric statistics \label{surrog_test}
@@
 sampling variation) the Spearman correlation of the empirical data.
+
+# Verifying Q1 results for multivariate datasets using nonparametric statistics \label{Q1multvarnonparam}
+
+For multivariate datasets, we calculated Spearman and Kendall correlations
+and the statistics 
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
+for all pairs of sampling locations, and we then computed means across the pairs
+for each statistic. We used a spatial resampling scheme 
+(Appendix \ref{spatial_resamp}) to calculate confidence intervals of these means.
+
+For the green spruce aphid abundance data, mean values across all pairs of 
+locations of the Spearman and Kendall correlations and the statistics 
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ were positive and
+confidence intervals excluded $0$ (Table \ref{tab_count_resamp}). 
+Mean values of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
+also were always positive and had confidence intervals that excluded $0$ 
+(Table \ref{tab_count_resamp}), confirming the results in the main text on 
+asymmetry of tail dependence and 
+stronger lower- than upper-tail dependence for this dataset. Each of the statistics 
+was plotted against geographic (great-circle) distances between sampling locations, 
+across all pairs of sampling locations (Fig. \ref{fig_aphid_count}).
+
+Asymmetry of tail dependence was also verified by our nonparametric methods for
+the other multivariate datasets. For instance, the $95\%$ confidence intervals of the
+mean over pairs of sampling locations of the statistic $\Ps_l-\Ps_u$ were
+$(-0.085,-0.041)$ for the leaf-curling plum aphid first flight data, 
+indicating greater upper-tail dependence, and consistent with
+<!--***DAN: Shya, pls make above results autolink-->
+the results of Table \ref{MT-tab_multivar_summary} (row 9). 
+For the *Ceratium furca* abundance data, confidence intervals were
+$(0.011,0.044)$, indicating greater lower-tail dependence, and again
+<!--***DAN: Shya, pls make above results autolink-->
+consistent with the results of Table \ref{MT-tab_multivar_summary} (row 9).
+For the methane flux data, confidence intervals were 
+$(-0.008,0.019)$.
 
@@
 [@bjornstadfalck2001 ; @walteretal2017].
-
-<!--***DAN: text all good up to here, except as marked (the data section). Also
-all figs and tables relevant to the stuff this far have been edited and look pretty
-good. This is as of 2018 01 10, 1255pm. So start editing from this point.-->
 

### 2019-03-22 SupportingInformation.Rmd
fixing multivar_summary table

@@
   m$`num.non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
-  m$`num.non-N`<-sum(r$info_ord_copcode[,,1]!=1,na.rm = T)
+  bcop<-r$info_ord_copcode[,,1] # matrix with best fit copula name
+  bcop[bcop==Inf]<-NaN # replacing nonInd pairs with NaN
+  m$`num.non-N`<-sum(bcop!=1,na.rm = T)
   r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN

### 2019-03-22 Paper.Rmd
finished moving most nonparam results to sup mat, the paper is shorter now

@@
              caption = "Summary of results for multivariate datasets.
-Rows 3-9 of the table were computed for the non-independent pairs (row 2) only; row 10 was for all location pairs.
+Rows 3-9 of the table were computed for the non-independent pairs (row 2) only.
 A well-fitting location pair (row 6) was one for which the best-fitting copula had $p$-values $>0.01$ for both the Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests.
@@
 in two places may bring about similarly 
-structured dependence between ecological variables that are influenced by the 
+structured dependence between ecological variables in those places that are influenced by the 
 environmental variables.
@@
 total plant community biomass or other community variable from an average value,
-or flux of a biogeochemical variable such as methane or CO$_2$..
+or flux of a biogeochemical variable such as methane or CO$_2$.
 
@@
 give rise to non-N copula structure and asymmetric tail dependence in the 
-ecological variable.
+ecological variable, in spite of the absence of such structure in the driver
+environmental variables.
 
@@
 $\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
-population model outputs; and likewise for SG and SJ copulas. Dynamics were
+population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}); 
+and likewise for SG and SJ copulas (Figs \ref{SM-fig_Cause4copula_SG_kend_sRicker_nsd_1}, 
+ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). Dynamics were
 strongly nonlinear for the model parameters we used, and nevertheless similarities
@@
 complex for $b=0.9$ or $-0.9$, indicating that strongly autocorrelated (or strongly 
-negatively autocorrelated dynamics) interact with the effect we have described.
+negatively autocorrelated) dynamics interact with the effect we have described.
 Ecological-variable copulas were strongly non-N for $g_1$, but not for $g_2$ 
@@
 \ref{SM-tab_summary_stat_asym_sens}-\ref{SM-tab_stat_npa_asym}).
+<!--***DAN: Shya, this has got to be an error. Results should be symmetric
+with respect to the choice of g_1 and g_2 because the model is symmetric. Let's discuss.
+I think there is likely a bug affecting Figs S51 and 51, and Tables S47-48, and 
+possibly all results based on the same model.-->
 
@@
 principal importance for 
-cases such as pest or resource abundance or CO$_2$ or methane flux.
+cases such as pest or resource abundance, for which extreme values (spikes) 
+in the spatial-mean
+quantity have disproportionately large effects.
 
@@
 component populations would occur simultaneously in many patches,
-reducing rescue effects. 
+reducing rescue effects. Here, we are assuming that low values of the environmental
+variable are "bad" for the populations being modelled, and high values are "good".
 We provided a test of the conceptual validity of this hypothesis using a metapopulation

### 2019-03-22 SupportingInformation.Rmd

--- 2019-03-22 SupportingInformation.Rmd
@@
   m$`num.non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
-  m$`num.non-N`<-sum(r$info_ord_copcode[,,1]!=1,na.rm = T)
+  bcop<-r$info_ord_copcode[,,1] # matrix with best fit copula name
+  bcop[bcop==Inf]<-NaN # replacing Independent pairs with NaN
+  m$`num.non-N`<-sum(bcop!=1,na.rm = T)
   r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN

### 2019-03-26 Paper.Rmd
editing first para of Discussion and made a list of ideas for Discussio section paragraphs

@@
 exceedingly common in datasets that range across multiple sub-disciplines
-of ecology. Via basic mathematical models, we hypothesized two probable
-mechanisms causing those non-N copula structures with asymmetric tail-dependence. One
-mechanism was "Moran"-like effect where dependence structure between ecological
-variables from two locations was shaped by the dependence structure of 
-the environmental variables at those patches. Second mechanism, a mimic of 
-Liebig's law of the minimum, explored
-the asymmetric influence of environmental variables which 
-could limit the growth of populations when below (or beyond)
-a certain threshold. Using model simulations and real ecological data,
-we concluded asymmetric tail dependence feature of 
-non-N copula could affect (1) the skewness (henceforth, the stability)
-of spatially averaged quantities over years, (2) the extinction risk 
-and (3) the spatial Taylor's law slope of metapopulations.
+and levels of organization in ecology, in spite of the fact that no published 
+ecological research we are aware of has ever
+taken this structure into account. Via simple mathematical models, we hypothesized two 
+mechanisms that may cause non-N copula structures with asymmetric tail dependence
+in ecological data. One
+mechanism was via a Moran-like effect by which the dependence structure between ecological
+variables, measured through time in different locations, is shaped by the dependence structure of 
+an environmental variable in those locations that influences the ecological variable. 
+A second mechanism gives rise to non-N copula structure and asymmetric tail dependence via
+asymmetric or nonlinear influence of environmental variables on ecological variables, 
+and relates to Liebigs law of the minimum. We discuss below how commonly these and other
+potential mechanisms may opperate. 
+<!--***DAN: make sure to do so-->
+We also demonstrated important consequences of 
+non-N and tail-dependent copula structure in ecology. First, the skewness, through time, 
+of spatially averaged time series is affected by the tail dependence of the relationships
+between the local quantities being averaged (such as local abundances of a pest or a resource). 
+Thus right-tail-dependent local quantities 
+tend to produce "spiky" spatially averaged time series, with large outbreaks. 
+Second, metapopulation extinction risk can be accentuated or reduced by the tail dependence
+properties of environmental variables that influence the populations. Finally, we presented
+simulation results in an Appendix showing that spatial Taylor's law can be strongly affected by
+tail dependence between local populations. In our view, our results show that
+many ecological phenomena cannot be completely understood without using a more complete, 
+copula-based approach to analyzing dependence between variables, and we advocate that
+such approaches become standard tools in ecology.
 
 Put ideas here for discussion paragraphs:
-1)
-
-<!--***DAN: some notes by which to revise this para, Shya:
-1) no need to restate methods at all, the methods are not among our main results in this paper
-2) No real need to say much about why it is important to study what we stdied, that was done in the intro.
-Except insofar as saying what our main results are should make it clear it is an important thng to study, and see also further notes below.
-3) Some additional points below
--->
-<!--
-Most studies in ecology have focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values.
-In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field.
--->
-<!--***DAN: Too discursive.
-We do not need to repeat that we introduced this.
-Just summarize our main results in pothy language!
-And then maybe emphasize the importance, or make any additional points about importance that could not be made in the Intro.
-This para needs to be hard-hitting and brief, saying all the important thigs we accomplished in short form, all in one place.
-See below for some examples.-->
-
-<!--We showed by examples of several real datasets (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence are common in nature and could tell us about the nature of interaction between different ecological variables.-->
-
-<!--***DAN: How about simply: "We showed that non-N copula structure and asymmetric tail dependence are exceedingly common in datasets that range across multiple sub-disciplines of ecology." See if you can write 1-3 pithy, impactful sentences like this one for Q2, and 1-3 for Q3 as well.
-Look at the first para of Discussion for some of my recent papers to get examples of the writing techniques for how we have done this in the past.-->
-
-<!--We applied two methods.
-First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family.
-Copula having highest AIC-weight came as best-fit from this approach.
-Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), we wanted to ensure the reliability of our model selection results.
-So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different nonparametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we defined and we found they were consistent with the inferences we drew from model selection results.
--->
-<!--***DAN: No need at all to summarize methods again, except in cases where the methods are 
-among the main products of the paper - not our case.
-So delete the above text.
-If you want to discuss our methods, do in later in the Discussion, not in this first paragraph, unless you may want to have a sentence saying something like "We synthesized and expanded upon methods for statistically assessing copula structure and tail dependence, and models for exploring the causes and consequences of copula strucuture in ecology and related fields."-->
-<!--Our approach could help to interpret the synchronous behavior of a species from several meta-populations.
-To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g.
-abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that meta-populations are simultaneously rare at those locations.
-Conversely, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of plankton could lead to algal-bloom in the ocean.
-Copulae could also give us information about the causal relationship between ecological variables.
-For example, if maximum rainfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be more strongly correlated than lower values.
--->
-<!--***DAN: I do not see how these examples go beyond the examples of the introduction. If they do, please clarify. If they don't, please remove.-->
-
-<!--***DAN: Shya, the below is more summary of results.
-Pls summarize very briefly in the first paragraph of the Discusion ONLY.
-So most of the below won't be needed, but some could be incorporated into the first Discussion para above.-->
-<!--We explained the possible causes which induced asymmetric tail-dependence in the copula.
-Borrowing the idea from "Moran's theorem", we demonstrated with simulations that two habitat patches when affected by bivariate noise with an asymmetric copula structure showed similar tail-dependence asymmetry in the population copula from those patches (see Fig.
-\ref{fig_Cause4copula_params_GOF_500} and Fig.
-\ref{fig_Cause4copula_scatter_C_kend}).
-We also showed the effect of global environmental fluctuation (behaving asymmetrically beyond a certain threshold) on the meta-population copula which was shaped with similar tail dependence as the initial global noise copula (Fig.
-\ref{fig_ms_asym_b_0.1}).
-We observed significant changes in skewness of spatially averaged time series and also in the curvature of spatial Taylor's law for copula having asymmetric tail-dependence than would have been expected for Normal copula.
-For example, Fig.
-\ref{fig_skewness_spearman}(A) showed green spruce aphid abundance time series had right skewed distribution but its skewness was significantly lower than was expected if it came from Spearman correlation preserving Normal surrogates.
-This result indicates less extreme outbreaks of green spruce aphid around UK than it would have been with Normal copula structure.
-Similarly, methane fluxes showing lower skewness than Normal surrogates mean that the high emissions are un-correlated between locations and monitoring this skewness over several decades may be helpful to understand the "green-house effect" on environment.
--->
+1) Maybe a para on how likely one is to get copula structure via the Liebig mechanism - considers
+what people have written on nonlinear effects and empirical evidence for Liebig's law and so on
+2) Maybe a para on how likely one is to get copula structure via the Moran mechanism - considers
+work on copula structure in environmental variables. Future work: copula structure of environmental
+variables known to be important in ecology, including how it may be changing and whether it has a 
+geography.
+3) a para on stability of the total biomass time series, and skewness - sets up Shya's later paper as
+future work
+4) We have only given two mechanisms, but there are likely to be others. Maybe I can think
+of some ideas and make a paragraph.
+5) If Van gives us some ideas, they may fit in the Discussion.
+6) Maybe give Holder a bunch of draws from a bivariate dist with standard normals marginals but with 
+some tail-dependent copula structure and ask him to simulate evolution down a mammal or bird tree and
+see what you get. Then give him the other tail dependence, and no tail dependence (bivariate standard
+normal). Maybe this will lead to more results, and also a Discussion para on why evolutionary steps 
+might be tail dependent. A result that tail dependent steps lead to tail dependent character dists
+in the end would replace one question (why are there tail-dependent characters) with another (why are
+there tail dependent steps). Of course it could also be the case that steps are not tail dependent and 
+the tail dependence of characters arises in another way. Would also lead in to alternative thinking in 
+evolutionary biology - Mark could write a para on why it would matter for evolutionary biol if there are 
+tail dependent steps.
+7) A para that sets up Shya's extinction risk paper - asks about importance of this effect for real 
+species and whether it needs to be incorporated into PVA, and in what cases.
+If the copula structure of environmental variables changes (is it changing?) what will be the impact on extinction risk?
+Also, there may be copula structure between vital rates, and it is not accounted for in typical stochastic matrix model based PVA.
+Does it need to be?
+8) A para about what is available in the copula and VineCopula packages.
+Maybe make it a broader para about emthods, and state in there that we developed some new surrogate methods that may be broadly useful for researches trying to incorporate copula approaches.
 

### 2019-03-26 SupportingInformation.Rmd
modified r chunk for asym sens results with b=+-0.1, +-0.5

@@
 #----------------------------------------------------------
-N<-5000
+N<-25000
 r<-0.8
 a<-0.2
-numkeep_last<-500
+numkeep_last<-2500
 temp<-get_noise_glob_loc_mat(N=N,r=r)
@@
 ans_b_neg0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
-ans_b_0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.9,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
-ans_b_neg0.9<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.9,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
-
-vlist_b_0.1_b_0.9<-c(ans_b_0.1,ans_b_neg0.1,ans_b_0.9,ans_b_neg0.9) # a list of 8 matrices
+ans_b_0.5<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.5,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
+ans_b_neg0.5<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.5,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
+
+vlist_b_0.1_b_0.5<-c(ans_b_0.1,ans_b_neg0.1,ans_b_0.5,ans_b_neg0.5) # a list of 8 matrices
 
@@
 #bg <- intToUtf8(0x03B2) 
-cnm<-c("b_0.1_g~1~","b_0.1_g~2~","b_-0.1_g~1~","b_-0.1_g~2~","b_0.9_g~1~","b_0.9_g~2~","b_-0.9_g~1~","b_-0.9_g~2~")
+cnm<-c("b_0.1_g~1~","b_0.1_g~2~","b_-0.1_g~1~","b_-0.1_g~2~","b_0.5_g~1~","b_0.5_g~2~","b_-0.5_g~1~","b_-0.5_g~2~")
 #cname<-paste0(bg,cnm)
@@
 
-ms_asym_b_0.1_b_0.9<-vector(mode = "list", length = 8) # initializing list to store model selection results for all possible b variation
-names(ms_asym_b_0.1_b_0.9)<-cnm
-
-stat_npa_asym_b_0.1_b_0.9<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible b variation
-names(stat_npa_asym_b_0.1_b_0.9)<-cnm
+ms_asym_b_0.1_b_0.5<-vector(mode = "list", length = 8) # initializing list to store model selection results for all possible b variation
+names(ms_asym_b_0.1_b_0.5)<-cnm
+
+stat_npa_asym_b_0.1_b_0.5<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible b variation
+names(stat_npa_asym_b_0.1_b_0.5)<-cnm
 
 for(ic in c(1:8)){
-  v<-vlist_b_0.1_b_0.9[[ic]]
+  v<-vlist_b_0.1_b_0.5[[ic]]
   res<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=c(1,3:10,13:14,16:20),level=0.05,
@@
 
-  ms_asym_b_0.1_b_0.9[[ic]]<-res
+  ms_asym_b_0.1_b_0.5[[ic]]<-res
 
@@
 
-  stat_npa_asym_b_0.1_b_0.9[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cnm[ic],numbin=10)
+  stat_npa_asym_b_0.1_b_0.5[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cnm[ic],numbin=10)
 
@@
 saveRDS(summary_stat,"./Results/asym_sens_results/b_variation_summary_stat.RDS")
-saveRDS(ms_asym_b_0.1_b_0.9,"./Results/asym_sens_results/ms_asym_b_0.1_b_0.9.RDS")
-saveRDS(stat_npa_asym_b_0.1_b_0.9,"./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
+saveRDS(ms_asym_b_0.1_b_0.5,"./Results/asym_sens_results/ms_asym_b_0.1_b_0.5.RDS")
+saveRDS(stat_npa_asym_b_0.1_b_0.5,"./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
 ```
@@
 pop<-ans_b_0.1$pop
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
 text(0.1,0.95,"A",cex=2)
 pop<-ans_b_0.1$pop_d
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,pch=20,col=rgb(0,0,0,0.5),cex=0.6)
 text(0.1,0.95,"B",cex=2)
@@
 pop<-ans_b_neg0.1$pop
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,pch=20,col=rgb(0,0,0,0.5),cex=0.6)
 text(0.1,0.95,"A",cex=2)
 pop<-ans_b_neg0.1$pop_d
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
 text(0.1,0.95,"B",cex=2)
@@
 
-pdf("./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf",height=3,width=6)
+pdf("./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf",height=3,width=6)
 op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
-pop<-ans_b_0.9$pop
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+pop<-ans_b_0.5$pop
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
 text(0.1,0.95,"A",cex=2)
-pop<-ans_b_0.9$pop_d
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+pop<-ans_b_0.5$pop_d
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
 text(0.1,0.95,"B",cex=2)
@@
 
-pdf("./Results/asym_sens_results/a_0.2_b_-0.9_r_0.8_asym_sens.pdf",height=3,width=6)
+pdf("./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf",height=3,width=6)
 op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
-pop<-ans_b_neg0.9$pop
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+pop<-ans_b_neg0.5$pop
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
 text(0.1,0.95,"A",cex=2)
-pop<-ans_b_neg0.9$pop_d
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,cex.axis=1,col="grey",pch=20,cex=0.6)
+pop<-ans_b_neg0.5$pop_d
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
 text(0.1,0.95,"B",cex=2)
@@
              format = "latex", linesep = "",
-             digits=3, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$",
+             digits=4, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$",
                                    "b=-0.1, g$_{1}$","b=-0.1, g$_{2}$",
-                                   "b=0.9, g$_{1}$","b=0.9, g$_{2}$",
-                                   "b=-0.9, g$_{1}$","b=-0.9, g$_{2}$"),
+                                   "b=0.5, g$_{1}$","b=0.5, g$_{2}$",
+                                   "b=-0.5, g$_{1}$","b=-0.5, g$_{2}$"),
       caption.short="Asymmetric sensitivity model, model selection results",
@@
 the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
-\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.9}) were subjected to the model 
+\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to the model 
 selection procedures detailed in the 
@@
 
-tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
+tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
 tab_stat_npa_asym<-c()
@@
 }
-gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.9, g$_{1}$","b=0.9, g$_{2}$","b=-0.9, g$_{1}$","b=-0.9, g$_{2}$")
-tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.9,0.9,-0.9,-0.9),each=3),
+gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, g$_{1}$","b=-0.5, g$_{2}$")
+tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),each=3),
   g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
@@
 the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
-\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.9}) were subjected to nonparamtric statistical 
+\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
 analyses decribed in the methods section for Q1. The listed statistic was computed for the
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.9_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.9$.\label{fig_ms_asym_b_0.9}}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.9$.\label{fig_ms_asym_b_0.5}}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.9_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.9$.\label{fig_ms_asym_b_-0.9}}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.9$.\label{fig_ms_asym_b_-0.5}}
 \end{center}

### 2019-03-26 Paper.Rmd
asym sens results explained

@@
 local noise, independent across time and sampling locations. We used $a=0.2$, and 
-$b=0.1, -0.1, 0.9, -0.9$ in different simulations. 
+$b=0.1, -0.1, 0.5, -0.5$ in different simulations. 
 Negative values of $b$ correspond to overcompensating dynamics 
@@
 
-For each value of $b$, we simulated the model once using $g_1$ and once using $g_2$, for 5000 time
-steps in both cases, retaining $E_1(t)$ and $E_2(t)$ for the final 500 time 
+For each value of $b$, we simulated the model once using $g_1$ and once using $g_2$, for $25000$ time
+steps in both cases, retaining $E_1(t)$ and $E_2(t)$ for the final $2500$ time 
 steps of each simulation. We applied our model selection algorithms and 
@@
 and likewise for SG and SJ copulas (Figs \ref{SM-fig_Cause4copula_SG_kend_sRicker_nsd_1}, 
-ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). Dynamics were
+\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). Dynamics were
 strongly nonlinear for the model parameters we used, and nevertheless similarities
@@
 asymmetric tail dependence and non-N copula structure were visually apparent.
+We also got similar results from simulations with $b=\pm 0.5$ (Figs \ref{SM-fig_ms_asym_b_0.5} - 
+\ref{SM-fig_ms_asym_b_-0.5}).
 Lower-tail (respectively, upper-tail) effects of the noise ($g_1$, resp., $g_2$) 
 produced lower-tail (resp., upper-tail) dependence in the ecological variable.
-Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) and
-our nonparametric statistics (Tables \ref{SM-tab_stat_npa_asym}) 
-statistically confirmed the visual result for $b=0.1, -0.1$. Results were more
-complex for $b=0.9$ or $-0.9$, indicating that strongly autocorrelated (or strongly 
-negatively autocorrelated) dynamics interact with the effect we have described.
-Ecological-variable copulas were strongly non-N for $g_1$, but not for $g_2$ 
-and asymmetries of tail dependence were not particularly noticable 
-(Figs \ref{SM-fig_ms_asym_b_0.9}-\ref{SM-fig_ms_asym_b_-0.9} and Tables 
-\ref{SM-tab_summary_stat_asym_sens}-\ref{SM-tab_stat_npa_asym}).
+Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
+non-Normal copula was always best fit for $b= \pm 0.1$ and $\pm 0.5$, but their model 
+AIC-wieght averaged tail dependence could not reflect the asymmetry for each case 
+when that dependnece was at too much extreme end.
+However, our nonparametric statistics (Tables \ref{SM-tab_stat_npa_asym}) focused on those extreme tips
+and overcome the shortcomings of model selection approach which fit data based on maximum likelihood.
+
 <!--***DAN: Shya, this has got to be an error. Results should be symmetric

### 2019-03-28 Paper.Rmd
working on Discussion

--- 2019-03-26 Paper.Rmd
@@
 and levels of organization in ecology, in spite of the fact that no published 
-ecological research we are aware of has ever
+ecological research we are aware of has 
 taken this structure into account. Via simple mathematical models, we hypothesized two 
@@
 mechanism was via a Moran-like effect by which the dependence structure between ecological
-variables, measured through time in different locations, is shaped by the dependence structure of 
-an environmental variable in those locations that influences the ecological variable. 
+variables (measured through time at different locations) is shaped by the dependence structure of 
+environmental variables (measured at the same times and locations). 
 A second mechanism gives rise to non-N copula structure and asymmetric tail dependence via
-asymmetric or nonlinear influence of environmental variables on ecological variables, 
-and relates to Liebigs law of the minimum. We discuss below how commonly these and other
+the asymmetric or nonlinear influence of environmental variables on ecological variables, 
+and relates to Liebig's law of the minimum. We discuss below how commonly these and other
 potential mechanisms may opperate. 
@@
 non-N and tail-dependent copula structure in ecology. First, the skewness, through time, 
-of spatially averaged time series is affected by the tail dependence of the relationships
-between the local quantities being averaged (such as local abundances of a pest or a resource). 
-Thus right-tail-dependent local quantities 
+of the spatial average of time series from different locations (e.g., local pest abundance
+time series) is affected by the tail dependence of the relationships
+between the local quantities. Thus right-tail-dependent local quantities 
 tend to produce "spiky" spatially averaged time series, with large outbreaks. 
 Second, metapopulation extinction risk can be accentuated or reduced by the tail dependence
-properties of environmental variables that influence the populations. Finally, we presented
-simulation results in an Appendix showing that spatial Taylor's law can be strongly affected by
-tail dependence between local populations. In our view, our results show that
-many ecological phenomena cannot be completely understood without using a more complete, 
-copula-based approach to analyzing dependence between variables, and we advocate that
-such approaches become standard tools in ecology.
-
-Put ideas here for discussion paragraphs:
-1) Maybe a para on how likely one is to get copula structure via the Liebig mechanism - considers
-what people have written on nonlinear effects and empirical evidence for Liebig's law and so on
-2) Maybe a para on how likely one is to get copula structure via the Moran mechanism - considers
-work on copula structure in environmental variables. Future work: copula structure of environmental
-variables known to be important in ecology, including how it may be changing and whether it has a 
-geography.
-3) a para on stability of the total biomass time series, and skewness - sets up Shya's later paper as
-future work
-4) We have only given two mechanisms, but there are likely to be others. Maybe I can think
+properties of environmental variables that influence the populations. Finally, 
+simulation results in an Appendix showed that spatial Taylor's law can be strongly influenced by
+tail dependence between local populations. In our view, our results show that, for many
+ecological phenomena, a complete understanding is impossible without 
+characterizing dependence between variables using copulas.
+
+<!--8) A para about what is available in the copula and VineCopula packages.
+Maybe make it a broader para about emthods, and state in there that we developed some new surrogate methods that may be broadly useful for researches trying to incorporate copula approaches.-->
+Copula approaches are
+well developed, statistically [@joe2014_dependence; @nelsen2006_copula], and 
+convenient computer implementations exist (e.g., the `copula` and 
+`VineCopula` packages in R). Ecologists need only to become aware of these tools
+and understand them conceptually to start applying them broadly to ecological
+problems. We created a variety of interrelated surrogate (randomization) procedures
+(Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
+that build upon existing methods and may be especially useful in ecological applications.
+
+<!--***DAN: some mechanism paragraphs follow-->
+
+<!--Maybe a para on how likely one is to get copula structure via the Moran mechanism - considers
+work on copula structure in environmental variables. Maybe mention possible future work: 
+study copula structure of environmental variables (such as temperature and precip) that are known 
+to be important in ecology, including how such copula structure may be changing and whether it 
+has a geography of the kind studied in Walter et al.-->
+<!--***DAN: Shya, please read or reread all the papers we have that say anything about 
+copula structure of dependence between an environmental variable measured in different places, 
+and then write this paragraph.-->
+One may expect our Moran-like mechanism for non-N copula structure and tail dependence to 
+operate commonly, because Moran effects are likely ubiquitous [@Liebhold2004], and 
+non-N and tail-dependent copula structures are common in environmental variables.
+<!--***DAN: Shya, this topic sentence may have to be modified if you find non-N and tail-dependent
+copula structures in environmental variables measured in different places are not so common,
+or have not been studied much. Butif you find some papers, and they show non-N structure, you 
+can just summarize their results here.-->
+<!--***DAN: Idea to work in if it fits: if large weather events (e.g., large rain events) 
+are also widespread, that should lead to right-tail dependence in, e.g., rainfall.-->
+
+<!--Maybe a para on how likely one is to get copula structure via the Liebig mechanism - considers
+what people have written on nonlinear effects and empirical evidence for Liebig's law and so on-->
+There are also substantial reasons to expect that our second mechanism operates commonly. 
+The second mechanism is the one that involves asymmetric or nonlinear influences of 
+environmental variables. Liebig's law of the minimum and the idea of the "limiting 
+nutrient" for organism or population growth are fundamental and dominant paradigms in ecology.
+Furthermore, innumerable studies document nonlinear influences of environmental
+variables on important ecological quantities at all levels of organization, e.g., the 
+influence of temperature on: 
+individual metabolism and autotroph growth rates 
+[@GarciaCarreras2018]; dispersal [@Oconnor2007]; population growth 
+[@Savage2004]; and microbial and ecosystem methane flux [@YvonDurocher2014].
+<!--***DAN: you may want to combine the above para with the previous one in the end, using an
+intro something like "There are good reasons to suspect that the two mechanisms we suggested
+operate extremely commonly in real ecosystems"-->
+
+<!--4) We have only given two mechanisms, but there are likely to be others, and discoverig
+additional mechanisms of copula structure is certain worthy of future work. Maybe I can think
 of some ideas and make a paragraph.
-5) If Van gives us some ideas, they may fit in the Discussion.
+5) If Van gives us some ideas, they may fit in the Discussion here.
 6) Maybe give Holder a bunch of draws from a bivariate dist with standard normals marginals but with 
@@
 tail dependent steps.
-7) A para that sets up Shya's extinction risk paper - asks about importance of this effect for real 
+-->
+
+<!--***DAN: some consequences paragraphs next-->
+
+<!--a para on stability of the total biomass time series in a community, and skewness - sets up Shya's later paper as future work-->
+
+<!--A para that sets up Shya's extinction risk paper - asks about importance of this effect for real 
 species and whether it needs to be incorporated into PVA, and in what cases.
@@
 Also, there may be copula structure between vital rates, and it is not accounted for in typical stochastic matrix model based PVA.
-Does it need to be?
-8) A para about what is available in the copula and VineCopula packages.
-Maybe make it a broader para about emthods, and state in there that we developed some new surrogate methods that may be broadly useful for researches trying to incorporate copula approaches.
+Does it need to be?-->
+Although we have demonstrated that tail dependence between environmental variables
+can influence extinction risk, substantial work remains to be done regarding the 
+circumstances under which this result holds, and its magnitude in different scenarios. 
+First, we used 
+a non-density dependent model to examine extinction risk; does tail dependence
+in environmental variables still influence extinction risk when density dependence
+plays a role? Second, we considered metapopulation extinction risk, but the whole field 
+of population viability analysis [@REF-Morris and Doak] stochastic matrix modelling
+
+<!--Maybe a para from Holder on his thoughts in the importance of copulas for
+macroevolutionary biology-->
+
+<!--Conclusion of some kind-->
 

### 2019-03-28 SupportingInformation.Rmd

@@
 plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
+     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
 text(0.1,0.95,"A",cex=2)
@@
 plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,pch=20,col=rgb(0,0,0,0.5),cex=0.6)
+     cex.axis=1,pch=20,col=rgb(0,0,0,0.2),cex=0.6)
 text(0.1,0.95,"B",cex=2)
@@
 plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,pch=20,col=rgb(0,0,0,0.5),cex=0.6)
+     cex.axis=1,pch=20,col=rgb(0,0,0,0.2),cex=0.6)
 text(0.1,0.95,"A",cex=2)
@@
 plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,col=rgb(0,0,0,0.5),pch=20,cex=0.6)
+     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
 text(0.1,0.95,"B",cex=2)
@@
 \includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.9$.\label{fig_ms_asym_b_0.5}}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.5$.\label{fig_ms_asym_b_0.5}}
 \end{center}
@@
 \includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.9$.\label{fig_ms_asym_b_-0.5}}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.5$.\label{fig_ms_asym_b_-0.5}}
 \end{center}

### 2019-03-28 Paper.Rmd
working on Discussion more

@@
 of population viability analysis [@REF-Morris and Doak] stochastic matrix modelling 
+<!--***DAN: Shya, don't work on this para, I am still working on it-->
 

### 2019-03-29 Paper.Rmd
further work on Discussion

--- 2019-03-28 Paper.Rmd
@@
 potential mechanisms may opperate. 
-<!--***DAN: make sure to do so-->
 We also demonstrated important consequences of 
@@
 
-<!--***DAN: some mechanism paragraphs follow-->
+<!--some mechanism paragraphs follow-->
 
@@
 
-<!--A para that sets up Shya's extinction risk paper - asks about importance of this effect for real 
-species and whether it needs to be incorporated into PVA, and in what cases.
-If the copula structure of environmental variables changes (is it changing?) what will be the impact on extinction risk?
-Also, there may be copula structure between vital rates, and it is not accounted for in typical stochastic matrix model based PVA.
-Does it need to be?-->
+<!--A para that sets up Shya's extinction risk paper-->
 Although we have demonstrated that tail dependence between environmental variables
@@
 First, we used 
-a non-density dependent model to examine extinction risk; does tail dependence
+a non-density dependent model to examine extinction risk. Does tail dependence
 in environmental variables still influence extinction risk when density dependence
-plays a role? Second, we considered metapopulation extinction risk, but the whole field 
-of population viability analysis [@REF-Morris and Doak] stochastic matrix modelling
+plays a role? Second, we considered metapopulation extinction risk, but the large field 
+of population viability analysis (PVA) via stochastic 
+matrix modelling [@Caswell2000; @MorrisDoak2002]
+uses a framework in which a single population's vital rates (e.g., life-stage-specific 
+fecundity and survival rates) are considered to
+vary stochastically through time due to environmental variation. Do relationships
+between different vital rates exhibit tail dependence, and does tail dependence 
+influence extinction risk of a single population?
+PVA is one of the routes by which a species can be added to the influential
+International Union for the Conservation of Nature (IUCN) red list of threatened
+species. It seems worthwhile for future research to investigate the extent to which 
+copula structure in dependences between vital rates may occur and may influence
+estimates of risk. Finally, is the copula structure of environmental variables
+changing through time, and, if so, how do such changes influence extinction risks?
+Climate change is known to amplify the factors that lead to extreme weather events
+[@Hansen2012]. Extreme events, by definition, produce values in the right tails of distributions
+of environmental variables. Extreme weather events also tend to be spatially extensive, so
+increased frequency of extreme weather events may correspond to changes in the tail dependence
+environmental variables measured in different places, and we showed such changes
+can affect extinction risk. It seems
+important for future work to explore whether such effects are of sufficient 
+commonness and magnitude to interest conservationists.
 
 <!--Maybe a para from Holder on his thoughts in the importance of copulas for
-macroevolutionary biology-->
-
-<!--Conclusion of some kind-->
+macro-scale evolutionary biology?-->
+
+<!--Brief conclusion of some kind-->
 
@@
 
-We thank the many contributors to the large datasets we have used; D.
+We thank the many contributors to the large datasets we used; D.
 Stevens and P.

### 2019-04-01 Paper.Rmd
Discussion : tasks done, added comment

@@
 has a geography of the kind studied in Walter et al.-->
+
 <!--***DAN: Shya, please read or reread all the papers we have that say anything about 
@@
 or have not been studied much. Butif you find some papers, and they show non-N structure, you 
-can just summarize their results here.-->
+can just summarize their results here.
+***Shya: Done but please check the last line of this para : 
+@Goswami2018 this paper tests model with 3 different copula : Frank, Gaussian, Plackett, as far as I know all are symmetric, 
+so non-N copula may be best fit but in this case I could not understand how tha tail-dep. thing would fit.--> 
+Using a bivariate mixed model approach with copula, 
+past study [@Serinaldi2008] examined the spatial dependence 
+of rainfall observations (7 years with 30 min. time resolution) 
+within a rain gauge network located in Central Italy and concluded that Gumbel or Student 2-copulas 
+were possible candidates for modelling bivariate dependence at several time scale. A relatively recent 
+study [@liu2016probabilistic] used copula 
+based approach in assessing the probability of vegetation related drought across China. They found 
+majority of non-Normal copula as best fit for the dependence between vegatation index (NDVI) 
+and climatic variables (e.g. precipitation and temperature) for each month of the growing season
+(April-October). Another long-term study (1950–2014) in the Loess Plateau (LP) of China [@She2018]
+showed that Gumbel copula (with upper tail dependence) can better model the spatial dependence 
+(in the 93.1$\%$ case of the 
+total grid considered) between the drought variables 
+(duration and severity) in the study area. Bivariate copula analysis was also used in forecasting the 
+co-occurrence of extreme events (flood or drought) 
+over North Sikkim Himalaya using spatial data set at different time scales [@Goswami2018].
+
 <!--***DAN: Idea to work in if it fits: if large weather events (e.g., large rain events) 

### 2019-04-01 Paper.Rmd
edits to Discussion

--- 2019-03-29 Paper.Rmd
@@
 non-N and tail-dependent copula structure in ecology. First, the skewness, through time, 
-of the spatial average of time series from different locations (e.g., local pest abundance
-time series) is affected by the tail dependence of the relationships
+of the spatial average of time series from different locations (e.g., average regional pest abundance computed from local pest abundances) is affected by the tail dependence of the relationships
 between the local quantities. Thus right-tail-dependent local quantities 
@@
 properties of environmental variables that influence the populations. Finally, 
-simulation results in an Appendix showed that spatial Taylor's law can be strongly influenced by
-tail dependence between local populations. In our view, our results show that, for many
-ecological phenomena, a complete understanding is impossible without 
+simulation results in Appendix \ref{SM-Consequence_TL_simulation} 
+showed that spatial Taylor's law can be strongly influenced by
+tail dependence between local populations. In our view, our results demonstrate that, for many
+ecological phenomena, complete understanding is impossible without 
 characterizing dependence between variables using copulas.
@@
 well developed, statistically [@joe2014_dependence; @nelsen2006_copula], and 
-convenient computer implementations exist (e.g., the `copula` and 
+computer implementations exist (e.g., the `copula` and 
 `VineCopula` packages in R). Ecologists need only to become aware of these tools
-and understand them conceptually to start applying them broadly to ecological
-problems. We created a variety of interrelated surrogate (randomization) procedures
+and understand them conceptually to start applying them broadly. 
+We created a variety of interrelated randomization procedures
 (Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
-that build upon existing methods and may be especially useful in ecological applications.
+that build upon existing methods and may be especially useful for ecological applications.
 

### 2019-04-01 Paper.Rmd
edits for Discussion

--- 2019-04-01 Paper.Rmd
@@
 computer implementations exist (e.g., the `copula` and 
-`VineCopula` packages in R). Ecologists need only to become aware of these tools
+`VineCopula` packages in R). Ecologists need only become aware of these tools
 and understand them conceptually to start applying them broadly. 
@@
 has a geography of the kind studied in Walter et al.-->
-<!--***DAN: Shya, please read or reread all the papers we have that say anything about 
-copula structure of dependence between an environmental variable measured in different places, 
-and then write this paragraph.-->
 One may expect our Moran-like mechanism for non-N copula structure and tail dependence to 
-operate commonly, because Moran effects are likely ubiquitous [@Liebhold2004], and 
-non-N and tail-dependent copula structures are common in environmental variables.
-<!--***DAN: Shya, this topic sentence may have to be modified if you find non-N and tail-dependent
-copula structures in environmental variables measured in different places are not so common,
-or have not been studied much. Butif you find some papers, and they show non-N structure, you 
-can just summarize their results here.-->
-<!--***DAN: Idea to work in if it fits: if large weather events (e.g., large rain events) 
-are also widespread, that should lead to right-tail dependence in, e.g., rainfall.-->
+operate commonly, because Moran effects are common [@Liebhold2004], and 
+non-N and tail-dependent copula structures have often been found in prior analyses of 
+environmental variables. If large meteorological events are also spatially widespread, then
+one would expect environmental variables associated with these events to take extreme
+values simultaneously across large spatial areas, producing tail dependence in measurements
+made at different locations. Non-extreme values, not associated with extreme events, may
+be more associated with local weather phenomena, and therefore 
+less prone to being correlated across large areas.
+<!--***Shya: Done but please check the last line of this para : 
+@Goswami2018 this paper tests model with 3 different copula : Frank, Gaussian, Plackett, as far as I know all are symmetric, 
+so non-N copula may be best fit but in this case I could not understand how tha tail-dep. thing would fit. 
+***DAN: that's OK, the purpose here is to give background on what has been done, it does not all have to pit perfectly. And also, the topic sentence just says non-N, which that is.--> 
+Using a bivariate mixed model approach with copulas, 
+@Serinaldi2008 examined the spatial dependence 
+of rainfall observations (7 years with 30 minute time resolution) 
+within a rain gauge network in Central Italy. Gumbel or Student 2-copulas 
+were possible candidates for modelling bivariate dependence at several time scales, and 
+neither of these is an N copula. @liu2016probabilistic used a copula 
+based approach in assessing the probability of vegetation related drought across China. 
+<!--***DAN: Shya, what do you mean by "vegetation related drought"? Is this a kind of drought
+that is somehow caused by the vegetation? That's what it sounds like, and it sounds like what
+they studied was the probability that such a drought occurred. If that's not what you mean, please
+rephrase.-->
+In a majority of their cases, they found 
+non-N copula structure in the dependence between a vegatation index 
+and climatic variables (e.g., precipitation and temperature). 
+Another long-term study (1950–2014) in the Loess Plateau of China [@She2018]
+showed that a Gumbel copula (which has upper tail dependence) could effectively model 
+the spatial dependence of drought variables 
+(duration and severity) in the study area. Bivariate copula analysis was also used in forecasting the 
+co-occurrence of extreme events (flood or drought) 
+over the North Sikkim Himalayas, India, using spatial data sets at different 
+time scales [@Goswami2018].
 
@@
 what people have written on nonlinear effects and empirical evidence for Liebig's law and so on-->
-There are also substantial reasons to expect that our second mechanism operates commonly. 
+There are also good reasons to expect that our second mechanism operates commonly. 
 The second mechanism is the one that involves asymmetric or nonlinear influences of 
-environmental variables. Liebig's law of the minimum and the idea of the "limiting 
-nutrient" for organism or population growth are fundamental and dominant paradigms in ecology.
+environmental variables. Liebig's law of the minimum and the idea of a "limiting 
+nutrient" of organism or population growth are fundamental and dominant paradigms in ecology.
 Furthermore, innumerable studies document nonlinear influences of environmental
@@
 [@Savage2004]; and microbial and ecosystem methane flux [@YvonDurocher2014].
-<!--***DAN: you may want to combine the above para with the previous one in the end, using an
-intro something like "There are good reasons to suspect that the two mechanisms we suggested
-operate extremely commonly in real ecosystems"-->
+Environmental variables influence not only ecological populations, but also physiological,
+community, and ecosystem functioning measurements. So our two mechanism may well operate
+to produce spatial copula structure in a wide variety of ecological measurements.
 
@@
 -->
+Other mechanisms that we have not considered likely also operate to produce copula structure
+in ecological datasets.
+For instance, we detected non-N structure and tail dependence in our 
+metabolic rate-versus-body mass datasets,
+but the mechanisms discussed above do not apply directly in a macroevolutionary context.
+We hope by enumerating two potential mechanisms we help inspire additional theoretical and
+empirical research on the
+potentially numerous mechanisms of copula structure that may apply in diverse ecological datasets.
+<!--***DAN: beef this up if/when we get inputs from Mark and Van-->
 
@@
 <!--a para on stability of the total biomass time series in a community, and skewness - sets up Shya's later paper as future work-->
+Our results showed that the skewness of the 
+spatial average of local time series from different locations is influenced 
+by their tail dependence. But the logic of this result should also apply to any collection of time
+series, whether associated with different locations in space or not, that are summed or averaged. 
+Another potential application is time series of all species from a single community, e.g., all
+plants in a quadrat or survey site (REF) or all herbivorous plankton in a lake (REF). A huge 
+literature has focused on synchrony versus compensatory dynamics between such time series (REFS), and 
+the influence of such interspecific relationships on the variability, through time, of total-community
+properties such as total biomass. Typically, variability through time of total-community
+properties is measured with the coefficient of variation, but the skewness of such a time series
+may also be of interest: it can characterize the extent to
+which community properties show "spikiness," i.e., uncharacteristically extreme values. 
+Therefore, future work on copula structure of intraspecific relationships in communities and its
+influence on community properties seems likely to us to be a potentially very exciting avenue.
 
@@
 
-<!--Brief conclusion of some kind-->
-
 # Acknowledgments

### 2019-04-01 Paper.Rmd
adding one line description before each r chunk stating what it does

@@
 $u_i+v_i<2u_b$ and the line $v=u$.
-<!--***DAN: Shyamolina, is this actually right, i.e., did we use the average, or did we 
-use the sum divided by n? Whichever way we did it is OK, but please check the code and revise the above 
-text if necessary and insert a comment answering this comment. I tried to check this, myself, 
-but I was having a hard time finding where the statistics are implemented. Please also
-indicate in your response comment what line of what file has the implementation of these
-statistics.
-***Shya: In code we used the sum of squared distances divided by n(=number of points lying within two boundaries) which I think okay to be called as "average-squared distance" or (to be precised) average of squared distances of points from 1:1 line.
-Reference : CopulaFunctions_flexible.R file, line 24.
-***DAN: OK, good, I am leaving these comments here as a reminder-->
+
 Unlike $\cor_{l_b,u_b}$ and 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
+\caption{Upper panels show raw data plots for (A) soil C and N data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
+for birds, (C) $\log_{10}$(BMR) vs.
+$\log_{10}$(body mass) for mammals, and (D) above-ground plant biomass vs.
+Shannon's index (H) from Cedar Creek.
+Bottom panels show the corresponding copula plots.
+See text for units used in upper panels.
+\label{fig_biv_multi_raw_cop}}
+\end{center}
+\end{figure}
+
 Across essentially all the sub-fields and levels of organization within ecology for which we 
@@
 Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}).
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
-\caption{Upper panels show raw data plots for (A) soil C and N data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
-for birds, (C) $\log_{10}$(BMR) vs.
-$\log_{10}$(body mass) for mammals, and (D) above-ground plant biomass vs.
-Shannon's index (H) from Cedar Creek.
-Bottom panels show the corresponding copula plots.
-See text for units used in upper panels.
-\label{fig_biv_multi_raw_cop}}
-\end{center}
-\end{figure}
 
@@
 rejected in every case. 
-<!--***DAN: Right, Shyamolina? I am assuming that because "Indep" does not 
-appear anywhere in Table S11, this is correct, but please confirm or correct.
-***Shya: Yes, you are right, all the pairwise sampling locations for aphid data rejected independence test and show copula structures.
-***DAN: Good, thanks, leaving these comments here as a reminder-->
 Model selection among copulas was carried out separately for each pair of locations.
@@
 aphid species for which we had data other than the green spruce and leaf-curling plum aphids,
-as well as for the 21 plankton taxa for which we had abundance data other 
+as well as for the 21 plankton taxa for which we had abundance data other than the
 \textit{Ceratium furca} (results not shown). The conclusions were generally supported that 
@@
 
-<!--
-```{r tab_bivar_summary, echo=F, results='markup'}
-tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
-knitr::kable(tab_bivar_summary, 
-             format='pandoc',
-             caption = "Summary of results for bivariate datasets : \\textbf{p.Independence.test}: p-value from `BiCopIndTest`, \\textbf{minimumAIC.bestfit.copula}: best fit copula based on minimum AIC, \\textbf{Bestfit.copula.AICw}: AIC-weight of that best fit copula, \\textbf{Normal.copula.AICw}: AIC-weight of Normal copula, \\textbf{p.CvM}: p-value from `BiCopGofTest` for Cramer-von Mises method, \\textbf{p.KS}: p-value from `BiCopGofTest` for Kolmogorov-Smirnov method,\\textbf{LT.model.avg}: AIC-weights model averaged lower tail dependence, \\textbf{UT.model.avg}: AIC-weights model averaged upper tail dependence, \\textbf{(LT-UT).model.avg}:  AIC-weights model averaged (lower-upper) tail dependence.
-Statistic $\\Ps_l$ - $\\Ps_u$ for the actual copula data and for Kendall-correlation preserving $1000$ Normal surrogates were computed and ranked in ascending order.
-Rank for real data \\textbf{Rank.(Pl-Pu)}, if were $>975$ or $<25$, then indicated a significantly ($95\\%$ CI) different lower or upper tail-dependent measures based on Normal-null hypothesis.
-\\label{tab_bivar_summary}",
-             booktabs=TRUE)
-```
-
-```{r tab_multivar_summary, echo=F, results='markup'}
-
-tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
-knitr::kable(tab_multivar_summary, 
-             format='pandoc',
-             caption = "Summary of results for multivariate datasets : \\textbf{num.pairs}: total number of location pairs (excluding self-comparison) considered, \\textbf{num.non-indep}: how many location pair copula are non-independent among them, \\textbf{num.non-N}: number of non-Normal copula as best fit (based on highest AIC-weight), \\textbf{Avg.best.AICw}: average AIC-weights of best fit copula from all location pair, \\textbf{Avg.N.AICw}: average AIC-weights of Normal copula from all location pair, \\textbf{num.good.fit}: number of location pair for which best fit copula have p-values $>$ 0.01
-(both Cramer-von Mises and Kolmogorov-Smirnov method), \\textbf{0.025-CI.LT}: lower and \\textbf{0.975-CI.LT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) lower tail dependence from all location pair, \\textbf{0.025-CI.UT}: lower and \\textbf{0.975-CI.UT}: upper quantile ($95\\%$CI) of model averaged (AIC-weight based) upper tail dependence from all location pair, \\textbf{num.LT-UT>0}: number of location pair which showed stronger lower tail dependence than upper, Last $6$ rows summarized the nonparametric analysis results for $\\Ps$ statistics : lower and upper quantile ($95\\%$CI) of $\\Ps_l$, $\\Ps_u$, $\\Ps_l-\\Ps_u$ based on resampling location pair.\\label{tab_multivar_summary}",
-             booktabs=TRUE)
-```
--->
-
-<!--bivariate table summary in latex format-->
+<!--bivariate table summary-->
 ```{r tab_bivar_summaryl, echo=F, results='asis',message=F}
@@
 
-<!--multivariate table summary in latex format-->
+<!--multivariate table summary-->
 ```{r tab_multivar_summaryl, echo=F, results='asis',message=F}
@@
 
-<!--***DAN: Shya, this has got to be an error. Results should be symmetric
-with respect to the choice of g_1 and g_2 because the model is symmetric. Let's discuss.
-I think there is likely a bug affecting Figs S51 and 51, and Tables S47-48, and 
-possibly all results based on the same model.-->
-
 \begin{figure}[!h]
@@
 less prone to being correlated across large areas.
-<!--***Shya: Done but please check the last line of this para : 
-@Goswami2018 this paper tests model with 3 different copula : Frank, Gaussian, Plackett, as far as I know all are symmetric, 
-so non-N copula may be best fit but in this case I could not understand how tha tail-dep. thing would fit. 
-***DAN: that's OK, the purpose here is to give background on what has been done, it does not all have to pit perfectly. And also, the topic sentence just says non-N, which that is.--> 
 Using a bivariate mixed model approach with copulas, 

### 2019-04-01 SupportingInformation.Rmd
adding one line description before each r chunk stating what it does

@@
 seed<-101
-# families<-c(1,3:10,13,14,16:20)
 source("mtime.R") #A function needed for caching
@@
 
-```{r cop_fig_intro,echo=F,results="hide",cache=T, cache.extra=list(seed,mtime("copsurrognd.R"))}
-library(copula)
-library(VineCopula)
-library(mvtnorm)
-set.seed(seed)
-
-source("copsurrognd.R")
-
-#x <- read.table("./Data/dataset/x.txt")
-#y <- read.table("./Data/dataset/y.txt")
-#mat<-cbind(x,y)
-#colnames(mat)<-c("x","y")
-
-#***constants
-N<-200 #length of time series
-n<-2 #number of sampling locations
-numdat<-1 #the number of datasets to get from the normal copula
-rho<-0.8 #normal-copula parameter for dependence between locations
-#this is a covariance, see below call to rmvnorm
-
-#***generate the starting data, which has normal copula structure
-#because it comes originally from a multivariate normal, before 
-#transformation
-sig<-matrix(rho,n,n)
-diag(sig)<-1
-m<-rmvnorm(N*numdat,mean=rep(0,n),sigma=sig)
-m<-array(m,c(N,numdat,n))
-m<-aperm(m,c(1,3,2)) # 2nd dim is normally distributed e.g. hist(m[,1,])
-gshape<-7.5
-gscale<-1 #gamma parameters
-
-cop_to_marg_trans<-function(x){
-  return(qgamma(x,shape=gshape,scale=gscale)) #this is where we choose the marginals,
-  #should be something skewed
-}
-# pnorm(m) makes normal to uniform
-# then qgamma(...) makes uniform to gamma distributed
-m<-cop_to_marg_trans(pnorm(m))
-
-ccop<-claytonCopula(2,2)
-csurrog<-array(NA,dim(m))
-csurrog[,,1]<-copsurrognd(m[,,1],targetcop=ccop,numsurrog=1)
-mat<-csurrog[,,1]
-
-pdf("./Results/rawdata.pdf",height = 4, width = 4)
-scale<-2
-xhist <- hist(mat[,1], breaks=30, plot=FALSE)
-yhist <- hist(mat[,2], breaks=30, plot=FALSE) 
-top <- max(c(xhist$counts, yhist$counts)) 
-xrange <- range(mat[,1])
-yrange <- range(mat[,2])
-nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
-op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0))  
-plot(mat[,1], mat[,2], xlim=xrange, ylim=yrange, xlab="x", ylab="y", cex.lab=1.5)
-legend("bottomright","raw data plot",bty="n",cex=1.5,text.col="magenta")
-op<-par(mar=c(0,3,1,1)) 
-barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
-op<-par(mar=c(3,0,1,1)) 
-barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
-par(op)
-dev.off()
-
-pdf("./Results/copdata.pdf",height = 4, width = 4)
-copmat<-pobs(mat)
-xhist <- hist(copmat[,1], breaks=30, plot=FALSE)
-yhist <- hist(copmat[,2], breaks=30, plot=FALSE) 
-top <- max(c(xhist$counts, yhist$counts)) 
-xrange <- c(0,1)
-yrange <- c(0,1)
-nf <- layout(matrix(c(2,0,1,3),2,2,byrow=TRUE), c(4,1), c(1,4), TRUE) 
-op<-par(mar=c(3,3,1,1),mgp=c(2, 1, 0)) 
-plot(copmat[,1], copmat[,2], xlim=xrange, ylim=yrange, xlab="u=rank(x)/(n+1)", ylab="v=rank(y)/(n+1)",cex.lab=1.5)
-legend("bottomright","copula plot",bty="n",cex=1.5,text.col="magenta")
-op<-par(mar=c(0,3,1,1)) 
-barplot(xhist$counts/scale, axes=FALSE, ylim=c(0, top), space=0) 
-op<-par(mar=c(3,0,1,1)) 
-barplot(yhist$counts/scale, axes=FALSE, xlim=c(0, top), space=0, horiz=TRUE)
-par(op)
-dev.off()
-```
-
+<!--generating plots for pedagog figures-->
 ```{r cop_pedagog_fig,echo=F,results="hide"}
@@
 
-<!--The raw data look like this:-->
-```{r load_data_Loecke, echo=F, results='hide'}
+<!--reading data and generaing plot for soilCN-->
+```{r load_and_plot_data_soilCN, echo=F, results='hide'}
 #fig.cap="Plot of the raw soil C and N data.\\label{fig_raw_cop_soilCN}(A)", fig.height=3, fig.width=3}
@@
 
-<!--The copula looks like this:-->
-```{r rankandplot_Loecke, echo=F, results='hide'}
+<!--generating copula plot with soilCN data-->
+```{r copula_plot_soilCN, echo=F, results='hide'}
 source("getcopula.R")
@@
 
+<!--generating plots with sampling locations for soilCN data-->
 ```{r plot_SoilCNmapUSA, echo=F, results='hide'}
@@
 
+<!--reading Birds' BMR data and generating raw, copula plots with it-->
 ```{r read_BMR_birds_and_copulaplot, echo=F, results='hide'}
@@
 
+<!--reading Mammals' BMR data and generating raw, copula plots with it-->
 ```{r read_BMR_mammals_and_copulaplot, echo=F, results='hide'}
@@
 
+<!--reading ceder creek data and generating raw and copula plots with it-->
 ```{r read_ceder_creek_data, echo=F, results="hide"}
@@
 
+<!--reading aphid and plankton data, then generating plots showing sampling locations-->
 ```{r map_plot_aphid_plankton,results="hide",echo=F}
@@
 
+<!--Generating a single plot containing raw and copula plots for all bivariate data-->
 ```{r make_data_plot_fig,echo=FALSE,results="hide"}
@@
 But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available).
-We used data from 14 locations which had at least 45 years time series data (see Figure \ref{SM-fig_plankton_map}).
+We used data from 14 locations which had at least 45 years time series data (see Figure \ref{fig_plankton_map}).
 
@@
 
+<!--generating table with information about 20 aphid species-->
 ```{r tab_aphid_info, echo=F, results='asis',message=F}
@@
 
+<!--generating table with information about 22 plankton species-->
 ```{r tab_plankton_info, echo=F, results='asis',message=F}
@@
 ```
-<!--***DAN: These names differ from those used in Emma's 2016 Glob Change Biol paper.
-Need to check with Chris Reid. Have to get this right.-->
-
+
+<!--reading multivariate methane flux data-->
 ```{r read_methane_data, echo=F, results="markup"}
@@
 methane_raw<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")
-#class(methane_raw) #it's a data.frame for multivariable data
-#knitr::kable(head(methane_raw,3),format='pandoc',
-#             caption = "First few rows of raw methane data.\\label{tab_rawdata_methane}",
-#             booktabs=T) # display first few rows
-```
-
+```
+
+<!--embedding plot with sampling locations : soilCN, aphids, planktons-->
 \begin{figure}[!h]
@@
 
-<!--The message=F and warning=F flags are to suppress warnings that come when you pass a parameter
+<!--Simulation for testing non-parametric statistics with different copulas :
+The message=F is to suppress warnings that come when you pass a parameter
 of 0 to the Frank or Clayton copulas. These are technicallly boundary cases for these families.
 They correspond to the independent copula. The packages we use provide the indepdent copula, but
-issue a warning. Since anyway we want the independent copula, we can suppress the warning. We ran this
-without the warning and it is OK to supress it.-->
-<!-- The lines you are talking about appeared as message not as warning(actualy no warning appered for this chunk), so we have to set message=F here-->
+issue a message.-->
 ```{r stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
@@
 \end{figure}
-
-<!--This r-chunk generates figure necessary for maintext Paper-->
-```{r fig_stat_testing_for_maintext, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"),mtime("StatTestNumericResults.RData"))}
-set.seed(seed) 
-source("fig_Paper_stat_testing.R")
-#load("./Results/stat_results/stat_testing/StatTestNumericResults.RData")
-resultsloc<-"./Results/stat_results/stat_testing/"
-
-copname<-"C"
-reslist<-res_pt35_rankT
-
-statname<-"CorlmCoru"
-call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
-statname<-"PlmPu"
-call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
-statname<-"D2umD2l"
-call_fig_Paper_stat_testing(resultsloc=resultsloc,statname=statname,copname=copname,reslist=reslist, numsims=100)
-```
 
@@
 
-<!--Display model selection results in a table-->
+<!--Display model selection results for soilCN data in a table-->
 ```{r tab_soilCNfit, echo=F, results="markup"}
@@
 
-<!--Now make and display a table, this one about asymmetry of tail dependence-->
+<!--Function to generate a table for npa results-summary with bivariate data, this one about asymmetry of tail dependence-->
 ```{r make_table_stat_fn, echo=F}
@@
   ptxt<-''
-  if (fracval>.5*numsurrog){                        # ?DCR why 0.5 factor?
+  if (fracval>.5*numsurrog){                        
     ptxt<-paste(">",fracval,sep='')
@@
 
+<!--Displaying a table with npa summary results for soilCN data-->
 ```{r tab_soilCN_asym,echo=F,results="markup"}
@@
 
+<!--Simulations for model selection results with Birds' BMR data-->
 ```{r bivmodselect_birds_BMR, echo=F, cache=T, cache.extra=list(seed,cop_b,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
+<!--This chunk is for npa testing results for Birds' BMR dataset-->
 ```{r stat_bmr_birds_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_b,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
@@
 
+<!--Displaying a table with npa summary results for Birds' BMR data-->
 ```{r tab_bmr_birds_asym, echo=F, results="markup"}
@@
 
+<!--Simulations for model selection results with Mammals' BMR data-->
 ```{r bivmodselect_mammals_BMR, echo=F, cache=T, cache.extra=list(seed,cop_m,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
+<!--This chunk is for npa testing results for Mammals' BMR dataset-->
 ```{r stat_bmr_mammals_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_m,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
@@
 
+<!--Displaying a table with npa summary results for Mammals' BMR data-->
 ```{r tab_bmr_mammals_asym, echo=F, results="markup"}
@@
 
+<!--Simulations for model selection results with Cedar creek data-->
 ```{r bivmodselect_cc, echo=F, cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
-<!--Display model selection results for Ceder-creek data for year 2000 in a table-->
+<!--Display model selection results for Cedar-creek data for year 2000 in a table-->
 ```{r tab_cc2000fit, echo=F, results='markup'}
@@
 
+<!--This chunk is for npa testing results for Cedar creek dataset-->
 ```{r stat_cc_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
@@
 
+<!--Displaying a table with npa summary results for Cedar creek data-->
 ```{r tab_cc_asym_2000, echo=F, results="markup"}
@@
 
-<!--Load the raw data-->
+<!--Load the raw data for aphid-->
 ```{r read_aphid_data, echo=F}
@@
 
+<!--Simulation for model selection and non-parametric analysis with aphids count data for all 20 sp.-->
 ```{r modelsel_and_npa_aphid_count_allsp, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 
-<!--Do the model selection for aphid count data and then showing results in Tables-->
+<!--Do the model selection for Green spruce aphid's abundance data and then showing results in Tables-->
 ```{r RES_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
+<!--table for Best-fitting copulas, green spruce aphid abundance data-->
 ```{r tab_count_bestcop, echo=F, results="markup"}
@@
 
+<!--table for AIC-weights of best-fitting copulas, green spruce aphid abundance data-->
 ```{r tab_count_AICw, echo=F, results='markup'}
@@
 
+<!--table for AIC-weights of Normal copula used for fitting, green spruce aphid abundance data-->
 ```{r tab_count_AICw_normal, echo=F, results='markup'}
@@
 
+<!--table for p-values from goodness of fit test (Cramer-von Mises-based test) with best-fit copula, green spruce aphid abundance data-->
 ```{r tab_count_p_CvM, echo=F, results='markup'}
@@
 
+<!--table for p-values from goodness of fit test (Kolmogorov-Smirnov-based test) with best-fit copula, green spruce aphid abundance data-->
 ```{r tab_count_p_KS, echo=F, results='markup'}
@@
 
-<!--\newpage-->
+<!--table for model averaged lower tail dep., green spruce aphid abundance data-->
 ```{r tab_count_LTdep_AICw, echo=F, results='markup'}
@@
 
-<!--\newpage-->
+<!--table for model averaged upper tail dep., green spruce aphid abundance data-->
 ```{r tab_count_UTdep_AICw, echo=F, results='markup'}
@@
 
-<!--\newpage-->
+<!--table for model averaged lower- minus upper-tail dep., green spruce aphid abundance data-->
 ```{r tab_count_LTmUT, echo=F, results='markup'}
@@
 
-<!--Do the main computations for nonparametric analysis of aphid count data-->
+<!--Do the main computations for nonparametric analysis of green spruce aphid abundance data-->
 ```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 
-<!--Show the statistical results in a table-->
+<!--Show the statistical results in a table for green spruce aphid abundance data-->
 ```{r tab_count_resamp,echo=F,results="markup"}
@@
 
+<!--Simulation for model selection and non-parametric analysis with aphids first flight data for all 20 sp.-->
 ```{r modelsel_and_npa_aphid_ff_allsp, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 
-<!--Do the model selection for aphid ff data and then showing results in Tables-->
+<!--Do the model selection for leaf plum curling aphid first flight data and then showing results in Tables-->
 ```{r RES_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
+<!--table for Best-fitting copulas, leaf plum curling aphid first flight data-->
 ```{r tab_ff_bestcop, echo=F, results="markup"}
@@
 
+<!--table for AIC-weights of best-fitting copulas, leaf plum curling aphid first flight data-->
 ```{r tab_ff_AICw, echo=F, results='markup'}
@@
 
-<!--\newpage-->
+<!--table for AIC-weights of Normal copula used for fitting, leaf plum curling aphid first flight data-->
 ```{r tab_ff_AICw_normal, echo=F, results='markup'}
@@
 
+<!--table for p-values from goodness of fit test (Cramer-von Mises-based test) with best-fit copula, 
+leaf plum curling aphid first flight data-->
 ```{r tab_ff_p_CvM, echo=F, results='markup'}
@@
 
+<!--table for p-values from goodness of fit test (Kolmogorov-Smirnov-based test) with best-fit copula, 
+leaf plum curling aphid first flight data-->
 ```{r tab_ff_p_KS, echo=F, results='markup'}
@@
 
+<!--table for model averaged lower tail dep., leaf plum curling aphid first flight data-->
 ```{r tab_ff_LTdep_AICw, echo=F, results='markup'}
@@
 
+<!--table for model averaged upper tail dep., leaf plum curling aphid first flight data-->
 ```{r tab_ff_UTdep_AICw, echo=F, results='markup'}
@@
 
+<!--table for model averaged lower- minus upper-tail dep., leaf plum curling aphid first flight data-->
 ```{r tab_ff_LTmUT, echo=F, results='markup'}
@@
 
-<!--Do the main computations for nonparametric analysis of aphid ff data-->
+<!--Do the main computations for nonparametric analysis of leaf plum curling aphid first flight data-->
 ```{r stat_aphid_ff, echo=F, cache=T,  cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 
-<!--display the main statistical results in a table-->
+<!--Show the statistical results in a table for leaf plum curling aphid first flight data-->
 ```{r tab_ff_resamp,echo=F,results="markup"}
@@
 
-<!--Load the data-->
+<!--Load the data : planktons-->
 ```{r read_plankton_north_sea_data, echo=F}
@@
 
+<!--Simulation for model selection and non-parametric analysis with plankton data for all 22 sp.-->
 ```{r modelsel_and_npa_pns_allsp, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,longs_pns,lats_pns,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 
-<!--Do the model selection for plankton data and then showing results in Tables-->
+<!--Do the model selection for plankton (Ceratium furca) and then showing results in Tables-->
 ```{r RES_plankton_north_sea, echo=F, cache=T, results="hide", cache.extra=list(seed,data_plankton_north_sea,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
+<!--table for Best-fitting copulas, Ceratium furca abundance data-->
 ```{r tab_pns_bestcop, echo=F, results="markup"}
@@
 
+<!--table for AIC-weights of best-fitting copulas, Ceratium furca abundance data-->
 ```{r tab_pns_AICw,echo=F,results="markup"}
@@
 
+<!--table for AIC-weights of Normal copula used for fitting, Ceratium furca abundance data-->
 ```{r tab_pns_AICw_normal,echo=F,results="markup"}
@@
 
+<!--table for p-values from goodness of fit test (Cramer-von Mises-based test) with best-fit copula, Ceratium furca abundance data-->
 ```{r tab_pns_p_CvM,echo=F,results="markup"}
@@
 
+<!--table for p-values from goodness of fit test (Kolmogorov-Smirnov-based test) with best-fit copula, Ceratium furca abundance data-->
 ```{r tab_pns_p_KS,echo=F,results="markup"}
@@
 
+<!--table for model averaged lower tail dep., Ceratium furca abundance data-->
 ```{r tab_pns_LTdep_AICw,echo=F,results="markup"}
@@
 
+<!--table for model averaged upper tail dep., Ceratium furca abundance data-->
 ```{r tab_pns_UTdep_AICw,echo=F,results="markup"}
@@
 
+<!--table for model averaged lower- minus upper-tail dep., Ceratium furca abundance data-->
 ```{r tab_pns_LTmUT,echo=F,results="markup"}
@@
 
-<!--Do the main computations for nonparametric analysis of plankton data-->
+<!--Do the main computations for nonparametric analysis of Ceratium furca abundance data-->
 ```{r stat_plankton_north_sea, echo=F, cache=T, cache=T,  cache.extra=list(seed,data_plankton_north_sea,longs_pns,lats_pns,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 
-<!--display the main statistical results in a table-->
+<!--Show the statistical results in a table for Ceratium furca abundance data-->
 ```{r tab_pns_resamp,echo=F,results="markup"}
@@
 
-<!--Load and format the data-->
+<!--Load and format the data : methane flux-->
 ```{r read_format_methane_data, echo=F}
@@
 
-<!--Do the main analysis-->
+<!--Do the model selection for methane flux data and then showing results in Tables-->
 ```{r RES_methane, echo=F, cache=T, results="hide", cache.extra=list(seed,methane_data,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
+<!--table for Best-fitting copulas, methane flux data-->
 ```{r tab_methane_bestcop, echo=F, results="markup"}
@@
 
+<!--table for AIC-weights of best-fitting copulas, methane flux data-->
 ```{r tab_methane_AICw,echo=F,results="markup"}
@@
 
+<!--table for AIC-weights of Normal copula used for fitting, methane flux data-->
 ```{r tab_methane_AICw_normal,echo=F,results="markup"}
@@
 
+<!--table for p-values from goodness of fit test (Cramer-von Mises-based test) with best-fit copula, methane flux data-->
 ```{r tab_methane_p_CvM,echo=F,results="markup"}
@@
 
+<!--table for p-values from goodness of fit test (Kolmogorov-Smirnov-based test) with best-fit copula, methane flux data-->
 ```{r tab_methane_p_KS,echo=F,results="markup"}
@@
 
+<!--table for model averaged lower tail dep., methane flux data-->
 ```{r tab_methane_LTdep_AICw,echo=F,results="markup"}
@@
 
+<!--table for model averaged upper tail dep., methane flux data-->
 ```{r tab_methane_UTdep_AICw,echo=F,results="markup"}
@@
 
+<!--table for model averaged lower- minus upper-tail dep., methane flux data-->
 ```{r tab_methane_LTmUT,echo=F,results="markup"}
@@
 
-<!--Do the main computations for nonparametric analysis of methane-flux data-->
+<!--Do the main computations for nonparametric analysis of methane flux data-->
 ```{r stat_methane, echo=F, cache=T, cache=T,  cache.extra=list(seed,methane_data,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 
-```{r read_stat_methane,echo=F}
-stat_methane<-readRDS("./Results/stat_results/stat_methane/stat_methane.RDS")
-```
-
-<!--display the main statistical results in a table-->
+<!--Show the statistical results in a table for methane flux data-->
 ```{r tab_methane_resamp,echo=F,results="markup"}
@@
 
+<!--Function to generate a table containing summary results for all bivariate datasets-->
 ```{r bivar_tab_summary_fn,echo=F}
@@
 
+<!--Function to generate a table containing summary results for all multivariate datasets-->
 ```{r multivar_tab_summary_fn,echo=F}
@@
 
-```{r sample_plot_aphid_count_ff,echo=F,results="hide"}
-#-------------------------------------------------------------------------------
-# code to plot raw data and copula plot for aphid count @location pair (4,11)
-# bestfit cop : SJ
-
-d1<-data_aphid_count$`Green spruce aphid`[[4]]
-d2<-data_aphid_count$`Green spruce aphid`[[11]]
-pdf("./Results/Aphid_count_raw_loc_4_11.pdf",height = 8,width=8)
-op<-par(mar=c(10,10, 0.2, 2.1),mgp=c(3,1,0))
-plot(d1$Dat,d2$Dat,type="p",xlab="",
-     ylab="Green-spruce aphid \ncount (loc. 11)",cex.lab=3.5,cex.axis=2.5)
-mtext(paste("Green-spruce aphid \ncount (loc. 4)"),line=-38,cex=3.5)
-par(op)
-dev.off()
-
-source("./vivj_matrix.R")
-m<-vivj_matrix(d_allsp = data_aphid_count,sp=10,i=4,j=11)
-pdf("./Results/Aphid_count_cop_loc_4_11.pdf",height = 8,width=8)
-op<-par(mar=c(6,6, 0.2, 2.1),mgp=c(3,1,0))
-plot(m[,1],m[,2],type="p",xlab="u",ylab="v",
-     cex.lab=4,cex.axis=3,
-     xlim=c(0,1),ylim=c(0,1),asp=1)
-rect(0,0,1,1)
-lines(c(0,1),c(0,1),type='l',col="grey")
-par(op)
-dev.off()
-
-# code to plot raw data and copula plot for aphid ff @location pair (2,5)
-# bestfit cop : J
-d1<-data_aphid_ff$`Leaf-curling plum aphid`[[2]]
-d2<-data_aphid_ff$`Leaf-curling plum aphid`[[5]]
-pdf("./Results/Aphid_ff_raw_loc_2_5.pdf",height = 8,width=8)
-op<-par(mar=c(10,10, 0.2, 2.1),mgp=c(3,1,0))
-plot(d1$Dat,d2$Dat,type="p",xlab="",ylab="Leaf-plum curling's \nfirst-flight (loc. 5)",cex.lab=3.5,cex.axis=2.5)
-mtext(paste("Leaf-plum curling's \nfirst-flight (loc. 2)"),line=-38,cex=3.5)
-par(op)
-dev.off()
-
-m<-vivj_matrix(d_allsp = data_aphid_ff,sp=11,i=2,j=5)
-pdf("./Results/Aphid_ff_cop_loc_2_5.pdf",height = 8,width=8)
-op<-par(mar=c(6,6, 0.2, 2.1),mgp=c(3,1,0))
-plot(m[,1],m[,2],type="p",xlab="u",ylab="v",
-     cex.lab=4,cex.axis=3,
-     xlim=c(0,1),ylim=c(0,1),asp=1)
-rect(0,0,1,1)
-lines(c(0,1),c(0,1),type='l',col="grey")
-par(op)
-dev.off()
-```
-
-```{r cause4copula_pGOF_500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-#source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-#pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_500.pdf", width=8, height=3)
-#op<-par(mfrow=c(1,2))
-
-#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=500,BS=1000)
-
-#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=500,BS=1000)
-
-#par(op)
-
-#op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
-#plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-#legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", #"magenta", "green2"),
-#       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
-#       bty = "n") 
-#par(op)
-#dev.off()
-```
-
-```{r cause4copula_pGOF_2500, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-#source("Cause4copula.R")
-#-----------------
-#set.seed(seed)
-#---------------------------------------
-
-#pdf("./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf", width=8, height=3)
-#op<-par(mfrow=c(1,2))
-
-#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "spearman",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="spearman",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "spearman",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="spearman",num_keep_last=2500,BS=1000)
-
-#Plotter_Cause4copula_GOF(N=5000,fcode=3,method = "kendall",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=13,method="kendall",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=4,method = "kendall",num_keep_last=2500,BS=1000)
-#Plotter_Cause4copula_GOF(N=5000,fcode=14,method="kendall",num_keep_last=2500,BS=1000)
-
-#op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 1, 0, 0), new = TRUE)
-#plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-#legend("top", c("noise copula,", "population copula,",  "Gof test : p-value(CvM),", "Gof test : p-value(KS)"), col = c("red","blue", #"magenta", "green2"),
-#       cex = 0.9, pch = c(2, 6, 16, 16), xpd = TRUE, horiz = TRUE, inset = c(0,0),
-#       bty = "n") 
-#par(op)
-#dev.off()
-```
-
+<!--Following two chunks examine Moran-like mechanism with clayton copula and then generate plots with the simulated results-->
 ```{r cause4copula_stat_ar1_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
-```{r cause4copula_stat_ar1_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-# This chunk stores results for ar1 model : survival clayton
-model<-"ar1"
-fcode<-13
-nsd<-1
-
-# for positive beta
-params<-c(0.5) # this parameter is beta in ar1 model
-
-resloc<-"./Results/Cause4copula_results/"
-tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-resloc<-paste(tempo,"/",sep="")
-# for survival clayton with spearman and kendall method
-ar1_SC_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
-                                            num_keep_last=500,resloc=resloc,
-                                            params=params,p0=c(0,0),model=model)
-
-saveRDS(ar1_SC_spear_beta0.5,paste(resloc,"ar1_SC_spear_beta0.5.RDS",sep=""))
-
-ar1_SC_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
-                                           num_keep_last=500,resloc=resloc,
-                                           params=params,p0=c(0,0),model=model)
-
-saveRDS(ar1_SC_kend_beta0.5,paste(resloc,"ar1_SC_kend_beta0.5.RDS",sep=""))
-# for negative beta
-params<-c(-0.5) # this parameter is beta in ar1 model
-
-resloc<-"./Results/Cause4copula_results/"
-tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-# for survival clayton with spearman and kendall method
-ar1_SC_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
-                                               num_keep_last=500,resloc=resloc,
-                                               params=params,p0=c(0,0),model=model)
-saveRDS(ar1_SC_spear_negbeta0.5,paste(resloc,"ar1_SC_spear_negbeta0.5.RDS",sep=""))
-
-ar1_SC_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
-                                              num_keep_last=500,resloc=resloc,
-                                              params=params,p0=c(0,0),model=model)
-saveRDS(ar1_SC_kend_negbeta0.5,paste(resloc,"ar1_SC_kend_negbeta0.5.RDS",sep=""))
-```
-
-```{r plot_cause4copula_stat_ar1_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SC_spear_beta0.5,ar1_SC_spear_negbeta0.5,ar1_SC_kend_beta0.5,ar1_SC_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-# This chunk plots results for ar1 model : clayton
-
-model<-"ar1"
-fcode<-13
-resloc0<-"./Results/Cause4copula_results/"
-
-# for positive beta
-params<-c(0.5) # this parameter is beta in ar1 model
-
-tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-resloc<-paste(tempo,"/",sep="")
-
-readres<-readRDS(paste(resloc,"ar1_SC_spear_beta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-readres<-readRDS(paste(resloc,"ar1_SC_kend_beta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-
-# for negative beta
-params<-c(-0.5) # this parameter is beta in ar1 model
-tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-resloc<-paste(tempo,"/",sep="")
-
-readres<-readRDS(paste(resloc,"ar1_SC_spear_negbeta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-readres<-readRDS(paste(resloc,"ar1_SC_kend_negbeta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-```
-
-```{r cause4copula_stat_ar1_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-# This chunk stores results for ar1 model : clayton
-
-model<-"ar1"
-fcode<-4
-nsd<-1
-
-# for positive beta
-params<-c(0.5) # this parameter is beta in ar1 model
-resloc<-"./Results/Cause4copula_results/"
-
-tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-resloc<-paste(tempo,"/",sep="")
-# for gumbel with spearman and kendall method
-ar1_G_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
-                                              num_keep_last=500,resloc=resloc,
-                                              params=params,p0=c(0,0),model=model)
-saveRDS(ar1_G_spear_beta0.5,paste(resloc,"ar1_G_spear_beta0.5.RDS",sep=""))
-
-ar1_G_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
-                                             num_keep_last=500,resloc=resloc,
-                                             params=params,p0=c(0,0),model=model)
-
-saveRDS(ar1_G_kend_beta0.5,paste(resloc,"ar1_G_kend_beta0.5.RDS",sep=""))
-
-# for negative beta
-resloc<-"./Results/Cause4copula_results/"
-params<-c(-0.5) # this parameter is beta in ar1 model
-tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-# for gumbel with spearman and kendall method
-ar1_G_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
-                                              num_keep_last=500,resloc=resloc,
-                                              params=params,p0=c(0,0),model=model)
-saveRDS(ar1_G_spear_negbeta0.5,paste(resloc,"ar1_G_spear_negbeta0.5.RDS",sep=""))
-
-ar1_G_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
-                                             num_keep_last=500,resloc=resloc,
-                                             params=params,p0=c(0,0),model=model)
-
-saveRDS(ar1_G_kend_negbeta0.5,paste(resloc,"ar1_G_kend_negbeta0.5.RDS",sep=""))
-```
-
-```{r plot_cause4copula_stat_ar1_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_G_spear_beta0.5,ar1_G_spear_negbeta0.5,ar1_G_kend_beta0.5,ar1_G_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-# This chunk plots results for ar1 model : clayton
-
-model<-"ar1"
-fcode<-4
-resloc0<-"./Results/Cause4copula_results/"
-
-# for positive beta
-params<-c(0.5) # this parameter is beta in ar1 model
-
-tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-resloc<-paste(tempo,"/",sep="")
-
-readres<-readRDS(paste(resloc,"ar1_G_spear_beta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-readres<-readRDS(paste(resloc,"ar1_G_kend_beta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-
-# for negative beta
-params<-c(-0.5) # this parameter is beta in ar1 model
-tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-resloc<-paste(tempo,"/",sep="")
-
-readres<-readRDS(paste(resloc,"ar1_G_spear_negbeta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-readres<-readRDS(paste(resloc,"ar1_G_kend_negbeta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-```
-
-```{r cause4copula_stat_ar1_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-# This chunk stores results for ar1 model : survival clayton
-model<-"ar1"
-fcode<-14
-nsd<-1
-
-# for positive beta
-params<-c(0.5) # this parameter is beta in ar1 model
-
-resloc<-"./Results/Cause4copula_results/"
-tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-resloc<-paste(tempo,"/",sep="")
-# for survival gumbel with spearman and kendall method
-ar1_SG_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
-                                              num_keep_last=500,resloc=resloc,
-                                              params=params,p0=c(0,0),model=model)
-saveRDS(ar1_SG_spear_beta0.5,paste(resloc,"ar1_SG_spear_beta0.5.RDS",sep=""))
-
-ar1_SG_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
-                                             num_keep_last=500,resloc=resloc,
-                                             params=params,p0=c(0,0),model=model)
-
-saveRDS(ar1_SG_kend_beta0.5,paste(resloc,"ar1_SG_kend_beta0.5.RDS",sep=""))
-
-# for negative beta
-params<-c(-0.5) # this parameter is beta in ar1 model
-
-resloc<-"./Results/Cause4copula_results/"
-tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-# for survival gumbel with spearman and kendall method
-ar1_SG_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
-                                              num_keep_last=500,resloc=resloc,
-                                              params=params,p0=c(0,0),model=model)
-saveRDS(ar1_SG_spear_negbeta0.5,paste(resloc,"ar1_SG_spear_negbeta0.5.RDS",sep=""))
-
-ar1_SG_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
-                                             num_keep_last=500,resloc=resloc,
-                                             params=params,p0=c(0,0),model=model)
-
-saveRDS(ar1_SG_kend_negbeta0.5,paste(resloc,"ar1_SG_kend_negbeta0.5.RDS",sep=""))
-```
-
-```{r plot_cause4copula_stat_ar1_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SG_spear_beta0.5,ar1_SG_spear_negbeta0.5,ar1_SG_kend_beta0.5,ar1_SG_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-# This chunk plots results for ar1 model : clayton
-
-model<-"ar1"
-fcode<-14
-resloc0<-"./Results/Cause4copula_results/"
-
-# for positive beta
-params<-c(0.5) # this parameter is beta in ar1 model
-
-tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-resloc<-paste(tempo,"/",sep="")
-
-readres<-readRDS(paste(resloc,"ar1_SG_spear_beta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-readres<-readRDS(paste(resloc,"ar1_SG_kend_beta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-
-# for negative beta
-params<-c(-0.5) # this parameter is beta in ar1 model
-tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-resloc<-paste(tempo,"/",sep="")
-
-readres<-readRDS(paste(resloc,"ar1_SG_spear_negbeta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-readres<-readRDS(paste(resloc,"ar1_SG_kend_negbeta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-```
-
-```{r cause4copula_stat_ar1_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-# This chunk stores results for ar1 model : clayton
-
-model<-"ar1"
-fcode<-6
-nsd<-1
-
-# for positive beta
-params<-c(0.5) # this parameter is beta in ar1 model
-resloc<-"./Results/Cause4copula_results/"
-
-tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-resloc<-paste(tempo,"/",sep="")
-# for joe with kendall method only
-
-ar1_J_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
-                                             num_keep_last=500,resloc=resloc,
-                                             params=params,p0=c(0,0),model=model)
-
-saveRDS(ar1_J_kend_beta0.5,paste(resloc,"ar1_J_kend_beta0.5.RDS",sep=""))
-
-# for negative beta
-resloc<-"./Results/Cause4copula_results/"
-params<-c(-0.5) # this parameter is beta in ar1 model
-tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-# for joe with kendall method only
-
-ar1_J_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
-                                             num_keep_last=500,resloc=resloc,
-                                             params=params,p0=c(0,0),model=model)
-
-saveRDS(ar1_J_kend_negbeta0.5,paste(resloc,"ar1_J_kend_negbeta0.5.RDS",sep=""))
-```
-
-```{r plot_cause4copula_stat_ar1_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_J_kend_beta0.5,ar1_J_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-# This chunk plots results for ar1 model : clayton
-
-model<-"ar1"
-fcode<-6
-resloc0<-"./Results/Cause4copula_results/"
-
-# for positive beta
-params<-c(0.5) # this parameter is beta in ar1 model
-
-tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-resloc<-paste(tempo,"/",sep="")
-
-readres<-readRDS(paste(resloc,"ar1_J_kend_beta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-
-# for negative beta
-params<-c(-0.5) # this parameter is beta in ar1 model
-tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-resloc<-paste(tempo,"/",sep="")
-
-readres<-readRDS(paste(resloc,"ar1_J_kend_negbeta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-```
-
-```{r cause4copula_stat_ar1_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-# This chunk stores results for ar1 model : survival clayton
-model<-"ar1"
-fcode<-16
-nsd<-1
-
-# for positive beta
-params<-c(0.5) # this parameter is beta in ar1 model
-
-resloc<-"./Results/Cause4copula_results/"
-tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-resloc<-paste(tempo,"/",sep="")
-# for survival joe with kendall method only
-
-ar1_SJ_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
-                                             num_keep_last=500,resloc=resloc,
-                                             params=params,p0=c(0,0),model=model)
-
-saveRDS(ar1_SJ_kend_beta0.5,paste(resloc,"ar1_SJ_kend_beta0.5.RDS",sep=""))
-
-# for negative beta
-params<-c(-0.5) # this parameter is beta in ar1 model
-
-resloc<-"./Results/Cause4copula_results/"
-tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-
-if (!dir.exists(tempo)){
-  dir.create(tempo)
-}
-
-resloc<-paste(tempo,"/",sep="")
-# for survival joe with kendall method only
-
-ar1_SJ_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
-                                             num_keep_last=500,resloc=resloc,
-                                             params=params,p0=c(0,0),model=model)
-
-saveRDS(ar1_SJ_kend_negbeta0.5,paste(resloc,"ar1_SJ_kend_negbeta0.5.RDS",sep=""))
-```
-
-```{r plot_cause4copula_stat_ar1_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SJ_kend_beta0.5,ar1_SJ_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-# This chunk plots results for ar1 model : clayton
-
-model<-"ar1"
-fcode<-16
-resloc0<-"./Results/Cause4copula_results/"
-
-# for positive beta
-params<-c(0.5) # this parameter is beta in ar1 model
-
-tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-resloc<-paste(tempo,"/",sep="")
-
-readres<-readRDS(paste(resloc,"ar1_SJ_kend_beta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-
-# for negative beta
-params<-c(-0.5) # this parameter is beta in ar1 model
-tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
-resloc<-paste(tempo,"/",sep="")
-
-readres<-readRDS(paste(resloc,"ar1_SJ_kend_negbeta0.5.RDS",sep=""))
-Plotter_Cause4copula_stat(res=readres)
-```
-
-```{r cause4copula_stat_ricker_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-# This chunk stores results for ar1 model : clayton
-tag_ricker_C<-"999C" #this is a caching tag used later to cache output from present chunk
-
-model<-"ricker"
-fcode<-3
-
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    if (!dir.exists(tempo)){
-      dir.create(tempo)
-    }
-    resloc<-paste(tempo,"/",sep="")
-
-  # for clayton with spearman and kendall method
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
-                                           num_keep_last=500,resloc=resloc,
-                                           params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_C_spear.RDS",sep=""))
-
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
-                                          num_keep_last=500,resloc=resloc,
-                                          params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_C_kend.RDS",sep=""))
-
-  }
- }
-}
-```
-
-```{r plot_cause4copula_stat_ricker_C, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_C,mtime("Plotter_Cause4copula.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-# This chunk plots results for ar1 model : clayton
-
-model<-"ricker"
-fcode<-3
-resloc0<-"./Results/Cause4copula_results/"
-
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    resloc<-paste(tempo,"/",sep="")
-
-    readres<-readRDS(paste(resloc,"ricker_C_spear.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-    readres<-readRDS(paste(resloc,"ricker_C_kend.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-
-  }
- }
-}
-```
-
-```{r cause4copula_stat_ricker_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-tag_ricker_SC<-"999SC" #this is a caching tag used later to cache output from present chunk
-
-# This chunk stores results for ar1 model : survival clayton
-
-model<-"ricker"
-fcode<-13
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    if (!dir.exists(tempo)){
-      dir.create(tempo)
-    }
-    resloc<-paste(tempo,"/",sep="")
-
-  # for survival clayton with spearman and kendall method
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
-                                           num_keep_last=500,resloc=resloc,
-                                           params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_SC_spear.RDS",sep=""))
-
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
-                                          num_keep_last=500,resloc=resloc,
-                                          params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_SC_kend.RDS",sep=""))
-  }
- }
-}
-```
-
-```{r plot_cause4copula_stat_ricker_SC, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SC,mtime("Plotter_Cause4copula.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-
-model<-"ricker"
-fcode<-13
-resloc0<-"./Results/Cause4copula_results/"
-
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    resloc<-paste(tempo,"/",sep="")
-
-    readres<-readRDS(paste(resloc,"ricker_SC_spear.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-    readres<-readRDS(paste(resloc,"ricker_SC_kend.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-
-  }
- }
-}
-```
-
-```{r cause4copula_stat_ricker_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-tag_ricker_G<-"999G" #this is a caching tag used later to cache output from present chunk
-
-# This chunk stores results for ar1 model : gumbel
-
-model<-"ricker"
-fcode<-4
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    if (!dir.exists(tempo)){
-      dir.create(tempo)
-    }
-    resloc<-paste(tempo,"/",sep="")
-
-  # for gumbel with spearman and kendall method
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
-                                           num_keep_last=500,resloc=resloc,
-                                           params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_G_spear.RDS",sep=""))
-
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
-                                          num_keep_last=500,resloc=resloc,
-                                          params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_G_kend.RDS",sep=""))
-  }
- }
-}
-```
-
-```{r plot_cause4copula_stat_ricker_G, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_G,mtime("Plotter_Cause4copula.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-
-model<-"ricker"
-fcode<-4
-resloc0<-"./Results/Cause4copula_results/"
-
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    resloc<-paste(tempo,"/",sep="")
-
-    readres<-readRDS(paste(resloc,"ricker_G_spear.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-    readres<-readRDS(paste(resloc,"ricker_G_kend.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-
-  }
- }
-}
-```
-
-```{r cause4copula_stat_ricker_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-tag_ricker_SG<-"999SG" #this is a caching tag used later to cache output from present chunk
-
-# This chunk stores results for ar1 model : survival gumbel
-
-model<-"ricker"
-fcode<-14
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    if (!dir.exists(tempo)){
-      dir.create(tempo)
-    }
-    resloc<-paste(tempo,"/",sep="")
-
-  # for survival gumbel with spearman and kendall method
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
-                                           num_keep_last=500,resloc=resloc,
-                                           params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_SG_spear.RDS",sep=""))
-
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
-                                          num_keep_last=500,resloc=resloc,
-                                          params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_SG_kend.RDS",sep=""))
-  }
- }
-}
-```
-
-```{r plot_cause4copula_stat_ricker_SG, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SG,mtime("Plotter_Cause4copula.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-
-model<-"ricker"
-fcode<-14
-resloc0<-"./Results/Cause4copula_results/"
-
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    resloc<-paste(tempo,"/",sep="")
-
-    readres<-readRDS(paste(resloc,"ricker_SG_spear.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-    readres<-readRDS(paste(resloc,"ricker_SG_kend.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-
-  }
- }
-}
-```
-
-```{r cause4copula_stat_ricker_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-tag_ricker_J<-"999J" #this is a caching tag used later to cache output from present chunk
-
-# This chunk stores results for ar1 model : joe
-
-model<-"ricker"
-fcode<-6
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    if (!dir.exists(tempo)){
-      dir.create(tempo)
-    }
-    resloc<-paste(tempo,"/",sep="")
-
-  # for joe with spearman and kendall method
-
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
-                                          num_keep_last=500,resloc=resloc,
-                                          params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_J_kend.RDS",sep=""))
-  }
- }
-}
-```
-
-```{r plot_cause4copula_stat_ricker_J, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_J,mtime("Plotter_Cause4copula.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-
-model<-"ricker"
-fcode<-6
-resloc0<-"./Results/Cause4copula_results/"
-
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    resloc<-paste(tempo,"/",sep="")
-
-    readres<-readRDS(paste(resloc,"ricker_J_kend.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-
-  }
- }
-}
-```
-
-```{r cause4copula_stat_ricker_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-tag_ricker_SJ<-"999SJ" #this is a caching tag used later to cache output from present chunk
-
-# This chunk stores results for ar1 model : survival joe
-
-model<-"ricker"
-fcode<-16
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    if (!dir.exists(tempo)){
-      dir.create(tempo)
-    }
-    resloc<-paste(tempo,"/",sep="")
-
-  # for survival joe with spearman and kendall method
-
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
-                                          num_keep_last=500,resloc=resloc,
-                                          params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_SJ_kend.RDS",sep=""))
-  }
- }
-}
-```
-
-```{r plot_cause4copula_stat_ricker_SJ, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SJ,mtime("Plotter_Cause4copula.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-
-model<-"ricker"
-fcode<-16
-resloc0<-"./Results/Cause4copula_results/"
-
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    resloc<-paste(tempo,"/",sep="")
-
-    readres<-readRDS(paste(resloc,"ricker_SJ_kend.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-
-  }
- }
-}
-```
-<!--
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/Cause4copula_results/compare_params_pGOF_num_keep_last_2500.pdf}
-\caption[Comparing noise-copula parameters and estimated population copula parameters with $2500$ sampling points]{Results of comparison between the parameters of environmental noise generated from a specific family of copula and estimated sample parameter (with some standard error) of the affected populations (Initially 5000 points are generated but we kept last \textbf{2500} points for both noise and population data for comparison).
-p-values $> 0.05$ (above horizontal dashed line) indicate good fit of population copula with same noise cpoula family.
-\label{fig_Cause4copula_params_GOF_2500}}
-\end{center}
-\end{figure}
--->
-
+<!--Plots for AR(1) with Clayton-->
 \begin{figure}[!h]
@@
 \end{figure}
-<!--***DAN: Shya, can you please change the top legend of this and the similar-format
-figures so instead of saying "noise copula" for the red points and "population copula" 
-for the blue points it says "input noise copula" for the red points and "model output 
-copula" for the blue ones? This is because one of our models is much more general than
-just a population model, so I have changed the langauge in the text to reflect that.
-Please be aware that when you make this change it will mean the next knit will take 
-several hours, unless you go to the effort of reorganizing your code to put the sims
-and the plots in separate chunks. So maybe make this change right before you go home,
-so you can knit over night. Also please let me know, so I can knit overnight, too.
-
-At the same time that you do that above, please see if you can fix the alignment
-between the bottom three panels of Fig S32 (and subsequent analogous figures) so they
-line up with the panels above that.
-***Shya: Done.-->
-
+
+<!--Following two chunks examine Moran-like mechanism with survival clayton copula noises in AR(1) model 
+and then generate plots with the simulated results-->
+```{r cause4copula_stat_ar1_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : survival clayton
+model<-"ar1"
+fcode<-13
+nsd<-1
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+resloc<-"./Results/Cause4copula_results/"
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+# for survival clayton with spearman and kendall method
+ar1_SC_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                            num_keep_last=500,resloc=resloc,
+                                            params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_SC_spear_beta0.5,paste(resloc,"ar1_SC_spear_beta0.5.RDS",sep=""))
+
+ar1_SC_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
+                                           params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_SC_kend_beta0.5,paste(resloc,"ar1_SC_kend_beta0.5.RDS",sep=""))
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+
+resloc<-"./Results/Cause4copula_results/"
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+# for survival clayton with spearman and kendall method
+ar1_SC_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                               num_keep_last=500,resloc=resloc,
+                                               params=params,p0=c(0,0),model=model)
+saveRDS(ar1_SC_spear_negbeta0.5,paste(resloc,"ar1_SC_spear_negbeta0.5.RDS",sep=""))
+
+ar1_SC_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                              num_keep_last=500,resloc=resloc,
+                                              params=params,p0=c(0,0),model=model)
+saveRDS(ar1_SC_kend_negbeta0.5,paste(resloc,"ar1_SC_kend_negbeta0.5.RDS",sep=""))
+```
+
+```{r plot_cause4copula_stat_ar1_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SC_spear_beta0.5,ar1_SC_spear_negbeta0.5,ar1_SC_kend_beta0.5,ar1_SC_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-13
+resloc0<-"./Results/Cause4copula_results/"
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SC_spear_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_SC_kend_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SC_spear_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_SC_kend_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+```
+
+<!--Plots for AR(1) with Survival Clayton-->
 \begin{figure}[!h]
@@
 
+<!--Following two chunks examine Moran-like mechanism with gumbel copula noises in AR(1) model 
+and then generate plots with the simulated results-->
+```{r cause4copula_stat_ar1_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-4
+nsd<-1
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+resloc<-"./Results/Cause4copula_results/"
+
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+# for gumbel with spearman and kendall method
+ar1_G_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                              num_keep_last=500,resloc=resloc,
+                                              params=params,p0=c(0,0),model=model)
+saveRDS(ar1_G_spear_beta0.5,paste(resloc,"ar1_G_spear_beta0.5.RDS",sep=""))
+
+ar1_G_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
+                                             params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_G_kend_beta0.5,paste(resloc,"ar1_G_kend_beta0.5.RDS",sep=""))
+
+# for negative beta
+resloc<-"./Results/Cause4copula_results/"
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+# for gumbel with spearman and kendall method
+ar1_G_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                              num_keep_last=500,resloc=resloc,
+                                              params=params,p0=c(0,0),model=model)
+saveRDS(ar1_G_spear_negbeta0.5,paste(resloc,"ar1_G_spear_negbeta0.5.RDS",sep=""))
+
+ar1_G_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
+                                             params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_G_kend_negbeta0.5,paste(resloc,"ar1_G_kend_negbeta0.5.RDS",sep=""))
+```
+
+```{r plot_cause4copula_stat_ar1_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_G_spear_beta0.5,ar1_G_spear_negbeta0.5,ar1_G_kend_beta0.5,ar1_G_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-4
+resloc0<-"./Results/Cause4copula_results/"
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_G_spear_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_G_kend_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_G_spear_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_G_kend_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+```
+
+<!--Plots for AR(1) with Gumbel-->
 \begin{figure}[!h]
@@
 
+<!--Following two chunks examine Moran-like mechanism with survival gumbel copula noises in AR(1) model 
+and then generate plots with the simulated results-->
+```{r cause4copula_stat_ar1_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : survival clayton
+model<-"ar1"
+fcode<-14
+nsd<-1
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+resloc<-"./Results/Cause4copula_results/"
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+# for survival gumbel with spearman and kendall method
+ar1_SG_spear_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                              num_keep_last=500,resloc=resloc,
+                                              params=params,p0=c(0,0),model=model)
+saveRDS(ar1_SG_spear_beta0.5,paste(resloc,"ar1_SG_spear_beta0.5.RDS",sep=""))
+
+ar1_SG_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
+                                             params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_SG_kend_beta0.5,paste(resloc,"ar1_SG_kend_beta0.5.RDS",sep=""))
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+
+resloc<-"./Results/Cause4copula_results/"
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+# for survival gumbel with spearman and kendall method
+ar1_SG_spear_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="spearman",lb=0,ub=0.1,
+                                              num_keep_last=500,resloc=resloc,
+                                              params=params,p0=c(0,0),model=model)
+saveRDS(ar1_SG_spear_negbeta0.5,paste(resloc,"ar1_SG_spear_negbeta0.5.RDS",sep=""))
+
+ar1_SG_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
+                                             params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_SG_kend_negbeta0.5,paste(resloc,"ar1_SG_kend_negbeta0.5.RDS",sep=""))
+```
+
+```{r plot_cause4copula_stat_ar1_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SG_spear_beta0.5,ar1_SG_spear_negbeta0.5,ar1_SG_kend_beta0.5,ar1_SG_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-14
+resloc0<-"./Results/Cause4copula_results/"
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SG_spear_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_SG_kend_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SG_spear_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+readres<-readRDS(paste(resloc,"ar1_SG_kend_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+```
+
+<!--Plots for AR(1) with Survival Gumbel-->
 \begin{figure}[!h]
@@
 
+<!--Following two chunks examine Moran-like mechanism with joe copula noises in AR(1) model 
+and then generate plots with the simulated results-->
+```{r cause4copula_stat_ar1_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-6
+nsd<-1
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+resloc<-"./Results/Cause4copula_results/"
+
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+# for joe with kendall method only
+
+ar1_J_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
+                                             params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_J_kend_beta0.5,paste(resloc,"ar1_J_kend_beta0.5.RDS",sep=""))
+
+# for negative beta
+resloc<-"./Results/Cause4copula_results/"
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+# for joe with kendall method only
+
+ar1_J_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
+                                             params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_J_kend_negbeta0.5,paste(resloc,"ar1_J_kend_negbeta0.5.RDS",sep=""))
+```
+
+```{r plot_cause4copula_stat_ar1_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_J_kend_beta0.5,ar1_J_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-6
+resloc0<-"./Results/Cause4copula_results/"
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_J_kend_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_J_kend_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+```
+
+<!--Plots for AR(1) with Joe-->
 \begin{figure}[!h]
@@
 
+<!--Following two chunks examine Moran-like mechanism with survival joe copula noises in AR(1) model
+and then generate plots with the simulated results-->
+```{r cause4copula_stat_ar1_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : survival clayton
+model<-"ar1"
+fcode<-16
+nsd<-1
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+resloc<-"./Results/Cause4copula_results/"
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+resloc<-paste(tempo,"/",sep="")
+# for survival joe with kendall method only
+
+ar1_SJ_kend_beta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
+                                             params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_SJ_kend_beta0.5,paste(resloc,"ar1_SJ_kend_beta0.5.RDS",sep=""))
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+
+resloc<-"./Results/Cause4copula_results/"
+tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+
+if (!dir.exists(tempo)){
+  dir.create(tempo)
+}
+
+resloc<-paste(tempo,"/",sep="")
+# for survival joe with kendall method only
+
+ar1_SJ_kend_negbeta0.5<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=nsd,method="kendall",lb=0,ub=0.1,
+                                             num_keep_last=500,resloc=resloc,
+                                             params=params,p0=c(0,0),model=model)
+
+saveRDS(ar1_SJ_kend_negbeta0.5,paste(resloc,"ar1_SJ_kend_negbeta0.5.RDS",sep=""))
+```
+
+```{r plot_cause4copula_stat_ar1_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,ar1_SJ_kend_beta0.5,ar1_SJ_kend_negbeta0.5,mtime("Plotter_Cause4copula_stat.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ar1"
+fcode<-16
+resloc0<-"./Results/Cause4copula_results/"
+
+# for positive beta
+params<-c(0.5) # this parameter is beta in ar1 model
+
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SJ_kend_beta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+
+# for negative beta
+params<-c(-0.5) # this parameter is beta in ar1 model
+tempo<-paste(resloc0,"Cause4copula_stat_results_",model,"_beta_",params,"_",BiCopName(fcode,short=F),sep="")
+resloc<-paste(tempo,"/",sep="")
+
+readres<-readRDS(paste(resloc,"ar1_SJ_kend_negbeta0.5.RDS",sep=""))
+Plotter_Cause4copula_stat(res=readres)
+```
+
+<!--Plots for AR(1) with Survival Joe-->
 \begin{figure}[!h]
@@
 
-<!--adding results for stochastic ricker with weak noise : nsd=0.1-->
+<!--Following two chunks examine Moran-like mechanism with clayton copula noises in Ricker model and 
+then generate plots with the simulated results-->
+```{r cause4copula_stat_ricker_clayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+# This chunk stores results for ar1 model : clayton
+tag_ricker_C<-"999C" #this is a caching tag used later to cache output from present chunk
+
+model<-"ricker"
+fcode<-3
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    if (!dir.exists(tempo)){
+      dir.create(tempo)
+    }
+    resloc<-paste(tempo,"/",sep="")
+
+  # for clayton with spearman and kendall method
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
+                                           params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_C_spear.RDS",sep=""))
+
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_C_kend.RDS",sep=""))
+
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_C, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_C,mtime("Plotter_Cause4copula.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+# This chunk plots results for ar1 model : clayton
+
+model<-"ricker"
+fcode<-3
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_C_spear.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+    readres<-readRDS(paste(resloc,"ricker_C_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
+  }
+ }
+}
+```
+
+<!--Plots for stochastic ricker with weak noise (Clayton copula) : nsd=0.1-->
 \begin{figure}[!h]
@@
 
+<!--Plots for stochastic ricker with strong noise (Clayton copula): nsd=1-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, strong noise, C copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, C copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_1}}
+\end{center}
+\end{figure}
+
+<!--Following two chunks examine Moran-like mechanism with survival clayton copula noises in Ricker model and 
+then generate plots with the simulated results-->
+```{r cause4copula_stat_ricker_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+tag_ricker_SC<-"999SC" #this is a caching tag used later to cache output from present chunk
+
+# This chunk stores results for ar1 model : survival clayton
+
+model<-"ricker"
+fcode<-13
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    if (!dir.exists(tempo)){
+      dir.create(tempo)
+    }
+    resloc<-paste(tempo,"/",sep="")
+
+  # for survival clayton with spearman and kendall method
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
+                                           params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_SC_spear.RDS",sep=""))
+
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_SC_kend.RDS",sep=""))
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_SC, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SC,mtime("Plotter_Cause4copula.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+
+model<-"ricker"
+fcode<-13
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_SC_spear.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+    readres<-readRDS(paste(resloc,"ricker_SC_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
+  }
+ }
+}
+```
+
+<!--Plots for stochastic ricker with weak noise (Survival Clayton copula) : nsd=0.1-->
 \begin{figure}[!h]
@@
 
+<!--Plots for stochastic ricker with strong noise (Survival Clayton copula): nsd=1-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, strong noise, SC copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, SC copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_1}}
+\end{center}
+\end{figure}
+
+<!--Following two chunks examine Moran-like mechanism with gumbel copula noises in Ricker model and 
+then generate plots with the simulated results-->
+```{r cause4copula_stat_ricker_gumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+tag_ricker_G<-"999G" #this is a caching tag used later to cache output from present chunk
+
+# This chunk stores results for ar1 model : gumbel
+
+model<-"ricker"
+fcode<-4
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    if (!dir.exists(tempo)){
+      dir.create(tempo)
+    }
+    resloc<-paste(tempo,"/",sep="")
+
+  # for gumbel with spearman and kendall method
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
+                                           params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_G_spear.RDS",sep=""))
+
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_G_kend.RDS",sep=""))
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_G, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_G,mtime("Plotter_Cause4copula.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+
+model<-"ricker"
+fcode<-4
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_G_spear.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+    readres<-readRDS(paste(resloc,"ricker_G_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
+  }
+ }
+}
+```
+
+<!--Plots for stochastic ricker with weak noise (Gumbel copula) : nsd=0.1-->
 \begin{figure}[!h]
@@
 
+<!--Plots for stochastic ricker with strong noise (Gumbel copula): nsd=1-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, strong noise, G copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, G copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details.\label{fig_Cause4copula_G_kend_sRicker_nsd_1}}
+\end{center}
+\end{figure}
+
+<!--Following two chunks examine Moran-like mechanism with survival gumbel copula noises in Ricker model and 
+then generate plots with the simulated results-->
+```{r cause4copula_stat_ricker_survivalgumbel, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+tag_ricker_SG<-"999SG" #this is a caching tag used later to cache output from present chunk
+
+# This chunk stores results for ar1 model : survival gumbel
+
+model<-"ricker"
+fcode<-14
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    if (!dir.exists(tempo)){
+      dir.create(tempo)
+    }
+    resloc<-paste(tempo,"/",sep="")
+
+  # for survival gumbel with spearman and kendall method
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
+                                           params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_SG_spear.RDS",sep=""))
+
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_SG_kend.RDS",sep=""))
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_SG, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SG,mtime("Plotter_Cause4copula.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+
+model<-"ricker"
+fcode<-14
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_SG_spear.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+    readres<-readRDS(paste(resloc,"ricker_SG_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
+  }
+ }
+}
+```
+
+<!--Plots for stochastic ricker with weak noise (Survival Gumbel copula)  : nsd=0.1-->
 \begin{figure}[!h]
@@
 
+<!--Plots for stochastic ricker with strong noise (Survival Gumbel copula): nsd=1-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, strong noise, SG copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, SG copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details. \label{fig_Cause4copula_SG_kend_sRicker_nsd_1}}
+\end{center}
+\end{figure}
+
+<!--Following two chunks examine Moran-like mechanism with joe copula noises in Ricker model and 
+then generate plots with the simulated results-->
+```{r cause4copula_stat_ricker_joe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+tag_ricker_J<-"999J" #this is a caching tag used later to cache output from present chunk
+
+# This chunk stores results for ar1 model : joe
+
+model<-"ricker"
+fcode<-6
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    if (!dir.exists(tempo)){
+      dir.create(tempo)
+    }
+    resloc<-paste(tempo,"/",sep="")
+
+  # for joe with spearman and kendall method
+
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_J_kend.RDS",sep=""))
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_J, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_J,mtime("Plotter_Cause4copula.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+
+model<-"ricker"
+fcode<-6
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_J_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
+  }
+ }
+}
+```
+
+<!--Plots for stochastic ricker with weak noise (Joe copula) : nsd=0.1-->
 \begin{figure}[!h]
@@
 
+<!--Plots for stochastic ricker with strong noise (Joe copula) : nsd=1-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, strong noise, J copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, J copula. Same format as Figs 
+\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
+see captions of those figures for details.\label{fig_Cause4copula_J_kend_sRicker_nsd_1}}
+\end{center}
+\end{figure}
+
+<!--Following two chunks examine Moran-like mechanism with survival joe copula noises in Ricker model and 
+then generate plots with the simulated results-->
+```{r cause4copula_stat_ricker_survivaljoe, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+tag_ricker_SJ<-"999SJ" #this is a caching tag used later to cache output from present chunk
+
+# This chunk stores results for ar1 model : survival joe
+
+model<-"ricker"
+fcode<-16
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    if (!dir.exists(tempo)){
+      dir.create(tempo)
+    }
+    resloc<-paste(tempo,"/",sep="")
+
+  # for survival joe with spearman and kendall method
+
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_SJ_kend.RDS",sep=""))
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_SJ, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SJ,mtime("Plotter_Cause4copula.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+
+model<-"ricker"
+fcode<-16
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_SJ_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
+  }
+ }
+}
+```
+
+<!--Plots for stochastic ricker with weak noise (Survival Joe): nsd=0.1-->
 \begin{figure}[!h]
@@
 
-<!--adding results for stochastic ricker with strong noise : nsd=1-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/common_legend_cause4copula_stat.pdf}\\
-\vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
-\hspace{-0.8cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, C copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, C copula. Same format as Figs 
-\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
-see captions of those figures for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_1}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
-\vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
-\hspace{-0.8cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, SC copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, SC copula. Same format as Figs 
-\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
-see captions of those figures for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_1}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/common_legend_cause4copula_stat.pdf}\\
-\vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Spearman_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Kendall_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Corl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Coru_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pu_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2u_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2l_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
-\hspace{-0.8cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Gumbel/Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, G copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, G copula. Same format as Figs 
-\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
-see captions of those figures for details.\label{fig_Cause4copula_G_kend_sRicker_nsd_1}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/common_legend_cause4copula_stat.pdf}\\
-\vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Spearman_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Kendall_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pearson_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Coru_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Corl-Coru_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pu_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_Pl-Pu_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2l_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_D2u-D2l_vs_Kendall's Tau.pdf}\\
-\hspace{-0.8cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Survival Gumbel/Survival Gumbel_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, SG copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, SG copula. Same format as Figs 
-\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
-see captions of those figures for details. \label{fig_Cause4copula_SG_kend_sRicker_nsd_1}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/common_legend_cause4copula_stat.pdf}\\
-\vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Spearman_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Kendall_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pearson_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Corl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Coru_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Corl-Coru_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pu_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_Pl-Pu_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2u_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2l_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_D2u-D2l_vs_Kendall's Tau.pdf}\\
-\hspace{-0.8cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_1_r_0.5_K_100_Joe/Joe_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, J copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, J copula. Same format as Figs 
-\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
-see captions of those figures for details.\label{fig_Cause4copula_J_kend_sRicker_nsd_1}}
-\end{center}
-\end{figure}
-
+<!--Plots for stochastic ricker with strong noise (Survival Joe): nsd=1-->
 \begin{figure}[!h]
@@
 
-```{r moran_effect_on_aphid_ff, echo=F, results="hide", cache=T, cache.extra=list(seed,temp_aphid,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-set.seed(seed)
-source("FittingCopula_selective_loc.R")  # top most hierarchy source file
-
-#cat(paste("start-time: ",Sys.time(),"\n"))
-RES_temp_aphid<-RES_single_sp(d_allsp=temp_aphid,sp=1,families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
-                              good_loc_given=T,good_loc=c(2:11))
-#cat(paste("stop-time: ",Sys.time(),"\n")) 
-resloc<-"./Results/fitting_results/"
-saveRDS(RES_temp_aphid,paste(resloc,file="AphidCopulaFit_selecloc_winter_neg_temp.RDS",sep=""))
-
-LTmUT_temp<-(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw)
-
-sum(LTmUT_temp>0,na.rm = T)
-sum(LTmUT_temp<0,na.rm = T) #66 out of 90 : suggests UT in leaf-plum curling aphids firstflight data (90 out of 90 UT)
-```
-
-<!--
-```{r tab_temp_aphid_bestcop, echo=F, results="markup"}
-knitr::kable(RES_temp_aphid$info_ord_copname[,,1], #longtable = T,
-             format='latex',
-             #caption.short = "Best-fitting copulas for each location pair from among those considered, according to AIC, for first flight data for the leaf-curling plum aphid.",
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for **negative average winter temperature data (1958-2013)**.
-Copulas' abbreviations were given in \\nameref{MT-Methods}.
-NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.
-\\label{tab_temp_aphid_bestcop}",
-             booktabs=TRUE
-             )
-```
-
-```{r tab_temp_aphid_AICw, echo=F, results='markup'}
-knitr::kable(RES_temp_aphid$info_ord_AICw[,,1], 
-             format='latex',
-             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_AICw}",
-             booktabs=TRUE
-             )
-```
-
-```{r tab_temp_aphid_AICw_normal, echo=F, results='markup'}
-source("normal_cop_AICw.R")
-knitr::kable(normal_cop_AICw_matrix(r=RES_temp_aphid),
-             format='latex',
-             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for **negative average winter temperature data**, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_temp_aphid_AICw_normal}",
-             booktabs=TRUE
-             )
-```
-
-<!--
-```{r tab_temp_aphid_p_CvM, echo=F, results='markup'}
-knitr::kable(RES_temp_aphid$gfc_p_CvM,
-             format='latex',
-             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**.
-Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_CvM}",
-             booktabs=TRUE
-             )
-```
--->
-
-<!--
-```{r tab_temp_aphid_p_KS, echo=F, results='markup'}
-knitr::kable(RES_temp_aphid$gfc_p_KS,
-             format='latex',
-             caption = "$P$-value results of a Kolmogorov-Smirnov-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for **negative average winter temperature data**.
-Tests were based on bootstrapping as described in Methods.\\label{tab_temp_aphid_p_KS}",
-             booktabs=TRUE
-             )
-```
--->
-
-<!--
-```{r tab_temp_aphid_LTdep_AICw, echo=F, results='markup'}
-knitr::kable(RES_temp_aphid$LTdep_AICw,
-             format='latex',
-             caption = "Model-averaged lower tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTdep_AICw}",
-             booktabs=TRUE
-             )
-```
--->
-
-<!--
-```{r tab_temp_aphid_UTdep_AICw, echo=F, results='markup'}
-knitr::kable(RES_temp_aphid$UTdep_AICw,
-             format='latex',
-             caption = "Model-averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_UTdep_AICw}",
-             booktabs=TRUE
-             )
-```
--->
-
-<!--
-```{r tab_temp_aphid_LTmUT, echo=F, results='markup'}
-knitr::kable(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw,
-             digits = 5,
-             format='latex',
-             caption = "Model averaged lower tail dependence minus model averaged upper tail dependence statistic for each location pair for **negative average winter temperature data**.\\label{tab_temp_aphid_LTmUT}",
-             booktabs=TRUE
-             )
-```
--->
-
+<!--Following two chunks examine asymetric dependence mechanism with model selection and non-parametric statistics approach-->
 ```{r asym_sens_b_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
@@
 
+<!--adding plots from asymmetric dependence analysis results-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$.\label{fig_ms_asym_b_-0.1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.5$.\label{fig_ms_asym_b_0.5}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.5$.\label{fig_ms_asym_b_-0.5}}
+\end{center}
+\end{figure}
+
+<!--Displaying summary table with model selection results from asymmetric dependence analysis-->
 ```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
@@
 
-```{r read_tab_stat_npa_asym,echo=F,results="asis"}
-#knitr::opts_knit$set(kable.force.latex = TRUE)
-#tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.9.RDS")
-
-#for(i in c(1:8)){
-#  temp<-make_table_stat_fn(data=tab_asym[[i]])
-#  temp2<-knitr::kable(temp, booktabs = T, 
-#             format = "latex",
-      #caption.short = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis.",
-#      caption =paste(cnm[i]," : A table entry $<N$(or $>N$) indicates the value of the given statistic on the data was less (or more) than its value on $N$ of the surrogates, so entries of the form $<X$ (or $>X$) for $X$ equal to $975$ or above indicate that upper-tail (or lower-tail) dependence was significantly stronger than lower-tail dependence."))
-#  print(temp2)
-#  cat("\\newpage")
-#}
-```
-
+<!--Displaying summary table with non-parametric approach results of asymmetric dependence analysis-->
 ```{r tab_stat_npa_asym,echo=F,results="markup"}
@@
 and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
-0.9 to 1.\\label{tab_stat_npa_asym}")#%>%
- #            kable_styling()%>%
- #            group_rows(index=setNames(c(3,3,3,3,3,3,3,3), gtext))
- #            group_rows("b=0.1, g$_{1}$", 2,4, escape=F) %>%
- #            group_rows("b=0.1, g$_{2}$", 5,7, escape=F) %>%
- #            group_rows("b=-0.1, g$_{1}$", 8,10, escape=F) %>%
- #            group_rows("b=-0.1, g$_{2}$", 11,13, escape=F) %>%
- #            group_rows("b=0.9, g$_{1}$", 14,16, escape=F) %>%
- #            group_rows("b=0.9, g$_{2}$", 17,19, escape=F) %>%
- #            group_rows("b=-0.9, g$_{1}$", 20,22, escape=F) %>%
- #            group_rows("b=-0.9, g$_{2}$", 23,25, escape=F)
-```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$.\label{fig_ms_asym_b_-0.1}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.5$.\label{fig_ms_asym_b_0.5}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.5$.\label{fig_ms_asym_b_-0.5}}
-\end{center}
-\end{figure}
-
+0.9 to 1.\\label{tab_stat_npa_asym}")
+```
+
+<!--Following 3 chunks examine the consequence of non-Normal copula atructure on skewness using aphid count, firstflight for 20 sp.
+and plankton count data for 22 sp. : these results are currently not showin in either main text or suppmat-->
 ```{r skewness_aphidcount_allsp_plot, echo=F, results="hide", cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
@@
 
+<!--Following chunk examine the consequence of non-Normal copula atructure on skewness using green spruce aphid abundance data-->
 ```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
@@
 
+<!--Following chunk examine the consequence of non-Normal copula atructure on skewness 
+using leaf plum curling aphids' first flight data-->
 ```{r skewness_ff_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
@@
 
+<!--Following chunk examine the consequence of non-Normal copula atructure on skewness 
+using ceratium furca's abundance data-->
 ```{r skewness_pns_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
@@
 
+<!--Following chunk examine the consequence of non-Normal copula atructure on skewness using methane flux data-->
 ```{r skewness_methane_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
@@
 
-<!--make surrogate datasets for TL analysis to follow-->
+<!--Following 4 chunks are examining consequence of non-Normal copula structure on Taylors law using made up data-->
+
+<!--first, make surrogate datasets for TL analysis to follow-->
 ```{r taylors_law_make_surrogs, echo=F, cache=T, cache.extra=list(seed,mtime("copsurrognd.R"))}
@@
 
-<!--do a TL analysis on the data just created-->
+<!--second, do a TL analysis on the data just created-->
 ```{r tl_analysis, echo=F, cache=T, cache.extra=list(kend.surrog,spea.surrog,mtime("mncurvat.R"),mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
@@
 
-<!--Now make some plots, these one for kendall surrogates-->
+<!--third, now make some plots, these one for kendall surrogates-->
 ```{r taylor_plots_kendall, echo=F, results="hide"}
@@
 
-<!--Now make some plots, these one for spearman surrogates-->
+<!--fourth, now make some plots, these one for spearman surrogates-->
 ```{r taylor_plots_spearman, echo=F, results="hide"}
@@
 
+<!--adding plots showing influences of non-Normal copula on spatial Taylors law-->
 \begin{figure}[!h]
@@
 
+<!--Following 2 chunks examine the consequence of non-Normal copula atructure on Taylors Law 
+using real data of aphid count with all 20 sp.-->
 ```{r spatialTL_on_realdata_aphid_count, results="hide", echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("TaylorsLawTools.R"),mtime("mncurvat.R"),mtime("ncsurrog.R"))}
@@
 
+<!--Following 2 chunks examine the effect of non-Normal asymmetric copula on extinction risk-->
 ```{r sim_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
@@
 
+<!--adding plots showing effects of asymmetric copula on extinction risk-->
 \begin{figure}[!h]

### 2019-04-02 SupportingInformation.Rmd
completed documentations for multiple files

@@
 library(captioner)
+library(parallel)
+busycores<-detectCores()-2 # for parallell computing with mclapply your computer will use all cores except two.
 ```
@@
 issue a message.-->
-```{r stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
+```{r stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,busycores,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
 source("TestStats_multivar_dataset.R")
@@
 coplist<-callfn$coplist
-ncores<-3
-res_pt35_rankF<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=F,mc.cores=ncores)
-res_pt35_rankT<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=T,mc.cores=ncores)
-res_pt1000_rankF<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=F,mc.cores=ncores)
-res_pt1000_rankT<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=T,mc.cores=ncores)
+
+res_pt35_rankF<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=F,mc.cores=busycores)
+res_pt35_rankT<-mclapply(X=coplist,FUN=worker,numpts=35,numsims=100,rank=T,mc.cores=busycores)
+res_pt1000_rankF<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=F,mc.cores=busycores)
+res_pt1000_rankT<-mclapply(X=coplist,FUN=worker,numpts=1000,numsims=100,rank=T,mc.cores=busycores)
 save(res_pt35_rankF,res_pt35_rankT,res_pt1000_rankF,res_pt1000_rankT,file=paste(resultsloc,"StatTestNumericResults.RData",sep=''))
@@
 <!--Simulation for model selection and non-parametric analysis with aphids count data for all 20 sp.-->
-```{r modelsel_and_npa_aphid_count_allsp, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+```{r modelsel_and_npa_aphid_count_allsp, echo=F, cache=T, cache.extra=list(seed,busycores,data_aphid_count,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
 
@@
                                  families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
-                                 good_loc_given=F,good_loc=NA,mc.cores=3)
+                                 good_loc_given=F,good_loc=NA,mc.cores=busycores)
 saveRDS(RES_all_sp_aphid_count,"./Results/fitting_results/RES_all_sp_aphid_count.RDS")
@@
 <!--Simulation for model selection and non-parametric analysis with aphids first flight data for all 20 sp.-->
-```{r modelsel_and_npa_aphid_ff_allsp, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+```{r modelsel_and_npa_aphid_ff_allsp, echo=F, cache=T, cache.extra=list(seed,busycores,data_aphid_ff,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
 
@@
                                  families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=30,
-                                 good_loc_given=F,good_loc=NA,mc.cores=3)
+                                 good_loc_given=F,good_loc=NA,mc.cores=busycores)
 saveRDS(RES_all_sp_aphid_ff,"./Results/fitting_results/RES_all_sp_aphid_ff.RDS")
@@
 <!--Simulation for model selection and non-parametric analysis with plankton data for all 22 sp.-->
-```{r modelsel_and_npa_pns_allsp, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,longs_pns,lats_pns,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+```{r modelsel_and_npa_pns_allsp, echo=F, cache=T, cache.extra=list(seed,busycores,data_plankton_north_sea,longs_pns,lats_pns,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("NonParamStat.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
 
@@
                               families=c(1,3:10,13,14,16:20),level=0.05,data_pt_thrs=45,
-                              good_loc_given=F,good_loc=NA,mc.cores=3)
+                              good_loc_given=F,good_loc=NA,mc.cores=busycores)
 saveRDS(RES_all_sp_pns,"./Results/fitting_results/RES_all_sp_pns.RDS")

### 2019-04-02 CopulaFunctions.R

@@ -1,5 +1,5 @@
 #-----------------------------------------------------------------------------------------------------------------------
-# THIS CODE CONTAINS STATISTICS FOR LOOKING AT TAIL DEPENDENCE
+# THIS CODE CONTAINS STATISTICS FOR LOOKING AT TAIL DEPENDENCE (using non-parametric statistics)
 #-----------------------------------------------------------------------------------------------------------------------
 
 source("CopulaFunctions_flexible.R")
@@ -14,7 +14,7 @@ CorlCoru<-function(vi,vj){
   return(c(Corl,Coru))
 }
 
-#--------------------------- Stat based on counting points a given distance from the diagonal in lower and upper triangle ---------------------------------------------------------
+#----------------- Stat based on counting points for a given distance from the diagonal in lower and upper triangle -------------
 
 PlPu<-function(vi,vj){
   
@@ -48,7 +48,7 @@ PlPu<-function(vi,vj){
   return(list(Sl_Su_Si_P,Pl,Pu)) 
 }
 
-#------------------------------------------------ stat based on distance from the diagonal -----------------------------------------------------------------------
+#-------------------------------- stat based on distance from the diagonal ---------------------------------------
 
 D2lD2u<-function(vi,vj)
 {
@@ -57,29 +57,4 @@ D2lD2u<-function(vi,vj)
   
   return(c(D2l,D2u))
 }
-
-#---------------------------------------------------------------------------------------------------------------------
-#                                                     CODE ENDS HERE
-#---------------------------------------------------------------------------------------------------------------------
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-#and the others
+#----------------------------------------------------------------------

### 2019-04-02 CopulaFunctions_flexible.R

@@ -1,4 +1,6 @@
-
+#---------------------------------------------------------------------------
+# Non-parametric stats to measure dependence for a bivariate copula
+#---------------------------------------------------------------------------
 #Calculates the mean squared distance between copula
 #points between two bounds and the main diagonal of
 #the unit square.
@@ -29,7 +31,7 @@ D2bds<-function(vi,vj,lb,ub)
   
   return(D2)
 }
-
+#-------------------------------------------------------------------------------------------
 #Calculates the portion of the Spearman correlation 
 #that is due to points in from a copula that are
 #between two bounds.
@@ -64,7 +66,8 @@ Corbds<-function(vi,vj,lb,ub)
   }
   return(res)  
 }
-
+#--------------------------------------------------------------------------------------------------
+# Calculates the P statistics for bivariate copula
 #Args
 #vi, vj       Coordinates of points from a copula
 #lb, ub       Lower and upper bounds between 0 and 1

### 2019-04-02 FittingCopula_selective_loc.R

@@ -1,15 +1,58 @@
-# THIS CODE TESTS FOR A GIVEN COPULA WHICH ARE THE POSSIBLE BEST FIT MODELS ?
-# It fits MULTIVARIATE COPULAS (PAIRWISE) based on MODEL SELECTION APPROACH
-# This code is written to test SPATIAL SYNCHRONY between different lacation pair for a specific species
-#----------------
+# This code fits MULTIVARIATE COPULAS (PAIRWISE) based on MODEL SELECTION APPROACH
+# This code can test SPATIAL SYNCHRONY between different lacation pair for a specific species
+#------------------------------------------------------------------------------------------------------
 library(copula)
 library(VineCopula)
 #-------------------------------------
 source("vivj_matrix.R")
 source("good_loclist.R")
 source(file="OurBiCopSelect.R")
-#---------------------------------------------------------------------------------------
-# This function generates result for single species
+#-----------------------------------------------------------------------------------------------------------------------
+# This function generates model selection results with timeseries taken from multiple locations for a single species
+# Args :
+#       sp : species id
+#       d_allsp : data in specified format as indicated in "vivj_matrix.R"      
+#                   a list (length = total no. of sp.) of 
+#                         a list (length = total no. of locations)
+#                             of a dataframe (with "Year" and "Dat" column)
+#       families : a vector indicating a set of copula families use for model selection approach
+#       level : Significance level to be used for the independence test (generally 0.05)
+#       data_pt_thrs : a threshold on data points available
+#       good_loc_given : logical , if F it will choose selected locations using file "good_loclist.R" 
+#       good_loc : a vector of ids for selected locations which you can supply externally
+
+# Output : a long list with following elements :
+#           num_indep_loc_pair = an integer : number of independent location pairs 
+#                               (e.g. if (loc1,loc2) and (loc2,loc1) bivar. copulas are indep. it will return value 1),
+#           num_neg_cor_loc_pair =an integer : number of negatively correlated location pairs 
+#                                 (e.g. if (loc3,loc4) and (loc4,loc3) bivar. copulas are -vely correlated it will return value 1),
+#           gfc_numBS = matrix with elements showing number of bootstraps used for goodness of fit test for each location pair
+#           gfc_numBS_success = matrix with elements showing number of bootstraps succeeded for goodness of fit test for each location pair
[…]

### 2019-04-02 NonParamStat.R

@@ -1,6 +1,6 @@
 # THIS CODE CALCULATES NON-PARAMETRIC STATS FOR MULTIVARIATE COPULAS (PAIRWISE)
-# This code is written to test synchrony between different LOCATIONS pair for a specific SPECIES
-#---------------------------------------------------------------
+# This code is written to test synchrony between different LOCATIONS pair for a specific SPECIES using non-parametric stats
+#----------------------------------------------------------------------------------------------------------------------------------
 source("CopulaFunctions.R")
 source("vivj_matrix.R")
 source("good_loclist.R")
@@ -9,7 +9,7 @@ library(ncf)
 #Processing function for a copula approach to synchrony
 #
 #Args
-#m : output from vivj_matrix.R
+#m : output from vivj_matrix.R : a bivariate copula matrix
 #
 #Output - A list with these elements
 #ranks        A dataframe with 2 columns, one for 
@@ -96,8 +96,7 @@ copsync<-function(m){
 #1) datmean - The mean of the matrix except for the diagonals
 #2) quantiles - as specified by prob
 #3) fracgt0 - fraction of resampling-based values greater than 0
-#4) fraclt0 - (will just be 1-fracgt0)
-#
+
 #Also generates the histogram plot if ploton=T
 
 resampmn<-function(M,numresamp=10000,prob=c(.025,0.975),ploton=F)
@@ -120,11 +119,7 @@ resampmn<-function(M,numresamp=10000,prob=c(.025,0.975),ploton=F)
   
   #get fracgt0 : fraction of resampling-based values greater than 0
   fracgt0<-length(which(resamp_stat>0))/length(resamp_stat)
-  
-  #get fraclt0 : fraction of resampling-based values greater than 0
-  #  fraclt0<-length(which(resamp_stat<0))/length(resamp_stat)
-  
-  
+
   #generate the plot
   if (ploton){
     hist(resamp_stat,breaks=100)
@@ -134,23 +129,18 @@ resampmn<-function(M,numresamp=10000,prob=c(.025,0.975),ploton=F)
     mtext(paste0("datmean=",round(datmean,4),", q2.5=",round(q2.5,4),", q97.5=",round(q97.5,4))) 
   }
   
-  #  return(list(datmean=datmean,quantiles=quantiles,fracgt0=fracgt0,fraclt0=fraclt0))
   return(list(datmean=datmean,quantiles=quantiles,fracgt0=fracgt0))
 }
 
 #---------------------------------------------------------------------------------------------------------------------------
[…]

### 2019-04-02 vivj_matrix.R

@@ -1,5 +1,13 @@
-# Select any two location pair [i,j] for a copula of species sp
-# This function gives you a matrix with vi and vj as two columns
+# This function takes two time series and makes a bivariate copula
+# Args:
+#     d_allsp : data in specified format, it is 
+#                   a list (length = total no. of sp.) of 
+#                         a list (length = total no. of locations)
+#                             of a dataframe (with "Year" and "Dat" column)
+#     sp : id of species
+#     i, j : integers indicating id of locations
+
+# Output : a bivariate copula as a 2 column matrix
 library(VineCopula)
 vivj_matrix<-function(d_allsp,sp,i,j){
   

### 2019-04-02 SupportingInformation.Rmd
minor edit

@@
 library(parallel)
-busycores<-detectCores()-2 # for parallell computing with mclapply your computer will use all cores except two.
+freecores<-2
+busycores<-detectCores()-freecores # for parallell computing with mclapply your computer will use all cores except the number of freecores.
 ```

### 2019-04-04 Paper.Rmd
Reply to Discussion edits for "vegetation-related drought"

@@
 they studied was the probability that such a drought occurred. If that's not what you mean, please
-rephrase.-->
+rephrase.
+***Shya: I used the exact term "vegetation-related drought" they used in the paper @liu2016probabilistic, I think they 
+wanted to mean that you can monitor the drought risk by tracking change in NDVI spatial greenness with precipitation or temperature (climatic variable), which can be considered as a proxy indicator of vegetation response to climate change.
+They also showed vegatation-drought stress analysis, which I also noticed in some other papers too.
+There are several studies using drought and NDVI data (mainly using correlation, regression) but they reported it as the first study using joint probability for drought-risk assessment.
+I have highlighted the info which may be important to us and sent you in mail.-->
 In a majority of their cases, they found 

### 2019-04-04 SupportingInformation.Rmd
Update SupportingInformation.Rmd
removing skewness results for all sp : aphids count, ff and plankton

@@
 Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016.
-**See and Smyth et al, pending for
+**See and** @smyth2019using, **pending for
 additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had at least 50 dates data.
@@
 
-<!--Following 3 chunks examine the consequence of non-Normal copula atructure on skewness using aphid count, firstflight for 20 sp.
-and plankton count data for 22 sp. : these results are currently not showin in either main text or suppmat-->
-```{r skewness_aphidcount_allsp_plot, echo=F, results="hide", cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
-
-# For sp =1, sp=4 and sp=6 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite
-# also it can't handle ties for "kendall" and code crashes
-# you can see those msg by deleting results="hide" and warning=F option in r chunk
-
-source("good_loclist.R")
-source("skewness_multivar_dataset.R")
-set.seed(seed)
-
-numsurrog<-10000
-
-splist<-c(1:20)
-skw_aphid_count_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
-colnames(skw_aphid_count_spearman)<-c("realskw","p_left","p_right")
-rownames(skw_aphid_count_spearman)<-paste("sp",splist,sep="")
-pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_spearman.pdf",height=25,width=25)
-op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
-cat("Now for spearman preserving surrogs : aphid count","\n")
-for(i in splist){
-  cat("sp=",i,"\n")
-  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
-  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
-                        numsurrog=numsurrog,
-                        ploton=T,
-                        corpres="spearman") 
-  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
-  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
-  skw_aphid_count_spearman$realskw[i]<-ans$realskw
-  skw_aphid_count_spearman$p_left[i]<-p_left
-  skw_aphid_count_spearman$p_right[i]<-p_right
-  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
-}
-par(op)
-dev.off()
-saveRDS(skw_aphid_count_spearman,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_spearman.RDS")
-
-#skw_aphid_count_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
-#colnames(skw_aphid_count_kendall)<-c("realskw","p_left","p_right")
-#rownames(skw_aphid_count_kendall)<-paste("sp",splist,sep="")
-#pdf("./Results/skewness_results/skewness_aphid_count/aphid_count_allsp_skewness_kendall.pdf",height=25,width=25)
-#op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
-#cat("Now for kendall preserving surrogs : aphid count","\n")
-#for(i in splist){
-#  cat("sp=",i,"\n")
-#  loclist<-good_loclist(d_allsp=data_aphid_count,sp=i,data_pt_thrs=30)
-#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_count),loclist=loclist,
-#                        numsurrog=numsurrog,
-#                        ploton=T,
-#                        corpres="kendall") 
-#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
-# p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
-#  skw_aphid_count_kendall$realskw[i]<-ans$realskw
-#  skw_aphid_count_kendall$p_left[i]<-p_left
-#  skw_aphid_count_kendall$p_right[i]<-p_right
-#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
-#}
-#par(op)
-#dev.off()
-#saveRDS(skw_aphid_count_kendall,"./Results/skewness_results/skewness_aphid_count/skw_aphid_count_kendall.RDS")
-```
-
-```{r skewness_aphidff_allsp_plot, echo=F, results="hide", cache=T, cache.extra=list(seed,data_aphid_ff,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
-
-# For sp =12 and sp=16 : kendall preserving surrogates gives warning : Error in ncsurrog: ncov is not positive semidefinite,
-# also it can't handle ties for "kendall" and code crashes
-# you can see those msg by deleting results="hide" and warning=F option in r chunk
-
-source("good_loclist.R")
-source("skewness_multivar_dataset.R")
-set.seed(seed)
-
-numsurrog<-10000
-
-splist<-c(1:20)
-skw_aphid_ff_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
-colnames(skw_aphid_ff_spearman)<-c("realskw","p_left","p_right")
-rownames(skw_aphid_ff_spearman)<-paste("sp",splist,sep="")
-pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_spearman.pdf",height=25,width=25)
-op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
-cat("Now for spearman preserving surrogs : aphid ff","\n")
-for(i in splist){
-  cat("sp=",i,"\n")
-  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
-  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
-                        numsurrog=numsurrog,
-                        ploton=T,
-                        corpres="spearman") 
-  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
-  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
-  skw_aphid_ff_spearman$realskw[i]<-ans$realskw
-  skw_aphid_ff_spearman$p_left[i]<-p_left
-  skw_aphid_ff_spearman$p_right[i]<-p_right
-  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
-}
-par(op)
-dev.off()
-saveRDS(skw_aphid_ff_spearman,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_spearman.RDS")
-
-#skw_aphid_ff_kendall<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
-#colnames(skw_aphid_ff_kendall)<-c("realskw","p_left","p_right")
-#rownames(skw_aphid_ff_kendall)<-paste("sp",splist,sep="")
-#pdf("./Results/skewness_results/skewness_aphid_ff/aphid_ff_allsp_skewness_kendall.pdf",height=25,width=25)
-#op<-par(mfrow=c(4,5),mar=c(5,5,1,1),mgp=c(3,1,0))
-#cat("Now for kendall preserving surrogs : aphid ff","\n")
-#for(i in splist){
-#  cat("sp=",i,"\n")
-#  loclist<-good_loclist(d_allsp=data_aphid_ff,sp=i,data_pt_thrs=30)
-#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_aphid_ff),loclist=loclist,
-#                        numsurrog=numsurrog,
-#                        ploton=T,
-#                        corpres="kendall") 
-#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
-#  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
-#  skw_aphid_ff_kendall$realskw[i]<-ans$realskw
-#  skw_aphid_ff_kendall$p_left[i]<-p_left
-#  skw_aphid_ff_kendall$p_right[i]<-p_right
-#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
-#}
-#par(op)
-#dev.off()
-#saveRDS(skw_aphid_ff_kendall,"./Results/skewness_results/skewness_aphid_ff/skw_aphid_ff_kendall.RDS")
-```
-
-```{r skewness_plankton_allsp_plot, echo=F, results="hide", cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
-
-# There is no error msg or warning coming from ncsurrog : ncov semidefinite matrix for any species with spearman or kendall preserving surrogates.
-# But there may be error handling ties with "kendall"
-# So,for safety, now we will only analyze with "spearman"
-
-source("good_loclist.R")
-source("skewness_multivar_dataset.R")
-set.seed(seed)
-
-splist<-c(1:22)
-skw_plankton_spearman<-as.data.frame(matrix(NA,nrow=length(splist),ncol=3))
-colnames(skw_plankton_spearman)<-c("realskw","p_left","p_right")
-rownames(skw_plankton_spearman)<-paste("sp",splist,sep="")
-#skw_plankton_kendall<-skw_plankton_spearman
-pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_spearman.pdf",height=30,width=30)
-op<-par(mfrow=c(4,6),mar=c(5,5,1,1),mgp=c(3,1,0))
-#cat("Now for spearman preserving surrogs : pns","\n")
-for(i in splist){
-  #cat("sp=",i,"\n")
-  loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
-  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
-                        numsurrog=numsurrog,
-                        ploton=T,
-                        corpres="spearman") 
-  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
-  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
-  skw_plankton_spearman$realskw[i]<-ans$realskw
-  skw_plankton_spearman$p_left[i]<-p_left
-  skw_plankton_spearman$p_right[i]<-p_right
-  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
-}
-par(op)
-dev.off()
-saveRDS(skw_plankton_spearman,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_spearman.RDS")
-
-#pdf("./Results/skewness_results/skewness_plankton_north_sea/plankton_allsp_skewness_kendall.pdf",height=30,width=30)
-#op<-par(mfrow=c(4,6),mar=c(5,5,1,1),mgp=c(3,1,0))
-#cat("Now for kendall preserving surrogs : pns","\n")
-#for(i in splist){
-  #cat("sp=",i,"\n")
- # loclist<-good_loclist(d_allsp=data_plankton_north_sea,sp=i,data_pt_thrs=45)
-#  ans<-skewness_testing(ts_matrix=sp_data(sp=i,d_allsp=data_plankton_north_sea),loclist=loclist,
-#                        numsurrog=numsurrog,
-#                        ploton=T,
-#                        corpres="kendall") 
-#  p_left<-sum(ans$surrogskw < ans$realskw)/numsurrog
-#  p_right<-sum(ans$surrogskw > ans$realskw)/numsurrog
-#  skw_plankton_kendall$realskw[i]<-ans$realskw
-#  skw_plankton_kendall$p_left[i]<-p_left
-#  skw_plankton_kendall$p_right[i]<-p_right
-#  mtext(paste0("sp = ",i,", p_left = ",p_left,", p_right = ",p_right),side=3,line=-0.5,col="navyblue",cex = 1)
-#}
-#par(op)
-#dev.off()
-#saveRDS(skw_plankton_kendall,"./Results/skewness_results/skewness_plankton_north_sea/skw_plankton_kendall.RDS")
-```
-
 <!--Following chunk examine the consequence of non-Normal copula atructure on skewness using green spruce aphid abundance data-->

### 2019-04-04 SupportingInformation.Rmd
removing Taylors law calculation with real data : aphid count

@@
 
-<!--Following 2 chunks examine the consequence of non-Normal copula atructure on Taylors Law 
-using real data of aphid count with all 20 sp.-->
-```{r spatialTL_on_realdata_aphid_count, results="hide", echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("TaylorsLawTools.R"),mtime("mncurvat.R"),mtime("ncsurrog.R"))}
-
-set.seed(seed)
-source("ncsurrog.R")
-source("TaylorsLawTools.R")
-source("good_loclist.R")
-
-TL_on_realdata<-function(data,numsurrog=1000,sp,data_pt_thrs,ploton,tag,labelon=T){
-
-  loclist<-good_loclist(d_allsp=data,sp=sp,data_pt_thrs=data_pt_thrs)
-  data<-data[[sp]]
-  df<-c()
-  for(i in loclist){
-    tem<-data[[i]]$Dat
-    df<-cbind(df,tem)
-  }
-  colnames(df)<-paste("loc",loclist,sep="")
-  df<-na.omit(df)
-
-  mat_real<-as.matrix(df)
-  check_rep<-function(y){all( abs(y - mean(y)) < 1e-8 )} #This fn checks for repeatation of same values in a vector
-  call_check_rep<-apply(mat_real,1,check_rep) # checks the same for mat_real matrix along all row
-
-  if(all(call_check_rep==F)==T){
-
-    realTL_stats<-tl_stats(m = mat_real)
-    real_spatialTLslope<-unname(realTL_stats$s.stats[5])
-    real_sTL_quadcoeff<-unname(realTL_stats$s.stats[6])
-    real_sTL_mncurv<-unname(realTL_stats$s.stats[7])
-
-    #generate the normal surrogates with spearman correlation preserving
-    array_surrog<-ncsurrog(m=mat_real,corpres="spearman",numsurrog=numsurrog)
-
-    bad_id<-c()
-    for(i in 1:numsurrog){
-      m<-array_surrog[,,i]
-      mean_m<-apply(X=m, MARGIN = 1, FUN = mean)
-      var_m<-apply(X=m, MARGIN = 1, FUN = var)
-      if(any(mean_m==0) | any(var_m==0)){
-        bad_id<-c(bad_id,i)
-      }
-    }
-
-    m<-array_surrog
-    if(length(bad_id)!=0){
-      m<-array_surrog[,,-bad_id]
-    }
-    ts<-apply(m,3,tl_stats)
-
-    good_surrog_slope<-c()
-    good_surrog_quadcoeff<-c()
-    good_surrog_mncurv<-c()
-    for(i in c(1:length(ts))){
-      good_surrog_slope<-c(good_surrog_slope,unname(ts[[i]]$s.stats[5]))
-      good_surrog_quadcoeff<-c(good_surrog_quadcoeff,unname(ts[[i]]$s.stats[6]))
-      good_surrog_mncurv<-c(good_surrog_mncurv,unname(ts[[i]]$s.stats[7]))
-    }
-
-    numsurrog_good<- numsurrog - length(bad_id)
-
-    p_left<-sum(good_surrog_slope < real_spatialTLslope)/numsurrog_good
-    p_right<-sum(good_surrog_slope > real_spatialTLslope)/numsurrog_good
-
-    p_left_quadcoeff<-sum(good_surrog_quadcoeff < real_sTL_quadcoeff)/numsurrog_good
-    p_right_quadcoeff<-sum(good_surrog_quadcoeff > real_sTL_quadcoeff)/numsurrog_good
-
-    p_left_mncurv<-sum(good_surrog_mncurv < real_sTL_mncurv)/numsurrog_good
-    p_right_mncurv<-sum(good_surrog_mncurv > real_sTL_mncurv)/numsurrog_good
-
-    if(ploton==T & tag=="slope"){
-      hist(good_surrog_slope,breaks=100,main="",xlab="surrogate_TL_slope",cex.axis=1.5,cex.lab=1.5)
-      points(x=real_spatialTLslope,y=0,col="red",pch=15)
-      if(labelon==T){
-        mtext(paste0("sp= ",sp,", p_left= ",round(p_left,3),", p_right=",round(p_right,3),sep=""),
-              side = 3, line=-2, adj=0, col="blue")
-      }
-    }
-
-    if(ploton==T & tag=="quadcoeff"){
-      hist(good_surrog_quadcoeff,breaks=100,main="",xlab="surrogate_TL_quadcoeff",cex.axis=1.5,cex.lab=1.5)
-      points(x=real_sTL_quadcoeff,y=0,col="red",pch=15)
-      if(labelon==T){
-        mtext(paste0("sp= ",sp,", p_left= ",round(p_left_quadcoeff,3),", p_right=",round(p_right_quadcoeff,3),sep=""),
-              side = 3, line=-2, adj=0, col="blue")
-      }
-    }
-
-    if(ploton==T & tag=="mncurv"){
-      hist(good_surrog_mncurv,breaks=100,main="",xlab="surrogate_TL_meanCurvature",cex.axis=1.5,cex.lab=1.5)
-      points(x=real_sTL_mncurv,y=0,col="red",pch=15)
-      if(labelon==T){
-        mtext(paste0("sp= ",sp,", p_left= ",round(p_left_mncurv,3),", p_right=",round(p_right_mncurv,3),sep=""),
-              side = 3, line=-2, adj=0, col="blue")
-      }
-    }
-
-  }else{
-    real_spatialTLslope<-NA
-    p_left<-NA
-    p_right<-NA
-    numsurrog_good<-NA
-    real_sTL_quadcoeff<-NA
-    p_left_quadcoeff<-NA
-    p_right_quadcoeff<-NA
-    real_sTL_mncurv<-NA
-    p_left_mncurv<-NA
-    p_right_mncurv<-NA
-  }
-  return(list(realTL_stats=realTL_stats,
-              real_spatialTLslope=real_spatialTLslope,
-              p_left=p_left,
-              p_right=p_right,
-              numsurrog_good=numsurrog_good,
-              real_sTL_quadcoeff=real_sTL_quadcoeff,
-              p_left_quadcoeff=p_left_quadcoeff,
-              p_right_quadcoeff=p_right_quadcoeff,
-              real_sTL_mncurv=real_sTL_mncurv,
-              p_left_mncurv=p_left_mncurv,
-              p_right_mncurv=p_right_mncurv,
-              surrogTL_stats=ts,
-              good_surrog_mncurv=good_surrog_mncurv))
-}
-
-#call on real data
-# For all sp
-splist<-c(1:20)
-tl_aphid_count<-as.data.frame(matrix(NA,nrow=length(splist),ncol=10))
-colnames(tl_aphid_count)<-c("real_sTL_slope","p_left","p_right","numsurrog_good",
-                            "real_sTL_quadcoeff","p_left_quadcoeff","p_right_quadcoeff",
-                            "real_sTL_mncurv","p_left_mncurv","p_right_mncurv")
-rownames(tl_aphid_count)<-paste("sp",splist,sep="")
-#pdf("aphid_count_allsp_spTLslope.pdf",height=20,width=20)
-#pdf("./Results/taylorslaw_results/aphid_count_allsp_sTL_quadcoeff.pdf",height=20,width=20)
-pdf("./Results/taylorslaw_results/aphid_count_allsp_sTL_mncurv.pdf",height=20,width=20)
-op<-par(mfrow=c(4,5),mar=c(4.2,5,1,1),mgp=c(3,1,0))
-for(i in splist){
-  #cat("sp=",i,"\n")
-  #ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="slope",labelon=T)
-  #ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="quadcoeff",labelon=T)
-  ans<-TL_on_realdata(data=data_aphid_count,numsurrog=10000,sp=i,data_pt_thrs = 30,ploton=T,tag="mncurv",labelon=T)
-  if(i==6){
-    saveRDS(ans,"./Results/taylorslaw_results/tl_aphid_count_sp6.RDS") #saving sp=6 results
-  }
-  tl_aphid_count$real_sTL_slope[i]<-ans$real_spatialTLslope
-  tl_aphid_count$p_left[i]<-ans$p_left
-  tl_aphid_count$p_right[i]<-ans$p_right
-  tl_aphid_count$numsurrog_good[i]<-ans$numsurrog_good
-  tl_aphid_count$real_sTL_quadcoeff[i]<-ans$real_sTL_quadcoeff
-  tl_aphid_count$p_left_quadcoeff[i]<-ans$p_left_quadcoeff
-  tl_aphid_count$p_right_quadcoeff[i]<-ans$p_right_quadcoeff
-  tl_aphid_count$real_sTL_mncurv[i]<-ans$real_sTL_mncurv
-  tl_aphid_count$p_left_mncurv[i]<-ans$p_left_mncurv
-  tl_aphid_count$p_right_mncurv[i]<-ans$p_right_mncurv
-}
-par(op)
-dev.off()
-saveRDS(tl_aphid_count,"./Results/taylorslaw_results/tl_aphid_count.RDS")
-```
-
-```{r spatialTL_on_realdata_aphid_count_sp6, results="hide", echo=F}
-#Now for sp=6
-
-ans<-readRDS("./Results/taylorslaw_results/tl_aphid_count_sp6.RDS")
-
-pdf("./Results/taylorslaw_results/aphid_count_sp6_sTL_mncurv.pdf",height=5,width=5)
-op<-par(mar=c(4,5,1,1),mgp=c(3,1,0))
-hist(ans$good_surrog_mncurv,breaks=100,main="",xlab="Mean curvature",cex.axis=2,cex.lab=2,cex=2,col="grey",border=F)
-points(x=ans$real_sTL_mncurv,y=0,col="black",pch=15,cex=2)
-mtext(paste0("p= ",ans$p_left_mncurv),side = 3, line=-2, adj=0.5, col="black",cex=2)
-par(op)
-dev.off()
-
-pdf("./Results/taylorslaw_results/aphid_count_sp6_logv_vs_logm.pdf",height=5,width=10)
-op<-par(mfrow=c(1,2),mar=c(4,5,1,1),mgp=c(3,1,0))
-plot(ans$realTL_stats$s.points,xlab=expression(log[10]("spatial mean")),
-     ylab=expression(log[10]("spatial variance")),col="grey",pch=20,cex=1.5,
-     cex.axis=2,cex.lab=2)
-mtext(paste0("mean curvature= ",round(unname(ans$realTL_stats$s.stats[7]),3)),side = 3, line=-2, adj=0, col="black",cex=2)
-
-target_mncurv<-0.1
-id<-which(abs(ans$good_surrog_mncurv-target_mncurv)<1e-4)[1]
-ans_s<-ans$surrogTL_stats
-plot(ans_s[[id]]$s.points,xlab=expression(log[10]("spatial mean")),
-     ylab=expression(log[10]("spatial variance")),col="grey",pch=20,cex=1.5,
-     cex.axis=2,cex.lab=2)
-mtext(paste0("mean curvature= ",round(ans$good_surrog_mncurv[id],3)),side = 3, line=-2, adj=0, col="black",cex=2)
-
-par(op)
-dev.off()
-```
-
 <!--Following 2 chunks examine the effect of non-Normal asymmetric copula on extinction risk-->

### 2019-04-04 Paper.Rmd
adding summary table of data info as suggested by PCReid

@@
 from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, 
-Ohio [@Holland1999; @Jarecke2016]. Each included location was measured on at least 50 dates. 
+Ohio [@Holland1999; @Jarecke2016; @smyth2019using]. Each included location was measured on at least 50 dates. 
 Sampling locations spanned a hydrologic gradient from upland to seasonal wetland.
@@
 copula structure can be studied for each pair of locations. Likewise the ecosystem 
-functioning (methane flux) dataset was  multivariate in this sense.
+functioning (methane flux) dataset was  multivariate in this sense (see \ref{tab_data_info}).
+
+<!--summary table with info of all data used in Paper-->
+```{r tab_summary_data_info, echo=F, results="asis",message=F}
+library(kableExtra)
+library(dplyr)
+tab_data_info<-as.data.frame(matrix(NA,nrow=8,ncol=4))
+colnames(tab_data_info)<-c("Data","Category","No. of measurements we used", "References")
+tab_data_info$Data<-c("Soil C and N","Bird masses and BMR", "Mammal masses and BMR", "Cedar Creek data", 
+                      "Green spruce aphid abundance", "Leaf-curling plum aphid first flight","Ceratium furca abundance","Methane flux")
+tab_data_info$Category<-c("Environmental","Species-trait","Species-trait","Community level",
+                          "Population","Phenological","Population","Ecosystem-functioning")
+tab_data_info$`No. of measurements we used`<-c(paste(nrow(d_soilCN)," locations",sep=""),
+                                               paste(nrow(d_birdsBMR)," birds",sep=""),
+                                               paste(nrow(d_mammalsBMR)," mammals",sep=""),
+                                               paste(nrow(d_cc[[5]])," plots",sep=""),
+                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
+                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
+                                               "14 locations with atleast 45 yrs. of timeseries for each 22 taxa",
+                                               "13 locations with atleast 50 dates of data")
+tab_data_info$References<-c("US Rapid Carbon Assessment database (RaCA), Wills et al. (2014)", "McNab 2009","McNab 2008",
+                            "Cedar Creek Ecosystem Science Reserve",
+                            "Rothamsted Insect Survey","Rothamsted Insect Survey","Continuous Plankton Recorder Survey",
+                            "Great Miami Wetland Mitigation Bank, Smyth et al. (2019)")
+
+knitr::kable(tab_data_info, 
+             format="latex", align="l",
+             caption = "Summary table for data we used, bolds are the multivariate data, rest are bivariates. \\label{tab_data_info}", 
+             booktabs = T,linesep = "\\addlinespace") %>%
+             column_spec(1:4,width="2.8 cm")%>%
+             row_spec(5:8, bold=T)
+```
 
@@
 
-<!--bivariate table summary-->
+<!--summary table : results with bivariate data-->
 ```{r tab_bivar_summaryl, echo=F, results='asis',message=F}
@@
 
-<!--multivariate table summary-->
+<!--summary table : results with multi-variate data-->
 ```{r tab_multivar_summaryl, echo=F, results='asis',message=F}

### 2019-04-06 Paper.Rmd
answering Reid's comment and adding notes for DAN

@@
 title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
-author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, P. Chris Reid, Daniel C. Reuman"
+author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, Daniel C. Reuman"
 date: ""
@@
 
-Reid: Marine Institute, Plymouth University, Drake Circus, Plymouth, United Kingdom; Mrine Biological Association of the UK, The Laboratory, Citadel Hill, Plymouth, United Kingdom
+Reid: School of Biological & Marine Sciences, University of Plymouth, Drake Circus, Plymouth PL4 8AA, United Kingdom
 
@@
 
-Very generally, all branches of ecology and related areas study relationships among environmental and biological 
+Very generally, all branches of ecology and related areas <!--DAN : Reid said " Is this text needed. 
+Ecology covers 
+all branches of the biological sciences?
+What do you mean by related areas? 
+You use related fields later.
+Needs clarifying."-->
+study relationships among environmental and biological 
 variables. However, commonly used correlation and regression 
 approaches to studying 
-such relationships are limited - common statistical
+such relationships are limited - such statistical
 approaches provide only a very partial description of a relationship between variables. 
@@
 is only one of numerous metrics for 
-studying relationships between variables, other common metrics also 
+studying relationships between variables, other common metrics <!--DAN : add example--> also 
 provide only partial information.
@@
 
-Well-developed statistical tools do exist, and are 
-applied widely in other fields, that provide a demonstrably complete 
+Other well-developed statistical tools exist, that are 
+applied widely in other fields (e.g. finance, neuroscience, bioinformatics etc.), 
+and provide a demonstrably complete 
 description of the dependence 
@@
 See @nelsen2006_copula and @joe2014_dependence
-for general background. An theorem of A. Sklar 
+for general background. A theorem of A. @sklar1959_theorem
 states that copulas give a unique and complete 
@@
 can be reconstructed from the 
-copula and the marginal distributions [@sklar1959_theorem]. 
+copula and the marginal distributions. 
 Copula theory is also well developed for
-multivariate distributions.
+multivariate distributions [@joe1997multivariate].
 
@@
 research is that prior work has demonstrated important copula 
-stucture in the spatial dependence of environental 
+stucture in the spatial dependence of environmental 
 variables [@Serinaldi2008; @She2018; @Goswami2018;
@@
 variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], finance, and other fields. 
-To our knowledge, these approaches have been used seldom or never in ecology. 
+To our knowledge, these approaches have never or seldom been used for ecological research. 
 We argue that benefits to ecology of using a complete statistical 
@@
 as measured in multiple locations has been 
-examined previously in the meteorological literature (e.g., @Serinaldi2008, @li2013). We therefore 
+examined previously in the meteorological literature [e.g., @Serinaldi2008; @li2013]. We therefore 
 examined environmental variables from the soil instead, using the Rapid Carbon 
-Assessment database (RaCA; @Wills2014).
+Assessment database [RaCA; @Wills2014].
 <!--***DAN: Add Wills and Loecke pending reference if published in time--> 
@@
 35 years 1976 to 2010. These time series were computed from the Rothamsted Insect 
-Survey suction-trap dataset [@Harrington2014; @Bell2015]. The first of our two 
+Survey (RIS) suction-trap dataset [@Harrington2014; @Bell2015]. The first of our two 
 population-level datasets was also derived from the RIS suction-trap data, and comprised 
@@
 that of @Sheppard2016,
-but we exlcuded one of the 11 sites they used so as to include only sites with at least 30 
+but we excluded one of the 11 sites they used in order to only use sites with at least 30 
 years data.  Our second population dataset comprised average annual abundance estimates 
@@
 56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. These data 
-were obtained from the Continuous Plankton Recorder (CPR) dataset maintained by the Sir 
-Alistair Hardy Foundation for Ocean Science (SAHFOS), now part of the Marine Biological 
-Association of the UK. Data pre-processing was similar but not identical to that 
+were obtained from the Continuous Plankton Recorder (CPR) dataset now operated and maintained by 
+the Marine Biological Association. Data pre-processing was similar but not identical to that 
 of @Sheppard2017; see Appendix \ref{SM-Data_details} for details.
-Community-level data were plant aboveground biomass and Shannon's diversity index for `r nrow(d_cc[[5]])` 
-plots, each 9m by 9m, maintained at Cedar Creek Ecosystem Science Reserve and sampled in 
-the year 2000 [@Tilman2001; @Tilman2006]. 
+Community-level data obtained from the Cedar Creek Ecosystem Science Reserve were <!--DAN : check this rephrasing-->
+plant aboveground biomass and plant percent cover data (we used this percent cover data to calculate 
+Shannon's diversity index) for `r nrow(d_cc[[5]])` plots, each 9m by 9m, as sampled in 
+the year 2000 [@Tilman2001; @Tilman2006] <!--actually they collected data in each year of 1996,1997,1998,1999,2000,2007-->. 
 Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
@@
 tab_data_info$References<-c("US Rapid Carbon Assessment database (RaCA), Wills et al. (2014)", "McNab 2009","McNab 2008",
-                            "Cedar Creek Ecosystem Science Reserve",
+                            "Cedar Creek Ecosystem Science Reserve [biodiversity experiment e120]",
                             "Rothamsted Insect Survey","Rothamsted Insect Survey","Continuous Plankton Recorder Survey",
@@
 We addressed Q1 principally via a model selection procedure in which the normal copula
-the alternative copulas listed above were fit to 
-our ecological datasets, and fits of the various copula models 
+the alternative copulas listed above were fitted to 
+our ecological datasets. The resulting fits of the various copula models 
 were compared via the Akaike and Bayesian 
@@
 $\log_{10}$(body mass) for mammals, and (D) above-ground plant biomass vs.
-Shannon's index (H) from Cedar Creek.
+Shannon's index \textit{(H)} from Cedar Creek.
 Bottom panels show the corresponding copula plots.
@@
 as well as for the 21 plankton taxa for which we had abundance data other than the
-\textit{Ceratium furca} (results not shown). The conclusions were generally supported that 
+\textit{Ceratium furca} (results not shown). The conclusions were generally supportive that 
 non-N copula structure and tail dependence, as well as asymmetry of tail dependence, are 
@@
                         "Leaf-curling plum aphid first flight",
-                        "Ceratium furca abundance","Methane flux")
+                        "\\textit{Ceratium furca} abundance","Methane flux")
              )%>%column_spec(2:5,width="2.5cm")%>%

### 2019-04-06 Paper.Rmd
Discussion edited for vegetation related drought

@@
 neither of these is an N copula. @liu2016probabilistic used a copula 
-based approach in assessing the probability of vegetation related drought across China. 
-<!--***DAN: Shya, what do you mean by "vegetation related drought"? Is this a kind of drought
-that is somehow caused by the vegetation? That's what it sounds like, and it sounds like what
-they studied was the probability that such a drought occurred. If that's not what you mean, please
-rephrase.
-***Shya: I used the exact term "vegetation-related drought" they used in the paper @liu2016probabilistic, I think they 
-wanted to mean that you can monitor the drought risk by tracking change in NDVI spatial greenness with precipitation or temperature (climatic variable), which can be considered as a proxy indicator of vegetation response to climate change.
-They also showed vegatation-drought stress analysis, which I also noticed in some other papers too.
-There are several studies using drought and NDVI data (mainly using correlation, regression) but they reported it as the first study using joint probability for drought-risk assessment.
-I have highlighted the info which may be important to us and sent you in mail.-->
+based approach in assessing the drought risk with joint probability distribution 
+between vegetation and climatic variables across China. 
 In a majority of their cases, they found 

### 2019-04-08 Paper.Rmd
editting as per PCReid's comment and adding notes for DAN

@@
 Fig. \ref{fig_cop_pedag1}E-F, which show the copula structure of 
-A-B, respectively. Ranking serves to uniformize the 
-marginal distributions of the underlying random variables, isolating the 
-dependence structure only.
+A-B, respectively. Ranking makes marginals of 
+the underlying random variables uniformly distributed, isolating the 
+dependence structure only<!--DAN : check this-->.
 Mathematically, if $(X,Y)$ is a continuous bivariate random
@@
 examined environmental variables from the soil instead, using the Rapid Carbon 
-Assessment database [RaCA; @Wills2014].
+Assessment database [[RaCA](https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/survey/?cid=nrcs142p2_054164); @Wills2014].
 <!--***DAN: Add Wills and Loecke pending reference if published in time--> 
@@
 locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
-Measurments were from the central pedon of a soil core, to 1 m depth.
+Measurments were from the central pedon of a soil core, to 1 m depth.<!--DAN: we should ask Terry to explain pedon/ R horizon-->
 <!--***DAN: Terry to check the wording and accuracy here-->
@@
 Model selection methods will give the *relative* support of several models, but will not,
-on their own, indicate whether any of the models selected is an objectively good fit: all 
+on their own, indicate whether any of the models selected are an objectively good fit: all 
 models considered might poorly describe the data. To test this, and as the final step of

### 2019-04-09 SupportingInformation.Rmd
answering Reid's comment and adding notes for DAN
code edited for cedar creek : include 168 plots as per Tilman's paper, species name screened

@@
 title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
-author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, P. Chris Reid, Daniel C. Reuman"
+author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, Daniel C. Reuman"
 geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"
@@
 #       dp : a dataframe : smaller subset of pce_120 data to calculate avg. shannon's H for those plots
+#       plots : a vector of given 168 plot ids as used in reference papers by Tilman
 #       year : the year for which you want to calculate biomass and shannon's index data
@@
 
-cop_HB<-function(db,dp,year,rawplot,ploton){
+cop_HB<-function(db,dp,plots,year,rawplot,ploton){
   # db : biomass dataset for a specific year
@@
 
-  plots<-sort(intersect(db$Plot,dp$Plot)) # common plots for both datasets for a specific year
   db<-subset(db,Plot%in%plots)
@@
    x<-subset(dp,Plot==plots[i])
+   x$Species<-trimws(x$Species) # to remove duplicated names of sp. (cutting white spaces) : e.g. 1999 : plot no 339
    uniq_sp<-sort(as.character(unique(x$Species)))  # These are the species in a particular plot
@@
    tab_each_sp<-data.frame(sp=uniq_sp,avg_cov=0,rel_avg_cov=0)
-
+   #cat("=====================================plot : ",plots[i],"===========================\n")
    for(sp in 1:length(uniq_sp)){
     my_sp<-uniq_sp[sp]
+    #cat("---------- sp = ",my_sp,"--------------------\n")
     tab_temp<-data.frame(Q=Q,abs_cov=0)
@@
     li<-length(ind)
+    #print(li)
     tab_temp$abs_cov[1:li]<-x$Abscover[ind]
@@
   tab_dbp$avg.H[i]<- -sum((tab_each_sp$rel_avg_cov)*log(tab_each_sp$rel_avg_cov))
-
-  }
-
+ }
   tab_dbp<-na.omit(tab_dbp)
@@
 b<-as.character(a$Var1)
-bad_sp<-c(b[51:53],b[156:158])
-
+bad_sp<-c(b[51:53],b[156:158],"Disturbance") 
+# bad sp : "Bare ground","Bare ground","BARE GROUND","Miscellaneous litter","Miscellaneous litter","MISCELLANEOUS LITTER"            
 ind<-which(r_pce$Species%in%bad_sp)
 r_pce_clean<-r_pce[-ind,]
+
+#----------------------------------------------------------------------------------------------
+# This sections selects 168 plots considered for both biomass and percent cover datasets in Tilman's papers 2001, 2006
+i<-5#for Yr=2000
+db<-subset(r_ple,Year==yr[i])
+dp<-subset(r_pce_clean,Year==yr[i])
+#------------------------------------------------------
+id_plot_nosp<-which(db$NumSp==0 & db$Biomass==0) # These plots have no sp. seeded initially and have no biomass data
+plots_no_biomass<-sort(unique(db$Plot[id_plot_nosp]))
+plots_no_biomass
+
+plots<-sort(intersect(db$Plot,dp$Plot)) # common plots for both datasets for a specific year
+length(plots)
+
+excluded_plots<-intersect(plots,plots_no_biomass)
+
+plots_168<-setdiff(plots,excluded_plots)
+
 #--------------------------------------------------------------------------------------------
@@
 
-  sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
+  sr<-cop_HB(db=db,dp=dp,year=yr[i],plots=plots_168,rawplot=T,ploton = T)
 }
@@
 dp<-subset(r_pce_clean,Year==yr[i])
-sr<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=T,ploton = T)
+sr<-cop_HB(db=db,dp=dp,year=yr[i],plots=plots_168,rawplot=T,ploton = T)
 saveRDS(sr,"./Results/Ceder_creek_results/raw_cop_cc_2000.RDS")
@@
 
-  s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
+  s<-cop_HB(db=db,dp=dp,year=yr[i],plots=plots_168,rawplot=F,ploton = T)
   cop_HB_all_yr[[i]]<-s$cop_tab_dbp[,2:3]
@@
 dp<-subset(r_pce_clean,Year==yr[i])
-s<-cop_HB(db=db,dp=dp,year=yr[i],rawplot=F,ploton = T)
+s<-cop_HB(db=db,dp=dp,year=yr[i],plots=plots_168,rawplot=F,ploton = T)
 cc_cop_2000<-s$cop_tab_dbp[,2:3]

### 2019-04-09 Paper.Rmd
Cedar creek results modified and explained accordingly

@@
 But our nonparametric statistics confirmed the same conclusions 
-(Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}).
-Detailed results for each dataset are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} 
-and Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.
+(Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}) for first three bivariate datasets.
+The fourth bivariate datasets we analyzed was above-ground biomass and percent cover data (i.e. 
+the data used to compute Shannon's diversity index) for the year 2000, but we did the same analysis 
+with Cedar creek data sampled over 168 plots from other available years (1996-1999, 2007) too. For 1996 and 
+2007 copula structures showed no dependence between two variables concerned, for other years (1997-1999), 
+we found similar results as of year 2000 (i.e. best fit non-Normal copula and overall lower tail-dependence).
+Due to smaller sample size relative to other three bivariate datasets, we chose a wider range with 
+non-parametric analysis (see Appendix \ref{SI-Q1bivarnonparam}).
+Detailed results for each dataset are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and Figs.
+\ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.
 

### 2019-04-09 SupportingInformation.Rmd
Cedar creek results modified and explained accordingly

@@
 #----------------------------------------------------------------------------------------------
+#=======================================================
 # This sections selects 168 plots considered for both biomass and percent cover datasets in Tilman's papers 2001, 2006
-i<-5#for Yr=2000
-db<-subset(r_ple,Year==yr[i])
-dp<-subset(r_pce_clean,Year==yr[i])
+plot_info<-read.delim("./Data/Ceder_creek_e120/edited_plots.txt")
+
+id_excluded_plots<-which(plot_info$Species=="Bargr") # there were 4 plots with no sp. seeded
+
+plots_168<-plot_info$Plot[-id_excluded_plots]
+
 #------------------------------------------------------
-id_plot_nosp<-which(db$NumSp==0 & db$Biomass==0) # These plots have no sp. seeded initially and have no biomass data
-plots_no_biomass<-sort(unique(db$Plot[id_plot_nosp]))
-plots_no_biomass
-
-plots<-sort(intersect(db$Plot,dp$Plot)) # common plots for both datasets for a specific year
-length(plots)
-
-excluded_plots<-intersect(plots,plots_no_biomass)
-
-plots_168<-setdiff(plots,excluded_plots)
-
+# You can check that here too
+s<-subset(r_ple,NumSp==0)
+sb<-sort(unique(s$Plot))
+
+s<-subset(r_pce_clean,NumSp==0)
+spc<-sort(unique(s$Plot))
+
+sb==spc # This should be all TRUE for 4 not seeded plots with id : 126, 166, 244, 248
+#=======================================================
 #--------------------------------------------------------------------------------------------
@@
 \end{figure}
+
+<!--This chunk is for model selection results of soilCN dataset-->
+```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("OurBiCopSelect.R")
+set.seed(seed)
+families_Loecke<-c(1,3:10,13:14,16:20) 
+BivMS_res_Loecke<-OurBiCopSelect(u1=v_CN[,1],u2=v_CN[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
+saveRDS(BivMS_res_Loecke,file="./Results/fitting_results/BivMS_soilCN.RDS")
+```
+
+<!--Display model selection results for soilCN data in a table-->
+```{r tab_soilCNfit, echo=F, results="markup"}
+#knitr::opts_knit$set(kable.force.latex = TRUE)
+h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+colnames(h)<-c("copname","AICw","BICw","LT","UT")
+knitr::kable(h, #longtable = T, 
+             format = "latex",linesep = "",
+      caption.short = "Model fitting results for soil C and N dataset",
+      caption = "Results of fitting our 16 copula families to the soil C and N dataset.
+Copula name abbreviations are given in the main text.
+AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.\\label{tab_soilCNfit}",
+      booktabs = TRUE)
+```
+
+<!--This chunk is for npa testing results of soilCN dataset-->
+```{r stat_soilCN_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_CN,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+source("bivfunctionplot.R")
+set.seed(seed) 
+stat_soilCN_ga<-bivfunctionplot(v=v_CN,resloc="./Results/stat_results/stat_soilCN/",nametag="soilCN",numbin=10)
+```
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
+\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for soil C and N data.
+As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
+Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
+For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates.
+When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level).
+When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.
+\label{fig_soilCN_nonparam}}
+\end{center}
+\end{figure}
+
+<!--Function to generate a table for npa results-summary with bivariate data, this one about asymmetry of tail dependence-->
+```{r make_table_stat_fn, echo=F}
+
+fractxt<-function(fracval,numsurrog){
+  ptxt<-''
+  if (fracval>.5*numsurrog){                        
+    ptxt<-paste(">",fracval,sep='')
+  } 
+  if (fracval<=.5*numsurrog){
+    ptxt<-paste("<",numsurrog-fracval,sep='')
+  }
+  return(ptxt)
+}
+
+make_table_stat_fn<-function(data){
+
+ #atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
+ #                               "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
+ #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
+ atdres<-data.frame(Statistic=c("cor$_{FS}$-cor$_{LS}$",
+                               "P$_{FS}$-P$_{LS}$",
+                                "D$_{LS}^2$-D$_{FS}^2$"),
+                   Kendall=rep("",3),
+                   Spearman=rep("",3),stringsAsFactors = F)
+
+ atdres[1,2]<-fractxt(data$corlmcoru_frac_K,numsurrog = data$numsurrog_success_K)
+ atdres[1,3]<-fractxt(data$corlmcoru_frac_S,numsurrog = data$numsurrog_success_S)
+ atdres[2,2]<-fractxt(data$PlmPu_frac_K,numsurrog = data$numsurrog_success_K)
+ atdres[2,3]<-fractxt(data$PlmPu_frac_S,numsurrog = data$numsurrog_success_S)
+ atdres[3,2]<-fractxt(data$D2umD2l_frac_K,numsurrog = data$numsurrog_success_K)
+ atdres[3,3]<-fractxt(data$D2umD2l_frac_S,numsurrog = data$numsurrog_success_S)
+
+ return(atdres)
+
+}
+```
+
+<!--Displaying a table with npa summary results for soilCN data-->
+```{r tab_soilCN_asym,echo=F,results="markup"}
+#knitr::opts_knit$set(kable.force.latex = TRUE)
+tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
+#tab_soilCN_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
+knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
+             format = "latex",linesep = "",
+      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
+      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
+The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons.
+A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
+```
+
+<!--Simulations for model selection results with Birds' BMR data-->
+```{r bivmodselect_birds_BMR, echo=F, cache=T, cache.extra=list(seed,cop_b,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("OurBiCopSelect.R")
+set.seed(seed)
+BivMS_birds<-OurBiCopSelect(u1=cop_b[,1],u2=cop_b[,2],families=c(1,3:10,13:14,16:20),
+                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
+                              gofnormal=F,status=F)
+saveRDS(BivMS_birds,file="./Results/BMR_results/birds/BivMS_birds.RDS")
+```
+
+<!--Display model selection results for BMR:birds data in a table-->
+```{r tab_bmr_birds_fit, echo=F, results='markup'}
+h<-BivMS_birds$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+colnames(h)<-c("copname","AICw","BICw","LT","UT")
+knitr::kable(h,
+             format='latex',linesep = "",
+             caption.short = "Model fitting results for the bird body mass and BMR dataset",
+             caption = "Results of fitting our 16 copula families to the bird body mass and BMR dataset.
+Analogous to Fig.
+\\ref{tab_soilCNfit}, see caption of that figure for details.
+\\label{tab_bmr_birds_fit}",
+             booktabs=T)
+```
+
+<!--This chunk is for npa testing results for Birds' BMR dataset-->
+```{r stat_bmr_birds_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_b,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+source("bivfunctionplot.R")
+set.seed(seed)
+stat_birds_ga<-bivfunctionplot(v=cop_b,resloc="./Results/BMR_results/birds/",nametag="birds_numbin_10",numbin = 10)
+```
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
+\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for bird body mass and BMR data.
+Same format as Fig.
+\ref{fig_soilCN_nonparam}, see caption of that figure.
+\label{fig_bmr_birds_nonparam}}
+\end{center}
+\end{figure}
+
+<!--Displaying a table with npa summary results for Birds' BMR data-->
+```{r tab_bmr_birds_asym, echo=F, results="markup"}
+tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
+#tab_bmr_birds_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
+knitr::kable(tab_bmr_birds_asym,
+             format="latex",linesep = "",escape=F,
+             caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
+             caption = "Bird body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
+Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_bmr_birds_asym}",
+             booktabs=T)
+```
+
+<!--Simulations for model selection results with Mammals' BMR data-->
+```{r bivmodselect_mammals_BMR, echo=F, cache=T, cache.extra=list(seed,cop_m,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("OurBiCopSelect.R")
+set.seed(seed)
+BivMS_mammals<-OurBiCopSelect(u1=cop_m[,1],u2=cop_m[,2],families=c(1,3:10,13:14,16:20),
+                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
+                              gofnormal=F,status=F)
+saveRDS(BivMS_mammals,file="./Results/BMR_results/mammals/BivMS_mammals.RDS")
+```
+
+<!--Display model selection results for BMR:mammals data in a table-->
+```{r tab_bmr_mammals_fit, echo=F, results='markup'}
+h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+colnames(h)<-c("copname","AICw","BICw","LT","UT")
+knitr::kable(h,
+             format='latex',linesep = "",
+             caption.short = "Model fitting results for the mammal body mass and BMR dataset",
+             caption = "Results of fitting our 16 copula families to the mammal body mass and BMR dataset.
+Analogous to Fig.
+\\ref{tab_soilCNfit}, see caption of that figure for details.
+\\label{tab_bmr_mammals_fit}",
+             booktabs=T)
+```
+
+<!--This chunk is for npa testing results for Mammals' BMR dataset-->
+```{r stat_bmr_mammals_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_m,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+source("bivfunctionplot.R")
+set.seed(seed)
+stat_mammals_ga<-bivfunctionplot(v=cop_m,resloc="./Results/BMR_results/mammals/",nametag="mammals_numbin_10",numbin = 10)
+```
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
+\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for mammal body mass and BMR data.
+Same format as Fig.
+\ref{fig_soilCN_nonparam}, see caption of that figure.
+\label{fig_bmr_mammals_nonparam}}
+\end{center}
+\end{figure}
+
+<!--Displaying a table with npa summary results for Mammals' BMR data-->
+```{r tab_bmr_mammals_asym, echo=F, results="markup"}
+tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
+#tab_bmr_mammals_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
+knitr::kable(tab_bmr_mammals_asym,
+             format="latex",linesep = "",escape=F,
+             caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
+             caption = "Mammal body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
+Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.
+\\label{tab_bmr_mammals_asym}",
+             booktabs=T)
+```
+
+<!--Simulations for model selection results with Cedar creek data-->
+```{r bivmodselect_cc, echo=F, cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("OurBiCopSelect.R")
+set.seed(seed)
+
+BivMS_res_cc<-vector("list",length(cop_HB_all_yr))
+names(BivMS_res_cc)<-names(cop_HB_all_yr)
+
+for(i in 1:length(BivMS_res_cc)){
+  BivMS_res_cc[[i]]<-OurBiCopSelect(u1=cop_HB_all_yr[[i]]$avg.H,u2=cop_HB_all_yr[[i]]$avg.Biomass,
+                                    families=c(1,3:10,13:14,16:20),
+                                    level=0.05,AICBIC="AIC",
+                                    numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
+}
+saveRDS(BivMS_res_cc,file="./Results/Ceder_creek_results/BivMS_res_cc.RDS")
+BivMS_res_cc<-readRDS("./Results/Ceder_creek_results/BivMS_res_cc.RDS")
+```
+
+<!--Display model selection results for Cedar-creek data for year 2000 in a table-->
+```{r tab_cc2000fit, echo=F, results='markup'}
+h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
+colnames(h)<-c("copname","AICw","BICw","LT","UT")
+knitr::kable(h,
+             format='latex',linesep = "",
+             caption.short = "Model fitting results for the Cedar Creek dataset",
+             caption = "Results of fitting our 16 copula families to the Cedar Creek dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_cc2000fit}",
+             booktabs=T)
+```
+
+<!--This chunk is for npa testing results for Cedar creek dataset-->
+```{r stat_cc_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+source("bivfunctionplot.R")
+set.seed(seed)
+
+stat_cc_ga<-vector("list",length(cop_HB_all_yr))
+names(stat_cc_ga)<-names(cop_HB_all_yr)
+
+for(i in 1:length(stat_cc_ga)){
+  temp<-bivfunctionplot(v=cop_HB_all_yr[[i]],
+                             resloc="./Results/Ceder_creek_results/",
+                             nametag=paste0(names(cop_HB_all_yr)[i],"cc_numbin_4",sep=""),
+                             numbin = 4)
+  stat_cc_ga[[i]]<-temp
+}
+
+saveRDS(stat_cc_ga,"./Results/Ceder_creek_results/stat_cc_ga.RDS")
+```
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_4_bivfunctionplot.pdf}
+\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data.
+As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$.
+If the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
+For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates.
+When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level).
+When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.\label{fig_cc2000_nonparam}}
+\end{center}
+\end{figure}
+
+<!--Displaying a table with npa summary results for Cedar creek data-->
+```{r tab_cc_asym_2000, echo=F, results="markup"}
+tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
+#tab_cc_asym_2000[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
+knitr::kable(tab_cc_asym_2000,
+             format="latex",linesep = "",escape=F,
+             caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
+             caption = "Cedar Creek dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
+The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.25}-\\cor_{0.75,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.25}-\\Ps_{0.75,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.75,1}-\\Dsq_{0,0.25}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons.
+A table entry $>X$ indicates the value of the given statistic on the data was higher than its value on $X$ of the surrogates, so entries of the form $>X$ for $X$ equal to $975$ or above would indicate that lower-tail dependence was significantly stronger than upper-tail dependence.\\label{tab_cc_asym_2000}",
+             booktabs=T)
+```
+
+<!--Function to generate a table containing summary results for all bivariate datasets-->
+```{r bivar_tab_summary_fn,echo=F}
+bivar_tab_summary_fn<-function(r,rs){
+
+  m<-matrix(NA,1,10)
+  colnames(m)<-c("p.Independence.test","minimumAIC.bestfit.copula","Bestfit.copula.AICw",
+                 "Normal.copula.AICw","p.CvM","p.KS",
+                 "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
+                 "Rank.(Pl-Pu)")
+  m<-as.data.frame(m)
+
+  m$p.Independence.test<-round(r$IndepTestRes,3)
+  m$minimumAIC.bestfit.copula<-r$InfCritRes$copname[which.min(r$InfCritRes$AIC)]
+  m$Bestfit.copula.AICw<-round(max(r$InfCritRes$AICw),3)
+  m$Normal.copula.AICw<-round(r$InfCritRes$AICw[1],3)
+  m$p.CvM<-r$GofRes_CvM
+  m$p.KS<-r$GofRes_KS
+  m$LT.model.avg<-round(r$relLTdep_AICw,3)
+  m$UT.model.avg<-round(r$relUTdep_AICw,3)
+  m$`(LT-UT).model.avg`<-m$LT.model.avg-m$UT.model.avg
+  m$`Rank.(Pl-Pu)`<-rs$Rank_PlmPu_K
+  mt<-as.data.frame(t(m))
+  return(mt)
+
+}
+#----------------------------------
+#calling the above function on bivariable datasets
+m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke,rs=stat_soilCN_ga)
+m2<-bivar_tab_summary_fn(r=BivMS_birds,rs=stat_birds_ga)
+m3<-bivar_tab_summary_fn(r=BivMS_mammals,rs=stat_mammals_ga)
+m4<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
+tab_bivar_summary<-cbind(m1,m2,m3,m4)
+colnames(tab_bivar_summary)<-c("SoilCN","Birds' BMR","Mammals' BMR","Ceder-creek")
+saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
+```
 
@@
 
-For all four bivariate datasets the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
-for 1000 Kendall-preserving surrogate datasets with N copulas, and the rank of the real-data value in 
-the distribution of surrogate values was computed. If ranks were $>975$ or $<25$, asymmetry of tail 
-dependence was significant compared to a N-copula null hypothesis. Ranks were 1, 1, 1, 998
-<!--***DAN: Shya, please make the quantities in the previous line automatic links-->
+For first three bivariate datasets in Table \ref{MT-tab_bivar_summary},
+the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
+for 1000 Kendall-preserving surrogate datasets with N copulas. Due to relatively smaller sample size,
+for Cedar creek data we used a wider range of boundaries $(l_b,u_b)$ to compute $P_{0,0.25}-P_{0.75,1}$.
+For each data, rank of the said statistics with real-data value in 
+the distribution of surrogate values was computed. 
+If ranks were $>975$ or $<25$, asymmetry of tail 
+dependence was significant compared to a N-copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
 for, respectively, the soil C and N data, bird masses and metabolic rates, mammal masses
 and metabolic rates, and the Cedar Creek data. Thus the first three datasets showed stronger 
-upper- than lower-tail dependence, and the Cedar Creek data showed the reverse, confirming the 
-results based on model selection presented in the main text.
+upper- than lower-tail dependence, and the Cedar Creek data which showed the reverse though not significantly with
+non-parametric analysis, confirming the 
+results based on model selection presented in the main text (see Table \ref{MT-tab_bivar_summary}, row 9 
+showed Cedar creek had relatively weak tail-dependence than the other three datasets).
 
@@
 
-# Verifying Q1 results for multivariate datasets using nonparametric statistics \label{Q1multvarnonparam}
-
-For multivariate datasets, we calculated Spearman and Kendall correlations
-and the statistics 
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
-for all pairs of sampling locations, and we then computed means across the pairs
-for each statistic. We used a spatial resampling scheme 
-(Appendix \ref{spatial_resamp}) to calculate confidence intervals of these means.
-
-For the green spruce aphid abundance data, mean values across all pairs of 
-locations of the Spearman and Kendall correlations and the statistics 
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ were positive and
-confidence intervals excluded $0$ (Table \ref{tab_count_resamp}). 
-Mean values of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
-also were always positive and had confidence intervals that excluded $0$ 
-(Table \ref{tab_count_resamp}), confirming the results in the main text on 
-asymmetry of tail dependence and 
-stronger lower- than upper-tail dependence for this dataset. Each of the statistics 
-was plotted against geographic (great-circle) distances between sampling locations, 
-across all pairs of sampling locations (Fig. \ref{fig_aphid_count}).
-
-Asymmetry of tail dependence was also verified by our nonparametric methods for
-the other multivariate datasets. For instance, the $95\%$ confidence intervals of the
-mean over pairs of sampling locations of the statistic $\Ps_l-\Ps_u$ were
-$(-0.085,-0.041)$ for the leaf-curling plum aphid first flight data, 
-indicating greater upper-tail dependence, and consistent with
-<!--***DAN: Shya, pls make above results autolink-->
-the results of Table \ref{MT-tab_multivar_summary} (row 9). 
-For the *Ceratium furca* abundance data, confidence intervals were
-$(0.011,0.044)$, indicating greater lower-tail dependence, and again
-<!--***DAN: Shya, pls make above results autolink-->
-consistent with the results of Table \ref{MT-tab_multivar_summary} (row 9).
-For the methane flux data, confidence intervals were 
-$(-0.008,0.019)$.
-
-# Spatial resampling with multivariate datasets for nonparametric statistics \label{spatial_resamp}
-
-The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
-widely used spline correlogram method; the same scheme was also used by @walteretal2017 
-as part of a method for 
-performing model selection among matrix regression models. For each pair of 
-locations $i \neq j$, the statistic
-of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
-$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
-$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
-values $s_{ij}$ were then arranged in
-a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
-$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
-average strength of 
-dependence, across all pairs of distinct locations, for data in the portion of the distributions 
-given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
-the mean characterizes average
-asymmetry of tail dependence across all pairs of locations. 
-Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
-distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
-revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
-or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
-N-copula null hypothesis, and answering question Q1 from the Introduction.
-Bootstrapping was performed by resampling the locations of measurement, with replacement. 
-This corresponds to 
-resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
-[@bjornstadfalck2001 ; @walteretal2017].
-
-<!--This chunk is for model selection results of soilCN dataset-->
-```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("OurBiCopSelect.R")
-set.seed(seed)
-families_Loecke<-c(1,3:10,13:14,16:20) 
-BivMS_res_Loecke<-OurBiCopSelect(u1=v_CN[,1],u2=v_CN[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
-saveRDS(BivMS_res_Loecke,file="./Results/fitting_results/BivMS_soilCN.RDS")
-```
-
-<!--Display model selection results for soilCN data in a table-->
-```{r tab_soilCNfit, echo=F, results="markup"}
-#knitr::opts_knit$set(kable.force.latex = TRUE)
-h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-colnames(h)<-c("copname","AICw","BICw","LT","UT")
-knitr::kable(h, #longtable = T, 
-             format = "latex",linesep = "",
-      caption.short = "Model fitting results for soil C and N dataset",
-      caption = "Results of fitting our 16 copula families to the soil C and N dataset.
-Copula name abbreviations are given in the main text.
-AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.\\label{tab_soilCNfit}",
-      booktabs = TRUE)
-```
-
-<!--This chunk is for npa testing results of soilCN dataset-->
-```{r stat_soilCN_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_CN,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
-source("bivfunctionplot.R")
-set.seed(seed) 
-stat_soilCN_ga<-bivfunctionplot(v=v_CN,resloc="./Results/stat_results/stat_soilCN/",nametag="soilCN",numbin=10)
-```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for soil C and N data.
-As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
-Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
-For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates.
-When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level).
-When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.
-\label{fig_soilCN_nonparam}}
-\end{center}
-\end{figure}
-
-<!--Function to generate a table for npa results-summary with bivariate data, this one about asymmetry of tail dependence-->
-```{r make_table_stat_fn, echo=F}
-
-fractxt<-function(fracval,numsurrog){
-  ptxt<-''
-  if (fracval>.5*numsurrog){                        
-    ptxt<-paste(">",fracval,sep='')
-  } 
-  if (fracval<=.5*numsurrog){
-    ptxt<-paste("<",numsurrog-fracval,sep='')
-  }
-  return(ptxt)
-}
-
-make_table_stat_fn<-function(data){
-
- #atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
- #                               "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
- #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
- atdres<-data.frame(Statistic=c("cor$_{FS}$-cor$_{LS}$",
-                               "P$_{FS}$-P$_{LS}$",
-                                "D$_{LS}^2$-D$_{FS}^2$"),
-                   Kendall=rep("",3),
-                   Spearman=rep("",3),stringsAsFactors = F)
-
- atdres[1,2]<-fractxt(data$corlmcoru_frac_K,numsurrog = data$numsurrog_success_K)
- atdres[1,3]<-fractxt(data$corlmcoru_frac_S,numsurrog = data$numsurrog_success_S)
- atdres[2,2]<-fractxt(data$PlmPu_frac_K,numsurrog = data$numsurrog_success_K)
- atdres[2,3]<-fractxt(data$PlmPu_frac_S,numsurrog = data$numsurrog_success_S)
- atdres[3,2]<-fractxt(data$D2umD2l_frac_K,numsurrog = data$numsurrog_success_K)
- atdres[3,3]<-fractxt(data$D2umD2l_frac_S,numsurrog = data$numsurrog_success_S)
-
- return(atdres)
-
-}
-```
-
-<!--Displaying a table with npa summary results for soilCN data-->
-```{r tab_soilCN_asym,echo=F,results="markup"}
-#knitr::opts_knit$set(kable.force.latex = TRUE)
-tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
-#tab_soilCN_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
-knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
-             format = "latex",linesep = "",
-      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
-      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
-The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons.
-A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
-```
-
-<!--Simulations for model selection results with Birds' BMR data-->
-```{r bivmodselect_birds_BMR, echo=F, cache=T, cache.extra=list(seed,cop_b,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("OurBiCopSelect.R")
-set.seed(seed)
-BivMS_birds<-OurBiCopSelect(u1=cop_b[,1],u2=cop_b[,2],families=c(1,3:10,13:14,16:20),
-                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
-                              gofnormal=F,status=F)
-saveRDS(BivMS_birds,file="./Results/BMR_results/birds/BivMS_birds.RDS")
-```
-
-<!--Display model selection results for BMR:birds data in a table-->
-```{r tab_bmr_birds_fit, echo=F, results='markup'}
-h<-BivMS_birds$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-colnames(h)<-c("copname","AICw","BICw","LT","UT")
-knitr::kable(h,
-             format='latex',linesep = "",
-             caption.short = "Model fitting results for the bird body mass and BMR dataset",
-             caption = "Results of fitting our 16 copula families to the bird body mass and BMR dataset.
-Analogous to Fig.
-\\ref{tab_soilCNfit}, see caption of that figure for details.
-\\label{tab_bmr_birds_fit}",
-             booktabs=T)
-```
-
-<!--This chunk is for npa testing results for Birds' BMR dataset-->
-```{r stat_bmr_birds_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_b,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
-source("bivfunctionplot.R")
-set.seed(seed)
-stat_birds_ga<-bivfunctionplot(v=cop_b,resloc="./Results/BMR_results/birds/",nametag="birds_numbin_10",numbin = 10)
-```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for bird body mass and BMR data.
-Same format as Fig.
-\ref{fig_soilCN_nonparam}, see caption of that figure.
-\label{fig_bmr_birds_nonparam}}
-\end{center}
-\end{figure}
-
-<!--Displaying a table with npa summary results for Birds' BMR data-->
-```{r tab_bmr_birds_asym, echo=F, results="markup"}
-tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
-#tab_bmr_birds_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
-knitr::kable(tab_bmr_birds_asym,
-             format="latex",linesep = "",escape=F,
-             caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
-             caption = "Bird body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
-Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_bmr_birds_asym}",
-             booktabs=T)
-```
-
-<!--Simulations for model selection results with Mammals' BMR data-->
-```{r bivmodselect_mammals_BMR, echo=F, cache=T, cache.extra=list(seed,cop_m,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("OurBiCopSelect.R")
-set.seed(seed)
-BivMS_mammals<-OurBiCopSelect(u1=cop_m[,1],u2=cop_m[,2],families=c(1,3:10,13:14,16:20),
-                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
-                              gofnormal=F,status=F)
-saveRDS(BivMS_mammals,file="./Results/BMR_results/mammals/BivMS_mammals.RDS")
-```
-
-<!--Display model selection results for BMR:mammals data in a table-->
-```{r tab_bmr_mammals_fit, echo=F, results='markup'}
-h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-colnames(h)<-c("copname","AICw","BICw","LT","UT")
-knitr::kable(h,
-             format='latex',linesep = "",
-             caption.short = "Model fitting results for the mammal body mass and BMR dataset",
-             caption = "Results of fitting our 16 copula families to the mammal body mass and BMR dataset.
-Analogous to Fig.
-\\ref{tab_soilCNfit}, see caption of that figure for details.
-\\label{tab_bmr_mammals_fit}",
-             booktabs=T)
-```
-
-<!--This chunk is for npa testing results for Mammals' BMR dataset-->
-```{r stat_bmr_mammals_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_m,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
-source("bivfunctionplot.R")
-set.seed(seed)
-stat_mammals_ga<-bivfunctionplot(v=cop_m,resloc="./Results/BMR_results/mammals/",nametag="mammals_numbin_10",numbin = 10)
-```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for mammal body mass and BMR data.
-Same format as Fig.
-\ref{fig_soilCN_nonparam}, see caption of that figure.
-\label{fig_bmr_mammals_nonparam}}
-\end{center}
-\end{figure}
-
-<!--Displaying a table with npa summary results for Mammals' BMR data-->
-```{r tab_bmr_mammals_asym, echo=F, results="markup"}
-tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
-#tab_bmr_mammals_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
-knitr::kable(tab_bmr_mammals_asym,
-             format="latex",linesep = "",escape=F,
-             caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
-             caption = "Mammal body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
-Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.
-\\label{tab_bmr_mammals_asym}",
-             booktabs=T)
-```
-
-<!--Simulations for model selection results with Cedar creek data-->
-```{r bivmodselect_cc, echo=F, cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("OurBiCopSelect.R")
-set.seed(seed)
-
-BivMS_res_cc<-vector("list",length(cop_HB_all_yr))
-names(BivMS_res_cc)<-names(cop_HB_all_yr)
-
-for(i in 1:length(BivMS_res_cc)){
-  BivMS_res_cc[[i]]<-OurBiCopSelect(u1=cop_HB_all_yr[[i]]$avg.H,u2=cop_HB_all_yr[[i]]$avg.Biomass,
-                                    families=c(1,3:10,13:14,16:20),
-                                    level=0.05,AICBIC="AIC",
-                                    numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
-}
-saveRDS(BivMS_res_cc,file="./Results/Ceder_creek_results/BivMS_res_cc.RDS")
-BivMS_res_cc<-readRDS("./Results/Ceder_creek_results/BivMS_res_cc.RDS")
-```
-
-<!--Display model selection results for Cedar-creek data for year 2000 in a table-->
-```{r tab_cc2000fit, echo=F, results='markup'}
-h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-colnames(h)<-c("copname","AICw","BICw","LT","UT")
-knitr::kable(h,
-             format='latex',linesep = "",
-             caption.short = "Model fitting results for the Cedar Creek dataset",
-             caption = "Results of fitting our 16 copula families to the Cedar Creek dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_cc2000fit}",
-             booktabs=T)
-```
-
-<!--This chunk is for npa testing results for Cedar creek dataset-->
-```{r stat_cc_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_HB_all_yr,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
-source("bivfunctionplot.R")
-set.seed(seed)
-
-stat_cc_ga<-vector("list",length(cop_HB_all_yr))
-names(stat_cc_ga)<-names(cop_HB_all_yr)
-
-for(i in 1:length(stat_cc_ga)){
-  temp<-bivfunctionplot(v=cop_HB_all_yr[[i]],
-                             resloc="./Results/Ceder_creek_results/",
-                             nametag=paste0(names(cop_HB_all_yr)[i],"cc_numbin_10",sep=""),
-                             numbin = 10)
-  stat_cc_ga[[i]]<-temp
-}
-
-saveRDS(stat_cc_ga,"./Results/Ceder_creek_results/stat_cc_ga.RDS")
-```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data.
-Same format as Fig.
-\ref{fig_soilCN_nonparam}, see caption of that figure.\label{fig_cc2000_nonparam}}
-\end{center}
-\end{figure}
-
-<!--Displaying a table with npa summary results for Cedar creek data-->
-```{r tab_cc_asym_2000, echo=F, results="markup"}
-tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
-#tab_cc_asym_2000[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
-knitr::kable(tab_cc_asym_2000,
-             format="latex",linesep = "",escape=F,
-             caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
-             caption = "Cedar Creek dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
-Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_cc_asym_2000}",
-             booktabs=T)
-```
-
 <!--Load the raw data for aphid-->
@@
 \end{figure}
-
-<!--Function to generate a table containing summary results for all bivariate datasets-->
-```{r bivar_tab_summary_fn,echo=F}
-bivar_tab_summary_fn<-function(r,rs){
-
-  m<-matrix(NA,1,10)
-  colnames(m)<-c("p.Independence.test","minimumAIC.bestfit.copula","Bestfit.copula.AICw",
-                 "Normal.copula.AICw","p.CvM","p.KS",
-                 "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
-                 "Rank.(Pl-Pu)")
-  m<-as.data.frame(m)
-
-  m$p.Independence.test<-round(r$IndepTestRes,3)
-  m$minimumAIC.bestfit.copula<-r$InfCritRes$copname[which.min(r$InfCritRes$AIC)]
-  m$Bestfit.copula.AICw<-round(max(r$InfCritRes$AICw),3)
-  m$Normal.copula.AICw<-round(r$InfCritRes$AICw[1],3)
-  m$p.CvM<-r$GofRes_CvM
-  m$p.KS<-r$GofRes_KS
-  m$LT.model.avg<-round(r$relLTdep_AICw,3)
-  m$UT.model.avg<-round(r$relUTdep_AICw,3)
-  m$`(LT-UT).model.avg`<-m$LT.model.avg-m$UT.model.avg
-  m$`Rank.(Pl-Pu)`<-rs$Rank_PlmPu_K
-  mt<-as.data.frame(t(m))
-  return(mt)
-
-}
-#----------------------------------
-#calling the above function on bivariable datasets
-m1<-bivar_tab_summary_fn(r=BivMS_res_Loecke,rs=stat_soilCN_ga)
-m2<-bivar_tab_summary_fn(r=BivMS_birds,rs=stat_birds_ga)
-m3<-bivar_tab_summary_fn(r=BivMS_mammals,rs=stat_mammals_ga)
-m4<-bivar_tab_summary_fn(r=BivMS_res_cc$`2000`,rs=stat_cc_ga$`2000`)
-tab_bivar_summary<-cbind(m1,m2,m3,m4)
-colnames(tab_bivar_summary)<-c("SoilCN","Birds' BMR","Mammals' BMR","Ceder-creek")
-saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
-```
 
@@
 ```
+
+# Verifying Q1 results for multivariate datasets using nonparametric statistics \label{Q1multvarnonparam}
+
+For multivariate datasets, we calculated Spearman and Kendall correlations
+and the statistics 
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
+for all pairs of sampling locations, and we then computed means across the pairs
+for each statistic. We used a spatial resampling scheme 
+(Appendix \ref{spatial_resamp}) to calculate confidence intervals of these means.
+
+For the green spruce aphid abundance data, mean values across all pairs of 
+locations of the Spearman and Kendall correlations and the statistics 
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ were positive and
+confidence intervals excluded $0$ (Table \ref{tab_count_resamp}). 
+Mean values of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
+also were always positive and had confidence intervals that excluded $0$ 
+(Table \ref{tab_count_resamp}), confirming the results in the main text on 
+asymmetry of tail dependence and 
+stronger lower- than upper-tail dependence for this dataset. Each of the statistics 
+was plotted against geographic (great-circle) distances between sampling locations, 
+across all pairs of sampling locations (Fig. \ref{fig_aphid_count}).
+
+Asymmetry of tail dependence was also verified by our nonparametric methods for
+the other multivariate datasets. For instance, the $95\%$ confidence intervals of the
+mean over pairs of sampling locations of the statistic $\Ps_l-\Ps_u$ were
+$(-0.085,-0.041)$ for the leaf-curling plum aphid first flight data, 
+indicating greater upper-tail dependence, and consistent with
+<!--***DAN: Shya, pls make above results autolink-->
+the results of Table \ref{MT-tab_multivar_summary} (row 9). 
+For the *Ceratium furca* abundance data, confidence intervals were
+$(0.011,0.044)$, indicating greater lower-tail dependence, and again
+<!--***DAN: Shya, pls make above results autolink-->
+consistent with the results of Table \ref{MT-tab_multivar_summary} (row 9).
+For the methane flux data, confidence intervals were 
+$(-0.008,0.019)$.
+
+# Spatial resampling with multivariate datasets for nonparametric statistics \label{spatial_resamp}
+
+The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
+widely used spline correlogram method; the same scheme was also used by @walteretal2017 
+as part of a method for 
+performing model selection among matrix regression models. For each pair of 
+locations $i \neq j$, the statistic
+of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
+$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
+$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
+values $s_{ij}$ were then arranged in
+a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
+$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterizes the 
+average strength of 
+dependence, across all pairs of distinct locations, for data in the portion of the distributions 
+given by the bounds $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, 
+the mean characterizes average
+asymmetry of tail dependence across all pairs of locations. 
+Confidence intervals for the value of $\mean(S)$ were then computed as quantiles of a 
+distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
+revealed by the confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
+or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
+N-copula null hypothesis, and answering question Q1 from the Introduction.
+Bootstrapping was performed by resampling the locations of measurement, with replacement. 
+This corresponds to 
+resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
+[@bjornstadfalck2001 ; @walteretal2017].
 

### 2019-04-10 SupportingInformation.Rmd
adding details for Cedar Creek data

@@
     tab_temp$abs_cov[1:li]<-x$Abscover[ind]
-    tab_each_sp$avg_cov[[sp]]<-mean(tab_temp$abs_cov) # avg. over 4Quadrates for a sp.
+    tab_each_sp$avg_cov[[sp]]<-mean(tab_temp$abs_cov) # avg. over Quadrats for a sp.
    }
@@
 <!--***DAN: This is a dumping ground at the moment, but needs to be organized into a section that gives all the necessary details about the data-->
+
+<!--additional details about Cedar creek data-->
+Above-ground biomass and percent cover data were obtained from long-term (? or short-term)
+biodiversity experiment (E120) conducted in Cedar Creek Ecosystem Science Reserve by 
+@Tilman2001; @Tilman2006. In 1994, 168 plots (each 9 m $\times$ 9 m) of 39, 35, 29, 30 and 35 replicates,
+were seeded with 1, 2, 4, 8, or 16 perennials, respectively, chosen randomly from a pool of 
+18 species (4 species, each, of C4 grasses, C3 grasses, legumes, non-legume forbs; 2 species of woody
+plants). All plots received, in total, 10 $gm^{-2}$ of pure live seed in May 1994 and 5 $gm^{-2}$ in May, 1995 
+with seed mass divided equally among species. Treatments were maintained by weeding manually 3-4 times a year 
+and sometimes selective herbicides were used to set up the monocultures either with only grasses or with only forbs.
+We downloaded the data on July, 2018 from Cedar Creek website \url{http://www.cedarcreek.umn.edu/research/experiments/e120} and 
+analyzed those [data](https://github.com/sghosh89/BIVAN/tree/master/Data/Ceder_creek_e120) for 6 common years 
+(1996-2000, 2007) when both biomass and percent cover data were available. Each plots were annually sampled by clipping 
+different locations each year. For each available year, we computed above-ground biomass ($gm^{-2}$) 
+and Shannon's biodiversity index (using percent cover data) averaged over the sampled strips/ quadrats taken from each 
+of 168 plots.
+
+<!--we need to add aphid and plankton data details, Lawrence will give those-->
 

### 2019-04-10 SupportingInformation.Rmd
minor change
testing git

@@
 
-<!--additional details about Cedar creek data-->
+<!--additional details about Cedar creek data, etc.-->
 Above-ground biomass and percent cover data were obtained from long-term (? or short-term)
-biodiversity experiment (E120) conducted in Cedar Creek Ecosystem Science Reserve by 
-@Tilman2001; @Tilman2006. In 1994, 168 plots (each 9 m $\times$ 9 m) of 39, 35, 29, 30 and 35 replicates,
+biodiversity experiment (E120) conducted in Cedar Creek Ecosystem Science Reserve by Tilman *et al*.
+In 1994, 168 plots (each 9 m $\times$ 9 m) of 39, 35, 29, 30 and 35 replicates,
 were seeded with 1, 2, 4, 8, or 16 perennials, respectively, chosen randomly from a pool of 

### 2019-04-12 SupportingInformation.Rmd
no change : just unique quadrats checked per plot in 1999

@@
    Q<-as.character(unique(x$Quadrat))
-
+   #cat("----------total unique quadrats = ",length(Q),"--------------------\n")
    tab_each_sp<-data.frame(sp=uniq_sp,avg_cov=0,rel_avg_cov=0)

### 2019-04-13 Paper.Rmd
comments added as suggested by PCR

@@
 However, ubiquitously used approaches to studying such relationships, based on correlation and regression, provide only a small slice of the complex information content of the relationships.
-Statistical tools do exist that provide a complete description of relationships, based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, but not in ecology.
+Other statistical tools exist that provide a complete description of relationships, based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, but not in ecology.
 We here examine the potential for copula approaches to improve our understanding of ecology.
@@
 and positive values correspond to undercompensating dynamics; larger-magnitude $b$
-mean slower return time of the model to equilibrium after a perturbation. 
+means that the model has a slower return time to equilibrium after a perturbation. 
 The $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
@@
 $g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
-$g_1$ represent environmental effects that occur only for negative 
-values of the environmental variables, and $g_2$ represent environmental effects 
-that occur only for positive values. Note that both $(\epsilon_1,\epsilon_2)$
+$g_1$ represents environmental effects that occur only for negative 
+values and $g_2$ only for positive values of the environmental variables.
+Note that both $(\epsilon_1,\epsilon_2)$
 and $(\delta_1,\delta_2)$ have N copula structure, being drawn from 
@@
 ecological-variable copula structure and tail dependencies can be reduced when the
-intrinsic dynamics of the ecological variable is strongly nonlinear.
+intrinsic dynamics of the ecological variable are strongly nonlinear.
 For simulations using AR(1) models (see section on methods for Q2), copula 
@@
 \ref{SM-fig_ms_asym_b_-0.5}).
-Lower-tail (respectively, upper-tail) effects of the noise ($g_1$, resp., $g_2$) 
-produced lower-tail (resp., upper-tail) dependence in the ecological variable.
+Lower-tail (respectively, upper-tail) effects of the noise ($g_1$, respectively, $g_2$) 
+produced lower-tail (respectively, upper-tail) dependence in the ecological variable.
 Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
@@
 <!--what we did and why is it important to study?-->
-We showed that non-N copula structure and asymmetric tail dependence are 
+We have shown that non-N copula structure and asymmetric tail dependence are 
 exceedingly common in datasets that range across multiple sub-disciplines
@@
 non-N and tail-dependent copula structures have often been found in prior analyses of 
-environmental variables. If large meteorological events are also spatially widespread, then
+environmental variables. A recent study [@agliardi2019relationship] found non-N 
+copulas (mainly Gumbel and Survival Clayton) were better fit in modelling the 
+upper tail dependence between extreme large values for global 
+temperature and annual mean green house gas concentrations for an extensively
+long datasets (1880-2017). If large meteorological events are also spatially widespread, then
 one would expect environmental variables associated with these events to take extreme
@@
 We thank James Bell of the Rothamsted Insect Survey (RIS).
-The Rothamsted Insect Survey, a National Capability, is funded by the Biotechnology and Biological Sciences Research Council under the Core Capability Grant BBS/E/C/000J0200.
+This survey, a UK Capability, is funded by the Biotechnology and Biological Sciences Research Council under the Core Capability Grant BBS/E/C/000J0200.
 SG, LWS and DCR were partly funded by US National Science Foundation grants 0075516 and 1442595 and the James S McDonnell Foundation.

### 2019-04-23 Data/BMR/simulations/text.tex
draft of text

%%%%%%%%%%%%%%%%%%%%%%%% Simulation text

%%%%%%%%%%%%%%%%%%%%%%%% Discussion text

Performing statistical tests for associations between continuous traits in different species
    was a primary motivating example of phylogenetic comparative methods.
For example, Felsenstein's method of phylogenetically independent contrasts 
    \citep{Felsenstein1985} is currently the second most-cited paper in the history
    of the journal {\em The American Naturalist} \citep{HueyGT2019}.
Frustratingly, the field still lacks a dependable procedure for dealing with branch lengths, 
    which are a crucial input to the method.
Ideally, the branch lengths used to correct for phylogenetic effects would represent
    the expected amounts of change for the characters that are being analyzed.
Because researchers almost never have a reliable method for providing such branch lengths,
    most researchers rely on ultrametric trees -- those for which the branch
    length can be treated as a proxy for the duration of the branch in time.
Frequently these branch lengths are transformed to assess sensitivity to different 
    assumptions about the degree of phylogenetic inertia displayed by the traits
    under study \citep[See recent reviews, ][]{Ives2018,Harmon2018}.
Even if one were able to simply use a time-based set of branch lengths, assign
    dates to nodes in phylogenies is difficult.
DNA sequence data can provide estimates of branch lengths, but these estimates
    are dependent on the adequacy of models which correct sequences for multiple
    substitutions occurring at the same location.
Biases in estimating the evolutionary distance can affect down stream analyses \citep[See][]{Phillips2009}.
Additionally, changes
    in the rate of molecular evolution make the estimation of dates difficult \citep[See][]{HeathM2014} even when branch lengths are accurately estimated.

Without reliable branch length estimates, it is difficult to interpret the significance
    of the magnitude of changes in traits across a tree.
Developing tests of association based on copula structure could lead the way to more
    robust methods for studying associations when we lack defensible estimates of
    branch lengths.
We note that our phylogenetic analyses here consisted merely of simulations to assess
    whether interesting copula structure in a simple evolutionary process could 
    leave a detectable signal on the data.
Significant work remains to be done before we have a copula-based method for 
    analyzing data on a phylogenetic tree.

%%%%%%%%%%%%%%%%%%%%%%%%  Refs

@article{Phillips2009,
title = "Branch-length estimation bias misleads molecular dating for a vertebrate mitochondrial phylogeny",
journal = "Gene",
volume = "441",
number = "1",
pages = "132 - 140",
year = "2009",
note = "Phylogenomics and its Future: Devoted to Masami Hasegawa",
issn = "0378-1119",
doi = "https://doi.org/10.1016/j.gene.2008.08.017",
url = "http://www.sciencedirect.com/science/article/pii/S0378111908004150",
author = "Matthew J. Phillips",
keywords = "Fossil calibration, Maximum likelihood, Mitochondrion, Model misspecification, RY-coding",
abstract = "Despite recent methodological advances in inferring the time-scale of biological evolution from molecular data, the fundamental question of whether our substitution models are sufficiently well specified to accurately estimate branch-lengths has received little attention. I examine this implicit assumption of all molecular dating methods, on a vertebrate mitochondrial protein-coding dataset. Comparison with analyses in which the data are RY-coded (AG→R; CT→Y) suggests that even rates-across-sites maximum likelihood greatly under-compensates for multiple substitutions among the standard (ACGT) NT-coded data, which has been subject to greater phylogenetic signal erosion. Accordingly, the fossil record indicates that branch-lengths inferred from the NT-coded data translate into divergence time overestimates when calibrated from deeper in the tree. Intriguingly, RY-coding led to the opposite result. The underlying NT and RY substitution model misspecifications likely relate respectively to “hidden” rate heterogeneity and changes in substitution processes across the tree, for which I provide simulated examples. Given the magnitude of the inferred molecular dating errors, branch-length estimation biases may partly explain current conflicts with some palaeontological dating estimates."
}

@misc{Ives2018,
  title={Mixed and phylogenetic models: a conceptual introduction to correlated data},
  author={Ives, AR},
  year={2018},
  publisher={Leanpub}
}

@article{Harmon2018,
  title={Phylogenetic comparative methods: learning from trees},
  author={Harmon, LJ},
  journal={Self published under a CC-BY-4.0 license},
  year={2018}
}

@article{HeathM2014,
  title={Bayesian inference of species divergence times},
  author={Heath, TA and Moore, BR},
  journal={Bayesian phylogenetics: methods, algorithms, and applications},
  pages={277--318},
  year={2014},
  publisher={Sinauer Associates, Sunderland, MA}
}

### 2019-04-23 Data/BMR/simulations/text.tex
some sim details

@@
 %%%%%%%%%%%%%%%%%%%%%%%% Simulation text
+We simulated bivariate character evolution on an estimate of 
+  phylogeny for 817 mammal species taken from \citep{GenoudIM2018}.
+The simulator was written in Python and used version 4.4 of 
+  the DendroPy \citep{DendroPy} package.
 
@@
 }
+
+@article{GenoudIM2018,
+author = {Genoud, Michel and Isler, Karin and Martin, Robert D.},
+title = {Comparative analyses of basal rate of metabolism in mammals: data selection does matter},
+journal = {Biological Reviews},
+volume = {93},
+number = {1},
+pages = {404-438},
+keywords = {basal rate of metabolism, BMR, brain, encephalization, allometric scaling, phylogenetic generalized least-squares regression, mammals},
+doi = {10.1111/brv.12350},
+url = {https://onlinelibrary.wiley.com/doi/abs/10.1111/brv.12350},
+eprint = {https://onlinelibrary.wiley.com/doi/pdf/10.1111/brv.12350},
+year = {2018}
+}
+
+@article{DendroPy,
+    author = {Sukumaran, Jeet and Holder, Mark T.},
+    title = "{DendroPy: a Python library for phylogenetic computing}",
+    journal = {Bioinformatics},
+    volume = {26},
+    number = {12},
+    pages = {1569-1571},
+    year = {2010},
+    month = {04},
+    abstract = "{Summary: DendroPy is a cross-platform library for the Python programming language that provides for object-oriented reading, writing, simulation and manipulation of phylogenetic data, with an emphasis on phylogenetic tree operations.
+DendroPy uses a splits-hash mapping to perform rapid calculations of tree distances, similarities and shape under various metrics.
+It contains rich simulation routines to generate trees under a number of different phylogenetic and coalescent models.
+DendroPy's data simulation and manipulation facilities, in conjunction with its support of a broad range of phylogenetic data formats (NEXUS, Newick, PHYLIP, FASTA, NeXML, etc.), allow it to serve a useful role in various phyloinformatics and phylogeographic pipelines.Availability: The stable release of the library is available for download and automated installation through the Python Package Index site (http://pypi.python.org/pypi/DendroPy), while the active development source code repository is available to the public from GitHub (http://github.com/jeetsukumaran/DendroPy).Contact:jeet@ku.edu}",
+    issn = {1367-4803},
+    doi = {10.1093/bioinformatics/btq228},
+    url = {https://doi.org/10.1093/bioinformatics/btq228},
+    eprint = {http://oup.prod.sis.lan/bioinformatics/article-pdf/26/12/1569/809362/btq228.pdf},
+}

### 2019-04-25 SupportingInformation.Rmd
log-log plot for soil CN modified

@@
 op<-par(mar=c(6, 6, 0.2, 2.1))
-plot(d[,1],d[,2],type="p",col="black",xlab="Soil C",ylab="Soil N",cex.lab=2.5,cex.axis=2.5)
+d_log<-log10(d[,c(1:2)])
+plot(d_log[,1],d_log[,2],type="p",col="black",xlab="log(Soil C)",ylab="log(Soil N)",cex.lab=2.5,cex.axis=2.5)
 par(op)
@@
 pdf("./Results/SoilCNcop.pdf")
-v_CN<-getcopula(d=d[,c(1:2)],rankon=T,ploton=T) 
+v_CN<-getcopula(d=d_log,rankon=T,ploton=T) 
 dev.off()

### 2019-04-26 Paper.Rmd
log-log plot for soil CN modified
edits beginning to incoporate Holder stuff

@@
 \includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
-\caption{Upper panels show raw data plots for (A) soil C and N data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
+\caption{Upper panels show raw data plots for (A) $\log_{10}$(soil C) and $\log_{10}$(soil N) data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
 for birds, (C) $\log_{10}$(BMR) vs.
@@
 Due to smaller sample size relative to other three bivariate datasets, we chose a wider range with 
-non-parametric analysis (see Appendix \ref{SI-Q1bivarnonparam}).
+non-parametric analysis (Appendix \ref{SM-Q1bivarnonparam}).
 Detailed results for each dataset are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and Figs.
@@
 Nonparametric statistics verified that tail dependence was asymmetric for the green 
-spruce aphid abundance data (Appendix \ref{SM-Q1multvarnonparam}).
+spruce aphid abundance data (Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}).
 
@@
 <!--(e.g., Table \ref{tab_multivar_summary}, row 10)-->
-(Appendix \ref{SM-Q1multvarnonparam}). Detailed results are in 
+(Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}). Detailed results are in 
 Tables \ref{SM-tab_count_bestcop}-\ref{SM-tab_methane_resamp} and 
@@
 features of ecological data, we addressed Q2 by exploring via mathematical 
-models two possible mechanisms of this structure which 
-seem likely to operate in diverse ecological scenarios. The first mechanism 
+models and simulations three possible mechanisms of this structure which 
+seem likely to operate in diverse ecological and evolutionary scenarios. The first mechanism 
 is an extension of the well known Moran
@@
 plots should be similar for populations and noises.
-
-<!--***DAN: Note to self, this model, and also the other one we study in this section,
-are actually much more general than population. Should replace P_i with E_i where E_i is
-just some ecological variable. Could be deviaitons from a pop from carrying capacity, could be a 
-biogeochemical variable, could maybe even be an evolutinary variable. Anything that is
-affected by an environmental variable and could reasonably be modelled by an AR process. 
-So the generalization to biogeochemical and community
-data seems pretty straightforward. Need to think this through and reword the corresponding
-part of the intro to encompass this greater generality. How would the evolutionary case work?
-Well, suppose the two "patches" are actually traits, and the perturbations epsilon are somehow
-the "drift events" or something like that? This does not really make sense, strictly speaking,
-now that I am thinking about it, but maybe the model could be modified to make it make sense?
-At least I have straightforward generalization to biogeochem and community cases. 
-More on evolutionary: suppose we model traits as a bivariate random walk where the perturbations
-in the traits are related with a copula with some tail dependence. Suppose you have a tree and 
-you model evolution down the tree in this way and then look at the bivariate distribution of
-trait on the tips of the tree. What do you get? I guess you get some copula structure? maybe
-this thinking can lead to some isnights into how copula structure we observed in the species
-trait data might have come about?-->
-
-<!--***DAN: note to self. Suppose deviations from the 
-regression line for log met rate v log M are 
-more costly at one end than at another, for whatever reason. That is a proximate
-mechanism for asymmetric tail dependence?-->
 
@@
 
+The third mechanism we explored is evolutionary, and pertains to bivariate 
+trait data across species, e.g., to our bird and mammal BMR versus body mass 
+data. This mechanism is a hypothetical explanation for the right tail dependence
+observed in those data (Fig. \ref{fig_biv_multi_raw_cop}F,G). 
+We simulated bivariate character evolution on an estimate of 
+phylogeny for $817$ mammal species taken from @GenoudIM2018.
+The root character state and the change 
+across each branch were randomly chosen from matrices of one million different
+bivariate draws from distributions showing one of five distinct types of copula 
+structure: 1) extreme left-tail dependence, 2) moderate left-tail dependence,
+3) symmetric tail dependence, 4) moderate right-tail dependence, and 5) extreme
+right-tail dependence (Appendix \ref{SM-evol_mech} for details). 
+All marginal distributions were standard normal, and distributions all had Spearman 
+correlation $0.875$, so simulations assess the impact of copula structure of character
+changes on evolutionary outcomes, controlling for marginals and correlations of character
+changes. For each of the five copula structures, mammalian character
+evolution was simulated 100 times. For each simulation, symmetry of tail dependence 
+of the two characters was assessed across the phylogeny tips.
+We hypothesized that cases 1 and 2 above would yield stronger left- than right-tail 
+dependence, and cases 4 and 5 would yield the reverse. 
+The simulator was written in Python and used version 4.4 of the DendroPy 
+package [@DendroPy].
+
 # Results for Q2: Moran-like effects and asymmetric dependencies produce non-normal copula structure\label{Results_Q2}
@@
 
-<!--***DAN: Maybe need to think about mechanisms for biogeochemical, evolutionary, and community data-->
+Our hypothesis was correct that asymmetric tail dependence in evolutionary changes in bivariate
+characters can give rise to similarly asymmetric tail dependence in character distributions across
+phylogeny tips (Appendix \ref{SM-evol_mech}). This topic is revisited in the Discussion.
 
@@
 
-<!--8) A para about what is available in the copula and VineCopula packages.
+<!--A para about what is available in the copula and VineCopula packages.
 Maybe make it a broader para about emthods, and state in there that we developed some new surrogate methods that may be broadly useful for researches trying to incorporate copula approaches.-->
@@
 
-<!--4) We have only given two mechanisms, but there are likely to be others, and discoverig
-additional mechanisms of copula structure is certain worthy of future work. Maybe I can think
-of some ideas and make a paragraph.
-5) If Van gives us some ideas, they may fit in the Discussion here.
-6) Maybe give Holder a bunch of draws from a bivariate dist with standard normals marginals but with 
-some tail-dependent copula structure and ask him to simulate evolution down a mammal or bird tree and
-see what you get. Then give him the other tail dependence, and no tail dependence (bivariate standard
-normal). Maybe this will lead to more results, and also a Discussion para on why evolutionary steps 
-might be tail dependent. A result that tail dependent steps lead to tail dependent character dists
-in the end would replace one question (why are there tail-dependent characters) with another (why are
-there tail dependent steps). Of course it could also be the case that steps are not tail dependent and 
-the tail dependence of characters arises in another way. Would also lead in to alternative thinking in 
-evolutionary biology - Mark could write a para on why it would matter for evolutionary biol if there are 
-tail dependent steps.
--->
+<!--Holder stuff-->
+Our simulations of character evolution suggest the hypothesis that evolutionary
+changes in bird and mammal BMR and body size may exhibit right-tail dependence. This is
+a hypothesis only, because the tail dependence of Fig. \ref{fig_biv_multi_raw_cop}F,G
+could have come about in another, unknown way: our simulations show that tail dependence 
+in evolutionary changes is sufficient, but may not be necessary, to produce tail
+dependence in characters of extant species. Even if the hypothesized mechanism
+is correct, our results simply replace one question, i.e., why do we see right-tail
+dependence between BMR and body mass for birds and mammals, with another question,
+i.e., why might we see right-tail dependence in evolutionary changes of these traits.
+See Appendix \ref{evol_mech} for additional thoughts on the potential importance of 
+copulas for studies of character evolution when phylogeny branch lengths are not available.
+
+<!-- We have only given three mechanisms, but there are likely to be others, and discoverig
+additional mechanisms of copula structure is certain worthy of future work. -->
 Other mechanisms that we have not considered likely also operate to produce copula structure
-in ecological datasets.
-For instance, we detected non-N structure and tail dependence in our 
-metabolic rate-versus-body mass datasets,
-but the mechanisms discussed above do not apply directly in a macroevolutionary context.
-We hope by enumerating two potential mechanisms we help inspire additional theoretical and
-empirical research on the
-potentially numerous mechanisms of copula structure that may apply in diverse ecological datasets.
-<!--***DAN: beef this up if/when we get inputs from Mark and Van-->
-
-<!--***DAN: some consequences paragraphs next-->
+in ecological and evolutionary datasets, and should be considered in future work. For instance, 
+measurement error may modify copula structure, though one would have to have information
+on the precise nature of measurement error in a dataset to explore its influence. 
+Not only did the models we used to explore mechanisms not consider measurement error, 
+they were extremely simple in other respects, as well, not including phenomena such as 
+delayed density dependence, dispersal, stage structure in population dynamics, and trophic
+interactions.
+We hope by enumerating three potential mechanisms we help inspire additional research on the
+potentially numerous mechanisms of copula structure that may apply in diverse datasets.
+
+<!--some consequences paragraphs next-->
 
@@
 
-<!--Maybe a para from Holder on his thoughts in the importance of copulas for
-macro-scale evolutionary biology?-->
-
 # Acknowledgments
 
-We thank the many contributors to the large datasets we used; D.
-Stevens and P.
-Verrier for data extraction; and Joel E.
-Cohen, Jonathan Walter, Thomas Anderson, and Lei Zhao for helpful suggestions.
-We thank James Bell of the Rothamsted Insect Survey (RIS).
-This survey, a UK Capability, is funded by the Biotechnology and Biological Sciences Research Council under the Core Capability Grant BBS/E/C/000J0200.
-SG, LWS and DCR were partly funded by US National Science Foundation grants 0075516 and 1442595 and the James S McDonnell Foundation.
+We thank the many contributors to the large datasets we used; D. Stevens and P. Verrier for data 
+extraction; and Joel E. Cohen, David Tilman, Lauren Hallet, Jim Bever, Jonathan Walter, Thomas 
+Anderson, and Lei Zhao for helpful suggestions. We thank James Bell of the Rothamsted Insect Survey 
+(RIS). 
+The RIS, a UK Capability, is funded by the Biotechnology and Biological Sciences Research Council 
+under the Core Capability Grant BBS/E/C/000J0200. SG, LWS and DCR were partly funded by US National 
+Science Foundation grants 0075516 and 1442595 and the James S McDonnell Foundation.
 

### 2019-04-26 SupportingInformation.Rmd
edits beginning to incoporate Holder stuff

@@
 
+# Hypothetical evolutionary mechanism for tail dependence between traits \label{evol_mech}
+
+\noindent The matrices from which root character states and changes 
+across phylogeny branches were randomly chosen were generated as follows.
+For extreme right-tail dependent noise, we first produced
+a $1000000$ by $2$ matrix, $m$, of 
+independent standard-normal random draws. Then, each row of $m$ 
+was identified for a first set of modifications randomly 
+and independently with a $50\%$ chance; 
+for those rows identified, every element of the row was replaced with the 
+absolute value of the first element of the row (including the first element itself).
+For other rows, every element in the row was replaced by the negative of its
+own absolute value. This produces right-tail dependent noise, normally distributed
+for each location. If left-tail dependent noise was desired, we took the negative of the
+entire matrix. The moderately left-tail dependent matrix was based on a C copula 
+with parameter chosen (using the `iRho` function in the `copula` package) so that
+the Spearman correlation would be the same as for the extreme tail dependence 
+datasets. The moderately right-tail dependent matrix and the symmetric tail dependence
+matrix were constructed similarly, but using an SC copula and a N copula,
+respectively.
+
+Once character evolution was simulated 100 times for each dependence structure, 
+we had 817 bivariate characters for each of 500 simulations. 
+<!--***DAN: Shya to insert as the rest of this paragraph a description of what 
+you did to each of these datasets to assess symmetry or asymmetry of tail dependence-->
+
+<!--***DAN: Shya, please insert a new para here with a description of results, including a new figure with 5 panels showing the 5 histograms you showed me.-->
+
+<!--***DAN: Dan does not understand the following two paragraphs yet, but is planning to
+speak to Mark about them-->
+Performing statistical tests for associations between continuous traits in different species
+    was a primary motivating example of phylogenetic comparative methods.
+For example, Felsenstein's method of phylogenetically independent contrasts 
+    [@Felsenstein1985] is currently the second most-cited paper in the history
+    of the journal the American Naturalist [@HueyGT2019].
+Frustratingly, the field still lacks a dependable procedure for dealing with branch lengths, 
+    which are a crucial input to the method.
+Ideally, the branch lengths used to correct for phylogenetic effects would represent
+    the expected amounts of change for the characters that are being analyzed.
+Because researchers almost never have a reliable method for providing such branch lengths,
+    most researchers rely on ultrametric trees -- those for which the branch
+    length can be treated as a proxy for the duration of the branch in time.
+Frequently these branch lengths are transformed to assess sensitivity to different 
+    assumptions about the degree of phylogenetic inertia displayed by the traits
+    under study [@Ives2018;@Harmon2018].
+Even if one were able to simply use a time-based set of branch lengths, assigning
+    dates to nodes in phylogenies is difficult.
+DNA sequence data can provide estimates of branch lengths, but these estimates
+    are dependent on the adequacy of models which correct sequences for multiple
+    substitutions occurring at the same location.
+Biases in estimating the evolutionary distance can affect down stream analyses [@Phillips2009].
+Additionally, changes
+    in the rate of molecular evolution make the estimation of dates difficult [@HeathM2014] even when branch lengths are accurately estimated.
+
+Without reliable branch length estimates, it is difficult to interpret the significance
+    of the magnitude of changes in traits across a tree.
+Developing tests of association based on copula structure could lead the way to more
+    robust methods for studying associations when we lack defensible estimates of
+    branch lengths.
+We note that our phylogenetic analyses here consisted merely of simulations to assess
+    whether interesting copula structure in a simple evolutionary process could 
+    leave a detectable signal on the data.
+Substantial work remains to be done before we have a copula-based method for 
+    analyzing data on a phylogenetic tree.
+
 <!--Following two chunks examine Moran-like mechanism with clayton copula and then generate plots with the simulated results-->

### 2019-04-26 SupportingInformation.Rmd
edits to Holder text in supp mat

@@
 Additionally, changes
-    in the rate of molecular evolution make the estimation of dates difficult [@HeathM2014] even when branch lengths are accurately estimated.
+    in the rate of molecular evolution make the estimation of dates difficult [@HeathM2014] even when branch lengths in time are accurately estimated.
 
@@
     of the magnitude of changes in traits across a tree.
-Developing tests of association based on copula structure could lead the way to more
+Developing tests of association based on copula structure may possibly lead the way to more
     robust methods for studying associations when we lack defensible estimates of
@@
     analyzing data on a phylogenetic tree. 
+But the nonparametric nature of approaches
+    based solely on data ranks, as some of our copula methods are, may be a promising avenue for 
+    avoiding inaccuracies that arise via the highly structured model assumptions implicit
+    in the method of phylogenetically independent contrasts and related methods.
 

### 2019-04-27 Paper.Rmd
results and description added for Holder's evol mechanism

@@
 i.e., why might we see right-tail dependence in evolutionary changes of these traits.
-See Appendix \ref{evol_mech} for additional thoughts on the potential importance of 
+See Appendix \ref{SM-evol_mech} for additional thoughts on the potential importance of 
 copulas for studies of character evolution when phylogeny branch lengths are not available.

### 2019-04-27 SupportingInformation.Rmd
results and description added for Holder's evol mechanism

@@
 Once character evolution was simulated 100 times for each dependence structure, 
-we had 817 bivariate characters for each of 500 simulations. 
-<!--***DAN: Shya to insert as the rest of this paragraph a description of what 
-you did to each of these datasets to assess symmetry or asymmetry of tail dependence-->
-
-<!--***DAN: Shya, please insert a new para here with a description of results, including a new figure with 5 panels showing the 5 histograms you showed me.-->
+we had 817 bivariate characters for each of 500 simulations. We generated 500 bivariate copulas 
+each made of 817 points. To assess the symmetry or asymmetry of tail dependence, we computed 
+our three non-parametric statistics ($\cor$, $\Ps$, $\Dsq$) on both extreme tips (lower, henceforth subscripted with `l` 
+and upper, henceforth subscripted with `u`) of each copula.
+To measure lower tail and upper tail statistics we used the bounds as $(l_b=0,u_b=0.2)$ and 
+$(l_b=0.8,u_b=1)$, respectively. Therefore, positive (or negative) difference of tail dependence statistics
+($\cor_l-\cor_u$, $\Ps_l - \Ps_u$ and $\Dsq_u - \Dsq_l$) would indicate asymmetry in dependence structure with
+stronger association in lower tail (or upper tail) of the bivariate copula.
+
+Five separate histograms were made with those 500 $\cor_l-\cor_u$ statistics, where each 100 
+sets were representative of 
+five different groups of simulations. These five groups of simulations used random draws from
+1) extreme left-tail, 2) moderate left-tail,
+3) symmetric tail (bivariate `N` copula), 4) moderate right-tail,
+and 5) extreme right-tail dependence copula, respectively, as two traits evolved through each branch.
+Five histograms each consisting 100 $\Ps_l - \Ps_u$ statistics and another five histograms each consisting
+100 $\Dsq_u - \Dsq_l$ statistics were produced in a similar way. Fig. \ref{fig_cop_evo} represents 
+three statistics for each five sets of dependence structure. If most of the statistics showed major
+occurrence on either side of zero line (vertical dashed line in Fig. \ref{fig_cop_evo}), then
+asymmetry in copula structure was confirmed. Fig. \ref{fig_cop_evo}(A-C, D-F) showing 
+stronger lower tail dependence with three statistics majorly distributed 
+on positive side were consistent with left tail-dependent copula as mentioned (1) and (2), respectively. 
+Similarly, Fig. \ref{fig_cop_evo}(J-L, M-O) showing 
+stronger upper tail dependence with three statistics majorly distributed 
+on negative side were consistent with right tail-dependent copula as mentioned (4) and (5), respectively.
+Evenly distributed tail dependence-difference statistics (see Fig. \ref{fig_cop_evo}G-I) 
+around zero line was indicative of symmetric dependence as emerged from bivariate normal copula.
 
@@
 ```
+
+<!--Following chunk is to examine the hypothesis for third probable mechanism : 
+asymmetric tail dependence in evolutionary changes in bivariate characters can give
+rise to similarly asymmetric tail dependence in character distributions across phylogeny tips-->
+
+```{r mechanism_evo_cop, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("explore_evo_cop.R"),mtime("CopulaFunctions_flexible.R"))}
+set.seed(seed)
+
+source("explore_evo_cop.R")
+```
+
+<!--adding plots showing results to confirm hypothesis behind third mechanism-->
+\begin{figure}[!h]
+\textbf{ \hspace{6.5 cm} (A) \hspace{2 cm} (B) \hspace{2 cm} (C)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeL_hist_taildep.pdf}\\
+\end{center}
+\textbf{ \hspace{6.5 cm} (D) \hspace{2 cm} (E) \hspace{2 cm} (F)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatL_hist_taildep.pdf}\\
+\end{center}
+\textbf{ \hspace{6.5 cm} (G) \hspace{2 cm} (H) \hspace{2 cm} (I)} \\
+\vspace{-0.7 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/BivNormal_hist_taildep.pdf}\\
+\end{center}
+\textbf{ \hspace{6.5 cm} (J) \hspace{2 cm} (K) \hspace{2 cm} (L)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatR_hist_taildep.pdf}\\
+\end{center}
+\textbf{ \hspace{6.5 cm} (M) \hspace{2 cm} (N) \hspace{2 cm} (O)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeR_hist_taildep.pdf}\\
+\caption[Non-parametric statistics to check probable mechanism behind asymmetric tail dependence in character distributions across phylogeny tips]{Three measures of asymmetric tail dependence using non-parametric statistics ($\cor,\Ps$ and $\Dsq$)
+on lower ($l_b=0,u_b=0.2$) and upper ($l_b=0.8,u_b=1$) tips confirms our hypothesis that copula structure in evolutionary changes in bivariate characters can give rise to similar tail dependent copula in character distributions across phylogeny tips (for details see Appendix \ref{evol_mech}).
+Stronger lower tail dependence (distributed more on positive values) was found in bivariate traits when (A-C) extreme left tail and (D-F) moderately left tail dependent copula was used in evolutionary shifts through each branch.
+Similarly, non-parametric statistics showed symmetric tail dependence for (G-I) normal copula and stronger upper tail dependence for (J-L) moderately right tail and (M-O) extreme right tail dependent copula, respectively.\label{fig_cop_evo}}
+\end{center}
+\end{figure}
 

### 2019-04-29 SupportingInformation.Rmd
minor edits in evo-cop histogram plot as instructed by DCR

@@
 Stronger lower tail dependence (distributed more on positive values) was found in bivariate traits when (A-C) extreme left tail and (D-F) moderately left tail dependent copula was used in evolutionary shifts through each branch.
-Similarly, non-parametric statistics showed symmetric tail dependence for (G-I) normal copula and stronger upper tail dependence for (J-L) moderately right tail and (M-O) extreme right tail dependent copula, respectively.\label{fig_cop_evo}}
+Similarly, non-parametric statistics showed symmetric tail dependence for (G-I) normal copula and stronger upper tail dependence for (J-L) moderately right tail and (M-O) extreme right tail dependent copula, respectively.
+The number $(n)$ in each subfigure designates how many statistics out of $100$ is equal or less than $0$, thus if $n$ is substantially lower (or higher) than $50$, it would indicate asymmetric left (or right) tail-dependence in evolved traits-copula.
+\label{fig_cop_evo}}
 \end{center}

### 2019-05-01 SupportingInformation.Rmd
edits, mostly for new holder materials

--- 2019-04-26 SupportingInformation.Rmd
@@
 datasets. The moderately right-tail dependent matrix and the symmetric tail dependence
-matrix were constructed similarly, but using an SC copula and a N copula,
+matrix were constructed similarly, but using an SC copula and an N copula,
 respectively.
@@
 Once character evolution was simulated 100 times for each dependence structure, 
-we had 817 bivariate characters for each of 500 simulations. 
-<!--***DAN: Shya to insert as the rest of this paragraph a description of what 
-you did to each of these datasets to assess symmetry or asymmetry of tail dependence-->
-
-<!--***DAN: Shya, please insert a new para here with a description of results, including a new figure with 5 panels showing the 5 histograms you showed me.-->
-
-<!--***DAN: Dan does not understand the following two paragraphs yet, but is planning to
-speak to Mark about them-->
+we had 817 bivariate characters for each of 500 simulations. We thus analyzed 500 corresponding 
+bivariate copulas (one for each simulation), each comprised of 817 points. 
+To assess the symmetry or asymmetry of tail dependence for each copula, we computed 
+the three non-parametric statistics $\cor_{0,0.2}-\cor{0.8,1}$, $\Ps_{0,0.2}-\Ps{0.8,1}$,
+and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ for each copula. For each statistic and for each
+dependence structure, we made a histogram (Fig. \ref{fig_cop_evo}). Values of these statistics less than 0
+indicate lower-tail dependence in characters of extant species, whereas values greater
+than 0 indicate upper-tail dependence. Asymmetry of tail dependence in evolutionary changes to characters
+was associated with similar asymmetry of tail dependence in extant characters.
+
 Performing statistical tests for associations between continuous traits in different species
@@
     whether interesting copula structure in a simple evolutionary process could 
-    leave a detectable signal on the data.
+    leave a detectable signal on the trait data for extant species.
 Substantial work remains to be done before we have a copula-based method for 
@@
 ```
+
+<!--Following chunk is to examine the hypothesis for third probable mechanism : 
+asymmetric tail dependence in evolutionary changes in bivariate characters can give
+rise to similarly asymmetric tail dependence in character distributions across phylogeny tips-->
+
+```{r mechanism_evo_cop, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("explore_evo_cop.R"),mtime("CopulaFunctions_flexible.R"))}
+set.seed(seed)
+
+source("explore_evo_cop.R")
+```
+
+<!--adding plots showing results to confirm hypothesis behind third mechanism-->
+\begin{figure}[!h]
+\textbf{ \hspace{6.5 cm} (A) \hspace{2 cm} (B) \hspace{2 cm} (C)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeL_hist_taildep.pdf}\\
+\end{center}
+\textbf{ \hspace{6.5 cm} (D) \hspace{2 cm} (E) \hspace{2 cm} (F)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatL_hist_taildep.pdf}\\
+\end{center}
+\textbf{ \hspace{6.5 cm} (G) \hspace{2 cm} (H) \hspace{2 cm} (I)} \\
+\vspace{-0.7 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/BivNormal_hist_taildep.pdf}\\
+\end{center}
+\textbf{ \hspace{6.5 cm} (J) \hspace{2 cm} (K) \hspace{2 cm} (L)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatR_hist_taildep.pdf}\\
+\end{center}
+\textbf{ \hspace{6.5 cm} (M) \hspace{2 cm} (N) \hspace{2 cm} (O)} \\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeR_hist_taildep.pdf}\\
+\caption[Non-parametric statistics on a mechanism of tail dependence in character distributions]{Three 
+measures of asymmetry of tail dependence between two characters, across extant species ($\cor_{0,0.2}-\cor_{0.8,1}$ (A, D, G, J, M), 
+$\Ps_{0,0.2}-\Ps_{0.8,1}$ (B, E, H, K, N), and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ (C, F, I, L, O)) were 
+computed for each of 100 simulations for each of five types of tail dependence between 
+evolutionary changes in the characters (extreme left-tail dependence (A-C), moderate left-tail dependence (D-F), 
+symmetry of tail dependence (G-I), moderate right-tail dependence (J-L), and extreme right-tail dependence 
+(M-O)). Asymmetry of tail dependence in evolutionary changes in characters tended to beget similar
+asymmetry in the characters across extant species. For additional details see 
+Appendix \ref{evol_mech}.\label{fig_cop_evo}}
+\end{center}
+\end{figure}
 

### 2019-05-01 SupportingInformation.Rmd

@@
 \caption[Non-parametric statistics on a mechanism of tail dependence in character distributions]{Three 
-measures of asymmetry of tail dependence between two characters, across extant species ($\cor_{0,0.2}-\cor_{0.8,1}$ (A, D, G, J, M), 
-$\Ps_{0,0.2}-\Ps_{0.8,1}$ (B, E, H, K, N), and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ (C, F, I, L, O)) were 
-computed for each of 100 simulations for each of five types of tail dependence between 
+measures of asymmetry of tail dependence ($\cor_{0,0.2}-\cor_{0.8,1}$ (A, D, G, J, M), 
+$\Ps_{0,0.2}-\Ps_{0.8,1}$ (B, E, H, K, N), and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ (C, F, I, L, O)) between two characters, across extant species,  were 
+computed for each of 100 simulations of mammalian evolution for each of five types of tail dependence between 
 evolutionary changes in the characters (extreme left-tail dependence (A-C), moderate left-tail dependence (D-F), 
@@
 asymmetry in the characters across extant species. For additional details see 
-Appendix \ref{evol_mech}.\label{fig_cop_evo}}
+Appendix \ref{evol_mech}.
+The number in the upper right of each panel indicates how many statistics out of $100$ were less than or equal to $0$.
+Values substantially lower (respectively, higher) than $50$ therefore indicate greater left (respectively, right) tail-dependence between characters.\label{fig_cop_evo}}
 \end{center}

### 2019-05-01 SupportingInformation.Rmd
typo fixed and plot label modified

@@
 To assess the symmetry or asymmetry of tail dependence for each copula, we computed 
-the three non-parametric statistics $\cor_{0,0.2}-\cor{0.8,1}$, $\Ps_{0,0.2}-\Ps{0.8,1}$,
+the three non-parametric statistics $\cor_{0,0.2}-\cor_{0.8,1}$, $\Ps_{0,0.2}-\Ps_{0.8,1}$,
 and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ for each copula. For each statistic and for each
-dependence structure, we made a histogram (Fig. \ref{fig_cop_evo}). Values of these statistics less than 0
-indicate lower-tail dependence in characters of extant species, whereas values greater
+dependence structure, we made a histogram (Fig. \ref{fig_cop_evo}). Values of these statistics greater than 0
+indicate lower-tail dependence in characters of extant species, whereas values less
 than 0 indicate upper-tail dependence. Asymmetry of tail dependence in evolutionary changes to characters
@@
 Appendix \ref{evol_mech}.
-The number in the upper right of each panel indicates how many statistics out of $100$ were less than or equal to $0$.
+The number above each panel indicates how many statistics out of $100$ were less than or equal to $0$.
 Values substantially lower (respectively, higher) than $50$ therefore indicate greater left (respectively, right) tail-dependence between characters.\label{fig_cop_evo}}

### 2019-05-01 explore_evo_cop.R

@@ -147,21 +147,24 @@ op<-par(mfrow=c(1,3),mar=c(5,5,2,1),mgp=c(3.5,1,0))
 res<-ans$CorlmCoru_list
 mres<-round(median(res),4)
 n<-sum(res<=0)
-hist(res,border=F,col="grey",breaks = 10,xlab=expression(cor[l]-cor[u]),cex.lab=2,cex.axis=2,main=paste("n = ",n,sep=""),cex.main=1.5)
+hist(res,border=F,col="grey",breaks = 10,xlab=expression(paste("cor" ["0,0.2"]-"cor"["0.8,1"])),
+     cex.lab=2,cex.axis=2,main=paste("n = ",n,sep=""),cex.main=1.5)
      #main=paste("median=",mres,sep=""))
 abline(v=0,lty="dashed")
 
 res<-ans$PlmPu_list
 mres<-round(median(res),4)
 n<-sum(res<=0)
-hist(res,border=F,col="grey",breaks = 10,xlab=expression(P[l]-P[u]),cex.lab=2,cex.axis=2,main=paste("n = ",n,sep=""),cex.main=1.5)
+hist(res,border=F,col="grey",breaks = 10,xlab=expression(paste("P" ["0,0.2"]-"P"["0.8,1"])),
+     cex.lab=2,cex.axis=2,main=paste("n = ",n,sep=""),cex.main=1.5)
 #     main=paste("median=",mres,sep=""))
 abline(v=0,lty="dashed")
 
 res<-ans$D2umD2l_list
 mres<-round(median(res),4)
 n<-sum(res<=0)
-hist(res,border=F,col="grey",breaks = 10,xlab=expression(D[u]^2 - D[l]^2),cex.lab=2,cex.axis=2,main=paste("n = ",n,sep=""),cex.main=1.5)
+hist(res,border=F,col="grey",breaks = 10,xlab=expression(paste("D" ["0.8,1"]^2-"D"["0,0.2"]^2)),
+     cex.lab=2,cex.axis=2,main=paste("n = ",n,sep=""),cex.main=1.5)
 #     main=paste("median=",mres,sep=""))
 abline(v=0,lty="dashed")
 
@@ -181,21 +184,24 @@ op<-par(mfrow=c(1,3),mar=c(5,5,2,1),mgp=c(3.5,1,0))
 res<-ans$CorlmCoru_list
 mres<-round(median(res),4)
 n<-sum(res<=0)
[…]

### 2019-05-02 Paper.Rmd
edits, mostly for new holder materials
various edits, getting ready for submission

@@
 title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
-author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, Daniel C. Reuman"
+author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
 date: ""
@@
 Reid: School of Biological & Marine Sciences, University of Plymouth, Drake Circus, Plymouth PL4 8AA, United Kingdom
+
+Bever: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
 
@@
 Datasets included environmental, species-trait, phenological, 
-population, community, and ecosystem functioning data. Copula structure describing 
+population, community, and ecosystem functioning data (Table \ref{tab_data_info}). Copula structure describing 
 the relationship between atmospheric weather variables such as rainfall or wind speed 
@@
 examined environmental variables from the soil instead, using the Rapid Carbon 
-Assessment database [[RaCA](https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/survey/?cid=nrcs142p2_054164); @Wills2014].
+Assessment database [RaCA; @Wills2014].
 <!--***DAN: Add Wills and Loecke pending reference if published in time--> 
@@
 locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
-Measurments were from the central pedon of a soil core, to 1 m depth.<!--DAN: we should ask Terry to explain pedon/ R horizon-->
-<!--***DAN: Terry to check the wording and accuracy here-->
-Species-trait data were average species basal metabolic rate (KJ per hour) and 
+Measurments were from the central pedon of a soil core, to 1 m depth.
+Species-trait data were average species basal metabolic rate (BMR, KJ per hour) and 
 body mass (grams) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
@@
 but we excluded one of the 11 sites they used in order to only use sites with at least 30 
-years data.  Our second population dataset comprised average annual abundance estimates 
+years data.  Our second population dataset comprised average annual plankton abundance estimates 
 for 14 locations (Fig. \ref{SM-fig_plankton_map})
-in the North Sea and British seas for 22 plankton taxa (Table \ref{SM-tab_plankton_info}) for the 
+in the North Sea and British seas for 22 taxa (Table \ref{SM-tab_plankton_info}) for the 
 56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. These data 
-were obtained from the Continuous Plankton Recorder (CPR) dataset now operated and maintained by 
-the Marine Biological Association. Data pre-processing was similar but not identical to that 
+were obtained from the Continuous Plankton Recorder (CPR) dataset, now operated and maintained by 
+the Marine Biological Association of the United Kingdom. Data pre-processing was similar but not identical to that 
 of @Sheppard2017; see Appendix \ref{SM-Data_details} for details.
-Community-level data obtained from the Cedar Creek Ecosystem Science Reserve were <!--DAN : check this rephrasing-->
-plant aboveground biomass and plant percent cover data (we used this percent cover data to calculate 
-Shannon's diversity index) for `r nrow(d_cc[[5]])` plots, each 9m by 9m, as sampled in 
-the year 2000 [@Tilman2001; @Tilman2006] <!--actually they collected data in each year of 1996,1997,1998,1999,2000,2007-->. 
+Community-level data, obtained from the Cedar Creek Ecosystem Science Reserve, were 
+plant aboveground biomass and Shannon's diversity index for `r nrow(d_cc[[5]])` plots, each 9m by 9m, as sampled in 
+the years 1996-2000 and 2007, each year analyzed separately [@Tilman2001; @Tilman2006]. 
+Shannon's index was computed from plant species percent cover data.  
 Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
@@
 copula structure can be studied for each pair of locations. Likewise the ecosystem 
-functioning (methane flux) dataset was  multivariate in this sense (see \ref{tab_data_info}).
+functioning (methane flux) dataset was  multivariate in this sense (Table \ref{tab_data_info}).
 
@@
 colnames(tab_data_info)<-c("Data","Category","No. of measurements we used", "References")
-tab_data_info$Data<-c("Soil C and N","Bird masses and BMR", "Mammal masses and BMR", "Cedar Creek data", 
-                      "Green spruce aphid abundance", "Leaf-curling plum aphid first flight","Ceratium furca abundance","Methane flux")
+tab_data_info$Data<-c("Soil C and N","Bird body masses and BMR", "Mammal body masses and BMR", "Cedar Creek data", 
+                      "Green spruce and other aphid species abundances", "Leaf-curling plum and other aphid first flight dates","Ceratium furca and other plankton taxa abundances","Methane flux")
 tab_data_info$Category<-c("Environmental","Species-trait","Species-trait","Community level",
-                          "Population","Phenological","Population","Ecosystem-functioning")
+                          "Population","Phenological","Population","Ecosystem functioning")
 tab_data_info$`No. of measurements we used`<-c(paste(nrow(d_soilCN)," locations",sep=""),
@@
                                                "13 locations with atleast 50 dates of data")
-tab_data_info$References<-c("US Rapid Carbon Assessment database (RaCA), Wills et al. (2014)", "McNab 2009","McNab 2008",
-                            "Cedar Creek Ecosystem Science Reserve [biodiversity experiment e120]",
+tab_data_info$References<-c("US Rapid Carbon Assessment database (RaCA), Wills et al. (2014)", "McNab (2009)","McNab (2008)",
+                            "Cedar Creek Ecosystem Science Reserve, biodiversity experiment e120",
                             "Rothamsted Insect Survey","Rothamsted Insect Survey","Continuous Plankton Recorder Survey",
@@
              format="latex", align="l",
-             caption = "Summary table for data we used, bolds are the multivariate data, rest are bivariates. \\label{tab_data_info}", 
+             caption = "Summary table for the data we used.
+Bolds entries are the multivariate data, the rest are bivariate datasets (see Data section).
+Basal metabolic rate=BMR \\label{tab_data_info}",
              booktabs = T,linesep = "\\addlinespace") %>%
@@
 We simulated bivariate character evolution on an estimate of 
-phylogeny for $817$ mammal species taken from @GenoudIM2018.
+the phylogeny, taken from @GenoudIM2018, of $817$ mammal species.
 The root character state and the change 
@@
 bivariate draws from distributions showing one of five distinct types of copula 
-structure: 1) extreme left-tail dependence, 2) moderate left-tail dependence,
+structures: 1) extreme left-tail dependence, 2) moderate left-tail dependence,
 3) symmetric tail dependence, 4) moderate right-tail dependence, and 5) extreme
@@
 All marginal distributions were standard normal, and distributions all had Spearman 
-correlation $0.875$, so simulations assess the impact of copula structure of character
-changes on evolutionary outcomes, controlling for marginals and correlations of character
-changes. For each of the five copula structures, mammalian character
+correlation $0.875$ between their two components, so our simulations assess 
+the impact only of copula structure in evolutionary character
+changes on evolutionary outcomes. For each of the five copula structures, mammalian character
 evolution was simulated 100 times. For each simulation, symmetry of tail dependence 
-of the two characters was assessed across the phylogeny tips.
+of the two characters across the phylogeny tips was assessed.
 We hypothesized that cases 1 and 2 above would yield stronger left- than right-tail 
@@
 Our simulations of character evolution suggest the hypothesis that evolutionary
-changes in bird and mammal BMR and body size may exhibit right-tail dependence. This is
-a hypothesis only, because the tail dependence of Fig. \ref{fig_biv_multi_raw_cop}F,G
+changes through time in bird and mammal BMR and body size may exhibit right-tail dependence. This is
+a hypothesis only, because the right-tail dependence of Fig. \ref{fig_biv_multi_raw_cop}F,G
 could have come about in another, unknown way: our simulations show that tail dependence 
 in evolutionary changes is sufficient, but may not be necessary, to produce tail
-dependence in characters of extant species. Even if the hypothesized mechanism
+dependence in characters of extant species. For instance, as discussed below and in
+Appendix \ref{SM-sect:missingdata} (see also Fig. \ref{SM-fig:missingdata}), 
+systematically missing data can produce tail dependence and may
+have influenced results for the BMR versus body mass datasets. Even if the hypothesized 
+evolutionary mechanism
 is correct, our results simply replace one question, i.e., why do we see right-tail
 dependence between BMR and body mass for birds and mammals, with another question,
-i.e., why might we see right-tail dependence in evolutionary changes of these traits.
+i.e., why might we see right-tail dependence in evolutionary changes of these traits?
 See Appendix \ref{SM-evol_mech} for additional thoughts on the potential importance of 
@@
 
-<!-- We have only given three mechanisms, but there are likely to be others, and discoverig
-additional mechanisms of copula structure is certain worthy of future work. -->
-Other mechanisms that we have not considered likely also operate to produce copula structure
+<!--Discuss stuff here on tradeoff and life history that Bever told me about-->
+The relationships we studied between BMR and body mass relate
+to a trade-off between mass-specific BMR (i.e., BMR per unit body mass) and body
+mass itself. Trade-offs between life history characteristics are a focal topic in 
+evolutionary ecology. There are reasons to believe copulas interrelate with 
+life-history theory in additional ways beyond what we could demonstrate 
+with the data we used.
+It is well known that energy allocation to a life function, X (e.g., reproduction)
+will reduce the energy that can be allocated to other functions, $Y_1$, $Y_2$, $Y_3$ 
+(e.g., growth, predation avoidance; this is sometimes called the principle of 
+allocation).
+But it is equally well known that X can trade-off against any or all of 
+the $Y_i$. Therefore, for large values of X which approach absolute 
+limitations, there may be a strong 
+association between X and $Y_1$. Whereas for small values of X there may be
+little association - resources not allocated to X can instead be allocated 
+to $Y_1$ or $Y_2$ or to some combination. This constitutes 
+asymmetric tail dependence between X and $Y_1$. For instance, \cite{Winemiller1992}
+described a three-way trade-off in North American fishes between age of reproductive
+maturity, juvenile survivorship, and fecundity. The trade-off should
+produce a tight association between age of maturity and fecundity for 
+fishes with low age of maturity, but little association between those
+variables for later-maturing
+fishes because those fishes may invest much or little in juvenile survival.
+Although copulas may be very useful for studying multi-dimensional trade-offs, 
+applications will 
+require careful attention to the possible consequences of biased sampling:
+if the degree of completeness of a dataset is associated with one or
+more of the quantities being studied then statistical artefacts
+may bias conclusions of copula analyses (Appendix \ref{SM-sect:missingdata}, 
+Fig. \ref{SM-fig:missingdata}).
+
+<!-- We have only given a few mechanisms, but there are likely to be others, and discoverig
+additional mechanisms of copula structure is certainly worthy of future work. -->
+Other mechanisms that we have not considered probably also operate to produce copula structure
 in ecological and evolutionary datasets, and should be considered in future work. For instance, 
@@
 Not only did the models we used to explore mechanisms not consider measurement error, 
-they were extremely simple in other respects, as well, not including phenomena such as 
+they were intentionally simple in other respects, as well, not including phenomena such as 
 delayed density dependence, dispersal, stage structure in population dynamics, and trophic
 interactions.
-We hope by enumerating three potential mechanisms we help inspire additional research on the
-potentially numerous mechanisms of copula structure that may apply in diverse datasets.
+We hope by enumerating a few potential mechanisms of copula structure in data 
+we help inspire additional research on the
+potentially numerous mechanisms that may operate in diverse datasets.
 
@@
 We thank the many contributors to the large datasets we used; D. Stevens and P. Verrier for data 
-extraction; and Joel E. Cohen, David Tilman, Lauren Hallet, Jim Bever, Jonathan Walter, Thomas 
+extraction; and Joel E. Cohen, David Tilman, Lauren Hallet, Jonathan Walter, Thomas 
 Anderson, and Lei Zhao for helpful suggestions. We thank James Bell of the Rothamsted Insect Survey 

### 2019-05-02 SupportingInformation.Rmd
various edits, getting ready for submission

--- 2019-05-01 SupportingInformation.Rmd
@@
 title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
-author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, Daniel C. Reuman"
+author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
 geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"
@@
 
-<!--***DAN: This is a dumping ground at the moment, but needs to be organized into a section that gives all the necessary details about the data-->
-
-<!--additional details about Cedar creek data, etc.-->
-Above-ground biomass and percent cover data were obtained from long-term (? or short-term)
-biodiversity experiment (E120) conducted in Cedar Creek Ecosystem Science Reserve by Tilman *et al*.
-In 1994, 168 plots (each 9 m $\times$ 9 m) of 39, 35, 29, 30 and 35 replicates,
+\textbf{Population data, plankton abundances.} 
+The CPR plankton data presented in @Sheppard2017 was aggregated in the following way. Fifty 
+$2^\circ$ by $2^\circ$ location boxes were examined between $-10^\circ$ and $10^\circ$ longitude 
+and between $50^\circ$ and $60^\circ$ latitude (Fig. \ref{fig_plankton_map}). For each species, for each box, 
+for each year, for each month, the average of any available CPR samples was taken to give monthly 
+time series of population densities with some missing values. Missing monthly values were then 
+replaced with the median value for that species for that location for that month. The data for each 
+species for each box for each year was then averaged over all twelve months, and 26 usable annual 
+time series were identified for wavelet analysis. The CPR plankton data used here is aggregated in 
+the same way, except that for any species, box and year for which more than 4 months data were missing, 
+missing values were not replaced with medians and those species, box, year combinations were not used.
+We used only data from the 14 locations which had at least 45 years of annual data subject to this 
+restriction, shown by blue markers in Fig. \ref{fig_plankton_map}. The CPR data are described in
+great detail elsewhere, e.g., by  @Batten2003 and @Raitsos2014.
+
+\textbf{Community-level data, Cedar Creek.} Above-ground biomass and percent cover data were 
+obtained from biodiversity experiment E120 conducted at the Cedar Creek Ecosystem Science Reserve 
+by Tilman *et al*. In 1994, 168 plots (each 9 m $\times$ 9 m) of 39, 35, 29, 30 and 35 replicates,
 were seeded with 1, 2, 4, 8, or 16 perennials, respectively, chosen randomly from a pool of 
-18 species (4 species, each, of C4 grasses, C3 grasses, legumes, non-legume forbs; 2 species of woody
+18 species (4 species, each, of C4 grasses, C3 grasses, legumes, non-legume forbs; and 2 species of woody
 plants). All plots received, in total, 10 $gm^{-2}$ of pure live seed in May 1994 and 5 $gm^{-2}$ in May, 1995 
 with seed mass divided equally among species. Treatments were maintained by weeding manually 3-4 times a year 
-and sometimes selective herbicides were used to set up the monocultures either with only grasses or with only forbs.
-We downloaded the data on July, 2018 from Cedar Creek website \url{http://www.cedarcreek.umn.edu/research/experiments/e120} and 
-analyzed those [data](https://github.com/sghosh89/BIVAN/tree/master/Data/Ceder_creek_e120) for 6 common years 
-(1996-2000, 2007) when both biomass and percent cover data were available. Each plots were annually sampled by clipping 
-different locations each year. For each available year, we computed above-ground biomass ($gm^{-2}$) 
-and Shannon's biodiversity index (using percent cover data) averaged over the sampled strips/ quadrats taken from each 
+and sometimes selective herbicides were used. We downloaded the data on July, 2018 from the Cedar Creek website \url{http://www.cedarcreek.umn.edu/research/experiments/e120} and 
+analyzed those data for 6 years 
+(1996-2000, 2007) for which both biomass and percent cover data were available. Plots were annually 
+sampled by clipping 
+different locations each year. For each available year, we analyzed above-ground biomass ($gm^{-2}$) 
+and Shannon's biodiversity index (computed using percent cover data) averaged over the 
+sampled strips/quadrats taken from each 
 of 168 plots.
 
-<!--we need to add aphid and plankton data details, Lawrence will give those-->
-
-<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
-unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
-location in the ocean.
-Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979].
-All the samples available for a given variable in a given area in a given month of a given year were
-averaged, producing a monthly time series for each year.
-Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).-->
-
- [@Sheppard2017] converted monthly values to yearly values by averaging over all months.
-But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available).
-We used data from 14 locations which had at least 45 years time series data (see Figure \ref{fig_plankton_map}).
-
- The Great Miami Wetland Mitigation Bank  was at $36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio.
-Additional information on the history of the site and soil types is detailed in @Jarecke2016.
-Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016.
-**See and** @smyth2019using, **pending for
-additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had at least 50 dates data.
+\textbf{Ecosystem functioning data on methane flux.} The Great Miami Wetland Mitigation Bank  was at 
+$36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W, in Trotwood, Montgomery County, Ohio. Additional information 
+on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method 
+was used [@Holland1999] to measure $\text{CH}_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to 
+seasonal wetland) at daily to weekly interval from September 2015 to September 2016. See @smyth2019using for
+additional details on greenhouse gas flux methods for this site. We considered 13 locations for which data were
+gathered for at least 50 dates.
 
@@
 and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ for each copula. For each statistic and for each
-dependence structure, we made a histogram (Fig. \ref{fig_cop_evo}). Values of these statistics less than 0
-indicate lower-tail dependence in characters of extant species, whereas values greater
-than 0 indicate upper-tail dependence. Asymmetry of tail dependence in evolutionary changes to characters
+dependence structure, we made a histogram of values 
+(Fig. \ref{fig_cop_evo}). Asymmetry of tail dependence in evolutionary changes to characters
 was associated with similar asymmetry of tail dependence in extant characters.
@@
     in the method of phylogenetically independent contrasts and related methods.
+
+Additionally, character evolution was simulated using one random draw from the relevant matrix
+per phylogeny branch. This was principally because branch lengths are often hugely
+uncertain. If branch lengths were well know, alternative simulation strategies may
+include selecting one draw from the matrix of evolutionary changes per unit time, 
+or selecting one draw for the whole branch but rescaling variance according to the 
+length of the branch. These choices amount to the same thing for normal copula
+structure, but not otherwise. Modelling choices such as these may have influenced
+our results. Additional research also seems warranted testing the realism 
+of our hypothesized mechanism and simulations.
 
@@
 rise to similarly asymmetric tail dependence in character distributions across phylogeny tips-->
-
 ```{r mechanism_evo_cop, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("explore_evo_cop.R"),mtime("CopulaFunctions_flexible.R"))}
@@
 $\Ps_{0,0.2}-\Ps_{0.8,1}$ (B, E, H, K, N), and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ (C, F, I, L, O)) between two characters, across extant species,  were 
-computed for each of 100 simulations of mammalian evolution for each of five types of tail dependence between 
+computed for each of 100 simulations of mammalian character evolution for each of five types of tail dependence between 
 evolutionary changes in the characters (extreme left-tail dependence (A-C), moderate left-tail dependence (D-F), 
@@
 
+# A means by which missing data can influence perceived copula structure\label{sect:missingdata}
+
+```{r missing_data, echo=FALSE, results="hide", cache=TRUE}
+set.seed(101)
+
+#generate data from a normal copula
+ncopparm<-0.8
+ncop<-normalCopula(ncopparm,2)
+d<-rCopula(5000,ncop)
+
+#censor it - randomly delete many of the points that are in the upper right 
+#triangle of the copula plot
+inds<-which(d[,2]>-d[,1]+1)
+fracdelete<-4/5
+inds<-inds[sample(1:length(inds),size=round(fracdelete*length(inds)),replace=FALSE)]
+dc<-d[-inds,]
+
+#reperform the ranking
+dc<-pobs(dc)
+
+#now make a plot
+panwd<-2
+panht<-panwd
+xaxht<-.6
+yaxwd<-xaxht
+gap<-0.1
+totwd<-yaxwd+2*panwd+2*gap
+totht<-xaxht+panht+gap
+pdf(file="./Results/MissingDataInfluencesCopula.pdf",width=totwd,height=totht)
+par(fig=c(yaxwd/totwd, #panel 1 - the original data
+          (yaxwd+panwd)/totwd,
+          (xaxht)/totht,
+          (xaxht+panht)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
+plot(d[,1],d[,2],pch=20,cex=.5,xlim=c(0,1),ylim=c(0,1))
+text(0,1,"A",adj=c(0,1),cex=1.5)
+mtext("x",1,1.5,cex=2)
+mtext("y",2,1.5,cex=2)
+par(fig=c((yaxwd+panwd+gap)/totwd, #panel 2 - the censored data, re-ranked
+          (yaxwd+2*panwd+gap)/totwd,
+          (xaxht)/totht,
+          (xaxht+panht)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=TRUE)
+plot(dc[,1],dc[,2],pch=20,cex=.5,xlim=c(0,1),ylim=c(0,1),yaxt="n")
+axis(2,labels=FALSE)
+text(0,1,"B",adj=c(0,1),cex=1.5)
+mtext("x",1,1.5,cex=2)
+dev.off()
+```
+
+To illustrate the potential for missing data to influence perceived
+copula structure, we first generated 5000 points from a bivariate N copula
+with parameter `r ncopparm` (Fig. \ref{fig:missingdata}A), and then 
+randomly deleted `r 100*fracdelete`% of the points that occurred in the
+upper-right triangle of the copula plot. Data were then re-ranked as
+described in the Introduction to produce a new copula plot based
+only on the censored dataset (Fig. \ref{fig:missingdata}B). Copula
+structure in the tails was affected. Although real datasets have 
+typically not been censored in the precise way we censored our
+artificial data, it may be common for species characters to be 
+measured on a subset of the extant species from the clade of interest,
+and the probability of a species having been included in the dataset
+may depend on one or both of the character values. 
+For instance, if larger or smaller bird or mammal species were more likely
+to have been measured for body mass or BMR, such accidental 
+censoring of the data could have influenced our measurements of tail 
+dependence. We have no reason to believe this is the case for our 
+BMR versus body mass datasets, but we admit the possibility. Any systematic 
+use of copulas to study life-history tradeoffs should take the possibility
+into account of artefacts such as described here.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=7in]{./Results/MissingDataInfluencesCopula.pdf}
+\caption[Missing data can influence perceived copula structure]{Missing
+data can influence perceived copula structure. Points were randomly 
+removed from the upper right portion of the copula dataset pictured in A, and
+then remaining values were re-ranked to make B. Thus removing data in a 
+non-random fashion can influence copula structure and tail dependence. See 
+Appendix \ref{sect:missingdata} for details. \label{fig:missingdata}}
+\end{center}
+\end{figure}
+
 \clearpage

### 2019-05-02 SupportingInformation.Rmd

@@
 To assess the symmetry or asymmetry of tail dependence for each copula, we computed 
-the three non-parametric statistics $\cor_{0,0.2}-\cor{0.8,1}$, $\Ps_{0,0.2}-\Ps{0.8,1}$,
+the three non-parametric statistics $\cor_{0,0.2}-\cor_{0.8,1}$, $\Ps_{0,0.2}-\Ps_{0.8,1}$,
 and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ for each copula. For each statistic and for each

### 2019-05-02 Paper.Rmd
minor edit

@@
 \newpage
-# Introduction \label{Intro}
+# Introduction\label{Intro}
 
@@
 to $Y_1$ or $Y_2$ or to some combination. This constitutes 
-asymmetric tail dependence between X and $Y_1$. For instance, \cite{Winemiller1992}
+asymmetric tail dependence between X and $Y_1$. For instance, @Winemiller1992
 described a three-way trade-off in North American fishes between age of reproductive

### 2019-05-02 Paper.Rmd
small edits

--- 2019-05-02 Paper.Rmd
@@
 Bolds entries are the multivariate data, the rest are bivariate datasets (see Data section).
-Basal metabolic rate=BMR \\label{tab_data_info}",
+Basal metabolic rate=BMR.
+\\label{tab_data_info}",
              booktabs = T,linesep = "\\addlinespace") %>%
@@
 and asymmetry of tail dependence. 
-Conceptually, a bivariate copula is essentially a probability density function (pdf)
+Conceptually, a bivariate copula is a probability density function (pdf)
 on the unit square with uniform maginals. 
-Strictly speaking, 
-the copula is actually defined as the bivariate 
-cdf of the distribution rather than as the pdf,
-but this makes little conceptual difference. 
+The copula is actually defined as the bivariate 
+cdf of the distribution rather than as the pdf itself,
+but this makes little difference. 
 The normal copula family is a one-parameter
@@
 We addressed Q1 principally via a model selection procedure in which the normal copula
-the alternative copulas listed above were fitted to 
+and the alternative copulas listed above were fitted to 
 our ecological datasets. The resulting fits of the various copula models 
@@
 to $0$.
-We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}).
+We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}). All code for this
+project is archived at \url{www.github.com/sghosh89/BIVAN}.
 
@@
 described a three-way trade-off in North American fishes between age of reproductive
-maturity, juvenile survivorship, and fecundity. The trade-off should
+maturity, juvenile survivorship, and fecundity. The trade-off should, in theory,
 produce a tight association between age of maturity and fecundity for 

### 2019-05-02 Paper.Rmd
editing for length

@@
 ---
-title: "The importance of a complete statistical description of dependence between variables for ecology and related fields"
+title: "The importance of a complete description of dependence between variables for ecology and related fields"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
@@
 
-*Affiliations:*
-
-Ghosh: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
-
-Sheppard: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
-
-Holder: Department of Ecology and Evolutionary Biology and Biodiversity Institute, University of Kansas, Lawrence, KS, 66045, USA
-
-Loecke: Environmental Studies Program and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA
-
-Reid: School of Biological & Marine Sciences, University of Plymouth, Drake Circus, Plymouth PL4 8AA, United Kingdom
-
-Bever: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
-
-Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA
-
-*Possible alternative titles:*
-
-- A complete statistical description of dependence between variables for ecology and related fields
-
-*Keywords:*
-
-*Type of Article*
-
-- Ideas and Perspectives
-
-*The number of words in the abstract*
-
-*The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends)*
-
-*The number of words in each text box*
-
-*The number of references*
-
-*The number of figures, tables, and text boxes*
-
-*The name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent*
+\noindent \emph{Affiliations:}
+
+\noindent Ghosh: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
+
+\noindent Sheppard: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
+
+\noindent Holder: Department of Ecology and Evolutionary Biology and Biodiversity Institute, University of Kansas, Lawrence, KS, 66045, USA
+
+\noindent Loecke: Environmental Studies Program and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA
+
+\noindent Reid: School of Biological & Marine Sciences, University of Plymouth, Drake Circus, Plymouth PL4 8AA, United Kingdom
+
+\noindent Bever: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
+
+\noindent Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA
+
+\noindent \emph{Possible alternative titles:}
+
+\noindent A complete statistical description of dependence between variables for ecology and related fields
+
+\noindent \emph{Keywords:}
+
+\noindent dependence, regression, correlation, copula, population, community
+
+\noindent \emph{Type of Article:}
+
+\noindent Ideas and Perspectives
+
+\noindent \emph{The number of words in the abstract:}
+
+\noindent 241
+
+\noindent \emph{The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends):}
+
+\noindent \emph{The number of words in each text box:}
+
+\noindent No text boxes
+
+\noindent \emph{The number of references:}
+
+\noindent \emph{The number of figures, tables, and text boxes:}
+
+\noindent 7 figures, 3 tables, 0 text boxes
+
+\noindent \emph{The name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent:}
+
+\noindent Daniel Reuman, 2101 Constant Ave, Lawrence, KS, 66047, reuman@ku.edu, 626 560 7084.
 
 \newpage
+
 # Abstract
@@
 All branches of ecology study relationships among environmental and biological variables.
-However, ubiquitously used approaches to studying such relationships, based on correlation and regression, provide only a small slice of the complex information content of the relationships.
-Other statistical tools exist that provide a complete description of relationships, based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, but not in ecology.
-We here examine the potential for copula approaches to improve our understanding of ecology.
+However, ubiquitously used approaches to studying such relationships, based on correlation and regression, provide only a small slice of the complex information contained in the relationships.
+Other statistical tools exist that provide a complete description of relationships between variables, based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, but not in ecology.
+We here explore the concepts that underpin copulas and examine the potential for those concepts to improve our understanding of ecology.
 We find that informative copula structure in dependences between variables is exceedingly common across all the environmental, species-trait, phenological, population, community, and ecosystem functioning dataset we considered.
@@
 Copula structure and tail dependence can help reveal causal relationships between variables.
-We describe two mechanisms by which observed copula structure and tail dependence can arise in ecological data: a Moran-like effect whereby dependence structures between environmental variables are inherited by ecological variables; and asymmetric or nonlinear influences of environments on ecological variables, such as under Liebig's law of the minimum.
+We describe mechanisms by which observed copula structure and tail dependence can arise in ecological data, including a Moran-like effect whereby dependence structures between environmental variables are inherited by ecological variables; and asymmetric or nonlinear influences of environments on ecological variables, such as under Liebig's law of the minimum.
 We describe consequences of copula structure, including impacts on extinction risk and the stability through time of ecosystem services.
-By documenting the importance of a complete statistical description of dependence, advancing conceptual frameworks, and demonstrating powerful methods, we aim to elevate copula approaches to widespread use in ecology.
+By documenting the importance of a complete description of dependence, advancing conceptual frameworks, and demonstrating a powerful approach, we aim to elevate copula reasoning to widespread use in ecology.
 
@@
 
-Very generally, all branches of ecology and related areas <!--DAN : Reid said " Is this text needed. 
-Ecology covers 
-all branches of the biological sciences?
-What do you mean by related areas? 
-You use related fields later.
-Needs clarifying."-->
+Very generally, all branches of ecology  
 study relationships among environmental and biological 
@@
 approaches to studying 
-such relationships are limited - such statistical
+such relationships are limited - such 
 approaches provide only a very partial description of a relationship between variables. 
@@
 which alters the Pearson, but not the Spearman correlation. Although correlation
-is only one of numerous metrics for 
-studying relationships between variables, other common metrics <!--DAN : add example--> also 
-provide only partial information.
+is only one way to 
+study relationships between variables, other common metrics also 
+provide partial information.
 
@@
 
-Other well-developed statistical tools exist, that are 
-applied widely in other fields (e.g. finance, neuroscience, bioinformatics etc.), 
-and provide a demonstrably complete 
-description of the dependence 
-between two or more variables [@joe2014_dependence; @nelsen2006_copula]. 
+Well-developed approaches exist, and are 
+applied widely in finance, neuroscience, bioinformatics and other fields, 
+that provide a description of the dependence 
+between two or more variables that can be mathematically proven 
+to be complete [@joe2014_dependence; @nelsen2006_copula]. 
 The goal of this paper is to illustrate the potential for 
-broader application of these
-tools in ecology and related fields. The tools are based on 
+broader application of the concepts behind these approaches 
+in ecology. The approaches are based on 
 the statistical concept of the *copula*, which we here 
@@
 variables), and the remaining information (which is solely about
-the dependence between the variables). From a data perspective, given 
+the dependence). From a data perspective, given 
 a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
 Fig. \ref{fig_cop_pedag1}A-B), the associated dependence
-or copula structure is revealed by the bivariate plot of $u_i$ versus $v_i$, where
+structure is revealed by the bivariate plot of $u_i$ versus $v_i$, where
 $u_i$ is the rank of $x_i$ in the set of the $x_j$ ($j=1,\ldots,n$), divided by
@@
 the underlying random variables uniformly distributed, isolating the 
-dependence structure only<!--DAN : check this-->.
-Mathematically, if $(X,Y)$ is a continuous bivariate random
+dependence structure only.
+Mathematically, if $(X,Y)$ is a continuous random
 variable and the cumulative distribution functions (cdfs) of $X$ and $Y$ are 
-$F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the dependence or copula structure
+$F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the dependence structure
 of $(X,Y)$ is revealed by the random variable $(F_X(X),F_Y(Y))$.
-This random variable takes values in the unit square, 
-$[0,1] \times [0,1]$, and has uniform marginal distributions. Probability density 
+This random variable takes values in $[0,1] \times [0,1]$, and has 
+uniform marginal distributions. Probability density 
 functions (pdfs) corresponding to the copulas associated 
@@
 See @nelsen2006_copula and @joe2014_dependence
-for general background. A theorem of A. @sklar1959_theorem
+for general background. A theorem of @sklar1959_theorem
 states that copulas give a unique and complete 
 representation of the dependence
-between variables, in the sense that the joint distribution of $(X,Y)$ 
-can be reconstructed from the 
-copula and the marginal distributions. 
-Copula theory is also well developed for
-multivariate distributions [@joe1997multivariate].
+between variables.
 
 A principal benefit of a copula approach is 
-that it can sensitively detect and model
-tail dependence and asymmetries of tail dependence. 
+that it can sensitively detect tail dependence and asymmetries of tail dependence. 
 Tail dependence is association between extreme values of variables. 
@@
 minimum. If, for instance, the growth of a plant depends on soil 
-nitrogen, N, as well
-as on other factors, a plot of plant growth rates versus soil nitrogen may be 
-expected to look more like Fig. \ref{fig_cop_pedag1}A or E than like 
+nitrogen, N, and other factors, a plot of plant growth rates versus N may 
+look more like Fig. \ref{fig_cop_pedag1}A or E than like 
 Fig. \ref{fig_cop_pedag1}B or F. Nitrogen 
 will control plant growth, producing a clear
-relationship, only when it is limiting, in the left portion of the N and 
-plant-growth-rate distributions. 
-Nitrogen will be less related or unrelated to growth rates in the right tails of 
-the distributions, as in Fig. \ref{fig_cop_pedag1}A and E. The portion
-of the distributions that dominate the association is visible in the copula
+relationship, only when it is limiting, in the left portions of the distributions. 
+The portion of the distributions that dominate the association is visible in the copula
 approach, but not via standard techniques.
 
-A second reason we may expect a copula-based approach to benefit 
+A second reason we may expect a copula approach to benefit 
 ecological research is that such an approach may help illuminate 
-the causal mechanisms by which variables are related. For instance, if two species
-X and Y are strong competitors, we may expect abundances of the two species
-across quadrats to be negatively related, as in 
+the causal mechanisms between variables. For instance, if two species
+X and Y are strong competitors, abundances of the two species
+across quadrats should be negatively related, as in 
 Fig. \ref{fig_cop_pedag1}I-J. If X is plotted on the horizontal axis
-and Y is plotted on the vertical, then the plot in Fig. \ref{fig_cop_pedag1}I 
-suggests X is the dominant 
-competitor: when X is abundant, Y is necessarily rare because it is suppressed 
-by X; whereas when X is rare, Y
+and Y is on the vertical, then Fig. \ref{fig_cop_pedag1}I 
+would suggest X is the dominant 
+competitor: when X is abundant, Y is necessarily rare because it is suppressed; 
+whereas when X is rare, Y
 may be abundant, or may also be relatively rare due to 
@@
 distributions are principally responsible for the association between variables, and this 
-detail may help support or refute alternative causal hypotheses about 
-the relationship between the variables.
+information can help support or refute alternative causal hypotheses.
 
@@
 Spatial dependence in an environmental variable 
-tends to beget dependence across space
+tends to beget spatial dependence 
 in fluctuations of populations or other ecosystem variables 
 (this is called \emph{spatial synchrony}) which are influenced by the environmental 
-variable. This process is called the \emph{Moran effect}. 
+variable. This is called the \emph{Moran effect}. 
 If copula structure is transmitted, 
@@
 from environmental to ecological variables, then we may expect important 
-copula structure to be a common feature of spatial synchrony of population, community, 
+copula structure to be a common feature of the spatial synchrony of population, community, 
 biogeochemical and other ecological variables.
@@
 fluxes are the sum of local fluxes; 
-and the total economic value of a fishery is the sum of catches from 
-local areas. We will explore how details, captured by copula methods, of the 
-nature of the dependence between local quantities can influence the spatial 
-mean or sum, and how this may influence human concerns and 
-higher organizational levels in 
-ecology. Illustrating the idea here, prior work demonstrated that the overall 
+and the total economic value of a fishery is the sum of local catches. 
+We will explore how details, captured by copulas, of the 
+dependence between local quantities can influence the spatial 
+mean or sum, and how this may influence higher organizational levels in 
+ecology and human concerns. Illustrating the idea here, prior work demonstrated that the overall 
 reliability of wind-generated electricity depends sensitively on details
@@
 pest species in different locations are all positively associated and also strongly
-related to each other in their right tails, then local outbreaks tend 
-to occur together, creating regional epidemics. If populations
-of a crop or other resource species are strongly 
-related in their left tails, then 
-episodes of local scarcity occur together, creating regional 
-shortages. On the other hand left-tail dependence in the pest species 
-or right-tail dependence in the crop, even if overall correlation were 
-unchanged, would have much more benign effects.
+related to each other in their right tails, then local outbreaks will tend 
+to occur together, creating regional epidemics. 
+On the other hand left-tail dependence in a pest species, even if overall 
+correlation were unchanged, would have much more benign effects.
 
 We approached our overall goal of exploring the potential of 
-copula approaches in ecology by addressing the 
-following specific questions. (Q1) Do datasets in ecology and related fields 
+copula approaches for ecology by addressing the 
+following specific questions. (Q1) Do datasets in ecology  
 have copula structure distinct from that of a Gaussian/normal distribution (here called 
@@
 non-normal copula structure? And (Q3) what are the consequences of non-normal 
-copula structure and tail dependence for our understanding of biology and for 
+copula structure and tail dependence for our understanding of ecology and for 
 applications? To address Q1, we analyze several datasets, including 
@@
 explore mechanisms. The seventh and eigth sections, which address Q3, use both
-the data of section two as well as models to explore a variety of consequences
-of copula structure and tail dependence for applications and for understanding
-ecology. The ninth section is the Discussion. 
-Copula approaches have been used to great benefit in neuroscience [@onken2009], 
+data and models to explore a variety of consequences. The ninth section is the Discussion. 
+Copula approaches have been used to great consequence in neuroscience [@onken2009], 
 bioinformatics 
 [@Kim2008], medical research [@Emura2016], direct study of environmental 
-variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], finance, and other fields. 
-To our knowledge, these approaches have never or seldom been used for ecological research. 
-We argue that benefits to ecology of using a complete statistical 
+variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000]. 
+To our knowledge, these concepts have never been used in ecology. 
+We argue that benefits to ecology of using a complete 
 desription of dependence will be substantial.
@@
 locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
-Measurments were from the central pedon of a soil core, to 1 m depth.
 Species-trait data were average species basal metabolic rate (BMR, KJ per hour) and 
-body mass (grams) data for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
+body mass (grams) for `r nrow(d_birdsBMR)`  species of birds [@Mcnab_2009] 
 and `r nrow(d_mammalsBMR)`  species of mammals [@Mcnab_2008].
@@
 population-level datasets was also derived from the RIS suction-trap data, and comprised 
-total counts of aphids trapped for the same locations, species, and years; these are 
-indicators of aphid migration counts. Data pre-processing for the aphid data was the same as 
-that of @Sheppard2016,
-but we excluded one of the 11 sites they used in order to only use sites with at least 30 
-years data.  Our second population dataset comprised average annual plankton abundance estimates 
+total counts of aphids trapped for the same locations, species, and years. 
+Our second population dataset comprised average annual plankton abundance estimates 
 for 14 locations (Fig. \ref{SM-fig_plankton_map})
 in the North Sea and British seas for 22 taxa (Table \ref{SM-tab_plankton_info}) for the 
-56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. These data 
-were obtained from the Continuous Plankton Recorder (CPR) dataset, now operated and maintained by 
-the Marine Biological Association of the United Kingdom. Data pre-processing was similar but not identical to that 
-of @Sheppard2017; see Appendix \ref{SM-Data_details} for details.
+56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. 
 Community-level data, obtained from the Cedar Creek Ecosystem Science Reserve, were 
@@
 from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, 
-Ohio [@Holland1999; @Jarecke2016; @smyth2019using]. Each included location was measured on at least 50 dates. 
-Sampling locations spanned a hydrologic gradient from upland to seasonal wetland.
-See Appendix \ref{SM-Data_details} for additional data details.
+Ohio [@Holland1999; @Jarecke2016; @smyth2019using]. Each included location was measured on at least 50 dates. See Appendix \ref{SM-Data_details} for additional data details.
 
@@
 called the "bivariate" datasets in the sense that they comprise two quantities measured at 
-different locations or for different species. The phenology and population datasets are 
+different locations or for different species (Table \ref{tab_data_info}). 
+The phenology, population, and ecosystem functioning datasets are 
 "multivariate" in that they comprise, for each taxon, measurements through time at multiple locations; 
-copula structure can be studied for each pair of locations. Likewise the ecosystem 
-functioning (methane flux) dataset was  multivariate in this sense (Table \ref{tab_data_info}).
+copula structure can be studied for each pair of locations.
+
+<!--***DAN: made it to here in my editing for length, 2019 05 02-->
 
@@
 
-## Results for Q3: Tail dependence influences skewness of spatial average, extinction risk, and Taylor's law
+# Results for Q3: Tail dependence influences skewness of spatial average, extinction risk, and Taylor's law
 

### 2019-05-02 SupportingInformation.Rmd
minor edit
editing for length

@@
 ---
-title: "Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields"
+title: "Supporting Information : The importance of a complete description of dependence between variables for ecology and related fields"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
@@
 
+\textbf{Environmental data, soil C and N.} Measurments were from the central pedon of a soil 
+core, to 1m depth.
+
+\textby{Phenological and population data on aphids.}
+Data pre-processing for the aphid data was the same as 
+that of @Sheppard2016,
+but we excluded one of the 11 sites they used in order to only use sites with at least 30 
+years data.
+
 \textbf{Population data, plankton abundances.}
+These data 
+were obtained from the Continuous Plankton Recorder (CPR) dataset, now operated and maintained by 
+the Marine Biological Association of the United Kingdom. Data pre-processing was similar but not identical to that 
+of @Sheppard2017. 
 The CPR plankton data presented in @Sheppard2017 was aggregated in the following way. Fifty 
@@
 upper-right triangle of the copula plot. Data were then re-ranked as
-described in the Introduction to produce a new copula plot based
+described in the \nameref{MT-Intro} to produce a new copula plot based
 only on the censored dataset (Fig. \ref{fig:missingdata}B). Copula

### 2019-05-02 SupportingInformation.Rmd
typo fixed

@@
 
-\textby{Phenological and population data on aphids.}
+\textbf{Phenological and population data on aphids.}
 Data pre-processing for the aphid data was the same as 

### 2019-05-03 Paper.Rmd
continuing edits

@@
 $u_i$ is the rank of $x_i$ in the set of the $x_j$ ($j=1,\ldots,n$), divided by
-$n+1$; and $v_i$ is defined in the same way but using the $y_i$. See, 
+$n+1$; and $v_i$ is defined in the same way but using the $y_i$. The $u_i$ and 
+$v_i$ are called *normalized ranks* of the $x_i$ and $y_i$. See, 
 for instance,
@@
 
-<!--***DAN: made it to here in my editing for length, 2019 05 02-->
-
 <!--summary table with info of all data used in Paper-->
@@
 To make results interpretable
-to readers unfamiliar with copulas, we first introduce the normal copula and 
-some other copulas we used; and we operationally introduce measures of tail dependence
-and asymmetry of tail dependence. 
+to unfamiliar readers, we first introduce the copula families we used, and 
+measures of tail dependence. 
 Conceptually, a bivariate copula is a probability density function (pdf)
 on the unit square with uniform maginals. 
-The copula is actually defined as the bivariate 
-cdf of the distribution rather than as the pdf itself,
-but this makes little difference. 
-The normal copula family is a one-parameter
+The copula is actually defined as the corresponding 
+cdf rather than as the pdf,
+but both functions contain essentially the same information. 
+The normal copula family (henceforth abreviated N) is a one-parameter
 family of copulas, with parameter corresponding to the correlation of the related 
@@
 (Fig. \ref{fig_cop_pedag3}A-B for an example).
-Left- and right-tail dependence are $0$ for the normal copula, and this 
+Left- and right-tail dependence are $0$ for the N copula, and this 
 serves as a benchmark for tail dependence in other copulas - positive values of our
-tail dependence measures correspond to dependence stronger than that of a normal
+tail dependence measures correspond to dependence stronger than a normal
 copula. The Clayton and 
@@
 the example of Fig. \ref{fig_cop_pedag1}G) and $0$ upper-tail dependence; the 
-survival-Clayton, which is a rotation of the Clayton by $180^\circ$, has 
-positive upper-tail dependence ($0.87$ in Fig. \ref{fig_cop_pedag1}H)
+survival-Clayton is a rotation of the Clayton by $180^\circ$ and has 
+positive upper-tail ($0.87$ in Fig. \ref{fig_cop_pedag1}H)
 and $0$ lower-tail dependence. For our model selection procedure (see below), we used 
-several one- and two-parameter families of copulas exhibiting a wide variety of 
+several one- and two-parameter families of copulas exhibiting a variety of 
 positive lower-tail dependence values 
 (Clayton, survival-Gumbel, survival-Joe, survival-BB6, henceforth
-abbreviated, respectively, C, SG, SJ, SBB6) and upper-tail dependence values
-(survival-Clayton, Gumbel, Joe, BB6, henceforth abbreviated, respectively, 
+abbreviated C, SG, SJ, SBB6) and upper-tail dependence values
+(survival-Clayton, Gumbel, Joe, BB6, henceforth abbreviated 
 SC, G, J, BB6). We also considered families that could exhibit positive upper- or lower-tail
 dependence, or both, depending on parameters (BB1, survival-BB1, BB7, survival-BB7, henceforth 
-abbreviated, respectively, BB1, SBB1, BB7, SBB7). 
-If LT and UT are used to denote 
-lower- and upper-tail dependence, their difference is a measure of asymmetry of 
+abbreviated BB1, SBB1, BB7, SBB7). 
+If LT and UT denote 
+lower- and upper-tail dependence measures, their difference measures asymmetry of 
 tail dependence. We also used two families that were asymmetric in their non-tail parts, 
-but that had zero tail dependence (BB8, survival-BB8, abbreviated BB8 and SBB8).
-In addition to the normal copula, we used another symmetric copula,
+but that had zero tail dependence (BB8, survival-BB8, abbreviated BB8 and SBB8), and
+a symmetric copula called 
 the Frank copula (abbreviated F; Fig. \ref{fig_cop_pedag3}C-D). 
 Copulas are implemented in the `copula` and `VineCopula` packages for the 
-R statistical computing language.
-Additional details are in introductory texts 
-on copulas [@nelsen2006_copula ; @joe2014_dependence] and in `VineCopula` documentation.
-Visualizations of the copulas we used are in Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}.
+R language. Visualizations of the copulas we used are in Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}. 
+See @nelsen2006_copula, @joe2014_dependence and the `VineCopula` documentation
+for more details.
 
@@
 \caption{Log-transformed pdfs (A, C) and samples from (B, D) example copulas.
-Shown are the normal copula with parameter $0.8$ and the Frank copula with 
-parameter $10$. See also Fig. \ref{fig_cop_pedag1}E-H for the Clayton and  
-survival-Clayton copulas with parameter 5. Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}
+Shown are the normal (N) copula with parameter $0.8$ and the Frank (F) copula with 
+parameter $10$. See also Fig. \ref{fig_cop_pedag1}E-H for the Clayton (C) and  
+survival-Clayton (SC) copulas with parameter 5. Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}
 have visualizations of all the copulas we used.\label{fig_cop_pedag3}}
@@
 
-We addressed Q1 principally via a model selection procedure in which the normal copula
-and the alternative copulas listed above were fitted to 
-our ecological datasets. The resulting fits of the various copula models 
-were compared via the Akaike and Bayesian 
+We addressed Q1 principally via a model selection procedure in which the 
+copulas listed above were fitted to 
+our ecological datasets and fits were compared via the Akaike and Bayesian 
 Information Criteria (AIC and BIC). For bivariate datasets 
 $(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
-First, as in the Introduction, we produced ranks 
-$(\tilde{u}_i,\tilde{v}_i)$ where $\tilde{u}_i$ is the rank of $x_i$ in the set 
-$\{x_i:i=1,...,n\}$ and $\tilde{v}_i$ is the rank of $y_i$ in the set $\{y_i:i=1,...,n\}$.
-Ranks were $1$ for the smallest element in a set. We then 
-set $u_i=\tilde{u}_i/(n+1)$ and $v_i=\tilde{v}_i/(n+1)$; subsequent analysis used these 
-*normalized ranks* $u_i$ and $v_i$. 
-Second, we did a test of the independence of the $u_i$ and $v_i$ 
-(implemented in `BiCopIndTest` in the `VineCopula` package). 
-Subsequent model 
-selection algorithms are ineffective if data could not be distinguished from independent data 
-since most or all bivariate copula families we considered include the independent copula in the 
+First, we produced normalized ranks 
+$u_i$ and $v_i$ as in the Introduction. 
+Second, we tested the independence of the $u_i$ and $v_i$ 
+(using `BiCopIndTest` in the `VineCopula` package). 
+Our model 
+selection algorithms were ineffective if data could not be distinguished from independent data 
+since many of the bivariate copula families we considered include the independent copula in the 
 family or as a boundary case. If independence could be rejected ($0.05$ significance level), model 
-selection proceeded. Third, we fit the bivariate copula families listed previously to the ranked data 
-$(u_i,v_i)$ via maximum likelihood, using the `BiCopEst` function in `VineCopula`, and 
-obtained AIC and BIC values and accompanying model weights, $\text{AIC}_{\text{w}}$ and 
-$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection]. The weight of a model represents the 
-relative support of data for the model; weights sum to 1 across models considered. The lower- and 
-upper-tail dependence of the best-fitting member of each family was also provided by `BiCopEst`. 
-Fourth, $\text{AIC}_{\text{w}}$ values for all fitted models were used to get
+selection proceeded. Third, we fit the bivariate copula families listed previously to the 
+normalized ranks via maximum likelihood, using the `BiCopEst` function in `VineCopula`, 
+then obtaining AIC and BIC values and accompanying model weights $\text{AIC}_{\text{w}}$ and 
+$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection]. `BiCopEst` also provided the lower- and 
+upper-tail dependence of the best-fitting member of each family. 
+Fourth, $\text{AIC}_{\text{w}}$ values were used to get
 model-averaged lower- and upper-tail dependence values using standard model averaging 
 formulas [@burnham2003_modelselection]; likewise for BIC.
-Model selection methods will give the *relative* support of several models, but will not,
-on their own, indicate whether any of the models selected are an objectively good fit: all 
-models considered might poorly describe the data. To test this, and as the final step of
+Model selection methods give the *relative* support of several models, but do not 
+indicate whether any of the models are an objectively good fit. To test that, 
+and as the final step of
 our procedure, we tested the goodness of fit of our AIC-best copula family using a 
-bootstrapping procedure implemented in `BiCopGofTest` of `VineCopula`. The method performed 
-two tests, one based on a Cramer-von Mises statistic, and one based on a 
+bootstrapping procedure (`BiCopGofTest` of `VineCopula`). The method performed 
+one test based on a Cramer-von Mises statistic and another based on a 
 Kolmogorov-Smirnov statistic. To keep computation times reasonable, 
-an initial run using $100$ bootstraps was performed, and if
-the $p$-value resulting from either test was less than $0.2$, tests were re-run with 
+an initial run using $100$ bootstraps was performed; if
+the $p$-value from either test was less than $0.2$, tests were re-run with 
 $1000$ boostraps.
@@
 The first part of Q1 was answered for a bivariate dataset 
-by comparing the AIC weights of the non-normal copula
-families with that of the normal family. 
+by comparing the AIC weights of the non-N copula
+families with that of the N family. 
 The second part of Q1 was answered by comparing 
 the model-averaged lower- and upper-tail dependence
-values to the lower- and upper-tail dependence values of a normal copula (which 
-are 0). The third part of Q1
+values to the lower- and upper-tail dependence values of an N copula (i.e., 0). The third part of Q1
 was answered by comparing the difference of the model-averaged lower- and upper-tail 
-dependence values to the analogous difference for a normal copula, again 0.
-For multivariate datasets,
-each of which consists of time series from multiple locations, we carried out the bivariate analysis
-for all possible pairwise combinations of locations and compared AIC weights and tail-dependence of 
-minimum-AIC copulas with that of the normal copula. We also counted the number of location pairs 
+dependence values to the analogous difference for an N copula, again 0.
+For multivariate datasets, we carried out the bivariate analysis
+for all possible pairwise combinations of locations, and compared AIC weights and tail-dependence of 
+minimum-AIC copulas with that of an N copula. We also counted the number of location pairs 
 for which model-averaged lower-tail dependence was greater than model-averaged upper-tail dependence 
@@
 
+<!--***DAN-->
+
 In addition to our model selection approach, we also used nonparametric approaches to 
-quantify lower- and upper-tail dependence and to answer Q1. Using two approaches can provide greater 
-confidence in results. We used three nonparametric
-statistics which can quantify the extent to which 
-two normalized ranks $u_i$ and $v_i$ are 
+provide greater 
+confidence in our answer to Q1. We used three 
+statistics which quantify the extent to which 
+the normalized ranks $u_i$ and $v_i$ are 
 related in any part of their distributions; we here describe the 
-statistics operationally, with details in Appendix \ref{SM-nonparam_stats}.
-The statistics were formulated with positively associated
+statistics conceptually, with details in Appendix \ref{SM-nonparam_stats}.
+The statistics are defined with positively associated
 variables in mind. All the variables we study were positively associated
-when they were significantly associated. So definitions and associated 
-diagrams will assume positive associations. Given two bounds 
-$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
+when they were significantly associated.  
+Given two bounds 
+$0 \leq l_b < u_b \leq 1$, define the lines $u+v=2l_b$ and 
 $u+v=2u_b$, which intersect the unit square (Fig. \ref{SM-fig_stat_image}). 
@@
 between $u_i$ and $v_i$ in the region bounded by these lines.
-Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
+Using $l_b=0$ and $u_b\leq 0.5$ gives information about 
 dependence in the left parts of the distributions of $u$ and $v$, and 
-using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
-the right parts of the distributions. The first statistic, denoted 
+using $u_b=1$ and $l_b \geq 0.5$ gives information about dependence in 
+the right parts. The first statistic, 
 $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
 correlation of $u_i$ and $v_i$ that is attributable to the points 
-lying in the region given by the bounds.
-Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
-between $u$ and $v$ in the region.
+between the bounds.
+Larger values indicate stronger positive association
+between between the bounds.
 We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
@@
 $u_i+v_i<2u_b$ and the line $v=u$.
-
 Unlike $\cor_{l_b,u_b}$ and 
 $\Ps_{l_b,u_b}$, for which large values inidicate strong dependence
-in the portions of the the distributions of $u$ and $v$ given by 
-the bounds, small values of 
+for points between the bounds, small values of 
 $\Dsq_{l_b,u_b}$ indicate strong dependence.
+
 For large datasets (large $n$),
-we could use $l_b$ and $u_b$ close together without incurring undue
-sampling variation in our statistics, and we could consider 
+we used $l_b$ and $u_b$ close together without incurring undue
+sampling variation in our statistics, and we considered 
 multiple bands $(l_b,u_b)$ to understand how dependence
 varies in different parts of the distributions. But for datasets with 
-smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
-We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
-and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
+smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$,
+abbreviating $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
+and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"). Likewise $\Ps_l=\Ps_{0,.5}$,
 $\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.
+
 To test for asymmetry of dependence in upper and lower portions of distributions,
-we used differences of upper and lower statistics: for smaller datasets, 
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ (note the opposite order
-in the last of these); for large datasets, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
+we used differences $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
+for smaller datasets, (note the opposite order
+in the last of these); and $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
 $\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
-$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ (again note the opposite order) with $u_b$ close 
-to $0$.
+$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ with $u_b$ close 
+to $0$ for large datasets.
 We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}). All code for this
@@
 \caption{Upper panels show raw data plots for (A) $\log_{10}$(soil C) and $\log_{10}$(soil N) data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
-for birds, (C) $\log_{10}$(BMR) vs.
-$\log_{10}$(body mass) for mammals, and (D) above-ground plant biomass vs.
+for birds, (C) the same for mammals, and (D) above-ground plant biomass vs.
 Shannon's index \textit{(H)} from Cedar Creek.
@@
 \end{figure}
+<!--***DAN: Shya, it say "Avg. H" and "Avg. biomass" on the axes of D. Reading the sup mat
+data section carefully, I presume this is because there were multiple strips per plot
+and these were averaged. However, the averaging is not mentioned in the main text, so I think
+this will be confusing for some readers. Can you please change the axes to read "Biomas" and "H"?-->
 

### 2019-05-03 Paper.Rmd
plot label : cedar creek fixed

@@
 \end{figure}
-<!--***DAN: Shya, it say "Avg. H" and "Avg. biomass" on the axes of D. Reading the sup mat
-data section carefully, I presume this is because there were multiple strips per plot
-and these were averaged. However, the averaging is not mentioned in the main text, so I think
-this will be confusing for some readers. Can you please change the axes to read "Biomas" and "H"?-->
 

### 2019-05-03 MakeBivarDataPlot.R

@@ -82,8 +82,8 @@ d4<-sr$tab_dbp[,c(2:3)]
 d4c<-sr$cop_tab_dbp[,c(2:3)]
 plot(d4$avg.H,d4$avg.Biomass,type='p',pch=pchval,cex=cexvalpts,
      col="grey")
-mtext("Avg. H",1,1.15)
-mtext("Avg. biomass",2,1.15)
+mtext("H",1,1.15)
+mtext("Biomass",2,1.15)
 mtext("D",3,-1.4,cex=cexvalpl,adj=.05)
 
 #***bottom row of panels : copula plots of corresponding figs of upper panel

### 2019-05-03 SupportingInformation.Rmd
rds to csv storage

@@
 ```{r load_and_plot_data_soilCN, echo=F, results='hide'}
-#fig.cap="Plot of the raw soil C and N data.\\label{fig_raw_cop_soilCN}(A)", fig.height=3, fig.width=3}
-d<-readRDS("Data/dataForShymolinaCoord.rds")
+#d_old<-readRDS("Data/dataForShymolinaCoord.rds")
+#write.csv(d_old,"./Data/SoilCN/SoilCNdata.csv")
+d<-read.csv("./Data/SoilCN/SoilCNdata.csv")
+rownames(d)<-d$X
+d<-d[,-1]
+#sapply(dCN,class)
+
 d<-d[,c("SOCstock100","TSNstock100","lat","long")]
+#identical(d_old[,c("SOCstock100","TSNstock100","lat","long")],d) # checked : they are identical :)
 d<-na.omit(d)
@@
       col.names = c("Plankton taxon name"))%>%row_spec(row=idn,italic=T)
-```
-
-<!--reading multivariate methane flux data-->
-```{r read_methane_data, echo=F, results="markup"}
-set.seed(seed)
-methane_raw<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")
 ```
@@
 ```{r read_format_methane_data, echo=F}
-
-methane_raw<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")
+#latlon_methane_old<-readRDS("./Data/Methane_data/chamberLocations.rds")
+#write.csv(latlon_methane_old,"./Data/Methane_data/chamberLocations.csv",row.names = F)
+latlon_methane<-read.csv("./Data/Methane_data/chamberLocations.csv")
+#identical(latlon_methane_old,latlon_methane)  # checked : they are identical :)
+
+#methane_raw_old<-readRDS("./Data/Methane_data/methane_data_for_Shyamolina.rds")
+#write.csv(methane_raw_old,"./Data/Methane_data/methane_data.csv",row.names = F)
+methane_raw<-read.csv("./Data/Methane_data/methane_data.csv")
+methane_raw$dates<-as.Date(methane_raw$dates) # converting factor class to date class
+#identical(methane_raw_old,methane_raw)  # checked : they are identical :)
 
@@
 <!--Do the main computations for nonparametric analysis of methane flux data-->
-```{r stat_methane, echo=F, cache=T, cache=T,  cache.extra=list(seed,methane_data,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
+```{r stat_methane, echo=F, cache=T, cache=T,  cache.extra=list(seed,methane_data,methane_raw,latlon_methane,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
 
@@
 
-latlon<-readRDS("./Data/Methane_data/chamberLocations.rds")
+latlon<-latlon_methane
 chamber_name<-colnames(methane_raw[,c(2:dim(methane_raw)[2])])
@@
 <!--Following chunk examine the consequence of non-Normal copula atructure on skewness using methane flux data-->
-```{r skewness_methane_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
+```{r skewness_methane_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,methane_data,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
 

### 2019-05-03 Paper.Rmd
edits

--- 2019-05-03 Paper.Rmd
@@
 
-<!--***DAN-->
-
 In addition to our model selection approach, we also used nonparametric approaches to 
@@
 
-Across essentially all the sub-fields and levels of organization within ecology for which we 
-examined a dataset, copula structures were typically non-normal and variables 
-commonly exhibited asymmetric tail 
-dependence. To make results easier to absorb for readers who are new to copulas: we first 
-present results in detail for an example bivariate dataset; then we summarize results
-for all bivariate datasets; then we present results in 
-detail for an example multivariate dataset; then we summarize results 
-for all multivariate datasets.
+Across essentially all our datasets, copula structures were typically non-normal and 
+commonly showed asymmetric tail 
+dependence. To make results easier to absorb: we first 
+present results in for an example bivariate dataset; then we summarize results
+for all bivariate datasets; then we present results for an example multivariate dataset; then we summarize for all multivariate datasets.
 
 <!--bivariate example-->
-We present the example of the soil C and N data (see Data), pictured in 
-Fig. \ref{fig_biv_multi_raw_cop}A (raw data) and E (copula plot). The variables were 
+For the soil C and N data (Data, Table \ref{tab_data_info}, 
+Fig. \ref{fig_biv_multi_raw_cop}A,E), variables were 
 non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
-from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent
-(Fig. \ref{fig_biv_multi_raw_cop}A,E).
-The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted the 16 copulas
-mentioned in the previous section to the 
-data pictured in Fig. \ref{fig_biv_multi_raw_cop}E; AIC and BIC weights and corresponding
-lower- and upper-tail statistics for the fitted distributions were computed (Table \ref{SM-tab_soilCNfit}). 
+from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent.
+The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
+(previous section) to the data and computed AIC and BIC weights and corresponding
+lower- and upper-tail statistics for the fitted distributions (Table \ref{SM-tab_soilCNfit}).
 `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the 
@@
 first part of Q1 for these data. 
-The model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
-and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for model averaging). 
+Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
+and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
 These values are distinct from what a normal copula would give, namely, 0, answering the second part 
-of Q1 for these data. The values also differed substantially 
+of Q1. The values also differed substantially 
 from each other, answering the third part of Q1. The numbers reflect stronger upper- than lower-tail
-dependence for the soil C and N data, and this can also be seen visually in extreme upper-right
+dependence, and this is also visible in the extreme upper-right
 and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).
 
-Although these results do convincingly show non-normal copula structure, 
+Although these results convincingly show non-normal copula structure, 
 we caution that the model-averaged tail-dependence statistics may have been biased 
@@
 Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
-available from `BiCopGofTest` in the `VineCopula` package). However, nonparametric statistics verified
-both that tail dependence was distinct from that of an N copula for the soil data, and that it 
+available from `BiCopGofTest`). However, nonparametric statistics verified
+both that tail dependence was distinct from that of an N copula, and that it 
 was asymmetric (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}, 
@@
 <!--bivariate summary-->
-Table \ref{tab_bivar_summary} summarizes results for Q1 for the bivariate datasets. For all $4$ 
-bivariate datasets, the two variables were always significantly related 
-(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
+Table \ref{tab_bivar_summary} summarizes Q1 results for the bivariate datasets. The two 
+variables were significantly related for all datasets 
+(Table \ref{tab_bivar_summary}, row 1). A non-N copula 
 always emerged as the lowest-AIC (highest-AIC-weight) copula 
@@
 (Table \ref{tab_bivar_summary}, rows 3-4), answering the first part of Q1. 
-Lower- and upper-tail statistics (Table \ref{tab_bivar_summary}, rows 7-8) 
-differed substantially from 0, the tail dependence statistics of 
-the N copula, answering the second part of Q1; and also differed substantially from each other
-(Table \ref{tab_bivar_summary}, row 9),
-answering the third part of Q1. The first three datasets had stronger upper- 
-than lower-tail dependence, and the last dataset had the reverse. 
+Lower- and upper-tail dependence statistics often differed substantially from 0 
+(Table \ref{tab_bivar_summary}, rows 7-8) and from each other (Table \ref{tab_bivar_summary}, row 9), 
+answering the second and third parts of Q1. The first three datasets had stronger upper- 
+than lower-tail dependence, and the last dataset had the reverse, though weakly. 
 Although these results convincingly show non-normal copula structure, the 
@@
 But our nonparametric statistics confirmed the same conclusions 
-(Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}) for first three bivariate datasets.
-The fourth bivariate datasets we analyzed was above-ground biomass and percent cover data (i.e. 
-the data used to compute Shannon's diversity index) for the year 2000, but we did the same analysis 
-with Cedar creek data sampled over 168 plots from other available years (1996-1999, 2007) too. For 1996 and 
-2007 copula structures showed no dependence between two variables concerned, for other years (1997-1999), 
-we found similar results as of year 2000 (i.e. best fit non-Normal copula and overall lower tail-dependence).
-Due to smaller sample size relative to other three bivariate datasets, we chose a wider range with 
-non-parametric analysis (Appendix \ref{SM-Q1bivarnonparam}).
-Detailed results for each dataset are in Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and Figs.
-\ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.
-
-We present our analysis of the green spruce aphid abundance data (see Data)
+(Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}).
+We did the same analysis 
+with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
+2007, biomass and Shannon's index were independent; for 1997-1999, results 
+were similar to 2000: the best-fitting copula was not N, and there was slightly stronger lower- 
+than upper-tail dependence. Detailed results for each dataset are in 
+Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and 
+Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.
+
+We present green spruce aphid abundance results (Table \ref{tab_data_info})
 as an example multivariate analysis.
-Independence was tested for each pair of the 10 sampling locations, and was 
-rejected in every case. 
-Model selection among copulas was carried out separately for each pair of locations.
-Best-fitting (lowest-AIC) copulas were non-normal
+Independence was rejected for each pair of the 10 sampling locations. 
+Best-fitting (lowest-AIC) copulas were non-N
 for the large majority of location pairs (Table \ref{SM-tab_count_bestcop}), and 
-AIC weights for the best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
+AIC weights for best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
 were typically substantially 
-larger than those for the N copula (Table \ref{SM-tab_count_AICw_normal}), answering 
-the first part of Q1 from the Introduction for these data.
-Goodness of fit tests based on Cramer-von Mises or Kolmogorov-Smirnov statistics 
+larger than N-copula weights (Table \ref{SM-tab_count_AICw_normal}), answering 
+the first part of Q1. Goodness of fit tests 
 (Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
@@
 an objectively adequate description of 
-the data, demonstrating that the collection of 
-copula families we used was sufficiently broad that the copulas of all pairs of 
+the data; the collection of 
+copula families we used was apparently sufficiently broad that all pairs of 
 locations could be characterized by at least one of them.
@@
 (Table \ref{SM-tab_count_UTdep_AICw}) dependence statistics were exclusively 
-positive and were frequently substantially large, thereby differing 
+positive and were frequently large, thereby differing 
 from what an N copula would give (i.e., 0), and answering the second part of 
-Q1 for these data. Apparently these aphid data have greater tail dependence 
-(both lower and upper) than would be expected from an N-copula null hypothesis. 
-Model-averaged lower-tail dependence statistics minus model-averaged 
-upper-tail dependence statistics (Table \ref{SM-tab_count_LTmUT}) were positive 
+Q1. These data have greater tail dependence 
+(lower and upper) than would be expected from an N-copula null hypothesis. 
+Model-averaged lower- minus upper-tail dependence statistics 
+(Table \ref{SM-tab_count_LTmUT}) were positive 
 for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
-out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`), 
-indicating that lower-tail dependence across space was much stronger than upper-tail 
-dependence. Thus the spatial synchrony of rarity in green spruce aphid is stronger 
-than the spatial synchrony of outbreak events. This also answers the third part of Q1.
+out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
+than the spatial synchrony of outbreaks. This also answers the third part of Q1.
 Nonparametric statistics verified that tail dependence was asymmetric for the green 
@@
 
+<!--***DAN: You could maybe move the next para to sup mat, though that would complicate the
+presentation of skewness results below. Only do it if you are desperate to cut.-->
 <!--multivariate summary-->
-Table \ref{tab_multivar_summary} summarizes some of our results using our
-multivariate data. In addition to summaries of the green spruce aphid abundance data results 
-elaborated in detail above, the table also summarizes results for 
-leaf-curling plum aphid first flight data, abundance data for the dinoflagellate
-\textit{Ceratium furca}, and methane-flux data (see Data). Results supported the 
+Table \ref{tab_multivar_summary} summarizes Q1 results from
+multivariate data. Results supported the 
 conclusions that non-N copula structure, non-N tail dependence, and asymmetric tail 
@@
 row 3), and AIC weights of best-fitting copulas were 
-substantially better than AIC weights of N copulas (Table \ref{tab_multivar_summary}, rows 4-5).
-For these data, best-fitting copulas were nearly always considered an adequate fit, 
-according to Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests 
+better than those of N copulas (Table \ref{tab_multivar_summary}, rows 4-5).
+Best-fitting copulas were nearly always considered an adequate fit 
 (Table \ref{tab_multivar_summary}, row 6). Some of the datasets (green spruce aphid
@@
 as well as for the 21 plankton taxa for which we had abundance data other than the
-\textit{Ceratium furca} (results not shown). The conclusions were generally supportive that 
-non-N copula structure and tail dependence, as well as asymmetry of tail dependence, are 
-very common.
+\textit{Ceratium furca} (results not shown); conclusions supported that 
+non-N copula structure and tail dependence, and asymmetry of tail dependence, are 
+common.
 
@@
 The $p$-values in rows 5-6 are for the minimum-AIC copula.
-Model averaging used for rows 7-9 was based on AIC weights.\\label{tab_bivar_summary}",
+Model averaging used for rows 7-9 was based on AIC weights.
+Cedar Creek results are for the year 2000.\\label{tab_bivar_summary}",
              booktabs=T, linesep = "\\addlinespace",
@@
 
-Having demonstrated that non-N copula structure and asymmetric tail dependence are common
-features of ecological data, we addressed Q2 by exploring via mathematical 
-models and simulations three possible mechanisms of this structure which 
-seem likely to operate in diverse ecological and evolutionary scenarios. The first mechanism 
+Having demonstrated that non-N copula structure and asymmetric tail dependence are 
+common, we addressed Q2 by exploring, using 
+models, three possible mechanisms. The first mechanism 
 is an extension of the well known Moran
@@
 It explores the hypothesis that non-N copula structure and asymmetric tail dependence 
-between environmental variables measured through time 
-in two places may bring about similarly 
-structured dependence between ecological variables in those places that are influenced by the 
-environmental variables.
-
+between environmental variables, measured through time 
+in two places, may beget similarly 
+structured dependence between ecological variables measured through time in those places. 
 Let $E_i(t)$ be an ecological variable 
-measured in location $i$ ($i=1,2$).
-We initially assumed the variable follows AR(1) dynamics within a location, 
-$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$. 
-We used $\beta=0.5$. Here the $\epsilon_i(t)$ represent environmental noise; they are 
+measured at location $i$ ($i=1,2$) at time $t$.
+We initially assume the AR(1) dynamics 
+$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
+The $\epsilon_i(t)$ represent environmental noise and are 
 standard-normal random variables that are independent for distinct times, $t$, but 
@@
 dependence across locations in different simulations (see below).
-The ecological variable, $E$, is very general, and could represent, for instance, deviations
-of a population density from a carrying capacity or average density, deviations of
-total plant community biomass or other community variable from an average value,
-or flux of a biogeochemical variable such as methane or CO$_2$.
-
-To explore the importance of nonlinear dynamics for our hypothesis, we also considered 
+The variable, $E$, is very general, and could represent deviations
+of a population density from a carrying capacity, deviations of
+total plant community biomass from an average value,
+flux of a biogeochemical variable such as methane or CO$_2$, or other quantities.
+
+To explore the importance of nonlinear dynamics for our proposed mechanism, we also considered 
 nonlinear population models based on $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right) \right)+\sigma\epsilon_i(t))$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
@@
 the weak-noise and strong-noise cases, though the importance of the noise here
-is that, when it is strong, it brings the nonlinearities of the model strongly into play. 
-The $\epsilon_i(t)$ were again standard
-normal and independent through time, but were related (see below) across locations.
-
-For each of the three model setups, for each of the copulas C, SC, G, SG, J, SJ,
-for each Kendall correlation value $\tau=0.1,0.2,\ldots,0.9$, and for each of
+is that, when it is strong, it brings the nonlinearities of the model into play.
+
+For each of the three model setups above, $\tau=0.1,0.2,\ldots,0.9$, and for each of
 $50$ replicate simulations, we generated 5000 noise pairs 
@@
 bivariate random variable with the given copula family and the given $\tau$.
-We then used the generated noise to drive the model, and retained both the
+We then used this noise to drive the model, and retained both the
 noise and population values for the final 500 
-time steps of the simulation. For each simulation, the following
-statistics were computed for both the noises and populations: 
+time steps. For each simulation, the following
+statistics were computed for both noises and populations: 
 Pearson, Spearman and Kendall correlations, $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, 
@@
 Values were plotted for noises and populations against $\tau$. If the hypothesis is
-correct that copula structure between locations of environmental noise can be 
-transmitted to produce similar dependence structure for ecological variables, 
+correct that copula structure in environmental noises begets similar dependence 
+structure for ecological variables, 
 plots should be similar for populations and noises.
 
-The second mechanism we explored, which was also summarized in the 
+Our second mechanism, also summarized in the 
 Introduction,
-relates to Liebig's law of the minimum, or, more generally, to the fact that 
-environmental variables can often influence ecological variables through nonlinear 
-relationships. If an environmental variable influences an ecological variable
-more strongly in one of its tails, we explored using a model 
-whether the ecological variable
-could then exhibit asymmetric tail dependence between measurements made in two
-locations. Let $E_i(t)$ again be an ecological variable measured in locations
-$i=1,2$. Assume $E_i(t)$ has dynamics that follow 
-$E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where $g$ is a function
-describing how the environmental variable $\epsilon_i(t)$ influences
-$E_i(t)$, and the $\delta_i(t)$ are standard-normally distributed 
-local noise, independent across time and sampling locations. We used $a=0.2$, and 
-$b=0.1, -0.1, 0.5, -0.5$ in different simulations. 
-Negative values of $b$ correspond to overcompensating dynamics 
-and positive values correspond to undercompensating dynamics; larger-magnitude $b$
-means that the model has a slower return time to equilibrium after a perturbation. 
-The $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
+relates to Liebig's law of the minimum and to nonlinear 
+influences of environmental variables on ecological variables. 
+If an environmental variable influences an ecological variable
+more strongly in one of its tails, we explored whether the ecological variable
+could then exhibit asymmetric tail dependence across space. 
+Let $E_i(t)$ again be an ecological variable, $i=1,2$. Assume the dynamics 
+$E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where 
+the $\delta_i(t)$ are standard-normally distributed 
+and independent across time and locations, $a=0.2$,
+$b=0.1, -0.1, 0.5, -0.5$ in different simulations, and 
+the $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
 time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
 $\cov(\epsilon_1,\epsilon_2)=0.8$.
-We used $g$ equal to $g_1$ or $g_2$ in two different simulations:
-$g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise; and 
+The function $g$ describes how the environment $\epsilon_i(t)$ influences
+$E_i(t)$. We used $g$ equal to $g_1$ or $g_2$ in different simulations:
+$g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise;
 $g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
 $g_1$ represents environmental effects that occur only for negative 
-values and $g_2$ only for positive values of the environmental variables.
-Note that both $(\epsilon_1,\epsilon_2)$
-and $(\delta_1,\delta_2)$ have N copula structure, being drawn from 
-bivariate normal distributions. Hence the Moran-like mechanism discussed
-previously cannot be operating in this model.
-
-For each value of $b$, we simulated the model once using $g_1$ and once using $g_2$, for $25000$ time
-steps in both cases, retaining $E_1(t)$ and $E_2(t)$ for the final $2500$ time 
-steps of each simulation. We applied our model selection algorithms and 
-non-parametric statistics, which were previously applied to 
-data to answer Q1, to the 
-simulation outputs to discover if the model could
-give rise to non-N copula structure and asymmetric tail dependence in the 
-ecological variable, in spite of the absence of such structure in the driver
-environmental variables.
-
-The third mechanism we explored is evolutionary, and pertains to bivariate 
-trait data across species, e.g., to our bird and mammal BMR versus body mass 
+values and $g_2$ only for positive values of $\epsilon$.
+Negative values of $b$ correspond to overcompensating dynamics 
+and positive values to undercompensating dynamics; larger $|b|$
+means slower return to equilibrium after a disturbance. 
+Because $(\epsilon_1,\epsilon_2)$
+and $(\delta_1,\delta_2)$ have N copula structure, the Moran-like 
+mechanism discussed previously does not operate here.
+
+For each $b$ and $g_i$, we simulated the model for $25000$ time
+steps and retained the $E_i(t)$ for the final $2500$ time 
+steps. We applied our model selection algorithms and 
+non-parametric statistics (methods for Q1) to these 
+outputs to discover if the model could
+produce non-N copula structure and asymmetric tail dependence in the 
+$E_i(t)$.
+
+Our third mechanism is evolutionary, and pertains to bivariate 
+trait data across species, e.g., our bird and mammal 
 data. This mechanism is a hypothetical explanation for the right tail dependence
@@
 The root character state and the change 
-across each branch were randomly chosen from matrices of one million different
+across each branch were randomly chosen from matrices of one million 
 bivariate draws from distributions showing one of five distinct types of copula 
-structures: 1) extreme left-tail dependence, 2) moderate left-tail dependence,
+structure: 1) extreme left-tail dependence, 2) moderate left-tail dependence,
 3) symmetric tail dependence, 4) moderate right-tail dependence, and 5) extreme
-right-tail dependence (Appendix \ref{SM-evol_mech} for details). 
-All marginal distributions were standard normal, and distributions all had Spearman 
-correlation $0.875$ between their two components, so our simulations assess 
-the impact only of copula structure in evolutionary character
-changes on evolutionary outcomes. For each of the five copula structures, mammalian character
+right-tail dependence (Appendix \ref{SM-evol_mech}). 
+All distributions had standard-normal marginals and Spearman 
+correlation $0.875$ between components, so our simulations assess 
+the impact of copula structure only. For each of the five copula structures, mammalian character
 evolution was simulated 100 times. For each simulation, symmetry of tail dependence 
-of the two characters across the phylogeny tips was assessed.
+of the two characters across phylogeny tips was assessed.
 We hypothesized that cases 1 and 2 above would yield stronger left- than right-tail 
-dependence, and cases 4 and 5 would yield the reverse. 
+dependence in tip characters, and cases 4 and 5 would yield the reverse. 
 The simulator was written in Python and used version 4.4 of the DendroPy 
@@
 
-The hypothesis that copula structure and tail dependence between environmental 
-variables measured through time in different locations can produce similar 
-structure in ecological variables influenced by the environmental variables was
-largely confirmed, though similarities between environmental- and 
-ecological-variable copula structure and tail dependencies can be reduced when the
-intrinsic dynamics of the ecological variable are strongly nonlinear.
-For simulations using AR(1) models (see section on methods for Q2), copula 
+The hypothesis that copula structure in ecological variables can arise through
+Moran-like effects was largely confirmed, though similarities between environmental- and 
+ecological-variable copula structure can be reduced when dynamics are strongly nonlinear.
+For simulations using AR(1) models, copula 
 statistics were always very similar for noise and populations 
@@
 See Fig. \ref{SM-fig_Cause4copula_C_kend_ar1} for additional results for the AR(1) model and Clayton copula, 
-Figs \ref{SM-fig_Cause4copula_SC_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1} for AR(1) results for other copula families, and 
+Figs \ref{SM-fig_Cause4copula_SC_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1} for AR(1) results with other copula families, and 
 Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1} for
-the other models described in the methods for Q2.\label{fig_Cause4copula_scatter_C_kend_ar1}}
+the nonlinear models described in the methods for Q2.\label{fig_Cause4copula_scatter_C_kend_ar1}}
 \end{center}
@@
 
-For weak noise and using our nonlinear model (methods for Q2), copula statistics
+For weak noise and using our nonlinear model, copula statistics
 were again very similar for noise and model outputs 
-(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_0.1}).
+(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-
+\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}).
 Though there were again significant but small differences for some statistics
-and simulations, the overall hypothesis that copula structure in the noise translates
-to similar but not identical copula structure in output populations, at least
-as measured with the tail dependence and correlation statistics we used, was
-supported. Since many ecological models are nonlinear, this result provides the 
+and simulations, the overall hypothesis was supported that copula structure in the noise translates
+to similar but not identical copula structure in output populations. 
+Since many ecological models are nonlinear, this result provides the 
 important expectation that there will be a Moran-effect-like correspondence 
@@
 one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
-at $K$ when $r<1$ (undercompensating dynamics), but exhibits 
-an oscillatory approach to $K$ when $r>1$ (overcompensating
-dynamics).
-
-For strong noise and using our nonlinear model (methods for Q2), 
+at $K$ when $r<1$ (undercompensating dynamics, e.g., the value $r=0.5$ used initially), 
+but exhibits an oscillatory approach to $K$ when $r>1$ (overcompensating
+dynamics, e.g., $r=1.3$).
+
+For strong noise and using our nonlinear model, 
 copula statistics, generally speaking, were approximately similar 
 between noise and model
-outputs; however, similarity was reduced relative to the AR(1) 
-and weak-noise cases, and, for a few simulations, tail dependence 
-asymmetry statistics
+outputs; however, similarity was reduced relative to previous results, 
+and for a few simulations, tail dependence asymmetry statistics
 even had opposite signs for noise and model outputs 
-(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). For instance, using a C copula with a large Kendall $\tau$,
+(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). For instance, using a C copula with a large Kendall $\tau$,
 $\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
@@
 and likewise for SG and SJ copulas (Figs \ref{SM-fig_Cause4copula_SG_kend_sRicker_nsd_1}, 
-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). Dynamics were
-strongly nonlinear for the model parameters we used, and nevertheless similarities
-between noise and model-output copula structure, according to the statistics 
-we used, were more noticable than discrepancies.
-We again repeated the analysis with $r=1.3$. Discrepancies between noise and population
+\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). Overall,
+similarities between noise and model-output copula structure
+were more noticable than discrepancies, in spite of nonlinearities.
+We repeated the analysis with $r=1.3$. Discrepancies between noise and population
 copula structure in this case were often glaring. Apparently noise of standard deviation 1
@@
 
-Our model with environmental effects which differed in the
-two tails could produce simulated ecological variables with visually apparent asymmetry 
-of tail dependence and non-N copula structure, depending on the value of $b$ used. 
+Our model with asymmetric environmental effects 
+could produce outputs with visually apparent asymmetry 
+of tail dependence, to an extent that depended on the value of $b$ used. 
 For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ (Fig. \ref{SM-fig_ms_asym_b_-0.1}), 
 for both $g_1$ and $g_2$,
-asymmetric tail dependence and non-N copula structure were visually apparent.
-We also got similar results from simulations with $b=\pm 0.5$ (Figs \ref{SM-fig_ms_asym_b_0.5} - 
+asymmetric tail dependence was strong; for 
+$b=\pm 0.5$, asymmetric tail dependence was weaker but still apparent
+(Figs \ref{SM-fig_ms_asym_b_0.5} - 
 \ref{SM-fig_ms_asym_b_-0.5}).
@@
 Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
-non-Normal copula was always best fit for $b= \pm 0.1$ and $\pm 0.5$, but their model 
-AIC-wieght averaged tail dependence could not reflect the asymmetry for each case 
-when that dependnece was at too much extreme end.
-However, our nonparametric statistics (Tables \ref{SM-tab_stat_npa_asym}) focused on those extreme tips
-and overcome the shortcomings of model selection approach which fit data based on maximum likelihood.
+ecological outputs had non-N copula structure, but did not always reflect the 
+asymmetry of tail dependence, perhaps because those methods measured
+dependence in the most extreme tails. Results using our nonparametric statistics, however,
+reflected the visually apparent asymmetry (Tables \ref{SM-tab_stat_npa_asym}).
 
@@
 tail dependence across space in the ecological variables. Shown are the last
-$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text.\label{fig_ms_asym_b_0.1}}
+$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text, $b=0.1$.\label{fig_ms_asym_b_0.1}}
 \end{center}
@@
 
-Our hypothesis was correct that asymmetric tail dependence in evolutionary changes in bivariate
-characters can give rise to similarly asymmetric tail dependence in character distributions across
+Our hypothesis was correct that asymmetric tail dependence in evolutionary changes 
+can produce similarly asymmetric tail dependence in character distributions across
 phylogeny tips (Appendix \ref{SM-evol_mech}). This topic is revisited in the Discussion.
@@
 
-We addressed Q3 using both models and data. An initial 
-idea was presented in the Introduction that
+An idea was presented in the Introduction that
 the distribution (through time) of a spatially averaged quantity should be influenced by the 
-nature of the dependence between the local quantities being averaged, including their
-copula structure and tail dependence. For instance, suppose $E_i(t)$ are ecological
-variables measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and suppose 
-the distribution and temporal dynamics of the spatial average $\sum_i E_i(t)/N$ is of
+dependence between the local quantities being averaged, including their
+copula structure and tail dependence. For instance, suppose an ecological variable 
+$E_i(t)$ is measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and  
+the distribution and temporal dynamics of the spatial mean $\sum_i E_i(t)/N$ is of
 interest. The $E_i(t)$ could be, for instance, local abundances of a pest or 
@@
 exceptionally large values tend to occur at the same time in all or most locations.
-This should tend to increase the skewness of the distribution through time of the
-spatial average. Similarly, left-tail dependence between local variables
-will tend to decreases skewness of the spatial mean. Positive skewness of the 
-spatial mean corresponds to the "spikiness" of the spatial-mean time series. 
+This should tend to increase the skewness of the distribution through time 
+of the spatial mean. Similarly, left-tail dependence between local variables
+will tend to decrease skewness of the spatial mean. Positive skewness of the 
+spatial mean corresponds to "spikiness" of the spatial-mean time series. 
 The spatial-mean time series and its skewness may be quantities of 
@@
 cases such as pest or resource abundance, for which extreme values (spikes) 
-in the spatial-mean
-quantity have disproportionately large effects.
-
-We tested these ideas using the focal multivariate datasets of our earlier analyses:
-green spruce aphid abundance data, leaf-curling plum aphid first flight data, \emph{Ceratium
-furca} abundance data, and methane flux data (see Data and Table \ref{tab_multivar_summary}).
-For each dataset we calculated the spatial mean, and then the skewness (through time) of that
-mean. Then, for each dataset separately, we compared the value obtained 
+in the spatial-mean have disproportionately large effects.
+
+We tested these ideas using our multivariate datasets (Table \ref{tab_multivar_summary}).
+For each dataset we calculated the spatial mean time series, and then 
+the skewness (through time) of that
+mean. Then, for each dataset, we compared the value obtained 
 to a distribution of values 
 of the same quantity calculated for each of 10000 surrogate datasets. Surrogate 
-datasets were produced from the empirical data by randomizing
-it in a special way so that surrogate/randomized dataset had the 
-copula structure of a multivariate-normal distribution
-(Appendix \ref{SM-Surrog_multi}). The surrogate datasets had exactly the same
-distributions of values for each sampling location as the original data. 
-They also had very similar Spearman correlations between pairs of sampling locations
-as the original data. 
+datasets were produced by randomizing the empirical data
+in a special way to have the 
+copula structure of a multivariate-normal distribution, but to retain exactly the same
+distributions of values for each sampling location as the original data
+(Appendix \ref{SM-Surrog_multi}). 
+Surrogates also had very similar Spearman correlations between pairs of sampling locations
+as the data. 
 Our comparisons therefore tested the null hypothesis that the skewness of the spatial 
@@
 (Table \ref{tab_multivar_summary}), we compared empirical and surrogate skewness values
-via a one-tailed test in the left tail: the $p$-value result was simply the fraction
+via a one-tailed test in the left tail: the $p$-value result was the fraction
 of surrogate skewnesses less than the skewness computed from the empirical data; the test
-examines whether lower-tail dependence made the spatial average have a significantly lower
+examines whether lower-tail dependence caused the spatial average to have significantly lower
 skewness than would have been expected without tail dependence. For
-leaf-curling plum aphid first flight data, because this dataset exhibited upper-tail 
+leaf-curling plum aphid data, because that dataset exhibited upper-tail 
 dependence (Table \ref{tab_multivar_summary}), we did the analogous one-tailed test
-in the right tail. The test examines whether upper-tail dependence made the spatial
-average of the first-flight data have significantly higher skewness than would have been
+in the right tail. The test examines whether upper-tail dependence caused the spatial
+average to have significantly higher skewness than would have been
 expected without tail dependence.
 
-We also examined whether the nature of the
-tail dependence of an environmental variable across space could influence the 
+We also examined whether spatial 
+tail dependence of an environmental variable could influence the 
 extinction risk of a metapopulation. We hypothesized that environmental noises
 exhibiting greater left-tail dependence across different habitat patches would 
-cause higher metapopulation extinction risks because then bad years for the 
+cause higher metapopulation extinction risks because then very bad years for the 
 component populations would occur simultaneously in many patches,
-reducing rescue effects. Here, we are assuming that low values of the environmental
+reducing rescue effects. Here we are assuming that low values of the environmental
 variable are "bad" for the populations being modelled, and high values are "good".
-We provided a test of the conceptual validity of this hypothesis using a metapopulation
+We tested the validity of this hypothesis using a metapopulation
 extension of the Lewontin-Cohen model,
@@
 $\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
-$\epsilon_i(t)$ represent environmental noises in each patch. They were standard-normally 
-distributed in each patch, were independent through time, and showed the same 
-spatial correlations (to within sampling error) for every simulation, 
+$\epsilon_i(t)$ represent environmental noises. They were standard-normally 
+distributed, were independent through time, and showed the same 
+spatial correlations for every simulation, 
 but were made to exhibit right- or left-tail dependence between patches in
@@
 to be arranged in a line, with dispersal occurring either to nearest neighbors only,
-or globally, in different simulations. The dispersal 
-rate was $d$ (Appendix \ref{SM-ext_risk_model}). After each time step, 
-if the population density in a patch was below $1$ (an extinction threshold), it was
+or globally, in different simulations (Appendix \ref{SM-ext_risk_model}). The dispersal 
+rate was $d$. After each time step, 
+if the population density in a patch was below $1$, it was
 set to $0$. We simulated the model 10000 times for each combination of parameters,
@@
 and calculated extinction risk after $25$ time steps.
+
+<!--***DAN: stopped here 2019 05 03 321pm-->
 

### 2019-05-03 SupportingInformation.Rmd
edits

--- 2019-05-02 SupportingInformation.Rmd
@@
              caption.short = "Model fitting results for the Cedar Creek dataset",
-             caption = "Results of fitting our 16 copula families to the Cedar Creek dataset. Analogous to Fig. \\ref{tab_soilCNfit}, see caption of that figure for details. \\label{tab_cc2000fit}",
+             caption = "Results of fitting our 16 copula families to the Cedar Creek dataset, year 2000.
+Analogous to Fig.
+\\ref{tab_soilCNfit}, see caption of that figure for details.
+\\label{tab_cc2000fit}",
              booktabs=T)
@@
 \includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_4_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data.
+\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data, year 2000.
 As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$.
@@
              caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
-             caption = "Cedar Creek dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
+             caption = "Cedar Creek dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis, year 2000.
 The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.25}-\\cor_{0.75,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.25}-\\Ps_{0.75,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.75,1}-\\Dsq_{0,0.25}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons.
@@
 the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
-correlation (the sorrogate algorithm had two versions, one for preserving each correlation coefficient), 
+correlation (the surrogate algorithm had two versions, one for preserving each correlation coefficient), 
 but had normal copula structure. Thus our comparisons tested the null hypothesis 
@@
 for 1000 Kendall-preserving surrogate datasets with N copulas. Due to relatively smaller sample size,
-for Cedar creek data we used a wider range of boundaries $(l_b,u_b)$ to compute $P_{0,0.25}-P_{0.75,1}$.
-For each data, rank of the said statistics with real-data value in 
+for Cedar Creek data we used $P_{0,0.25}-P_{0.75,1}$.
+For each dataset, the rank of the statistic for real-data in 
 the distribution of surrogate values was computed. 
 If ranks were $>975$ or $<25$, asymmetry of tail 
-dependence was significant compared to a N-copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
-for, respectively, the soil C and N data, bird masses and metabolic rates, mammal masses
-and metabolic rates, and the Cedar Creek data. Thus the first three datasets showed stronger 
-upper- than lower-tail dependence, and the Cedar Creek data which showed the reverse though not significantly with
-non-parametric analysis, confirming the 
-results based on model selection presented in the main text (see Table \ref{MT-tab_bivar_summary}, row 9 
-showed Cedar creek had relatively weak tail-dependence than the other three datasets).
+dependence was significant compared to an N-copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
+for, respectively, the soil C and N dataset, the bird body masses and metabolic rates dataset, 
+the mammal body masses and metabolic rates dataset, and the Cedar Creek data for the year 2000. 
+Thus the first three datasets showed stronger 
+upper- than lower-tail dependence, and the Cedar Creek data showed the reverse 
+(though not significantly), confirming the 
+results based on model selection presented in the main text (see 
+Table \ref{MT-tab_bivar_summary}, row 9). Model selection results for Cedar Creek showed 
+only slightly stronger lower- than upper-tail dependence, consistent with the non-significant 
+tendency toward lower-tail dependence also revealed by the non-parametric analysis for that
+dataset.
 

### 2019-05-06 Paper.Rmd
finished a round of editing the main text, we are now within Ecol Lett length limits

@@
 
-<!--***DAN: stopped here 2019 05 03 321pm-->
-
 # Results for Q3: Tail dependence influences skewness of spatial average, extinction risk, and Taylor's law
+
+\noindent Skewness results were generally consistent with hypotheses. For the datasets that
+exhibited strong lower-tail (respectively, upper-tail) 
+dependence in earlier analyses (namely, the green spruce 
+aphid abundance data for lower-tail, and the leaf-curling plum aphid first flight data
+for upper-tail, Table 
+\ref{tab_multivar_summary}), the skewness of the spatial average 
+was less than (respectively, greater than) a significant fraction of the
+surrogate skewness values 
+(Fig. \ref{fig_skewness_spearman}A,B). For the datasets with moderate lower-tail 
+dependence (\emph{Ceratium furca} abundance and methane data), the skewness of the spatial
+average showed a non-significant or marginally significant tendency 
+toward being less than surrogate skewnesses (Fig. \ref{fig_skewness_spearman}C,D).
 
@@
 
-\noindent Skewness results were generally consistent with hypotheses. For the datasets that
-exhibited strong lower-tail (respectively, upper-tail) 
-dependence in earlier analyses (namely, the green spruce 
-aphid abundance data for lower-tail, and the leaf-curling plum aphid first flight data
-for upper-tail, Table 
-\ref{tab_multivar_summary}), the skewness, through time, of the spatial average of the 
-empirical data was less than (respectively, greater than) a significant fraction (i.e.,
-more than $95\%$) of the
-surrogate skewness values 
-(Fig. \ref{fig_skewness_spearman}A,B). For the datasets that demonstrated a 
-more moderate lower-tail dependence in earlier analyses (the 
-\emph{Ceratium furca} abundance and the methane data), skewness of the spatial
-average showed a non-significant tendency toward being less than surrogate skewnesses
-(Fig. \ref{fig_skewness_spearman}C), or was less than a marginally significant 
-fraction of surrogate skewnesses (Fig. \ref{fig_skewness_spearman}D).
-
 As hypothesized, left-tail dependent environmental fluctuations increased metapopulation
@@
 (Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
-dispersal. We also examined the influence of non-N copula structure and tail dependence
-on Taylor's law in Appendices \ref{SM-Consequence_TL_simulation} and 
-\ref{SM-sect:copsurrognd}.
+dispersal. We also showed (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd})
+that Taylor's law can be strongly influenced by
+tail dependence between populations.
 
@@
 <!--what we did and why is it important to study?-->
-We have shown that non-N copula structure and asymmetric tail dependence are 
-exceedingly common in datasets that range across multiple sub-disciplines
-and levels of organization in ecology, in spite of the fact that no published 
-ecological research we are aware of has 
-taken this structure into account. Via simple mathematical models, we hypothesized two 
-mechanisms that may cause non-N copula structures with asymmetric tail dependence
-in ecological data. One
-mechanism was via a Moran-like effect by which the dependence structure between ecological
-variables (measured through time at different locations) is shaped by the dependence structure of 
-environmental variables (measured at the same times and locations). 
-A second mechanism gives rise to non-N copula structure and asymmetric tail dependence via
-the asymmetric or nonlinear influence of environmental variables on ecological variables, 
-and relates to Liebig's law of the minimum. We discuss below how commonly these and other
+We showed that non-N copula structure and asymmetric tail dependence are 
+exceedingly common in datasets across multiple sub-disciplines
+and levels of organization in ecology, although no published 
+research we are aware of has 
+taken this structure into account. Via simple models, we hypothesized three 
+mechanisms that may cause non-N copula structures. One
+mechanism was via a Moran-like effect, whereby dependencies between environmental variables 
+beget similar dependencies between ecological variables in different locations.
+A second mechanism produces tail dependence via nonlinear environmental influences
+on ecological variables. We discuss below how commonly these and other
 potential mechanisms may opperate. 
 We also demonstrated important consequences of 
-non-N and tail-dependent copula structure in ecology. First, the skewness, through time, 
-of the spatial average of time series from different locations (e.g., average regional pest abundance computed from local pest abundances) is affected by the tail dependence of the relationships
-between the local quantities. Thus right-tail-dependent local quantities 
-tend to produce "spiky" spatially averaged time series, with large outbreaks. 
-Second, metapopulation extinction risk can be accentuated or reduced by the tail dependence
-properties of environmental variables that influence the populations. Finally, 
-simulation results in Appendix \ref{SM-Consequence_TL_simulation} 
-showed that spatial Taylor's law can be strongly influenced by
-tail dependence between local populations. In our view, our results demonstrate that, for many
-ecological phenomena, complete understanding is impossible without 
-characterizing dependence between variables using copulas.
-
-<!--A para about what is available in the copula and VineCopula packages.
-Maybe make it a broader para about emthods, and state in there that we developed some new surrogate methods that may be broadly useful for researches trying to incorporate copula approaches.-->
+non-N and tail-dependent copula structure in ecology. First, the skewness
+of a spatial-average time series is influenced by tail dependencies between its 
+constituent time series. Thus right-tail-dependent local time series 
+lead to "spiky" spatially averaged time series, with large outbreaks.
+These observations could have implications for numerous exploited and pest species,
+for which the spatial average time series have applied importance.
+Second, metapopulation extinction risk can be accentuated or reduced by tail dependence
+properties of environmental variables. We also showed effects of copula 
+structure on Taylor's law, which is has been applied in many fields.
+In our view, our results demonstrate that comprehensive understanding 
+of many ecological phenomena is impossible without 
+a complete characterization of dependence between variables. 
 Copula approaches are
@@
 computer implementations exist (e.g., the `copula` and 
-`VineCopula` packages in R). Ecologists need only become aware of these tools
-and understand them conceptually to start applying them broadly. 
-We created a variety of interrelated randomization procedures
+`VineCopula` packages in R); ecologists can start applying these tools
+immediately. We created a variety of interrelated randomization procedures
 (Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
@@
 
-<!--some mechanism paragraphs follow-->
-
-<!--Maybe a para on how likely one is to get copula structure via the Moran mechanism - considers
-work on copula structure in environmental variables. Maybe mention possible future work: 
-study copula structure of environmental variables (such as temperature and precip) that are known 
-to be important in ecology, including how such copula structure may be changing and whether it 
-has a geography of the kind studied in Walter et al.-->
-One may expect our Moran-like mechanism for non-N copula structure and tail dependence to 
+<!--para on how likely one is to get copula structure via the Moran mechanism-->
+Our Moran-like mechanism for copula structure and tail dependence seems likely to
 operate commonly, because Moran effects are common [@Liebhold2004], and 
-non-N and tail-dependent copula structures have often been found in prior analyses of 
-environmental variables. A recent study [@agliardi2019relationship] found non-N 
-copulas (mainly Gumbel and Survival Clayton) were better fit in modelling the 
-upper tail dependence between extreme large values for global 
-temperature and annual mean green house gas concentrations for an extensively
-long datasets (1880-2017). If large meteorological events are also spatially widespread, then
+non-N and tail-dependent copula structures are often found in 
+environmental variables. 
+If large meteorological events are 
+also spatially widespread, then
 one would expect environmental variables associated with these events to take extreme
 values simultaneously across large spatial areas, producing tail dependence in measurements
-made at different locations. Non-extreme values, not associated with extreme events, may
-be more associated with local weather phenomena, and therefore 
+made through time at different locations. Non-extreme values, not associated with extreme, 
+spatially extensive weather events, may
+be more associated with local phenomena, and therefore 
 less prone to being correlated across large areas.
-Using a bivariate mixed model approach with copulas, 
-@Serinaldi2008 examined the spatial dependence 
-of rainfall observations (7 years with 30 minute time resolution) 
-within a rain gauge network in Central Italy. Gumbel or Student 2-copulas 
-were possible candidates for modelling bivariate dependence at several time scales, and 
-neither of these is an N copula. @liu2016probabilistic used a copula 
-based approach in assessing the drought risk with joint probability distribution 
-between vegetation and climatic variables across China. 
-In a majority of their cases, they found 
-non-N copula structure in the dependence between a vegatation index 
-and climatic variables (e.g., precipitation and temperature). 
-Another long-term study (1950–2014) in the Loess Plateau of China [@She2018]
-showed that a Gumbel copula (which has upper tail dependence) could effectively model 
-the spatial dependence of drought variables 
-(duration and severity) in the study area. Bivariate copula analysis was also used in forecasting the 
+For instance, @Serinaldi2008 examined the spatial dependence 
+of rainfall observations in Central Italy. G or Student 2-copulas 
+were possible candidates for modelling dependence, and 
+neither of these is an N copula.
+A long-term study (1950–2014) in the Loess Plateau of China [@She2018]
+showed that a G copula (which has upper tail dependence) could effectively model 
+the spatial dependence of drought variables in the study area.
+Bivariate copula analysis was also used in forecasting the 
 co-occurrence of extreme events (flood or drought) 
-over the North Sikkim Himalayas, India, using spatial data sets at different 
-time scales [@Goswami2018].
-
-<!--Maybe a para on how likely one is to get copula structure via the Liebig mechanism - considers
-what people have written on nonlinear effects and empirical evidence for Liebig's law and so on-->
-There are also good reasons to expect that our second mechanism operates commonly. 
+over the North Sikkim Himalayas, India, using spatial data sets [@Goswami2018].
+
+<!--a para on how likely one is to get copula structure via the Liebig mechanism-->
+There are also reasons to expect that our second mechanism operates commonly. 
 The second mechanism is the one that involves asymmetric or nonlinear influences of 
-environmental variables. Liebig's law of the minimum and the idea of a "limiting 
-nutrient" of organism or population growth are fundamental and dominant paradigms in ecology.
+environmental variables. Liebig's law and the idea of a "limiting 
+nutrient" of growth are dominant paradigms in ecology.
 Furthermore, innumerable studies document nonlinear influences of environmental
-variables on important ecological quantities at all levels of organization, e.g., the 
-influence of temperature on: 
-individual metabolism and autotroph growth rates 
-[@GarciaCarreras2018]; dispersal [@Oconnor2007]; population growth 
-[@Savage2004]; and microbial and ecosystem methane flux [@YvonDurocher2014].
-Environmental variables influence not only ecological populations, but also physiological,
-community, and ecosystem functioning measurements. So our two mechanism may well operate
-to produce spatial copula structure in a wide variety of ecological measurements.
+variables on ecological quantities at all levels of organization.
 
@@
 Our simulations of character evolution suggest the hypothesis that evolutionary
-changes through time in bird and mammal BMR and body size may exhibit right-tail dependence. This is
+changes through time in bird and mammal BMR and body size may exhibit right-tail dependence,
+contrary to standard assumptions of models of character evolution. This is
 a hypothesis only, because the right-tail dependence of Fig. \ref{fig_biv_multi_raw_cop}F,G
@@
 dependence in characters of extant species. For instance, as discussed below and in
-Appendix \ref{SM-sect:missingdata} (see also Fig. \ref{SM-fig:missingdata}), 
+Appendix \ref{SM-sect:missingdata} (also Fig. \ref{SM-fig:missingdata}), 
 systematically missing data can produce tail dependence and may
-have influenced results for the BMR versus body mass datasets. Even if the hypothesized 
+have influenced results for the BMR-body mass datasets. Even if the hypothesized 
 evolutionary mechanism
@@
 dependence between BMR and body mass for birds and mammals, with another question,
-i.e., why might we see right-tail dependence in evolutionary changes of these traits?
+i.e., why might we see right-tail dependence in evolutionary changes in these traits?
 See Appendix \ref{SM-evol_mech} for additional thoughts on the potential importance of 
-copulas for studies of character evolution when phylogeny branch lengths are not available.
+copulas for studies of character evolution.
 
@@
 mass itself. Trade-offs between life history characteristics are a focal topic in 
-evolutionary ecology. There are reasons to believe copulas interrelate with 
+ecology. Copulas probably interrelate with 
 life-history theory in additional ways beyond what we could demonstrate 
@@
 will reduce the energy that can be allocated to other functions, $Y_1$, $Y_2$, $Y_3$ 
-(e.g., growth, predation avoidance; this is sometimes called the principle of 
+(e.g., growth, predation avoidance; this is called the principle of 
 allocation).
@@
 little association - resources not allocated to X can instead be allocated 
-to $Y_1$ or $Y_2$ or to some combination. This constitutes 
+to $Y_1$, $Y_2$, $Y_3$ or to some combination. This constitutes 
 asymmetric tail dependence between X and $Y_1$. For instance, @Winemiller1992
@@
 if the degree of completeness of a dataset is associated with one or
-more of the quantities being studied then statistical artefacts
-may bias conclusions of copula analyses (Appendix \ref{SM-sect:missingdata}, 
+more of the characters being studied then statistical artefacts
+may bias conclusions (Appendix \ref{SM-sect:missingdata}, 
 Fig. \ref{SM-fig:missingdata}).
@@
 additional mechanisms of copula structure is certainly worthy of future work. -->
-Other mechanisms that we have not considered probably also operate to produce copula structure
-in ecological and evolutionary datasets, and should be considered in future work. For instance, 
-measurement error may modify copula structure, though one would have to have information
-on the precise nature of measurement error in a dataset to explore its influence. 
-Not only did the models we used to explore mechanisms not consider measurement error, 
-they were intentionally simple in other respects, as well, not including phenomena such as 
+Other mechanisms that we have not considered probably also operate, 
+and should be considered in future work. For instance, 
+measurement error may modify copula structure. 
+Our models were intentionally simple in other respects, as well, not including phenomena such as 
 delayed density dependence, dispersal, stage structure in population dynamics, and trophic
 interactions.
-We hope by enumerating a few potential mechanisms of copula structure in data 
-we help inspire additional research on the
+We hope by enumerating a few potential mechanisms of copula structure,
+we inspire additional research on the
 potentially numerous mechanisms that may operate in diverse datasets.
@@
 
+<!--***DAN: could remove the next para if you need to save words-->
 <!--a para on stability of the total biomass time series in a community, and skewness - sets up Shya's later paper as future work-->
@@
 Another potential application is time series of all species from a single community, e.g., all
-plants in a quadrat or survey site (REF) or all herbivorous plankton in a lake (REF). A huge 
-literature has focused on synchrony versus compensatory dynamics between such time series (REFS), and 
-the influence of such interspecific relationships on the variability, through time, of total-community
+plants in a quadrat surveyed repeatedly over time. A huge 
+literature has focused on synchrony versus 
+compensatory dynamics between such time series, and 
+the influence of such interspecific relationships on the 
+variability, through time, of total-community
 properties such as total biomass. Typically, variability through time of total-community
 properties is measured with the coefficient of variation, but the skewness of such a time series
-may also be of interest: it can characterize the extent to
-which community properties show "spikiness," i.e., uncharacteristically extreme values. 
-Therefore, future work on copula structure of intraspecific relationships in communities and its
-influence on community properties seems likely to us to be a potentially very exciting avenue.
-
+may also be of interest because it can characterize the extent to
+which community properties show "spikiness".  
+Future work on copula structure of interspecific relationships in communities and its
+implications for community stability seems likely to be an exciting avenue of research.
+
+<!--***DAN: could remove the next para if you need to save words-->
 <!--A para that sets up Shya's extinction risk paper-->
-Although we have demonstrated that tail dependence between environmental variables
-can influence extinction risk, substantial work remains to be done regarding the 
+Although we demonstrated that tail dependence between environmental variables
+can influence extinction risk, substantial work remains regarding the 
 circumstances under which this result holds, and its magnitude in different scenarios. 
 First, we used 
-a non-density dependent model to examine extinction risk. Does tail dependence
-in environmental variables still influence extinction risk when density dependence
+a non-density-dependent model to examine extinction risk. Do similar results pertain
+when density dependence
 plays a role? Second, we considered metapopulation extinction risk, but the large field 
@@
 PVA is one of the routes by which a species can be added to the influential
-International Union for the Conservation of Nature (IUCN) red list of threatened
-species. It seems worthwhile for future research to investigate the extent to which 
-copula structure in dependences between vital rates may occur and may influence
-estimates of risk. Finally, is the copula structure of environmental variables
+IUCN red list of threatened species. Finally, is the copula structure of environmental variables
 changing through time, and, if so, how do such changes influence extinction risks?
@@
 increased frequency of extreme weather events may correspond to changes in the tail dependence
-environmental variables measured in different places, and we showed such changes
-can affect extinction risk. It seems
-important for future work to explore whether such effects are of sufficient 
-commonness and magnitude to interest conservationists.
+of environmental variables measured in different places.
 

### 2019-05-06 Paper.Rmd
presentation format modified

@@
 
-\begin{figure}
+\begin{figure}[!h]
 \begin{center}
@@
 co-occurrence of extreme events (flood or drought) 
-over the North Sikkim Himalayas, India, using spatial data sets [@Goswami2018].
+over the North Sikkim Himalayas, India, using spatial datasets [@Goswami2018].
 

### 2019-05-06 SupportingInformation.Rmd
finished a round of editing the main text, we are now within Ecol Lett length limits
presentation format modified

@@
 
-<!--Do the model selection for methane flux data and then showing results in Tables-->
+<!--Do the model selection for methane-flux data and then showing results in Tables-->
 ```{r RES_methane, echo=F, cache=T, results="hide", cache.extra=list(seed,methane_data,mtime("FittingCopula_selective_loc.R"), mtime("vivj_matrix.R"),mtime("good_loclist.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
-<!--table for Best-fitting copulas, methane flux data-->
+<!--table for Best-fitting copulas, methane-flux data-->
 ```{r tab_methane_bestcop, echo=F, results="markup"}
@@
              format='latex',linesep = "",
-             caption.short = "Best-fitting copulas, methane flux data",
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane flux data.
+             caption.short = "Best-fitting copulas, methane-flux data",
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.
 See methods for Q1 for copula abbreviations.
@@
 
-<!--table for AIC-weights of best-fitting copulas, methane flux data-->
+<!--table for AIC-weights of best-fitting copulas, methane-flux data-->
 ```{r tab_methane_AICw,echo=F,results="markup"}
@@
              digits=3,
-             caption.short="AIC weights, best-fitting copulas, methane flux data",
-             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for methane flux data.\\label{tab_methane_AICw}",
+             caption.short="AIC weights, best-fitting copulas, methane-flux data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the best-fitting (lowest AIC) copula family for each location pair for methane-flux data.\\label{tab_methane_AICw}",
              booktabs=TRUE)
@@
 
-<!--table for AIC-weights of Normal copula used for fitting, methane flux data-->
+<!--table for AIC-weights of Normal copula used for fitting, methane-flux data-->
 ```{r tab_methane_AICw_normal,echo=F,results="markup"}
@@
              digits=3,
-             caption.short="AIC weights, N copulas, methane flux data",
-             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for methane flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
+             caption.short="AIC weights, N copulas, methane-flux data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for methane-flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
              booktabs=TRUE)
@@
 
-<!--table for p-values from goodness of fit test (Cramer-von Mises-based test) with best-fit copula, methane flux data-->
+<!--table for p-values from goodness of fit test (Cramer-von Mises-based test) with best-fit copula, methane-flux data-->
 ```{r tab_methane_p_CvM,echo=F,results="markup"}
@@
              digits=3,
-             caption.short="Goodness of fit of best-fitting copulas, methane flux data",
-             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for methane flux data.\\label{tab_methane_p_CvM}",
+             caption.short="Goodness of fit of best-fitting copulas, methane-flux data",
+             caption = "$P$-value results of a Cramer-von Mises-based test of the goodness of fit of the lowest-AIC copula family for each pair of locations for methane-flux data.\\label{tab_methane_p_CvM}",
              booktabs=TRUE)
@@
 
-<!--table for p-values from goodness of fit test (Kolmogorov-Smirnov-based test) with best-fit copula, methane flux data-->
+<!--table for p-values from goodness of fit test (Kolmogorov-Smirnov-based test) with best-fit copula, methane-flux data-->
 ```{r tab_methane_p_KS,echo=F,results="markup"}
@@
              digits=3,
-             caption.short="Goodness of fit of best-fitting copulas, methane flux data",
+             caption.short="Goodness of fit of best-fitting copulas, methane-flux data",
              caption = "$P$-value results of a Kolmogorov-Smirnov-based test of
 the goodness of fit of the lowest-AIC copula family for each pair of
-locations for methane flux data.\\label{tab_methane_p_KS}",
+locations for methane-flux data.\\label{tab_methane_p_KS}",
              booktabs=TRUE)
@@
 
-<!--table for model averaged lower tail dep., methane flux data-->
+<!--table for model averaged lower tail dep., methane-flux data-->
 ```{r tab_methane_LTdep_AICw,echo=F,results="markup"}
@@
              digits=3,
-             caption.short="Lower-tail dependence, methane flux data",
-             caption = "Model-averaged lower-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_LTdep_AICw}",
+             caption.short="Lower-tail dependence, methane-flux data",
+             caption = "Model-averaged lower-tail dependence statistics for each location pair for methane-flux data.\\label{tab_methane_LTdep_AICw}",
              booktabs=TRUE)
@@
 
-<!--table for model averaged upper tail dep., methane flux data-->
+<!--table for model averaged upper tail dep., methane-flux data-->
 ```{r tab_methane_UTdep_AICw,echo=F,results="markup"}
@@
              digits=3,
-             caption.short="Upper-tail dependence, methane flux data",
-             caption = "Model-averaged upper-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_UTdep_AICw}",
+             caption.short="Upper-tail dependence, methane-flux data",
+             caption = "Model-averaged upper-tail dependence statistics for each location pair for methane-flux data.\\label{tab_methane_UTdep_AICw}",
              booktabs=TRUE)
@@
 
-<!--table for model averaged lower- minus upper-tail dep., methane flux data-->
+<!--table for model averaged lower- minus upper-tail dep., methane-flux data-->
 ```{r tab_methane_LTmUT,echo=F,results="markup"}
@@
              digits=3,
-             caption.short="Lower- minus upper-tail dependence, methane flux data",
-             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for methane flux data.\\label{tab_methane_LTmUT}",
+             caption.short="Lower- minus upper-tail dependence, methane-flux data",
+             caption = "Model-averaged lower-tail dependence statistics minus model-averaged upper-tail dependence statistics for each location pair for methane-flux data.\\label{tab_methane_LTmUT}",
              booktabs=TRUE)
@@
 
-<!--Do the main computations for nonparametric analysis of methane flux data-->
+<!--Do the main computations for nonparametric analysis of methane-flux data-->
 ```{r stat_methane, echo=F, cache=T, cache=T,  cache.extra=list(seed,methane_data,methane_raw,latlon_methane,mtime("good_loclist.R"),mtime("NonParamStat.R"),mtime("vivj_matrix.R"),mtime("CopulaFunctions.R"), mtime("CopulaFunctions_flexible.R"))}
@@
 
-<!--Show the statistical results in a table for methane flux data-->
+<!--Show the statistical results in a table for methane-flux data-->
 ```{r tab_methane_resamp,echo=F,results="markup"}
@@
              col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
-             caption.short="Results for nonparametric statistics, methane flux data",
-             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for methane flux data.\\label{tab_methane_resamp}",
+             caption.short="Results for nonparametric statistics, methane-flux data",
+             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for methane-flux data.\\label{tab_methane_resamp}",
              booktabs=TRUE)
@@
 consistent with the results of Table \ref{MT-tab_multivar_summary} (row 9).
-For the methane flux data, confidence intervals were 
+For the methane-flux data, confidence intervals were 
 $(-0.008,0.019)$.
@@
 
-<!--Following chunk examine the consequence of non-Normal copula atructure on skewness using methane flux data-->
+<!--Following chunk examine the consequence of non-Normal copula atructure on skewness using methane-flux data-->
 ```{r skewness_methane_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,methane_data,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
@@
 
-\clearpage
-
 # References
+\setlength{\parindent}{-0.2in}
+\setlength{\leftskip}{0.2in}
+\setlength{\parskip}{1pt}
+\noindent
 

### 2019-05-09 Paper.Rmd
small changes
the commit right befor this one was the start of making changes to respond to Lawrence's comments and those of John Drake after the preproposal

@@
 
-updated on `r Sys.Date()`
-
 ```{r setup_Paper, echo=F}
@@
 
-\noindent A complete statistical description of dependence between variables for ecology and related fields
+The importance of a complete copula description of dependence between variables for ecology and related fields
+
+Complex, informative, and previously ignored dependencies propogate throughout ecology and can be revealed with copulas
+
+Complex, informative, and previously ignored dependencies propogate throughout ecology and are revealed using copulas
+
+Copulas reveal complex, informative, and previously ignored dependence
+structures between variables across ecological disciplines
+
+Copulas reveal complex, informative, and previously ignored dependence
+structures which in turn illuminate multiple ecological disciplines
+
+Copulas reveal complex, informative, and previously ignored dependence
+structures which in turn can illuminate every subdiscipline of ecology
+
+Copulas reveal complex, informative, and previously ignored dependence
+structures which in turn can illuminate every subfield of ecology
+
+Copulas reveal complex, informative and previously ignored dependence
+structure between variables throughout ecology
+
+The complexity of dependencies between ecological variables, previously
+ignored, is revealed by copulas and in turn illuminates multiple 
+disciplines of ecology
 
@@
 colnames(tab_data_info)<-c("Data","Category","No. of measurements we used", "References")
-tab_data_info$Data<-c("Soil C and N","Bird body masses and BMR", "Mammal body masses and BMR", "Cedar Creek data", 
-                      "Green spruce and other aphid species abundances", "Leaf-curling plum and other aphid first flight dates","Ceratium furca and other plankton taxa abundances","Methane flux")
-tab_data_info$Category<-c("Environmental","Species-trait","Species-trait","Community level",
-                          "Population","Phenological","Population","Ecosystem functioning")
+tab_data_info$Data<-c("Soil C and N","Bird body masses and BMR", "Mammal body masses and BMR", 
+                      "Green spruce and other aphid species abundances", "Leaf-curling plum and other aphid first flight dates","Ceratium furca and other plankton taxa abundances", "Plant diversity and aboveground biomass","Methane flux")
+tab_data_info$Category<-c("Environmental","Species-trait","Species-trait",
+                          "Population","Phenological","Population","Community level","Ecosystem functioning")
 tab_data_info$`No. of measurements we used`<-c(paste(nrow(d_soilCN)," locations",sep=""),
@@
                                                paste(nrow(d_mammalsBMR)," mammals",sep=""),
-                                               paste(nrow(d_cc[[5]])," plots",sep=""),
                                                "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
@@
                                                "14 locations with atleast 45 yrs. of timeseries for each 22 taxa",
+                                               paste(nrow(d_cc[[5]])," plots",sep=""),
                                                "13 locations with atleast 50 dates of data")
 tab_data_info$References<-c("US Rapid Carbon Assessment database (RaCA), Wills et al. (2014)", "McNab (2009)","McNab (2008)",
+                            "Rothamsted Insect Survey","Rothamsted Insect Survey","Continuous Plankton Recorder Survey",
                             "Cedar Creek Ecosystem Science Reserve, biodiversity experiment e120",
-                            "Rothamsted Insect Survey","Rothamsted Insect Survey","Continuous Plankton Recorder Survey",
                             "Great Miami Wetland Mitigation Bank, Smyth et al. (2019)")
@@
              column_spec(1:4,width="2.8 cm")%>%
-             row_spec(5:8, bold=T)
+             row_spec(c(4,5,6,8), bold=T)
 ```

### 2019-05-10 Paper.Rmd
PCR:09may2019-edits included

@@
 
-\noindent Reid: School of Biological & Marine Sciences, University of Plymouth, Drake Circus, Plymouth PL4 8AA, United Kingdom
+\noindent Reid: School of Biological & Marine Sciences, University of Plymouth, Drake Circus, Plymouth PL4 8AA, UK; Continuous Plankton Recorder Survey, The Marine Biological Association, The Laboratory, Citadel Hill, Plymouth PL1 2PB, UK
 
@@
 We here explore the concepts that underpin copulas and examine the potential for those concepts to improve our understanding of ecology.
-We find that informative copula structure in dependences between variables is exceedingly common across all the environmental, species-trait, phenological, population, community, and ecosystem functioning dataset we considered.
+We find that informative copula structure in dependences between variables is exceedingly common across all the environmental, species-trait, phenological, population, community, and ecosystem functioning datasets we considered.
 Many datasets exhibited asymmetric tail dependence, whereby two variables are more strongly related in their left compared to right tails, or vice versa.
@@
 We argue that benefits to ecology of using a complete 
-desription of dependence will be substantial.
+description of dependence will be substantial.
 
@@
 tab_data_info$Data<-c("Soil C and N","Bird body masses and BMR", "Mammal body masses and BMR", 
-                      "Green spruce and other aphid species abundances", "Leaf-curling plum and other aphid first flight dates","Ceratium furca and other plankton taxa abundances", "Plant diversity and aboveground biomass","Methane flux")
+                      "Green spruce and other aphid species abundances", "Leaf-curling plum and other aphid first flight dates","\\textit{Ceratium furca} and other plankton taxa abundances", "Plant diversity and aboveground biomass","Methane-flux")
 tab_data_info$Category<-c("Environmental","Species-trait","Species-trait",
@@
 knitr::kable(tab_data_info, 
-             format="latex", align="l",
+             format="latex", align="l",escape=F,
              caption = "Summary table for the data we used.
-Bolds entries are the multivariate data, the rest are bivariate datasets (see Data section).
+Bold entries are the multivariate data, the rest are bivariate datasets (see Data section).
 Basal metabolic rate=BMR.
@@
 dependence. To make results easier to absorb: we first 
-present results in for an example bivariate dataset; then we summarize results
+present results for an example bivariate dataset; then we summarize results
 for all bivariate datasets; then we present results for an example multivariate dataset; then we summarize for all multivariate datasets.
@@
 (Table \ref{SM-tab_count_LTmUT}) were positive 
-for all but a few locations pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
+for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
 out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
@@
 conclusions that non-N copula structure, non-N tail dependence, and asymmetric tail 
-dependence were common. Most locations pairs were non-independent 
+dependence were common. Most location pairs were non-independent 
 (Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
@@
 (Table \ref{tab_multivar_summary}, row 6). Some of the datasets (green spruce aphid
-abundance, \textit{Ceratium furca} abundance, methane flux) showed stronger 
+abundance, \textit{Ceratium furca} abundance, methane-flux) showed stronger 
 or marginally stronger lower- than upper-tail dependence 
@@
                         "Leaf-curling plum aphid first flight",
-                        "\\textit{Ceratium furca} abundance","Methane flux")
+                        "\\textit{Ceratium furca} abundance","Methane-flux")
              )%>%column_spec(2:5,width="2.5cm")%>%
@@
 properties of environmental variables. We also showed effects of copula 
-structure on Taylor's law, which is has been applied in many fields.
+structure on Taylor's law, which has been applied in many fields.
 In our view, our results demonstrate that comprehensive understanding 
@@
 which community properties show "spikiness".
-Future work on copula structure of interspecific relationships in communities and its
-implications for community stability seems likely to be an exciting avenue of research.
+Future work on copula structure of interspecific relationships in communities and its implications for community stability seems likely to be an exciting avenue of research.
 

### 2019-05-10 SupportingInformation.Rmd
PCR:09may2019-edits included

@@
 
-\textbf{Ecosystem functioning data on methane flux.} The Great Miami Wetland Mitigation Bank  was at 
+\textbf{Ecosystem functioning data on methane-flux.} The Great Miami Wetland Mitigation Bank  was at 
 $36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W, in Trotwood, Montgomery County, Ohio. Additional information 
@@
 
-<!--Load and format the data : methane flux-->
+<!--Load and format the data : methane-flux-->
 ```{r read_format_methane_data, echo=F}

### 2019-05-10 Paper.Rmd
editing to address Drake comments at presubmission

--- 2019-05-09 Paper.Rmd
@@
 ---
-title: "The importance of a complete description of dependence between variables for ecology and related fields"
+title: "Complex, informative, and previously ignored dependencies propogate throughout ecology and can be revealed using copulas"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
@@
 
-Complex, informative, and previously ignored dependencies propogate throughout ecology and can be revealed with copulas
+The importance of a complete description of dependence between variables for ecology and related fields
 
@@
 
-\noindent 241
+\noindent
 
@@
 
-\noindent 7 figures, 3 tables, 0 text boxes
+\noindent X figures, X tables, 0 text boxes
 
@@
 
-Very generally, all branches of ecology 
+All branches of ecology 
 study relationships among environmental and biological 
@@
 approaches to studying 
-such relationships are limited - such 
+such relationships are limited - standard 
 approaches provide only a very partial description of a relationship between variables. 
@@
 (respectively, Fig. \ref{fig_cop_pedag1}B) are principally 
-related in the left (respectively, right) portions of their distributions, an asymmetric 
-pattern of dependence that may
-have ecological significance, as discussed below, but is not captured by 
-correlation. Panels C and D 
-were obtained from A and B through monotonic transformation 
-(Appendix \ref{SM-pedagog1_details} for details),
-which alters the Pearson, but not the Spearman correlation. Although correlation
+related in the left (respectively, right) portions of their distributions, 
+an asymmetric pattern of dependence that can
+have profound ecological significance, as discussed below, but is not captured by 
+correlation. 
+Monotonic transformations can alter the Pearson, but not the Spearman
+correlation (Appendix \ref{SM-pedagog1_details}, 
+Fig. \ref{SM-fig_transf_pearson_spearman}). Although correlation
 is only one way to 
@@
 \includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
-\caption{(A-F, I, J) Bivariate datasets showing diverse relationships between 
+\caption{(A-D) Bivariate datasets showing diverse relationships between 
 variables; many of these datasets nevertheless have the same Pearson (P)
-or Spearman (S) correlation coefficients. Panels C and E (respectively, D and F) were 
-obtained from A (respectively, B) via monotonic functions that transformed the 
-standard-normal marginal distributions of A and B to gamma (C, D) or 
-uniform (E, F) marginals (details in Appendix \ref{SM-pedagog1_details}). (G, H) Contours 
-of log-scale density functions of Clayton (G) and survival-Clayton 
-(H) copulas with parameter 5 (see section on methods for Q1). (I,J) Reflections
-of E, F, respectively, about a vertical axis. Subscripts h and v 
+or Spearman (S) correlation coefficients. (C, D) Contours 
+of density functions of Clayton (C) and survival-Clayton 
+(D) copulas with parameter 5 (see section on methods for OH), together
+with samples from the corresponding distributions. (E,F) Reflections
+of C, D, respectively, about a vertical axis. (G) Contours of the
+density function of a normal copula (methods for OH) with 
+parameter chosen to
+have the same Spearman correlation as A-D, together with a sample.
+(H) The portion of the Spearman correlation within a band (diagonal lines
+show two bands) can be computed 
+for any band (methods for OH) to describe how dependence
+varies from one part of the two distributions to another.
+Subscripts h and v 
 stand for "horizontal" and "vertical". \label{fig_cop_pedag1}}
@@
 
-Well-developed approaches exist, and are 
+Well-developed approaches do exist, however, and are 
 applied widely in finance, neuroscience, bioinformatics and other fields, 
@@
 to be complete [@joe2014_dependence; @nelsen2006_copula]. 
-The goal of this paper is to illustrate the potential for 
-broader application of the concepts behind these approaches 
-in ecology. The approaches are based on 
+The approaches are based on 
 the statistical concept of the *copula*, which we here 
@@
 for instance,
-Fig. \ref{fig_cop_pedag1}E-F, which show the copula structure of 
-A-B, respectively. Ranking makes marginals of 
-the underlying random variables uniformly distributed, isolating the 
+Fig. \ref{fig_cop_pedag1}C-D (the points), which show the copula structures of 
+A-B, respectively. Ranking makes the marginal distributions of 
+the underlying random variables uniform, isolating the 
 dependence structure only.
@@
 functions (pdfs) corresponding to the copulas associated 
-with Fig. \ref{fig_cop_pedag1}E-F are in Fig.
-\ref{fig_cop_pedag1}G-H, respectively. 
+with Fig. \ref{fig_cop_pedag1}C-D are superimposed on those panels. 
 See @nelsen2006_copula and @joe2014_dependence
@@
 have stronger lower- or left-tail dependence than right- or upper-tail dependence
-(Fig. \ref{fig_cop_pedag1}A, E);
+(Fig. \ref{fig_cop_pedag1}A, C);
 and \emph{vice versa} if larger values are more strongly associated than smaller
-ones (Fig. \ref{fig_cop_pedag1}B, F). Datasets of the same 
+ones (Fig. \ref{fig_cop_pedag1}B, D). Datasets of the same 
 correlation can have a range of tail dependence 
 characteristics (Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see section on
-methods for Q1).
+methods for OH).
+
+The goal of this paper is to illustrate the potential for 
+broader application of copulas in ecology. 
+Our results provide initial support for the following 
+overall hypothesis: environmental, ecological and evolutionary processes 
+generate and transmit copula structure and tail dependence, with substantial 
+consequences for theoretical and conceptual understanding and for 
+applications across ecology; analyzing copula structure 
+provides more detailed tests of ecological theories and more revealing 
+desciptions of ecological datasets. This is a very broad hypothesis, and 
+even the substantial empirical and theoretical
+evidence we assembled here has only begun to test it. So we propose the hypothesis
+principally for further evaluation by the research community.
+The overall hypothesis is examined by assessing several more specific 
+hypotheses, described below and outlined in Fig. 
+\ref{MasterFigure}, all of which bear upon the overall hypothesis.
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7in]{./MasterFigure.pdf}
+\caption{Guide for hypotheses addressed in the text and how each instantiates the
+overall hypothesis. Labels are defined 
+in the text to refer to parts of the figure. CaH="causal hypothesis" (a hypothesis about 
+causes of copula structure), CoH="consequential hypothesis" (a hypothesis about 
+consequences of copula structure) and OH="observational hypothesis" (a hypothesis about 
+copula structure of data).\label{MasterFigure}}
+\end{center}
+\end{figure}
 
@@
 The portion of the distributions that dominate the association is visible in the copula
-approach, but not via standard techniques.
+approach, but not via standard techniques. We therefore frame "causal hypothesis" H1: 
+copulas of dependencies between environmental and 
+ecological variables will sensitively reveal 
+nonlinear, asymmetric and threshold effects for population, community, 
+and ecosystem functioning variables. Copulas between 
+environmental and ecological variables
+will commonly show tail dependence. This is labeled CaH1 in Fig. \ref{MasterFigure};
+we use the term "causal hypothesis" because it is about a prospective cause of 
+copula structure.
 
@@
 across quadrats should be negatively related, as in 
-Fig. \ref{fig_cop_pedag1}I-J. If X is plotted on the horizontal axis
-and Y is on the vertical, then Fig. \ref{fig_cop_pedag1}I 
+Fig. \ref{fig_cop_pedag1}E-F. If X is plotted on the horizontal axis
+and Y on the vertical, then Fig. \ref{fig_cop_pedag1}E 
 would suggest X is the dominant 
@@
 limiting factors other than X. 
-Alternatively, Fig. \ref{fig_cop_pedag1}J suggests that Y
-is the dominant competitor. The copula approach again reveals which portions of the
+Alternatively, Fig. \ref{fig_cop_pedag1}F suggests that Y
+is dominant. The copula approach again reveals which portions of the
 distributions are principally responsible for the association between variables, and this 
-information can help support or refute alternative causal hypotheses.
+information can help support or refute alternative causal hypotheses. We
+frame causal hypothesis H2 (CaH2 on Fig. \ref{MasterFigure}): copulas of abundances
+through time or space of two competing or otherwise interacting species will
+reflect the asymmetries in the interaction.
 
@@
 Synchrony is a phenomenon of great interest in 
-ecology [@Liebhold2004; @Sheppard2016; @walteretal2017].
+ecology [@Liebhold2004; @Sheppard2016; @walteretal2017]. We therefore frame 
+causal hypothesis 3 (CaH3 on Fig. \ref{MasterFigure}): Moran effects
+transmit copula structure in spatial dependence from environmental to 
+ecological variables, and hence copula structure and tail dependence will be 
+common features of spatial synchrony in ecological variables. We will later
+frame CaH4, which concerns character evolution.
 
@@
 and the total economic value of a fishery is the sum of local catches. 
-We will explore how details, captured by copulas, of the 
+We therefore frame "consequential hypothesis" H1 (CoH1 on Fig. \ref{MasterFigure}):
+details, revealed by copulas, of the 
 dependence between local quantities can influence the spatial 
-mean or sum, and how this may influence higher organizational levels in 
-ecology and human concerns. Illustrating the idea here, prior work demonstrated that the overall 
+mean or sum, and this may influence higher organizational levels in 
+ecology and human concerns. We use the term "consequential hypothesis"
+because it is about a prospective consequence of copula structure.
+Illustrating the idea here, prior work demonstrated that the overall 
 reliability of wind-generated electricity depends sensitively on details
@@
 On the other hand left-tail dependence in a pest species, even if overall 
-correlation were unchanged, would have much more benign effects.
-
-We approached our overall goal of exploring the potential of 
-copula approaches for ecology by addressing the 
-following specific questions. (Q1) Do datasets in ecology 
-have copula structure distinct from that of a Gaussian/normal distribution (here called 
-non-normal copula structure)? Do they show tail dependence distinct from that 
-of a normal distribution, and in particular do they show asymmetric tail 
-dependence? (Q2) If so, what are some possible causes/mechanisms of this 
-non-normal copula structure? And (Q3) what are the consequences of non-normal 
-copula structure and tail dependence for our understanding of ecology and for 
-applications? To address Q1, we analyze several datasets, including 
+correlation were unchanged, would have much more benign effects. We will
+later frame CoH2, about extinction risk, and CoH3, about Taylor's law.
+
+Our causal and consequential hypotheses imply, or rely on, an "observational
+hypothesis," OH, which we here frame: datasets in ecology will commonly have copula structure
+distinct from that of a Gaussian/normal distribution (here called 
+non-normal copula structure). They will commonly show tail dependence 
+distinct from that of a normal distribution. In particular they will
+show asymmetric tail dependence. We use the term "observational hypothesis"
+because it pertains to what copula structures will be observed in data.
+
+Because OH bears upon all the causal
+and consequential hypotheses, we test OH first, in sections 2-4 of the paper.
+We analyze several datasets, including 
 environmental, species-trait, phenological, population, community, and 
-ecosystem functioning data. We selected these datasets to span multiple 
-sub-fields and organizational levels of ecology. The second section of the 
-paper introduces the data, and the third and fourth sections use the data to 
-address Q1. The fifth and sixth sections, which address Q2, use simple models to 
-explore mechanisms. The seventh and eigth sections, which address Q3, use both
-data and models to explore a variety of consequences. The ninth section is the Discussion. 
-Copula approaches have been used to great consequence in neuroscience [@onken2009], 
-bioinformatics 
-[@Kim2008], medical research [@Emura2016], direct study of environmental 
+ecosystem functioning data, selected to span multiple 
+sub-fields and organizational levels of ecology. 
+Sections 5 and 6 test the causal hypotheses, and sections 7 and 8 test the
+consequential hypotheses. Section 9 is the Discussion.
+Copula approaches have been used to great effect in neuroscience [@onken2009], 
+bioinformatics [@Kim2008], medical research [@Emura2016], direct study of environmental 
 variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000]. 
@@
 We argue that benefits to ecology of using a complete 
-desription of dependence will be substantial.
+copula desription of dependence will be substantial.
 
@@
                                                paste(nrow(d_mammalsBMR)," mammals",sep=""),
-                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
-                                               "10 locations with atleast 30 yrs. of timeseries for each 20 sp.",
-                                               "14 locations with atleast 45 yrs. of timeseries for each 22 taxa",
+                                               "10 locations with at least 30-yr. timeseries for each 20 sp.",
+                                               "10 locations with at least 30-yr. timeseries for each 20 sp.",
+                                               "14 locations with at least 45-yr. timeseries for each 22 taxa",
                                                paste(nrow(d_cc[[5]])," plots",sep=""),
-                                               "13 locations with atleast 50 dates of data")
+                                               "13 locations with at least 50 dates of data")
 tab_data_info$References<-c("US Rapid Carbon Assessment database (RaCA), Wills et al. (2014)", "McNab (2009)","McNab (2008)",
@@
              caption = "Summary table for the data we used.
-Bolds entries are the multivariate data, the rest are bivariate datasets (see Data section).
+Bold entries are the multivariate data, the rest are bivariate datasets (see Data section).
 Basal metabolic rate=BMR.
@@
 
-# Background, concepts and methods for Q1: Detecting copula structure in data \label{Methods}
+# Background, concepts and methods for OH: Detecting copula structure in data \label{Methods}
 
@@
 
-We addressed Q1 principally via a model selection procedure in which the 
+We addressed OH principally via a model selection procedure in which the 
 copulas listed above were fitted to 
@@
 
-The first part of Q1 was answered for a bivariate dataset 
+The first part of OH was answered for a bivariate dataset 
 by comparing the AIC weights of the non-N copula
 families with that of the N family. 
-The second part of Q1 was answered by comparing 
+The second part of OH was answered by comparing 
 the model-averaged lower- and upper-tail dependence
-values to the lower- and upper-tail dependence values of an N copula (i.e., 0). The third part of Q1
+values to the lower- and upper-tail dependence values of an N copula (i.e., 0). The third part of OH
 was answered by comparing the difference of the model-averaged lower- and upper-tail 
@@
 provide greater 
-confidence in our answer to Q1. We used three 
+confidence in our answer to OH. We used three 
 statistics which quantify the extent to which 
@@
 
-# Results for Q1: Ecological datasets often have non-normal copula structure and asymmetric tail dependence \label{Results}
+# Results for OH: Ecological datasets often have non-normal copula structure and asymmetric tail dependence \label{Results}
 
@@
 normal copula had AIC weight essentially 0, answering the 
-first part of Q1 for these data. 
+first part of OH for these data. 
 Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
@@
 These values are distinct from what a normal copula would give, namely, 0, answering the second part 
-of Q1. The values also differed substantially 
-from each other, answering the third part of Q1. The numbers reflect stronger upper- than lower-tail
+of OH. The values also differed substantially 
+from each other, answering the third part of OH. The numbers reflect stronger upper- than lower-tail
 dependence, and this is also visible in the extreme upper-right
@@
 <!--bivariate summary-->
-Table \ref{tab_bivar_summary} summarizes Q1 results for the bivariate datasets. The two 
+Table \ref{tab_bivar_summary} summarizes OH results for the bivariate datasets. The two 
 variables were significantly related for all datasets 
@@
 The N copula was always a poor fit compared to the highest-AIC-weight copula 
-(Table \ref{tab_bivar_summary}, rows 3-4), answering the first part of Q1. 
+(Table \ref{tab_bivar_summary}, rows 3-4), answering the first part of OH. 
 Lower- and upper-tail dependence statistics often differed substantially from 0 
 (Table \ref{tab_bivar_summary}, rows 7-8) and from each other (Table \ref{tab_bivar_summary}, row 9), 
-answering the second and third parts of Q1. The first three datasets had stronger upper- 
+answering the second and third parts of OH. The first three datasets had stronger upper- 
 than lower-tail dependence, and the last dataset had the reverse, though weakly. 
@@
 larger than N-copula weights (Table \ref{SM-tab_count_AICw_normal}), answering 
-the first part of Q1. Goodness of fit tests 
+the first part of OH. Goodness of fit tests 
 (Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
@@
 from what an N copula would give (i.e., 0), and answering the second part of 
-Q1. These data have greater tail dependence 
+OH. These data have greater tail dependence 
 (lower and upper) than would be expected from an N-copula null hypothesis. 
@@
 out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
-than the spatial synchrony of outbreaks. This also answers the third part of Q1.
+than the spatial synchrony of outbreaks. This also answers the third part of OH.
 Nonparametric statistics verified that tail dependence was asymmetric for the green 
@@
 <!--multivariate summary-->
-Table \ref{tab_multivar_summary} summarizes Q1 results from
+Table \ref{tab_multivar_summary} summarizes OH results from
 multivariate data. Results supported the 
@@
              format="latex",
-             caption = "Summary of Q1 results for bivariate datasets.
+             caption = "Summary of OH results for bivariate datasets.
 The $p$-values in rows 5-6 are for the minimum-AIC copula.
@@
 <!--Question 2 analyses-->
-# Background, concepts and methods for Q2: Causes and mechanisms of non-normal copula structure \label{Causes}
+# Background, concepts and methods for CaH: Causes and mechanisms of non-normal copula structure \label{Causes}
 
 Having demonstrated that non-N copula structure and asymmetric tail dependence are 
-common, we addressed Q2 by exploring, using 
+common, we addressed CaH by exploring, using 
 models, three possible mechanisms. The first mechanism 
@@
 steps. We applied our model selection algorithms and 
-non-parametric statistics (methods for Q1) to these 
+non-parametric statistics (methods for OH) to these 
 outputs to discover if the model could
@@
 
-# Results for Q2: Moran-like effects and asymmetric dependencies produce non-normal copula structure\label{Results_Q2}
+# Results for CaH: Moran-like effects and asymmetric dependencies produce non-normal copula structure\label{Results_Q2}
 
@@
 for this model. The 
-AR(1) model and a Clayton copula were used (see methods section for Q2).
+AR(1) model and a Clayton copula were used (see methods section for CaH).
 Each point is the mean across 50 replicate simulations for which the same 
@@
 Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1} for
-the nonlinear models described in the methods for Q2.\label{fig_Cause4copula_scatter_C_kend_ar1}}
+the nonlinear models described in the methods for CaH.\label{fig_Cause4copula_scatter_C_kend_ar1}}
 \end{center}
@@
 <!--Question 3 analyses-->
-# Background, concepts and methods for Q3: Consequences of non-normal copula structure for our understanding of ecology \label{Consequences}
+# Background, concepts and methods for CoH: Consequences of non-normal copula structure for our understanding of ecology \label{Consequences}
 
@@
 
-# Results for Q3: Tail dependence influences skewness of spatial average, extinction risk, and Taylor's law
+# Results for CoH: Tail dependence influences skewness of spatial average, extinction risk, and Taylor's law
 

### 2019-05-10 SupportingInformation.Rmd
editing to address Drake comments at presubmission

--- 2019-05-06 SupportingInformation.Rmd
@@
 ---
-title: "Supporting Information : The importance of a complete description of dependence between variables for ecology and related fields"
+title: "Supporting Information : Complex, informative, and previously ignored dependencies propogate throughout ecology and can be revealed using copulas"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
@@
 
-# Transformations used in Fig. \ref{MT-fig_cop_pedag1} \label{pedagog1_details}
+# Transformation alters Pearson but not Spearman correlation \label{pedagog1_details}
 
 The data of Fig. \ref{MT-fig_cop_pedag1}A-B were generated to have standard normal marginals.
-Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
-of both variables by the cdf of the standard normal distribution, and then by the inverse of 
+Fig. \ref{fig_transf_pearson_spearman}A can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
+of both variables by the cummulative distribution function (cdf) of the standard normal distribution, and then by the inverse of 
 the cdf of the gamma distribution with shape and scale parameters 2. 
-The function `qgamma` in R was used.
+The functions `pnorm` and `qgamma` in R were used.
+Fig. \ref{fig_transf_pearson_spearman}B can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
+The Pearson correlations differ substantially after transformation, 
+but the Spearman correlations are unchanged (Figs \ref{MT-fig_cop_pedag1}A-B, \ref{fig_transf_pearson_spearman}).
+Fig. \ref{MT-fig_cop_pedag1}C (the points, not the contours) can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
+of both variables by the cdf of the standard normal distribution.
 Fig. \ref{MT-fig_cop_pedag1}D can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
-Fig. \ref{MT-fig_cop_pedag1}E can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
-of both variables by the cdf of the standard normal distribution.
-Fig. \ref{MT-fig_cop_pedag1}F can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=10 cm]{./Results/PedagogFig_SMExport.pdf}
+\caption[Transformation alters Pearson but not Spearman correlation]{Transformation alters Pearson but not Spearman correlation. 
+Displayed data are transformed from the data pictured on
+Fig. \ref{MT-fig_cop_pedag1}A-B. See Appendix \ref{pedagog1_details} for 
+details. P=Pearson, S=Spearman 
+correlation.\label{fig_transf_pearson_spearman}}
+\end{center}
+\end{figure}
 

### 2019-05-10 Paper.Rmd

@@
 knitr::kable(tab_data_info, 
-             format="latex", align="l",
+             #format="latex", align="l",
+             format="latex", align="l",escape=F,
              caption = "Summary table for the data we used.

### 2019-05-13 Paper.Rmd
still editing

@@
 ---
-title: "Complex, informative, and previously ignored dependencies propogate throughout ecology and can be revealed using copulas"
+title: "Complex and informative dependencies propagate throughout ecology and can be revealed using copulas"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
@@
 
+Complex, informative, and previously ignored dependencies propagate throughout ecology and can be revealed using copulas
+
+The complexity of ecological relationships is revealed by copulas
+
+The true complexity of ecological relationships is revealed by copulas
+
+The true complexity of ecological relationships is only revealed by copulas
+
 Copulas reveal complex, informative, and previously ignored dependence
+structures between variables across ecological disciplines
+
+Copulas reveal complex and informative dependence
 structures between variables across ecological disciplines
@@
 of density functions of Clayton (C) and survival-Clayton 
-(D) copulas with parameter 5 (see section on methods for OH), together
+(D) copulas with parameter 5 (see section on methods for the observational hypothesis), together
 with samples from the corresponding distributions. (E,F) Reflections
 of C, D, respectively, about a vertical axis. (G) Contours of the
-density function of a normal copula (methods for OH) with 
+density function of a normal copula (methods for the observational hypothesis) with 
 parameter chosen to
@@
 (H) The portion of the Spearman correlation within a band (diagonal lines
-show two bands) can be computed 
-for any band (methods for OH) to describe how dependence
+show two bands, the data in the left/lower band showing stronger
+dependence than the right/upper band) can be computed 
+for any band (methods for the observational hypothesis) to describe how dependence
 varies from one part of the two distributions to another.
@@
 dependence structure only.
-Mathematically, if $(X,Y)$ is a continuous random
-variable and the cumulative distribution functions (cdfs) of $X$ and $Y$ are 
-$F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the dependence structure
-of $(X,Y)$ is revealed by the random variable $(F_X(X),F_Y(Y))$.
-This random variable takes values in $[0,1] \times [0,1]$, and has 
-uniform marginal distributions. Probability density 
+Probability density 
 functions (pdfs) corresponding to the copulas associated 
@@
 characteristics (Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see section on
-methods for OH).
+methods for the observational hypothesis).
 
@@
 generate and transmit copula structure and tail dependence, with substantial 
-consequences for theoretical and conceptual understanding and for 
-applications across ecology; analyzing copula structure 
-provides more detailed tests of ecological theories and more revealing 
-desciptions of ecological datasets. This is a very broad hypothesis, and 
+consequences for theoretical and conceptual understanding, for inference, and for 
+applications across ecology. This is a broad hypothesis, and 
 even the substantial empirical and theoretical
-evidence we assembled here has only begun to test it. So we propose the hypothesis
-principally for further evaluation by the research community.
-The overall hypothesis is examined by assessing several more specific 
+evidence we present here has only begun to test it. So we hope our work stimulates further
+evaluation of the hypothesis by the research community.
+Our overall hypothesis is examined by assessing several more specific 
 hypotheses, described below and outlined in Fig. 
@@
 overall hypothesis. Labels are defined 
-in the text to refer to parts of the figure. CaH="causal hypothesis" (a hypothesis about 
-causes of copula structure), CoH="consequential hypothesis" (a hypothesis about 
-consequences of copula structure) and OH="observational hypothesis" (a hypothesis about 
-copula structure of data).\label{MasterFigure}}
+in the text to refer to parts of the figure. "Causal hypotheses" (i.e., hypotheses about 
+causes of copula structure) are labeled with letters A-D. "Consequential 
+hypothesis" (i.e., hypothesis about 
+consequences of copula structure) are labeled with letters X-Z. The "observational 
+hypothesis" (i.e., our hypothesis about 
+copula structure of data) is labeled with O.\label{MasterFigure}}
 \end{center}
@@
 relationship, only when it is limiting, in the left portions of the distributions. 
-The portion of the distributions that dominate the association is visible in the copula
-approach, but not via standard techniques. We therefore frame "causal hypothesis" H1: 
+The portions of the distributions that dominate the association is visible in the copula
+approach, but not via standard techniques. We therefore frame "causal hypothesis" A 
+(labeled in Fig. \ref{MasterFigure}): 
 copulas of dependencies between environmental and 
@@
 nonlinear, asymmetric and threshold effects for population, community, 
-and ecosystem functioning variables. Copulas between 
+and ecosystem functioning variables; copulas between 
 environmental and ecological variables
-will commonly show tail dependence. This is labeled CaH1 in Fig. \ref{MasterFigure};
-we use the term "causal hypothesis" because it is about a prospective cause of 
-copula structure.
-
-A second reason we may expect a copula approach to benefit 
-ecological research is that such an approach may help illuminate 
-the causal mechanisms between variables. For instance, if two species
-X and Y are strong competitors, abundances of the two species
-across quadrats should be negatively related, as in 
-Fig. \ref{fig_cop_pedag1}E-F. If X is plotted on the horizontal axis
-and Y on the vertical, then Fig. \ref{fig_cop_pedag1}E 
-would suggest X is the dominant 
-competitor: when X is abundant, Y is necessarily rare because it is suppressed; 
-whereas when X is rare, Y
-may be abundant, or may also be relatively rare due to 
-limiting factors other than X. 
-Alternatively, Fig. \ref{fig_cop_pedag1}F suggests that Y
-is dominant. The copula approach again reveals which portions of the
-distributions are principally responsible for the association between variables, and this 
-information can help support or refute alternative causal hypotheses. We
-frame causal hypothesis H2 (CaH2 on Fig. \ref{MasterFigure}): copulas of abundances
-through time or space of two competing or otherwise interacting species will
-reflect the asymmetries in the interaction.
-
-A third reason we may expect a copula approach to benefit ecological
+will commonly show tail dependence; such tail dependence will lead to 
+tail dependence across space in ecological variables.
+
+A second reason we may expect a copula approach to benefit ecological
 research is that prior work has demonstrated important copula 
@@
 ecology [@Liebhold2004; @Sheppard2016; @walteretal2017]. We therefore frame 
-causal hypothesis 3 (CaH3 on Fig. \ref{MasterFigure}): Moran effects
+causal hypothesis B (again labeled in Fig. \ref{MasterFigure}): Moran effects
 transmit copula structure in spatial dependence from environmental to 
 ecological variables, and hence copula structure and tail dependence will be 
-common features of spatial synchrony in ecological variables. We will later
-frame CaH4, which concerns character evolution.
-
-A fourth reason we may expect copula structure to be important
+common features of spatial synchrony in ecological variables.
+
+A third reason we may expect a copula approach to benefit 
+ecological research is that such an approach may help illuminate 
+the causal mechanisms between variables. For instance, if two species
+Sp1 and Sp2 are strong competitors, abundances of the two species
+across quadrats should be negatively related, as in 
+Fig. \ref{fig_cop_pedag1}E-F. If Sp1 is plotted on the horizontal axis
+and Sp2 on the vertical, then Fig. \ref{fig_cop_pedag1}E 
+would suggest Sp1 is the dominant 
+competitor: when Sp1 is abundant, Sp2 is necessarily rare because it is suppressed; 
+whereas when Sp1 is rare, Sp2
+may be abundant, or may also be relatively rare due to 
+limiting factors other than Sp1. 
+Alternatively, Fig. \ref{fig_cop_pedag1}F suggests that Sp2
+is dominant. The copula approach again reveals which portions of the
+distributions are principally responsible for the association between variables, and this 
+information can help support or refute alternative causal hypotheses. We
+frame causal hypothesis C (Fig. \ref{MasterFigure}): copulas of abundances
+through time or space of two competing or otherwise interacting species will
+reflect any asymmetries in the interaction. We will later
+frame causal hypothesis D, which concerns character evolution.
+
+Another reason we may expect copula structure to be important
 in ecology has to do with spatially aggregated or averaged quantities. 
@@
 and the total economic value of a fishery is the sum of local catches. 
-We therefore frame "consequential hypothesis" H1 (CoH1 on Fig. \ref{MasterFigure}):
+We therefore frame "consequential hypothesis" X (labeled on Fig. \ref{MasterFigure}):
 details, revealed by copulas, of the 
@@
 mean or sum, and this may influence higher organizational levels in 
-ecology and human concerns. We use the term "consequential hypothesis"
-because it is about a prospective consequence of copula structure.
+ecology and human concerns.
 Illustrating the idea here, prior work demonstrated that the overall 
@@
 On the other hand left-tail dependence in a pest species, even if overall 
-correlation were unchanged, would have much more benign effects. We will
-later frame CoH2, about extinction risk, and CoH3, about Taylor's law.
+correlation were unchanged, would have more benign effects. We will
+later frame consequential hypotheses Y, about extinction risk, and Z, about Taylor's law.
 
 Our causal and consequential hypotheses imply, or rely on, an "observational
-hypothesis," OH, which we here frame: datasets in ecology will commonly have copula structure
+hypothesis" (O in Fig. \ref{MasteFigure}), which we here frame: datasets in 
+ecology will commonly have copula structure
 distinct from that of a Gaussian/normal distribution (here called 
-non-normal copula structure). They will commonly show tail dependence 
-distinct from that of a normal distribution. In particular they will
-show asymmetric tail dependence. We use the term "observational hypothesis"
-because it pertains to what copula structures will be observed in data.
-
-Because OH bears upon all the causal
-and consequential hypotheses, we test OH first, in sections 2-4 of the paper.
+non-normal copula structure); they will commonly show tail dependence 
+distinct from that of a normal distribution; in particular they will
+show asymmetric tail dependence.
+
+Because the observational hypothesis bears upon all the causal
+and consequential hypotheses, we test it first, in sections 2-4 of the paper.
 We analyze several datasets, including 
@@
 
-# Background, concepts and methods for OH: Detecting copula structure in data \label{Methods}
-
-To make results interpretable
-to unfamiliar readers, we first introduce the copula families we used, and 
+# Background, concepts and methods for testing the observational hypothesis \label{Methods}
+
+We introduce the copula families we used, and 
 measures of tail dependence. 
-Conceptually, a bivariate copula is a probability density function (pdf)
+Conceptually, a bivariate copula is a pdf
 on the unit square with uniform maginals. 
-The copula is actually defined as the corresponding 
-cdf rather than as the pdf,
-but both functions contain essentially the same information. 
+The copula is actually defined instead as the corresponding 
+cummulative dstribution function (cdf), but both functions contain the same information.
+If $(X,Y)$ is a continuous random
+variable and the cdfs of $X$ and $Y$ are 
+$F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the dependence structure
+of $(X,Y)$ is revealed by the random variable $(F_X(X),F_Y(Y))$.
+This random variable takes values in $[0,1] \times [0,1]$, and has 
+uniform marginal distributions. 
 The normal copula family (henceforth abreviated N) is a one-parameter
@@
 bivariate normal distribtion, and controlling the degree of dependence 
-(Fig. \ref{fig_cop_pedag3}A-B for an example).
+(Fig. \ref{fig_cop_pedag1}G for an example).
 Left- and right-tail dependence are $0$ for the N copula, and this 
 serves as a benchmark for tail dependence in other copulas - positive values of our
-tail dependence measures correspond to dependence stronger than a normal
+tail dependence measures correspond to dependence stronger than an N
 copula. The Clayton and 
-survival-Clayton families (Fig. \ref{fig_cop_pedag1}G-H) are two more one-parameter 
+survival-Clayton families (Fig. \ref{fig_cop_pedag1}C-D) are two more one-parameter 
 families. The Clayton has positive lower-tail dependence ($0.87$ for
-the example of Fig. \ref{fig_cop_pedag1}G) and $0$ upper-tail dependence; the 
+the example of Fig. \ref{fig_cop_pedag1}C) and $0$ upper-tail dependence; the 
 survival-Clayton is a rotation of the Clayton by $180^\circ$ and has 
-positive upper-tail ($0.87$ in Fig. \ref{fig_cop_pedag1}H)
+positive upper-tail ($0.87$ in Fig. \ref{fig_cop_pedag1}D)
 and $0$ lower-tail dependence. For our model selection procedure (see below), we used 
 several one- and two-parameter families of copulas exhibiting a variety of 
-positive lower-tail dependence values 
-(Clayton, survival-Gumbel, survival-Joe, survival-BB6, henceforth
-abbreviated C, SG, SJ, SBB6) and upper-tail dependence values
-(survival-Clayton, Gumbel, Joe, BB6, henceforth abbreviated 
-SC, G, J, BB6). We also considered families that could exhibit positive upper- or lower-tail
-dependence, or both, depending on parameters (BB1, survival-BB1, BB7, survival-BB7, henceforth 
-abbreviated BB1, SBB1, BB7, SBB7). 
+lower- and upper-tail dependence characteristics 
+(Clayton, survival-Clayton, Gumbel, survival-Gumbel, Joe, survival-Joe, Frank,
+BB1, survival-BB1,
+BB6, survival-BB6, BB7, survival-BB7, BB8, survival-BB8, henceforth
+abbreviated C, SC, G, SG, J, SJ, F, BB1, SBB1, BB6, SBB6, BB7, SBB7, BB8, SBB8).
+Copulas are implemented in the `copula` and `VineCopula` packages for the 
+R language. Visualizations of the copulas we used are in 
+Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}. 
 If LT and UT denote 
 lower- and upper-tail dependence measures, their difference measures asymmetry of 
-tail dependence. We also used two families that were asymmetric in their non-tail parts, 
-but that had zero tail dependence (BB8, survival-BB8, abbreviated BB8 and SBB8), and
-a symmetric copula called 
-the Frank copula (abbreviated F; Fig. \ref{fig_cop_pedag3}C-D). 
-Copulas are implemented in the `copula` and `VineCopula` packages for the 
-R language. Visualizations of the copulas we used are in Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}. 
+tail dependence.
 See @nelsen2006_copula, @joe2014_dependence and the `VineCopula` documentation
-for more details.
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=3.5in]{./Results/PedagogFig3.pdf}
-\caption{Log-transformed pdfs (A, C) and samples from (B, D) example copulas.
-Shown are the normal (N) copula with parameter $0.8$ and the Frank (F) copula with 
-parameter $10$. See also Fig. \ref{fig_cop_pedag1}E-H for the Clayton (C) and  
-survival-Clayton (SC) copulas with parameter 5. Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}
-have visualizations of all the copulas we used.\label{fig_cop_pedag3}}
-\end{center}
-\end{figure}
-
-We addressed OH principally via a model selection procedure in which the 
+for further background.
+
+We tested the observational hypothesis (O in Fig. \ref{MasterFigure}) 
+principally via a model selection procedure in which the 
 copulas listed above were fitted to 
@@
 family or as a boundary case. If independence could be rejected ($0.05$ significance level), model 
-selection proceeded. Third, we fit the bivariate copula families listed previously to the 
+selection proceeded. Third, we fit the 16 bivariate copula families listed previously to the 
 normalized ranks via maximum likelihood, using the `BiCopEst` function in `VineCopula`, 
@@
 
-The first part of OH was answered for a bivariate dataset 
+The first part of the observational hypothesis was tested for a bivariate dataset 
 by comparing the AIC weights of the non-N copula
 families with that of the N family. 
-The second part of OH was answered by comparing 
+The second part was tested by comparing 
 the model-averaged lower- and upper-tail dependence
-values to the lower- and upper-tail dependence values of an N copula (i.e., 0). The third part of OH
-was answered by comparing the difference of the model-averaged lower- and upper-tail 
+values to the lower- and upper-tail dependence values of an N copula (i.e., 0). 
+The third part of the observational hypothesis
+was tested by comparing the difference of the model-averaged lower- and upper-tail 
 dependence values to the analogous difference for an N copula, again 0.
@@
 provide greater 
-confidence in our answer to OH. We used three 
+confidence in our tests of the observational hypothesis. We used three 
 statistics which quantify the extent to which 
@@
 $0 \leq l_b < u_b \leq 1$, define the lines $u+v=2l_b$ and 
-$u+v=2u_b$, which intersect the unit square (Fig. \ref{SM-fig_stat_image}). 
+$u+v=2u_b$, which intersect the unit square (Figs \ref{fig_cop_pedag1}H, 
+\ref{SM-fig_stat_image}). 
 Our statistics quantify the dependence
@@
 Larger values indicate stronger positive association
-between between the bounds.
+between the bounds.
 We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
@@
 to $0$ for large datasets.
-We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}). All code for this
+We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}). Code and data for this
 project is archived at \url{www.github.com/sghosh89/BIVAN}.
 
-# Results for OH: Ecological datasets often have non-normal copula structure and asymmetric tail dependence \label{Results}
+# Results of testing the observational hypothesis: Ecological datasets often have non-normal copula structure and asymmetric tail dependence \label{Results}
 
@@
 for birds, (C) the same for mammals, and (D) above-ground plant biomass vs.
-Shannon's index \textit{(H)} from Cedar Creek.
+Shannon's index (\textit{H}) from Cedar Creek.
 Bottom panels show the corresponding copula plots.
-See text for units used in upper panels.
+See Data for units used in upper panels.
 \label{fig_biv_multi_raw_cop}}
@@
 
-Across essentially all our datasets, copula structures were typically non-normal and 
+Across essentially all our datasets, copula structures were typically non-N and 
 commonly showed asymmetric tail 
-dependence. To make results easier to absorb: we first 
-present results for an example bivariate dataset; then we summarize results
-for all bivariate datasets; then we present results for an example multivariate dataset; then we summarize for all multivariate datasets.
+dependence: the observational hypothesis was supported (Fig. \ref{MasterFigure}). To make results easier to absorb, we first 
+present results for an example bivariate dataset, then we summarize results
+for all bivariate datasets, then we present results for an example multivariate dataset, then we summarize for all multivariate datasets.
 
@@
 The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
-(previous section) to the data and computed AIC and BIC weights and corresponding
+to the data and computed AIC and BIC weights and corresponding
 lower- and upper-tail statistics for the fitted distributions (Table \ref{SM-tab_soilCNfit}).
@@
 minimum-AIC copula, with AIC weight `r round(max(BivMS_res_Loecke$InfCritRes$AICw),2)`, whereas the 
-normal copula had AIC weight essentially 0, answering the 
-first part of OH for these data. 
+N copula had AIC weight essentially 0, confirming the 
+first part of the observational hypothesis for these data. 
 Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
 and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
-These values are distinct from what a normal copula would give, namely, 0, answering the second part 
-of OH. The values also differed substantially 
-from each other, answering the third part of OH. The numbers reflect stronger upper- than lower-tail
+These values are distinct from what an N copula would give, namely, 0, confirming the second part 
+of the observational hypothesis. The values also differed substantially 
+from each other, confirming the third part of the observational hypothesis. 
+The numbers reflect stronger upper- than lower-tail
 dependence, and this is also visible in the extreme upper-right
@@
 
-Although these results convincingly show non-normal copula structure, 
+Although these results convincingly show non-N copula structure, 
 we caution that the model-averaged tail-dependence statistics may have been biased 
@@
 <!--bivariate summary-->
-Table \ref{tab_bivar_summary} summarizes OH results for the bivariate datasets. The two 
+Table \ref{tab_bivar_summary} summarizes results of testing the observational hypothesis 
+using the bivariate datasets. The two 
 variables were significantly related for all datasets 
@@
 The N copula was always a poor fit compared to the highest-AIC-weight copula 
-(Table \ref{tab_bivar_summary}, rows 3-4), answering the first part of OH. 
+(Table \ref{tab_bivar_summary}, rows 3-4), confirming 
+the first part of the observational hypothesis. 
 Lower- and upper-tail dependence statistics often differed substantially from 0 
 (Table \ref{tab_bivar_summary}, rows 7-8) and from each other (Table \ref{tab_bivar_summary}, row 9), 
-answering the second and third parts of OH. The first three datasets had stronger upper- 
+confirming the second and third parts of the observational hypothesis. 
+The first three datasets had stronger upper- 
 than lower-tail dependence, and the last dataset had the reverse, though weakly. 
-Although these results convincingly show non-normal copula structure, the 
+Although these results convincingly show non-N copula structure, the 
 model-averaged tail-dependence statistics may have been biased 
@@
 
-We present green spruce aphid abundance results (Table \ref{tab_data_info})
+We present green spruce aphid abundance (Table \ref{tab_data_info}) results
 as an example multivariate analysis.
@@
 were typically substantially 
-larger than N-copula weights (Table \ref{SM-tab_count_AICw_normal}), answering 
-the first part of OH. Goodness of fit tests 
+larger than N-copula weights (Table \ref{SM-tab_count_AICw_normal}), confirming 
+the first part of the observational hypothesis. Goodness of fit tests 
 (Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
@@
 the data; the collection of 
-copula families we used was apparently sufficiently broad that all pairs of 
-locations could be characterized by at least one of them.
+copula families we used was sufficiently broad to characterize these data.
 Model-averaged lower- (Table \ref{SM-tab_count_LTdep_AICw}) and upper-tail
@@
 positive and were frequently large, thereby differing 
-from what an N copula would give (i.e., 0), and answering the second part of 
-OH. These data have greater tail dependence 
+from what an N copula would give (i.e., 0), and confirming the second part of 
+the observational hypothesis. These data have greater tail dependence 
 (lower and upper) than would be expected from an N-copula null hypothesis. 
@@
 out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
-than the spatial synchrony of outbreaks. This also answers the third part of OH.
+than the spatial synchrony of outbreaks. This also confirms the third part of the 
+observational hypothesis.
 Nonparametric statistics verified that tail dependence was asymmetric for the green 
@@
 <!--multivariate summary-->
-Table \ref{tab_multivar_summary} summarizes OH results from
+Table \ref{tab_multivar_summary} summarizes results of testing the observational hypothesis for
 multivariate data. Results supported the 
 conclusions that non-N copula structure, non-N tail dependence, and asymmetric tail 
-dependence were common. Most location pairs were non-independent 
+dependence were common, supporting the observational hypothesis. 
+Most location pairs were non-independent 
 (Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
-pairs had best-fitting copula that was not the N copula (Table \ref{tab_multivar_summary}, 
+pairs had best-fitting copulas that were not the N copula (Table \ref{tab_multivar_summary}, 
 row 3), and AIC weights of best-fitting copulas were 
@@
 \textit{Ceratium furca} (results not shown); conclusions supported that 
-non-N copula structure and tail dependence, and asymmetry of tail dependence, are 
+non-N copula structure and tail dependence and asymmetry of tail dependence are 
 common.
@@
 <!--Question 2 analyses-->
-# Background, concepts and methods for CaH: Causes and mechanisms of non-normal copula structure \label{Causes}
+# Background, concepts and methods for testing the causal hypotheses \label{Causes}
 
 Having demonstrated that non-N copula structure and asymmetric tail dependence are 
-common, we addressed CaH by exploring, using 
-models, three possible mechanisms. The first mechanism 
-is an extension of the well known Moran
-effect, and was summarized conceptually in the Introduction. 
-It explores the hypothesis that non-N copula structure and asymmetric tail dependence 
-between environmental variables, measured through time 
-in two places, may beget similarly 
-structured dependence between ecological variables measured through time in those places. 
+common, we addressed the causal hypotheses (Fig. \ref{MasterFigure}, A-D)
+using models. We did not systematically 
+test the first two parts of causal hypothesis A (Fig. \ref{MasterFigure})
+because nonlinear and thresold effects of 
+environmental variables on ecological variables are already well documented in
+the literature (however, see Discussion material about the green spruce aphid). To test 
+the third part of causal hypothesis A, we formulated a model in which 
+an environmental variable influences an ecological variable
+more strongly in one of its tails, and explored whether the ecological variable
+could then exhibit asymmetric tail dependence across space.
 Let $E_i(t)$ be an ecological variable 
 measured at location $i$ ($i=1,2$) at time $t$.
-We initially assume the AR(1) dynamics 
-$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
-The $\epsilon_i(t)$ represent environmental noise and are 
-standard-normal random variables that are independent for distinct times, $t$, but 
-that were made to exhibit different kinds of 
-dependence across locations in different simulations (see below).
-The variable, $E$, is very general, and could represent deviations
-of a population density from a carrying capacity, deviations of
-total plant community biomass from an average value,
-flux of a biogeochemical variable such as methane or CO$_2$, or other quantities.
-
-To explore the importance of nonlinear dynamics for our proposed mechanism, we also considered 
-nonlinear population models based on $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right) \right)+\sigma\epsilon_i(t))$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
-$\sigma=1$. When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
-$K$, and the nonlinearities of the model have limited influence on dynamics. When
-$\sigma=1$, model dynamics are strongly nonlinear. We refer to these as
-the weak-noise and strong-noise cases, though the importance of the noise here
-is that, when it is strong, it brings the nonlinearities of the model into play.
-
-For each of the three model setups above, $\tau=0.1,0.2,\ldots,0.9$, and for each of
-$50$ replicate simulations, we generated 5000 noise pairs 
-$(\epsilon_1(t),\epsilon_2(t))$ from the
-bivariate random variable with the given copula family and the given $\tau$.
-We then used this noise to drive the model, and retained both the
-noise and population values for the final 500 
-time steps. For each simulation, the following
-statistics were computed for both noises and populations: 
-Pearson, Spearman and Kendall correlations, $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, 
-$\Dsq_l$, $\Dsq_u$, $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$. 
-Values were plotted for noises and populations against $\tau$. If the hypothesis is
-correct that copula structure in environmental noises begets similar dependence 
-structure for ecological variables, 
-plots should be similar for populations and noises.
-
-Our second mechanism, also summarized in the 
-Introduction,
-relates to Liebig's law of the minimum and to nonlinear 
-influences of environmental variables on ecological variables. 
-If an environmental variable influences an ecological variable
-more strongly in one of its tails, we explored whether the ecological variable
-could then exhibit asymmetric tail dependence across space. 
-Let $E_i(t)$ again be an ecological variable, $i=1,2$. Assume the dynamics 
+Assume the dynamics 
 $E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where 
@@
 Because $(\epsilon_1,\epsilon_2)$
-and $(\delta_1,\delta_2)$ have N copula structure, the Moran-like 
-mechanism discussed previously does not operate here.
+and $(\delta_1,\delta_2)$ have N copula structure, the Moran 
+mechanism analyzed below (causal hypotheis B on Fig. \ref{MasterFigure}) 
+does not operate here.
 
@@
 steps. We applied our model selection algorithms and 
-non-parametric statistics (methods for OH) to these 
+non-parametric statistics (methods for the observational hypothesis) to these 
 outputs to discover if the model could
@@
 
-Our third mechanism is evolutionary, and pertains to bivariate 
+Causal hypothesis B is an extension of the well known Moran
+effect, and was summarized conceptually in the Introduction. 
+Let $E_i(t)$ again be the ecological variable, $i=1,2$.
+We initially assume the AR(1) dynamics 
+$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
+The $\epsilon_i(t)$ represent environmental noise and are 
+standard-normal random variables that are independent for distinct times, $t$, but 
+that were made to exhibit different kinds of 
+dependence across locations in different simulations (see below).
+The variable $E$ is very general, and could represent deviations
+of a population density from a carrying capacity, deviations of
+total plant community biomass from an average value,
+flux of a biogeochemical variable such as methane or CO$_2$, or other quantities.
+
+To explore the importance of nonlinear dynamics for causal hypothesis B, we also considered 
+nonlinear population models based on $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right) \right)+\sigma\epsilon_i(t))$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
+$\sigma=1$. When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
+$K$, and the nonlinearities of the model have limited influence on dynamics. When
+$\sigma=1$, model dynamics are strongly nonlinear. We refer to these as
+the weak-noise and strong-noise cases, though the importance of the noise here
+is that, when it is strong, it brings the nonlinearities of the model into play.
+
+For each of the three model setups above, $\tau=0.1,0.2,\ldots,0.9$, and for each of
+$50$ replicate simulations, we generated 5000 noise pairs 
+$(\epsilon_1(t),\epsilon_2(t))$ from the
+bivariate random variable with the given copula family and the given $\tau$.
+We then used this noise to drive the model, and retained both the
+noise and population values for the final 500 
+time steps. For each simulation, the following
+statistics were computed for both noises and populations: 
+Pearson, Spearman and Kendall correlations, $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, 
+$\Dsq_l$, $\Dsq_u$, $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$. 
+Values were plotted against $\tau$ for noises and populations. If causal hypothesis B is
+correct, plots should be similar for populations and noises.
+
+Causal hypothesis C was not tested. Our next 
+causal hypothesis, D on Fig. \ref{MasterFigure}, 
+is evolutionary, and pertains to bivariate 
 trait data across species, e.g., our bird and mammal 
-data. This mechanism is a hypothetical explanation for the right tail dependence
+data. This proposed mechanism is a hypothetical explanation for the right tail dependence
 observed in those data (Fig. \ref{fig_biv_multi_raw_cop}F,G). 
+The hypothesis is that tail dependence occurs in evolutionary changes
+in bivariate characters, and gives rise to tail dependence in observed character values
+across extant species.
 We simulated bivariate character evolution on an estimate of 
@@
 
-# Results for CaH: Moran-like effects and asymmetric dependencies produce non-normal copula structure\label{Results_Q2}
-
-The hypothesis that copula structure in ecological variables can arise through
-Moran-like effects was largely confirmed, though similarities between environmental- and 
-ecological-variable copula structure can be reduced when dynamics are strongly nonlinear.
+# Results of testing the causal hypotheses: Moran effects and asymmetric dependencies produce non-normal copula structure\label{Results_Q2}
+
+Our model with asymmetric environmental effects 
+could produce outputs with visually apparent asymmetry 
+of tail dependence, to an extent that depended on the value of $b$ used,
+thus confirming the third part of causal hypothesis A for the model we used. 
+For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ (Fig. \ref{SM-fig_ms_asym_b_-0.1}), 
+for both $g_1$ and $g_2$,
+asymmetry of tail dependence was strong; for 
+$b=\pm 0.5$, asymmetry of tail dependence was weaker but still apparent
+(Figs \ref{SM-fig_ms_asym_b_0.5} - 
+\ref{SM-fig_ms_asym_b_-0.5}).
+Lower-tail (respectively, upper-tail) spatial dependence in 
+the noise ($g_1$, respectively, $g_2$) 
+produced lower-tail (respectively, upper-tail) spatial dependence in the ecological variable.
+Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
+ecological outputs had non-N copula structure, but did not always reflect the 
+asymmetry of tail dependence, perhaps because those methods measured
+dependence in the most extreme tails. Results using our nonparametric statistics, however,
+reflected the visually apparent asymmetry (Tables \ref{SM-tab_stat_npa_asym}).
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
+\caption{If environmental effects operate asymmetrically in their tails on
+ecological variables, it can result in non-N copula structure and asymmetric 
+tail dependence across space in the ecological variables. Shown are the last
+$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text, $b=0.1$.\label{fig_ms_asym_b_0.1}}
+\end{center}
+\end{figure}
+
+Causal hypothesis B (Fig. \ref{MasterFigure}), i.e. that copula structure 
+in the spatial dependence of ecological variables can arise through
+Moran effects, was largely confirmed; however, similarities between environmental- and 
+ecological-variable copula structure were reduced when dynamics were strongly nonlinear.
 For simulations using AR(1) models, copula 
@@
 for this model. The 
-AR(1) model and a Clayton copula were used (see methods section for CaH).
+AR(1) model and a Clayton copula were used (see methods section for the causal hypotheses).
 Each point is the mean across 50 replicate simulations for which the same 
@@
 Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1} for
-the nonlinear models described in the methods for CaH.\label{fig_Cause4copula_scatter_C_kend_ar1}}
+the nonlinear models described in the methods for the causal hypotheses.\label{fig_Cause4copula_scatter_C_kend_ar1}}
 \end{center}
@@
 
-Our model with asymmetric environmental effects 
-could produce outputs with visually apparent asymmetry 
-of tail dependence, to an extent that depended on the value of $b$ used. 
-For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ (Fig. \ref{SM-fig_ms_asym_b_-0.1}), 
-for both $g_1$ and $g_2$,
-asymmetric tail dependence was strong; for 
-$b=\pm 0.5$, asymmetric tail dependence was weaker but still apparent
-(Figs \ref{SM-fig_ms_asym_b_0.5} - 
-\ref{SM-fig_ms_asym_b_-0.5}).
-Lower-tail (respectively, upper-tail) effects of the noise ($g_1$, respectively, $g_2$) 
-produced lower-tail (respectively, upper-tail) dependence in the ecological variable.
-Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
-ecological outputs had non-N copula structure, but did not always reflect the 
-asymmetry of tail dependence, perhaps because those methods measured
-dependence in the most extreme tails. Results using our nonparametric statistics, however,
-reflected the visually apparent asymmetry (Tables \ref{SM-tab_stat_npa_asym}).
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
-\caption{If environmental effects operate asymmetrically in their tails on
-ecological variables, it can result in non-N copula structure and asymmetric 
-tail dependence across space in the ecological variables. Shown are the last
-$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text, $b=0.1$.\label{fig_ms_asym_b_0.1}}
-\end{center}
-\end{figure}
-
-Our hypothesis was correct that asymmetric tail dependence in evolutionary changes 
+For the model we examined, causal hypothesis D (Fig. \ref{MasterFigure}) was 
+correct that asymmetric tail dependence in evolutionary changes 
 can produce similarly asymmetric tail dependence in character distributions across
@@
 <!--Question 3 analyses-->
-# Background, concepts and methods for CoH: Consequences of non-normal copula structure for our understanding of ecology \label{Consequences}
-
-An idea was presented in the Introduction that
+# Background, concepts and methods for testing the consequential hypotheses\label{Consequences}
+
+The hypothesis was presented in the Introduction that
 the distribution (through time) of a spatially averaged quantity should be influenced by the 
-dependence between the local quantities being averaged, including their
-copula structure and tail dependence. For instance, suppose an ecological variable 
+dependence between the local quantities, including their
+copula structure and tail dependence (consequential hypothesis 
+X on Fig. \ref{MasterFigure}). 
+To refine the hypothesis, suppose an ecological variable 
 $E_i(t)$ is measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and
@@
 exceptionally large values tend to occur at the same time in all or most locations.
-This should tend to increase the skewness of the distribution through time 
+We therefore hypothesize that this 
+should tend to increase the skewness of the distribution through time 
 of the spatial mean. Similarly, left-tail dependence between local variables
-will tend to decrease skewness of the spatial mean. Positive skewness of the 
+should tend to decrease skewness of the spatial mean. Strong positive skewness of the 
 spatial mean corresponds to "spikiness" of the spatial-mean time series. 
@@
 
-We tested these ideas using our multivariate datasets (Table \ref{tab_multivar_summary}).
+We tested consequential hypothesis X 
+using our multivariate datasets (Table \ref{tab_multivar_summary}).
 For each dataset we calculated the spatial mean time series, and then 
@@
 mean took values on the empirical data no different from what one would expect if the copula
-structure of the data were multivariate normal, but the data were otherwise statistically
+structure of the data were multivariate N, but the data were otherwise statistically
 unchanged. Significant differences indicate that non-N copula structure in the data 
@@
 
-For green spruce aphid abundance data, *Ceratium furca* abundance data, and methane-flux
+For green spruce aphid abundance data, *C. furca* abundance data, and methane-flux
 data, because these datasets exhibited lower-tail dependence 
@@
 
-We also examined whether spatial 
+We also examined consequential hypothesis Y on Fig. \ref{MasterFigure}, i.e., that spatial 
 tail dependence of an environmental variable could influence the 
@@
 component populations would occur simultaneously in many patches,
-reducing rescue effects. Here we are assuming that low values of the environmental
+reducing rescue effects. Here we are assuming, for simplicity, 
+that low values of the environmental
 variable are "bad" for the populations being modelled, and high values are "good".
-We tested the validity of this hypothesis using a metapopulation
+We tested the reasonableness of Y using a metapopulation
 extension of the Lewontin-Cohen model,
@@
 
-# Results for CoH: Tail dependence influences skewness of spatial average, extinction risk, and Taylor's law
-
-\noindent Skewness results were generally consistent with hypotheses. For the datasets that
+# Results of testing the consequential hypotheses: Tail dependence influences skewness of the spatial average, extinction risk, and Taylor's law
+
+\noindent Skewness results were generally consistent with consequential 
+hypothesis X. For the datasets that
 exhibited strong lower-tail (respectively, upper-tail) 
@@
 
-As hypothesized, left-tail dependent environmental fluctuations increased metapopulation
+Consistent with consequential hypothesis Y, 
+left-tail dependent environmental fluctuations increased metapopulation
 extinction risk for the spatially 
@@
 (Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
-dispersal. We also showed (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd})
-that Taylor's law can be strongly influenced by
+dispersal. We also substantiated using models (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd})
+a consequential hypothesis Z (Fig. \ref{MasterFigure}), that 
+Taylor's law can be strongly influenced by
 tail dependence between populations.
@@
 We showed that non-N copula structure and asymmetric tail dependence are 
-exceedingly common in datasets across multiple sub-disciplines
-and levels of organization in ecology, although no published 
+common in datasets across multiple sub-disciplines
+and levels of organization in ecology (Fig. \ref{MasterFigure}, O), although no published 
 research we are aware of has 
-taken this structure into account. Via simple models, we hypothesized three 
-mechanisms that may cause non-N copula structures. One
-mechanism was via a Moran-like effect, whereby dependencies between environmental variables 
+taken this structure into account. We hypothesized four 
+mechanisms (Fig. \ref{MasterFigure}, A-D) that may cause non-N copula structures. One
+mechanism (A) was via nonlinear environmental influences
+on ecological variables. A second mechanism (B) produces tail dependence via
+a Moran effect, whereby dependencies between environmental variables 
 beget similar dependencies between ecological variables in different locations. 
-A second mechanism produces tail dependence via nonlinear environmental influences
-on ecological variables. We discuss below how commonly these and other
+We discuss below how commonly these and other
 potential mechanisms may opperate. 
-We also demonstrated important consequences of 
-non-N and tail-dependent copula structure in ecology. First, the skewness
+We also hypothesized and demonstrated important consequences of 
+non-N and tail-dependent copula structure in ecology 
+(Fig. \ref{MasterFigure}, X-Z). First (X), the skewness
 of a spatial-average time series is influenced by tail dependencies between its 
@@
 for which the spatial average time series have applied importance.
-Second, metapopulation extinction risk can be accentuated or reduced by tail dependence
+Second (Y), metapopulation extinction risk can be accentuated or reduced by tail dependence
 properties of environmental variables. We also showed effects of copula 
-structure on Taylor's law, which has been applied in many fields.
-In our view, our results demonstrate that comprehensive understanding 
+structure on Taylor's law (Z), which has been applied in many fields.
+In our view, our results make our "overall hypothesis," framed in the Introduction, quite 
+reasonable and worthy of broader testing by the research community. We have illustrated 
+extensively how such tests might be carried out by performing them ourselves on 
+example datasets and models. Further suggestions future possible tests are below. 
+<!--***DAN: make sure you have enough of this, Dan-->
+It already seems likely that comprehensive understanding 
 of many ecological phenomena is impossible without 
-a complete characterization of dependence between variables. 
+a complete copula characterization of dependence between variables.
+
 Copula approaches are
@@
 computer implementations exist (e.g., the `copula` and 
-`VineCopula` packages in R); ecologists can start applying these tools
+`VineCopula` packages in R). Ecologists can start applying these tools
 immediately. We created a variety of interrelated randomization procedures
@@
 
+<!--a para on how likely one is to get copula structure via the Liebig mechanism-->
+Our first causal mechanism (Fig. \ref{MasterFigure}, A) seems likely to operate commonly.
+Liebig's law and the idea of a "limiting 
+nutrient" of growth are dominant paradigms in ecology.
+Although we did not systematically test the first two parts of causal 
+hypothesis A, many studies document nonlinear influences of environmental
+variables on ecological quantities at all levels of organization. 
+These parts of A could be tested
+straightforwardly for numerous datasets by examing the copulas of environmental and ecological
+variables (see below discussion about the green spruce aphid, and 
+Appendix \ref{SM-sect:greenspruce}, which illstrate how to do this).
+
+<!--***DAN: stoped here and went back to put in Shya/Lawrence recent results with 
+aphid counts and winter temp-->
+
 <!--para on how likely one is to get copula structure via the Moran mechanism-->
-Our Moran-like mechanism for copula structure and tail dependence seems likely to
+There are also reasons to expect that our Moran mechanism (Fig. \ref{MasterFigure}, B) for 
+copula structure and tail dependence will
 operate commonly, because Moran effects are common [@Liebhold2004], and 
@@
 over the North Sikkim Himalayas, India, using spatial datasets [@Goswami2018].
-
-<!--a para on how likely one is to get copula structure via the Liebig mechanism-->
-There are also reasons to expect that our second mechanism operates commonly. 
-The second mechanism is the one that involves asymmetric or nonlinear influences of 
-environmental variables. Liebig's law and the idea of a "limiting 
-nutrient" of growth are dominant paradigms in ecology.
-Furthermore, innumerable studies document nonlinear influences of environmental
-variables on ecological quantities at all levels of organization.
 
@@
 measurement error may modify copula structure. 
-Our models were intentionally simple in other respects, as well, not including phenomena such as 
+Our models were intentionally simple in other respects, 
+as well, not including phenomena such as 
 delayed density dependence, dispersal, stage structure in population dynamics, and trophic
@@
 <!--some consequences paragraphs next-->
+
+<!--***DAN: green spruce aphid all the way from causes to consequences-->
+Our hypotheses cover propagation of copula structure from causes, 
+to observable data, to consequences for human concerns.
+We here take a closer look at the green spruce aphid, which illustrates 
+all these stages in one example.
+Green spruce aphid abundance, as detected in the suction trap data we use (Data,
+Appendix \ref{SM-Data_details}),
+is known to be strongly influenced by temperature the previous winter (@Sheppard2016,
+supplementary fig. 6).
+For each of our 10 sampling locations, we examined (Appendix \ref{SM-sect:greenspruce}, 
+Fig. \ref{}) the copula of winter temperature
+and green spruce aphid abundance time series for the location, finding 
+substantial left-tail dependence: apparently winter temperature has an asymmetric
+influence on abundance in the sense that cold winters produce low abundances
+but warm winters do not yield especially higher abundances than moderate winters
+(in passing, this is evidence for the first two parts of causal mechanism A).
 

### 2019-05-13 SupportingInformation.Rmd
still editing

@@
 ---
-title: "Supporting Information : Complex, informative, and previously ignored dependencies propogate throughout ecology and can be revealed using copulas"
+title: "Supporting Information : Complex and informative dependencies propogate throughout ecology and can be revealed using copulas"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
@@
 \includegraphics[width=16 cm,height=8 cm]{./Results/copula_pedagog_fig_BB1.pdf}
-\caption[Lower- and upper-tail dependence can vary independently of correlation]{Lower- and upper-tail dependence can vary independently of correlation.
-(A-C) Data generated using the BB1 copula (see section on methods for Q1)
+\caption[Lower- and upper-tail dependence can vary independently of correlation]{Lower- and 
+upper-tail dependence can vary independently of correlation. (A-C) Data generated using the BB1 
+copula (see section on methods for the observational hypothesis) 
 with three different 
 sets of parameters chosen to produce a variety of 
-lower-tail (LT) and upper-tail (UT) dependence strengths (methods for Q1), but the same 
+lower-tail (LT) and upper-tail (UT) dependence strengths, but the same 
 Kendall correlation (K) (to two digits) in all cases. (D-F) Similar to A-C, but
@@
 
+# From causes to consequences for green spruce aphid abundance data\label{sect:greenspruce}
+Describe the winter temp data
+Show the copulas with winter temp, 10 locations
+
+<!--***DAN: Shya, please insert here a figure showing the 10 copulas between winter
+temp and green aphid count data as 10 panels. Insert the code that generates that figure 
+in an appropriate place so it is part of the overall workflow. Please use the label 
+fig:greenspruce, since that is what I refer to this fig by in the main text.-->
+
 # A means by which missing data can influence perceived copula structure\label{sect:missingdata}

### 2019-05-13 Paper.Rmd
green spruce aphid-winter temp copula figures, table added

@@
 For each of our 10 sampling locations, we examined (Appendix \ref{SM-sect:greenspruce}, 
-Fig. \ref{}) the copula of winter temperature
+Fig. \ref{SM-fig:greenspruce}) the copula of winter temperature
 and green spruce aphid abundance time series for the location, finding 

### 2019-05-13 SupportingInformation.Rmd
added another request for Shya, for a table on the green spruce aphid
green spruce aphid-winter temp copula figures, table added

@@
 
-# Aphid raw data for flight duration   
-#d0_flightduration<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidtsflightduration141117.csv",header=F))
-
-# Time series analysis aphids winter avg temp (here taken as -ve)
+# Time series analysis aphids winter avg temp 
 temp_aphid_raw<-as.matrix(read.csv("./Data/Aphid_data/APHID_DATA_RAW/aphidmonthlytempdata.csv",header = F))
@@
 
-aphid_winter_neg_temp<-vector("list",11) #initializing
-names(aphid_winter_neg_temp)<-rownames(winter_avg_temp)
+aphid_winter_temp<-vector("list",11) #initializing
+names(aphid_winter_temp)<-rownames(winter_avg_temp)
 
 for(i in 1:11){
-  aphid_winter_neg_temp[[i]]<-data.frame(Year=c(1976:2010),Dat=-winter_avg_temp[i,])
-}
-
-temp_aphid<-list(aphid_winter_neg_temp)
-names(temp_aphid)<-"aphid_winter_neg_temp"
-saveRDS(temp_aphid,"./Results/temp_aphid_winter_neg_temp_avg_data.RDS")
+  aphid_winter_temp[[i]]<-data.frame(Year=c(1976:2010),Dat=winter_avg_temp[i,])
+}
+
+temp_aphid<-list(aphid_winter_temp)
+
+#-------------------------------------------------------------------------
+# append winter avg temp data as sp 21 in aphid's count data
+
+data_aphid_count_with_temp<-append(data_aphid_count,temp_aphid)
+names(data_aphid_count_with_temp)[[21]]<-"winter_avg_temp"
+
+# create res folder
+if(!dir.exists("./Results/aphid_count_winter_temp_results")){
+  dir.create("./Results/aphid_count_winter_temp_results/")
+}
+
+saveRDS(data_aphid_count_with_temp,"./Results/aphid_count_winter_temp_results/data_aphid_count_with_temp.RDS")
 ```
@@
 # From causes to consequences for green spruce aphid abundance data\label{sect:greenspruce}
-Describe the winter temp data
-Show the copulas with winter temp, 10 locations
-
-<!--***DAN: Shya, please insert here a figure showing the 10 copulas between winter
-temp and green aphid count data as 10 panels. Insert the code that generates that figure 
-in an appropriate place so it is part of the overall workflow. Please use the label 
-fig:greenspruce, since that is what I refer to this fig by in the main text.-->
+
+For each locations (selected based on a threshold of 30 years of data, see \ref{Data_details}), 
+we computed yearwise average temperature of winter season (spanning over 4 months from December, previous 
+year to March, present year). With this winter temperature time series and green spruce aphid abundance time series data 
+a bivariate copula was made for each selected locations and shown in Fig. \ref{fig:greenspruce} whenever it was 
+significantly dependent with a p-value $<0.05$. Stronger lower tail dependence
+was visually prominent in those copulas. We used our model selection and non-parametric statistical tools 
+(see Table \ref{tab:greenspruce}) to confirm our hypothesis that winter temperature 
+was the main driver in "lower tail dependence" of the green
+spruce aphid abundance copula from any two location pairs (see Table \ref{tab_count_LTmUT}).
+
+```{r copulaplot_greenspruceaphid_with_temp,echo=F,results="hide"}
+source("./getcopula.R")
+source("./good_loclist.R")
+sp<-10 #only for green spruce aphid
+
+aphid_temp_coplist<-vector("list",1) 
+names(aphid_temp_coplist)<-c("sp10_Green_spruce_aphid")
+
+loclist<-good_loclist(d_allsp=data_aphid_count_with_temp,sp=sp,data_pt_thrs=30)
+
+pdf("./Results/aphid_count_winter_temp_results/copulaplot_greenspruceaphidcount_temp.pdf",height=4,width = 10)
+op<-par(mfrow=c(2,5),mar=c(3,3,3,3), mgp=c(1.5,0.5,0)) # each row for each sp for 2:11 locations
+
+  gsa_temp_coplist<-vector("list",10) # for 10 selective locations which had atleast 30 datapoints
+  names(gsa_temp_coplist)<-paste("loc",loclist,sep="")
+
+  for(iloc in c(1:length(loclist))){
+    loc<-loclist[iloc]
+    s1<-cbind(data_aphid_count_with_temp[[sp]][[loc]]$Dat,data_aphid_count_with_temp[[21]][[loc]]$Dat)
+    c1<-getcopula(d=s1,rankon=T,ploton=F)
+    vi<-c1[,1]
+    vj<-c1[,2]
+
+    IndepTestRes<-VineCopula::BiCopIndTest(vi,vj)$p.value
+    ct<-cor.test(vi,vj,alternative = "two.sided",method="kendall",exact=F)
+    tauval<-unname(ct$estimate)
+
+      if(IndepTestRes<0.05 && tauval>0){
+        plot(vi,vj,type='p',col=rgb(0,0,0,0.3),pch=19,xlim=c(0,1),ylim=c(0,1),
+             xlab=names(data_aphid_count_with_temp)[sp],ylab="winter temp.",cex.lab=1.5)
+        mtext(paste0("Loc = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
+      }else if(IndepTestRes<0.05 && tauval<0){
+        plot(vi,vj,type='p',col=rgb(0,1,0,0.3),pch=19,xlim=c(0,1),ylim=c(0,1),
+             xlab=names(data_aphid_count_with_temp)[sp],ylab="winter temp.",cex.lab=1.5)
+        mtext(paste0("Location = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
+      }else{
+        plot(-1,0,xlim=c(0,1),ylim=c(0,1),xlab=names(data_aphid_count_with_temp)[sp],ylab="winter temp.",cex.lab=1.5)
+        text(0.5,0.5,"Indep.",adj=c(0.5,.5),cex=2)
+        mtext(paste0("Location = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
+      }
+
+    gsa_temp_coplist[[iloc]]<-c1
+  }
+aphid_temp_coplist[[1]]<- gsa_temp_coplist
+par(op)
+dev.off()
+```
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/aphid_count_winter_temp_results/copulaplot_greenspruceaphidcount_temp.pdf}
+\caption[Copula plot for green spruce aphid abundance data with winter temperature]{Copula plot for green spruce 
+aphid abundance data with winter temperature for each selected locations. \label{fig:greenspruce}}
+\end{center}
+\end{figure}
+
+```{r ms_npa_greenspruceaphid_count_temp, echo=F, results="hide", cache=T, cache.extra=list(seed,aphid_temp_coplist,mtime("CopulaFunctions_flexible.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+
+set.seed(seed)
+source("./OurBiCopSelect.R")
+source("./CopulaFunctions_flexible.R")
+
+#====================== FUNCTION TO GET MS RESULTS FOR Green spruce aphid abundance with temp copula from all 10 locs===========================
+
+get_ms_aphid_temp<-function(sp,d_copsp){
+
+  cop_eachsp_allloc<-d_copsp[[sp]]
+
+  lengoodloc<-length(cop_eachsp_allloc)
+
+  # model selection with a copula (made of aphid-count and winter avg temp ts) from each location 2:11
+  ms_summary_allloc<-c()
+  ms_res_allloc<-vector("list",lengoodloc)
+  names(ms_res_allloc)<-paste("Loc",c(2:11),sep="")
+
+  for(loc in 1:lengoodloc){
+
+    cop_eachloc<-cop_eachsp_allloc[[loc]]
+    u1<-cop_eachloc[,1]
+    u2<-cop_eachloc[,2]
+
+    #----------------- model selection -----------------------------
+    ms_res<-OurBiCopSelect(u1=u1,u2=u2,families=c(1,3:10,13:14,16:20),
+                           level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
+                           gofnormal=F,status=F)
+    ms_res_allloc[[loc]]<-ms_res
+
+    m<-as.data.frame(matrix(NA,1,12))
+    colnames(m)<-c("p.Indep","minAIC.bestfit.cop","Bestfit.cop.AICw",
+                   "Normal.cop.AICw","p.CvM","p.KS",
+                   "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
+                   "CorlmCoru","PlmPu","D2umD2l")
+    m$p.Indep<-ms_res$IndepTestRes
+
+    if(ms_res$IndepTestRes<0.05){
+      #------------------------------ non-param analysis ---------------------------
+      #Cor stat
+      Corl<-Corbds(vi=u1,vj=u2,lb=0,ub=0.5)
+      Coru<-Corbds(vi=u1,vj=u2,lb=0.5,ub=1)
+      CorlmCoru<-Corl-Coru
+
+      #P stat
+      res_l<-Pbds(vi=u1,vj=u2,lb=0,ub=0.5)
+      Pl<-res_l$abs_res
+
+      res_u<-Pbds(vi=u1,vj=u2,lb=0.5,ub=1)
+      Pu<-res_u$abs_res
+      PlmPu<-Pl-Pu
+
+      #D2 stat
+      D2l<-D2bds(vi=u1,vj=u2,lb=0,ub=0.5)
+      D2u<-D2bds(vi=u1,vj=u2,lb=0.5,ub=1)
+      D2umD2l<-D2u-D2l
+
+      m$minAIC.bestfit.cop<-ms_res$InfCritRes$copname[which.min(ms_res$InfCritRes$AIC)]
+      m$Bestfit.cop.AICw<-round(max(ms_res$InfCritRes$AICw),3)
+      m$Normal.cop.AICw<-round(ms_res$InfCritRes$AICw[1],3)
+      m$p.CvM<-ms_res$GofRes_CvM
+      m$p.KS<-ms_res$GofRes_KS
+      m$LT.model.avg<-round(ms_res$relLTdep_AICw,3)
+      m$UT.model.avg<-round(ms_res$relUTdep_AICw,3)
+      m$`(LT-UT).model.avg`<-m$LT.model.avg-m$UT.model.avg
+      m$CorlmCoru<-CorlmCoru
+      m$PlmPu<-PlmPu
+      m$D2umD2l<-D2umD2l
+    }
+
+    ms_summary_allloc<-rbind(ms_summary_allloc,m)
+    #cat("loc=",loc,"=========ok=========","\n")
+
+  }
+  rownames(ms_summary_allloc)<-names(ms_res_allloc)
+  return(list(ms_res_allloc=ms_res_allloc,
+              ms_summary_allloc=ms_summary_allloc))
+}
+
+#===================== call above function for green spruce aphid abundance and winter temp data =================
+
+d_copsp<-aphid_temp_coplist
+splist<-c(1:length(d_copsp))
+ms_npa_RES_greenspruceaphid_count_temp<-vector("list",length(splist))
+names(ms_npa_RES_greenspruceaphid_count_temp)<-c("sp10_Green_spruce_aphid")
+
+for(isp in splist){
+  #cat("--------------sp =",isp,"-----------------------","\n")
+  res<-get_ms_aphid_temp(sp=isp,d_copsp=d_copsp)
+  ms_npa_RES_greenspruceaphid_count_temp[[isp]]<-res
+}
+
+saveRDS(ms_npa_RES_greenspruceaphid_count_temp,"./Results/aphid_count_winter_temp_results/ms_npa_RES_greenspruceaphid_count_temp.RDS")
+```
+
+```{r tab_greenspruceaphid_temp_summary, echo=F, results='asis',message=F}
+library(kableExtra)
+set.seed(seed)
+
+res<-ms_npa_RES_greenspruceaphid_count_temp$sp10_Green_spruce_aphid$ms_summary_allloc
+#row.names(res)<-NULL
+#res<-cbind(c(2:11),res)
+colnames(res)<-c("p-value from indep. test","bestfit copula",
+                 "bestfit copula AICw","Normal Copula AICw","p-value from CvM stats","p value from KS stats","LT dependence","UT dependence","(LT - UT) dependence","cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$")
+res<-res[,-1]
+knitr::kable(res,"latex", booktabs = T, linesep = "",digits=3, escape=F,longtable=T,
+             caption.short="Summary results for copula made with time-series of \\textbf{green spruce aphid} abundance data and winter temperature for each selected locations",
+      caption = "Summary results for copula made with time-series of green spruce aphid abundance data and winter temperature for each selected locations, tail dependence were calculated using model averaged AIC-weights, NA represents copula was independent at given location as p-value from independence test was beyond the threshold 0.05.
+Last three columns indicate non-parametric statistics computed on $(0,0.5)$ and $(0.5,1)$ bounds for lower and upper statistics, respectively.\\label{tab:greenspruce}") %>% column_spec(1:12,width="1.2 cm")#%>%
+#kable_styling(latex_options = c("scale_down"))
+```
 

### 2019-05-13 SupportingInformation.Rmd
minor edits

@@
 knitr::kable(res,"latex", booktabs = T, linesep = "",digits=3, escape=F,longtable=T,
-             caption.short="Summary results for copula made with time-series of \\textbf{green spruce aphid} abundance data and winter temperature for each selected locations",
+             caption.short="Summary results for copula made with time-series of green spruce aphid abundance data and winter temperature for each selected locations",
       caption = "Summary results for copula made with time-series of green spruce aphid abundance data and winter temperature for each selected locations, tail dependence were calculated using model averaged AIC-weights, NA represents copula was independent at given location as p-value from independence test was beyond the threshold 0.05.

### 2019-05-13 Paper.Rmd
more edits

--- 2019-05-13 Paper.Rmd
@@
 
-The importance of a complete copula description of dependence between variables for ecology and related fields
-
-The importance of a complete description of dependence between variables for ecology and related fields
-
-Complex, informative, and previously ignored dependencies propogate throughout ecology and are revealed using copulas
-
-Complex, informative, and previously ignored dependencies propagate throughout ecology and can be revealed using copulas
-
-The complexity of ecological relationships is revealed by copulas
-
-The true complexity of ecological relationships is revealed by copulas
-
-The true complexity of ecological relationships is only revealed by copulas
-
-Copulas reveal complex, informative, and previously ignored dependence
-structures between variables across ecological disciplines
-
-Copulas reveal complex and informative dependence
-structures between variables across ecological disciplines
-
-Copulas reveal complex, informative, and previously ignored dependence
-structures which in turn illuminate multiple ecological disciplines
-
-Copulas reveal complex, informative, and previously ignored dependence
-structures which in turn can illuminate every subdiscipline of ecology
-
-Copulas reveal complex, informative, and previously ignored dependence
-structures which in turn can illuminate every subfield of ecology
-
-Copulas reveal complex, informative and previously ignored dependence
-structure between variables throughout ecology
-
-The complexity of dependencies between ecological variables, previously
-ignored, is revealed by copulas and in turn illuminates multiple 
-disciplines of ecology
-
-\noindent \emph{Keywords:}
-
-\noindent dependence, regression, correlation, copula, population, community
-
-\noindent \emph{Type of Article:}
-
-\noindent Ideas and Perspectives
+Copulas and their importance for ecology
+
+Copulas and their importance throughout ecology
+
+Copulas and their potential for ecology
+
+Copulas and their potential throughout ecology
+
+Copulas and their importance
+
+Copulas in ecology
+
+Copulas
+
+\noindent \emph{Keywords:} dependence, regression, correlation, copula, population, community
+
+\noindent \emph{Type of Article:} Ideas and Perspectives
 
@@
 
-\noindent
-
 \noindent \emph{The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends):}
 
-\noindent \emph{The number of words in each text box:}
-
-\noindent No text boxes
+\noindent \emph{The number of words in each text box:} No text boxes
 
@@
 
-\noindent \emph{The number of figures, tables, and text boxes:}
-
-\noindent X figures, X tables, 0 text boxes
-
-\noindent \emph{The name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent:}
-
-\noindent Daniel Reuman, 2101 Constant Ave, Lawrence, KS, 66047, reuman@ku.edu, 626 560 7084.
+\noindent \emph{The number of figures, tables, and text boxes:} 7 figures, 3 tables, 0 text boxes
+
+\noindent \emph{The name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent:} Daniel Reuman, 2101 Constant Ave, Lawrence, KS, 66047, reuman@ku.edu, 626 560 7084.
 
@@
 approaches to studying 
-such relationships are limited - standard 
-approaches provide only a very partial description of a relationship between variables. 
+such relationships are limited, and 
+provide only a partial description of a relationship between variables. 
 For instance, datasets showing markedly different 
-relationships may nevertheless have the same correlation coefficient value 
+relationships may have the same correlation coefficient
 (Fig. \ref{fig_cop_pedag1}A-B). The variables of Fig. \ref{fig_cop_pedag1}A 
@@
 an asymmetric pattern of dependence that can
-have profound ecological significance, as discussed below, but is not captured by 
+have profound significance, as discussed below, but is not captured by 
 correlation. 
@@
 or Spearman (S) correlation coefficients. (C, D) Contours 
-of density functions of Clayton (C) and survival-Clayton 
+of probability density functions (pdf) of Clayton (C) and survival-Clayton 
 (D) copulas with parameter 5 (see section on methods for the observational hypothesis), together
@@
 of C, D, respectively, about a vertical axis. (G) Contours of the
-density function of a normal copula (methods for the observational hypothesis) with 
+pdf function of a normal copula (methods for the observational hypothesis) with 
 parameter chosen to
@@
 Well-developed approaches do exist, however, and are 
-applied widely in finance, neuroscience, bioinformatics and other fields, 
+applied widely in other fields, 
 that provide a description of the dependence 
@@
 The approaches are based on 
-the statistical concept of the *copula*, which we here 
-introduce. Copula approaches separate the information contained in a bivariate 
-random variable into two distinct parts: the information contained in the 
+the concept of the *copula*, which we here 
+introduce. Copula approaches separate the information of a bivariate 
+random variable into two distinct parts: the information in the 
 marginal distributions (which says nothing about the dependence between the 
@@
 a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
-Fig. \ref{fig_cop_pedag1}A-B), the associated dependence
+Fig. \ref{fig_cop_pedag1}A-B), the dependence
 structure is revealed by the bivariate plot of $u_i$ versus $v_i$, where
-$u_i$ is the rank of $x_i$ in the set of the $x_j$ ($j=1,\ldots,n$), divided by
+$u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
 $n+1$; and $v_i$ is defined in the same way but using the $y_i$. The $u_i$ and 
@@
 the underlying random variables uniform, isolating the 
-dependence structure only.
-Probability density 
-functions (pdfs) corresponding to the copulas associated 
-with Fig. \ref{fig_cop_pedag1}C-D are superimposed on those panels. 
+dependence structure.
 See @nelsen2006_copula and @joe2014_dependence
@@
 
-A principal benefit of a copula approach is 
+A main benefit of a copula approach is 
 that it can sensitively detect tail dependence and asymmetries of tail dependence. 
@@
 correlation can have a range of tail dependence 
-characteristics (Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see section on
+(Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see section on
 methods for the observational hypothesis).
@@
 The goal of this paper is to illustrate the potential for 
-broader application of copulas in ecology. 
+applications of copulas in ecology. 
 Our results provide initial support for the following 
@@
 evaluation of the hypothesis by the research community.
-Our overall hypothesis is examined by assessing several more specific 
+We examine our overall hypothesis by assessing several more specific 
 hypotheses, described below and outlined in Fig. 
@@
 hypothesis" (i.e., hypothesis about 
-consequences of copula structure) are labeled with letters X-Z. The "observational 
+consequences of copula structure) are labeled with X-Z. The "observational 
 hypothesis" (i.e., our hypothesis about 
-copula structure of data) is labeled with O.\label{MasterFigure}}
+copula structure of data) is labeled O.\label{MasterFigure}}
 \end{center}
@@
 We argue that benefits to ecology of using a complete 
-copula desription of dependence will be substantial.
+copula description of dependence will be substantial.
 
@@
              format="latex",
-             caption = "Summary of OH results for bivariate datasets.
+             caption = "Summary of results of testing the observational hypothesis for bivariate datasets.
 The $p$-values in rows 5-6 are for the minimum-AIC copula.
@@
 can produce similarly asymmetric tail dependence in character distributions across
-phylogeny tips (Appendix \ref{SM-evol_mech}). This topic is revisited in the Discussion.
+phylogeny tips (Appendix \ref{SM-evol_mech} fro details). This topic is revisited in the Discussion.
 
@@
 mechanisms (Fig. \ref{MasterFigure}, A-D) that may cause non-N copula structures. One
-mechanism (A) was via nonlinear environmental influences
-on ecological variables. A second mechanism (B) produces tail dependence via
+mechanism (Fig. \ref{MasterFigure}, A) was via nonlinear environmental influences
+on ecological variables. A second mechanism (Fig. \ref{MasterFigure}, B) produces tail dependence via
 a Moran effect, whereby dependencies between environmental variables 
@@
 non-N and tail-dependent copula structure in ecology 
-(Fig. \ref{MasterFigure}, X-Z). First (X), the skewness
+(Fig. \ref{MasterFigure}, X-Z). First (Fig. \ref{MasterFigure}, X), the skewness
 of a spatial-average time series is influenced by tail dependencies between its 
@@
 lead to "spiky" spatially averaged time series, with large outbreaks.
-These observations could have implications for numerous exploited and pest species,
+These observations could have implications for numerous exploited and pest species
 for which the spatial average time series have applied importance.
-Second (Y), metapopulation extinction risk can be accentuated or reduced by tail dependence
+Second (Fig. \ref{MasterFigure}, Y), metapopulation extinction risk can be accentuated or reduced by tail dependence
 properties of environmental variables. We also showed effects of copula 
-structure on Taylor's law (Z), which has been applied in many fields.
-In our view, our results make our "overall hypothesis," framed in the Introduction, quite 
+structure on Taylor's law (Fig. \ref{MasterFigure}, Z), which has been applied in many fields.
+In our view, our results make our "overall hypothesis," framed in the Introduction,
 reasonable and worthy of broader testing by the research community. We have illustrated 
 extensively how such tests might be carried out by performing them ourselves on 
-example datasets and models. Further suggestions future possible tests are below. 
+example datasets and models. Further suggestions for future possible tests are below. 
 <!--***DAN: make sure you have enough of this, Dan-->
@@
 Our first causal mechanism (Fig. \ref{MasterFigure}, A) seems likely to operate commonly.
-Liebig's law and the idea of a "limiting 
-nutrient" of growth are dominant paradigms in ecology.
+Liebig's law and the idea of a limiting 
+nutrient of growth are dominant paradigms in ecology.
 Although we did not systematically test the first two parts of causal 
@@
 These parts of A could be tested
-straightforwardly for numerous datasets by examing the copulas of environmental and ecological
+straightforwardly for numerous datasets by examing copulas describing dependencies
+between environmental and ecological
 variables (see below discussion about the green spruce aphid, and 
 Appendix \ref{SM-sect:greenspruce}, which illstrate how to do this).
-
-<!--***DAN: stoped here and went back to put in Shya/Lawrence recent results with 
-aphid counts and winter temp-->
+Causal mechanism C, which we also did not test, could be tested straightforwardly
+by analyzing copulas of abundances of competing species, across space or
+through time.
 
@@
 values simultaneously across large spatial areas, producing tail dependence in measurements
-made through time at different locations. Non-extreme values, not associated with extreme, 
+made through time at different locations. Non-extreme values, not associated with 
 spatially extensive weather events, may
@@
 less prone to being correlated across large areas.
-For instance, @Serinaldi2008 examined the spatial dependence 
-of rainfall observations in Central Italy. G or Student 2-copulas 
+@Serinaldi2008 examined the spatial dependence 
+of rainfall observations in Central Italy; G or Student 2-copulas 
 were possible candidates for modelling dependence, and 
@@
 have influenced results for the BMR-body mass datasets. Even if the hypothesized 
-evolutionary mechanism
+evolutionary mechanism (Fig. \ref{MasterFigure}, D)
 is correct, our results simply replace one question, i.e., why do we see right-tail
@@
 allocation).
-But it is equally well known that X can trade-off against any or all of 
+But it is equally well known that X can trade off against any or all of 
 the $Y_i$. Therefore, for large values of X which approach absolute 
@@
 little association - resources not allocated to X can instead be allocated 
-to $Y_1$, $Y_2$, $Y_3$ or to some combination. This constitutes 
+to $Y_1$, $Y_2$, $Y_3$ or to any combination thereof. This constitutes 
 asymmetric tail dependence between X and $Y_1$. For instance, @Winemiller1992
@@
 variables for later-maturing
-fishes because those fishes may invest much or little in juvenile survival.
+fishes because such fishes may invest much or little in juvenile survival.
 Although copulas may be very useful for studying multi-dimensional trade-offs, 
@@
 
-<!--***DAN: green spruce aphid all the way from causes to consequences-->
+<!--green spruce aphid all the way from causes to consequences-->
 Our hypotheses cover propagation of copula structure from causes, 
@@
 Appendix \ref{SM-Data_details}),
-is known to be strongly influenced by temperature the previous winter (@Sheppard2016,
+is strongly influenced by temperature the previous winter (@Sheppard2016,
 supplementary fig. 6).
-For each of our 10 sampling locations, we examined (Appendix \ref{SM-sect:greenspruce}, 
-Fig. \ref{}) the copula of winter temperature
+For each of our 10 sampling locations, we therefore 
+examined (Appendix \ref{SM-sect:greenspruce}, 
+Fig. \ref{SM-fig:greenspruce}, Table \ref{SM-tab:greenspruce}) the copula of winter temperature
 and green spruce aphid abundance time series for the location, finding 
@@
 influence on abundance in the sense that cold winters produce low abundances
-but warm winters do not yield especially higher abundances than moderate winters
-(in passing, this is evidence for the first two parts of causal mechanism A).
+but warm winters do not yield especially higher abundances than moderate winters.
+Although we did not systematically test the first two 
+parts of causal hypothesis A due to much existing evidence, this result 
+illustrates how parts one and two of the hypothesis can be tested using copula
+techniques. The third part of causal hypothesis A, which our modelling results
+supported, now suggests spatial dependence between green spruce aphid counts
+should exhibit left-tail dependence - exactly what was observed 
+(Table \ref{tab_multivar_summary}). And the consequences of that spatial tail 
+dependence for skewness of spatially averaged aphid counts was described already 
+(Fig. \ref{fig_skewness_spearman}A). Thus the asymmetric influence of winter 
+temperature ultimately causes spatially averaged green spruce aphid abundances to have
+lower skew than they would otherwise, because negative impacts of cold winter
+are spatially synchronous but positive impacts of warmer winters and other factors
+are less spatially or temporally consistent.
 
@@
 compensatory dynamics between such time series, and 
-the influence of such interspecific relationships on the 
-variability, through time, of total-community
+the influence of interspecific relationships on the 
+variability of total-community
 properties such as total biomass. Typically, variability through time of total-community
 properties is measured with the coefficient of variation, but the skewness of such a time series
-may also be of interest because it can characterize the extent to
-which community properties show "spikiness".
-Future work on copula structure of interspecific relationships in communities and its implications for community stability seems likely to be an exciting avenue of research.
+may also be of interest because it can help characterize "spikiness". Future work on copula 
+structure of interspecific relationships in communities and its implications for community 
+stability seems likely to be an exciting avenue of research.
 
@@
 Although we demonstrated that tail dependence between environmental variables
-can influence extinction risk, substantial work remains regarding the 
-circumstances under which this result holds, and its magnitude in different scenarios. 
-First, we used 
+can influence extinction risk, substantial work remains to determine the magnitude 
+of this effect in different scenarios. First, we used 
 a non-density-dependent model to examine extinction risk. Do similar results pertain

### 2019-05-13 SupportingInformation.Rmd
more edits

--- 2019-05-13 SupportingInformation.Rmd
@@
 
-# Verifying Q1 results for bivariate datasets using nonparametric statistics \label{Q1bivarnonparam}
+# Verifying model selection results of testing the observational hypothesis on bivariate datasets using nonparametric statistics \label{Q1bivarnonparam}
 
@@
 Kendall (or Spearman)
-correlation. Thus the comparison helps answer the first two parts of Q1
+correlation. Thus the comparison addresses the first two parts of the observational hypothesis
 for bivariate datasets. 
@@
 as computed on the empirical data, to the distribution of its values on many surrogate datasets 
-provides a test of asymmetry of dependence. Thus this comparison helps answer the
-third part of Q1 for bivariate 
+provides a test of asymmetry of dependence. Thus this comparison helps address the
+third part of the observational hypothesis for bivariate 
 datasets.
@@
 considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
-survival Joe, survival BB1, survival BB6, survival BB7, and survival BB8 copulas, with abbreviations in the methods for Q1.
+survival Joe, survival BB1, survival BB6, survival BB7, and survival BB8 copulas, with abbreviations in the methods for the observational hypothesis.
 NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
@@
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for leaf-curling plum aphid first flight data.
-Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations in the methods for Q1.
+Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations in the methods for the observational hypothesis.
 NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.
@@
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for \\textit{Ceratium furca} abundance data.
-See methods for Q1 for copula abbreviations.
+See methods for the observational hypothesis for copula abbreviations.
 Indep.=Independent.
@@
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.
-See methods for Q1 for copula abbreviations.
+See methods for the observational hypothesis for copula abbreviations.
 Indep.=Independent.
@@
 
-# Verifying Q1 results for multivariate datasets using nonparametric statistics \label{Q1multvarnonparam}
+# Verifying model selection results of testing the observational hypothesis on multivariate datasets using nonparametric statistics \label{Q1multvarnonparam}
 
@@
 or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
-N-copula null hypothesis, and answering question Q1 from the Introduction.
+N-copula null hypothesis.
 Bootstrapping was performed by resampling the locations of measurement, with replacement. 
@@
 copula structure as measured by the listed statistic differs significant
-between noise inputs and model outputs. See methods section for Q2 
+between noise inputs and model outputs. See methods section for the causal hypotheses 
 for additional details.\label{fig_Cause4copula_C_kend_ar1}}
@@
 The output of simulations from the second model detailed in 
-the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
+the methods section for the causal hypotheses (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
 \\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to the model 
 selection procedures detailed in the 
-methods section for Q1. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
+methods section for the observational hypothesis. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
 N-copula AICw: the AIC weight of the N copula family; Model-avg. LT: the lower-tail 
@@
 The output of simulations from the second model detailed in 
-the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
+the methods section for the causal hypotheses (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
 \\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
-analyses decribed in the methods section for Q1. The listed statistic was computed for the
+analyses decribed in the methods section for the observational hypothesis. The listed statistic was computed for the
 output of the model with the indicated b and g, and the value was compared to 1000 values
 of the same statistic computed on surrogate time series randomized to have N-copula structure
-(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
+(see methods for the observational hypothesis). A table entry $<X$ (respectively, $>X$) indicates the value of the 
 given statistic on the data was less (respectively, more) than its value on $X$ of the 
@@
 
-# From causes to consequences for green spruce aphid abundance data\label{sect:greenspruce}
-Describe the winter temp data
-Show the copulas with winter temp, 10 locations
+# From causes to consequences for green spruce aphid abundance, details \label{sect:greenspruce}
+Winter temperature for year $t$ was for December of year $t-1$ to March of 
+year $t$. Data, which pertained to the same locations and years as the aphid data,
+were obtained and processed exactly as in [@Sheppard2016]. Copulas
+with winter temperature are in Fig. \ref{fig:greenspruce}. Results of analyses of these
+copulas that paralleled the analyses applied to bivariate ecological datasets 
+(see section on methods for the observational hypotheses) are in Table \ref{tab:greenspruce}.
 
@@
 fig:greenspruce, since that is what I refer to this fig by in the main text.-->
+
+<!--***DAN: Shya, please insert here a table like Table 2 in your recent doc "explore_aphid_abund_wintertemp_dep". Please label it tab:greenspruce. Note that 
+currently there are two NA values in the first column, but there should be none.-->
 

### 2019-05-13 SupportingInformation.Rmd

--- 2019-05-13 SupportingInformation.Rmd
@@
 
-# From causes to consequences for green spruce aphid abundance data\label{sect:greenspruce}
-
-For each locations (selected based on a threshold of 30 years of data, see \ref{Data_details}), 
-we computed yearwise average temperature of winter season (spanning over 4 months from December, previous 
-year to March, present year). With this winter temperature time series and green spruce aphid abundance time series data 
-a bivariate copula was made for each selected locations and shown in Fig. \ref{fig:greenspruce} whenever it was 
-significantly dependent with a p-value $<0.05$. Stronger lower tail dependence
-was visually prominent in those copulas. We used our model selection and non-parametric statistical tools 
-(see Table \ref{tab:greenspruce}) to confirm our hypothesis that winter temperature 
-was the main driver in "lower tail dependence" of the green
-spruce aphid abundance copula from any two location pairs (see Table \ref{tab_count_LTmUT}).
+# From causes to consequences for green spruce aphid abundance, details \label{sect:greenspruce}
+Winter temperature for year $t$ was for December of year $t-1$ to March of 
+year $t$. Data, which pertained to the same locations and years as the aphid data,
+were obtained and processed exactly as in [@Sheppard2016]. Copulas
+with winter temperature are in Fig. \ref{fig:greenspruce}. Results of analyses of these
+copulas that paralleled the analyses applied to bivariate ecological datasets 
+(see section on methods for the observational hypotheses) are in Table \ref{tab:greenspruce}.
 

### 2019-05-13 Paper.Rmd
about to try to switch the order of some sup mat figs, committing in case somethig goes wrong

@@
 Our causal and consequential hypotheses imply, or rely on, an "observational
-hypothesis" (O in Fig. \ref{MasteFigure}), which we here frame: datasets in 
+hypothesis" (O in Fig. \ref{MasterFigure}), which we here frame: datasets in 
 ecology will commonly have copula structure
@@
 can produce similarly asymmetric tail dependence in character distributions across
-phylogeny tips (Appendix \ref{SM-evol_mech} fro details). This topic is revisited in the Discussion.
+phylogeny tips (Appendix \ref{SM-evol_mech} for details). This topic is revisited in the Discussion.
 

### 2019-05-13 SupportingInformation.Rmd
about to try to switch the order of some sup mat figs, committing in case somethig goes wrong

@@
 Winter temperature for year $t$ was for December of year $t-1$ to March of 
-year $t$. Data, which pertained to the same locations and years as the aphid data,
-were obtained and processed exactly as in [@Sheppard2016]. Copulas
-with winter temperature are in Fig. \ref{fig:greenspruce}. Results of analyses of these
+year $t$. Temperature data, which pertained to the same locations and years as the aphid data,
+were obtained and processed exactly as in @Sheppard2016. Copulas
+describing the relationship between green spruce aphid abundance and winter 
+temperature are in Fig. \ref{fig:greenspruce}. Results of analyses of these
 copulas that paralleled the analyses applied to bivariate ecological datasets 
@@
         plot(vi,vj,type='p',col=rgb(0,0,0,0.3),pch=19,xlim=c(0,1),ylim=c(0,1),
-             xlab=names(data_aphid_count_with_temp)[sp],ylab="winter temp.",cex.lab=1.5)
-        mtext(paste0("Loc = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
+             xlab="Aphid abundance",ylab="winter temp.",cex.lab=1.5)
+        mtext(paste0("Location = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
       }else if(IndepTestRes<0.05 && tauval<0){
         plot(vi,vj,type='p',col=rgb(0,1,0,0.3),pch=19,xlim=c(0,1),ylim=c(0,1),
-             xlab=names(data_aphid_count_with_temp)[sp],ylab="winter temp.",cex.lab=1.5)
+             xlab="Aphid abundance",ylab="winter temp.",cex.lab=1.5)
         mtext(paste0("Location = ",loc," , p = ",round(IndepTestRes,3)),side = 3, line=0.15, adj=0.5)
@@
 \includegraphics[width=16 cm]{./Results/aphid_count_winter_temp_results/copulaplot_greenspruceaphidcount_temp.pdf}
-\caption[Copula plot for green spruce aphid abundance data with winter temperature]{Copula plot for green spruce 
-aphid abundance data with winter temperature for each selected locations. \label{fig:greenspruce}}
+\caption[Copula plots for winter temperature versus green spruce aphid abundance]{Copula plots 
+for winter temperature versus green spruce 
+aphid abundance for each of the 10 sampling locations. The $p$-values in the header are from the test of independence (see section on methods for the observational hypothesis).\label{fig:greenspruce}}
 \end{center}
@@
 #res<-cbind(c(2:11),res)
-colnames(res)<-c("p-value from indep. test","bestfit copula",
-                 "bestfit copula AICw","Normal Copula AICw","p-value from CvM stats","p value from KS stats","LT dependence","UT dependence","(LT - UT) dependence","cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$")
+colnames(res)<-c("p-value, independence test","Best-fit copula",
+                 "Best-fit copula AICw","N-copula AICw","p, CvM","p, KS","Model-avg. LT","Model-avg. UT","Model-avg. LT minus UT","cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$")
 res<-res[,-1]
 knitr::kable(res,"latex", booktabs = T, linesep = "",digits=3, escape=F,longtable=T,
-             caption.short="Summary results for copula made with time-series of green spruce aphid abundance data and winter temperature for each selected locations",
-      caption = "Summary results for copula made with time-series of green spruce aphid abundance data and winter temperature for each selected locations, tail dependence were calculated using model averaged AIC-weights, NA represents copula was independent at given location as p-value from independence test was beyond the threshold 0.05.
-Last three columns indicate non-parametric statistics computed on $(0,0.5)$ and $(0.5,1)$ bounds for lower and upper statistics, respectively.\\label{tab:greenspruce}") %>% column_spec(1:12,width="1.2 cm")#%>%
+             caption.short="Summary results for copulas for winter temperature versus green spruce aphid abundance",
+      caption = "Summary results for analysis of copulas for winter temperature versus green spruce 
+aphid abundance for each of the 10 sampling locations. The column labeled p, CvM is the p-value 
+result for the goodness of fit test using the Cramer-von Mises statistic (see section on 
+methods for the observational hypothesis). The 
+column labeled p, KS is the p-value 
+result for the goodness of fit test using the Kolmogorov-Smirnov statistic. Entries with an NA are because model selection and other statistics
+were only employed for copulas for which independence was rejected (5 percent significance level).\\label{tab:greenspruce}") %>% column_spec(1:12,width="1.2 cm")#%>%
 #kable_styling(latex_options = c("scale_down"))

### 2019-05-13 Paper.Rmd
more editing

@@
 can produce similarly asymmetric tail dependence in character distributions across
-phylogeny tips (Appendix \ref{SM-evol_mech} for details). This topic is revisited in the Discussion.
+phylogeny tips (Appendix \ref{SM-evol_mech}, Fig. \ref{SM-fig_cop_evo} for details). This topic 
+is revisited in the Discussion.
 
@@
 (Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
-dispersal. We also substantiated using models (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd})
+dispersal. We also substantiated using models (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd}, Fig. \ref{SM-fig_spTL})
 a consequential hypothesis Z (Fig. \ref{MasterFigure}), that 
@@
 common in datasets across multiple sub-disciplines
-and levels of organization in ecology (Fig. \ref{MasterFigure}, O), although no published 
+and levels of organization in ecology, although no published 
 research we are aware of has 
 taken this structure into account. We hypothesized four 
-mechanisms (Fig. \ref{MasterFigure}, A-D) that may cause non-N copula structures. One
-mechanism (Fig. \ref{MasterFigure}, A) was via nonlinear environmental influences
-on ecological variables. A second mechanism (Fig. \ref{MasterFigure}, B) produces tail dependence via
-a Moran effect, whereby dependencies between environmental variables 
-beget similar dependencies between ecological variables in different locations. 
-We discuss below how commonly these and other
+mechanisms that may cause non-N copula structures.
+We discuss below how commonly some of our
 potential mechanisms may opperate. 
 We also hypothesized and demonstrated important consequences of 
-non-N and tail-dependent copula structure in ecology 
-(Fig. \ref{MasterFigure}, X-Z). First (Fig. \ref{MasterFigure}, X), the skewness
+non-N copula structure and tail-dependence in ecology. 
+For instance, the skewness
 of a spatial-average time series is influenced by tail dependencies between its 
-constituent time series. Thus right-tail-dependent local time series 
+constituent time series: right-tail-dependent local time series 
 lead to "spiky" spatially averaged time series, with large outbreaks.
-These observations could have implications for numerous exploited and pest species
-for which the spatial average time series have applied importance.
-Second (Fig. \ref{MasterFigure}, Y), metapopulation extinction risk can be accentuated or reduced by tail dependence
-properties of environmental variables. We also showed effects of copula 
-structure on Taylor's law (Fig. \ref{MasterFigure}, Z), which has been applied in many fields.
-In our view, our results make our "overall hypothesis," framed in the Introduction,
-reasonable and worthy of broader testing by the research community. We have illustrated 
-extensively how such tests might be carried out by performing them ourselves on 
-example datasets and models. Further suggestions for future possible tests are below. 
-<!--***DAN: make sure you have enough of this, Dan-->
+Thus tail dependence could have implications for numerous exploited and pest species
+for which the spatial average has applied importance.
+Extinction risk can be accentuated or reduced by tail dependence
+properties of environmental variables. 
+In our view, our results make our "overall hypothesis" from the Introduction
+reasonable and worthy of broader testing. We illustrated 
+how such tests might be carried out by performing them ourselves on 
+example datasets and models. Further suggestions for future tests are below.
 It already seems likely that comprehensive understanding 
 of many ecological phenomena is impossible without 
-a complete copula characterization of dependence between variables.
+a complete copula characterization of dependence.
 
 Copula approaches are
-well developed, statistically [@joe2014_dependence; @nelsen2006_copula], and 
-computer implementations exist (e.g., the `copula` and 
-`VineCopula` packages in R). Ecologists can start applying these tools
-immediately. We created a variety of interrelated randomization procedures
+well developed [@joe2014_dependence; @nelsen2006_copula], and 
+computer implementations exist (the `copula` and 
+`VineCopula` packages). Ecologists can start applying these tools
+immediately. We created several interrelated randomization procedures
 (Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
-that build upon existing methods and may be especially useful for ecological applications.
+that build upon existing methods and may be especially useful for ecology.
 
@@
 variables (see below discussion about the green spruce aphid, and 
-Appendix \ref{SM-sect:greenspruce}, which illstrate how to do this).
+Appendix \ref{SM-sect:greenspruce}, which illustrate how to do this).
 Causal mechanism C, which we also did not test, could be tested straightforwardly

### 2019-05-14 SupportingInformation.Rmd
more editing
adding auto link

@@
 mean over pairs of sampling locations of the statistic $\Ps_l-\Ps_u$ were 
-$(-0.085,-0.041)$ for the leaf-curling plum aphid first flight data, 
+(`r unname(round(stat_aphid_ff$numericdf[10,c(3,4)],3))`)
+for the leaf-curling plum aphid first flight data, 
 indicating greater upper-tail dependence, and consistent with
-<!--***DAN: Shya, pls make above results autolink-->
 the results of Table \ref{MT-tab_multivar_summary} (row 9). 
 For the *Ceratium furca* abundance data, confidence intervals were
-$(0.011,0.044)$, indicating greater lower-tail dependence, and again
-<!--***DAN: Shya, pls make above results autolink-->
+(`r unname(round(stat_plankton_north_sea$numericdf[10,c(3,4)],3))`), 
+indicating greater lower-tail dependence, and again 
 consistent with the results of Table \ref{MT-tab_multivar_summary} (row 9).
 For the methane-flux data, confidence intervals were 
-$(-0.008,0.019)$.
+(`r unname(round(stat_methane$numericdf[10,c(3,4)],3))`).
 
@@
 of our hypothesized mechanism and simulations.
+
+<!--Following two chunks examine asymetric dependence mechanism with model selection and non-parametric statistics approach-->
+```{r asym_sens_b_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
+set.seed(seed)
+#-----------------------------------------------------------------------------------
+source("OurBiCopSelect.R")
+source("bivfunctionplot.R")
+source("asym_sensitivity.R")
+#-----------------------------------------------------------------------------------
+# This function takes 
+# Input :
+#       mg : global noise matrix (dim = N by 2)
+#       ml : local noise matrix (dim = N by 2)
+#       a : a number : local noise coeeficient
+#       b : growth factor of pop
+#       decrease : logical (controls the nature of global noise)
+#       numkeep_last : number of points you want to keep for the pop matrix from the end
+#       ploton : logical for optional plot
+
+# Output :
+#     a list of 4 pop matrices 
+#                         1. (for b=+ve, decrease=F)
+#                         2. (for b=+ve, decrease=T)
+#                         3. (for b=-ve, decrease=F)
+#                         4. (for b=-ve, decrease=T)
+
+plotter_asym_sens_b_varn<-function(mg,ml,a,b,numkeep_last,ploton,resloc){
+
+  pop_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
+  pop_d_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
+
+  if(ploton==T){
+    pdf(paste0(resloc,"a_",a,"_b_",b,"_r_",r,"_asym_sens.pdf",sep=""),height=3,width=6)
+    op<-par(mfrow=c(1,2),mar=c(4,4.2,1.5,1),mgp=c(2.5,0.6,0))
+    #plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
+    pop<-pop_g$pop
+    geps<-pop_g$g_noise
+    #plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=1")
+    #plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=2")
+    plot(pop[,1],pop[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2,col="grey",pch=20,cex=1.4)
+    #title(expression(paste("g = ",g[1]," , ")), adj=0.06,cex.main=1.5)
+    #mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)
+
+    pop_d<-pop_d_g$pop
+    geps<-pop_d_g$g_noise
+    #plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=1")
+    #plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=2")
+    plot(pop_d[,1],pop_d[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2,col="grey",pch=20,cex=1.4)
+    #title(expression(paste("g = ",g[2]," , ")), adj=0.06,cex.main=1.5)
+    #mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)
+
+    par(op)
+    dev.off()
+  }
+
+  return(list(pop=pop,
+              pop_d=pop_d))
+
+}
+#----------------------------------------------------------
+N<-25000
+r<-0.8
+a<-0.2
+numkeep_last<-2500
+temp<-get_noise_glob_loc_mat(N=N,r=r)
+mg<-temp$mg
+ml<-temp$ml
+
+resloc<-"./Results/asym_sens_results/"
+ans_b_0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
+ans_b_neg0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
+ans_b_0.5<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.5,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
+ans_b_neg0.5<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.5,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
+
+vlist_b_0.1_b_0.5<-c(ans_b_0.1,ans_b_neg0.1,ans_b_0.5,ans_b_neg0.5) # a list of 8 matrices
+
+summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
+rownames(summary_stat)<-c("bestcop_AICw","N_AICw","relLTdep_AICw","relUTdep_AICw","p_CvM","p_KS","BS_success%")
+#bg <- intToUtf8(0x03B2) 
+cnm<-c("b_0.1_g~1~","b_0.1_g~2~","b_-0.1_g~1~","b_-0.1_g~2~","b_0.5_g~1~","b_0.5_g~2~","b_-0.5_g~1~","b_-0.5_g~2~")
+#cname<-paste0(bg,cnm)
+colnames(summary_stat)<-cnm
+
+ms_asym_b_0.1_b_0.5<-vector(mode = "list", length = 8) # initializing list to store model selection results for all possible b variation
+names(ms_asym_b_0.1_b_0.5)<-cnm
+
+stat_npa_asym_b_0.1_b_0.5<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible b variation
+names(stat_npa_asym_b_0.1_b_0.5)<-cnm
+
+for(ic in c(1:8)){
+  v<-vlist_b_0.1_b_0.5[[ic]]
+  res<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=c(1,3:10,13:14,16:20),level=0.05,
+                      AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
+                      gofnormal=F,status=F)
+
+  ms_asym_b_0.1_b_0.5[[ic]]<-res
+
+  summary_stat[1,ic]<-max(res$InfCritRes$AICw)
+  summary_stat[2,ic]<-res$InfCritRes$AICw[1]
+  summary_stat[3,ic]<-res$relLTdep_AICw
+  summary_stat[4,ic]<-res$relUTdep_AICw
+  summary_stat[5,ic]<-res$GofRes_CvM
+  summary_stat[6,ic]<-res$GofRes_KS
+  summary_stat[7,ic]<-(res$Numboot_success/res$Numboot)*100
+
+  stat_npa_asym_b_0.1_b_0.5[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cnm[ic],numbin=10)
+
+}
+
+saveRDS(summary_stat,"./Results/asym_sens_results/b_variation_summary_stat.RDS")
+saveRDS(ms_asym_b_0.1_b_0.5,"./Results/asym_sens_results/ms_asym_b_0.1_b_0.5.RDS")
+saveRDS(stat_npa_asym_b_0.1_b_0.5,"./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
+```
+
+```{r overwritting_plots_asym_cop,echo=F,results="hide"}
+pdf("./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf",height=3,width=6)
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
+pop<-ans_b_0.1$pop
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
+text(0.1,0.95,"A",cex=2)
+pop<-ans_b_0.1$pop_d
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,pch=20,col=rgb(0,0,0,0.2),cex=0.6)
+text(0.1,0.95,"B",cex=2)
+par(op)
+dev.off()
+
+pdf("./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf",height=3,width=6)
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
+pop<-ans_b_neg0.1$pop
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,pch=20,col=rgb(0,0,0,0.2),cex=0.6)
+text(0.1,0.95,"A",cex=2)
+pop<-ans_b_neg0.1$pop_d
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
+text(0.1,0.95,"B",cex=2)
+par(op)
+dev.off()
+
+pdf("./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf",height=3,width=6)
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
+pop<-ans_b_0.5$pop
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
+text(0.1,0.95,"A",cex=2)
+pop<-ans_b_0.5$pop_d
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
+text(0.1,0.95,"B",cex=2)
+par(op)
+dev.off()
+
+pdf("./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf",height=3,width=6)
+op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
+pop<-ans_b_neg0.5$pop
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
+text(0.1,0.95,"A",cex=2)
+pop<-ans_b_neg0.5$pop_d
+plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
+     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
+text(0.1,0.95,"B",cex=2)
+par(op)
+dev.off()
+```
+
+<!--adding plots from asymmetric dependence analysis results-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$.\label{fig_ms_asym_b_-0.1}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.5$.\label{fig_ms_asym_b_0.5}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.5$.\label{fig_ms_asym_b_-0.5}}
+\end{center}
+\end{figure}
+
+<!--Displaying summary table with model selection results from asymmetric dependence analysis-->
+```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
+summary_stat<-summary_stat[c(1:6),]
+rownames(summary_stat)<-c("Best-fit copula AICw","N-copula AICw","Model-avg. LT","Model-avg. UT","p, CvM","p, KS")
+knitr::kable(summary_stat, booktabs = T, escape=F,
+             format = "latex", linesep = "",
+             digits=4, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$",
+                                   "b=-0.1, g$_{1}$","b=-0.1, g$_{2}$",
+                                   "b=0.5, g$_{1}$","b=0.5, g$_{2}$",
+                                   "b=-0.5, g$_{1}$","b=-0.5, g$_{2}$"),
+      caption.short="Asymmetric sensitivity model, model selection results",
+      caption="Asymmetric sensitivity model, model selection results. 
+The output of simulations from the second model detailed in 
+the methods section for the causal hypotheses (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
+\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to the model 
+selection procedures detailed in the 
+methods section for the observational hypothesis. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
+N-copula AICw: the AIC weight of the N copula family; Model-avg. LT: the lower-tail 
+statistic, weighted-averaged across models using AICw; Model-avg. UT: same, but for the 
+upper-tail statistic; p, CvM: p-value for goodness of fit test of the best-fitting copula
+family using the test based on the Cramer-von Mises statistic; p, KS: p-value for goodness of fit 
+test of the best-fitting copula family using the test based on the Kolmogorov-Smirnov 
+statistic.\\label{tab_summary_stat_asym_sens}")
+```
+
+<!--Displaying summary table with non-parametric approach results of asymmetric dependence analysis-->
+```{r tab_stat_npa_asym,echo=F,results="markup"}
+library(kableExtra)
+
+tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
+tab_stat_npa_asym<-c()
+for(i in c(1:8)){
+  temp<-make_table_stat_fn(data=tab_asym[[i]])
+  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
+}
+gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, g$_{1}$","b=-0.5, g$_{2}$")
+tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),each=3),
+  g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
+knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
+             format = "latex",linesep = "",escape=F,
+             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
+             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
+The output of simulations from the second model detailed in 
+the methods section for the causal hypotheses (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
+\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
+analyses decribed in the methods section for the observational hypothesis. The listed statistic was computed for the
+output of the model with the indicated b and g, and the value was compared to 1000 values
+of the same statistic computed on surrogate time series randomized to have N-copula structure
+(see methods for the observational hypothesis). A table entry $<X$ (respectively, $>X$) indicates the value of the 
+given statistic on the data was less (respectively, more) than its value on $X$ of the 
+surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
+indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
+lower-tail dependence compared to an N-copula null hypothesis. `FS' stands for `first slice'
+and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
+0.9 to 1.\\label{tab_stat_npa_asym}")
+```
 
@@
 
-<!--Following two chunks examine asymetric dependence mechanism with model selection and non-parametric statistics approach-->
-```{r asym_sens_b_variation, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"),mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
-set.seed(seed)
-#-----------------------------------------------------------------------------------
-source("OurBiCopSelect.R")
-source("bivfunctionplot.R")
-source("asym_sensitivity.R")
-#-----------------------------------------------------------------------------------
-# This function takes 
-# Input :
-#       mg : global noise matrix (dim = N by 2)
-#       ml : local noise matrix (dim = N by 2)
-#       a : a number : local noise coeeficient
-#       b : growth factor of pop
-#       decrease : logical (controls the nature of global noise)
-#       numkeep_last : number of points you want to keep for the pop matrix from the end
-#       ploton : logical for optional plot
-
-# Output :
-#     a list of 4 pop matrices 
-#                         1. (for b=+ve, decrease=F)
-#                         2. (for b=+ve, decrease=T)
-#                         3. (for b=-ve, decrease=F)
-#                         4. (for b=-ve, decrease=T)
-
-plotter_asym_sens_b_varn<-function(mg,ml,a,b,numkeep_last,ploton,resloc){
-
-  pop_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=F,numkeep_last=numkeep_last)
-  pop_d_g<-Simulator_asym_sens(mg=mg,ml=ml,a=a,b=b,p0=c(0,0),decrease=T,numkeep_last=numkeep_last)
-
-  if(ploton==T){
-    pdf(paste0(resloc,"a_",a,"_b_",b,"_r_",r,"_asym_sens.pdf",sep=""),height=3,width=6)
-    op<-par(mfrow=c(1,2),mar=c(4,4.2,1.5,1),mgp=c(2.5,0.6,0))
-    #plot(pop[,1],pop[,2],col="red",main = bquote(beta ==.(beta)))
-    pop<-pop_g$pop
-    geps<-pop_g$g_noise
-    #plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=1")
-    #plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[1])),main="location=2")
-    plot(pop[,1],pop[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2,col="grey",pch=20,cex=1.4)
-    #title(expression(paste("g = ",g[1]," , ")), adj=0.06,cex.main=1.5)
-    #mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)
-
-    pop_d<-pop_d_g$pop
-    geps<-pop_d_g$g_noise
-    #plot(mg[,1],geps[,1],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=1")
-    #plot(mg[,2],geps[,2],cex.lab=2,cex.axis=2,xlab=expression(epsilon),ylab=expression(paste(g[2])),main="location=2")
-    plot(pop_d[,1],pop_d[,2],xlab=expression(paste(P[1])), ylab=expression(paste(P[2])),cex.lab=2,cex.axis=2,col="grey",pch=20,cex=1.4)
-    #title(expression(paste("g = ",g[2]," , ")), adj=0.06,cex.main=1.5)
-    #mtext(paste("a = ",a,", b = ",b),side = 3, line=0.2, adj=0.9)
-
-    par(op)
-    dev.off()
-  }
-
-  return(list(pop=pop,
-              pop_d=pop_d))
-
-}
-#----------------------------------------------------------
-N<-25000
-r<-0.8
-a<-0.2
-numkeep_last<-2500
-temp<-get_noise_glob_loc_mat(N=N,r=r)
-mg<-temp$mg
-ml<-temp$ml
-
-resloc<-"./Results/asym_sens_results/"
-ans_b_0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
-ans_b_neg0.1<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.1,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
-ans_b_0.5<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=0.5,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
-ans_b_neg0.5<-plotter_asym_sens_b_varn(mg=mg,ml=ml,a=a,b=-0.5,numkeep_last=numkeep_last,ploton=T,resloc=resloc)
-
-vlist_b_0.1_b_0.5<-c(ans_b_0.1,ans_b_neg0.1,ans_b_0.5,ans_b_neg0.5) # a list of 8 matrices
-
-summary_stat<-as.data.frame(matrix(NA,7,8),stringsAsFactors = F) # Table for model selection summary
-rownames(summary_stat)<-c("bestcop_AICw","N_AICw","relLTdep_AICw","relUTdep_AICw","p_CvM","p_KS","BS_success%")
-#bg <- intToUtf8(0x03B2) 
-cnm<-c("b_0.1_g~1~","b_0.1_g~2~","b_-0.1_g~1~","b_-0.1_g~2~","b_0.5_g~1~","b_0.5_g~2~","b_-0.5_g~1~","b_-0.5_g~2~")
-#cname<-paste0(bg,cnm)
-colnames(summary_stat)<-cnm
-
-ms_asym_b_0.1_b_0.5<-vector(mode = "list", length = 8) # initializing list to store model selection results for all possible b variation
-names(ms_asym_b_0.1_b_0.5)<-cnm
-
-stat_npa_asym_b_0.1_b_0.5<-vector(mode = "list", length = 8) # initializing list to store NPA results for all possible b variation
-names(stat_npa_asym_b_0.1_b_0.5)<-cnm
-
-for(ic in c(1:8)){
-  v<-vlist_b_0.1_b_0.5[[ic]]
-  res<-OurBiCopSelect(u1=v[,1],u2=v[,2],families=c(1,3:10,13:14,16:20),level=0.05,
-                      AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
-                      gofnormal=F,status=F)
-
-  ms_asym_b_0.1_b_0.5[[ic]]<-res
-
-  summary_stat[1,ic]<-max(res$InfCritRes$AICw)
-  summary_stat[2,ic]<-res$InfCritRes$AICw[1]
-  summary_stat[3,ic]<-res$relLTdep_AICw
-  summary_stat[4,ic]<-res$relUTdep_AICw
-  summary_stat[5,ic]<-res$GofRes_CvM
-  summary_stat[6,ic]<-res$GofRes_KS
-  summary_stat[7,ic]<-(res$Numboot_success/res$Numboot)*100
-
-  stat_npa_asym_b_0.1_b_0.5[[ic]]<-bivfunctionplot(v=v,resloc="./Results/asym_sens_results/",nametag=cnm[ic],numbin=10)
-
-}
-
-saveRDS(summary_stat,"./Results/asym_sens_results/b_variation_summary_stat.RDS")
-saveRDS(ms_asym_b_0.1_b_0.5,"./Results/asym_sens_results/ms_asym_b_0.1_b_0.5.RDS")
-saveRDS(stat_npa_asym_b_0.1_b_0.5,"./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
-```
-
-```{r overwritting_plots_asym_cop,echo=F,results="hide"}
-pdf("./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf",height=3,width=6)
-op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
-pop<-ans_b_0.1$pop
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
-text(0.1,0.95,"A",cex=2)
-pop<-ans_b_0.1$pop_d
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,pch=20,col=rgb(0,0,0,0.2),cex=0.6)
-text(0.1,0.95,"B",cex=2)
-par(op)
-dev.off()
-
-pdf("./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf",height=3,width=6)
-op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
-pop<-ans_b_neg0.1$pop
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,pch=20,col=rgb(0,0,0,0.2),cex=0.6)
-text(0.1,0.95,"A",cex=2)
-pop<-ans_b_neg0.1$pop_d
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
-text(0.1,0.95,"B",cex=2)
-par(op)
-dev.off()
-
-pdf("./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf",height=3,width=6)
-op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
-pop<-ans_b_0.5$pop
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
-text(0.1,0.95,"A",cex=2)
-pop<-ans_b_0.5$pop_d
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
-text(0.1,0.95,"B",cex=2)
-par(op)
-dev.off()
-
-pdf("./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf",height=3,width=6)
-op<-par(mfrow=c(1,2),mar=c(3.5,4,1,1),mgp=c(2.2,0.5,0))
-pop<-ans_b_neg0.5$pop
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
-text(0.1,0.95,"A",cex=2)
-pop<-ans_b_neg0.5$pop_d
-plot(pop[,1],pop[,2],xlab=expression(paste(E[1])), ylab=expression(paste(E[2])),cex.lab=1.5,
-     cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
-text(0.1,0.95,"B",cex=2)
-par(op)
-dev.off()
-```
-
-<!--adding plots from asymmetric dependence analysis results-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$.\label{fig_ms_asym_b_-0.1}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.5$.\label{fig_ms_asym_b_0.5}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.5$.\label{fig_ms_asym_b_-0.5}}
-\end{center}
-\end{figure}
-
-<!--Displaying summary table with model selection results from asymmetric dependence analysis-->
-```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
-summary_stat<-summary_stat[c(1:6),]
-rownames(summary_stat)<-c("Best-fit copula AICw","N-copula AICw","Model-avg. LT","Model-avg. UT","p, CvM","p, KS")
-knitr::kable(summary_stat, booktabs = T, escape=F,
-             format = "latex", linesep = "",
-             digits=4, col.names=c("b=0.1, g$_{1}$","b=0.1, g$_{2}$",
-                                   "b=-0.1, g$_{1}$","b=-0.1, g$_{2}$",
-                                   "b=0.5, g$_{1}$","b=0.5, g$_{2}$",
-                                   "b=-0.5, g$_{1}$","b=-0.5, g$_{2}$"),
-      caption.short="Asymmetric sensitivity model, model selection results",
-      caption="Asymmetric sensitivity model, model selection results. 
-The output of simulations from the second model detailed in 
-the methods section for the causal hypotheses (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
-\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to the model 
-selection procedures detailed in the 
-methods section for the observational hypothesis. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
-N-copula AICw: the AIC weight of the N copula family; Model-avg. LT: the lower-tail 
-statistic, weighted-averaged across models using AICw; Model-avg. UT: same, but for the 
-upper-tail statistic; p, CvM: p-value for goodness of fit test of the best-fitting copula
-family using the test based on the Cramer-von Mises statistic; p, KS: p-value for goodness of fit 
-test of the best-fitting copula family using the test based on the Kolmogorov-Smirnov 
-statistic.\\label{tab_summary_stat_asym_sens}")
-```
-
-<!--Displaying summary table with non-parametric approach results of asymmetric dependence analysis-->
-```{r tab_stat_npa_asym,echo=F,results="markup"}
-library(kableExtra)
-
-tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
-tab_stat_npa_asym<-c()
-for(i in c(1:8)){
-  temp<-make_table_stat_fn(data=tab_asym[[i]])
-  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
-}
-gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, g$_{1}$","b=-0.5, g$_{2}$")
-tab_stat_npa_asym<-cbind(b=rep(c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),each=3),
-  g=rep(c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),each=3),tab_stat_npa_asym)
-knitr::kable(tab_stat_npa_asym, booktabs = T, longtable=T,
-             format = "latex",linesep = "",escape=F,
-             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
-             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
-The output of simulations from the second model detailed in 
-the methods section for the causal hypotheses (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
-\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
-analyses decribed in the methods section for the observational hypothesis. The listed statistic was computed for the
-output of the model with the indicated b and g, and the value was compared to 1000 values
-of the same statistic computed on surrogate time series randomized to have N-copula structure
-(see methods for the observational hypothesis). A table entry $<X$ (respectively, $>X$) indicates the value of the 
-given statistic on the data was less (respectively, more) than its value on $X$ of the 
-surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
-indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
-lower-tail dependence compared to an N-copula null hypothesis. `FS' stands for `first slice'
-and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
-0.9 to 1.\\label{tab_stat_npa_asym}")
-```
-
 <!--Following chunk is to examine the hypothesis for third probable mechanism : 
@@
 ```
+
+<!--Following 2 chunks examine the effect of non-Normal asymmetric copula on extinction risk-->
+```{r sim_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
+
+source("mydispersal.R")
+set.seed(seed)
+
+model<-"lcohen"
+numsims<-10000
+numsteps<-25
+params<-0
+ext_thrs<-1
+scl<-1
+p0<-50
+
+sl5<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
+saveRDS(sl5,"./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs5.RDS")
+
+sl25<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
+saveRDS(sl25,"./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs25.RDS")
+
+sg5<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
+saveRDS(sg5,"./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs5.RDS")
+
+sg25<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
+                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
+saveRDS(sg25,"./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs25.RDS")
+```
+
+```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,sl5,sl25,sg5,sg25)}
+sl5<-readRDS("./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs5.RDS")
+sl25<-readRDS("./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs25.RDS")
+sg5<-readRDS("./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs5.RDS")
+sg25<-readRDS("./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs25.RDS")
+
+# plot for local dispersal in D matrix 
+pdf("./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf",heigh=5,width=10)
+op<-par(mfrow=c(1,2),mar=c(5,5,1.8,1),mgp=c(3,1,0))
+plot(sl5$d_seq[1:6],sl5$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
+     xlab="d",ylab="Extinction risk",
+     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
+lines(sl5$d_seq[1:6],sl5$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
+legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 5"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
+       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
+       horiz=F, bty='n')
+
+plot(sl25$d_seq[1:6],sl25$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
+     xlab="d",ylab="Extinction risk",
+     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
+lines(sl25$d_seq[1:6],sl25$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
+legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 25"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
+       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
+       horiz=F, bty='n')
+
+par(op)
+dev.off()
+
+# plot for global dispersal in D matrix 
+pdf("./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf",heigh=5,width=10)
+op<-par(mfrow=c(1,2),mar=c(5,5,1.8,1),mgp=c(3,1,0))
+plot(sg5$d_seq[1:6],sg5$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
+     xlab="d",ylab="Extinction risk",
+     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
+lines(sg5$d_seq[1:6],sg5$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
+legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 5"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
+       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
+       horiz=F, bty='n')
+
+plot(sg25$d_seq[1:6],sg25$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
+     xlab="d",ylab="Extinction risk",
+     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
+lines(sg25$d_seq[1:6],sg25$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
+legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 25"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
+       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
+       horiz=F, bty='n')
+par(op)
+dev.off()
+```
+
+<!--adding plots showing effects of asymmetric copula on extinction risk-->
+\begin{figure}[!h]
+\textbf{ \hspace{7.3 cm} (A) \hspace{4 cm} (B)} \\
+\vspace{-0.8 cm}
+\begin{center}
+\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
+\caption{Analogous to Fig. \ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal.\label{fig_extrisk_lcohen_global}}
+\end{center}
+\end{figure}
 
@@
 simulation and plotting details.\label{fig_spTL}}
-\end{center}
-\end{figure}
-
-<!--Following 2 chunks examine the effect of non-Normal asymmetric copula on extinction risk-->
-```{r sim_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
-
-source("mydispersal.R")
-set.seed(seed)
-
-model<-"lcohen"
-numsims<-10000
-numsteps<-25
-params<-0
-ext_thrs<-1
-scl<-1
-p0<-50
-
-sl5<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
-                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
-saveRDS(sl5,"./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs5.RDS")
-
-sl25<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
-                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=F,ploton=F)
-saveRDS(sl25,"./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs25.RDS")
-
-sg5<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=5,p0=p0,params=params,
-                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
-saveRDS(sg5,"./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs5.RDS")
-
-sg25<-varying_d(numsims=numsims,numsteps =numsteps,numlocs=25,p0=p0,params=params,
-                 ext_thrs=ext_thrs,scl=scl,model=model,disp_everywhere=T,ploton=F)
-saveRDS(sg25,"./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs25.RDS")
-```
-
-```{r fig_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,sl5,sl25,sg5,sg25)}
-sl5<-readRDS("./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs5.RDS")
-sl25<-readRDS("./Results/ext_risk_copula/ext_risk_local_lcohen_numlocs25.RDS")
-sg5<-readRDS("./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs5.RDS")
-sg25<-readRDS("./Results/ext_risk_copula/ext_risk_global_lcohen_numlocs25.RDS")
-
-# plot for local dispersal in D matrix 
-pdf("./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf",heigh=5,width=10)
-op<-par(mfrow=c(1,2),mar=c(5,5,1.8,1),mgp=c(3,1,0))
-plot(sl5$d_seq[1:6],sl5$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
-     xlab="d",ylab="Extinction risk",
-     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
-lines(sl5$d_seq[1:6],sl5$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
-legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 5"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
-       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
-       horiz=F, bty='n')
-
-plot(sl25$d_seq[1:6],sl25$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
-     xlab="d",ylab="Extinction risk",
-     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
-lines(sl25$d_seq[1:6],sl25$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
-legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 25"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
-       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
-       horiz=F, bty='n')
-
-par(op)
-dev.off()
-
-# plot for global dispersal in D matrix 
-pdf("./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf",heigh=5,width=10)
-op<-par(mfrow=c(1,2),mar=c(5,5,1.8,1),mgp=c(3,1,0))
-plot(sg5$d_seq[1:6],sg5$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
-     xlab="d",ylab="Extinction risk",
-     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
-lines(sg5$d_seq[1:6],sg5$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
-legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 5"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
-       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
-       horiz=F, bty='n')
-
-plot(sg25$d_seq[1:6],sg25$risk_left[1:6],xlim=c(0,0.5),ylim=c(0,1),type="b",col="black",
-     xlab="d",ylab="Extinction risk",
-     cex.lab=2,cex.axis=1.8,cex=1.5,pch=16,lwd=1.5)
-lines(sg25$d_seq[1:6],sg25$risk_right[1:6],type="b",lwd=1.5,cex=1.5,pch=16,col="grey")
-legend(x=0.05,y=0.9, title=paste0("r = 0 , N = 25"), c("Left tail-dep. in noise","Right tail-dep. in noise"), 
-       pch=c(16,16), lty=c(1,1), cex=1.4, pt.cex=c(1.5,1.5), col=c("black","grey"),
-       horiz=F, bty='n')
-par(op)
-dev.off()
-```
-
-<!--adding plots showing effects of asymmetric copula on extinction risk-->
-\begin{figure}[!h]
-\textbf{ \hspace{7.3 cm} (A) \hspace{4 cm} (B)} \\
-\vspace{-0.8 cm}
-\begin{center}
-\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal.\label{fig_extrisk_lcohen_global}}
 \end{center}
@@
 
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/aphid_count_winter_temp_results/copulaplot_greenspruceaphidcount_temp.pdf}
-\caption[Copula plots for winter temperature versus green spruce aphid abundance]{Copula plots 
-for winter temperature versus green spruce 
-aphid abundance for each of the 10 sampling locations. The $p$-values in the header are from the test of independence (see section on methods for the observational hypothesis).\label{fig:greenspruce}}
-\end{center}
-\end{figure}
-
 ```{r ms_npa_greenspruceaphid_count_temp, echo=F, results="hide", cache=T, cache.extra=list(seed,aphid_temp_coplist,mtime("CopulaFunctions_flexible.R"),mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
@@
 
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/aphid_count_winter_temp_results/copulaplot_greenspruceaphidcount_temp.pdf}
+\caption[Copula plots for winter temperature versus green spruce aphid abundance]{Copula plots 
+for winter temperature versus green spruce 
+aphid abundance for each of the 10 sampling locations. The $p$-values in the header are from the test of independence (see section on methods for the observational hypothesis).\label{fig:greenspruce}}
+\end{center}
+\end{figure}
+
 # References

### 2019-05-14 Paper.Rmd
finally got the length down to journal limits again

@@
 
+Copulas and their importance in ecology
+
 Copulas and their importance throughout ecology
@@
 
+The importance of copulas
+
+The importance of copulas in ecology
+
 \noindent \emph{Keywords:} dependence, regression, correlation, copula, population, community
@@
 
-\noindent \emph{The number of words in the abstract:}
-
-\noindent \emph{The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends):}
+\noindent \emph{The number of words in the abstract:} 250
+
+\noindent \emph{The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends):} 7493
 
@@
 
-\noindent \emph{The number of references:}
+\noindent \emph{The number of references:} 31
 
@@
 
-All branches of ecology study relationships among environmental and biological variables.
-However, ubiquitously used approaches to studying such relationships, based on correlation and regression, provide only a small slice of the complex information contained in the relationships.
-Other statistical tools exist that provide a complete description of relationships between variables, based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, but not in ecology.
-We here explore the concepts that underpin copulas and examine the potential for those concepts to improve our understanding of ecology.
-We find that informative copula structure in dependences between variables is exceedingly common across all the environmental, species-trait, phenological, population, community, and ecosystem functioning datasets we considered.
-Many datasets exhibited asymmetric tail dependence, whereby two variables are more strongly related in their left compared to right tails, or vice versa.
-Copula structure and tail dependence can help reveal causal relationships between variables.
-We describe mechanisms by which observed copula structure and tail dependence can arise in ecological data, including a Moran-like effect whereby dependence structures between environmental variables are inherited by ecological variables; and asymmetric or nonlinear influences of environments on ecological variables, such as under Liebig's law of the minimum.
-We describe consequences of copula structure, including impacts on extinction risk and the stability through time of ecosystem services.
-By documenting the importance of a complete description of dependence, advancing conceptual frameworks, and demonstrating a powerful approach, we aim to elevate copula reasoning to widespread use in ecology.
+All branches of ecology study relationships among environmental and biological variables. However, 
+ubiquitously used approaches to studying such relationships, based on correlation and regression, 
+provide only a small slice of the complex information contained in the relationships. Other 
+statistical approaches exist that provide a complete description of relationships between variables, 
+based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, 
+but not in ecology. We here explore the concepts that underpin copulas and examine the potential 
+for those concepts to improve our understanding of ecology. We find that informative copula 
+structure in dependencies between variables is exceedingly common across all the environmental, 
+species-trait, phenological, population, community, and ecosystem functioning datasets we 
+considered. Many datasets exhibited asymmetric tail dependence, whereby two variables are more 
+strongly related in their left compared to right tails, or vice versa. We describe mechanisms by which 
+observed copula structure and tail dependence can arise in ecological data, including a Moran-like 
+effect whereby dependence structures between environmental variables are inherited by ecological 
+variables; and asymmetric or nonlinear influences of environments on ecological variables, such as 
+under Liebig's law of the minimum. Copula structure and tail 
+dependence can also reveal causal relationships between variables. We describe consequences of copula structure, including impacts 
+on extinction risk and the stability through time of ecosystem services. By documenting the 
+importance of a complete description of dependence, advancing conceptual frameworks, and 
+demonstrating a powerful approach, we aim to elevate copula reasoning to widespread use in ecology.
 
@@
 
+Conceptually, a bivariate copula is a random variable 
+on the unit square, $[0,1] \times [0,1]$, with uniform maginals. 
+If $(X,Y)$ is a continuous random
+variable and the cummulative distribution functions of $X$ and $Y$ are 
+$F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the dependence structure
+of $(X,Y)$ is revealed by the random variable $(F_X(X),F_Y(Y))$,
+which takes values in $[0,1] \times [0,1]$ and has 
+uniform marginals. 
 We introduce the copula families we used, and 
 measures of tail dependence.
-Conceptually, a bivariate copula is a pdf
-on the unit square with uniform maginals. 
-The copula is actually defined instead as the corresponding 
-cummulative dstribution function (cdf), but both functions contain the same information.
-If $(X,Y)$ is a continuous random
-variable and the cdfs of $X$ and $Y$ are 
-$F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the dependence structure
-of $(X,Y)$ is revealed by the random variable $(F_X(X),F_Y(Y))$.
-This random variable takes values in $[0,1] \times [0,1]$, and has 
-uniform marginal distributions. 
-The normal copula family (henceforth abreviated N) is a one-parameter
-family of copulas, with parameter corresponding to the correlation of the related 
+The normal copula family (henceforth N) is a one-parameter
+family, with parameter corresponding to the correlation of the related 
 bivariate normal distribtion, and controlling the degree of dependence 
@@
 and $0$ lower-tail dependence. For our model selection procedure (see below), we used 
-several one- and two-parameter families of copulas exhibiting a variety of 
+several families of copulas exhibiting various 
 lower- and upper-tail dependence characteristics 
@@
 abbreviated C, SC, G, SG, J, SJ, F, BB1, SBB1, BB6, SBB6, BB7, SBB7, BB8, SBB8).
-Copulas are implemented in the `copula` and `VineCopula` packages for the 
-R language. Visualizations of the copulas we used are in 
+Copulas are implemented in the R packages `copula` and `VineCopula`. 
+Visualizations are in 
 Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}. 
@@
 tail dependence.
-See @nelsen2006_copula, @joe2014_dependence and the `VineCopula` documentation
-for further background.
-
-We tested the observational hypothesis (O in Fig. \ref{MasterFigure}) 
+See @nelsen2006_copula, @joe2014_dependence for further background.
+
+We tested the observational hypothesis (Fig. \ref{MasterFigure}, O) 
 principally via a model selection procedure in which the 
@@
 selection algorithms were ineffective if data could not be distinguished from independent data 
-since many of the bivariate copula families we considered include the independent copula in the 
-family or as a boundary case. If independence could be rejected ($0.05$ significance level), model 
+since many of the copula families we considered include the independent copula. 
+If independence could be rejected ($0.05$ significance level), model 
 selection proceeded. Third, we fit the 16 bivariate copula families listed previously to the 
-normalized ranks via maximum likelihood, using the `BiCopEst` function in `VineCopula`, 
+normalized ranks via maximum likelihood, using `BiCopEst` in `VineCopula`, 
 then obtaining AIC and BIC values and accompanying model weights $\text{AIC}_{\text{w}}$ and 
@@
 indicate whether any of the models are an objectively good fit. To test that, 
-and as the final step of
-our procedure, we tested the goodness of fit of our AIC-best copula family using a 
-bootstrapping procedure (`BiCopGofTest` of `VineCopula`). The method performed 
+and as our final step, we tested the goodness of fit of our AIC-best copula family using a 
+bootstrapping procedure (`BiCopGofTest` of `VineCopula`). The procedure performed 
 one test based on a Cramer-von Mises statistic and another based on a 
@@
 values to the lower- and upper-tail dependence values of an N copula (i.e., 0). 
-The third part of the observational hypothesis
+The third part 
 was tested by comparing the difference of the model-averaged lower- and upper-tail 
@@
 For multivariate datasets, we carried out the bivariate analysis
-for all possible pairwise combinations of locations, and compared AIC weights and tail-dependence of 
+for all possible pairwise combinations of locations, and 
+compared AIC weights and tail-dependence of 
 minimum-AIC copulas with that of an N copula. We also counted the number of location pairs 
-for which model-averaged lower-tail dependence was greater than model-averaged upper-tail dependence 
-or vice-versa, and reported the percentage of pairs showing stronger lower-tail dependence out of 
-the total number of non-independent location pairs considered.
+for which model-averaged lower-tail dependence was 
+greater than model-averaged upper-tail dependence 
+or vice-versa, and reported the percentage of non-independent 
+location pairs showing stronger lower-tail dependence.
 
@@
 The statistics are defined with positively associated
-variables in mind. All the variables we study were positively associated
+variables in mind. All our variables were positively associated
 when they were significantly associated.
@@
 between the bounds.
-Larger values indicate stronger positive association
-between the bounds.
+Larger values indicate stronger positive association.
 We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
@@
 $\Ps_{l_b,u_b}$, for which large values inidicate strong dependence
-for points between the bounds, small values of 
+between the bounds, small values of 
 $\Dsq_{l_b,u_b}$ indicate strong dependence.
@@
 and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"). Likewise $\Ps_l=\Ps_{0,.5}$,
-$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.
+$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, $\Dsq_u=\Dsq_{0.5,1}$.
 
@@
              caption.short = "Summary of results for multivariate datasets",
-             caption = "Summary of results for multivariate datasets.
+             caption = "Summary of results of testing the observational hypothesis for multivariate datasets.
 Rows 3-9 of the table were computed for the non-independent pairs (row 2) only.
@@
 using models. We did not systematically 
-test the first two parts of causal hypothesis A (Fig. \ref{MasterFigure})
+test the first two parts of A 
 because nonlinear and thresold effects of 
-environmental variables on ecological variables are already well documented in
-the literature (however, see Discussion material about the green spruce aphid). To test 
-the third part of causal hypothesis A, we formulated a model in which 
+environmental variables on ecological variables are already well documented 
+(however, see Discussion material about the green spruce aphid). To test 
+the third part of hypothesis A, we formulated a model in which 
 an environmental variable influences an ecological variable
@@
 and $(\delta_1,\delta_2)$ have N copula structure, the Moran 
-mechanism analyzed below (causal hypotheis B on Fig. \ref{MasterFigure}) 
+mechanism analyzed below (B on Fig. \ref{MasterFigure}) 
 does not operate here.
@@
 
-Causal hypothesis B is an extension of the well known Moran
+Hypothesis B is an extension of the well known Moran
 effect, and was summarized conceptually in the Introduction. 
-Let $E_i(t)$ again be the ecological variable, $i=1,2$.
+Let $E_i(t)$ again be an ecological variable, $i=1,2$.
 We initially assume the AR(1) dynamics 
 $E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
-The $\epsilon_i(t)$ represent environmental noise and are 
+The environmental noises $\epsilon_i(t)$ are 
 standard-normal random variables that are independent for distinct times, $t$, but 
-that were made to exhibit different kinds of 
+exhibited different kinds of 
 dependence across locations in different simulations (see below).
@@
 total plant community biomass from an average value,
-flux of a biogeochemical variable such as methane or CO$_2$, or other quantities.
-
-To explore the importance of nonlinear dynamics for causal hypothesis B, we also considered 
+flux of a biogeochemical variable such as methane, or other quantities.
+
+To explore the importance of nonlinear dynamics for hypothesis B, we also considered 
 nonlinear population models based on $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right) \right)+\sigma\epsilon_i(t))$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
@@
 
-For each of the three model setups above, $\tau=0.1,0.2,\ldots,0.9$, and for each of
+For each of the three model setups above, for each of several copula families, for each $\tau=0.1,0.2,\ldots,0.9$, and for each of
 $50$ replicate simulations, we generated 5000 noise pairs 
@@
 
-Causal hypothesis C was not tested. Our next 
+Hypothesis C was not tested. Our next 
 causal hypothesis, D on Fig. \ref{MasterFigure}, 
@@
 trait data across species, e.g., our bird and mammal 
-data. This proposed mechanism is a hypothetical explanation for the right tail dependence
+data. This mechanism is a hypothetical explanation for the right tail dependence
 observed in those data (Fig. \ref{fig_biv_multi_raw_cop}F,G). 
@@
 bivariate draws from distributions showing one of five distinct types of copula 
-structure: 1) extreme left-tail dependence, 2) moderate left-tail dependence,
-3) symmetric tail dependence, 4) moderate right-tail dependence, and 5) extreme
+structure: 1) extreme or 2) moderate left-tail dependence,
+3) symmetric tail dependence, or 4) moderate or 5) extreme
 right-tail dependence (Appendix \ref{SM-evol_mech}). 
@@
 correlation $0.875$ between components, so our simulations assess 
-the impact of copula structure only. For each of the five copula structures, mammalian character
+the impact of copula structure only. For each of the five copulas, mammalian character
 evolution was simulated 100 times. For each simulation, symmetry of tail dependence 
@@
 could produce outputs with visually apparent asymmetry 
-of tail dependence, to an extent that depended on the value of $b$ used,
-thus confirming the third part of causal hypothesis A for the model we used. 
+of tail dependence, to an extent that depended on the value of $b$,
+thus confirming the third part of hypothesis A for our model. 
 For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ (Fig. \ref{SM-fig_ms_asym_b_-0.1}), 
@@
 asymmetry of tail dependence was strong; for 
-$b=\pm 0.5$, asymmetry of tail dependence was weaker but still apparent
+$b=\pm 0.5$, asymmetry was weaker but still apparent
 (Figs \ref{SM-fig_ms_asym_b_0.5} - 
@@
 Lower-tail (respectively, upper-tail) spatial dependence in 
-the noise ($g_1$, respectively, $g_2$) 
+effects of noise ($g_1$, respectively, $g_2$) 
 produced lower-tail (respectively, upper-tail) spatial dependence in the ecological variable.
 Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
-ecological outputs had non-N copula structure, but did not always reflect the 
+ecological outputs had non-N copula structure, but did not always reflect 
 asymmetry of tail dependence, perhaps because those methods measured
-dependence in the most extreme tails. Results using our nonparametric statistics, however,
+dependence in the extreme tails. Results using our nonparametric statistics, however,
 reflected the visually apparent asymmetry (Tables \ref{SM-tab_stat_npa_asym}).
@@
 
-Causal hypothesis B (Fig. \ref{MasterFigure}), i.e. that copula structure 
-in the spatial dependence of ecological variables can arise through
-Moran effects, was largely confirmed; however, similarities between environmental- and 
+Causal hypothesis B (Fig. \ref{MasterFigure}) was largely confirmed; 
+however, similarities between environmental- and 
 ecological-variable copula structure were reduced when dynamics were strongly nonlinear.
@@
 
-For weak noise and using our nonlinear model, copula statistics
+For weak noise and our nonlinear model, copula statistics
 were again very similar for noise and model outputs 
@@
 Though there were again significant but small differences for some statistics
-and simulations, the overall hypothesis was supported that copula structure in the noise translates
+and simulations, the hypothesis was supported that copula structure in the noise translates
 to similar but not identical copula structure in output populations. 
 Since many ecological models are nonlinear, this result provides the 
-important expectation that there will be a Moran-effect-like correspondence 
+expectation that there will be a Moran-effect-like correspondence 
 between noise and model-output copula structure for typical ecological dynamics,
-as long as the noise is small enough that dynamics stay relatively close
+as long as noise is small enough that dynamics stay relatively close
 to the model equilibrium. Results that hold for "weak noise" in this sense
@@
 
-The hypothesis was presented in the Introduction that
-the distribution (through time) of a spatially averaged quantity should be influenced by the 
-dependence between the local quantities, including their
-copula structure and tail dependence (consequential hypothesis 
-X on Fig. \ref{MasterFigure}). 
-To refine the hypothesis, suppose an ecological variable 
+The hypothesis was presented that
+the distribution (through time) of a spatially averaged quantity should be influenced by 
+dependencies between the local quantities, including their
+copula structure and tail dependence (Fig. \ref{MasterFigure}, X). 
+To refine this hypothesis, suppose an ecological variable 
 $E_i(t)$ is measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and
-the distribution and temporal dynamics of the spatial mean $\sum_i E_i(t)/N$ is of
+the spatial mean $\sum_i E_i(t)/N$ is of
 interest. The $E_i(t)$ could be, for instance, local abundances of a pest or 
 exploited species, or local fluxes of a greenhouse gas. If  $E_i$ and $E_j$ are 
-strongly right-tail dependent for all or most location pairs $i$ and $j$, then
-exceptionally large values tend to occur at the same time in all or most locations.
-We therefore hypothesize that this 
-should tend to increase the skewness of the distribution through time 
+right-tail dependent for most location pairs $i$ and $j$, then
+exceptionally large values tend to occur at the same time in most locations.
+We hypothesize that this 
+will tend to increase the skewness of the distribution 
 of the spatial mean. Similarly, left-tail dependence between local variables
-should tend to decrease skewness of the spatial mean. Strong positive skewness of the 
+should tend to decrease skewness. Strong positive skewness of the 
 spatial mean corresponds to "spikiness" of the spatial-mean time series. 
@@
 principal importance for 
-cases such as pest or resource abundance, for which extreme values (spikes) 
-in the spatial-mean have disproportionately large effects.
-
-We tested consequential hypothesis X 
+pest or resource abundance, for which extreme values (spikes) 
+in the spatial-mean have large effects.
+
+We tested hypothesis X 
 using our multivariate datasets (Table \ref{tab_multivar_summary}).
 For each dataset we calculated the spatial mean time series, and then 
-the skewness (through time) of that
+the skewness through time of that
 mean. Then, for each dataset, we compared the value obtained 
 to a distribution of values 
-of the same quantity calculated for each of 10000 surrogate datasets. Surrogate 
+of the same quantity for each of 10000 surrogate datasets. Surrogate 
 datasets were produced by randomizing the empirical data
@@
 unchanged. Significant differences indicate that non-N copula structure in the data 
-contributed to the skewness of the spatial mean, i.e., to the "spikiness" of that
-time series.
+contributed to the skewness of the spatial mean, i.e., to its "spikiness".
 
@@
 (Table \ref{tab_multivar_summary}), we compared empirical and surrogate skewness values
-via a one-tailed test in the left tail: the $p$-value result was the fraction
-of surrogate skewnesses less than the skewness computed from the empirical data; the test
+via a one-tailed test in the left tail: the $p$-value was the fraction
+of surrogate skewnesses less than the skewness for the empirical data; the test
 examines whether lower-tail dependence caused the spatial average to have significantly lower
@@
 in the right tail. The test examines whether upper-tail dependence caused the spatial
-average to have significantly higher skewness than would have been
+average to have significantly higher skewness than 
 expected without tail dependence.
 
-We also examined consequential hypothesis Y on Fig. \ref{MasterFigure}, i.e., that spatial 
-tail dependence of an environmental variable could influence the 
+We also examined hypothesis Y on Fig. \ref{MasterFigure}, i.e., that spatial 
+tail dependence of an environmental variable can influence the 
 extinction risk of a metapopulation. We hypothesized that environmental noises
@@
 component populations would occur simultaneously in many patches,
-reducing rescue effects. Here we are assuming, for simplicity, 
+reducing rescue effects. Here we assume, for simplicity, 
 that low values of the environmental
-variable are "bad" for the populations being modelled, and high values are "good".
+variable are "bad" for the populations and high values are "good".
 We tested the reasonableness of Y using a metapopulation
@@
 $\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
-where $\vec{P}(t)$ is a vector of population densities, 
-the $i^{th}$ component $P_i(t)$ representing
-the density in the $i^{th}$ habitat patch ($i=1,\ldots,N$) at time $t$. 
+where the $i^{th}$ component of the length-$N$ vector 
+$\vec{P}(t)$ represents
+population density in the $i^{th}$ habitat patch at time $t$. 
 The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
@@
 different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
-matrix $D$ was a dispersal matrix. For simplicity, we considered habitat patches 
-to be arranged in a line, with dispersal occurring either to nearest neighbors only,
-or globally, in different simulations (Appendix \ref{SM-ext_risk_model}). The dispersal 
-rate was $d$. After each time step, 
-if the population density in a patch was below $1$, it was
+matrix $D$ was a dispersal matrix
+modelling local or global dispersal at rate $d$, in different simulations
+(Appendix \ref{SM-ext_risk_model}).
+After each step, 
+if the density in a patch was $<1$, it was
 set to $0$. We simulated the model 10000 times for each combination of parameters,
-starting from initial populations $p_0 = 50$ in each patch, 
+starting from $p_0 = 50$ in each patch, 
 and calculated extinction risk after $25$ time steps.
@@
 
-\noindent Skewness results were generally consistent with consequential 
-hypothesis X. For the datasets that
+\noindent Skewness results were consistent with 
+consequential hypothesis X. For datasets that
 exhibited strong lower-tail (respectively, upper-tail) 
-dependence in earlier analyses (namely, the green spruce 
-aphid abundance data for lower-tail, and the leaf-curling plum aphid first flight data
+dependence in earlier analyses (green spruce 
+aphid abundance data for lower-tail, and leaf-curling plum aphid first flight data
 for upper-tail, Table 
-\ref{tab_multivar_summary}), the skewness of the spatial average 
-was less than (respectively, greater than) a significant fraction of the
+\ref{tab_multivar_summary}), skewness of the spatial average 
+was less than (respectively, greater than) a significant fraction of 
 surrogate skewness values 
-(Fig. \ref{fig_skewness_spearman}A,B). For the datasets with moderate lower-tail 
-dependence (\emph{Ceratium furca} abundance and methane data), the skewness of the spatial
+(Fig. \ref{fig_skewness_spearman}A,B). For datasets with moderate lower-tail 
+dependence (\emph{Ceratium furca} abundance and methane data), skewness of the spatial
 average showed a non-significant or marginally significant tendency 
@@
 
-Consistent with consequential hypothesis Y, 
+Consistent with hypothesis Y, 
 left-tail dependent environmental fluctuations increased metapopulation
-extinction risk for the spatially 
-extended Lewontin-Cohen model (Fig. \ref{fig_extrisk_lcohen_local}). 
-This was true for both $N=5$ and $N=25$ habitat patches, and for both local 
+extinction risk for the spatial 
+Lewontin-Cohen model (Fig. \ref{fig_extrisk_lcohen_local}), 
+for $N=5$ and $N=25$ habitat patches, and for local 
 (Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
@@
 a consequential hypothesis Z (Fig. \ref{MasterFigure}), that 
-Taylor's law can be strongly influenced by
-tail dependence between populations.
+Taylor's law can be influenced by
+tail dependence.
 
@@
 We showed that non-N copula structure and asymmetric tail dependence are 
-common in datasets across multiple sub-disciplines
-and levels of organization in ecology, although no published 
-research we are aware of has 
+common across multiple sub-disciplines
+in ecology, although no 
+research we know of has 
 taken this structure into account. We hypothesized four 
@@
 We discuss below how commonly some of our
-potential mechanisms may opperate. 
+mechanisms may opperate. 
 We also hypothesized and demonstrated important consequences of 
-non-N copula structure and tail-dependence in ecology. 
+non-N copula structure and tail-dependence. 
 For instance, the skewness
@@
 lead to "spiky" spatially averaged time series, with large outbreaks.
-Thus tail dependence could have implications for numerous exploited and pest species
-for which the spatial average has applied importance.
-Extinction risk can be accentuated or reduced by tail dependence
+Thus tail dependence could have implications for exploited and pest species.
+Extinction risk can be altered by tail dependence
 properties of environmental variables. 
-In our view, our results make our "overall hypothesis" from the Introduction
+In our view, our results make our "overall hypothesis" (Introduction)
 reasonable and worthy of broader testing. We illustrated 
-how such tests might be carried out by performing them ourselves on 
-example datasets and models. Further suggestions for future tests are below.
-It already seems likely that comprehensive understanding 
+how such tests might be carried out by performing tests ourselves on 
+example datasets and models. Further suggestions for tests are below.
+It seems likely that comprehensive understanding 
 of many ecological phenomena is impossible without 
@@
 well developed [@joe2014_dependence; @nelsen2006_copula], and 
-computer implementations exist (the `copula` and 
-`VineCopula` packages). Ecologists can start applying these tools
+computer implementations exist (`copula` and 
+`VineCopula` packages). Ecologists can apply these tools
 immediately. We created several interrelated randomization procedures
 (Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
-that build upon existing methods and may be especially useful for ecology.
+that build upon existing methods.
 
 <!--a para on how likely one is to get copula structure via the Liebig mechanism-->
-Our first causal mechanism (Fig. \ref{MasterFigure}, A) seems likely to operate commonly.
-Liebig's law and the idea of a limiting 
-nutrient of growth are dominant paradigms in ecology.
-Although we did not systematically test the first two parts of causal 
-hypothesis A, many studies document nonlinear influences of environmental
-variables on ecological quantities at all levels of organization. 
-These parts of A could be tested
-straightforwardly for numerous datasets by examing copulas describing dependencies
-between environmental and ecological
-variables (see below discussion about the green spruce aphid, and 
-Appendix \ref{SM-sect:greenspruce}, which illustrate how to do this).
-Causal mechanism C, which we also did not test, could be tested straightforwardly
-by analyzing copulas of abundances of competing species, across space or
-through time.
+Our first causal mechanism (Fig. \ref{MasterFigure}, A) probably operates commonly,
+for two reasons.
+First, Liebig's law and the idea of limiting 
+nutrients are dominant paradigms in ecology.
+Second, although we did not systematically test the first two parts of 
+hypothesis A, many studies have documented nonlinear or threshold influences of environmental
+variables on ecological quantities. 
+These first two parts of A could be tested for numerous 
+datasets by examining copulas for environmental and ecological
+variables. See below discussion about the green spruce aphid, and 
+Appendix \ref{SM-sect:greenspruce}, which illustrate how to do this.
+Causal mechanism C, which we also did not test, could be tested
+by analyzing copulas of abundances of competing species, sampled across space or
+time.
 
 <!--para on how likely one is to get copula structure via the Moran mechanism-->
-There are also reasons to expect that our Moran mechanism (Fig. \ref{MasterFigure}, B) for 
-copula structure and tail dependence will
-operate commonly, because Moran effects are common [@Liebhold2004], and 
+There are also reasons to expect our Moran mechanism (Fig. \ref{MasterFigure}, B) 
+operates commonly: Moran effects are common [@Liebhold2004], and 
 non-N and tail-dependent copula structures are often found in 
-environmental variables. 
-If large meteorological events are 
-also spatially widespread, then
-one would expect environmental variables associated with these events to take extreme
+environmental variables. If intense meteorological events are 
+also widespread, then
+environmental variables associated with these events should take extreme
 values simultaneously across large spatial areas, producing tail dependence in measurements
-made through time at different locations. Non-extreme values, not associated with 
-spatially extensive weather events, may
-be more associated with local phenomena, and therefore 
-less prone to being correlated across large areas.
+made through time at different locations. Non-extreme values may instead
+be associated with local phenomena, and therefore may be
+less correlated across large areas.
 @Serinaldi2008 examined the spatial dependence 
-of rainfall observations in Central Italy; G or Student 2-copulas 
-were possible candidates for modelling dependence, and 
+of rainfall in Central Italy; G or Student 2-copulas 
+were candidates for modelling dependence, and 
 neither of these is an N copula.
 A long-term study (1950–2014) in the Loess Plateau of China [@She2018]
-showed that a G copula (which has upper tail dependence) could effectively model 
-the spatial dependence of drought variables in the study area.
+showed that a G copula effectively modeled 
+the spatial dependence of drought variables.
 Bivariate copula analysis was also used in forecasting the 
 co-occurrence of extreme events (flood or drought) 
-over the North Sikkim Himalayas, India, using spatial datasets [@Goswami2018].
+over the North Sikkim Himalayas using spatial datasets [@Goswami2018].
 
 <!--Holder stuff-->
-Our simulations of character evolution suggest the hypothesis that evolutionary
-changes through time in bird and mammal BMR and body size may exhibit right-tail dependence,
-contrary to standard assumptions of models of character evolution. This is
+Our simulations of character evolution suggest the hypothesis that
+changes through evolutionary time in bird and 
+mammal BMR and body size may exhibit right-tail dependence,
+contrary to standard normality assumptions of character evolution models. This is
 a hypothesis only, because the right-tail dependence of Fig. \ref{fig_biv_multi_raw_cop}F,G
-could have come about in another, unknown way: our simulations show that tail dependence 
+could have come about in another, unknown way. Our simulations show that tail dependence 
 in evolutionary changes is sufficient, but may not be necessary, to produce tail
-dependence in characters of extant species. For instance, as discussed below and in
-Appendix \ref{SM-sect:missingdata} (also Fig. \ref{SM-fig:missingdata}), 
-systematically missing data can produce tail dependence and may
+dependence in characters of extant species. For instance, (see below, 
+Appendix \ref{SM-sect:missingdata}, Fig. \ref{SM-fig:missingdata}), 
+systematically missing data can also produce tail dependence and may
 have influenced results for the BMR-body mass datasets. Even if the hypothesized 
 evolutionary mechanism (Fig. \ref{MasterFigure}, D)
-is correct, our results simply replace one question, i.e., why do we see right-tail
-dependence between BMR and body mass for birds and mammals, with another question,
+is correct, our results only replace one question, i.e., why do we see right-tail
+dependence between BMR and body mass, with another,
 i.e., why might we see right-tail dependence in evolutionary changes in these traits?
-See Appendix \ref{SM-evol_mech} for additional thoughts on the potential importance of 
-copulas for studies of character evolution.
+See Appendix \ref{SM-evol_mech} for further thoughts on potential importance of 
+copulas for character evolution.
 
 <!--Discuss stuff here on tradeoff and life history that Bever told me about-->
-The relationships we studied between BMR and body mass relate
-to a trade-off between mass-specific BMR (i.e., BMR per unit body mass) and body
-mass itself. Trade-offs between life history characteristics are a focal topic in 
-ecology. Copulas probably interrelate with 
-life-history theory in additional ways beyond what we could demonstrate 
-with the data we used.
-It is well known that energy allocation to a life function, X (e.g., reproduction)
-will reduce the energy that can be allocated to other functions, $Y_1$, $Y_2$, $Y_3$ 
-(e.g., growth, predation avoidance; this is called the principle of 
-allocation).
-But it is equally well known that X can trade off against any or all of 
-the $Y_i$. Therefore, for large values of X which approach absolute 
+Relationships between BMR and body mass relate
+to a trade-off between mass-specific BMR (BMR per unit body mass) and body
+mass itself. Copulas probably interrelate with 
+life-history trade-off theory in additional ways beyond what we demonstrated.
+For instance, it is well known that energy allocation to a life function, F (e.g., reproduction)
+will reduce the energy that can be allocated to other functions, $G_1$, $G_2$, $G_3$ 
+(e.g., growth, predation avoidance). This is the principle of 
+allocation. But F can trade off against any or all of 
+the $G_i$. Therefore, for large F, approaching absolute 
 limitations, there may be a strong 
-association between X and $Y_1$. Whereas for small values of X there may be
-little association - resources not allocated to X can instead be allocated 
-to $Y_1$, $Y_2$, $Y_3$ or to any combination thereof. This constitutes 
-asymmetric tail dependence between X and $Y_1$. For instance, @Winemiller1992
-described a three-way trade-off in North American fishes between age of reproductive
+association between F and $G_1$. For small F, there may be
+little association because resources not allocated to F can instead be allocated 
+to any combination of the $G_i$. This constitutes 
+asymmetric tail dependence between F and $G_1$. @Winemiller1992
+described a three-way trade-off in fishes between age of reproductive
 maturity, juvenile survivorship, and fecundity. The trade-off should, in theory,
 produce a tight association between age of maturity and fecundity for 
-fishes with low age of maturity, but little association between those
-variables for later-maturing
-fishes because such fishes may invest much or little in juvenile survival.
-Although copulas may be very useful for studying multi-dimensional trade-offs, 
-applications will 
+fishes with low age of maturity, but little such association for later-maturing
+fishes because those fishes may invest the resources not invested in maturing quickly into
+either fecundity or juvenile survival.
+These ideas suggest that copulas will be 
+useful for studying multi-dimensional trade-offs. But applications will 
 require careful attention to the possible consequences of biased sampling:
 if the degree of completeness of a dataset is associated with one or
-more of the characters being studied then statistical artefacts
-may bias conclusions (Appendix \ref{SM-sect:missingdata}, 
+more of the characters, then statistical artefacts
+can bias conclusions (Appendix \ref{SM-sect:missingdata}, 
 Fig. \ref{SM-fig:missingdata}).
@@
 additional mechanisms of copula structure is certainly worthy of future work. -->
-Other mechanisms that we have not considered probably also operate, 
-and should be considered in future work. For instance, 
+Other mechanisms probably also operate, 
+and should be considered. For instance, 
 measurement error may modify copula structure. 
 Our models were intentionally simple in other respects, 
-as well, not including phenomena such as 
-delayed density dependence, dispersal, stage structure in population dynamics, and trophic
-interactions.
+too, not including
+delayed density dependence, dispersal, population stage structure, trophic
+interactions, and other factors.
 We hope by enumerating a few potential mechanisms of copula structure,
@@
 Our hypotheses cover propagation of copula structure from causes, 
-to observable data, to consequences for human concerns.
-We here take a closer look at the green spruce aphid, which illustrates 
-all these stages in one example.
-Green spruce aphid abundance, as detected in the suction trap data we use (Data,
-Appendix \ref{SM-Data_details}),
-is strongly influenced by temperature the previous winter (@Sheppard2016,
+to data, to consequences.
+We here take a closer look at the green spruce aphid, which simultaneously illustrates 
+all these stages.
+Green spruce aphid abundance, as measured in the data we use,
+is strongly influenced by temperature of the previous winter (@Sheppard2016,
 supplementary fig. 6).
@@
 Fig. \ref{SM-fig:greenspruce}, Table \ref{SM-tab:greenspruce}) the copula of winter temperature
-and green spruce aphid abundance time series for the location, finding 
-substantial left-tail dependence: apparently winter temperature has an asymmetric
-influence on abundance in the sense that cold winters produce low abundances
-but warm winters do not yield especially higher abundances than moderate winters.
+and aphid abundance time series for the location, finding 
+left-tail dependence: apparently winter temperature has an asymmetric
+influence on abundance in that cold winters produce low abundances
+but warm winters often do not yield higher abundances than moderate winters.
 Although we did not systematically test the first two 
-parts of causal hypothesis A due to much existing evidence, this result 
-illustrates how parts one and two of the hypothesis can be tested using copula
-techniques. The third part of causal hypothesis A, which our modelling results
-supported, now suggests spatial dependence between green spruce aphid counts
-should exhibit left-tail dependence - exactly what was observed 
-(Table \ref{tab_multivar_summary}). And the consequences of that spatial tail 
-dependence for skewness of spatially averaged aphid counts was described already 
+parts of hypothesis A due to much existing evidence, this result 
+illustrates how those parts can be tested using copula
+techniques. The third part of hypothesis A, which our modelling results
+supported, therefore suggests spatial dependence between green spruce aphid counts
+in different locations should show left-tail dependence - exactly as observed 
+(Table \ref{tab_multivar_summary}). And the consequences of such tail 
+dependence for the skewness of spatially averaged aphid counts was described previously 
 (Fig. \ref{fig_skewness_spearman}A). Thus the asymmetric influence of winter 
-temperature ultimately causes spatially averaged green spruce aphid abundances to have
+temperature ultimately causes spatially averaged aphid abundances to have
 lower skew than they would otherwise, because negative impacts of cold winter
-are spatially synchronous but positive impacts of warmer winters and other factors
-are less spatially or temporally consistent.
+are spatially synchronous.
 
@@
 Our results showed that the skewness of the 
-spatial average of local time series from different locations is influenced 
-by their tail dependence. But the logic of this result should also apply to any collection of time
-series, whether associated with different locations in space or not, that are summed or averaged. 
+spatial average of local time series is influenced 
+by their tail dependence. But the same logic should also apply to any collection of time
+series, whether associated with locations in space or not. 
 Another potential application is time series of all species from a single community, e.g., all
@@
 the influence of interspecific relationships on the 
-variability of total-community
-properties such as total biomass. Typically, variability through time of total-community
-properties is measured with the coefficient of variation, but the skewness of such a time series
-may also be of interest because it can help characterize "spikiness". Future work on copula 
+variability of community
+properties such as total biomass. Typically, variability of community biomass 
+is measured with the coefficient of variation, but skewness
+may also be of interest because it can help characterize "spikiness" 
+through time. Future work on copula 
 structure of interspecific relationships in communities and its implications for community 
-stability seems likely to be an exciting avenue of research.
+variability is likely to be valuable.
 
@@
 Although we demonstrated that tail dependence between environmental variables
-can influence extinction risk, substantial work remains to determine the magnitude 
-of this effect in different scenarios. First, we used 
-a non-density-dependent model to examine extinction risk. Do similar results pertain
+can influence extinction risk, substantial work remains to determine the importance 
+of this effect. First, we used 
+a non-density-dependent model. Do similar results pertain
 when density dependence
-plays a role? Second, we considered metapopulation extinction risk, but the large field 
+acts? Second, we considered metapopulation extinction risk, but the huge field 
 of population viability analysis (PVA) via stochastic 
@@
 between different vital rates exhibit tail dependence, and does tail dependence 
-influence extinction risk of a single population?
-PVA is one of the routes by which a species can be added to the influential
-IUCN red list of threatened species. Finally, is the copula structure of environmental variables
+influence extinction risk in this context?
+Finally, is the copula structure of environmental variables
 changing through time, and, if so, how do such changes influence extinction risks?
 Climate change is known to amplify the factors that lead to extreme weather events
-[@Hansen2012]. Extreme events, by definition, produce values in the right tails of distributions
-of environmental variables. Extreme weather events also tend to be spatially extensive, so
-increased frequency of extreme weather events may correspond to changes in the tail dependence
-of environmental variables measured in different places.
+[@Hansen2012] and hence may alter tail dependence.
 

### 2019-05-14 SupportingInformation.Rmd
finally got the length down to journal limits again

--- 2019-05-13 SupportingInformation.Rmd
@@
 For extreme right-tail dependent noise, we first produced
-a $1000000$ by $2$ matrix, $m$, of 
+a 1,000,000 by $2$ matrix, $m$, of 
 independent standard-normal random draws. Then, each row of $m$ 
@@
     substitutions occurring at the same location.
-Biases in estimating the evolutionary distance can affect down stream analyses [@Phillips2009].
+Biases in estimating the evolutionary distance can affect downstream analyses [@Phillips2009].
 Additionally, changes
@@
 per phylogeny branch. This was principally because branch lengths are often hugely
-uncertain. If branch lengths were well know, alternative simulation strategies may
+uncertain. If branch lengths were well known, alternative simulation strategies may
 include selecting one draw from the matrix of evolutionary changes per unit time, 
-or selecting one draw for the whole branch but rescaling variance according to the 
+or selecting one draw for the whole branch but rescaling the variances of the selected
+character changes according to the 
 length of the branch. These choices amount to the same thing for normal copula

### 2019-05-14 Paper.Rmd

@@
 ---
-title: "Complex and informative dependencies propagate throughout ecology and can be revealed using copulas"
+title: "Copulas reveal complex and informative dependencies propagating throughout ecology"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
@@
 
-\noindent \emph{Possible alternative titles:}
+<!--\noindent \emph{Possible alternative titles:}
+
+Complex and informative dependencies propagate throughout ecology and can be revealed using copulas
 
@@
 
-The importance of copulas in ecology
+The importance of copulas in ecology-->
 

### 2019-05-14 SupportingInformation.Rmd

@@
 ---
-title: "Supporting Information : Complex and informative dependencies propogate throughout ecology and can be revealed using copulas"
+title: "Supporting information: Copulas reveal complex and informative dependencies propagating throughout ecology"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"

### 2019-05-15 SupportingInformation.Rmd
small chnages suggested by Shya, Lawrence, Van

@@
 \textbf{Phenological and population data on aphids.}
-Data pre-processing for the aphid data was the same as 
-that of @Sheppard2016,
-but we excluded one of the 11 sites they used in order to only use sites with at least 30 
-years data.
+Data pre-processing for the aphid data was the same as that of @Sheppard2016, 
+except that years for which data was unavailable were excluded rather than being 
+replaced by median values. We also excluded one of the 11 sites they used in 
+order to only use sites with at least 30 years data.
 

### 2019-05-15 Paper.Rmd
small chnages suggested by Shya, Lawrence, Van
cedar creek data refs added

@@
 Community-level data, obtained from the Cedar Creek Ecosystem Science Reserve, were 
-plant aboveground biomass and Shannon's diversity index for `r nrow(d_cc[[5]])` plots, each 9m by 9m, as sampled in 
+plant aboveground biomass [@ccdataple120] and Shannon's diversity index 
+for `r nrow(d_cc[[5]])` plots[@ccdataplotse120], each 9m by 9m, as sampled in 
 the years 1996-2000 and 2007, each year analyzed separately [@Tilman2001; @Tilman2006]. 
-Shannon's index was computed from plant species percent cover data. 
+Shannon's index was computed from plant species percent cover data[@ccdatapce120]. 
 Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 
@@
 
-# Background, concepts and methods for testing the observational hypothesis \label{Methods}
+# Background, concepts and methods for testing the observational hypothesis\label{Methods}
 
@@
 \noindent
+\sloppy

### 2019-05-15 SupportingInformation.Rmd
cedar creek data refs added

--- 2019-05-14 SupportingInformation.Rmd
@@
 
-\textbf{Community-level data, Cedar Creek.} Above-ground biomass and percent cover data were 
+\textbf{Community-level data, Cedar Creek.} Above-ground biomass [@ccdataple120] and percent cover data [@ccdatapce120] were 
 obtained from biodiversity experiment E120 conducted at the Cedar Creek Ecosystem Science Reserve 
-by Tilman *et al*. In 1994, 168 plots (each 9 m $\times$ 9 m) of 39, 35, 29, 30 and 35 replicates,
+by Tilman *et al*. In 1994, 168 plots [@ccdataplotse120] (each 9 m $\times$ 9 m) of 39, 35, 29, 30 and 35 replicates,
 were seeded with 1, 2, 4, 8, or 16 perennials, respectively, chosen randomly from a pool of 
@@
 \noindent
+\sloppy
 

### 2019-05-15 Paper.Rmd
minor edits

@@
 $u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
-$n+1$; and $v_i$ is defined in the same way but using the $y_i$. The $u_i$ and 
+$n+1$; and $v_i$ is defined in the same way but using the $y_i$. The rank
+of the smallest element in a set is $1$. The $u_i$ and 
 $v_i$ are called *normalized ranks* of the $x_i$ and $y_i$. See, 
@@
 Community-level data, obtained from the Cedar Creek Ecosystem Science Reserve, were 
-plant aboveground biomass [@ccdataple120] and Shannon's diversity index 
-for `r nrow(d_cc[[5]])` plots[@ccdataplotse120], each 9m by 9m, as sampled in 
-the years 1996-2000 and 2007, each year analyzed separately [@Tilman2001; @Tilman2006]. 
-Shannon's index was computed from plant species percent cover data[@ccdatapce120]. 
+plant aboveground biomass [@ccdataple120] and Shannon's diversity index (computed from plant 
+species percent cover data [@ccdatapce120]) for `r nrow(d_cc[[5]])` plots [@ccdataplotse120], 
+each 9m by 9m, as sampled in the years 1996-2000 and 2007, each year 
+analyzed separately [@Tilman2001; @Tilman2006]. 
 Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or 

### 2019-05-16 Paper.Rmd
updated length stats
reading as csv

@@
 ```{r read_somedata, echo=F, results='hide'}
-d_soilCN<-readRDS("Data/dataForShymolinaCoord.rds")
+d_soilCN<-read.csv("./Data/SoilCN/SoilCNdata.csv")
+rownames(d_soilCN)<-d_soilCN$X
+d_soilCN<-d_soilCN[,-1]
 d_soilCN<-na.omit(d_soilCN)

### 2019-05-16 Paper.Rmd
reading as csv

@@
 
-\noindent \emph{The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends):} 7493
+\noindent \emph{The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends):} 7509
 
@@
 
-\noindent \emph{The number of references:} 31
+\noindent \emph{The number of references:} 34
 
@@
 ```{r read_BivMS_soilCN_data,echo=F}
-d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
+d<-read.csv("./Data/SoilCN/SoilCNdata.csv")
+rownames(d)<-d$X
+d<-d[,-1]
 d<-d[,c("SOCstock100","TSNstock100")] #raw soilCN data

### 2019-05-16 Paper.Rmd
minor edits from one final read before submitting to Ecol Lett

--- 2019-05-15 Paper.Rmd
@@
 
-\noindent \emph{The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends):} 7493
+\noindent \emph{The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends):} 7497
 
@@
 of C, D, respectively, about a vertical axis. (G) Contours of the
-pdf function of a normal copula (methods for the observational hypothesis) with 
+pdf of a normal copula (methods for the observational hypothesis) with 
 parameter chosen to
 have the same Spearman correlation as A-D, together with a sample.
-(H) The portion of the Spearman correlation within a band (diagonal lines
+(H) The portion of the Spearman correlation within a band 
+can be computed
+for any band (methods for the observational hypothesis) to describe how dependence
+varies from one part of the two distributions to another. Diagonal lines
 show two bands, the data in the left/lower band showing stronger
-dependence than the right/upper band) can be computed 
-for any band (methods for the observational hypothesis) to describe how dependence
-varies from one part of the two distributions to another.
+dependence than the right/upper band.
 Subscripts h and v 
@@
 variables), and the remaining information (which is solely about
-the dependence). From a data perspective, given 
+dependence). From a data perspective, given 
 a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
@@
 causes of copula structure) are labeled with letters A-D. "Consequential 
-hypothesis" (i.e., hypothesis about 
+hypotheses" (i.e., hypotheses about 
 consequences of copula structure) are labeled with X-Z. The "observational 
@@
 nitrogen, N, and other factors, a plot of plant growth rates versus N may 
-look more like Fig. \ref{fig_cop_pedag1}A or E than like 
-Fig. \ref{fig_cop_pedag1}B or F. Nitrogen 
+look more like Fig. \ref{fig_cop_pedag1}A or C than like 
+Fig. \ref{fig_cop_pedag1}B or D. Nitrogen 
 will control plant growth, producing a clear
 relationship, only when it is limiting, in the left portions of the distributions. 
-The portions of the distributions that dominate the association is visible in the copula
+The portion of the distributions that dominate the association is visible in the copula
 approach, but not via standard techniques. We therefore frame "causal hypothesis" A 
@@
 Our causal and consequential hypotheses imply, or rely on, an "observational
-hypothesis" (O in Fig. \ref{MasterFigure}), which we here frame: datasets in 
+hypothesis" (O in Fig. \ref{MasterFigure}), which we here frame. Datasets in 
 ecology will commonly have copula structure
@@
 uniform marginals.
+
 We introduce the copula families we used, and 
@@
 Kolmogorov-Smirnov statistic. To keep computation times reasonable, 
-an initial run using $100$ bootstraps was performed; if
+a run using $100$ bootstraps was performed; if
 the $p$-value from either test was less than $0.2$, tests were re-run with 
@@
 dependence values to the analogous difference for an N copula, again 0.
-For multivariate datasets, we carried out the bivariate analysis
+For multivariate datasets, we performed the bivariate analysis
 for all possible pairwise combinations of locations, and 
@@
 we used differences $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
-for smaller datasets, (note the opposite order
+for smaller datasets (note the opposite order
 in the last of these); and $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
@@
 commonly showed asymmetric tail 
-dependence: the observational hypothesis was supported (Fig. \ref{MasterFigure}). To make results easier to absorb, we first 
+dependence: the observational hypothesis (Fig. \ref{MasterFigure}) was supported. To make results easier to absorb, we first 
 present results for an example bivariate dataset, then we summarize results
@@
 and $(\delta_1,\delta_2)$ have N copula structure, the Moran 
-mechanism analyzed below (B on Fig. \ref{MasterFigure}) 
+mechanism analyzed below (Fig. \ref{MasterFigure}, B) 
 does not operate here.
@@
 
-For each of the three model setups above, for each of several copula families, for each $\tau=0.1,0.2,\ldots,0.9$, and for each of
+For each of the model setups above, for each of several copula families, for each $\tau=0.1,0.2,\ldots,0.9$, and for each of
 $50$ replicate simulations, we generated 5000 noise pairs 
@@
 Hypothesis C was not tested. Our next 
-causal hypothesis, D on Fig. \ref{MasterFigure}, 
+causal hypothesis, Fig. \ref{MasterFigure}, D, 
 is evolutionary, and pertains to bivariate 
@@
 and simulations.
+For weak noise and our nonlinear model, copula statistics
+were again very similar for noise and model outputs 
+(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-
+\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}).
+Though there were again significant but small differences for some statistics
+and simulations, the hypothesis was supported that copula structure in the noise translates
+to similar but not identical copula structure in output populations. 
+Since many ecological models are nonlinear, this result provides the 
+expectation that there will be a Moran-effect-like correspondence 
+between noise and model-output copula structure for typical ecological dynamics,
+as long as noise is small enough that dynamics stay relatively close
+to the model equilibrium. Results that hold for "weak noise" in this sense
+are common in ecology. 
+We repeated the analysis with $r=1.3$, with similar results. The deterministic 
+one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
+at $K$ when $r<1$ (undercompensating dynamics, e.g., the value $r=0.5$ used initially), 
+but exhibits an oscillatory approach to $K$ when $r>1$ (overcompensating
+dynamics, e.g., $r=1.3$).
 
@@
 
-For weak noise and our nonlinear model, copula statistics
-were again very similar for noise and model outputs 
-(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-
-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}).
-Though there were again significant but small differences for some statistics
-and simulations, the hypothesis was supported that copula structure in the noise translates
-to similar but not identical copula structure in output populations. 
-Since many ecological models are nonlinear, this result provides the 
-expectation that there will be a Moran-effect-like correspondence 
-between noise and model-output copula structure for typical ecological dynamics,
-as long as noise is small enough that dynamics stay relatively close
-to the model equilibrium. Results that hold for "weak noise" in this sense
-are common in ecology. 
-We repeated the analysis with $r=1.3$, with similar results. The deterministic 
-one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
-at $K$ when $r<1$ (undercompensating dynamics, e.g., the value $r=0.5$ used initially), 
-but exhibits an oscillatory approach to $K$ when $r>1$ (overcompensating
-dynamics, e.g., $r=1.3$).
-
 For strong noise and using our nonlinear model, 
@@
 copula structure in this case were often glaring. Apparently noise of standard deviation 1
-interacted very strongly with model nonlinearities when the model was in its overcompensatory 
+interacted strongly with model nonlinearities when the model was in its overcompensatory 
 regime.
 
-For the model we examined, causal hypothesis D (Fig. \ref{MasterFigure}) was 
+For our model, hypothesis D (Fig. \ref{MasterFigure}) was 
 correct that asymmetric tail dependence in evolutionary changes 
@@
 pest or resource abundance, for which extreme values (spikes) 
-in the spatial-mean have large effects.
+in the spatial mean have large effects.
 
@@
 
-We also examined hypothesis Y on Fig. \ref{MasterFigure}, i.e., that spatial 
+We also examined hypothesis Y (Fig. \ref{MasterFigure}), i.e., that spatial 
 tail dependence of an environmental variable can influence the 
@@
 \noindent Skewness results were consistent with 
-consequential hypothesis X. For datasets that
+hypothesis X. For datasets that
 exhibited strong lower-tail (respectively, upper-tail) 
@@
 Extinction risk can be altered by tail dependence
-properties of environmental variables. 
-In our view, our results make our "overall hypothesis" (Introduction)
+in environmental variables. 
+In our view, our results make our overall hypothesis (Introduction)
 reasonable and worthy of broader testing. We illustrated 
@@
 in evolutionary changes is sufficient, but may not be necessary, to produce tail
-dependence in characters of extant species. For instance, (see below, 
+dependence in characters of extant species. For instance (see below, 
 Appendix \ref{SM-sect:missingdata}, Fig. \ref{SM-fig:missingdata}), 
@@
 Our models were intentionally simple in other respects, 
-too, not including
+too, and did not include
 delayed density dependence, dispersal, population stage structure, trophic
-interactions, and other factors.
+interactions, etc.
 We hope by enumerating a few potential mechanisms of copula structure,
@@
 temperature ultimately causes spatially averaged aphid abundances to have
-lower skew than they would otherwise, because negative impacts of cold winter
+lower skew than they would otherwise, because negative impacts of cold winters
 are spatially synchronous.

### 2019-06-12 Paper.Rmd
starting revisions post Ecol Lett

@@
 ---
-title: "Copulas reveal complex and informative dependencies propagating throughout ecology"
+title: "Copulas and their potential for ecology"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
 date: ""
-geometry: "left=1.5cm,right=1.5cm,top=2.5cm,bottom=2.8cm"
+fontsize: 12pt
+geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm"
 
@@
 
-# Front matter
-
 \noindent \emph{Affiliations:}
 
-\noindent Ghosh: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
-
-\noindent Sheppard: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
+\noindent Ghosh, Sheppard, Bever: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
 
@@
 
-\noindent Bever: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66045, USA
-
 \noindent Reuman: Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas, Lawrence, KS, 66047, USA; Laboratory of Populations, Rockefeller University, 1230 York Ave, New York, NY, 10065, USA
@@
 
+Copulas reveal complex and informative dependencies propagating throughout ecology
+
 Copulas and their importance for ecology
@@
 
-Copulas and their potential for ecology
-
 Copulas and their potential throughout ecology
@@
 
-\noindent \emph{Keywords:} dependence, regression, correlation, copula, population, community
-
-\noindent \emph{Type of Article:} Ideas and Perspectives
-
-\noindent \emph{The number of words in the abstract:} 250
-
-\noindent \emph{The number of words in the main text (excluding abstract, acknowledgements, references, table and figure legends):} 7497
-
-\noindent \emph{The number of words in each text box:} No text boxes
-
-\noindent \emph{The number of references:} 34
-
-\noindent \emph{The number of figures, tables, and text boxes:} 7 figures, 3 tables, 0 text boxes
-
-\noindent \emph{The name and complete mailing address (including telephone and fax numbers and e-mail address) of the person to whom correspondence should be sent:} Daniel Reuman, 2101 Constant Ave, Lawrence, KS, 66047, reuman@ku.edu, 626 560 7084.
-
-\newpage
+\noindent \emph{Correspondence:} Daniel Reuman, 2101 Constant Ave, Lawrence, KS, 66047, reuman@ku.edu, 626 560 7084.
+
+\noindent \emph{Type of Submission:} Article
+
+\noindent \emph{The number of words in the abstract:}
 
@@
 
-\newpage
+\noindent \emph{Keywords:} dependence, regression, correlation, copula, population, community
+
 # Introduction\label{Intro}

### 2019-06-12 Paper.Rmd
edits to the Intro and master figure

@@
 
-The goal of this paper is to illustrate the potential for 
+The goal of this paper is to explore the potential for 
 applications of copulas in ecology. 
-Our results provide initial support for the following 
-overall hypothesis: environmental, ecological and evolutionary processes 
-generate and transmit copula structure and tail dependence, with substantial 
+Our results provide support for the idea that environmental, ecological and evolutionary processes 
+generate and transmit nontrivial copula structure and tail dependence, with substantial 
 consequences for theoretical and conceptual understanding, for inference, and for 
-applications across ecology. This is a broad hypothesis, and 
-even the substantial empirical and theoretical
-evidence we present here has only begun to test it. So we hope our work stimulates further
+applications across ecology. The work we present here has begun to test this
+hypothesis, and we hope our results stimulate further
 evaluation of the hypothesis by the research community.
-We examine our overall hypothesis by assessing several more specific 
-hypotheses, described below and outlined in Fig. 
-\ref{MasterFigure}, all of which bear upon the overall hypothesis.
 
@@
 \includegraphics[width=7in]{./MasterFigure.pdf}
-\caption{Guide for hypotheses addressed in the text and how each instantiates the
-overall hypothesis. Labels are defined 
-in the text to refer to parts of the figure. "Causal hypotheses" (i.e., hypotheses about 
-causes of copula structure) are labeled with letters A-D. "Consequential 
-hypotheses" (i.e., hypotheses about 
-consequences of copula structure) are labeled with X-Z. The "observational 
-hypothesis" (i.e., our hypothesis about 
-copula structure of data) is labeled O.\label{MasterFigure}}
+\caption{Summary and guide for analyses presented in the text. Middle boxes
+correspond to ecological datasets for which Q1 was examined (in the 
+second through
+fourth sections of the text, see also Table \ref{tab_data_info}). Upper boxes 
+correspond to potential causes of non-normal (NN) copula structure that
+were examined (in the fifth and sixth sections of the text). Lower boxes 
+correspond to potential consequences we examined of NN copula structure for
+ecological understanding and for applications (in the seventh and eighth
+sections of the text). Arrow labels (A-D for analyses pertaining 
+to causes, Q2; and X-Z for analyses pertaining to consequences, Q3) 
+link to locations in the text.\label{MasterFigure}}
 \end{center}
@@
 
-One way we may expect a copula approach 
+One way we may expect, *a priori*, a copula approach 
 to better reflect ecological relationships relates to Liebig's law of the
@@
 The portion of the distributions that dominate the association is visible in the copula
-approach, but not via standard techniques. We therefore frame "causal hypothesis" A 
-(labeled in Fig. \ref{MasterFigure}): 
-copulas of dependencies between environmental and 
-ecological variables will sensitively reveal 
-nonlinear, asymmetric and threshold effects for population, community, 
-and ecosystem functioning variables; copulas between 
-environmental and ecological variables
-will commonly show tail dependence; such tail dependence will lead to 
-tail dependence across space in ecological variables.
+approach, but not via standard techniques.
 
@@
 Synchrony is a phenomenon of great interest in 
-ecology [@Liebhold2004; @Sheppard2016; @walteretal2017]. We therefore frame 
-causal hypothesis B (again labeled in Fig. \ref{MasterFigure}): Moran effects
-transmit copula structure in spatial dependence from environmental to 
-ecological variables, and hence copula structure and tail dependence will be 
-common features of spatial synchrony in ecological variables.
+ecology [@Liebhold2004; @Sheppard2016; @walteretal2017].
 
@@
 distributions are principally responsible for the association between variables, and this 
-information can help support or refute alternative causal hypotheses. We
-frame causal hypothesis C (Fig. \ref{MasterFigure}): copulas of abundances
-through time or space of two competing or otherwise interacting species will
-reflect any asymmetries in the interaction. We will later
-frame causal hypothesis D, which concerns character evolution.
-
-Another reason we may expect copula structure to be important
+information can help support or refute alternative causal hypotheses.
+
+A fourth reason we may expect copula structure to be important
 in ecology has to do with spatially aggregated or averaged quantities. 
@@
 and the total economic value of a fishery is the sum of local catches. 
-We therefore frame "consequential hypothesis" X (labeled on Fig. \ref{MasterFigure}):
-details, revealed by copulas, of the 
-dependence between local quantities can influence the spatial 
-mean or sum, and this may influence higher organizational levels in 
-ecology and human concerns.
+We will explore how details, captured by copulas, of the dependence between local
+quantities can influence the spatial mean or sum, and how this may influence 
+higher organizational levels in eoclogy and human concerns.
 Illustrating the idea here, prior work demonstrated that the overall 
@@
 On the other hand left-tail dependence in a pest species, even if overall 
-correlation were unchanged, would have more benign effects. We will
-later frame consequential hypotheses Y, about extinction risk, and Z, about Taylor's law.
-
-Our causal and consequential hypotheses imply, or rely on, an "observational
-hypothesis" (O in Fig. \ref{MasterFigure}), which we here frame. Datasets in 
-ecology will commonly have copula structure
-distinct from that of a Gaussian/normal distribution (here called 
-non-normal copula structure); they will commonly show tail dependence 
-distinct from that of a normal distribution; in particular they will
-show asymmetric tail dependence.
-
-Because the observational hypothesis bears upon all the causal
-and consequential hypotheses, we test it first, in sections 2-4 of the paper.
-We analyze several datasets, including 
+correlation were unchanged, would have more benign effects.
+
+We approached our overall goal of exploring the potential of copula approaches 
+for ecology by addressing the following specific questions. (Q1) Do datasets
+in ecology have copula structure distinct from that of a Gaussian/normal
+distribution (here called non-normal copula structure)? Do they show tail 
+dependence distinct from that of a normal distribution, and in particular
+do they show asymmetric tail dependence? Normal copula structure is assumed
+by standard approaches that use multivariate 
+normal distributions or distributions obtained by transforming
+the marginals of normal distribution. Such 
+approaches do not utilize the power of copulas, so Q1 asks whether ecological
+data contain dependency information distinct from what can be modelled with
+standard approaches. (Q2) If the answer to
+Q1 is "yes," what are some possible causes/mechanisms of non-normal
+copula structure? And (Q3) what are the consequences of non-normal copula
+structure and tail dependence for our understanding of ecology and for 
+applications? To address Q1 (second through fourth sections of the paper), 
+we analyze several datasets, including 
 environmental, species-trait, phenological, population, community, and 
@@
 sub-fields and organizational levels of ecology. 
-Sections 5 and 6 test the causal hypotheses, and sections 7 and 8 test the
-consequential hypotheses. Section 9 is the Discussion.
+The fifth and sixth sections of the paper 
+address Q2 using simple models. The seventh and eigths sections address
+Q3 using both data and models. The ninth section is the Discussion.
+Multiple analyses are brought to bear on each question, and these
+are summarized diagramatically in Fig. \ref{MasterFigure}, which can 
+serve as a *post hoc* guide to the paper. 
 Copula approaches have been used to great effect in neuroscience [@onken2009], 
@@
 variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000]. 
-To our knowledge, these concepts have never been used in ecology. 
+To our knowledge, non-normal copulas have never been used in ecology. 
 We argue that benefits to ecology of using a complete 

### 2019-06-12 Paper.Rmd
still editing

@@
 \noindent \emph{The number of words in the abstract:}
+
+\newpage
 
@@
 considered. Many datasets exhibited asymmetric tail dependence, whereby two variables are more 
-strongly related in their left compared to right tails, or vice versa. We describe mechanisms by which 
+strongly related in their left compared to right tails, or *vice versa*. We describe mechanisms by which 
 observed copula structure and tail dependence can arise in ecological data, including a Moran-like 
@@
 \noindent \emph{Keywords:} dependence, regression, correlation, copula, population, community
+
+\newpage
 
@@
 of probability density functions (pdf) of Clayton (C) and survival-Clayton 
-(D) copulas with parameter 5 (see section on methods for the observational hypothesis), together
+(D) copulas with parameter 5 (see section on methods for Q1), together
 with samples from the corresponding distributions. (E,F) Reflections
 of C, D, respectively, about a vertical axis. (G) Contours of the
-pdf of a normal copula (methods for the observational hypothesis) with 
+pdf of a normal copula (methods for Q1) with 
 parameter chosen to
@@
 can be computed
-for any band (methods for the observational hypothesis) to describe how dependence
+for any band (methods for Q1) to describe how dependence
 varies from one part of the two distributions to another. Diagonal lines
@@
 (Fig. \ref{fig_cop_pedag1}A, C);
-and \emph{vice versa} if larger values are more strongly associated than smaller
+and *vice versa* if larger values are more strongly associated than smaller
 ones (Fig. \ref{fig_cop_pedag1}B, D). Datasets of the same 
@@
 (Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see section on
-methods for the observational hypothesis).
+methods for Q1).
 
@@
 applications of copulas in ecology. 
-Our results provide support for the idea that environmental, ecological and evolutionary processes 
+Our results provide support for the hypothesis that environmental, ecological and evolutionary processes 
 generate and transmit nontrivial copula structure and tail dependence, with substantial 
@@
 fourth sections of the text, see also Table \ref{tab_data_info}). Upper boxes 
-correspond to potential causes of non-normal (NN) copula structure that
+correspond to potential causes of non-normal (non-N) copula structure that
 were examined (in the fifth and sixth sections of the text). Lower boxes 
-correspond to potential consequences we examined of NN copula structure for
+correspond to potential consequences we examined of non-N copula structure for
 ecological understanding and for applications (in the seventh and eighth
@@
 
-A fourth reason we may expect copula structure to be important
-in ecology has to do with spatially aggregated or averaged quantities. 
+A fourth reason we may expect a copula approach to be important
+for ecology has to do with spatially aggregated or averaged quantities. 
 Many ecological variables of applied importance 
@@
 quantities can influence the spatial mean or sum, and how this may influence 
-higher organizational levels in eoclogy and human concerns.
+higher organizational levels in ecology and human concerns.
 Illustrating the idea here, prior work demonstrated that the overall 
@@
 of the dependence between wind speeds at multiple generator sites [@li2013]. 
-Spatially aggregated ecological variables may show a similar dependence, with
+Spatially aggregated ecological variables may be subject to a similar 
+effect, with
 potentially substantial consequences. For instance, if populations of a 
@@
 normal distributions or distributions obtained by transforming
-the marginals of normal distribution. Such 
+the marginals of a normal distribution. Such 
 approaches do not utilize the power of copulas, so Q1 asks whether ecological
@@
 structure and tail dependence for our understanding of ecology and for 
-applications? To address Q1 (second through fourth sections of the paper), 
+applications? To address Q1 (the second through fourth sections of the paper), 
 we analyze several datasets, including 
@@
 sub-fields and organizational levels of ecology. 
-The fifth and sixth sections of the paper 
-address Q2 using simple models. The seventh and eigths sections address
-Q3 using both data and models. The ninth section is the Discussion.
+To address Q2 (the fifth and sixth sections of the paper) we use 
+simple models. To address Q3 (the seventh and eigths sections) we use 
+both data and models. The ninth section is the Discussion.
 Multiple analyses are brought to bear on each question, and these
@@
 variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000]. 
-To our knowledge, non-normal copulas have never been used in ecology. 
+To our knowledge, non-normal copulas have not been used in ecology. 
 We argue that benefits to ecology of using a complete 
@@
 Datasets included environmental, species-trait, phenological, 
-population, community, and ecosystem functioning data (Table \ref{tab_data_info}). Copula structure describing 
+population, community, and ecosystem functioning data (Table \ref{tab_data_info}), selected to span multiple fields and levels
+of organization within ecology. Copula structure describing 
 the relationship between atmospheric weather variables such as rainfall or wind speed 
@@
 
-The environmental, species-trait, and community datasets are henceforth 
-called the "bivariate" datasets in the sense that they comprise two quantities measured at 
+Our environmental, species-trait, and community datasets happen to be 
+"bivariate" datasets in the sense that they comprise two quantities measured at 
 different locations or for different species (Table \ref{tab_data_info}). 
-The phenology, population, and ecosystem functioning datasets are 
+Our phenology, population, and ecosystem functioning datasets are 
 "multivariate" in that they comprise, for each taxon, measurements through time at multiple locations; 
@@
 
-# Background, concepts and methods for testing the observational hypothesis\label{Methods}
+# Concepts and methods for Q1 \label{Methods}
 
@@
 
-We introduce the copula families we used, and 
-measures of tail dependence.
+We introduce the copula families and 
+measures of tail dependence we used.
 The normal copula family (henceforth N) is a one-parameter
@@
 Copulas are implemented in the R packages `copula` and `VineCopula`. 
-Visualizations are in 
-Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8}. 
+Visualizations of these families (Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8})
+can be compared with normalized rank plots to estimate visually 
+which families may effectively model a given bivariate dataset. 
 If LT and UT denote 
@@
 
-We tested the observational hypothesis (Fig. \ref{MasterFigure}, O) 
-principally via a model selection procedure in which the 
+We addressed Q1 principally via a model selection procedure in which the 
 copulas listed above were fitted to 
@@
 
-The first part of the observational hypothesis was tested for a bivariate dataset 
+The first part of Q1 was answered for a bivariate dataset 
 by comparing the AIC weights of the non-N copula
 families with that of the N family. 
-The second part was tested by comparing 
+The second part of Q1 was answered by comparing 
 the model-averaged lower- and upper-tail dependence
 values to the lower- and upper-tail dependence values of an N copula (i.e., 0). 
-The third part 
-was tested by comparing the difference of the model-averaged lower- and upper-tail 
+The third part of Q1
+was answered by comparing the difference of the model-averaged lower- and upper-tail 
 dependence values to the analogous difference for an N copula, again 0.
@@
 
-In addition to our model selection approach, we also used nonparametric approaches to 
-provide greater 
-confidence in our tests of the observational hypothesis. We used three 
+In addition to our model selection approach, we also used 
+nonparametric approaches to provide greater 
+confidence in our answer to Q1. We used three 
 statistics which quantify the extent to which 
@@
 
-# Results of testing the observational hypothesis: Ecological datasets often have non-normal copula structure and asymmetric tail dependence \label{Results}
+# Results for Q1: Ecological datasets often have non-normal copula structure and asymmetric tail dependence \label{Results}
 
@@
 commonly showed asymmetric tail 
-dependence: the observational hypothesis (Fig. \ref{MasterFigure}) was supported. To make results easier to absorb, we first 
+dependence, answering Q1 in the affirmative. 
+To make results easier to absorb, we first 
 present results for an example bivariate dataset, then we summarize results
@@
 minimum-AIC copula, with AIC weight `r round(max(BivMS_res_Loecke$InfCritRes$AICw),2)`, whereas the 
-N copula had AIC weight essentially 0, confirming the 
-first part of the observational hypothesis for these data. 
+N copula had AIC weight essentially 0, answering the 
+first part of Q1 for these data. 
 Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
 and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
-These values are distinct from what an N copula would give, namely, 0, confirming the second part 
-of the observational hypothesis. The values also differed substantially 
-from each other, confirming the third part of the observational hypothesis. 
+These values are distinct from what an N copula would give, namely, 0, answering the second part 
+Q2 for these data. The values also differed substantially 
+from each other, answering the third part of Q1. 
 The numbers reflect stronger upper- than lower-tail
@@
 was asymmetric (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}, 
-Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}).
+Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}). The 
+affirmative, empirically based answer to Q1 for the soil C and N data is 
+represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
+in the right-most box in the middle row.
 
 <!--bivariate summary-->
-Table \ref{tab_bivar_summary} summarizes results of testing the observational hypothesis 
-using the bivariate datasets. The two 
+Table \ref{tab_bivar_summary} summarizes Q1 results 
+for the bivariate datasets. The two 
 variables were significantly related for all datasets 
@@
 The N copula was always a poor fit compared to the highest-AIC-weight copula 
-(Table \ref{tab_bivar_summary}, rows 3-4), confirming 
-the first part of the observational hypothesis. 
+(Table \ref{tab_bivar_summary}, rows 3-4), answering the first
+part of Q1. 
 Lower- and upper-tail dependence statistics often differed substantially from 0 
 (Table \ref{tab_bivar_summary}, rows 7-8) and from each other (Table \ref{tab_bivar_summary}, row 9), 
-confirming the second and third parts of the observational hypothesis. 
+answering the second and third parts of Q1. 
 The first three datasets had stronger upper- 
@@
 (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}).
+The 
+affirmative, empirically based answer to Q1 for the bivariate datasets is 
+represented in Fig. \ref{MasterFigure} as solid outlines around the names of those
+datasets in the middle row of boxes.
 We did the same analysis 
@@
 were typically substantially 
-larger than N-copula weights (Table \ref{SM-tab_count_AICw_normal}), confirming 
-the first part of the observational hypothesis. Goodness of fit tests 
+larger than N-copula weights (Table \ref{SM-tab_count_AICw_normal}), answering the first part of Q1. Goodness of fit tests 
 (Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
@@
 positive and were frequently large, thereby differing 
-from what an N copula would give (i.e., 0), and confirming the second part of 
-the observational hypothesis. These data have greater tail dependence 
+from what an N copula would give (i.e., 0), and answering the second
+part of Q1. These data have greater tail dependence 
 (lower and upper) than would be expected from an N-copula null hypothesis. 
@@
 out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
-than the spatial synchrony of outbreaks. This also confirms the third part of the 
-observational hypothesis.
+than the spatial synchrony of outbreaks. This also answers the third part of Q1.
 Nonparametric statistics verified that tail dependence was asymmetric for the green 
@@
 
-<!--***DAN: You could maybe move the next para to sup mat, though that would complicate the
-presentation of skewness results below. Only do it if you are desperate to cut.-->
 <!--multivariate summary-->
-Table \ref{tab_multivar_summary} summarizes results of testing the observational hypothesis for
+Table \ref{tab_multivar_summary} summarizes Q1 results for
 multivariate data. Results supported the 
 conclusions that non-N copula structure, non-N tail dependence, and asymmetric tail 
-dependence were common, supporting the observational hypothesis. 
+dependence were common, answering Q1 in the affirmative for these data. 
 Most location pairs were non-independent 
@@
 Figs \ref{SM-fig_aphid_count}-\ref{SM-fig_methane}.
+The 
+affirmative, empirically based answer to Q1 for the multivariate datasets is 
+represented in Fig. \ref{MasterFigure} as solid outlines around the names
+of those datasets in the middle row of boxes.
 We also carried out the same analyses for abundance and first-flight data for the 18 
@@
              format="latex",
-             caption = "Summary of results of testing the observational hypothesis for bivariate datasets.
+             caption = "Summary of Q1 results for bivariate datasets.
 The $p$-values in rows 5-6 are for the minimum-AIC copula.
@@
              caption.short = "Summary of results for multivariate datasets",
-             caption = "Summary of results of testing the observational hypothesis for multivariate datasets.
+             caption = "Summary of Q1 results for multivariate datasets.
 Rows 3-9 of the table were computed for the non-independent pairs (row 2) only.
@@
 <!--Question 2 analyses-->
-# Background, concepts and methods for testing the causal hypotheses \label{Causes}
+# Concepts and methods for Q2 \label{Causes}
+
+<!--**DAN: stopped here-->
 
 Having demonstrated that non-N copula structure and asymmetric tail dependence are 
-common, we addressed the causal hypotheses (Fig. \ref{MasterFigure}, A-D)
-using models. We did not systematically 
+common in ecological variables, we 
+addressed Q2 by exploring, using models, three possible mechanisms
+by which these phenomena may arise. 
+The first mechanism, summarized conceptually in the Introduction, relates to 
+Liebig's law of the minimum and to nonlinear influences of environmental
+variables on ecological variables.
+
+We did not systematically 
 test the first two parts of A 
@@
 
-# Results of testing the causal hypotheses: Moran effects and asymmetric dependencies produce non-normal copula structure\label{Results_Q2}
+# Results for Q2: Moran effects and asymmetric dependencies produce non-normal copula structure\label{Results_Q2}
 
@@
 <!--Question 3 analyses-->
-# Background, concepts and methods for testing the consequential hypotheses\label{Consequences}
+# Concepts and methods for Q3 \label{Consequences}
 
@@
 
-# Results of testing the consequential hypotheses: Tail dependence influences skewness of the spatial average, extinction risk, and Taylor's law
+# Results for Q3: Tail dependence influences skewness of the spatial average, extinction risk, and Taylor's law
 
@@
 \clearpage
+\newpage
 

### 2019-06-13 Paper.Rmd
done with round 1 of prepping for Ecol Mongr, going through it one more time now

@@
 
-\noindent \emph{Type of Submission:} Article
+\noindent \emph{Type of Submission:} Articles, or Concepts and Synthesis
 
@@
 All branches of ecology study relationships among environmental and biological variables. However, 
-ubiquitously used approaches to studying such relationships, based on correlation and regression, 
+standard approaches to studying such relationships, based on correlation and regression, 
 provide only a small slice of the complex information contained in the relationships. Other 
@@
 under Liebig's law of the minimum. Copula structure and tail 
-dependence can also reveal causal relationships between variables. We describe consequences of copula structure, including impacts 
+dependence can also help reveal causal relationships between variables. We also describe consequences of copula structure for ecological phenomena, including impacts 
 on extinction risk and the stability through time of ecosystem services. By documenting the 
 importance of a complete description of dependence, advancing conceptual frameworks, and 
-demonstrating a powerful approach, we aim to elevate copula reasoning to widespread use in ecology.
-
-\noindent \emph{Keywords:} dependence, regression, correlation, copula, population, community
+demonstrating a powerful approach, we aim to elevate copula reasoning to widespread use in ecology, benefiting the discipline.
+
+\noindent \emph{Keywords:} dependence, regression, correlation, copula, population, community, ecosystem functioning
 
@@
 copula description of dependence will be substantial.
+
+\newpage
 
@@
              booktabs = T,linesep = "\\addlinespace") %>%
-             column_spec(1:4,width="2.8 cm")%>%
+             column_spec(1:4,width="3.4 cm")%>%
              row_spec(c(4,5,6,8), bold=T)
 ```
+
+\newpage
 
@@
 
-# Results for Q1: Ecological datasets often have non-normal copula structure and asymmetric tail dependence \label{Results}
+\newpage
+
+# Results for Q1: Ecological datasets have non-normal copula structure and asymmetric tail dependence \label{Results}
 
@@
 
-two_tab_in1 <- as.data.frame(do.call(cbind, lapply(1:ncol(tab1), function(i) paste0(" (",tab1[c(7:8,10:12),i], " , ",tab2[ , i], ")") )))   # see carefully if you rearrange tab_multivar_summary's row in suppmat this will change
+two_tab_in1 <- as.data.frame(do.call(cbind, lapply(1:ncol(tab1), function(i) paste0(" (",tab1[c(7:8,10:12),i], ", ",tab2[ , i], ")") )))   # see carefully if you rearrange tab_multivar_summary's row in suppmat this will change
 
@@
              )%>%column_spec(2:5,width="2.5cm")%>%
-             column_spec(1, width = "7cm")%>%kable_styling(latex_options = "basic",position = "center")
+             column_spec(1, width = "6cm")%>%kable_styling(latex_options = "basic",position = "center")
 ```
+
+\newpage
 
@@
 # Concepts and methods for Q2 \label{Causes}
-
-<!--**DAN: stopped here-->
 
@@
 by which these phenomena may arise. 
-The first mechanism, summarized conceptually in the Introduction, relates to 
-Liebig's law of the minimum and to nonlinear influences of environmental
+The first mechanism relates to the ideas in the Introduction about 
+Liebig's law of the minimum, and to nonlinear influences of environmental
 variables on ecological variables.
-
-We did not systematically 
-test the first two parts of A 
-because nonlinear and thresold effects of 
-environmental variables on ecological variables are already well documented 
-(however, see Discussion material about the green spruce aphid). To test 
-the third part of hypothesis A, we formulated a model in which 
-an environmental variable influences an ecological variable
-more strongly in one of its tails, and explored whether the ecological variable
-could then exhibit asymmetric tail dependence across space.
+If an environmental variables influences an ecological variable 
+more strongly in one of its tails, we explored whether the ecological
+variable could then exhibit asymmetric tail dependence across space.
 Let $E_i(t)$ be an ecological variable 
@@
 means slower return to equilibrium after a disturbance.
-Because $(\epsilon_1,\epsilon_2)$
-and $(\delta_1,\delta_2)$ have N copula structure, the Moran 
-mechanism analyzed below (Fig. \ref{MasterFigure}, B) 
-does not operate here.
 
@@
 steps. We applied our model selection algorithms and 
-non-parametric statistics (methods for the observational hypothesis) to these 
+non-parametric statistics (methods for Q1) to these 
 outputs to discover if the model could
 produce non-N copula structure and asymmetric tail dependence in the 
-$E_i(t)$.
-
-Hypothesis B is an extension of the well known Moran
+$E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
+and $(\delta_1,\delta_2)$ have N copula structure, the Moran 
+mechanism analyzed below 
+does not operate here.
+
+Our second mechanism is an extension of the well known Moran
 effect, and was summarized conceptually in the Introduction. 
@@
 
-To explore the importance of nonlinear dynamics for hypothesis B, we also considered 
-nonlinear population models based on $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right) \right)+\sigma\epsilon_i(t))$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
+To explore the importance of nonlinear dynamics for our proposed mechanism, 
+we also considered 
+nonlinear population models based on $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
 $\sigma=1$. When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
@@
 $(\epsilon_1(t),\epsilon_2(t))$ from the
-bivariate random variable with the given copula family and the given $\tau$.
+bivariate random variable with the given copula family and the given Kendall
+correlation $\tau$.
 We then used this noise to drive the model, and retained both the
@@
 time steps. For each simulation, the following
-statistics were computed for both noises and populations: 
+statistics were then computed for both noises and populations: 
 Pearson, Spearman and Kendall correlations, $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, 
 $\Dsq_l$, $\Dsq_u$, $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$. 
-Values were plotted against $\tau$ for noises and populations. If causal hypothesis B is
-correct, plots should be similar for populations and noises.
-
-Hypothesis C was not tested. Our next 
-causal hypothesis, Fig. \ref{MasterFigure}, D, 
-is evolutionary, and pertains to bivariate 
+Values were plotted against $\tau$ for noises and populations. If the 
+hypothesis was correct that characteristics of 
+the copula structure of spatial dependence in an
+ecological variable can be inherited from the copula structure of spatial 
+dependence in an environmental variables, then 
+plots should be similar for populations and noises.
+
+Our next mechanism is evolutionary, and pertains to bivariate 
 trait data across species, e.g., our bird and mammal 
@@
 The hypothesis is that tail dependence occurs in evolutionary changes
-in bivariate characters, and gives rise to tail dependence in observed character values
-across extant species.
+in bivariate characters, and gives rise to tail dependence in the 
+character values observed across extant species.
 We simulated bivariate character evolution on an estimate of 
@@
 correlation $0.875$ between components, so our simulations assess 
-the impact of copula structure only. For each of the five copulas, mammalian character
+the impact of copula structure only. For 
+each of the five copulas, mammalian character
 evolution was simulated 100 times. For each simulation, symmetry of tail dependence 
@@
 
+\newpage
+
 # Results for Q2: Moran effects and asymmetric dependencies produce non-normal copula structure\label{Results_Q2}
@@
 could produce outputs with visually apparent asymmetry 
-of tail dependence, to an extent that depended on the value of $b$,
-thus confirming the third part of hypothesis A for our model. 
-For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ (Fig. \ref{SM-fig_ms_asym_b_-0.1}), 
+of tail dependence, to an extent that depended on the value of $b$;
+thus this is a mechanism that may produce non-N copula structure and
+asymmetric tail dependence across space in ecological variables.
+This result is represented in Fig. \ref{MasterFigure} as the solid box around 
+"Nonlinear environmental effects, Liebig's law" and the arrows labelled "A".
+It is explained in the Discussion why the box and some of those arrows 
+are solid, instead of dashed, although the results we 
+present in this section are theoretical.
+For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
+(Fig. \ref{SM-fig_ms_asym_b_-0.1}), 
 for both $g_1$ and $g_2$,
@@
 effects of noise ($g_1$, respectively, $g_2$) 
-produced lower-tail (respectively, upper-tail) spatial dependence in the ecological variable.
+produced lower-tail (respectively, upper-tail) spatial 
+dependence in the ecological variable.
 Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
@@
 
-Causal hypothesis B (Fig. \ref{MasterFigure}) was largely confirmed; 
+The hypothesis that characteristics of 
+spatial copula structure and tail dependence
+in an ecological variable can be inherited from an environmental 
+variable was largely confirmed; 
 however, similarities between environmental- and 
-ecological-variable copula structure were reduced when dynamics were strongly nonlinear.
+ecological-variable copula structure were reduced when 
+dynamics were strongly nonlinear.
+Thus this is a mechanism that may produce non-N copula structure and
+asymmetric tail dependence across space in ecological variables.
+This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
+"Moran effects" and the arrows labelled "B".
 For simulations using AR(1) models, copula 
@@
 Though there were again significant but small differences for some statistics
-and simulations, the hypothesis was supported that copula structure in the noise translates
+and simulations, the hypothesis was supported that copula 
+structure in the noise translates
 to similar but not identical copula structure in output populations. 
@@
 expectation that there will be a Moran-effect-like correspondence 
-between noise and model-output copula structure for typical ecological dynamics,
+between noise and model-output copula structure across space 
+for typical ecological dynamics,
 as long as noise is small enough that dynamics stay relatively close
@@
 for this model. The 
-AR(1) model and a Clayton copula were used (see methods section for the causal hypotheses).
+AR(1) model and a Clayton copula were used (see methods section for Q2).
 Each point is the mean across 50 replicate simulations for which the same 
@@
 Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1} for
-the nonlinear models described in the methods for the causal hypotheses.\label{fig_Cause4copula_scatter_C_kend_ar1}}
+the nonlinear models described in the methods section for Q2.\label{fig_Cause4copula_scatter_C_kend_ar1}}
 \end{center}
@@
 and for a few simulations, tail dependence asymmetry statistics
-even had opposite signs for noise and model outputs 
+had opposite signs for noise and model outputs 
 (Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). For instance, using a C copula with a large Kendall $\tau$,
@@
 \ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). Overall,
-similarities between noise and model-output copula structure
+similarities between noise and model-output copula statistics
 were more noticable than discrepancies, in spite of nonlinearities.
 We repeated the analysis with $r=1.3$. Discrepancies between noise and population
-copula structure in this case were often glaring. Apparently noise of standard deviation 1
-interacted strongly with model nonlinearities when the model was in its overcompensatory 
+copula structure in this case were often glaring. Apparently 
+noise of standard deviation 1 interacted especially strongly with 
+model nonlinearities when the model was in its overcompensatory 
 regime.
 
-For our model, hypothesis D (Fig. \ref{MasterFigure}) was 
+For our evolutionary model, the hypothesis was 
 correct that asymmetric tail dependence in evolutionary changes 
-can produce similarly asymmetric tail dependence in character distributions across
-phylogeny tips (Appendix \ref{SM-evol_mech}, Fig. \ref{SM-fig_cop_evo} for details). This topic 
-is revisited in the Discussion.
+can produce similarly asymmetric tail dependence between characters across
+phylogeny tips (Appendix \ref{SM-evol_mech}, Fig. \ref{SM-fig_cop_evo} 
+for details). 
+Thus this is a mechanism that may produce non-N copula structure and
+asymmetric tail dependence between characters.
+This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
+"If character evolution has copula structure" and the arrow labelled "D".
+This topic is revisited in the Discussion, as is the box "Asymmetric species
+interactions" and the arrow labelled "C" in Fig. \ref{MasterFigure}.
+
+\newpage
 
@@
 
-The hypothesis was presented that
-the distribution (through time) of a spatially averaged quantity should be influenced by 
+The hypothesis was presented in the Introduction that
+the distribution (through time) of a spatially 
+averaged quantity should be influenced by 
 dependencies between the local quantities, including their
-copula structure and tail dependence (Fig. \ref{MasterFigure}, X). 
+copula structure and tail dependence. 
 To refine this hypothesis, suppose an ecological variable 
@@
 
-We tested hypothesis X 
+To help answer Q3, we tested the above hypothesis
 using our multivariate datasets (Table \ref{tab_multivar_summary}).
@@
 in a special way to have the 
-copula structure of a multivariate-normal distribution, but to retain exactly the same
+copula structure of a multivariate-normal distribution, but 
+to retain exactly the same
 distributions of values for each sampling location as the original data
 (Appendix \ref{SM-Surrog_multi}). 
-Surrogates also had very similar Spearman correlations between pairs of sampling locations
+Surrogates also had very similar Spearman correlations 
+between pairs of sampling locations
 as the data. 
-Our comparisons therefore tested the null hypothesis that the skewness of the spatial 
-mean took values on the empirical data no different from what one would expect if the copula
+Our comparisons therefore tested the null 
+hypothesis that the skewness of the spatial 
+mean took values on the empirical data no different 
+from what one would expect if the copula
 structure of the data were multivariate N, but the data were otherwise statistically
 unchanged. Significant differences indicate that non-N copula structure in the data 
-contributed to the skewness of the spatial mean, i.e., to its "spikiness".
+contributed to the skewness of the spatial mean time series, 
+i.e., to its "spikiness".
 
@@
 
-We also examined hypothesis Y (Fig. \ref{MasterFigure}), i.e., that spatial 
+To further address Q3, we also examined a hypothesis that spatial 
 tail dependence of an environmental variable can influence the 
@@
 variable are "bad" for the populations and high values are "good".
-We tested the reasonableness of Y using a metapopulation
+We tested the reasonableness of this hypothesis using a metapopulation
 extension of the Lewontin-Cohen model,
@@
 
+\newpage
+
 # Results for Q3: Tail dependence influences skewness of the spatial average, extinction risk, and Taylor's law
 
-\noindent Skewness results were consistent with 
-hypothesis X. For datasets that
+\noindent Results were consistent with the hypothesis that the skewness
+of a spatial-average time series is influenced by the tail dependence
+between the local quantities that were averaged. Thus copula 
+structure and tail dependence are important for spatially averaged 
+quantities. This result is represented in Fig. \ref{MasterFigure}
+as the solid box around "Instability/skewness of mean or total 
+time series" and the solid arrows labelled "X". For datasets that
 exhibited strong lower-tail (respectively, upper-tail) 
@@
 (Fig. \ref{fig_skewness_spearman}A,B). For datasets with moderate lower-tail 
-dependence (\emph{Ceratium furca} abundance and methane data), skewness of the spatial
+dependence (\emph{Ceratium furca} abundance and methane data), 
+skewness of the spatial
 average showed a non-significant or marginally significant tendency 
@@
 
-Consistent with hypothesis Y, 
-left-tail dependent environmental fluctuations increased metapopulation
+Consistent with our extinction risk hypothesis, 
+left-tail-dependent environmental fluctuations increased metapopulation
 extinction risk for the spatial 
@@
 (Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
-dispersal. We also substantiated using models (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd}, Fig. \ref{SM-fig_spTL})
-a consequential hypothesis Z (Fig. \ref{MasterFigure}), that 
+dispersal. This result is represented in Fig. \ref{MasterFigure}
+as the dashed box around "Extinction risk" and the dashed arrow labelled "Y".
+We also substantiated using models (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd}, Fig. \ref{SM-fig_spTL})
+a hypothesis (Fig. \ref{MasterFigure}, arrow labelled "Z"), that 
 Taylor's law can be influenced by
@@
 
+\newpage
+
 # Discussion \label{Discussion}
@@
 common across multiple sub-disciplines
-in ecology, although no 
-research we know of has 
-taken this structure into account. We hypothesized four 
-mechanisms that may cause non-N copula structures.
-We discuss below how commonly some of our
+in ecology, although no research we know of has 
+taken this structure into account. We hypothesized 
+mechanisms that may cause non-N copula structures and tail dependence;
+we discuss below how commonly some of our
 mechanisms may opperate. 
 We also hypothesized and demonstrated important consequences of 
-non-N copula structure and tail-dependence. 
+non-N copula structure and tail-dependence for the field of ecology. 
 For instance, the skewness
@@
 Thus tail dependence could have implications for exploited and pest species.
-Extinction risk can be altered by tail dependence
+Extinction risk can also be altered by tail dependence across space
 in environmental variables. 
-In our view, our results make our overall hypothesis (Introduction)
-reasonable and worthy of broader testing. We illustrated 
-how such tests might be carried out by performing tests ourselves on 
-example datasets and models. Further suggestions for tests are below.
-It seems likely that comprehensive understanding 
-of many ecological phenomena is impossible without 
+In our view, our results make it reasonable to suggest 
+that comprehensive understanding 
+of many ecological phenomena may be impossible without 
 a complete copula characterization of dependence.
@@
 (Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
-that build upon existing methods.
+that build upon existing copula methods. 
+To our knowledge, only N copulas have 
+been used in ecology [@Valpine2014; @Anderson2018; @Popovic2018].
+Such use is equivalent to using normal distributions with transformed
+marginals; i.e., it is equivalent to an approach not using copulas at all, and
+therefore does not capitalize on the power of copula techniques.
 
@@
 First, Liebig's law and the idea of limiting 
-nutrients are dominant paradigms in ecology.
-Second, although we did not systematically test the first two parts of 
-hypothesis A, many studies have documented nonlinear or threshold influences of environmental
+nutrients are dominant paradigms in ecology, and 
+many studies have documented nonlinear or 
+threshold influences of environmental
 variables on ecological quantities. 
-These first two parts of A could be tested for numerous 
-datasets by examining copulas for environmental and ecological
-variables. See below discussion about the green spruce aphid, and 
-Appendix \ref{SM-sect:greenspruce}, which illustrate how to do this.
-Causal mechanism C, which we also did not test, could be tested
-by analyzing copulas of abundances of competing species, sampled across space or
-time.
+Second, fluctuations in environmental variables through time
+are very commonly correlated across space. 
+Because these factors, which are the essential ingredients of
+the mechanism, are so common, it is reasonable to suppose the mechanism
+opperates commonly. It may be a dominant cause of 
+tail depencence and non-N copula struture of ecological dependencies 
+across space.
+We provide some empirical support for the mechanism 
+in our discussion of green sprice aphids and
+winter temperature below.
 
 <!--para on how likely one is to get copula structure via the Moran mechanism-->
-There are also reasons to expect our Moran mechanism (Fig. \ref{MasterFigure}, B) 
+There are also reasons to expect that 
+our Moran mechanism (Fig. \ref{MasterFigure}, B) 
 operates commonly: Moran effects are common [@Liebhold2004], and 
@@
 over the North Sikkim Himalayas using spatial datasets [@Goswami2018].
+
+<!--Causal mechanism C, which we also did not test-->
+We suggested in the Introduction that asymmetric competitive relationships
+between species could yield tail dependence between abundance data for the 
+species. This is another theoretical mechanism for non-N copula
+structure and tail dependence, represented in Fig. \ref{MasterFigure} 
+by the box "Asymmetric species interactions" and the arrow labelled "C".
+It could be tested by analyzing copulas of abundances of competing 
+species, sampled across space or time.
 
@@
 delayed density dependence, dispersal, population stage structure, trophic
-interactions, etc.
+interactions, etc., and we did not comprehensively 
+explore parameter space.
 We hope by enumerating a few potential mechanisms of copula structure,
 we inspire additional research on the
-potentially numerous mechanisms that may operate in diverse datasets.
+potentially numerous mechanisms that may operate in diverse datasets,
+and their relative importance under different circumstances.
 
@@
 <!--green spruce aphid all the way from causes to consequences-->
-Our hypotheses cover propagation of copula structure from causes, 
-to data, to consequences.
-We here take a closer look at the green spruce aphid, which simultaneously illustrates 
-all these stages.
+Our hypotheses and results cover the presence, causes, and consequences
+of non-N copula structure and tail dependence in ecological 
+datasets (Fig. \ref{MasterFigure}).
+We here take a closer look at the green spruce aphid, which simultaneously 
+illustrates causes and consequences.
 Green spruce aphid abundance, as measured in the data we use,
 is strongly influenced by temperature of the previous winter (@Sheppard2016,
-supplementary fig. 6).
+their supplementary fig. 6).
 For each of our 10 sampling locations, we therefore 
 examined (Appendix \ref{SM-sect:greenspruce}, 
-Fig. \ref{SM-fig:greenspruce}, Table \ref{SM-tab:greenspruce}) the copula of winter temperature
+Fig. \ref{SM-fig:greenspruce}, Table \ref{SM-tab:greenspruce}) the 
+copula of winter temperature
 and aphid abundance time series for the location, finding 
-left-tail dependence: apparently winter temperature has an asymmetric
-influence on abundance in that cold winters produce low abundances
+left-tail dependence. Apparently winter temperature has an asymmetric
+influence on aphid abundance in that cold winters produce low abundances
 but warm winters often do not yield higher abundances than moderate winters.
-Although we did not systematically test the first two 
-parts of hypothesis A due to much existing evidence, this result 
-illustrates how those parts can be tested using copula
-techniques. The third part of hypothesis A, which our modelling results
-supported, therefore suggests spatial dependence between green spruce aphid counts
-in different locations should show left-tail dependence - exactly as observed 
-(Table \ref{tab_multivar_summary}). And the consequences of such tail 
-dependence for the skewness of spatially averaged aphid counts was described previously 
+One of our hypothesized mechanisms (Fig. \ref{MasterFigure}, A), 
+which our modelling results
+supported (section on results for Q2), therefore suggests that spatial 
+dependence between green spruce aphid counts
+in different locations should show left-tail dependence. This is 
+exactly what was observed (Table \ref{tab_multivar_summary}), providing 
+empirical evidence supporting the mechanism. The consequences of such tail 
+dependence for the skewness of spatially averaged aphid counts was 
+described previously 
 (Fig. \ref{fig_skewness_spearman}A). Thus the asymmetric influence of winter 
-temperature ultimately causes spatially averaged aphid abundances to have
-lower skew than they would otherwise, because negative impacts of cold winters
-are spatially synchronous.
+temperature ultimately causes spatially averaged aphid abundance time series 
+to have lower skewness (i.e., less spikiness, greater stability through time) 
+than they would otherwise.
 
@@
 
-<!--***DAN: could remove the next para if you need to save words-->
 <!--A para that sets up Shya's extinction risk paper-->
@@
 matrix modelling [@Caswell2000; @MorrisDoak2002]
-uses a framework in which a single population's vital rates (e.g., life-stage-specific 
+uses a framework in which a single population's 
+vital rates (e.g., life-stage-specific 
 fecundity and survival rates) are considered to
@@
 Climate change is known to amplify the factors that lead to extreme weather events
-[@Hansen2012] and hence may alter tail dependence.
+[@Hansen2012] and hence may alter spatial tail dependence for weather variables.
+
+\newpage
 
@@
 # References
-\setlength{\parindent}{-0.2in}
+
+<div id="refs"></div>
+
+<!--\setlength{\parindent}{-0.2in}
 \setlength{\leftskip}{0.2in}
@@
 \noindent
-\sloppy
+\sloppy-->
+
+\clearpage
+\newpage
+
+# Figure captions
+
+\noindent \textbf{Figure 1.} (A-D) Bivariate datasets showing diverse relationships between 
+variables; many of these datasets nevertheless have the same Pearson (P)
+or Spearman (S) correlation coefficients. (C, D) Contours 
+of probability density functions (pdf) of Clayton (C) and survival-Clayton 
+(D) copulas with parameter 5 (see section on methods for Q1), together
+with samples from the corresponding distributions. (E,F) Reflections
+of C, D, respectively, about a vertical axis. (G) Contours of the
+pdf of a normal copula (methods for Q1) with 
+parameter chosen to
+have the same Spearman correlation as A-D, together with a sample.
+(H) The portion of the Spearman correlation within a band 
+can be computed
+for any band (methods for Q1) to describe how dependence
+varies from one part of the two distributions to another. Diagonal lines
+show two bands, the data in the left/lower band showing stronger
+dependence than the right/upper band.
+Subscripts h and v 
+stand for "horizontal" and "vertical".
+
+\hspace{1in}
+
+\noindent \textbf{Figure 2.} Summary and guide for analyses presented in the text. Middle boxes
+correspond to ecological datasets for which Q1 was examined (in the 
+second through
+fourth sections of the text, see also Table \ref{tab_data_info}). Upper boxes 
+correspond to potential causes of non-normal (non-N) copula structure that
+were examined (in the fifth and sixth sections of the text). Lower boxes 
+correspond to potential consequences we examined of non-N copula structure for
+ecological understanding and for applications (in the seventh and eighth
+sections of the text). Arrow labels (A-D for analyses pertaining 
+to causes, Q2; and X-Z for analyses pertaining to consequences, Q3) 
+link to locations in the text.
+
+\hspace{1in}
+
+\noindent \textbf{Figure 3.} Upper panels show raw data plots for (A) $\log_{10}$(soil C) and $\log_{10}$(soil N) data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
+for birds, (C) the same for mammals, and (D) above-ground plant biomass vs.
+Shannon's index (\textit{H}) from Cedar Creek.
+Bottom panels show the corresponding copula plots.
+See Data for units used in upper panels.
+
+\hspace{1in}
+
+\noindent \textbf{Figure 4.} If environmental effects operate asymmetrically in their tails on
+ecological variables, it can result in non-N copula structure and asymmetric 
+tail dependence across space in the ecological variables. Shown are the last
+$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text, $b=0.1$.
+
+\hspace{1in}
+
+\noindent \textbf{Figure 5.} Example results showing similarity of copula statistics for 
+environmental-noise inputs and ecological-variable outputs of a dynamical model.  
+Asymmetric tail dependence was similar in both noise inputs and model outputs
+for this model. The 
+AR(1) model and a Clayton copula were used (see methods section for Q2).
+Each point is the mean across 50 replicate simulations for which the same 
+Kendall $\tau$ value was used. Error bars are standard errors 
+and panel headers give Pearson correlation results for
+the points. The regression line through the points (black line) was similar
+to the 1-1 line (green line). 
+See Fig. \ref{SM-fig_Cause4copula_C_kend_ar1} for additional results for the AR(1) model and Clayton copula, 
+Figs \ref{SM-fig_Cause4copula_SC_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1} for AR(1) results with other copula families, and 
+Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1} for
+the nonlinear models described in the methods section for Q2.
+
+\hspace{1in}
+
+\noindent \textbf{Figure 6.} Skewness of spatially averaged green spruce aphid abundance $(A)$, leaf-curling plum 
+aphid first flight dates $(B)$, \emph{Ceratium furca} abundance $(C)$, and methane-flux
+$(D)$ time series compared to an multivariate 
+N-copula null hypothesis. Black dots are empirical 
+skewnesses; see text 
+and Appendix \ref{SM-Surrog_multi} for details of the null hypothesis. Results show a 
+tendency for skewness of the spatial average to be affected as hypothesized 
+by tail dependence.
+
+\hspace{1in}
+
+\noindent \textbf{Figure 7.} Extinction risk for the metapopulation extension of the Lewontin-Cohen model, after 
+$25$ time steps, was higher for left-tail-dependent environmental noise. Dispersal was local,
+with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
+for $N=5$ (A) and $N=25$ (B).

### 2019-06-13 SupportingInformation.Rmd
done with round 1 of prepping for Ecol Mongr, going through it one more time now

@@
 ---
-title: "Supporting information: Copulas reveal complex and informative dependencies propagating throughout ecology"
+title: "Supporting information: Copulas and their potential for ecology"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
@@
 upper-tail dependence can vary independently of correlation. (A-C) Data generated using the BB1 
-copula (see section on methods for the observational hypothesis) 
+copula (see section on methods for Q1) 
 with three different 
@@
 
-# Verifying model selection results of testing the observational hypothesis on bivariate datasets using nonparametric statistics \label{Q1bivarnonparam}
+# Verifying Q1 results for bivariate datasets using nonparametric statistics \label{Q1bivarnonparam}
 
@@
 Kendall (or Spearman)
-correlation. Thus the comparison addresses the first two parts of the observational hypothesis
+correlation. Thus the comparison addresses the first two parts of Q1
 for bivariate datasets. 
@@
 provides a test of asymmetry of dependence. Thus this comparison helps address the
-third part of the observational hypothesis for bivariate 
+third part of Q1 for bivariate 
 datasets.
@@
 considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
-survival Joe, survival BB1, survival BB6, survival BB7, and survival BB8 copulas, with abbreviations in the methods for the observational hypothesis.
+survival Joe, survival BB1, survival BB6, survival BB7, and survival BB8 copulas, with abbreviations in the methods for Q1.
 NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
@@
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for leaf-curling plum aphid first flight data.
-Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations in the methods for the observational hypothesis.
+Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations in the methods for Q1.
 NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.
@@
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for \\textit{Ceratium furca} abundance data.
-See methods for the observational hypothesis for copula abbreviations.
+See methods for Q1 for copula abbreviations.
 Indep.=Independent.
@@
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.
-See methods for the observational hypothesis for copula abbreviations.
+See methods for Q1 for copula abbreviations.
 Indep.=Independent.
@@
 
-# Verifying model selection results of testing the observational hypothesis on multivariate datasets using nonparametric statistics \label{Q1multvarnonparam}
+# Verifying Q1 results for multivariate datasets using nonparametric statistics \label{Q1multvarnonparam}
 
@@
       caption="Asymmetric sensitivity model, model selection results. 
-The output of simulations from the second model detailed in 
-the methods section for the causal hypotheses (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
+The output of simulations from the first model detailed in 
+the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
 \\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to the model 
 selection procedures detailed in the 
-methods section for the observational hypothesis. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
+methods section for Q1. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
 N-copula AICw: the AIC weight of the N copula family; Model-avg. LT: the lower-tail 
@@
              caption ="Asymmetric sensitivity model, nonparametric statistics results. 
-The output of simulations from the second model detailed in 
-the methods section for the causal hypotheses (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
+The output of simulations from the first model detailed in 
+the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
 \\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
-analyses decribed in the methods section for the observational hypothesis. The listed statistic was computed for the
+analyses decribed in the methods section for Q1. The listed statistic was computed for the
 output of the model with the indicated b and g, and the value was compared to 1000 values
 of the same statistic computed on surrogate time series randomized to have N-copula structure
-(see methods for the observational hypothesis). A table entry $<X$ (respectively, $>X$) indicates the value of the 
+(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
 given statistic on the data was less (respectively, more) than its value on $X$ of the 
@@
 copula structure as measured by the listed statistic differs significant
-between noise inputs and model outputs. See methods section for the causal hypotheses 
+between noise inputs and model outputs. See methods section for Q2 
 for additional details.\label{fig_Cause4copula_C_kend_ar1}}
@@
 copulas that paralleled the analyses applied to bivariate ecological datasets 
-(see section on methods for the observational hypotheses) are in Table \ref{tab:greenspruce}.
+(see section on methods for Q1) are in Table \ref{tab:greenspruce}.
 
@@
 result for the goodness of fit test using the Cramer-von Mises statistic (see section on 
-methods for the observational hypothesis). The 
+methods for Q1). The 
 column labeled p, KS is the p-value 
@@
 for winter temperature versus green spruce 
-aphid abundance for each of the 10 sampling locations. The $p$-values in the header are from the test of independence (see section on methods for the observational hypothesis).\label{fig:greenspruce}}
+aphid abundance for each of the 10 sampling locations. The $p$-values in the header are from the test of independence (see section on methods for Q1).\label{fig:greenspruce}}
 \end{center}

### 2019-06-14 Paper.Rmd
further edits

@@
 (Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see section on
-methods for Q1).
+methods for Q1 below).
 
@@
 minimum. If, for instance, the growth of a plant depends on soil 
-nitrogen, N, and other factors, a plot of plant growth rates versus N may 
+nitrogen, N, and other factors, a plot of plant growth rates versus soil N may 
 look more like Fig. \ref{fig_cop_pedag1}A or C than like 
@@
 relationship, only when it is limiting, in the left portions of the distributions. 
-The portion of the distributions that dominate the association is visible in the copula
-approach, but not via standard techniques.
+The portions of the distributions that dominate the association is visible in the copula
+approach, but may not be revealed by standard techniques.
 
@@
 variables [@Serinaldi2008; @She2018; @Goswami2018;
-@li2013]. One may expect copula structure in environmental variables 
-to take the form of strong tail dependence if
+@li2013]. One may expect an environmental variable 
+measured through time in two locations to show strong tail dependence between
+the locations if
 intense meteorological events are also widespread in their effects, as seems 
-frequently to be the case.
+frequently to be the case: extreme values are associated with intense events,
+so happen in both places at the same time, whereas moderate values of the 
+environmental variable are associated with local phenomena, which differ
+between the locations.
 Spatial dependence in an environmental variable 
@@
 (this is called \emph{spatial synchrony}) which are influenced by the environmental 
-variable. This is called the \emph{Moran effect}. 
+variable. This influence is called the \emph{Moran effect}. 
 If copula structure is transmitted, 
@@
 from environmental to ecological variables, then we may expect important 
-copula structure to be a common feature of the spatial synchrony of population, community, 
-biogeochemical and other ecological variables.
-Synchrony is a phenomenon of great interest in 
+copula structure to be a common feature of the 
+spatial synchrony of population, community, 
+biogeochemical and other environmentally influenced ecological variables.
+Synchrony is a phenomenon of substantial interest in 
 ecology [@Liebhold2004; @Sheppard2016; @walteretal2017].
@@
 Alternatively, Fig. \ref{fig_cop_pedag1}F suggests that Sp2
-is dominant. The copula approach again reveals which portions of the
-distributions are principally responsible for the association between variables, and this 
-information can help support or refute alternative causal hypotheses.
+is the dominant competitor. The copula approach again reveals which portions of the
+distributions are principally responsible for the association between 
+variables, and this 
+information can then help support or refute alternative causal hypotheses
+about which species is driving the dynamics of the other.
 
@@
 and the total economic value of a fishery is the sum of local catches. 
-We will explore how details, captured by copulas, of the dependence between local
-quantities can influence the spatial mean or sum, and how this may influence 
+We will explore how details, captured by copulas, of the dependence between 
+fluctuations of local
+quantities can influence fluctuations of 
+the spatial mean or sum, and how this may influence 
 higher organizational levels in ecology and human concerns.
@@
 for ecology by addressing the following specific questions. (Q1) Do datasets
-in ecology have copula structure distinct from that of a Gaussian/normal
+in ecology have copula structure distinct from that of a multivariate 
+Gaussian/normal
 distribution (here called non-normal copula structure)? Do they show tail 

### 2019-06-14 SupportingInformation.Rmd
further edits

@@
 upper-tail dependence can vary independently of correlation. (A-C) Data generated using the BB1 
-copula (see section on methods for Q1) 
+copula (see section on methods for Q1 in the main text) 
 with three different 
 sets of parameters chosen to produce a variety of 
-lower-tail (LT) and upper-tail (UT) dependence strengths, but the same 
+lower-tail (LT) and upper-tail (UT) dependence strengths, but with the same 
 Kendall correlation (K) (to two digits) in all cases. (D-F) Similar to A-C, but

### 2019-06-16 Paper.Rmd
further edits, increasingly minor

@@
 Q1 is "yes," what are some possible causes/mechanisms of non-normal
-copula structure? And (Q3) what are the consequences of non-normal copula
+copula structure and tail dependence? And (Q3) what are the consequences of non-normal copula
 structure and tail dependence for our understanding of ecology and for 
-applications? To address Q1 (the second through fourth sections of the paper), 
-we analyze several datasets, including 
+applications? We address Q1, in the second through fourth sections of the paper, 
+by analyzing several datasets, including 
 environmental, species-trait, phenological, population, community, and 
@@
 sub-fields and organizational levels of ecology. 
-To address Q2 (the fifth and sixth sections of the paper) we use 
-simple models. To address Q3 (the seventh and eigths sections) we use 
+We address Q2, in the fifth and sixth sections of the paper, using 
+simple models. We address Q3, in the seventh and eigths sections of the paper, using 
 both data and models. The ninth section is the Discussion.
@@
 of organization within ecology. Copula structure describing 
-the relationship between atmospheric weather variables such as rainfall or wind speed 
-as measured in multiple locations has been 
+the relationship between atmospheric weather variables such as rainfall or wind speed, 
+as measured in multiple locations through time, has been 
 examined previously in the meteorological literature [e.g., @Serinaldi2008; @li2013]. We therefore 
@@
                                                paste(nrow(d_mammalsBMR)," mammals",sep=""),
-                                               "10 locations with at least 30-yr. timeseries for each 20 sp.",
-                                               "10 locations with at least 30-yr. timeseries for each 20 sp.",
-                                               "14 locations with at least 45-yr. timeseries for each 22 taxa",
+                                               "10 locations with at least 30-yr. timeseries for each of 20 sp.",
+                                               "10 locations with at least 30-yr. timeseries for each of 20 sp.",
+                                               "14 locations with at least 45-yr. timeseries for each of 22 taxa",
                                                paste(nrow(d_cc[[5]])," plots",sep=""),
@@
 nonparametric approaches to provide greater 
-confidence in our answer to Q1. We used three 
+confidence in our answers to Q1. We used three 
 statistics which quantify the extent to which 
@@
 Bottom panels show the corresponding copula plots.
-See Data for units used in upper panels.
+See Data section for units used in upper panels.
 \label{fig_biv_multi_raw_cop}}
@@
 <!--bivariate example-->
-For the soil C and N data (Data, Table \ref{tab_data_info}, 
-Fig. \ref{fig_biv_multi_raw_cop}A,E), variables were 
+For the soil C and N data (Data section, Table \ref{tab_data_info}, 
+Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
 non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
@@
 The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
-to the data and computed AIC and BIC weights and corresponding
+to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
+and computed AIC and BIC weights and corresponding
 lower- and upper-tail statistics for the fitted distributions (Table \ref{SM-tab_soilCNfit}).
@@
 These values are distinct from what an N copula would give, namely, 0, answering the second part 
-Q2 for these data. The values also differed substantially 
+Q1 for these data. The values also differed substantially 
 from each other, answering the third part of Q1. 
@@
 was asymmetric (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}, 
-Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}). The 
+Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}). Our 
 affirmative, empirically based answer to Q1 for the soil C and N data is 
@@
 (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}).
-The 
+Our 
 affirmative, empirically based answer to Q1 for the bivariate datasets is 
 represented in Fig. \ref{MasterFigure} as solid outlines around the names of those
-datasets in the middle row of boxes.
+datasets, in the middle row of boxes.
 We did the same analysis 
@@
 an objectively adequate description of 
-the data; the collection of 
+the data; i.e., the collection of 
 copula families we used was sufficiently broad to characterize these data.
@@
 from what an N copula would give (i.e., 0), and answering the second
-part of Q1. These data have greater tail dependence 
+part of Q1: these data have greater tail dependence 
 (lower and upper) than would be expected from an N-copula null hypothesis. 
@@
 Figs \ref{SM-fig_aphid_count}-\ref{SM-fig_methane}.
-The 
+Our 
 affirmative, empirically based answer to Q1 for the multivariate datasets is 
 represented in Fig. \ref{MasterFigure} as solid outlines around the names
-of those datasets in the middle row of boxes.
+of those datasets, in the middle row of boxes.
 We also carried out the same analyses for abundance and first-flight data for the 18 
 aphid species for which we had data other than the green spruce and leaf-curling plum aphids,
-as well as for the 21 plankton taxa for which we had abundance data other than the
-\textit{Ceratium furca} (results not shown); conclusions supported that 
-non-N copula structure and tail dependence and asymmetry of tail dependence are 
+as well as for the 21 plankton taxa for which we had abundance data other than
+\textit{Ceratium furca} (results not shown); results supported the conclusion that 
+non-N copula structure and tail dependence, and asymmetry of tail dependence are 
 common.
@@
 $g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
-$g_1$ represents environmental effects that occur only for negative 
-values and $g_2$ only for positive values of $\epsilon$.
+$g_1$ represents environmental effects that occur only below a threshold of $\epsilon=0$, 
+and $g_2$ represents effects that occur only above a threshold of $\epsilon=0$.
 Negative values of $b$ correspond to overcompensating dynamics 
@@
 steps. We applied our model selection algorithms and 
-non-parametric statistics (methods for Q1) to these 
+non-parametric statistics (see section on methods for Q1) to these 
 outputs to discover if the model could
@@
 $E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
-The environmental noises $\epsilon_i(t)$ are 
-standard-normal random variables that are independent for distinct times, $t$, but 
+The environmental noises $\epsilon_i(t)$ were 
+standard-normal random variables that were independent for distinct times, $t$, but 
 exhibited different kinds of 
@@
 To explore the importance of nonlinear dynamics for our proposed mechanism, 
-we also considered 
-nonlinear population models based on $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
+we also considered a
+nonlinear population model $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
 $\sigma=1$. When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
@@
 
-Our next mechanism is evolutionary, and pertains to bivariate 
+The next mechanism we investigated is evolutionary, and pertains to bivariate 
 trait data across species, e.g., our bird and mammal 
@@
 The hypothesis is that tail dependence occurs in evolutionary changes
-in bivariate characters, and gives rise to tail dependence in the 
-character values observed across extant species.
+in bivariate characters, and gives rise to tail dependence between the 
+two character values across extant species.
 We simulated bivariate character evolution on an estimate of 
@@
 across each branch were randomly chosen from matrices of one million 
-bivariate draws from distributions showing one of five distinct types of copula 
+independent draws from bivariate distributions showing one of five distinct types of copula 
 structure: 1) extreme or 2) moderate left-tail dependence,
@@
 Our model with asymmetric environmental effects 
-could produce outputs with visually apparent asymmetry 
+produces outputs with visually apparent asymmetry 
 of tail dependence, to an extent that depended on the value of $b$;
-thus this is a mechanism that may produce non-N copula structure and
+thus this is a mechanism that can produce non-N copula structure and
 asymmetric tail dependence across space in ecological variables.
@@
 dynamics were strongly nonlinear.
-Thus this is a mechanism that may produce non-N copula structure and
+Thus this is a mechanism that can produce non-N copula structure and
 asymmetric tail dependence across space in ecological variables.
@@
 were again very similar for noise and model outputs 
-(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-
-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}).
+(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}).
 Though there were again significant but small differences for some statistics
@@
 averaged quantity should be influenced by 
-dependencies between the local quantities, including their
+dependencies between the local quantities being averaged, including their
 copula structure and tail dependence. 
@@
 should tend to decrease skewness. Strong positive skewness of the 
-spatial mean corresponds to "spikiness" of the spatial-mean time series. 
+spatial-mean time series corresponds to "spikiness" of that time series,
+which corresponds to instability through time. 
 The spatial-mean time series and its skewness may be quantities of 
@@
 To help answer Q3, we tested the above hypothesis
-using our multivariate datasets (Table \ref{tab_multivar_summary}).
+using our multivariate datasets (Table \ref{tab_data_info}).
 For each dataset we calculated the spatial mean time series, and then 
@@
 contributed to the skewness of the spatial mean time series, 
-i.e., to its "spikiness".
+i.e., to its instability and "spikiness" through time.
 
@@
 \noindent Results were consistent with the hypothesis that the skewness
-of a spatial-average time series is influenced by the tail dependence
-between the local quantities that were averaged. Thus copula 
+of a spatial-average time series is influenced by tail dependence
+between the local quantities being averaged. Thus copula 
 structure and tail dependence are important for spatially averaged 
@@
 as the solid box around "Instability/skewness of mean or total 
-time series" and the solid arrows labelled "X". For datasets that
+time series" and the solid arrows labeled "X". For datasets that
 exhibited strong lower-tail (respectively, upper-tail) 
@@
 aphid first flight dates $(B)$, \emph{Ceratium furca} abundance $(C)$, and methane-flux
-$(D)$ time series compared to an multivariate 
+$(D)$ time series compared to a multivariate 
 N-copula null hypothesis. Black dots are empirical 
@@
 We also substantiated using models (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd}, Fig. \ref{SM-fig_spTL})
-a hypothesis (Fig. \ref{MasterFigure}, arrow labelled "Z"), that 
+a hypothesis (Fig. \ref{MasterFigure}, arrow labelled "Z") that 
 Taylor's law can be influenced by
@@
 exactly what was observed (Table \ref{tab_multivar_summary}), providing 
-empirical evidence supporting the mechanism. The consequences of such tail 
+empirical evidence supporting the mechanism (this is why the box around "Nonlinear 
+environmental effects, Liebig's law" and some of the arrows labelled "A" are solid
+instead of dashed in Fig. \ref{MasterFigure}). The consequences of such tail 
 dependence for the skewness of spatially averaged aphid counts was 
@@
 temperature ultimately causes spatially averaged aphid abundance time series 
-to have lower skewness (i.e., less spikiness, greater stability through time) 
+to have lower skewness (i.e., less spikiness, and greater stability through time) 
 than they would otherwise.
@@
 Bottom panels show the corresponding copula plots.
-See Data for units used in upper panels.
+See Data section for units used in upper panels.
 

### 2019-06-16 SupportingInformation.Rmd
further edits, increasingly minor

@@
 For extreme right-tail dependent noise, we first produced
-a 1,000,000 by $2$ matrix, $m$, of 
+a 1,000,000 by $2$ matrix, $m$, filled with 2,000,000 
 independent standard-normal random draws. Then, each row of $m$ 
@@
 own absolute value. This produces right-tail dependent noise, normally distributed
-for each location. If left-tail dependent noise was desired, we took the negative of the
+for each of the two locations. 
+If left-tail dependent noise was desired, we took the negative of the
 entire matrix. The moderately left-tail dependent matrix was based on a C copula 

### 2019-06-25 Paper.Rmd
submitted this version to Ecological Monographs yesterday
changed an erroneous grant number

@@
 
+\raggedright
+\setlength\parindent{2em}
+\setlength{\parskip}{6pt}
+
 \noindent \emph{Affiliations:}
@@
 
-\noindent \emph{The number of words in the abstract:}
+\noindent \emph{The number of words in the abstract:} 266
 
@@
 under the Core Capability Grant BBS/E/C/000J0200. SG, LWS and DCR were partly funded by US National 
-Science Foundation grants 0075516 and 1442595 and the James S McDonnell Foundation.
+Science Foundation grants 1714195 and 1442595 and the James S McDonnell Foundation.
 
@@
 
-<div id="refs"></div>
-
-<!--\setlength{\parindent}{-0.2in}
+\setlength{\parindent}{-0.2in}
 \setlength{\leftskip}{0.2in}
@@
 \noindent
-\sloppy-->
+\sloppy
+
+<div id="refs"></div>
 
@@
 # Figure captions
+
+\setlength{\parindent}{0in}
+\setlength{\leftskip}{0in}
+\setlength{\parskip}{6pt}
 

### 2019-08-23 Paper.Rmd
starting to address referee comments from Ecol Mongr, only inserted comments into Paper.Rmd so far, but about to start actually editing

@@
 an asymmetric pattern of dependence that can
-have profound significance, as discussed below, but is not captured by 
+have significance, as discussed below, but is not captured by 
 correlation. 
@@
 
+<!--***DANr: reword this para so people don't think the ranks *are* the 
+copulas, and to try to pre-defend against some of what ref 1 said. Also,
+state that these are conceptual introductions only and are not formal,
+a more comprehensive introduction is in xyz-->
 Well-developed approaches do exist, however, and are 
@@
 dependence structure.
+<!--***DANr: Maybe move the below two sentences into the new section on Background-->
 See @nelsen2006_copula and @joe2014_dependence
@@
 A main benefit of a copula approach is 
-that it can sensitively detect tail dependence and asymmetries of tail dependence. 
+that it can detect tail dependence and asymmetries of tail dependence. 
 Tail dependence is association between extreme values of variables. 
@@
 applications of copulas in ecology. 
+<!--***DANr: Maybe emphasize here that we provide some, but not a comprehensive,
+intro to copulas - for that go elsewhere. Instead we are trying to determine
+to what extent they may be useful in ecology.-->
 Our results provide support for the hypothesis that environmental, ecological and evolutionary processes 
-generate and transmit nontrivial copula structure and tail dependence, with substantial 
+generate and transmit nontrivial 
+<!--***DANr: nontrivial? what does it mean-->
+copula structure and tail dependence, with substantial 
 consequences for theoretical and conceptual understanding, for inference, and for 
@@
 
+<!--***DANr: Maybe use the words tail dependence more and the word copula
+less in the below? Generally the words "copula structure" may be rubbing some
+people wrong. Replacement words? What do you actually mean here? Maybe call it
+asymmetric dependence, or complex dependence structure?-->
 One way we may expect, *a priori*, a copula approach 
@@
 
+<!--***DANr: Need some caveating here to accommodate the people who slavishly
+follow the old maxim that correlation is not causation.-->
 A third reason we may expect a copula approach to benefit 
@@
 
+# Background on copulas
+
+<!--***DANr: Add a new section above this one called Background on copulas.
+It can have the following kinds of stuff (probably not exhaustive, and probably
+not in order, and some of this material is currently below):
+The definition
+The explanation of how ranks relate to estimating the copula parameters
+How to estimate carry out fitting (basically, MLE fitting on ranks)
+What some of the different copula families look like
+What tail dependence actually is
+-->
+
 # Concepts and methods for Q1 \label{Methods}

### 2019-08-23 Paper.Rmd
edited the Intro to address referee concerns

@@
 # Abstract
-
+<!--***DANr: I have not edited this abstract at all since Ecol Monogr-->
 All branches of ecology study relationships among environmental and biological variables. However, 
@@
 an asymmetric pattern of dependence that can
-have significance, as discussed below, but is not captured by 
-correlation. 
-Monotonic transformations can alter the Pearson, but not the Spearman
-correlation (Appendix \ref{SM-pedagog1_details}, 
-Fig. \ref{SM-fig_transf_pearson_spearman}). Although correlation
-is only one way to 
+have ecological significance, as discussed below, but is not captured by 
+correlation. Although correlation is only one way to 
 study relationships between variables, other common metrics also 
-provide partial information.
+provide only partial information.
 
@@
 
-<!--***DANr: reword this para so people don't think the ranks *are* the 
+<!--DONE ***DANr: reword this para so people don't think the ranks *are* the 
 copulas, and to try to pre-defend against some of what ref 1 said. Also,
@@
 The approaches are based on 
-the concept of the *copula*, which we here 
-introduce. Copula approaches separate the information of a bivariate 
+the concept of the *copula*; copula approaches separate the information of a bivariate 
 random variable into two distinct parts: the information in the 
@@
 variables), and the remaining information (which is solely about
-dependence). From a data perspective, given 
+dependence). We here introduce some simple concepts, based on taking ranks,
+that relate to copulas and give a conceptual 
+flavor of copula ideas to be introduced 
+more formally in the section on Background below. Given 
 a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
-Fig. \ref{fig_cop_pedag1}A-B), the dependence
-structure is revealed by the bivariate plot of $u_i$ versus $v_i$, where
+Fig. \ref{fig_cop_pedag1}A-B), information about the structure of the dependence
+between $x$ and $y$ can be separated from information contained 
+in the marginal distributions by 
+considering the bivariate plot of $u_i$ versus $v_i$, where
 $u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
@@
 of the smallest element in a set is $1$. The $u_i$ and 
-$v_i$ are called *normalized ranks* of the $x_i$ and $y_i$. See, 
-for instance,
-Fig. \ref{fig_cop_pedag1}C-D (the points), which show the copula structures of 
+$v_i$ are called *normalized ranks* of the $x_i$ and $y_i$; they relate to 
+the empirical cummulative distribution function (cdf) of the $x_i$ and $y_i$,
+respectively. See, for instance,
+Fig. \ref{fig_cop_pedag1}C-D (the points), which show the normalized ranks of 
 A-B, respectively. Ranking makes the marginal distributions of 
-the underlying random variables uniform, isolating the 
-dependence structure.
-<!--***DANr: Maybe move the below two sentences into the new section on Background-->
-See @nelsen2006_copula and @joe2014_dependence
-for general background. A theorem of @sklar1959_theorem
-states that copulas give a unique and complete 
-representation of the dependence
-between variables.
+the component datasets uniform, isolating the 
+dependence structure. Dependence structure and marginals can then be studied 
+separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D}
+relate to copulas in a way to be described in the section on Background below, 
+after a definition of copulas
+is provided. Note that Pearson correlation, even though it is the most commonly
+used measure of relationships between variables, is modified by monotonic
+tranformation of the component variables (Appendix 
+\ref{SM-pedagog1_details}, Fig. \ref{SM-fig_transf_pearson_spearman}),
+and therefore reflects not only dependence information, but also
+information on the marginals. Spearman and Kendall correlations, being rank-based, 
+are not influenced by monotonic transformations of the component variables; 
+they provide information solely on the dependence between those variables.
 
@@
 that it can detect tail dependence and asymmetries of tail dependence. 
-Tail dependence is association between extreme values of variables. 
+Tail dependence (defined precisely in the section on Background, below) 
+is association between extreme values of variables. 
 If smaller values of two positively associated variables are more 
@@
 have stronger lower- or left-tail dependence than right- or upper-tail dependence
-(Fig. \ref{fig_cop_pedag1}A, C);
+(Fig. \ref{fig_cop_pedag1}C);
 and *vice versa* if larger values are more strongly associated than smaller
-ones (Fig. \ref{fig_cop_pedag1}B, D). Datasets of the same 
-correlation can have a range of tail dependence 
+ones (Fig. \ref{fig_cop_pedag1}D). Datasets of the same 
+Spearman or Kendall correlation can have a range of tail dependence 
 (Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see section on
-methods for Q1 below).
+methods for Q1 below). Although copulas can be used to model the entire
+complex dependence structure of data, we have found tail dependence to be
+a useful aspect of the full structure of dependence, so tail dependence 
+is one of the main foci of this paper.
 
 The goal of this paper is to explore the potential for 
-applications of copulas in ecology. 
-<!--***DANr: Maybe emphasize here that we provide some, but not a comprehensive,
-intro to copulas - for that go elsewhere. Instead we are trying to determine
-to what extent they may be useful in ecology.-->
-Our results provide support for the hypothesis that environmental, ecological and evolutionary processes 
-generate and transmit nontrivial 
-<!--***DANr: nontrivial? what does it mean-->
-copula structure and tail dependence, with substantial 
-consequences for theoretical and conceptual understanding, for inference, and for 
-applications across ecology. The work we present here has begun to test this
-hypothesis, and we hope our results stimulate further
-evaluation of the hypothesis by the research community.
+applications of copulas in ecology. We provide a sketch of an introduction
+to copula concepts in the Background section below, but this is a sketch
+only - readers can find comprehensive introductions elsewhere
+[@nelsen2006_copula; @joe2014_dependence]. Instead our goal is to 
+help determine to what extent ecological understanding may benefit from
+using copulas to help understand complex dependence 
+structures between variables.
+Our results provide support for the hypothesis that environmental, 
+ecological and evolutionary processes 
+generate and transmit complex dependence structures, including tail dependence, 
+with consequences for theory and 
+applications. We hope our results will stimulate further
+evaluation of this hypothesis by other researchers.
 
@@
 
-<!--***DANr: Maybe use the words tail dependence more and the word copula
-less in the below? Generally the words "copula structure" may be rubbing some
+<!--DONE ***DANr: Maybe use the words tail dependence more and the word copula
+less in the below few paras? Generally the words "copula structure" may be rubbing some
 people wrong. Replacement words? What do you actually mean here? Maybe call it
 asymmetric dependence, or complex dependence structure?-->
-One way we may expect, *a priori*, a copula approach 
+One way we may expect, *a priori*, a copula 
+study of complex dependence structures between
+ecological variables
 to better reflect ecological relationships relates to Liebig's law of the
-minimum. If, for instance, the growth of a plant depends on soil 
+minimum (the idea that growth is controlled not by total resources but by the scarcest 
+resources). If, for instance, the growth of a plant depends on soil 
 nitrogen, N, and other factors, a plot of plant growth rates versus soil N may 
@@
 relationship, only when it is limiting, in the left portions of the distributions. 
-The portions of the distributions that dominate the association is visible in the copula
-approach, but may not be revealed by standard techniques.
-
-A second reason we may expect a copula approach to benefit ecological
-research is that prior work has demonstrated important copula 
+The portions of the distributions that dominate the association is visible in rank-based
+plot, but may not be revealed by standard correlation techniques.
+
+A second reason we may expect understanding of complex dependence structures
+to benefit ecological
+research is that prior work has demonstrated (using copulas) complex 
 stucture in the spatial dependence of environmental 
@@
 so happen in both places at the same time, whereas moderate values of the 
-environmental variable are associated with local phenomena, which differ
+environmental variable may be associated with local phenomena, which differ
 between the locations.
@@
 variable. This influence is called the \emph{Moran effect}. 
-If copula structure is transmitted, 
+If complex dependence structure is transmitted, 
 possibly in modified form, 
-from environmental to ecological variables, then we may expect important 
-copula structure to be a common feature of the 
+from environmental to ecological variables, then we may expect complex dependence 
+structure to be a common feature of the 
 spatial synchrony of population, community, 
@@
 
-<!--***DANr: Need some caveating here to accommodate the people who slavishly
+<!--DONE ***DANr: Need some caveating here to accommodate the people who slavishly
 follow the old maxim that correlation is not causation.-->
-A third reason we may expect a copula approach to benefit 
+A third reason we may expect an understanding of complex 
+dependence structures to benefit 
 ecological research is that such an approach may help illuminate 
@@
 and Sp2 on the vertical, then Fig. \ref{fig_cop_pedag1}E 
-would suggest Sp1 is the dominant 
+may suggest Sp1 is the dominant 
 competitor: when Sp1 is abundant, Sp2 is necessarily rare because it is suppressed; 
@@
 limiting factors other than Sp1. 
-Alternatively, Fig. \ref{fig_cop_pedag1}F suggests that Sp2
-is the dominant competitor. The copula approach again reveals which portions of the
-distributions are principally responsible for the association between 
-variables, and this 
-information can then help support or refute alternative causal hypotheses
-about which species is driving the dynamics of the other.
-
-A fourth reason we may expect a copula approach to be important
+Alternatively, Fig. \ref{fig_cop_pedag1}F may suggest that Sp2
+is the dominant competitor. Although other causal hypotheses may produce 
+similar dependence structure, and it is usually impossible to obtain
+complete information on causal pathways from analyses which are
+fundamentally based on association data, more detailed 
+examination of the patterns of dependence of variables may
+well be combined with biological and other reasoning to rule out
+some causal hypotheses which could not be eliminated solely 
+with standard correlation approaches.
+
+A fourth reason we may expect a copula understanding of complex 
+dependence structures to be useful
 for ecology has to do with spatially aggregated or averaged quantities. 
@@
 and the total economic value of a fishery is the sum of local catches. 
-We will explore how details, captured by copulas, of the dependence between 
-fluctuations of local
+We will explore how details, which can be captured using copulas, 
+of the dependence between fluctuations of local
 quantities can influence fluctuations of 
@@
 effect, with
-potentially substantial consequences. For instance, if populations of a 
+possible consequences. For instance, if populations of a 
 pest species in different locations are all positively associated and also strongly
@@
 for ecology by addressing the following specific questions. (Q1) Do datasets
-in ecology have copula structure distinct from that of a multivariate 
+in ecology have dependence structure distinct from that of a multivariate 
 Gaussian/normal
@@
 normal distributions or distributions obtained by transforming
-the marginals of a normal distribution. Such 
-approaches do not utilize the power of copulas, so Q1 asks whether ecological
+the marginals of a normal distribution. So Q1 asks whether ecological
 data contain dependency information distinct from what can be modelled with
@@
 Q1 is "yes," what are some possible causes/mechanisms of non-normal
-copula structure and tail dependence? And (Q3) what are the consequences of non-normal copula
-structure and tail dependence for our understanding of ecology and for 
-applications? We address Q1, in the second through fourth sections of the paper, 
+copula structure and tail dependence? And (Q3) what are the 
+consequences of non-normal copula
+structure and tail dependence for ecological understanding and
+applications? After a Background section (the second section), we address Q1 
+(third through fifth sections of the paper) 
 by analyzing several datasets, including 
@@
 sub-fields and organizational levels of ecology. 
-We address Q2, in the fifth and sixth sections of the paper, using 
-simple models. We address Q3, in the seventh and eigths sections of the paper, using 
-both data and models. The ninth section is the Discussion.
+We address Q2 (sixth and seventh sections of the paper), using 
+simple models. We address Q3 (eighth and ninth sections of the paper) using 
+both data and models. The final section is the Discussion.
 Multiple analyses are brought to bear on each question, and these
@@
 bioinformatics [@Kim2008], medical research [@Emura2016], direct study of environmental 
-variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000]. 
-To our knowledge, non-normal copulas have not been used in ecology. 
-We argue that benefits to ecology of using a complete 
-copula description of dependence will be substantial.
-
-\newpage
+variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000],
+but have been used only rarely in ecology. 
+We argue that benefits of using a complete 
+copula description of dependence in ecology will be substantial.
+
+# Background on copulas
+
+<!--***DANr: Add a new section above this one called Background on copulas.
+It can have the following kinds of stuff (probably not exhaustive, and probably
+not in order, and some of this material is currently below):
+The definition
+The explanation of how ranks relate to estimating the copula parameters
+How to estimate carry out fitting (basically, MLE fitting on ranks)
+What some of the different copula families look like
+What tail dependence actually is
+-->
+
+<!--***DANr: Maybe move the below two sentences into the new section on Background-->
+See @nelsen2006_copula and @joe2014_dependence
+for general background. A theorem of @sklar1959_theorem
+states that copulas give a unique and complete 
+representation of the dependence
+between variables.
 
@@
 
-\newpage
-
-# Background on copulas
-
-<!--***DANr: Add a new section above this one called Background on copulas.
-It can have the following kinds of stuff (probably not exhaustive, and probably
-not in order, and some of this material is currently below):
-The definition
-The explanation of how ranks relate to estimating the copula parameters
-How to estimate carry out fitting (basically, MLE fitting on ranks)
-What some of the different copula families look like
-What tail dependence actually is
--->
-
 # Concepts and methods for Q1 \label{Methods}

### 2019-09-07 Paper.Rmd
some progress writing a Background section

@@
 between two or more variables that can be mathematically proven 
-to be complete [@joe2014_dependence; @nelsen2006_copula]. 
+to be complete [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. 
 The approaches are based on 
@@
 dependence structure. Dependence structure and marginals can then be studied 
-separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D}
+separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D
 relate to copulas in a way to be described in the section on Background below, 
@@
 The goal of this paper is to explore the potential for 
-applications of copulas in ecology. We provide a sketch of an introduction
+applications of copulas in ecology. We provide a basic introduction
 to copula concepts in the Background section below, but this is a sketch
-only - readers can find comprehensive introductions elsewhere
-[@nelsen2006_copula; @joe2014_dependence]. Instead our goal is to 
-help determine to what extent ecological understanding may benefit from
-using copulas to help understand complex dependence 
-structures between variables.
-Our results provide support for the hypothesis that environmental, 
+only - readers can find abbreviated [@Anderson2018] 
+and comprehensive introductions elsewhere
+[@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. Instead our goal is to 
+help determine to what extent ecological understanding may benefit generally from
+using copulas. Copulas have already been used very effectively 
+in a few studies to illuminate ecological phenomena 
+[@Valpine2014; @Anderson2018; @Popovic2018], but such usage is
+still rare. Our results suggest that environmental, 
 ecological and evolutionary processes 
-generate and transmit complex dependence structures, including tail dependence, 
-with consequences for theory and 
-applications. We hope our results will stimulate further
-evaluation of this hypothesis by other researchers.
+may commonly generate and transmit complex dependence structures, including tail 
+dependence, and that greater use of copulas can illuminate the underlying processes.
 
@@
 to better reflect ecological relationships relates to Liebig's law of the
-minimum (the idea that growth is controlled not by total resources but by the scarcest 
-resources). If, for instance, the growth of a plant depends on soil 
+minimum. Liebig's law is 
+the idea that growth is controlled not by total resources but by the resources which
+are scarcest relative to organism needs. 
+If, for instance, the growth of a plant depends on soil 
 nitrogen, N, and other factors, a plot of plant growth rates versus soil N may 
@@
 relationship, only when it is limiting, in the left portions of the distributions. 
-The portions of the distributions that dominate the association is visible in rank-based
+The portions of the distributions that dominate the association are visible in a rank-based
 plot, but may not be revealed by standard correlation techniques.
@@
 complete information on causal pathways from analyses which are
-fundamentally based on association data, more detailed 
+fundamentally based on associations, more detailed 
 examination of the patterns of dependence of variables may
@@
 some causal hypotheses which could not be eliminated solely 
-with standard correlation approaches.
+with standard correlation approaches. @Popovic2018 has previously
+used copulas to help illuminate causal aspects of relationships
+between species.
 
@@
 variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000],
-but have been used only rarely in ecology. 
-We argue that benefits of using a complete 
-copula description of dependence in ecology will be substantial.
+but have been used only rarely so far in ecology (but see @Valpine2014, 
+@Anderson2018 and @Popovic2018). 
+We argue that benefits wider usage of complete 
+copula descriptions of dependence in ecology will be substantial.
 
@@
 
-<!--***DANr: Add a new section above this one called Background on copulas.
+We give a brief introduction to copulas, focusing, for simplicity, on bivariate copulas
+and on concepts needed
+for the rest of the paper. See @nelsen2006_copula, @joe2014_dependence, and
+@MaiScherer2017 for general background, and see also @Anderson2018, part of 
+which is a short introduction to 
+copulas for ecologists. A bivariate *copula* can be defined as a bivariate distribution
+function with both margins uniform on $(0,1)$ (@joe2014_dependence, p. 7). 
+This will be the
+distribution function of a bivariate random variable with uniform marginals, and 
+the terminology *copula* is sometimes also applied to the random variable. 
+A foundational theorem of
+@sklar1959_theorem (see also @MaiScherer2017, p. 16) 
+says that if $F$ is the bivariate distribution function
+of a random vector $(X,Y)$, with margins $F_X$ and $F_Y$, then there 
+exists a copula $C$
+such that for all $(x,y)$ in the Euclidean plane, $F(x,y)=C(F_X(x),F_Y(y))$.
+The theorem also says $C$ is unique if $F_X$ and $F_Y$ are continuous.
+Thus $C$ couples the bivariate distribution, $F$, with the two marginal distributions. 
+Finally, Sklar's theorem says that if $D$ is any bivariate copula and 
+$G_A$ and $G_B$ are univariate distribution functions, then 
+$D(G_A(x),G_B(y))$ is a distribution function.
+The applications of this study fall within the case in which univariate marginal 
+distributions have continuous, strictly 
+monotonic distribution functions, and this case is simpler, so we 
+henceforth make such an assumption. 
+Then, "the" copula (now guaranteed unique) 
+associated with $(X,Y)$ is the distribution
+function of the random variable $(U,V)=(F_X(X),F_Y(Y))$ 
+(Appendix \ref{ContStrictMonoton}). And 
+$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
+$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ is the random
+vector with distribution function $D$ and
+$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
+(Appendix \ref{ContStrictMonoton}).
+
+Thus Sklar's theorem implies that
+any random vector can be constructed from a copula and marginal 
+distributions. And conversely,
+any copula and any univariate distribution functions give rise to a random vector. 
+Marginals contain no information about the
+dependence structure of a random vector, so the copula contains
+all such information - it is a complete description of the dependence
+between variables. Sklar's theorem makes it possible to construct
+bivariate distributions in two separate steps: by specifying the marginal 
+distributions and by specifying the dependence structure. The theorem
+also make it possible to study and model dependence independently 
+from marginal distributions. For instance, in a pedagogical 
+example, [@Anderson2018] modelled the dependence between abundance
+data for two fish species, red moki and black angelfish, by first 
+modelling the marginal distributions and then separately modelling 
+the dependence using Gaussian copulas (their section 2).
+See the Discussion for a few words on multivariate copulas, which can
+be developed in the same way as above [@nelsen2006_copula; @joe2014_dependence;
+@MaiScherer2017; @Anderson2018].
+<!--***DAN: dont forget to put these few words in!-->
+
+<!--***DAN: I still feel like we have to make this more intuitive for people
+at this point. Maybe a figure with two panels showing two copulas on
+the left, and two panels shwoing two sets of marginals at the top,
+and then four panels showing what you get when each of these combine?
+Good idea. Steps: split the current Fig 1 so the new Fig 1 has the first four panels,
+but C-D should only have the points. Then make a Fig 2 as above, and edit the
+text in the above para to use that figure. 
+The rest of the panels of the current Fig 1 will make a new Fig 3 later.-->
+
+<!--***DANr: Add a new section called Background on copulas.
 It can have the following kinds of stuff (probably not exhaustive, and probably
 not in order, and some of this material is currently below):
-The definition
-The explanation of how ranks relate to estimating the copula parameters
-How to estimate carry out fitting (basically, MLE fitting on ranks)
+1) DONE - The definition, Sklar's theorem
+2) What good are copulas - intuitive idea, basically, but also have to be linked to 
+the formalities of the definition
+
 What some of the different copula families look like
+The explanation of how ranks relate to copulas and estimating the copula parameters
+How to carry out fitting (basically, MLE fitting on ranks - Genest et al paper)
 What tail dependence actually is
@@
 
-<!--***DANr: Maybe move the below two sentences into the new section on Background-->
-See @nelsen2006_copula and @joe2014_dependence
-for general background. A theorem of @sklar1959_theorem
-states that copulas give a unique and complete 
-representation of the dependence
-between variables.
+<!--***DAN: In the Discussion, you need to point out not only that we
+use only bivariate copulas (and there are multivariate ones), but also
+that we use only continuous marginals. Probably an adequate approximation
+for our data, but see the other papers I've seen recently which use count data,
+and discrete marginals. Approaches such as that will be necessary for some 
+datasets-->
+
+We introduce the copula families and 
+measures of tail dependence we used.
+The normal copula family (henceforth N) is a one-parameter
+family, with parameter corresponding to the correlation of the related 
+bivariate normal distribtion, and controlling the degree of dependence 
+(Fig. \ref{fig_cop_pedag1}G for an example).
+Left- and right-tail dependence are $0$ for the N copula, and this 
+serves as a benchmark for tail dependence in other copulas - positive values of our
+tail dependence measures correspond to dependence stronger than an N
+copula. The Clayton and 
+survival-Clayton families (Fig. \ref{fig_cop_pedag1}C-D) are two more one-parameter 
+families. The Clayton has positive lower-tail dependence ($0.87$ for
+the example of Fig. \ref{fig_cop_pedag1}C) and $0$ upper-tail dependence; the 
+survival-Clayton is a rotation of the Clayton by $180^\circ$ and has 
+positive upper-tail ($0.87$ in Fig. \ref{fig_cop_pedag1}D)
+and $0$ lower-tail dependence. For our model selection procedure (see below), we used 
+several families of copulas exhibiting various 
+lower- and upper-tail dependence characteristics 
+(Clayton, survival-Clayton, Gumbel, survival-Gumbel, Joe, survival-Joe, Frank,
+BB1, survival-BB1,
+BB6, survival-BB6, BB7, survival-BB7, BB8, survival-BB8, henceforth
+abbreviated C, SC, G, SG, J, SJ, F, BB1, SBB1, BB6, SBB6, BB7, SBB7, BB8, SBB8).
+Copulas are implemented in the R packages `copula` and `VineCopula`. 
+Visualizations of these families (Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8})
+can be compared with normalized rank plots to estimate visually 
+which families may effectively model a given bivariate dataset. 
+If LT and UT denote 
+lower- and upper-tail dependence measures, their difference measures asymmetry of 
+tail dependence.
+See @nelsen2006_copula, @joe2014_dependence for further background.
 
@@
 
-Conceptually, a bivariate copula is a random variable 
-on the unit square, $[0,1] \times [0,1]$, with uniform maginals. 
-If $(X,Y)$ is a continuous random
-variable and the cummulative distribution functions of $X$ and $Y$ are 
-$F_X(x)=P[X\leq x]$ and $F_Y(y)=P[Y\leq y]$, then the dependence structure
-of $(X,Y)$ is revealed by the random variable $(F_X(X),F_Y(Y))$,
-which takes values in $[0,1] \times [0,1]$ and has 
-uniform marginals.
-
-We introduce the copula families and 
-measures of tail dependence we used.
-The normal copula family (henceforth N) is a one-parameter
-family, with parameter corresponding to the correlation of the related 
-bivariate normal distribtion, and controlling the degree of dependence 
-(Fig. \ref{fig_cop_pedag1}G for an example).
-Left- and right-tail dependence are $0$ for the N copula, and this 
-serves as a benchmark for tail dependence in other copulas - positive values of our
-tail dependence measures correspond to dependence stronger than an N
-copula. The Clayton and 
-survival-Clayton families (Fig. \ref{fig_cop_pedag1}C-D) are two more one-parameter 
-families. The Clayton has positive lower-tail dependence ($0.87$ for
-the example of Fig. \ref{fig_cop_pedag1}C) and $0$ upper-tail dependence; the 
-survival-Clayton is a rotation of the Clayton by $180^\circ$ and has 
-positive upper-tail ($0.87$ in Fig. \ref{fig_cop_pedag1}D)
-and $0$ lower-tail dependence. For our model selection procedure (see below), we used 
-several families of copulas exhibiting various 
-lower- and upper-tail dependence characteristics 
-(Clayton, survival-Clayton, Gumbel, survival-Gumbel, Joe, survival-Joe, Frank,
-BB1, survival-BB1,
-BB6, survival-BB6, BB7, survival-BB7, BB8, survival-BB8, henceforth
-abbreviated C, SC, G, SG, J, SJ, F, BB1, SBB1, BB6, SBB6, BB7, SBB7, BB8, SBB8).
-Copulas are implemented in the R packages `copula` and `VineCopula`. 
-Visualizations of these families (Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8})
-can be compared with normalized rank plots to estimate visually 
-which families may effectively model a given bivariate dataset. 
-If LT and UT denote 
-lower- and upper-tail dependence measures, their difference measures asymmetry of 
-tail dependence.
-See @nelsen2006_copula, @joe2014_dependence for further background.
+<!--***DANr: I moved some text from here and have not smoothed the transtion,
+if it needs smoothing-->
 

### 2019-09-07 SupportingInformation.Rmd
some progress writing a Background section

@@
 \end{figure}
+
+# Proofs of statements related to Skalr's theorem in the case of continuous, strictly monotonic margins \label{ContStrictMonoton}
+
+Let $F$ be the bivariate distribution function
+of a random vector $(X,Y)$, with margins $F_X$ and $F_Y$ assumed to be 
+continuous and strictly monotonic. We show the copula associated with
+$(X,Y)$, which exists and is unique by Sklar's theorem, is the distrbution function
+of the random variable $(U,V)=(F_X(X),F_Y(Y))$. Let $C$ denote the distribution
+function of $(U,V)$, which has uniform margins. By the uniqueness
+statement of Sklar's theorem, it suffices to show that $F(x,y)=C(F_X(x),F_Y(y))$
+for all $(x,y)$ in the Euclidean plane. But 
+$C(F_X(x),F_Y(y)) = P[U \leq F_X(x), V \leq F_Y(y)]
+= P[F_X(X) \leq F_X(x), F_Y(Y) \leq F_Y(y)]
+= P[X \leq x, Y \leq y]
+= F(x,y),$
+where the third equality follows from the strict monotonicity assumption. In fact, 
+that equality should hold even with non-strict monotonicity, so a stronger 
+statement is possible (see below for more on this topic). The continuity assumption 
+for $F_X$ and $F_Y$ is necessary, however: without it the random variables 
+$U$ and $V$ are not uniformly distributed, so their joint distribution function 
+is not even a copula.
+
+Now let $D$ be any bivariate copula and let 
+$G_A$ and $G_B$ be univariate distribution functions, so that 
+$D(G_A(x),G_B(y))$ is guaranteed by Sklar's theorem to be a a distribution function.
+We here show that if $G_A$ and $G_B$ and continuous and strictly monotonic, then
+$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
+$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ is the random
+vector with uniform marginals associated with $D$ and
+$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$.
+The inverses exist because of the strict monotonicity assumption.
+To see this, note that the distribution function of 
+$(G_A^{-1}(U),G_B^{-1}(V))$ is $P[G_A^{-1}(U) \leq x, G_B^{-1}(V) \leq y]$,
+which equals $P[U \leq G_A(x), V \leq G_B(y)]$, by strict monotonicity.
+But this is $D(G_A(x),G_B(y))$ because $D$ is the distribution function of 
+$(U,V)$, which proves the statement. In fact, a stronger statement assuming
+neither continuity nor strict monotonicity is possible, if one uses generalized
+inverses for $G_A^{-1}$ and $G_B^{-1}$ [@MaiScherer2017]. We adopt the weaker
+statements for simplicity and because our applications are solely under the case
+of continuous, strictly monotonic margins. But see @Anderson2018 and 
+@Popovic2018 for applications of copulas outside this case.
 

### 2019-09-08 Paper.Rmd
adjusting the first pedagog fig

@@
 \includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
-\caption{(A-D) Bivariate datasets showing diverse relationships between 
-variables; many of these datasets nevertheless have the same Pearson (P)
-or Spearman (S) correlation coefficients. (C, D) Contours 
+\caption{Bivariate datasets showing diverse relationships between 
+variables; some of these datasets nevertheless have the same Pearson (P)
+or Spearman (S) correlation coefficients. Subscripts h and v 
+in the axis labels stand for "horizontal" and "vertical".\label{fig_cop_pedag1}}
+\end{center}
+\end{figure}
+<!--***DAN: When I cut panels out of this figure (saving them for use in a 
+later figure), I also cut the caption, and here is the cut stuff for later use:
+(C, D) Contours 
 of probability density functions (pdf) of Clayton (C) and survival-Clayton 
@@
 Subscripts h and v 
-stand for "horizontal" and "vertical". \label{fig_cop_pedag1}}
-\end{center}
-\end{figure}
+stand for "horizontal" and "vertical".
+-->
 

### 2019-09-08 FirstPedagogFig.R

@@ -43,7 +43,7 @@ gap<-0.1
 pan_wd<-1
 pan_ht<-pan_wd
 tot_wd<-ymarg_wd+2*gap+2*pan_wd+2*numsp
-tot_ht<-xmarg_ht+4*gap+4*pan_ht+4*numsp
+tot_ht<-xmarg_ht+2*gap+2*pan_ht+2*numsp
 pchval<-20
 cexvalpts<-0.25
 cexvaltxt<-0.75
@@ -53,8 +53,8 @@ pdf(file="./Results/PedagogFig1.pdf",width=tot_wd,height=tot_ht)
 #upper-left panel, normal marginals, left-tail dependence
 par(fig=c((ymarg_wd+numsp)/tot_wd,
           (ymarg_wd+numsp+pan_wd)/tot_wd,
-          (xmarg_ht+3*pan_ht+3*gap+4*numsp)/tot_ht,
-          (xmarg_ht+4*pan_ht+3*gap+4*numsp)/tot_ht),
+          (xmarg_ht+1*pan_ht+1*gap+2*numsp)/tot_ht,
+          (xmarg_ht+2*pan_ht+1*gap+2*numsp)/tot_ht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
 rg<-max(abs(dcc5_nm))
 plot(dcc5_nm[,1],dcc5_nm[,2],type='p',pch=pchval,cex=cexvalpts,xlim=c(-rg,rg),ylim=c(-rg,rg),col="grey")
@@ -68,8 +68,8 @@ text(rg,-rg,paste0("S=",round(cor(dcc5_nm[,1],dcc5_nm[,2],method="spearman"),2))
 #upper-right panel, normal marginals, right-tail dependence
 par(fig=c((ymarg_wd+2*numsp+pan_wd+gap)/tot_wd,
           (ymarg_wd+2*numsp+2*pan_wd+gap)/tot_wd,
-          (xmarg_ht+3*pan_ht+3*gap+4*numsp)/tot_ht,
-          (xmarg_ht+4*pan_ht+3*gap+4*numsp)/tot_ht),
+          (xmarg_ht+1*pan_ht+1*gap+2*numsp)/tot_ht,
+          (xmarg_ht+2*pan_ht+1*gap+2*numsp)/tot_ht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
 rg<-max(abs(dsc5_nm))
[…]

### 2019-09-08 Paper.Rmd
deleted the separate-page figure captions that had been required at previous submission

@@
 <div id="refs"></div>
-
-\clearpage
-\newpage
-
-# Figure captions
-
-\setlength{\parindent}{0in}
-\setlength{\leftskip}{0in}
-\setlength{\parskip}{6pt}
-
-\noindent \textbf{Figure 1.} (A-D) Bivariate datasets showing diverse relationships between 
-variables; many of these datasets nevertheless have the same Pearson (P)
-or Spearman (S) correlation coefficients. (C, D) Contours 
-of probability density functions (pdf) of Clayton (C) and survival-Clayton 
-(D) copulas with parameter 5 (see section on methods for Q1), together
-with samples from the corresponding distributions. (E,F) Reflections
-of C, D, respectively, about a vertical axis. (G) Contours of the
-pdf of a normal copula (methods for Q1) with 
-parameter chosen to
-have the same Spearman correlation as A-D, together with a sample.
-(H) The portion of the Spearman correlation within a band 
-can be computed
-for any band (methods for Q1) to describe how dependence
-varies from one part of the two distributions to another. Diagonal lines
-show two bands, the data in the left/lower band showing stronger
-dependence than the right/upper band.
-Subscripts h and v 
-stand for "horizontal" and "vertical".
-
-\hspace{1in}
-
-\noindent \textbf{Figure 2.} Summary and guide for analyses presented in the text. Middle boxes
-correspond to ecological datasets for which Q1 was examined (in the 
-second through
-fourth sections of the text, see also Table \ref{tab_data_info}). Upper boxes 
-correspond to potential causes of non-normal (non-N) copula structure that
-were examined (in the fifth and sixth sections of the text). Lower boxes 
-correspond to potential consequences we examined of non-N copula structure for
-ecological understanding and for applications (in the seventh and eighth
-sections of the text). Arrow labels (A-D for analyses pertaining 
-to causes, Q2; and X-Z for analyses pertaining to consequences, Q3) 
-link to locations in the text.
-
-\hspace{1in}
-
-\noindent \textbf{Figure 3.} Upper panels show raw data plots for (A) $\log_{10}$(soil C) and $\log_{10}$(soil N) data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
-for birds, (C) the same for mammals, and (D) above-ground plant biomass vs.
-Shannon's index (\textit{H}) from Cedar Creek.
-Bottom panels show the corresponding copula plots.
-See Data section for units used in upper panels.
-
-\hspace{1in}
-
-\noindent \textbf{Figure 4.} If environmental effects operate asymmetrically in their tails on
-ecological variables, it can result in non-N copula structure and asymmetric 
-tail dependence across space in the ecological variables. Shown are the last
-$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text, $b=0.1$.
-
-\hspace{1in}
-
-\noindent \textbf{Figure 5.} Example results showing similarity of copula statistics for 
-environmental-noise inputs and ecological-variable outputs of a dynamical model.
-Asymmetric tail dependence was similar in both noise inputs and model outputs
-for this model. The 
-AR(1) model and a Clayton copula were used (see methods section for Q2).
-Each point is the mean across 50 replicate simulations for which the same 
-Kendall $\tau$ value was used. Error bars are standard errors 
-and panel headers give Pearson correlation results for
-the points. The regression line through the points (black line) was similar
-to the 1-1 line (green line). 
-See Fig. \ref{SM-fig_Cause4copula_C_kend_ar1} for additional results for the AR(1) model and Clayton copula, 
-Figs \ref{SM-fig_Cause4copula_SC_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1} for AR(1) results with other copula families, and 
-Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1} for
-the nonlinear models described in the methods section for Q2.
-
-\hspace{1in}
-
-\noindent \textbf{Figure 6.} Skewness of spatially averaged green spruce aphid abundance $(A)$, leaf-curling plum 
-aphid first flight dates $(B)$, \emph{Ceratium furca} abundance $(C)$, and methane-flux
-$(D)$ time series compared to an multivariate 
-N-copula null hypothesis. Black dots are empirical 
-skewnesses; see text 
-and Appendix \ref{SM-Surrog_multi} for details of the null hypothesis. Results show a 
-tendency for skewness of the spatial average to be affected as hypothesized 
-by tail dependence.
-
-\hspace{1in}
-
-\noindent \textbf{Figure 7.} Extinction risk for the metapopulation extension of the Lewontin-Cohen model, after 
-$25$ time steps, was higher for left-tail-dependent environmental noise. Dispersal was local,
-with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
-for $N=5$ (A) and $N=25$ (B).

### 2019-09-10 Paper.Rmd
new pedagog fig created, changes made to old pedagog fig

@@
 \includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
-\caption{Bivariate datasets showing diverse relationships between 
+\caption{(A-D) Bivariate datasets showing diverse relationships between 
 variables; some of these datasets nevertheless have the same Pearson (P)
-or Spearman (S) correlation coefficients. Subscripts h and v 
+or Spearman (S) correlation coefficients. C and D show normalized ranks of A, B, 
+respectively. (E,F) Reflections
+of C, D, respectively, about a vertical axis. Subscripts h and v 
 in the axis labels stand for "horizontal" and "vertical".\label{fig_cop_pedag1}}
@@
 respectively. See, for instance,
-Fig. \ref{fig_cop_pedag1}C-D (the points), which show the normalized ranks of 
+Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
 A-B, respectively. Ranking makes the marginal distributions of 
@@
 and then four panels showing what you get when each of these combine?
-Good idea. Steps: split the current Fig 1 so the new Fig 1 has the first four panels,
-but C-D should only have the points. Then make a Fig 2 as above, and edit the
-text in the above para to use that figure. 
-The rest of the panels of the current Fig 1 will make a new Fig 3 later.-->
+Good idea. Steps: 
+DONE split the current Fig 1 so the new Fig 1 has the first 6 panels,
+but C-D should only have the points. 
+Then make a Fig 2 as above, and edit the text in the above para to use that figure. 
+The rest of the panels of the current Fig 1 will make a new Fig 3 later.
+-->
 

### 2019-09-10 FirstPedagogFig.R

@@ -43,7 +43,7 @@ gap<-0.1
 pan_wd<-1
 pan_ht<-pan_wd
 tot_wd<-ymarg_wd+2*gap+2*pan_wd+2*numsp
-tot_ht<-xmarg_ht+2*gap+2*pan_ht+2*numsp
+tot_ht<-xmarg_ht+3*gap+3*pan_ht+3*numsp
 pchval<-20
 cexvalpts<-0.25
 cexvaltxt<-0.75
@@ -53,8 +53,8 @@ pdf(file="./Results/PedagogFig1.pdf",width=tot_wd,height=tot_ht)
 #upper-left panel, normal marginals, left-tail dependence
 par(fig=c((ymarg_wd+numsp)/tot_wd,
           (ymarg_wd+numsp+pan_wd)/tot_wd,
-          (xmarg_ht+1*pan_ht+1*gap+2*numsp)/tot_ht,
-          (xmarg_ht+2*pan_ht+1*gap+2*numsp)/tot_ht),
+          (xmarg_ht+2*pan_ht+2*gap+3*numsp)/tot_ht,
+          (xmarg_ht+3*pan_ht+2*gap+3*numsp)/tot_ht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
 rg<-max(abs(dcc5_nm))
 plot(dcc5_nm[,1],dcc5_nm[,2],type='p',pch=pchval,cex=cexvalpts,xlim=c(-rg,rg),ylim=c(-rg,rg),col="grey")
@@ -68,8 +68,8 @@ text(rg,-rg,paste0("S=",round(cor(dcc5_nm[,1],dcc5_nm[,2],method="spearman"),2))
 #upper-right panel, normal marginals, right-tail dependence
 par(fig=c((ymarg_wd+2*numsp+pan_wd+gap)/tot_wd,
           (ymarg_wd+2*numsp+2*pan_wd+gap)/tot_wd,
-          (xmarg_ht+1*pan_ht+1*gap+2*numsp)/tot_ht,
-          (xmarg_ht+2*pan_ht+1*gap+2*numsp)/tot_ht),
+          (xmarg_ht+2*pan_ht+2*gap+3*numsp)/tot_ht,
+          (xmarg_ht+3*pan_ht+2*gap+3*numsp)/tot_ht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
[…]

### 2019-09-10 PedagogFigCopThry.R

@@ -1,0 +1,371 @@
+#***
+#This fle generates a pedagogical figure for showing people how different copulas
+#can be combined with different marginals to give different distributions
+#***
+
+rm(list=ls())
+
+library(copula)
+
+#***make the two copulas
+
+#**Gaussian copula
+ncop<-normalCopula(.7)
+
+#**Clayton copula with same Spearman
+ccop<-claytonCopula(2)
+ccop<-claytonCopula(iRho(ccop,rho(ncop)))
+
+#***make the marginals
+dnmarg<-function(x){return(dnorm(x,mean=0,sd=1))}
+pnmarg<-function(x){return(pnorm(x,mean=0,sd=1))}
+qnmarg<-function(x){return(qnorm(x,mean=0,sd=1))}
+
+shape<-5
+scale<-1
+dgmarg<-function(x){return(dgamma(x,shape=shape,scale=scale))}
+pgmarg<-function(x){return(pgamma(x,shape=shape,scale=scale))}
+qgmarg<-function(x){return(qgamma(x,shape=shape,scale=scale))}
+
+#***other stuff
+numsamps<-50
+
+#***make the plot
+
+#**set the plot up, units inches
+totwd<-6.5
+biggap<-.2
+smallgap<-.1
+xmarht<-.5
+ymarwd<-xmarht
+panwd<-(totwd-3*ymarwd-2*biggap-4*smallgap)/(4)
+margpansm<-panwd/3
+panht<-panwd
+margpanbg<-panht
+totht<-2*xmarht+2*panht+2*margpansm+biggap+3*smallgap
+if (!isTRUE(all.equal(totwd,3*ymarwd+3*panwd+3*margpansm+2*biggap+4*smallgap)))
+{
+  stop("Error in figure dimensions")
+}
+pdf(file="./Results/PedagogFigCopThry.pdf",width=totwd,height=totht)
+
+#**panel for contours for the normal cop
+par(fig=c((ymarwd)/totwd,
+          (ymarwd+panwd)/totwd,
+          (2*xmarht+panht+margpansm+smallgap+biggap)/totht,
+          (2*xmarht+2*panht+margpansm+smallgap+biggap)/totht),
[…]

### 2019-09-10 Paper.Rmd
still working on Background section

@@
 an asymmetric pattern of dependence that can
-have ecological significance, as discussed below, but is not captured by 
+have ecological significance, as discussed below, but that is not captured by 
 correlation. Although correlation is only one way to 
@@
 of probability density functions (pdf) of Clayton (C) and survival-Clayton 
-(D) copulas with parameter 5 (see section on methods for Q1), together
+(D) copulas with parameter 5 (see sections on methods for Q1), together
 with samples from the corresponding distributions. (E,F) Reflections
@@
 Spearman or Kendall correlation can have a range of tail dependence 
-(Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see section on
+(Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see sections on Background and
 methods for Q1 below). Although copulas can be used to model the entire
@@
 the marginals of a normal distribution. So Q1 asks whether ecological
-data contain dependency information distinct from what can be modelled with
-standard approaches. (Q2) If the answer to
-Q1 is "yes," what are some possible causes/mechanisms of non-normal
-copula structure and tail dependence? And (Q3) what are the 
+data contain dependency information distinct from that standard or default cases. 
+(Q2) If the answer to
+Q1 is "yes," then what are some possible causes/mechanisms of non-normal
+copula structure and tail dependence in ecology? And (Q3) what are the 
 consequences of non-normal copula
@@
 variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000],
-but have been used only rarely so far in ecology (but see @Valpine2014, 
-@Anderson2018 and @Popovic2018). 
-We argue that benefits wider usage of complete 
+and have been used very effectively, but rarely so far in ecology [@Valpine2014; 
+@Anderson2018; @Popovic2018]. 
+We argue that benefits of wider usage of complete 
 copula descriptions of dependence in ecology will be substantial.
@@
 associated with $(X,Y)$ is the distribution
-function of the random variable $(U,V)=(F_X(X),F_Y(Y))$ 
-(Appendix \ref{ContStrictMonoton}). And 
+function of the random variable $(F_X(X),F_Y(Y))$ 
+(Appendix \ref{SM-ContStrictMonoton}). And 
 $D(G_A(x),G_B(y))$ is the distribution function of the random variable 
@@
 $G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
-(Appendix \ref{ContStrictMonoton}).
+(Appendix \ref{SM-ContStrictMonoton}). Application of the 
+distributions functions of the component random variables translates
+between the uniform-marginal context, which isolates the 
+dependence information, and the context of the original random variable.
 
@@
 bivariate distributions in two separate steps: by specifying the marginal 
-distributions and by specifying the dependence structure. The theorem
-also make it possible to study and model dependence independently 
+distributions and by specifying the dependence structure. For instance,
+construction of several bivariate distributions is carried out in this way in
+Fig. \ref{PedagogFigCopThry}. The contrast between Fig. \ref{PedagogFigCopThry}C and D
+illustrates in particular how a bivariate distribution can be changed, while retaining 
+the same marginals, by changing the copula. Fig. \ref{PedagogFigCopThry}C is a 
+bivariate normal distribution, but Fig. \ref{PedagogFigCopThry}D clearly 
+is not, although its marginals
+are the same. Fig. \ref{PedagogFigCopThry}D shows stronger left- than right-tail dependence. 
+Sklar's theorem
+also makes it possible to study and model dependence separately 
 from marginal distributions. For instance, in a pedagogical 
-example, [@Anderson2018] modelled the dependence between abundance
+example, @Anderson2018 modelled the dependence between abundance
 data for two fish species, red moki and black angelfish, by first 
-modelling the marginal distributions and then separately modelling 
+modelling the marginal distributions and then modelling 
 the dependence using Gaussian copulas (their section 2).
@@
 but C-D should only have the points. 
-Then make a Fig 2 as above, and edit the text in the above para to use that figure. 
-The rest of the panels of the current Fig 1 will make a new Fig 3 later.
+DONE Then make a Fig 2 as above, 
+DONE and edit the text in the above para to use that figure. 
+STILL HAVE TO DO The rest of the panels of the current Fig 1 will make a new Fig 3 later.
 -->
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=6.5in]{./Results/PedagogFigCopThry.pdf}
+\caption{Normal (A) and Clayton (B) copulas were combined with standard normal 
+marginals (C, D) and gamma marginals (E, F) via Sklar's theorem to produce 
+bivariate distributions. Note that a normal copula is distinct from a normal marginal 
+distribution: a normal copula is the copula of a bivariate normal distribution.
+See the Background section for more information on normal, Clayton,
+and other copulas. Each copula was used with both sets of marginals and each
+set of marginals was used with both copulas, to demonstrate that: both copula and 
+marginals contribute fundamentally to the resulting distribution, the copula 
+contributing the information on the dependence between the variables; and that 
+one can select the copula and marginals independently. Bivariate
+distributions, including the copulas, are depicted via their log-scale probability density functions (pdfs), 
+and marginals are depicted via their linear-scale pdfs. Grey dots are
+50 random samples from each distribution. The parameter $0.7$ was used for the normal 
+copula, the parameter $2$ for the Clayton 
+copula, and shape and scale parameters $5$ and $1$ for the gamma marginals. 
+Variables $u$ and $v$ were used for copulas
+and $x$ and $y$ for distributions created by pairing a copula with
+(non-uniform) marginals.
+\label{PedagogFigCopThry}}
+\end{center}
+\end{figure}
 
@@
 1) DONE - The definition, Sklar's theorem
-2) What good are copulas - intuitive idea, basically, but also have to be linked to 
+2) DONE What good are copulas - intuitive idea, basically, but also have to be linked to 
 the formalities of the definition
-
-What some of the different copula families look like
+3) What some of the different copula families look like - in particular you will use BB1 in Fig S2 and cite Background, so mention it
+  Three new figures: Normal, Clayton, BB1
+4) What tail dependence actually is
+
 The explanation of how ranks relate to copulas and estimating the copula parameters
 How to carry out fitting (basically, MLE fitting on ranks - Genest et al paper)
-What tail dependence actually is
+
+You are going to have to rearrange the order of some figs in Sup Mat because now the data section comes after Bacground, in which the different copula famlies are introduced
 -->
@@
 datasets-->
+
+We introduce some standard copula families, as example illustran
 
@@
 See @nelsen2006_copula, @joe2014_dependence for further background.
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Normal.pdf}
+\caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
+S is Spearman correlation; p is the value of the parameter for the N 
+family (it is a 
+one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively (see Background section for definition).\label{figped_N}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Clayton.pdf}
+\caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C 
+copulas. S is Spearman correlation; p is the value of the parameter for the C family (it 
+is a one-parameter family); and LT and UT are the measures of lower- and 
+upper-tail dependence, 
+respectively (see Background section for definition).\label{figped_C}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB1.pdf}
+\caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
+denote the two parameters of the family (it is a 
+two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
+respectively (see Background section for definition).\label{figped_BB1}}
+\end{center}
+\end{figure}
 

### 2019-09-10 SupportingInformation.Rmd
still working on Background section

@@
 source("./MakeCopulaPedagogFigs.R")
+source("./PedagogFigCopThry.R")
 ```
@@
 but the Spearman correlations are unchanged (Figs \ref{MT-fig_cop_pedag1}A-B, \ref{fig_transf_pearson_spearman}).
-Fig. \ref{MT-fig_cop_pedag1}C (the points, not the contours) can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
+A scatterploy very simlar to Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
 of both variables by the cdf of the standard normal distribution.
-Fig. \ref{MT-fig_cop_pedag1}D can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
+Likewise a scatterplot similar to Fig. \ref{MT-fig_cop_pedag1}D can be obtained from Fig. \ref{MT-fig_cop_pedag1}B in the same way.
 
@@
 upper-tail dependence can vary independently of correlation. (A-C) Data generated using the BB1 
-copula (see section on methods for Q1 in the main text) 
+copula (see Background section in the main text) 
 with three different 
@@
 
-# Proofs of statements related to Skalr's theorem in the case of continuous, strictly monotonic margins \label{ContStrictMonoton}
+# Proofs of statements related to Sklar's theorem in the case of continuous, strictly monotonic margins \label{ContStrictMonoton}
 
@@
 $D(G_A(x),G_B(y))$ is the distribution function of the random variable 
-$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ is the random
+$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ now denotes the random
 vector with uniform marginals associated with $D$ and
@@
 The inverses exist because of the strict monotonicity assumption.
-To see this, note that the distribution function of 
+To get to the desired result, note that the distribution function of 
 $(G_A^{-1}(U),G_B^{-1}(V))$ is $P[G_A^{-1}(U) \leq x, G_B^{-1}(V) \leq y]$,
@@
 But this is $D(G_A(x),G_B(y))$ because $D$ is the distribution function of 
-$(U,V)$, which proves the statement. In fact, a stronger statement assuming
+$(U,V)$. In fact, a stronger statement assuming
 neither continuity nor strict monotonicity is possible, if one uses generalized
@@
 
-<!--Now the copula figures-->
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Normal.pdf}
-\caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
-S is Spearman correlation; p is the value of the parameter for the N 
-family (it is a 
-one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively.\label{figped_N}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Clayton.pdf}
-\caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C copulas. S is Spearman correlation; p is the value of the parameter for the C family (it is a 
-one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
-respectively.\label{figped_C}}
-\end{center}
-\end{figure}
+<!--Now the copula figures that were not presented in the main text-->
 
@@
 respectively.\label{figped_F}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB1.pdf}
-\caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
-denote the two parameters of the family (it is a 
-two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
-respectively.\label{figped_BB1}}
 \end{center}

### 2019-09-10 PedagogFigCopThry.R

@@ -3,8 +3,6 @@
 #can be combined with different marginals to give different distributions
 #***
 
-rm(list=ls())
-
 library(copula)
 
 #***make the two copulas
@@ -38,20 +36,21 @@ biggap<-.2
 smallgap<-.1
 xmarht<-.5
 ymarwd<-xmarht
-panwd<-(totwd-3*ymarwd-2*biggap-4*smallgap)/(4)
+textspace<-0.2
+panwd<-(totwd-3*ymarwd-2*biggap-4*smallgap-textspace)/(4)
 margpansm<-panwd/3
 panht<-panwd
 margpanbg<-panht
-totht<-2*xmarht+2*panht+2*margpansm+biggap+3*smallgap
-if (!isTRUE(all.equal(totwd,3*ymarwd+3*panwd+3*margpansm+2*biggap+4*smallgap)))
+totht<-2*xmarht+2*panht+2*margpansm+biggap+3*smallgap+textspace
+if (!isTRUE(all.equal(totwd,3*ymarwd+3*panwd+3*margpansm+2*biggap+4*smallgap+textspace)))
 {
   stop("Error in figure dimensions")
 }
 pdf(file="./Results/PedagogFigCopThry.pdf",width=totwd,height=totht)
 
 #**panel for contours for the normal cop
-par(fig=c((ymarwd)/totwd,
-          (ymarwd+panwd)/totwd,
+par(fig=c((textspace+ymarwd)/totwd,
+          (textspace+ymarwd+panwd)/totwd,
           (2*xmarht+panht+margpansm+smallgap+biggap)/totht,
           (2*xmarht+2*panht+margpansm+smallgap+biggap)/totht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
@@ -69,6 +68,7 @@ contour(x,y,log10(z),levels=levs,lty=0,add=TRUE)
 mtext("u",1,1.2)
 mtext("v",2,1.2)
 mtext("A",3,.8,at=c(-.35))
+mtext("NORMAL COPULA",2,2.4)
 
 #***add samples
 set.seed(101)
@@ -76,8 +76,8 @@ samps_ncop<-rCopula(numsamps,ncop)
 points(samps_ncop[,1],samps_ncop[,2],type="p",pch=20,col="grey",cex=.5)
 
 #**normal copula plot, top marginal
-par(fig=c((ymarwd)/totwd,
-          (ymarwd+margpanbg)/totwd,
[…]

### 2019-09-11 Paper.Rmd
still working on Background

@@
 
+<!--***DAN: In the Discussion, you need to point out not only that we
+use only bivariate copulas (and there are multivariate ones), but also
+that we use only continuous marginals. Probably an adequate approximation
+for our data, but see the other papers I've seen recently which use count data,
+and discrete marginals. Approaches such as that will be necessary for some 
+datasets-->
+
+We introduce some standard copula families, as examples of copulas and also
+because we will fit these families to data later. The bivariate normal
+copula family (henceforth abbreviated N), which consists of copulas of bivariate normal
+distributions, is a one-parameter family. The parameter, which we here call $p$,
+corresponds to the correlation
+of the related bivariate normal distribution, and controls the degree of dependence
+between the variables. An N copula with $p=0.7$ was already introduced (Fig.
+\ref{PedagogFigCopThry}A). The formula for the
+copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
+cdf of a (univariate) standard normal and $\Phi_{2,p}$ is the joint cdf of a
+bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
+and $p$ in the off diagonal positions. However, formulas for copulas are often not as 
+directly informative about copula properties as pdf diagrams; Fig. \ref{figped_N} has 
+example pdfs correspondng to bivariate N copulas for various values of $p$.
+Note that the pdfs are symmetric
+around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
+tail dependence (see
+below for a more precise defintion of tail dependence). 
+The Clayton copula family (henceforth abbreviated C),
+of which we also already pictured an example (Fig.
+\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to N copulas, C copulas
+have stronger left- than right-tail dependence. The formula is 
+$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
+less direct intuition than do example pdfs (Fig. \ref{figped_C}). The survival
+Clayton family (here abbreviated SC) is the symmetric opposite of the C family, showing
+stronger right- than left tail dependence (Fig. \ref{SM-figped_C}). The BB1 copula 
+family is a two-parameter family, thereby providing more flexibility: for some parameters
+BB1 copulas have stronger left- than right-tail dependence, and for others the reverse
+(Fig. \ref{figped_BB1}). See @nelsen2006_copula and @joe2014_dependence for further
+information on these and other copula families.
+
+Next intro tail dependence
+
 <!--***DANr: Add a new section called Background on copulas.
@@
 the formalities of the definition
-3) What some of the different copula families look like - in particular you will use BB1 in Fig S2 and cite Background, so mention it
+3) DONE What some of the different copula families look like - in particular you will use BB1 in Fig S2 and cite Background, so mention it
   Three new figures: Normal, Clayton, BB1
@@
 
-<!--***DAN: In the Discussion, you need to point out not only that we
-use only bivariate copulas (and there are multivariate ones), but also
-that we use only continuous marginals. Probably an adequate approximation
-for our data, but see the other papers I've seen recently which use count data,
-and discrete marginals. Approaches such as that will be necessary for some 
-datasets-->
-
-We introduce some standard copula families, as example illustran
-
-We introduce the copula families and 
-measures of tail dependence we used.
-The normal copula family (henceforth N) is a one-parameter
-family, with parameter corresponding to the correlation of the related 
-bivariate normal distribtion, and controlling the degree of dependence 
-(Fig. \ref{fig_cop_pedag1}G for an example).
+<!--***DAN: text storage for later:
 Left- and right-tail dependence are $0$ for the N copula, and this 
@@
 positive upper-tail ($0.87$ in Fig. \ref{fig_cop_pedag1}D)
-and $0$ lower-tail dependence. For our model selection procedure (see below), we used 
+and $0$ lower-tail dependence. -->
+
+<!--***DAN: put this text in the methods. It won't help people here in the ms to have all
+these copulas intro'd. The sole purpose of them is give a variety of options to have a robust model
+selection and multi-model inference procedure for tail dependence, so mention them in that context in 
+the Methods in a way that make clear we don't need to give a damn about these things:
+For our model selection procedure (see below), we used 
 several families of copulas exhibiting various 
@@
 tail dependence.
-See @nelsen2006_copula, @joe2014_dependence for further background.
+See @nelsen2006_copula, @joe2014_dependence for further background. -->
 
@@
 \begin{center}
-\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Normal.pdf}
+\includegraphics[width=\textwidth]{./Results/PedagogSuppMat_Normal.pdf}
 \caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
@@
 \begin{center}
-\includegraphics[width=7.5in]{./Results/PedagogSuppMat_Clayton.pdf}
+\includegraphics[width=\textwidth]{./Results/PedagogSuppMat_Clayton.pdf}
 \caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C 
@@
 \begin{center}
-\includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB1.pdf}
+\includegraphics[width=\textwidth]{./Results/PedagogSuppMat_BB1.pdf}
 \caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2

### 2019-09-12 Paper.Rmd
continued work on Background

@@
 
-Next intro tail dependence
+Having introduced tail dependence in a conceptual way and referred conceptually to it
+in examples, we now need a more precise definition. 
+<!--***DAN: fill in-->
+
+Plots using normalized ranks were proposed in the Introduction as being conceptually
+similar to copulas. We here explain the connection. As stated above, the copula associated
+with a random variable $(X,Y)$ is the joint distribution function of the random variable
+$(F_X(X),F_Y(Y))$. Given a sample $(x_i,y_i)$, $i=1,\ldots,n$, from $(X,Y)$, let $\hat{F}_X$ and
+$\hat{F}_Y$ be the empirical cdfs associated with the $x_i$ and $y_i$, respectively.
+These are step functions of $x$ and $y$, respectively, that start at $0$ for low values of 
+$x$ and $y$ and jump by $1/n$
+at each of the data points; for large $n$ they approximate the cdfs $F_X$ and $F_Y$. 
+But $\hat{F}_X(x_i)$ equals the normalized rank of $x_i$ times
+$n/(n+1)$, and likewise $\hat{F}_Y(y_i)$ equals the normalized rank of $y_i$ times
+$n/(n+1)$. So the normalized rank in turn approximates the empirical cdf and therefore 
+the cdf. Thus the normalized rank pairs $(u_i,v_i)$ for $i=1,\ldots,n$ can be 
+regarded, for large $n$, as a sample from the bivariate pdf associated with the copula of $(X,Y)$,
+and the scatterplot of $v_i$ versus $u_i$ can be used to infer aspects of copula structure
+(see section on methods for Q1 below).
 
@@
   Three new figures: Normal, Clayton, BB1
-4) What tail dependence actually is
-
-The explanation of how ranks relate to copulas and estimating the copula parameters
+STILL HAVE TO DO 4) What tail dependence actually is
+DONE 5) The explanation of how ranks relate to copulas and estimating the copula parameters
 How to carry out fitting (basically, MLE fitting on ranks - Genest et al paper)
-
+-->
+
+<!--***DAN:
 You are going to have to rearrange the order of some figs in Sup Mat because now the data section comes after Bacground, in which the different copula famlies are introduced

### 2019-09-14 Paper.Rmd
background section complete (though still may need some editing)

@@
 # Abstract
-<!--***DANr: I have not edited this abstract at all since Ecol Monogr-->
+<!--***DANr: I have not edited this abstract at all since Ecol Monogr, it needs to be redone-->
 All branches of ecology study relationships among environmental and biological variables. However, 
@@
 
-<!--DONE ***DANr: reword this para so people don't think the ranks *are* the 
-copulas, and to try to pre-defend against some of what ref 1 said. Also,
-state that these are conceptual introductions only and are not formal,
-a more comprehensive introduction is in xyz-->
 Well-developed approaches do exist, however, and are 
@@
 A main benefit of a copula approach is 
-that it can detect tail dependence and asymmetries of tail dependence. 
-Tail dependence (defined precisely in the section on Background, below) 
+that it can detect associations in the tails of distributions, and asymmetries of such associations. 
+Tail association (introduced formally in the section on Background, below) 
 is association between extreme values of variables. 
@@
 strongly associated than are larger values, the variables are said to
-have stronger lower- or left-tail dependence than right- or upper-tail dependence
+have stronger lower- or left-tail association than right- or upper-tail association
 (Fig. \ref{fig_cop_pedag1}C);
@@
 ones (Fig. \ref{fig_cop_pedag1}D). Datasets of the same 
-Spearman or Kendall correlation can have a range of tail dependence 
+Spearman or Kendall correlation can have a range of tail association 
 (Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see sections on Background and
 methods for Q1 below). Although copulas can be used to model the entire
-complex dependence structure of data, we have found tail dependence to be
-a useful aspect of the full structure of dependence, so tail dependence 
+complex dependence structure of data, we have found tail associations to be
+a useful aspect of the full structure of dependence, so tail associations 
 is one of the main foci of this paper.
@@
 correspond to ecological datasets for which Q1 was examined (in the 
-second through
-fourth sections of the text, see also Table \ref{tab_data_info}). Upper boxes 
+third through
+fifth sections of the text, see also Table \ref{tab_data_info}). Upper boxes 
 correspond to potential causes of non-normal (non-N) copula structure that
-were examined (in the fifth and sixth sections of the text). Lower boxes 
+were examined (in the sixth and seventh sections of the text). Lower boxes 
 correspond to potential consequences we examined of non-N copula structure for
-ecological understanding and for applications (in the seventh and eighth
+ecological understanding and for applications (in the eighth and ninth
 sections of the text). Arrow labels (A-D for analyses pertaining 
@@
 
-<!--DONE ***DANr: Maybe use the words tail dependence more and the word copula
-less in the below few paras? Generally the words "copula structure" may be rubbing some
-people wrong. Replacement words? What do you actually mean here? Maybe call it
-asymmetric dependence, or complex dependence structure?-->
 One way we may expect, *a priori*, a copula 
@@
 @li2013]. One may expect an environmental variable 
-measured through time in two locations to show strong tail dependence between
+measured through time in two locations to show strong tail associations between
 the locations if
@@
 
-<!--DONE ***DANr: Need some caveating here to accommodate the people who slavishly
-follow the old maxim that correlation is not causation.-->
 A third reason we may expect an understanding of complex 
@@
 possible consequences. For instance, if populations of a 
-pest species in different locations are all positively associated and also strongly
+pest species in different locations are all positively associated and also more strongly
 related to each other in their right tails, then local outbreaks will tend 
 to occur together, creating regional epidemics. 
-On the other hand left-tail dependence in a pest species, even if overall 
+On the other hand, stronger left-tail associations in a pest species, even if overall 
 correlation were unchanged, would have more benign effects.
@@
 distribution (here called non-normal copula structure)? Do they show tail 
-dependence distinct from that of a normal distribution, and in particular
-do they show asymmetric tail dependence? Normal copula structure is assumed
+associations distinct from those of a normal distribution, and in particular
+do they show asymmetric tail associations? Normal copula structure is assumed
 by standard approaches that use multivariate 
@@
 Q1 is "yes," then what are some possible causes/mechanisms of non-normal
-copula structure and tail dependence in ecology? And (Q3) what are the 
+copula structure and asymmetric tail associations in ecology? And (Q3) what are the 
 consequences of non-normal copula
-structure and tail dependence for ecological understanding and
+structure and asymmetric tail associations for ecological understanding and
 applications? After a Background section (the second section), we address Q1 
@@
 is not, although its marginals
-are the same. Fig. \ref{PedagogFigCopThry}D shows stronger left- than right-tail dependence. 
-Sklar's theorem
+are the same. Fig. \ref{PedagogFigCopThry}D shows stronger association of the two variables
+in the left than in the right tails. Sklar's theorem
 also makes it possible to study and model dependence separately 
@@
 around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
-tail dependence (see
-below for a more precise defintion of tail dependence). 
+associations between the two variables in the left and right tails (see
+below for a more precise defintion of tail associations). 
 The Clayton copula family (henceforth abbreviated C),
@@
 \ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to N copulas, C copulas
-have stronger left- than right-tail dependence. The formula is 
+have stronger left- than right-tail association. The formula is 
 $\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
@@
 Clayton family (here abbreviated SC) is the symmetric opposite of the C family, showing
-stronger right- than left tail dependence (Fig. \ref{SM-figped_C}). The BB1 copula 
+stronger right- than left tail association (Fig. \ref{SM-figped_C}). The BB1 copula 
 family is a two-parameter family, thereby providing more flexibility: for some parameters
-BB1 copulas have stronger left- than right-tail dependence, and for others the reverse
+BB1 copulas have stronger left- than right-tail association, and for others the reverse
 (Fig. \ref{figped_BB1}). See @nelsen2006_copula and @joe2014_dependence for further
@@
 
-Having introduced tail dependence in a conceptual way and referred conceptually to it
-in examples, we now need a more precise definition. 
-<!--***DAN: fill in-->
+<!--***DAN: Note: I am using tail dependence for the specific thing defined on p 62 of Joe,
+and tail associations as a more general term. I have changed this in parts of the ms
+upsteam from this point, but need to remember to change it as I move downstream.-->
+
+Having introduced tail association conceptually and referred to it
+in examples, we now need a more precise definition of a measure of tail association.
+We use the term *tail association* to describe the general idea of the strength of
+associations between two variables in the tails of their distributions. One of the
+ways this is measured is called *tail dependence* (@joe2014_dependence, section 2.13; 
+@nelsen2006_copula, section 5.4), which we here define. Given a random vector $(X,Y)$
+with margins $F_X$, $F_Y$ and defining $U=F_X(X)$, $V=F_Y(Y)$, the upper-tail
+dependence of $X$ and $Y$ is defined as 
+$\lambda_U = \lim_{u \rightarrow 1^-} P[Y>F_Y^{-1}(u) | X>F_X^{-1}(u)]$. This equals
+$\lim_{u \rightarrow 1^-} P[U>u | V>u]$, which in turn equals 
+$\lim_{u \rightarrow 1^-} P[U>u, V>u]/P[V>u]=\lim_{u \rightarrow 1^-} P[U>u, V>u]/(1-u)$,
+which shows upper-tail dependence is defined symmetrically in the two random variables.
+Lower tail dependence is defined analogously as 
+$\lambda_L = \lim_{u \rightarrow 1^-} P[Y \leq F_Y^{-1}(u) | X \leq F_X^{-1}(u)]$.
+Tail dependence is a property of the random variable $(U,V)$, so depends only
+on the copula and not on the marginals combined with it. It can be shown 
+(@nelsen2006_copula, section 5.4) that $\lambda_L=\lambda_U=0$ for the N copula,
+$\lambda_L=2^{-1/p}$ and $\lambda_U=0$ for a C copula with parameter $p>0$,
+and $\lambda_L=0$ and $\lambda_U=2^{-1/p}$ for an SC copula with parameter $p>0$.
+In the section on methods for Q1 we will introduce
+ways tail dependence can be applied to data, by first fitting copulas to the data. We will also 
+introduce nonparametric measures of tail association that apply directly to data.
 
@@
 
-<!--***DANr: Add a new section called Background on copulas.
-It can have the following kinds of stuff (probably not exhaustive, and probably
-not in order, and some of this material is currently below):
-1) DONE - The definition, Sklar's theorem
-2) DONE What good are copulas - intuitive idea, basically, but also have to be linked to 
-the formalities of the definition
-3) DONE What some of the different copula families look like - in particular you will use BB1 in Fig S2 and cite Background, so mention it
-  Three new figures: Normal, Clayton, BB1
-STILL HAVE TO DO 4) What tail dependence actually is
-DONE 5) The explanation of how ranks relate to copulas and estimating the copula parameters
-How to carry out fitting (basically, MLE fitting on ranks - Genest et al paper)
--->
-
 <!--***DAN:
@@
 -->
-
-<!--***DAN: text storage for later:
-Left- and right-tail dependence are $0$ for the N copula, and this 
-serves as a benchmark for tail dependence in other copulas - positive values of our
-tail dependence measures correspond to dependence stronger than an N
-copula. The Clayton and 
-survival-Clayton families (Fig. \ref{fig_cop_pedag1}C-D) are two more one-parameter 
-families. The Clayton has positive lower-tail dependence ($0.87$ for
-the example of Fig. \ref{fig_cop_pedag1}C) and $0$ upper-tail dependence; the 
-survival-Clayton is a rotation of the Clayton by $180^\circ$ and has 
-positive upper-tail ($0.87$ in Fig. \ref{fig_cop_pedag1}D)
-and $0$ lower-tail dependence. -->
 

### 2019-09-14 SupportingInformation.Rmd
background section complete (though still may need some editing)

@@
 with stronger correlation. Asymmetry of tail dependence (LT-UT) is the same 
-within each column.\label{fig_cop_pedag}}
+within each column. See Background section in the main text for a formal description of 
+tail dependence and other measures of association in the tails.\label{fig_cop_pedag}}
 \end{center}

### 2019-09-19 Paper.Rmd
done a roudn of edits up through Concepts and Methods for Q1

@@
 \end{figure}
-<!--***DAN: When I cut panels out of this figure (saving them for use in a 
-later figure), I also cut the caption, and here is the cut stuff for later use:
-(C, D) Contours 
-of probability density functions (pdf) of Clayton (C) and survival-Clayton 
-(D) copulas with parameter 5 (see sections on methods for Q1), together
-with samples from the corresponding distributions. (E,F) Reflections
-of C, D, respectively, about a vertical axis. (G) Contours of the
-pdf of a normal copula (methods for Q1) with 
-parameter chosen to
-have the same Spearman correlation as A-D, together with a sample.
-(H) The portion of the Spearman correlation within a band 
-can be computed
-for any band (methods for Q1) to describe how dependence
-varies from one part of the two distributions to another. Diagonal lines
-show two bands, the data in the left/lower band showing stronger
-dependence than the right/upper band.
-Subscripts h and v 
-stand for "horizontal" and "vertical".
--->
 
@@
 and therefore reflects not only dependence information, but also
-information on the marginals. Spearman and Kendall correlations, being rank-based, 
+information on the marginals \cite{Genest2007}. 
+Spearman and Kendall correlations, being rank-based, 
 are not influenced by monotonic transformations of the component variables; 
@@
 to copula concepts in the Background section below, but this is a sketch
-only - readers can find abbreviated [@Anderson2018] 
+only - readers can find abbreviated [@Anderson2018; @Genest2007] 
 and comprehensive introductions elsewhere
@@
 the marginals of a normal distribution. So Q1 asks whether ecological
-data contain dependency information distinct from that standard or default cases. 
+data contain dependency information distinct from the standard or default case. 
 (Q2) If the answer to
@@
 and on concepts needed
-for the rest of the paper. See @nelsen2006_copula, @joe2014_dependence, and
+for the rest of the paper. See @nelsen2006_copula, @Genest2007, @joe2014_dependence, and
 @MaiScherer2017 for general background, and see also @Anderson2018, part of 
@@
 
-<!--***DAN: I still feel like we have to make this more intuitive for people
-at this point. Maybe a figure with two panels showing two copulas on
-the left, and two panels shwoing two sets of marginals at the top,
-and then four panels showing what you get when each of these combine?
-Good idea. Steps: 
-DONE split the current Fig 1 so the new Fig 1 has the first 6 panels,
-but C-D should only have the points. 
-DONE Then make a Fig 2 as above, 
-DONE and edit the text in the above para to use that figure. 
-STILL HAVE TO DO The rest of the panels of the current Fig 1 will make a new Fig 3 later.
--->
-
 \begin{figure}[!h]
@@
 directly informative about copula properties as pdf diagrams; Fig. \ref{figped_N} has 
-example pdfs correspondng to bivariate N copulas for various values of $p$.
+example pdfs corresponding to bivariate N copulas for various values of $p$.
 Note that the pdfs are symmetric
@@
 Clayton family (here abbreviated SC) is the symmetric opposite of the C family, showing
-stronger right- than left tail association (Fig. \ref{SM-figped_C}). The BB1 copula 
+stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
 family is a two-parameter family, thereby providing more flexibility: for some parameters
@@
 which shows upper-tail dependence is defined symmetrically in the two random variables.
-Lower tail dependence is defined analogously as 
+Lower tail dependence is defined analogously, as 
 $\lambda_L = \lim_{u \rightarrow 1^-} P[Y \leq F_Y^{-1}(u) | X \leq F_X^{-1}(u)]$.
@@
 the cdf. Thus the normalized rank pairs $(u_i,v_i)$ for $i=1,\ldots,n$ can be 
-regarded, for large $n$, as a sample from the bivariate pdf associated with the copula of $(X,Y)$,
+regarded as approximate samples from the bivariate pdf 
+associated with the copula of $(X,Y)$,
 and the scatterplot of $v_i$ versus $u_i$ can be used to infer aspects of copula structure
 (see section on methods for Q1 below).
-
-<!--***DAN:
-You are going to have to rearrange the order of some figs in Sup Mat because now the data section comes after Bacground, in which the different copula famlies are introduced
--->
-
-<!--***DAN: put this text in the methods. It won't help people here in the ms to have all
-these copulas intro'd. The sole purpose of them is give a variety of options to have a robust model
-selection and multi-model inference procedure for tail dependence, so mention them in that context in 
-the Methods in a way that make clear we don't need to give a damn about these things:
-For our model selection procedure (see below), we used 
-several families of copulas exhibiting various 
-lower- and upper-tail dependence characteristics 
-(Clayton, survival-Clayton, Gumbel, survival-Gumbel, Joe, survival-Joe, Frank,
-BB1, survival-BB1,
-BB6, survival-BB6, BB7, survival-BB7, BB8, survival-BB8, henceforth
-abbreviated C, SC, G, SG, J, SJ, F, BB1, SBB1, BB6, SBB6, BB7, SBB7, BB8, SBB8).
-Copulas are implemented in the R packages `copula` and `VineCopula`. 
-Visualizations of these families (Figs \ref{SM-figped_N}-\ref{SM-figped_SBB8})
-can be compared with normalized rank plots to estimate visually 
-which families may effectively model a given bivariate dataset. 
-If LT and UT denote 
-lower- and upper-tail dependence measures, their difference measures asymmetry of 
-tail dependence.
-See @nelsen2006_copula, @joe2014_dependence for further background. -->
 
@@
 \caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
-S is Spearman correlation; p is the value of the parameter for the N 
+S is Spearman correlation; $p$ is the value of the parameter for the N 
 family (it is a 
-one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively (see Background section for definition).\label{figped_N}}
+one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively (see Background section for definition). The 
+parameter range for the family is $p \in [-1,1]$, and lower- and upper-tail
+dependence are always $0$.\label{figped_N}}
 \end{center}
@@
 \caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C 
-copulas. S is Spearman correlation; p is the value of the parameter for the C family (it 
+copulas. S is Spearman correlation; $p$ is the value of the parameter for the C family (it 
 is a one-parameter family); and LT and UT are the measures of lower- and 
 upper-tail dependence, 
-respectively (see Background section for definition).\label{figped_C}}
+respectively (see Background section for definition). The parameter range for the family is $p \in (0,\infty)$, lower-tail dependence is $2^{-1/p}$ and upper-tail
+dependence is $0$.\label{figped_C}}
 \end{center}
@@
 \includegraphics[width=\textwidth]{./Results/PedagogSuppMat_BB1.pdf}
-\caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; p1 and p2
+\caption[Example BB1 copulas]{Log-transformed pdfs for example BB1 copulas. K is Kendall correlation; $p_1$ and $p_2$
 denote the two parameters of the family (it is a 
 two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
-respectively (see Background section for definition).\label{figped_BB1}}
+respectively (see Background section for definition). The parameter ranges for 
+the family are $p_1 \in (0,\infty)$ and $p_2 \in [1,\infty)$, 
+lower-tail dependence is $2^{-1/(p_1 p_2)}$ and upper-tail
+dependence is $2-2^{1/p_2}$.\label{figped_BB1}}
 \end{center}
@@
 
-<!--***DANr: I moved some text from here and have not smoothed the transtion,
-if it needs smoothing-->
-
-We addressed Q1 principally via a model selection procedure in which the 
-copulas listed above were fitted to 
+We addressed Q1 first via a model selection procedure in which several
+families of copulas were fitted to 
 our ecological datasets and fits were compared via the Akaike and Bayesian 
-Information Criteria (AIC and BIC). For bivariate datasets 
+Information Criteria (AIC and BIC). One of the fitted copulas was the N
+copula, making possible comparisons of the degree to which the N copula versus
+other copula families were good descriptions of data.
+
+For bivariate datasets 
 $(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
@@
 Second, we tested the independence of the $u_i$ and $v_i$
-(using `BiCopIndTest` in the `VineCopula` package). 
-Our model 
-selection algorithms were ineffective if data could not be distinguished from independent data 
+using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
+is Kendall's tau for the data \cite{Genest2007}. We used the 
+implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
+in the R programming language. 
+We tested for independence because our model selection 
+algorithms were ineffective if data could not 
+be distinguished from independent data 
 since many of the copula families we considered include the independent copula. 
 If independence could be rejected ($0.05$ significance level), model 
-selection proceeded. Third, we fit the 16 bivariate copula families listed previously to the 
-normalized ranks via maximum likelihood, using `BiCopEst` in `VineCopula`, 
-then obtaining AIC and BIC values and accompanying model weights $\text{AIC}_{\text{w}}$ and 
-$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection]. `BiCopEst` also provided the lower- and 
+selection proceeded. Third, we fit 16 bivariate copula families (see below for 
+information on the families used) to the 
+normalized ranks via maximum likelihood. The approach of fitting
+copula families to normalized ranks was studied in depth
+and recommended by @Genest1995 and @Shih1995. Their estimator of copula
+parameters, which we use, was shown to be consistent, asymptotically normal,
+and fully efficient at indepencence [@Genest1995]. @Genest2007
+recommend carrying out inferences of dependence structures (which was our
+goal here) using ranks.
+We used the fitting implementation given in `BiCopEst` in `VineCopula`.
+Fourth, we obtained AIC and BIC values and accompanying model weights 
+$\text{AIC}_{\text{w}}$ and 
+$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection] for each fitted copula. 
+`BiCopEst` also 
+provided lower- and 
 upper-tail dependence of the best-fitting member of each family. 
-Fourth, $\text{AIC}_{\text{w}}$ values were used to get
+$\text{AIC}_{\text{w}}$ values were used to get
 model-averaged lower- and upper-tail dependence values using standard model averaging 
 formulas [@burnham2003_modelselection]; likewise for BIC.
+
+Thus for each bivariate dataset, 
+the end products of the procedure just described were threefold, corresponding to the
+three parts of Q1 listed in the Introduction: A) the AIC, BIC, $\text{AIC}_{\text{w}}$ and 
+$\text{BIC}_{\text{w}}$ values for each of our 16 copula families (see below for list),
+including the normal family, providing an inference as to whether the normal
+copula or an alternative was better supported by 
+data; B) lower- and upper-tail dependence measures for each fitted copula and 
+model-averaged tail dependence measures, providing information on whether,
+and to what extent, tail dependence differed from the tail dependence of 
+a normal copula (i.e., 0); and C) the difference of lower- and upper-tail
+dependence measures for each fitted copula and model averages of those 
+quantities, providing information on whether, and to what extent, tail
+dependence was asymmetric.
+
 Model selection methods give the *relative* support of several models, but do not 
 indicate whether any of the models are an objectively good fit. To test that, 
-and as our final step, we tested the goodness of fit of our AIC-best copula family using a 
-bootstrapping procedure (`BiCopGofTest` of `VineCopula`). The procedure performed 
+we tested the goodness of fit of our AIC-best copula family using a 
+bootstrapping procedure developed and studied by @Wang2000 and @Genest2006
+and implemented as `BiCopGofTest` of `VineCopula`. The procedure performed 
 one test based on a Cramer-von Mises statistic and another based on a 
@@
 
-The first part of Q1 was answered for a bivariate dataset 
-by comparing the AIC weights of the non-N copula
-families with that of the N family. 
-The second part of Q1 was answered by comparing 
-the model-averaged lower- and upper-tail dependence
-values to the lower- and upper-tail dependence values of an N copula (i.e., 0). 
-The third part of Q1
-was answered by comparing the difference of the model-averaged lower- and upper-tail 
-dependence values to the analogous difference for an N copula, again 0.
 For multivariate datasets, we performed the bivariate analysis
@@
 or vice-versa, and reported the percentage of non-independent 
-location pairs showing stronger lower-tail dependence.
-
-In addition to our model selection approach, we also used 
-nonparametric approaches to provide greater 
+location pairs showing stronger lower-tail dependence. We carried out
+pairwise bivariate analyses for simplicity and because they were
+sufficient to answer our research questions. But see the Discussion
+for a few words on multivariate copulas.
+
+We fit 16 bivariate copula families, exhibiting a variety of 
+lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
+of using a large collection of families was solely to include a 
+variety of alternative dependence structures to have a robust model
+selection and multi-model inference procedure. For that purpose, it 
+is not important for the reader to
+understand the details of these copulas,
+and, additonally, these copulas have been 
+described formally elsewhere [@Joe1997]; so 
+we only briefly identify each family, say a few words about tail associations 
+of the families, and provide pictorial descriptions in the Appendices.
+We used several families that can exhibit positive lower-tail dependence 
+(of strength depending on parameters) and 
+zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
+and survival BB6 familes. We used several families that can exhibit positive
+upper-tail dependence (strength depends on parameters) and zero lower-tail
+dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
+families that show zero upper- and lower-tail dependence: the normal
+and Frank families. These families have pdfs symmetric about the line
+$v=-u+1$. We also used several families that can show both upper- and
+lower-tail dependence, in relative amounts that depend on parameter
+values: the BB1, survival BB1, BB7 and survival BB7 families. 
+We also used the BB8 copula, which shows zero lower-tail dependence,
+and zero upper-tail dependence except for a boundary case for the parameters.
+We used the survival BB8 copula, which shows zero upper-tail dependence, 
+and zero lower-tail dependence except for a boundary case for the 
+paremeters.
+"Survival" families are rotations of the copula with the similar name by
+180 degrees. 
+See @Joe1997 for details on all these families. We used the implementations
+of these families provided in the `VineCopula` package for the R programming
+language. See Figs \ref{figped_N}-\ref{figped_BB1} and \ref{SM-figped_SC}-\ref{SM-figped_SBB8} 
+for visual depictions of the pdfs of these copulas and how
+pdfs and tail dependence are influenced by the parameters.
+
+<!--***DAN: Info on families kept here for convenience, see also the file
+CopulaFamilyParamRangesAndTailDependence.jpg, stored in the BIVAN folder:
+lower not upper dependence:
+Clayton (3), survival Gumbel (14), survival Joe (16),
+SBB6 (18)
+zero both, and symmetric:
+  normal (1), Frank (5)
+upper not lower dependence:
+  SC (13), Gumbel (4), Joe (6)
+  BB6 (8)
+both, in relative amounts that depend on parameters:
+  BB1 (7) - maybe need to look at formulas for this one to understand what the   
+    limits are, since I only chose some random values of parameters in my pdf 
+    pictures
+  SBB1 (17)
+  BB7 (9), SBB7 (19)
+BB8 (10) - 0 LT dep, UT dep 0 except in a boundary case for the parameters
+SBB8 (20) - reverse of BB8-->
+
+<!--***DAN: You decided to reverse the decision to use pithy abbreviations for each
+copula. Now the format is longer, so this might just annoy people. So this will take
+some work to replace all those throughout the main text and sup mat,
+including in figures and table and captions - everywhere.-->
+
+In addition to our model selection approach, we also used a
+nonparametric approach, to provide greater 
 confidence in our answers to Q1. We used three 
@@
 related in any part of their distributions; we here describe the 
-statistics conceptually, with details in Appendix \ref{SM-nonparam_stats}.
+statistics, with some additional details in Appendix \ref{SM-nonparam_stats}.
 The statistics are defined with positively associated
@@
 $0 \leq l_b < u_b \leq 1$, define the lines $u+v=2l_b$ and 
-$u+v=2u_b$, which intersect the unit square (Figs \ref{fig_cop_pedag1}H, 
-\ref{SM-fig_stat_image}). 
-Our statistics quantify the dependence
+$u+v=2u_b$, which intersect the unit square (Fig. \ref{NonparamStatsFig}). 
+Our statistics quantify the association
 between $u_i$ and $v_i$ in the region bounded by these lines.
 Using $l_b=0$ and $u_b\leq 0.5$ gives information about 
-dependence in the left parts of the distributions of $u$ and $v$, and 
-using $u_b=1$ and $l_b \geq 0.5$ gives information about dependence in 
-the right parts. The first statistic, 
-$\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
+association in the left parts of the distributions of $u$ and $v$, and 
+using $u_b=1$ and $l_b \geq 0.5$ gives information about association in 
+the right parts. The first statistic, a partial Spearman correlation
+\begin{equation}
+\cor_{l_b,u_b}(u,v)=\frac{\sum 
+   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}},
+\end{equation}
+\noindent is the portion of the Spearman 
 correlation of $u_i$ and $v_i$ that is attributable to the points 
 between the bounds. 
-Larger values indicate stronger positive association.
+Here sample means and variances are computed using all $n$ data 
+points, but the
+sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
+$u_i+v_i < 2u_b$.
+Larger values of the partial Spearman correlation indicate stronger positive association.
+The sum of 
+$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
+$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
+interval $(0,1)$) equals the standard Spearman correlation
+as long as no points fall directly on the bounds.
 We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
-which has a similar 
-interpretation to $\cor_{l_{b},u_{b}}(u,v)$.
-Our third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
+which has a similar qualitative
+interpretation to $\cor_{l_{b},u_{b}}$.
+Our third statistic, $\Dsq_{l_b,u_b}$, is the average squared 
 distance between points satisfying $u_i+v_i>2l_b$ and 
@@
 Unlike $\cor_{l_b,u_b}$ and 
-$\Ps_{l_b,u_b}$, for which large values inidicate strong dependence
+$\Ps_{l_b,u_b}$, for which large values inidicate strong association
 between the bounds, small values of 
-$\Dsq_{l_b,u_b}$ indicate strong dependence.
+$\Dsq_{l_b,u_b}$ indicate strong association.
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=3.5in]{./Results/NonparamStatsFig.pdf}
+\caption{The partial Spearman correlation, $\cor_{l_{b},u_{b}}(u,v)$, 
+within a band can be computed
+for any band (methods for Q1) to describe how the strength of association
+between $u$ and $v$
+varies from one part of the two distributions to another, 
+as can the statistics $\Ps_{l_b,u_b}(u,v)$ and $\Dsq_{l_b,u_b}(u,v)$. Diagonal lines
+show two bands, the data in the left/lower band showing stronger
+association than those in the right/upper band.\label{NonparamStatsFig}}
+\end{center}
+\end{figure}
 
@@
 
-\newpage
+<!--***DAN: got to here 9/19-->
 
@@
 
-\newpage
-
 <!--Question 2 analyses-->
@@
 package [@DendroPy].
-
-\newpage
 
@@
 
-\newpage
-
 <!--Question 3 analyses-->
@@
 
-\newpage
-
 # Results for Q3: Tail dependence influences skewness of the spatial average, extinction risk, and Taylor's law
@@
 \end{figure}
-
-\newpage
 
@@
 
-\newpage
-
 # Acknowledgments

### 2019-09-19 SupportingInformation.Rmd
done a roudn of edits up through Concepts and Methods for Q1

@@
 <!--Now the copula figures that were not presented in the main text-->
-
 \begin{figure}[!h]
@@
 family (it is a one-parameter family); and LT and UT are the measures of lower- and 
-upper-tail dependence, respectively.\label{figped_SC}}
+upper-tail dependence, respectively. The parameter range for the family is $p \in (0,\infty)$, upper-tail dependence is $2^{-1/p}$ and lower-tail
+dependence is $0$.\label{figped_SC}}
 \end{center}
@@
 \caption[Example G copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
-Spearman correlation; p is the value of the parameter for the G family (it is a 
+Spearman correlation; $p$ is the value of the parameter for the G family (it is a 
 one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
-respectively.\label{figped_G}}
+respectively. The parameter range for the family is $p \in [1,\infty)$, lower-tail dependence is $0$ and upper-tail
+dependence is $2-2^{1/p}$.\label{figped_G}}
 \end{center}
@@
 \caption[Example SG copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
-copulas. S is Spearman correlation; p is the value of the parameter for the SG 
+copulas. S is Spearman correlation; $p$ is the value of the parameter for the SG 
 family (it is a one-parameter family); and LT and UT are the measures of lower- and 
-upper-tail dependence, respectively.\label{figped_SG}}
+upper-tail dependence, respectively. The parameter range for the family is $p \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
+dependence is $2-2^{1/p}$.\label{figped_SG}}
 \end{center}
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Joe.pdf}
-\caption[Example J copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; p is the value of the parameter for the J family (it is a 
+\caption[Example J copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; $p$ is the value of the parameter for the J family (it is a 
 one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
-respectively.\label{figped_J}}
+respectively. The parameter range for the family is $p \in [1,\infty)$, lower-tail dependence is $0$ and upper-tail
+dependence is $2-2^{1/p}$.\label{figped_J}}
 \end{center}
@@
 \caption[Example SJ copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
-copulas. K is Spearman correlation; p is the value of the parameter for the SJ 
+copulas. K is Spearman correlation; $p$ is the value of the parameter for the SJ 
 family (it is a one-parameter family); and LT and UT are the measures of lower- and 
-upper-tail dependence, respectively.\label{figped_SJ}}
+upper-tail dependence, respectively. The parameter range for the family is $p \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
+dependence is $2-2^{1/p}$.\label{figped_SJ}}
 \end{center}
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
-\caption[Example F copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; p is the value of the parameter for the F family (it is a 
+\caption[Example F copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; $p$ is the value of the parameter for the F family (it is a 
 one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
-respectively.\label{figped_F}}
+respectively. The parameter range for the family is $p \in (0,\infty)$, lower-tail dependence is $0$ and upper-tail
+dependence is $0$.\label{figped_F}}
 \end{center}
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
-\caption[Example SBB1 copulas]{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; p1 ad p2 
+\caption[Example SBB1 copulas]{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
 denote the two parameters of the family (it is a 
-two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
-respectively.\label{figped_SBB1}}
+two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively. The parameter ranges for the family are $p_1 \in (0,\infty)$
+and $p_2 \in [1,\infty)$, lower-tail dependence is $2-2^{1/p_2}$ and upper-tail
+dependence is $2^{-1/(p_1 p_2)}$.\label{figped_SBB1}}
 \end{center}
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB6.pdf}
-\caption[Example BB6 copulas]{Log-transformed pdfs for example BB6 copulas. K is Kendall correlation; p1 and p2
+\caption[Example BB6 copulas]{Log-transformed pdfs for example BB6 copulas. K is Kendall correlation; $p_1$ and $p_2$
 denote the two parameters of the family (it is a 
-two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
-respectively.\label{figped_BB6}}
+two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
+and $p_2 \in [1,\infty)$, lower-tail dependence is $0$ and upper-tail
+dependence is $2-2^{1/(p_1 p_2)}$. \label{figped_BB6}}
 \end{center}
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
-\caption[Example SBB6 copulas]{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; p1 ad p2 
+\caption[Example SBB6 copulas]{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
 denote the two parameters of the family (it is a 
-two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
-respectively.\label{figped_SBB6}}
+two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
+and $p_2 \in [1,\infty)$, upper-tail dependence is $0$ and lower-tail
+dependence is $2-2^{1/(p_1 p_2)}$.\label{figped_SBB6}}
 \end{center}
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB7.pdf}
-\caption[Example BB7 copulas]{Log-transformed pdfs for example BB7 copulas. K is Kendall correlation; p1 and p2
+\caption[Example BB7 copulas]{Log-transformed pdfs for example BB7 copulas. K is Kendall correlation; $p_1$ and $p_2$
 denote the two parameters of the family (it is a 
-two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
-respectively.\label{figped_BB7}}
+two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
+and $p_2 \in (0,\infty)$, lower-tail dependence is $2^{-1/p_2}$ and upper-tail
+dependence is $2-2^{1/p_1}$.\label{figped_BB7}}
 \end{center}
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
-\caption[Example SBB7 copulas]{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; p1 ad p2 
+\caption[Example SBB7 copulas]{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
 denote the two parameters of the family (it is a 
-two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
-respectively.\label{figped_SBB7}}
+two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
+and $p_2 \in (0,\infty)$, upper-tail dependence is $2^{-1/p_2}$ and lower-tail
+dependence is $2-2^{1/p_1}$.\label{figped_SBB7}}
 \end{center}
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_BB8.pdf}
-\caption[Example BB8 copulas]{Log-transformed pdfs for example BB8 copulas. K is Kendall correlation; p1 and p2
+\caption[Example BB8 copulas]{Log-transformed pdfs for example BB8 copulas. K is Kendall correlation; $p_1$ and $p_2$
 denote the two parameters of the family (it is a 
-two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
-respectively.\label{figped_BB8}}
+two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
+and $p_2 \in (0,1]$, lower-tail dependence is $0$ and upper-tail
+dependence is $2-2^{1/p_1}$ of $p_2=1$ and $0$ otherwise.\label{figped_BB8}}
 \end{center}
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
-\caption[Example SBB8 copulas]{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; p1 ad p2 
+\caption[Example SBB8 copulas]{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
 denote the two parameters of the family (it is a 
-two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
-respectively.\label{figped_SBB8}}
+two-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
+respectively. The parameter ranges for the family are $p_1 \in [1,\infty)$
+and $p_2 \in (0,1]$, upper-tail dependence is $0$ and lower-tail
+dependence is $2-2^{1/p_1}$ of $p_2=1$ and $0$ otherwise.\label{figped_SBB8}}
 \end{center}
@@
 
-\noindent We here define
-\begin{equation}\label{eq.Cor}
-   \cor_{l_b,u_b}(u,v) = \frac{\sum 
-   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}}.
-\end{equation}
-\noindent Here means and sample variances are computed using all $n$ data 
-points, but the
-sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
-$u_i+v_i < 2u_b$. 
-The sum of 
-$\cor_{0,0.5}(u,v)$ and $\cor_{0.5,1}(u,v)$ (or some other choice of 
-$\cor_{l_{b_k},u_{b_k}}(u,v)$ for bounds $l_{b_k},u_{b_k}$ that partition the
-interval $(0,1)$) equals the standard Spearman correlation.
-
-We also define $\Ps_{l_b,u_b}$, using the bounds
+\noindent We define $\Ps_{l_b,u_b}$, using the bounds
 $u+v=2l_b$ and $u+v=2u_b$ defined in the main text, as follows. For a distance $h$, we define
@@
 
+<!--Decided we do not need this fig.
 \begin{figure}[!h]

### 2019-09-19 FirstPedagogFig.R

@@ -205,6 +205,28 @@ mtext(expression(z[h]),1,line=1.15)
 
 dev.off()
 
+#***A pedagog fig for intro-ing the nonparam stats
+
+tot_wd<-ymarg_wd+numsp+pan_wd+gap
+tot_ht<-xmarg_ht+numsp+pan_ht+gap
+pdf(file="./Results/NonparamStatsFig.pdf",width=tot_wd,height=tot_ht)
+
+par(fig=c((ymarg_wd+numsp)/tot_wd,
+          (ymarg_wd+numsp+pan_wd)/tot_wd,
+          (xmarg_ht+numsp)/tot_ht,
+          (xmarg_ht+numsp+pan_ht)/tot_ht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
+plot(dcc5[,1],dcc5[,2],type='p',pch=pchval,cex=cexvalpts,xlim=c(rg1,rg2),ylim=c(rg1,rg2),col="grey")
+lines(c(0,.4),c(.4,0),type='l')
+lines(c(0,.65),c(.65,0),type='l')
+lines(c(0,1.6),c(1.6,0),type='l')
+lines(c(0,1.35),c(1.35,0),type='l')
+mtext("u",1,line=1.15)
+mtext("v",2,line=1.15)
+#text(rg1,rg2,"H",cex=cexvalpl,adj=c(0,1))
+
+dev.off()
+
 #***a pedagog fig for the sup mat about monotonic transformations
 
 #plot dimensions, units inches
@@ -254,3 +276,5 @@ text(rg2,rg1,paste0("S=",round(cor(dsc5_gm[,1],dsc5_gm[,2],method="spearman"),2)
 mtext(expression(z[h]),1,line=1.15)
 
 dev.off()
+
+

### 2019-09-20 Paper.Rmd
Beginning edits of Q1 results. Listed some things for Shya to do on line 893

@@
 
-<!--***DAN: got to here 9/19-->
-
 # Results for Q1: Ecological datasets have non-normal copula structure and asymmetric tail dependence \label{Results}
@@
 
+<!--***SHYAMOLINA: Please do the following to help me revise this section, so far only covering changes in the presentation of the bivariate results.
+Please let me know if any of this is unclear or you want to discuss anything.
+1) Replace the pithy abbreviations for the different copulas (N, C, SC, G, etc.) with their full names throughout the main text and sup mat, including in figure and table captions.
+This will not affect BB1, BB6, etc., but it will affect SBB1, SBB6, etc., which will now be called survival BB1, survival BB6, etc.
+2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc). Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
+3) Sort the rows of Table S3 (and all analogous tables, e.g., S5, S7, etc.) by AIC, putting the lowest AIC at the top. This should mean the highest AICw is also at the top.
+3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text. Same for Fig. S21 and Table S4.
+7) Combine tables S6, S8, S10 into one table, and fix it so
+whenever we currently cite one of those, change it so we cite the combined table instead.
+8) Put the cor results only, for Spearman only (or is Kendall more consistent with what we do elsewhere - what do you think?) from tables S4, S6, S8, S10 into Table 2 as a new row.
+-->
+
+<!--***DAN: Notes on what needs to be done to this section, so far only getting up through the presentation of the bivariate results:
+1) be more nuanced in the presentation, don't give the appearance of sweeping results that dont agree with the general conclusions under the rug
+2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc). Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
+3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text.
+4) Sort the rows of Table S2 (and all analogous tables) AIC, putting the lowest
+AIC at the top. This should mean the highest AICw is at the top.
+5) move some of the content from appendices S6 and S7 into methods, and move fig S21 and table S4 into main text. These are non-parametric results for
+soil C and N. So we will be presenting the nonparam results more comprehensively in the main text.
+6) That means you also need to summarize the nonparm results for the other 3 bivariate systems in Table 2.
+How to do this?
+Maybe put the cor results only, for Spearman only (or Kendall, not sure which to choose) into Table 2.
+7) Probably ought to combine tables S6, S8, S10 into one table, and fix it so
+whenever we currently cite one of those, change it so we cite the combined table instead.
+-->
+
 \begin{figure}[!h]
@@
 Shannon's index (\textit{H}) from Cedar Creek.
-Bottom panels show the corresponding copula plots.
+Bottom panels show the corresponding normalized rank plots.
 See Data section for units used in upper panels.

### 2019-09-20 Paper.Rmd
edits done as suggested by DCR: see my response

@@
 and therefore reflects not only dependence information, but also
-information on the marginals \cite{Genest2007}. 
+information on the marginals [@Genest2007]. 
 Spearman and Kendall correlations, being rank-based, 
@@
 were examined (in the sixth and seventh sections of the text). Lower boxes 
-correspond to potential consequences we examined of non-N copula structure for
+correspond to potential consequences we examined of non-Normalcopula structure for
 ecological understanding and for applications (in the eighth and ninth
@@
 of the related bivariate normal distribution, and controls the degree of dependence
-between the variables. An N copula with $p=0.7$ was already introduced (Fig.
+between the variables. An Normal copula with $p=0.7$ was already introduced (Fig.
 \ref{PedagogFigCopThry}A). The formula for the
@@
 directly informative about copula properties as pdf diagrams; Fig. \ref{figped_N} has 
-example pdfs corresponding to bivariate N copulas for various values of $p$.
+example pdfs corresponding to bivariate Normalcopulas for various values of $p$.
 Note that the pdfs are symmetric
@@
 of which we also already pictured an example (Fig.
-\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to N copulas, C copulas
+\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to Normalcopulas, Clayton copulas
 have stronger left- than right-tail association. The formula is 
@@
 less direct intuition than do example pdfs (Fig. \ref{figped_C}). The survival
-Clayton family (here abbreviated SC) is the symmetric opposite of the C family, showing
+Clayton family (here abbreviated SC) is the symmetric opposite of the Claytonfamily, showing
 stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
@@
 on the copula and not on the marginals combined with it. It can be shown 
-(@nelsen2006_copula, section 5.4) that $\lambda_L=\lambda_U=0$ for the N copula,
-$\lambda_L=2^{-1/p}$ and $\lambda_U=0$ for a C copula with parameter $p>0$,
-and $\lambda_L=0$ and $\lambda_U=2^{-1/p}$ for an SC copula with parameter $p>0$.
+(@nelsen2006_copula, section 5.4) that $\lambda_L=\lambda_U=0$ for the Normalcopula,
+$\lambda_L=2^{-1/p}$ and $\lambda_U=0$ for a Clayton copula with parameter $p>0$,
+and $\lambda_L=0$ and $\lambda_U=2^{-1/p}$ for an survival Clayton  copula with parameter $p>0$.
 In the section on methods for Q1 we will introduce
@@
 \includegraphics[width=\textwidth]{./Results/PedagogSuppMat_Normal.pdf}
-\caption[Example N copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example N copulas. 
-S is Spearman correlation; $p$ is the value of the parameter for the N 
+\caption[Example Normalcopulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Normalcopulas. 
+S is Spearman correlation; $p$ is the value of the parameter for the Normal
 family (it is a 
@@
 \includegraphics[width=\textwidth]{./Results/PedagogSuppMat_Clayton.pdf}
-\caption[Example C copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example C 
-copulas. S is Spearman correlation; $p$ is the value of the parameter for the C family (it 
+\caption[Example Clayton  copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Clayton  
+copulas. S is Spearman correlation; $p$ is the value of the parameter for the Clayton  family (it 
 is a one-parameter family); and LT and UT are the measures of lower- and 
@@
 Information Criteria (AIC and BIC). One of the fitted copulas was the N
-copula, making possible comparisons of the degree to which the N copula versus
+copula, making possible comparisons of the degree to which the Normalcopula versus
 other copula families were good descriptions of data.
@@
 using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
-is Kendall's tau for the data \cite{Genest2007}. We used the 
+is Kendall's tau for the data [@Genest2007]. We used the 
 implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
@@
 compared AIC weights and tail-dependence of 
-minimum-AIC copulas with that of an N copula. We also counted the number of location pairs 
+minimum-AIC copulas with that of an Normalcopula. We also counted the number of location pairs 
 for which model-averaged lower-tail dependence was 
@@
 We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}). Code and data for this
-project is archived at \url{www.github.com/sghosh89/BIVAN}.
+project are archived at \url{www.github.com/sghosh89/BIVAN}.
 
@@
 This will not affect BB1, BB6, etc., but it will affect SBB1, SBB6, etc., which will now be called survival BB1, survival BB6, etc.
+
+>>>> Done. I don't know whether it should be called as a Normal copula or an Normal copula? Now, we use "an".
+
 2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc). Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
 3) Sort the rows of Table S3 (and all analogous tables, e.g., S5, S7, etc.) by AIC, putting the lowest AIC at the top. This should mean the highest AICw is also at the top.
+
+>>>> Added, sorting based on min AIC.
+
 3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text. Same for Fig. S21 and Table S4.
@@
 8) Put the cor results only, for Spearman only (or is Kendall more consistent with what we do elsewhere - what do you think?) from tables S4, S6, S8, S10 into Table 2 as a new row.
+
+>>>> All done.
 -->
@@
 
-Across essentially all our datasets, copula structures were typically non-N and 
+Across essentially all our datasets, copula structures were typically non-Normaland 
 commonly showed asymmetric tail 
@@
 and computed AIC and BIC weights and corresponding
-lower- and upper-tail statistics for the fitted distributions (Table \ref{SM-tab_soilCNfit}).
+lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
 `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the 
 minimum-AIC copula, with AIC weight `r round(max(BivMS_res_Loecke$InfCritRes$AICw),2)`, whereas the 
-N copula had AIC weight essentially 0, answering the 
+Normal copula had AIC weight essentially 0, answering the 
 first part of Q1 for these data. 
@@
 and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
-These values are distinct from what an N copula would give, namely, 0, answering the second part 
+These values are distinct from what an Normal copula would give, namely, 0, answering the second part 
 Q1 for these data. The values also differed substantially 
@@
 
-Although these results convincingly show non-N copula structure, 
+Although these results convincingly show non-Normal copula structure, 
 we caution that the model-averaged tail-dependence statistics may have been biased 
@@
 available from `BiCopGofTest`). However, nonparametric statistics verified
-both that tail dependence was distinct from that of an N copula, and that it 
+both that tail dependence was distinct from that of an Normal copula, and that it 
 was asymmetric (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}, 
-Fig. \ref{SM-fig_soilCN_nonparam}, Table \ref{SM-tab_soilCN_asym}). Our 
+Fig. \ref{fig_soilCN_nonparam}, Table \ref{tab_soilCN_asym}). Our 
 affirmative, empirically based answer to Q1 for the soil C and N data is 
@@
 variables were significantly related for all datasets 
-(Table \ref{tab_bivar_summary}, row 1). A non-N copula 
+(Table \ref{tab_bivar_summary}, row 1). A non-Normal copula 
 always emerged as the lowest-AIC (highest-AIC-weight) copula 
 (Table \ref{tab_bivar_summary}, row 2). 
-The N copula was always a poor fit compared to the highest-AIC-weight copula 
+The Normal copula was always a poor fit compared to the highest-AIC-weight copula 
 (Table \ref{tab_bivar_summary}, rows 3-4), answering the first
@@
 than lower-tail dependence, and the last dataset had the reverse, though weakly. 
-Although these results convincingly show non-N copula structure, the 
+Although these results convincingly show non-Normal copula structure, the 
 model-averaged tail-dependence statistics may have been biased 
@@
 than upper-tail dependence. Detailed results for each dataset are in 
-Tables \ref{SM-tab_soilCNfit} - \ref{SM-tab_cc_asym_2000} and 
-Figs. \ref{SM-fig_soilCN_nonparam} - \ref{SM-fig_cc2000_nonparam}.
+Tables \ref{tab_soilCNfit} - \ref{tab_soilCN_asym},  \ref{SM-tab_bmr_birds_fit} - \ref{SM-tab_3biv_npa_summary} and 
+Figs. \ref{fig_soilCN_nonparam},  \ref{SM-fig_bmr_birds_nonparam} - \ref{SM-fig_cc2000_nonparam}.
 
@@
 positive and were frequently large, thereby differing 
-from what an N copula would give (i.e., 0), and answering the second
+from what an Normal copula would give (i.e., 0), and answering the second
 part of Q1: these data have greater tail dependence 
@@
 multivariate data. Results supported the 
-conclusions that non-N copula structure, non-N tail dependence, and asymmetric tail 
+conclusions that non-Normal copula structure, non-Normal tail dependence, and asymmetric tail 
 dependence were common, answering Q1 in the affirmative for these data. 
@@
 (Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
-pairs had best-fitting copulas that were not the N copula (Table \ref{tab_multivar_summary}, 
+pairs had best-fitting copulas that were not the Normal copula (Table \ref{tab_multivar_summary}, 
 row 3), and AIC weights of best-fitting copulas were 
-better than those of N copulas (Table \ref{tab_multivar_summary}, rows 4-5).
+better than those of Normal copulas (Table \ref{tab_multivar_summary}, rows 4-5).
 Best-fitting copulas were nearly always considered an adequate fit 
@@
 \textit{Ceratium furca} (results not shown); results supported the conclusion that 
-non-N copula structure and tail dependence, and asymmetry of tail dependence are 
+non-Normal copula structure and tail dependence, and asymmetry of tail dependence are 
 common.
+
+<!--bivariate results for soil CN data-->
+
+<!--Display model selection results for soilCN data in a table-->
+```{r tab_soilCNfit, echo=F, results="asis"}
+#knitr::opts_knit$set(kable.force.latex = TRUE)
+BivMS_res_Loecke<-readRDS("./Results/fitting_results/BivMS_soilCN.RDS")
+h<-BivMS_res_Loecke$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
+colnames(h)<-c("copula","AIC","AICw","BIC","BICw","LT","UT")
+h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC
+knitr::kable(h_ord, #longtable = T, 
+             format = "latex",linesep = "",row.names = F, digits = c(4,2,2,2,2,4,4),
+      caption.short = "Model fitting results for soil C and N dataset",
+      caption = "Results of fitting our 16 copula families to the soil C and N dataset.
+Copula name abbreviations are given in the main text.
+AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.
+Wights and tail dependences are rounded upto 2 and 4 digits, respectively.
+\\label{tab_soilCNfit}",
+      booktabs = TRUE)
+```
+
+<!--Displaying a table with npa summary results for soilCN data-->
+```{r tab_soilCN_asym,echo=F,results="asis"}
+tab_biv_npa_summary<-readRDS("./Results/stat_results/tab_biv_npa_summary.RDS")
+tab_soilCN_asym<-tab_biv_npa_summary[1:3,2:4]
+knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
+             format = "latex",linesep = "",
+      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
+      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
+The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving Normal surrogates, in seperate comparisons.
+A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
+```
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
+\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for soil C and N data.
+As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving Normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
+Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
+For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates.
+When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level).
+When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.
+\label{fig_soilCN_nonparam}}
+\end{center}
+\end{figure}
 
@@
 tab_bivar_summary<-readRDS("./Results/tab_bivar_summary.RDS")
+tab_biv_npa_summary<-readRDS("./Results/stat_results/tab_biv_npa_summary.RDS")
+
+# adding npa stat summary res: spearman cor. preserving corl-coru for 4 bivariate data
+my_mat<- matrix(tab_biv_npa_summary[c(1,4,7,10),4],nrow=1,ncol=4)
+colnames(my_mat)<-colnames(tab_bivar_summary) # this line is needed to apply rbind later: colnames should be matched
+
+tab_bivar_summary<-rbind(tab_bivar_summary,my_mat)
 row.names(tab_bivar_summary)<-NULL
+
 library(kableExtra)
@@
            "3. AIC weight of minimum-AIC copula",
-           "4. AIC weight of N copula",
+           "4. AIC weight of Normal copula",
            "5. p-value, Cramer-von Mises goodness of fit test",
@@
            "9. Model averaged lower- minus upper-tail dependence",
-           "10. Asymmetry rank in surrogates")
+           "10. Asymmetry rank in surrogates", # this row will be commented out, that's why the numbering starts from 10 in the next line
+           "10. How many times Non-parametric tail asymmetry ($Cor_{FS}-Cor_{LS}$) on given data are less (or greater) than the Spearman-preserving Normal surrogates")
 )
 tab_bivar_summary<-cbind(dt,tab_bivar_summary)
-knitr::kable(tab_bivar_summary[1:9,], 
-             format="latex",
+tab_bivar_summary<-tab_bivar_summary[c(1:9,11),]
+
+knitr::kable(tab_bivar_summary, 
+             format="latex", escape=F,row.names=F,
              caption = "Summary of Q1 results for bivariate datasets.
-The $p$-values in rows 5-6 are for the minimum-AIC copula.
+The $p$-values (rows 5-6) are for the minimum-AIC copula.
 Model averaging used for rows 7-9 was based on AIC weights.
-Cedar Creek results are for the year 2000.\\label{tab_bivar_summary}",
+First 3 bivariate data use the default range for first slice (FS) is (0,0.1) and for last slice (LS) is (0.9, 1), whereas Cedar Creek result uses range (0, 0.25) for FS and (0.75, 1) for LS for the year 2000.\\label{tab_bivar_summary}",
              booktabs=T, linesep = "\\addlinespace",
@@
            "2. Number of non-independent pairs",
-           "3. Number of pairs with non-N copula as best fit",
+           "3. Number of pairs with non-Normal copula as best fit",
            "4. Average AIC-weights of best-fitting copulas across location pairs",
-           "5. Average AIC-weights of N copula across pairs",
+           "5. Average AIC-weights of Normal copula across pairs",
            "6. Number of well-fitting location pairs",
@@
 
-Having demonstrated that non-N copula structure and asymmetric tail dependence are 
+Having demonstrated that non-Normal copula structure and asymmetric tail dependence are 
 common in ecological variables, we 
@@
 outputs to discover if the model could
-produce non-N copula structure and asymmetric tail dependence in the 
+produce non-Normal copula structure and asymmetric tail dependence in the 
 $E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
-and $(\delta_1,\delta_2)$ have N copula structure, the Moran 
+and $(\delta_1,\delta_2)$ have Normal copula structure, the Moran 
 mechanism analyzed below 
@@
 of tail dependence, to an extent that depended on the value of $b$;
-thus this is a mechanism that can produce non-N copula structure and
+thus this is a mechanism that can produce non-Normal copula structure and
 asymmetric tail dependence across space in ecological variables.
@@
 Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
-ecological outputs had non-N copula structure, but did not always reflect 
+ecological outputs had non-Normal copula structure, but did not always reflect 
 asymmetry of tail dependence, perhaps because those methods measured
@@
 \caption{If environmental effects operate asymmetrically in their tails on
-ecological variables, it can result in non-N copula structure and asymmetric 
+ecological variables, it can result in non-Normal copula structure and asymmetric 
 tail dependence across space in the ecological variables. Shown are the last
@@
 dynamics were strongly nonlinear.
-Thus this is a mechanism that can produce non-N copula structure and
+Thus this is a mechanism that can produce non-Normal copula structure and
 asymmetric tail dependence across space in ecological variables.
@@
 had opposite signs for noise and model outputs 
-(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). For instance, using a C copula with a large Kendall $\tau$,
+(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). For instance, using a Clayton  copula with a large Kendall $\tau$,
 $\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
 population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}); 
-and likewise for SG and SJ copulas (Figs \ref{SM-fig_Cause4copula_SG_kend_sRicker_nsd_1}, 
+and likewise for survival Gumbel and survival Joe copulas (Figs \ref{SM-fig_Cause4copula_SG_kend_sRicker_nsd_1}, 
 \ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). Overall,
@@
 for details). 
-Thus this is a mechanism that may produce non-N copula structure and
+Thus this is a mechanism that may produce non-Normal copula structure and
 asymmetric tail dependence between characters.
@@
 structure of the data were multivariate N, but the data were otherwise statistically
-unchanged. Significant differences indicate that non-N copula structure in the data 
+unchanged. Significant differences indicate that non-Normal copula structure in the data 
 contributed to the skewness of the spatial mean time series, 
@@
 <!--what we did and why is it important to study?-->
-We showed that non-N copula structure and asymmetric tail dependence are 
+We showed that non-Normal copula structure and asymmetric tail dependence are 
 common across multiple sub-disciplines
@@
 taken this structure into account. We hypothesized 
-mechanisms that may cause non-N copula structures and tail dependence;
+mechanisms that may cause non-Normal copula structures and tail dependence;
 we discuss below how commonly some of our
@@
 We also hypothesized and demonstrated important consequences of 
-non-N copula structure and tail-dependence for the field of ecology. 
+non-Normal copula structure and tail-dependence for the field of ecology. 
 For instance, the skewness
@@
 that build upon existing copula methods. 
-To our knowledge, only N copulas have 
+To our knowledge, only Normal copulas have 
 been used in ecology [@Valpine2014; @Anderson2018; @Popovic2018].
@@
 opperates commonly. It may be a dominant cause of 
-tail depencence and non-N copula struture of ecological dependencies 
+tail depencence and non-Normal copula struture of ecological dependencies 
 across space.
@@
 operates commonly: Moran effects are common [@Liebhold2004], and 
-non-N and tail-dependent copula structures are often found in 
+non-Normal and tail-dependent copula structures are often found in 
 environmental variables. If intense meteorological events are 
@@
 @Serinaldi2008 examined the spatial dependence 
-of rainfall in Central Italy; G or Student 2-copulas 
+of rainfall in Central Italy; Gumbel or Student 2-copulas 
 were candidates for modelling dependence, and 
-neither of these is an N copula.
+neither of these is an Normal copula.
 A long-term study (1950–2014) in the Loess Plateau of China [@She2018]
-showed that a G copula effectively modeled 
+showed that a Gumbel copula effectively modeled 
 the spatial dependence of drought variables.
@@
 between species could yield tail dependence between abundance data for the 
-species. This is another theoretical mechanism for non-N copula
+species. This is another theoretical mechanism for non-Normal copula
 structure and tail dependence, represented in Fig. \ref{MasterFigure} 
@@
 Our hypotheses and results cover the presence, causes, and consequences
-of non-N copula structure and tail dependence in ecological 
+of non-Normal copula structure and tail dependence in ecological 
 datasets (Fig. \ref{MasterFigure}).

### 2019-09-20 SupportingInformation.Rmd
edits done as suggested by DCR: see my response

@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvClayton.pdf}
-\caption[Example SC copulas]{Log-transformed pdfs (A-E) and samples (F-J) 
-from example SC 
-copulas. S is Spearman correlation; p is the value of the parameter for the SC 
+\caption[Example survival Clayton copulas]{Log-transformed pdfs (A-E) and samples (F-J) 
+from example survival Clayton 
+copulas. S is Spearman correlation; p is the value of the parameter for the survival Clayton 
 family (it is a one-parameter family); and LT and UT are the measures of lower- and 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
-\caption[Example G copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example G copulas. S is 
-Spearman correlation; $p$ is the value of the parameter for the G family (it is a 
+\caption[Example Gumbel copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Gumbel copulas. S is 
+Spearman correlation; $p$ is the value of the parameter for the Gumbel family (it is a 
 one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvGumbel.pdf}
-\caption[Example SG copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SG 
-copulas. S is Spearman correlation; $p$ is the value of the parameter for the SG 
+\caption[Example survival Gumbel copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example survival Gumbel 
+copulas. S is Spearman correlation; $p$ is the value of the parameter for the survival Gumbel 
 family (it is a one-parameter family); and LT and UT are the measures of lower- and 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SurvJoe.pdf}
-\caption[Example SJ copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example SJ 
-copulas. K is Spearman correlation; $p$ is the value of the parameter for the SJ 
+\caption[Example survival Joe  copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example survival Joe  
+copulas. K is Spearman correlation; $p$ is the value of the parameter for the survival Joe  
 family (it is a one-parameter family); and LT and UT are the measures of lower- and 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
-\caption[Example F copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example F copulas. S is Kendall correlation; $p$ is the value of the parameter for the F family (it is a 
+\caption[Example Frank copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Frank copulas. S is Kendall correlation; $p$ is the value of the parameter for the Frank family (it is a 
 one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB1.pdf}
-\caption[Example SBB1 copulas]{Log-transformed pdfs for example SBB1 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
+\caption[Example survival BB1 copulas]{Log-transformed pdfs for example survival BB1 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
 denote the two parameters of the family (it is a 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB6.pdf}
-\caption[Example SBB6 copulas]{Log-transformed pdfs for example SBB6 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
+\caption[Example survival BB6 copulas]{Log-transformed pdfs for example survival BB6 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
 denote the two parameters of the family (it is a 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB7.pdf}
-\caption[Example SBB7 copulas]{Log-transformed pdfs for example SBB7 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
+\caption[Example survival BB7 copulas]{Log-transformed pdfs for example survival BB7 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
 denote the two parameters of the family (it is a 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_SBB8.pdf}
-\caption[Example SBB8 copulas]{Log-transformed pdfs for example SBB8 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
+\caption[Example survival BB8 copulas]{Log-transformed pdfs for example survival BB8 copulas. K is Kendall correlation; $p_1$ and $p_2$ 
 denote the two parameters of the family (it is a 
@@
 
-N and F copula families 
+Normal and Frank copula families 
 are known to have symmetric tail dependence,
@@
 in these families [@nelsen2006_copula]. C copulas are known to have stronger lower- than 
-upper-tail dependence, and SC copulas, being rotations of C copulas by 
+upper-tail dependence, and survival Clayton copulas, being rotations of C copulas by 
 $180^\circ$, have stronger upper- than lower tail dependence [@nelsen2006_copula].
-Using the `iRho` function from the `copula` package, for each of the four families N, 
-F, C, and SC, we obtained parameters for which the expected value of 
+Using the `iRho` function from the `copula` package, for each of the four families Normal, 
+Frank, Clayton, and survival Clayton, we obtained parameters for which the expected value of 
 Spearman's $\rho$ was, separately, $0,0.1,0.2,...,0.9$. For each copula family and for each 
@@
 the expectation being that the difference statistics would be positive for 
-C copulas, negative for SC copulas (except for the case of $\rho=0$, a 
+C copulas, negative for survival Clayton copulas (except for the case of $\rho=0$, a 
 boundary case of these families corresponding to independence), and not significantly different 
-from $0$ for F and N 
+from $0$ for Frank and Normal 
 copulas.
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
-\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas.
+\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the Normal and Frank copulas.
 See Appendix  \ref{Test_npa_stat} for details.
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
-\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the C and SC copulas, and the known symmetry of tail dependence of the N and F copulas.
+\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the Normal and Frank copulas.
 See Appendix  \ref{Test_npa_stat} for details.
@@
 
-<!--Display model selection results for soilCN data in a table-->
-```{r tab_soilCNfit, echo=F, results="markup"}
-#knitr::opts_knit$set(kable.force.latex = TRUE)
-h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-colnames(h)<-c("copname","AICw","BICw","LT","UT")
-knitr::kable(h, #longtable = T, 
-             format = "latex",linesep = "",
-      caption.short = "Model fitting results for soil C and N dataset",
-      caption = "Results of fitting our 16 copula families to the soil C and N dataset.
-Copula name abbreviations are given in the main text.
-AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.\\label{tab_soilCNfit}",
-      booktabs = TRUE)
-```
-
 <!--This chunk is for npa testing results of soilCN dataset-->
@@
 stat_soilCN_ga<-bivfunctionplot(v=v_CN,resloc="./Results/stat_results/stat_soilCN/",nametag="soilCN",numbin=10)
-```
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for soil C and N data.
-As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
-Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
-For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates.
-When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level).
-When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.
-\label{fig_soilCN_nonparam}}
-\end{center}
-\end{figure}
-
-<!--Function to generate a table for npa results-summary with bivariate data, this one about asymmetry of tail dependence-->
-```{r make_table_stat_fn, echo=F}
-
-fractxt<-function(fracval,numsurrog){
-  ptxt<-''
-  if (fracval>.5*numsurrog){                        
-    ptxt<-paste(">",fracval,sep='')
-  } 
-  if (fracval<=.5*numsurrog){
-    ptxt<-paste("<",numsurrog-fracval,sep='')
-  }
-  return(ptxt)
-}
-
-make_table_stat_fn<-function(data){
-
- #atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
- #                               "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
- #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
- atdres<-data.frame(Statistic=c("cor$_{FS}$-cor$_{LS}$",
-                               "P$_{FS}$-P$_{LS}$",
-                                "D$_{LS}^2$-D$_{FS}^2$"),
-                   Kendall=rep("",3),
-                   Spearman=rep("",3),stringsAsFactors = F)
-
- atdres[1,2]<-fractxt(data$corlmcoru_frac_K,numsurrog = data$numsurrog_success_K)
- atdres[1,3]<-fractxt(data$corlmcoru_frac_S,numsurrog = data$numsurrog_success_S)
- atdres[2,2]<-fractxt(data$PlmPu_frac_K,numsurrog = data$numsurrog_success_K)
- atdres[2,3]<-fractxt(data$PlmPu_frac_S,numsurrog = data$numsurrog_success_S)
- atdres[3,2]<-fractxt(data$D2umD2l_frac_K,numsurrog = data$numsurrog_success_K)
- atdres[3,3]<-fractxt(data$D2umD2l_frac_S,numsurrog = data$numsurrog_success_S)
-
- return(atdres)
-
-}
-```
-
-<!--Displaying a table with npa summary results for soilCN data-->
-```{r tab_soilCN_asym,echo=F,results="markup"}
-#knitr::opts_knit$set(kable.force.latex = TRUE)
-tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
-#tab_soilCN_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
-knitr::kable(tab_soilCN_asym, booktabs = T, escape=F,
-             format = "latex",linesep = "",
-      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
-      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
-The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons.
-A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
 ```
@@
 ```{r tab_bmr_birds_fit, echo=F, results='markup'}
-h<-BivMS_birds$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-colnames(h)<-c("copname","AICw","BICw","LT","UT")
-knitr::kable(h,
-             format='latex',linesep = "",
+h<-BivMS_birds$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
+colnames(h)<-c("copula","AIC","AICw","BIC","BICw","LT","UT")
+h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC
+knitr::kable(h_ord,
+             format='latex',linesep = "",row.names = F, digits = c(4,2,2,4,4),
              caption.short = "Model fitting results for the bird body mass and BMR dataset",
@@
 Analogous to Fig.
-\\ref{tab_soilCNfit}, see caption of that figure for details.
+\\ref{MT-tab_soilCNfit}, see caption of that figure for details.
 \\label{tab_bmr_birds_fit}",
@@
 Same format as Fig.
-\ref{fig_soilCN_nonparam}, see caption of that figure.
+\ref{MT-fig_soilCN_nonparam}, see caption of that figure.
 \label{fig_bmr_birds_nonparam}}
@@
 \end{figure}
-
-<!--Displaying a table with npa summary results for Birds' BMR data-->
-```{r tab_bmr_birds_asym, echo=F, results="markup"}
-tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
-#tab_bmr_birds_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
-knitr::kable(tab_bmr_birds_asym,
-             format="latex",linesep = "",escape=F,
-             caption.short = "Nonparametric asymmetry tests against an N-copula null, bird body mass and BMR",
-             caption = "Bird body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
-Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.\\label{tab_bmr_birds_asym}",
-             booktabs=T)
-```
 
@@
 ```{r tab_bmr_mammals_fit, echo=F, results='markup'}
-h<-BivMS_mammals$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-colnames(h)<-c("copname","AICw","BICw","LT","UT")
-knitr::kable(h,
-             format='latex',linesep = "",
+h<-BivMS_mammals$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
+colnames(h)<-c("copula","AIC","AICw","BIC","BICw","LT","UT")
+h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC
+knitr::kable(h_ord,
+             format='latex',linesep = "",row.names = F, digits = c(4,2,2,4,4),
              caption.short = "Model fitting results for the mammal body mass and BMR dataset",
@@
 Analogous to Fig.
-\\ref{tab_soilCNfit}, see caption of that figure for details.
+\\ref{MT-tab_soilCNfit}, see caption of that figure for details.
 \\label{tab_bmr_mammals_fit}",
@@
 Same format as Fig.
-\ref{fig_soilCN_nonparam}, see caption of that figure.
+\ref{MT-fig_soilCN_nonparam}, see caption of that figure.
 \label{fig_bmr_mammals_nonparam}}
@@
 \end{figure}
-
-<!--Displaying a table with npa summary results for Mammals' BMR data-->
-```{r tab_bmr_mammals_asym, echo=F, results="markup"}
-tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
-#tab_bmr_mammals_asym[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
-knitr::kable(tab_bmr_mammals_asym,
-             format="latex",linesep = "",escape=F,
-             caption.short = "Nonparametric asymmetry tests against an N-copula null, mammal body mass and BMR",
-             caption = "Mammal body mass and BMR dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
-Same format as Table \\ref{tab_soilCN_asym}, see caption of that figure.
-\\label{tab_bmr_mammals_asym}",
-             booktabs=T)
-```
 
@@
 ```{r tab_cc2000fit, echo=F, results='markup'}
-h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
-colnames(h)<-c("copname","AICw","BICw","LT","UT")
-knitr::kable(h,
-             format='latex',linesep = "",
+h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
+colnames(h)<-c("copula","AIC","AICw","BIC","BICw","LT","UT")
+h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC
+knitr::kable(h_ord,
+             format='latex',linesep = "",row.names = F, digits = c(4,2,2,4,4),
              caption.short = "Model fitting results for the Cedar Creek dataset",
@@
 Analogous to Fig.
-\\ref{tab_soilCNfit}, see caption of that figure for details.
+\\ref{MT-tab_soilCNfit}, see caption of that figure for details.
 \\label{tab_cc2000fit}",
@@
 \caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data, year 2000.
-As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$.
+As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving Normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$.
 If the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
@@
 \end{figure}
-
-<!--Displaying a table with npa summary results for Cedar creek data-->
-```{r tab_cc_asym_2000, echo=F, results="markup"}
-tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
-#tab_cc_asym_2000[,1]<-c("corFS-corLS","PFS-PLS","D2LS-D2FS")
-knitr::kable(tab_cc_asym_2000,
-             format="latex",linesep = "",escape=F,
-             caption.short = "Nonparametric asymmetry tests against an N-copula null, Cedar Creek data",
-             caption = "Cedar Creek dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis, year 2000.
-The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.25}-\\cor_{0.75,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.25}-\\Ps_{0.75,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.75,1}-\\Dsq_{0,0.25}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving N surrogates, in seperate comparisons.
-A table entry $>X$ indicates the value of the given statistic on the data was higher than its value on $X$ of the surrogates, so entries of the form $>X$ for $X$ equal to $975$ or above would indicate that lower-tail dependence was significantly stronger than upper-tail dependence.\\label{tab_cc_asym_2000}",
-             booktabs=T)
-```
 
@@
 saveRDS(tab_bivar_summary,"./Results/tab_bivar_summary.RDS")
+```
+
+<!--Function to generate a table for npa results-summary with bivariate data, this one about asymmetry of tail dependence-->
+```{r make_table_stat_fn, echo=F}
+fractxt<-function(fracval,numsurrog){
+  ptxt<-''
+  if (fracval>.5*numsurrog){                        
+    ptxt<-paste(">",fracval,sep='')
+  } 
+  if (fracval<=.5*numsurrog){
+    ptxt<-paste("<",numsurrog-fracval,sep='')
+  }
+  return(ptxt)
+}
+
+make_table_stat_fn<-function(data){
+
+  #atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
+  #                               "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
+  #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
+  atdres<-data.frame(Statistic=c("cor$_{FS}$-cor$_{LS}$",
+                                 "P$_{FS}$-P$_{LS}$",
+                                 "D$_{LS}^2$-D$_{FS}^2$"),
+                     Kendall=rep("",3),
+                     Spearman=rep("",3),stringsAsFactors = F)
+
+  atdres[1,2]<-fractxt(data$corlmcoru_frac_K,numsurrog = data$numsurrog_success_K)
+  atdres[1,3]<-fractxt(data$corlmcoru_frac_S,numsurrog = data$numsurrog_success_S)
+  atdres[2,2]<-fractxt(data$PlmPu_frac_K,numsurrog = data$numsurrog_success_K)
+  atdres[2,3]<-fractxt(data$PlmPu_frac_S,numsurrog = data$numsurrog_success_S)
+  atdres[3,2]<-fractxt(data$D2umD2l_frac_K,numsurrog = data$numsurrog_success_K)
+  atdres[3,3]<-fractxt(data$D2umD2l_frac_S,numsurrog = data$numsurrog_success_S)
+
+  return(atdres)
+
+}
+```
+
+<!--Displaying a table with npa summary results for soilCN data-->
+```{r tab_biv_npa_summary,echo=F,results="asis"}
+library(kableExtra)
+library(dplyr)
+library(tibble)
+
+datatype <- tibble(Data = c("Soil CN","","","Bird masses and BMR","","","Mammal masses and BMR","","","Cedar Creek","",""))
+
+tab_soilCN_asym<-make_table_stat_fn(data=stat_soilCN_ga)
+tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
+tab_bmr_mammals_asym<-make_table_stat_fn(data=stat_mammals_ga)
+tab_cc_asym_2000<-make_table_stat_fn(data=stat_cc_ga$`2000`)
+tab_biv_npa_summary<-rbind(tab_soilCN_asym,tab_bmr_birds_asym, tab_bmr_mammals_asym, tab_cc_asym_2000)
+
+tab_biv_npa_summary<-cbind(datatype,tab_biv_npa_summary)
+saveRDS(tab_biv_npa_summary,"./Results/stat_results/tab_biv_npa_summary.RDS")
+
+tab_3biv_npa_summary<-tab_biv_npa_summary[c(4:12),]
+
+knitr::kable(tab_3biv_npa_summary, booktabs = T, escape=F,
+             format = "latex",linesep = "", row.names=F,
+      caption.short = "Nonparametric asymmetry tests against an Normal-copula null model with birds, mammals and Cedar Creek data",
+      caption = "Nonparametric asymmetry tests against an Normal-copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence).
+For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for $(0,0.25)$ and $(0.75,1)$ range of lower and upper bounds.
+Though the non-parametric result does not indicate significantly stronger lower tail dependence for Cedar Creek data, the
+greater than (>) sign before the numbers do show the opposite direction in tail-dependence compared to other bivariate datasets.\\label{tab_3biv_npa_summary}")
 ```
@@
 C and N data
-were compared to distributions of their values on 1000 Kendall- or Spearman-preserving N surrogates
+were compared to distributions of their values on 1000 Kendall- or Spearman-preserving Normal surrogates
 of the data,
 separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
-(Fig. \ref{fig_soilCN_nonparam}). Results show that tail dependence of data is stronger in 
+(Fig. \ref{MT-fig_soilCN_nonparam}). Results show that tail dependence of data is stronger in 
 both the lower and upper tails than would be expected under a null hypothesis of a normal copula.
@@
 $\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
-on 1000 Kendall- or Spearman-preserving N surrogates, separately in two 
-comparisons (Table \ref{tab_soilCN_asym}). Results showed that
+on 1000 Kendall- or Spearman-preserving Normal surrogates, separately in two 
+comparisons (Table \ref{MT-tab_soilCN_asym}). Results showed that
 upper-tail dependence was significantly stronger than lower-tail dependence.
@@
 the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
-for 1000 Kendall-preserving surrogate datasets with N copulas. Due to relatively smaller sample size,
+for 1000 Kendall-preserving surrogate datasets with Normal copulas. Due to relatively smaller sample size,
 for Cedar Creek data we used $P_{0,0.25}-P_{0.75,1}$.
@@
              format='latex',linesep = "",digits=3,
-             caption.short = "AIC weights, N copulas, green spruce aphid abundance data",
-             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for green spruce aphid abundance data, for comparison with Table \\ref{tab_count_AICw}.
+             caption.short = "AIC weights, Normal copulas, green spruce aphid abundance data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the Normal copula family for each location pair for green spruce aphid abundance data, for comparison with Table \\ref{tab_count_AICw}.
 \\label{tab_count_AICw_normal}",
@@
              format='latex',linesep = "",digits=3,
-             caption.short="AIC weights, N copulas, leaf-curling plum aphid first flight data",
+             caption.short="AIC weights, Normal copulas, leaf-curling plum aphid first flight data",
              caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for leaf-curlung plum aphid first flight data, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
@@
              digits=3,
-             caption.short="AIC weights, N copulas, \\textit{Ceratium furca} abundance data",
-             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for \\textit{Ceratium furca} abundance data, for comparison with Table \\ref{tab_pns_AICw}.\\label{tab_pns_AICw_normal}",
+             caption.short="AIC weights, Normal copulas, \\textit{Ceratium furca} abundance data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the Normal copula family for each location pair for \\textit{Ceratium furca} abundance data, for comparison with Table \\ref{tab_pns_AICw}.\\label{tab_pns_AICw_normal}",
              booktabs=TRUE)
@@
              digits=3,
-             caption.short="AIC weights, N copulas, methane-flux data",
-             caption = "$\\text{AIC}_{\\text{w}}$ for the N copula family for each location pair for methane-flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
+             caption.short="AIC weights, Normal copulas, methane-flux data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the Normal copula family for each location pair for methane-flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
              booktabs=TRUE)
@@
 If left-tail dependent noise was desired, we took the negative of the
-entire matrix. The moderately left-tail dependent matrix was based on a C copula 
+entire matrix. The moderately left-tail dependent matrix was based on a Clayton copula 
 with parameter chosen (using the `iRho` function in the `copula` package) so that
@@
 datasets. The moderately right-tail dependent matrix and the symmetric tail dependence
-matrix were constructed similarly, but using an SC copula and an N copula,
+matrix were constructed similarly, but using an survival Clayton copula and an Normal copula,
 respectively.
@@
 methods section for Q1. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
-N-copula AICw: the AIC weight of the N copula family; Model-avg. LT: the lower-tail 
+N-copula AICw: the AIC weight of the Normal copula family; Model-avg. LT: the lower-tail 
 statistic, weighted-averaged across models using AICw; Model-avg. UT: same, but for the 
@@
 \caption[Comparison of copula structure of noise and model output, AR(1) model,
-C copula.]{Comparison of copula structure of noise inputs and model 
-outputs, AR(1) model, C copula. Red and blue points give 
+Clayton copula.]{Comparison of copula structure of noise inputs and model 
+outputs, AR(1) model, Clayton copula. Red and blue points give 
 means over 50 replicate simulations of the statistics listed on the left axes, 
@@
 \caption[Comparison of copula structure of noise and model output, AR(1) model, 
-SC copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
-model, SC copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
+survival Clayton copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+model, survival Clayton copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
 \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
@@
 \caption[Comparison of copula structure of noise and model output, AR(1) model, 
-G copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
-model, G copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
+Gumbelcopula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+model, Gumbelcopula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
 \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
@@
 \caption[Comparison of copula structure of noise and model output, AR(1) model, 
-SG copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
-model, SG copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
+survival Gumbel copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+model, survival Gumbel copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
 \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
@@
 \caption[Comparison of copula structure of noise and model output, AR(1) model, 
-SJ copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
-model, SJ copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
+survival Joe  copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+model, survival Joe  copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
 \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, weak noise, C copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, weak noise, C copula. Same format as Figs 
+model, weak noise, Clayton copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, Clayton copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, C copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, C copula. Same format as Figs 
+model, strong noise, Clayton copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, Clayton copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, weak noise, SC copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, weak noise, SC copula. Same format as Figs 
+model, weak noise, survival Clayton copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, survival Clayton copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, SC copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, SC copula. Same format as Figs 
+model, strong noise, survival Clayton copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, survival Clayton copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, weak noise, G copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, weak noise, G copula. Same format as Figs 
+model, weak noise, Gumbel copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, Gumbel copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, G copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, G copula. Same format as Figs 
+model, strong noise, Gumbel copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, Gumbel copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, weak noise, SG copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, weak noise, SG copula. Same format as Figs 
+model, weak noise, survival Gumbel copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, survival Gumbel copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, SG copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, SG copula. Same format as Figs 
+model, strong noise, survival Gumbel copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, survival Gumbel copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, weak noise, SJ copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, weak noise, SJ copula. Same format as Figs 
+model, weak noise, survival Joe  copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, survival Joe  copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, SJ copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, SJ copula. Same format as Figs 
+model, strong noise, survival Joe  copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, survival Joe copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
@@
 
-# Effects of non-N copula structure on Taylor's law \label{Consequence_TL_simulation}
+# Effects of non-Normal copula structure on Taylor's law \label{Consequence_TL_simulation}
 
@@
 the N-copula population array in all ways except the new
-arrays were constructed to have C (respectively, F, SC) copula structure for 
+arrays were constructed to have Clayton (respectively, Frank, survival Clayton) copula structure for 
 the dependences between 
@@
 constructed as follows; the others were constructed analogously. First, we constructed an
-$n$-dimensional C copula with parameter chosen so that the 
+$n$-dimensional Clayton copula with parameter chosen so that the 
 Spearman correlations between each pair of the $n$ components were the same as those
-of the N distribution constructed above. This was done using the `rho` and `iRho` functions 
-in the `copula` package in R. Because the C copula is a one-paremeter family, this 
+of the Normal distribution constructed above. This was done using the `rho` and `iRho` functions 
+in the `copula` package in R. Because the Clayton copula is a one-paremeter family, this 
 step relies on the fact that all pairs of locations were 
@@
 surrogate/randomized dataset had the copula structure
-of the C copula just constructed. The randomization procedure worked 
+of the Clayton copula just constructed. The randomization procedure worked 
 by permuting time series of values in the N-copula array, separately for each 
@@
 across the values $x$. The distrubutions of all these statistics across 1000 replicate 
-simulations were compared for the C, N, F, and SC simulations to determine if copula structure 
+simulations were compared for the Clayton, Normal, Frank, and survival Clayton simulations to determine if copula structure 
 influences Taylor's law.
@@
 Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
-linearity and homoskedasticity, by non-N copula structure. For N copula structure 
+linearity and homoskedasticity, by non-Normal copula structure. For Normal copula structure 
 (i.e., for simulations in our N-copula population array), $p$-values for the linearty
@@
 errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
-But linearity or homoskedasticity were violated more frequently for non-N copula structure
+But linearity or homoskedasticity were violated more frequently for non-Normal copula structure
 (Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
@@
 To illustrate the potential for missing data to influence perceived
-copula structure, we first generated 5000 points from a bivariate N copula
+copula structure, we first generated 5000 points from a bivariate Normal copula
 with parameter `r ncopparm` (Fig. \ref{fig:missingdata}A), and then 

### 2019-09-21 Paper.Rmd
edits missed before

@@
 were examined (in the sixth and seventh sections of the text). Lower boxes 
-correspond to potential consequences we examined of non-Normalcopula structure for
+correspond to potential consequences we examined of non-normal copula structure for
 ecological understanding and for applications (in the eighth and ninth
@@
 (Appendix \ref{SM-ContStrictMonoton}). Application of the 
-distributions functions of the component random variables translates
+distribution functions of the component random variables translates
 between the uniform-marginal context, which isolates the 
@@
 because we will fit these families to data later. The bivariate normal
-copula family (henceforth abbreviated N), which consists of copulas of bivariate normal
+copula family, which consists of copulas of bivariate normal
 distributions, is a one-parameter family. The parameter, which we here call $p$,
@@
 of the related bivariate normal distribution, and controls the degree of dependence
-between the variables. An Normal copula with $p=0.7$ was already introduced (Fig.
+between the variables. A normal copula with $p=0.7$ was already introduced (Fig.
 \ref{PedagogFigCopThry}A). The formula for the
@@
 directly informative about copula properties as pdf diagrams; Fig. \ref{figped_N} has 
-example pdfs corresponding to bivariate Normalcopulas for various values of $p$.
+example pdfs corresponding to bivariate normal copulas for various values of $p$.
 Note that the pdfs are symmetric
@@
 below for a more precise defintion of tail associations). 
-The Clayton copula family (henceforth abbreviated C),
+The Clayton copula family,
 of which we also already pictured an example (Fig.
-\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to Normalcopulas, Clayton copulas
+\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
 have stronger left- than right-tail association. The formula is 
@@
 less direct intuition than do example pdfs (Fig. \ref{figped_C}). The survival
-Clayton family (here abbreviated SC) is the symmetric opposite of the Claytonfamily, showing
+Clayton family is the symmetric opposite of the Clayton family, showing
 stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
@@
 on the copula and not on the marginals combined with it. It can be shown 
-(@nelsen2006_copula, section 5.4) that $\lambda_L=\lambda_U=0$ for the Normalcopula,
+(@nelsen2006_copula, section 5.4) that $\lambda_L=\lambda_U=0$ for the normal copula,
 $\lambda_L=2^{-1/p}$ and $\lambda_U=0$ for a Clayton copula with parameter $p>0$,
-and $\lambda_L=0$ and $\lambda_U=2^{-1/p}$ for an survival Clayton  copula with parameter $p>0$.
+and $\lambda_L=0$ and $\lambda_U=2^{-1/p}$ for a survival Clayton  copula with parameter $p>0$.
 In the section on methods for Q1 we will introduce
@@
 \includegraphics[width=\textwidth]{./Results/PedagogSuppMat_Normal.pdf}
-\caption[Example Normalcopulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Normalcopulas. 
-S is Spearman correlation; $p$ is the value of the parameter for the Normal
+\caption[Example normal copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example normal copulas. 
+S is Spearman correlation; $p$ is the value of the parameter for the normal
 family (it is a 
@@
 our ecological datasets and fits were compared via the Akaike and Bayesian 
-Information Criteria (AIC and BIC). One of the fitted copulas was the N
-copula, making possible comparisons of the degree to which the Normalcopula versus
+Information Criteria (AIC and BIC). One of the fitted copulas was the normal
+copula, making possible comparisons of the degree to which the normal copula versus
 other copula families were good descriptions of data.
@@
 compared AIC weights and tail-dependence of 
-minimum-AIC copulas with that of an Normalcopula. We also counted the number of location pairs 
+minimum-AIC copulas with that of a normal copula. We also counted the number of location pairs 
 for which model-averaged lower-tail dependence was 
@@
 
-Across essentially all our datasets, copula structures were typically non-Normaland 
+Across essentially all our datasets, copula structures were typically non-normal and 
 commonly showed asymmetric tail 
@@
 minimum-AIC copula, with AIC weight `r round(max(BivMS_res_Loecke$InfCritRes$AICw),2)`, whereas the 
-Normal copula had AIC weight essentially 0, answering the 
+normal copula had AIC weight essentially 0, answering the 
 first part of Q1 for these data. 
@@
 and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
-These values are distinct from what an Normal copula would give, namely, 0, answering the second part 
+These values are distinct from what a normal copula would give, namely, 0, answering the second part 
 Q1 for these data. The values also differed substantially 
@@
 
-Although these results convincingly show non-Normal copula structure, 
+Although these results convincingly show non-normal copula structure, 
 we caution that the model-averaged tail-dependence statistics may have been biased 
@@
 available from `BiCopGofTest`). However, nonparametric statistics verified
-both that tail dependence was distinct from that of an Normal copula, and that it 
+both that tail dependence was distinct from that of a normal copula, and that it 
 was asymmetric (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}, 
@@
 variables were significantly related for all datasets 
-(Table \ref{tab_bivar_summary}, row 1). A non-Normal copula 
+(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
 always emerged as the lowest-AIC (highest-AIC-weight) copula 
 (Table \ref{tab_bivar_summary}, row 2). 
-The Normal copula was always a poor fit compared to the highest-AIC-weight copula 
+The normal copula was always a poor fit compared to the highest-AIC-weight copula 
 (Table \ref{tab_bivar_summary}, rows 3-4), answering the first
@@
 than lower-tail dependence, and the last dataset had the reverse, though weakly. 
-Although these results convincingly show non-Normal copula structure, the 
+Although these results convincingly show non-normal copula structure, the 
 model-averaged tail-dependence statistics may have been biased 
@@
 were typically substantially 
-larger than N-copula weights (Table \ref{SM-tab_count_AICw_normal}), answering the first part of Q1. Goodness of fit tests 
+larger than normal copula weights (Table \ref{SM-tab_count_AICw_normal}), answering the first part of Q1. Goodness of fit tests 
 (Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
@@
 positive and were frequently large, thereby differing 
-from what an Normal copula would give (i.e., 0), and answering the second
+from what a normal copula would give (i.e., 0), and answering the second
 part of Q1: these data have greater tail dependence 
@@
 multivariate data. Results supported the 
-conclusions that non-Normal copula structure, non-Normal tail dependence, and asymmetric tail 
+conclusions that non-normal copula structure, non-normal tail dependence, and asymmetric tail 
 dependence were common, answering Q1 in the affirmative for these data. 
@@
 (Table \ref{tab_multivar_summary}, row 2). The large majority of non-independent location 
-pairs had best-fitting copulas that were not the Normal copula (Table \ref{tab_multivar_summary}, 
+pairs had best-fitting copulas that were not the normal copula (Table \ref{tab_multivar_summary}, 
 row 3), and AIC weights of best-fitting copulas were 
-better than those of Normal copulas (Table \ref{tab_multivar_summary}, rows 4-5).
+better than those of normal copulas (Table \ref{tab_multivar_summary}, rows 4-5).
 Best-fitting copulas were nearly always considered an adequate fit 
@@
 \textit{Ceratium furca} (results not shown); results supported the conclusion that 
-non-Normal copula structure and tail dependence, and asymmetry of tail dependence are 
+non-normal copula structure and tail dependence, and asymmetry of tail dependence are 
 common.
@@
       caption.short = "Model fitting results for soil C and N dataset",
-      caption = "Results of fitting our 16 copula families to the soil C and N dataset.
-Copula name abbreviations are given in the main text.
+      caption = "Model fitting results for soil C and N dataset using 16 copula families: copula name abbreviations are given in the main text.
 AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.
@@
       caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
-      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to an N-copula null hypothesis.
-The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving Normal surrogates, in seperate comparisons.
+      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to a normal copula null hypothesis.
+The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving normal surrogates, in seperate comparisons.
 A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for soil C and N data.
-As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving Normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
+\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for soil C and N data.
+As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
 Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
@@
 When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level).
-When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.
+When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal copula null hypothesis.
 \label{fig_soilCN_nonparam}}
@@
            "3. AIC weight of minimum-AIC copula",
-           "4. AIC weight of Normal copula",
+           "4. AIC weight of normal copula",
            "5. p-value, Cramer-von Mises goodness of fit test",
@@
            "10. Asymmetry rank in surrogates", # this row will be commented out, that's why the numbering starts from 10 in the next line
-           "10. How many times Non-parametric tail asymmetry ($Cor_{FS}-Cor_{LS}$) on given data are less (or greater) than the Spearman-preserving Normal surrogates")
+           "10. How many times non-parametric tail asymmetry ($Cor_{FS}-Cor_{LS}$) on given data are less (or greater) than the Spearman-preserving normal surrogates")
 )
@@
            "2. Number of non-independent pairs",
-           "3. Number of pairs with non-Normal copula as best fit",
+           "3. Number of pairs with non-normal copula as best fit",
            "4. Average AIC-weights of best-fitting copulas across location pairs",
-           "5. Average AIC-weights of Normal copula across pairs",
+           "5. Average AIC-weights of normal copula across pairs",
            "6. Number of well-fitting location pairs",
@@
 
-Having demonstrated that non-Normal copula structure and asymmetric tail dependence are 
+Having demonstrated that non-normal copula structure and asymmetric tail dependence are 
 common in ecological variables, we 
@@
 outputs to discover if the model could
-produce non-Normal copula structure and asymmetric tail dependence in the 
+produce non-normal copula structure and asymmetric tail dependence in the 
 $E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
-and $(\delta_1,\delta_2)$ have Normal copula structure, the Moran 
+and $(\delta_1,\delta_2)$ have normal copula structure, the Moran 
 mechanism analyzed below 
@@
 of tail dependence, to an extent that depended on the value of $b$;
-thus this is a mechanism that can produce non-Normal copula structure and
+thus this is a mechanism that can produce non-normal copula structure and
 asymmetric tail dependence across space in ecological variables.
@@
 Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
-ecological outputs had non-Normal copula structure, but did not always reflect 
+ecological outputs had non-normal copula structure, but did not always reflect 
 asymmetry of tail dependence, perhaps because those methods measured
@@
 \caption{If environmental effects operate asymmetrically in their tails on
-ecological variables, it can result in non-Normal copula structure and asymmetric 
+ecological variables, it can result in non-normal copula structure and asymmetric 
 tail dependence across space in the ecological variables. Shown are the last
@@
 dynamics were strongly nonlinear.
-Thus this is a mechanism that can produce non-Normal copula structure and
+Thus this is a mechanism that can produce non-normal copula structure and
 asymmetric tail dependence across space in ecological variables.
@@
 for details). 
-Thus this is a mechanism that may produce non-Normal copula structure and
+Thus this is a mechanism that may produce non-normal copula structure and
 asymmetric tail dependence between characters.
@@
 structure of the data were multivariate N, but the data were otherwise statistically
-unchanged. Significant differences indicate that non-Normal copula structure in the data 
+unchanged. Significant differences indicate that non-normal copula structure in the data 
 contributed to the skewness of the spatial mean time series, 
@@
 <!--what we did and why is it important to study?-->
-We showed that non-Normal copula structure and asymmetric tail dependence are 
+We showed that non-normal copula structure and asymmetric tail dependence are 
 common across multiple sub-disciplines
@@
 taken this structure into account. We hypothesized 
-mechanisms that may cause non-Normal copula structures and tail dependence;
+mechanisms that may cause non-normal copula structures and tail dependence;
 we discuss below how commonly some of our
@@
 We also hypothesized and demonstrated important consequences of 
-non-Normal copula structure and tail-dependence for the field of ecology. 
+non-normal copula structure and tail-dependence for the field of ecology. 
 For instance, the skewness
@@
 that build upon existing copula methods. 
-To our knowledge, only Normal copulas have 
+To our knowledge, only normal copulas have 
 been used in ecology [@Valpine2014; @Anderson2018; @Popovic2018].
@@
 opperates commonly. It may be a dominant cause of 
-tail depencence and non-Normal copula struture of ecological dependencies 
+tail depencence and non-normal copula struture of ecological dependencies 
 across space.
@@
 operates commonly: Moran effects are common [@Liebhold2004], and 
-non-Normal and tail-dependent copula structures are often found in 
+non-normal and tail-dependent copula structures are often found in 
 environmental variables. If intense meteorological events are 
@@
 were candidates for modelling dependence, and 
-neither of these is an Normal copula.
+neither of these is a normal copula.
 A long-term study (1950–2014) in the Loess Plateau of China [@She2018]
@@
 between species could yield tail dependence between abundance data for the 
-species. This is another theoretical mechanism for non-Normal copula
+species. This is another theoretical mechanism for non-normal copula
 structure and tail dependence, represented in Fig. \ref{MasterFigure} 
@@
 Our hypotheses and results cover the presence, causes, and consequences
-of non-Normal copula structure and tail dependence in ecological 
+of non-normal copula structure and tail dependence in ecological 
 datasets (Fig. \ref{MasterFigure}).

### 2019-09-21 SupportingInformation.Rmd
edits missed before

@@
 but the Spearman correlations are unchanged (Figs \ref{MT-fig_cop_pedag1}A-B, \ref{fig_transf_pearson_spearman}).
-A scatterploy very simlar to Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
+A scatterplot very simlar to Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
 of both variables by the cdf of the standard normal distribution.
@@
 $G_A$ and $G_B$ be univariate distribution functions, so that 
-$D(G_A(x),G_B(y))$ is guaranteed by Sklar's theorem to be a a distribution function.
-We here show that if $G_A$ and $G_B$ and continuous and strictly monotonic, then
+$D(G_A(x),G_B(y))$ is guaranteed by Sklar's theorem to be a distribution function.
+We here show that if $G_A$ and $G_B$ are continuous and strictly monotonic, then
 $D(G_A(x),G_B(y))$ is the distribution function of the random variable 
@@
 i.e., strengths of dependence in the lower and upper tails are the same, for all copulas
-in these families [@nelsen2006_copula]. C copulas are known to have stronger lower- than 
-upper-tail dependence, and survival Clayton copulas, being rotations of C copulas by 
+in these families [@nelsen2006_copula]. Clayton copulas are known to have stronger lower- than 
+upper-tail dependence, and survival Clayton copulas, being rotations of Clayton copulas by 
 $180^\circ$, have stronger upper- than lower tail dependence [@nelsen2006_copula].
-Using the `iRho` function from the `copula` package, for each of the four families Normal, 
+Using the `iRho` function from the `copula` package, for each of the four families normal, 
 Frank, Clayton, and survival Clayton, we obtained parameters for which the expected value of 
@@
 the expectation being that the difference statistics would be positive for 
-C copulas, negative for survival Clayton copulas (except for the case of $\rho=0$, a 
+Clayton copulas, negative for survival Clayton copulas (except for the case of $\rho=0$, a 
 boundary case of these families corresponding to independence), and not significantly different 
-from $0$ for Frank and Normal 
+from $0$ for Frank and normal 
 copulas.
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt35_rankT.pdf}
-\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the Normal and Frank copulas.
+\caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the normal and Frank copulas.
 See Appendix  \ref{Test_npa_stat} for details.
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_testing/res_pt1000_rankT.pdf}
-\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the Normal and Frank copulas.
+\caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the normal and Frank copulas.
 See Appendix  \ref{Test_npa_stat} for details.
@@
 \includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for bird body mass and BMR data.
+\caption[Nonparametric tests against a normal copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for bird body mass and BMR data.
 Same format as Fig.
@@
 \includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for mammal body mass and BMR data.
+\caption[Nonparametric tests against a normal copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for mammal body mass and BMR data.
 Same format as Fig.
@@
 \includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_4_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from N-copula structure, for Cedar Creek data, year 2000.
-As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving Normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$.
+\caption[Nonparametric tests against a normal copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for Cedar Creek data, year 2000.
+As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$.
 If the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
@@
              format = "latex",linesep = "", row.names=F,
-      caption.short = "Nonparametric asymmetry tests against an Normal-copula null model with birds, mammals and Cedar Creek data",
-      caption = "Nonparametric asymmetry tests against an Normal-copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence).
+      caption.short = "Nonparametric asymmetry tests against a normal copula null model with birds, mammals and Cedar Creek data",
+      caption = "Nonparametric asymmetry tests against an normal copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence).
 For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for $(0,0.25)$ and $(0.75,1)$ range of lower and upper bounds.
@@
 C and N data
-were compared to distributions of their values on 1000 Kendall- or Spearman-preserving Normal surrogates
+were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
 of the data,
@@
 $\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
-on 1000 Kendall- or Spearman-preserving Normal surrogates, separately in two 
+on 1000 Kendall- or Spearman-preserving normal surrogates, separately in two 
 comparisons (Table \ref{MT-tab_soilCN_asym}). Results showed that
@@
 the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
-for 1000 Kendall-preserving surrogate datasets with Normal copulas. Due to relatively smaller sample size,
+for 1000 Kendall-preserving surrogate datasets with normal copulas. Due to relatively smaller sample size,
 for Cedar Creek data we used $P_{0,0.25}-P_{0.75,1}$.
@@
 If ranks were $>975$ or $<25$, asymmetry of tail 
-dependence was significant compared to an N-copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
+dependence was significant compared to a normal copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
 for, respectively, the soil C and N dataset, the bird body masses and metabolic rates dataset, 
@@
 
-<!--table for AIC-weights of Normal copula used for fitting, green spruce aphid abundance data-->
+<!--table for AIC-weights of normal copula used for fitting, green spruce aphid abundance data-->
 ```{r tab_count_AICw_normal, echo=F, results='markup'}
@@
              format='latex',linesep = "",digits=3,
-             caption.short = "AIC weights, Normal copulas, green spruce aphid abundance data",
-             caption = "$\\text{AIC}_{\\text{w}}$ for the Normal copula family for each location pair for green spruce aphid abundance data, for comparison with Table \\ref{tab_count_AICw}.
+             caption.short = "AIC weights, normal copulas, green spruce aphid abundance data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for green spruce aphid abundance data, for comparison with Table \\ref{tab_count_AICw}.
 \\label{tab_count_AICw_normal}",
@@
 
-<!--table for AIC-weights of Normal copula used for fitting, leaf plum curling aphid first flight data-->
+<!--table for AIC-weights of normal copula used for fitting, leaf plum curling aphid first flight data-->
 ```{r tab_ff_AICw_normal, echo=F, results='markup'}
@@
              format='latex',linesep = "",digits=3,
-             caption.short="AIC weights, Normal copulas, leaf-curling plum aphid first flight data",
+             caption.short="AIC weights, normal copulas, leaf-curling plum aphid first flight data",
              caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for leaf-curlung plum aphid first flight data, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
@@
 
-<!--table for AIC-weights of Normal copula used for fitting, Ceratium furca abundance data-->
+<!--table for AIC-weights of normal copula used for fitting, Ceratium furca abundance data-->
 ```{r tab_pns_AICw_normal,echo=F,results="markup"}
@@
              digits=3,
-             caption.short="AIC weights, Normal copulas, \\textit{Ceratium furca} abundance data",
-             caption = "$\\text{AIC}_{\\text{w}}$ for the Normal copula family for each location pair for \\textit{Ceratium furca} abundance data, for comparison with Table \\ref{tab_pns_AICw}.\\label{tab_pns_AICw_normal}",
+             caption.short="AIC weights, normal copulas, \\textit{Ceratium furca} abundance data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for \\textit{Ceratium furca} abundance data, for comparison with Table \\ref{tab_pns_AICw}.\\label{tab_pns_AICw_normal}",
              booktabs=TRUE)
@@
 
-<!--table for AIC-weights of Normal copula used for fitting, methane-flux data-->
+<!--table for AIC-weights of normal copula used for fitting, methane-flux data-->
 ```{r tab_methane_AICw_normal,echo=F,results="markup"}
@@
              digits=3,
-             caption.short="AIC weights, Normal copulas, methane-flux data",
-             caption = "$\\text{AIC}_{\\text{w}}$ for the Normal copula family for each location pair for methane-flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
+             caption.short="AIC weights, normal copulas, methane-flux data",
+             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for methane-flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
              booktabs=TRUE)
@@
 or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
-N-copula null hypothesis.
+normal copula null hypothesis.
 Bootstrapping was performed by resampling the locations of measurement, with replacement. 
@@
 datasets. The moderately right-tail dependent matrix and the symmetric tail dependence
-matrix were constructed similarly, but using an survival Clayton copula and an Normal copula,
+matrix were constructed similarly, but using an survival Clayton copula and an normal copula,
 respectively.
@@
 summary_stat<-summary_stat[c(1:6),]
-rownames(summary_stat)<-c("Best-fit copula AICw","N-copula AICw","Model-avg. LT","Model-avg. UT","p, CvM","p, KS")
+rownames(summary_stat)<-c("Best-fit copula AICw","Normal copula AICw","Model-avg. LT","Model-avg. UT","p, CvM","p, KS")
 knitr::kable(summary_stat, booktabs = T, escape=F,
@@
 methods section for Q1. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
-N-copula AICw: the AIC weight of the Normal copula family; Model-avg. LT: the lower-tail 
+Normal copula AICw: the AIC weight of the normal copula family; Model-avg. LT: the lower-tail 
 statistic, weighted-averaged across models using AICw; Model-avg. UT: same, but for the 
@@
 output of the model with the indicated b and g, and the value was compared to 1000 values
-of the same statistic computed on surrogate time series randomized to have N-copula structure
+of the same statistic computed on surrogate time series randomized to have normal copula structure
 (see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
@@
 indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
-lower-tail dependence compared to an N-copula null hypothesis. `FS' stands for `first slice'
+lower-tail dependence compared to a normal copula null hypothesis. `FS' stands for `first slice'
 and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
@@
 
-<!--Following chunk examine the consequence of non-Normal copula atructure on skewness using green spruce aphid abundance data-->
+<!--Following chunk examine the consequence of non-normal copula atructure on skewness using green spruce aphid abundance data-->
 ```{r skewness_count_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
@@
 
-<!--Following chunk examine the consequence of non-Normal copula atructure on skewness 
+<!--Following chunk examine the consequence of non-normal copula atructure on skewness 
 using leaf plum curling aphids' first flight data-->
@@
 
-<!--Following chunk examine the consequence of non-Normal copula atructure on skewness 
+<!--Following chunk examine the consequence of non-normal copula atructure on skewness 
 using ceratium furca's abundance data-->
@@
 
-<!--Following chunk examine the consequence of non-Normal copula atructure on skewness using methane-flux data-->
+<!--Following chunk examine the consequence of non-normal copula atructure on skewness using methane-flux data-->
 ```{r skewness_methane_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,methane_data,mtime("good_loclist.R"),mtime("skewness_multivar_dataset.R"),mtime("ncsurrog.R"),mtime("SkewnessAnd3CentMom.R"))}
@@
 
-<!--Following 4 chunks are examining consequence of non-Normal copula structure on Taylors law using made up data-->
+<!--Following 4 chunks are examining consequence of non-normal copula structure on Taylors law using made up data-->
 
@@
 
-<!--Following 2 chunks examine the effect of non-Normal asymmetric copula on extinction risk-->
+<!--Following 2 chunks examine the effect of non-normal asymmetric copula on extinction risk-->
 ```{r sim_ext_risk, echo=F,  results="hide", cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R"))}
@@
 
-<!--adding plots showing influences of non-Normal copula on spatial Taylors law-->
+<!--adding plots showing influences of non-normal copula on spatial Taylors law-->
 \begin{figure}[!h]
@@
 
-# Effects of non-Normal copula structure on Taylor's law \label{Consequence_TL_simulation}
+# Effects of non-normal copula structure on Taylor's law \label{Consequence_TL_simulation}
 
@@
 Spearman correlations between each pair of the $n$ components were the same as those
-of the Normal distribution constructed above. This was done using the `rho` and `iRho` functions 
+of the normal distribution constructed above. This was done using the `rho` and `iRho` functions 
 in the `copula` package in R. Because the Clayton copula is a one-paremeter family, this 
@@
 across the values $x$. The distrubutions of all these statistics across 1000 replicate 
-simulations were compared for the Clayton, Normal, Frank, and survival Clayton simulations to determine if copula structure 
+simulations were compared for the Clayton, normal, Frank, and survival Clayton simulations to determine if copula structure 
 influences Taylor's law.
@@
 Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
-linearity and homoskedasticity, by non-Normal copula structure. For Normal copula structure 
+linearity and homoskedasticity, by non-normal copula structure. For normal copula structure 
 (i.e., for simulations in our N-copula population array), $p$-values for the linearty
@@
 errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
-But linearity or homoskedasticity were violated more frequently for non-Normal copula structure
+But linearity or homoskedasticity were violated more frequently for non-normal copula structure
 (Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
@@
 linear regressions do not always adequately represent the $\log(v_t)$ versus $\log(m_t)$
-relationship when copula structure was not N.
+relationship when copula structure was not normal.
 
@@
 colnames(res)<-c("p-value, independence test","Best-fit copula",
-                 "Best-fit copula AICw","N-copula AICw","p, CvM","p, KS","Model-avg. LT","Model-avg. UT","Model-avg. LT minus UT","cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$")
+                 "Best-fit copula AICw","Normal copula AICw","p, CvM","p, KS","Model-avg.
+LT","Model-avg.
+UT","Model-avg.
+LT minus UT","cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$")
 res<-res[,-1]
@@
 To illustrate the potential for missing data to influence perceived
-copula structure, we first generated 5000 points from a bivariate Normal copula
+copula structure, we first generated 5000 points from a bivariate normal copula
 with parameter `r ncopparm` (Fig. \ref{fig:missingdata}A), and then 

### 2019-09-22 Paper.Rmd
edits missed before

@@
 part of Q1: these data have greater tail dependence 
-(lower and upper) than would be expected from an N-copula null hypothesis. 
+(lower and upper) than would be expected from a normal-copula null hypothesis. 
 Model-averaged lower- minus upper-tail dependence statistics 
@@
              format = "latex",linesep = "",
-      caption.short = "Nonparametric asymmetry tests against an N-copula null, soil C and N",
-      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to a normal copula null hypothesis.
+      caption.short = "Nonparametric asymmetry tests against a normal-copula null, soil C and N",
+      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to a normal-copula null hypothesis.
 The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving normal surrogates, in seperate comparisons.
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
-\caption[Nonparametric tests against an N-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for soil C and N data.
+\caption[Nonparametric tests against a normal-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for soil C and N data.
 As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
@@
 When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level).
-When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal copula null hypothesis.
+When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.
 \label{fig_soilCN_nonparam}}
@@
 $(D)$ time series compared to a multivariate 
-N-copula null hypothesis. Black dots are empirical 
+normal-copula null hypothesis. Black dots are empirical 
 skewnesses; see text 

### 2019-09-22 SupportingInformation.Rmd
edits missed before

@@
 \includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against a normal copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for bird body mass and BMR data.
+\caption[Nonparametric tests against a normal-copula null, bird body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for bird body mass and BMR data.
 Same format as Fig.
@@
 \includegraphics[width=16 cm]{./Results/BMR_results/mammals/mammals_numbin_10_bivfunctionplot.pdf}
-\caption[Nonparametric tests against a normal copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for mammal body mass and BMR data.
+\caption[Nonparametric tests against a normal-copula null, mammal body mass and BMR data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for mammal body mass and BMR data.
 Same format as Fig.
@@
 \includegraphics[width=16 cm]{./Results/Ceder_creek_results/2000cc_numbin_4_bivfunctionplot.pdf}
-\caption[Nonparametric tests against a normal copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for Cedar Creek data, year 2000.
+\caption[Nonparametric tests against a normal-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for Cedar Creek data, year 2000.
 As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$.
@@
              format = "latex",linesep = "", row.names=F,
-      caption.short = "Nonparametric asymmetry tests against a normal copula null model with birds, mammals and Cedar Creek data",
-      caption = "Nonparametric asymmetry tests against an normal copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence).
+      caption.short = "Nonparametric asymmetry tests against a normal-copula null model with birds, mammals and Cedar Creek data",
+      caption = "Nonparametric asymmetry tests against a normal-copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence).
 For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for $(0,0.25)$ and $(0.75,1)$ range of lower and upper bounds.
@@
 If ranks were $>975$ or $<25$, asymmetry of tail 
-dependence was significant compared to a normal copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
+dependence was significant compared to a normal-copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
 for, respectively, the soil C and N dataset, the bird body masses and metabolic rates dataset, 
@@
 or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
-normal copula null hypothesis.
+normal-copula null hypothesis.
 Bootstrapping was performed by resampling the locations of measurement, with replacement. 
@@
 indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
-lower-tail dependence compared to a normal copula null hypothesis. `FS' stands for `first slice'
+lower-tail dependence compared to a normal-copula null hypothesis. `FS' stands for `first slice'
 and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds

### 2019-09-22 Paper.Rmd
edits suggested by Shyamolina made

@@
 \caption[Example normal copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example normal copulas. 
-S is Spearman correlation; $p$ is the value of the parameter for the normal
+K is Kendall correlation; $p$ is the value of the parameter for the normal
 family (it is a 
@@
 \caption[Example Clayton copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Clayton  
-copulas. S is Spearman correlation; $p$ is the value of the parameter for the Clayton family (it 
+copulas. K is Kendall correlation; $p$ is the value of the parameter for the Clayton family (it 
 is a one-parameter family); and LT and UT are the measures of lower- and 
@@
 This will not affect BB1, BB6, etc., but it will affect SBB1, SBB6, etc., which will now be called survival BB1, survival BB6, etc.
-
 >>>> Done. I don't know whether it should be called as a Normal copula or an Normal copula? Now, we use "an".
-
 2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc). Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
 3) Sort the rows of Table S3 (and all analogous tables, e.g., S5, S7, etc.) by AIC, putting the lowest AIC at the top. This should mean the highest AICw is also at the top.
-
 >>>> Added, sorting based on min AIC.
-
 3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text. Same for Fig. S21 and Table S4.
@@
 8) Put the cor results only, for Spearman only (or is Kendall more consistent with what we do elsewhere - what do you think?) from tables S4, S6, S8, S10 into Table 2 as a new row.
-
 >>>> All done.

### 2019-09-22 SupportingInformation.Rmd
edits suggested by Shyamolina made

@@
 from example survival Clayton 
-copulas. S is Spearman correlation; p is the value of the parameter for the survival Clayton 
+copulas. K is Kendall correlation; p is the value of the parameter for the survival Clayton 
 family (it is a one-parameter family); and LT and UT are the measures of lower- and 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Gumbel.pdf}
-\caption[Example Gumbel copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Gumbel copulas. S is 
-Spearman correlation; $p$ is the value of the parameter for the Gumbel family (it is a 
+\caption[Example Gumbel copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Gumbel copulas. K is Kendall 
+correlation; $p$ is the value of the parameter for the Gumbel family (it is a 
 one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
@@
 \caption[Example survival Gumbel copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example survival Gumbel 
-copulas. S is Spearman correlation; $p$ is the value of the parameter for the survival Gumbel 
+copulas. K is Kendall correlation; $p$ is the value of the parameter for the survival Gumbel 
 family (it is a one-parameter family); and LT and UT are the measures of lower- and 
@@
 \caption[Example survival Joe copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example survival Joe  
-copulas. K is Spearman correlation; $p$ is the value of the parameter for the survival Joe  
+copulas. K is Kendall correlation; $p$ is the value of the parameter for the survival Joe  
 family (it is a one-parameter family); and LT and UT are the measures of lower- and 
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Frank.pdf}
-\caption[Example Frank copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Frank copulas. S is Kendall correlation; $p$ is the value of the parameter for the Frank family (it is a 
+\caption[Example Frank copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Frank copulas. K is Kendall correlation; $p$ is the value of the parameter for the Frank family (it is a 
 one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 

### 2019-09-22 MakeCopulaPedagogFigs.R

@@ -57,7 +57,7 @@ muvcpp<-function(coplist,numpts,corvals,cortype,thetavals,tot_wd,filename)
   axis(1,labels=FALSE)
   mtext("v",2,1.15)
   mtext("A",3,.05,cex=cexvalpl,adj=0)
-  mtext(paste0(cortype,"=",signif(spearvals[px+1],2),"; p=",signif(thetavals[px+1],2)),3,1.55,cex=cexvaltxt)
+  mtext(paste0(cortype,"=",signif(corvals[px+1],2),"; p=",signif(thetavals[px+1],2)),3,1.55,cex=cexvaltxt)
   lbds<-unname(lambda(coplist[[px+1]]))
   mtext(paste0("LT=",signif(lbds[1],2),"; UT=",signif(lbds[2],2)),3,1,cex=cexvaltxt)
   
@@ -76,7 +76,7 @@ muvcpp<-function(coplist,numpts,corvals,cortype,thetavals,tot_wd,filename)
   axis(1,labels=FALSE)
   axis(2,labels=FALSE)
   mtext("B",3,.05,cex=cexvalpl,adj=0)
-  mtext(paste0(cortype,"=",signif(spearvals[px+1],2),"; p=",signif(thetavals[px+1],2)),3,1.55,cex=cexvaltxt)
+  mtext(paste0(cortype,"=",signif(corvals[px+1],2),"; p=",signif(thetavals[px+1],2)),3,1.55,cex=cexvaltxt)
   lbds<-unname(lambda(coplist[[px+1]]))
   mtext(paste0("LT=",signif(lbds[1],2),"; UT=",signif(lbds[2],2)),3,1,cex=cexvaltxt)
   
@@ -95,7 +95,7 @@ muvcpp<-function(coplist,numpts,corvals,cortype,thetavals,tot_wd,filename)
   axis(1,labels=FALSE)
   axis(2,labels=FALSE)
   mtext("C",3,.05,cex=cexvalpl,adj=0)
-  mtext(paste0(cortype,"=",signif(spearvals[px+1],2),"; p=",signif(thetavals[px+1],2)),3,1.55,cex=cexvaltxt)
+  mtext(paste0(cortype,"=",signif(corvals[px+1],2),"; p=",signif(thetavals[px+1],2)),3,1.55,cex=cexvaltxt)
   lbds<-unname(lambda(coplist[[px+1]]))
[…]

### 2019-09-22 PedagogFigCopThry.R

@@ -63,7 +63,6 @@ z<-dCopula(dcarg,ncop,log=FALSE)
 z<-matrix(z,101,101)
 levs<-c(-3,-2,-1,0)
 contour(x,y,log10(z),levels=levs,xlim=c(0,1),ylim=c(0,1),col="gray",labels="")
-contour(x,y,log10(z),levels=levs,lty=0,add=TRUE)
 
 mtext("u",1,1.2)
 mtext("v",2,1.2)
@@ -74,6 +73,8 @@ mtext("NORMAL COPULA",2,2.4)
 set.seed(101)
 samps_ncop<-rCopula(numsamps,ncop)
 points(samps_ncop[,1],samps_ncop[,2],type="p",pch=20,col="grey",cex=.5)
+
+contour(x,y,log10(z),levels=levs,lty=0,add=TRUE)
 
 #**normal copula plot, top marginal
 par(fig=c((textspace+ymarwd)/totwd,
@@ -119,7 +120,6 @@ z<-dCopula(cbind(pnmarg(dcarg[,1]),pnmarg(dcarg[,2])),ncop,log=FALSE)*dnmarg(dca
 z<-matrix(z,length(x),length(y))
 levs<-c(-8,-6,-4,-2,-1)
 contour(x,y,log10(z),xlim=c(-3,3),ylim=c(-3,3),levels=levs,col="gray",labels="")
-contour(x,y,log10(z),levels=levs,lty=0,add=TRUE)
 
 mtext("x",1,1.2)
 mtext("y",2,1.2)
@@ -128,7 +128,9 @@ mtext("C",3,.8,at=c(-3-6*.35))
 #**add samples
 samps<-qnmarg(samps_ncop)
 points(samps[,1],samps[,2],type='p',pch=20,cex=0.5,col="grey")
-  
+
+contour(x,y,log10(z),levels=levs,lty=0,add=TRUE)
+
 #**first normal copula example, top marginal
 par(fig=c((textspace+2*ymarwd+panwd+biggap+margpansm)/totwd,
           (textspace+2*ymarwd+2*panwd+biggap+margpansm)/totwd,
@@ -173,7 +175,6 @@ z<-dCopula(cbind(pgmarg(dcarg[,1]),pgmarg(dcarg[,2])),ncop,log=FALSE)*dgmarg(dca
 z<-matrix(z,length(x),length(y))
 levs<-c(-10,-5,-4,-3,-2)
[…]

### 2019-09-22 Paper.Rmd
a comment changed
copula abbreviation added in table caption

@@
 whenever we currently cite one of those, change it so we cite the combined table instead.
+8) The nonparam stats are not estimators of tail dependence, they measure 
+something else (though related). Make it clear we know this. It turns out there are estimators
+of tail dependence. Do we want to mention that?
 -->
@@
 h<-BivMS_res_Loecke$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
-colnames(h)<-c("copula","AIC","AICw","BIC","BICw","LT","UT")
+colnames(h)<-c("Copula","AIC","AICw","BIC","BICw","LT","UT")
 h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC
+
+#---- This line forces the formating which is not always true by using digits arg in kable function ----
+h_ord$AIC<-formatC(h_ord$AIC,2,format="f")
+h_ord$AICw<-formatC(h_ord$AICw,2,format="f")
+h_ord$BIC<-formatC(h_ord$BIC,2,format="f")
+h_ord$BICw<-formatC(h_ord$BICw,2,format="f")
+h_ord$LT<-formatC(h_ord$LT,4,format="f")
+h_ord$UT<-formatC(h_ord$UT,4,format="f")
+#-------------------------------------------------
+
 knitr::kable(h_ord, #longtable = T, 
-             format = "latex",linesep = "",row.names = F, digits = c(4,2,2,2,2,4,4),
+             format = "latex",linesep = "",row.names = F, #digits = c(4,2,2,2,2,4,4), 
       caption.short = "Model fitting results for soil C and N dataset",
-      caption = "Model fitting results for soil C and N dataset using 16 copula families: copula name abbreviations are given in the main text.
-AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.
-Wights and tail dependences are rounded upto 2 and 4 digits, respectively.
+      caption = "Model fitting results for soil C and N dataset using 16 copula families: normal (N), Frank (F), Clayton (C), survival Clayton (SC), Gumbel (G), survival Gumbel (SG), Joe (J), survival Joe (SJ), BB1, survival BB1 (SBB1), BB6, survival BB6 (SBB6), BB7, survival BB7 (SBB7), BB8, survival BB8 (SBB8).
+Copulas are sorted according to their AIC values (2nd column).
+AICw = AIC weight; BIC = BIC value; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.
+AIC, BIC and their corresponding weights are rounded upto 2 decimal place whereas tail dependences are rounded upto 4 decimal places.
 \\label{tab_soilCNfit}",

### 2019-09-22 SupportingInformation.Rmd
copula abbreviation added in table caption

@@
 h<-BivMS_birds$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
-colnames(h)<-c("copula","AIC","AICw","BIC","BICw","LT","UT")
+colnames(h)<-c("Copula","AIC","AICw","BIC","BICw","LT","UT")
 h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC
+
+#---- This line forces the formating which is not always true by using digits arg in kable function ----
+h_ord$AIC<-formatC(h_ord$AIC,2,format="f")
+h_ord$AICw<-formatC(h_ord$AICw,2,format="f")
+h_ord$BIC<-formatC(h_ord$BIC,2,format="f")
+h_ord$BICw<-formatC(h_ord$BICw,2,format="f")
+h_ord$LT<-formatC(h_ord$LT,4,format="f")
+h_ord$UT<-formatC(h_ord$UT,4,format="f")
+#-------------------------------------------------
+
 knitr::kable(h_ord,
-             format='latex',linesep = "",row.names = F, digits = c(4,2,2,4,4),
+             format='latex',linesep = "",row.names = F, #digits = c(4,2,2,2,2,4,4),
              caption.short = "Model fitting results for the bird body mass and BMR dataset",
@@
 h<-BivMS_mammals$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
-colnames(h)<-c("copula","AIC","AICw","BIC","BICw","LT","UT")
+colnames(h)<-c("Copula","AIC","AICw","BIC","BICw","LT","UT")
 h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC
+
+#---- This line forces the formating which is not always true by using digits arg in kable function ----
+h_ord$AIC<-formatC(h_ord$AIC,2,format="f")
+h_ord$AICw<-formatC(h_ord$AICw,2,format="f")
+h_ord$BIC<-formatC(h_ord$BIC,2,format="f")
+h_ord$BICw<-formatC(h_ord$BICw,2,format="f")
+h_ord$LT<-formatC(h_ord$LT,4,format="f")
+h_ord$UT<-formatC(h_ord$UT,4,format="f")
+#-------------------------------------------------
+
 knitr::kable(h_ord,
-             format='latex',linesep = "",row.names = F, digits = c(4,2,2,4,4),
+             format='latex',linesep = "",row.names = F, #digits = c(4,2,2,2,2,4,4),
              caption.short = "Model fitting results for the mammal body mass and BMR dataset",
@@
 h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AIC","AICw","BIC","BICw","LTdep","UTdep")]
-colnames(h)<-c("copula","AIC","AICw","BIC","BICw","LT","UT")
+colnames(h)<-c("Copula","AIC","AICw","BIC","BICw","LT","UT")
 h_ord <- h[order(h$AIC),] #sorting data: min AIC to max AIC
+
+#---- This line forces the formating which is not always true by using digits arg in kable function ----
+h_ord$AIC<-formatC(h_ord$AIC,2,format="f")
+h_ord$AICw<-formatC(h_ord$AICw,2,format="f")
+h_ord$BIC<-formatC(h_ord$BIC,2,format="f")
+h_ord$BICw<-formatC(h_ord$BICw,2,format="f")
+h_ord$LT<-formatC(h_ord$LT,4,format="f")
+h_ord$UT<-formatC(h_ord$UT,4,format="f")
+#-------------------------------------------------
+
 knitr::kable(h_ord,
-             format='latex',linesep = "",row.names = F, digits = c(4,2,2,4,4),
+             format='latex',linesep = "",row.names = F, #digits = c(4,2,2,2,2,4,4),
              caption.short = "Model fitting results for the Cedar Creek dataset",
@@
              caption.short = "Best-fitting copulas, green spruce aphid abundance data",
-             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for green spruce aphid abundance data. Copulas 
-considered were the normal, Clayton, Gumbel, Frank, Joe, BB1, BB6, BB7, BB8, survival Clayton, survival Gumbel, 
-survival Joe, survival BB1, survival BB6, survival BB7, and survival BB8 copulas, with abbreviations in the methods for Q1.
+             caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for green spruce aphid abundance data.
+Copulas considered were the normal (N), Clayton (C), Gumbel (G), Frank (F), Joe (J), BB1, BB6, BB7, BB8, survival Clayton (SC), survival Gumbel (SG), survival Joe (SJ), survival BB1 (SBB1), survival BB6 (SBB6), survival BB7 (SBB7), and survival BB8 (SBB8) copulas.
 NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.\\label{tab_count_bestcop}",
@@
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for leaf-curling plum aphid first flight data.
-Copulas considered were as in Table \\ref{tab_count_bestcop}, with abbreviations in the methods for Q1.
+Copulas considered were as in Table \\ref{tab_count_bestcop}.
 NA occurs along the diagonal of the matrix because the diagonal represents a sampling location compared to itself.
@@
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for \\textit{Ceratium furca} abundance data.
-See methods for Q1 for copula abbreviations.
+Copula families were considered as in Table \\ref{tab_count_bestcop}.
 Indep.=Independent.
@@
              caption = "Best-fitting copulas for each location pair from among those considered, according to AIC, for methane-flux data.
-See methods for Q1 for copula abbreviations.
+Copula families were considered as in Table \\ref{tab_count_bestcop}.
 Indep.=Independent.

### 2019-09-22 SupportingInformation.Rmd
replace copula abbreviations with names in appendix S13

@@
 important factor in Taylor's law [@Cohen2015; @Reuman2017]. We refer henceforth to the 
-resulting $N \times n \times 1000$ array as our *N-copula population array* 
+resulting $N \times n \times 1000$ array as our *normal-copula population array* 
 for this simulation exercise;
 copula structure of dependences between populations in different locations
-was normal (N) for this array.
-
-We then generated an series of arrays that were the same dimensions ($N \times n \times 1000$)
-as the N-copula population array, and that were statistically similar to 
-the N-copula population array in all ways except the new
+was normal for this array.
+
+We then generated a series of arrays that were the same dimensions ($N \times n \times 1000$)
+as the normal-copula population array, and that were statistically similar to 
+the normal-copula population array in all ways except the new
 arrays were constructed to have Clayton (respectively, Frank, survival Clayton) copula structure for 
@@
 refer to the new arrays as our
-*C-copula*, *F-copula*, and *SC-copula population arrays*, respectively. The C-copula array was
+*Clayton-copula*, *Frank-copula*, and *Survival Clayton-copula population arrays*, respectively. The Clayton-copula array was
 constructed as follows; the others were constructed analogously. First, we constructed an
@@
 constructed above. We then randomized each simulation in 
-the N-copula population array in a special way 
+the normal-copula population array in a special way 
 (see Appendix \ref{sect:copsurrognd}) so that the resulting 
@@
 of the Clayton copula just constructed. The randomization procedure worked 
-by permuting time series of values in the N-copula array, separately for each 
+by permuting time series of values in the normal-copula array, separately for each 
 sampling location and simulation. So population marginal distrbutions for each location
-and simulation were *exactly* the same as those in the N-copula population array. 
+and simulation were *exactly* the same as those in the normal-copula population array. 
 Permutations were done in such a way
-that resulting pairwise Spearman correlations were the same as the C-copula 
-constructed above, and hence were the same as the N-copula array, 
+that resulting pairwise Spearman correlations were the same as the Clayton copula 
+constructed above, and hence were the same as the normal-copula array, 
 except for sampling variation. Thus, to summarize using mathematical notation,
-if $m^{(\text{N})}$ is our N-copula population array and $m^{(\text{C})}$ is our 
-C-copula population array, then,
+if $m^{(\text{N})}$ is our normal-copula population array and $m^{(\text{C})}$ is our 
+Clayton-copula population array, then,
 for any given $i$ and $j$, the sets $\{m^{(\text{N})}_{tij}: t=1,\ldots,N\}$ and 
@@
 linearity and homoskedasticity, by non-normal copula structure. For normal copula structure 
-(i.e., for simulations in our N-copula population array), $p$-values for the linearty
+(i.e., for simulations in our normal-copula population array), $p$-values for the linearty
 and homoskedasticity tests were roughly uniformly distributed across replicate 

### 2019-09-24 Paper.Rmd
editing results for Q1

@@
 
-For multivariate datasets, we performed the bivariate analysis
-for all possible pairwise combinations of locations, and 
-compared AIC weights and tail-dependence of 
-minimum-AIC copulas with that of a normal copula. We also counted the number of location pairs 
-for which model-averaged lower-tail dependence was 
-greater than model-averaged upper-tail dependence 
-or vice-versa, and reported the percentage of non-independent 
-location pairs showing stronger lower-tail dependence. We carried out
-pairwise bivariate analyses for simplicity and because they were
-sufficient to answer our research questions. But see the Discussion
-for a few words on multivariate copulas.
-
 We fit 16 bivariate copula families, exhibiting a variety of 
@@
 
-<!--***DAN: You decided to reverse the decision to use pithy abbreviations for each
-copula. Now the format is longer, so this might just annoy people. So this will take
-some work to replace all those throughout the main text and sup mat,
-including in figures and table and captions - everywhere.-->
+For multivariate datasets, we performed the bivariate analysis
+for all possible pairwise combinations of locations, and 
+compared AIC weights and tail-dependence of 
+minimum-AIC copulas with that of a normal copula. We also counted the number of location pairs 
+for which model-averaged lower-tail dependence was 
+greater than model-averaged upper-tail dependence 
+or vice-versa, and reported the percentage of non-independent 
+location pairs showing stronger lower-tail dependence. We carried out
+pairwise bivariate analyses for simplicity and because they were
+sufficient to answer our research questions. But see the Discussion
+for a few words on multivariate copulas.
 
@@
 between the bounds, small values of 
-$\Dsq_{l_b,u_b}$ indicate strong association.
+$\Dsq_{l_b,u_b}$ indicate strong association. These statistics are not estimators of the 
+tail dependence quantities
+defined previously, but rather are conceptually similar measures of associations in
+the tail portions of the distributions when appropriate values of $l_b$ and $u_b$ are
+selected.
 
@@
 $\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, $\Dsq_u=\Dsq_{0.5,1}$.
-
 To test for asymmetry of dependence in upper and lower portions of distributions,
@@
 to $0$ for large datasets.
-We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}). Code and data for this
+We tested our statistics in Appendix \ref{SM-Test_npa_stat} (see also 
+Figs \ref{SM-fig_stat_testing35}-\ref{SM-fig_stat_testing1000}).
+
+For bivariate datasets, we compared values 
+of the statistics $\cor_{l_b,u_b}$,
+$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
+$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
+$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ to 
+distributions of values of the same statistics computed on surrogate datasets that were produced from 
+the empirical data by randomizing it in a special way to have no 
+tail dependence (Appendix \ref{surrog_test}). 
+The surrogate/randomized datasets had exactly 
+the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
+correlation (the surrogate algorithm had two versions, one for preserving each correlation coefficient), 
+but had normal copula structure. Thus our comparisons tested the null hypothesis 
+that our statistics took values on the empirical data no different from what would have been expected 
+if the copula structure of the data were normal, but the data were otherwise 
+statistically unchanged. The comparison of one of the
+statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
+empirical data, to the distribution
+of its values computed on surrogate datasets 
+provides a test of whether association between the variables in the part of the distributions specified 
+by $l_b$ and $u_b$ is different from
+what would be expected from a null hypothesis of normal copula structure. 
+Thus the comparison addresses the first two parts of Q1
+for bivariate datasets. 
+Significant deviations correspond to deviations from normal 
+copula structure. In particular, deviations 
+using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower-tail associations 
+different from that of a normal copula;
+likewise, using $l_b=0.9$ and $u_b=1$ tests for upper-tail associations different 
+from that of a normal copula.
+The comparison of one of the statistics 
+$\cor_{0,u_b}-\cor_{1-u_b,1}$, 
+$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
+$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$,
+as computed on the empirical data, to the distribution of its values on many surrogate datasets 
+provides a test of asymmetry of tail associations. Thus this comparison helps address the
+third part of Q1 for bivariate 
+datasets.
+
+For multivariate datasets, we calculated Spearman and Kendall correlations
+and the statistics 
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
+$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
+for all pairs of sampling locations, and we then computed means across the pairs
+for each statistic. We used a spatial resampling scheme 
+(Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals of these means.
+The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
+widely used spline correlogram method; the same scheme was also used by @walteretal2017 
+as part of a method for 
+performing model selection among matrix regression models.
+Code and data for this
 project are archived at \url{www.github.com/sghosh89/BIVAN}.
@@
 1) be more nuanced in the presentation, don't give the appearance of sweeping results that dont agree with the general conclusions under the rug
-2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc). Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
+DONE 2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc).
+Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
 3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text.
-4) Sort the rows of Table S2 (and all analogous tables) AIC, putting the lowest
+DONE 4) Sort the rows of Table S2 (and all analogous tables) AIC, putting the lowest
 AIC at the top. This should mean the highest AICw is at the top.
-5) move some of the content from appendices S6 and S7 into methods, and move fig S21 and table S4 into main text. These are non-parametric results for
+DONE 5a) move some of the content from appendices S6 and S7 into methods, 
+5b) and move fig S21 and table S4 into main text. These are non-parametric results for
 soil C and N. So we will be presenting the nonparam results more comprehensively in the main text.
@@
 whenever we currently cite one of those, change it so we cite the combined table instead.
-8) The nonparam stats are not estimators of tail dependence, they measure 
+DONE 8) The nonparam stats are not estimators of tail dependence, they measure 
 something else (though related). Make it clear we know this. It turns out there are estimators
@@
 
-Across essentially all our datasets, copula structures were typically non-normal and 
+Across datasets, copula structures were often non-normal and 
 commonly showed asymmetric tail 
@@
 present results for an example bivariate dataset, then we summarize results
-for all bivariate datasets, then we present results for an example multivariate dataset, then we summarize for all multivariate datasets.
-
-<!--bivariate example-->
+for all bivariate datasets, then we present results for an example multivariate dataset, then we 
+summarize for all multivariate datasets.
+
+<!--bivariate example, model selection-->
 For the soil C and N data (Data section, Table \ref{tab_data_info}, 
@@
 and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
-These values are distinct from what a normal copula would give, namely, 0, answering the second part 
+These values are distinct from what a normal copula would give, namely, 0, helping to answer 
+the second part 
 Q1 for these data. The values also differed substantially 
-from each other, answering the third part of Q1. 
+from each other, helping answer the third part of Q1. 
 The numbers reflect stronger upper- than lower-tail
@@
 and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).
-
-Although these results convincingly show non-normal copula structure, 
-we caution that the model-averaged tail-dependence statistics may have been biased 
+Our model selection procedures do not reveal whether the tail dependence values are 
+*significantly* different from $0$ and from each other; @burnham2003_modelselection do
+not recommend attempts to make statements about statistical significance of 
+information-criterion-based, model selection results. 
+Furthermore, we caution that, while our model selection results do convincingly show 
+non-normal copula structure, model-averaged tail-dependence statistics may have been biased 
 because even the lowest-AIC copula 
@@
 Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
-available from `BiCopGofTest`). However, nonparametric statistics verified
+available from `BiCopGofTest`).
+Our alternative nonparametric
+approach, results detailed below, provides informaton about significance of tail
+associations and asymmetries of tail associations.
+
+<!--***DAN: Stopped here. Now do the nonpram results.-->
+
+However, nonparametric statistics verified
 both that tail dependence was distinct from that of a normal copula, and that it 
@@
       caption.short = "Model fitting results for soil C and N dataset",
-      caption = "Model fitting results for soil C and N dataset using 16 copula families: normal (N), Frank (F), Clayton (C), survival Clayton (SC), Gumbel (G), survival Gumbel (SG), Joe (J), survival Joe (SJ), BB1, survival BB1 (SBB1), BB6, survival BB6 (SBB6), BB7, survival BB7 (SBB7), BB8, survival BB8 (SBB8).
-Copulas are sorted according to their AIC values (2nd column).
-AICw = AIC weight; BIC = BIC value; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family, with fitted parameters; UT = the same for upper-tail dependence.
-AIC, BIC and their corresponding weights are rounded upto 2 decimal place whereas tail dependences are rounded upto 4 decimal places.
+      caption = "Model fitting results for soil C and N dataset using 16 copula families: normal (N), Frank (F), Clayton (C), survival Clayton (SC), Gumbel (G), survival Gumbel (SG), Joe (J), survival Joe (SJ), BB1, survival BB1 (SBB1), BB6, survival BB6 (SBB6), BB7, survival BB7 (SBB7), BB8, and survival BB8 (SBB8).
+Table rows are sorted according by AIC.
+AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family with fitted parameters; UT = the same for upper-tail dependence.
 \\label{tab_soilCNfit}",

### 2019-09-24 SupportingInformation.Rmd
editing results for Q1

@@
 
-For bivariate datasets, we compared values 
-of the statistics $\cor_{l_b,u_b}$,
-$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
-$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
-$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ to 
-distributions of values of the same statistics computed on surrogate datasets that were produced from 
-the empirical data by randomizing it in a special way to have no 
-tail dependence (Appendix \ref{surrog_test}). 
-The surrogate/randomized datasets had exactly 
-the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
-correlation (the surrogate algorithm had two versions, one for preserving each correlation coefficient), 
-but had normal copula structure. Thus our comparisons tested the null hypothesis 
-that our statistics took values on the empirical data no different from what would have been expected 
-if the copula structure of the data were normal, but the data were otherwise 
-statistically unchanged. The comparison of one of the
-statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
-empirical data, to the distribution
-of its values computed on many surrogate datasets 
-provides a test of whether dependence in the part of the distribution specified 
-by $l_b$ and $u_b$ is different from
-what would be expected from a null hypothesis of normal copula structure but 
-unchanged marginals and similar 
-Kendall (or Spearman)
-correlation. Thus the comparison addresses the first two parts of Q1
-for bivariate datasets. 
-Significant deviations correspond to deviations from normal 
-copula structure. In particular, deviations 
-using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
-different from that of a normal copula;
-likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
-from that of a normal copula.
-The comparison of one of the statistics 
-$\cor_{0,u_b}-\cor_{1-u_b,1}$, 
-$\Ps_{0,u_b}-\Ps_{1-u_b,1}$ and
-$\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$,
-as computed on the empirical data, to the distribution of its values on many surrogate datasets 
-provides a test of asymmetry of dependence. Thus this comparison helps address the
-third part of Q1 for bivariate 
-datasets.
+<!--***DAN: This appendix will eventually be eliminated when we move all the content to the 
+main text. The next appendix (S7) will be retained pretty much as is. Probably S8 will all be moved to
+the main text and S9 will be kept as it is.-->
 
@@
 
-For multivariate datasets, we calculated Spearman and Kendall correlations
-and the statistics 
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
-$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
-for all pairs of sampling locations, and we then computed means across the pairs
-for each statistic. We used a spatial resampling scheme 
-(Appendix \ref{spatial_resamp}) to calculate confidence intervals of these means.
-
 For the green spruce aphid abundance data, mean values across all pairs of 
@@
 
-The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
-widely used spline correlogram method; the same scheme was also used by @walteretal2017 
-as part of a method for 
-performing model selection among matrix regression models. For each pair of 
+For each pair of 
 locations $i \neq j$, the statistic

### 2019-09-29 Paper.Rmd
part way through edits to results Q1

@@
 For multivariate datasets, we performed the bivariate analysis
-for all possible pairwise combinations of locations, and 
+for all possible pairwise combinations of distinct locations. We carried out
+pairwise bivariate analyses instead of trying to fit a multivariate copula,
+for simplicity and because that approach was
+sufficient to answer our research questions; but see the Discussion
+for a few words on multivariate copulas. We present
+the number of pairs for which a non-normal copula
+was the AIC-best copula. We also computed the model-averaged lower- and
+upper-tail statistics for each pair of locations, and computed the
+mean.
+
+We used a spatial resampling scheme 
+(Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals of these means.
+The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
+widely used spline correlogram method; the same scheme was also used by @walteretal2017 
+as part of a method for 
+performing model selection among matrix regression models.
+Code and data for this
+project are archived at \url{www.github.com/sghosh89/BIVAN}.
+
 compared AIC weights and tail-dependence of 
@@
 or vice-versa, and reported the percentage of non-independent 
-location pairs showing stronger lower-tail dependence. We carried out
-pairwise bivariate analyses for simplicity and because they were
-sufficient to answer our research questions. But see the Discussion
-for a few words on multivariate copulas.
+location pairs showing stronger lower-tail dependence.
 
@@
 the empirical data by randomizing it in a special way to have no 
-tail dependence (Appendix \ref{surrog_test}). 
+tail dependence (Appendix \ref{SM-surrog_test}). 
 The surrogate/randomized datasets had exactly 
@@
 
-# Results for Q1: Ecological datasets have non-normal copula structure and asymmetric tail dependence \label{Results}
+# Results for Q1: Ecological datasets have non-normal copula structure and asymmetric tail associations \label{Results}
 
@@
 ```
-
-<!--***SHYAMOLINA: Please do the following to help me revise this section, so far only covering changes in the presentation of the bivariate results.
-Please let me know if any of this is unclear or you want to discuss anything.
-1) Replace the pithy abbreviations for the different copulas (N, C, SC, G, etc.) with their full names throughout the main text and sup mat, including in figure and table captions.
-This will not affect BB1, BB6, etc., but it will affect SBB1, SBB6, etc., which will now be called survival BB1, survival BB6, etc.
->>>> Done. I don't know whether it should be called as a Normal copula or an Normal copula? Now, we use "an".
-2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc). Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
-3) Sort the rows of Table S3 (and all analogous tables, e.g., S5, S7, etc.) by AIC, putting the lowest AIC at the top. This should mean the highest AICw is also at the top.
->>>> Added, sorting based on min AIC.
-3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text. Same for Fig. S21 and Table S4.
-7) Combine tables S6, S8, S10 into one table, and fix it so
-whenever we currently cite one of those, change it so we cite the combined table instead.
-8) Put the cor results only, for Spearman only (or is Kendall more consistent with what we do elsewhere - what do you think?) from tables S4, S6, S8, S10 into Table 2 as a new row.
->>>> All done.
--->
 
@@
 
-Across datasets, copula structures were often non-normal and 
+Copula structures were often non-normal and 
 commonly showed asymmetric tail 
@@
 and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).
-Our model selection procedures do not reveal whether the tail dependence values are 
-*significantly* different from $0$ and from each other; @burnham2003_modelselection do
-not recommend attempts to make statements about statistical significance of 
-information-criterion-based, model selection results. 
+
+Our model selection procedures do not reveal whether tail dependence parameters are 
+*significantly* different from $0$ and from each other. 
 Furthermore, we caution that, while our model selection results do convincingly show 
@@
 Our alternative nonparametric
-approach, results detailed below, provides informaton about significance of tail
+approach, results detailed next, provides informaton about significance of tail
 associations and asymmetries of tail associations.
 
-<!--***DAN: Stopped here. Now do the nonpram results.-->
-
-However, nonparametric statistics verified
-both that tail dependence was distinct from that of a normal copula, and that it 
-was asymmetric (Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}, 
-Fig. \ref{fig_soilCN_nonparam}, Table \ref{tab_soilCN_asym}). Our 
-affirmative, empirically based answer to Q1 for the soil C and N data is 
+The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
+C and N data
+were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
+of the data,
+separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
+(Fig. \ref{fig_soilCN_nonparam}). Results showed that tail associations of data are stronger in 
+both the lower and upper tails than would be expected under a null hypothesis of a normal copula (Fig. \ref{fig_soilCN_nonparam}B-F).
+The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
+$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
+on surrogates, separately in two 
+comparisons (Table \ref{tab_soilCN_asym}). Results showed that
+upper-tail association was significantly stronger than lower-tail association. These results
+are consistent with the model selection results, but go beyond them by providing information
+about statistical significance. Thus our results provide an
+affirmative answer to Q1 for the soil C and N data; this answer is 
 represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
 in the right-most box in the middle row.
+<!--***DAN: you may need to change the box around the Cedar Creek data-->
 
@@
 Table \ref{tab_bivar_summary} summarizes Q1 results 
-for the bivariate datasets. The two 
+for the bivariate datasets, with some details in the Appendices. The two 
 variables were significantly related for all datasets 
 (Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
-always emerged as the lowest-AIC (highest-AIC-weight) copula 
+always emerged as the lowest-AIC copula 
 (Table \ref{tab_bivar_summary}, row 2). 
-The normal copula was always a poor fit compared to the highest-AIC-weight copula 
-(Table \ref{tab_bivar_summary}, rows 3-4), answering the first
-part of Q1. 
-Lower- and upper-tail dependence statistics often differed substantially from 0 
-(Table \ref{tab_bivar_summary}, rows 7-8) and from each other (Table \ref{tab_bivar_summary}, row 9), 
-answering the second and third parts of Q1. 
+The normal copula was a poor fit compared to the lowest-AIC copula 
+(Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
+difference between these fits was not meaningfully large (Table \ref{SM-tab_cc2000fit}).
+Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
+Often, either lower- and upper-tail dependence statistics differed substantially from 0 
+(Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
+helping to answering the second and third parts of Q1. Though again the model
+selection results do not provide 
+information on statistical significance of these differences, and are subject to the
+caveat that, for some datasets, even the lowest-AIC copula was not an objectively good fit
+(Table \ref{tab_bivar_summary}, rows 5 and 6), nonparametric results 
+(Table \ref{tab_bivar_summary}, row 10) showed that tail associations
+deviated significantly from what would be expected from a normal-copula
+null model, and were also significantly asymmetric, except for the Cedar Creek data.
+See Figs \ref{SM-fig_bmr_birds_nonparam}-\ref{SM-fig_cc2000_nonparam} for analogues to 
+Fig. \ref{fig_soilCN_nonparam} for the bivariate datasets other than the soil C and N
+data, and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
 The first three datasets had stronger upper- 
-than lower-tail dependence, and the last dataset had the reverse, though weakly. 
-Although these results convincingly show non-normal copula structure, the 
-model-averaged tail-dependence statistics may have been biased 
-because even the lowest-AIC copula was a poor fit in some cases 
-(Table \ref{tab_bivar_summary}, rows 5-6).
-But our nonparametric statistics confirmed the same conclusions 
-(Appendices \ref{SM-Q1bivarnonparam}, \ref{SM-surrog_test}).
-Our 
-affirmative, empirically based answer to Q1 for the bivariate datasets is 
-represented in Fig. \ref{MasterFigure} as solid outlines around the names of those
+than lower-tail dependence, and the last dataset had the reverse, though non-sgnificantly. Our 
+generally affirmative, empirically based answer to Q1 for the bivariate datasets is 
+represented in Fig. \ref{MasterFigure} as solid outlines around the names of the soil C and N and bird and mammal body mass versus metabolic rate
 datasets, in the middle row of boxes.
+<!--***DAN: change the box for Cedar Creek-->
 We did the same analysis 
@@
 2007, biomass and Shannon's index were independent; for 1997-1999, results 
-were similar to 2000: the best-fitting copula was not N, and there was slightly stronger lower- 
-than upper-tail dependence. Detailed results for each dataset are in 
-Tables \ref{tab_soilCNfit} - \ref{tab_soilCN_asym},  \ref{SM-tab_bmr_birds_fit} - \ref{SM-tab_3biv_npa_summary} and 
-Figs. \ref{fig_soilCN_nonparam},  \ref{SM-fig_bmr_birds_nonparam} - \ref{SM-fig_cc2000_nonparam}.
-
-We present green spruce aphid abundance (Table \ref{tab_data_info}) results
+were similar to 2000.
+
+<!--***DAN: to do for dealing with multivariate results
+1) Here is all you want to present for each system
+a) number of total pairs and nonindep pairs
+b) number of pairs for which non-N was best; skip AICw values, i.e., rows 4 and 5 of Table 5
+c) number of well-fitting locs
+d) also keep rows 7-9 of table 5
+
+Get rid of the statement in the methods that says we should not try to combine
+model selection results with p-values
+
+Delete the following stuff from sup mat:
+Figures S24-27
+Tables S7-14, keep Table S15
+Tables S16-23, keep S24
+S25-32, keep S33
+S34-42, keep S42
+
+You will have to go back to the methods and revise those parts
+-->
+
+We present green spruce aphid abundance (Data section; Table \ref{tab_data_info}) results
 as an example multivariate analysis.
@@
       caption = "Model fitting results for soil C and N dataset using 16 copula families: normal (N), Frank (F), Clayton (C), survival Clayton (SC), Gumbel (G), survival Gumbel (SG), Joe (J), survival Joe (SJ), BB1, survival BB1 (SBB1), BB6, survival BB6 (SBB6), BB7, survival BB7 (SBB7), BB8, and survival BB8 (SBB8).
-Table rows are sorted according by AIC.
+Table rows are sorted by AIC.
 AICw = AIC weight; BICw = BIC weight; LT = the lower-tail dependence statistic for the indicated copula family with fitted parameters; UT = the same for upper-tail dependence.
@@
       caption.short = "Nonparametric asymmetry tests against a normal-copula null, soil C and N",
-      caption = "Soil C and N dataset results for nonparametric tests of whether asymmetry of tail dependence was significant, compared to a normal-copula null hypothesis.
+      caption = "Soil C and N results for nonparametric tests of whether asymmetry of tail association was significant, compared to a normal-copula null hypothesis.
 The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving normal surrogates, in seperate comparisons.
-A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.\\label{tab_soilCN_asym}")
+A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.
+Results were significant only for the $P$ and $D^2$ statistics.\\label{tab_soilCN_asym}")
 ```
@@
 \includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
-\caption[Nonparametric tests against a normal-copula null, soil C and N data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for soil C and N data.
+\caption[Nonparametric tests against a normal-copula null, soil C and N data]{Nonparametric tests for tail association and other deviations from normal copula structure, for soil C and N data.
 As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
@@
 When the statistic was greater than 975 or less than 975 surrogate values, it indicates significance ($95\%$ confidence level).
-When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means dependence in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.
+When $\cor$ or $\Ps$ values (respectively, $\Dsq$ values) were greater than surrogates, it means association in that part of the distributions was stronger than (respectively, weaker than) expected from a normal-copula null hypothesis.
 \label{fig_soilCN_nonparam}}
@@
            "10. Asymmetry rank in surrogates", # this row will be commented out, that's why the numbering starts from 10 in the next line
-           "10. How many times non-parametric tail asymmetry ($Cor_{FS}-Cor_{LS}$) on given data are less (or greater) than the Spearman-preserving normal surrogates")
+           "10. Placement of cor$_{FS}$-cor$_{LS}$ in surrogate distribution (see caption)")
 )
@@
 Model averaging used for rows 7-9 was based on AIC weights.
-First 3 bivariate data use the default range for first slice (FS) is (0,0.1) and for last slice (LS) is (0.9, 1), whereas Cedar Creek result uses range (0, 0.25) for FS and (0.75, 1) for LS for the year 2000.\\label{tab_bivar_summary}",
+Row 10 shows values as in Table \\ref{tab_soilCN_asym}, upper right table entry.
+Although the result shown in row 10 for the soil C and N data is non-significant, see Table \\ref{tab_soilCN_asym} for significant results using the $P$ and $D^2$ statistics.
+The first 3 datasets use $(l_b,u_b)=(0,0.1)$ for the first slice (FS) and $(0.9,1)$ for last slice (LS), whereas $(0, 0.25)$ and $(0.75,1)$ were used for Cedar Creek due to fewer data for that system.
+Results for Cedar Creek for the year 2000 are shown.
+See Table \\ref{tab_soilCNfit} for copula family abbreviations.\\label{tab_bivar_summary}",
              booktabs=T, linesep = "\\addlinespace",

### 2019-09-29 SupportingInformation.Rmd
part way through edits to results Q1

@@
       caption.short = "Nonparametric asymmetry tests against a normal-copula null model with birds, mammals and Cedar Creek data",
-      caption = "Nonparametric asymmetry tests against a normal-copula null model with birds, mammals and Cedar creek data: for birds and mammals data, results should be interpreted as of soil C and N results (see Table \\ref{MT-tab_soilCN_asym} which shows stronger upper tail dependence).
-For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for $(0,0.25)$ and $(0.75,1)$ range of lower and upper bounds.
-Though the non-parametric result does not indicate significantly stronger lower tail dependence for Cedar Creek data, the
-greater than (>) sign before the numbers do show the opposite direction in tail-dependence compared to other bivariate datasets.\\label{tab_3biv_npa_summary}")
-```
-
-# Verifying Q1 results for bivariate datasets using nonparametric statistics \label{Q1bivarnonparam}
-
-<!--***DAN: This appendix will eventually be eliminated when we move all the content to the 
-main text. The next appendix (S7) will be retained pretty much as is. Probably S8 will all be moved to
-the main text and S9 will be kept as it is.-->
-
-The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
-C and N data
-were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
-of the data,
-separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
-(Fig. \ref{MT-fig_soilCN_nonparam}). Results show that tail dependence of data is stronger in 
-both the lower and upper tails than would be expected under a null hypothesis of a normal copula.
-The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
-$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
-on 1000 Kendall- or Spearman-preserving normal surrogates, separately in two 
-comparisons (Table \ref{MT-tab_soilCN_asym}). Results showed that
-upper-tail dependence was significantly stronger than lower-tail dependence.
-
-For first three bivariate datasets in Table \ref{MT-tab_bivar_summary},
-the statistic $P_{0,0.1}-P_{0.9,1}$ was computed for real data and 
-for 1000 Kendall-preserving surrogate datasets with normal copulas. Due to relatively smaller sample size,
-for Cedar Creek data we used $P_{0,0.25}-P_{0.75,1}$.
-For each dataset, the rank of the statistic for real-data in 
-the distribution of surrogate values was computed. 
-If ranks were $>975$ or $<25$, asymmetry of tail 
-dependence was significant compared to a normal-copula null hypothesis. Ranks were `r unname(tab_bivar_summary[10,])`
-for, respectively, the soil C and N dataset, the bird body masses and metabolic rates dataset, 
-the mammal body masses and metabolic rates dataset, and the Cedar Creek data for the year 2000. 
-Thus the first three datasets showed stronger 
-upper- than lower-tail dependence, and the Cedar Creek data showed the reverse 
-(though not significantly), confirming the 
-results based on model selection presented in the main text (see 
-Table \ref{MT-tab_bivar_summary}, row 9). Model selection results for Cedar Creek showed 
-only slightly stronger lower- than upper-tail dependence, consistent with the non-significant 
-tendency toward lower-tail dependence also revealed by the non-parametric analysis for that
-dataset.
+      caption = "Nonparametric asymmetry tests against a normal-copula null model with birds, mammals and Cedar Creek data.
+For bird and mammal data, results should be interpreted in the same way as soil C and N results (Table \\ref{MT-tab_soilCN_asym}).
+For Cedar Creek data, first slice ($FS$) and last slice ($LS$) stand for the ranges $(0,0.25)$ and $(0.75,1)$ for the lower and upper bounds.\\label{tab_3biv_npa_summary}")
+```
 
@@
 
+<!--***DAN: I'm inclined to eliminate this appendix and put the material
+in the main text-->
+
 For the green spruce aphid abundance data, mean values across all pairs of 
@@
 # Spatial resampling with multivariate datasets for nonparametric statistics \label{spatial_resamp}
+<!--***DAN: leave this section exactly as it is, but it may be out of order
+now that we describe how this is done in the methods. In fact, I think I
+have to do this more generally.-->
 

### 2019-09-29 Paper.Rmd
table fixed as suggested by DCR

@@
 dt <- tibble(Description = c("1. p-value, independence test",
-           "2. Minimum-AIC copula",
-           "3. AIC weight of minimum-AIC copula",
-           "4. AIC weight of normal copula",
+           "2. Minimum-AIC bestfit copula",
+           "3. AIC for bestfit copula",
+           "4. AIC for normal copula",
            "5. p-value, Cramer-von Mises goodness of fit test",
@@
              col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar Creek data")
-             )%>%column_spec(2:5,width="1.4cm")%>%
-             column_spec(1, width = "8.5cm")%>%kable_styling(latex_options = "basic",position = "center")
+             )%>%column_spec(2:5,width="1.5cm")%>%
+             column_spec(1, width = "8.4cm")%>%kable_styling(latex_options = "basic",position = "center")
 ```
@@
 
-tab_multivar_summary<-rbind(tab_multivar_summary[c(1:6),],
+tab_multivar_summary<-rbind(tab_multivar_summary[c(1:3,5,6),],
                             two_tab_in1[1:2,], #CI of LT, CI of UT
@@
            "3. Number of pairs with non-normal copula as best fit",
-           "4. Average AIC-weights of best-fitting copulas across location pairs",
-           "5. Average AIC-weights of normal copula across pairs",
-           "6. Number of well-fitting location pairs",
-           "7. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. lower-tail dependence",
-           "8. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. upper-tail dependence",
-           "9. Percent pairs showing stronger lower tail dependence",
-           "10. 2.5$^{th}$ and 97.5$^{th}$ quantiles of $<$P$_{l}$ - P${_u}$$>$ across all location pair")
+           #"4. Avg. of normal-AICw minus bestfit-copula AICw across location pairs",
+           "4. Average of AIC differences (normal minus bestfit-copula) across location pairs",
+           "5. Number of well-fitting location pairs",
+           "6. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. lower-tail dependence",
+           "7. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. upper-tail dependence",
+           "8. Percent pairs showing stronger lower tail dependence",
+           "9. 2.5$^{th}$ and 97.5$^{th}$ quantiles of $<$P$_{l}$ - P${_u}$$>$ across all location pair")
            )
@@
 
-knitr::kable(tab_multivar_summary[1:9,], 
+knitr::kable(tab_multivar_summary[1:8,], 
              format="latex", escape = F,
@@
              caption = "Summary of Q1 results for multivariate datasets.
-Rows 3-9 of the table were computed for the non-independent pairs (row 2) only.
-A well-fitting location pair (row 6) was one for which the best-fitting copula had $p$-values $>0.01$ for both the Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests.
+Rows 3-8 of the table were computed for the non-independent pairs (row 2) only.
+A well-fitting location pair (row 5) was one for which the best-fitting copula had $p$-values $>0.01$ for both the Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests.
 \\label{tab_multivar_summary}",
@@
                         "\\textit{Ceratium furca} abundance","Methane-flux")
-             )%>%column_spec(2:5,width="2.5cm")%>%
+             )%>%column_spec(2:5,width="2.4cm")%>%
              column_spec(1, width = "6cm")%>%kable_styling(latex_options = "basic",position = "center")

### 2019-09-29 SupportingInformation.Rmd
table fixed as suggested by DCR

@@
   m<-matrix(NA,1,10)
-  colnames(m)<-c("p.Independence.test","minimumAIC.bestfit.copula","Bestfit.copula.AICw",
-                 "Normal.copula.AICw","p.CvM","p.KS",
+  colnames(m)<-c("p.Independence.test","minimumAIC.bestfit.copula","Bestfit.copula.AIC",
+                 "Normal.copula.AIC","p.CvM","p.KS",
                  "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
@@
   m$minimumAIC.bestfit.copula<-r$InfCritRes$copname[which.min(r$InfCritRes$AIC)]
-  m$Bestfit.copula.AICw<-round(max(r$InfCritRes$AICw),3)
-  m$Normal.copula.AICw<-round(r$InfCritRes$AICw[1],3)
+  m$Bestfit.copula.AIC<-round(min(r$InfCritRes$AIC),2)
+  m$Normal.copula.AIC<-round(r$InfCritRes$AIC[1],2)
   m$p.CvM<-r$GofRes_CvM
   m$p.KS<-r$GofRes_KS
-  m$LT.model.avg<-round(r$relLTdep_AICw,3)
-  m$UT.model.avg<-round(r$relUTdep_AICw,3)
+  m$LT.model.avg<-round(r$relLTdep_AICw,4)
+  m$UT.model.avg<-round(r$relUTdep_AICw,4)
   m$`(LT-UT).model.avg`<-m$LT.model.avg-m$UT.model.avg
@@
 source("normal_cop_AICw.R")
-knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
+h<-normal_cop_AICw_matrix(r=RES_aphid_count)
+h<-h$normal_cop_AICw
+h[is.infinite(h)]<-NA
+knitr::kable(h,
              format='latex',linesep = "",digits=3,
@@
 source("normal_cop_AICw.R")
-knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
+h<-normal_cop_AICw_matrix(r=RES_aphid_ff)
+h<-h$normal_cop_AICw
+h[is.infinite(h)]<-NA
+knitr::kable(h,
              format='latex',linesep = "",digits=3,
@@
 h<-normal_cop_AICw_matrix(r=RES_plankton_north_sea)
+h<-h$normal_cop_AICw
 h[is.infinite(h)]<-NA
@@
 h<-normal_cop_AICw_matrix(r=RES_methane)
+h<-h$normal_cop_AICw
 h[is.infinite(h)]<-NA
@@
   colnames(m)<-c("num.pairs","num.non-indep","num.non-N",
-                 "Avg.best.AICw","Avg.N.AICw","num.good.fit",
+                 "Avg.N_minus_bestfit.AICw","Avg.N_minus_bestfit.AIC","num.good.fit",
                  "0.025-CI.LT","0.975-CI.LT","0.025-CI.UT","0.975-CI.UT",
@@
 
-  normal_AICw<-normal_cop_AICw_matrix(r=r)
+  call_AICs_N<-normal_cop_AICw_matrix(r=r)
+  normal_AICw<-call_AICs_N$normal_cop_AICw
+  normal_AICw[normal_AICw==Inf]<-NaN
+  normal_AIC<-call_AICs_N$normal_cop_AIC
+  normal_AIC[normal_AIC==Inf]<-NaN
 
@@
   m$`num.non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
+
   bcop<-r$info_ord_copcode[,,1] # matrix with best fit copula name
@@
   m$`num.non-N`<-sum(bcop!=1,na.rm = T)
-  r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN
-  m$Avg.best.AICw<-round(mean(r$info_ord_AICw[,,1],na.rm = T),3)
-  normal_AICw[normal_AICw==Inf]<-NaN
-  m$Avg.N.AICw<-round(mean(normal_AICw,na.rm=T),3)
+
+  r$info_ord_AIC[r$info_ord_AIC==Inf] <- NaN # replacing non-ind loc pair value by NaN
+  r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN # replacing non-ind loc pair value by NaN
+
+  bestfit.AIC<-r$info_ord_AIC[,,1]
+  bestfit.AICw<-r$info_ord_AICw[,,1]
+
+  m$Avg.N_minus_bestfit.AICw<-round(mean(normal_AICw - bestfit.AICw, na.rm = T),3)
+  m$Avg.N_minus_bestfit.AIC<-round(mean(normal_AIC - bestfit.AIC, na.rm = T),3)
+
   pcvm<-r$gfc_p_CvM

### 2019-09-29 normal_cop_AICw.R

@@ -1,9 +1,11 @@
-# This function gives you a matrix where its [i,j] element is the AICw for a fitted normal copula for the location pair [i,j]
+# This function gives you a list of two matrices where its [i,j] element are the AICw and AIC 
+# for a fitted normal copula for the location pair [i,j]
 # Input : 
 #       r is the RDS file which came as an output from FittingCopula_selective_loc.R
 
 normal_cop_AICw_matrix<-function(r){
   normal_cop_AICw<-r$gfc_numBS
+  normal_cop_AIC<-normal_cop_AICw
   #matrix(NA, nrow=dim(r$gfc_p_CvM)[1], ncol=dim(r$gfc_p_CvM)[2])
   #colnames(normal_cop_AICw)<-colnames(r$gfc_p_CvM)
   #rownames(normal_cop_AICw)<-rownames(r$gfc_p_CvM)
@@ -15,11 +17,14 @@ normal_cop_AICw_matrix<-function(r){
         norm_ind<-which(x==1)
         norm_AICw<-r$info_ord_AICw[i,j,norm_ind]
         normal_cop_AICw[i,j]<-norm_AICw
+        norm_AIC<-r$info_ord_AIC[i,j,norm_ind]
+        normal_cop_AIC[i,j]<-norm_AIC
       }
     }
   } 
   
-  return(normal_cop_AICw)
+  return(list(normal_cop_AICw=normal_cop_AICw,
+              normal_cop_AIC=normal_cop_AIC))
   
 }
 

### 2019-09-30 Paper.Rmd
edits to Q1 methods and results

@@
 \noindent \emph{Correspondence:} Daniel Reuman, 2101 Constant Ave, Lawrence, KS, 66047, reuman@ku.edu, 626 560 7084.
-
-\noindent \emph{Type of Submission:} Articles, or Concepts and Synthesis
 
@@
 recommend carrying out inferences of dependence structures (which was our
-goal here) using ranks.
+goal here) using normalized ranks.
 We used the fitting implementation given in `BiCopEst` in `VineCopula`.
@@
 values: the BB1, survival BB1, BB7 and survival BB7 families. 
-We also used the BB8 copula, which shows zero lower-tail dependence,
+We also used the BB8 family, which shows zero lower-tail dependence,
 and zero upper-tail dependence except for a boundary case for the parameters.
-We used the survival BB8 copula, which shows zero upper-tail dependence, 
-and zero lower-tail dependence except for a boundary case for the 
-paremeters.
+And we used the survival BB8 copula, which shows zero upper-tail dependence, 
+and zero lower-tail dependence except for a boundary case.
 "Survival" families are rotations of the copula with the similar name by
@@
 sufficient to answer our research questions; but see the Discussion
-for a few words on multivariate copulas. We present
+for a few words on multivariate copulas. 
+<!--***DAN: dont forget to put these in!-->
+We present
 the number of pairs for which a non-normal copula
-was the AIC-best copula. We also computed the model-averaged lower- and
-upper-tail statistics for each pair of locations, and computed the
-mean.
-
-We used a spatial resampling scheme 
-(Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals of these means.
-The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
-widely used spline correlogram method; the same scheme was also used by @walteretal2017 
-as part of a method for 
-performing model selection among matrix regression models.
-Code and data for this
-project are archived at \url{www.github.com/sghosh89/BIVAN}.
-
-compared AIC weights and tail-dependence of 
-minimum-AIC copulas with that of a normal copula. We also counted the number of location pairs 
-for which model-averaged lower-tail dependence was 
-greater than model-averaged upper-tail dependence 
-or vice-versa, and reported the percentage of non-independent 
-location pairs showing stronger lower-tail dependence.
+was the AIC-best copula, and we characterize AIC differences between
+the normal and AIC-best copulas across location pairs. 
+We also computed the model-averaged lower- and
+upper-tail statistics, and differences
+between these, for each pair of locations, and we characterize
+the distribution of these values across location pairs.
 
@@
 
-Copula structures were often non-normal and 
-commonly showed asymmetric tail 
-dependence, answering Q1 in the affirmative. 
+Copula structures were non-normal and 
+showed asymmetric tail 
+associations for most, but not all datasets we examined, 
+answering Q1 in the affirmative. 
 To make results easier to absorb, we first 
@@
 `r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the 
-minimum-AIC copula, with AIC weight `r round(max(BivMS_res_Loecke$InfCritRes$AICw),2)`, whereas the 
-normal copula had AIC weight essentially 0, answering the 
-first part of Q1 for these data. 
+minimum-AIC copula, with AIC `r round(min(BivMS_res_Loecke$InfCritRes$AIC),2)`, whereas the 
+normal copula had a much higher AIC, $-6333.16$, 
+<!--***DAN: Shyamolina, please make this an autolink. Please make 
+sure to round to two digits after the decimal, as with the AIC best value
+presented just above.-->
+answering the 
+first part of Q1 for these data. Recall that AIC differences of $2$
+or $3$ are considered meaningful and differences of $8$ or greater
+are considered diffinitive, so that data provide essentially no support
+for the higher-AIC model in that case [@burnham2003_modelselection].
 Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
@@
 Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
-Often, either lower- and upper-tail dependence statistics differed substantially from 0 
+Often, either lower- or upper-tail dependence statistics differed substantially from 0 
 (Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
@@
 The first three datasets had stronger upper- 
-than lower-tail dependence, and the last dataset had the reverse, though non-sgnificantly. Our 
-generally affirmative, empirically based answer to Q1 for the bivariate datasets is 
-represented in Fig. \ref{MasterFigure} as solid outlines around the names of the soil C and N and bird and mammal body mass versus metabolic rate
+than lower-tail dependence, and the last dataset had the reverse, though non-significantly. Our 
+generally affirmative (except for Cedar Creek), 
+empirically based answer to Q1 for the bivariate datasets is 
+represented in Fig. \ref{MasterFigure} as solid outlines around the names of 
+the soil C and N and bird and mammal body mass versus metabolic rate
 datasets, in the middle row of boxes.
@@
 Delete the following stuff from sup mat:
-Figures S24-27
-Tables S7-14, keep Table S15
-Tables S16-23, keep S24
-S25-32, keep S33
-S34-42, keep S42
-
-You will have to go back to the methods and revise those parts
+DONE Figures S24-27
+DONE Tables S7-14, keep Table S15
+DONE Tables S16-23, keep S24
+DONE S25-32, keep S33
+DONE S34-42, keep S42
+
+DONE You will have to go back to the methods and revise those parts
 -->
@@
 Independence was rejected for each pair of the 10 sampling locations. 
-Best-fitting (lowest-AIC) copulas were non-N
-for the large majority of location pairs (Table \ref{SM-tab_count_bestcop}), and 
-AIC weights for best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
-were typically substantially 
-larger than normal copula weights (Table \ref{SM-tab_count_AICw_normal}), answering the first part of Q1. Goodness of fit tests 
-(Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
-failed to reject the hypothesis that the AIC-best copula family was also 
+Best-fitting (lowest-AIC) copulas were non-normal
+for the large majority of location pairs (72 of 90 pairs), 
+<!--***DAN: Shyamolina, please make these numbers autolinks-->
+and the AIC
+of the normal copula minus the minimum AIC, averaged across 
+location pairs, was $2.732$.
+<!--***DAN: Shyamolina, please make this an autolink, so the value is
+automatically filled in even if it changes-->
+Whereas $2.732$
+<!--***DAN: Shyamolna, again, autolink, please-->
+would be only a marginally meaningful AIC difference for a single
+location pair, it is more meaningful as an average across many
+pairs. Relatedly, and illustrating the concept here, 
+the chances of getting 72 or more 
+non-normal location pairs if the chances were equal of getting
+a normal versus a non-normal result (and taking into account that
+the location pair (A,B) will necessarily 
+produce the same result as the pair
+(B,A)) is low, $`r 1-pbinom(72/2-1,45,.5)`$. 
+<!--***DAN: Shyamolina, please replace the 72 in the above expression
+by an autolink, and make this show
+3 significant digits, i.e., it should show 3.29e-5, not 3.287332e-05-->
+These results help answer the first part of Q1: non-normal
+copula structure appears to be meaningfully common in
+these data, even if not universal.
+Goodness of fit tests in every case 
+failed to reject the hypothesis that the AIC-best copula 
+family was also 
 an objectively adequate description of 
 the data; i.e., the collection of 
-copula families we used was sufficiently broad to characterize these data.
-Model-averaged lower- (Table \ref{SM-tab_count_LTdep_AICw}) and upper-tail
-(Table \ref{SM-tab_count_UTdep_AICw}) dependence statistics were exclusively 
-positive and were frequently large, thereby differing 
-from what a normal copula would give (i.e., 0), and answering the second
+copula families we used was sufficiently broad to characterize 
+these data. Model-averaged lower- and upper-tail
+dependence statistics had $2.5$th and $97.5$th quantiles
+$(0.123,0.721)$ and $(0.002,0.41)$, respectively,
+<!--***DAN: Shyamolina, autolink pls -->
+thereby commonly differing 
+from what a normal copula would give (i.e., 0), and helping to
+answer the second
 part of Q1: these data have greater tail dependence 
-(lower and upper) than would be expected from a normal-copula null hypothesis. 
+(lower and upper) than would be expected from a normal-copula null hypothesis. We note however that our model selection 
+procedures again do not reveal whether tail-dependence 
+parameters are significantly different from $0$, and we refer 
+the reader to nonparametric results below for information
+on significance. 
 Model-averaged lower- minus upper-tail dependence statistics 
-(Table \ref{SM-tab_count_LTmUT}) were positive 
-for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
-out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
+were positive for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
+out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). The chances
+of getting `r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
+or more positive values, here, under a null hypothesis of equal chances 
+for positive and negative values (and again accounting for the fact that location pairs (A,B) and (B,A) will show the same result) was
+again low, $`r 1-pbinom(86/2-1,45,.5)`$. 
+<!--***DAN: Shyamolina, please autolink this, and use 3 sig figs as 
+before-->
+Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
 than the spatial synchrony of outbreaks. This also answers the third part of Q1.
+
+<!--***DAN: got to here. Next step, put in the actual nonparam stats results which are currently in the Appendices-->
+
 Nonparametric statistics verified that tail dependence was asymmetric for the green 
@@
 dt <- tibble(Description = c("1. p-value, independence test",
-           "2. Minimum-AIC bestfit copula",
-           "3. AIC for bestfit copula",
+           "2. Minimum-AIC copula",
+           "3. AIC for best copula",
            "4. AIC for normal copula",
@@
 Although the result shown in row 10 for the soil C and N data is non-significant, see Table \\ref{tab_soilCN_asym} for significant results using the $P$ and $D^2$ statistics.
-The first 3 datasets use $(l_b,u_b)=(0,0.1)$ for the first slice (FS) and $(0.9,1)$ for last slice (LS), whereas $(0, 0.25)$ and $(0.75,1)$ were used for Cedar Creek due to fewer data for that system.
+The first 3 datasets use $(l_b,u_b)=(0,0.1)$ for the first slice (FS) and $(0.9,1)$ for the last slice (LS), whereas $(0, 0.25)$ and $(0.75,1)$ were used for Cedar Creek due to fewer data for that system.
 Results for Cedar Creek for the year 2000 are shown.
@@
            #"4. Avg. of normal-AICw minus bestfit-copula AICw across location pairs",
-           "4. Average of AIC differences (normal minus bestfit-copula) across location pairs",
+           "4. Average of AIC differences (normal AIC minus min AIC) across location pairs",
            "5. Number of well-fitting location pairs",

### 2019-09-30 SupportingInformation.Rmd
edits to Q1 methods and results

@@
 <!--table for Best-fitting copulas, green spruce aphid abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_count_bestcop, echo=F, results="markup"}
@@
 ```
+-->
 
 <!--table for AIC-weights of best-fitting copulas, green spruce aphid abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_count_AICw, echo=F, results='markup'}
@@
 ```
+-->
 
 <!--table for AIC-weights of normal copula used for fitting, green spruce aphid abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_count_AICw_normal, echo=F, results='markup'}
@@
 ```
+-->
 
 <!--table for p-values from goodness of fit test (Cramer-von Mises-based test) with best-fit copula, green spruce aphid abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_count_p_CvM, echo=F, results='markup'}
@@
 ```
+-->
 
 <!--table for p-values from goodness of fit test (Kolmogorov-Smirnov-based test) with best-fit copula, green spruce aphid abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_count_p_KS, echo=F, results='markup'}
@@
 ```
+-->
 
 <!--table for model averaged lower tail dep., green spruce aphid abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_count_LTdep_AICw, echo=F, results='markup'}
@@
 ```
+-->
 
 <!--table for model averaged upper tail dep., green spruce aphid abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_count_UTdep_AICw, echo=F, results='markup'}
@@
 ```
+-->
 
 <!--table for model averaged lower- minus upper-tail dep., green spruce aphid abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_count_LTmUT, echo=F, results='markup'}
@@
 ```
+-->
 
@@
 
+<!--***DAN: decided to remove this for AER
 \begin{figure}[!h]
@@
 <!--table for Best-fitting copulas, leaf plum curling aphid first flight data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_ff_bestcop, echo=F, results="markup"}
@@
 ```
+-->
 
 <!--table for AIC-weights of best-fitting copulas, leaf plum curling aphid first flight data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_ff_AICw, echo=F, results='markup'}
@@
 ```
+-->
 
 <!--table for AIC-weights of normal copula used for fitting, leaf plum curling aphid first flight data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_ff_AICw_normal, echo=F, results='markup'}
@@
 leaf plum curling aphid first flight data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_ff_p_CvM, echo=F, results='markup'}
@@
 leaf plum curling aphid first flight data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_ff_p_KS, echo=F, results='markup'}
@@
 ```
+-->
 
 <!--table for model averaged lower tail dep., leaf plum curling aphid first flight data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_ff_LTdep_AICw, echo=F, results='markup'}
@@
 ```
+-->
 
 <!--table for model averaged upper tail dep., leaf plum curling aphid first flight data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_ff_UTdep_AICw, echo=F, results='markup'}
@@
 ```
+-->
 
 <!--table for model averaged lower- minus upper-tail dep., leaf plum curling aphid first flight data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_ff_LTmUT, echo=F, results='markup'}
@@
 <!--table for Best-fitting copulas, Ceratium furca abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_pns_bestcop, echo=F, results="markup"}
@@
 ```
+-->
 
 <!--table for AIC-weights of best-fitting copulas, Ceratium furca abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_pns_AICw,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for AIC-weights of normal copula used for fitting, Ceratium furca abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_pns_AICw_normal,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for p-values from goodness of fit test (Cramer-von Mises-based test) with best-fit copula, Ceratium furca abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_pns_p_CvM,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for p-values from goodness of fit test (Kolmogorov-Smirnov-based test) with best-fit copula, Ceratium furca abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_pns_p_KS,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for model averaged lower tail dep., Ceratium furca abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_pns_LTdep_AICw,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for model averaged upper tail dep., Ceratium furca abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_pns_UTdep_AICw,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for model averaged lower- minus upper-tail dep., Ceratium furca abundance data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_pns_LTmUT,echo=F,results="markup"}
@@
 <!--table for Best-fitting copulas, methane-flux data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_methane_bestcop, echo=F, results="markup"}
@@
 ```
+-->
 
 <!--table for AIC-weights of best-fitting copulas, methane-flux data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_methane_AICw,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for AIC-weights of normal copula used for fitting, methane-flux data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_methane_AICw_normal,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for p-values from goodness of fit test (Cramer-von Mises-based test) with best-fit copula, methane-flux data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_methane_p_CvM,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for p-values from goodness of fit test (Kolmogorov-Smirnov-based test) with best-fit copula, methane-flux data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_methane_p_KS,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for model averaged lower tail dep., methane-flux data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_methane_LTdep_AICw,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for model averaged upper tail dep., methane-flux data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_methane_UTdep_AICw,echo=F,results="markup"}
@@
 ```
+-->
 
 <!--table for model averaged lower- minus upper-tail dep., methane-flux data-->
+<!--***DAN: decided to remove this for AER
 ```{r tab_methane_LTmUT,echo=F,results="markup"}

### 2019-09-30 Paper.Rmd
just about done editing Q1 methods and results for this round, just want to go through it one more time after Shyamolina makes some changes I requested in an email

@@
 
-<!--***DAN: to do for dealing with multivariate results
-1) Here is all you want to present for each system
-a) number of total pairs and nonindep pairs
-b) number of pairs for which non-N was best; skip AICw values, i.e., rows 4 and 5 of Table 5
-c) number of well-fitting locs
-d) also keep rows 7-9 of table 5
-
-Get rid of the statement in the methods that says we should not try to combine
-model selection results with p-values
-
-Delete the following stuff from sup mat:
-DONE Figures S24-27
-DONE Tables S7-14, keep Table S15
-DONE Tables S16-23, keep S24
-DONE S25-32, keep S33
-DONE S34-42, keep S42
-
-DONE You will have to go back to the methods and revise those parts
--->
-
 We present green spruce aphid abundance (Data section; Table \ref{tab_data_info}) results
@@
 $(0.123,0.721)$ and $(0.002,0.41)$, respectively,
-<!--***DAN: Shyamolina, autolink pls -->
+<!--***DAN: Shyamolina, autolinks pls -->
 thereby commonly differing 
@@
 procedures again do not reveal whether tail-dependence 
-parameters are significantly different from $0$, and we refer 
+parameters are *significantly* different from $0$, and we refer 
 the reader to nonparametric results below for information
@@
 Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
-than the spatial synchrony of outbreaks. This also answers the third part of Q1.
-
-<!--***DAN: got to here. Next step, put in the actual nonparam stats results which are currently in the Appendices-->
-
-Nonparametric statistics verified that tail dependence was asymmetric for the green 
-spruce aphid abundance data (Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}).
+than the spatial synchrony of outbreaks. This answers the third part of Q1.
+
+Nonparametric statistics verified that tail associations were asymmetric for the green 
+spruce aphid abundance data. Mean values across all pairs of 
+locations of the Spearman and Kendall correlations and the statistics 
+$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ were positive and
+confidence intervals excluded $0$ (Table \ref{tab_count_resamp}). 
+Mean values of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
+also were always positive and had confidence intervals that 
+excluded $0$ 
+(Table \ref{tab_count_resamp}).
+
+<!--***DAN: Shyamolina, can you please insert code to make what
+is currently Table S7 appear here? Please also make it disappear from
+the sup mat. I want to move it from there to here, please. I am
+talking about the table with label tab_count_resamp. Please also
+search both the main text and the supp mat for that label, and 
+fix references - you may have to fix some prefixes (SM-, MT-), since
+the table will have moved from sup mat to main text.-->
 
@@
 conclusions that non-normal copula structure, non-normal tail dependence, and asymmetric tail 
-dependence were common, answering Q1 in the affirmative for these data. 
+dependence were common, though not universal, answering 
+Q1 in the affirmative for these data. 
 Most location pairs were non-independent 
@@
 pairs had best-fitting copulas that were not the normal copula (Table \ref{tab_multivar_summary}, 
-row 3), and AIC weights of best-fitting copulas were 
-better than those of normal copulas (Table \ref{tab_multivar_summary}, rows 4-5).
-Best-fitting copulas were nearly always considered an adequate fit 
-(Table \ref{tab_multivar_summary}, row 6). Some of the datasets (green spruce aphid
+row 3), and AIC values of best-fitting copulas were, on average
+across location pairs, between $2.714$ (*Ceratium furca* abundance data)
+and $5.764$ (leaf-curling plum aphid first-flight data) lower than
+AIC values for the normal copula (Table \ref{tab_multivar_summary}, 
+row 4). Best-fitting copulas were nearly always considered an adequate fit 
+(Table \ref{tab_multivar_summary}, row 5). Some of the datasets (green spruce aphid
 abundance, \textit{Ceratium furca} abundance, methane-flux) showed stronger 
-or marginally stronger lower- than upper-tail dependence 
-(Table \ref{tab_multivar_summary}, rows 7-9), whereas leaf-curling plum aphid first flight
-data showed the reverse. Asymmetry results were verified by nonparametric approaches 
-<!--(e.g., Table \ref{tab_multivar_summary}, row 10)-->
-(Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}). Detailed results are in 
-Tables \ref{SM-tab_count_bestcop}-\ref{SM-tab_methane_resamp} and 
-Figs \ref{SM-fig_aphid_count}-\ref{SM-fig_methane}.
-Our 
+lower- than upper-tail dependence 
+(Table \ref{tab_multivar_summary}, rows 6-8), whereas leaf-curling plum aphid first flight
+data showed the reverse. All of the values in Table \ref{tab_multivar_summary}, rows 8 deviated highly significantly
+from what would have been expected under a null hypothesis 
+of equal chances for positive and negative values.
+
+Asymmetry results were generally verified by nonparametric approaches,
+with the exception of the methane-flux data. 
+For instance, the $95\%$ confidence intervals of the
+mean over pairs of sampling locations of the statistic $\cor_l-\cor_u$ were 
+(`r unname(round(stat_aphid_ff$numericdf[10,c(3,4)],3))`)
+for the leaf-curling plum aphid first flight data
+(Table \ref{tab_multivar_summary}, row 9), 
+indicating greater upper-tail dependence, and consistent with
+the results of Table \ref{MT-tab_multivar_summary}, row 8. 
+For the *Ceratium furca* abundance data, confidence intervals were
+(`r unname(round(stat_plankton_north_sea$numericdf[10,c(3,4)],3))`), 
+indicating greater lower-tail dependence, and again 
+consistent with the results of Table \ref{MT-tab_multivar_summary} (row 9).
+For the methane-flux data, confidence intervals were 
+(`r unname(round(stat_methane$numericdf[10,c(3,4)],3))`);
+the asymmetry in tail dependence revealed by the model selection
+results for the methane-flux data was apparently not strong enough to
+also be revealed by the nonparametric analyses. 
+Analogues to Table \ref{tab_count_resamp} are in Tables
+\ref{SM-tab_ff_resamp}-\ref{SM-tab_methane_resamp}.
+Our generally
 affirmative, empirically based answer to Q1 for the multivariate datasets is 

### 2019-09-30 SupportingInformation.Rmd
just about done editing Q1 methods and results for this round, just want to go through it one more time after Shyamolina makes some changes I requested in an email

@@
 ```
-
-# Verifying Q1 results for multivariate datasets using nonparametric statistics \label{Q1multvarnonparam}
-
-<!--***DAN: I'm inclined to eliminate this appendix and put the material
-in the main text-->
-
-For the green spruce aphid abundance data, mean values across all pairs of 
-locations of the Spearman and Kendall correlations and the statistics 
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ were positive and
-confidence intervals excluded $0$ (Table \ref{tab_count_resamp}). 
-Mean values of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
-also were always positive and had confidence intervals that excluded $0$ 
-(Table \ref{tab_count_resamp}), confirming the results in the main text on 
-asymmetry of tail dependence and 
-stronger lower- than upper-tail dependence for this dataset. Each of the statistics 
-was plotted against geographic (great-circle) distances between sampling locations, 
-across all pairs of sampling locations (Fig. \ref{fig_aphid_count}).
-
-Asymmetry of tail dependence was also verified by our nonparametric methods for
-the other multivariate datasets. For instance, the $95\%$ confidence intervals of the
-mean over pairs of sampling locations of the statistic $\Ps_l-\Ps_u$ were 
-(`r unname(round(stat_aphid_ff$numericdf[10,c(3,4)],3))`)
-for the leaf-curling plum aphid first flight data, 
-indicating greater upper-tail dependence, and consistent with
-the results of Table \ref{MT-tab_multivar_summary} (row 9). 
-For the *Ceratium furca* abundance data, confidence intervals were
-(`r unname(round(stat_plankton_north_sea$numericdf[10,c(3,4)],3))`), 
-indicating greater lower-tail dependence, and again 
-consistent with the results of Table \ref{MT-tab_multivar_summary} (row 9).
-For the methane-flux data, confidence intervals were 
-(`r unname(round(stat_methane$numericdf[10,c(3,4)],3))`).
 

### 2019-10-01 Paper.Rmd
Tables fixed, autolink added

@@
 knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
-options(scipen = 1, digits = 5) #This option round all numbers appeared in the inline r code upto 5th digit
+options(scipen = 1, digits = 2) #This option round all numbers appeared in the inline r code upto specified digit
 #seed<-101
@@
 RES_aphid_count<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_count_species_10.RDS")
+#------------ also reading some extra stuffs -------
+stat_aphid_count<-readRDS("./Results/stat_results/stat_aphid_count/stat_aphid_count_sp_10.RDS")
+stat_aphid_ff<-readRDS("./Results/stat_results/stat_aphid_ff/stat_aphid_ff_sp_11.RDS")
+stat_plankton_north_sea<-readRDS("./Results/stat_results/stat_plankton_north_sea/stat_plankton_north_sea_sp_16.RDS")
+stat_methane<-readRDS("./Results/stat_results/stat_methane/stat_methane.RDS")
+tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
 ```
@@
 lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
-`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the 
+`r BivMS_res_Loecke$InfCritRes$copname[which.min(BivMS_res_Loecke$InfCritRes$AIC)]` was the 
 minimum-AIC copula, with AIC `r round(min(BivMS_res_Loecke$InfCritRes$AIC),2)`, whereas the 
-normal copula had a much higher AIC, $-6333.16$, 
-<!--***DAN: Shyamolina, please make this an autolink. Please make 
-sure to round to two digits after the decimal, as with the AIC best value
-presented just above.-->
-answering the 
+normal copula had a much higher AIC, `r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)`, answering the 
 first part of Q1 for these data. Recall that AIC differences of $2$
@@
 Best-fitting (lowest-AIC) copulas were non-normal
-for the large majority of location pairs (72 of 90 pairs), 
-<!--***DAN: Shyamolina, please make these numbers autolinks-->
+for the large majority of location pairs (`r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` of `r tab_multivar_summary$Green_spruce_aphid_count$num.pairs` pairs), 
 and the AIC
 of the normal copula minus the minimum AIC, averaged across 
-location pairs, was $2.732$.
-<!--***DAN: Shyamolina, please make this an autolink, so the value is
-automatically filled in even if it changes-->
-Whereas $2.732$
-<!--***DAN: Shyamolna, again, autolink, please-->
+location pairs, was `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`.
+Whereas `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`
 would be only a marginally meaningful AIC difference for a single
@@
 pairs. Relatedly, and illustrating the concept here, 
-the chances of getting 72 or more 
+the chances of getting `r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` or more 
 non-normal location pairs if the chances were equal of getting
@@
 produce the same result as the pair
-(B,A)) is low, $`r 1-pbinom(72/2-1,45,.5)`$. 
-<!--***DAN: Shyamolina, please replace the 72 in the above expression
-by an autolink, and make this show
-3 significant digits, i.e., it should show 3.29e-5, not 3.287332e-05-->
+(B,A)) is low, 
+$`r 1-pbinom(tab_multivar_summary$Green_spruce_aphid_count$num.non_N/2-1,45,0.5)`$. 
 These results help answer the first part of Q1: non-normal
@@
 these data. Model-averaged lower- and upper-tail
-dependence statistics had $2.5$th and $97.5$th quantiles
-$(0.123,0.721)$ and $(0.002,0.41)$, respectively,
-<!--***DAN: Shyamolina, autolinks pls -->
+dependence statistics had $2.5^{th}$ and $97.5^{th}$ quantiles
+(`r tab_multivar_summary$Green_spruce_aphid_count$LT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$LT_0.975_CI`) and 
+(`r tab_multivar_summary$Green_spruce_aphid_count$UT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$UT_0.975_CI`), respectively,
 thereby commonly differing 
@@
 for positive and negative values (and again accounting for the fact that location pairs (A,B) and (B,A) will show the same result) was
-again low, $`r 1-pbinom(86/2-1,45,.5)`$. 
-<!--***DAN: Shyamolina, please autolink this, and use 3 sig figs as 
-before-->
+again low, 
+$`r 1-pbinom(sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)/2-1,45,0.5)`$.
 Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
@@
 the table will have moved from sup mat to main text.-->
+<!--Show the statistical results in a table for green spruce aphid abundance data-->
+
+```{r tab_count_resamp,echo=F,results="markup"}
+s<-stat_aphid_count$numericdf[,c(3,2,4)]
+sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
+                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
+s<-cbind(sc,s)
+knitr::kable(s, digits = 3,
+             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
+             format = "latex",linesep = "", escape = F,align="c",
+             caption.short="Results for nonparametric statistics, green spruce aphid abundance data",
+             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for green spruce aphid abundance data.\\label{tab_count_resamp}",
+             booktabs=TRUE)
+```
 
@@
 indicating greater upper-tail dependence, and consistent with
-the results of Table \ref{MT-tab_multivar_summary}, row 8. 
+the results of Table \ref{tab_multivar_summary}, row 8. 
 For the *Ceratium furca* abundance data, confidence intervals were
@@
 indicating greater lower-tail dependence, and again 
-consistent with the results of Table \ref{MT-tab_multivar_summary} (row 9).
+consistent with the results of Table \ref{tab_multivar_summary} (row 9).
 For the methane-flux data, confidence intervals were 
@@
 ```{r tab_multivar_summaryl, echo=F, results='asis',message=F}
-tab_multivar_summary<-readRDS("./Results/tab_multivar_summary.RDS")
-tab1<-tab_multivar_summary[c(1:7,9,11,12,14,16),] # see carefully if you rearrange tab_multivar_summary's row
+tab_multivar_summaryl<-tab_multivar_summary[-c(4,12:15),] #exclude extra rows for AICw, Corl, Coru
+
+tab1<-tab_multivar_summaryl[c(6,8,12),] # see carefully if you rearrange tab_multivar_summary's row
                                                    #in suppmat this will change
-tab2<-tab_multivar_summary[c(8,10,13,15,17),]  # see carefully if you rearrange tab_multivar_summary's row
+tab2<-tab_multivar_summaryl[c(7,9,13),]  # see carefully if you rearrange tab_multivar_summary's row
                                                    #in suppmat this will change
 
-two_tab_in1 <- as.data.frame(do.call(cbind, lapply(1:ncol(tab1), function(i) paste0(" (",tab1[c(7:8,10:12),i], ", ",tab2[ , i], ")") )))   # see carefully if you rearrange tab_multivar_summary's row in suppmat this will change
-
-colnames(two_tab_in1)<-colnames(tab_multivar_summary)
-rownames(two_tab_in1)<-c("CI_LT","CI_UT","CI_Pl","CI_Pu","CI_Pl-Pu")
-
-tab_multivar_summary<-rbind(tab_multivar_summary[c(1:3,5,6),],
-                            two_tab_in1[1:2,], #CI of LT, CI of UT
-                            tab_multivar_summary[11,],
-                            two_tab_in1[5,] #CI of Pl-Pu
+two_tab_in1 <- as.data.frame(do.call(cbind, lapply(1:ncol(tab1), function(i) paste0(" (",tab1[,i], ", ", tab2[ , i], ")") )))   # see carefully if you rearrange tab_multivar_summary's row in suppmat this will change
+
+two_tab_in1_again <- as.data.frame(do.call(cbind, lapply(1:ncol(two_tab_in1), 
+                                                         function(i) paste0(tab_multivar_summaryl[11,i], ", ", two_tab_in1[3,i]) )))
+
+colnames(two_tab_in1)<-colnames(two_tab_in1_again)<-colnames(tab_multivar_summaryl)
+
+compiled_tab<-rbind(two_tab_in1[1:2,],two_tab_in1_again)
+rownames(compiled_tab)<-c("CI_LT","CI_UT","Avg_CI_Corl-Coru")
+
+tab_multivar_summary_ed<-rbind(tab_multivar_summaryl[c(1:5),],
+                            compiled_tab[1:2,], #CI of LT, CI of UT
+                            tab_multivar_summaryl[10,],
+                            compiled_tab[3,] #avg and CI of Corl-Coru
                            )
 
-row.names(tab_multivar_summary)<-NULL
+row.names(tab_multivar_summary_ed)<-NULL
 library(kableExtra)
@@
            "3. Number of pairs with non-normal copula as best fit",
-           #"4. Avg. of normal-AICw minus bestfit-copula AICw across location pairs",
            "4. Average of AIC differences (normal AIC minus min AIC) across location pairs",
@@
            "8. Percent pairs showing stronger lower tail dependence",
-           "9. 2.5$^{th}$ and 97.5$^{th}$ quantiles of $<$P$_{l}$ - P${_u}$$>$ across all location pair")
+           "9. Mean with 2.5$^{th}$ and 97.5$^{th}$ quantiles of Cor$_{l}$ - Cor${_u}$ across all location pair")
            )
 
-tab_multivar_summary<-cbind(dt,tab_multivar_summary)
-
-knitr::kable(tab_multivar_summary[1:8,], 
+tab_multivar_summary_ed<-cbind(dt,tab_multivar_summary_ed)
+
+knitr::kable(tab_multivar_summary_ed, 
              format="latex", escape = F,
@@
                         "\\textit{Ceratium furca} abundance","Methane-flux")
-             )%>%column_spec(2:5,width="2.4cm")%>%
+             )%>%column_spec(2:5,width="2.5cm")%>%
              column_spec(1, width = "6cm")%>%kable_styling(latex_options = "basic",position = "center")

### 2019-10-01 SupportingInformation.Rmd
Tables fixed, autolink added

@@
 
-<!--Show the statistical results in a table for green spruce aphid abundance data-->
-```{r tab_count_resamp,echo=F,results="markup"}
-s<-stat_aphid_count$numericdf[,c(3,2,4)]
-sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
-                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
-s<-cbind(sc,s)
-knitr::kable(s, digits = 3,
-             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
-             format = "latex",linesep = "", escape = F,align="c",
-             caption.short="Results for nonparametric statistics, green spruce aphid abundance data",
-             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for green spruce aphid abundance data.\\label{tab_count_resamp}",
-             booktabs=TRUE)
-```
-
 <!--***DAN: decided to remove this for AER
@@
 
-  m<-matrix(NA,1,17)
-  colnames(m)<-c("num.pairs","num.non-indep","num.non-N",
+  m<-matrix(NA,1,18)
+  colnames(m)<-c("num.pairs","num.non.indep","num.non_N",
                  "Avg.N_minus_bestfit.AICw","Avg.N_minus_bestfit.AIC","num.good.fit",
-                 "0.025-CI.LT","0.975-CI.LT","0.025-CI.UT","0.975-CI.UT",
-                 "percent.LT-UT>0",
-                 "0.025-CI.Pl","0.975-CI-Pl","0.025-CI.Pu","0.975-CI.Pu","0.025-CI.Pl-Pu","0.975-CI.Pl-Pu")
+                 "LT_0.025_CI","LT_0.975_CI","UT_0.025_CI","UT_0.975_CI",
+                 "percent.LT_minus_UT_gt0",
+                 "Corl_0.025_CI","Corl_0.975_CI","Coru_0.025_CI","Coru_0.975_CI",
+                 "avg.Corl_minus_Coru","Corl_minus_Coru_0.025_CI","Corl_minus_Coru_0.975_CI")
   m<-as.data.frame(m)
@@
   m$num.pairs<-(d^2)-d 
-  m$`num.non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
+  m$num.non.indep<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
 
@@
   bcop[bcop==Inf]<-NaN # replacing Independent pairs with NaN
-  m$`num.non-N`<-sum(bcop!=1,na.rm = T)
+  m$num.non_N<-sum(bcop!=1,na.rm = T)
 
@@
   CI_LT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
-  m$`0.025-CI.LT`<-round(CI_LT[1],3)
-  m$`0.975-CI.LT`<-round(CI_LT[2],3)
+  m$LT_0.025_CI<-round(CI_LT[1],3)
+  m$LT_0.975_CI<-round(CI_LT[2],3)
   t<-r$UTdep_AICw[is.finite(r$UTdep_AICw)]
   CI_UT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
-  m$`0.025-CI.UT`<-round(CI_UT[1],3)
-  m$`0.975-CI.UT`<-round(CI_UT[2],3)
-  m$`percent.LT-UT>0`<-round((sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)/((d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)))*100,2)
-  m$`0.025-CI.Pl`<-round(rs$numericdf[5,3],3)
-  m$`0.975-CI-Pl`<-round(rs$numericdf[5,4],3)
-  m$`0.025-CI.Pu`<-round(rs$numericdf[6,3],3)
-  m$`0.975-CI.Pu`<-round(rs$numericdf[6,4],3)
-  m$`0.025-CI.Pl-Pu`<-round(rs$numericdf[10,3],3)
-  m$`0.975-CI.Pl-Pu`<-list(round(rs$numericdf[10,4],3))
+  m$UT_0.025_CI<-round(CI_UT[1],3)
+  m$UT_0.975_CI<-round(CI_UT[2],3)
+  m$percent.LT_minus_UT_gt0<-round((sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)/((d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)))*100,2)
+  m$Corl_0.025_CI<-round(rs$numericdf[3,3],3)
+  m$Corl_0.975_CI<-round(rs$numericdf[3,4],3)
+  m$Coru_0.025_CI<-round(rs$numericdf[4,3],3)
+  m$Coru_0.975_CI<-round(rs$numericdf[4,4],3)
+  m$avg.Corl_minus_Coru<-round(rs$numericdf[9,2],3)
+  m$Corl_minus_Coru_0.025_CI<-round(rs$numericdf[9,3],3)
+  m$Corl_minus_Coru_0.975_CI<-list(round(rs$numericdf[9,4],3))
   #m$CI_Pl<-list(c(round(rs$numericdf[5,3],3),round(rs$numericdf[5,4],3)))
@@
 tab_multivar_summary<-cbind(m1,m2,m3,m4)
-colnames(tab_multivar_summary)<-c("Green-spruce aphid_count","Leaf-plum curling aphid_firstflight","C. furca_count","Methane-flux")
+colnames(tab_multivar_summary)<-c("Green_spruce_aphid_count","Leaf_plum_curling_aphid_firstflight","C_furca_count","Methane_flux")
 saveRDS(tab_multivar_summary,"./Results/tab_multivar_summary.RDS")

### 2019-10-01 Paper.Rmd
finished editing (for now) the Q1 sections, but Shyanolina needs to fix a couple things

@@
 of these families provided in the `VineCopula` package for the R programming
-language. See Figs \ref{figped_N}-\ref{figped_BB1} and \ref{SM-figped_SC}-\ref{SM-figped_SBB8} 
+language. See Figs \ref{figped_N}-\ref{figped_BB1} and \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
 for visual depictions of the pdfs of these copulas and how
@@
 
-<!--***DAN: Notes on what needs to be done to this section, so far only getting up through the presentation of the bivariate results:
-1) be more nuanced in the presentation, don't give the appearance of sweeping results that dont agree with the general conclusions under the rug
-DONE 2) Add AIC and BIC columsn to table S3, and all analogous figures (e.g., S5, S7, S9, etc).
-Also reduce the number of digits shown after the decimal to 2 for AIC and BIC and 4 for the other columns.
-3) Move Table S3 to an appropriate place in the main text. Choose the place based on where it is first cited in the main text.
-DONE 4) Sort the rows of Table S2 (and all analogous tables) AIC, putting the lowest
-AIC at the top. This should mean the highest AICw is at the top.
-DONE 5a) move some of the content from appendices S6 and S7 into methods, 
-5b) and move fig S21 and table S4 into main text. These are non-parametric results for
-soil C and N. So we will be presenting the nonparam results more comprehensively in the main text.
-6) That means you also need to summarize the nonparm results for the other 3 bivariate systems in Table 2.
-How to do this?
-Maybe put the cor results only, for Spearman only (or Kendall, not sure which to choose) into Table 2.
-7) Probably ought to combine tables S6, S8, S10 into one table, and fix it so
-whenever we currently cite one of those, change it so we cite the combined table instead.
-DONE 8) The nonparam stats are not estimators of tail dependence, they measure 
-something else (though related). Make it clear we know this. It turns out there are estimators
-of tail dependence. Do we want to mention that?
--->
-
 \begin{figure}[!h]
@@
 in the right-most box in the middle row.
-<!--***DAN: you may need to change the box around the Cedar Creek data-->
+<!--***DAN: you need to change the box around the Cedar Creek data-->
 
@@
 (Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
-difference between these fits was not meaningfully large (Table \ref{SM-tab_cc2000fit}).
+difference between these fits was not meaningfully large.
 Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
@@
 Fig. \ref{fig_soilCN_nonparam} for the bivariate datasets other than the soil C and N
-data, and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
+data, see Tables \ref{SM-tab_bmr_birds_fit}-\ref{SM-tab_cc2000fit} for analogues to Table \ref{tab_soilCNfit},
+and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
 The first three datasets had stronger upper- 
@@
 these data. Model-averaged lower- and upper-tail
-dependence statistics had $2.5^{th}$ and $97.5^{th}$ quantiles
+dependence statistics had $2.5^{th}$ and $97.5^{th}$ quantiles (across
+location pairs)
 (`r tab_multivar_summary$Green_spruce_aphid_count$LT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$LT_0.975_CI`) and 
@@
 
-<!--***DAN: Shyamolina, can you please insert code to make what
-is currently Table S7 appear here? Please also make it disappear from
-the sup mat. I want to move it from there to here, please. I am
-talking about the table with label tab_count_resamp. Please also
-search both the main text and the supp mat for that label, and 
-fix references - you may have to fix some prefixes (SM-, MT-), since
-the table will have moved from sup mat to main text.-->
-<!--Show the statistical results in a table for green spruce aphid abundance data-->
-
-```{r tab_count_resamp,echo=F,results="markup"}
-s<-stat_aphid_count$numericdf[,c(3,2,4)]
-sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
-                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
-s<-cbind(sc,s)
-knitr::kable(s, digits = 3,
-             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
-             format = "latex",linesep = "", escape = F,align="c",
-             caption.short="Results for nonparametric statistics, green spruce aphid abundance data",
-             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for green spruce aphid abundance data.\\label{tab_count_resamp}",
-             booktabs=TRUE)
-```
-
 <!--multivariate summary-->
@@
 conclusions that non-normal copula structure, non-normal tail dependence, and asymmetric tail 
-dependence were common, though not universal, answering 
+associations were common, though not universal, answering 
 Q1 in the affirmative for these data. 
@@
 (`r unname(round(stat_aphid_ff$numericdf[10,c(3,4)],3))`)
+<!--***DAN: Shyamolina, these do not match with the value in table 6!
+Please check BOTH values and make them correct, and make them agree!
+You seemed to have used the values for the P statistic by mistake.-->
 for the leaf-curling plum aphid first flight data
@@
 (`r unname(round(stat_plankton_north_sea$numericdf[10,c(3,4)],3))`),
+<!--***DAN: Shyamolina, these do not match with the value in table 6!
+Please check BOTH values and make them correct, and make them agree!
+You seemed to have used the values for the P statistic by mistake.-->
 indicating greater lower-tail dependence, and again 
@@
 (`r unname(round(stat_methane$numericdf[10,c(3,4)],3))`);
+<!--***DAN: Shyamolina, these do not match with the value in table 6!
+Please check BOTH values and make them correct, and make them agree!
+You seemed to have used the values for the P statistic by mistake.-->
 the asymmetry in tail dependence revealed by the model selection
@@
 \textit{Ceratium furca} (results not shown); results supported the conclusion that 
-non-normal copula structure and tail dependence, and asymmetry of tail dependence are 
+non-normal copula structure and tail dependence, and asymmetry of tail associations, are 
 common.
@@
              column_spec(1, width = "8.4cm")%>%kable_styling(latex_options = "basic",position = "center")
+```
+
+```{r tab_count_resamp,echo=F,results="markup"}
+s<-stat_aphid_count$numericdf[,c(3,2,4)]
+sc<-as.data.frame(c("Spearman","Kendall","cor$_{l}$","cor$_{u}$","P$_{l}$","P$_{u}$","D$_{l}^2$","D$_{u}^2$",
+                    "cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$"))
+s<-cbind(sc,s)
+knitr::kable(s, digits = 3,
+             col.names=c("","2.5$^{th}$ quantile", "Mean", "97.5$^{th}$ quantile"),
+             format = "latex",linesep = "", escape = F,align="c",
+             caption.short="Results for nonparametric statistics, green spruce aphid abundance data",
+             caption = "Average values of statistics across all locations pairs and confidence intervals based on spatial resampling for green spruce aphid abundance data.\\label{tab_count_resamp}",
+             booktabs=TRUE)
 ```
@@
 
-<!--***DAN: could remove the next para if you need to save words-->
 <!--a para on stability of the total biomass time series in a community, and skewness - sets up Shya's later paper as future work-->

### 2019-10-01 Paper.Rmd
error fixed in table

@@
 knitr::opts_chunk$set(echo = TRUE, fig.pos = "H")
-options(scipen = 1, digits = 2) #This option round all numbers appeared in the inline r code upto specified digit
+options(scipen = 1, digits = 3) #This option round all numbers appeared in the inline r code upto specified digit
 #seed<-101
@@
 mean over pairs of sampling locations of the statistic $\cor_l-\cor_u$ were 
-(`r unname(round(stat_aphid_ff$numericdf[10,c(3,4)],3))`)
-<!--***DAN: Shyamolina, these do not match with the value in table 6!
-Please check BOTH values and make them correct, and make them agree!
-You seemed to have used the values for the P statistic by mistake.-->
+(`r unname(round(stat_aphid_ff$numericdf[9,c(3,4)],3))`)
 for the leaf-curling plum aphid first flight data
@@
 For the *Ceratium furca* abundance data, confidence intervals were
-(`r unname(round(stat_plankton_north_sea$numericdf[10,c(3,4)],3))`),
-<!--***DAN: Shyamolina, these do not match with the value in table 6!
-Please check BOTH values and make them correct, and make them agree!
-You seemed to have used the values for the P statistic by mistake.-->
+(`r unname(round(stat_plankton_north_sea$numericdf[9,c(3,4)],3))`),
 indicating greater lower-tail dependence, and again 
@@
 For the methane-flux data, confidence intervals were 
-(`r unname(round(stat_methane$numericdf[10,c(3,4)],3))`);
-<!--***DAN: Shyamolina, these do not match with the value in table 6!
-Please check BOTH values and make them correct, and make them agree!
-You seemed to have used the values for the P statistic by mistake.-->
+(`r unname(round(stat_methane$numericdf[9,c(3,4)],3))`);
 the asymmetry in tail dependence revealed by the model selection
@@
            "8. Percent pairs showing stronger lower tail dependence",
-           "9. Mean with 2.5$^{th}$ and 97.5$^{th}$ quantiles of Cor$_{l}$ - Cor${_u}$ across all location pair")
+           "9. Mean with 2.5$^{th}$ and 97.5$^{th}$ quantiles of cor$_{l}$ - cor${_u}$ across all location pair")
            )

### 2019-10-01 Paper.Rmd
comments in prep for revisions for the Q2 sections#

--- 2019-10-01 Paper.Rmd
@@
 
-Having demonstrated that non-normal copula structure and asymmetric tail dependence are 
+Having demonstrated that non-normal copula structure and asymmetric tail associations are 
 common in ecological variables, we 
@@
 by which these phenomena may arise. 
+<!--***DAN: Maybe insert some text indicating these are first, exploratory models, and are not intended to comprehensively explore
+a mechanism (that can be done in future work), but rather to get a sense
+for whether observed copula structure may be partly due to the
+mechanism, and whether the mechanism is worthy of more comprehensive
+exploration in that connection in future work. One way these are noncomprehensive
+explorations is that I chose particular limited sets of parameters.-->
 The first mechanism relates to the ideas in the Introduction about 
@@
 $g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
-$g_1$ represents environmental effects that occur only below a threshold of $\epsilon=0$, 
-and $g_2$ represents effects that occur only above a threshold of $\epsilon=0$.
+$g_1$ represents environmental effects that negatively impact populations,
+and that occur only below a threshold of $\epsilon=0$; 
+and $g_2$ represents effects that positively impact populations and 
+that occur only above a threshold of $\epsilon=0$.
 Negative values of $b$ correspond to overcompensating dynamics 
@@
 steps. We applied our model selection algorithms and 
-non-parametric statistics (see section on methods for Q1) to these 
+nonparametric statistics (see section on methods for Q1) to these 
 outputs to discover if the model could
@@
 $E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
-and $(\delta_1,\delta_2)$ have normal copula structure, the Moran 
+and $(\delta_1,\delta_2)$ have normal copula structure (they were drawn
+from bivariate normal distributions), the Moran 
 mechanism analyzed below 
 does not operate here.
+<!--***DAN: You may also want to have some additional commentary in the Discussion on
+how the models and related models can be explored more comprehensively. In fact you may already have some of that, though you may want to beef it up. You may
+want to insert a point at the end of the Methids section here again emphasizing
+that these are simple models for an initial exploration, and we discuss in the 
+Discussion how...-->
 
@@
 is that, when it is strong, it brings the nonlinearities of the model into play. 
-
-For each of the model setups above, for each of several copula families, for each $\tau=0.1,0.2,\ldots,0.9$, and for each of
+<!--***DAN: Didn't we use another value of r as well, and present results
+on it? If so, probably should mention it here, in part so the exploration
+seems a bit broader?-->
+
+For each of the model setups above, for each of several copula families, 
+<!--***DAN: I currently use C, SC, G, SG, J, SJ. Do I really need all those? 
+Surely it would be enough to just do C and SC and then state that the others 
+give similar results? And if I move one of the result figure from this section
+into the main text, that will mean only one such fig in the sup mat, at least
+for AR(1)? I mean, right now we have 18 figure in sup mat from this analysis,
+and only part of one of them in main text! See if the results are actually similar across
+S/SC, G/SG, J/SJ, and if so consider keeping only the S/SC results in either the
+main text or the sup mat, and just stating the others were similar. Then you would presumably
+move the full C fig into the main text, replacing the current partial version.-->
+for each $\tau=0.1,0.2,\ldots,0.9$, and for each of
 $50$ replicate simulations, we generated 5000 noise pairs 
@@
 Values were plotted against $\tau$ for noises and populations. If the 
-hypothesis was correct that characteristics of 
+hypothesis from the Introduction was correct that characteristics of 
 the copula structure of spatial dependence in an
@@
 
+<!--***DAN: How much of the material from Appendix S8 should be moved into
+the main text? I am thinking a fair amount of it. Currrently there
+is really no results on that topic at all in the main text, and that seems
+a bit ridiculous and probably is why the referee said we treated this 
+material in a "breezy" way.-->
 The next mechanism we investigated is evolutionary, and pertains to bivariate 
@@
 asymmetric tail dependence across space in ecological variables.
+<!--***DAN: move the following to the end of this part-->
 This result is represented in Fig. \ref{MasterFigure} as the solid box around 
@@
 present in this section are theoretical.
+<!--***DAN: move the above to the end of this part-->
 For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
@@
 \ref{SM-fig_ms_asym_b_-0.5}).
+<!--***DAN: combin all 3 sup mat figs on this subject into one sup mat fig-->
 Lower-tail (respectively, upper-tail) spatial dependence in 
@@
 reflected the visually apparent asymmetry (Tables \ref{SM-tab_stat_npa_asym}). 
+<!--***DAN: Simplify as follows, maybe:
+1) get rid of table S10 (you will have to change the methods so it no longer says we
+do out model selection analysis here)
+2) simplify table S11 so it only shows the cor results (also means a change to methods)
+and then maybe bring the resulting table to main text?
+-->
 
@@
 variable was largely confirmed; 
+<!--***DAN: You are using language about confirming hypotheses, here and elsewhere,
+but this may rub some people the wrong way because we have not confirmed anything
+empirically, we are just doing models, and only a cursory exploration of each
+model, at that! So change the language.-->
 however, similarities between environmental- and 
@@
 Thus this is a mechanism that can produce non-normal copula structure and
-asymmetric tail dependence across space in ecological variables.
+asymmetric tail dependence 
+<!--****DAN: Don't firget to replace @tail dependence@ by @tail asociation@ as appropriate-->
+across space in ecological variables.
+<!--***DAN: move the below to end of this part-->
 This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
 "Moran effects" and the arrows labelled "B".
+<!--***DAN: move the above to end of this part-->
 For simulations using AR(1) models, copula 

### 2019-10-01 SupportingInformation.Rmd
finished editing (for now) the Q1 sections, but Shyanolina needs to fix a couple things
comments in prep for revisions for the Q2 sections#

@@
 
-<!--embedding plot with sampling locations : soilCN, aphids, planktons-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm,height=16 cm]{./Results/SoilCNmapUSA.pdf} 
-\caption[Sampling locations for C and N data]{Locations of central pedons sampled 
-for soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil 
-surface) at 5907 sites across the coterminous United States. \label{fig_SoilCNmapUSA}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm,height=16 cm]{./Results/aphid_map_plot.pdf} 
-\caption[Sampling locations for aphid data]{Locations of the 
-Rothamsted Insect Survey 
-suction traps from which data were available are indicated by the red circles, 
-out of which only blue boxes had at least 30 years of data and were 
-used for our study. \label{fig_aphid_map}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=16 cm,height=16 cm]{./Results/plankton_map_plot.pdf} 
-\caption[Sampling regions for CPR plankton data]{CPR data were available from 
-26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ regions on the map) around UK 
-seas (red circles), out of which only blue boxes had at least 45 years of data and
-were used for our study. \label{fig_plankton_map}}
-\end{center}
-\end{figure}
-
-<!--Now the copula figures that were not presented in the main text-->
+<!--survival clayton copula-->
 \begin{figure}[!h]
@@
 
+<!--embedding plot with sampling locations : soilCN, aphids, planktons-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm,height=16 cm]{./Results/SoilCNmapUSA.pdf} 
+\caption[Sampling locations for C and N data]{Locations of central pedons sampled 
+for soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil 
+surface) at 5907 sites across the coterminous United States. \label{fig_SoilCNmapUSA}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm,height=16 cm]{./Results/aphid_map_plot.pdf} 
+\caption[Sampling locations for aphid data]{Locations of the 
+Rothamsted Insect Survey 
+suction traps from which data were available are indicated by the red circles, 
+out of which only blue boxes had at least 30 years of data and were 
+used for our study. \label{fig_aphid_map}}
+\end{center}
+\end{figure}
+
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=16 cm,height=16 cm]{./Results/plankton_map_plot.pdf} 
+\caption[Sampling regions for CPR plankton data]{CPR data were available from 
+26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ regions on the map) around UK 
+seas (red circles), out of which only blue boxes had at least 45 years of data and
+were used for our study. \label{fig_plankton_map}}
+\end{center}
+\end{figure}
+
+<!--Now the copula figures that were not presented in the main text-->
 \begin{figure}[!h]
@@
 \includegraphics[width=7.5in]{./Results/PedagogSuppMat_Joe.pdf}
-\caption[Example J copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example J copulas. K is Kendall correlation; $p$ is the value of the parameter for the J family (it is a 
+\caption[Example Joe copulas]{Log-transformed pdfs (A-E) and samples (F-J) from example Joe copulas. K is Kendall correlation; $p$ is the value of the parameter for the Joe family (it is a 
 one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, 
@@
 \caption[Comparison of copula structure of noise and model output, AR(1) model, 
-Gumbelcopula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+Gumbel copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
 model, Gumbelcopula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
@@
 \caption[Comparison of copula structure of noise and model output, AR(1) model, 
-J copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
-model, J copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
+Joe copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+model, Joe copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
 \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, weak noise, J copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, weak noise, J copula. Same format as Figs 
+model, weak noise, Joe copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, Joe copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, J copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, J copula. Same format as Figs 
+model, strong noise, Joe copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, strong noise, Joe copula. Same format as Figs 
 \ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 

### 2019-10-01 Paper.Rmd
more comments refining what I will next do to Q2 sections

@@
 on it? If so, probably should mention it here, in part so the exploration
-seems a bit broader?-->
+seems a bit broader? Well, we mentioned that we did those and we verbally summarize
+results but we did not show any. So maybe leave things as is in this respect.-->
 
@@
 main text or the sup mat, and just stating the others were similar. Then you would presumably
-move the full C fig into the main text, replacing the current partial version.-->
+move the full C fig into the main text, replacing the current partial version.
+So instead of having 18 figs in sup mat and half a fig in main text, you would
+have 5 figs in sup mat and 1 in main text, and then you would just say "we did the
+same thing for G, SG, J, SJ and got similar results."
+So let's check, are results, broadly speaking, independent of which of the 6 copulas
+I used?:
+A) for AR(1) models, yes
+B) for Ricker, weak noise, yes
+C) for Ricker, strong noise, its more complicated. It's generally true that you get
+more approximate agreement between noise and pop stats, with some rather glaring disagreement.
+But which stats disagree the most depends on which cop we used!
+So I guess we can still just display the C and SC results, as planned above, and explain
+in text.
+-->
 for each $\tau=0.1,0.2,\ldots,0.9$, and for each of
@@
 \ref{SM-fig_ms_asym_b_-0.5}).
-<!--***DAN: combin all 3 sup mat figs on this subject into one sup mat fig-->
+<!--***DAN: Shyamolina, can you please combine Figs S24-S26 into one single figure
+with 6 panels? Please also search for citations in the main text and sup mat that 
+used to refer to the three figures and replace them with a citation to the
+new, combined figure.-->
 Lower-tail (respectively, upper-tail) spatial dependence in 
@@
 1) get rid of table S10 (you will have to change the methods so it no longer says we
-do out model selection analysis here)
+do our model selection analysis here)
 2) simplify table S11 so it only shows the cor results (also means a change to methods)
@@
 -->
+
+<!--***DAN: Shyamolina, can you please simplify Table S11 so it only shows the cor
+results? That also means the third column will be unnecessary. Please do this as follows:
+1) keep but comment out the code that produces the old table; 2) write new code (presumably
+starting from a copy of the old code) that makes the new, simpler version of the table. 
+That will make it very easy to produce the old version again if we ever want to.-->
 

### 2019-10-01 Paper.Rmd
table and figure simplified as suggested by DCR

@@
 For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
-(Fig. \ref{SM-fig_ms_asym_b_-0.1}), 
+(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}A,B), 
 for both $g_1$ and $g_2$,
@@
 $b=\pm 0.5$, asymmetry was weaker but still apparent
-(Figs \ref{SM-fig_ms_asym_b_0.5} - 
-\ref{SM-fig_ms_asym_b_-0.5}).
+(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}C-F).
 <!--***DAN: Shyamolina, can you please combine Figs S24-S26 into one single figure
@@
 used to refer to the three figures and replace them with a citation to the
-new, combined figure.-->
+new, combined figure.
+***Shya: Done.-->
 Lower-tail (respectively, upper-tail) spatial dependence in 
@@
 starting from a copy of the old code) that makes the new, simpler version of the table. 
-That will make it very easy to produce the old version again if we ever want to.-->
+That will make it very easy to produce the old version again if we ever want to.
+***Shya: Done.-->
 

### 2019-10-01 SupportingInformation.Rmd
table and figure simplified as suggested by DCR

@@
      cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
-text(0.1,0.95,"A",cex=2)
+text(0.1,0.95,"C",cex=2)
 pop<-ans_b_0.5$pop_d
@@
      cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
-text(0.1,0.95,"B",cex=2)
+text(0.1,0.95,"D",cex=2)
 par(op)
@@
      cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
-text(0.1,0.95,"A",cex=2)
+text(0.1,0.95,"E",cex=2)
 pop<-ans_b_neg0.5$pop_d
@@
      cex.axis=1,col=rgb(0,0,0,0.2),pch=20,cex=0.6)
-text(0.1,0.95,"B",cex=2)
+text(0.1,0.95,"F",cex=2)
 par(op)
@@
 <!--adding plots from asymmetric dependence analysis results-->
+
 \begin{figure}[!h]
@@
 \includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.1_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$.\label{fig_ms_asym_b_-0.1}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
 \includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.5_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=0.5$.\label{fig_ms_asym_b_0.5}}
-\end{center}
-\end{figure}
-
-\begin{figure}[!h]
-\begin{center}
 \includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.5$.\label{fig_ms_asym_b_-0.5}}
+\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$ (A-B), $b=0.5$ (C-D) and $b=-0.5$ (E-F).\label{fig_ms_asym_b_-0.1_0.5_-0.5}}
 \end{center}
@@
 the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
-\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to the model 
+\\ref{fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to the model 
 selection procedures detailed in the 
@@
 
-<!--Displaying summary table with non-parametric approach results of asymmetric dependence analysis-->
+<!--*** Keeping old complex table for future use:
+Displaying summary table with non-parametric approach results of asymmetric dependence analysis
 ```{r tab_stat_npa_asym,echo=F,results="markup"}
@@
 the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
-\\ref{fig_ms_asym_b_-0.1}-\\ref{fig_ms_asym_b_-0.5}) were subjected to nonparamtric statistical 
+\\ref{fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to nonparamtric statistical 
 analyses decribed in the methods section for Q1. The listed statistic was computed for the
+output of the model with the indicated b and g, and the value was compared to 1000 values
+of the same statistic computed on surrogate time series randomized to have normal copula structure
+(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
+given statistic on the data was less (respectively, more) than its value on $X$ of the 
+surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
+indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
+lower-tail dependence compared to a normal-copula null hypothesis. `FS' stands for `first slice'
+and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
+0.9 to 1.\\label{tab_stat_npa_asym}")
+```
+-->
+
+<!---Simpler version: Displaying summary table with non-parametric approach results of asymmetric dependence analysis--->
+```{r tab_stat_npa_asym_simple,echo=F,results="markup"}
+library(kableExtra)
+
+tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
+tab_stat_npa_asym<-c()
+for(i in c(1:8)){
+  temp<-make_table_stat_fn(data=tab_asym[[i]])
+  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
+}
+
+ind_corstat<-which(tab_stat_npa_asym$Statistic %in% c("cor$_{FS}$-cor$_{LS}$"))
+tab_corstat_asym<-tab_stat_npa_asym[ind_corstat,]
+tab_corstat_asym<-tab_corstat_asym[,-1]
+
+#gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, #g$_{1}$","b=-0.5, g$_{2}$")
+#tab_stat_npa_asym<-cbind(gtext,tab_corstat_asym)
+
+tab_corstat_asym<-cbind(b=c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),
+  g=c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),tab_corstat_asym)
+
+rownames(tab_corstat_asym)<-NULL
+
+knitr::kable(tab_corstat_asym, booktabs = T, 
+             format = "latex",linesep = "",escape=F, 
+             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
+             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
+The output of simulations from the first model detailed in 
+the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
+\\ref{fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to nonparamtric statistical 
+analyses decribed in the methods section for Q1. The statistic ($cor_{FS} - cor_{LS}$) was computed for the
 output of the model with the indicated b and g, and the value was compared to 1000 values

### 2019-10-03 Paper.Rmd
edits to Q2 sections, through first model methods and results

@@
 by which these phenomena may arise. 
-<!--***DAN: Maybe insert some text indicating these are first, exploratory models, and are not intended to comprehensively explore
-a mechanism (that can be done in future work), but rather to get a sense
-for whether observed copula structure may be partly due to the
-mechanism, and whether the mechanism is worthy of more comprehensive
-exploration in that connection in future work. One way these are noncomprehensive
-explorations is that I chose particular limited sets of parameters.-->
+Models presented are initial explorations, only, of whether the proposed mechanisms have the 
+potential to explain observed patterns. As such, simple models were used, 
+and comprehensive explorations of
+model parameter space and alternative model formulations were left for future work.
+
 The first mechanism relates to the ideas in the Introduction about 
@@
 If an environmental variables influences an ecological variable 
-more strongly in one of its tails, we explored whether the ecological
+disproportionately strongly in one of its tails, we explored whether the ecological
 variable could then exhibit asymmetric tail dependence across space.
@@
 time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
-$\cov(\epsilon_1,\epsilon_2)=0.8$.
-The function $g$ describes how the environment $\epsilon_i(t)$ influences
+$\cov(\epsilon_1,\epsilon_2)=0.8$. Thus basic ecological dynamics follow the
+very simple AR(1) formulation, influenced by a "regional" environmental factor which
+is correlated across locatons ($\epsilon$) and by a local factor ($\delta$). 
+The function $g$ describes how the environmental factor $\epsilon_i(t)$ influences
 $E_i(t)$. We used $g$ equal to $g_1$ or $g_2$ in different simulations:
@@
 that occur only above a threshold of $\epsilon=0$.
-Negative values of $b$ correspond to overcompensating dynamics 
+The values of $b$ selected provide a modicum of exploration of the potential
+for ecological dynamics to influence how complex dependencies between
+the $E_i(t)$ may arise: 
+negative values of $b$ correspond to overcompensating dynamics 
 and positive values to undercompensating dynamics; larger $|b|$
@@
 steps and retained the $E_i(t)$ for the final $2500$ time 
-steps. We applied our model selection algorithms and 
-nonparametric statistics (see section on methods for Q1) to these 
+steps. We applied our 
+nonparametric statistic $\cor_{0,0.1}-\cor_{0.9,1}$ 
+and our Spearman- or Kendall-preserving normal-copula 
+surrogates (see section on methods for Q1) to these 
 outputs to discover if the model could
-produce non-normal copula structure and asymmetric tail dependence in the 
-$E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
+produce asymmetric tail associations (and therefore non-normal copula structure)
+between the $E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
 and $(\delta_1,\delta_2)$ have normal copula structure (they were drawn
 from bivariate normal distributions), the Moran 
-mechanism analyzed below 
-does not operate here.
+mechanism analyzed below does not operate here.
 <!--***DAN: You may also want to have some additional commentary in the Discussion on
@@
 Discussion how...-->
+
+<!--***DAN: stopped here in Q2 methods in order to go on to the correspondng
+section of Q2 results-->
 
@@
 Our model with asymmetric environmental effects 
-produces outputs with visually apparent asymmetry 
-of tail dependence, to an extent that depended on the value of $b$;
-thus this is a mechanism that can produce non-normal copula structure and
+produced outputs with visually apparent asymmetry 
+of tail associations between the ecological variables $E_i$ in the two locations, 
+to an extent that depended on the value of $b$.
+For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
+(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}A,B), 
+for both $g_1$ and $g_2$,
+asymmetry of tail association was strong; for 
+$b=\pm 0.5$, asymmetry was weaker but still apparent
+(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}C-F).
+Lower-tail (respectively, upper-tail) spatial associations in the 
+effects of noise ($g_1$, respectively, $g_2$) 
+produced lower-tail (respectively, upper-tail) spatial 
+associations in the ecological variable, $E_i$.
+Results using our statistic $\cor_{0,0.1}-\cor_{0.9,1}$ 
+strongly reflected the visually apparent asymmetry (Tables \ref{tab_stat_npa_asym}). 
+Thus asymmetry of environmental effects is a mechanism that may be partly 
+responsible for non-normal copula structure and
 asymmetric tail dependence across space in ecological variables.
-<!--***DAN: move the following to the end of this part-->
 This result is represented in Fig. \ref{MasterFigure} as the solid box around 
@@
 are solid, instead of dashed, although the results we 
-present in this section are theoretical.
-<!--***DAN: move the above to the end of this part-->
-For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
-(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}A,B), 
-for both $g_1$ and $g_2$,
-asymmetry of tail dependence was strong; for 
-$b=\pm 0.5$, asymmetry was weaker but still apparent
-(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}C-F).
-<!--***DAN: Shyamolina, can you please combine Figs S24-S26 into one single figure
-with 6 panels? Please also search for citations in the main text and sup mat that 
-used to refer to the three figures and replace them with a citation to the
-new, combined figure.
-***Shya: Done.-->
-Lower-tail (respectively, upper-tail) spatial dependence in 
-effects of noise ($g_1$, respectively, $g_2$) 
-produced lower-tail (respectively, upper-tail) spatial 
-dependence in the ecological variable.
-Model selection results (Table \ref{SM-tab_summary_stat_asym_sens}) confirmed that
-ecological outputs had non-normal copula structure, but did not always reflect 
-asymmetry of tail dependence, perhaps because those methods measured
-dependence in the extreme tails. Results using our nonparametric statistics, however,
-reflected the visually apparent asymmetry (Tables \ref{SM-tab_stat_npa_asym}). 
-<!--***DAN: Simplify as follows, maybe:
-1) get rid of table S10 (you will have to change the methods so it no longer says we
-do our model selection analysis here)
-2) simplify table S11 so it only shows the cor results (also means a change to methods)
-and then maybe bring the resulting table to main text?
--->
-
-<!--***DAN: Shyamolina, can you please simplify Table S11 so it only shows the cor
-results? That also means the third column will be unnecessary. Please do this as follows:
-1) keep but comment out the code that produces the old table; 2) write new code (presumably
-starting from a copy of the old code) that makes the new, simpler version of the table. 
-That will make it very easy to produce the old version again if we ever want to.
-***Shya: Done.-->
+present so far in this section are theoretical.
 
@@
 
+```{r tab_stat_npa_asym_simple,echo=F,results="markup"}
+library(kableExtra)
+
+tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
+tab_stat_npa_asym<-c()
+for(i in c(1:8)){
+  temp<-make_table_stat_fn(data=tab_asym[[i]])
+  tab_stat_npa_asym<-rbind(tab_stat_npa_asym,temp)
+}
+
+ind_corstat<-which(tab_stat_npa_asym$Statistic %in% c("cor$_{FS}$-cor$_{LS}$"))
+tab_corstat_asym<-tab_stat_npa_asym[ind_corstat,]
+tab_corstat_asym<-tab_corstat_asym[,-1]
+
+#gtext<-c("b=0.1, g$_{1}$","b=0.1, g$_{2}$","b=-0.1, g$_{1}$","b=-0.1, g$_{2}$","b=0.5, g$_{1}$","b=0.5, g$_{2}$","b=-0.5, #g$_{1}$","b=-0.5, g$_{2}$")
+#tab_stat_npa_asym<-cbind(gtext,tab_corstat_asym)
+
+tab_corstat_asym<-cbind(b=c(0.1,0.1,-0.1,-0.1,0.5,0.5,-0.5,-0.5),
+  g=c("g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$","g$_{1}$","g$_{2}$"),tab_corstat_asym)
+
+rownames(tab_corstat_asym)<-NULL
+
+knitr::kable(tab_corstat_asym, booktabs = T, 
+             format = "latex",linesep = "",escape=F, 
+             caption.short = "Asymmetric sensitivity model, nonparametric statistics results",
+             caption ="Asymmetric sensitivity model, nonparametric statistics results. 
+Outputs of simulations from the first model detailed in 
+the methods section for Q2 (i.e., Figs \\ref{fig_ms_asym_b_0.1} and 
+\\ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to nonparamtric statistical 
+analyses decribed in the methods section for Q1. The statistic cor$_{FS}$ - cor$_{LS}$ was computed for the
+output of the model with the indicated b and g, and the value was compared to 1000 values
+of the same statistic computed on surrogate time series randomized to have normal copula structure
+(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
+given statistic on the data was less (respectively, more) than its value on $X$ of the 
+surrogates, so entries of the form $<X$ (respectively, $>X$) for $X$ equal to $975$ or above 
+indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
+lower-tail dependence compared to a normal-copula null hypothesis. `FS' stands for `first slice'
+and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
+0.9 to 1.\\label{tab_stat_npa_asym}")
+```
+
+<!--***DAN: got to here so far in editing the result for Q2-->
+
+<!--Second mechanism, the Moran one-->
 The hypothesis that characteristics of 

### 2019-10-03 SupportingInformation.Rmd
edits to Q2 sections, through first model methods and results

@@
 \includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_-0.5_r_0.8_asym_sens.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$ (A-B), $b=0.5$ (C-D) and $b=-0.5$ (E-F).\label{fig_ms_asym_b_-0.1_0.5_-0.5}}
+\caption[Analogous to Fig.
+\ref{MT-fig_ms_asym_b_0.1}, but using different model parameters]{Analogous to Fig.
+\ref{MT-fig_ms_asym_b_0.1}, but using $b=-0.1$ (A-B), $b=0.5$ (C-D) and $b=-0.5$ (E-F).\label{fig_ms_asym_b_-0.1_0.5_-0.5}}
 \end{center}
@@
 <!--Displaying summary table with model selection results from asymmetric dependence analysis-->
+<!--Decided to remove this from the AER version
 ```{r tab_summary_stat_asym_sens,echo=F,results="markup"}
@@
 ```
-
-<!--*** Keeping old complex table for future use:
+-->
+
+<!--*** Keeping old complex table for future use, simpler version now displayed in main text:
 Displaying summary table with non-parametric approach results of asymmetric dependence analysis
@@
 <!---Simpler version: Displaying summary table with non-parametric approach results of asymmetric dependence analysis--->
+<!--
 ```{r tab_stat_npa_asym_simple,echo=F,results="markup"}
@@
              caption ="Asymmetric sensitivity model, nonparametric statistics results. 
-The output of simulations from the first model detailed in 
+Outputs of simulations from the first model detailed in 
 the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
 \\ref{fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to nonparamtric statistical 
-analyses decribed in the methods section for Q1. The statistic ($cor_{FS} - cor_{LS}$) was computed for the
+analyses decribed in the methods section for Q1. The statistic cor$_{FS}$ - cor$_{LS}$ was computed for the
 output of the model with the indicated b and g, and the value was compared to 1000 values

### 2019-10-03 Paper.Rmd
working on Q2 stuff

@@
 very simple AR(1) formulation, influenced by a "regional" environmental factor which
-is correlated across locatons ($\epsilon$) and by a local factor ($\delta$). 
+is correlated across locatons ($\epsilon$) and by an entirely local factor ($\delta$). 
 The function $g$ describes how the environmental factor $\epsilon_i(t)$ influences
@@
 and our Spearman- or Kendall-preserving normal-copula 
-surrogates (see section on methods for Q1) to these 
+surrogate comparison methods (see section on methods for Q1) to these 
 outputs to discover if the model could
@@
 
-<!--***DAN: stopped here in Q2 methods in order to go on to the correspondng
-section of Q2 results-->
-
 Our second mechanism is an extension of the well known Moran
 effect, and was summarized conceptually in the Introduction. 
-Let $E_i(t)$ again be an ecological variable, $i=1,2$.
-We initially assume the AR(1) dynamics 
+We consider a linear model, as well as 
+two parameterizations of a nonlinear population model which includes density dependence
+of population growth. For the linear model, let $E_i(t)$ 
+again be an ecological variable, $i=1,2$.
+We use AR(1) dynamics, 
 $E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
@@
 flux of a biogeochemical variable such as methane, or other quantities.
-
-To explore the importance of nonlinear dynamics for our proposed mechanism, 
-we also considered a
-nonlinear population model $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
-$\sigma=1$. When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
-$K$, and the nonlinearities of the model have limited influence on dynamics. When
-$\sigma=1$, model dynamics are strongly nonlinear. We refer to these as
+The nonlinear population model was $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$ for $i=1,2$, using $r=0.5$, $K=100$, $\epsilon_i(t)$
+as above, and $\sigma=0.1$ or 
+$\sigma=1$ in different simulations. 
+When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
+$K$, where the model equation can be well approximated by a linear equation,
+and the nonlinearities of the model therefore have limited influence on dynamics. When
+$\sigma=1$, model dynamics are strongly nonlinear because the stochastic 
+component of the model causes populations to stray far from the carrying capacity, $K$. 
+We refer to these as
 the weak-noise and strong-noise cases, though the importance of the noise here
 is that, when it is strong, it brings the nonlinearities of the model into play.
-<!--***DAN: Didn't we use another value of r as well, and present results
-on it? If so, probably should mention it here, in part so the exploration
-seems a bit broader? Well, we mentioned that we did those and we verbally summarize
-results but we did not show any. So maybe leave things as is in this respect.-->
-
-For each of the model setups above, for each of several copula families, 
-<!--***DAN: I currently use C, SC, G, SG, J, SJ. Do I really need all those? 
-Surely it would be enough to just do C and SC and then state that the others 
-give similar results? And if I move one of the result figure from this section
-into the main text, that will mean only one such fig in the sup mat, at least
-for AR(1)? I mean, right now we have 18 figure in sup mat from this analysis,
-and only part of one of them in main text! See if the results are actually similar across
-S/SC, G/SG, J/SJ, and if so consider keeping only the S/SC results in either the
-main text or the sup mat, and just stating the others were similar. Then you would presumably
-move the full C fig into the main text, replacing the current partial version.
-So instead of having 18 figs in sup mat and half a fig in main text, you would
-have 5 figs in sup mat and 1 in main text, and then you would just say "we did the
-same thing for G, SG, J, SJ and got similar results."
-So let's check, are results, broadly speaking, independent of which of the 6 copulas
-I used?:
-A) for AR(1) models, yes
-B) for Ricker, weak noise, yes
-C) for Ricker, strong noise, its more complicated. It's generally true that you get
-more approximate agreement between noise and pop stats, with some rather glaring disagreement.
-But which stats disagree the most depends on which cop we used!
-So I guess we can still just display the C and SC results, as planned above, and explain
-in text.
--->
+
+For each of the three model setups above (the linear model and the nonlinear model
+with weak and strong stochasticity), for the Clayton and survival Clayton copula families, 
 for each $\tau=0.1,0.2,\ldots,0.9$, and for each of
@@
 $(\epsilon_1(t),\epsilon_2(t))$ from the
-bivariate random variable with the given copula family and the given Kendall
+bivariate random variable with standard-normal marginals and with 
+the given copula family and the given Kendall
 correlation $\tau$.
@@
 Values were plotted against $\tau$ for noises and populations. If the 
-hypothesis from the Introduction was correct that characteristics of 
+hypothesis from the Introduction was reasonable that characteristics of 
 the copula structure of spatial dependence in an
-ecological variable can be inherited from the copula structure of spatial 
-dependence in an environmental variables, then 
+ecological variable may be inherited from characteristics of spatial 
+dependence in an environmental variable through a Moran-like effect, then 
 plots should be similar for populations and noises.
+
+<!--***DAN: stopped here for methods of Q2-->
 
@@
 Results using our statistic $\cor_{0,0.1}-\cor_{0.9,1}$ 
-strongly reflected the visually apparent asymmetry (Tables \ref{tab_stat_npa_asym}). 
+strongly reflected the visually apparent asymmetry (Table \ref{tab_stat_npa_asym}). 
 Thus asymmetry of environmental effects is a mechanism that may be partly 
 responsible for non-normal copula structure and
-asymmetric tail dependence across space in ecological variables.
+asymmetric tail associations across space in ecological variables.
 This result is represented in Fig. \ref{MasterFigure} as the solid box around 
 "Nonlinear environmental effects, Liebig's law" and the arrows labelled "A".
-It is explained in the Discussion why the box and some of those arrows 
-are solid, instead of dashed, although the results we 
-present so far in this section are theoretical.
+It is explained in the Discussion why the box and some of the arrows 
+are solid, instead of dashed, although the results of this section are theoretical.
 
@@
 tail dependence across space in the ecological variables. Shown are the last
-$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text, $b=0.1$.\label{fig_ms_asym_b_0.1}}
+$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text, $b=0.1$. 
+Asymmetric tail associations are visually apparent, but were also elaborated statistically in 
+Table \ref{tab_stat_npa_asym}. \label{fig_ms_asym_b_0.1}}
 \end{center}
@@
 
-<!--***DAN: got to here so far in editing the result for Q2-->
-
 <!--Second mechanism, the Moran one-->
 The hypothesis that characteristics of 
-spatial copula structure and tail dependence
+the copula structure of spatial dependence 
 in an ecological variable can be inherited from an environmental 
-variable was largely confirmed; 
-<!--***DAN: You are using language about confirming hypotheses, here and elsewhere,
-but this may rub some people the wrong way because we have not confirmed anything
-empirically, we are just doing models, and only a cursory exploration of each
-model, at that! So change the language.-->
+variable was found to be reasonable, because it held for the models we examined; 
 however, similarities between environmental- and 
@@
 dynamics were strongly nonlinear.
-Thus this is a mechanism that can produce non-normal copula structure and
-asymmetric tail dependence 
-<!--****DAN: Don't firget to replace @tail dependence@ by @tail asociation@ as appropriate-->
-across space in ecological variables.
-<!--***DAN: move the below to end of this part-->
-This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
-"Moran effects" and the arrows labelled "B".
-<!--***DAN: move the above to end of this part-->
-For simulations using AR(1) models, copula 
-statistics were always very similar for noise and populations 
-(Fig \ref{fig_Cause4copula_scatter_C_kend_ar1}, 
-\ref{SM-fig_Cause4copula_C_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1}),
-though there were significant but minor differences for some statistics
-and simulations.
-For weak noise and our nonlinear model, copula statistics
-were again very similar for noise and model outputs 
-(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}).
-Though there were again significant but small differences for some statistics
-and simulations, the hypothesis was supported that copula 
-structure in the noise translates
-to similar but not identical copula structure in output populations. 
+For simulations using AR(1) models, our correlation and tail asymmetry statistics 
+were always similar for noise and populations 
+(Fig. \ref{fig_Cause4copula_C_kend_ar1} for the Clayon copula, 
+Fig. \ref{SM-fig_Cause4copula_SC_kend_ar1} for the survival Clayton copula).
+<!--***DAN: reorder sup mat figs so they are cited in order-->
+Though there were significant differences for many statistics
+and simulations, these were small compared to the overall tendency for
+larger (respectively, smaller) values of our statistics, as applied to
+noise, to be paired with larger (respectively, smaller) values of the same
+statistics as applied to model outputs. 
+For our nonlinear model with weak noise, values of the statistics
+were again quite similar for noise and model outputs 
+(Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1} for Clayton, 
+Fig. \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_0.1} for survival Clayton).
+Though there were again significant differences for many statistics
+and simulations, these were again small relative to overall variation of 
+values of the statistics.
 Since many ecological models are nonlinear, this result provides the 
-expectation that there will be a Moran-effect-like correspondence 
-between noise and model-output copula structure across space 
+reasonable expectation that there may be a Moran-effect-like correspondence 
+between noise and model-output dependence structure across space 
 for typical ecological dynamics,
-as long as noise is small enough that dynamics stay relatively close
+as long as environmental noise is small enough that dynamics stay relatively close
 to the model equilibrium. Results that hold for "weak noise" in this sense
 are common in ecology.
-We repeated the analysis with $r=1.3$, with similar results. The deterministic 
+For strong noise and using our nonlinear model, 
+our correlation and tail-asymmetry statistics, generally 
+speaking, were approximately similar between noise and model
+outputs; however, similarity was reduced relative to previous results, 
+and for a few simulations, asymmetry statistics
+had opposite signs for noise and model outputs 
+(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1} and \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_1}). For instance, 
+using a Clayton  copula with a large Kendall $\tau$,
+$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
+population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}).
+
+We repeated our analyses using the nonlinear model with $r=1.3$. The deterministic 
 one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
@@
 but exhibits an oscillatory approach to $K$ when $r>1$ (overcompensating
-dynamics, e.g., $r=1.3$).
-
+dynamics, e.g., $r=1.3$). For weak noise, similarities were again dominant between
+values of our statistics on noise and population time series. 
+For strong noise, however, discrepancies were often glaring. Apparently 
+noise of standard deviation 1 interacted especially strongly with 
+model nonlinearities when the model was in its overcompensatory 
+regime. We repeated all analyses using the Gumbel, survival Gumbel, Joe, and 
+survival Joe copulas, with substantially similar main conclusions (not shown).
+
+Thus it is reasonable to hypothesize that a Moran-effect-like 
+mechanism may produce non-normal copula structure and
+asymmetric tail associations across space in ecological variables.
+This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
+"Moran effects" and the arrows labelled "B".
+
+<!--***DAN: Decided to remove this simplified version of the figure for AER
+and replace by the bigger one which was in the supp mat, now below.
 \begin{figure}[!h]
@@
 \end{figure}
-
-For strong noise and using our nonlinear model, 
-copula statistics, generally speaking, were approximately similar 
-between noise and model
-outputs; however, similarity was reduced relative to previous results, 
-and for a few simulations, tail dependence asymmetry statistics
-had opposite signs for noise and model outputs 
-(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). For instance, using a Clayton  copula with a large Kendall $\tau$,
-$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
-population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}); 
-and likewise for survival Gumbel and survival Joe copulas (Figs \ref{SM-fig_Cause4copula_SG_kend_sRicker_nsd_1}, 
-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). Overall,
-similarities between noise and model-output copula statistics
-were more noticable than discrepancies, in spite of nonlinearities.
-We repeated the analysis with $r=1.3$. Discrepancies between noise and population
-copula structure in this case were often glaring. Apparently 
-noise of standard deviation 1 interacted especially strongly with 
-model nonlinearities when the model was in its overcompensatory 
-regime.
+-->
+
+<!--Plots for AR(1) with Clayton-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf} \\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption{Comparison of correlation and tail association statistics between 
+environmental-noise inputs and ecological-variable 
+outputs, AR(1) model, Clayton copula. Red and blue points give 
+means over 50 replicate simulations of the listed statistics, 
+red points are for noise inputs and blue points are for
+model outputs. Error bars are standard errors. 
+$P$-values (purple points, right axis) are for a 
+paired $t$-test of the null hypothesis that the distributions have the same mean. 
+For bottom panels, headers give Pearson correlations for
+the points. The regression line through the points (black) was similar
+to the 1-1 line (green). 
+See methods section for Q2 
+for details.\label{fig_Cause4copula_C_kend_ar1}}
+\end{center}
+\end{figure}
+<!--***DAN: Shyamolina, the panels in the bottom row do not line up with the others. 
+I remember you fixed this before for the sup mat figs, can you plase fix it again here?
+Also, can you make the panels smaller or more tightly packed or soemthing
+so the whole figure plus caption fits on a page?-->
+
+<!--***DAN: Got to here in results for Q2-->
 

### 2019-10-03 SupportingInformation.Rmd
working on Q2 stuff

@@
 
-<!--Plots for AR(1) with Clayton-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/common_legend_cause4copula_stat.pdf}\\
-\vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf}
-\caption[Comparison of copula structure of noise and model output, AR(1) model,
-Clayton copula.]{Comparison of copula structure of noise inputs and model 
-outputs, AR(1) model, Clayton copula. Red and blue points give 
-means over 50 replicate simulations of the statistics listed on the left axes, 
-red points are values for noise inputs and blue points are for
-model outputs. Error bars are standard errors. 
-$P$-values (purple points, right axis) are for a 
-paired $t$-test of the null hypothesis that the distributions of 50 noise and 
-output values have the same mean; $p$-values less than $0.05$ indicate that 
-copula structure as measured by the listed statistic differs significant
-between noise inputs and model outputs. See methods section for Q2 
-for additional details.\label{fig_Cause4copula_C_kend_ar1}}
-\end{center}
-\end{figure}
-
 <!--Following two chunks examine Moran-like mechanism with survival clayton copula noises in AR(1) model 
@@
 survival Clayton copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
-model, survival Clayton copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
-\ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
+model, survival Clayton copula. Same format as Fig. \ref{MT-fig_Cause4copula_C_kend_ar1}, 
+see caption of that figure for 
 details.\label{fig_Cause4copula_SC_kend_ar1}}
@@
 <!--Plots for AR(1) with Gumbel-->
+<!--Removed from AER version
 \begin{figure}[!h]
@@
 Gumbel copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
-model, Gumbelcopula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
+model, Gumbel copula. Same format as Figs \ref{fig_Cause4copula_C_kend_ar1} and 
 \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, see captions of those figures for 
@@
 <!--Plots for AR(1) with Survival Gumbel-->
+<!--Removed from AER version
 \begin{figure}[!h]
@@
 <!--Plots for AR(1) with Joe-->
+<!--Removed from AER version
 \begin{figure}[!h]
@@
 <!--Plots for AR(1) with Survival Joe-->
+<!--Removed from AER version
 \begin{figure}[!h]
@@
 model, weak noise, Clayton copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, weak noise, Clayton copula. Same format as Figs 
-\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
-see captions of those figures for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_0.1}}
+outputs, stochastic Ricker model, weak noise, Clayton copula. Same format as Fig. 
+\ref{MT-fig_Cause4copula_C_kend_ar1}, 
+see caption of that figure for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_0.1}}
+\end{center}
+\end{figure}
+
+<!--Following two chunks examine Moran-like mechanism with survival clayton copula noises in Ricker model and 
+then generate plots with the simulated results-->
+```{r cause4copula_stat_ricker_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
+source("Cause4copula.R")
+#-----------------
+set.seed(seed)
+#---------------------------------------
+tag_ricker_SC<-"999SC" #this is a caching tag used later to cache output from present chunk
+
+# This chunk stores results for ar1 model : survival clayton
+
+model<-"ricker"
+fcode<-13
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    if (!dir.exists(tempo)){
+      dir.create(tempo)
+    }
+    resloc<-paste(tempo,"/",sep="")
+
+  # for survival clayton with spearman and kendall method
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
+                                           num_keep_last=500,resloc=resloc,
+                                           params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_SC_spear.RDS",sep=""))
+
+    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
+                                          num_keep_last=500,resloc=resloc,
+                                          params=params,p0=p0,model=model)
+
+    saveRDS(myres,paste(resloc,"ricker_SC_kend.RDS",sep=""))
+  }
+ }
+}
+```
+
+```{r plot_cause4copula_stat_ricker_SC, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SC,mtime("Plotter_Cause4copula.R"))}
+
+set.seed(seed)
+source("Plotter_Cause4copula_stat.R")
+#---------------------------------------
+
+model<-"ricker"
+fcode<-13
+resloc0<-"./Results/Cause4copula_results/"
+
+nsdlist<-c(0.1,1)
+Klist<-c(100)
+
+# for varying r (@r=1 transition from under- to over-compensatory dynamics )
+rlist<-seq(from=0.1,to=1.3,by=0.2)
+
+for(insd in nsdlist){
+  for(iK in Klist){
+   for(i in rlist){
+    params<-c(i,iK) # this parameter is a vector with r and K
+    p0<-c(params[2],params[2])
+
+    resloc<-"./Results/Cause4copula_results/"
+
+    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
+
+    resloc<-paste(tempo,"/",sep="")
+
+    readres<-readRDS(paste(resloc,"ricker_SC_spear.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+    readres<-readRDS(paste(resloc,"ricker_SC_kend.RDS",sep=""))
+    Plotter_Cause4copula_stat(res=readres)
+
+  }
+ }
+}
+```
+
+<!--Plots for stochastic ricker with weak noise (Survival Clayton copula) : nsd=0.1-->
+\begin{figure}[!h]
+\begin{center}
+\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
+\vspace{-0.4 cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
+\hspace{-0.8cm}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
+model, weak noise, survival Clayton copula.]{Comparison of copula structure of noise inputs and model 
+outputs, stochastic Ricker model, weak noise, survival Clayton copula. Same format as Fig. 
+\ref{MT-fig_Cause4copula_C_kend_ar1}, 
+see caption of that figure for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_0.1}}
 \end{center}
@@
 model, strong noise, Clayton copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, Clayton copula. Same format as Figs 
-\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
-see captions of those figures for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_1}}
-\end{center}
-\end{figure}
-
-<!--Following two chunks examine Moran-like mechanism with survival clayton copula noises in Ricker model and 
-then generate plots with the simulated results-->
-```{r cause4copula_stat_ricker_survivalclayton, echo=F, results="hide", cache=T, cache.extra=list(seed,mtime("Cause4copula.R"),mtime("CopulaFunctions_flexible.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
-source("Cause4copula.R")
-#-----------------
-set.seed(seed)
-#---------------------------------------
-tag_ricker_SC<-"999SC" #this is a caching tag used later to cache output from present chunk
-
-# This chunk stores results for ar1 model : survival clayton
-
-model<-"ricker"
-fcode<-13
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    if (!dir.exists(tempo)){
-      dir.create(tempo)
-    }
-    resloc<-paste(tempo,"/",sep="")
-
-  # for survival clayton with spearman and kendall method
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="spearman",lb=0,ub=0.1,
-                                           num_keep_last=500,resloc=resloc,
-                                           params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_SC_spear.RDS",sep=""))
-
-    myres<-Sim_Cause4copula_stat(N=5000,numsim=50,fcode=fcode,nsd=insd,method="kendall",lb=0,ub=0.1,
-                                          num_keep_last=500,resloc=resloc,
-                                          params=params,p0=p0,model=model)
-
-    saveRDS(myres,paste(resloc,"ricker_SC_kend.RDS",sep=""))
-  }
- }
-}
-```
-
-```{r plot_cause4copula_stat_ricker_SC, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SC,mtime("Plotter_Cause4copula.R"))}
-
-set.seed(seed)
-source("Plotter_Cause4copula_stat.R")
-#---------------------------------------
-
-model<-"ricker"
-fcode<-13
-resloc0<-"./Results/Cause4copula_results/"
-
-nsdlist<-c(0.1,1)
-Klist<-c(100)
-
-# for varying r (@r=1 transition from under- to over-compensatory dynamics )
-rlist<-seq(from=0.1,to=1.3,by=0.2)
-
-for(insd in nsdlist){
-  for(iK in Klist){
-   for(i in rlist){
-    params<-c(i,iK) # this parameter is a vector with r and K
-    p0<-c(params[2],params[2])
-
-    resloc<-"./Results/Cause4copula_results/"
-
-    tempo<-paste(resloc,"Cause4copula_stat_results_",model,"_nsd_",insd,"_r_",params[1],"_K_",params[2],"_",BiCopName(fcode,short=F),sep="")
-
-    resloc<-paste(tempo,"/",sep="")
-
-    readres<-readRDS(paste(resloc,"ricker_SC_spear.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-    readres<-readRDS(paste(resloc,"ricker_SC_kend.RDS",sep=""))
-    Plotter_Cause4copula_stat(res=readres)
-
-  }
- }
-}
-```
-
-<!--Plots for stochastic ricker with weak noise (Survival Clayton copula) : nsd=0.1-->
-\begin{figure}[!h]
-\begin{center}
-\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/common_legend_cause4copula_stat.pdf}\\
-\vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Spearman_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Kendall_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pearson_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Coru_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pu_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2l_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_D2u-D2l_vs_Kendall's Tau.pdf}\\
-\hspace{-0.8cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ricker_nsd_0.1_r_0.5_K_100_Survival Clayton/Survival Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
-\caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, weak noise, survival Clayton copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, weak noise, survival Clayton copula. Same format as Figs 
-\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
-see captions of those figures for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_0.1}}
+outputs, stochastic Ricker model, strong noise, Clayton copula. Same format as Fig. 
+\ref{MT-fig_Cause4copula_C_kend_ar1}, 
+see caption of that figure for details.\label{fig_Cause4copula_C_kend_sRicker_nsd_1}}
 \end{center}
@@
 model, strong noise, survival Clayton copula.]{Comparison of copula structure of noise inputs and model 
-outputs, stochastic Ricker model, strong noise, survival Clayton copula. Same format as Figs 
-\ref{fig_Cause4copula_C_kend_ar1} and \ref{MT-fig_Cause4copula_scatter_C_kend_ar1}, 
-see captions of those figures for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_1}}
+outputs, stochastic Ricker model, strong noise, survival Clayton copula. Same format as Fig. 
+\ref{MT-fig_Cause4copula_C_kend_ar1}, 
+see caption of that figure for details. \label{fig_Cause4copula_SC_kend_sRicker_nsd_1}}
 \end{center}
@@
 <!--Plots for stochastic ricker with weak noise (Gumbel copula) : nsd=0.1-->
+<!--Removed from AER version
 \begin{figure}[!h]
@@
 \end{figure}
+-->
 
 <!--Plots for stochastic ricker with strong noise (Gumbel copula): nsd=1-->
+<!--Removed from AER version
 \begin{figure}[!h]
@@
 <!--Plots for stochastic ricker with weak noise (Survival Gumbel copula)  : nsd=0.1-->
+<!--Removed from AER version
 \begin{figure}[!h]
@@
 \end{figure}
+-->
 
 <!--Plots for stochastic ricker with strong noise (Survival Gumbel copula): nsd=1-->
+<!--Removed from AER version
 \begin{figure}[!h]
@@
 <!--Plots for stochastic ricker with weak noise (Joe copula) : nsd=0.1-->
+<!--Removed from AER version
 \begin{figure}[!h]
@@
 \end{figure}
+-->
 
 <!--Plots for stochastic ricker with strong noise (Joe copula) : nsd=1-->
+<!--Removed from AER version
 \begin{figure}[!h]
@@
 <!--Plots for stochastic ricker with weak noise (Survival Joe): nsd=0.1-->
+<!--Removed from AER version
 \begin{figure}[!h]
@@
 \end{figure}
+-->
 
 <!--Plots for stochastic ricker with strong noise (Survival Joe): nsd=1-->
+<!--Removed from AER version
 \begin{figure}[!h]

### 2019-10-03 Paper.Rmd
figure fixed within a page: downsized

@@
 library(kableExtra)
-
+source("./make_table_stat_fn.R")
 tab_asym<-readRDS("./Results/asym_sens_results/stat_npa_asym_b_0.1_b_0.5.RDS")
@@
 \vspace{-0.4 cm}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf} \\
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
-\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
+\vspace{-0.2 cm}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
+\vspace{-0.2 cm}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
+\vspace{-0.2 cm}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf} \\
+\hspace{-0.6 cm}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
+\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
 \caption{Comparison of correlation and tail association statistics between 
@@
 Also, can you make the panels smaller or more tightly packed or soemthing
-so the whole figure plus caption fits on a page?-->
+so the whole figure plus caption fits on a page?
+***Shya: Ok, now I fixed it. See the margin specifications are different for Paper and Suuppmat, 
+that's why I have to downsize the figure here in the maintext.-->
 

### 2019-10-03 SupportingInformation.Rmd
figure fixed within a page: downsized

@@
 
-<!--Function to generate a table for npa results-summary with bivariate data, this one about asymmetry of tail dependence-->
-```{r make_table_stat_fn, echo=F}
-fractxt<-function(fracval,numsurrog){
-  ptxt<-''
-  if (fracval>.5*numsurrog){                        
-    ptxt<-paste(">",fracval,sep='')
-  } 
-  if (fracval<=.5*numsurrog){
-    ptxt<-paste("<",numsurrog-fracval,sep='')
-  }
-  return(ptxt)
-}
-
-make_table_stat_fn<-function(data){
-
-  #atdres<-data.frame(Statistic=c("$\\cor_{0,0.1}-\\cor_{0.9,1}$",
-  #                               "$\\Ps_{0,0.1}-\\Ps_{0.9,1}$",
-  #                               "$\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$"),
-  atdres<-data.frame(Statistic=c("cor$_{FS}$-cor$_{LS}$",
-                                 "P$_{FS}$-P$_{LS}$",
-                                 "D$_{LS}^2$-D$_{FS}^2$"),
-                     Kendall=rep("",3),
-                     Spearman=rep("",3),stringsAsFactors = F)
-
-  atdres[1,2]<-fractxt(data$corlmcoru_frac_K,numsurrog = data$numsurrog_success_K)
-  atdres[1,3]<-fractxt(data$corlmcoru_frac_S,numsurrog = data$numsurrog_success_S)
-  atdres[2,2]<-fractxt(data$PlmPu_frac_K,numsurrog = data$numsurrog_success_K)
-  atdres[2,3]<-fractxt(data$PlmPu_frac_S,numsurrog = data$numsurrog_success_S)
-  atdres[3,2]<-fractxt(data$D2umD2l_frac_K,numsurrog = data$numsurrog_success_K)
-  atdres[3,3]<-fractxt(data$D2umD2l_frac_S,numsurrog = data$numsurrog_success_S)
-
-  return(atdres)
-
-}
+<!--Sourcing Function to generate a table for npa results-summary with bivariate data, this one about asymmetry of tail dependence-->
+```{r source_make_table_stat_fn, echo=F}
+source("./make_table_stat_fn.R")
 ```

### 2019-10-04 Paper.Rmd
done editing Q2 sections, just one thing for Shyamolina to do

@@
 
-<!--***DAN: stopped here for methods of Q2-->
-
-<!--***DAN: How much of the material from Appendix S8 should be moved into
-the main text? I am thinking a fair amount of it. Currrently there
-is really no results on that topic at all in the main text, and that seems
-a bit ridiculous and probably is why the referee said we treated this 
-material in a "breezy" way.-->
 The next mechanism we investigated is evolutionary, and pertains to bivariate 
 trait data across species, e.g., our bird and mammal 
-data. This mechanism is a hypothetical explanation for the right tail dependence
+data. This mechanism is a hypothetical explanation for the bias toward right tail association
 observed in those data (Fig. \ref{fig_biv_multi_raw_cop}F,G). 
-The hypothesis is that tail dependence occurs in evolutionary changes
-in bivariate characters, and gives rise to tail dependence between the 
+The hypothesis is that asymmetric tail association occurs in evolutionary changes
+in bivariate characters, and gives rise to asymmtric tail association between the 
 two character values across extant species.
@@
 evolution was simulated 100 times. For each simulation, symmetry of tail dependence 
-of the two characters across phylogeny tips was assessed.
+of the two characters across phylogeny tips was assessed using our nonparametric
+statistics.
 We hypothesized that cases 1 and 2 above would yield stronger left- than right-tail 
-dependence in tip characters, and cases 4 and 5 would yield the reverse. 
+associations in tip characters, and cases 4 and 5 would yield the reverse. 
 The simulator was written in Python and used version 4.4 of the DendroPy 
@@
 <!--***DAN: Decided to remove this simplified version of the figure for AER
-and replace by the bigger one which was in the supp mat, now below.
+and replace by the bigger one which was in the supp mat, now below. Keeping the
+code the simpler one here in case we ever need it.
 \begin{figure}[!h]
@@
 \end{figure}
-<!--***DAN: Shyamolina, the panels in the bottom row do not line up with the others. 
-I remember you fixed this before for the sup mat figs, can you plase fix it again here?
-Also, can you make the panels smaller or more tightly packed or soemthing
-so the whole figure plus caption fits on a page?
-***Shya: Ok, now I fixed it. See the margin specifications are different for Paper and Suuppmat, 
-that's why I have to downsize the figure here in the maintext.-->
-
-<!--***DAN: Got to here in results for Q2-->
 
 For our evolutionary model, the hypothesis was 
-correct that asymmetric tail dependence in evolutionary changes 
-can produce similarly asymmetric tail dependence between characters across
-phylogeny tips (Appendix \ref{SM-evol_mech}, Fig. \ref{SM-fig_cop_evo} 
-for details). 
-Thus this is a mechanism that may produce non-normal copula structure and
-asymmetric tail dependence between characters.
+correct that asymmetric tail association in evolutionary changes 
+can produce similarly asymmetric tail association between characters across
+phylogeny tips.
+Once character evolution was simulated 100 times for each of the
+dependence structure we used on evolutionary changes, 
+we had 817 bivariate characters for each of 500 simulations (see methods 
+section for Q2). We 
+computed our nonparametric asymmetry statistics 
+$\cor_{0,0.2}-\cor_{0.8,1}$, $\Ps_{0,0.2}-\Ps_{0.8,1}$,
+and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ for each simulation output and 
+produced a histogram for each statistic and for each
+dependence structure (Fig. \ref{fig_cop_evo}). Results showed that 
+asymmetries of tail associations in evolutionary changes
+were associated with similar asymmetries of tail associations in extant characters.
+Thus, based on our simulations, we cannot reject the possibility that
+this is a mechanism behind observed asymmetric tail associations
+in our bird and mammal data.
 This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
@@
 interactions" and the arrow labelled "C" in Fig. \ref{MasterFigure}.
+
+<!--results for third mechanism-->
+\begin{figure}[!h]
+\textbf{ \hspace{3.5 cm} (A) \hspace{2 cm} (F) \hspace{2 cm} (K)} \\
+\vspace{-1 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeL_hist_taildep.pdf}\\
+\end{center}
+\vspace{-.33 cm}
+\textbf{ \hspace{3.5 cm} (B) \hspace{2 cm} (G) \hspace{2 cm} (L)} \\
+\vspace{-1 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatL_hist_taildep.pdf}\\
+\end{center}
+\vspace{-.33 cm}
+\textbf{ \hspace{3.5 cm} (C) \hspace{2 cm} (H) \hspace{2 cm} (M)} \\
+\vspace{-1 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/BivNormal_hist_taildep.pdf}\\
+\end{center}
+\vspace{-.33 cm}
+\textbf{ \hspace{3.5 cm} (D) \hspace{2 cm} (I) \hspace{2 cm} (N)} \\
+\vspace{-1 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatR_hist_taildep.pdf}\\
+\end{center}
+\vspace{-.33 cm}
+\textbf{ \hspace{3.5 cm} (E) \hspace{2 cm} (J) \hspace{2 cm} (O)} \\
+\vspace{-1 cm}
+\begin{center}
+\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeR_hist_taildep.pdf}\\
+\caption[Non-parametric statistics on a mechanism of tail dependence in character distributions]{Three 
+measures of asymmetry of tail associations ($\cor_{0,0.2}-\cor_{0.8,1}$, A-E; 
+$\Ps_{0,0.2}-\Ps_{0.8,1}$, F-J; $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ K-O) between two characters, across extant species,  were 
+computed for each of 100 simulations of mammalian character evolution for each of five types of tail dependence between 
+evolutionary changes in the characters (extreme left-tail dependence (A, F, K), moderate left-tail dependence (B, G, L), 
+symmetry of tail dependence (C, H, M), moderate right-tail dependence (D, I, N), and extreme right-tail dependence 
+(E, J, O)).
+See the methods section for Q2 for details.
+The number above each panel indicates how many statistics out of $100$ were less than or equal to $0$.
+Values substantially lower (respectively, higher) than $50$ indicate greater left (respectively, right) tail-dependence between characters.\label{fig_cop_evo}}
+\end{center}
+\end{figure}
+<!--***DAN: Shyamolina, can you please make this figure and its caption fit on one page,
+somehow?-->
 

### 2019-10-04 SupportingInformation.Rmd
done editing Q2 sections, just one thing for Shyamolina to do

@@
 
+<!--***Methods for generating evolutionary changes, keep this here-->
 \noindent The matrices from which root character states and changes 
@@
 
-Once character evolution was simulated 100 times for each dependence structure, 
-we had 817 bivariate characters for each of 500 simulations. We thus analyzed 500 corresponding 
-bivariate copulas (one for each simulation), each comprised of 817 points. 
-To assess the symmetry or asymmetry of tail dependence for each copula, we computed 
-the three non-parametric statistics $\cor_{0,0.2}-\cor_{0.8,1}$, $\Ps_{0,0.2}-\Ps_{0.8,1}$,
-and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ for each copula. For each statistic and for each
-dependence structure, we made a histogram of values 
-(Fig. \ref{fig_cop_evo}). Asymmetry of tail dependence in evolutionary changes to characters
-was associated with similar asymmetry of tail dependence in extant characters.
-
+<!--Move all the rest of it into the Discussion, I think, but can decide when I get to the Discussion. If it is mved, then you may want to retitle this appendix to reflect the 
+fact it is onl methods-->
 Performing statistical tests for associations between continuous traits in different species
@@
 \caption[Comparison of copula structure of noise and model output, AR(1) model, 
-survival Clayton copula.]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
+survival Clayton copula]{Comparison of copula structure of noise inputs and model outputs, AR(1) 
 model, survival Clayton copula. Same format as Fig. \ref{MT-fig_Cause4copula_C_kend_ar1}, 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, weak noise, Clayton copula.]{Comparison of copula structure of noise inputs and model 
+model, weak noise, Clayton copula]{Comparison of copula structure of noise inputs and model 
 outputs, stochastic Ricker model, weak noise, Clayton copula. Same format as Fig. 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, weak noise, survival Clayton copula.]{Comparison of copula structure of noise inputs and model 
+model, weak noise, survival Clayton copula]{Comparison of copula structure of noise inputs and model 
 outputs, stochastic Ricker model, weak noise, survival Clayton copula. Same format as Fig. 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, Clayton copula.]{Comparison of copula structure of noise inputs and model 
+model, strong noise, Clayton copula]{Comparison of copula structure of noise inputs and model 
 outputs, stochastic Ricker model, strong noise, Clayton copula. Same format as Fig. 
@@
 \caption[Comparison of copula structure of noise and model output, stochastic Ricker 
-model, strong noise, survival Clayton copula.]{Comparison of copula structure of noise inputs and model 
+model, strong noise, survival Clayton copula]{Comparison of copula structure of noise inputs and model 
 outputs, stochastic Ricker model, strong noise, survival Clayton copula. Same format as Fig. 
@@
 
-<!--adding plots showing results to confirm hypothesis behind third mechanism-->
-\begin{figure}[!h]
-\textbf{ \hspace{6.5 cm} (A) \hspace{2 cm} (B) \hspace{2 cm} (C)} \\
-\vspace{-0.5 cm}
-\begin{center}
-\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeL_hist_taildep.pdf}\\
-\end{center}
-\textbf{ \hspace{6.5 cm} (D) \hspace{2 cm} (E) \hspace{2 cm} (F)} \\
-\vspace{-0.5 cm}
-\begin{center}
-\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatL_hist_taildep.pdf}\\
-\end{center}
-\textbf{ \hspace{6.5 cm} (G) \hspace{2 cm} (H) \hspace{2 cm} (I)} \\
-\vspace{-0.7 cm}
-\begin{center}
-\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/BivNormal_hist_taildep.pdf}\\
-\end{center}
-\textbf{ \hspace{6.5 cm} (J) \hspace{2 cm} (K) \hspace{2 cm} (L)} \\
-\vspace{-0.5 cm}
-\begin{center}
-\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatR_hist_taildep.pdf}\\
-\end{center}
-\textbf{ \hspace{6.5 cm} (M) \hspace{2 cm} (N) \hspace{2 cm} (O)} \\
-\vspace{-0.5 cm}
-\begin{center}
-\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeR_hist_taildep.pdf}\\
-\caption[Non-parametric statistics on a mechanism of tail dependence in character distributions]{Three 
-measures of asymmetry of tail dependence ($\cor_{0,0.2}-\cor_{0.8,1}$ (A, D, G, J, M), 
-$\Ps_{0,0.2}-\Ps_{0.8,1}$ (B, E, H, K, N), and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ (C, F, I, L, O)) between two characters, across extant species,  were 
-computed for each of 100 simulations of mammalian character evolution for each of five types of tail dependence between 
-evolutionary changes in the characters (extreme left-tail dependence (A-C), moderate left-tail dependence (D-F), 
-symmetry of tail dependence (G-I), moderate right-tail dependence (J-L), and extreme right-tail dependence 
-(M-O)). Asymmetry of tail dependence in evolutionary changes in characters tended to beget similar
-asymmetry in the characters across extant species. For additional details see 
-Appendix \ref{evol_mech}.
-The number above each panel indicates how many statistics out of $100$ were less than or equal to $0$.
-Values substantially lower (respectively, higher) than $50$ therefore indicate greater left (respectively, right) tail-dependence between characters.\label{fig_cop_evo}}
-\end{center}
-\end{figure}
-
 <!--Following chunk examine the consequence of non-normal copula atructure on skewness using green spruce aphid abundance data-->
@@
 \includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
-\caption{Analogous to Fig. \ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal.\label{fig_extrisk_lcohen_global}}
+\caption[Analogous to Fig.
+\ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal]{Analogous to Fig.
+\ref{MT-fig_extrisk_lcohen_local} in the main text, but with global dispersal.\label{fig_extrisk_lcohen_global}}
 \end{center}

### 2019-10-06 Paper.Rmd
working on Q3 edits

@@
 Fig. \ref{SM-fig_Cause4copula_SC_kend_ar1} for the survival Clayton copula).
-<!--***DAN: reorder sup mat figs so they are cited in order-->
 Though there were significant differences for many statistics
@@
 dependencies between the local quantities being averaged, including their
-copula structure and tail dependence. 
+copula structure and tail associations. 
 To refine this hypothesis, suppose an ecological variable 
@@
 exploited species, or local fluxes of a greenhouse gas. If  $E_i$ and $E_j$ are 
-right-tail dependent for most location pairs $i$ and $j$, then
+associated primarily in their right tails for most location pairs $i$ and $j$, then
 exceptionally large values tend to occur at the same time in most locations.
@@
 will tend to increase the skewness of the distribution 
-of the spatial mean. Similarly, left-tail dependence between local variables
+of the spatial mean. Similarly, left-tail associations between local variables
 should tend to decrease skewness. Strong positive skewness of the 
@@
 pest or resource abundance, for which extreme values (spikes) 
-in the spatial mean have large effects.
+in the spatial mean may have large effects.
 
@@
 from what one would expect if the copula
-structure of the data were multivariate N, but the data were otherwise statistically
+structure of the data were multivariate normal (i.e., the same copula as a 
+multivariate normal distribtion), but the data were otherwise statistically
 unchanged. Significant differences indicate that non-normal copula structure in the data 
@@
 For green spruce aphid abundance data, *C. furca* abundance data, and methane-flux
-data, because these datasets exhibited lower-tail dependence 
+data, because these datasets exhibited stronger lower- than upper-tail associations 
 (Table \ref{tab_multivar_summary}), we compared empirical and surrogate skewness values
 via a one-tailed test in the left tail: the $p$-value was the fraction
-of surrogate skewnesses less than the skewness for the empirical data; the test
-examines whether lower-tail dependence caused the spatial average to have significantly lower
-skewness than would have been expected without tail dependence. For
-leaf-curling plum aphid data, because that dataset exhibited upper-tail 
-dependence (Table \ref{tab_multivar_summary}), we did the analogous one-tailed test
-in the right tail. The test examines whether upper-tail dependence caused the spatial
+of surrogate skewnesses less than the skewness for the empirical data. The test
+examines whether stronger lower-tail associations between local time series 
+caused the spatial average to have significantly lower
+skewness than would have been expected with symmetric tail associations. For
+leaf-curling plum aphid data, because that dataset exhibited stronger upper-
+than lower-tail 
+associations (Table \ref{tab_multivar_summary}), we did the analogous one-tailed test
+in the right tail. The test examines whether upper-tail associations caused the spatial
 average to have significantly higher skewness than 
-expected without tail dependence.
-
-To further address Q3, we also examined a hypothesis that spatial 
-tail dependence of an environmental variable can influence the 
+would have been expected with symmetric tail associations.
+
+To further address Q3, we also examined a hypothesis that 
+asymmetric tail associations across space of an environmental variable can influence the 
 extinction risk of a metapopulation. We hypothesized that environmental noises
-exhibiting greater left-tail dependence across different habitat patches would 
+exhibiting greater left-tail associations across different habitat patches would 
 cause higher metapopulation extinction risks because then very bad years for the 
@@
 spatial correlations for every simulation, 
-but were made to exhibit right- or left-tail dependence between patches in
+but were made to exhibit stronger right- or left-tail associations between patches in
 different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
@@
 
+<!--***DAN: if you do decide to move the TL stuff to main text, this is where the 
+methods go. Other than this, I think I am done revising methods for Q3.-->
+
+<!--***DAN: Just ike in the Causes secton you added a comment saying these models
+were for initial exploration only and so we did not explore parameter space
+nor look at model variants, likewise we also need some commentary in this section 
+indicating that the modelling is simple and not comprehensively explored, because
+the purpose is just to get an idea of what kinds of consequences of copula 
+structure *may* pertain; further work is needed to understand to what extent these
+consequences are realized in real systems.-->
+
+Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
+tested whether the copula structure of the dependences between population variables 
+measured in different locations influences the spatial version of Taylor's
+law. Taylor's law is a commonly observed and widely studied 
+[@Taylor1961; @Taylor1988; @Cohen2013; @Xu2016; @Reuman2017] 
+empirical pattern that relates the 
+variances of groups, $g$, of population measurements to the means of the 
+groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
+$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
+the Taylor's law *exponent* or 
+*slope*, and $\log(a)$ is the Taylor's law intercept. 
+There are several versions of Taylor's law. For spatial Taylor's law,
+given population density or abundance data $x_i(t)$ measured in locations
+$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
+all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
+So means $m_g$ and variances $v_g$ are computed across space. We consider that 
+Taylor's law is validated for a dataset if the 
+$\log(v_g)$ versus $\log(m_g)$ scatter plot 
+for the dataset shows a linear, homoskedastic pattern.
+Taylor's law has been 
+verified empirically for hundreds of taxa and has been applied in numerous fields
+including fisheries management, estimation of species persistence times, and
+agriculture [@Cohen2015; @Reuman2017].
+
 # Results for Q3: Tail dependence influences skewness of the spatial average, extinction risk, and Taylor's law
@@
 \noindent Results were consistent with the hypothesis that the skewness
-of a spatial-average time series is influenced by tail dependence
-between the local quantities being averaged. Thus copula 
-structure and tail dependence are important for spatially averaged 
-quantities. This result is represented in Fig. \ref{MasterFigure}
-as the solid box around "Instability/skewness of mean or total 
-time series" and the solid arrows labeled "X". For datasets that
-exhibited strong lower-tail (respectively, upper-tail) 
-dependence in earlier analyses (green spruce 
-aphid abundance data for lower-tail, and leaf-curling plum aphid first flight data
-for upper-tail, Table 
+of a spatial-average time series is influenced by tail associations
+between the local quantities being averaged. 
+For datasets that
+exhibited highly asymmetric tail associations in earlier analyses (green spruce 
+aphid abundance data had stronger lower-tail association, and, 
+respectively,  leaf-curling plum 
+aphid first flight data had stronger upper-tail association, Table 
 \ref{tab_multivar_summary}), skewness of the spatial average 
@@
 surrogate skewness values 
-(Fig. \ref{fig_skewness_spearman}A,B). For datasets with moderate lower-tail 
+(Fig. \ref{fig_skewness_spearman}A,B). For datasets with moderately stronger 
+lower than upper-tail 
 dependence (\emph{Ceratium furca} abundance and methane data), 
@@
 toward being less than surrogate skewnesses (Fig. \ref{fig_skewness_spearman}C,D). 
+Thus copula 
+structure and asymmetric tail associations are important for spatially averaged 
+quantities. This result is represented in Fig. \ref{MasterFigure}
+as the solid box around "Instability/skewness of mean or total 
+time series" and the solid arrows labeled "X".
 
@@
 tendency for skewness of the spatial average to be affected as hypothesized 
-by tail dependence. \label{fig_skewness_spearman}}
+by asymmetric tail associations. \label{fig_skewness_spearman}}
 \end{center}
@@
 Consistent with our extinction risk hypothesis, 
-left-tail-dependent environmental fluctuations increased metapopulation
+left-tail-associated environmental fluctuations increased metapopulation
 extinction risk for the spatial 
-Lewontin-Cohen model (Fig. \ref{fig_extrisk_lcohen_local}), 
+Lewontin-Cohen model, 
 for $N=5$ and $N=25$ habitat patches, and for local 
-(Fig. \ref{fig_extrisk_lcohen_local}) and global (Fig. \ref{SM-fig_extrisk_lcohen_global})
-dispersal. This result is represented in Fig. \ref{MasterFigure}
+and global
+dispersal (Fig. \ref{fig_extrisk_lcohen_local}). This result 
+is represented in Fig. \ref{MasterFigure}
 as the dashed box around "Extinction risk" and the dashed arrow labelled "Y".
+
+<!--***DAN: the below is on TL and needs to be augmented if we keep TL results-->
 We also substantiated using models (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd}, Fig. \ref{SM-fig_spTL})
@@
 \begin{figure}[!h]
-\textbf{ \hspace{6.6 cm} (A) \hspace{4.2 cm} (B)} \\
+\textbf{ \hspace{3.6 cm} (A) \hspace{4.2 cm} (B)} \\
 \vspace{-0.8 cm}
@@
 \includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_local_disp.pdf}
+\end{center}
+\textbf{ \hspace{3.6 cm} (C) \hspace{4.2 cm} (D)} \\
+\vspace{-0.8 cm}
+\begin{center}
+\includegraphics[width=10 cm]{./Results/ext_risk_copula/lcohen_ext_risk_global_disp.pdf}
+\end{center}
 \caption{Extinction risk for the metapopulation extension of the Lewontin-Cohen model, after 
-$25$ time steps, was higher for left-tail-dependent environmental noise. Dispersal was local,
+$25$ time steps, was higher for environmental noise with stronger left-tail association
+across space. Dispersal was local (A, B) or global (C, D),
 with dispersal rate $d$ (Appendix \ref{SM-ext_risk_model}). Simulations used $N$ patches
-for $N=5$ (A) and $N=25$ (B).\label{fig_extrisk_lcohen_local}}
-\end{center}
+for $N=5$ (A, C) and $N=25$ (B, D).\label{fig_extrisk_lcohen_local}}
 \end{figure}

### 2019-10-06 SupportingInformation.Rmd
working on Q3 edits

@@
 
-<!--***Methods for generating evolutionary changes, keep this here-->
 \noindent The matrices from which root character states and changes 
@@
 
-<!--Move all the rest of it into the Discussion, I think, but can decide when I get to the Discussion. If it is mved, then you may want to retitle this appendix to reflect the 
+<!--***DAN: Move all the rest of it into the Discussion, I think, but can decide when I get to the Discussion. If it is mved, then you may want to retitle this appendix to reflect the 
 fact it is onl methods-->
@@
 <!--adding plots showing effects of asymmetric copula on extinction risk-->
+<!--Combined this with the main text fig for AER
 \begin{figure}[!h]
@@
 # Effects of non-normal copula structure on Taylor's law \label{Consequence_TL_simulation}
-
-\noindent Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
-tested whether the copula structure of the dependences between population variables 
-measured in different locations influences the spatial version of Taylor's
-law. Taylor's law is an extremely commonly observed and widely studied 
-[@Taylor1961; @Taylor1988; @Cohen2013; @Xu2016] empirical pattern that relates the 
-variances of groups, $g$, of population measurements to the means of the 
-groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
-$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
-the Taylor's law *exponent* or 
-*slope*, and we refer to $\log(a)$ as the Taylor's law intercept. 
-There are several versions of Taylor's law. For spatial Taylor's law,
-given population density or abundance data $x_i(t)$ measured in locations
-$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
-all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
-So means $m_g$ and variances $v_g$ are computed across space. We view the assumptions
-of Taylor's law to be validated for a dataset if the 
-$\log(v_g)$ versus $\log(m_g)$ scatter plot 
-for the dataset shows a linear, homoskedastic pattern.
-Taylor's law has been 
-verified empirically for hundreds of taxa and has been applied in numerous fields
-including fisheries management, estimation of species persistence times, and
-agriculture [@Cohen2015; @Reuman2017].
 

### 2019-10-06 Paper.Rmd
part way thrugh adding TL stuff to Q3 sections

@@
 
+We next addressed Q3 by exploring, using both data and models, three hypothesized
+ecological consequences of non-normal copula structure and asymmetric tail
+associations. As for Q2, models used are intended as initial explorations, only,
+of whether copula structure may possibly have a meaningful influence on important ecological
+phenomena. Therefore only simple models are used, and comprehensive explorations of the  
+sensitivity of results to model strucuture and parameters are left for future work.
+
 The hypothesis was presented in the Introduction that
@@
 
-<!--***DAN: if you do decide to move the TL stuff to main text, this is where the 
-methods go. Other than this, I think I am done revising methods for Q3.-->
-
-<!--***DAN: Just ike in the Causes secton you added a comment saying these models
-were for initial exploration only and so we did not explore parameter space
-nor look at model variants, likewise we also need some commentary in this section 
-indicating that the modelling is simple and not comprehensively explored, because
-the purpose is just to get an idea of what kinds of consequences of copula 
-structure *may* pertain; further work is needed to understand to what extent these
-consequences are realized in real systems.-->
-
 Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
@@
 So means $m_g$ and variances $v_g$ are computed across space. We consider that 
-Taylor's law is validated for a dataset if the 
+Taylor's law holds true for a dataset if the 
 $\log(v_g)$ versus $\log(m_g)$ scatter plot 
@@
 agriculture [@Cohen2015; @Reuman2017].
+
+<!--***DAN: continung to add TL stuff here, and then next in the following results section-->
 

### 2019-10-06 Paper.Rmd
added TL stuff to Q3 sections

@@
 
-We next addressed Q3 by exploring, using both data and models, three hypothesized
+We addressed Q3 by exploring, using both data and models, three hypothesized
 ecological consequences of non-normal copula structure and asymmetric tail
@@
 
-To help answer Q3, we tested the above hypothesis
+We tested the above hypothesis
 using our multivariate datasets (Table \ref{tab_data_info}).
@@
 
-Pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
+Finally, and pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
 tested whether the copula structure of the dependences between population variables 
-measured in different locations influences the spatial version of Taylor's
+measured in different locations has consequences for the spatial version of Taylor's
 law. Taylor's law is a commonly observed and widely studied 
@@
 all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
-So means $m_g$ and variances $v_g$ are computed across space. We consider that 
+So means $m_g$ and variances $v_g$ are computed across space. We refer to the matrix
+with $x_i(t)$ in the $t$th row and the $i$th column as a *population matrix*. 
+We consider that 
 Taylor's law holds true for a dataset if the 
@@
 including fisheries management, estimation of species persistence times, and
-agriculture [@Cohen2015; @Reuman2017].
-
-<!--***DAN: continung to add TL stuff here, and then next in the following results section-->
+agriculture [@Cohen2015; @Reuman2017], so the question of whether copula
+structure of spatial dependence influences Taylor's law may have applied
+significance.
+
+To explore the influence of copula structure and tail associations on spatial Taylor's
+law, we carried out a series of simulations that generated sets of population 
+matrices that had 
+distinct copula structure between locations but that 
+were otherwise statistically similar. For $k=1,\ldots,1000$
+and X representing a Clayton, normal, Frank, or a survival Clayton copula family,
+we generated $N \times n$ population matrices $m^{(\text{X},k)}$ with the following properties:
+1) for any given values of $j$ and $k$, the values 
+$\{ m_{t,j}^{(\text{X},k)} : t=1,\ldots,N  \}$ were exactly the same, as unordered
+sets, for all X, i.e., the same actual population values were used for location $j$ 
+in simulation $k$, regardless of X;
+2) for any given $k$, $j_1 \neq j_2$, the Spearman correlations (computed through time)
+$\cor_t(m_{t,j_1}^{(\text{X},k)},m_{t,j_2}^{(\text{X},k)})$ were the same, to within sampling
+variation, i.e., correlations through time of populations in two locations were the same,
+up to sampling variation, regardless of X;
+3) the copula for the dependence between $m_{t,j_1}^{(\text{X},k)}$ and $m_{t,j_2}^{(\text{X},k)}$ was X.
+We used $N=50$ and $n=25$. Details of how these matrices were generated are in 
+Appendices \ref{SM-Consequence_TL_simulation} and \ref{SM-sect:copsurrognd}.
+Thus for each $k$, the poulation matrices $m^{(\text{X},k)}$
+for X taking the values Clayton, normal, Frank, and survival Clayton
+were statistically similar except for the copula structure between locations, which was
+X. Thus comparing how Taylor's law may manifest differently for $m^{(\text{X},k)}$
+for different values of X constitutes a test of the influence of copula
+structure on Taylor's law. We selected X equal to the Clayton, normal, Frank, and survival 
+Clayton copulas to explore a range of tail association patterns.
+
+A variety of Taylor's law statistics were computed for each populaton matrix
+$m^{(\text{X},k)}$. First, the $N=50$ 
+spatial means,
+$m_g$, and variances, $v_g$, were computed, and $\log(v_g)$ vs. $\log(m_g)$ plots 
+were considered. Linearity of the $\log(v_g)$ vs. $\log(m_g)$ relationship was 
+tested for each simulation by comparing the linear regression
+through these $50$ points to a quadratic alternative via an $F$-test, 
+producing a $p$-value result. If 
+these $p$-values were uniformly distributed across the unit interval for the 1000 replicate 
+simulations which were generated for X, it supported the 
+linearity assumption of Taylor's law for X; whereas if 
+they were clustered toward smaller values
+the test tended to reject that assumption. We also tested, for each simulation, 
+the assumption of Taylor's 
+law that the $\log(v_g)$ vs. $\log(m_g)$ plot was homoskedastic: we regressed the absolute 
+residuals of the 
+linear regression of $\log(v_g)$ versus $\log(m_g)$ against the predictions of that 
+regression. A significant $p$-value
+result of this test indicates heteroskedasticity. For each simulation we also computed the
+root mean squared error of $\log(v_g)$ vs. $\log(m_g)$ data from the linear regression
+of $\log(v_g)$ versus $\log(m_g)$, as well as the 
+intercept and slope of the linear regression.
+Finally we recorded the quadratic coefficient of the regression of $\log(v_g)$ against
+$\log(m_g)$ and $(\log(m_g))^2$, and the mean curvature of 
+the quadratic regression equation itself
+across the values $\log(m_g)$. The distrubutions of all these 
+statistics across the 1000 replicate 
+simulations were compared for X the Clayton, normal, Frank, and survival Clayton copulas, 
+to determine if copula structure influences Taylor's law.
 
@@
 
-<!--***DAN: the below is on TL and needs to be augmented if we keep TL results-->
-We also substantiated using models (Appendix \ref{SM-Consequence_TL_simulation}, \ref{SM-sect:copsurrognd}, Fig. \ref{SM-fig_spTL})
-a hypothesis (Fig. \ref{MasterFigure}, arrow labelled "Z") that 
-Taylor's law can be influenced by
-tail dependence.
+Also consistent with hypothesis, copula structure had a substantial effect on
+Taylor's law. 
+Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
+linearity and homoskedasticity, by non-normal copula structure. For normal copula structure 
+(i.e., for simulations that gave normal-copula dependence between populatons
+in different locations), $p$-values for the linearty
+and homoskedasticity tests were roughly uniformly distributed across replicate 
+simulations and Taylor's law appeared visually to be a reasonable approximation
+of the $\log(v_g)$ versus $\log(m_g)$ relationship (Fig. \ref{fig_spTL}B,E,F). 
+Furthermore, quadratic coefficients
+and curvature values were close to $0$ (Fig. \ref{fig_spTL}J,K), and root mean squared
+errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
+But linearity or homoskedasticity were violated more frequently for non-normal 
+copula structure
+(Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
+non-zero (Fig. \ref{fig_spTL}J,K); and root mean squared
+errors from the linear regression were much higher (Fig. \ref{fig_spTL}G). Slopes 
+and intercepts of the linear regression were also strongly affected by copula
+structure (Fig. \ref{fig_spTL}H,I), though some of the effect here was because
+linear regressions do not always adequately represent the $\log(v_g)$ versus $\log(m_g)$
+relationship when copula structure was not normal.  
+Thus our results substantiated the hypothesis, at least for the models we used,
+that Taylor's law can be influenced by copula structure and asymmetric
+tail associations. This is represented in Fig. \ref{MasterFigure} by the 
+arrow labelled "Z".
 
@@
 \end{figure}
+
+<!--adding plots showing influences of non-normal copula on spatial Taylors law-->
+\begin{figure}[!h]
+\textbf{ \hspace{1.4cm} (A) \hspace{2.92 cm} (B) \hspace{2.92 cm} (C) \hspace{2.92 cm} (D)} \\
+\vspace{-0.5 cm} 
+\begin{center}
+\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png}\\
+\end{center}
+\textbf{ \hspace{0.6 cm} (E) \hspace{1.4 cm} (F) \hspace{1.4 cm} (G) \hspace{1.4 cm} (H) \hspace{1.4 cm} (I) \hspace{1.4 cm} (J) \hspace{1.4 cm} (K)}\\
+\vspace{-0.5 cm}
+\begin{center}
+\includegraphics[width=16 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
+\caption[Spatial Taylor's law results]{Spatial Taylor's law results. Log(spatial variance) vs. 
+log(spatial mean) relationships for $1000$ simulations over $50$ time steps for $25$ locations 
+using Clayton (A), normal (B), Frank (C), and survival Clayton (D) copula structures between
+locations. Results from all $1000$ 
+simulations were plotted on the same axes for each panel, but two example simulations are 
+shown in red and green to help assess to what extent variation on the plots was between 
+simulations or within simulations. For each simulation, we tested the linearity 
+(E) and homoskedasticity (F) of the log(variance) vs. log(mean) plot for that simulation, 
+and quantified the root mean squared error of points from the linear regression line (G).
+We also calculated the linear regression intercept (H) and slope (I), the quadratic
+term of a quadratic regression through the points on the log(variance) vs. log(mean) plot
+(J), and the curvature of that quadratic regression (K). Distributions of values 
+across all 1000 simulations are displayed. Spearman-preserving surrogates were used, 
+though results using Kendall-preserving surrogates were similar. See the methods
+section for Q3 and Appendices \ref{SM-Consequence_TL_simulation} and 
+\ref{SM-sect:copsurrognd} for details.\label{fig_spTL}}
+\end{center}
+\end{figure}
+<!--***DAN: Shyamolina, the spacing of the panel labels (A, B, C, etc.) is not quite 
+matched up with the spacing of the panels themselves, can you fix it please?-->
 

### 2019-10-06 SupportingInformation.Rmd
added TL stuff to Q3 sections

@@
 -->
-
-<!--adding plots showing influences of non-normal copula on spatial Taylors law-->
-\begin{figure}[!h]
-\textbf{ \hspace{2.8 cm} (A) \hspace{3.6 cm} (B) \hspace{3.8 cm} (C) \hspace{3.6 cm} (D)} \\
-\vspace{-0.5 cm} 
-\begin{center}
-\includegraphics[width=18 cm]{./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png}\\
-\end{center}
-\textbf{ \hspace{1.8 cm} (E) \hspace{1.6 cm} (F) \hspace{1.8 cm} (G) \hspace{1.8 cm} (H) \hspace{1.8 cm} (I) \hspace{1.8 cm} (J) \hspace{1.8 cm} (K)}\\
-\vspace{-0.5 cm}
-\begin{center}
-\includegraphics[width=18 cm]{./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf}
-\caption[Spatial Taylor's law results]{Spatial Taylor's law results. Log(spatial variance) vs. 
-log(spatial mean) relationships for $1000$ simulations over $50$ time steps for $25$ locations 
-using Clayton (A), normal (B), Frank (C), and survival Clayton (D) copula structures between
-locations. Results from all $1000$ 
-simulations were plotted on the same axes for each panel, but two example simulations are 
-shown in red and green to help assess to what extent variation on the plots was between 
-simulations or within simulations. For each simulation, we tested the linearity 
-(E) and homoskedasticity (F) of the log(variance) vs. log(mean) plot for that simulation, 
-and quantified the root mean squared error of points from the linear regression line (G).
-We also calculated the linear regression intercept (H) and slope (I), the quadratic
-term of a quadratic regression through the points on the log(variance) vs. log(mean) plot
-(J), and the curvature of that quadratic regression (K). Distributions of values 
-across all 1000 simulations are displayed. Spearman-preserving surrogates were used, 
-though results using Kendall-preserving surrogates were very similar. See 
-Appendix \ref{Consequence_TL_simulation} for 
-simulation and plotting details.\label{fig_spTL}}
-\end{center}
-\end{figure}
 

### 2019-10-07 Paper.Rmd
fig fixed: see response, please

@@
 \begin{figure}[!h]
-\textbf{ \hspace{3.5 cm} (A) \hspace{2 cm} (F) \hspace{2 cm} (K)} \\
-\vspace{-1 cm}
 \begin{center}
-\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeL_hist_taildep.pdf}\\
-\end{center}
-\vspace{-.33 cm}
-\textbf{ \hspace{3.5 cm} (B) \hspace{2 cm} (G) \hspace{2 cm} (L)} \\
-\vspace{-1 cm}
-\begin{center}
-\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatL_hist_taildep.pdf}\\
-\end{center}
-\vspace{-.33 cm}
-\textbf{ \hspace{3.5 cm} (C) \hspace{2 cm} (H) \hspace{2 cm} (M)} \\
-\vspace{-1 cm}
-\begin{center}
-\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/BivNormal_hist_taildep.pdf}\\
-\end{center}
-\vspace{-.33 cm}
-\textbf{ \hspace{3.5 cm} (D) \hspace{2 cm} (I) \hspace{2 cm} (N)} \\
-\vspace{-1 cm}
-\begin{center}
-\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/SomewhatR_hist_taildep.pdf}\\
-\end{center}
-\vspace{-.33 cm}
-\textbf{ \hspace{3.5 cm} (E) \hspace{2 cm} (J) \hspace{2 cm} (O)} \\
-\vspace{-1 cm}
-\begin{center}
-\includegraphics[width=9 cm]{./Results/BMR_results/evo_cop/ExtremeR_hist_taildep.pdf}\\
+\includegraphics[width=15 cm]{./Results/BMR_results/evo_cop/singlepage_hist_taildep.pdf}\\
 \caption[Non-parametric statistics on a mechanism of tail dependence in character distributions]{Three 
@@
 \end{figure}
-<!--***DAN: Shyamolina, can you please make this figure and its caption fit on one page,
-somehow?-->
 
@@
 \begin{figure}[!h]
-\textbf{ \hspace{6.6 cm} (A) \hspace{4.6 cm} (B)} \\
-\vspace{-0.4 cm}
+\textbf{ \hspace{5.6 cm} (A) \hspace{4 cm} (B)} \\
+\vspace{-0.6 cm}
 \begin{center}
@@
 \vspace{0.2 cm}
-\textbf{ \hspace{6.6 cm} (C) \hspace{4.6 cm} (D)} \\
-\vspace{-0.4 cm}
+\textbf{ \hspace{5.6 cm} (C) \hspace{4 cm} (D)} \\
+\vspace{-0.6 cm}
 \begin{center}
@@
 \includegraphics[width=5 cm]{./Results/skewness_results/skewness_methane/skewness_methane_spearman.pdf}
-\caption{Skewness of spatially averaged green spruce aphid abundance $(A)$, leaf-curling plum 
-aphid first flight dates $(B)$, \emph{Ceratium furca} abundance $(C)$, and methane-flux
-$(D)$ time series compared to a multivariate 
+\caption{Skewness of spatially averaged green spruce aphid abundance (A), leaf-curling plum 
+aphid first flight dates (B), \emph{Ceratium furca} abundance (C), and methane-flux
+(D) time series compared to a multivariate 
 normal-copula null hypothesis. Black dots are empirical 
@@
 \begin{figure}[!h]
-\textbf{ \hspace{3.6 cm} (A) \hspace{4.2 cm} (B)} \\
-\vspace{-0.8 cm}
+\textbf{ \hspace{5.6 cm} (A) \hspace{4 cm} (B)} \\
+\vspace{-1 cm}
 \begin{center}
@@
 \end{center}
-\textbf{ \hspace{3.6 cm} (C) \hspace{4.2 cm} (D)} \\
-\vspace{-0.8 cm}
+\textbf{ \hspace{5.6 cm} (C) \hspace{4 cm} (D)} \\
+\vspace{-1 cm}
 \begin{center}
@@
 \begin{figure}[!h]
-\textbf{ \hspace{1.4cm} (A) \hspace{2.92 cm} (B) \hspace{2.92 cm} (C) \hspace{2.92 cm} (D)} \\
-\vspace{-0.5 cm} 
+\textbf{ \hspace{2cm} (A) \hspace{2.92 cm} (B) \hspace{2.92 cm} (C) \hspace{2.92 cm} (D)} \\
+\vspace{-0.8 cm} 
 \begin{center}
@@
 \end{center}
-\textbf{ \hspace{0.6 cm} (E) \hspace{1.4 cm} (F) \hspace{1.4 cm} (G) \hspace{1.4 cm} (H) \hspace{1.4 cm} (I) \hspace{1.4 cm} (J) \hspace{1.4 cm} (K)}\\
-\vspace{-0.5 cm}
+\textbf{ \hspace{1 cm} (E) \hspace{1.4 cm} (F) \hspace{1.2 cm} (G) \hspace{1.2 cm} (H) \hspace{1.4 cm} (I) \hspace{1.4 cm} (J) \hspace{1.2 cm} (K)}\\
+\vspace{-0.8 cm}
 \begin{center}
@@
 <!--***DAN: Shyamolina, the spacing of the panel labels (A, B, C, etc.) is not quite 
-matched up with the spacing of the panels themselves, can you fix it please?-->
+matched up with the spacing of the panels themselves, can you fix it please?
+***Shya: check now, if is it ok?-->
 
@@
 
-<div id="refs"></div>
+<!--<div id="refs"></div>-->

### 2019-10-07 SupportingInformation.Rmd
fig fixed: see response, please

@@
 png("./Results/taylorslaw_results/SpatialTL_CNFSC_kendall.png",width=16,height=4,units="in",res=400)
-op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
+op<-par(mfrow=c(1,4),mar=c(8,8,2,1),mgp=c(4,1,0))
 ##plot TL for clayton
@@
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=1.5,cex.lab=1.5)  
+     cex.axis=2,cex.lab=3)  
 points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
-mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black")
+mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black",cex=2)
 
@@
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=1.5,cex.lab=1.5)  
+     cex.axis=2,cex.lab=3)  
 points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
-mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black")
+mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black",cex=2)
 
@@
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=1.5,cex.lab=1.5)  
+     cex.axis=2,cex.lab=3)  
 points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
-mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black")
+mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black",cex=2)
 
@@
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=1.5,cex.lab=1.5)  
+     cex.axis=2,cex.lab=3)  
 points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
-mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black")
+mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black",cex=2)
 
@@
 
-pdf("./Results/taylorslaw_results/SpatialTL_7stats_kendall.pdf",height=3,width=21)
-op<-par(mfrow=c(1,7),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+pdf("./Results/taylorslaw_results/SpatialTL_7stats_kendall.pdf",height=3,width=22)
+op<-par(mfrow=c(1,7),mar=c(8,8,2,1),mgp=c(3,1,0))
 #***DAN: x axis needs to be redone to show the copula names used in the surrogates
@@
 #}
-boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="Linearity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="p-vals", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Linearity test", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.hom distr for each set of surrogates
-boxplot(s.hom , col=rainbow(4), xlab="Copula family", ylab="Homoskedasticity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.hom , col=rainbow(4), xlab="Copula family", ylab="p-vals", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Homoskedasticity test", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.rmse distr for each set of surrogates
-boxplot(s.rmse , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL rmse", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.rmse , col=rainbow(4) , xlab="Copula family", ylab="rmse", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Spatial TL", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.int distr for each set of surrogates
-boxplot(s.int , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL intercept", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.int , col=rainbow(4) , xlab="Copula family", ylab="intercept", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Spatial TL", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.slope distr for each set of surrogates
-boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL slope", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="slope", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Spatial TL", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.quadcoeff distr for each set of surrogates
-boxplot(s.quadcoeff , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL quadratic coefficient", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.quadcoeff , col=rainbow(4) , xlab="Copula family", ylab="quadratic \n coefficient", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Spatial TL", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.mncurv distr for each set of surrogates
-boxplot(s.mncurv , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL mean curvature", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.mncurv , col=rainbow(4) , xlab="Copula family", ylab="mean \n curvature", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Spatial TL", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
@@
 png("./Results/taylorslaw_results/SpatialTL_CNFSC_spearman.png",height=4,width=16,units="in",res=400)
-op<-op<-par(mfrow=c(1,4),mar=c(4,5,1,1),mgp=c(3,1,0))
+op<-par(mfrow=c(1,4),mar=c(8,8,2,1),mgp=c(4,1,0))
 ##plot TL for clayton
@@
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=1.5,cex.lab=1.5)  
+     cex.axis=2,cex.lab=3)  
 points(spea.points.res[[1]][,1,1],spea.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(spea.points.res[[1]][,1,2],spea.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
-mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black")
+mtext(paste("Clayton"),side = 3, line=-2.5, adj=0.1, col="black",cex=2)
 
@@
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=1.5,cex.lab=1.5)  
+     cex.axis=2,cex.lab=3)  
 points(spea.points.res[[2]][,1,1],spea.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(spea.points.res[[2]][,1,2],spea.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
-mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black")
+mtext(paste("Normal"),side = 3, line=-2.5, adj=0.1, col="black",cex=2)
 
@@
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=1.5,cex.lab=1.5)  
+     cex.axis=2,cex.lab=3)  
 points(spea.points.res[[3]][,1,1],spea.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(spea.points.res[[3]][,1,2],spea.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
-mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black")
+mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black",cex=2)
 
@@
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=1.5,cex.lab=1.5)  
+     cex.axis=2,cex.lab=3)  
 points(spea.points.res[[4]][,1,1],spea.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
 points(spea.points.res[[4]][,1,2],spea.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
-mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black")
+mtext(paste("Survival Clayton"),side=3, line=-2.5, adj=0.1, col="black",cex=2)
 
@@
 
-pdf("./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf",height=3,width=21)
-op<-par(mfrow=c(1,7),mar=c(4.2,5,1,1),mgp=c(3,1,0))
+pdf("./Results/taylorslaw_results/SpatialTL_7stats_spearman.pdf",height=3,width=24)
+op<-par(mfrow=c(1,7),mar=c(8,9,4,1),mgp=c(4,1,0))
 x2<-c("C", "N", "F", "SC")
@@
 #plot s.lin distr for each set of surrogates
-boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="Linearity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="p-vals", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Linearity test", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.hom distr for each set of surrogates
-boxplot(s.hom , col=rainbow(4), xlab="Copula family", ylab="Homoskedasticity test p-vals", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.hom , col=rainbow(4), xlab="Copula family", ylab="p-vals", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Homoskedasticity test", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.rmse distr for each set of surrogates
-boxplot(s.rmse , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL rmse", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.rmse , col=rainbow(4) , xlab="Copula family", ylab="rmse", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Spatial TL", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.int distr for each set of surrogates
-boxplot(s.int , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL intercept", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.int , col=rainbow(4) , xlab="Copula family", ylab="intercept", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Spatial TL", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.slope distr for each set of surrogates
-boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL slope", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="slope", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Spatial TL", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.quadcoeff distr for each set of surrogates
-boxplot(s.quadcoeff , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL quadratic coefficient", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.quadcoeff , col=rainbow(4) , xlab="Copula family", ylab="quadratic \n coefficient", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Spatial TL", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 
 #plot s.mncurv distr for each set of surrogates
-boxplot(s.mncurv , col=rainbow(4) , xlab="Copula family", ylab="Spatial TL mean curvature", cex.axis=1.5,cex.lab=1.5,xaxt="n")
-axis(1, at = c(1:4), x2, cex.axis=1.5)
+boxplot(s.mncurv , col=rainbow(4) , xlab="Copula family", ylab="mean \n curvature", cex.axis=2,cex.lab=3,xaxt="n")
+mtext("Spatial TL", side=3, cex=1.8,line=1.5)
+axis(1, at = c(1:4), x2, cex.axis=2)
 

### 2019-10-07 explore_evo_cop.R

@@ -319,6 +319,69 @@ abline(v=0,lty="dashed")
 par(op)
 dev.off()
 
+#==================== single page figure ===============
+# Alternate plot
+
+resloc<-"./Results/BMR_results/evo_cop/"
+
+nametag<-"BivNormal"
+ansN<-readRDS(paste(resloc,nametag,"_NPA_results.RDS",sep=""))
+
+nametag<-"ExtremeL"
+ansXL<-readRDS(paste(resloc,nametag,"_NPA_results.RDS",sep=""))
+
+nametag<-"ExtremeR"
+ansXR<-readRDS(paste(resloc,nametag,"_NPA_results.RDS",sep=""))
+
+nametag<-"SomewhatL"
+ansSL<-readRDS(paste(resloc,nametag,"_NPA_results.RDS",sep=""))
+
+nametag<-"SomewhatR"
+ansSR<-readRDS(paste(resloc,nametag,"_NPA_results.RDS",sep=""))
+
+pdf(paste(resloc,"singlepage_hist_taildep.pdf",sep=""),height=9,width=15)
+op<-par(mfrow=c(3,5),mar=c(6,7,4,2),mgp=c(4.5,1,0))
+
+ans_list<-list(ansXL=ansXL,ansSL=ansSL,ansN=ansN,ansSR=ansSR,ansXR=ansXR)
+
+for(i in 1:5){
+  ans<-ans_list[[i]]
+  res<-ans$CorlmCoru_list
+  mres<-round(median(res),4)
+  n<-sum(res<=0)
+  h<-hist(res,border=F,col="grey",breaks = 10,xlab=expression(paste("cor" ["0,0.2"]-"cor"["0.8,1"])),
+          cex.lab=2.5,cex.axis=2,main=paste("n = ",n,sep=""),cex.main=2.5,xlim=c(-0.1,0.1))
+  abline(v=0,lty="dashed")
+  legend(x=-0.15,y=max(h$counts),legend=LETTERS[i],bty="n",cex=3)
+}
+
+for(i in 1:5){
+  ans<-ans_list[[i]]
+  res<-ans$PlmPu_list
+  mres<-round(median(res),4)
+  n<-sum(res<=0)
+  h<-hist(res,border=F,col="grey",breaks = 10,xlab=expression(paste("P" ["0,0.2"]-"P"["0.8,1"])),
[…]

### 2019-10-07 Paper.Rmd
finished an edit of Discussion, made a list of what else needs to be done

@@
 
+<!--***DAN: remianing to do:
+1) Edit abstract
+2) Go through the whole thing one more time
+3) send it out to coauthors
+4)make sure to say something about tropical systems
+5) maybe write to the editors saying this is coming soon?
+6) In the Discussion, you maybe need to point out not only that we
+use only bivariate copulas (and there are multivariate ones), but also
+that we use only continuous marginals. Probably an adequate approximation
+for our data, but see the other papers I've seen recently which use count data,
+and discrete marginals. Approaches such as that will be necessary for some 
+datasets
+7) change the box around the Cedar Creek data in the main fig
+-->
+
 # Abstract
-<!--***DANr: I have not edited this abstract at all since Ecol Monogr, it needs to be redone-->
+<!--***DAN: I have not edited this abstract at all since Ecol Monogr, it needs to be redone-->
 All branches of ecology study relationships among environmental and biological variables. However, 
@@
 @MaiScherer2017; @Anderson2018].
-<!--***DAN: dont forget to put these few words in!-->
 
@@
 \end{figure}
-
-<!--***DAN: In the Discussion, you need to point out not only that we
-use only bivariate copulas (and there are multivariate ones), but also
-that we use only continuous marginals. Probably an adequate approximation
-for our data, but see the other papers I've seen recently which use count data,
-and discrete marginals. Approaches such as that will be necessary for some 
-datasets-->
 
@@
 information on these and other copula families.
-
-<!--***DAN: Note: I am using tail dependence for the specific thing defined on p 62 of Joe,
-and tail associations as a more general term. I have changed this in parts of the ms
-upsteam from this point, but need to remember to change it as I move downstream.-->
 
@@
 for a few words on multivariate copulas. 
-<!--***DAN: dont forget to put these in!-->
 We present
@@
 mechanism analyzed below does not operate here.
-<!--***DAN: You may also want to have some additional commentary in the Discussion on
-how the models and related models can be explored more comprehensively. In fact you may already have some of that, though you may want to beef it up. You may
-want to insert a point at the end of the Methids section here again emphasizing
-that these are simple models for an initial exploration, and we discuss in the 
-Discussion how...-->
 
@@
 \includegraphics[width=15 cm]{./Results/BMR_results/evo_cop/singlepage_hist_taildep.pdf}\\
-\caption[Non-parametric statistics on a mechanism of tail dependence in character distributions]{Three 
-measures of asymmetry of tail associations ($\cor_{0,0.2}-\cor_{0.8,1}$, A-E; 
-$\Ps_{0,0.2}-\Ps_{0.8,1}$, F-J; $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ K-O) between two characters, across extant species,  were 
+\caption[Non-parametric statistics on a mechanism of tail dependence in 
+character distributions]{Three measures of asymmetry of tail 
+associations ($\cor_{0,0.2}-\cor_{0.8,1}$, A-E; 
+$\Ps_{0,0.2}-\Ps_{0.8,1}$, F-J; $\Dsq_{0.8,1}-\Dsq_{0,0.2}$, K-O) between two characters, across extant species,  were 
 computed for each of 100 simulations of mammalian character evolution for each of five types of tail dependence between 
@@
 in simulation $k$, regardless of X;
-2) for any given $k$, $j_1 \neq j_2$, the Spearman correlations (computed through time)
+2) for any given $k$ and $j_1$ and $j_2$ such that $j_1 \neq j_2$, the Spearman correlations (computed through time)
 $\cor_t(m_{t,j_1}^{(\text{X},k)},m_{t,j_2}^{(\text{X},k)})$ were the same, to within sampling
@@
 $\log(m_g)$ and $(\log(m_g))^2$, and the mean curvature of 
-the quadratic regression equation itself
+the quadratic regression equation 
 across the values $\log(m_g)$. The distrubutions of all these 
@@
 \end{figure}
-<!--***DAN: Shyamolina, the spacing of the panel labels (A, B, C, etc.) is not quite 
-matched up with the spacing of the panels themselves, can you fix it please?
-***Shya: check now, if is it ok?-->
 
@@
 <!--what we did and why is it important to study?-->
-We showed that non-normal copula structure and asymmetric tail dependence are 
+We showed that non-normal copula structure and asymmetric tail associations are 
 common across multiple sub-disciplines
-in ecology, although no research we know of has 
-taken this structure into account. We hypothesized 
-mechanisms that may cause non-normal copula structures and tail dependence;
+in ecology, although these facets of ecological data are only ocassionally 
+taken into account [@Valpine2014; @Anderson2018; @Popovic2018]. We hypothesized 
+mechanisms that may cause non-normal copula structures and asymmetric tail associations;
 we discuss below how commonly some of our
@@
 We also hypothesized and demonstrated important consequences of 
-non-normal copula structure and tail-dependence for the field of ecology. 
+non-normal copula structure and asymmetric tail associations for the field of ecology. 
 For instance, the skewness
-of a spatial-average time series is influenced by tail dependencies between its 
-constituent time series: right-tail-dependent local time series 
-lead to "spiky" spatially averaged time series, with large outbreaks.
-Thus tail dependence could have implications for exploited and pest species.
-Extinction risk can also be altered by tail dependence across space
+of a spatial-average time series is influenced by asymmetric tail associations between its 
+constituent time series: predominantly right-tail-associated local time series 
+can lead to "spiky" spatially averaged time series, with large outbreaks.
+Thus tail associations could have implications for exploited and pest species.
+Extinction risk and Taylor's law can also be altered by tail association patterns across space
 in environmental variables. 
 In our view, our results make it reasonable to suggest 
-that comprehensive understanding 
-of many ecological phenomena may be impossible without 
-a complete copula characterization of dependence.
-
+that a more comprehensive understanding 
+of many ecological phenomena may be possible if a
+complete copula characterization of dependence is employed. 
 Copula approaches are
-well developed [@joe2014_dependence; @nelsen2006_copula], and 
-computer implementations exist (`copula` and 
-`VineCopula` packages). Ecologists can apply these tools
+well developed [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], statistically, 
+and have been introduced in accessible formats [@Anderson2018; @Genest2007]. Furthermore, 
+open-source computer implementations exist (e.g., the `copula` and 
+`VineCopula` packages in the R programming language). Ecologists can apply these tools
 immediately. We created several interrelated randomization procedures
@@
 that build upon existing copula methods.
-To our knowledge, only normal copulas have 
-been used in ecology [@Valpine2014; @Anderson2018; @Popovic2018].
-Such use is equivalent to using normal distributions with transformed
-marginals; i.e., it is equivalent to an approach not using copulas at all, and
-therefore does not capitalize on the power of copula techniques.
+
+<!--Stuff about the mechanisms-->
 
 <!--a para on how likely one is to get copula structure via the Liebig mechanism-->
-Our first causal mechanism (Fig. \ref{MasterFigure}, A) probably operates commonly,
+Our first proposed causal mechanism (Fig. \ref{MasterFigure}, A) may well operate commonly,
 for two reasons.
@@
 Because these factors, which are the essential ingredients of
-the mechanism, are so common, it is reasonable to suppose the mechanism
-opperates commonly. It may be a dominant cause of 
-tail depencence and non-normal copula struture of ecological dependencies 
-across space.
-We provide some empirical support for the mechanism 
-in our discussion of green sprice aphids and
+the mechanism, are so common, it is reasonable to hypothesize that the mechanism
+may opperate commonly. It may be a dominant cause of 
+asymmetric tail associations and non-normal copula struture of ecological 
+dependencies across space.
+We provide some further support for the mechanism 
+in our discussion of green spruce aphids and
 winter temperature below.
@@
 <!--para on how likely one is to get copula structure via the Moran mechanism-->
-There are also reasons to expect that 
+There are also reasons to hypothesize that 
 our Moran mechanism (Fig. \ref{MasterFigure}, B) 
-operates commonly: Moran effects are common [@Liebhold2004], and 
-non-normal and tail-dependent copula structures are often found in 
+may operate commonly: Moran effects are common [@Liebhold2004], and 
+non-normal copula structures and asymmetric tail associations are often found in 
 environmental variables. If intense meteorological events are 
@@
 showed that a Gumbel copula effectively modeled 
-the spatial dependence of drought variables.
+the spatial dependence of drought variables; the Gumbel copula demonstrates 
+asymmetric tail associations.
 Bivariate copula analysis was also used in forecasting the 
@@
 We suggested in the Introduction that asymmetric competitive relationships
-between species could yield tail dependence between abundance data for the 
+between species could yield asymmetric tail associations between abundance measurements for the 
 species. This is another theoretical mechanism for non-normal copula
-structure and tail dependence, represented in Fig. \ref{MasterFigure} 
+structure and asymmetric tail associations, represented in Fig. \ref{MasterFigure} 
 by the box "Asymmetric species interactions" and the arrow labelled "C".
@@
 changes through evolutionary time in bird and 
-mammal BMR and body size may exhibit right-tail dependence,
+mammal BMR and body size may exhibit greater right- than left-tail association,
 contrary to standard normality assumptions of character evolution models. This is
-a hypothesis only, because the right-tail dependence of Fig. \ref{fig_biv_multi_raw_cop}F,G
-could have come about in another, unknown way. Our simulations show that tail dependence 
-in evolutionary changes is sufficient, but may not be necessary, to produce tail
-dependence in characters of extant species. For instance (see below, 
-Appendix \ref{SM-sect:missingdata}, Fig. \ref{SM-fig:missingdata}), 
+a hypothesis only, because the greater right-tail association shown in 
+Fig. \ref{fig_biv_multi_raw_cop}F,G
+could have come about in another, unknown way rather than via the mechanism we suggested
+which implicated asymmetric tail associations in evolutionary change. 
+Our simulations show that asymmetric tail associations 
+in evolutionary changes are sufficient, but may not be necessary, to produce 
+the observed asymmetric tail
+associations in characters of extant species. For instance (see below, and Appendix
+\ref{SM-sect:missingdata} and Fig. \ref{SM-fig:missingdata}), 
 systematically missing data can also produce tail dependence and may
@@
 evolutionary mechanism (Fig. \ref{MasterFigure}, D)
-is correct, our results only replace one question, i.e., why do we see right-tail
-dependence between BMR and body mass, with another,
-i.e., why might we see right-tail dependence in evolutionary changes in these traits?
-See Appendix \ref{SM-evol_mech} for further thoughts on potential importance of 
-copulas for character evolution.
+is correct, our results only replace one question, i.e., why do we see greater right-
+than left-tail
+association between BMR and body mass, with another,
+i.e., why might we see greater right- than left-tail association in evolutionary 
+changes in these traits?
+
+Performing statistical tests for associations between continuous traits in different species
+    was a primary motivating example of phylogenetic comparative methods.
+For example, Felsenstein's method of phylogenetically independent contrasts 
+    [@Felsenstein1985] is currently the second most-cited paper in the history
+    of the journal the American Naturalist [@HueyGT2019].
+Frustratingly, the field still lacks a dependable procedure for dealing with branch lengths, 
+    which are a crucial input to the method.
+Ideally, the branch lengths used to correct for phylogenetic effects would represent
+    the expected amounts of change for the characters that are being analyzed.
+Because researchers almost never have a reliable method for providing such branch lengths,
+    most researchers rely on ultrametric trees -- those for which 
+    the duration of the branch in time can be treated as a proxy for the branch length.
+Frequently these branch lengths are transformed to assess sensitivity to different 
+    assumptions about the degree of phylogenetic inertia displayed by the traits
+    under study [@Ives2018;@Harmon2018].
+Even if one were able to simply use a time-based set of branch lengths, assigning
+    dates to nodes in phylogenies is difficult.
+DNA sequence data can provide estimates of branch lengths, but these estimates
+    are dependent on the adequacy of models which correct sequences for multiple
+    substitutions occurring at the same location.
+Biases in estimating the evolutionary distance can affect downstream analyses [@Phillips2009].
+Additionally, changes
+    in the rate of molecular evolution make the estimation of dates difficult [@HeathM2014] even when branch lengths in time are accurately estimated.
+
+Without reliable branch length estimates, it is difficult to interpret the significance
+    of the magnitude of changes in traits across a tree.
+Developing tests of association based on copula structure may possibly lead the way to more
+    robust methods for studying associations when we lack defensible estimates of
+    branch lengths. 
+We note that our phylogenetic analyses here consisted merely of simulations to assess
+    whether interesting copula structure in a simple evolutionary process could 
+    leave a detectable signal on the trait data for extant species.
+Substantial work remains to be done before we have a copula-based method for 
+    analyzing data on a phylogenetic tree. 
+But the nonparametric nature of approaches
+    based solely on data ranks, as many of our copula methods are, may be a promising avenue for 
+    avoiding inaccuracies that arise via the highly structured model assumptions implicit
+    in the method of phylogenetically independent contrasts and related methods.
+
+Additionally, character evolution was simulated using one random draw from the relevant matrix
+per phylogeny branch. This was principally because branch lengths are often hugely
+uncertain. If branch lengths were well known, alternative simulation strategies may
+include selecting one draw from the matrix of evolutionary changes per unit time, 
+or selecting one draw for the whole branch but rescaling the variances of the selected
+character changes according to the 
+length of the branch. These choices amount to the same thing for normal copula
+structure, but not otherwise. Modelling choices such as these may have influenced
+our results. Additional research seems warranted testing the realism 
+of our hypothesized mechanism and simulations.
 
@@
 limitations, there may be a strong 
-association between F and $G_1$. For small F, there may be
+association between F and $G_1$, for instance. For small F, there may be
 little association because resources not allocated to F can instead be allocated 
@@
 either fecundity or juvenile survival.
-These ideas suggest that copulas will be 
+These ideas suggest that copulas may be 
 useful for studying multi-dimensional trade-offs. But applications will 
@@
 can bias conclusions (Appendix \ref{SM-sect:missingdata}, 
-Fig. \ref{SM-fig:missingdata}).
+Fig. \ref{SM-fig:missingdata}). Multi-dimensional copulas may also be the 
+appropriate copula approach
+to studying multi-dimensional trade-offs. We used only bivariate 
+analyses in this study solely in order to have at least one
+way to constrain the complexity of what is already a complex piece of work.
+But the statistical theory on multi-dimensional copulas is well developed
+[@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], 
+and open-source computer implementations exist for multivariate as well as 
+bivariate copulas (e.g., the `VineCopula`
+package for the R programming language). Such approaches may be an 
+useful next step in life-history theory and other areas.
 
@@
 additional mechanisms of copula structure is certainly worthy of future work. -->
-Other mechanisms probably also operate, 
-and should be considered. For instance, 
+Additional mechanisms of non-normal copula structure and asymmetric tail associations
+probably also operate, and should be considered. For instance, 
 measurement error may modify copula structure. 
 Our models were intentionally simple in other respects, 
-too, and did not include
+too, and did not include factors such as
 delayed density dependence, dispersal, population stage structure, trophic
-interactions, etc., and we did not comprehensively 
-explore parameter space.
-We hope by enumerating a few potential mechanisms of copula structure,
-we inspire additional research on the
+interactions, etc.; and we did not comprehensively 
+explore parameter space. We re-emphasize that fuller explorations in future work
+of some of our models, and variant model, are likely to be informative. 
+We hope by enumerating a few potential mechanisms of copula structure
+we will inspire additional research on the
 potentially numerous mechanisms that may operate in diverse datasets,
@@
 Our hypotheses and results cover the presence, causes, and consequences
-of non-normal copula structure and tail dependence in ecological 
-datasets (Fig. \ref{MasterFigure}).
+of non-normal copula structure and asymmetric tail associations in ecological 
+systems (Fig. \ref{MasterFigure}), but this is done using a variety of 
+datasets and models.
 We here take a closer look at the green spruce aphid, which simultaneously 
-illustrates causes and consequences.
+illustrates causes and consequences with one system.
 Green spruce aphid abundance, as measured in the data we use,
-is strongly influenced by temperature of the previous winter (@Sheppard2016,
-their supplementary fig. 6).
+is strongly influenced by the temperature of the previous winter (@Sheppard2016,
+their supplementary figure 6).
+Winter temperature for year $t$ was here taken to be an average for December of year $t-1$ through
+March of year $t$, was available for the locations of aphid sampling, 
+and was processed in the same way as @Sheppard2016.
 For each of our 10 sampling locations, we therefore 
-examined (Appendix \ref{SM-sect:greenspruce}, 
-Fig. \ref{SM-fig:greenspruce}, Table \ref{SM-tab:greenspruce}) the 
-copula of winter temperature
+examined the copula of winter temperature
 and aphid abundance time series for the location, finding 
-left-tail dependence. Apparently winter temperature has an asymmetric
-influence on aphid abundance in that cold winters produce low abundances
+stronger left- than right-tail associations in 7 of the 8 locations for which 
+independence of winter temperature and aphid abundance could be rejected, according to 
+the data specific to the location (Table \ref{tab:greenspruce}).
+Apparently winter temperature has an asymmetric
+influence on aphid abundance in that cold winters generally produce low abundances
 but warm winters often do not yield higher abundances than moderate winters.
@@
 dependence between green spruce aphid counts
-in different locations should show left-tail dependence. This is 
+in different locations should show stronger left- than right-tail associations. This is 
 exactly what was observed (Table \ref{tab_multivar_summary}), providing 
@@
 
+<!--***DAN: Shyamolina, can you please move what is currently Table S10 (Summary results
+for winter temp and green spruce aphid abundance) to here? Please also replace the two
+columns that show AICw values by the corresponding AIC values, consistent with the switches
+elsewhere in the paper of AICw for AIC.-->
+
 <!--a para on stability of the total biomass time series in a community, and skewness - sets up Shya's later paper as future work-->
@@
 spatial average of local time series is influenced 
-by their tail dependence. But the same logic should also apply to any collection of time
+by their tail associations. But the same logic should also apply to any collection of time
 series, whether associated with locations in space or not. 
-Another potential application is time series of all species from a single community, e.g., all
-plants in a quadrat surveyed repeatedly over time. A huge 
+Another potential application is time series of abundances of 
+all species from a single community, e.g., all
+plants in a quadrat surveyed repeatedly over time. A large 
 literature has focused on synchrony versus 
@@
 variability of community
-properties such as total biomass. Typically, variability of community biomass 
+properties such as total biomass (e.g., @Doak1998, @Tilman1999, @Tilman2006,
+@Gonzalez2009). Typically, variability of community biomass 
 is measured with the coefficient of variation, but skewness
@@
 structure of interspecific relationships in communities and its implications for community 
-variability is likely to be valuable.
+variability is likely to be valuable, in our opinion.
 
 <!--A para that sets up Shya's extinction risk paper-->
-Although we demonstrated that tail dependence between environmental variables
+Although we demonstrated that tail associations between environmental variables
 can influence extinction risk, substantial work remains to determine the importance 
@@
 when density dependence
-acts? Second, we considered metapopulation extinction risk, but the huge field 
+is involved? Second, we considered metapopulation extinction risk, but the large field 
 of population viability analysis (PVA) via stochastic 
@@
 vary stochastically through time due to environmental variation. Do relationships
-between different vital rates exhibit tail dependence, and does tail dependence 
-influence extinction risk in this context?
-Finally, is the copula structure of environmental variables
+between different vital rates exhibit asymmetric tail associations, 
+and do tail associations influence extinction risk in this context?
+Finally, is the copula structure of environmental variables or vital rates
 changing through time, and, if so, how do such changes influence extinction risks?
 Climate change is known to amplify the factors that lead to extreme weather events
-[@Hansen2012] and hence may alter spatial tail dependence for weather variables.
+[@Hansen2012] and hence may alter spatial tail associations for weather variables.
 
@@
 
-We thank the many contributors to the large datasets we used; D. Stevens and P. Verrier for data 
-extraction; and Joel E. Cohen, David Tilman, Lauren Hallet, Jonathan Walter, Thomas 
-Anderson, and Lei Zhao for helpful suggestions. We thank James Bell of the Rothamsted Insect Survey 
-(RIS). 
+We thank Joel E. Cohen for making us aware of copulas and their potential utility in the 
+first place. We thank the many contributors to the large datasets 
+we used; D. Stevens and P. Verrier for data 
+extraction; and David Tilman, Lauren Hallet, Jonathan Walter, Thomas 
+Anderson, Lei Zhao, and editors and annonymous reviewers for helpful suggestions. 
+We thank James Bell of the Rothamsted Insect Survey (RIS). 
 The RIS, a UK Capability, is funded by the Biotechnology and Biological Sciences Research Council 

### 2019-10-07 SupportingInformation.Rmd
finished an edit of Discussion, made a list of what else needs to be done

@@
 
-# Hypothetical evolutionary mechanism for tail dependence between traits \label{evol_mech}
+# Methods for the hypothetical evolutionary mechanism for tail dependence between traits \label{evol_mech}
 
@@
 respectively.
-
-<!--***DAN: Move all the rest of it into the Discussion, I think, but can decide when I get to the Discussion. If it is mved, then you may want to retitle this appendix to reflect the 
-fact it is onl methods-->
-Performing statistical tests for associations between continuous traits in different species
-    was a primary motivating example of phylogenetic comparative methods.
-For example, Felsenstein's method of phylogenetically independent contrasts 
-    [@Felsenstein1985] is currently the second most-cited paper in the history
-    of the journal the American Naturalist [@HueyGT2019].
-Frustratingly, the field still lacks a dependable procedure for dealing with branch lengths, 
-    which are a crucial input to the method.
-Ideally, the branch lengths used to correct for phylogenetic effects would represent
-    the expected amounts of change for the characters that are being analyzed.
-Because researchers almost never have a reliable method for providing such branch lengths,
-    most researchers rely on ultrametric trees -- those for which the branch
-    length can be treated as a proxy for the duration of the branch in time.
-Frequently these branch lengths are transformed to assess sensitivity to different 
-    assumptions about the degree of phylogenetic inertia displayed by the traits
-    under study [@Ives2018;@Harmon2018].
-Even if one were able to simply use a time-based set of branch lengths, assigning
-    dates to nodes in phylogenies is difficult.
-DNA sequence data can provide estimates of branch lengths, but these estimates
-    are dependent on the adequacy of models which correct sequences for multiple
-    substitutions occurring at the same location.
-Biases in estimating the evolutionary distance can affect downstream analyses [@Phillips2009].
-Additionally, changes
-    in the rate of molecular evolution make the estimation of dates difficult [@HeathM2014] even when branch lengths in time are accurately estimated.
-
-Without reliable branch length estimates, it is difficult to interpret the significance
-    of the magnitude of changes in traits across a tree.
-Developing tests of association based on copula structure may possibly lead the way to more
-    robust methods for studying associations when we lack defensible estimates of
-    branch lengths. 
-We note that our phylogenetic analyses here consisted merely of simulations to assess
-    whether interesting copula structure in a simple evolutionary process could 
-    leave a detectable signal on the trait data for extant species.
-Substantial work remains to be done before we have a copula-based method for 
-    analyzing data on a phylogenetic tree. 
-But the nonparametric nature of approaches
-    based solely on data ranks, as some of our copula methods are, may be a promising avenue for 
-    avoiding inaccuracies that arise via the highly structured model assumptions implicit
-    in the method of phylogenetically independent contrasts and related methods.
-
-Additionally, character evolution was simulated using one random draw from the relevant matrix
-per phylogeny branch. This was principally because branch lengths are often hugely
-uncertain. If branch lengths were well known, alternative simulation strategies may
-include selecting one draw from the matrix of evolutionary changes per unit time, 
-or selecting one draw for the whole branch but rescaling the variances of the selected
-character changes according to the 
-length of the branch. These choices amount to the same thing for normal copula
-structure, but not otherwise. Modelling choices such as these may have influenced
-our results. Additional research also seems warranted testing the realism 
-of our hypothesized mechanism and simulations.
 
@@
 
-We examined using simulations whether copula structure influences spatial Taylor's law, as 
-follows. First, an $n \times n$ covariance matrix $\Sigma$ was created
+As described in the main text, we examined using simulations 
+whether copula structure influences spatial Taylor's law; details are provided here. 
+First, an $n \times n$ covariance matrix $\Sigma$ was created
 with unit diagonal entries 
@@
 $n$ locations over $N$ time steps. We used $N=50$.
-The cumulative distribution function
-(cdf) of the standard normal distribution was applied to these numbers, followed by the 
+The cdf of the standard normal distribution was applied to these numbers, followed by the 
 inverse of the cdf of a gamma distribution with shape and scale parameters $7.5$ and $1$. 
@@
 copula structure of dependences between populations in different locations
-was normal for this array.
+was normal for this array. The array contains $1000$ population matrices.
 
@@
 except for sampling variation. Thus, to summarize using mathematical notation,
-if $m^{(\text{N})}$ is our normal-copula population array and $m^{(\text{C})}$ is our 
+if $m^{(\text{N},k)}_{t,j}$ is our normal-copula population array (the three dimensions 
+of the array are indexed, respectively, by $t$, $j$ and $k$, and $m^{(\text{C},k)}_{t,j}$ is our 
 Clayton-copula population array, then,
-for any given $i$ and $j$, the sets $\{m^{(\text{N})}_{tij}: t=1,\ldots,N\}$ and 
-$\{m^{(\text{C})}_{tij}: t=1,\ldots,N\}$ are exactly the same (when considered 
+for any given $j$ and $k$, the sets $\{m^{(\text{N},k)}_{t,j}: t=1,\ldots,N\}$ and 
+$\{m^{(\text{C},k)}_{t,j}: t=1,\ldots,N\}$ are exactly the same (when considered 
 as unordered sets); 
-and for any given $i_1$, $i_2$ and $j$, the Spearman correlations $\cor_t(m^{(\text{C})}_{ti_1j},m^{(\text{C})}_{ti_2j})$
-and $\cor_t(m^{(\text{N})}_{ti_1j},m^{(\text{N})}_{ti_2j})$ are the same, to within 
+and for any given $j_1$, $j_2$ and $k$ with $j_1 \neq j_2$, the Spearman correlations $\cor_t(m^{(\text{C},k)}_{t,j_1},m^{(\text{C},k)}_{t,j_2})$
+and $\cor_t(m^{(\text{N},k)}_{t,j_1},m^{(\text{N},k)}_{t,j_2})$ are the same, to within 
 sampling variation. The whole exercise was repeated preserving Kendall instead of 
 Spearman correlations.
-
-A variety of Taylor's law statistics were computed for each simulation $m^{(X)}_{tij}$ 
-($i=1,\ldots,n$, $t=1,\ldots,N$) for each copula structure $X$ used. First, the $N=50$ 
-spatial means,
-$m_t$, and variances, $v_t$, were computed, and $\log(v_t)$ vs. $\log(m_t)$ plots 
-were considered. As a shorthand, denote $y=\log(v_t)$ and $x=\log(m_t)$ for a 
-single simulation. Linearity of the $y$ vs. $x$ relationship was 
-tested for each simulation by comparing the linear regression
-through these $50$ points to a quadratic alternative via an $F$-test, 
-producing a $p$-value result. If 
-these $p$-values were uniformly distributed across the unit interval for the 1000 replicate 
-simulations, it supported the linearity assumption of Taylor's law; whereas if 
-they were clustered toward smaller values
-the test tended to reject that assumption. We also tested, for each simulation, 
-the assumption of Taylor's 
-law that the $y$ vs. $x$ plot was homoskedastic: we regressed the absolute residuals of the 
-linear regression of $y$ versus $x$ against the predictions of that regression. 
-A significant $p$-value
-result of this test indicates heteroskedasticity. For each simulation we also computed the
-root mean squared error of $y$ vs. $x$ data from the linear regression
-of $y$ versus $x$, as well as the intercept and slope of the linear regression.
-Finally we recorded the quadratic coefficient of the regression of $y$ against
-$x$ and $x^2$, and the mean curvature of the quadratic regression equation itself
-across the values $x$. The distrubutions of all these statistics across 1000 replicate 
-simulations were compared for the Clayton, normal, Frank, and survival Clayton simulations to determine if copula structure 
-influences Taylor's law.
-
-Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
-linearity and homoskedasticity, by non-normal copula structure. For normal copula structure 
-(i.e., for simulations in our normal-copula population array), $p$-values for the linearty
-and homoskedasticity tests were roughly uniformly distributed across replicate 
-simulations and Taylor's law appeared visually to be a reasonable approximation
-of the $\log(v_t)$ versus $\log(m_t)$ relationship (Fig. \ref{fig_spTL}B,E,F). 
-Furthermore, quadratic coefficients
-and curvature values were close to $0$ (Fig. \ref{fig_spTL}J,K), and root mean squared
-errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
-But linearity or homoskedasticity were violated more frequently for non-normal copula structure
-(Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
-non-zero (Fig. \ref{fig_spTL}J,K); and root mean squared
-errors from the linear regression were much higher (Fig. \ref{fig_spTL}G). Slopes 
-and intercepts of the linear regression were also strongly affected by copula
-structure (Fig. \ref{fig_spTL}H,I), though some of the effect here was because
-linear regressions do not always adequately represent the $\log(v_t)$ versus $\log(m_t)$
-relationship when copula structure was not normal.
 
@@
 
+<!--Moved this stuff to main text for AER
 # From causes to consequences for green spruce aphid abundance, details \label{sect:greenspruce}
@@
 (see section on methods for Q1) are in Table \ref{tab:greenspruce}.
+-->
 
@@
       caption = "Summary results for analysis of copulas for winter temperature versus green spruce 
-aphid abundance for each of the 10 sampling locations. The column labeled p, CvM is the p-value 
-result for the goodness of fit test using the Cramer-von Mises statistic (see section on 
-methods for Q1). The 
+aphid abundance for each of the 10 sampling locations, using the same model selection
+and nonparametric methods detailed in the section on methods for Q1. The column labeled p, CvM is the p-value 
+result for the goodness of fit test using the Cramer-von Mises statistic. The 
 column labeled p, KS is the p-value 
@@
 
+<!--Decided to cut this for AER
 \begin{figure}[!h]

### 2019-10-07 Paper.Rmd
table fixed

@@
 
-<!--***DAN: Shyamolina, can you please move what is currently Table S10 (Summary results
-for winter temp and green spruce aphid abundance) to here? Please also replace the two
-columns that show AICw values by the corresponding AIC values, consistent with the switches
-elsewhere in the paper of AICw for AIC.-->
+```{r tab_greenspruceaphid_temp_summary, echo=F, results='asis',message=F}
+library(kableExtra)
+ms_npa_RES_greenspruceaphid_count_temp<-readRDS("./Results/aphid_count_winter_temp_results/ms_npa_RES_greenspruceaphid_count_temp.RDS")
+res<-ms_npa_RES_greenspruceaphid_count_temp$sp10_Green_spruce_aphid$ms_summary_allloc
+#row.names(res)<-NULL
+#res<-cbind(c(2:11),res)
+colnames(res)<-c("p-value, independence test","Best-fit copula",
+                 "Best-fit AIC","Normal AIC","p, CvM","p, KS","Model-avg. LT","Model-avg. UT","Model-avg. LT minus UT","cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$")
+res<-res[,-1]
+res<-cbind("Site"=c(2:11),res)
+knitr::kable(res,"latex", booktabs = T, linesep = "",digits=3, escape=F,longtable=T,row.names = F,
+             caption.short="Summary results for copulas for winter temperature versus green spruce aphid abundance",
+      caption = "Summary results for analysis of copulas for winter temperature versus green spruce 
+aphid abundance for each of the 10 sampling locations, using the same model selection
+and nonparametric methods detailed in the section on methods for Q1. The column labeled p, CvM is the p-value 
+result for the goodness of fit test using the Cramer-von Mises statistic. The 
+column labeled p, KS is the p-value 
+result for the goodness of fit test using the Kolmogorov-Smirnov statistic. Entries with an NA are because model selection and other statistics
+were only employed for copulas for which independence was rejected (5 percent significance level).\\label{tab:greenspruce}") %>% column_spec(1:12,width="1.1 cm")#%>%
+#kable_styling(latex_options = c("scale_down"))
+```
 

### 2019-10-07 SupportingInformation.Rmd
table fixed

@@
     m<-as.data.frame(matrix(NA,1,12))
-    colnames(m)<-c("p.Indep","minAIC.bestfit.cop","Bestfit.cop.AICw",
-                   "Normal.cop.AICw","p.CvM","p.KS",
+    colnames(m)<-c("p.Indep","minAIC.bestfit.cop",
+                   "Bestfit.cop.AIC", "Normal.cop.AIC",
+                   "p.CvM","p.KS",
                    "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
@@
       m$minAIC.bestfit.cop<-ms_res$InfCritRes$copname[which.min(ms_res$InfCritRes$AIC)]
-      m$Bestfit.cop.AICw<-round(max(ms_res$InfCritRes$AICw),3)
-      m$Normal.cop.AICw<-round(ms_res$InfCritRes$AICw[1],3)
+
+      m$Bestfit.cop.AIC<-round(min(ms_res$InfCritRes$AIC),3)
+      m$Normal.cop.AIC<-round(ms_res$InfCritRes$AIC[1],3)
+
       m$p.CvM<-ms_res$GofRes_CvM
@@
 saveRDS(ms_npa_RES_greenspruceaphid_count_temp,"./Results/aphid_count_winter_temp_results/ms_npa_RES_greenspruceaphid_count_temp.RDS")
-```
-
-```{r tab_greenspruceaphid_temp_summary, echo=F, results='asis',message=F}
-library(kableExtra)
-set.seed(seed)
-
-res<-ms_npa_RES_greenspruceaphid_count_temp$sp10_Green_spruce_aphid$ms_summary_allloc
-#row.names(res)<-NULL
-#res<-cbind(c(2:11),res)
-colnames(res)<-c("p-value, independence test","Best-fit copula",
-                 "Best-fit copula AICw","Normal copula AICw","p, CvM","p, KS","Model-avg.
-LT","Model-avg.
-UT","Model-avg.
-LT minus UT","cor$_{l}$ - cor$_{u}$","P$_{l}$ - P$_{u}$","D$_{u}^2$ - D$_{l}^2$")
-res<-res[,-1]
-knitr::kable(res,"latex", booktabs = T, linesep = "",digits=3, escape=F,longtable=T,
-             caption.short="Summary results for copulas for winter temperature versus green spruce aphid abundance",
-      caption = "Summary results for analysis of copulas for winter temperature versus green spruce 
-aphid abundance for each of the 10 sampling locations, using the same model selection
-and nonparametric methods detailed in the section on methods for Q1. The column labeled p, CvM is the p-value 
-result for the goodness of fit test using the Cramer-von Mises statistic. The 
-column labeled p, KS is the p-value 
-result for the goodness of fit test using the Kolmogorov-Smirnov statistic. Entries with an NA are because model selection and other statistics
-were only employed for copulas for which independence was rejected (5 percent significance level).\\label{tab:greenspruce}") %>% column_spec(1:12,width="1.2 cm")#%>%
-#kable_styling(latex_options = c("scale_down"))
 ```

### 2019-10-09 Paper.Rmd
edits to to-do list

@@
 2) Go through the whole thing one more time
-3) send it out to coauthors
+3) send it out to coauthors (done) and maybe get feedback
 4)make sure to say something about tropical systems
@@
 and discrete marginals. Approaches such as that will be necessary for some 
-datasets
+datasets.
 7) change the box around the Cedar Creek data in the main fig
+8) say in Discussion that you could have used non-ranks for the nonparam stats - this
+is the stuff the Ecol Mongr referee talked about. Maybe this should go in Background instead?
+9) maybe go back and read the Ecol Monogr reviews again
 -->

### 2019-10-10 Paper.Rmd
fixed typo in definition of lower-tail dependence
edited the abstract

@@
 <!--***DAN: remianing to do:
-1) Edit abstract
+DONE 1) Edit abstract
 2) Go through the whole thing one more time
@@
 # Abstract
-<!--***DAN: I have not edited this abstract at all since Ecol Monogr, it needs to be redone-->
 All branches of ecology study relationships among environmental and biological variables. However, 
@@
 based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, 
-but not in ecology. We here explore the concepts that underpin copulas and examine the potential 
+but rarely in ecology. We here explore the concepts that underpin copulas and examine the potential 
 for those concepts to improve our understanding of ecology. We find that informative copula 
-structure in dependencies between variables is exceedingly common across all the environmental, 
+structure in dependencies between variables is common across all the environmental, 
 species-trait, phenological, population, community, and ecosystem functioning datasets we 
-considered. Many datasets exhibited asymmetric tail dependence, whereby two variables are more 
+considered. Many datasets exhibited asymmetric tail associations, whereby two variables are more 
 strongly related in their left compared to right tails, or *vice versa*. We describe mechanisms by which 
-observed copula structure and tail dependence can arise in ecological data, including a Moran-like 
+observed copula structure and asymmetric tail associations can arise in ecological data, including a Moran-like 
 effect whereby dependence structures between environmental variables are inherited by ecological 
 variables; and asymmetric or nonlinear influences of environments on ecological variables, such as 
-under Liebig's law of the minimum. Copula structure and tail 
-dependence can also help reveal causal relationships between variables. We also describe consequences of copula structure for ecological phenomena, including impacts 
-on extinction risk and the stability through time of ecosystem services. By documenting the 
-importance of a complete description of dependence, advancing conceptual frameworks, and 
-demonstrating a powerful approach, we aim to elevate copula reasoning to widespread use in ecology, benefiting the discipline.
+under Liebig's law of the minimum. We also describe consequences of copula structure for 
+ecological phenomena, including impacts 
+on extinction risk, Taylor's law, the stability through time of ecosystem services. By documenting the 
+importance of a complete description of dependence between variables, advancing conceptual frameworks, and 
+demonstrating a powerful approach, we aim to encourage widespread use of copulas in ecology, which we
+believe can benefit the discipline.
 
@@
 Lower tail dependence is defined analogously, as 
-$\lambda_L = \lim_{u \rightarrow 1^-} P[Y \leq F_Y^{-1}(u) | X \leq F_X^{-1}(u)]$.
+$\lambda_L = \lim_{u \rightarrow 0^+} P[Y \leq F_Y^{-1}(u) | X \leq F_X^{-1}(u)]$.
 Tail dependence is a property of the random variable $(U,V)$, so depends only

### 2019-10-10 Paper.Rmd
added discussion paras

@@
 4)make sure to say something about tropical systems
-5) maybe write to the editors saying this is coming soon?
+DONE 5) maybe write to the editors saying this is coming soon?
 6) In the Discussion, you maybe need to point out not only that we
@@
 datasets.
-7) change the box around the Cedar Creek data in the main fig
+DONE 7) change the box around the Cedar Creek data in the main fig
 8) say in Discussion that you could have used non-ranks for the nonparam stats - this
@@
 \cor_{l_b,u_b}(u,v)=\frac{\sum 
-   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}},
+   (u_i-\mean(u)) (v_i-\mean(v))}{(n-1)\sqrt{\var(u)\var(v)}},\label{eq:partialspearman}
 \end{equation}
@@
 
+<!--a para on stability of the total biomass time series in a community, and skewness - sets up Shya's later paper as future work-->
+We also elaborated potential consequences of copula structure for ecological understanding. 
+Our results showed that the skewness of the 
+spatial average of local time series is influenced 
+by their tail associations. But the same logic should also apply to any collection of time
+series, whether associated with locations in space or not. 
+Another potential application is time series of abundances of 
+all species from a single community, e.g., all
+plants in a quadrat surveyed repeatedly over time. A large 
+literature has focused on synchrony versus 
+compensatory dynamics between such time series, and 
+the influence of interspecific relationships on the 
+variability of community
+properties such as total biomass (e.g., @Doak1998, @Tilman1999, @Tilman2006,
+@Gonzalez2009). Typically, variability of community biomass 
+is measured with the coefficient of variation, but skewness
+may also be of interest because it can help characterize "spikiness" 
+through time. Future work on copula 
+structure of interspecific relationships in communities and its implications for community 
+variability is likely to be valuable, in our opinion.
+
+<!--A para that sets up Shya's extinction risk paper-->
+Although we demonstrated that tail associations between environmental variables
+can influence extinction risk, substantial work remains to determine the importance 
+of this effect. First, we used 
+a non-density-dependent model. Do similar results pertain
+when density dependence
+is involved? Second, we considered metapopulation extinction risk, but the large field 
+of population viability analysis (PVA) via stochastic 
+matrix modelling [@Caswell2000; @MorrisDoak2002]
+uses a framework in which a single population's 
+vital rates (e.g., life-stage-specific 
+fecundity and survival rates) are considered to
+vary stochastically through time due to environmental variation. Do relationships
+between different vital rates exhibit asymmetric tail associations, 
+and do tail associations influence extinction risk in this context?
+Finally, is the copula structure of environmental variables or vital rates
+changing through time, and, if so, how do such changes influence extinction risks?
+Climate change is known to amplify the factors that lead to extreme weather events
+[@Hansen2012] and hence may alter spatial tail associations for weather variables.
+
 <!--green spruce aphid all the way from causes to consequences-->
@@
 
-<!--a para on stability of the total biomass time series in a community, and skewness - sets up Shya's later paper as future work-->
-Our results showed that the skewness of the 
-spatial average of local time series is influenced 
-by their tail associations. But the same logic should also apply to any collection of time
-series, whether associated with locations in space or not. 
-Another potential application is time series of abundances of 
-all species from a single community, e.g., all
-plants in a quadrat surveyed repeatedly over time. A large 
-literature has focused on synchrony versus 
-compensatory dynamics between such time series, and 
-the influence of interspecific relationships on the 
-variability of community
-properties such as total biomass (e.g., @Doak1998, @Tilman1999, @Tilman2006,
-@Gonzalez2009). Typically, variability of community biomass 
-is measured with the coefficient of variation, but skewness
-may also be of interest because it can help characterize "spikiness" 
-through time. Future work on copula 
-structure of interspecific relationships in communities and its implications for community 
-variability is likely to be valuable, in our opinion.
-
-<!--A para that sets up Shya's extinction risk paper-->
-Although we demonstrated that tail associations between environmental variables
-can influence extinction risk, substantial work remains to determine the importance 
-of this effect. First, we used 
-a non-density-dependent model. Do similar results pertain
-when density dependence
-is involved? Second, we considered metapopulation extinction risk, but the large field 
-of population viability analysis (PVA) via stochastic 
-matrix modelling [@Caswell2000; @MorrisDoak2002]
-uses a framework in which a single population's 
-vital rates (e.g., life-stage-specific 
-fecundity and survival rates) are considered to
-vary stochastically through time due to environmental variation. Do relationships
-between different vital rates exhibit asymmetric tail associations, 
-and do tail associations influence extinction risk in this context?
-Finally, is the copula structure of environmental variables or vital rates
-changing through time, and, if so, how do such changes influence extinction risks?
-Climate change is known to amplify the factors that lead to extreme weather events
-[@Hansen2012] and hence may alter spatial tail associations for weather variables.
+<!--Alternative methods-->
+
+<!--Partial Person correlation-->
+We carried out most of our analyses on data ranks, $u_i$ and $v_i$, for good reasons 
+int he Inrtoduction, Background, and methods sections and 
+reviewed here; but under some circumstances it may also be appropriate to use 
+techniques which parallel our nonparametric statistics but that use unranked data, 
+$x_i$, $y_i$. As mentioned previously, @Genest2007 and others
+recommend carrying out inferences about dependence structures (which was our
+goal) using normalized ranks, stating explicitly "statistical inference concerning
+dependence structures should always be based on ranks." The reason for this is
+that measures of dependence which use unranked data conflate information about
+the maginal distributions of variables with information about the dependence
+between variables. To provide an ecological example, suppose the true population densities $p_{1,t}$
+and $p_{2,t}$ of two species of fish in a lake in year $t$ are unknown,
+but are assessed using catch per unit effort (CPUE), a standard approach in fisheries
+science. The CPUE measurements, which we denote $c_{1,t}$ and $c_{2,t}$ may be more
+or less correlated through time than the true densities would be, if they were known,
+if the function $f$ relating $p_{i,t}$ and $c_{i,t}$ is nonlinear and if Pearson
+correlation is used. The function $f$ modifies the marginal distributions, which modifies
+measures such as Pearson correlation that conflate dependence and marginal information.
+Rank-based measures such as Spearman or Kendall correlation, or any of our copula approaches,
+will be unaffected by nonlinear functions such as $f$. The same difficulty will pertain in any
+case for which a measurement of an ecological quantity is a nonlinear index of the true quantity
+of interest. Nevertheless, if one is certain that measurements are linearly related to 
+the true quantities being measured, statistics based on unranked data may be more appropriate 
+under some circumstances, in part because ecological mechanisms which produce relationships 
+between variables are influenced not only by the ordinal informatio in ranks but also 
+by the cardinal information in the unranked data. Judgements of what statistics
+are appropriate depend, of course, on the purpose of the analysis; the goal of this 
+study was to study dependence in isolation, so we used ranks. If one is interested
+in tail associations using cardinal data, a partial Pearson correlation can developed
+straightforwardly by replacing $u$ and $v$ in equation \ref{eq:partialspearman}
+by the $x$ and $y$, but still computing the sum in that equation over data points
+with ranks, $u_i$ and $v_i$ constrained by the bounds $u_i+v_i>2l_b$ and 
+$u_i+v_i<2u_b$. This would result in a different approach to
+tail associations that would conflate dependence and marginal information but 
+that may also be useful for understanding ecological relationships.
+
+<!--6) In the Discussion, you maybe need to point out not only that we
+use only bivariate copulas (and there are multivariate ones), but also
+that we use only continuous marginals. Probably an adequate approximation
+for our data, but see the other papers I've seen recently which use count data,
+and discrete marginals. Approaches such as that will be necessary for some 
+datasets.-->
+Copula methods are numerous, and go well beyond the cases we have considered
+in this introductory paper; we hope our work helps inspire applications of copulas
+in ecology beyond the specific tools we have employed. For instance, multivariate copulas are useful for 
+studying relationships between multiple interacting variables, and have been used in finance
+to model tail risk for portfolio optimization problems [@Czado2019; @Joe2011]. 
+Analogous ecological applications
+exist that may be amenable to the same approach, e.g., ecosystem functioning variables
+such as total primary productivity or carbon flux are the sum of the contributions of
+multiple species in the same way that athe value of an investment portfolio is the sum 
+of the values of its constituent assets. Ecologists may likewise be interested in
+managing the risk that an ecosystem functioning variable will take an extreme value. 
+As another example, we assumed for this paper that univariate marginal 
+distributions have continuous, strictly 
+monotonic distribution functions; this was for simplicity and because such a simpler
+approach was sufficient for our research questions. But count data in ecology have been analyzed
+with approaches not making this assumption [@Anderson2018].
 

### 2019-10-11 Paper.Rmd
adding some stuff to Discussion

@@
 2) Go through the whole thing one more time
-3) send it out to coauthors (done) and maybe get feedback
+3) send it out to coauthors (done) and maybe get feedback, ESP FROM lAWRENCE
 4)make sure to say something about tropical systems
 DONE 5) maybe write to the editors saying this is coming soon?
-6) In the Discussion, you maybe need to point out not only that we
+DONE 6) In the Discussion, you maybe need to point out not only that we
 use only bivariate copulas (and there are multivariate ones), but also
@@
 datasets.
-DONE 7) change the box around the Cedar Creek data in the main fig
-8) say in Discussion that you could have used non-ranks for the nonparam stats - this
+7) change the box around the Cedar Creek data in the main fig
+DONE 8) say in Discussion that you could have used non-ranks for the nonparam stats - this
 is the stuff the Ecol Mongr referee talked about. Maybe this should go in Background instead?
@@
 We carried out most of our analyses on data ranks, $u_i$ and $v_i$, for good reasons 
-int he Inrtoduction, Background, and methods sections and 
+mentioned in the Inrtoduction, Background, and methods sections and 
 reviewed here; but under some circumstances it may also be appropriate to use 
@@
 Rank-based measures such as Spearman or Kendall correlation, or any of our copula approaches,
-will be unaffected by nonlinear functions such as $f$. The same difficulty will pertain in any
+will be unaffected by nonlinear monotonic functions such as $f$. The same difficulty will pertain in any
 case for which a measurement of an ecological quantity is a nonlinear index of the true quantity
@@
 
-<!--6) In the Discussion, you maybe need to point out not only that we
-use only bivariate copulas (and there are multivariate ones), but also
-that we use only continuous marginals. Probably an adequate approximation
-for our data, but see the other papers I've seen recently which use count data,
-and discrete marginals. Approaches such as that will be necessary for some 
-datasets.-->
+<!--additional copula methods-->
 Copula methods are numerous, and go well beyond the cases we have considered
 in this introductory paper; we hope our work helps inspire applications of copulas
-in ecology beyond the specific tools we have employed. For instance, multivariate copulas are useful for 
+in ecology both of and 
+beyond the specific tools we have employed. For instance, multivariate copulas are useful for 
 studying relationships between multiple interacting variables, and have been used in finance
@@
 such as total primary productivity or carbon flux are the sum of the contributions of
-multiple species in the same way that athe value of an investment portfolio is the sum 
+multiple species in the same way that the value of an investment portfolio is the sum 
 of the values of its constituent assets. Ecologists may likewise be interested in
@@
 distributions have continuous, strictly 
-monotonic distribution functions; this was for simplicity and because such a simpler
+monotonic distribution functions; this was for simplicity and because the simpler
 approach was sufficient for our research questions. But count data in ecology have been analyzed

### 2019-10-11 Paper.Rmd
working on stuff

@@
 datasets.
-7) change the box around the Cedar Creek data in the main fig
+DONE 7) change the box around the Cedar Creek data in the main fig
 DONE 8) say in Discussion that you could have used non-ranks for the nonparam stats - this
@@
 dependence, and that greater use of copulas can illuminate the underlying processes.
+We believe copula approaches are among the tools all ecologists, including tropical
+ecologists, should be considering for analysis of their data in the 21st century.
 
@@
 in the right-most box in the middle row.
-<!--***DAN: you need to change the box around the Cedar Creek data-->
 
@@
 (Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
-helping to answering the second and third parts of Q1. Though again the model
+helping to answer the second and third parts of Q1. Though again the model
 selection results do not provide 
@@
 datasets, in the middle row of boxes.
-<!--***DAN: change the box for Cedar Creek-->
-We did the same analysis 
+We did the same analyses 
 with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
@@
 were similar to 2000. 
+<!--***DAN: cehcking the below with Shya-->
+We did the same analyses with bird or mammal body mass versus BMR datasets filtered
+to include only tropical species. Answers to Q1-Q3 were substantially the same.
 
@@
 
+<!--***DAN: still working on this-->
+Approaches used and insights gained via our work should apply equally well to 
+tropical and temperate ecosystem; we believe copulas are among the tools 
+that all ecologists should consider for analysis of their data in the 21st century.
+There is certainly no reason to expect that non-normal copula structure and 
+asymmetric tail dependence are special properties of datasets from temperate 
+regions; indeed both of these features were found 
+in mammal and bird body mass versus BMR data filtered to tropical
+species. The Moran-effect and Liebig's law potential mechanisms that we explored
+for copula structure and tail associations (Fig. \ref{MasterFigure}) also seem as 
+likely to hold in tropical as in temperate systems; these are general mechanisms.
+Indeed, Moran effects are widespread, and Liebig's law is a dominant paradigm,
+certainly operating in a wide variety of tropical and temperate ecosystems.
+
 <!--Stuff about the mechanisms-->

### 2019-10-11 Paper.Rmd
added something on tropical

@@
 3) send it out to coauthors (done) and maybe get feedback, ESP FROM lAWRENCE
-4)make sure to say something about tropical systems
+DONE 4)make sure to say something about tropical systems
 DONE 5) maybe write to the editors saying this is coming soon?
@@
 were similar to 2000.
-<!--***DAN: cehcking the below with Shya-->
-We did the same analyses with bird or mammal body mass versus BMR datasets filtered
-to include only tropical species. Answers to Q1-Q3 were substantially the same.
 
@@
 
-<!--***DAN: still working on this-->
-Approaches used and insights gained via our work should apply equally well to 
-tropical and temperate ecosystem; we believe copulas are among the tools 
-that all ecologists should consider for analysis of their data in the 21st century.
-There is certainly no reason to expect that non-normal copula structure and 
-asymmetric tail dependence are special properties of datasets from temperate 
-regions; indeed both of these features were found 
-in mammal and bird body mass versus BMR data filtered to tropical
-species. The Moran-effect and Liebig's law potential mechanisms that we explored
-for copula structure and tail associations (Fig. \ref{MasterFigure}) also seem as 
-likely to hold in tropical as in temperate systems; these are general mechanisms.
-Indeed, Moran effects are widespread, and Liebig's law is a dominant paradigm,
-certainly operating in a wide variety of tropical and temperate ecosystems.
+<!--tropical ecosystems-->
+The approaches we have demonstrated should apply equally well to 
+data from tropical or temperate ecosystems.
+There is no reason to expect that non-normal copula structure and 
+asymmetric tail dependence should be special properties of datasets from temperate 
+regions. The mechanisms we have proposed of non-normal copula structure and asymmetric tail 
+dependence seem equally likely to apply anywhere. For instance, the Moran effect, which underpins 
+one of our proposed mechanisms (Fig. \ref{MasterFigure}B), is a standard mechanism
+that occurs whenever environmental variables influence populations. And Liebig's
+law and nonlinear environmental influences on ecosystems, which underpin 
+another one of our proposed mechanisms (Fig. \ref{MasterFigureA}), are widely
+demonstrated paradigms that apply to all biomes.
 

### 2019-10-13 Paper.Rmd
added a reference

@@
 We carried out most of our analyses on data ranks, $u_i$ and $v_i$, for good reasons 
-mentioned in the Inrtoduction, Background, and methods sections and 
+mentioned in the Introduction, Background, and methods sections and 
 reviewed here; but under some circumstances it may also be appropriate to use 
@@
 the maginal distributions of variables with information about the dependence
-between variables. To provide an ecological example, suppose the true population densities $p_{1,t}$
+between variables. For instance, suppose the true population densities $p_{1,t}$
 and $p_{2,t}$ of two species of fish in a lake in year $t$ are unknown,
 but are assessed using catch per unit effort (CPUE), a standard approach in fisheries
-science. The CPUE measurements, which we denote $c_{1,t}$ and $c_{2,t}$ may be more
+science [@Zale2013]. The CPUE measurements, which we denote $c_{1,t}$ and $c_{2,t}$ may be more
 or less correlated through time than the true densities would be, if they were known,
@@
 under some circumstances, in part because ecological mechanisms which produce relationships 
-between variables are influenced not only by the ordinal informatio in ranks but also 
+between variables are influenced not only by the ordinal information in ranks but also 
 by the cardinal information in the unranked data. Judgements of what statistics
@@
 tail associations that would conflate dependence and marginal information but 
-that may also be useful for understanding ecological relationships.
+that may also be useful for understanding some ecological relationships.
 

### 2019-10-15 Paper.Rmd
implemented Lawrence's suggestions

@@
 $v_i$ are called *normalized ranks* of the $x_i$ and $y_i$; they relate to 
-the empirical cummulative distribution function (cdf) of the $x_i$ and $y_i$,
-respectively. See, for instance,
+the empirical cumulative distribution function of the $x_i$ and $y_i$,
+respectively. (We henceforth abbreviate "cumulative distribution function" as "distribution
+function" or "cdf".) See, for instance,
 Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
@@
 $\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
-less direct intuition than do example pdfs (Fig. \ref{figped_C}). The survival
+less direct intuition than do example pdfs (Fig. \ref{figped_C}). Higher values of the parameter,
+$p$ produce copulas from the Clayton family with higher Kendall or Spearman correlation. The survival
 Clayton family is the symmetric opposite of the Clayton family, showing
@@
 using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
-is Kendall's tau for the data [@Genest2007]. We used the 
+is Kendall's tau for the data. @Genest2007 argue that this statistic is approximately standard normally
+distributed. We used the 
 implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
@@
 Our alternative nonparametric
-approach, results detailed next, provides informaton about significance of tail
+approach, results detailed next, provides information about significance of tail
 associations and asymmetries of tail associations.
@@
 comparisons (Table \ref{tab_soilCN_asym}). Results showed that
-upper-tail association was significantly stronger than lower-tail association. These results
+upper-tail association was significantly stronger than lower-tail association; i.e., C and N values are
+more related when high than when low. These results
 are consistent with the model selection results, but go beyond them by providing information
@@
 $`r 1-pbinom(tab_multivar_summary$Green_spruce_aphid_count$num.non_N/2-1,45,0.5)`$. 
+The average AIC difference `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC` was
+much less than the AIC difference between `r round(min(BivMS_res_Loecke$InfCritRes$AIC),2)` and 
+`r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)` that was obtained for the soil C and N data
+at least in part because the soil C and N data were much more numerous (Table \ref{tab_data_info}).
 These results help answer the first part of Q1: non-normal
@@
 very simple AR(1) formulation, influenced by a "regional" environmental factor which
-is correlated across locatons ($\epsilon$) and by an entirely local factor ($\delta$). 
+is correlated across locations ($\epsilon$) and by an entirely local factor ($\delta$). 
 The function $g$ describes how the environmental factor $\epsilon_i(t)$ influences
@@
 flux of a biogeochemical variable such as methane, or other quantities.
-The nonlinear population model was $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$ for $i=1,2$, using $r=0.5$, $K=100$, $\epsilon_i(t)$
+The nonlinear population model was a stochastic, multi-habitat-patch version of the Ricer model, 
+$P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$ for $i=1,2$, using $r=0.5$, 
+$K=100$, $\epsilon_i(t)$
 as above, and $\sigma=0.1$ or 
@@
 linearity and homoskedasticity, by non-normal copula structure. For normal copula structure 
-(i.e., for simulations that gave normal-copula dependence between populatons
+(i.e., for simulations that gave normal-copula dependence between populations
 in different locations), $p$-values for the linearty
@@
 common across multiple sub-disciplines
-in ecology, although these facets of ecological data are only ocassionally 
+in ecology, although these facets of ecological data are only occasionally 
 taken into account [@Valpine2014; @Anderson2018; @Popovic2018]. We hypothesized 
@@
 we discuss below how commonly some of our
-mechanisms may opperate. 
+mechanisms may operate. 
 We also hypothesized and demonstrated important consequences of 
@@
 by the cardinal information in the unranked data. Judgements of what statistics
-are appropriate depend, of course, on the purpose of the analysis; the goal of this 
+are appropriate depend, of course, on the purpose of the analysis. The goal of this 
 study was to study dependence in isolation, so we used ranks. If one is interested
-in tail associations using cardinal data, a partial Pearson correlation can developed
-straightforwardly by replacing $u$ and $v$ in equation \ref{eq:partialspearman}
-by the $x$ and $y$, but still computing the sum in that equation over data points
+in tail associations using cardinal data, a partial Pearson correlation can be developed
+straightforwardly, by replacing $u$ and $v$ in equation \ref{eq:partialspearman}
+by $x$ and $y$, but still computing the sum in that equation over data points
 with ranks, $u_i$ and $v_i$ constrained by the bounds $u_i+v_i>2l_b$ and 
@@
 tail associations that would conflate dependence and marginal information but 
-that may also be useful for understanding some ecological relationships.
+that may also have utility for some applications.
 
@@
 distributions have continuous, strictly 
-monotonic distribution functions; this was for simplicity and because the simpler
+monotonic distribution functions. Yhis was for simplicity and because the simpler
 approach was sufficient for our research questions. But count data in ecology have been analyzed

### 2019-10-15 FirstPedagogFig.R
fig edited

@@ -48,6 +48,7 @@ pchval<-20
 cexvalpts<-0.25
 cexvaltxt<-0.75
 cexvalpl<-1.5
+cexaxs<-0.9
 pdf(file="./Results/PedagogFig1.pdf",width=tot_wd,height=tot_ht)
 
 #upper-left panel, normal marginals, left-tail dependence
@@ -57,7 +58,7 @@ par(fig=c((ymarg_wd+numsp)/tot_wd,
           (xmarg_ht+3*pan_ht+2*gap+3*numsp)/tot_ht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
 rg<-max(abs(dcc5_nm))
-plot(dcc5_nm[,1],dcc5_nm[,2],type='p',pch=pchval,cex=cexvalpts,xlim=c(-rg,rg),ylim=c(-rg,rg),col="grey")
+plot(dcc5_nm[,1],dcc5_nm[,2],type='p',pch=pchval,cex=cexvalpts,xlim=c(-rg,rg),ylim=c(-rg,rg),col="grey",cex.axis=cexaxs)
 mtext(expression(z[v]),2,line=1.15)
 text(-rg,rg,"A",cex=cexvalpl,adj=c(0,1))
 text(rg,-rg+.25*rg,paste0("P=",round(cor(dcc5_nm[,1],dcc5_nm[,2]),2)),
@@ -72,7 +73,7 @@ par(fig=c((ymarg_wd+2*numsp+pan_wd+gap)/tot_wd,
           (xmarg_ht+3*pan_ht+2*gap+3*numsp)/tot_ht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
 rg<-max(abs(dsc5_nm))
-plot(dsc5_nm[,1],dsc5_nm[,2],type='p',pch=pchval,cex=cexvalpts,xlim=c(-rg,rg),ylim=c(-rg,rg),col="grey")
+plot(dsc5_nm[,1],dsc5_nm[,2],type='p',pch=pchval,cex=cexvalpts,xlim=c(-rg,rg),ylim=c(-rg,rg),col="grey",cex.axis=cexaxs)
 text(-rg,rg,"B",cex=cexvalpl,adj=c(0,1))
[…]

### 2019-10-15 Paper.Rmd
fig edited
changes to acknowledgments

@@
 
-We thank Joel E. Cohen for making us aware of copulas and their potential utility in the 
+We thank Joel E. Cohen for making us aware of copulas and their potential ecological utility in the 
 first place. We thank the many contributors to the large datasets 
@@
 extraction; and David Tilman, Lauren Hallet, Jonathan Walter, Thomas 
-Anderson, Lei Zhao, and editors and annonymous reviewers for helpful suggestions. 
+Anderson, Lei Zhao, Andrew Rypel, and editors and annonymous reviewers for helpful suggestions. 
 We thank James Bell of the Rothamsted Insect Survey (RIS). 

### 2019-10-16 Paper.Rmd
edits to Background section, for clarity

@@
 2) Go through the whole thing one more time
-3) send it out to coauthors (done) and maybe get feedback, ESP FROM lAWRENCE
+3) send it out to coauthors (done) and maybe get feedback
+DONE 3.5) ESP FROM lAWRENCE
 DONE 4)make sure to say something about tropical systems
@@
 9) maybe go back and read the Ecol Monogr reviews again
+10) try to emphasize early on that the focus is on positively associated variables, so 'lower tail' means the lower tail of both variables and it is not necessary to worry about negative relationships (in practice we would always transform these for ease).
+The dfinitions of tail dependence we use cannot usefully be applied to untransformed data in fig 1e,f.
 -->
@@
 ecological phenomena, including impacts 
-on extinction risk, Taylor's law, the stability through time of ecosystem services. By documenting the 
+on extinction risk, Taylor's law, and the stability through time of ecosystem services. By documenting the 
 importance of a complete description of dependence between variables, advancing conceptual frameworks, and 
@@
 \caption{(A-D) Bivariate datasets showing diverse relationships between 
-variables; some of these datasets nevertheless have the same Pearson (P)
+variables. Some of these datasets nevertheless have the same Pearson (P)
 or Spearman (S) correlation coefficients. C and D show normalized ranks of A, B, 
@@
 The approaches are based on 
-the concept of the *copula*; copula approaches separate the information of a bivariate 
+the concept of the *copula*. Copula approaches separate the information of a bivariate 
 random variable into two distinct parts: the information in the 
@@
 the empirical cumulative distribution function of the $x_i$ and $y_i$,
-respectively. (We henceforth abbreviate "cumulative distribution function" as "distribution
-function" or "cdf".) See, for instance,
+respectively. We henceforth abbreviate "cumulative distribution function" as "distribution
+function" or "cdf". See, for instance,
 Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
@@
 Spearman and Kendall correlations, being rank-based, 
-are not influenced by monotonic transformations of the component variables; 
-they provide information solely on the dependence between those variables.
+are not influenced by monotonic transformations of the component variables. 
+They provide information solely on the dependence between those variables.
 
@@
 methods for Q1 below). Although copulas can be used to model the entire
-complex dependence structure of data, we have found tail associations to be
-a useful aspect of the full structure of dependence, so tail associations 
+complex dependence structure of data, we have found tail association to be
+a useful aspect of the full structure of dependence, so tail association 
 is one of the main foci of this paper.
@@
 only - readers can find abbreviated [@Anderson2018; @Genest2007] 
-and comprehensive introductions elsewhere
-[@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. Instead our goal is to 
+and comprehensive [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]
+introductions elsewhere. Instead our goal is to 
 help determine to what extent ecological understanding may benefit generally from
@@
 ecological and evolutionary processes 
-may commonly generate and transmit complex dependence structures, including tail 
-dependence, and that greater use of copulas can illuminate the underlying processes.
+may commonly generate and transmit complex dependence structures, including asymmetric tail 
+associations, and that greater use of copulas can illuminate the underlying processes.
 We believe copula approaches are among the tools all ecologists, including tropical
@@
 to benefit ecological
-research is that prior work has demonstrated (using copulas) complex 
+research is that prior work has demonstrated, using copulas, complex 
 stucture in the spatial dependence of environmental 
@@
 possible consequences. For instance, if populations of a 
-pest species in different locations are all positively associated and also more strongly
+pest species in different locations are all positively associated and are also more strongly
 related to each other in their right tails, then local outbreaks will tend 
@@
 the marginals of a normal distribution. So Q1 asks whether ecological
-data contain dependency information distinct from the standard or default case. 
+data contain dependency information distinct from a standard or default case. 
 (Q2) If the answer to
@@
 variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000],
-and have been used very effectively, but rarely so far in ecology [@Valpine2014; 
+and have also been used very effectively, but rarely so far, in ecology [@Valpine2014; 
 @Anderson2018; @Popovic2018]. 
@@
 
-We give a brief introduction to copulas, focusing, for simplicity, on bivariate copulas
-and on concepts needed
-for the rest of the paper. See @nelsen2006_copula, @Genest2007, @joe2014_dependence, and
-@MaiScherer2017 for general background, and see also @Anderson2018, part of 
+We give a brief introduction to copulas, focusing, for simplicity, on bivariate 
+copulas and on concepts needed
+for the rest of the paper. We assume familiarity with the basic 
+language of probability distributions and random variables.
+See @nelsen2006_copula, @Genest2007, @joe2014_dependence, and
+@MaiScherer2017 for general background on copulas, 
+and see also @Anderson2018, part of 
 which is a short introduction to 
@@
 distribution function of a bivariate random variable with uniform marginals, and 
-the terminology *copula* is sometimes also applied to the random variable. 
+the terminology *copula* is sometimes also applied to this random vector. 
 A foundational theorem of
@@
 The theorem also says $C$ is unique if $F_X$ and $F_Y$ are continuous.
-Thus $C$ couples the bivariate distribution, $F$, with the two marginal distributions. 
+Thus $C$ couples the bivariate distribution function, $F$, with the 
+distribution functions of the marginals. 
 Finally, Sklar's theorem says that if $D$ is any bivariate copula and 
 $G_A$ and $G_B$ are univariate distribution functions, then 
-$D(G_A(x),G_B(y))$ is a distribution function.
+$D(G_A(x),G_B(y))$ is a valid distribution function 
+of some bivariate random variable.
 The applications of this study fall within the case in which univariate marginal 
 distributions have continuous, strictly 
-monotonic distribution functions, and this case is simpler, so we 
+monotonic distribution functions, and this case is simpler. So we 
 henceforth make such an assumption.
-Then, "the" copula (now guaranteed unique) 
-associated with $(X,Y)$ is the distribution
-function of the random variable $(F_X(X),F_Y(Y))$ 
-(Appendix \ref{SM-ContStrictMonoton}). And 
-$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
-$(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ is the random
-vector with distribution function $D$ and
-$G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
-(Appendix \ref{SM-ContStrictMonoton}). Application of the 
-distribution functions of the component random variables translates
-between the uniform-marginal context, which isolates the 
-dependence information, and the context of the original random variable.
-
-Thus Sklar's theorem implies that
-any random vector can be constructed from a copula and marginal 
-distributions. And conversely,
-any copula and any univariate distribution functions give rise to a random vector. 
+
+Sklar's theorem implies that
+any random vector can be constructed from a unique copula and marginal 
+distribution functions, and, furthermore,
+any copula and any univariate distribution functions give rise to a random vector;
+so Sklar's theorem provides a correspondence between random vectors
+and pairs consisting of a copula and two univariate distribution functions,
+thereby making it possible to study a random vector by studying these two
+constituents.
 Marginals contain no information about the
@@
 all such information - it is a complete description of the dependence
-between variables. Sklar's theorem makes it possible to construct
+between variables.
+<!--***-->
+The univariate distribution functions associated with a random vector 
+$(X,Y)$ are its margins, $F_X$ and $F_Y$, and the 
+associated copula is the copula, $C$, guaranteed by Sklar's theorem
+(and guaranteed unique, thanks to the assumptions of continuity 
+and monotonicity made above). In fact, and making Sklar's correspondence
+more concrete, we show in Appendix \ref{SM-ContStrictMonoton}
+that $C$ is the bivariate distribution
+function of the random variable $(F_X(X),F_Y(Y))$. 
+<!--***-->
+Conversely, given a copula
+$D$ and univariate distribution functions $G_A$ and $G_B$,
+Sklar's theorem guarantees that $D(G_A(x),G_B(y))$ is a 
+distribution function, so the random variable with this 
+distribution function is the random variable corresponding
+to $D$, $G_A$ and $G_B$ under Sklar's correspondence. 
+Again making Sklar's correspondence
+more concrete, we show in Appendix \ref{SM-ContStrictMonoton}
+that this random variable is $(G_A^{-1}(U),G_B^{-1}(V))$, 
+where $G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
+and $(U,V)$ is the random variable with distribution function $D$
+(which has uniform marginals).
+<!--***-->
+If we conflate a copula with the random vector of 
+which the copula is the distribution function, then Sklar's correspondence
+is between random variables, $(U,V)$, with uniform marginals 
+and random variables, $(X,Y)$, with arbitrary continuous, strictly monotonic
+marginals. The correspondence is simply via the application of the
+univariate distribution functions, or their inverses: $U=F_X(X)$,
+$V=F_Y(Y)$, $X=F^{-1}_X(U)$, $Y=F^{-1}_Y(V)$.
+<!--***-->
+Application of the 
+distribution functions of the component random variables translates
+between the uniform-marginal context, $(U,V)$, which isolates the 
+dependence information, and the context of the original random variable,
+$(X,Y)$.
+
+Thus Sklar's theorem makes it possible to construct
 bivariate distributions in two separate steps: by specifying the marginal 
@@
 around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
-associations between the two variables in the left and right tails (see
-below for a more precise defintion of tail associations). 
+associations between the two variables in the left and right tails. 
 The Clayton copula family,
@@
 the cdf. Thus the normalized rank pairs $(u_i,v_i)$ for $i=1,\ldots,n$ can be 
-regarded as approximate samples from the bivariate pdf 
-associated with the copula of $(X,Y)$,
-and the scatterplot of $v_i$ versus $u_i$ can be used to infer aspects of copula structure
-(see section on methods for Q1 below).
+regarded as approximate samples from the random variable $(F_X(X),F_Y(Y))$, 
+which is the random variable associated with the copula of $(X,Y)$.
+The scatterplot of $v_i$ versus $u_i$ can be used to infer aspects of copula 
+structure (see section on methods for Q1 below).
 
@@
 distributions have continuous, strictly 
-monotonic distribution functions. Yhis was for simplicity and because the simpler
+monotonic distribution functions. This was for simplicity and because the simpler
 approach was sufficient for our research questions. But count data in ecology have been analyzed

### 2019-10-16 SupportingInformation.Rmd
edits to Background section, for clarity

@@
 continuous and strictly monotonic. We show the copula associated with
-$(X,Y)$, which exists and is unique by Sklar's theorem, is the distrbution function
+$(X,Y)$, which exists and is unique by Sklar's theorem, is, in fact, 
+the distribution function
 of the random variable $(U,V)=(F_X(X),F_Y(Y))$. Let $C$ denote the distribution

### 2019-10-17 Paper.Rmd
edits to Q2 sections and Master figure

@@
 between these, for each pair of locations, and we characterize
-the distribution of these values across location pairs.
+the distributions of these values across location pairs.
 
@@
 the normalized ranks $u_i$ and $v_i$ are 
-related in any part of their distributions; we here describe the 
+related in any part of their distributions. We here describe the 
 statistics, with some additional details in Appendix \ref{SM-nonparam_stats}.
@@
 using $u_b=1$ and $l_b \geq 0.5$ gives information about association in 
-the right parts. The first statistic, a partial Spearman correlation
+the right parts. The first statistic, a partial Spearman correlation,
 \begin{equation}
@@
 interval $(0,1)$) equals the standard Spearman correlation
-as long as no points fall directly on the bounds.
+as long as no points fall exactly on the bounds.
 We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
@@
 Unlike $\cor_{l_b,u_b}$ and 
-$\Ps_{l_b,u_b}$, for which large values inidicate strong association
+$\Ps_{l_b,u_b}$, for which large values indicate strong association
 between the bounds, small values of 
@@
 sampling variation in our statistics, and we considered 
-multiple bands $(l_b,u_b)$ to understand how dependence
-varies in different parts of the distributions. But for datasets with 
+multiple bands $(l_b,u_b)$ to understand how association
+varies across different parts of the distributions. But for datasets with 
 smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$,
@@
 The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
-widely used spline correlogram method; the same scheme was also used by @walteretal2017 
+widely used spline correlogram method. The same scheme was also used by @walteretal2017 
 as part of a method for 
@@
 non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
-from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent.
+from `BiCopIndTest` in the `VineCopula` package). Non-independence is also visually apparent.
 The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
@@
 (Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
-difference between these fits was not meaningfully large.
+difference between these fits was marginal.
 Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
@@
 with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
-2007, biomass and Shannon's index were independent; for 1997-1999, results 
-were similar to 2000.
+2007, independence of biomass and Shannon's index could not be rejected; 
+for 1997-1999, results were similar to 2000.
 
@@
 indicating greater lower-tail dependence, and again 
-consistent with the results of Table \ref{tab_multivar_summary} (row 9).
+consistent with the results of Table \ref{tab_multivar_summary}, row 8.
 For the methane-flux data, confidence intervals were 
@@
 as well as for the 21 plankton taxa for which we had abundance data other than
-\textit{Ceratium furca} (results not shown); results supported the conclusion that 
+\textit{Ceratium furca} (results not shown). Results supported the conclusion that 
 non-normal copula structure and tail dependence, and asymmetry of tail associations, are 
@@
 Having demonstrated that non-normal copula structure and asymmetric tail associations are 
-common in ecological variables, we 
 addressed Q2 by exploring, using models, three possible mechanisms
@@
 Models presented are initial explorations, only, of whether the proposed mechanisms have the 
-potential to explain observed patterns. As such, simple models were used, 
-and comprehensive explorations of
+potential to explain observed patterns. As such, simple models were used. 
+Comprehensive explorations of
 model parameter space and alternative model formulations were left for future work.
@@
 If an environmental variables influences an ecological variable 
-disproportionately strongly in one of its tails, we explored whether the ecological
-variable could then exhibit asymmetric tail dependence across space.
+disproportionately in one of its tails, we explored whether the ecological
+variable could then exhibit asymmetric tail associations across space.
 Let $E_i(t)$ be an ecological variable 
@@
 The values of $b$ selected provide a modicum of exploration of the potential
-for ecological dynamics to influence how complex dependencies between
+for ecological dynamics to also influence how complex dependencies between
 the $E_i(t)$ may arise: 
@@
 flux of a biogeochemical variable such as methane, or other quantities.
-The nonlinear population model was a stochastic, multi-habitat-patch version of the Ricer model, 
-$P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$ for $i=1,2$, using $r=0.5$, 
+The nonlinear population model was a stochastic, multi-habitat-patch version 
+of the Ricker model, 
+$P_i(t+1)=P_i(t)\exp \left[r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right]$ for $i=1,2$, using $r=0.5$, 
 $K=100$, $\epsilon_i(t)$
@@
 each of the five copulas, mammalian character
-evolution was simulated 100 times. For each simulation, symmetry of tail dependence 
+evolution was simulated 100 times. For each simulation, symmetry of tail associations 
 of the two characters across phylogeny tips was assessed using our nonparametric
@@
 indicate that upper-tail (respectively, lower-tail) dependence was significantly stronger than 
-lower-tail dependence compared to a normal-copula null hypothesis. `FS' stands for `first slice'
+lower-tail (respectively, upper-tail) associations compared to a 
+normal-copula null hypothesis. `FS' stands for `first slice'
 and refers to the bounds 0 to 0.1; `LS' stands for `last slice' and refers to the bounds
@@
 as long as environmental noise is small enough that dynamics stay relatively close
-to the model equilibrium. Results that hold for "weak noise" in this sense
+to the model equilibrium. Theoretical results that hold for "weak noise" in this sense
 are common in ecology.
 For strong noise and using our nonlinear model, 
-our correlation and tail-asymmetry statistics, generally 
+our correlation and tail-association-asymmetry statistics, generally 
 speaking, were approximately similar between noise and model
@@
 For our evolutionary model, the hypothesis was 
-correct that asymmetric tail association in evolutionary changes 
+correct, for the models we employed, that asymmetric tail association in evolutionary changes 
 can produce similarly asymmetric tail association between characters across
@@
 Once character evolution was simulated 100 times for each of the
-dependence structure we used on evolutionary changes, 
+dependence structures we considered for evolutionary changes, 
 we had 817 bivariate characters for each of 500 simulations (see methods 
@@
 Thus, based on our simulations, we cannot reject the possibility that
-this is a mechanism behind observed asymmetric tail associations
+this is a proximate mechanism behind observed asymmetric tail associations
 in our bird and mammal data.
@@
 computed for each of 100 simulations of mammalian character evolution for each of five types of tail dependence between 
-evolutionary changes in the characters (extreme left-tail dependence (A, F, K), moderate left-tail dependence (B, G, L), 
-symmetry of tail dependence (C, H, M), moderate right-tail dependence (D, I, N), and extreme right-tail dependence 
-(E, J, O)).
+evolutionary changes in the characters (extreme left-tail dependence, A, F, K; moderate left-tail dependence, B, G, L; 
+symmetry of tail dependence, C, H, M; moderate right-tail dependence, D, I, N; and extreme right-tail dependence, 
+E, J, O).
 See the methods section for Q2 for details.

### 2019-10-18 Paper.Rmd
small edits on final read through, part done

@@
 $x$ and $y$ and jump by $1/n$
-at each of the data points; for large $n$ they approximate the cdfs $F_X$ and $F_Y$. 
+at each of the data points. Therefore $\hat{F}_X(x_j)$ is the rank of $x_j$ in
+the set $\{x_i : i=1,\ldots,n\}$, dividided by $n$, 
+and analogously $\hat{F}_Y(y_j)$ is the rank of $y_j$ in the set 
+$\{y_i : i=1,\dots,n\}$, divided
+by $n$. For large $n$, $\hat{F}_X(x_i)$ and $\hat{F}_Y(y_i)$ 
+approximate the cdfs $F_X$ and $F_Y$, respectively. 
 But $\hat{F}_X(x_i)$ equals the normalized rank of $x_i$ times
@@
 ecological consequences of non-normal copula structure and asymmetric tail
-associations. As for Q2, models used are intended as initial explorations, only,
-of whether copula structure may possibly have a meaningful influence on important ecological
+associations. As for Q2, the models we used are intended as initial explorations, only,
+of whether copula structure may have a meaningful influence on important ecological
 phenomena. Therefore only simple models are used, and comprehensive explorations of the 
@@
 copula structure and tail associations. 
-To refine this hypothesis, suppose an ecological variable 
+To make this hypothesis more precise, suppose an ecological variable 
 $E_i(t)$ is measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and
@@
 spatial-mean time series corresponds to "spikiness" of that time series,
+i.e., occassional very large values, 
 which corresponds to instability through time. 
@@
 up to sampling variation, regardless of X;
-3) the copula for the dependence between $m_{t,j_1}^{(\text{X},k)}$ and $m_{t,j_2}^{(\text{X},k)}$ was X.
+3) the copula for the dependence between $m_{t,j_1}^{(\text{X},k)}$ and $m_{t,j_2}^{(\text{X},k)}$ was  from the family X.
 We used $N=50$ and $n=25$. Details of how these matrices were generated are in 
@@
 were statistically similar except for the copula structure between locations, which was
-X. Thus comparing how Taylor's law may manifest differently for $m^{(\text{X},k)}$
+X. Thus comparing how the $log(v_g)$ versus $\log(m_g)$ relationship may
+manifest differently for $m^{(\text{X},k)}$
 for different values of X constitutes a test of the influence of copula
@@
 
-\noindent Results were consistent with the hypothesis that the skewness
+\noindent Empirical results were consistent with the hypothesis that the skewness
 of a spatial-average time series is influenced by tail associations
@@
 Also consistent with hypothesis, copula structure had a substantial effect on
-Taylor's law. 
+Taylor's law for the models we considered. 
 Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
@@
 constituent time series: predominantly right-tail-associated local time series 
-can lead to "spiky" spatially averaged time series, with large outbreaks.
-Thus tail associations could have implications for exploited and pest species.
-Extinction risk and Taylor's law can also be altered by tail association patterns across space
-in environmental variables. 
+can lead to "spiky" spatially averaged time series, with large outbreaks; and 
+predominantly left-tail-associated local time series can lead to spatially
+averaged time series showing accentuated "crashes".
+Thus tail associations could have implications for pest and exploited species.
+Extinction risk and Taylor's law can also be altered by tail association 
+patterns across space. 
 In our view, our results make it reasonable to suggest 
@@
 of many ecological phenomena may be possible if a
-complete copula characterization of dependence is employed. 
+complete copula characterization of associations between variables is employed. 
 Copula approaches are
-well developed [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], statistically, 
+well developed, statistically [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], 
 and have been introduced in accessible formats [@Anderson2018; @Genest2007]. Furthermore, 
@@
 There is no reason to expect that non-normal copula structure and 
-asymmetric tail dependence should be special properties of datasets from temperate 
+asymmetric tail associations should be special properties of datasets from temperate 
 regions. The mechanisms we have proposed of non-normal copula structure and asymmetric tail 
-dependence seem equally likely to apply anywhere. For instance, the Moran effect, which underpins 
-one of our proposed mechanisms (Fig. \ref{MasterFigure}B), is a standard mechanism
+associations seem equally likely to apply anywhere. 
+For instance, the Moran effect, which underpins 
+one of our proposed mechanisms (Fig. \ref{MasterFigure}, B), is a standard mechanism
 that occurs whenever environmental variables influence populations. And Liebig's
 law and nonlinear environmental influences on ecosystems, which underpin 
-another one of our proposed mechanisms (Fig. \ref{MasterFigureA}), are widely
-demonstrated paradigms that apply to all biomes.
+another one of our proposed mechanisms (Fig. \ref{MasterFigure}, A), are widely
+demonstrated phenomena that apply in all biomes.
 
@@
 our Moran mechanism (Fig. \ref{MasterFigure}, B) 
-may operate commonly: Moran effects are common [@Liebhold2004], and 
+may operate commonly: Moran effects are common [@Liebhold2004; @Sheppard2016; 
+@Defriez2017_ocean; @Defriez2017_land], and 
 non-normal copula structures and asymmetric tail associations are often found in 
@@
 environmental variables associated with these events should take extreme
-values simultaneously across large spatial areas, producing tail dependence in measurements
+values simultaneously across large spatial areas, producing tail associations in measurements
 made through time at different locations. Non-extreme values may instead
@@
 @Serinaldi2008 examined the spatial dependence 
-of rainfall in Central Italy; Gumbel or Student 2-copulas 
+of rainfall in Central Italy. Gumbel or Student 2-copulas 
 were candidates for modelling dependence, and 
@@
 showed that a Gumbel copula effectively modeled 
-the spatial dependence of drought variables; the Gumbel copula demonstrates 
+the spatial dependence of drought variables. The Gumbel copula demonstrates 
 asymmetric tail associations.
@@
 over the North Sikkim Himalayas using spatial datasets [@Goswami2018].
+
+<!--***DAN: made it to here 2019 10 18-->
 

### 2019-10-19 Paper.Rmd
finshed reading it through and implemented Shya comments

@@
 
-\noindent \emph{The number of words in the abstract:} 266
+\noindent \emph{The number of words in the abstract:} 261
 
@@
 DONE 1) Edit abstract
-2) Go through the whole thing one more time
+DONE 2) Go through the whole thing one more time
 3) send it out to coauthors (done) and maybe get feedback
 DONE 3.5) ESP FROM lAWRENCE
+DONE 3.6) Incorporate Shya feedback
 DONE 4)make sure to say something about tropical systems
@@
 9) maybe go back and read the Ecol Monogr reviews again
-10) try to emphasize early on that the focus is on positively associated variables, so 'lower tail' means the lower tail of both variables and it is not necessary to worry about negative relationships (in practice we would always transform these for ease).
+DONE 10) try to emphasize early on that the focus is on positively associated variables, so 'lower tail' means the lower tail of both variables and it is not necessary to worry about negative relationships (in practice we would always transform these for ease).
 The dfinitions of tail dependence we use cannot usefully be applied to untransformed data in fig 1e,f.
@@
 species-trait, phenological, population, community, and ecosystem functioning datasets we 
-considered. Many datasets exhibited asymmetric tail associations, whereby two variables are more 
+considered. Many datasets exhibited asymmetric tail associations, whereby two variables were more 
 strongly related in their left compared to right tails, or *vice versa*. We describe mechanisms by which 
@@
 Alternatively, Fig. \ref{fig_cop_pedag1}F may suggest that Sp2
-is the dominant competitor. Although other causal hypotheses may produce 
+is the dominant competitor. Other causal hypotheses may produce 
 similar dependence structure, and it is usually impossible to obtain
 complete information on causal pathways from analyses which are
-fundamentally based on associations, more detailed 
+fundamentally based on associations. However, more detailed 
 examination of the patterns of dependence of variables may
@@
 Gaussian/normal
-distribution (here called non-normal copula structure)? Do they show tail 
+distribution (here called non-normal copula structure)? Do positvely associated 
+ecological variables show tail 
 associations distinct from those of a normal distribution, and in particular
@@
 corresponds to the correlation
-of the related bivariate normal distribution, and controls the degree of dependence
+of the related bivariate normal distribution, and controls the degree of association
 between the variables. A normal copula with $p=0.7$ was already introduced (Fig.
@@
 which shows upper-tail dependence is defined symmetrically in the two random variables.
-Lower tail dependence is defined analogously, as 
+All the variables we consider were positively associated when they were significantly
+associated, so we think of positively associated variables when conceptualizing the
+definitions presented here. Lower tail dependence is defined analogously, as 
 $\lambda_L = \lim_{u \rightarrow 0^+} P[Y \leq F_Y^{-1}(u) | X \leq F_X^{-1}(u)]$.
@@
 described formally elsewhere [@Joe1997]; so 
-we only briefly identify each family, say a few words about tail associations 
+we only briefly identify each family, say a few words about tail dependence 
 of the families, and provide pictorial descriptions in the Appendices.
@@
 $\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, $\Dsq_u=\Dsq_{0.5,1}$.
-To test for asymmetry of dependence in upper and lower portions of distributions,
+To test for asymmetry of association in upper and lower portions of distributions,
 we used differences $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$
@@
 So means $m_g$ and variances $v_g$ are computed across space. We refer to the matrix
-with $x_i(t)$ in the $t$th row and the $i$th column as a *population matrix*. 
+with $x_i(t)$ in the $t^{\text{th}}$ row and the $i^{\text{th}}$ 
+column as a *population matrix*. 
 We consider that 
@@
 
-<!--***DAN: made it to here 2019 10 18-->
-
 <!--Causal mechanism C, which we also did not test-->
@@
 It could be tested by analyzing copulas of abundances of competing 
-species, sampled across space or time.
+species, sampled across space or time. We note that, whereas all the datasets
+we studied here have been positively associated when they were significantly
+associated, for negatively associated variables such as abundances of competing species,
+the definitions of left- and
+right-tail association no longer apply, strictly speaking: the left tail
+of one distribution corresponds to the right tail of the other. One must
+be careful with terminology, but it is still possible to study asymmetries of
+association.
 
@@
 
-Performing statistical tests for associations between continuous traits in different species
+Performing statistical tests for associations between continuous traits across different species
     was a primary motivating example of phylogenetic comparative methods.
@@
 to any combination of the $G_i$. This constitutes 
-asymmetric tail dependence between F and $G_1$. @Winemiller1992
+asymmetric tail association between F and $G_1$. @Winemiller1992
 described a three-way trade-off in fishes between age of reproductive
@@
 to studying multi-dimensional trade-offs. We used only bivariate 
-analyses in this study solely in order to have at least one
-way to constrain the complexity of what is already a complex piece of work.
-But the statistical theory on multi-dimensional copulas is well developed
+analyses in this study solely for simplicity.
+But statistical theory on multi-dimensional copulas is also well developed
 [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], 
 and open-source computer implementations exist for multivariate as well as 
-bivariate copulas (e.g., the `VineCopula`
-package for the R programming language). Such approaches may be an 
+bivariate copula methods (e.g., the `VineCopula`
+package for the R programming language). Such approaches may be a 
 useful next step in life-history theory and other areas.
@@
 measurement error may modify copula structure. 
-Our models were intentionally simple in other respects, 
+Our models investigating potential causes of copula structure 
+were intentionally simple in other respects, 
 too, and did not include factors such as
@@
 interactions, etc.; and we did not comprehensively 
-explore parameter space. We re-emphasize that fuller explorations in future work
-of some of our models, and variant model, are likely to be informative. 
+explore parameter space for our models. We re-emphasize 
+that fuller explorations, in future work,
+of some of our models and of variant model may be informative. 
 We hope by enumerating a few potential mechanisms of copula structure
@@
 <!--a para on stability of the total biomass time series in a community, and skewness - sets up Shya's later paper as future work-->
-We also elaborated potential consequences of copula structure for ecological understanding. 
+We also elaborated potential consequences of copula structure for ecological 
+phenomena and understanding. 
 Our results showed that the skewness of the 
@@
 Green spruce aphid abundance, as measured in the data we use,
-is strongly influenced by the temperature of the previous winter (@Sheppard2016,
+is strongly positively associated with the temperature of the previous winter (@Sheppard2016,
 their supplementary figure 6).
@@
 March of year $t$, was available for the locations of aphid sampling, 
-and was processed in the same way as @Sheppard2016.
+and was preprocessed in the same way as @Sheppard2016.
 For each of our 10 sampling locations, we therefore 
@@
 to have lower skewness (i.e., less spikiness, and greater stability through time) 
-than they would otherwise.
+than they would otherwise. It seems likely that asymmetric influence of winter 
+termperature on populations may be a common phenomenon, so effects such as 
+we have documented for green spruce aphid may be common.
 
@@
 recommend carrying out inferences about dependence structures (which was our
-goal) using normalized ranks, stating explicitly "statistical inference concerning
+goal) using normalized ranks, stating explicitly that "statistical inference concerning
 dependence structures should always be based on ranks." The reason for this is
 that measures of dependence which use unranked data conflate information about
-the maginal distributions of variables with information about the dependence
+the marginal distributions of variables with information about the dependence
 between variables. For instance, suppose the true population densities $p_{1,t}$
@@
 but are assessed using catch per unit effort (CPUE), a standard approach in fisheries
-science [@Zale2013]. The CPUE measurements, which we denote $c_{1,t}$ and $c_{2,t}$ may be more
-or less correlated through time than the true densities would be, if they were known,
+science [@Zale2013]. The CPUE measurements, which we denote $c_{1,t}$ and $c_{2,t}$ may be 
+differently correlated through time than the true densities would be, if they were known,
 if the function $f$ relating $p_{i,t}$ and $c_{i,t}$ is nonlinear and if Pearson
@@
 under some circumstances, in part because ecological mechanisms which produce relationships 
-between variables are influenced not only by the ordinal information in ranks but also 
-by the cardinal information in the unranked data. Judgements of what statistics
+between variables are influenced not only by the ordering information in ranks but also 
+by the relative-size information in the unranked data. Judgements of what statistics
 are appropriate depend, of course, on the purpose of the analysis. The goal of this 
 study was to study dependence in isolation, so we used ranks. If one is interested
-in tail associations using cardinal data, a partial Pearson correlation can be developed
+in tail associations using unranked data, a partial Pearson correlation can be developed
 straightforwardly, by replacing $u$ and $v$ in equation \ref{eq:partialspearman}
 by $x$ and $y$, but still computing the sum in that equation over data points
-with ranks, $u_i$ and $v_i$ constrained by the bounds $u_i+v_i>2l_b$ and 
+with ranks $u_i$ and $v_i$ constrained by the bounds $u_i+v_i>2l_b$ and 
 $u_i+v_i<2u_b$. This would result in a different approach to
 tail associations that would conflate dependence and marginal information but 
-that may also have utility for some applications.
+that may also have its own utility for some applications.
 
@@
 Copula methods are numerous, and go well beyond the cases we have considered
-in this introductory paper; we hope our work helps inspire applications of copulas
+in this introductory paper. We hope our work helps inspire applications of copulas
 in ecology both of and 

### 2019-10-22 Paper.Rmd
ready to submit, I think

@@
 title: "Copulas and their potential for ecology"
-author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
+author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance D. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
 date: ""
@@
   pdf_document:
-    number_sections: no
+    number_sections: yes
     keep_tex: yes
@@
 
-\noindent \emph{The number of words in the abstract:} 261
+\noindent \emph{The number of words in the abstract:} 263
 
@@
 DONE 2) Go through the whole thing one more time
-3) send it out to coauthors (done) and maybe get feedback
+DONE 3) send it out to coauthors (done) and maybe get feedback
 DONE 3.5) ESP FROM lAWRENCE
 DONE 3.6) Incorporate Shya feedback
+DONE 3.7) Incorporate Reid feedback
 DONE 4)make sure to say something about tropical systems
@@
 is the stuff the Ecol Mongr referee talked about. Maybe this should go in Background instead?
-9) maybe go back and read the Ecol Monogr reviews again
+DONE 9) maybe go back and read the Ecol Monogr reviews again
 DONE 10) try to emphasize early on that the focus is on positively associated variables, so 'lower tail' means the lower tail of both variables and it is not necessary to worry about negative relationships (in practice we would always transform these for ease).
 The dfinitions of tail dependence we use cannot usefully be applied to untransformed data in fig 1e,f.
+DONE 11) Number the sections and subsections in the main text and change all the section 
+references to use those numbers. Search for "section" (DONE) also for "methods" (DONE)
 -->
 
-# Abstract
-All branches of ecology study relationships among environmental and biological variables. However, 
+\noindent \textbf{Abstract}
+
+\noindent All branches of ecology study relationships among and between environmental and biological variables. However, 
 standard approaches to studying such relationships, based on correlation and regression, 
@@
 All branches of ecology 
-study relationships among environmental and biological 
+study relationships among biological variables and relationships between 
+environmental and biological 
 variables. However, commonly used correlation and regression 
@@
 flavor of copula ideas to be introduced 
-more formally in the section on Background below. Given 
+more formally in the "Background on copulas" section below (section \ref{Background}). Given 
 a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
@@
 separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D
-relate to copulas in a way to be described in the section on Background below, 
-after a definition of copulas
+relate to copulas in a way to be described in the "Background on copulas" section below
+(section \ref{Background}), after a definition of copulas
 is provided. Note that Pearson correlation, even though it is the most commonly
@@
 that it can detect associations in the tails of distributions, and asymmetries of such associations. 
-Tail association (introduced formally in the section on Background, below) 
+Tail association (introduced formally in the "Background on copulas" section, below, section 
+\ref{Background}) 
 is association between extreme values of variables. 
@@
 ones (Fig. \ref{fig_cop_pedag1}D). Datasets of the same 
-Spearman or Kendall correlation can have a range of tail association 
+Spearman or Kendall correlation can have a range of tail associations 
 (Fig. \ref{SM-fig_cop_pedag}), which can be quantified (see sections on Background and
-methods for Q1 below). Although copulas can be used to model the entire
-complex dependence structure of data, we have found tail association to be
-a useful aspect of the full structure of dependence, so tail association 
+methods for Q1 below, sections \ref{Background} and \ref{Methods}). 
+Although copulas can be used to model the entire
+complex dependence structure of data, we have found tail associations to
+usefully describe the full structure of dependence, so tail association 
 is one of the main foci of this paper.
@@
 The goal of this paper is to explore the potential for 
-applications of copulas in ecology. We provide a basic introduction
-to copula concepts in the Background section below, but this is a sketch
-only - readers can find abbreviated [@Anderson2018; @Genest2007] 
+applications of copulas in ecology and to 
+help determine to what extent ecological understanding may benefit generally from
+using copulas. We provide a basic introduction
+to copula concepts in the "Background on copulas" section below (section \ref{Background}). 
+Readers can find additional abbreviated [@Anderson2018; @Genest2007] 
 and comprehensive [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]
-introductions elsewhere. Instead our goal is to 
-help determine to what extent ecological understanding may benefit generally from
-using copulas. Copulas have already been used very effectively 
+introductions elsewhere. Copulas have already been used very effectively 
 in a few studies to illuminate ecological phenomena 
@@
 \caption{Summary and guide for analyses presented in the text. Middle boxes
-correspond to ecological datasets for which Q1 was examined (in the 
-third through
-fifth sections of the text, see also Table \ref{tab_data_info}). Upper boxes 
+correspond to ecological datasets for which Q1 was examined (in sections
+\ref{Data}-\ref{Results}, see also Table \ref{tab_data_info}). Upper boxes 
 correspond to potential causes of non-normal (non-N) copula structure that
-were examined (in the sixth and seventh sections of the text). Lower boxes 
+were examined (in sections \ref{Causes}-\ref{Results_Q2} of the text). Lower boxes 
 correspond to potential consequences we examined of non-normal copula structure for
-ecological understanding and for applications (in the eighth and ninth
-sections of the text). Arrow labels (A-D for analyses pertaining 
+ecological understanding and for applications (in sections 
+\ref{Consequences}-\ref{Results_Q3} of the text). Arrow labels (A-D for analyses pertaining 
 to causes, Q2; and X-Z for analyses pertaining to consequences, Q3) 
@@
 
-A second reason we may expect understanding of complex dependence structures
+A second reason why we may expect understanding of complex dependence structures
 to benefit ecological
@@
 
-A third reason we may expect an understanding of complex 
+A third reason why we may expect an understanding of complex 
 dependence structures to benefit 
@@
 
-A fourth reason we may expect a copula understanding of complex 
+A fourth reason why we may expect a copula understanding of complex 
 dependence structures to be useful
@@
 higher organizational levels in ecology and human concerns.
-Illustrating the idea here, prior work demonstrated that the overall 
+To illustrate this idea we cite @li2013, who have demonstrated that the overall 
 reliability of wind-generated electricity depends sensitively on details
-of the dependence between wind speeds at multiple generator sites [@li2013]. 
+of the dependence between wind speeds at multiple generator sites. 
 Spatially aggregated ecological variables may be subject to a similar 
@@
 Q1 is "yes," then what are some possible causes/mechanisms of non-normal
-copula structure and asymmetric tail associations in ecology? And (Q3) what are the 
+copula structure and asymmetric tail associations in ecology? And (Q3), what are the 
 consequences of non-normal copula
 structure and asymmetric tail associations for ecological understanding and
-applications? After a Background section (the second section), we address Q1 
-(third through fifth sections of the paper) 
+applications? After section \ref{Background}, 'Background on copulas', we address Q1 
+(sections \ref{Data} through \ref{Results} of the paper) 
 by analyzing several datasets, including 
@@
 sub-fields and organizational levels of ecology. 
-We address Q2 (sixth and seventh sections of the paper), using 
-simple models. We address Q3 (eighth and ninth sections of the paper) using 
-both data and models. The final section is the Discussion.
+We address Q2 (sections \ref{Causes} through \ref{Results_Q2} of the paper), using 
+simple models. We address Q3 (sections \ref{Consequences} through \ref{Results_Q3} 
+of the paper) using 
+both data and models. Section \ref{Discussion} is the Discussion.
 Multiple analyses are brought to bear on each question, and these
@@
 
-# Background on copulas
+# Background on copulas\label{Background}
 
@@
 distribution: a normal copula is the copula of a bivariate normal distribution.
-See the Background section for more information on normal, Clayton,
+See section \ref{Background} for more information on normal, Clayton,
 and other copulas. Each copula was used with both sets of marginals and each
@@
 copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
-cdf of a (univariate) standard normal and $\Phi_{2,p}$ is the joint cdf of a
+cummulative dstribution function (cdf) of a (univariate) standard 
+normal and $\Phi_{2,p}$ is the joint cdf of a
 bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
 and $p$ in the off diagonal positions. However, formulas for copulas are often not as 
-directly informative about copula properties as pdf diagrams; Fig. \ref{figped_N} has 
+directly informative about copula properties as probability density function (pdf) 
+pictures; Fig. \ref{figped_N} has 
 example pdfs corresponding to bivariate normal copulas for various values of $p$.
@@
 The Clayton copula family,
-of which we also already pictured an example (Fig.
+of which we have already pictured an example (Fig.
 \ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
@@
 and $\lambda_L=0$ and $\lambda_U=2^{-1/p}$ for a survival Clayton  copula with parameter $p>0$.
-In the section on methods for Q1 we will introduce
+In section \ref{Methods} we will introduce
 ways tail dependence can be applied to data, by first fitting copulas to the data. We will also 
@@
 The scatterplot of $v_i$ versus $u_i$ can be used to infer aspects of copula 
-structure (see section on methods for Q1 below).
+structure (see section \ref{Methods} below).
 
@@
 family (it is a 
-one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively (see Background section for definition). The 
+one-parameter family); and LT and UT are the measures of lower- and upper-tail dependence, respectively (for the definitions of these, see the section on Background on copulas, 
+section \ref{Background}). The 
 parameter range for the family is $p \in [-1,1]$, and lower- and upper-tail
@@
 upper-tail dependence, 
-respectively (see Background section for definition). The parameter range for the family is $p \in (0,\infty)$, lower-tail dependence is $2^{-1/p}$ and upper-tail
+respectively (for the definitions of these, see the section on Background on copulas, 
+section \ref{Background}). The parameter range for the family 
+is $p \in (0,\infty)$, lower-tail dependence is $2^{-1/p}$ and upper-tail
 dependence is $0$.\label{figped_C}}
@@
 two-parameter family); and LT and UT are our measures of lower- and upper-tail dependence, 
-respectively (see Background section for definition). The parameter ranges for 
+respectively (for the definitions of these, see the section on Background on copulas, 
+section \ref{Background}). The parameter ranges for 
 the family are $p_1 \in (0,\infty)$ and $p_2 \in [1,\infty)$, 
@@
 
-Datasets included environmental, species-trait, phenological, 
+The datasets we used included environmental, species-trait, phenological, 
 population, community, and ecosystem functioning data (Table \ref{tab_data_info}), selected to span multiple fields and levels
@@
 56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells.
+These data were computed from the Continuous Plankton Recorder survey hosted by the 
+UK Marine Biological Survey.
 Community-level data, obtained from the Cedar Creek Ecosystem Science Reserve, were 
@@
              caption = "Summary table for the data we used.
-Bold entries are the multivariate data, the rest are bivariate datasets (see Data section).
+Bold entries are the multivariate data, the rest are bivariate datasets (see section \ref{Data}).
 Basal metabolic rate=BMR.
@@
 and, additonally, these copulas have been 
-described formally elsewhere [@Joe1997]; so 
+described in more detail elsewhere [@Joe1997]; so 
 we only briefly identify each family, say a few words about tail dependence 
@@
 zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
-and survival BB6 familes. We used several families that can exhibit positive
+and survival BB6 families. We used several families that can exhibit positive
 upper-tail dependence (strength depends on parameters) and zero lower-tail
@@
 within a band can be computed
-for any band (methods for Q1) to describe how the strength of association
+for any band (section \ref{Methods}) to describe how the strength of association
 between $u$ and $v$
@@
 Bottom panels show the corresponding normalized rank plots.
-See Data section for units used in upper panels.
+See section \ref{Data} for units used in upper panels.
 \label{fig_biv_multi_raw_cop}}
@@
 <!--bivariate example, model selection-->
-For the soil C and N data (Data section, Table \ref{tab_data_info}, 
+For the soil C and N data (section \ref{Data}, Table \ref{tab_data_info}, 
 Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
@@
 or $3$ are considered meaningful and differences of $8$ or greater
-are considered diffinitive, so that data provide essentially no support
+are considered definitive, so that data provide essentially no support
 for the higher-AIC model in that case [@burnham2003_modelselection].
@@
 
-We present green spruce aphid abundance (Data section; Table \ref{tab_data_info}) results
+We present green spruce aphid abundance (section \ref{Data}; Table \ref{tab_data_info}) results
 as an example multivariate analysis.
@@
 variables on ecological variables.
-If an environmental variables influences an ecological variable 
+If an environmental variable influences an ecological variable 
 disproportionately in one of its tails, we explored whether the ecological
@@
 and our Spearman- or Kendall-preserving normal-copula 
-surrogate comparison methods (see section on methods for Q1) to these 
+surrogate comparison methods (see section \ref{Methods}) to these 
 outputs to discover if the model could
@@
 Outputs of simulations from the first model detailed in 
-the methods section for Q2 (i.e., Figs \\ref{fig_ms_asym_b_0.1} and 
+section \\ref{Causes} (i.e., Figs \\ref{fig_ms_asym_b_0.1} and 
 \\ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to nonparamtric statistical 
-analyses decribed in the methods section for Q1. The statistic cor$_{FS}$ - cor$_{LS}$ was computed for the
+analyses decribed in section \\ref{Methods}. The statistic cor$_{FS}$ - cor$_{LS}$ was computed for the
 output of the model with the indicated b and g, and the value was compared to 1000 values
 of the same statistic computed on surrogate time series randomized to have normal copula structure
-(see methods for Q1). A table entry $<X$ (respectively, $>X$) indicates the value of the 
+(see section \\ref{Methods}). A table entry $<X$ (respectively, $>X$) indicates the value of the 
 given statistic on the data was less (respectively, more) than its value on $X$ of the 
@@
 were always similar for noise and populations 
-(Fig. \ref{fig_Cause4copula_C_kend_ar1} for the Clayon copula, 
+(Fig. \ref{fig_Cause4copula_C_kend_ar1} for the Clayton copula, 
 Fig. \ref{SM-fig_Cause4copula_SC_kend_ar1} for the survival Clayton copula).
@@
 for this model. The 
-AR(1) model and a Clayton copula were used (see methods section for Q2).
+AR(1) model and a Clayton copula were used (see section \\ref{Causes}).
 Each point is the mean across 50 replicate simulations for which the same 
@@
 to the 1-1 line (green). 
-See methods section for Q2 
-for details.\label{fig_Cause4copula_C_kend_ar1}}
+See section \ref{Causes} for details.\label{fig_Cause4copula_C_kend_ar1}}
 \end{center}
@@
 dependence structures we considered for evolutionary changes, 
-we had 817 bivariate characters for each of 500 simulations (see methods 
-section for Q2). We 
+we had 817 bivariate characters for each of 500 simulations (see section 
+\ref{Causes}). We 
 computed our nonparametric asymmetry statistics 
@@
 E, J, O).
-See the methods section for Q2 for details.
+See section \ref{Causes} for details.
 The number above each panel indicates how many statistics out of $100$ were less than or equal to $0$.
@@
 structure of the data were multivariate normal (i.e., the same copula as a 
-multivariate normal distribtion), but the data were otherwise statistically
+multivariate normal distribution), but the data were otherwise statistically
 unchanged. Significant differences indicate that non-normal copula structure in the data 
@@
 
-# Results for Q3: Tail dependence influences skewness of the spatial average, extinction risk, and Taylor's law
+# Results for Q3: Tail dependence influences skewness of the spatial average, extinction risk, and Taylor's law \label{Results_Q3}
 
@@
 (i.e., for simulations that gave normal-copula dependence between populations
-in different locations), $p$-values for the linearty
+in different locations), $p$-values for the linearity
 and homoskedasticity tests were roughly uniformly distributed across replicate 
@@
 across all 1000 simulations are displayed. Spearman-preserving surrogates were used, 
-though results using Kendall-preserving surrogates were similar. See the methods
-section for Q3 and Appendices \ref{SM-Consequence_TL_simulation} and 
+though results using Kendall-preserving surrogates were similar. See section \ref{Consequences}
+and Appendices \ref{SM-Consequence_TL_simulation} and 
 \ref{SM-sect:copsurrognd} for details.\label{fig_spTL}}
@@
 (Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
-that build upon existing copula methods.
+that have built upon existing copula methods.
 
@@
 the mechanism, are so common, it is reasonable to hypothesize that the mechanism
-may opperate commonly. It may be a dominant cause of 
+may operate commonly. It may be a dominant cause of 
 asymmetric tail associations and non-normal copula struture of ecological 
@@
     [@Felsenstein1985] is currently the second most-cited paper in the history
-    of the journal the American Naturalist [@HueyGT2019].
+    of The American Naturalist journal [@HueyGT2019].
 Frustratingly, the field still lacks a dependable procedure for dealing with branch lengths, 
@@
 which our modelling results
-supported (section on results for Q2), therefore suggests that spatial 
+supported (section \ref{Results_Q2}), therefore suggests that spatial 
 dependence between green spruce aphid counts
@@
 aphid abundance for each of the 10 sampling locations, using the same model selection
-and nonparametric methods detailed in the section on methods for Q1. The column labeled p, CvM is the p-value 
+and nonparametric methods detailed in section \\ref{Methods}. The column labeled p, CvM is the p-value 
 result for the goodness of fit test using the Cramer-von Mises statistic. The 
@@
 We carried out most of our analyses on data ranks, $u_i$ and $v_i$, for good reasons 
-mentioned in the Introduction, Background, and methods sections and 
+mentioned in sections \ref{Intro}, \ref{Background} and \ref{Methods} and 
 reviewed here; but under some circumstances it may also be appropriate to use 

### 2019-10-22 SupportingInformation.Rmd
ready to submit, I think

@@
 title: "Supporting information: Copulas and their potential for ecology"
-author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance E. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
+author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance D. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
 geometry: "left=1cm,right=1cm,top=2.2cm,bottom=2.2cm"
@@
 upper-tail dependence can vary independently of correlation. (A-C) Data generated using the BB1 
-copula (see Background section in the main text) 
+copula (see section \ref{MT-Background} in the main text) 
 with three different 
@@
 with stronger correlation. Asymmetry of tail dependence (LT-UT) is the same 
-within each column. See Background section in the main text for a formal description of 
+within each column. See section \ref{Background} in the main text for a formal description of 
 tail dependence and other measures of association in the tails.\label{fig_cop_pedag}}
@@
 # Spatial resampling with multivariate datasets for nonparametric statistics \label{spatial_resamp}
-<!--***DAN: leave this section exactly as it is, but it may be out of order
-now that we describe how this is done in the methods. In fact, I think I
-have to do this more generally.-->
 
@@
 The output of simulations from the first model detailed in 
-the methods section for Q2 (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
+section \\ref{Causes} of the main text (i.e., Figs \\ref{MT-fig_ms_asym_b_0.1} and 
 \\ref{fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to the model 
-selection procedures detailed in the 
-methods section for Q1. Best-fit copula AICw: the AIC weight of the best-fitting copula family;
+selection procedures detailed in section \\ref{Methods} of the main text. 
+Best-fit copula AICw: the AIC weight of the best-fitting copula family;
 Normal copula AICw: the AIC weight of the normal copula family; Model-avg. LT: the lower-tail 

### 2019-12-13 AER_Responses.tex
new responses file and working on some changes in response to ref 1

Hi Dan,
I have sent your Adv. Ecol. Res. submission to two peers for review. One has already returned their review (below) and I have spoken to the other who assure me it will be submitted shortly and that they are overall positive about the work. Thus I'm happy to accept this paper pending you addressing the referees' comments and thought I would send the first of these over now so that you can get cracking. I'll forward the other referee's comments shortly once I receive them. I hope this is not too onerous a task, and look forward to receiving your revisions. The final deadline for publication is January 8th 2020 - so please could you submit the final revised version then. Please note minor edits can easily be taken care of in the proof stage and it is more important to get a paper to the publishers by the 8th and then edited the proofs than it is to submit a perfect version late. I hope this makes sense. You'll also shortly receive an email from the online submission system letting you know you MS has been returned to you, this open the submission portal for the revision to be submitted. Once again, thanks for your contribution - the volume is shaping up nicely!
Cheers,
Alex
REFEREE 1:
Review of Ghosh et al., "Copulas and their potential for ecology"

%DAN: started here

The paper by Ghosh et al provides a introduction and evaluation of a modelling technique, 
copula modelling, in ecology and environmental science. The authors provide some basic 
background theory as to what copulas are and what information they can summarise, and through 
a series of case studies address multiple questions relating to the suitability of this technique 
for modelling and understanding various ecological phenomena. As a reader who was not previously 
familiar with copula modelling, I think the paper is a very useful introduction and has certainly 
convinced me of the usefulness of copula modelling techniques - I will be exploring this family 
of statistics further.

>>>Thank you for the positive feedback! It was our goal to get readers to see the usefulness of 
copulas and consider them for their own research, so we are very happy to hear this!

I did however find myself a little lost in certain parts, and therefore most of my comments relate 
to guiding an audience who may be, statistically competent rather than confident. Overall, I 
think the paper would make a nice addition to "Advances in Ecological Research" series.

>>>Thank you for the positive feedback!

- When describing the properties of the different copula families, it might be helpful to have a 
figure that allows for quick and easy comparisons.

>>>Thank you for this suggestion. There were 16 families, and the properties 
of copulas in each family depend on one or two parameters, depending on the 
family, thus requiring multiple panels per family. This volume of pictorial 
information would be extremely hard or impossible to summarize effectively in a 
single figure, so we instead have several figures, each of which visually 
shows properties of one family. These can be compared to give the functionality that 
the referee seems to be looking for. See Figs 4-6, S3 and S7-S18. See also the new
table described in our next response.

- A minor point, but I think that subdividing the descriptions of copula properties by family could 
help researchers to quickly identify the maths/parameter descriptions relevant to their chosen copula 
family.

>>>We were not entirely sure what the referee was getting at here, but we
thank the referee for the comment because it spurred us to create a new table, 
table X, which summarizes the properties of the different families.
%***DAN: Replace X with whatever the number of the new table turns out to be-->

- line 318 - units should be mg

>>>Actually Mg, meaning megagrams, is correct. We clarified by replacing "Mg" with "megagrams".

- What are the data requirements for robust copula modelling? I appreciate this is a broad question, but generally do these methods work well for small sample sizes, or do we need large n.

>>>

- Concepts and methods sections would be easier to read if the questions were included, so that the reader is not forced to backtrack through the manuscript.

>>>

- The manuscript is very long and quite intimidating for a lay reader such as myself. I wonder whether reducing the number of case studies might help?

>>>

- When discussing tail-associations, it might be helpful to briefly suggest what a left or right tail association would mean in ecological terms.

>>>

- Building on the above point, I think the entire results section would benefit from "touching base" with ecology more often, mainly so that readers who perhaps aren't as confident following the maths/stats can still derive the key ecological points from your work.

>>>

### 2019-12-13 Paper.Rmd
abstract word count:253
new responses file and working on some changes in response to ref 1

@@
 
-\noindent \emph{The number of words in the abstract:} 263
+\noindent \emph{The number of words in the abstract:} 253
 
@@
 The database comprises measurements of soil organic carbon and total soil 
-nitrogen (Mg C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
+nitrogen (megagrams C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
 locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
@@
 we only briefly identify each family, say a few words about tail dependence 
-of the families, and provide pictorial descriptions in the Appendices.
+of the families, and provide pictorial descriptions (Figs 
+\ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8}). 
+The pictorial decriptions can also be used to aid quick comparisons of copula
+families when alternative families are being considered for future applications. 
 We used several families that can exhibit positive lower-tail dependence 
@@
 of these families provided in the `VineCopula` package for the R programming
-language. See Figs \ref{figped_N}-\ref{figped_BB1} and \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
+language. See Figs \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
 for visual depictions of the pdfs of these copulas and how
-pdfs and tail dependence are influenced by the parameters.
+pdfs and tail dependence are influenced by the parameters. See 
+Table X for a summary of some properties of each family.
+<!--***DAN: Shyamolina, please make a table, and replace the X above with 
+a citation to that table. Please place the table chunk in such a way that 
+this citation becomes table 2, since prior to this point the only main text
+table that has been cited is table 1, and we want the table citations to 
+come in order. The columns of the table should include: 1) copula family name,
+2) number of parameters for the family; 3) info on tail dependence that is
+in the above para; 4) a reference to the figure (in the main text or sup mat)
+where there is a picture of this family. Anything else? There should be a 
+row for each of our 16 copula families.-->
 

### 2019-12-13 Paper.Rmd
some more working on responses for AER

@@
 
+<!--para on data requirements of the methods, requested by the AER referees-->
+Data requirements for copula methods will vary depending on multiple factors,
+but our results show that data requirements are not beyond what is common 
+in ecology. With copula methods as with any statistical analysis, the data required to detect 
+an effect or a phenomenon depends on the strength of the effect. For instance,
+detection of mild asymmetry of tail association will require more data than
+detection of strong asymmetry. Different copula or copula-related methods will
+also require different amount of data. For instance, our nonparametric methods 
+such as the partial Spearman correlation should be effective with datasets 
+only slightly bigger than typical guidelines for standard correlation 
+methods; whereas our model-selection approaches may be most effective with more data.
+Some of our datasets were large, but others were much smaller, e.g., aphid datasets
+comprised 30 points, and our analyses still provided robust results. Thus
+many copula methods will be suitable for datasets of sizes that are common in ecology.
+The partial Spearman correlation seems particularly widely applicable.
+
 <!--additional copula methods-->

### 2019-12-14 Paper.Rmd
table added

@@
              caption = "Summary table for the data we used.
-Bold entries are the multivariate data, the rest are bivariate datasets (see section \ref{Data}).
+Bold entries are the multivariate data, the rest are bivariate datasets (see section \\ref{Data}).
 Basal metabolic rate=BMR.
@@
              row_spec(c(4,5,6,8), bold=T)
+```
+
+```{r tab_summary_copula_info, echo=F, results="asis",message=F}
+library(kableExtra)
+library(dplyr)
+tab_copula_info<-as.data.frame(matrix(NA,nrow=16,ncol=4))
+colnames(tab_copula_info)<-c("Copula family","No. of parameters","Stronger dependent tail", "Reference")
+
+tab_copula_info$`Copula family`<-c("Normal",
+                                   "Clayton","Survival Clayton",
+                                   "Gumbel","Survival Gumbel",
+                                   "Joe","Survival Joe",
+                                   "Frank", 
+                                   "BB1","Survival BB1",
+                                   "BB6","Survival BB6", 
+                                   "BB7","Survival BB7", 
+                                   "BB8","Survival BB8")
+
+tab_copula_info$`No. of parameters`<-c("1",
+                                       "1","1",
+                                       "1","1",
+                                       "1","1",
+                                       "1",
+                                       "2","2",
+                                       "2","2",
+                                       "2","2",
+                                       "2","2")
+
+tab_copula_info$`Stronger dependent tail`<-c("None",
+                                             "Lower","Upper",
+                                             "Upper","Lower",
+                                             "Upper","Lower",
+                                             "None",
+                                             "Any (depending on the parameters)","Any (depending on the parameters)",
+                                             "Upper","Lower",
+                                             "Any (depending on the parameters)","Any (depending on the parameters)",
+                                             "Either upper or none (depending on the parameters)","Either lower or none (depending on the parameters)")
+
+# caution: this fig. numbers are not auto-linked
+tab_copula_info$Reference<-c("Fig. 4",
+                              "Fig. 5","Fig. S3",
+                              "Fig. S7","Fig. S8",
+                              "Fig. S9","Fig. S10",
+                              "Fig. S11",
+                              "Fig. 6","Fig. S12",
+                              "Fig. S13","Fig. S14",
+                              "Fig. S15","Fig. S16",
+                              "Fig. S17","Fig. S18")
+
+knitr::kable(tab_copula_info, 
+             format="latex", align="l",escape=F,
+             caption = "Summary table for the sixteen copula families we used (see section \\ref{Background}). \\label{tab_copula_info}", 
+             booktabs = T,linesep = "\\addlinespace") %>%
+             column_spec(2,width="2 cm")
 ```
@@
 pdfs and tail dependence are influenced by the parameters. See 
-Table X for a summary of some properties of each family.
+Table \ref{tab_copula_info} for a summary of some properties of each family.
 <!--***DAN: Shyamolina, please make a table, and replace the X above with 

### 2019-12-15 AER_Responses.tex
some more working on responses for AER
a bit more progress on responses to AER refs

@@
 thank the referee for the comment because it spurred us to create a new table, 
-table X, which summarizes the properties of the different families.
-%***DAN: Replace X with whatever the number of the new table turns out to be-->
+table 2, which summarizes the properties of the different families.
 
@@
 
->>>Actually Mg, meaning megagrams, is correct. We clarified by replacing "Mg" with "megagrams".
+>>>Actually Mg, meaning megagrams, is correct. We clarified by replacing "Mg" 
+with "megagrams".
 
-- What are the data requirements for robust copula modelling? I appreciate this is a broad question, but generally do these methods work well for small sample sizes, or do we need large n.
+- What are the data requirements for robust copula modelling? I appreciate this 
+is a broad question, but generally do these methods work well for small sample 
+sizes, or do we need large n.
 
->>>
+>>>Thank you for asking this question, as it prompted us to we add a new paragraph
+on this topic to the Discussion. The new paragraph begins "Data requirements for 
+copula methods..." and is the second-to-last paragraph of the main text.
 
-- Concepts and methods sections would be easier to read if the questions were included, so that the reader is not forced to backtrack through the manuscript.
+- Concepts and methods sections would be easier to read if the questions were 
+included, so that the reader is not forced to backtrack through the manuscript.
 
->>>
+>>>Thanks for the suggestion, we have made this change.
 
@@
 
->>>
+>>>Thank you for this suggestion. We have considered at various times the possibility 
+of splitting the manuscript up
+and/or removing some of its content. In the end we decided it is best to keep it all together,
+for several reasons. First, the parts of the manuscript (theoretical and empirical, the three
+research question, the multiple case studies) reinforce each other, providing better evidence
+for the points we are making than any of these parts would on its own. Second,
+we selected datasets in an attempt to span multiple disciplines of ecology.
+Removing some now would create a less representative collection of datasets across the field, and would weaken
+our claim that copula structure seems to be a common phenomenon, broadly, across ecology. 
+Finally, we selected datasets prior to analysis and we present all results. This was done
+to be sure to fairly evaluate the hypothesis that non-Gaussian copula structures are
+common in ecology. We did *not* carry out additional analyses on other datasets and
+only present the results with interesting copula structure. We would prefer not to 
+modify the manuscript in such a way as to invalidate this strength of our work.
 

### 2019-12-15 Paper.Rmd
a bit more progress on responses to AER refs

@@
 tab_copula_info<-as.data.frame(matrix(NA,nrow=16,ncol=4))
-colnames(tab_copula_info)<-c("Copula family","No. of parameters","Stronger dependent tail", "Reference")
+colnames(tab_copula_info)<-c("Copula family","No. of parameters","Tails with stronger association", "Reference")
 
@@
 
-tab_copula_info$`Stronger dependent tail`<-c("None",
+tab_copula_info$`Tails with stronger association`<-c("Neither (symmetric)",
                                              "Lower","Upper",
@@
                                              "Upper","Lower",
-                                             "None",
-                                             "Any (depending on the parameters)","Any (depending on the parameters)",
+                                             "Neither (symmetric)",
+                                             "Either (depending on params)","Either (depending on params)",
                                              "Upper","Lower",
-                                             "Any (depending on the parameters)","Any (depending on the parameters)",
-                                             "Either upper or none (depending on the parameters)","Either lower or none (depending on the parameters)")
+                                             "Either (depending on params)","Either (depending on params)",
+                                             "Upper or symmetric (depending on paras)","Lower or symmetric (depending on params)")
 
@@
              format="latex", align="l",escape=F,
-             caption = "Summary table for the sixteen copula families we used (see section \\ref{Background}). \\label{tab_copula_info}", 
+             caption = "Summary table for the sixteen copula families we used (see section \\ref{Background}).
+For the parameters we consider, all copula families model positive association between variables.
+\\label{tab_copula_info}",
              booktabs = T,linesep = "\\addlinespace") %>%
@@
 
+Recall that Q1 has three parts: Do datasets
+in ecology have dependence structure distinct from that of a multivariate 
+Gaussian/normal
+distribution (here called non-normal copula structure)? Do positvely associated 
+ecological variables show tail 
+associations distinct from those of a normal distribution? And in particular
+do they show asymmetric tail associations? 
 We addressed Q1 first via a model selection procedure in which several
@@
 
+Recall that Q2 is: what are some possible causes/mechanisms of non-normal
+copula structure and asymmetric tail associations in ecology?
 Having demonstrated that non-normal copula structure and asymmetric tail associations are 
-addressed Q2 by exploring, using models, three possible mechanisms
+common in ecology, we addressed Q2 by exploring, using models, three possible mechanisms
 by which these phenomena may arise. 
@@
 
+Recall that Q3 is: what are the consequences of non-normal copula
+structure and asymmetric tail associations for ecological understanding and
+applications?
 We addressed Q3 by exploring, using both data and models, three hypothesized

### 2019-12-16 AER_Responses.tex
working on referee comments, second referee now included

@@
 >>>
+
+Referee 2
+
+Review of Ghosh et al.: Copulas and the Potential for Ecology
+I would like to thank the authors for their manuscript entitled "Copulas and their potential for ecology".
+The manuscript provides an introduction to the concept of copulas and their utility for application in studying the natures of variable dependency in ecology.
+The authors demonstrate that the most commonly used models of inference, correlation and regression, only encapsulate certain features of the dependency between variables and the author introduce a more complete method of inference to an ecologist audience.
+I must first state that I must give fair warning that, whilst I have heard about copulas before, I have never applied them to my own research and I certainly cannot claim to be an expert in this area.
+I am therefore probably the target audience of this paper.
+This means that whilst I can give constructive feedback on the author's introduction of the concepts to an ecologist readership and can follow the mathematical background that they have provided, I cannot necessarily state definitively whether the authors have reproduced a fair portrayal of the theory of copulas in this manuscript.
+However, despite not being a specialist on the subject, I found the presentation of the manuscript reasonably easy to follow and the authors should be commended for making what is quite a dense subject approachable.
+Each theoretical claim is backed by well-documented supplementary materials which allows for digestion of the key concepts without being side-tracked by dense mathematical derivations.
+
+>>>Thank you for the positive feedback.
+
+The authors detail a large number of example ecological applications and detail a strong case for the use of copula families with left- or right- tail associations.
+This is a very good demonstration where the use of copulas can benefit the ecological investigator far above the traditional methods.
+
+>>>Thank you for the positive feedback.
+
+I would however appreciate a discussion of how copula-based methods compare with regression models with a heteroscedastic error term.
+It seems that these models would also be able to cope with left- or right- tail associations in a bivariate application.
+Given the rise of hierarchical Bayesian models in ecology, before reading this paper, my (maybe somewhat naiive) approach would be to apply a standard regression model but allow for the variance in the error distribution to also be determined by the predictor variable.
+
+>>><DAN: to add some discussion material>
+
+In my view this paper represents an impressive contribution to statistical ecology and I am very happy to be introduced to a new and exciting avenue of analysis.
+Indeed, the thing I hate most about the manuscript is that it has spurred me on to spend hundreds of dollars on new academic books on copula modelling.
+
+>>>Thank you for the positive feedback. This was exactly our goal!
+
+What follows is a small set of very minor points that are all highly subjective and whether or not they are addressed does not, in my view, preclude the manuscript of being published:
+o    The authors use the term "distribution function" to represent "cumulative distribution function".
+My feeling is that ecologists in general are more familiar with the term "cumulative distribution function" (which thereafter can be abbreviated to "CDF") as the former is more likely to be mistaken for the density function.
+
+>>><DAN to do>
+
+o    The manuscript could benefit from a little more detail in the introduction as to how this material goes further than that in other ecological introductions to copulas.
+For example, Anderson et al.
+2018 also contains an introduction to the theory of copulas so if might be worth explicitly pointing out how this paper goes further in explaining and introducing the concept to an ecologist audience.
+
+>>><DAN to do>
+
+o    A table giving a run-down of the notation used in the manuscript might be useful for the more mathematically-adverse reader.
+A sort of notation cheat-sheet for easy notation lookup would be useful for readers so that they can concentrate on the new concepts rather than looking up notation definitions.
+
+>>>
+
+o    The authors make reference to many families of copula models (a total of 16 families) and, whilst there is a graphical illustration of the shapes of the copulas, we don't see the functional forms of these copulas.
+A table either in the main text or supplementary materials giving these would be appreciated.
+
+>>>

### 2019-12-16 Paper.Rmd
working on referee comments, second referee now included

@@
 <!--***DAN: remianing to do:
-DONE 1) Edit abstract
-DONE 2) Go through the whole thing one more time
-DONE 3) send it out to coauthors (done) and maybe get feedback
-DONE 3.5) ESP FROM lAWRENCE
-DONE 3.6) Incorporate Shya feedback
-DONE 3.7) Incorporate Reid feedback
-DONE 4)make sure to say something about tropical systems
-DONE 5) maybe write to the editors saying this is coming soon?
-DONE 6) In the Discussion, you maybe need to point out not only that we
-use only bivariate copulas (and there are multivariate ones), but also
-that we use only continuous marginals. Probably an adequate approximation
-for our data, but see the other papers I've seen recently which use count data,
-and discrete marginals. Approaches such as that will be necessary for some 
-datasets.
-DONE 7) change the box around the Cedar Creek data in the main fig
-DONE 8) say in Discussion that you could have used non-ranks for the nonparam stats - this
-is the stuff the Ecol Mongr referee talked about. Maybe this should go in Background instead?
-DONE 9) maybe go back and read the Ecol Monogr reviews again
-DONE 10) try to emphasize early on that the focus is on positively associated variables, so 'lower tail' means the lower tail of both variables and it is not necessary to worry about negative relationships (in practice we would always transform these for ease).
-The dfinitions of tail dependence we use cannot usefully be applied to untransformed data in fig 1e,f.
-DONE 11) Number the sections and subsections in the main text and change all the section 
-references to use those numbers. Search for "section" (DONE) also for "methods" (DONE)
 -->
@@
 have stronger left- than right-tail association. The formula is 
+<!--***DAN: is it clear implicitly from what came before that this is the cdf or what?-->
 $\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 

### 2019-12-17 SupportingInformation.Rmd
upto S20 BnW fig version added

@@
 map("usa")
-points(x = d$long, y = d$lat, col="red",pch=16,cex=0.5)
+points(x = d$long, y = d$lat, col=rgb(0,0,0,0.2),pch=16,cex=0.5)
 dev.off()
@@
 map.axes(cex.axis=2.5)
-points(x = df_loc_info_org$longs, y = df_loc_info_org$lats, col="red",pch=16)
-points(x = df_loc_info$longs, y = df_loc_info$lats, col="blue", pch=0, cex=1.5)
-text(df_loc_info_org$loc_no, x = df_loc_info_org$longs, y = df_loc_info_org$lats, col="red",  pos=1)
+points(x = df_loc_info_org$longs, y = df_loc_info_org$lats, col="black",pch=16)
+points(x = df_loc_info$longs, y = df_loc_info$lats, col="black", pch=0, cex=1.5)
+text(df_loc_info_org$loc_no, x = df_loc_info_org$longs, y = df_loc_info_org$lats, col="black",  pos=1)
 dev.off()
@@
 abline(h=seq(from=ylim[1],to=ylim[2],by=2), v=seq(from=xlim[1],to=xlim[2],by=2), col="black", lty=2)
-points(x = df_loc_info_org$longs+1, y = df_loc_info_org$lats+1, col="red",pch=16)
-points(x = df_loc_info$longs+1, y = df_loc_info$lats+1, col="blue", pch=0, cex=1.5)
-text(df_loc_info_org$loc_no, x = df_loc_info_org$longs+1, y = df_loc_info_org$lats+1, col="red",  pos=1)
+points(x = df_loc_info_org$longs+1, y = df_loc_info_org$lats+1, col="black",pch=16)
+points(x = df_loc_info$longs+1, y = df_loc_info$lats+1, col="black", pch=0, cex=1.5)
+text(df_loc_info_org$loc_no, x = df_loc_info_org$longs+1, y = df_loc_info_org$lats+1, col="black",  pos=1)
 dev.off()
@@
 Rothamsted Insect Survey 
-suction traps from which data were available are indicated by the red circles, 
-out of which only blue boxes had at least 30 years of data and were 
+suction traps from which data were available are indicated by the circles, 
+out of which only squares had at least 30 years of data and were 
 used for our study. \label{fig_aphid_map}}
@@
 26 locations ($2^{\circ}$ $\times$ $2^{\circ}$ regions on the map) around UK 
-seas (red circles), out of which only blue boxes had at least 45 years of data and
+seas (circles), out of which only squares had at least 45 years of data and
 were used for our study. \label{fig_plankton_map}}
@@
 issue a message.-->
-```{r stat_testing_save_and_plot_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,busycores,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
+```{r stat_testing_save_result, message=F, results="hide", echo=F, cache=T, cache.extra=list(seed,busycores,mtime("TestStats_multivar_dataset.R"),mtime("CopulaFunctions.R"),mtime("CopulaFunctions_flexible.R"))}
 source("TestStats_multivar_dataset.R")
@@
 b<-callfn$paramlist
-plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b,resultsloc,spearvals)
-plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b,resultsloc,spearvals)
-plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b,resultsloc,spearvals)
-plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b,resultsloc,spearvals)
+
+plotter_stat_testing(reslist=res_pt35_rankF,filename="res_pt35_rankF",xaxparams=b,resultsloc,spearvals)
+plotter_stat_testing(reslist=res_pt35_rankT,filename="res_pt35_rankT",xaxparams=b,resultsloc,spearvals)
+plotter_stat_testing(reslist=res_pt1000_rankF,filename="res_pt1000_rankF",xaxparams=b,resultsloc,spearvals)
+plotter_stat_testing(reslist=res_pt1000_rankT,filename="res_pt1000_rankT",xaxparams=b,resultsloc,spearvals)
 ```
@@
 \caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the normal and Frank copulas.
+p-values (indicated by the triangles) when less than $0.05$ (below the dashed line) mean there is significant tail asymmetry.
 See Appendix  \ref{Test_npa_stat} for details.
@@
 \caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the normal and Frank copulas.
+p-values (indicated by the triangles) when less than $0.05$ (below the dashed line) mean there is significant tail asymmetry.
 See Appendix  \ref{Test_npa_stat} for details.

### 2019-12-17 Paper.Rmd
fig in greyscale fig_s25-s29, fig_15

@@
 Gaussian/normal
-distribution (here called non-normal copula structure)? Do positvely associated 
+distribution (here called non-normal copula structure)? Do positively associated 
 ecological variables show tail 
@@
 environmental-noise inputs and ecological-variable 
-outputs, AR(1) model, Clayton copula. Red and blue points give 
+outputs, AR(1) model, Clayton copula. Grey and black points give 
 means over 50 replicate simulations of the listed statistics, 
-red points are for noise inputs and blue points are for
+grey points are for noise inputs and black points are for
 model outputs. Error bars are standard errors. 
@@
 For bottom panels, headers give Pearson correlations for
-the points. The regression line through the points (black) was similar
-to the 1-1 line (green). 
+the points. The regression line through the points (solid) was similar
+to the 1-1 line (dashed). 
 See section \ref{Causes} for details.\label{fig_Cause4copula_C_kend_ar1}}
@@
 simulations were plotted on the same axes for each panel, but two example simulations are 
-shown in red and green to help assess to what extent variation on the plots was between 
+shown in solid and void black dots to help assess to what extent variation on the plots was between 
 simulations or within simulations. For each simulation, we tested the linearity 

### 2019-12-17 AER_Responses.tex
working on ref comments editorially

@@
-Hi Dan,
-I have sent your Adv.
-Ecol.
-Res.
-submission to two peers for review.
-One has already returned their review (below) and I have spoken to the other who assure me it will be submitted shortly and that they are overall positive about the work.
-Thus I'm happy to accept this paper pending you addressing the referees' comments and thought I would send the first of these over now so that you can get cracking.
-I'll forward the other referee's comments shortly once I receive them.
-I hope this is not too onerous a task, and look forward to receiving your revisions.
-The final deadline for publication is January 8th 2020 - so please could you submit the final revised version then.
-Please note minor edits can easily be taken care of in the proof stage and it is more important to get a paper to the publishers by the 8th and then edited the proofs than it is to submit a perfect version late.
-I hope this makes sense.
-You'll also shortly receive an email from the online submission system letting you know you MS has been returned to you, this open the submission portal for the revision to be submitted.
-Once again, thanks for your contribution - the volume is shaping up nicely!
-Cheers,
-Alex
 REFEREE 1:
+
 Review of Ghosh et al., "Copulas and their potential for ecology"
-
-%DAN: started here
 
@@
 >>>Thank you for the positive feedback! It was our goal to get readers to see the usefulness of
-copulas and consider them for their own research, so we are very happy to hear this!
+copulas and consider them for their own research, so we are very happy to hear we have succeeded!
 
@@
 
->>>Thanks for the suggestion, we have made this change.
+>>>Thanks for the suggestion, we made this change.
 
@@
 we selected datasets in an attempt to span multiple disciplines of ecology.
-Removing some now would create a less representative collection of datasets across the field, and would weaken
-our claim that copula structure seems to be a common phenomenon, broadly, across ecology.
+Removing some now would create a less representative collection 
+of datasets across the field, and would weaken
+our claim that non-Gaussian and informative 
+copula structure seems to be a common phenomenon, broadly, across ecology.
 Finally, we selected datasets prior to analysis and we present all results. This was done
@@
 common in ecology. We did *not* carry out additional analyses on other datasets and
-only present the results with interesting copula structure. We would prefer not to
+then only present the results with interesting copula structure. We would prefer not to
 modify the manuscript in such a way as to invalidate this strength of our work.
@@
 
->>>
+>>>Tail associations are introduced in the third paragraph of the Introduction, and four 
+conceptual examples of the ecological implications/meaning of copula structure and especially 
+of tail associations are given in paragraphs 5-8 of the Introduction. These paragraphs 
+serve to provide an initial idea of what a left- or right-tail association would mean 
+in ecological terms. Because there is a paragraph between paragraphs 3 and 5, we added 
+a note to paragraph 3 indicating that "We give four examples below which illustrate some of the 
+potential ecological meaning of tail associations."
 
@@
 
-Referee 2
+REFEREE 2
 
@@
 
->>>
+>>><Shyamolina is building a preliminary version of the table>
 
@@
 
->>>
+>>><Shyamolina is collecting the formulas>

### 2019-12-17 Paper.Rmd
working on ref comments editorially

--- 2019-12-16 Paper.Rmd
@@
 usefully describe the full structure of dependence, so tail association 
-is one of the main foci of this paper.
+is one of the main foci of this paper. We give four examples below which
+illustrate some of the potential ecological meaning of tail associations.
 
@@
 
-One way we may expect, *a priori*, a copula 
-study of complex dependence structures between
+One way we may expect, *a priori*, a study of tail associations and of the 
+complex copula dependence structures between
 ecological variables
@@
 look more like Fig. \ref{fig_cop_pedag1}A or C than like 
-Fig. \ref{fig_cop_pedag1}B or D. Nitrogen 
+Fig. \ref{fig_cop_pedag1}B or D, i.e., the two variables may show 
+left-tail association. Nitrogen 
 will control plant growth, producing a clear
@@
 variable. This influence is called the \emph{Moran effect}. 
-If complex dependence structure is transmitted, 
+If asymmetric tail associations, or other complex dependence structure, is transmitted, 
 possibly in modified form, 
 from environmental to ecological variables, then we may expect complex dependence 
-structure to be a common feature of the 
+structure and tail associations to be a common feature of the 
 spatial synchrony of population, community, 
@@
 A third reason why we may expect an understanding of complex 
-dependence structures to benefit 
+dependence structures and tail associations to benefit 
 ecological research is that such an approach may help illuminate 
@@
 fundamentally based on associations. However, more detailed 
-examination of the patterns of dependence of variables may
+examination of the tail associations and potentially other 
+patterns of dependence of variables may
 well be combined with biological and other reasoning to rule out
@@
 
-A fourth reason why we may expect a copula understanding of complex 
+A fourth reason why we may expect an understanding of tail associations and of complex 
 dependence structures to be useful
@@
 and the total economic value of a fishery is the sum of local catches. 
-We will explore how details, which can be captured using copulas, 
+We will explore how details, which can be captured using copulas and which
+relate to tail associations, 
 of the dependence between fluctuations of local
@@
 have stronger left- than right-tail association. The formula is 
-<!--***DAN: is it clear implicitly from what came before that this is the cdf or what?-->
 $\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 

### 2019-12-17 SupportingInformation.Rmd

@@
 
-```{r plot_cause4copula_stat_ricker_C, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_C,mtime("Plotter_Cause4copula.R"))}
+```{r plot_cause4copula_stat_ricker_C, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_C,mtime("Plotter_Cause4copula_stat.R"))}
 
@@
 
-```{r plot_cause4copula_stat_ricker_SC, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SC,mtime("Plotter_Cause4copula.R"))}
+```{r plot_cause4copula_stat_ricker_SC, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SC,mtime("Plotter_Cause4copula_stat.R"))}
 
@@
 
-```{r plot_cause4copula_stat_ricker_G, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_G,mtime("Plotter_Cause4copula.R"))}
+```{r plot_cause4copula_stat_ricker_G, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_G,mtime("Plotter_Cause4copula_stat.R"))}
 
@@
 
-```{r plot_cause4copula_stat_ricker_SG, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SG,mtime("Plotter_Cause4copula.R"))}
+```{r plot_cause4copula_stat_ricker_SG, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SG,mtime("Plotter_Cause4copula_stat.R"))}
 
@@
 
-```{r plot_cause4copula_stat_ricker_J, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_J,mtime("Plotter_Cause4copula.R"))}
+```{r plot_cause4copula_stat_ricker_J, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_J,mtime("Plotter_Cause4copula_stat.R"))}
 
@@
 
-```{r plot_cause4copula_stat_ricker_SJ, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SJ,mtime("Plotter_Cause4copula.R"))}
+```{r plot_cause4copula_stat_ricker_SJ, echo=F, results="hide",cache=T, cache.extra=list(seed,tag_ricker_SJ,mtime("Plotter_Cause4copula_stat.R"))}
 
@@
 ##plot TL for clayton
-plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+plot(kend.points.res[[1]][,1,],kend.points.res[[1]][,2,],type='p',pch=19,cex=.75,xlim=xlimits,ylim=ylimits,
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=2,cex.lab=3)  
-points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
-points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
+     cex.axis=2,cex.lab=3,col="grey")  
+points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=19,cex=.75,col='black')
+points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=1,cex=.75,col='black')
 mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black",cex=2)
@@
 ##plot TL for normal copula
-plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+plot(kend.points.res[[2]][,1,],kend.points.res[[2]][,2,],type='p',pch=19,cex=.75,xlim=xlimits,ylim=ylimits,
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=2,cex.lab=3)  
-points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
-points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
+     cex.axis=2,cex.lab=3,col="grey")  
+points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=19,cex=.75,col='black')
+points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=1,cex=.75,col='black')
 mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black",cex=2)
@@
 #plot TL for frank
-plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+plot(kend.points.res[[3]][,1,],kend.points.res[[3]][,2,],type='p',pch=19,cex=.75,xlim=xlimits,ylim=ylimits,
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=2,cex.lab=3)  
-points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
-points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
+     cex.axis=2,cex.lab=3,col="grey")  
+points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=19,cex=.75,col='black')
+points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=1,cex=.75,col='black')
 mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black",cex=2)
@@
 #plot TL for survival clayton
-plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+plot(kend.points.res[[4]][,1,],kend.points.res[[4]][,2,],type='p',pch=19,cex=.75,xlim=xlimits,ylim=ylimits,
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=2,cex.lab=3)  
-points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
-points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
+     cex.axis=2,cex.lab=3,col="grey")  
+points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=19,cex=.75,col='black')
+points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=1,cex=.75,col='black')
 mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black",cex=2)
@@
 #}
-boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="p-vals", cex.axis=2,cex.lab=3,xaxt="n")
+boxplot(s.lin , col="grey", xlab="Copula family", ylab="p-vals", cex.axis=2,cex.lab=3,xaxt="n")
 mtext("Linearity test", side=3, cex=1.8,line=1.5)
@@
 #plot s.hom distr for each set of surrogates
-boxplot(s.hom , col=rainbow(4), xlab="Copula family", ylab="p-vals", cex.axis=2,cex.lab=3,xaxt="n")
+boxplot(s.hom , col="grey", xlab="Copula family", ylab="p-vals", cex.axis=2,cex.lab=3,xaxt="n")
 mtext("Homoskedasticity test", side=3, cex=1.8,line=1.5)
@@
 #plot s.rmse distr for each set of surrogates
-boxplot(s.rmse , col=rainbow(4) , xlab="Copula family", ylab="rmse", cex.axis=2,cex.lab=3,xaxt="n")
+boxplot(s.rmse , col="grey" , xlab="Copula family", ylab="rmse", cex.axis=2,cex.lab=3,xaxt="n")
 mtext("Spatial TL", side=3, cex=1.8,line=1.5)
@@
 #plot s.int distr for each set of surrogates
-boxplot(s.int , col=rainbow(4) , xlab="Copula family", ylab="intercept", cex.axis=2,cex.lab=3,xaxt="n")
+boxplot(s.int , col="grey" , xlab="Copula family", ylab="intercept", cex.axis=2,cex.lab=3,xaxt="n")
 mtext("Spatial TL", side=3, cex=1.8,line=1.5)
@@
 #plot s.slope distr for each set of surrogates
-boxplot(s.slope , col=rainbow(4) , xlab="Copula family", ylab="slope", cex.axis=2,cex.lab=3,xaxt="n")
+boxplot(s.slope , col="grey" , xlab="Copula family", ylab="slope", cex.axis=2,cex.lab=3,xaxt="n")
 mtext("Spatial TL", side=3, cex=1.8,line=1.5)
@@
 #plot s.quadcoeff distr for each set of surrogates
-boxplot(s.quadcoeff , col=rainbow(4) , xlab="Copula family", ylab="quadratic \n coefficient", cex.axis=2,cex.lab=3,xaxt="n")
+boxplot(s.quadcoeff , col="grey" , xlab="Copula family", ylab="quadratic \n coefficient", cex.axis=2,cex.lab=3,xaxt="n")
 mtext("Spatial TL", side=3, cex=1.8,line=1.5)
@@
 #plot s.mncurv distr for each set of surrogates
-boxplot(s.mncurv , col=rainbow(4) , xlab="Copula family", ylab="mean \n curvature", cex.axis=2,cex.lab=3,xaxt="n")
+boxplot(s.mncurv , col="grey" , xlab="Copula family", ylab="mean \n curvature", cex.axis=2,cex.lab=3,xaxt="n")
 mtext("Spatial TL", side=3, cex=1.8,line=1.5)
@@
 ##plot TL for clayton
-plot(spea.points.res[[1]][,1,],spea.points.res[[1]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+plot(spea.points.res[[1]][,1,],spea.points.res[[1]][,2,],type='p',pch=19,cex=.75,xlim=xlimits,ylim=ylimits,
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=2,cex.lab=3)  
-points(spea.points.res[[1]][,1,1],spea.points.res[[1]][,2,1],type='p',pch=20,cex=.75,col='red')
-points(spea.points.res[[1]][,1,2],spea.points.res[[1]][,2,2],type='p',pch=20,cex=.75,col='green')
+     cex.axis=2,cex.lab=3,col="grey")  
+points(spea.points.res[[1]][,1,1],spea.points.res[[1]][,2,1],type='p',pch=19,cex=.75,col='black')
+points(spea.points.res[[1]][,1,2],spea.points.res[[1]][,2,2],type='p',pch=1,cex=.75,col='black')
 mtext(paste("Clayton"),side = 3, line=-2.5, adj=0.1, col="black",cex=2)
@@
 ##plot TL for normal copula
-plot(spea.points.res[[2]][,1,],spea.points.res[[2]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+plot(spea.points.res[[2]][,1,],spea.points.res[[2]][,2,],type='p',pch=19,cex=.75,xlim=xlimits,ylim=ylimits,
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=2,cex.lab=3)  
-points(spea.points.res[[2]][,1,1],spea.points.res[[2]][,2,1],type='p',pch=20,cex=.75,col='red')
-points(spea.points.res[[2]][,1,2],spea.points.res[[2]][,2,2],type='p',pch=20,cex=.75,col='green')
+     cex.axis=2,cex.lab=3,col="grey")  
+points(spea.points.res[[2]][,1,1],spea.points.res[[2]][,2,1],type='p',pch=19,cex=.75,col='black')
+points(spea.points.res[[2]][,1,2],spea.points.res[[2]][,2,2],type='p',pch=1,cex=.75,col='black')
 mtext(paste("Normal"),side = 3, line=-2.5, adj=0.1, col="black",cex=2)
@@
 #plot TL for frank
-plot(spea.points.res[[3]][,1,],spea.points.res[[3]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+plot(spea.points.res[[3]][,1,],spea.points.res[[3]][,2,],type='p',pch=19,cex=.75,xlim=xlimits,ylim=ylimits,
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=2,cex.lab=3)  
-points(spea.points.res[[3]][,1,1],spea.points.res[[3]][,2,1],type='p',pch=20,cex=.75,col='red')
-points(spea.points.res[[3]][,1,2],spea.points.res[[3]][,2,2],type='p',pch=20,cex=.75,col='green')
+     cex.axis=2,cex.lab=3,col="grey")  
+points(spea.points.res[[3]][,1,1],spea.points.res[[3]][,2,1],type='p',pch=19,cex=.75,col='black')
+points(spea.points.res[[3]][,1,2],spea.points.res[[3]][,2,2],type='p',pch=1,cex=.75,col='black')
 mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black",cex=2)
@@
 #plot TL for survival clayton
-plot(spea.points.res[[4]][,1,],spea.points.res[[4]][,2,],type='p',pch=20,cex=.75,xlim=xlimits,ylim=ylimits,
+plot(spea.points.res[[4]][,1,],spea.points.res[[4]][,2,],type='p',pch=19,cex=.75,xlim=xlimits,ylim=ylimits,
      xlab=expression(log[10]("spatial mean")),ylab=expression(log[10]("spatial variance")),
-     cex.axis=2,cex.lab=3)  
-points(spea.points.res[[4]][,1,1],spea.points.res[[4]][,2,1],type='p',pch=20,cex=.75,col='red')
-points(spea.points.res[[4]][,1,2],spea.points.res[[4]][,2,2],type='p',pch=20,cex=.75,col='green')
+     cex.axis=2,cex.lab=3,col="grey")  
+points(spea.points.res[[4]][,1,1],spea.points.res[[4]][,2,1],type='p',pch=19,cex=.75,col='black')
+points(spea.points.res[[4]][,1,2],spea.points.res[[4]][,2,2],type='p',pch=1,cex=.75,col='black')
 mtext(paste("Survival Clayton"),side=3, line=-2.5, adj=0.1, col="black",cex=2)
@@
 #plot s.lin distr for each set of surrogates
-boxplot(s.lin , col=rainbow(4), xlab="Copula family", ylab="p-vals", cex.axis=2,cex.lab=3,xaxt="n")
+boxplot(s.lin , col="grey", xlab="Copula family", ylab="p-vals", cex.axis=2,cex.lab=3,xaxt="n")
 mtext("Linearity test", side=3, cex=1.8,line=1.5)

### 2019-12-17 Paper.Rmd
caption edited

@@
 \begin{center}
-\includegraphics[width=16 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
+\includegraphics[width=13 cm]{./Results/stat_results/stat_soilCN/soilCN_bivfunctionplot.pdf}
 \caption[Nonparametric tests against a normal-copula null, soil C and N data]{Nonparametric tests for tail association and other deviations from normal copula structure, for soil C and N data.
-As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
-Whenever the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
+As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (dark and light grey x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$.
+Whenever the cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
 For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates.
@@
 model outputs. Error bars are standard errors. 
-$P$-values (purple points, right axis) are for a 
+$P$-values (triangles, right axis) are for a 
 paired $t$-test of the null hypothesis that the distributions have the same mean. 

### 2019-12-17 SupportingInformation.Rmd
caption edited

@@
 \caption[Nonparametric tests against a normal-copula null, Cedar Creek data]{Nonparametric tests for tail dependence and other deviations from normal copula structure, for Cedar Creek data, year 2000.
-As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (red crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (blue and green x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$.
-If the red cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
+As described in the main text, the values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for real data (crosses) were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates of the data (dark and light grey x's show 0.025 and 0.975 quantiles), separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.25),\ldots,(0.75,1)$.
+If the cross was outside the range given by the x's, text at the top of panels indicates the number of surrogate values the real-data value was greater than or less than.
 For instance, a value $>N$ (respectively, $<N$) means the value of the statistic on real data was greater than (respectively, less than) its value on $N$ surrogates.

### 2019-12-17 Paper.Rmd
caption edited

@@
 simulations were plotted on the same axes for each panel, but two example simulations are 
-shown in solid and void black dots to help assess to what extent variation on the plots was between 
+shown in solid and void black circles to help assess to what extent variation on the plots was between 
 simulations or within simulations. For each simulation, we tested the linearity 

### 2019-12-17 SupportingInformation.Rmd
caption edited

@@
      cex.axis=2,cex.lab=3,col="grey")  
-points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=19,cex=.75,col='black')
-points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=1,cex=.75,col='black')
+points(kend.points.res[[1]][,1,1],kend.points.res[[1]][,2,1],type='p',pch=19,cex=2,col='black')
+points(kend.points.res[[1]][,1,2],kend.points.res[[1]][,2,2],type='p',pch=1,cex=2,col='black')
 mtext(paste("Clayton"),side = 3, line=-2, adj=0.1, col="black",cex=2)
@@
      cex.axis=2,cex.lab=3,col="grey")  
-points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=19,cex=.75,col='black')
-points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=1,cex=.75,col='black')
+points(kend.points.res[[2]][,1,1],kend.points.res[[2]][,2,1],type='p',pch=19,cex=2,col='black')
+points(kend.points.res[[2]][,1,2],kend.points.res[[2]][,2,2],type='p',pch=1,cex=2,col='black')
 mtext(paste("Normal"),side = 3, line=-2, adj=0.1, col="black",cex=2)
@@
      cex.axis=2,cex.lab=3,col="grey")  
-points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=19,cex=.75,col='black')
-points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=1,cex=.75,col='black')
+points(kend.points.res[[3]][,1,1],kend.points.res[[3]][,2,1],type='p',pch=19,cex=2,col='black')
+points(kend.points.res[[3]][,1,2],kend.points.res[[3]][,2,2],type='p',pch=1,cex=2,col='black')
 mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black",cex=2)
@@
      cex.axis=2,cex.lab=3,col="grey")  
-points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=19,cex=.75,col='black')
-points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=1,cex=.75,col='black')
+points(kend.points.res[[4]][,1,1],kend.points.res[[4]][,2,1],type='p',pch=19,cex=2,col='black')
+points(kend.points.res[[4]][,1,2],kend.points.res[[4]][,2,2],type='p',pch=1,cex=2,col='black')
 mtext(paste("Survival Clayton"),side=3, line=-2, adj=0.1, col="black",cex=2)
@@
      cex.axis=2,cex.lab=3,col="grey")  
-points(spea.points.res[[1]][,1,1],spea.points.res[[1]][,2,1],type='p',pch=19,cex=.75,col='black')
-points(spea.points.res[[1]][,1,2],spea.points.res[[1]][,2,2],type='p',pch=1,cex=.75,col='black')
+points(spea.points.res[[1]][,1,1],spea.points.res[[1]][,2,1],type='p',pch=19,cex=2,col='black')
+points(spea.points.res[[1]][,1,2],spea.points.res[[1]][,2,2],type='p',pch=1,cex=2,col='black')
 mtext(paste("Clayton"),side = 3, line=-2.5, adj=0.1, col="black",cex=2)
@@
      cex.axis=2,cex.lab=3,col="grey")  
-points(spea.points.res[[2]][,1,1],spea.points.res[[2]][,2,1],type='p',pch=19,cex=.75,col='black')
-points(spea.points.res[[2]][,1,2],spea.points.res[[2]][,2,2],type='p',pch=1,cex=.75,col='black')
+points(spea.points.res[[2]][,1,1],spea.points.res[[2]][,2,1],type='p',pch=19,cex=2,col='black')
+points(spea.points.res[[2]][,1,2],spea.points.res[[2]][,2,2],type='p',pch=1,cex=2,col='black')
 mtext(paste("Normal"),side = 3, line=-2.5, adj=0.1, col="black",cex=2)
@@
      cex.axis=2,cex.lab=3,col="grey")  
-points(spea.points.res[[3]][,1,1],spea.points.res[[3]][,2,1],type='p',pch=19,cex=.75,col='black')
-points(spea.points.res[[3]][,1,2],spea.points.res[[3]][,2,2],type='p',pch=1,cex=.75,col='black')
+points(spea.points.res[[3]][,1,1],spea.points.res[[3]][,2,1],type='p',pch=19,cex=2,col='black')
+points(spea.points.res[[3]][,1,2],spea.points.res[[3]][,2,2],type='p',pch=1,cex=2,col='black')
 mtext(paste("Frank"),side = 3, line=-2, adj=0.1, col="black",cex=2)
@@
      cex.axis=2,cex.lab=3,col="grey")  
-points(spea.points.res[[4]][,1,1],spea.points.res[[4]][,2,1],type='p',pch=19,cex=.75,col='black')
-points(spea.points.res[[4]][,1,2],spea.points.res[[4]][,2,2],type='p',pch=1,cex=.75,col='black')
+points(spea.points.res[[4]][,1,1],spea.points.res[[4]][,2,1],type='p',pch=19,cex=2,col='black')
+points(spea.points.res[[4]][,1,2],spea.points.res[[4]][,2,2],type='p',pch=1,cex=2,col='black')
 mtext(paste("Survival Clayton"),side=3, line=-2.5, adj=0.1, col="black",cex=2)

### 2019-12-17 Paper.Rmd
continued eidtorial to satisfy referee comments

--- 2019-12-17 Paper.Rmd
@@
 normal copula had a much higher AIC, `r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)`, answering the 
-first part of Q1 for these data. Recall that AIC differences of $2$
+first part of Q1 for these data: these data have a dependence structure markedly distinct 
+from that of a normal copula. Recall that AIC differences of $2$
 or $3$ are considered meaningful and differences of $8$ or greater

### 2019-12-18 AER_Responses.tex
changed distribution function to cdf

@@
 
->>>
+>>><***DAN to fill in>
 
@@
 Review of Ghosh et al.: Copulas and the Potential for Ecology
-I would like to thank the authors for their manuscript entitled "Copulas and their potential for ecology".
-The manuscript provides an introduction to the concept of copulas and their utility for application in studying the natures of variable dependency in ecology.
-The authors demonstrate that the most commonly used models of inference, correlation and regression, only encapsulate certain features of the dependency between variables and the author introduce a more complete method of inference to an ecologist audience.
-I must first state that I must give fair warning that, whilst I have heard about copulas before, I have never applied them to my own research and I certainly cannot claim to be an expert in this area.
-I am therefore probably the target audience of this paper.
-This means that whilst I can give constructive feedback on the author's introduction of the concepts to an ecologist readership and can follow the mathematical background that they have provided, I cannot necessarily state definitively whether the authors have reproduced a fair portrayal of the theory of copulas in this manuscript.
-However, despite not being a specialist on the subject, I found the presentation of the manuscript reasonably easy to follow and the authors should be commended for making what is quite a dense subject approachable.
-Each theoretical claim is backed by well-documented supplementary materials which allows for digestion of the key concepts without being side-tracked by dense mathematical derivations.
+I would like to thank the authors for their manuscript entitled "Copulas and their potential 
+for ecology".  The manuscript provides an introduction to the concept of copulas and their 
+utility for application in studying the natures of variable dependency in ecology.  The 
+authors demonstrate that the most commonly used models of inference, correlation and regression, 
+only encapsulate certain features of the dependency between variables and the author introduce a 
+more complete method of inference to an ecologist audience.
+
+I must first state that I must give fair warning that, whilst I have heard about copulas before, 
+I have never applied them to my own research and I certainly cannot claim to be an expert in this 
+area.  I am therefore probably the target audience of this paper.  This means that whilst I can 
+give constructive feedback on the author's introduction of the concepts to an ecologist readership 
+and can follow the mathematical background that they have provided, I cannot necessarily state 
+definitively whether the authors have reproduced a fair portrayal of the theory of copulas in this 
+manuscript.
+
+However, despite not being a specialist on the subject, I found the presentation of the manuscript 
+reasonably easy to follow and the authors should be commended for making what is quite a dense 
+subject approachable.
+
+>>>Many thanks for the positive feedback! We worked hard for clarity.
+
+Each theoretical claim is backed by well-documented supplementary materials which allows for 
+digestion of the key concepts without being side-tracked by dense mathematical derivations.
 
@@
 
-The authors detail a large number of example ecological applications and detail a strong case for the use of copula families with left- or right- tail associations.
-This is a very good demonstration where the use of copulas can benefit the ecological investigator far above the traditional methods.
+The authors detail a large number of example ecological applications and detail a strong case 
+for the use of copula families with left- or right- tail associations.  This is a very good 
+demonstration where the use of copulas can benefit the ecological investigator far above the 
+traditional methods.
 
@@
 
-I would however appreciate a discussion of how copula-based methods compare with regression models with a heteroscedastic error term.
-It seems that these models would also be able to cope with left- or right- tail associations in a bivariate application.
-Given the rise of hierarchical Bayesian models in ecology, before reading this paper, my (maybe somewhat naiive) approach would be to apply a standard regression model but allow for the variance in the error distribution to also be determined by the predictor variable.
+I would however appreciate a discussion of how copula-based methods compare with regression 
+models with a heteroscedastic error term.  It seems that these models would also be able to 
+cope with left- or right- tail associations in a bivariate application.  Given the rise of 
+hierarchical Bayesian models in ecology, before reading this paper, my (maybe somewhat naiive) 
+approach would be to apply a standard regression model but allow for the variance in the error 
+distribution to also be determined by the predictor variable.
 
->>><DAN: to add some discussion material>
+>>>We have added a paragraph to the Discussion on this topic. Whereas certainly 
+one could use the approach of regression models with heteroskedastic error variance to 
+explore tail associations, probably gaining insight, copulas have theoretical 
+and practical advantages sketched in the new paragraph.
 
-In my view this paper represents an impressive contribution to statistical ecology and I am very happy to be introduced to a new and exciting avenue of analysis.
-Indeed, the thing I hate most about the manuscript is that it has spurred me on to spend hundreds of dollars on new academic books on copula modelling.
+In my view this paper represents an impressive contribution to statistical ecology and I am 
+very happy to be introduced to a new and exciting avenue of analysis.  Indeed, the thing I 
+hate most about the manuscript is that it has spurred me on to spend hundreds of dollars on 
+new academic books on copula modelling.
 
@@
 
-What follows is a small set of very minor points that are all highly subjective and whether or not they are addressed does not, in my view, preclude the manuscript of being published:
-o    The authors use the term "distribution function" to represent "cumulative distribution function".
-My feeling is that ecologists in general are more familiar with the term "cumulative distribution function" (which thereafter can be abbreviated to "CDF") as the former is more likely to be mistaken for the density function.
+What follows is a small set of very minor points that are all highly subjective and whether 
+or not they are addressed does not, in my view, preclude the manuscript of being published:
+
+>>>Thank you for giving us some reasonable wiggle room here, which helps in light of 
+other constraints.
+
+The authors use the term "distribution function" to represent "cumulative distribution 
+function".  My feeling is that ecologists in general are more familiar with the term 
+"cumulative distribution function" (which thereafter can be abbreviated to "CDF") as the 
+former is more likely to be mistaken for the density function.
+
+>>>We made the change as requested, thanks for the suggestion.
+
+The manuscript could benefit from a little more detail in the introduction as to how 
+this material goes further than that in other ecological introductions to copulas.  
+For example, Anderson et al. 2018 also contains an introduction to the theory of copulas 
+so if might be worth explicitly pointing out how this paper goes further in explaining 
+and introducing the concept to an ecologist audience.
 
@@
 
-o    The manuscript could benefit from a little more detail in the introduction as to how this material goes further than that in other ecological introductions to copulas.
-For example, Anderson et al.
-2018 also contains an introduction to the theory of copulas so if might be worth explicitly pointing out how this paper goes further in explaining and introducing the concept to an ecologist audience.
-
->>><DAN to do>
-
-o    A table giving a run-down of the notation used in the manuscript might be useful for the more mathematically-adverse reader.
-A sort of notation cheat-sheet for easy notation lookup would be useful for readers so that they can concentrate on the new concepts rather than looking up notation definitions.
+A table giving a run-down of the notation used in the manuscript might be useful for the 
+more mathematically-adverse reader.  A sort of notation cheat-sheet for easy notation 
+lookup would be useful for readers so that they can concentrate on the new concepts rather 
+than looking up notation definitions.
 
@@
 
-o    The authors make reference to many families of copula models (a total of 16 families) and, whilst there is a graphical illustration of the shapes of the copulas, we don't see the functional forms of these copulas.
-A table either in the main text or supplementary materials giving these would be appreciated.
+The authors make reference to many families of copula models (a total of 16 families) and, 
+whilst there is a graphical illustration of the shapes of the copulas, we don't see the 
+functional forms of these copulas.  A table either in the main text or supplementary materials 
+giving these would be appreciated.
 
->>><Shyamolina is collecting the formulas>
+>>>Thank you for this suggestion. Exactly such a table was already provided by Brechmann 
+and Schepsmeier (2013, Journal of Statistical Software), so rather than reproduce it, we
+now include a reference to that work. We also feel like such a table would alienate most
+of the audience we are trying to reach here. In fact we have not found the formulas
+for the copula families to be particularly useful knowledge for standard ecological 
+application (though they will be intimidating to some readers). Rather, it helps to 
+know the properties of the copula families (which we describe) and what they look 
+like (which we show). Of course mathematical applications will need the formulas, but 
+I think most readers will be quantitative ecologists who want to use these families 
+but, in our estimation based on our experience of using the families ourselves, will 
+have almost no occassion to use the formulas.

### 2019-12-18 Paper.Rmd
changed distribution function to cdf

@@
 the empirical cumulative distribution function of the $x_i$ and $y_i$,
-respectively. We henceforth abbreviate "cumulative distribution function" as "distribution
-function" or "cdf". See, for instance,
+respectively. We henceforth abbreviate "cumulative distribution function" as "cdf". 
+See, for instance,
 Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
@@
 which is a short introduction to 
-copulas for ecologists. A bivariate *copula* can be defined as a bivariate distribution
-function with both margins uniform on $(0,1)$ (@joe2014_dependence, p. 7). 
+copulas for ecologists. A bivariate *copula* can be defined as a bivariate cdf 
+with both margins uniform on $(0,1)$ (@joe2014_dependence, p. 7). 
 This will be the 
-distribution function of a bivariate random variable with uniform marginals, and 
+cdf of a bivariate random variable with uniform marginals, and 
 the terminology *copula* is sometimes also applied to this random vector. 
@@
 @sklar1959_theorem (see also @MaiScherer2017, p. 16) 
-says that if $F$ is the bivariate distribution function
+says that if $F$ is the cdf
 of a random vector $(X,Y)$, with margins $F_X$ and $F_Y$, then there 
@@
 The theorem also says $C$ is unique if $F_X$ and $F_Y$ are continuous.
-Thus $C$ couples the bivariate distribution function, $F$, with the 
-distribution functions of the marginals. 
+Thus $C$ couples the bivariate cdf, $F$, with the 
+cdfs of the marginals. 
 Finally, Sklar's theorem says that if $D$ is any bivariate copula and 
-$G_A$ and $G_B$ are univariate distribution functions, then 
-$D(G_A(x),G_B(y))$ is a valid distribution function 
+$G_A$ and $G_B$ are univariate cdfs, then 
+$D(G_A(x),G_B(y))$ is a valid cdf 
 of some bivariate random variable.
@@
 distributions have continuous, strictly 
-monotonic distribution functions, and this case is simpler. So we 
+monotonic cdfs, and this case is simpler. So we 
 henceforth make such an assumption.
@@
 any random vector can be constructed from a unique copula and marginal 
-distribution functions, and, furthermore,
-any copula and any univariate distribution functions give rise to a random vector;
+cdfs, and, furthermore,
+any copula and any univariate cdfs give rise to a random vector;
 so Sklar's theorem provides a correspondence between random vectors
-and pairs consisting of a copula and two univariate distribution functions,
+and pairs consisting of a copula and two univariate cdfs,
 thereby making it possible to study a random vector by studying these two
@@
 <!--***-->
-The univariate distribution functions associated with a random vector 
+The univariate cdfs associated with a random vector 
 $(X,Y)$ are its margins, $F_X$ and $F_Y$, and the 
@@
 Conversely, given a copula
-$D$ and univariate distribution functions $G_A$ and $G_B$,
+$D$ and univariate cdfs $G_A$ and $G_B$,
 Sklar's theorem guarantees that $D(G_A(x),G_B(y))$ is a 
-distribution function, so the random variable with this 
-distribution function is the random variable corresponding
+cdf, so the random variable with this 
+cdf is the random variable corresponding
 to $D$, $G_A$ and $G_B$ under Sklar's correspondence. 
@@
 where $G_A^{-1}$ and $G_B^{-1}$ are the inverses of $G_A$ and $G_B$
-and $(U,V)$ is the random variable with distribution function $D$
+and $(U,V)$ is the random variable with cdf $D$
 (which has uniform marginals).
-<!--***-->
 If we conflate a copula with the random vector of 
-which the copula is the distribution function, then Sklar's correspondence
+which the copula is the cdf, then Sklar's correspondence
 is between random variables, $(U,V)$, with uniform marginals 
@@
 marginals. The correspondence is simply via the application of the
-univariate distribution functions, or their inverses: $U=F_X(X)$,
+univariate cdfs, or their inverses: $U=F_X(X)$,
 $V=F_Y(Y)$, $X=F^{-1}_X(U)$, $Y=F^{-1}_Y(V)$.
-<!--***-->
 Application of the 
-distribution functions of the component random variables translates
+cdfs of the component random variables translates
 between the uniform-marginal context, $(U,V)$, which isolates the 
@@
 cummulative dstribution function (cdf) of a (univariate) standard 
-normal and $\Phi_{2,p}$ is the joint cdf of a
+normal and $\Phi_{2,p}$ is the cdf of a
 bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
-and $p$ in the off diagonal positions. However, formulas for copulas are often not as 
+and $p$ in the off diagonal positions. Formulas for the normal copula and all the 
+other copula families we use were provided by @Brechmann2013. 
+However, formulas for copulas are often not as 
 directly informative about copula properties as probability density function (pdf) 
@@
 BB1 copulas have stronger left- than right-tail association, and for others the reverse
-(Fig. \ref{figped_BB1}). See @nelsen2006_copula and @joe2014_dependence for further
+(Fig. \ref{figped_BB1}). See @nelsen2006_copula, @joe2014_dependence and 
+@Brechmann2013 for further
 information on these and other copula families.
@@
 similar to copulas. We here explain the connection. As stated above, the copula associated
-with a random variable $(X,Y)$ is the joint distribution function of the random variable
+with a random variable $(X,Y)$ is the cdf of the random variable
 $(F_X(X),F_Y(Y))$. Given a sample $(x_i,y_i)$, $i=1,\ldots,n$, from $(X,Y)$, let $\hat{F}_X$ and
@@
 and, additonally, these copulas have been 
-described in more detail elsewhere [@Joe1997]; so 
+described in more detail elsewhere [@Joe1997; @Brechmann2013]; so 
 we only briefly identify each family, say a few words about tail dependence 
@@
 180 degrees. 
-See @Joe1997 for details on all these families. We used the implementations
+See @Joe1997 and @Brechmann2013 for details on all these families. We used the implementations
 of these families provided in the `VineCopula` package for the R programming
@@
 Table \ref{tab_copula_info} for a summary of some properties of each family.
-<!--***DAN: Shyamolina, please make a table, and replace the X above with 
-a citation to that table. Please place the table chunk in such a way that 
-this citation becomes table 2, since prior to this point the only main text
-table that has been cited is table 1, and we want the table citations to 
-come in order. The columns of the table should include: 1) copula family name,
-2) number of parameters for the family; 3) info on tail dependence that is
-in the above para; 4) a reference to the figure (in the main text or sup mat)
-where there is a picture of this family. Anything else? There should be a 
-row for each of our 16 copula families.-->
 
@@
 simulations were plotted on the same axes for each panel, but two example simulations are 
-shown in solid and void black circles to help assess to what extent variation on the plots was between 
+shown in solid and hollow black circles to help assess to what extent variation on the plots was between 
 simulations or within simulations. For each simulation, we tested the linearity 
@@
 
+<!--referee-requested para comparing to regression with heteroskedastic error term-->
+Non-copula methods may be useful under some circumstances for exploring
+asymmetries of tail association, but copula approaches have theoretical
+and practical advantages. It was suggested to us that asymmetries of
+tail association can also be explored using, for instance, a regression 
+model with heteroskedastic error term, $y=mx+b+\epsilon(x)$. Whereas this 
+approach seems capable of yielding insight for some datasets, it implicitly assumes
+a particular causal relationship between the variable, an untenable assumption for our
+synchrony applications and many other applications. Furthermore, no equivalent
+of Sklar's theorem exists under this approach, so information on marginal 
+distributions cannot be separated from association information. Sklar's theorem
+means that the copula associated with two variables contains all the information
+about the association. Such statements of 
+mathematical completeness will not be avilable for others methods.
+
 <!--para on data requirements of the methods, requested by the AER referees-->
@@
 distributions have continuous, strictly 
-monotonic distribution functions. This was for simplicity and because the simpler
+monotonic cdfs. This was for simplicity and because the simpler
 approach was sufficient for our research questions. But count data in ecology have been analyzed

### 2019-12-18 SupportingInformation.Rmd
changed distribution function to cdf

@@
 
-Let $F$ be the bivariate distribution function
+Let $F$ be the cdf
 of a random vector $(X,Y)$, with margins $F_X$ and $F_Y$ assumed to be 
@@
 $(X,Y)$, which exists and is unique by Sklar's theorem, is, in fact, 
-the distribution function
+the bivariate cdf
 of the random variable $(U,V)=(F_X(X),F_Y(Y))$. Let $C$ denote the distribution
@@
 for $F_X$ and $F_Y$ is necessary, however: without it the random variables 
-$U$ and $V$ are not uniformly distributed, so their joint distribution function 
+$U$ and $V$ are not uniformly distributed, so their joint cdf 
 is not even a copula.
@@
 Now let $D$ be any bivariate copula and let 
-$G_A$ and $G_B$ be univariate distribution functions, so that 
-$D(G_A(x),G_B(y))$ is guaranteed by Sklar's theorem to be a distribution function.
+$G_A$ and $G_B$ be univariate cdfs, so that 
+$D(G_A(x),G_B(y))$ is guaranteed by Sklar's theorem to be a cdf.
 We here show that if $G_A$ and $G_B$ are continuous and strictly monotonic, then
-$D(G_A(x),G_B(y))$ is the distribution function of the random variable 
+$D(G_A(x),G_B(y))$ is the cdf of the random variable 
 $(G_A^{-1}(U),G_B^{-1}(V))$, where $(U,V)$ now denotes the random
@@
 The inverses exist because of the strict monotonicity assumption.
-To get to the desired result, note that the distribution function of 
+To get to the desired result, note that the cdf of 
 $(G_A^{-1}(U),G_B^{-1}(V))$ is $P[G_A^{-1}(U) \leq x, G_B^{-1}(V) \leq y]$,
 which equals $P[U \leq G_A(x), V \leq G_B(y)]$, by strict monotonicity.
-But this is $D(G_A(x),G_B(y))$ because $D$ is the distribution function of 
+But this is $D(G_A(x),G_B(y))$ because $D$ is the cdf of 
 $(U,V)$. In fact, a stronger statement assuming

### 2019-12-18 Paper.Rmd
further progress

@@
 [@Valpine2014; @Anderson2018; @Popovic2018], but such usage is
-still rare. Our results suggest that environmental, 
+still rare. Whereas @Anderson2018 provide a specific methodological pathway, based partly 
+on copulas, for the important problem of the 
+analysis of multivariate ecological population count data;
+we instead advocate more broadly for the application of copulas across ecology.
+Our results suggest that environmental, 
 ecological and evolutionary processes 
@@
 and practical advantages. It was suggested to us that asymmetries of
-tail association can also be explored using, for instance, a regression 
+tail association can also be explored using a regression 
 model with heteroskedastic error term, $y=mx+b+\epsilon(x)$. Whereas this 
 approach seems capable of yielding insight for some datasets, it implicitly assumes
-a particular causal relationship between the variable, an untenable assumption for our
-synchrony applications and many other applications. Furthermore, no equivalent
+a particular causal relationship between $x$ and $y$, an untenable assumption for 
+many applications. Furthermore, no equivalent
 of Sklar's theorem exists under this approach, so information on marginal 
-distributions cannot be separated from association information. Sklar's theorem
-means that the copula associated with two variables contains all the information
+distributions cannot be separated from association information, and the two types
+of information will both influence regression results. Sklar's theorem
+means that the copula associated with two variables contains all and only the information
 about the association. Such statements of 
-mathematical completeness will not be avilable for others methods.
+mathematical completeness will not be avilable for other methods.
 
@@
 
-We thank Joel E. Cohen for making us aware of copulas and their potential ecological utility in the 
+We thank Joel E. Cohen for making us aware of copulas in the 
 first place. We thank the many contributors to the large datasets 
@@
 under the Core Capability Grant BBS/E/C/000J0200. SG, LWS and DCR were partly funded by US National 
-Science Foundation grants 1714195 and 1442595 and the James S McDonnell Foundation.
+Science Foundation grants 1714195 and 1442595, the James S McDonnell Foundation, and the 
+California Delta Science Program.
+.
 

### 2019-12-18 AER_Responses.tex
further progress
almost done addressing referee comments, except for possible need for a notation table

@@
 
->>><***DAN to fill in>
+>>>Thank you for this suggestion, which is a very good one. We did as requested and find 
+the results sections seem more readable now to a less technicaly oriented audience.
 
@@
 
->>><DAN to do>
+>>>We added the sentence "Whereas Anderson et al (2018) provide a specific methodological 
+pathway, based partly on copulas, for the important problem of theanalysis of multivariate 
+ecological population count data; we instead advocate more broadly for the application of 
+copulas across ecology."
 

### 2019-12-18 Paper.Rmd
almost done addressing referee comments, except for possible need for a notation table

@@
 normal copula had a much higher AIC, `r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)`, answering the 
-first part of Q1 for these data: these data have a dependence structure markedly distinct 
+first part of Q1 for these data: the data have a dependence structure markedly distinct 
 from that of a normal copula. Recall that AIC differences of $2$
@@
 These values are distinct from what a normal copula would give, namely, 0, helping to answer 
-the second part 
+the second part of
 Q1 for these data. The values also differed substantially 
@@
 and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).
-
-Our model selection procedures do not reveal whether tail dependence parameters are 
+Apparently C and N are more strongly associated with each other in soils
+for which they are both plentiful than in soils for which they are 
+both less abundant, and C and N are more strongly associated in both
+tails than would be modelled using a normal copula model or methods that
+assume such a model.
+
+However, our model selection procedures do not reveal whether tail dependence parameters are 
 *significantly* different from $0$ and from each other. 
@@
 separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
-(Fig. \ref{fig_soilCN_nonparam}). Results showed that tail associations of data are stronger in 
+(Fig. \ref{fig_soilCN_nonparam}). Results confirmed that tail associations of data are stronger in 
 both the lower and upper tails than would be expected under a null hypothesis of a normal copula (Fig. \ref{fig_soilCN_nonparam}B-F).
@@
 on surrogates, separately in two 
-comparisons (Table \ref{tab_soilCN_asym}). Results showed that
+comparisons (Table \ref{tab_soilCN_asym}). Results confirmed that
 upper-tail association was significantly stronger than lower-tail association; i.e., C and N values are
@@
 The first three datasets had stronger upper- 
-than lower-tail dependence, and the last dataset had the reverse, though non-significantly. Our 
-generally affirmative (except for Cedar Creek), 
+than lower-tail dependence, and the last dataset had the reverse, though non-significantly. 
+Apparently species body masses and metabolic rates, for both birds and mammals, are more strongly 
+associated with each other when both these quantities are large than when they are 
+both small. Our generally affirmative (except for Cedar Creek), 
 empirically based answer to Q1 for the bivariate datasets is 
@@
 (Table \ref{tab_multivar_summary}, rows 6-8), whereas leaf-curling plum aphid first flight
-data showed the reverse. All of the values in Table \ref{tab_multivar_summary}, rows 8 deviated highly significantly
+data showed the reverse. 
+Thus the spatial synchrony of rarity was stronger than that of outbreaks for 
+green spruce aphid and \textit{Ceratium furca} abundance, whereas leaf-curling aphid
+first flights were more spatially related to each other when these flights happened
+late than when they happened early.
+All of the values in Table \ref{tab_multivar_summary}, rows 8 deviated highly significantly
 from what would have been expected under a null hypothesis 
@@
 responsible for non-normal copula structure and
-asymmetric tail associations across space in ecological variables.
+asymmetric tail associations across space in ecological variables:
+when populations or other ecological quantities are influenced principally by
+low (respectively, high) values of an environmental variable, one may expect 
+left-tail (respectively, right-tail) associations across space for the ecological variable. 
 This result is represented in Fig. \ref{MasterFigure} as the solid box around 
@@
 in an ecological variable can be inherited from an environmental 
-variable was found to be reasonable, because it held for the models we examined; 
+variable was found to be reasonable, because it held for the models we examined -
+the Moran effect seems to extend to copula structure and tail associations - 
 however, similarities between environmental- and 
@@
 Since many ecological models are nonlinear, this result provides the 
-reasonable expectation that there may be a Moran-effect-like correspondence 
+reasonable expectation for a Moran-effect-like correspondence 
 between noise and model-output dependence structure across space 
@@
 structure and asymmetric tail associations are important for spatially averaged 
-quantities. This result is represented in Fig. \ref{MasterFigure}
+quantities and can influence the temporal stability of those quantities. 
+This result is represented in Fig. \ref{MasterFigure}
 as the solid box around "Instability/skewness of mean or total 

### 2019-12-18 Paper.Rmd
working on final trim

@@
 
-<!--***DAN: remianing to do:
--->
-
 \noindent \textbf{Abstract}
@@
 standard approaches to studying such relationships, based on correlation and regression, 
-provide only a small slice of the complex information contained in the relationships. Other 
+provide only some of the complex information contained in the relationships. Other 
 statistical approaches exist that provide a complete description of relationships between variables, 
-based on the concept of the *copula*; they are applied in finance, neuroscience and other fields, 
-but rarely in ecology. We here explore the concepts that underpin copulas and examine the potential 
+based on the concept of the *copula*; they are applied in finance, neuroscience and elsewhere, 
+but rarely in ecology. We explore the concepts that underpin copulas and the potential 
 for those concepts to improve our understanding of ecology. We find that informative copula 
@@
 strongly related in their left compared to right tails, or *vice versa*. We describe mechanisms by which 
-observed copula structure and asymmetric tail associations can arise in ecological data, including a Moran-like 
-effect whereby dependence structures between environmental variables are inherited by ecological 
-variables; and asymmetric or nonlinear influences of environments on ecological variables, such as 
+observed copula structure and tail associations can arise in ecological data, including a Moran-like 
+effect whereby dependence structures are inherited from environmental variables; and asymmetric or nonlinear influences of environments on ecological variables, such as 
 under Liebig's law of the minimum. We also describe consequences of copula structure for 
 ecological phenomena, including impacts 
-on extinction risk, Taylor's law, and the stability through time of ecosystem services. By documenting the 
+on extinction risk, Taylor's law, and the temporal stability of ecosystem services. By documenting the 
 importance of a complete description of dependence between variables, advancing conceptual frameworks, and 
-demonstrating a powerful approach, we aim to encourage widespread use of copulas in ecology, which we
+demonstrating a powerful approach, we encourage widespread use of copulas in ecology, which we
 believe can benefit the discipline.
@@
 such relationships are limited, and 
-provide only a partial description of a relationship between variables. 
-For instance, datasets showing markedly different 
+provide only a partial description the relationship. 
+For instance, datasets showing different 
 relationships may have the same correlation coefficient
@@
 related in the left (respectively, right) portions of their distributions, 
-an asymmetric pattern of dependence that can
+an asymmetric pattern of association that can
 have ecological significance, as discussed below, but that is not captured by 
@@
 applied widely in other fields, 
-that provide a description of the dependence 
-between two or more variables that can be mathematically proven 
-to be complete [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. 
+that provide a description of the association 
+between two or more variables that is mathematically provably 
+complete [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. 
 The approaches are based on 
-the concept of the *copula*. Copula approaches separate the information of a bivariate 
-random variable into two distinct parts: the information in the 
-marginal distributions (which says nothing about the dependence between the 
+the concept of the *copula*. Copula approaches separate information in a bivariate 
+random variable into two parts: the information in the 
+marginal distributions (which says nothing about the association between the 
 variables), and the remaining information (which is solely about
-dependence). We here introduce some simple concepts, based on taking ranks,
+the association). We here introduce some simple concepts, based on ranks,
 that relate to copulas and give a conceptual 
 flavor of copula ideas to be introduced 
-more formally in the "Background on copulas" section below (section \ref{Background}). Given 
+formally in the "Background" section below (section \ref{Background}). Given 
 a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
-Fig. \ref{fig_cop_pedag1}A-B), information about the structure of the dependence
+Fig. \ref{fig_cop_pedag1}A-B), information about the structure of the association
 between $x$ and $y$ can be separated from information contained 
 in the marginal distributions by 
-considering the bivariate plot of $u_i$ versus $v_i$, where
+considering the plot of $u_i$ versus $v_i$, where
 $u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
@@
 Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
-A-B, respectively. Ranking makes the marginal distributions of 
+A-B. Ranking makes the marginal distributions of 
 the component datasets uniform, isolating the 
@@
 separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D
-relate to copulas in a way to be described in the "Background on copulas" section below
-(section \ref{Background}), after a definition of copulas
-is provided. Note that Pearson correlation, even though it is the most commonly
-used measure of relationships between variables, is modified by monotonic
+relate to copulas in a way to be described 
+(section \ref{Background}). Note that Pearson correlation, even though it is the most commonly
+used measure of association, is modified by monotonic
 tranformation of the component variables (Appendix 
@@
 Spearman and Kendall correlations, being rank-based, 
-are not influenced by monotonic transformations of the component variables. 
+are not influenced by monotonic transformations of the variables. 
 They provide information solely on the dependence between those variables.
@@
 that it can detect associations in the tails of distributions, and asymmetries of such associations. 
-Tail association (introduced formally in the "Background on copulas" section, below, section 
-\ref{Background}) 
-is association between extreme values of variables. 
+Tail association (introduced formally in section 
+\ref{Background}) is association between extreme values of variables. 
 If smaller values of two positively associated variables are more 
@@
 (Fig. \ref{fig_cop_pedag1}C);
-and *vice versa* if larger values are more strongly associated than smaller
-ones (Fig. \ref{fig_cop_pedag1}D). Datasets of the same 
+and *vice versa* if larger values are more strongly associated 
+(Fig. \ref{fig_cop_pedag1}D). Datasets of the same 
 Spearman or Kendall correlation can have a range of tail associations 
@@
 complex dependence structure of data, we have found tail associations to
-usefully describe the full structure of dependence, so tail association 
-is one of the main foci of this paper. We give four examples below which
-illustrate some of the potential ecological meaning of tail associations.
+be a useful component of that structure, so tail association 
+is a main focus of this paper. We give four examples below which
+illustrate some of the ecological meaning of tail associations.
 
@@
 applications of copulas in ecology and to 
-help determine to what extent ecological understanding may benefit generally from
-using copulas. We provide a basic introduction
-to copula concepts in the "Background on copulas" section below (section \ref{Background}). 
+estimate to what extent ecological understanding may benefit from
+using copulas. We provide an introduction
+to copula concepts in section \ref{Background}. 
 Readers can find additional abbreviated [@Anderson2018; @Genest2007] 
 and comprehensive [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]
-introductions elsewhere. Copulas have already been used very effectively 
+introductions elsewhere. Copulas have already been used effectively 
 in a few studies to illuminate ecological phenomena 
 [@Valpine2014; @Anderson2018; @Popovic2018], but such usage is
-still rare. Whereas @Anderson2018 provide a specific methodological pathway, based partly 
-on copulas, for the important problem of the 
-analysis of multivariate ecological population count data;
+still rare. Whereas @Anderson2018 provide a specific method, based
+on copulas, for important problems of the 
+analysis of multivariate population count data,
 we instead advocate more broadly for the application of copulas across ecology.
@@
 ecological and evolutionary processes 
-may commonly generate and transmit complex dependence structures, including asymmetric tail 
-associations, and that greater use of copulas can illuminate the underlying processes.
-We believe copula approaches are among the tools all ecologists, including tropical
-ecologists, should be considering for analysis of their data in the 21st century.
+may commonly generate complex dependence structures, including asymmetric tail 
+associations, and that greater use of copulas can illuminate underlying processes.
+We believe copula approaches are among the tools all ecologists 
+should be considering for analysis of their data in the 21st century.
 
@@
 \end{figure}
+
+<!--***DAN: got to here with final trim-->
 

### 2019-12-18 Paper.Rmd
continuing to trim words here and there

@@
 
-<!--***DAN: got to here with final trim-->
-
-One way we may expect, *a priori*, a study of tail associations and of the 
-complex copula dependence structures between
-ecological variables
+One way we may expect, *a priori*, a study of tail associations and 
+copula dependence structures 
 to better reflect ecological relationships relates to Liebig's law of the
 minimum. Liebig's law is 
-the idea that growth is controlled not by total resources but by the resources which
-are scarcest relative to organism needs. 
+the idea that growth is controlled not by total resources but by the resource which
+is scarcest relative to organism needs. 
 If, for instance, the growth of a plant depends on soil 
-nitrogen, N, and other factors, a plot of plant growth rates versus soil N may 
+nitrogen, N, and other factors, a plot of growth rates versus soil N may 
 look more like Fig. \ref{fig_cop_pedag1}A or C than like 
 Fig. \ref{fig_cop_pedag1}B or D, i.e., the two variables may show 
-left-tail association. Nitrogen 
-will control plant growth, producing a clear
-relationship, only when it is limiting, in the left portions of the distributions. 
-The portions of the distributions that dominate the association are visible in a rank-based
-plot, but may not be revealed by standard correlation techniques.
-
-A second reason why we may expect understanding of complex dependence structures
+left-tail association: N 
+controls plant growth, producing a clear
+relationship, only when it is limiting. 
+This aspect of the association would be visible in a rank-based
+plot, but may not be revealed by correlation.
+
+A second reason why we may expect detailed understanding of dependence structures
 to benefit ecological
-research is that prior work has demonstrated, using copulas, complex 
+research is that prior work demonstrated, using copulas, complex 
 stucture in the spatial dependence of environmental 
 variables [@Serinaldi2008; @She2018; @Goswami2018;
-@li2013]. One may expect an environmental variable 
-measured through time in two locations to show strong tail associations between
+@li2013]. An environmental variable 
+measured through time in two locations may show strong tail associations between
 the locations if
-intense meteorological events are also widespread in their effects, as seems 
+intense meteorological events are also widespread, as seems 
 frequently to be the case: extreme values are associated with intense events,
@@
 in fluctuations of populations or other ecosystem variables 
-(this is called \emph{spatial synchrony}) which are influenced by the environmental 
-variable. This influence is called the \emph{Moran effect}. 
-If asymmetric tail associations, or other complex dependence structure, is transmitted, 
-possibly in modified form, 
-from environmental to ecological variables, then we may expect complex dependence 
+(this is called \emph{spatial synchrony}) influenced by the environmental 
+variable. This is called the \emph{Moran effect}. 
+If asymmetric tail associations, or other complex dependence structure, is transmitted 
+from environmental to ecological variables, then we would expect complex dependence 
 structure and tail associations to be a common feature of the 
@@
 biogeochemical and other environmentally influenced ecological variables.
-Synchrony is a phenomenon of substantial interest in 
+Synchrony attracts major interest in 
 ecology [@Liebhold2004; @Sheppard2016; @walteretal2017].
@@
 dependence structures and tail associations to benefit 
-ecological research is that such an approach may help illuminate 
-the causal mechanisms between variables. For instance, if two species
-Sp1 and Sp2 are strong competitors, abundances of the two species
+ecology is that such an approach may help illuminate 
+causal mechanisms between variables. If two species,
+Sp1 and Sp2, are strong competitors, abundances of the two species
 across quadrats should be negatively related, as in 
@@
 limiting factors other than Sp1. 
-Alternatively, Fig. \ref{fig_cop_pedag1}F may suggest that Sp2
+Alternatively, Fig. \ref{fig_cop_pedag1}F may suggest Sp2
 is the dominant competitor. Other causal hypotheses may produce 
@@
 complete information on causal pathways from analyses which are
-fundamentally based on associations. However, more detailed 
-examination of the tail associations and potentially other 
-patterns of dependence of variables may
-well be combined with biological and other reasoning to rule out
-some causal hypotheses which could not be eliminated solely 
-with standard correlation approaches. @Popovic2018 has previously
-used copulas to help illuminate causal aspects of relationships
+fundamentally based on associations. However, 
+examination of tail associations may
+be combined with biological reasoning to rule out
+some causal hypotheses which could not previously be eliminated. 
+@Popovic2018 used copulas to illuminate causal relationships
 between species.
 
-A fourth reason why we may expect an understanding of tail associations and of complex 
+A fourth reason why we may expect an understanding of tail associations and complex 
 dependence structures to be useful
@@
 and the total economic value of a fishery is the sum of local catches. 
-We will explore how details, which can be captured using copulas and which
-relate to tail associations, 
-of the dependence between fluctuations of local
+We will explore how tail associations between fluctuations of local
 quantities can influence fluctuations of 
@@
 higher organizational levels in ecology and human concerns.
-To illustrate this idea we cite @li2013, who have demonstrated that the overall 
+To illustrate this idea we cite @li2013, who demonstrated that the overall 
 reliability of wind-generated electricity depends sensitively on details
@@
 Spatially aggregated ecological variables may be subject to a similar 
-effect, with
-possible consequences. For instance, if populations of a 
+effect. For instance, if populations of a 
 pest species in different locations are all positively associated and are also more strongly
@@
 to occur together, creating regional epidemics. 
-On the other hand, stronger left-tail associations in a pest species, even if overall 
-correlation were unchanged, would have more benign effects.
+Stronger left-tail associations in a pest species, even if overall 
+correlation were the same, would have more benign effects.
 
@@
 sub-fields and organizational levels of ecology. 
-We address Q2 (sections \ref{Causes} through \ref{Results_Q2} of the paper), using 
-simple models. We address Q3 (sections \ref{Consequences} through \ref{Results_Q3} 
-of the paper) using 
+We address Q2 (sections \ref{Causes} through \ref{Results_Q2}), using 
+simple models. We address Q3 (sections \ref{Consequences} through \ref{Results_Q3}) using 
 both data and models. Section \ref{Discussion} is the Discussion.
@@
 variables [@Serinaldi2008; @She2018; @Goswami2018; @li2013], and finance [@Li2000],
-and have also been used very effectively, but rarely so far, in ecology [@Valpine2014; 
+and have also been used effectively, but rarely so far, in ecology [@Valpine2014; 
 @Anderson2018; @Popovic2018]. 
-We argue that benefits of wider usage of complete 
+We argue that benefits of wider use of
 copula descriptions of dependence in ecology will be substantial.
@@
 Again making Sklar's correspondence
-more concrete, we show in Appendix \ref{SM-ContStrictMonoton}
+concrete, we show in Appendix \ref{SM-ContStrictMonoton}
 that this random variable is $(G_A^{-1}(U),G_B^{-1}(V))$, 
@@
 $V=F_Y(Y)$, $X=F^{-1}_X(U)$, $Y=F^{-1}_Y(V)$.
-Application of the 
-cdfs of the component random variables translates
-between the uniform-marginal context, $(U,V)$, which isolates the 
-dependence information, and the context of the original random variable,
-$(X,Y)$.
 
@@
 bivariate distributions in two separate steps: by specifying the marginal 
-distributions and by specifying the dependence structure. For instance,
-construction of several bivariate distributions is carried out in this way in
+distributions and by specifying the dependence structure.
+Construction of several bivariate distributions is carried out in this way in
 Fig. \ref{PedagogFigCopThry}. The contrast between Fig. \ref{PedagogFigCopThry}C and D
@@
 
-We introduce some standard copula families, as examples of copulas and also
-because we will fit these families to data later. The bivariate normal
+We introduce some standard copula families, as examples, and also
+because we will fit these families to data. The bivariate normal
 copula family, which consists of copulas of bivariate normal
-distributions, is a one-parameter family. The parameter, which we here call $p$,
+distributions, is a one-parameter family. The parameter, $p$,
 corresponds to the correlation
@@
 copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
-cummulative dstribution function (cdf) of a (univariate) standard 
+cdf of a (univariate) standard 
 normal and $\Phi_{2,p}$ is the cdf of a
 bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
-and $p$ in the off diagonal positions. Formulas for the normal copula and all the 
-other copula families we use were provided by @Brechmann2013. 
+and $p$ in the off diagonal positions. Formulas for all the 
+copula families we use were provided by @Brechmann2013. 
 However, formulas for copulas are often not as 
@@
 The Clayton copula family,
-of which we have already pictured an example (Fig.
+of which we already pictured an example (Fig.
 \ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
@@
 less direct intuition than do example pdfs (Fig. \ref{figped_C}). Higher values of the parameter,
-$p$ produce copulas from the Clayton family with higher Kendall or Spearman correlation. The survival
+$p$ produce copulas with higher Kendall or Spearman correlation. The survival
 Clayton family is the symmetric opposite of the Clayton family, showing
@@
 @Brechmann2013 for further
-information on these and other copula families.
+information on these and other families.
 
 Having introduced tail association conceptually and referred to it
-in examples, we now need a more precise definition of a measure of tail association.
+in examples, we now need a precise definition of a measure of tail association.
 We use the term *tail association* to describe the general idea of the strength of
-associations between two variables in the tails of their distributions. One of the
-ways this is measured is called *tail dependence* (@joe2014_dependence, section 2.13; 
-@nelsen2006_copula, section 5.4), which we here define. Given a random vector $(X,Y)$
+association between two variables in the tails of their distributions. One 
+way this is measured is called *tail dependence* (@joe2014_dependence, section 2.13; 
+@nelsen2006_copula, section 5.4), defined here. Given a random vector $(X,Y)$
 with margins $F_X$, $F_Y$ and defining $U=F_X(X)$, $V=F_Y(Y)$, the upper-tail
@@
 $\lim_{u \rightarrow 1^-} P[U>u, V>u]/P[V>u]=\lim_{u \rightarrow 1^-} P[U>u, V>u]/(1-u)$,
-which shows upper-tail dependence is defined symmetrically in the two random variables.
+which shows upper-tail dependence is defined symmetrically in the two variables.
 All the variables we consider were positively associated when they were significantly
 associated, so we think of positively associated variables when conceptualizing the
-definitions presented here. Lower tail dependence is defined analogously, as
+definitions here. Lower tail dependence is defined analogously, as
 $\lambda_L = \lim_{u \rightarrow 0^+} P[Y \leq F_Y^{-1}(u) | X \leq F_X^{-1}(u)]$.
-Tail dependence is a property of the random variable $(U,V)$, so depends only
+Tail dependence is a property of $(U,V)$, so depends only
 on the copula and not on the marginals combined with it. It can be shown 
@@
 ways tail dependence can be applied to data, by first fitting copulas to the data. We will also 
-introduce nonparametric measures of tail association that apply directly to data.
+introduce nonparametric measures of tail association.
 
 Plots using normalized ranks were proposed in the Introduction as being conceptually
-similar to copulas. We here explain the connection. As stated above, the copula associated
+similar to copulas; we explain the connection. As stated above, the copula associated
 with a random variable $(X,Y)$ is the cdf of the random variable
@@
 The datasets we used included environmental, species-trait, phenological, 
-population, community, and ecosystem functioning data (Table \ref{tab_data_info}), selected to span multiple fields and levels
-of organization within ecology. Copula structure describing 
-the relationship between atmospheric weather variables such as rainfall or wind speed, 
+population, community, and ecosystem functioning data (Table \ref{tab_data_info}), 
+selected to span multiple fields and levels
+of organization within ecology. Copula structure 
+between atmospheric weather variables such as rainfall or wind speed, 
 as measured in multiple locations through time, has been 
@@
 nitrogen (megagrams C or N per hectare of soil surface) from `r nrow(d_soilCN)` 
-locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for sampling locations).
+locations across the coterminous United States (Fig. \ref{SM-fig_SoilCNmapUSA} for locations).
 Species-trait data were average species basal metabolic rate (BMR, KJ per hour) and 
@@
 These data have been much studied, but to our knowledge the copula structure 
-of the dependence between metabolic rate and body mass has not been examined. 
+of the association has not been examined. 
 Species-trait data such as these reflect the coevolution of the two traits.
@@
 56 years 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells.
-These data were computed from the Continuous Plankton Recorder survey hosted by the 
+These data were computed from the Continuous Plankton Recorder survey of the 
 UK Marine Biological Survey.
@@
 Our phenology, population, and ecosystem functioning datasets are 
-"multivariate" in that they comprise, for each taxon, measurements through time at multiple locations; 
-copula structure can be studied for each pair of locations.
+"multivariate" in that they comprise, for each taxon, measurements through 
+time at multiple locations; 
+copula structure was studied for each location pair.
 
@@
 Recall that Q1 has three parts: Do datasets
-in ecology have dependence structure distinct from that of a multivariate 
-Gaussian/normal
-distribution (here called non-normal copula structure)? Do positively associated 
+in ecology have non-normal copula structure? Do positively associated 
 ecological variables show tail 
@@
 Information Criteria (AIC and BIC). One of the fitted copulas was the normal
-copula, making possible comparisons of the degree to which the normal copula versus
+copula, making possible comparisons of the degree to which the normal versus
 other copula families were good descriptions of data.
@@
 Second, we tested the independence of the $u_i$ and $v_i$
-using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
+using the statistic $\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
 is Kendall's tau for the data. @Genest2007 argue that this statistic is approximately standard normally
 distributed. We used the 
-implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
-in the R programming language. 
-We tested for independence because our model selection 
+implementation of this test in `BiCopIndTest` in the `VineCopula` package
+in R. We tested for independence because our model selection 
 algorithms were ineffective if data could not 
-be distinguished from independent data 
-since many of the copula families we considered include the independent copula. 
-If independence could be rejected ($0.05$ significance level), model 
-selection proceeded. Third, we fit 16 bivariate copula families (see below for 
-information on the families used) to the 
+be distinguished from independent data, 
+since many copula families include the independent copula. 
+If independence could be rejected ($0.05$ level), model 
+selection proceeded. Third, we fit 16 bivariate copula families (see below) to the 
 normalized ranks via maximum likelihood. The approach of fitting
-copula families to normalized ranks was studied in depth
-and recommended by @Genest1995 and @Shih1995. Their estimator of copula
-parameters, which we use, was shown to be consistent, asymptotically normal,
+copula families to normalized ranks was recommended 
+by @Genest1995 and @Shih1995. Their estimator of copula
+parameters, which we use, is consistent, asymptotically normal,
 and fully efficient at indepencence [@Genest1995]. @Genest2007
@@
 Thus for each bivariate dataset, 
-the end products of the procedure just described were threefold, corresponding to the
+the end products of our procedure were threefold, corresponding to the
 three parts of Q1 listed in the Introduction: A) the AIC, BIC, $\text{AIC}_{\text{w}}$ and 
@@
 dependence measures for each fitted copula and model averages of those 
-quantities, providing information on whether, and to what extent, tail
+quantities, providing information on whether tail
 dependence was asymmetric.
@@
 we tested the goodness of fit of our AIC-best copula family using a 
-bootstrapping procedure developed and studied by @Wang2000 and @Genest2006
-and implemented as `BiCopGofTest` of `VineCopula`. The procedure performed 
+bootstrapping procedure of @Wang2000 and @Genest2006,
+implemented as `BiCopGofTest` in `VineCopula`. The procedure performed 
 one test based on a Cramer-von Mises statistic and another based on a 
 Kolmogorov-Smirnov statistic. To keep computation times reasonable, 
-a run using $100$ bootstraps was performed; if
-the $p$-value from either test was less than $0.2$, tests were re-run with 
+a run using $100$ bootstraps was performed. If
+the $p$-value from either test was $<0.2$, tests were re-run with 
 $1000$ boostraps.
@@
 lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
-of using a large collection of families was solely to include a 
+of using a large collection of families was to include a 
 variety of alternative dependence structures to have a robust model
@@
 and, additonally, these copulas have been 
-described in more detail elsewhere [@Joe1997; @Brechmann2013]; so 
-we only briefly identify each family, say a few words about tail dependence 
-of the families, and provide pictorial descriptions (Figs 
+described in detail elsewhere [@Joe1997; @Brechmann2013]. So 
+we only briefly identify each family, say a few words about its tail 
+dependence, and provide pictorial descriptions (Figs 
 \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8}). 
@@
 $v=-u+1$. We also used several families that can show both upper- and
-lower-tail dependence, in relative amounts that depend on parameter
+lower-tail dependence, in relative amounts depending on parameter
 values: the BB1, survival BB1, BB7 and survival BB7 families. 
@@
 180 degrees. 
-See @Joe1997 and @Brechmann2013 for details on all these families. We used the implementations
-of these families provided in the `VineCopula` package for the R programming
+See @Joe1997 and @Brechmann2013 for details on all families. We used the implementations
+of provided in the `VineCopula` package for the R programming
 language. See Figs \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
@@
 pdfs and tail dependence are influenced by the parameters. See 
-Table \ref{tab_copula_info} for a summary of some properties of each family.
+Table \ref{tab_copula_info} for a summary.
 
@@
 related in any part of their distributions. We here describe the 
-statistics, with some additional details in Appendix \ref{SM-nonparam_stats}.
+statistics, with additional details in Appendix \ref{SM-nonparam_stats}.
 The statistics are defined with positively associated
@@
 We also defined a statistic $\Ps_{l_b,u_b}$ (Appendix \ref{SM-nonparam_stats}), 
-which has a similar qualitative
-interpretation to $\cor_{l_{b},u_{b}}$.
+which has a similar interpretation to $\cor_{l_{b},u_{b}}$.
 Our third statistic, $\Dsq_{l_b,u_b}$, is the average squared 
@@
 For bivariate datasets, we compared values 
-of the statistics $\cor_{l_b,u_b}$,
+of $\cor_{l_b,u_b}$,
 $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_{0,u_b}-\cor_{1-u_b,1}$, 
@@
 $\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$ to 
-distributions of values of the same statistics computed on surrogate datasets that were produced from 
+distributions of the same statistics computed on surrogate datasets that were produced from 
 the empirical data by randomizing it in a special way to have no 
@@
 of its values computed on surrogate datasets 
-provides a test of whether association between the variables in the part of the distributions specified 
+provides a test of whether association between the variables in the 
+part of the distributions specified 
 by $l_b$ and $u_b$ is different from
@@
 likewise, using $l_b=0.9$ and $u_b=1$ tests for upper-tail associations different 
-from that of a normal copula.
+from a normal copula.
 The comparison of one of the statistics 
@@
 $\Dsq_{1-u_b,1}-\Dsq_{0,u_b}$,
-as computed on the empirical data, to the distribution of its values on many surrogate datasets 
+as computed on the empirical data, to the distribution of its values on surrogate datasets 
 provides a test of asymmetry of tail associations. Thus this comparison helps address the
-third part of Q1 for bivariate 
-datasets.
+third part of Q1.
 
 For multivariate datasets, we calculated Spearman and Kendall correlations
-and the statistics 
-$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
+and $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$,
 $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$
@@
 (Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals of these means.
-The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
-widely used spline correlogram method. The same scheme was also used by @walteretal2017 
-as part of a method for 
-performing model selection among matrix regression models.
-Code and data for this
+The scheme is identical to that proposed by @bjornstadfalck2001 and also 
+used by @walteretal2017. Code and data for this
 project are archived at \url{www.github.com/sghosh89/BIVAN}.
@@
 answering Q1 in the affirmative. 
-To make results easier to absorb, we first 
-present results for an example bivariate dataset, then we summarize results
-for all bivariate datasets, then we present results for an example multivariate dataset, then we 
-summarize for all multivariate datasets.
+To make results easier to absorb, we
+present results first for an example bivariate dataset, then in summary
+for all bivariate datasets, then for an example multivariate dataset, then 
+in summary for all multivariate datasets.
 
@@
 minimum-AIC copula, with AIC `r round(min(BivMS_res_Loecke$InfCritRes$AIC),2)`, whereas the 
-normal copula had a much higher AIC, `r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)`, answering the 
+normal copula had much higher AIC, `r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)`, answering the 
 first part of Q1 for these data: the data have a dependence structure markedly distinct 
-from that of a normal copula. Recall that AIC differences of $2$
+from a normal copula. Recall that AIC differences of $2$
 or $3$ are considered meaningful and differences of $8$ or greater
-are considered definitive, so that data provide essentially no support
+are definitive, so that data provide essentially no support
 for the higher-AIC model in that case [@burnham2003_modelselection].
@@
 for which they are both plentiful than in soils for which they are 
-both less abundant, and C and N are more strongly associated in both
+both scarce, and C and N are more strongly associated in both
 tails than would be modelled using a normal copula model or methods that
@@
 available from `BiCopGofTest`).
-Our alternative nonparametric
-approach, results detailed next, provides information about significance of tail
+Our nonparametric results, detailed next, provide information about significance of tail
 associations and asymmetries of tail associations.
 
-The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
+The values of $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
 C and N data
-were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
-of the data,
+were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates,
 separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
 (Fig. \ref{fig_soilCN_nonparam}). Results confirmed that tail associations of data are stronger in 
-both the lower and upper tails than would be expected under a null hypothesis of a normal copula (Fig. \ref{fig_soilCN_nonparam}B-F).
+both the lower and upper tails than would be expected under a normal-copula null hypothesis (Fig. \ref{fig_soilCN_nonparam}B-F).
 The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
 $\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
-on surrogates, separately in two 
-comparisons (Table \ref{tab_soilCN_asym}). Results confirmed that
+on surrogates (Table \ref{tab_soilCN_asym}). Results confirmed that
 upper-tail association was significantly stronger than lower-tail association; i.e., C and N values are
@@
 are consistent with the model selection results, but go beyond them by providing information
-about statistical significance. Thus our results provide an
+about significance. Thus our results provide an
 affirmative answer to Q1 for the soil C and N data; this answer is 
@@
 selection results do not provide 
-information on statistical significance of these differences, and are subject to the
+information on significance of these differences, and are subject to the
 caveat that, for some datasets, even the lowest-AIC copula was not an objectively good fit
@@
 associated with each other when both these quantities are large than when they are 
-both small. Our generally affirmative (except for Cedar Creek), 
+small. Our generally affirmative (except for Cedar Creek), 
 empirically based answer to Q1 for the bivariate datasets is 
@@
 parameters are *significantly* different from $0$, and we refer 
-the reader to nonparametric results below for information
-on significance. 
+the reader to nonparametric results below for that information. 
 Model-averaged lower- minus upper-tail dependence statistics 
@@
 also were always positive and had confidence intervals that 
-excluded $0$ 
-(Table \ref{tab_count_resamp}).
+excluded $0$ (Table \ref{tab_count_resamp}).
 
@@
 Thus the spatial synchrony of rarity was stronger than that of outbreaks for 
-green spruce aphid and \textit{Ceratium furca} abundance, whereas leaf-curling aphid
+green spruce aphid and \textit{Ceratium furca} abundance, for instance, whereas leaf-curling aphid
 first flights were more spatially related to each other when these flights happened
@@
       caption = "Soil C and N results for nonparametric tests of whether asymmetry of tail association was significant, compared to a normal-copula null hypothesis.
-The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving normal surrogates, in seperate comparisons.
+The values of the listed statistics (cor$_{FS}$-cor$_{LS}$ = $\\cor_{0,0.1}-\\cor_{0.9,1}$, P$_{FS}$-P$_{LS}$ = $\\Ps_{0,0.1}-\\Ps_{0.9,1}$, D$_{LS}^2$-D$_{FS}^2$ = $\\Dsq_{0.9,1}-\\Dsq_{0,0.1}$; here FS stands for `first slice' and LS for `last slice') for real data were compared to their values for 1000 Kendall- or Spearman-preserving normal surrogates, in separate comparisons.
 A table entry $<X$ indicates the value of the given statistic on the data was less than its value on $X$ of the surrogates, so entries of the form $<X$ for $X$ equal to $975$ or above indicate that upper-tail dependence was significantly stronger than lower-tail dependence.
@@
 Having demonstrated that non-normal copula structure and asymmetric tail associations are 
-common in ecology, we addressed Q2 by exploring, using models, three possible mechanisms
-by which these phenomena may arise. 
-Models presented are initial explorations, only, of whether the proposed mechanisms have the 
-potential to explain observed patterns. As such, simple models were used. 
+common, we addressed Q2 by exploring, using models, three possible mechanisms
+for these phenomena. 
+Our models presented are initial explorations, only, of whether the proposed 
+mechanisms may explain observed patterns. As such, simple models were used. 
 Comprehensive explorations of
@@
 very simple AR(1) formulation, influenced by a "regional" environmental factor which
-is correlated across locations ($\epsilon$) and by an entirely local factor ($\delta$). 
-The function $g$ describes how the environmental factor $\epsilon_i(t)$ influences
+is correlated across locations ($\epsilon$) and by a local factor ($\delta$). 
+The function $g$ describes how $\epsilon_i(t)$ influences
 $E_i(t)$. We used $g$ equal to $g_1$ or $g_2$ in different simulations:
@@
 $g_1$ represents environmental effects that negatively impact populations,
-and that occur only below a threshold of $\epsilon=0$; 
-and $g_2$ represents effects that positively impact populations and 
-that occur only above a threshold of $\epsilon=0$.
-The values of $b$ selected provide a modicum of exploration of the potential
-for ecological dynamics to also influence how complex dependencies between
-the $E_i(t)$ may arise: 
-negative values of $b$ correspond to overcompensating dynamics 
-and positive values to undercompensating dynamics; larger $|b|$
+but only below the threshold $\epsilon=0$; 
+and $g_2$ represents effects that positively impact populations, but 
+only above $\epsilon=0$.
+The values of $b$ provide a modicum of exploration of whether ecological 
+dynamics may also influence how complex dependencies between
+the $E_i(t)$ arise: 
+negative $b$ corresponds to overcompensating dynamics 
+and positive $b$ to undercompensating dynamics; larger $|b|$
 means slower return to equilibrium after a disturbance.
@@
 dependence across locations in different simulations (see below).
-The variable $E$ is very general, and could represent deviations
+The variable $E$ is general. It could represent deviations
 of a population density from a carrying capacity, deviations of
@@
 $\sigma=1$, model dynamics are strongly nonlinear because the stochastic 
-component of the model causes populations to stray far from the carrying capacity, $K$. 
+component of the model causes populations to stray far from the carrying capacity. 
 We refer to these as
 the weak-noise and strong-noise cases, though the importance of the noise here
-is that, when it is strong, it brings the nonlinearities of the model into play.
+is that, when strong, it brings the nonlinearities of the model into play.
 
@@
 time steps. For each simulation, the following
-statistics were then computed for both noises and populations: 
+statistics were then computed for noises and populations: 
 Pearson, Spearman and Kendall correlations, $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, 
@@
 trait data across species, e.g., our bird and mammal 
-data. This mechanism is a hypothetical explanation for the bias toward right tail association
+data. This mechanism is a hypothetical explanation for the bias toward right-tail association
 observed in those data (Fig. \ref{fig_biv_multi_raw_cop}F,G). 
@@
 the phylogeny, taken from @GenoudIM2018, of $817$ mammal species.
-The root character state and the change 
+The root character state and change 
 across each branch were randomly chosen from matrices of one million 
@@
 of tail associations between the ecological variables $E_i$ in the two locations, 
-to an extent that depended on the value of $b$.
+to an extent that depended on $b$.
 For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
@@
 It is explained in the Discussion why the box and some of the arrows 
-are solid, instead of dashed, although the results of this section are theoretical.
+are solid, instead of dashed, although the results here are theoretical.
 
@@
 section \\ref{Causes} (i.e., Figs \\ref{fig_ms_asym_b_0.1} and 
-\\ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to nonparamtric statistical 
-analyses decribed in section \\ref{Methods}. The statistic cor$_{FS}$ - cor$_{LS}$ was computed for the
+\\ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}) were subjected to nonparametric statistical 
+analyses described in section \\ref{Methods}. The statistic cor$_{FS}$ - cor$_{LS}$ was computed for the
 output of the model with the indicated b and g, and the value was compared to 1000 values
@@
 in an ecological variable can be inherited from an environmental 
-variable was found to be reasonable, because it held for the models we examined -
+variable was found to be reasonable, because it held for our models -
 the Moran effect seems to extend to copula structure and tail associations - 
@@
 noise, to be paired with larger (respectively, smaller) values of the same
-statistics as applied to model outputs. 
+statistics for model outputs. 
 For our nonlinear model with weak noise, values of the statistics

### 2019-12-18 SupportingInformation.Rmd
continuing to trim words here and there

@@
 ---
-title: "Supporting information: Copulas and their potential for ecology"
+title: "Copulas and their potential for ecology: Appendices"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance D. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
@@
 with stronger correlation. Asymmetry of tail dependence (LT-UT) is the same 
-within each column. See section \ref{Background} in the main text for a formal description of 
+within each column. See section \ref{MT-Background} in the main text for a formal description of 
 tail dependence and other measures of association in the tails.\label{fig_cop_pedag}}

### 2019-12-19 Paper.Rmd
finished trimming the main text

@@
 speaking, were approximately similar between noise and model
-outputs; however, similarity was reduced relative to previous results, 
+outputs; however, similarity was reduced, 
 and for a few simulations, asymmetry statistics
@@
 $\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
-population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}).
+population outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}).
 
@@
 at $K$ when $r<1$ (undercompensating dynamics, e.g., the value $r=0.5$ used initially), 
-but exhibits an oscillatory approach to $K$ when $r>1$ (overcompensating
+but exhibits an oscillatory approach when $r>1$ (overcompensating
 dynamics, e.g., $r=1.3$). For weak noise, similarities were again dominant between
@@
 were associated with similar asymmetries of tail associations in extant characters.
-Thus, based on our simulations, we cannot reject the possibility that
+Thus we cannot reject the possibility that
 this is a proximate mechanism behind observed asymmetric tail associations
@@
 We addressed Q3 by exploring, using both data and models, three hypothesized
-ecological consequences of non-normal copula structure and asymmetric tail
-associations. As for Q2, the models we used are intended as initial explorations, only,
-of whether copula structure may have a meaningful influence on important ecological
-phenomena. Therefore only simple models are used, and comprehensive explorations of the 
+consequences. As for Q2, the models we used are intended as initial explorations, only. 
+Therefore only simple models are used, and comprehensive explorations of the 
 sensitivity of results to model strucuture and parameters are left for future work.
@@
 exceptionally large values tend to occur at the same time in most locations.
-We hypothesize that this 
-will tend to increase the skewness of the distribution 
+We hypothesize that this can increase the skewness of the distribution 
 of the spatial mean. Similarly, left-tail associations between local variables
-should tend to decrease skewness. Strong positive skewness of the 
+should decrease skewness. Strong positive skewness of the 
 spatial-mean time series corresponds to "spikiness" of that time series,
@@
 
-To further address Q3, we also examined a hypothesis that 
+We also examined a hypothesis that 
 asymmetric tail associations across space of an environmental variable can influence the 
 extinction risk of a metapopulation. We hypothesized that environmental noises
-exhibiting greater left-tail associations across different habitat patches would 
+exhibiting greater left-tail associations across habitat patches would 
 cause higher metapopulation extinction risks because then very bad years for the 
-component populations would occur simultaneously in many patches,
+component populations occur simultaneously in many patches,
 reducing rescue effects. Here we assume, for simplicity, 
@@
 
-Finally, and pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
+Finally, and pursuing ideas from @Reuman2017, we 
 tested whether the copula structure of the dependences between population variables 
@@
 in simulation $k$, regardless of X;
-2) for any given $k$ and $j_1$ and $j_2$ such that $j_1 \neq j_2$, the Spearman correlations (computed through time)
+2) for any given $k$, and $j_1$ and $j_2$ such that $j_1 \neq j_2$, the 
+Spearman correlations (computed through time)
 $\cor_t(m_{t,j_1}^{(\text{X},k)},m_{t,j_2}^{(\text{X},k)})$ were the same, to within sampling
@@
 common across multiple sub-disciplines
-in ecology, although these facets of ecological data are only occasionally 
-taken into account [@Valpine2014; @Anderson2018; @Popovic2018]. We hypothesized 
+in ecology, although these facets of data are only occasionally 
+accounted for [@Valpine2014; @Anderson2018; @Popovic2018]. We hypothesized 
 mechanisms that may cause non-normal copula structures and asymmetric tail associations;
@@
 mechanisms may operate. 
-We also hypothesized and demonstrated important consequences of 
-non-normal copula structure and asymmetric tail associations for the field of ecology. 
+We also demonstrated important consequences of 
+non-normal copula structure and asymmetric tail associations for ecology. 
 For instance, the skewness
@@
 averaged time series showing accentuated "crashes".
-Thus tail associations could have implications for pest and exploited species.
+Thus tail associations could have implications for pests and exploited species.
 Extinction risk and Taylor's law can also be altered by tail association 
@@
 complete copula characterization of associations between variables is employed. 
-Copula approaches are
-well developed, statistically [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], 
+Copula statistics are
+well developed [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], 
 and have been introduced in accessible formats [@Anderson2018; @Genest2007]. Furthermore, 
 open-source computer implementations exist (e.g., the `copula` and 
-`VineCopula` packages in the R programming language). Ecologists can apply these tools
+`VineCopula` packages in R). Ecologists can apply these tools
 immediately. We created several interrelated randomization procedures
 (Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
-that have built upon existing copula methods.
+that built upon existing copula methods.
 
 <!--tropical ecosystems-->
-The approaches we have demonstrated should apply equally well to 
+The approaches we demonstrated should apply equally well to 
 data from tropical or temperate ecosystems.
@@
 asymmetric tail associations should be special properties of datasets from temperate 
-regions. The mechanisms we have proposed of non-normal copula structure and asymmetric tail 
+regions. The mechanisms we proposed of non-normal copula structure and asymmetric tail 
 associations seem equally likely to apply anywhere. 
@@
 law and nonlinear environmental influences on ecosystems, which underpin 
-another one of our proposed mechanisms (Fig. \ref{MasterFigure}, A), are widely
-demonstrated phenomena that apply in all biomes.
+another one of our mechanisms (Fig. \ref{MasterFigure}, A), are widely
+demonstrated phenomena.
 
@@
 Because these factors, which are the essential ingredients of
-the mechanism, are so common, it is reasonable to hypothesize that the mechanism
-may operate commonly. It may be a dominant cause of 
+the mechanism, are common, it is reasonable to hypothesize that the mechanism
+may operate commonly and may be a dominant cause of 
 asymmetric tail associations and non-normal copula struture of ecological 
 dependencies across space.
-We provide some further support for the mechanism 
+We provide further support for the mechanism 
 in our discussion of green spruce aphids and
@@
 showed that a Gumbel copula effectively modeled 
-the spatial dependence of drought variables. The Gumbel copula demonstrates 
-asymmetric tail associations.
+the spatial dependence of drought variables. The Gumbel copula has 
+asymmetric tail association.
 Bivariate copula analysis was also used in forecasting the 
@@
 But the nonparametric nature of approaches
-    based solely on data ranks, as many of our copula methods are, may be a promising avenue for 
+    based solely on data ranks, as many of our methods are, may be a promising avenue for 
     avoiding inaccuracies that arise via the highly structured model assumptions implicit
@@
 fishes with low age of maturity, but little such association for later-maturing
-fishes because those fishes may invest the resources not invested in maturing quickly into
+fishes because those species may invest the resources not invested in maturing quickly into
 either fecundity or juvenile survival.
 These ideas suggest that copulas may be 
-useful for studying multi-dimensional trade-offs. But applications will 
+useful for studying multi-dimensional life-history trade-offs. But applications will 
 require careful attention to the possible consequences of biased sampling:
@@
 bivariate copula methods (e.g., the `VineCopula`
-package for the R programming language). Such approaches may be a 
-useful next step in life-history theory and other areas.
+package for R). Such approaches may be a 
+useful next step in life-history theory.
 
@@
 Additional mechanisms of non-normal copula structure and asymmetric tail associations
-probably also operate, and should be considered. For instance, 
+probably also operate. For instance, 
 measurement error may modify copula structure. 
@@
 were intentionally simple in other respects, 
-too, and did not include factors such as
+too, not including factors such as
 delayed density dependence, dispersal, population stage structure, trophic
@@
 structure of interspecific relationships in communities and its implications for community 
-variability is likely to be valuable, in our opinion.
+variability seems likely to be valuable.
 
@@
 of non-normal copula structure and asymmetric tail associations in ecological 
-systems (Fig. \ref{MasterFigure}), but this is done using a variety of 
+systems (Fig. \ref{MasterFigure}), but this was done using a variety of 
 datasets and models.
-We here take a closer look at the green spruce aphid, which simultaneously 
-illustrates causes and consequences with one system.
+We here take a closer look at the green spruce aphid, because it simultaneously 
+illustrates causes and consequences within one system.
 Green spruce aphid abundance, as measured in the data we use,
@@
 that measures of dependence which use unranked data conflate information about
-the marginal distributions of variables with information about the dependence
+the marginal distributions of variables with information about the association
 between variables. For instance, suppose the true population densities $p_{1,t}$
@@
 but are assessed using catch per unit effort (CPUE), a standard approach in fisheries
-science [@Zale2013]. The CPUE measurements, which we denote $c_{1,t}$ and $c_{2,t}$ may be 
+science [@Zale2013]. The CPUE measurements, which we denote $c_{1,t}$ and $c_{2,t}$, may be 
 differently correlated through time than the true densities would be, if they were known,
@@
 under some circumstances, in part because ecological mechanisms which produce relationships 
-between variables are influenced not only by the ordering information in ranks but also 
+between variables are influenced not only by ranks but also 
 by the relative-size information in the unranked data. Judgements of what statistics
@@
 Data requirements for copula methods will vary depending on multiple factors,
-but our results show that data requirements are not beyond what is common 
+but our results show that requirements are not beyond what is common 
 in ecology. With copula methods as with any statistical analysis, the data required to detect 
@@
 detection of strong asymmetry. Different copula or copula-related methods will
-also require different amount of data. For instance, our nonparametric methods 
+also require different amounts of data. For instance, our nonparametric methods 
 such as the partial Spearman correlation should be effective with datasets 
@@
 many copula methods will be suitable for datasets of sizes that are common in ecology.
-The partial Spearman correlation seems particularly widely applicable.
+Partial Spearman correlation seems particularly widely applicable.
 
@@
 in ecology both of and 
-beyond the specific tools we have employed. For instance, multivariate copulas are useful for 
+beyond the specific tools we employed. For instance, multivariate copulas are useful for 
 studying relationships between multiple interacting variables, and have been used in finance

### 2019-12-19 Paper.Rmd
what to do about the notation table

@@
 monotonic cdfs, and this case is simpler. So we 
-henceforth make such an assumption.
+henceforth make such an assumption. Table SX provides a summary
+of notation.
+<!--***DAN: Shyamolina, once I email you an edited version of the
+table you produced, please make a table in the sup mat Rmd and cause it to be
+cited here in place of the SX above. Please place the table in such
+a way that SX is replaced with S1,  since I believe this is now going
+to be the first supplementary table we are citing.-->
 

### 2019-12-20 Paper.Rmd
eqn edited upto sec8 in maintext

@@
 used measure of association, is modified by monotonic
-tranformation of the component variables (Appendix 
+transformation of the component variables (Appendix 
 \ref{SM-pedagog1_details}, Fig. \ref{SM-fig_transf_pearson_spearman}),
@@
 research is that prior work demonstrated, using copulas, complex 
-stucture in the spatial dependence of environmental 
+structure in the spatial dependence of environmental 
 variables [@Serinaldi2008; @She2018; @Goswami2018;
@@
 (Q2) If the answer to
-Q1 is "yes," then what are some possible causes/mechanisms of non-normal
+Q1 is "yes", then what are some possible causes/mechanisms of non-normal
 copula structure and asymmetric tail associations in ecology? And (Q3), what are the 
@@
 Multiple analyses are brought to bear on each question, and these
-are summarized diagramatically in Fig. \ref{MasterFigure}, which can 
+are summarized diagrammatically in Fig. \ref{MasterFigure}, which can 
 serve as a *post hoc* guide to the paper. 
@@
 at each of the data points. Therefore $\hat{F}_X(x_j)$ is the rank of $x_j$ in
-the set $\{x_i : i=1,\ldots,n\}$, dividided by $n$, 
+the set $\{x_i : i=1,\ldots,n\}$, divided by $n$, 
 and analogously $\hat{F}_Y(y_j)$ is the rank of $y_j$ in the set 
@@
 the $p$-value from either test was $<0.2$, tests were re-run with 
-$1000$ boostraps.
+$1000$ bootstraps.
 
@@
 understand the details of these copulas,
-and, additonally, these copulas have been 
+and, additionally, these copulas have been 
 described in detail elsewhere [@Joe1997; @Brechmann2013]. So 
@@
 \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8}). 
-The pictorial decriptions can also be used to aid quick comparisons of copula
+The pictorial descriptions can also be used to aid quick comparisons of copula
 families when alternative families are being considered for future applications. 
@@
 abbreviating $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
-and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"). Likewise $\Ps_l=\Ps_{0,.5}$,
+and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"). Likewise $\Ps_l=\Ps_{0,0.5}$,
 $\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, $\Dsq_u=\Dsq_{0.5,1}$.
@@
 Assume the dynamics 
-$E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where 
+$E_i(t+1)=bE_i(t)+g(\varepsilon_i(t))+a \delta_i(t)$, where 
 the $\delta_i(t)$ are standard-normally distributed 
@@
 $b=0.1, -0.1, 0.5, -0.5$ in different simulations, and 
-the $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
-time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
-$\cov(\epsilon_1,\epsilon_2)=0.8$. Thus basic ecological dynamics follow the
+the $(\varepsilon_1(t),\varepsilon_2(t))$ were drawn, independently through
+time, from a bivariate normal distribution with $\var(\varepsilon_i)=1$ and 
+$\cov(\varepsilon_1,\varepsilon_2)=0.8$. Thus basic ecological dynamics follow the
 very simple AR(1) formulation, influenced by a "regional" environmental factor which
-is correlated across locations ($\epsilon$) and by a local factor ($\delta$). 
-The function $g$ describes how $\epsilon_i(t)$ influences
+is correlated across locations ($\varepsilon$) and by a local factor ($\delta$). 
+The function $g$ describes how $\varepsilon_i(t)$ influences
 $E_i(t)$. We used $g$ equal to $g_1$ or $g_2$ in different simulations:
-$g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise;
-$g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
+$g_1(\varepsilon)$ equals $\varepsilon$ if $\varepsilon<0$ and $0$ otherwise;
+$g_2(\varepsilon)$ equals $\varepsilon$ if $\varepsilon>0$ and $0$ otherwise. Thus
 $g_1$ represents environmental effects that negatively impact populations,
-but only below the threshold $\epsilon=0$; 
+but only below the threshold $\varepsilon=0$; 
 and $g_2$ represents effects that positively impact populations, but 
-only above $\epsilon=0$.
+only above $\varepsilon=0$.
 The values of $b$ provide a modicum of exploration of whether ecological 
@@
 produce asymmetric tail associations (and therefore non-normal copula structure)
-between the $E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
+between the $E_i(t)$. Because $(\varepsilon_1,\varepsilon_2)$
 and $(\delta_1,\delta_2)$ have normal copula structure (they were drawn
@@
 We use AR(1) dynamics, 
-$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
-The environmental noises $\epsilon_i(t)$ were 
+$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \varepsilon_i(t)$, with $\beta=0.5$. 
+The environmental noises $\varepsilon_i(t)$ were 
 standard-normal random variables that were independent for distinct times, $t$, but 
@@
 of the Ricker model, 
-$P_i(t+1)=P_i(t)\exp \left[r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right]$ for $i=1,2$, using $r=0.5$, 
-$K=100$, $\epsilon_i(t)$
+$P_i(t+1)=P_i(t)\exp \left[r\left(1-P_i(t)/K  \right)+\sigma\varepsilon_i(t)\right]$ for $i=1,2$, using $r=0.5$, 
+$K=100$, $\varepsilon_i(t)$
 as above, and $\sigma=0.1$ or 
@@
 $50$ replicate simulations, we generated 5000 noise pairs 
-$(\epsilon_1(t),\epsilon_2(t))$ from the
+$(\varepsilon_1(t),\varepsilon_2(t))$ from the
 bivariate random variable with standard-normal marginals and with 
@@
 The hypothesis is that asymmetric tail association occurs in evolutionary changes
-in bivariate characters, and gives rise to asymmtric tail association between the 
+in bivariate characters, and gives rise to asymmetric tail association between the 
 two character values across extant species.
@@
 Therefore only simple models are used, and comprehensive explorations of the 
-sensitivity of results to model strucuture and parameters are left for future work.
+sensitivity of results to model structure and parameters are left for future work.
 
@@
 spatial-mean time series corresponds to "spikiness" of that time series,
-i.e., occassional very large values, 
+i.e., occasional very large values, 
 which corresponds to instability through time. 
@@
 The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
-$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
-$\epsilon_i(t)$ represent environmental noises. They were standard-normally 
+$\exp(r+\varepsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
+$\varepsilon_i(t)$ represent environmental noises. They were standard-normally 
 distributed, were independent through time, and showed the same 
@@
 Appendices \ref{SM-Consequence_TL_simulation} and \ref{SM-sect:copsurrognd}.
-Thus for each $k$, the poulation matrices $m^{(\text{X},k)}$
+Thus for each $k$, the population matrices $m^{(\text{X},k)}$
 for X taking the values Clayton, normal, Frank, and survival Clayton
@@
 
-A variety of Taylor's law statistics were computed for each populaton matrix
+A variety of Taylor's law statistics were computed for each population matrix
 $m^{(\text{X},k)}$. First, the $N=50$ 
@@
 the quadratic regression equation 
-across the values $\log(m_g)$. The distrubutions of all these 
+across the values $\log(m_g)$. The distributions of all these 
 statistics across the 1000 replicate 
@@
 tail association can also be explored using a regression 
-model with heteroskedastic error term, $y=mx+b+\epsilon(x)$. Whereas this 
+model with heteroskedastic error term, $y=mx+b+\varepsilon(x)$. Whereas this 
 approach seems capable of yielding insight for some datasets, it implicitly assumes

### 2019-12-20 Paper.Rmd
changed the N in the statistic for independence to n

--- 2019-12-19 Paper.Rmd
@@
 Second, we tested the independence of the $u_i$ and $v_i$
-using the statistic $\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
+using the statistic $\sqrt{(9n(n-1))/(4n+10)}|\tau|$, where $\tau$
 is Kendall's tau for the data. @Genest2007 argue that this statistic is approximately standard normally

### 2019-12-20 Paper.Rmd
fixed a log that was rendered correctly

@@
 were statistically similar except for the copula structure between locations, which was
-X. Thus comparing how the $log(v_g)$ versus $\log(m_g)$ relationship may
+X. Thus comparing how the $\log(v_g)$ versus $\log(m_g)$ relationship may
 manifest differently for $m^{(\text{X},k)}$

### 2019-12-20 Paper.Rmd
refs checked

@@
 or Spearman (S) correlation coefficients. C and D show normalized ranks of A, B, 
-respectively. (E,F) Reflections
+respectively. (E, F) Reflections
 of C, D, respectively, about a vertical axis. Subscripts h and v 
@@
 we tested the goodness of fit of our AIC-best copula family using a 
-bootstrapping procedure of @Wang2000 and @Genest2006,
+bootstrapping procedure of @Wang2000 and @genest2006goodness,
 implemented as `BiCopGofTest` in `VineCopula`. The procedure performed 
@@
 and, additionally, these copulas have been 
-described in detail elsewhere [@Joe1997; @Brechmann2013]. So 
+described in detail elsewhere [@joe1997multivariate; @Brechmann2013]. So 
 we only briefly identify each family, say a few words about its tail 
@@
 180 degrees. 
-See @Joe1997 and @Brechmann2013 for details on all families. We used the implementations
+See @joe1997multivariate and @Brechmann2013 for details on all families. We used the implementations
 of provided in the `VineCopula` package for the R programming
@@
 model outputs. Error bars are standard errors. 
-$P$-values (triangles, right axis) are for a 
+$p$-values (triangles, right axis) are for a 
 paired $t$-test of the null hypothesis that the distributions have the same mean. 
@@
 surrogate skewness values 
-(Fig. \ref{fig_skewness_spearman}A,B). For datasets with moderately stronger 
+(Fig. \ref{fig_skewness_spearman}A, B). For datasets with moderately stronger 
 lower than upper-tail 
@@
 average showed a non-significant or marginally significant tendency 
-toward being less than surrogate skewnesses (Fig. \ref{fig_skewness_spearman}C,D). 
+toward being less than surrogate skewnesses (Fig. \ref{fig_skewness_spearman}C, D). 
 Thus copula 
@@
 simulations and Taylor's law appeared visually to be a reasonable approximation
-of the $\log(v_g)$ versus $\log(m_g)$ relationship (Fig. \ref{fig_spTL}B,E,F). 
+of the $\log(v_g)$ versus $\log(m_g)$ relationship (Fig. \ref{fig_spTL}B, E, F). 
 Furthermore, quadratic coefficients
-and curvature values were close to $0$ (Fig. \ref{fig_spTL}J,K), and root mean squared
+and curvature values were close to $0$ (Fig. \ref{fig_spTL}J, K), and root mean squared
 errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
@@
 copula structure
-(Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
-non-zero (Fig. \ref{fig_spTL}J,K); and root mean squared
+(Fig. \ref{fig_spTL}A, C, D, E, F); quadratic coefficients and curvatures were frequently 
+non-zero (Fig. \ref{fig_spTL}J, K); and root mean squared
 errors from the linear regression were much higher (Fig. \ref{fig_spTL}G). Slopes 
 and intercepts of the linear regression were also strongly affected by copula
-structure (Fig. \ref{fig_spTL}H,I), though some of the effect here was because
+structure (Fig. \ref{fig_spTL}H, I), though some of the effect here was because
 linear regressions do not always adequately represent the $\log(v_g)$ versus $\log(m_g)$
@@
 may operate commonly and may be a dominant cause of 
-asymmetric tail associations and non-normal copula struture of ecological 
+asymmetric tail associations and non-normal copula structure of ecological 
 dependencies across space.
@@
 a hypothesis only, because the greater right-tail association shown in 
-Fig. \ref{fig_biv_multi_raw_cop}F,G
+Fig. \ref{fig_biv_multi_raw_cop}F, G
 could have come about in another, unknown way rather than via the mechanism we suggested
@@
     assumptions about the degree of phylogenetic inertia displayed by the traits
-    under study [@Ives2018;@Harmon2018].
+    under study [@Ives2018; @Harmon2018].
 Even if one were able to simply use a time-based set of branch lengths, assigning
@@
 than they would otherwise. It seems likely that asymmetric influence of winter 
-termperature on populations may be a common phenomenon, so effects such as 
+temperature on populations may be a common phenomenon, so effects such as 
 we have documented for green spruce aphid may be common.
@@
 goal) using normalized ranks, stating explicitly that "statistical inference concerning
-dependence structures should always be based on ranks." The reason for this is
+dependence structures should always be based on ranks". The reason for this is
 that measures of dependence which use unranked data conflate information about
@@
 extraction; and David Tilman, Lauren Hallet, Jonathan Walter, Thomas 
-Anderson, Lei Zhao, Andrew Rypel, and editors and annonymous reviewers for helpful suggestions. 
+Anderson, Lei Zhao, Andrew Rypel, and editors and anonymous reviewers for helpful suggestions. 
 We thank James Bell of the Rothamsted Insect Survey (RIS). 
@@
 California Delta Science Program.
-.
 

### 2019-12-20 SupportingInformation.Rmd
suppmat notation table need to fix, created with NA as of now

@@
 gathered for at least 50 dates.
+
+<!--table with summarized notation used in maintext-->
+```{r tab_summary_notation, echo=F, results='asis',message=F}
+library(tinytex)
+library(tibble)
+library(kableExtra)
+library(dplyr)
+
+tab_summary_notation<- tibble(Notation=c("xi,yi", "$ui,vi$",
+                                 "(X,Y)", "F",
+                                 "FX, FY","C",
+                                 "D", "GA,GB",
+                                 "$\\lambda_{L}$, $\\lambda_{U}$",
+                                 "l$_{b}$, u$_{b}$",
+                                 "cor$_{lb,ub}$",
+                                 "cor$_{l}$","cor$_{u}$",
+                                 "P$_{lb,ub}$, P$_{l}$,P$_{u}$, D$_{lb,ub}^2$, D$_{l}^2$, D$_{u}^2$",
+                                 "b", "g,g$_{1}$,g$_{2}$",
+                                 "E","P"), 
+                              Meaning=c("Variables representing raw bivariate data",
+                                "Normalized ranks of $(x_{i}$,$y_{i})$",
+                                "A bivariate random vector",
+                                "The bivariate cdf of $(X,Y)$",
+                                "The cdfs of the marginals, $X$ and $Y$, respectively",
+                                "The copula for $(X,Y)$",
+                                "A bivariate copula",
+                                "Two univariate cdfs",
+                                "Lower and upper tail dependence of a copula",
+                                "A lower and upper bound satisfying $0 \\leq l_{b} \\lt u_{b} \\leq 1$",
+                                "partial Spearman correlation",
+                                "cor$_{lb,ub}$ when $l_{b} = 0, u_{b} = 0.5$",
+                                "cor$_{lb,ub}$ when $l_{b} = 0.5, u_{b} = 1$",
+                                "Related nonparametric statistics",
+                                "AR$(1)$ coefficient in the model related to Leibigs law",
+                                "Function describing environmental influence in the model related to Leibigs law",
+                                "An ecological variable",
+                                "A population variable"))
+
+# Shya need to fix this error
+
+knitr::kable(NA, "latex", booktabs = T, linesep = "", escape=F,
+      caption = "Summary of principal notation.
+Many notations are defined and used only within a small part of the paper.
+This table summarizes notation that is used across larger portions of the paper.
+Section headers refer to the section in which the notation was originally defined.
+cdf = cumulative distribution function.
+\\label{tab_summary_notation}")
+```
 

### 2019-12-20 SupportingInformation.Rmd
adding short notation
test case: endline editing, committing from linux for change made in windows

@@
 ---
-title: "Copulas and their potential for ecology: Appendices"
+title: "Copulas and their potential for ecology: Appendices for main text"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance D. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
@@
 knitr::kable(NA, "latex", booktabs = T, linesep = "", escape=F,
+             caption.short = "Summary of principal notation",
       caption = "Summary of principal notation.

### 2019-12-20 SupportingInformation.Rmd
suppmat checked, need to fig out table s1

@@
 ---
-title: "Copulas and their potential for ecology: Appendices for main text"
+title: "Copulas and their potential for ecology: Appendices"
 author: "Shyamolina Ghosh, Lawrence W. Sheppard, Mark T. Holder, Terrance D. Loecke, Philip C. Reid, James D. Bever, Daniel C. Reuman"
@@
 Fig. \ref{fig_transf_pearson_spearman}A can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
-of both variables by the cummulative distribution function (cdf) of the standard normal distribution, and then by the inverse of 
+of both variables by the cumulative distribution function (cdf) of the standard normal distribution, and then by the inverse of 
 the cdf of the gamma distribution with shape and scale parameters 2. 
@@
 but the Spearman correlations are unchanged (Figs \ref{MT-fig_cop_pedag1}A-B, \ref{fig_transf_pearson_spearman}).
-A scatterplot very simlar to Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
+A scatterplot very similar to Fig. \ref{MT-fig_cop_pedag1}C can be obtained from Fig. \ref{MT-fig_cop_pedag1}A via transformation
 of both variables by the cdf of the standard normal distribution.
@@
 
-\textbf{Environmental data, soil C and N.} Measurments were from the central pedon of a soil 
+\textbf{Environmental data, soil C and N.} Measurements were from the central pedon of a soil 
 core, to 1m depth.
@@
 This table summarizes notation that is used across larger portions of the paper.
-Section headers refer to the section in which the notation was originally defined.
+Third coloumns to the section in which the notation was originally defined.
 cdf = cumulative distribution function.
@@
 datasets. The moderately right-tail dependent matrix and the symmetric tail dependence
-matrix were constructed similarly, but using an survival Clayton copula and an normal copula,
+matrix were constructed similarly, but using a survival Clayton copula and a normal copula,
 respectively.
@@
 $k$th-smallest (non-`NA`) element of the time series 
-$a_i(t)$, $t=1,\ldots,T$ by the $k$th smallest (non-`NA`) 
+$a_i(t)$, $t=1,\ldots,T$ by the $k^{th}$ smallest (non-`NA`) 
 element of $x_i(t)$, $t=1,\ldots,T$.
@@
 by permuting time series of values in the normal-copula array, separately for each 
-sampling location and simulation. So population marginal distrbutions for each location
+sampling location and simulation. So population marginal distributions for each location
 and simulation were *exactly* the same as those in the normal-copula population array. 
@@
 draws from $C$.
-- For each $k$ and $i$, replace the $k$th-smallest element of the time series $a_i(t)$, 
-$t=1,\ldots,T$ by the $k$th smallest element of $x_i(t)$, $t=1,\ldots,T$. Call the 
+- For each $k$ and $i$, replace the $k^{th}$-smallest element of the time series $a_i(t)$, 
+$t=1,\ldots,T$ by the $k^{th}$ smallest element of $x_i(t)$, $t=1,\ldots,T$. Call the 
 result $y_i(t)$. This is the surrogate dataset.

### 2019-12-23 Paper.Rmd
suppmat notation table need to fix, created with NA as of now
minor corrections made while checking Shyas conversion to Word

@@
 
-\noindent \emph{The number of words in the abstract:} 253
+\noindent \emph{The number of words in the abstract:} 241
 
@@
 such relationships are limited, and 
-provide only a partial description the relationship. 
+provide only a partial description of the relationship. 
 For instance, datasets showing different 
@@
 monotonic cdfs, and this case is simpler. So we 
-henceforth make such an assumption. Table SX provides a summary
+henceforth make such an assumption. Table \ref{SM-tab_summary_notation} provides a summary
 of notation.
@@
 See @joe1997multivariate and @Brechmann2013 for details on all families. We used the implementations
-of provided in the `VineCopula` package for the R programming
+provided in the `VineCopula` package for the R programming
 language. See Figs \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
@@
            "8. Percent pairs showing stronger lower tail dependence",
-           "9. Mean with 2.5$^{th}$ and 97.5$^{th}$ quantiles of cor$_{l}$ - cor${_u}$ across all location pair")
+           "9. Mean with 2.5$^{th}$ and 97.5$^{th}$ quantiles of cor$_{l}$ - cor${_u}$ across all location pairs")
            )
@@
 
-Our second mechanism is an extension of the well known Moran
+Our second mechanism is an extension of the well-known Moran
 effect, and was summarized conceptually in the Introduction. 
@@
 about the association. Such statements of 
-mathematical completeness will not be avilable for other methods.
+mathematical completeness will not be available for other methods.
 

### 2019-12-23 SupportingInformation.Rmd
minor corrections made while checking Shyas conversion to Word

@@
 18 species (4 species, each, of C4 grasses, C3 grasses, legumes, non-legume forbs; and 2 species of woody
-plants). All plots received, in total, 10 $gm^{-2}$ of pure live seed in May 1994 and 5 $gm^{-2}$ in May, 1995 
+plants). All plots received, in total, 10 gm$^{-2}$ of pure live seed in May 1994 and 5 gm$^{-2}$ in May, 1995 
 with seed mass divided equally among species. Treatments were maintained by weeding manually 3-4 times a year 
@@
 sampled by clipping 
-different locations each year. For each available year, we analyzed above-ground biomass ($gm^{-2}$) 
+different locations each year. For each available year, we analyzed above-ground biomass (gm$^{-2}$) 
 and Shannon's biodiversity index (computed using percent cover data) averaged over the 
@@
 This table summarizes notation that is used across larger portions of the paper.
-Third coloumns to the section in which the notation was originally defined.
+Third column refers to the section in which the notation was originally defined.
 cdf = cumulative distribution function.
@@
 \caption[Sampling locations for C and N data]{Locations of central pedons sampled 
-for soil organic carbon and total soil nitrogen (Mg C or N per hectare of soil 
+for soil organic carbon and total soil nitrogen (megagrams C or N per hectare of soil 
 surface) at 5907 sites across the coterminous United States. \label{fig_SoilCNmapUSA}}
@@
 and $p_2 \in (0,1]$, lower-tail dependence is $0$ and upper-tail
-dependence is $2-2^{1/p_1}$ of $p_2=1$ and $0$ otherwise.\label{figped_BB8}}
+dependence is $2-2^{1/p_1}$ if $p_2=1$ and $0$ otherwise.\label{figped_BB8}}
 \end{center}
@@
 and $p_2 \in (0,1]$, upper-tail dependence is $0$ and lower-tail
-dependence is $2-2^{1/p_1}$ of $p_2=1$ and $0$ otherwise.\label{figped_SBB8}}
+dependence is $2-2^{1/p_1}$ if $p_2=1$ and $0$ otherwise.\label{figped_SBB8}}
 \end{center}
@@
 We use both statistics 
-because $\cor_{l_{b_k},u_{b_k}}(u,v)$ is more familiar, but 
+because $\cor_{l_{b},u_{b}}(u,v)$ is more familiar, but 
 in some instances $\Ps_{l_b,u_b}$ appears to have more power to
@@
 \caption[Testing nonparametric statistics with 35 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the normal and Frank copulas.
-p-values (indicated by the triangles) when less than $0.05$ (below the dashed line) mean there is significant tail asymmetry.
+p-values (indicated by the triangles), when less than $0.05$ (below the dashed line), mean there is significant tail asymmetry.
 See Appendix  \ref{Test_npa_stat} for details.
@@
 \caption[Testing nonparametric statistics with 1000 data points]{Results of tests of whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ reveal the known asymmetry of tail dependence of the Clayton and survival Clayton copulas, and the known symmetry of tail dependence of the normal and Frank copulas.
-p-values (indicated by the triangles) when less than $0.05$ (below the dashed line) mean there is significant tail asymmetry.
+p-values (indicated by the triangles), when less than $0.05$ (below the dashed line), mean there is significant tail asymmetry.
 See Appendix  \ref{Test_npa_stat} for details.
@@
 $k$th-smallest (non-`NA`) element of the time series 
-$a_i(t)$, $t=1,\ldots,T$ by the $k^{th}$ smallest (non-`NA`) 
+$a_i(t)$, $t=1,\ldots,T$ by the $k^{\text{th}}$ smallest (non-`NA`) 
 element of $x_i(t)$, $t=1,\ldots,T$.
@@
 because the Kendall correlation in the very commonly used `cor` function in R 
-is the Kendall tau-b coefficient in the event of ties, and index which 
+is the Kendall tau-b coefficient in the event of ties, an index which 
 differs in its definition from the commonly used tau-a that works when there are
@@
 of the normal distribution constructed above. This was done using the `rho` and `iRho` functions 
-in the `copula` package in R. Because the Clayton copula is a one-paremeter family, this 
+in the `copula` package in R. Because the Clayton copula is a one-parameter family, this 
 step relies on the fact that all pairs of locations were 
@@
 if $m^{(\text{N},k)}_{t,j}$ is our normal-copula population array (the three dimensions 
-of the array are indexed, respectively, by $t$, $j$ and $k$, and $m^{(\text{C},k)}_{t,j}$ is our 
+of the array are indexed, respectively, by $t$, $j$ and $k$), and $m^{(\text{C},k)}_{t,j}$ is our 
 Clayton-copula population array, then,
@@
 draws from $C$.
-- For each $k$ and $i$, replace the $k^{th}$-smallest element of the time series $a_i(t)$, 
-$t=1,\ldots,T$ by the $k^{th}$ smallest element of $x_i(t)$, $t=1,\ldots,T$. Call the 
+- For each $k$ and $i$, replace the $k^{\text{th}}$-smallest element of the time series $a_i(t)$, 
+$t=1,\ldots,T$ by the $k^{\text{th}}$ smallest element of $x_i(t)$, $t=1,\ldots,T$. Call the 
 result $y_i(t)$. This is the surrogate dataset.

### 2019-12-30 SupportingInformation.Rmd
summary notation table in suppmat fixed

@@
 <!--table with summarized notation used in maintext-->
-```{r tab_summary_notation, echo=F, results='asis',message=F}
-library(tinytex)
+```{r tab_summary_notation, echo=F, results='markup',message=F}
 library(tibble)
@@
 
-tab_summary_notation<- tibble(Notation=c("xi,yi", "$ui,vi$",
-                                 "(X,Y)", "F",
-                                 "FX, FY","C",
-                                 "D", "GA,GB",
+tab_summary_notation<- tibble(Notation=c("$x_{i}$,$y_{i}$", "$u_{i}$,$v_{i}$",
+                                 "$(X,Y)$", "F",
+                                 "$F_{X}$, $F_{Y}$","$C$",
+                                 "$D$", "$G_{A}$,$G_{B}$",
                                  "$\\lambda_{L}$, $\\lambda_{U}$",
-                                 "l$_{b}$, u$_{b}$",
+                                 "$l_{b}$, $u_{b}$",
                                  "cor$_{lb,ub}$",
@@
                                  "P$_{lb,ub}$, P$_{l}$,P$_{u}$, D$_{lb,ub}^2$, D$_{l}^2$, D$_{u}^2$",
-                                 "b", "g,g$_{1}$,g$_{2}$",
-                                 "E","P"), 
+                                 "$b$", "$g,g_{1},g_{2}$",
+                                 "$E$","$P$"), 
                               Meaning=c("Variables representing raw bivariate data",
-                                "Normalized ranks of $(x_{i}$,$y_{i})$",
+                                "Normalized ranks of ($x_{i}$,$y_{i}$)",
                                 "A bivariate random vector",
@@
                                 "Lower and upper tail dependence of a copula",
-                                "A lower and upper bound satisfying $0 \\leq l_{b} \\lt u_{b} \\leq 1$",
+                                "A lower and upper bound satisfying $0 \\leq l_{b} \\leq u_{b} \\leq 1$",
                                 "partial Spearman correlation",
-                                "cor$_{lb,ub}$ when $l_{b} = 0, u_{b} = 0.5$",
-                                "cor$_{lb,ub}$ when $l_{b} = 0.5, u_{b} = 1$",
+                                "cor$_{lb,ub}$ when $l_{b}$ = 0, $u_{b}$ = 0.5",
+                                "cor$_{lb,ub}$ when $l_{b}$ = 0.5, $u_{b}$ = 1",
                                 "Related nonparametric statistics",
-                                "AR$(1)$ coefficient in the model related to Leibigs law",
+                                "AR$(1)$ coefficient in the model related to Leibig's law",
                                 "Function describing environmental influence in the model related to Leibigs law",
                                 "An ecological variable",
-                                "A population variable"))
-
-# Shya need to fix this error
-
-knitr::kable(NA, "latex", booktabs = T, linesep = "", escape=F,
+                                "A population variable"),
+                                "Section in maintext"=c(1,1,
+                                                      2,2,2,2,2,2,2,
+                                                      4,4,4,4,4,
+                                                      6,6,6,6))
+
+knitr::kable(tab_summary_notation, "latex", booktabs = T, linesep = "", escape=F,
              caption.short = "Summary of principal notation",
@@
 cdf = cumulative distribution function.
-\\label{tab_summary_notation}")
+\\label{tab_summary_notation}")%>%
+     column_spec(column = 2,width="4.5 cm")
 ```

### 2020-02-03 Paper.Rmd
correction for proof started

@@
 variables. Some of these datasets nevertheless have the same Pearson (P)
-or Spearman (S) correlation coefficients. C and D show normalized ranks of A, B, 
+or Spearman (S) correlation coefficients. (C) and (D) show normalized ranks of (A), (B), 
 respectively. (E, F) Reflections
-of C, D, respectively, about a vertical axis. Subscripts h and v 
+of (C), (D), respectively, about a vertical axis. Subscripts h and v 
 in the axis labels stand for "horizontal" and "vertical".\label{fig_cop_pedag1}}
@@
 considering the plot of $u_i$ versus $v_i$, where
-$u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
+$u_i$ is the rank of $x_i$ in the set of $x_j$ ($j=1,\ldots,n$), divided by
 $n+1$; and $v_i$ is defined in the same way but using the $y_i$. The rank
@@
 be a useful component of that structure, so tail association 
-is a main focus of this paper. We give four examples below which
+is a main focus of this paper. We give four examples, below, which
 illustrate some of the ecological meaning of tail associations.
@@
 in a few studies to illuminate ecological phenomena 
-[@Valpine2014; @Anderson2018; @Popovic2018], but such usage is
+[@Valpine2014; @Anderson2018; @Popovic2019], but such usage is
 still rare. Whereas @Anderson2018 provide a specific method, based
@@
 some causal hypotheses which could not previously be eliminated. 
-@Popovic2018 used copulas to illuminate causal relationships
+@Popovic2019 used copulas to illuminate causal relationships
 between species.
@@
 sub-fields and organizational levels of ecology. 
-We address Q2 (sections \ref{Causes} through \ref{Results_Q2}), using 
+We address Q2 (sections \ref{Causes} through \ref{Results_Q2}) using 
 simple models. We address Q3 (sections \ref{Consequences} through \ref{Results_Q3}) using 
@@
 and have also been used effectively, but rarely so far, in ecology [@Valpine2014; 
-@Anderson2018; @Popovic2018]. 
+@Anderson2018; @Popovic2019]. 
 We argue that benefits of wider use of
@@
 Clayton family is the symmetric opposite of the Clayton family, showing
-stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
+stronger right- than left-tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
 family is a two-parameter family, thereby providing more flexibility: for some parameters
@@
 with a random variable $(X,Y)$ is the cdf of the random variable
-$(F_X(X),F_Y(Y))$. Given a sample $(x_i,y_i)$, $i=1,\ldots,n$, from $(X,Y)$, let $\hat{F}_X$ and
+$(F_X(X),F_Y(Y))$. Given a sample $(x_i,y_i)$, $i=1,\ldots,n$ from $(X,Y)$, let $\hat{F}_X$ and
 $\hat{F}_Y$ be the empirical cdfs associated with the $x_i$ and $y_i$, respectively.
@@
 regarded as approximate samples from the random variable $(F_X(X),F_Y(Y))$, 
-which is the random variable associated with the copula of $(X,Y)$.
-The scatterplot of $v_i$ versus $u_i$ can be used to infer aspects of copula 
+which is the random variable associated with the copula of $(X,Y)$. The scatterplot 
+of $v_i$ versus $u_i$ can be used to infer aspects of copula 
 structure (see section \ref{Methods} below).
@@
 These data were computed from the Continuous Plankton Recorder survey of the 
-UK Marine Biological Survey.
+UK Marine Biological Association.
 Community-level data, obtained from the Cedar Creek Ecosystem Science Reserve, were 
@@
 in ecology, although these facets of data are only occasionally 
-accounted for [@Valpine2014; @Anderson2018; @Popovic2018]. We hypothesized 
+accounted for [@Valpine2014; @Anderson2018; @Popovic2019]. We hypothesized 
 mechanisms that may cause non-normal copula structures and asymmetric tail associations;
@@
 that fuller explorations, in future work,
-of some of our models and of variant model may be informative. 
+of some of our models and of variant models may be informative. 
 We hope by enumerating a few potential mechanisms of copula structure

### 2020-02-03 FirstPedagogFig.R

@@ -217,7 +217,7 @@ par(fig=c((ymarg_wd+numsp)/tot_wd,
           (xmarg_ht+numsp)/tot_ht,
           (xmarg_ht+numsp+pan_ht)/tot_ht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
-plot(dcc5[,1],dcc5[,2],type='p',pch=pchval,cex=cexvalpts,xlim=c(rg1,rg2),ylim=c(rg1,rg2),col="grey",cex.axis=cexaxs)
+plot(dcc5[,1],dcc5[,2],type='p',pch=pchval,cex=cexvalpts,xlim=c(rg1,rg2),ylim=c(rg1,rg2),col="grey",cex.axis=cexaxs,xaxs="i",yaxs="i")
 lines(c(0,.4),c(.4,0),type='l')
 lines(c(0,.65),c(.65,0),type='l')
 lines(c(0,1.6),c(1.6,0),type='l')

### 2020-02-04 Plotter_Cause4copula_stat.R
legend color modified

@@ -437,8 +437,8 @@ Plotter_Cause4copula_stat<-function(res){
   pdf(paste0(resloc,"common_legend_cause4copula_stat.pdf",sep=""),height=1,width=15)
   op<-par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(0, 0, 0, 0))
   plot(0, 0, type = "n", bty = "n", xaxt = "n", yaxt = "n")
-  legend("center", c("input noise copula","model output copula", "p-value(paired t-test)"), col = c("dimgrey", "black", "black"),
-         cex = 2.5, pch = c(1, 1, 2), xpd = TRUE, horiz = T, inset = c(0,0), text.col = c("dimgrey", "black", "black"),
+  legend("center", c("input noise copula","model output copula", "p-value(paired t-test)"), col = c("grey70", "black", "black"),
+         pt.lwd=4.5, cex = 2.5, pch = c(1, 1, 2), xpd = TRUE, horiz = T, inset = c(0,0), text.col = c("grey70", "black", "black"),
          bty = "n")
   par(op)
   dev.off()

### 2020-03-02 Paper.Rmd
legend color modified
small changes on correcting of AER proofs

@@
 But $\hat{F}_X(x_i)$ equals the normalized rank of $x_i$ times
-$n/(n+1)$, and likewise $\hat{F}_Y(y_i)$ equals the normalized rank of $y_i$ times
-$n/(n+1)$. So the normalized rank in turn approximates the empirical cdf and therefore 
+$(n+1)/n$, and likewise $\hat{F}_Y(y_i)$ equals the normalized rank of $y_i$ times
+$(n+1)/n$. So the normalized rank in turn approximates the empirical cdf and therefore 
 the cdf. Thus the normalized rank pairs $(u_i,v_i)$ for $i=1,\ldots,n$ can be 
@@
 \includegraphics[width=6.5in]{./Results/BivarDataPlot.pdf}
-\caption{Upper panels show raw data plots for (A) $\log_{10}$(soil C) and $\log_{10}$(soil N) data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
+\caption{Upper panels show raw data plots for (A) $\log_{10}$(soil N) vs. $\log_{10}$(soil C) data, (B) $\log_{10}$(basal metabolic rate (BMR)) vs. $\log_{10}$(body mass) 
 for birds, (C) the same for mammals, and (D) above-ground plant biomass vs.
-Shannon's index (\textit{H}) from Cedar Creek.
+Shannon's index from Cedar Creek.
 Bottom panels show the corresponding normalized rank plots.
